### 2020-06-23 latex/chapter-hydras.tex
first commit

\chapter{Hydras and Hydra Games}
\label{sec:orgheadline91}

This chapter is dedicated to the representation of hydras and rules of the hydra game in \coq's specification language:\gallina.

Technically, a \emph{hydra} is just a finite ordered tree, each node of which 
has any number of sons. Note that, contrary to the computer science tradition, we will show the hydras 
with the heads up and the foot (i.e., the root of the tree) down.
Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the module 
\url{../src/html/hydras.Hydra.Hydra_Examples.html}).

\begin{figure}[h]
\begin{tikzpicture}[very thick, scale=0.6]

\node (foot) at (2,0) {$\bullet$};
\node (N1) at (2,2) {$\bullet$};
\node (N2) at (2,4) {$\bullet$};
\node (N3) at (2,6) {$\bullet$};
\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
\node (H4) at (4,2) {$\Smiley[2][vertfluo]$};
\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
\draw (foot) -- (N1)[very thick] ;
\draw (N1) -- (N2);
\draw (N2) -- (N3);
\draw (N3) to [bend left= 10]  (H1) ;
\draw (N3) to [bend right= 16] (H2);
\draw (foot) to [bend left= 10]  (H0) ;
\draw (foot) to [bend right = 10] (H4) ;
\draw (N1) to [bend right= 16] (H5);
\end{tikzpicture}
\caption{The hydra Hy \label{fig:Hy}}
\end{figure}

We use a specific vocabulary for talking about hydras. Table~\ref{tab:hyd2tree} shows the correspondance between our terminology and the usual vocabulary for trees in computer science.

\begin{center}
\begin{tabular}{ll}
Hydras & Finite rooted trees\\
\hline
foot & root\\
head & leaf\\
node & node\\
segment  & (directed) edge \\
sub-hydra & subtree\\
daughter & immediate subtree\\
\end{tabular}
\captionof{table}{\label{tab:hyd2tree}
Translation from hydras to trees}
\end{center}

The hydra \texttt{Hy} has a \emph{foot} (below), five \emph{heads}, and eight \emph{segments}. 
We leave it to the reader to define various parameters such as the height, the size, the highest arity (number of sons of a node) of a hydra. In our example, these parameters have the respective values : $4$, $9$ and $3$.

\subsection{The rules of the game}

\label{sec:orgheadline44}
A \emph{hydra battle} is a fight between Hercules and the Hydra. 
More formally, a  battle is a sequence of \emph{rounds}.
At each round:
\begin{itemize}
\item If the hydra is composed of just one head, the battle is finished
and  Hercules is the winner
\item Otherwise, Hercules chops off \emph{one} head of the hydra

\begin{itemize}
\item If the head is at distance 1 from the foot, the head is just lost by the hydra
\item Otherwise, let us denote by \(r\) the node that was at distance \(2\) from 
the removed head in the direction of the foot,  and consider the  sub-hydra \(h'\) of \(h\), whose  root is \(r\) \footnote{$h'$ will be called ``the wounded part of the hydra'' in the subsequent text. In Figures~\vref{fig:Hy2} and ~\vref{fig:Hy4}, this sub-hydra  is displayed in red.}. Let $n$ be some natural number.
Then $h'$ is replaced by  $n+1$ of copies of \(h'\) which share the same root $r$.
 The \emph{replication number} $n$ may be different (and generally is)   at each round of the fight.
It may be chosen by the hydra, according to its strategy, or imposed by some 
particular rule. In many presentations of hydra battles, this number is increased by $1$ at each round. In the following presentation, we will also consider battles where the hydra is free to chose its number of replication at every round of the battle\footnote{Let us recall that, if the chopped-off head was at distance 1 from the foot, the replication factor is meaningless.}.
\end{itemize}
\end{itemize}

Note that the description given in~\cite{KP82} of the replication process in hydra battles is also  semi-formal.

\label{original-rules}

\begin{quote}
  ``From the node that used to be attached to the head which was just chopped off, traverse one 
segment towards the root until the next node is reached. From this node sprout $n$ replicas of 
that part of the hydra (after decapitation) which is ``above'' the segment just traversed, i.e., those 
nodes and segments from which, in order to reach the root, this segment would have to be 
traversed. If the head just chopped off had the root of its nodes, no new head is grown. ''
\end{quote}

Moreover, we note that this description is in \emph{imperative} terms. In order to build a formal  study of the properties of hydra battles, we prefer to use mathematical language, i.e., graphs, relations, functions, etc.
Thus, the replication process will be represented as a binary relation on a data type \texttt{Hydra},
linking the state of the hydra \emph{before} and \emph{after} the transformation.
A battle will thus be represented as a sequence of terms of type \texttt{Hydra}.

\subsection{Example}
Let us start a battle between Hercules and the hydra \texttt{Hy} of Fig.~\ref{fig:Hy}.

At the first round, Hercules choses to chop off the rightmost head of \texttt{Hy}.
Since this head is near the floor, the hydra loses this head. Let us call 
 \texttt{Hy'} the resulting state of the hydra, represented in Fig.~\vref{fig:Hy-prime}.

Next, assume Hercules choses to chop off one of the two highest heads of \texttt{Hy'}, for instance the rightmost one. Fig.~\vref{fig:Hy2} represents the rotten neck in dashed lines, and the part that will be replicated in red. Assume also that the hydra decides to add 4 copies of the red part. We obtain a new state \texttt{Hy''} depicted in Fig.~\ref{fig:Hy3}.

\begin{figure}[h]
\begin{tikzpicture}[very thick, scale=0.6]

\node (foot) at (2,0) {$\bullet$};
\node (N1) at (2,2) {$\bullet$};
\node (N2) at (2,4) {$\bullet$};
\node (N3) at (2,6) {$\bullet$};
\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
\node (H4) at (6,0) {$\Xey[2][lightgray]$};
\draw (foot) -- (N1)[very thick] ;
\draw (N1) -- (N2);
\draw (N2) -- (N3) ;
\draw (N3) to [bend left= 10]  (H1) ;
\draw (N3) to [bend right= 16] (H2);
\draw (foot) to [bend left= 10]  (H0) ;
\draw (N1) to [bend right= 16] (H5);
\end{tikzpicture}

\caption{Hy': the state  of Hy after one round \label{fig:Hy-prime}}
\end{figure}

\begin{figure}[hp]

\begin{tikzpicture}[very thick, scale=0.5]

\node (foot) at (2,0) {$\bullet$};
\node (N1) at (2,2) {$\bullet$};
\node (N2) at (2,4)  {{\color{lightred}$\bullet$}};
\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
\node (H1) at (0,8) {$\Sey[2][lightred]$};
\node (H2) at (5,0) {$\Xey[2][lightgray]$};
\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
\node (ex) at (5,8) {};
\draw (foot) -- (N1)[very thick] ;
\draw (N1) -- (N2);
\draw  (N2) -- (N3)[draw=lightred];
\draw (N3) to   [bend left= 10](H1) [draw=lightred];
\draw [dashed] (N3) to [bend left= 10](ex);
\draw (foot) to [bend left= 10]  (H0) ;
\draw (N1) to [bend right= 16] (H5);
\end{tikzpicture}
\caption{A beheading}
\label{fig:Hy2}
\end{figure}

\begin{figure}[hp]
\begin{tikzpicture}[very thick, scale=0.6]

\node (foot) at (2,0) {$\bullet$};
\node (N1) at (2,2) {$\bullet$};
\node (N2) at (2,4) {$\bullet$};
\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
\node (H11) at (2,8) {$\Smiley[2][vertfluo]$};
\node (H12) at (4,8) {$\Smiley[2][vertfluo]$};
\node (H13) at (6,8) {$\Smiley[2][vertfluo]$};
\node (H14) at (8,8) {$\Smiley[2][vertfluo]$};

\node (N3) at (1,6) {$\bullet$};
\node (N31) at (2,6) {$\bullet$};
\node (N32) at (3,6) {$\bullet$};
\node (N33) at (4,6) {$\bullet$};
\node (N34) at (5,6) {$\bullet$};

\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
\draw (foot) -- (N1)[very thick] ;
\draw (N1) -- (N2);
\draw (N2) -- (N3);
\draw (N2) -- (N31);
\draw (N2) -- (N32);
\draw (N2) -- (N33);
\draw (N2) -- (N34);
\draw (N3) to   [bend left= 10](H1) ;
\draw (N31) to   [bend left= 10](H11) ;
\draw (N32) to   [bend left= 10](H12) ;
\draw (N33) to   [bend left= 10](H13) ;
\draw (N34) to   [bend left= 10](H14) ;
\draw (foot) to [bend left= 10]  (H0) ;
\draw (N1) to [bend left= 10]  (H5) ;
\end{tikzpicture}
\caption{Hy'', the state of Hy after two rounds \label{fig:Hy3}}
\end{figure}

Figs.~\ref{fig:Hy4} and~\vref{fig:Hy5} represent a possible third round of the battle, with a replication factor equal to $2$. Let us call \texttt{Hy'''} the state of the hydra after that third round.

\begin{figure}[hp]
\begin{tikzpicture}[very thick, scale=0.6]

\node (foot) at (2,0)  {{\color{lightred}$\bullet$}};
\node (N1) at (2,2) {{\color{lightred}$\bullet$}};
\node (N2) at (2,4) {{\color{lightred}$\bullet$}};
\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
\node (exN4) at (4,4) {};
\node (H1) at (0,8) {$\Sey[2][lightred]$};
\node (H11) at (2,8) {$\Sey[2][lightred]$};
\node (H12) at (4,8) {$\Sey[2][lightred]$};
\node (H13) at (6,8) {$\Sey[2][lightred]$};
\node (H14) at (8,8) {$\Sey[2][lightred]$};

\node (N3) at (1,6) {{\color{lightred}$\bullet$}};
\node (N31) at (2,6) {{\color{lightred}$\bullet$}};
\node (N32) at (3,6) {{\color{lightred}$\bullet$}};
\node (N33) at (4,6) {{\color{lightred}$\bullet$}};
\node (N34) at (5,6) {{\color{lightred}$\bullet$}};

\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
\node (H5) at (4,0) {$\Xey[2][lightgray]$};
\draw (foot) -- (N1)[very thick,draw=lightred] ;
\draw (N1) -- (N2)[draw=lightred];
\draw (N2) -- (N3)[draw=lightred];
\draw (N2) -- (N31)[draw=lightred];
\draw (N2) -- (N32)[draw=lightred];
\draw (N2) -- (N33)[draw=lightred];
\draw (N2) -- (N34)[draw=lightred];
\draw (N3) to   [bend left= 10](H1) [draw=lightred];
\draw (N31) to   [bend left= 10](H11) [draw=lightred];
\draw (N32) to   [bend left= 10](H12) [draw=lightred];
\draw (N33) to   [bend left= 10](H13) [draw=lightred];
\draw (N34) to   [bend left= 10](H14) [draw=lightred];
\draw (foot) to [bend left= 10]  (H0) ;
\draw [dashed] (N1) to  [bend left= 10](exN4);
\end{tikzpicture}
\caption{A third beheading (wounded part in red) \label{fig:Hy4}}
\end{figure}

\begin{figure}[hp]
\begin{tikzpicture}[very thick, scale=0.4]

\node (foot) at (10,0) {$\bullet$};

\node (N1) at (2,2) {$\bullet$};
\node (N2) at (2,4) {$\bullet$};
\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
\node (H1) at (0,8) {$\Smiley[1][vertfluo]$};
\node (H11) at (2,8) {$\Smiley[1][vertfluo]$};
\node (H12) at (4,8) {$\Smiley[1][vertfluo]$};
\node (H13) at (6,8) {$\Smiley[1][vertfluo]$};
\node (H14) at (8,8) {$\Smiley[1][vertfluo]$};

\node (N3) at (1,6) {$\bullet$};
\node (N31) at (2,6) {$\bullet$};
\node (N32) at (3,6) {$\bullet$};
\node (N33) at (4,6) {$\bullet$};
\node (N34) at (5,6) {$\bullet$};

\node (H0) at (-3,3) {$\Smiley[1][vertfluo]$};

\draw (foot) to [bend left=10] (N1)[very thick] ;
\draw (N1) -- (N2);
\draw (N2) -- (N3);
\draw (N2) -- (N31);
\draw (N2) -- (N32);
\draw (N2) -- (N33);
\draw (N2) -- (N34);
\draw (N3) to   [bend left= 10](H1) ;
\draw (N31) to   [bend left= 10](H11) ;
\draw (N32) to   [bend left= 10](H12) ;
\draw (N33) to   [bend left= 10](H13) ;
\draw (N34) to   [bend left= 10](H14) ;
\draw (foot) to [bend left = 15]  (H0) ;

% second copy 
\node (N01) at (12,2) {$\bullet$};
\node (N02) at (12,4) {$\bullet$};
\node (N03) at (12,6) {{\color{lightred}$\bullet$}};
\node (H001) at (10,8) {$\Smiley[1][vertfluo]$};
\node (H0011) at (12,8) {$\Smiley[1][vertfluo]$};
\node (H0012) at (14,8) {$\Smiley[1][vertfluo]$};
\node (H0013) at (16,8) {$\Smiley[1][vertfluo]$};
\node (H0014) at (18,8) {$\Smiley[1][vertfluo]$};

\node (N03) at (11,6) {$\bullet$};
\node (N031) at (12,6) {$\bullet$};
\node (N032) at (13,6) {$\bullet$};
\node (N033) at (14,6) {$\bullet$};
\node (N034) at (15,6) {$\bullet$};

\draw (foot) -- (N01)[very thick] ;
\draw (N01) -- (N02);
\draw (N02) -- (N03);
\draw (N02) -- (N031);
\draw (N02) -- (N032);
\draw (N02) -- (N033);
\draw (N02) -- (N034);
\draw (N03) to   [bend left= 10](H001) ;
\draw (N031) to   [bend left= 10](H0011) ;
\draw (N032) to   [bend left= 10](H0012) ;
\draw (N033) to   [bend left= 10](H0013) ;
\draw (N034) to   [bend left= 10](H0014) ;

% third copy 
\node (N001) at (22,2) {$\bullet$};
\node (N002) at (22,4) {$\bullet$};
\node (N003) at (22,6) {{\color{lightred}$\bullet$}};
\node (H001) at (20,8) {$\Smiley[1][vertfluo]$};
\node (H0011) at (22,8) {$\Smiley[1][vertfluo]$};
\node (H0012) at (24,8) {$\Smiley[1][vertfluo]$};
\node (H0013) at (26,8) {$\Smiley[1][vertfluo]$};
\node (H0014) at (28,8) {$\Smiley[1][vertfluo]$};

\node (N003) at (21,6) {$\bullet$};
\node (N0031) at (22,6) {$\bullet$};
\node (N0032) at (23,6) {$\bullet$};
\node (N0033) at (24,6) {$\bullet$};
\node (N0034) at (25,6) {$\bullet$};

\draw (foot) -- (N001)[very thick] ;
\draw (N001) -- (N002);
\draw (N002) -- (N003);
\draw (N002) -- (N0031);
\draw (N002) -- (N0032);
\draw (N002) -- (N0033);
\draw (N002) -- (N0034);
\draw (N003) to   [bend left= 10](H001) ;
\draw (N0031) to   [bend left= 10](H0011) ;
\draw (N0032) to   [bend left= 10](H0012) ;
\draw (N0033) to   [bend left= 10](H0013) ;
\draw (N0034) to   [bend left= 10](H0014) ;
\end{tikzpicture}
\caption{The configuration Hy''' of Hy \label{fig:Hy5}}
\end{figure}
\FloatBarrier

We leave it to the reader  to guess the following  rounds of the battle \dots
 % Please keep in mind that, in this 
% the hydra is free to chose any number of replications at each time, whereas
% Hercules chops only one head per round.

% Let us precise that, in this game, Hercules wins if the hydra is eventually reduced 
% to a single head. 
% We know from~\cite{KP82} that, whichever the initial configuration of the
% hydra, and the strategies of both players, Hercules eventually wins. The 
% aforementionned paper shows also that there do not exist any \emph{simple} proof of this result.

\section{Hydras and their representation in \emph{Coq}}
\label{sec:orgheadline48}

Various \emph{Coq} libraries  represent trees the nodes of which 
have a non-fixed number of sons. The reader can, for instance, look at
Chapter 14, pages 400-406 of \cite{BC04}. 
Our representation uses two mutual inductive types: \texttt{Hydra} to represent
 any hydra, and \texttt{Hydrae} to represent finite sequences of hydras.

\index{Types!Hydra}

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Hydra}{Hydra.Hydra\_Definitions}}

\begin{Coqsrc}
Inductive Hydra : Set :=
|  node :  Hydrae -> Hydra
with Hydrae : Set :=
| hnil : Hydrae
| hcons : Hydra -> Hydrae -> Hydrae.
\end{Coqsrc}

\index{Projects}

\begin{project}[**]

 Another very similar representation could use the \texttt{list} type family instead of the specific 
type \texttt{Hydrae}:

\begin{Coqalt}
Module Alt.

Inductive Hydra: Set :=
  hnode (daughters : list Hydra).

End Alt.
\end{Coqalt}

Using this representation, re-define all the constructions of this chapter.
You will probably have to use patterns described for instance in~\cite{BC04} or the archives of the Coq-club~\cite{Coq}.

\end{project}

\index{Projects}

\begin{project}
The type \texttt{Hydra} above describes hydras as \emph{plane trees}, i.e., as drawn on a sheet of paper or computer screen. Thus, hydras are \emph{oriented},
and it is appropriate to consider a \emph{leftmost} or \emph{rightmost} head of
the beast. It could be interesting to consider another representation, in which
every non-leaf node has a \emph{multi-set} -- not an ordered list -- of daughters.
\end{project}

\subsubsection{Abbreviations}

We provide several notations for ``patterns'' which occur often in our developments.

\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#head}{Hydra.Hydra\_Definitions}}

\begin{Coqsrc}
(** heads *)
Notation head := (node hnil).

(** nodes  with 1, 2 or 3 daughters *)
Notation hyd1 h := (node (hcons h hnil)).
Notation hyd2 h h' := (node (hcons h (hcons h' hnil))).
Notation hyd3 h h' h'' := 
                   (node (hcons h (hcons h' (hcons h'' hnil)))).
\end{Coqsrc}

For instance, the hydra \texttt{Hy}  of Figure~\vref{fig:Hy} is defined in \emph{Gallina} as follows:

\vspace{4mm}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html\#Hy}{Hydra.Hydra\_Examples}}

\begin{Coqsrc}
Example Hy := hyd3 head
                   (hyd2
                      (hyd1 
                         (hyd2 head head))
                      head) 
                   head.
\end{Coqsrc}

Hydras quite frequently contain  multiple  copies of the same pattern. The following functions
will help us to describe and reason about replications in hydra battles.

\vspace{4pt}

\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#hcons_mult}{Hydra.Hydra\_Definitions}}

\begin{Coqsrc}
Fixpoint hcons_mult (h:Hydra)(n:nat)(s:Hydrae):Hydrae :=
  match n with 
  | O => s
  | S p => hcons h (hcons_mult h p s)
  end.

(** hydra with n copies of the same daughter *)

Definition hyd_mult h n :=
  node (hcons_mult h n hnil).
\end{Coqsrc}

\vspace{4mm}

For instance, the hydra $Hy''$ of Fig~\vref{fig:Hy3}  can be defined in \coq{} as follows:

\vspace{4pt}

\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}

\begin{Coqsrc}
Example Hy'' := 
     hyd2 head
          (hyd2 (hyd_mult (hyd1 head) 5)
                head).
\end{Coqsrc}

\subsubsection{Recursive functions on type \texttt{Hydra}}
\label{sec:orgheadline41}
\label{sec:hsize-def}

When defining a recursive function over the type \texttt{Hydra}, one has to consider the three constructors 
\texttt{node}, \texttt{hnil} and \texttt{hcons} of the mutually inductive types \texttt{Hydra} and \texttt{Hydrae}. 
Let us define for instance the function that  computes the number of nodes of any hydra:

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}

\begin{Coqsrc}
Fixpoint hsize (h:Hydra) : nat :=
  match h with node l => S (lhsize l)
  end
with lhsize l : nat :=
  match l with hnil => 0
            | hcons h hs => hsize h + lhsize hs 
  end.

 Compute hsize Hy.
\end{Coqsrc}

\begin{Coqanswer}
 = 9
     : nat 
\end{Coqanswer}

Likewise, the \emph{height} (maximum distance between the foot and a head) 
is defined by mutual recursion:

\begin{Coqsrc}
Fixpoint height  (h:Hydra) : nat :=
  match h with node l => lheight l
  end
with lheight l : nat :=
  match l with 
  | hnil => 0
  | hcons h hs => Max.max (S (height h)) (lheight hs)
  end.
\end{Coqsrc}

\begin{Coqsrc}
Compute height Hy.
\end{Coqsrc}

\begin{Coqanswer}
 = 4
     : nat  
\end{Coqanswer}

\index{Exercises}

\begin{exercise}
Define a function \texttt{max\_degree: Hydra $\arrow$ nat} which  returns the highest degree of a node in any hydra. For instance, the evaluation of the term \texttt{(max\_degree Hy)} should return $3$.
\end{exercise}

\subsection{Induction principles for hydras}
\label{sec:orgheadline42}

In this section, we show how induction principles are used to prove properties on the type 
\texttt{Hydra}. Let us consider for instance the following statement:
\begin{quote}
  `` The height of any hydra is strictly less than its size. ''
\end{quote}

\subsubsection{A failed attempt}

One may try to use the default tactic of proof by induction, that corresponds to an application of the automatically  generated  induction principle for  type \texttt{Hydra}:

\begin{Coqanswer}
Hydra_ind :
forall P : Hydra -> Prop,
(forall h : Hydrae, P (node h)) -> forall h : Hydra, P h
\end{Coqanswer}

Ler us start a simple proof by induction.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}

\begin{Coqbad}
Module Bad.

Lemma height_lt_size (h:Hydra) :
  height h <= hsize h.
Proof.
  induction h as [s].
\end{Coqbad}

\begin{Coqanswer}
1 subgoal, subgoal 1 (ID 11)

  s : Hydrae
  ============================
   height (node s) <= hsize (node s)
\end{Coqanswer}

We might be tempted to do an induction on the sequence \texttt{s}:

% \begin{Coqbad}
% induction  s as [ | h s'].
%   -   cbn; auto with arith.
%   - 
% \end{Coqbad}

\begin{Coqanswer}
 1 focused subgoal
(unfocused: 0), subgoal 1 (ID 19)

  h : Hydra
  s' : Hydrae
  IHs' : height (node s') <= hsize (node s')
  ============================
   height (node (hcons h s')) <= hsize (node (hcons h s'))

\end{Coqanswer}

Note that the displayed subgoal does not contain any assumption on \texttt{h}, thus there is no way to 
infer any property about the height and size of the hydra \texttt{(hcons h t)}.

\begin{Coqbad}
Abort.

End Bad.
\end{Coqbad}

\subsubsection{A principle for  mutual induction}
In order to get an appropriate induction scheme for the types 
\texttt{Hydra} and \texttt{Hydrae}, we can use  \coq{}'s  command \texttt{Scheme}.

\index{Coq!Techniques!Mutual induction}
\index{Coq!Commands!Scheme}

\begin{Coqsrc}
Scheme Hydra_rect2 := Induction for Hydra Sort Type
with Hydrae_rect2 := Induction for Hydrae Sort Type.
\end{Coqsrc}

\begin{Coqsrc}
Check Hydra_rect2.
\end{Coqsrc}

\begin{Coqanswer}
Hydra_rect2
 : forall (P : Hydra -> Type) (P0 : Hydrae -> Type),
   (forall h : Hydrae, P0 h -> P (node h)) ->
    P0 hnil ->
    (forall h : Hydra, P h -> 
             forall h0 : Hydrae, P0 h0 -> P0 (hcons h h0)) ->
    forall h : Hydra, P h
\end{Coqanswer}

\subsubsection{A correct proof}

Let us now use \texttt{Hydra\_rect2} for proving that the height of any hydra is strictly less than its size.
Using this scheme requires an auxiliary predicate, called \texttt{P0} in \texttt{Hydra\_rect2}'s statement. 
Let us begin by defining an ad-hoc version of \texttt{List.Forall}.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}

\begin{Coqsrc}
(** All elements of s satisfy P *)

Fixpoint h_forall (P: Hydra -> Prop) (s: Hydrae) :=
  match s with
      hnil => True
    | hcons h s' => P h /\ h_forall P s'
  end.
 \end{Coqsrc}

 \begin{Coqsrc}
Lemma height_lt_size (h:Hydra) :
 height h < hsize h.
Proof.
  induction h using Hydra_rect2  with 
  (P0 :=  h_forall (fun h =>  height h < hsize h)).
 \end{Coqsrc}

\begin{enumerate}
\item The first subgoal is as follows:

\begin{Coqanswer}

  h: Hydrae
  IHh : h_forall (fun h : Hydra => height h < hsize h) h
  ============================
   height (node s) < hsize (node s)

\end{Coqanswer}

This goal is easily solvable, using some arithmetic. We let the reader look at the source of this development.

\item The second subgoal is trivial:

\begin{Coqanswer}

  ============================
    h_forall (fun h : Hydra => height h < hsize h) hnil

\end{Coqanswer}

\begin{Coqsrc}
  reflexivity.
\end{Coqsrc}

\item Finally, the last subgoal is also easy to solve:

\begin{Coqanswer}
  h : Hydra
  h0 : Hydrae
  IHh : height h < hsize h
  IHh0 : h_forall (fun h : Hydra => height h < hsize h) h0
  ============================
   h_forall (fun h1 : Hydra => height h1 < hsize h1) 
                 (hcons h h0)
\end{Coqanswer}

\begin{Coqsrc}
 split;auto. 
Qed.   
\end{Coqsrc}

\end{enumerate}

\index{Exercises}

\begin{exercise}
It happens very often that, in the proof of  a proposition of the form 
(\texttt{$\forall\,$ h:Hydra, $P$ h}), the predicate \texttt{P0}
is  (\texttt{h\_forall $P$}).  Design a tactic for induction on hydras that frees the user from binding explicitly \texttt{P0},  and solves trivial subgoals. Apply it for writing  a shorter proof of \texttt{height\_lt\_size}.
\end{exercise}

\section{Relational description of hydra battles}

In this section, we represent the rules of hydra battles as a binary relation associated with
a \emph{round}, i.e., an interaction composed of the two following actions:
\begin{enumerate}
\item Hercules chops off one head of the hydra
\item Then, the  hydra replicates the wounded part (if the head is at distance $\geq 2$ from the foot).
\end{enumerate}
The relation associated with each round of the battle is \emph{parameterized}  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot).

In our description,  we will apply the following naming convention: if $h$ represents the configuration of the hydra before a round, then the configuration of $h$ after this round will be called $h'$.
 Thus, we are going to define a proposition  (\texttt{round\_n $n\;h\;h'$})  whose intended meaning will be `` the hydra $h$  is transformed into $h'$  in a single round of a battle, with the expected replication factor $n$ ''.

Since the replication of parts of the hydra depends on the distance of the chopped head from  the foot, we  decompose our description into several cases, under the form of a bunch of [mutually] inductive predicates over the types \texttt{Hydra} and \texttt{Hydrae}.

We decompose the relation associated with each  round in two cases: 
\begin{itemize}
\item \textbf{R1}: The chopped off head was at distance 1 from the foot.
\item \textbf{R2}: The chopped off head was at a distance greater than or equal to  $2$ from the foot.
\end{itemize}

\subsection{Chopping off a head at distance 1 from the foot (relation  R1)}

If Hercules chops off a head near the floor, there is no replication at all. We use an auxiliary predicate, associated with the removing of one head from a sequence of hydras.

\pagebreak
\vspace{4pt}\emph{From Module\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}

\begin{Coqsrc}
Inductive S0 :  relation Hydrae :=
| S0_first : forall s, S0  (hcons head s) s
| S0_rest : forall  h s s',  S0  s s' ->
                             S0  (hcons h s) (hcons h s').

Inductive R1  :  Hydra -> Hydra -> Prop :=
| R1_intro : forall s s', S0 s s' -> R1 (node s) (node s').
\end{Coqsrc}

\subsubsection{Example}
\label{sec:orgheadline45}

Let us represent in \coq{}   the transformation of the hydra of Fig.~\vref{fig:Hy} into
the configuration represented in Fig.~\ref{fig:Hy-prime}.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}

\begin{Coqsrc}
Example Hy_1 : R1 Hy Hy'.
Proof. 
  split; right; right; left.
Qed.
\end{Coqsrc}

\subsection{Chopping of a head at distance \texorpdfstring{$\geq 2$}{>= 2} from the foot (relation R2) }

Let us now consider beheadings  where the chopped-off head is at distance greater than or equal to $2$ from the foot. All the following relations are parameterized by the number $n$ of new copies added by the hydra.

 Let $s$ be a sequence of hydras. 
The proposition (\texttt{S1 n s s'}) holds if $s'$ is obtained by replacing some element $h$ of $s$ by 
$n+1$ copies of $h'$, where  the proposition (\texttt{R1 h h'}) holds, in other words, \texttt{h'} is just \texttt{h}, without the chopped-off  head. \texttt{S1} is an inductive relation with two constructors that allow us to choose the position in $s'$ of the wounded sub-hydra $h$.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#S1}{Hydra.Hydra\_Definitions}}

\begin{Coqsrc}
Inductive S1 (n:nat)  : Hydrae -> Hydrae -> Prop :=
| S1_first : forall s h h' ,   R1 h h' -> 
                  S1 n (hcons h s) (hcons_mult h' (S n) s)
| S1_next : forall h s s',  S1 n s s' ->
                   S1 n (hcons h s) (hcons h s').
\end{Coqsrc}

The rest of the definition is structured as two mutually inductive relations on hydras and sequences of hydras. The first constructor of \texttt{R2} describes the case where the chopped head is exactly at height $2$. The others constructors allow us to consider beheadings at height strictly greater than $2$.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions}}

\begin{Coqsrc}
Inductive R2 (n:nat)  :  Hydra -> Hydra -> Prop :=
| R2_intro : forall s s', S1 n s s' -> R2 n (node s) (node s')
| R2_intro_2 : forall s s', S2 n s s' -> R2 n (node s) (node s')

with S2 (n:nat) :  Hydrae -> Hydrae -> Prop :=
|  S2_first : forall h h' s ,
               R2 n  h h'  -> 
               S2  n (hcons h s) (hcons h'  s)
|  S2_next  : forall h   r r',
               S2 n   r r' ->
               S2 n (hcons h r) (hcons h r').                  
\end{Coqsrc}

\subsubsection{Example}
Let us prove the transformation of \texttt{Hy'} into \texttt{Hy''} (see Fig.~\vref{fig:Hy3}).

\vspace{4pt}\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}.

\begin{Coqsrc}
Example R2_example:  R2 4 Hy' Hy''.
Proof.
  right. right. left.  right. left. left. left. split. left.
Qed.
\end{Coqsrc}

\emph{Please consider exercise~\vref{exo:interactive-battle}.}

\subsection{Description of any round}

We combine the two cases above into a single relation.
First,  we define the  relation \texttt{(round\_n n h h')} where \texttt{n} is the expected number of  replications (irrelevant in the case of an \texttt{R1}-transformation).

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#round_n}{Hydra.Hydra\_Definitions}}

\index{Predicates!round\_n}

\begin{Coqsrc}
Definition round_n n h h' := R1 h h' \/ R2 n h h'.  
\end{Coqsrc}

By abstraction over \texttt{n}, we define a \emph{round} (small step) of a battle:

\index{Predicates!round}

\begin{Coqsrc}
Definition round h h' := exists n,  round_n n h h'.

Infix "-1->" := round (at level 60).
\end{Coqsrc}

\index{Projects}

\begin{project}
Give a direct translation of Kirby and Paris's description of hydra battles (quoted on page~\pageref{original-rules}) and prove that our relational description is consistent with theirs.
\end{project}

\subsection{Rounds and battles}

Using library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}, we define \texttt{round\_plus},  the transitive closure of \texttt{round}, and \texttt{round\_star},  the reflexive and transitive closure of \texttt{round}.

\begin{Coqsrc}
Definition round_plus := clos_trans_1n Hydra round.
Infix "-+->" := rounds (at level 60).

Definition round_star h h' := h = h' \/ round_plus h h'.
Infix "-*->" := round_star (at level 60).
\end{Coqsrc}

\index{Notations!''-+->''}
\index{Notations!''-*->''}

\index{Exercises}

\begin{exercise}
Prove the following lemma:

\begin{Coqsrc}
Lemma rounds_height : forall h h', 
   h -+-> h' -> height h' <= height h.  
\end{Coqsrc}

\end{exercise}

\begin{remark}
\coq's library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Coq.Relations.Relation\_Operators} 
contains three logically equivalent definitions of the transitive closure of a binary relation. This equivalence is proved in 
\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} .

Why three definitions for a single mathematical concept?
Each definition generates an associated induction principle. 
 According to the form of statement one would like to prove, there is a ``best choice'':

\begin{itemize}
\item For proving $\forall y, x\,R^+\,y \;\arrow\; P\,y$, prefer 
\texttt{clos\_trans\_n1}
\item For proving $\forall x,\,x\,R^+\,y \;\arrow\; P\,x$, prefer \texttt{clos\_trans\_1n}
\item For proving $\forall x\,y, \,x\,R^+\,y \;\arrow\;P\,x\,y$,  
prefer \texttt{clos\_trans},
\end{itemize}
But there is no ``wrong choice'' at all: the equivalence lemmas in \linebreak 
\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} 
 allow the user
to convert any one of the three closures into another one before applying the corresponding elimination tactic.
The same remark also holds for reflexive and transitive closures. 
\end{remark}

\index{Exercises}

\begin{exercise}
Define a restriction of \coqsimple{round},  where Hercules always chops off
the leftmost among the lowest heads.

Prove that, if $h$ is not a simple head, then there exists a unique $h'$ such that \texttt{h}  is transformed into \texttt{ h'} in one round, according to this restriction.

\end{exercise}

\index{Exercises}

\begin{exercise}[Interactive battles]
\label{exo:interactive-battle}
Given a hydra \texttt{h}, the specification of a hydra battle for \texttt{h} is the type 
\Verb@{h':Hydra | h -*-> h'}@. In order to avoid long sequences of \texttt{split}, \texttt{left}, and 
\texttt{right}, design a set of dedicated tactics for the interactive building of a battle.
Your tactics will have the following functionalities:
\begin{itemize}
\item  Chose to stop a battle, or continue
\item Chose an expected number of replications
\item Navigate in a hydra, looking for a head to chop off.
\end{itemize}

Use your tactics for simulating a small part of a hydra battle, for instance the rounds which lead from
\texttt{Hy} to \texttt{Hy'''}  (Fig.~\vref{fig:Hy5}).

\textbf{Hint:} Please keep in mind that the last  configuration of your interactively built battle is known only at the end of the battle. Thus, you will have to create and solve subgoals with existential variables. For that purpose, the tactic \texttt{eexists}, applied to the 
goal \Verb@{h':Hydra | h -*-> h'}@ generates the subgoal \Verb|h -*-> ?h'|.
\end{exercise}

\subsection{Classes of battles}
\label{sect:battle-classes}

In some presentations of hydra battles, e.g.~\cite{KP82, bauer2008}, the transformation associated with the $i$-th round may depend on $i$. For instance, in these articles, the replication factor at the $i$-th round is equal to $i$. In other examples, one can allow the hydra to apply any replication factor at any time. In order to be the most general as possible, we define the type of predicates which relate the state of the hydra before and after the $i$-th round of a battle.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
\index{Coq!Techniques!Type classes}

\begin{Coqsrc}
Definition dep_round_t := nat -> Hydra -> Hydra -> Prop.

Class Battle :=  {battle_r : dep_round_t;
                  battle_inclusion : forall i h h',
                      battle_r i h h' -> round h h'}.

\end{Coqsrc}

The most general class of battles is \texttt{free}, which allows the hydra to chose any replication factor at every step:

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#free}{Hydra.Hydra\_Definitions}}

\begin{Coqsrc}
Program Instance free : Battle :=
  (Build_Battle ( fun _  h h' => round h h') _).
\end{Coqsrc}

We chosed to call \emph{standard} the kind of battles which appear  most often in the litterature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#standard}{Hydra.Hydra\_Definitions}}

\begin{Coqsrc}
Program Instance standard : Battle := (Build_Battle round_n _).
Next Obligation.
  now exists i.  
Defined.
\end{Coqsrc}

\subsection{Fights}

Let $B$ be some instance of class \texttt{Battle}. It is easy to define inductively the relation between the $i$-th and the $j$-th steps of a battle of type $B$.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}

\begin{Coqsrc}
Inductive fight (B:Battle) : nat -> Hydra -> nat -> Hydra -> Prop :=
|  fight_1 : forall i h  h', battle_r   B i  h h' -> 
                            fight B i h (S i) h'
| fight_n : forall i h  j h' h'',  battle_r  B i h h''  ->
                                   fight B (S i) h'' j h'  ->
                                   fight B i h j h'.
\end{Coqsrc}

% \begin{remark}
%  The class \texttt{free} is strongly related with the transitive closure  \texttt{round\_plus}, as expressed by the following lemmas.

% \vspace{4pt}
% \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}

%  \begin{Coqsrc}
%  Lemma fight_free_equiv1 : forall i j h h',  
%              fight free i h j h' ->   h -+-> h'.

%  Lemma fight_free_equiv2 : forall h h',
%      h -+-> h' ->
%     forall i, exists j,  fight free i h j h'.
%  \end{Coqsrc}

% \end{remark}

\section{A long battle}
\label{sect:big-battle}

In this section we show how long a simple hydra battle can last.  The following example
considers a very small hydra, shown on figure~\vref{fig:hinit}.

\begin{figure}[h]
  \centering
  \begin{tikzpicture}[thick, scale=0.30]
 \node (foot) at (6,0) {$\bullet$};
\node (n1) at  (3,3) {$\bullet$};
\node (h1) at  (1,6) {$\Smiley[1][green]$};
\node (h2) at  (3,6) {$\Smiley[1][green]$};
\node (h3) at  (6,6) {$\Smiley[1][green]$};
\node (h4) at  (6,6) {$\Smiley[1][green]$};
\node (h5) at  (6,3) {$\Smiley[1][green]$};
\node (h6) at  (9,3) {$\Smiley[1][green]$};
\draw (foot) -- (n1);
\draw (n1) to   [bend left=20] (h1);
\draw (n1) to   (h2);
\draw (n1) to   [bend right=20] (h3);
\draw (foot) -- (h5);
\draw (foot) to  [bend right=20] (h6);
\end{tikzpicture}

  \caption{The hydra hinit}
  \label{fig:hinit}
\end{figure}

\begin{Coqsrc}
Definition hinit := hyd3 (hyd_mult head 3)  head head.  
\end{Coqsrc}

Moreover, in order to \emph{compute} the length of a battle, we set Hercules strategy : to chop the rightmost head at each round, and the hydra's strategy : at the $i$-th round, the replication factor is exactly $i$ (which means that the battle is
a \emph{standard} battle, see section~\vref{sect:battle-classes}).

The  battle is so long that no \emph{test} can give us an estimation of its length, and we do need the expressive power of logic to compute this length. However, in order to  guess this length, we made some experiments, computing with \gallina{}, \coq{}'s  functional programming language.
Thus, we can consider this development as a collaboration of proof with computation.

We want to verify whether Hercules wins a standard battle starting from \texttt{hinit} at time $t=0$ , and in this case, in how many rounds. We assume Hercules' 
strategy is to chop off the rightmost head at every round.

The complete proof is in file \url{../src/html/hydras.Hydra.BigBattle.html}.

During the two first rounds, our hydra loses its two rightmost heads. Thus just before the third round, it looks like in figure~\vref{fig:hinit-plus2}.

\begin{figure}[h]
  \centering
  \begin{tikzpicture}[thick, scale=0.30]
 \node (foot) at (3,0) {$\bullet$};
\node (n1) at  (3,3) {$\bullet$};
\node (h1) at  (1,6) {$\Smiley[1][green]$};
\node (h2) at  (3,6) {$\Smiley[1][green]$};
\node (h3) at  (6,6) {$\Smiley[1][green]$};
\node (h4) at  (6,6) {$\Smiley[1][green]$};
\draw (foot) -- (n1);
\draw (n1) to   [bend left=20] (h1);
\draw (n1) to   (h2);
\draw (n1) to   [bend right=20] (h3);
\end{tikzpicture}

  \caption{The hydra (hyd1 h3)}
  \label{fig:hinit-plus2}
\end{figure}

The following lemma  is a formal statement about the two first rounds of the battle.

\begin{Coqsrc}
Lemma L_0_2 : fight standard 0 hinit 2 (hyd1 h3).   
\end{Coqsrc}

A first study with pencil and paper suggested us that, after three rounds, the hydra always looks like in figure~\vref{fig:hinit-plusn} (with a variable number of 
subtrees of height 1 or 0).
Thus, we introduce handy notations.

\begin{Coqsrc}
Notation h3 := (hyd_mult head 3).
Notation h2 := (hyd_mult head 2).
Notation h1 := (hyd1 head).

Definition hyd a b c := 
  node (hcons_mult h2  a
             (hcons_mult h1  b
                         (hcons_mult head c hnil))).
\end{Coqsrc}

For instance Fig~\vref{fig:hinit-plusn} shows the hydra (\texttt{hyd 3 4 2}). The hydra (\texttt{hyd 0 0 0})  is the ``final'' hydra of any terminating battle, {i.e.},
a tree whith exactly one node and no edge.

\begin{figure}[h]
  \centering
  \begin{tikzpicture}[thick, scale=0.30]
 \node (foot) at (15,0) {$\bullet$};
\node (a) at  (3,4) {$\bullet$};
\node (b) at  (6,4) {$\bullet$};
\node (c) at  (9,4) {$\bullet$};
\node (d) at  (13,4) {$\bullet$};
\node (e) at  (16,4) {$\bullet$};
\node (f) at  (19,4) {$\bullet$};
\node (g) at  (22,4) {$\bullet$};
\node (h) at  (25,4) {$\Smiley[1][green]$};
\node (i) at  (28,4) {$\Smiley[1][green]$};
\node (aa) at  (2.5,8) {$\Smiley[1][green]$};
\node (ab) at  (3.5,8) {$\Smiley[1][green]$};
\node (ba) at  (5.5,8) {$\Smiley[1][green]$};
\node (bb) at  (6.5,8) {$\Smiley[1][green]$};
\node (ca) at  (8.5,8) {$\Smiley[1][green]$};
\node (cb) at  (9.5,8) {$\Smiley[1][green]$};
\node (da) at  (13,8) {$\Smiley[1][green]$};
\node (ea) at  (16,8) {$\Smiley[1][green]$};
\node (fa) at  (19,8) {$\Smiley[1][green]$};
\node (ga) at  (22,8) {$\Smiley[1][green]$};
\draw (foot) -- (a);
\draw (foot) -- (b);
\draw (foot) -- (c);
\draw (foot) -- (d);
\draw (foot) -- (e);
\draw (foot) -- (f);
\draw (foot) -- (g);
\draw (foot) -- (h);
\draw (foot) -- (i);
\draw (a) -- (aa);
\draw (a) -- (ab);
\draw (b) -- (ba);
\draw (b) -- (bb);
\draw (c) -- (ca);
\draw (c) -- (cb);
\draw (d) -- (da);
\draw (e) -- (ea);
\draw (f) -- (fa);
\draw (g) -- (ga);
% \node (a) at  (3,4) {$\bullet$};
% \node (h1) at  (1,6) 
% \node (h2) at  (3,6) {$\Smiley[1][green]$};
% \node (h3) at  (6,6) {$\Smiley[1][green]$};
% \node (h4) at  (6,6) {$\Smiley[1][green]$};
% \draw (foot) -- (n1);
% \draw (n1) to   [bend left=20] (h1);
% \draw (n1) to   (h2);
% \draw (n1) to   [bend right=20] (h3);
\end{tikzpicture}

  \caption{The hydra (hyd 3 4 2)}
  \label{fig:hinit-plusn}
\end{figure}

With these notations, we get a formal description of the first three rounds.

\begin{Coqsrc}
Lemma L_2_3 : fight standard 2 (hyd1 h3)  3 (hyd 3 0 0).

Lemma L_0_3 : fight standard 0 hinit 3 (hyd 3 0 0).
\end{Coqsrc}

We would like to have a formal proof that the length of the considered battle is 
equal to some integer $N$, but we have first  to \emph{guess} this number.
In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple datatype for representing the states as tuples composed of
the round number, and the respective number of daughters  \texttt{h2}, \texttt{h1}, and heads
of the current hydra.

\begin{Coqsrc}
 Record state : Type :=
    mks {round: nat ; n2 : nat ; n1 : nat ; nh : nat}.
\end{Coqsrc}

The following function returns the next configurarion of the game. 
Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle will only start with the lemma
\texttt{lemma:step-fight}, page~\pageref{lemma:step-fight}.

\begin{Coqsrc}
Definition next (s : state) :=
  match s with
  | mks round a b (S c) => mks (S round) a b c
  | mks round a (S b) 0 => mks (S round) a b (S round)
  | mks round (S a) 0 0 => mks (S round) a (S round) 0
  | _ => s
  end.
\end{Coqsrc}

We can make bigger steps through iterations of \texttt{next}.

\begin{Coqsrc}
Fixpoint iterate {A:Type} (f : A -> A) (n:nat) (x: A) :=
 match n with 
  | 0 => x
  | S p => f (iterate f p x)
 end.
\end{Coqsrc}

The following function computes the state of the battle at the $n$-th round.

\begin{Coqsrc}
Definition test n := iterate next (n-3) (mks 3 3 0 0).

Compute test 3.
   (**
     = {| round := 3; n2 := 3; n1 := 0; nh := 0 |}
     : state
    *)

Compute test 4.
 (*
  = {| round := 4; n2 := 2; n1 := 4; nh := 0 |}
     : state
 *)

Compute test 5.
 (*
   = {| round := 5; n2 := 2; n1 := 3; nh := 5 |}
     : state
 *)

Compute test 2000.
(*
  = {| round := 2000; n2 := 1; n1 := 90; nh := 1102 |}
     : state
*)
\end{Coqsrc}

The battle we study seems to be awfully long. Let us concentrate our
tests on some particular events : the states where $\texttt{nh}=0$

From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} will be
equal to zero.

 \begin{Coqsrc}
 Compute test 10.
(*

    = {| round := 10; n2 := 2; n1 := 3; nh := 0 |}
     : state
*)
 \end{Coqsrc}

Thus, $ (1 + 11)$ rounds later, the \texttt{n1} field will be equal to $2$, and 
\texttt{nh}  will equal to $0$.

\begin{Coqsrc}
Compute test 22.
(*

 = {| round := 22; n2 := 2; n1 := 2; nh := 0 |}
     : state
*)
\end{Coqsrc}

\begin{Coqsrc}
 Compute test 46.
(*

 = {| round := 46; n2 := 2; n1 := 1; nh := 0 |}
     : state

*)

Compute test 94.

(*

 = {| round := 94; n2 := 2; n1 := 0; nh := 0 |}
     : state

*) 
\end{Coqsrc}

Next round, we decrement \texttt{n2} and set \texttt{n1} to $95$.

\begin{Coqsrc}
 Compute test 95.

(*

  = {| round := 95; n2 := 1; n1 := 95; nh := 0 |}
     : state

*)
\end{Coqsrc}

We now have some intuition of the sequence.
It looks like the next ``\texttt{nh}=0'' event will happen at the $192=2(95+1)$-th round, then at the $2(192+1)$-th round.

\begin{Coqsrc}
 Definition doubleS (n : nat) := 2 * (S n).

Compute test (doubleS 95).

(**
 = {| round := 192; n2 := 1; n1 := 94; nh := 0 |}
     : state
 *)

Compute test (iterate doubleS 2 95).

(*
  = {| round := 386; n2 := 1; n1 := 93; nh := 0 |}
     : state
*)
\end{Coqsrc}

We are now able to reason about the sequence of transitions defined by our hydra battle. Instead of using the data-type \texttt{state} we study the relationship
between different configurations of the battle.

Let us define a binary relation associated with every round of the battle.
In the following definition \texttt{i} is associated with the round number (or date, if we consider a discrete time), and \texttt{a}, \texttt{b}, \texttt{c} respectively associated with the number of \texttt{h2}, \texttt{h1} and heads connected to the hydra's foot.

\begin{Coqsrc}
Inductive one_step (i: nat) :
  nat -> nat -> nat -> nat -> nat -> nat -> Prop :=
| step1 : forall a b c,  one_step i a b (S c) a b c
| step2 : forall a b ,  one_step i a (S b) 0 a b (S i)
| step3  : forall a, one_step i (S a) 0 0 a (S i) 0.
\end{Coqsrc}

The relation between \texttt{one\_step} and the rules of hydra battle is asserted by the following lemma.

\label{lemma:step-fight}

\begin{Coqsrc}
Lemma step_fight : forall i a b c a' b' c', 
   one_step i a b c a' b' c' ->
   fight standard i (hyd  a b c)  (S i) (hyd a' b' c').
\end{Coqsrc}

Next, we define ``big steps'' as the transitive closure of \texttt{one\_step},
and reachability (from the initial configuration of figure~\ref{fig:hinit} at time $0$).

\begin{Coqsrc}
 Inductive steps : nat -> nat -> nat -> nat ->
                  nat -> nat -> nat -> nat -> Prop :=
| steps1 : forall i a b c a' b' c',
    one_step i a b c a' b' c' -> steps i a b c (S i) a' b' c'
| steps_S : forall i a b c j a' b' c' k a'' b'' c'',
    steps i a b c j a' b' c' ->
    steps j a' b' c' k a'' b'' c'' ->
    steps i a b c k  a'' b'' c''.

Definition reachable (i a b c : nat) : Prop :=
  steps 3 3 0 0 i a b c.
\end{Coqsrc}

The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{fight}.

\begin{Coqsrc}
 Lemma steps_fight : forall i a b c j a' b' c', 
   steps i a b c j a' b' c' ->
   fight standard i (hyd  a b c)   j  (hyd a' b' c').
\end{Coqsrc}

Thus, any result about \texttt{steps} will be applicable to standard fights.
Using the predicate \texttt{steps} our study of the length of the considered battle
can  be decomposed into three parts:

\begin{enumerate}
\item  Characterization of regularities of some events
\item Study of the beginning of the battle
\item Computing the exact length of the battle.
\end{enumerate}

First, we prove that, if at round $i$ the hydra is equal to
(\texttt{hyd a (S b) 0}), then it will be equal to (\texttt{hyd a b 0}) at the $2(i+1)$-th round.

\begin{Coqsrc}
Lemma LS : forall c a b i,  steps i a b (S c) (i + S c) a b 0.
Proof.
  induction c.
 -   intros;  replace (i + 1) with (S i).
     + repeat constructor.
     + ring.
 -  intros; eapply  steps_S.
   +   eleft;   apply rule1.
   +   replace (i + S (S c)) with (S i + S c) by ring;  apply IHc.
Qed.
\end{Coqsrc}

\begin{Coqsrc}
Lemma doubleS_law : forall  a b i, steps i a (S b) 0 (doubleS i) a b 0.
Proof.
  intros;  eapply steps_S.
  +   eleft;   apply step2.
  +   unfold doubleS; replace (2 * S i) with (S i + S i) by ring; 
        apply LS.
Qed.
\end{Coqsrc}

\begin{Coqsrc}
Lemma reachable_S  : forall i a b, reachable i a (S b) 0 ->
                                   reachable (doubleS i) a b 0.
Proof.
  intros; right with  (1 := H); apply doubleS_law.
Qed.
\end{Coqsrc}

From now on, the lemma \texttt{reachable\_S} allows us to watch larger steps of 
the fight.

\begin{Coqsrc}
 Lemma L4 : reachable 4 2 4 0.
Proof.
  left; constructor.
Qed.
\end{Coqsrc}

\begin{Coqsrc}
Lemma L10 : reachable 10 2 3 0.
Proof.
  change 10 with (doubleS 4).
  apply reachable_S, L4.
Qed.
\end{Coqsrc}

\begin{Coqsrc}
Lemma L22 : reachable 22 2 2 0.
Proof.
  change 22 with (doubleS 10).
  apply reachable_S, L10.
Qed.
\end{Coqsrc}

\begin{Coqsrc}
Lemma L46 : reachable 46 2 1 0.
Proof.
  change 46 with (doubleS 22); apply  reachable_S, L22.
Qed.
\end{Coqsrc}

\begin{Coqsrc}
Lemma L94 : reachable 94 2 0 0.
Proof.
  change 94 with (doubleS 46); apply reachable_S, L46.
Qed.
\end{Coqsrc}

\begin{Coqsrc}
Lemma L95 : reachable 95 1 95 0.
Proof.
  eapply steps_S.
  -  eexact L94.
  -  repeat constructor.
Qed.
\end{Coqsrc}

We are now able to make giant steps in the simulation of the battle.
First, we iterate the lemma \texttt{reachable\_S}.

\begin{Coqsrc}
Lemma Bigstep : forall b i a , reachable i a b 0 ->
                               reachable (iterate doubleS b i) a 0 0.
 Proof.
  induction b.
  -  trivial.
  -  intros;  simpl;   apply reachable_S in H.
     rewrite <- iterate_comm; now apply IHb.
 Qed.
\end{Coqsrc}

Applying lemmas \texttt{BigStep} and \texttt{L95} we make a first jump.

\begin{Coqsrc}
 Definition M := (iterate doubleS 95 95).

Lemma L2_95 : reachable M 1 0 0.
Proof.
  apply Bigstep,  L95.
Qed.
\end{Coqsrc}

Figure~\ref{fig:HM}  represents the hydra at the $M$-th round.
At the $(M+1)$-th round, it will look like in fig~\ref{fig:HM-plus1}.

\begin{figure}[htb]
\centering
\begin{tikzpicture}[very thick, scale=0.5]
\node (foot) at (2,0) {$\bullet$};
\node (N1) at (2,2) {$\bullet$};
\node (N2) at (3,4) {$\Smiley[2][green]$};
\node (N3) at (1,4) {$\Smiley[2][green]$};
\draw (foot) -- (N1);
\draw (N1) to [bend right =15] (N2);
\draw (N1) to  [bend left=15](N3);
\end{tikzpicture}
\caption{\label{fig:HM}}
The state of the hydra after $M$ rounds.
% The hydra \texttt{h} of the proof that \(\omega^2\) is too small for proving Hercules' victory

\end{figure}

\begin{figure}[htb]
\centering
\begin{tikzpicture}[very thick, scale=0.5]
\node (foot) at (10,0) {$\bullet$};
\node (N1) at (0,5) {$\bullet$};
\node (N12) at (0,8) {$\Smiley[2][green]$};
\node (N2) at (2,5) {$\bullet$};
\node (N22) at (2,8) {$\Smiley[2][green]$};
\node (N3) at (4,5) {$\bullet$};
\node (N32) at (4,8) {$\Smiley[2][green]$};
\node (N4) at (6,5) {$\bullet$};
\node (N42) at (6,8) {$\Smiley[2][green]$};

\node (Ndots) at (12,8) {\Huge $\dots$};
\node (Ndots2) at (12,5) {\Huge $\dots$};

\node (N8) at (18,5) {$\bullet$};
\node (N82) at (18,8) {$\Smiley[2][green]$};
\node (N9) at (20,5) {$\bullet$};
\node (N92) at (20,8) {$\Smiley[2][green]$};

\draw (foot) -- (N1);
\draw (foot) -- (N2);
\draw (foot) -- (N3);
\draw (foot) -- (N4);
\draw (foot) -- (N8);
\draw (foot) -- (N9);
\draw (N1) to  (N12);
\draw (N2) to  (N22);
\draw (N3) to  (N32);
\draw (N4) to  (N42);
\draw (N8) to  (N82);
\draw (N9) to  (N92);
\end{tikzpicture}
\caption{\label{fig:HM-plus1}}
The state of the hydra after $M+1$ rounds (with $M+1$ heads).

\end{figure}

\begin{Coqsrc}
Lemma L2_95_S : reachable (S M) 0 (S M) 0.
Proof.
  eright.
  - apply L2_95.
  -  left; constructor 3.
Qed.
\end{Coqsrc}

Then, applying once more the lemma \texttt{BigStep}, we get the exact time when
Hercules wins!

\begin{Coqsrc}
Definition N :=   iterate doubleS (S M) (S M).

Theorem   SuperbigStep : reachable N  0 0 0 .
Proof.
  apply Bigstep, L2_95_S.
Qed.
\end{Coqsrc}

We are now able to prove formally that the considered fight is 
composed of $N$ steps.

\begin{Coqsrc}
Lemma Almost_done :
  fight standard 3 (hyd 3 0 0) N (hyd 0 0 0).
Proof. 
  apply steps_fight, SuperbigStep.
Qed.

Theorem  Done :
  fight standard 0 hinit N head.
Proof.
  eapply fight_trans.
  -   apply Almost_done.
  -  apply L_0_3.
Qed.
\end{Coqsrc}

Now, we would like to get an intuition of  how big the number $N$ is.
For that purpose, we use a minoration of the function \texttt{doubleS} by the
function (\texttt{fun n => 2 * n}).

\begin{Coqsrc}
Definition exp2 n := iterate (fun n => 2 * n) n 1.
\end{Coqsrc}
Using some facts (proven in \url{../src/html/BigBattle.html}),we get several  minorations.

\begin{Coqsrc}
Lemma minoration_0 : forall n,  2 * n <= doubleS n.

Lemma minoration_1 : forall n x, exp2 n * x <= iterate doubleS n x.

Lemma minoration_2 : exp2 95 * 95 <= M.

Lemma minoration_3 : exp2 (S M) * S M <= N.

Lemma minoration : exp2 (exp2 95 * 95) <= N.
\end{Coqsrc}

The number $N$ is greater than or  equal to $2^{2^{95}\times 95}.$ If we wrote $N$ in base $10$, $N$ would require at least $10^{30}$ digits!

\section{Reasoning about any battle}

The example we just studied shows that the termination of any battle may take a very long time. If we want to study hydra battles in general, we have to consider 
any hydra and any strategy, both for Hercules and the hydra itself. So, we first  give some definitions, generally borrowed from transition systems vocabulary (see~\cite{tel_2000} for instance).

\subsection{Termination of all battles}

The termination of all battles is naturally expressed by the predicate \texttt{well\_founded} defined in the module \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf} 
 of the Standard Library.

\index{Predicates!Termination}

\begin{Coqsrc}
Definition Termination :=  well_founded (transp _ round).
\end{Coqsrc}

% \subsection{Reachability}

% Let $b$ be an instance of \texttt{battle}.  We say that a configuration $(i,h)$ is 
% \emph{reachable} if there exists some initial hydra $h_0$ such that the transformation from $(0,h_0)$ into $(i,h)$ belongs to $b$. The following inductive definition considers also the case where $i=0$ and $h=h_0$.

% \vspace{4pt}
% \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}

% \begin{Coqsrc}
% Inductive reachable (b : Battle)
%   : nat -> Hydra ->  Prop  :=
% | reachable_0 : forall h1, reachable b 0 h1
% | reachable_S : forall i h h0,  fight b  0 h0 i h ->
%                                 reachable b i h.
% \end{Coqsrc}

% A nice property of the classes \texttt{free} and \texttt{standard} is that every configuration $(i,h)$ is reachable. It is sufficient to add $i$ heads to $h$'s foot and start the battle at time $0$.

% \vspace{4pt}
% \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}

% \begin{Coqsrc}
% Lemma reachable_free : forall i  h, reachable free i h.

% Lemma reachable_standard  : forall h i, reachable standard i h.
% \end{Coqsrc}

\subsection{Liveliness}

We say that a kind  $B$ of battles is \emph{alive} if for any configuration $(i,h)$, where $h$ is not a head, there exists a further step in class $B$.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}

\begin{Coqsrc}
Definition Alive (B : Battle) :=
  forall i h, 
     h <> head -> {h' : Hydra |  B i h h'}.
\end{Coqsrc}

The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
\url{../src/html/hydras.Hydra.Hydra_Theorems.html} show that the classes \texttt{free} and \texttt{standard} satisfy this property.

\begin{Coqsrc}
Theorem Alive_free: Alive free.

Theorem Alive_standard: Alive standard.  
\end{Coqsrc}

Both theorems are proved with the help of the  following strongly specified function:

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}

\begin{Coqsrc}
Definition  next_round_dec n :
 forall h , (h = head) + {h' : Hydra & {R1 h h'} + {R2  n h  h'}}.
\end{Coqsrc}

\section{Termination}

In this section, we are interested in proofs of termination of all battles of a  given kind, and specifically proofs by \emph{variants}.

Let $b$ be an instance of class \texttt{Battle}. A \emph{variant} for $b$ consists
in a well-founded relation $<$  on some type \texttt{A}, and a function
(also called a \emph{measure}) \texttt{m:Hydra->A} such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $b$, the inequality $m(h')<m(h)$ holds.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
\index{Coq!Techniques!Type classes}

\begin{Coqsrc}
Class Hvariant {A:Type}{Lt:relation A}(Wf: well_founded Lt)(b : Battle)
                          (m: Hydra -> A): Prop :=
 {variant_decr :forall i h h', 
     h <> head -> b i  h h' -> Lt (m h') (m h)}.
\end{Coqsrc}

\index{Exercises}

\begin{exercise}
 Prove that, if there is an instance of \texttt{Hvariant Lt wf\_Lt b m}, then there exists no infinite battle in  \texttt{b}.
\end{exercise}

\subsection{Some failed attempts}
The class \texttt{Hvariant} is parameterized by a well-founded order $<$ on some type $A$, a class of battles $B$ and a measure $m$. When $<$ or $m$ are 
badly given for a given class $B$, one may fail to find a variant.

In this section, we present some naive attempts, where we fail to build proof of termination of free battles.

\subsubsection{Using the Hydra's height as a variant}
A first plan to prove termination of all hydra battles is to use a simple
measure that maps any hydra to a natural number. For instance, let us check whether 
 the function \emph{height} could be such  a variant.

Unfortunately, that does not work.
Please consider the hydras of Fig.~\vref{fig:h-omega2-small} and
\ref{fig:h-omega-omega}. The former can be transformed into the latter in one round, but their heights are equal.

\begin{figure}[htb]
\centering
\begin{tikzpicture}[very thick, scale=0.5]
\node (foot) at (2,0) {$\bullet$};
\node (N1) at (2,2) {$\bullet$};
\node (N2) at (2,4) {$\Smiley[2][green]$};
\draw (foot) -- (N1);
\draw (N1) -- (N2) ;
\end{tikzpicture}
\caption{\label{fig:h-omega-omega}}
\end{figure}

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Omega_Small.html\#height_bad}{Hydra.Hydra.Omega\_Small}}

\begin{Coqsrc}
Lemma height_bad :  ~ Hvariant lt_wf free height.
Proof.
  intros [H];
   specialize (H 1 (hyd1 (hyd2 head head))  (hyd1 (hyd1 head)));
   apply (lt_irrefl 2), H.
  -  apply reachable_free.
  -  discriminate.
  -  exists 0; right;  R2_here 0;  left. 
Qed.
\end{Coqsrc}

\subsubsection{Using any variant defined on \texorpdfstring{nat}{nat}}
\label{omega-case}
We could imagine that our previous failure is due to the choice of \texttt{height} as a measure, and that a more complex function would work.
In fact, we can prove that 
\emph{no} instance of class \texttt{WfVariant round Peano.lt $m$} can be build, where
$m$ is \emph{any} function of type \texttt{Hydra $\arrow$ nat}.

Let us present the main steps of that proof, the script of which  is in the module ~\href{../src/html/hydras.Hydra.Omega_Small.html}{Hydra/Omega\_Small.v} \footnote{ The name of this file means ``the ordinal $\omega$ is too small for proving the termination of [free] hydra battles ''. In effect, the elements of $\omega$, considered as a set, are just the natural numbers.}.

%\subsubsection{Preliminaries}

Let us assume there is a variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
    the  termination of all hydra battles.

\begin{Coqsrc}
Section Impossibility_Proof.
 Variable m : Hydra -> nat.
 Hypothesis Hvar : Hvariant lt_wf free m.
\end{Coqsrc}

We define an injection from the type \texttt{nat} into \texttt{Hydra}.
For any natural number $i$, $\iota(i)$ is the hydra composed of a foot and
$i+1$ heads at height $1$. For instance, Fig.~\ref{fig:flower} represents the hydra $\iota(3)$.

\begin{figure}[htb]
\centering
\begin{tikzpicture}[very thick, scale=0.5]
\node (foot) at (4,0) {$\bullet$};
\node (N1) at (2,2) {$\Smiley[2][green]$};
\node (N2) at (4,2) {$\Smiley[2][green]$};
\node (N3) at (6,2) {$\Smiley[2][green]$};
\node (N4) at (8,2) {$\Smiley[2][green]$};
\draw (foot) to [bend left =25] (N1);
\draw (foot) to [bend left =15] (N2);
\draw (foot) to [bend right =15] (N3);
\draw (foot) to [bend right =25] (N4);
\end{tikzpicture}
\caption{\label{fig:flower}
The hydra $\iota(3)$}
\end{figure}

  \begin{Coqsrc}
  Let iota (i: nat) := hyd_mult head (S i).    
  \end{Coqsrc}

Let us consider now some hydra \texttt{big\_h} out of the range of the injection $\iota$ (see Fig.~\vref{fig:h-omega-omega}).

 \begin{Coqsrc}
  Let big_h := hyd1 (hyd1 head).
 \end{Coqsrc}

 Using the functions $m$ and $\iota$, we define a second hydra \texttt{small\_h}, and show
 there is a one-round battle that transforms \texttt{big\_h} into \texttt{small\_h}. Please note that,
due to the hypothesis \texttt{Hvar}, we are interested in the termination of \emph{free} battles. 
There is no problem to consider a round with \texttt{m big\_h} as replication factor.

  \begin{Coqsrc}
 Let small_h := iota (m big_h).

 Fact big_to_small : big_h -1-> small_h.
 Proof.
      exists (m big_h); right; repeat constructor.     
 Qed.
    \end{Coqsrc}

But, by hypothesis, $m$ is a variant. Hence, we infer the following inequality.

 \begin{Coqsrc}
Lemma m_lt : m small_h < m big_h.
 \end{Coqsrc}

In order to get a contradiction, it suffices to  prove the inequality
\texttt{m big\_h <= m small\_h}, i.e.,  \texttt{m big\_h <= m (iota (m big\_h))}.

More generally, we prove the following lemma:

\begin{Coqsrc}
Lemma m_ge : forall i:nat , i <= m (iota i).
\end{Coqsrc}

Intuitively, it means that, from any hydra of the form (\texttt{iota $i$}), the battle will 
take (at least) $i$ rounds. Thus the associated measure cannot be less than $i$.
Technically, we prove this lemma by Peano induction on $i$.

\begin{itemize}
\item The base case $i=0$ is trivial
\item Otherwise, let $i$ be any natural number and assume  the inequality
  $i \leq m(\iota(i))$.
  \begin{enumerate}
  \item  But the hydra $\iota(S(i))$ can be transformed in one round into
    $\iota(i)$ (by losing its righmost head, for instance)
  \item Since $m$ is a variant, we have $m(\iota(i)) < m(\iota(S(i)))$,
    hence  $i< m(\iota(S(i)))$, which implies  $S(i)\leq  m(\iota(S(i)))$.
  \end{enumerate}
\end{itemize}

 Then our proof is almost finished.

   \begin{Coqsrc}
Theorem Contradiction : False.
Proof.
 apply (Nat.lt_irrefl (m big_h));
   apply  Lt.le_lt_trans with (m small_h).
  - apply m_ge.
  - apply m_lt.
Qed.

End Impossibility_Proof.
\end{Coqsrc}

\index{Exercises}

\begin{exercise}
Prove that there exists no variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
    the  termination of all \emph{standard} battles.
\end{exercise}

\subsubsection{Lexicographic order on \texorpdfstring{nat*nat}{nat*nat}}
\label{omega2-case}

We  prove now that even the type \texttt{nat * nat}, provided with the lexicographic product of \texttt{(nat,<)} by itself is too simple for proving the termination of all hydra battles. This impossibility result will prevent us from considering measures like the following one:

  \begin{Coqsrc}
  Let m h = (height h, hsize h).
  \end{Coqsrc}

The proof we are going to develop has exactly the same structure as the previous one. Nevertheless, the proof of technical  lemmas is a little more complex, due to 
 the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
Consider for instance that there exists an infinite number of pairs between
$(1,0)$ and $(2,0)$.

\begin{remark}
  The order structure we consider in this section is also known as the ordinal
  $\omega^2$. We identify any pair $(i,j)\in \mathbb{N}\times\mathbb{N}$ with the ordinal $\omega\times i + j$. Thus the three kinds of ordinals in $\omega^2$
  are represented as follows:
  \begin{description}
  \item[null ordinal] : the pair $(0,0)$
  \item[successor ordinal] : any pair $(i,j)$ where $j>0$
  \item[limit ordinal] : any pair $(i,0)$ where $i>0$
  \end{description}
\end{remark}

The detailed  proof script is in the file \url{../src/html/hydras.Hydra.Omega2_Small.html}.

\subsubsection{Preliminaries}
Let us assume there is a variant from \texttt{Hydra} into \texttt{nat*nat} (with the
   lexicographic ordering)  for proving the   termination of all hydra battles.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}

\begin{Coqsrc}
  Section Impossibility_Proof.

  Let t := (nat * nat)%type.

  (** non-dependent lexicographic strict ordering on nat*nat *)

  Let lt2 : relation t := lexico Peano.lt Peano.lt.

  Infix "<" := lt2.

  (** reflexive closure of lt2 *)
  Let le2 := clos_refl _ lt2.
  Infix "<=" := le2.

  Variable m : Hydra -> t.

  Context (Hvar : Hvariant lt2_wf free m).
\end{Coqsrc}

Let us follow the same pattern as in Sect.~\ref{omega-case}.
First, we define an injection from type \texttt{t} into \texttt{Hydra}.
We associate with any pair $(i,j)$ the hydra with $i$ branches of length $2$ and
$j$ branches of length $1$.

%% revenir ici

\vspace{4pt}
\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}

\begin{Coqsrc}
  Let iota (p: t) := 
    node (hcons_mult (hyd1 head) (fst p)
                     (hcons_mult head (snd p) hnil)).
\end{Coqsrc}

For instance, Figure~\vref{fig:essai2} shows the hydra associated to the pair
$(3,5)$.

\begin{figure}[htb]
\centering
\begin{tikzpicture}[very thick, scale=0.4]
\node (foot) at (6,0) {$\bullet$};
\node (N1) at (1,3) {$\bullet$};
\node (N2) at (3,3) {$\bullet$};
\node (N3) at (5,3) {$\bullet$};
\node (N4) at (8,3) {$\Smiley[2][green]$};
\node (N5) at (11,3) {$\Smiley[2][green]$};
\node (N6) at (14,3) {$\Smiley[2][green]$};
\node (N7) at (17,3){$\Smiley[2][green]$};
\node (N8) at (20,3){$\Smiley[2][green]$};
\node  (N9) at (0,5) {$\Smiley[2][green]$};
\node (N10) at (2,5) {$\Smiley[2][green]$};
\node (N11) at (4,5) {$\Smiley[2][green]$};
\draw (foot) to [bend left=10] (N1);
\draw (foot) -- (N2);
\draw (foot) -- (N3);
\draw (foot) -- (N4);
\draw (foot) -- (N5);
\draw (foot) -- (N6);
\draw (foot) to [bend right=10] (N7);
\draw (foot) to [bend right=15] (N8);
\draw (N1) to [bend left=10] (N9);
\draw (N2) -- (N10);
\draw (N3) -- (N11);
\end{tikzpicture}
\caption{\label{fig:essai2}
The hydra $\iota(3,5)$}
\end{figure}

Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).

\begin{figure}[htb]
\centering
\begin{tikzpicture}[very thick, scale=0.5]
\node (foot) at (2,0) {$\bullet$};
\node (N1) at (2,2) {$\bullet$};
\node (N2) at (3,4) {$\Smiley[2][green]$};
\node (N3) at (1,4) {$\Smiley[2][green]$};
\draw (foot) -- (N1);
\draw (N1) to [bend right =15] (N2);
\draw (N1) to  [bend left=15](N3);
\end{tikzpicture}
\caption{\label{fig:h-omega2-small}}
 The hydra \texttt{big\_h}.
\end{figure}
\begin{Coqsrc}
   Let big_h := hyd1 (hyd2 head head).  
 \end{Coqsrc}

 In a second step, we build a ``smaller'' hydra.

\begin{Coqsrc}
   Let small_h := iota (m big_h).
\end{Coqsrc}

Like in Sect.~\ref{omega-case}, we prove the double inequality \texttt{m big\_h <= m small\_h < m big\_h}, which is impossible.

\subsubsection{Proof of the inequality \texttt{m small\_h < m big\_h}}

For proving the inequality  \texttt{m\_lt: m small\_h < m big\_h}, it suffices to
build a fight transforming \texttt{big\_h} into \texttt{small\_h}.

First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h}
into the pair $(i,j)$.
If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the folowing result.

\begin{Coqsrc}
 Lemma big_to_small: big_h -+-> small_h.
\end{Coqsrc}

Since $m$ is a variant, we infer the following inequality:

\begin{Coqsrc}
Corollary m_lt : m small_h < m big_h.
\end{Coqsrc}

\subsubsection{Proof of the inequality \texttt{m big\_h <= m small\_h} }

The proof of the inequality \texttt{m big\_h <= m small\_h} is quite more complex than in Sect~\ref{omega-case}.  If we consider some pair $(i,j)$, where $i>0$, there exists an infinite number of
pairs stricly less than $(i,j)$, and there exists an infinite number of battles that start from
$\iota(i,j)$. In effect, at any configuration $\iota(k,0)$, the hydra can freely chose any replication number. Intuitively, the measure of such a hydra must be large enough for taking into account
all the possible battles issued from that hydra.
Let us give more technical details.

\begin{itemize}
\item The proof of the lemma \texttt{m\_ge : forall p : t,   p <= m (iota p)} uses well-founded induction on $p$, and not structural induction on natural numbers

\item For any pair $p$, we have to distinguish between three cases, according to the value of $p$'s components.
  \begin{itemize}
  \item $p=(0,0)$
  \item $p=(i,0)$, where $i>0$\,: $p$ corresponds to a limit ordinal
  \item $p=(i,j)$, where $j>0$\,: $p$ is the successor of $(i,j-1)$.
  \end{itemize}
\end{itemize}

Before starting the proof, we have to express the notion of \emph{limit} in terms of
least upper bounds, through the following logical equivalence.

\vspace{4pt}
\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#limit_is_lub}{Hydra.Omega2\_Small.v}}.

\begin{Coqsrc}
Lemma limit_is_lub : forall i p, 
   (forall j, (i,j) < p) <-> (S i, 0) <= p.  
\end{Coqsrc}

Let us define the notion of elementary ``step'' of decreasing sequences in
\texttt{t}

\begin{Coqsrc}
Inductive step : t -> t -> Prop :=
| succ_step : forall i j,  step (i, S j) (i, j)
| limit_step : forall i j, step (S i, 0) (i, j).
\end{Coqsrc}

The following lemma establishes a correspondance between the relation
\texttt{step} and hydra fights.

\begin{Coqsrc}
Lemma step_to_fight : forall p q, step p q -> iota p -+-> iota q.
\end{Coqsrc}

\index{Maths!Transfinite induction}

Thus, starting from any inequality $q < p$ on type \texttt{t}, we can build 
by transfinite induction over \texttt{p} a fight 
that transforms the hydra $\iota(p)$ into $\iota(q)$.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}

\begin{Coqsrc}
Lemma m_ge : forall p : t,   p <= m (iota p).
Proof.
  intro p ; pattern p;
    apply  well_founded_induction with 
               (R := lt2) (1:= wf_lexico lt_wf lt_wf);
     intros (i,j) IHij (* rest of proof skipped *)
\end{Coqsrc}

\begin{Coqanswer}
  i, j : nat
  IHij : forall y : t, y < (i, j) -> y <= m (iota y)
  ============================
  (i, j) <= m (iota (i, j)) 
\end{Coqanswer}

Then we have  three cases to consider, according to the values of $i$ and $j$.
\begin{itemize}
\item If $p=(0,0)$ then obviously, $\iota(p)\geq p = (0,0)$
\item If  $p=(i+1,0)$ for some $i\in\mathbb{N}$, we
 remark  that $p$ is strictly greater than any pair $ (i, j)$, where $j$ 
is any natural number.

Applying the battle rules, for any $j$, we have $\iota(i+1,j)  {\round} \iota(i, j) $, thus $m(\iota(p)) > m(\iota(i,j)$ since  $m$ is assumed to be a variant.

Applying the induction hypothesis, we get the inequality
 $ m(\iota(i,j)) \geq (i,j)$ for any $j$.

Thus, $m(\iota(p)) > (i,j)$ for any $j$.
Applying the lemma \texttt{limit\_is\_lub}, we get  the inequality
$\iota(i+1,0)\geq (i+1,0)$

\item If $p=(i,j+1)$ with $j\in\mathbb{N}$, we have  $\iota(p)  {\round} \iota(i, j) $,
hence $m(\iota(p))> m(\iota(i,j)) \geq (i,j)$, thus $m(\iota(p))\geq (i,j+1)=p$

\end{itemize}

\subsubsection{End of the proof}

Since $<$ is a strict order (irreflexive  and transitive) on \texttt{nat*nat}, we can conclude that there is no
variant for termination on the lexicographic square of $(\mathbb{N},<)$.

\begin{enumerate}
\item From \texttt{m\_lt}, we infer the strict inequality 
\texttt{m small\_h < m big\_h}
\item from \texttt{m\_ge}, we get \texttt{m big\_h <= m (iota (m big\_h)) = m small\_h} 
\end{enumerate}

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}

\begin{Coqsrc}
Theorem Impossible : False.
Proof.
  destruct (StrictOrder_Irreflexive  (m big_h)).
  apply le2_lt2_trans with (m small_h).
  -  unfold small_h;  apply m_ge.
  -  apply m_lt. 
Qed.

End Impossibility_Proof.
\end{Coqsrc}

\index{Exercises}

\begin{exercise}
Prove that there exists no variant $m$ from \texttt{Hydra} into \texttt{nat*nat} for proving
    the  termination of all \emph{standard} battles.
\end{exercise}

\begin{remark}
In Chapter~\ref{ks-chapter}, we will prove a much more general  theorem
than the two previous results about $\mathbb{N}$ and $\mathbb{N}\times\mathbb{N}$. The proof of that general  result will share the same structure, but will 
require a lot of technical results. 
\end{remark}

\begin{exercise}

\label{sec:orgheadline63}
Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.

\begin{itemize}
\item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
  lexicographic ordering
\item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.

For instance , the sequence $\langle 4,4,2 \rangle$ is strictly greater than 
$\langle 4,3,3,3,3,3,3,2,2,2 \rangle$.
\end{itemize}

\end{exercise}

### 2020-06-23 latex/chapter-ks.tex
first commit

s]\chapter{Inside \texorpdfstring{$\epsilon_0$}{epsilon0}: The Ketonen-Solovay machinery\label{ks-chapter}}
\label{chap:ketonen}
\index{Maths!Ketonen-Solovay machinery}

\section{Introduction}
The reader may think that our proof of termination in the previous  chapter requires a lot of mathematical tools and may be too  complex. So, the question is ``is there  any  simpler proof'' ?

In their article~\cite{KP82}, Kirby and Paris show that this result cannot be proved in Peano arithmetic. Their proof uses some knowledge about model theory and non-standard models of Peano arithmetic. In this chapter, we focus on a specific class of proofs of termination of hydra battles: construction of some variant mapping the type \texttt{Hydra} into some segment of ordinals. Our proof relies only on the Calculus of Inductive Constructions and is a natural complement of the results proven in the previous chapter.

\begin{itemize}
\item There is no variant mapping the type \texttt{Hydra} into the interval $[0,\omega^2[$ (section ~\vref{omega2-case}), and a fortiori 
$[0,\omega[$

\item There is a variant that maps the type \texttt{Hydra} into the
interval $[0,\epsilon_0[$ (theorem \texttt{every\_battle\_terminates}, in section~\vref{hydra-variant}).
\end{itemize}

Thus, a very natural question is the following one:
\begin{quote}
  `` Is there  any variant from
\texttt{Hydra} into some interval $[0,\mu[$, where $\mu<\epsilon_0$, for proving the termination of all hydra battles ?''
\end{quote}

We prove in \coq{} the following results:

\begin{quote}
There is no variant for proving the termination of all hydra battles
from \texttt{Hydra} into the interval $[0..\mu[$, where
$\mu< \epsilon_0$.
The same impossibility holds even if we consider only standard battles (with the successive replication factors $0,1,2,\dots,t,t+1,\dots$).
\end{quote}

Our proofs are  constructive and require no axioms: they are  closed terms of the CIC, and are mainly composed on function definitions and proofs of properties of these functions. 
They  share much theoretical material with Kirby and Paris', although they do not use any knowledge about Peano arithmetic nor model  theory.  The combinatorial arguments we use and implement
come from 
 an article by J.~Ketonen and R.~Solovay~\cite{KS81}, already  cited in the work
 by L.~Kirby et J.~Paris on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
 Section $2$ of this article: ''A hierarchy of probably recursive functions'', contains a systematic study of \emph{canonical sequences}, which are closely related to
rounds of hydra battles.

\section{Canonical Sequences}
\label{ketonen-solovay-sect}
\index{Maths!Canonical sequences}

Canonical sequences are functions that associate an ordinal $\canonseq{\alpha}{i}$ to every ordinal $\alpha<\epsilon_0$ and positive integer $i$. They satisfy nice properties :

\index{Maths!Transfinite induction}
\begin{itemize}
\item If $\alpha\not=0$, then $\canonseq{\alpha}{i}<\alpha$. Thus canonical sequences can be used for proofs by transfinite induction or function definition by transfinite recursion
\item If $\lambda$ is a limit ordinal, then $\lambda$ is the least upper bound of the set 
$\{\canonseq{\lambda}{i}\;|\,i\in\mathbb{N}_1\}$

\item If $\beta<\alpha<\epsilon_0$, then there is a ``path'' from $\alpha$ to $\beta$, \emph{i.e.} a
sequence $\alpha_0=\alpha, \alpha_1, \dots, \alpha_n=\beta$, where for every $k<n$, there exists some $i_k$ such that $\alpha_{k+1}=\canonseq{\alpha_k}{i_k}$
\item Canonical sequences correspond tightly to rounds of hydra battles: if $\alpha\not=0$,
then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1}$ in one round with
the replication factor $i$ (Lemma \href{../src/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
\item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) fight from $\iota(\alpha)$ to $\iota(\beta)$.
\end{itemize}

\paragraph{Remark}
In~\cite{KS81}, canonical sequences are defined for any ordinal $\alpha <\epsilon_0$,
by stating that if $\alpha$ is a successor ordinal, the sequence associated with 
$\alpha$ is simply the constant sequence whose terms are equal to the predecessor of $\alpha$. Likewise, we define the canonical sequence of $0$ as the
sequence whose all terms are equal to $0$.

This convention allows us to make total the function that maps any ordinal $\alpha$ and natural number $i$ to the $i$-th item of the canonical sequence associated with $\alpha$.

First, let us recall how canonical sequences are defined in~\cite{KS81}. For efficiency's sake, we decided not to implement directly K.\&S's definitions, but to define in \gallina{} simply typed structurally recursive functions which share the abstract properties which are used in the mathematical proofs\footnote{With a small diffrence: the $0$-th term of the canonical sequence is not the same in our development as in~\cite{KS81}.}

\subsubsection{Mathematical definition of canonical sequences}

In~\cite{KS81} the definition of $\canonseq{\alpha}{i}$ is based on the following remark:
\begin{quote}
Any non-zero ordinal $\alpha$ can be decomposed in a unique way as the product
$\omega^\beta\times (\gamma+1)$.
\end{quote}

Thus the $\canonseq{\alpha}{i}$\,s are defined in terms of this decomposition:
\begin{definition}[Canonical sequences: mathematical definition]
\label{def:canonseq-math}

\end{definition}
\begin{mathframe}
  \begin{itemize}
\item Let $\lambda<\epsilon_0$ be a limit ordinal

\begin{itemize}
\item If $\lambda=\omega^{\alpha+1}\times (\beta+1)$, then 
$\canonseq{\lambda}{i}= \omega^{\alpha+1}\times\beta +  \omega^\alpha \times i$
\item If $\lambda=\omega^{\gamma}\times (\beta+1)$, where $\gamma<\lambda$ is a limit ordinal, then 
$\canonseq{\lambda}{i}=\omega^{\gamma}\times \beta + \omega^{\canonseq{\gamma}{i}}$
\end{itemize}

\item For successor ordinals, we have $\canonseq{\alpha+1}{i}= \alpha$

\item Finally, $\canonseq{0}{i}= \alpha$.
\end{itemize}
\end{mathframe}

\subsubsection{Canonical sequences in Coq}

Our definition may look more complex than the mathematical one, but
uses plain structural recursion over the type \coqsimple{T1}. Thus, tactics like
\coqsimple{cbn}, \coqsimple{simpl}, \coqsimple{compute}, etc., are applicable. For simplicity's sake, 
we used an auxiliary function \texttt{canonS} of type \texttt{nat -> T1 -> T1} such that
\texttt{canonS i $\alpha$} is equal to $\canonseq{\alpha}{i+1}$.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.Canon.html\#canonS}{Epsilon0.Canon}}

\index{Functions!canonS canonS (helper for canon)}
\index{Functions!canon @ canon (canonical sequence)}
\begin{Coqsrc}
Fixpoint canonS (i:nat) alpha :=
  match alpha with
      zero => zero
    | ocons zero 0 zero => zero
    | ocons zero (S k) zero => FS k
    | ocons gamma 0 zero =>
      match pred gamma with
          Some gamma' => ocons gamma' i zero
        | None => ocons (canonS i gamma) 0 zero
      end
    | ocons gamma (S n) zero =>
       match pred gamma with
           Some gamma' => ocons gamma n (ocons gamma' i zero)
         | None => ocons gamma n (ocons (canonS i gamma) 0 zero)
       end
    | ocons alpha n beta => ocons alpha n (canonS i beta)
  end.
\end{Coqsrc}

The following function computes $\canonseq{\alpha}{i}$, except for the case $i=0$, where it simply returns $0$\;\footnote{This restriction did not prevent us from proving all the main theorems of~\cite{KS81, KP82}. Nevertheless, in a future version of this development, we may define $\canonseq{\alpha}{0}$ exactly as 
in~\cite{KS81}. This would  be done at the cost of making proofs much more complex.}.

\begin{Coqsrc}
 Definition canon i alpha := 
   match i with 0 => zero | S j => canonS j alpha end. 
\end{Coqsrc}

For instance \coq's computing facilities allow us to verify the equalities\linebreak 
\mathcolor{$\canonseq{\omega^\omega}{3} = \omega^3$} and
\mathcolor{$\canonseq{\omega^\omega*3}{42} = \omega^\omega*2 + \omega^{42}$}.

\begin{Coqsrc}
Compute (canon 3 (omega ^ omega)).
\end{Coqsrc}

\begin{Coqanswer}
  = phi0 (FS 2) : T1
\end{Coqanswer}

\begin{Coqsrc}
Example canon3 :  canon 3 (omega ^ omega) = omega ^ 3.
Proof. reflexivity. Qed.
\end{Coqsrc}

\begin{Coqsrc}
Compute pp (canon 42 (omega ^ omega * 3)).  
\end{Coqsrc}

\begin{Coqanswer}
    = (omega ^ omega * 2 + omega ^ 42)%pT1
     : ppT1
\end{Coqanswer}

\subsection{Basic properties of canonical sequences}

We did not  try to prove that our definition really implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
~\href{../src/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}

Proving the equality $\canonseq{\alpha+1}{i}=\alpha$ is not 
as simple as suggested by the equations of definition~\ref{def:canonseq-math}\,.
Nevertheless, we could prove it by  structural induction on $\alpha$.

\begin{Coqsrc}
Lemma canonS_succ i alpha :
  nf alpha ->  canonS i (succ alpha) = alpha.
Proof.
 induction alpha.
 (* ... *)
\end{Coqsrc}

\subsubsection{Canonical sequences and the order $<$}

\index{Maths!Transfinite induction}

First, we prove by transfinite induction over $\alpha$ that $\canonseq{\alpha}{i+1}$ is an ordinal strictly less than $\alpha$ (assuming $\alpha\not=0$). This property allows us to use the function \texttt{canonS} and its derivates in function definitions by transfinite recursion.

\label{lemma:canonS_LT}
\begin{Coqsrc}
Lemma canonS_LT i alpha :
  nf alpha -> alpha <> zero -> canonS i alpha <  alpha.
\end{Coqsrc}

\subsubsection{Limit ordinals are really limits}
The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
is the limit of the sequence \showmath{\canonseq{\lambda}{i}\;(1\le i)}.

Note the use of \coq's \texttt{sig} type in the theorem's statement, which
relates the boolean function \texttt{is\_limit} defined on the \texttt{T1} data-type with a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}

\begin{Coqsrc}
Lemma canonS_limit_strong (lambda : T1) : 
     nf lambda ->
     is_limit lambda  ->
     forall beta, beta < lambda ->
                  {i:nat | beta < canonS i lambda}.

Proof.
  transfinite_induction_LT lambda.
  (* ... *)
Defined.
\end{Coqsrc}

\label{lemma:canonS-limit}

\begin{Coqsrc}
Lemma canonS_limit_lub (lambda : T1) :
  nf lambda -> is_limit lambda  ->
  strict_lub (fun i => canonS i lambda) lambda.
\end{Coqsrc}

\index{Exercises}

\begin{exercise}\label{exo:simply-typed-canonseq}
Instead of using the \texttt{sig} type, define a simply typed function that, given two ordinals $\alpha$ and $\beta$, returns a natural number $i$ such that, if $\alpha$ is a limit ordinal and $\beta<\alpha$, then $\beta< \canonseq{\alpha}{i+1}$. Of course, you will have to prove the correctness of your function.
\end{exercise}

\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0} : Paths}
\index{Maths!Accessibility inside epsilon0}

Let us consider a kind of accessibility problem inside $\epsilon_0$: given two ordinals $\alpha$ and $\beta$, where $\beta<\alpha<\epsilon_0$, find a \emph{path} consisting of a finite sequence $\gamma_0=\alpha,\dots,\gamma_l=\beta$,
where, for every $i<l$, $\gamma_i \not= 0$ and there exists some strictly positive integer $s_i$
such that $\gamma_{i+1}=\canonseq{\gamma}{s_i})$,

\subsection{Definition}

We shall denote by $\alpha\xrightarrow [s]{}\beta$
  the proposition 
``there is a path from $\alpha$ to $\beta$ with $s$ as the sequence of indices
$s_0,s_1,\dots,s_{l-1}$ ''.

For instance, we have $\omega*2 \xrightarrow[2,2,2]{}\omega$, through the 
sequence of ordinals $\omega\times 2, \omega+2,\omega+1,\omega$.

Note that, given $\alpha$ and $\beta$, where $\beta < \alpha$, the sequence $s$ which leads from $\alpha$ to $\beta$ is not unique.

Indeed, if $\alpha$ is a limit ordinal, the first element of $s$ can be any integer $i$ such that $\beta<\canonseq{\alpha}(i)$, and if $\alpha$ is a successor ordinal,
then the sequence $s$ can start with any positive integer.

For instance, we have also 
$\omega*2 \xrightarrow[3,4,5,6]{}\omega$. 
Likewise,
$\omega*2 \xrightarrow[{[3,14]}]{} 0$ and
$\omega*2 \xrightarrow[s] 0$, where $s$ is the sequence containing eight times the number $3$.

\label{path-to-definition}

In \coq{}, the notion of path can be simply defined as an inductive predicate 
parameterized by the destination $\beta$.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}

\begin{Coqsrc}
(Definition transition_S i : relation T1 :=
  fun alpha beta =>  alpha <> zero /\ beta = canonS i alpha.

Definition transition i : relation T1 :=
  match i with 0 => fun _ _ => False | S j => transition_S j end.
\end{Coqsrc}

\index{Predicates!path\_to}

\begin{Coqsrc}
Inductive path_to (beta: T1) : list nat -> T1 -> Prop :=
  path_to_1 : forall (i:nat) alpha , 
    i <> 0 ->
    transition i alpha beta ->
    path_to beta (i::nil) alpha
| path_to_cons : forall i alpha s gamma,
    i <> 0 ->
    transition i alpha gamma ->
    path_to beta  s gamma ->
    path_to beta  (i::s) alpha.
\end{Coqsrc}

\begin{remark}
In the present version of our library, we use a variant \texttt{path\_toS} of
\texttt{path\_to}, where the proposition
(\texttt{path\_toS $\beta$ $s$ $\alpha$}) is equivalent to
(\texttt{path\_to $\beta$ (shift $s$) $\alpha$}). This variant is scheduled to be deprecated.
\end{remark}

\index{Exercises}

\begin{exercise}
Write a tactic for solving goals of the form (\texttt{path\_to $\beta$ $s$ $\alpha$})
where $\alpha$, $\beta$ and $s$ are closed terms. 
You should solve automatically the following goals:

\begin{Coqsrc}
 path_to omega (2::2::2::nil) (omega * 2).

 path_to omega (3::4::5::6::nil) (omega * 2).

 path_to zero (interval 3 14) (omega * 2).

 path_to zero (repeat 3 8) (omega * 2).
\end{Coqsrc}

\end{exercise}

\subsection{Existence of a path}

\index{Maths!Transfinite induction}

By transfinite induction on $\alpha$, we prove that for any $\beta<\alpha$, 
on can build a path from $\alpha$ to $\beta$.

\begin{Coqsrc}
Lemma LT_path_to (alpha beta : T1) :
  beta < alpha -> {s : list nat | path_to beta s alpha}.
\end{Coqsrc}

\index{Exercises}

\begin{exercise}[Continuation of exercise~\vref{exo:simply-typed-canonseq}]
Define a simply typed function for computing a path from $\alpha$ to $\beta$.
\end{exercise}

% ici
\noindent 
From the lemma \texttt{canonS\_LT}~\vref{lemma:canonS_LT}, we convert any path into an inequality on ordinals (by induction on paths).

\begin{Coqsrc}
Lemma path_to_LT beta s alpha :
  path_to beta s alpha -> nf alpha -> beta < alpha.
\end{Coqsrc}

\subsection{Paths and hydra battles}
\label{KS-o2h}

In order to apply our knowledge about  ordinal numbers (less than $\epsilon_0$) to the study of hydra battles, we define an injection
from the interval $[0,\epsilon_0[$ into the type \texttt{Hydra}.

\vspace{4pt}

\emph{From Module~\href{../src/html/hydras.Hydra.O2H.html}{Hydra.O2H}}

\begin{Coqsrc}
Fixpoint iota (alpha : T1) : Hydra :=
  match alpha with
  | T1.zero => head
  | ocons gamma n beta => 
         node (hcons_mult (iota gamma) (S n) (iotas beta))
  end 
with iotas (alpha : T1) :  Hydrae :=
       match alpha with
       | T1.zero => hnil
       | ocons alpha0 n beta  => 
           hcons_mult (iota alpha0) (S n)
                                (iotas beta)
       end.
\end{Coqsrc}

For instance Fig.~\ref{fig:iota-example} shows the image by $\iota$ of the ordinal  \textcolor{black}{$\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$}

  \begin{figure}[htb]
\centering
\begin{tikzpicture}[very thick, scale=0.3]
\node (foot) at (10,0) {$\bullet$};
\node (N1) at (2,2) {$\bullet$};
\node (N2) at (10,2) {$\bullet$};
\node (N22) at (7,2) {$\bullet$};
\node (N3) at (14,2) {$\bullet$};
\node (N4) at (18,2) {$\Smiley[2][green]$};
\node (N5) at (0,4) {$\bullet$};
\node (N6) at (2,5) {$\Smiley[2][green]$};
\node (N7) at (4,6) {$\Smiley[2][green]$};
\node (N88) at (7,4) {$\bullet$};
\node (N8) at (10,4) {$\bullet$};
\node (N9) at (14,6) {$\Smiley[2][green]$};
\node (N10) at (0,8) {$\Smiley[2][green]$};
\node (N11) at (10,7) {$\Smiley[2][green]$};
\node (N111) at (7,7) {$\Smiley[2][green]$};
\draw (foot) to [bend left=10] (N1);
\draw (foot) -- (N2);
\draw (foot) -- (N22);
\draw (foot) -- (N3);
\draw (foot) -- (N4);
\draw (N1) to  (N5);
\draw (N1) to   [bend left=10] (N6);
\draw (N1) to   [bend right=20] (N7);
\draw (N2) to  (N8);
\draw (N22) to  (N88);
\draw (N8) to  (N11);
\draw (N88) to  (N111);
\draw (N3) to  (N9);
\draw (N5) to  (N10);
\end{tikzpicture}
\caption{The hydra $\iota(\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$) \label{fig:iota-example}}

\end{figure}

The following lemma (proved in ~\href{../src/html/hydras.Hydra.O2H.html}{Hydra.O2H.v}) maps  canonical sequences to rounds of hydra battles.

\label{lemma:canonS-iota}

\begin{Coqsrc}
Lemma canonS_iota i alpha :
    nf alpha -> alpha <> 0 ->
    iota alpha -1-> iota (canonS i alpha).
\end{Coqsrc}

The next step of our development is to extend this relationship to
the order $<$ on $[0,\epsilon_0[$ on one side, and hydra fights on the other side.

\begin{Coqsrc}
Lemma path_to_fight alpha s beta :
  path_to  beta  s alpha -> nf alpha ->
  iota alpha -+-> iota beta.
\end{Coqsrc}

As a corollary, we are now able to transform any inequality $\beta<\alpha<\epsilon_0$ into a (free) fight.

\begin{Coqsrc}
Lemma LT_to_fight alpha beta :
    beta < alpha ->  iota alpha -+-> iota beta.
\end{Coqsrc}

\section{A first proof of impossibility}

We have got all the tools for proving there is no variant bounded by some $\mu<\epsilon_0$ for proving the termination   of all battles. The proof we are going to show is a proof by contradiction. It  can
 be considered as a generalization of the
proofs described in  sections~\vref{omega-case} and \vref{omega2-case}.
We advise the reader to compare the three proofs step by step, lemma by lemma.

In the module \href{../src/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$. This part of the development is parameterized by some class $B$ of battles, which will be instantiated later to \texttt{free} or \texttt{standard}.

\index{Coq!Techniques!Type classes}

\begin{Coqsrc}
Class BoundedVariant  (B:Battle) :=
  {
    mu:T1 ;
    m: Hydra -> T1;
    mu_nf : nf mu;
    Hvar : Hvariant T1_wf B m;
    m_bounded : forall h, m h < mu
  }.
\end{Coqsrc}

Let us assume there exists such a variant:

\begin{Coqsrc}
Section Bounded.
  Context (B: Battle)  (Hy : BoundedVariant B).

  Hypothesis m_decrease : forall  i h h',
        round_n i h h' -> m h' < m h.
\end{Coqsrc}

\begin{remark}
  The hypothesis \texttt{m\_decrease} is false in general, but is satisfied by
the  \texttt{free} and \texttt{standard} kinds of battles. This trick allows to 
``factorize'' our proofs our proofs of impossibility.
\end{remark}

\index{Maths!Transfinite induction}

First, we prove that $m(\iota(\alpha))$ is always greater than or equal to $\alpha$, by  transfinite induction over $\alpha$.

\begin{Coqsrc}
Lemma m_ge_0 alpha:  nf alpha -> alpha <= m (iota alpha).
\end{Coqsrc}

\begin{itemize}
\item If $\alpha=0$, the inequality trivially holds
\item If $\alpha$ is the successor of  some ordinal $\beta$, the inequality $\beta \leq m(\iota(\beta))$ holds (by induction hypothesis). But the hydra $\iota(\alpha)$ is transformed in one round into 
$\iota(\beta)$, thus $m(\iota(\beta))<m(\iota(\alpha))$. Hence $\beta<m(\iota(\alpha))$, which implies $\alpha \leq m(\iota(\alpha))$
\item If $\alpha$ is a limit ordinal, then $\alpha$ is the least upper bound of the set
of all  the $\canonseq{\alpha}{i}$.  Thus, we have just to prove that $\canonseq{\alpha}{i}< m(\iota(\alpha))$ for any $i$. 
\begin{itemize}
\item Let $i$ be some natural number.
By the induction hypothesis, we have $\canonseq{\alpha}{i} \leq m(\iota(\canonseq{\alpha}{i}))$. But the hydra $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i})$ in one round, thus $m(\iota(\canonseq{\alpha}{i})) < m(\iota(\alpha))$, by our hypothesis \texttt{m\_decrease}.
\end{itemize}
\end{itemize}

Please note that the impossibility proofs of 
sections~\vref{omega-case} and \vref{omega2-case} contain a similar lemma, also called \texttt{m\_ge}.
We are now able to build a counter-example.

\begin{Coqsrc}
  Definition big_h := iota mu.
  Definition beta_h := m big_h.
  Definition small_h := iota beta_h.
\end{Coqsrc}

From Lemma \texttt{m\_ge\_0} we infer the following inequality :

\begin{Coqsrc}
    Corollary m_ge_generic : m big_h <= m small_h.
 \end{Coqsrc}

The (big) rest of the proof is dedicated to prove formally the converse inequality 
\texttt{m small\_h < m big\_h}.

\subsection{The case of free battles}
\label{sec:free-battles-case}
Let us now consider that $B$ is instantiated to \texttt{free} (which means that we are considering proofs of termination of \emph{all} battles). The following lemmas are proved in Module~\href{../src/html/hydras.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
The case $B=\texttt{standard}$ is studied in section~\vref{std-case}.

\begin{Coqsrc}
Section Impossibility_Proof.

  Context (Var : BoundedVariant free ).
  \end{Coqsrc}

\begin{enumerate}
\item The following lemma is an application of \texttt{m\_ge\_generic}, since \texttt{free}
satisfies trivially the hypothesis \texttt{m\_decrease}.

\begin{Coqsrc}
Lemma m_ge : m big_h <= m small_h.
  Proof.
    apply m_ge_generic.
   (* ... *)
\end{Coqsrc}

\item From the hypothesis \texttt{m\_bounded}, we have \texttt{m big\_h < mu}
\item By Lemma \texttt{LT\_to\_fight}, we get a (free) battle from
\texttt{big\_h = iota mu} to \texttt{small\_h = iota (m big\_h)}.

\begin{Coqsrc}
  Lemma  big_to_small : big_h  -+-> small_h.
\end{Coqsrc}
\item From the hypotheses on $m$, we infer:

\begin{Coqsrc}
Lemma m_lt : m small_h < m big_h.
\end{Coqsrc}

\item From lemmas \texttt{m\_ge} and \texttt{m\_lt}, and the irreflexivity of $<$, we get a contradiction.

  \begin{Coqsrc}
Theorem Impossibility_free : False.

End Impossibility_Proof.
\end{Coqsrc}

\end{enumerate}

We have now proved there exists no bounded variant for the class of free battles.

\begin{Coqsrc}
Check Impossibility_free.
\end{Coqsrc}

\begin{Coqanswer}
  Impossibility_free
     : BoundedVariant free -> False
\end{Coqanswer}
%%% ICI

\section{The case of standard battles}
\label{sec:standard-intro}\label{std-case}
One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
Module~\href{../src/html/hydras.Hydra.O2H.html}{Hydra.O2H}.

\begin{Coqsrc}
Lemma LT_to_round_plus alpha beta :
    beta < alpha ->  iota alpha -+-> iota beta.
\end{Coqsrc}

This lemma builds a battle out of any inequality $\beta<\alpha$. 
It is a straightforward application of \texttt{LT\_path\_to} of
Module~\href{../src/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}:

\begin{Coqsrc}
Lemma LT_path_to (alpha beta : T1) :
  beta < alpha -> {s : list nat | path_to beta s alpha}.
\end{Coqsrc}

The sequence $s$, used to build the sequence of replication factors of the battle depends on 
$\beta$, so we cannot be sure that the generated battle is a genuine standard battle.

The tool we need to use is once again in Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
$\alpha_0=\alpha,\alpha_1,\dots,\alpha_k=\beta$ where $\alpha_{j+1}$ is equal
to $\canonseq{\alpha_j}{i_j}$ for some $i_j$, 
we will consider various constraints on these sequences.
In particular, a path is called \emph{standard} if $i_{j+1} = i_j + 1$ for every $j<k$.
It  corresponds to a ``segment'' of some standard battles. 
Please note that the vocabulary on paths is ours, but all the concepts come really from~\cite{KS81}.

In \coq{}, standard paths can be defined as follows.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.KS.html}{Epsilon0.KS}}

\begin{Coqsrc}
(**  standard path from (i, alpha) to (j, beta) *)

Inductive standard_pathR(j:nat)( beta:T1):  nat -> T1 -> Prop :=
  std_1 : forall i alpha, 
       beta = canon i alpha -> j = S i ->
       standard_pathR j beta i  alpha
| std_S : forall i alpha, 
      standard_pathR j beta (S i) (canon i alpha)  ->
      standard_pathR j beta i alpha.

Definition standard_path  i alpha j beta := 
   standard_pathR j beta i alpha.
\end{Coqsrc}

In the mathematical text and figures, we shall use the notation 
$\alpha \xrightarrow[i,j]{}\beta$ for the proposition 
\texttt{standard\_path $i$ $\alpha$ $j$ $\beta$}.
In~\cite{KS81} the notation is
$\alpha \xrightarrow[i]{*}\beta$
for 
the proposition  $\exists j, i<j \wedge \alpha \xrightarrow[i,j]{} \beta$.

Our goal is now  to transform any inequality $\beta<\alpha<\epsilon_0$ into a standard path $\alpha \xrightarrow[i,j]{} \beta$ for some $i$ and $j$, then into a standard battle
from $\iota(\alpha+i)$ to $\iota(\beta)$. 
Following~\cite{KS81}, we proceed in two stages:
\begin{enumerate}
\item we simulate plain (free) paths from $\alpha$ to $\beta$ with
paths made of steps $(\gamma,\canonseq{\gamma}{n})$, \emph{with the same $n$ all along the path}
\item we simulate any such path by a standard path.
\end{enumerate}

\subsection{Paths with constant index}

First of all, paths with a constant index 
enjoy nice properties. They are defined as paths where all the $i_j$ are equal to the same natural number $i$, for some $i>0$.

Like in~\cite{KS81}, we shall use the notation $\alpha \xrightarrow[i]{} \beta$ for denoting such a path, also called an $i$-path.

\begin{Coqsrc}
Definition const_pathS i :=
    clos_trans_1n T1 (fun alpha beta => beta = canonS i alpha).

Definition const_path i alpha beta :=
  match i with
    0 => False
  | S j => const_pathS j alpha beta
end.
\end{Coqsrc}

% Paths with a given index can be effectively computed.
% Given $i$, $\alpha$ and $l$, the following function returns the ordinal $\beta$ such that there exists a path 
% $\alpha \xrightarrow [i+1] {} \beta$ of length $l$.

% \begin{Coqsrc}
% Fixpoint const_funS (i:nat)(alpha : T1)(l:nat):  T1  :=
%   match l
%   with
%   | 0 => alpha
%   | S m => const_funS i (canonS i alpha) m
%   end.
% \end{Coqsrc}

% The following computations show  applications of \texttt{constS\_fun} to the 
% ordinal $\omega^\omega$, with various values of $i$ and $l$.

% \begin{Coqsrc}
% Compute  (const_funS 2 (omega ^omega)  55).
% \end{Coqsrc}

% \begin{Coqanswer}
%   = zero
%      : T1 
% \end{Coqanswer}

% \begin{Coqsrc}
% Compute pp (const_funS 2 (omega ^omega) 15).
% \end{Coqsrc}

%   \begin{Coqanswer}
%  = (omega ^ 2 * 2)%pT1
%      : ppT1   
%   \end{Coqanswer}

% \begin{Coqsrc}
% Compute pp (const_funS 4 (omega^omega)  100).
% \end{Coqsrc}

% \begin{Coqanswer}
% = (omega ^ 4 * 4 + omega ^ 3 * 4 + omega ^ 2 + omega * 4 + 4)%pT1
%      : ppT1
% \end{Coqanswer}

A most interesting property of $i$-paths is that we can ``upgrade'' their index, as stated by K.\&S.'s Corollary 12.

\index{Maths!Transfinite induction}

\begin{Coqsrc}
Corollary Cor12 (alpha : T1) :  nf alpha ->
         forall beta i n, beta  < alpha  ->
                (i < n)%nat ->
                 const_pathS i alpha beta ->
                 const_pathS n alpha beta.
Proof.
  transfinite_induction_lt alpha.
  (* (long) proof skipped *)
\end{Coqsrc}

We shall also use a version of \texttt{Cor12} with large inequalities.

\begin{Coqsrc}
Corollary Cor12' (alpha : T1) :  nf alpha ->
         forall beta i n, (beta < alpha)%t1 ->
               (i <= n)%nat ->
               const_pathS i alpha beta ->
               const_pathS n alpha beta.
\end{Coqsrc}

\subsubsection{Sketch of proof of \texttt{Cor12}}
\index{Maths!Transfinite induction}

We prove this lemma by transfinite induction on $\alpha$.
Let us consider a path $\alpha \xrightarrow [i]{} \beta$ $(i>0)$. Its first step is
the pair $(\alpha,\canonseq{\alpha}{i})$, We have $\canonseq{\alpha}{i}<\alpha$ and
$\canonseq{\alpha}{i} \xrightarrow [i]{} \beta$. 
Let $n$ be any natural number such that $n>i$.
By the induction hypothesis, there exists a path $\canonseq{\alpha}{n} \xrightarrow[i]{} \beta$.
\begin{itemize}
\item  If $\alpha$ is a successor ordinal $\gamma+1$, then $\canonseq{\alpha}{n} =
\canonseq{\alpha}{i}=\gamma$. Thus we have a path 
$\alpha  \xrightarrow [n]{}  \gamma \xrightarrow [n]{} \beta$
\item If $\alpha$ is a limit ordinal, we apply the following theorem (numbered \texttt{2.4} in Ketonen and Solovay's article).

%   \begin{theorem}
% Let $\lambda$ be a limit ordinal, then for any pair of indices $0<i<j$, there is a path $\canonseq{\lambda}{j} \xrightarrow[1]{} \canonseq{\lambda}{i}$.    
%   \end{theorem}

  \begin{Coqsrc}
Theorem Theorem_2_4 (lambda : T1) :
   nf lambda ->
   is_limit lambda  ->
   forall i j, (i < j)%nat ->
               const_pathS 0 (canonS j lambda)
                             (canonS i lambda). 
  \end{Coqsrc}

 We build the following paths :

 \begin{enumerate}
 \item $\alpha \xrightarrow n \canonseq{\alpha}{n}$
 \item $\canonseq{\alpha}{n} \xrightarrow[1]{} \canonseq{\alpha}{i}$ (by \texttt{Theorem\_2\_4}),
\item $\canonseq{\alpha}{n} \xrightarrow[n]{} \canonseq{\alpha}{i}$ (applying the induction hypothesis to the preceding path);
\item $\canonseq{\alpha}{i} \xrightarrow[n]{} \beta$ (applying the induction hypothesis)\item $\alpha \xrightarrow[n]{} \beta$ (by composition of 1, 3, and 4).

 \end{enumerate}

\end{itemize}

\begin{remark}
 \texttt{Cor12} ``casts'' $i$-paths into $n$-paths for any $n>i$.
But the obtained $n$-path can be much longer than the original $i$-path.
The following exercise will give an idea of this increase. 
\end{remark}

\index{Exercises}
\begin{exercise}
  Prove that  the length of the $i+1$-path from
  $\omega^\omega$ to $\omega^i$ is $1 + (i+1)^{(i+1)}$, for any $i$. Note that the $i$-path from
  $\omega^\omega$ to $\omega^i$ is only one step long.
 \end{exercise}

Why is \texttt{Cor12} so useful? 
Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying collorary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
where $n$ is the maximum of the indices $i$ met in the induction.

 Lemma 1, Section 2.6 of~\cite{KS81} is naturally expressed in terms of \coq's
\verb@sig@ construct.

\label{lemma:L-2_6-1}
\index{Coq!Techniques!Sigma types}

\begin{Coqsrc}
Lemma Lemma2_6_1 (alpha : T1) :  
  nf alpha ->
  forall beta,  beta < alpha  ->
                       {n:nat | const_pathS n alpha beta}.
Proof.
  transfinite_induction alpha.
  (* rest of proof skipped *)
\end{Coqsrc}

Intuitively, lemma   \texttt{L2\_6\_1}  shows that if $\beta<\alpha<\epsilon_0$, then there exists  a battle from $\iota(\alpha)$ to $\iota(\beta)$ where the replication factor is constant, although large enough.

\subsection{Casting paths with constant index into standard paths}

%%% traduire la v.f.  (voir %%% A traduire %%%% )

The article~\cite{KS81} contains 
the following lemma, the proof of which is quite complex, which allows to simulate $i$-paths by $[i+1,j]$-paths, where $j$ is large enough.

\begin{Coqsrc}
(* Lemma 1 page 300 of [KS] *)

Lemma constant_to_standard_path 
  (alpha beta : T1) (i : nat):
  nf alpha -> const_pathS i alpha beta -> zero  < alpha ->
  {l:nat | standard_path (S i) alpha j beta}.
\end{Coqsrc}

\subsubsection{Sketch of proof of \texttt{constant\_to\_standard\_path}}

Our proof follows the proof by Ketonen and Solovay, including its organization as a sequence of lemma.  Since it is a non-trivial proof, we will comment its main steps below.

\subsubsection*{Préliminaries}

Please note that, given an ordinal $\alpha:\texttt{T1}$, and two natural numbers $i$ and $l$, there exists at most a standard path $\alpha \xrightarrow [i,i+l]{*} \beta$.
The following function computes $\beta$ from $\alpha$, $i$ and $l$.

\begin{Coqsrc}
Fixpoint standard_gnaw (i:nat)(alpha : T1)(l:nat):  T1  :=
  match l with
  | 0 => alpha
  | S m => standard_gnaw (S i) (canon i alpha) m
  end.
\end{Coqsrc}

\begin{Coqsrc}
  Compute standard_fun 2 omega 15.
(*   = zero
     : T1 *)
Compute pp (standard_gnaw 2 (omega^omega)  10).
(*
= (omega + 7)%pT1
     : ppT1
*)
Compute pp (standard_gnaw 4 (omega^omega)  100).
(*
 = (omega ^ 3 * 4 + omega ^ 2 * 5 + omega * 3 + 39)%pT1
     : ppT1 *)
\end{Coqsrc}

\index{Maths!Transfinite induction}

By transfinite induction over  $\alpha$, we prove that the ordinal $0$ is reachable from any ordinal $\alpha<\epsilon_0$ by some standard path.

\begin{Coqsrc}
Lemma standard_path_to_zero :
  forall  alpha i, nf alpha ->
                   {j: nat | standard_path (S i) alpha j zero}.
\end{Coqsrc}

\paragraph*{}
Let us consider two ordinals  $\beta<\alpha<\epsilon_0$.  Let $p$  be some $(n+1)$-path from $\alpha$ to $\beta$.

\begin{Coqsrc}
 Section Constant_to_standard_Proof.

  Variables (alpha beta: T1) (n : nat).
  Hypotheses (Halpha: nf alpha) (Hpos : zero <  beta)
             (p : const_pathS n alpha  beta).
\end{Coqsrc}

Applying \texttt{standard\_path\_to\_zero}, $0$ is reachable from $\alpha$ by some standard path  (see figure~\vref{fig:belle-preuve-1}).

\begin{figure}[h]
  \centering

\begin{tikzpicture}[very thick, scale=0.25]
\node (alpha) at (0,0) {$\alpha$};
    \node (beta) at (32, 0){$\beta$};

  \draw[->, very thick,blue] (alpha)-- node [below]{$n+1$} node [above] {$+$} (beta);

  \node (alpha1) at (5,5) {};
  \node (alpha2) at (13,5) {};
    \node (alpha3) at (20,5) {};
  \node (alphalast) at (35,5) {};
  \node (zero) at (45,0) {$0$};
  \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate=40]{$n+1$}  (alpha1);
  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
   \draw [->, dashed,very thick,blue] (alpha2)-- node [below]{$n+3$}  (alpha3);

  \node (dots) at (24,5) {$\dots$};
  \draw [->, dashed, very thick,blue] (alphalast)-- node [below, rotate=-26]{$n+p+1$}  (zero);

\end{tikzpicture}
\caption{A nice proof (1)}
  \label{fig:belle-preuve-1}
\end{figure}

\paragraph*{}

Since comparison on \texttt{T1} is decidable, one can compute the last step $\gamma$ of the standard path from $(\alpha,n+1)$  such that $\beta\leq \gamma$.
Let $l$ be the length of the path from $\alpha$ to $\gamma$.

% We use for that purpose a function that computes the first index $i$ where $n\leq i <p$ and $P(i+1)$ is
% \texttt{false}, assuming that $P(n)=\texttt{true}$ and $P(p)=\texttt{false}$.

% \begin{Coqsrc}
% first_toggle  (P : nat -> bool) (n p : nat) :
% (n < p)%nat -> P n = true -> P p = false ->
% {l : nat | (n <= l)%nat /\ (l < p)%nat /\
%      (forall i : nat, (n <= i)%nat -> (i <= l)%nat -> P i = true) /\
%       P (S l) = false}
% \end{Coqsrc}

% \begin{Coqsrc}
% Let P (i:nat) := le_b beta (standard_fun (S n) alpha i).

% Remark Rem2 : {t: nat | standard_fun (S n) alpha t < beta}.

% Let t := proj1_sig Rem2.
%  Remark Rem04 : P 0.

% Remark Rem05 : P t = false.

% Remark Rem06 : (0 <  t)%nat.

%  Let l_def :=  first_toggle P 0 t  Rem06 Rem04 Rem05.

% Let l := proj1_sig Rem07.

% Let gamma := standard_fun (S n) alpha l.
% \end{Coqsrc}

This step of the proof is illustrated in figure~\vref{fig:belle-preuve-2}.

\begin{figure}[h]
  \centering

\begin{tikzpicture}[very thick, scale=0.25]
\node (alpha) at (0,0) {$\alpha$};
    \node (beta) at (32, 0){$\beta$};

  \node (alpha1) at (5,5) {};
  \node (alpha2) at (13,5) {};
  \node (dots) at (17,5) {$\ldots$};
    \node (alpha3) at (20,5) {};
    \node (gamma) at (24,0) {$\gamma$};
    \node (delta) at (38,0) {$\delta$};
    \draw [->, dashed,very thick,blue] (alpha)-- node [below,rotate=35]{$n+1$}  (alpha1);
  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
   \draw [->, dashed,very thick,blue] (alpha3)-- node [below,rotate = -48]{\tiny $n+l$}  (gamma);
   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{$n+l+1$} (delta);
   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{$n+1$} node [above] {$+$} (beta);
   \draw[thick] (alpha)--  (gamma);
   \draw[thick] (gamma)--  node [above] {$\geq$} (beta);
    \draw[thick] (beta)--  node [above] {$>$} (delta);
\end{tikzpicture}

\caption{A nice proof (2)}
  \label{fig:belle-preuve-2}
\end{figure}

\paragraph*{}

\begin{itemize}
\item If $\beta=\gamma$, its OK! We have got a standard path
from  
$\alpha$ to $\beta$ with successive indices  $n+1, n+2, \dots, n+l+1$

\item Otherwise,  $\beta < \gamma$.  Let us consider  $\delta=\canonseq{\gamma}{n+l+2}$.
By applying several times lemma \texttt{Cor12},  one converts  every path of Fig~\ref{fig:belle-preuve-2} into
 a $n+l+1$-path  (see figure~\ref{fig:belle-preuve-3}).

But $\gamma$ is on the $n+l+1$-path from $\alpha$ to $\beta$.
As shown by figure~\vref{fig:fin-belle-preuve}), the ordinal $\delta$, reachable from
$\gamma$ in one single step,  must be greater than or equal to $\beta$, which contradicts our  hypothesis $\beta < \gamma$.

\end{itemize}
 The only remaining case is  $\beta=\gamma$, thus we have got a standard path
from $\alpha$ to $\beta$.

\begin{Coqsrc}
 Lemma constant_to_standard_0 : 
    {l : nat | standard_fun (S n) alpha l = beta}.

End Constant_to_standard_Proof.

Lemma constant_to_standard_path 
  (alpha beta : T1) (i : nat):
  nf alpha -> const_pathS i alpha beta -> zero  < alpha ->
  {j:nat | standard_path (S i) alpha j beta}.
\end{Coqsrc}

\begin{figure}[h]
  \centering

\begin{tikzpicture}[very thick, scale=0.25]
\node (alpha) at (0,0) {$\alpha$};
    \node (beta) at (32, 0){$\beta$};
    \node (alpha1) at (5,5) {};
  \node (alpha2) at (13,5) {};
  \node (dots) at (17,5) {$\ldots$};
    \node (alpha3) at (20,5) {};
    \node (gamma) at (24,0) {$\gamma$};
    \node (delta) at (38,0) {$\delta$};
     \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -48]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\tiny $n+l+1$} node [above]{\color{red} $1$} (delta);
   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
    \draw[thick] (gamma)--   node [above]{\color{red} $>$}(beta);
   \draw[thick] (alpha)--  (gamma);

    \draw[thick] (beta)--  node [above] {$>$} (delta);

\end{tikzpicture}

\caption{A nice proof (3)}
  \label{fig:belle-preuve-3}
\end{figure}

\begin{figure}[h]
  \centering
\begin{tikzpicture}[very thick, scale=0.25]
\node (alpha) at (0,0) {$\alpha$};
    \node (beta) at (32, 0){$\beta$};

  \node (alpha1) at (5,5) {};
  \node (alpha2) at (13,5) {};
  \node (dots) at (15,5) {$\ldots$};
    \node (alpha3) at (18,5) {};
    \node (gamma) at (24,0) {$\gamma$};
    \node (delta) at (42,0) {$\delta$};
    \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -36]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\small $n+l+1$} node [above]{\color{red} $1$} (delta);
   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
   \draw[thick] (alpha)--  (gamma);

    \draw[thick] (gamma)--  node [below]{\tiny $n+l+1$} node [above]{\color{red} $+$}(beta);
    \draw[thick] (beta)--  node [above] {$>$} (delta);

\end{tikzpicture}

\caption{A nice proof (4)}
  \label{fig:fin-belle-preuve}
\end{figure}

Applying \texttt{Lemma2\_6\_1} and \texttt{constant\_to\_standard\_path}, we get the following corollary.

\begin{Coqsrc}
Corollary  LT_to_standard_path 
      (alpha beta : T1) :
  beta   < alpha ->
  {n : nat & {j:nat | standard_path (S n) alpha j beta}}.
\end{Coqsrc}

\subsection{Back to hydras}
\label{sec:standard-battles-cases}
We are now able to complete our proof that there exists no bounded variant for proving the termination of standard hydra battles. The proof we are going to comment can
be consulted in the module 
\url{../src/html/hydras.Hydra.Epsilon0_Needed_Std.html}.
Please note that it has the same global structure as in section\ref{sec:free-battles-case} 
% ICI !
Applying the  lemmas  \texttt{Lemma2\_6\_1} of the module 
\href{../src/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
{\texttt{Lemma2\_6\_1}}   and 
\href{../src/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
{\texttt{constant\_to\_standard\_path}},
we can convert any inequality $\beta<\alpha<\epsilon_0$ into a standard path from
$\alpha$ to  $\beta$, then into a fragment of a standard battle from 
$\iota(\alpha)$ to $\iota(\beta)$.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Epsilon0_Needed_Std.html\#LT_to_standard_battle}{Hydra.Epsilon0\_Needed\_Std}}

\begin{Coqsrc}
Lemma LT_to_standard_battle :
    forall alpha beta,
      beta < alpha ->
      exists n i,  fight standard  n (iota alpha) i (iota beta).
\end{Coqsrc}

Next, please consider the following context:

\begin{Coqsrc}
Section Impossibility_Proof.

  Context (Var : BoundedVariant standard).
 \end{Coqsrc}

In the same way as for free battles, we import a large inequality from 
the module \url{../src/html/hydras.Hydra.Epsilon0_Needed_Generic.html} .

\begin{Coqsrc}
 Lemma m_ge : m big_h <= m small_h.
\end{Coqsrc}

\paragraph*{} If remains to prove the following strict inequality, in order to have a contradiction.

\begin{Coqsrc}
Lemma m_lt : m small_h  < m big_h.
\end{Coqsrc}

\paragraph*{Proof:} Let us recall that $\texttt{big\_h} = \iota(\mu)$
 and $\texttt{small\_h} = \iota (m (\texttt{big\_h}))$.

Since $m(\texttt{big\_h })< \mu$, there exists a standard path from $\mu$ to
$m(\texttt{big\_h})$, hence a   standard battle from $\iota(\mu)$  to
$\iota(m(\texttt{big\_h}))$,  i.e. from \texttt{big\_h} to \texttt{small\_h}.

Since $m$ is assumed to be a variant for standard battles, we get the inequality  $m(\texttt{small\_h}) < m(\texttt{big\_h})$.

\subsection{Remarks}

We are grateful to 
 J. Ketonen and R. Solovay  for the high quality of their explanations and proof details.
Our proof follows tightly the sequence of lemmas in their article, with a focus on 
constructive aspects.
Roughly steaking, our implementation \emph{builds}, out of a hypothetic 
  variant $m$, bounded by some ordinal $\mu<\epsilon_0$, a hydra \texttt{big\_h} which verifies the impossible inequality  $m(\texttt{big\_h})< m(\texttt{big\_h})$.

On may ask whether the preceding results are not too restrictive, since they 
refer to a particular data type \texttt{T1}.
In fact, our representation of ordinals strictly less than 
 $\epsilon_0$ is faithful to their mathematical definition, at least 
Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
(please see also \href{../src/html/hydras.Schutte.Injection_E0.html}{the module \texttt{Ordinals.Schutte.Injection\_E0}}).

Thus, we can infer that our theorems can be applied to any well order.

\index{Projects}
\begin{project}
Study a possible modification of the definition of a variant  (for  standard battles).

\begin{itemize}
\item The variant is assumed to be strictly decreasing \emph{on configurations 
reachable from some initial configuration where the replication factor is equal to $0$}
\item The variant may depend on the number of the current round.
\end{itemize}

In other words, its type should be \texttt{nat -> Hydra -> T1}, and it must 
verify the inequality $m\, (S\,i)\, h' < m\,i\, h$ whenever the configuration 
$(i,h)$ is reachable from some initial configuration $(0,h_0)$
and \texttt{h} is transformed into \texttt{h'} in the considered round.

Can we still prove the theorems of section~\ref{std-case} with this new definition?

\end{project}

### 2020-06-23 latex/chapter-large-sets.tex
first commit

\chapter{Large sets and rapidly growing functions}\label{chap:alpha-large}

\section{Introduction}

In this chapter, we try to feel how long a standard battle can be.
To be precise, for any ordinal $\alpha<\epsilon0$ and any positive integer $k$,
we give a minoration of the number of steps of a standard fight which
starts with the hydra $\iota(\alpha)$ and the replication factor $k$.

We express this number in terms of the Hardy hierarchy of fast-growing 
functions~\cite{BW85, Wainer1970, KS81, Promel2013}.
 From the \coq{} user's point of view, such  functions are  very 
attractive:  they are defined as functions  in \gallina{}, and we can apply rhem \emph{in theory}, but they are so complex that you will never be able to look at the result of the computation.
 Thus, our knowledge on these functions must rely on \emph{proofs}. In our development, we use often the rewriting rules generated by \coq's \texttt{Equations} plug-in.

\section{Large  sequences}

\begin{definition}
Let $0<\alpha<\epsilon_0$ be any ordinal, and $s$ a sequence of strictly positive natural numbers. We say that $s$ is an  \emph{$\alpha$-large} 
sequence  if $s$ determines a path to the ordinal $0$.
We say that $s$ is \emph{minimally $\alpha$-large} (abbreviated in
\emph{$\alpha$-mlarge} if $s$ if $s$ is $\alpha$-large 
 and every strict prefix of $s$ leads to a non-zero ordinal.

\index{Maths!Large sequences}
\index{Maths!Minimal large sequences}

\end{definition}

\begin{remark}
  Ketonen and Solovay~\cite{KS81} consider  large finite \emph{sets} of natural numbers,  but they are mainly used as sequences. Thus, we chosed to represent them explicitely as (sorted) lists. 
\end{remark}

\vspace{4pt}

\noindent
\emph{From Module~ \href{../src/html/hydras.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}

\index{Predicates!mlarge @ mlarge (minimal large sequences)}
\begin{Coqsrc}
Definition mlarge alpha s := path_to zero s alpha.
\end{Coqsrc}

Let us consider two integers $k$ and $l$, such that $0<k<l$. In order to check whether the interval $[k,l]$ is minimally large for $\alpha$, it is enough to
follow from $\alpha$ the path associated with the interval $[k,l[$ and verify that the last ordinal we obtain is equal to $1$.

This verification can be done by a simple computation. First, we define a function for following a path, by successively computing canonical sequences.

\noindent
\emph{From Module~\href{../src/html/hydras.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}

\begin{Coqsrc}
Fixpoint gnaw (alpha : T1) (s: list nat) :=
  match s with
    | nil => alpha
    | (0::s') => gnaw  alpha s'
    | (S i :: s')  =>  gnaw (canonS i alpha) s'
  end.
\end{Coqsrc}

For instance the interval $[6,70]$ leads $\omega^2$ to $\omega\times 2 + 56$. Thus this interval is not $\omega^2$-mlarge.

\begin{Coqsrc}
Compute pp (gnaw (omega * omega) (interval 6 70)).
\end{Coqsrc}

\begin{Coqanswer}
 = (omega * 2 + 56)%pT1
     : ppT1
\end{Coqanswer}

Let us try another computation.

\begin{Coqsrc}
Compute (gnaw (omega * omega) (interval 6 700)).
\end{Coqsrc}

\begin{Coqanswer}
 = zero : T1
\end{Coqanswer}

We may say that the interval $[6,700]$ is $\omega^2$-large, since it leads to $0$, but nothing assures us that the condition of minimality is satisfied.

The following lemma relates minimal largeness with the function 
\texttt{gnaw}.

\begin{Coqsrc}
Lemma mlarge_iff alpha x (s:list nat) :
  s <> nil -> ~ In 0 (x::s) ->
  mlarge alpha (x::s) <-> gnaw alpha (but_last x s) = one.
 \end{Coqsrc}

For instance, we can verify that the interval $[6,509]$ is $\omega^2$-mlarge.

 \noindent
\emph{From Module~\href{../src/html/hydras.Epsilon0.Large_Sets_Examples.html}{Epsilon0.Large\_Sets\_Examples}}
\begin{Coqsrc}
Example Ex1 : mlarge (omega * omega) (interval 6 510).
\end{Coqsrc}

Now, consider any natural number $k>0$. We would like to compute
a number $l$ such that the interval $[k,l]$ is $\alpha$-mlarge. So, 
the standard battle starting with $\iota(\alpha)$ and the replication factor $k$ will end after $(l-k+1)$ steps.

First, we notice that this  number $l$ exists, thanks to the following theorems:

\begin{Coqsrc}
canonS_LT:
  forall (i : nat) (alpha : T1),
  nf alpha -> alpha <> zero -> canonS i alpha < alpha.

T1_wf: well_founded LT.
\end{Coqsrc}

Moreover, it is unique:

\begin{Coqsrc}
Lemma mlarge_unicity alpha k l l' : 
  mlarge alpha (interval (S k) l) ->
  mlarge alpha (interval (S k) l') ->
  l = l'.
\end{Coqsrc}

Thus, it seems obvious that there should be a function, parameterized by $\alpha$ which associates to any  strictly positive integer $k$ the number $l$ such that
the interval $[k,l]$ is $\alpha$-large. It would be fine to write in \gallina{} a definition like this:

\begin{Coqbad}
Function L  (alpha: E0) (i:nat) :  nat := ...
\end{Coqbad}

But we do not know how to fill the dots yet \dots{}   In the next section, we will 
use \coq{} to reason  about the \emph{specification} of \texttt{L},
prove properties of any function which satisfies this specification, and finally define a correct function with the help of the \texttt{coq-equations} plug-in.

\section{Looking for a function to define}

Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider the functions which associate to each stritly positive integer $k$ the number $l$, where 
the interval $[k,l-1]$\;\footnote{also denoted by $[k,l[$}  is $\alpha$-mlarge.

\begin{remark}
The upper bound of the considered interval has been chosen to be  $l-1$ and not $l$, in order to simplify some statements and proofs in composition lemmas associated to ordinals of the form $\alpha\times i$ and 
$\omega^\alpha\times i + \beta$.
In order to consider any ordinal below $\epsilon_0$, we consider a special case for $\alpha=0$.
\end{remark}

\subsection{Formal specification}

Our specification of the function \texttt{L} is as follows:

\begin{Coqsrc}
Inductive L_spec : T1 -> (nat -> nat) -> Prop :=
  L_spec0 :
    forall f, (forall k, f k = k) ->  L_spec zero f
| L_spec1 : forall alpha f,
    alpha <> zero ->
    (forall k, mlarge alpha (interval (S k) (Nat.pred (f (S k))))) ->
    L_spec alpha f.
\end{Coqsrc}

Note that, for $\alpha\not=0$, the value of $f(0)$ is not specified.
Nevertheless, the restriction of $f$ to the set of strictly positive integers is unique (up to extensionnality).

\begin{Coqsrc}
Lemma L_spec_unicity alpha f g :
  L_spec alpha f -> L_spec alpha g ->
  forall k, f (S k) = g (S k).
\end{Coqsrc}

\subsection{Properties derived from \texttt{L\_spec}}

Let us now prove properties of any function $f$ which satisfies 
\texttt{L\_spec}. We are looking for properties which could be used for writing \emph{equations} and prove the correctness of the function generated by the \texttt{coq-equations} plugin. Moreover, they will give us some examples of
$L_\alpha$ for small values of $\alpha$.

Our exploration of the $L_\alpha$s will follow the usual scheme : transfinite induction, and proof-by-cases : zero, successors and limit ordinals.

\index{Maths!Transfinite induction}

\subsubsection{The  ordinal zero}

The base case is directly a consequence of the specification.

\begin{Coqsrc}
Lemma L_zero_inv f : L_spec zero f -> forall k, f (S k) = S k.
\end{Coqsrc}

\subsubsection{Successor ordinals}

Let $\beta$ be some ordinal, and assume the arithmetic function $f$ satisfies 
the specification $\texttt{L\_spec}\;\beta$.  Let $k$ be any natural number.
Any path from $\texttt{succ}\,\beta$ to $0$ starting at $k+1$ can be decomposed into a first step from $\texttt{succ}\,\beta$ to $\beta$, then a path from
$\beta$ at $k+2$ to $0$. 
By hypothesis the interval $[k+2, f(k+2)-1]$ is $\beta$-mlarge.

If we assume the inequality $k+1\leq f(k+2)$, we infer that the interval 
$[k+1, f(k+2)]$ is $\beta+1$-mlarge. This new hypothesis allows us 
to apply the following lemma 
(from~\href{../src/html/hydras.Prelude.MoreLists.html\#interval_app}{Prelude.MoreLists}).

\begin{Coqsrc}
Lemma interval_app (i j k:nat):
  (i <= j)%nat -> (j <= k)%nat ->
  interval i k = interval i j ++ interval (S j) k.
\end{Coqsrc}

For ensuring this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <= f}, or \texttt{(fun\_le S f)} (defined 
in~\href{../src/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}).

\begin{Coqsrc}
Definition fun_le f g  := forall n:nat,  f n <=  g n.
\end{Coqsrc}

It looks also natural to show that the functions we consider are strictly monotonous. The section on successor ordinals has thus the following structure.

\begin{Coqsrc}
Section succ.
   Variables (beta : T1) (f : nat -> nat).

   Hypotheses (Hbeta : nf beta)
              (f_mono : strict_mono f)
              (f_Sle : S <<= f)
              (f_ok : L_spec beta f).

   Definition L_succ := fun k => f (S k).

   Lemma L_succ_mono : strict_mono L_succ.

   Lemma L_succ_Sle : S <<= L_succ.

   Lemma L_succ_ok : L_spec (succ beta) L_succ.

End succ.

\end{Coqsrc}

\subsubsection{Limit ordinals}

Let $\lambda<\epsilon_0$ be any limit ordinal. In a similar way as for successors, we decompose any path from $\lambda$ (at $k$) into a step to
$\canonseq{\lambda}{k}$, then to $0$. In the following section, we assume that there exists à correct function for $\canonseq{\lambda}{k}$, \emph{for any strictly positive $k$}.

\begin{Coqsrc}
Section lim.
  Variables (lambda : T1)
            (Hnf : nf lambda)
            (Hlim : is_limit lambda)
            (f : nat -> nat -> nat)
            (H : forall k, L_spec (canonS k lambda) (f (S k))).

  Let L_lim k := f k (S k).

  Lemma L_lim_ok : L_spec lambda L_lim.

End lim.
\end{Coqsrc}

\subsection{First results}

Applying the previous lemmas on successors and limit ordinals, 
we get several correct implementations of \texttt{(L\_spec $\alpha$)} for small values of $\alpha$.

\subsubsection{Finite ordinals}

By iterating the functional \texttt{L\_succ}, we get a realization of
\texttt{(L\_spec (fin $i$))} for any natural number $i$.

\begin{Coqsrc}
Definition L_fin i := fun k => (i + k)%nat.

Lemma L_fin_ok i : L_spec (fin i) (L_fin i).
\end{Coqsrc}

\subsubsection{The first limit ordinal \texorpdfstring{$\omega$}{omega}}

The lemmas \texttt{L\_fin\_ok} and \texttt{L\_lim\_ok}   allow us to get 
by diagonalization a correct implementation for 
\texttt{L\_spec omega}.

\begin{Coqsrc}
Definition L_omega k := S (2 * k)%nat.

Lemma L_omega_ok : L_spec omega L_omega.
\end{Coqsrc}

\subsubsection{Towards  \texorpdfstring{$\omega^2$}{omega*omega}}

We would like to get exact formulas for the ordinal $\omega^2$, a.k.a.
$\varphi_0(2)$. This ordinal is the limit of the sequence $\omega\times i\;(i \in \mathbb{N}$. Thus, we have to study ordinals of this form, then use 
our lemma on limits.

The following lemma establishes a path from $\omega\times ( i+1)$ to
$\omega \times i$.

\begin{Coqsrc}
Lemma path_to_omega_mult (i k:nat) :
  path_to (omega * i)
                (interval (S k) (2 * (S k)))
                (omega * (S i)).
\end{Coqsrc}

Let us consider a path from  $\omega\times(i+1)$ to $0$ starting at $k+1$.
A first ``big step'' will lead to $\omega\times i$ at $2(k+1)$. If $i>0$, the
next jump leads to $\omega\times(i-1)$ at $2(2(k+1))+1$, etc.

The functional \texttt{iterate}, similar to Standard Library's \texttt{Nat.iter},
is defined and studied in~\href{../src/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates})

\begin{Coqsrc}
Fixpoint iterate {A:Type}(f : A -> A) (n: nat)(x:A) :=
  match n with
  | 0 => x
  | S p => f (iterate  f p x)
  end.
\end{Coqsrc}

The following lemma expresses the length of the mlarge sequences associated with finite multiples of $\omega$.

\begin{Coqsrc}
Lemma omega_mult_mlarge_0 i  : forall k,
    mlarge  (omega * (S i))
            (interval (S k)
                      (Nat.pred (iterate (fun p =>  S (2 * p)%nat)
                                         (S i)
                                         (S k)))).
\end{Coqsrc}

Thus, we infer the following result:

\emph{From Module~ \href{../src/html/hydras.Epsilon0.Large_Sets.html\#L_omega_mult}{Epsilon0.Large\_Sets}}

\begin{Coqsrc}
Definition L_omega_mult i (x:nat) :=  iterate L_omega i x.

Lemma L_omega_mult_ok (i: nat) :  L_spec (omega * i) (L_omega_mult i).
\end{Coqsrc}

For instance, let us consider the ordinal $\omega\times 8$, and a sequence 
starting at $k=5$.

\begin{Coqsrc}
Compute L_omega_mult 8 5.
\end{Coqsrc}

\begin{Coqanswer}
= 1535
     : nat
\end{Coqanswer}

More generally, we prove the equality $L_{\omega\times i}(k)=2^i\times(k+1)-1$.

\begin{Coqsrc}
Lemma L_omega_mult_eqn (i : nat) :
  forall (k : nat),  (0 < k)%nat  ->
                     L_omega_mult i k = (exp2 i * S k - 1)%nat.
\end{Coqsrc}

By diagonalization, we obtain a simple formula for $L_{\omega^2}$.

\begin{Coqsrc}
Definition L_omega_square k := iterate (fun z => S (2 * z)%nat)
                                        k
                                        (S k).

Lemma L_omega_square_eqn k :
  (0 < k)%nat ->
  L_omega_square k = (exp2 k * (k + 2) - 1)%nat.

Lemma L_omega_square_ok: L_spec (omega * omega) 
          L_omega_square.

Compute L_omega_square 8.
\end{Coqsrc}

\begin{Coqanswer}
 = 2559
     : nat
\end{Coqanswer}

%%%% ICI

\subsubsection{Going further}
Let us consider a last example, ``computing'' $L_{\omega^3}$.
Since the canonical sequence associated with this ordinal is composed of the
$\omega^2\times i\;(i\in\mathbb{N}_1)$, we have to study this sequence.

To this end, we prove a generic lemma, which expresses $L_{\omega^\alpha\times i}$ in terms of $L_{\omega^\alpha}$. Note that in this correction lemma, we assume that the fonction associated with $\alpha$ is stritly monotonous and 
greater or equal than the successor function.

\begin{Coqsrc}
Section phi0_mult.
 Variables (alpha : T1) (f : nat -> nat).
 Hypotheses (Halpha : nf alpha)
            (f_mono : strict_mono f)
            (f_Sle : S <<= f)
            (f_ok : L_spec (phi0 alpha) f).

 Definition L_phi0_mult i := iterate f i.

Lemma L_phi0_mult_ok i: 
  L_spec (ocons alpha i zero)  (L_phi0_mult (S i)).

 Lemma L_phi0_mult_smono i: strict_mono (L_phi0_mult i).

 Lemma L_phi0_mult_Sle i: S <<=  L_phi0_mult (S i).

End phi0_mult.
\end{Coqsrc}

Let us look 
at the ordinal $\omega^2\times i$, using \texttt{L\_phi0\_mult}

\begin{Coqsrc}
 Definition L_omega_square_times i :=  iterate L_omega_square i.

 Lemma L_omega_square_times_ok i : 
    L_spec (ocons 2 i zero) (L_omega_square_times (S i)).
 Proof.
  apply L_phi0_mult_ok.
  -  auto with T1.
  -  apply L_omega_square_Sle.
  -  apply L_omega_square_ok.
 Qed.
\end{Coqsrc}

We are now ready to get an exact formula for $L_{\omega^3}$. 
\begin{Coqsrc}
Definition L_omega_cube  := L_lim  L_omega_square_times .

Lemma L_omega_cube_ok : L_spec (phi0 3) L_omega_cube.
\end{Coqsrc}

The function  $L_{\omega^3}$  is just obtained by diagonalization upon $L_{\omega^2\times i}$.

\begin{Coqsrc}
Lemma L_omega_cube_eqn i : L_omega_cube i = L_omega_square_times i (S i).
Proof.   reflexivity. Qed.
\end{Coqsrc}

Thus, for instance, $L_{\omega^3}(3)=L_{\omega^2\times 4}(3)$.
So, we obtain an exact expression of this number.

\begin{Coqsrc}
Lemma L_omega_cube_3_eq:
   let N := exp2 95 in
   let P := (N * 97 - 1)%nat in
    L_omega_cube 3  =  (exp2 P * (P + 2) - 1)%nat.
\end{Coqsrc}

This number is quite big. Using \texttt{Ocaml}'s \texttt{float} arithmetic,
we can [under-]approximate it by $2^{3.8\times10^{30}}\times 3.8\times{10^{30}}$.

\begin{Coqsrc}
# let exp2 x = 2.0 ** x;;

val exp2 : float -> float = <fun>
#   exp2 95.0 *. 97.0 -. 1.0;;
- : float = 3.84256588194182037e+30
# let n = exp2 95.0 ;;
# let p = n *. 97.0 -. 1.0;;
val p : float = 3.84256588194182037e+30

Estimation :
2 ** (3.84 e+30) * 3.84 e+30.
\end{Coqsrc}

\section{Using \texttt{Equations}}
\label{sect:equations}

Note that we did not define any function $L_\alpha$ \emph{for any $\alpha<\epsilon_0$} yet. We have got no more than a collection of proved realizations of $\texttt{L\_spec}\;\alpha$ for several values of $\alpha$.

\index{Coq!Plug-ins!Equations}

Using the \texttt{coq-equations} plug-in by 
M. Sozeau~\cite{sozeau:hal-01671777}, we will now define a function \texttt{L\_} which maps any ordinal  $\alpha<\epsilon_0$ to a proven realization of 
$\texttt{L\_spec}\;\alpha$.   
\subsection{Definition}

In order to get a total function, we use our type \texttt{E0} of well-formed ordinal terms,(see Sect~\vref{sect:E0-def}).  Our definition is structured along a well-founded recursion and
a case-study (zero, limit and successor ordinals).

\vspace{4pt}XS
\noindent
\emph{From Module~\href{../src/html/hydras.Epsilon0.L_alpha.html\#L_}{L\_alpha}).}

\begin{Coqsrc}
From Equations Require Import Equations.
Require Import ArithRing Lia.

Instance Olt : WellFounded Lt := Lt_wf.
Hint Resolve Olt : E0.

(** Using Coq-Equations for building a function which satisfies 
    Large_sets.L_spec *)

Equations  L_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
  L_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Is_Limit alpha) :=
          { | left _ =>  L_ (Canon i alpha)  (S i) ;
            | right notlimit =>  L_ (Pred alpha) (S i)}}.

Solve All Obligations with auto with E0.
\end{Coqsrc}

It is worth looking at the answer from \texttt{Equations} and check (with \texttt{About} ) all the lemmas this plug-in gives you for free. We show here only a part of \coq's anwer.

\begin{Coqanswer}
L__obligations_obligation_1 is defined
L__obligations_obligation_2 is defined
L__obligations is defined
L__clause_1 is defined
L__functional is defined
L_ is defined
...
L__equation_1 is defined
L__graph_mut is defined
L__graph_rect is recursively defined
L__graph_correct is defined
L__elim has type-checked, generating 1 obligation
L__elim is defined
FunctionalElimination_L_ is defined
FunctionalInduction_L_ is defined
\end{Coqanswer}

Sometimes, these automatically generated statements may look cryptic.

\begin{Coqsrc}
About L__equation_1.
\end{Coqsrc}

\begin{Coqanswer}
L__equation_1 :
forall (alpha : E0) (i : nat),
L_ alpha i = L__unfold_clause_1 alpha (E0_eq_dec alpha Zero) i
\end{Coqanswer}

In most cases, it may be useful to write human-readable  paraphrases of these statements.

\begin{Coqsrc}
Lemma L_zero_eqn : forall i, L_ Zero i = i.
Proof. intro i; now rewrite L__equation_1. Qed.

Lemma L_lim_eqn alpha i : Is_Limit alpha -> L_ alpha i =
                                        L_ (Canon i alpha) (S i).

Lemma L_succ_eqn alpha i :  L_ (Succ alpha) i = L_  alpha (S i).
\end{Coqsrc}

Using these three lemmas as rewrite rules, we can prove more properties of the \texttt{L\_} family of functions, including its correctness w.r.t. the specification
\texttt{L\_spec}.

\begin{Coqsrc}
Lemma L_ge_S alpha : alpha <> Zero -> S <<= L_ alpha.

Lemma L_omega : forall k, L_ omega%e0 k = S (2 * k)%nat.

Theorem L_correct alpha : L_spec (cnf alpha) (L_ alpha).
\end{Coqsrc}

\subsubsection{Back to Hydra battles}

Theorem \texttt{battle\_length\_std } of
Module~\href{../src/html/hydras.Hydra.Hydra_Theorems.html\#battle_length_std}{Hydra.Hydra\_Theorems} relates the length of standard battles with the functions $L\_alpha$.

\begin{Coqsrc}
Theorem battle_length_std (alpha : E0)  :
  alpha <> Zero ->
  forall k, (1 <= k)%nat ->
            fight_length standard k (iota (cnf alpha))
                         (L_ alpha (S k) - k).
\end{Coqsrc}

\section{The Wainer-Hardy hierarchy}

In order to give an idea of the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the so called \emph{Wainer-Hardy hierarchy} of fast growing functions,
presented for instance in~\cite{Promel2013}. 
\index{Maths!Hardy Hierarchy}

For each ordinal $\alpha$ below $\epsilon_0$, $H_\alpha$ is a 
total arithmetic function, defined  by  transfinite recursion on $\alpha$, according to three cases:

\index{Maths!Transfinite induction}

\begin{itemize}
\item If $\alpha=0$, then $\H_\alpha (k)= k$ for any natural number $k$.
\item If $\alpha=\textrm{succ}(\beta)$, then 
$H_\alpha(k)=H_\beta(k+1)$ for any $k \in \mathbb{N}$
\item If $\alpha$ is a limit ordinal, then 
$\H_\alpha(k) = H_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
\end{itemize}

\subsection{Hardy functions in \texttt{Coq}}

We define a function \texttt{H\_} of type \texttt{E0 -> nat -> nat} by transfinite induction over the type \texttt{E0} of the well formed ordinals below $\epsilon_0$.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.H_alpha.html\#H_}{Epsilon0.H\_alpha}}

\index{Coq!Plug-ins!Equations}
\index{Functions!H@H\_ (Hardy hierarchy)}

\begin{Coqsrc}
Equations H_ (alpha: E0) (i:nat) :  nat  by wf  alpha lt :=
  H_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Is_Limit alpha) :=
          { | left _ =>  H_ (Canon (S i) alpha)  i ;
            | right notlimit =>  H_ (Pred alpha) (S i)}}.

Solve All Obligations with auto with E0.
\end{Coqsrc}

\begin{Coqsrc}
Lemma H_zero_eqn : forall i, H_ Zero i = i.
Proof.   intro i; now rewrite H__equation_1.  Qed.

Lemma H_eq2 alpha i : Is_Succ alpha ->
                      H_ alpha i = H_ (Pred alpha) (S i).

Lemma H_lim_eqn alpha i : Is_Limit alpha ->
                      H_ alpha i =  H_ (Canon (S i) alpha) i.

Lemma H_eq4  alpha i :  H_ (Succ alpha) i = H_ alpha (S i).
\end{Coqsrc}

\subsection{First  steps of the Hardy hierarchy}
Using rewrite rules from \texttt{H\_eq1} to \texttt{H\_eq4}, we can explore the functions $H_\alpha$ for some small values of $\alpha$.

\subsubsection{Finite ordinals}

By induction on $i$, we prove a simple expression of \texttt{H\_ (OF i)}, where 
\texttt{OF $i$}  is the $i$-th finite ordinal.

\begin{Coqsrc}
Lemma H_Fin : forall i k: nat,  H_ i k = (i+k)%nat.
Proof with eauto with E0.
  induction i.
  - intros; simpl OF; simpl; autorewrite with H_rw E0_rw ... 
  - intros ;simpl; autorewrite with H_rw E0_rw ... 
    rewrite IHi; lia. 
Qed.
\end{Coqsrc}

\subsubsection{Multiples of \texorpdfstring{$\omega$}{omega}}

Since the canonical sequence of $\omega$ is composed of finite ordinals, 
it is easy to get the formula associated with $H_\omega$.

\begin{Coqsrc}
Lemma H_omega : forall k, H_ Omega k = S (2 * k)%nat.
Proof with auto with E0.
  intro k; rewrite H_lim_eqn ...
  - replace (Canon (S k) omega) with (Fin (S k)).
    + rewrite H_Fin; lia.
    +  now autorewrite with E0_rw.
Qed.
\end{Coqsrc}

Before going further, we prove a useful rewriting lemma:

\begin{Coqsrc}
Lemma H_Plus_Fin alpha : forall i k : nat,
    H_ (alpha + i)%e0 k = H_ alpha (i + k)%nat.
(* Proof by induction on i *)
\end{Coqsrc}

Then, we get easily formulas for $H_{\omega+i}$, and $H_{\omega\times i}$ for any natural number $i$.

\begin{Coqsrc}
Lemma H_omega_double k : H_ (omega * 2)%e0 k =  (4 * k + 3)%nat.
Proof.
 rewrite H_lim_eqn; simpl Canon.
 - ochange  (CanonS  k (omega * 2)%e0)  (omega + (S k))%e0.
  + rewrite H_Plus_Fin, H_omega;  lia.
  -  now compute.
Qed.

Lemma H_omega_3 k : H_ (omega * 3)%e0 k = (8 * k + 7)%nat.

Lemma H_omega_4 k : H_ (omega * 4)%e0 k = (16 * k + 15)%nat.

Lemma H_omega_i i  : forall k,
    H_ (omega * i)%e0 k = (exp2 i * k + Nat.pred (exp2 i))%nat.
\end{Coqsrc}

Crossing a new limit, we prove the following equality: 
$$H_{\omega^2} (k) = 2 ^ {k+1} \times (k+1) - 1$$.

\begin{Coqsrc}

Lemma H_omega_sqr : forall k,
    H_ (Phi0  2)%e0 k = (exp2 (S k ) *  S k - 1)%nat.
Proof.
  intro k; 
   rewrite H_lim_eqn; auto with E0.
  - ochange (Canon (S k) (Phi0 2)) (omega * (S k))%e0.
    +  rewrite H_omega_i; simpl (exp2 (S k)).
       *  rewrite Nat.add_pred_r.
          -- lia. 
          --   generalize (exp2_not_zero k);  lia.
    + cbn; f_equal; lia.
Qed.
\end{Coqsrc}

\subsubsection{New limits}

Our next step would be to prove an exact formula for $H_{\omega^\omega}(k)$.
Since the canonical sequence of $\omega^\omega$ is composed of all the
$\omega^i$, we first need to express $H_{\omega^i}$ for every natural number $i$.

Let $i$ and $k$ be two natural numbers. 
The ordinal $\canonseq{\omega^(i+1)}{k}$ is the product
$\omega^i \times k$, so we need also to consider ordinals of this form.

\begin{enumerate}
\item First,  we express $H_{\omega^\alpha \times (i+2)}$ in terms of
$H_{\omega^\alpha \times (i+1)}$.

\begin{Coqsrc}
Lemma H_Omega_term_1 : alpha <> Zero -> forall  k,  
    H_ (Omega_term alpha (S i)) k =
    H_ (Omega_term alpha i) (H_ (Phi0 alpha) k).
\end{Coqsrc}

\item
Then, we prove by induction on $i$ that $H_{\omega^\alpha \times (i+1)}$ is just the
$(i+1)$-th iterate of $H_{\omega^\alpha}$.

\begin{Coqsrc}
Lemma H_Omega_term (alpha : E0)  :
alpha <> Zero -> forall i k, 
  H_ (Omega_term alpha i) k = iterate  (H_ (Phi0 alpha)) (S i) k.
\end{Coqsrc}

\item In particular, we have got a formula for $H_{\omega^{i+1}}$.

\begin{Coqsrc}
Definition H_succ_fun f k := iterate f (S k) k.

Lemma H_Phi0_succ alpha  : alpha <> Zero -> forall k,
      H_ (Phi0 (Succ alpha)) k = H_succ_fun (H_ (Phi0 alpha)) k.

Lemma H_Phi0_Si : forall i k,
      H_ (Phi0 (S i)) k = iterate H_succ_fun i (H_ omega) k. 
\end{Coqsrc}

\end{enumerate}
We get now a  formula for $H_{\omega^3}$:

\begin{Coqsrc}
Lemma H_omega_cube : forall k,
    H_ (Phi0 3)%e0 k = iterate (H_ (Phi0 2))  (S k) k.
Proof.
  intro k; rewrite <-FinS_eq, -> Fin_Succ, H_Phi0_succ; auto.
  compute; injection 1; discriminate.
Qed.
\end{Coqsrc}

t seems hard to capture the complexity of this function by looking only at this
``exact'' formula. 
Let us consider a simple example, the number $H_{\omega^3}(3)$.

\begin{Coqsrc}
Section H_omega_cube_3.

Let f k :=   (exp2 (S k ) * (S k) - 1)%nat.

Remark R0 k :  H_ (Phi0 3)%e0 k = iterate f (S k) k.
\end{Coqsrc}

Thus, the number $H_{\omega^3}(3)$ can be written as four nested applications of $f$.

\begin{Coqsrc}
Fact F0 : H_ (Phi0 3) 3 = f (f (f (f 3))).
 rewrite R0; reflexivity. 
Qed.
\end{Coqsrc}

In order to make this statement more readable, we can introduce a local définition.

\begin{Coqsrc}
Let N := (exp2 64 * 64 - 1)%nat.
\end{Coqsrc}

This number looks quite big; let us compute an approximation in \texttt{Ocaml}:

\begin{Coqsrc}
# (2.0 ** 64.0 *. 64.0 -. 1.0);; 
\end{Coqsrc}

\begin{Coqanswer}
- : float = 1.1805916207174113e+21
\end{Coqanswer}

\begin{Coqsrc}
Fact F1 : H_ (Phi0 3) 3 = f (f N).
Proof.
 rewrite H_omega_cube_0; reflexivity. 
Qed.

Lemma F1_simpl : H_ (Phi0 3) 3 =
                 (exp2 (exp2 (S N) * S N) * (exp2 (S N) * S N) - 1)%nat.

\end{Coqsrc}

In a more classical writing, this number is displayed as follows:

{\Large
$$
H_{\omega^3}(3) =  2 ^ {(2 ^ {N + 1} \times (N+1) } )  \times  (2 ^ {N+1} \times ( N +1) ) - 1
$$
}

We leave as an exercise to determine the best approximation as possible of
 the size of this number (for instance its number of digits).  For instance, if
we do not take into account the multiplications in the formula above,
we obtain that, in base $2$, the number $H_{\omega^3}(3)$ has at least
$2^{10^{21}}$  digits. But it is still an under-approximation !

\begin{Coqsrc}
End H_omega_cube_3.
\end{Coqsrc}

Now, we have got at last an exact formula for $H_{\omega^\omega}$.

\begin{Coqsrc}
Lemma H_Phi0_omega : forall k, H_ (Phi0 omega) k =
                               iterate H_succ_fun  k (H_ omega) k.
Proof with auto with E0.
  intro k; rewrite H_lim_eqn,  <- H_Phi0_Si ...
  -  rewrite CanonS_Canon, CanonS_Phi0_lim;  f_equal ...
Qed.
\end{Coqsrc}

Using extensionality of the functional \texttt{iterate}, we can get a closed formula.

\begin{Coqsrc}
Lemma H_Phi0_omega_closed_formula k :
  H_ (Phi0 omega) k =
  iterate (fun (f: nat -> nat) (l : nat) => iterate  f (S l) l)
               k
               (fun k : nat => S (2 * k)%nat)
               k.
\end{Coqsrc}

Note that this short formula contains two occurences of the functional \texttt{iterate}, the outer one is in fact a second-order iteration (on type \texttt{nat -> nat)}
and the inner one  first-order (on type \texttt{nat}).

\subsection{Abstract properties of H-functions}
~\label{sect:H-alpha-prop}

Since pure computation seems to be useless for dealing with expressions of the form $H_\alpha(k)$, even for small values of $\alpha$ and $k$, we need to prove theorems for comparing $H_\alpha(k)$ and $H_\beta(l)$, in terms of comparison
between $\alpha$ and $\beta$ on the one hand, $k$ and $l$ on the other hand.

But beware of non-theorems! For instance, one could believe that $H$ is monotonous in its first argument. The following proof shows this is false.

\begin{Coqsrc}
Remark H_non_mono1 :
  ~ (forall alpha beta k, (alpha <= beta)%e0 ->
                          (H_ alpha k <= H_ beta k)%nat).
Proof.
 intros H ;specialize (H 42 omega 3).
 assert (H0 :(42 <= omega)%e0) by (repeat split; auto).  
 apply H in H0; rewrite H_Fin, H_omega  in H0; lia.
Qed.
\end{Coqsrc}

On the contrary, the fonctions of the Hardy hierarchy have the following five properties. %%% ICI lien avec KS81

For any $\alpha < \epsilon_0$,
\begin{itemize}
\item the function $H_\alpha$ is strictly monotonous :
      For all $n,p \in\mathbb{N}, n < p \Rightarrow H_\alpha(n)< H_\alpha(p)$.
\item If $\alpha \not= 0$, then for every $n$, $n<H_\alpha(n)$.
\item The function $H_\alpha$ is pointwise less or equal than $H_{\alpha+1}$

\item For any $n\geq 1$, $H_\alpha(n)<H_{\alpha+1}(n)$.
\emph{We say that $H_{\alpha+1}$ dominates $H_\alpha$ from $1$}.
\item For any $n$ and $\beta$, if $\alpha \xrightarrow[n]{} \beta$, then
$H_\beta(n)\leq H_\alpha(n)$.
\end{itemize}

\index{Maths!Transfinite induction}

In \coq{}, we follow the  proof in~\cite{KS81}. This proof is mainly a single  proof by transfinite induction on $\alpha$ of the conjonction of the five properties.
For each $\alpha$, the three cases : $\alpha=0$, $\alpha$ is a limit, and 
$\alpha$ is a successor are considered. Inside each case, the five sub-properties are proved sequentially.

\begin{Coqsrc}
Section Proof_of_Abstract_Properties.
  Record P (alpha:E0) : Prop :=
    mkP {
        PA : strict_mono (H_ alpha);
        PB : alpha <> Zero -> forall n,  (n < H_ alpha n)%nat;
        PC : H_ alpha <<= H_ (Succ alpha);
        PD : dominates_from 1 (H_ (Succ alpha)) (H_ alpha);
        PE : forall beta n, Canon_plus n alpha beta -> 
                            (H_ beta n <= H_ alpha n)%nat}.

Theorem P_alpha : forall alpha, P alpha.
  Proof.
    intro alpha; apply well_founded_induction with lt.
   (* rest of proof skipped *)

Section Proof_of_Abstract_Properties.
\end{Coqsrc}

\subsection{Comparison between \texttt{L\_} and \texttt{H\_} }

By well-founded induction on $\alpha$, we prove that our $L$ hierachy is ``almost'' the Hardy hierarchy (up to a small shift).

\emph{From Module~\href{../src/html/hydras.Epsilon0.L_alpha.html\#H_L_}{Epsilon0.L\_alpha}}

\begin{Coqsrc}
 Theorem H_L_ alpha: forall i:nat,  (H_ alpha i <= L_ alpha (S i))%nat.
\end{Coqsrc}

\subsubsection{Back to hydras}

The following theorem relates the length of (standard) battles with the Hardy hierarchy.

\emph{From Module~\href{../src/html/hydras.Epsilon0.L_alpha.html}{Epsilon0.L\_alpha}}

\begin{Coqsrc}
Theorem battle_length_std_Hardy (alpha : E0) :
  alpha <> Zero ->
  forall k , 1 <= k -> exists l: nat,  
       H_ alpha k - k <= l /\
       fight_length standard k (iota (cnf alpha)) l.    
\end{Coqsrc}

\section{The Wainer hierarchy (functions \texorpdfstring{$F_\alpha$}{F\_alpha})}

\index{Maths!Wainer Hierarchy}

The Wainer hierarchy~\cite{BW85, Wainer1970, KS81}, is also a family of fast growing functions, indexed by ordinals below $\epsilon_0$, by the following equations:

\label{F_equations}
\begin{itemize}
\item $F_0(i)=i+1$
\item $F_{\beta+1}(i)= (F_\beta)^{(i+1)}(i)$, where $f^{(i)}$ is the $i$-th iterate of $f$.
\item $F_\alpha(i) = F_{\canonseq{\alpha}{i}} (i)$ if $\alpha$ is a limit ordinal.
\end{itemize}

A first attempt is to write a definition of $F_\alpha$ by equations, in the same as for $H\_alpha$.  We use the functional \texttt{iterate} defined in 
Module~\href{../src/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.

\begin{Coqsrc}
Fixpoint iterate {A:Type}(f : A -> A) (n: nat)(x:A) :=
  match n with
  | 0 => x
  | S p => f (iterate  f p x)
  end.
\end{Coqsrc}

The following code comes from 
 \url{../src/html/hydras.Epsilon0.F_alpha.html}.

\index{Coq!Plug-ins!Equations}
\begin{Coqsrc}
Fail Equations F_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
  F_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Is_Limit alpha) :=
          { | left _ =>  F_ (Canon i alpha)  i ;
            | right notlimit =>  iterate (F_ (Pred alpha))  (S i) i}}.
\end{Coqsrc}

\begin{Coqanswer}
The command has indeed failed with message:
In environment
alpha : E0
notlimit : Is_Limit alpha = false
nonzero : alpha <> Zero
i : nat
F_ : forall x : E0, nat -> x < alpha -> nat
The term "F_ (Pred alpha) ?x" has type "Pred alpha < alpha -> nat"
while it is expected to have type "Pred alpha < alpha -> Pred alpha < alpha"
(cannot unify "nat" and "Pred alpha < alpha").
\end{Coqanswer}

We presume that this error comes from the recursive call of \texttt{F\_} inside
an application of \texttt{iterate}. The workaround we propose is to define first 
the iteration of \texttt{F\_}  as an helper $F^*$, then to define the function $F$ as a ``iterating $F^*$ once''.

\texttt{Equations} accepts the following definition, relying on  lexicographic ordering on pairs $(\alpha,n)$.

\label{sect:F-equations}

\index{Coq!Plug-ins!Equations}
\index{Functions!F\_@F\_ (Wainer hierarchy)}
\index{Maths! Fast growing functions}

\begin{Coqsrc}
Definition call_lt (c c' : E0 * nat) :=
  lexico Lt (Peano.lt) c c'.

Lemma call_lt_wf : well_founded call_lt.
  unfold call_lt; apply Inverse_Image.wf_inverse_image,  wf_lexico.
  -  apply E0.Lt_wf.
  -  unfold Peano.lt; apply Nat.lt_wf_0. 
Qed.

Instance WF : WellFounded call_lt := call_lt_wf.

Equations  F_star (c: E0 * nat) (i:nat) :  nat by wf  c call_lt :=
    F_star (alpha, 0) i := i;
    F_star (alpha, 1) i
      with E0_eq_dec alpha Zero :=
           { | left _ => S i ;
             | right nonzero
                 with Utils.dec (Is_Limit alpha) :=
                 { | left _ => F_star (Canon i alpha,1) i ;
                   | right notlimit =>
                     F_star (Pred alpha, S i)  i}};
    F_star (alpha,(S (S n))) i :=
               F_star (alpha, 1) (F_star (alpha, (S n)) i).

(* Finally, F_ alpha is defined as its first iterate ! *)

Definition F_  alpha i := F_star (alpha, 1) i.
\end{Coqsrc}

It is quite easy to prove that our function \texttt{F\_} satisfies the equations on page~\pageref{sect:F-equations}.

\begin{Coqsrc}
Lemma F_zero_eqn : forall i, F_ Zero i = S i.

Lemma F_lim_eqn : forall alpha i,  Is_Limit alpha ->
                               F_ alpha i = F_ (Canon i alpha) i.

Lemma F_succ_eqn : forall alpha i,
    F_ (Succ alpha) i = iterate (F_ alpha) (S i) i.
\end{Coqsrc}

As for the Hardy functions, we can use these equalities as rewrite rules for
``computing'' some values of $F_\alpha(i)$, for small values of $\alpha$.

\begin{Coqsrc}
Lemma LF1 : forall n,  F_ 1 n = S (2 * n).

Lemma LF2 : forall i, (exp2 i * i < F_ 2 i)%nat.
\end{Coqsrc}

Like in Sect~\ref{sect:H-alpha-prop}, we prove by induction the following properties (see~\cite{KS81}).

\begin{Coqsrc}
Theorem F_alpha_mono alpha : strict_mono (F_ alpha).

Theorem F_alpha_ge_S alpha : forall n, (n < F_ alpha n)%nat.

Theorem F_alpha_Succ_le alpha : F_ alpha <<= F_ (Succ alpha).

Theorem F_alpha_dom alpha : dominates_from 1 (F_ (Succ alpha)) (F_ alpha).

Theorem F_alpha_beta alpha : forall beta n, Canon_plus n alpha beta -> 
                                        (F_ beta n <= F_ alpha n)%nat.
\end{Coqsrc}

As a corollary, we prove the following proposition, p. 284 of~\cite{KS81}.

\begin{quote}
  If $\beta<\alpha$, $F_\alpha$ dominates $F_\beta$.

($f$ is said to \emph{dominate} $g$ if $f(p)>g(p)$ for any all sufficiently large $p$).
\end{quote}

\begin{Coqsrc}
Lemma Propp284 : forall alpha beta : E0, 
   beta < alpha -> dominates (F_ alpha) (F_ beta).
\end{Coqsrc}

\index{Exercises}

\begin{exercise}
Let us quote a theorem from ~\cite{KS81} (page 297).

\begin{quote}
\begin{align*}
  H_{\omega^\alpha}(n+1) &\geq F_{\alpha}(n) \quad (n\geq 1, \alpha<\epsilon_0) \\
 F_{\alpha}(n+1) &\geq H_{\omega^\alpha}(n) \quad (n\geq 1, \alpha<\epsilon_0) 
\end{align*}
Thus $H_{\omega^\alpha}$ and $F_{\alpha}$ have essentially the same order of growth.

\end{quote}

 But, before trying to prove these facts, look at the definition of function $H$ in Ketonen and Solovay's paper ! Is it really the same as the definition we quote from Pr{\H o}mel's chapter~\cite{Promel2013},
whereas \cite{KS81} define $H_\alpha(n)$ as ``the least integer $k$ such that $[n,k]$ is $\alpha$-large''. Thus, it may be useful to adapt the statement above.

\end{exercise}

\index{Exercises}

\begin{exercise}
Prove the following result~\cite{KS81}(p. 298).

\begin{quote}
   For $n\geq 2$ and $\alpha \geq 3$, $F_\alpha(n+1)\geq 2^{F_\alpha(n)}$.
\end{quote}
\end{exercise}

%  \subsection{Gnawing ordinals}

% \begin{definition}[After~\cite{KS81}]
%   Let $S$ be a finite set of positive integers, and $\alpha$ be an ordinal strictly less than $\epsilon_0$.  Let us denote by $s=s_1,s_2, \dots, s_N$ the sequence of 
% the elements of $S$, enumerated in strictly increasing order.

% We consider the sequence of ordinals $\alpha_o=\alpha, \alpha_1=\canonseq{\alpha_0}{s_1},\dots,\alpha_{i+1}=\canonseq{\alpha_i}{s_{i+1}},\dots, 
% \alpha_{N}=\canonseq{\alpha_{N-1}}{s_N}$.
% We denote by $\gnaw{s}{\alpha}$ the last ordinal of the sequence, \emph{i.e.}  $\alpha_N$.
% \end{definition}

% The following function computes $\gnaw{s}{\alpha}$ by recursion on $s$.
% \vspace{4pt}

% For instance, let us consider the ordinal $\omega^\omega$, and try some sequence of integers.

% \begin{Coqsrc}
% Compute pp (gnaw (omega ^ omega) (1::nil)).  
% \end{Coqsrc}

% \begin{Coqanswer}
%   = omega%pT1 : ppT1
% \end{Coqanswer}

% \begin{Coqsrc}
% Compute pp (gnaw (omega ^ omega) (1::2::nil)).  
% \end{Coqsrc}

% \begin{Coqanswer}
% = P_fin 2
%      : ppT1
% \end{Coqanswer}

% Likewise, we can verify that $\gnaw{\omega^\omega}{\langle 1,2,3 \rangle}=1$
% and $\gnaw{\omega^\omega}{\langle 1,2,3,4 \rangle}=0$.

% \begin{Coqsrc}
% Example omega_omega_1_4 : gnaw (omega ^omega) (interval 1 4) = 0.
% Proof. trivial. Qed.

% Example omega_omega_1_3 : gnaw (omega ^omega) (interval 1 3) = 1.
% Proof. trivial. Qed.
% \end{Coqsrc}

% \begin{remark}
% \label{remark:gnaw-vs-battles}
% Let us consider an hydra battle, where Hercules always chops off the rightmost head of the hydra. Let $\alpha<\epsilon_0$ be an ordinal number, and 
% $s=\langle i_1<i_2<\dots<i_N\rangle$ be  any finite sequence of positive integers.
% Then the battle initiated by the hydra $\iota(\alpha)$, with $s$ as the sequence of successive replication factors, leads to the hydra $\iota(\gnaw{s}{\alpha})$ as the final state.
% \end{remark}

% \subsection{Large sequences}

% \begin{remark}
% In their article~\cite{KS81}, Ketonen and Solovay use the appellation ``large set'' instead of ``large sequence'', but their definitions use an enumeration of the elements in increasing order. Thus, we shall use the term ``sequence'' when referring to our implementation in \coq{}, and ``set'' when referring to the statements of~\cite{KS81}.
% \end{remark}

% \begin{definition}[After\cite{KS81}]
% The sequence $s$ is said to be \emph{$\alpha$-large} if $\gnaw{\alpha}{s}=0$.  
% \end{definition}

% \begin{Coqsrc}
% Definition largeb (alpha : T1) (s: list nat) :=
%   match gnaw alpha s with
%     | zero => true
%     | _ => false
%   end.

% Definition large (alpha : T1) (s : list nat) : Prop :=
%   largeb alpha s.
% \end{Coqsrc}

% For instance, the sequence $\langle 1,2,3,4 \rangle$ is $\omega^\omega$-large but not
% $\omega^{\omega+1}$-large, since $\gnaw{\omega^{\omega+1}}{\langle 1,2,3,4\rangle}$ is equal to $\omega \times 2 + 4$.

% %\omega^\omega+\omega^2\times 3 + \omega\times 4+ 5$.

% \begin{Coqsrc}
% Compute largeb (omega^omega) (interval 1 4).  
% \end{Coqsrc}

% \begin{Coqanswer}
% = true : bool
% \end{Coqanswer}

% \begin{Coqsrc}
% Compute largeb (omega^(omega+1)) (interval 1 4).
% \end{Coqsrc}

% \begin{Coqanswer}
% = false : bool
% \end{Coqanswer}

% \begin{remark}
% A sequence $s$ is $\alpha$-large if (still considering Hercules ``righmost-head'', tactic), it leads to Hercules' victory.
% \end{remark}

% \subsection{A little game}

% Let $\alpha<\epsilon_0$ be an ordinal, and $i$ a positive integer. We want to guess the least natural number  $j$  such that the interval $[i,i+j]$ is $\alpha$-large.
% Equivalently, we should have $\gnaw{\alpha}{[i, i+j-1]}=1$.

% The following functions takes three arguments: an ordinal $\alpha$, and two positive natural numbers  $i$ and $j$ (we assume, but not verify that $i$ and $j$ are strictly positive). It returns one of the three possible answers:

% \begin{itemize}
% \item \texttt{Ok} if $j$ is the smallest integer such that the interval $[i,i+j]$ is $\alpha$-large
% \item \texttt{Too\_far} if $[i,i+j]$ is  $\alpha$-large, but $j$ is not the smallest such positive integer
% \item \texttt{(Remaining $\beta$)} if $j$ is too small, and gnawing $\alpha$ with
%   $[i,i+j]$ is still equal to $\beta$, instead of $0$
% \end{itemize}

% \vspace{4pt}
% \noindent
% \emph{From Module~\href{../src/html/hydras.Epsilon0.Large_Sets_Demo.html}{Ordinals.Epsilon0.Large\_Sets\_Demo}}

% \begin{Coqsrc}
% Inductive answer : Set := You_won | Too_far | Remaining (rest : ppT1).

% Definition large_set_check alpha i j :=
%   let beta := gnaw alpha (interval i (Nat.pred j))
%   in match beta with
%      | one => Ok
%      | zero => Too_far
%      |  _ => Remaining (pp (canonseq j beta))
%      end.
% \end{Coqsrc}

% \subsubsection{\texorpdfstring{$\omega$}{omega}-large intervals}

% For instance, let us consider the ordinal $\omega$ and start with $i=1$.

% \begin{Coqsrc}
% Compute large_set_check omega 1 2.
% \end{Coqsrc}

% \begin{Coqanswer}
%      = Ok
%      : answer
% \end{Coqanswer}

% Let us give greater values of $i$, still with the ordinal $\omega$.

% \begin{Coqsrc}
% Compute large_set_check omega 2 3.
% \end{Coqsrc}

% \begin{Coqanswer}
%   = Remaining 1
%      : answer
% \end{Coqanswer}

% \begin{Coqsrc}
% Compute large_set_check omega 3 3.
% \end{Coqsrc}

% \begin{Coqanswer}
%   = Remaining 3
%      : answer
% \end{Coqanswer}

% \begin{Coqsrc}
% Compute large_set_check omega 2 4.
% \end{Coqsrc}

% \begin{Coqanswer}
%   = Ok
%      : answer
% \end{Coqanswer}

% \begin{Coqsrc}
% Compute large_set_check omega 3 6.
% \end{Coqsrc}

% \begin{Coqanswer}
%   = Ok
%      : answer
% \end{Coqanswer}

% It looks like every request to compute (\texttt{large\_set\_check omega $i$ $2\times i$})  will succeed. Let us try an example.

% \begin{Coqsrc}
% Compute large_set_check omega 49 98.
% \end{Coqsrc}

% \begin{Coqanswer}
%   = Ok
%      : answer
% \end{Coqanswer}

% \subsubsection{\texorpdfstring{$\omega^2$}{omega\^{}2}-large intervals}

% Still using \texttt{Compute} and \texttt{large\_set\_check}, we obtained the 
% following values of $j$, the least integer such that the interval $[i,j]$ is $\omega^2$ large.

% $$
% \begin{array}{|c|c|}
% \hline
%   i & j \\
% \hline 
% 1 & 4 \\
% 2 & 14 \\
% 3 & 38 \\
% 4 & 94 \\
% 5 & 222 \\
% % 6 & 510 
% \hline
% \end{array}
% $$

% \begin{exercise}
% Please give the 6-th and 7-th line of the array above.
% \end{exercise}

% \subsubsection{The limits of (pure) computation}

% Let us now play with bigger ordinals, for instance $\alpha=\omega^\omega$ or
% $\alpha=\omega^{\omega + 1}$. We notice that, even for small values of $i$, it is hard
% to guess values of $j$ such that $[i,i+j]$ is $\alpha$-large.

% \begin{Coqsrc}
% Compute large_set_check (omega ^ omega) 1 4.
% \end{Coqsrc}

% \begin{Coqanswer}
%  = Ok
%      : answer
% \end{Coqanswer}

% \begin{Coqsrc}
% Compute large_set_check (omega ^ omega) 2 38.
% \end{Coqsrc}

% \begin{Coqanswer}
%  = Ok
%      : answer
% \end{Coqanswer}

% \begin{Coqsrc}
% Compute large_set_check (omega ^ omega) 3 1000.
% \end{Coqsrc}

% \begin{Coqanswer}
%   = Remaining (omega ^ 2 * 2 + omega * 220 + 798)%pT1
%      : answer
% \end{Coqanswer}

% \begin{Coqsrc}
% Compute large_set_check (omega ^ omega) 3 1798.
% \end{Coqsrc}

% \begin{Coqanswer}
%   = Remaining (omega ^ 2 * 2 + omega * 220)%pT1
%      : answer
% \end{Coqanswer}

% \begin{Coqsrc}
% Compute large_set_check (omega ^ omega) 3 5000.
% \end{Coqsrc}

% \begin{Coqanswer}
%   = Remaining (omega ^ 2 * 2 + omega * 218 + 2198)%pT1
%      : answer
% \end{Coqanswer}

% \begin{Coqsrc}
% Compute large_set_check (omega ^ (omega + 1)) 3 5000.
% \end{Coqsrc}

% \begin{Coqanswer}
%  = Remaining
%          (omega ^ omega * 2 + omega ^ 3 * 4 + 
%           omega ^ 2 * 4 + omega * 1148 +  4222)%pT1
%      : answer  
% \end{Coqanswer}

% \begin{Coqsrc}
% Compute large_set_check (omega ^ (omega + 1)) 3 10000.
% \end{Coqsrc}

% \begin{Coqanswer}
% Warning: To avoid stack overflow, large numbers in nat are interpreted as
% applications of Init.Nat.of_uint. [abstract-large-number,numbers]

%     = Remaining
%          (omega ^ omega * 2 + omega ^ 3 * 4 + 
%           omega ^ 2 * 4 + omega * 1147 +
%           8446)%pT1
%      : answer
% \end{Coqanswer}

% Since computation is not enough, let us comme back to proofs, or, better, proofs \emph{and} computations.

% \subsection{Proving largeness}

% %%% ICI %%%

% \subsection{$n$-large and $\omega$-large intervals }

% A finite ordinal is just made out of \texttt{zero} and \texttt{succ} constructions.
% Thus any sequence of strictly positive integers of length greater than  or equal to $n$ will completely gnaw the ordinal $n$.

% Concerning the first limit ordinal $\omega$ the largeness of a sequence $s$ depends only on its first element and the length of the rest of the sequence (please keep in mind that the argument $i$ of $\canonseq{\alpha}{i}$ is meaningful only if $\alpha$ is a limit ordinal).
% The following proposition is labeled $4.2$ in~\cite{KS81}.

% \begin{proposition}
%   For $n<\omega$, a set $X$id $n$-large if and only if $|X|\geq n$.
% A finite set $X$ is $\omega$-large if $|X|>\min{X}$.
% \end{proposition}

% Rewritten in terms of list of strictly positive integers, we get the following statements:

% \begin{Coqsrc}
% Lemma large_n_iff : forall s (n:nat) , ~ In 0 s  ->
%                                   large n s  <-> (n <= List.length s)%nat.
% \end{Coqsrc}

% \begin{Coqsrc}
% Lemma large_omega_iff : forall s n,  ~In 0 (n::s) -> 
%                                             large omega (n::s) <->
%                                             (n <=  List.length s)%nat.
% \end{Coqsrc}

% % \begin{Coqsrc}
% % (* sorted list of natural numbers greater than or equal to n *)

% % Inductive sorted_ge (n: nat) : list nat -> Prop :=
% % | sorted_ge_nil : sorted_ge n nil
% % | sorted_ge_one : forall p, n<=p ->
% %                             sorted_ge n (p::nil)
% % | sorted_ge_cons: forall p q s,  n<=p -> p<q ->
% %                                  sorted_ge p (q::s) ->
% %                                  sorted_ge n (p::q::s).
% % \end{Coqsrc}

% \subsection{Mimimal large sequences}

% Let us consider \emph{minimal} large sequences, \emph{i.e.} large sequences 
% the strict prefix of which do not lead to $0$. In other words, only the last ordinal 
% defined by the sequence is null.  For this purpose, we use the predicate \texttt{path\_to} introduced page~\pageref{path-to-definition}.

% \begin{Coqsrc}
% Definition mlarge alpha s := path_to zero s alpha.
% \end{Coqsrc}

### 2020-06-23 latex/chapter-schutte.tex
first commit

\chapter[Countable Ordinals (after Sch\"{u}tte)]{Kurt Schütte's axiomatic Definition of countable Ordinals}

\label{chap:schutte}

In the present chapter, we  compare our implementation of the segment $[0,\epsilon_0[$ with a mathematical text in order to ``validate'' our constructions.
We chose as reference the axiomatic definition of the set of countable ordinals,
in chapter V of Kurt Schütte's book `` Proof Theory ''~\cite{schutte}.

\begin{remark}
\emph{In all this chapter, the word ``ordinal'' will be considered as a synonymous of
``countable ordinal''}  
\end{remark}

Schütte's definition of countable ordinal relies on the following three axioms.

\begin{enumerate}
\item There exists a well-ordered set $(\mathbb{O},<)$
\item Every bounded subset of $\mathbb{O}$  is countable
\item Every countable subset of $\mathbb{O}$  is bounded.
\end{enumerate}

Starting with these three axioms, Schütte re-defines the vocabulary about ordinal numbers: the null ordinal $0$, limits and successors, the addition of ordinals, the infinite ordinals $\omega$, $\epsilon_0$, $\Gamma_0$, etc.

This chapter describes an adaptation to \coq{} of Schütte's axiomatization. 
 Unlike the rest of our libraries, our library
\href{../src/html/hydras.Schutte.Schutte.html}{Ordinals.Schutte}
is not constructive, and relies on several axioms.

\begin{itemize}
\item First, please keep in mind  that the set of countable ordinals is not countable. Thus, we cannot hope to represent all countable ordinals as finite terms of an inductive type, which was possible with  the set of ordinals strictly less than $\epsilon_0$ (resp. $\Gamma_0$)
\item We tried to be as close as possible to K. Schütte's text, which uses ``classical'' mathematics : excluded middle, Hilbert's $\epsilon$ (choice) and Russel's $\iota$ (definite description) operators. Both operators allow us to write definitions close to the natural mathematical language, such as ``$\textrm{succ}$ is \emph{the} least ordinal strictly greater than $\alpha$''
\item Please note that only the sub-library \href{../src/html/hydras.Schutte.Schutte.html}{Ordinals.Schutte.Schutte/} is ``contaminated'' by axioms, and that the rest of our libraries remain constructive.

\end{itemize}

\section{Declarations and Axioms}

Let us declare a type 
\texttt{ON} for representing countable ordinals, and a binary relation
 \texttt{lt}. Note that, in our development, \texttt{ON} is a type, while the \emph{set} of countable ordinals (called $\mathbb{O}$ by Schütte) 
is the full set over the type \texttt{ON}.

\index{Types!ON@ON (countable ordinals)}

\vspace{4pt}

\emph{From Module\href{../src/html/hydras.Schutte.Schutte_basics.html}%
{\texttt{Schutte.Schutte\_basics}}}

\begin{Coqsrc}
Parameter ON : Type.
Parameter lt : relation ON.
Notation "a < b" := (lt a b): schutte_scope.

Definition ordinal := Full_set ON.
\end{Coqsrc}

Schütte's first axiom tells that \texttt{lt} is a well order on the set 
\texttt{ordinal} (The  class \texttt{WO} is defined in
Module~\href{../src/html/hydras.Schutte.Well_Orders.html}%
{\texttt{Well\_Orders.v}}).

\index{Coq!Techniques!Type classes}

\begin{Coqsrc}
Variables (M:Type)
         (Lt : relation M).

Class WO : Type:=
    {
      Lt_trans : Transitive  Lt;
      Lt_irreflexive : forall a:M, ~ (Lt a a);
      well_order : forall (X:Ensemble M)(a:M),
          In X a ->
          exists a0:M, least_member  X a0
    }.
\end{Coqsrc}

\begin{Coqsrc}
  Axiom AX1 : WO lt.
\end{Coqsrc}

The second and third axioms say that a subset $X$ of $\mathbb{O}$ is
(strictly) bounded if and only if it is countable. We use Florian Hatat's library on countable sets, written as he was a student of  \emph{\'Ecole Normale Supérieure de Lyon}.

\begin{Coqsrc}
Axiom AX2 : forall X: Ensemble ON, 
   (exists a,  (forall y, In X y -> y < a)) ->
   countable X.

Axiom AX3 : forall X : Ensemble ON,
              countable X -> 
              exists a,  forall y, In X y -> y < a.
\end{Coqsrc}

\texttt{AX2} and \texttt{AX3} could have been replaced by a single axiom (using the \texttt{iff} connector), but we decide to respect as most as possible the structure of Schütte's definitions.

\section{Additional  axioms}

The adaptation of Schütte's mathematical discourse to \coq{} led us to
import a few axioms from the standard library. We encourage the reader to consult \coq{}'s FAQ about the safe use of axioms
 \url{https://github.com/coq/coq/wiki/The-Logic-of-Coq#axioms}.

\subsubsection{Classical logic}

In order to work with classical logic, we import the module
\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Classical.html}{Coq.Logic.Classical}  of \coq{}'s standard library, specifially the following axiom:

\begin{Coqsrc}
 Axiom classic : forall P:Prop, P \/ ~P.
\end{Coqsrc}

\subsubsection{Description operators}

In order to respect Schütte's style, we imported also the library 
\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Epsilon.html}{\texttt{Coq.Logic.Epsilon}}.  The rest of this section presents a few examples of
how Hilbert's choice operator and Church's definite description allow us
 to write understandable definitions (close to the mathematical natural language).

\subsubsection{The definition of zero}

According to the  definition of a well order, every non-empty subset of \texttt{ON} has a least element. Furthermore, this least element is unique.

\begin{Coqsrc}
Remark R : exists! z : ON, least_member lt  ordinal z.
Proof.
  destruct inh_ON as [a]; apply (well_order (WO:=AX1)) with a .
  split.
Qed.
\end{Coqsrc}

Assume we want to call this element  \texttt{zero}.

\begin{Coqsrc}
Definition zero : ON.
Proof.
  Fail destruct R.
\end{Coqsrc}

\begin{Coqanswer}
The command has indeed failed with message:
Case analysis on sort Type is not allowed for inductive 
definition ex.
\end{Coqanswer}

Indeed, the basic logic of  \coq{} does not allow us to eliminate a proof of a proposition 
$\exists\,x:A,\,P(x)$ for building a term whose type lies in the sort \texttt{Type}. 
The reasons for this impossibility are explained in many documents~\cite{BC04, chlipalacpdt2011, Coq}.

Let us import the library \texttt{Coq.Logic.Epsilon}, which contains the following axiom.

\begin{Coqsrc}
Axiom epsilon_statement:
  forall (A : Type) (P : A->Prop), inhabited A ->
    {x : A | (exists x, P x) -> P x}.
\end{Coqsrc}

Hilbert's $\epsilon$ \emph{operator} is derived from this  axiom.

\begin{Coqsrc}
  Definition epsilon (A : Type) (i:inhabited A) (P : A->Prop) : A
  := proj1_sig (epsilon_statement P i).

Lemma constructive_indefinite_description :
  forall (A : Type) (P : A->Prop),
    (exists x, P x) -> { x : A | P x }.
\end{Coqsrc}

If we consider the \emph{unique existential} quantifier $\exists!$, we obtain
Church's \emph{definite description operator}.

\index{Functions!iota description operator}

\begin{Coqsrc}
Definition iota (A : Type) (i:inhabited A) (P : A->Prop) : A
  := proj1_sig (iota_statement P i).
\end{Coqsrc}

\begin{Coqsrc}
 Lemma constructive_definite_description :
  forall (A : Type) (P : A->Prop),
    (exists! x, P x) -> { x : A | P x }.
\end{Coqsrc}

\begin{Coqsrc}
Definition iota_spec (A : Type) (i:inhabited A) (P : A->Prop) :
  (exists! x:A, P x) -> P (iota i P)
  := proj2_sig (iota_statement P i).
\end{Coqsrc}

Indeed, the operators \texttt{epsilon} and \texttt{iota} allowed us to make our definitions 
quite close to Schütte's text. Our libraries \href{../src/html/hydras.Schutte.MoreEpsilonIota.html}%
{\texttt{Schutte.MoreEpsilonIota}}
and
\href{../src/html/hydras.Schutte.PartialFun.html}%
{\texttt{Schutte.PartialFun}} are extensions of \texttt{Coq.logic.Epsilon} for making easier 
such definitions. See also an article in french~\cite{PCiota}.

\index{Coq!Techniques!Type classes}

\begin{Coqsrc}
Class InH (A: Type) : Prop :=
   InHWit : inhabited A.

Definition some {A:Type} {H : InH A} (P: A -> Prop) := 
   epsilon (@InHWit A H) P.

Definition the {A:Type} {H : InH A} (P: A -> Prop) := 
   iota (@InHWit A H) P.
\end{Coqsrc}

In order to use these tools,  we had to tell \coq{}  that the type \texttt{ON} is not empty:

\begin{Coqsrc}
Axiom inh_ON : inhabited ON.
\end{Coqsrc}

We are now able de define \texttt{zero} as the least ordinal. For this purpose,
we define a function returning the least element of any [non-empty]  set.

\begin{Coqsrc}
Definition the_least {M: Type} {Lt}
           {inh : InH M} {WO: WO Lt} (X: Ensemble M)  : M :=
  the  (least_member  Lt X ).
\end{Coqsrc}

\vspace{4pt}

From Module \href{../src/html/hydras.Schutte.Schutte_basics.html}%
{\texttt{~Schutte.Schutte\_basics}}

\index{Constants!zero:ON}

\begin{Coqsrc}
Definition zero: ON :=the_least ordinal.
\end{Coqsrc}

We want to prove now that zero is less than or equal to any ordinal number.

\begin{Coqsrc}
Lemma zero_le (alpha : ON) :  zero <= alpha.
Proof.
  unfold zero, the_least, the; apply iota_ind.
\end{Coqsrc}

According to the use of the description operator \texttt{iota}, we have to solve  two trivial sub-goals.
\begin{enumerate}
\item Prove that there exists a unique least member of \texttt{ON}
\item Prove that being a least member of \texttt{ON} entails the announced inequality 
\end{enumerate}

\begin{Coqanswer}
2 subgoals (ID 155)

  alpha : ON
  ============================
  exists ! x : ON, least_member lt ordinal x

subgoal 2 (ID 156) is:
 forall a : ON, unique (least_member lt ordinal) a -> 
                a <= alpha
\end{Coqanswer}

\begin{Coqsrc}
  -  apply the_least_unicity, Inh_ord.
  -  destruct 1 as [[_ H1] _]; apply H1; split. 
Qed.
\end{Coqsrc}

\subsubsection{Remarks on \texttt{epsilon} and \texttt{iota}}

 What would happen in case of a misuse of \texttt{epsilon} or \texttt{iota} ?
For instance, one could give a unsatisfiable specification to \texttt{epsilon} or 
a specification for \texttt{iota} that admits several realizations.

Let us consider an example:

\begin{Coqbad}
Module Bad.

 Definition bottom := the_least (Empty_set ON).
\end{Coqbad}

\begin{Coqanswer}
 bottom is defined
\end{Coqanswer}

Since we won't be able to prove the proposition
\linebreak \Verb|{exists! a: ON, least_member (Empty_set ON) a|, the only properties we would be able to prove about \texttt{bottom} would be \emph{trivial} properties, 
\emph{i.e.}, satisfied by \emph{any} element of type \texttt{ON}, like for instance
\texttt{bottom = bottom}, or \texttt{zero <= bottom}.

\begin{Coqbad}
Lemma le_zero_bottom : zero <= bottom. 
Proof. apply zero_le. Qed.

Lemma bottom_eq : bottom = bottom.
Proof. trivial. Qed.

Lemma le_bottom_zero : bottom <= zero.
Proof.
   unfold bottom, the_least, the; apply iota_ind.
\end{Coqbad}

\begin{Coqanswer}
2 subgoals (ID 413)

  ============================
  exists ! x : ON, least_member lt (Empty_set ON) x

subgoal 2 (ID 414) is:
 forall a : ON, unique (least_member lt (Empty_set ON)) a -> 
      a <= zero
\end{Coqanswer}

\begin{Coqbad}
Abort.
End Bad.
\end{Coqbad}

In short, using \texttt{epsilon} and \texttt{iota} in our implementation of countable ordinals after Schütte has two main advantages.

\begin{itemize}
\item It allows us to give a \emph{name} (using \texttt{Definition}) two witnesses 
of existential quantifiers (let us recall that, in classical logic, one may consider non-constructive proofs of existential statements)
\item By separating definitions from proofs of [unique] existence, one may make definitions  more concise and readable. Look for instance at 
the definitions of  \texttt{zero}, \texttt{succ}, \texttt{plus}, etc. in the rest of this chapter.
\end{itemize}
%%%% ICI ICI

\section{The  successor function}

The definition of the function \texttt{succ:ON -> ON} is very concise. The successor of any ordinal $\alpha$ is the smallest ordinal strictly greater than $\alpha$.

\index{Functions!succ (in type ON)}

\begin{Coqsrc}
Definition succ (alpha : ON) := the_least (fun beta => alpha < beta).
\end{Coqsrc}

Using \texttt{succ}, we define the folloing predicates.

\begin{Coqsrc}
Definition is_succ (alpha:ON) := exists beta, alpha = succ beta.

Definition is_limit (alpha:ON) := alpha <> zero /\ ~ is_succ alpha.
\end{Coqsrc}

It is also easy to define recursively the finite ordinals.

\begin{Coqsrc}
Reserved Notation "'F' n" (at level 29) .

Fixpoint finite (i:nat) : ON :=
  match i with 
            | 0 => zero
            | S i => succ (F i)
  end
where "'F' i" := (finite i)  : schutte_scope.

Coercion finite : nat >-> ON.
\end{Coqsrc}

\begin{remark}
Please look at remark~\vref{warning:coercions}.
\end{remark}

How do we prove properties of the successor function?
First, we make its specification explicit.

\begin{Coqsrc}
Definition succ_spec (alpha:ON) :=
  least_member   lt (fun z => alpha < z).
\end{Coqsrc}

Then, we prove that our function \texttt{succ} meets this specification.

\begin{Coqsrc}
Lemma succ_ok : forall alpha,  succ_spec alpha  (succ alpha).
Proof.
  intros; unfold succ, the_least, the;  apply iota_spec.
\end{Coqsrc}

\begin{Coqanswer}
1 subgoal (ID 172)

  alpha : ON
  ============================
  exists ! x : ON, succ_spec alpha x
\end{Coqanswer}

We have now to prove that the set of all ordinals strictly greater than $\alpha$ has a unique least element. But the singleton set $\{\alpha\}$ is countable, hence  bounded (by the axiom \texttt{AX3}). Hence; the set $\{\beta\in\mathbb{O}|\alpha < \beta\}$ is not empty
and therefore has a unique least element.

The \coq{} proof script is quite short.

\begin{Coqsrc}
  destruct (@AX3 (Singleton _ alpha)).
  - apply countable_singleton.
  -  unfold succ_spec; apply the_least_unicity;  exists x; intuition.
Qed.     
\end{Coqsrc}

We can ``uncap'' the description operator for proving properties of the
\texttt{succ} function.

\begin{Coqsrc}
Lemma lt_succ (alpha : ON) :  alpha < succ alpha.
Proof.
  destruct  (succ_ok  alpha);  tauto.
Qed.

Hint Resolve lt_succ : schutte.

Lemma lt_succ_le (alpha beta : ON):
  alpha < beta -> succ alpha <= beta.
Proof with eauto with schutte.
  intros  H;  pattern (succ alpha); apply the_least_ok ... 
  exists (succ alpha); red;apply lt_succ ...
Qed.
\end{Coqsrc}

\begin{Coqsrc}
Lemma lt_succ_le_2 (alpha beta : ON):
  alpha < succ beta -> alpha <= beta.

Lemma succ_mono (alpha beta : ON):
  alpha < beta -> succ alpha < succ beta.

Lemma succ_monoR (alpha beta : ON) :
 succ alpha < succ beta -> alpha < beta.

Lemma lt_succ_lt (alpha beta : ON) :
  is_limit beta ->  alpha < beta -> succ alpha < beta.
\end{Coqsrc}

\section{The definition of \texttt{omega}}
In order to define $\omega$, the first infinite ordinal, we use an operator which
``returns'' the least upper bound (if it exists) of a subset $X\subseteq \mathbb{O}$.
For that purpose, we first use a predicate:
(\texttt{is\_lub $D$ \textit{lt} $X$ $a$}) if $a$ belongs to $D$ and is the least 
upper bound  of $X$ (with respect to \textit{lt}).

\begin{Coqsrc}
Definition is_lub (M:Type)
                  (D : Ensemble M)
                  (lt : relation M)
                  (X:Ensemble M)
                  (a:M) :=
   In _ D a  /\ upper_bound  D lt X a  /\
   (forall y, In _ D y -> upper_bound  D lt X y  -> 
                  y = a \/ lt a y).
\end{Coqsrc}

\begin{Coqsrc}
Definition sup_spec X lambda := is_lub ordinal lt X lambda.

Definition sup (X: Ensemble ON) : ON  := the  (sup_spec X).

Notation "'|_|' X" := (sup X) (at level 29) : schutte_scope.
\end{Coqsrc}

Then, we define the function \texttt{omega\_limit} which returns the least upper bound 
of the  (denumerable) range of any sequence \texttt{s: nat -> ON}. 
By \texttt{AX3} this range is bounded, hence the set of its upper bounds is not empty and has a least element.

\begin{Coqsrc}
Definition omega_limit (s:nat->ON) : ON 
  := |_| (seq_range s).
\end{Coqsrc}

Then we define \texttt{omega} as the limit of the sequence of finite ordinals.

\index{Notations!omega@omega (the least infinite ordinal)}

\begin{Coqsrc}
Definition _omega := omega_limit finite.

Notation "'omega'" := (_omega) : schutte_scope.
\end{Coqsrc}

Among the numerous properties of the ordinal $\omega$, les us quote the following ones
(proved in Module 
\href{../src/html/hydras.Schutte.Schutte_basics.html\#finite_lt_omega}{\texttt{Schutte.Schutte\_basics}})

\begin{Coqsrc}
Lemma finite_lt_omega : forall i: nat,  i < omega.

Lemma lt_omega_finite alpha : ON) : 
  alpha < omega ->  exists i:nat, alpha =  i.

Lemma is_limit_omega : is_limit omega.
\end{Coqsrc}

\subsection{Ordering functions and ordinal addition}

After having defined the finite ordinals and the infinite ordinal $\omega$, we  define the sum $\alpha+\beta$ of two countable ordinals.
Schütte's definition looks like the following one:

\begin{quote}
``$\alpha+\beta$ is the $\beta$-th ordinal greater than or equal to $\alpha$''
\end{quote}

The purpose of this section is to give a meaning to the construction
``the $\alpha$-th element of $X$''  where $X$ is any non-empty subset of $\mathbb{O}$.
We follow Schütte's approach, by defining the notion of \emph{ordering functions},
a way to associate a unique ordinal to each element of $X$.
Complete definitions and proofs can be found in Module
 \href{../src/html/hydras.Schutte.Ordering_Functions.html}%
{\texttt{Schutte.Ordering\_Functions}} ).

\subsection{Definitions}

A \emph{segment} is a set $A$ of ordinals such that, whenever  $\alpha\in A$ and
$\beta<\alpha$, then $\beta\in A$; a segment is  \emph{proper} if it strictly included in $\mathbb{O}$.

\begin{Coqsrc}
 Definition segment (A: Ensemble ON) :=
  forall alpha beta, In A alpha -> beta < alpha -> In A  beta.

Definition proper_segment (A: Ensemble ON) :=
  segment A /\ ~ Same_set A ordinal.
\end{Coqsrc}

Let  $A$ be a segment, and $B$ a subset of $\mathbb{O}$ : an \emph{ordering function for $A$ and  $B$} is a strictly increasing bijection from $A$ to $B$.
The set $B$ is said to be an \emph{ordering segment} of $A$.
Our definition in \coq{} is a direct translation of the mathematical text of~\cite{schutte}.

\index{Maths!Ordering functions}

\begin{Coqsrc}
Definition ordering_function (f : ON -> ON)(A B : Ensemble ON) :=
 segment A /\
 (forall a, In A a -> In B (f a)) /\
 (forall b, B b -> exists a, In A a /\ f a = b) /\
 forall a b, In A a -> In A b -> a < b ->  f a < f b.

Definition ordering_segment (A B : Ensemble ON) :=
  exists f : ON -> ON, ordering_function f A B.
\end{Coqsrc}

We are now able to associate with any subset $B$ of $\mathbb{O}$ its ordering segment and ordering function.

\begin{Coqsrc}
Definition the_ordering_segment (B : Ensemble ON) :=
  the  (fun x => ordering_segment x B).

Definition ord  (B : Ensemble ON) := 
  some (fun f => ordering_function f (the_ordering_segment B) B).
\end{Coqsrc}

Thus (\texttt{ord $B \;\alpha$}) is the $\alpha$-th element of $B$.
Please note that the last definition uses the epsilon-based operator \texttt{some} and
not \texttt{the}. This is due to the fact that we cannot prove the unicity (w.r.t. Leibniz' equality) of the ordering function of a given set. 
By contrast, we admit the axiom  \texttt{Extensionality\_Ensembles}, from the library 
\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Sets.Ensembles.html}{Coq.Sets.Ensembles}, so we use the operator \texttt{the} in the definition of
\texttt{the\_ordering\_segment}.

One of the main theorems of
\href{../src/html/hydras.Schutte.Ordering_Functions.html\#ordering_function_ex}%
{\texttt{Ordering\_Functions}} 
associates a unique segment and a unique (up to extensionality) ordering function to every subset $B$ of $\mathbb{O}$.

\begin{Coqsrc}
About ordering_function_ex.
\end{Coqsrc}

\begin{Coqanswer}
forall B : Ensemble ON,
 exists ! S : Ensemble ON, 
      exists f : ON -> ON, ordering_function f S B
\end{Coqanswer}

\begin{Coqanswer}
ordering_function_unicity :
forall (B S1 S2 : Ensemble ON) (f1 f2 : ON -> ON),
ordering_function f1 S B ->
ordering_function f2 S2 B -> 
fun_equiv f1 f2 S1 S2
\end{Coqanswer}

Thus,  our function \texttt{ord}  which enumerates the elements of $B$ is defined in a non-ambiguous way.
Let us quote the following theorems (see Library
\href{../src/html/hydras.Schutte.Ordering_Functions.html}%
{\texttt{Schutte.Ordering\_Functions}} for more details).

\begin{Coqsrc}
Theorem ordering_le : forall f A B,
    ordering_function f A B ->
    forall alpha, In A alpha -> alpha  <= f alpha.

Th_13_5_2 :
forall (A B : Ensemble ON) (f : ON -> ON),
ordering_function f A B -> Closed B -> continuous f A B
\end{Coqsrc}

\subsection{Ordinal addition}

We are now ready to define and study addition on the type \texttt{ON}.
The following definitions and proofs can be consulted in Module
\href{../src/html/hydras.Schutte.Addition.html}%
{\texttt{Schutte.Addition.v}}.

\begin{Coqsrc}
Definition plus alpha := ord  (ge alpha).
Notation "alpha + beta " := (plus alpha beta) : schutte_scope.
\end{Coqsrc}

In other words,  $\alpha + \beta$ is the  $\beta$-th ordinal greater than or equal to $\alpha$. 
Thanks to generic properties of ordering functions, we can show the following 
properties of addition on $\mathbb{O}$. First, we prove a useful lemma:

\begin{Coqsrc}
Lemma plus_elim (alpha : ON) :
  forall P : (ON->ON)->Prop,
    (forall f: ON->ON, 
        ordering_function f ordinal (ge alpha)-> P f) ->
    P (plus alpha).
\end{Coqsrc}

\begin{Coqsrc}
Lemma alpha_plus_zero (alpha: ON): alpha + zero = alpha.
Proof.
 pattern  (plus alpha); apply plus_elim; eauto.
 \end{Coqsrc}

 \begin{Coqanswer}
 1 subgoal (ID 24)

  alpha : ON
  ============================
  forall f : ON -> ON,
  ordering_function f ordinal (ge alpha) -> 
  f zero = alpha
 \end{Coqanswer}

 \begin{Coqsrc}
 (* rest of proof skipped *)
 \end{Coqsrc}

The following lemmas are proved the same way.

 \begin{Coqsrc}

Lemma zero_plus_alpha (alpha : ON) : zero + alpha = alpha.

Lemma le_plus_l (alpha beta : ON) : alpha <= alpha + beta.

Lemma le_plus_r (alpha beta : ON) :  beta <= alpha + beta.

Lemma plus_mono_r (alpha beta gamma : ON) : 
    beta < gamma -> alpha + beta < alpha + gamma.

Lemma plus_of_succ (alpha beta : ON) :
    alpha + (succ beta) = succ (alpha + beta).

Theorem plus_assoc (alpha beta gamma : ON) :
  alpha + (beta + gamma) = (alpha + beta) + gamma.

Lemma one_plus_omega :  1 + omega = omega.

Lemma finite_plus_ge_omega (n : nat) (alpha : ON) :
  omega <= alpha -> n + alpha = alpha.
\end{Coqsrc}

It isinteresting to compare the proof of these lemmas with the 
computational proofs of the corresponding statements in Module
\href{../src/html/hydras.Epsilon0.T1.html}%
{\texttt{Epsilon0.T1}}. 
For instance, the proof of the lemma 
\texttt{one\_plus\_omega} uses the continuity of ordering functions (applied to  \texttt{(plus 1)}) and compares the limit of the $\omega$-sequences $i_{(i \in \mathbb{N})}$ and
$(1+i)i_{(i \in \mathbb{N})}$, whereas in the library  \texttt{Epsilon0/T1}, the equality 
$1+\omega=\omega$ is just proved with \texttt{reflexivity}!

\subsubsection{Multiplication by a natural number}

The multiplication of an ordinal by a natural number is defined in terms of addition.
This operation is useful for the study of Cantor normal forms.

\begin{Coqsrc}
Fixpoint mult_Sn (alpha:ON)(n:nat){struct n} :ON :=
 match n with 
            | 0 => alpha
            | S p => mult_Sn alpha p + alpha
 end.

Definition mult_n alpha n :=
  match n with
      0 => zero
    | S p => mult_Sn alpha p
  end.

Notation "alpha * n" := (mult_n alpha n) : schutte_scope.
\end{Coqsrc}

\section{The exponential of basis \texorpdfstring{$\omega$}{omega}}

In this section, we define the function which maps any $\alpha\in\mathbb{O}$ to
the ordinal  $\omega^\alpha$, also written 
$\varphi_0(\alpha)$. 
It is an opportunity to apply the definitions and results of the preceding section. 
Indeed,  Schütte first defines a subset of $\mathbb{O}$: the set of additive principal ordinals, and $\varphi_0$  is just defined as the ordering function of this set.

\subsection{Additive principal ordinals}

\index{Maths!Additive principal ordinals}

\begin{definition}
A non-zero ordinal  $\alpha$ is said to be \emph{additive principal} if, for all  $\beta<\alpha$, $\beta+\alpha$ is equal to  $\alpha$.
We call \texttt{AP} the set of additive principal ordinals.

\end{definition}

\noindent\emph{From Module \href{../src/html/hydras.Schutte.AP.html}%
{\texttt{Schutte.AP}}}

\begin{Coqsrc}
Definition AP : Ensemble ON :=
  fun alpha => 
  zero < alpha /\
  (forall beta, beta < alpha ->  beta + alpha = alpha).
\end{Coqsrc}

\subsection{The function \texttt{phi0}}

Let us call  $\varphi_0$ the ordering function of \texttt{AP}.
In the mathematical text, we shall use indifferently the notations  $\omega^\alpha$ and$\varphi_0(\alpha)$.

\index{Functions!phi0@phi0 (exponential of base omega)}

\begin{Coqsrc}
Definition phi0 := ord AP.

Notation "'omega^'" := phi0 (only parsing) : schutte_scope.
\end{Coqsrc}

\subsection{Omega-towers and the ordinal \texorpdfstring{$\epsilon_0$}{epsilon0}}

Using $\varphi_0$, we can define recursively the set of finite omega-towers.

\begin{Coqsrc}
Fixpoint omega_tower (i : nat) : ON :=
  match i with
    0 =>  1
  | S j => phi0 (omega_tower j)
  end.
\end{Coqsrc}

Then, the ordinal  $\epsilon_0$ is defined as the limit of the sequence of all finite towers (a kind of infinite tower).

\begin{Coqsrc}
Definition epsilon0 := omega_limit omega_tower.
\end{Coqsrc}

The rest of our library \texttt{AP} is devoted to the proof of properties of additive principal ordinals, hence of the ordering function  $\varphi0$ and the ordinal $\epsilon_0$ (which we could not express within the type \texttt{T1}).

\subsection{Properties of the set  \texttt{AP}}

The set of additive principal ordinals is not empty: it contains at least the ordinals  $1$ and  $\omega$.

\begin{Coqsrc}
Lemma AP_one : In AP 1.

Lemma AP_omega : In AP omega.
\end{Coqsrc}

Moreover, $1$ is the least principal ordinal and $\omega$ is the second element of
\texttt{AP}.

\begin{Coqsrc}
Lemma least_AP: least_member  lt AP 1.

Lemma omega_second_AP :
  least_member   lt 
                  (fun alpha => 1 < alpha /\ In AP alpha)
                  omega.
\end{Coqsrc}

The set  \texttt{AP} is  \emph{closed} under addition, and unbounded.

\begin{Coqsrc}
Lemma AP_plus_closed (alpha beta gamma : ON): 
     In AP alpha -> beta < alpha -> gamma < alpha -> beta + gamma < alpha.

Theorem AP_unbounded : Unbounded AP.
\end{Coqsrc}

Finally, \texttt{AP} is (topologically) \emph{closed} and ordered by the segment of all countable ordinals.

\index{Predicates!Closed}

\begin{Coqsrc} 
Definition Closed (B : Ensemble ON) : Prop := 
  forall M, Included M B -> Inhabited _ M -> 
                 countable M -> In B (|_| M).
\end{Coqsrc}

\begin{Coqsrc}
Theorem AP_closed : Closed AP.

Lemma AP_o_segment :  the_ordering_segment AP = ordinal.
\end{Coqsrc}

\subsubsection{Properties of the function \texorpdfstring{$\varphi_0$}{phi0}}

The ordering function $\varphi_0$ of the set \texttt{AP} is defined on the full set $\mathbb{O}$ and is continuous (Schütte says that this function is \emph{normal}).

\begin{Coqsrc}
Theorem normal_phi0 : normal phi0 AP.
\end{Coqsrc}

The following properties come from  the definition of $\varphi_0$ as the ordering function of \texttt{AP}. It may be interesting to compare these proofs with the computational ones described in Chapter ~\ref{chap:T1}.

\begin{Coqsrc}
Lemma AP_phi0 (alpha : ON) : In AP (phi0 alpha).

Lemma phi0_zero : phi0 zero =  1.

Lemma phi0_mono (alpha beta : ON) :
  alpha < beta ->  phi0 alpha < phi0 beta.

Lemma phi0_inj (alpha beta : ON) :
    phi0 alpha = phi0 beta -> alpha = beta.

Lemma phi0_sup : forall (U: Ensemble ON),
   Inhabited _ U ->   countable U ->  phi0 (|_| U) = |_| (image U phi0).

Lemma is_limit_phi0 (alpha : ON) :
  zero < alpha ->  is_limit (phi0 alpha).

Lemma omega_eq : omega = phi0 1.

Lemma phi0_le (alpha : ON) : alpha <= phi0 alpha.
\end{Coqsrc}

Please note that the lemma \texttt{omega\_eq} above, is consistent with the interpretation of the ordering function $\varphi_0$ as the exponential of basis $\omega$. Indeed we could have written this lemma with our alternative notation:

\begin{Coqsrc}
 Lemma omega_eq : omega = omega^ 1.
\end{Coqsrc}

\section{More about \texorpdfstring{$\epsilon_0$}{\texttt{epsilon0}}}

Let us recall that the limit ordinal  $\epsilon_0$ cannot be written within the type \texttt{T1}. Since we are now considering the set of all countable ordinals, we can now prove some properties of this ordinal.

We prove the inequality  $\alpha<\omega^\alpha$ whenever $\alpha < \epsilon_0$.
\emph{Note that this condition was implicit in Module \href{../src/html/hydras.Epsilon0/T1.html\#lt_phi0}%
{\texttt{Epsilon0.T1}}.}

\begin{Coqsrc}
Lemma lt_phi0 (alpha : ON):
  alpha < epsilon0 -> alpha < phi0 alpha.
\end{Coqsrc}

The proof is as follows:
\begin{enumerate}
\item Since $\alpha<\epsilon_0$, consider the least $i$ such that $\alpha$ is strictly less than the omega-tower of height $i$.
\item
  \begin{itemize}
  \item If $i=0$, then the result is trivial (because $\alpha=0$)
 \item  Otherwise let $i=j+1$; 
          $\alpha$ is greater than or equal to the omega-tower of height $j$.
         By monotonicity,  $\varphi_0(\alpha)$ is greater than or equal to 
        the omega-tower of height $j+1$, thus strictly greater than $\alpha$
  \end{itemize}
 \end{enumerate}

Moreover,  $\epsilon_0$ is the least ordinal $\alpha$ that verifies the equality 
$\alpha = \omega^\alpha$, in other words the least fixpoint of the function  $\varphi_0$.

\begin{Coqsrc}
Theorem epsilon0_lfp : least_fixpoint lt phi0 epsilon0.
\end{Coqsrc}

\section{Critical ordinals}

\index{Maths!Critical ordinals}

For any  (countable) ordinal $\alpha$, the set $\textit{Cr}(\alpha)$ is inductively defined 
as follows by Schütte (p.81 of~\cite{schutte}).

\begin{quote}
  \begin{itemize}
  \item $\textit{Cr}(0)$ is the set \textit{AP} of additive principal ordinals.
  \item If $0<\alpha$, then $\textit{Cr}(\alpha)$ is the intersection of all the sets of fixpoints of the $\textit{Cr}(\beta)$ for $\beta<\alpha$.
  \end{itemize}
\end{quote}

This definition is translated in \coq{} in 
Module \href{../src/html/hydras.Schutte.Critical.html}%
{\texttt{Schutte.Critical}}, as the least fixpoint of a functional.

\begin{Coqsrc}
Definition Cr_fun : forall alpha : ON,
       (forall beta : ON, beta < alpha -> Ensemble ON) ->
        Ensemble ON 
:= 
   fun (alpha :ON)
        (Cr : forall beta, 
                beta < alpha -> Ensemble ON) 
        (x : ON) => (
       (alpha = zero /\ AP x) \/
       (zero < alpha /\
        forall beta (H:beta < alpha),
          the_ordering_segment (Cr beta H) x /\ ord (Cr  beta H) x = x)).

Definition Cr (alpha : ON) : Ensemble ON := 
    (Fix  all_ord_acc (fun (_:ON) => Ensemble ON) Cr_fun) alpha.
\end{Coqsrc}

\begin{Coqsrc}
Definition phi (alpha : ON) : ON -> ON 
    :=  ord (Cr alpha).

Definition A (alpha : ON) : Ensemble ON :=
  the_ordering_segment (Cr alpha).
\end{Coqsrc}

For instance,  we prove that $\textit{Cr}(0)$ is the set of additive principals and that $\epsilon_0$
belongs to $\textit{Cr}(1)$.

\begin{Coqsrc}
Lemma Cr_zero_AP :  Cr 0 = AP

Lemma epsilon0_Cr1 : In (Cr 1) epsilon0.
\end{Coqsrc}

\index{Exercises}

\begin{exercise}
 Prove that $\epsilon_0$ is the least element of $\textit{Cr}(1)$.
\end{exercise}

\subsection{A flavor of infinity}

The family of the $\textit{Cr}(\alpha)$s is made of infinitely many unbounded (hence infinite) sets.
Let us quote Lemma 5, p. 82  of~\cite{schutte}:
\begin{quote}
  For all $\alpha$, the set $\textit{Cr}(\alpha)$ is closed (for the least upper bound of non-empty countable sets) and unbounded.
\end{quote}

We prove this result by transfinite induction on $\alpha$ of both properties.

The proof is still quite long, by transfinite induction over $\alpha$.

\index{Coq!Techniques!Simultaneous induction}
\index{Coq!Techniques!Transfinite induction}

\begin{Coqsrc}
Section Proof_of_Lemma5.
  Let P (alpha:ON) := Unbounded (Cr alpha) /\ Closed (Cr alpha).

 Lemma Lemma5 : forall alpha, P alpha.
(* ... *)
 End Proof_of_Lemma5.

Corollary Unbounded_Cr alpha : Unbounded (Cr alpha).
Proof.
  now destruct (Lemma5 alpha).
Qed.

Corollary Closed_Cr alpha : Closed (Cr alpha).
Proof.
  now destruct (Lemma5 alpha).
Qed.
\end{Coqsrc}

\section{Cantor normal form}

The notion of Cantor normal form is defined for all countable ordinals.
Nevertheless, note that, contrary to the implementation based on type \texttt{T1},
the Cantor normal form of an ordinal $\alpha$ may contain $\alpha$ as a 
sub-term\footnote{This would prevent us from trying to represent Cantor normal forms as finite trees (like in Sect.~\ref{sec:T1-inductive-def})}.

We represent  Cantor normal forms as lists of ordinals.
A  list $l$ is a Cantor normal form of a given ordinal $\alpha$ if it satisfies two conditions:

\begin{itemize}
\item The list  $l$ is sorted (in decreasing order) w.r.t. the order $\leq$
\item The sum of all the  $\omega^{\beta_i}$ where the $\beta_i$ are the terms of $l$ (in this order) is equal to $\alpha$.
\end{itemize}

\vspace{4pt}

\noindent\emph{From \href{../src/html/hydras.Schutte.CNF.html\#cnf_t}%
{\texttt{Schutte.CNF}}}

\begin{Coqsrc}
 Definition cnf_t := list ON.

Fixpoint eval (l : cnf_t) : ON :=
  match l with nil => zero
              | beta :: l' => phi0 beta + eval l'
  end.

Definition sorted (l: cnf_t) :=
  LocallySorted (fun alpha beta => beta <= alpha) l.

Definition is_cnf_of (alpha : ON)(l : cnf_t) : Prop :=
  sorted l /\ alpha = eval l.
\end{Coqsrc}

\index{Maths!Transfinite induction}

By transfinite induction on $\alpha$, we prove that every countable ordinal $\alpha$ 
 has at least a Cantor normal form.

\begin{Coqsrc}
Theorem cnf_exists (alpha : ON) :
  exists l: cnf_t, is_cnf_of alpha l.
\end{Coqsrc}

By structural induction on lists, we prove that this normal form is unique.

\begin{Coqsrc}
 Lemma cnf_unicity : forall l alpha, 
   is_cnf_of alpha l -> 
   forall l',  is_cnf_of alpha l' -> l=l'.
Proof.
 induction l.
 (*  ...  *)

Theorem cnf_exists_unique (alpha:ON) :
  exists! l: cnf_t, is_cnf_of alpha l.
\end{Coqsrc}

Finally, the following two lemmas relate  $\epsilon_0$ with Cantor normal forms.

If $\alpha<\epsilon_0$, then the Cantor normal form of $\alpha$ is made of ordinals strictly less than $\alpha$.

\begin{Coqsrc}
Lemma cnf_lt_epsilon0 : 
 forall l alpha, 
   is_cnf_of alpha l ->  alpha < epsilon0 ->
   Forall (fun beta =>  beta < alpha) l.
\end{Coqsrc}

\index{Exercises}

\begin{exercise}
Please consider the following statement :

\begin{Coqsrc}
Lemma cnf_lt_epsilon0_iff : 
 forall l alpha, 
   is_cnf_of alpha l ->  
   (alpha < epsilon0 <->  Forall (fun beta =>  beta < alpha) l).
\end{Coqsrc}

Is it true ?
\end{exercise}

Finally, the Cantor normal form of $\epsilon_0$ is just $\omega^{\epsilon_0}$.

\begin{Coqsrc}
Lemma cnf_of_epsilon0 : is_cnf_of epsilon0 (epsilon0 :: nil).
Proof.
  split.
  - constructor.  
  - simpl;  now rewrite alpha_plus_zero, epsilon0_fxp.
Qed.
\end{Coqsrc}

\index{Projects}

\begin{project}
Implement pages 82 to 85 of~\cite{schutte} (critical, strongly critical, maximal critical ordinals, Feferman's ordinal $\Gamma_0$).
\end{project}

\begin{remark}
The sub-directory \href{../src/html/hydras.Gamma0.html}%
{\texttt{src/Gamma0}} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form. 
\end{remark}

\section{An embedding of \texttt{T1} into \texttt{ON}}

Our library 
\href{../src/html/hydras.Schutte.Injection_from_T1.html}%
{\texttt{Schutte.Injection\_from\_T1.v}} establishes the link between two very different modelizations of ordinal numbers. In other words, it ``validates'' a data structure in terms of
a classical mathematical discourse considered as a model. 
First, we define a function from \texttt{T1} into  \texttt{ON} by structural recursion.

\begin{Coqsrc}
Fixpoint inject (t:T1) : ON :=
 match t with 
     | T1.zero => zero
     | T1.ocons a n b =>  AP.phi0 (inject a) * S n + inject b
 end.  
\end{Coqsrc}

This function enjoys good commutation properties with respect to the main operations which
allow us to build Cantor normal form.

\begin{Coqsrc}
Theorem inject_of_zero : inject T1.zero = zero.

Theorem inject_of_finite (n : nat):
  inject (T1.fin n) =  n.

Theorem inject_of_phi0 (alpha : T1):
  inject (phi0 alpha) = AP.phi0 (inject alpha).

Theorem inject_plus (alpha beta : T1): nf alpha -> nf beta ->
  inject (alpha + beta)%t1 = inject alpha + inject beta.

Theorem inject_mult_n (alpha : T1)  :
  nf alpha -> forall n:nat , inject (alpha *  n)%t1 =  inject alpha * n.

Theorem inject_mono (beta gamma : T1) :
  T1.lt  beta gamma -> 
  T1.nf beta -> T1.nf gamma -> 
  inject beta < inject gamma.

Theorem inject_injective (beta gamma : T1) : nf beta -> nf gamma ->
  inject beta = inject gamma -> beta = gamma.
\end{Coqsrc}

Finally, we prove that \texttt{inject} is a bijection from the set of all terms of \texttt{T1} in normal form to the set 
\texttt{members epsilon0} of the elements of \texttt{ON} strictly less than  $\epsilon_0$.

\begin{Coqsrc}
 Theorem inject_lt_epsilon0 (alpha : T1):
      inject alpha < epsilon0.

Theorem embedding : 
     fun_bijection (nf: Ensemble T1)  (members epsilon0) inject.
 \end{Coqsrc}
\subsection{Remarks}
Let us recall that the library \href{../src/html/hydras.Schutte.Schutte.html}%
{\texttt{Schutte}} depends on five \emph{axioms} and lies explicitly in the  
framework of classical logic with a weak version of the axiom of choice
(please look at the documentation of
\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.ChoiceFacts.html}{\texttt{Coq.Logic.ChoiceFacts}}).
Nevertheless, the other modules:
\href{../src/html/hydras.Epsilon0.Epsilon0.html}%
{\texttt{Epsilon0}},
\href{../src/html/hydras.Hydra.Hydra.html}%
{\texttt{Hydra}}, et 
\href{../src/html/hydras.Gamma0.Gamma0.html}%
{\texttt{Gamma0}}
do not import any axioms and are really constructive.

\section{Related work}

In~\cite{grimm:hal-00911710}, José Grimm establishes the consistency between or ordinal notations (\texttt{T1} and \texttt{T2} (Veblen normal form) and his implementation
of ordinal numbers after Bourbaki's set theory.

### 2020-06-23 latex/chapter-termination.tex
first commit

\chapter[A proof of termination, using epsilon0]{A proof of termination, using ordinals below \texorpdfstring{$\epsilon_0$}{epsilon0}}

\label{cnf-math-def}
\label{chap:proof-term}
\label{chap:T1}

In this chapter, we adapt to \coq{} the well-known~\cite{KP82}  proof that Hercules eventually wins every battle, whichever the strategy  of each player.
In other words, we present  a formal and self contained proof of termination  of all [free] hydra battles.
First, we take from Manolios and Vroon~\cite{Manolios2005} a representation of the ordinal $\epsilon_0$ as terms in Cantor normal form. Then, we define a variant for hydra battles as a measure that maps any hydra to some ordinal strictly less than $\epsilon_0$.

\section{Ordinal numbers}
\index{Maths!Ordinal numbers}

The proof of termination of all hydra battles presented in~\cite{KP82} is based
on \emph{ordinal numbers}.
From a mathematical point of view, an ordinal is a representant of an equivalence class for isomorphims of strict, total and well-founded orders.

We can also associate to every ordinal $\alpha$ a set whose elements are all ordinals strictly less than $\alpha$. Thus, it is meaningful  to consider \emph{finite}, \emph{infinite}, \emph{demunerable} and \emph{non-countable} ordinals.
The relation $<$ on ordinals is well-founded, and the order $\leq$ associated with
$<$ is total.

We cannot cite all the litterature published on ordinals since Cantor's book 
\cite{cantorbook}, and 
leave it to the reader to explore the bibliography. Let us cite the book by Schütte~\cite{schutte} which contains an axiomatic definition of the set of countable ordinals we used as a mathematical specification of our implementaion in \coq{}~\cite{CantorContrib}.

Out of respect of the tradition, the meta-variables for ordinals will be 
 $\alpha$, $\beta$, $\gamma$, etc.

\subsection{Definitions}

\begin{itemize}
\item Let  $\alpha$ be an ordinal; we say that  $\alpha$ is a \emph{successor} if there exists some ordinal  $\beta$ such that 
$\alpha$ is is the least ordinal strictly greater than  $\beta$.

\item We say that an ordinal $\lambda$ is a \emph{limit ordinal} is $\lambda$  is the least upper bound of a stricly increasing sequence of ordinals.
The meta-variable $\lambda$ will be used for denoting  limit ordinals.

\item \index{Maths!Transfinite induction}
An ordinal is either $0$, a limit ordinal or a successor ordinal. This case analysis, as well as \emph{transfinite} ({i.e.}, well-founded) induction is used in many proofs about ordinal numbers.

\item The segment of finite ordinals is isomorphic to the set of natural numbers.
 The first infinite ordinal is the limit ordinal $\omega$.

\item The operations $+$, $\times$ and exponentiation on $\mathbb{N}$ are extended on ordinals numbers. Note that these  extensions are not commutative any more.  For instance $\omega = 1 + \omega \not= \omega + 1$
and $\omega = 2 \times \omega \not= \omega \times 2$.

\item The ordinal $\epsilon_0$ is the least solution of the equation
 \(\alpha=\omega^{\alpha}\).
\end{itemize}

Please note that the set of countable ordinals is not countable.

 Unless otherwise specified, we will only consider ordinals strictly less than  $\epsilon_0$ in this chapter.

% \subsection{Ordinal notations}
% Whenever possible, define a representation of ordinals by finite expressions.
% Of course, since the set of all countable ordinals is not countable, it cannot be represented by a language of finite terms.

% \subsubsection{Notations systems for small ordinals}
% \begin{itemize}
% \item $\omega$
% \item $\omega^2$
% \item $\omega^n$
% \item  $\omega^\omega$
% \end{itemize}

\section{The ordinal \texorpdfstring{\(\epsilon_0\)}{epsilon0}}
\label{sec:epsilon0-intro}

\subsection{Ordinal Notations}

In a proof assistant like \coq{},  it is useful to represent ordinals through some data-type, and make arithmetical operations and comparison effectively implemented  through certified functions.
Our user contribution~\cite{CantorContrib} represents 
the set of ordinals strictly less than $\epsilon_0$ in Cantor normal form as in~\cite{Manolios2005}, and also the set
of ordinals strictly  less than $\Gamma_0$ in Veblen normal form.

The ordinal \(\epsilon_0\) is the least ordinal number that satisfies 
the equation \(\alpha = \omega^\alpha\), where \(\omega\) is 
the least infinite ordinal. Thus, we can consider \(\epsilon_0\) as an
\emph{infinite} \(\omega\)-tower.

\index{Maths!Cantor normal form}

Any ordinal strictly less that \(\epsilon_0\) 
can be represented by a unique  \emph{Cantor normal form}, 
that is, an expression  which is either  the ordinal \(0\) or 
a sum  \(\omega^{\alpha_1} \times n_1 + \omega^{\alpha_2} \times n_2 + 
  \dots + \omega^{\alpha_p} \times n_p\) where all the \(\alpha_i\) 
are ordinals in Cantor  normal form, \(\alpha_1 > \alpha_2 > \alpha_p\), 
and all the \(n_i\) are positive integers.

An example of Cantor normal form is displayed in Fig \ref{fig:cnf-example}:
Note that  any ordinal of
the form \(\omega^0 \times i + 0\) is just written \(i\).

\begin{figure}[htb]
\centering
\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
\node[color=blue]{$\omega^{(\omega^\omega\,+\, \omega^2 \times 8 \,+\, \omega)}+ \omega^\omega + \omega^4+ 6$};
\end{tikzpicture}
\caption{\label{fig:cnf-example}
An ordinal in Cantor normal form}
\end{figure}

In the rest of this section, we define an inductive type for representing in \texttt{Coq}
all the ordinals strictly  less than  \(\epsilon_0\), then extend some arithmetic operations
to this type, and finally prove that our representation fits well with 
the expected mathematical properties: the order we define is a well order, 
and the decomposition into Cantor normal form  is consistent 
with the implementation of the arithmetic operations of exponentiation of base \(\omega\) 
and addition.

\paragraph*{Remark}
\label{sec:orgheadline65}
Unless explicitly mentionned, the term ``ordinal" will be used instead of
``ordinal strictly less than \(\epsilon_0\)" (except in Chapter~\ref{chap:schutte} where it stands for ``countable ordinal'').

\subsection{A data type for  Cantor normal forms}
\label{sec:orgheadline72}
\label{sec:T1-inductive-def}

    Let us define an inductive type whose 
constructors are respectively associated
with the ways to build Cantor normal forms:

\begin{itemize}
\item the ordinal \(0\)
\item the construction \((\alpha,\, n,\,\beta)  \mapsto \omega^\alpha \times (n + 1)+ \beta \quad (n\in\mathbb{N})\)
\end{itemize}

\vspace{4pt}
\noindent\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#T1}{Epsilon0.T1}}

\index{Types!T1@T1 (ordinal terms below $\epsilon_0$)}
\index{Constants!zero:T1}

\begin{Coqsrc}
Inductive T1 : Set  :=
| zero : T1
| ocons : T1 -> nat -> T1 -> T1.
\end{Coqsrc}

\subsubsection*{Remark}
\label{sec:orgheadline66}

The name \texttt{T1} we gave to this data-type  is proper to this development and refers
to a hierarchy of ordinal notations. For instance, in~\cite{CantorContrib}, the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{schutte}).

\index{Types!T2@T2 (ordinals in Veblen normal form)}
\emph{Please look also at sub-directory
href{../src/html/hydras.Gamma0.T2.html}{Gamma0.T2.html}}.

\begin{Coqsrc}
Inductive T2 : Set :=
  zero : T2
| gcons : T2 -> T2 -> nat -> T2 -> T2.
\end{Coqsrc}

Another useful ordinal notation could be used for denoting the ordinals strictly less than $\omega^\omega$, but \texttt{list nat} could be used as well.

\begin{Coqsrc}
Inductive T0 : Set := 
  zero : T0
| cons : nat -> T0 -> T0.
\end{Coqsrc}

\subsubsection{Example}

\label{alpha0-def}
For instance, the ordinal  $\omega^\omega+\omega^3\times 5+2$ is represented by the following term:

\begin{Coqsrc}
Example alpha_0 : T1 :=
  ocons (ocons (ocons zero 0 zero)
               0
               zero)
        0
       (ocons (ocons zero 2 zero)
              4
              (ocons zero 1 zero)).
\end{Coqsrc}

\begin{figure}[htb]
\centering
\begin{tikzpicture}[very thick, scale=0.5, level 1/.style={sibling distance=6cm},
level 2/.style={sibling distance=35mm},  
level 3/.style={sibling distance=17mm}]
\node  {ocons}
  child {  node {ocons}
            child { node {ocons} child {node {zero}} child {node{0}} child{node{zero}}}
         child {node {0}}
         child {node {zero}}}
    child {node {0}}
   child {node {ocons} 
 child { node {ocons} child {node {zero}} child {node{2}} child{node{zero}}}
  child {node {4}}
         child {node {ocons} child {node {zero}} child {node{1}} child{node{zero}}}};

\end{tikzpicture}

\caption{The tree-like representation of the ordinal $\omega^\omega+\omega^3\times 5 +2$\label{fig:cnf-tree}}

\end{figure}

\paragraph{Remark}
For simplicity's sake, we chosed to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the contruction (\texttt{ocons $\alpha$ $n$ $\beta$} is intented to represent the
ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
In a future version, we should replace  the type \texttt{nat} with \texttt{positive} in \texttt{T1}'s 
definition. But this replacement would take a lot of time \dots.

\subsection{Abbreviations}

Some abbreviations may help to write more consisely complex ordinal terms.

\subsubsection{Finite ordinals}
\label{sec:orgheadline67}

For representing finite ordinals, \emph{i.e.} natural numbers, we first introduce a notation for terms of the form $n+1$, then define a coercion from type \texttt{nat} into \texttt{T1}.

\begin{Coqsrc}
Notation "'FS' n" :=
     (ocons zero n zero) (at level 10) : t1_scope.
\end{Coqsrc}

\begin{Coqsrc}
Definition fin (n:nat) : T1 := 
    match n with 0 => zero | S p => FS p end.

Coercion fin  : nat >-> T1.

Example ten : T1 := 10.   
\end{Coqsrc}

\index{Coq!Techniques!Coercions}
\index{Functions!Coercions@Coercions (from nat to ordinal types)}
\begin{remark}
\label{warning:coercions}
The use of coercions like \texttt{fin} allow us to be close to the mathematical tradition where natural numbers are ordinals too.
Nevertheless, it may happen that a goal like \texttt{3 < 5} could be 
interpreted as \texttt{(lt (fin 3) (fin 5))},  depending on the current notation scope.  
When this misinterpretation happens, tactics like \texttt{auto with arith}, \texttt{lia} do not work!
Thus, it is useful to write \texttt{(3 < 5)\%nat}  an inequality between two natural numbers. 
\end{remark}

\subsubsection{The ordinal \(\omega\)}
\label{sec:orgheadline68}

  Since \(\omega\)'s Cantor normal form is
i.e. \(\omega^{\omega^0}\times 1+ 0\), we can define the following abbreviation:

\index{Notations!omega@omega (the least infinite ordinal)}
\begin{Coqsrc}
Notation omega := (ocons (ocons zero 0 zero) 0 zero).
\end{Coqsrc}

Note that \texttt{omega} is not an identifier, thus any tactic like \texttt{unfold omega} would fail.

\subsubsection{The ordinal \(\omega^\alpha\), a.k.a. \(\varphi_0(\alpha)\)}
\label{sec:orgheadline70}

We provide also a notation for ordinals of the form $\omega^\alpha$.

\index{Notations!phi0@phi0 (exponential of base omega)}

\begin{Coqsrc}
Notation "'phi0' alpha" := (ocons alpha 0 zero) (at level 29) : t1_scope.
\end{Coqsrc}

\index{Maths!Additive principal ordinals}

\begin{remark}
\label{sec:orgheadline69}
The name \(\varphi_0\)
   comes from ordinal numbers theory. In~\cite{schutte}, Schütte defines 
$\varphi_0$  as the ordering (\emph{i.e.} enumerating) function of the set  of \emph{additive principal ordinals} \emph{i.e.} strictly positive ordinals $\alpha$ that verify $\forall \beta<\alpha, \beta+\alpha=\alpha$. For Schütte,  $\omega^\alpha$ is just a notation for $\varphi_0(\alpha)$.  See also Chapter~\vref{chap:schutte}.
\end{remark}

\subsubsection{The hierarchy of \(\omega\)-towers:}
\label{sec:orgheadline71}

The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximed by the sequance of $\omega$-towers.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}

\begin{Coqsrc}
Fixpoint omega_tower (height:nat) : T1 := 
 match height with 
 | 0 =>  1 
 | S h => phi0 (omega_tower h)
 end.
\end{Coqsrc}

For instance, Figure~\ref{fig:tower7} represents  the ordinal returned by the
 evaluation of the term \texttt{omega\_tower 7}.

\begin{figure}[htb]
\centering
\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
\node[color=blue]{$\omega^{{{\omega}^{{{\omega}}^{{{\omega}}^{{\omega^{{\omega}^{\omega}}}}}}}}$};
\end{tikzpicture}
\caption{\label{fig:tower7}
The $\omega$-tower of height 7}
\end{figure}

\subsection{Comparison between ordinal terms}
\label{sec:orgheadline73}

% Our formalisation of Cantor Normal Form will take two steps:
% 1 Definition of a strict order \texttt{<} on the type \texttt{T1}, 
% 2 Using \texttt{<} for characterizing terms in normal form.

In order to compare two terms of type \texttt{T1}, we define a recursive function \texttt{compare} that maps two ordinals $\alpha$ and $\beta$ to a value of type \texttt{comparison}. This type is defined in \coq's standard library 
\texttt{Init.Datatypes} and
contains three constructors:  \texttt{Lt} (less than), \texttt{Eq} (equal), and
\texttt{Gt} (greater than).

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#compare}{Epsilon0.T1}}

\begin{Coqsrc}
Fixpoint compare (alpha alpha':T1):comparison :=
  match alpha, alpha' with
    zero, zero => Eq
  | zero, ocons a' n' b' => Lt
  | _   , zero => Gt
  | (ocons a n b),(ocons a' n' b') =>
      (match compare a a' with 
          | Lt => Lt
          | Gt => Gt
          | Eq => (match lt_eq_lt_dec n n'
                   with
                       inleft  (left _) => Lt
                     | inright _ => Gt
                     |   _ => compare b b'
                   end)
       end)
  end.
\end{Coqsrc}

It is now easy to define the boolean predicate \texttt{lt\_b $\alpha$ $\beta$}: 
`` $\alpha$ is strictly less than $\beta$ ''. By coercion to sort \texttt{Prop} we define also the predicate \texttt{lt}.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}

\begin{Coqsrc}
Definition lt_b alpha beta : bool :=
  match compare alpha beta with
      Lt => true
    | _ => false
  end.

Definition lt alpha beta : Prop := lt_b alpha beta.
\end{Coqsrc}

\index{Predicates!lt @ lt (total, strict, non-well founded order on type T1)}
Please note that this definition of \texttt{lt} makes it easy to write proofs by reflection, as shown by the following examples.

\begin{Coqsrc}
Example E1 : lt (ocons omega 56 zero) (tower 3).
Proof. reflexivity. Qed.

Example E2 : ~ lt (tower 3) (tower 3).
Proof.  discriminate.  Qed.
\end{Coqsrc}

The following lemmas establish relations between \texttt{compare}, 
the predicate \texttt{lt} and Leibniz equality \texttt{eq}.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#compare_refl}{Epsilon0.T1}}

\begin{Coqsrc}
Lemma compare_refl : forall alpha, compare alpha alpha =  Eq.
\end{Coqsrc}

\begin{Coqsrc}
Lemma compare_reflect : forall alpha beta,
    match compare alpha beta with
    |   Lt => lt alpha  beta
    |   Eq => alpha = beta
    |   Gt => lt beta  alpha
    end.
\end{Coqsrc}

We prove also that the relation \texttt{lt} is a strict total order.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#lt_irrefl}{Epsilon0.T1}}

\begin{Coqsrc}
Theorem lt_irrefl : forall alpha, ~ lt alpha  alpha.

Theorem lt_trans :
forall alpha beta: T1,
 lt alpha  beta -> 
 forall gamma, lt beta gamma -> lt alpha gamma.

Definition lt_eq_lt_dec  :
   forall alpha beta : T1, 
          {lt alpha  beta} + {alpha = beta} + {lt beta alpha}.
\end{Coqsrc}

Note that the order \texttt{lt} is not reflected 
in the structure (size and/or height) of the terms of \texttt{T1}. 
For instance the ordinal of Fig \ref{fig:cnf-example} is strictly less
than the structurally simpler \(\omega^{\omega^\omega}\times 2\).

\subsubsection{A Predicate for characterizing normal forms}
\label{sec:orgheadline74}
\label{sec:orgheadline75}
We note that our data-type \texttt{T1} allows us to write expressions that
are not properly in Cantor normal form as specified in Section \ref{sec:epsilon0-intro}.
For instance, consider the following term of type  \texttt{T1}.

\begin{Coqbad}
Example bad_term  : T1 := ocons 1 1 (ocons omega 2 zero).
\end{Coqbad}

This term would have been written \(\omega^1\times 2 + \omega^\omega \times 3\) in the usual mathematical notation. We note that the exponents of $\omega$ are not in the right (strictly decreasing) order.

With the help of the order \texttt{lt} on \texttt{T1}, we are now able to characterize
the set of all well-formed ordinal terms:

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#nf_b}{Epsilon0.T1}}

\index{Predicates!nf @ nf (being in Cantor normal form)}

\begin{Coqsrc}
Fixpoint nf_b (alpha : T1) : bool :=
  match alpha with
    | zero => true
    | ocons a n zero => nf_b a
    | ocons a n ((ocons a' n' b') as b) =>
      (nf_b a && nf_b b && lt_b a' a)%bool
  end.

Definition nf alpha :Prop := nf_b alpha.
\end{Coqsrc}

\begin{Coqsrc}
 Compute nf_b alpha_0.
\end{Coqsrc}

\begin{Coqanswer}
   = true 
     : bool
\end{Coqanswer}

\begin{Coqsrc}
 Compute nf_b bad_term.
\end{Coqsrc}

\begin{Coqanswer}
   = false 
     : bool
\end{Coqanswer}

\subsection{Making normality implicit}
  We would like to get rid of terms of type \texttt{T1} which are not in Cantor normal form.
A simple way to do this is to consider statements of the form 
\texttt{forall alpha: T1, nf alpha -> $P$ alpha}, where $P$ is a predicate over type \texttt{T1}, lile in the following lemma.

\begin{Coqsrc}
Lemma plus_is_zero alpha beta :
  nf alpha -> nf beta ->
  alpha + beta  = zero -> alpha = zero /\  beta = zero.
\end{Coqsrc}

But this style leads to clumsy statements, and many sub-goals in interactive proofs (although often solved with \texttt{auto} or \texttt{eauto}).

Another way is encapsulate conditions of the form \texttt{(nf $\alpha$)} in
often used predicates. For instance, we introduced the restriction of \texttt{lt} to terms in normal form, and provided a handy notation for this restriction.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Prelude.Restriction.html}{Ordinals.Prelude.Restriction}}

\begin{Coqsrc}
Definition restrict {A:Type}(E: Ensemble A)(R: relation A) :=
    fun a b => E a /\ R a b /\ E b.
 \end{Coqsrc}

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#LT}{Epsilon0.T1}}

\begin{Coqsrc}
Definition LT := restrict nf lt.
Infix "<" := LT : t1_scope.

Definition LE := restrict nf le.
Notation "alpha <= beta" := (LE alpha beta) : t1_scope.
\end{Coqsrc}

\index{Predicates!LT @LT (partial, well-founded strict order on type T1)}

For instance, in the following lemma, the condition that $\alpha$ is in normal form is included in the condition $\alpha< 1$.

\begin{Coqsrc}
Lemma LT_one : forall alpha, alpha < one -> alpha = zero.
\end{Coqsrc}

\subsection{A first-class type for \texorpdfstring{$\epsilon_0$}{epsilon0}}

In Sect~\vref{sect:equations}, we chosed to define total functions over the set of ordinals less than $\epsilon_0$, using M. Sozeau's plug-in
\texttt{coq-equations}\cite{sozeau:hal-01671777}.

To this end, we define a type for representing terms in Cantor normal form,
as a \coq{} type-class.

\label{sect:E0-def}
\index{Types!E0@ EO (ordinal terms below $\epsilon_0$ in normal form)}

\emph{From Module~\href{../src/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}

\index{Coq!Techniques!Type classes}

\begin{Coqsrc}
Class E0 : Type := t1_2o {cnf : T1; cnf_ok : nf cnf}.
\end{Coqsrc}

Many constructs : types, predicates, functions, notations, etc., on type \texttt{T1} are adapted to \texttt{E0}.

First, we declare a notation scope for \texttt{E0}.

\begin{Coqsrc}
Declare Scope E0_scope.
Delimit Scope E0_scope with e0.
Open Scope E0_scope.
\end{Coqsrc}

Then we redefine the predicates of comparison.

\index{Predicates!Lt @ Lt (Total, strict, well-founded order on type E0)}

\begin{Coqsrc}
Definition Lt (alpha beta : E0) := T1.LT (@cnf alpha) (@cnf beta).
Definition Le (alpha beta : E0) := T1.LE (@cnf alpha) (@cnf beta).

Infix "<" := Lt : E0_scope.
Infix "<=" := Le : E0_scope.
\end{Coqsrc}

Equality in \texttt{E0} is just Leibniz equality. Note that, since \texttt{nf} is
defined by a Boolean function, for  any term $\alpha:\texttt{T1}$, there exists at most one proof of \texttt{nf $\alpha$}, thus two ordinals of type \texttt{E0} are
equal if and only iff their projection to \texttt{T1} are equal.

\index{Coq!Techniques!Unicity of  equality proofs}

\begin{Coqsrc}
Require Import Logic.Eqdep_dec.

Lemma nf_proof_unicity :
  forall (alpha:T1) (H H': nf alpha), H = H'.

Lemma E0_eq_iff alpha beta : alpha = beta <-> cnf alpha = cnf beta.
\end{Coqsrc}

For upgrading constants and fonctions of \texttt{T1}, we have to prove that 
the term they build is in normal form.
For instance, let us represent the ordinals $0$ $\omega$ as instances of the class \texttt{E0}.

\index{Notations!omega@omega (the least infinite ordinal)}
\index{Constants!zero:T1}

\begin{Coqsrc}
Instance Zero : E0.
Proof.
  refine (@mkord T1.zero _);  now compute. 
Defined.

Instance _Omega : E0.
Proof.  now exists omega. Defined.

Notation "'omega'"  := _Omega : E0_scope.
\end{Coqsrc}

 Likewise, we define an addition in type \texttt{E0}, using the function{T1.plus}
and its properties.

\begin{Coqsrc}
Search   (nf (_ + _)%t1).
\end{Coqsrc}

\begin{Coqanswer}
plus_nf: forall a : T1, nf a -> forall b : T1, nf b -> nf (a + b)
\end{Coqanswer}

\begin{Coqsrc}
Instance plus (alpha beta : E0) : E0.
Proof.
  refine (@mkord (T1.plus (@cnf alpha) (@cnf beta))_ );
    apply plus_nf; apply cnf_ok.
Defined.

Infix "+" := plus : E0_scope.

Check omega + omega.
\end{Coqsrc}

\begin{Coqanswer}
omega + omega
     : E0
\end{Coqanswer}

\subsection{Syntactic definition of limit and successor ordinals}

Pattern matching and structural recursion allow us to define the notions of successor and limit ordinal with the help of  boolean functions on type \texttt{T1}.

 \vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#is_succ}{Epsilon0.T1}}

\begin{Coqsrc}
  Fixpoint is_succ alpha :=
  match alpha with
      zero => false
    | ocons zero _ _ => true
    | ocons alpha n beta => is_succ beta
  end.

Fixpoint is_limit alpha :=
  match alpha with
      zero => false
    | ocons zero _ _ => false
    | ocons alpha n zero => true
    | ocons alpha n beta => is_limit beta
  end.
\end{Coqsrc}

\begin{Coqsrc}
  Compute is_limit omega.
\end{Coqsrc}

\begin{Coqanswer}
  = true
     : bool
\end{Coqanswer}

\begin{Coqsrc}
Compute is_succ 42.
\end{Coqsrc}

\begin{Coqanswer}
  = true
     : bool
   \end{Coqanswer}

The correctness of these definitions with respect to the mathematical notions of
limit and successor ordinals is established through several lemmas. For instance,
Lemma \texttt{canonS\_limit}, page~\pageref{lemma:canonS-limit}, shows that
if $\alpha$ is (syntactically) a limit ordinal, then it is the least upper bound of
a strictly increasing sequence of ordinals.

   The following function is very useful in constructions by cases (proofs and function definitions).

\begin{Coqsrc}
  Definition zero_succ_limit (alpha: T1) :
  {is_succ alpha} + {is_limit alpha} +  {alpha=zero}.
  (* ... *)
Defined.
 \end{Coqsrc}

\subsection{Arithmetic on \texorpdfstring{$\epsilon_0$}{epsilon0}}
\subsubsection{Successor}

The successor of any ordinal $\alpha< \epsilon_0$ is defined by structural 
recursion on its Cantor normal form.

\index{Functions!T1.succ}

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}

\begin{Coqsrc}
Fixpoint succ (alpha:T1) : T1 :=
  match alpha with zero => 1
             | ocons zero n _ => ocons zero (S n) zero
             | ocons beta n gamma => ocons beta n (succ gamma)
 end.
\end{Coqsrc}

The following lemma establishes the connection between the function
\texttt{succ} and the predicate \texttt{is\_succ}.

\begin{Coqsrc}
 Lemma is_succ_iff alpha (Halpha : nf alpha) :
  is_succ alpha <-> exists beta : T1, nf beta /\ alpha = succ  beta.
\end{Coqsrc}

\subsubsection{Addition and multiplication}

Ordinal addition and multiplication are also defined by structural recursion over the type \texttt{T1}. Please note that they use the \texttt{compare} function on some subterms of their arguments.

\begin{Coqsrc}
Fixpoint plus (alpha beta : T1) : T1 :=
  match alpha,beta with
 |  zero, y  => y
 |  x, zero  => x
 |  ocons a n b, ocons a' n' b' =>
    (match compare a a' with
     | Lt => ocons a' n' b'
     | Gt => (ocons a n (plus b (ocons a' n' b')))
     | Eq  => (ocons a (S(n+n')) b')
     end)
  end
where "alpha + beta" := (plus alpha beta) : t1_scope.
\end{Coqsrc}

\begin{Coqsrc}
Fixpoint mult (alpha beta : T1) :T1 :=
  match alpha,beta with
 |  zero, y  => zero
 |  x, zero => zero
 |  ocons zero n _, ocons zero n' _ => 
                 ocons zero (Peano.pred((S n) * (S n'))) zero
 |  ocons a n b, ocons zero n' b' =>  
                 ocons a (Peano.pred((S n) * (S n'))) b
 |  ocons a n b, ocons a' n' b' =>
     ocons (a + a') n' ((ocons a n b) * b')
 end
where  "alpha * beta" := (mult alpha beta) : t1_scope.
\end{Coqsrc}

\subsubsection{Examples}

The following examples are instances of \emph{proofs by computation}. Please note that  addition and multiplication on \texttt{T1}
are not commutative. Moreover,  both operations fail to be strictly monotonous in their first argument.

\begin{Coqsrc}
Example e2 : 6 + omega = omega.
Proof. reflexivity. Qed.

Example e'2 : omega < omega + 6.
Proof. now compute. Qed.

Example e''2 : 6 * omega = omega.
Proof. reflexivity. Qed.

Example e'''2 : omega < omega * 6.
Proof. now compute. Qed.
\end{Coqsrc}

\begin{Coqsrc}
Lemma plus_not_monotonous : exists alpha beta gamma : T1,
    alpha < beta /\ alpha + gamma = beta + gamma.
Proof.
  exists 3, 5, omega;  now  compute.
Qed.

Lemma mult_not_monotonous :  exists alpha beta gamma : T1,
      alpha < beta /\ alpha * gamma = beta * gamma.
Proof.
  exists 3, 5, omega; now compute.
Qed.
\end{Coqsrc}

\subsection{Pretty printing Ordinals in Cantor normal Form}

Let us consider again the ordinal $\alpha_0$ defined in section~\vref{alpha0-def}

If we ask \coq{} to print the normal form of \texttt{alpha\_0}, we get a hardly readable term of type \texttt{T1}.

\begin{Coqsrc}
Compute alpha_0.
\end{Coqsrc}

\begin{Coqanswer}
  = ocons omega 0 (ocons (FS 2) 4 (FS 1))
     : T1
\end{Coqanswer}

The following data type defines a abstract syntax fore more readable ordinals in Cantor normal form:

\index{Types!ppT1@ppT1 (pretty printed ordinal terms)}
\index{Functions!pp@ pp (pretty printing terms in Cantor normal form)}

\begin{Coqsrc}
Inductive ppT1 : Set :=
    P_fin : nat -> ppT1
  | P_add : ppT1 -> ppT1 -> ppT1
  | P_mult : ppT1 -> nat -> ppT1
  | P_exp : ppT1 -> ppT1 -> ppT1
  | P_omega : ppT1
\end{Coqsrc}

The function \texttt{pp: T1 -> ppT1} converts any closed term of type \texttt{T1} into a human-readable expression. For instance, let us convert the term \texttt{alpha\_0} of~\pageref{alpha0-def}.

\begin{Coqsrc}
Compute pp alpha_0.
\end{Coqsrc}

\begin{Coqanswer}
     = (omega ^ omega + omega ^ 3 * 5 + 2)%pT1
     : ppT1
\end{Coqanswer}

\section{Well-foundedness and transfinite induction}

\index{Maths!Transfinite induction}

\subsection{About  well-foundedness}
\label{sec:orgheadline82}
   In order to use \texttt{T1} for proving termination results,
we need to prove that  our representation of ordinals strictly less than \(\epsilon_0\)
makes our order \texttt{<} well-founded. Then we will get \emph{transfinite induction} for free.

The proof of well-foundedness of the strict order $<$ on Cantor normal forms is already 
available in the Cantor contribution by Castéran and Contéjean~\cite{CantorContrib}. That proof relies on a library on recursive path orderings written by
E. Contéjean. We present also a direct proof of the same result, which does not require any knowledge on r.p.o.s.

\index{Exercises}

\begin{exercise}
Prove that the \emph{total} order \texttt{lt} on \texttt{T1} is not well-founded. 
\textbf{Hint:}  You will have to build a counter-example with terms of type \texttt{T1}
which are not in Cantor normal form.
\end{exercise}

% \subsubsection{The total order \texttt{lt} on \texttt{T1} is \emph{not} well-founded}
% \label{sec:orgheadline76}

% Let us recall that the data type \texttt{T1} contains too many inhabitants, including
% terms which are not in Cantor normal form. Thus, the following result is not 
% very surprising.

% \begin{Coqsrc}
% Section lt_not_well_founded.

%   (* let us build the sequence of terms :
%         omega + omega + .... + omega ^ 2   *)
%   Let f := (fix f (i:nat): T1 :=
%             match i with 0 => phi0 2
%                        | S i => ocons 1 1 (f  i)
%             end).

%  Lemma  f_decreases : forall i, f (S i) <  f i.
%  Proof.
%   induction i; compute; auto with T1.
%  Qed.

%  Theorem lt_not_wf : ~  well_founded lt.
%  Proof. 
%    intro wf; case (not_decreasing _ lt);auto.
%    exists f; apply f_decreases.
%  Qed.

% End lt_not_well_founded.
% \end{Coqsrc}

% Thus, we have

\subsubsection{A first attempt}
\label{sec:orgheadline77}
\index{Coq!Techniques!Well founded relations}

It is natural to try to prove by structural induction over \texttt{T1} 
that every term in normal form is accessible through \texttt{LT}.

Unfortunately, it won't work. Let us consider some well-formed term
 $\alpha=\texttt{ocons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are accessible
 through \texttt{LT}. For proving the accessibility of $\alpha$, we have to consider
any well formed term \(\delta\) such that \(\delta<\alpha\). 
But nothing guarantees that \(\delta\)  is strictly  less than \(\beta\) nor \(\gamma\), and we cannot use the induction hypotheses on   \(\beta\) nor \(\gamma\).

\begin{Coqbad}
Section First_attempt.

 Lemma wf_LT : forall alpha,  nf alpha -> Acc LT alpha. 
 Proof.
  induction alpha as [| beta IHbeta n gamma IHgamma].
  - split.
    inversion 1.
    destruct H2 as [H3 _];not_neg H3.
  -  split; intros delta Hdelta.
\end{Coqbad}

\begin{Coqanswer}
1 subgoal (ID 560)

  beta : T1
  n : nat
  gamma : T1
  IHbeta : nf beta -> Acc LT beta
  IHgamma : nf gamma -> Acc LT gamma
  H : nf (ocons beta n gamma)
  delta : T1
  Hdelta : delta < ocons beta n gamma
  ============================
  Acc LT delta
 \end{Coqanswer}

\begin{Coqbad}
  Abort.
\end{Coqbad}

The problem comes from that \(\delta\) may be bigger that \(\beta\) or
\(\gamma\);
for instance \(\delta\) may be of the form
\texttt{ocons $\beta'$ $p'$  $\gamma'$},
where  \(\beta' \leq  \beta\) and  \(p' < n\).
Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.

\subsubsection{Using a stronger inductive predicate.}
\label{sec:orgheadline78}
  Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
through \texttt{LT}, we propose to show first that any well formed 
term of the form \(\omega^\alpha\times(n+1)+\beta\) is accessible (which is a stronger result).

\begin{Coqsrc}
 Let Acc_strong (alpha:T1) :=
      forall n beta, 
        nf (ocons alpha n beta) -> Acc LT (ocons alpha  n beta).
\end{Coqsrc}

The following lemma is an application of the strict inequality 
\showmath{\alpha < \omega ^\alpha}. If \showmath{\omega^\alpha} is accessible,
then \showmath{\alpha} is \emph{a fortori} accessible.

\begin{Coqsrc}
 Lemma Acc_strong_stronger : forall alpha, 
     nf alpha -> Acc_strong  alpha -> Acc LT  alpha.
 Proof.
  intros alpha H H0; apply acc_imp with (phi0 alpha).
  - repeat split; trivial.
    + now apply lt_a_phi0_a.
  -  apply H0;  now apply single_nf.
Qed.
\end{Coqsrc}

Thus, it remains to prove that every ordinal strictly less than \showmath{\epsilon_0} 
is strongly accessible.

% \subsubsection{Structure of the proof of well-foundedness of \texttt{LT}}

\label{sec:orgheadline81}
\label{proof-wf-epsilon0}
\paragraph{A helper}
\label{sec:orgheadline79}

First, we prove that, for  any \texttt{LT}-accessible term \showmath{\alpha}, any well formed
term (\texttt{ocons $\alpha$ $n$ $\beta$})  is accessible too.

\begin{Coqsrc}
Lemma Acc_implies_Acc_strong : 
   forall alpha, Acc LT  alpha -> Acc_strong alpha.
\end{Coqsrc}

The proof is structured as an induction on \showmath{\alpha}'s accessibility. Let us consider
an accessible term $\alpha$.

\begin{Coqanswer}
  subgoal 1

  alpha : T1
  Aalpha : forall y : T1,  y < alpha -> Acc LT y
  IHalpha : forall y : T1,
       LT y alpha ->
       forall (n : nat) (beta : T1),
       nf (ocons y n beta) -> Acc LT (ocons y n beta)
  ============================
   forall (n : nat) (beta : T1),
   nf (ocons alpha n beta) -> Acc LT (ocons alpha n beta)
\end{Coqanswer}

Let \texttt{n:nat} and \texttt{beta:T1} such that \texttt{ocons alpha n beta} is in normal form. 
We prove first that \texttt{beta} is accessible,  which allows us to by well-founded induction on \texttt{beta}, 
and natural induction on \texttt{n}, that (\texttt{ocons alpha n beta}) is accessible.
The proof, quite long, can be consulted in \url{../src/html/hydras.Epsilon0.T1.html}

\paragraph{Accessibility of any well-formed ordinal term}
\label{sec:orgheadline80}

Our goal is still to prove accessibility of any well formed ordinal term.
Thanks to our previous lemmas, we are almost done.

\begin{Coqsrc}
(* A (last) structural induction *)

Theorem nf_Acc : forall alpha, nf alpha -> Acc LT  alpha.
Proof.
 induction alpha.
-  intro; apply Acc_zero.
 -  intros; eapply Acc_implies_Acc_strong;auto.
    apply IHalpha1;eauto.
    apply nf_inv1 in H; auto. 
Defined.

Corollary T1_wf : well_founded LT.
\end{Coqsrc}

And we have now our transfinite recursor:
\index{Maths!Transfinite induction}

\begin{Coqsrc}

Definition transfinite_recursor :
 forall (P:T1 -> Type),
   (forall x:T1, 
     (forall y:T1, nf x -> nf y ->  lt y  x -> P y) -> P x) ->
    forall alpha:T1, P alpha.
Proof.
 intros; apply well_founded_induction_type with LT.
 -  exact T1_wf;auto.
 - intros. apply X. intros; apply X0. repeat split;auto. 
Defined.
\end{Coqsrc}

The following tactic tries to do a  transfinite induction on any ordinal \mathcolor{$\alpha<\epsilon_0$}.

\begin{Coqsrc}
Ltac transfinite_induction alpha :=
  pattern alpha; apply transfinite_recursor;[ | try assumption].
\end{Coqsrc}

\begin{remark}
The proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering is available in the module \href{../src/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
 \end{remark}

\section{A variant for hydra battles}

\subsection{Natural sum (a.k.a. Hessenberg's  sum)}
\label{sec:orgheadline87}
\label{hydra-variant}

Natural sum (Hessenberg's  sum) is a commutative and monotonous version of
addition. It is used as an auxiliary operation  for defining variants
for hydra battles, where Hercules is allowed to chop off any  head of the hydra.

In the litterature, the natural sum of ordinals \(\alpha\) and \(\beta\) 
is often denoted by \(\alpha \# \beta\)  or  \(\alpha \oplus  \beta\).
Thus we called \texttt{oplus} the associated \emph{Coq} function.

\subsubsection{Definition of \texttt{oplus}}
\label{sec:orgheadline84}
\index{Functions!oplus @ oplus (Hessenberg commutative sum)}

The definition of \texttt{oplus} is recursive in both of its 
arguments, which makes a structural recursive definition a little 
complex.
We used the same pattern as for the \texttt{merge} function on lists of library
\texttt{Coq.Sorting.Mergesort}.

\begin{enumerate}
\item Define a nested recursive function, using the \texttt{Fix} 
    construct

\item Build a principle of induction dedicated to \texttt{oplus}

\item Establish equations associated to each case of the definition.
\end{enumerate}

\paragraph{Nested recursive definition}
\label{sec:orgheadline83}

The following definition is composed of 
\begin{itemize}
\item A main function \texttt{oplus}, structurally recursive in its 
first argument \texttt{alpha}
\item An auxiliary function \texttt{oplus\_aux} within the scope of \texttt{alpha},
structurally recursive in its argument \texttt{beta};  \texttt{oplus\_aux beta} 
   is supposed to compute  \texttt{oplus alpha beta}.
\end{itemize}

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}

\begin{Coqsrc}
Fixpoint oplus (alpha beta : T1) : T1 :=
  let fix oplus_aux beta {struct beta} :=
      match alpha, beta with
        | zero, _ => beta
        | _,  zero => alpha
        | ocons a1 n1 b1, ocons a2 n2 b2 =>
          match compare a1 a2 with
            |  Gt => ocons a1 n1 (oplus b1 beta)
            |  Lt => ocons a2 n2 (oplus_aux b2)
            |  Eq => ocons a1 (S (n1 + n2)%nat) (oplus b1 b2)
          end
      end
  in oplus_aux beta.

Infix "o+" := oplus  (at level 50, left associativity).
\end{Coqsrc}

The reader will note that each recursive call of the functions
\texttt{oplus} and \texttt{oplus\_aux} satisfies \emph{Coq}'s constraint
on recursive definitions. The function \texttt{oplus} is recursively called on a sub-term of its first argument,
and \texttt{oplus\_aux} on a sub-term of its unique argument.
Thus, \texttt{oplus}'s definition is accepted by \coq{} as a structurally recursive function.

\subsubsection{Rewriting lemmas}
\label{sec:orgheadline86}

\emph{Coq}'s constraints on recursive definitions resulted in 
the quite  complex form of \texttt{oplus}'s definition.
For making easier proof of properties of this function, 
it is helpful to \emph{derive} lemmas that will help to simplify 
expressions of the form (\texttt{oplus $a$ $ b$}).

A first set of lemmas correspond to the various cases of \texttt{oplus}'s 
definition. They can be proved almost immediately, using \texttt{cbn} 
and \texttt{rewrite} tactics.

\begin{Coqsrc}
Lemma oplus_alpha_0 (alpha : T1) : alpha o+ zero = alpha.
Proof.
  destruct alpha; reflexivity.
Qed.

Lemma oplus_0_beta (beta : T1): zero o+ beta = beta.
Proof.
  destruct beta; reflexivity.
Qed.
\end{Coqsrc}

\subsubsection{A hand-made induction principle}
\label{sec:orgheadline85}

\index{Coq!Commands!Functional Scheme}

\emph{Coq} contains a command  \texttt{Functional Scheme} that 
generates induction principles which correspond to recursive functions.
Unfortunately, the current version ( \texttt{8.11.0} ) doesn't work on \texttt{oplus},
probably because of the inner \texttt{Fix}.

\begin{Coqsrc}
Functional Scheme oplus_ind := Induction for oplus Sort Prop.
\end{Coqsrc}

\begin{Coqanswer}
Error: Anomaly "todo." Please report at http://coq.inria.fr/bugs/.
\end{Coqanswer}

Fortunately, it's a good exercise for a semi-experienced user, to write
her/him-self induction principles similar to the ones returned by
\texttt{Functional Scheme}.

\begin{itemize}
\item First, we chose to write a version for sort \texttt{Type}, since versions
for sorts \texttt{Prop} and \texttt{Set} can be easily derived from
the former one. According to \emph{Coq}'s naming politics, we will call our 
principle \texttt{oplus\_rect}

\item The conclusion of \texttt{oplus\_rect} will be (\texttt{$P$ a b (oplus a b)}),
where $P$ is an arbitrary function of type 
\texttt{T1 -> T1 -> T1 -> Type}

\item The premises of \texttt{oplus\_rect} will describe how to build an induction 
on the graph of \texttt{oplus}.
\end{itemize}

We are now ready to state and prove \texttt{oplus\_rect}, and the reader
will note that the statement is longer than the proof script itself,
which is a standard proof by induction, simplification and case-analysis 
that follows  \texttt{oplus}'s definition.

We associate also a tactic to the application of \texttt{oplus\_rect}.

\begin{Coqsrc}
 Lemma oplus_rect:
      forall P: T1 -> T1 -> T1 -> Type, 
        (forall a:T1, P zero a a) ->
        (forall a: T1, P a zero a) ->
        (forall a1 n1 b1 a2 n2 b2 o,
           compare a1 a2 = Gt ->
           P b1 (ocons a2 n2 b2) o ->
           P (ocons a1 n1 b1) (ocons a2 n2 b2)
             (ocons a1 n1 o)) ->
        (forall a1 n1 b1 a2 n2 b2 o,
           compare a1 a2 = Lt ->
           P (ocons a1 n1 b1) b2 o ->
           P (ocons a1 n1 b1) (ocons a2 n2 b2) 
           (ocons a2 n2 o)) ->
        (forall a1 n1 b1 a2 n2 b2 o,
           compare a1 a2 = Eq ->
           P b1 b2 o ->
          P (ocons a1 n1 b1) (ocons a2 n2 b2)
            (ocons a1 (S (n1 + n2)%nat) o)) ->
         forall a b, P a b (oplus a b).
Proof with auto.
   induction a.
   -    intro; simpl; destruct b;auto.
   -   induction b.
       + apply X0.
       + case_eq (compare a1 b1).
         * intro Comp; unfold oplus; rewrite Comp.
           cbn; apply X3 ...
         * intro Comp; cbn; rewrite Comp; apply X2...
         * intro Comp; cbn; rewrite Comp ...
 Defined.

Ltac oplus_induction a b:= pattern (oplus a b); apply oplus_rect.
\end{Coqsrc}

\index{Exercises}

\begin{exercise}
The induction principle \texttt{oplus\_rect} is still unused in our development. 
Please build some nice examples of application.
\end{exercise}

\subsection{More theorems on Hessenberg's sum}

We need to prove some properties of $\oplus$, particularly about 
its relation with the order $<$ on \texttt{T1}.

\subsubsection{Boundedness}
If $\alpha$ and $\beta$ are both strictly  less than  $\omega^\gamma$, then so is their natural sum
$\alpha \oplus \beta$. This result can be proved by structural induction on $\gamma$.

\begin{Coqsrc}
Lemma lt_phi0_oplus : forall gamma alpha beta,
                        lt_phi0 alpha gamma ->
                        lt_phi0 beta gamma ->
                        lt_phi0 (alpha o+ beta) gamma.

Proof with auto.
  induction gamma; destruct alpha, beta.  
(* ... *)
\end{Coqsrc}

\subsubsection{Commutativity, associativity}

We prove  the commutativity of $\oplus$ in two steps.

First, we prove by transfinite induction on $\alpha$ that the restriction of $\oplus$ to the
interval $[0..\alpha[$ is commutative.

\index{Maths!Transfinite induction}

\begin{Coqsrc}
Lemma oplus_comm_0 : forall alpha, nf alpha ->
     forall a b,  nf a -> nf b ->
                  lt a alpha ->
                  lt b alpha ->
                  a o+ b = b o+ a.
 Proof with eauto with T1.
    intros alpha Halph; transfinite_induction alpha.
(* rest of proof omitted *)  
\end{Coqsrc}

Then, we infer  $\oplus$'s commutativity for any pair of ordinals:
Let $\alpha$ and $\beta$ be two ordinals strictly less than $\epsilon_0$. Both ordinals $\alpha$ and $\beta$ are
strictly less than $\textrm{max}(\alpha,\beta)+1$.

    Thus, we have just to apply the lemma \coqsimple{oplus\_comm\_0}.

\begin{Coqsrc}
  Lemma oplus_comm : forall alpha beta, 
      nf alpha -> nf beta ->
      alpha o+ beta =  beta o+ alpha.
  Proof with eauto with T1.
    intros alpha beta Halpha Hbeta;
    apply oplus_comm_0 with (succ (max alpha beta)) ...  
  (* rest of proof omitted *)
\end{Coqsrc}

The associativity of Hessenberg's sum is proved the same way.

\begin{Coqsrc}
 Lemma oplus_assoc_0 :
    forall alpha,
      nf alpha ->
      forall a b c,  nf a -> nf b -> nf c ->
                      lt a alpha ->
                      lt b alpha -> lt c alpha ->
                      a o+ (b o+ c) = (a o+ b) o+ c.
  Proof with eauto with T1.
    intros alpha Halpha.
    transfinite_induction alpha.
(* rest of proof omitted *) 
\end{Coqsrc}

\begin{Coqsrc}
 Lemma oplus_assoc : forall alpha beta gamma,
                        nf alpha -> nf beta -> nf gamma ->
                                    alpha o+ (beta o+ gamma) =
                                    alpha o+ beta o+ gamma.
 Proof with eauto with T1.
    intros;
    apply oplus_assoc_0 with (succ (max alpha (max beta gamma))) ...
(* rest of proof omitted *)   
\end{Coqsrc}

\subsubsection{Monotonicity}

At last, we prove that $\oplus$ is strictly monotonous in both of its arguments.

\begin{Coqsrc}
Lemma oplus_strict_mono_LT_l (alpha beta gamma : T1) :
  nf gamma   -> alpha  < beta ->
  alpha o+ gamma < beta o+ gamma.

Lemma oplus_strict_mono_LT_r (alpha beta gamma : T1) :
  nf alpha -> beta < gamma ->
  alpha o+ beta < alpha o+ gamma.
\end{Coqsrc}

\subsection{A measure for hydra-battle termination}
\label{sec:orgheadline90}
\label{sec:hydra-measure}

Let us define a measure from type \texttt{Hydra} into \texttt{T1}.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Termination.html\#m}{Hydra.Hydra\_Termination}}

\begin{Coqsrc}
Fixpoint m (h:Hydra) : T1 :=
  match h with head => zero
             | node hs => ms hs
end 
with ms (s:Hydrae) :  T1 :=
  match s with  hnil => zero
              | hcons h s' => phi0 (m h) o+  ms s'
 end.  
\end{Coqsrc}

First, we prove that the measure $m(h)$  of any hydra $h$ is a well-formed ordinal term of type \texttt{T1}.

\begin{Coqsrc}
Theorem m_nf : forall h, nf (m h).
Proof.
 intro h; elim h using Hydra_rect2 
            with (P0 := fun s =>  nf (ms s)).
 (* ... *)

Theorem ms_nf : forall s, nf (ms s).
Proof with auto with T1.
(* ... *)
\end{Coqsrc}

For proving the termination of all hydra battles, we have to prove that
\texttt{m} is a variant. First, a few technical lemmas follow the decomposition of \texttt{round} into several relations. Then the lemma \texttt{round\_decr} gathers all the cases.

\begin{Coqsrc}
Lemma S0_decr :
  forall s s', S0  s s' -> ms s' < ms s.
\end{Coqsrc}

\begin{Coqsrc}
Lemma R1_decr : forall h h',
                  R1 h h' -> m h' < m h.
\end{Coqsrc}

\begin{Coqsrc}
Lemma S1_decr n:
  forall s s', S1 n s s' -> ms s' <  ms s.
\end{Coqsrc}

\begin{Coqsrc}
Lemma R2_decr n : forall h h', R2 n h h' -> m h'  < m h.
\end{Coqsrc}

\begin{Coqsrc}
Lemma round_decr : forall h h', h -1-> h' -> m h' < m h.
Proof.
   destruct 1 as [n H]; destruct H. 
   -  now apply R1_decr.
   -  now apply R2_decr with n.
Qed.
\end{Coqsrc}

Finally, we prove the termination of all (free) battles.

\begin{Coqsrc}
Global Instance HVariant : Hvariant lt_wf free var.
Proof.
 split; intros; eapply round_decr; eauto.
Qed.

Theorem every_battle_terminates: Termination.
Proof. 
  red; apply Inclusion.wf_incl with 
         (R2 := fun h h' =>  m h < m h').
   red; intros;  now apply round_decr.
   apply Inverse_Image.wf_inverse_image, T1_wf.
Qed.
\end{Coqsrc}

\section*{Conclusion}

Let us recall three results we have proved so far.
\begin{itemize}
\item There exists a strictly decreasing variant mapping \texttt{Hydra} into 
$[0,\epsilon_0[$ for proving the termination of any hydra battle
\item There exists \emph{no} such variant from \texttt{Hydra} into $[0,\omega^2[$ and
a fortiori into $[0,\omega[$.
\end{itemize}

So, a  natural question is `` Does there exist any strictly decreasing variant mapping
type \texttt{Hydra} into some interval $[0,\alpha[$ (where $\alpha <\epsilon_0$) for proving the termination of all hydra battles''. The next chapter is dedicated to a formal proof that there exists no such $\alpha$, even if we consider a restriction to the set of ``standard'' battles.

%\include{epsilon0}

%\include{impossibility-proofs}

### 2020-06-23 latex/hydras.tex
first commit

\documentclass{book}
\usepackage{alltt}
\usepackage{makeidx}
\usepackage{hyperref}
\usepackage{url}
\usepackage{mdframed}
\usepackage{amsmath}
\usepackage{varioref}
\usepackage{synttree, proof, centernot}
\usepackage[firstpage]{draftwatermark}
\usepackage{verbatim}
\usepackage[note]{marginote}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{dsfont}
\setcounter{tocdepth}{1}
\input{coqmacros}

\author{Pierre Castéran\\ LaBRI, Univ. Bordeaux, CNRS UMR 5800, Bordeaux-INP}
\date{\today}
\title{The hydra and the rooster}
\makeindex
\begin{document}
\maketitle

\cleardoublepage

% \mbox{~}
% \clearpage
% \newpage
% \mbox{~}
\vspace{10cm}
\begin{quote} 
{ \Large {\it
``I start from one point and go as far as possible.'' }

John Coltrane.}
\end{quote}

\tableofcontents

\include{introduction}
\include{chapter-hydras}
\include{chapter-termination}
\include{chapter-ks}
\include{chapter-large-sets}
\include{chapter-schutte}

\chapter{Appendices}

\bibliographystyle{alpha}
\bibliography{thebib, thebib2}

\section{Future work (projects)}
This document and the proof scripts are far from being complete.

First, there must be a lot of typos to correct, references and index items to add. Many proofs are too complex and should be simplified, etc.

The following extensions are planned, but help is needed:

\begin{itemize}
\item Semi automatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
\item Extension to $\Gamma_0$ (in Veblen normal form)
\item More lemmas about hierarchies of rapidly growing functions, and their relationship 
    with primitive recursive functions and provability in Peano arithmetic 
(following~\cite{KS81, KP82}).
\item From \coq's point of view, this development could be used as an illustration of the evolution of the software, every time new libraries and sets of tactics could help to simplify the proofs.
\end{itemize}

\section{How to install the libraries}
\label{sec:orgheadline4}
\begin{itemize}
\item The present distribution has been checked with version 8.11.0 of the Coq proof assistant, with the plug-ins \texttt{coq-paramcoq} and \texttt{coq-equations}.

\item just go into the \texttt{src} directory, and type "\texttt{make}"
\end{itemize}

\section{Index}
\printindex

% \section{Main notations}

% \subsubsection*{Intervals}

% Let $\leq$ be a total order on a type $A$. The intervals $\{k\in A| i\leq k\leq j\}$ is written $[i,j]$. Likewise, $\{k\in A| i\leq k<j\}$ is written
% $[i,j[$.

\end{document}

### 2020-06-23 latex/introduction.tex
first commit

\chapter{Introduction}

\vspace{16pt}

Proof assistants are excellent tools for exploring the structure of mathematical proofs,
studying  which hypotheses are really needed, and which proof patterns are useful and/or
necessary. Since the development of a theory is represented as a bunch of computer files,
everyone is able to read the proofs with an arbitrary level of detail, or to play with the theory by building alternate proofs or definitions.

Among all the theorems proved with the help of proof assistants like \coq{}, \isabelle{}, \hol{}, etc.,
several statements and proofs  share some interesting features:
\begin{itemize}
\item Their statements are easy to understand, even by non-mathematicians
\item Their proof requires some non-trivial mathematical tools
\item Their mechanization on computer presents some methodological interest.
\end{itemize}

This is obviously the case of the four-color theorem~\cite{fourcolors}  and the Kepler conjecture~\cite{flyspeck2015}. We do not mention impressive works like the proof of the odd-order theorem ~\cite{oddorderthm}, since understanding its statement requires some mathematical culture.

Hydra games appeared in an article published in 1982 by two mathematicians:
L. Kirby and J. Paris~\cite{KP82}: \emph{Accessible Independence Results for Peano Arithmetic}. 
Although the mathematical contents of this 
paper are quite advanced, hydra games (a.k.a. \emph{hydra battles}) are very easy to understand. There are now several sites on Internet where you can find tutorials on hydra games, together with simulators you can play with. See, for instance, the page written by Andrej Bauer~\cite{bauer2008}.

Hydra battles, as well as Goodstein Sequences~\cite{goodstein_1944, KP82}
are a nice way to present complex termination problems.
The article by Kirby and Paris presents a proof of termination
based on ordinal numbers, as well as a proof that this termination is not
provable in Peano arithmetic. In the book dedicated to 
J.P. ~Jouannaud \cite{HommageJPJ}, N.~Dershowitz and G.~Moser  give a thorough survey on this topic~\cite{Dershowitz2007}.

Here, we present a development  for the \coq{} proof assistant, after the work of Kirby and Paris. This formalization contains the following main parts:

\begin{itemize}
\item Representation in \coq{} of hydras and hydra battles
\item A proof that every battle is finite and won by Hercules. This proof is based on a \emph{variant} which maps any hydra to an ordinal strictly less than $\epsilon_0$ and is strictly decreasing along any battle.

\item Using a combinatorial toolkit designed by J.~Ketonen and R.~Solovay~\cite{KS81}, we prove that, for any ordinal $\mu<\epsilon_0$, there exists no such variant mapping any hydra to an ordinal stricly less than $\mu$. Thus, the complexity of $\epsilon_0$ is really needed in the previous proof.

\item We prove a relation between the length of a ``classic''  class of  battles
and the Wainer-Hardy hierarchy of ``rapidly growing functions'' $H_\alpha$~\cite{Wainer1970}. The considered class of battles, which we call \emph{standard}  is the most considered one in the scientific  litterature(including popularization).
\end{itemize}

Simply put, this document tries to combines the scientific interest of two articles~\cite{KP82, KS81} and a book~\cite{schutte} with the playful activity of truly proving theorems.
We hope  that such a work, besides exploring a nice piece of discrete maths, 
will show how \coq{} and its standard library are well fitted to help us to understand some non-trivial mathematical developments, and also to experiment the constructive parts of  the proof through functional programming.

\begin{figure}[h]
  \centering
  \includegraphics[width=11cm]{epsilon0.jpg}
  \caption{Mathematics and formal proofs}
  \label{fig:gamma0}
\end{figure}
 We also hope to provide highlights on infinity (both potential and actual) through the notions of function, computation, limit,
 types and proofs.

\subsection{Remarks}
In~\cite{KP82}, Kirby and Paris showed  that there is no proof of termination of all hydra battles in Peano Arithmetic (PA).
Since we are used to writing proofs in higher order logic, the restriction to PA was quite unnatural for us. So we chosed to prove another statement without any reference to PA, by considering a class of proofs indexed by ordinal numbers upto $\epsilon_0$.

Unlike mathematical literature, where definitions and proofs are spread over many articles and books,
the whole proof is now inside your computer. It is composed of the \texttt{.v} files you downloaded and 
parts of \coq's standard library. Thus, there is no ambiguity in our definitions and the premises of the theorems. Furthermore, you will be able to navigate through the development, using your favourite editor or IDE, and some commands like \texttt{Search}, \texttt{Locate}, \texttt{Print Assumptions}, etc.

Except in the \texttt{Schutte} library, dedicated to an axiomatic presentation of the set of countable ordinal numbers, all our development is axiom-free, and respects the rules of intuitionistic logic. Note that we also use the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777} for defining several rapidly growing hierarchy of functions, in Chap.~\ref{chap:alpha-large}. This plug-in imports the known-as-harmless following axiom:

\begin{Coqsrc}
FunctionalExtensionality.functional_extensionality_dep : 
forall (A : Type) (B : A -> Type) (f g : forall x : A, B x),
(forall x : A, f x = g x) -> f = g
\end{Coqsrc}

\paragraph{Main references}

In our development, we adapt the definitions and prove many theorems which
we found in the following articles. 
\begin{itemize}
\item ``Accessible independence results for Peano arithmetic''  by Laurie Kirby and Jeff Paris~\cite{KP82}
\item ''Rapidly growing Ramsey Functions'' by Jussi Ketonen and Robert Solovay~\cite{KS81}
\item ``The Termite and the Tower'', by Will Sladek~\cite{Sladek07thetermite}

\end{itemize}

\subsubsection*{Structure of this document}
\begin{itemize}
\item Whenever several implementations are possible, we will discuss the pros and cons of every possible choice
\item Most of the proofs we present are \emph{constructive}. Whenever possible, we provide the user with an associated function, which she or he can apply in \gallina{} or \ocaml{} in order to get a ``concrete'' feeling of the meaning of the considered theorem.
For instance, in Chapter~\vref{chap:ketonen}, the notion of \emph{limit ordinal} is
made more ``concrete'' thanks to a function \texttt{canonseq} that computes every item of a sequence which converges on a given limit ordinal $\alpha$. This simply typed function allows the user/reader to make her/his own experimentations.
For instance, one can very easily compute the $42$-nd item of a sequence which converges towards $\omega^{\omega^\omega}$.

\item We found it interesting to present several  implementions of a given concept. After some discussions of the pros and cons of each solution, we will choose to develop only one of them, leaving the others  as exercises or projects (i.e., big or difficult exercises).
In order to discuss which assumptions are really needed for proving a theorem, we will also present 
several aborted proofs.

 \end{itemize}

\paragraph*{Warning:}

This document is \emph{not} an introductory text for \coq{}, and there are many aspects of this proof assistant that are not covered. 
 The reader should already have some basic experience with the \coq{} system. The Reference Manual and several tutorials are available on \coq{} page~\cite{Coq}.  First chapters of textbooks like \emph{Interactive Theorem Proving and Program Development}~\cite{BC04}, \emph{Software Foundations}~\cite{SF} or  \emph{Certified Programming with Dependent Types} ~\cite{chlipalacpdt2011} will give you the right background.

\subsubsection*{Contributions are welcome}
\label{sec:orgheadline2}

Any form of contribution  is welcome: correction of errors, improvement of
Coq scripts, proposition of inclusion of new chapters, and generally any
comment or proposition that would help us. The text contains several \emph{projects} which, when completed, may improve the present work.
Please do not hesitate to bring your contribution!

\subsubsection{Acknowledgements}
\label{sec:orgheadline5}
    Many thanks to David Ilcinkas, Sylvain Salvati, Alan Schmitt and Théo Zimmerman for their help on the elaboration of this document, and to the
 members of the \emph{Formal Methods} team at laBRI for their helpful comments 
on an oral presentation of this work.

Many thanks also to the Coq development team, Yves Bertot, and members of the \emph{Coq Club} for interesting discussions about the \coq{} system and the Calculus of Inductive Constructions.

The author of the present document wishes to express his gratitude to the late Patrick Dehornoy, whose talk  was determinant for our desire to work on this topic.

I owe my interest in discrete mathematics and their relation to formal proofs and functional programming  to Srecko Brlek.  Equally, there is W. H. Burge's book ``\emph{Recursive Programming Techniques}'' ~\cite{burge} which was a great  source of inspiration.

\subsubsection{Typographical conventions}

Quotations from our  \coq{} source are displayed as follows:

  \begin{Coqsrc}
 Require Import Arith.

 Definition square (n:nat) := n * n.

 Lemma square_double : exists n:nat, n + n = square n.
 Proof.
    exists 2. 
  \end{Coqsrc}

Answers from \coq{} (including sub-goals, error messages, etc.) are displayed in slanted style
with a different background color.

 \begin{Coqanswer}
 1 subgoal, subgoal 1 (ID 5)

  ============================
   2 + 2 = square 2

 \end{Coqanswer}

 \begin{Coqsrc}
   reflexivity.
Qed.
 \end{Coqsrc}

\subsubsection{Alternative or bad definitions}

Finally, we decided to include definitions or lemma statements, as well as tactics,  that lead to
dead-ends or to too complex developments, with the following coloring.
Bad definitions 
 and encapsulation in modules called \texttt{Bad}, \texttt{Bad1}, etc.

 \begin{Coqbad}
Module Bad.

Definition double (n:nat)  := n + 2.

Lemma lt_double : forall n:nat, n < double  n.
Proof.
   unfold double; omega.
Qed.

End Bad.
  \end{Coqbad}

Likewise, alternative, but still unexplored definitions will be presented in modules
\texttt{Alt}, \texttt{Alt1}, etc. Using these definitions is left as an implicit exercise.

\begin{Coqalt}
Module Alt.

  Definition double (n : nat) := 2 * n.

End Alt.
\end{Coqalt}

\begin{Coqsrc}
Lemma alt_double_ok n : Nat.double n = Alt.double n.
Proof.
  unfold Alt.double, Nat.double.
  omega.
Qed.
\end{Coqsrc}

% \subsubsection{Links to the Coq source}

% Active links towards our \coq{} modules may be incorrect if you got this \texttt{pdf} document otherwise than by compiling the distribution available in
% \url{http://www.labri.fr/casteran/CoqArt/le_teaser/Teaser.tar.gz}.

### 2020-06-27 latex/chapter-skippable.tex
new chapter-skippable.tex

\chapter[Skippable chapter]{A skippable introduction to ordinal numbers and the Coq proof assistant}

blah.    blah.

\section{blah}

### 2020-06-30 latex/chapter-ks.tex
standradizing order of arguments of function canon and derivates

@@
 \coqsimple{cbn}, \coqsimple{simpl}, \coqsimple{compute}, etc., are applicable. For simplicity's sake, 
-we used an auxiliary function \texttt{canonS} of type \texttt{nat -> T1 -> T1} such that
-\texttt{canonS i $\alpha$} is equal to $\canonseq{\alpha}{i+1}$.
+we used an auxiliary function \texttt{canonS} of type \texttt{T1 -> nat  -> T1} such that
+\texttt{canonS  $\alpha$ $i$} is equal to $\canonseq{\alpha}{i+1}$.
 
@@
 \begin{Coqsrc}
-Fixpoint canonS (i:nat) alpha :=
+Fixpoint canonS alpha (i:nat) :=
   match alpha with
@@
           Some gamma' => ocons gamma' i zero
-        | None => ocons (canonS i gamma) 0 zero
+        | None => ocons (canonS gamma i) 0 zero
       end
@@
            Some gamma' => ocons gamma n (ocons gamma' i zero)
-         | None => ocons gamma n (ocons (canonS i gamma) 0 zero)
+         | None => ocons gamma n (ocons (canonS gamma i) 0 zero)
        end
-    | ocons alpha n beta => ocons alpha n (canonS i beta)
+    | ocons alpha n beta => ocons alpha n (canonS beta i)
   end.
@@
 \begin{Coqsrc}
- Definition canon i alpha := 
-   match i with 0 => zero | S j => canonS j alpha end. 
+ Definition canon alpha i := 
+   match i with 0 => zero | S j => canonS alpha j end. 
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Compute (canon 3 (omega ^ omega)).
+Compute (canon (omega ^ omega) 3).
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Example canon3 :  canon 3 (omega ^ omega) = omega ^ 3.
+Example canon3 :  canon (omega ^ omega) 3 = omega ^ 3.
 Proof. reflexivity. Qed.
@@
 \begin{Coqsrc}
-Compute pp (canon 42 (omega ^ omega * 3)).  
+Compute pp (canon (omega ^ omega * 3) 42).  
 \end{Coqsrc}
@@
 Lemma canonS_succ i alpha :
-  nf alpha ->  canonS i (succ alpha) = alpha.
+  nf alpha ->  canonS (succ alpha) i = alpha.
 Proof.
@@
 Lemma canonS_LT i alpha :
-  nf alpha -> alpha <> zero -> canonS i alpha <  alpha.
+  nf alpha -> alpha <> zero -> canonS alpha i <  alpha.
 \end{Coqsrc}
@@
      forall beta, beta < lambda ->
-                  {i:nat | beta < canonS i lambda}.
+                  {i:nat | beta < canonS lambda i}.
 
@@
   nf lambda -> is_limit lambda  ->
-  strict_lub (fun i => canonS i lambda) lambda.
+  strict_lub (fun i => canonS lambda i) lambda.
 \end{Coqsrc}
@@
 (Definition transition_S i : relation T1 :=
-  fun alpha beta =>  alpha <> zero /\ beta = canonS i alpha.
+  fun alpha beta =>  alpha <> zero /\ beta = canonS alpha i.
 
@@
     nf alpha -> alpha <> 0 ->
-    iota alpha -1-> iota (canonS i alpha).
+    iota alpha -1-> iota (canonS alpha i).
 \end{Coqsrc}
@@
   std_1 : forall i alpha, 
-       beta = canon i alpha -> j = S i ->
+       beta = canon alpha i -> j = S i ->
        standard_pathR j beta i  alpha
 | std_S : forall i alpha, 
-      standard_pathR j beta (S i) (canon i alpha)  ->
+      standard_pathR j beta (S i) (canon alpha i)  ->
       standard_pathR j beta i alpha.
@@
 Definition const_pathS i :=
-    clos_trans_1n T1 (fun alpha beta => beta = canonS i alpha).
+    clos_trans_1n T1 (fun alpha beta => beta = canonS alpha i).
 
@@
    forall i j, (i < j)%nat ->
-               const_pathS 0 (canonS j lambda)
-                             (canonS i lambda). 
+               const_pathS 0 (canonS lambda j)
+                             (canonS lambda i). 
   \end{Coqsrc}
@@
   | 0 => alpha
-  | S m => standard_gnaw (S i) (canon i alpha) m
+  | S m => standard_gnaw (S i) (canon alpha i) m
   end.

### 2020-06-30 latex/chapter-large-sets.tex
standradizing order of arguments of function canon and derivates

@@
   forall (i : nat) (alpha : T1),
-  nf alpha -> alpha <> zero -> canonS i alpha < alpha.
+  nf alpha -> alpha <> zero -> canonS alpha i < alpha.
 
@@
             (f : nat -> nat -> nat)
-            (H : forall k, L_spec (canonS k lambda) (f (S k))).
+            (H : forall k, L_spec (canonS lambda k) (f (S k))).
 
@@
           with Utils.dec (Is_Limit alpha) :=
-          { | left _ =>  L_ (Canon i alpha)  (S i) ;
+          { | left _ =>  L_ (Canon alpha i)  (S i) ;
             | right notlimit =>  L_ (Pred alpha) (S i)}}.
@@
 Lemma L_lim_eqn alpha i : Is_Limit alpha -> L_ alpha i =
-                                        L_ (Canon i alpha) (S i).
+                                        L_ (Canon alpha i) (S i).
 
@@
           with Utils.dec (Is_Limit alpha) :=
-          { | left _ =>  H_ (Canon (S i) alpha)  i ;
+          { | left _ =>  H_ (Canon alpha (S i))  i ;
             | right notlimit =>  H_ (Pred alpha) (S i)}}.
@@
 Lemma H_lim_eqn alpha i : Is_Limit alpha ->
-                      H_ alpha i =  H_ (Canon (S i) alpha) i.
+                      H_ alpha i =  H_ (Canon alpha (S i)) i.
 
@@
   intro k; rewrite H_lim_eqn ...
-  - replace (Canon (S k) omega) with (Fin (S k)).
+  - replace (Canon omega (S k)) with (Fin (S k)).
     + rewrite H_Fin; lia.
@@
  rewrite H_lim_eqn; simpl Canon.
- - ochange  (CanonS  k (omega * 2)%e0)  (omega + (S k))%e0.
+ - ochange  (CanonS  (omega * 2)%e0 k)  (omega + (S k))%e0.
   + rewrite H_Plus_Fin, H_omega;  lia.
@@
    rewrite H_lim_eqn; auto with E0.
-  - ochange (Canon (S k) (Phi0 2)) (omega * (S k))%e0.
+  - ochange (Canon (Phi0 2) (S k)) (omega * (S k))%e0.
     +  rewrite H_omega_i; simpl (exp2 (S k)).
@@
           with Utils.dec (Is_Limit alpha) :=
-          { | left _ =>  F_ (Canon i alpha)  i ;
+          { | left _ =>  F_ (Canon alpha i)  i ;
             | right notlimit =>  iterate (F_ (Pred alpha))  (S i) i}}.
@@
                  with Utils.dec (Is_Limit alpha) :=
-                 { | left _ => F_star (Canon i alpha,1) i ;
+                 { | left _ => F_star (Canon alpha i,1) i ;
                    | right notlimit =>
@@
 Lemma F_lim_eqn : forall alpha i,  Is_Limit alpha ->
-                               F_ alpha i = F_ (Canon i alpha) i.
+                               F_ alpha i = F_ (Canon alpha i) i.
 

### 2020-07-01 latex/chapter-skippable.tex
generic ipred and Omega_limit

@@
-\chapter[Skippable chapter]{A skippable introduction to ordinal numbers and the Coq proof assistant}
+\chapter{Introduction to ordinal numbers and the Coq proof assistant}
 
-blah.    blah.
+In this chapter, we introduce on a simple example the main concepts which are developped later.
+We assume the reader to have a very basic experience with the \coq{} proof assistant, for instance with one of the numerous tutorial one can consult on Internet
+(see for instance \url{https://coq.inria.fr/documentation}).
 
-\section{blah}
+The definitions  and theorems we are going to show are just simpler versions (\emph{i.e.} restrictions) of the ones contained in the forthcoming chapters.
+We hope that this simplicity will help to understand how we represent, compute and reason about ordinal notations.
+
+Beginners in \coq{} often play with natural numbers (the type \texttt{nat} of \coq's 
+standard library), so we considered the ordinal  $\omega$ as ``well-known'' and chosed 
+$\omega^2$ as a basis for our examples.
+
+\section{Ordinal numbers}
+\index{Maths!Ordinal numbers}
+
+The proof of termination of all hydra battles presented in~\cite{KP82} is based
+on \emph{ordinal numbers}.
+From a mathematical point of view, an ordinal is a representant of an equivalence class for isomorphims of strict, total and well-founded orders.
+
+We can also associate to every ordinal $\alpha$ a set whose elements are all ordinals strictly less than $\alpha$.
+Thus, it is meaningful  to consider \emph{finite}, \emph{infinite}, \emph{demunerable} and \emph{non-countable} ordinals.
+The relation $<$ on ordinals is well-founded, and the order $\leq$ associated with
+$<$ is total.
+
+We cannot cite all the litterature published on ordinals since Cantor's book 
+\cite{cantorbook}, and 
+leave it to the reader to explore the bibliography.
+Let us cite the book by Schütte~\cite{schutte} which contains an axiomatic definition of the set of countable ordinals we used as a mathematical specification of our implementaion in \coq{}~\cite{CantorContrib}.
+
+Out of respect of the tradition, the meta-variables for ordinals will be 
+ $\alpha$, $\beta$, $\gamma$, etc.
+
+\subsection{Definitions}
+
+\begin{itemize}
+\item Let  $\alpha$ be an ordinal; we say that  $\alpha$ is a \emph{successor} if there exists some ordinal  $\beta$ such that 
+$\alpha$ is is the least ordinal strictly greater than  $\beta$.
+
+\item We say that an ordinal $\lambda$ is a \emph{limit ordinal} is $\lambda$  is the least upper bound of a stricly increasing sequence of ordinals.
+The meta-variable $\lambda$ will be used for denoting  limit ordinals.
+
+\item \index{Maths!Transfinite induction}
+An ordinal is either $0$, a limit ordinal or a successor ordinal. This case analysis, as well as \emph{transfinite} ({i.e.}, well-founded) induction is used in many proofs about ordinal numbers.
+
+\item The segment of finite ordinals is isomorphic to the set of natural numbers.
+ The first infinite ordinal is the limit ordinal $\omega$.
+
+\item The operations $+$, $\times$ and exponentiation on $\mathbb{N}$ are extended on ordinals numbers.
+Note that these  extensions are not commutative any more.
+For instance $\omega = 1 + \omega \not= \omega + 1$
+and $\omega = 2 \times \omega \not= \omega \times 2$.
+
+\item The ordinal $\epsilon_0$ is the least solution of the equation
+ \(\alpha=\omega^{\alpha}\).
+\end{itemize}
+
+Please note that the set of countable ordinals is not countable.
+
+\subsection{Ordinal notations}
+In a proof assistant like \coq{},  it is useful to represent ordinals through some data-type, and make arithmetical operations and comparison effectively implemented  through certified functions.
+
+Of course, since the set of all countable ordinals is not countable, it cannot be represented by a language of finite terms.
+
+\subsubsection{Notations systems for small ordinals}
+
+The following list shows five popular notation systems for several ordinals (shown in increasing order).
+Please keep in mind that a notation system for a given ordinal $\alpha$ represents 
+ordinals \emph{strictly below} $\alpha$.
+
+\begin{itemize}
+\item $i\;(i\in\mathbb{N})$ \coq's type \verb@{j:nat | j < i}@ (see also
+\href{https://www.math.nagoya-u.ac.jp/~garrigue/lecture/2018_AW/mathcomp-1.7.0/htmldoc/mathcomp.ssreflect.fintype.html#Ordinal}{finite ordinals in Mathcomp}).
+\item $\omega$: \coq's type \texttt{nat}, with the order \texttt{Nat.lt}.
+\item $\omega^n$ (for some  integer  $n\geq 2$) : the type of $n$-uples of natural numbers, with the lexicographic product of $n$ copies of \texttt{Nat.lt}.
+\item  $\omega^\omega$: the set of nonincreasing sequences of natural numbers, lexicographically ordered (also the set of finite multisets of natural numbers).
+\item $\epsilon_0$: The set of terms in Cantor normal form (see Chap.~\ref{chap/T1}).
+\item $\Gamma_0$: The set of terms in Veblen normal form.
+\end{itemize}

### 2020-07-01 latex/chapter-termination.tex
generic ipred and Omega_limit

@@
 
-\section{Ordinal numbers}
-\index{Maths!Ordinal numbers}
-
-The proof of termination of all hydra battles presented in~\cite{KP82} is based
-on \emph{ordinal numbers}.
-From a mathematical point of view, an ordinal is a representant of an equivalence class for isomorphims of strict, total and well-founded orders.
-
-We can also associate to every ordinal $\alpha$ a set whose elements are all ordinals strictly less than $\alpha$.
-Thus, it is meaningful  to consider \emph{finite}, \emph{infinite}, \emph{demunerable} and \emph{non-countable} ordinals.
-The relation $<$ on ordinals is well-founded, and the order $\leq$ associated with
-$<$ is total.
-
-We cannot cite all the litterature published on ordinals since Cantor's book 
-\cite{cantorbook}, and 
-leave it to the reader to explore the bibliography.
-Let us cite the book by Schütte~\cite{schutte} which contains an axiomatic definition of the set of countable ordinals we used as a mathematical specification of our implementaion in \coq{}~\cite{CantorContrib}.
-
-Out of respect of the tradition, the meta-variables for ordinals will be 
- $\alpha$, $\beta$, $\gamma$, etc.
-
-\subsection{Definitions}
-
-\begin{itemize}
-\item Let  $\alpha$ be an ordinal; we say that  $\alpha$ is a \emph{successor} if there exists some ordinal  $\beta$ such that 
-$\alpha$ is is the least ordinal strictly greater than  $\beta$.
-
-\item We say that an ordinal $\lambda$ is a \emph{limit ordinal} is $\lambda$  is the least upper bound of a stricly increasing sequence of ordinals.
-The meta-variable $\lambda$ will be used for denoting  limit ordinals.
-
-\item \index{Maths!Transfinite induction}
-An ordinal is either $0$, a limit ordinal or a successor ordinal. This case analysis, as well as \emph{transfinite} ({i.e.}, well-founded) induction is used in many proofs about ordinal numbers.
-
-\item The segment of finite ordinals is isomorphic to the set of natural numbers.
- The first infinite ordinal is the limit ordinal $\omega$.
-
-\item The operations $+$, $\times$ and exponentiation on $\mathbb{N}$ are extended on ordinals numbers.
-Note that these  extensions are not commutative any more.
-For instance $\omega = 1 + \omega \not= \omega + 1$
-and $\omega = 2 \times \omega \not= \omega \times 2$.
-
-\item The ordinal $\epsilon_0$ is the least solution of the equation
- \(\alpha=\omega^{\alpha}\).
-\end{itemize}
-
-Please note that the set of countable ordinals is not countable.
-
- Unless otherwise specified, we will only consider ordinals strictly less than  $\epsilon_0$ in this chapter.
-
-% \subsection{Ordinal notations}
-% Whenever possible, define a representation of ordinals by finite expressions.
-% Of course, since the set of all countable ordinals is not countable, it cannot be represented by a language of finite terms.
-
-% \subsubsection{Notations systems for small ordinals}
-% \begin{itemize}
-% \item $\omega$
-% \item $\omega^2$
-% \item $\omega^n$
-% \item  $\omega^\omega$
-% \end{itemize}
-
 \section{The ordinal \texorpdfstring{\(\epsilon_0\)}{epsilon0}}
@@
 
-\subsection{Ordinal Notations}
-
-In a proof assistant like \coq{},  it is useful to represent ordinals through some data-type, and make arithmetical operations and comparison effectively implemented  through certified functions.
-Our user contribution~\cite{CantorContrib} represents 
-the set of ordinals strictly less than $\epsilon_0$ in Cantor normal form as in~\cite{Manolios2005}, and also the set
-of ordinals strictly  less than $\Gamma_0$ in Veblen normal form.
+\subsection{Cantor normal form}
 
@@
 
-\subsection{A data type for  Cantor normal forms}
+\subsection{A data type for  ordinals in Cantor normal form}
 \label{sec:orgheadline72}
 \label{sec:T1-inductive-def}
+
+% Our user contribution~\cite{CantorContrib} represents 
+% the set of ordinals strictly less than $\epsilon_0$ in Cantor normal form as in~\cite{Manolios2005}, and also the set
+% of ordinals strictly  less than $\Gamma_0$ in Veblen normal form.
 

### 2020-07-03 latex/chapter-hydras.tex
redaction Finite_ordinals

@@
 
-Let $B$ be some instance of class \texttt{Battle}. It is easy to define inductively the relation between the $i$-th and the $j$-th steps of a battle of type $B$.
+Let $B$ be any instance of class \texttt{Battle}. It is easy to define inductively the relation between the $i$-th and the $j$-th steps of a battle of type $B$.
 
@@
 % \end{remark}
+
+\section{Generic properties}
+
+\subsection{Liveliness}
+
+Let $B$ be an instance of \texttt{Battle}. We say that $B$ is \emph{alive} if
+for any configuration $(i,h)$, where $h$ is not a head, there exists a further step in class $B$.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
+
+\begin{Coqsrc}
+Definition Alive (B : Battle) :=
+  forall i h, 
+     h <> head -> {h' : Hydra |  B i h h'}.
+\end{Coqsrc}
+
+The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
+\url{../src/html/hydras.Hydra.Hydra_Theorems.html} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
+
+\begin{Coqsrc}
+Theorem Alive_free: Alive free.
+
+Theorem Alive_standard: Alive standard.  
+\end{Coqsrc}
+
+Both theorems are proved with the help of the  following strongly specified function:
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
+
+\begin{Coqsrc}
+Definition  next_round_dec n :
+ forall h , (h = head) + {h' : Hydra & {R1 h h'} + {R2  n h  h'}}.
+\end{Coqsrc}
+
+\subsection{Termination}
+
+The termination of all battles is naturally expressed by the predicate \texttt{well\_founded} defined in the module \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf} 
+ of the Standard Library.
+
+\index{Predicates!Termination}
+
+\begin{Coqsrc}
+Definition Termination :=  well_founded (transp _ round).
+\end{Coqsrc}
+
+Let $b$ be an instance of class \texttt{Battle}. A \emph{variant} for $b$ consists
+in a well-founded relation $<$  on some type \texttt{A}, and a function
+(also called a \emph{measure}) \texttt{m:Hydra->A} such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $b$, the inequality $m(h')<m(h)$ holds.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
+\index{Coq!Techniques!Type classes}
+
+\begin{Coqsrc}
+Class Hvariant {A:Type}{Lt:relation A}(Wf: well_founded Lt)(b : Battle)
+                          (m: Hydra -> A): Prop :=
+ {variant_decr :forall i h h', 
+     h <> head -> b i  h h' -> Lt (m h') (m h)}.
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+ Prove that, if there is an instance of \texttt{Hvariant Lt wf\_Lt b m}, then there exists no infinite battle in  \texttt{b}.
+\end{exercise}
+
+The example we just studied shows that the termination of any battle may take a very long time. If we want to study hydra battles in general, we have to consider 
+any hydra and any strategy, both for Hercules and the hydra itself. So, we first  give some definitions, generally borrowed from transition systems vocabulary (see~\cite{tel_2000} for instance).
+
+\subsection{Termination of all battles}
+
+% \subsection{Reachability}
+
+% Let $b$ be an instance of \texttt{battle}.  We say that a configuration $(i,h)$ is 
+% \emph{reachable} if there exists some initial hydra $h_0$ such that the transformation from $(0,h_0)$ into $(i,h)$ belongs to $b$.
+The following inductive definition considers also the case where $i=0$ and $h=h_0$.
+
+% \vspace{4pt}
+% \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
+
+% \begin{Coqsrc}
+% Inductive reachable (b : Battle)
+%   : nat -> Hydra ->  Prop  :=
+% | reachable_0 : forall h1, reachable b 0 h1
+% | reachable_S : forall i h h0,  fight b  0 h0 i h ->
+%                                 reachable b i h.
+% \end{Coqsrc}
+
+% A nice property of the classes \texttt{free} and \texttt{standard} is that every configuration $(i,h)$ is reachable. It is sufficient to add $i$ heads to $h$'s foot and start the battle at time $0$.
+
+% \vspace{4pt}
+% \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
+
+% \begin{Coqsrc}
+% Lemma reachable_free : forall i  h, reachable free i h.
+
+% Lemma reachable_standard  : forall h i, reachable standard i h.
+% \end{Coqsrc}
+
+\subsection{Liveliness}
+
+We say that a kind  $B$ of battles is \emph{alive} if for any configuration $(i,h)$, where $h$ is not a head, there exists a further step in class $B$.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
+
+\begin{Coqsrc}
+Definition Alive (B : Battle) :=
+  forall i h, 
+     h <> head -> {h' : Hydra |  B i h h'}.
+\end{Coqsrc}
 
@@
 
-\section{Reasoning about any battle}
-
-The example we just studied shows that the termination of any battle may take a very long time. If we want to study hydra battles in general, we have to consider 
-any hydra and any strategy, both for Hercules and the hydra itself. So, we first  give some definitions, generally borrowed from transition systems vocabulary (see~\cite{tel_2000} for instance).
-
-\subsection{Termination of all battles}
-
-The termination of all battles is naturally expressed by the predicate \texttt{well\_founded} defined in the module \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf} 
- of the Standard Library.
-
-\index{Predicates!Termination}
-
-\begin{Coqsrc}
-Definition Termination :=  well_founded (transp _ round).
-\end{Coqsrc}
-
-% \subsection{Reachability}
-
-% Let $b$ be an instance of \texttt{battle}.  We say that a configuration $(i,h)$ is 
-% \emph{reachable} if there exists some initial hydra $h_0$ such that the transformation from $(0,h_0)$ into $(i,h)$ belongs to $b$.
-The following inductive definition considers also the case where $i=0$ and $h=h_0$.
-
-% \vspace{4pt}
-% \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-% \begin{Coqsrc}
-% Inductive reachable (b : Battle)
-%   : nat -> Hydra ->  Prop  :=
-% | reachable_0 : forall h1, reachable b 0 h1
-% | reachable_S : forall i h h0,  fight b  0 h0 i h ->
-%                                 reachable b i h.
-% \end{Coqsrc}
-
-% A nice property of the classes \texttt{free} and \texttt{standard} is that every configuration $(i,h)$ is reachable. It is sufficient to add $i$ heads to $h$'s foot and start the battle at time $0$.
-
-% \vspace{4pt}
-% \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-
-% \begin{Coqsrc}
-% Lemma reachable_free : forall i  h, reachable free i h.
-
-% Lemma reachable_standard  : forall h i, reachable standard i h.
-% \end{Coqsrc}
-
-\subsection{Liveliness}
-
-We say that a kind  $B$ of battles is \emph{alive} if for any configuration $(i,h)$, where $h$ is not a head, there exists a further step in class $B$.
-
-\vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
-
-\begin{Coqsrc}
-Definition Alive (B : Battle) :=
-  forall i h, 
-     h <> head -> {h' : Hydra |  B i h h'}.
-\end{Coqsrc}
-
-The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
-\url{../src/html/hydras.Hydra.Hydra_Theorems.html} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
-
-\begin{Coqsrc}
-Theorem Alive_free: Alive free.
-
-Theorem Alive_standard: Alive standard.  
-\end{Coqsrc}
-
-Both theorems are proved with the help of the  following strongly specified function:
-
-\vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
-
-\begin{Coqsrc}
-Definition  next_round_dec n :
- forall h , (h = head) + {h' : Hydra & {R1 h h'} + {R2  n h  h'}}.
-\end{Coqsrc}
-
-\section{Termination}
-
-In this section, we are interested in proofs of termination of all battles of a  given kind, and specifically proofs by \emph{variants}.
-
-Let $b$ be an instance of class \texttt{Battle}. A \emph{variant} for $b$ consists
-in a well-founded relation $<$  on some type \texttt{A}, and a function
-(also called a \emph{measure}) \texttt{m:Hydra->A} such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $b$, the inequality $m(h')<m(h)$ holds.
-
-\vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
-\index{Coq!Techniques!Type classes}
-
-\begin{Coqsrc}
-Class Hvariant {A:Type}{Lt:relation A}(Wf: well_founded Lt)(b : Battle)
-                          (m: Hydra -> A): Prop :=
- {variant_decr :forall i h h', 
-     h <> head -> b i  h h' -> Lt (m h') (m h)}.
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
- Prove that, if there is an instance of \texttt{Hvariant Lt wf\_Lt b m}, then there exists no infinite battle in  \texttt{b}.
-\end{exercise}
-
 \subsection{Some failed attempts}

### 2020-07-03 latex/chapter-skippable.tex
Finite_ordinals added
redaction Finite_ordinals

@@
-\chapter{Introduction to ordinal numbers and the Coq proof assistant}
+\chapter{Introduction to ordinal numbers with the Coq proof assistant}
 
-In this chapter, we introduce on a simple example the main concepts which are developped later.
-We assume the reader to have a very basic experience with the \coq{} proof assistant, for instance with one of the numerous tutorial one can consult on Internet
-(see for instance \url{https://coq.inria.fr/documentation}).
+In this chapter, we introduce on  simple examples the main concepts which are developped later.
+We assume the reader to have a very basic experience with the \coq{} proof assistant, for instance with one of the numerous tutorials one can consult on Internet
+(see for instance \url{https://coq.inria.fr/documentation}). A basic experience  with Peano numbers (library \texttt{Arith}) would be useful.
 
@@
 We hope that this simplicity will help to understand how we represent, compute and reason about ordinal notations.
-
-Beginners in \coq{} often play with natural numbers (the type \texttt{nat} of \coq's 
-standard library), so we considered the ordinal  $\omega$ as ``well-known'' and chosed 
-$\omega^2$ as a basis for our examples.
 
@@
 
-\item The segment of finite ordinals is isomorphic to the set of natural numbers.
+\item Let $\alpha$ be any ordinal, the \emph{segment} $\mathbb{O}_\alpha$ is the
+interval $[0,\alpha[\,=\,\{\beta|0\leq\beta<\alpha\}$.
+
+\item The set  of finite ordinals is a segment isomorphic to the set of natural numbers.
  The first infinite ordinal is the limit ordinal $\omega$.
@@
 
-\subsection{Ordinal notations}
+Without ambiguity we will denote the segment $\mathbb{O}_\alpha$ by just $\alpha$.
+
+\section{Ordinal notations}
 In a proof assistant like \coq{},  it is useful to represent ordinals through some data-type, and make arithmetical operations and comparison effectively implemented  through certified functions.
 
-Of course, since the set of all countable ordinals is not countable, it cannot be represented by a language of finite terms.
+Of course, since the set of all countable ordinals is not countable, it cannot be represented by a language of finite terms. So, we will study ordinal notations just  for countable segments.
 
-\subsubsection{Notations systems for small ordinals}
+\subsection{Notation systems for some small ordinals}
 
-The following list shows five popular notation systems for several ordinals (shown in increasing order).
+The following enumeration  shows  popular notation systems for several ordinals (shown in increasing order with respect to inclusion), and their possible representation in \coq{}.
 Please keep in mind that a notation system for a given ordinal $\alpha$ represents 
 ordinals \emph{strictly below} $\alpha$.
+
+\subsubsection{Finite ordinals}
+Let $n$ be some natural number. The segment associated with $n$ is the interval 
+$[0,n[\,=\,\{0,1,\dots,n-1\}$.
+
+One may represent the ordinal $n$ by a sigma type.
+
+\vspace{4pt}
+\noindent\emph{From Module~\href{../src/html/hydras.Prelude.Finite_ordinals.html}{Prelude.Finite\_ordinals}}
+
+\begin{Coqsrc}
+Coercion is_true: bool >-> Sortclass.
+
+Definition t (n:nat) := {i:nat | Nat.ltb i  n}.
+\end{Coqsrc}
+
+The order on type \texttt{t $n$} is defined through the projection on \texttt{nat}.
+
+\begin{Coqsrc}
+Definition lt {n:nat} : relation (t n) :=
+  fun alpha beta => Nat.ltb ( proj1_sig alpha) (proj1_sig beta).
+\end{Coqsrc}
+
+For instance, let us build two elements of the segment $[0, 7[$, \emph{i.e.} two
+inhabitants of   type (\texttt{t 7}), and prove a simple  inequality (see Fig.~\ref{fig:O7}).
+
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}[very thick, scale=0.6]
+
+\node (N0) at (0,0) {$\bullet$};
+\node (i0) at (0,1) {$0$};
+\node (N1) at (2,0) {$\bullet$};
+\node (i1) at (2,1) {$1$};
+\node (N2) at (4,0) {$\bullet$};
+\node (i2) at (4,1) {$2$};
+\node (N3) at (6,0) {$\bullet$};
+\node (i3) at (6,1) {$3$};
+\node (N4) at (8,0) {$\bullet$};
+\node (i4) at (8,1) {$4$};
+\node (N5) at (10,0) {$\bullet$};
+\node (i5) at (10,1) {$5$};
+\node (N6) at (12,0) {$\bullet$};
+\node (i6) at (12,1) {$6$};
+\node(alpha1) at (4,-1) {$\alpha_1$};
+\node(alpha2) at (10,-1) {$\beta_1$};
+\end{tikzpicture}
+
+\caption{The segment $\mathbb{O}_7$\label{fig:O7}}
+\end{figure}
+
+\index{Coq!Commands!Program Definition}
+
+\begin{Coqsrc}
+Program Example alpha1 : t 7 := 2.
+
+Program Example beta1 : t 7 := 5.
+
+Example i1 : lt  alpha1 beta1.
+Proof.   now compute. Qed.
+\end{Coqsrc}
+
+Note that the type \texttt{t 0} is empty, and that \texttt{Program} generates an obligation
+for verifying that the constraint $i<n$ is fullfilled when one tries to build the $i$-th ordinal in type \texttt{t $n$}.
+
+\begin{Coqsrc}
+Lemma t0_empty (alpha: t 0): False.
+Proof.
+  destruct alpha.
+  destruct x; cbn in i; discriminate.
+Qed.
+
+Program Definition bad : t 10 := 10.
+Next Obligation.
+  compute.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+1 subgoal (ID 162)
+
+  ============================
+  false = true
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Abort.
+\end{Coqsrc}
+
+Note also that attempting to compare an ordinal of type \texttt{t $n$}  with an ordinal of
+type \texttt{t $p$}  leads to an error if $n$ and $p$ are not convertible.
+
+\begin{Coqsrc}
+
+Program Example gamma1 : t 8 := 7.
+
+Fail Goal lt alpha1 gamma1.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+ The command has indeed failed with message:
+The term "gamma1" has type "t 8" while it is expected to have type "t 7".
+\end{Coqanswer}
 

### 2020-07-04 latex/chapter-skippable.tex
Compatibility with Datatypes.CompareSpec

@@
 
-Without ambiguity we will denote the segment $\mathbb{O}_\alpha$ by just $\alpha$.
+Without ambiguity we will often denote the segment $\mathbb{O}_\alpha$ by just $\alpha$.
 
@@
 
+\begin{todo}
+Comments on \texttt{cast, compare, bigO}
+
+Show that there is no interesting arihmetic, due to the lack of total succesor, addition, etc.
+\end{todo}
 \begin{itemize}

### 2020-07-06 latex/chapter-skippable.tex
work on type classes for ordinal notations

@@
 
+\subsection{A type class for ordinal notations}
+
+Let $<$ be some relation on a given type $A$. In order to build an ordinal  notation 
+system out of $(A,<)$, we require $<$ to be a strict, well-founded order, and that 
+the reflexive closure of $<$ is a total order on $A$.
+
+\index{Coq!Techniques!Type classes}
+
+Using \coq's standard library and type classes, we propose the following definition.
+
+\vspace{4pt}
+\noindent\emph{From Module~\href{../src/html/hydras.Prelude.Ordinal_generic.html}{Prelude.Ordinal\_generic}}
+
+\begin{Coqsrc}
+Class OrdinalNotation {A:Type}{lt: relation A}(sto:StrictOrder lt)
+      (compare : A -> A -> comparison) :=
+  { compare_correct :
+      forall alpha beta:A,
+        CompareSpec (alpha=beta) (lt alpha beta) (lt beta alpha)
+                                 (compare alpha beta);
+    wf : well_founded lt}.  
+\end{Coqsrc}
+
 \subsection{Notation systems for some small ordinals}
@@
 
+\label{def: Finite-ord-type}
 \begin{Coqsrc}
@@
 
+In order to build an instance of \texttt{OrdinalNotation}, we define a comparison function, by delegation to standard library's  \texttt{Nat.compare}, and prove its correction.
+
+\begin{Coqsrc}
+Definition compare {n:nat} (alpha beta : t n) :=
+  Nat.compare (proj1_sig alpha) (proj1_sig beta).
+
+Lemma compare_correct {n} (alpha beta : t n) :
+  CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
+              (compare alpha beta).
+\end{Coqsrc}
+
+\begin{remark}
+ The proof of \texttt{compare\_correct} uses a well-know pattern of \coq{}.
+Let us consider  the following subgoal.
+
+\begin{Coqanswer}
+ 1 subgoal (ID 110)
+
+  n, x0 : nat
+  i, i0 : x0 <? S n
+  ============================
+  exist (fun i1 : nat => i1 <=? n) x0 i =
+  exist (fun i1 : nat => i1 <=? n) x0 i0
+\end{Coqanswer}
+
+Applying the tactic \texttt{f\_equal} generates a simpler subgoal.
+
+\begin{Coqanswer}
+1 subgoal (ID 112)
+
+  n, x0 : nat
+  i, i0 : x0 <? S n
+  ============================
+  i = i0
+\end{Coqanswer}
+
+We have now to prove that two proofs of \texttt{Nat.ltb x0 (S n)} are equal.
+
+This is not obvious, but  a consequence of the following lemma of library 
+\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Eqdep_dec.html}{Coq.Logic.Eqdep\_dec}.
+
+\index{Coq!Techniques!Unicity of equality proofs}
+
+\begin{Coqanswer}
+eq_proofs_unicity_on :
+forall (A : Type) (x : A),
+(forall y : A, x = y \/ x <> y) -> forall (y : A) (p1 p2 : x = y), p1 = p2
+\end{Coqanswer}
+
+Thus unicity of proofs of \texttt{Nat.ltb x0 (S n)}  comes from the decidability of
+equality on type \texttt{bool}.
+
+This is why we used the boolean function \texttt{Nat.ltb} instead of the inductive predicate \texttt{Nat.lt} in the definition of type \texttt{t $n$} (see page~\pageref{def: Finite-ord-type}).
+
+For more information about this pattern, please lokk at the numerous mailing lists and 
+FAQs on \coq{}).
+
+\end{remark}
+
+Applying lemmas of the libraries \texttt{Coq.Wellfounded.Inverse\_Image},
+ \texttt{Coq.Wellfounded.Inclusion}, and \texttt{Coq.Arith.Wf\_nat}, we prove that our
+relation \texttt{wf} is well founded.
+
+\begin{Coqsrc}
+Lemma lt_wf (n:nat) : well_founded (@lt n).
+\end{Coqsrc}
+
+Now we can build our instance of \texttt{OrdinalNotation}.
+
+\begin{Coqsrc}
+Global Instance sto n : StrictOrder (@lt n).
+
+Global Instance FinOrd (n:nat) : OrdinalNotation (sto n) compare.
+Proof.
+  split.
+  - apply compare_correct.
+  - apply lt_wf.
+Qed.
+\end{Coqsrc}
+
+Finally, it is interesting to show that the segment $[0,n[$ is a ``sub-segment'' of
+$[0,n+1[$. Since types $t\;n$ are $t\;(n+1)$ are not convertible, we consider a ``cast'' 
+function from $t\;n$ to $t\;(n+1)$, and prove this function is compatible with
+comparison function of both types, and that it is a bijection from $t\;n$ to
+the segment $[0,n-1[$ of $t\;(n+1)$.
+
+In other words, \texttt{FinOrd $n$}  is a subnotation of \texttt{FinOrd $(n+1)$}.
+
+\begin{Coqsrc}
+Class  SubNotation {A:Type}
+       {ltA: relation A}{stoA:StrictOrder ltA}
+       {compareA : A -> A -> comparison}
+       (notA : OrdinalNotation stoA compareA)
+       {B:Type}{ltB: relation B}{stoB:StrictOrder ltB}
+       {compareB : B -> B -> comparison}
+       (notB : OrdinalNotation stoB compareB)
+       (bound :  B)
+       (cast : A -> B):=
+  {
+  commutation :forall a a' : A,  compareB (cast a) (cast a') =
+                                 compareA a a';
+  incl : forall a, ltB (cast a) bound;
+  surj : forall b, ltB b bound -> exists a:A, cast a = b}.
+\end{Coqsrc}
+
+Coming back to finite ordinals, the conversion function from $t\;i$ to $t\;j$ is
+defined with \texttt{Program}.
+
+\begin{Coqsrc}
+Program Definition cast {i j : nat}  (H: i < j) (alpha: t i) : t j :=
+  alpha.
+
+Lemma cast_compare_commute (i j  :nat)(H : i < j) :
+  forall alpha beta, compare alpha beta = compare (cast  H alpha)
+                                                  (cast  H beta).
+
+\end{Coqsrc}
+
+We are now able to build an instance of \texttt{SubNotation}
+
+\begin{Coqsrc}
+(** biggest element of t (S n) *)
+
+Program Definition biggest_S (n:nat) : t (S n) :=n.
+Next Obligation. 
+  red;rewrite  Nat.ltb_lt; auto with arith.
+Defined.
+
+Instance F_incl n : SubNotation  (FinOrd n) (FinOrd (S n)) (biggest_S n)
+                                 (cast (Nat.lt_succ_diag_r n)).
+\end{Coqsrc}
+
 \begin{todo}

### 2020-07-07 latex/chapter-skippable.tex
work on type classes for ordinal notations

@@
 $[0,n+1[$. Since types $t\;n$ are $t\;(n+1)$ are not convertible, we consider a ``cast'' 
-function from $t\;n$ to $t\;(n+1)$, and prove this function is compatible with
-comparison function of both types, and that it is a bijection from $t\;n$ to
+function $\iota$ from $t\;n$ to $t\;(n+1)$, and prove this function is  a bijection  from $t\;n$ to
 the segment $[0,n-1[$ of $t\;(n+1)$.
 
-In other words, \texttt{FinOrd $n$}  is a subnotation of \texttt{FinOrd $(n+1)$}.
+This notion is defined generically as a class in
+ Module~\href{../src/html/hydras.Prelude.Ordinal-generic.html}{Prelude.Ordinal\_generic} (see Fig~\vref{figure:subsegment}).
 
 \begin{Coqsrc}
-Class  SubNotation {A:Type}
-       {ltA: relation A}{stoA:StrictOrder ltA}
-       {compareA : A -> A -> comparison}
-       (notA : OrdinalNotation stoA compareA)
-       {B:Type}{ltB: relation B}{stoB:StrictOrder ltB}
-       {compareB : B -> B -> comparison}
-       (notB : OrdinalNotation stoB compareB)
-       (bound :  B)
-       (cast : A -> B):=
+(** The segment associated with nA is isomorphic to
+    the interval [0,b[ *)
+
+Class  Subsegment 
+       `(nA : @OrdinalNotation A ltA stoA compareA)
+       `(nB : @OrdinalNotation B ltB stoB compareB)
+       (b :  B)
+       (iota : A -> B):=
   {
-  commutation :forall a a' : A,  compareB (cast a) (cast a') =
-                                 compareA a a';
-  incl : forall a, ltB (cast a) bound;
-  surj : forall b, ltB b bound -> exists a:A, cast a = b}.
+  subseg_compare :forall x y : A,  compareB (iota x) (iota y) =
+                                 compareA x y;
+  subseg_incl : forall x, ltB (iota x) b;
+  subseg_onto : forall y, ltB y b  -> exists x:A, iota x = y}.
 \end{Coqsrc}
+
+ \begin{figure}[h]
+   \centering
+   \begin{tikzpicture}[very thick, scale=0.6]
+\begin{scope}[color=blue]
+\node (A) at (0,0) {$A$};
+\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
+\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
+\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
+\node (Adots) at (6,0) {$\ldots$};
+\end{scope}
+\begin{scope}[color=red]
+\node (B) at (0,2) {$B$};
+\node(B0) at (2,2)[label=above:$0$]{$\bullet$};
+\node(B1) at (3,2)[label=above:$1$]{$\bullet$};
+\node(B2) at (4,2)[label=above:$2$]{$\bullet$};
+\node (Bdots) at (6,2) {$\ldots$};
+\node (b) at (8,2) [label=above:$b$]{$\bullet$};
+\node (bsucc) at (9,2) [label=above:$b+1$]{$\bullet$};
+\node (Bdots2) at (10,2) {$\ldots$};
+\end{scope}
+\begin{scope}[color=red!50!blue]
+\draw [->,thin] (A0) -- node [auto] {$\iota$} (B0);
+\draw [->,thin] (A1) -- node [auto] {$\iota$} (B1);
+\draw [->,thin] (A2) -- node [auto] {$\iota$} (B2);
+\draw [->,thin] (Adots) -- node [auto] {$\iota$} (Bdots);
+\end{scope}
+\end{tikzpicture}
+   \caption{\textcolor{blue}{$A$} is a sub-segment  of \textcolor{red}{$B$}}
+   \label{fig:subsegment}
+ \end{figure}
 
@@
 \begin{Coqsrc}
-Program Definition cast {i j : nat}  (H: i < j) (alpha: t i) : t j :=
+Program Definition iota {i j : nat}  (H: i < j) (alpha: t i) : t j :=
   alpha.
 
-Lemma cast_compare_commute (i j  :nat)(H : i < j) :
-  forall alpha beta, compare alpha beta = compare (cast  H alpha)
-                                                  (cast  H beta).
+Lemma iota_compare_commute (i j  :nat)(H : i < j) :
+  forall alpha beta, compare alpha beta = compare (iota  H alpha)
+                                                  (iota  H beta).
+\end{Coqsrc}
+
+We are now able to build an instance of \texttt{Subsegment}
+
+\begin{Coqsrc}
+Section Inclusion_ij.
+
+  Variables i j : nat.
+  Hypothesis Hij : (i < j)%nat.
+
+   Program Definition iota  (alpha: t i) : t j :=  alpha.
+
+   Let b : t j := exist _ i Ltb_ij.
+
+  Instance F_incl_ij  : SubSegment  (FinOrd i) (FinOrd j) b iota.
+
+  End Inclusion_ij.
 
@@
 
-We are now able to build an instance of \texttt{SubNotation}
+\begin{remark}
+ There is no interesting arihmetic on finite ordinals, since functions like successor, addition;, etc.,  cannot represented in \coq{} as \emph{total} functions.
+\end{remark}
 
-\begin{Coqsrc}
-(** biggest element of t (S n) *)
-
-Program Definition biggest_S (n:nat) : t (S n) :=n.
-Next Obligation. 
-  red;rewrite  Nat.ltb_lt; auto with arith.
-Defined.
-
-Instance F_incl n : SubNotation  (FinOrd n) (FinOrd (S n)) (biggest_S n)
-                                 (cast (Nat.lt_succ_diag_r n)).
-\end{Coqsrc}
-
-\begin{todo}
-Comments on \texttt{cast, compare, bigO}
-
-Show that there is no interesting arihmetic, due to the lack of total succesor, addition, etc.
-\end{todo}
 \begin{itemize}

### 2020-07-07 latex/chapter-skippable.tex
restructuration of Gamma0

@@
 
-Finally, it is interesting to show that the segment $[0,n[$ is a ``sub-segment'' of
-$[0,n+1[$. Since types $t\;n$ are $t\;(n+1)$ are not convertible, we consider a ``cast'' 
-function $\iota$ from $t\;n$ to $t\;(n+1)$, and prove this function is  a bijection  from $t\;n$ to
-the segment $[0,n-1[$ of $t\;(n+1)$.
+Finally, it is interesting to show that, if $i<j$, then the segment $[0,i[$ is a ``sub-segment'' of
+$[0,j[$. Since the terms  $t\;i$ and $t\;j$ are not convertible, we consider a ``cast'' 
+function $\iota$ from $t\;i$ to $t\;j$, and prove this function is  a bijection  from $t\;i$ to
+the segment $[0,i[$ of $t\;j$.
 
@@
 
-Coming back to finite ordinals, the conversion function from $t\;i$ to $t\;j$ is
-defined with \texttt{Program}.
+\index{Coq!Commands!Program}
 
-\begin{Coqsrc}
-Program Definition iota {i j : nat}  (H: i < j) (alpha: t i) : t j :=
-  alpha.
-
-Lemma iota_compare_commute (i j  :nat)(H : i < j) :
-  forall alpha beta, compare alpha beta = compare (iota  H alpha)
-                                                  (iota  H beta).
-\end{Coqsrc}
-
-We are now able to build an instance of \texttt{Subsegment}
+We are now able to build an instance of \texttt{Subsegment}.
 
@@
 
-   Program Definition iota  (alpha: t i) : t j :=  alpha.
+   Remark Ltb_ij : Nat.ltb i j.
+   Program Definition iota_ij  (alpha: t i) : t j :=  alpha.
 
@@
 
-  Instance F_incl_ij  : SubSegment  (FinOrd i) (FinOrd j) b iota.
+   Global Instance F_incl_ij  : SubSegment  (FinOrd i) (FinOrd j) b iota_ij.
+  (* ... *)
 

### 2020-07-08 latex/chapter-skippable.tex
add Omega_ordinal

@@
 
-\subsection{Notation systems for some small ordinals}
+\section{Notation systems for some small ordinals}
 
@@
 
-\subsubsection{Finite ordinals}
+\subsection{Finite ordinals}
 Let $n$ be some natural number. The segment associated with $n$ is the interval 
@@
 \begin{remark}
- There is no interesting arihmetic on finite ordinals, since functions like successor, addition;, etc.,  cannot represented in \coq{} as \emph{total} functions.
+ There is no interesting arihmetic on finite ordinals, since functions like successor, addition, etc.,  cannot be represented in \coq{} as \emph{total} functions.
 \end{remark}
 
+\paragraph{Related work}
+Finite ordinals are also formailzed in MathComp~\cite{SSR}.  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.
+
+\subsection{The first infinite ordinal : \texorpdfstring{$\omega$}{omega}}
+
+Mathematically speaking, the ordinal $\omega$ is (set-theoretically) considered as the well-ordered set of
+natural numbers. 
+Every beginner in \coq{} must have seen examples using the type \texttt{nat} of natural numbers. \coq's standard 
+library already contains a lot of functions and lemmas on Peano arithmetic.
+
+In order to build a notation system for $\omega$, we have just to look for lemmas about strict orders and well-foundedness in type \texttt{nat}.
+
+\begin{Coqsrc}
+Require Import Arith Compare_dec Lia Simple_LexProd Ordinal_generic.
+Import Relations.
+
+Search (@StrictOrder nat).
+Search @Nat.compare.
+Search (@well_founded nat).
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Instance Omega : OrdinalNotation Nat.lt_strorder Nat.compare.
+Proof.
+ split.
+ - apply Nat.compare_spec.
+ - apply lt_wf.
+Qed.
+\end{Coqsrc}
+
 \begin{itemize}
-\item $i\;(i\in\mathbb{N})$ \coq's type \verb@{j:nat | j < i}@ (see also
-\href{https://www.math.nagoya-u.ac.jp/~garrigue/lecture/2018_AW/mathcomp-1.7.0/htmldoc/mathcomp.ssreflect.fintype.html#Ordinal}{finite ordinals in Mathcomp}).
 \item $\omega$: \coq's type \texttt{nat}, with the order \texttt{Nat.lt}.

### 2020-07-08 latex/chapter-skippable.tex
redaction chapter on ordinal notations

@@
 \paragraph{Related work}
-Finite ordinals are also formailzed in MathComp~\cite{SSR}.  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.
+Finite ordinals are also formalized in MathComp~\cite{SSR}.  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.
 
@@
 
+We can also prove that any finite ordinal $i$ can be considered as a sub-segment of $\omega$.
+
+\begin{Coqsrc}
+Global Instance FinOrd_Omega (i:nat) :
+  SubSegment (FinOrd i) Omega i 
+             (fun alpha =>  proj1_sig alpha).
+\end{Coqsrc}
+
+\subsection{The ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
+
+The ordinal $\omega+\omega$ (also known as $\omega\times 2$) may be represented as the concatenation 
+of two copies of $\omega$ (Figure~\ref{fig:omega-plus-omega}).
+
+\begin{figure}[h]
+   \centering
+   \begin{tikzpicture}[very thick, scale=0.5]
+\begin{scope}[color=blue]
+\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
+\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
+\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
+\node (Adots) at (6,0) {$\ldots$};
+\node(An) at (8,0)[label=below:$n$]{$\bullet$};
+\node(A2) at (10,0)[label=below:$n+1$]{$\bullet$};
+\node (Adots1) at (12,0) {$\ldots$};
+\end{scope}
+\begin{scope}[color=red]
+\node(B0) at (14,0)[label=below:$0$,label=above:$\omega$]{$\bullet$};
+\node(B1) at (16,0)[label=below:$1$, label=above:$\omega+1$]{$\bullet$};
+\node(B2) at (18,0)[label=below:$2$,label=above:$\omega+2$]{$\bullet$};
+\node (Bdots) at (20,0) {$\ldots$};
+\node (Bn) at (22,0) [label=below:$p$, label=above:$\omega+p$]{$\bullet$};
+\node (Bdots2) at (24,0) {$\ldots$};
+\end{scope}
+\end{tikzpicture}
+   \caption{\textcolor{blue}{$\omega+{\color{red}\omega}$}}
+   \label{fig:omega-plus-omega}
+ \end{figure}
+
+In \coq{}, we use the disjoint union \texttt{(nat+nat)\%type} for representing the elements of $\omega+\omega$.
+The blue ordinals of Figure~\ref{fig:omega-plus-omega} will be represented by terms of the form \texttt{(inl $i$)},
+the red ones by \texttt{(inr $j$)}.
+
+\vspace{4pt}
+\noindent\emph{From Module~\href{../src/html/hydras.Prelude.Omega_plus_omega.html}{Prelude.Omega\_plus\_omega}}
+
+\begin{Coqsrc}
+Declare Scope oo_scope.
+Delimit Scope oo_scope with oo.
+Open Scope oo_scope.
+
+Definition t := (nat + nat)%type.
+Arguments inl  {A B} _.
+Arguments inr  {A B} _.
+\end{Coqsrc}
+
 \begin{itemize}
-\item $\omega$: \coq's type \texttt{nat}, with the order \texttt{Nat.lt}.
 \item $\omega^n$ (for some  integer  $n\geq 2$) : the type of $n$-uples of natural numbers, with the lexicographic product of $n$ copies of \texttt{Nat.lt}.

### 2020-07-09 latex/chapter-ks.tex
renaming is_limit and is_succ

@@
      nf lambda ->
-     is_limit lambda  ->
+     limitb lambda  ->
      forall beta, beta < lambda ->
@@
 Lemma canonS_limit_lub (lambda : T1) :
-  nf lambda -> is_limit lambda  ->
+  nf lambda -> limitb lambda  ->
   strict_lub (fun i => canonS lambda i) lambda.
@@
    nf lambda ->
-   is_limit lambda  ->
+   limitb lambda  ->
    forall i j, (i < j)%nat ->

### 2020-07-09 latex/chapter-large-sets.tex
renaming is_limit and is_succ

@@
             (Hnf : nf lambda)
-            (Hlim : is_limit lambda)
+            (Hlim : limitb lambda)
             (f : nat -> nat -> nat)
@@
       | right nonzero
-          with Utils.dec (Is_Limit alpha) :=
+          with Utils.dec (Limitb alpha) :=
           { | left _ =>  L_ (Canon alpha i)  (S i) ;
@@
 
-Lemma L_lim_eqn alpha i : Is_Limit alpha -> L_ alpha i =
+Lemma L_lim_eqn alpha i : Limitb alpha -> L_ alpha i =
                                         L_ (Canon alpha i) (S i).
@@
       | right nonzero
-          with Utils.dec (Is_Limit alpha) :=
+          with Utils.dec (Limitb alpha) :=
           { | left _ =>  H_ (Canon alpha (S i))  i ;
@@
 
-Lemma H_lim_eqn alpha i : Is_Limit alpha ->
+Lemma H_lim_eqn alpha i : Limitb alpha ->
                       H_ alpha i =  H_ (Canon alpha (S i)) i.
@@
       | right nonzero
-          with Utils.dec (Is_Limit alpha) :=
+          with Utils.dec (Limitb alpha) :=
           { | left _ =>  F_ (Canon alpha i)  i ;
@@
 alpha : E0
-notlimit : Is_Limit alpha = false
+notlimit : Limitb alpha = false
 nonzero : alpha <> Zero
@@
              | right nonzero
-                 with Utils.dec (Is_Limit alpha) :=
+                 with Utils.dec (Limitb alpha) :=
                  { | left _ => F_star (Canon alpha i,1) i ;
@@
 
-Lemma F_lim_eqn : forall alpha i,  Is_Limit alpha ->
+Lemma F_lim_eqn : forall alpha i,  Limitb alpha ->
                                F_ alpha i = F_ (Canon alpha i) i.

### 2020-07-09 latex/chapter-skippable.tex
renaming is_limit and is_succ

@@
 This notion is defined generically as a class in
- Module~\href{../src/html/hydras.Prelude.Ordinal-generic.html}{Prelude.Ordinal\_generic} (see Fig~\vref{figure:subsegment}).
+ Module~\href{../src/html/hydras.Prelude.Ordinal-generic.html}{Prelude.Ordinal\_generic} see Fig~\vref{figure:subsegment}).
 
@@
 
+The order on type \texttt{t} is already defined in
+Library~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}.
+
+\begin{Coqanswer}
+Inductive
+le_AsB (A B : Type) (leA : A -> A -> Prop) (leB : B -> B -> Prop)
+  : A + B -> A + B -> Prop :=
+    le_aa : forall x y : A, leA x y -> le_AsB A B leA leB (inl x) (inl y)
+  | le_ab : forall (x : A) (y : B), le_AsB A B leA leB (inl x) (inr y)
+  | le_bb : forall x y : B, leB x y -> le_AsB A B leA leB (inr x) (inr y)
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Definition lt : relation t := le_AsB _ _ Peano.lt Peano.lt.
+Infix "<" := lt : oo_scope.
+
+Definition le := clos_refl _ lt.
+
+Infix "<=" := le : oo_scope.
+\end{Coqsrc}
+
+\subsubsection{Finite ordinals as members of \texorpdfstring{$\omega+\omega$}{omega+omega}}
+
+It is easy do define a coercion from type \texttt{nat} into our type \texttt{t}.
+
+\begin{Coqsrc}
+Definition fin (n:nat) : t := inl n.
+Coercion fin : nat >-> t.
+\end{Coqsrc}
+
+The range of \texttt{fin} is called the set of \emph{finite} ordinals.
+
+\index{Coq!Techniques!Coercions} 
+\begin{remark}
+Beware of coercions and notation scopes!
+
+Let us consider the following goal:
+
+\begin{Coqsrc}
+ Goal (6 < 8).
+ auto with arith.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+1 subgoal (ID 9)
+
+  ============================
+  6 < 8
+\end{Coqanswer}
+
+Please keep in mind that the current notation scope interprets the infix \texttt{``<''} as the predicate \texttt{Omega\_plus\_omega.lt} and not \texttt{Nat.lt}.
+More,  the coercion mechanism converts the terms \texttt{6:nat} [resp.
+\texttt{8:nat} ]
+into \texttt{inl 6} [resp. \texttt{inl 8}].  So, the initial goal is correctly interpreted by \coq{}, but not as an inequality between two natural numbers.
+
+Anyway, the initial goal is provable, using \texttt{le\_AsB} first constructor.
+
+\begin{Coqsrc}
+  constructor; auto with arith.
+Qed.
+\end{Coqsrc}
+
+\end{remark}
+\subsubsection{Definition of \texorpdfstring{$\omega$}{omega}}
+
+We define $\omega$ as the least ordinal which is larger than any finite ordinal.
+
+\begin{Coqsrc}
+ Notation  "'omega'"  := (inr 0) : oo_scope. 
+\end{Coqsrc}
+
+\begin{exercise}
+Show that, for any $\alpha:t$, $\alpha$ is finite iff $\alpha<\omega$.
+\end{exercise}
+
+We say also that $\alpha$ is the \emph{limit} of the sequence $\textit{fin}\;i, i\in\mathbb{N}$, according to the followi,g definition (in~\href{../src/html/hydras.Prelude.Ordinal-generic.html}{Prelude.Ordinal\_generic}).
+$\alpha$ is the $\omega$-limit of the sequence $s$ if $\alpha$ is [strictly] larger than all the $s(i)$, and for any
+$\beta<\alpha$, $\beta$ is dimonated by at least an element of  the sequence $s$ (in fact $\beta$ is dominated by an infinite number of elements of the sequence (\emph{exercise!})).
+
+\begin{Coqsrc}
+Definition  Omega_limit
+            {A:Type}{lt : relation A}
+           {sto : StrictOrder lt} (s: nat -> A) (alpha:A)  :=
+  (forall i: nat, lt (s i) alpha) /\
+  (forall beta, lt beta  alpha -> exists i:nat, lt beta (s i)).
+
+Lemma omega_is_limit:
+   Omega_limit fin omega.
+\end{Coqsrc}
+
 \begin{itemize}
@@
 \item  $\omega^\omega$: the set of nonincreasing sequences of natural numbers, lexicographically ordered (also the set of finite multisets of natural numbers).
-\item $\epsilon_0$: The set of terms in Cantor normal form (see Chap.~\ref{chap/T1}).
+\item $\epsilon_0$: The set of terms in Cantor normal form (see Chap.~\ref{chap:T1}).
 \item $\Gamma_0$: The set of terms in Veblen normal form.

### 2020-07-09 latex/chapter-termination.tex
renaming is_limit and is_succ

@@
  \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#is_succ}{Epsilon0.T1}}
-
-\begin{Coqsrc}
-  Fixpoint is_succ alpha :=
+\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}
+
+\begin{Coqsrc}
+  Fixpoint succb alpha :=
   match alpha with
@@
     | ocons zero _ _ => true
-    | ocons alpha n beta => is_succ beta
+    | ocons alpha n beta => succb beta
   end.
 
-Fixpoint is_limit alpha :=
+Fixpoint limitb alpha :=
   match alpha with
@@
     | ocons alpha n zero => true
-    | ocons alpha n beta => is_limit beta
+    | ocons alpha n beta => limitb beta
   end.
@@
 \begin{Coqsrc}
-  Compute is_limit omega.
+  Compute limitb omega.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Compute is_succ 42.
+Compute succb 42.
 \end{Coqsrc}
@@
   Definition zero_succ_limit (alpha: T1) :
-  {is_succ alpha} + {is_limit alpha} +  {alpha=zero}.
+  {succb alpha} + {limitb alpha} +  {alpha=zero}.
   (* ... *)
@@
 \begin{Coqsrc}
- Lemma is_succ_iff alpha (Halpha : nf alpha) :
-  is_succ alpha <-> exists beta : T1, nf beta /\ alpha = succ  beta.
+ Lemma succb_iff alpha (Halpha : nf alpha) :
+  succb alpha <-> exists beta : T1, nf beta /\ alpha = succ  beta.
 \end{Coqsrc}

### 2020-07-10 latex/chapter-skippable.tex
better definoition of Ipred

@@
 
-We say also that $\alpha$ is the \emph{limit} of the sequence $\textit{fin}\;i, i\in\mathbb{N}$, according to the followi,g definition (in~\href{../src/html/hydras.Prelude.Ordinal-generic.html}{Prelude.Ordinal\_generic}).
+\subsection{Limit and successor ordinals}
+
+\index{Maths!Limit ordinals}
+
+We say also that $\omega$ is the \emph{limit} of the sequence $\textit{fin}\;i, i\in\mathbb{N}$, according to the followi,g definition (in~\href{../src/html/hydras.Prelude.Ordinal_generic.html}{Prelude.Ordinal\_generic}).
 $\alpha$ is the $\omega$-limit of the sequence $s$ if $\alpha$ is [strictly] larger than all the $s(i)$, and for any
@@
 
+\index{Exercises}
+
+\begin{exercise}
+Prove that $\omega$ is the unique   limit ordinal \emph{(in this precise notation system! In larger segments of ordinal, there will be \emph{a lot} of limit ordinals}).
+\end{exercise}
+
+Let $\alpha$ and $\beta$ be some ordinals, we say that \emph{$\alpha$ is an immediate predecessor of $\beta$}
+(or that \emph{$\beta$ is a  successor of $\alpha$ }) if $\alpha<\$beta$ and there is no ordinal $\gamma$ between $\alpha$ and $\gamma$.
+
+\noindent\emph{From Module~\href{../src/html/hydras.Prelude.Ordinal_generic.html}{Prelude.Ordinal\_generic}}
+
+\begin{Coqsrc}
+Definition Ipred {A:Type}{lt : relation A}
+           {sto : StrictOrder lt} (x y : A):=
+  lt x y /\ (forall z,  lt z y -> clos_refl _ lt z x).
+
+Definition Successor {A:Type}{lt : relation A}
+           {sto : StrictOrder lt} (y x : A):= Ipred x y.
+\end{Coqsrc}
+
 \begin{itemize}

### 2020-07-11 latex/chapter-hydras.tex
renaming zero_succ_limit-dec

@@
 % \end{remark}
-
-\section{Generic properties}
-
-\subsection{Liveliness}
-
-Let $B$ be an instance of \texttt{Battle}. We say that $B$ is \emph{alive} if
-for any configuration $(i,h)$, where $h$ is not a head, there exists a further step in class $B$.
-
-\vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
-
-\begin{Coqsrc}
-Definition Alive (B : Battle) :=
-  forall i h, 
-     h <> head -> {h' : Hydra |  B i h h'}.
-\end{Coqsrc}
-
-The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
-\url{../src/html/hydras.Hydra.Hydra_Theorems.html} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
-
-\begin{Coqsrc}
-Theorem Alive_free: Alive free.
-
-Theorem Alive_standard: Alive standard.  
-\end{Coqsrc}
-
-Both theorems are proved with the help of the  following strongly specified function:
-
-\vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
-
-\begin{Coqsrc}
-Definition  next_round_dec n :
- forall h , (h = head) + {h' : Hydra & {R1 h h'} + {R2  n h  h'}}.
-\end{Coqsrc}
-
-\subsection{Termination}
-
-The termination of all battles is naturally expressed by the predicate \texttt{well\_founded} defined in the module \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf} 
- of the Standard Library.
-
-\index{Predicates!Termination}
-
-\begin{Coqsrc}
-Definition Termination :=  well_founded (transp _ round).
-\end{Coqsrc}
-
-Let $b$ be an instance of class \texttt{Battle}. A \emph{variant} for $b$ consists
-in a well-founded relation $<$  on some type \texttt{A}, and a function
-(also called a \emph{measure}) \texttt{m:Hydra->A} such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $b$, the inequality $m(h')<m(h)$ holds.
-
-\vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
-\index{Coq!Techniques!Type classes}
-
-\begin{Coqsrc}
-Class Hvariant {A:Type}{Lt:relation A}(Wf: well_founded Lt)(b : Battle)
-                          (m: Hydra -> A): Prop :=
- {variant_decr :forall i h h', 
-     h <> head -> b i  h h' -> Lt (m h') (m h)}.
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
- Prove that, if there is an instance of \texttt{Hvariant Lt wf\_Lt b m}, then there exists no infinite battle in  \texttt{b}.
-\end{exercise}
-
-The example we just studied shows that the termination of any battle may take a very long time. If we want to study hydra battles in general, we have to consider 
-any hydra and any strategy, both for Hercules and the hydra itself. So, we first  give some definitions, generally borrowed from transition systems vocabulary (see~\cite{tel_2000} for instance).
-
-\subsection{Termination of all battles}
-
-% \subsection{Reachability}
-
-% Let $b$ be an instance of \texttt{battle}.  We say that a configuration $(i,h)$ is 
-% \emph{reachable} if there exists some initial hydra $h_0$ such that the transformation from $(0,h_0)$ into $(i,h)$ belongs to $b$.
-The following inductive definition considers also the case where $i=0$ and $h=h_0$.
-
-% \vspace{4pt}
-% \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-% \begin{Coqsrc}
-% Inductive reachable (b : Battle)
-%   : nat -> Hydra ->  Prop  :=
-% | reachable_0 : forall h1, reachable b 0 h1
-% | reachable_S : forall i h h0,  fight b  0 h0 i h ->
-%                                 reachable b i h.
-% \end{Coqsrc}
-
-% A nice property of the classes \texttt{free} and \texttt{standard} is that every configuration $(i,h)$ is reachable. It is sufficient to add $i$ heads to $h$'s foot and start the battle at time $0$.
-
-% \vspace{4pt}
-% \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-
-% \begin{Coqsrc}
-% Lemma reachable_free : forall i  h, reachable free i h.
-
-% Lemma reachable_standard  : forall h i, reachable standard i h.
-% \end{Coqsrc}
-
-\subsection{Liveliness}
-
-We say that a kind  $B$ of battles is \emph{alive} if for any configuration $(i,h)$, where $h$ is not a head, there exists a further step in class $B$.
-
-\vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
-
-\begin{Coqsrc}
-Definition Alive (B : Battle) :=
-  forall i h, 
-     h <> head -> {h' : Hydra |  B i h h'}.
-\end{Coqsrc}
 
@@
 
-\subsection{Some failed attempts}
+\section{Generic properties}
+
+\subsection{Liveliness}
+
+Let $B$ be an instance of \texttt{Battle}. We say that $B$ is \emph{alive} if
+for any configuration $(i,h)$, where $h$ is not a head, there exists a further step in class $B$.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
+
+\begin{Coqsrc}
+Definition Alive (B : Battle) :=
+  forall i h, 
+     h <> head -> {h' : Hydra |  B i h h'}.
+\end{Coqsrc}
+
+The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
+\url{../src/html/hydras.Hydra.Hydra_Theorems.html} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
+
+\begin{Coqsrc}
+Theorem Alive_free: Alive free.
+
+Theorem Alive_standard: Alive standard.  
+\end{Coqsrc}
+
+Both theorems are proved with the help of the  following strongly specified function:
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
+
+\begin{Coqsrc}
+Definition  next_round_dec n :
+ forall h , (h = head) + {h' : Hydra & {R1 h h'} + {R2  n h  h'}}.
+\end{Coqsrc}
+
+\subsection{Termination}
+
+The termination of all battles is naturally expressed by the predicate \texttt{well\_founded} defined in the module \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf} 
+ of the Standard Library.
+
+\index{Predicates!Termination}
+
+\begin{Coqsrc}
+Definition Termination :=  well_founded (transp _ round).
+\end{Coqsrc}
+
+Let $b$ be an instance of class \texttt{Battle}. A \emph{variant} for $b$ consists
+in a well-founded relation $<$  on some type \texttt{A}, and a function
+(also called a \emph{measure}) \texttt{m:Hydra->A} such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $b$, the inequality $m(h')<m(h)$ holds.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
+\index{Coq!Techniques!Type classes}
+
+\begin{Coqsrc}
+Class Hvariant {A:Type}{Lt:relation A}(Wf: well_founded Lt)(b : Battle)
+                          (m: Hydra -> A): Prop :=
+ {variant_decr :forall i h h', 
+     h <> head -> b i  h h' -> Lt (m h') (m h)}.
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+ Prove that, if there is an instance of \texttt{Hvariant Lt wf\_Lt b m}, then there exists no infinite battle in  \texttt{b}.
+\end{exercise}
+
+The example we just studied shows that the termination of any battle may take a very long time. If we want to study hydra battles in general, we have to consider 
+any hydra and any strategy, both for Hercules and the hydra itself. So, we first  give some definitions, generally borrowed from transition systems vocabulary (see~\cite{tel_2000} for instance).
+
+\subsection{Some failed attempts to prove termination}
 The class \texttt{Hvariant} is parameterized by a well-founded order $<$ on some type $A$, a class of battles $B$ and a measure $m$. When $<$ or $m$ are 
@@
 Let us present the main steps of that proof, the script of which  is in the module ~\href{../src/html/hydras.Hydra.Omega_Small.html}{Hydra/Omega\_Small.v} \footnote{ The name of this file means ``the ordinal $\omega$ is too small for proving the termination of [free] hydra battles ''.
-In effect, the elements of $\omega$, considered as a set, are just the natural numbers.}.
+In effect, the elements of $\omega$, considered as a set, are just the natural numbers (see next chapter for more details)}.
 

### 2020-07-11 latex/chapter-ordinals.tex
renaming zero_succ_limit-dec

--- 2020-07-10 latex/chapter-skippable.tex
@@
 \begin{exercise}
-Prove that $\omega$ is the unique   limit ordinal \emph{(in this precise notation system! In larger segments of ordinal, there will be \emph{a lot} of limit ordinals}).
+Prove that $\omega$ is the unique   limit ordinal \emph{(in this precise notation system! In larger segments of 
+ordinals, there will be \emph{a lot} of limit ordinals}).
 \end{exercise}
@@
            {sto : StrictOrder lt} (x y : A):=
-  lt x y /\ (forall z,  lt z y -> clos_refl _ lt z x).
+  lt x y /\ (forall z,  lt x z -> lt z y -> False).
 
@@
 
+\begin{todo}
+trichotomy zero/succ/limit
+boolean functions limitb, succb
+
+\end{todo}
 \begin{itemize}

### 2020-07-13 latex/chapter-hydras.tex
more text on omega plus opmega

@@
 
+The example we just studied shows that the termination of any battle may take a very long time. If we want to study hydra battles in general, we have to consider 
+any hydra and any strategy, both for Hercules and the hydra itself. So, we first  give some definitions, generally borrowed from transition systems vocabulary (see~\cite{tel_2000} for instance).
+
 \subsection{Liveliness}
@@
 
-The example we just studied shows that the termination of any battle may take a very long time. If we want to study hydra battles in general, we have to consider 
-any hydra and any strategy, both for Hercules and the hydra itself. So, we first  give some definitions, generally borrowed from transition systems vocabulary (see~\cite{tel_2000} for instance).
-
-\subsection{Some failed attempts to prove termination}
+\subsection{A first theorem of impossibility}
+
 The class \texttt{Hvariant} is parameterized by a well-founded order $<$ on some type $A$, a class of battles $B$ and a measure $m$. When $<$ or $m$ are 
@@
 
-In this section, we present some naive attempts, where we fail to build proof of termination of free battles.
+We show that this is the case of the type \texttt{nat} of Peano natural numbers, provided with the usual strict order $<$.
 
 \subsubsection{Using the Hydra's height as a variant}
-A first plan to prove termination of all hydra battles is to use a simple
-measure that maps any hydra to a natural number. For instance, let us check whether 
- the function \emph{height} could be such  a variant.
-
-Unfortunately, that does not work.
+Trivially, a measure like the height will not work for building a variant.
+
 Please consider the hydras of Fig.~\vref{fig:h-omega2-small} and
@@
 \subsubsection{Using any variant defined on \texorpdfstring{nat}{nat}}
-\label{omega-case}
+\label{sect:omega-case}
 We could imagine that our previous failure is due to the choice of \texttt{height} as a measure, and that a more complex function would work.
@@
 
-\subsubsection{Lexicographic order on \texorpdfstring{nat*nat}{nat*nat}}
-\label{omega2-case}
-
-We  prove now that even the type \texttt{nat * nat}, provided with the lexicographic product of \texttt{(nat,<)} by itself is too simple for proving the termination of all hydra battles.
-This impossibility result will prevent us from considering measures like the following one:
-
-  \begin{Coqsrc}
-  Let m h = (height h, hsize h).
-  \end{Coqsrc}
-
-The proof we are going to develop has exactly the same structure as the previous one. Nevertheless, the proof of technical  lemmas is a little more complex, due to 
- the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
-Consider for instance that there exists an infinite number of pairs between
-$(1,0)$ and $(2,0)$.
-
-\begin{remark}
-  The order structure we consider in this section is also known as the ordinal
-  $\omega^2$. We identify any pair $(i,j)\in \mathbb{N}\times\mathbb{N}$ with the ordinal $\omega\times i + j$. Thus the three kinds of ordinals in $\omega^2$
-  are represented as follows:
-  \begin{description}
-  \item[null ordinal] : the pair $(0,0)$
-  \item[successor ordinal] : any pair $(i,j)$ where $j>0$
-  \item[limit ordinal] : any pair $(i,0)$ where $i>0$
-  \end{description}
-\end{remark}
-
-The detailed  proof script is in the file \url{../src/html/hydras.Hydra.Omega2_Small.html}.
-
-\subsubsection{Preliminaries}
-Let us assume there is a variant from \texttt{Hydra} into \texttt{nat*nat} (with the
-   lexicographic ordering)  for proving the   termination of all hydra battles.
-
-\vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
-
-\begin{Coqsrc}
-  Section Impossibility_Proof.
-
-  Let t := (nat * nat)%type.
-
-  (** non-dependent lexicographic strict ordering on nat*nat *)
-
-  Let lt2 : relation t := lexico Peano.lt Peano.lt.
-
-  Infix "<" := lt2.
-
-  (** reflexive closure of lt2 *)
-  Let le2 := clos_refl _ lt2.
-  Infix "<=" := le2.
-
-  Variable m : Hydra -> t.
-
-  Context (Hvar : Hvariant lt2_wf free m).
-\end{Coqsrc}
-
-Let us follow the same pattern as in Sect.~\ref{omega-case}.
-First, we define an injection from type \texttt{t} into \texttt{Hydra}.
-We associate with any pair $(i,j)$ the hydra with $i$ branches of length $2$ and
-$j$ branches of length $1$.
-
-%% revenir ici
-
-\vspace{4pt}
-\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
-
-\begin{Coqsrc}
-  Let iota (p: t) := 
-    node (hcons_mult (hyd1 head) (fst p)
-                     (hcons_mult head (snd p) hnil)).
-\end{Coqsrc}
-
-For instance, Figure~\vref{fig:essai2} shows the hydra associated to the pair
-$(3,5)$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-\node (foot) at (6,0) {$\bullet$};
-\node (N1) at (1,3) {$\bullet$};
-\node (N2) at (3,3) {$\bullet$};
-\node (N3) at (5,3) {$\bullet$};
-\node (N4) at (8,3) {$\Smiley[2][green]$};
-\node (N5) at (11,3) {$\Smiley[2][green]$};
-\node (N6) at (14,3) {$\Smiley[2][green]$};
-\node (N7) at (17,3){$\Smiley[2][green]$};
-\node (N8) at (20,3){$\Smiley[2][green]$};
-\node  (N9) at (0,5) {$\Smiley[2][green]$};
-\node (N10) at (2,5) {$\Smiley[2][green]$};
-\node (N11) at (4,5) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N5);
-\draw (foot) -- (N6);
-\draw (foot) to [bend right=10] (N7);
-\draw (foot) to [bend right=15] (N8);
-\draw (N1) to [bend left=10] (N9);
-\draw (N2) -- (N10);
-\draw (N3) -- (N11);
-\end{tikzpicture}
-\caption{\label{fig:essai2}
-The hydra $\iota(3,5)$}
-\end{figure}
-
-Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:h-omega2-small}}
- The hydra \texttt{big\_h}.
-\end{figure}
-\begin{Coqsrc}
-   Let big_h := hyd1 (hyd2 head head).  
- \end{Coqsrc}
-
- In a second step, we build a ``smaller'' hydra.
-
-\begin{Coqsrc}
-   Let small_h := iota (m big_h).
-\end{Coqsrc}
-
-Like in Sect.~\ref{omega-case}, we prove the double inequality \texttt{m big\_h <= m small\_h < m big\_h}, which is impossible.
-
-\subsubsection{Proof of the inequality \texttt{m small\_h < m big\_h}}
-
-For proving the inequality  \texttt{m\_lt: m small\_h < m big\_h}, it suffices to
-build a fight transforming \texttt{big\_h} into \texttt{small\_h}.
-
-First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h}
-into the pair $(i,j)$.
-If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
-If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the folowing result.
-
-\begin{Coqsrc}
- Lemma big_to_small: big_h -+-> small_h.
-\end{Coqsrc}
-
-Since $m$ is a variant, we infer the following inequality:
-
-\begin{Coqsrc}
-Corollary m_lt : m small_h < m big_h.
-\end{Coqsrc}
-
-\subsubsection{Proof of the inequality \texttt{m big\_h <= m small\_h} }
-
-The proof of the inequality \texttt{m big\_h <= m small\_h} is quite more complex than in Sect~\ref{omega-case}.  If we consider some pair $(i,j)$, where $i>0$, there exists an infinite number of
-pairs stricly less than $(i,j)$, and there exists an infinite number of battles that start from
-$\iota(i,j)$. In effect, at any configuration $\iota(k,0)$, the hydra can freely chose any replication number. Intuitively, the measure of such a hydra must be large enough for taking into account
-all the possible battles issued from that hydra.
-Let us give more technical details.
-
-\begin{itemize}
-\item The proof of the lemma \texttt{m\_ge : forall p : t,   p <= m (iota p)} uses well-founded induction on $p$, and not structural induction on natural numbers
-
-\item For any pair $p$, we have to distinguish between three cases, according to the value of $p$'s components.
-  \begin{itemize}
-  \item $p=(0,0)$
-  \item $p=(i,0)$, where $i>0$\,: $p$ corresponds to a limit ordinal
-  \item $p=(i,j)$, where $j>0$\,: $p$ is the successor of $(i,j-1)$.
-  \end{itemize}
-\end{itemize}
-
-Before starting the proof, we have to express the notion of \emph{limit} in terms of
-least upper bounds, through the following logical equivalence.
-
-\vspace{4pt}
-\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#limit_is_lub}{Hydra.Omega2\_Small.v}}.
-
-\begin{Coqsrc}
-Lemma limit_is_lub : forall i p, 
-   (forall j, (i,j) < p) <-> (S i, 0) <= p.  
-\end{Coqsrc}
-
-Let us define the notion of elementary ``step'' of decreasing sequences in
-\texttt{t}
-
-\begin{Coqsrc}
-Inductive step : t -> t -> Prop :=
-| succ_step : forall i j,  step (i, S j) (i, j)
-| limit_step : forall i j, step (S i, 0) (i, j).
-\end{Coqsrc}
-
-The following lemma establishes a correspondance between the relation
-\texttt{step} and hydra fights.
-
-\begin{Coqsrc}
-Lemma step_to_fight : forall p q, step p q -> iota p -+-> iota q.
-\end{Coqsrc}
-
-\index{Maths!Transfinite induction}
-
-Thus, starting from any inequality $q < p$ on type \texttt{t}, we can build 
-by transfinite induction over \texttt{p} a fight 
-that transforms the hydra $\iota(p)$ into $\iota(q)$.
-
-\vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}
-
-\begin{Coqsrc}
-Lemma m_ge : forall p : t,   p <= m (iota p).
-Proof.
-  intro p ; pattern p;
-    apply  well_founded_induction with 
-               (R := lt2) (1:= wf_lexico lt_wf lt_wf);
-     intros (i,j) IHij (* rest of proof skipped *)
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  i, j : nat
-  IHij : forall y : t, y < (i, j) -> y <= m (iota y)
-  ============================
-  (i, j) <= m (iota (i, j)) 
-\end{Coqanswer}
-
-Then we have  three cases to consider, according to the values of $i$ and $j$.
-\begin{itemize}
-\item If $p=(0,0)$ then obviously, $\iota(p)\geq p = (0,0)$
-\item If  $p=(i+1,0)$ for some $i\in\mathbb{N}$, we
- remark  that $p$ is strictly greater than any pair $ (i, j)$, where $j$ 
-is any natural number.
-
-Applying the battle rules, for any $j$, we have $\iota(i+1,j)  {\round} \iota(i, j) $, thus $m(\iota(p)) > m(\iota(i,j)$ since  $m$ is assumed to be a variant.
-
-Applying the induction hypothesis, we get the inequality
- $ m(\iota(i,j)) \geq (i,j)$ for any $j$.
-
-Thus, $m(\iota(p)) > (i,j)$ for any $j$.
-Applying the lemma \texttt{limit\_is\_lub}, we get  the inequality
-$\iota(i+1,0)\geq (i+1,0)$
-
-\item If $p=(i,j+1)$ with $j\in\mathbb{N}$, we have  $\iota(p)  {\round} \iota(i, j) $,
-hence $m(\iota(p))> m(\iota(i,j)) \geq (i,j)$, thus $m(\iota(p))\geq (i,j+1)=p$
-
-\end{itemize}
-
-\subsubsection{End of the proof}
-
-Since $<$ is a strict order (irreflexive  and transitive) on \texttt{nat*nat}, we can conclude that there is no
-variant for termination on the lexicographic square of $(\mathbb{N},<)$.
-
-\begin{enumerate}
-\item From \texttt{m\_lt}, we infer the strict inequality 
-\texttt{m small\_h < m big\_h}
-\item from \texttt{m\_ge}, we get \texttt{m big\_h <= m (iota (m big\_h)) = m small\_h} 
-\end{enumerate}
-
-\vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
-
-\begin{Coqsrc}
-Theorem Impossible : False.
-Proof.
-  destruct (StrictOrder_Irreflexive  (m big_h)).
-  apply le2_lt2_trans with (m small_h).
-  -  unfold small_h;  apply m_ge.
-  -  apply m_lt. 
-Qed.
-
-End Impossibility_Proof.
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into \texttt{nat*nat} for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\begin{remark}
-In Chapter~\ref{ks-chapter}, we will prove a much more general  theorem
-than the two previous results about $\mathbb{N}$ and $\mathbb{N}\times\mathbb{N}$. The proof of that general  result will share the same structure, but will 
-require a lot of technical results. 
-\end{remark}
-
-\begin{exercise}
-
-\label{sec:orgheadline63}
-Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.
-
-\begin{itemize}
-\item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
-  lexicographic ordering
-\item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.
-
-For instance , the sequence $\langle 4,4,2 \rangle$ is strictly greater than 
-$\langle 4,3,3,3,3,3,3,2,2,2 \rangle$.
-\end{itemize}
-
-\end{exercise}
+\subsubsection{Conclusion/Teaser}
+
+In order to build a variant for proving the termination of all hydra battles, we need to consider order structures more complex than the usual order on type \texttt{nat}.
+The next chapter will present the notion of \emph{ordinal notation} from the point view of the \coq{} user: inductive types, functions, type classes, etc.
+The ordinal $\epsilon_0$, the first one allowing us to build our desired variant, will be studied in detail in 
+Chapter~\ref{chap:T1}.

### 2020-07-13 latex/chapter-ordinals.tex
more text on omega plus opmega

@@
 \chapter{Introduction to ordinal numbers with the Coq proof assistant}
+
+\centerline{\LARGE {\bf work in progress!}}
 
@@
 
+\begin{remark}
+It is important to keep in mind  that the integer $n$ is not an ``element'' of \texttt{FinOrd $n$}.
+In set-theoretic presentations of ordinals, the set associated with the ordinal $n$ is $\{0,1,\dots,n-1\}$.
+In our formalization, the interpretation of an ordinal as a set is realized by the following definition
+(in ~\href{../src/html/hydras.Prelude.Ordinal_generic.html}{Prelude.Ordinal\_generic}).
+
+\begin{Coqsrc}
+Definition bigO `{nA : @OrdinalNotation A ltA stoA compareA}
+           (a: A) : Ensemble A :=
+  fun x: A => ltA x a.
+\end{Coqsrc}
+\end{remark}
+
 Finally, it is interesting to show that, if $i<j$, then the segment $[0,i[$ is a ``sub-segment'' of
@@
 \end{remark}
+
 \subsubsection{Definition of \texorpdfstring{$\omega$}{omega}}
@@
 
-\subsection{Limit and successor ordinals}
+\subsubsection{Building an instance of OrdinalNotation}
+
+The class \texttt{OrdinalNotation} is parameterized by the strictorder $<$ and a correct comparison function.
+
+First, we prove that \texttt{lt} is an instance of \texttt{StrictOrder}.
+
+\begin{Coqsrc}
+Instance lt_strorder : StrictOrder lt.
+Qed.
+\end{Coqsrc}
+
+Then, we define a comparison function by pattern matching over the sum type $t$.
+
+\begin{Coqsrc}
+Definition compare (alpha beta: t) : comparison :=
+   match alpha, beta with
+     inl _, inr _ => Lt
+   | inl n, inl p | inr n, inr p => Nat.compare n p
+   | inr _, inl _ => Gt
+  end.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Compute compare 7 omega.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+= Lt : comparison
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Compute compare  (inr 2) 42.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+= Gt : comparison
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Compute compare (inr 0) omega.  
+\end{Coqsrc}
+
+ \begin{Coqanswer}
+= Eq : comparison
+\end{Coqanswer}
+
+Correctness of the function \texttt{compare} is asserted by the following lemma:
+
+\begin{Coqsrc}
+Lemma compare_correct alpha beta :
+    CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
+              (compare alpha beta).
+\end{Coqsrc}
+
+%%%% ICI %%%%
+
+\subsubsection{Limit and successor ordinals}
+
+The notions of limit and successor are defined in the module 
+\href{../src/html/hydras.Prelude.Limits_and_Successors.html}{Prelude.Limits\_and\_Successors} for any instance of
+\texttt{StrictOrder}.
 
@@
 
-We say also that $\omega$ is the \emph{limit} of the sequence $\textit{fin}\;i, i\in\mathbb{N}$, according to the followi,g definition (in~\href{../src/html/hydras.Prelude.Ordinal_generic.html}{Prelude.Ordinal\_generic}).
-$\alpha$ is the $\omega$-limit of the sequence $s$ if $\alpha$ is [strictly] larger than all the $s(i)$, and for any
-$\beta<\alpha$, $\beta$ is dimonated by at least an element of  the sequence $s$ (in fact $\beta$ is dominated by an infinite number of elements of the sequence (\emph{exercise!})).
+Let $<$ be a strict order on a type $A$ .
+We say $x:A$ is an  $\omega$-\emph{limit} of  a sequence $s(i), i\in\mathbb{N}$, if
+ if $x$ is [strictly] larger than all the $s(i)$, and for any
+$y<x$, $y$ is dominated by at least an element of  the sequence $s$.
+
+\vspace{4pt}
+
+\noindent\emph{From Module~\href{../src/html/hydras.Prelude.Limits_and_Successors.html}{Prelude.Limits\_and\_Successors}}
 
@@
             {A:Type}{lt : relation A}
-           {sto : StrictOrder lt} (s: nat -> A) (alpha:A)  :=
-  (forall i: nat, lt (s i) alpha) /\
-  (forall beta, lt beta  alpha -> exists i:nat, lt beta (s i)).
+           {sto : StrictOrder lt} (s: nat -> A) (x:A)  :=
+  (forall i: nat, lt (s i) x) /\
+  (forall y, lt y  x -> exists i:nat, lt y (s i)).
 
-Lemma omega_is_limit:
-   Omega_limit fin omega.
+Definition Limit   {A:Type}{lt : relation A}
+           {sto : StrictOrder lt}  (x:A)  :=
+  exists s : nat -> A, Omega_limit s x.
  \end{Coqsrc}
 
-\index{Exercises}
-
-\begin{exercise}
-Prove that $\omega$ is the unique   limit ordinal \emph{(in this precise notation system! In larger segments of 
-ordinals, there will be \emph{a lot} of limit ordinals}).
-\end{exercise}
-
-Let $\alpha$ and $\beta$ be some ordinals, we say that \emph{$\alpha$ is an immediate predecessor of $\beta$}
-(or that \emph{$\beta$ is a  successor of $\alpha$ }) if $\alpha<\$beta$ and there is no ordinal $\gamma$ between $\alpha$ and $\gamma$.
-
-\noindent\emph{From Module~\href{../src/html/hydras.Prelude.Ordinal_generic.html}{Prelude.Ordinal\_generic}}
+Let $x$ and $y$ be two terms of type $A$, we say that \emph{$x$ is an immediate predecessor of $y$}
+(or that \emph{$y$ is a  successor of $x$ }) if $x<y$ and there is no $z$ between $x$ and $y$.
 
@@
 \end{Coqsrc}
+
+For instance, 
+we prove that $\omega$ is the limit of the sequence  of finite ordinals.
+
+  \begin{Coqsrc}
+Lemma omega_is_limit: Omega_limit fin omega.
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+Prove that $\omega$ is the unique   limit ordinal within the notation system for $\omega+\omega$
+\emph{(in this precise notation system! In larger segments of 
+ordinals, there will be \emph{a lot} of limit ordinals}).
+\end{exercise}
+
+\subsection{Computing in \texorpdfstring{$\omega+\omega$}{omega+omega}}
+
+The simple data type we use for representing the ordinals less than $\omega+\omega$ allows us to compute successors, and to decide properties like \texttt{Successor} or \texttt{Limit}.
 
@@
 \end{itemize}
+
+\begin{todo}
+The following section has moved from chaper on hydras.
+Adapt the redaction to the $\omega^2$ ordinl segment.
+\end{todo}
+
+\subsubsection{Lexicographic order on \texorpdfstring{nat*nat}{nat*nat}}
+\label{omega2-case}
+
+We  prove now that even the type \texttt{nat * nat}, provided with the lexicographic product of \texttt{(nat,<)} by itself is too simple for proving the termination of all hydra battles.
+This impossibility result will prevent us from considering measures like the following one:
+
+  \begin{Coqsrc}
+  Let m h = (height h, hsize h).
+  \end{Coqsrc}
+
+The proof we are going to develop has exactly the same structure as in Section~\ref{sect:omega:case}
+
+previous one. Nevertheless, the proof of technical  lemmas is a little more complex, due to 
+ the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
+Consider for instance that there exists an infinite number of pairs between
+$(1,0)$ and $(2,0)$.
+
+\begin{remark}
+  The order structure we consider in this section is also known as the ordinal
+  $\omega^2$. We identify any pair $(i,j)\in \mathbb{N}\times\mathbb{N}$ with the ordinal $\omega\times i + j$. Thus the three kinds of ordinals in $\omega^2$
+  are represented as follows:
+  \begin{description}
+  \item[null ordinal] : the pair $(0,0)$
+  \item[successor ordinal] : any pair $(i,j)$ where $j>0$
+  \item[limit ordinal] : any pair $(i,0)$ where $i>0$
+  \end{description}
+\end{remark}
+
+The detailed  proof script is in the file \url{../src/html/hydras.Hydra.Omega2_Small.html}.
+
+\subsubsection{Preliminaries}
+Let us assume there is a variant from \texttt{Hydra} into \texttt{nat*nat} (with the
+   lexicographic ordering)  for proving the   termination of all hydra battles.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
+
+\begin{Coqsrc}
+  Section Impossibility_Proof.
+
+  Let t := (nat * nat)%type.
+
+  (** non-dependent lexicographic strict ordering on nat*nat *)
+
+  Let lt2 : relation t := lexico Peano.lt Peano.lt.
+
+  Infix "<" := lt2.
+
+  (** reflexive closure of lt2 *)
+  Let le2 := clos_refl _ lt2.
+  Infix "<=" := le2.
+
+  Variable m : Hydra -> t.
+
+  Context (Hvar : Hvariant lt2_wf free m).
+\end{Coqsrc}
+
+Let us follow the same pattern as in Sect.~\ref{omega-case}.
+First, we define an injection from type \texttt{t} into \texttt{Hydra}.
+We associate with any pair $(i,j)$ the hydra with $i$ branches of length $2$ and
+$j$ branches of length $1$.
+
+%% revenir ici
+
+\vspace{4pt}
+\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
+
+\begin{Coqsrc}
+  Let iota (p: t) := 
+    node (hcons_mult (hyd1 head) (fst p)
+                     (hcons_mult head (snd p) hnil)).
+\end{Coqsrc}
+
+For instance, Figure~\vref{fig:essai2} shows the hydra associated to the pair
+$(3,5)$.
+
+\begin{figure}[htb]
+\centering
+\begin{tikzpicture}[very thick, scale=0.4]
+\node (foot) at (6,0) {$\bullet$};
+\node (N1) at (1,3) {$\bullet$};
+\node (N2) at (3,3) {$\bullet$};
+\node (N3) at (5,3) {$\bullet$};
+\node (N4) at (8,3) {$\Smiley[2][green]$};
+\node (N5) at (11,3) {$\Smiley[2][green]$};
+\node (N6) at (14,3) {$\Smiley[2][green]$};
+\node (N7) at (17,3){$\Smiley[2][green]$};
+\node (N8) at (20,3){$\Smiley[2][green]$};
+\node  (N9) at (0,5) {$\Smiley[2][green]$};
+\node (N10) at (2,5) {$\Smiley[2][green]$};
+\node (N11) at (4,5) {$\Smiley[2][green]$};
+\draw (foot) to [bend left=10] (N1);
+\draw (foot) -- (N2);
+\draw (foot) -- (N3);
+\draw (foot) -- (N4);
+\draw (foot) -- (N5);
+\draw (foot) -- (N6);
+\draw (foot) to [bend right=10] (N7);
+\draw (foot) to [bend right=15] (N8);
+\draw (N1) to [bend left=10] (N9);
+\draw (N2) -- (N10);
+\draw (N3) -- (N11);
+\end{tikzpicture}
+\caption{\label{fig:essai2}
+The hydra $\iota(3,5)$}
+\end{figure}
+
+Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).
+
+\begin{figure}[htb]
+\centering
+\begin{tikzpicture}[very thick, scale=0.5]
+\node (foot) at (2,0) {$\bullet$};
+\node (N1) at (2,2) {$\bullet$};
+\node (N2) at (3,4) {$\Smiley[2][green]$};
+\node (N3) at (1,4) {$\Smiley[2][green]$};
+\draw (foot) -- (N1);
+\draw (N1) to [bend right =15] (N2);
+\draw (N1) to  [bend left=15](N3);
+\end{tikzpicture}
+\caption{\label{fig:h-omega2-small}}
+ The hydra \texttt{big\_h}.
+\end{figure}
+\begin{Coqsrc}
+   Let big_h := hyd1 (hyd2 head head).  
+ \end{Coqsrc}
+
+ In a second step, we build a ``smaller'' hydra.
+
+\begin{Coqsrc}
+   Let small_h := iota (m big_h).
+\end{Coqsrc}
+
+Like in Sect.~\ref{omega-case}, we prove the double inequality \texttt{m big\_h <= m small\_h < m big\_h}, which is impossible.
+
+\subsubsection{Proof of the inequality \texttt{m small\_h < m big\_h}}
+
+For proving the inequality  \texttt{m\_lt: m small\_h < m big\_h}, it suffices to
+build a fight transforming \texttt{big\_h} into \texttt{small\_h}.
+
+First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h}
+into the pair $(i,j)$.
+If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
+If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the folowing result.
+
+\begin{Coqsrc}
+ Lemma big_to_small: big_h -+-> small_h.
+\end{Coqsrc}
+
+Since $m$ is a variant, we infer the following inequality:
+
+\begin{Coqsrc}
+Corollary m_lt : m small_h < m big_h.
+\end{Coqsrc}
+
+\subsubsection{Proof of the inequality \texttt{m big\_h <= m small\_h} }
+
+The proof of the inequality \texttt{m big\_h <= m small\_h} is quite more complex than in Sect~\ref{omega-case}.  If we consider some pair $(i,j)$, where $i>0$, there exists an infinite number of
+pairs stricly less than $(i,j)$, and there exists an infinite number of battles that start from
+$\iota(i,j)$. In effect, at any configuration $\iota(k,0)$, the hydra can freely chose any replication number. Intuitively, the measure of such a hydra must be large enough for taking into account
+all the possible battles issued from that hydra.
+Let us give more technical details.
+
+\begin{itemize}
+\item The proof of the lemma \texttt{m\_ge : forall p : t,   p <= m (iota p)} uses well-founded induction on $p$, and not structural induction on natural numbers
+
+\item For any pair $p$, we have to distinguish between three cases, according to the value of $p$'s components.
+  \begin{itemize}
+  \item $p=(0,0)$
+  \item $p=(i,0)$, where $i>0$\,: $p$ corresponds to a limit ordinal
+  \item $p=(i,j)$, where $j>0$\,: $p$ is the successor of $(i,j-1)$.
+  \end{itemize}
+\end{itemize}
+
+Before starting the proof, we have to express the notion of \emph{limit} in terms of
+least upper bounds, through the following logical equivalence.
+
+\vspace{4pt}
+\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#limit_is_lub}{Hydra.Omega2\_Small.v}}.
+
+\begin{Coqsrc}
+Lemma limit_is_lub : forall i p, 
+   (forall j, (i,j) < p) <-> (S i, 0) <= p.  
+\end{Coqsrc}
+
+Let us define the notion of elementary ``step'' of decreasing sequences in
+\texttt{t}
+
+\begin{Coqsrc}
+Inductive step : t -> t -> Prop :=
+| succ_step : forall i j,  step (i, S j) (i, j)
+| limit_step : forall i j, step (S i, 0) (i, j).
+\end{Coqsrc}
+
+The following lemma establishes a correspondance between the relation
+\texttt{step} and hydra fights.
+
+\begin{Coqsrc}
+Lemma step_to_fight : forall p q, step p q -> iota p -+-> iota q.
+\end{Coqsrc}
+
+\index{Maths!Transfinite induction}
+
+Thus, starting from any inequality $q < p$ on type \texttt{t}, we can build 
+by transfinite induction over \texttt{p} a fight 
+that transforms the hydra $\iota(p)$ into $\iota(q)$.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}
+
+\begin{Coqsrc}
+Lemma m_ge : forall p : t,   p <= m (iota p).
+Proof.
+  intro p ; pattern p;
+    apply  well_founded_induction with 
+               (R := lt2) (1:= wf_lexico lt_wf lt_wf);
+     intros (i,j) IHij (* rest of proof skipped *)
+\end{Coqsrc}
+
+\begin{Coqanswer}
+  i, j : nat
+  IHij : forall y : t, y < (i, j) -> y <= m (iota y)
+  ============================
+  (i, j) <= m (iota (i, j)) 
+\end{Coqanswer}
+
+Then we have  three cases to consider, according to the values of $i$ and $j$.
+\begin{itemize}
+\item If $p=(0,0)$ then obviously, $\iota(p)\geq p = (0,0)$
+\item If  $p=(i+1,0)$ for some $i\in\mathbb{N}$, we
+ remark  that $p$ is strictly greater than any pair $ (i, j)$, where $j$ 
+is any natural number.
+
+Applying the battle rules, for any $j$, we have $\iota(i+1,j)  {\round} \iota(i, j) $, thus $m(\iota(p)) > m(\iota(i,j)$ since  $m$ is assumed to be a variant.
+
+Applying the induction hypothesis, we get the inequality
+ $ m(\iota(i,j)) \geq (i,j)$ for any $j$.
+
+Thus, $m(\iota(p)) > (i,j)$ for any $j$.
+Applying the lemma \texttt{limit\_is\_lub}, we get  the inequality
+$\iota(i+1,0)\geq (i+1,0)$
+
+\item If $p=(i,j+1)$ with $j\in\mathbb{N}$, we have  $\iota(p)  {\round} \iota(i, j) $,
+hence $m(\iota(p))> m(\iota(i,j)) \geq (i,j)$, thus $m(\iota(p))\geq (i,j+1)=p$
+
+\end{itemize}
+
+\subsubsection{End of the proof}
+
+Since $<$ is a strict order (irreflexive  and transitive) on \texttt{nat*nat}, we can conclude that there is no
+variant for termination on the lexicographic square of $(\mathbb{N},<)$.
+
+\begin{enumerate}
+\item From \texttt{m\_lt}, we infer the strict inequality 
+\texttt{m small\_h < m big\_h}
+\item from \texttt{m\_ge}, we get \texttt{m big\_h <= m (iota (m big\_h)) = m small\_h} 
+\end{enumerate}
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
+
+\begin{Coqsrc}
+Theorem Impossible : False.
+Proof.
+  destruct (StrictOrder_Irreflexive  (m big_h)).
+  apply le2_lt2_trans with (m small_h).
+  -  unfold small_h;  apply m_ge.
+  -  apply m_lt. 
+Qed.
+
+End Impossibility_Proof.
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+Prove that there exists no variant $m$ from \texttt{Hydra} into \texttt{nat*nat} for proving
+    the  termination of all \emph{standard} battles.
+\end{exercise}
+
+\begin{remark}
+In Chapter~\ref{ks-chapter}, we will prove a much more general  theorem
+than the two previous results about $\mathbb{N}$ and $\mathbb{N}\times\mathbb{N}$. The proof of that general  result will share the same structure, but will 
+require a lot of technical results. 
+\end{remark}
+
+\begin{exercise}
+
+\label{sec:orgheadline63}
+Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.
+
+\begin{itemize}
+\item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
+  lexicographic ordering
+\item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.
+
+For instance , the sequence $\langle 4,4,2 \rangle$ is strictly greater than 
+$\langle 4,3,3,3,3,3,3,2,2,2 \rangle$.
+\end{itemize}
+
+\end{exercise}

### 2020-07-14 latex/chapter-ordinals.tex
Limits_and_Successors now autonomous

@@
 
+\begin{Coqsrc}
+Definition succ (alpha : t) := match alpha with
+                                 inl i => inl (S i)
+                               | inr i => inr (S i)
+                               end.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma Successor_correct alpha beta : Successor beta alpha <-> beta = succ alpha.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Definition succb (alpha: t) := match alpha with
+                                   inr (S  _) | inl (S _) => true
+                                 | _ => false
+                                 end.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma succb_correct alpha : succb alpha <-> exists beta,  alpha = succ beta.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Definition limitb (alpha : t) := match alpha with
+                                     (inr 0) => true
+                                   | _ => false
+                                   end.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma omega_is_limit:
+   Omega_limit fin omega.
+Proof.
+  split.
+  - constructor.
+  - inversion_clear  1.
+    + exists (S x); constructor; auto with arith.
+    + lia.
+Qed.
+\end{Coqsrc}
+
 \begin{todo}

### 2020-07-14 latex/chapter-ordinals.tex
Limits_and_Successors now autonomous

@@
 
+\begin{todo}
+  Well foundedness; instance of OrdinalNotation.
+\end{todo}
+
 %%%% ICI %%%%
@@
 
+\subsubsection{Computing successors}
+
+The following function computes the successor of any ordinal less than $\omega+\omega$, by a simple pattern matching.
+
 \begin{Coqsrc}
@@
                                end.
-\end{Coqsrc}
+
+Compute succ omega.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+  = inr 1
+     : nat + nat
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Lemma lt_succ alpha : alpha < succ alpha.
+Proof.
+ destruct alpha; constructor; auto with arith. 
+Qed.
+\end{Coqsrc}
+
+The correctness  of the \texttt{succ} function is asserted by the following lemma.
 
@@
 \end{Coqsrc}
+
+Being a successor is decidable, thanks to the following boolean function.
 
@@
                                  end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
+
 Lemma succb_correct alpha : succb alpha <-> exists beta,  alpha = succ beta.
 \end{Coqsrc}
+
+\subsubsection{Deciding whether a given ordinal is a limit}
+
+In the same way as successors, one can decide by pattern matching wether a given ordinal (dont' forget less than $\omega+\omega$ !) is a limit.
 
@@
                                    end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma omega_is_limit:
-   Omega_limit fin omega.
-Proof.
-  split.
-  - constructor.
-  - inversion_clear  1.
-    + exists (S x); constructor; auto with arith.
-    + lia.
-Qed.
+
+Lemma limitb_correct alpha  : limitb alpha <-> 
+                              exists s: nat -> t, Omega_limit s alpha.
 \end{Coqsrc}

### 2020-07-16 latex/chapter-ordinals.tex
redaction intro ordinals

@@
 
-In this chapter, we introduce on  simple examples the main concepts which are developped later.
-We assume the reader to have a very basic experience with the \coq{} proof assistant, for instance with one of the numerous tutorials one can consult on Internet
-(see for instance \url{https://coq.inria.fr/documentation}). A basic experience  with Peano numbers (library \texttt{Arith}) would be useful.
-
-The definitions  and theorems we are going to show are just simpler versions (\emph{i.e.} restrictions) of the ones contained in the forthcoming chapters.
-We hope that this simplicity will help to understand how we represent, compute and reason about ordinal notations.
+The proof of termination of all hydra battles presented in~\cite{KP82} is based
+on \emph{ordinal numbers}.
+From a mathematical point of view, an ordinal is a representant of an equivalence class for isomorphims of strict, total and well-founded ordered sets.
+
+Chapter~\ref{chap:schutte} presents an adaptation to \coq{} of an axiomatization in classical logic of the set of countable ordinals by K.
+Schütte~\cite{schutte}.
+That formalization is quite complex and technical, so we put it at the end of this document.
+Fortunately, the ordinals we need to study hydra battles are much simpler than Schütte's, and can be represented as rather simple data types in \gallina.
+On the other hand, we hope that chapter~\ref{chap:schutte} will be used
+as a mathematical reference for our constructions: inductive types, functions and predicates.
+
+\section{Ordered types}
+
+The library ~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Classes.RelationClasses.html}{%
+Classes.RelationClasses} contains some definitions and facts about binary relations, among them strict orders.
+
+\begin{Coqsrc}
+Variable A: Type.
+
+  Class StrictOrder (R : relation A) : Prop := {
+    StrictOrder_Irreflexive :> Irreflexive R ;
+    StrictOrder_Transitive :> Transitive R }.
+\end{Coqsrc}
+
+Let \texttt{lt} be a strict order on type $A$.
+
+\begin{Coqsrc}
+Section A_given.
+  Variables (A : Type)
+            (lt: relation A).
+
+Local Infix "<" := lt.
+Local Infix "<=" := (clos_refl _ lt).
+\end{Coqsrc}
+
+We define three notions in 
+\href{../src/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
+
+First, we characterize a least element of $(A,\texttt{lt})$.
+
+\begin{Coqsrc}
+Definition Least `{lt : relation A}
+           {sto : StrictOrder lt} (x : A):=
+  forall y,  clos_refl A lt x y.
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+Prove that there exists at most one least element w.r.t. $(A,<)$.  Give instances where no such element exists.
+\end{exercise}
+
+We say that $y$ is a successor of $x$ if $x<y$ and there is no $z$ such that $x<z<y$.
+
+\begin{Coqsrc}
+Definition Successor {sto : StrictOrder lt} (y x : A):=
+  x < y /\ (forall z,  x < z ->  z <  y -> False).
+\end{Coqsrc}
+
+An element of $A$ which is neither the least element nor a successor is called a \emph{limit}.
+
+\begin{Coqsrc}
+Definition Limit {sto : StrictOrder lt}  (x:A)  :=
+  (exists w:A,  w < x) /\
+  (forall y:A, y < x -> exists z:A, y < z /\  z < x).
+\end{Coqsrc}
 
@@
 
-The proof of termination of all hydra battles presented in~\cite{KP82} is based
-on \emph{ordinal numbers}.
-From a mathematical point of view, an ordinal is a representant of an equivalence class for isomorphims of strict, total and well-founded orders.
+Let us consider how these notions are represented in \coq{}.
 
@@
 leave it to the reader to explore the bibliography.
-Let us cite the book by Schütte~\cite{schutte} which contains an axiomatic definition of the set of countable ordinals we used as a mathematical specification of our implementaion in \coq{}~\cite{CantorContrib}.
+Let us cite the book by Schütte~\cite{schutte} which contains an axiomatic definition of the set of countable ordinals we used as a mathematical specification of our implementation in \coq{}~\cite{CantorContrib}.
 
@@
 
-\subsection{Definitions}
+\subsection{Preliminary Definitions}
+
+The type class \texttt{StrictOrder} is already defined in Standard Library
+(module.
 
@@
 
-Let $x$ and $y$ be two terms of type $A$, we say that \emph{$x$ is an immediate predecessor of $y$}
-(or that \emph{$y$ is a  successor of $x$ }) if $x<y$ and there is no $z$ between $x$ and $y$.
-
-\begin{Coqsrc}
-Definition Ipred {A:Type}{lt : relation A}
-           {sto : StrictOrder lt} (x y : A):=
+Let $x$ and $y$ be two terms of type $A$, we say that \emph{$y$ is a successor of $x$}
+if $x<y$ and there is no $z$ strictly greater than $x$ and  strictly less than $y$.
+
+\begin{Coqsrc}
+Definition Successor `{lt : relation A}
+           {sto : StrictOrder lt} (y x : A):=
   lt x y /\ (forall z,  lt x z -> lt z y -> False).
-
-Definition Successor {A:Type}{lt : relation A}
-           {sto : StrictOrder lt} (y x : A):= Ipred x y.
-\end{Coqsrc}
-
-For instance, 
-we prove that $\omega$ is the limit of the sequence  of finite ordinals.
-
-  \begin{Coqsrc}
-Lemma omega_is_limit: Omega_limit fin omega.
 \end{Coqsrc}

### 2020-07-17 latex/chapter-ordinals.tex
redaction intro ordinals

@@
 
-Chapter~\ref{chap:schutte} presents an adaptation to \coq{} of an axiomatization in classical logic of the set of countable ordinals by K.
-Schütte~\cite{schutte}.
-That formalization is quite complex and technical, so we put it at the end of this document.
-Fortunately, the ordinals we need to study hydra battles are much simpler than Schütte's, and can be represented as rather simple data types in \gallina.
-On the other hand, we hope that chapter~\ref{chap:schutte} will be used
-as a mathematical reference for our constructions: inductive types, functions and predicates.
-
-\section{Ordered types}
+\section{The mathematical point of view}
+
+\subsection{Well-ordered sets}
+Let us start with some definitions.
+A  \emph{well-ordered set} is a set provided with a binary relation $<$ which has the following properties.
+\begin{description}
+\item[irreflexivity] : $\forall x\in A, x\not< x$
+\item[transitivity] : $\forall x\,y\,z\in A, x<y \Rightarrow y<z \Rightarrow x<z$
+\item[trichotomy]: $\forall x\,y\in A, x<y \vee x = y \vee y < x$
+\item[well foundedness] $<$ is well-founded (every element of $A$ is accessible)\footnote{In classical mathematics, we would say that there is no infinite sequence $a_1>a_2> \dots a_n> a_{n+1}\dots$ in $A$.
+Please refer to any documenation on \coq{} for having more details on well-foundedness and accessibility.}.
+
+\end{description}
+
+The best known examples of well-ordered sets are the set $\mathbb{N}$ of natural numbers (with the usual order $<$), as well as any finite segment $[0,i[=\{j\in\mathbb{N}|j<i\}$.
+The disjoint union of two copies of $\mathbb{N}$, \emph{i.e.} the set $\{0,1\}\times\mathbb{N}$ is also well-ordered,
+with respect to the order below:
+
+\begin{align*}
+(i,j) < (i,k) & \;\textbf{if} \; j < k\\
+(0,k) < (1,l) & \;\textbf{for\,any}\;k \;\textbf{and} \; l
+\end{align*}
+
+\subsection{Ordinal numbers}
+
+Let $(A,<_A)$ and $(B,<_B)$ two well-ordered sets. $A$ and $B$ are said to have \emph{the same order type} if 
+there exists a strictly monotonous bijection $b$ from $A$ to $B$, \emph{i.e.} which verifies the proposition
+$\forall x\,y\in A, x <_A y \Rightarrow b(x) <_B  b(y)$.
+
+Having the same order type is an equivalence relation between well-ordered sets. Ordinal numbers (in short \emph{ordinals}) are descriptions (\emph{names}) of the equivalence classes.
+For instance, the order type of $(\mathbb{N},<)$ is associated with the ordinal called  $\omega$, and the order we considered on 
+the disjoint union of $\mathbb{N}$ and itself is named $\omega+\omega$.
+
+If we consider an ordinal $\alpha$ as a well-ordered set, then its elements are just the ordinals strictly less than $\alpha$, \emph{i.e.} the \emph{segment} $\mathbb{O}_\alpha=[0, \alpha[$.
+So, we can speak about \emph{finite}, \emph{infinite}, \emph{countable}, etc., ordinals.
+
+For instance, the elements of $\omega$ are the finite ordinals. $\omega$ is the first infinite ordinal.
+
+We cannot cite all the litterature published on ordinals since Cantor's book 
+\cite{cantorbook}, and 
+leave it to the reader to explore the bibliography.
+
+Chapter~\ref{chap:schutte} presents an adaptation to \coq{} of an axiomatization in classical logic of the set of countable ordinals by K. Schütte~\cite{schutte}. 
+That formalization is quite complex and technical and unshamedly non-constructive,  so we put its description  in the last chapter of this document.
+
+Please note that Schütte considers the (uncountable) set $\mathbb{O}$ of all countable ordinals.
+This set is well ordered (which is one of Schütte's axioms), and any ordinal $\alpha$ is associated with the segment $\mathbb{O}_\alpha$
+of all ordinals strictly less than $\alpha$.
+
+Fortunately, the ordinals we need for  studying hydra battles are much simpler than Schütte's, and can be represented as rather simple data types in \gallina.
+So, we will use \emph{ordinal notations}: Let $\alpha$ be some countable ordinal; In \coq{} terms, an ordinal notation for $\alpha$ will be a data type for representing all ordinals below $\alpha$ (excluding $\alpha$) and a bunch of functions for comparing ordinals, doing some arithmetic, inspecting the topology of the order (limits, successors), and proofs of well-foundedness).
+These functions should be proved correct, with respect to
+Schütte's model (work in progress).
+
+\section{Ordinal notations in Coq}
+
+Our definition of ordinal notation borrows much material from \coq's standard library.
+
+\subsection{Ordered types}
 
@@
 
-Let \texttt{lt} be a strict order on type $A$.
-
-\begin{Coqsrc}
-Section A_given.
-  Variables (A : Type)
-            (lt: relation A).
+It is usual to classify ordinal numbers in three pairwise disjoint categories: 
+\begin{itemize}
+\item The category containing only the least ordinal (called $0$),
+\item The \emph{successors}: $y$ is a successor of $x$ if $x<y$ and there is no $z$ such that $x<z<y$. 
+\item \emph{Limit}: $x\not=0$ is a limit if $x$ is the least upper bound of the set $\{y :\mathbb{O} | y < x\}$.
+\end{itemize}
+
+These notions are defined in the module 
+\href{../src/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}
+ (for any strict order).
+
+\begin{Coqsrc}
+Variable lt: relation A.
 
@@
  Local Infix "<=" := (clos_refl _ lt).
-\end{Coqsrc}
-
-We define three notions in 
-\href{../src/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
-
-First, we characterize a least element of $(A,\texttt{lt})$.
-
-\begin{Coqsrc}
-Definition Least `{lt : relation A}
-           {sto : StrictOrder lt} (x : A):=
-  forall y,  clos_refl A lt x y.
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
-Prove that there exists at most one least element w.r.t. $(A,<)$.  Give instances where no such element exists.
-\end{exercise}
-
-We say that $y$ is a successor of $x$ if $x<y$ and there is no $z$ such that $x<z<y$.
-
-\begin{Coqsrc}
+
+ Definition Least  {sto : StrictOrder lt} (x : A):= forall y,  clos_refl A lt x y.
+
 Definition Successor {sto : StrictOrder lt} (y x : A):=
   x < y /\ (forall z,  x < z ->  z <  y -> False).
-\end{Coqsrc}
-
-An element of $A$ which is neither the least element nor a successor is called a \emph{limit}.
-
-\begin{Coqsrc}
+
 Definition Limit {sto : StrictOrder lt}  (x:A)  :=
@@
 \end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+Prove in \coq{} that the three predicates above are pairwise exclusive.  
+\end{exercise}
 
@@
 
-We cannot cite all the litterature published on ordinals since Cantor's book 
-\cite{cantorbook}, and 
-leave it to the reader to explore the bibliography.
 Let us cite the book by Schütte~\cite{schutte} which contains an axiomatic definition of the set of countable ordinals we used as a mathematical specification of our implementation in \coq{}~\cite{CantorContrib}.

### 2020-07-18 latex/chapter-ordinals.tex
redaction intro ordinals

@@
 
+\index{Maths!Ordinal numbers}
+
 Let $(A,<_A)$ and $(B,<_B)$ two well-ordered sets. $A$ and $B$ are said to have \emph{the same order type} if 
@@
 
-These notions are defined in the module 
-\href{../src/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}
- (for any strict order).
+These notions are defined (for any strict order) in the module 
+\href{../src/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
 
@@
 
-\section{Ordinal numbers}
-\index{Maths!Ordinal numbers}
+\subsection{A class for ordinal notations}
+
+Our definition of ordinal notation is inspired by our use of \coq{}: defining inductive data types, predicates, functions, etc, and proving properties of these functions. 
+The following type class describes a well-ordered type, and specifies functions for comparing two ordinals, and checking wether a given  ordinal is $0$, a limit or a successor.
+
+\vspace{4pt}
+\noindent\emph{From Library~\href{../src/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}}
+
+\begin{Coqsrc}
+Class OrdinalNotation {A:Type}{lt: relation A}(sto:StrictOrder lt)
+      (compare : A -> A -> comparison)  :=
+  {   wf : well_founded lt;
+      compare_correct :
+         forall alpha beta:A,
+          CompareSpec (alpha=beta) (lt alpha beta) (lt beta alpha)
+                                   (compare alpha beta);
+      ZeroLimitSucc_dec : forall alpha,
+        {Least alpha} +
+        {Limit alpha} +
+        {beta: A | Successor alpha beta}
+  }. 
+\end{Coqsrc}
+
+%%%% ICI %%%%%
 
@@
 Without ambiguity we will often denote the segment $\mathbb{O}_\alpha$ by just $\alpha$.
-
-\section{Ordinal notations}
-In a proof assistant like \coq{},  it is useful to represent ordinals through some data-type, and make arithmetical operations and comparison effectively implemented  through certified functions.
-
-Of course, since the set of all countable ordinals is not countable, it cannot be represented by a language of finite terms. So, we will study ordinal notations just  for countable segments.
-
-\subsection{A type class for ordinal notations}
-
-Let $<$ be some relation on a given type $A$. In order to build an ordinal  notation 
-system out of $(A,<)$, we require $<$ to be a strict, well-founded order, and that 
-the reflexive closure of $<$ is a total order on $A$.
-
-\index{Coq!Techniques!Type classes}
-
-Using \coq's standard library and type classes, we propose the following definition.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../src/html/hydras.Prelude.Ordinal_generic.html}{Prelude.Ordinal\_generic}}
-
-\begin{Coqsrc}
-Class OrdinalNotation {A:Type}{lt: relation A}(sto:StrictOrder lt)
-      (compare : A -> A -> comparison) :=
-  { compare_correct :
-      forall alpha beta:A,
-        CompareSpec (alpha=beta) (lt alpha beta) (lt beta alpha)
-                                 (compare alpha beta);
-    wf : well_founded lt}.  
-\end{Coqsrc}
 

### 2020-07-18 latex/chapter-ordinals.tex
simplification of class OrdinalNotation

@@
 Our definition of ordinal notation is inspired by our use of \coq{}: defining inductive data types, predicates, functions, etc, and proving properties of these functions.
-The following type class describes a well-ordered type, and specifies functions for comparing two ordinals, and checking wether a given  ordinal is $0$, a limit or a successor.
 
@@
                                    (compare alpha beta);
-      ZeroLimitSucc_dec : forall alpha,
-        {Least alpha} +
-        {Limit alpha} +
-        {beta: A | Successor alpha beta}
   }. 
@@
 %%%% ICI %%%%%
+
+\begin{todo}
+Present the method ZeroLimitSucc 
+\end{todo}
 

### 2020-07-19 latex/chapter-ordinals.tex
simplification of class OrdinalNotation

@@
 
-It is usual to classify ordinal numbers in three pairwise disjoint categories: 
-\begin{itemize}
-\item The category containing only the least ordinal (called $0$),
-\item The \emph{successors}: $y$ is a successor of $x$ if $x<y$ and there is no $z$ such that $x<z<y$. 
-\item \emph{Limit}: $x\not=0$ is a limit if $x$ is the least upper bound of the set $\{y :\mathbb{O} | y < x\}$.
-\end{itemize}
-
-These notions are defined (for any strict order) in the module 
-\href{../src/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
-
-\begin{Coqsrc}
-Variable lt: relation A.
-
- Local Infix "<" := lt.
- Local Infix "<=" := (clos_refl _ lt).
-
- Definition Least  {sto : StrictOrder lt} (x : A):= forall y,  clos_refl A lt x y.
-
-Definition Successor {sto : StrictOrder lt} (y x : A):=
-  x < y /\ (forall z,  x < z ->  z <  y -> False).
-
-Definition Limit {sto : StrictOrder lt}  (x:A)  :=
-  (exists w:A,  w < x) /\
-  (forall y:A, y < x -> exists z:A, y < z /\  z < x).
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
-Prove in \coq{} that the three predicates above are pairwise exclusive.  
-\end{exercise}
-
 \subsection{A class for ordinal notations}
@@
 
-%%%% ICI %%%%%
+% ICI
+
+\begin{todo}
+Present instances of Ordinal notation.  
+\end{todo}
+
+It is usual to classify ordinal numbers in three pairwise disjoint categories: 
+\begin{itemize}
+\item The category containing only the least ordinal (called $0$),
+\item The \emph{successors}: $y$ is a successor of $x$ if $x<y$ and there is no $z$ such that $x<z<y$. 
+\item \emph{Limit}: $x\not=0$ is a limit if $x$ is the least upper bound of the set $\{y :\mathbb{O} | y < x\}$.
+\end{itemize}
+
+These notions are defined (for any strict order) in the module 
+\href{../src/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
+
+\begin{Coqsrc}
+Variable lt: relation A.
+
+ Local Infix "<" := lt.
+ Local Infix "<=" := (clos_refl _ lt).
+
+ Definition Least  {sto : StrictOrder lt} (x : A):= forall y,  clos_refl A lt x y.
+
+Definition Successor {sto : StrictOrder lt} (y x : A):=
+  x < y /\ (forall z,  x < z ->  z <  y -> False).
+
+Definition Limit {sto : StrictOrder lt}  (x:A)  :=
+  (exists w:A,  w < x) /\
+  (forall y:A, y < x -> exists z:A, y < z /\  z < x).
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+Prove in \coq{} that the three predicates above are pairwise exclusive.  
+\end{exercise}
 
@@
 \end{todo}
-
-Let us consider how these notions are represented in \coq{}.
-
-We can also associate to every ordinal $\alpha$ a set whose elements are all ordinals strictly less than $\alpha$.
-Thus, it is meaningful  to consider \emph{finite}, \emph{infinite}, \emph{demunerable} and \emph{non-countable} ordinals.
-The relation $<$ on ordinals is well-founded, and the order $\leq$ associated with
-$<$ is total.
-
-Let us cite the book by Schütte~\cite{schutte} which contains an axiomatic definition of the set of countable ordinals we used as a mathematical specification of our implementation in \coq{}~\cite{CantorContrib}.
-
-Out of respect of the tradition, the meta-variables for ordinals will be 
- $\alpha$, $\beta$, $\gamma$, etc.
 

### 2020-07-21 latex/chapter-ordinals.tex
various renamings: not yet corrected in latex

@@
   }.
+
+Definition on_t  {A:Type}{lt: relation A}
+            {compare : A -> A -> comparison}
+            {on : OrdinalNotation lt compare} := A.
+
+Definition on_compare {A:Type}{lt: relation A}
+            {compare : A -> A -> comparison}
+            {on : OrdinalNotation lt compare} := compare.
+
+Definition on_lt {A:Type}{lt: relation A}
+           {compare : A -> A -> comparison}
+           {on : OrdinalNotation lt compare} := lt.
+
+Definition on_le  {A:Type}{lt: relation A}
+           {compare : A -> A -> comparison}
+           {on : OrdinalNotation lt compare} :=
+  clos_refl _ on_lt.
+
 \end{Coqsrc}

### 2020-07-21 latex/chapter-hydras.tex
corrections in the pdf documentation

@@
 with the heads up and the foot (i.e., the root of the tree) down.
-Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the module 
-\url{../src/html/hydras.Hydra.Hydra_Examples.html}).
+Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the 
+module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}).
 
 \begin{figure}[h]
+\centering
 \begin{tikzpicture}[very thick, scale=0.6]
@@
 
-\begin{center}
+\begin{figure}[h]
+  \centering
   \begin{tabular}{ll}
@@
 \end{tabular}
-\captionof{table}{\label{tab:hyd2tree}
-Translation from hydras to trees}
-\end{center}
+  \caption{Translation from hydras to trees}
+  \label{tab:hyd2tree}
+\end{figure}
 
@@
 \item If the hydra is composed of just one head, the battle is finished
-and  Hercules is the winner
-\item Otherwise, Hercules chops off \emph{one} head of the hydra
+and  Hercules is the winner.
+\item Otherwise, Hercules chops off \emph{one} head of the hydra,
 
 \begin{itemize}
-\item If the head is at distance 1 from the foot, the head is just lost by the hydra
+\item If the head is at distance 1 from the foot, the head is just lost by the hydra, with no more reaction.
 \item Otherwise, let us denote by \(r\) the node that was at distance \(2\) from 
@@
 Moreover, we note that this description is in \emph{imperative} terms.
-In order to build a formal  study of the properties of hydra battles, we prefer to use mathematical language, i.e., graphs, relations, functions, etc.
+In order to build a formal  study of the properties of hydra battles, we prefer to use a mathematical vocabulary, i.e., graphs, relations, functions, etc.
 Thus, the replication process will be represented as a binary relation on a data type \texttt{Hydra},
 linking the state of the hydra \emph{before} and \emph{after} the transformation.
-A battle will thus be represented as a sequence of terms of type \texttt{Hydra}.
+A battle will thus be represented as a sequence of terms of type \texttt{Hydra}, respecting the rules of the game.
 
@@
 Fig.~\vref{fig:Hy2} represents the rotten neck in dashed lines, and the part that will be replicated in red.
-Assume also that the hydra decides to add 4 copies of the red part.
+Assume also that the hydra decides to add 4 copies of the red part\footnote{In other words, the replication factor at this round is equal to $4$.}.
 We obtain a new state \texttt{Hy''} depicted in Fig.~\ref{fig:Hy3}.
@@
 \begin{figure}[h]
+\centering
 \begin{tikzpicture}[very thick, scale=0.6]
@@
 \begin{figure}[hp]
-
+\centering
 \begin{tikzpicture}[very thick, scale=0.5]
@@
 \end{tikzpicture}
-\caption{A beheading}
+\caption{A second beheading}
 \label{fig:Hy2}
@@
 \begin{figure}[hp]
+\centering
 \begin{tikzpicture}[very thick, scale=0.6]
@@
 \begin{figure}[hp]
+\centering
 \begin{tikzpicture}[very thick, scale=0.4]
@@
 
-Various \emph{Coq} libraries  represent trees the nodes of which 
-have a non-fixed number of sons. The reader can, for instance, look at
-Chapter 14, pages 400-406 of \cite{BC04}. 
-Our representation uses two mutual inductive types: \texttt{Hydra} to represent
- any hydra, and \texttt{Hydrae} to represent finite sequences of hydras.
-
+In order to describe trees where each node can have an arbitrary (but finite) number of sons, it usual to define a type where each node carries a \emph{forest}, \emph{i.e} a list of trees
+(see for instance Chapter 14, pages 400-406 of \cite{BC04}).
+
+For this purpose, we define two mutual inductive types.
 \index{Types!Hydra}

### 2020-07-21 latex/chapter-introduction.tex
corrections in the pdf documentation

--- 2020-06-23 latex/introduction.tex
@@
 This is obviously the case of the four-color theorem~\cite{fourcolors}  and the Kepler conjecture~\cite{flyspeck2015}.
-We do not mention impressive works like the proof of the odd-order theorem ~\cite{oddorderthm}, since understanding its statement requires some mathematical culture.
+We do not mention impressive works like the proof of the odd-order theorem ~\cite{oddorderthm}, since understanding its statement requires a quite good mathematical culture.
 
-Hydra games appeared in an article published in 1982 by two mathematicians:
+Hydra games (a.k.a. \emph{Hydra battles}) appeared in an article published in 1982 by two mathematicians:
 L. Kirby and J. Paris~\cite{KP82}: \emph{Accessible Independence Results for Peano Arithmetic}. 
 Although the mathematical contents of this 
-paper are quite advanced, hydra games (a.k.a.
-\emph{hydra battles}) are very easy to understand.
+paper are quite advanced, the rules of hydra battles are very easy to understand.
 There are now several sites on Internet where you can find tutorials on hydra games, together with simulators you can play with.
@@
   \includegraphics[width=11cm]{epsilon0.jpg}
-  \caption{Mathematics and formal proofs}
+  \caption{Discrete mathematics and formal proofs}
   \label{fig:gamma0}
 \end{figure}
- We also hope to provide highlights on infinity (both potential and actual) through the notions of function, computation, limit,
+ We also hope to provide a little clarification on infinity (both potential and actual) through the notions of function, computation, limit,
  types and proofs.
 
-\subsection{Remarks}
+\subsection*{Remarks}
 In~\cite{KP82}, Kirby and Paris showed  that there is no proof of termination of all hydra battles in Peano Arithmetic (PA).
@@
 Thus, there is no ambiguity in our definitions and the premises of the theorems.
-Furthermore, you will be able to navigate through the development, using your favourite editor or IDE, and some commands like \texttt{Search}, \texttt{Locate}, \texttt{Print Assumptions}, etc.
+Furthermore, you will be able to navigate through the development, using your favourite text editor or IDE, and some commands like \texttt{Search}, \texttt{Locate}, \texttt{Print Assumptions}, etc.
 
@@
 
-\paragraph{Main references}
+\paragraph*{Main references}
 
@@
 \item ``The Termite and the Tower'', by Will Sladek~\cite{Sladek07thetermite}
-
+\item Chapter V of ``Proof Theory'' by Kurt Schütte~\cite{Schutte}
 \end{itemize}
@@
 For instance, in Chapter~\vref{chap:ketonen}, the notion of \emph{limit ordinal} is
-made more ``concrete'' thanks to a function \texttt{canonseq} that computes every item of a sequence which converges on a given limit ordinal $\alpha$.
+made more ``concrete'' thanks to a function \texttt{canon} which computes every item of a sequence which converges on a given limit ordinal $\alpha$.
 This simply typed function allows the user/reader to make her/his own experimentations.
@@
 
+\subsubsection*{State of the development}
+The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on ordinal notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
+We added new material : axiomatic definition of countable ordinals after Schütte~\cite{Schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, equations, etc.
+
+We are now working in order to make clumsy proofs more readable, simplify definitions, 
+to ``factorize'' proofs as much as possible.
+
+Many ``todo''s or projects are put in this text, suggesting possible improvements.
+
 \subsubsection*{Contributions are welcome}
@@
 
-\subsubsection{Acknowledgements}
+\subsubsection*{Acknowledgements}
 \label{sec:orgheadline5}
@@
 
-\subsubsection{Typographical conventions}
+\subsection*{Typographical conventions}
 
@@
 
-\subsubsection{Alternative or bad definitions}
+\subsection*{Alternative or bad definitions}
 

### 2020-07-22 latex/chapter-hydras.tex
renaming fight --> battle

@@
 
-For this purpose, we define two mutual inductive types.
+For this purpose, we define two mutual \emph{ad-hoc}  inductive types, where \texttt{Hydra} is the main type, and \texttt{Hydrae} is a helper for describing finite sequences of hydra.
 \index{Types!Hydra}
@@
 \end{Coqsrc}
+
+\index{To do}
+
+\begin{todo}
+Look for an existing library on trees with nodes of arbitrary arity, iin order to replace  this ad-hoc type with something more generic.
+\end{todo}
 
@@
 \end{enumerate}
-The relation associated with each round of the battle is \emph{parameterized}  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot).
+The relation associated with each round of the battle is \emph{parameterized}  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot,
+but present for consistency's sake).
 
@@
 
-Since the replication of parts of the hydra depends on the distance of the chopped head from  the foot, we  decompose our description into several cases, under the form of a bunch of [mutually] inductive predicates over the types \texttt{Hydra} and \texttt{Hydrae}.
-
-We decompose the relation associated with each  round in two cases: 
+Since the replication of parts of the hydra depends on the distance of the chopped head from  the foot, we  decompose our description into two main  cases, under the form of a bunch of [mutually] inductive predicates over the types \texttt{Hydra} and \texttt{Hydrae}.
+
+The mutually exclusive cases we consider are the following.
 \begin{itemize}
@@
 
-\emph{Please consider exercise~\vref{exo:interactive-battle}.}
+The reader is encouraged to look at all the intermediary subgoals of this example.
+\emph{Please consider also exercise~\vref{exo:interactive-battle}.}
 
@@
 
-We chosed to call \emph{standard} the kind of battles which appear  most often in the litterature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$.
+We chosed to call \emph{standard} the kind of battles which appear  most often in the litterature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$
 
@@
 
-\subsection{Fights}
+\subsection{Big steps}
 
@@
 \begin{Coqsrc}
-Inductive fight (B:Battle) : nat -> Hydra -> nat -> Hydra -> Prop :=
-|  fight_1 : forall i h  h', battle_r   B i  h h' -> 
-                            fight B i h (S i) h'
-| fight_n : forall i h  j h' h'',  battle_r  B i h h''  ->
-                                   fight B (S i) h'' j h'  ->
-                                   fight B i h j h'.
+Inductive battle (B:Battle) : nat -> Hydra -> nat -> Hydra -> Prop :=
+|  battle_1 : forall i h  h', battle_r   B i  h h' -> 
+                            battle B i h (S i) h'
+| battle_n : forall i h  j h' h'',  battle_r  B i h h''  ->
+                                   battle B (S i) h'' j h'  ->
+                                   battle B i h j h'.
 \end{Coqsrc}
@@
 %  \begin{Coqsrc}
-%  Lemma fight_free_equiv1 : forall i j h h',  
-%              fight free i h j h' ->   h -+-> h'.
-
-%  Lemma fight_free_equiv2 : forall h h',
+%  Lemma battle_free_equiv1 : forall i j h h',  
+%              battle free i h j h' ->   h -+-> h'.
+
+%  Lemma battle_free_equiv2 : forall h h',
 %      h -+-> h' ->
-%     forall i, exists j,  fight free i h j h'.
+%     forall i, exists j,  battle free i h j h'.
 %  \end{Coqsrc}
@@
 \begin{Coqsrc}
-Lemma L_0_2 : fight standard 0 hinit 2 (hyd1 h3).   
+Lemma L_0_2 : battle standard 0 hinit 2 (hyd1 h3).   
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Lemma L_2_3 : fight standard 2 (hyd1 h3)  3 (hyd 3 0 0).
-
-Lemma L_0_3 : fight standard 0 hinit 3 (hyd 3 0 0).
+Lemma L_2_3 : battle standard 2 (hyd1 h3)  3 (hyd 3 0 0).
+
+Lemma L_0_3 : battle standard 0 hinit 3 (hyd 3 0 0).
 \end{Coqsrc}
@@
 Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle will only start with the lemma
-\texttt{lemma:step-fight}, page~\pageref{lemma:step-fight}.
+\texttt{lemma:step-battle}, page~\pageref{lemma:step-battle}.
 
@@
 
-\label{lemma:step-fight}
-
-\begin{Coqsrc}
-Lemma step_fight : forall i a b c a' b' c', 
+\label{lemma:step-battle}
+
+\begin{Coqsrc}
+Lemma step_battle : forall i a b c a' b' c', 
    one_step i a b c a' b' c' ->
-   fight standard i (hyd  a b c)  (S i) (hyd a' b' c').
+   battle standard i (hyd  a b c)  (S i) (hyd a' b' c').
 \end{Coqsrc}
@@
 
-The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{fight}.
-
-\begin{Coqsrc}
- Lemma steps_fight : forall i a b c j a' b' c', 
+The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{battle}.
+
+\begin{Coqsrc}
+ Lemma steps_battle : forall i a b c j a' b' c', 
    steps i a b c j a' b' c' ->
-   fight standard i (hyd  a b c)   j  (hyd a' b' c').
-\end{Coqsrc}
-
-Thus, any result about \texttt{steps} will be applicable to standard fights.
+   battle standard i (hyd  a b c)   j  (hyd a' b' c').
+\end{Coqsrc}
+
+Thus, any result about \texttt{steps} will be applicable to standard battles.
 Using the predicate \texttt{steps} our study of the length of the considered battle
@@
 From now on, the lemma \texttt{reachable\_S} allows us to watch larger steps of 
-the fight.
+the battle.
 
@@
 
-We are now able to prove formally that the considered fight is 
+We are now able to prove formally that the considered battle is 
 composed of $N$ steps.
@@
 Lemma Almost_done :
-  fight standard 3 (hyd 3 0 0) N (hyd 0 0 0).
-Proof. 
-  apply steps_fight, SuperbigStep.
+  battle standard 3 (hyd 3 0 0) N (hyd 0 0 0).
+Proof. 
+  apply steps_battle, SuperbigStep.
 Qed.
@@
 Theorem  Done :
-  fight standard 0 hinit N head.
-Proof.
-  eapply fight_trans.
+  battle standard 0 hinit N head.
+Proof.
+  eapply battle_trans.
   -   apply Almost_done.

### 2020-07-22 latex/chapter-ks.tex
renaming fight --> battle

@@
 the replication factor $i$ (Lemma \href{../src/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
-\item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) fight from $\iota(\alpha)$ to $\iota(\beta)$.
+\item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) battle from $\iota(\alpha)$ to $\iota(\beta)$.
 \end{itemize}
@@
 The next step of our development is to extend this relationship to
-the order $<$ on $[0,\epsilon_0[$ on one side, and hydra fights on the other side.
-
-\begin{Coqsrc}
-Lemma path_to_fight alpha s beta :
+the order $<$ on $[0,\epsilon_0[$ on one side, and hydra battles on the other side.
+
+\begin{Coqsrc}
+Lemma path_to_battle alpha s beta :
   path_to  beta  s alpha -> nf alpha ->
@@
 
-As a corollary, we are now able to transform any inequality $\beta<\alpha<\epsilon_0$ into a (free) fight.
-
-\begin{Coqsrc}
-Lemma LT_to_fight alpha beta :
+As a corollary, we are now able to transform any inequality $\beta<\alpha<\epsilon_0$ into a (free) battle.
+
+\begin{Coqsrc}
+Lemma LT_to_battle alpha beta :
     beta < alpha ->  iota alpha -+-> iota beta.
@@
 \item From the hypothesis \texttt{m\_bounded}, we have \texttt{m big\_h < mu}
-\item By Lemma \texttt{LT\_to\_fight}, we get a (free) battle from
+\item By Lemma \texttt{LT\_to\_battle}, we get a (free) battle from
 \texttt{big\_h = iota mu} to \texttt{small\_h = iota (m big\_h)}.
@@
       beta < alpha ->
-      exists n i,  fight standard  n (iota alpha) i (iota beta).
+      exists n i,  battle standard  n (iota alpha) i (iota beta).
 \end{Coqsrc}

### 2020-07-22 latex/chapter-large-sets.tex
renaming fight --> battle

@@
 To be precise, for any ordinal $\alpha<\epsilon0$ and any positive integer $k$,
-we give a minoration of the number of steps of a standard fight which
+we give a minoration of the number of steps of a standard battle which
 starts with the hydra $\iota(\alpha)$ and the replication factor $k$.
@@
   forall k, (1 <= k)%nat ->
-            fight_length standard k (iota (cnf alpha))
+            battle_length standard k (iota (cnf alpha))
                          (L_ alpha (S k) - k).
@@
        H_ alpha k - k <= l /\
-       fight_length standard k (iota (cnf alpha)) l.    
+       battle_length standard k (iota (cnf alpha)) l.    
 \end{Coqsrc}

### 2020-07-22 latex/chapter-ordinals.tex
renaming fight --> battle

@@
 For proving the inequality  \texttt{m\_lt: m small\_h < m big\_h}, it suffices to
-build a fight transforming \texttt{big\_h} into \texttt{small\_h}.
+build a battle transforming \texttt{big\_h} into \texttt{small\_h}.
 
@@
 The following lemma establishes a correspondance between the relation
-\texttt{step} and hydra fights.
-
-\begin{Coqsrc}
-Lemma step_to_fight : forall p q, step p q -> iota p -+-> iota q.
+\texttt{step} and hydra battles.
+
+\begin{Coqsrc}
+Lemma step_to_battle : forall p q, step p q -> iota p -+-> iota q.
 \end{Coqsrc}
@@
 Thus, starting from any inequality $q < p$ on type \texttt{t}, we can build 
-by transfinite induction over \texttt{p} a fight 
+by transfinite induction over \texttt{p} a battle 
 that transforms the hydra $\iota(p)$ into $\iota(q)$.

### 2020-07-23 latex/chapter-hydras.tex
redaction introduction to ordinals

@@
 
-\subsection{Big steps}
+\subsection{Formal definition of battles}
 
@@
 
-In this section we show how long a simple hydra battle can last.  The following example
-considers a very small hydra, shown on figure~\vref{fig:hinit}.
+In this section we consider a simple example of battle, starting with a small hydra,
+shown on figure~\vref{fig:hinit}, with a simple strategy for both players:
+
+\begin{itemize}
+\item Hercules chops always the rightmost head of the hydra.
+\item The battle is standard: at the round number $i$, the expected replication is $i$.
+\end{itemize}
 
@@
 
-Moreover, in order to \emph{compute} the length of a battle, we set Hercules strategy : to chop the rightmost head at each round, and the hydra's strategy : at the $i$-th round, the replication factor is exactly $i$ (which means that the battle is
-a \emph{standard} battle, see section~\vref{sect:battle-classes}).
-
-The  battle is so long that no \emph{test} can give us an estimation of its length, and we do need the expressive power of logic to compute this length.
+The lemma we would like to prove is ``The considered battle lasts exactly $N$ rounds'',
+with $N$ being a natural number we gave to guess.
+
+But the  battle is so long that no \emph{test} can give us an estimation of its length, and we do need the expressive power of logic to compute this length.
 However, in order to  guess this length, we made some experiments, computing with \gallina{}, \coq{}'s  functional programming language.
@@
 
-We want to verify whether Hercules wins a standard battle starting from \texttt{hinit} at time $t=0$ , and in this case, in how many rounds. We assume Hercules' 
-strategy is to chop off the rightmost head at every round.
+In the following lines, we try to show faithfully how we found the value of the number $N$.
 
@@
 
+\subsection{The beginning of hostilities}
 During the two first rounds, our hydra loses its two rightmost heads. Thus just before the third round, it looks like in figure~\vref{fig:hinit-plus2}.
@@
 
-The following lemma  is a formal statement about the two first rounds of the battle.
+The following lemma  is a formal statement about these first rounds, in terms of the
+\texttt{battle} predicate.
 
@@
 \end{Coqsrc}
+
+\subsection{Looking for regularities}
 
@@
 
-We would like to have a formal proof that the length of the considered battle is 
-equal to some integer $N$, but we have first  to \emph{guess} this number.
+\subsection{Computing \dots}
 In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple datatype for representing the states as tuples composed of
@@
 
+\subsection{Proving \dots}
 We are now able to reason about the sequence of transitions defined by our hydra battle. Instead of using the data-type \texttt{state} we study the relationship
@@
 
-We are now able to make giant steps in the simulation of the battle.
+\subsection{Giant steps}
+
+We are now able to make bigger steps in the simulation of the battle.
 First, we iterate the lemma \texttt{reachable\_S}.
@@
 
+\subsection{A minoration lemma}
+
 Now, we would like to get an intuition of  how big the number $N$ is.
@@
 
-Let $b$ be an instance of class \texttt{Battle}. A \emph{variant} for $b$ consists
+Let $B$ be an instance of class \texttt{Battle}. A \emph{variant} for $B$ consists
 in a well-founded relation $<$  on some type \texttt{A}, and a function
-(also called a \emph{measure}) \texttt{m:Hydra->A} such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $b$, the inequality $m(h')<m(h)$ holds.
+(also called a \emph{measure}) \texttt{m:Hydra->A} such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $B$, the inequality $m(h')<m(h)$ holds.
 
@@
 \begin{Coqsrc}
-Class Hvariant {A:Type}{Lt:relation A}(Wf: well_founded Lt)(b : Battle)
+Class Hvariant {A:Type}{Lt:relation A}(Wf: well_founded Lt)(B : Battle)
   (m: Hydra -> A):   Prop :=
   {variant_decr :forall i h h',
-     h <> head -> b i  h h' -> Lt (m h') (m h)}.
+      h <> head ->
+      battle_r  B i  h h' -> Lt (m h') (m h)}.
 \end{Coqsrc}
@@
 \begin{exercise}
- Prove that, if there is an instance of \texttt{Hvariant Lt wf\_Lt b m}, then there exists no infinite battle in  \texttt{b}.
+ Prove that, if there is an instance of \texttt{Hvariant Lt wf\_Lt $B$ $m$}, then there exists no infinite battle in  $B$.
 \end{exercise}
@@
 
-The class \texttt{Hvariant} is parameterized by a well-founded order $<$ on some type $A$, a class of battles $B$ and a measure $m$. When $<$ or $m$ are 
-badly given for a given class $B$, one may fail to find a variant.
-
-We show that this is the case of the type \texttt{nat} of Peano natural numbers, provided with the usual strict order $<$.
-
-\subsubsection{Using the Hydra's height as a variant}
-Trivially, a measure like the height will not work for building a variant.
-
-Please consider the hydras of Fig.~\vref{fig:h-omega2-small} and
-\ref{fig:h-omega-omega}. The former can be transformed into the latter in one round, but their heights are equal.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) -- (N2) ;
-\end{tikzpicture}
-\caption{\label{fig:h-omega-omega}}
-\end{figure}
-
-\vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Omega_Small.html\#height_bad}{Hydra.Hydra.Omega\_Small}}
-
-\begin{Coqsrc}
-Lemma height_bad :  ~ Hvariant lt_wf free height.
-Proof.
-  intros [H];
-   specialize (H 1 (hyd1 (hyd2 head head))  (hyd1 (hyd1 head)));
-   apply (lt_irrefl 2), H.
-  -  apply reachable_free.
-  -  discriminate.
-  -  exists 0; right;  R2_here 0;  left. 
-Qed.
-\end{Coqsrc}
-
-\subsubsection{Using any variant defined on \texorpdfstring{nat}{nat}}
-\label{sect:omega-case}
-We could imagine that our previous failure is due to the choice of \texttt{height} as a measure, and that a more complex function would work.
-In fact, we can prove that 
-\emph{no} instance of class \texttt{WfVariant round Peano.lt $m$} can be build, where
+The best known well-founded order is the natural order on the set $\mathbb{N}$ of natural numbers (the type \texttt{nat} of Standard library).
+It would be interesting to look for some measure $m:\texttt{nat}\arrow\texttt{nat}$ and prove it is a variant.
+
+Unfortunately, we can prove that 
+\emph{no} instance of class \texttt{WfVariant round Peano.lt $m$} can be built, where
 $m$ is \emph{any} function of type \texttt{Hydra $\arrow$ nat}.
@@
 In order to build a variant for proving the termination of all hydra battles, we need to consider order structures more complex than the usual order on type \texttt{nat}. 
-The next chapter will present the notion of \emph{ordinal notation} from the point view of the \coq{} user: inductive types, functions, type classes, etc.
-The ordinal $\epsilon_0$, the first one allowing us to build our desired variant, will be studied in detail in 
-Chapter~\ref{chap:T1}.
+The notion of \emph{ordinal number} provides a catalogue of well-founded order types.
+For a reasonably large bunch of ordinal numbers, \emph{ordinal notations} are data-types which allow the \coq{} user to define functions, to compute and prove some properties, for instance by reflection.
+
+So, the next chapters will present ordinal numbers and ordinal notations, and we will
+look for a variant complex enough for proving the termination of all battles, but not too complex if possible.

### 2020-07-23 latex/chapter-ordinals.tex
redaction introduction to ordinals

--- 2020-07-21 latex/chapter-ordinals.tex
@@
-\chapter{Introduction to ordinal numbers with the Coq proof assistant}
-
-\centerline{\LARGE {\bf work in progress!}}
+\chapter{Introduction to ordinal numbers}
 
@@
 From a mathematical point of view, an ordinal is a representant of an equivalence class for isomorphims of strict, total and well-founded ordered sets.
+
+More intuitively, let us have a look at Figure~\ref{fig:ordinal-sequence}. It presents a sequence of ordinal numbers, as an extension of the sequence of natural numbers.
+
+\begin{figure}[h]
+  \centering
+\fbox{\Large
+  \begin{minipage}{1.0\linewidth}
+  \begin{align*}
+     &\textcolor{blue}{0},\,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,\ldots\\
+&\textcolor{red}{\omega},\,\omega+1,\omega+2,\omega+3,\ldots\\
+&\textcolor{red}{\omega\times 2},\,\omega\times 2+1,\ldots, \textcolor{red}{\omega\times 3},\,\omega\times 3+1,\ldots, \textcolor{red}{\omega\times 4},\ldots,\\
+&\textcolor{red}{\omega^2},\ldots, \textcolor{red}{\omega^2\times 42},\ldots,\textcolor{red}{\omega^3},\ldots, \textcolor{red}{\omega^4},\omega^4+1,\ldots,\\
+&\textcolor{red}{\omega^\omega2},\ldots, \textcolor{red}{\omega^\omega\times 2},\ldots,\\
+&\textcolor{red}{\omega^{\omega^\omega}},\ldots, \textcolor{red}{\omega^{\omega^\omega}+\omega^\omega\times 42+ \omega^{55}+\omega}, \ldots, \textcolor{red}{\omega^{\omega^{\omega+1}}}, \omega^{\omega^{\omega+1}}+1,\dots\\
+& \textcolor{red}{\epsilon_0 (= \omega^\epsilon_0)}, \epsilon_0+1, \ldots,\\
+& \textcolor{red}{\epsilon_1}, \ldots, \textcolor{red}{\epsilon_2}, \ldots, \textcolor{red}{\epsilon_\omega},\ldots \\
+& \textcolor{red}{\Gamma_0}, \Gamma_0+1, \Gamma_0+2, \Gamma_0+3,\ldots, \textcolor{red}{\Gamma_0+\omega}, \ldots,\\
+&\ldots
+  \end{align*}   
+  \end{minipage}}
+
+  \caption{A short overview of the sequence of ordinal numbers}
+  \label{fig:ordinal-sequence}
+\end{figure}
+
+Let us comment some features of this table.
+
+\begin{itemize}
+\item The ordinals are listed in a strictly increasing order. 
+\item Dots : ``$\ldots$'' stand for an infinite sequence of ordinals, not shown for lack of space. For instance, the ordinal $42$ is not shown in the first line, but it exists, betaween $41$ and $43$.
+\item Each ordinal in black is an immediate successor of anotter ordinal.
+We call it a \emph{successor} ordinal.
+For instance, $43$ is the successor of $42$, and $\omega^4+1$ the successor of $\omega^4$.
+\item Ordinals that follow dots (written in red) are called \emph{limit ordinals}.
+With respect to the order induced by this sequence, any limit ordinal $\alpha$ is the least upper bound of  the set $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
+For instance $\omega$ is the least upper bound of the set of all finite ordinals (in the first line). $\omega$ is also the first limit ordinal, and the first infinite ordinal, in the sense that 
+the set $\mathbb{O}_\omega$ is infinite.
+\item The ordinal $\epsilon_0$ is the first number which is equal to its own exponential of base $\omega$.
+It plays an important role in proof theory, and is particularly studied in chapters~\ref{chap:T1} to \ref{chap:alpha-large}.
+\end{itemize}
 

### 2020-07-23 latex/chapter-ks.tex
redaction introduction to ordinals

--- 2020-07-09 latex/chapter-ks.tex
@@
 \begin{itemize}
-\item There is no variant mapping the type \texttt{Hydra} into the interval $[0,\omega^2[$ (section ~\vref{omega2-case}), and a fortiori 
-$[0,\omega[$
+\item There is no variant mapping the type \texttt{Hydra} into the interval $[0,\omega^2)$ (section ~\vref{omega2-case}), and a fortiori 
+$[0,\omega)$
 
 \item There is a variant that maps the type \texttt{Hydra} into the
-interval $[0,\epsilon_0[$ (theorem \texttt{every\_battle\_terminates}, in section~\vref{hydra-variant}).
+interval $[0,\epsilon_0)$ (theorem \texttt{every\_battle\_terminates}, in section~\vref{hydra-variant}).
 \end{itemize}
@@
   `` Is there  any variant from
-\texttt{Hydra} into some interval $[0,\mu[$, where $\mu<\epsilon_0$, for proving the termination of all hydra battles ?''
+\texttt{Hydra} into some interval $[0,\mu)$, where $\mu<\epsilon_0$, for proving the termination of all hydra battles ?''
 \end{quote}
@@
 There is no variant for proving the termination of all hydra battles
-from \texttt{Hydra} into the interval $[0..\mu[$, where
+from \texttt{Hydra} into the interval $[0..\mu)$, where
 $\mu< \epsilon_0$.
@@
 In order to apply our knowledge about  ordinal numbers (less than $\epsilon_0$) to the study of hydra battles, we define an injection
-from the interval $[0,\epsilon_0[$ into the type \texttt{Hydra}.
+from the interval $[0,\epsilon_0)$ into the type \texttt{Hydra}.
 
@@
 The next step of our development is to extend this relationship to
-the order $<$ on $[0,\epsilon_0[$ on one side, and hydra fights on the other side.
+the order $<$ on $[0,\epsilon_0)$ on one side, and hydra fights on the other side.
 

### 2020-07-23 latex/chapter-large-sets.tex
redaction introduction to ordinals

--- 2020-07-09 latex/chapter-large-sets.tex
@@
 Let us consider two integers $k$ and $l$, such that $0<k<l$. In order to check whether the interval $[k,l]$ is minimally large for $\alpha$, it is enough to
-follow from $\alpha$ the path associated with the interval $[k,l[$ and verify that the last ordinal we obtain is equal to $1$.
+follow from $\alpha$ the path associated with the interval $[k,l($ and verify that the last ordinal we obtain is equal to $1$.
 
@@
 Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider the functions which associate to each stritly positive integer $k$ the number $l$, where 
-the interval $[k,l-1]$\;\footnote{also denoted by $[k,l[$}  is $\alpha$-mlarge.
+the interval $[k,l-1]$\;\footnote{also denoted by $[k,l)$}  is $\alpha$-mlarge.
 

### 2020-07-23 latex/chapter-ordinals.tex
redaction introduction to ordinals

@@
 &\textcolor{red}{\omega^2},\ldots, \textcolor{red}{\omega^2\times 42},\ldots,\textcolor{red}{\omega^3},\ldots, \textcolor{red}{\omega^4},\omega^4+1,\ldots,\\
-&\textcolor{red}{\omega^\omega2},\ldots, \textcolor{red}{\omega^\omega\times 2},\ldots,\\
+&\textcolor{red}{\omega^\omega},\ldots, \textcolor{red}{\omega^\omega+\omega^7\times 8},\ldots\textcolor{red}{\omega^\omega\times 2},\omega^\omega\times 2+1, \ldots,\\
 &\textcolor{red}{\omega^{\omega^\omega}},\ldots, \textcolor{red}{\omega^{\omega^\omega}+\omega^\omega\times 42+ \omega^{55}+\omega}, \ldots, \textcolor{red}{\omega^{\omega^{\omega+1}}}, \omega^{\omega^{\omega+1}}+1,\dots\\
-& \textcolor{red}{\epsilon_0 (= \omega^\epsilon_0)}, \epsilon_0+1, \ldots,\\
+& \textcolor{red}{\epsilon_0 (= \omega^{\epsilon_0)}}, \epsilon_0+1, \epsilon_0+2, \epsilon_0+3, \ldots,\\
 & \textcolor{red}{\epsilon_1}, \ldots, \textcolor{red}{\epsilon_2}, \ldots, \textcolor{red}{\epsilon_\omega},\ldots \\
@@
 For instance, $43$ is the successor of $42$, and $\omega^4+1$ the successor of $\omega^4$.
-\item Ordinals that follow dots (written in red) are called \emph{limit ordinals}.
+\item Ordinals that  follow dots immediately (written in red) are called \emph{limit ordinals}.
 With respect to the order induced by this sequence, any limit ordinal $\alpha$ is the least upper bound of  the set $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
+\item
 For instance $\omega$ is the least upper bound of the set of all finite ordinals (in the first line). $\omega$ is also the first limit ordinal, and the first infinite ordinal, in the sense that 
@@
 It plays an important role in proof theory, and is particularly studied in chapters~\ref{chap:T1} to \ref{chap:alpha-large}.
+\item All the ordinals shown in this list are either the ordinal \textcolor{blue}{$0$},
+a successor ordinal, or a \textcolor{red}{limit ordinal}.
 \end{itemize}
@@
 
-The best known examples of well-ordered sets are the set $\mathbb{N}$ of natural numbers (with the usual order $<$), as well as any finite segment $[0,i[=\{j\in\mathbb{N}|j<i\}$.
+The best known examples of well-ordered sets are the set $\mathbb{N}$ of natural numbers (with the usual order $<$), as well as any finite segment $[0,i)=\{j\in\mathbb{N}|j<i\}$.
 The disjoint union of two copies of $\mathbb{N}$, \emph{i.e.} the set $\{0,1\}\times\mathbb{N}$ is also well-ordered,
@@
 
-If we consider an ordinal $\alpha$ as a well-ordered set, then its elements are just the ordinals strictly less than $\alpha$, \emph{i.e.} the \emph{segment} $\mathbb{O}_\alpha=[0, \alpha[$.
-So, we can speak about \emph{finite}, \emph{infinite}, \emph{countable}, etc., ordinals.
-
-For instance, the elements of $\omega$ are the finite ordinals. $\omega$ is the first infinite ordinal.
+In a set-theoretic framework, one can consider any ordinal $\alpha$ as a well-ordered set, whose  elements are just the ordinals strictly less than $\alpha$, \emph{i.e.} the \emph{segment} $\mathbb{O}_\alpha=[0, \alpha)$.
+So, one can speak about \emph{finite}, \emph{infinite}, \emph{countable}, etc., ordinals.
+Nevertheless, since we work with type theory,
+we do not identify ordinals as sets of ordinals, but the correspondance between ordinals and sets of ordinals is the function that maps $\alpha$ to $\mathbb{O}_\alpha$.
+For instance $\mathbb{O}_\omega=\mathbb{N}$, and $\mathbb{O}_7=\{0,1,2,3,4,5,6\}$.
 
@@
 
-Chapter~\ref{chap:schutte} presents an adaptation to \coq{} of an axiomatization in classical logic of the set of countable ordinals by K. Schütte~\cite{schutte}. 
+\section{Ordinal numbers in Coq}
+
+\subsection{Countable ordinals}
+
+Chapter~\ref{chap:schutte} of this document presents an adaptation to \coq{} of an axiomatization in classical logic of the set of countable ordinals by K. Schütte~\cite{schutte}. 
 That formalization is quite complex and technical and unshamedly non-constructive,  so we put its description  in the last chapter of this document.
@@
 Please note that Schütte considers the (uncountable) set $\mathbb{O}$ of all countable ordinals.
-This set is well ordered (which is one of Schütte's axioms), and any ordinal $\alpha$ is associated with the segment $\mathbb{O}_\alpha$
-of all ordinals strictly less than $\alpha$.
+This set is well ordered (which is one of Schütte's axioms), and associates to any ordinal $\alpha$ the segment $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
+
+In our adaptation to \coq{}, we declare a type \texttt{Ord}, a binary relation \texttt{lt} (with infix notation \texttt{``\_<\_''}, and assume Schütte's axiom. In Chapter~\ref{chap:schutte},
+we derive some interesting properties of countable ordinals from these axioms.
+
+\subsection{Ordinal notations}
 
@@
 So, we will use \emph{ordinal notations}: Let $\alpha$ be some countable ordinal; In \coq{} terms, an ordinal notation for $\alpha$ will be a data type for representing all ordinals below $\alpha$ (excluding $\alpha$) and a bunch of functions for comparing ordinals, doing some arithmetic, inspecting the topology of the order (limits, successors), and proofs of well-foundedness).
-These functions should be proved correct, with respect to
-Schütte's model (work in progress).
-
-\section{Ordinal notations in Coq}
+These functions should be proved correct, with respect to Schütte's model.
 
@@
 
-Our definition of ordinal notation is inspired by our use of \coq{}: defining inductive data types, predicates, functions, etc, and proving properties of these functions.
+The following class definition, parameterized with a type $A$, a binary relation \texttt{lt} on $A$,
 
@@
 \item Let $\alpha$ be any ordinal, the \emph{segment} $\mathbb{O}_\alpha$ is the
-interval $[0,\alpha[\,=\,\{\beta|0\leq\beta<\alpha\}$.
+interval $[0,\alpha)\,=\,\{\beta|0\leq\beta<\alpha\}$.
 
@@
 Let $n$ be some natural number. The segment associated with $n$ is the interval 
-$[0,n[\,=\,\{0,1,\dots,n-1\}$.
+$)0,n[\,=\,\{0,1,\dots,n-1\}$.
 
@@
 
-For instance, let us build two elements of the segment $[0, 7[$, \emph{i.e.} two
+For instance, let us build two elements of the segment $[0, 7)$, \emph{i.e.} two
 inhabitants of   type (\texttt{t 7}), and prove a simple  inequality (see Fig.~\ref{fig:O7}).
@@
 
-Finally, it is interesting to show that, if $i<j$, then the segment $[0,i[$ is a ``sub-segment'' of
-$[0,j[$. Since the terms  $t\;i$ and $t\;j$ are not convertible, we consider a ``cast'' 
+Finally, it is interesting to show that, if $i<j$, then the segment $[0,i)$ is a ``sub-segment'' of
+$[0,j)$. Since the terms  $t\;i$ and $t\;j$ are not convertible, we consider a ``cast'' 
 function $\iota$ from $t\;i$ to $t\;j$, and prove this function is  a bijection  from $t\;i$ to
-the segment $[0,i[$ of $t\;j$.
+the segment $[0,i)$ of $t\;j$.
 
@@
 (** The segment associated with nA is isomorphic to
-    the interval [0,b[ *)
+    the interval [0,b) *)
 

### 2020-07-23 latex/chapter-schutte.tex
redaction introduction to ordinals

@@
 
-In the present chapter, we  compare our implementation of the segment $[0,\epsilon_0[$ with a mathematical text in order to ``validate'' our constructions.
+In the present chapter, we  compare our implementation of the segment $[0,\epsilon_0)$ with a mathematical text in order to ``validate'' our constructions.
 We chose as reference the axiomatic definition of the set of countable ordinals,

### 2020-07-23 latex/chapter-schutte.tex
redaction introduction to ordinals

@@
 \label{chap:schutte} 
+%ON
 
@@
 Let us declare a type 
-\texttt{ON} for representing countable ordinals, and a binary relation
- \texttt{lt}. Note that, in our development, \texttt{ON} is a type, while the \emph{set} of countable ordinals (called $\mathbb{O}$ by Schütte) 
-is the full set over the type \texttt{ON}.
-
-\index{Types!ON@ON (countable ordinals)}
+\texttt{Ord} for representing countable ordinals, and a binary relation
+ \texttt{lt}. Note that, in our development, \texttt{Ord} is a type, while the \emph{set} of countable ordinals (called $\mathbb{O}$ by Schütte) 
+is the full set over the type \texttt{Ord}.
+
+\index{Types!Ord@Ord (countable ordinals)}
 
@@
 \begin{Coqsrc}
-Parameter ON : Type.
-Parameter lt : relation ON.
+Parameter Ord : Type.
+Parameter lt : relation Ord.
 Notation "a < b" := (lt a b): schutte_scope.
 
-Definition ordinal := Full_set ON.
+Definition ordinal := Full_set Ord.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Axiom AX2 : forall X: Ensemble ON, 
+Axiom AX2 : forall X: Ensemble Ord, 
    (exists a,  (forall y, In X y -> y < a)) ->
@@
 
-Axiom AX3 : forall X : Ensemble ON,
+Axiom AX3 : forall X : Ensemble Ord,
               countable X -> 
@@
 
-According to the  definition of a well order, every non-empty subset of \texttt{ON} has a least element. Furthermore, this least element is unique.
-
-\begin{Coqsrc}
-Remark R : exists! z : ON, least_member lt  ordinal z.
+According to the  definition of a well order, every non-empty subset of \texttt{Ord} has a least element. Furthermore, this least element is unique.
+
+\begin{Coqsrc}
+Remark R : exists! z : Ord, least_member lt  ordinal z.
 Proof.
-  destruct inh_ON as [a]; apply (well_order (WO:=AX1)) with a .
+  destruct inh_Ord as [a]; apply (well_order (WO:=AX1)) with a .
   split.
@@
 \begin{Coqsrc}
-Definition zero : ON.
+Definition zero : Ord.
 Proof.
@@
 
-In order to use these tools,  we had to tell \coq{}  that the type \texttt{ON} is not empty:
-
-\begin{Coqsrc}
-Axiom inh_ON : inhabited ON.
+In order to use these tools,  we had to tell \coq{}  that the type \texttt{Ord} is not empty:
+
+\begin{Coqsrc}
+Axiom inh_Ord : inhabited Ord.
 \end{Coqsrc}
@@
 
-\index{Constants!zero:ON}
-
-\begin{Coqsrc}
-Definition zero: ON :=the_least ordinal.
+\index{Constants!zero:Ord}
+
+\begin{Coqsrc}
+Definition zero: Ord :=the_least ordinal.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Lemma zero_le (alpha : ON) :  zero <= alpha.
+Lemma zero_le (alpha : Ord) :  zero <= alpha.
 Proof.
@@
 \begin{enumerate}
-\item Prove that there exists a unique least member of \texttt{ON}
-\item Prove that being a least member of \texttt{ON} entails the announced inequality 
+\item Prove that there exists a unique least member of \texttt{Ord}
+\item Prove that being a least member of \texttt{Ord} entails the announced inequality 
 \end{enumerate}
@@
 
-  alpha : ON
+  alpha : Ord
   ============================
-  exists ! x : ON, least_member lt ordinal x
+  exists ! x : Ord, least_member lt ordinal x
 
 subgoal 2 (ID 156) is:
- forall a : ON, unique (least_member lt ordinal) a -> 
+ forall a : Ord, unique (least_member lt ordinal) a -> 
                 a <= alpha
@@
 
- Definition bottom := the_least (Empty_set ON).
+ Definition bottom := the_least (Empty_set Ord).
 \end{Coqbad}
@@
 Since we won't be able to prove the proposition
-\linebreak \Verb|{exists! a: ON, least_member (Empty_set ON) a|, the only properties we would be able to prove about \texttt{bottom} would be \emph{trivial} properties, 
-\emph{i.e.}, satisfied by \emph{any} element of type \texttt{ON}, like for instance
+\linebreak \Verb|{exists! a: Ord, least_member (Empty_set Ord) a|, the only properties we would be able to prove about \texttt{bottom} would be \emph{trivial} properties, 
+\emph{i.e.}, satisfied by \emph{any} element of type \texttt{Ord}, like for instance
 \texttt{bottom = bottom}, or \texttt{zero <= bottom}.
@@
   ============================
-  exists ! x : ON, least_member lt (Empty_set ON) x
+  exists ! x : Ord, least_member lt (Empty_set Ord) x
 
 subgoal 2 (ID 414) is:
- forall a : ON, unique (least_member lt (Empty_set ON)) a -> 
+ forall a : Ord, unique (least_member lt (Empty_set Ord)) a -> 
       a <= zero
@@
 
-The definition of the function \texttt{succ:ON -> ON} is very concise. The successor of any ordinal $\alpha$ is the smallest ordinal strictly greater than $\alpha$.
-
-\index{Functions!succ (in type ON)}
-
-\begin{Coqsrc}
-Definition succ (alpha : ON) := the_least (fun beta => alpha < beta).
+The definition of the function \texttt{succ:Ord -> Ord} is very concise. The successor of any ordinal $\alpha$ is the smallest ordinal strictly greater than $\alpha$.
+
+\index{Functions!succ (in type Ord)}
+
+\begin{Coqsrc}
+Definition succ (alpha : Ord) := the_least (fun beta => alpha < beta).
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Definition is_succ (alpha:ON) := exists beta, alpha = succ beta.
-
-Definition is_limit (alpha:ON) := alpha <> zero /\ ~ is_succ alpha.
+Definition is_succ (alpha:Ord) := exists beta, alpha = succ beta.
+
+Definition is_limit (alpha:Ord) := alpha <> zero /\ ~ is_succ alpha.
 \end{Coqsrc}
@@
 
-Fixpoint finite (i:nat) : ON :=
+Fixpoint finite (i:nat) : Ord :=
   match i with 
@@
 
-Coercion finite : nat >-> ON.
+Coercion finite : nat >-> Ord.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Definition succ_spec (alpha:ON) :=
+Definition succ_spec (alpha:Ord) :=
   least_member   lt (fun z => alpha < z).
@@
 
-  alpha : ON
+  alpha : Ord
   ============================
-  exists ! x : ON, succ_spec alpha x
+  exists ! x : Ord, succ_spec alpha x
 \end{Coqanswer}
@@
 \begin{Coqsrc}
-Lemma lt_succ (alpha : ON) :  alpha < succ alpha.
+Lemma lt_succ (alpha : Ord) :  alpha < succ alpha.
 Proof.
@@
 
-Lemma lt_succ_le (alpha beta : ON):
+Lemma lt_succ_le (alpha beta : Ord):
   alpha < beta -> succ alpha <= beta.
@@
 \begin{Coqsrc}
-Lemma lt_succ_le_2 (alpha beta : ON):
+Lemma lt_succ_le_2 (alpha beta : Ord):
   alpha < succ beta -> alpha <= beta.
 
-Lemma succ_mono (alpha beta : ON):
+Lemma succ_mono (alpha beta : Ord):
   alpha < beta -> succ alpha < succ beta.
 
-Lemma succ_monoR (alpha beta : ON) :
+Lemma succ_monoR (alpha beta : Ord) :
  succ alpha < succ beta -> alpha < beta.
 
-Lemma lt_succ_lt (alpha beta : ON) :
+Lemma lt_succ_lt (alpha beta : Ord) :
   is_limit beta ->  alpha < beta -> succ alpha < beta.
@@
 
-Definition sup (X: Ensemble ON) : ON  := the  (sup_spec X).
+Definition sup (X: Ensemble Ord) : Ord  := the  (sup_spec X).
 
@@
 Then, we define the function \texttt{omega\_limit} which returns the least upper bound 
-of the  (denumerable) range of any sequence \texttt{s: nat -> ON}. 
+of the  (denumerable) range of any sequence \texttt{s: nat -> Ord}. 
 By \texttt{AX3} this range is bounded, hence the set of its upper bounds is not empty and has a least element.
@@
 \begin{Coqsrc}
-Definition omega_limit (s:nat->ON) : ON 
+Definition omega_limit (s:nat->Ord) : Ord 
   := |_| (seq_range s).
@@
 
-Lemma lt_omega_finite alpha : ON) : 
+Lemma lt_omega_finite alpha : Ord) : 
   alpha < omega ->  exists i:nat, alpha =  i.
@@
 \begin{Coqsrc}
- Definition segment (A: Ensemble ON) :=
+ Definition segment (A: Ensemble Ord) :=
   forall alpha beta, In A alpha -> beta < alpha -> In A  beta.
 
-Definition proper_segment (A: Ensemble ON) :=
+Definition proper_segment (A: Ensemble Ord) :=
   segment A /\ ~ Same_set A ordinal.
@@
 \begin{Coqsrc}
-Definition ordering_function (f : ON -> ON)(A B : Ensemble ON) :=
+Definition ordering_function (f : Ord -> Ord)(A B : Ensemble Ord) :=
  segment A /\
@@
 
-Definition ordering_segment (A B : Ensemble ON) :=
-  exists f : ON -> ON, ordering_function f A B.
+Definition ordering_segment (A B : Ensemble Ord) :=
+  exists f : Ord -> Ord, ordering_function f A B.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Definition the_ordering_segment (B : Ensemble ON) :=
+Definition the_ordering_segment (B : Ensemble Ord) :=
   the  (fun x => ordering_segment x B).
 
-Definition ord  (B : Ensemble ON) := 
+Definition ord  (B : Ensemble Ord) := 
   some (fun f => ordering_function f (the_ordering_segment B) B).
@@
 \begin{Coqanswer}
-forall B : Ensemble ON,
- exists ! S : Ensemble ON, 
-      exists f : ON -> ON, ordering_function f S B
+forall B : Ensemble Ord,
+ exists ! S : Ensemble Ord, 
+      exists f : Ord -> Ord, ordering_function f S B
 \end{Coqanswer}
@@
 ordering_function_unicity :
-forall (B S1 S2 : Ensemble ON) (f1 f2 : ON -> ON),
+forall (B S1 S2 : Ensemble Ord) (f1 f2 : Ord -> Ord),
 ordering_function f1 S B ->
@@
 Th_13_5_2 :
-forall (A B : Ensemble ON) (f : ON -> ON),
+forall (A B : Ensemble Ord) (f : Ord -> Ord),
 ordering_function f A B -> Closed B -> continuous f A B
@@
 
-We are now ready to define and study addition on the type \texttt{ON}.
+We are now ready to define and study addition on the type \texttt{Ord}.
 The following definitions and proofs can be consulted in Module
@@
 \begin{Coqsrc}
-Lemma plus_elim (alpha : ON) :
-  forall P : (ON->ON)->Prop,
-    (forall f: ON->ON, 
+Lemma plus_elim (alpha : Ord) :
+  forall P : (Ord->Ord)->Prop,
+    (forall f: Ord->Ord, 
         ordering_function f ordinal (ge alpha)-> P f) ->
@@
 \begin{Coqsrc}
-Lemma alpha_plus_zero (alpha: ON): alpha + zero = alpha.
+Lemma alpha_plus_zero (alpha: Ord): alpha + zero = alpha.
 Proof.
@@
 
-  alpha : ON
+  alpha : Ord
   ============================
-  forall f : ON -> ON,
+  forall f : Ord -> Ord,
   ordering_function f ordinal (ge alpha) -> 
@@
 
-Lemma zero_plus_alpha (alpha : ON) : zero + alpha = alpha.
-
-Lemma le_plus_l (alpha beta : ON) : alpha <= alpha + beta.
-
-Lemma le_plus_r (alpha beta : ON) :  beta <= alpha + beta.
-
-Lemma plus_mono_r (alpha beta gamma : ON) : 
+Lemma zero_plus_alpha (alpha : Ord) : zero + alpha = alpha.
+
+Lemma le_plus_l (alpha beta : Ord) : alpha <= alpha + beta.
+
+Lemma le_plus_r (alpha beta : Ord) :  beta <= alpha + beta.
+
+Lemma plus_mono_r (alpha beta gamma : Ord) : 
     beta < gamma -> alpha + beta < alpha + gamma.
 
-Lemma plus_of_succ (alpha beta : ON) :
+Lemma plus_of_succ (alpha beta : Ord) :
     alpha + (succ beta) = succ (alpha + beta).
 
-Theorem plus_assoc (alpha beta gamma : ON) :
+Theorem plus_assoc (alpha beta gamma : Ord) :
   alpha + (beta + gamma) = (alpha + beta) + gamma.
@@
 
-Lemma finite_plus_ge_omega (n : nat) (alpha : ON) :
+Lemma finite_plus_ge_omega (n : nat) (alpha : Ord) :
   omega <= alpha -> n + alpha = alpha.
@@
 \begin{Coqsrc}
-Fixpoint mult_Sn (alpha:ON)(n:nat){struct n} :ON :=
+Fixpoint mult_Sn (alpha:Ord)(n:nat){struct n} :Ord :=
  match n with 
@@
 \begin{Coqsrc}
-Definition AP : Ensemble ON :=
+Definition AP : Ensemble Ord :=
   fun alpha => 
@@
 \begin{Coqsrc}
-Fixpoint omega_tower (i : nat) : ON :=
+Fixpoint omega_tower (i : nat) : Ord :=
   match i with
@@
 \begin{Coqsrc}
-Lemma AP_plus_closed (alpha beta gamma : ON): 
+Lemma AP_plus_closed (alpha beta gamma : Ord): 
      In AP alpha -> beta < alpha -> gamma < alpha -> beta + gamma < alpha.
@@
 \begin{Coqsrc} 
-Definition Closed (B : Ensemble ON) : Prop := 
+Definition Closed (B : Ensemble Ord) : Prop := 
   forall M, Included M B -> Inhabited _ M -> 
@@
 \begin{Coqsrc}
-Lemma AP_phi0 (alpha : ON) : In AP (phi0 alpha).
+Lemma AP_phi0 (alpha : Ord) : In AP (phi0 alpha).
 
@@
 
-Lemma phi0_mono (alpha beta : ON) :
+Lemma phi0_mono (alpha beta : Ord) :
   alpha < beta ->  phi0 alpha < phi0 beta.
 
-Lemma phi0_inj (alpha beta : ON) :
+Lemma phi0_inj (alpha beta : Ord) :
     phi0 alpha = phi0 beta -> alpha = beta.
 
-Lemma phi0_sup : forall (U: Ensemble ON),
+Lemma phi0_sup : forall (U: Ensemble Ord),
    Inhabited _ U ->   countable U ->  phi0 (|_| U) = |_| (image U phi0).
 
-Lemma is_limit_phi0 (alpha : ON) :
+Lemma is_limit_phi0 (alpha : Ord) :
   zero < alpha ->  is_limit (phi0 alpha).
@@
 
-Lemma phi0_le (alpha : ON) : alpha <= phi0 alpha.
+Lemma phi0_le (alpha : Ord) : alpha <= phi0 alpha.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Lemma lt_phi0 (alpha : ON):
+Lemma lt_phi0 (alpha : Ord):
   alpha < epsilon0 -> alpha < phi0 alpha.
@@
 \begin{Coqsrc}
-Definition Cr_fun : forall alpha : ON,
-       (forall beta : ON, beta < alpha -> Ensemble ON) ->
-        Ensemble ON 
+Definition Cr_fun : forall alpha : Ord,
+       (forall beta : Ord, beta < alpha -> Ensemble Ord) ->
+        Ensemble Ord 
 := 
-   fun (alpha :ON)
+   fun (alpha :Ord)
         (Cr : forall beta, 
-                beta < alpha -> Ensemble ON) 
-        (x : ON) => (
+                beta < alpha -> Ensemble Ord) 
+        (x : Ord) => (
        (alpha = zero /\ AP x) \/
@@
 
-Definition Cr (alpha : ON) : Ensemble ON := 
-    (Fix  all_ord_acc (fun (_:ON) => Ensemble ON) Cr_fun) alpha.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Definition phi (alpha : ON) : ON -> ON 
+Definition Cr (alpha : Ord) : Ensemble Ord := 
+    (Fix  all_ord_acc (fun (_:Ord) => Ensemble Ord) Cr_fun) alpha.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Definition phi (alpha : Ord) : Ord -> Ord 
     :=  ord (Cr alpha).
 
-Definition A (alpha : ON) : Ensemble ON :=
+Definition A (alpha : Ord) : Ensemble Ord :=
   the_ordering_segment (Cr alpha).
@@
 Section Proof_of_Lemma5.
-  Let P (alpha:ON) := Unbounded (Cr alpha) /\ Closed (Cr alpha).
+  Let P (alpha:Ord) := Unbounded (Cr alpha) /\ Closed (Cr alpha).
 
@@
 \begin{Coqsrc}
- Definition cnf_t := list ON.
-
-Fixpoint eval (l : cnf_t) : ON :=
+ Definition cnf_t := list Ord.
+
+Fixpoint eval (l : cnf_t) : Ord :=
   match l with nil => zero
@@
 
-Definition is_cnf_of (alpha : ON)(l : cnf_t) : Prop :=
+Definition is_cnf_of (alpha : Ord)(l : cnf_t) : Prop :=
   sorted l /\ alpha = eval l.
@@
 \begin{Coqsrc}
-Theorem cnf_exists (alpha : ON) :
+Theorem cnf_exists (alpha : Ord) :
   exists l: cnf_t, is_cnf_of alpha l.
@@
 
-Theorem cnf_exists_unique (alpha:ON) :
+Theorem cnf_exists_unique (alpha:Ord) :
   exists! l: cnf_t, is_cnf_of alpha l.
@@
 
-\section{An embedding of \texttt{T1} into \texttt{ON}}
+\section{An embedding of \texttt{T1} into \texttt{Ord}}
 
@@
 a classical mathematical discourse considered as a model. 
-First, we define a function from \texttt{T1} into  \texttt{ON} by structural recursion.
-
-\begin{Coqsrc}
-Fixpoint inject (t:T1) : ON :=
+First, we define a function from \texttt{T1} into  \texttt{Ord} by structural recursion.
+
+\begin{Coqsrc}
+Fixpoint inject (t:T1) : Ord :=
  match t with 
@@
 Finally, we prove that \texttt{inject} is a bijection from the set of all terms of \texttt{T1} in normal form to the set 
-\texttt{members epsilon0} of the elements of \texttt{ON} strictly less than  $\epsilon_0$.
+\texttt{members epsilon0} of the elements of \texttt{Ord} strictly less than  $\epsilon_0$.
 

### 2020-07-24 latex/chapter-ordinals.tex
redaction introduction to ordinals

@@
 
-The following class definition, parameterized with a type $A$, a binary relation \texttt{lt} on $A$,
+The following class definition, parameterized with a type $A$, a binary relation \texttt{lt} on $A$, specifies that \texttt{lt} is a well-founded strict order.
+The reflexive closure of \texttt{lt}, (called \texttt{le}, for ``less  or equal than'') is a total decidable order, implemented through a comparison function \texttt{compare}.
 
@@
 \begin{Coqsrc}
-Class OrdinalNotation {A:Type}{lt: relation A}(sto:StrictOrder lt)
+Class ON {A:Type}(lt: relation A)
       (compare : A -> A -> comparison)  :=
-  {   wf : well_founded lt;
+  {
+  sto :> StrictOrder lt;
+  wf : well_founded lt;
   compare_correct :
@@
   }.
-
+\end{Coqsrc}
+
+The following definitions allow us to make implicit several guessable arguments.
+
+\begin{Coqsrc}
 Definition on_t  {A:Type}{lt: relation A}
             {compare : A -> A -> comparison}
-            {on : OrdinalNotation lt compare} := A.
+            {on : ON lt compare} := A.
 
@@
             {compare : A -> A -> comparison}
-            {on : OrdinalNotation lt compare} := compare.
+            {on : ON lt compare} := compare.
 
@@
            {compare : A -> A -> comparison}
-           {on : OrdinalNotation lt compare} := lt.
+           {on : ON lt compare} := lt.
+Infix "<" := on_lt : ON_scope.
 
@@
            {compare : A -> A -> comparison}
-           {on : OrdinalNotation lt compare} :=
+           {on : ON lt compare} :=
   clos_refl _ on_lt.
 
-\end{Coqsrc}
+Infix "<=" := on_le : ON_scope.
+\end{Coqsrc}
+
+\subsection{Comparing Ordinal Notations}
+
+It is sometimes useful to compare two ordinal notations with respect to expressive power
+(the segment of ordinals one they represent).
+
+The following class specifies a strict inclusion of segments.
+The notation \texttt{OA} describes a segment $[0,alpha($, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$).
+We require that the comparison functions of the two notation systems are compatible.
+
+\begin{Coqsrc}
+Class  SubON 
+       `(OA : @ON A ltA  compareA)
+       `(OB : @ON B ltB  compareB)
+       (alpha :  B)
+       (iota : A -> B):=
+  {
+  subseg_compare :forall x y : A,  compareB (iota x) (iota y) =
+                                 compareA x y;
+  subseg_incl : forall x, ltB (iota x) alpha;
+  subseg_onto : forall y, ltB y alpha  -> exists x:A, iota x = y}.
+\end{Coqsrc}
+
+\begin{todo}
+pointers to examples 
+\end{todo}
+
+In some cases we want to show that two notation systems describe the same segment (for instance $3+\omega$ and $\omega$).
+For that purpose, one may prove that the two notation systems are order-isomorphic.
+
+\begin{Coqsrc}
+Class  ON_Isomorphic 
+       `(OA : @ON A ltA compareA)
+       `(OB : @ON B ltB  compareB)
+       (f : A -> B)
+       (g : B -> A):=
+  {
+  iso_compare_comm :forall x y : A,  compareB (f x) (f y) =
+                                 compareA x y;
+  iso_inv1 : forall a, g (f a)= a;
+  iso_inv2 : forall b, f (g b) = b}.
+\end{Coqsrc}
+
+Finally, it may be interesting to compare an ordinal notation with the more theoretical model from Schütte (well, at least with our formalization of thhat model).
+This would be a relative proof of correctenss of our ordinal  notations.
+
+The following class specifies that a notation \texttt{OA} describes a segment $[0,\alpha)$,
+where $\alpha$ is a countable ordinal à la Schütte.
+
+\begin{Coqsrc}
+Class ON_for `(alpha : Schutte_basics.Ord)
+     `(OA : @ON A ltA  compareA)
+      (iota : A -> Schutte_basics.Ord) :=
+  { conform_inj : forall a, Schutte_basics.lt (iota a) alpha;
+    conform_surj : forall beta, Schutte_basics.lt beta alpha ->
+                                exists b, iota b = beta;
+    conform_compare : forall a b:A,
+        match compareA a b with
+          Datatypes.Lt => Schutte_basics.lt (iota a) (iota b)
+        | Datatypes.Eq => iota a = iota b
+        | Datatypes.Gt => Schutte_basics.lt (iota b) (iota a)
+        end}.
+\end{Coqsrc}
+
+For instance, the following theorem tells that \texttt{Epsilon0}, our otation system for the segment $[0,\epsilon0)$ is correct with respect to the countqable ordinal $\epsilon_0$
+(see chapter~\ref{chap:schutte} for more details).
+
+\begin{Coqsrc}
+Instance Epsilon0_correct :
+  ON_for epsilon0 Epsilon0  (fun alpha => inject (cnf alpha)).
+\end{Coqsrc}
+
+\subsection{Examples}
+
+The simplest example of ordinal notation is
 

### 2020-07-24 latex/chapter-ordinals.tex
redaction introduction to ordinals

@@
 Fortunately, the ordinals we need for  studying hydra battles are much simpler than Schütte's, and can be represented as rather simple data types in \gallina.
-So, we will use \emph{ordinal notations}: Let $\alpha$ be some countable ordinal; In \coq{} terms, an ordinal notation for $\alpha$ will be a data type for representing all ordinals below $\alpha$ (excluding $\alpha$) and a bunch of functions for comparing ordinals, doing some arithmetic, inspecting the topology of the order (limits, successors), and proofs of well-foundedness).
+So, we will use \emph{ordinal notations} (also called \emph{ordinal notation systems}).
+Let $\alpha$ be some countable ordinal; In \coq{} terms, an ordinal notation for $\alpha$ will be a data type for representing all ordinals below $\alpha$ (excluding $\alpha$) and a bunch of functions for comparing ordinals, doing some arithmetic, inspecting the topology of the order (limits, successors), and proofs of well-foundedness).
 These functions should be proved correct, with respect to Schütte's model.
@@
 
+\subsubsection{Examples}
+
+The simplest example of ordinal notation is built over the type \texttt{nat} of \coq's standard library. We have only to apply already proven lemmas about Peano numbers.
+
+\vspace{4pt}
+\noindent\emph{From Library~\href{../src/html/hydras.OrdinalNotations/Omega_ON.html}{OrdinalNotations.Omega\_ON}}
+
+\begin{Coqsrc}
+Global Instance Omega : ON  Peano.lt Nat.compare.
+Proof.
+ split.
+ - apply Nat.lt_strorder.
+ - apply Wf_nat.lt_wf.
+ - apply Nat.compare_spec.
+Qed.
+\end{Coqsrc}
+
+More complex examples will be developped in further sections or chapters. For instance,
+a notation system for ordinals less than $\epsilon_0$ is presented in chapter~\ref{chap:T1}.
+Notation systems for $\omega+\omega$, $\omega^2$, etc. will be given in the next paragraphs.
+
 \subsection{Comparing Ordinal Notations}
@@
 
-In some cases we want to show that two notation systems describe the same segment (for instance $3+\omega$ and $\omega$).
-For that purpose, one may prove that the two notation systems are order-isomorphic.
+In some cases we want to show that two notation systems describe the same segment (for instance $[0,3+\omega)$ and $[0,\omega($).
+For this purpose, one may prove that the two notation systems are order-isomorphic.
 
@@
   {
-  iso_compare_comm :forall x y : A,  compareB (f x) (f y) =
+  iso_compare: forall x y : A,  compareB (f x) (f y) =
                                  compareA x y;
@@
       (iota : A -> Schutte_basics.Ord) :=
-  { conform_inj : forall a, Schutte_basics.lt (iota a) alpha;
-    conform_surj : forall beta, Schutte_basics.lt beta alpha ->
+  { ON_for_inj : forall a, Schutte_basics.lt (iota a) alpha;
+    ON_for_onto : forall beta, Schutte_basics.lt beta alpha ->
                                 exists b, iota b = beta;
-    conform_compare : forall a b:A,
+    On_compare_spec : forall a b:A,
         match compareA a b with
@@
 \end{Coqsrc}
-
-\subsection{Examples}
-
-The simplest example of ordinal notation is
 

### 2020-07-25 latex/chapter-ordinals.tex
redaction introduction to ordinals

@@
 
-\subsubsection{Examples}
+\subsection{Examples of Notation systems}
+
+\subsubsection{The ordinal \texorpdfstring{$\omega$}{omega}}
 
@@
 
+\subsection{The ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
+
+The ordinal $\omega+\omega$ (also known as $\omega\times 2$) may be represented as the concatenation 
+of two copies of $\omega$ (Figure~\ref{fig:omega-plus-omega}).
+
+\begin{figure}[h]
+   \centering
+   \begin{tikzpicture}[very thick, scale=0.5]
+\begin{scope}[color=blue]
+\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
+\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
+\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
+\node (Adots) at (6,0) {$\ldots$};
+\node(An) at (8,0)[label=below:$n$]{$\bullet$};
+\node(A2) at (10,0)[label=below:$n+1$]{$\bullet$};
+\node (Adots1) at (12,0) {$\ldots$};
+\end{scope}
+\begin{scope}[color=red]
+\node(B0) at (14,0)[label=below:$0$,label=above:$\omega$]{$\bullet$};
+\node(B1) at (16,0)[label=below:$1$, label=above:$\omega+1$]{$\bullet$};
+\node(B2) at (18,0)[label=below:$2$,label=above:$\omega+2$]{$\bullet$};
+\node (Bdots) at (20,0) {$\ldots$};
+\node (Bn) at (22,0) [label=below:$p$, label=above:$\omega+p$]{$\bullet$};
+\node (Bdots2) at (24,0) {$\ldots$};
+\end{scope}
+\end{tikzpicture}
+   \caption{\textcolor{blue}{$\omega+{\color{red}\omega}$}}
+   \label{fig:omega-plus-omega}
+ \end{figure}
+
+In \coq{}, we use the disjoint union \texttt{(nat+nat)\%type} for representing the elements of $\omega+\omega$.
+The blue ordinals of Figure~\ref{fig:omega-plus-omega} will be represented by terms of the form \texttt{(inl $i$)},
+the red ones by \texttt{(inr $j$)}.
+
+\vspace{4pt}
+\noindent\emph{From Module~\href{../src/html/hydras.Prelude.Omega_plus_omega.html}{Prelude.Omega\_plus\_omega}}
+
+\begin{Coqsrc}
+Declare Scope oo_scope.
+Delimit Scope oo_scope with oo.
+Open Scope oo_scope.
+
+Definition t := (nat + nat)%type.
+Arguments inl  {A B} _.
+Arguments inr  {A B} _.
+\end{Coqsrc}
+
+The order on type \texttt{t} is already defined in
+Library~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}.
+
+\begin{Coqanswer}
+Inductive
+le_AsB (A B : Type) (leA : A -> A -> Prop) (leB : B -> B -> Prop)
+  : A + B -> A + B -> Prop :=
+    le_aa : forall x y : A, leA x y -> le_AsB A B leA leB (inl x) (inl y)
+  | le_ab : forall (x : A) (y : B), le_AsB A B leA leB (inl x) (inr y)
+  | le_bb : forall x y : B, leB x y -> le_AsB A B leA leB (inr x) (inr y)
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Definition lt : relation t := le_AsB _ _ Peano.lt Peano.lt.
+Infix "<" := lt : oo_scope.
+
+Definition le := clos_refl _ lt.
+
+Infix "<=" := le : oo_scope.
+\end{Coqsrc}
+
+In order to build an instance of \texttt{ON}, we have to verify that \texttt{lt} is a well-founded strict order and define a correct comparison function.
+
+\begin{Coqsrc}
+Instance lt_strorder : StrictOrder lt.
+
+Lemma lt_wf : well_founded lt.
+Proof. apply wf_disjoint_sum; apply lt_wf. Qed.
+
+Definition compare (alpha beta: t) : comparison :=
+   match alpha, beta with
+     inl _, inr _ => Lt
+   | inl n, inl p | inr n, inr p => Nat.compare n p
+   | inr _, inl _ => Gt
+  end.
+
+Lemma compare_correct alpha beta :
+    CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
+              (compare alpha beta).
+
+Instance OmegaPlusOmega : ON lt compare.
+Proof.
+  split.
+  - apply lt_strorder.
+  - apply lt_wf.
+  - apply compare_correct.
+Qed.
+
+\end{Coqsrc}
+
 More complex examples will be developped in further sections or chapters. For instance,
 a notation system for ordinals less than $\epsilon_0$ is presented in chapter~\ref{chap:T1}.
-Notation systems for $\omega+\omega$, $\omega^2$, etc. will be given in the next paragraphs.
 
@@
 
-\subsection{The ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
-
-The ordinal $\omega+\omega$ (also known as $\omega\times 2$) may be represented as the concatenation 
-of two copies of $\omega$ (Figure~\ref{fig:omega-plus-omega}).
-
-\begin{figure}[h]
-   \centering
-   \begin{tikzpicture}[very thick, scale=0.5]
-\begin{scope}[color=blue]
-\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
-\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
-\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
-\node (Adots) at (6,0) {$\ldots$};
-\node(An) at (8,0)[label=below:$n$]{$\bullet$};
-\node(A2) at (10,0)[label=below:$n+1$]{$\bullet$};
-\node (Adots1) at (12,0) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red]
-\node(B0) at (14,0)[label=below:$0$,label=above:$\omega$]{$\bullet$};
-\node(B1) at (16,0)[label=below:$1$, label=above:$\omega+1$]{$\bullet$};
-\node(B2) at (18,0)[label=below:$2$,label=above:$\omega+2$]{$\bullet$};
-\node (Bdots) at (20,0) {$\ldots$};
-\node (Bn) at (22,0) [label=below:$p$, label=above:$\omega+p$]{$\bullet$};
-\node (Bdots2) at (24,0) {$\ldots$};
-\end{scope}
-\end{tikzpicture}
-   \caption{\textcolor{blue}{$\omega+{\color{red}\omega}$}}
-   \label{fig:omega-plus-omega}
- \end{figure}
-
-In \coq{}, we use the disjoint union \texttt{(nat+nat)\%type} for representing the elements of $\omega+\omega$.
-The blue ordinals of Figure~\ref{fig:omega-plus-omega} will be represented by terms of the form \texttt{(inl $i$)},
-the red ones by \texttt{(inr $j$)}.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../src/html/hydras.Prelude.Omega_plus_omega.html}{Prelude.Omega\_plus\_omega}}
-
-\begin{Coqsrc}
-Declare Scope oo_scope.
-Delimit Scope oo_scope with oo.
-Open Scope oo_scope.
-
-Definition t := (nat + nat)%type.
-Arguments inl  {A B} _.
-Arguments inr  {A B} _.
-\end{Coqsrc}
-
-The order on type \texttt{t} is already defined in
-Library~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}.
-
-\begin{Coqanswer}
-Inductive
-le_AsB (A B : Type) (leA : A -> A -> Prop) (leB : B -> B -> Prop)
-  : A + B -> A + B -> Prop :=
-    le_aa : forall x y : A, leA x y -> le_AsB A B leA leB (inl x) (inl y)
-  | le_ab : forall (x : A) (y : B), le_AsB A B leA leB (inl x) (inr y)
-  | le_bb : forall x y : B, leB x y -> le_AsB A B leA leB (inr x) (inr y)
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Definition lt : relation t := le_AsB _ _ Peano.lt Peano.lt.
-Infix "<" := lt : oo_scope.
-
-Definition le := clos_refl _ lt.
-
-Infix "<=" := le : oo_scope.
-\end{Coqsrc}
-
 \subsubsection{Finite ordinals as members of \texorpdfstring{$\omega+\omega$}{omega+omega}}

### 2020-07-26 latex/chapter-hydras.tex

@@
 
-\subsection{Formal definition of battles}
+\subsection{Big steps}
 
@@
 Inductive battle (B:Battle) : nat -> Hydra -> nat -> Hydra -> Prop :=
+
 | battle_1 : forall i h  h', battle_r   B i  h h' -> 

### 2020-07-26 latex/chapter-introduction.tex

@@
 
-\subsubsection*{Contributions are welcome}
+\subsubsection*{Contributions}
+Many thanks to \'Evelyne Contejean and Florian Hatat for their contribution.
+\'Evelyne contributed libraries on the recusrive path ordering (\emph{rpo}) for proving the well-foundedness of our representation of $\epsilon_0$ and $\Gamma_0$.
+Florian Hatat proved many useful lemmas on countable sets, which we used in our adaptation of Schütte's formalization of countable ordinals.
+
 \label{sec:orgheadline2}

### 2020-07-26 latex/chapter-ks.tex

@@
 The next step of our development is to extend this relationship to
-the order $<$ on $[0,\epsilon_0)$ on one side, and hydra fights on the other side.
-
-\begin{Coqsrc}
-Lemma path_to_fight alpha s beta :
+the order $<$ on $[0,\epsilon_0)$ on one side, and hydra battles on the other side.
+
+\begin{Coqsrc}
+Lemma path_to_battle alpha s beta :
   path_to  beta  s alpha -> nf alpha ->

### 2020-07-27 latex/chapter-ordinals.tex
redaction introduction to ordinals

@@
 Fortunately, the ordinals we need for  studying hydra battles are much simpler than Schütte's, and can be represented as rather simple data types in \gallina.
-So, we will use \emph{ordinal notations} (also called \emph{ordinal notation systems}).
+So, we will use \emph{ordinal notations} (also called \emph{[ordinal] notation systems}).
 Let $\alpha$ be some countable ordinal; In \coq{} terms, an ordinal notation for $\alpha$ will be a data type for representing all ordinals below $\alpha$ (excluding $\alpha$) and a bunch of functions for comparing ordinals, doing some arithmetic, inspecting the topology of the order (limits, successors), and proofs of well-foundedness).
@@
 
-\subsection{Examples of Notation systems}
+\subsection{Instances of notation systems}
 
@@
 \vspace{4pt}
-\noindent\emph{From Library~\href{../src/html/hydras.OrdinalNotations/Omega_ON.html}{OrdinalNotations.Omega\_ON}}
+\noindent\emph{From Library~\href{../src/html/hydras.OrdinalNotations/ON_Omega.html}{OrdinalNotations.ON\_Omega}}
 
@@
  - apply Nat.compare_spec.
+Qed.
+\end{Coqsrc}
+
+\subsection{Sum of two ordinal notations}
+
+Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective types \texttt{A} and \texttt{B}.
+
+ We consider a new strict order
+on the disjoint sum of the associated types, by putting all elements of \texttt{A} before the elements of \texttt{B} (thanks to Standard Library's relation operator \texttt{le\_AsB}).
+(please look at Figure~\ref{fig:omega-plus-omega} for an example).
+
+\emph{From Library~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}}.
+
+\begin{Coqanswer}
+Inductive
+le_AsB (A B : Type) (leA : A -> A -> Prop) (leB : B -> B -> Prop)
+  : A + B -> A + B -> Prop :=
+    le_aa : forall x y : A, leA x y -> le_AsB A B leA leB (inl x) (inl y)
+  | le_ab : forall (x : A) (y : B), le_AsB A B leA leB (inl x) (inr y)
+  | le_bb : forall x y : B, leB x y -> le_AsB A B leA leB (inr x) (inr y)
+\end{Coqanswer}
+
+\vspace{4pt}
+\noindent\emph{From Library~\href{../src/html/hydras.OrdinalNotations/ON_plus.html}{OrdinalNotations.ON\_plus}}
+
+\begin{Coqsrc}
+Section Defs.
+
+  Context `(ltA: relation A)
+          (compareA : A -> A -> comparison)
+          (NA: ON ltA compareA).
+  Context `(ltB: relation B)
+          (compareB : B -> B -> comparison)
+          (NB: ON ltB compareB).
+
+Definition t := (A + B)%type.
+Arguments inl  {A B} _.
+Arguments inr  {A B} _.
+
+Definition lt : relation t := le_AsB _ _ ltA ltB.
+\end{Coqsrc}
+
+In order to build an instance of \texttt{ON}, we have to define a comparison function.
+
+\begin{Coqsrc}
+Definition compare (alpha beta: t) : comparison :=
+   match alpha, beta with
+     inl _, inr _ => Lt
+   | inl a, inl a' => compareA a a'
+   | inr b, inr b' => compareB b b'
+   | inr _, inl _ => Gt
+  end.
+
+Lemma compare_correct alpha beta :
+    CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
+              (compare alpha beta).
+\end{Coqsrc}
+
+The Lemma \texttt{Wellfounded.Disjoint\_Union.wf\_disjoint\_sum} of Standard Library
+helps us to prove that our order \texttt{lt} is well-founded.
+
+\begin{Coqsrc}
+Global Instance ON_plus : ON lt compare.
+Proof.
+  split.
+  - apply lt_strorder.
+  -  apply lt_wf.
+  - apply compare_correct.
 Qed.
@@
 
-In \coq{}, we use the disjoint union \texttt{(nat+nat)\%type} for representing the elements of $\omega+\omega$.
-The blue ordinals of Figure~\ref{fig:omega-plus-omega} will be represented by terms of the form \texttt{(inl $i$)},
-the red ones by \texttt{(inr $j$)}.
+We can define this notation in \coq{} as an instance of \texttt{ON\_plus}.
 
 \vspace{4pt}
-\noindent\emph{From Module~\href{../src/html/hydras.Prelude.Omega_plus_omega.html}{Prelude.Omega\_plus\_omega}}
-
-\begin{Coqsrc}
-Declare Scope oo_scope.
-Delimit Scope oo_scope with oo.
-Open Scope oo_scope.
-
-Definition t := (nat + nat)%type.
-Arguments inl  {A B} _.
-Arguments inr  {A B} _.
-\end{Coqsrc}
-
-The order on type \texttt{t} is already defined in
-Library~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}.
-
-\begin{Coqanswer}
-Inductive
-le_AsB (A B : Type) (leA : A -> A -> Prop) (leB : B -> B -> Prop)
-  : A + B -> A + B -> Prop :=
-    le_aa : forall x y : A, leA x y -> le_AsB A B leA leB (inl x) (inl y)
-  | le_ab : forall (x : A) (y : B), le_AsB A B leA leB (inl x) (inr y)
-  | le_bb : forall x y : B, leB x y -> le_AsB A B leA leB (inr x) (inr y)
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Definition lt : relation t := le_AsB _ _ Peano.lt Peano.lt.
-Infix "<" := lt : oo_scope.
-
-Definition le := clos_refl _ lt.
-
-Infix "<=" := le : oo_scope.
-\end{Coqsrc}
-
-In order to build an instance of \texttt{ON}, we have to verify that \texttt{lt} is a well-founded strict order and define a correct comparison function.
-
-\begin{Coqsrc}
-Instance lt_strorder : StrictOrder lt.
-
-Lemma lt_wf : well_founded lt.
-Proof. apply wf_disjoint_sum; apply lt_wf. Qed.
-
-Definition compare (alpha beta: t) : comparison :=
-   match alpha, beta with
-     inl _, inr _ => Lt
-   | inl n, inl p | inr n, inr p => Nat.compare n p
-   | inr _, inl _ => Gt
-  end.
-
-Lemma compare_correct alpha beta :
-    CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
-              (compare alpha beta).
-
-Instance OmegaPlusOmega : ON lt compare.
-Proof.
-  split.
-  - apply lt_strorder.
-  - apply lt_wf.
-  - apply compare_correct.
-Qed.
-
-\end{Coqsrc}
+\noindent\emph{From Module~\href{../src/html/hydras.Prelude.ON_Omega_plus_omega.html}{Prelude.ON\_Omega\_plus\_omega}}
+
+\begin{Coqsrc}
+Definition Omega_plus_Omega := ON_plus Omega Omega.
+
+Existing Instance Omega_plus_Omega.
+Definition t := @ON_plus.t nat nat.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Example ex1 : inl 7 < inr 0.
+Proof. constructor. Qed.
+\end{Coqsrc}
+
+We can now define abbreviations.
+For instance, the finite ordinals will be represented by the terms built with  the constructor \texttt{inl}, and the first infinite ordinal $\omega$ by the term \texttt{(inr 0)}.
+
+\begin{Coqsrc}
+Definition fin (i:nat) : t := inl i.
+Coercion fin : nat >-> t.
+
+Notation "'omega'" := (inr  0:t).
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Example ex1' : fin 7 < omega.
+Proof. constructor. Qed.
+\end{Coqsrc}
+
+\section{Limits and Successors}
+
+Let us look again at our implementation of $\omega+\omega$. We can distinguish between the three kinds of ordinals seen in Fig~\ref{fig:ordinal-sequence}:
+
+\begin{itemize}
+\item The least ordinal, \texttt{(inl 0)}, also written \texttt{(fin 0)}.
+\item The limit ordinal $\omega$.
+\item The successor ordinals, either of the form \texttt{(inl (S $i$))} or \texttt{(inr (S $i$))}
+\end{itemize}
+
+It would be interesting to specify at the most generic level, what is a zero, a successor or a limit ordinal.
+
+%%%% ICI ICI %%%%%%%%%%%
 
@@
 
-\subsection{Comparing Ordinal Notations}
+\section{Comparing Ordinal Notations}
 

### 2020-07-28 latex/chapter-ordinals.tex
redaction introduction to ordinals

@@
 
+\index{Coq!Techniques!Coercions} 
+\begin{remark}
+Beware of coercions and notation scopes!
+
+Let us consider the following goal:
+
+\begin{Coqsrc}
+ Goal (6 < 8).
+ auto with arith.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+1 subgoal (ID 9)
+
+  ============================
+  6 < 8
+\end{Coqanswer}
+
+Please keep in mind that the current notation scope interprets the infix \texttt{``<''} as the predicate \texttt{Omega\_plus\_omega.lt} and not \texttt{Nat.lt}.
+More,  the coercion mechanism converts the terms \texttt{6:nat} [resp.
+\texttt{8:nat} ]
+into \texttt{inl 6} [resp. \texttt{inl 8}].  So, the initial goal is correctly interpreted by \coq{}, but not as an inequality between two natural numbers.
+
+\begin{Coqsrc}
+Set Printing All.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+1 subgoal (ID 337)
+
+  ============================
+  @on_lt nat Peano.lt Nat.compare Omega (S (S (S (S (S (S O))))))
+    (S (S (S (S (S (S (S (S O))))))))
+\end{Coqanswer}
+
+Anyway, the initial goal is provable, using \texttt{le\_AsB} first constructor.
+
+\begin{Coqsrc}
+  constructor; auto with arith.
+Qed.
+\end{Coqsrc}
+
+\end{remark}
+
 \section{Limits and Successors}
@@
 
-It would be interesting to specify at the most generic level, what is a zero, a successor or a limit ordinal.
-
-%%%% ICI ICI %%%%%%%%%%%
-
-More complex examples will be developped in further sections or chapters. For instance,
-a notation system for ordinals less than $\epsilon_0$ is presented in chapter~\ref{chap:T1}.
+It would be interesting to specify at the most generic level, what is a zero, a successor or a limit ordinal. Let $<$ be a strict order on a type $A$.
+
+\begin{itemize}
+\item A \emph{least} element is a minorant (in the large sense) of the full set  on $A$,
+\item $y$ is a \emph{successor} of $x$ if $x<y$ and there is no element between $x$ and $y$,
+\item $x$ is a \emph{limit} if $x$ is not a least element, and for any $y$ such that $y<x$,
+ there exists some $z$ such that $y<z<x$.
+\end{itemize}
+
+The following definitions are in Library\href{../src/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
+
+\begin{Coqsrc}
+Section A_given.
+  Variables (A : Type)  (lt: relation A).
+
+Local Infix "<" := lt.
+Local Infix "<=" := (clos_refl _ lt).
+
+Definition Least {sto : StrictOrder lt} (x : A):=
+  forall y,  x <= y.
+
+Definition Successor {sto : StrictOrder lt} (y x : A):=
+  x < y /\ (forall z,  x < z ->  z <  y -> False).
+
+Definition Limit {sto : StrictOrder lt}  (x:A)  :=
+  (exists w:A,  w < x) /\
+  (forall y:A, y < x -> exists z:A, y < z /\ z < x).
+\end{Coqsrc}
+
+Using these definitions , we can prove the following lemmas:
+
+\vspace{4pt}
+
+\noindent\emph{From Module~\href{../src/html/hydras.Prelude.ON_Omega_plus_omega.html}{Prelude.ON\_Omega\_plus\_omega}}
+
+\begin{Coqsrc}
+ Lemma omega_is_limit : Limit omega.
+
+Lemma limit_is_omega alpha : Limit alpha -> alpha = omega.
+
+\end{Coqsrc}
+
+Regarding successors, let us introduce the following definition:
+
+\begin{Coqsrc}
+Definition succ (alpha : t) :=
+  match alpha with
+    inl n => inl (S n)
+  | inr n => inr (S n)
+  end.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma Successor_succ alpha : Successor (succ alpha) alpha.
+
+Lemma Successor_correct alpha beta : Successor beta alpha <->
+                                     beta = succ alpha.
+\end{Coqsrc}
+
+We can also check whether an ordinal is a successor:
+
+\begin{Coqsrc}
+Definition succb (alpha: t) := match alpha with
+                                   inr (S  _) | inl (S _) => true
+                                 | _ => false
+                                 end.
+
+Lemma succb_correct alpha : succb alpha <-> exists beta,  alpha = succ beta.
+\end{Coqsrc}
+
+Finally the nature of any ordinal (inside this notation system) is decidable:
+
+\begin{Coqsrc}
+Lemma ZLS_dec (alpha : t) :
+  {alpha = 0} +
+  {Limit alpha} +
+  {beta : t | Successor alpha beta}.
+\end{Coqsrc}
 
@@
 It is sometimes useful to compare two ordinal notations with respect to expressive power
-(the segment of ordinals one they represent).
+(the segment of ordinals  they represent).
 
 The following class specifies a strict inclusion of segments.
-The notation \texttt{OA} describes a segment $[0,alpha($, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$).
-We require that the comparison functions of the two notation systems are compatible.
+The notation \texttt{OA} describes a segment $[0,alpha($, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented by any term of type $A$).
+We require also  that the comparison functions of the two notation systems are compatible.
 
@@
 
-\begin{todo}
-pointers to examples 
-\end{todo}
-
-In some cases we want to show that two notation systems describe the same segment (for instance $[0,3+\omega)$ and $[0,\omega($).
-For this purpose, one may prove that the two notation systems are order-isomorphic.
-
-\begin{Coqsrc}
-Class  ON_Isomorphic 
-       `(OA : @ON A ltA compareA)
-       `(OB : @ON B ltB  compareB)
-       (f : A -> B)
-       (g : B -> A):=
-  {
-  iso_compare: forall x y : A,  compareB (f x) (f y) =
-                                 compareA x y;
-  iso_inv1 : forall a, g (f a)= a;
-  iso_inv2 : forall b, f (g b) = b}.
-\end{Coqsrc}
-
-Finally, it may be interesting to compare an ordinal notation with the more theoretical model from Schütte (well, at least with our formalization of thhat model).
+For instance, we prove that \texttt{Omega} is a sub-notation of
+\texttt{Omega\_plus\_Omega} (with $\omega$ as the first ``new'' ordinal, and \texttt{fin} as the injection).
+
+\begin{Coqsrc}
+Instance Incl : SubON Omega Omega_plus_Omega omega fin.
+\end{Coqsrc}
+
+\subsection{Comparing an ordinal notation with Schütte's model}
+
+Finally, it may be interesting to compare an ordinal notation with the more theoretical model from Schütte (well, at least with our formalization of that model).
 This would be a relative proof of correctenss of our ordinal  notations.
@@
 
-For instance, the following theorem tells that \texttt{Epsilon0}, our otation system for the segment $[0,\epsilon0)$ is correct with respect to the countqable ordinal $\epsilon_0$
+For instance, the following theorem tells that \texttt{Epsilon0}, our notation system for the segment $[0,\epsilon0)$ is correct with respect to the countqable ordinal $\epsilon_0$
 (see chapter~\ref{chap:schutte} for more details).
@@
 
-% ICI
-
-\begin{todo}
-Present instances of Ordinal notation.  
-\end{todo}
-
-It is usual to classify ordinal numbers in three pairwise disjoint categories: 
-\begin{itemize}
-\item The category containing only the least ordinal (called $0$),
-\item The \emph{successors}: $y$ is a successor of $x$ if $x<y$ and there is no $z$ such that $x<z<y$. 
-\item \emph{Limit}: $x\not=0$ is a limit if $x$ is the least upper bound of the set $\{y :\mathbb{O} | y < x\}$.
-\end{itemize}
-
-These notions are defined (for any strict order) in the module 
-\href{../src/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
-
-\begin{Coqsrc}
-Variable lt: relation A.
-
- Local Infix "<" := lt.
- Local Infix "<=" := (clos_refl _ lt).
-
- Definition Least  {sto : StrictOrder lt} (x : A):= forall y,  clos_refl A lt x y.
-
-Definition Successor {sto : StrictOrder lt} (y x : A):=
-  x < y /\ (forall z,  x < z ->  z <  y -> False).
-
-Definition Limit {sto : StrictOrder lt}  (x:A)  :=
-  (exists w:A,  w < x) /\
-  (forall y:A, y < x -> exists z:A, y < z /\  z < x).
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
-Prove in \coq{} that the three predicates above are pairwise exclusive.  
-\end{exercise}
-
-\begin{todo}
-Present the method ZeroLimitSucc 
-\end{todo}
-
-\subsection{Preliminary Definitions}
-
-The type class \texttt{StrictOrder} is already defined in Standard Library
-(module.
-
-\begin{itemize}
-\item Let  $\alpha$ be an ordinal; we say that  $\alpha$ is a \emph{successor} if there exists some ordinal  $\beta$ such that 
-$\alpha$ is is the least ordinal strictly greater than  $\beta$.
-
-\item We say that an ordinal $\lambda$ is a \emph{limit ordinal} is $\lambda$  is the least upper bound of a stricly increasing sequence of ordinals.
-The meta-variable $\lambda$ will be used for denoting  limit ordinals.
-
-\item \index{Maths!Transfinite induction}
-An ordinal is either $0$, a limit ordinal or a successor ordinal. This case analysis, as well as \emph{transfinite} ({i.e.}, well-founded) induction is used in many proofs about ordinal numbers.
-
-\item Let $\alpha$ be any ordinal, the \emph{segment} $\mathbb{O}_\alpha$ is the
-interval $[0,\alpha)\,=\,\{\beta|0\leq\beta<\alpha\}$.
-
-\item The set  of finite ordinals is a segment isomorphic to the set of natural numbers.
- The first infinite ordinal is the limit ordinal $\omega$.
-
-\item The operations $+$, $\times$ and exponentiation on $\mathbb{N}$ are extended on ordinals numbers.
-Note that these  extensions are not commutative any more.
-For instance $\omega = 1 + \omega \not= \omega + 1$
-and $\omega = 2 \times \omega \not= \omega \times 2$.
-
-\item The ordinal $\epsilon_0$ is the least solution of the equation
- \(\alpha=\omega^{\alpha}\).
-\end{itemize}
-
-Please note that the set of countable ordinals is not countable.
-
-Without ambiguity we will often denote the segment $\mathbb{O}_\alpha$ by just $\alpha$.
-
-\section{Notation systems for some small ordinals}
-
-The following enumeration  shows  popular notation systems for several ordinals (shown in increasing order with respect to inclusion), and their possible representation in \coq{}.
-Please keep in mind that a notation system for a given ordinal $\alpha$ represents 
-ordinals \emph{strictly below} $\alpha$.
-
-\subsection{Finite ordinals}
+\index{Projects}
+
+\begin{project}
+  When you have read Chapter~\ref{chap:schutte}, prove that the sum of two ordinal notations \texttt{ON\_plus} implements the addition of ordinals.
+\end{project}
+
+\section{A notation for finite ordinals}
+
 Let $n$ be some natural number. The segment associated with $n$ is the interval 
-$)0,n[\,=\,\{0,1,\dots,n-1\}$.
+$[0,n)\,=\,\{0,1,\dots,n-1\}$.
 
@@
 
-For more information about this pattern, please lokk at the numerous mailing lists and 
+For more information about this pattern, please look at the numerous mailing lists and 
 FAQs on \coq{}).
@@
 
-\subsection{The first infinite ordinal : \texorpdfstring{$\omega$}{omega}}
-
-Mathematically speaking, the ordinal $\omega$ is (set-theoretically) considered as the well-ordered set of
-natural numbers. 
-Every beginner in \coq{} must have seen examples using the type \texttt{nat} of natural numbers. \coq's standard 
-library already contains a lot of functions and lemmas on Peano arithmetic.
-
-In order to build a notation system for $\omega$, we have just to look for lemmas about strict orders and well-foundedness in type \texttt{nat}.
-
-\begin{Coqsrc}
-Require Import Arith Compare_dec Lia Simple_LexProd Ordinal_generic.
-Import Relations.
-
-Search (@StrictOrder nat).
-Search @Nat.compare.
-Search (@well_founded nat).
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Instance Omega : OrdinalNotation Nat.lt_strorder Nat.compare.
-Proof.
- split.
- - apply Nat.compare_spec.
- - apply lt_wf.
-Qed.
-\end{Coqsrc}
-
-We can also prove that any finite ordinal $i$ can be considered as a sub-segment of $\omega$.
-
-\begin{Coqsrc}
-Global Instance FinOrd_Omega (i:nat) :
-  SubSegment (FinOrd i) Omega i 
-             (fun alpha =>  proj1_sig alpha).
-\end{Coqsrc}
-
-\subsubsection{Finite ordinals as members of \texorpdfstring{$\omega+\omega$}{omega+omega}}
-
-It is easy do define a coercion from type \texttt{nat} into our type \texttt{t}.
-
-\begin{Coqsrc}
-Definition fin (n:nat) : t := inl n.
-Coercion fin : nat >-> t.
-\end{Coqsrc}
-
-The range of \texttt{fin} is called the set of \emph{finite} ordinals.
-
-\index{Coq!Techniques!Coercions} 
-\begin{remark}
-Beware of coercions and notation scopes!
-
-Let us consider the following goal:
-
-\begin{Coqsrc}
- Goal (6 < 8).
- auto with arith.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-1 subgoal (ID 9)
-
-  ============================
-  6 < 8
-\end{Coqanswer}
-
-Please keep in mind that the current notation scope interprets the infix \texttt{``<''} as the predicate \texttt{Omega\_plus\_omega.lt} and not \texttt{Nat.lt}.
-More,  the coercion mechanism converts the terms \texttt{6:nat} [resp.
-\texttt{8:nat} ]
-into \texttt{inl 6} [resp. \texttt{inl 8}].  So, the initial goal is correctly interpreted by \coq{}, but not as an inequality between two natural numbers.
-
-Anyway, the initial goal is provable, using \texttt{le\_AsB} first constructor.
-
-\begin{Coqsrc}
-  constructor; auto with arith.
-Qed.
-\end{Coqsrc}
-
-\end{remark}
-
-\subsubsection{Definition of \texorpdfstring{$\omega$}{omega}}
-
-We define $\omega$ as the least ordinal which is larger than any finite ordinal.
-
-\begin{Coqsrc}
- Notation  "'omega'"  := (inr 0) : oo_scope. 
-\end{Coqsrc}
+\section{Isomorphism of ordinal notations}
+
+In some cases we want to show that two notation systems describe the same segment (for instance $[0,3+\omega)$ and $[0,\omega($).
+For this purpose, one may prove that the two notation systems are order-isomorphic.
+
+\begin{Coqsrc}
+Class  ON_Isomorphic 
+       `(OA : @ON A ltA compareA)
+       `(OB : @ON B ltB  compareB)
+       (f : A -> B)
+       (g : B -> A):=
+  {
+  iso_compare: forall x y : A,  compareB (f x) (f y) =
+                                 compareA x y;
+  iso_inv1 : forall a, g (f a)= a;
+  iso_inv2 : forall b, f (g b) = b}.
+\end{Coqsrc}
+
+\index{Exercises}
 
 \begin{exercise}
-Show that, for any $\alpha:t$, $\alpha$ is finite iff $\alpha<\omega$.
+Let $i$ be some natural number. Prove that the notation systems 
+\texttt{Omega} and \texttt{ON\_plus (OrdFin $i$) Omega} are isomorphic.
+
+{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$ we will prove in larger notation systems, as well as in Schütte's model}
 \end{exercise}
 
-\subsubsection{Building an instance of OrdinalNotation}
-
-The class \texttt{OrdinalNotation} is parameterized by the strictorder $<$ and a correct comparison function.
-
-First, we prove that \texttt{lt} is an instance of \texttt{StrictOrder}.
-
-\begin{Coqsrc}
-Instance lt_strorder : StrictOrder lt.
-Qed.
-\end{Coqsrc}
-
-Then, we define a comparison function by pattern matching over the sum type $t$.
-
-\begin{Coqsrc}
-Definition compare (alpha beta: t) : comparison :=
-   match alpha, beta with
-     inl _, inr _ => Lt
-   | inl n, inl p | inr n, inr p => Nat.compare n p
-   | inr _, inl _ => Gt
-  end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Compute compare 7 omega.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-= Lt : comparison
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Compute compare  (inr 2) 42.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-= Gt : comparison
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Compute compare (inr 0) omega.  
-\end{Coqsrc}
-
- \begin{Coqanswer}
-= Eq : comparison
-\end{Coqanswer}
-
-Correctness of the function \texttt{compare} is asserted by the following lemma:
-
-\begin{Coqsrc}
-Lemma compare_correct alpha beta :
-    CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
-              (compare alpha beta).
-\end{Coqsrc}
-
-\begin{todo}
-  Well foundedness; instance of OrdinalNotation.
-\end{todo}
-
-%%%% ICI %%%%
-
-\subsubsection{Limit and successor ordinals}
-
-The notions of limit and successor are defined in the module 
-\href{../src/html/hydras.Prelude.Limits_and_Successors.html}{Prelude.Limits\_and\_Successors} for any instance of
-\texttt{StrictOrder}.
-
-\index{Maths!Limit ordinals}
-
-Let $<$ be a strict order on a type $A$ .
-We say $x:A$ is an  $\omega$-\emph{limit} of  a sequence $s(i), i\in\mathbb{N}$, if
- if $x$ is [strictly] larger than all the $s(i)$, and for any
-$y<x$, $y$ is dominated by at least an element of  the sequence $s$.
-
-\vspace{4pt}
-
-\noindent\emph{From Module~\href{../src/html/hydras.Prelude.Limits_and_Successors.html}{Prelude.Limits\_and\_Successors}}
-
-\begin{Coqsrc}
-Definition  Omega_limit
-            {A:Type}{lt : relation A}
-           {sto : StrictOrder lt} (s: nat -> A) (x:A)  :=
-  (forall i: nat, lt (s i) x) /\
-  (forall y, lt y  x -> exists i:nat, lt y (s i)).
-
-Definition Limit   {A:Type}{lt : relation A}
-           {sto : StrictOrder lt}  (x:A)  :=
-  exists s : nat -> A, Omega_limit s x.
- \end{Coqsrc}
-
-Let $x$ and $y$ be two terms of type $A$, we say that \emph{$y$ is a successor of $x$}
-if $x<y$ and there is no $z$ strictly greater than $x$ and  strictly less than $y$.
-
-\begin{Coqsrc}
-Definition Successor `{lt : relation A}
-           {sto : StrictOrder lt} (y x : A):=
-  lt x y /\ (forall z,  lt x z -> lt z y -> False).
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
-Prove that $\omega$ is the unique   limit ordinal within the notation system for $\omega+\omega$
-\emph{(in this precise notation system! In larger segments of 
-ordinals, there will be \emph{a lot} of limit ordinals}).
-\end{exercise}
-
-\subsection{Computing in \texorpdfstring{$\omega+\omega$}{omega+omega}}
-
-The simple data type we use for representing the ordinals less than $\omega+\omega$ allows us to compute successors, and to decide properties like \texttt{Successor} or \texttt{Limit}.
-
-\subsubsection{Computing successors}
-
-The following function computes the successor of any ordinal less than $\omega+\omega$, by a simple pattern matching.
-
-\begin{Coqsrc}
-Definition succ (alpha : t) := match alpha with
-                                 inl i => inl (S i)
-                               | inr i => inr (S i)
-                               end.
-
-Compute succ omega.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = inr 1
-     : nat + nat
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Lemma lt_succ alpha : alpha < succ alpha.
-Proof.
- destruct alpha; constructor; auto with arith. 
-Qed.
-\end{Coqsrc}
-
-The correctness  of the \texttt{succ} function is asserted by the following lemma.
-
-\begin{Coqsrc}
-Lemma Successor_correct alpha beta : Successor beta alpha <-> beta = succ alpha.
-\end{Coqsrc}
-
-Being a successor is decidable, thanks to the following boolean function.
-
-\begin{Coqsrc}
-Definition succb (alpha: t) := match alpha with
-                                   inr (S  _) | inl (S _) => true
-                                 | _ => false
-                                 end.
-
-Lemma succb_correct alpha : succb alpha <-> exists beta,  alpha = succ beta.
-\end{Coqsrc}
-
-\subsubsection{Deciding whether a given ordinal is a limit}
-
-In the same way as successors, one can decide by pattern matching wether a given ordinal (dont' forget less than $\omega+\omega$ !) is a limit.
-
-\begin{Coqsrc}
-Definition limitb (alpha : t) := match alpha with
-                                     (inr 0) => true
-                                   | _ => false
-                                   end.
-
-Lemma limitb_correct alpha  : limitb alpha <-> 
-                              exists s: nat -> t, Omega_limit s alpha.
-\end{Coqsrc}
-
-\begin{todo}
-trichotomy zero/succ/limit
-boolean functions limitb, succb
-
-\end{todo}
+%%%% ICI ICI
+
 \begin{itemize}
@@
 \begin{todo}
-The following section has moved from chaper on hydras.
-Adapt the redaction to the $\omega^2$ ordinl segment.
+The following section has moved from chapter on hydras.
+Adapt the redaction to the $\omega^2$ ordinal segment.
 \end{todo}

### 2020-07-28 latex/chapter-hydras.tex
redaction introduction to ordinals

@@
 
+\subsubsection{Lexicographic order on \texorpdfstring{nat*nat}{nat*nat}}
+\label{omega2-case}
+
+We  prove now that even the type \texttt{nat * nat}, provided with the lexicographic product of \texttt{(nat,<)} by itself is too simple for proving the termination of all hydra battles.
+This impossibility result will prevent us from considering measures like the following one:
+
+  \begin{Coqsrc}
+  Let m h = (height h, hsize h).
+  \end{Coqsrc}
+
+The proof we are going to develop has exactly the same structure as in Section~\ref{sect:omega:case}
+
+previous one. Nevertheless, the proof of technical  lemmas is a little more complex, due to 
+ the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
+Consider for instance that there exists an infinite number of pairs between
+$(1,0)$ and $(2,0)$.
+
+\begin{remark}
+  The order structure we consider in this section is also known as the ordinal
+  $\omega^2$. We identify any pair $(i,j)\in \mathbb{N}\times\mathbb{N}$ with the ordinal $\omega\times i + j$. Thus the three kinds of ordinals in $\omega^2$
+  are represented as follows:
+  \begin{description}
+  \item[null ordinal] : the pair $(0,0)$
+  \item[successor ordinal] : any pair $(i,j)$ where $j>0$
+  \item[limit ordinal] : any pair $(i,0)$ where $i>0$
+  \end{description}
+\end{remark}
+
+The detailed  proof script is in the file \url{../src/html/hydras.Hydra.Omega2_Small.html}.
+
+\subsubsection{Preliminaries}
+Let us assume there is a variant from \texttt{Hydra} into \texttt{nat*nat} (with the
+   lexicographic ordering)  for proving the   termination of all hydra battles.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
+
+\begin{Coqsrc}
+  Section Impossibility_Proof.
+
+  Let t := (nat * nat)%type.
+
+  (** non-dependent lexicographic strict ordering on nat*nat *)
+
+  Let lt2 : relation t := lexico Peano.lt Peano.lt.
+
+  Infix "<" := lt2.
+
+  (** reflexive closure of lt2 *)
+  Let le2 := clos_refl _ lt2.
+  Infix "<=" := le2.
+
+  Variable m : Hydra -> t.
+
+  Context (Hvar : Hvariant lt2_wf free m).
+\end{Coqsrc}
+
+Let us follow the same pattern as in Sect.~\ref{omega-case}.
+First, we define an injection from type \texttt{t} into \texttt{Hydra}.
+We associate with any pair $(i,j)$ the hydra with $i$ branches of length $2$ and
+$j$ branches of length $1$.
+
+%% revenir ici
+
+\vspace{4pt}
+\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
+
+\begin{Coqsrc}
+  Let iota (p: t) := 
+    node (hcons_mult (hyd1 head) (fst p)
+                     (hcons_mult head (snd p) hnil)).
+\end{Coqsrc}
+
+For instance, Figure~\vref{fig:essai2} shows the hydra associated to the pair
+$(3,5)$.
+
+\begin{figure}[htb]
+\centering
+\begin{tikzpicture}[very thick, scale=0.4]
+\node (foot) at (6,0) {$\bullet$};
+\node (N1) at (1,3) {$\bullet$};
+\node (N2) at (3,3) {$\bullet$};
+\node (N3) at (5,3) {$\bullet$};
+\node (N4) at (8,3) {$\Smiley[2][green]$};
+\node (N5) at (11,3) {$\Smiley[2][green]$};
+\node (N6) at (14,3) {$\Smiley[2][green]$};
+\node (N7) at (17,3){$\Smiley[2][green]$};
+\node (N8) at (20,3){$\Smiley[2][green]$};
+\node  (N9) at (0,5) {$\Smiley[2][green]$};
+\node (N10) at (2,5) {$\Smiley[2][green]$};
+\node (N11) at (4,5) {$\Smiley[2][green]$};
+\draw (foot) to [bend left=10] (N1);
+\draw (foot) -- (N2);
+\draw (foot) -- (N3);
+\draw (foot) -- (N4);
+\draw (foot) -- (N5);
+\draw (foot) -- (N6);
+\draw (foot) to [bend right=10] (N7);
+\draw (foot) to [bend right=15] (N8);
+\draw (N1) to [bend left=10] (N9);
+\draw (N2) -- (N10);
+\draw (N3) -- (N11);
+\end{tikzpicture}
+\caption{\label{fig:essai2}
+The hydra $\iota(3,5)$}
+\end{figure}
+
+Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).
+
+\begin{figure}[htb]
+\centering
+\begin{tikzpicture}[very thick, scale=0.5]
+\node (foot) at (2,0) {$\bullet$};
+\node (N1) at (2,2) {$\bullet$};
+\node (N2) at (3,4) {$\Smiley[2][green]$};
+\node (N3) at (1,4) {$\Smiley[2][green]$};
+\draw (foot) -- (N1);
+\draw (N1) to [bend right =15] (N2);
+\draw (N1) to  [bend left=15](N3);
+\end{tikzpicture}
+\caption{\label{fig:h-omega2-small}}
+ The hydra \texttt{big\_h}.
+\end{figure}
+\begin{Coqsrc}
+   Let big_h := hyd1 (hyd2 head head).  
+ \end{Coqsrc}
+
+ In a second step, we build a ``smaller'' hydra.
+
+\begin{Coqsrc}
+   Let small_h := iota (m big_h).
+\end{Coqsrc}
+
+Like in Sect.~\ref{omega-case}, we prove the double inequality \texttt{m big\_h <= m small\_h < m big\_h}, which is impossible.
+
+\subsubsection{Proof of the inequality \texttt{m small\_h < m big\_h}}
+
+For proving the inequality  \texttt{m\_lt: m small\_h < m big\_h}, it suffices to
+build a battle transforming \texttt{big\_h} into \texttt{small\_h}.
+
+First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h}
+into the pair $(i,j)$.
+If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
+If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the folowing result.
+
+\begin{Coqsrc}
+ Lemma big_to_small: big_h -+-> small_h.
+\end{Coqsrc}
+
+Since $m$ is a variant, we infer the following inequality:
+
+\begin{Coqsrc}
+Corollary m_lt : m small_h < m big_h.
+\end{Coqsrc}
+
+\subsubsection{Proof of the inequality \texttt{m big\_h <= m small\_h} }
+
+The proof of the inequality \texttt{m big\_h <= m small\_h} is quite more complex than in Sect~\ref{omega-case}.  If we consider some pair $(i,j)$, where $i>0$, there exists an infinite number of
+pairs stricly less than $(i,j)$, and there exists an infinite number of battles that start from
+$\iota(i,j)$. In effect, at any configuration $\iota(k,0)$, the hydra can freely chose any replication number. Intuitively, the measure of such a hydra must be large enough for taking into account
+all the possible battles issued from that hydra.
+Let us give more technical details.
+
+\begin{itemize}
+\item The proof of the lemma \texttt{m\_ge : forall p : t,   p <= m (iota p)} uses well-founded induction on $p$, and not structural induction on natural numbers
+
+\item For any pair $p$, we have to distinguish between three cases, according to the value of $p$'s components.
+  \begin{itemize}
+  \item $p=(0,0)$
+  \item $p=(i,0)$, where $i>0$\,: $p$ corresponds to a limit ordinal
+  \item $p=(i,j)$, where $j>0$\,: $p$ is the successor of $(i,j-1)$.
+  \end{itemize}
+\end{itemize}
+
+Before starting the proof, we have to express the notion of \emph{limit} in terms of
+least upper bounds, through the following logical equivalence.
+
+\vspace{4pt}
+\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#limit_is_lub}{Hydra.Omega2\_Small.v}}.
+
+\begin{Coqsrc}
+Lemma limit_is_lub : forall i p, 
+   (forall j, (i,j) < p) <-> (S i, 0) <= p.  
+\end{Coqsrc}
+
+Let us define the notion of elementary ``step'' of decreasing sequences in
+\texttt{t}
+
+\begin{Coqsrc}
+Inductive step : t -> t -> Prop :=
+| succ_step : forall i j,  step (i, S j) (i, j)
+| limit_step : forall i j, step (S i, 0) (i, j).
+\end{Coqsrc}
+
+The following lemma establishes a correspondance between the relation
+\texttt{step} and hydra battles.
+
+\begin{Coqsrc}
+Lemma step_to_battle : forall p q, step p q -> iota p -+-> iota q.
+\end{Coqsrc}
+
+\index{Maths!Transfinite induction}
+
+Thus, starting from any inequality $q < p$ on type \texttt{t}, we can build 
+by transfinite induction over \texttt{p} a battle 
+that transforms the hydra $\iota(p)$ into $\iota(q)$.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}
+
+\begin{Coqsrc}
+Lemma m_ge : forall p : t,   p <= m (iota p).
+Proof.
+  intro p ; pattern p;
+    apply  well_founded_induction with 
+               (R := lt2) (1:= wf_lexico lt_wf lt_wf);
+     intros (i,j) IHij (* rest of proof skipped *)
+\end{Coqsrc}
+
+\begin{Coqanswer}
+  i, j : nat
+  IHij : forall y : t, y < (i, j) -> y <= m (iota y)
+  ============================
+  (i, j) <= m (iota (i, j)) 
+\end{Coqanswer}
+
+Then we have  three cases to consider, according to the values of $i$ and $j$.
+\begin{itemize}
+\item If $p=(0,0)$ then obviously, $\iota(p)\geq p = (0,0)$
+\item If  $p=(i+1,0)$ for some $i\in\mathbb{N}$, we
+ remark  that $p$ is strictly greater than any pair $ (i, j)$, where $j$ 
+is any natural number.
+
+Applying the battle rules, for any $j$, we have $\iota(i+1,j)  {\round} \iota(i, j) $, thus $m(\iota(p)) > m(\iota(i,j)$ since  $m$ is assumed to be a variant.
+
+Applying the induction hypothesis, we get the inequality
+ $ m(\iota(i,j)) \geq (i,j)$ for any $j$.
+
+Thus, $m(\iota(p)) > (i,j)$ for any $j$.
+Applying the lemma \texttt{limit\_is\_lub}, we get  the inequality
+$\iota(i+1,0)\geq (i+1,0)$
+
+\item If $p=(i,j+1)$ with $j\in\mathbb{N}$, we have  $\iota(p)  {\round} \iota(i, j) $,
+hence $m(\iota(p))> m(\iota(i,j)) \geq (i,j)$, thus $m(\iota(p))\geq (i,j+1)=p$
+
+\end{itemize}
+
+\subsubsection{End of the proof}
+
+Since $<$ is a strict order (irreflexive  and transitive) on \texttt{nat*nat}, we can conclude that there is no
+variant for termination on the lexicographic square of $(\mathbb{N},<)$.
+
+\begin{enumerate}
+\item From \texttt{m\_lt}, we infer the strict inequality 
+\texttt{m small\_h < m big\_h}
+\item from \texttt{m\_ge}, we get \texttt{m big\_h <= m (iota (m big\_h)) = m small\_h} 
+\end{enumerate}
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
+
+\begin{Coqsrc}
+Theorem Impossible : False.
+Proof.
+  destruct (StrictOrder_Irreflexive  (m big_h)).
+  apply le2_lt2_trans with (m small_h).
+  -  unfold small_h;  apply m_ge.
+  -  apply m_lt. 
+Qed.
+
+End Impossibility_Proof.
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+Prove that there exists no variant $m$ from \texttt{Hydra} into \texttt{nat*nat} for proving
+    the  termination of all \emph{standard} battles.
+\end{exercise}
+
+\begin{remark}
+In Chapter~\ref{ks-chapter}, we will prove a much more general  theorem
+than the two previous results about $\mathbb{N}$ and $\mathbb{N}\times\mathbb{N}$. The proof of that general  result will share the same structure, but will 
+require a lot of technical results. 
+\end{remark}
+
+\begin{exercise}
+
+\label{sec:orgheadline63}
+Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.
+
+\begin{itemize}
+\item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
+  lexicographic ordering
+\item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.
+
+For instance , the sequence $\langle 4,4,2 \rangle$ is strictly greater than 
+$\langle 4,3,3,3,3,3,3,2,2,2 \rangle$.
+\end{itemize}
+
+\end{exercise}
+
 \subsubsection{Conclusion/Teaser}

### 2020-07-28 latex/chapter-ordinals.tex
redaction introduction to ordinals

@@
 \end{itemize}
-
-\begin{todo}
-The following section has moved from chapter on hydras.
-Adapt the redaction to the $\omega^2$ ordinal segment.
-\end{todo}
-
-\subsubsection{Lexicographic order on \texorpdfstring{nat*nat}{nat*nat}}
-\label{omega2-case}
-
-We  prove now that even the type \texttt{nat * nat}, provided with the lexicographic product of \texttt{(nat,<)} by itself is too simple for proving the termination of all hydra battles.
-This impossibility result will prevent us from considering measures like the following one:
-
-  \begin{Coqsrc}
-  Let m h = (height h, hsize h).
-  \end{Coqsrc}
-
-The proof we are going to develop has exactly the same structure as in Section~\ref{sect:omega:case}
-
-previous one. Nevertheless, the proof of technical  lemmas is a little more complex, due to 
- the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
-Consider for instance that there exists an infinite number of pairs between
-$(1,0)$ and $(2,0)$.
-
-\begin{remark}
-  The order structure we consider in this section is also known as the ordinal
-  $\omega^2$. We identify any pair $(i,j)\in \mathbb{N}\times\mathbb{N}$ with the ordinal $\omega\times i + j$. Thus the three kinds of ordinals in $\omega^2$
-  are represented as follows:
-  \begin{description}
-  \item[null ordinal] : the pair $(0,0)$
-  \item[successor ordinal] : any pair $(i,j)$ where $j>0$
-  \item[limit ordinal] : any pair $(i,0)$ where $i>0$
-  \end{description}
-\end{remark}
-
-The detailed  proof script is in the file \url{../src/html/hydras.Hydra.Omega2_Small.html}.
-
-\subsubsection{Preliminaries}
-Let us assume there is a variant from \texttt{Hydra} into \texttt{nat*nat} (with the
-   lexicographic ordering)  for proving the   termination of all hydra battles.
-
-\vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
-
-\begin{Coqsrc}
-  Section Impossibility_Proof.
-
-  Let t := (nat * nat)%type.
-
-  (** non-dependent lexicographic strict ordering on nat*nat *)
-
-  Let lt2 : relation t := lexico Peano.lt Peano.lt.
-
-  Infix "<" := lt2.
-
-  (** reflexive closure of lt2 *)
-  Let le2 := clos_refl _ lt2.
-  Infix "<=" := le2.
-
-  Variable m : Hydra -> t.
-
-  Context (Hvar : Hvariant lt2_wf free m).
-\end{Coqsrc}
-
-Let us follow the same pattern as in Sect.~\ref{omega-case}.
-First, we define an injection from type \texttt{t} into \texttt{Hydra}.
-We associate with any pair $(i,j)$ the hydra with $i$ branches of length $2$ and
-$j$ branches of length $1$.
-
-%% revenir ici
-
-\vspace{4pt}
-\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
-
-\begin{Coqsrc}
-  Let iota (p: t) := 
-    node (hcons_mult (hyd1 head) (fst p)
-                     (hcons_mult head (snd p) hnil)).
-\end{Coqsrc}
-
-For instance, Figure~\vref{fig:essai2} shows the hydra associated to the pair
-$(3,5)$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-\node (foot) at (6,0) {$\bullet$};
-\node (N1) at (1,3) {$\bullet$};
-\node (N2) at (3,3) {$\bullet$};
-\node (N3) at (5,3) {$\bullet$};
-\node (N4) at (8,3) {$\Smiley[2][green]$};
-\node (N5) at (11,3) {$\Smiley[2][green]$};
-\node (N6) at (14,3) {$\Smiley[2][green]$};
-\node (N7) at (17,3){$\Smiley[2][green]$};
-\node (N8) at (20,3){$\Smiley[2][green]$};
-\node  (N9) at (0,5) {$\Smiley[2][green]$};
-\node (N10) at (2,5) {$\Smiley[2][green]$};
-\node (N11) at (4,5) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N5);
-\draw (foot) -- (N6);
-\draw (foot) to [bend right=10] (N7);
-\draw (foot) to [bend right=15] (N8);
-\draw (N1) to [bend left=10] (N9);
-\draw (N2) -- (N10);
-\draw (N3) -- (N11);
-\end{tikzpicture}
-\caption{\label{fig:essai2}
-The hydra $\iota(3,5)$}
-\end{figure}
-
-Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:h-omega2-small}}
- The hydra \texttt{big\_h}.
-\end{figure}
-\begin{Coqsrc}
-   Let big_h := hyd1 (hyd2 head head).  
- \end{Coqsrc}
-
- In a second step, we build a ``smaller'' hydra.
-
-\begin{Coqsrc}
-   Let small_h := iota (m big_h).
-\end{Coqsrc}
-
-Like in Sect.~\ref{omega-case}, we prove the double inequality \texttt{m big\_h <= m small\_h < m big\_h}, which is impossible.
-
-\subsubsection{Proof of the inequality \texttt{m small\_h < m big\_h}}
-
-For proving the inequality  \texttt{m\_lt: m small\_h < m big\_h}, it suffices to
-build a battle transforming \texttt{big\_h} into \texttt{small\_h}.
-
-First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h}
-into the pair $(i,j)$.
-If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
-If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the folowing result.
-
-\begin{Coqsrc}
- Lemma big_to_small: big_h -+-> small_h.
-\end{Coqsrc}
-
-Since $m$ is a variant, we infer the following inequality:
-
-\begin{Coqsrc}
-Corollary m_lt : m small_h < m big_h.
-\end{Coqsrc}
-
-\subsubsection{Proof of the inequality \texttt{m big\_h <= m small\_h} }
-
-The proof of the inequality \texttt{m big\_h <= m small\_h} is quite more complex than in Sect~\ref{omega-case}.  If we consider some pair $(i,j)$, where $i>0$, there exists an infinite number of
-pairs stricly less than $(i,j)$, and there exists an infinite number of battles that start from
-$\iota(i,j)$. In effect, at any configuration $\iota(k,0)$, the hydra can freely chose any replication number. Intuitively, the measure of such a hydra must be large enough for taking into account
-all the possible battles issued from that hydra.
-Let us give more technical details.
-
-\begin{itemize}
-\item The proof of the lemma \texttt{m\_ge : forall p : t,   p <= m (iota p)} uses well-founded induction on $p$, and not structural induction on natural numbers
-
-\item For any pair $p$, we have to distinguish between three cases, according to the value of $p$'s components.
-  \begin{itemize}
-  \item $p=(0,0)$
-  \item $p=(i,0)$, where $i>0$\,: $p$ corresponds to a limit ordinal
-  \item $p=(i,j)$, where $j>0$\,: $p$ is the successor of $(i,j-1)$.
-  \end{itemize}
-\end{itemize}
-
-Before starting the proof, we have to express the notion of \emph{limit} in terms of
-least upper bounds, through the following logical equivalence.
-
-\vspace{4pt}
-\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#limit_is_lub}{Hydra.Omega2\_Small.v}}.
-
-\begin{Coqsrc}
-Lemma limit_is_lub : forall i p, 
-   (forall j, (i,j) < p) <-> (S i, 0) <= p.  
-\end{Coqsrc}
-
-Let us define the notion of elementary ``step'' of decreasing sequences in
-\texttt{t}
-
-\begin{Coqsrc}
-Inductive step : t -> t -> Prop :=
-| succ_step : forall i j,  step (i, S j) (i, j)
-| limit_step : forall i j, step (S i, 0) (i, j).
-\end{Coqsrc}
-
-The following lemma establishes a correspondance between the relation
-\texttt{step} and hydra battles.
-
-\begin{Coqsrc}
-Lemma step_to_battle : forall p q, step p q -> iota p -+-> iota q.
-\end{Coqsrc}
-
-\index{Maths!Transfinite induction}
-
-Thus, starting from any inequality $q < p$ on type \texttt{t}, we can build 
-by transfinite induction over \texttt{p} a battle 
-that transforms the hydra $\iota(p)$ into $\iota(q)$.
-
-\vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}
-
-\begin{Coqsrc}
-Lemma m_ge : forall p : t,   p <= m (iota p).
-Proof.
-  intro p ; pattern p;
-    apply  well_founded_induction with 
-               (R := lt2) (1:= wf_lexico lt_wf lt_wf);
-     intros (i,j) IHij (* rest of proof skipped *)
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  i, j : nat
-  IHij : forall y : t, y < (i, j) -> y <= m (iota y)
-  ============================
-  (i, j) <= m (iota (i, j)) 
-\end{Coqanswer}
-
-Then we have  three cases to consider, according to the values of $i$ and $j$.
-\begin{itemize}
-\item If $p=(0,0)$ then obviously, $\iota(p)\geq p = (0,0)$
-\item If  $p=(i+1,0)$ for some $i\in\mathbb{N}$, we
- remark  that $p$ is strictly greater than any pair $ (i, j)$, where $j$ 
-is any natural number.
-
-Applying the battle rules, for any $j$, we have $\iota(i+1,j)  {\round} \iota(i, j) $, thus $m(\iota(p)) > m(\iota(i,j)$ since  $m$ is assumed to be a variant.
-
-Applying the induction hypothesis, we get the inequality
- $ m(\iota(i,j)) \geq (i,j)$ for any $j$.
-
-Thus, $m(\iota(p)) > (i,j)$ for any $j$.
-Applying the lemma \texttt{limit\_is\_lub}, we get  the inequality
-$\iota(i+1,0)\geq (i+1,0)$
-
-\item If $p=(i,j+1)$ with $j\in\mathbb{N}$, we have  $\iota(p)  {\round} \iota(i, j) $,
-hence $m(\iota(p))> m(\iota(i,j)) \geq (i,j)$, thus $m(\iota(p))\geq (i,j+1)=p$
-
-\end{itemize}
-
-\subsubsection{End of the proof}
-
-Since $<$ is a strict order (irreflexive  and transitive) on \texttt{nat*nat}, we can conclude that there is no
-variant for termination on the lexicographic square of $(\mathbb{N},<)$.
-
-\begin{enumerate}
-\item From \texttt{m\_lt}, we infer the strict inequality 
-\texttt{m small\_h < m big\_h}
-\item from \texttt{m\_ge}, we get \texttt{m big\_h <= m (iota (m big\_h)) = m small\_h} 
-\end{enumerate}
-
-\vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
-
-\begin{Coqsrc}
-Theorem Impossible : False.
-Proof.
-  destruct (StrictOrder_Irreflexive  (m big_h)).
-  apply le2_lt2_trans with (m small_h).
-  -  unfold small_h;  apply m_ge.
-  -  apply m_lt. 
-Qed.
-
-End Impossibility_Proof.
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into \texttt{nat*nat} for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\begin{remark}
-In Chapter~\ref{ks-chapter}, we will prove a much more general  theorem
-than the two previous results about $\mathbb{N}$ and $\mathbb{N}\times\mathbb{N}$. The proof of that general  result will share the same structure, but will 
-require a lot of technical results. 
-\end{remark}
-
-\begin{exercise}
-
-\label{sec:orgheadline63}
-Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.
-
-\begin{itemize}
-\item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
-  lexicographic ordering
-\item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.
-
-For instance , the sequence $\langle 4,4,2 \rangle$ is strictly greater than 
-$\langle 4,3,3,3,3,3,3,2,2,2 \rangle$.
-\end{itemize}
-
-\end{exercise}

### 2020-07-28 latex/chapter-termination.tex
redaction introduction to ordinals

@@
 
-\subsection{A first-class type for \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-In Sect~\vref{sect:equations}, we chosed to define total functions over the set of ordinals less than $\epsilon_0$, using M. Sozeau's plug-in
-\texttt{coq-equations}\cite{sozeau:hal-01671777}.
-
-To this end, we define a type for representing terms in Cantor normal form,
-as a \coq{} type-class.
+\subsection{An ordinal notation for \texorpdfstring{$\epsilon_0$}{epsilon0}}
+
+The type \texttt{T1} is not a correct ordinal notation, since it contains terms that are not in Cantor normal form.
+In order to define a type whose inhabitants represent just ordinals, we build a type gathering a term of type \texttt{T1} and a proof that this term is in normal form.
 
@@
 \end{Coqanswer}
+
+\index{Projects}
+\begin{project}
+The class \texttt{ON} of ordinal notations has been defined long after this 
+chapter, and is not used yet in the development of the type \texttt{E0}. 
+A better integration of both notions should simplify the development on ordinals in Cantor normal form. This integration is planned for the future versions.
+
+Nevertheless, we built an instance of \texttt{ON}, and proved its correction w.r.t. Schutte's model.
+
+\begin{Coqsrc}
+Instance Epsilon0 : ON Lt compare.  
+\end{Coqsrc}
+
+\emph{From Module~\href{../src/html/hydras.Schutte.Injection_from_T1.html}{Schutte.Injection\_from\_T1}}
+
+\begin{Coqsrc}
+Instance Epsilon0_correct :
+  ON_for epsilon0 Epsilon0  (fun alpha => inject (cnf alpha)).
+\end{Coqsrc}
+
+\end{project}
 

### 2020-07-30 latex/chapter-ordinals.tex
cleanup Schutte/Countable; replace omega tactic with lia

@@
 
-Applying lemmas of the libraries \texttt{Coq.Wellfounded.Inverse\_Image},
+Applying lemmas of the libraries \texttt{Coq.Wellfounded.Inverse\_Image}, \linebreak
  \texttt{Coq.Wellfounded.Inclusion}, and \texttt{Coq.Arith.Wf\_nat}, we prove that our
@@
 the segment $[0,i)$ of $t\;j$.
-
-This notion is defined generically as a class in
- Module~\href{../src/html/hydras.Prelude.Ordinal-generic.html}{Prelude.Ordinal\_generic} see Fig~\vref{figure:subsegment}).
-
-\begin{Coqsrc}
-(** The segment associated with nA is isomorphic to
-    the interval [0,b) *)
-
-Class  Subsegment 
-       `(nA : @OrdinalNotation A ltA stoA compareA)
-       `(nB : @OrdinalNotation B ltB stoB compareB)
-       (b :  B)
-       (iota : A -> B):=
-  {
-  subseg_compare :forall x y : A,  compareB (iota x) (iota y) =
-                                 compareA x y;
-  subseg_incl : forall x, ltB (iota x) b;
-  subseg_onto : forall y, ltB y b  -> exists x:A, iota x = y}.
-\end{Coqsrc}
 
@@
 
-   Global Instance F_incl_ij  : SubSegment  (FinOrd i) (FinOrd j) b iota_ij.
+   Global Instance F_incl_ij  : SubON  (FinOrd i) (FinOrd j) b iota_ij.
   (* ... *)
@@
 
+\index{Projects}
+\begin{project}
+Define in \coq{} the product of two ordinal notations $N_A$ and $N_B$.
+If $A$ [resp. $B$] is the underlying type of $N_A$ [resp. $N_B$], the
+product \texttt{ON\_mult $N_A$ $N_B$} is implemented over the cartesian product $B\times A$ (with the lexicographic ordering).
+
+For instance, the
+elements of the product \texttt{ON\_mult Omega (FinOrd 3)} are ordered as follows.
+\[(0,0),(0,1),(0,2),(0,3),(0,4),\dots,{\color{red}(1,0),} (1,1),(1,2),\dots, {\color{red}(2,0)},(2,1),(2,2),\dots\]
+
+Note that the elements of  \texttt{ON\_mult (FinOrd 3) Omega} are differently ordered (without limit ordinals):
+\[(0,0),(1,0),(2,0),(0,1),(1,1),(2,1),(0,2),(1,2),(2,2),(0,3),\dots\]
+(no limit ordinals).
+
+Prove that \texttt{ON\_mult (FinOrd $i$) Omega} is isomorphic to
+\texttt{Omega}  whilst
+\texttt{Omega}  is a sub-notation of \texttt{ON\_mult Omega (FinOrd $i$)},
+for any strictly positive $i$.
+\end{project}
+
 %%%% ICI ICI
 
-\begin{itemize}
-\item $\omega^n$ (for some  integer  $n\geq 2$) : the type of $n$-uples of natural numbers, with the lexicographic product of $n$ copies of \texttt{Nat.lt}.
-\item  $\omega^\omega$: the set of nonincreasing sequences of natural numbers, lexicographically ordered (also the set of finite multisets of natural numbers).
-\item $\epsilon_0$: The set of terms in Cantor normal form (see Chap.~\ref{chap:T1}).
-\item $\Gamma_0$: The set of terms in Veblen normal form.
-\end{itemize}
+\section*{Other ordinal notations}
+
+\index{Projects}
+\begin{project}
+Let $N_A$ be a notation system for ordinals strictly less than $\alpha$, 
+with the strict order $(A,<_A)$. Please build the notation system
+\texttt{ON\_Expl $N_A$}, on the type of multisets of elements of $A$
+(or, if preferred, the type of non-increasing finite sequences on $A$,
+provided with the lexicographic ordering on lists).
+
+For instance, let us take $N_A=\texttt{Omega}$, and take $\alpha=(4,4,2,1,0)$,
+ $\beta=(4,3,3,3,3,3,2)$, and $\gamma=(5)$. Then $\beta<\alpha<\gamma$.
+
+In contrast the list $(5,6,3,3)$ is not non-increasing (\emph{i.e.} sorted w.r.t. $\geq$), so it is not to be considered.
+
+Note that if the notation $N_A$ implements the ordinal 
+$\alpha$,  the new notation $\omega^{N_A}$ must implement the ordinal $\varphi_0(\alpha)$, a.k.a. $\omega^\alpha$ (see chapter~\ref{chap:schutte})
+
+\end{project}
+
+\begin{remark}
+ The set of ordinal terms in Cantor normal form (see Chap.~\ref{chap:T1}) and 
+in Veblen normal form (see 
+\href{../src/html/hydras.Gamma0.Gamma0.html}{Gamma0.Gamma0}) are shown to be ordinal notation systems, but there is a lot of work to be done in order to unify ad-hoc  definitions and proofs which were written before the definition of the \texttt{ON} type class.
+\end{remark}

### 2020-08-03 latex/chapter-hydras.tex
correction on pdf file

@@
 \index{Coq!Techniques!Type classes}
+\label{sect:hvariant-def}
 
@@
 \begin{exercise}
- Prove that, if there is an instance of \texttt{Hvariant Lt wf\_Lt $B$ $m$}, then there exists no infinite battle in  $B$.
+ Prove that, if there is an instance of (\texttt{Hvariant Lt wf\_Lt $B$ $m$}), then there exists no infinite battle in  $B$.
 \end{exercise}
 
-\subsection{A first theorem of impossibility}
+\subsection{Two small proofs of impossibility}
+\index{Maths!Proofs of impossibility}
+
+When one wants to prove a termination theorem with the help of a variant, 
+one has to consider first a well-founded set $(A,<)$, then a strictly decreasing measure on this set.
+The following two lemmas show that if  the order structure $(A,<)$ is too simple, it is useless to look for a convenient measure, which simply no exists.
+Such kind of result is useful, because it saves you time and effort.
+
+\subsubsection{No variant on Peano.lt}
+\label{omega-case}
 
@@
 
+\begin{figure}[htb]
+\centering
+\begin{tikzpicture}[very thick, scale=0.5]
+\node (foot) at (2,0) {$\bullet$};
+\node (N1) at (2,2) {$\bullet$};
+\node (N2) at (2,4) {$\Smiley[2][green]$};
+\draw (foot) -- (N1);
+\draw (N1) to  (N2);
+\end{tikzpicture}
+\caption{\label{fig:h-omega-omega}}
+ The hydra \texttt{big\_h}.
+\end{figure}
+
  \begin{Coqsrc}
@@
 Let us follow the same pattern as in Sect.~\ref{omega-case}.
-First, we define an injection from type \texttt{t} into \texttt{Hydra}.
-We associate with any pair $(i,j)$ the hydra with $i$ branches of length $2$ and
+First, we define an injection from type \texttt{t} into \texttt{Hydra}, by
+ associating to  any pair $(i,j)$ the hydra with $i$ branches of length $2$ and
 $j$ branches of length $1$.
@@
 
+\index{Exercises}
 \begin{exercise}
@@
 
-For instance , the sequence $\langle 4,4,2 \rangle$ is strictly greater than 
-$\langle 4,3,3,3,3,3,3,2,2,2 \rangle$.
+For instance, the following inequality holds:
+\[\langle 4,3,3,3,3,3,3,2,2,2 \rangle\,<\,\langle 4,4,2 \rangle\]
 \end{itemize}
@@
 
-\subsubsection{Conclusion/Teaser}
+\subsubsection{Conclusion}
 

### 2020-08-03 latex/chapter-ordinals.tex
correction on pdf file

@@
-\chapter{Introduction to ordinal numbers}
+\chapter{Introduction to ordinal numbers and ordinal notations}
 
@@
 &\textcolor{red}{\omega^2},\ldots, \textcolor{red}{\omega^2\times 42},\ldots,\textcolor{red}{\omega^3},\ldots, \textcolor{red}{\omega^4},\omega^4+1,\ldots,\\
-&\textcolor{red}{\omega^\omega},\ldots, \textcolor{red}{\omega^\omega+\omega^7\times 8},\ldots\textcolor{red}{\omega^\omega\times 2},\omega^\omega\times 2+1, \ldots,\\
+&\textcolor{red}{\omega^\omega},\ldots, \textcolor{red}{\omega^\omega+\omega^7\times 8},\ldots,\textcolor{red}{\omega^\omega\times 2},\omega^\omega\times 2+1, \ldots,\\
 &\textcolor{red}{\omega^{\omega^\omega}},\ldots, \textcolor{red}{\omega^{\omega^\omega}+\omega^\omega\times 42+ \omega^{55}+\omega}, \ldots, \textcolor{red}{\omega^{\omega^{\omega+1}}}, \omega^{\omega^{\omega+1}}+1,\dots\\
@@
 
-Let us comment some features of this table.
+Let us comment some features of this table:
 
@@
 For instance, $43$ is the successor of $42$, and $\omega^4+1$ the successor of $\omega^4$.
-\item Ordinals that  follow dots immediately (written in red) are called \emph{limit ordinals}.
+\item Ordinals that  follow dots immediately (displayed in red) are called \emph{limit ordinals}.
 With respect to the order induced by this sequence, any limit ordinal $\alpha$ is the least upper bound of  the set $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
@@
 \begin{align*}
-(i,j) < (i,k) & \;\textbf{if} \; j < k\\
-(0,k) < (1,l) & \;\textbf{for\,any}\;k \;\textbf{and} \; l
+(i,j) < (i,k) & \;\textit{if} \; j < k\\
+(0,k) < (1,l) & \;\textit{for\,any}\;k \;\textit{and} \; l
 \end{align*}
@@
 
-Fortunately, the ordinals we need for  studying hydra battles are much simpler than Schütte's, and can be represented as rather simple data types in \gallina.
+Fortunately, the ordinals we need for  studying hydra battles are much simpler than Schütte's, and can be represented as quite simple data types in \gallina.
 So, we will use \emph{ordinal notations} (also called \emph{[ordinal] notation systems}).
-Let $\alpha$ be some countable ordinal; In \coq{} terms, an ordinal notation for $\alpha$ will be a data type for representing all ordinals below $\alpha$ (excluding $\alpha$) and a bunch of functions for comparing ordinals, doing some arithmetic, inspecting the topology of the order (limits, successors), and proofs of well-foundedness).
-These functions should be proved correct, with respect to Schütte's model.
 
-Our definition of ordinal notation borrows much material from \coq's standard library.
+Let $\alpha$ be some (countable) ordinal; 
+in \coq{} terms, we call ordinal notation for $\alpha$ a structure composed 
+of:
+\begin{itemize}
+\item A data type $A$ for representing all ordinals strictly below $\alpha$,
+\item A well founded order $<$ on $A$, 
+\item A Function for comparing two ordinals, which implies that the reflexive closure of $<$ is a \emph{total order}.
+\end{itemize}
+
+Such a structure can be proved correct relatively to another ordinal notation or
+to Schütte's model.
 
@@
 The library ~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Classes.RelationClasses.html}{%
-Classes.RelationClasses} contains some definitions and facts about binary relations, among them strict orders.
+Coq.Classes.RelationClasses} contains some definitions and facts about binary relations, among them strict orders.
 
@@
   : A + B -> A + B -> Prop :=
-    le_aa : forall x y : A, leA x y -> le_AsB A B leA leB (inl x) (inl y)
+| le_aa : forall x y : A, leA x y -> le_AsB A B leA leB (inl x) (inl y)
 | le_ab : forall (x : A) (y : B), le_AsB A B leA leB (inl x) (inr y)
@@
 \vspace{4pt}
-\noindent\emph{From Module~\href{../src/html/hydras.Prelude.ON_Omega_plus_omega.html}{Prelude.ON\_Omega\_plus\_omega}}
+\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
 
@@
 
+\label{warning:coercions}
 \index{Coq!Techniques!Coercions} 
@@
 
-Using these definitions , we can prove the following lemmas:
+Using these definitions , we can prove the following lemma:
 
@@
 
-\noindent\emph{From Module~\href{../src/html/hydras.Prelude.ON_Omega_plus_omega.html}{Prelude.ON\_Omega\_plus\_omega}}
+\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
 
 \begin{Coqsrc}
- Lemma omega_is_limit : Limit omega.
-
-Lemma limit_is_omega alpha : Limit alpha -> alpha = omega.
-
+Lemma limit_iff (alpha : t) : Limit alpha <-> alpha = omega.
 \end{Coqsrc}
@@
 
+\index{Projects}
+\begin{project}
+Adapt the definition of \texttt{Hvariant} (Sect.~\ref{sect:hvariant-def}) in order to
+have an ordinal notation as argument. Prove that if $O_A$ is a sub-notation of $O_B$, then any variant defined on  $O_A$ can be automatically transformed into 
+a variant on $O_B$.
+\end{project}
+
 \subsection{Comparing an ordinal notation with Schütte's model}
@@
 The following class specifies that a notation \texttt{OA} describes a segment $[0,\alpha)$,
-where $\alpha$ is a countable ordinal à la Schütte.
+where $\alpha$ is a countable ordinal \emph{à la}  Schütte.
 
@@
 
-For instance, the following theorem tells that \texttt{Epsilon0}, our notation system for the segment $[0,\epsilon0)$ is correct with respect to the countqable ordinal $\epsilon_0$
+For instance, the following theorem tells that \texttt{Epsilon0}, our notation system for the segment $[0,\epsilon0)$ is correct with respect to the countable ordinal $\epsilon_0$
 (see chapter~\ref{chap:schutte} for more details).
@@
 \begin{Coqsrc}
-Class  ON_Isomorphic 
+Class  ON_Iso 
        `(OA : @ON A ltA compareA)
@@
 
-\section*{Other ordinal notations}
+\section{Other ordinal notations}
 

### 2020-08-03 latex/chapter-schutte.tex
correction on pdf file

@@
 Our library 
-\href{../src/html/hydras.Schutte.Injection_from_T1.html}%
-{\texttt{Schutte.Injection\_from\_T1.v}} establishes the link between two very different modelizations of ordinal numbers. In other words, it ``validates'' a data structure in terms of
+\href{../src/html/hydras.Schutte.correctness_E0.html}%
+{\texttt{Schutte.Correctness\_E0}} establishes the link between two very different modelizations of ordinal numbers. In other words, it ``validates'' a data structure in terms of
 a classical mathematical discourse considered as a model. 

### 2020-08-03 latex/chapter-termination.tex
correction on pdf file

@@
 \index{Types!T2@T2 (ordinals in Veblen normal form)}
-\emph{Please look also at sub-directory
-href{../src/html/hydras.Gamma0.T2.html}{Gamma0.T2.html}}.
+\emph{Please look also at the library
+\href{../src/html/hydras.Gamma0.T2.html}{Gamma0.T2.html}}.
 
@@
 | gcons : T2 -> T2 -> nat -> T2 -> T2.
-\end{Coqsrc}
-
-Another useful ordinal notation could be used for denoting the ordinals strictly less than $\omega^\omega$, but \texttt{list nat} could be used as well.
-
-\begin{Coqsrc}
-Inductive T0 : Set := 
-  zero : T0
-| cons : nat -> T0 -> T0.
 \end{Coqsrc}
@@
 In a future version, we should replace  the type \texttt{nat} with \texttt{positive} in \texttt{T1}'s 
-definition. But this replacement would take a lot of time \dots.
+definition. But this replacement would take a lot of time \dots{}
 
@@
 \begin{remark}
-\label{warning:coercions}
-The use of coercions like \texttt{fin} allow us to be close to the mathematical tradition where natural numbers are ordinals too.
-Nevertheless, it may happen that a goal like \texttt{3 < 5} could be 
-interpreted as \texttt{(lt (fin 3) (fin 5))},  depending on the current notation scope.  
-When this misinterpretation happens, tactics like \texttt{auto with arith}, \texttt{lia} do not work!
-Thus, it is useful to write \texttt{(3 < 5)\%nat}  an inequality between two natural numbers. 
+Please refer to the remark~\pageref{waening:coercions} about the use of coercions.
+% The use of coercions like \texttt{fin} allow us to be close to the mathematical tradition where natural numbers are ordinals too.
+% Nevertheless, it may happen that a goal like \texttt{3 < 5} could be 
+% interpreted as \texttt{(lt (fin 3) (fin 5))},  depending on the current notation scope.  
+% When this misinterpretation happens, tactics like \texttt{auto with arith}, \texttt{lia} do not work!
+% Thus, it is useful to write \texttt{(3 < 5)\%nat}  an inequality between two natural numbers. 
 \end{remark}
@@
 \begin{Coqsrc}
-Notation omega := (ocons (ocons zero 0 zero) 0 zero).
+Notation omega := (ocons (ocons zero 0 zero) 0 zero): t1_scope.
 \end{Coqsrc}
@@
 For Schütte,  $\omega^\alpha$ is just a notation for $\varphi_0(\alpha)$.
-See also Chapter~\vref{chap:schutte}.
+See also Chapter~\ref{chap:schutte} of this document.
 \end{remark}
@@
 
-The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximed by the sequance of $\omega$-towers.
+The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximed by the sequence of $\omega$-towers.
 
@@
 \begin{Coqsrc}
-Theorem lt_irrefl : forall alpha, ~ lt alpha  alpha.
-
-Theorem lt_trans :
-forall alpha beta: T1,
- lt alpha  beta -> 
- forall gamma, lt beta gamma -> lt alpha gamma.
+Theorem lt_irrefl  alpha :  ~ lt alpha  alpha.
+
+Theorem lt_trans (alpha beta gamma : T1) :
+  lt alpha  beta -> lt beta gamma -> lt alpha gamma.
 
@@
 
-\subsubsection{A Predicate for characterizing normal forms}
+\subsubsection{A predicate for characterizing normal forms}
 \label{sec:orgheadline74}
@@
 
-Definition nf alpha :Prop := nf_b alpha.
+Definition nf alpha: Prop := nf_b alpha.
 \end{Coqsrc}
@@
 
-\emph{From Module~\href{../src/html/hydras.Schutte.Injection_from_T1.html}{Schutte.Injection\_from\_T1}}
+\emph{From Module~\href{../src/html/hydras.Schutte.Schutte.Correctness_E0.html}{Schutte.Schutte.Correctness\_E0}}
 

### 2020-08-06 latex/chapter-ks.tex
correction on pdf file

@@
  an article by J.~Ketonen and R.~Solovay~\cite{KS81}, already  cited in the work
- by L.~Kirby et J.~Paris on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
+ by L.~Kirby et J.~Paris.% on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
  Section $2$ of this article: ''A hierarchy of probably recursive functions'', contains a systematic study of \emph{canonical sequences}, which are closely related to
@@
 
-\paragraph{Remark}
+\begin{remark}
   In~\cite{KS81}, canonical sequences are defined for any ordinal $\alpha <\epsilon_0$,
-by stating that if $\alpha$ is a successor ordinal, the sequence associated with 
-$\alpha$ is simply the constant sequence whose terms are equal to the predecessor of $\alpha$. Likewise, we define the canonical sequence of $0$ as the
-sequence whose all terms are equal to $0$.
-
-This convention allows us to make total the function that maps any ordinal $\alpha$ and natural number $i$ to the $i$-th item of the canonical sequence associated with $\alpha$.
+by stating that if $\alpha$ is a successor ordinal $\beta+1$,  the sequence associated with 
+$\alpha$ is simply the constant sequence whose terms are equal to $\beta$.
+Likewise, we define the canonical sequence of $0$ as the 
+sequence all terms of which are equal to $0$.
+
+This convention allows us to make total the function that maps any ordinal $\alpha$ and natural number $i$ to the ordinal $\canonseq{\alpha}{i}$.
+
+\end{remark}
 
 First, let us recall how canonical sequences are defined in~\cite{KS81}.
-For efficiency's sake, we decided not to implement directly K.\&S's definitions, but to define in \gallina{} simply typed structurally recursive functions which share the abstract properties which are used in the mathematical proofs\footnote{With a small diffrence: the $0$-th term of the canonical sequence is not the same in our development as in~\cite{KS81}.}
+For efficiency's sake, we decided not to implement directly K.\&S's definitions, but to define in \gallina{} simply typed structurally recursive functions which share the abstract properties which are used in the mathematical proofs\footnote{With a small diffrence: the $0$-th term of the canonical sequence is not the same in our development as in~\cite{KS81}.}.
 
@@
 we used an auxiliary function \texttt{canonS} of type \texttt{T1 -> nat  -> T1} such that
-\texttt{canonS  $\alpha$ $i$} is equal to $\canonseq{\alpha}{i+1}$.
+(\texttt{canonS  $\alpha$ $i$}) is equal to $\canonseq{\alpha}{i+1}$.
 
@@
 
-\subsection{Definition}
-
-We shall denote by $\alpha\xrightarrow [s]{}\beta$
-  the proposition 
-``there is a path from $\alpha$ to $\beta$ with $s$ as the sequence of indices
-$s_0,s_1,\dots,s_{l-1}$ ''.
+Let $s$ be the sequence $\langle s_0,s_1,\dots, s_{l-1} \rangle$. We describe the
+existence of such a path with the notation $\alpha\xrightarrow [s]{}\beta$.
 
 For instance, we have $\omega*2 \xrightarrow[2,2,2]{}\omega$, through the 
-sequence of ordinals $\omega\times 2, \omega+2,\omega+1,\omega$.
+path $\omega\times 2, \omega+2,\omega+1,\omega$.
+
+\begin{remark}
 
@@
 
-Indeed, if $\alpha$ is a limit ordinal, the first element of $s$ can be any integer $i$ such that $\beta<\canonseq{\alpha}(i)$, and if $\alpha$ is a successor ordinal,
+Indeed, if $\alpha$ is a limit ordinal, the first element of $s$ can be any integer $i$ such that $\beta<\canonseq{\alpha}{i}$, and if $\alpha$ is a successor ordinal,
 then the sequence $s$ can start with any positive integer.
@@
 $\omega*2 \xrightarrow[{[3,14]}]{} 0$ and
-$\omega*2 \xrightarrow[s] 0$, where $s$ is the sequence containing eight times the number $3$.
+$\omega*2 \xrightarrow[3,3,3,3,3,3,3,3]{} 0$.
+\end{remark}
+
+\subsection{Formal definition}
 
@@
 Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
-(please see also \href{../src/html/hydras.Schutte.Injection_E0.html}{the module \texttt{Ordinals.Schutte.Injection\_E0}}).
+(please see also \href{../src/html/hydras.Schutte.Correctness_E0.html}{the module \texttt{Ordinals.Schutte.Correctness\_E0}}).
 

### 2020-08-06 latex/chapter-termination.tex
correction on pdf file

@@
 
-Another way is encapsulate conditions of the form \texttt{(nf $\alpha$)} in
-often used predicates. For instance, we introduced the restriction of \texttt{lt} to terms in normal form, and provided a handy notation for this restriction.
+One may encapsulate conditions of the form \texttt{(nf $\alpha$)} in
+the most used predicates. For instance, we introduce the restriction of \texttt{lt} to terms in normal form, and provide a handy notation for this restriction.
 
@@
 
-\subsection{An ordinal notation for \texorpdfstring{$\epsilon_0$}{epsilon0}}
+\subsubsection{A sigma type for \texorpdfstring{$\epsilon_0$}{epsilon0}}
 
@@
 In order to define a type whose inhabitants represent just ordinals, we build a type gathering a term of type \texttt{T1} and a proof that this term is in normal form.
+This representation will be used in section~\ref{sect:equations} in order to define recursive functions with the plug-in \texttt{coq-equations}, and also to build an instance of the class \texttt{ON} of ordinal notations.
 
@@
 the term they build is in normal form.
-For instance, let us represent the ordinals $0$ $\omega$ as instances of the class \texttt{E0}.
+For instance, let us represent the ordinals $0$ and $\omega$ as instances of the class \texttt{E0}.
 
@@
 \end{Coqanswer}
-
-\index{Projects}
-\begin{project}
-The class \texttt{ON} of ordinal notations has been defined long after this 
-chapter, and is not used yet in the development of the type \texttt{E0}. 
-A better integration of both notions should simplify the development on ordinals in Cantor normal form. This integration is planned for the future versions.
-
-Nevertheless, we built an instance of \texttt{ON}, and proved its correction w.r.t. Schutte's model.
-
-\begin{Coqsrc}
-Instance Epsilon0 : ON Lt compare.  
-\end{Coqsrc}
-
-\emph{From Module~\href{../src/html/hydras.Schutte.Schutte.Correctness_E0.html}{Schutte.Schutte.Correctness\_E0}}
-
-\begin{Coqsrc}
-Instance Epsilon0_correct :
-  ON_for epsilon0 Epsilon0  (fun alpha => inject (cnf alpha)).
-\end{Coqsrc}
-
-\end{project}
 
@@
 Fixpoint succ (alpha:T1) : T1 :=
-  match alpha with zero => 1
+  match alpha with 
+   | zero => 1
    | ocons zero n _ => ocons zero (S n) zero
@@
 
-The function \texttt{pp: T1 -> ppT1} converts any closed term of type \texttt{T1} into a human-readable expression. For instance, let us convert the term \texttt{alpha\_0} of~\pageref{alpha0-def}.
+The function \texttt{pp: T1 -> ppT1} converts any closed term of type \texttt{T1} into a human-readable expression. For instance, let us convert the term \texttt{alpha\_0}.
 
@@
  \end{remark}
+
+\subsection{An ordinal notation for  \texorpdfstring{$\epsilon_0$}{epsilon0}}
+
+We built an instance of \texttt{ON}, and proved its correction w.r.t. Schutte's model.
+
+\begin{Coqsrc}
+Instance Epsilon0 : ON Lt compare.  
+\end{Coqsrc}
+
+\emph{From Module~\href{../src/html/hydras.Schutte.Schutte.Correctness_E0.html}{Schutte.Schutte.Correctness\_E0}}
+
+\begin{Coqsrc}
+Instance Epsilon0_correct :
+  ON_for epsilon0 Epsilon0 (fun alpha => inject (cnf alpha)).
+\end{Coqsrc}
+
+\index{Projects}
+\begin{project}
+ \emph{This exercise is a continuation of Project~\vref{exo:ON-mult}.}
+Define an ordinal notation \texttt{Omega2} for $\omega^2=\omega\times\omega$ (using
+\texttt{ON\_mult}).
+
+Prove that \texttt{Omega2} is a sub-notation of \texttt{Epsilon0}.
+
+Define on \texttt{Omega2} an addition compatible with the addition on \texttt{Epsilon0}.
+
+\textbf{Hint}. You may use the following definition (in 
+\href{../src/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}).
+
+\begin{Coqsrc}
+Definition SubON_same_op  `{OA : @ON A ltA  compareA}
+       `{OB : @ON B ltB  compareB}
+       {iota : A -> B} 
+       {alpha: B}
+       {_ : SubON OA OB alpha iota}
+       (f : A -> A -> A)
+       (g : B -> B -> B)
+  :=
+  forall x y,  iota (f x y) = g (iota x) (iota y).
+\end{Coqsrc}
+
+\end{project}
+
+\index{Projects}
+\begin{project}
+The class \texttt{ON} of ordinal notations has been defined long after this 
+chapter, and is not used yet in the development of the type \texttt{E0}. 
+A better integration of both notions should simplify the development on ordinals in Cantor normal form. This integration is planned for the future versions.
+
+\end{project}
 
@@
 First, we prove by transfinite induction on $\alpha$ that the restriction of $\oplus$ to the
-interval $[0..\alpha[$ is commutative.
+interval $[0..\alpha)$ is commutative.
 
@@
 strictly less than $\textrm{max}(\alpha,\beta)+1$.
-
     Thus, we have just to apply the lemma \coqsimple{oplus\_comm\_0}.
@@
 \begin{Coqsrc}
-Theorem m_nf : forall h, nf (m h).
+Lemma m_nf : forall h, nf (m h).
 Proof.
@@
 
-Theorem ms_nf : forall s, nf (ms s).
+Lemma ms_nf : forall s, nf (ms s).
 Proof with auto with T1.
@@
 \begin{itemize}
-\item There exists a strictly decreasing variant mapping \texttt{Hydra} into 
-$[0,\epsilon_0[$ for proving the termination of any hydra battle
-\item There exists \emph{no} such variant from \texttt{Hydra} into $[0,\omega^2[$ and
-a fortiori into $[0,\omega[$.
+\item There exists a strictly decreasing variant which maps \texttt{Hydra} into 
+the segment $[0,\epsilon_0)$ for proving the termination of any hydra battle
+\item There exists \emph{no} such variant from \texttt{Hydra} into $[0,\omega^2)$ and
+a fortiori into $[0,\omega)$.
 \end{itemize}

### 2020-08-06 latex/chapter-ks.tex
corrections on pdf file

@@
        | ocons alpha0 n beta  => 
-           hcons_mult (iota alpha0) (S n)
-                                (iotas beta)
+           hcons_mult (iota alpha0) (S n) (iotas beta)
        end.
@@
     m: Hydra -> T1;
-    mu_nf : nf mu;
-    Hvar : Hvariant T1_wf B m;
-    m_bounded : forall h, m h < mu
+    mu_nf: nf mu;
+    Hvar: Hvariant T1_wf B m;
+    m_bounded: forall h, m h < mu
   }.
@@
 $\alpha \xrightarrow[i,j]{}\beta$ for the proposition 
-\texttt{standard\_path $i$ $\alpha$ $j$ $\beta$}.
+(\texttt{standard\_path $i$ $\alpha$ $j$ $\beta$}).
 In~\cite{KS81} the notation is
@@
  \begin{enumerate}
- \item $\alpha \xrightarrow n \canonseq{\alpha}{n}$
+ \item $\alpha \xrightarrow[n]{} \canonseq{\alpha}{n}$
  \item $\canonseq{\alpha}{n} \xrightarrow[1]{} \canonseq{\alpha}{i}$ (by \texttt{Theorem\_2\_4}),
@@
 Lemma Lemma2_6_1 (alpha : T1) :  
-  nf alpha ->
-  forall beta,  beta < alpha  ->
+  nf alpha -> forall beta,  beta < alpha  ->
   {n:nat | const_pathS n alpha beta}.
@@
 \begin{Coqsrc}
-  Compute standard_fun 2 omega 15.
+  Compute standard_gnaw 2 omega 15.
 (*   = zero
@@
 Let $l$ be the length of the path from $\alpha$ to $\gamma$.  
-
-% We use for that purpose a function that computes the first index $i$ where $n\leq i <p$ and $P(i+1)$ is
-% \texttt{false}, assuming that $P(n)=\texttt{true}$ and $P(p)=\texttt{false}$.
-
-% \begin{Coqsrc}
-% first_toggle  (P : nat -> bool) (n p : nat) :
-% (n < p)%nat -> P n = true -> P p = false ->
-% {l : nat | (n <= l)%nat /\ (l < p)%nat /\
-%      (forall i : nat, (n <= i)%nat -> (i <= l)%nat -> P i = true) /\
-%       P (S l) = false}
-% \end{Coqsrc}
-
-% \begin{Coqsrc}
-% Let P (i:nat) := le_b beta (standard_fun (S n) alpha i).
-
-% Remark Rem2 : {t: nat | standard_fun (S n) alpha t < beta}.
-
-% Let t := proj1_sig Rem2.
-%  Remark Rem04 : P 0.
-
-% Remark Rem05 : P t = false.
-
-% Remark Rem06 : (0 <  t)%nat.
-
-%  Let l_def :=  first_toggle P 0 t  Rem06 Rem04 Rem05.
-
-% Let l := proj1_sig Rem07.
-
-% Let gamma := standard_fun (S n) alpha l.
-% \end{Coqsrc}
-
 This step of the proof is illustrated in figure~\vref{fig:belle-preuve-2}.
@@
 \begin{Coqsrc}
-Corollary  LT_to_standard_path 
-      (alpha beta : T1) :
+Corollary  LT_to_standard_path  (alpha beta : T1) :
   beta < alpha ->
@@
 
-In the same way as for free battles, we import a large inequality from 
-the module \url{../src/html/hydras.Hydra.Epsilon0_Needed_Generic.html} .
+In the same way as for free battles, we import a large inequality 
+from 
+the module \href{../src/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}.
 
@@
 
-\paragraph*{Proof:} Let us recall that $\texttt{big\_h} = \iota(\mu)$
+\paragraph*{Sketch of proof:} Let us recall that $\texttt{big\_h} = \iota(\mu)$
  and $\texttt{small\_h} = \iota (m (\texttt{big\_h}))$.
@@
 and \texttt{h} is transformed into \texttt{h'} in the considered round.
-
 Can we still prove the theorems of section~\ref{std-case} with this new definition?

### 2020-08-08 latex/chapter-hydras.tex
corrections on pdf file

@@
 
-  Infix "<" := lt2.
+  Infix "o<" := lt2.
 
@@
   Let le2 := clos_refl _ lt2.
-  Infix "<=" := le2.
+  Infix "o<=" := le2.
 
@@
 
-Like in Sect.~\ref{omega-case}, we prove the double inequality \texttt{m big\_h <= m small\_h < m big\_h}, which is impossible.
-
-\subsubsection{Proof of the inequality \texttt{m small\_h < m big\_h}}
-
-For proving the inequality  \texttt{m\_lt: m small\_h < m big\_h}, it suffices to
+Like in Sect.~\ref{omega-case}, we prove the double inequality \texttt{m big\_h o<= m small\_h o< m big\_h}, which is impossible.
+
+\subsubsection{Proof of the inequality \texttt{m small\_h o< m big\_h}}
+
+For proving the inequality  \texttt{m\_lt: m small\_h o< m big\_h}, it suffices to
 build a battle transforming \texttt{big\_h} into \texttt{small\_h}.
@@
 \begin{Coqsrc}
-Corollary m_lt : m small_h < m big_h.
-\end{Coqsrc}
-
-\subsubsection{Proof of the inequality \texttt{m big\_h <= m small\_h} }
-
-The proof of the inequality \texttt{m big\_h <= m small\_h} is quite more complex than in Sect~\ref{omega-case}.  If we consider some pair $(i,j)$, where $i>0$, there exists an infinite number of
+Corollary m_lt : m small_h o< m big_h.
+\end{Coqsrc}
+
+\subsubsection{Proof of the inequality \texttt{m big\_h o<= m small\_h} }
+
+The proof of the inequality \texttt{m big\_h o<= m small\_h} is quite more complex than in Sect~\ref{omega-case}.  If we consider some pair $(i,j)$, where $i>0$, there exists an infinite number of
 pairs stricly less than $(i,j)$, and there exists an infinite number of battles that start from
@@
 \begin{itemize}
-\item The proof of the lemma \texttt{m\_ge : forall p : t,   p <= m (iota p)} uses well-founded induction on $p$, and not structural induction on natural numbers
+\item The proof of the lemma \texttt{m\_ge : forall p : t,   p o<= m (iota p)} uses well-founded induction on $p$, and not structural induction on natural numbers
 
@@
 Lemma limit_is_lub : forall i p, 
-   (forall j, (i,j) < p) <-> (S i, 0) <= p.  
+   (forall j, (i,j) o< p) <-> (S i, 0) o<= p.  
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Lemma m_ge : forall p : t,   p <= m (iota p).
+Lemma m_ge : forall p : t,   p o<= m (iota p).
 Proof.
@@
   i, j : nat
-  IHij : forall y : t, y < (i, j) -> y <= m (iota y)
+  IHij : forall y : t, y o< (i, j) -> y o<= m (iota y)
   ============================
-  (i, j) <= m (iota (i, j)) 
+  (i, j) o<= m (iota (i, j)) 
 \end{Coqanswer}
@@
 \item From \texttt{m\_lt}, we infer the strict inequality 
-\texttt{m small\_h < m big\_h}
-\item from \texttt{m\_ge}, we get \texttt{m big\_h <= m (iota (m big\_h)) = m small\_h} 
+\texttt{m small\_h o< m big\_h}
+\item from \texttt{m\_ge}, we get \texttt{m big\_h o<= m (iota (m big\_h)) = m small\_h} 
 \end{enumerate}

### 2020-08-08 latex/chapter-ks.tex
corrections on pdf file

@@
-s]\chapter{Inside \texorpdfstring{$\epsilon_0$}{epsilon0}: The Ketonen-Solovay machinery\label{ks-chapter}}
+\chapter[The Ketonen-Solovay machinery]{Strolling inside \texorpdfstring{$\epsilon_0$}{epsilon0}: The Ketonen-Solovay machinery\label{ks-chapter}}
 \label{chap:ketonen}
@@
 rounds of hydra battles. 
+Nevertheless, they have the same global structure as the simple proofs described in
+sections~\vref{omega-case} and \vref{omega2-case}. 
+We invite the reader to compare the three proofs step by step, lemma by lemma.
 
@@
 Lemma canonS_LT i alpha :
-  nf alpha -> alpha <> zero -> canonS alpha i <  alpha.
+  nf alpha -> alpha <> zero -> canonS alpha i o<  alpha.
 \end{Coqsrc}
@@
      limitb lambda  ->
-     forall beta, beta < lambda ->
-                  {i:nat | beta < canonS lambda i}.
+     forall beta, beta o< lambda ->
+                  {i:nat | beta o< canonS lambda i}.
 
@@
 Lemma LT_path_to (alpha beta : T1) :
-  beta < alpha -> {s : list nat | path_to beta s alpha}.
+  beta o< alpha -> {s : list nat | path_to beta s alpha}.
 \end{Coqsrc}
@@
 Lemma path_to_LT beta s alpha :
-  path_to beta s alpha -> nf alpha -> beta < alpha.
+  path_to beta s alpha -> nf alpha -> beta o< alpha.
 \end{Coqsrc}
@@
 Lemma LT_to_battle alpha beta :
-    beta < alpha ->  iota alpha -+-> iota beta.
-\end{Coqsrc}
-
-\section{A first proof of impossibility}
+    beta o< alpha ->  iota alpha -+-> iota beta.
+\end{Coqsrc}
+
+\section{A  proof of impossibility}
 
@@
 proofs described in  sections~\vref{omega-case} and \vref{omega2-case}.
-We advise the reader to compare the three proofs step by step, lemma by lemma.
 
@@
     Hvar: Hvariant T1_wf B m;
-    m_bounded: forall h, m h < mu
+    m_bounded: forall h, m h o< mu
   }.
@@
   Hypothesis m_decrease : forall  i h h',
-        round_n i h h' -> m h' < m h.
+        round_n i h h' -> m h' o< m h.
 \end{Coqsrc}
@@
 the  \texttt{free} and \texttt{standard} kinds of battles. This trick allows to 
-``factorize'' our proofs our proofs of impossibility.
+``factorize'' our proofs  of impossibility.
 \end{remark}
@@
 \begin{Coqsrc}
-Lemma m_ge_0 alpha:  nf alpha -> alpha <= m (iota alpha).
+Lemma m_ge_0 alpha:  nf alpha -> alpha o<= m (iota alpha).
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-    Corollary m_ge_generic : m big_h <= m small_h.
+    Corollary m_ge_generic : m big_h o<= m small_h.
  \end{Coqsrc}
@@
 The (big) rest of the proof is dedicated to prove formally the converse inequality 
-\texttt{m small\_h < m big\_h}.
+\texttt{m small\_h o< m big\_h}.
 
@@
 \begin{Coqsrc}
-Lemma m_ge : m big_h <= m small_h.
+Lemma m_ge : m big_h o<= m small_h.
   Proof.
@@
 
-\item From the hypothesis \texttt{m\_bounded}, we have \texttt{m big\_h < mu}
+\item From the hypothesis \texttt{m\_bounded}, we have \texttt{m big\_h o< mu}
 \item By Lemma \texttt{LT\_to\_battle}, we get a (free) battle from
@@
 \begin{Coqsrc}
-Lemma m_lt : m small_h < m big_h.
+Lemma m_lt : m small_h o< m big_h.
 \end{Coqsrc}
@@
 Lemma LT_to_round_plus alpha beta :
-    beta < alpha ->  iota alpha -+-> iota beta.
+    beta o< alpha ->  iota alpha -+-> iota beta.
 \end{Coqsrc}
@@
 Corollary Cor12 (alpha : T1) :  nf alpha ->
-         forall beta i n, beta  < alpha  ->
+         forall beta i n, beta  o< alpha  ->
                 (i < n)%nat ->
@@
 Corollary Cor12' (alpha : T1) :  nf alpha ->
-         forall beta i n, (beta < alpha)%t1 ->
+         forall beta i n, (beta o< alpha)%t1 ->
                (i <= n)%nat ->
@@
 Lemma Lemma2_6_1 (alpha : T1) :  
-  nf alpha -> forall beta,  beta < alpha  ->
+  nf alpha -> forall beta,  beta o< alpha  ->
   {n:nat | const_pathS n alpha beta}.
@@
   (alpha beta : T1) (i : nat):
-  nf alpha -> const_pathS i alpha beta -> zero  < alpha ->
+  nf alpha -> const_pathS i alpha beta -> zero  o< alpha ->
   {l:nat | standard_path (S i) alpha j beta}.
@@
   Variables (alpha beta: T1) (n : nat).
-  Hypotheses (Halpha: nf alpha) (Hpos : zero <  beta)
+  Hypotheses (Halpha: nf alpha) (Hpos : zero o<  beta)
              (p : const_pathS n alpha  beta).
@@
   (alpha beta : T1) (i : nat):
-  nf alpha -> const_pathS i alpha beta -> zero  < alpha ->
+  nf alpha -> const_pathS i alpha beta -> zero  o< alpha ->
   {j:nat | standard_path (S i) alpha j beta}.
@@
 Corollary  LT_to_standard_path  (alpha beta : T1) :
-  beta < alpha ->
+  beta o< alpha ->
   {n : nat & {j:nat | standard_path (S n) alpha j beta}}.
@@
     forall alpha beta,
-      beta < alpha ->
+      beta o< alpha ->
       exists n i,  battle standard  n (iota alpha) i (iota beta).
@@
 \begin{Coqsrc}
- Lemma m_ge : m big_h <= m small_h.
+ Lemma m_ge : m big_h o<= m small_h.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Lemma m_lt : m small_h  < m big_h.
+Lemma m_lt : m small_h  o< m big_h.
 \end{Coqsrc}

### 2020-08-08 latex/chapter-large-sets.tex
corrections on pdf file

@@
   forall (i : nat) (alpha : T1),
-  nf alpha -> alpha <> zero -> canonS alpha i < alpha.
+  nf alpha -> alpha <> zero -> canonS alpha i o< alpha.
 
@@
 Remark H_non_mono1 :
-  ~ (forall alpha beta k, (alpha <= beta)%e0 ->
+  ~ (forall alpha beta k, (alpha o<= beta)%e0 ->
                           (H_ alpha k <= H_ beta k)%nat).
@@
  intros H ;specialize (H 42 omega 3).
- assert (H0 :(42 <= omega)%e0) by (repeat split; auto).  
+ assert (H0 :(42 o<= omega)%e0) by (repeat split; auto).  
  apply H in H0; rewrite H_Fin, H_omega  in H0; lia.
@@
 i : nat
-F_ : forall x : E0, nat -> x < alpha -> nat
-The term "F_ (Pred alpha) ?x" has type "Pred alpha < alpha -> nat"
-while it is expected to have type "Pred alpha < alpha -> Pred alpha < alpha"
-(cannot unify "nat" and "Pred alpha < alpha").
+F_ : forall x : E0, nat -> x o< alpha -> nat
+The term "F_ (Pred alpha) ?x" has type "Pred alpha o< alpha -> nat"
+while it is expected to have type "Pred alpha o< alpha -> Pred alpha o< alpha"
+(cannot unify "nat" and "Pred alpha o< alpha").
 \end{Coqanswer}
@@
 Lemma Propp284 : forall alpha beta : E0, 
-   beta < alpha -> dominates (F_ alpha) (F_ beta).
+   beta o< alpha -> dominates (F_ alpha) (F_ beta).
 \end{Coqsrc}

### 2020-08-08 latex/chapter-ordinals.tex
correction on pdf file
corrections on pdf file

@@
            {on : ON lt compare} := lt.
-Infix "<" := on_lt : ON_scope.
+Infix "o<" := on_lt : ON_scope.
 
@@
 
-Infix "<=" := on_le : ON_scope.
+Infix "o<=" := on_le : ON_scope.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Example ex1 : inl 7 < inr 0.
+Example ex1 : inl 7 o< inr 0.
 Proof. constructor. Qed.
@@
 \begin{Coqsrc}
-Example ex1' : fin 7 < omega.
+Example ex1' : fin 7 o< omega.
 Proof. constructor. Qed.
@@
 
-\label{warning:coercions}
-\index{Coq!Techniques!Coercions} 
-\begin{remark}
-Beware of coercions and notation scopes!
+% \label{warning:coercions}
+% \index{Coq!Techniques!Coercions} 
+% \begin{remark}
+% Beware of coercions and notation scopes!
+% Let us consider the following goal:
 
-Let us consider the following goal:
+% \begin{Coqsrc}
+%  Goal (6 o< 8).
+%  auto with arith.
+% \end{Coqsrc}
 
-\begin{Coqsrc}
- Goal (6 < 8).
- auto with arith.
-\end{Coqsrc}
+% \begin{Coqanswer}
+% 1 subgoal (ID 9)
 
-\begin{Coqanswer}
-1 subgoal (ID 9)
+%   ============================
+%   6 o< 8
+% \end{Coqanswer}
 
-  ============================
-  6 < 8
-\end{Coqanswer}
-
-Please keep in mind that the current notation scope interprets the infix \texttt{``<''} as the predicate \texttt{Omega\_plus\_omega.lt} and not \texttt{Nat.lt}.
+% Please keep in mind that the current notation scope interprets the infix \texttt{``<''} as the predicate \texttt{Omega\_plus\_omega.lt} and not \texttt{Nat.lt}.
 More,  the coercion mechanism converts the terms \texttt{6:nat} [resp.
 \texttt{8:nat} ]
-into \texttt{inl 6} [resp. \texttt{inl 8}].  So, the initial goal is correctly interpreted by \coq{}, but not as an inequality between two natural numbers.
+% into \texttt{inl 6} [resp. \texttt{inl 8}].  So, the initial goal is correctly interpreted by \coq{}, but not as an inequality between two natural numbers.
 
-\begin{Coqsrc}
-Set Printing All.
-\end{Coqsrc}
+% \begin{Coqsrc}
+% Set Printing All.
+% \end{Coqsrc}
 
-\begin{Coqanswer}
-1 subgoal (ID 337)
+% \begin{Coqanswer}
+% 1 subgoal (ID 337)
 
-  ============================
-  @on_lt nat Peano.lt Nat.compare Omega (S (S (S (S (S (S O))))))
-    (S (S (S (S (S (S (S (S O))))))))
-\end{Coqanswer}
+%   ============================
+%   @on_lt nat Peano.lt Nat.compare Omega (S (S (S (S (S (S O))))))
+%     (S (S (S (S (S (S (S (S O))))))))
+% \end{Coqanswer}
 
-Anyway, the initial goal is provable, using \texttt{le\_AsB} first constructor.
+% Anyway, the initial goal is provable, using \texttt{le\_AsB}'s first constructor.
 
-\begin{Coqsrc}
-  constructor; auto with arith.
-Qed.
-\end{Coqsrc}
+% \begin{Coqsrc}
+%   constructor; auto with arith.
+% Qed.
+% \end{Coqsrc}
 
-\end{remark}
+% \end{remark}
+%
 
@@
 \item $y$ is a \emph{successor} of $x$ if $x<y$ and there is no element between $x$ and $y$,
-\item $x$ is a \emph{limit} if $x$ is not a least element, and for any $y$ such that $y<x$,
+\item $x$ is a \emph{limit} if $x$ is not a least element, and for any $y$ such that $yo<x$,
  there exists some $z$ such that $y<z<x$.
@@
 \index{Projects}
+\label{exo:ON-mult}
 \begin{project}
@@
 
-For instance, let us take $N_A=\texttt{Omega}$, and take $\alpha=(4,4,2,1,0)$,
- $\beta=(4,3,3,3,3,3,2)$, and $\gamma=(5)$. Then $\beta<\alpha<\gamma$.
+For instance, let us take $N_A=\texttt{Omega}$, and take $\alpha=\langle 4,4,2,1,0\rangle$,
+ $\beta=\langle 4,3,3,3,3,3,2\rangle$, and $\gamma=\langle 5\rangle$. Then $\beta<\alpha<\gamma$.
 
-In contrast the list $(5,6,3,3)$ is not non-increasing (\emph{i.e.} sorted w.r.t. $\geq$), so it is not to be considered.
+In contrast the list $\langle5,6,3,3\rangle$ is not non-increasing (\emph{i.e.} sorted w.r.t. $\geq$), so it is not to be considered.
 

### 2020-08-08 latex/chapter-schutte.tex
corrections on pdf file

@@
 Parameter lt : relation Ord.
-Notation "a < b" := (lt a b): schutte_scope.
+Notation "a o< b" := (lt a b): schutte_scope.
 
@@
 Axiom AX2 : forall X: Ensemble Ord, 
-   (exists a,  (forall y, In X y -> y < a)) ->
+   (exists a,  (forall y, In X y -> y o< a)) ->
    countable X.
@@
               countable X -> 
-              exists a,  forall y, In X y -> y < a.
+              exists a,  forall y, In X y -> y o< a.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Lemma zero_le (alpha : Ord) :  zero <= alpha.
+Lemma zero_le (alpha : Ord) :  zero o<= alpha.
 Proof.
@@
  forall a : Ord, unique (least_member lt ordinal) a -> 
-                a <= alpha
+                a o<= alpha
 \end{Coqanswer}
@@
 \emph{i.e.}, satisfied by \emph{any} element of type \texttt{Ord}, like for instance
-\texttt{bottom = bottom}, or \texttt{zero <= bottom}.
+\texttt{bottom = bottom}, or \texttt{zero o<= bottom}.
 
 \begin{Coqbad}
-Lemma le_zero_bottom : zero <= bottom. 
+Lemma le_zero_bottom : zero o<= bottom. 
 Proof. apply zero_le. Qed.
@@
 
-Lemma le_bottom_zero : bottom <= zero.
+Lemma le_bottom_zero : bottom o<= zero.
 Proof.
@@
  forall a : Ord, unique (least_member lt (Empty_set Ord)) a -> 
-      a <= zero
+      a o<= zero
 \end{Coqanswer}
@@
 \begin{Coqsrc}
-Definition succ (alpha : Ord) := the_least (fun beta => alpha < beta).
+Definition succ (alpha : Ord) := the_least (fun beta => alpha o< beta).
 \end{Coqsrc}
@@
 Definition succ_spec (alpha:Ord) :=
-  least_member   lt (fun z => alpha < z).
+  least_member   lt (fun z => alpha o< z).
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Lemma lt_succ (alpha : Ord) :  alpha < succ alpha.
+Lemma lt_succ (alpha : Ord) :  alpha o< succ alpha.
 Proof.
@@
 Lemma lt_succ_le (alpha beta : Ord):
-  alpha < beta -> succ alpha <= beta.
+  alpha o< beta -> succ alpha o<= beta.
 Proof with eauto with schutte.
@@
 Lemma lt_succ_le_2 (alpha beta : Ord):
-  alpha < succ beta -> alpha <= beta.
+  alpha o< succ beta -> alpha o<= beta.
 
 Lemma succ_mono (alpha beta : Ord):
-  alpha < beta -> succ alpha < succ beta.
+  alpha o< beta -> succ alpha o< succ beta.
 
 Lemma succ_monoR (alpha beta : Ord) :
- succ alpha < succ beta -> alpha < beta.
+ succ alpha o< succ beta -> alpha o< beta.
 
 Lemma lt_succ_lt (alpha beta : Ord) :
-  is_limit beta ->  alpha < beta -> succ alpha < beta.
+  is_limit beta ->  alpha o< beta -> succ alpha o< beta.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Lemma finite_lt_omega : forall i: nat,  i < omega.
+Lemma finite_lt_omega : forall i: nat,  i o< omega.
 
 Lemma lt_omega_finite alpha : Ord) : 
-  alpha < omega ->  exists i:nat, alpha =  i.
+  alpha o< omega ->  exists i:nat, alpha =  i.
 
@@
  Definition segment (A: Ensemble Ord) :=
-  forall alpha beta, In A alpha -> beta < alpha -> In A  beta.
+  forall alpha beta, In A alpha -> beta o< alpha -> In A  beta.
 
@@
  (forall b, B b -> exists a, In A a /\ f a = b) /\
- forall a b, In A a -> In A b -> a < b ->  f a < f b.
+ forall a b, In A a -> In A b -> a o< b ->  f a o< f b.
 
@@
     ordering_function f A B ->
-    forall alpha, In A alpha -> alpha  <= f alpha.
+    forall alpha, In A alpha -> alpha  o<= f alpha.
 
@@
 
-Lemma le_plus_l (alpha beta : Ord) : alpha <= alpha + beta.
-
-Lemma le_plus_r (alpha beta : Ord) :  beta <= alpha + beta.
+Lemma le_plus_l (alpha beta : Ord) : alpha o<= alpha + beta.
+
+Lemma le_plus_r (alpha beta : Ord) :  beta o<= alpha + beta.
 
 Lemma plus_mono_r (alpha beta gamma : Ord) : 
-    beta < gamma -> alpha + beta < alpha + gamma.
+    beta o< gamma -> alpha + beta o< alpha + gamma.
 
@@
 Lemma finite_plus_ge_omega (n : nat) (alpha : Ord) :
-  omega <= alpha -> n + alpha = alpha.
+  omega o<= alpha -> n + alpha = alpha.
 \end{Coqsrc}
@@
   fun alpha => 
-  zero < alpha /\
-  (forall beta, beta < alpha ->  beta + alpha = alpha).
+  zero o< alpha /\
+  (forall beta, beta o< alpha ->  beta + alpha = alpha).
 \end{Coqsrc}
@@
   least_member   lt 
-                  (fun alpha => 1 < alpha /\ In AP alpha)
+                  (fun alpha => 1 o< alpha /\ In AP alpha)
                   omega.
@@
 Lemma AP_plus_closed (alpha beta gamma : Ord): 
-     In AP alpha -> beta < alpha -> gamma < alpha -> beta + gamma < alpha.
+     In AP alpha -> beta o< alpha -> gamma o< alpha -> beta + gamma o< alpha.
 
@@
 Lemma phi0_mono (alpha beta : Ord) :
-  alpha < beta ->  phi0 alpha < phi0 beta.
+  alpha o< beta ->  phi0 alpha o< phi0 beta.
 
@@
 Lemma is_limit_phi0 (alpha : Ord) :
-  zero < alpha ->  is_limit (phi0 alpha).
+  zero o< alpha ->  is_limit (phi0 alpha).
 
@@
 
-Lemma phi0_le (alpha : Ord) : alpha <= phi0 alpha.
+Lemma phi0_le (alpha : Ord) : alpha o<= phi0 alpha.
 \end{Coqsrc}
@@
 Lemma lt_phi0 (alpha : Ord):
-  alpha < epsilon0 -> alpha < phi0 alpha.
+  alpha o< epsilon0 -> alpha o< phi0 alpha.
 \end{Coqsrc}
@@
 Definition Cr_fun : forall alpha : Ord,
-       (forall beta : Ord, beta < alpha -> Ensemble Ord) ->
+       (forall beta : Ord, beta o< alpha -> Ensemble Ord) ->
         Ensemble Ord 
@@
         (Cr : forall beta, 
-                beta < alpha -> Ensemble Ord) 
+                beta o< alpha -> Ensemble Ord) 
         (x : Ord) => (
        (alpha = zero /\ AP x) \/
-       (zero < alpha /\
-        forall beta (H:beta < alpha),
+       (zero o< alpha /\
+        forall beta (H:beta o< alpha),
           the_ordering_segment (Cr beta H) x /\ ord (Cr  beta H) x = x)).
@@
 Definition sorted (l: cnf_t) :=
-  LocallySorted (fun alpha beta => beta <= alpha) l.
+  LocallySorted (fun alpha beta => beta o<= alpha) l.
 
@@
  forall l alpha, 
-   is_cnf_of alpha l ->  alpha < epsilon0 ->
-   Forall (fun beta =>  beta < alpha) l.
+   is_cnf_of alpha l ->  alpha o< epsilon0 ->
+   Forall (fun beta =>  beta o< alpha) l.
 \end{Coqsrc}
@@
    is_cnf_of alpha l ->  
-   (alpha < epsilon0 <->  Forall (fun beta =>  beta < alpha) l).
+   (alpha o< epsilon0 <->  Forall (fun beta =>  beta o< alpha) l).
 \end{Coqsrc}
@@
   T1.nf beta -> T1.nf gamma -> 
-  inject beta < inject gamma.
+  inject beta o< inject gamma.
 
@@
  Theorem inject_lt_epsilon0 (alpha : T1):
-      inject alpha < epsilon0.
+      inject alpha o< epsilon0.
 

### 2020-08-08 latex/chapter-termination.tex
specific infix for ordinal inequalities : first wave
corrections on pdf file

@@
 % Our formalisation of Cantor Normal Form will take two steps:
-% 1 Definition of a strict order \texttt{<} on the type \texttt{T1}, 
-% 2 Using \texttt{<} for characterizing terms in normal form.
+% 1 Definition of a strict order \texttt{o<} on the type \texttt{T1}, 
+% 2 Using \texttt{o<} for characterizing terms in normal form.
 
@@
 Definition LT := restrict nf lt.
-Infix "<" := LT : t1_scope.
+Infix "o<" := LT : t1_scope.
 
 Definition LE := restrict nf le.
-Notation "alpha <= beta" := (LE alpha beta) : t1_scope.
-\end{Coqsrc}
+Notation "alpha o<= beta" := (LE alpha beta) : t1_scope.
+\end{Coqsrc}
+
+\begin{todo}
+Infix notation for inequalities on ordinals have changed to \texttt{o<} and \texttt{o<=}.  This document is not updated yet, but the \coq{} scripts are correct.
+\end{todo}
 
@@
 \begin{Coqsrc}
-Lemma LT_one : forall alpha, alpha < one -> alpha = zero.
+Lemma LT_one : forall alpha, alpha o< one -> alpha = zero.
 \end{Coqsrc}
@@
 
-Infix "<" := Lt : E0_scope.
-Infix "<=" := Le : E0_scope.
-\end{Coqsrc}
+Infix "o<" := Lt : E0_scope.
+Infix "o<=" := Le : E0_scope.
+\end{Coqsrc}
+
+\begin{todo}
+Infix notation for inequalities on ordinals have changed to \texttt{o<} and \texttt{o<=}.  This document is not updated yet, but the \coq{} scripts are correct.
+\end{todo}
 
@@
 
-Example e'2 : omega < omega + 6.
+Example e'2 : omega o< omega + 6.
 Proof. now compute. Qed.
@@
 
-Example e'''2 : omega < omega * 6.
+Example e'''2 : omega o< omega * 6.
 Proof. now compute. Qed.
@@
 Lemma plus_not_monotonous : exists alpha beta gamma : T1,
-    alpha < beta /\ alpha + gamma = beta + gamma.
+    alpha o< beta /\ alpha + gamma = beta + gamma.
 Proof.
@@
 Lemma mult_not_monotonous :  exists alpha beta gamma : T1,
-      alpha < beta /\ alpha * gamma = beta * gamma.
+      alpha o< beta /\ alpha * gamma = beta * gamma.
 Proof.
@@
   delta : T1
-  Hdelta : delta < ocons beta n gamma
+  Hdelta : delta o< ocons beta n gamma
   ============================
@@
   alpha : T1
-  Aalpha : forall y : T1,  y < alpha -> Acc LT y
+  Aalpha : forall y : T1,  y o< alpha -> Acc LT y
   IHalpha : forall y : T1,
@@
 Lemma oplus_strict_mono_LT_l (alpha beta gamma : T1) :
-  nf gamma   -> alpha  < beta ->
-  alpha o+ gamma < beta o+ gamma.
+  nf gamma   -> alpha  o< beta ->
+  alpha o+ gamma o< beta o+ gamma.
 
 Lemma oplus_strict_mono_LT_r (alpha beta gamma : T1) :
-  nf alpha -> beta < gamma ->
-  alpha o+ beta < alpha o+ gamma.
+  nf alpha -> beta o< gamma ->
+  alpha o+ beta o< alpha o+ gamma.
 \end{Coqsrc}
@@
 Lemma S0_decr :
-  forall s s', S0  s s' -> ms s' < ms s.
+  forall s s', S0  s s' -> ms s' o< ms s.
 \end{Coqsrc}
@@
 Lemma R1_decr : forall h h',
-                  R1 h h' -> m h' < m h.
+                  R1 h h' -> m h' o< m h.
 \end{Coqsrc}
@@
 Lemma S1_decr n:
-  forall s s', S1 n s s' -> ms s' <  ms s.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma R2_decr n : forall h h', R2 n h h' -> m h'  < m h.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma round_decr : forall h h', h -1-> h' -> m h' < m h.
+  forall s s', S1 n s s' -> ms s' o<  ms s.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma R2_decr n : forall h h', R2 n h h' -> m h'  o< m h.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma round_decr : forall h h', h -1-> h' -> m h' o< m h.
 Proof.
@@
   red; apply Inclusion.wf_incl with 
-         (R2 := fun h h' =>  m h < m h').
+         (R2 := fun h h' =>  m h o< m h').
    red; intros;  now apply round_decr.

### 2020-08-12 latex/chapter-hydras.tex
proof simplification in L_alpha

@@
 \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-\index{Coq!Techniques!Type classes}
+\index{Coq!Type classes}
 
@@
 \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
-\index{Coq!Techniques!Type classes}
+
+\index{Coq!Type classes}
 \label{sect:hvariant-def}
@@
 \end{Coqsrc}
+
+\label{remark-infix}
+\begin{remark}
+We use in \coq{} scripts the infix notations \verb@o<@ and \verb@o<=@ for
+the orders $<$ and $\leq$ on ordinals. This convention allows us to define
+implicit coercions from \texttt{nat} into various types for representing ordinals.
+For instance a proposition like \verb@ 6 o< 8@ will be interpreted as 
+``the finite ordinal 6 is strictly less than the finite ordinal 8'', whereas 
+\verb@ 6 o< 8@ is a statement about natural numbers. Since many lemmas talk about both ordinals and natural numbers, we wished to avoid too many scope changes inside propositions.
+\end{remark}
 

### 2020-08-12 latex/chapter-large-sets.tex
proof simplification in L_alpha

@@
 
-\section{Introduction}
+%\section{Introduction}
 
@@
  From the \coq{} user's point of view, such  functions are  very 
-attractive:  they are defined as functions  in \gallina{}, and we can apply rhem \emph{in theory}, but they are so complex that you will never be able to look at the result of the computation.
+attractive:  they are defined as functions  in \gallina{}, and we can apply them \emph{in theory}, but they are so complex that you will never be able to look at the result of the computation.
  Thus, our knowledge on these functions must rely on \emph{proofs}. In our development, we use often the rewriting rules generated by \coq's \texttt{Equations} plug-in.
@@
 
+\subsection{Definition}
+
 \begin{definition}
-Let $0<\alpha<\epsilon_0$ be any ordinal, and $s$ a sequence of strictly positive natural numbers. We say that $s$ is an  \emph{$\alpha$-large} 
-sequence  if $s$ determines a path to the ordinal $0$.
-We say that $s$ is \emph{minimally $\alpha$-large} (abbreviated in
-\emph{$\alpha$-mlarge} if $s$ if $s$ is $\alpha$-large 
- and every strict prefix of $s$ leads to a non-zero ordinal.
+Let $0<\alpha<\epsilon_0$ be any ordinal, and $s$ a sequence of strictly positive natural numbers. 
+We say that $s$ is \emph{minimally $\alpha$-large} (in short:
+\emph{$\alpha$-mlarge}) if $s$ if $s$ is $\alpha$-large 
+ and every strict prefix of $s$ leads to a non-zero ordinal (\emph{cf} Sect.~\vref{sect:path-to-def}).
 
@@
 
+\vspace{4pt}
+
+\noindent
+\emph{From Module~ \href{../src/html/hydras.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}
+
+\index{Predicates!mlarge @ mlarge (minimal large sequences)}
+\begin{Coqsrc}
+Definition mlarge alpha (s:list nat) := path_to zero s alpha.
+\end{Coqsrc}
+
 \begin{remark}
   Ketonen and Solovay~\cite{KS81} consider  large finite \emph{sets} of natural numbers,  but they are mainly used as sequences. Thus, we chosed to represent them explicitely as (sorted) lists.
-\end{remark}
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~ \href{../src/html/hydras.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}
-
-\index{Predicates!mlarge @ mlarge (minimal large sequences)}
-\begin{Coqsrc}
-Definition mlarge alpha s := path_to zero s alpha.
-\end{Coqsrc}
-
-Let us consider two integers $k$ and $l$, such that $0<k<l$. In order to check whether the interval $[k,l]$ is minimally large for $\alpha$, it is enough to
-follow from $\alpha$ the path associated with the interval $[k,l($ and verify that the last ordinal we obtain is equal to $1$.
-
-This verification can be done by a simple computation. First, we define a function for following a path, by successively computing canonical sequences.
+
+They also consider large (but not minimally large) sets. They can be defined in
+\coq{} as follows:
 
@@
   end.
-\end{Coqsrc}
+
+Definition large alpha (s:list nat) := gnaw alpha s = zero.
+\end{Coqsrc}
+\end{remark}
+
+Let us consider two integers $k$ and $l$, such that $0<k<l$. In order to check whether the interval $[k,l]$ is minimally large for $\alpha$, it is enough to
+follow from $\alpha$ the path associated with the interval $[k,l($ and verify that the last ordinal we obtain is equal to $1$.
+
+\subsection{Example}
 
@@
 
-For instance, we can verify that the interval $[6,509]$ is $\omega^2$-mlarge.
+For instance, we can verify that the interval $[6,510]$ is $\omega^2$-mlarge.
 
@@
 \end{Coqsrc}
+
+\section{Computing mlarge sequences}
 
@@
 
-First, we notice that this  number $l$ exists, thanks to the following theorems:
-
-\begin{Coqsrc}
-canonS_LT:
-  forall (i : nat) (alpha : T1),
-  nf alpha -> alpha <> zero -> canonS alpha i o< alpha.
-
-T1_wf: well_founded LT.
-\end{Coqsrc}
-
+First, we notice that this  number $l$ exists, since the segment $[0,\epsilon_0)$ is well-founded and $\canonseq{\alpha}{i}<\alpha$ for any $i$ and $\alpha>0$.
 Moreover, it is unique:
 
+\vspace{4pt}
+\noindent
+\emph{From Module~\href{../src/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
 \begin{Coqsrc}
@@
 
-Thus, it seems obvious that there should be a function, parameterized by $\alpha$ which associates to any  strictly positive integer $k$ the number $l$ such that
-the interval $[k,l]$ is $\alpha$-large. It would be fine to write in \gallina{} a definition like this:
+Thus, it seems obvious that there must exist a function, parameterized by $\alpha$ which associates to any  strictly positive integer $k$ the number $l$ such that
+the interval $[k,l]$ is $\alpha$-mlarge. It would be fine to write in \gallina{} a definition like this:
 
@@
 use \coq{} to reason  about the \emph{specification} of \texttt{L},
-prove properties of any function which satisfies this specification, and finally define a correct function with the help of the \texttt{coq-equations} plug-in.
-
-\section{Looking for a function to define}
+prove properties of any function which satisfies this specification.
+In Sect.~\ref{sect:L-equaions}, we use the \texttt{coq-equations} plug-in
+to define a function \texttt{L\_}, and prove its correctness w.r.t. its specification.
+
+\subsection{Looking for a function definition}
 
 Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider the functions which associate to each stritly positive integer $k$ the number $l$, where 
-the interval $[k,l-1]$\;\footnote{also denoted by $[k,l)$}  is $\alpha$-mlarge.
+the interval $[k,l)$ is $\alpha$-mlarge.
 
@@
 
-Let us now prove properties of any function $f$ which satisfies 
+Let us now prove properties of any function $f$ (if any) which satisfies 
 \texttt{L\_spec}.
@@
 
-Our exploration of the $L_\alpha$s will follow the usual scheme : transfinite induction, and proof-by-cases : zero, successors and limit ordinals.
+Our exploration of the $L_\alpha$s  follows the usual scheme : transfinite induction, and proof-by-cases : zero, successors and limit ordinals.
 
@@
 Let $\beta$ be some ordinal, and assume the arithmetic function $f$ satisfies 
-the specification $\texttt{L\_spec}\;\beta$.  Let $k$ be any natural number.
+the specification $(\texttt{L\_spec}\;\beta)$.  Let $k$ be any natural number.
 Any path from $\texttt{succ}\,\beta$ to $0$ starting at $k+1$ can be decomposed into a first step from $\texttt{succ}\,\beta$ to $\beta$, then a path from
@@
 By hypothesis the interval $[k+2, f(k+2)-1]$ is $\beta$-mlarge.
-
-If we assume the inequality $k+1\leq f(k+2)$, we infer that the interval 
-$[k+1, f(k+2)]$ is $\beta+1$-mlarge. This new hypothesis allows us 
-to apply the following lemma 
-(from~\href{../src/html/hydras.Prelude.MoreLists.html\#interval_app}{Prelude.MoreLists}).
-
-\begin{Coqsrc}
-Lemma interval_app (i j k:nat):
-  (i <= j)%nat -> (j <= k)%nat ->
-  interval i k = interval i j ++ interval (S j) k.
-\end{Coqsrc}
-
-For ensuring this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <= f}, or \texttt{(fun\_le S f)} (defined 
+But the interval $[k+1, f(k+2)-1]$ is the concatenation of the singleton
+$\{k+1\}$ and the interval $[k+2, f(k+2)-1]$.
+So, the function $\lambda\,k.\,f(k+1)$ satisfies the specification $\texttt{L\_spec}\,\beta$.
+
+Note that our decomposition of intervals works only if the intervals we consider are not empty.
+In order to ensure this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <= f}, or \texttt{(fun\_le S f)} (defined
 in~\href{../src/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}).
@@
 \section{Using \texttt{Equations}}
-\label{sect:equations}
+\label{sect:L-equations}
 
@@
 
+\begin{todo}
+The proof of \texttt{L\_alpha.L\_ok\_lim} seems to be redundant with
+\texttt{Large\_Sets.L\_lim\_ok}.  
+\end{todo}
+
 \begin{Coqsrc}
@@
 \section{The Wainer-Hardy hierarchy}
-
+\label{sect:hardy}
 In order to give an idea of the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the so called \emph{Wainer-Hardy hierarchy} of fast growing functions,
@@
 \section{The Wainer hierarchy (functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
+\label{sect:wainer}
 

### 2020-08-12 latex/chapter-termination.tex
proof simplification in L_alpha

@@
 
-\index{Types!T2@T2 (ordinals in Veblen normal form)}
+\index{Types!T2@T2 (ordinals terms in Veblen normal form)}
 \emph{Please look also at the library
@@
 
-The type \texttt{T1} is not a correct ordinal notation, since it contains terms that are not in Cantor normal form.
+The type \texttt{T1} is not a correct ordinal notation, since it contains terms that are not in Cantor normal form. Moreover,
+in Sections~\ref{sect:equations}, \ref{sect:hardy},
+and \ref{sect:wainer}, we needed to define total recursive functions on well-formed ordinal terms less  than $\epsilon_0$, using the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777}.
  In order to define a type whose inhabitants represent just ordinals, we build a type gathering a term of type \texttt{T1} and a proof that this term is in normal form.
-This representation will be used in section~\ref{sect:equations} in order to define recursive functions with the plug-in \texttt{coq-equations}, and also to build an instance of the class \texttt{ON} of ordinal notations.
 
@@
 
-\index{Coq!Techniques!Type classes}
+\index{Coq!Type classes}
 
@@
 \end{Coqsrc}
+
+\index{Exercises}
+ \begin{exercise}
+Prove in \coq{} that for any ordinal $\alpha<\epsilon_0$, $\alpha$ is a limit if 
+and only if for all $\beta<\alpha$, the interval $[\beta,\alpha)$ is infinite.
+ \end{exercise}
 
@@
 Instance Epsilon0 : ON Lt compare.  
+(* ... *)
 \end{Coqsrc}

### 2020-08-13 latex/chapter-large-sets.tex
proof simplification in L_alpha

@@
 \subsubsection{The  ordinal zero}
-
+\label{sect:L-spec-zero}
 The base case is directly a consequence of the specification.
@@
 \subsubsection{Successor ordinals}
-
+\label{sect:L-spec-succ}
 Let $\beta$ be some ordinal, and assume the arithmetic function $f$ satisfies 
@@
 \subsubsection{Limit ordinals}
+\label{sect:L-spec-lim}
 
@@
 Lemma L_succ_eqn alpha i :  L_ (Succ alpha) i = L_  alpha (S i).
-\end{Coqsrc}
-
-Using these three lemmas as rewrite rules, we can prove more properties of the \texttt{L\_} family of functions, including its correctness w.r.t. the specification
-\texttt{L\_spec}.
-
-\begin{todo}
-The proof of \texttt{L\_alpha.L\_ok\_lim} seems to be redundant with
-\texttt{Large\_Sets.L\_lim\_ok}.  
-\end{todo}
+
+Hint Rewrite L_zero_eqn L_succ_eqn : L_rw.
+\end{Coqsrc}
+
+Using these three lemmas as rewrite rules, we can prove more properties of the functions \texttt{L\_ $\alpha$}.
+
+\begin{Coqsrc}
+Lemma L_finite : forall i k :nat,  L_ i k = (i+k)%nat.  
+(* Proof by induction on i, using L_zero_eqn and L_succ_eqn *)
+
+Lemma L_omega : forall k, L_ omega%e0 k = S (2 * k)%nat.
+(* Proof using L_finite and L_lim_eqn *)
+\end{Coqsrc}
+
+By  well-founded induction on $\alpha$, we prove the following lemmas:
 
@@
 
-Lemma L_omega : forall k, L_ omega%e0 k = S (2 * k)%nat.
-
 Theorem L_correct alpha : L_spec (cnf alpha) (L_ alpha).
@@
 
+Please note that the proof of \texttt{L\_correct} applies the lemmas proven in Sections~\ref{sect:L-spec-zero}, ~\ref{sect:L-spec-succ} and ~\ref{sect:L-spec-lim}.
+Our previous study of \texttt{L\_spec} allowed us to pave the way for the definition by \texttt{Equations} and the correctness proof.
+
 \subsubsection{Back to Hydra battles}
@@
 Theorem \texttt{battle\_length\_std } of
-Module~\href{../src/html/hydras.Hydra.Hydra_Theorems.html\#battle_length_std}{Hydra.Hydra\_Theorems} relates the length of standard battles with the functions $L\_alpha$.
+Module~\href{../src/html/hydras.Hydra.Hydra_Theorems.html\#battle_length_std}{Hydra.Hydra\_Theorems} relates the length of standard battles with the functions $L_\alpha$.
 
@@
 
-By induction on $i$, we prove a simple expression of \texttt{H\_ (OF i)}, where 
-\texttt{OF $i$}  is the $i$-th finite ordinal.
-
-\begin{Coqsrc}
-Lemma H_Fin : forall i k: nat,  H_ i k = (i+k)%nat.
+By induction on $i$, we prove a simple expression of \texttt{H\_ (Fin i)}, where 
+\texttt{Fin $i$}  is the $i$-th finite ordinal.
+
+\begin{Coqsrc}
+Lemma H_Fin : forall i k: nat,  H_ (Fin i) k = (i+k)%nat.
 Proof with eauto with E0.
@@
 Since the canonical sequence of $\omega^\omega$ is composed of all the
-$\omega^i$, we first need to express $H_{\omega^i}$ for every natural number $i$.
+$\omega^i$, we first need to express $H_{\omega^i}$ for any natural number $i$.
 
@@
 
-t seems hard to capture the complexity of this function by looking only at this
+\subsubsection{A numerical example}
+
+It seems hard to capture the complexity of this function by looking only at this
 ``exact'' formula. 
@@
 
-On the contrary, the fonctions of the Hardy hierarchy have the following five properties. %%% ICI lien avec KS81
-
-For any $\alpha < \epsilon_0$,
+On the contrary, the fonctions of the Hardy hierarchy have the following five properties~\cite{KS81}: for any $\alpha < \epsilon_0$,
 \begin{itemize}

### 2020-08-16 latex/chapter-introduction.tex
new exercise

@@
 \item ``The Termite and the Tower'', by Will Sladek~\cite{Sladek07thetermite}
-\item Chapter V of ``Proof Theory'' by Kurt Schütte~\cite{Schutte}
+\item Chapter V of ``Proof Theory'' by Kurt Schütte~\cite{schutte}
 \end{itemize}
@@
 The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on ordinal notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
-We added new material : axiomatic definition of countable ordinals after Schütte~\cite{Schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, equations, etc.
+We added new material : axiomatic definition of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, equations, etc.
 

### 2020-08-16 latex/chapter-large-sets.tex
new exercise

@@
 prove properties of any function which satisfies this specification.
-In Sect.~\ref{sect:L-equaions}, we use the \texttt{coq-equations} plug-in
+In Sect.~\ref{sect:L-equations}, we use the \texttt{coq-equations} plug-in
 to define a function \texttt{L\_}, and prove its correctness w.r.t. its specification.
@@
 \begin{itemize}
-\item If $\alpha=0$, then $\H_\alpha (k)= k$ for any natural number $k$.
+\item If $\alpha=0$, then $H_\alpha (k)= k$ for any natural number $k$.
 \item If $\alpha=\textrm{succ}(\beta)$, then 
@@
 \item If $\alpha$ is a limit ordinal, then 
-$\H_\alpha(k) = H_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
+$H_\alpha(k) = H_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
 \end{itemize}

### 2020-08-16 latex/chapter-termination.tex
new exercise

@@
 
-\index{Coq!Techniques!Coercions}
-\index{Functions!Coercions@Coercions (from nat to ordinal types)}
-\begin{remark}
-Please refer to the remark~\pageref{waening:coercions} about the use of coercions.
-% The use of coercions like \texttt{fin} allow us to be close to the mathematical tradition where natural numbers are ordinals too.
-% Nevertheless, it may happen that a goal like \texttt{3 < 5} could be 
-% interpreted as \texttt{(lt (fin 3) (fin 5))},  depending on the current notation scope.  
-% When this misinterpretation happens, tactics like \texttt{auto with arith}, \texttt{lia} do not work!
-% Thus, it is useful to write \texttt{(3 < 5)\%nat}  an inequality between two natural numbers. 
-\end{remark}
+% \index{Coq!Techniques!Coercions}
+% \index{Functions!Coercions@Coercions (from nat to ordinal types)}
+% \begin{remark}
+% Please refer to the remark~\pageref{warning:coercions} about the use of coercions.
+% % The use of coercions like \texttt{fin} allow us to be close to the mathematical tradition where natural numbers are ordinals too.
+% % Nevertheless, it may happen that a goal like \texttt{3 < 5} could be 
+% % interpreted as \texttt{(lt (fin 3) (fin 5))},  depending on the current notation scope.  
+% % When this misinterpretation happens, tactics like \texttt{auto with arith}, \texttt{lia} do not work!
+% % Thus, it is useful to write \texttt{(3 < 5)\%nat}  an inequality between two natural numbers. 
+% \end{remark}
 
@@
 The type \texttt{T1} is not a correct ordinal notation, since it contains terms that are not in Cantor normal form. Moreover,
-in Sections~\ref{sect:equations}, \ref{sect:hardy},
+in Sections~\ref{sect:L-equations}, \ref{sect:hardy},
 and \ref{sect:wainer}, we needed to define total recursive functions on well-formed ordinal terms less  than $\epsilon_0$, using the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777}.
@@
 \end{Coqsrc}
-
- Likewise, we define an addition in type \texttt{E0}, using the function{T1.plus}
-and its properties.
-
-\begin{Coqsrc}
-Search   (nf (_ + _)%t1).
-\end{Coqsrc}
-
-\begin{Coqanswer}
-plus_nf: forall a : T1, nf a -> forall b : T1, nf b -> nf (a + b)
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Instance plus (alpha beta : E0) : E0.
-Proof.
-  refine (@mkord (T1.plus (@cnf alpha) (@cnf beta))_ );
-    apply plus_nf; apply cnf_ok.
-Defined.
-
-Infix "+" := plus : E0_scope.
-
-Check omega + omega.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-omega + omega
-     : E0
-\end{Coqanswer}
 
@@
 \end{Coqanswer}
+
+\subsubsection{Arithmetic on type \texttt{E0}}
+
+ We define an addition in type \texttt{E0}, using the function{T1.plus}
+and its properties.
+
+\begin{Coqsrc}
+Search   (nf (_ + _)%t1).
+\end{Coqsrc}
+
+\begin{Coqanswer}
+plus_nf: forall a : T1, nf a -> forall b : T1, nf b -> nf (a + b)
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Instance plus (alpha beta : E0) : E0.
+Proof.
+  refine (@mkord (T1.plus (@cnf alpha) (@cnf beta))_ );
+    apply plus_nf; apply cnf_ok.
+Defined.
+
+Infix "+" := plus : E0_scope.
+
+Check omega + omega.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+omega + omega
+     : E0
+\end{Coqanswer}
+
+\index{Exercises}
+ \begin{exercise}
+Let $\alpha$ be an ordinal. We say that $\alpha$ is \emph{infinite} iff the segment $[0,\alpha)$ is an infinite set. 
+\begin{enumerate}
+\item Adapt this definition to the type \texttt{E0}.
+\item Prove that \emph{being infinite} is decidable on type \texttt{E0}.
+\item Prove that $\alpha$ is infinite if and only if for all finite ordinal $i$,
+$i+\alpha=\alpha$.
+\end{enumerate}
+ \end{exercise}
 

### 2020-08-19 latex/chapter-ordinals.tex
proof simplification in L_alpha
add a new project/exercise
minor corrections

@@
 
+\index{Coq!Type classes}
+
 \begin{Coqsrc}
@@
 \noindent\emph{From Library~\href{../src/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}}
+
+\index{Coq!Type classes}
 
@@
 \begin{Coqsrc}
-Lemma Successor_succ alpha : Successor (succ alpha) alpha.
-
 Lemma Successor_correct alpha beta : Successor beta alpha <->
@@
 
-We can also check whether an ordinal is a successor:
+We can also check whether an ordinal is a successor by a simple pattern matching:
 
@@
 
-Lemma succb_correct alpha : succb alpha <-> exists beta,  alpha = succ beta.
+Lemma succb_correct (alpha : t) :
+    succb alpha <->  exists beta: t,   alpha = succ beta.
 \end{Coqsrc}
@@
 We require also  that the comparison functions of the two notation systems are compatible.
+
+\index{Coq!Type classes}
 
@@
 $[0,n)\,=\,\{0,1,\dots,n-1\}$. 
-
 One may represent the ordinal $n$ by a sigma type.
@@
 
-The order on type \texttt{t $n$} is defined through the projection on \texttt{nat}.
+The order on type (\texttt{t $n$}) is defined through the projection on \texttt{nat}.
 
@@
 
-Note that the type \texttt{t 0} is empty, and that \texttt{Program} generates an obligation
-for verifying that the constraint $i<n$ is fullfilled when one tries to build the $i$-th ordinal in type \texttt{t $n$}.
+Note that the type (\texttt{t 0}) is empty, and that \texttt{Program} generates an obligation
+for verifying that the constraint $i<n$ is fullfilled when one tries to build the $i$-th ordinal inside the type (\texttt{t $n$}).
 
@@
 
+\index{Projects}
+\begin{project}
+Add to the class \texttt{ON} the requirement that for any $\alpha$ it is decidable whether $\alpha$ is $0$, a successor pr a mimit ordinal.
+\end{project}
+
 %%%% ICI ICI

### 2020-08-20 latex/chapter-ordinals.tex
typo corrections in *.tex

@@
 \begin{remark}
- The proof of \texttt{compare\_correct} uses a well-know pattern of \coq{}.
+ The proof of \texttt{compare\_correct} uses a well-known pattern of \coq{}.
 Let us consider  the following subgoal.
@@
 
-We have now to prove that two proofs of \texttt{Nat.ltb x0 (S n)} are equal.
-
-This is not obvious, but  a consequence of the following lemma of library 
+We have now to prove that there exists at most one  proof of \texttt{Nat.ltb x0 (S n)}. This is not obvious, but  a consequence of the following lemma of library 
 \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Eqdep_dec.html}{Coq.Logic.Eqdep\_dec}.
@@
 equality on type \texttt{bool}.
-
 This is why we used the boolean function \texttt{Nat.ltb} instead of the inductive predicate \texttt{Nat.lt} in the definition of type \texttt{t $n$} (see page~\pageref{def: Finite-ord-type}).
-
 For more information about this pattern, please look at the numerous mailing lists and 
@@
  \texttt{Coq.Wellfounded.Inclusion}, and \texttt{Coq.Arith.Wf\_nat}, we prove that our
-relation \texttt{wf} is well founded.
+relation \texttt{lt} is well founded.
 
@@
 Finally, it is interesting to show that, if $i<j$, then the segment $[0,i)$ is a ``sub-segment'' of
-$[0,j)$. Since the terms  $t\;i$ and $t\;j$ are not convertible, we consider a ``cast'' 
-function $\iota$ from $t\;i$ to $t\;j$, and prove this function is  a bijection  from $t\;i$ to
-the segment $[0,i)$ of $t\;j$.
+$[0,j)$. Since the terms  ($t\;i$) and ($t\;j$) are not convertible, we consider a ``cast'' 
+function $\iota$ from ($t\;i$) into ($t\;j$), and prove that this function is  a monotonous bijection  from ($t\;i$) to
+the segment $[0,i)$ of ($t\;j$).
 
@@
   {
-  iso_compare: forall x y : A,  compareB (f x) (f y) =
-                                 compareA x y;
+  iso_compare: forall x y : A, 
+      compareB (f x) (f y) = compareA x y;
   iso_inv1 : forall a, g (f a)= a;

### 2020-08-20 latex/chapter-hydras.tex
new exercise
typo corrections in *.tex

@@
 
-\pagebreak
 \vspace{4pt}\emph{From Module\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
@@
 
-The reader is encouraged to look at all the intermediary subgoals of this example.
+The reader is encouraged to look at all the successive subgoals of this example.
 \emph{Please consider also exercise~\vref{exo:interactive-battle}.}
@@
 
-The proof we are going to develop has exactly the same structure as in Section~\ref{sect:omega:case}
-
-previous one. Nevertheless, the proof of technical  lemmas is a little more complex, due to 
+The proof we are going to develop has exactly the same structure as in Section~\ref{omega-case}.
+ Nevertheless, the proof of technical  lemmas is a little more complex, due to 
  the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 

### 2020-08-20 latex/chapter-ks.tex
proof simplification in L_alpha
typo corrections in *.tex

@@
 \index{Maths!Accessibility inside epsilon0}
+\label{sect:pathes-intro}
 
@@
 \index{Predicates!path\_to}
+\label{sect:path-to-def}
 
@@
 
-\item Otherwise,  $\beta < \gamma$.  Let us consider  $\delta=\canonseq{\gamma}{n+l+2}$.
+\item Otherwise,  $\beta < \gamma$.  Let us consider  $\delta=\canonseq{\gamma}{n+l+1}$.
 By applying several times lemma \texttt{Cor12},  one converts  every path of Fig~\ref{fig:belle-preuve-2} into
@@
 But $\gamma$ is on the $n+l+1$-path from $\alpha$ to $\beta$.
-As shown by figure~\vref{fig:fin-belle-preuve}), the ordinal $\delta$, reachable from
+As shown by figure~\vref{fig:fin-belle-preuve}, the ordinal $\delta$, reachable from
 $\gamma$ in one single step,  must be greater than or equal to $\beta$, which contradicts our  hypothesis $\beta < \gamma$.
-
-\end{itemize}
- The only remaining case is  $\beta=\gamma$, thus we have got a standard path
-from $\alpha$ to $\beta$.
-
-\begin{Coqsrc}
- Lemma constant_to_standard_0 : 
-    {l : nat | standard_fun (S n) alpha l = beta}.
-
-End Constant_to_standard_Proof.
-
-Lemma constant_to_standard_path 
-  (alpha beta : T1) (i : nat):
-  nf alpha -> const_pathS i alpha beta -> zero  o< alpha ->
-  {j:nat | standard_path (S i) alpha j beta}.
-\end{Coqsrc}
 
@@
 
+\end{itemize}
+ The only possible case is  thus $\beta=\gamma$, so we have got a standard path  from $\alpha$ to $\beta$.
+
+\begin{Coqsrc}
+ Lemma constant_to_standard_0 : 
+    {l : nat | standard_fun (S n) alpha l = beta}.
+
+End Constant_to_standard_Proof.
+\end{Coqsrc}
+
+Here is the full statement of the conversion from constant to standard paths.
+
+\begin{Coqsrc}
+Lemma constant_to_standard_path 
+  (alpha beta : T1) (i : nat):
+  nf alpha -> const_pathS i alpha beta -> zero  o< alpha ->
+  {j:nat | standard_path (S i) alpha j beta}.
+\end{Coqsrc}
+
 Applying \texttt{Lemma2\_6\_1} and \texttt{constant\_to\_standard\_path}, we get the following corollary.
@@
 
-Since $m(\texttt{big\_h })< \mu$, there exists a standard path from $\mu$ to
+Since $m(\texttt{big\_h})< \mu$, there exists a standard path from $\mu$ to
 $m(\texttt{big\_h})$, hence a   standard battle from $\iota(\mu)$  to

### 2020-08-21 latex/chapter-large-sets.tex
typo corrections in *.tex

@@
 
+\index{Projects}
+\begin{project}
+Instead of considering standard paths and battles, consider ``constant'' paths and the corresponding battles.
+Please use \texttt{Equations} in order to define the function that computes the length of the $k$-path which leads  from $\alpha$ to $0$.
+Prove a few  exact formulas and minoration lemmas.
+\end{project}
+
 \section{The Wainer-Hardy hierarchy}

### 2020-08-24 latex/chapter-introduction.tex
corrections in hydras.tex

@@
 
-\subsection*{Remarks}
+\section{Remarks}
+
+\subsection{Difference from Kirby and Paris's work}
 In~\cite{KP82}, Kirby and Paris showed  that there is no proof of termination of all hydra battles in Peano Arithmetic (PA).
@@
 So we chosed to prove another statement without any reference to PA, by considering a class of proofs indexed by ordinal numbers upto $\epsilon_0$.
+
+\subsection{Trust in our proofs?}
+\label{sect:trust-in-proofs}
 
@@
 Thus, there is no ambiguity in our definitions and the premises of the theorems.
-Furthermore, you will be able to navigate through the development, using your favourite text editor or IDE, and some commands like \texttt{Search}, \texttt{Locate}, \texttt{Print Assumptions}, etc.
+Furthermore, you will be able to navigate through the development, using your favourite text editor or IDE, and some commands like \texttt{Search}, \texttt{Locate},  etc.
 
@@
 
-\paragraph*{Main references}
+\index{Coq!Commands!Print Assumptions}
+
+At any place of our development, you may use the command \texttt{Print Assumptions {\it ident}} in order to verify on which axiom the theorem {\it ident} may depend.
+The following example is extracted from
+Library~\href{../src/html/hydras.Epsilon0.F_alpha.html}{hydras.Epsilon0.F\_alpha}, we uses the \texttt{coq-equations} plug-in (see Sect.~\vref{sect:wainer}).
+
+\begin{Coqsrc}
+About F_zero_eqn.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+F_zero_eqn : forall i : nat, F_ Zero i = S i
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Print Assumptions F_zero_eqn. 
+\end{Coqsrc}
+
+\begin{Coqanswer}
+Axioms:
+FunctionalExtensionality.functional_extensionality_dep
+  : forall (A : Type) (B : A -> Type) (f g : forall x : A, B x),
+    (forall x : A, f x = g x) -> f = g
+Eqdep.Eq_rect_eq.eq_rect_eq
+  : forall (U : Type) (p : U) (Q : U -> Type) (x : Q p) (h : p = p),
+    x = eq_rect p Q x p h
+\end{Coqanswer}
+
+\subsection{Main references}
 
@@
 
-\subsubsection*{Structure of this document}
+\subsection{Structure of this document}
 \begin{itemize}
@@
 
-\subsubsection*{State of the development}
+\subsection{State of the development}
 The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on ordinal notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
@@
 
-\subsubsection*{Contributions}
+\subsection{Contributions}
 Many thanks to \'Evelyne Contejean and Florian Hatat for their contribution.
@@
 
-\subsubsection*{Acknowledgements}
+\subsection*{Acknowledgements}
 \label{sec:orgheadline5}
@@
 
-\subsection*{Typographical conventions}
+\subsection{Typographical conventions}
 
@@
 
-\subsection*{Alternative or bad definitions}
+\subsection{Alternative or bad definitions}
 

### 2020-08-24 latex/chapter-schutte.tex
proof simplification in L_alpha
corrections index

@@
 
-\vspace{4pt}
+We use Florian Hatat's library on countable sets, written as he was a student of  \emph{\'Ecole Normale Supérieure de Lyon}.
+A set $A$ is countable if there is an injective function from $A$ to $\mathbb{N}$ (see
+Library \href{../src/html/hydras.Schutte.Countable.html}%
+{\texttt{Schutte.Countable}}).
+
+\vspace{6pt}
 
@@
 The second and third axioms say that a subset $X$ of $\mathbb{O}$ is
-(strictly) bounded if and only if it is countable. We use Florian Hatat's library on countable sets, written as he was a student of  \emph{\'Ecole Normale Supérieure de Lyon}.
+(strictly) bounded if and only if it is countable.
 
@@
 
-\begin{remark}
-Please look at remark~\vref{warning:coercions}.
-\end{remark}
+% \begin{remark}
+% Please look at remark~\vref{warning:coercions}.
+% \end{remark}
 
@@
 
+\index{Projects}
+\begin{project}
+There is no construction of ordinal multiplication in~\cite{schutte}. 
+It would be interesting to derive this operation from Schütte's axioms,
+and prove its consistence with multiplication in ordinal notations for 
+$\epsilon_0$ and $\Gamma_0$.
+\end{project}
+
 \section{Related work}

### 2020-08-27 latex/chapter-introduction.tex
corrections introduction

@@
 
+\subsection{Assumed redundancy}
+
+It may often happen that several definitions of a given concept, or several proofs of a given theorem are possible.
+If all the versions present some interest, we will make them available, since each one may be of some methodological
+interest (by illustrating some tactic of proof pattern, for instance).
+We may use \coq's module system to make several proofs of a givent theorem co-exists in our libraries (see also Sect~\vref{sect:alt-proofs}).
+After some discussions of the pros and cons of each solution, we will choose to develop only one of them, leaving the others  as exercises or projects (i.e., big or difficult exercises).
+In order to discuss which assumptions are really needed for proving a theorem, we will also present 
+several aborted proofs.
+Of course, do not hesitate to contribute nice proofs or alternative definitions !
+
+It may also happen that some proof looks to be useless, because the proven theorem is a trivial consequence of another (proven too) result.
+For instance, let us consider the three following statements:
+\begin{enumerate}
+\item There is no measure into $\mathbb{N}$ for proving the termination of all hydra battles (Sect~\vref{omega-case}).
+\item There is no measure into $[0,\omega^2)$ for proving the termination of all hydra battle (Sect~\vref{omega2-case}).
+\item There is no measure into $[0,\mu)$ for proving the termination of all hydra battles, for any $\mu<\epsilon_0$ (Sect\vref{sec:free-battles-case}).
+\end{enumerate}
+
+Obviously, the third theorem implies the second one, which implies the first one.
+So, theoretically, a library would contain only a proof of $(3)$ and remarks for $(2)$ and $(1)$.
+But we found it interesting to make all the three proofs available, allowing the reader to compare their common structure and notice their technical differences.
+In particular, the proof of $(3)$ uses several non-trivial combinatorial properties of ordinal numbers~\cite{KS81}, whilst $(1)$ and $(2)$ use simple properties of $\mathbb{N}$ and $\mathbb{N}^2$.
+
+\subsection{About Logic}
+
+Most of the proofs we present are \emph{constructive}.
+Whenever possible, we provide the user with an associated function, which she or he can apply in \gallina{} or \ocaml{} in order to get a ``concrete'' feeling of the meaning of the considered theorem.
+For instance, in Chapter~\vref{chap:ketonen}, the notion of \emph{limit ordinal} is
+made more ``concrete'' thanks to a function \texttt{canon} which computes every item of a sequence which converges on a given limit ordinal $\alpha$.
+This simply typed function allows the user/reader to make her/his own experimentations.
+For instance, one can very easily compute the $42$-nd item of a sequence which converges towards $\omega^{\omega^\omega}$.
+
 \subsection{Main references}
@@
 \item Chapter V of ``Proof Theory'' by Kurt Schütte~\cite{schutte}
-\end{itemize}
-
-\subsection{Structure of this document}
-\begin{itemize}
-\item Whenever several implementations are possible, we will discuss the pros and cons of every possible choice
-\item Most of the proofs we present are \emph{constructive}.
-Whenever possible, we provide the user with an associated function, which she or he can apply in \gallina{} or \ocaml{} in order to get a ``concrete'' feeling of the meaning of the considered theorem.
-For instance, in Chapter~\vref{chap:ketonen}, the notion of \emph{limit ordinal} is
-made more ``concrete'' thanks to a function \texttt{canon} which computes every item of a sequence which converges on a given limit ordinal $\alpha$.
-This simply typed function allows the user/reader to make her/his own experimentations.
-For instance, one can very easily compute the $42$-nd item of a sequence which converges towards $\omega^{\omega^\omega}$.
-
-\item We found it interesting to present several  implementions of a given concept.
-After some discussions of the pros and cons of each solution, we will choose to develop only one of them, leaving the others  as exercises or projects (i.e., big or difficult exercises).
-In order to discuss which assumptions are really needed for proving a theorem, we will also present 
-several aborted proofs.
-
 \end{itemize}
@@
 Please do not hesitate to bring your contribution!
-
-\subsection*{Acknowledgements}
-\label{sec:orgheadline5}
-    Many thanks to David Ilcinkas, Sylvain Salvati, Alan Schmitt and Théo Zimmerman for their help on the elaboration of this document, and to the
- members of the \emph{Formal Methods} team at laBRI for their helpful comments 
-on an oral presentation of this work.
-
-Many thanks also to the Coq development team, Yves Bertot, and members of the \emph{Coq Club} for interesting discussions about the \coq{} system and the Calculus of Inductive Constructions.
-
-The author of the present document wishes to express his gratitude to the late Patrick Dehornoy, whose talk  was determinant for our desire to work on this topic.
-
-I owe my interest in discrete mathematics and their relation to formal proofs and functional programming  to Srecko Brlek.
-Equally, there is W.
-H.
-Burge's book ``\emph{Recursive Programming Techniques}'' ~\cite{burge} which was a great  source of inspiration.
 
@@
 
+\subsection{Active links}
+The  links which appear in this pdf  document lead are of three possible kinds of destination:
+\begin{itemize}
+\item Local links to the document itself,
+\item External links, mainly to \coq's page,
+\item Local links to pages generated by \texttt{coqdoc}. According to the current makefile,
+  we assume that the page generated from a library \texttt{XXX/YYY.v} is stored as
+the relative address \texttt{../src/html/hydras.XXX.YYY.html} (from the location of the pdf).
+Thus,  active links towards our \coq{} modules may be incorrect if you got this \texttt{pdf} document otherwise than by compiling the distribution available in
+\url{https://framagit.org/casteran/hydra-battles-in-coq}.
+
+\end{itemize}
+
 \subsection{Alternative or bad definitions}
-
+\label{sect:alt-proofs}
 Finally, we decided to include definitions or lemma statements, as well as tactics,  that lead to
@@
 
+\section{Acknowledgements}
+\label{sec:orgheadline5}
+    Many thanks to David Ilcinkas, Sylvain Salvati, Alan Schmitt and Théo Zimmerman for their help on the elaboration of this document, and to the
+ members of the \emph{Formal Methods} team at laBRI for their helpful comments 
+on an oral presentation of this work.
+
+Many thanks also to the Coq development team, Yves Bertot, and members of the \emph{Coq Club} for interesting discussions about the \coq{} system and the Calculus of Inductive Constructions.
+
+The author of the present document wishes to express his gratitude to the late Patrick Dehornoy, whose talk  was determinant for our desire to work on this topic.
+
+I owe my interest in discrete mathematics and their relation to formal proofs and functional programming  to Srecko Brlek.
+Equally, there is W.
+H.
+Burge's book ``\emph{Recursive Programming Techniques}'' ~\cite{burge} which was a great  source of inspiration.
+
 % \subsubsection{Links to the Coq source}
-
-% Active links towards our \coq{} modules may be incorrect if you got this \texttt{pdf} document otherwise than by compiling the distribution available in
-% \url{http://www.labri.fr/casteran/CoqArt/le_teaser/Teaser.tar.gz}.

### 2020-08-31 latex/chapter-ks.tex
typo corrections in *.tex
new table of notations

@@
 
-\index{Functions!canonS canonS (helper for canon)}
-\index{Functions!canon @ canon (canonical sequence)}
+\index{Functions!canonS@canonS (helper for canon)}
+\index{Functions!canon @canon (canonical sequence)}
 \begin{Coqsrc}
@@
 Let us consider a kind of accessibility problem inside $\epsilon_0$: given two ordinals $\alpha$ and $\beta$, where $\beta<\alpha<\epsilon_0$, find a \emph{path} consisting of a finite sequence $\gamma_0=\alpha,\dots,\gamma_l=\beta$,
-where, for every $i<l$, $\gamma_i \not= 0$ and there exists some strictly positive integer $s_i$
-such that $\gamma_{i+1}=\canonseq{\gamma}{s_i})$,
+where, for every $i<l$, $\gamma_i \not= 0$ \footnote{This condition allows us to ignore paths which end by a lot of useless $0$.} and there exists some strictly positive integer $s_i$
+such that $\gamma_{i+1}=\canonseq{\gamma}{s_i}$,
 
@@
 
-\begin{exercise}[Continuation of exercise~\vref{exo:simply-typed-canonseq}]
+\begin{exercise}[continuation of exercise~\vref{exo:simply-typed-canonseq}]
 Define a simply typed function for computing a path from $\alpha$ to $\beta$.
@@
 \noindent 
-From the lemma \texttt{canonS\_LT}~\vref{lemma:canonS_LT}, we convert any path into an inequality on ordinals (by induction on paths).
+From the lemma \texttt{canonS\_LT}~\vref{lemma:canonS_LT}, we can convert any path into an inequality on ordinals (by induction on paths).
 

### 2020-08-31 latex/chapter-ordinals.tex
corrections index
new table of notations

@@
 The following definitions allow us to make implicit several guessable arguments.
-
+\label{sect:on-lt-notation}
+\label{sect:on-le-notation}
 \begin{Coqsrc}
@@
 \end{Coqsrc}
+
+\begin{remark}
+The infix notations \texttt{o<} and \texttt{o<=} were defined in order to make apparent the distinction between the various notation scopes that may co-exist in a statement.
+In the informal and semi-formal statements, we still use $<$ and $\leq$.
+\end{remark}
 
@@
 
+\subsection{Definitions}
 It would be interesting to specify at the most generic level, what is a zero, a successor or a limit ordinal. Let $<$ be a strict order on a type $A$.
@@
 
-Using these definitions , we can prove the following lemma:
+\index{Exercises}
+\begin{exercise}
+Prove, that, in any ordinal notation system, that if $\beta$ is a successor of $\alpha$,
+then for any $\gamma$, $\gamma<\beta$ implies 
+$\gamma\leq\alpha$.
+\end{exercise}
+
+\subsection{Limits and successors in \texorpdfstring{$\omega+\omega$}{omega+omega}}
+
+Using thedefinitions above, we can prove the following lemma:
 
@@
 
-\index{Coq!Commands!Program Definition}
+\index{Coq!Commands!Program}
 
@@
 
-Note also that attempting to compare an ordinal of type \texttt{t $n$}  with an ordinal of
-type \texttt{t $p$}  leads to an error if $n$ and $p$ are not convertible.
+Note also that attempting to compare a term  of type (\texttt{t $n$}) with a term of
+type (\texttt{t $p$})  leads to an error if $n$ and $p$ are not convertible.
 

### 2020-08-31 latex/chapter-termination.tex
minor corrections
corrections index
new table of notations

@@
 \begin{remark}
-The proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering is available in the module \href{../src/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
+The alternate proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering is available in the library \href{../src/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
  \end{remark}
@@
 \emph{From Module~\href{../src/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}
+
+\label{sect:infix-oplus}
 
@@
 
+\index{Projects}
+\begin{project}
+Compare \texttt{oplus}'s definition (with inner fixpoint) with other possibilities
+(\texttt{coq-equations}, \texttt{Function}, etc.).
+\end{project}
 \subsection{More theorems on Hessenberg's sum}

### 2020-08-31 latex/hydras.tex
new chapter-skippable.tex
renaming zero_succ_limit-dec
simplification of class OrdinalNotation
corrections in the pdf documentation
correction on pdf file
proof simplification in L_alpha
added old Denumerable.v (to adapt)
new table of notations

@@
-\documentclass{book}
+\documentclass[a4paper]{book}
 \usepackage{alltt}
@@
 \usepackage{dsfont}
-\setcounter{tocdepth}{1}
+\setcounter{tocdepth}{2}
 \input{coqmacros}
 
-\author{Pierre Castéran\\ LaBRI, Univ. Bordeaux, CNRS UMR 5800, Bordeaux-INP}
+\author{Pierre Castéran\\ LaBRI, Univ. Bordeaux, CNRS UMR 5800, Bordeaux-INP\\with contributions by \'Evelyne Contejean and Florian Hatat}
 \date{\today}
-\title{The hydra and the rooster}
+\title{Hydra battles in Coq}
 \makeindex
@@
 
-\include{introduction}
+\include{chapter-introduction}
 \include{chapter-hydras}
+\include{chapter-ordinals}
 \include{chapter-termination}
@@
 \bibliographystyle{alpha}
-\bibliography{thebib, thebib2}
+\bibliography{thebib}
 
 \section{Future work (projects)}
+\index{Projects}
+
 This document and the proof scripts are far from being complete.
@@
 \section{Index}
+\subsection{Infix notations}
+
+\begin{tabular}{|c|c|c|c|l|}
+\hline
+Name & Ascii&Math& Description& Where \\\hline
+ \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality & \vref{sect:on-lt-notation}\\
+ \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \vref{sect:on-lt-notation}\\
+\texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \vref{sect:infix-oplus} \\
+\texttt{round} & \texttt{h -1-> h'} & & round of a battle & \vref{sect:infix-round} \\
+\hline
+\end{tabular}
+
 \printindex

### 2020-08-31 latex/hydras.tex
table of notations

@@
 \section{Index}
+Some notations may belong to several scopes. For instance, ``\texttt{o<}'' is
+bound in \texttt{ON\_scope}, \texttt{E0\_scope}, \texttt{t1\_scope}, etc., and locally in several libraries.
+
 \subsection{Infix notations}
+
+\vspace{4pt}
 
@@
 \hline
-Name & Ascii&Math& Description& Where \\\hline
- \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality & \vref{sect:on-lt-notation}\\
- \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \vref{sect:on-lt-notation}\\
-\texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \vref{sect:infix-oplus} \\
-\texttt{round} & \texttt{h -1-> h'} & & round of a battle & \vref{sect:infix-round} \\
+Name & Ascii&Math& Description& Page \\\hline
+ \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
+ \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
+\texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & Ordinal addition & \pageref{sect:infix-oplus} \\
+\texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\
+\texttt{round} & \texttt{h -1-> h'} & & one round of a battle & \pageref{sect:infix-round} \\
+\texttt{rounds} & \texttt{h -+-> h'} & & one or more  rounds of a battle & \pageref{sect:infix-rounds} \\
+\texttt{round\_star} & \texttt{h -*-> h'} & & any number of rounds of a battle & \pageref{sect:infix-rounds} \\
+
+\hline
+\end{tabular}
+
+\subsection{Other notations}
+
+\begin{tabular}{|c|c|c|c|l|}
+\hline
+Name & Ascii&Math& Description& Page \\\hline
+omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega},\dots\\
+phi0     & \texttt{phi0 alpha} & $\varphi_0(\alpha),\; \omega^\alpha$&exponential of base $\omega$ & \pageref{sect:notation-phi0}\\
+
 \hline

### 2020-09-01 latex/chapter-ks.tex
table of notations
corrections typo

@@
 First, let us recall how canonical sequences are defined in~\cite{KS81}.
-For efficiency's sake, we decided not to implement directly K.\&S's definitions, but to define in \gallina{} simply typed structurally recursive functions which share the abstract properties which are used in the mathematical proofs\footnote{With a small diffrence: the $0$-th term of the canonical sequence is not the same in our development as in~\cite{KS81}.}.
+For efficiency's sake, we decided not to implement directly K.\&S's definitions, but to define in \gallina{} simply typed structurally recursive functions which share the abstract properties which are used in the mathematical proofs\footnote{With a small difference: the $0$-th term of the canonical sequence is not the same in our development as in~\cite{KS81}.}.
 
@@
 
-The tool we need to use is once again in Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
+The tool we  use comes  once again from Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
 $\alpha_0=\alpha,\alpha_1,\dots,\alpha_k=\beta$ where $\alpha_{j+1}$ is equal
-to $\canonseq{\alpha_j}{i_j}$ for some $i_j$, 
-we will consider various constraints on these sequences.
+to $\canonseq{\alpha_j}{i_j}$ where $i_j$ is \emph{any} natural number, 
+we consider various constraints on these sequences.
 In particular, a path is called \emph{standard} if $i_{j+1} = i_j + 1$ for every $j<k$.

### 2020-09-01 latex/chapter-large-sets.tex
minor corrections

@@
 Note that our decomposition of intervals works only if the intervals we consider are not empty.
-In order to ensure this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <= f}, or \texttt{(fun\_le S f)} (defined
+In order to ensure this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <<= f}, or \texttt{(fun\_le S f)} (defined
 in~\href{../src/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}).
@@
 The functional \texttt{iterate}, similar to Standard Library's \texttt{Nat.iter},
-is defined and studied in~\href{../src/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates})
+is defined and studied in~\href{../src/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}).
 

### 2020-09-01 latex/chapter-termination.tex
table of notations
corrections typo
minor corrections

@@
 We provide also a notation for ordinals of the form $\omega^\alpha$.
+
+\label{sect:notation-phi0}
 
@@
 
-The following tactic tries to do a  transfinite induction on any ordinal \mathcolor{$\alpha<\epsilon_0$}.
+The following tactic starts a proof by  transfinite induction on any ordinal \mathcolor{$\alpha<\epsilon_0$}.
 
@@
 Proof.
-   destruct 1 as [n H]; destruct H. 
+   destruct 1 as [n [H | H]].
    -  now apply R1_decr.
@@
 the segment $[0,\epsilon_0)$ for proving the termination of any hydra battle
-\item There exists \emph{no} such variant from \texttt{Hydra} into $[0,\omega^2)$ and
-a fortiori into $[0,\omega)$.
+\item There exists \emph{no} such variant from \texttt{Hydra} into 
+$[0,\omega^2)$, \emph{a fortiori} into $[0,\omega)$.
 \end{itemize}

### 2020-09-01 latex/hydras.tex
minor corrections

@@
 Name & Ascii&Math& Description& Page \\\hline
+F & \texttt{F $n$} & $n$ & The $n$-th finite ordinal &  
+\pageref{sect:notation-F}, \pageref{sect:notation-F_sch}\\ 
 omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega},\dots\\

### 2020-09-02 latex/chapter-hydras.tex
new table of notations
table of notations
minor corrections

@@
 For this purpose, we define two mutual \emph{ad-hoc}  inductive types, where \texttt{Hydra} is the main type, and \texttt{Hydrae} is a helper for describing finite sequences of hydra.
-\index{Types!Hydra}
+\label{types:Hydra}
+\label{types:Hydrae}
 
@@
 \index{Predicates!round}
-
+\label{sect:infix-round}
 \begin{Coqsrc}
@@
 
+\label{sect:infix-rounds}
+
 \begin{Coqsrc}
@@
 \end{Coqsrc}
-
-\index{Notations!''-+->''}
-\index{Notations!''-*->''}
 
@@
 \index{Coq!Type classes}
+\label{types:Battle}
 
@@
 
-\begin{Coqsrc}
-Fixpoint iterate {A:Type} (f : A -> A) (n:nat) (x: A) :=
+The functional \texttt{iterate}, similar to Standard Library's \texttt{Nat.iter},
+is defined and studied in~\href{../src/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}).
+
+\label{Functions:iterate}
+
+\begin{Coqsrc}
+Fixpoint iterate {A:Type}(f : A -> A) (n: nat)(x:A) :=
   match n with

### 2020-09-02 latex/hydras.tex
minor corrections

--- 2020-08-31 latex/hydras.tex
@@
 \usepackage{dsfont}
+\usepackage{threeparttable}
 \setcounter{tocdepth}{2}
@@
 \section{Index}
-Some notations may belong to several scopes. For instance, ``\texttt{o<}'' is
-bound in \texttt{ON\_scope}, \texttt{E0\_scope}, \texttt{t1\_scope}, etc., and locally in several libraries.
 
-\subsection{Infix notations}
+\begin{table}[h]
+  \centering
+  \begin{threeparttable}
+    \caption{Main types and type classes}
+\begin{tabular}{|c|c|c|l|}
+\hline
+Name & Description& Page \\\hline
+Battle & Type class of battle types & \pageref{types:Battle} \\
+E0 & (Well formed) ordinals below $\epsilon_0$ & \pageref{types:E0}\\
+Hvariant & variants for proving terminatio of hydra battles & \pageref{sect:hvariant-def} \\
+Hydra & tree-like representation of hydras & \pageref{types:Hydra} \\
+Hydrae & finite sequences of hydras & \pageref{types:Hydrae} \\
+ON & ordinal  notations & \pageref{types:ON} \\
+ON\_for & comparison of an ordinal notation with Schütte's model &  \pageref{types:ON-for}\\
+ON\_iso & isomorphism of ordinal notations  & \pageref{types:ON-iso}\\
+
+Ord & countable ordinals (after Schütte) & \pageref{types:Ord} \\
+ppT1 & pretty printed version of T1 & \pageref{types:ppT1} \\
+SubON & comparison of ordinal notations & \pageref{types:SubON} \\
+T1 & ordinal terms below $\epsilon_0$ & \pageref{types:T1} \\
+T2 & ordinal terms below $\Gamma_0$ & \pageref{types:T2} \\
+WO & well orders (Schütte's definition) & \pageref{types:WO} \\
+\hline
+\end{tabular}
+\end{threeparttable}
+\end{table}
+
+\begin{table}[h]
+  \centering
+  \begin{threeparttable}
+    \caption{Main functions}
+\begin{tabular}{|r | c|c|c|c|l|}
+\hline
+Name & Gallina & Math & Description & Page \\ \hline
+canon & canon alpha i & $\canonseq{\alpha}{i}$& Canonical sequence & \pageref{Functions:canon}\\
+canonS & canonS alpha i & $\canonseq{\alpha}{i+1}$& Helper for canon & \pageref{Functions:canonS}\\
+F\_ & F\_ alpha n & $F_\alpha(n)$ & Wainer's \texttt{F} fast growing hierarchy & \pageref{Functions:F-alpha} \\
+iterate & iterate $f$ $n$ $x$ & $f^{(n)}(x)$ & Functional iteration & \pageref{Functions:iterate} \\
+\hline
+\end{tabular}
+\end{threeparttable}
+\end{table}
 
 \vspace{4pt}
-
+\begin{table}[h]
+  \centering
+  \begin{threeparttable}
+    \caption{Infix notations}
 \begin{tabular}{|c|c|c|c|l|}
 \hline
-Name & Ascii&Math& Description& Page \\\hline
- \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
+Name & Gallina&Math& Description& Page \\\hline
+ \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality \tnote{1} & \pageref{sect:on-lt-notation}\\
  \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
@@
 
-\subsection{Other notations}
+\begin{tablenotes}
+ \item [1]
+Some notations may belong to several scopes. For instance, ``\texttt{o<}'' is
+bound in \texttt{ON\_scope}, \texttt{E0\_scope}, \texttt{t1\_scope}, etc., and locally in several libraries.
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
 
+\begin{table}[h]
+  \centering
+  \begin{threeparttable}
+    \caption{Other notations}
 \begin{tabular}{|c|c|c|c|l|}
 \hline
-Name & Ascii&Math& Description& Page \\\hline
-omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega},\dots\\
+Name & Gallina&Math& Description& Page \\\hline
+F & \texttt{F $n$} & $n$ & The $n$-th finite ordinal &  
+\pageref{sect:notation-F}, \pageref{sect:notation-F-sch}\\ 
+FS & \texttt{FS $n$} & $n+1$ & The $n+1$-th finite ordinal  \tnote{2} &  
+\pageref{sect:notation-FS}\\ 
+omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega}, \pageref{sect:omega-T1}, \pageref{sect:omega-notation2}, \dots\\
 phi0     & \texttt{phi0 alpha} & $\varphi_0(\alpha),\; \omega^\alpha$&exponential of base $\omega$ & \pageref{sect:notation-phi0}\\
@@
 \end{tabular}
+
+\begin{tablenotes}
+  \item [2] Note that there exist also various coercions from \texttt{nat} to types of ordinal.
+Depending on the current scope and  \coq's syntactic analysis algorithm, \texttt{F} may be left implicit.
+\end{tablenotes}
+  \end{threeparttable}
+
+\end{table}
 

### 2020-09-03 latex/chapter-schutte.tex
table of notations
minor corrections

@@
 
-\index{Types!Ord@Ord (countable ordinals)}
+\label{types:Ord}
 
@@
 \index{Coq!Type classes}
+\label{types:WO}
 
@@
 
-Let us import the library \texttt{Coq.Logic.Epsilon}, which contains the following axiom.
+Let us import the library \texttt{Coq.Logic.Epsilon}, which contains the following axiom and lemmas.
 
@@
 Church's \emph{definite description operator}.
-
-\index{Functions!iota description operator}
 
@@
 
-\index{Functions!succ (in type Ord)}
+\label{Functions:succ-sch}
 
@@
 It is also easy to define recursively the finite ordinals.
+
+\label{sect:notation-F-sch}
 
@@
 
-\index{Notations!omega@omega (the least infinite ordinal)}
-
+\label{sect:notation-omega}
 \begin{Coqsrc}
@@
 In the mathematical text, we shall use indifferently the notations  $\omega^\alpha$ and$\varphi_0(\alpha)$.
-
-\index{Functions!phi0@phi0 (exponential of base omega)}
 

### 2020-09-03 latex/chapter-termination.tex
minor corrections

@@
 
-\index{Types!T1@T1 (ordinal terms below $\epsilon_0$)}
+\label{types:T1}
 \index{Constants!zero:T1}
@@
 
-\index{Types!T2@T2 (ordinals terms in Veblen normal form)}
+\label{types:T2}
 \emph{Please look also at the library
@@
 For representing finite ordinals, \emph{i.e.} natural numbers, we first introduce a notation for terms of the form $n+1$, then define a coercion from type \texttt{nat} into \texttt{T1}.
+\label{sect:notation-FS}
 
@@
 
+\label{sect:notation-F}
+
 \begin{Coqsrc}
@@
 Coercion fin  : nat >-> T1.
+
+Notation "'F' n" := (fin n) (at level 10) : t1_scope.
 
@@
 
-\index{Notations!omega@omega (the least infinite ordinal)}
+\label{sect:omega-notation2}
 \begin{Coqsrc}
@@
 A simple way to do this is to consider statements of the form 
-\texttt{forall alpha: T1, nf alpha -> $P$ alpha}, where $P$ is a predicate over type \texttt{T1}, lile in the following lemma.
+\texttt{forall alpha: T1, nf alpha -> $P$ alpha}, where $P$ is a predicate over type \texttt{T1}, like in the following lemma \footnote{Ordinal addition is formally defined a little later (page~\ref{sect:infix-plus-T1})}.
 
@@
 \label{sect:E0-def}
-\index{Types!E0@ EO (ordinal terms below $\epsilon_0$ in normal form)}
+\label{types:E0}
 
@@
 defined by a Boolean function, for  any term $\alpha:\texttt{T1}$, there exists at most one proof of \texttt{nf $\alpha$}, thus two ordinals of type \texttt{E0} are
-equal if and only iff their projection to \texttt{T1} are equal.
+equal if and only iff their projection to \texttt{T1} are equal (see also Sect.~\vref{sect:eq-proof-unicity}).
 
@@
 
-\index{Notations!omega@omega (the least infinite ordinal)}
+\label{sect:omega-T1}
 \index{Constants!zero:T1}
@@
 
-\index{Functions!T1.succ}
+\label{Functions:succ-T1}
 
@@
 Ordinal addition and multiplication are also defined by structural recursion over the type \texttt{T1}. Please note that they use the \texttt{compare} function on some subterms of their arguments.
+
+\label{sect:infix-plus-T1}
 
@@
 
-\index{Types!ppT1@ppT1 (pretty printed ordinal terms)}
+\label{types:ppT1}
 \index{Functions!pp@ pp (pretty printing terms in Cantor normal form)}

### 2020-09-03 latex/hydras.tex
minor corrections

@@
 E0 & (Well formed) ordinals below $\epsilon_0$ & \pageref{types:E0}\\
-Hvariant & variants for proving terminatio of hydra battles & \pageref{sect:hvariant-def} \\
+Hvariant & variants for proving termination of hydra battles & \pageref{sect:hvariant-def} \\
 Hydra & tree-like representation of hydras & \pageref{types:Hydra} \\
@@
 F\_ & F\_ alpha n & $F_\alpha(n)$ & Wainer's \texttt{F} fast growing hierarchy & \pageref{Functions:F-alpha} \\
+H\_ & H\_ alpha n & $H_\alpha(n)$ & Hardy's \texttt{H} fast growing hierarchy & \pageref{Functions:H-alpha} \\
 iterate & iterate $f$ $n$ $x$ & $f^{(n)}(x)$ & Functional iteration & \pageref{Functions:iterate} \\
+L\_ & L\_ alpha k & $L_\alpha(k)$ & final step of a standard path & \pageref{Functions:L-alpha} \\
+succ & succ alpha & & Successor & \pageref{Functions:succ-T1}, \pageref{Functions:succ-sch} \dots \\
 \hline
@@
  \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
-\texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & Ordinal addition & \pageref{sect:infix-oplus} \\
+\texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & ordinal addition & \pageref{sect:infix-plus-T1} ,\dots\\
 \texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\

### 2020-09-04 latex/chapter-termination.tex
restructuration of the index

@@
 
-\index{Predicates!lt @ lt (total, strict, non-well founded order on type T1)}
-Please note that this definition of \texttt{lt} makes it easy to write proofs by reflection, as shown by the following examples.
+\label{Predicates:lt-T1}
+Please note that this definition of \texttt{lt} makes it easy to write proofs by reflection, as shown by the following exampgles.
 
@@
 
-\index{Predicates!nf @ nf (being in Cantor normal form)}
+\label{Predicates:nf-T1}
 
@@
 
-\index{Predicates!LT @LT (partial, well-founded strict order on type T1)}
+\index{Predicates!LT-T1}
 
@@
 
-\index{Predicates!Lt @ Lt (Total, strict, well-founded order on type E0)}
+\label{Predicates:Lt-E0}
 

### 2020-09-04 latex/hydras.tex
restructuration of the index

@@
   \begin{threeparttable}
-    \caption{Main functions}
+    \caption{Main functions and constants}
 \begin{tabular}{|r | c|c|c|c|l|}
@@
 succ & succ alpha & & Successor & \pageref{Functions:succ-T1}, \pageref{Functions:succ-sch} \dots \\
+zero:Ord & & $0$ & The least ordinal (Schütte's model) & \pageref{Constants:zero:Ord} \\
+
 \hline
 \end{tabular}
+\end{threeparttable}
+\end{table}
+
+\begin{table}[h]
+  \centering
+  \begin{threeparttable}
+    \caption{Main predicates}
+\begin{tabular}{|r | c|c|c|c|l|}
+\hline
+Name & Gallina&Math& Description& Page \\\hline
+\texttt{lt : T1->T1->Prop}& lt alpha beta & $\alpha < \beta$& strict order on type \texttt{T1} \tnote{1} & \pageref{Predicates:lt-T1}\\
+\texttt{LT: T1->T1->Prop}& alpha o< beta & $\alpha < \beta$& strict order on type \texttt{T1}   \tnote{2} & \pageref{Predicates:LT-T1}\\
+\texttt{Lt : E0->E0->Prop} & alpha o< beta & $\alpha < \beta$& strict order on type \texttt{E0} \tnote{3} & \pageref{Predicates:Lt-E0} \\
+\texttt{nf: T1->Prop} & \texttt{nf alpha} && alpha is in Cantor normal form & \pageref{Predicates:nf-T1}\\
+\hline
+\end{tabular}
+\begin{tablenotes}
+  \item[1] This order is total, but not well-founded, because of not well formed terms.
+\item[2] Restriction of \texttt{lt} to terms in normal form; this order is partial, but well-founded.
+\item[3] This order is total \emph{and} well-founded.
+\end{tablenotes}
 \end{threeparttable}

### 2020-09-09 latex/chapter-ordinals.tex
minor corrections

@@
 \index{Coq!Type classes}
+\label{types:ON}
 
@@
 \index{Coq!Type classes}
+\label{types:SubON}
 
@@
 \index{Coq!Type classes}
+\label{types:ON-for}
 
@@
 \index{Coq!Techniques!Unicity of equality proofs}
+\label{sect:eq-proof-unicity}
 
@@
 \index{Coq!Type classes}
-
+\label{types:ON-iso} 
 \begin{Coqsrc}
@@
 \begin{project}
-Add to the class \texttt{ON} the requirement that for any $\alpha$ it is decidable whether $\alpha$ is $0$, a successor pr a mimit ordinal.
+\label{project:succ-limit-dec}
+Add to the class \texttt{ON} the requirement that for any $\alpha$ it is decidable whether $\alpha$ is $0$, a successor or a limit ordinal.
+\end{project}
+
+\index{Projects}
+\begin{project}
+\label{project:on-setoid}
+Reconsider the  class \texttt{ON}, with an equivalence instead of Leibniz equality.
 \end{project}
@@
 \index{Projects}
+
+\begin{project}
+The directory \texttt{src/OmegaOmega} contains an ad-hoc formalization of $\omega^\omega$, contributed by Pascal Manoury.
+Every ordinal $\alpha$ is represented by a list $l$ whose elements are the coefficients of $\omega$ in  the Cantor normal form of $\alpha$ (in reverse order).
+For instance, the ordinal
+$\omega^{8}\times 5 + \omega^{6}\times 8 + \omega^2\times 10 + \omega + 7$ is represented by the list \texttt{[5;0;8;0;0.0;10,1,7]}.
+
+ Develop this representation and compare it with the other ordinal notations.
+
+\end{project}
+
+\index{Projects}
+
 \begin{project}

### 2020-09-13 doc/chapter-introduction.tex
update README.md; move latex/ to doc/

--- 2020-08-27 latex/chapter-introduction.tex
@@
 
-Except in the \texttt{Schutte} library, dedicated to an axiomatic presentation of the set of countable ordinal numbers, all our development is axiom-free, and respects the rules of intuitionistic logic.
-Note that we also use the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777} for defining several rapidly growing hierarchy of functions, in Chap.~\ref{chap:alpha-large}.
-This plug-in imports the known-as-harmless following axiom:
-
-\begin{Coqsrc}
-FunctionalExtensionality.functional_extensionality_dep : 
-forall (A : Type) (B : A -> Type) (f g : forall x : A, B x),
-(forall x : A, f x = g x) -> f = g
-\end{Coqsrc}
-
-\index{Coq!Commands!Print Assumptions}
-
-At any place of our development, you may use the command \texttt{Print Assumptions {\it ident}} in order to verify on which axiom the theorem {\it ident} may depend.
-The following example is extracted from
-Library~\href{../src/html/hydras.Epsilon0.F_alpha.html}{hydras.Epsilon0.F\_alpha}, we uses the \texttt{coq-equations} plug-in (see Sect.~\vref{sect:wainer}).
-
-\begin{Coqsrc}
-About F_zero_eqn.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-F_zero_eqn : forall i : nat, F_ Zero i = S i
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Print Assumptions F_zero_eqn. 
-\end{Coqsrc}
-
-\begin{Coqanswer}
-Axioms:
-FunctionalExtensionality.functional_extensionality_dep
-  : forall (A : Type) (B : A -> Type) (f g : forall x : A, B x),
-    (forall x : A, f x = g x) -> f = g
-Eqdep.Eq_rect_eq.eq_rect_eq
-  : forall (U : Type) (p : U) (Q : U -> Type) (x : Q p) (h : p = p),
-    x = eq_rect p Q x p h
-\end{Coqanswer}
-
 \subsection{Assumed redundancy}
@@
 
+Except in the \texttt{Schutte} library, dedicated to an axiomatic presentation of the set of countable ordinal numbers, all our development is axiom-free, and respects the rules of intuitionistic logic.
+Note that we also use the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777} in the definitition of  several rapidly growing hierarchy of functions, in Chap.~\ref{chap:alpha-large}.
+This plug-in imports several known-as-harmless  axioms.
+
+% \begin{Coqsrc}
+% FunctionalExtensionality.functional_extensionality_dep : 
+% forall (A : Type) (B : A -> Type) (f g : forall x : A, B x),
+% (forall x : A, f x = g x) -> f = g
+% \end{Coqsrc}
+
+\index{Coq!Commands!Print Assumptions}
+
+At any place of our development, you may use the command \texttt{Print Assumptions {\it ident}} in order to verify on which axiom the theorem {\it ident} may depend.
+The following example is extracted from
+Library~\href{../src/html/hydras.Epsilon0.F_alpha.html}{hydras.Epsilon0.F\_alpha}, where we use the \texttt{coq-equations} plug-in (see Sect.~\vref{sect:wainer}).
+
+\begin{Coqsrc}
+About F_zero_eqn.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+F_zero_eqn : forall i : nat, F_ Zero i = S i
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Print Assumptions F_zero_eqn. 
+\end{Coqsrc}
+
+\begin{Coqanswer}
+Axioms:
+FunctionalExtensionality.functional_extensionality_dep
+  : forall (A : Type) (B : A -> Type) (f g : forall x : A, B x),
+    (forall x : A, f x = g x) -> f = g
+Eqdep.Eq_rect_eq.eq_rect_eq
+  : forall (U : Type) (p : U) (Q : U -> Type) (x : Q p) (h : p = p),
+    x = eq_rect p Q x p h
+\end{Coqanswer}
+
 \subsection{Main references}
@@
 Many thanks to \'Evelyne Contejean and Florian Hatat for their contribution.
-\'Evelyne contributed libraries on the recusrive path ordering (\emph{rpo}) for proving the well-foundedness of our representation of $\epsilon_0$ and $\Gamma_0$.
+\'Evelyne contributed libraries on the recursive path ordering (\emph{rpo}) for proving the well-foundedness of our representation of $\epsilon_0$ and $\Gamma_0$.
 Florian Hatat proved many useful lemmas on countable sets, which we used in our adaptation of Schütte's formalization of countable ordinals.

### 2020-09-13 doc/chapter-ordinals.tex
update README.md; move latex/ to doc/

--- 2020-09-09 latex/chapter-ordinals.tex
@@
 In our formalization, the interpretation of an ordinal as a set is realized by the following definition
-(in ~\href{../src/html/hydras.Prelude.Ordinal_generic.html}{Prelude.Ordinal\_generic}).
+(in ~\href{../src/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}).
 
@@
 
-{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$ we will prove in larger notation systems, as well as in Schütte's model}
+{\it \textbf{Note:s} This property reflects the equality $i+\omega=\omega$ we will prove in larger notation systems, as well as in Schütte's model}
+
+This exercise is partially solved for $i=3$ (in ~\href{../src/html/hydras.OrdinalNotations.Example_3PlusOmega.html}{OrdinalNotations.Example\_3PlusOmega}).
+
 \end{exercise}

### 2020-09-13 doc/hydras.tex
update README.md; move latex/ to doc/

--- 2020-09-04 latex/hydras.tex
@@
 \end{itemize}
+
+\section{Comments on exercises and projects}
+
+Although we do not plan to include complete solutions to the exercises, 
+we think it would be useful to include comments and hints, and questions/answers from the users. In constrast, ``projects'' are supposed, once completed, to be included in the repository.
+
+\subsubsection{Project \vref{project:succ-limit-dec}}
+\paragraph*{Hint:} 
+  Beware of the instances associated with sum and product of notations !
+  You may consider additional fields such as ``there is a greatest element'' 
+to make the sum and product of notations ``compositional''.
 

### 2020-09-14 doc/chapter-ordinals.tex
minor modifications

@@
 on \emph{ordinal numbers}.
-From a mathematical point of view, an ordinal is a representant of an equivalence class for isomorphims of strict, total and well-founded ordered sets.
+From a mathematical point of view, an ordinal is a representant of an equivalence class for isomorphims of  totally ordered well-founded sets.
 
-More intuitively, let us have a look at Figure~\ref{fig:ordinal-sequence}. It presents a sequence of ordinal numbers, as an extension of the sequence of natural numbers.
+More intuitively, let us have a look at Figure~\ref{fig:ordinal-sequence}. It presents a small sequence of ordinal numbers, which extends the sequence of natural numbers.
 
@@
 \item The ordinals are listed in a strictly increasing order. 
-\item Dots : ``$\ldots$'' stand for an infinite sequence of ordinals, not shown for lack of space. For instance, the ordinal $42$ is not shown in the first line, but it exists, betaween $41$ and $43$.
-\item Each ordinal in black is an immediate successor of anotter ordinal.
+\item Dots : ``$\ldots$'' stand for an infinite sequence of ordinals, not shown for lack of space.
+For instance, the ordinal $42$ is not shown in the first line, but it exists, between $17$ and $\omega$.
+\item Each ordinal printed in black is the immediate successor of another ordinal.
 We call it a \emph{successor} ordinal.
-For instance, $43$ is the successor of $42$, and $\omega^4+1$ the successor of $\omega^4$.
+For instance, $12$ is the successor of $11$, and $\omega^4+1$ the successor of $\omega^4$.
 \item Ordinals that  follow dots immediately (displayed in red) are called \emph{limit ordinals}.
@@
 \item
-For instance $\omega$ is the least upper bound of the set of all finite ordinals (in the first line). $\omega$ is also the first limit ordinal, and the first infinite ordinal, in the sense that 
+For instance $\omega$ is the least upper bound of the set of all finite ordinals (in the first line). It is also the first limit ordinal, and the first infinite ordinal, in the sense that 
 the set $\mathbb{O}_\omega$ is infinite.
@@
 
-\subsection{Instances of notation systems}
+\section{Instances of notation systems}
 
-\subsubsection{The ordinal \texorpdfstring{$\omega$}{omega}}
+\subsection{The ordinal \texorpdfstring{$\omega$}{omega}}
 
@@
 
+\subsection{The ordinal \texorpdfstring{$\omega^2$}{omega^2}}
+
+The ordinal $\omega^2$ (also called $\varphi_0(2)$, see Chap.~\ref{chap:schutte}), is represented by the lexicographically ordered cartesian product $\mathbb{N}\times\mathbb{N}$.
+
+\vspace{4pt}
+\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Omega2.html}{ON\_Omega2}}
+
+\begin{Coqsrc}
+Definition t := (nat * nat)%type.
+
+Definition lt : relation t := lexico Peano.lt Peano.lt.
+Definition le := clos_refl _ lt.
+
+Infix "o<" := lt : o2_scope.
+Infix "o<=" := le : o2_scope.
+\end{Coqsrc}
+
+It is easy to represent the finite ordinals, successors and limits.
+
+\begin{Coqsrc}
+Definition zero: t := (0,0).
+
+Definition fin (n:nat) : t := (0, n).
+
+Notation "'F'" := fin : o2_scope.
+
+Coercion fin : nat >-> t.
+
+Definition succ (alpha : t) := (fst alpha, S (snd alpha)).
+
+Notation "'omega'" := (1,0) : o2_scope.
+\end{Coqsrc}
+
 \section{Comparing Ordinal Notations}
@@
 \vspace{4pt}
-\noindent\emph{From Module~\href{../src/html/hydras.Prelude.Finite_ordinals.html}{Prelude.Finite\_ordinals}}
+\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
 
@@
 
-We have now to prove that there exists at most one  proof of \texttt{Nat.ltb x0 (S n)}. This is not obvious, but  a consequence of the following lemma of library 
+We have now to prove that there exists at most one  proof of (\texttt{Nat.ltb x0 (S n)}). This is not obvious, but  a consequence of the following lemma of library 
 \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Eqdep_dec.html}{Coq.Logic.Eqdep\_dec}.
@@
 forall (A : Type) (x : A),
-(forall y : A, x = y \/ x <> y) -> forall (y : A) (p1 p2 : x = y), p1 = p2
+(forall y : A, x = y \/ x <> y) -> 
+forall (y : A) (p1 p2 : x = y), p1 = p2
 \end{Coqanswer}

### 2020-09-15 doc/chapter-ordinals.tex
minor modifications

@@
 
-\subsection{Countable ordinals}
+Two kinds of representation of ordinals are defined herein.
+
+\begin{itemize}
+\item A ``mathematical'' representation of the set of countable ordinal numbers, afer Kurt Schütte~\cite{schutte}.
+This representation uses several (hopefully harmess) axioms.
+It is used as a reference for proving the correctness of ordinal notations.
+\item A family of \emph{ordinal notations}, \emph{i.e.} data types used to represent segments $[0,\mu)$, where $\mu$ is some countable ordinal.
+Each ordinal notation is defined inside the Calculus of Inductive Constructions (without axioms).
+Many functions are defined, allowing proofs by computation.
+Note that proofs of
+correctness of a given ordinal notation with respect to Schütte's model use axioms.
+Please execute the \texttt{Print Assumptions} command in case of doubt.
+\end{itemize}
+
+\section{Countable ordinals}
 
@@
 
-\subsection{Ordinal notations}
+\section{Ordinal notations}
 
@@
 \item A well founded order $<$ on $A$, 
-\item A Function for comparing two ordinals, which implies that the reflexive closure of $<$ is a \emph{total order}.
+\item A function for comparing two ordinals, whose correctnes implies that the reflexive closure of $<$ is a \emph{total order}.
 \end{itemize}
@@
 
-\subsection{Ordered types}
+\subsubsection*{Ordered types}
 
@@
 \begin{remark}
-The infix notations \texttt{o<} and \texttt{o<=} were defined in order to make apparent the distinction between the various notation scopes that may co-exist in a statement.
-In the informal and semi-formal statements, we still use $<$ and $\leq$.
+The infix notations \texttt{o<} and \texttt{o<=} were defined in order to make apparent the distinction between the various notation scopes that may co-exist in a same statement.
+So the infix \texttt{<} and \texttt{<=} are reserved to the natural numbers.
+In the mathematical formulas, we still use $<$ and $\leq$ for comparing ordinals.
 \end{remark}
 
-\section{Instances of notation systems}
+\section{Examples of ordinal notations}
 
@@
 
-In order to build an instance of \texttt{ON}, we have to define a comparison function.
+Before building an instance of \texttt{ON}, we have to define a comparison function.
 
@@
 We can now define abbreviations.
-For instance, the finite ordinals will be represented by the terms built with  the constructor \texttt{inl}, and the first infinite ordinal $\omega$ by the term \texttt{(inr 0)}.
+For instance, the finite ordinals are represented by terms built with  the constructor \texttt{inl}, and the first infinite ordinal $\omega$ by the term \texttt{(inr 0)}.
 
@@
 \item A \emph{least} element is a minorant (in the large sense) of the full set  on $A$,
-\item $y$ is a \emph{successor} of $x$ if $x<y$ and there is no element between $x$ and $y$,
+\item $y$ is a \emph{successor} of $x$ if $x<y$ and there is no element between $x$ and $y$. We will also say that $x$ is a \emph{predecessor} of $y$.
 \item $x$ is a \emph{limit} if $x$ is not a least element, and for any $y$ such that $yo<x$,
@@
 
-The following definitions are in Library\href{../src/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
+The following definitions are in Library \href{../src/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
 
@@
 \begin{exercise}
+Prove, that, in any ordinal notation system, that every ordinal has at most one predecessor, and at most one successor. 
+\end{exercise}
+
+\index{Exercises}
+\begin{exercise}
 Prove, that, in any ordinal notation system, that if $\beta$ is a successor of $\alpha$,
@@
 
-Using thedefinitions above, we can prove the following lemma:
+Using the definitions above, we can prove the following lemma:
 
@@
 Notation "'omega'" := (1,0) : o2_scope.
+
+Definition limitb (alpha : t) := 
+ match alpha with
+ |  (S _, 0) => true
+ | _ => false
+end.
+\end{Coqsrc}
+
+In order to build an ordinal notation out the type \texttt{t}, we have to define a comparison function, and prove that the order \texttt{lt} is well-founded.
+
+\begin{Coqsrc}
+Definition compare (alpha beta: t) : comparison :=
+  match Nat.compare (fst alpha) (fst beta) with
+    Eq => Nat.compare (snd alpha) (snd beta)
+  | c => c
+  end.
+
+Lemma compare_correct alpha beta :
+    CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
+                (compare alpha beta).
+
+Instance Omega2 : ON  lt compare.
 \end{Coqsrc}
@@
 The following class specifies a strict inclusion of segments.
-The notation \texttt{OA} describes a segment $[0,alpha($, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented by any term of type $A$).
+The notation \texttt{OA} describes a segment $[0,\alpha($, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented in \texttt{OA}).
 We require also  that the comparison functions of the two notation systems are compatible.
@@
   {
-  subseg_compare :forall x y : A,  compareB (iota x) (iota y) =
+  subon_compare :forall x y : A,  compareB (iota x) (iota y) =
                                  compareA x y;
-  subseg_incl : forall x, ltB (iota x) alpha;
-  subseg_onto : forall y, ltB y alpha  -> exists x:A, iota x = y}.
+  subon_incl : forall x, ltB (iota x) alpha;
+  subon_onto : forall y, ltB y alpha  -> exists x:A, iota x = y}.
 \end{Coqsrc}
@@
 
-\subsection{Comparing an ordinal notation with Schütte's model}
+\section{Comparing an ordinal notation with Schütte's model}
 
@@
 
-We are now able to build an instance of \texttt{Subsegment}.
+We are now able to build an instance of \texttt{SubON}.
 

### 2020-09-16 doc/chapter-ordinals.tex
minor modifications

@@
 \end{remark}
+
+\subsection{Ordinal notations and measures for proving termination}
+\label{sect:measure-ON}
+
+The following lemma (together with the type class mechnism) allows us to use simply  measures towards an ordinal notation.
+It is just an application of  the libraries \texttt{Coq.Wellfounded.Inverse\_Image}
+and  \texttt{Coq.Wellfounded.Inclusion}.
+
+\begin{Coqsrc}
+Definition m_lt {A:Type}{lt: relation A}
+            {compare : A -> A -> comparison}
+            {on : ON lt compare}
+            {B : Type}
+  (m : B -> A) : relation B :=
+  fun x y =>  on_lt (m x) (m y).
+
+Lemma wf_measure  {A:Type}(lt: relation A)
+            {compare : A -> A -> comparison}
+            {on : ON lt compare}
+            {B : Type}
+            (m : B -> A) :
+  well_founded (m_lt m). 
+\end{Coqsrc}
+
+A simple example of application is given in Sect.~\vref{sect:merge-example}.
 
@@
 
-Finally the nature of any ordinal (inside this notation system) is decidable:
-
-\begin{Coqsrc}
-Lemma ZLS_dec (alpha : t) :
-  {alpha = 0} +
+Finally, the nature of any ordinal is decidable(inside this notation system) :
+
+\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}}
+\begin{Coqsrc}
+  Definition ZeroLimitSucc_dec {A:Type}{lt: relation A}
+           {compare : A -> A -> comparison}
+           {on : ON lt compare} :=
+  forall alpha,
+    {Least alpha} +
     {Limit alpha} +
-  {beta : t | Successor alpha beta}.
+    {beta: A | Successor alpha beta}.
+\end{Coqsrc}
+
+\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
+
+\begin{Coqsrc}
+Definition Zero_limit_succ_dec : ZeroLimitSucc_dec.
 \end{Coqsrc}
@@
 Instance Omega2 : ON  lt compare.
+\end{Coqsrc}
+
+\subsection{Arithmetic of \texorpdfstring{$\omega^2$}{omega^2}}
+
+\subsubsection{Addition}
+
+We can define on \texttt{ON\_omega2} an addition which extends the addition on \texttt{nat}.
+
+\begin{Coqsrc}
+Definition  plus (alpha beta : t) : t :=
+  match alpha,beta with
+  | (0, b), (0, b') => (0, b + b')
+  | (0,0), y  => y
+  | x, (0,0)  => x
+  | (0, b), (S n', b') => (S n', b')
+  | (S n, b), (S n', b') => (S n + S n', b')
+  | (S n, b), (0, b') => (S n, b + b')
+   end.
+
+Infix "+" := plus : o2_scope.
+\end{Coqsrc}
+
+Please note that this operation is not commutative:
+
+\begin{Coqsrc}
+Example non_commutativity_of_plus :  omega + 3 <> 3 + omega.
+Proof.
+  cbn.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+1 subgoal (ID 237)
+
+  ============================
+(1, 3) <> omega.
+\end{Coqanswer}
+
+\begin{Coqsrc}
+ discriminate.
+Qed.
+\end{Coqsrc}
+
+\subsubsection{Multiplication}
+
+The restriction of ordinal multiplication to the segment $[0,\omega^2)$ is not a total function.
+For instance $\omega\times\omega= \omega^2$ is outside the set of represented values.
+Nevertheless, we can define two operations mixing natural numbers and ordinals.
+
+\begin{Coqsrc}
+(** multiplication of an ordinal and a natural number *)
+
+Definition mult_fin  (alpha : t) (p : nat): t :=
+  match alpha, p with
+ |  (0,0), _  => zero
+ |  _, 0 => zero
+ |  (0, n), p => (0, n * p)
+ |  ( n, b),  n' => ( n *  n', b)
+               end.
+Infix "*" := mult_fin : o2_scope.
+
+(** multiplication of  a natural number and an ordinal *)
+
+Definition fin_mult (n:nat)(beta : t) : t :=
+  match n, beta with
+ |  0, _  => zero
+ |  _, (0,0) => zero
+ |   n , (0,n') => (0, (n*n')%nat)
+ |  n, (n',p') => (n', (n * p')%nat)
+ end.
+
+Example e1 : fin_mult 3 (omega * 7 + 15) = omega * 7 + 45.
+Proof. reflexivity. Qed.
+
+Example e2 :  omega * 2 + 2 + omega * 3 + 10 = omega * 5 + 10.
+Proof. reflexivity. Qed.
+\end{Coqsrc}
+
+Multiplication with a finite ordinal and addition are related through the following lemma:
+
+\begin{Coqsrc}
+Lemma unique_decomposition alpha : 
+    exists! i j: nat,  alpha = omega * i + j.
+\end{Coqsrc}
+
+\subsection{A proof a termination using \texorpdfstring{$\omega^2$}{omega^2}}
+
+Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily the function which merges two
+
+\begin{Coqsrc}
+(* adapted from Pascal Manoury et al. *)
+
+Require Import Coq.Program.Wf.
+
+Section Merge.
+
+  Variable A: Type.
+
+  Local Definition m (p : list A * list A) :=
+    omega * length (fst p) + length (snd p).
+
+  Program Fixpoint  merge  (ltb: A -> A -> bool)
+          (xys: list A * list A)
+          {wf (m_lt m) xys} :
+    list A :=
+    match xys with
+      (nil, ys) => ys
+    | (xs, nil) => xs
+    | (x :: xs, y :: ys) =>
+      if ltb x y then x :: merge  ltb (xs, (y :: ys))
+      else y :: merge  ltb ((x :: xs), ys)
+    end.
+
+  Next Obligation.
+    unfold m, m_lt; destruct xs; simpl; left; lia.
+  Defined.
+
+  Next Obligation.
+    unfold m, m_lt ;cbn; destruct ys; simpl; right; lia.
+  Defined.
+
+End Merge.
 \end{Coqsrc}
@@
 \end{Coqsrc}
+
+\index{Projects}
+\begin{project}
+Prove that \texttt{Omega\_plus\_Omega} cannot be a sub-notation of \texttt{Omega}.
+\end{project}
 

### 2020-09-16 doc/chapter-ordinals.tex
minor modifications

@@
 \begin{Coqsrc}
-Definition m_lt {A:Type}{lt: relation A}
+Definition measure_lt {A:Type}{lt: relation A}
             {compare : A -> A -> comparison}
@@
             (m : B -> A) :
-  well_founded (m_lt m). 
+  well_founded (measure_lt m). 
 \end{Coqsrc}
@@
 \subsection{A proof a termination using \texorpdfstring{$\omega^2$}{omega^2}} 
-
-Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily the function which merges two
+\label{sect:merge-example}.
+
+Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily the function which merges two lists.
 
@@
 
-Require Import Coq.Program.Wf.
+Require Import Coq.Program.Wf List.
+Require Import FunInd Recdef.
 
@@
 
-  Program Fixpoint  merge  (ltb: A -> A -> bool)
+  Function  merge  (ltb: A -> A -> bool)
           (xys: list A * list A)
-          {wf (m_lt m) xys} :
+          {wf (measure_lt m) xys} :
     list A :=
@@
 
-  Next Obligation.
-    unfold m, m_lt; destruct xs; simpl; left; lia.
+  - intros.  unfold m, measure_lt;  cbn; destruct xs0; simpl; left; lia.
+  - intros;    unfold m, measure_lt ;cbn; destruct ys0; simpl; right; lia.
+   - auto.
   Defined.
 
-  Next Obligation.
-    unfold m, m_lt ;cbn; destruct ys; simpl; right; lia.
-  Defined.
-
 End Merge.
+
+Goal forall l,  merge nat Nat.leb (nil, l) = l.
+  intro; now rewrite merge_equation.
+Qed.
 \end{Coqsrc}

### 2020-09-18 doc/chapter-hydras.tex
modification in *.tex

--- 2020-09-02 latex/chapter-hydras.tex
@@
 
-\index{To do}
+%\index{To do}
 
 \begin{todo}
-Look for an existing library on trees with nodes of arbitrary arity, iin order to replace  this ad-hoc type with something more generic.
+Look for an existing library on trees with nodes of arbitrary arity, in order to replace  this ad-hoc type with something more generic.
 \end{todo}
@@
 
-\subsubsection{Lexicographic order on \texorpdfstring{nat*nat}{nat*nat}}
-\label{omega2-case}
-
-We  prove now that even the type \texttt{nat * nat}, provided with the lexicographic product of \texttt{(nat,<)} by itself is too simple for proving the termination of all hydra battles.
-This impossibility result will prevent us from considering measures like the following one:
-
-  \begin{Coqsrc}
-  Let m h = (height h, hsize h).
-  \end{Coqsrc}
-
-The proof we are going to develop has exactly the same structure as in Section~\ref{omega-case}.
- Nevertheless, the proof of technical  lemmas is a little more complex, due to 
- the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
-Consider for instance that there exists an infinite number of pairs between
-$(1,0)$ and $(2,0)$.
-
-\begin{remark}
-  The order structure we consider in this section is also known as the ordinal
-  $\omega^2$. We identify any pair $(i,j)\in \mathbb{N}\times\mathbb{N}$ with the ordinal $\omega\times i + j$. Thus the three kinds of ordinals in $\omega^2$
-  are represented as follows:
-  \begin{description}
-  \item[null ordinal] : the pair $(0,0)$
-  \item[successor ordinal] : any pair $(i,j)$ where $j>0$
-  \item[limit ordinal] : any pair $(i,0)$ where $i>0$
-  \end{description}
-\end{remark}
-
-The detailed  proof script is in the file \url{../src/html/hydras.Hydra.Omega2_Small.html}.
-
-\subsubsection{Preliminaries}
-Let us assume there is a variant from \texttt{Hydra} into \texttt{nat*nat} (with the
-   lexicographic ordering)  for proving the   termination of all hydra battles.
-
-\vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
-
-\begin{Coqsrc}
-  Section Impossibility_Proof.
-
-  Let t := (nat * nat)%type.
-
-  (** non-dependent lexicographic strict ordering on nat*nat *)
-
-  Let lt2 : relation t := lexico Peano.lt Peano.lt.
-
-  Infix "o<" := lt2.
-
-  (** reflexive closure of lt2 *)
-  Let le2 := clos_refl _ lt2.
-  Infix "o<=" := le2.
-
-  Variable m : Hydra -> t.
-
-  Context (Hvar : Hvariant lt2_wf free m).
-\end{Coqsrc}
-
-\label{remark-infix}
-\begin{remark}
-We use in \coq{} scripts the infix notations \verb@o<@ and \verb@o<=@ for
-the orders $<$ and $\leq$ on ordinals. This convention allows us to define
-implicit coercions from \texttt{nat} into various types for representing ordinals.
-For instance a proposition like \verb@ 6 o< 8@ will be interpreted as 
-``the finite ordinal 6 is strictly less than the finite ordinal 8'', whereas 
-\verb@ 6 o< 8@ is a statement about natural numbers. Since many lemmas talk about both ordinals and natural numbers, we wished to avoid too many scope changes inside propositions.
-\end{remark}
-
-Let us follow the same pattern as in Sect.~\ref{omega-case}.
-First, we define an injection from type \texttt{t} into \texttt{Hydra}, by
- associating to  any pair $(i,j)$ the hydra with $i$ branches of length $2$ and
-$j$ branches of length $1$.
-
-%% revenir ici
-
-\vspace{4pt}
-\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
-
-\begin{Coqsrc}
-  Let iota (p: t) := 
-    node (hcons_mult (hyd1 head) (fst p)
-                     (hcons_mult head (snd p) hnil)).
-\end{Coqsrc}
-
-For instance, Figure~\vref{fig:essai2} shows the hydra associated to the pair
-$(3,5)$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-\node (foot) at (6,0) {$\bullet$};
-\node (N1) at (1,3) {$\bullet$};
-\node (N2) at (3,3) {$\bullet$};
-\node (N3) at (5,3) {$\bullet$};
-\node (N4) at (8,3) {$\Smiley[2][green]$};
-\node (N5) at (11,3) {$\Smiley[2][green]$};
-\node (N6) at (14,3) {$\Smiley[2][green]$};
-\node (N7) at (17,3){$\Smiley[2][green]$};
-\node (N8) at (20,3){$\Smiley[2][green]$};
-\node  (N9) at (0,5) {$\Smiley[2][green]$};
-\node (N10) at (2,5) {$\Smiley[2][green]$};
-\node (N11) at (4,5) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N5);
-\draw (foot) -- (N6);
-\draw (foot) to [bend right=10] (N7);
-\draw (foot) to [bend right=15] (N8);
-\draw (N1) to [bend left=10] (N9);
-\draw (N2) -- (N10);
-\draw (N3) -- (N11);
-\end{tikzpicture}
-\caption{\label{fig:essai2}
-The hydra $\iota(3,5)$}
-\end{figure}
-
-Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:h-omega2-small}}
- The hydra \texttt{big\_h}.
-\end{figure}
-\begin{Coqsrc}
-   Let big_h := hyd1 (hyd2 head head).  
- \end{Coqsrc}
-
- In a second step, we build a ``smaller'' hydra.
-
-\begin{Coqsrc}
-   Let small_h := iota (m big_h).
-\end{Coqsrc}
-
-Like in Sect.~\ref{omega-case}, we prove the double inequality \texttt{m big\_h o<= m small\_h o< m big\_h}, which is impossible.
-
-\subsubsection{Proof of the inequality \texttt{m small\_h o< m big\_h}}
-
-For proving the inequality  \texttt{m\_lt: m small\_h o< m big\_h}, it suffices to
-build a battle transforming \texttt{big\_h} into \texttt{small\_h}.
-
-First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h}
-into the pair $(i,j)$.
-If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
-If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the folowing result.
-
-\begin{Coqsrc}
- Lemma big_to_small: big_h -+-> small_h.
-\end{Coqsrc}
-
-Since $m$ is a variant, we infer the following inequality:
-
-\begin{Coqsrc}
-Corollary m_lt : m small_h o< m big_h.
-\end{Coqsrc}
-
-\subsubsection{Proof of the inequality \texttt{m big\_h o<= m small\_h} }
-
-The proof of the inequality \texttt{m big\_h o<= m small\_h} is quite more complex than in Sect~\ref{omega-case}.  If we consider some pair $(i,j)$, where $i>0$, there exists an infinite number of
-pairs stricly less than $(i,j)$, and there exists an infinite number of battles that start from
-$\iota(i,j)$. In effect, at any configuration $\iota(k,0)$, the hydra can freely chose any replication number. Intuitively, the measure of such a hydra must be large enough for taking into account
-all the possible battles issued from that hydra.
-Let us give more technical details.
-
-\begin{itemize}
-\item The proof of the lemma \texttt{m\_ge : forall p : t,   p o<= m (iota p)} uses well-founded induction on $p$, and not structural induction on natural numbers
-
-\item For any pair $p$, we have to distinguish between three cases, according to the value of $p$'s components.
-  \begin{itemize}
-  \item $p=(0,0)$
-  \item $p=(i,0)$, where $i>0$\,: $p$ corresponds to a limit ordinal
-  \item $p=(i,j)$, where $j>0$\,: $p$ is the successor of $(i,j-1)$.
-  \end{itemize}
-\end{itemize}
-
-Before starting the proof, we have to express the notion of \emph{limit} in terms of
-least upper bounds, through the following logical equivalence.
-
-\vspace{4pt}
-\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#limit_is_lub}{Hydra.Omega2\_Small.v}}.
-
-\begin{Coqsrc}
-Lemma limit_is_lub : forall i p, 
-   (forall j, (i,j) o< p) <-> (S i, 0) o<= p.  
-\end{Coqsrc}
-
-Let us define the notion of elementary ``step'' of decreasing sequences in
-\texttt{t}
-
-\begin{Coqsrc}
-Inductive step : t -> t -> Prop :=
-| succ_step : forall i j,  step (i, S j) (i, j)
-| limit_step : forall i j, step (S i, 0) (i, j).
-\end{Coqsrc}
-
-The following lemma establishes a correspondance between the relation
-\texttt{step} and hydra battles.
-
-\begin{Coqsrc}
-Lemma step_to_battle : forall p q, step p q -> iota p -+-> iota q.
-\end{Coqsrc}
-
-\index{Maths!Transfinite induction}
-
-Thus, starting from any inequality $q < p$ on type \texttt{t}, we can build 
-by transfinite induction over \texttt{p} a battle 
-that transforms the hydra $\iota(p)$ into $\iota(q)$.
-
-\vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}
-
-\begin{Coqsrc}
-Lemma m_ge : forall p : t,   p o<= m (iota p).
-Proof.
-  intro p ; pattern p;
-    apply  well_founded_induction with 
-               (R := lt2) (1:= wf_lexico lt_wf lt_wf);
-     intros (i,j) IHij (* rest of proof skipped *)
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  i, j : nat
-  IHij : forall y : t, y o< (i, j) -> y o<= m (iota y)
-  ============================
-  (i, j) o<= m (iota (i, j)) 
-\end{Coqanswer}
-
-Then we have  three cases to consider, according to the values of $i$ and $j$.
-\begin{itemize}
-\item If $p=(0,0)$ then obviously, $\iota(p)\geq p = (0,0)$
-\item If  $p=(i+1,0)$ for some $i\in\mathbb{N}$, we
- remark  that $p$ is strictly greater than any pair $ (i, j)$, where $j$ 
-is any natural number.
-
-Applying the battle rules, for any $j$, we have $\iota(i+1,j)  {\round} \iota(i, j) $, thus $m(\iota(p)) > m(\iota(i,j)$ since  $m$ is assumed to be a variant.
-
-Applying the induction hypothesis, we get the inequality
- $ m(\iota(i,j)) \geq (i,j)$ for any $j$.
-
-Thus, $m(\iota(p)) > (i,j)$ for any $j$.
-Applying the lemma \texttt{limit\_is\_lub}, we get  the inequality
-$\iota(i+1,0)\geq (i+1,0)$
-
-\item If $p=(i,j+1)$ with $j\in\mathbb{N}$, we have  $\iota(p)  {\round} \iota(i, j) $,
-hence $m(\iota(p))> m(\iota(i,j)) \geq (i,j)$, thus $m(\iota(p))\geq (i,j+1)=p$
-
-\end{itemize}
-
-\subsubsection{End of the proof}
-
-Since $<$ is a strict order (irreflexive  and transitive) on \texttt{nat*nat}, we can conclude that there is no
-variant for termination on the lexicographic square of $(\mathbb{N},<)$.
-
-\begin{enumerate}
-\item From \texttt{m\_lt}, we infer the strict inequality 
-\texttt{m small\_h o< m big\_h}
-\item from \texttt{m\_ge}, we get \texttt{m big\_h o<= m (iota (m big\_h)) = m small\_h} 
-\end{enumerate}
-
-\vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
-
-\begin{Coqsrc}
-Theorem Impossible : False.
-Proof.
-  destruct (StrictOrder_Irreflexive  (m big_h)).
-  apply le2_lt2_trans with (m small_h).
-  -  unfold small_h;  apply m_ge.
-  -  apply m_lt. 
-Qed.
-
-End Impossibility_Proof.
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into \texttt{nat*nat} for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\begin{remark}
-In Chapter~\ref{ks-chapter}, we will prove a much more general  theorem
-than the two previous results about $\mathbb{N}$ and $\mathbb{N}\times\mathbb{N}$. The proof of that general  result will share the same structure, but will 
-require a lot of technical results. 
-\end{remark}
-
-\index{Exercises}
-\begin{exercise}
-
-\label{sec:orgheadline63}
-Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.
-
-\begin{itemize}
-\item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
-  lexicographic ordering
-\item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.
-
-For instance, the following inequality holds:
-\[\langle 4,3,3,3,3,3,3,2,2,2 \rangle\,<\,\langle 4,4,2 \rangle\]
-\end{itemize}
-
-\end{exercise}
-
 \subsubsection{Conclusion}

### 2020-09-18 doc/chapter-ordinals.tex
update the index
modification in *.tex

@@
 
-\subsection{The ordinal \texorpdfstring{$\omega^2$}{omega^2}}
+\section{The ordinal \texorpdfstring{$\omega^2$}{omega^2}}
 
@@
 Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily the function which merges two lists.
+
+\index{Coq!Commands!Function}
 
@@
 
-  - intros.  unfold m, measure_lt;  cbn; destruct xs0; simpl; left; lia.
-  - intros;    unfold m, measure_lt ;cbn; destruct ys0; simpl; right; lia.
+  - intros; unfold m, measure_lt; cbn; destruct xs0; simpl; left; lia.
+  - intros; unfold m, measure_lt; cbn; destruct ys0; simpl; right; lia.
    - auto.
@@
 
+\subsection{Yet another  proof of impossibility}
+\label{omega2-case}
+
+In Sect.~\vref{omega-case}, we proved that there exists no variant towards \texttt{nat}
+(\emph{i.e.} the ordinal $\omega$) for proving the termination of all hydra battles.
+We  prove now that even the ordinal $\omega^2$ is also insufficient for this purpose.
+
+The proof we are going to develop has exactly the same structure as in Section~\ref{omega-case}.
+ Nevertheless, the proof of technical  lemmas is a little more complex, due to 
+ the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
+Consider for instance that there exists an infinite number of ordinals  between
+$\omega$ and $\omega\times 2$.
+
+The detailed  proof script is in the file \url{../src/html/hydras.Hydra.Omega2_Small.html}.
+
+\subsubsection{Preliminaries}
+Let us assume there is a variant from \texttt{Hydra} into $\omega^2$  for proving the   termination of all hydra battles.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
+
+\begin{Coqsrc}
+  Section Impossibility_Proof.
+
+ Variable m : Hydra -> ON_Omega2.t.
+
+ Context (Hvar : Hvariant lt_wf free m).
+\end{Coqsrc}
+
+Let us follow the same pattern as in Sect.~\ref{omega-case}.
+First, we define an injection from type \texttt{t} into \texttt{Hydra}, by
+ associating to  each ordinal $\omega\times i+ j = (i,j)$ the hydra with $i$ branches of length $2$ and
+$j$ branches of length $1$.
+
+%% revenir ici
+
+\vspace{4pt}
+\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
+
+\begin{Coqsrc}
+Let iota (p: ON_Omega2.t) := 
+    node (hcons_mult (hyd1 head) (fst p)
+                     (hcons_mult head (snd p) hnil)).
+  \end{Coqsrc}
+
+For instance, Figure~\vref{fig:essai2} shows the hydra associated to the ordinal 
+$(3,5)$, a.k.a. $\omega\times 3 + 5$.
+
+\begin{figure}[htb]
+\centering
+\begin{tikzpicture}[very thick, scale=0.4]
+\node (foot) at (6,0) {$\bullet$};
+\node (N1) at (1,3) {$\bullet$};
+\node (N2) at (3,3) {$\bullet$};
+\node (N3) at (5,3) {$\bullet$};
+\node (N4) at (8,3) {$\Smiley[2][green]$};
+\node (N5) at (11,3) {$\Smiley[2][green]$};
+\node (N6) at (14,3) {$\Smiley[2][green]$};
+\node (N7) at (17,3){$\Smiley[2][green]$};
+\node (N8) at (20,3){$\Smiley[2][green]$};
+\node  (N9) at (0,5) {$\Smiley[2][green]$};
+\node (N10) at (2,5) {$\Smiley[2][green]$};
+\node (N11) at (4,5) {$\Smiley[2][green]$};
+\draw (foot) to [bend left=10] (N1);
+\draw (foot) -- (N2);
+\draw (foot) -- (N3);
+\draw (foot) -- (N4);
+\draw (foot) -- (N5);
+\draw (foot) -- (N6);
+\draw (foot) to [bend right=10] (N7);
+\draw (foot) to [bend right=15] (N8);
+\draw (N1) to [bend left=10] (N9);
+\draw (N2) -- (N10);
+\draw (N3) -- (N11);
+\end{tikzpicture}
+\caption{\label{fig:essai2}
+The hydra $\iota(\omega\times 3+5)$}
+\end{figure}
+
+Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).
+
+\begin{figure}[htb]
+\centering
+\begin{tikzpicture}[very thick, scale=0.5]
+\node (foot) at (2,0) {$\bullet$};
+\node (N1) at (2,2) {$\bullet$};
+\node (N2) at (3,4) {$\Smiley[2][green]$};
+\node (N3) at (1,4) {$\Smiley[2][green]$};
+\draw (foot) -- (N1);
+\draw (N1) to [bend right =15] (N2);
+\draw (N1) to  [bend left=15](N3);
+\end{tikzpicture}
+\caption{\label{fig:h-omega2-small}}
+ The hydra \texttt{big\_h}.
+\end{figure}
+
+\begin{Coqsrc}
+   Let big_h := hyd1 (hyd2 head head).  
+\end{Coqsrc}
+
+ In a second step, we build a ``smaller'' hydra.
+
+\begin{Coqsrc}
+   Let small_h := iota (m big_h).
+\end{Coqsrc}
+
+Like in Sect.~\ref{omega-case}, we prove the double inequality \texttt{m big\_h o<= m small\_h o< m big\_h}, which is impossible.
+
+\subsubsection{Proof of the inequality \texttt{m small\_h o< m big\_h}}
+
+In order to prove the inequality  \texttt{m\_lt: m small\_h o< m big\_h}, it suffices to
+build a battle transforming \texttt{big\_h} into \texttt{small\_h}.
+
+First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h} as $(i,j)$.
+If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
+If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the following result.
+
+\begin{Coqsrc}
+Lemma big_to_small: big_h -+-> small_h.
+\end{Coqsrc}
+
+Since $m$ is a variant, we infer the following inequality:
+
+\begin{Coqsrc}
+Corollary m_lt : m small_h o< m big_h.
+\end{Coqsrc}
+
+\subsubsection{Proof of the inequality \texttt{m big\_h o<= m small\_h} }
+
+The proof of the inequality \texttt{m big\_h o<= m small\_h} is quite more complex than in Sect~\ref{omega-case}.
+If we consider some ordinal $\alpha=(i,j)$, where $i>0$, there exists an infinite number of
+ordinals stricly less than $\alpha$, and there exists an infinite number of battles that start from
+$\iota(\alpha)$.
+In effect, at any configuration $\iota(k,0)$, where $k>0$, the hydra can freely chose any replication number.
+Intuitively, the measure of such a hydra must be large enough for taking into account
+all the possible battles issued from that hydra.
+Let us give more technical details.
+
+\begin{itemize}
+\item The proof of the lemma \texttt{m\_ge : m big\_h  o<= m small\_h} uses well-founded induction on \texttt{big\_h}.
+
+\item For any pair $p$, we have to distinguish between three cases, according to the value of $p$'s components.
+  \begin{itemize}
+  \item $p=(0,0)$
+  \item $p=(i,0)$, where $i>0$\,: $p$ corresponds to a limit ordinal
+  \item $p=(i,j)$, where $j>0$\,: $p$ is the successor of $(i,j-1)$.
+  \end{itemize}
+\end{itemize}
+
+Let us define the notion of elementary ``step'' of decreasing sequences in
+\texttt{t}
+
+\begin{Coqsrc}
+Inductive step : t -> t -> Prop :=
+| succ_step : forall i j,  step (i, S j) (i, j)
+| limit_step : forall i j, step (S i, 0) (i, j).
+\end{Coqsrc}
+
+The following lemma establishes a correspondance between the relation
+\texttt{step} and hydra battles.
+
+\begin{Coqsrc}
+Lemma step_to_battle : forall p q, step p q -> iota p -+-> iota q.
+\end{Coqsrc}
+
+\index{Maths!Transfinite induction}
+
+Thus, starting from any inequality $q < p$ on type \texttt{t}, we can build 
+by transfinite induction over \texttt{p} a battle 
+that transforms the hydra $\iota(p)$ into $\iota(q)$.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}
+
+\begin{Coqsrc}
+Lemma m_ge : forall p : t,   p o<= m (iota p).
+Proof.
+  unfold small_h; pattern (m big_h) .   
+   apply  well_founded_induction with (R := lt) (1:= lt_wf).
+  intro p ; pattern p;
+    apply  well_founded_induction with 
+               (R := lt2) (1:= wf_lexico lt_wf lt_wf);
+     intros (i,j) IHij. 
+\end{Coqsrc}
+
+\begin{Coqanswer}
+  i, j : nat
+  IHij : forall y : t, y o< (i, j) -> y o<= m (iota y)
+  ============================
+  (i, j) o<= m (iota (i, j)) 
+\end{Coqanswer}
+
+Then we have  three cases to consider, according to the values of $i$ and $j$.
+\begin{itemize}
+\item If $p=(0,0)$ then obviously, $\iota(p)\geq p = (0,0)$
+\item If  $p=(i+1,0)$ for some $i\in\mathbb{N}$, we
+ remark  that $p$ is strictly greater than any pair $ (i, j)$, where $j$ 
+is any natural number.
+
+Applying the battle rules, for any $j$, we have $\iota(i+1,j)  {\round} \iota(i, j) $, thus $m(\iota(p)) > m(\iota(i,j)$ since  $m$ is assumed to be a variant.
+
+Applying the induction hypothesis, we get the inequality
+ $ m(\iota(i,j)) \geq (i,j)$ for any $j$.
+
+Thus, $m(\iota(p)) > (i,j)$ for any $j$.
+Applying the lemma \texttt{limit\_is\_lub}, we get  the inequality
+$\iota(i+1,0)\geq (i+1,0)$
+
+\item If $p=(i,j+1)$ with $j\in\mathbb{N}$, we have  $\iota(p)  {\round} \iota(i, j) $,
+hence $m(\iota(p))> m(\iota(i,j)) \geq (i,j)$, thus $m(\iota(p))\geq (i,j+1)=p$
+
+\end{itemize}
+
+\begin{Coqsrc}
+  (* ... *)
+Qed.
+\end{Coqsrc}
+
+\subsubsection{End of the proof}
+
+Since $<$ is a strict order (irreflexive  and transitive) on \texttt{nat*nat}, we can conclude that there is no
+variant for termination on the lexicographic square of $(\mathbb{N},<)$.
+
+\begin{enumerate}
+\item From \texttt{m\_lt}, we infer the strict inequality 
+\texttt{m small\_h o< m big\_h}
+\item from \texttt{m\_ge}, we get \texttt{m big\_h o<= m (iota (m big\_h)) = m small\_h} 
+\end{enumerate}
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
+
+\begin{Coqsrc}
+Theorem Impossible : False.
+Proof.
+  destruct (StrictOrder_Irreflexive  (m big_h)).
+  apply le2_lt2_trans with (m small_h).
+  -  unfold small_h;  apply m_ge.
+  -  apply m_lt. 
+Qed.
+
+End Impossibility_Proof.
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+Prove that there exists no variant $m$ from \texttt{Hydra} into $\omega^2$ for proving
+    the  termination of all \emph{standard} battles.
+\end{exercise}
+
+\begin{remark}
+In Chapter~\ref{ks-chapter}, we  prove a generalization of the impossibility lemmas of
+Sect.~\ref{omega-case} and this section, with the same proof structure, but with much more 
+complex technical details.
+ \end{remark}
+
+\index{Exercises}
+\begin{exercise}
+
+\label{sec:orgheadline63}
+Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.
+
+\begin{itemize}
+\item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
+  lexicographic ordering
+\item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.
+
+For instance, the following inequality holds:
+\[\langle 4,3,3,3,3,3,3,2,2,2 \rangle\,<\,\langle 4,4,2 \rangle\]
+\end{itemize}
+
+\end{exercise}
+
 \section{Comparing Ordinal Notations}
@@
 We require also  that the comparison functions of the two notation systems are compatible.
+
+If \texttt{OB} is presumed to be correct, the we can consider that \texttt{OA} ``inherits'' its correcteness from the bigger notation system \texttt{OB}.
 
@@
   {
-  subon_compare :forall x y : A,  compareB (iota x) (iota y) =
+  subon_compare: forall x y : A,  compareB (iota x) (iota y) =
                                  compareA x y;
@@
 \begin{Coqsrc}
-Class ON_for `(alpha : Schutte_basics.Ord)
+Class ON_correct `(alpha : Schutte_basics.Ord)
      `(OA : @ON A ltA  compareA)
       (iota : A -> Schutte_basics.Ord) :=
-  { ON_for_inj : forall a, Schutte_basics.lt (iota a) alpha;
-    ON_for_onto : forall beta, Schutte_basics.lt beta alpha ->
+  { ON_correct_inj : forall a, Schutte_basics.lt (iota a) alpha;
+    ON_correct_onto : forall beta, Schutte_basics.lt beta alpha ->
                                 exists b, iota b = beta;
@@
 Instance Epsilon0_correct :
-  ON_for epsilon0 Epsilon0  (fun alpha => inject (cnf alpha)).
+  ON_correct epsilon0 Epsilon0  (fun alpha => inject (cnf alpha)).
 \end{Coqsrc}
@@
 
-{\it \textbf{Note:s} This property reflects the equality $i+\omega=\omega$ we will prove in larger notation systems, as well as in Schütte's model}
+{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$ we will prove in larger notation systems, as well as in Schütte's model}
 

### 2020-09-23 doc/hydras.tex
minor modifications

@@
 
-\include{chapter-introduction}
-\include{chapter-hydras}
-\include{chapter-ordinals}
-\include{chapter-termination}
-\include{chapter-ks}
-\include{chapter-large-sets}
-\include{chapter-schutte}
+%-------------------------------------------------------------------
+
+\chapter{Introduction}
+
+\vspace{16pt}
+
+Proof assistants are excellent tools for exploring the structure of mathematical proofs,
+studying  which hypotheses are really needed, and which proof patterns are useful and/or
+necessary. Since the development of a theory is represented as a bunch of computer files,
+everyone is able to read the proofs with an arbitrary level of detail, or to play with the theory by building alternate proofs or definitions.
+
+Among all the theorems proved with the help of proof assistants like \coq{}, \isabelle{}, \hol{}, etc.,
+several statements and proofs  share some interesting features:
+\begin{itemize}
+\item Their statements are easy to understand, even by non-mathematicians
+\item Their proof requires some non-trivial mathematical tools
+\item Their mechanization on computer presents some methodological interest.
+\end{itemize}
+
+This is obviously the case of the four-color theorem~\cite{fourcolors}  and the Kepler conjecture~\cite{flyspeck2015}.
+We do not mention impressive works like the proof of the odd-order theorem ~\cite{oddorderthm}, since understanding its statement requires a quite good mathematical culture.
+
+Hydra games (a.k.a. \emph{Hydra battles}) appeared in an article published in 1982 by two mathematicians:
+L. Kirby and J. Paris~\cite{KP82}: \emph{Accessible Independence Results for Peano Arithmetic}. 
+Although the mathematical contents of this 
+paper are quite advanced, the rules of hydra battles are very easy to understand.
+There are now several sites on Internet where you can find tutorials on hydra games, together with simulators you can play with.
+See, for instance, the page written by Andrej Bauer~\cite{bauer2008}.
+
+Hydra battles, as well as Goodstein Sequences~\cite{goodstein_1944, KP82}
+are a nice way to present complex termination problems.
+The article by Kirby and Paris presents a proof of termination
+based on ordinal numbers, as well as a proof that this termination is not
+provable in Peano arithmetic. In the book dedicated to 
+J.P. ~Jouannaud \cite{HommageJPJ}, N.~Dershowitz and G.~Moser  give a thorough survey on this topic~\cite{Dershowitz2007}.
+
+Here, we present a development  for the \coq{} proof assistant, after the work of Kirby and Paris. This formalization contains the following main parts:
+
+\begin{itemize}
+\item Representation in \coq{} of hydras and hydra battles
+\item A proof that every battle is finite and won by Hercules.
+This proof is based on a \emph{variant} which maps any hydra to an ordinal strictly less than $\epsilon_0$ and is strictly decreasing along any battle.
+
+\item Using a combinatorial toolkit designed by J.~Ketonen and R.~Solovay~\cite{KS81}, we prove that, for any ordinal $\mu<\epsilon_0$, there exists no such variant mapping any hydra to an ordinal stricly less than $\mu$.
+Thus, the complexity of $\epsilon_0$ is really needed in the previous proof.
+
+\item We prove a relation between the length of a ``classic''  class of  battles
+and the Wainer-Hardy hierarchy of ``rapidly growing functions'' $H_\alpha$~\cite{Wainer1970}.
+The considered class of battles, which we call \emph{standard}  is the most considered one in the scientific  litterature(including popularization).
+\end{itemize}
+
+Simply put, this document tries to combines the scientific interest of two articles~\cite{KP82, KS81} and a book~\cite{schutte} with the playful activity of truly proving theorems.
+We hope  that such a work, besides exploring a nice piece of discrete maths, 
+will show how \coq{} and its standard library are well fitted to help us to understand some non-trivial mathematical developments, and also to experiment the constructive parts of  the proof through functional programming.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=11cm]{epsilon0.jpg}
+  \caption{Discrete mathematics and formal proofs}
+  \label{fig:gamma0}
+\end{figure}
+ We also hope to provide a little clarification on infinity (both potential and actual) through the notions of function, computation, limit,
+ types and proofs.
+
+\section{Remarks}
+
+\subsection{Difference from Kirby and Paris's work}
+In~\cite{KP82}, Kirby and Paris showed  that there is no proof of termination of all hydra battles in Peano Arithmetic (PA).
+Since we are used to writing proofs in higher order logic, the restriction to PA was quite unnatural for us.
+So we chosed to prove another statement without any reference to PA, by considering a class of proofs indexed by ordinal numbers upto $\epsilon_0$.
+
+\subsection{Trust in our proofs?}
+\label{sect:trust-in-proofs}
+
+Unlike mathematical literature, where definitions and proofs are spread over many articles and books,
+the whole proof is now inside your computer. It is composed of the \texttt{.v} files you downloaded and 
+parts of \coq's standard library.
+Thus, there is no ambiguity in our definitions and the premises of the theorems.
+Furthermore, you will be able to navigate through the development, using your favourite text editor or IDE, and some commands like \texttt{Search}, \texttt{Locate},  etc.
+
+\subsection{Assumed redundancy}
+
+It may often happen that several definitions of a given concept, or several proofs of a given theorem are possible.
+If all the versions present some interest, we will make them available, since each one may be of some methodological
+interest (by illustrating some tactic of proof pattern, for instance).
+We may use \coq's module system to make several proofs of a givent theorem co-exists in our libraries (see also Sect~\vref{sect:alt-proofs}).
+After some discussions of the pros and cons of each solution, we will choose to develop only one of them, leaving the others  as exercises or projects (i.e., big or difficult exercises).
+In order to discuss which assumptions are really needed for proving a theorem, we will also present 
+several aborted proofs.
+Of course, do not hesitate to contribute nice proofs or alternative definitions !
+
+It may also happen that some proof looks to be useless, because the proven theorem is a trivial consequence of another (proven too) result.
+For instance, let us consider the three following statements:
+\begin{enumerate}
+\item There is no measure into $\mathbb{N}$ for proving the termination of all hydra battles (Sect~\vref{omega-case}).
+\item There is no measure into $[0,\omega^2)$ for proving the termination of all hydra battle (Sect~\vref{omega2-case}).
+\item There is no measure into $[0,\mu)$ for proving the termination of all hydra battles, for any $\mu<\epsilon_0$ (Sect\vref{sec:free-battles-case}).
+\end{enumerate}
+
+Obviously, the third theorem implies the second one, which implies the first one.
+So, theoretically, a library would contain only a proof of $(3)$ and remarks for $(2)$ and $(1)$.
+But we found it interesting to make all the three proofs available, allowing the reader to compare their common structure and notice their technical differences.
+In particular, the proof of $(3)$ uses several non-trivial combinatorial properties of ordinal numbers~\cite{KS81}, whilst $(1)$ and $(2)$ use simple properties of $\mathbb{N}$ and $\mathbb{N}^2$.
+
+\subsection{About Logic}
+
+Most of the proofs we present are \emph{constructive}.
+Whenever possible, we provide the user with an associated function, which she or he can apply in \gallina{} or \ocaml{} in order to get a ``concrete'' feeling of the meaning of the considered theorem.
+For instance, in Chapter~\vref{chap:ketonen}, the notion of \emph{limit ordinal} is
+made more ``concrete'' thanks to a function \texttt{canon} which computes every item of a sequence which converges on a given limit ordinal $\alpha$.
+This simply typed function allows the user/reader to make her/his own experimentations.
+For instance, one can very easily compute the $42$-nd item of a sequence which converges towards $\omega^{\omega^\omega}$.
+
+Except in the \texttt{Schutte} library, dedicated to an axiomatic presentation of the set of countable ordinal numbers, all our development is axiom-free, and respects the rules of intuitionistic logic.
+Note that we also use the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777} in the definitition of  several rapidly growing hierarchy of functions, in Chap.~\ref{chap:alpha-large}.
+This plug-in imports several known-as-harmless  axioms.
+
+% \begin{Coqsrc}
+% FunctionalExtensionality.functional_extensionality_dep : 
+% forall (A : Type) (B : A -> Type) (f g : forall x : A, B x),
+% (forall x : A, f x = g x) -> f = g
+% \end{Coqsrc}
+
+\index{Coq!Commands!Print Assumptions}
+
+At any place of our development, you may use the command \texttt{Print Assumptions {\it ident}} in order to verify on which axiom the theorem {\it ident} may depend.
+The following example is extracted from
+Library~\href{../src/html/hydras.Epsilon0.F_alpha.html}{hydras.Epsilon0.F\_alpha}, where we use the \texttt{coq-equations} plug-in (see Sect.~\vref{sect:wainer}).
+
+\begin{Coqsrc}
+About F_zero_eqn.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+F_zero_eqn : forall i : nat, F_ Zero i = S i
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Print Assumptions F_zero_eqn. 
+\end{Coqsrc}
+
+\begin{Coqanswer}
+Axioms:
+FunctionalExtensionality.functional_extensionality_dep
+  : forall (A : Type) (B : A -> Type) (f g : forall x : A, B x),
+    (forall x : A, f x = g x) -> f = g
+Eqdep.Eq_rect_eq.eq_rect_eq
+  : forall (U : Type) (p : U) (Q : U -> Type) (x : Q p) (h : p = p),
+    x = eq_rect p Q x p h
+\end{Coqanswer}
+
+\subsection{Main references}
+
+In our development, we adapt the definitions and prove many theorems which
+we found in the following articles. 
+\begin{itemize}
+\item ``Accessible independence results for Peano arithmetic''  by Laurie Kirby and Jeff Paris~\cite{KP82}
+\item ''Rapidly growing Ramsey Functions'' by Jussi Ketonen and Robert Solovay~\cite{KS81}
+\item ``The Termite and the Tower'', by Will Sladek~\cite{Sladek07thetermite}
+\item Chapter V of ``Proof Theory'' by Kurt Schütte~\cite{schutte}
+\end{itemize}
+
+\paragraph*{Warning:}
+
+This document is \emph{not} an introductory text for \coq{}, and there are many aspects of this proof assistant that are not covered. 
+ The reader should already have some basic experience with the \coq{} system.
+The Reference Manual and several tutorials are available on \coq{} page~\cite{Coq}.
+First chapters of textbooks like \emph{Interactive Theorem Proving and Program Development}~\cite{BC04}, \emph{Software Foundations}~\cite{SF} or  \emph{Certified Programming with Dependent Types} ~\cite{chlipalacpdt2011} will give you the right background.
+
+\subsection{State of the development}
+The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on ordinal notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
+We added new material : axiomatic definition of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, equations, etc.
+
+We are now working in order to make clumsy proofs more readable, simplify definitions, 
+to ``factorize'' proofs as much as possible.
+
+Many ``todo''s or projects are put in this text, suggesting possible improvements.
+
+\subsection{Contributions}
+Many thanks to \'Evelyne Contejean and Florian Hatat for their contribution.
+\'Evelyne contributed libraries on the recursive path ordering (\emph{rpo}) for proving the well-foundedness of our representation of $\epsilon_0$ and $\Gamma_0$.
+Florian Hatat proved many useful lemmas on countable sets, which we used in our adaptation of Schütte's formalization of countable ordinals.
+
+\label{sec:orgheadline2}
+
+Any form of contribution  is welcome: correction of errors, improvement of
+Coq scripts, proposition of inclusion of new chapters, and generally any
+comment or proposition that would help us. The text contains several \emph{projects} which, when completed, may improve the present work.
+Please do not hesitate to bring your contribution!
+
+\subsection{Typographical conventions}
+
+Quotations from our  \coq{} source are displayed as follows:
+
+  \begin{Coqsrc}
+ Require Import Arith.
+
+ Definition square (n:nat) := n * n.
+
+ Lemma square_double : exists n:nat, n + n = square n.
+ Proof.
+    exists 2. 
+  \end{Coqsrc}
+
+Answers from \coq{} (including sub-goals, error messages, etc.) are displayed in slanted style
+with a different background color.
+
+ \begin{Coqanswer}
+ 1 subgoal, subgoal 1 (ID 5)
+
+  ============================
+   2 + 2 = square 2
+
+ \end{Coqanswer}
+
+ \begin{Coqsrc}
+   reflexivity.
+Qed.
+ \end{Coqsrc}
+
+\subsection{Active links}
+The  links which appear in this pdf  document lead are of three possible kinds of destination:
+\begin{itemize}
+\item Local links to the document itself,
+\item External links, mainly to \coq's page,
+\item Local links to pages generated by \texttt{coqdoc}. According to the current makefile,
+  we assume that the page generated from a library \texttt{XXX/YYY.v} is stored as
+the relative address \texttt{../src/html/hydras.XXX.YYY.html} (from the location of the pdf).
+Thus,  active links towards our \coq{} modules may be incorrect if you got this \texttt{pdf} document otherwise than by compiling the distribution available in
+\url{https://framagit.org/casteran/hydra-battles-in-coq}.
+
+\end{itemize}
+
+\subsection{Alternative or bad definitions}
+\label{sect:alt-proofs}
+Finally, we decided to include definitions or lemma statements, as well as tactics,  that lead to
+dead-ends or to too complex developments, with the following coloring.
+Bad definitions 
+ and encapsulation in modules called \texttt{Bad}, \texttt{Bad1}, etc.
+
+\begin{Coqbad}
+Module Bad.
+
+Definition double (n:nat)  := n + 2.
+
+Lemma lt_double : forall n:nat, n < double  n.
+Proof.
+   unfold double; omega.
+Qed.
+
+End Bad.
+\end{Coqbad}
+
+Likewise, alternative, but still unexplored definitions will be presented in modules
+\texttt{Alt}, \texttt{Alt1}, etc. Using these definitions is left as an implicit exercise.
+
+\begin{Coqalt}
+Module Alt.
+
+  Definition double (n : nat) := 2 * n.
+
+End Alt.
+\end{Coqalt}
+
+\begin{Coqsrc}
+Lemma alt_double_ok n : Nat.double n = Alt.double n.
+Proof.
+  unfold Alt.double, Nat.double.
+  omega.
+Qed.
+\end{Coqsrc}
+
+\section{Acknowledgements}
+\label{sec:orgheadline5}
+    Many thanks to David Ilcinkas, Sylvain Salvati, Alan Schmitt and Théo Zimmerman for their help on the elaboration of this document, and to the
+ members of the \emph{Formal Methods} team at laBRI for their helpful comments 
+on an oral presentation of this work.
+
+Many thanks also to the Coq development team, Yves Bertot, and members of the \emph{Coq Club} for interesting discussions about the \coq{} system and the Calculus of Inductive Constructions.
+
+The author of the present document wishes to express his gratitude to the late Patrick Dehornoy, whose talk  was determinant for our desire to work on this topic.
+
+I owe my interest in discrete mathematics and their relation to formal proofs and functional programming  to Srecko Brlek.
+Equally, there is W.
+H.
+Burge's book ``\emph{Recursive Programming Techniques}'' ~\cite{burge} which was a great  source of inspiration.
+
+% \subsubsection{Links to the Coq source}
+
+%-----------------------------------------------------------
+
+\chapter{Hydras and Hydra Games}
+\label{sec:orgheadline91}
+
+This chapter is dedicated to the representation of hydras and rules of the hydra game in \coq's specification language:\gallina.
+
+Technically, a \emph{hydra} is just a finite ordered tree, each node of which 
+has any number of sons. Note that, contrary to the computer science tradition, we will show the hydras 
+with the heads up and the foot (i.e., the root of the tree) down.
+Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the 
+module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}).
+
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}[very thick, scale=0.6]
+
+\node (foot) at (2,0) {$\bullet$};
+\node (N1) at (2,2) {$\bullet$};
+\node (N2) at (2,4) {$\bullet$};
+\node (N3) at (2,6) {$\bullet$};
+\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
+\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
+\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
+\node (H4) at (4,2) {$\Smiley[2][vertfluo]$};
+\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
+\draw (foot) -- (N1)[very thick] ;
+\draw (N1) -- (N2);
+\draw (N2) -- (N3);
+\draw (N3) to [bend left= 10]  (H1) ;
+\draw (N3) to [bend right= 16] (H2);
+\draw (foot) to [bend left= 10]  (H0) ;
+\draw (foot) to [bend right = 10] (H4) ;
+\draw (N1) to [bend right= 16] (H5);
+\end{tikzpicture}
+\caption{The hydra Hy \label{fig:Hy}}
+\end{figure}
+
+We use a specific vocabulary for talking about hydras. Table~\ref{tab:hyd2tree} shows the correspondance between our terminology and the usual vocabulary for trees in computer science.
+
+\begin{figure}[h]
+  \centering
+  \begin{tabular}{ll}
+Hydras & Finite rooted trees\\
+\hline
+foot & root\\
+head & leaf\\
+node & node\\
+segment  & (directed) edge \\
+sub-hydra & subtree\\
+daughter & immediate subtree\\
+\end{tabular}
+  \caption{Translation from hydras to trees}
+  \label{tab:hyd2tree}
+\end{figure}
+
+The hydra \texttt{Hy} has a \emph{foot} (below), five \emph{heads}, and eight \emph{segments}. 
+We leave it to the reader to define various parameters such as the height, the size, the highest arity (number of sons of a node) of a hydra.
+In our example, these parameters have the respective values : $4$, $9$ and $3$.
+
+\subsection{The rules of the game}
+
+\label{sec:orgheadline44}
+A \emph{hydra battle} is a fight between Hercules and the Hydra. 
+More formally, a  battle is a sequence of \emph{rounds}.
+At each round:
+\begin{itemize}
+\item If the hydra is composed of just one head, the battle is finished
+and  Hercules is the winner.
+\item Otherwise, Hercules chops off \emph{one} head of the hydra,
+
+\begin{itemize}
+\item If the head is at distance 1 from the foot, the head is just lost by the hydra, with no more reaction.
+\item Otherwise, let us denote by \(r\) the node that was at distance \(2\) from 
+the removed head in the direction of the foot,  and consider the  sub-hydra \(h'\) of \(h\), whose  root is \(r\) \footnote{$h'$ will be called ``the wounded part of the hydra'' in the subsequent text.
+In Figures~\vref{fig:Hy2} and ~\vref{fig:Hy4}, this sub-hydra  is displayed in red.}.
+Let $n$ be some natural number.
+Then $h'$ is replaced by  $n+1$ of copies of \(h'\) which share the same root $r$.
+ The \emph{replication number} $n$ may be different (and generally is)   at each round of the fight.
+It may be chosen by the hydra, according to its strategy, or imposed by some 
+particular rule.
+In many presentations of hydra battles, this number is increased by $1$ at each round.
+In the following presentation, we will also consider battles where the hydra is free to chose its number of replication at every round of the battle\footnote{Let us recall that, if the chopped-off head was at distance 1 from the foot, the replication factor is meaningless.}.
+\end{itemize}
+\end{itemize}
+
+Note that the description given in~\cite{KP82} of the replication process in hydra battles is also  semi-formal.
+
+\label{original-rules}
+
+\begin{quote}
+  ``From the node that used to be attached to the head which was just chopped off, traverse one 
+segment towards the root until the next node is reached. From this node sprout $n$ replicas of 
+that part of the hydra (after decapitation) which is ``above'' the segment just traversed, i.e., those 
+nodes and segments from which, in order to reach the root, this segment would have to be 
+traversed. If the head just chopped off had the root of its nodes, no new head is grown. ''
+\end{quote}
+
+Moreover, we note that this description is in \emph{imperative} terms.
+In order to build a formal  study of the properties of hydra battles, we prefer to use a mathematical vocabulary, i.e., graphs, relations, functions, etc.
+Thus, the replication process will be represented as a binary relation on a data type \texttt{Hydra},
+linking the state of the hydra \emph{before} and \emph{after} the transformation.
+A battle will thus be represented as a sequence of terms of type \texttt{Hydra}, respecting the rules of the game.
+
+\subsection{Example}
+Let us start a battle between Hercules and the hydra \texttt{Hy} of Fig.~\ref{fig:Hy}.
+
+At the first round, Hercules choses to chop off the rightmost head of \texttt{Hy}.
+Since this head is near the floor, the hydra loses this head. Let us call 
+ \texttt{Hy'} the resulting state of the hydra, represented in Fig.~\vref{fig:Hy-prime}.
+
+Next, assume Hercules choses to chop off one of the two highest heads of \texttt{Hy'}, for instance the rightmost one.
+Fig.~\vref{fig:Hy2} represents the rotten neck in dashed lines, and the part that will be replicated in red.
+Assume also that the hydra decides to add 4 copies of the red part\footnote{In other words, the replication factor at this round is equal to $4$.}.
+We obtain a new state \texttt{Hy''} depicted in Fig.~\ref{fig:Hy3}.
+
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}[very thick, scale=0.6]
+
+\node (foot) at (2,0) {$\bullet$};
+\node (N1) at (2,2) {$\bullet$};
+\node (N2) at (2,4) {$\bullet$};
+\node (N3) at (2,6) {$\bullet$};
+\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
+\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
+\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
+\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
+\node (H4) at (6,0) {$\Xey[2][lightgray]$};
+\draw (foot) -- (N1)[very thick] ;
+\draw (N1) -- (N2);
+\draw (N2) -- (N3) ;
+\draw (N3) to [bend left= 10]  (H1) ;
+\draw (N3) to [bend right= 16] (H2);
+\draw (foot) to [bend left= 10]  (H0) ;
+\draw (N1) to [bend right= 16] (H5);
+\end{tikzpicture}
+
+\caption{Hy': the state  of Hy after one round \label{fig:Hy-prime}}
+\end{figure}
+
+\begin{figure}[hp]
+\centering
+\begin{tikzpicture}[very thick, scale=0.5]
+
+\node (foot) at (2,0) {$\bullet$};
+\node (N1) at (2,2) {$\bullet$};
+\node (N2) at (2,4)  {{\color{lightred}$\bullet$}};
+\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
+\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
+\node (H1) at (0,8) {$\Sey[2][lightred]$};
+\node (H2) at (5,0) {$\Xey[2][lightgray]$};
+\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
+\node (ex) at (5,8) {};
+\draw (foot) -- (N1)[very thick] ;
+\draw (N1) -- (N2);
+\draw  (N2) -- (N3)[draw=lightred];
+\draw (N3) to   [bend left= 10](H1) [draw=lightred];
+\draw [dashed] (N3) to [bend left= 10](ex);
+\draw (foot) to [bend left= 10]  (H0) ;
+\draw (N1) to [bend right= 16] (H5);
+\end{tikzpicture}
+\caption{A second beheading}
+\label{fig:Hy2}
+\end{figure}
+
+\begin{figure}[hp]
+\centering
+\begin{tikzpicture}[very thick, scale=0.6]
+
+\node (foot) at (2,0) {$\bullet$};
+\node (N1) at (2,2) {$\bullet$};
+\node (N2) at (2,4) {$\bullet$};
+\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
+\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
+\node (H11) at (2,8) {$\Smiley[2][vertfluo]$};
+\node (H12) at (4,8) {$\Smiley[2][vertfluo]$};
+\node (H13) at (6,8) {$\Smiley[2][vertfluo]$};
+\node (H14) at (8,8) {$\Smiley[2][vertfluo]$};
+
+\node (N3) at (1,6) {$\bullet$};
+\node (N31) at (2,6) {$\bullet$};
+\node (N32) at (3,6) {$\bullet$};
+\node (N33) at (4,6) {$\bullet$};
+\node (N34) at (5,6) {$\bullet$};
+
+\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
+\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
+\draw (foot) -- (N1)[very thick] ;
+\draw (N1) -- (N2);
+\draw (N2) -- (N3);
+\draw (N2) -- (N31);
+\draw (N2) -- (N32);
+\draw (N2) -- (N33);
+\draw (N2) -- (N34);
+\draw (N3) to   [bend left= 10](H1) ;
+\draw (N31) to   [bend left= 10](H11) ;
+\draw (N32) to   [bend left= 10](H12) ;
+\draw (N33) to   [bend left= 10](H13) ;
+\draw (N34) to   [bend left= 10](H14) ;
+\draw (foot) to [bend left= 10]  (H0) ;
+\draw (N1) to [bend left= 10]  (H5) ;
+\end{tikzpicture}
+\caption{Hy'', the state of Hy after two rounds \label{fig:Hy3}}
+\end{figure}
+
+Figs.~\ref{fig:Hy4} and~\vref{fig:Hy5} represent a possible third round of the battle, with a replication factor equal to $2$. Let us call \texttt{Hy'''} the state of the hydra after that third round.
+
+\begin{figure}[hp]
+\centering
+\begin{tikzpicture}[very thick, scale=0.6]
+
+\node (foot) at (2,0)  {{\color{lightred}$\bullet$}};
+\node (N1) at (2,2) {{\color{lightred}$\bullet$}};
+\node (N2) at (2,4) {{\color{lightred}$\bullet$}};
+\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
+\node (exN4) at (4,4) {};
+\node (H1) at (0,8) {$\Sey[2][lightred]$};
+\node (H11) at (2,8) {$\Sey[2][lightred]$};
+\node (H12) at (4,8) {$\Sey[2][lightred]$};
+\node (H13) at (6,8) {$\Sey[2][lightred]$};
+\node (H14) at (8,8) {$\Sey[2][lightred]$};
+
+\node (N3) at (1,6) {{\color{lightred}$\bullet$}};
+\node (N31) at (2,6) {{\color{lightred}$\bullet$}};
+\node (N32) at (3,6) {{\color{lightred}$\bullet$}};
+\node (N33) at (4,6) {{\color{lightred}$\bullet$}};
+\node (N34) at (5,6) {{\color{lightred}$\bullet$}};
+
+\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
+\node (H5) at (4,0) {$\Xey[2][lightgray]$};
+\draw (foot) -- (N1)[very thick,draw=lightred] ;
+\draw (N1) -- (N2)[draw=lightred];
+\draw (N2) -- (N3)[draw=lightred];
+\draw (N2) -- (N31)[draw=lightred];
+\draw (N2) -- (N32)[draw=lightred];
+\draw (N2) -- (N33)[draw=lightred];
+\draw (N2) -- (N34)[draw=lightred];
+\draw (N3) to   [bend left= 10](H1) [draw=lightred];
+\draw (N31) to   [bend left= 10](H11) [draw=lightred];
+\draw (N32) to   [bend left= 10](H12) [draw=lightred];
+\draw (N33) to   [bend left= 10](H13) [draw=lightred];
+\draw (N34) to   [bend left= 10](H14) [draw=lightred];
+\draw (foot) to [bend left= 10]  (H0) ;
+\draw [dashed] (N1) to  [bend left= 10](exN4);
+\end{tikzpicture}
+\caption{A third beheading (wounded part in red) \label{fig:Hy4}}
+\end{figure}
+
+\begin{figure}[hp]
+\centering
+\begin{tikzpicture}[very thick, scale=0.4]
+
+\node (foot) at (10,0) {$\bullet$};
+
+\node (N1) at (2,2) {$\bullet$};
+\node (N2) at (2,4) {$\bullet$};
+\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
+\node (H1) at (0,8) {$\Smiley[1][vertfluo]$};
+\node (H11) at (2,8) {$\Smiley[1][vertfluo]$};
+\node (H12) at (4,8) {$\Smiley[1][vertfluo]$};
+\node (H13) at (6,8) {$\Smiley[1][vertfluo]$};
+\node (H14) at (8,8) {$\Smiley[1][vertfluo]$};
+
+\node (N3) at (1,6) {$\bullet$};
+\node (N31) at (2,6) {$\bullet$};
+\node (N32) at (3,6) {$\bullet$};
+\node (N33) at (4,6) {$\bullet$};
+\node (N34) at (5,6) {$\bullet$};
+
+\node (H0) at (-3,3) {$\Smiley[1][vertfluo]$};
+
+\draw (foot) to [bend left=10] (N1)[very thick] ;
+\draw (N1) -- (N2);
+\draw (N2) -- (N3);
+\draw (N2) -- (N31);
+\draw (N2) -- (N32);
+\draw (N2) -- (N33);
+\draw (N2) -- (N34);
+\draw (N3) to   [bend left= 10](H1) ;
+\draw (N31) to   [bend left= 10](H11) ;
+\draw (N32) to   [bend left= 10](H12) ;
+\draw (N33) to   [bend left= 10](H13) ;
+\draw (N34) to   [bend left= 10](H14) ;
+\draw (foot) to [bend left = 15]  (H0) ;
+
+% second copy 
+\node (N01) at (12,2) {$\bullet$};
+\node (N02) at (12,4) {$\bullet$};
+\node (N03) at (12,6) {{\color{lightred}$\bullet$}};
+\node (H001) at (10,8) {$\Smiley[1][vertfluo]$};
+\node (H0011) at (12,8) {$\Smiley[1][vertfluo]$};
+\node (H0012) at (14,8) {$\Smiley[1][vertfluo]$};
+\node (H0013) at (16,8) {$\Smiley[1][vertfluo]$};
+\node (H0014) at (18,8) {$\Smiley[1][vertfluo]$};
+
+\node (N03) at (11,6) {$\bullet$};
+\node (N031) at (12,6) {$\bullet$};
+\node (N032) at (13,6) {$\bullet$};
+\node (N033) at (14,6) {$\bullet$};
+\node (N034) at (15,6) {$\bullet$};
+
+\draw (foot) -- (N01)[very thick] ;
+\draw (N01) -- (N02);
+\draw (N02) -- (N03);
+\draw (N02) -- (N031);
+\draw (N02) -- (N032);
+\draw (N02) -- (N033);
+\draw (N02) -- (N034);
+\draw (N03) to   [bend left= 10](H001) ;
+\draw (N031) to   [bend left= 10](H0011) ;
+\draw (N032) to   [bend left= 10](H0012) ;
+\draw (N033) to   [bend left= 10](H0013) ;
+\draw (N034) to   [bend left= 10](H0014) ;
+
+% third copy 
+\node (N001) at (22,2) {$\bullet$};
+\node (N002) at (22,4) {$\bullet$};
+\node (N003) at (22,6) {{\color{lightred}$\bullet$}};
+\node (H001) at (20,8) {$\Smiley[1][vertfluo]$};
+\node (H0011) at (22,8) {$\Smiley[1][vertfluo]$};
+\node (H0012) at (24,8) {$\Smiley[1][vertfluo]$};
+\node (H0013) at (26,8) {$\Smiley[1][vertfluo]$};
+\node (H0014) at (28,8) {$\Smiley[1][vertfluo]$};
+
+\node (N003) at (21,6) {$\bullet$};
+\node (N0031) at (22,6) {$\bullet$};
+\node (N0032) at (23,6) {$\bullet$};
+\node (N0033) at (24,6) {$\bullet$};
+\node (N0034) at (25,6) {$\bullet$};
+
+\draw (foot) -- (N001)[very thick] ;
+\draw (N001) -- (N002);
+\draw (N002) -- (N003);
+\draw (N002) -- (N0031);
+\draw (N002) -- (N0032);
+\draw (N002) -- (N0033);
+\draw (N002) -- (N0034);
+\draw (N003) to   [bend left= 10](H001) ;
+\draw (N0031) to   [bend left= 10](H0011) ;
+\draw (N0032) to   [bend left= 10](H0012) ;
+\draw (N0033) to   [bend left= 10](H0013) ;
+\draw (N0034) to   [bend left= 10](H0014) ;
+\end{tikzpicture}
+\caption{The configuration Hy''' of Hy \label{fig:Hy5}}
+\end{figure}
+\FloatBarrier
+
+We leave it to the reader  to guess the following  rounds of the battle \dots
+ % Please keep in mind that, in this 
+% the hydra is free to chose any number of replications at each time, whereas
+% Hercules chops only one head per round.
+
+% Let us precise that, in this game, Hercules wins if the hydra is eventually reduced 
+% to a single head. 
+% We know from~\cite{KP82} that, whichever the initial configuration of the
+% hydra, and the strategies of both players, Hercules eventually wins. The 
+% aforementionned paper shows also that there do not exist any \emph{simple} proof of this result.
+
+\section{Hydras and their representation in \emph{Coq}}
+\label{sec:orgheadline48}
+
+In order to describe trees where each node can have an arbitrary (but finite) number of sons, it usual to define a type where each node carries a \emph{forest}, \emph{i.e} a list of trees
+(see for instance Chapter 14, pages 400-406 of \cite{BC04}).
+
+For this purpose, we define two mutual \emph{ad-hoc}  inductive types, where \texttt{Hydra} is the main type, and \texttt{Hydrae} is a helper for describing finite sequences of hydra.
+\label{types:Hydra}
+\label{types:Hydrae}
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Hydra}{Hydra.Hydra\_Definitions}}
+
+\begin{Coqsrc}
+Inductive Hydra : Set :=
+|  node :  Hydrae -> Hydra
+with Hydrae : Set :=
+| hnil : Hydrae
+| hcons : Hydra -> Hydrae -> Hydrae.
+\end{Coqsrc}
+
+%\index{To do}
+
+\begin{todo}
+Look for an existing library on trees with nodes of arbitrary arity, in order to replace  this ad-hoc type with something more generic.
+\end{todo}
+
+\index{Projects}
+
+\begin{project}[**]
+
+ Another very similar representation could use the \texttt{list} type family instead of the specific 
+type \texttt{Hydrae}:
+
+\begin{Coqalt}
+Module Alt.
+
+Inductive Hydra: Set :=
+  hnode (daughters : list Hydra).
+
+End Alt.
+\end{Coqalt}
+
+Using this representation, re-define all the constructions of this chapter.
+You will probably have to use patterns described for instance in~\cite{BC04} or the archives of the Coq-club~\cite{Coq}.
+
+\end{project}
+
+\index{Projects}
+
+\begin{project}
+The type \texttt{Hydra} above describes hydras as \emph{plane trees}, i.e., as drawn on a sheet of paper or computer screen. Thus, hydras are \emph{oriented},
+and it is appropriate to consider a \emph{leftmost} or \emph{rightmost} head of
+the beast. It could be interesting to consider another representation, in which
+every non-leaf node has a \emph{multi-set} -- not an ordered list -- of daughters.
+\end{project}
+
+\subsubsection{Abbreviations}
+
+We provide several notations for ``patterns'' which occur often in our developments.
+
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#head}{Hydra.Hydra\_Definitions}}
+
+\begin{Coqsrc}
+(** heads *)
+Notation head := (node hnil).
+
+(** nodes  with 1, 2 or 3 daughters *)
+Notation hyd1 h := (node (hcons h hnil)).
+Notation hyd2 h h' := (node (hcons h (hcons h' hnil))).
+Notation hyd3 h h' h'' := 
+                   (node (hcons h (hcons h' (hcons h'' hnil)))).
+\end{Coqsrc}
+
+For instance, the hydra \texttt{Hy}  of Figure~\vref{fig:Hy} is defined in \emph{Gallina} as follows:
+
+\vspace{4mm}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html\#Hy}{Hydra.Hydra\_Examples}}
+
+\begin{Coqsrc}
+Example Hy := hyd3 head
+                   (hyd2
+                      (hyd1 
+                         (hyd2 head head))
+                      head) 
+                   head.
+\end{Coqsrc}
+
+Hydras quite frequently contain  multiple  copies of the same pattern. The following functions
+will help us to describe and reason about replications in hydra battles.
+
+\vspace{4pt}
+
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#hcons_mult}{Hydra.Hydra\_Definitions}}
+
+\begin{Coqsrc}
+Fixpoint hcons_mult (h:Hydra)(n:nat)(s:Hydrae):Hydrae :=
+  match n with 
+  | O => s
+  | S p => hcons h (hcons_mult h p s)
+  end.
+
+(** hydra with n copies of the same daughter *)
+
+Definition hyd_mult h n :=
+  node (hcons_mult h n hnil).
+\end{Coqsrc}
+
+\vspace{4mm}
+
+For instance, the hydra $Hy''$ of Fig~\vref{fig:Hy3}  can be defined in \coq{} as follows:
+
+\vspace{4pt}
+
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
+
+\begin{Coqsrc}
+Example Hy'' := 
+     hyd2 head
+          (hyd2 (hyd_mult (hyd1 head) 5)
+                head).
+\end{Coqsrc}
+
+\subsubsection{Recursive functions on type \texttt{Hydra}}
+\label{sec:orgheadline41}
+\label{sec:hsize-def}
+
+When defining a recursive function over the type \texttt{Hydra}, one has to consider the three constructors 
+\texttt{node}, \texttt{hnil} and \texttt{hcons} of the mutually inductive types \texttt{Hydra} and \texttt{Hydrae}. 
+Let us define for instance the function that  computes the number of nodes of any hydra:
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
+
+\begin{Coqsrc}
+Fixpoint hsize (h:Hydra) : nat :=
+  match h with node l => S (lhsize l)
+  end
+with lhsize l : nat :=
+  match l with hnil => 0
+            | hcons h hs => hsize h + lhsize hs 
+  end.
+
+ Compute hsize Hy.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+ = 9
+     : nat 
+\end{Coqanswer}
+
+Likewise, the \emph{height} (maximum distance between the foot and a head) 
+is defined by mutual recursion:
+
+\begin{Coqsrc}
+Fixpoint height  (h:Hydra) : nat :=
+  match h with node l => lheight l
+  end
+with lheight l : nat :=
+  match l with 
+  | hnil => 0
+  | hcons h hs => Max.max (S (height h)) (lheight hs)
+  end.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Compute height Hy.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+ = 4
+     : nat  
+\end{Coqanswer}
+
+\index{Exercises}
+
+\begin{exercise}
+Define a function \texttt{max\_degree: Hydra $\arrow$ nat} which  returns the highest degree of a node in any hydra.
+For instance, the evaluation of the term \texttt{(max\_degree Hy)} should return $3$.
+\end{exercise}
+
+\subsection{Induction principles for hydras}
+\label{sec:orgheadline42}
+
+In this section, we show how induction principles are used to prove properties on the type 
+\texttt{Hydra}. Let us consider for instance the following statement:
+\begin{quote}
+  `` The height of any hydra is strictly less than its size. ''
+\end{quote}
+
+\subsubsection{A failed attempt}
+
+One may try to use the default tactic of proof by induction, that corresponds to an application of the automatically  generated  induction principle for  type \texttt{Hydra}:
+
+\begin{Coqanswer}
+Hydra_ind :
+forall P : Hydra -> Prop,
+(forall h : Hydrae, P (node h)) -> forall h : Hydra, P h
+\end{Coqanswer}
+
+Ler us start a simple proof by induction.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
+
+\begin{Coqbad}
+Module Bad.
+
+Lemma height_lt_size (h:Hydra) :
+  height h <= hsize h.
+Proof.
+  induction h as [s].
+\end{Coqbad}
+
+\begin{Coqanswer}
+1 subgoal, subgoal 1 (ID 11)
+
+  s : Hydrae
+  ============================
+   height (node s) <= hsize (node s)
+\end{Coqanswer}
+
+We might be tempted to do an induction on the sequence \texttt{s}:
+
+% \begin{Coqbad}
+% induction  s as [ | h s'].
+%   -   cbn; auto with arith.
+%   - 
+% \end{Coqbad}
+
+\begin{Coqanswer}
+ 1 focused subgoal
+(unfocused: 0), subgoal 1 (ID 19)
+
+  h : Hydra
+  s' : Hydrae
+  IHs' : height (node s') <= hsize (node s')
+  ============================
+   height (node (hcons h s')) <= hsize (node (hcons h s'))
+
+\end{Coqanswer}
+
+Note that the displayed subgoal does not contain any assumption on \texttt{h}, thus there is no way to 
+infer any property about the height and size of the hydra \texttt{(hcons h t)}.
+
+\begin{Coqbad}
+Abort.
+
+End Bad.
+\end{Coqbad}
+
+\subsubsection{A principle for  mutual induction}
+In order to get an appropriate induction scheme for the types 
+\texttt{Hydra} and \texttt{Hydrae}, we can use  \coq{}'s  command \texttt{Scheme}.
+
+\index{Coq!Techniques!Mutual induction}
+\index{Coq!Commands!Scheme}
+
+\begin{Coqsrc}
+Scheme Hydra_rect2 := Induction for Hydra Sort Type
+with Hydrae_rect2 := Induction for Hydrae Sort Type.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Check Hydra_rect2.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+Hydra_rect2
+ : forall (P : Hydra -> Type) (P0 : Hydrae -> Type),
+   (forall h : Hydrae, P0 h -> P (node h)) ->
+    P0 hnil ->
+    (forall h : Hydra, P h -> 
+             forall h0 : Hydrae, P0 h0 -> P0 (hcons h h0)) ->
+    forall h : Hydra, P h
+\end{Coqanswer}
+
+\subsubsection{A correct proof}
+
+Let us now use \texttt{Hydra\_rect2} for proving that the height of any hydra is strictly less than its size.
+Using this scheme requires an auxiliary predicate, called \texttt{P0} in \texttt{Hydra\_rect2}'s statement. 
+Let us begin by defining an ad-hoc version of \texttt{List.Forall}.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
+
+\begin{Coqsrc}
+(** All elements of s satisfy P *)
+
+Fixpoint h_forall (P: Hydra -> Prop) (s: Hydrae) :=
+  match s with
+      hnil => True
+    | hcons h s' => P h /\ h_forall P s'
+  end.
+ \end{Coqsrc}
+
+ \begin{Coqsrc}
+Lemma height_lt_size (h:Hydra) :
+ height h < hsize h.
+Proof.
+  induction h using Hydra_rect2  with 
+  (P0 :=  h_forall (fun h =>  height h < hsize h)).
+ \end{Coqsrc}
+
+\begin{enumerate}
+\item The first subgoal is as follows:
+
+\begin{Coqanswer}
+
+  h: Hydrae
+  IHh : h_forall (fun h : Hydra => height h < hsize h) h
+  ============================
+   height (node s) < hsize (node s)
+
+\end{Coqanswer}
+
+This goal is easily solvable, using some arithmetic. We let the reader look at the source of this development.
+
+\item The second subgoal is trivial:
+
+\begin{Coqanswer}
+
+  ============================
+    h_forall (fun h : Hydra => height h < hsize h) hnil
+
+\end{Coqanswer}
+
+\begin{Coqsrc}
+  reflexivity.
+\end{Coqsrc}
+
+\item Finally, the last subgoal is also easy to solve:
+
+\begin{Coqanswer}
+  h : Hydra
+  h0 : Hydrae
+  IHh : height h < hsize h
+  IHh0 : h_forall (fun h : Hydra => height h < hsize h) h0
+  ============================
+   h_forall (fun h1 : Hydra => height h1 < hsize h1) 
+                 (hcons h h0)
+\end{Coqanswer}
+
+\begin{Coqsrc}
+ split;auto. 
+Qed.   
+\end{Coqsrc}
+
+\end{enumerate}
+
+\index{Exercises}
+
+\begin{exercise}
+It happens very often that, in the proof of  a proposition of the form 
+(\texttt{$\forall\,$ h:Hydra, $P$ h}), the predicate \texttt{P0}
+is  (\texttt{h\_forall $P$}).
+Design a tactic for induction on hydras that frees the user from binding explicitly \texttt{P0},  and solves trivial subgoals.
+Apply it for writing  a shorter proof of \texttt{height\_lt\_size}.
+\end{exercise}
+
+\section{Relational description of hydra battles}
+
+In this section, we represent the rules of hydra battles as a binary relation associated with
+a \emph{round}, i.e., an interaction composed of the two following actions:
+\begin{enumerate}
+\item Hercules chops off one head of the hydra
+\item Then, the  hydra replicates the wounded part (if the head is at distance $\geq 2$ from the foot).
+\end{enumerate}
+The relation associated with each round of the battle is \emph{parameterized}  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot,
+but present for consistency's sake).
+
+In our description,  we will apply the following naming convention: if $h$ represents the configuration of the hydra before a round, then the configuration of $h$ after this round will be called $h'$.
+ Thus, we are going to define a proposition  (\texttt{round\_n $n\;h\;h'$})  whose intended meaning will be `` the hydra $h$  is transformed into $h'$  in a single round of a battle, with the expected replication factor $n$ ''.
+
+Since the replication of parts of the hydra depends on the distance of the chopped head from  the foot, we  decompose our description into two main  cases, under the form of a bunch of [mutually] inductive predicates over the types \texttt{Hydra} and \texttt{Hydrae}.
+
+The mutually exclusive cases we consider are the following.
+\begin{itemize}
+\item \textbf{R1}: The chopped off head was at distance 1 from the foot.
+\item \textbf{R2}: The chopped off head was at a distance greater than or equal to  $2$ from the foot.
+\end{itemize}
+
+\subsection{Chopping off a head at distance 1 from the foot (relation  R1)}
+
+If Hercules chops off a head near the floor, there is no replication at all. We use an auxiliary predicate, associated with the removing of one head from a sequence of hydras.
+
+\vspace{4pt}\emph{From Module\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
+
+\begin{Coqsrc}
+Inductive S0 :  relation Hydrae :=
+| S0_first : forall s, S0  (hcons head s) s
+| S0_rest : forall  h s s',  S0  s s' ->
+                             S0  (hcons h s) (hcons h s').
+
+Inductive R1  :  Hydra -> Hydra -> Prop :=
+| R1_intro : forall s s', S0 s s' -> R1 (node s) (node s').
+\end{Coqsrc}
+
+\subsubsection{Example}
+\label{sec:orgheadline45}
+
+Let us represent in \coq{}   the transformation of the hydra of Fig.~\vref{fig:Hy} into
+the configuration represented in Fig.~\ref{fig:Hy-prime}.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
+
+\begin{Coqsrc}
+Example Hy_1 : R1 Hy Hy'.
+Proof. 
+  split; right; right; left.
+Qed.
+\end{Coqsrc}
+
+\subsection{Chopping of a head at distance \texorpdfstring{$\geq 2$}{>= 2} from the foot (relation R2) }
+
+Let us now consider beheadings  where the chopped-off head is at distance greater than or equal to $2$ from the foot.
+All the following relations are parameterized by the number $n$ of new copies added by the hydra.
+
+ Let $s$ be a sequence of hydras. 
+The proposition (\texttt{S1 n s s'}) holds if $s'$ is obtained by replacing some element $h$ of $s$ by 
+$n+1$ copies of $h'$, where  the proposition (\texttt{R1 h h'}) holds, in other words, \texttt{h'} is just \texttt{h}, without the chopped-off  head.
+\texttt{S1} is an inductive relation with two constructors that allow us to choose the position in $s'$ of the wounded sub-hydra $h$.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#S1}{Hydra.Hydra\_Definitions}}
+
+\begin{Coqsrc}
+Inductive S1 (n:nat)  : Hydrae -> Hydrae -> Prop :=
+| S1_first : forall s h h' ,   R1 h h' -> 
+                  S1 n (hcons h s) (hcons_mult h' (S n) s)
+| S1_next : forall h s s',  S1 n s s' ->
+                   S1 n (hcons h s) (hcons h s').
+\end{Coqsrc}
+
+The rest of the definition is structured as two mutually inductive relations on hydras and sequences of hydras.
+The first constructor of \texttt{R2} describes the case where the chopped head is exactly at height $2$.
+The others constructors allow us to consider beheadings at height strictly greater than $2$.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions}}
+
+\begin{Coqsrc}
+Inductive R2 (n:nat)  :  Hydra -> Hydra -> Prop :=
+| R2_intro : forall s s', S1 n s s' -> R2 n (node s) (node s')
+| R2_intro_2 : forall s s', S2 n s s' -> R2 n (node s) (node s')
+
+with S2 (n:nat) :  Hydrae -> Hydrae -> Prop :=
+|  S2_first : forall h h' s ,
+               R2 n  h h'  -> 
+               S2  n (hcons h s) (hcons h'  s)
+|  S2_next  : forall h   r r',
+               S2 n   r r' ->
+               S2 n (hcons h r) (hcons h r').                  
+\end{Coqsrc}
+
+\subsubsection{Example}
+Let us prove the transformation of \texttt{Hy'} into \texttt{Hy''} (see Fig.~\vref{fig:Hy3}).
+
+\vspace{4pt}\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}.
+
+\begin{Coqsrc}
+Example R2_example:  R2 4 Hy' Hy''.
+Proof.
+  right. right. left.  right. left. left. left. split. left.
+Qed.
+\end{Coqsrc}
+
+The reader is encouraged to look at all the successive subgoals of this example.
+\emph{Please consider also exercise~\vref{exo:interactive-battle}.}
+
+\subsection{Description of any round}
+
+We combine the two cases above into a single relation.
+First,  we define the  relation \texttt{(round\_n n h h')} where \texttt{n} is the expected number of  replications (irrelevant in the case of an \texttt{R1}-transformation).
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#round_n}{Hydra.Hydra\_Definitions}}
+
+\index{Predicates!round\_n}
+
+\begin{Coqsrc}
+Definition round_n n h h' := R1 h h' \/ R2 n h h'.  
+\end{Coqsrc}
+
+By abstraction over \texttt{n}, we define a \emph{round} (small step) of a battle:
+
+\index{Predicates!round}
+\label{sect:infix-round}
+\begin{Coqsrc}
+Definition round h h' := exists n,  round_n n h h'.
+
+Infix "-1->" := round (at level 60).
+\end{Coqsrc}
+
+\index{Projects}
+
+\begin{project}
+Give a direct translation of Kirby and Paris's description of hydra battles (quoted on page~\pageref{original-rules}) and prove that our relational description is consistent with theirs.
+\end{project}
+
+\subsection{Rounds and battles}
+
+Using library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}, we define \texttt{round\_plus},  the transitive closure of \texttt{round}, and \texttt{round\_star},  the reflexive and transitive closure of \texttt{round}.
+
+\label{sect:infix-rounds}
+
+\begin{Coqsrc}
+Definition round_plus := clos_trans_1n Hydra round.
+Infix "-+->" := rounds (at level 60).
+
+Definition round_star h h' := h = h' \/ round_plus h h'.
+Infix "-*->" := round_star (at level 60).
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+Prove the following lemma:
+
+\begin{Coqsrc}
+Lemma rounds_height : forall h h', 
+   h -+-> h' -> height h' <= height h.  
+\end{Coqsrc}
+
+\end{exercise}
+
+\begin{remark}
+\coq's library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Coq.Relations.Relation\_Operators} 
+contains three logically equivalent definitions of the transitive closure of a binary relation. This equivalence is proved in 
+\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} .
+
+Why three definitions for a single mathematical concept?
+Each definition generates an associated induction principle. 
+ According to the form of statement one would like to prove, there is a ``best choice'':
+
+\begin{itemize}
+\item For proving $\forall y, x\,R^+\,y \;\arrow\; P\,y$, prefer 
+\texttt{clos\_trans\_n1}
+\item For proving $\forall x,\,x\,R^+\,y \;\arrow\; P\,x$, prefer \texttt{clos\_trans\_1n}
+\item For proving $\forall x\,y, \,x\,R^+\,y \;\arrow\;P\,x\,y$,  
+prefer \texttt{clos\_trans},
+\end{itemize}
+But there is no ``wrong choice'' at all: the equivalence lemmas in \linebreak 
+\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} 
+ allow the user
+to convert any one of the three closures into another one before applying the corresponding elimination tactic.
+The same remark also holds for reflexive and transitive closures. 
+\end{remark}
+
+\index{Exercises}
+
+\begin{exercise}
+Define a restriction of \coqsimple{round},  where Hercules always chops off
+the leftmost among the lowest heads.
+
+Prove that, if $h$ is not a simple head, then there exists a unique $h'$ such that \texttt{h}  is transformed into \texttt{ h'} in one round, according to this restriction.
+
+\end{exercise}
+
+\index{Exercises}
+
+\begin{exercise}[Interactive battles]
+\label{exo:interactive-battle}
+Given a hydra \texttt{h}, the specification of a hydra battle for \texttt{h} is the type 
+\Verb@{h':Hydra | h -*-> h'}@. In order to avoid long sequences of \texttt{split}, \texttt{left}, and 
+\texttt{right}, design a set of dedicated tactics for the interactive building of a battle.
+Your tactics will have the following functionalities:
+\begin{itemize}
+\item  Chose to stop a battle, or continue
+\item Chose an expected number of replications
+\item Navigate in a hydra, looking for a head to chop off.
+\end{itemize}
+
+Use your tactics for simulating a small part of a hydra battle, for instance the rounds which lead from
+\texttt{Hy} to \texttt{Hy'''}  (Fig.~\vref{fig:Hy5}).
+
+\textbf{Hint:} Please keep in mind that the last  configuration of your interactively built battle is known only at the end of the battle.
+Thus, you will have to create and solve subgoals with existential variables.
+For that purpose, the tactic \texttt{eexists}, applied to the
+goal \Verb@{h':Hydra | h -*-> h'}@ generates the subgoal \Verb|h -*-> ?h'|.
+\end{exercise}
+
+\subsection{Classes of battles}
+\label{sect:battle-classes}
+
+In some presentations of hydra battles, e.g.~\cite{KP82, bauer2008}, the transformation associated with the $i$-th round may depend on $i$.
+For instance, in these articles, the replication factor at the $i$-th round is equal to $i$.
+In other examples, one can allow the hydra to apply any replication factor at any time.
+In order to be the most general as possible, we define the type of predicates which relate the state of the hydra before and after the $i$-th round of a battle.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
+\index{Coq!Type classes}
+\label{types:Battle}
+
+\begin{Coqsrc}
+Definition dep_round_t := nat -> Hydra -> Hydra -> Prop.
+
+Class Battle :=  {battle_r : dep_round_t;
+                  battle_inclusion : forall i h h',
+                      battle_r i h h' -> round h h'}.
+
+\end{Coqsrc}
+
+The most general class of battles is \texttt{free}, which allows the hydra to chose any replication factor at every step:
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#free}{Hydra.Hydra\_Definitions}}
+
+\begin{Coqsrc}
+Program Instance free : Battle :=
+  (Build_Battle ( fun _  h h' => round h h') _).
+\end{Coqsrc}
+
+We chosed to call \emph{standard} the kind of battles which appear  most often in the litterature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#standard}{Hydra.Hydra\_Definitions}}
+
+\begin{Coqsrc}
+Program Instance standard : Battle := (Build_Battle round_n _).
+Next Obligation.
+  now exists i.  
+Defined.
+\end{Coqsrc}
+
+\subsection{Big steps}
+
+Let $B$ be any instance of class \texttt{Battle}. It is easy to define inductively the relation between the $i$-th and the $j$-th steps of a battle of type $B$.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}
+
+\begin{Coqsrc}
+Inductive battle (B:Battle) : nat -> Hydra -> nat -> Hydra -> Prop :=
+
+| battle_1 : forall i h  h', battle_r   B i  h h' -> 
+                            battle B i h (S i) h'
+| battle_n : forall i h  j h' h'',  battle_r  B i h h''  ->
+                                   battle B (S i) h'' j h'  ->
+                                   battle B i h j h'.
+\end{Coqsrc}
+
+% \begin{remark}
+%  The class \texttt{free} is strongly related with the transitive closure  \texttt{round\_plus}, as expressed by the following lemmas.
+
+% \vspace{4pt}
+% \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
+
+%  \begin{Coqsrc}
+%  Lemma battle_free_equiv1 : forall i j h h',  
+%              battle free i h j h' ->   h -+-> h'.
+
+%  Lemma battle_free_equiv2 : forall h h',
+%      h -+-> h' ->
+%     forall i, exists j,  battle free i h j h'.
+%  \end{Coqsrc}
+
+% \end{remark}
+
+\section{A long battle}
+\label{sect:big-battle}
+
+In this section we consider a simple example of battle, starting with a small hydra,
+shown on figure~\vref{fig:hinit}, with a simple strategy for both players:
+
+\begin{itemize}
+\item Hercules chops always the rightmost head of the hydra.
+\item The battle is standard: at the round number $i$, the expected replication is $i$.
+\end{itemize}
+
+\begin{figure}[h]
+  \centering
+  \begin{tikzpicture}[thick, scale=0.30]
+ \node (foot) at (6,0) {$\bullet$};
+\node (n1) at  (3,3) {$\bullet$};
+\node (h1) at  (1,6) {$\Smiley[1][green]$};
+\node (h2) at  (3,6) {$\Smiley[1][green]$};
+\node (h3) at  (6,6) {$\Smiley[1][green]$};
+\node (h4) at  (6,6) {$\Smiley[1][green]$};
+\node (h5) at  (6,3) {$\Smiley[1][green]$};
+\node (h6) at  (9,3) {$\Smiley[1][green]$};
+\draw (foot) -- (n1);
+\draw (n1) to   [bend left=20] (h1);
+\draw (n1) to   (h2);
+\draw (n1) to   [bend right=20] (h3);
+\draw (foot) -- (h5);
+\draw (foot) to  [bend right=20] (h6);
+\end{tikzpicture}
+
+  \caption{The hydra hinit}
+  \label{fig:hinit}
+\end{figure}
+
+\begin{Coqsrc}
+Definition hinit := hyd3 (hyd_mult head 3)  head head.  
+\end{Coqsrc}
+
+The lemma we would like to prove is ``The considered battle lasts exactly $N$ rounds'',
+with $N$ being a natural number we gave to guess.
+
+But the  battle is so long that no \emph{test} can give us an estimation of its length, and we do need the expressive power of logic to compute this length.
+However, in order to  guess this length, we made some experiments, computing with \gallina{}, \coq{}'s  functional programming language.
+Thus, we can consider this development as a collaboration of proof with computation.
+
+In the following lines, we try to show faithfully how we found the value of the number $N$.
+
+The complete proof is in file \url{../src/html/hydras.Hydra.BigBattle.html}.
+
+\subsection{The beginning of hostilities}
+During the two first rounds, our hydra loses its two rightmost heads. Thus just before the third round, it looks like in figure~\vref{fig:hinit-plus2}.
+
+\begin{figure}[h]
+  \centering
+  \begin{tikzpicture}[thick, scale=0.30]
+ \node (foot) at (3,0) {$\bullet$};
+\node (n1) at  (3,3) {$\bullet$};
+\node (h1) at  (1,6) {$\Smiley[1][green]$};
+\node (h2) at  (3,6) {$\Smiley[1][green]$};
+\node (h3) at  (6,6) {$\Smiley[1][green]$};
+\node (h4) at  (6,6) {$\Smiley[1][green]$};
+\draw (foot) -- (n1);
+\draw (n1) to   [bend left=20] (h1);
+\draw (n1) to   (h2);
+\draw (n1) to   [bend right=20] (h3);
+\end{tikzpicture}
+
+  \caption{The hydra (hyd1 h3)}
+  \label{fig:hinit-plus2}
+\end{figure}
+
+The following lemma  is a formal statement about these first rounds, in terms of the
+\texttt{battle} predicate.
+
+\begin{Coqsrc}
+Lemma L_0_2 : battle standard 0 hinit 2 (hyd1 h3).   
+\end{Coqsrc}
+
+\subsection{Looking for regularities}
+
+A first study with pencil and paper suggested us that, after three rounds, the hydra always looks like in figure~\vref{fig:hinit-plusn} (with a variable number of 
+subtrees of height 1 or 0).
+Thus, we introduce handy notations.
+
+\begin{Coqsrc}
+Notation h3 := (hyd_mult head 3).
+Notation h2 := (hyd_mult head 2).
+Notation h1 := (hyd1 head).
+
+Definition hyd a b c := 
+  node (hcons_mult h2  a
+             (hcons_mult h1  b
+                         (hcons_mult head c hnil))).
+\end{Coqsrc}
+
+For instance Fig~\vref{fig:hinit-plusn} shows the hydra (\texttt{hyd 3 4 2}). The hydra (\texttt{hyd 0 0 0})  is the ``final'' hydra of any terminating battle, {i.e.},
+a tree whith exactly one node and no edge.
+
+\begin{figure}[h]
+  \centering
+  \begin{tikzpicture}[thick, scale=0.30]
+ \node (foot) at (15,0) {$\bullet$};
+\node (a) at  (3,4) {$\bullet$};
+\node (b) at  (6,4) {$\bullet$};
+\node (c) at  (9,4) {$\bullet$};
+\node (d) at  (13,4) {$\bullet$};
+\node (e) at  (16,4) {$\bullet$};
+\node (f) at  (19,4) {$\bullet$};
+\node (g) at  (22,4) {$\bullet$};
+\node (h) at  (25,4) {$\Smiley[1][green]$};
+\node (i) at  (28,4) {$\Smiley[1][green]$};
+\node (aa) at  (2.5,8) {$\Smiley[1][green]$};
+\node (ab) at  (3.5,8) {$\Smiley[1][green]$};
+\node (ba) at  (5.5,8) {$\Smiley[1][green]$};
+\node (bb) at  (6.5,8) {$\Smiley[1][green]$};
+\node (ca) at  (8.5,8) {$\Smiley[1][green]$};
+\node (cb) at  (9.5,8) {$\Smiley[1][green]$};
+\node (da) at  (13,8) {$\Smiley[1][green]$};
+\node (ea) at  (16,8) {$\Smiley[1][green]$};
+\node (fa) at  (19,8) {$\Smiley[1][green]$};
+\node (ga) at  (22,8) {$\Smiley[1][green]$};
+\draw (foot) -- (a);
+\draw (foot) -- (b);
+\draw (foot) -- (c);
+\draw (foot) -- (d);
+\draw (foot) -- (e);
+\draw (foot) -- (f);
+\draw (foot) -- (g);
+\draw (foot) -- (h);
+\draw (foot) -- (i);
+\draw (a) -- (aa);
+\draw (a) -- (ab);
+\draw (b) -- (ba);
+\draw (b) -- (bb);
+\draw (c) -- (ca);
+\draw (c) -- (cb);
+\draw (d) -- (da);
+\draw (e) -- (ea);
+\draw (f) -- (fa);
+\draw (g) -- (ga);
+% \node (a) at  (3,4) {$\bullet$};
+% \node (h1) at  (1,6) 
+% \node (h2) at  (3,6) {$\Smiley[1][green]$};
+% \node (h3) at  (6,6) {$\Smiley[1][green]$};
+% \node (h4) at  (6,6) {$\Smiley[1][green]$};
+% \draw (foot) -- (n1);
+% \draw (n1) to   [bend left=20] (h1);
+% \draw (n1) to   (h2);
+% \draw (n1) to   [bend right=20] (h3);
+\end{tikzpicture}
+
+  \caption{The hydra (hyd 3 4 2)}
+  \label{fig:hinit-plusn}
+\end{figure}
+
+With these notations, we get a formal description of the first three rounds.
+
+\begin{Coqsrc}
+Lemma L_2_3 : battle standard 2 (hyd1 h3)  3 (hyd 3 0 0).
+
+Lemma L_0_3 : battle standard 0 hinit 3 (hyd 3 0 0).
+\end{Coqsrc}
+
+\subsection{Computing \dots}
+In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple datatype for representing the states as tuples composed of
+the round number, and the respective number of daughters  \texttt{h2}, \texttt{h1}, and heads
+of the current hydra.
+
+\begin{Coqsrc}
+ Record state : Type :=
+    mks {round: nat ; n2 : nat ; n1 : nat ; nh : nat}.
+\end{Coqsrc}
+
+The following function returns the next configurarion of the game. 
+Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle will only start with the lemma
+\texttt{lemma:step-battle}, page~\pageref{lemma:step-battle}.
+
+\begin{Coqsrc}
+Definition next (s : state) :=
+  match s with
+  | mks round a b (S c) => mks (S round) a b c
+  | mks round a (S b) 0 => mks (S round) a b (S round)
+  | mks round (S a) 0 0 => mks (S round) a (S round) 0
+  | _ => s
+  end.
+\end{Coqsrc}
+
+We can make bigger steps through iterations of \texttt{next}.
+
+The functional \texttt{iterate}, similar to Standard Library's \texttt{Nat.iter},
+is defined and studied in~\href{../src/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}).
+
+\label{Functions:iterate}
+
+\begin{Coqsrc}
+Fixpoint iterate {A:Type}(f : A -> A) (n: nat)(x:A) :=
+  match n with
+  | 0 => x
+  | S p => f (iterate  f p x)
+  end.
+\end{Coqsrc}
+
+The following function computes the state of the battle at the $n$-th round.
+
+\begin{Coqsrc}
+Definition test n := iterate next (n-3) (mks 3 3 0 0).
+
+Compute test 3.
+   (**
+     = {| round := 3; n2 := 3; n1 := 0; nh := 0 |}
+     : state
+    *)
+
+Compute test 4.
+ (*
+  = {| round := 4; n2 := 2; n1 := 4; nh := 0 |}
+     : state
+ *)
+
+Compute test 5.
+ (*
+   = {| round := 5; n2 := 2; n1 := 3; nh := 5 |}
+     : state
+ *)
+
+Compute test 2000.
+(*
+  = {| round := 2000; n2 := 1; n1 := 90; nh := 1102 |}
+     : state
+*)
+\end{Coqsrc}
+
+The battle we study seems to be awfully long. Let us concentrate our
+tests on some particular events : the states where $\texttt{nh}=0$
+
+From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} will be
+equal to zero.
+
+ \begin{Coqsrc}
+ Compute test 10.
+(*
+
+    = {| round := 10; n2 := 2; n1 := 3; nh := 0 |}
+     : state
+*)
+ \end{Coqsrc}
+
+Thus, $ (1 + 11)$ rounds later, the \texttt{n1} field will be equal to $2$, and 
+\texttt{nh}  will equal to $0$.
+
+\begin{Coqsrc}
+Compute test 22.
+(*
+
+ = {| round := 22; n2 := 2; n1 := 2; nh := 0 |}
+     : state
+*)
+\end{Coqsrc}
+
+\begin{Coqsrc}
+ Compute test 46.
+(*
+
+ = {| round := 46; n2 := 2; n1 := 1; nh := 0 |}
+     : state
+
+*)
+
+Compute test 94.
+
+(*
+
+ = {| round := 94; n2 := 2; n1 := 0; nh := 0 |}
+     : state
+
+*) 
+\end{Coqsrc}
+
+Next round, we decrement \texttt{n2} and set \texttt{n1} to $95$.
+
+\begin{Coqsrc}
+ Compute test 95.
+
+(*
+
+  = {| round := 95; n2 := 1; n1 := 95; nh := 0 |}
+     : state
+
+*)
+\end{Coqsrc}
+
+We now have some intuition of the sequence.
+It looks like the next ``\texttt{nh}=0'' event will happen at the $192=2(95+1)$-th round, then at the $2(192+1)$-th round.
+
+\begin{Coqsrc}
+ Definition doubleS (n : nat) := 2 * (S n).
+
+Compute test (doubleS 95).
+
+(**
+ = {| round := 192; n2 := 1; n1 := 94; nh := 0 |}
+     : state
+ *)
+
+Compute test (iterate doubleS 2 95).
+
+(*
+  = {| round := 386; n2 := 1; n1 := 93; nh := 0 |}
+     : state
+*)
+\end{Coqsrc}
+
+\subsection{Proving \dots}
+We are now able to reason about the sequence of transitions defined by our hydra battle. Instead of using the data-type \texttt{state} we study the relationship
+between different configurations of the battle.
+
+Let us define a binary relation associated with every round of the battle.
+In the following definition \texttt{i} is associated with the round number (or date, if we consider a discrete time), and \texttt{a}, \texttt{b}, \texttt{c} respectively associated with the number of \texttt{h2}, \texttt{h1} and heads connected to the hydra's foot.
+
+\begin{Coqsrc}
+Inductive one_step (i: nat) :
+  nat -> nat -> nat -> nat -> nat -> nat -> Prop :=
+| step1 : forall a b c,  one_step i a b (S c) a b c
+| step2 : forall a b ,  one_step i a (S b) 0 a b (S i)
+| step3  : forall a, one_step i (S a) 0 0 a (S i) 0.
+\end{Coqsrc}
+
+The relation between \texttt{one\_step} and the rules of hydra battle is asserted by the following lemma.
+
+\label{lemma:step-battle}
+
+\begin{Coqsrc}
+Lemma step_battle : forall i a b c a' b' c', 
+   one_step i a b c a' b' c' ->
+   battle standard i (hyd  a b c)  (S i) (hyd a' b' c').
+\end{Coqsrc}
+
+Next, we define ``big steps'' as the transitive closure of \texttt{one\_step},
+and reachability (from the initial configuration of figure~\ref{fig:hinit} at time $0$).
+
+\begin{Coqsrc}
+ Inductive steps : nat -> nat -> nat -> nat ->
+                  nat -> nat -> nat -> nat -> Prop :=
+| steps1 : forall i a b c a' b' c',
+    one_step i a b c a' b' c' -> steps i a b c (S i) a' b' c'
+| steps_S : forall i a b c j a' b' c' k a'' b'' c'',
+    steps i a b c j a' b' c' ->
+    steps j a' b' c' k a'' b'' c'' ->
+    steps i a b c k  a'' b'' c''.
+
+Definition reachable (i a b c : nat) : Prop :=
+  steps 3 3 0 0 i a b c.
+\end{Coqsrc}
+
+The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{battle}.
+
+\begin{Coqsrc}
+ Lemma steps_battle : forall i a b c j a' b' c', 
+   steps i a b c j a' b' c' ->
+   battle standard i (hyd  a b c)   j  (hyd a' b' c').
+\end{Coqsrc}
+
+Thus, any result about \texttt{steps} will be applicable to standard battles.
+Using the predicate \texttt{steps} our study of the length of the considered battle
+can  be decomposed into three parts:
+
+\begin{enumerate}
+\item  Characterization of regularities of some events
+\item Study of the beginning of the battle
+\item Computing the exact length of the battle.
+\end{enumerate}
+
+First, we prove that, if at round $i$ the hydra is equal to
+(\texttt{hyd a (S b) 0}), then it will be equal to (\texttt{hyd a b 0}) at the $2(i+1)$-th round.
+
+\begin{Coqsrc}
+Lemma LS : forall c a b i,  steps i a b (S c) (i + S c) a b 0.
+Proof.
+  induction c.
+ -   intros;  replace (i + 1) with (S i).
+     + repeat constructor.
+     + ring.
+ -  intros; eapply  steps_S.
+   +   eleft;   apply rule1.
+   +   replace (i + S (S c)) with (S i + S c) by ring;  apply IHc.
+Qed.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma doubleS_law : forall  a b i, steps i a (S b) 0 (doubleS i) a b 0.
+Proof.
+  intros;  eapply steps_S.
+  +   eleft;   apply step2.
+  +   unfold doubleS; replace (2 * S i) with (S i + S i) by ring; 
+        apply LS.
+Qed.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma reachable_S  : forall i a b, reachable i a (S b) 0 ->
+                                   reachable (doubleS i) a b 0.
+Proof.
+  intros; right with  (1 := H); apply doubleS_law.
+Qed.
+\end{Coqsrc}
+
+From now on, the lemma \texttt{reachable\_S} allows us to watch larger steps of 
+the battle.
+
+\begin{Coqsrc}
+ Lemma L4 : reachable 4 2 4 0.
+Proof.
+  left; constructor.
+Qed.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma L10 : reachable 10 2 3 0.
+Proof.
+  change 10 with (doubleS 4).
+  apply reachable_S, L4.
+Qed.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma L22 : reachable 22 2 2 0.
+Proof.
+  change 22 with (doubleS 10).
+  apply reachable_S, L10.
+Qed.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma L46 : reachable 46 2 1 0.
+Proof.
+  change 46 with (doubleS 22); apply  reachable_S, L22.
+Qed.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma L94 : reachable 94 2 0 0.
+Proof.
+  change 94 with (doubleS 46); apply reachable_S, L46.
+Qed.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma L95 : reachable 95 1 95 0.
+Proof.
+  eapply steps_S.
+  -  eexact L94.
+  -  repeat constructor.
+Qed.
+\end{Coqsrc}
+
+\subsection{Giant steps}
+
+We are now able to make bigger steps in the simulation of the battle.
+First, we iterate the lemma \texttt{reachable\_S}.
+
+\begin{Coqsrc}
+Lemma Bigstep : forall b i a , reachable i a b 0 ->
+                               reachable (iterate doubleS b i) a 0 0.
+ Proof.
+  induction b.
+  -  trivial.
+  -  intros;  simpl;   apply reachable_S in H.
+     rewrite <- iterate_comm; now apply IHb.
+ Qed.
+\end{Coqsrc}
+
+Applying lemmas \texttt{BigStep} and \texttt{L95} we make a first jump.
+
+\begin{Coqsrc}
+ Definition M := (iterate doubleS 95 95).
+
+Lemma L2_95 : reachable M 1 0 0.
+Proof.
+  apply Bigstep,  L95.
+Qed.
+\end{Coqsrc}
+
+Figure~\ref{fig:HM}  represents the hydra at the $M$-th round.
+At the $(M+1)$-th round, it will look like in fig~\ref{fig:HM-plus1}.
+
+\begin{figure}[htb]
+\centering
+\begin{tikzpicture}[very thick, scale=0.5]
+\node (foot) at (2,0) {$\bullet$};
+\node (N1) at (2,2) {$\bullet$};
+\node (N2) at (3,4) {$\Smiley[2][green]$};
+\node (N3) at (1,4) {$\Smiley[2][green]$};
+\draw (foot) -- (N1);
+\draw (N1) to [bend right =15] (N2);
+\draw (N1) to  [bend left=15](N3);
+\end{tikzpicture}
+\caption{\label{fig:HM}}
+The state of the hydra after $M$ rounds.
+% The hydra \texttt{h} of the proof that \(\omega^2\) is too small for proving Hercules' victory
+
+\end{figure}
+
+\begin{figure}[htb]
+\centering
+\begin{tikzpicture}[very thick, scale=0.5]
+\node (foot) at (10,0) {$\bullet$};
+\node (N1) at (0,5) {$\bullet$};
+\node (N12) at (0,8) {$\Smiley[2][green]$};
+\node (N2) at (2,5) {$\bullet$};
+\node (N22) at (2,8) {$\Smiley[2][green]$};
+\node (N3) at (4,5) {$\bullet$};
+\node (N32) at (4,8) {$\Smiley[2][green]$};
+\node (N4) at (6,5) {$\bullet$};
+\node (N42) at (6,8) {$\Smiley[2][green]$};
+
+\node (Ndots) at (12,8) {\Huge $\dots$};
+\node (Ndots2) at (12,5) {\Huge $\dots$};
+
+\node (N8) at (18,5) {$\bullet$};
+\node (N82) at (18,8) {$\Smiley[2][green]$};
+\node (N9) at (20,5) {$\bullet$};
+\node (N92) at (20,8) {$\Smiley[2][green]$};
+
+\draw (foot) -- (N1);
+\draw (foot) -- (N2);
+\draw (foot) -- (N3);
+\draw (foot) -- (N4);
+\draw (foot) -- (N8);
+\draw (foot) -- (N9);
+\draw (N1) to  (N12);
+\draw (N2) to  (N22);
+\draw (N3) to  (N32);
+\draw (N4) to  (N42);
+\draw (N8) to  (N82);
+\draw (N9) to  (N92);
+\end{tikzpicture}
+\caption{\label{fig:HM-plus1}}
+The state of the hydra after $M+1$ rounds (with $M+1$ heads).
+
+\end{figure}
+
+\begin{Coqsrc}
+Lemma L2_95_S : reachable (S M) 0 (S M) 0.
+Proof.
+  eright.
+  - apply L2_95.
+  -  left; constructor 3.
+Qed.
+\end{Coqsrc}
+
+Then, applying once more the lemma \texttt{BigStep}, we get the exact time when
+Hercules wins!
+
+\begin{Coqsrc}
+Definition N :=   iterate doubleS (S M) (S M).
+
+Theorem   SuperbigStep : reachable N  0 0 0 .
+Proof.
+  apply Bigstep, L2_95_S.
+Qed.
+\end{Coqsrc}
+
+We are now able to prove formally that the considered battle is 
+composed of $N$ steps.
+
+\begin{Coqsrc}
+Lemma Almost_done :
+  battle standard 3 (hyd 3 0 0) N (hyd 0 0 0).
+Proof. 
+  apply steps_battle, SuperbigStep.
+Qed.
+
+Theorem  Done :
+  battle standard 0 hinit N head.
+Proof.
+  eapply battle_trans.
+  -   apply Almost_done.
+  -  apply L_0_3.
+Qed.
+\end{Coqsrc}
+
+\subsection{A minoration lemma}
+
+Now, we would like to get an intuition of  how big the number $N$ is.
+For that purpose, we use a minoration of the function \texttt{doubleS} by the
+function (\texttt{fun n => 2 * n}).
+
+\begin{Coqsrc}
+Definition exp2 n := iterate (fun n => 2 * n) n 1.
+\end{Coqsrc}
+Using some facts (proven in \url{../src/html/BigBattle.html}),we get several  minorations.
+
+\begin{Coqsrc}
+Lemma minoration_0 : forall n,  2 * n <= doubleS n.
+
+Lemma minoration_1 : forall n x, exp2 n * x <= iterate doubleS n x.
+
+Lemma minoration_2 : exp2 95 * 95 <= M.
+
+Lemma minoration_3 : exp2 (S M) * S M <= N.
+
+Lemma minoration : exp2 (exp2 95 * 95) <= N.
+\end{Coqsrc}
+
+The number $N$ is greater than or  equal to $2^{2^{95}\times 95}.$ If we wrote $N$ in base $10$, $N$ would require at least $10^{30}$ digits!
+
+\section{Generic properties}
+
+The example we just studied shows that the termination of any battle may take a very long time. If we want to study hydra battles in general, we have to consider 
+any hydra and any strategy, both for Hercules and the hydra itself. So, we first  give some definitions, generally borrowed from transition systems vocabulary (see~\cite{tel_2000} for instance).
+
+\subsection{Liveliness}
+
+Let $B$ be an instance of \texttt{Battle}. We say that $B$ is \emph{alive} if
+for any configuration $(i,h)$, where $h$ is not a head, there exists a further step in class $B$.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
+
+\begin{Coqsrc}
+Definition Alive (B : Battle) :=
+  forall i h, 
+     h <> head -> {h' : Hydra |  B i h h'}.
+\end{Coqsrc}
+
+The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
+\url{../src/html/hydras.Hydra.Hydra_Theorems.html} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
+
+\begin{Coqsrc}
+Theorem Alive_free: Alive free.
+
+Theorem Alive_standard: Alive standard.  
+\end{Coqsrc}
+
+Both theorems are proved with the help of the  following strongly specified function:
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
+
+\begin{Coqsrc}
+Definition  next_round_dec n :
+ forall h , (h = head) + {h' : Hydra & {R1 h h'} + {R2  n h  h'}}.
+\end{Coqsrc}
+
+\subsection{Termination}
+
+The termination of all battles is naturally expressed by the predicate \texttt{well\_founded} defined in the module \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf} 
+ of the Standard Library.
+
+\index{Predicates!Termination}
+
+\begin{Coqsrc}
+Definition Termination :=  well_founded (transp _ round).
+\end{Coqsrc}
+
+Let $B$ be an instance of class \texttt{Battle}. A \emph{variant} for $B$ consists
+in a well-founded relation $<$  on some type \texttt{A}, and a function
+(also called a \emph{measure}) \texttt{m:Hydra->A} such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $B$, the inequality $m(h')<m(h)$ holds.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
+
+\index{Coq!Type classes}
+\label{sect:hvariant-def}
+
+\begin{Coqsrc}
+Class Hvariant {A:Type}{Lt:relation A}(Wf: well_founded Lt)(B : Battle)
+  (m: Hydra -> A):   Prop :=
+  {variant_decr :forall i h h',
+      h <> head ->
+      battle_r  B i  h h' -> Lt (m h') (m h)}.
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+ Prove that, if there is an instance of (\texttt{Hvariant Lt wf\_Lt $B$ $m$}), then there exists no infinite battle in  $B$.
+\end{exercise}
+
+\subsection{Two small proofs of impossibility}
+\index{Maths!Proofs of impossibility}
+
+When one wants to prove a termination theorem with the help of a variant, 
+one has to consider first a well-founded set $(A,<)$, then a strictly decreasing measure on this set.
+The following two lemmas show that if  the order structure $(A,<)$ is too simple, it is useless to look for a convenient measure, which simply no exists.
+Such kind of result is useful, because it saves you time and effort.
+
+\subsubsection{No variant on Peano.lt}
+\label{omega-case}
+
+The best known well-founded order is the natural order on the set $\mathbb{N}$ of natural numbers (the type \texttt{nat} of Standard library).
+It would be interesting to look for some measure $m:\texttt{nat}\arrow\texttt{nat}$ and prove it is a variant.
+
+Unfortunately, we can prove that 
+\emph{no} instance of class \texttt{WfVariant round Peano.lt $m$} can be built, where
+$m$ is \emph{any} function of type \texttt{Hydra $\arrow$ nat}.
+
+Let us present the main steps of that proof, the script of which  is in the module ~\href{../src/html/hydras.Hydra.Omega_Small.html}{Hydra/Omega\_Small.v} \footnote{ The name of this file means ``the ordinal $\omega$ is too small for proving the termination of [free] hydra battles ''.
+In effect, the elements of $\omega$, considered as a set, are just the natural numbers (see next chapter for more details)}.
+
+%\subsubsection{Preliminaries}
+
+Let us assume there is a variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
+    the  termination of all hydra battles.
+
+\begin{Coqsrc}
+Section Impossibility_Proof.
+ Variable m : Hydra -> nat.
+ Hypothesis Hvar : Hvariant lt_wf free m.
+\end{Coqsrc}
+
+We define an injection from the type \texttt{nat} into \texttt{Hydra}.
+For any natural number $i$, $\iota(i)$ is the hydra composed of a foot and
+$i+1$ heads at height $1$. For instance, Fig.~\ref{fig:flower} represents the hydra $\iota(3)$.
+
+\begin{figure}[htb]
+\centering
+\begin{tikzpicture}[very thick, scale=0.5]
+\node (foot) at (4,0) {$\bullet$};
+\node (N1) at (2,2) {$\Smiley[2][green]$};
+\node (N2) at (4,2) {$\Smiley[2][green]$};
+\node (N3) at (6,2) {$\Smiley[2][green]$};
+\node (N4) at (8,2) {$\Smiley[2][green]$};
+\draw (foot) to [bend left =25] (N1);
+\draw (foot) to [bend left =15] (N2);
+\draw (foot) to [bend right =15] (N3);
+\draw (foot) to [bend right =25] (N4);
+\end{tikzpicture}
+\caption{\label{fig:flower}
+The hydra $\iota(3)$}
+\end{figure}
+
+  \begin{Coqsrc}
+  Let iota (i: nat) := hyd_mult head (S i).    
+  \end{Coqsrc}
+
+Let us consider now some hydra \texttt{big\_h} out of the range of the injection $\iota$ (see Fig.~\vref{fig:h-omega-omega}).
+
+\begin{figure}[htb]
+\centering
+\begin{tikzpicture}[very thick, scale=0.5]
+\node (foot) at (2,0) {$\bullet$};
+\node (N1) at (2,2) {$\bullet$};
+\node (N2) at (2,4) {$\Smiley[2][green]$};
+\draw (foot) -- (N1);
+\draw (N1) to  (N2);
+\end{tikzpicture}
+\caption{\label{fig:h-omega-omega}}
+ The hydra \texttt{big\_h}.
+\end{figure}
+
+ \begin{Coqsrc}
+  Let big_h := hyd1 (hyd1 head).
+ \end{Coqsrc}
+
+ Using the functions $m$ and $\iota$, we define a second hydra \texttt{small\_h}, and show
+ there is a one-round battle that transforms \texttt{big\_h} into \texttt{small\_h}. Please note that,
+due to the hypothesis \texttt{Hvar}, we are interested in the termination of \emph{free} battles. 
+There is no problem to consider a round with \texttt{m big\_h} as replication factor.
+
+  \begin{Coqsrc}
+ Let small_h := iota (m big_h).
+
+ Fact big_to_small : big_h -1-> small_h.
+ Proof.
+      exists (m big_h); right; repeat constructor.     
+ Qed.
+    \end{Coqsrc}
+
+But, by hypothesis, $m$ is a variant. Hence, we infer the following inequality.
+
+ \begin{Coqsrc}
+Lemma m_lt : m small_h < m big_h.
+ \end{Coqsrc}
+
+In order to get a contradiction, it suffices to  prove the inequality
+\texttt{m big\_h <= m small\_h}, i.e.,  \texttt{m big\_h <= m (iota (m big\_h))}.
+
+More generally, we prove the following lemma:
+
+\begin{Coqsrc}
+Lemma m_ge : forall i:nat , i <= m (iota i).
+\end{Coqsrc}
+
+Intuitively, it means that, from any hydra of the form (\texttt{iota $i$}), the battle will 
+take (at least) $i$ rounds. Thus the associated measure cannot be less than $i$.
+Technically, we prove this lemma by Peano induction on $i$.
+
+\begin{itemize}
+\item The base case $i=0$ is trivial
+\item Otherwise, let $i$ be any natural number and assume  the inequality
+  $i \leq m(\iota(i))$.
+  \begin{enumerate}
+  \item  But the hydra $\iota(S(i))$ can be transformed in one round into
+    $\iota(i)$ (by losing its righmost head, for instance)
+  \item Since $m$ is a variant, we have $m(\iota(i)) < m(\iota(S(i)))$,
+    hence  $i< m(\iota(S(i)))$, which implies  $S(i)\leq  m(\iota(S(i)))$.
+  \end{enumerate}
+\end{itemize}
+
+ Then our proof is almost finished.
+
+   \begin{Coqsrc}
+Theorem Contradiction : False.
+Proof.
+ apply (Nat.lt_irrefl (m big_h));
+   apply  Lt.le_lt_trans with (m small_h).
+  - apply m_ge.
+  - apply m_lt.
+Qed.
+
+End Impossibility_Proof.
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+Prove that there exists no variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
+    the  termination of all \emph{standard} battles.
+\end{exercise}
+
+\subsubsection{Conclusion}
+
+In order to build a variant for proving the termination of all hydra battles, we need to consider order structures more complex than the usual order on type \texttt{nat}. 
+The notion of \emph{ordinal number} provides a catalogue of well-founded order types.
+For a reasonably large bunch of ordinal numbers, \emph{ordinal notations} are data-types which allow the \coq{} user to define functions, to compute and prove some properties, for instance by reflection.
+
+So, the next chapters will present ordinal numbers and ordinal notations, and we will
+look for a variant complex enough for proving the termination of all battles, but not too complex if possible.
+
+%--------------------------------------------------------------
+
+\chapter{Introduction to ordinal numbers and ordinal notations}
+
+The proof of termination of all hydra battles presented in~\cite{KP82} is based
+on \emph{ordinal numbers}.
+From a mathematical point of view, an ordinal is a representant of an equivalence class for isomorphims of  totally ordered well-founded sets.
+
+More intuitively, let us have a look at Figure~\ref{fig:ordinal-sequence}. It presents a small sequence of ordinal numbers, which extends the sequence of natural numbers.
+
+\begin{figure}[h]
+  \centering
+\fbox{\Large
+  \begin{minipage}{1.0\linewidth}
+  \begin{align*}
+     &\textcolor{blue}{0},\,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,\ldots\\
+&\textcolor{red}{\omega},\,\omega+1,\omega+2,\omega+3,\ldots\\
+&\textcolor{red}{\omega\times 2},\,\omega\times 2+1,\ldots, \textcolor{red}{\omega\times 3},\,\omega\times 3+1,\ldots, \textcolor{red}{\omega\times 4},\ldots,\\
+&\textcolor{red}{\omega^2},\ldots, \textcolor{red}{\omega^2\times 42},\ldots,\textcolor{red}{\omega^3},\ldots, \textcolor{red}{\omega^4},\omega^4+1,\ldots,\\
+&\textcolor{red}{\omega^\omega},\ldots, \textcolor{red}{\omega^\omega+\omega^7\times 8},\ldots,\textcolor{red}{\omega^\omega\times 2},\omega^\omega\times 2+1, \ldots,\\
+&\textcolor{red}{\omega^{\omega^\omega}},\ldots, \textcolor{red}{\omega^{\omega^\omega}+\omega^\omega\times 42+ \omega^{55}+\omega}, \ldots, \textcolor{red}{\omega^{\omega^{\omega+1}}}, \omega^{\omega^{\omega+1}}+1,\dots\\
+& \textcolor{red}{\epsilon_0 (= \omega^{\epsilon_0)}}, \epsilon_0+1, \epsilon_0+2, \epsilon_0+3, \ldots,\\
+& \textcolor{red}{\epsilon_1}, \ldots, \textcolor{red}{\epsilon_2}, \ldots, \textcolor{red}{\epsilon_\omega},\ldots \\
+& \textcolor{red}{\Gamma_0}, \Gamma_0+1, \Gamma_0+2, \Gamma_0+3,\ldots, \textcolor{red}{\Gamma_0+\omega}, \ldots,\\
+&\ldots
+  \end{align*}   
+  \end{minipage}}
+
+  \caption{A short overview of the sequence of ordinal numbers}
+  \label{fig:ordinal-sequence}
+\end{figure}
+
+Let us comment some features of this table:
+
+\begin{itemize}
+\item The ordinals are listed in a strictly increasing order. 
+\item Dots : ``$\ldots$'' stand for an infinite sequence of ordinals, not shown for lack of space.
+For instance, the ordinal $42$ is not shown in the first line, but it exists, between $17$ and $\omega$.
+\item Each ordinal printed in black is the immediate successor of another ordinal.
+We call it a \emph{successor} ordinal.
+For instance, $12$ is the successor of $11$, and $\omega^4+1$ the successor of $\omega^4$.
+\item Ordinals that  follow dots immediately (displayed in red) are called \emph{limit ordinals}.
+With respect to the order induced by this sequence, any limit ordinal $\alpha$ is the least upper bound of  the set $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
+\item
+For instance $\omega$ is the least upper bound of the set of all finite ordinals (in the first line). It is also the first limit ordinal, and the first infinite ordinal, in the sense that 
+the set $\mathbb{O}_\omega$ is infinite.
+\item The ordinal $\epsilon_0$ is the first number which is equal to its own exponential of base $\omega$.
+It plays an important role in proof theory, and is particularly studied in chapters~\ref{chap:T1} to \ref{chap:alpha-large}.
+\item All the ordinals shown in this list are either the ordinal \textcolor{blue}{$0$},
+a successor ordinal, or a \textcolor{red}{limit ordinal}.
+\end{itemize}
+
+\section{The mathematical point of view}
+
+\subsection{Well-ordered sets}
+Let us start with some definitions.
+A  \emph{well-ordered set} is a set provided with a binary relation $<$ which has the following properties.
+\begin{description}
+\item[irreflexivity] : $\forall x\in A, x\not< x$
+\item[transitivity] : $\forall x\,y\,z\in A, x<y \Rightarrow y<z \Rightarrow x<z$
+\item[trichotomy]: $\forall x\,y\in A, x<y \vee x = y \vee y < x$
+\item[well foundedness] $<$ is well-founded (every element of $A$ is accessible)\footnote{In classical mathematics, we would say that there is no infinite sequence $a_1>a_2> \dots a_n> a_{n+1}\dots$ in $A$.
+Please refer to any documenation on \coq{} for having more details on well-foundedness and accessibility.}.
+
+\end{description}
+
+The best known examples of well-ordered sets are the set $\mathbb{N}$ of natural numbers (with the usual order $<$), as well as any finite segment $[0,i)=\{j\in\mathbb{N}|j<i\}$.
+The disjoint union of two copies of $\mathbb{N}$, \emph{i.e.} the set $\{0,1\}\times\mathbb{N}$ is also well-ordered,
+with respect to the order below:
+
+\begin{align*}
+(i,j) < (i,k) & \;\textit{if} \; j < k\\
+(0,k) < (1,l) & \;\textit{for\,any}\;k \;\textit{and} \; l
+\end{align*}
+
+\subsection{Ordinal numbers}
+
+\index{Maths!Ordinal numbers}
+
+Let $(A,<_A)$ and $(B,<_B)$ two well-ordered sets. $A$ and $B$ are said to have \emph{the same order type} if 
+there exists a strictly monotonous bijection $b$ from $A$ to $B$, \emph{i.e.} which verifies the proposition
+$\forall x\,y\in A, x <_A y \Rightarrow b(x) <_B  b(y)$.
+
+Having the same order type is an equivalence relation between well-ordered sets. Ordinal numbers (in short \emph{ordinals}) are descriptions (\emph{names}) of the equivalence classes.
+For instance, the order type of $(\mathbb{N},<)$ is associated with the ordinal called  $\omega$, and the order we considered on 
+the disjoint union of $\mathbb{N}$ and itself is named $\omega+\omega$.
+
+In a set-theoretic framework, one can consider any ordinal $\alpha$ as a well-ordered set, whose  elements are just the ordinals strictly less than $\alpha$, \emph{i.e.} the \emph{segment} $\mathbb{O}_\alpha=[0, \alpha)$.
+So, one can speak about \emph{finite}, \emph{infinite}, \emph{countable}, etc., ordinals.
+Nevertheless, since we work with type theory,
+we do not identify ordinals as sets of ordinals, but the correspondance between ordinals and sets of ordinals is the function that maps $\alpha$ to $\mathbb{O}_\alpha$.
+For instance $\mathbb{O}_\omega=\mathbb{N}$, and $\mathbb{O}_7=\{0,1,2,3,4,5,6\}$.
+
+We cannot cite all the litterature published on ordinals since Cantor's book 
+\cite{cantorbook}, and 
+leave it to the reader to explore the bibliography.
+
+\section{Ordinal numbers in Coq}
+
+Two kinds of representation of ordinals are defined herein.
+
+\begin{itemize}
+\item A ``mathematical'' representation of the set of countable ordinal numbers, afer Kurt Schütte~\cite{schutte}.
+This representation uses several (hopefully harmess) axioms.
+It is used as a reference for proving the correctness of ordinal notations.
+\item A family of \emph{ordinal notations}, \emph{i.e.} data types used to represent segments $[0,\mu)$, where $\mu$ is some countable ordinal.
+Each ordinal notation is defined inside the Calculus of Inductive Constructions (without axioms).
+Many functions are defined, allowing proofs by computation.
+Note that proofs of
+correctness of a given ordinal notation with respect to Schütte's model use axioms.
+Please execute the \texttt{Print Assumptions} command in case of doubt.
+\end{itemize}
+
+\section{Countable ordinals}
+
+Chapter~\ref{chap:schutte} of this document presents an adaptation to \coq{} of an axiomatization in classical logic of the set of countable ordinals by K. Schütte~\cite{schutte}. 
+That formalization is quite complex and technical and unshamedly non-constructive,  so we put its description  in the last chapter of this document.
+
+Please note that Schütte considers the (uncountable) set $\mathbb{O}$ of all countable ordinals.
+This set is well ordered (which is one of Schütte's axioms), and associates to any ordinal $\alpha$ the segment $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
+
+In our adaptation to \coq{}, we declare a type \texttt{Ord}, a binary relation \texttt{lt} (with infix notation \texttt{``\_<\_''}, and assume Schütte's axiom. In Chapter~\ref{chap:schutte},
+we derive some interesting properties of countable ordinals from these axioms.
+
+\section{Ordinal notations}
+
+Fortunately, the ordinals we need for  studying hydra battles are much simpler than Schütte's, and can be represented as quite simple data types in \gallina.
+So, we will use \emph{ordinal notations} (also called \emph{[ordinal] notation systems}).
+
+Let $\alpha$ be some (countable) ordinal; 
+in \coq{} terms, we call ordinal notation for $\alpha$ a structure composed 
+of:
+\begin{itemize}
+\item A data type $A$ for representing all ordinals strictly below $\alpha$,
+\item A well founded order $<$ on $A$, 
+\item A function for comparing two ordinals, whose correctnes implies that the reflexive closure of $<$ is a \emph{total order}.
+\end{itemize}
+
+Such a structure can be proved correct relatively to another ordinal notation or
+to Schütte's model.
+
+\subsubsection*{Ordered types}
+
+The library ~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Classes.RelationClasses.html}{%
+Coq.Classes.RelationClasses} contains some definitions and facts about binary relations, among them strict orders.
+
+\index{Coq!Type classes}
+
+\begin{Coqsrc}
+Variable A: Type.
+
+  Class StrictOrder (R : relation A) : Prop := {
+    StrictOrder_Irreflexive :> Irreflexive R ;
+    StrictOrder_Transitive :> Transitive R }.
+\end{Coqsrc}
+
+\subsection{A class for ordinal notations}
+
+The following class definition, parameterized with a type $A$, a binary relation \texttt{lt} on $A$, specifies that \texttt{lt} is a well-founded strict order.
+The reflexive closure of \texttt{lt}, (called \texttt{le}, for ``less  or equal than'') is a total decidable order, implemented through a comparison function \texttt{compare}.
+
+\vspace{4pt}
+\noindent\emph{From Library~\href{../src/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}}
+
+\index{Coq!Type classes}
+\label{types:ON}
+
+\begin{Coqsrc}
+Class ON {A:Type}(lt: relation A)
+      (compare : A -> A -> comparison)  :=
+  {
+  sto :> StrictOrder lt;
+  wf : well_founded lt;
+  compare_correct :
+    forall alpha beta:A,
+      CompareSpec (alpha=beta) (lt alpha beta) (lt beta alpha)
+                  (compare alpha beta);
+  }.
+\end{Coqsrc}
+
+The following definitions allow us to make implicit several guessable arguments.
+\label{sect:on-lt-notation}
+\label{sect:on-le-notation}
+\begin{Coqsrc}
+Definition on_t  {A:Type}{lt: relation A}
+            {compare : A -> A -> comparison}
+            {on : ON lt compare} := A.
+
+Definition on_compare {A:Type}{lt: relation A}
+            {compare : A -> A -> comparison}
+            {on : ON lt compare} := compare.
+
+Definition on_lt {A:Type}{lt: relation A}
+           {compare : A -> A -> comparison}
+           {on : ON lt compare} := lt.
+Infix "o<" := on_lt : ON_scope.
+
+Definition on_le  {A:Type}{lt: relation A}
+           {compare : A -> A -> comparison}
+           {on : ON lt compare} :=
+  clos_refl _ on_lt.
+
+Infix "o<=" := on_le : ON_scope.
+\end{Coqsrc}
+
+\begin{remark}
+The infix notations \texttt{o<} and \texttt{o<=} were defined in order to make apparent the distinction between the various notation scopes that may co-exist in a same statement.
+So the infix \texttt{<} and \texttt{<=} are reserved to the natural numbers.
+In the mathematical formulas, we still use $<$ and $\leq$ for comparing ordinals.
+\end{remark}
+
+\subsection{Ordinal notations and measures for proving termination}
+\label{sect:measure-ON}
+
+The following lemma (together with the type class mechnism) allows us to use simply  measures towards an ordinal notation.
+It is just an application of  the libraries \texttt{Coq.Wellfounded.Inverse\_Image}
+and  \texttt{Coq.Wellfounded.Inclusion}.
+
+\begin{Coqsrc}
+Definition measure_lt {A:Type}{lt: relation A}
+            {compare : A -> A -> comparison}
+            {on : ON lt compare}
+            {B : Type}
+  (m : B -> A) : relation B :=
+  fun x y =>  on_lt (m x) (m y).
+
+Lemma wf_measure  {A:Type}(lt: relation A)
+            {compare : A -> A -> comparison}
+            {on : ON lt compare}
+            {B : Type}
+            (m : B -> A) :
+  well_founded (measure_lt m). 
+\end{Coqsrc}
+
+A simple example of application is given in Sect.~\vref{sect:merge-example}.
+
+\section{Examples of ordinal notations}
+
+\subsection{The ordinal \texorpdfstring{$\omega$}{omega}}
+
+The simplest example of ordinal notation is built over the type \texttt{nat} of \coq's standard library. We have only to apply already proven lemmas about Peano numbers.
+
+\vspace{4pt}
+\noindent\emph{From Library~\href{../src/html/hydras.OrdinalNotations/ON_Omega.html}{OrdinalNotations.ON\_Omega}}
+
+\begin{Coqsrc}
+Global Instance Omega : ON  Peano.lt Nat.compare.
+Proof.
+ split.
+ - apply Nat.lt_strorder.
+ - apply Wf_nat.lt_wf.
+ - apply Nat.compare_spec.
+Qed.
+\end{Coqsrc}
+
+\subsection{Sum of two ordinal notations}
+
+Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective types \texttt{A} and \texttt{B}.
+
+ We consider a new strict order
+on the disjoint sum of the associated types, by putting all elements of \texttt{A} before the elements of \texttt{B} (thanks to Standard Library's relation operator \texttt{le\_AsB}).
+(please look at Figure~\ref{fig:omega-plus-omega} for an example).
+
+\emph{From Library~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}}.
+
+\begin{Coqanswer}
+Inductive
+le_AsB (A B : Type) (leA : A -> A -> Prop) (leB : B -> B -> Prop)
+  : A + B -> A + B -> Prop :=
+| le_aa : forall x y : A, leA x y -> le_AsB A B leA leB (inl x) (inl y)
+| le_ab : forall (x : A) (y : B), le_AsB A B leA leB (inl x) (inr y)
+| le_bb : forall x y : B, leB x y -> le_AsB A B leA leB (inr x) (inr y)
+\end{Coqanswer}
+
+\vspace{4pt}
+\noindent\emph{From Library~\href{../src/html/hydras.OrdinalNotations/ON_plus.html}{OrdinalNotations.ON\_plus}}
+
+\begin{Coqsrc}
+Section Defs.
+
+  Context `(ltA: relation A)
+          (compareA : A -> A -> comparison)
+          (NA: ON ltA compareA).
+  Context `(ltB: relation B)
+          (compareB : B -> B -> comparison)
+          (NB: ON ltB compareB).
+
+Definition t := (A + B)%type.
+Arguments inl  {A B} _.
+Arguments inr  {A B} _.
+
+Definition lt : relation t := le_AsB _ _ ltA ltB.
+\end{Coqsrc}
+
+Before building an instance of \texttt{ON}, we have to define a comparison function.
+
+\begin{Coqsrc}
+Definition compare (alpha beta: t) : comparison :=
+   match alpha, beta with
+     inl _, inr _ => Lt
+   | inl a, inl a' => compareA a a'
+   | inr b, inr b' => compareB b b'
+   | inr _, inl _ => Gt
+  end.
+
+Lemma compare_correct alpha beta :
+    CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
+                            (compare alpha beta).
+\end{Coqsrc}
+
+The Lemma \texttt{Wellfounded.Disjoint\_Union.wf\_disjoint\_sum} of Standard Library
+helps us to prove that our order \texttt{lt} is well-founded.
+
+\begin{Coqsrc}
+Global Instance ON_plus : ON lt compare.
+Proof.
+  split.
+  - apply lt_strorder.
+  -  apply lt_wf.
+  - apply compare_correct.
+Qed.
+\end{Coqsrc}
+
+\subsection{The ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
+
+The ordinal $\omega+\omega$ (also known as $\omega\times 2$) may be represented as the concatenation 
+of two copies of $\omega$ (Figure~\ref{fig:omega-plus-omega}).
+
+\begin{figure}[h]
+   \centering
+   \begin{tikzpicture}[very thick, scale=0.5]
+\begin{scope}[color=blue]
+\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
+\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
+\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
+\node (Adots) at (6,0) {$\ldots$};
+\node(An) at (8,0)[label=below:$n$]{$\bullet$};
+\node(A2) at (10,0)[label=below:$n+1$]{$\bullet$};
+\node (Adots1) at (12,0) {$\ldots$};
+\end{scope}
+\begin{scope}[color=red]
+\node(B0) at (14,0)[label=below:$0$,label=above:$\omega$]{$\bullet$};
+\node(B1) at (16,0)[label=below:$1$, label=above:$\omega+1$]{$\bullet$};
+\node(B2) at (18,0)[label=below:$2$,label=above:$\omega+2$]{$\bullet$};
+\node (Bdots) at (20,0) {$\ldots$};
+\node (Bn) at (22,0) [label=below:$p$, label=above:$\omega+p$]{$\bullet$};
+\node (Bdots2) at (24,0) {$\ldots$};
+\end{scope}
+\end{tikzpicture}
+   \caption{\textcolor{blue}{$\omega+{\color{red}\omega}$}}
+   \label{fig:omega-plus-omega}
+ \end{figure}
+
+We can define this notation in \coq{} as an instance of \texttt{ON\_plus}.
+
+\vspace{4pt}
+\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
+
+\begin{Coqsrc}
+Definition Omega_plus_Omega := ON_plus Omega Omega.
+
+Existing Instance Omega_plus_Omega.
+Definition t := @ON_plus.t nat nat.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Example ex1 : inl 7 o< inr 0.
+Proof. constructor. Qed.
+\end{Coqsrc}
+
+We can now define abbreviations.
+For instance, the finite ordinals are represented by terms built with  the constructor \texttt{inl}, and the first infinite ordinal $\omega$ by the term \texttt{(inr 0)}.
+
+\begin{Coqsrc}
+Definition fin (i:nat) : t := inl i.
+Coercion fin : nat >-> t.
+
+Notation "'omega'" := (inr  0:t).
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Example ex1' : fin 7 o< omega.
+Proof. constructor. Qed.
+\end{Coqsrc}
+
+% \label{warning:coercions}
+% \index{Coq!Techniques!Coercions} 
+% \begin{remark}
+% Beware of coercions and notation scopes!
+% Let us consider the following goal:
+
+% \begin{Coqsrc}
+%  Goal (6 o< 8).
+%  auto with arith.
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+% 1 subgoal (ID 9)
+
+%   ============================
+%   6 o< 8
+% \end{Coqanswer}
+
+% Please keep in mind that the current notation scope interprets the infix \texttt{``<''} as the predicate \texttt{Omega\_plus\_omega.lt} and not \texttt{Nat.lt}.
+More,  the coercion mechanism converts the terms \texttt{6:nat} [resp.
+\texttt{8:nat} ]
+% into \texttt{inl 6} [resp. \texttt{inl 8}].  So, the initial goal is correctly interpreted by \coq{}, but not as an inequality between two natural numbers.
+
+% \begin{Coqsrc}
+% Set Printing All.
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+% 1 subgoal (ID 337)
+
+%   ============================
+%   @on_lt nat Peano.lt Nat.compare Omega (S (S (S (S (S (S O))))))
+%     (S (S (S (S (S (S (S (S O))))))))
+% \end{Coqanswer}
+
+% Anyway, the initial goal is provable, using \texttt{le\_AsB}'s first constructor.
+
+% \begin{Coqsrc}
+%   constructor; auto with arith.
+% Qed.
+% \end{Coqsrc}
+
+% \end{remark}
+%
+
+\section{Limits and Successors}
+
+Let us look again at our implementation of $\omega+\omega$. We can distinguish between the three kinds of ordinals seen in Fig~\ref{fig:ordinal-sequence}:
+
+\begin{itemize}
+\item The least ordinal, \texttt{(inl 0)}, also written \texttt{(fin 0)}.
+\item The limit ordinal $\omega$.
+\item The successor ordinals, either of the form \texttt{(inl (S $i$))} or \texttt{(inr (S $i$))}
+\end{itemize}
+
+\subsection{Definitions}
+It would be interesting to specify at the most generic level, what is a zero, a successor or a limit ordinal. Let $<$ be a strict order on a type $A$.
+
+\begin{itemize}
+\item A \emph{least} element is a minorant (in the large sense) of the full set  on $A$,
+\item $y$ is a \emph{successor} of $x$ if $x<y$ and there is no element between $x$ and $y$. We will also say that $x$ is a \emph{predecessor} of $y$.
+\item $x$ is a \emph{limit} if $x$ is not a least element, and for any $y$ such that $yo<x$,
+ there exists some $z$ such that $y<z<x$.
+\end{itemize}
+
+The following definitions are in Library \href{../src/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
+
+\begin{Coqsrc}
+Section A_given.
+  Variables (A : Type)  (lt: relation A).
+
+Local Infix "<" := lt.
+Local Infix "<=" := (clos_refl _ lt).
+
+Definition Least {sto : StrictOrder lt} (x : A):=
+  forall y,  x <= y.
+
+Definition Successor {sto : StrictOrder lt} (y x : A):=
+  x < y /\ (forall z,  x < z ->  z <  y -> False).
+
+Definition Limit {sto : StrictOrder lt}  (x:A)  :=
+  (exists w:A,  w < x) /\
+  (forall y:A, y < x -> exists z:A, y < z /\ z < x).
+\end{Coqsrc}
+
+\index{Exercises}
+\begin{exercise}
+Prove, that, in any ordinal notation system, that every ordinal has at most one predecessor, and at most one successor. 
+\end{exercise}
+
+\index{Exercises}
+\begin{exercise}
+Prove, that, in any ordinal notation system, that if $\beta$ is a successor of $\alpha$,
+then for any $\gamma$, $\gamma<\beta$ implies 
+$\gamma\leq\alpha$.
+\end{exercise}
+
+\subsection{Limits and successors in \texorpdfstring{$\omega+\omega$}{omega+omega}}
+
+Using the definitions above, we can prove the following lemma:
+
+\vspace{4pt}
+
+\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
+
+\begin{Coqsrc}
+Lemma limit_iff (alpha : t) : Limit alpha <-> alpha = omega.
+\end{Coqsrc}
+
+Regarding successors, let us introduce the following definition:
+
+\begin{Coqsrc}
+Definition succ (alpha : t) :=
+  match alpha with
+    inl n => inl (S n)
+  | inr n => inr (S n)
+  end.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma Successor_correct alpha beta : Successor beta alpha <->
+                                     beta = succ alpha.
+\end{Coqsrc}
+
+We can also check whether an ordinal is a successor by a simple pattern matching:
+
+\begin{Coqsrc}
+Definition succb (alpha: t) := match alpha with
+                                   inr (S  _) | inl (S _) => true
+                                 | _ => false
+                                 end.
+
+Lemma succb_correct (alpha : t) :
+    succb alpha <->  exists beta: t,   alpha = succ beta.
+\end{Coqsrc}
+
+Finally, the nature of any ordinal is decidable(inside this notation system) :
+
+\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}}
+\begin{Coqsrc}
+  Definition ZeroLimitSucc_dec {A:Type}{lt: relation A}
+           {compare : A -> A -> comparison}
+           {on : ON lt compare} :=
+  forall alpha,
+    {Least alpha} +
+    {Limit alpha} +
+    {beta: A | Successor alpha beta}.
+\end{Coqsrc}
+
+\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
+
+\begin{Coqsrc}
+Definition Zero_limit_succ_dec : ZeroLimitSucc_dec.
+\end{Coqsrc}
+
+\section{The ordinal \texorpdfstring{$\omega^2$}{omega^2}}
+
+The ordinal $\omega^2$ (also called $\varphi_0(2)$, see Chap.~\ref{chap:schutte}), is represented by the lexicographically ordered cartesian product $\mathbb{N}\times\mathbb{N}$.
+
+\vspace{4pt}
+\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Omega2.html}{ON\_Omega2}}
+
+\begin{Coqsrc}
+Definition t := (nat * nat)%type.
+
+Definition lt : relation t := lexico Peano.lt Peano.lt.
+Definition le := clos_refl _ lt.
+
+Infix "o<" := lt : o2_scope.
+Infix "o<=" := le : o2_scope.
+\end{Coqsrc}
+
+It is easy to represent the finite ordinals, successors and limits.
+
+\begin{Coqsrc}
+Definition zero: t := (0,0).
+
+Definition fin (n:nat) : t := (0, n).
+
+Notation "'F'" := fin : o2_scope.
+
+Coercion fin : nat >-> t.
+
+Definition succ (alpha : t) := (fst alpha, S (snd alpha)).
+
+Notation "'omega'" := (1,0) : o2_scope.
+
+Definition limitb (alpha : t) := 
+ match alpha with
+ |  (S _, 0) => true
+ | _ => false
+end.
+\end{Coqsrc}
+
+In order to build an ordinal notation out the type \texttt{t}, we have to define a comparison function, and prove that the order \texttt{lt} is well-founded.
+
+\begin{Coqsrc}
+Definition compare (alpha beta: t) : comparison :=
+  match Nat.compare (fst alpha) (fst beta) with
+    Eq => Nat.compare (snd alpha) (snd beta)
+  | c => c
+  end.
+
+Lemma compare_correct alpha beta :
+    CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
+                (compare alpha beta).
+
+Instance Omega2 : ON  lt compare.
+\end{Coqsrc}
+
+\subsection{Arithmetic of \texorpdfstring{$\omega^2$}{omega^2}}
+
+\subsubsection{Addition}
+
+We can define on \texttt{ON\_omega2} an addition which extends the addition on \texttt{nat}.
+
+\begin{Coqsrc}
+Definition  plus (alpha beta : t) : t :=
+  match alpha,beta with
+  | (0, b), (0, b') => (0, b + b')
+  | (0,0), y  => y
+  | x, (0,0)  => x
+  | (0, b), (S n', b') => (S n', b')
+  | (S n, b), (S n', b') => (S n + S n', b')
+  | (S n, b), (0, b') => (S n, b + b')
+   end.
+
+Infix "+" := plus : o2_scope.
+\end{Coqsrc}
+
+Please note that this operation is not commutative:
+
+\begin{Coqsrc}
+Example non_commutativity_of_plus :  omega + 3 <> 3 + omega.
+Proof.
+  cbn.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+1 subgoal (ID 237)
+
+  ============================
+(1, 3) <> omega.
+\end{Coqanswer}
+
+\begin{Coqsrc}
+ discriminate.
+Qed.
+\end{Coqsrc}
+
+\subsubsection{Multiplication}
+
+The restriction of ordinal multiplication to the segment $[0,\omega^2)$ is not a total function.
+For instance $\omega\times\omega= \omega^2$ is outside the set of represented values.
+Nevertheless, we can define two operations mixing natural numbers and ordinals.
+
+\begin{Coqsrc}
+(** multiplication of an ordinal and a natural number *)
+
+Definition mult_fin  (alpha : t) (p : nat): t :=
+  match alpha, p with
+ |  (0,0), _  => zero
+ |  _, 0 => zero
+ |  (0, n), p => (0, n * p)
+ |  ( n, b),  n' => ( n *  n', b)
+               end.
+Infix "*" := mult_fin : o2_scope.
+
+(** multiplication of  a natural number and an ordinal *)
+
+Definition fin_mult (n:nat)(beta : t) : t :=
+  match n, beta with
+ |  0, _  => zero
+ |  _, (0,0) => zero
+ |   n , (0,n') => (0, (n*n')%nat)
+ |  n, (n',p') => (n', (n * p')%nat)
+ end.
+
+Example e1 : fin_mult 3 (omega * 7 + 15) = omega * 7 + 45.
+Proof. reflexivity. Qed.
+
+Example e2 :  omega * 2 + 2 + omega * 3 + 10 = omega * 5 + 10.
+Proof. reflexivity. Qed.
+\end{Coqsrc}
+
+Multiplication with a finite ordinal and addition are related through the following lemma:
+
+\begin{Coqsrc}
+Lemma unique_decomposition alpha : 
+    exists! i j: nat,  alpha = omega * i + j.
+\end{Coqsrc}
+
+\subsection{A proof a termination using \texorpdfstring{$\omega^2$}{omega^2}} 
+\label{sect:merge-example}.
+
+Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily the function which merges two lists.
+
+\index{Coq!Commands!Function}
+
+\begin{Coqsrc}
+(* adapted from Pascal Manoury et al. *)
+
+Require Import Coq.Program.Wf List.
+Require Import FunInd Recdef.
+
+Section Merge.
+
+  Variable A: Type.
+
+  Local Definition m (p : list A * list A) :=
+    omega * length (fst p) + length (snd p).
+
+  Function  merge  (ltb: A -> A -> bool)
+          (xys: list A * list A)
+          {wf (measure_lt m) xys} :
+    list A :=
+    match xys with
+      (nil, ys) => ys
+    | (xs, nil) => xs
+    | (x :: xs, y :: ys) =>
+      if ltb x y then x :: merge  ltb (xs, (y :: ys))
+      else y :: merge  ltb ((x :: xs), ys)
+    end.
+
+  - intros; unfold m, measure_lt; cbn; destruct xs0; simpl; left; lia.
+  - intros; unfold m, measure_lt; cbn; destruct ys0; simpl; right; lia.
+   - auto.
+  Defined.
+
+End Merge.
+
+Goal forall l,  merge nat Nat.leb (nil, l) = l.
+  intro; now rewrite merge_equation.
+Qed.
+\end{Coqsrc}
+
+\subsection{Yet another  proof of impossibility}
+\label{omega2-case}
+
+In Sect.~\vref{omega-case}, we proved that there exists no variant towards \texttt{nat}
+(\emph{i.e.} the ordinal $\omega$) for proving the termination of all hydra battles.
+We  prove now that even the ordinal $\omega^2$ is also insufficient for this purpose.
+
+The proof we are going to develop has exactly the same structure as in Section~\ref{omega-case}.
+ Nevertheless, the proof of technical  lemmas is a little more complex, due to 
+ the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
+Consider for instance that there exists an infinite number of ordinals  between
+$\omega$ and $\omega\times 2$.
+
+The detailed  proof script is in the file \url{../src/html/hydras.Hydra.Omega2_Small.html}.
+
+\subsubsection{Preliminaries}
+Let us assume there is a variant from \texttt{Hydra} into $\omega^2$  for proving the   termination of all hydra battles.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
+
+\begin{Coqsrc}
+  Section Impossibility_Proof.
+
+ Variable m : Hydra -> ON_Omega2.t.
+
+ Context (Hvar : Hvariant lt_wf free m).
+\end{Coqsrc}
+
+Let us follow the same pattern as in Sect.~\ref{omega-case}.
+First, we define an injection from type \texttt{t} into \texttt{Hydra}, by
+ associating to  each ordinal $\omega\times i+ j = (i,j)$ the hydra with $i$ branches of length $2$ and
+$j$ branches of length $1$.
+
+%% revenir ici
+
+\vspace{4pt}
+\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
+
+\begin{Coqsrc}
+Let iota (p: ON_Omega2.t) := 
+    node (hcons_mult (hyd1 head) (fst p)
+                     (hcons_mult head (snd p) hnil)).
+  \end{Coqsrc}
+
+For instance, Figure~\vref{fig:essai2} shows the hydra associated to the ordinal 
+$(3,5)$, a.k.a. $\omega\times 3 + 5$.
+
+\begin{figure}[htb]
+\centering
+\begin{tikzpicture}[very thick, scale=0.4]
+\node (foot) at (6,0) {$\bullet$};
+\node (N1) at (1,3) {$\bullet$};
+\node (N2) at (3,3) {$\bullet$};
+\node (N3) at (5,3) {$\bullet$};
+\node (N4) at (8,3) {$\Smiley[2][green]$};
+\node (N5) at (11,3) {$\Smiley[2][green]$};
+\node (N6) at (14,3) {$\Smiley[2][green]$};
+\node (N7) at (17,3){$\Smiley[2][green]$};
+\node (N8) at (20,3){$\Smiley[2][green]$};
+\node  (N9) at (0,5) {$\Smiley[2][green]$};
+\node (N10) at (2,5) {$\Smiley[2][green]$};
+\node (N11) at (4,5) {$\Smiley[2][green]$};
+\draw (foot) to [bend left=10] (N1);
+\draw (foot) -- (N2);
+\draw (foot) -- (N3);
+\draw (foot) -- (N4);
+\draw (foot) -- (N5);
+\draw (foot) -- (N6);
+\draw (foot) to [bend right=10] (N7);
+\draw (foot) to [bend right=15] (N8);
+\draw (N1) to [bend left=10] (N9);
+\draw (N2) -- (N10);
+\draw (N3) -- (N11);
+\end{tikzpicture}
+\caption{\label{fig:essai2}
+The hydra $\iota(\omega\times 3+5)$}
+\end{figure}
+
+Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).
+
+\begin{figure}[htb]
+\centering
+\begin{tikzpicture}[very thick, scale=0.5]
+\node (foot) at (2,0) {$\bullet$};
+\node (N1) at (2,2) {$\bullet$};
+\node (N2) at (3,4) {$\Smiley[2][green]$};
+\node (N3) at (1,4) {$\Smiley[2][green]$};
+\draw (foot) -- (N1);
+\draw (N1) to [bend right =15] (N2);
+\draw (N1) to  [bend left=15](N3);
+\end{tikzpicture}
+\caption{\label{fig:h-omega2-small}}
+ The hydra \texttt{big\_h}.
+\end{figure}
+
+\begin{Coqsrc}
+   Let big_h := hyd1 (hyd2 head head).  
+\end{Coqsrc}
+
+ In a second step, we build a ``smaller'' hydra.
+
+\begin{Coqsrc}
+   Let small_h := iota (m big_h).
+\end{Coqsrc}
+
+Like in Sect.~\ref{omega-case}, we prove the double inequality \texttt{m big\_h o<= m small\_h o< m big\_h}, which is impossible.
+
+\subsubsection{Proof of the inequality \texttt{m small\_h o< m big\_h}}
+
+In order to prove the inequality  \texttt{m\_lt: m small\_h o< m big\_h}, it suffices to
+build a battle transforming \texttt{big\_h} into \texttt{small\_h}.
+
+First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h} as $(i,j)$.
+If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
+If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the following result.
+
+\begin{Coqsrc}
+Lemma big_to_small: big_h -+-> small_h.
+\end{Coqsrc}
+
+Since $m$ is a variant, we infer the following inequality:
+
+\begin{Coqsrc}
+Corollary m_lt : m small_h o< m big_h.
+\end{Coqsrc}
+
+\subsubsection{Proof of the inequality \texttt{m big\_h o<= m small\_h} }
+
+The proof of the inequality \texttt{m big\_h o<= m small\_h} is quite more complex than in Sect~\ref{omega-case}.
+If we consider some ordinal $\alpha=(i,j)$, where $i>0$, there exists an infinite number of
+ordinals stricly less than $\alpha$, and there exists an infinite number of battles that start from
+$\iota(\alpha)$.
+In effect, at any configuration $\iota(k,0)$, where $k>0$, the hydra can freely chose any replication number.
+Intuitively, the measure of such a hydra must be large enough for taking into account
+all the possible battles issued from that hydra.
+Let us give more technical details.
+
+\begin{itemize}
+\item The proof of the lemma \texttt{m\_ge : m big\_h  o<= m small\_h} uses well-founded induction on \texttt{big\_h}.
+
+\item For any pair $p$, we have to distinguish between three cases, according to the value of $p$'s components.
+  \begin{itemize}
+  \item $p=(0,0)$
+  \item $p=(i,0)$, where $i>0$\,: $p$ corresponds to a limit ordinal
+  \item $p=(i,j)$, where $j>0$\,: $p$ is the successor of $(i,j-1)$.
+  \end{itemize}
+\end{itemize}
+
+Let us define the notion of elementary ``step'' of decreasing sequences in
+\texttt{t}
+
+\begin{Coqsrc}
+Inductive step : t -> t -> Prop :=
+| succ_step : forall i j,  step (i, S j) (i, j)
+| limit_step : forall i j, step (S i, 0) (i, j).
+\end{Coqsrc}
+
+The following lemma establishes a correspondance between the relation
+\texttt{step} and hydra battles.
+
+\begin{Coqsrc}
+Lemma step_to_battle : forall p q, step p q -> iota p -+-> iota q.
+\end{Coqsrc}
+
+\index{Maths!Transfinite induction}
+
+Thus, starting from any inequality $q < p$ on type \texttt{t}, we can build 
+by transfinite induction over \texttt{p} a battle 
+that transforms the hydra $\iota(p)$ into $\iota(q)$.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}
+
+\begin{Coqsrc}
+Lemma m_ge : forall p : t,   p o<= m (iota p).
+Proof.
+  unfold small_h; pattern (m big_h) .   
+   apply  well_founded_induction with (R := lt) (1:= lt_wf).
+  intro p ; pattern p;
+    apply  well_founded_induction with 
+               (R := lt2) (1:= wf_lexico lt_wf lt_wf);
+     intros (i,j) IHij. 
+\end{Coqsrc}
+
+\begin{Coqanswer}
+  i, j : nat
+  IHij : forall y : t, y o< (i, j) -> y o<= m (iota y)
+  ============================
+  (i, j) o<= m (iota (i, j)) 
+\end{Coqanswer}
+
+Then we have  three cases to consider, according to the values of $i$ and $j$.
+\begin{itemize}
+\item If $p=(0,0)$ then obviously, $\iota(p)\geq p = (0,0)$
+\item If  $p=(i+1,0)$ for some $i\in\mathbb{N}$, we
+ remark  that $p$ is strictly greater than any pair $ (i, j)$, where $j$ 
+is any natural number.
+
+Applying the battle rules, for any $j$, we have $\iota(i+1,j)  {\round} \iota(i, j) $, thus $m(\iota(p)) > m(\iota(i,j)$ since  $m$ is assumed to be a variant.
+
+Applying the induction hypothesis, we get the inequality
+ $ m(\iota(i,j)) \geq (i,j)$ for any $j$.
+
+Thus, $m(\iota(p)) > (i,j)$ for any $j$.
+Applying the lemma \texttt{limit\_is\_lub}, we get  the inequality
+$\iota(i+1,0)\geq (i+1,0)$
+
+\item If $p=(i,j+1)$ with $j\in\mathbb{N}$, we have  $\iota(p)  {\round} \iota(i, j) $,
+hence $m(\iota(p))> m(\iota(i,j)) \geq (i,j)$, thus $m(\iota(p))\geq (i,j+1)=p$
+
+\end{itemize}
+
+\begin{Coqsrc}
+  (* ... *)
+Qed.
+\end{Coqsrc}
+
+\subsubsection{End of the proof}
+
+Since $<$ is a strict order (irreflexive  and transitive) on \texttt{nat*nat}, we can conclude that there is no
+variant for termination on the lexicographic square of $(\mathbb{N},<)$.
+
+\begin{enumerate}
+\item From \texttt{m\_lt}, we infer the strict inequality 
+\texttt{m small\_h o< m big\_h}
+\item from \texttt{m\_ge}, we get \texttt{m big\_h o<= m (iota (m big\_h)) = m small\_h} 
+\end{enumerate}
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
+
+\begin{Coqsrc}
+Theorem Impossible : False.
+Proof.
+  destruct (StrictOrder_Irreflexive  (m big_h)).
+  apply le2_lt2_trans with (m small_h).
+  -  unfold small_h;  apply m_ge.
+  -  apply m_lt. 
+Qed.
+
+End Impossibility_Proof.
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+Prove that there exists no variant $m$ from \texttt{Hydra} into $\omega^2$ for proving
+    the  termination of all \emph{standard} battles.
+\end{exercise}
+
+\begin{remark}
+In Chapter~\ref{ks-chapter}, we  prove a generalization of the impossibility lemmas of
+Sect.~\ref{omega-case} and this section, with the same proof structure, but with much more 
+complex technical details.
+ \end{remark}
+
+\index{Exercises}
+\begin{exercise}
+
+\label{sec:orgheadline63}
+Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.
+
+\begin{itemize}
+\item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
+  lexicographic ordering
+\item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.
+
+For instance, the following inequality holds:
+\[\langle 4,3,3,3,3,3,3,2,2,2 \rangle\,<\,\langle 4,4,2 \rangle\]
+\end{itemize}
+
+\end{exercise}
+
+\section{Comparing Ordinal Notations}
+
+It is sometimes useful to compare two ordinal notations with respect to expressive power
+(the segment of ordinals  they represent).
+
+The following class specifies a strict inclusion of segments.
+The notation \texttt{OA} describes a segment $[0,\alpha($, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented in \texttt{OA}).
+We require also  that the comparison functions of the two notation systems are compatible.
+
+If \texttt{OB} is presumed to be correct, the we can consider that \texttt{OA} ``inherits'' its correcteness from the bigger notation system \texttt{OB}.
+
+\index{Coq!Type classes}
+\label{types:SubON}
+
+\begin{Coqsrc}
+Class  SubON 
+       `(OA : @ON A ltA  compareA)
+       `(OB : @ON B ltB  compareB)
+       (alpha :  B)
+       (iota : A -> B):=
+  {
+  subon_compare: forall x y : A,  compareB (iota x) (iota y) =
+                                 compareA x y;
+  subon_incl : forall x, ltB (iota x) alpha;
+  subon_onto : forall y, ltB y alpha  -> exists x:A, iota x = y}.
+\end{Coqsrc}
+
+For instance, we prove that \texttt{Omega} is a sub-notation of
+\texttt{Omega\_plus\_Omega} (with $\omega$ as the first ``new'' ordinal, and \texttt{fin} as the injection).
+
+\begin{Coqsrc}
+Instance Incl : SubON Omega Omega_plus_Omega omega fin.
+\end{Coqsrc}
+
+\index{Projects}
+\begin{project}
+Prove that \texttt{Omega\_plus\_Omega} cannot be a sub-notation of \texttt{Omega}.
+\end{project}
+
+\index{Projects}
+\begin{project}
+Adapt the definition of \texttt{Hvariant} (Sect.~\ref{sect:hvariant-def}) in order to
+have an ordinal notation as argument. Prove that if $O_A$ is a sub-notation of $O_B$, then any variant defined on  $O_A$ can be automatically transformed into 
+a variant on $O_B$.
+\end{project}
+
+\section{Comparing an ordinal notation with Schütte's model}
+
+Finally, it may be interesting to compare an ordinal notation with the more theoretical model from Schütte (well, at least with our formalization of that model).
+This would be a relative proof of correctenss of our ordinal  notations.
+
+The following class specifies that a notation \texttt{OA} describes a segment $[0,\alpha)$,
+where $\alpha$ is a countable ordinal \emph{à la}  Schütte.
+
+\index{Coq!Type classes}
+\label{types:ON-for}
+
+\begin{Coqsrc}
+Class ON_correct `(alpha : Schutte_basics.Ord)
+     `(OA : @ON A ltA  compareA)
+      (iota : A -> Schutte_basics.Ord) :=
+  { ON_correct_inj : forall a, Schutte_basics.lt (iota a) alpha;
+    ON_correct_onto : forall beta, Schutte_basics.lt beta alpha ->
+                                exists b, iota b = beta;
+    On_compare_spec : forall a b:A,
+        match compareA a b with
+          Datatypes.Lt => Schutte_basics.lt (iota a) (iota b)
+        | Datatypes.Eq => iota a = iota b
+        | Datatypes.Gt => Schutte_basics.lt (iota b) (iota a)
+        end}.
+\end{Coqsrc}
+
+For instance, the following theorem tells that \texttt{Epsilon0}, our notation system for the segment $[0,\epsilon0)$ is correct with respect to the countable ordinal $\epsilon_0$
+(see chapter~\ref{chap:schutte} for more details).
+
+\begin{Coqsrc}
+Instance Epsilon0_correct :
+  ON_correct epsilon0 Epsilon0  (fun alpha => inject (cnf alpha)).
+\end{Coqsrc}
+
+\index{Projects}
+
+\begin{project}
+  When you have read Chapter~\ref{chap:schutte}, prove that the sum of two ordinal notations \texttt{ON\_plus} implements the addition of ordinals.
+\end{project}
+
+\section{A notation for finite ordinals}
+
+Let $n$ be some natural number. The segment associated with $n$ is the interval 
+$[0,n)\,=\,\{0,1,\dots,n-1\}$. 
+One may represent the ordinal $n$ by a sigma type.
+
+\vspace{4pt}
+\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
+
+\label{def: Finite-ord-type}
+\begin{Coqsrc}
+Coercion is_true: bool >-> Sortclass.
+
+Definition t (n:nat) := {i:nat | Nat.ltb i n}.
+\end{Coqsrc}
+
+The order on type (\texttt{t $n$}) is defined through the projection on \texttt{nat}.
+
+\begin{Coqsrc}
+Definition lt {n:nat} : relation (t n) :=
+  fun alpha beta => Nat.ltb ( proj1_sig alpha) (proj1_sig beta).
+\end{Coqsrc}
+
+For instance, let us build two elements of the segment $[0, 7)$, \emph{i.e.} two
+inhabitants of   type (\texttt{t 7}), and prove a simple  inequality (see Fig.~\ref{fig:O7}).
+
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}[very thick, scale=0.6]
+
+\node (N0) at (0,0) {$\bullet$};
+\node (i0) at (0,1) {$0$};
+\node (N1) at (2,0) {$\bullet$};
+\node (i1) at (2,1) {$1$};
+\node (N2) at (4,0) {$\bullet$};
+\node (i2) at (4,1) {$2$};
+\node (N3) at (6,0) {$\bullet$};
+\node (i3) at (6,1) {$3$};
+\node (N4) at (8,0) {$\bullet$};
+\node (i4) at (8,1) {$4$};
+\node (N5) at (10,0) {$\bullet$};
+\node (i5) at (10,1) {$5$};
+\node (N6) at (12,0) {$\bullet$};
+\node (i6) at (12,1) {$6$};
+\node(alpha1) at (4,-1) {$\alpha_1$};
+\node(alpha2) at (10,-1) {$\beta_1$};
+\end{tikzpicture}
+
+\caption{The segment $\mathbb{O}_7$\label{fig:O7}}
+\end{figure}
+
+\index{Coq!Commands!Program}
+
+\begin{Coqsrc}
+Program Example alpha1 : t 7 := 2.
+
+Program Example beta1 : t 7 := 5.
+
+Example i1 : lt  alpha1 beta1.
+Proof. now compute. Qed.
+\end{Coqsrc}
+
+Note that the type (\texttt{t 0}) is empty, and that \texttt{Program} generates an obligation
+for verifying that the constraint $i<n$ is fullfilled when one tries to build the $i$-th ordinal inside the type (\texttt{t $n$}).
+
+\begin{Coqsrc}
+Lemma t0_empty (alpha: t 0): False.
+Proof.
+  destruct alpha.
+  destruct x; cbn in i; discriminate.
+Qed.
+
+Program Definition bad : t 10 := 10.
+Next Obligation.
+  compute.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+1 subgoal (ID 162)
+
+  ============================
+  false = true
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Abort.
+\end{Coqsrc}
+
+Note also that attempting to compare a term  of type (\texttt{t $n$}) with a term of
+type (\texttt{t $p$})  leads to an error if $n$ and $p$ are not convertible.
+
+\begin{Coqsrc}
+
+Program Example gamma1 : t 8 := 7.
+
+Fail Goal lt alpha1 gamma1.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+ The command has indeed failed with message:
+The term "gamma1" has type "t 8" while it is expected to have type "t 7".
+\end{Coqanswer}
+
+In order to build an instance of \texttt{OrdinalNotation}, we define a comparison function, by delegation to standard library's  \texttt{Nat.compare}, and prove its correction.
+
+\begin{Coqsrc}
+Definition compare {n:nat} (alpha beta : t n) :=
+  Nat.compare (proj1_sig alpha) (proj1_sig beta).
+
+Lemma compare_correct {n} (alpha beta : t n) :
+  CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
+              (compare alpha beta).
+\end{Coqsrc}
+
+\begin{remark}
+ The proof of \texttt{compare\_correct} uses a well-known pattern of \coq{}.
+Let us consider  the following subgoal.
+
+\begin{Coqanswer}
+ 1 subgoal (ID 110)
+
+  n, x0 : nat
+  i, i0 : x0 <? S n
+  ============================
+  exist (fun i1 : nat => i1 <=? n) x0 i =
+  exist (fun i1 : nat => i1 <=? n) x0 i0
+\end{Coqanswer}
+
+Applying the tactic \texttt{f\_equal} generates a simpler subgoal.
+
+\begin{Coqanswer}
+1 subgoal (ID 112)
+
+  n, x0 : nat
+  i, i0 : x0 <? S n
+  ============================
+  i = i0
+\end{Coqanswer}
+
+We have now to prove that there exists at most one  proof of (\texttt{Nat.ltb x0 (S n)}). This is not obvious, but  a consequence of the following lemma of library 
+\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Eqdep_dec.html}{Coq.Logic.Eqdep\_dec}.
+
+\index{Coq!Techniques!Unicity of equality proofs}
+\label{sect:eq-proof-unicity}
+
+\begin{Coqanswer}
+eq_proofs_unicity_on :
+forall (A : Type) (x : A),
+(forall y : A, x = y \/ x <> y) -> 
+forall (y : A) (p1 p2 : x = y), p1 = p2
+\end{Coqanswer}
+
+Thus unicity of proofs of \texttt{Nat.ltb x0 (S n)}  comes from the decidability of
+equality on type \texttt{bool}.
+This is why we used the boolean function \texttt{Nat.ltb} instead of the inductive predicate \texttt{Nat.lt} in the definition of type \texttt{t $n$} (see page~\pageref{def: Finite-ord-type}).
+For more information about this pattern, please look at the numerous mailing lists and 
+FAQs on \coq{}).
+
+\end{remark}
+
+Applying lemmas of the libraries \texttt{Coq.Wellfounded.Inverse\_Image}, \linebreak
+ \texttt{Coq.Wellfounded.Inclusion}, and \texttt{Coq.Arith.Wf\_nat}, we prove that our
+relation \texttt{lt} is well founded.
+
+\begin{Coqsrc}
+Lemma lt_wf (n:nat) : well_founded (@lt n).
+\end{Coqsrc}
+
+Now we can build our instance of \texttt{OrdinalNotation}.
+
+\begin{Coqsrc}
+Global Instance sto n : StrictOrder (@lt n).
+
+Global Instance FinOrd (n:nat) : OrdinalNotation (sto n) compare.
+Proof.
+  split.
+  - apply compare_correct.
+  - apply lt_wf.
+Qed.
+\end{Coqsrc}
+
+\begin{remark}
+It is important to keep in mind  that the integer $n$ is not an ``element'' of \texttt{FinOrd $n$}.
+In set-theoretic presentations of ordinals, the set associated with the ordinal $n$ is $\{0,1,\dots,n-1\}$.
+In our formalization, the interpretation of an ordinal as a set is realized by the following definition
+(in ~\href{../src/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}).
+
+\begin{Coqsrc}
+Definition bigO `{nA : @OrdinalNotation A ltA stoA compareA}
+           (a: A) : Ensemble A :=
+  fun x: A => ltA x a.
+\end{Coqsrc}
+\end{remark}
+
+Finally, it is interesting to show that, if $i<j$, then the segment $[0,i)$ is a ``sub-segment'' of
+$[0,j)$. Since the terms  ($t\;i$) and ($t\;j$) are not convertible, we consider a ``cast'' 
+function $\iota$ from ($t\;i$) into ($t\;j$), and prove that this function is  a monotonous bijection  from ($t\;i$) to
+the segment $[0,i)$ of ($t\;j$).
+
+ \begin{figure}[h]
+   \centering
+   \begin{tikzpicture}[very thick, scale=0.6]
+\begin{scope}[color=blue]
+\node (A) at (0,0) {$A$};
+\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
+\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
+\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
+\node (Adots) at (6,0) {$\ldots$};
+\end{scope}
+\begin{scope}[color=red]
+\node (B) at (0,2) {$B$};
+\node(B0) at (2,2)[label=above:$0$]{$\bullet$};
+\node(B1) at (3,2)[label=above:$1$]{$\bullet$};
+\node(B2) at (4,2)[label=above:$2$]{$\bullet$};
+\node (Bdots) at (6,2) {$\ldots$};
+\node (b) at (8,2) [label=above:$b$]{$\bullet$};
+\node (bsucc) at (9,2) [label=above:$b+1$]{$\bullet$};
+\node (Bdots2) at (10,2) {$\ldots$};
+\end{scope}
+\begin{scope}[color=red!50!blue]
+\draw [->,thin] (A0) -- node [auto] {$\iota$} (B0);
+\draw [->,thin] (A1) -- node [auto] {$\iota$} (B1);
+\draw [->,thin] (A2) -- node [auto] {$\iota$} (B2);
+\draw [->,thin] (Adots) -- node [auto] {$\iota$} (Bdots);
+\end{scope}
+\end{tikzpicture}
+   \caption{\textcolor{blue}{$A$} is a sub-segment  of \textcolor{red}{$B$}}
+   \label{fig:subsegment}
+ \end{figure}
+
+\index{Coq!Commands!Program}
+
+We are now able to build an instance of \texttt{SubON}.
+
+\begin{Coqsrc}
+Section Inclusion_ij.
+
+  Variables i j : nat.
+  Hypothesis Hij : (i < j)%nat.
+
+   Remark Ltb_ij : Nat.ltb i j.
+   Program Definition iota_ij  (alpha: t i) : t j :=  alpha.
+
+   Let b : t j := exist _ i Ltb_ij.
+
+   Global Instance F_incl_ij  : SubON  (FinOrd i) (FinOrd j) b iota_ij.
+  (* ... *)
+
+  End Inclusion_ij.
+
+\end{Coqsrc}
+
+\begin{remark}
+ There is no interesting arihmetic on finite ordinals, since functions like successor, addition, etc.,  cannot be represented in \coq{} as \emph{total} functions.
+\end{remark}
+
+\paragraph{Related work}
+Finite ordinals are also formalized in MathComp~\cite{SSR}.  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.
+
+\section{Isomorphism of ordinal notations}
+
+In some cases we want to show that two notation systems describe the same segment (for instance $[0,3+\omega)$ and $[0,\omega($).
+For this purpose, one may prove that the two notation systems are order-isomorphic.
+
+\index{Coq!Type classes}
+\label{types:ON-iso} 
+\begin{Coqsrc}
+Class  ON_Iso 
+       `(OA : @ON A ltA compareA)
+       `(OB : @ON B ltB  compareB)
+       (f : A -> B)
+       (g : B -> A):=
+  {
+  iso_compare: forall x y : A, 
+      compareB (f x) (f y) = compareA x y;
+  iso_inv1 : forall a, g (f a)= a;
+  iso_inv2 : forall b, f (g b) = b}.
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+Let $i$ be some natural number. Prove that the notation systems 
+\texttt{Omega} and \texttt{ON\_plus (OrdFin $i$) Omega} are isomorphic.
+
+{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$ we will prove in larger notation systems, as well as in Schütte's model}
+
+This exercise is partially solved for $i=3$ (in ~\href{../src/html/hydras.OrdinalNotations.Example_3PlusOmega.html}{OrdinalNotations.Example\_3PlusOmega}).
+
+\end{exercise}
+
+\index{Projects}
+\label{exo:ON-mult}
+\begin{project}
+Define in \coq{} the product of two ordinal notations $N_A$ and $N_B$.
+If $A$ [resp. $B$] is the underlying type of $N_A$ [resp. $N_B$], the
+product \texttt{ON\_mult $N_A$ $N_B$} is implemented over the cartesian product $B\times A$ (with the lexicographic ordering).
+
+For instance, the
+elements of the product \texttt{ON\_mult Omega (FinOrd 3)} are ordered as follows.
+\[(0,0),(0,1),(0,2),(0,3),(0,4),\dots,{\color{red}(1,0),} (1,1),(1,2),\dots, {\color{red}(2,0)},(2,1),(2,2),\dots\]
+
+Note that the elements of  \texttt{ON\_mult (FinOrd 3) Omega} are differently ordered (without limit ordinals):
+\[(0,0),(1,0),(2,0),(0,1),(1,1),(2,1),(0,2),(1,2),(2,2),(0,3),\dots\]
+(no limit ordinals).
+
+Prove that \texttt{ON\_mult (FinOrd $i$) Omega} is isomorphic to
+\texttt{Omega}  whilst
+\texttt{Omega}  is a sub-notation of \texttt{ON\_mult Omega (FinOrd $i$)},
+for any strictly positive $i$.
+\end{project}
+
+\index{Projects}
+\begin{project}
+\label{project:succ-limit-dec}
+Add to the class \texttt{ON} the requirement that for any $\alpha$ it is decidable whether $\alpha$ is $0$, a successor or a limit ordinal.
+\end{project}
+
+\index{Projects}
+\begin{project}
+\label{project:on-setoid}
+Reconsider the  class \texttt{ON}, with an equivalence instead of Leibniz equality.
+\end{project}
+
+%%%% ICI ICI
+
+\section{Other ordinal notations}
+
+\index{Projects}
+
+\begin{project}
+The directory \texttt{src/OmegaOmega} contains an ad-hoc formalization of $\omega^\omega$, contributed by Pascal Manoury.
+Every ordinal $\alpha$ is represented by a list $l$ whose elements are the coefficients of $\omega$ in  the Cantor normal form of $\alpha$ (in reverse order).
+For instance, the ordinal
+$\omega^{8}\times 5 + \omega^{6}\times 8 + \omega^2\times 10 + \omega + 7$ is represented by the list \texttt{[5;0;8;0;0.0;10,1,7]}.
+
+ Develop this representation and compare it with the other ordinal notations.
+
+\end{project}
+
+\index{Projects}
+
+\begin{project}
+Let $N_A$ be a notation system for ordinals strictly less than $\alpha$, 
+with the strict order $(A,<_A)$. Please build the notation system
+\texttt{ON\_Expl $N_A$}, on the type of multisets of elements of $A$
+(or, if preferred, the type of non-increasing finite sequences on $A$,
+provided with the lexicographic ordering on lists).
+
+For instance, let us take $N_A=\texttt{Omega}$, and take $\alpha=\langle 4,4,2,1,0\rangle$,
+ $\beta=\langle 4,3,3,3,3,3,2\rangle$, and $\gamma=\langle 5\rangle$. Then $\beta<\alpha<\gamma$.
+
+In contrast the list $\langle5,6,3,3\rangle$ is not non-increasing (\emph{i.e.} sorted w.r.t. $\geq$), so it is not to be considered.
+
+Note that if the notation $N_A$ implements the ordinal 
+$\alpha$,  the new notation $\omega^{N_A}$ must implement the ordinal $\varphi_0(\alpha)$, a.k.a. $\omega^\alpha$ (see chapter~\ref{chap:schutte})
+
+\end{project}
+
+\begin{remark}
+ The set of ordinal terms in Cantor normal form (see Chap.~\ref{chap:T1}) and 
+in Veblen normal form (see 
+\href{../src/html/hydras.Gamma0.Gamma0.html}{Gamma0.Gamma0}) are shown to be ordinal notation systems, but there is a lot of work to be done in order to unify ad-hoc  definitions and proofs which were written before the definition of the \texttt{ON} type class.
+\end{remark}
+
+%------------------------------------------------------------------------
+
+\chapter[A proof of termination, using epsilon0]{A proof of termination, using ordinals below \texorpdfstring{$\epsilon_0$}{epsilon0}}
+
+\label{cnf-math-def}
+\label{chap:proof-term}
+\label{chap:T1}
+
+In this chapter, we adapt to \coq{} the well-known~\cite{KP82}  proof that Hercules eventually wins every battle, whichever the strategy  of each player.
+In other words, we present  a formal and self contained proof of termination  of all [free] hydra battles.
+First, we take from Manolios and Vroon~\cite{Manolios2005} a representation of the ordinal $\epsilon_0$ as terms in Cantor normal form.
+Then, we define a variant for hydra battles as a measure that maps any hydra to some ordinal strictly less than $\epsilon_0$.
+
+\section{The ordinal \texorpdfstring{\(\epsilon_0\)}{epsilon0}}
+\label{sec:epsilon0-intro}
+
+\subsection{Cantor normal form}
+
+The ordinal \(\epsilon_0\) is the least ordinal number that satisfies 
+the equation \(\alpha = \omega^\alpha\), where \(\omega\) is 
+the least infinite ordinal. Thus, we can consider \(\epsilon_0\) as an
+\emph{infinite} \(\omega\)-tower.
+
+\index{Maths!Cantor normal form}
+
+Any ordinal strictly less that \(\epsilon_0\) 
+can be represented by a unique  \emph{Cantor normal form}, 
+that is, an expression  which is either  the ordinal \(0\) or 
+a sum  \(\omega^{\alpha_1} \times n_1 + \omega^{\alpha_2} \times n_2 + 
+  \dots + \omega^{\alpha_p} \times n_p\) where all the \(\alpha_i\) 
+are ordinals in Cantor  normal form, \(\alpha_1 > \alpha_2 > \alpha_p\), 
+and all the \(n_i\) are positive integers.
+
+An example of Cantor normal form is displayed in Fig \ref{fig:cnf-example}:
+Note that  any ordinal of
+the form \(\omega^0 \times i + 0\) is just written \(i\).
+
+\begin{figure}[htb]
+\centering
+\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
+\node[color=blue]{$\omega^{(\omega^\omega\,+\, \omega^2 \times 8 \,+\, \omega)}+ \omega^\omega + \omega^4+ 6$};
+\end{tikzpicture}
+\caption{\label{fig:cnf-example}
+An ordinal in Cantor normal form}
+\end{figure}
+
+In the rest of this section, we define an inductive type for representing in \texttt{Coq}
+all the ordinals strictly  less than  \(\epsilon_0\), then extend some arithmetic operations
+to this type, and finally prove that our representation fits well with 
+the expected mathematical properties: the order we define is a well order, 
+and the decomposition into Cantor normal form  is consistent 
+with the implementation of the arithmetic operations of exponentiation of base \(\omega\) 
+and addition.
+
+\paragraph*{Remark}
+\label{sec:orgheadline65}
+Unless explicitly mentionned, the term ``ordinal" will be used instead of
+``ordinal strictly less than \(\epsilon_0\)" (except in Chapter~\ref{chap:schutte} where it stands for ``countable ordinal'').
+
+\subsection{A data type for  ordinals in Cantor normal form}
+\label{sec:orgheadline72}
+\label{sec:T1-inductive-def}
+
+% Our user contribution~\cite{CantorContrib} represents 
+% the set of ordinals strictly less than $\epsilon_0$ in Cantor normal form as in~\cite{Manolios2005}, and also the set
+% of ordinals strictly  less than $\Gamma_0$ in Veblen normal form.
+
+    Let us define an inductive type whose 
+constructors are respectively associated
+with the ways to build Cantor normal forms:
+
+\begin{itemize}
+\item the ordinal \(0\)
+\item the construction \((\alpha,\, n,\,\beta)  \mapsto \omega^\alpha \times (n + 1)+ \beta \quad (n\in\mathbb{N})\)
+\end{itemize}
+
+\vspace{4pt}
+\noindent\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#T1}{Epsilon0.T1}}
+
+\label{types:T1}
+\index{Constants!zero:T1}
+
+\begin{Coqsrc}
+Inductive T1 : Set  :=
+| zero : T1
+| ocons : T1 -> nat -> T1 -> T1.
+\end{Coqsrc}
+
+\subsubsection*{Remark}
+\label{sec:orgheadline66}
+
+The name \texttt{T1} we gave to this data-type  is proper to this development and refers
+to a hierarchy of ordinal notations.
+For instance, in~\cite{CantorContrib}, the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{schutte}).
+
+\label{types:T2}
+\emph{Please look also at the library
+\href{../src/html/hydras.Gamma0.T2.html}{Gamma0.T2.html}}.
+
+\begin{Coqsrc}
+Inductive T2 : Set :=
+  zero : T2
+| gcons : T2 -> T2 -> nat -> T2 -> T2.
+\end{Coqsrc}
+
+\subsubsection{Example}
+
+\label{alpha0-def}
+For instance, the ordinal  $\omega^\omega+\omega^3\times 5+2$ is represented by the following term:
+
+\begin{Coqsrc}
+Example alpha_0 : T1 :=
+  ocons (ocons (ocons zero 0 zero)
+               0
+               zero)
+        0
+       (ocons (ocons zero 2 zero)
+              4
+              (ocons zero 1 zero)).
+\end{Coqsrc}
+
+\begin{figure}[htb]
+\centering
+\begin{tikzpicture}[very thick, scale=0.5, level 1/.style={sibling distance=6cm},
+level 2/.style={sibling distance=35mm},  
+level 3/.style={sibling distance=17mm}]
+\node  {ocons}
+  child {  node {ocons}
+            child { node {ocons} child {node {zero}} child {node{0}} child{node{zero}}}
+         child {node {0}}
+         child {node {zero}}}
+    child {node {0}}
+   child {node {ocons} 
+ child { node {ocons} child {node {zero}} child {node{2}} child{node{zero}}}
+  child {node {4}}
+         child {node {ocons} child {node {zero}} child {node{1}} child{node{zero}}}};
+
+\end{tikzpicture}
+
+\caption{The tree-like representation of the ordinal $\omega^\omega+\omega^3\times 5 +2$\label{fig:cnf-tree}}
+
+\end{figure}
+
+\paragraph{Remark}
+For simplicity's sake, we chosed to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the contruction (\texttt{ocons $\alpha$ $n$ $\beta$} is intented to represent the
+ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
+In a future version, we should replace  the type \texttt{nat} with \texttt{positive} in \texttt{T1}'s 
+definition. But this replacement would take a lot of time \dots{}
+
+\subsection{Abbreviations}
+
+Some abbreviations may help to write more consisely complex ordinal terms.
+
+\subsubsection{Finite ordinals}
+\label{sec:orgheadline67}
+
+For representing finite ordinals, \emph{i.e.} natural numbers, we first introduce a notation for terms of the form $n+1$, then define a coercion from type \texttt{nat} into \texttt{T1}.
+\label{sect:notation-FS}
+
+\begin{Coqsrc}
+Notation "'FS' n" :=
+     (ocons zero n zero) (at level 10) : t1_scope.
+\end{Coqsrc}
+
+\label{sect:notation-F}
+
+\begin{Coqsrc}
+Definition fin (n:nat) : T1 := 
+    match n with 0 => zero | S p => FS p end.
+
+Coercion fin  : nat >-> T1.
+
+Notation "'F' n" := (fin n) (at level 10) : t1_scope.
+
+Example ten : T1 := 10.   
+\end{Coqsrc}
+
+% \index{Coq!Techniques!Coercions}
+% \index{Functions!Coercions@Coercions (from nat to ordinal types)}
+% \begin{remark}
+% Please refer to the remark~\pageref{warning:coercions} about the use of coercions.
+% % The use of coercions like \texttt{fin} allow us to be close to the mathematical tradition where natural numbers are ordinals too.
+% % Nevertheless, it may happen that a goal like \texttt{3 < 5} could be 
+% % interpreted as \texttt{(lt (fin 3) (fin 5))},  depending on the current notation scope.  
+% % When this misinterpretation happens, tactics like \texttt{auto with arith}, \texttt{lia} do not work!
+% % Thus, it is useful to write \texttt{(3 < 5)\%nat}  an inequality between two natural numbers. 
+% \end{remark}
+
+\subsubsection{The ordinal \(\omega\)}
+\label{sec:orgheadline68}
+
+  Since \(\omega\)'s Cantor normal form is
+i.e. \(\omega^{\omega^0}\times 1+ 0\), we can define the following abbreviation:
+
+\label{sect:omega-notation2}
+\begin{Coqsrc}
+Notation omega := (ocons (ocons zero 0 zero) 0 zero): t1_scope.
+\end{Coqsrc}
+
+Note that \texttt{omega} is not an identifier, thus any tactic like \texttt{unfold omega} would fail.
+
+\subsubsection{The ordinal \(\omega^\alpha\), a.k.a. \(\varphi_0(\alpha)\)}
+\label{sec:orgheadline70}
+
+We provide also a notation for ordinals of the form $\omega^\alpha$.
+
+\label{sect:notation-phi0}
+
+\index{Notations!phi0@phi0 (exponential of base omega)}
+
+\begin{Coqsrc}
+Notation "'phi0' alpha" := (ocons alpha 0 zero) (at level 29) : t1_scope.
+\end{Coqsrc}
+
+\index{Maths!Additive principal ordinals}
+
+\begin{remark}
+\label{sec:orgheadline69}
+The name \(\varphi_0\)
+   comes from ordinal numbers theory. In~\cite{schutte}, Schütte defines 
+$\varphi_0$  as the ordering (\emph{i.e.} enumerating) function of the set  of \emph{additive principal ordinals} \emph{i.e.} strictly positive ordinals $\alpha$ that verify $\forall \beta<\alpha, \beta+\alpha=\alpha$.
+For Schütte,  $\omega^\alpha$ is just a notation for $\varphi_0(\alpha)$.
+See also Chapter~\ref{chap:schutte} of this document.
+\end{remark}
+
+\subsubsection{The hierarchy of \(\omega\)-towers:}
+\label{sec:orgheadline71}
+
+The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximed by the sequence of $\omega$-towers.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}
+
+\begin{Coqsrc}
+Fixpoint omega_tower (height:nat) : T1 := 
+ match height with 
+ | 0 =>  1 
+ | S h => phi0 (omega_tower h)
+ end.
+\end{Coqsrc}
+
+For instance, Figure~\ref{fig:tower7} represents  the ordinal returned by the
+ evaluation of the term \texttt{omega\_tower 7}.
+
+\begin{figure}[htb]
+\centering
+\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
+\node[color=blue]{$\omega^{{{\omega}^{{{\omega}}^{{{\omega}}^{{\omega^{{\omega}^{\omega}}}}}}}}$};
+\end{tikzpicture}
+\caption{\label{fig:tower7}
+The $\omega$-tower of height 7}
+\end{figure}
+
+\subsection{Comparison between ordinal terms}
+\label{sec:orgheadline73}
+
+% Our formalisation of Cantor Normal Form will take two steps:
+% 1 Definition of a strict order \texttt{o<} on the type \texttt{T1}, 
+% 2 Using \texttt{o<} for characterizing terms in normal form.
+
+In order to compare two terms of type \texttt{T1}, we define a recursive function \texttt{compare} that maps two ordinals $\alpha$ and $\beta$ to a value of type \texttt{comparison}.
+This type is defined in \coq's standard library
+\texttt{Init.Datatypes} and
+contains three constructors:  \texttt{Lt} (less than), \texttt{Eq} (equal), and
+\texttt{Gt} (greater than).
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#compare}{Epsilon0.T1}}
+
+\begin{Coqsrc}
+Fixpoint compare (alpha alpha':T1):comparison :=
+  match alpha, alpha' with
+    zero, zero => Eq
+  | zero, ocons a' n' b' => Lt
+  | _   , zero => Gt
+  | (ocons a n b),(ocons a' n' b') =>
+      (match compare a a' with 
+          | Lt => Lt
+          | Gt => Gt
+          | Eq => (match lt_eq_lt_dec n n'
+                   with
+                       inleft  (left _) => Lt
+                     | inright _ => Gt
+                     |   _ => compare b b'
+                   end)
+       end)
+  end.
+\end{Coqsrc}
+
+It is now easy to define the boolean predicate \texttt{lt\_b $\alpha$ $\beta$}: 
+`` $\alpha$ is strictly less than $\beta$ ''. By coercion to sort \texttt{Prop} we define also the predicate \texttt{lt}.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}
+
+\begin{Coqsrc}
+Definition lt_b alpha beta : bool :=
+  match compare alpha beta with
+      Lt => true
+    | _ => false
+  end.
+
+Definition lt alpha beta : Prop := lt_b alpha beta.
+\end{Coqsrc}
+
+\label{Predicates:lt-T1}
+Please note that this definition of \texttt{lt} makes it easy to write proofs by reflection, as shown by the following exampgles.
+
+\begin{Coqsrc}
+Example E1 : lt (ocons omega 56 zero) (tower 3).
+Proof. reflexivity. Qed.
+
+Example E2 : ~ lt (tower 3) (tower 3).
+Proof.  discriminate.  Qed.
+\end{Coqsrc}
+
+The following lemmas establish relations between \texttt{compare}, 
+the predicate \texttt{lt} and Leibniz equality \texttt{eq}.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#compare_refl}{Epsilon0.T1}}
+
+\begin{Coqsrc}
+Lemma compare_refl : forall alpha, compare alpha alpha =  Eq.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma compare_reflect : forall alpha beta,
+    match compare alpha beta with
+    |   Lt => lt alpha  beta
+    |   Eq => alpha = beta
+    |   Gt => lt beta  alpha
+    end.
+\end{Coqsrc}
+
+We prove also that the relation \texttt{lt} is a strict total order.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#lt_irrefl}{Epsilon0.T1}}
+
+\begin{Coqsrc}
+Theorem lt_irrefl  alpha :  ~ lt alpha  alpha.
+
+Theorem lt_trans (alpha beta gamma : T1) :
+  lt alpha  beta -> lt beta gamma -> lt alpha gamma.
+
+Definition lt_eq_lt_dec  :
+   forall alpha beta : T1, 
+          {lt alpha  beta} + {alpha = beta} + {lt beta alpha}.
+\end{Coqsrc}
+
+Note that the order \texttt{lt} is not reflected 
+in the structure (size and/or height) of the terms of \texttt{T1}. 
+For instance the ordinal of Fig \ref{fig:cnf-example} is strictly less
+than the structurally simpler \(\omega^{\omega^\omega}\times 2\).
+
+\subsubsection{A predicate for characterizing normal forms}
+\label{sec:orgheadline74}
+\label{sec:orgheadline75}
+We note that our data-type \texttt{T1} allows us to write expressions that
+are not properly in Cantor normal form as specified in Section \ref{sec:epsilon0-intro}.
+For instance, consider the following term of type  \texttt{T1}.
+
+\begin{Coqbad}
+Example bad_term  : T1 := ocons 1 1 (ocons omega 2 zero).
+\end{Coqbad}
+
+This term would have been written \(\omega^1\times 2 + \omega^\omega \times 3\) in the usual mathematical notation.
+We note that the exponents of $\omega$ are not in the right (strictly decreasing) order.
+
+With the help of the order \texttt{lt} on \texttt{T1}, we are now able to characterize
+the set of all well-formed ordinal terms:
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#nf_b}{Epsilon0.T1}}
+
+\label{Predicates:nf-T1}
+
+\begin{Coqsrc}
+Fixpoint nf_b (alpha : T1) : bool :=
+  match alpha with
+    | zero => true
+    | ocons a n zero => nf_b a
+    | ocons a n ((ocons a' n' b') as b) =>
+      (nf_b a && nf_b b && lt_b a' a)%bool
+  end.
+
+Definition nf alpha: Prop := nf_b alpha.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+ Compute nf_b alpha_0.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+   = true 
+     : bool
+\end{Coqanswer}
+
+\begin{Coqsrc}
+ Compute nf_b bad_term.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+   = false 
+     : bool
+\end{Coqanswer}
+
+\subsection{Making normality implicit}
+  We would like to get rid of terms of type \texttt{T1} which are not in Cantor normal form.
+A simple way to do this is to consider statements of the form 
+\texttt{forall alpha: T1, nf alpha -> $P$ alpha}, where $P$ is a predicate over type \texttt{T1}, like in the following lemma \footnote{Ordinal addition is formally defined a little later (page~\ref{sect:infix-plus-T1})}.
+
+\begin{Coqsrc}
+Lemma plus_is_zero alpha beta :
+  nf alpha -> nf beta ->
+  alpha + beta  = zero -> alpha = zero /\  beta = zero.
+\end{Coqsrc}
+
+But this style leads to clumsy statements, and many sub-goals in interactive proofs (although often solved with \texttt{auto} or \texttt{eauto}).
+
+One may encapsulate conditions of the form \texttt{(nf $\alpha$)} in
+the most used predicates. For instance, we introduce the restriction of \texttt{lt} to terms in normal form, and provide a handy notation for this restriction.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Prelude.Restriction.html}{Ordinals.Prelude.Restriction}}
+
+\begin{Coqsrc}
+Definition restrict {A:Type}(E: Ensemble A)(R: relation A) :=
+    fun a b => E a /\ R a b /\ E b.
+ \end{Coqsrc}
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#LT}{Epsilon0.T1}}
+
+\begin{Coqsrc}
+Definition LT := restrict nf lt.
+Infix "o<" := LT : t1_scope.
+
+Definition LE := restrict nf le.
+Notation "alpha o<= beta" := (LE alpha beta) : t1_scope.
+\end{Coqsrc}
+
+\begin{todo}
+Infix notation for inequalities on ordinals have changed to \texttt{o<} and \texttt{o<=}.  This document is not updated yet, but the \coq{} scripts are correct.
+\end{todo}
+
+\label{Predicates:LT-T1}
+
+For instance, in the following lemma, the condition that $\alpha$ is in normal form is included in the condition $\alpha< 1$.
+
+\begin{Coqsrc}
+Lemma LT_one : forall alpha, alpha o< one -> alpha = zero.
+\end{Coqsrc}
+
+\subsubsection{A sigma type for \texorpdfstring{$\epsilon_0$}{epsilon0}}
+
+The type \texttt{T1} is not a correct ordinal notation, since it contains terms that are not in Cantor normal form. Moreover,
+in Sections~\ref{sect:L-equations}, \ref{sect:hardy},
+and \ref{sect:wainer}, we needed to define total recursive functions on well-formed ordinal terms less  than $\epsilon_0$, using the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777}.
+ In order to define a type whose inhabitants represent just ordinals, we build a type gathering a term of type \texttt{T1} and a proof that this term is in normal form.
+
+\label{sect:E0-def}
+\label{types:E0}
+
+\emph{From Module~\href{../src/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
+
+\index{Coq!Type classes}
+
+\begin{Coqsrc}
+Class E0 : Type := t1_2o {cnf : T1; cnf_ok : nf cnf}.
+\end{Coqsrc}
+
+Many constructs : types, predicates, functions, notations, etc., on type \texttt{T1} are adapted to \texttt{E0}.
+
+First, we declare a notation scope for \texttt{E0}.
+
+\begin{Coqsrc}
+Declare Scope E0_scope.
+Delimit Scope E0_scope with e0.
+Open Scope E0_scope.
+\end{Coqsrc}
+
+Then we redefine the predicates of comparison.
+
+\label{Predicates:Lt-E0}
+
+\begin{Coqsrc}
+Definition Lt (alpha beta : E0) := T1.LT (@cnf alpha) (@cnf beta).
+Definition Le (alpha beta : E0) := T1.LE (@cnf alpha) (@cnf beta).
+
+Infix "o<" := Lt : E0_scope.
+Infix "o<=" := Le : E0_scope.
+\end{Coqsrc}
+
+\begin{todo}
+Infix notation for inequalities on ordinals have changed to \texttt{o<} and \texttt{o<=}.  This document is not updated yet, but the \coq{} scripts are correct.
+\end{todo}
+
+Equality in \texttt{E0} is just Leibniz equality. Note that, since \texttt{nf} is
+defined by a Boolean function, for  any term $\alpha:\texttt{T1}$, there exists at most one proof of \texttt{nf $\alpha$}, thus two ordinals of type \texttt{E0} are
+equal if and only iff their projection to \texttt{T1} are equal (see also Sect.~\vref{sect:eq-proof-unicity}).
+
+\index{Coq!Techniques!Unicity of equality proofs}
+
+\begin{Coqsrc}
+Require Import Logic.Eqdep_dec.
+
+Lemma nf_proof_unicity :
+  forall (alpha:T1) (H H': nf alpha), H = H'.
+
+Lemma E0_eq_iff alpha beta : alpha = beta <-> cnf alpha = cnf beta.
+\end{Coqsrc}
+
+For upgrading constants and fonctions of \texttt{T1}, we have to prove that 
+the term they build is in normal form.
+For instance, let us represent the ordinals $0$ and $\omega$ as instances of the class \texttt{E0}.
+
+\label{sect:omega-T1}
+\index{Constants!zero:T1}
+
+\begin{Coqsrc}
+Instance Zero : E0.
+Proof.
+  refine (@mkord T1.zero _);  now compute. 
+Defined.
+
+Instance _Omega : E0.
+Proof.  now exists omega. Defined.
+
+Notation "'omega'"  := _Omega : E0_scope.
+\end{Coqsrc}
+
+\subsection{Syntactic definition of limit and successor ordinals}
+
+Pattern matching and structural recursion allow us to define the notions of successor and limit ordinal with the help of  boolean functions on type \texttt{T1}.
+
+ \vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}
+
+\begin{Coqsrc}
+  Fixpoint succb alpha :=
+  match alpha with
+      zero => false
+    | ocons zero _ _ => true
+    | ocons alpha n beta => succb beta
+  end.
+
+Fixpoint limitb alpha :=
+  match alpha with
+      zero => false
+    | ocons zero _ _ => false
+    | ocons alpha n zero => true
+    | ocons alpha n beta => limitb beta
+  end.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+  Compute limitb omega.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+  = true
+     : bool
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Compute succb 42.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+  = true
+     : bool
+   \end{Coqanswer}
+
+The correctness of these definitions with respect to the mathematical notions of
+limit and successor ordinals is established through several lemmas. For instance,
+Lemma \texttt{canonS\_limit}, page~\pageref{lemma:canonS-limit}, shows that
+if $\alpha$ is (syntactically) a limit ordinal, then it is the least upper bound of
+a strictly increasing sequence of ordinals.
+
+   The following function is very useful in constructions by cases (proofs and function definitions).
+
+\begin{Coqsrc}
+  Definition zero_succ_limit (alpha: T1) :
+  {succb alpha} + {limitb alpha} +  {alpha=zero}.
+  (* ... *)
+Defined.
+ \end{Coqsrc}
+
+\subsection{Arithmetic on \texorpdfstring{$\epsilon_0$}{epsilon0}}
+\subsubsection{Successor}
+
+The successor of any ordinal $\alpha< \epsilon_0$ is defined by structural 
+recursion on its Cantor normal form.
+
+\label{Functions:succ-T1}
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}
+
+\begin{Coqsrc}
+Fixpoint succ (alpha:T1) : T1 :=
+  match alpha with 
+   | zero => 1
+   | ocons zero n _ => ocons zero (S n) zero
+   | ocons beta n gamma => ocons beta n (succ gamma)
+ end.
+\end{Coqsrc}
+
+The following lemma establishes the connection between the function
+\texttt{succ} and the predicate \texttt{is\_succ}.
+
+\begin{Coqsrc}
+ Lemma succb_iff alpha (Halpha : nf alpha) :
+  succb alpha <-> exists beta : T1, nf beta /\ alpha = succ  beta.
+\end{Coqsrc}
+
+\index{Exercises}
+ \begin{exercise}
+Prove in \coq{} that for any ordinal $\alpha<\epsilon_0$, $\alpha$ is a limit if 
+and only if for all $\beta<\alpha$, the interval $[\beta,\alpha)$ is infinite.
+ \end{exercise}
+
+\subsubsection{Addition and multiplication}
+
+Ordinal addition and multiplication are also defined by structural recursion over the type \texttt{T1}. Please note that they use the \texttt{compare} function on some subterms of their arguments.
+
+\label{sect:infix-plus-T1}
+
+\begin{Coqsrc}
+Fixpoint plus (alpha beta : T1) : T1 :=
+  match alpha,beta with
+ |  zero, y  => y
+ |  x, zero  => x
+ |  ocons a n b, ocons a' n' b' =>
+    (match compare a a' with
+     | Lt => ocons a' n' b'
+     | Gt => (ocons a n (plus b (ocons a' n' b')))
+     | Eq  => (ocons a (S(n+n')) b')
+     end)
+  end
+where "alpha + beta" := (plus alpha beta) : t1_scope.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Fixpoint mult (alpha beta : T1) :T1 :=
+  match alpha,beta with
+ |  zero, y  => zero
+ |  x, zero => zero
+ |  ocons zero n _, ocons zero n' _ => 
+                 ocons zero (Peano.pred((S n) * (S n'))) zero
+ |  ocons a n b, ocons zero n' b' =>  
+                 ocons a (Peano.pred((S n) * (S n'))) b
+ |  ocons a n b, ocons a' n' b' =>
+     ocons (a + a') n' ((ocons a n b) * b')
+ end
+where  "alpha * beta" := (mult alpha beta) : t1_scope.
+\end{Coqsrc}
+
+\subsubsection{Examples}
+
+The following examples are instances of \emph{proofs by computation}. Please note that  addition and multiplication on \texttt{T1}
+are not commutative. Moreover,  both operations fail to be strictly monotonous in their first argument.
+
+\begin{Coqsrc}
+Example e2 : 6 + omega = omega.
+Proof. reflexivity. Qed.
+
+Example e'2 : omega o< omega + 6.
+Proof. now compute. Qed.
+
+Example e''2 : 6 * omega = omega.
+Proof. reflexivity. Qed.
+
+Example e'''2 : omega o< omega * 6.
+Proof. now compute. Qed.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma plus_not_monotonous : exists alpha beta gamma : T1,
+    alpha o< beta /\ alpha + gamma = beta + gamma.
+Proof.
+  exists 3, 5, omega;  now  compute.
+Qed.
+
+Lemma mult_not_monotonous :  exists alpha beta gamma : T1,
+      alpha o< beta /\ alpha * gamma = beta * gamma.
+Proof.
+  exists 3, 5, omega; now compute.
+Qed.
+\end{Coqsrc}
+
+\subsection{Pretty printing Ordinals in Cantor normal Form}
+
+Let us consider again the ordinal $\alpha_0$ defined in section~\vref{alpha0-def}
+
+If we ask \coq{} to print the normal form of \texttt{alpha\_0}, we get a hardly readable term of type \texttt{T1}.
+
+\begin{Coqsrc}
+Compute alpha_0.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+  = ocons omega 0 (ocons (FS 2) 4 (FS 1))
+     : T1
+\end{Coqanswer}
+
+The following data type defines a abstract syntax fore more readable ordinals in Cantor normal form:
+
+\label{types:ppT1}
+\index{Functions!pp@ pp (pretty printing terms in Cantor normal form)}
+
+\begin{Coqsrc}
+Inductive ppT1 : Set :=
+    P_fin : nat -> ppT1
+  | P_add : ppT1 -> ppT1 -> ppT1
+  | P_mult : ppT1 -> nat -> ppT1
+  | P_exp : ppT1 -> ppT1 -> ppT1
+  | P_omega : ppT1
+\end{Coqsrc}
+
+The function \texttt{pp: T1 -> ppT1} converts any closed term of type \texttt{T1} into a human-readable expression. For instance, let us convert the term \texttt{alpha\_0}.
+
+\begin{Coqsrc}
+Compute pp alpha_0.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+     = (omega ^ omega + omega ^ 3 * 5 + 2)%pT1
+     : ppT1
+\end{Coqanswer}
+
+\subsubsection{Arithmetic on type \texttt{E0}}
+
+ We define an addition in type \texttt{E0}, using the function{T1.plus}
+and its properties.
+
+\begin{Coqsrc}
+Search   (nf (_ + _)%t1).
+\end{Coqsrc}
+
+\begin{Coqanswer}
+plus_nf: forall a : T1, nf a -> forall b : T1, nf b -> nf (a + b)
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Instance plus (alpha beta : E0) : E0.
+Proof.
+  refine (@mkord (T1.plus (@cnf alpha) (@cnf beta))_ );
+    apply plus_nf; apply cnf_ok.
+Defined.
+
+Infix "+" := plus : E0_scope.
+
+Check omega + omega.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+omega + omega
+     : E0
+\end{Coqanswer}
+
+\index{Exercises}
+ \begin{exercise}
+Let $\alpha$ be an ordinal. We say that $\alpha$ is \emph{infinite} iff the segment $[0,\alpha)$ is an infinite set. 
+\begin{enumerate}
+\item Adapt this definition to the type \texttt{E0}.
+\item Prove that \emph{being infinite} is decidable on type \texttt{E0}.
+\item Prove that $\alpha$ is infinite if and only if for all finite ordinal $i$,
+$i+\alpha=\alpha$.
+\end{enumerate}
+ \end{exercise}
+
+\section{Well-foundedness and transfinite induction}
+
+\index{Maths!Transfinite induction}
+
+\subsection{About  well-foundedness}
+\label{sec:orgheadline82}
+   In order to use \texttt{T1} for proving termination results,
+we need to prove that  our representation of ordinals strictly less than \(\epsilon_0\)
+makes our order \texttt{<} well-founded. Then we will get \emph{transfinite induction} for free.
+
+The proof of well-foundedness of the strict order $<$ on Cantor normal forms is already 
+available in the Cantor contribution by Castéran and Contéjean~\cite{CantorContrib}. That proof relies on a library on recursive path orderings written by
+E. Contéjean. We present also a direct proof of the same result, which does not require any knowledge on r.p.o.s.
+
+\index{Exercises}
+
+\begin{exercise}
+Prove that the \emph{total} order \texttt{lt} on \texttt{T1} is not well-founded. 
+\textbf{Hint:}  You will have to build a counter-example with terms of type \texttt{T1}
+which are not in Cantor normal form.
+\end{exercise}
+
+% \subsubsection{The total order \texttt{lt} on \texttt{T1} is \emph{not} well-founded}
+% \label{sec:orgheadline76}
+
+% Let us recall that the data type \texttt{T1} contains too many inhabitants, including
+% terms which are not in Cantor normal form. Thus, the following result is not 
+% very surprising.
+
+% \begin{Coqsrc}
+% Section lt_not_well_founded.
+
+%   (* let us build the sequence of terms :
+%         omega + omega + .... + omega ^ 2   *)
+%   Let f := (fix f (i:nat): T1 :=
+%             match i with 0 => phi0 2
+%                        | S i => ocons 1 1 (f  i)
+%             end).
+
+%  Lemma  f_decreases : forall i, f (S i) <  f i.
+%  Proof.
+%   induction i; compute; auto with T1.
+%  Qed.
+
+%  Theorem lt_not_wf : ~  well_founded lt.
+%  Proof. 
+%    intro wf; case (not_decreasing _ lt);auto.
+%    exists f; apply f_decreases.
+%  Qed.
+
+% End lt_not_well_founded.
+% \end{Coqsrc}
+
+% Thus, we have
+
+\subsubsection{A first attempt}
+\label{sec:orgheadline77}
+\index{Coq!Techniques!Well founded relations}
+
+It is natural to try to prove by structural induction over \texttt{T1} 
+that every term in normal form is accessible through \texttt{LT}.
+
+Unfortunately, it won't work. Let us consider some well-formed term
+ $\alpha=\texttt{ocons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are accessible
+ through \texttt{LT}. For proving the accessibility of $\alpha$, we have to consider
+any well formed term \(\delta\) such that \(\delta<\alpha\). 
+But nothing guarantees that \(\delta\)  is strictly  less than \(\beta\) nor \(\gamma\), and we cannot use the induction hypotheses on   \(\beta\) nor \(\gamma\).
+
+\begin{Coqbad}
+Section First_attempt.
+
+ Lemma wf_LT : forall alpha,  nf alpha -> Acc LT alpha. 
+ Proof.
+  induction alpha as [| beta IHbeta n gamma IHgamma].
+  - split.
+    inversion 1.
+    destruct H2 as [H3 _];not_neg H3.
+  -  split; intros delta Hdelta.
+\end{Coqbad}
+
+\begin{Coqanswer}
+1 subgoal (ID 560)
+
+  beta : T1
+  n : nat
+  gamma : T1
+  IHbeta : nf beta -> Acc LT beta
+  IHgamma : nf gamma -> Acc LT gamma
+  H : nf (ocons beta n gamma)
+  delta : T1
+  Hdelta : delta o< ocons beta n gamma
+  ============================
+  Acc LT delta
+ \end{Coqanswer}
+
+\begin{Coqbad}
+  Abort.
+\end{Coqbad}
+
+The problem comes from that \(\delta\) may be bigger that \(\beta\) or
+\(\gamma\);
+for instance \(\delta\) may be of the form
+\texttt{ocons $\beta'$ $p'$  $\gamma'$},
+where  \(\beta' \leq  \beta\) and  \(p' < n\).
+Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.
+
+\subsubsection{Using a stronger inductive predicate.}
+\label{sec:orgheadline78}
+  Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
+through \texttt{LT}, we propose to show first that any well formed 
+term of the form \(\omega^\alpha\times(n+1)+\beta\) is accessible (which is a stronger result).
+
+\begin{Coqsrc}
+ Let Acc_strong (alpha:T1) :=
+      forall n beta, 
+        nf (ocons alpha n beta) -> Acc LT (ocons alpha  n beta).
+\end{Coqsrc}
+
+The following lemma is an application of the strict inequality 
+\showmath{\alpha < \omega ^\alpha}. If \showmath{\omega^\alpha} is accessible,
+then \showmath{\alpha} is \emph{a fortori} accessible.
+
+\begin{Coqsrc}
+ Lemma Acc_strong_stronger : forall alpha, 
+     nf alpha -> Acc_strong  alpha -> Acc LT  alpha.
+ Proof.
+  intros alpha H H0; apply acc_imp with (phi0 alpha).
+  - repeat split; trivial.
+    + now apply lt_a_phi0_a.
+  -  apply H0;  now apply single_nf.
+Qed.
+\end{Coqsrc}
+
+Thus, it remains to prove that every ordinal strictly less than \showmath{\epsilon_0} 
+is strongly accessible.
+
+% \subsubsection{Structure of the proof of well-foundedness of \texttt{LT}}
+
+\label{sec:orgheadline81}
+\label{proof-wf-epsilon0}
+\paragraph{A helper}
+\label{sec:orgheadline79}
+
+First, we prove that, for  any \texttt{LT}-accessible term \showmath{\alpha}, any well formed
+term (\texttt{ocons $\alpha$ $n$ $\beta$})  is accessible too.
+
+\begin{Coqsrc}
+Lemma Acc_implies_Acc_strong : 
+   forall alpha, Acc LT  alpha -> Acc_strong alpha.
+\end{Coqsrc}
+
+The proof is structured as an induction on \showmath{\alpha}'s accessibility. Let us consider
+an accessible term $\alpha$.
+
+\begin{Coqanswer}
+  subgoal 1
+
+  alpha : T1
+  Aalpha : forall y : T1,  y o< alpha -> Acc LT y
+  IHalpha : forall y : T1,
+       LT y alpha ->
+       forall (n : nat) (beta : T1),
+       nf (ocons y n beta) -> Acc LT (ocons y n beta)
+  ============================
+   forall (n : nat) (beta : T1),
+   nf (ocons alpha n beta) -> Acc LT (ocons alpha n beta)
+\end{Coqanswer}
+
+Let \texttt{n:nat} and \texttt{beta:T1} such that \texttt{ocons alpha n beta} is in normal form. 
+We prove first that \texttt{beta} is accessible,  which allows us to by well-founded induction on \texttt{beta}, 
+and natural induction on \texttt{n}, that (\texttt{ocons alpha n beta}) is accessible.
+The proof, quite long, can be consulted in \url{../src/html/hydras.Epsilon0.T1.html}
+
+\paragraph{Accessibility of any well-formed ordinal term}
+\label{sec:orgheadline80}
+
+Our goal is still to prove accessibility of any well formed ordinal term.
+Thanks to our previous lemmas, we are almost done.
+
+\begin{Coqsrc}
+(* A (last) structural induction *)
+
+Theorem nf_Acc : forall alpha, nf alpha -> Acc LT  alpha.
+Proof.
+ induction alpha.
+-  intro; apply Acc_zero.
+ -  intros; eapply Acc_implies_Acc_strong;auto.
+    apply IHalpha1;eauto.
+    apply nf_inv1 in H; auto. 
+Defined.
+
+Corollary T1_wf : well_founded LT.
+\end{Coqsrc}
+
+And we have now our transfinite recursor:
+\index{Maths!Transfinite induction}
+
+\begin{Coqsrc}
+
+Definition transfinite_recursor :
+ forall (P:T1 -> Type),
+   (forall x:T1, 
+     (forall y:T1, nf x -> nf y ->  lt y  x -> P y) -> P x) ->
+    forall alpha:T1, P alpha.
+Proof.
+ intros; apply well_founded_induction_type with LT.
+ -  exact T1_wf;auto.
+ - intros. apply X. intros; apply X0. repeat split;auto. 
+Defined.
+\end{Coqsrc}
+
+The following tactic starts a proof by  transfinite induction on any ordinal \mathcolor{$\alpha<\epsilon_0$}.
+
+\begin{Coqsrc}
+Ltac transfinite_induction alpha :=
+  pattern alpha; apply transfinite_recursor;[ | try assumption].
+\end{Coqsrc}
+
+\begin{remark}
+The alternate proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering is available in the library \href{../src/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
+ \end{remark}
+
+\subsection{An ordinal notation for  \texorpdfstring{$\epsilon_0$}{epsilon0}}
+
+We built an instance of \texttt{ON}, and proved its correction w.r.t. Schutte's model.
+
+\begin{Coqsrc}
+Instance Epsilon0 : ON Lt compare.  
+(* ... *)
+\end{Coqsrc}
+
+\emph{From Module~\href{../src/html/hydras.Schutte.Schutte.Correctness_E0.html}{Schutte.Schutte.Correctness\_E0}}
+
+\begin{Coqsrc}
+Instance Epsilon0_correct :
+  ON_correct epsilon0 Epsilon0 (fun alpha => inject (cnf alpha)).
+\end{Coqsrc}
+
+\index{Projects}
+\begin{project}
+ \emph{This exercise is a continuation of Project~\vref{exo:ON-mult}.}
+Define an ordinal notation \texttt{Omega2} for $\omega^2=\omega\times\omega$ (using
+\texttt{ON\_mult}).
+
+Prove that \texttt{Omega2} is a sub-notation of \texttt{Epsilon0}.
+
+Define on \texttt{Omega2} an addition compatible with the addition on \texttt{Epsilon0}.
+
+\textbf{Hint}. You may use the following definition (in 
+\href{../src/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}).
+
+\begin{Coqsrc}
+Definition SubON_same_op  `{OA : @ON A ltA  compareA}
+       `{OB : @ON B ltB  compareB}
+       {iota : A -> B} 
+       {alpha: B}
+       {_ : SubON OA OB alpha iota}
+       (f : A -> A -> A)
+       (g : B -> B -> B)
+  :=
+  forall x y,  iota (f x y) = g (iota x) (iota y).
+\end{Coqsrc}
+
+\end{project}
+
+\index{Projects}
+\begin{project}
+The class \texttt{ON} of ordinal notations has been defined long after this 
+chapter, and is not used yet in the development of the type \texttt{E0}. 
+A better integration of both notions should simplify the development on ordinals in Cantor normal form. This integration is planned for the future versions.
+
+\end{project}
+
+\section{A variant for hydra battles}
+
+\subsection{Natural sum (a.k.a. Hessenberg's  sum)}
+\label{sec:orgheadline87}
+\label{hydra-variant}
+
+Natural sum (Hessenberg's  sum) is a commutative and monotonous version of
+addition. It is used as an auxiliary operation  for defining variants
+for hydra battles, where Hercules is allowed to chop off any  head of the hydra.
+
+In the litterature, the natural sum of ordinals \(\alpha\) and \(\beta\) 
+is often denoted by \(\alpha \# \beta\)  or  \(\alpha \oplus  \beta\).
+Thus we called \texttt{oplus} the associated \emph{Coq} function.
+
+\subsubsection{Definition of \texttt{oplus}}
+\label{sec:orgheadline84}
+\index{Functions!oplus @ oplus (Hessenberg commutative sum)}
+
+The definition of \texttt{oplus} is recursive in both of its 
+arguments, which makes a structural recursive definition a little 
+complex.
+We used the same pattern as for the \texttt{merge} function on lists of library
+\texttt{Coq.Sorting.Mergesort}.
+
+\begin{enumerate}
+\item Define a nested recursive function, using the \texttt{Fix} 
+    construct
+
+\item Build a principle of induction dedicated to \texttt{oplus}
+
+\item Establish equations associated to each case of the definition.
+\end{enumerate}
+
+\paragraph{Nested recursive definition}
+\label{sec:orgheadline83}
+
+The following definition is composed of 
+\begin{itemize}
+\item A main function \texttt{oplus}, structurally recursive in its 
+first argument \texttt{alpha}
+\item An auxiliary function \texttt{oplus\_aux} within the scope of \texttt{alpha},
+structurally recursive in its argument \texttt{beta};  \texttt{oplus\_aux beta} 
+   is supposed to compute  \texttt{oplus alpha beta}.
+\end{itemize}
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}
+
+\label{sect:infix-oplus}
+
+\begin{Coqsrc}
+Fixpoint oplus (alpha beta : T1) : T1 :=
+  let fix oplus_aux beta {struct beta} :=
+      match alpha, beta with
+        | zero, _ => beta
+        | _,  zero => alpha
+        | ocons a1 n1 b1, ocons a2 n2 b2 =>
+          match compare a1 a2 with
+            |  Gt => ocons a1 n1 (oplus b1 beta)
+            |  Lt => ocons a2 n2 (oplus_aux b2)
+            |  Eq => ocons a1 (S (n1 + n2)%nat) (oplus b1 b2)
+          end
+      end
+  in oplus_aux beta.
+
+Infix "o+" := oplus  (at level 50, left associativity).
+\end{Coqsrc}
+
+The reader will note that each recursive call of the functions
+\texttt{oplus} and \texttt{oplus\_aux} satisfies \emph{Coq}'s constraint
+on recursive definitions. The function \texttt{oplus} is recursively called on a sub-term of its first argument,
+and \texttt{oplus\_aux} on a sub-term of its unique argument.
+Thus, \texttt{oplus}'s definition is accepted by \coq{} as a structurally recursive function.
+
+\subsubsection{Rewriting lemmas}
+\label{sec:orgheadline86}
+
+\emph{Coq}'s constraints on recursive definitions resulted in 
+the quite  complex form of \texttt{oplus}'s definition.
+For making easier proof of properties of this function, 
+it is helpful to \emph{derive} lemmas that will help to simplify 
+expressions of the form (\texttt{oplus $a$ $ b$}).
+
+A first set of lemmas correspond to the various cases of \texttt{oplus}'s 
+definition. They can be proved almost immediately, using \texttt{cbn} 
+and \texttt{rewrite} tactics.
+
+\begin{Coqsrc}
+Lemma oplus_alpha_0 (alpha : T1) : alpha o+ zero = alpha.
+Proof.
+  destruct alpha; reflexivity.
+Qed.
+
+Lemma oplus_0_beta (beta : T1): zero o+ beta = beta.
+Proof.
+  destruct beta; reflexivity.
+Qed.
+\end{Coqsrc}
+
+\subsubsection{A hand-made induction principle}
+\label{sec:orgheadline85}
+
+\index{Coq!Commands!Functional Scheme}
+
+\emph{Coq} contains a command  \texttt{Functional Scheme} that 
+generates induction principles which correspond to recursive functions.
+Unfortunately, the current version ( \texttt{8.11.0} ) doesn't work on \texttt{oplus},
+probably because of the inner \texttt{Fix}.
+
+\begin{Coqsrc}
+Functional Scheme oplus_ind := Induction for oplus Sort Prop.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+Error: Anomaly "todo." Please report at http://coq.inria.fr/bugs/.
+\end{Coqanswer}
+
+Fortunately, it's a good exercise for a semi-experienced user, to write
+her/him-self induction principles similar to the ones returned by
+\texttt{Functional Scheme}.
+
+\begin{itemize}
+\item First, we chose to write a version for sort \texttt{Type}, since versions
+for sorts \texttt{Prop} and \texttt{Set} can be easily derived from
+the former one. According to \emph{Coq}'s naming politics, we will call our 
+principle \texttt{oplus\_rect}
+
+\item The conclusion of \texttt{oplus\_rect} will be (\texttt{$P$ a b (oplus a b)}),
+where $P$ is an arbitrary function of type 
+\texttt{T1 -> T1 -> T1 -> Type}
+
+\item The premises of \texttt{oplus\_rect} will describe how to build an induction 
+on the graph of \texttt{oplus}.
+\end{itemize}
+
+We are now ready to state and prove \texttt{oplus\_rect}, and the reader
+will note that the statement is longer than the proof script itself,
+which is a standard proof by induction, simplification and case-analysis 
+that follows  \texttt{oplus}'s definition.
+
+We associate also a tactic to the application of \texttt{oplus\_rect}.
+
+\begin{Coqsrc}
+ Lemma oplus_rect:
+      forall P: T1 -> T1 -> T1 -> Type, 
+        (forall a:T1, P zero a a) ->
+        (forall a: T1, P a zero a) ->
+        (forall a1 n1 b1 a2 n2 b2 o,
+           compare a1 a2 = Gt ->
+           P b1 (ocons a2 n2 b2) o ->
+           P (ocons a1 n1 b1) (ocons a2 n2 b2)
+             (ocons a1 n1 o)) ->
+        (forall a1 n1 b1 a2 n2 b2 o,
+           compare a1 a2 = Lt ->
+           P (ocons a1 n1 b1) b2 o ->
+           P (ocons a1 n1 b1) (ocons a2 n2 b2) 
+           (ocons a2 n2 o)) ->
+        (forall a1 n1 b1 a2 n2 b2 o,
+           compare a1 a2 = Eq ->
+           P b1 b2 o ->
+          P (ocons a1 n1 b1) (ocons a2 n2 b2)
+            (ocons a1 (S (n1 + n2)%nat) o)) ->
+         forall a b, P a b (oplus a b).
+Proof with auto.
+   induction a.
+   -    intro; simpl; destruct b;auto.
+   -   induction b.
+       + apply X0.
+       + case_eq (compare a1 b1).
+         * intro Comp; unfold oplus; rewrite Comp.
+           cbn; apply X3 ...
+         * intro Comp; cbn; rewrite Comp; apply X2...
+         * intro Comp; cbn; rewrite Comp ...
+ Defined.
+
+Ltac oplus_induction a b:= pattern (oplus a b); apply oplus_rect.
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+The induction principle \texttt{oplus\_rect} is still unused in our development. 
+Please build some nice examples of application.
+\end{exercise}
+
+\index{Projects}
+\begin{project}
+Compare \texttt{oplus}'s definition (with inner fixpoint) with other possibilities
+(\texttt{coq-equations}, \texttt{Function}, etc.).
+\end{project}
+\subsection{More theorems on Hessenberg's sum}
+
+We need to prove some properties of $\oplus$, particularly about 
+its relation with the order $<$ on \texttt{T1}.
+
+\subsubsection{Boundedness}
+If $\alpha$ and $\beta$ are both strictly  less than  $\omega^\gamma$, then so is their natural sum
+$\alpha \oplus \beta$. This result can be proved by structural induction on $\gamma$.
+
+\begin{Coqsrc}
+Lemma lt_phi0_oplus : forall gamma alpha beta,
+                        lt_phi0 alpha gamma ->
+                        lt_phi0 beta gamma ->
+                        lt_phi0 (alpha o+ beta) gamma.
+
+Proof with auto.
+  induction gamma; destruct alpha, beta.  
+(* ... *)
+\end{Coqsrc}
+
+\subsubsection{Commutativity, associativity}
+
+We prove  the commutativity of $\oplus$ in two steps.
+
+First, we prove by transfinite induction on $\alpha$ that the restriction of $\oplus$ to the
+interval $[0..\alpha)$ is commutative.
+
+\index{Maths!Transfinite induction}
+
+\begin{Coqsrc}
+Lemma oplus_comm_0 : forall alpha, nf alpha ->
+     forall a b,  nf a -> nf b ->
+                  lt a alpha ->
+                  lt b alpha ->
+                  a o+ b = b o+ a.
+ Proof with eauto with T1.
+    intros alpha Halph; transfinite_induction alpha.
+(* rest of proof omitted *)  
+\end{Coqsrc}
+
+Then, we infer  $\oplus$'s commutativity for any pair of ordinals:
+Let $\alpha$ and $\beta$ be two ordinals strictly less than $\epsilon_0$. Both ordinals $\alpha$ and $\beta$ are
+strictly less than $\textrm{max}(\alpha,\beta)+1$.
+    Thus, we have just to apply the lemma \coqsimple{oplus\_comm\_0}.
+
+\begin{Coqsrc}
+  Lemma oplus_comm : forall alpha beta, 
+      nf alpha -> nf beta ->
+      alpha o+ beta =  beta o+ alpha.
+  Proof with eauto with T1.
+    intros alpha beta Halpha Hbeta;
+    apply oplus_comm_0 with (succ (max alpha beta)) ...  
+  (* rest of proof omitted *)
+\end{Coqsrc}
+
+The associativity of Hessenberg's sum is proved the same way.
+
+\begin{Coqsrc}
+ Lemma oplus_assoc_0 :
+    forall alpha,
+      nf alpha ->
+      forall a b c,  nf a -> nf b -> nf c ->
+                      lt a alpha ->
+                      lt b alpha -> lt c alpha ->
+                      a o+ (b o+ c) = (a o+ b) o+ c.
+  Proof with eauto with T1.
+    intros alpha Halpha.
+    transfinite_induction alpha.
+(* rest of proof omitted *) 
+\end{Coqsrc}
+
+\begin{Coqsrc}
+ Lemma oplus_assoc : forall alpha beta gamma,
+                        nf alpha -> nf beta -> nf gamma ->
+                                    alpha o+ (beta o+ gamma) =
+                                    alpha o+ beta o+ gamma.
+ Proof with eauto with T1.
+    intros;
+    apply oplus_assoc_0 with (succ (max alpha (max beta gamma))) ...
+(* rest of proof omitted *)   
+\end{Coqsrc}
+
+\subsubsection{Monotonicity}
+
+At last, we prove that $\oplus$ is strictly monotonous in both of its arguments.
+
+\begin{Coqsrc}
+Lemma oplus_strict_mono_LT_l (alpha beta gamma : T1) :
+  nf gamma   -> alpha  o< beta ->
+  alpha o+ gamma o< beta o+ gamma.
+
+Lemma oplus_strict_mono_LT_r (alpha beta gamma : T1) :
+  nf alpha -> beta o< gamma ->
+  alpha o+ beta o< alpha o+ gamma.
+\end{Coqsrc}
+
+\subsection{A measure for hydra-battle termination}
+\label{sec:orgheadline90}
+\label{sec:hydra-measure}
+
+Let us define a measure from type \texttt{Hydra} into \texttt{T1}.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Termination.html\#m}{Hydra.Hydra\_Termination}}
+
+\begin{Coqsrc}
+Fixpoint m (h:Hydra) : T1 :=
+  match h with head => zero
+             | node hs => ms hs
+end 
+with ms (s:Hydrae) :  T1 :=
+  match s with  hnil => zero
+              | hcons h s' => phi0 (m h) o+  ms s'
+ end.  
+\end{Coqsrc}
+
+First, we prove that the measure $m(h)$  of any hydra $h$ is a well-formed ordinal term of type \texttt{T1}.
+
+\begin{Coqsrc}
+Lemma m_nf : forall h, nf (m h).
+Proof.
+ intro h; elim h using Hydra_rect2 
+            with (P0 := fun s =>  nf (ms s)).
+ (* ... *)
+
+Lemma ms_nf : forall s, nf (ms s).
+Proof with auto with T1.
+(* ... *)
+\end{Coqsrc}
+
+For proving the termination of all hydra battles, we have to prove that
+\texttt{m} is a variant. First, a few technical lemmas follow the decomposition of \texttt{round} into several relations. Then the lemma \texttt{round\_decr} gathers all the cases.
+
+\begin{Coqsrc}
+Lemma S0_decr :
+  forall s s', S0  s s' -> ms s' o< ms s.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma R1_decr : forall h h',
+                  R1 h h' -> m h' o< m h.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma S1_decr n:
+  forall s s', S1 n s s' -> ms s' o<  ms s.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma R2_decr n : forall h h', R2 n h h' -> m h'  o< m h.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma round_decr : forall h h', h -1-> h' -> m h' o< m h.
+Proof.
+   destruct 1 as [n [H | H]].
+   -  now apply R1_decr.
+   -  now apply R2_decr with n.
+Qed.
+\end{Coqsrc}
+
+Finally, we prove the termination of all (free) battles.
+
+\begin{Coqsrc}
+Global Instance HVariant : Hvariant lt_wf free var.
+Proof.
+ split; intros; eapply round_decr; eauto.
+Qed.
+
+Theorem every_battle_terminates: Termination.
+Proof. 
+  red; apply Inclusion.wf_incl with 
+         (R2 := fun h h' =>  m h o< m h').
+   red; intros;  now apply round_decr.
+   apply Inverse_Image.wf_inverse_image, T1_wf.
+Qed.
+\end{Coqsrc}
+
+\section*{Conclusion}
+
+Let us recall three results we have proved so far.
+\begin{itemize}
+\item There exists a strictly decreasing variant which maps \texttt{Hydra} into 
+the segment $[0,\epsilon_0)$ for proving the termination of any hydra battle
+\item There exists \emph{no} such variant from \texttt{Hydra} into 
+$[0,\omega^2)$, \emph{a fortiori} into $[0,\omega)$.
+\end{itemize}
+
+So, a  natural question is `` Does there exist any strictly decreasing variant mapping
+type \texttt{Hydra} into some interval $[0,\alpha[$ (where $\alpha <\epsilon_0$) for proving the termination of all hydra battles''.
+The next chapter is dedicated to a formal proof that there exists no such $\alpha$, even if we consider a restriction to the set of ``standard'' battles.
+
+%\include{epsilon0}
+
+%\include{impossibility-proofs}
+
+%-------------------------------------------------------------------
+
+\chapter[The Ketonen-Solovay machinery]{Strolling inside \texorpdfstring{$\epsilon_0$}{epsilon0}: The Ketonen-Solovay machinery\label{ks-chapter}}
+\label{chap:ketonen}
+\index{Maths!Ketonen-Solovay machinery}
+
+\section{Introduction}
+The reader may think that our proof of termination in the previous  chapter requires a lot of mathematical tools and may be too  complex. So, the question is ``is there  any  simpler proof'' ?
+
+In their article~\cite{KP82}, Kirby and Paris show that this result cannot be proved in Peano arithmetic.
+Their proof uses some knowledge about model theory and non-standard models of Peano arithmetic.
+In this chapter, we focus on a specific class of proofs of termination of hydra battles: construction of some variant mapping the type \texttt{Hydra} into some segment of ordinals.
+Our proof relies only on the Calculus of Inductive Constructions and is a natural complement of the results proven in the previous chapter.
+
+\begin{itemize}
+\item There is no variant mapping the type \texttt{Hydra} into the interval $[0,\omega^2)$ (section ~\vref{omega2-case}), and a fortiori 
+$[0,\omega)$
+
+\item There is a variant that maps the type \texttt{Hydra} into the
+interval $[0,\epsilon_0)$ (theorem \texttt{every\_battle\_terminates}, in section~\vref{hydra-variant}).
+\end{itemize}
+
+Thus, a very natural question is the following one:
+\begin{quote}
+  `` Is there  any variant from
+\texttt{Hydra} into some interval $[0,\mu)$, where $\mu<\epsilon_0$, for proving the termination of all hydra battles ?''
+\end{quote}
+
+We prove in \coq{} the following results:
+
+\begin{quote}
+There is no variant for proving the termination of all hydra battles
+from \texttt{Hydra} into the interval $[0..\mu)$, where
+$\mu< \epsilon_0$.
+The same impossibility holds even if we consider only standard battles (with the successive replication factors $0,1,2,\dots,t,t+1,\dots$).
+\end{quote}
+
+Our proofs are  constructive and require no axioms: they are  closed terms of the CIC, and are mainly composed on function definitions and proofs of properties of these functions. 
+They  share much theoretical material with Kirby and Paris', although they do not use any knowledge about Peano arithmetic nor model  theory.  The combinatorial arguments we use and implement
+come from 
+ an article by J.~Ketonen and R.~Solovay~\cite{KS81}, already  cited in the work
+ by L.~Kirby et J.~Paris.% on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
+ Section $2$ of this article: ''A hierarchy of probably recursive functions'', contains a systematic study of \emph{canonical sequences}, which are closely related to
+rounds of hydra battles. 
+Nevertheless, they have the same global structure as the simple proofs described in
+sections~\vref{omega-case} and \vref{omega2-case}. 
+We invite the reader to compare the three proofs step by step, lemma by lemma.
+
+\section{Canonical Sequences}
+\label{ketonen-solovay-sect}
+\index{Maths!Canonical sequences}
+
+Canonical sequences are functions that associate an ordinal $\canonseq{\alpha}{i}$ to every ordinal $\alpha<\epsilon_0$ and positive integer $i$. They satisfy nice properties :
+
+\index{Maths!Transfinite induction}
+\begin{itemize}
+\item If $\alpha\not=0$, then $\canonseq{\alpha}{i}<\alpha$. Thus canonical sequences can be used for proofs by transfinite induction or function definition by transfinite recursion
+\item If $\lambda$ is a limit ordinal, then $\lambda$ is the least upper bound of the set 
+$\{\canonseq{\lambda}{i}\;|\,i\in\mathbb{N}_1\}$
+
+\item If $\beta<\alpha<\epsilon_0$, then there is a ``path'' from $\alpha$ to $\beta$, \emph{i.e.} a
+sequence $\alpha_0=\alpha, \alpha_1, \dots, \alpha_n=\beta$, where for every $k<n$, there exists some $i_k$ such that $\alpha_{k+1}=\canonseq{\alpha_k}{i_k}$
+\item Canonical sequences correspond tightly to rounds of hydra battles: if $\alpha\not=0$,
+then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1}$ in one round with
+the replication factor $i$ (Lemma \href{../src/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
+\item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) battle from $\iota(\alpha)$ to $\iota(\beta)$.
+\end{itemize}
+
+\begin{remark}
+  In~\cite{KS81}, canonical sequences are defined for any ordinal $\alpha <\epsilon_0$,
+by stating that if $\alpha$ is a successor ordinal $\beta+1$,  the sequence associated with 
+$\alpha$ is simply the constant sequence whose terms are equal to $\beta$.
+Likewise, we define the canonical sequence of $0$ as the 
+sequence all terms of which are equal to $0$.
+
+This convention allows us to make total the function that maps any ordinal $\alpha$ and natural number $i$ to the ordinal $\canonseq{\alpha}{i}$.
+
+\end{remark}
+
+First, let us recall how canonical sequences are defined in~\cite{KS81}.
+For efficiency's sake, we decided not to implement directly K.\&S's definitions, but to define in \gallina{} simply typed structurally recursive functions which share the abstract properties which are used in the mathematical proofs\footnote{With a small difference: the $0$-th term of the canonical sequence is not the same in our development as in~\cite{KS81}.}.
+
+\subsubsection{Mathematical definition of canonical sequences}
+
+In~\cite{KS81} the definition of $\canonseq{\alpha}{i}$ is based on the following remark:
+\begin{quote}
+Any non-zero ordinal $\alpha$ can be decomposed in a unique way as the product
+$\omega^\beta\times (\gamma+1)$.
+\end{quote}
+
+Thus the $\canonseq{\alpha}{i}$\,s are defined in terms of this decomposition:
+\begin{definition}[Canonical sequences: mathematical definition]
+\label{def:canonseq-math}
+
+\end{definition}
+\begin{mathframe}
+  \begin{itemize}
+\item Let $\lambda<\epsilon_0$ be a limit ordinal
+
+\begin{itemize}
+\item If $\lambda=\omega^{\alpha+1}\times (\beta+1)$, then 
+$\canonseq{\lambda}{i}= \omega^{\alpha+1}\times\beta +  \omega^\alpha \times i$
+\item If $\lambda=\omega^{\gamma}\times (\beta+1)$, where $\gamma<\lambda$ is a limit ordinal, then 
+$\canonseq{\lambda}{i}=\omega^{\gamma}\times \beta + \omega^{\canonseq{\gamma}{i}}$
+\end{itemize}
+
+\item For successor ordinals, we have $\canonseq{\alpha+1}{i}= \alpha$
+
+\item Finally, $\canonseq{0}{i}= \alpha$.
+\end{itemize}
+\end{mathframe}
+
+\subsubsection{Canonical sequences in Coq}
+
+Our definition may look more complex than the mathematical one, but
+uses plain structural recursion over the type \coqsimple{T1}. Thus, tactics like
+\coqsimple{cbn}, \coqsimple{simpl}, \coqsimple{compute}, etc., are applicable. For simplicity's sake, 
+we used an auxiliary function \texttt{canonS} of type \texttt{T1 -> nat  -> T1} such that
+(\texttt{canonS  $\alpha$ $i$}) is equal to $\canonseq{\alpha}{i+1}$.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Epsilon0.Canon.html\#canonS}{Epsilon0.Canon}}
+
+\label{Functions:canonS}
+\label{Functions:canon}
+\begin{Coqsrc}
+Fixpoint canonS alpha (i:nat) :=
+  match alpha with
+      zero => zero
+    | ocons zero 0 zero => zero
+    | ocons zero (S k) zero => FS k
+    | ocons gamma 0 zero =>
+      match pred gamma with
+          Some gamma' => ocons gamma' i zero
+        | None => ocons (canonS gamma i) 0 zero
+      end
+    | ocons gamma (S n) zero =>
+       match pred gamma with
+           Some gamma' => ocons gamma n (ocons gamma' i zero)
+         | None => ocons gamma n (ocons (canonS gamma i) 0 zero)
+       end
+    | ocons alpha n beta => ocons alpha n (canonS beta i)
+  end.
+\end{Coqsrc}
+
+The following function computes $\canonseq{\alpha}{i}$, except for the case $i=0$, where it simply returns $0$\;\footnote{This restriction did not prevent us from proving all the main theorems of~\cite{KS81, KP82}.
+Nevertheless, in a future version of this development, we may define $\canonseq{\alpha}{0}$ exactly as
+in~\cite{KS81}. This would  be done at the cost of making proofs much more complex.}.
+
+\begin{Coqsrc}
+ Definition canon alpha i := 
+   match i with 0 => zero | S j => canonS alpha j end. 
+\end{Coqsrc}
+
+For instance \coq's computing facilities allow us to verify the equalities\linebreak 
+\mathcolor{$\canonseq{\omega^\omega}{3} = \omega^3$} and
+\mathcolor{$\canonseq{\omega^\omega*3}{42} = \omega^\omega*2 + \omega^{42}$}.
+
+\begin{Coqsrc}
+Compute (canon (omega ^ omega) 3).
+\end{Coqsrc}
+
+\begin{Coqanswer}
+  = phi0 (FS 2) : T1
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Example canon3 :  canon (omega ^ omega) 3 = omega ^ 3.
+Proof. reflexivity. Qed.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Compute pp (canon (omega ^ omega * 3) 42).  
+\end{Coqsrc}
+
+\begin{Coqanswer}
+    = (omega ^ omega * 2 + omega ^ 42)%pT1
+     : ppT1
+\end{Coqanswer}
+
+\subsection{Basic properties of canonical sequences}
+
+We did not  try to prove that our definition really implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
+abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
+~\href{../src/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}
+
+Proving the equality $\canonseq{\alpha+1}{i}=\alpha$ is not 
+as simple as suggested by the equations of definition~\ref{def:canonseq-math}\,.
+Nevertheless, we could prove it by  structural induction on $\alpha$.
+
+\begin{Coqsrc}
+Lemma canonS_succ i alpha :
+  nf alpha ->  canonS (succ alpha) i = alpha.
+Proof.
+ induction alpha.
+ (* ... *)
+\end{Coqsrc}
+
+\subsubsection{Canonical sequences and the order $<$}
+
+\index{Maths!Transfinite induction}
+
+First, we prove by transfinite induction over $\alpha$ that $\canonseq{\alpha}{i+1}$ is an ordinal strictly less than $\alpha$ (assuming $\alpha\not=0$).
+This property allows us to use the function \texttt{canonS} and its derivates in function definitions by transfinite recursion.
+
+\label{lemma:canonS_LT}
+\begin{Coqsrc}
+Lemma canonS_LT i alpha :
+  nf alpha -> alpha <> zero -> canonS alpha i o<  alpha.
+\end{Coqsrc}
+
+\subsubsection{Limit ordinals are really limits}
+The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
+is the limit of the sequence \showmath{\canonseq{\lambda}{i}\;(1\le i)}.
+
+Note the use of \coq's \texttt{sig} type in the theorem's statement, which
+relates the boolean function \texttt{is\_limit} defined on the \texttt{T1} data-type with a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}
+
+\begin{Coqsrc}
+Lemma canonS_limit_strong (lambda : T1) : 
+     nf lambda ->
+     limitb lambda  ->
+     forall beta, beta o< lambda ->
+                  {i:nat | beta o< canonS lambda i}.
+
+Proof.
+  transfinite_induction_LT lambda.
+  (* ... *)
+Defined.
+\end{Coqsrc}
+
+\label{lemma:canonS-limit}
+
+\begin{Coqsrc}
+Lemma canonS_limit_lub (lambda : T1) :
+  nf lambda -> limitb lambda  ->
+  strict_lub (fun i => canonS lambda i) lambda.
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}\label{exo:simply-typed-canonseq}
+Instead of using the \texttt{sig} type, define a simply typed function that, given two ordinals $\alpha$ and $\beta$, returns a natural number $i$ such that, if $\alpha$ is a limit ordinal and $\beta<\alpha$, then $\beta< \canonseq{\alpha}{i+1}$.
+Of course, you will have to prove the correctness of your function.
+\end{exercise}
+
+\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0} : Paths}
+\index{Maths!Accessibility inside epsilon0}
+\label{sect:pathes-intro}
+
+Let us consider a kind of accessibility problem inside $\epsilon_0$: given two ordinals $\alpha$ and $\beta$, where $\beta<\alpha<\epsilon_0$, find a \emph{path} consisting of a finite sequence $\gamma_0=\alpha,\dots,\gamma_l=\beta$,
+where, for every $i<l$, $\gamma_i \not= 0$ \footnote{This condition allows us to ignore paths which end by a lot of useless $0$.} and there exists some strictly positive integer $s_i$
+such that $\gamma_{i+1}=\canonseq{\gamma}{s_i}$,
+
+Let $s$ be the sequence $\langle s_0,s_1,\dots, s_{l-1} \rangle$. We describe the
+existence of such a path with the notation $\alpha\xrightarrow [s]{}\beta$.
+
+For instance, we have $\omega*2 \xrightarrow[2,2,2]{}\omega$, through the 
+path $\omega\times 2, \omega+2,\omega+1,\omega$.
+
+\begin{remark}
+
+Note that, given $\alpha$ and $\beta$, where $\beta < \alpha$, the sequence $s$ which leads from $\alpha$ to $\beta$ is not unique.
+
+Indeed, if $\alpha$ is a limit ordinal, the first element of $s$ can be any integer $i$ such that $\beta<\canonseq{\alpha}{i}$, and if $\alpha$ is a successor ordinal,
+then the sequence $s$ can start with any positive integer.
+
+For instance, we have also 
+$\omega*2 \xrightarrow[3,4,5,6]{}\omega$. 
+Likewise,
+$\omega*2 \xrightarrow[{[3,14]}]{} 0$ and
+$\omega*2 \xrightarrow[3,3,3,3,3,3,3,3]{} 0$.
+\end{remark}
+
+\subsection{Formal definition}
+
+\label{path-to-definition}
+
+In \coq{}, the notion of path can be simply defined as an inductive predicate 
+parameterized by the destination $\beta$.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
+
+\begin{Coqsrc}
+(Definition transition_S i : relation T1 :=
+  fun alpha beta =>  alpha <> zero /\ beta = canonS alpha i.
+
+Definition transition i : relation T1 :=
+  match i with 0 => fun _ _ => False | S j => transition_S j end.
+\end{Coqsrc}
+
+\index{Predicates!path\_to}
+\label{sect:path-to-def}
+
+\begin{Coqsrc}
+Inductive path_to (beta: T1) : list nat -> T1 -> Prop :=
+  path_to_1 : forall (i:nat) alpha , 
+    i <> 0 ->
+    transition i alpha beta ->
+    path_to beta (i::nil) alpha
+| path_to_cons : forall i alpha s gamma,
+    i <> 0 ->
+    transition i alpha gamma ->
+    path_to beta  s gamma ->
+    path_to beta  (i::s) alpha.
+\end{Coqsrc}
+
+\begin{remark}
+In the present version of our library, we use a variant \texttt{path\_toS} of
+\texttt{path\_to}, where the proposition
+(\texttt{path\_toS $\beta$ $s$ $\alpha$}) is equivalent to
+(\texttt{path\_to $\beta$ (shift $s$) $\alpha$}). This variant is scheduled to be deprecated.
+\end{remark}
+
+\index{Exercises}
+
+\begin{exercise}
+Write a tactic for solving goals of the form (\texttt{path\_to $\beta$ $s$ $\alpha$})
+where $\alpha$, $\beta$ and $s$ are closed terms. 
+You should solve automatically the following goals:
+
+\begin{Coqsrc}
+ path_to omega (2::2::2::nil) (omega * 2).
+
+ path_to omega (3::4::5::6::nil) (omega * 2).
+
+ path_to zero (interval 3 14) (omega * 2).
+
+ path_to zero (repeat 3 8) (omega * 2).
+\end{Coqsrc}
+
+\end{exercise}
+
+\subsection{Existence of a path}
+
+\index{Maths!Transfinite induction}
+
+By transfinite induction on $\alpha$, we prove that for any $\beta<\alpha$, 
+on can build a path from $\alpha$ to $\beta$.
+
+\begin{Coqsrc}
+Lemma LT_path_to (alpha beta : T1) :
+  beta o< alpha -> {s : list nat | path_to beta s alpha}.
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}[continuation of exercise~\vref{exo:simply-typed-canonseq}]
+Define a simply typed function for computing a path from $\alpha$ to $\beta$.
+\end{exercise}
+
+% ici
+\noindent 
+From the lemma \texttt{canonS\_LT}~\vref{lemma:canonS_LT}, we can convert any path into an inequality on ordinals (by induction on paths).
+
+\begin{Coqsrc}
+Lemma path_to_LT beta s alpha :
+  path_to beta s alpha -> nf alpha -> beta o< alpha.
+\end{Coqsrc}
+
+\subsection{Paths and hydra battles}
+\label{KS-o2h}
+
+In order to apply our knowledge about  ordinal numbers (less than $\epsilon_0$) to the study of hydra battles, we define an injection
+from the interval $[0,\epsilon_0)$ into the type \texttt{Hydra}.
+
+\vspace{4pt}
+
+\emph{From Module~\href{../src/html/hydras.Hydra.O2H.html}{Hydra.O2H}}
+
+\begin{Coqsrc}
+Fixpoint iota (alpha : T1) : Hydra :=
+  match alpha with
+  | T1.zero => head
+  | ocons gamma n beta => 
+         node (hcons_mult (iota gamma) (S n) (iotas beta))
+  end 
+with iotas (alpha : T1) :  Hydrae :=
+       match alpha with
+       | T1.zero => hnil
+       | ocons alpha0 n beta  => 
+           hcons_mult (iota alpha0) (S n) (iotas beta)
+       end.
+\end{Coqsrc}
+
+For instance Fig.~\ref{fig:iota-example} shows the image by $\iota$ of the ordinal  \textcolor{black}{$\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$}
+
+  \begin{figure}[htb]
+\centering
+\begin{tikzpicture}[very thick, scale=0.3]
+\node (foot) at (10,0) {$\bullet$};
+\node (N1) at (2,2) {$\bullet$};
+\node (N2) at (10,2) {$\bullet$};
+\node (N22) at (7,2) {$\bullet$};
+\node (N3) at (14,2) {$\bullet$};
+\node (N4) at (18,2) {$\Smiley[2][green]$};
+\node (N5) at (0,4) {$\bullet$};
+\node (N6) at (2,5) {$\Smiley[2][green]$};
+\node (N7) at (4,6) {$\Smiley[2][green]$};
+\node (N88) at (7,4) {$\bullet$};
+\node (N8) at (10,4) {$\bullet$};
+\node (N9) at (14,6) {$\Smiley[2][green]$};
+\node (N10) at (0,8) {$\Smiley[2][green]$};
+\node (N11) at (10,7) {$\Smiley[2][green]$};
+\node (N111) at (7,7) {$\Smiley[2][green]$};
+\draw (foot) to [bend left=10] (N1);
+\draw (foot) -- (N2);
+\draw (foot) -- (N22);
+\draw (foot) -- (N3);
+\draw (foot) -- (N4);
+\draw (N1) to  (N5);
+\draw (N1) to   [bend left=10] (N6);
+\draw (N1) to   [bend right=20] (N7);
+\draw (N2) to  (N8);
+\draw (N22) to  (N88);
+\draw (N8) to  (N11);
+\draw (N88) to  (N111);
+\draw (N3) to  (N9);
+\draw (N5) to  (N10);
+\end{tikzpicture}
+\caption{The hydra $\iota(\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$) \label{fig:iota-example}}
+
+\end{figure}
+
+The following lemma (proved in ~\href{../src/html/hydras.Hydra.O2H.html}{Hydra.O2H.v}) maps  canonical sequences to rounds of hydra battles.
+
+\label{lemma:canonS-iota}
+
+\begin{Coqsrc}
+Lemma canonS_iota i alpha :
+    nf alpha -> alpha <> 0 ->
+    iota alpha -1-> iota (canonS alpha i).
+\end{Coqsrc}
+
+The next step of our development is to extend this relationship to
+the order $<$ on $[0,\epsilon_0)$ on one side, and hydra battles on the other side.
+
+\begin{Coqsrc}
+Lemma path_to_battle alpha s beta :
+  path_to  beta  s alpha -> nf alpha ->
+  iota alpha -+-> iota beta.
+\end{Coqsrc}
+
+As a corollary, we are now able to transform any inequality $\beta<\alpha<\epsilon_0$ into a (free) battle.
+
+\begin{Coqsrc}
+Lemma LT_to_battle alpha beta :
+    beta o< alpha ->  iota alpha -+-> iota beta.
+\end{Coqsrc}
+
+\section{A  proof of impossibility}
+
+We have got all the tools for proving there is no variant bounded by some $\mu<\epsilon_0$ for proving the termination   of all battles.
+The proof we are going to show is a proof by contradiction.
+It  can
+ be considered as a generalization of the
+proofs described in  sections~\vref{omega-case} and \vref{omega2-case}.
+
+In the module \href{../src/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
+This part of the development is parameterized by some class $B$ of battles, which will be instantiated later to \texttt{free} or \texttt{standard}.
+
+\index{Coq!Type classes}
+
+\begin{Coqsrc}
+Class BoundedVariant (B:Battle) :=
+  {
+    mu:T1 ;
+    m: Hydra -> T1;
+    mu_nf: nf mu;
+    Hvar: Hvariant T1_wf B m;
+    m_bounded: forall h, m h o< mu
+  }.
+\end{Coqsrc}
+
+Let us assume there exists such a variant:
+
+\begin{Coqsrc}
+Section Bounded.
+  Context (B: Battle) (Hy : BoundedVariant B).
+
+  Hypothesis m_decrease : forall  i h h',
+        round_n i h h' -> m h' o< m h.
+\end{Coqsrc}
+
+\begin{remark}
+  The hypothesis \texttt{m\_decrease} is false in general, but is satisfied by
+the  \texttt{free} and \texttt{standard} kinds of battles. This trick allows to 
+``factorize'' our proofs  of impossibility.
+\end{remark}
+
+\index{Maths!Transfinite induction}
+
+First, we prove that $m(\iota(\alpha))$ is always greater than or equal to $\alpha$, by  transfinite induction over $\alpha$.
+
+\begin{Coqsrc}
+Lemma m_ge_0 alpha:  nf alpha -> alpha o<= m (iota alpha).
+\end{Coqsrc}
+
+\begin{itemize}
+\item If $\alpha=0$, the inequality trivially holds
+\item If $\alpha$ is the successor of  some ordinal $\beta$, the inequality $\beta \leq m(\iota(\beta))$ holds (by induction hypothesis).
+But the hydra $\iota(\alpha)$ is transformed in one round into
+$\iota(\beta)$, thus $m(\iota(\beta))<m(\iota(\alpha))$. Hence $\beta<m(\iota(\alpha))$, which implies $\alpha \leq m(\iota(\alpha))$
+\item If $\alpha$ is a limit ordinal, then $\alpha$ is the least upper bound of the set
+of all  the $\canonseq{\alpha}{i}$.  Thus, we have just to prove that $\canonseq{\alpha}{i}< m(\iota(\alpha))$ for any $i$. 
+\begin{itemize}
+\item Let $i$ be some natural number.
+By the induction hypothesis, we have $\canonseq{\alpha}{i} \leq m(\iota(\canonseq{\alpha}{i}))$.
+But the hydra $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i})$ in one round, thus $m(\iota(\canonseq{\alpha}{i})) < m(\iota(\alpha))$, by our hypothesis \texttt{m\_decrease}.
+\end{itemize}
+\end{itemize}
+
+Please note that the impossibility proofs of 
+sections~\vref{omega-case} and \vref{omega2-case} contain a similar lemma, also called \texttt{m\_ge}.
+We are now able to build a counter-example.
+
+\begin{Coqsrc}
+  Definition big_h := iota mu.
+  Definition beta_h := m big_h.
+  Definition small_h := iota beta_h.
+\end{Coqsrc}
+
+From Lemma \texttt{m\_ge\_0} we infer the following inequality :
+
+\begin{Coqsrc}
+    Corollary m_ge_generic : m big_h o<= m small_h.
+ \end{Coqsrc}
+
+The (big) rest of the proof is dedicated to prove formally the converse inequality 
+\texttt{m small\_h o< m big\_h}.
+
+\subsection{The case of free battles}
+\label{sec:free-battles-case}
+Let us now consider that $B$ is instantiated to \texttt{free} (which means that we are considering proofs of termination of \emph{all} battles).
+The following lemmas are proved in Module~\href{../src/html/hydras.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
+The case $B=\texttt{standard}$ is studied in section~\vref{std-case}.
+
+\begin{Coqsrc}
+Section Impossibility_Proof.
+
+  Context (Var : BoundedVariant free ).
+  \end{Coqsrc}
+
+\begin{enumerate}
+\item The following lemma is an application of \texttt{m\_ge\_generic}, since \texttt{free}
+satisfies trivially the hypothesis \texttt{m\_decrease}.
+
+\begin{Coqsrc}
+Lemma m_ge : m big_h o<= m small_h.
+  Proof.
+    apply m_ge_generic.
+   (* ... *)
+\end{Coqsrc}
+
+\item From the hypothesis \texttt{m\_bounded}, we have \texttt{m big\_h o< mu}
+\item By Lemma \texttt{LT\_to\_battle}, we get a (free) battle from
+\texttt{big\_h = iota mu} to \texttt{small\_h = iota (m big\_h)}.
+
+\begin{Coqsrc}
+  Lemma  big_to_small : big_h  -+-> small_h.
+\end{Coqsrc}
+\item From the hypotheses on $m$, we infer:
+
+\begin{Coqsrc}
+Lemma m_lt : m small_h o< m big_h.
+\end{Coqsrc}
+
+\item From lemmas \texttt{m\_ge} and \texttt{m\_lt}, and the irreflexivity of $<$, we get a contradiction.
+
+  \begin{Coqsrc}
+Theorem Impossibility_free : False.
+
+End Impossibility_Proof.
+\end{Coqsrc}
+
+\end{enumerate}
+
+We have now proved there exists no bounded variant for the class of free battles.
+
+\begin{Coqsrc}
+Check Impossibility_free.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+  Impossibility_free
+     : BoundedVariant free -> False
+\end{Coqanswer}
+%%% ICI
+
+\section{The case of standard battles}
+\label{sec:standard-intro}\label{std-case}
+One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
+Module~\href{../src/html/hydras.Hydra.O2H.html}{Hydra.O2H}.
+
+\begin{Coqsrc}
+Lemma LT_to_round_plus alpha beta :
+    beta o< alpha ->  iota alpha -+-> iota beta.
+\end{Coqsrc}
+
+This lemma builds a battle out of any inequality $\beta<\alpha$. 
+It is a straightforward application of \texttt{LT\_path\_to} of
+Module~\href{../src/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}:
+
+\begin{Coqsrc}
+Lemma LT_path_to (alpha beta : T1) :
+  beta o< alpha -> {s : list nat | path_to beta s alpha}.
+\end{Coqsrc}
+
+The sequence $s$, used to build the sequence of replication factors of the battle depends on 
+$\beta$, so we cannot be sure that the generated battle is a genuine standard battle.
+
+The tool we  use comes  once again from Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
+$\alpha_0=\alpha,\alpha_1,\dots,\alpha_k=\beta$ where $\alpha_{j+1}$ is equal
+to $\canonseq{\alpha_j}{i_j}$ where $i_j$ is \emph{any} natural number, 
+we consider various constraints on these sequences.
+In particular, a path is called \emph{standard} if $i_{j+1} = i_j + 1$ for every $j<k$.
+It  corresponds to a ``segment'' of some standard battles. 
+Please note that the vocabulary on paths is ours, but all the concepts come really from~\cite{KS81}.
+
+In \coq{}, standard paths can be defined as follows.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Epsilon0.KS.html}{Epsilon0.KS}}
+
+\begin{Coqsrc}
+(**  standard path from (i, alpha) to (j, beta) *)
+
+Inductive standard_pathR(j:nat)( beta:T1):  nat -> T1 -> Prop :=
+  std_1 : forall i alpha, 
+       beta = canon alpha i -> j = S i ->
+       standard_pathR j beta i  alpha
+| std_S : forall i alpha, 
+      standard_pathR j beta (S i) (canon alpha i)  ->
+      standard_pathR j beta i alpha.
+
+Definition standard_path  i alpha j beta := 
+   standard_pathR j beta i alpha.
+\end{Coqsrc}
+
+In the mathematical text and figures, we shall use the notation 
+$\alpha \xrightarrow[i,j]{}\beta$ for the proposition 
+(\texttt{standard\_path $i$ $\alpha$ $j$ $\beta$}).
+In~\cite{KS81} the notation is
+$\alpha \xrightarrow[i]{*}\beta$
+for 
+the proposition  $\exists j, i<j \wedge \alpha \xrightarrow[i,j]{} \beta$.
+
+Our goal is now  to transform any inequality $\beta<\alpha<\epsilon_0$ into a standard path $\alpha \xrightarrow[i,j]{} \beta$ for some $i$ and $j$, then into a standard battle
+from $\iota(\alpha+i)$ to $\iota(\beta)$. 
+Following~\cite{KS81}, we proceed in two stages:
+\begin{enumerate}
+\item we simulate plain (free) paths from $\alpha$ to $\beta$ with
+paths made of steps $(\gamma,\canonseq{\gamma}{n})$, \emph{with the same $n$ all along the path}
+\item we simulate any such path by a standard path.
+\end{enumerate}
+
+\subsection{Paths with constant index}
+
+First of all, paths with a constant index 
+enjoy nice properties. They are defined as paths where all the $i_j$ are equal to the same natural number $i$, for some $i>0$.
+
+Like in~\cite{KS81}, we shall use the notation $\alpha \xrightarrow[i]{} \beta$ for denoting such a path, also called an $i$-path.
+
+\begin{Coqsrc}
+Definition const_pathS i :=
+    clos_trans_1n T1 (fun alpha beta => beta = canonS alpha i).
+
+Definition const_path i alpha beta :=
+  match i with
+    0 => False
+  | S j => const_pathS j alpha beta
+end.
+\end{Coqsrc}
+
+% Paths with a given index can be effectively computed.
+% Given $i$, $\alpha$ and $l$, the following function returns the ordinal $\beta$ such that there exists a path 
+% $\alpha \xrightarrow [i+1] {} \beta$ of length $l$.
+
+% \begin{Coqsrc}
+% Fixpoint const_funS (i:nat)(alpha : T1)(l:nat):  T1  :=
+%   match l
+%   with
+%   | 0 => alpha
+%   | S m => const_funS i (canonS i alpha) m
+%   end.
+% \end{Coqsrc}
+
+% The following computations show  applications of \texttt{constS\_fun} to the 
+% ordinal $\omega^\omega$, with various values of $i$ and $l$.
+
+% \begin{Coqsrc}
+% Compute  (const_funS 2 (omega ^omega)  55).
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+%   = zero
+%      : T1 
+% \end{Coqanswer}
+
+% \begin{Coqsrc}
+% Compute pp (const_funS 2 (omega ^omega) 15).
+% \end{Coqsrc}
+
+%   \begin{Coqanswer}
+%  = (omega ^ 2 * 2)%pT1
+%      : ppT1   
+%   \end{Coqanswer}
+
+% \begin{Coqsrc}
+% Compute pp (const_funS 4 (omega^omega)  100).
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+% = (omega ^ 4 * 4 + omega ^ 3 * 4 + omega ^ 2 + omega * 4 + 4)%pT1
+%      : ppT1
+% \end{Coqanswer}
+
+A most interesting property of $i$-paths is that we can ``upgrade'' their index, as stated by K.\&S.'s Corollary 12.
+
+\index{Maths!Transfinite induction}
+
+\begin{Coqsrc}
+Corollary Cor12 (alpha : T1) :  nf alpha ->
+         forall beta i n, beta  o< alpha  ->
+                (i < n)%nat ->
+                 const_pathS i alpha beta ->
+                 const_pathS n alpha beta.
+Proof.
+  transfinite_induction_lt alpha.
+  (* (long) proof skipped *)
+\end{Coqsrc}
+
+We shall also use a version of \texttt{Cor12} with large inequalities.
+
+\begin{Coqsrc}
+Corollary Cor12' (alpha : T1) :  nf alpha ->
+         forall beta i n, (beta o< alpha)%t1 ->
+               (i <= n)%nat ->
+               const_pathS i alpha beta ->
+               const_pathS n alpha beta.
+\end{Coqsrc}
+
+\subsubsection{Sketch of proof of \texttt{Cor12}}
+\index{Maths!Transfinite induction}
+
+We prove this lemma by transfinite induction on $\alpha$.
+Let us consider a path $\alpha \xrightarrow [i]{} \beta$ $(i>0)$. Its first step is
+the pair $(\alpha,\canonseq{\alpha}{i})$, We have $\canonseq{\alpha}{i}<\alpha$ and
+$\canonseq{\alpha}{i} \xrightarrow [i]{} \beta$. 
+Let $n$ be any natural number such that $n>i$.
+By the induction hypothesis, there exists a path $\canonseq{\alpha}{n} \xrightarrow[i]{} \beta$.
+\begin{itemize}
+\item  If $\alpha$ is a successor ordinal $\gamma+1$, then $\canonseq{\alpha}{n} =
+\canonseq{\alpha}{i}=\gamma$. Thus we have a path 
+$\alpha  \xrightarrow [n]{}  \gamma \xrightarrow [n]{} \beta$
+\item If $\alpha$ is a limit ordinal, we apply the following theorem (numbered \texttt{2.4} in Ketonen and Solovay's article).
+
+%   \begin{theorem}
+% Let $\lambda$ be a limit ordinal, then for any pair of indices $0<i<j$, there is a path $\canonseq{\lambda}{j} \xrightarrow[1]{} \canonseq{\lambda}{i}$.    
+%   \end{theorem}
+
+  \begin{Coqsrc}
+Theorem Theorem_2_4 (lambda : T1) :
+   nf lambda ->
+   limitb lambda  ->
+   forall i j, (i < j)%nat ->
+               const_pathS 0 (canonS lambda j)
+                             (canonS lambda i). 
+  \end{Coqsrc}
+
+ We build the following paths :
+
+ \begin{enumerate}
+ \item $\alpha \xrightarrow[n]{} \canonseq{\alpha}{n}$
+ \item $\canonseq{\alpha}{n} \xrightarrow[1]{} \canonseq{\alpha}{i}$ (by \texttt{Theorem\_2\_4}),
+\item $\canonseq{\alpha}{n} \xrightarrow[n]{} \canonseq{\alpha}{i}$ (applying the induction hypothesis to the preceding path);
+\item $\canonseq{\alpha}{i} \xrightarrow[n]{} \beta$ (applying the induction hypothesis)\item $\alpha \xrightarrow[n]{} \beta$ (by composition of 1, 3, and 4).
+
+ \end{enumerate}
+
+\end{itemize}
+
+\begin{remark}
+ \texttt{Cor12} ``casts'' $i$-paths into $n$-paths for any $n>i$.
+But the obtained $n$-path can be much longer than the original $i$-path.
+The following exercise will give an idea of this increase. 
+\end{remark}
+
+\index{Exercises}
+\begin{exercise}
+  Prove that  the length of the $i+1$-path from
+  $\omega^\omega$ to $\omega^i$ is $1 + (i+1)^{(i+1)}$, for any $i$. Note that the $i$-path from
+  $\omega^\omega$ to $\omega^i$ is only one step long.
+ \end{exercise}
+
+Why is \texttt{Cor12} so useful? 
+Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
+we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying collorary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
+where $n$ is the maximum of the indices $i$ met in the induction.
+
+ Lemma 1, Section 2.6 of~\cite{KS81} is naturally expressed in terms of \coq's
+\verb@sig@ construct.
+
+\label{lemma:L-2_6-1}
+\index{Coq!Techniques!Sigma types}
+
+\begin{Coqsrc}
+Lemma Lemma2_6_1 (alpha : T1) :  
+  nf alpha -> forall beta,  beta o< alpha  ->
+  {n:nat | const_pathS n alpha beta}.
+Proof.
+  transfinite_induction alpha.
+  (* rest of proof skipped *)
+\end{Coqsrc}
+
+Intuitively, lemma   \texttt{L2\_6\_1}  shows that if $\beta<\alpha<\epsilon_0$, then there exists  a battle from $\iota(\alpha)$ to $\iota(\beta)$ where the replication factor is constant, although large enough.
+
+\subsection{Casting paths with constant index into standard paths}
+
+%%% traduire la v.f.  (voir %%% A traduire %%%% )
+
+The article~\cite{KS81} contains 
+the following lemma, the proof of which is quite complex, which allows to simulate $i$-paths by $[i+1,j]$-paths, where $j$ is large enough.
+
+\begin{Coqsrc}
+(* Lemma 1 page 300 of [KS] *)
+
+Lemma constant_to_standard_path 
+  (alpha beta : T1) (i : nat):
+  nf alpha -> const_pathS i alpha beta -> zero  o< alpha ->
+  {l:nat | standard_path (S i) alpha j beta}.
+\end{Coqsrc}
+
+\subsubsection{Sketch of proof of \texttt{constant\_to\_standard\_path}}
+
+Our proof follows the proof by Ketonen and Solovay, including its organization as a sequence of lemma.  Since it is a non-trivial proof, we will comment its main steps below.
+
+\subsubsection*{Préliminaries}
+
+Please note that, given an ordinal $\alpha:\texttt{T1}$, and two natural numbers $i$ and $l$, there exists at most a standard path $\alpha \xrightarrow [i,i+l]{*} \beta$.
+The following function computes $\beta$ from $\alpha$, $i$ and $l$.
+
+\begin{Coqsrc}
+Fixpoint standard_gnaw (i:nat)(alpha : T1)(l:nat):  T1  :=
+  match l with
+  | 0 => alpha
+  | S m => standard_gnaw (S i) (canon alpha i) m
+  end.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+  Compute standard_gnaw 2 omega 15.
+(*   = zero
+     : T1 *)
+Compute pp (standard_gnaw 2 (omega^omega)  10).
+(*
+= (omega + 7)%pT1
+     : ppT1
+*)
+Compute pp (standard_gnaw 4 (omega^omega)  100).
+(*
+ = (omega ^ 3 * 4 + omega ^ 2 * 5 + omega * 3 + 39)%pT1
+     : ppT1 *)
+\end{Coqsrc}
+
+\index{Maths!Transfinite induction}
+
+By transfinite induction over  $\alpha$, we prove that the ordinal $0$ is reachable from any ordinal $\alpha<\epsilon_0$ by some standard path.
+
+\begin{Coqsrc}
+Lemma standard_path_to_zero :
+  forall  alpha i, nf alpha ->
+                   {j: nat | standard_path (S i) alpha j zero}.
+\end{Coqsrc}
+
+\paragraph*{}
+Let us consider two ordinals  $\beta<\alpha<\epsilon_0$.  Let $p$  be some $(n+1)$-path from $\alpha$ to $\beta$.
+
+\begin{Coqsrc}
+ Section Constant_to_standard_Proof.
+
+  Variables (alpha beta: T1) (n : nat).
+  Hypotheses (Halpha: nf alpha) (Hpos : zero o<  beta)
+             (p : const_pathS n alpha  beta).
+\end{Coqsrc}
+
+Applying \texttt{standard\_path\_to\_zero}, $0$ is reachable from $\alpha$ by some standard path  (see figure~\vref{fig:belle-preuve-1}).
+
+\begin{figure}[h]
+  \centering
+
+\begin{tikzpicture}[very thick, scale=0.25]
+\node (alpha) at (0,0) {$\alpha$};
+    \node (beta) at (32, 0){$\beta$};
+
+  \draw[->, very thick,blue] (alpha)-- node [below]{$n+1$} node [above] {$+$} (beta);
+
+  \node (alpha1) at (5,5) {};
+  \node (alpha2) at (13,5) {};
+    \node (alpha3) at (20,5) {};
+  \node (alphalast) at (35,5) {};
+  \node (zero) at (45,0) {$0$};
+  \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate=40]{$n+1$}  (alpha1);
+  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
+   \draw [->, dashed,very thick,blue] (alpha2)-- node [below]{$n+3$}  (alpha3);
+
+  \node (dots) at (24,5) {$\dots$};
+  \draw [->, dashed, very thick,blue] (alphalast)-- node [below, rotate=-26]{$n+p+1$}  (zero);
+
+\end{tikzpicture}
+\caption{A nice proof (1)}
+  \label{fig:belle-preuve-1}
+\end{figure}
+
+\paragraph*{}
+
+Since comparison on \texttt{T1} is decidable, one can compute the last step $\gamma$ of the standard path from $(\alpha,n+1)$  such that $\beta\leq \gamma$.
+Let $l$ be the length of the path from $\alpha$ to $\gamma$.  
+This step of the proof is illustrated in figure~\vref{fig:belle-preuve-2}.
+
+\begin{figure}[h]
+  \centering
+
+\begin{tikzpicture}[very thick, scale=0.25]
+\node (alpha) at (0,0) {$\alpha$};
+    \node (beta) at (32, 0){$\beta$};
+
+  \node (alpha1) at (5,5) {};
+  \node (alpha2) at (13,5) {};
+  \node (dots) at (17,5) {$\ldots$};
+    \node (alpha3) at (20,5) {};
+    \node (gamma) at (24,0) {$\gamma$};
+    \node (delta) at (38,0) {$\delta$};
+    \draw [->, dashed,very thick,blue] (alpha)-- node [below,rotate=35]{$n+1$}  (alpha1);
+  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
+   \draw [->, dashed,very thick,blue] (alpha3)-- node [below,rotate = -48]{\tiny $n+l$}  (gamma);
+   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{$n+l+1$} (delta);
+   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{$n+1$} node [above] {$+$} (beta);
+   \draw[thick] (alpha)--  (gamma);
+   \draw[thick] (gamma)--  node [above] {$\geq$} (beta);
+    \draw[thick] (beta)--  node [above] {$>$} (delta);
+\end{tikzpicture}
+
+\caption{A nice proof (2)}
+  \label{fig:belle-preuve-2}
+\end{figure}
+
+\paragraph*{}
+
+\begin{itemize}
+\item If $\beta=\gamma$, its OK! We have got a standard path
+from  
+$\alpha$ to $\beta$ with successive indices  $n+1, n+2, \dots, n+l+1$
+
+\item Otherwise,  $\beta < \gamma$.  Let us consider  $\delta=\canonseq{\gamma}{n+l+1}$.
+By applying several times lemma \texttt{Cor12},  one converts  every path of Fig~\ref{fig:belle-preuve-2} into
+ a $n+l+1$-path  (see figure~\ref{fig:belle-preuve-3}).
+
+But $\gamma$ is on the $n+l+1$-path from $\alpha$ to $\beta$.
+As shown by figure~\vref{fig:fin-belle-preuve}, the ordinal $\delta$, reachable from
+$\gamma$ in one single step,  must be greater than or equal to $\beta$, which contradicts our  hypothesis $\beta < \gamma$.
+
+\begin{figure}[h]
+  \centering
+
+\begin{tikzpicture}[very thick, scale=0.25]
+\node (alpha) at (0,0) {$\alpha$};
+    \node (beta) at (32, 0){$\beta$};
+    \node (alpha1) at (5,5) {};
+  \node (alpha2) at (13,5) {};
+  \node (dots) at (17,5) {$\ldots$};
+    \node (alpha3) at (20,5) {};
+    \node (gamma) at (24,0) {$\gamma$};
+    \node (delta) at (38,0) {$\delta$};
+     \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
+  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
+   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -48]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
+   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\tiny $n+l+1$} node [above]{\color{red} $1$} (delta);
+   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
+    \draw[thick] (gamma)--   node [above]{\color{red} $>$}(beta);
+   \draw[thick] (alpha)--  (gamma);
+
+    \draw[thick] (beta)--  node [above] {$>$} (delta);
+
+\end{tikzpicture}
+
+\caption{A nice proof (3)}
+  \label{fig:belle-preuve-3}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+\begin{tikzpicture}[very thick, scale=0.25]
+\node (alpha) at (0,0) {$\alpha$};
+    \node (beta) at (32, 0){$\beta$};
+
+  \node (alpha1) at (5,5) {};
+  \node (alpha2) at (13,5) {};
+  \node (dots) at (15,5) {$\ldots$};
+    \node (alpha3) at (18,5) {};
+    \node (gamma) at (24,0) {$\gamma$};
+    \node (delta) at (42,0) {$\delta$};
+    \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
+  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
+   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -36]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
+   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\small $n+l+1$} node [above]{\color{red} $1$} (delta);
+   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
+   \draw[thick] (alpha)--  (gamma);
+
+    \draw[thick] (gamma)--  node [below]{\tiny $n+l+1$} node [above]{\color{red} $+$}(beta);
+    \draw[thick] (beta)--  node [above] {$>$} (delta);
+
+\end{tikzpicture}
+
+\caption{A nice proof (4)}
+  \label{fig:fin-belle-preuve}
+\end{figure}
+
+\end{itemize}
+ The only possible case is  thus $\beta=\gamma$, so we have got a standard path  from $\alpha$ to $\beta$.
+
+\begin{Coqsrc}
+ Lemma constant_to_standard_0 : 
+    {l : nat | standard_fun (S n) alpha l = beta}.
+
+End Constant_to_standard_Proof.
+\end{Coqsrc}
+
+Here is the full statement of the conversion from constant to standard paths.
+
+\begin{Coqsrc}
+Lemma constant_to_standard_path 
+  (alpha beta : T1) (i : nat):
+  nf alpha -> const_pathS i alpha beta -> zero  o< alpha ->
+  {j:nat | standard_path (S i) alpha j beta}.
+\end{Coqsrc}
+
+Applying \texttt{Lemma2\_6\_1} and \texttt{constant\_to\_standard\_path}, we get the following corollary.
+
+\begin{Coqsrc}
+Corollary  LT_to_standard_path  (alpha beta : T1) :
+  beta o< alpha ->
+  {n : nat & {j:nat | standard_path (S n) alpha j beta}}.
+\end{Coqsrc}
+
+\subsection{Back to hydras}
+\label{sec:standard-battles-cases}
+We are now able to complete our proof that there exists no bounded variant for proving the termination of standard hydra battles. The proof we are going to comment can
+be consulted in the module 
+\url{../src/html/hydras.Hydra.Epsilon0_Needed_Std.html}.
+Please note that it has the same global structure as in section\ref{sec:free-battles-case} 
+% ICI !
+Applying the  lemmas  \texttt{Lemma2\_6\_1} of the module 
+\href{../src/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
+{\texttt{Lemma2\_6\_1}}   and 
+\href{../src/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
+{\texttt{constant\_to\_standard\_path}},
+we can convert any inequality $\beta<\alpha<\epsilon_0$ into a standard path from
+$\alpha$ to  $\beta$, then into a fragment of a standard battle from 
+$\iota(\alpha)$ to $\iota(\beta)$.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Hydra.Epsilon0_Needed_Std.html\#LT_to_standard_battle}{Hydra.Epsilon0\_Needed\_Std}}
+
+\begin{Coqsrc}
+Lemma LT_to_standard_battle :
+    forall alpha beta,
+      beta o< alpha ->
+      exists n i,  battle standard  n (iota alpha) i (iota beta).
+\end{Coqsrc}
+
+Next, please consider the following context:
+
+\begin{Coqsrc}
+Section Impossibility_Proof.
+
+  Context (Var : BoundedVariant standard).
+ \end{Coqsrc}
+
+In the same way as for free battles, we import a large inequality 
+from 
+the module \href{../src/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}.
+
+\begin{Coqsrc}
+ Lemma m_ge : m big_h o<= m small_h.
+\end{Coqsrc}
+
+\paragraph*{} If remains to prove the following strict inequality, in order to have a contradiction.
+
+\begin{Coqsrc}
+Lemma m_lt : m small_h  o< m big_h.
+\end{Coqsrc}
+
+\paragraph*{Sketch of proof:} Let us recall that $\texttt{big\_h} = \iota(\mu)$
+ and $\texttt{small\_h} = \iota (m (\texttt{big\_h}))$.
+
+Since $m(\texttt{big\_h})< \mu$, there exists a standard path from $\mu$ to
+$m(\texttt{big\_h})$, hence a   standard battle from $\iota(\mu)$  to
+$\iota(m(\texttt{big\_h}))$,  i.e. from \texttt{big\_h} to \texttt{small\_h}.
+
+Since $m$ is assumed to be a variant for standard battles, we get the inequality  $m(\texttt{small\_h}) < m(\texttt{big\_h})$.
+
+\subsection{Remarks}
+
+We are grateful to 
+ J. Ketonen and R. Solovay  for the high quality of their explanations and proof details.
+Our proof follows tightly the sequence of lemmas in their article, with a focus on 
+constructive aspects.
+Roughly steaking, our implementation \emph{builds}, out of a hypothetic 
+  variant $m$, bounded by some ordinal $\mu<\epsilon_0$, a hydra \texttt{big\_h} which verifies the impossible inequality  $m(\texttt{big\_h})< m(\texttt{big\_h})$.
+
+On may ask whether the preceding results are not too restrictive, since they 
+refer to a particular data type \texttt{T1}.
+In fact, our representation of ordinals strictly less than 
+ $\epsilon_0$ is faithful to their mathematical definition, at least 
+Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
+(please see also \href{../src/html/hydras.Schutte.Correctness_E0.html}{the module \texttt{Ordinals.Schutte.Correctness\_E0}}).
+
+Thus, we can infer that our theorems can be applied to any well order.
+
+\index{Projects}
+\begin{project}
+Study a possible modification of the definition of a variant  (for  standard battles).
+
+\begin{itemize}
+\item The variant is assumed to be strictly decreasing \emph{on configurations 
+reachable from some initial configuration where the replication factor is equal to $0$}
+\item The variant may depend on the number of the current round.
+\end{itemize}
+
+In other words, its type should be \texttt{nat -> Hydra -> T1}, and it must 
+verify the inequality $m\, (S\,i)\, h' < m\,i\, h$ whenever the configuration 
+$(i,h)$ is reachable from some initial configuration $(0,h_0)$
+and \texttt{h} is transformed into \texttt{h'} in the considered round.
+Can we still prove the theorems of section~\ref{std-case} with this new definition?
+
+\end{project}
+
+%---------------------------------------------------------------------
+\chapter{Large sets and rapidly growing functions}\label{chap:alpha-large}
+
+%\section{Introduction}
+
+In this chapter, we try to feel how long a standard battle can be.
+To be precise, for any ordinal $\alpha<\epsilon0$ and any positive integer $k$,
+we give a minoration of the number of steps of a standard battle which
+starts with the hydra $\iota(\alpha)$ and the replication factor $k$.
+
+We express this number in terms of the Hardy hierarchy of fast-growing 
+functions~\cite{BW85, Wainer1970, KS81, Promel2013}.
+ From the \coq{} user's point of view, such  functions are  very 
+attractive:  they are defined as functions  in \gallina{}, and we can apply them \emph{in theory}, but they are so complex that you will never be able to look at the result of the computation.
+ Thus, our knowledge on these functions must rely on \emph{proofs}. In our development, we use often the rewriting rules generated by \coq's \texttt{Equations} plug-in.
+
+\section{Large  sequences}
+
+\subsection{Definition}
+
+\begin{definition}
+Let $0<\alpha<\epsilon_0$ be any ordinal, and $s$ a sequence of strictly positive natural numbers. 
+We say that $s$ is \emph{minimally $\alpha$-large} (in short:
+\emph{$\alpha$-mlarge}) if $s$ if $s$ is $\alpha$-large 
+ and every strict prefix of $s$ leads to a non-zero ordinal (\emph{cf} Sect.~\vref{sect:path-to-def}).
+
+\index{Maths!Large sequences}
+\index{Maths!Minimal large sequences}
+
+\end{definition}
+
+\vspace{4pt}
+
+\noindent
+\emph{From Module~ \href{../src/html/hydras.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}
+
+\index{Predicates!mlarge @ mlarge (minimal large sequences)}
+\begin{Coqsrc}
+Definition mlarge alpha (s:list nat) := path_to zero s alpha.
+\end{Coqsrc}
+
+\begin{remark}
+  Ketonen and Solovay~\cite{KS81} consider  large finite \emph{sets} of natural numbers,  but they are mainly used as sequences. Thus, we chosed to represent them explicitely as (sorted) lists.
+
+They also consider large (but not minimally large) sets. They can be defined in
+\coq{} as follows:
+
+\noindent
+\emph{From Module~\href{../src/html/hydras.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}
+
+\begin{Coqsrc}
+Fixpoint gnaw (alpha : T1) (s: list nat) :=
+  match s with
+    | nil => alpha
+    | (0::s') => gnaw  alpha s'
+    | (S i :: s')  =>  gnaw (canonS i alpha) s'
+  end.
+
+Definition large alpha (s:list nat) := gnaw alpha s = zero.
+\end{Coqsrc}
+\end{remark}
+
+Let us consider two integers $k$ and $l$, such that $0<k<l$. In order to check whether the interval $[k,l]$ is minimally large for $\alpha$, it is enough to
+follow from $\alpha$ the path associated with the interval $[k,l($ and verify that the last ordinal we obtain is equal to $1$.
+
+\subsection{Example}
+
+For instance the interval $[6,70]$ leads $\omega^2$ to $\omega\times 2 + 56$. Thus this interval is not $\omega^2$-mlarge.
+
+\begin{Coqsrc}
+Compute pp (gnaw (omega * omega) (interval 6 70)).
+\end{Coqsrc}
+
+\begin{Coqanswer}
+ = (omega * 2 + 56)%pT1
+     : ppT1
+\end{Coqanswer}
+
+Let us try another computation.
+
+\begin{Coqsrc}
+Compute (gnaw (omega * omega) (interval 6 700)).
+\end{Coqsrc}
+
+\begin{Coqanswer}
+ = zero : T1
+\end{Coqanswer}
+
+We may say that the interval $[6,700]$ is $\omega^2$-large, since it leads to $0$, but nothing assures us that the condition of minimality is satisfied.
+
+The following lemma relates minimal largeness with the function 
+\texttt{gnaw}.
+
+\begin{Coqsrc}
+Lemma mlarge_iff alpha x (s:list nat) :
+  s <> nil -> ~ In 0 (x::s) ->
+  mlarge alpha (x::s) <-> gnaw alpha (but_last x s) = one.
+ \end{Coqsrc}
+
+For instance, we can verify that the interval $[6,510]$ is $\omega^2$-mlarge.
+
+ \noindent
+\emph{From Module~\href{../src/html/hydras.Epsilon0.Large_Sets_Examples.html}{Epsilon0.Large\_Sets\_Examples}}
+\begin{Coqsrc}
+Example Ex1 : mlarge (omega * omega) (interval 6 510).
+\end{Coqsrc}
+
+\section{Computing mlarge sequences}
+
+Now, consider any natural number $k>0$. We would like to compute
+a number $l$ such that the interval $[k,l]$ is $\alpha$-mlarge. So, 
+the standard battle starting with $\iota(\alpha)$ and the replication factor $k$ will end after $(l-k+1)$ steps.
+
+First, we notice that this  number $l$ exists, since the segment $[0,\epsilon_0)$ is well-founded and $\canonseq{\alpha}{i}<\alpha$ for any $i$ and $\alpha>0$.
+Moreover, it is unique:
+
+\vspace{4pt}
+\noindent
+\emph{From Module~\href{../src/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
+\begin{Coqsrc}
+Lemma mlarge_unicity alpha k l l' : 
+  mlarge alpha (interval (S k) l) ->
+  mlarge alpha (interval (S k) l') ->
+  l = l'.
+\end{Coqsrc}
+
+Thus, it seems obvious that there must exist a function, parameterized by $\alpha$ which associates to any  strictly positive integer $k$ the number $l$ such that
+the interval $[k,l]$ is $\alpha$-mlarge. It would be fine to write in \gallina{} a definition like this:
+
+\begin{Coqbad}
+Function L (alpha: E0) (i:nat) :  nat := ...
+\end{Coqbad}
+
+But we do not know how to fill the dots yet \dots{}   In the next section, we will 
+use \coq{} to reason  about the \emph{specification} of \texttt{L},
+prove properties of any function which satisfies this specification.
+In Sect.~\ref{sect:L-equations}, we use the \texttt{coq-equations} plug-in
+to define a function \texttt{L\_}, and prove its correctness w.r.t. its specification.
+
+\subsection{Looking for a function definition}
+
+Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider the functions which associate to each stritly positive integer $k$ the number $l$, where 
+the interval $[k,l)$ is $\alpha$-mlarge.
+
+\begin{remark}
+The upper bound of the considered interval has been chosen to be  $l-1$ and not $l$, in order to simplify some statements and proofs in composition lemmas associated to ordinals of the form $\alpha\times i$ and
+$\omega^\alpha\times i + \beta$.
+In order to consider any ordinal below $\epsilon_0$, we consider a special case for $\alpha=0$.
+\end{remark}
+
+\subsection{Formal specification}
+
+Our specification of the function \texttt{L} is as follows:
+
+\begin{Coqsrc}
+Inductive L_spec : T1 -> (nat -> nat) -> Prop :=
+  L_spec0 :
+    forall f, (forall k, f k = k) ->  L_spec zero f
+| L_spec1 : forall alpha f,
+    alpha <> zero ->
+    (forall k, mlarge alpha (interval (S k) (Nat.pred (f (S k))))) ->
+    L_spec alpha f.
+\end{Coqsrc}
+
+Note that, for $\alpha\not=0$, the value of $f(0)$ is not specified.
+Nevertheless, the restriction of $f$ to the set of strictly positive integers is unique (up to extensionnality).
+
+\begin{Coqsrc}
+Lemma L_spec_unicity alpha f g :
+  L_spec alpha f -> L_spec alpha g ->
+  forall k, f (S k) = g (S k).
+\end{Coqsrc}
+
+\subsection{Properties derived from \texttt{L\_spec}}
+
+Let us now prove properties of any function $f$ (if any) which satisfies 
+\texttt{L\_spec}.
+We are looking for properties which could be used for writing \emph{equations} and prove the correctness of the function generated by the \texttt{coq-equations} plugin.
+Moreover, they will give us some examples of
+$L_\alpha$ for small values of $\alpha$.
+
+Our exploration of the $L_\alpha$s  follows the usual scheme : transfinite induction, and proof-by-cases : zero, successors and limit ordinals.
+
+\index{Maths!Transfinite induction}
+
+\subsubsection{The  ordinal zero}
+\label{sect:L-spec-zero}
+The base case is directly a consequence of the specification.
+
+\begin{Coqsrc}
+Lemma L_zero_inv f : L_spec zero f -> forall k, f (S k) = S k.
+\end{Coqsrc}
+
+\subsubsection{Successor ordinals}
+\label{sect:L-spec-succ}
+Let $\beta$ be some ordinal, and assume the arithmetic function $f$ satisfies 
+the specification $(\texttt{L\_spec}\;\beta)$.  Let $k$ be any natural number.
+Any path from $\texttt{succ}\,\beta$ to $0$ starting at $k+1$ can be decomposed into a first step from $\texttt{succ}\,\beta$ to $\beta$, then a path from
+$\beta$ at $k+2$ to $0$. 
+By hypothesis the interval $[k+2, f(k+2)-1]$ is $\beta$-mlarge.
+But the interval $[k+1, f(k+2)-1]$ is the concatenation of the singleton
+$\{k+1\}$ and the interval $[k+2, f(k+2)-1]$.
+So, the function $\lambda\,k.\,f(k+1)$ satisfies the specification $\texttt{L\_spec}\,\beta$.
+
+Note that our decomposition of intervals works only if the intervals we consider are not empty.
+In order to ensure this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <<= f}, or \texttt{(fun\_le S f)} (defined
+in~\href{../src/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}).
+
+\begin{Coqsrc}
+Definition fun_le f g  := forall n:nat,  f n <=  g n.
+\end{Coqsrc}
+
+It looks also natural to show that the functions we consider are strictly monotonous. The section on successor ordinals has thus the following structure.
+
+\begin{Coqsrc}
+Section succ.
+   Variables (beta : T1) (f : nat -> nat).
+
+   Hypotheses (Hbeta : nf beta)
+              (f_mono : strict_mono f)
+              (f_Sle : S <<= f)
+              (f_ok : L_spec beta f).
+
+   Definition L_succ := fun k => f (S k).
+
+   Lemma L_succ_mono : strict_mono L_succ.
+
+   Lemma L_succ_Sle : S <<= L_succ.
+
+   Lemma L_succ_ok : L_spec (succ beta) L_succ.
+
+End succ.
+
+\end{Coqsrc}
+
+\subsubsection{Limit ordinals}
+\label{sect:L-spec-lim}
+
+Let $\lambda<\epsilon_0$ be any limit ordinal. In a similar way as for successors, we decompose any path from $\lambda$ (at $k$) into a step to
+$\canonseq{\lambda}{k}$, then to $0$. In the following section, we assume that there exists à correct function for $\canonseq{\lambda}{k}$, \emph{for any strictly positive $k$}.
+
+\begin{Coqsrc}
+Section lim.
+  Variables (lambda : T1)
+            (Hnf : nf lambda)
+            (Hlim : limitb lambda)
+            (f : nat -> nat -> nat)
+            (H : forall k, L_spec (canonS lambda k) (f (S k))).
+
+  Let L_lim k := f k (S k).
+
+  Lemma L_lim_ok : L_spec lambda L_lim.
+
+End lim.
+\end{Coqsrc}
+
+\subsection{First results}
+
+Applying the previous lemmas on successors and limit ordinals, 
+we get several correct implementations of \texttt{(L\_spec $\alpha$)} for small values of $\alpha$.
+
+\subsubsection{Finite ordinals}
+
+By iterating the functional \texttt{L\_succ}, we get a realization of
+\texttt{(L\_spec (fin $i$))} for any natural number $i$.
+
+\begin{Coqsrc}
+Definition L_fin i := fun k => (i + k)%nat.
+
+Lemma L_fin_ok i : L_spec (fin i) (L_fin i).
+\end{Coqsrc}
+
+\subsubsection{The first limit ordinal \texorpdfstring{$\omega$}{omega}}
+
+The lemmas \texttt{L\_fin\_ok} and \texttt{L\_lim\_ok}   allow us to get 
+by diagonalization a correct implementation for 
+\texttt{L\_spec omega}.
+
+\begin{Coqsrc}
+Definition L_omega k := S (2 * k)%nat.
+
+Lemma L_omega_ok : L_spec omega L_omega.
+\end{Coqsrc}
+
+\subsubsection{Towards  \texorpdfstring{$\omega^2$}{omega*omega}}
+
+We would like to get exact formulas for the ordinal $\omega^2$, a.k.a.
+$\varphi_0(2)$. This ordinal is the limit of the sequence $\omega\times i\;(i \in \mathbb{N}$. Thus, we have to study ordinals of this form, then use 
+our lemma on limits.
+
+The following lemma establishes a path from $\omega\times ( i+1)$ to
+$\omega \times i$.
+
+\begin{Coqsrc}
+Lemma path_to_omega_mult (i k:nat) :
+  path_to (omega * i)
+                (interval (S k) (2 * (S k)))
+                (omega * (S i)).
+\end{Coqsrc}
+
+Let us consider a path from  $\omega\times(i+1)$ to $0$ starting at $k+1$.
+A first ``big step'' will lead to $\omega\times i$ at $2(k+1)$. If $i>0$, the
+next jump leads to $\omega\times(i-1)$ at $2(2(k+1))+1$, etc.
+
+The following lemma expresses the length of the mlarge sequences associated with finite multiples of $\omega$.
+
+\begin{Coqsrc}
+Lemma omega_mult_mlarge_0 i  : forall k,
+    mlarge  (omega * (S i))
+            (interval (S k)
+                      (Nat.pred (iterate (fun p =>  S (2 * p)%nat)
+                                         (S i)
+                                         (S k)))).
+\end{Coqsrc}
+
+Thus, we infer the following result:
+
+\emph{From Module~ \href{../src/html/hydras.Epsilon0.Large_Sets.html\#L_omega_mult}{Epsilon0.Large\_Sets}}
+
+\begin{Coqsrc}
+Definition L_omega_mult i (x:nat) :=  iterate L_omega i x.
+
+Lemma L_omega_mult_ok (i: nat) :  L_spec (omega * i) (L_omega_mult i).
+\end{Coqsrc}
+
+For instance, let us consider the ordinal $\omega\times 8$, and a sequence 
+starting at $k=5$.
+
+\begin{Coqsrc}
+Compute L_omega_mult 8 5.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+= 1535
+     : nat
+\end{Coqanswer}
+
+More generally, we prove the equality $L_{\omega\times i}(k)=2^i\times(k+1)-1$.
+
+\begin{Coqsrc}
+Lemma L_omega_mult_eqn (i : nat) :
+  forall (k : nat),  (0 < k)%nat  ->
+                     L_omega_mult i k = (exp2 i * S k - 1)%nat.
+\end{Coqsrc}
+
+By diagonalization, we obtain a simple formula for $L_{\omega^2}$.
+
+\begin{Coqsrc}
+Definition L_omega_square k := iterate (fun z => S (2 * z)%nat)
+                                        k
+                                        (S k).
+
+Lemma L_omega_square_eqn k :
+  (0 < k)%nat ->
+  L_omega_square k = (exp2 k * (k + 2) - 1)%nat.
+
+Lemma L_omega_square_ok: L_spec (omega * omega) 
+          L_omega_square.
+
+Compute L_omega_square 8.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+ = 2559
+     : nat
+\end{Coqanswer}
+
+%%%% ICI
+
+\subsubsection{Going further}
+Let us consider a last example, ``computing'' $L_{\omega^3}$.
+Since the canonical sequence associated with this ordinal is composed of the
+$\omega^2\times i\;(i\in\mathbb{N}_1)$, we have to study this sequence.
+
+To this end, we prove a generic lemma, which expresses $L_{\omega^\alpha\times i}$ in terms of $L_{\omega^\alpha}$.
+Note that in this correction lemma, we assume that the fonction associated with $\alpha$ is stritly monotonous and
+greater or equal than the successor function.
+
+\begin{Coqsrc}
+Section phi0_mult.
+ Variables (alpha : T1) (f : nat -> nat).
+ Hypotheses (Halpha : nf alpha)
+            (f_mono : strict_mono f)
+            (f_Sle : S <<= f)
+            (f_ok : L_spec (phi0 alpha) f).
+
+ Definition L_phi0_mult i := iterate f i.
+
+Lemma L_phi0_mult_ok i: 
+  L_spec (ocons alpha i zero)  (L_phi0_mult (S i)).
+
+ Lemma L_phi0_mult_smono i: strict_mono (L_phi0_mult i).
+
+ Lemma L_phi0_mult_Sle i: S <<=  L_phi0_mult (S i).
+
+End phi0_mult.
+\end{Coqsrc}
+
+Let us look 
+at the ordinal $\omega^2\times i$, using \texttt{L\_phi0\_mult}
+
+\begin{Coqsrc}
+ Definition L_omega_square_times i :=  iterate L_omega_square i.
+
+ Lemma L_omega_square_times_ok i : 
+    L_spec (ocons 2 i zero) (L_omega_square_times (S i)).
+ Proof.
+  apply L_phi0_mult_ok.
+  -  auto with T1.
+  -  apply L_omega_square_Sle.
+  -  apply L_omega_square_ok.
+ Qed.
+\end{Coqsrc}
+
+We are now ready to get an exact formula for $L_{\omega^3}$. 
+\begin{Coqsrc}
+Definition L_omega_cube  := L_lim  L_omega_square_times .
+
+Lemma L_omega_cube_ok : L_spec (phi0 3) L_omega_cube.
+\end{Coqsrc}
+
+The function  $L_{\omega^3}$  is just obtained by diagonalization upon $L_{\omega^2\times i}$.
+
+\begin{Coqsrc}
+Lemma L_omega_cube_eqn i : L_omega_cube i = L_omega_square_times i (S i).
+Proof.   reflexivity. Qed.
+\end{Coqsrc}
+
+Thus, for instance, $L_{\omega^3}(3)=L_{\omega^2\times 4}(3)$.
+So, we obtain an exact expression of this number.
+
+\begin{Coqsrc}
+Lemma L_omega_cube_3_eq:
+   let N := exp2 95 in
+   let P := (N * 97 - 1)%nat in
+    L_omega_cube 3  =  (exp2 P * (P + 2) - 1)%nat.
+\end{Coqsrc}
+
+This number is quite big. Using \texttt{Ocaml}'s \texttt{float} arithmetic,
+we can [under-]approximate it by $2^{3.8\times10^{30}}\times 3.8\times{10^{30}}$.
+
+\begin{Coqsrc}
+# let exp2 x = 2.0 ** x;;
+
+val exp2 : float -> float = <fun>
+#   exp2 95.0 *. 97.0 -. 1.0;;
+- : float = 3.84256588194182037e+30
+# let n = exp2 95.0 ;;
+# let p = n *. 97.0 -. 1.0;;
+val p : float = 3.84256588194182037e+30
+
+Estimation :
+2 ** (3.84 e+30) * 3.84 e+30.
+\end{Coqsrc}
+
+\section{Using \texttt{Equations}}
+\label{sect:L-equations}
+
+Note that we did not define any function $L_\alpha$ \emph{for any $\alpha<\epsilon_0$} yet.
+We have got no more than a collection of proved realizations of $\texttt{L\_spec}\;\alpha$ for several values of $\alpha$.
+
+\index{Coq!Plug-ins!Equations}
+
+Using the \texttt{coq-equations} plug-in by 
+M. Sozeau~\cite{sozeau:hal-01671777}, we will now define a function \texttt{L\_} which maps any ordinal  $\alpha<\epsilon_0$ to a proven realization of 
+$\texttt{L\_spec}\;\alpha$.   
+\subsection{Definition}
+
+In order to get a total function, we use our type \texttt{E0} of well-formed ordinal terms,(see Sect~\vref{sect:E0-def}).  Our definition is structured along a well-founded recursion and
+a case-study (zero, limit and successor ordinals).
+
+\vspace{4pt}XS
+\noindent
+\emph{From Module~\href{../src/html/hydras.Epsilon0.L_alpha.html\#L_}{L\_alpha}).}
+
+\label{Functions:L-alpha}
+
+\begin{Coqsrc}
+From Equations Require Import Equations.
+Require Import ArithRing Lia.
+
+Instance Olt : WellFounded Lt := Lt_wf.
+Hint Resolve Olt : E0.
+
+(** Using Coq-Equations for building a function which satisfies 
+    Large_sets.L_spec *)
+
+Equations  L_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
+  L_ alpha  i with E0_eq_dec alpha Zero :=
+    { | left _ =>  i ;
+      | right nonzero
+          with Utils.dec (Limitb alpha) :=
+          { | left _ =>  L_ (Canon alpha i)  (S i) ;
+            | right notlimit =>  L_ (Pred alpha) (S i)}}.
+
+Solve All Obligations with auto with E0.
+\end{Coqsrc}
+
+It is worth looking at the answer from \texttt{Equations} and check (with \texttt{About} ) all the lemmas this plug-in gives you for free. We show here only a part of \coq's anwer.
+
+\begin{Coqanswer}
+L__obligations_obligation_1 is defined
+L__obligations_obligation_2 is defined
+L__obligations is defined
+L__clause_1 is defined
+L__functional is defined
+L_ is defined
+...
+L__equation_1 is defined
+L__graph_mut is defined
+L__graph_rect is recursively defined
+L__graph_correct is defined
+L__elim has type-checked, generating 1 obligation
+L__elim is defined
+FunctionalElimination_L_ is defined
+FunctionalInduction_L_ is defined
+\end{Coqanswer}
+
+Sometimes, these automatically generated statements may look cryptic.
+
+\begin{Coqsrc}
+About L__equation_1.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+L__equation_1 :
+forall (alpha : E0) (i : nat),
+L_ alpha i = L__unfold_clause_1 alpha (E0_eq_dec alpha Zero) i
+\end{Coqanswer}
+
+In most cases, it may be useful to write human-readable  paraphrases of these statements.
+
+\begin{Coqsrc}
+Lemma L_zero_eqn : forall i, L_ Zero i = i.
+Proof. intro i; now rewrite L__equation_1. Qed.
+
+Lemma L_lim_eqn alpha i : Limitb alpha -> L_ alpha i =
+                                        L_ (Canon alpha i) (S i).
+
+Lemma L_succ_eqn alpha i :  L_ (Succ alpha) i = L_  alpha (S i).
+
+Hint Rewrite L_zero_eqn L_succ_eqn : L_rw.
+\end{Coqsrc}
+
+Using these three lemmas as rewrite rules, we can prove more properties of the functions \texttt{L\_ $\alpha$}.
+
+\begin{Coqsrc}
+Lemma L_finite : forall i k :nat,  L_ i k = (i+k)%nat.  
+(* Proof by induction on i, using L_zero_eqn and L_succ_eqn *)
+
+Lemma L_omega : forall k, L_ omega%e0 k = S (2 * k)%nat.
+(* Proof using L_finite and L_lim_eqn *)
+\end{Coqsrc}
+
+By  well-founded induction on $\alpha$, we prove the following lemmas:
+
+\begin{Coqsrc}
+Lemma L_ge_S alpha : alpha <> Zero -> S <<= L_ alpha.
+
+Theorem L_correct alpha : L_spec (cnf alpha) (L_ alpha).
+\end{Coqsrc}
+
+Please note that the proof of \texttt{L\_correct} applies the lemmas proven in Sections~\ref{sect:L-spec-zero}, ~\ref{sect:L-spec-succ} and ~\ref{sect:L-spec-lim}.
+Our previous study of \texttt{L\_spec} allowed us to pave the way for the definition by \texttt{Equations} and the correctness proof.
+
+\subsubsection{Back to Hydra battles}
+
+Theorem \texttt{battle\_length\_std } of
+Module~\href{../src/html/hydras.Hydra.Hydra_Theorems.html\#battle_length_std}{Hydra.Hydra\_Theorems} relates the length of standard battles with the functions $L_\alpha$.
+
+\begin{Coqsrc}
+Theorem battle_length_std (alpha : E0)  :
+  alpha <> Zero ->
+  forall k, (1 <= k)%nat ->
+            battle_length standard k (iota (cnf alpha))
+                         (L_ alpha (S k) - k).
+\end{Coqsrc}
+
+\index{Projects}
+\begin{project}
+Instead of considering standard paths and battles, consider ``constant'' paths and the corresponding battles.
+Please use \texttt{Equations} in order to define the function that computes the length of the $k$-path which leads  from $\alpha$ to $0$.
+Prove a few  exact formulas and minoration lemmas.
+\end{project}
+
+\section{The Wainer-Hardy hierarchy}
+\label{sect:hardy}
+In order to give an idea of the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the so called \emph{Wainer-Hardy hierarchy} of fast growing functions,
+presented for instance in~\cite{Promel2013}. 
+\index{Maths!Hardy Hierarchy}
+
+For each ordinal $\alpha$ below $\epsilon_0$, $H_\alpha$ is a 
+total arithmetic function, defined  by  transfinite recursion on $\alpha$, according to three cases:
+
+\index{Maths!Transfinite induction}
+
+\begin{itemize}
+\item If $\alpha=0$, then $H_\alpha (k)= k$ for any natural number $k$.
+\item If $\alpha=\textrm{succ}(\beta)$, then 
+$H_\alpha(k)=H_\beta(k+1)$ for any $k \in \mathbb{N}$
+\item If $\alpha$ is a limit ordinal, then 
+$H_\alpha(k) = H_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
+\end{itemize}
+
+\subsection{Hardy functions in \texttt{Coq}}
+
+We define a function \texttt{H\_} of type \texttt{E0 -> nat -> nat} by transfinite induction over the type \texttt{E0} of the well formed ordinals below $\epsilon_0$.
+
+\vspace{4pt}
+\emph{From Module~\href{../src/html/hydras.Epsilon0.H_alpha.html\#H_}{Epsilon0.H\_alpha}}
+
+\index{Coq!Plug-ins!Equations}
+\label{Functions:H-alpha}
+
+\begin{Coqsrc}
+Equations H_ (alpha: E0) (i:nat) :  nat  by wf  alpha lt :=
+  H_ alpha  i with E0_eq_dec alpha Zero :=
+    { | left _ =>  i ;
+      | right nonzero
+          with Utils.dec (Limitb alpha) :=
+          { | left _ =>  H_ (Canon alpha (S i))  i ;
+            | right notlimit =>  H_ (Pred alpha) (S i)}}.
+
+Solve All Obligations with auto with E0.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma H_zero_eqn : forall i, H_ Zero i = i.
+Proof.   intro i; now rewrite H__equation_1.  Qed.
+
+Lemma H_eq2 alpha i : Is_Succ alpha ->
+                      H_ alpha i = H_ (Pred alpha) (S i).
+
+Lemma H_lim_eqn alpha i : Limitb alpha ->
+                      H_ alpha i =  H_ (Canon alpha (S i)) i.
+
+Lemma H_eq4  alpha i :  H_ (Succ alpha) i = H_ alpha (S i).
+\end{Coqsrc}
+
+\subsection{First  steps of the Hardy hierarchy}
+Using rewrite rules from \texttt{H\_eq1} to \texttt{H\_eq4}, we can explore the functions $H_\alpha$ for some small values of $\alpha$.
+
+\subsubsection{Finite ordinals}
+
+By induction on $i$, we prove a simple expression of \texttt{H\_ (Fin i)}, where 
+\texttt{Fin $i$}  is the $i$-th finite ordinal.
+
+\begin{Coqsrc}
+Lemma H_Fin : forall i k: nat,  H_ (Fin i) k = (i+k)%nat.
+Proof with eauto with E0.
+  induction i.
+  - intros; simpl OF; simpl; autorewrite with H_rw E0_rw ... 
+  - intros ;simpl; autorewrite with H_rw E0_rw ... 
+    rewrite IHi; lia. 
+Qed.
+\end{Coqsrc}
+
+\subsubsection{Multiples of \texorpdfstring{$\omega$}{omega}}
+
+Since the canonical sequence of $\omega$ is composed of finite ordinals, 
+it is easy to get the formula associated with $H_\omega$.
+
+\begin{Coqsrc}
+Lemma H_omega : forall k, H_ Omega k = S (2 * k)%nat.
+Proof with auto with E0.
+  intro k; rewrite H_lim_eqn ...
+  - replace (Canon omega (S k)) with (Fin (S k)).
+    + rewrite H_Fin; lia.
+    +  now autorewrite with E0_rw.
+Qed.
+\end{Coqsrc}
+
+Before going further, we prove a useful rewriting lemma:
+
+\begin{Coqsrc}
+Lemma H_Plus_Fin alpha : forall i k : nat,
+    H_ (alpha + i)%e0 k = H_ alpha (i + k)%nat.
+(* Proof by induction on i *)
+\end{Coqsrc}
+
+Then, we get easily formulas for $H_{\omega+i}$, and $H_{\omega\times i}$ for any natural number $i$.
+
+\begin{Coqsrc}
+Lemma H_omega_double k : H_ (omega * 2)%e0 k =  (4 * k + 3)%nat.
+Proof.
+ rewrite H_lim_eqn; simpl Canon.
+ - ochange  (CanonS  (omega * 2)%e0 k)  (omega + (S k))%e0.
+  + rewrite H_Plus_Fin, H_omega;  lia.
+  -  now compute.
+Qed.
+
+Lemma H_omega_3 k : H_ (omega * 3)%e0 k = (8 * k + 7)%nat.
+
+Lemma H_omega_4 k : H_ (omega * 4)%e0 k = (16 * k + 15)%nat.
+
+Lemma H_omega_i i  : forall k,
+    H_ (omega * i)%e0 k = (exp2 i * k + Nat.pred (exp2 i))%nat.
+\end{Coqsrc}
+
+Crossing a new limit, we prove the following equality: 
+$$H_{\omega^2} (k) = 2 ^ {k+1} \times (k+1) - 1$$.
+
+\begin{Coqsrc}
+
+Lemma H_omega_sqr : forall k,
+    H_ (Phi0  2)%e0 k = (exp2 (S k ) *  S k - 1)%nat.
+Proof.
+  intro k; 
+   rewrite H_lim_eqn; auto with E0.
+  - ochange (Canon (Phi0 2) (S k)) (omega * (S k))%e0.
+    +  rewrite H_omega_i; simpl (exp2 (S k)).
+       *  rewrite Nat.add_pred_r.
+          -- lia. 
+          --   generalize (exp2_not_zero k);  lia.
+    + cbn; f_equal; lia.
+Qed.
+\end{Coqsrc}
+
+\subsubsection{New limits}
+
+Our next step would be to prove an exact formula for $H_{\omega^\omega}(k)$.
+Since the canonical sequence of $\omega^\omega$ is composed of all the
+$\omega^i$, we first need to express $H_{\omega^i}$ for any natural number $i$.
+
+Let $i$ and $k$ be two natural numbers. 
+The ordinal $\canonseq{\omega^(i+1)}{k}$ is the product
+$\omega^i \times k$, so we need also to consider ordinals of this form.
+
+\begin{enumerate}
+\item First,  we express $H_{\omega^\alpha \times (i+2)}$ in terms of
+$H_{\omega^\alpha \times (i+1)}$.
+
+\begin{Coqsrc}
+Lemma H_Omega_term_1 : alpha <> Zero -> forall  k,  
+    H_ (Omega_term alpha (S i)) k =
+    H_ (Omega_term alpha i) (H_ (Phi0 alpha) k).
+\end{Coqsrc}
+
+\item
+Then, we prove by induction on $i$ that $H_{\omega^\alpha \times (i+1)}$ is just the
+$(i+1)$-th iterate of $H_{\omega^\alpha}$.
+
+\begin{Coqsrc}
+Lemma H_Omega_term (alpha : E0)  :
+alpha <> Zero -> forall i k, 
+  H_ (Omega_term alpha i) k = iterate  (H_ (Phi0 alpha)) (S i) k.
+\end{Coqsrc}
+
+\item In particular, we have got a formula for $H_{\omega^{i+1}}$.
+
+\begin{Coqsrc}
+Definition H_succ_fun f k := iterate f (S k) k.
+
+Lemma H_Phi0_succ alpha  : alpha <> Zero -> forall k,
+      H_ (Phi0 (Succ alpha)) k = H_succ_fun (H_ (Phi0 alpha)) k.
+
+Lemma H_Phi0_Si : forall i k,
+      H_ (Phi0 (S i)) k = iterate H_succ_fun i (H_ omega) k. 
+\end{Coqsrc}
+
+\end{enumerate}
+We get now a  formula for $H_{\omega^3}$:
+
+\begin{Coqsrc}
+Lemma H_omega_cube : forall k,
+    H_ (Phi0 3)%e0 k = iterate (H_ (Phi0 2))  (S k) k.
+Proof.
+  intro k; rewrite <-FinS_eq, -> Fin_Succ, H_Phi0_succ; auto.
+  compute; injection 1; discriminate.
+Qed.
+\end{Coqsrc}
+
+\subsubsection{A numerical example}
+
+It seems hard to capture the complexity of this function by looking only at this
+``exact'' formula. 
+Let us consider a simple example, the number $H_{\omega^3}(3)$.
+
+\begin{Coqsrc}
+Section H_omega_cube_3.
+
+Let f k :=   (exp2 (S k ) * (S k) - 1)%nat.
+
+Remark R0 k :  H_ (Phi0 3)%e0 k = iterate f (S k) k.
+\end{Coqsrc}
+
+Thus, the number $H_{\omega^3}(3)$ can be written as four nested applications of $f$.
+
+\begin{Coqsrc}
+Fact F0 : H_ (Phi0 3) 3 = f (f (f (f 3))).
+ rewrite R0; reflexivity. 
+Qed.
+\end{Coqsrc}
+
+In order to make this statement more readable, we can introduce a local définition.
+
+\begin{Coqsrc}
+Let N := (exp2 64 * 64 - 1)%nat.
+\end{Coqsrc}
+
+This number looks quite big; let us compute an approximation in \texttt{Ocaml}:
+
+\begin{Coqsrc}
+# (2.0 ** 64.0 *. 64.0 -. 1.0);; 
+\end{Coqsrc}
+
+\begin{Coqanswer}
+- : float = 1.1805916207174113e+21
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Fact F1 : H_ (Phi0 3) 3 = f (f N).
+Proof.
+ rewrite H_omega_cube_0; reflexivity. 
+Qed.
+
+Lemma F1_simpl : H_ (Phi0 3) 3 =
+                 (exp2 (exp2 (S N) * S N) * (exp2 (S N) * S N) - 1)%nat.
+
+\end{Coqsrc}
+
+In a more classical writing, this number is displayed as follows:
+
+{\Large
+$$
+H_{\omega^3}(3) =  2 ^ {(2 ^ {N + 1} \times (N+1) } )  \times  (2 ^ {N+1} \times ( N +1) ) - 1
+$$
+}
+
+We leave as an exercise to determine the best approximation as possible of
+ the size of this number (for instance its number of digits).  For instance, if
+we do not take into account the multiplications in the formula above,
+we obtain that, in base $2$, the number $H_{\omega^3}(3)$ has at least
+$2^{10^{21}}$  digits. But it is still an under-approximation !
+
+\begin{Coqsrc}
+End H_omega_cube_3.
+\end{Coqsrc}
+
+Now, we have got at last an exact formula for $H_{\omega^\omega}$.
+
+\begin{Coqsrc}
+Lemma H_Phi0_omega : forall k, H_ (Phi0 omega) k =
+                               iterate H_succ_fun  k (H_ omega) k.
+Proof with auto with E0.
+  intro k; rewrite H_lim_eqn,  <- H_Phi0_Si ...
+  -  rewrite CanonS_Canon, CanonS_Phi0_lim;  f_equal ...
+Qed.
+\end{Coqsrc}
+
+Using extensionality of the functional \texttt{iterate}, we can get a closed formula.
+
+\begin{Coqsrc}
+Lemma H_Phi0_omega_closed_formula k :
+  H_ (Phi0 omega) k =
+  iterate (fun (f: nat -> nat) (l : nat) => iterate  f (S l) l)
+               k
+               (fun k : nat => S (2 * k)%nat)
+               k.
+\end{Coqsrc}
+
+Note that this short formula contains two occurences of the functional \texttt{iterate}, the outer one is in fact a second-order iteration (on type \texttt{nat -> nat)}
+and the inner one  first-order (on type \texttt{nat}).
+
+\subsection{Abstract properties of H-functions}
+~\label{sect:H-alpha-prop}
+
+Since pure computation seems to be useless for dealing with expressions of the form $H_\alpha(k)$, even for small values of $\alpha$ and $k$, we need to prove theorems for comparing $H_\alpha(k)$ and $H_\beta(l)$, in terms of comparison
+between $\alpha$ and $\beta$ on the one hand, $k$ and $l$ on the other hand.
+
+But beware of non-theorems! For instance, one could believe that $H$ is monotonous in its first argument. The following proof shows this is false.
+
+\begin{Coqsrc}
+Remark H_non_mono1 :
+  ~ (forall alpha beta k, (alpha o<= beta)%e0 ->
+                          (H_ alpha k <= H_ beta k)%nat).
+Proof.
+ intros H ;specialize (H 42 omega 3).
+ assert (H0 :(42 o<= omega)%e0) by (repeat split; auto).  
+ apply H in H0; rewrite H_Fin, H_omega  in H0; lia.
+Qed.
+\end{Coqsrc}
+
+On the contrary, the fonctions of the Hardy hierarchy have the following five properties~\cite{KS81}: for any $\alpha < \epsilon_0$,
+\begin{itemize}
+\item the function $H_\alpha$ is strictly monotonous :
+      For all $n,p \in\mathbb{N}, n < p \Rightarrow H_\alpha(n)< H_\alpha(p)$.
+\item If $\alpha \not= 0$, then for every $n$, $n<H_\alpha(n)$.
+\item The function $H_\alpha$ is pointwise less or equal than $H_{\alpha+1}$
+
+\item For any $n\geq 1$, $H_\alpha(n)<H_{\alpha+1}(n)$.
+\emph{We say that $H_{\alpha+1}$ dominates $H_\alpha$ from $1$}.
+\item For any $n$ and $\beta$, if $\alpha \xrightarrow[n]{} \beta$, then
+$H_\beta(n)\leq H_\alpha(n)$.
+\end{itemize}
+
+\index{Maths!Transfinite induction}
+
+In \coq{}, we follow the  proof in~\cite{KS81}. This proof is mainly a single  proof by transfinite induction on $\alpha$ of the conjonction of the five properties.
+For each $\alpha$, the three cases : $\alpha=0$, $\alpha$ is a limit, and 
+$\alpha$ is a successor are considered. Inside each case, the five sub-properties are proved sequentially.
+
+\begin{Coqsrc}
+Section Proof_of_Abstract_Properties.
+  Record P (alpha:E0) : Prop :=
+    mkP {
+        PA : strict_mono (H_ alpha);
+        PB : alpha <> Zero -> forall n,  (n < H_ alpha n)%nat;
+        PC : H_ alpha <<= H_ (Succ alpha);
+        PD : dominates_from 1 (H_ (Succ alpha)) (H_ alpha);
+        PE : forall beta n, Canon_plus n alpha beta -> 
+                            (H_ beta n <= H_ alpha n)%nat}.
+
+Theorem P_alpha : forall alpha, P alpha.
+  Proof.
+    intro alpha; apply well_founded_induction with lt.
+   (* rest of proof skipped *)
+
+Section Proof_of_Abstract_Properties.
+\end{Coqsrc}
+
+\subsection{Comparison between \texttt{L\_} and \texttt{H\_} }
+
+By well-founded induction on $\alpha$, we prove that our $L$ hierachy is ``almost'' the Hardy hierarchy (up to a small shift).
+
+\emph{From Module~\href{../src/html/hydras.Epsilon0.L_alpha.html\#H_L_}{Epsilon0.L\_alpha}}
+
+\begin{Coqsrc}
+ Theorem H_L_ alpha: forall i:nat,  (H_ alpha i <= L_ alpha (S i))%nat.
+\end{Coqsrc}
+
+\subsubsection{Back to hydras}
+
+The following theorem relates the length of (standard) battles with the Hardy hierarchy.
+
+\emph{From Module~\href{../src/html/hydras.Epsilon0.L_alpha.html}{Epsilon0.L\_alpha}}
+
+\begin{Coqsrc}
+Theorem battle_length_std_Hardy (alpha : E0) :
+  alpha <> Zero ->
+  forall k , 1 <= k -> exists l: nat,  
+       H_ alpha k - k <= l /\
+       battle_length standard k (iota (cnf alpha)) l.    
+\end{Coqsrc}
+
+\section{The Wainer hierarchy (functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
+\label{sect:wainer}
+
+\index{Maths!Wainer Hierarchy}
+
+The Wainer hierarchy~\cite{BW85, Wainer1970, KS81}, is also a family of fast growing functions, indexed by ordinals below $\epsilon_0$, by the following equations:
+
+\label{F_equations}
+\begin{itemize}
+\item $F_0(i)=i+1$
+\item $F_{\beta+1}(i)= (F_\beta)^{(i+1)}(i)$, where $f^{(i)}$ is the $i$-th iterate of $f$.
+\item $F_\alpha(i) = F_{\canonseq{\alpha}{i}} (i)$ if $\alpha$ is a limit ordinal.
+\end{itemize}
+
+A first attempt is to write a definition of $F_\alpha$ by equations, in the same as for $H\_alpha$.  We use the functional \texttt{iterate} defined in 
+Module~\href{../src/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
+
+\begin{Coqsrc}
+Fixpoint iterate {A:Type}(f : A -> A) (n: nat)(x:A) :=
+  match n with
+  | 0 => x
+  | S p => f (iterate  f p x)
+  end.
+\end{Coqsrc}
+
+The following code comes from 
+ \url{../src/html/hydras.Epsilon0.F_alpha.html}.
+
+\index{Coq!Plug-ins!Equations}
+\begin{Coqsrc}
+Fail Equations F_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
+  F_ alpha  i with E0_eq_dec alpha Zero :=
+    { | left _ =>  i ;
+      | right nonzero
+          with Utils.dec (Limitb alpha) :=
+          { | left _ =>  F_ (Canon alpha i)  i ;
+            | right notlimit =>  iterate (F_ (Pred alpha))  (S i) i}}.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+The command has indeed failed with message:
+In environment
+alpha : E0
+notlimit : Limitb alpha = false
+nonzero : alpha <> Zero
+i : nat
+F_ : forall x : E0, nat -> x o< alpha -> nat
+The term "F_ (Pred alpha) ?x" has type "Pred alpha o< alpha -> nat"
+while it is expected to have type "Pred alpha o< alpha -> Pred alpha o< alpha"
+(cannot unify "nat" and "Pred alpha o< alpha").
+\end{Coqanswer}
+
+We presume that this error comes from the recursive call of \texttt{F\_} inside
+an application of \texttt{iterate}. The workaround we propose is to define first 
+the iteration of \texttt{F\_}  as an helper $F^*$, then to define the function $F$ as a ``iterating $F^*$ once''.
+
+\texttt{Equations} accepts the following definition, relying on  lexicographic ordering on pairs $(\alpha,n)$.
+
+\label{sect:F-equations}
+
+\index{Coq!Plug-ins!Equations}
+\label{Functions:F-alpha}
+\index{Maths! Fast growing functions}
+
+\begin{Coqsrc}
+Definition call_lt (c c' : E0 * nat) :=
+  lexico Lt (Peano.lt) c c'.
+
+Lemma call_lt_wf : well_founded call_lt.
+  unfold call_lt; apply Inverse_Image.wf_inverse_image,  wf_lexico.
+  -  apply E0.Lt_wf.
+  -  unfold Peano.lt; apply Nat.lt_wf_0. 
+Qed.
+
+Instance WF : WellFounded call_lt := call_lt_wf.
+
+Equations  F_star (c: E0 * nat) (i:nat) :  nat by wf  c call_lt :=
+    F_star (alpha, 0) i := i;
+    F_star (alpha, 1) i
+      with E0_eq_dec alpha Zero :=
+           { | left _ => S i ;
+             | right nonzero
+                 with Utils.dec (Limitb alpha) :=
+                 { | left _ => F_star (Canon alpha i,1) i ;
+                   | right notlimit =>
+                     F_star (Pred alpha, S i)  i}};
+    F_star (alpha,(S (S n))) i :=
+               F_star (alpha, 1) (F_star (alpha, (S n)) i).
+
+(* Finally, F_ alpha is defined as its first iterate ! *)
+
+Definition F_  alpha i := F_star (alpha, 1) i.
+\end{Coqsrc}
+
+It is quite easy to prove that our function \texttt{F\_} satisfies the equations on page~\pageref{sect:F-equations}.
+
+\begin{Coqsrc}
+Lemma F_zero_eqn : forall i, F_ Zero i = S i.
+
+Lemma F_lim_eqn : forall alpha i,  Limitb alpha ->
+                               F_ alpha i = F_ (Canon alpha i) i.
+
+Lemma F_succ_eqn : forall alpha i,
+    F_ (Succ alpha) i = iterate (F_ alpha) (S i) i.
+\end{Coqsrc}
+
+As for the Hardy functions, we can use these equalities as rewrite rules for
+``computing'' some values of $F_\alpha(i)$, for small values of $\alpha$.
+
+\begin{Coqsrc}
+Lemma LF1 : forall n,  F_ 1 n = S (2 * n).
+
+Lemma LF2 : forall i, (exp2 i * i < F_ 2 i)%nat.
+\end{Coqsrc}
+
+Like in Sect~\ref{sect:H-alpha-prop}, we prove by induction the following properties (see~\cite{KS81}).
+
+\begin{Coqsrc}
+Theorem F_alpha_mono alpha : strict_mono (F_ alpha).
+
+Theorem F_alpha_ge_S alpha : forall n, (n < F_ alpha n)%nat.
+
+Theorem F_alpha_Succ_le alpha : F_ alpha <<= F_ (Succ alpha).
+
+Theorem F_alpha_dom alpha : dominates_from 1 (F_ (Succ alpha)) (F_ alpha).
+
+Theorem F_alpha_beta alpha : forall beta n, Canon_plus n alpha beta -> 
+                                        (F_ beta n <= F_ alpha n)%nat.
+\end{Coqsrc}
+
+As a corollary, we prove the following proposition, p. 284 of~\cite{KS81}.
+
+\begin{quote}
+  If $\beta<\alpha$, $F_\alpha$ dominates $F_\beta$.
+
+($f$ is said to \emph{dominate} $g$ if $f(p)>g(p)$ for any all sufficiently large $p$).
+\end{quote}
+
+\begin{Coqsrc}
+Lemma Propp284 : forall alpha beta : E0, 
+   beta o< alpha -> dominates (F_ alpha) (F_ beta).
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+Let us quote a theorem from ~\cite{KS81} (page 297).
+
+\begin{quote}
+\begin{align*}
+  H_{\omega^\alpha}(n+1) &\geq F_{\alpha}(n) \quad (n\geq 1, \alpha<\epsilon_0) \\
+ F_{\alpha}(n+1) &\geq H_{\omega^\alpha}(n) \quad (n\geq 1, \alpha<\epsilon_0) 
+\end{align*}
+Thus $H_{\omega^\alpha}$ and $F_{\alpha}$ have essentially the same order of growth.
+
+\end{quote}
+
+ But, before trying to prove these facts, look at the definition of function $H$ in Ketonen and Solovay's paper !
+Is it really the same as the definition we quote from Pr{\H o}mel's chapter~\cite{Promel2013},
+whereas \cite{KS81} define $H_\alpha(n)$ as ``the least integer $k$ such that $[n,k]$ is $\alpha$-large''. Thus, it may be useful to adapt the statement above.
+
+\end{exercise}
+
+\index{Exercises}
+
+\begin{exercise}
+Prove the following result~\cite{KS81}(p. 298).
+
+\begin{quote}
+   For $n\geq 2$ and $\alpha \geq 3$, $F_\alpha(n+1)\geq 2^{F_\alpha(n)}$.
+\end{quote}
+\end{exercise}
+
+%  \subsection{Gnawing ordinals}
+
+% \begin{definition}[After~\cite{KS81}]
+%   Let $S$ be a finite set of positive integers, and $\alpha$ be an ordinal strictly less than $\epsilon_0$.  Let us denote by $s=s_1,s_2, \dots, s_N$ the sequence of 
+% the elements of $S$, enumerated in strictly increasing order.
+
+% We consider the sequence of ordinals $\alpha_o=\alpha, \alpha_1=\canonseq{\alpha_0}{s_1},\dots,\alpha_{i+1}=\canonseq{\alpha_i}{s_{i+1}},\dots, 
+% \alpha_{N}=\canonseq{\alpha_{N-1}}{s_N}$.
+% We denote by $\gnaw{s}{\alpha}$ the last ordinal of the sequence, \emph{i.e.}  $\alpha_N$.
+% \end{definition}
+
+% The following function computes $\gnaw{s}{\alpha}$ by recursion on $s$.
+% \vspace{4pt}
+
+% For instance, let us consider the ordinal $\omega^\omega$, and try some sequence of integers.
+
+% \begin{Coqsrc}
+% Compute pp (gnaw (omega ^ omega) (1::nil)).  
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+%   = omega%pT1 : ppT1
+% \end{Coqanswer}
+
+% \begin{Coqsrc}
+% Compute pp (gnaw (omega ^ omega) (1::2::nil)).  
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+% = P_fin 2
+%      : ppT1
+% \end{Coqanswer}
+
+% Likewise, we can verify that $\gnaw{\omega^\omega}{\langle 1,2,3 \rangle}=1$
+% and $\gnaw{\omega^\omega}{\langle 1,2,3,4 \rangle}=0$.
+
+% \begin{Coqsrc}
+% Example omega_omega_1_4 : gnaw (omega ^omega) (interval 1 4) = 0.
+% Proof. trivial. Qed.
+
+% Example omega_omega_1_3 : gnaw (omega ^omega) (interval 1 3) = 1.
+% Proof. trivial. Qed.
+% \end{Coqsrc}
+
+% \begin{remark}
+% \label{remark:gnaw-vs-battles}
+% Let us consider an hydra battle, where Hercules always chops off the rightmost head of the hydra. Let $\alpha<\epsilon_0$ be an ordinal number, and 
+% $s=\langle i_1<i_2<\dots<i_N\rangle$ be  any finite sequence of positive integers.
+% Then the battle initiated by the hydra $\iota(\alpha)$, with $s$ as the sequence of successive replication factors, leads to the hydra $\iota(\gnaw{s}{\alpha})$ as the final state.
+% \end{remark}
+
+% \subsection{Large sequences}
+
+% \begin{remark}
+% In their article~\cite{KS81}, Ketonen and Solovay use the appellation ``large set'' instead of ``large sequence'', but their definitions use an enumeration of the elements in increasing order.
+Thus, we shall use the term ``sequence'' when referring to our implementation in \coq{}, and ``set'' when referring to the statements of~\cite{KS81}.
+% \end{remark}
+
+% \begin{definition}[After\cite{KS81}]
+% The sequence $s$ is said to be \emph{$\alpha$-large} if $\gnaw{\alpha}{s}=0$.  
+% \end{definition}
+
+% \begin{Coqsrc}
+% Definition largeb (alpha : T1) (s: list nat) :=
+%   match gnaw alpha s with
+%     | zero => true
+%     | _ => false
+%   end.
+
+% Definition large (alpha : T1) (s : list nat) : Prop :=
+%   largeb alpha s.
+% \end{Coqsrc}
+
+% For instance, the sequence $\langle 1,2,3,4 \rangle$ is $\omega^\omega$-large but not
+% $\omega^{\omega+1}$-large, since $\gnaw{\omega^{\omega+1}}{\langle 1,2,3,4\rangle}$ is equal to $\omega \times 2 + 4$.
+
+% %\omega^\omega+\omega^2\times 3 + \omega\times 4+ 5$.
+
+% \begin{Coqsrc}
+% Compute largeb (omega^omega) (interval 1 4).  
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+% = true : bool
+% \end{Coqanswer}
+
+% \begin{Coqsrc}
+% Compute largeb (omega^(omega+1)) (interval 1 4).
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+% = false : bool
+% \end{Coqanswer}
+
+% \begin{remark}
+% A sequence $s$ is $\alpha$-large if (still considering Hercules ``righmost-head'', tactic), it leads to Hercules' victory.
+% \end{remark}
+
+% \subsection{A little game}
+
+% Let $\alpha<\epsilon_0$ be an ordinal, and $i$ a positive integer. We want to guess the least natural number  $j$  such that the interval $[i,i+j]$ is $\alpha$-large.
+% Equivalently, we should have $\gnaw{\alpha}{[i, i+j-1]}=1$.
+
+% The following functions takes three arguments: an ordinal $\alpha$, and two positive natural numbers  $i$ and $j$ (we assume, but not verify that $i$ and $j$ are strictly positive).
+It returns one of the three possible answers:
+
+% \begin{itemize}
+% \item \texttt{Ok} if $j$ is the smallest integer such that the interval $[i,i+j]$ is $\alpha$-large
+% \item \texttt{Too\_far} if $[i,i+j]$ is  $\alpha$-large, but $j$ is not the smallest such positive integer
+% \item \texttt{(Remaining $\beta$)} if $j$ is too small, and gnawing $\alpha$ with
+%   $[i,i+j]$ is still equal to $\beta$, instead of $0$
+% \end{itemize}
+
+% \vspace{4pt}
+% \noindent
+% \emph{From Module~\href{../src/html/hydras.Epsilon0.Large_Sets_Demo.html}{Ordinals.Epsilon0.Large\_Sets\_Demo}}
+
+% \begin{Coqsrc}
+% Inductive answer : Set := You_won | Too_far | Remaining (rest : ppT1).
+
+% Definition large_set_check alpha i j :=
+%   let beta := gnaw alpha (interval i (Nat.pred j))
+%   in match beta with
+%      | one => Ok
+%      | zero => Too_far
+%      |  _ => Remaining (pp (canonseq j beta))
+%      end.
+% \end{Coqsrc}
+
+% \subsubsection{\texorpdfstring{$\omega$}{omega}-large intervals}
+
+% For instance, let us consider the ordinal $\omega$ and start with $i=1$.
+
+% \begin{Coqsrc}
+% Compute large_set_check omega 1 2.
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+%      = Ok
+%      : answer
+% \end{Coqanswer}
+
+% Let us give greater values of $i$, still with the ordinal $\omega$.
+
+% \begin{Coqsrc}
+% Compute large_set_check omega 2 3.
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+%   = Remaining 1
+%      : answer
+% \end{Coqanswer}
+
+% \begin{Coqsrc}
+% Compute large_set_check omega 3 3.
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+%   = Remaining 3
+%      : answer
+% \end{Coqanswer}
+
+% \begin{Coqsrc}
+% Compute large_set_check omega 2 4.
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+%   = Ok
+%      : answer
+% \end{Coqanswer}
+
+% \begin{Coqsrc}
+% Compute large_set_check omega 3 6.
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+%   = Ok
+%      : answer
+% \end{Coqanswer}
+
+% It looks like every request to compute (\texttt{large\_set\_check omega $i$ $2\times i$})  will succeed. Let us try an example.
+
+% \begin{Coqsrc}
+% Compute large_set_check omega 49 98.
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+%   = Ok
+%      : answer
+% \end{Coqanswer}
+
+% \subsubsection{\texorpdfstring{$\omega^2$}{omega\^{}2}-large intervals}
+
+% Still using \texttt{Compute} and \texttt{large\_set\_check}, we obtained the 
+% following values of $j$, the least integer such that the interval $[i,j]$ is $\omega^2$ large.
+
+% $$
+% \begin{array}{|c|c|}
+% \hline
+%   i & j \\
+% \hline 
+% 1 & 4 \\
+% 2 & 14 \\
+% 3 & 38 \\
+% 4 & 94 \\
+% 5 & 222 \\
+% % 6 & 510 
+% \hline
+% \end{array}
+% $$
+
+% \begin{exercise}
+% Please give the 6-th and 7-th line of the array above.
+% \end{exercise}
+
+% \subsubsection{The limits of (pure) computation}
+
+% Let us now play with bigger ordinals, for instance $\alpha=\omega^\omega$ or
+% $\alpha=\omega^{\omega + 1}$. We notice that, even for small values of $i$, it is hard
+% to guess values of $j$ such that $[i,i+j]$ is $\alpha$-large.
+
+% \begin{Coqsrc}
+% Compute large_set_check (omega ^ omega) 1 4.
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+%  = Ok
+%      : answer
+% \end{Coqanswer}
+
+% \begin{Coqsrc}
+% Compute large_set_check (omega ^ omega) 2 38.
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+%  = Ok
+%      : answer
+% \end{Coqanswer}
+
+% \begin{Coqsrc}
+% Compute large_set_check (omega ^ omega) 3 1000.
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+%   = Remaining (omega ^ 2 * 2 + omega * 220 + 798)%pT1
+%      : answer
+% \end{Coqanswer}
+
+% \begin{Coqsrc}
+% Compute large_set_check (omega ^ omega) 3 1798.
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+%   = Remaining (omega ^ 2 * 2 + omega * 220)%pT1
+%      : answer
+% \end{Coqanswer}
+
+% \begin{Coqsrc}
+% Compute large_set_check (omega ^ omega) 3 5000.
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+%   = Remaining (omega ^ 2 * 2 + omega * 218 + 2198)%pT1
+%      : answer
+% \end{Coqanswer}
+
+% \begin{Coqsrc}
+% Compute large_set_check (omega ^ (omega + 1)) 3 5000.
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+%  = Remaining
+%          (omega ^ omega * 2 + omega ^ 3 * 4 + 
+%           omega ^ 2 * 4 + omega * 1148 +  4222)%pT1
+%      : answer  
+% \end{Coqanswer}
+
+% \begin{Coqsrc}
+% Compute large_set_check (omega ^ (omega + 1)) 3 10000.
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+% Warning: To avoid stack overflow, large numbers in nat are interpreted as
+% applications of Init.Nat.of_uint. [abstract-large-number,numbers]
+
+%     = Remaining
+%          (omega ^ omega * 2 + omega ^ 3 * 4 + 
+%           omega ^ 2 * 4 + omega * 1147 +
+%           8446)%pT1
+%      : answer
+% \end{Coqanswer}
+
+% Since computation is not enough, let us comme back to proofs, or, better, proofs \emph{and} computations.
+
+% \subsection{Proving largeness}
+
+% %%% ICI %%%
+
+% \subsection{$n$-large and $\omega$-large intervals }
+
+% A finite ordinal is just made out of \texttt{zero} and \texttt{succ} constructions.
+% Thus any sequence of strictly positive integers of length greater than  or equal to $n$ will completely gnaw the ordinal $n$.
+
+% Concerning the first limit ordinal $\omega$ the largeness of a sequence $s$ depends only on its first element and the length of the rest of the sequence (please keep in mind that the argument $i$ of $\canonseq{\alpha}{i}$ is meaningful only if $\alpha$ is a limit ordinal).
+% The following proposition is labeled $4.2$ in~\cite{KS81}.
+
+% \begin{proposition}
+%   For $n<\omega$, a set $X$id $n$-large if and only if $|X|\geq n$.
+% A finite set $X$ is $\omega$-large if $|X|>\min{X}$.
+% \end{proposition}
+
+% Rewritten in terms of list of strictly positive integers, we get the following statements:
+
+% \begin{Coqsrc}
+% Lemma large_n_iff : forall s (n:nat) , ~ In 0 s  ->
+%                                   large n s  <-> (n <= List.length s)%nat.
+% \end{Coqsrc}
+
+% \begin{Coqsrc}
+% Lemma large_omega_iff : forall s n,  ~In 0 (n::s) -> 
+%                                             large omega (n::s) <->
+%                                             (n <=  List.length s)%nat.
+% \end{Coqsrc}
+
+% % \begin{Coqsrc}
+% % (* sorted list of natural numbers greater than or equal to n *)
+
+% % Inductive sorted_ge (n: nat) : list nat -> Prop :=
+% % | sorted_ge_nil : sorted_ge n nil
+% % | sorted_ge_one : forall p, n<=p ->
+% %                             sorted_ge n (p::nil)
+% % | sorted_ge_cons: forall p q s,  n<=p -> p<q ->
+% %                                  sorted_ge p (q::s) ->
+% %                                  sorted_ge n (p::q::s).
+% % \end{Coqsrc}
+
+% \subsection{Mimimal large sequences}
+
+% Let us consider \emph{minimal} large sequences, \emph{i.e.} large sequences 
+% the strict prefix of which do not lead to $0$. In other words, only the last ordinal 
+% defined by the sequence is null.  For this purpose, we use the predicate \texttt{path\_to} introduced page~\pageref{path-to-definition}.
+
+% \begin{Coqsrc}
+% Definition mlarge alpha s := path_to zero s alpha.
+% \end{Coqsrc}
+
+%----------------------------------------------------------------------
+\chapter[Countable Ordinals (after Sch\"{u}tte)]{Kurt Schütte's axiomatic Definition of countable Ordinals}
+
+\label{chap:schutte} 
+%ON
+
+In the present chapter, we  compare our implementation of the segment $[0,\epsilon_0)$ with a mathematical text in order to ``validate'' our constructions.
+We chose as reference the axiomatic definition of the set of countable ordinals,
+in chapter V of Kurt Schütte's book `` Proof Theory ''~\cite{schutte}.
+
+\begin{remark}
+\emph{In all this chapter, the word ``ordinal'' will be considered as a synonymous of
+``countable ordinal''}  
+\end{remark}
+
+Schütte's definition of countable ordinal relies on the following three axioms.
+
+\begin{enumerate}
+\item There exists a well-ordered set $(\mathbb{O},<)$
+\item Every bounded subset of $\mathbb{O}$  is countable
+\item Every countable subset of $\mathbb{O}$  is bounded.
+\end{enumerate}
+
+Starting with these three axioms, Schütte re-defines the vocabulary about ordinal numbers: the null ordinal $0$, limits and successors, the addition of ordinals, the infinite ordinals $\omega$, $\epsilon_0$, $\Gamma_0$, etc.
+
+This chapter describes an adaptation to \coq{} of Schütte's axiomatization. 
+ Unlike the rest of our libraries, our library
+\href{../src/html/hydras.Schutte.Schutte.html}{Ordinals.Schutte}
+is not constructive, and relies on several axioms.
+
+\begin{itemize}
+\item First, please keep in mind  that the set of countable ordinals is not countable.
+Thus, we cannot hope to represent all countable ordinals as finite terms of an inductive type, which was possible with  the set of ordinals strictly less than $\epsilon_0$ (resp.
+$\Gamma_0$)
+\item We tried to be as close as possible to K.
+Schütte's text, which uses ``classical'' mathematics : excluded middle, Hilbert's $\epsilon$ (choice) and Russel's $\iota$ (definite description) operators.
+Both operators allow us to write definitions close to the natural mathematical language, such as ``$\textrm{succ}$ is \emph{the} least ordinal strictly greater than $\alpha$''
+\item Please note that only the sub-library \href{../src/html/hydras.Schutte.Schutte.html}{Ordinals.Schutte.Schutte/} is ``contaminated'' by axioms, and that the rest of our libraries remain constructive.
+
+\end{itemize}
+
+\section{Declarations and Axioms}
+
+Let us declare a type 
+\texttt{Ord} for representing countable ordinals, and a binary relation
+ \texttt{lt}. Note that, in our development, \texttt{Ord} is a type, while the \emph{set} of countable ordinals (called $\mathbb{O}$ by Schütte) 
+is the full set over the type \texttt{Ord}.
+
+\label{types:Ord}
+
+We use Florian Hatat's library on countable sets, written as he was a student of  \emph{\'Ecole Normale Supérieure de Lyon}.
+A set $A$ is countable if there is an injective function from $A$ to $\mathbb{N}$ (see
+Library \href{../src/html/hydras.Schutte.Countable.html}%
+{\texttt{Schutte.Countable}}).
+
+\vspace{6pt}
+
+\emph{From Module\href{../src/html/hydras.Schutte.Schutte_basics.html}%
+{\texttt{Schutte.Schutte\_basics}}}
+
+\begin{Coqsrc}
+Parameter Ord : Type.
+Parameter lt : relation Ord.
+Notation "a o< b" := (lt a b): schutte_scope.
+
+Definition ordinal := Full_set Ord.
+\end{Coqsrc}
+
+Schütte's first axiom tells that \texttt{lt} is a well order on the set 
+\texttt{ordinal} (The  class \texttt{WO} is defined in
+Module~\href{../src/html/hydras.Schutte.Well_Orders.html}%
+{\texttt{Well\_Orders.v}}).
+
+\label{types:WO}
+
+\begin{Coqsrc}
+Variables (M:Type)
+         (Lt : relation M).
+
+Class WO : Type:=
+    {
+      Lt_trans : Transitive  Lt;
+      Lt_irreflexive : forall a:M, ~ (Lt a a);
+      well_order : forall (X:Ensemble M)(a:M),
+          In X a ->
+          exists a0:M, least_member  X a0
+    }.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+  Axiom AX1 : WO lt.
+\end{Coqsrc}
+
+The second and third axioms say that a subset $X$ of $\mathbb{O}$ is
+(strictly) bounded if and only if it is countable.
+
+\begin{Coqsrc}
+Axiom AX2 : forall X: Ensemble Ord, 
+   (exists a,  (forall y, In X y -> y o< a)) ->
+   countable X.
+
+Axiom AX3 : forall X : Ensemble Ord,
+              countable X -> 
+              exists a,  forall y, In X y -> y o< a.
+\end{Coqsrc}
+
+\texttt{AX2} and \texttt{AX3} could have been replaced by a single axiom (using the \texttt{iff} connector), but we decide to respect as most as possible the structure of Schütte's definitions.
+
+\section{Additional  axioms}
+
+The adaptation of Schütte's mathematical discourse to \coq{} led us to
+import a few axioms from the standard library. We encourage the reader to consult \coq{}'s FAQ about the safe use of axioms
+ \url{https://github.com/coq/coq/wiki/The-Logic-of-Coq#axioms}.
+
+\subsubsection{Classical logic}
+
+In order to work with classical logic, we import the module
+\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Classical.html}{Coq.Logic.Classical}  of \coq{}'s standard library, specifially the following axiom:
+
+\begin{Coqsrc}
+ Axiom classic : forall P:Prop, P \/ ~P.
+\end{Coqsrc}
+
+\subsubsection{Description operators}
+
+In order to respect Schütte's style, we imported also the library 
+\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Epsilon.html}{\texttt{Coq.Logic.Epsilon}}.  The rest of this section presents a few examples of
+how Hilbert's choice operator and Church's definite description allow us
+ to write understandable definitions (close to the mathematical natural language).
+
+\subsubsection{The definition of zero}
+
+According to the  definition of a well order, every non-empty subset of \texttt{Ord} has a least element. Furthermore, this least element is unique.
+
+\begin{Coqsrc}
+Remark R : exists! z : Ord, least_member lt  ordinal z.
+Proof.
+  destruct inh_Ord as [a]; apply (well_order (WO:=AX1)) with a .
+  split.
+Qed.
+\end{Coqsrc}
+
+Assume we want to call this element  \texttt{zero}.
+
+\begin{Coqsrc}
+Definition zero : Ord.
+Proof.
+  Fail destruct R.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+The command has indeed failed with message:
+Case analysis on sort Type is not allowed for inductive 
+definition ex.
+\end{Coqanswer}
+
+Indeed, the basic logic of  \coq{} does not allow us to eliminate a proof of a proposition 
+$\exists\,x:A,\,P(x)$ for building a term whose type lies in the sort \texttt{Type}. 
+The reasons for this impossibility are explained in many documents~\cite{BC04, chlipalacpdt2011, Coq}.
+
+Let us import the library \texttt{Coq.Logic.Epsilon}, which contains the following axiom and lemmas.
+
+\begin{Coqsrc}
+Axiom epsilon_statement:
+  forall (A : Type) (P : A->Prop), inhabited A ->
+    {x : A | (exists x, P x) -> P x}.
+\end{Coqsrc}
+
+Hilbert's $\epsilon$ \emph{operator} is derived from this  axiom.
+
+\begin{Coqsrc}
+  Definition epsilon (A : Type) (i:inhabited A) (P : A->Prop) : A
+  := proj1_sig (epsilon_statement P i).
+
+Lemma constructive_indefinite_description :
+  forall (A : Type) (P : A->Prop),
+    (exists x, P x) -> { x : A | P x }.
+\end{Coqsrc}
+
+If we consider the \emph{unique existential} quantifier $\exists!$, we obtain
+Church's \emph{definite description operator}.
+
+\begin{Coqsrc}
+Definition iota (A : Type) (i:inhabited A) (P : A->Prop) : A
+  := proj1_sig (iota_statement P i).
+\end{Coqsrc}
+
+\begin{Coqsrc}
+ Lemma constructive_definite_description :
+  forall (A : Type) (P : A->Prop),
+    (exists! x, P x) -> { x : A | P x }.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Definition iota_spec (A : Type) (i:inhabited A) (P : A->Prop) :
+  (exists! x:A, P x) -> P (iota i P)
+  := proj2_sig (iota_statement P i).
+\end{Coqsrc}
+
+Indeed, the operators \texttt{epsilon} and \texttt{iota} allowed us to make our definitions 
+quite close to Schütte's text. Our libraries \href{../src/html/hydras.Schutte.MoreEpsilonIota.html}%
+{\texttt{Schutte.MoreEpsilonIota}}
+and
+\href{../src/html/hydras.Schutte.PartialFun.html}%
+{\texttt{Schutte.PartialFun}} are extensions of \texttt{Coq.logic.Epsilon} for making easier 
+such definitions. See also an article in french~\cite{PCiota}.
+
+\index{Coq!Type classes}
+
+\begin{Coqsrc}
+Class InH (A: Type) : Prop :=
+   InHWit : inhabited A.
+
+Definition some {A:Type} {H : InH A} (P: A -> Prop) := 
+   epsilon (@InHWit A H) P.
+
+Definition the {A:Type} {H : InH A} (P: A -> Prop) := 
+   iota (@InHWit A H) P.
+\end{Coqsrc}
+
+In order to use these tools,  we had to tell \coq{}  that the type \texttt{Ord} is not empty:
+
+\begin{Coqsrc}
+Axiom inh_Ord : inhabited Ord.
+\end{Coqsrc}
+
+We are now able de define \texttt{zero} as the least ordinal. For this purpose,
+we define a function returning the least element of any [non-empty]  set.
+
+\begin{Coqsrc}
+Definition the_least {M: Type} {Lt}
+           {inh : InH M} {WO: WO Lt} (X: Ensemble M)  : M :=
+  the  (least_member  Lt X ).
+\end{Coqsrc}
+
+\vspace{4pt}
+
+From Module \href{../src/html/hydras.Schutte.Schutte_basics.html}%
+{\texttt{~Schutte.Schutte\_basics}}
+
+\label{Constants:zero:Ord}
+
+\begin{Coqsrc}
+Definition zero: Ord :=the_least ordinal.
+\end{Coqsrc}
+
+We want to prove now that zero is less than or equal to any ordinal number.
+
+\begin{Coqsrc}
+Lemma zero_le (alpha : Ord) :  zero o<= alpha.
+Proof.
+  unfold zero, the_least, the; apply iota_ind.
+\end{Coqsrc}
+
+According to the use of the description operator \texttt{iota}, we have to solve  two trivial sub-goals.
+\begin{enumerate}
+\item Prove that there exists a unique least member of \texttt{Ord}
+\item Prove that being a least member of \texttt{Ord} entails the announced inequality 
+\end{enumerate}
+
+\begin{Coqanswer}
+2 subgoals (ID 155)
+
+  alpha : Ord
+  ============================
+  exists ! x : Ord, least_member lt ordinal x
+
+subgoal 2 (ID 156) is:
+ forall a : Ord, unique (least_member lt ordinal) a -> 
+                a o<= alpha
+\end{Coqanswer}
+
+\begin{Coqsrc}
+  -  apply the_least_unicity, Inh_ord.
+  -  destruct 1 as [[_ H1] _]; apply H1; split. 
+Qed.
+\end{Coqsrc}
+
+\subsubsection{Remarks on \texttt{epsilon} and \texttt{iota}}
+
+ What would happen in case of a misuse of \texttt{epsilon} or \texttt{iota} ?
+For instance, one could give a unsatisfiable specification to \texttt{epsilon} or 
+a specification for \texttt{iota} that admits several realizations.
+
+Let us consider an example:
+
+\begin{Coqbad}
+Module Bad.
+
+ Definition bottom := the_least (Empty_set Ord).
+\end{Coqbad}
+
+\begin{Coqanswer}
+ bottom is defined
+\end{Coqanswer}
+
+Since we won't be able to prove the proposition
+\linebreak \Verb|{exists! a: Ord, least_member (Empty_set Ord) a|, the only properties we would be able to prove about \texttt{bottom} would be \emph{trivial} properties, 
+\emph{i.e.}, satisfied by \emph{any} element of type \texttt{Ord}, like for instance
+\texttt{bottom = bottom}, or \texttt{zero o<= bottom}.
+
+\begin{Coqbad}
+Lemma le_zero_bottom : zero o<= bottom. 
+Proof. apply zero_le. Qed.
+
+Lemma bottom_eq : bottom = bottom.
+Proof. trivial. Qed.
+
+Lemma le_bottom_zero : bottom o<= zero.
+Proof.
+   unfold bottom, the_least, the; apply iota_ind.
+\end{Coqbad}
+
+\begin{Coqanswer}
+2 subgoals (ID 413)
+
+  ============================
+  exists ! x : Ord, least_member lt (Empty_set Ord) x
+
+subgoal 2 (ID 414) is:
+ forall a : Ord, unique (least_member lt (Empty_set Ord)) a -> 
+      a o<= zero
+\end{Coqanswer}
+
+\begin{Coqbad}
+Abort.
+End Bad.
+\end{Coqbad}
+
+In short, using \texttt{epsilon} and \texttt{iota} in our implementation of countable ordinals after Schütte has two main advantages.
+
+\begin{itemize}
+\item It allows us to give a \emph{name} (using \texttt{Definition}) two witnesses 
+of existential quantifiers (let us recall that, in classical logic, one may consider non-constructive proofs of existential statements)
+\item By separating definitions from proofs of [unique] existence, one may make definitions  more concise and readable. Look for instance at 
+the definitions of  \texttt{zero}, \texttt{succ}, \texttt{plus}, etc. in the rest of this chapter.
+\end{itemize}
+%%%% ICI ICI
+
+\section{The  successor function}
+
+The definition of the function \texttt{succ:Ord -> Ord} is very concise. The successor of any ordinal $\alpha$ is the smallest ordinal strictly greater than $\alpha$.
+
+\label{Functions:succ-sch}
+
+\begin{Coqsrc}
+Definition succ (alpha : Ord) := the_least (fun beta => alpha o< beta).
+\end{Coqsrc}
+
+Using \texttt{succ}, we define the folloing predicates.
+
+\begin{Coqsrc}
+Definition is_succ (alpha:Ord) := exists beta, alpha = succ beta.
+
+Definition is_limit (alpha:Ord) := alpha <> zero /\ ~ is_succ alpha.
+\end{Coqsrc}
+
+It is also easy to define recursively the finite ordinals.
+
+\label{sect:notation-F-sch}
+
+\begin{Coqsrc}
+Reserved Notation "'F' n" (at level 29) .
+
+Fixpoint finite (i:nat) : Ord :=
+  match i with 
+            | 0 => zero
+            | S i => succ (F i)
+  end
+where "'F' i" := (finite i)  : schutte_scope.
+
+Coercion finite : nat >-> Ord.
+\end{Coqsrc}
+
+% \begin{remark}
+% Please look at remark~\vref{warning:coercions}.
+% \end{remark}
+
+How do we prove properties of the successor function?
+First, we make its specification explicit.
+
+\begin{Coqsrc}
+Definition succ_spec (alpha:Ord) :=
+  least_member   lt (fun z => alpha o< z).
+\end{Coqsrc}
+
+Then, we prove that our function \texttt{succ} meets this specification.
+
+\begin{Coqsrc}
+Lemma succ_ok : forall alpha,  succ_spec alpha  (succ alpha).
+Proof.
+  intros; unfold succ, the_least, the;  apply iota_spec.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+1 subgoal (ID 172)
+
+  alpha : Ord
+  ============================
+  exists ! x : Ord, succ_spec alpha x
+\end{Coqanswer}
+
+We have now to prove that the set of all ordinals strictly greater than $\alpha$ has a unique least element.
+But the singleton set $\{\alpha\}$ is countable, hence  bounded (by the axiom \texttt{AX3}).
+Hence; the set $\{\beta\in\mathbb{O}|\alpha < \beta\}$ is not empty
+and therefore has a unique least element.
+
+The \coq{} proof script is quite short.
+
+\begin{Coqsrc}
+  destruct (@AX3 (Singleton _ alpha)).
+  - apply countable_singleton.
+  -  unfold succ_spec; apply the_least_unicity;  exists x; intuition.
+Qed.     
+\end{Coqsrc}
+
+We can ``uncap'' the description operator for proving properties of the
+\texttt{succ} function.
+
+\begin{Coqsrc}
+Lemma lt_succ (alpha : Ord) :  alpha o< succ alpha.
+Proof.
+  destruct  (succ_ok  alpha);  tauto.
+Qed.
+
+Hint Resolve lt_succ : schutte.
+
+Lemma lt_succ_le (alpha beta : Ord):
+  alpha o< beta -> succ alpha o<= beta.
+Proof with eauto with schutte.
+  intros  H;  pattern (succ alpha); apply the_least_ok ... 
+  exists (succ alpha); red;apply lt_succ ...
+Qed.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma lt_succ_le_2 (alpha beta : Ord):
+  alpha o< succ beta -> alpha o<= beta.
+
+Lemma succ_mono (alpha beta : Ord):
+  alpha o< beta -> succ alpha o< succ beta.
+
+Lemma succ_monoR (alpha beta : Ord) :
+ succ alpha o< succ beta -> alpha o< beta.
+
+Lemma lt_succ_lt (alpha beta : Ord) :
+  is_limit beta ->  alpha o< beta -> succ alpha o< beta.
+\end{Coqsrc}
+
+\section{The definition of \texttt{omega}}
+In order to define $\omega$, the first infinite ordinal, we use an operator which
+``returns'' the least upper bound (if it exists) of a subset $X\subseteq \mathbb{O}$.
+For that purpose, we first use a predicate:
+(\texttt{is\_lub $D$ \textit{lt} $X$ $a$}) if $a$ belongs to $D$ and is the least 
+upper bound  of $X$ (with respect to \textit{lt}).
+
+\begin{Coqsrc}
+Definition is_lub (M:Type)
+                  (D : Ensemble M)
+                  (lt : relation M)
+                  (X:Ensemble M)
+                  (a:M) :=
+   In _ D a  /\ upper_bound  D lt X a  /\
+   (forall y, In _ D y -> upper_bound  D lt X y  -> 
+                  y = a \/ lt a y).
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Definition sup_spec X lambda := is_lub ordinal lt X lambda.
+
+Definition sup (X: Ensemble Ord) : Ord  := the  (sup_spec X).
+
+Notation "'|_|' X" := (sup X) (at level 29) : schutte_scope.
+\end{Coqsrc}
+
+Then, we define the function \texttt{omega\_limit} which returns the least upper bound 
+of the  (denumerable) range of any sequence \texttt{s: nat -> Ord}. 
+By \texttt{AX3} this range is bounded, hence the set of its upper bounds is not empty and has a least element.
+
+\begin{Coqsrc}
+Definition omega_limit (s:nat->Ord) : Ord 
+  := |_| (seq_range s).
+\end{Coqsrc}
+
+Then we define \texttt{omega} as the limit of the sequence of finite ordinals.
+
+\label{sect:notation-omega}
+\begin{Coqsrc}
+Definition _omega := omega_limit finite.
+
+Notation "'omega'" := (_omega) : schutte_scope.
+\end{Coqsrc}
+
+Among the numerous properties of the ordinal $\omega$, les us quote the following ones
+(proved in Module 
+\href{../src/html/hydras.Schutte.Schutte_basics.html\#finite_lt_omega}{\texttt{Schutte.Schutte\_basics}})
+
+\begin{Coqsrc}
+Lemma finite_lt_omega : forall i: nat,  i o< omega.
+
+Lemma lt_omega_finite alpha : Ord) : 
+  alpha o< omega ->  exists i:nat, alpha =  i.
+
+Lemma is_limit_omega : is_limit omega.
+\end{Coqsrc}
+
+\subsection{Ordering functions and ordinal addition}
+
+After having defined the finite ordinals and the infinite ordinal $\omega$, we  define the sum $\alpha+\beta$ of two countable ordinals.
+Schütte's definition looks like the following one:
+
+\begin{quote}
+``$\alpha+\beta$ is the $\beta$-th ordinal greater than or equal to $\alpha$''
+\end{quote}
+
+The purpose of this section is to give a meaning to the construction
+``the $\alpha$-th element of $X$''  where $X$ is any non-empty subset of $\mathbb{O}$.
+We follow Schütte's approach, by defining the notion of \emph{ordering functions},
+a way to associate a unique ordinal to each element of $X$.
+Complete definitions and proofs can be found in Module
+ \href{../src/html/hydras.Schutte.Ordering_Functions.html}%
+{\texttt{Schutte.Ordering\_Functions}} ).
+
+\subsection{Definitions}
+
+A \emph{segment} is a set $A$ of ordinals such that, whenever  $\alpha\in A$ and
+$\beta<\alpha$, then $\beta\in A$; a segment is  \emph{proper} if it strictly included in $\mathbb{O}$.
+
+\begin{Coqsrc}
+ Definition segment (A: Ensemble Ord) :=
+  forall alpha beta, In A alpha -> beta o< alpha -> In A  beta.
+
+Definition proper_segment (A: Ensemble Ord) :=
+  segment A /\ ~ Same_set A ordinal.
+\end{Coqsrc}
+
+Let  $A$ be a segment, and $B$ a subset of $\mathbb{O}$ : an \emph{ordering function for $A$ and  $B$} is a strictly increasing bijection from $A$ to $B$.
+The set $B$ is said to be an \emph{ordering segment} of $A$.
+Our definition in \coq{} is a direct translation of the mathematical text of~\cite{schutte}.
+
+\index{Maths!Ordering functions}
+
+\begin{Coqsrc}
+Definition ordering_function (f : Ord -> Ord)(A B : Ensemble Ord) :=
+ segment A /\
+ (forall a, In A a -> In B (f a)) /\
+ (forall b, B b -> exists a, In A a /\ f a = b) /\
+ forall a b, In A a -> In A b -> a o< b ->  f a o< f b.
+
+Definition ordering_segment (A B : Ensemble Ord) :=
+  exists f : Ord -> Ord, ordering_function f A B.
+\end{Coqsrc}
+
+We are now able to associate with any subset $B$ of $\mathbb{O}$ its ordering segment and ordering function.
+
+\begin{Coqsrc}
+Definition the_ordering_segment (B : Ensemble Ord) :=
+  the  (fun x => ordering_segment x B).
+
+Definition ord  (B : Ensemble Ord) := 
+  some (fun f => ordering_function f (the_ordering_segment B) B).
+\end{Coqsrc}
+
+Thus (\texttt{ord $B \;\alpha$}) is the $\alpha$-th element of $B$.
+Please note that the last definition uses the epsilon-based operator \texttt{some} and
+not \texttt{the}. This is due to the fact that we cannot prove the unicity (w.r.t. Leibniz' equality) of the ordering function of a given set. 
+By contrast, we admit the axiom  \texttt{Extensionality\_Ensembles}, from the library 
+\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Sets.Ensembles.html}{Coq.Sets.Ensembles}, so we use the operator \texttt{the} in the definition of
+\texttt{the\_ordering\_segment}.
+
+One of the main theorems of
+\href{../src/html/hydras.Schutte.Ordering_Functions.html\#ordering_function_ex}%
+{\texttt{Ordering\_Functions}} 
+associates a unique segment and a unique (up to extensionality) ordering function to every subset $B$ of $\mathbb{O}$.
+
+\begin{Coqsrc}
+About ordering_function_ex.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+forall B : Ensemble Ord,
+ exists ! S : Ensemble Ord, 
+      exists f : Ord -> Ord, ordering_function f S B
+\end{Coqanswer}
+
+\begin{Coqanswer}
+ordering_function_unicity :
+forall (B S1 S2 : Ensemble Ord) (f1 f2 : Ord -> Ord),
+ordering_function f1 S B ->
+ordering_function f2 S2 B -> 
+fun_equiv f1 f2 S1 S2
+\end{Coqanswer}
+
+Thus,  our function \texttt{ord}  which enumerates the elements of $B$ is defined in a non-ambiguous way.
+Let us quote the following theorems (see Library
+\href{../src/html/hydras.Schutte.Ordering_Functions.html}%
+{\texttt{Schutte.Ordering\_Functions}} for more details).
+
+\begin{Coqsrc}
+Theorem ordering_le : forall f A B,
+    ordering_function f A B ->
+    forall alpha, In A alpha -> alpha  o<= f alpha.
+
+Th_13_5_2 :
+forall (A B : Ensemble Ord) (f : Ord -> Ord),
+ordering_function f A B -> Closed B -> continuous f A B
+\end{Coqsrc}
+
+\subsection{Ordinal addition}
+
+We are now ready to define and study addition on the type \texttt{Ord}.
+The following definitions and proofs can be consulted in Module
+\href{../src/html/hydras.Schutte.Addition.html}%
+{\texttt{Schutte.Addition.v}}.
+
+\begin{Coqsrc}
+Definition plus alpha := ord  (ge alpha).
+Notation "alpha + beta " := (plus alpha beta) : schutte_scope.
+\end{Coqsrc}
+
+In other words,  $\alpha + \beta$ is the  $\beta$-th ordinal greater than or equal to $\alpha$. 
+Thanks to generic properties of ordering functions, we can show the following 
+properties of addition on $\mathbb{O}$. First, we prove a useful lemma:
+
+\begin{Coqsrc}
+Lemma plus_elim (alpha : Ord) :
+  forall P : (Ord->Ord)->Prop,
+    (forall f: Ord->Ord, 
+        ordering_function f ordinal (ge alpha)-> P f) ->
+    P (plus alpha).
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma alpha_plus_zero (alpha: Ord): alpha + zero = alpha.
+Proof.
+ pattern  (plus alpha); apply plus_elim; eauto.
+ \end{Coqsrc}
+
+ \begin{Coqanswer}
+ 1 subgoal (ID 24)
+
+  alpha : Ord
+  ============================
+  forall f : Ord -> Ord,
+  ordering_function f ordinal (ge alpha) -> 
+  f zero = alpha
+ \end{Coqanswer}
+
+ \begin{Coqsrc}
+ (* rest of proof skipped *)
+ \end{Coqsrc}
+
+The following lemmas are proved the same way.
+
+ \begin{Coqsrc}
+
+Lemma zero_plus_alpha (alpha : Ord) : zero + alpha = alpha.
+
+Lemma le_plus_l (alpha beta : Ord) : alpha o<= alpha + beta.
+
+Lemma le_plus_r (alpha beta : Ord) :  beta o<= alpha + beta.
+
+Lemma plus_mono_r (alpha beta gamma : Ord) : 
+    beta o< gamma -> alpha + beta o< alpha + gamma.
+
+Lemma plus_of_succ (alpha beta : Ord) :
+    alpha + (succ beta) = succ (alpha + beta).
+
+Theorem plus_assoc (alpha beta gamma : Ord) :
+  alpha + (beta + gamma) = (alpha + beta) + gamma.
+
+Lemma one_plus_omega :  1 + omega = omega.
+
+Lemma finite_plus_infinite (n : nat) (alpha : Ord) :
+  omega o<= alpha -> n + alpha = alpha.
+\end{Coqsrc}
+
+It isinteresting to compare the proof of these lemmas with the 
+computational proofs of the corresponding statements in Module
+\href{../src/html/hydras.Epsilon0.T1.html}%
+{\texttt{Epsilon0.T1}}. 
+For instance, the proof of the lemma 
+\texttt{one\_plus\_omega} uses the continuity of ordering functions (applied to  \texttt{(plus 1)}) and compares the limit of the $\omega$-sequences $i_{(i \in \mathbb{N})}$ and
+$(1+i)i_{(i \in \mathbb{N})}$, whereas in the library  \texttt{Epsilon0/T1}, the equality 
+$1+\omega=\omega$ is just proved with \texttt{reflexivity}!
+
+\subsubsection{Multiplication by a natural number}
+
+The multiplication of an ordinal by a natural number is defined in terms of addition.
+This operation is useful for the study of Cantor normal forms.
+
+\begin{Coqsrc}
+Fixpoint mult_Sn (alpha:Ord)(n:nat){struct n} :Ord :=
+ match n with 
+            | 0 => alpha
+            | S p => mult_Sn alpha p + alpha
+ end.
+
+Definition mult_n alpha n :=
+  match n with
+      0 => zero
+    | S p => mult_Sn alpha p
+  end.
+
+Notation "alpha * n" := (mult_n alpha n) : schutte_scope.
+\end{Coqsrc}
+
+\section{The exponential of basis \texorpdfstring{$\omega$}{omega}}
+
+In this section, we define the function which maps any $\alpha\in\mathbb{O}$ to
+the ordinal  $\omega^\alpha$, also written 
+$\varphi_0(\alpha)$. 
+It is an opportunity to apply the definitions and results of the preceding section. 
+Indeed,  Schütte first defines a subset of $\mathbb{O}$: the set of additive principal ordinals, and $\varphi_0$  is just defined as the ordering function of this set.
+
+\subsection{Additive principal ordinals}
+
+\index{Maths!Additive principal ordinals}
+
+\begin{definition}
+A non-zero ordinal  $\alpha$ is said to be \emph{additive principal} if, for all  $\beta<\alpha$, $\beta+\alpha$ is equal to  $\alpha$.
+We call \texttt{AP} the set of additive principal ordinals.
+
+\end{definition}
+
+\noindent\emph{From Module \href{../src/html/hydras.Schutte.AP.html}%
+{\texttt{Schutte.AP}}}
+
+\begin{Coqsrc}
+Definition AP : Ensemble Ord :=
+  fun alpha => 
+  zero o< alpha /\
+  (forall beta, beta o< alpha ->  beta + alpha = alpha).
+\end{Coqsrc}
+
+\subsection{The function \texttt{phi0}}
+
+Let us call  $\varphi_0$ the ordering function of \texttt{AP}.
+In the mathematical text, we shall use indifferently the notations  $\omega^\alpha$ and$\varphi_0(\alpha)$.
+
+\begin{Coqsrc}
+Definition phi0 := ord AP.
+
+Notation "'omega^'" := phi0 (only parsing) : schutte_scope.
+\end{Coqsrc}
+
+\subsection{Omega-towers and the ordinal \texorpdfstring{$\epsilon_0$}{epsilon0}}
+
+Using $\varphi_0$, we can define recursively the set of finite omega-towers.
+
+\begin{Coqsrc}
+Fixpoint omega_tower (i : nat) : Ord :=
+  match i with
+    0 =>  1
+  | S j => phi0 (omega_tower j)
+  end.
+\end{Coqsrc}
+
+Then, the ordinal  $\epsilon_0$ is defined as the limit of the sequence of all finite towers (a kind of infinite tower).
+
+\begin{Coqsrc}
+Definition epsilon0 := omega_limit omega_tower.
+\end{Coqsrc}
+
+The rest of our library \texttt{AP} is devoted to the proof of properties of additive principal ordinals, hence of the ordering function  $\varphi0$ and the ordinal $\epsilon_0$ (which we could not express within the type \texttt{T1}).
+
+\subsection{Properties of the set  \texttt{AP}}
+
+The set of additive principal ordinals is not empty: it contains at least the ordinals  $1$ and  $\omega$.
+
+\begin{Coqsrc}
+Lemma AP_one : In AP 1.
+
+Lemma AP_omega : In AP omega.
+\end{Coqsrc}
+
+Moreover, $1$ is the least principal ordinal and $\omega$ is the second element of
+\texttt{AP}.
+
+\begin{Coqsrc}
+Lemma least_AP: least_member  lt AP 1.
+
+Lemma omega_second_AP :
+  least_member   lt 
+                  (fun alpha => 1 o< alpha /\ In AP alpha)
+                  omega.
+\end{Coqsrc}
+
+The set  \texttt{AP} is  \emph{closed} under addition, and unbounded.
+
+\begin{Coqsrc}
+Lemma AP_plus_closed (alpha beta gamma : Ord): 
+     In AP alpha -> beta o< alpha -> gamma o< alpha -> beta + gamma o< alpha.
+
+Theorem AP_unbounded : Unbounded AP.
+\end{Coqsrc}
+
+Finally, \texttt{AP} is (topologically) \emph{closed} and ordered by the segment of all countable ordinals.
+
+\index{Predicates!Closed}
+
+\begin{Coqsrc} 
+Definition Closed (B : Ensemble Ord) : Prop := 
+  forall M, Included M B -> Inhabited _ M -> 
+                 countable M -> In B (|_| M).
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Theorem AP_closed : Closed AP.
+
+Lemma AP_o_segment :  the_ordering_segment AP = ordinal.
+\end{Coqsrc}
+
+\subsubsection{Properties of the function \texorpdfstring{$\varphi_0$}{phi0}}
+
+The ordering function $\varphi_0$ of the set \texttt{AP} is defined on the full set $\mathbb{O}$ and is continuous (Schütte says that this function is \emph{normal}).
+
+\begin{Coqsrc}
+Theorem normal_phi0 : normal phi0 AP.
+\end{Coqsrc}
+
+The following properties come from  the definition of $\varphi_0$ as the ordering function of \texttt{AP}.
+It may be interesting to compare these proofs with the computational ones described in Chapter ~\ref{chap:T1}.
+
+\begin{Coqsrc}
+Lemma AP_phi0 (alpha : Ord) : In AP (phi0 alpha).
+
+Lemma phi0_zero : phi0 zero =  1.
+
+Lemma phi0_mono (alpha beta : Ord) :
+  alpha o< beta ->  phi0 alpha o< phi0 beta.
+
+Lemma phi0_inj (alpha beta : Ord) :
+    phi0 alpha = phi0 beta -> alpha = beta.
+
+Lemma phi0_sup : forall (U: Ensemble Ord),
+   Inhabited _ U ->   countable U ->  phi0 (|_| U) = |_| (image U phi0).
+
+Lemma is_limit_phi0 (alpha : Ord) :
+  zero o< alpha ->  is_limit (phi0 alpha).
+
+Lemma omega_eq : omega = phi0 1.
+
+Lemma phi0_le (alpha : Ord) : alpha o<= phi0 alpha.
+\end{Coqsrc}
+
+Please note that the lemma \texttt{omega\_eq} above, is consistent with the interpretation of the ordering function $\varphi_0$ as the exponential of basis $\omega$.
+Indeed we could have written this lemma with our alternative notation:
+
+\begin{Coqsrc}
+ Lemma omega_eq : omega = omega^ 1.
+\end{Coqsrc}
+
+\section{More about \texorpdfstring{$\epsilon_0$}{\texttt{epsilon0}}}
+
+Let us recall that the limit ordinal  $\epsilon_0$ cannot be written within the type \texttt{T1}.
+Since we are now considering the set of all countable ordinals, we can now prove some properties of this ordinal.
+
+We prove the inequality  $\alpha<\omega^\alpha$ whenever $\alpha < \epsilon_0$.
+\emph{Note that this condition was implicit in Module \href{../src/html/hydras.Epsilon0/T1.html\#lt_phi0}%
+{\texttt{Epsilon0.T1}}.}
+
+\begin{Coqsrc}
+Lemma lt_phi0 (alpha : Ord):
+  alpha o< epsilon0 -> alpha o< phi0 alpha.
+\end{Coqsrc}
+
+The proof is as follows:
+\begin{enumerate}
+\item Since $\alpha<\epsilon_0$, consider the least $i$ such that $\alpha$ is strictly less than the omega-tower of height $i$.
+\item
+  \begin{itemize}
+  \item If $i=0$, then the result is trivial (because $\alpha=0$)
+ \item  Otherwise let $i=j+1$; 
+          $\alpha$ is greater than or equal to the omega-tower of height $j$.
+         By monotonicity,  $\varphi_0(\alpha)$ is greater than or equal to 
+        the omega-tower of height $j+1$, thus strictly greater than $\alpha$
+  \end{itemize}
+ \end{enumerate}
+
+Moreover,  $\epsilon_0$ is the least ordinal $\alpha$ that verifies the equality 
+$\alpha = \omega^\alpha$, in other words the least fixpoint of the function  $\varphi_0$.
+
+\begin{Coqsrc}
+Theorem epsilon0_lfp : least_fixpoint lt phi0 epsilon0.
+\end{Coqsrc}
+
+\section{Critical ordinals}
+
+\index{Maths!Critical ordinals}
+
+For any  (countable) ordinal $\alpha$, the set $\textit{Cr}(\alpha)$ is inductively defined 
+as follows by Schütte (p.81 of~\cite{schutte}).
+
+\begin{quote}
+  \begin{itemize}
+  \item $\textit{Cr}(0)$ is the set \textit{AP} of additive principal ordinals.
+  \item If $0<\alpha$, then $\textit{Cr}(\alpha)$ is the intersection of all the sets of fixpoints of the $\textit{Cr}(\beta)$ for $\beta<\alpha$.
+  \end{itemize}
+\end{quote}
+
+This definition is translated in \coq{} in 
+Module \href{../src/html/hydras.Schutte.Critical.html}%
+{\texttt{Schutte.Critical}}, as the least fixpoint of a functional.
+
+\begin{Coqsrc}
+Definition Cr_fun : forall alpha : Ord,
+       (forall beta : Ord, beta o< alpha -> Ensemble Ord) ->
+        Ensemble Ord 
+:= 
+   fun (alpha :Ord)
+        (Cr : forall beta, 
+                beta o< alpha -> Ensemble Ord) 
+        (x : Ord) => (
+       (alpha = zero /\ AP x) \/
+       (zero o< alpha /\
+        forall beta (H:beta o< alpha),
+          the_ordering_segment (Cr beta H) x /\ ord (Cr  beta H) x = x)).
+
+Definition Cr (alpha : Ord) : Ensemble Ord := 
+    (Fix  all_ord_acc (fun (_:Ord) => Ensemble Ord) Cr_fun) alpha.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Definition phi (alpha : Ord) : Ord -> Ord 
+    :=  ord (Cr alpha).
+
+Definition A (alpha : Ord) : Ensemble Ord :=
+  the_ordering_segment (Cr alpha).
+\end{Coqsrc}
+
+For instance,  we prove that $\textit{Cr}(0)$ is the set of additive principals and that $\epsilon_0$
+belongs to $\textit{Cr}(1)$.
+
+\begin{Coqsrc}
+Lemma Cr_zero_AP :  Cr 0 = AP
+
+Lemma epsilon0_Cr1 : In (Cr 1) epsilon0.
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+ Prove that $\epsilon_0$ is the least element of $\textit{Cr}(1)$.
+\end{exercise}
+
+\subsection{A flavor of infinity}
+
+The family of the $\textit{Cr}(\alpha)$s is made of infinitely many unbounded (hence infinite) sets.
+Let us quote Lemma 5, p. 82  of~\cite{schutte}:
+\begin{quote}
+  For all $\alpha$, the set $\textit{Cr}(\alpha)$ is closed (for the least upper bound of non-empty countable sets) and unbounded.
+\end{quote}
+
+We prove this result by transfinite induction on $\alpha$ of both properties.
+
+The proof is still quite long, by transfinite induction over $\alpha$.
+
+\index{Coq!Techniques!Simultaneous induction}
+\index{Coq!Techniques!Transfinite induction}
+
+\begin{Coqsrc}
+Section Proof_of_Lemma5.
+  Let P (alpha:Ord) := Unbounded (Cr alpha) /\ Closed (Cr alpha).
+
+ Lemma Lemma5 : forall alpha, P alpha.
+(* ... *)
+ End Proof_of_Lemma5.
+
+Corollary Unbounded_Cr alpha : Unbounded (Cr alpha).
+Proof.
+  now destruct (Lemma5 alpha).
+Qed.
+
+Corollary Closed_Cr alpha : Closed (Cr alpha).
+Proof.
+  now destruct (Lemma5 alpha).
+Qed.
+\end{Coqsrc}
+
+\section{Cantor normal form}
+
+The notion of Cantor normal form is defined for all countable ordinals.
+Nevertheless, note that, contrary to the implementation based on type \texttt{T1},
+the Cantor normal form of an ordinal $\alpha$ may contain $\alpha$ as a 
+sub-term\footnote{This would prevent us from trying to represent Cantor normal forms as finite trees (like in Sect.~\ref{sec:T1-inductive-def})}.
+
+We represent  Cantor normal forms as lists of ordinals.
+A  list $l$ is a Cantor normal form of a given ordinal $\alpha$ if it satisfies two conditions:
+
+\begin{itemize}
+\item The list  $l$ is sorted (in decreasing order) w.r.t. the order $\leq$
+\item The sum of all the  $\omega^{\beta_i}$ where the $\beta_i$ are the terms of $l$ (in this order) is equal to $\alpha$.
+\end{itemize}
+
+\vspace{4pt}
+
+\noindent\emph{From \href{../src/html/hydras.Schutte.CNF.html\#cnf_t}%
+{\texttt{Schutte.CNF}}}
+
+\begin{Coqsrc}
+ Definition cnf_t := list Ord.
+
+Fixpoint eval (l : cnf_t) : Ord :=
+  match l with nil => zero
+              | beta :: l' => phi0 beta + eval l'
+  end.
+
+Definition sorted (l: cnf_t) :=
+  LocallySorted (fun alpha beta => beta o<= alpha) l.
+
+Definition is_cnf_of (alpha : Ord)(l : cnf_t) : Prop :=
+  sorted l /\ alpha = eval l.
+\end{Coqsrc}
+
+\index{Maths!Transfinite induction}
+
+By transfinite induction on $\alpha$, we prove that every countable ordinal $\alpha$ 
+ has at least a Cantor normal form.
+
+\begin{Coqsrc}
+Theorem cnf_exists (alpha : Ord) :
+  exists l: cnf_t, is_cnf_of alpha l.
+\end{Coqsrc}
+
+By structural induction on lists, we prove that this normal form is unique.
+
+\begin{Coqsrc}
+ Lemma cnf_unicity : forall l alpha, 
+   is_cnf_of alpha l -> 
+   forall l',  is_cnf_of alpha l' -> l=l'.
+Proof.
+ induction l.
+ (*  ...  *)
+
+Theorem cnf_exists_unique (alpha:Ord) :
+  exists! l: cnf_t, is_cnf_of alpha l.
+\end{Coqsrc}
+
+Finally, the following two lemmas relate  $\epsilon_0$ with Cantor normal forms.
+
+If $\alpha<\epsilon_0$, then the Cantor normal form of $\alpha$ is made of ordinals strictly less than $\alpha$.
+
+\begin{Coqsrc}
+Lemma cnf_lt_epsilon0 : 
+ forall l alpha, 
+   is_cnf_of alpha l ->  alpha o< epsilon0 ->
+   Forall (fun beta =>  beta o< alpha) l.
+\end{Coqsrc}
+
+\index{Exercises}
+
+\begin{exercise}
+Please consider the following statement :
+
+\begin{Coqsrc}
+Lemma cnf_lt_epsilon0_iff : 
+ forall l alpha, 
+   is_cnf_of alpha l ->  
+   (alpha o< epsilon0 <->  Forall (fun beta =>  beta o< alpha) l).
+\end{Coqsrc}
+
+Is it true ?
+\end{exercise}
+
+Finally, the Cantor normal form of $\epsilon_0$ is just $\omega^{\epsilon_0}$.
+
+\begin{Coqsrc}
+Lemma cnf_of_epsilon0 : is_cnf_of epsilon0 (epsilon0 :: nil).
+Proof.
+  split.
+  - constructor.  
+  - simpl;  now rewrite alpha_plus_zero, epsilon0_fxp.
+Qed.
+\end{Coqsrc}
+
+\index{Projects}
+
+\begin{project}
+Implement pages 82 to 85 of~\cite{schutte} (critical, strongly critical, maximal critical ordinals, Feferman's ordinal $\Gamma_0$).
+\end{project}
+
+\begin{remark}
+The sub-directory \href{../src/html/hydras.Gamma0.html}%
+{\texttt{src/Gamma0}} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form. 
+\end{remark}
+
+\section{An embedding of \texttt{T1} into \texttt{Ord}}
+
+Our library 
+\href{../src/html/hydras.Schutte.correctness_E0.html}%
+{\texttt{Schutte.Correctness\_E0}} establishes the link between two very different modelizations of ordinal numbers. In other words, it ``validates'' a data structure in terms of
+a classical mathematical discourse considered as a model. 
+First, we define a function from \texttt{T1} into  \texttt{Ord} by structural recursion.
+
+\begin{Coqsrc}
+Fixpoint inject (t:T1) : Ord :=
+ match t with 
+     | T1.zero => zero
+     | T1.ocons a n b =>  AP.phi0 (inject a) * S n + inject b
+ end.  
+\end{Coqsrc}
+
+This function enjoys good commutation properties with respect to the main operations which
+allow us to build Cantor normal form.
+
+\begin{Coqsrc}
+Theorem inject_of_zero : inject T1.zero = zero.
+
+Theorem inject_of_finite (n : nat):
+  inject (T1.fin n) =  n.
+
+Theorem inject_of_phi0 (alpha : T1):
+  inject (phi0 alpha) = AP.phi0 (inject alpha).
+
+Theorem inject_plus (alpha beta : T1): nf alpha -> nf beta ->
+  inject (alpha + beta)%t1 = inject alpha + inject beta.
+
+Theorem inject_mult_n (alpha : T1)  :
+  nf alpha -> forall n:nat , inject (alpha *  n)%t1 =  inject alpha * n.
+
+Theorem inject_mono (beta gamma : T1) :
+  T1.lt  beta gamma -> 
+  T1.nf beta -> T1.nf gamma -> 
+  inject beta o< inject gamma.
+
+Theorem inject_injective (beta gamma : T1) : nf beta -> nf gamma ->
+  inject beta = inject gamma -> beta = gamma.
+\end{Coqsrc}
+
+Finally, we prove that \texttt{inject} is a bijection from the set of all terms of \texttt{T1} in normal form to the set 
+\texttt{members epsilon0} of the elements of \texttt{Ord} strictly less than  $\epsilon_0$.
+
+\begin{Coqsrc}
+ Theorem inject_lt_epsilon0 (alpha : T1):
+      inject alpha o< epsilon0.
+
+Theorem embedding : 
+     fun_bijection (nf: Ensemble T1)  (members epsilon0) inject.
+ \end{Coqsrc}
+\subsection{Remarks}
+Let us recall that the library \href{../src/html/hydras.Schutte.Schutte.html}%
+{\texttt{Schutte}} depends on five \emph{axioms} and lies explicitly in the  
+framework of classical logic with a weak version of the axiom of choice
+(please look at the documentation of
+\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.ChoiceFacts.html}{\texttt{Coq.Logic.ChoiceFacts}}).
+Nevertheless, the other modules:
+\href{../src/html/hydras.Epsilon0.Epsilon0.html}%
+{\texttt{Epsilon0}},
+\href{../src/html/hydras.Hydra.Hydra.html}%
+{\texttt{Hydra}}, et 
+\href{../src/html/hydras.Gamma0.Gamma0.html}%
+{\texttt{Gamma0}}
+do not import any axioms and are really constructive.
+
+\index{Projects}
+\begin{project}
+There is no construction of ordinal multiplication in~\cite{schutte}. 
+It would be interesting to derive this operation from Schütte's axioms,
+and prove its consistence with multiplication in ordinal notations for 
+$\epsilon_0$ and $\Gamma_0$.
+\end{project}
+
+\section{Related work}
+
+In~\cite{grimm:hal-00911710}, José Grimm establishes the consistency between or ordinal notations (\texttt{T1} and \texttt{T2} (Veblen normal form) and his implementation
+of ordinal numbers after Bourbaki's set theory.
 

### 2020-09-24 doc/hydras.tex
minor simplifications
merge omegaomega into master
better name consistency in selectors of types ON and SubON

@@
 
-\author{Pierre Castéran\\ LaBRI, Univ. Bordeaux, CNRS UMR 5800, Bordeaux-INP\\with contributions by \'Evelyne Contejean and Florian Hatat}
+\author{Pierre Castéran\\ LaBRI, Univ. Bordeaux, CNRS UMR 5800, Bordeaux-INP\\with contributions by \'Evelyne Contejean,  Florian Hatat and Pascal Manoury}
 \date{\today}
@@
 
-Definition on_compare {A:Type}{lt: relation A}
+Definition ON_compare {A:Type}{lt: relation A}
             {compare : A -> A -> comparison}
@@
 
-Definition on_lt {A:Type}{lt: relation A}
+Definition ON_lt {A:Type}{lt: relation A}
            {compare : A -> A -> comparison}
            {on : ON lt compare} := lt.
-Infix "o<" := on_lt : ON_scope.
-
-Definition on_le  {A:Type}{lt: relation A}
+Infix "o<" := ON_lt : ON_scope.
+
+Definition ON_le  {A:Type}{lt: relation A}
            {compare : A -> A -> comparison}
            {on : ON lt compare} :=
-  clos_refl _ on_lt.
-
-Infix "o<=" := on_le : ON_scope.
+  clos_refl _ ON_lt.
+
+Infix "o<=" := ON_le : ON_scope.
 \end{Coqsrc}
@@
 \begin{scope}[color=red]
-\node(B0) at (14,0)[label=below:$0$,label=above:$\omega$]{$\bullet$};
+\node(B0) at (14,0)[label=below:$0$,label=above:\textcolor{red}{$\omega$}]{$\bullet$};
 \node(B1) at (16,0)[label=below:$1$, label=above:$\omega+1$]{$\bullet$};
@@
 Applying the lemma \texttt{limit\_is\_lub}, we get  the inequality
-$\iota(i+1,0)\geq (i+1,0)$
+$m(\iota(i+1,0))\geq (i+1,0)$
 
@@
   {
-  subon_compare: forall x y : A,  compareB (iota x) (iota y) =
+  SubON_compare: forall x y : A,  compareB (iota x) (iota y) =
                                  compareA x y;
-  subon_incl : forall x, ltB (iota x) alpha;
-  subon_onto : forall y, ltB y alpha  -> exists x:A, iota x = y}.
+  SubON_incl : forall x, ltB (iota x) alpha;
+  SubON_onto : forall y, ltB y alpha  -> exists x:A, iota x = y}.
 \end{Coqsrc}
@@
 Proof.
-  refine (@mkord T1.zero _);  now compute. 
+  now exists T1.zero.
 Defined.
@@
 Instance _Omega : E0.
-Proof.  now exists omega. Defined.
+Proof.  now exists omega%t1. 
+Defined.
 

### 2020-09-25 doc/hydras.tex
various infix operators for ordinal inequalities

@@
 \begin{todo}
-Infix notation for inequalities on ordinals have changed to \texttt{o<} and \texttt{o<=}.  This document is not updated yet, but the \coq{} scripts are correct.
+Infix notation for inequalities on type \texttt{T1} have changed to \texttt{t1<} and \texttt{t1<=}.  This document is not updated yet, but the \coq{} scripts are correct.
 \end{todo}
@@
 
+\index{Exercises}
+\begin{exercise}
+In earlier versions of this development, the predicate \texttt{nf} was defined  inductively, with various constructors describing all possible cases.
+\begin{enumerate}
+\item Please give such a definition, in a dedicated module.
+\item Prove the logical equivalence between your definition and ours.
+\item Define a variant of the type \texttt{E0} (with your definition of \texttt{nf}).
+\item Can you still prove a lemma like \texttt{E0\_eq\_iff} ? With the help of an axiom of the standard library ?
+\end{enumerate}
+\end{exercise}
 For upgrading constants and fonctions of \texttt{T1}, we have to prove that 
@@
 The following lemma establishes the connection between the  function
-\texttt{succ} and the predicate \texttt{is\_succ}.
+\texttt{succ} and the Boolean predicate \texttt{succb}.
 
@@
 
-Example e'2 : omega o< omega + 6.
+Example e'2 : omega t1< omega + 6.
 Proof. now compute. Qed.
@@
 
-Example e'''2 : omega o< omega * 6.
+Example e'''2 : omega t1< omega * 6.
 Proof. now compute. Qed.
@@
 Lemma plus_not_monotonous : exists alpha beta gamma : T1,
-    alpha o< beta /\ alpha + gamma = beta + gamma.
+    alpha t1< beta /\ alpha + gamma = beta + gamma.
 Proof.
@@
 Lemma mult_not_monotonous :  exists alpha beta gamma : T1,
-      alpha o< beta /\ alpha * gamma = beta * gamma.
+      alpha t1< beta /\ alpha * gamma = beta * gamma.
 Proof.
@@
   delta : T1
-  Hdelta : delta o< ocons beta n gamma
+  Hdelta : delta t1< ocons beta n gamma
   ============================
@@
   alpha : T1
-  Aalpha : forall y : T1,  y o< alpha -> Acc LT y
+  Aalpha : forall y : T1,  y t1< alpha -> Acc LT y
   IHalpha : forall y : T1,
@@
 Lemma oplus_strict_mono_LT_l (alpha beta gamma : T1) :
-  nf gamma   -> alpha  o< beta ->
-  alpha o+ gamma o< beta o+ gamma.
+  nf gamma   -> alpha  t1< beta ->
+  alpha o+ gamma  t1< beta o+ gamma.
 
 Lemma oplus_strict_mono_LT_r (alpha beta gamma : T1) :
-  nf alpha -> beta o< gamma ->
-  alpha o+ beta o< alpha o+ gamma.
+  nf alpha -> beta t1< gamma ->
+  alpha o+ beta t1< alpha o+ gamma.
 \end{Coqsrc}
@@
 Lemma S0_decr :
-  forall s s', S0  s s' -> ms s' o< ms s.
+  forall s s', S0  s s' -> ms s' t1< ms s.
 \end{Coqsrc}
@@
 Lemma R1_decr : forall h h',
-                  R1 h h' -> m h' o< m h.
+                  R1 h h' -> m h' t1< m h.
 \end{Coqsrc}
@@
 Lemma S1_decr n:
-  forall s s', S1 n s s' -> ms s' o<  ms s.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma R2_decr n : forall h h', R2 n h h' -> m h'  o< m h.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma round_decr : forall h h', h -1-> h' -> m h' o< m h.
+  forall s s', S1 n s s' -> ms s' t1<  ms s.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma R2_decr n : forall h h', R2 n h h' -> m h'  t1< m h.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma round_decr : forall h h', h -1-> h' -> m h' t1< m h.
 Proof.
@@
   red; apply Inclusion.wf_incl with 
-         (R2 := fun h h' =>  m h o< m h').
+         (R2 := fun h h' =>  m h t1< m h').
    red; intros;  now apply round_decr.
@@
 Lemma canonS_LT i alpha :
-  nf alpha -> alpha <> zero -> canonS alpha i o<  alpha.
+  nf alpha -> alpha <> zero -> canonS alpha i t1<  alpha.
 \end{Coqsrc}
@@
      limitb lambda  ->
-     forall beta, beta o< lambda ->
-                  {i:nat | beta o< canonS lambda i}.
+     forall beta, beta t1< lambda ->
+                  {i:nat | beta t1< canonS lambda i}.
 
@@
 Lemma LT_path_to (alpha beta : T1) :
-  beta o< alpha -> {s : list nat | path_to beta s alpha}.
+  beta t1< alpha -> {s : list nat | path_to beta s alpha}.
 \end{Coqsrc}
@@
 Lemma path_to_LT beta s alpha :
-  path_to beta s alpha -> nf alpha -> beta o< alpha.
+  path_to beta s alpha -> nf alpha -> beta t1< alpha.
 \end{Coqsrc}
@@
 Lemma LT_to_battle alpha beta :
-    beta o< alpha ->  iota alpha -+-> iota beta.
+    beta t1< alpha ->  iota alpha -+-> iota beta.
 \end{Coqsrc}
@@
     Hvar: Hvariant T1_wf B m;
-    m_bounded: forall h, m h o< mu
+    m_bounded: forall h, m h t1< mu
   }.
@@
   Hypothesis m_decrease : forall  i h h',
-        round_n i h h' -> m h' o< m h.
+        round_n i h h' -> m h' t1< m h.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Lemma m_ge_0 alpha:  nf alpha -> alpha o<= m (iota alpha).
+Lemma m_ge_0 alpha:  nf alpha -> alpha t1<= m (iota alpha).
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-    Corollary m_ge_generic : m big_h o<= m small_h.
+    Corollary m_ge_generic : m big_h t1<= m small_h.
  \end{Coqsrc}
@@
 The (big) rest of the proof is dedicated to prove formally the converse inequality 
-\texttt{m small\_h o< m big\_h}.
+\texttt{m small\_h t1< m big\_h}.
 
@@
 \begin{Coqsrc}
-Lemma m_ge : m big_h o<= m small_h.
+Lemma m_ge : m big_h t1<= m small_h.
   Proof.
@@
 
-\item From the hypothesis \texttt{m\_bounded}, we have \texttt{m big\_h o< mu}
+\item From the hypothesis \texttt{m\_bounded}, we have \texttt{m big\_h t1< mu}
 \item By Lemma \texttt{LT\_to\_battle}, we get a (free) battle from
@@
 \begin{Coqsrc}
-Lemma m_lt : m small_h o< m big_h.
+Lemma m_lt : m small_h t1< m big_h.
 \end{Coqsrc}
@@
 Lemma LT_to_round_plus alpha beta :
-    beta o< alpha ->  iota alpha -+-> iota beta.
+    beta t1< alpha ->  iota alpha -+-> iota beta.
 \end{Coqsrc}
@@
 Corollary Cor12 (alpha : T1) :  nf alpha ->
-         forall beta i n, beta  o< alpha  ->
+         forall beta i n, beta  t1< alpha  ->
                 (i < n)%nat ->
@@
 Corollary Cor12' (alpha : T1) :  nf alpha ->
-         forall beta i n, (beta o< alpha)%t1 ->
+         forall beta i n, (beta t1< alpha)%t1 ->
                (i <= n)%nat ->
@@
 Lemma Lemma2_6_1 (alpha : T1) :  
-  nf alpha -> forall beta,  beta o< alpha  ->
+  nf alpha -> forall beta,  beta t1< alpha  ->
   {n:nat | const_pathS n alpha beta}.
@@
   (alpha beta : T1) (i : nat):
-  nf alpha -> const_pathS i alpha beta -> zero  o< alpha ->
+  nf alpha -> const_pathS i alpha beta -> zero  t1< alpha ->
   {l:nat | standard_path (S i) alpha j beta}.
@@
   Variables (alpha beta: T1) (n : nat).
-  Hypotheses (Halpha: nf alpha) (Hpos : zero o<  beta)
+  Hypotheses (Halpha: nf alpha) (Hpos : zero t1<  beta)
              (p : const_pathS n alpha  beta).
@@
   (alpha beta : T1) (i : nat):
-  nf alpha -> const_pathS i alpha beta -> zero  o< alpha ->
+  nf alpha -> const_pathS i alpha beta -> zero  t1< alpha ->
   {j:nat | standard_path (S i) alpha j beta}.
@@
 Corollary  LT_to_standard_path  (alpha beta : T1) :
-  beta o< alpha ->
+  beta t1< alpha ->
   {n : nat & {j:nat | standard_path (S n) alpha j beta}}.
@@
     forall alpha beta,
-      beta o< alpha ->
+      beta t1< alpha ->
       exists n i,  battle standard  n (iota alpha) i (iota beta).
@@
 \begin{Coqsrc}
- Lemma m_ge : m big_h o<= m small_h.
+ Lemma m_ge : m big_h t1<= m small_h.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Lemma m_lt : m small_h  o< m big_h.
+Lemma m_lt : m small_h  t1< m big_h.
 \end{Coqsrc}

### 2020-09-28 doc/hydras.tex
correcting typos

@@
 \usepackage{threeparttable}
-\setcounter{tocdepth}{2}
+\setcounter{tocdepth}{1}
 \input{coqmacros}
@@
 necessary. Since the development of a theory is represented as a bunch of computer files,
-everyone is able to read the proofs with an arbitrary level of detail, or to play with the theory by building alternate proofs or definitions.
+everyone is able to read the proofs with an arbitrary level of detail, or to play with the theory by writing alternate proofs or definitions.
 
@@
 
-\item We prove a relation between the length of a ``classic''  class of  battles
+\item We prove a relation between the length of a ``classic''  class of  battles \footnote{This class is also called \emph{standard} in this document (text and proofs).
+The \emph{replication factor} of the hydra is exactly $i$ at the $i$-th round of the battle (see Sect~\vref{sect:replication-def}).}
 and the Wainer-Hardy hierarchy of ``rapidly growing functions'' $H_\alpha$~\cite{Wainer1970}.
@@
 interest (by illustrating some tactic of proof pattern, for instance).
-We may use \coq's module system to make several proofs of a givent theorem co-exists in our libraries (see also Sect~\vref{sect:alt-proofs}).
-After some discussions of the pros and cons of each solution, we will choose to develop only one of them, leaving the others  as exercises or projects (i.e., big or difficult exercises).
+We may use \coq's module system to make several proofs of a given theorem co-exist in our libraries (see also Sect~\vref{sect:alt-proofs}).
+After some discussions of the pros and cons of each solution, we develop only one of them, leaving the others  as exercises or projects (i.e., big or difficult exercises).
 In order to discuss which assumptions are really needed for proving a theorem, we will also present 
@@
 But we found it interesting to make all the three proofs available, allowing the reader to compare their common structure and notice their technical differences.
-In particular, the proof of $(3)$ uses several non-trivial combinatorial properties of ordinal numbers~\cite{KS81}, whilst $(1)$ and $(2)$ use simple properties of $\mathbb{N}$ and $\mathbb{N}^2$.
+In particular, the proof of $(3)$ uses several non-trivial combinatorial properties of ordinal numbersup to $\epsilon_0$~\cite{KS81}, whilst $(1)$ and $(2)$ use simple properties of $\mathbb{N}$ and $\mathbb{N}^2$.
 
@@
 
-At any place of our development, you may use the command \texttt{Print Assumptions {\it ident}} in order to verify on which axiom the theorem {\it ident} may depend.
+At any place of our development, you may use the  \texttt{Print Assumptions {\it ident}} command in order to verify on which axiom the theorem {\it ident} may depend.
 The following example is extracted from
@@
 \subsection{Contributions}
-Many thanks to \'Evelyne Contejean and Florian Hatat for their contribution.
+Many thanks to \'Evelyne Contejean,  Florian Hatat and Pascal Manoury  for their contribution.
 \'Evelyne contributed libraries on the recursive path ordering (\emph{rpo}) for proving the well-foundedness of our representation of $\epsilon_0$ and $\Gamma_0$.
 Florian Hatat proved many useful lemmas on countable sets, which we used in our adaptation of Schütte's formalization of countable ordinals.
+Pascal Manoury is integrating the ordinal $\omega^\omega$ into our hierarchy of ordinal notations.
 
@@
 
-Any form of contribution  is welcome: correction of errors, improvement of
+Any form of contribution  is welcome: correction of errors (typos and more serious mistakes), improvement of
 Coq scripts, proposition of inclusion of new chapters, and generally any
@@
 Finally, we decided to include definitions or lemma statements, as well as tactics,  that lead to
-dead-ends or to too complex developments, with the following coloring.
+dead-ends or too complex developments, with the following coloring.
 Bad definitions 
@@
 \label{sec:orgheadline44}
+\label{sect:replication-def}
+
 A \emph{hydra battle} is a fight between Hercules and the Hydra. 
@@
 Then $h'$ is replaced by  $n+1$ of copies of \(h'\) which share the same root $r$.
- The \emph{replication number} $n$ may be different (and generally is)   at each round of the fight.
+ The \emph{replication factor} $n$ may be different (and generally is)   at each round of the fight.
 It may be chosen by the hydra, according to its strategy, or imposed by some 
@@
 In many presentations of hydra battles, this number is increased by $1$ at each round.
-In the following presentation, we will also consider battles where the hydra is free to chose its number of replication at every round of the battle\footnote{Let us recall that, if the chopped-off head was at distance 1 from the foot, the replication factor is meaningless.}.
+In the following presentation, we will also consider battles where the hydra is free to choose its ~replication factor at each round of the battle\footnote{Let us recall that, if the chopped-off head was at distance 1 from the foot, the replication factor is meaningless.}.
 \end{itemize}
@@
 
-This goal is easily solvable, using some arithmetic. We let the reader look at the source of this development.
+This goal is easily solvable, using some arithmetic. We let the reader look at the source.
 
@@
 
-The mutually exclusive cases we consider are the following.
+The mutually exclusive cases we consider are the following:
 \begin{itemize}
@@
 
-If Hercules chops off a head near the floor, there is no replication at all. We use an auxiliary predicate, associated with the removing of one head from a sequence of hydras.
+If Hercules chops off a head near the floor, there is no replication at all. We use an auxiliary 
+predicate \texttt{S0}, associated with the removing of one head from a sequence of hydras.
 
@@
 
-Let us now consider beheadings  where the chopped-off head is at distance greater than or equal to $2$ from the foot.
-All the following relations are parameterized by the number $n$ of new copies added by the hydra.
+Let us now consider beheadings  where the chopped-off head is at distance greater than or equal to $2$ from the foot. All the following relations are parameterized by the replication factor  $n$.
 
@@
 
-\subsection{Description of any round}
+\subsection{Relation associated with a round}
 
@@
 
-\textbf{Hint:} Please keep in mind that the last  configuration of your interactively built battle is known only at the end of the battle.
+\textbf{Hints:} 
+\begin{itemize}
+
+\item Please keep in mind that the last  configuration of your interactively built battle is known only at the end of the battle.
 Thus, you will have to create and solve subgoals with existential variables.
@@
 goal \Verb@{h':Hydra | h -*-> h'}@ generates the subgoal \Verb|h -*-> ?h'|.
+\item You may use Gérard Huet's \emph{zipper} data structure~\cite{zipper} for writing tactics associated with Hercule's  interactive search of a head to chop off.
+\end{itemize}
+
 \end{exercise}
@@
 \begin{itemize}
-\item Hercules chops always the rightmost head of the hydra.
+\item Hercules chops off always the rightmost head of the hydra.
 \item The battle is standard: at the round number $i$, the expected replication is $i$.
@@
 
-The following lemma  is a formal statement about these first rounds, in terms of the
+The following lemma  is a formal description of these first rounds, in terms of the
 \texttt{battle} predicate.
@@
 We can make bigger steps through iterations of \texttt{next}.
-
 The functional \texttt{iterate}, similar to Standard Library's \texttt{Nat.iter},
-is defined and studied in~\href{../src/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}).
+is defined and studied in~\href{../src/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
 
@@
 
-The battle we study seems to be awfully long. Let us concentrate our
+The battle we are studying seems to be awfully long. Let us concentrate our
 tests on some particular events : the states where $\texttt{nh}=0$
@@
 We now have some intuition of the sequence.
-It looks like the next ``\texttt{nh}=0'' event will happen at the $192=2(95+1)$-th round, then at the $2(192+1)$-th round.
+It looks like the next ``\texttt{nh}=0'' event will happen at the $192=2(95+1)$-th round, then at the $2(192+1)$-th round, etc.
 
@@
   nat -> nat -> nat -> nat -> nat -> nat -> Prop :=
-| step1 : forall a b c,  one_step i a b (S c) a b c
-| step2 : forall a b ,  one_step i a (S b) 0 a b (S i)
-| step3  : forall a, one_step i (S a) 0 0 a (S i) 0.
+| step1: forall a b c, one_step i a b (S c) a b c
+| step2:  forall a b, one_step i a (S b) 0 a b (S i)
+| step3: forall a, one_step i (S a) 0 0 a (S i) 0.
 \end{Coqsrc}
@@
 
-\subsection{Two small proofs of impossibility}
+\subsection{A  small proof of impossibility}
 \index{Maths!Proofs of impossibility}
+
+\label{omega-case}
 
@@
 
-\subsubsection{No variant on Peano.lt}
-\label{omega-case}
-
 The best known well-founded order is the natural order on the set $\mathbb{N}$ of natural numbers (the type \texttt{nat} of Standard library).
@@
 Unfortunately, we can prove that 
-\emph{no} instance of class \texttt{WfVariant round Peano.lt $m$} can be built, where
+\emph{no} instance of class (\texttt{WfVariant round Peano.lt $m$}) can be built, where
 $m$ is \emph{any} function of type \texttt{Hydra $\arrow$ nat}.
@@
 
-Let us assume there is a variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
+Let us assume there exists some variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
     the  termination of all hydra battles.
@@
 due to the hypothesis \texttt{Hvar}, we are interested in the termination of \emph{free} battles. 
-There is no problem to consider a round with \texttt{m big\_h} as replication factor.
+There is no problem to consider a round with (\texttt{m big\_h}) as the replication factor.
 
@@
 \begin{Coqsrc}
-Lemma m_ge : forall i:nat , i <= m (iota i).
+Lemma m_ge : forall i:nat, i <= m (iota i).
 \end{Coqsrc}
@@
 
-\subsection{A proof a termination using \texorpdfstring{$\omega^2$}{omega^2}} 
+\subsection{A proof of termination using \texorpdfstring{$\omega^2$}{omega^2}} 
 \label{sect:merge-example}.
@@
 Add to the class \texttt{ON} the requirement that for any $\alpha$ it is decidable whether $\alpha$ is $0$, a successor or a limit ordinal.
+
+\textbf{Hint:}   Beware of the instances associated with sum and product of notations !
+  You may consider additional fields such as ``there is a greatest element'' 
+to make the sum and product of notations ``compositional''.
+
 \end{project}
@@
 
-\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0} : Paths}
+\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0} : paths}
 \index{Maths!Accessibility inside epsilon0}
@@
 
-\section{Large  sequences}
-
-\subsection{Definition}
+\section{Definitions}
+
+%\subsection{Definition}
 
@@
 
-\section{Computing mlarge sequences}
+\section{The length of minimal large sequences}
 
@@
 
-\subsection{Looking for a function definition}
+%\subsection{Looking for a function definition}
 
@@
 
-\subsection{Properties derived from \texttt{L\_spec}}
+\subsection{Abstract properties}
 
@@
 
-\section{Using \texttt{Equations}}
+\subsection{Using \texttt{Equations}}
 \label{sect:L-equations}
@@
 
-\section{The Wainer-Hardy hierarchy}
+\section{The Wainer-Hardy hierarchy (functions \texorpdfstring{$H_\alpha$}{\texttt{H\_alpha}})}
+
 \label{sect:hardy}
@@
 we think it would be useful to include comments and hints, and questions/answers from the users. In constrast, ``projects'' are supposed, once completed, to be included in the repository.
-
-\subsubsection{Project \vref{project:succ-limit-dec}}
-\paragraph*{Hint:} 
-  Beware of the instances associated with sum and product of notations !
-  You may consider additional fields such as ``there is a greatest element'' 
-to make the sum and product of notations ``compositional''.
 

### 2020-09-29 doc/hydras.tex
correcting typos

@@
 
-Let us comment some features of this table:
+Let us comment some features of this figure:
 
@@
 For instance, $12$ is the successor of $11$, and $\omega^4+1$ the successor of $\omega^4$.
-\item Ordinals that  follow dots immediately (displayed in red) are called \emph{limit ordinals}.
+\item Ordinals (displayed in red)  that  follow immediately dots are called \emph{limit ordinals}.
 With respect to the order induced by this sequence, any limit ordinal $\alpha$ is the least upper bound of  the set $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
@@
 It plays an important role in proof theory, and is particularly studied in chapters~\ref{chap:T1} to \ref{chap:alpha-large}.
-\item All the ordinals shown in this list are either the ordinal \textcolor{blue}{$0$},
+\item Any ordinal is  either the ordinal \textcolor{blue}{$0$},
 a successor ordinal, or a \textcolor{red}{limit ordinal}.
@@
 \item[trichotomy]: $\forall x\,y\in A, x<y \vee x = y \vee y < x$
-\item[well foundedness] $<$ is well-founded (every element of $A$ is accessible)\footnote{In classical mathematics, we would say that there is no infinite sequence $a_1>a_2> \dots a_n> a_{n+1}\dots$ in $A$.
+\item[well foundedness]: $<$ is well-founded (every element of $A$ is accessible)\footnote{In classical mathematics, we would say that there is no infinite sequence $a_1>a_2> \dots a_n> a_{n+1}\dots$ in $A$.
 Please refer to any documenation on \coq{} for having more details on well-foundedness and accessibility.}.
@@
 
-The best known examples of well-ordered sets are the set $\mathbb{N}$ of natural numbers (with the usual order $<$), as well as any finite segment $[0,i)=\{j\in\mathbb{N}|j<i\}$.
+The best known examples of well-ordered sets are the set $\mathbb{N}$ of natural numbers (with the usual order $<$), as well as any finite segment $[0,i)=\{j\in\mathbb{N}\,|\,j<i\}$.
 The disjoint union of two copies of $\mathbb{N}$, \emph{i.e.} the set $\{0,1\}\times\mathbb{N}$ is also well-ordered,
@@
 So, one can speak about \emph{finite}, \emph{infinite}, \emph{countable}, etc., ordinals.
-Nevertheless, since we work with type theory,
+Nevertheless, since we work within type theory,
 we do not identify ordinals as sets of ordinals, but the correspondance between ordinals and sets of ordinals is the function that maps $\alpha$ to $\mathbb{O}_\alpha$.
@@
 \item A ``mathematical'' representation of the set of countable ordinal numbers, afer Kurt Schütte~\cite{schutte}.
-This representation uses several (hopefully harmess) axioms.
-It is used as a reference for proving the correctness of ordinal notations.
+This representation uses several (hopefully harmless) axioms.
+We use it as a reference for proving the correctness of ordinal notations.
 \item A family of \emph{ordinal notations}, \emph{i.e.} data types used to represent segments $[0,\mu)$, where $\mu$ is some countable ordinal.
@@
 Note that proofs of
-correctness of a given ordinal notation with respect to Schütte's model use axioms.
+correctness of a given ordinal notation with respect to Schütte's model obviously use axioms.
 Please execute the \texttt{Print Assumptions} command in case of doubt.
@@
 Chapter~\ref{chap:schutte} of this document presents an adaptation to \coq{} of an axiomatization in classical logic of the set of countable ordinals by K. Schütte~\cite{schutte}. 
-That formalization is quite complex and technical and unshamedly non-constructive,  so we put its description  in the last chapter of this document.
+That formalization is quite complex, technical and unshamedly non-constructive,  so we put its description  in the last chapter of this document.
 
@@
 
-In our adaptation to \coq{}, we declare a type \texttt{Ord}, a binary relation \texttt{lt} (with infix notation \texttt{``\_<\_''}, and assume Schütte's axiom. In Chapter~\ref{chap:schutte},
+In our adaptation to \coq{}, we declare a type \texttt{Ord}, a binary relation \texttt{lt} (with infix notation \texttt{"\_<\_"}, and assume Schütte's axiom. In Chapter~\ref{chap:schutte},
 we derive some interesting properties of countable ordinals from these axioms.
+It is interesting to compare proofs of a given property (for instance the associativity of addition) in the computatuinal framework of some ordinal notation, and in the axiomatic model of Schütte.
 
@@
 \item A well founded order $<$ on $A$, 
-\item A function for comparing two ordinals, whose correctnes implies that the reflexive closure of $<$ is a \emph{total order}.
+\item A correct function for comparing two ordinals. Note  that the reflexive closure of $<$ is thus a \emph{total order}.
 \end{itemize}
@@
 The reflexive closure of \texttt{lt}, (called \texttt{le}, for ``less  or equal than'') is a total decidable order, implemented through a comparison function \texttt{compare}.
+The correctness of this function is expressed through Stdlib's type
+\texttt{Datatypes.CompareSpec}.
+
+\begin{Coqsrc}
+Inductive CompareSpec (Peq Plt Pgt : Prop) : comparison -> Prop :=
+    CompEq : Peq -> CompareSpec Peq Plt Pgt Eq
+  | CompLt : Plt -> CompareSpec Peq Plt Pgt Lt
+  | CompGt : Pgt -> CompareSpec Peq Plt Pgt Gt
+\end{Coqsrc}
 
@@
             {on : ON lt compare}
-            {B : Type}
-  (m : B -> A) : relation B :=
+            {B : Type} (m : B -> A) : relation B :=
              fun x y => on_lt (m x) (m y).
@@
             {B : Type}
-            (m : B -> A) :
-  well_founded (measure_lt m). 
+            (m : B -> A):  well_founded (measure_lt m). 
 \end{Coqsrc}
@@
 on the disjoint sum of the associated types, by putting all elements of \texttt{A} before the elements of \texttt{B} (thanks to Standard Library's relation operator \texttt{le\_AsB}).
-(please look at Figure~\ref{fig:omega-plus-omega} for an example).
 
@@
 Proof. constructor. Qed.
+
+Lemma lt_omega alpha : 
+     alpha o< omega <-> exists n:nat,  alpha = fin n.
+(* ... *)
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Definition succb (alpha: t) := match alpha with
-                                   inr (S  _) | inl (S _) => true
+Definition succb (alpha: t) : bool := match alpha with
+                                 | inr (S  _) | inl (S _) => true
                                  | _ => false
@@
     {beta: A | Successor alpha beta}.
+(* ... *)
 \end{Coqsrc}
@@
 
-In order to build an ordinal notation out the type \texttt{t}, we have to define a comparison function, and prove that the order \texttt{lt} is well-founded.
-
-\begin{Coqsrc}
+In order to build an ordinal notation out of the type \texttt{t}, we have to define a comparison function, and prove that the order \texttt{lt} is well-founded.
+
+\begin{Coqsrc}
+Instance lt_strorder : StrictOrder lt.
+(* ... *)
+
 Definition compare (alpha beta: t) : comparison :=
@@
 
+Lemma lt_wf : well_founded lt.
+
 Instance Omega2 : ON  lt compare.
+Proof.
+  split.
+  - apply lt_strorder.
+  - apply lt_wf.
+  - apply compare_correct.
+Qed.
 \end{Coqsrc}
@@
 
-Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily the function which merges two lists.
+Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily a total function which merges two lists.
 
@@
 (\emph{i.e.} the ordinal $\omega$) for proving the termination of all hydra battles.
-We  prove now that even the ordinal $\omega^2$ is also insufficient for this purpose.
+We  prove now that  the ordinal $\omega^2$ is also insufficient for this purpose.
 
@@
 The proof of the inequality \texttt{m big\_h o<= m small\_h} is quite more complex than in Sect~\ref{omega-case}.
-If we consider some ordinal $\alpha=(i,j)$, where $i>0$, there exists an infinite number of
+If we consider any ordinal $\alpha=(i,j)$, where $i>0$, there exists an infinite number of
 ordinals stricly less than $\alpha$, and there exists an infinite number of battles that start from
 $\iota(\alpha)$.
-In effect, at any configuration $\iota(k,0)$, where $k>0$, the hydra can freely chose any replication number.
+Indeed, at any configuration $\iota(k,0)$, where $k>0$, the hydra can freely choose any replication number.
 Intuitively, the measure of such a hydra must be large enough for taking into account
 all the possible battles issued from that hydra.
-Let us give more technical details.
+Let us now give more technical details.
 
@@
 Thus, starting from any inequality $q < p$ on type \texttt{t}, we can build 
-by transfinite induction over \texttt{p} a battle 
+by \emph{transfinite induction} (\emph{i.e.} well-founded) over \texttt{p} a battle 
 that transforms the hydra $\iota(p)$ into $\iota(q)$.
@@
 
-Since $<$ is a strict order (irreflexive  and transitive) on \texttt{nat*nat}, we can conclude that there is no
+Since $<$ is a strict order (irreflexive  and transitive) on \texttt{nat*nat}, we infer that there is no
 variant for termination on the lexicographic square of $(\mathbb{N},<)$.
@@
 \end{exercise}
-
-\section{Comparing Ordinal Notations}
-
-It is sometimes useful to compare two ordinal notations with respect to expressive power
-(the segment of ordinals  they represent).
-
-The following class specifies a strict inclusion of segments.
-The notation \texttt{OA} describes a segment $[0,\alpha($, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented in \texttt{OA}).
-We require also  that the comparison functions of the two notation systems are compatible.
-
-If \texttt{OB} is presumed to be correct, the we can consider that \texttt{OA} ``inherits'' its correcteness from the bigger notation system \texttt{OB}.
-
-\index{Coq!Type classes}
-\label{types:SubON}
-
-\begin{Coqsrc}
-Class  SubON 
-       `(OA : @ON A ltA  compareA)
-       `(OB : @ON B ltB  compareB)
-       (alpha :  B)
-       (iota : A -> B):=
-  {
-  SubON_compare: forall x y : A,  compareB (iota x) (iota y) =
-                                 compareA x y;
-  SubON_incl : forall x, ltB (iota x) alpha;
-  SubON_onto : forall y, ltB y alpha  -> exists x:A, iota x = y}.
-\end{Coqsrc}
-
-For instance, we prove that \texttt{Omega} is a sub-notation of
-\texttt{Omega\_plus\_Omega} (with $\omega$ as the first ``new'' ordinal, and \texttt{fin} as the injection).
-
-\begin{Coqsrc}
-Instance Incl : SubON Omega Omega_plus_Omega omega fin.
-\end{Coqsrc}
-
-\index{Projects}
-\begin{project}
-Prove that \texttt{Omega\_plus\_Omega} cannot be a sub-notation of \texttt{Omega}.
-\end{project}
-
-\index{Projects}
-\begin{project}
-Adapt the definition of \texttt{Hvariant} (Sect.~\ref{sect:hvariant-def}) in order to
-have an ordinal notation as argument. Prove that if $O_A$ is a sub-notation of $O_B$, then any variant defined on  $O_A$ can be automatically transformed into 
-a variant on $O_B$.
-\end{project}
-
-\section{Comparing an ordinal notation with Schütte's model}
-
-Finally, it may be interesting to compare an ordinal notation with the more theoretical model from Schütte (well, at least with our formalization of that model).
-This would be a relative proof of correctenss of our ordinal  notations.
-
-The following class specifies that a notation \texttt{OA} describes a segment $[0,\alpha)$,
-where $\alpha$ is a countable ordinal \emph{à la}  Schütte.
-
-\index{Coq!Type classes}
-\label{types:ON-for}
-
-\begin{Coqsrc}
-Class ON_correct `(alpha : Schutte_basics.Ord)
-     `(OA : @ON A ltA  compareA)
-      (iota : A -> Schutte_basics.Ord) :=
-  { ON_correct_inj : forall a, Schutte_basics.lt (iota a) alpha;
-    ON_correct_onto : forall beta, Schutte_basics.lt beta alpha ->
-                                exists b, iota b = beta;
-    On_compare_spec : forall a b:A,
-        match compareA a b with
-          Datatypes.Lt => Schutte_basics.lt (iota a) (iota b)
-        | Datatypes.Eq => iota a = iota b
-        | Datatypes.Gt => Schutte_basics.lt (iota b) (iota a)
-        end}.
-\end{Coqsrc}
-
-For instance, the following theorem tells that \texttt{Epsilon0}, our notation system for the segment $[0,\epsilon0)$ is correct with respect to the countable ordinal $\epsilon_0$
-(see chapter~\ref{chap:schutte} for more details).
-
-\begin{Coqsrc}
-Instance Epsilon0_correct :
-  ON_correct epsilon0 Epsilon0  (fun alpha => inject (cnf alpha)).
-\end{Coqsrc}
-
-\index{Projects}
-
-\begin{project}
-  When you have read Chapter~\ref{chap:schutte}, prove that the sum of two ordinal notations \texttt{ON\_plus} implements the addition of ordinals.
-\end{project}
 
@@
 
-Finally, it is interesting to show that, if $i<j$, then the segment $[0,i)$ is a ``sub-segment'' of
+\begin{remark}
+ There is no interesting arihmetic on finite ordinals, since functions like successor, addition, etc.,  cannot be represented in \coq{} as \emph{total} functions.
+\end{remark}
+
+\begin{remark}
+Finite ordinals are also formalized in MathComp~\cite{SSR}.  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.  
+\end{remark}
+
+%%%
+
+\section{Comparing two ordinal notations}
+
+It is sometimes useful to compare two ordinal notations with respect to expressive power
+(the segment of ordinals  they represent).
+
+The following class specifies a strict inclusion of segments.
+The notation \texttt{OA} describes a segment $[0,\alpha($, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented in \texttt{OA}).
+We require also  that the comparison functions of the two notation systems are compatible.
+
+If \texttt{OB} is presumed to be correct, the we can consider that \texttt{OA} ``inherits'' its correcteness from the bigger notation system \texttt{OB}.
+
+\index{Coq!Type classes}
+\label{types:SubON}
+
+\begin{Coqsrc}
+Class  SubON 
+       `(OA : @ON A ltA  compareA)
+       `(OB : @ON B ltB  compareB)
+       (alpha :  B)
+       (iota : A -> B):=
+  {
+  SubON_compare: forall x y : A,  compareB (iota x) (iota y) =
+                                 compareA x y;
+  SubON_incl : forall x, ltB (iota x) alpha;
+  SubON_onto : forall y, ltB y alpha  -> exists x:A, iota x = y}.
+\end{Coqsrc}
+
+For instance, we prove that \texttt{Omega} is a sub-notation of
+\texttt{Omega\_plus\_Omega} (with $\omega$ as the first ``new'' ordinal, and \texttt{fin} as the injection).
+
+\begin{Coqsrc}
+Instance Incl : SubON Omega Omega_plus_Omega omega fin.
+\end{Coqsrc}
+
+We can also show that, if $i<j$, then the segment $[0,i)$ is a ``sub-segment'' of
 $[0,j)$. Since the terms  ($t\;i$) and ($t\;j$) are not convertible, we consider a ``cast'' 
@@
 
-\begin{remark}
- There is no interesting arihmetic on finite ordinals, since functions like successor, addition, etc.,  cannot be represented in \coq{} as \emph{total} functions.
-\end{remark}
-
-\paragraph{Related work}
-Finite ordinals are also formalized in MathComp~\cite{SSR}.  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.
+\index{Projects}
+\begin{project}
+Prove that \texttt{Omega\_plus\_Omega} cannot be a sub-notation of \texttt{Omega}.
+\end{project}
+
+\index{Projects}
+\begin{project}
+Adapt the definition of \texttt{Hvariant} (Sect.~\ref{sect:hvariant-def}) in order to
+have an ordinal notation as argument. Prove that if $O_A$ is a sub-notation of $O_B$, then any variant defined on  $O_A$ can be automatically transformed into 
+a variant on $O_B$.
+\end{project}
+
+\section{Comparing an ordinal notation with Schütte's model}
+
+Finally, it may be interesting to compare an ordinal notation with the more theoretical model from Schütte (well, at least with our formalization of that model).
+This would be a relative proof of correctenss of the considered  ordinal  notation.
+
+The following class specifies that a notation \texttt{OA} describes a segment $[0,\alpha)$,
+where $\alpha$ is a countable ordinal \emph{à la}  Schütte.
+
+\index{Coq!Type classes}
+\label{types:ON-for}
+
+\begin{Coqsrc}
+Class ON_correct `(alpha : Schutte_basics.Ord)
+     `(OA : @ON A ltA  compareA)
+      (iota : A -> Schutte_basics.Ord) :=
+  { ON_correct_inj : forall a, Schutte_basics.lt (iota a) alpha;
+    ON_correct_onto : forall beta, Schutte_basics.lt beta alpha ->
+                                exists b, iota b = beta;
+    On_compare_spec : forall a b:A,
+        match compareA a b with
+          Datatypes.Lt => Schutte_basics.lt (iota a) (iota b)
+        | Datatypes.Eq => iota a = iota b
+        | Datatypes.Gt => Schutte_basics.lt (iota b) (iota a)
+        end}.
+\end{Coqsrc}
+
+For instance, the following theorem tells that \texttt{Epsilon0}, our notation system for the segment $[0,\epsilon0)$ is a correct implementation of the theoretically defined  ordinal $\epsilon_0$
+(see chapter~\ref{chap:schutte} for more details).
+
+\begin{Coqsrc}
+Instance Epsilon0_correct :
+  ON_correct epsilon0 Epsilon0  (fun alpha => inject (cnf alpha)).
+\end{Coqsrc}
+
+\index{Projects}
+
+\begin{project}
+  When you have read Chapter~\ref{chap:schutte}, prove that the sum of two ordinal notations \texttt{ON\_plus} implements the addition of ordinals.
+\end{project}
 
@@
 
-In some cases we want to show that two notation systems describe the same segment (for instance $[0,3+\omega)$ and $[0,\omega($).
+In some cases we want to show that two notation systems describe the same segment (for instance $[0,3+\omega)$ and $[0,\omega($\;).
 For this purpose, one may prove that the two notation systems are order-isomorphic.
@@
 
-{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$ we will prove in larger notation systems, as well as in Schütte's model}
+{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$, that we will prove in larger notation systems, as well as in Schütte's model.}
 
@@
 
-\textbf{Hint:}   Beware of the instances associated with sum and product of notations !
+\textbf{Hint:}   Beware of the instances associated with sum and product of notations!
   You may consider additional fields such as ``there is a greatest element'' 
@@
  \end{Coqsrc}
+
 \subsection{Remarks}

### 2020-10-01 doc/hydras.tex
minor modifications

@@
 
-Instance Omega2 : ON  lt compare.
+Instance Omega2: ON lt compare.
 Proof.
@@
 \begin{Coqsrc}
-(** multiplication of an ordinal and a natural number *)
+(** multiplication of an ordinal by a natural number *)
 
@@
 
-(** multiplication of  a natural number and an ordinal *)
+(** multiplication of  a natural number by an ordinal *)
 
@@
 \subsection{Cantor normal form}
+\index{Maths!Cantor normal form}
 
@@
 \emph{infinite} \(\omega\)-tower.
-
-\index{Maths!Cantor normal form}
-
-Any ordinal strictly less that \(\epsilon_0\) 
-can be represented by a unique  \emph{Cantor normal form}, 
+Nevertheless, 
+any ordinal strictly less that \(\epsilon_0\) 
+can be finitely represented by a unique  \emph{Cantor normal form}, 
 that is, an expression  which is either  the ordinal \(0\) or 
@@
 \paragraph{Remark}
-For simplicity's sake, we chosed to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the contruction (\texttt{ocons $\alpha$ $n$ $\beta$} is intented to represent the
+For simplicity's sake, we chosed to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the contruction (\texttt{ocons $\alpha$ $n$ $\beta$}) is intented to represent the
 ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
@@
 
-Some abbreviations may help to write more consisely complex ordinal terms.
+Some abbreviations may help to write more concisely complex ordinal terms.
 
@@
 Coercion fin  : nat >-> T1.
-
-Notation "'F' n" := (fin n) (at level 10) : t1_scope.
 
@@
 
-The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximed by the sequence of $\omega$-towers.
+The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximed by the sequence of $\omega$-towers (see also Sect~\vref{sect:epsilon0-as-limit} ).
 
@@
 \begin{Coqsrc}
-Theorem lt_irrefl  alpha :  ~ lt alpha  alpha.
+Theorem lt_irrefl (alpha: T1):  ~ lt alpha alpha.
 
@@
 \label{sec:orgheadline75}
-We note that our data-type \texttt{T1} allows us to write expressions that
+Our data-type \texttt{T1} allows us to write expressions that
 are not properly in Cantor normal form as specified in Section \ref{sec:epsilon0-intro}.
@@
 
-But this style leads to clumsy statements, and many sub-goals in interactive proofs (although often solved with \texttt{auto} or \texttt{eauto}).
+But this style leads to clumsy statements, and generates too many sub-goals in interactive proofs (although often solved with \texttt{auto} or \texttt{eauto}).
 
@@
 Definition LT := restrict nf lt.
-Infix "o<" := LT : t1_scope.
+Infix "t1<" := LT : t1_scope.
 
 Definition LE := restrict nf le.
-Notation "alpha o<= beta" := (LE alpha beta) : t1_scope.
-\end{Coqsrc}
-
-\begin{todo}
-Infix notation for inequalities on ordinals have changed to \texttt{o<} and \texttt{o<=}.  This document is not updated yet, but the \coq{} scripts are correct.
-\end{todo}
+Infix "t1<=" := LE : t1_scope.
+\end{Coqsrc}
 
@@
 \begin{Coqsrc}
-Lemma LT_one : forall alpha, alpha o< one -> alpha = zero.
+Lemma LT_one : forall alpha, alpha t1< one -> alpha = zero.
 \end{Coqsrc}
@@
 \end{Coqsrc}
-
-\begin{todo}
-Infix notation for inequalities on type \texttt{T1} have changed to \texttt{t1<} and \texttt{t1<=}.  This document is not updated yet, but the \coq{} scripts are correct.
-\end{todo}
 
@@
 
+\label{sect:epsilon0-as-limit}
 Then, the ordinal  $\epsilon_0$ is defined as the limit of the sequence of all finite towers (a kind of infinite tower).

### 2020-10-04 doc/hydras.tex
minor modifications

@@
 \subsubsection{A predicate for characterizing normal forms}
+\label{sect:t1-nf}
+
 \label{sec:orgheadline74}
@@
 
-The type \texttt{T1} is not a correct ordinal notation, since it contains terms that are not in Cantor normal form. Moreover,
-in Sections~\ref{sect:L-equations}, \ref{sect:hardy},
-and \ref{sect:wainer}, we needed to define total recursive functions on well-formed ordinal terms less  than $\epsilon_0$, using the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777}.
+As we noticed in Sect.~\ref{sect:t1-nf}, the type \texttt{T1} is not a correct ordinal notation, since it contains terms that are not in Cantor normal form.
+In certain contexts (for instance in Sections~\ref{sect:L-equations}, \ref{sect:hardy},
+and \ref{sect:wainer}),  we need to define total recursive functions on well-formed ordinal terms less  than $\epsilon_0$, using the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777}.
  In order to define a type whose inhabitants represent just ordinals, we build a type gathering a term of type \texttt{T1} and a proof that this term is in normal form.
@@
 \item Define a variant of the type \texttt{E0} (with your definition of \texttt{nf}).
-\item Can you still prove a lemma like \texttt{E0\_eq\_iff} ? With the help of an axiom of the standard library ?
+\item Can you still prove a lemma like \texttt{E0\_eq\_iff} ? With the help of an axiom from some module of the standard library ?
 \end{enumerate}
@@
 \begin{Coqsrc}
-Lemma plus_not_monotonous : exists alpha beta gamma : T1,
+Lemma plus_not_monotonous_l : exists alpha beta gamma : T1,
     alpha t1< beta /\ alpha + gamma = beta + gamma.
@@
 Let us consider again the ordinal $\alpha_0$ defined in section~\vref{alpha0-def}
-
-If we ask \coq{} to print the normal form of \texttt{alpha\_0}, we get a hardly readable term of type \texttt{T1}.
+If we ask \coq{} to print its  normal form, we get a hardly readable term of type \texttt{T1}.
 
@@
 
-The following data type defines a abstract syntax fore more readable ordinals in Cantor normal form:
+The following data type defines a abstract syntax for more readable ordinals in Cantor normal form:
 
@@
    In order to use \texttt{T1} for proving termination results,
-we need to prove that  our representation of ordinals strictly less than \(\epsilon_0\)
-makes our order \texttt{<} well-founded. Then we will get \emph{transfinite induction} for free.
+we need to prove that  our order \texttt{<} is well-founded. Then we will get \emph{transfinite induction} for free.
 
 The proof of well-foundedness of the strict order $<$ on Cantor normal forms is already 
-available in the Cantor contribution by Castéran and Contéjean~\cite{CantorContrib}. That proof relies on a library on recursive path orderings written by
-E. Contéjean. We present also a direct proof of the same result, which does not require any knowledge on r.p.o.s.
+available in the Cantor contribution by Castéran and Contejean~\cite{CantorContrib}. That proof relies on a library on recursive path orderings written by
+E. Contejean. We present here  a direct proof of the same result, which does not require any knowledge on r.p.o.s.
 
@@
 
-The problem comes from that \(\delta\) may be bigger that \(\beta\) or
+The problem comes from the hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
 \(\gamma\);
@@
 \showmath{\alpha < \omega ^\alpha}. If \showmath{\omega^\alpha} is accessible,
-then \showmath{\alpha} is \emph{a fortori} accessible.
+then \showmath{\alpha} is \emph{a fortiori} accessible.
 
@@
 
-First, we prove that, for  any \texttt{LT}-accessible term \showmath{\alpha}, any well formed
-term (\texttt{ocons $\alpha$ $n$ $\beta$})  is accessible too.
+First, we prove that, for  any \texttt{LT}-accessible term \showmath{\alpha}, \showmath{\alpha} is 
+strongly accessible too (\emph{i.e.} any well formed
+term (\texttt{ocons $\alpha$ $n$ $\beta$})  is accessible).
 
@@
 
-We built an instance of \texttt{ON}, and proved its correction w.r.t. Schutte's model.
+We build an instance of \texttt{ON}, and prove its correction w.r.t. Schutte's model.
 
@@
 
-\emph{Coq}'s constraints on recursive definitions resulted in 
+\emph{Coq}'s constraints on recursive definitions result in 
 the quite  complex form of \texttt{oplus}'s definition.
-For making easier proof of properties of this function, 
-it is helpful to \emph{derive} lemmas that will help to simplify 
+Proofs of properties of this function can be simpler if we
+ \emph{derive} rewriting lemmas that will help to simplify 
 expressions of the form (\texttt{oplus $a$ $ b$}).
@@
 definition. They can be proved almost immediately, using \texttt{cbn} 
-and \texttt{rewrite} tactics.
+and \texttt{reflexivity} tactics.
 
@@
     apply oplus_comm_0 with (succ (max alpha beta)) ...  
-  (* rest of proof omitted *)
+  (* ... *)
 \end{Coqsrc}
@@
     transfinite_induction alpha.
-(* rest of proof omitted *) 
+    (* ... *)
 \end{Coqsrc}
@@
     apply oplus_assoc_0 with (succ (max alpha (max beta gamma))) ...
-(* rest of proof omitted *)   
+    (* ... *)   
 \end{Coqsrc}
@@
 \subsection{A measure for hydra-battle termination}
-\label{sec:orgheadline90}
+
 \label{sec:hydra-measure}
@@
 
+\label{thm:every-battle-terminates}
+
 \begin{Coqsrc}
@@
 Their proof uses some knowledge about model theory and non-standard models of Peano arithmetic.
-In this chapter, we focus on a specific class of proofs of termination of hydra battles: construction of some variant mapping the type \texttt{Hydra} into some segment of ordinals.
+In this chapter, we focus on a specific class of proofs of termination of hydra battles: construction of some variant mapping the type \texttt{Hydra} into a given initial  segment of ordinals.
 Our proof relies only on the Calculus of Inductive Constructions and is a natural complement of the results proven in the previous chapter.
@@
 \item There is no variant mapping the type \texttt{Hydra} into the interval $[0,\omega^2)$ (section ~\vref{omega2-case}), and a fortiori 
-$[0,\omega)$
-
-\item There is a variant that maps the type \texttt{Hydra} into the
-interval $[0,\epsilon_0)$ (theorem \texttt{every\_battle\_terminates}, in section~\vref{hydra-variant}).
+$[0,\omega)$ (section ~\vref{omega-case}).
+
+\item There exists a variant which maps the type \texttt{Hydra} into the
+interval $[0,\epsilon_0)$ (theorem \texttt{every\_battle\_terminates}, in section~\vref{thm:every-battle-terminates}).
 \end{itemize}
@@
 
-Canonical sequences are functions that associate an ordinal $\canonseq{\alpha}{i}$ to every ordinal $\alpha<\epsilon_0$ and positive integer $i$. They satisfy nice properties :
+Canonical sequences are functions that associate an ordinal $\canonseq{\alpha}{i}$ to every ordinal $\alpha<\epsilon_0$ and positive integer $i$. They satisfy several nice properties:
 
@@
 \item Canonical sequences correspond tightly to rounds of hydra battles: if $\alpha\not=0$,
-then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1}$ in one round with
+then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1})$ in one round with
 the replication factor $i$ (Lemma \href{../src/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
@@
 $\alpha$ is simply the constant sequence whose terms are equal to $\beta$.
-Likewise, we define the canonical sequence of $0$ as the 
-sequence all terms of which are equal to $0$.
+Likewise, the canonical sequence of $0$ maps any natural number to $0$.
 

### 2020-10-05 doc/hydras.tex
minor modifications

@@
 
-We prove in \coq{} the following results:
+We prove in \coq{} the following result:
 
@@
 \coqsimple{cbn}, \coqsimple{simpl}, \coqsimple{compute}, etc., are applicable. For simplicity's sake, 
-we used an auxiliary function \texttt{canonS} of type \texttt{T1 -> nat  -> T1} such that
+we use an auxiliary function \texttt{canonS} of type \texttt{T1 -> nat  -> T1} such that
 (\texttt{canonS  $\alpha$ $i$}) is equal to $\canonseq{\alpha}{i+1}$.
@@
 Nevertheless, in a future version of this development, we may define $\canonseq{\alpha}{0}$ exactly as
-in~\cite{KS81}. This would  be done at the cost of making proofs much more complex.}.
+in~\cite{KS81}. But we are afraid this would  be done at the cost of making some proofs much more complex.}.
 
@@
 as simple as suggested by the equations of definition~\ref{def:canonseq-math}\,.
-Nevertheless, we could prove it by  structural induction on $\alpha$.
+Nevertheless, we can prove it by  plain structural induction on $\alpha$.
 
@@
 
-First, we prove by transfinite induction over $\alpha$ that $\canonseq{\alpha}{i+1}$ is an ordinal strictly less than $\alpha$ (assuming $\alpha\not=0$).
+We prove by transfinite induction over $\alpha$ that $\canonseq{\alpha}{i+1}$ is an ordinal strictly less than $\alpha$ (assuming $\alpha\not=0$).
 This property allows us to use the function \texttt{canonS} and its derivates in function definitions by transfinite recursion.
@@
 Note the use of \coq's \texttt{sig} type in the theorem's statement, which
-relates the boolean function \texttt{is\_limit} defined on the \texttt{T1} data-type with a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.
+relates the boolean function \texttt{limitb} defined on the \texttt{T1} data-type with a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.
 

### 2020-10-06 doc/hydras.tex
minor modifications

@@
 \begin{remark}
+\label{remark:transitive-closure}
 \coq's library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Coq.Relations.Relation\_Operators} 
@@
 
+\index{Projects}
+\begin{project}
+Many lemmas presented in this chapter were stated and proved before the introduction of 
+the type class \texttt{ON} of ordinal notations, and in particular its  instance \texttt{Epsilon0}.
+Thus definitions and lemmas refer to the type \texttt{T1} of possibly not well-formed terms.
+This should be fixed in  a future version.
+\end{project}
+
 \subsection{Basic properties of canonical sequences}
@@
 
-Note the use of \coq's \texttt{sig} type in the theorem's statement, which
-relates the boolean function \texttt{limitb} defined on the \texttt{T1} data-type with a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.
-
 \vspace{4pt}
@@
 
+Note the use of \coq's \texttt{sig} type in the theorem's statement, which
+relates the boolean function \texttt{limitb} defined on the \texttt{T1} data-type with a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.
+We can also state directly that $\lambda$ is a (strict) least upper bound of the elements of its canonical sequence.
+
 \begin{Coqsrc}
@@
 Let us consider a kind of accessibility problem inside $\epsilon_0$: given two ordinals $\alpha$ and $\beta$, where $\beta<\alpha<\epsilon_0$, find a \emph{path} consisting of a finite sequence $\gamma_0=\alpha,\dots,\gamma_l=\beta$,
-where, for every $i<l$, $\gamma_i \not= 0$ \footnote{This condition allows us to ignore paths which end by a lot of useless $0$.} and there exists some strictly positive integer $s_i$
+where, for every $i<l$, $\gamma_i \not= 0$ \footnote{This condition allows us to ignore paths which end by a lot of useless $0$s.} and there exists some strictly positive integer $s_i$
 such that $\gamma_{i+1}=\canonseq{\gamma}{s_i}$,
@@
 
-For instance, we have $\omega*2 \xrightarrow[2,2,2]{}\omega$, through the 
-path $\omega\times 2, \omega+2,\omega+1,\omega$.
+For instance, we have $\omega*2 \xrightarrow[2,2,2,4,5]{}3$, through the 
+path $\langle\omega\times 2, \omega+2,\omega+1,\omega,4,3\rangle$.
 
@@
 Likewise,
-$\omega*2 \xrightarrow[{[3,14]}]{} 0$ and
+$\omega*2 \xrightarrow[1,2,1,4]{} 0$ and
 $\omega*2 \xrightarrow[3,3,3,3,3,3,3,3]{} 0$.
@@
 
+\index{Predicates!path\_to}
+\label{sect:path-to-def}
+
 \begin{Coqsrc}
@@
   match i with 0 => fun _ _ => False | S j => transition_S j end.
-\end{Coqsrc}
-
-\index{Predicates!path\_to}
-\label{sect:path-to-def}
-
-\begin{Coqsrc}
+
 Inductive path_to (beta: T1) : list nat -> T1 -> Prop :=
@@
 \begin{remark}
+The definition above is parameterized with the \emph{destination} of the path and indexed by the origin, hence the name \texttt{path\_to}.
+The rationale behind this choice is a personal preference of the developer  for the kind of eliminators generated by \coq{} in this case.
+The symmetric option could have been also considered (see also Remark~\vref{remark:transitive-closure}).
+\end{remark}
+
+\begin{remark}
 In the present version of our library, we use a variant \texttt{path\_toS} of
@@
 By transfinite induction on $\alpha$, we prove that for any $\beta<\alpha$, 
-on can build a path from $\alpha$ to $\beta$.
+one can build a path from $\alpha$ to $\beta$ (in other terms, $\beta$ is accessible from $\alpha$).
 
@@
 
-% ici
 \noindent 
@@
 
-In order to apply our knowledge about  ordinal numbers (less than $\epsilon_0$) to the study of hydra battles, we define an injection
+In order to apply our knowledge about  ordinal numbers less than $\epsilon_0$ to the study of hydra battles, we define an injection
 from the interval $[0,\epsilon_0)$ into the type \texttt{Hydra}.
@@
 
-The next step of our development is to extend this relationship to
+The next step of our development extends this relationship to
 the order $<$ on $[0,\epsilon_0)$ on one side, and hydra battles on the other side.
@@
 
-We have got all the tools for proving there is no variant bounded by some $\mu<\epsilon_0$ for proving the termination   of all battles.
+We now have  the tools for proving that  there exists no variant bounded by some $\mu<\epsilon_0$ for proving the termination   of all battles.
 The proof we are going to show is a proof by contradiction.
@@
 
+\label{remark:m-decrease}
 \begin{remark}
-  The hypothesis \texttt{m\_decrease} is false in general, but is satisfied by
+  The hypothesis \texttt{m\_decrease} is not provable  in general, but is satisfied by
 the  \texttt{free} and \texttt{standard} kinds of battles. This trick allows to 
@@
 \item The following lemma is an application of \texttt{m\_ge\_generic}, since \texttt{free}
-satisfies trivially the hypothesis \texttt{m\_decrease}.
+satisfies trivially the hypothesis \texttt{m\_decrease} (see page~\pageref{remark:m-decrease}).
 
@@
 
-The tool we  use comes  once again from Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
+The solution of this issue comes  once again from Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
 $\alpha_0=\alpha,\alpha_1,\dots,\alpha_k=\beta$ where $\alpha_{j+1}$ is equal
@@
 
-We shall also use a version of \texttt{Cor12} with large inequalities.
+We  also use a version of \texttt{Cor12} with large inequalities.
 
@@
   transfinite_induction alpha.
-  (* rest of proof skipped *)
+  (* ... *)
 \end{Coqsrc}
@@
 \paragraph*{}
-Let us consider two ordinals  $\beta<\alpha<\epsilon_0$.  Let $p$  be some $(n+1)$-path from $\alpha$ to $\beta$.
+Noq, let us consider two ordinals  $\beta<\alpha<\epsilon_0$.  Let $p$  be some $(n+1)$-path from $\alpha$ to $\beta$.
 
@@
     {l : nat | standard_fun (S n) alpha l = beta}.
+ (* ... *)
 
@@
 \label{sec:standard-battles-cases}
-We are now able to complete our proof that there exists no bounded variant for proving the termination of standard hydra battles. The proof we are going to comment can
+We are now able to complete our proof that there exists no bounded variant for proving the termination of standard hydra battles. This proof can
 be consulted in the module 
@@
 In this chapter, we try to feel how long a standard battle can be.
-To be precise, for any ordinal $\alpha<\epsilon0$ and any positive integer $k$,
+To be precise, for any ordinal $\alpha<\epsilon_0$ and any positive integer $k$,
 we give a minoration of the number of steps of a standard battle which
@@
 Inductive L_spec : T1 -> (nat -> nat) -> Prop :=
-  L_spec0 :
-    forall f, (forall k, f k = k) ->  L_spec zero f
+| L_spec0 : forall f, (forall k, f k = k) ->  L_spec zero f
 | L_spec1 : forall alpha f,
@@
 Lemma L_spec_unicity alpha f g :
-  L_spec alpha f -> L_spec alpha g ->
-  forall k, f (S k) = g (S k).
+  L_spec alpha f -> L_spec alpha g -> forall k, f (S k) = g (S k).
 \end{Coqsrc}
@@
 \texttt{L\_spec}.
-We are looking for properties which could be used for writing \emph{equations} and prove the correctness of the function generated by the \texttt{coq-equations} plugin.
+We are looking for properties which could be used for writing \emph{equations} and prove the correctness of the function generated by the \texttt{coq-equations} plug-in.
 Moreover, they will give us some examples of
@@
 Lemma path_to_omega_mult (i k:nat) :
-  path_to (omega * i)
-                (interval (S k) (2 * (S k)))
-                (omega * (S i)).
+  path_to (omega * i) (interval (S k) (2 * (S k))) (omega * (S i)).
 \end{Coqsrc}
@@
 
-The following lemma expresses the length of the mlarge sequences associated with finite multiples of $\omega$.
+The following lemma expresses the length of the mlarge sequences associated with the finite multiples of $\omega$.
 
@@
 
-To this end, we prove a generic lemma, which expresses $L_{\omega^\alpha\times i}$ in terms of $L_{\omega^\alpha}$.
-Note that in this correction lemma, we assume that the fonction associated with $\alpha$ is stritly monotonous and
-greater or equal than the successor function.
+To this end, we prove a generic lemma, which expresses $L_{\omega^\alpha\times i}$ as an iterate of $L_{\omega^\alpha}$.
+Note that in this lemma, we assume that the fonction associated with $\alpha$ is stritly monotonous and
+greater or equal than the successor function, and prove that $L_{\omega^\alpha\times i}$satisfies  the same properties.
 
@@
 \begin{Coqsrc}
-Lemma L_omega_cube_eqn i : L_omega_cube i = L_omega_square_times i (S i).
+Lemma L_omega_cube_eqn i : 
+   L_omega_cube i = L_omega_square_times i (S i).
 Proof. reflexivity. Qed.
@@
 Thus, for instance, $L_{\omega^3}(3)=L_{\omega^2\times 4}(3)$.
-So, we obtain an exact expression of this number.
+Thus, we obtain an exact expression of this number.
 
@@
 
-\vspace{4pt}XS
+\vspace{4pt}
 \noindent
@@
 
-Using these three lemmas as rewrite rules, we can prove more properties of the functions \texttt{L\_ $\alpha$}.
+Using these three lemmas as rewrite rules, we can prove more properties of the functions \texttt{L\_$\alpha$}.
 
@@
 \begin{Coqsrc}
-Lemma H_zero_eqn : forall i, H_ Zero i = i.
+Lemma H_eq1 : forall i, H_ Zero i = i.
 Proof.   intro i; now rewrite H__equation_1.  Qed.
@@
 
-Lemma H_lim_eqn alpha i : Limitb alpha ->
+Lemma H_eq3 alpha i : Limitb alpha ->
                       H_ alpha i =  H_ (Canon alpha (S i)) i.
@@
 Proof with auto with E0.
-  intro k; rewrite H_lim_eqn ...
+  intro k; rewrite H_eq3 ...
   - replace (Canon omega (S k)) with (Fin (S k)).
@@
 
+\begin{definition}
+ Let $f$ zand $g$ be two arithmetic  functions; $f$ is said to \emph{dominate} $g$ if $f(p)>g(p)$ for any all sufficiently large $p$.
+\end{definition}
+
 \begin{Coqsrc}
@@
   If $\beta<\alpha$, $F_\alpha$ dominates $F_\beta$.
-
-($f$ is said to \emph{dominate} $g$ if $f(p)>g(p)$ for any all sufficiently large $p$).
 \end{quote}

### 2020-10-07 doc/hydras.tex
minor modifications

@@
 
-Schütte's definition of countable ordinal relies on the following three axioms.
-
+Schütte's definition of countable ordinals relies on the following three axioms:
+
+There  exists a strictly ordered set , such that
 \begin{enumerate}
-\item There exists a well-ordered set $(\mathbb{O},<)$
+\item  $(\mathbb{O},<)$ is well-ordered
 \item Every bounded subset of $\mathbb{O}$  is countable
@@
 Both operators allow us to write definitions close to the natural mathematical language, such as ``$\textrm{succ}$ is \emph{the} least ordinal strictly greater than $\alpha$''
-\item Please note that only the sub-library \href{../src/html/hydras.Schutte.Schutte.html}{Ordinals.Schutte.Schutte/} is ``contaminated'' by axioms, and that the rest of our libraries remain constructive.
-
+\item Please note that only the library \href{../src/html/hydras.Schutte.Schutte.html}{Schutte/*.v} is ``contaminated'' by axioms, and that the rest of our libraries remain constructive.
 \end{itemize}
@@
 Parameter lt : relation Ord.
-Notation "a o< b" := (lt a b): schutte_scope.
+Infix "<" := lt : schutte_scope.
 
@@
 Axiom AX2 : forall X: Ensemble Ord, 
-   (exists a,  (forall y, In X y -> y o< a)) ->
+   (exists a,  (forall y, In X y -> y < a)) ->
    countable X.
@@
               countable X -> 
-              exists a,  forall y, In X y -> y o< a.
+              exists a,  forall y, In X y -> y < a.
 \end{Coqsrc}
@@
 
-We are now able de define \texttt{zero} as the least ordinal. For this purpose,
-we define a function returning the least element of any [non-empty]  set.
+We are now able to define \texttt{zero} as the least ordinal. For this purpose,
+we define a function returning the least element of any [non-empty]  subset.
 
@@
 \begin{Coqsrc}
-Lemma zero_le (alpha : Ord) :  zero o<= alpha.
+Lemma zero_le (alpha : Ord) :  zero <= alpha.
 Proof.
@@
  forall a : Ord, unique (least_member lt ordinal) a -> 
-                a o<= alpha
+                a <= alpha
 \end{Coqanswer}
@@
 \emph{i.e.}, satisfied by \emph{any} element of type \texttt{Ord}, like for instance
-\texttt{bottom = bottom}, or \texttt{zero o<= bottom}.
+\texttt{bottom = bottom}, or \texttt{zero <= bottom}.
 
 \begin{Coqbad}
-Lemma le_zero_bottom : zero o<= bottom. 
+Lemma le_zero_bottom : zero <= bottom. 
 Proof. apply zero_le. Qed.
@@
 
-Lemma le_bottom_zero : bottom o<= zero.
+Lemma le_bottom_zero : bottom <= zero.
 Proof.
@@
  forall a : Ord, unique (least_member lt (Empty_set Ord)) a -> 
-      a o<= zero
+      a <= zero
 \end{Coqanswer}
@@
 \begin{Coqsrc}
-Definition succ (alpha : Ord) := the_least (fun beta => alpha o< beta).
+Definition succ (alpha : Ord) := the_least (fun beta => alpha < beta).
 \end{Coqsrc}
@@
 
-It is also easy to define recursively the finite ordinals.
+% \begin{remark}
+% Please look at remark~\vref{warning:coercions}.
+% \end{remark}
+
+How do we prove properties of the successor function?
+First, we make its specification explicit.
+
+\begin{Coqsrc}
+Definition succ_spec (alpha:Ord) :=
+  least_member   lt (fun z => alpha < z).
+\end{Coqsrc}
+
+Then, we prove that our function \texttt{succ} meets this specification.
+
+\begin{Coqsrc}
+Lemma succ_ok : forall alpha,  succ_spec alpha  (succ alpha).
+Proof.
+  intros; unfold succ, the_least, the;  apply iota_spec.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+1 subgoal (ID 172)
+
+  alpha : Ord
+  ============================
+  exists ! x : Ord, succ_spec alpha x
+\end{Coqanswer}
+
+We have now to prove that the set of all ordinals strictly greater than $\alpha$ has a unique least element.
+But the singleton set $\{\alpha\}$ is countable, hence  bounded (by the axiom \texttt{AX3}).
+Hence; the set $\{\beta\in\mathbb{O}|\alpha < \beta\}$ is not empty
+and therefore has a unique least element.
+
+The \coq{} proof script is quite short.
+
+\begin{Coqsrc}
+  destruct (@AX3 (Singleton _ alpha)).
+  - apply countable_singleton.
+  -  unfold succ_spec; apply the_least_unicity;  exists x; intuition.
+Qed.     
+\end{Coqsrc}
+
+We can ``uncap'' the description operator for proving properties of the
+\texttt{succ} function.
+
+\begin{Coqsrc}
+Lemma lt_succ (alpha : Ord) :  alpha < succ alpha.
+Proof.
+  destruct  (succ_ok  alpha);  tauto.
+Qed.
+
+Hint Resolve lt_succ : schutte.
+
+Lemma lt_succ_le (alpha beta : Ord):
+  alpha < beta -> succ alpha <= beta.
+Proof with eauto with schutte.
+  intros  H;  pattern (succ alpha); apply the_least_ok ... 
+  exists (succ alpha); red;apply lt_succ ...
+Qed.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma lt_succ_le_2 (alpha beta : Ord):
+  alpha < succ beta -> alpha <= beta.
+
+Lemma succ_mono (alpha beta : Ord):
+  alpha < beta -> succ alpha < succ beta.
+
+Lemma succ_monoR (alpha beta : Ord) :
+ succ alpha < succ beta -> alpha < beta.
+
+Lemma lt_succ_lt (alpha beta : Ord) :
+  is_limit beta ->  alpha < beta -> succ alpha < beta.
+\end{Coqsrc}
+
+\section{Finite ordinals}
+
+Using \texttt{succ}, it is now easy to define recursively all the finite ordinals.
 
@@
 Coercion finite : nat >-> Ord.
-\end{Coqsrc}
-
-% \begin{remark}
-% Please look at remark~\vref{warning:coercions}.
-% \end{remark}
-
-How do we prove properties of the successor function?
-First, we make its specification explicit.
-
-\begin{Coqsrc}
-Definition succ_spec (alpha:Ord) :=
-  least_member   lt (fun z => alpha o< z).
-\end{Coqsrc}
-
-Then, we prove that our function \texttt{succ} meets this specification.
-
-\begin{Coqsrc}
-Lemma succ_ok : forall alpha,  succ_spec alpha  (succ alpha).
-Proof.
-  intros; unfold succ, the_least, the;  apply iota_spec.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-1 subgoal (ID 172)
-
-  alpha : Ord
-  ============================
-  exists ! x : Ord, succ_spec alpha x
-\end{Coqanswer}
-
-We have now to prove that the set of all ordinals strictly greater than $\alpha$ has a unique least element.
-But the singleton set $\{\alpha\}$ is countable, hence  bounded (by the axiom \texttt{AX3}).
-Hence; the set $\{\beta\in\mathbb{O}|\alpha < \beta\}$ is not empty
-and therefore has a unique least element.
-
-The \coq{} proof script is quite short.
-
-\begin{Coqsrc}
-  destruct (@AX3 (Singleton _ alpha)).
-  - apply countable_singleton.
-  -  unfold succ_spec; apply the_least_unicity;  exists x; intuition.
-Qed.     
-\end{Coqsrc}
-
-We can ``uncap'' the description operator for proving properties of the
-\texttt{succ} function.
-
-\begin{Coqsrc}
-Lemma lt_succ (alpha : Ord) :  alpha o< succ alpha.
-Proof.
-  destruct  (succ_ok  alpha);  tauto.
-Qed.
-
-Hint Resolve lt_succ : schutte.
-
-Lemma lt_succ_le (alpha beta : Ord):
-  alpha o< beta -> succ alpha o<= beta.
-Proof with eauto with schutte.
-  intros  H;  pattern (succ alpha); apply the_least_ok ... 
-  exists (succ alpha); red;apply lt_succ ...
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma lt_succ_le_2 (alpha beta : Ord):
-  alpha o< succ beta -> alpha o<= beta.
-
-Lemma succ_mono (alpha beta : Ord):
-  alpha o< beta -> succ alpha o< succ beta.
-
-Lemma succ_monoR (alpha beta : Ord) :
- succ alpha o< succ beta -> alpha o< beta.
-
-Lemma lt_succ_lt (alpha beta : Ord) :
-  is_limit beta ->  alpha o< beta -> succ alpha o< beta.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Lemma finite_lt_omega : forall i: nat,  i o< omega.
+Lemma finite_lt_omega : forall i: nat,  i < omega.
 
 Lemma lt_omega_finite alpha : Ord) : 
-  alpha o< omega ->  exists i:nat, alpha =  i.
+  alpha < omega ->  exists i:nat, alpha =  i.
 
@@
  Definition segment (A: Ensemble Ord) :=
-  forall alpha beta, In A alpha -> beta o< alpha -> In A  beta.
+  forall alpha beta, In A alpha -> beta < alpha -> In A  beta.
 
@@
  (forall b, B b -> exists a, In A a /\ f a = b) /\
- forall a b, In A a -> In A b -> a o< b ->  f a o< f b.
+ forall a b, In A a -> In A b -> a < b ->  f a < f b.
 
@@
     ordering_function f A B ->
-    forall alpha, In A alpha -> alpha  o<= f alpha.
+    forall alpha, In A alpha -> alpha <= f alpha.
 
@@
 
-Lemma le_plus_l (alpha beta : Ord) : alpha o<= alpha + beta.
-
-Lemma le_plus_r (alpha beta : Ord) :  beta o<= alpha + beta.
+Lemma le_plus_l (alpha beta : Ord) : alpha <= alpha + beta.
+
+Lemma le_plus_r (alpha beta : Ord) :  beta <= alpha + beta.
 
 Lemma plus_mono_r (alpha beta gamma : Ord) : 
-    beta o< gamma -> alpha + beta o< alpha + gamma.
+    beta < gamma -> alpha + beta < alpha + gamma.
 
@@
 Lemma finite_plus_infinite (n : nat) (alpha : Ord) :
-  omega o<= alpha -> n + alpha = alpha.
+  omega <= alpha -> n + alpha = alpha.
 \end{Coqsrc}
@@
   fun alpha => 
-  zero o< alpha /\
-  (forall beta, beta o< alpha ->  beta + alpha = alpha).
+  zero < alpha /\
+  (forall beta, beta < alpha ->  beta + alpha = alpha).
 \end{Coqsrc}
@@
   least_member   lt 
-                  (fun alpha => 1 o< alpha /\ In AP alpha)
+                  (fun alpha => 1 < alpha /\ In AP alpha)
                   omega.
@@
 Lemma AP_plus_closed (alpha beta gamma : Ord): 
-     In AP alpha -> beta o< alpha -> gamma o< alpha -> beta + gamma o< alpha.
+     In AP alpha -> beta < alpha -> gamma < alpha -> beta + gamma < alpha.
 
@@
 Lemma phi0_mono (alpha beta : Ord) :
-  alpha o< beta ->  phi0 alpha o< phi0 beta.
+  alpha < beta ->  phi0 alpha < phi0 beta.
 
@@
 Lemma is_limit_phi0 (alpha : Ord) :
-  zero o< alpha ->  is_limit (phi0 alpha).
+  zero < alpha ->  is_limit (phi0 alpha).
 
@@
 
-Lemma phi0_le (alpha : Ord) : alpha o<= phi0 alpha.
+Lemma phi0_le (alpha : Ord) : alpha <= phi0 alpha.
 \end{Coqsrc}
@@
 Lemma lt_phi0 (alpha : Ord):
-  alpha o< epsilon0 -> alpha o< phi0 alpha.
+  alpha < epsilon0 -> alpha < phi0 alpha.
 \end{Coqsrc}
@@
 Definition Cr_fun : forall alpha : Ord,
-       (forall beta : Ord, beta o< alpha -> Ensemble Ord) ->
+       (forall beta : Ord, beta < alpha -> Ensemble Ord) ->
         Ensemble Ord 
@@
         (Cr : forall beta, 
-                beta o< alpha -> Ensemble Ord) 
+                beta < alpha -> Ensemble Ord) 
         (x : Ord) => (
        (alpha = zero /\ AP x) \/
-       (zero o< alpha /\
-        forall beta (H:beta o< alpha),
+       (zero < alpha /\
+        forall beta (H:beta < alpha),
           the_ordering_segment (Cr beta H) x /\ ord (Cr  beta H) x = x)).
@@
 Definition sorted (l: cnf_t) :=
-  LocallySorted (fun alpha beta => beta o<= alpha) l.
+  LocallySorted (fun alpha beta => beta <= alpha) l.
 
@@
  forall l alpha, 
-   is_cnf_of alpha l ->  alpha o< epsilon0 ->
-   Forall (fun beta =>  beta o< alpha) l.
+   is_cnf_of alpha l ->  alpha < epsilon0 ->
+   Forall (fun beta =>  beta < alpha) l.
 \end{Coqsrc}
@@
    is_cnf_of alpha l ->  
-   (alpha o< epsilon0 <->  Forall (fun beta =>  beta o< alpha) l).
+   (alpha < epsilon0 <->  Forall (fun beta =>  beta < alpha) l).
 \end{Coqsrc}
@@
   T1.nf beta -> T1.nf gamma -> 
-  inject beta o< inject gamma.
+  inject beta < inject gamma.
 
@@
 Theorem inject_lt_epsilon0 (alpha : T1):
-      inject alpha o< epsilon0.
+      inject alpha < epsilon0.
 

### 2020-10-14 doc/hydras.tex
minor modifications

@@
 But we found it interesting to make all the three proofs available, allowing the reader to compare their common structure and notice their technical differences.
-In particular, the proof of $(3)$ uses several non-trivial combinatorial properties of ordinal numbersup to $\epsilon_0$~\cite{KS81}, whilst $(1)$ and $(2)$ use simple properties of $\mathbb{N}$ and $\mathbb{N}^2$.
+In particular, the proof of $(3)$ uses several non-trivial combinatorial properties of ordinal numbers up to $\epsilon_0$~\cite{KS81}, whilst $(1)$ and $(2)$ use simple properties of $\mathbb{N}$ and $\mathbb{N}^2$.
 
@@
 \subsection{State of the development}
-The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on ordinal notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
+The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on  notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
 We added new material : axiomatic definition of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, equations, etc.
 
-We are now working in order to make clumsy proofs more readable, simplify definitions, 
-to ``factorize'' proofs as much as possible.
-
-Many ``todo''s or projects are put in this text, suggesting possible improvements.
+We are now working in order to make clumsy proofs more readable, simplify definitions, and ``factorize'' proofs as much as possible. 
+Many possible improvements are suggested as ``todo''s or ``projects'' in this text.
 

### 2020-10-18 doc/hydras.tex
removing the useless/too-weak oplus_rect

@@
 \textbf{Hint:}   Beware of the instances associated with sum and product of notations!
-  You may consider additional fields such as ``there is a greatest element'' 
+  You may consider additional fields 
 to make the sum and product of notations ``compositional''.
@@
 
-\subsubsection{A hand-made induction principle}
-\label{sec:orgheadline85}
-
-\index{Coq!Commands!Functional Scheme}
-
-\emph{Coq} contains a command  \texttt{Functional Scheme} that 
-generates induction principles which correspond to recursive functions.
-Unfortunately, the current version ( \texttt{8.11.0} ) doesn't work on \texttt{oplus},
-probably because of the inner \texttt{Fix}.
-
-\begin{Coqsrc}
-Functional Scheme oplus_ind := Induction for oplus Sort Prop.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-Error: Anomaly "todo." Please report at http://coq.inria.fr/bugs/.
-\end{Coqanswer}
-
-Fortunately, it's a good exercise for a semi-experienced user, to write
-her/him-self induction principles similar to the ones returned by
-\texttt{Functional Scheme}.
-
-\begin{itemize}
-\item First, we chose to write a version for sort \texttt{Type}, since versions
-for sorts \texttt{Prop} and \texttt{Set} can be easily derived from
-the former one. According to \emph{Coq}'s naming politics, we will call our 
-principle \texttt{oplus\_rect}
-
-\item The conclusion of \texttt{oplus\_rect} will be (\texttt{$P$ a b (oplus a b)}),
-where $P$ is an arbitrary function of type 
-\texttt{T1 -> T1 -> T1 -> Type}
-
-\item The premises of \texttt{oplus\_rect} will describe how to build an induction 
-on the graph of \texttt{oplus}.
-\end{itemize}
-
-We are now ready to state and prove \texttt{oplus\_rect}, and the reader
-will note that the statement is longer than the proof script itself,
-which is a standard proof by induction, simplification and case-analysis 
-that follows  \texttt{oplus}'s definition.
-
-We associate also a tactic to the application of \texttt{oplus\_rect}.
-
-\begin{Coqsrc}
- Lemma oplus_rect:
-      forall P: T1 -> T1 -> T1 -> Type, 
-        (forall a:T1, P zero a a) ->
-        (forall a: T1, P a zero a) ->
-        (forall a1 n1 b1 a2 n2 b2 o,
-           compare a1 a2 = Gt ->
-           P b1 (ocons a2 n2 b2) o ->
-           P (ocons a1 n1 b1) (ocons a2 n2 b2)
-             (ocons a1 n1 o)) ->
-        (forall a1 n1 b1 a2 n2 b2 o,
-           compare a1 a2 = Lt ->
-           P (ocons a1 n1 b1) b2 o ->
-           P (ocons a1 n1 b1) (ocons a2 n2 b2) 
-           (ocons a2 n2 o)) ->
-        (forall a1 n1 b1 a2 n2 b2 o,
-           compare a1 a2 = Eq ->
-           P b1 b2 o ->
-          P (ocons a1 n1 b1) (ocons a2 n2 b2)
-            (ocons a1 (S (n1 + n2)%nat) o)) ->
-         forall a b, P a b (oplus a b).
-Proof with auto.
-   induction a.
-   -    intro; simpl; destruct b;auto.
-   -   induction b.
-       + apply X0.
-       + case_eq (compare a1 b1).
-         * intro Comp; unfold oplus; rewrite Comp.
-           cbn; apply X3 ...
-         * intro Comp; cbn; rewrite Comp; apply X2...
-         * intro Comp; cbn; rewrite Comp ...
- Defined.
-
-Ltac oplus_induction a b:= pattern (oplus a b); apply oplus_rect.
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
-The induction principle \texttt{oplus\_rect} is still unused in our development. 
-Please build some nice examples of application.
-\end{exercise}
+% \subsubsection{A hand-made induction principle}
+% \label{sec:orgheadline85}
+
+% \index{Coq!Commands!Functional Scheme}
+
+% \emph{Coq} contains a command  \texttt{Functional Scheme} that 
+% generates induction principles which correspond to recursive functions.
+% Unfortunately, the current version ( \texttt{8.11.0} ) doesn't work on \texttt{oplus},
+% probably because of the inner \texttt{Fix}.
+
+% \begin{Coqsrc}
+% Functional Scheme oplus_ind := Induction for oplus Sort Prop.
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+% Error: Anomaly "todo." Please report at http://coq.inria.fr/bugs/.
+% \end{Coqanswer}
+
+% Fortunately, it's a good exercise for a semi-experienced user, to write
+% her/him-self induction principles similar to the ones returned by
+% \texttt{Functional Scheme}.
+
+% \begin{itemize}
+% \item First, we chose to write a version for sort \texttt{Type}, since versions
+% for sorts \texttt{Prop} and \texttt{Set} can be easily derived from
+% the former one. According to \emph{Coq}'s naming politics, we will call our 
+% principle \texttt{oplus\_rect}
+
+% \item The conclusion of \texttt{oplus\_rect} will be (\texttt{$P$ a b (oplus a b)}),
+% where $P$ is an arbitrary function of type 
+% \texttt{T1 -> T1 -> T1 -> Type}
+
+% \item The premises of \texttt{oplus\_rect} will describe how to build an induction 
+% on the graph of \texttt{oplus}.
+% \end{itemize}
+
+% We are now ready to state and prove \texttt{oplus\_rect}, and the reader
+% will note that the statement is longer than the proof script itself,
+% which is a standard proof by induction, simplification and case-analysis 
+% that follows  \texttt{oplus}'s definition.
+
+% We associate also a tactic to the application of \texttt{oplus\_rect}.
+
+% \begin{Coqsrc}
+%  Lemma oplus_rect:
+%       forall P: T1 -> T1 -> T1 -> Type, 
+%         (forall a:T1, P zero a a) ->
+%         (forall a: T1, P a zero a) ->
+%         (forall a1 n1 b1 a2 n2 b2 o,
+%            compare a1 a2 = Gt ->
+%            P b1 (ocons a2 n2 b2) o ->
+%            P (ocons a1 n1 b1) (ocons a2 n2 b2)
+%              (ocons a1 n1 o)) ->
+%         (forall a1 n1 b1 a2 n2 b2 o,
+%            compare a1 a2 = Lt ->
+%            P (ocons a1 n1 b1) b2 o ->
+%            P (ocons a1 n1 b1) (ocons a2 n2 b2) 
+%            (ocons a2 n2 o)) ->
+%         (forall a1 n1 b1 a2 n2 b2 o,
+%            compare a1 a2 = Eq ->
+%            P b1 b2 o ->
+%           P (ocons a1 n1 b1) (ocons a2 n2 b2)
+%             (ocons a1 (S (n1 + n2)%nat) o)) ->
+%          forall a b, P a b (oplus a b).
+% Proof with auto.
+%    induction a.
+%    -    intro; simpl; destruct b;auto.
+%    -   induction b.
+%        + apply X0.
+%        + case_eq (compare a1 b1).
+%          * intro Comp; unfold oplus; rewrite Comp.
+%            cbn; apply X3 ...
+%          * intro Comp; cbn; rewrite Comp; apply X2...
+%          * intro Comp; cbn; rewrite Comp ...
+%  Defined.
+
+% Ltac oplus_induction a b:= pattern (oplus a b); apply oplus_rect.
+% \end{Coqsrc}
+
+% \index{Exercises}
+
+% \begin{exercise}
+% The induction principle \texttt{oplus\_rect} is still unused in our development. 
+% Please build some nice examples of application.
+% \end{exercise}
 

### 2020-10-19 doc/hydras.tex
new tactics for navigating in the hydra

@@
 \end{quote}
+
+% \begin{verse}
+% Hydra hydra hydram \\
+
+% Hydrae hydrae hydra \\
+
+% Hydrae hydrae hydras \\
+
+% Hydrarum hydris hydris\\
+% \end{verse}
 
@@
 Let us prove the transformation of \texttt{Hy'} into \texttt{Hy''} (see Fig.~\vref{fig:Hy3}).
+We use an experimental set of tactics for specifying place where the interaction between Hercules and the hydra holds.
 
@@
 Proof.
-  right. right. left.  right. left. left. left. split. left.
+    (** move to 2nd sub-hydra (0-based indices) *) R2_up 1. 
+    (** move to first sub-hydra *)  R2_up 0.
+    (** we're at distance 2 from the to-be-chopped-off head 
+        let's go to the first daughter, then chop-off the leftmost head *)
+    last_two 0 0. 
 Qed.
@@
 
+\index{Projects}
+
+\begin{project}
+The library \texttt{Hessenberg} looks too long (proof scripts and compilation).
+Please try to make it simpler and more efficient!
+Thanks!
+\end{project}
+
 \subsection{A measure for hydra-battle termination}

### 2020-10-20 doc/hydras.tex
new tactics for navigating in the hydra
last push before move

@@
 \subsubsection{Example}
-Let us prove the transformation of \texttt{Hy'} into \texttt{Hy''} (see Fig.~\vref{fig:Hy3}).
-We use an experimental set of tactics for specifying place where the interaction between Hercules and the hydra holds.
+Let us prove the transformation of \texttt{Hy'} into \texttt{Hy''} (see Fig.~\vref{fig:Hy3}). We use an experimental set of tactics for specifying the place where the 
+interaction between Hercules and the hydra holds.
 
@@
     (** we're at distance 2 from the to-be-chopped-off head 
-        let's go to the first daughter, then chop-off the leftmost head *)
-    last_two 0 0. 
+        let's go to the first daughter, 
+        then chop-off the leftmost head *)  r2_d2  0 0. 
 Qed.

### 2020-10-21 doc/hydras.tex
initialization of a documentation on Gamma0; strongly critical ordinals

@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=11cm]{epsilon0.jpg}
-  \caption{Discrete mathematics and formal proofs}
-  \label{fig:gamma0}
-\end{figure}
  We also hope to provide a little clarification on infinity (both potential and actual) through the notions of function, computation, limit,
@@
 
+\chapter{The ordinal \texorpdfstring{$\Gamma_0$}{Gamma0}}
+
+\textbf{To write!}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=11cm]{epsilon0.jpg}
+  \caption{Discrete mathematics and formal proofs}
+  \label{fig:gamma0}
+\end{figure}
+
 \chapter{Appendices}

### 2020-10-22 doc/hydras.tex
redaction of chapter on Gamma0 (partial)

@@
 
-In~\cite{grimm:hal-00911710}, José Grimm establishes the consistency between or ordinal notations (\texttt{T1} and \texttt{T2} (Veblen normal form) and his implementation
+In~\cite{grimm:hal-00911710}, José Grimm establishes the consistency between our ordinal notations (\texttt{T1} and \texttt{T2} (Veblen normal form) and his implementation
 of ordinal numbers after Bourbaki's set theory.
@@
 
-\textbf{To write!}
+\section{Introduction}
+We present a notation system for the ordinal $\Gamma_0$, following Chapter V, Section 14 of~\cite{schutte}: ``A notation system fir the ordinals $<\Gamma_0$''.
+We try to be as close as possible to Schütte's text and usual practices of \coq{} developments.
+
+The ordinal $\Gamma_0$ is defined in Section 13 of ~\cite{schutte} as the least \emph{strongly critical ordinal}. It is widely known as the \emph{Feferman-Schütte ordinal}.
+Section V, 13 of~\cite{schutte} defines \emph{strongly critical} and
+\emph{maximal $\alpha$-critical} ordinals.
+
+\noindent\emph{From \href{../src/html/hydras.Schutte.Critical.html\#strongly_critical}%
+{\texttt{Schutte.Critical}}}
+
+\begin{Coqsrc}
+Definition strongly_critical alpha := In (Cr alpha) alpha.
+
+Definition maximal_critical alpha : Ensemble Ord :=
+  fun gamma =>
+    In (Cr alpha) gamma /\
+    forall ksi, alpha < ksi -> ~ In (Cr alpha) ksi.
+\end{Coqsrc}
+
+\index{Projects}
+
+\begin{project}
+In order to define $\Gamma_0$ as a n inhabitant of type \texttt{Ord}, you should prove that the set of strongly critical ordinal is not empty, hence has a least element.
+\end{project}
+
+In the present version of this development, we will only study $\Gamma_0$ as a notation system, much more powerful than the ordinal notation for $\epsilon_0$.
+
+%\index{Projects}
+%
+%\begin{project}
+% Schûtte's section 13 of~\cite{schutte} contain several definitions and lemmas which give another view on $\Gamma_0$.  We leave it as a project to implement them in \coq{}.  
+%\end{project}
 
@@
 \end{figure}
+
+\section{The type \texttt{T2} of ordinal terms}
+
+The notation system for ordinals less than $\gamma_0$ comes from the following theorem of~\cite{schutte}, where $\psi\,\alpha$ is the ordering function 
+of the set of maximal $\alpha$-critical ordinals.
+
+\index{Projects}
+\begin{project}
+This theorem is not formally proved in this development yet. It should be!
+\end{project}
+
+\begin{quote}
+  Any ordinal $\not= 0$ which is not strongly critical can be expressed in terms of $+$ and $\psi$.
+\end{quote}
+
+Like in Chapter~\ref{chap:T1}, we define an inductive type with two constructors, one for $0$, the other for the construction $\psi(\alpha,\beta)\times(n+1)+\gamma$, adapting a Manolios-Vroon-like notation~\cite{Manolios2005} to
+\emph{Veblen normal forms}.
+
+\noindent\emph{From \href{../src/html/hydras.Gamma0.T2.html\#T2}%
+{\texttt{Gamma0.T2}}}
+\begin{Coqsrc}
+(**  [gcons alpha beta n gamma] is : [psi(alpha,beta)*(S n)+ gamma]  *)
+
+Inductive T2 : Set :=
+| zero : T2
+| gcons : T2 -> T2  -> nat -> T2 -> T2.
+
+Notation "[ alpha , beta ]" := (gcons alpha beta 0 zero)
+                                 (at level 0): t2_scope.
+\end{Coqsrc}
+
+Like in chapter~\ref{chap:T1}, we get familiar with the type \texttt{T2} by recognising simple constructs like finite ordinals, $\omega$, etc., as inhabitants of ÷texttt{T2}.
+
+\begin{Coqsrc}
+Notation  "'one'"  := [zero,zero] : T2_scope.
+
+Notation "'FS' n" := (gcons zero zero n zero) (at level 10) : T2_scope.
+
+(** the [n]-th ordinal  *)
+
+Definition fin (n:nat) := match n with 0 => zero | S p => FS p end.
+
+Notation "'omega'"  := [zero,one] : T2_scope.
+\end{Coqsrc}
+
+We are also able to give a term to $\epsilon_0$, and even to generalize it!.
+
+\begin{Coqsrc}
+Definition epsilon0  := [one,zero].
+
+Definition epsilon alpha := [one, alpha].
+\end{Coqsrc}
+
+The justification of the definition of \texttt{epsilon0} will be given in further sections.
+
+\section{How big is \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}?}
+
+Let us define a strict order on type \texttt{T2}. The following definition is 
+an adaptation of Schütte's, taking into account the multiplications by a natural number (inspired by~\cite{Manolios2005}, and also present in \texttt{T1}).
+
+\begin{Coqsrc}
+Inductive lt : T2 -> T2 -> Prop :=
+| (* 1 *) 
+ lt_1 : forall alpha beta n gamma,  zero t2< gcons alpha beta n gamma
+| (* 2 *)
+ lt_2 : forall alpha1 alpha2 beta1 beta2 n1 n2 gamma1 gamma2, 
+                alpha1 t2< alpha2 ->
+                beta1 t2< gcons alpha2 beta2 0 zero ->
+               gcons alpha1 beta1 n1 gamma1 t2<
+               gcons alpha2 beta2 n2 gamma2
+| (* 3 *)
+ lt_3 : forall alpha1  beta1 beta2 n1 n2 gamma1 gamma2, 
+               beta1 t2< beta2 ->
+               gcons alpha1 beta1 n1 gamma1 t2<
+               gcons alpha1 beta2 n2 gamma2
+
+| (* 4 *)
+ lt_4 : forall alpha1 alpha2 beta1 beta2 n1 n2 gamma1 gamma2, 
+               alpha2 t2< alpha1 ->
+               gcons alpha1 beta1 0 zero t2< beta2 ->
+               gcons alpha1 beta1 n1 gamma1 t2<
+               gcons alpha2 beta2 n2 gamma2
+
+| (* 5 *)
+lt_5 : forall alpha1 alpha2 beta1 n1 n2 gamma1 gamma2, 
+               alpha2 t2< alpha1 ->
+               gcons alpha1 beta1 n1 gamma1 t2<
+               gcons alpha2  (gcons alpha1 beta1 0 zero) n2 gamma2
+
+| (* 6 *)
+lt_6 : forall alpha1 beta1  n1  n2 gamma1 gamma2,  (n1 < n2)%nat ->
+                                    gcons alpha1 beta1 n1 gamma1 t2< 
+                                    gcons alpha1 beta1 n2 gamma2
+
+| (* 7 *)
+  lt_7 : forall alpha1 beta1 n1   gamma1 gamma2,  gamma1 t2< gamma2 ->
+                                      gcons alpha1 beta1 n1 gamma1 t2<
+                                      gcons alpha1 beta1 n1 gamma2
+where  "o1 t2< o2" := (lt o1 o2): T2_scope.
+\end{Coqsrc}
+
+Here are two simple verifications: $\omega$ is less than $\epsilon_0$, whic is less than $\psi(2,1)$.
+
+\begin{Coqsrc}
+Goal omega t2< epsilon0.
+Proof.
+  unfold epsilon0; auto with T2.
+Qed.
+
+Goal epsilon0 t2< [2,1].
+Proof.
+  unfold epsilon0;  apply lt_2.
+  - apply lt_6; auto with arith.
+  - auto with T2.
+Qed.
+\end{Coqsrc}
+
+\index{Projects}
+\begin{project}
+Write a tactic that solves automatically goals of the form (\texttt{$\alpha$ t2< $\beta$}), where $\alpha$ and $\beta$ are closed terms of type \texttt{T2}.
+\end{project}
 

### 2020-10-23 doc/hydras.tex
Writing a chapter on Gamma0

@@
 \begin{remark}
-The alternate proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering is available in the library \href{../src/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
+The alternate proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the library \href{../src/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
  \end{remark}
@@
 \begin{Coqsrc}
-Definition epsilon0  := [one,zero].
+Notation "'epsilon0'"  := ([one,zero]) : T2_scope.
 
@@
 an adaptation of Schütte's, taking into account the multiplications by a natural number (inspired by~\cite{Manolios2005}, and also present in \texttt{T1}).
+
+\label{sect:t2-lt-def}
 
@@
                alpha2 t2< alpha1 ->
-               gcons alpha1 beta1 0 zero t2< beta2 ->
+               [alpha1, beta1] t2< beta2 ->
                gcons alpha1 beta1 n1 gamma1 t2<
@@
                gcons alpha1 beta1 n1 gamma1 t2<
-               gcons alpha2  (gcons alpha1 beta1 0 zero) n2 gamma2
+               gcons alpha2  [alpha1, beta1] n2 gamma2
 
@@
 where  "o1 t2< o2" := (lt o1 o2): T2_scope.
-\end{Coqsrc}
-
-Here are two simple verifications: $\omega$ is less than $\epsilon_0$, whic is less than $\psi(2,1)$.
-
-\begin{Coqsrc}
-Goal omega t2< epsilon0.
+
+Hint Constructors lt : T2.
+\end{Coqsrc}
+
+\subsection{Examples}
+Seven constructors! In order to get accustomed with this definition, let us look at a small set of examples, covering all the constructors of \texttt{lt}.
+
+\subsubsection*{Proof of $0<\epsilon_0$}
+
+\begin{Coqsrc}
+Example Ex1: 0 t2< epsilon0.
+Proof.  constructor 1. Qed.
+\end{Coqsrc}
+
+\subsubsection*{Proof of $\omega<\epsilon_0$}
+
+\begin{Coqsrc}
+Example Ex2: omega t2< epsilon0.
+Proof. info_auto with T2. (* uses lt_1 and lt_2 *) Qed.
+\end{Coqsrc}
+
+\subsubsection*{Proof of $\psi(\omega,8]\times 13+56 < \psi(\omega,8]\times 13+57 $}
+
+\begin{Coqsrc}
+Example Ex3: gcons omega 8 12 56 t2<  gcons omega 8 12 57.
 Proof.
-  unfold epsilon0; auto with T2.
+  constructor 7; constructor 6; auto with arith.
 Qed.
-
-Goal epsilon0 t2< [2,1].
+\end{Coqsrc}
+
+\subsubsection*{Proof of $\epsilon_0<\psi(2,1)$}
+
+\begin{Coqsrc}
+Example Ex4: epsilon0 t2< [2,1].
 Proof.
-  unfold epsilon0;  apply lt_2.
-  - apply lt_6; auto with arith.
-  - auto with T2.
+   constructor 2; auto with T2.
+   - constructor 6; auto with arith.
+Qed.
+\end{Coqsrc}
+
+\subsubsection*{Proof of $\psi(2,1)<\psi(2,3)$}
+
+\begin{Coqsrc}
+Example Ex5 : [2,1] t2< [2,3].
+Proof.
+  constructor 3; auto with T2.
+  - constructor 6; auto with arith.
+Qed.
+\end{Coqsrc}
+
+\subsubsection*{Proof of $\psi(1,0)\times 13+ \omega < \psi(0,\psi(2,1))$}
+
+\begin{Coqsrc}
+Example Ex6 : gcons 1 0 12 omega t2< [0,[2,1]].
+Proof.
+  constructor 4.
+  - constructor 1.
+  - constructor 2.
+    + constructor 6; auto with arith.
+    + constructor 1.
+Qed.
+\end{Coqsrc}
+
+\subsubsection*{Proof of $\psi(2,1)\times 43 + \epsilon_0 < \psi(1,\psi(2,1))$}
+
+\begin{Coqsrc}
+Example Ex7 : gcons 2 1 42 epsilon0 t2< [1, [2,1]].
+Proof.
+ constructor 5.
+ constructor 6; auto with arith.
 Qed.
@@
 
+\section{Veblen normal forms}
+\begin{definition}
+  A term of the form $\psi(\alpha_1,\beta_1)\times n_1+ \psi(\alpha_2,\beta_2)\times n_2+\dots+\psi(\alpha_k,\beta_k)\times n_k$ is said to be in
+ \emph{[Veblen] normal form} if for every $i<n$, $\psi(\alpha_i,\beta_i)<\psi(\alpha_{i+1},\beta_{i+1})$, all the $\alpha_i$ and $\beta_i$ are in normal form, and all the $n_i$ are strictly positive integers.
+\end{definition}
+
+\begin{Coqsrc}
+Inductive nf : T2 -> Prop :=
+| zero_nf : nf zero
+| single_nf : forall a b n, nf a ->  nf b -> nf (gcons a b n zero)
+| gcons_nf : forall a b n a' b' n' c', 
+                      [a', b'] t2< [a, b]  -> 
+                      nf a -> nf b -> 
+                      nf(gcons a' b' n' c')-> 
+                      nf(gcons a b n (gcons a' b' n' c')).
+\end{Coqsrc}
+
+Let us look at some positive examples (we have to prove some inversion lemmas before proving counter-examples).
+
+\begin{Coqsrc}
+Lemma  nf_finite i : nf (fin (S i)).
+Proof.
+   constructor 2; auto with T2.
+Qed.
+
+Lemma nf_epsilon0 : nf epsilon0.
+Proof. constructor 2; auto with T2. Qed.
+
+Example Ex8: nf (gcons 2 1 42 epsilon0).
+Proof.
+  constructor 3; auto with T2.
+  - apply Ex4.
+  - apply nf_finite.
+  - apply nf_finite.
+Qed.
+
+\end{Coqsrc}
 \chapter{Appendices}

### 2020-10-24 doc/hydras.tex
First draft of the chapter on Gamma0 (to be reviewed).

@@
 \begin{remark}
+\label{remark:a3pat}
 The alternate proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the library \href{../src/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
@@
 
+\label{instance-epsilon0}
 \begin{Coqsrc}
@@
 
+\label{sect:phi-schutte}
+
 \begin{Coqsrc}
@@
 
-\chapter{The ordinal \texorpdfstring{$\Gamma_0$}{Gamma0}}
+\chapter{The ordinal \texorpdfstring{$\Gamma_0$}{Gamma0} first draft)}
 
@@
 \subsubsection*{Proof of $\psi(1,0)\times 13+ \omega < \psi(0,\psi(2,1))$}
+\label{sect:ex6-first-proof}
 
@@
 \begin{Coqsrc}
-Lemma  nf_finite i : nf (fin (S i)).
+Lemma  nf_fin i : nf (fin i).
 Proof.
-   constructor 2; auto with T2.
+  destruct i.
+  - auto with T2.
+  - constructor 2; auto with T2.
 Qed.
+
+Lemma nf_omega : nf omega.
+Proof.  compute; auto with T2. Qed.
 
@@
 Proof. constructor 2; auto with T2. Qed.
+
+Lemma nf_epsilon : forall alpha, nf alpha -> nf (epsilon alpha).
+Proof. compute; auto with T2. Qed.
 
@@
   - apply Ex4.
-  - apply nf_finite.
-  - apply nf_finite.
+  - apply nf_fin.
+  - apply nf_fin.
 Qed.
-
-\end{Coqsrc}
+\end{Coqsrc}
+
+\subsection{Length of a term}
+
+The notion of \emph{term length} is introduced by Schütte as a helper for proving (at least) the \emph{trichotomy} property and transitivity of the strict order \texttt{lt} on \texttt{T2}.
+These properties are proved by induction on length.
+
+\begin{Coqsrc}
+Fixpoint nbterms (t:T2) : nat :=
+  match t with zero => 0
+             | gcons a b n v => (S n) + nbterms v
+  end.
+
+Fixpoint t2_length (t:T2) : nat :=
+  match t  with 
+    zero => 0
+  | gcons a b n v => 
+       nbterms (gcons a b n v) + 
+      2 * (Max.max (t2_length a)
+                              (Max.max (t2_length b) 
+                                                (t2_length_aux v)))
+  end
+with t2_length_aux (t:T2) : nat :=
+ match t with zero => 0
+            | gcons a b n v =>
+               Max.max (t2_length a) 
+                                (Max.max (t2_length b) (t2_length_aux v))
+ end.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Compute t2_length (gcons 2 1 42 epsilon0).
+\end{Coqsrc}
+
+\begin{Coqanswer}
+ = 48 : nat
+\end{Coqanswer}
+
+\subsection{Trichotomy}
+
+\emph{Trichotomy} is another name for the well-known property of decidable total ordering (like Standard Library's \texttt{Compare\_dec.lt\_eq\_lt\_dec}).
+
+We first prove by induction on $l$ the following lemma:
+
+\noindent\emph{From \href{../src/html/hydras.Gamma0.Gamma0\#tricho_aux}%
+{\texttt{Gamma0.Gamma0}}}
+
+\begin{Coqsrc}
+Lemma tricho_aux (l: nat) : forall t t' :T2,
+      t2_length t + t2_length t' < l  ->
+      {t t2< t'} + {t = t'} + {t' t2<  t}.
+\end{Coqsrc}
+
+Then we get our version of \texttt{lt\_eq\_lt\_dec}, and derive a comparison function;
+
+\begin{Coqsrc}
+Definition lt_eq_lt_dec (t t': T2) : {t t2< t'}+{t = t'}+{t' t2<  t}.
+Proof.
+  eapply tricho_aux.
+  eapply lt_n_Sn.
+Defined.
+
+Definition compare (t1 t2 : T2) : comparison := 
+  match lt_eq_lt_dec t1 t2 with
+  | inleft (left _) => Lt
+  | inleft (right _) => Eq
+  | inright _ => Gt
+  end.
+\end{Coqsrc}
+
+With the help of \texttt{compare}, we get a boolean version of \texttt{nf}
+(being in Veblen normal form).
+
+\begin{Coqsrc}
+Fixpoint nfb (alpha : T2) : bool :=
+  match alpha with
+    zero => true
+  | gcons a b n zero => andb (nfb a) (nfb b)
+  | gcons a b n ((gcons a' b' n' c') as c) =>
+    match compare [a', b'] [a, b] with
+           Lt => andb (nfb a) (andb (nfb b) (nfb c))
+           | _ => false
+           end
+end.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Compute compare (gcons 2 1 42 epsilon0) [2,2].
+\end{Coqsrc}
+
+\begin{Coqanswer}
+   = Lt
+     : comparison
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Compute nfb  (gcons 2 1 42 epsilon0).
+\end{Coqsrc}
+
+\begin{Coqanswer}
+   = true
+     : bool
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Compute nfb (gcons 2 1 42 (gcons 2 2 4 epsilon0)).
+\end{Coqsrc}
+
+\begin{Coqanswer}
+   = false
+     : bool
+\end{Coqanswer}
+
+\begin{remark}
+The connexion between the predicate \texttt{nf} and the relation \texttt{lt} on one part, and the functions \texttt{nfb} and \texttt{compare} on the other, is expressed by the following lemmas:
+
+\begin{Coqsrc}
+Lemma nfb_equiv gamma : nfb gamma = true <-> nf gamma.
+
+Lemma compare_correct alpha beta :
+  CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
+              (compare alpha beta).
+\end{Coqsrc}
+
+The function \texttt{compare} helps to make easier proofs of inequalities of
+closed terms of type \texttt{T2}.
+
+First, we prove a lemma:
+
+\begin{Coqsrc}
+Lemma compare_Lt : forall alpha beta, compare alpha beta = Lt -> 
+                                         alpha t2< beta.
+Proof.
+  intros alpha beta; destruct (compare_correct alpha beta);
+    trivial; discriminate. 
+Qed.
+\end{Coqsrc}
+
+Then, we give another version of the proof of Sect.~\vref{sect:ex6-first-proof}.
+
+\begin{Coqsrc}
+Example Ex6 : gcons 1 0 12 omega t2< [0,[2,1]].
+Proof. now apply compare_Lt. Qed.
+\end{Coqsrc}
+
+\end{remark}
+
+\section{Main functions on \texttt{T2}}
+
+\subsection{Successor}
+The successor function is defined by structural recursion.
+
+\noindent\emph{From \href{../src/html/hydras.Gamma0.T2.html\#succ}%
+{\texttt{Gamma0.T2}}}
+\begin{Coqsrc}
+Fixpoint succ (a:T2) : T2 :=
+ match a with zero => one
+             | gcons zero zero n c => fin (S (S n))
+             | gcons a b n c => gcons a b n (succ c)
+ end.
+\end{Coqsrc}
+
+\subsection{Addition}
+
+Like for Cantor normal forms (see Sect.\~ref{sect:infix-plus-T1})  the definition of addition in \texttt{T2}  requires comparison between ordinal terms.
+
+\begin{Coqsrc}
+Fixpoint plus (t1 t2 : T2) {struct t1}:T2 :=
+  match t1,t2 with
+  |  zero, y  => y
+  |  x, zero => x
+  |  gcons a b n c, gcons a' b' n' c' =>
+     (match compare (gcons a b 0 zero)
+                    (gcons a' b' 0 zero)
+      with | Lt => gcons a' b' n' c'
+      | Gt =>
+        (gcons a b n
+              (c +
+               (gcons a' b' n' c')))
+      | Eq  => (gcons a b (S(n+n')) c')
+      end)
+  end
+where "alpha + beta" := (plus alpha beta): T2_scope.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Example Ex7 : 3 + epsilon0 = epsilon0.
+Proof. trivial. Qed.
+\end{Coqsrc}
+
+\subsection{The Veblen function \texorpdfstring{$\varphi$}{\texttt{phi}}}
+
+The enumeration function of critical ordinals, presented in Sect.\~vref{sect:phi-schutte}, is recursively defined in type \texttt{T2}.
+
+\begin{Coqsrc}
+Definition  phi (alpha beta : T2) : T2 :=
+  match beta with zero => [alpha, beta] 
+             | [b1, b2] => 
+               (match compare alpha b1
+                with Datatypes.Lt => [b1, b2 ]
+                | _ => [alpha,[b1, b2]]
+                end)
+             | gcons b1 b2 0 (gcons zero zero  n zero) => 
+               (match compare alpha b1
+                with  Datatypes.Lt => 
+                      [alpha, (gcons b1 b2 0 (fin n))]
+                | _ =>  [alpha, (gcons b1 b2 0 (fin (S n)))]
+                end)
+             | any_beta => [alpha, any_beta]
+  end.
+\end{Coqsrc}
+
+Despite its complexity, the function \texttt{phi} is well adapted to proofs by simplification or computation.
+\begin{Coqsrc}
+Example Ex8:  phi 1 (succ epsilon0) = [1, [1,0] + 1].
+Proof. reflexivity. Qed.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+(**  All epsilons are fixpoints of phi 0 *)
+
+Theorem epsilon_fxp : forall beta, phi zero (epsilon beta) =
+                                   epsilon beta.
+Proof. reflexivity. Qed.
+
+Theorem epsilon0_fxp : epsilon0 = phi zero epsilon0.
+Proof. apply epsilon_fxp. Qed.
+\end{Coqsrc}
+
+The relation between the constructor $\psi$ and the function $\phi$ is
+studied in~\cite{schutte}, and partially implemented in this development.
+\emph{Please contribute!}
+
+For instance, the following theorem states that, if $\gamma$ is the sum of a limit ordinal $\beta$ and a finite ordinal $n$, and $\beta$ is a fixpoint of
+$\varphi(\alpha)$, then $\psi(\alpha,\gamma)=\phi(\alpha, \gamma+1)$.
+
+\begin{Coqanswer}
+phi_psi :
+forall (alpha : T2) [beta gamma : T2] [n : nat],
+nf gamma ->
+limit_plus_fin beta n gamma ->
+phi alpha beta = beta -> [alpha, gamma] = phi alpha (succ gamma)
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Example Ex9 : [zero, epsilon 2 + 4] = phi 0 (epsilon 2 + 5).
+Proof. trivial. Qed.
+\end{Coqsrc}
+
+On the other hand, $\varphi$ can be expressed in terms of $\psi$.
+
+\begin{Coqanswer}
+phi_of_psi:
+  forall a b1 b2 : T2,
+  phi a [b1, b2] = (if lt_ge_dec a b1 then [b1, b2] else [a, [b1, b2]])
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Example Ex10 : phi omega [epsilon0, 5] = [epsilon0, 5].
+Proof. reflexivity. Qed.
+\end{Coqsrc}
+
+\section{An ordinal notation for \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}}
+
+In order to consider type \texttt{T2} as an ordinal notation, we have to build an instance of class \texttt{ON} (See Definition page~\pageref{types:ON}).
+
+First, we define a type that contains only terms in Veblen normal form, and redefine \texttt{lt} and \texttt{compare} by delegation.
+
+\begin{Coqsrc}
+Class G0 := mkg0 {vnf : T2; vnf_ok : nfb vnf}.
+
+Definition lt (alpha beta : G0) := T2.lt (@vnf alpha) (@vnf beta).
+
+Definition compare alpha beta := Gamma0.compare (@vnf alpha) (@vnf beta).
+\end{Coqsrc}
+
+Then we, prove that \texttt{lt} is a well-founded strict order and that the
+function \texttt{compare} is correct.
+
+\begin{Coqsrc}
+Instance lt_sto : StrictOrder lt.
+
+Lemma lt_wf : well_founded lt.
+
+Lemma compare_correct alpha beta :
+  CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
+              (compare alpha beta).
+
+Instance Gamma0: ON lt  compare.
+Proof.
+  split.
+  - apply lt_sto.
+  - apply lt_wf. 
+  - apply compare_correct.
+Qed.
+\end{Coqsrc}
+
+\begin{remark}
+The proof of \texttt{lt\_wf} has been written by \'Evelyne Contejean, using her library on the recursive path ordering (see also remark~\vref{remark:a3pat}).
+\end{remark}
+
+\index{Projects}
+\begin{project}
+Prove that \texttt{Epsilon0} (page~\pageref{instance-epsilon0})
+is a sub-notation system of \texttt{Gamma0}.
+
+Prove that the implemantations of \texttt{succ}, \texttt{+}, $\varphi_0$, etc.
+are compatible in both notation systems.
+
+Note that a function \texttt{T1\_inj} from \texttt{T1} to \texttt{T2} has already been defined. It may help to complete the task.
+\end{project}
+
 \chapter{Appendices}

### 2020-10-24 doc/hydras.tex
Corrections to Gamma0 (chapter and scripts)

@@
 
-\chapter{The ordinal \texorpdfstring{$\Gamma_0$}{Gamma0} first draft)}
+\chapter{The ordinal \texorpdfstring{$\Gamma_0$}{Gamma0} (first draft)}
+
+\emph{This chapter and the files it presents are still very incomplete, considering the impressive properties of $\Gamma_0$~\cite{Gallier91}.
+We hope to add new material soon, and accept contributions!}
 
 \section{Introduction}
-We present a notation system for the ordinal $\Gamma_0$, following Chapter V, Section 14 of~\cite{schutte}: ``A notation system fir the ordinals $<\Gamma_0$''.
+We present a notation system for the ordinal $\Gamma_0$, following Chapter V, Section 14 of~\cite{schutte}: ``A notation system for the ordinals $<\Gamma_0$''.
 We try to be as close as possible to Schütte's text and usual practices of \coq{} developments.
@@
      (match compare (gcons a b 0 zero)
-                    (gcons a' b' 0 zero)
-      with | Lt => gcons a' b' n' c'
-      | Gt =>
-        (gcons a b n
-              (c +
-               (gcons a' b' n' c')))
-      | Eq  => (gcons a b (S(n+n')) c')
+                    (gcons a' b' 0 zero) with
+      | Lt => gcons a' b' n' c'
+      | Gt => gcons a b n (c + gcons a' b' n' c')
+      | Eq => gcons a b (S(n+n')) c'
       end)

### 2020-10-25 doc/hydras.tex
Document gamma0 (#5)
* initialization of a documentation on Gamma0; strongly critical ordinals

* redaction of chapter on Gamma0 (partial)

* Writing a chapter on Gamma0

* First draft of the chapter on Gamma0 (to be reviewed).

* Corrections to Gamma0 (chapter and scripts)

* First draft of the chapter on Gamma0

@@
 
-The ordinal $\omega^2$ (also called $\varphi_0(2)$, see Chap.~\ref{chap:schutte}), is represented by the lexicographically ordered cartesian product $\mathbb{N}\times\mathbb{N}$.
+The ordinal $\omega^2$ (also called $\phi_0(2)$, see Chap.~\ref{chap:schutte}), is represented by the lexicographically ordered cartesian product $\mathbb{N}\times\mathbb{N}$.
 
@@
 Note that if the notation $N_A$ implements the ordinal 
-$\alpha$,  the new notation $\omega^{N_A}$ must implement the ordinal $\varphi_0(\alpha)$, a.k.a. $\omega^\alpha$ (see chapter~\ref{chap:schutte})
+$\alpha$,  the new notation $\omega^{N_A}$ must implement the ordinal $\phi_0(\alpha)$, a.k.a. $\omega^\alpha$ (see chapter~\ref{chap:schutte})
 
@@
 
-\subsubsection{The ordinal \(\omega^\alpha\), a.k.a. \(\varphi_0(\alpha)\)}
+\subsubsection{The ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
 \label{sec:orgheadline70}
@@
 \label{sec:orgheadline69}
-The name \(\varphi_0\)
+The name \(\phi_0\)
    comes from ordinal numbers theory. In~\cite{schutte}, Schütte defines 
-$\varphi_0$  as the ordering (\emph{i.e.} enumerating) function of the set  of \emph{additive principal ordinals} \emph{i.e.} strictly positive ordinals $\alpha$ that verify $\forall \beta<\alpha, \beta+\alpha=\alpha$.
-For Schütte,  $\omega^\alpha$ is just a notation for $\varphi_0(\alpha)$.
+$\phi_0$  as the ordering (\emph{i.e.} enumerating) function of the set  of \emph{additive principal ordinals} \emph{i.e.} strictly positive ordinals $\alpha$ that verify $\forall \beta<\alpha, \beta+\alpha=\alpha$.
+For Schütte,  $\omega^\alpha$ is just a notation for $\phi_0(\alpha)$.
 See also Chapter~\ref{chap:schutte} of this document.
@@
 \subsection{Pretty printing Ordinals in Cantor normal Form}
+\label{sect:ppT1}
 
@@
 \end{Coqanswer}
+
+\index{Projects}
+\begin{project}
+Design  (in \ocaml?) a set of tools for systematically pretty printing ordinal terms in Cantor normal form.
+\end{project}
 
@@
 We would like to get exact formulas for the ordinal $\omega^2$, a.k.a.
-$\varphi_0(2)$. This ordinal is the limit of the sequence $\omega\times i\;(i \in \mathbb{N}$. Thus, we have to study ordinals of this form, then use 
+$\phi_0(2)$. This ordinal is the limit of the sequence $\omega\times i\;(i \in \mathbb{N}$. Thus, we have to study ordinals of this form, then use 
 our lemma on limits.
@@
 the ordinal  $\omega^\alpha$, also written 
-$\varphi_0(\alpha)$. 
+$\phi_0(\alpha)$. 
 It is an opportunity to apply the definitions and results of the preceding section. 
-Indeed,  Schütte first defines a subset of $\mathbb{O}$: the set of additive principal ordinals, and $\varphi_0$  is just defined as the ordering function of this set.
+Indeed,  Schütte first defines a subset of $\mathbb{O}$: the set of additive principal ordinals, and $\phi_0$  is just defined as the ordering function of this set.
 
@@
 
-Let us call  $\varphi_0$ the ordering function of \texttt{AP}.
-In the mathematical text, we shall use indifferently the notations  $\omega^\alpha$ and$\varphi_0(\alpha)$.
+Let us call  $\phi_0$ the ordering function of \texttt{AP}.
+In the mathematical text, we shall use indifferently the notations  $\omega^\alpha$ and$\phi_0(\alpha)$.
 
@@
 
-Using $\varphi_0$, we can define recursively the set of finite omega-towers.
+Using $\phi_0$, we can define recursively the set of finite omega-towers.
 
@@
 
-The rest of our library \texttt{AP} is devoted to the proof of properties of additive principal ordinals, hence of the ordering function  $\varphi0$ and the ordinal $\epsilon_0$ (which we could not express within the type \texttt{T1}).
+The rest of our library \texttt{AP} is devoted to the proof of properties of additive principal ordinals, hence of the ordering function  $\phi0$ and the ordinal $\epsilon_0$ (which we could not express within the type \texttt{T1}).
 
@@
 
-\subsubsection{Properties of the function \texorpdfstring{$\varphi_0$}{phi0}}
-
-The ordering function $\varphi_0$ of the set \texttt{AP} is defined on the full set $\mathbb{O}$ and is continuous (Schütte says that this function is \emph{normal}).
+\subsubsection{Properties of the function \texorpdfstring{$\phi_0$}{phi0}}
+
+The ordering function $\phi_0$ of the set \texttt{AP} is defined on the full set $\mathbb{O}$ and is continuous (Schütte says that this function is \emph{normal}).
 
@@
 
-The following properties come from  the definition of $\varphi_0$ as the ordering function of \texttt{AP}.
+The following properties come from  the definition of $\phi_0$ as the ordering function of \texttt{AP}.
 It may be interesting to compare these proofs with the computational ones described in Chapter ~\ref{chap:T1}.
@@
 
-Please note that the lemma \texttt{omega\_eq} above, is consistent with the interpretation of the ordering function $\varphi_0$ as the exponential of basis $\omega$.
+Please note that the lemma \texttt{omega\_eq} above, is consistent with the interpretation of the ordering function $\phi_0$ as the exponential of basis $\omega$.
 Indeed we could have written this lemma with our alternative notation:
@@
           $\alpha$ is greater than or equal to the omega-tower of height $j$.
-         By monotonicity,  $\varphi_0(\alpha)$ is greater than or equal to 
+         By monotonicity,  $\phi_0(\alpha)$ is greater than or equal to 
         the omega-tower of height $j+1$, thus strictly greater than $\alpha$
@@
 Moreover,  $\epsilon_0$ is the least ordinal $\alpha$ that verifies the equality 
-$\alpha = \omega^\alpha$, in other words the least fixpoint of the function  $\varphi_0$.
+$\alpha = \omega^\alpha$, in other words the least fixpoint of the function  $\phi_0$.
 
@@
 The ordinal $\Gamma_0$ is defined in Section 13 of ~\cite{schutte} as the least \emph{strongly critical ordinal}. It is widely known as the \emph{Feferman-Schütte ordinal}.
+
 Section V, 13 of~\cite{schutte} defines \emph{strongly critical} and
 \emph{maximal $\alpha$-critical} ordinals.
+
+We give their definition in \coq's syntax.
+
+\vspace{4pt}
 
@@
     forall ksi, alpha < ksi -> ~ In (Cr alpha) ksi.
+
+Definition Gamma0 := the_least strongly_critical.
 \end{Coqsrc}
@@
 \index{Projects}
-
 \begin{project}
-In order to define $\Gamma_0$ as a n inhabitant of type \texttt{Ord}, you should prove that the set of strongly critical ordinal is not empty, hence has a least element.
+Prove that a (countable)  ordinal $\alpha$ is strongly critical iff $\phi \alpha 0=\alpha$ (Theorem 13.13 of~\cite{schutte} ). 
+\end{project}
+
+\index{Projects}
+\begin{project}
+Prove that the set of strongly critical ordinals is unbounded and closed (Theorem 13.14 of~\cite{schutte} ).
+Thus this set is not empty,  hence has a least element.
+Otherwise, the definition of $\Gamma_0$ above would be useless.
 \end{project}
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=11cm]{epsilon0.jpg}
-  \caption{Discrete mathematics and formal proofs}
-  \label{fig:gamma0}
-\end{figure}
-
 \section{The type \texttt{T2} of ordinal terms}
@@
 of the set of maximal $\alpha$-critical ordinals.
+
+\begin{quote}
+  Any ordinal $\not= 0$ which is not strongly critical can be expressed in terms of $+$ and $\psi$.
+\end{quote}
 
@@
 
-\begin{quote}
-  Any ordinal $\not= 0$ which is not strongly critical can be expressed in terms of $+$ and $\psi$.
-\end{quote}
-
 Like in Chapter~\ref{chap:T1}, we define an inductive type with two constructors, one for $0$, the other for the construction $\psi(\alpha,\beta)\times(n+1)+\gamma$, adapting a Manolios-Vroon-like notation~\cite{Manolios2005} to
@@
 
-Like in chapter~\ref{chap:T1}, we get familiar with the type \texttt{T2} by recognising simple constructs like finite ordinals, $\omega$, etc., as inhabitants of ÷texttt{T2}.
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=11cm]{epsilon0.jpg}
+  \caption{Veblen normal form}
+  \label{fig:gamma0}
+\end{figure}
+
+Like in chapter~\ref{chap:T1}, we get familiar with the type \texttt{T2} by recognising simple constructs like finite ordinals, $\omega$, etc., as inhabitants of \texttt{T2}.
 
@@
 
+(** The (n+1)-th finite ordinal *)
 Notation "'FS' n" := (gcons zero zero n zero) (at level 10) : T2_scope.
@@
 (** the [n]-th ordinal  *)
-
 Definition fin (n:nat) := match n with 0 => zero | S p => FS p end.
@@
 
-We are also able to give a term to $\epsilon_0$, and even to generalize it!.
-
 \begin{Coqsrc}
@@
 \end{Coqsrc}
-
-The justification of the definition of \texttt{epsilon0} will be given in further sections.
 
@@
 
+Seven constructors! In order to get accustomed with this definition, let us look at a small set of examples, covering all the constructors of \texttt{lt}.
+
 \subsection{Examples}
-Seven constructors! In order to get accustomed with this definition, let us look at a small set of examples, covering all the constructors of \texttt{lt}.
 
@@
 with t2_length_aux (t:T2) : nat :=
- match t with zero => 0
+ match t with 
+ | zero => 0
   | gcons a b n v =>
@@
 We first prove by induction on $l$ the following lemma:
+
+\vspace{4pt}
 
@@
 
-\subsection{The Veblen function \texorpdfstring{$\varphi$}{\texttt{phi}}}
+\subsection{The Veblen function \texorpdfstring{$\phi$}{\texttt{phi}}}
 
@@
 For instance, the following theorem states that, if $\gamma$ is the sum of a limit ordinal $\beta$ and a finite ordinal $n$, and $\beta$ is a fixpoint of
-$\varphi(\alpha)$, then $\psi(\alpha,\gamma)=\phi(\alpha, \gamma+1)$.
+$\phi(\alpha)$, then $\psi(\alpha,\gamma)=\phi_\alpha(\gamma+1)$.
 
@@
 
-On the other hand, $\varphi$ can be expressed in terms of $\psi$.
+On the other hand, $\phi$ can be expressed in terms of $\psi$.
 
@@
 \end{Coqsrc}
+
+\index{Projects}
+\begin{project}
+Please study a way to pretty print ordinal terms in Veblen normal form (see Section~\vref{sect:ppT1}).
+\end{project}
 
@@
 
-Prove that the implemantations of \texttt{succ}, \texttt{+}, $\varphi_0$, etc.
+Prove that the implemantations of \texttt{succ}, \texttt{+}, $\phi_0$, etc.
 are compatible in both notation systems.
@@
 Note that a function \texttt{T1\_inj} from \texttt{T1} to \texttt{T2} has already been defined. It may help to complete the task.
+
+\noindent\emph{From \href{../src/html/hydras.Gamma0.T2.html\#T1_to_T2}%
+{\texttt{Gamma0.T2}}}
+\begin{Coqsrc}
+(* injection from T1 *)
+
+Fixpoint T1_to_T2 (alpha :T1) : T2 :=
+  match alpha  with
+  | T1.zero => zero
+  | T1.ocons a n b => gcons zero (T1_to_T2 a) n (T1_to_T2 b)
+  end.
+\end{Coqsrc}
+
+\end{project}
+
+\begin{project}
+Prove that the notation system \texttt{Gamma0} is a correct implementation 
+of the segment $[0,\Gamma_0)$ of the set of countable ordinals.
 \end{project}
@@
 omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega}, \pageref{sect:omega-T1}, \pageref{sect:omega-notation2}, \dots\\
-phi0     & \texttt{phi0 alpha} & $\varphi_0(\alpha),\; \omega^\alpha$&exponential of base $\omega$ & \pageref{sect:notation-phi0}\\
+phi0     & \texttt{phi0 alpha} & $\phi_0(\alpha),\; \omega^\alpha$&exponential of base $\omega$ & \pageref{sect:notation-phi0}\\
 

### 2020-10-26 doc/hydras.tex
typos in the index

@@
 \label{types:T1}
-\index{Constants!zero:T1}
+%\index{Constants!zero:T1}
 
@@
 
-\label{types:T2}
 \emph{Please look also at the library
@@
 \label{sect:omega-T1}
-\index{Constants!zero:T1}
+%\index{Constants!zero:T1}
 
@@
 \emph{Veblen normal forms}.
+\label{types:T2}
 
@@
 
-The enumeration function of critical ordinals, presented in Sect.\~vref{sect:phi-schutte}, is recursively defined in type \texttt{T2}.
+The enumeration function of critical ordinals, presented in Sect.~\vref{sect:phi-schutte}, is recursively defined in type \texttt{T2}.
 
@@
 \begin{Coqsrc}
+Module G0.
+
 Class G0 := mkg0 {vnf : T2; vnf_ok : nfb vnf}.

### 2020-10-27 doc/hydras.tex
correction of a fake definition of maximal-alpha-critical ordinals

@@
 
-Hydra games (a.k.a. \emph{Hydra battles}) appeared in an article published in 1982 by two mathematicians:
+Hydra games (a.k.a. \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians:
 L. Kirby and J. Paris~\cite{KP82}: \emph{Accessible Independence Results for Peano Arithmetic}. 
@@
 \begin{Coqbad}
+Require Import Lia.
+
 Module Bad.
@@
 Proof.
-   unfold double; omega.
+   unfold double; lia.
 Qed.
@@
 \begin{Coqalt}
+Require Import Arith Lia Iterates.
 Module Alt.
-
-  Definition double (n : nat) := 2 * n.
-
+  Definition double (n : nat) := iterate S n n.
 End Alt.
@@
 Proof.
-  unfold Alt.double, Nat.double.
-  omega.
+  unfold Alt.double, Nat.double; induction n; cbn.
+  -  trivial.
+  -  rewrite <- iterate_rw, iterate_S_eqn, <- IHn; lia.
 Qed.
@@
 %\index{To do}
-
-\begin{todo}
+\index{Projects}
+
+\begin{project}
 Look for an existing library on trees with nodes of arbitrary arity, in order to replace  this ad-hoc type with something more generic.
-\end{todo}
+\end{project}
 
@@
 
-\begin{project}[**]
+\begin{project}
 
@@
 
-We provide several notations for ``patterns'' which occur often in our developments.
+We provide several notations for hydra patterns  which occur often in our developments.
 
@@
 \texttt{node}, \texttt{hnil} and \texttt{hcons} of the mutually inductive types \texttt{Hydra} and \texttt{Hydrae}. 
-Let us define for instance the function that  computes the number of nodes of any hydra:
+Let us define for instance the function which  computes the number of nodes of any hydra:
 
 \vspace{4pt}
+\noindent
 \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
@@
 
-One may try to use the default tactic of proof by induction, that corresponds to an application of the automatically  generated  induction principle for  type \texttt{Hydra}:
+One may try to use the default tactic of proof by induction, which corresponds to an application of the automatically  generated  induction principle for  type \texttt{Hydra}:
 
@@
 \vspace{4pt}
+\noindent
 \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
@@
 
-\index{Coq!Techniques!Mutual induction}
+\index{Coq!Techniques!Mutually inductive types}
 \index{Coq!Commands!Scheme}
@@
 \end{enumerate}
-The relation associated with each round of the battle is \emph{parameterized}  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot,
+The relation associated with each round of the battle is parameterized  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot,
 but present for consistency's sake).
@@
 
-\subsection{Chopping of a head at distance \texorpdfstring{$\geq 2$}{>= 2} from the foot (relation R2) }
+\subsection{Chopping off a head at distance \texorpdfstring{$\geq 2$}{>= 2} from the foot (relation R2) }
 
@@
 The proposition (\texttt{S1 n s s'}) holds if $s'$ is obtained by replacing some element $h$ of $s$ by 
-$n+1$ copies of $h'$, where  the proposition (\texttt{R1 h h'}) holds, in other words, \texttt{h'} is just \texttt{h}, without the chopped-off  head.
+$n+1$ copies of $h'$, where  the proposition (\texttt{R1 h h'}) holds, in other words, $h'$ is just $h$, without the chopped-off  head.
 \texttt{S1} is an inductive relation with two constructors that allow us to choose the position in $s'$ of the wounded sub-hydra $h$.
@@
 \vspace{4pt}
+\noindent
 \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#S1}{Hydra.Hydra\_Definitions}}
@@
 \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-\index{Coq!Type classes}
 \label{types:Battle}
@@
 
-\index{Coq!Type classes}
 \label{sect:hvariant-def}
@@
 
-\index{Coq!Type classes}
-
 \begin{Coqsrc}
@@
 
-\index{Coq!Type classes}
 \label{types:ON}
@@
 
-\index{Coq!Type classes}
 \label{types:SubON}
@@
 
-\index{Coq!Type classes}
 \label{types:ON-for}
@@
 
-\index{Coq!Type classes}
 \label{types:ON-iso} 
@@
 \label{sec:orgheadline77}
-\index{Coq!Techniques!Well founded relations}
+\index{Coq!Techniques!Well-founded induction}
 
@@
 
-\index{Coq!Techniques!Simultaneous induction}
 \index{Coq!Techniques!Transfinite induction}
@@
 Section V, 13 of~\cite{schutte} defines \emph{strongly critical} and
-\emph{maximal $\alpha$-critical} ordinals.
-
-We give their definition in \coq's syntax.
+\emph{maximal $\alpha$-critical} ordinals:
+
+\begin{itemize}
+\item $\alpha$ is strongly critical if
+$\alpha$ is $\alpha$-critical,
+\item $\gamma$ is maximal $\alpha$-critical if $\gamma$ is $\alpha$-critical, and, for all $\xi>\alpha$, $\gamma$ is not $\xi$-critical.
+
+\end{itemize}
 
@@
     In (Cr alpha) gamma /\
-    forall ksi, alpha < ksi -> ~ In (Cr alpha) ksi.
+    forall xi, alpha < xi -> ~ In (Cr xi) gamma.
 
@@
 
-First, we define a type that contains only terms in Veblen normal form, and redefine \texttt{lt} and \texttt{compare} by delegation.
+First, we define a type that contains only terms in Veblen normal form, and redefine \texttt{lt} and \texttt{compare} by delegation (see for comparison the construction of type \texttt{E0} in Sect.~\vref{sect:E0-def}).
 
@@
 
-Then we, prove that \texttt{lt} is a well-founded strict order and that the
+Then, we prove that \texttt{lt} is a well-founded strict order and that the
 function \texttt{compare} is correct.
@@
 Name & Description& Page \\\hline
-Battle & Type class of battle types & \pageref{types:Battle} \\
-E0 & (Well formed) ordinals below $\epsilon_0$ & \pageref{types:E0}\\
+Battle & type class of battle types & \pageref{types:Battle} \\
+E0 & (well formed) ordinals below $\epsilon_0$ & \pageref{types:E0}\\
 Hvariant & variants for proving termination of hydra battles & \pageref{sect:hvariant-def} \\

### 2020-10-28 doc/hydras.tex
typos in the pdf file

@@
 \date{\today}
-\title{Hydra battles in Coq}
+\title{Hydra Battles in Coq}
 \makeindex
@@
 
-\subsection{Difference from Kirby and Paris's work}
+\subsection{Difference from Kirby and Paris's Work}
 In~\cite{KP82}, Kirby and Paris showed  that there is no proof of termination of all hydra battles in Peano Arithmetic (PA).
@@
 
-\subsection{Trust in our proofs?}
+\subsection{Trust in our Proofs?}
 \label{sect:trust-in-proofs}
@@
 
-\subsection{Assumed redundancy}
+\subsection{Assumed Redundancy}
 
@@
 
-\subsection{Main references}
+\subsection{Main References}
 
@@
 
-\subsection{State of the development}
+\subsection{State of the Development}
 The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on  notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
@@
 
-\subsection{Typographical conventions}
+\subsection{Typographical Conventions}
 
@@
 
-\subsection{Active links}
+\subsection{Active Links}
 The  links which appear in this pdf  document lead are of three possible kinds of destination:
@@
 
-\subsection{Alternative or bad definitions}
+\subsection{Alternative or Bad Definitions}
 \label{sect:alt-proofs}
@@
 
-\subsection{The rules of the game}
+\subsection{The Rules of the Game}
 
@@
 
-\section{Hydras and their representation in \emph{Coq}}
+\section{Hydras and their Representation in \emph{Coq}}
 \label{sec:orgheadline48}
@@
 
-\subsubsection{Recursive functions on type \texttt{Hydra}}
+\subsubsection{Recursive Functions on type \texttt{Hydra}}
 \label{sec:orgheadline41}
@@
 
-\subsection{Induction principles for hydras}
+\subsection{Induction Principles for Hydras}
 \label{sec:orgheadline42}
@@
 
-\subsubsection{A failed attempt}
+\subsubsection{A failed Attempt}
 
@@
 
-\subsubsection{A principle for  mutual induction}
+\subsubsection{A Principle of Mutual Induction}
 In order to get an appropriate induction scheme for the types 
@@
 
-\subsubsection{A correct proof}
+\subsubsection{A Correct Proof}
 
@@
 
-\section{Relational description of hydra battles}
+\section{Relational Description of Hydra Battles}
 
@@
 
-\subsection{Chopping off a head at distance 1 from the foot (relation  R1)}
+\subsection{Chopping off a Head at Distance 1 from the Foot (Relation  R1)}
 
@@
 
-\subsection{Chopping off a head at distance \texorpdfstring{$\geq 2$}{>= 2} from the foot (relation R2) }
+\subsection{Chopping off a Head at Distance \texorpdfstring{$\geq 2$}{>= 2} from the Foot (relation R2) }
 
@@
 
-\subsection{Relation associated with a round}
+\subsection{Relation Associated with a Round}
 
@@
 
-\subsection{Rounds and battles}
+\subsection{Rounds and Battles}
 
@@
 
-\subsection{Classes of battles}
+\subsection{Classes of Battles}
 \label{sect:battle-classes}
@@
 
-\subsection{Big steps}
+\subsection{Big Steps}
 
@@
 
-\section{A long battle}
+\section{A Long Battle}
 \label{sect:big-battle}
@@
 
-\subsection{The beginning of hostilities}
+\subsection{The beginning of Hostilities}
 During the two first rounds, our hydra loses its two rightmost heads. Thus just before the third round, it looks like in figure~\vref{fig:hinit-plus2}.
@@
 
-\subsection{Looking for regularities}
+\subsection{Looking for Regularities}
 
@@
 
-\subsection{Giant steps}
+\subsection{Giant Steps}
 
@@
 
-\subsection{A minoration lemma}
+\subsection{A Minoration Lemma}
 
@@
 
-\section{Generic properties}
+\section{Generic Properties}
 
@@
 
-\subsection{A  small proof of impossibility}
+\subsection{A  Small Proof of Impossibility}
 \index{Maths!Proofs of impossibility}
@@
 
-\chapter{Introduction to ordinal numbers and ordinal notations}
+\chapter{Introduction to Ordinal Numbers and Ordinal Notations}
 
@@
 
-\section{The mathematical point of view}
-
-\subsection{Well-ordered sets}
+\section{The Mathematical Point of View}
+
+\subsection{Well-ordered Sets}
 Let us start with some definitions.
@@
 
-\subsection{Ordinal numbers}
+\subsection{Ordinal Numbers}
 
@@
 
-\section{Ordinal numbers in Coq}
+\section{Ordinal Numbers in Coq}
 
@@
 
-\section{Countable ordinals}
+\section{Countable Ordinals}
 
@@
 
-\section{Ordinal notations}
+\section{Ordinal Notations}
 
@@
 
-\subsubsection*{Ordered types}
+\subsubsection*{Ordered Types}
 
@@
 
-\subsection{A class for ordinal notations}
+\subsection{A Class for Ordinal Notations}
 
@@
 
-\subsection{Ordinal notations and measures for proving termination}
+\subsection{Ordinal Notations and Measures for Proving Termination}
 \label{sect:measure-ON}
@@
 
-\section{Examples of ordinal notations}
-
-\subsection{The ordinal \texorpdfstring{$\omega$}{omega}}
+\section{Examples of Ordinal Notations}
+
+\subsection{The Ordinal \texorpdfstring{$\omega$}{omega}}
 
@@
 
-\subsection{Sum of two ordinal notations}
+\subsection{Sum of two Ordinal Notations}
 
@@
 
-\subsection{The ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
+\subsection{The Ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
 
@@
 
-\subsection{Limits and successors in \texorpdfstring{$\omega+\omega$}{omega+omega}}
+\subsection{Limits and Successors in \texorpdfstring{$\omega+\omega$}{omega+omega}}
 
@@
 
-\section{The ordinal \texorpdfstring{$\omega^2$}{omega^2}}
+\section{The Ordinal \texorpdfstring{$\omega^2$}{omega^2}}
 
@@
 
-\subsection{A proof of termination using \texorpdfstring{$\omega^2$}{omega^2}} 
+\subsection{A Proof of Termination using \texorpdfstring{$\omega^2$}{omega^2}} 
 \label{sect:merge-example}.
@@
 
-\subsection{Yet another  proof of impossibility}
+\subsection{Yet Another  Proof of Impossibility}
 \label{omega2-case}
@@
 
-\subsubsection{Proof of the inequality \texttt{m small\_h o< m big\_h}}
+\subsubsection{Proof of the Inequality \texttt{m small\_h o< m big\_h}}
 
@@
 
-\subsubsection{Proof of the inequality \texttt{m big\_h o<= m small\_h} }
+\subsubsection{Proof of the Inequality \texttt{m big\_h o<= m small\_h} }
 
@@
 
-\subsubsection{End of the proof}
+\subsubsection{End of the Proof}
 
@@
 
-\section{A notation for finite ordinals}
+\section{A Notation for Finite Ordinals}
 
@@
 
-\section{Comparing two ordinal notations}
+\section{Comparing two Ordinal Notations}
 
@@
 
-\section{Comparing an ordinal notation with Schütte's model}
+\section{Comparing an Ordinal Notation with Schütte's Model}
 
@@
 
-\section{Isomorphism of ordinal notations}
+\section{Isomorphism of Ordinal Notations}
 
@@
 
-\section{Other ordinal notations}
+\section{Other Ordinal Notations}
 
@@
 
-\chapter[A proof of termination, using epsilon0]{A proof of termination, using ordinals below \texorpdfstring{$\epsilon_0$}{epsilon0}}
+\chapter[A proof of termination, using epsilon0]{A Proof of Termination, using Ordinals below \texorpdfstring{$\epsilon_0$}{Epsilon0}}
 
@@
 
-\section{The ordinal \texorpdfstring{\(\epsilon_0\)}{epsilon0}}
+\section{The Ordinal \texorpdfstring{\(\epsilon_0\)}{epsilon0}}
 \label{sec:epsilon0-intro}
 
-\subsection{Cantor normal form}
+\subsection{Cantor Normal Form}
 \index{Maths!Cantor normal form}
@@
 
-\subsection{A data type for  ordinals in Cantor normal form}
+\subsection{A Data Type for  Ordinals in Cantor Normal Form}
 \label{sec:orgheadline72}
@@
 
-\subsubsection{Finite ordinals}
+\subsubsection{Finite Ordinals}
 \label{sec:orgheadline67}
@@
 
-\subsubsection{The ordinal \(\omega\)}
+\subsubsection{The Ordinal \(\omega\)}
 \label{sec:orgheadline68}
@@
 
-\subsubsection{The ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
+\subsubsection{The Ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
 \label{sec:orgheadline70}
@@
 
-\subsubsection{The hierarchy of \(\omega\)-towers:}
+\subsubsection{The Hierarchy of \(\omega\)-towers:}
 \label{sec:orgheadline71}
@@
 
-\subsection{Comparison between ordinal terms}
+\subsection{Comparison between Ordinal Terms}
 \label{sec:orgheadline73}
@@
 
-\subsubsection{A predicate for characterizing normal forms}
+\subsubsection{A Predicate for Characterizing Normal Forms}
 \label{sect:t1-nf}
@@
 
-\subsection{Making normality implicit}
+\subsection{Making Normality Implicit}
   We would like to get rid of terms of type \texttt{T1} which are not in Cantor normal form.
@@
 
-\subsubsection{A sigma type for \texorpdfstring{$\epsilon_0$}{epsilon0}}
+\subsubsection{A Sigma Type for \texorpdfstring{$\epsilon_0$}{epsilon0}}
 
@@
 
-\subsection{Syntactic definition of limit and successor ordinals}
+\subsection{Syntactic Definition of Limit and Successor Ordinals}
 
@@
 
-\subsubsection{Addition and multiplication}
+\subsubsection{Addition and Multiplication}
 
@@
 
-\subsection{Pretty printing Ordinals in Cantor normal Form}
+\subsection{Pretty Printing Ordinals in Cantor Normal Form}
 \label{sect:ppT1}
@@
 
-The following data type defines a abstract syntax for more readable ordinals in Cantor normal form:
+The following data type defines an abstract syntax for more readable ordinals terms in Cantor normal form:
 
@@
 
-\subsubsection{Arithmetic on type \texttt{E0}}
+\subsubsection{Arithmetic on Type \texttt{E0}}
 
@@
 
-\section{Well-foundedness and transfinite induction}
+\section{Well-foundedness and Transfinite Induction}
 
@@
 
-\subsection{About  well-foundedness}
+\subsection{About  Well-foundedness}
 \label{sec:orgheadline82}
@@
 
-\subsubsection{A first attempt}
+\subsubsection{A First Attempt}
 \label{sec:orgheadline77}
@@
 
-\subsubsection{Using a stronger inductive predicate.}
+\subsubsection{Using a Stronger Inductive Predicate.}
 \label{sec:orgheadline78}
@@
 
-\subsection{An ordinal notation for  \texorpdfstring{$\epsilon_0$}{epsilon0}}
+\subsection{An Ordinal Notation for  \texorpdfstring{$\epsilon_0$}{epsilon0}}
 
@@
 
-\section{A variant for hydra battles}
-
-\subsection{Natural sum (a.k.a. Hessenberg's  sum)}
+\section{A Variant for Hydra Battles}
+
+\subsection{Natural Sum (a.k.a. Hessenberg's  Sum)}
 \label{sec:orgheadline87}
@@
 
-\subsubsection{Rewriting lemmas}
+\subsubsection{Rewriting Lemmas}
 \label{sec:orgheadline86}
@@
 \end{project}
-\subsection{More theorems on Hessenberg's sum}
+\subsection{More Theorems on Hessenberg's Sum}
 
@@
 
-\subsubsection{Commutativity, associativity}
+\subsubsection{Commutativity, Associativity}
 
@@
 
-\subsection{A measure for hydra-battle termination}
+\subsection{A Measure for Hydra-battle Termination}
 
@@
 
-\chapter[The Ketonen-Solovay machinery]{Strolling inside \texorpdfstring{$\epsilon_0$}{epsilon0}: The Ketonen-Solovay machinery\label{ks-chapter}}
+\chapter[The Ketonen-Solovay machinery]{Strolling inside \texorpdfstring{$\epsilon_0$}{Epsilon0}: The Ketonen-Solovay Machinery\label{ks-chapter}}
 \label{chap:ketonen}
@@
 
-\subsubsection{Mathematical definition of canonical sequences}
+\subsubsection{Mathematical Definition of Canonical Sequences}
 
@@
 
-\subsubsection{Canonical sequences in Coq}
+\subsubsection{Canonical Sequences in Coq}
 
@@
 
-\subsection{Basic properties of canonical sequences}
+\subsection{Basic Properties of Canonical Sequences}
 
@@
 
-\subsubsection{Canonical sequences and the order $<$}
+\subsubsection{Canonical Sequences and the Order $<$}
 
@@
 
-\subsubsection{Limit ordinals are really limits}
+\subsubsection{Limit Ordinals are Really Limits}
 The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
@@
 
-\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0} : paths}
+\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0} : Paths}
 \index{Maths!Accessibility inside epsilon0}
@@
 
-\subsection{Formal definition}
+\subsection{Formal Definition}
 
@@
 
-\subsection{Existence of a path}
+\subsection{Existence of a Path}
 
@@
 
-\subsection{Paths and hydra battles}
+\subsection{Paths and Hydra Battles}
 \label{KS-o2h}
@@
 
-\section{A  proof of impossibility}
+\section{A  Proof of Impossibility}
 
@@
 
-\subsection{The case of free battles}
+\subsection{The case of Free Battles}
 \label{sec:free-battles-case}
@@
 
-\section{The case of standard battles}
+\section{The Case of Standard Battles}
 \label{sec:standard-intro}\label{std-case}
@@
 
-\subsection{Paths with constant index}
+\subsection{Paths with a Constant Index}
 
@@
 
-\subsubsection{Sketch of proof of \texttt{Cor12}}
+\subsubsection{Sketch of Proof of \texttt{Cor12}}
 \index{Maths!Transfinite induction}
@@
 
-\subsection{Casting paths with constant index into standard paths}
+\subsection{Casting Paths with a Constant Index into a Standard Path}
 
@@
 
-\subsubsection{Sketch of proof of \texttt{constant\_to\_standard\_path}}
+\subsubsection{Sketch of Proof of \texttt{constant\_to\_standard\_path}}
 
@@
 
-\subsection{Back to hydras}
+\subsection{Back to Hydras}
 \label{sec:standard-battles-cases}
@@
 %---------------------------------------------------------------------
-\chapter{Large sets and rapidly growing functions}\label{chap:alpha-large}
+\chapter{Large Sets and Rapidly Growing Functions}\label{chap:alpha-large}
 
@@
 
-\section{The length of minimal large sequences}
+\section{The Length of Minimal Large Sequences}
 
@@
 
-\subsection{Formal specification}
+\subsection{Formal Specification}
 
@@
 
-\subsection{Abstract properties}
+\subsection{Abstract Properties}
 
@@
 
-\subsubsection{The  ordinal zero}
+\subsubsection{The  Ordinal Zero}
 \label{sect:L-spec-zero}
@@
 
-\subsubsection{Successor ordinals}
+\subsubsection{Successor Ordinals}
 \label{sect:L-spec-succ}
@@
 
-\subsubsection{Limit ordinals}
+\subsubsection{Limit Ordinals}
 \label{sect:L-spec-lim}
@@
 
-\subsection{First results}
+\subsection{First Results}
 
@@
 
-\subsubsection{Finite ordinals}
+\subsubsection{Finite Ordinals}
 
@@
 
-\subsubsection{The first limit ordinal \texorpdfstring{$\omega$}{omega}}
+\subsubsection{The First Limit Ordinal  \texorpdfstring{$\omega$}{omega}}
 
@@
 
-\subsubsection{Going further}
+\subsubsection{Going Further}
 Let us consider a last example, ``computing'' $L_{\omega^3}$.
@@
 
-\subsubsection{Back to Hydra battles}
+\subsubsection{Back to Hydra Battles}
 
@@
 
-\section{The Wainer-Hardy hierarchy (functions \texorpdfstring{$H_\alpha$}{\texttt{H\_alpha}})}
+\section{The Wainer-Hardy Hierarchy (Functions \texorpdfstring{$H_\alpha$}{\texttt{H\_alpha}})}
 
@@
 
-\subsection{Hardy functions in \texttt{Coq}}
+\subsection{Hardy Functions in \texttt{Coq}}
 
@@
 
-\subsection{First  steps of the Hardy hierarchy}
+\subsection{First  Steps of the Hardy hierarchy}
 Using rewrite rules from \texttt{H\_eq1} to \texttt{H\_eq4}, we can explore the functions $H_\alpha$ for some small values of $\alpha$.
 
-\subsubsection{Finite ordinals}
+\subsubsection{Finite Ordinals}
 
@@
 
-\subsubsection{New limits}
+\subsubsection{New Limits}
 
@@
 
-\subsubsection{A numerical example}
+\subsubsection{A Numerical Example}
 
@@
 
-\subsection{Abstract properties of H-functions}
+\subsection{Abstract Properties of H-functions}
 ~\label{sect:H-alpha-prop}
@@
 
-\subsubsection{Back to hydras}
+\subsubsection{Back to Hydras}
 
@@
 
-\section{The Wainer hierarchy (functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
+\section{The Wainer Hierarchy (Functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
 \label{sect:wainer}
@@
 %----------------------------------------------------------------------
-\chapter[Countable Ordinals (after Sch\"{u}tte)]{Kurt Schütte's axiomatic Definition of countable Ordinals}
+\chapter[Countable Ordinals (after Sch\"{u}tte)]{Kurt Schütte's Axiomatic Definition of Countable Ordinals}
 
@@
 
-\section{Additional  axioms}
+\section{Additional  Axioms}
 
@@
 
-\subsubsection{Classical logic}
+\subsubsection{Classical Logic}
 
@@
 
-\subsubsection{Description operators}
+\subsubsection{Description Operators}
 
@@
 
-\subsubsection{The definition of zero}
+\subsubsection{The Definition of zero}
 
@@
 
-\section{The  successor function}
+\section{The  Successor Function}
 
@@
 
-\section{Finite ordinals}
+\section{Finite Ordinals}
 
@@
 
-\section{The definition of \texttt{omega}}
+\section{The Definition of \texttt{omega}}
 In order to define $\omega$, the first infinite ordinal, we use an operator which
@@
 
-\subsection{Ordering functions and ordinal addition}
+\subsection{Ordering Functions and Ordinal Addition}
 
@@
 
-\subsection{Ordinal addition}
+\subsection{Ordinal Addition}
 
@@
 
-\subsubsection{Multiplication by a natural number}
+\subsubsection{Multiplication by a Natural Number}
 
@@
 
-\section{The exponential of basis \texorpdfstring{$\omega$}{omega}}
+\section{The Exponential of Basis \texorpdfstring{$\omega$}{omega}}
 
@@
 
-\subsection{Additive principal ordinals}
+\subsection{Additive Principal Ordinals}
 
@@
 
-\subsection{The function \texttt{phi0}}
+\subsection{The Function \texttt{phi0}}
 
@@
 
-\subsection{Omega-towers and the ordinal \texorpdfstring{$\epsilon_0$}{epsilon0}}
+\subsection{Omega-towers and the Ordinal \texorpdfstring{$\epsilon_0$}{epsilon0}}
 
@@
 
-\subsection{Properties of the set  \texttt{AP}}
+\subsection{Properties of the Set  \texttt{AP}}
 
@@
 
-\subsubsection{Properties of the function \texorpdfstring{$\phi_0$}{phi0}}
+\subsubsection{Properties of the Function \texorpdfstring{$\phi_0$}{phi0}}
 
@@
 
-\section{Critical ordinals}
+\section{Critical Ordinals}
 
@@
 
-\subsection{A flavor of infinity}
+\subsection{A flavor of Infinity}
 
@@
 
-\section{Cantor normal form}
+\section{Cantor Normal Form}
 
@@
 
-\section{An embedding of \texttt{T1} into \texttt{Ord}}
+\section{An Embedding of \texttt{T1} into \texttt{Ord}}
 
@@
 
-\section{Related work}
+\section{Related Work}
 
@@
 
-\chapter{The ordinal \texorpdfstring{$\Gamma_0$}{Gamma0} (first draft)}
+\chapter{The Ordinal \texorpdfstring{$\Gamma_0$}{Gamma0} (first draft)}
 
@@
 \begin{project}
-Prove that a (countable)  ordinal $\alpha$ is strongly critical iff $\phi \alpha 0=\alpha$ (Theorem 13.13 of~\cite{schutte} ). 
+Prove that a (countable)  ordinal $\alpha$ is strongly critical iff 
+$\phi_\alpha(0)=\alpha$ (Theorem 13.13 of~\cite{schutte} ). 
 \end{project}
@@
 
-In the present version of this development, we will only study $\Gamma_0$ as a notation system, much more powerful than the ordinal notation for $\epsilon_0$.
+In the present version of this development, we  only study $\Gamma_0$ as a notation system, much more powerful than the ordinal notation for $\epsilon_0$.
 
@@
 
-\section{The type \texttt{T2} of ordinal terms}
+\section{The Type \texttt{T2} of Ordinal Terms}
 
@@
 
-\section{How big is \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}?}
+\section{How Big is \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}?}
 
@@
 
-\section{Veblen normal forms}
+\section{Veblen Normal Forms}
 \begin{definition}
@@
 
-\subsection{Length of a term}
+\subsection{Length of a Term}
 
@@
 
-\section{Main functions on \texttt{T2}}
+\section{Main Functions on \texttt{T2}}
 
@@
 
-Like for Cantor normal forms (see Sect.\~ref{sect:infix-plus-T1})  the definition of addition in \texttt{T2}  requires comparison between ordinal terms.
+Like for Cantor normal forms (see Sect.~\ref{sect:infix-plus-T1}),  the definition of addition in \texttt{T2}  requires comparison between ordinal terms.
 
@@
 
-\subsection{The Veblen function \texorpdfstring{$\phi$}{\texttt{phi}}}
+\subsection{The Veblen Function \texorpdfstring{$\phi$}{\texttt{phi}}}
 
@@
 
-\section{An ordinal notation for \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}}
+\section{An Ordinal Notation for \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}}
 
@@
 
-\section{Future work (projects)}
+\section{Future Work (projects)}
 \index{Projects}
@@
 
-\section{How to install the libraries}
+\section{How to Install the Libraries}
 \label{sec:orgheadline4}
@@
 
-\section{Comments on exercises and projects}
+\section{Comments on Exercises and Projects}
 

### 2020-10-29 doc/hydras.tex
corrections to the documentation (pdf)

@@
 \item External links, mainly to \coq's page,
-\item Local links to pages generated by \texttt{coqdoc}. According to the current makefile,
+\item Local links to pages generated by \texttt{coqdoc}. According to the current makefile (through the commands \texttt{make html} and \texttt{make pdf}), 
   we assume that the page generated from a library \texttt{XXX/YYY.v} is stored as
-the relative address \texttt{../src/html/hydras.XXX.YYY.html} (from the location of the pdf).
+the relative address \texttt{../src/html/hydras.XXX.YYY.html} (from the location of the pdf)
 Thus,  active links towards our \coq{} modules may be incorrect if you got this \texttt{pdf} document otherwise than by compiling the distribution available in
-\url{https://framagit.org/casteran/hydra-battles-in-coq}.
+\url{https://github.com/coq-community/hydra-battles}.
 
@@
 \node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\node (H4) at (6,0) {$\Xey[2][lightgray]$};
+%\node (H4) at (6,0) {$\Xey[2][lightgray]$};
 \draw (foot) -- (N1)[very thick] ;
@@
 \node (H1) at (0,8) {$\Sey[2][lightred]$};
-\node (H2) at (5,0) {$\Xey[2][lightgray]$};
+%\node (H2) at (5,0) {$\Xey[2][lightgray]$};
 \node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
@@
 \node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,0) {$\Xey[2][lightgray]$};
+%\node (H5) at (4,0) {$\Xey[2][lightgray]$};
 \draw (foot) -- (N1)[very thick,draw=lightred] ;
@@
 
-So, the next chapters will present ordinal numbers and ordinal notations, and we will
-look for a variant complex enough for proving the termination of all battles, but not too complex if possible.
+The next chapter is dedicated to a generic implementation of ordinal notations, and chapter~\ref{chap:T1} to a proof of termination of all hydra battles with the help of an ordinal notation for the interval $[0,\epsilon_0)$.
 
@@
 \label{cnf-math-def}
-\label{chap:proof-term}
 \label{chap:T1}

### 2020-10-30 doc/hydras.tex
correction of typos

@@
 
-This chapter is dedicated to the representation of hydras and rules of the hydra game in \coq's specification language:\gallina.
+This chapter is dedicated to the representation of hydras and rules of the hydra game in \coq's specification language:~\gallina.
 
@@
 The battle we are studying seems to be awfully long. Let us concentrate our
-tests on some particular events : the states where $\texttt{nh}=0$
-
-From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} will be
-equal to zero.
+tests on some particular events : the states where $\texttt{nh}=0$.
+From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} is equal to zero.
 
@@
 
-Thus, $ (1 + 11)$ rounds later, the \texttt{n1} field will be equal to $2$, and 
-\texttt{nh}  will equal to $0$.
+Thus, $ (1 + 11)$ rounds later, the \texttt{n1} field is equal to $2$, and 
+\texttt{nh}   to $0$.
 
@@
 \end{Coqsrc}
-Using some facts (proven in \url{../src/html/BigBattle.html}),we get several  minorations.
+Using some facts (proven in 
+\href{../src/html/hydras.Hydra.BigBattle.html}{hydras.Hydra.BigBattle}),we get several  minorations.
 
@@
 
-It is easy to represent the finite ordinals, successors and limits.
+It is easy to represent  finite ordinals, successors and limits.
 
@@
 
-Definition mult_fin  (alpha : t) (p : nat): t :=
+Definition mult_alpha_n  (alpha : t) (p : nat): t :=
   match alpha, p with
@@
                end.
-Infix "*" := mult_fin : o2_scope.
+Infix "*" := mult_alpha_n : o2_scope.
 
@@
 
-Definition fin_mult (n:nat)(beta : t) : t :=
-  match n, beta with
+Definition mult_n_alpha (n:nat)(alpha : t) : t :=
+  match n, alpha with
  |  0, _  => zero
@@
 
-Example e1 : fin_mult 3 (omega * 7 + 15) = omega * 7 + 45.
+Example e1 : (omega * 7 + 15) * 3 = omega * 21 + 15.
 Proof. reflexivity. Qed.
 
-Example e2 :  omega * 2 + 2 + omega * 3 + 10 = omega * 5 + 10.
+Example e2 :  mult_n_alpha 3 (omega * 7 + 15) = omega * 7 + 45.
 Proof. reflexivity. Qed.
@@
 
-Note that the type (\texttt{t 0}) is empty, and that \texttt{Program} generates an obligation
-for verifying that the constraint $i<n$ is fullfilled when one tries to build the $i$-th ordinal inside the type (\texttt{t $n$}).
+Note that the type (\texttt{t 0}) is empty, and that, for any natural number
+ $n$, $n$ does not belong to (\texttt{t $n$}).
 
@@
 \begin{project}
+Consider two isomorphic ordinal notations \texttt{OA} and \texttt{OB}.
+Prove that, if \texttt{OA} [resp. \texttt{OB}] is a correct implementation 
+of $\alpha$ [resp. $\beta$], then $\alpha=\beta$.
+\end{project}
+
+\index{Projects}
+\begin{project}
 \label{project:succ-limit-dec}
@@
 
-And we have now our transfinite recursor:
 \index{Maths!Transfinite induction}
@@
 \label{sec:orgheadline84}
-\index{Functions!oplus @ oplus (Hessenberg commutative sum)}
+%\index{Functions!oplus @ oplus (Hessenberg commutative sum)}
 
@@
          forall beta i n, beta  t1< alpha  ->
-                (i < n)%nat ->
+                i < n ->
                  const_pathS i alpha beta ->
@@
 \begin{Coqsrc}
-Corollary Cor12' (alpha : T1) :  nf alpha ->
-         forall beta i n, (beta t1< alpha)%t1 ->
-               (i <= n)%nat ->
+Corollary Cor12_1 (alpha : T1) :
+  nf alpha ->
+  forall beta i n, beta t1< alpha ->
+                   i <= n ->
                    const_pathS i alpha beta ->
@@
 
+\vspace{4pt}
  \noindent
@@
 Lemma AP_plus_closed (alpha beta gamma : Ord): 
-     In AP alpha -> beta < alpha -> gamma < alpha -> beta + gamma < alpha.
+     In AP alpha -> beta < alpha -> gamma < alpha ->
+     beta + gamma < alpha.
 
@@
 
-The ordering function $\phi_0$ of the set \texttt{AP} is defined on the full set $\mathbb{O}$ and is continuous (Schütte says that this function is \emph{normal}).
+The ordering function $\phi_0$ of the set \texttt{AP} is defined on the full set $\mathbb{O}$ and is continuous (Schütte calls such a  function  \emph{normal}).
 

### 2020-11-02 doc/hydras.tex
Minor corrections

@@
 
-Definition mult_alpha_n  (alpha : t) (p : nat): t :=
+Definition mult_fin_r  (alpha : t) (p : nat): t :=
   match alpha, p with
@@
                end.
-Infix "*" := mult_alpha_n : o2_scope.
+Infix "*" := mult_fin_r : o2_scope.
 
@@
 
-Definition mult_n_alpha (n:nat)(alpha : t) : t :=
+Definition mult_fin_l (n:nat)(alpha : t) : t :=
   match n, alpha with
@@
 
-Example e2 :  mult_n_alpha 3 (omega * 7 + 15) = omega * 7 + 45.
+Example e2 :  mult_fin_l 3 (omega * 7 + 15) = omega * 7 + 45.
 Proof. reflexivity. Qed.
@@
 
+\vspace{4pt}
+\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
+
 \begin{Coqsrc}
@@
   Remark Ltb_ij : Nat.ltb i j.
+
   Program Definition iota_ij  (alpha: t i) : t j :=  alpha.
@@
   End Inclusion_ij.
-
 \end{Coqsrc}
@@
 
-Note that the elements of  \texttt{ON\_mult (FinOrd 3) Omega} are differently ordered (without limit ordinals):
+Note that the elements of  (\texttt{ON\_mult (FinOrd 3) Omega}) are differently ordered (without limit ordinals):
 \[(0,0),(1,0),(2,0),(0,1),(1,1),(2,1),(0,2),(1,2),(2,2),(0,3),\dots\]
-(no limit ordinals).
 
@@
 % \begin{itemize}
-% \item First, we chose to write a version for sort \texttt{Type}, since versions
+% \item First, we choose to write a version for sort \texttt{Type}, since versions
 % for sorts \texttt{Prop} and \texttt{Set} can be easily derived from
@@
 In the present chapter, we  compare our implementation of the segment $[0,\epsilon_0)$ with a mathematical text in order to ``validate'' our constructions.
-We chose as reference the axiomatic definition of the set of countable ordinals,
+Our reference here is the axiomatic definition of the set of countable ordinals,
 in chapter V of Kurt Schütte's book `` Proof Theory ''~\cite{schutte}.
@@
 \begin{Coqsrc}
-Fixpoint mult_Sn (alpha:Ord)(n:nat){struct n} :Ord :=
+Fixpoint mult_Sn (alpha:Ord)(n:nat){struct n}: Ord :=
  match n with 
@@
 
-Definition mult_n alpha n :=
+Definition mult_fin_r alpha n :=
   match n with
@@
 
-Notation "alpha * n" := (mult_n alpha n) : schutte_scope.
+Notation "alpha * n" := (mult_fin_r alpha n) : schutte_scope.
 \end{Coqsrc}
@@
 
-Theorem inject_mult_n (alpha : T1)  :
+Theorem inject_mult_fin_r (alpha : T1)  :
   nf alpha -> forall n:nat , inject (alpha *  n)%t1 =  inject alpha * n.
@@
 
-\subsubsection*{Proof of $\psi(\omega,8]\times 13+56 < \psi(\omega,8]\times 13+57 $}
+\subsubsection*{Proof of $\psi(\omega,8)\times 13+56 < \psi(\omega,8)\times 13+57 $}
 
@@
  \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
-\texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & ordinal addition & \pageref{sect:infix-plus-T1} ,\dots\\
+\texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & ordinal addition & \pageref{sect:infix-plus-T1}, \dots\\
 \texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\

### 2020-11-02 doc/hydras.tex
removing inefficient eauto with T1

@@
 
-\index{Projects}
-\begin{project}
+\index{Exercises}
+\begin{exercise}
 Prove that \texttt{Omega\_plus\_Omega} cannot be a sub-notation of \texttt{Omega}.
-\end{project}
+\end{exercise}
 
@@
  \emph{This exercise is a continuation of Project~\vref{exo:ON-mult}.}
-Define an ordinal notation \texttt{Omega2} for $\omega^2=\omega\times\omega$ (using
-\texttt{ON\_mult}).
+Use \texttt{ON\_mult} to define an ordinal notation \texttt{Omega2} for $\omega^2=\omega\times\omega$.
 

### 2020-11-04 doc/hydras.tex
Proposition to rename the directory src/ into the more canonical theories/ (#7)

@@
 The following example is extracted from
-Library~\href{../src/html/hydras.Epsilon0.F_alpha.html}{hydras.Epsilon0.F\_alpha}, where we use the \texttt{coq-equations} plug-in (see Sect.~\vref{sect:wainer}).
+Library~\href{../theories/html/hydras.Epsilon0.F_alpha.html}{hydras.Epsilon0.F\_alpha}, where we use the \texttt{coq-equations} plug-in (see Sect.~\vref{sect:wainer}).
 
@@
   we assume that the page generated from a library \texttt{XXX/YYY.v} is stored as
-the relative address \texttt{../src/html/hydras.XXX.YYY.html} (from the location of the pdf)
+the relative address \texttt{../theories/html/hydras.XXX.YYY.html} (from the location of the pdf)
 Thus,  active links towards our \coq{} modules may be incorrect if you got this \texttt{pdf} document otherwise than by compiling the distribution available in
@@
 Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the 
-module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}).
+module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}).
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Hydra}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hydra}{Hydra.Hydra\_Definitions}}
 
@@
 
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#head}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#head}{Hydra.Hydra\_Definitions}}
 
@@
 \vspace{4mm}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html\#Hy}{Hydra.Hydra\_Examples}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html\#Hy}{Hydra.Hydra\_Examples}}
 
@@
 
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#hcons_mult}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#hcons_mult}{Hydra.Hydra\_Definitions}}
 
@@
 
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
 
@@
 \noindent
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
 
@@
 \noindent
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
 
@@
 
-\vspace{4pt}\emph{From Module\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
+\vspace{4pt}\emph{From Module\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
 
@@
 \noindent
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#S1}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#S1}{Hydra.Hydra\_Definitions}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions}}
 
@@
 
-\vspace{4pt}\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}.
+\vspace{4pt}\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}.
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#round_n}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#round_n}{Hydra.Hydra\_Definitions}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
 \label{types:Battle}
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#free}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#free}{Hydra.Hydra\_Definitions}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#standard}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#standard}{Hydra.Hydra\_Definitions}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}
 
@@
 % \vspace{4pt}
-% \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
+% \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
 
@@
 
-The complete proof is in file \url{../src/html/hydras.Hydra.BigBattle.html}.
+The complete proof is in file \url{../theories/html/hydras.Hydra.BigBattle.html}.
 
@@
 The functional \texttt{iterate}, similar to Standard Library's \texttt{Nat.iter},
-is defined and studied in~\href{../src/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
+is defined and studied in~\href{../theories/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
 
@@
 Using some facts (proven in 
-\href{../src/html/hydras.Hydra.BigBattle.html}{hydras.Hydra.BigBattle}),we get several  minorations.
+\href{../theories/html/hydras.Hydra.BigBattle.html}{hydras.Hydra.BigBattle}),we get several  minorations.
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
 
@@
 The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
-\url{../src/html/hydras.Hydra.Hydra_Theorems.html} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
+\url{../theories/html/hydras.Hydra.Hydra_Theorems.html} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
 
@@
 
-Let us present the main steps of that proof, the script of which  is in the module ~\href{../src/html/hydras.Hydra.Omega_Small.html}{Hydra/Omega\_Small.v} \footnote{ The name of this file means ``the ordinal $\omega$ is too small for proving the termination of [free] hydra battles ''.
+Let us present the main steps of that proof, the script of which  is in the module ~\href{../theories/html/hydras.Hydra.Omega_Small.html}{Hydra/Omega\_Small.v} \footnote{ The name of this file means ``the ordinal $\omega$ is too small for proving the termination of [free] hydra battles ''.
 In effect, the elements of $\omega$, considered as a set, are just the natural numbers (see next chapter for more details)}.
@@
 \vspace{4pt}
-\noindent\emph{From Library~\href{../src/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}}
+\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}}
 
@@
 \vspace{4pt}
-\noindent\emph{From Library~\href{../src/html/hydras.OrdinalNotations/ON_Omega.html}{OrdinalNotations.ON\_Omega}}
+\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations/ON_Omega.html}{OrdinalNotations.ON\_Omega}}
 
@@
 \vspace{4pt}
-\noindent\emph{From Library~\href{../src/html/hydras.OrdinalNotations/ON_plus.html}{OrdinalNotations.ON\_plus}}
+\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations/ON_plus.html}{OrdinalNotations.ON\_plus}}
 
@@
 \vspace{4pt}
-\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
+\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
 
@@
 
-The following definitions are in Library \href{../src/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
+The following definitions are in Library \href{../theories/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
 
@@
 
-\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
+\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
 
@@
 
-\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}}
+\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}}
 \begin{Coqsrc}
@@
 \vspace{4pt}
-\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Omega2.html}{ON\_Omega2}}
+\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{ON\_Omega2}}
 
@@
 
-The detailed  proof script is in the file \url{../src/html/hydras.Hydra.Omega2_Small.html}.
+The detailed  proof script is in the file \url{../theories/html/hydras.Hydra.Omega2_Small.html}.
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
 
@@
 \vspace{4pt}
-\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
+\emph{From Module ~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}
 
@@
 \vspace{4pt}
-\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
+\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
 
@@
 In our formalization, the interpretation of an ordinal as a set is realized by the following definition
-(in ~\href{../src/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}).
+(in ~\href{../theories/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}).
 
@@
 
-This exercise is partially solved for $i=3$ (in ~\href{../src/html/hydras.OrdinalNotations.Example_3PlusOmega.html}{OrdinalNotations.Example\_3PlusOmega}).
+This exercise is partially solved for $i=3$ (in ~\href{../theories/html/hydras.OrdinalNotations.Example_3PlusOmega.html}{OrdinalNotations.Example\_3PlusOmega}).
 
@@
 \begin{project}
-The directory \texttt{src/OmegaOmega} contains an ad-hoc formalization of $\omega^\omega$, contributed by Pascal Manoury.
+The directory \texttt{theories/OmegaOmega} contains an ad-hoc formalization of $\omega^\omega$, contributed by Pascal Manoury.
 Every ordinal $\alpha$ is represented by a list $l$ whose elements are the coefficients of $\omega$ in  the Cantor normal form of $\alpha$ (in reverse order).
@@
 in Veblen normal form (see 
-\href{../src/html/hydras.Gamma0.Gamma0.html}{Gamma0.Gamma0}) are shown to be ordinal notation systems, but there is a lot of work to be done in order to unify ad-hoc  definitions and proofs which were written before the definition of the \texttt{ON} type class.
+\href{../theories/html/hydras.Gamma0.Gamma0.html}{Gamma0.Gamma0}) are shown to be ordinal notation systems, but there is a lot of work to be done in order to unify ad-hoc  definitions and proofs which were written before the definition of the \texttt{ON} type class.
 \end{remark}
@@
 \vspace{4pt}
-\noindent\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#T1}{Epsilon0.T1}}
+\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#T1}{Epsilon0.T1}}
 
@@
 \emph{Please look also at the library
-\href{../src/html/hydras.Gamma0.T2.html}{Gamma0.T2.html}}.
+\href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2.html}}.
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#compare}{Epsilon0.T1}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare}{Epsilon0.T1}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#compare_refl}{Epsilon0.T1}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare_refl}{Epsilon0.T1}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#lt_irrefl}{Epsilon0.T1}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#lt_irrefl}{Epsilon0.T1}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#nf_b}{Epsilon0.T1}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#nf_b}{Epsilon0.T1}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Prelude.Restriction.html}{Ordinals.Prelude.Restriction}}
+\emph{From Module~\href{../theories/html/hydras.Prelude.Restriction.html}{Ordinals.Prelude.Restriction}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#LT}{Epsilon0.T1}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#LT}{Epsilon0.T1}}
 
@@
 
-\emph{From Module~\href{../src/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
 
@@
  \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}
 
@@
 and natural induction on \texttt{n}, that (\texttt{ocons alpha n beta}) is accessible.
-The proof, quite long, can be consulted in \url{../src/html/hydras.Epsilon0.T1.html}
+The proof, quite long, can be consulted in \url{../theories/html/hydras.Epsilon0.T1.html}
 
@@
 \label{remark:a3pat}
-The alternate proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the library \href{../src/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
+The alternate proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
  \end{remark}
@@
 
-\emph{From Module~\href{../src/html/hydras.Schutte.Schutte.Correctness_E0.html}{Schutte.Schutte.Correctness\_E0}}
+\emph{From Module~\href{../theories/html/hydras.Schutte.Schutte.Correctness_E0.html}{Schutte.Schutte.Correctness\_E0}}
 
@@
 \textbf{Hint}. You may use the following definition (in 
-\href{../src/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}).
+\href{../theories/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}).
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Termination.html\#m}{Hydra.Hydra\_Termination}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Termination.html\#m}{Hydra.Hydra\_Termination}}
 
@@
 then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1})$ in one round with
-the replication factor $i$ (Lemma \href{../src/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
+the replication factor $i$ (Lemma \href{../theories/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
 \item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) battle from $\iota(\alpha)$ to $\iota(\beta)$.
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Epsilon0.Canon.html\#canonS}{Epsilon0.Canon}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS}{Epsilon0.Canon}}
 
@@
 abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
-~\href{../src/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}
+~\href{../theories/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
 
@@
 
-\emph{From Module~\href{../src/html/hydras.Hydra.O2H.html}{Hydra.O2H}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}}
 
@@
 
-The following lemma (proved in ~\href{../src/html/hydras.Hydra.O2H.html}{Hydra.O2H.v}) maps  canonical sequences to rounds of hydra battles.
+The following lemma (proved in ~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H.v}) maps  canonical sequences to rounds of hydra battles.
 
@@
 
-In the module \href{../src/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
+In the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
 This part of the development is parameterized by some class $B$ of battles, which will be instantiated later to \texttt{free} or \texttt{standard}.
@@
 Let us now consider that $B$ is instantiated to \texttt{free} (which means that we are considering proofs of termination of \emph{all} battles).
-The following lemmas are proved in Module~\href{../src/html/hydras.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
+The following lemmas are proved in Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
 The case $B=\texttt{standard}$ is studied in section~\vref{std-case}.
@@
 One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
-Module~\href{../src/html/hydras.Hydra.O2H.html}{Hydra.O2H}.
+Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}.
 
@@
 It is a straightforward application of \texttt{LT\_path\_to} of
-Module~\href{../src/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}:
+Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}:
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Epsilon0.KS.html}{Epsilon0.KS}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.KS.html}{Epsilon0.KS}}
 
@@
 be consulted in the module 
-\url{../src/html/hydras.Hydra.Epsilon0_Needed_Std.html}.
+\url{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}.
 Please note that it has the same global structure as in section\ref{sec:free-battles-case} 
@@
 Applying the  lemmas  \texttt{Lemma2\_6\_1} of the module 
-\href{../src/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
+\href{../theories/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
 {\texttt{Lemma2\_6\_1}}   and 
-\href{../src/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
+\href{../theories/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
 {\texttt{constant\_to\_standard\_path}},
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Hydra.Epsilon0_Needed_Std.html\#LT_to_standard_battle}{Hydra.Epsilon0\_Needed\_Std}}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html\#LT_to_standard_battle}{Hydra.Epsilon0\_Needed\_Std}}
 
@@
 from 
-the module \href{../src/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}.
+the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}.
 
@@
 Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
-(please see also \href{../src/html/hydras.Schutte.Correctness_E0.html}{the module \texttt{Ordinals.Schutte.Correctness\_E0}}).
+(please see also \href{../theories/html/hydras.Schutte.Correctness_E0.html}{the module \texttt{Ordinals.Schutte.Correctness\_E0}}).
 
@@
 \noindent
-\emph{From Module~ \href{../src/html/hydras.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}
+\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}
 
@@
 \noindent
-\emph{From Module~\href{../src/html/hydras.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}
 
@@
  \noindent
-\emph{From Module~\href{../src/html/hydras.Epsilon0.Large_Sets_Examples.html}{Epsilon0.Large\_Sets\_Examples}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html}{Epsilon0.Large\_Sets\_Examples}}
 \begin{Coqsrc}
@@
 \noindent
-\emph{From Module~\href{../src/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
 \begin{Coqsrc}
@@
 In order to ensure this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <<= f}, or \texttt{(fun\_le S f)} (defined
-in~\href{../src/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}).
+in~\href{../theories/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}).
 
@@
 
-\emph{From Module~ \href{../src/html/hydras.Epsilon0.Large_Sets.html\#L_omega_mult}{Epsilon0.Large\_Sets}}
+\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#L_omega_mult}{Epsilon0.Large\_Sets}}
 
@@
 \noindent
-\emph{From Module~\href{../src/html/hydras.Epsilon0.L_alpha.html\#L_}{L\_alpha}).}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#L_}{L\_alpha}).}
 
@@
 Theorem \texttt{battle\_length\_std } of
-Module~\href{../src/html/hydras.Hydra.Hydra_Theorems.html\#battle_length_std}{Hydra.Hydra\_Theorems} relates the length of standard battles with the functions $L_\alpha$.
+Module~\href{../theories/html/hydras.Hydra.Hydra_Theorems.html\#battle_length_std}{Hydra.Hydra\_Theorems} relates the length of standard battles with the functions $L_\alpha$.
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/hydras.Epsilon0.H_alpha.html\#H_}{Epsilon0.H\_alpha}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.H_alpha.html\#H_}{Epsilon0.H\_alpha}}
 
@@
 
-\emph{From Module~\href{../src/html/hydras.Epsilon0.L_alpha.html\#H_L_}{Epsilon0.L\_alpha}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#H_L_}{Epsilon0.L\_alpha}}
 
@@
 
-\emph{From Module~\href{../src/html/hydras.Epsilon0.L_alpha.html}{Epsilon0.L\_alpha}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0.L\_alpha}}
 
@@
 A first attempt is to write a definition of $F_\alpha$ by equations, in the same as for $H\_alpha$.  We use the functional \texttt{iterate} defined in 
-Module~\href{../src/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
+Module~\href{../theories/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
 
@@
 The following code comes from 
- \url{../src/html/hydras.Epsilon0.F_alpha.html}.
+ \url{../theories/html/hydras.Epsilon0.F_alpha.html}.
 
@@
 % \noindent
-% \emph{From Module~\href{../src/html/hydras.Epsilon0.Large_Sets_Demo.html}{Ordinals.Epsilon0.Large\_Sets\_Demo}}
+% \emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets_Demo.html}{Ordinals.Epsilon0.Large\_Sets\_Demo}}
 
@@
  Unlike the rest of our libraries, our library
-\href{../src/html/hydras.Schutte.Schutte.html}{Ordinals.Schutte}
+\href{../theories/html/hydras.Schutte.Schutte.html}{Ordinals.Schutte}
 is not constructive, and relies on several axioms.
@@
 Both operators allow us to write definitions close to the natural mathematical language, such as ``$\textrm{succ}$ is \emph{the} least ordinal strictly greater than $\alpha$''
-\item Please note that only the library \href{../src/html/hydras.Schutte.Schutte.html}{Schutte/*.v} is ``contaminated'' by axioms, and that the rest of our libraries remain constructive.
+\item Please note that only the library \href{../theories/html/hydras.Schutte.Schutte.html}{Schutte/*.v} is ``contaminated'' by axioms, and that the rest of our libraries remain constructive.
 \end{itemize}
@@
 A set $A$ is countable if there is an injective function from $A$ to $\mathbb{N}$ (see
-Library \href{../src/html/hydras.Schutte.Countable.html}%
+Library \href{../theories/html/hydras.Schutte.Countable.html}%
 {\texttt{Schutte.Countable}}).
@@
 
-\emph{From Module\href{../src/html/hydras.Schutte.Schutte_basics.html}%
+\emph{From Module\href{../theories/html/hydras.Schutte.Schutte_basics.html}%
 {\texttt{Schutte.Schutte\_basics}}}
@@
 \texttt{ordinal} (The  class \texttt{WO} is defined in
-Module~\href{../src/html/hydras.Schutte.Well_Orders.html}%
+Module~\href{../theories/html/hydras.Schutte.Well_Orders.html}%
 {\texttt{Well\_Orders.v}}).
@@
 Indeed, the operators \texttt{epsilon} and \texttt{iota} allowed us to make our definitions 
-quite close to Schütte's text. Our libraries \href{../src/html/hydras.Schutte.MoreEpsilonIota.html}%
+quite close to Schütte's text. Our libraries \href{../theories/html/hydras.Schutte.MoreEpsilonIota.html}%
 {\texttt{Schutte.MoreEpsilonIota}}
 and
-\href{../src/html/hydras.Schutte.PartialFun.html}%
+\href{../theories/html/hydras.Schutte.PartialFun.html}%
 {\texttt{Schutte.PartialFun}} are extensions of \texttt{Coq.logic.Epsilon} for making easier 
@@
 
-From Module \href{../src/html/hydras.Schutte.Schutte_basics.html}%
+From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
 {\texttt{~Schutte.Schutte\_basics}}
@@
 (proved in Module 
-\href{../src/html/hydras.Schutte.Schutte_basics.html\#finite_lt_omega}{\texttt{Schutte.Schutte\_basics}})
+\href{../theories/html/hydras.Schutte.Schutte_basics.html\#finite_lt_omega}{\texttt{Schutte.Schutte\_basics}})
 
@@
 Complete definitions and proofs can be found in Module
- \href{../src/html/hydras.Schutte.Ordering_Functions.html}%
+ \href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
 {\texttt{Schutte.Ordering\_Functions}} ).
@@
 One of the main theorems of
-\href{../src/html/hydras.Schutte.Ordering_Functions.html\#ordering_function_ex}%
+\href{../theories/html/hydras.Schutte.Ordering_Functions.html\#ordering_function_ex}%
 {\texttt{Ordering\_Functions}} 
@@
 Let us quote the following theorems (see Library
-\href{../src/html/hydras.Schutte.Ordering_Functions.html}%
+\href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
 {\texttt{Schutte.Ordering\_Functions}} for more details).
@@
 The following definitions and proofs can be consulted in Module
-\href{../src/html/hydras.Schutte.Addition.html}%
+\href{../theories/html/hydras.Schutte.Addition.html}%
 {\texttt{Schutte.Addition.v}}.
@@
 computational proofs of the corresponding statements in Module
-\href{../src/html/hydras.Epsilon0.T1.html}%
+\href{../theories/html/hydras.Epsilon0.T1.html}%
 {\texttt{Epsilon0.T1}}. 
@@
 
-\noindent\emph{From Module \href{../src/html/hydras.Schutte.AP.html}%
+\noindent\emph{From Module \href{../theories/html/hydras.Schutte.AP.html}%
 {\texttt{Schutte.AP}}}
@@
 We prove the inequality  $\alpha<\omega^\alpha$ whenever $\alpha < \epsilon_0$.
-\emph{Note that this condition was implicit in Module \href{../src/html/hydras.Epsilon0/T1.html\#lt_phi0}%
+\emph{Note that this condition was implicit in Module \href{../theories/html/hydras.Epsilon0/T1.html\#lt_phi0}%
 {\texttt{Epsilon0.T1}}.}
@@
 This definition is translated in \coq{} in 
-Module \href{../src/html/hydras.Schutte.Critical.html}%
+Module \href{../theories/html/hydras.Schutte.Critical.html}%
 {\texttt{Schutte.Critical}}, as the least fixpoint of a functional.
@@
 
-\noindent\emph{From \href{../src/html/hydras.Schutte.CNF.html\#cnf_t}%
+\noindent\emph{From \href{../theories/html/hydras.Schutte.CNF.html\#cnf_t}%
 {\texttt{Schutte.CNF}}}
@@
 \begin{remark}
-The sub-directory \href{../src/html/hydras.Gamma0.html}%
-{\texttt{src/Gamma0}} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form. 
+The sub-directory \href{../theories/html/hydras.Gamma0.html}%
+{\texttt{theories/Gamma0}} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form. 
 \end{remark}
@@
 Our library 
-\href{../src/html/hydras.Schutte.correctness_E0.html}%
+\href{../theories/html/hydras.Schutte.correctness_E0.html}%
 {\texttt{Schutte.Correctness\_E0}} establishes the link between two very different modelizations of ordinal numbers. In other words, it ``validates'' a data structure in terms of
@@
 \subsection{Remarks}
-Let us recall that the library \href{../src/html/hydras.Schutte.Schutte.html}%
+Let us recall that the library \href{../theories/html/hydras.Schutte.Schutte.html}%
 {\texttt{Schutte}} depends on five \emph{axioms} and lies explicitly in the  
@@
 Nevertheless, the other modules:
-\href{../src/html/hydras.Epsilon0.Epsilon0.html}%
+\href{../theories/html/hydras.Epsilon0.Epsilon0.html}%
 {\texttt{Epsilon0}},
-\href{../src/html/hydras.Hydra.Hydra.html}%
+\href{../theories/html/hydras.Hydra.Hydra.html}%
 {\texttt{Hydra}}, et 
-\href{../src/html/hydras.Gamma0.Gamma0.html}%
+\href{../theories/html/hydras.Gamma0.Gamma0.html}%
 {\texttt{Gamma0}}
@@
 
-\noindent\emph{From \href{../src/html/hydras.Schutte.Critical.html\#strongly_critical}%
+\noindent\emph{From \href{../theories/html/hydras.Schutte.Critical.html\#strongly_critical}%
 {\texttt{Schutte.Critical}}}
@@
 
-\noindent\emph{From \href{../src/html/hydras.Gamma0.T2.html\#T2}%
+\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T2}%
 {\texttt{Gamma0.T2}}}
@@
 
-\noindent\emph{From \href{../src/html/hydras.Gamma0.Gamma0\#tricho_aux}%
+\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0\#tricho_aux}%
 {\texttt{Gamma0.Gamma0}}}
@@
 
-\noindent\emph{From \href{../src/html/hydras.Gamma0.T2.html\#succ}%
+\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#succ}%
 {\texttt{Gamma0.T2}}}
@@
 
-\noindent\emph{From \href{../src/html/hydras.Gamma0.T2.html\#T1_to_T2}%
+\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T1_to_T2}%
 {\texttt{Gamma0.T2}}}
@@
 
-\item just go into the \texttt{src} directory, and type "\texttt{make}"
+\item just go into the \texttt{theories} directory, and type "\texttt{make}"
 \end{itemize}

### 2020-11-05 doc/hydras.tex
corrections in documentation
correction of typos

@@
 
+\section{How to Install the Libraries}
+\label{sec:orgheadline4}
+\begin{itemize}
+\item The present distribution has been checked with version 8.11.0 of the Coq proof assistant, with the plug-ins \texttt{coq-paramcoq} and \texttt{coq-equations}.
+
+\item Just go into the root directory, and type "\texttt{make}"
+\end{itemize}
+
+\section{Comments on Exercises and Projects}
+
+Although we do not plan to include complete solutions to the exercises, 
+we think it would be useful to include comments and hints, and questions/answers from the users. In constrast, ``projects'' are supposed, once completed, to be included in the repository.
+
 % \subsubsection{Links to the Coq source}
+
+\section{Future Work (projects)}
+\index{Projects}
+
+This document and the proof scripts are far from being complete.
+
+First, there must be a lot of typos to correct, references and index items to add. Many proofs are too complex and should be simplified, etc.
+
+The following extensions are planned, but help is needed:
+
+\begin{itemize}
+\item Semi automatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
+\item Extension to $\Gamma_0$ (in Veblen normal form)
+\item More lemmas about hierarchies of rapidly growing functions, and their relationship 
+    with primitive recursive functions and provability in Peano arithmetic 
+(following~\cite{KS81, KP82}).
+\item From \coq's point of view, this development could be used as an illustration of the evolution of the software, every time new libraries and sets of tactics could help to simplify the proofs.
+\end{itemize}
 
@@
 \begin{itemize}
-\item Hercules chops off always the rightmost head of the hydra.
+\item At each round, Hercules chops off the rightmost head of the hydra.
 \item The battle is standard: at the round number $i$, the expected replication is $i$.
@@
 
-We define an injection from the type \texttt{nat} into \texttt{Hydra}.
+We define an injection $\iota$ from the type \texttt{nat} into \texttt{Hydra}.
 For any natural number $i$, $\iota(i)$ is the hydra composed of a foot and
@@
 Let us follow the same pattern as in Sect.~\ref{omega-case}.
-First, we define an injection from type \texttt{t} into \texttt{Hydra}, by
+First, we define an injection $\iota$ from type \texttt{t} into \texttt{Hydra}, by
  associating to  each ordinal $\omega\times i+ j = (i,j)$ the hydra with $i$ branches of length $2$ and
@@
 
-Then we have  three cases to consider, according to the values of $i$ and $j$.
+Then we have  three cases to consider, according to the value of $p$.
 \begin{itemize}
@@
 \subsubsection{End of the Proof}
-
-Since $<$ is a strict order (irreflexive  and transitive) on \texttt{nat*nat}, we infer that there is no
-variant for termination on the lexicographic square of $(\mathbb{N},<)$.
-
-\begin{enumerate}
-\item From \texttt{m\_lt}, we infer the strict inequality 
-\texttt{m small\_h o< m big\_h}
-\item from \texttt{m\_ge}, we get \texttt{m big\_h o<= m (iota (m big\_h)) = m small\_h} 
-\end{enumerate}
+From \texttt{m\_ge}, we get \texttt{m big\_h o<= m small\_h = m (iota (m big\_h)) }. 
+Since $<$ is a strict order (irreflexive  and transitive), this inequality is incompatible with the strict inequality  \texttt{m small\_h o< m big\_h} (lemma \texttt{m\_lt}).
 
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
-
-\begin{Coqsrc}
-Theorem Impossible : False.
+\noindent
+In \coq (Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#Impossible}{Hydra.Omega2\_Small}):
+
+\begin{Coqsrc}
+Theorem Impossible: False.
 Proof.
@@
 Qed. 
-
 End Impossibility_Proof.
@@
 
-\section{Future Work (projects)}
-\index{Projects}
-
-This document and the proof scripts are far from being complete.
-
-First, there must be a lot of typos to correct, references and index items to add. Many proofs are too complex and should be simplified, etc.
-
-The following extensions are planned, but help is needed:
-
-\begin{itemize}
-\item Semi automatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
-\item Extension to $\Gamma_0$ (in Veblen normal form)
-\item More lemmas about hierarchies of rapidly growing functions, and their relationship 
-    with primitive recursive functions and provability in Peano arithmetic 
-(following~\cite{KS81, KP82}).
-\item From \coq's point of view, this development could be used as an illustration of the evolution of the software, every time new libraries and sets of tactics could help to simplify the proofs.
-\end{itemize}
-
-\section{How to Install the Libraries}
-\label{sec:orgheadline4}
-\begin{itemize}
-\item The present distribution has been checked with version 8.11.0 of the Coq proof assistant, with the plug-ins \texttt{coq-paramcoq} and \texttt{coq-equations}.
-
-\item just go into the \texttt{theories} directory, and type "\texttt{make}"
-\end{itemize}
-
-\section{Comments on Exercises and Projects}
-
-Although we do not plan to include complete solutions to the exercises, 
-we think it would be useful to include comments and hints, and questions/answers from the users. In constrast, ``projects'' are supposed, once completed, to be included in the repository.
-
 \section{Index}

### 2020-11-08 doc/hydras.tex
minor corrections

@@
 
-In some cases we want to show that two notation systems describe the same segment (for instance $[0,3+\omega)$ and $[0,\omega($\;).
+In some cases we want to show that two notation systems describe the same segment (for instance $[0,3+\omega)$ and $[0,\omega)$\;).
 For this purpose, one may prove that the two notation systems are order-isomorphic.
@@
 Let $i$ be some natural number. Prove that the notation systems 
-\texttt{Omega} and \texttt{ON\_plus (OrdFin $i$) Omega} are isomorphic.
-
-{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$, that we will prove in larger notation systems, as well as in Schütte's model.}
+\texttt{Omega} and (\texttt{ON\_plus (OrdFin $i$) Omega}) are isomorphic.
+
+{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$, that we prove amso in larger notation systems, as well as in Schütte's model.}
 
@@
 For instance, the
-elements of the product \texttt{ON\_mult Omega (FinOrd 3)} are ordered as follows.
+elements of the product (\texttt{ON\_mult Omega (FinOrd 3)}) are ordered as follows.
 \[(0,0),(0,1),(0,2),(0,3),(0,4),\dots,{\color{red}(1,0),} (1,1),(1,2),\dots, {\color{red}(2,0)},(2,1),(2,2),\dots\]

### 2020-11-08 doc/hydras.tex
correction of typos

@@
 
-{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$, that we prove amso in larger notation systems, as well as in Schütte's model.}
+{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$, that we prove also in larger notation systems, as well as in Schütte's model.}
 

### 2020-11-10 doc/hydras.tex
Added multiplication of ordinal notations (first redaction)

@@
 
-\section{Examples of Ordinal Notations}
-
-\subsection{The Ordinal \texorpdfstring{$\omega$}{omega}}
+\section{Example: The Ordinal \texorpdfstring{$\omega$}{omega}}
 
@@
 Qed.
-\end{Coqsrc}
-
-\subsection{Sum of two Ordinal Notations}
+
+Compute ON_compare 6 9.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+     = Lt
+     : comparison
+\end{Coqanswer}
+
+\section{Sum of  Ordinal Notations}
 
@@
 
+\noindent
 \emph{From Library~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}}.
@@
 
+\section{Product of Ordinal Notations}
+
+Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective  ordered types \texttt{A} and \texttt{B}.
+The product of \texttt{NA} and \texttt{NB} is considered as the concatenation of $B$ copies of $A$, ordered by the lexicographic order on $B\times A$.
+
+\vspace{4pt}
+\noindent
+\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_mult.html}{OrdinalNotations.ON\_mult}}
+
+\begin{Coqsrc}
+Section Defs.
+
+  Context `(ltA: relation A)
+          (compareA : A -> A -> comparison)
+          (NA: ON ltA compareA).
+  Context `(ltB: relation B)
+          (compareB : B -> B -> comparison)
+          (NB: ON ltB compareB).
+
+Definition t := (B * A)%type.
+Definition lt : relation t := lexico ltB ltA.
+Definition le := clos_refl _ lt.
+
+Definition compare (alpha beta: t) : comparison :=
+  match compareB (fst alpha) (fst beta) with
+  |  Eq => compareA (snd alpha) (snd beta)
+  | c => c
+  end.
+
+Lemma compare_reflect alpha beta :
+  match (compare alpha beta)
+  with
+    Lt => lt alpha  beta
+  | Eq => alpha = beta
+  | Gt => lt beta  alpha
+  end.
+
+Global Instance ON_mult : ON lt compare.
+
+End Defs.
+\end{Coqsrc}
+
 \section{The Ordinal \texorpdfstring{$\omega^2$}{omega^2}}
 
-The ordinal $\omega^2$ (also called $\phi_0(2)$, see Chap.~\ref{chap:schutte}), is represented by the lexicographically ordered cartesian product $\mathbb{N}\times\mathbb{N}$.
+The ordinal $\omega^2$ (also called $\phi_0(2)$, see Chap.~\ref{chap:schutte}), is an instance of the multiplication presented in the last section.
 
 \vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{ON\_Omega2}}
-
-\begin{Coqsrc}
-Definition t := (nat * nat)%type.
-
-Definition lt : relation t := lexico Peano.lt Peano.lt.
-Definition le := clos_refl _ lt.
-
-Infix "o<" := lt : o2_scope.
-Infix "o<=" := le : o2_scope.
-\end{Coqsrc}
-
-It is easy to represent  finite ordinals, successors and limits.
+\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
+
+\begin{Coqsrc}
+Definition Omega2 := ON_mult Omega Omega.
+Existing Instance Omega2.
+Definition t := ON_mult.t nat nat.
+\end{Coqsrc}
 
@@
 
-Notation "'F'" := fin : o2_scope.
-
 Coercion fin : nat >-> t.
 
-Definition succ (alpha : t) := (fst alpha, S (snd alpha)).
-
-Notation "'omega'" := (1,0) : o2_scope.
+Notation "'omega'" := (1,0) : ON_scope.
 
@@
 
-In order to build an ordinal notation out of the type \texttt{t}, we have to define a comparison function, and prove that the order \texttt{lt} is well-founded.
-
-\begin{Coqsrc}
-Instance lt_strorder : StrictOrder lt.
-(* ... *)
-
-Definition compare (alpha beta: t) : comparison :=
-  match Nat.compare (fst alpha) (fst beta) with
-    Eq => Nat.compare (snd alpha) (snd beta)
-  | c => c
-  end.
-
-Lemma compare_correct alpha beta :
-    CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
-                (compare alpha beta).
-
-Lemma lt_wf : well_founded lt.
-
-Instance Omega2: ON lt compare.
-Proof.
-  split.
-  - apply lt_strorder.
-  - apply lt_wf.
-  - apply compare_correct.
-Qed.
-\end{Coqsrc}
-
 \subsection{Arithmetic of \texorpdfstring{$\omega^2$}{omega^2}}
 
+\subsubsection{Successor}
+
+The successor of any ordinal is defined by a simple pattern-matching.
+This function is proved to be correct w.r.t. the \texttt{Successor} predicate.
+
+\begin{Coqsrc}
+Definition succ (alpha : t) := (fst alpha, S (snd alpha)).
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Lemma lt_succ_le alpha beta : alpha o< beta <-> succ alpha o<= beta.
+
+Lemma lt_succ alpha : alpha o< succ alpha.
+
+Lemma succ_ok alpha beta : Successor beta alpha <-> beta = succ alpha.
+\end{Coqsrc}
+
 \subsubsection{Addition}
 
-We can define on \texttt{ON\_omega2} an addition which extends the addition on \texttt{nat}.
+We can define on \texttt{Omega2} an addition which extends the addition on \texttt{nat}.
 
@@
 
- Context (Hvar : Hvariant lt_wf free m).
+ Context (Hvar : Hvariant (Generic.wf (ON:=Omega2))  free m).
 \end{Coqsrc}
@@
 
-\index{Exercises}
-\begin{exercise}
-
-\label{sec:orgheadline63}
-Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.
-
-\begin{itemize}
-\item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
-  lexicographic ordering
-\item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.
-
-For instance, the following inequality holds:
-\[\langle 4,3,3,3,3,3,3,2,2,2 \rangle\,<\,\langle 4,4,2 \rangle\]
-\end{itemize}
-
-\end{exercise}
+% \index{Exercises}
+% \begin{exercise}
+
+% \label{sec:orgheadline63}
+% Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.
+
+% \begin{itemize}
+% \item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
+%   lexicographic ordering
+% \item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.
+
+% For instance, the following inequality holds:
+% \[\langle 4,3,3,3,3,3,3,2,2,2 \rangle\,<\,\langle 4,4,2 \rangle\]
+% \end{itemize}
+
+% \end{exercise}
 
@@
 \begin{project}
-Define in \coq{} the product of two ordinal notations $N_A$ and $N_B$.
-If $A$ [resp. $B$] is the underlying type of $N_A$ [resp. $N_B$], the
-product \texttt{ON\_mult $N_A$ $N_B$} is implemented over the cartesian product $B\times A$ (with the lexicographic ordering).
+% Define in \coq{} the product of two ordinal notations $N_A$ and $N_B$.
+% If $A$ [resp. $B$] is the underlying type of $N_A$ [resp. $N_B$], the
+% product \texttt{ON\_mult $N_A$ $N_B$} is implemented over the cartesian product $B\times A$ (with the lexicographic ordering).
+
+This exercise is about the non-commutativity of the multiplication of ordinals, reflected in ordinal notations.
 
@@
 
-Prove that \texttt{ON\_mult (FinOrd $i$) Omega} is isomorphic to
+Prove formally  that \texttt{ON\_mult (FinOrd $i$) Omega} is isomorphic to
 \texttt{Omega}  whilst

### 2020-11-15 doc/chapter-powers.tex
restructuration of latex files

\chapter{Smart Computation of \texorpdfstring{$x^n$}{Powers}}
\label{chapter-powers}
\section{Introduction}

Nothing looks simpler than writing some function for computing $x^n$.
On the contrary, this simple programming exercise allows us to address
advanced programming techniques such as:
\begin{itemize}
\item monadic programming, and continuation passing style
\item type classes, and generalized rewriting
\item proof engineering, in particular proof re-using
\item proof by reflection
\item polymorphism and parametricity
\item composition of correct programs, etc.
\end{itemize}

\section{Some basic implementations}
\label{sect:linear-naive}
Let us start with a very naive way of computing the $n$-th power of $x$, where
$n$ is a natural number and $x$ belongs to some type for which a multiplication and an identity element are defined.

\emph{From Module 
\href{../src/html/addition.power.FirstSteps.html}{\texttt{Powers.FirstSteps}}}
\label{sect: power-definitions}

\begin{Coqsrc}
Section Definitions.

Variables (A: Type)
           (mult: A -> A -> A)
           (one: A).
Local Infix "*" := mult.
Local Notation "1" := one.
\end{Coqsrc}

\begin{Coqsrc}
Fixpoint power (x:A)(n:nat) : A :=
  match n with 0%nat => 1
             | S p =>   x * x ^ p
  end
where "x ^ n" := (power x n).

\end{Coqsrc}

\begin{Coqsrc}
Compute power Z.mul 1%Z 2%Z 10.  
\end{Coqsrc}
\begin{Coqanswer}
   = 1024%Z
     : Z
\end{Coqanswer}

\begin{Coqsrc}
Open Scope string_scope.
Compute power append  "" "ab"  12.
\end{Coqsrc}

\begin{Coqanswer}
 = "abababababababababababab"
     : string
\end{Coqanswer}

The number of multiplications needed to compute $x^n$ via this function is linear with respect to $n$. Despite this lack of efficiency, and thanks to its simplicity, we keep it as a specification for more efficient and complex exponentiation algorithms.
A function will be considered a \emph{correct} exponentiation function if we can prove it is extensionally equivalent to \texttt{power}.

% \subsection{A semi-naive algorithm}

% In versions up to \texttt{V8.9.1}, the exponentiation function on type \texttt{Z} was defined as follows,
% (in modules \texttt{Coq.PArith.BinPosDef.Pos} and \texttt{Coq.ZArith.BinIntDef.Z}.

% \begin{Coqsrc}
% (** ** Iteration of a function over a positive number *)

% Definition iter {A} (f:A -> A) : A -> positive -> A :=
%   fix iter_fix x n := match n with
%     | xH => f x
%     | xO n' => iter_fix (iter_fix x n') n'
%     | xI n' => f (iter_fix (iter_fix x n') n')
%   end.

% Definition pow (x:positive) := iter (mul x) 1.
% \end{Coqsrc}

% \begin{Coqsrc}
% Definition pow_pos (z:Z) := Pos.iter (mul z) 1.

% Definition pow x y :=
%   match y with
%     | pos p => pow_pos x p
%     | 0 => 1
%     | neg _ => 0
%   end.

% Infix "^" := pow : Z_scope.
% \end{Coqsrc}

% At first sight, the function \texttt{Pos.pow} seems to be logarithmic because of the recursive structure of the help function \texttt{iter\_fix}. Unfortunately, it is obvious that a call to 
% \texttt{iter f x n} will apply $n$ times the function $f$. Thus, these exponentiation functions with binary exponents are in fact linear!

% \label{sect:slow-computation}

% \begin{Coqsrc}
% Time Compute (1 ^ 56666667)%N.
% \end{Coqsrc}

% \begin{Coqanswer}
% Finished transaction in 3.604 secs (3.587u,0.007s)   
% \end{Coqanswer}

\subsection{A logarithmic exponentiation  function}

Using the following equations, we can easily define a polymorphic exponentiation whose application requires only a logarithmic number of multiplications.

\begin{align}
x^1 &= x \label{binary-eq1}\\
x^{2p} &= (x^2)^p \label{binary-eq2}\\
x^{2p+1} &= (x^2)^p \times x \label{binary-eq3}\\
x^1 \times a &= x \times a \label{binary-eq4}\\
x^{2p} \times a  &= (x^2)^p \times a\label{binary-eq5}\\
x^{2p+1} \times a  &= (x^2)^p \times (a\times x)\label{binary-eq6}
\end{align}

In equalities \ref{binary-eq4} to \ref{binary-eq6}, the variable $a$ plays the role
of an \emph{accumulator} whose initial value (set by \ref{binary-eq3}) is $x$.
This accumulator helps us to get a tail-recursive implementation.

For instance, the computation of $2^{14}$ can be decomposed as follows:
\begin{align*}
2^{14} &= 4^{7} \\
      &= 16^3 \times 4 \\
      &= 256^1 \times (4 \times 16) \\
      &= 16384  
\end{align*}

With the same notations as in Sect~\vref{sect:linear-naive}, we can implement this algorithm in \gallina. The following definitions are still within the scope of the 
section open in~\vref{sect: power-definitions}.

\label{polymorhic-binary_exp}

%%% ICI (presenter la fonction binaire de First_Steps )
%%%  Reprendre des explications placees ci-dessous

\vspace{4pt}

\emph{From Module
\href{../src/html/addition.power.FirstSteps.html}{Powers.FirstSteps}}

\begin{Coqsrc}
Fixpoint binary_power_mult (x a:A)(p:positive) : A 
  :=
  match p with
    | xH =>  a * x
    | xO q => binary_power_mult  (x * x) a q
    | xI q =>  binary_power_mult  (x * x) (a * x) q
  end.

Fixpoint Pos_bpow (x:A)(p:positive) :=
 match p with
  | xH => x
  | xO q => Pos_bpow  (x * x) q
  | xI q => binary_power_mult   (x * x) x q
end.

Definition N_bpow x (n:N) := 
  match n with 
  | 0%N => 1
  | Npos p => Pos_bpow x p
  end.

End Definitions.
\end{Coqsrc}

Let us  close the section \texttt{Definitions} and mark the argument \texttt{A} as implicit.

\begin{Coqsrc}
End Definitions.

Arguments N_bpow {A}.
Arguments power {A}.
\end{Coqsrc}

\begin{remark}
Our function \texttt{Pos\_bpow} can be considered as a tail recursive variant
of the following function defined in \texttt{Coq.PArith.BinPosDef}.

\begin{Coqsrc}
Definition iter_op {A}(op:A->A->A) :=
  fix iter (p:positive)(a:A) : A :=
  match p with
    | 1 => a
    | p~0 => iter p (op a a)
    | p~1 => op a (iter p (op a a))
  end.
\end{Coqsrc}

This scheme is used in \texttt{Coq.ZArith.Zpow\_alt} in order to define a logarithmic exponentiation \texttt{Zpower\_alt} on \texttt{Z} (notation : $x\,\texttt{\^{}\^{}}\,p$).

\end{remark}

\paragraph*{Remark}
Note that closing the section \texttt{Definitions} makes us lose the
handy notations \texttt{\_ * \_} and \texttt{one}. Fortunately, \emph{operational type classes} will help us to define nice infix notations for polymorphic functions (Sect.~\vref{op-classes}).

\subsection{Examples of computation}
It is now possible to test our functions with various interpretations of
$\times$ and $1$:

\begin{Coqsrc}
Compute N_bpow Z.mul 1%Z 2%Z 10. 
\end{Coqsrc}

\begin{Coqanswer}
 = 1024%Z
     : Z 
\end{Coqanswer}

\begin{Coqsrc}
Require Import String.
Open Scope string_scope.

Compute N_bpow append "" "ab"  12.
\end{Coqsrc}

\begin{Coqanswer}
  = "abababababababababababab"
     : string 
\end{Coqanswer}

\subsubsection{Exponentiation on $2\times 2$ matrices}
\label{naive-matrix}
Our second example is a definition of $M^n$ where $M$ is a $2\times 2$ matrix
over any ``scalar''  type $A$, assuming one can provide $A$ with a semi-ring structure~\cite{Coq}.

%\subsubsection{Representation of $2\times 2$ matrices}

A $2\times 2$ matrix will be simply represented by a structure with four fields;
each field \texttt{c$ij$} is associated with the $i$-th line and $j$-th column of the considered matrix.

\begin{Coqsrc}
Module M2.
Section Definitions.

  Variables (A: Type) (zero one : A)  (plus mult  : A -> A -> A).

  Variable rt : semi_ring_theory  zero one plus mult   (@eq A).
  Add Ring Aring : rt.

  Notation "0" := zero.  
  Notation "1" := one.
  Notation "x + y" := (plus x y).  
  Notation "x * y " := (mult x y).

  Structure t : Type := mat{c00 : A;  c01 : A;  c10 : A;  c11 : A}.
\end{Coqsrc}

%\subsubsection{Matrix Multiplication}

The structure type \texttt{M2.t} allows us to define the product
of two matrices.

\begin{Coqsrc}
Definition M2_mult (M M':t) : t := mat
  (c00 M * c00 M' + c01 M * c10 M') (c00 M * c01 M' + c01 M * c11 M')
  (c10 M * c00 M' + c11 M * c10 M') (c10 M * c01 M' + c11 M * c11 M').
\end{Coqsrc}

The neutral element for \texttt{M2\_mult} is the identity matrix.

\begin{Coqsrc}
Definition Id2 : t := mat 1 0 0 1.

End M2_Definitions.
End M2.
\end{Coqsrc}

Matrix exponentiation is a well-known method for computing Fibonacci numbers:
\index{Fibonacci numbers!Matrix exponentiation}

\begin{Coqsrc}

Import M2.

Arguments M2_mult {A} plus mult  _  _.
Arguments mat {A} _ _ _ _.
Arguments Id2 {A}  _ _.

Definition fibonacci (n:N) :=
 c00 N  (N_bpow  (M2_mult Nplus Nmult) 
                 (Id2  0%N 1%N)
                 (mat  1 1 1 0)%N 
                 n).

Compute fibonacci 20.
\end{Coqsrc}

\begin{Coqanswer}
= 10946%N
     : N  
\end{Coqanswer}

\subsubsection{Remark}
\label{sect:faster}

Our function \texttt{N\_bpow} is really logarithmic. Let us make a comparative 
test with Standard Library's exponentiation function on type \texttt{N} (see section~\vref{sect:slow-computation}).

\begin{Coqsrc}
Time Compute (N_bpow N.mul 1 1 56666667)%N.  
\end{Coqsrc}

\begin{Coqanswer}
Finished transaction in 0. secs (0.u,0.s) (successful)  
\end{Coqanswer}

\subsection{Formal specification of an exponentiation function: a first attempt}

Let us compare the functions \texttt{power} and \texttt{N\_bpow}.
The first one is obviously correct, since it is a straightforward translation of the mathematical definition.
The second one is much more efficient, but it is not obvious  that its 18-line long definition is bug-free.
Thus, we must prove that the two functions are extensionally equal (taking into account conversions
between \texttt{N} and \texttt{nat}).

More abstractly, we can define a predicate that characterizes any correct implementation 
of \texttt{power}, this ``naive''  function being a \emph{specification} of any polymorphic
exponentiation function.

First, we define a type for any such function.

\begin{Coqsrc}
Definition power_t := forall (A:Type)
                             (mult : A -> A -> A)
                             (one:A)
                             (x:A)
                             (n:N), A.
\end{Coqsrc}

Then, we would say that a function \texttt{f:power\_t} is a correct exponentiation function if it
is extensionally equal to \texttt{power}.

\begin{Coqbad}
Module Bad.

 Definition correct_expt_function (f : power_t) : Prop :=
  forall A (mult : A -> A -> A) (one:A)
            (x:A) (n:N), 
            power mult one x (N.to_nat n) = f A mult one x n.
\end{Coqbad}

Unfortunately, our definition of \texttt{correct\_expt} is too general. It suffices to build 
an interpretation where the multiplication is not associative or \texttt{one} is not a neutral
element to obtain different results through the two functions.

\begin{Coqbad}
Section CounterExample.
    Let mul (n p : nat) := n + 2 * p.
    Let one := 0.

    Remark mul_not_associative :
      exists  n p q,  mul n (mul p q) <> mul (mul n p) q.
    Proof. 
        exists 1, 1, 1; discriminate. 
    Qed.

    Remark one_not_neutral  :
      exists n : nat, mul one n <> n.
    Proof.
      exists 1; discriminate.
    Qed.

    Lemma correct_expt_too_strong : 
          ~ correct_expt_function (@N_bpow).
    Proof.
      intro H; specialize (H _ mul one 1  7%N).
      discriminate H.    
    Qed.

End CounterExample.
End Bad.
 \end{Coqbad}

So, we will have to improve our definition of correctness, by restricting  the universal quantification to associative operations and neutral elements, \emph{i.e.}, by considering \emph{monoids}.
An exponentiation  function will be considered as correct if it returns always the same result as \texttt{power} \emph{in any monoid}.

\section{Representing Monoids in \coq \label{monoid-class-def}}

In this section, we present a ``minimal'' algebraic framework in which  exponentiation can be defined and efficiently implemented.

Exponentiation is built on multiplication, and many properties of 
this operation are derived from the associativity of multiplication. 
Furthermore, if we allow the exponent to be any natural number, including $0$, 
then we need to consider a neutral element for multiplication.

The structure on which we define exponentiation is called a \emph{monoid}.
It is composed of a \emph{carrier} $A$, an associative binary operation $\times$ on $A$, and a neutral element $\mathds{1}$ for $\times$ . The required properties of $\times$ and
$\mathds{1}$ are expressed by the following equations:

\begin{align}
  \label{eq}
  \forall x\,y\,z\,:A,\, x\times (y \times z) &= (x\times y) \times z
  \\
\forall x:A,\, x \times \mathds{1}  &= \mathds{1}  \times x = x
\end{align}

In \coq{}, we define the monoid structure in terms of 
\emph{type classes}\cite{MS08,BS2011}. The tutorial on type classes \cite{PCMS} gives more details on type classes and
operational type classes, also illustrated with the monoid structure.

First, we define a class and a notation for representing multiplication operators, then we use
these definitions for defining the \texttt{Monoid} type class.

\subsection{A common notation for multiplication}
\label{op-classes}
\index{Type Classes!Operational Type Classes}
\emph{Operational type classes}~\cite{BS2011}
allow us to define a common notation 
for multiplication in any algebraic structure. 
First, we associate a class to the notion of \emph{multiplication} 
on any type $A$.

\emph{From Module \href{../src/html/addition.power.Monoid_def.html}{Powers/Monoid\_def.v}.}

\begin{Coqsrc}
Class Mult_op (A:Type) := mult_op : A -> A -> A.  
\end{Coqsrc}

From the type theoretic point of view, the term (\texttt{Mult\_op $A$}) is 
$\beta\delta$-reducible to \texttt{$A\arrow A \arrow A$}, and
if \texttt{\it op} has type (\texttt{Mult\_op $A$}), then 
(\texttt{@mult\_op A {\it op}}) is convertible with \texttt{\it op}.

We are now ready to define a new notation scope, in which the notation
\texttt{x * y} will be interpreted as an application of the function
\texttt{mult\_op}.

\begin{Coqsrc}
Delimit Scope M_scope with M.
Infix "*" := mult_op : M_scope.
Open Scope M_scope.  
\end{Coqsrc}

 Let us show two examples of use of the
notation scope \texttt{M\_scope}. Each example consists in declaring an 
instance of \texttt{Mult\_op}, then type checking or evaluating
a term of the form \texttt{x * y} in \texttt{M\_scope}.

Note that, since the reserved notation \texttt{"\_ * \_ "} is 
present in several scopes such as  \texttt{nat\_scope}, \texttt{Z\_scope},
\texttt{N\_scope}, etc., in addition to  \texttt{M\_scope},  the user should
take care of which scopes are active --- and with  which precedence --- in a \gallina{} term.
In case of doubt, explicit scope delimiters should be used.

\subsubsection{Multiplication on Peano Numbers}

Multiplication  on type \texttt{nat}, called \texttt{Nat.mul} in
Standard Library, has  type \linebreak \texttt{nat -> nat -> nat}, which is
convertible  with \texttt{Mult\_op nat}. Thus the following definition is
accepted:

\begin{Coqsrc}
Instance nat_mult_op : Mult_op nat  := Nat.mul.
\end{Coqsrc}

Inside \texttt{M\_scope}, the expression \texttt{3 * 4} is 
correctly read as an application of \texttt{mult\_op}. Nevertheless 
this term is convertible with \texttt{Nat.mul 3 4}, as shown by the 
interaction below.

\emph{From Module \href{../src/html/addition.power.Monoid_def.html}{Powers.Monoid\_def}}

\begin{Coqsrc}
Set Printing All.
Check  3 * 4. 
\end{Coqsrc}
\begin{Coqanswer}
@mult_op nat nat_mult_op (S (S (S O))) (S (S (S (S O))))
     : nat  
\end{Coqanswer}
\begin{Coqsrc}
Unset Printing All.
Compute 3 * 4.   
\end{Coqsrc}
\begin{Coqanswer}
 = 12  : nat  
\end{Coqanswer}

\subsubsection{String Concatenation}
We can use the notation \texttt{"\_ * \_ "} for other types than numbers.
In the following example,  the expression \texttt{"abc" * "def"} is interpreted
as \linebreak \texttt{@mult\_op string  {\color{darkred}?X} "abc"  "def"}, then the type  class mechanism replaces the unknown  {\color{darkred}?X} with 
\texttt{string\_op}.

\emph{From Module \href{../src/html/addition.power.Monoid_def.html}{Powers.Monoid\_def}}

\begin{Coqsrc}
Require Import String.

Instance string_op : Mult_op string := append.
Open Scope string_scope.

Example ex_string : "ab" * "cde" = "abcde".
Proof. reflexivity. Qed.
\end{Coqsrc}

\subsubsection{Solving Ambiguities}
Let $A$ be some type, and let us assume there are several instances of
\texttt{Mult\_op $A$}. For solving ambiguity issues, one can
add a \emph{precedence} to each instance declaration of  
\texttt{Mult\_op $A$}. In any case, such ambiguity  can be addressed
by explicitly providing  some arguments of \texttt{mult\_op}.
For instance, in Sect.~\vref{nat-monoids}, we consider various monoids on types
\texttt{nat} and \texttt{N}.

\subsection{The Monoid Type Class}
\index{Type Classes}
We are now ready to  give a definition of the \texttt{Monoid} class, using
\texttt{*} as an infix operator in scope \coqscope{M} for the monoid  multiplication.

The following class definition, from Module \href{../src/html/addition.power.Monoid_def.html}{Powers.Monoid\_def},
is parameterized with some type $A$,
a multiplication (called \texttt{op} in the definition), and a neutral element
$\mathds{1}$ (called \texttt{one} in the definition).

\begin{Coqsrc}
Class Monoid {A:Type}(op : Mult_op A)(one : A) : Prop :=
{
    op_assoc : forall x y z:A, x * (y * z) = x * y * z;
    one_left : forall x, one * x = x;
    one_right : forall x, x * one = x
}.
\end{Coqsrc}

\subsection{Building Instances of \texttt{Monoid}}
Let \texttt{$A$} be some type, \texttt{{\it op}} an instance of 
\texttt{Mult\_op $A$} and \texttt{\it one: $A$}.
In order to build an instance of (\texttt{Monoid $A$ {\it op} {\it one}}),
one has to provide proofs of ``monoid axioms'' \texttt{ op\_assoc},
\texttt{one\_left} and \texttt{one\_right}.

Let us show various instances, which will be used in further proofs and examples.
Complete definitions and proofs are given in 
File~\href{../src/html/addition.power.Monoid_instances.html}{Powers/Monoid\_instances.v}.

\subsubsection{Monoid on \texttt{Z}}
The following monoid allows us to compute powers of integers of arbitrary size, 
using type \texttt{Z} from standard library:

\begin{Coqsrc}
Instance Z_mult_op : Mult_op Z := Z.mul.

Instance ZMult : Monoid  Z_mult_op 1.
Proof. 
  split.
\end{Coqsrc}

\begin{Coqanswer}
3 subgoals, subgoal 1 (ID 8)

  ============================
   forall x y z : Z, (x * (y * z))%M = (x * y * z)%M

subgoal 2 (ID 9) is:
 forall x : Z, (1 * x)%M = x
subgoal 3 (ID 10) is:
 forall x : Z, (x * 1)%M = x}
\end{Coqanswer}

\begin{Coqsrc}
    all: unfold Z_mult_op, mult_op;intros;ring.
Qed.
\end{Coqsrc}

\subsubsection{Monoids on type \texttt{nat} and \texttt{N}}
\label{nat-monoids}
% ~~\\
% \noindent

We define two monoids on type \texttt{nat}:
\begin{itemize}
\item The ``natural'' monoid $(\mathbb{N},\times, 1)$ :

  \begin{Coqsrc}
Instance nat_mult_op : Mult_op nat | 5 := Nat.mul.

Instance  Natmult : Monoid nat_mult_op  1%nat | 5
Proof.
   split;unfold nat_mult_op, mult_op; intros; ring.
Qed.
\end{Coqsrc}

\item The ``additive''  monoid $(\mathbb{N},+, 0)$.
This monoid will play an important role in correctness proofs of complex
exponentiation algorithms. Its most important property is that the $n$-th 
power of $1$ is equal to $n$. See Sect.~\vref{correctness-for-free} for more details.

\begin{Coqsrc}
Instance nat_plus_op : Mult_op nat | 12 := Nat.add.

Instance Natplus : Monoid nat_plus_op  0%nat | 12.
(* Proof omitted *)
\end{Coqsrc}
\end{itemize}

Similarly, instances \texttt{NMult} and  \texttt{NPlus}  are built for type \texttt{N}, and
\texttt{PMult} for type \texttt{positive}.

\subsubsection{Machine integers}

Cyclic numeric types are  good candidates for testing exponentiations
with big exponents, since the size of data is bounded.

The type \texttt{int31} is defined  in Module
\textbf{Coq.Numbers.Cyclic.Int31.Int31} of \coq's standard library. The tactic \texttt{ring} works 
with this type, and helps us to register an instance \texttt{Int31Mult} of class  \texttt{Monoid int31\_mult\_op 1}.

\begin{Coqsrc}
Instance int31_mult_op : Mult_op int31 := mul31.

Instance  Int31mult : Monoid int31_mult_op  1.
Proof.
   split;unfold int31_mult_op, mult_op; intros; ring.
Qed.
\end{Coqsrc}

Beware that machine integers are not natural numbers!

\begin{Coqbad}
Module Bad.

Fixpoint int31_from_nat (n:nat) :=
  match n with
  | O => 1
  | S p => 1 + int31_from_nat p
  end.

Coercion int31_from_nat : nat >-> int31.

Fixpoint fact (n:nat) := 
  match n with
   | O => 1
   | S p => n * fact p
  end.

Example fact_zero : exists n:nat, fact n = 0.
Proof.  now exists 40%nat.  Qed.

End Bad.
\end{Coqbad}

\subsection{Matrices on a semi-ring}

In Sect.\vref{naive-matrix}, we defined a function for computing powers of any $2\times 2$ 
matrix over any semi-ring. For proving a simple property of matrix exponentiation, we had
 to prove that matrix multiplication is associative and admits the identity matrix as a neutral element. These properties are easily expressed within the type class framework, by defining a \emph{family} of monoids.
It suffices to define an instance of \texttt{Monoid} within the scope of an hypothesis
of type \texttt{semi\_ring\_theory}

\begin{Coqsrc}
Section M2_def.
Variables (A:Type)
           (zero one : A) 
           (plus mult  : A -> A -> A).

 Variable rt : semi_ring_theory  zero one plus mult  (@eq A).
 Add  Ring Aring : rt.
\end{Coqsrc}

\begin{Coqsrc}
Structure M2 : Type := {c00 : A;  c01 : A;
                        c10 : A;  c11 : A}.

Definition Id2 : M2 := Build_M2 1 0 0 1.

Definition M2_mult (m m':M2) : M2 :=
 Build_M2 
          (c00 m * c00 m' + c01 m * c10 m')
          (c00 m * c01 m' + c01 m * c11 m')
          (c10 m * c00 m' + c11 m * c10 m')
          (c10 m * c01 m' + c11 m * c11 m').

Global Instance M2_op : Mult_op M2 := M2_mult.
\end{Coqsrc}

\begin{Coqsrc}
Global Instance M2_Monoid : Monoid   M2_op Id2.
(* Proof omitted *)

End M2_def.

Arguments M2_Monoid {A zero one plus mult} rt.
\end{Coqsrc}

\subsection{Monoids and Equivalence Relations}
\index{Generalized rewriting}
\index{Equivalence relations}

In some contexts, the ``axioms'' of the \texttt{Monoid} class  may be too restrictive.
For instance, consider multiplication in $\mathds{Z}/m\mathds{Z}$ where
 $1<m$.
Although it could be possible to compute with values of the dependent 
type \texttt{\{n:N | n < m\}}, 
it looks simpler to compute with numbers of type
\texttt{N} and consider the multiplication $x \times y \mod{m}$.

It is easy to prove that this operation is associative, using library \texttt{NArith}. Unfortunately, the following proposition is false in general (left as an exercise).

$$\forall x:N, (1 * x) \mod{m} = x$$

Thus, we define a more general class, parameterized by an equivalence
relation \texttt{Aeq}  on a type \texttt{A}, compatible with the multiplication \texttt{*}. The laws of associativity and neutral element
are not expressed as Leibniz equalities but as equivalence statements:

First, let us define an operational type class for equivalence relations:

\vspace{4pt}

\noindent
\emph{From Module \href{../src/html/addition.power.Monoid_def.html}{Powers.Monoid\_def}}

\begin{Coqsrc}
Class Equiv A := equiv : relation A.

Infix "==" := equiv (at level 70) : type_scope.
\end{Coqsrc}

The definition of class \texttt{EMonoid} looks like \texttt{Monoid}'s definition, 
plus some constraints on \texttt{E\_eq}.

Please look for instance at our tutorial on type classes and relations~\cite{PCMS} 
for understanding the use of  type classes \texttt{Equivalence}, \texttt{Reflexive}, \texttt{Proper}, etc, in relation with tactics like \texttt{rewrite}, \texttt{reflexivity}, etc., in proofs which involve  equivalence relations instead of equality.

\index{Type Classes!Proper}
\label{EMonoid-def}

%\todo{link to Proper in stdlib : Coq.Classes.Morphisms and Coq.Classes.CMorphisms}

\begin{Coqsrc}
Class EMonoid (A:Type)(E_op : Mult_op A)(E_one : A) 
      (E_eq: Equiv A): Prop :=
  {
    Eq_equiv :> Equivalence equiv;
    Eop_proper :> Proper (equiv ==> equiv ==> equiv) E_op;
    Eop_assoc : forall x y z, x * (y * z) == x * y * z;
    Eone_left : forall x,  E_one * x == x;
    Eone_right : forall x,  x * E_one ==  x
  }.
\end{Coqsrc}

\subsubsection{Coercion from Monoid to EMonoid} 
Every instance of class  \texttt{Monoid} can be transformed into an instance of
\texttt{EMonoid}, considering Leibniz' equality \texttt{eq}.
Thus, our  definitions and theorems about exponentiation will take place as 
much as possible within the more generic framework of \texttt{EMonoid}s.

\index{Coercions}

\begin{Coqsrc}
Global Instance eq_equiv {A} : Equiv A := eq.

Global Instance Monoid_EMonoid `(M:@Monoid A op one) :
        EMonoid  op one eq_equiv.
Proof.
split; unfold eq_equiv, equiv in *.
 - apply eq_equivalence.
 - intros x y H z t H0; now subst.
 - intros; now rewrite (op_assoc).
 - intro; now rewrite one_left.
 - intro; now rewrite one_right.
Defined.
\end{Coqsrc}

\begin{remark}
In the definition of \texttt{Monoid\_EMonoid}, the free variables  \texttt{A}, 
\texttt{op} and \texttt{one} are automatically generalized thanks to the \emph{backquote} syntax (see the section about implicit generalization in the reference manual~\cite{Coq}).
\end{remark}

Thanks to the following \emph{coercion}, every instance of \texttt{Monoid} can 
now be considered as an instance of \texttt{EMonoid}. For more details, please look at the section \emph{Implicit Coercions} of \coq's reference manual~\cite{Coq}.

\begin{Coqsrc}
Coercion Monoid_EMonoid : Monoid >-> EMonoid.
\end{Coqsrc}

%ici%

\emph{From Module \href{../src/html/addition.power.Monoid_instances.html}{Powers.Monoid\_instances}}

\begin{Coqsrc}
Check NMult : EMonoid  N.mul 1%N eq.
\end{Coqsrc}

\begin{Coqanswer}
  NMult:EMonoid N.mul 1%N eq
     : EMonoid N.mul 1%N eq
\end{Coqanswer}

\subsubsection{Example : Arithmetic  modulo $m$}

The following instance of \texttt{EMonoid} describes the set of integers modulo
$m$, where $m$ is any integer greater than or equal to $2$.
For simplicity's sake, we represent such values using the \texttt{N} type,
and consider ``equivalence modulo \texttt{$m$}'' instead of equality.
Note that the law of associativity has been stated as Leibniz' equality.

\begin{Coqsrc}
Section Nmodulo.
  Variable m : N.
  Hypothesis m_gt_1 : 1 < m.

  Definition mult_mod ( x y : N) := (x * y) mod m.
  Definition mod_eq ( x y: N) := x mod m = y mod m.

  Global Instance mod_equiv : Equiv N := mod_eq.

  Global Instance mod_op : Mult_op N := mult_mod.

  Global Instance mod_Equiv : Equivalence mod_equiv.
  (* Proof omitted *)

  Global Instance mult_mod_proper : 
  Proper (mod_equiv ==> mod_equiv ==> mod_equiv)  mod_op.
  (* Proof omitted *)

  Local Open Scope M_Scope.

  Lemma mult_mod_associative :  
  forall x y z,  x * (y * z) = x * y * z.
  (* Proof omitted *)

  Lemma one_mod_neutral_l  : forall x, 1 * x ==  x.
  (* Proof omitted *)

  Lemma one_mod_neutral_r  : forall x, x * 1 == x.
  (* Proof omitted *)

  Global Instance Nmod_Monoid : EMonoid  mod_op 1 mod_equiv.
  (* Proof omitted *)

End Nmodulo.

\end{Coqsrc}

\paragraph{Example}
In the following interaction, we show how to instantiate the parameter \texttt{$m$} to a 
concrete value, for instance \texttt{$256$}.

\begin{Coqsrc}
Section S256.
Let mod256 :=  mod_op 256.
Local Existing Instance mod256 | 1.

Compute (211 * 67)
\end{Coqsrc}

\begin{Coqanswer}
= 57 : N  
\end{Coqanswer}

\begin{Coqsrc}
End S256.
\end{Coqsrc}

Outside the section \texttt{S256}, the term \texttt{(211 * 67)\%M} is interpreted as a plain multiplication in type \texttt{N}:

\begin{Coqsrc}
Compute (211 * 67)%M.
\end{Coqsrc}

\begin{Coqanswer}
= 14137 : N   
\end{Coqanswer}

\section{Computing Powers in any EMonoid}

The  module \href{../src/html/addition.power.Pow.html}{Powers.Pow} defines two functions for exponentiation on any 
\texttt{EMonoid}  on carrier $A$.
They are essentially the same as in Sect.~\vref{sect: power-definitions}. The main difference lies in the arguments of the functions, which now contain
 an instance~\texttt{M} of class \texttt{EMonoid}. 
Thus, the arguments associated with the multiplication,
the neutral element and the equivalence relation associated with \texttt{M}
are left implicit.

\subsection{The naive (linear) Algorithm}
The new version of the linear exponentiation function is as follows:

\begin{Coqsrc}
Fixpoint power`{M: @EMonoid A  E_op E_one E_eq} 
               (x:A) (n:nat) :=
match n with 
| 0%nat => E_one
| S p =>   x * x ^ p
end
where "x ^ n" := (power x n) : M_scope.
\end{Coqsrc}

The three following lemmas will be used by the \texttt{rewrite} tactic in further
correctness proofs.
Note  that the first two lemmas are strong
(\emph{i.e.}, Leibniz) equalities, whilst \texttt{power\_eq3}  is only an equivalence statement, because its proof uses one of the \texttt{EMonoid} laws, namely
\texttt{Eone\_right}.

\begin{Coqsrc}
Lemma power_eq1 {A:Type} `{M: @EMonoid A  E_op E_one E_eq} 
               (x:A) :  x ^ 0 = E_one.
Proof. reflexivity. Qed.

Lemma power_eq2 {A:Type} `{M: @EMonoid A  E_op E_one E_eq}
                (x:A) (n:nat) :
                x ^ (S n)  = x * x ^ n.
Proof. reflexivity. Qed.

Lemma power_eq3 {A:Type} `{M: @EMonoid A  E_op E_one E_eq}
                (x:A) : x ^ 1 == x.
Proof. cbn; rewrite Eone_right; reflexivity. Qed.
\end{Coqsrc}

\subsubsection{Examples of computation}

In the following computations, we first show an exponentiation in $\mathds{Z}$, then in
the type of 31-bit machine integers.\footnote{\texttt{phi} and \texttt{phi\_inv} are 
standard library's conversion
functions between types \texttt{Z} and \texttt{int31}, used for making it possible to read  and print values of type \texttt{int31}.}

\vspace{4pt}

From Module~\href{../src/html/addition.power.Demo_power.html}{Powers.Demo\_power}

\begin{Coqsrc}
Open Scope M_scope.

Compute 22%Z ^ 20.
\end{Coqsrc}

\begin{Coqanswer}
= 705429498686404044207947776%Z
\end{Coqanswer}

\begin{Coqsrc}
Import Int31.
Coercion phi_inv : Z >-> int31.

Compute (22%int31 ^ 20).
\end{Coqsrc}

\begin{Coqanswer}
   = 2131755008%int31
     : int31
\end{Coqanswer}

\subsection{The Binary Exponentiation Algorithm}

Please find below the implementation of binary exponentiation using type classes
(to be compared with the version in~\vref{polymorhic-binary_exp}).

% It takes the form of an auxiliary function  \texttt{binary\_power\_mult}
% associated with equalities \ref{binary-eq4} to \ref{binary-eq6} and a main function \texttt{Pos\_bpow} associated with equalities \ref{binary-eq1} to \ref{binary-eq3}.

\vspace{4pt}
\emph{From Module~\href{../src/html/addition.power.Pow.html}{Powers.Pow}}

\begin{Coqsrc}
Fixpoint binary_power_mult `{M: @EMonoid A E_op E_one E_eq}
             (x a:A)(p:positive) : A 
  :=
  match p with
    | xH =>    a * x
    | xO q => binary_power_mult  (x * x) a q
    | xI q =>   binary_power_mult (x * x) (a * x) q
  end.

Fixpoint Pos_bpow  `{M: @EMonoid A E_op E_one E_eq} 
         (x:A)(p:positive) :=
 match p with
  | xH => x
  | xO q => Pos_bpow  (x * x) q
  | xI q => binary_power_mult (x * x) x q
end.
\end{Coqsrc}

It is easy to extend \texttt{Pos\_bpow}'s domain to the type of all 
natural numbers:

\vspace{4pt}
From Module~\href{../src/html/addition.power.Pow.html}{Powers.Pow}

\begin{Coqsrc}
Definition N_bpow {A} `{M: @EMonoid A E_op E_one E_eq} x (n:N) := 
  match n with 
  | 0%N => E_one
  | Npos p => Pos_bpow x p
  end.

Infix "^b" := N_bpow (at level 30, right associativity): M_scope.
\end{Coqsrc}

\subsection{Refinement and Correctness}
We have got two functions for computing powers in any monoid. 
So, it is interesting to ask oneself whether this duplication is useful, and which would be the respective role of \texttt{N\_bpow} and \texttt{power}.

\begin{itemize}
\item The function \texttt{power}, although very inefficient, is a direct 
translation of the mathematical definition, as shown by  lemmas \texttt{power\_eq1} to \linebreak \texttt{power\_eq3}. Moreover, its structural recursion over type \texttt{nat} allows simple proofs by induction over the exponent. 
Thus, we will consider \texttt{power} as a \emph{specification} of any exponentiation algorithm.

\item Functions \texttt{N\_bpow} and \texttt{Pos\_bpow} are more efficient, but less readable than \texttt{power}, and we cannot use these functions before 
having proved their correctness. In fact, the correctness of 
\texttt{N\_bpow} and \texttt{Pos\_bpow} will mean ``being extensionally equivalent to \texttt{power}''.
For instance \texttt{N\_bpow}'s correctness is expressed by the following
statement (in the context of an \texttt{EMonoid} on type \texttt{A}).

\vspace{4pt}
From Module~\href{../src/html/addition.power.Pow.html}{Powers.Pow}

\begin{Coqsrc}
Lemma N_bpow_ok : 
forall (x:A) (n:N),   x ^b n  == x ^ N.to_nat n.
(* Proof omitted *)
\end{Coqsrc}

\end{itemize}

The relationship between \texttt{power} and \texttt{N\_bpow} can be considered
as a kind of \emph{refinement} as in the \texttt{B}-method~\cite{b-book}. Note
that the two representations of natural numbers and the function \texttt{N.to\_nat}
form a kind of  \emph{data refinement} \cite{Abrial:2010:MES:1855020, cohen:hal-01113453}.

\subsection{Proof of correctness of binary exponentiation w.r.t. the function \texttt{power}}
Section \texttt{M\_given} of Module 
~\href{../src/html/addition.power.Pow.html}{Powers.Pow} is devoted to the proof 
of properties of the functions above.
Note that properties of \texttt{power} refer to the \emph{specification} of exponentiation, and can be applied for proving correctness of any implementation.

In this section, we consider an arbitrary instance  \texttt{M} of class \texttt{EMonoid}.

\begin{Coqsrc}
Section M_given.
 Variables (A:Type) (E_op : Mult_op A)(E_one:A) (E_eq : Equiv A).
 Context (M:EMonoid  E_op E_one E_eq).
\end{Coqsrc}

\subsubsection{Properties of exponentiation}
We establish a few well-known properties of exponentiation, and define some basic tactics for simplifying proof search.

\begin{Coqsrc}
Ltac monoid_rw :=
    rewrite Eone_left  ||
    rewrite Eone_right  || 
    rewrite Eop_assoc .

Ltac monoid_simpl := repeat monoid_rw.

Section About_power.
\end{Coqsrc}

In order to make possible proof by rewriting on expressions which contain
the exponentiation operator, we have to prove that, whenever \texttt{$x$ == $y$},
the equality \texttt{$x^n$ == $y^n$} holds for any exponent \texttt{$n$}. 
For this purpose, we use the \texttt{Proper} class of module
\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Classes.Morphisms.html}{Coq.Classes.Morphisms}
\index{Type Classes!Proper}

\begin{Coqsrc}
Global Instance power_proper :
     Proper (equiv ==> eq ==> equiv) power.
(* Proof omitted *)
\end{Coqsrc}

In the following proofs, we note how notations, type classes and generalized 
rewriting can be used  to write algebraic properties in a nice way.

\begin{Coqsrc}
Lemma power_of_plus :   forall x n p, x ^ (n + p) ==  x ^ n *  x ^ p.
(* Proof omitted *)

Ltac power_simpl := 
    repeat (monoid_rw || rewrite <- power_x_plus).
\end{Coqsrc}

  Please note that the following two lemmas \emph{do not require} 
the operation~\texttt{*} to be commutative.

\begin{Coqsrc}
Lemma power_commute : 
    forall x n p, x ^ n * x ^ p ==  x ^ p * x ^ n. 
(* Proof omitted *)

Lemma power_commute_with_x : 
    forall x n,  x * x ^ n == x ^ n * x.
(* Proof omitted *)

Lemma power_of_power : 
   forall x n p,  (x ^ n) ^ p == x ^ (p * n).
(* Proof omitted *)

\end{Coqsrc}

The following two equalities are auxiliary lemmas for proving correctness of the binary exponentiation functions.

\begin{Coqsrc}
Lemma sqr_def : forall x, x ^ 2 ==  x * x.
(* Proof omitted *)

Lemma power_of_square : 
  forall x n, (x * x) ^ n ==  x ^ n * x ^ n.
(* Proof omitted *) 
\end{Coqsrc}

\subsection{Equivalence of the two exponentiation functions}

Since \texttt{binary\_power\_mult} is defined by structural recursion on the
exponent \texttt{p:positive}, its basic properties are proved by induction
along \texttt{positive}'s constructors.

\vspace{4pt}
\emph{From Module~\href{../src/html/addition.power.Pow.html}{Powers.Pow}}

\begin{Coqsrc}
Lemma binary_power_mult_ok :
  forall p a x,   binary_power_mult  x a p  ==  
                  a * x ^ Pos.to_nat p.
Proof.
  induction p as [q IHq | q IHq| ].
 (* Rest of proof omitted *)
 \end{Coqsrc}

 \begin{Coqsrc}
Lemma Pos_bpow_ok : 
  forall p x, Pos_bpow x p == x ^ Pos.to_nat p.
(* Proof omitted *)

Lemma N_bpow_ok : 
  forall n x,  x ^b n  == x ^ N.to_nat n.
(* Proof omitted *)
\end{Coqsrc}

\begin{Coqsrc}
Lemma N_bpow_ok_R : 
  forall n x, x ^b (N.of_nat n)   ==  x ^  n.
(* Proof omitted *)

Lemma Pos_bpow_ok_R : 
   forall p x, p <> 0 ->
                      Pos_bpow x  (Pos.of_nat p)   ==  x ^  p.
(* Proof omitted *)

End About_power.  
\end{Coqsrc}

\subsubsection{Remark}
The preceding lemmas can be applied for deriving properties of the binary exponentiation 
functions:

\begin{Coqsrc}
Lemma N_bpow_commute : forall x n p,  
                        x ^b n *  x ^b p ==  
                        x ^b p *  x ^b n.
Proof.
 intros x n p; repeat rewrite N_bpow_ok.
 rewrite power_commute; reflexivity.
Qed.  
\end{Coqsrc}

\section{Comparing Exponentiation Algorithms with respect to Efficiency}

It looks obvious that  the binary exponentiation algorithm is more efficient than the 
naive one. Can we study \emph{within \coq{}} the respective efficiency of both functions?
Let us take a simple example with the exponent $17$,  in any \texttt{EMonoid}.

\begin{Coqsrc}
Eval simpl in   fun (x:A) => x ^b 17.
\end{Coqsrc}

\begin{Coqanswer}
 = fun x : A =>
       x *
       (x * x * (x * x) * (x * x * (x * x)) *
        (x * x * (x * x) * (x * x * (x * x))))
     : A -> A  
\end{Coqanswer}

Therefore, we note that the term (\Verb|fun (x:A) =>x ^b 17|)  is
convertible, --- \emph{thus logically indistinguishable} ---, with a function that performs 16 multiplications.

Likewise, let us simplify the term (\Verb|fun (x:A) =>x ^ 17|):

\begin{Coqsrc}
Eval simpl in   fun x =>  x ^ 17.  
\end{Coqsrc}

\begin{Coqanswer}
= fun x : A =>
   x * (x *  (x *  (x *  (x *   (x *  (x *  (x *
    (x * (x * (x * (x * (x * (x * (x * (x * (x * one)))))
  )))))))))))  
\end{Coqanswer}

From these tests, we may infer that  representing exponentiation algorithms as \coq{}  functions hides
information about the real structure of the computations, particularly the sharing on intermediate computations.

Thus, we propose to define a data structure that makes explicit the sequence of multiplications that lead to the computation of $x^n$. For instance, the values of  
\texttt{x * x} and
\texttt{x * x * (x * x)}  are used
twice in the  computation of $x^{17}$ with the binary algorithm. This information should 
appear explicitly in the data structure chosen for representing exponentiation 
algorithms.

It is well known that local variables can be used to store intermediate results.
In an \texttt{ISWIM} - \texttt{ML} style, the function computing $x^{17}$ could be written as follows:

\begin{Coqsrc}
Definition pow_17  (x:A) :=
  let x2 := x * x in
  let x4 := x2 * x2 in
  let x8 := x4 * x4 in
  let x16 := x8 * x8 in
  x16 * x.
\end{Coqsrc}
\label{pow-17-let-in}

Unfortunately, \coq's \textbf{let-in} construct is useless for our purpose, since $\zeta$-conversion 
would make the sharing of computations disappear.

\begin{Coqsrc}
Eval cbv  zeta beta delta [pow_17]  in  pow_17.
\end{Coqsrc}

\begin{Coqanswer}
 = fun x : A =>
       x * x * (x * x) * (x * x * (x * x)) *
       (x * x * (x * x) * (x * x * (x * x))) * x
     : A -> A 
\end{Coqanswer}

In the next section, we propose to use a \emph{data structure} for representing 
the computations that lead to the evaluation of some power $x^n$, where
intermediary results are explicitly named for further use in the rest of the computation.

\subsection{Addition chains}
An \emph{addition chain} (In short : \emph{a chain})~\cite{brauer1939} is a representation of a sequence of
intermediate steps that lead to the evaluation of some $x^n$, under the 
assumption that each of these steps is a computation of  a power $x^i$, with 
$i<n$.

In articles from the combinatorist  community, 
\emph{e.g.}~\cite{brauer1939,DBLP:journals/ipl/BerstelB87},  addition chains 
are represented as sequences of positive integers, each member of which 
is either $1$ or  the sum of two previous elements.
For instance, the three following sequences are addition chains for the exponent $87$:

\begin{align}
c_{87} &= (1,2,3,6,7,10,20,40,80,87) \\
c'_{87}&=(1,2,3,4,7,8,16,23,32,64,87) \\
c''_{87}&=(1,2,4,8,16,32,64,80,84,86,87)
\end{align}

It is possible to associate to any addition chain a directed acyclic graph:
whenever $i=j+k$, there is an arc from $x^j$ to $x^i$ and an arc
from $x^k$ to $x^i$. Figures~\ref{fig:chain-87-eucl}  and 
\ref{fig:chain-87-bin} show the graphical representations of 
 $c_{87}$  and $c'_{87}$. 
Please note that some chains may be represented by various different dags.
For instance, we can associate four different dags to the chain $(1,2,3,4,6,9,13)$.

\begin{figure}[h]
  \centering

  \caption{Graphical representation of $c_{87}$ (9 multiplications)}
  \label{fig:chain-87-eucl}
\begin{tikzpicture}
\node (X) at (0,0) {$x$};
\node (X2) at (1,0) {$x^2$};
\node (X3) at (2,0) {$x^3$};
\node (X6) at (3,0) {$x^6$};
\node (X7) at (4,0) {$x^7$};
\node (X10) at (5.5,0) {$x^{10}$};
\node (X20) at (6.5,0) {$x^{20}$};
\node (X40) at (7.5,0) {$x^{40}$};
\node (X80) at (8.5,0) {$x^{80}$};
\node (X87) at (9.5,0) {$x^{87}$};
\draw [->, >=latex](X) -- (X2);
\draw [->, >=latex](X2) -- (X3);
\draw [->, >=latex](X3) -- (X6);
\draw [->, >=latex](X6) -- (X7);
\draw [->, >=latex](X7) -- (X10);
\draw [->, >=latex](X10) -- (X20);
\draw [->, >=latex](X20) -- (X40);
\draw [->, >=latex](X40) -- (X80);
\draw [->, >=latex](X80) -- (X87);
\draw [->, >=latex](X) to [bend left] (X3);
\draw [->, >=latex](X) to [bend left] (X7);
\draw [->, >=latex](X3) to [bend left] (X10);
\draw [->, >=latex](X7) to [bend left] (X87);
\end{tikzpicture}
\end{figure}

\begin{figure}[h]
  \centering

  \caption{Graphical representation of $c'_{87}$ (10 multiplications)}
  \label{fig:chain-87-bin}
\begin{tikzpicture}
\node (X) at (0,0) {$x$};
\node (X2) at (1,0) {$x^2$};
\node (X3) at (2,0) {$x^3$};
\node (X4) at (3,0) {$x^4$};
\node (X7) at (4.5,0) {$x^7$};
\node (X8) at (5.5,0) {$x^8$};
\node (X16) at (6.5,0) {$x^{16}$};
\node (X23) at (8,0) {$x^{23}$};
\node (X32) at (9.5,0) {$x^{32}$};
\node (X64) at (10.5,0) {$x^{64}$};
\node (X87) at (11.5,0) {$x^{87}$};
\draw [->, >=latex](X) -- (X2);
\draw [->, >=latex](X) to [bend left] (X3);
\draw [->, >=latex](X2) to [bend right] (X4);
\draw [->, >=latex](X2) -- (X3);
\draw [->, >=latex](X4) -- (X7);
\draw [->, >=latex](X3) to [bend left] (X7);
\draw [->, >=latex](X4) to [bend right] (X8);
\draw [->, >=latex](X7) to [bend left] (X23);
\draw [->, >=latex](X8) -- (X16);
\draw [->, >=latex](X16) -- (X23);
\draw [->, >=latex](X16) to [bend right] (X32);
\draw [->, >=latex](X32) -- (X64);
\draw [->, >=latex](X64) -- (X87);
\draw [->, >=latex](X23) to [bend left] (X87);
\end{tikzpicture}
\end{figure}

Let us assume that the efficiency of an exponentiation algorithm is proportional
to the number of multiplications it requires. This assumption looks reasonable 
when the data size is bounded (for instance : machine integers, arithmetic modulo $m$, etc.). 
Let us define the \emph{length} of a chain $c$ as its number $|c|$ of exponents
(without counting the initial $1$). 
This length is the number of multiplications needed for 
computing the $x^i$s by applying the following algorithm:

\begin{quote}
For any item $i$ of $c$, there exists $j$ and $k$ in $c$, where
$i=j+k$, and $x^j$ and $x^k$ are already computed.

Thus, compute $x^i = x^j \times x^k$.
\end{quote}

In our little example, we have 
$|c_{87}| = 9 < 10 = |c'_{87}|$. 
In the rest of this chapter, we will try to focus on the following aspects:
\begin{itemize}
\item Define a representation of addition chains that allows to compute
  efficiently $x^n$ in any monoid, for quite large exponents $n$;
\item Certify that our representation of chains is correct, 
    \emph{i.e.}, determines a computation of $x^n$ for a given $n$;
\item Define and certify functions for automatically  generating 
    correct and shortest as possible chains.
\end{itemize}

In a previous work~\cite{DBLP:journals/ita/BrlekCHM95, DBLP:conf/tapsoft/BrlekCS91,AdditionsContrib},  addition chains were represented so as to allow
efficient computations of powers and certification of a family of
automatic chain generators.
  We present here a new implementation that takes into account some
advances in the way we use \coq{}: generalized rewriting, type classes,
parametricity, etc.

\subsection{A type for addition chains}

Let us recall that we want to represent some algorithms of the form
described in section~\ref{pow-17-let-in}, but avoiding to represent
intermediate results by \textbf{let-in}  constructs.
We describe below the main design choices we made:

\begin{itemize}
\item Continuation Passing Style (CPS) \index{Programming Styles!Continuation Passing Style (CPS)} \cite{reynolds93}
is a way to make explicit the 
     control in the evaluation of an expression, in a purely functional way. 
    For every intermediate computation step, the result is sent
    to a \emph{continuation} that executes the further continuations.
   When the continuation is a lambda-abstraction, its bound variable 
   gives a \emph{name} to this result

\item Like in Parametric Higher Order Abstract Syntax (PHOAS)~\cite{PHOAS}, \index{Parametric Higher-Order Abstract Syntax (PHOAS)}
     the local variables associated to intermediate results are
     represented by variables of  type $A$, where $A$ is the underlying type
  of the considered monoid.
\end{itemize}

\subsubsection{Definition}
\label{computation-def}
Let  \texttt{A} be some type;  a \emph{computation} on \texttt{A} is 
\begin{itemize}
\item  either a final step, returning some value of type \texttt{A}
\item or the multiplication of two values of type  \texttt{A}, with a  \emph{continuation}
  that takes as argument the result of this multiplication, then starts a new
  computation.
\end{itemize}

  In the following inductive type definition, the intended meaning 
  of the contruct (\texttt{Mult $x$ $y$ $k$})  is \emph{``multiply \texttt{x} with 
\texttt{y}, then send  the result of this multiplication to 
  the continuation  \texttt{k}''}.

From Module~\href{../src/html/addition.power.Chains.html}{Powers.Chains}

\begin{Coqsrc}
Inductive computation {A:Type}  : Type :=
| Return (a : A)
| Mult (x y : A) (k : A -> computation).    
\end{Coqsrc}
\subsubsection{Monadic Notation}

The following \emph{monadic} 
notation makes terms of type \texttt{computation} look like
expressions of a small programming language dedicated to sequences of mutiplications.
Please look at \emph{CPDT}~\cite{chlipalacpdt2011} for more details on monadic notations in \coq.

\begin{Coqsrc}
Notation "z '<---'  x 'times' y ';' e2 " :=
  (Mult x y  (fun z => e2))
    (right associativity, at level 60).
\end{Coqsrc}

The \texttt{computation} type family is able to express sharing of intermediate computations. For instance, the computation of $2^7$ depicted in Figure~\ref{fig:dag7} is described by  the following term:

\begin{Coqsrc}
Example comp7 : computation  :=
  x <--- 2 times 2;
  y <--- x times 2;
  z <--- y times y ;
  t <--- 2 times z ;
  Return t.  
\end{Coqsrc}

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
  \node (leaf) at (2,0){\texttt{$2$}};
  \node (x2) at (3.5,1){\texttt{$x$}};
  \node (x3) at (3.5,2){\texttt{$y$}};
 \node (x6) at (3.5,4){\texttt{$z$}};
  \node  (root) at (2,5.5) {\texttt{$t$}};
\draw [->, >=latex, bend left] (leaf) to  node[midway, left] {\tiny{$2$}} 
 (root);
  \draw [->, >=latex,bend left] (leaf) to  node[midway, left] {\tiny{$2$}}  
(x2);
  \draw [->, >=latex,bend right] (leaf) to  node[midway, right] {\tiny{$2$}} 
 (x2);
 \draw [->, >=latex,bend left] (leaf) to  node[midway, left] {\tiny{$2$}} 
 (x3);
\draw [->, >=latex,bend right] (x2) to  node[midway, right] {\tiny{$2^2$}} 
 (x3);
\draw [->, >=latex,bend left] (x3) to  node[midway, left] {\tiny{$2^3$}} 
 (x6);
\draw [->, >=latex,bend right] (x3) to  node[midway, right] {\tiny{$2^3$}} 
 (x6);
\draw [->, >=latex, bend right] (x6) to  node[midway, right] {\tiny{$2^6$}} 
 (root);
  \end{tikzpicture}
  \caption{The dag associated to a computation of $2^7$}
  \label{fig:dag7}
\end{figure}

\subsubsection{Definition}
\label{chain-def}

Thanks to the  \texttt{computation} type family, we can associate a type
to the kind of computation schemes described in Figures~\ref{fig:chain-87-eucl} and ~\ref{fig:chain-87-bin}.

We define 
 \emph{addition chains} (in short  \emph{chains}) as functions that map
 any
 type \texttt{A} and any value \texttt{a} of type \texttt{A}  into a computation 
on \texttt{A}:

\begin{Coqsrc}
Definition chain := forall A:Type, A -> @computation A.   
\end{Coqsrc}

Thus, terms of type \texttt{chain} describe polymorphic 
exponentiation algorithms.

For instance, Fig~\vref{fig:C87} shows a definition of the chain  of Figure~\ref{fig:chain-87-eucl}, for the exponent $87$.
Note that, like in PHOAS, bound variables associated with the 
intermediary results are \coq{} variables of type $A$.
\begin{figure}[h]
  \centering
  \begin{Coqsrc}
Example  C87 : chain :=
 fun A (x : A) =>
  x2 <--- x times x ;
  x3 <--- x2 times x ;
  x6 <--- x3 times x3 ;
  x7 <--- x6 times x ;
  x10 <--- x7 times x3 ;
  x20 <--- x10 times x10 ;
  x40 <--- x20 times x20 ;
  x80 <--- x40 times x40 ;  
  x87 <--- x80 times x7 ;
  Return x87.
 \end{Coqsrc}
  \caption{A chain for raising x to its $87$-th power}
  \label{fig:C87}
\end{figure}

The structure of the definition of types \texttt{computation}   and \texttt{chain} suggest that basic definitions over \texttt{chain} will have the following structure:
\begin{itemize}
\item A recursive function on type \texttt{computation $A$} (for a given
    type $A$)
\item A main function on type \texttt{chain} that calls the previous one on 
any \texttt{$A$:Type}.
\end{itemize}

For instance, the following function computes the length of any chain,
\emph{i.e.}, the number of multiplications of the associated computation.
Note that the function \texttt{chain\_length} calls the auxiliary function
\texttt{computation\_length}, with the variable \texttt{A} instantiated to the singleton type  \texttt{unit}.

Any other type in \coq{} would have fitted our needs, but \texttt{unit} and
its unique inhabitant \texttt{tt} was the simplest  solution.

\label{C87-length}
\begin{Coqsrc}
Fixpoint computation_length {A} (a:A)(m : @computation A) 
  : nat :=
match m with
  | Mult _ _ k => S (computation_length a (k a))
  | _ => 0%nat
end.

Definition chain_length (c:chain) 
   := computation_length tt (c _ tt).

Compute chain_length C87.
\end{Coqsrc}
\begin{Coqanswer}
 = 9 : nat  
\end{Coqanswer}

\subsection{Chains as a (small) programming language}

The \texttt{chain} type can be considered as a tiny programming language dedicated to compute powers in any \texttt{EMonoid}. Thus, we have to define a semantics for this language. This semantics is defined in two parts:
\begin{itemize}
\item A structurally recursive function,  --- parameterized with an \texttt{EMonoid} \texttt{M} on a given type \texttt{A} ---, that computes the value associated with any computation on \texttt{M}
\item A polymorphic function that takes as arguments  a  chain \texttt{c},
 a type \texttt{A},  an \texttt{EMonoid} on \texttt{A}, and 
   a value \texttt{x:A},
  then executes the computation \texttt{(c A x)}.
\end{itemize}

\begin{Coqsrc}
Fixpoint computation_execute  {A:Type} (op: Mult_op A) 
                              (c : computation) :=
match c with 
| Return x => x 
| Mult x y k => computation_execute op (k (x * y))
end.

Definition chain_execute (c:chain) {A} op  (a:A) :=
  computation_execute op (c A a).
\end{Coqsrc}

\begin{Coqsrc}
Definition computation_eval `{M:@EMonoid A E_op E_one E_eq}
           (c : computation) : A := computation_execute E_op c.

Definition chain_apply (c:chain) 
    {M:@EMonoid A E_op E_one E_eq} a : A :=
    computation_eval (c A a).
\end{Coqsrc}

\index{Projects}
\begin{project}
Study how  to compile efficiently such data structures.
\end{project}

\subsubsection*{Examples:} 
The following interactions show how to apply the chain \texttt{C87} 
for exponentiation within two different monoids:

\begin{Coqsrc}
Compute  chain_apply C87 3%Z.
\end{Coqsrc}

\begin{Coqanswer}
 =  323257909929174534292273980721360271853387%Z
     : Z
\end{Coqanswer}

\begin{Coqsrc}
Compute chain_apply C87 (M:=M2N) (Build_M2 1 1 1 0)%N.
\end{Coqsrc}

\begin{Coqanswer}
  = {|
      c00 := 1100087778366101931%N;
      c01 := 679891637638612258%N;
      c10 := 679891637638612258%N;
      c11 := 420196140727489673%N |}
     : M2 N
\end{Coqanswer}

\subsubsection{Chain Correctness and Optimality}

A chain is said to be \emph{correct} with respect to a positive
integer \texttt{p} if its execution in any monoid computes $p$-th powers.

\label{chain-correct-def}
\begin{Coqsrc}
Definition chain_correct_nat (c: chain) (n:nat) := 
  n <> 0 /\
  forall `(M:@EMonoid  A E_op E_one E_eq) (x:A), 
      chain_apply c x ==   x ^ n.

Definition chain_correct (c: chain) (p:positive) :=
  chain_correct_nat c (Pos.to_nat p). 
\end{Coqsrc}

\begin{definition}
A chain $c$ is \emph{optimal} for a given exponent $p$ if its length is less 
than or equal to
the length of any chain correct for $p$.  
\end{definition}

 \begin{Coqsrc}
Definition optimal (p:positive) (c : chain) :=
 forall c', chain_correct p c' -> 
            (chain_length c <= chain_length c')%nat.
 \end{Coqsrc}

\section{Proving a chain's correctness}
\label{chain-correctness-sect}
In this section, we present various ways of proving that a given chain is 
correct w.r.t. a given exponent. First, we just try to apply 
the definition in Section~\vref{chain-correct-def}, but this method is very 
inefficient, even for small exponents. In a second step, we use more sophisticated techniques such as reflection and parametricity. Automatic generation of correct chains will be treated in Sect.~\vref{chain-generation}.

\subsection{Proof by rewriting}
Let us show how to prove  the correctness of some chains, using
the \texttt{EMonoid} laws shown in Sect.~\vref{EMonoid-def}.

\begin{Coqsrc}
Ltac slow_chain_correct_tac :=
  match goal with 
      [ |- chain_correct ?c ?p ] =>
      let A := fresh "A" in
      let op := fresh "op" in
      let one := fresh "one" in
      let eqv := fresh "eqv" in
      let M := fresh "M" in
      let x := fresh "x"
      in  split;
        [discriminate | 
         unfold c, chain_apply, computation_eval; simpl;
         intros A op one eq M x; monoid_simpl M; reflexivity]
  end.

Example C7_ok : chain_correct C7 7.
Proof.
   slow_chain_correct_tac.
Qed.
\end{Coqsrc}

Unfortunately, this approach is terribly inefficient, even for quite small exponents:

\begin{Coqsrc}
Example C87_ok : chain_correct C87 87.
Proof.
 Time  slow_chain_correct_tac. 
\end{Coqsrc}

\begin{Coqanswer}
 Finished transaction in 62.808 secs (62.677u,0.085s) (successful)   
 \end{Coqanswer}

\begin{Coqsrc}
Qed.
\end{Coqsrc}

In addition to this big computation time, this approach 
generates a huge proof term. Just try to execute the command 
``\texttt{Print C87\_ok}'' to get a measure of its size.
In order to understand this poor performance, let us consider an intermediate
subgoal of the previous proof generated after a sequence of unfoldings and simplifications. This goal is presented below.

\begin{Coqanswer}
1 subgoal, subgoal 1 (ID 219)

  A : Type
  E_op : Mult_op A
  E_one : A
  E_eq : Equiv A
  M : EMonoid E_op E_one E_eq
  x : A
  ============================
   x * x * x * (x * x * x) * x * (x * x * x) *
   (x * x * x * (x * x * x) * x * (x * x * x)) *
   (x * x * x * (x * x * x) * x * (x * x * x) *
    (x * x * x * (x * x * x) * x * (x * x * x))) *
   (x * x * x * (x * x * x) * x * (x * x * x) *
    (x * x * x * (x * x * x) * x * (x * x * x)) *
    (x * x * x * (x * x * x) * x * (x * x * x) *
     (x * x * x * (x * x * x) * x * (x * x * x)))) *
   (x * x * x * (x * x * x) * x) ==
   x *
   (x *
    (x *
     (x *
      (x *
       (x *
        (x *
         (x *
          (x *
           (x *
            (x *
             (x *
              (x *
               (x * (x * (x * (x * (x * (x * (x * (x * (x * (x * 
                ..))))))))))))))))))))))
\end{Coqanswer}
\label{fig:big-goal}

This goal is solved by the following tactic composition:

\begin{Coqsrc}
monoid_simpl M; reflexivity. 
\end{Coqsrc}

This inefficiency certainly comes from the cost of setoid rewriting.
At every application of an \texttt{EMonoid} law, the system must
verify that the context of this rewriting is compatible  with the equivalence
relation associated with the current \texttt{EMonoid}.
The rest of this chapter is devoted to the  presentation of more efficient 
 methods for proving chain correctness.

\subsection{Correctness Proofs by Reflection}
\label{reflection-section}
\index{Advanced proof techniques!Proof by reflection}
Instead of letting the tactic \texttt{rewrite} look for contexts in which
setoid rewriting is possible, we propose to use (deterministic) computations for
obtaining a ``canonical'' form for terms generated from a variable \texttt{x}
by contructors associated with monoid multiplication and neutral element.

The reader will find general explanations about proofs by reflection in \coq{},
for instance in Chapter 16 of Coq'Art\cite{BC04} and the numerous examples (including the \texttt{ring} tactic) 
in \coq's reference manual.

\subsubsection{How does reflection work}
Let us consider again the subgoal on page~\pageref{fig:big-goal}, the conclusion of which has the form \texttt{|$a_1\,==\,a_2$|}, where \texttt{|$a_1$|} and
\texttt{|$a_2$|} are terms of  type \texttt{A}.
Instead of spending space and time in setoid rewritings, we would like to
normalize the terms \texttt{|$a_1$|} and \texttt{|$a_2$|} and verify that 
the associated normal forms are equal.

Defining such a normalization function is possible on an inductive type.
The following type describes expressions composed of monoid operations and inhabitants of a given type $A$.

\begin{Coqsrc}
(** Binary trees of multiplications over A *)

Inductive Monoid_Exp (A:Type) : Type :=
 Mul_node (t t' : Monoid_Exp A) | One_node | A_node (a:A).

Arguments Mul_node {A} _ _.
Arguments One_node {A} .
Arguments A_node {A} _ .
\end{Coqsrc}

Thus, the main steps of a correctness proof of a given chain, \emph{e.g.}
\texttt{C87} will be the following ones:
\begin{enumerate}
\item generate a subgoal as in page~\pageref{fig:big-goal},
\item express each term of the equivalence as the image of a term
     of type \texttt{Monoid\_Exp $A$},
\item normalize both terms and verify that their normal forms are equal.
\end{enumerate}

The rest of this section is devoted to the definition of the normalization 
function on \texttt{Monoid\_Exp $A$}, and the proofs of lemmas that
link equivalence on type \texttt{A} and equality of normal forms
of terms of type \texttt{Monoid\_Exp $A$}.

\subsubsection{Linearization function}

The following functions help to transform any term of type
\texttt{Monoid\_Exp $A$} into a flat ``normal form''.

\begin{Coqsrc}

Fixpoint flatten_aux {A:Type} (t fin : Monoid_Exp A) 
  : Monoid_Exp A :=
match t with Mul_node  t t' =>
              flatten_aux t (flatten_aux t' fin)
           | One_node  => fin
           |  x => Mul_node  x fin
end.

Fixpoint flatten {A:Type} (t: Monoid_Exp A) : Monoid_Exp A :=
match t with
| Mul_node t t' => flatten_aux t (flatten t')
| One_node => One_node
| X => Mul_node X One_node
end.
\end{Coqsrc}

\subsubsection{Interpretation function}

The function \texttt{eval} maps any term of type \texttt{Monoid\_Exp $A$}
into a term of type \texttt{$A$}.

\begin{Coqsrc}
Function eval {A:Type} {op one eqv}
         (M: @EMonoid A op one eqv)
         (t: Monoid_Exp A) : A :=
 match t with 
            | Mul_node t1 t2 => (eval M t1 * eval M t2)%M
            | One_node => one
            | A_node a => a
end.
\end{Coqsrc}

The following two lemmas relate the linearization function \texttt{flatten}
with the interpretation function \texttt{eval}.

\begin{Coqsrc}
Lemma flatten_valid {A} `(M: @EMonoid A op one eqv):
forall t , eval M t == eval M (flatten t).
(* Proof omitted *)

Lemma flatten_valid_2 {A} `(M: @EMonoid A op one eqv):
forall t t' , eval  M (flatten t) == eval M (flatten t')  ->
     eval M t == eval M t'.
(* Proof omitted *)
\end{Coqsrc}

\subsubsection{Transforming a multiplication into a tree}
Let us now build a tool for building terms of type  \texttt{Monoid\_Exp $A$} out
of terms of type \texttt{A} containing multiplications of the form 
\Verb|(_ * _)%M| and the variable \texttt{one}. 
In fact, what we want to  define is an inverse of the function \texttt{flatten}.

Since \texttt{mult\_op} is not a constructor (see Sect.~\ref{op-classes}), 
the transformation of  
a product of type \texttt{A} into a term of type \texttt{Monoid\_Exp A}
is done with the help of a tactic:

\begin{Coqsrc}
(** "Quote" tactic *)

Ltac model A  op one v :=
match v with 
| (?x  * ?y)%M => let r1 := model A op one x
                  with r2 := model A op one y 
                  in  constr:(@Mul_node A r1 r2)
| one => constr:(@One_node A)
| ?x => constr:(@A_node A x)
end.
\end{Coqsrc}

For instance, the term \texttt{(x * x * x * (x * x * x) * x)} is
transformed by \texttt{model} in the following term of type \texttt{Monoid\_Exp $A$}

\begin{Coqsrc}
(eval M
   (Mul_node
     (Mul_node 
        (Mul_node (Mul_node (A_node x) (A_node x)) (A_node x))
        (Mul_node (Mul_node (A_node x) (A_node x)) (A_node x))) 
     (A_node x)))  
\end{Coqsrc}

\subsection{reflection tactic}
The tactic \texttt{monoid\_eq\_A} converts a goal of the form 
(\texttt{E\_eq $X$ $Y$}), where
\texttt{$X$} and \texttt{$Y$} are terms of type $A$, into
(\texttt{E\_eq (eval M  (model X)) (eval M  (model Y))}). This last goal is intended to be solved thanks 
to the lemma \texttt{flatten\_valid\_2}.

\begin{Coqsrc}
Ltac monoid_eq_A A op one E_eq M  :=
match goal with 
| [ |- E_eq  ?X ?Y ] =>
  let tX := model A op one X with
      tY := model A op one Y in
      (change (E_eq (eval M tX) (eval M tY)))
end.
\end{Coqsrc}

\subsubsection{Main reflection tactic}

The tactic \texttt{reflection\_correct\_tac} tries to prove a chain's 
correctness by a comparison of two terms of type \texttt{Monoid\_Exp $A$}:
one being obtained from the chain's definition, the other one by expansion
of the naive exponentiation definition.

\begin{Coqsrc}
Ltac reflection_correct_tac :=
match goal with
[ |- chain_correct ?c ?n ] =>
 split; [try discriminate |
         let A := fresh "A"
         in let op := fresh "op"
         in let one := fresh "one" 
         in let E_eq := fresh "eq" 
         in let M := fresh "M"
         in let x := fresh "x" 
         in  (try unfold c); unfold chain_apply;
           simpl; red; intros  A op one E_eq M x;
           unfold computation_eval;simpl;
           monoid_eq_A A op one E_eq M;
           apply flatten_valid_2;try reflexivity
        ]
end. 
\end{Coqsrc}

\subsubsection{Example}
The following dialogue clearly shows the efficiency gain over naive setoid rewriting.

\begin{Coqsrc}
Example C87_ok : chain_correct C87 87.
Proof.
  Time reflection_correct_tac.
\end{Coqsrc}

\begin{Coqanswer}
 Finished transaction in 0.038 secs (0.038u,0.s) (successful)
\end{Coqanswer}

\begin{Coqsrc}
Qed. 
\end{Coqsrc}

This tactic is not adapted to much bigger exponents. In \linebreak
 Module~\href{../src/html/addition.power.Euclidean_Chains.html}{Euclidean\_Chains},
 for instance, we tried to apply this tactic for proving the correctness 
of a chain associated with the exponent $45319$. 
 We had to interrupt the prover, which 
was trying to build a linear tree of $2\times  45319 + 1$ nodes!
Indeed, using \texttt{reflection\_correct\_tac} is like doing a 
symbolic evaluation of an inefficient (linear) exponentiation algorithm.

In the next section, we present a solution that avoids doing such a lot of computations.

\subsection{Chain correctness for ---practically --- free!}
\label{correctness-for-free}

% Let us consider again the chain \texttt{C87} of Fig.~\vref{fig:C87}.
% Every bound variable of type \texttt{A} is either the argument \texttt{x}
% or a variable introduced by the abstraction corresponding to the
% continuation argument of constructor \texttt{Mult} (hidden by the monadic notation). Thus, it seems obvious that during the execution of some  computation
% \texttt{C87 A a}, each of this variable will be bound to some power of 
% \texttt{a}.

% Thus, we would like to prove that  every chain \texttt{c} has this property,
% which would be a great step for proving any chains's correctness.

\subsubsection{About parametricity}
\index{Advanced proof techniques!Parametricity}
Let us now present another tactic for proving chain correctness,
in the tradition of works on \emph{parametricity} and its use for 
proving properties on programs.
Strachey~\cite{Strachey:2000:FCP:609150.609208}
explores the nature of \emph{parametric
polymorphism}: ``\emph{Polymorphic functions behave uniformly for all types}''
then Reynolds~\cite{REYNOLDS83} formalizes this notion through binary relations.
Wadler~\cite{Wadler1989}, then Cohen \emph{et al.}~\cite{Cohen2013}
use this relation for deriving
 theorems about functions that operate on parametric
polymorphic types.

Let us look again at the definitions of type family \texttt{computation}
and the type \texttt{chain}:

\begin{Coqsrc}
Inductive computation {A:Type}  : Type :=
| Return (a : A)
| Mult (x y : A) (k : A -> computation).

Definition chain := forall A:Type, A -> @computation A.
\end{Coqsrc}

Let $c$ be a closed term of type 
\texttt{chain}; $c$ is  of the form \linebreak
\texttt{fun (A:Type)(a:A) => $t_a$}, where $t_a$ is a term of type
\texttt{@computation A}.
\label{obvious-remark}
Obviously,  in every subterm of {$t_a$} of type \texttt{A}, 
the two first arguments of constructor \texttt{Mult} or the
argument of \texttt{Return} are either \texttt{a} or a variable 
introduced as the formal argument of a continuation \texttt{k}.
In effect, there is no other way to build terms of type \texttt{A} in the considered context.

Marc Lasson's \textbf{paramcoq} plug-in~(available as  \texttt{opam} package 
\texttt{coq-paramcoq}) generates  a family of binary relations definitions
from \texttt{computation}'s definition.

\begin{Coqanswer}
Inductive
computation_R (A B : Type) (R : A -> B -> Type)
  : computation -> computation -> Type :=
 |    computation_R_Return_R : 
       forall (a1 : A) (a2 : B),  R a1 a2 ->
         computation_R A B R (Return a1) (Return a2)
  | computation_R_Mult_R : forall (x1 : A) (x2 : B),
         R x1 x2 ->
          forall (y1 : A) (y2 : B),
            R y1 y2 ->
            forall (k1 : A -> computation)
                   (k2 : B -> computation),
                      (forall (H : A) (H0 : B),
                          R H H0 -> 
                          computation_R A B R (k1 H) (k2 H0)) ->
                       computation_R A  B R 
                          (z <--- x1 times y1; k1 z)
                          (z <--- x2 times y2; k2 z)
\end{Coqanswer}

Let $A$ and $B$  be two types, and $R: A \arrow B \arrow \typesort$ 
a relation.
Two computations \texttt{cA: @computation A} and \texttt{cB: @computation B}
are related \emph{w.r.t.} \texttt{computation\_R} if every pair of 
arguments of \texttt{Mult} and \texttt{Return} at the same position 
are related \emph{w.r.t.} \texttt{R}.

\subsubsection{Definition}
A chain $c$ is \emph{parametric} if it has the same behaviour for any
pair of types $A$  and $B$, any relation $R$
between  $A$ and $B$ and any $R$-related pair of 
arguments $a$ and $b$:

\begin{Coqsrc}
Definition parametric (c:chain) :=
  forall A B (R: A -> B -> Type) (a:A) (b:B),
   R a b -> computation_R  R (c A a) (c B b).
\end{Coqsrc}

\subsubsection{How to use these definitions?}
Let us use parametricity for proving easily 
a given chain's correctness.
In other words, 
let $c$ be a chain and \texttt{$p$:positive} be a given exponent.
Consider some instance of \texttt{EMonoid} over a type $A$.
We want to prove that the application of the chain $c$ to 
any value $a$ of type $A$ returns the value \texttt{$a^p$}.

We first use \coq's computation facilities for ``guessing'' the exponent associated with any given chain. It suffices to instantiate ``monoid multiplication'' with addition on positive integers.

\begin{Coqsrc}
Definition the_exponent_nat (c:chain) : nat :=
 chain_apply c (M:=Natplus) 1%nat.

Definition the_exponent (c:chain) : positive :=
  chain_execute c Pos.add  1%positive.

Compute the_exponent C87.
\end{Coqsrc}

\begin{Coqanswer}
 = 87%positive
     : positive  
\end{Coqanswer}

We show how to \emph{prove} that  a given  chain $c$,
applied to any $a$, really computes $a^p$, where $p=\textrm{the\_exponent}\;c$.
Parametricity allows us to compare executions on any monoid $M$ 
with executions on \texttt{NatPlus}.
Let us consider the following mathematical relation 
$$\{(x,n)\in M\times\mathbb{N}\,|\, 0<n \wedge x=a^n\}$$

\begin{Coqsrc}
Definition power_R  (a:A) :=
  fun (x:A)(n:nat) => n <> 0 /\ x == a ^ n.  
\end{Coqsrc}

First, we prove the following lemma, that relates \texttt{computation\_R}
with the result  of the  executions of the corresponding computations:

\begin{Coqsrc}
Lemma  power_R_is_a_refinemnt (a:A) :
  forall(gamma : @computation A)
        (gamma_nat : @computation nat),
    computation_R  (power_R a) gamma gamma_nat -> 
     power_R a (computation_eval gamma)
               (computation_eval (M:= Natplus) gamma_nat).    
(* Proof omitted *)
\end{Coqsrc}

Thus, if \texttt{$c$:chain} is parametric, this refinement lemma allows us
to prove a correctness result:

\begin{Coqsrc}
Lemma param_correctness_nat :
 forall c:chain, parametric c ->  
               chain_correct_nat c  (the_exponent_nat c).
(* Proof omitted *)
\end{Coqsrc}

A similar result can be proven with the exponent in \texttt{positive}.
First we instantiate the parameter \texttt{R} of \texttt{computation\_R},
with the relation that links the representations of natural numbers
on respective types \texttt{nat} and \texttt{positive}.
Then we use our lemmas for rewriting under the assumption that the
considered chain is parametric. Please note how our approach is related with
\emph{data refinement} (see also~\cite{Cohen2013}).
The reader may also consult a survey by D. Brown on the most important contributions to 
the notion of parametricity~\cite{DanBrown-survey}.

\begin{Coqsrc}
Lemma exponent_pos2nat : forall c: chain,  parametric c -> 
  the_exponent_nat c = Pos.to_nat (the_exponent c).

Lemma exponent_pos_of_nat : forall c: chain,  parametric c -> 
  the_exponent c = Pos.of_nat (the_exponent_nat c).

Lemma param_correctness (c:chain) :
   parametric c -> 
   chain_correct c (the_exponent c). 
Proof.
  intros; rewrite  exponent_pos_of_nat; auto.
  red;  rewrite  exponent_pos2nat;auto.
  rewrite Pos2Nat.id,  <- exponent_pos2nat;auto.
  apply param_correctness_nat; auto.
Qed.
\end{Coqsrc}

Lemma \texttt{param\_correctness} suggests us a method for verifying 
that a given chain $c$ is correct \emph{w.r.t.} some positive exponent $p$:

\begin{enumerate}
\item Verify that $c$ is parametric.
\item Verify that $p$ is equal to (\texttt{the\_exponent $c$}).
\end{enumerate}

\subsubsection{How to prove a chain's parametricity}
Despite the apparent complexity of \texttt{computation\_R}'s definition,
it is very simple to prove that a given chain is parametric. The following tactics
proceed as follows:

\begin{enumerate}
\item Given a chain $c$, consider two types \texttt{A} and
\texttt{B}, and any relation \texttt{R:A->B->Prop}, 
\item Push into the context declarations of \texttt{a:A}, \texttt{b:B}
and an hypothesis assuming \texttt{R a b}.
\item Then the tactic crosses in parallel the terms (\texttt{c A a}) and
(\texttt{c B b}) (of the same structure),
\begin{itemize}
\item On a pair of terms of the form 
\texttt{Mult xA yA (fun zA => tA)} and \linebreak \texttt{Mult xB yB (fun zB => tB)}, the tactic checks whether 
   \texttt{R xA xB} and \texttt{R yA yB} are already assumed in the context,
 then  pushes into the context the declaration of \texttt{zA} and \texttt{zB}
and the hypothesis \linebreak \texttt{Hz: R zA zB}, then crosses the terms \texttt{tA} and
 \texttt{tB}
\item On a pair of terms  of the form   (\texttt{Return xA}) and (\texttt{Return xB}),
 the tactic just checks whether (\texttt{R xA xB}) is assumed.
\end{itemize}

\end{enumerate}

The tactic itself is simpler than its explanation.

\begin{Coqsrc}
Ltac parametric_tac  := 
match goal with [ |- parametric ?c] =>
   red ; intros;
   repeat (right;[assumption | assumption | ]);  
   left; assumption
end.

Example P87 : parametric C87.
Proof. Time parametric_tac. 
\end{Coqsrc}

\begin{Coqanswer}
Finished transaction in 0.005 secs (0.005u,0.s) (successful)
\end{Coqanswer}

\begin{Coqsrc}
Qed. 
\end{Coqsrc}

\subsubsection{Proving a chain's correctness}
\label{C87-param-ok}
Finally, for proving that a given chain $c$ is correct with respect to an exponent $p$, it suffices to check that $c$ is parametric, and
to apply the lemma \texttt{param\_correctness}. 
The reader will note how this computation-less method is much more efficient
than our reflection tactic.

\begin{Coqsrc}
Ltac param_chain_correct :=
match goal with 
[|- chain_correct ?c ?p ] => 
apply param_correctness; parametric_tac
end.

Lemma C87_ok' : chain_correct C87 87.
Time param_chain_correct.
\end{Coqsrc}

\begin{Coqanswer}
Finished transaction in 0.005 secs (0.005u,0.s) (successful)
\end{Coqanswer}

\begin{Coqsrc}
Qed.
\end{Coqsrc}

\subsubsection{Remark}
For the reasons exposed in Section~\vref{obvious-remark}, 
it seems obvious that any well-written chain is parametric.
Unfortunately, we cannot prove this property  in \coq{},
for instance by induction on \texttt{c}, 
since \texttt{chain} is a product type and not an inductive type.

\begin{Coqbad}
Definition any_chain_parametric : Type :=
 forall c:chain, parametric c.

Goal any_chain_parametric.
Proof.
intros c A B R a b ; induction c.
\end{Coqbad}

\begin{Coqanswer}
 2 subgoals, subgoal 1 (ID 556)

  c : chain
  A : Type
  B : Type
  R : A -> B -> Type
  a : A
  b : B
  a0 : A
  ============================
   R a b -> computation_R R (Return a0) (c B b)

...
\end{Coqanswer}

\begin{Coqbad}
Abort.
\end{Coqbad}

Given this situation, we could  admit (as an axiom) that 
any chain is parametric. Nevertheless, if a chain is under the form of a 
closed term, using \texttt{parametric\_tac} is so efficient than we prefer to 
 avoid
a shameful introduction of an axiom in our development.

\section{Certified Chain Generators}
\label{chain-generation}

In this section, we are interested in the \emph{correct by construction} paradigm.
We just want to give a positive exponent to \coq{} and get a (hopefully)  correct and  efficient chain for this exponent.

We first define the notion of \emph{chain generator}, then present a certified generator that simulates the binary exponentiation algorithm. Last, we present a better chain generator based on integer division.

\subsection{Definitions}

We call \emph{chain generator} any function that takes as argument 
any positive integer and returns a chain.

\begin{Coqsrc}
Definition chain_generator := positive -> chain.  
\end{Coqsrc}

A generator $g$  is \emph{correct} it it returns a correct chain
for any exponent:

\begin{Coqsrc}
Definition correct_generator (g : positive -> chain) :=
 forall p, chain_correct p (g p).
\end{Coqsrc}

Correct generators can be used for computing powers 
on the fly, thanks to the following functions:

\begin{Coqsrc}
Definition cpower_pos (g : chain_generator)  p
           `{M:@EMonoid A E_op E_one E_eq} a :=
  chain_apply (g p) (M:=M) a.

Definition cpower (g : chain_generator)  n
           `{M:@EMonoid A E_op E_one E_eq} a :=
  match n with 0%N => E_one 
             | Npos p => cpower_pos  g p a
  end.  
\end{Coqsrc}

Note also that the use of chain generators is independent from  the techniques presented in Sect.~\ref{chain-correctness-sect}:
Designing an efficient and correct chain generator may be a long and hard task.
On the other hand, once a generator is certified, we are assured of the correctness of  
all its outputs.
Finally, we say that a generator $g$ is \emph{optimal} if it returns chains whose length are less than or
equal to any chain returned by any correct generator:

\begin{Coqsrc}

Definition optimal_generator (g : positive -> chain) :=
 forall p:positive, optimal p (g p).
\end{Coqsrc}

\subsection{The binary chain generator}

Let us reinterpret the  binary exponentiation algorithms in the framework 
of addition chains.
Instead of directly computing $x^n$ for some base $x$ and exponent $n$,
we build chains that describe the computations associated with the binary exponentiation method.
Not surprisingly, this chain generation will be described in terms of recursive
functions, once the underlying monoid is fixed.

As for the ``classical'' binary exponentiation algorithm,
we define an auxiliary computation generator for  the
product of an accumulator $a$ with an arbitrary power of some value $x$.
Then, the main function builds a computation for any positive exponent:

\begin{Coqsrc}
Fixpoint axp_scheme  {A} p : A -> A -> @computation A   :=
 match p with
   | xH =>  (fun a x => y <--- a  times x ; Return y)
   | xO q => (fun a x => x2 <--- x times  x ; axp_scheme q a x2)
   | xI q => (fun a x => ax <--- a times x ;
                         x2 <--- x times x ;
                         axp_scheme q ax x2)
end.

Fixpoint  bin_pow_scheme {A} (p:positive)  
: A -> @computation A:=
  match p with 
  |  xH => fun x => Return x
  | xI q  => fun x => x2 <--- x times x; axp_scheme q x x2
  | xO q => fun x => x2 <--- x times x ; bin_pow_scheme q x2
  end.

\end{Coqsrc}

The following function associates  a chain to any positive exponent:

\begin{Coqsrc}
Definition binary_chain (p:positive) : chain :=
  fun A => bin_pow_scheme p.

Compute binary_chain 87.
\end{Coqsrc}

\begin{Coqanswer}
    = fun (A : Type) (x : A) =>
       x0 <--- x times x;
       x1 <--- x times x0;
       x2 <--- x0 times x0;
       x3 <--- x1 times x2;
       x4 <--- x2 times x2;
       x5 <--- x4 times x4;
       x6 <--- x3 times x5;
       x7 <--- x5 times x5;
       x8 <--- x7 times x7; 
       x9 <--- x6 times x8; 
       Return x9
     : chain
\end{Coqanswer}

\subsubsection{Proof of \texttt{binary\_chain}'s correctness}

Let us now prove that \texttt{binary\_chain} always returns correct chains.
First, due to the structure of this generator's definition, we study the
properties of the auxiliary functions that operate \emph{on a given monoid $M$}.

\begin{Coqsrc}
Section binary_power_proof.

Variables (A: Type)
         (E_op : Mult_op A)
         (E_one : A)
         (E_eq: Equiv A).

Context (M : EMonoid  E_op E_one E_eq).

Existing Instance Eop_proper.
\end{Coqsrc}

\begin{Coqsrc}
Lemma axp_correct : forall p a x,
 computation_eval (axp_scheme p a x) == a *  x ^ (Pos.to_nat p).
(*  Proof by induction on p *)

Lemma binary_correct : 
  forall p x,
        computation_eval  (bin_pow_scheme p (A:=A) x) ==
        x ^ (Pos.to_nat p).
(* Proof by induction on p *)

End  binary_power_proof.
\end{Coqsrc}

\begin{Coqsrc}
Lemma binary_generator_correct : correct_generator binary_chain.
Proof.
  red;unfold chain_correct,  binary_chain, chain_apply;
  split; [auto| intros A op one Eq M x; apply binary_correct].
Qed.
\end{Coqsrc}

\subsubsection{The binary method is not optimal}

It is easy to prove by contradiction  that the binary method is not the most efficient 
for computing powers. 
    First, let us  assume that \texttt{binary\_chain} is optimal:

\begin{Coqsrc}
Section non_optimality_proof.

 Hypothesis binary_opt : optimal binary_chain.
\end{Coqsrc}

Then, let us consider for instance the binary chain generated for the exponent 87.

\begin{Coqsrc}
Compute chain_length (binary_chain 87).
\end{Coqsrc}

\begin{Coqanswer}
   = 10 : nat 
\end{Coqanswer}

Let us recall that \texttt{C87}'s length has been evaluated to $9$ (Sect~\ref{C87-length},
and that this chain is correct (Sect~\vref{C87-param-ok}). Thus, it is very easy
to finish our proof:

\begin{Coqsrc}
Lemma binary_generator_not_optimal : False.
Proof.
  generalize (binary_opt gen _  _ C87_ok); 
  compute; omega.
Qed.

End non_optimality_proof.
\end{Coqsrc}

\index{Exercises}
\begin{exercise}
Prove that for any positive integer $p$,  the length of any optimal chain 
for $p$ is less  than twice the number of digits of the binary representation of $p$.
\end{exercise}

\section{Euclidean Chains}
\label{euclide-sect}

In this section, we present an efficient chain generator. The chains built by this generator
are never longer than the chains built by the binary generator. Moreover, for an 
infinite number of exponents, the chains it builds are strictly shorter than the chain
returned by \texttt{binary\_chain}. 
Euclidean chains are based on the following idea: 
\begin{quote}
For generating a chain that computes $x^n$, one may choose some natural number
$0<p<n$, and build a chain that computes first $x^p$ \textbf{then} uses this value
for computing $x^n$. 
\end{quote}

For instance, a  computation of $x^{42}$ can be decomposed into a computation 
of $y=x^3$, then a computation of $y^{14}$. The efficiency of the chain built with this
methods depends heavily on the choice of $p$. See~\cite{DBLP:journals/ita/BrlekCHM95} for details.

Considering chain generators and their correctness, we may consider the dual of 
decomposition of exponents: we would like to write \emph{composable} correct 
chain generators. For instance, we want to build some object that, ``composed''  
with any correct chain for $n$, returns a correct chain for $3n$.

\paragraph{Note:}
All the \coq{} material described in this section is available on 
 Module~\href{../src/html/addition.power.Euclidean_Chains.html}{Powers/Euclidean\_Chains.v}

\subsection{Chains and Continuations : f-chains}

Please consider the following small example:

\begin{Coqsrc}
Example C3 : chain :=
 fun A (x:A) =>
  x2 <--- x times x;
  x3 <--- x2 times x ;
  Return x3.
\end{Coqsrc}

The execution of this chain on  some value $x:A$ stops after 
computing \texttt{$x^3$}, because of the \texttt{Return} ``statement''.
However, we would like to compose the instructions of \texttt{C3} 
with a chain for another exponent $n$, in order to generate a chain for 
the exponent $3\times n$.

Since \texttt{computation} is an inductive family of types, 
it could be possible to define a composition operator that 
works like list appending (replacing the \texttt{Return y} of the first
computation with the second computation).
\emph{This approach is left as an exercise.}  The solution we present is based on functional programming and the concept of continuation.

\index{Exercises}
\begin{exercise}
Develop the approach suggested in the previous paragraph.  
\end{exercise}

\subsubsection{Type definition of  f-chains}

Let us   consider \emph{incomplete} or \emph{open} chains.
Such an object waits for another chain to resume  a computation.

Figure~\ref{fig:F3-as-dag} represents an  f-chain associated with the exponent $3$, as a dag with an input and one output the edges of which are depicted as thick arrows.

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
  \node (entree) at (0,0) {};
  \node (X) at (1,0) {$x$};
\node (X2) at (2,0) {$x^2$};
  \node (X3) at (3,0) {$x^3$};
\node (sortie) at (4,0) {};
\draw [>-,   thick](entree) -- (X);
\draw [->, >=latex, ](X) -- (X2);
\draw [->, >=latex, ](X2) -- (X3);
\draw [->, >=latex,  thick](X3) -- (sortie);
\draw [->, >=latex](X) to [bend left] (X3);
\draw[dashed] (0.5,-0.3) rectangle (3.5,0.6);
  \end{tikzpicture}
  \caption{Graphical representation of \texttt{F3}}
  \label{fig:F3-as-dag}
\end{figure}

In other words, this kind of objects can be considered as \emph{functions}
from chains to chains. So, we called their type \texttt{Fchain}.

First, we define a type of \emph{continuations},
\emph{i.e.},  functions  that wait for some value $x$, then 
build  a computation for raising {$x$} to some  given exponent.

\begin{Coqsrc}
Definition Fkont (A:Type) := A -> @computation A.
\end{Coqsrc}

An \texttt{f-chain} is just a polymorphic function that combines  a 
continuation and en element into a computation:

\begin{Coqsrc}
Definition Fchain := forall A, Fkont A -> A -> @computation A.
\end{Coqsrc}

\subsubsection{Examples}

Let us define a chain for computing the cube of some $x$, then sending 
the result to a continuation $k$.

\begin{Coqsrc}
Definition F3 : Fchain := 
 fun  A k  (x:A) =>
  y <--- x times x ;
  z <--- y times x ;
  k  z.
\end{Coqsrc}

Any f-chain can be converted into a chain by the help of the following function:

\begin{Coqsrc}
Definition F2C (f : Fchain) : chain :=
 fun (A:Type) => f A Return.

Compute the_exponent (F2C F3).
\end{Coqsrc}

\begin{Coqanswer}
= 3%nat  
\end{Coqanswer}

In the rest of this chapter, we will use two other f-chains, respectively associated with the exponents $1$ and $2$. Chains \texttt{F1}, \texttt{F2} and
\texttt{F3} will form a basis to generate  chains for many exponents
by \emph{composition of correct functions}.

\begin{Coqsrc}
Definition F1 : Fchain := 
 fun A k (x:A) => k x.

Definition F2 : Fchain := 
fun  A k  (x:A) =>
  y <--- x times x ;
  k  y.
\end{Coqsrc}

\subsubsection{F-chain application and composition}

The following definition allows us to consider any value {$f$} 
of type 
\texttt{Fchain} as a function of type \texttt{chain $\arrow$ chain}.

\begin{Coqsrc}
Definition Fapply (f : Fchain) (c: chain) : chain  :=
 fun A  x =>  f  A (fun y => c A  y) x.
\end{Coqsrc}

In a similar way, \emph{composition} of \texttt{f-chain}s is easily defined
(see Figure~\vref{fig:Fcompose}).

\begin{Coqsrc}
Definition Fcompose (f1 f2: Fchain) : Fchain  :=
 fun   A k x =>  f1  A (fun y => f2 A k y) x.

Lemma F1_neutral_l : forall f, Fcompose F1 f = f.
Proof. reflexivity. Qed.

Lemma F1_neutral_r : forall f, Fcompose f F1 = f.
Proof. reflexivity. Qed.
\end{Coqsrc}

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
 \node (input1)  at (0,0) {};  
 \node [draw] (F1)  at (1,0) {$f_1$};  
 \node (output1)  at (2.0,0) {};  
 \node (input2)  at (2.6,0) {};  
 \node [draw] (F2)  at (3.5,0) {$f_2$};  
 \node (output2)  at (4.5,0) {};  
 \draw [>-,   thick](input1) -- (F1);
 \draw [->, >=latex, thick](F1) -- (output1);
\draw [dotted, ](output1) -- (input2);
\draw [>-,   thick](input2) -- (F2);
\draw [->, >=latex, thick](F2) -- (output2);
\draw[dashed] (0.4,-1) rectangle (4,1);
  \end{tikzpicture}
  \caption{Composition  of f-chains $f_1$ and $f_2$ (\texttt{Fcompose})}
  \label{fig:Fcompose}
\end{figure}
\subsubsection{Examples}

The following examples show that the apparent complexity of the previous 
definition is counterbalanced with the simplicity of using \texttt{Fapply}
and \texttt{Fcompose}.

\begin{Coqsrc}
Example F9 := Fcompose F3 F3.

Compute F9.
\end{Coqsrc}

\begin{Coqanswer}
= fun (A : Type) (x : Fkont A) (x0 : A) =>
       x1 <--- x0 times x0;
       x2 <--- x1 times x0; x3 <--- x2 times x2; 
       x4 <--- x3 times x2; 
       x x4
     : Fchain  
\end{Coqanswer}

\begin{Coqsrc}
Remark F9_correct :chain_correct (F2C F9) 9.
Proof.
  apply param_correctness_pos;  lazy;  parametric_tac.
Qed.
\end{Coqsrc}

\begin{figure}[h]
  \centering
  \begin{tikzpicture}

 \node (inputx)  at (-0.2,0) {};  
\node (x)  at (1,0) {$x$};  
 \node (x2)  at (2,0) {$x^2$};  
 \node (x3)  at (3,0) {$x^3$};
 \node (output1)  at (4,0) {};  
 \node (inputy)  at (5.5,0) {};  
\node (y)  at (6.5,0) {$y$};  
 \node (y2)  at (7.5,0) {$y^2$};  
 \node (y3)  at (8.5,0) {$y^3$};
 \node (output2)  at (9.6,0) {};  
 \draw [>-,   thick](inputx) -- (x);
 \draw [->,   >=latex](x) -- (x2);
\draw [->,   >=latex](x2) -- (x3);
\draw [->, >=latex](x) to [bend left] (x3);
\draw [->, >=latex, thick](x3) -- (output1);
\draw [->, dotted ](output1) -- (inputy) node [midway, above] {\tiny{$y:=x$}};
\draw [>-,   thick](inputy) -- (y);
 \draw [->,   >=latex](y) -- (y2);
\draw [->,   >=latex](y2) -- (y3);
\draw [->, >=latex](y) to [bend left] (y3);
\draw [->, >=latex, thick](y3) -- (output2);
\draw[dashed] (0.4,-1) rectangle (3.5,1);
\draw[dashed] (6,-1) rectangle (9,1);
\draw[dashed] (0.2,-1.2) rectangle (9.2,1.2);
  \end{tikzpicture}
  \caption{Composition  of F-chains: F9}
  \label{fig:F9}
\end{figure}

Using structural recursion and the operator \texttt{FCompose},
we build a chain for any exponent of the form $2^n$:

\begin{Coqsrc}
 Fixpoint  Fexp2_of_nat (n:nat) : Fchain :=
 match n with O => F1
            | S p => Fcompose F2 (Fexp2_of_nat p)
 end.

Definition Fexp2 (p:positive) : Fchain :=
  Fexp2_of_nat (Pos.to_nat p).

Compute Fexp2  4.
\end{Coqsrc}

\begin{Coqanswer}
 = fun (A : Type) (x : Fkont A) (x0 : A) =>
       x1 <--- x0 times x0;
       x2 <--- x1 times x1; x3 <--- x2 times x2; 
       x4 <--- x3 times x3; x x4
     : Fchain
\end{Coqanswer}
%%% ICI

\subsection{F-chain correctness}
Let \texttt{f} be some term of type \texttt{Fchain}, and \texttt{n:nat}.
We would like to say that \texttt{f} is correct \emph{w.r.t.} \texttt{n:nat}
if for any continuation \texttt{k} and \texttt{a}, the application of 
\texttt{f} to \texttt{k} and \texttt{a} computes \texttt{$k(a^n)$}.

\begin{Coqbad}
Module Bad.

Definition Fchain_correct  (f : Fchain) (n:nat) :=
  forall A `(M : @EMonoid A op E_one E_equiv) k (a:A),
    computation_execute op (f A k  a)==
    computation_execute op (k  (a ^ n)).
\end{Coqbad}

Let us now try to prove that \texttt{F3} is correct \emph{w.r.t.} $3$.

\begin{Coqbad}
Theorem F3_correct : Fchain_correct F3 3.
Proof.  
  intros    A op E_one E_equiv M k  a ; simpl.
  monoid_simpl M.  
\end{Coqbad}

\begin{Coqanswer}
  A : Type
  op : Mult_op A
  E_one : A
  E_equiv : Equiv A
  M : EMonoid op E_one E_equiv
  k : Fkont A
  a : A
  H : Proper (equiv ==> equiv ==> equiv) op
  ============================
   computation_execute op (k (a * a * a)) ==
   computation_execute op (k (a * (a * (a * E_one))))
\end{Coqanswer}

\begin{Coqbad}
Abort.  
End Bad.
\end{Coqbad}

This failure is due to a lack of an assumption that the continuation
\texttt{k} is \emph{proper} with respect to the equivalence \texttt{equiv}.
Thus, \coq{} is unable to infer from the equivalence 
\texttt{(a * a * a) == (a * (a * (a * E\_one)))} \linebreak that 
\texttt{k (a * a * a)} and \texttt{k (a * (a * (a * E\_one)))} are 
equivalent computations.

\subsubsection{Definition:} 
\index{Type Classes!Proper}

A continuation \texttt{k:Fkont A} is \emph{proper}
if, whenever \linebreak[3] \texttt{x == y} holds, the computations \texttt{k x} and 
\texttt{k y} are equivalent.

\begin{Coqsrc}
Class Fkont_proper
      `(M : @EMonoid A op E_one E_equiv) (k: Fkont A )  :=
  Fkont_proper_prf:
    Proper (equiv ==> computation_equiv op E_equiv) k.
\end{Coqsrc}

We are now able to improve our definition of correctness, taking only
proper continuations into account.

\begin{Coqsrc}
Definition Fchain_correct_nat (f : Fchain) (n:nat) :=
 forall A `(M : @EMonoid A op E_one E_equiv) k
        (Hk :Fkont_proper M k)
        (a : A) ,
 computation_execute op (f A k  a) ==
 computation_execute op (k  (a ^ n)).

Definition Fchain_correct (f : Fchain) (p:positive) :=
 Fchain_correct_nat f (Pos.to_nat p).
\end{Coqsrc}

\subsubsection{Examples}

Let us show some manual correctness proofs for small f-chains:

\begin{Coqsrc}
Lemma F1_correct : Fchain_correct F1 1.
Proof.
  intros until M ; intros k Hk a ; unfold F1; simpl.
  apply Hk; monoid_simpl M; reflexivity.
Qed.
\end{Coqsrc}

 While proving \texttt{F3}'s correctness, we will have to apply
 the properness hypothesis on \texttt{k}:

\begin{Coqsrc}
Theorem F3_correct : Fchain_correct F3 3.
Proof. 
  intros until M; intros k Hk a; simpl.
\end{Coqsrc}

\begin{Coqanswer}
 A : Type
  op : Mult_op A
  E_one : A
  E_equiv : Equiv A
  M : EMonoid op E_one E_equiv
  k : Fkont A
  Hk : Fkont_proper M k
  a : A
  ============================
   computation_execute op (k (a * a * a)) ==
   computation_execute op (k (a * (a * (a * E_one))))}
\end{Coqanswer}

\begin{Coqsrc}
apply Hk. 
\end{Coqsrc}

\begin{Coqanswer}
...
============================
   a * a * a == a * (a * (a * E_one))}
\end{Coqanswer}

\begin{Coqsrc}
  monoid_simpl M; reflexivity.
Qed.
\end{Coqsrc}

Correctness of \texttt{F2} is proved the same way:

\begin{Coqsrc}
Theorem F2_correct : Fchain_correct F2 2.
Proof. 
  intros until M; intros k Hk a; simpl;
  apply Hk;  monoid_simpl M;  reflexivity.
Qed.
\end{Coqsrc}

\subsubsection{Composition of correct f-chains: a first attempt}

We are now looking for a way to generate correct chains for any positive 
number. It seems obvious that we could use \texttt{Fcompose} for building 
a correct f-chain for $n\times p$ by composition of a correct f-chain for 
$n$ and a correct f-chain for $p$.

Let us try to certify this construction:

\begin{Coqbad}
Module Bad2.

Lemma Fcompose_correct_attempt :
  forall f1 f2 n1 n2, Fchain_correct f1 n1 ->
                      Fchain_correct f2 n2 ->
                      Fchain_correct (Fcompose f1 f2)
                                       (n1 * n2).

(* Beginning of proof omitted *)
\end{Coqbad}

\begin{Coqanswer}
  Hk : Fkont_proper M k
  a, x, y : A
  Hxy : x == y
  ============================
   computation_execute op (f2 A k x) == 
   computation_execute op (f2 A k y) 
\end{Coqanswer}

No hypothesis guarantees us that the execution of \texttt{f2} respects the equivalence
\texttt{x == y}.

\begin{Coqbad}
Abort.
\end{Coqbad}

 Thus, we need to define also a  notion of properness for f-chains. 
A first attempt would be :

\begin{Coqbad}
Module Bad3.

Class Fchain_proper_ (fc : Fchain) := Fchain_proper_prf : 
 forall  `(M : @EMonoid A op E_one E_equiv) k  ,
    Fkont_proper M k 
    forall x y, x == y ->
        @computation_equiv _ op E_equiv (fc A k x) (fc A k y).
\end{Coqbad}

This definition is powerful enough for proving that properness is 
preserved by composition:

\begin{Coqbad}
Instance Fcompose_proper_ (f1 f2 : Fchain)
                             (_ : Fchain_proper_simple f1)
                             (_ : Fchain_proper_simple f2) :
 Fchain_proper_ (Fcompose f1 f2).
Proof. 
 intros until M;intros k Hk x y Hxy; unfold Fcompose;cbn. 
 apply (H _ _ _ _ M); auto.
 intros u v Huv;apply (H0 _ _ _ _ M);auto.
Qed.
\end{Coqbad}

Nevertheless, we had to throw away  this definition of properness:
In further 
developments (Sect.~\vref{Kkonts-section})  we shall  have to compare
executions of the form \texttt{fc A $k_x$ x} and \texttt{fc A $k_y$ y}
where \texttt{x == y} and {$k_x$} and {$k_y$} are 
``equivalent''
but not \emph{convertible} continuations.

\begin{Coqbad}
End Bad3.
\end{Coqbad}

\subsubsection{A better definition of properness}

 The following  generalization will allow us to consider continuations that are
different (according to Leibniz equality) but lead to equivalent
computations and results.

\begin{Coqsrc}
Definition Fkont_equiv  `(M : @EMonoid A op E_one E_equiv)
 (k k': Fkont A )  := 
 forall x y : A, x == y ->
                 computation_equiv op E_equiv  (k x)  (k' y).

Class Fchain_proper (fc : Fchain) := Fchain_proper_prf : 
 forall  `(M : @EMonoid A op E_one E_equiv) k k' ,
    Fkont_proper M k -> Fkont_proper M k' ->    
    Fkont_equiv M k k' ->
    forall x y, x == y ->
        @computation_equiv _ op E_equiv
                           (fc A k x)
                           (fc A k' y).
\end{Coqsrc}

\subsubsection{Examples}
The definition above allows us  to build simply several instances of the class \linebreak
\texttt{Fchain\_proper}:

\begin{Coqsrc}
Instance F1_proper : Fchain_proper F1.
Proof.
  intros until M ; intros k k' Hk Hk' H a b H0; unfold F1; cbn;
  now apply H.  
Qed.
\end{Coqsrc}

\begin{Coqsrc}
Ltac add_op_proper M H := 
 let h := fresh H in
   generalize (@Eop_proper _ _ _ _ M); intro h.

Instance F3_proper : Fchain_proper F3.
Proof.
  intros  A op one equiv M  k k' Hk Hk'  Hkk' x y Hxy;  
  apply Hkk'; add_op_proper M H; repeat rewrite Hxy;
  reflexivity.
Qed.
\end{Coqsrc}

We are now able to prove  \texttt{Fexp2 $n$}'s correctness by induction 
on $n$:

\begin{Coqsrc}
Instance Fexp2_nat_proper (n:nat) : 
                           Fchain_proper (Fexp2_of_nat n).
Proof.
 induction n; cbn.
   - apply F1_proper.
   - apply Fcompose_proper ; [apply F2_proper | apply IHn].
Qed.
\end{Coqsrc}

\begin{Coqsrc}
Lemma  Fexp2_nat_correct (n:nat) : 
         Fchain_correct_nat (Fexp2_of_nat n) (2  ^ n).
Proof.
  induction n; cbn.
 - apply F1_correct.
 -  rewrite Nat.add_0_r;
   replace (2 ^ n + 2 ^ n)%nat with (2 * 2 ^n)%nat by  omega;
   apply Fcompose_correct_nat;auto.
   +  apply F2_correct.
   +  apply  Fexp2_nat_proper.
Qed.
\end{Coqsrc}

\begin{Coqsrc}
Lemma  Fexp2_correct (p:positive) : 
                           Fchain_correct (Fexp2 p) (2  ^ p).
(* Proof omitted *)

Instance  Fexp2_proper (p:positive) : Fchain_proper (Fexp2 p).
(* Proof omitted *)

\end{Coqsrc}

We are now  able to build chains for any exponent of the form 
$2^k\times 3^p$, using \texttt{Fcompose}. Les us look at a simple example:

\begin{Coqsrc}
Hint Resolve F1_correct F1_proper
     F3_correct F3_proper Fcompose_correct Fcompose_proper
     Fexp2_correct Fexp2_proper .

Example F144:  {f : Fchain | Fchain_correct f 144 /\
                                Fchain_proper f}.
Proof.
 change 144 with ( (3 * 3) * (2 ^ 4))%positive.
 exists (Fcompose (Fcompose F3 F3) (Fexp2 4)); auto.
Defined.

Compute proj1_sig F144.
\end{Coqsrc}

\begin{Coqanswer}
= fun (A : Type) (x : Fkont A) (x0 : A) =>
       x1 <--- x0 times x0;
       x2 <--- x1 times x0;
       x3 <--- x2 times x2;
       x4 <--- x3 times x2;
       x5 <--- x4 times x4;
       x6 <--- x5 times x5; 
       x7 <--- x6 times x6; 
       x8 <--- x7 times x7; 
       x x8
     : Fchain  
\end{Coqanswer}

\subsection{Building chains for two distinct exponents : k-chains  \label{Kkonts-section}}

\subsubsection{Introduction}
Not every chain can be built efficiently  with \texttt{Fcompose}.
 For instance, consider the exponent $n= 23 = 3 + 2^4 + 2^2$.

One may attempt to define a new operator  for combining f-chains for 
$n$ and $p$ into an f-chain for $n+p$.

\begin{Coqbad}
Definition Fplus (f1 f2 : Fchain) : Fchain :=
  fun A k x => 
   f1 A (fun y => 
            f2 A (fun z => t <--- z times y; k t) x) 
            x.
\end{Coqbad}

For instance, we can define a chain for $23$:

\begin{Coqbad}
Let F23 := Fplus F3 (Fplus (Fexp2 4) (Fexp2 2)).  
\end{Coqbad}

Unfortunately, our construct is still very inefficient, since it results in 
duplications of computations, as shown by the normal form of \texttt{F23}.

\begin{Coqbad}
Compute F23
\end{Coqbad}

\begin{Coqanswer}
 = fun (A : Type) (k : Fkont A) (x0 : A) =>
       x1 <--- x0 times x0;
       x2 <--- x1 times x0;
       x3 <--- x0 times x0;
       x4 <--- x3 times x3;
       x5 <--- x4 times x4;
       x6 <--- x5 times x5;
       x7 <--- x0 times x0;
       x8 <--- x7 times x7; 
       x9 <--- x8 times x6; 
       x10 <--- x9 times x2; 
       k x10
\end{Coqanswer}

We observe that the variables \texttt{x3} and \texttt{x7} are 
useless, since
they will have the same value as \texttt{x1}. Likewise, computing
\texttt{x8} (same value as \texttt{x4}) is a waste of time.

 A better scheme for computing $x^{23}$ would be the following one:

 \begin{enumerate}
 \item Compute $x$, $x^2$, $x^3$, \textbf{and} $x^6 = {(x^3)}^2$, then  $x^7$,
 \item Compute $x^{10} = x^7 \times x^3$, then $x^{20}$
 \item Finally, return  $x^{23} = x^{20} \times x^3$
 \end{enumerate}

In fact, the first step of this sequence  computes \emph{two}
values: $x^7$ and $x^3$, that are re-used by the rest of the computation.

  Like in some programming languages
 that allow  ``multiple values'', like \texttt{Scheme} and \texttt{Common Lisp}, we can  express this feature 
 in terms of continuations that accept two arguments.
Thus, we extend our previous definitions to chains that return two 
different powers of their argument\footnote{The name \texttt{Kchain} comes from previous versions of this development. It may be changed later.}.

\index{Programming Styles!Continuation Passing Style (CPS)}

\begin{Coqsrc}
Definition Kkont A:=  A -> A -> @computation A.

Definition Kchain :=  forall A, Kkont A -> A -> @computation A.
\end{Coqsrc}

\subsubsection{Examples}

The chain \texttt{k3\_1} sends both values $x$ and $x^3$ to its continuation.
Likewise, \texttt{k7\_3} ``returns''  $x^7$ and $x^3$.

\begin{Coqsrc}
Example k3_1 : Kchain := fun A (k:Kkont A) (x:A) =>
  x2 <--- x times x ;
  x3 <--- x2 times x ;
  k x3 x.

Example k7_3 : Kchain := fun A (k:Kkont A)  (x:A) =>
  x2 <--- x times x;
  x3 <--- x2 times x ;
  x6 <--- x3 times x3 ;
  x7 <--- x6 times x ;
  k  x7 x3.
\end{Coqsrc}

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
  \node (entree) at (0,0) {};
  \node (X) at (1,0) {$x$};
\node (X2) at (2,0) {$x^2$};
  \node (X3) at (3,0) {$x^3$};
\node (sortieX) at (4.5,0.6) {};
\node(beforeSortieX) at (2,0.6){};
\node (sortieX3) at (4.5,0) {};
\draw [>-,   thick](entree) -- (X) node [near start, above] {\tiny{$x$}};
\draw [->, >=latex, ](X) -- (X2);
\draw [->, >=latex, ](X2) -- (X3);
\draw [->,  thick](X3) -- (sortieX3) node [at end, above] {\tiny{$x^3$}};
\draw [->, >=latex](X) to [bend left] (X3);
\draw [->, >=latex](X) to [bend left] (X3);
%\draw [thick](X) to [bend left] (beforeSortieX)
\draw [->, thick](X) -- (2,0.6) --  (sortieX) node [at end, above] {\tiny{$x$}};
\draw[dashed] (0.6,-0.4) rectangle (3.7,1);
  \end{tikzpicture}
  \caption{Graphical representation of \texttt{K3\_1}}
  \label{fig:K3-1-as-dag}
\end{figure}

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
  \node (entree) at (-0.4,0) {};
  \node (X) at (1,0) {$x$};
  \node (X2) at (2,0) {$x^2$};
\node (X3) at (3.5,0) {$x^3$};
  \node (X6) at (4.5,0) {$x^6$};
\node (X7) at (6,0) {$x^7$};
  \draw [>-,   thick](entree) -- (X) node [near start, above] {\tiny{$x$}};
 \draw [->](X) -- (X2);
\draw [->](X2) -- (X3);
\draw [->, bend left](X) to (X3) ;
\draw [->, bend left](X) to (X7) ;
\draw [->](X3) -- (X6) ;
\draw [->](X6) -- (X7) ;
  \node (sortieX7) at (7.5,0) {};
  \node (sortieX3) at (7.5,0.8) {};
 \draw [->,   thick](X7) -- (sortieX7) node [at end, above] {\tiny{$x^7$}};
 \draw [->,   thick](X3) -- (5.5,0.8) -- (sortieX3)  node [at end, above] {\tiny{$x^3$}};
\draw  [dashed] (0.5,-0.7) rectangle (6.5,1.4);
  \end{tikzpicture}
  \caption{Graphical representation of \texttt{K7\_3}}
  \label{fig:K7-3-as-dag}
\end{figure}

\subsubsection{Definitions}

First, we have to adapt to k-chains our definitions of correctness and properness. 
\begin{Coqsrc}
Definition Kkont_proper `(M : @EMonoid A op E_one E_equiv)
           (k : Kkont A) :=
 Proper (equiv ==> equiv ==> computation_equiv op E_equiv) k .

Definition Kkont_equiv  `(M : @EMonoid A op E_one E_equiv)
           (k k': Kkont A )  := 
 forall x y : A, x == y -> forall z t, z == t -> 
         computation_equiv op E_equiv   (k  x z) (k' y t).
 \end{Coqsrc}

A k-chain is correct with respect to two exponents $n$ and $p$ 
  if it computes $x ^ n$ and $x ^ p$ for any $x$ in any monoid $M$.

 \begin{Coqsrc}
Definition Kchain_correct_nat (kc : Kchain) (n p : nat) :=
  forall `(M : @EMonoid A op E_one E_equiv)
        (k : Kkont A), 
        Kkont_proper M k ->
        forall  (x : A) ,
              computation_execute op (kc  A k  x) ==
              computation_execute op (k  (x ^ n) (x ^ p)).

\end{Coqsrc}

\begin{Coqsrc}
Definition Kchain_correct (kc : Kchain) (n p : positive) :=
  Kchain_correct_nat kc (Pos.to_nat n) (Pos.to_nat p).

Class Kchain_proper (kc : Kchain) :=
Kchain_proper_prf : 
 forall `(M : @EMonoid A op E_one E_equiv) k k' x y ,
   Kkont_proper M k ->
   Kkont_proper M k' -> 
   Kkont_equiv M k k' ->
   E_equiv x y ->
   computation_equiv op E_equiv (kc A k x) (kc A k' y).
\end{Coqsrc}

\subsubsection{Example}
For instance, let us prove that \texttt{k7\_3} is proper and correct for the exponents  $7$ and $3$.

\begin{Coqsrc}
Instance k7_3_proper : Kchain_proper k7_3.
Proof.
  intros until M; intros; red; unfold k7_3; cbn;
  add_op_proper M H3; apply H1;  rewrite H2;   reflexivity. 
Qed.

Lemma k7_3_correct : Kchain_correct k7_3 7 3.
Proof.
  intros until M; intros; red; unfold k7_3; simpl.
  apply H; monoid_simpl M;  reflexivity.
Qed. 
\end{Coqsrc}

\subsection{Systematic construction of  correct f-chains and k-chains}

We are now ready to define various operators on f- and k-chains, and prove these
operators preserve correcness and properness. We will also show that 
these operators allow to generate easily correct chains for any positive 
exponent. They will be used to generate chains for
numbers of the form $n=bq+r$ where $0\leq r < b$, assuming the previous
construction of correct chains for $r$, $b$ and $q$.
For instance, Figure~\ref{fig:K7-3-decomposition} shows how \texttt{K7\_3} is built
as a composition of \texttt{K3\_1} and \texttt{F2}.

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
  \node (entree) at (-0.4,0) {};
  \draw [dashed] (1,-0.5) rectangle (2,0.5);
  \node at (1.5,0){$K3\_1$};
  \draw [>-,   thick](entree) -- (1,0) node [at start, above] {\tiny{$x$}};
  \node (sortiex3) at (2,0) {};
  \node (sortiex) at (2,0.4) {};
  \draw [dashed] (4.5,-0.5) rectangle (5.5,0.5);

  \node at (5,0){$F2$};
  \node (entreeF2) at (4.5,0) {};
\node (sortieF2) at (5.5,0) {};
  \draw[->, ] (sortiex3) -- (entreeF2) 
       node [midway, above] {\tiny{$x^3$}};
  \node (sortiex3global) at (7,1) {};
  \node (sortiex3global) at (8.4,1)  {};
  \draw [->, thick, bend left] (3.6,0) -- (4,0.6) -- (5,1) --  node [at end, above]
 {\tiny{$x^3$}} (sortiex3global) ;
  \node (join) at (7,0) {$\times$};
  \draw [->, ] (sortieF2) -- (join)
    node [midway, above] {\tiny{$x^6$}};
\draw [->, bend left] (sortiex) to node [midway, above] {\tiny{$x$}} (join);

 \draw [->,   thick](join) -- (8.4,0) node [at end, above] {\tiny{$x^7$}};
% \draw [->,   thick](sortiex3global) --  (8.4,1);
\draw  [dashed] (0.5,-0.7) rectangle (7.5,1.4);
  \end{tikzpicture}
  \caption{Decomposition of  \texttt{K7\_3}}
  \label{fig:K7-3-decomposition}
\end{figure}

\subsubsection{Conversion from k-chains into f-chains}

Any k-chain for $n$ and $p$ can be converted into an f-chain, just by applying it to a continuation that 
ignores its second argument.

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
\node (input) at (-0.5,0){};
\node(inputkc) at (1.5,0){};
\node(outputkc1) at (2.5,0){};
\node(outputkc2) at (2.5,0.4){};
\node (ignore) at (3,0.4){{$\circ$}};
\node (output) at (4.4,0){};
\draw [dashed] (1.2,-0.5) rectangle (2.5,0.5) ;
\draw [dashed] (0.5,-0.7) rectangle (3.3,0.8) ;
\node (knp) at (2,0) {$k_{n,p}$};
\draw[>->] (input) -- node [near start, above] {\tiny{$x$}} (inputkc);
\draw[thick, ->,>=latex] (outputkc1) +(-0.1,0) -- (output) node [near end,above] {\tiny{$x^n$}};
\draw[thick] (outputkc2) +(-0.1,0)  -- (ignore);
\end{tikzpicture}
  \caption{The \texttt{K2F (knp)} construction}
  \label{fig:K2F}
\end{figure}

\begin{Coqsrc}
Definition K2F (knp : Kchain) : Fchain :=
  fun A (k:Fkont A) => kc A (fun  y _ => k y).

Lemma K2F_correct :
  forall knp n p, Kchain_correct kc n p ->
                 Fchain_correct (K2F knp) n.
(* Proof omitted *)

Instance K2F_proper (kc : Kchain)(_ : Kchain_proper kc) :
                                  Fchain_proper (K2F kc).

(* Proof omitted} *)
\end{Coqsrc}

\subsubsection{Construction associated with Euclidean division with a positive rest}

Let $n=bq+r$, with $0<r<b$. Then, for any $x$,  $x^n= (x^{b})^q \times x^r$. Thus, we can 
compose an chain that computes $x^b$ and $x^r$ with a chain that raises
any $y$ to its $q$-th power for obtaining a chain that computes $x^n$.

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
  \node (entree) at (-0.4,0) {};
  \draw [dashed] (1,-0.5) rectangle (2,0.5);
  \node at (1.5,0){$K_{b,r}$};
  \draw [>-,   thick](entree) -- (1,0) node [above, near start] {\tiny{$x$}};
  \node (sortiex3) at (2,0) {};
  \node (sortiex) at (2,0.4) {};
  \draw [dashed] (4.5,-0.5) rectangle (5.5,0.5);
  \node at (5,0){$F_q$};
  \node (entreeF2) at (4.5,0) {};
\node (sortieF2) at (5.5,0) {};
  \draw[->, ] (sortiex3) -- (entreeF2) 
       node [midway, above] {\tiny{$x^b$}};
  %\node (sortiex3global) at (7,1) {};
  \node (sortiex3global) at (7,1)  {};
  \draw [->, , bend left] (3.6,0) -- (4,0.6) -- (5,1) -- node [near end, above] {\tiny{$x^b$}} (8.7,1)  ;
  \node [draw] (multipl) at (7,0) {$\times$};
\draw [->] (2,0.4) to (3,1.5) -- node [midway,above] {\tiny{$x^r$}} (6,1.5) to (multipl);
  \draw [->, ] (sortieF2)  -- (multipl)   node [midway, above] {\tiny{$x^{bq}$}};
 \draw [->,   thick](multipl) -- (8.7,0) node [above, near end] {\tiny{$x^{bq+r}$}};
\draw  [dashed] (0.5,-0.7) rectangle (7.5,2);
  \end{tikzpicture}
  \caption{The KFK combinator}
  \label{fig:KFK}
\end{figure}

\begin{Coqsrc}
Definition KFK (kbr : Kchain) (fq : Fchain) : Kchain  :=
  fun A k a =>
    kbr A  (fun xb xr =>
              fq A (fun y =>
                      z <--- y times xr; k z xb) xb) a.

Lemma KFK_correct :
  forall (b q r : positive) (kbr : Kchain) (fq : Fchain),
    Kchain_correct kbr b r ->
    Fchain_correct fq q ->
    Kchain_proper kbr ->
    Fchain_proper fq -> 
    Kchain_correct (KFK kbr fq) (b * q + r) b.
(* Proof omitted *)

Instance KFK_proper :
forall (kbr : Kchain) (fq : Fchain),
  Kchain_proper kbr -> 
  Fchain_proper fq -> 
  Kchain_proper (KFK kbr fq)
(* Proof omitted *)
\end{Coqsrc}

%\subsection{More certified operators on chains}

\subsubsection{Ignoring the remainder}

Let $n=bq+r$, with $0<r<b$. The following construction computes
$x^r$ and $x^b$, then $x^{bq}$, and finally sends $x^{bq+r}$ to the continuation,
throwing away $x^b$.

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
  \node (entree) at (-0.4,0) {};
  \draw [dashed] (1,-0.5) rectangle (2,0.5);
  \node at (1.5,0){$K_{b,r}$};
  \draw [>-,   thick](entree) -- (1,0) node [above, near start] {\tiny{$x$}};
  \node (sortiex3) at (2,0) {};
  \node (sortiex) at (2,0.4) {};
  \draw [dashed] (4.5,-0.5) rectangle (5.5,0.5);
  \node at (5,0){$F_q$};
  \node (entreeF2) at (4.5,0) {};
\node (sortieF2) at (5.5,0) {};
  \draw[->, ] (sortiex3) -- (entreeF2) 
       node [midway, above] {\tiny{$x^b$}};
  %\node (sortiex3global) at (7,1) {};
  %\node (sortiex3global) at (7,1)  {};
  % \draw [->, , bend left] (3.6,0) -- (4,0.6) -- (5,1) -- node [near end, above] {\tiny{$x^b$}} (8.7,1)  ;
  \node [draw] (multipl) at (7,0) {$\times$};
\draw [->] (2,0.4) to (3,1.5) -- node [midway,above] {\tiny{$x^r$}} (6,1.5) to (multipl);
  \draw [->, ] (sortieF2)  -- (multipl)   node [midway, above] {\tiny{$x^{bq}$}};
 \draw [->,   thick](multipl) -- (8.7,0) node [above, near end] {\tiny{$x^{bq+r}$}};
\draw  [dashed] (0.5,-0.7) rectangle (7.5,2);
  \end{tikzpicture}
  \caption{The KFF combinator}
  \label{fig:KFF}
\end{figure}

\begin{Coqsrc}
Definition KFF (kbr : Kchain) (fq : Fchain) : Fchain :=
  K2F (KFK kbr fq).

Lemma KFF_correct :
forall (b q r : positive) (kbr : Kchain) (fq : Fchain),
Kchain_correct kbr b r ->
Fchain_correct fq q ->
Kchain_proper kbr ->
Fchain_proper fq -> Fchain_correct (KFF kbr fq) (b * q + r).
(* Proof omitted *)

Instance KFF_proper :
forall (kbr : Kchain) (fq : Fchain),
Kchain_proper kbr -> Fchain_proper fq -> Fchain_proper (KFF kbr fq).
(* Proof omitted *)
\end{Coqsrc}

\subsubsection{Conversion of an f-chain into a k-chain}
The following conversion is useful when a chain generation algorithm
needs to build a k-chain for exponents $p$ and $1$:

\begin{Coqsrc}

Definition FK (f : Fchain) : Kchain :=
  fun (A : Type) (k : Kkont A) (a : A) =>
    f A (fun y => k y a) a.

Lemma FK_correct : forall (p: positive) (Fp : Fchain),
                     Fchain_correct  Fp p ->
                     Fchain_proper Fp ->
                     Kchain_correct (FK Fp) p 1.
(* Proof omitted *)

Instance  FK_proper  (Fp : Fchain) (_ : Fchain_proper Fp):
  Kchain_proper (FK Fp).
(* Proof omitted *)
\end{Coqsrc}

\subsubsection{Computing $x^p$ \emph{and} $x^{pq}$}

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
  \node (entree) at (-0.4,0) {};
  \draw [dashed] (1,-0.5) rectangle (2,0.5);
  \node at (1.5,0){$F_p$};
  \draw [>-,   thick](entree) -- (1,0) node [above, near start] {\tiny{$x$}};
  \node (sortiexp) at (2,0) {};
  \draw [dashed] (4.5,-0.5) rectangle (5.5,0.5);
  \node at (5,0){$F_q$};
  \node (entreeF2) at (4.5,0) {};
\node (sortieF2) at (5.5,0) {};
  \draw[->, ] (sortiex3) -- (entreeF2) 
       node [midway, above] {\tiny{$x^p$}};
  \draw [->, , bend left] (3.6,0) -- (4,0.6) -- (5,1) -- node [near end, above] {\tiny{$x^p$}} (7.7,1)  ;
 \draw [->,   thick](sortieF2) -- (7.7,0) node [above, near end] {\tiny{$x^{pq}$}};
\draw  [dashed] (0.5,-0.7) rectangle (6.5,2);
  \end{tikzpicture}
  \caption{The FFK combinator}
  \label{fig:FFK}
\end{figure}

\begin{Coqsrc}
Definition FFK (fp fq : Fchain) : Kchain :=
  fun A k a =>  fp A (fun xb  => fq A (fun y => k y xb) xb) a.

Lemma FFK_correct :
  forall (p q  : positive) (fp fq : Fchain),
    Fchain_correct fp p  ->
    Fchain_correct fq q ->
    Fchain_proper fp ->
    Fchain_proper fq -> Kchain_correct (FFK fp fq) (p * q ) p.
(* Proof omitted *)

Instance FFK_proper 
   (fp: Fchain) (fq : Fchain)
    (_ :   Fchain_proper fp)
    (_ :  Fchain_proper fq) :  Kchain_proper (FFK fp fq) .
(* Proof omitted *)
\end{Coqsrc}

\subsubsection{A correct-by-construction chain}

A simple example will show us how to build correct chains 
for any positive exponent, using the operators above.

\begin{Coqsrc}
Hint Resolve KFF_correct KFF_proper KFK_correct KFK_proper.

Definition F87 :=
 let k7_3 :=  KFK k3_1 (Fexp2 1) in
 let k10_7 := KFK k7_3 F1 in
 KFF k10_7 (Fexp2 3).

Lemma OK87 : Fchain_correct F87 87.
Proof.
 unfold F87; change 87 with (10 * (2 ^ 3) + 7)%positive.
 apply KFF_correct;auto.
 change 10 with (7 * 1 + 3); apply KFK_correct;auto.
 change 7 with (3 * 2 ^ 1 + 1)%positive;  apply KFK_correct;auto.
Qed.
\end{Coqsrc}

Note that this method of construction still requires some  
interaction from the user. 
In the next section, we build a \emph{function} that maps any 
positive number $n$ into a correct and proper chain for $n$.
Thus correct chain generation will be fully automated.

\subsection{Automatic chain generation by Euclidean division}

The goal of this section is to write a function 
\texttt{make\_chain (p:positive): chain} that builds a correct chain for $p$, using
the Euclidean method above. In other words, we want to get correct chains
by computation. The correctness of the result of this computation should be
asserted by a  theorem:

\begin{Coqsrc}
Theorem make_chain_correct : 
   forall p, chain_correct (make_chain p) p.  
\end{Coqsrc}

In the previous section, we  considered two different kinds of objects:
f-chains, associated with a single exponent, and k-chains, associated with two exponents. We would expect that the function \texttt{make\_chain} we want to build and certify is structured as a pair of mutually recursive functions.
 In \coq{} , various ways of building such functions are available:
 \begin{itemize}
 \item Structural [mutual] recursion with \texttt{Fixpoint}
 \item  Using \texttt{Program Fixpoint}
 \item Using   \texttt{Function}.
 \end{itemize}

Since our construction is based on Euclidean division, we could not
define our chain generator by structural recursion. 
For simplicity's sake, we chosed to avoid dependent elimination
 and used \texttt{Function}  with a decreasing measure.

 For this purpose, we define a single data-type for associated with
 the generation of F- and K-chains.

We had two slight technical problems to consider:
\begin{itemize}
\item The generation of a k-chain for $n$ and $p$ is meaningfull only if $p < n$. Thus, in order to avoid a clumsy  dependent pattern-matching, we chosed to represent 
     a pair $(n,p)$ where $0<p<n$ by a pair of positive numbers $(p,d)$ where 
     $d=n-p$
\item In order to avoid to deal explicitly with mutual recursion, we
     defined a type called \texttt{signature} for representing both
     forms of function calls.
     Thus, it is easy to define a decreasing measure on type 
     \texttt{signature} for proving termination. 
    Likewise, correctness and properness statements are also indexed by 
    this type.

\end{itemize}

\begin{Coqsrc}
Inductive signature : Type :=
|  (** Fchain for the exponent n *)
     gen_F (n:positive) 
| (** Kchain for the exponents p+d  and p *) 
     gen_K (p d: positive).
\end{Coqsrc}

The following dependently-typed functions will help us to specify  formally
any correct chain generator.
\index{Dependently Typed Functions}

\begin{Coqsrc}
(** 
  exponent associated with a signature:
*)
Definition signature_exponent (s:signature) : positive :=
 match s with 
| gen_F n => n 
| gen_K p d  =>  p + d
end.
\end{Coqsrc}

\begin{Coqsrc}
(**
Type of the associated continuation 
*)

Definition kont_type (s: signature)(A:Type) : Type :=
match s with 
| gen_F _  => Fkont A 
| gen_K _ _   => Kkont A
end.

Definition chain_type (s: signature) : Type :=
 match s with 
| gen_F _   => Fchain
|  gen_K _ _  => Kchain
end.

\end{Coqsrc}

\begin{Coqsrc}
Definition correctness_statement (s: signature) : 
chain_type s -> Prop :=
match s  with
  | gen_F p => fun ch => Fchain_correct ch p
  | gen_K p d   => fun ch => Kchain_correct ch (p + d) p
end.

Definition proper_statement (s: signature) : 
chain_type s -> Prop :=
match s  with
  | gen_F p => fun ch => Fchain_proper ch 
  | gen_K p d   => fun ch => Kchain_proper ch 
end.

(**  Full correctness *)

Definition  OK (s: signature) 
  := fun c: chain_type s => 
       correctness_statement s c /\
       proper_statement s c.

\end{Coqsrc}

\subsection{The dichotomic strategy}

Assume we want to build automatically a correct  f-chain for some 
positive integer $n$.
If $n$ equals to $1$, $3$, or $2^p$ for some positive integer  $p$,
this task is immediate, thanks to the constants \texttt{F1}, 
\texttt{F3} and \texttt{Fexp2}.
Otherwise, like in \cite{DBLP:journals/ita/BrlekCHM95}, we decompose 
$n$ into $bq+r$, where $1<b<n$, and compose the recursively built
chains for $q$ and $r$ on one side, and $q$ on the other side.

The efficiency of this method depends on the choice of $b$.
In \cite{DBLP:journals/ita/BrlekCHM95}, the function that maps $n$ into $b$
is called a \emph{strategy}. In this chapter, we concentrate
on the so-called \emph{dichotomic strategy}.

$\delta(n) = n \div {2^k}$ where $k=\floor{(\log_2{n})/2}$.

Intuitively, it corresponds to splitting the binary representation of a positive
integer into two halves. For instance, consider $n=87$ its binary representation
is \texttt{1010111}. The number $\floor{(\log_2{n})/2}$ is equal to $3$.
Dividing $n$ by $2^3$ gives the decomposition $n=10 \times 2^3 + 7$.
Thus, a chain for $n=87$ can be built from a chain computing both $x^7$ and $x^{10}$,
and a chain that raises its argument to its $8-th$ power.

Module ~\href{../src/html/addition.power.Dichotomy.html}{Powers.Dichotomy} contains a definition of the function \texttt{delta}, and proofs that if 
$n>3$ then $1<\delta(n)<n$.

\subsection{Main chain generation function}
We are now able to define a function that generates a correct chain 
for any signature. We use the \texttt{Recdef} module of Standard Library,
with an appropriate \emph{measure}.

\begin{Coqsrc}
Definition signature_measure (s : signature) : nat :=
match s with
  | gen_F n => 2 * Pos.to_nat n 
  | gen_K p d => 2 * Pos.to_nat (p + d) +1
end.
\end{Coqsrc}

The following function definition generates 9 sub-goals,
for proving that the measure on signatures is strictly decreasing along
the recursive calls. They are solved with the help of Standard Library's lemmas 
on arithemtic of \texttt{positive} numebers end Euclidean division.

%\todo{Is Function (from Recdef) obsolete ? Use Equations instead ? }

\begin{Coqsrc}
Function chain_gen  (s:signature) {measure signature_measure}
:  chain_type s :=
  match s  return chain_type s with
    | gen_F i =>
      if pos_eq_dec i 1 then F1 else
        if pos_eq_dec i 3
        then F3
        else 
          match exact_log2 i with
              Some p => Fexp2 p
            | _ =>
              match N.pos_div_eucl i (Npos (dicho i))
              with
                | (q, 0%N) => 
                  Fcompose  (chain_gen (gen_F (dicho i)))
                            (chain_gen (gen_F (N2Pos q)))
                | (q,r)  => KFF (chain_gen
                                   (gen_K (N2Pos r)
                                          (dicho i - N2Pos r)))
                                (chain_gen (gen_F (N2Pos q)))

              end end
\end{Coqsrc}

\begin{Coqsrc}
    | gen_K p d =>
      if pos_eq_dec p 1 then FK (chain_gen (gen_F (1 + d)))
      else
        match N.pos_div_eucl (p + d)  (Npos p) with
          | (q, 0%N) => FFK   (chain_gen (gen_F p))
                              (chain_gen (gen_F (N2Pos q)))
          | (q,r)  => KFK (chain_gen (gen_K (N2Pos r)
                                            (p - N2Pos r)))
                          (chain_gen (gen_F (N2Pos q)))
        end
  end.
(* A lot of arithmetic proofs omitted *)
Defined.

Definition make_chain (n:positive) : chain :=
 F2C (chain_gen (gen_F n)).

\end{Coqsrc}

Thanks to the \texttt{Recdef} package, we are now able to get automatically 
built chains using the dichotomic strategy.

\begin{Coqsrc}
Compute make_chain 87.
\end{Coqsrc}

\begin{Coqanswer}
 =  fun (A : Type) (x : A) =>
       x0 <--- x times x;
       x1 <--- x0 times x;
       x2 <--- x1 times x1;
       x3 <--- x2 times x;
       x4 <--- x3 times x1;
       x5 <--- x4 times x4;
       x6 <--- x5 times x5;
       x7 <--- x6 times x6; 
       x8 <--- x7 times x3; 
       Return x8
     : chain 
\end{Coqanswer}

\subsubsection{Correctness of the Euclidean chain generator}

\texttt{Recdef}'s \texttt{functional induction} tactic allows us to
prove that every value returned by (\texttt{chain\_gen $s$}) is correct w.r.t. 
\texttt{$s$} and proper.
The proof obligations are solved thanks to the previous lemmas on 
the composition operators on chains: \texttt{Fcompose}, \texttt{KFK}, etc.
Unfortunately, a lot of interaction is still needed or proving properties of
Euclidean division and binary logarithm.

\begin{Coqsrc}
Lemma chain_gen_OK : forall s:signature, OK  s (chain_gen  s).
intro s; functional induction chain_gen s.
Proof.
(* A lot of arithmetic proofs omitted *)

Theorem make_chain_correct :
    forall p, chain_correct (make_chain p) p.
Proof.
 intro p; destruct (chain_gen_OK (gen_F p)).
 unfold make_chain; apply F2C_correct; apply H.
Qed.
\end{Coqsrc}

\subsubsection{A last example}
\label{ex45319}

Let us compute  $67777^{6145319}$ with 32 bits integers:

\begin{Coqsrc}

Ltac compute_chain ch := 
   let X := fresh "x" in 
   let Y := fresh "y" in
   let X := constr:ch in 
   let Y := (eval vm_compute in X) in 
   exact Y.

Let big_chain := ltac:(compute_chain  (make_chain 6145319)).

Print big_chain.
\end{Coqsrc}

\begin{Coqanswer}
big_chain = 
fun (A : Type) (x : A) =>
x0 <--- x times x; x1 <--- x0 times x0;
x2 <--- x1 times x1; x3 <--- x2 times x1;
x4 <--- x3 times x3; x5 <--- x4 times x;
x6 <--- x5 times x5; x7 <--- x6 times x6;
x8 <--- x7 times x1; x9 <--- x8 times x5;
x10 <--- x9 times x8; x11 <--- x10 times x9;
x12 <--- x11 times x11; x13 <--- x12 times x11;
x14 <--- x13 times x10; x15 <--- x14 times x14;
x16 <--- x15 times x11; x17 <--- x16 times x16;
x18 <--- x17 times x17; x19 <--- x18 times x18;
x20 <--- x19 times x19; x21 <--- x20 times x20;
x22 <--- x21 times x21; x23 <--- x22 times x22;
x24 <--- x23 times x23; x25 <--- x24 times x24; 
x26 <--- x25 times x25; x27 <--- x26 times x26; 
x28 <--- x27 times x14;  Return x28
     : forall A : Type, A -> computation

\end{Coqanswer}

\begin{Coqsrc}
Time   Compute  Int31.phi 
     (chain_apply big_chain (snd (positive_to_int31  67777))).
\end{Coqsrc}
\begin{Coqanswer}
= 2014111041%Z
     : Z
Finished transaction in 0.005 secs (0.005u,0.s) (successful)}  
\end{Coqanswer}

\begin{Coqsrc}
Compute chain_length big_chain.
\end{Coqsrc}

\begin{Coqanswer}
= 29%nat
     : nat  
\end{Coqanswer}

\section{Projects}

\index{Projects}
\begin{project}[Optimality and relative efficiency]

\vspace{3pt}

\noindent

\begin{enumerate}
\item  Prove that the chain generated by \texttt{Fexp2} is optimal.
\item Prove that  the length of any optimal chain for $n$ is
greater than or equal to $\floor{\log_2{n}}$.
\item Prove that, for any positive $n$, the length of any Euclidean chain generated by the 
  dichotomic strategy  is always less than or equal to
  the length of \texttt{binary\_chain $n$}, and for an infinite number
of positive integers $n$, the first chain  is strictly shorter
than  the latter.
\item Prove that our implementation of the dichotomic strategy describes
 the same function as in the litterature (for instance ~\cite{DBLP:journals/ita/BrlekCHM95}.)
This is important if we want to follow the complexity analyses in this and similar articles.
\item Study how to \emph{compile} a chain into imperative code, using a register allocation strategy (it may be useful  to define \emph{chain width} ).

\paragraph*{Remark:} The first two questions of the list above should involve a 
universal quantification on type
\texttt{chain}. It may be necessary (but we're not sure) to consider  some 
restriction on parametric chains.

\end{enumerate}
\end{project}

%\subsection{Optimality and Relative Efficiency}

% \begin{project}[Proof techniques]
% \noindent

% \begin{enumerate}
% \item Improve automated proofs on types \texttt{positive} and \texttt{N}.
% \item Compare  \texttt{Program Fixpoint} and \texttt{Function} for
% writing \texttt{make\_chain}. Consider measure \emph{vs} well-founded 
% relations, mutual recursion, possibility of using sigma-types, etc.
% \item Chains are always associated with strictly positive exponents. 
% Thus, many lemmas about chain correctness  can be proved using semi-groups instead of
% monoids. Define type classes for semi-groups and use them whenever possible.
% \end{enumerate}  
% \end{project}

% \section{Exponentiation in \coq's standard library}

% Exponentiation is already defined for several types in Standard Library\footnote{The following information was checked according to \texttt{8.5beta2} version of \coq{}.}

% \subsection{Peano numbers}
% Exponentation is defined in \texttt{Coq.Init.Nat} as a binary operation on type nat. Basically, it is a structurally recursive definition with respect 
% to its second argument.

% \begin{verbatim}
% Fixpoint pow n m :=
%   match m with
%     | 0 => 1
%     | S m => n * (n ^ m)
%   end

% where "n ^ m" := (pow n m) : nat_scope.
% \end{verbatim}

% \subsection{Binary positive numbers}
% Exponentiation is defined in \texttt{Coq.PArith.BinPosDef} as a binary operation on type \texttt{positive}.
% Its definition uses a polymorphic functional that allows to iterate 
% a unary function on any type $A$.

% \begin{verbatim}
% Definition iter {A} (f:A -> A) : A -> positive -> A :=
%   fix iter_fix x n := match n with
%     | xH => f x
%     | xO n' => iter_fix (iter_fix x n') n'
%     | xI n' => f (iter_fix (iter_fix x n') n')
%   end.

% (** ** Power *)

% Definition pow (x:positive) := iter (mul x) 1.

% Infix "^" := pow : positive_scope.
% \end{verbatim}

% Although the definition of \texttt{iter} takes advantage of the binary
% representation of \texttt{n}, the computation of some $x^n$ amounts
% to $n$ iterations of the function \texttt{(mul x)}. Thus, the function
% \texttt{Pos.pow} is linear with respect to the exponent.

% \begin{alltt}
% Time Compute Pos.pow 1\%positive 56666667\%positive.\it
% Finished transaction in 5.679 secs (5.693u,0.s) (successful)
% \end{alltt}

% \subsection{Binary integers}

% Types \texttt{N} [resp. \texttt{Z}] are also provided with binary operations
% that allow to compute $x^n$ (for $x,n$ of type \texttt{N} [resp.   \texttt{Z}].

% In \texttt{Coq-8.5beta2}, the functions \texttt{Coq.NArith.BinNat.N.pow}
% and \linebreak \texttt{Coq.ZArith.BinInt.Z.pow} are also defined as iterations
% of \texttt{(mul x)} using the same iteration functional as for the \texttt{positive} type. Thus they are also linear with respect to the exponent.

### 2020-11-15 doc/hydras.tex
restructuration of latex files

--- 2020-11-08 doc/hydras.tex
@@
 
-\author{Pierre Castéran\\ LaBRI, Univ. Bordeaux, CNRS UMR 5800, Bordeaux-INP\\with contributions by \'Evelyne Contejean,  Florian Hatat and Pascal Manoury}
+\author{Pierre Castéran\\ LaBRI, Univ. Bordeaux, CNRS UMR 5800, Bordeaux-INP\\with contributions by Yves Bertot, \'Evelyne Contejean,  Florian Hatat and Pascal Manoury}
 \date{\today}
-\title{Hydra Battles in Coq}
+\title{Hydra Battles and Cie}
 \makeindex
@@
 
-Among all the theorems proved with the help of proof assistants like \coq{}, \isabelle{}, \hol{}, etc.,
+Among all the theorems proved with the help of proof assistants like \coq{}, \isabelle{}~\cite{isabelle}, \hol{}~\cite{HOL}, etc.,
 several statements and proofs  share some interesting features:
@@
 
-Hydra games (a.k.a. \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians:
+In this document, we present two examples which seem to have the above properties.
+
+\begin{itemize}
+\item Hydra games (a.k.a. \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians:
 L. Kirby and J. Paris~\cite{KP82}: \emph{Accessible Independence Results for Peano Arithmetic}. 
@@
 
-Here, we present a development  for the \coq{} proof assistant, after the work of Kirby and Paris. This formalization contains the following main parts:
-
-\begin{itemize}
-\item Representation in \coq{} of hydras and hydra battles
-\item A proof that every battle is finite and won by Hercules.
-This proof is based on a \emph{variant} which maps any hydra to an ordinal strictly less than $\epsilon_0$ and is strictly decreasing along any battle.
-
-\item Using a combinatorial toolkit designed by J.~Ketonen and R.~Solovay~\cite{KS81}, we prove that, for any ordinal $\mu<\epsilon_0$, there exists no such variant mapping any hydra to an ordinal stricly less than $\mu$.
-Thus, the complexity of $\epsilon_0$ is really needed in the previous proof.
-
-\item We prove a relation between the length of a ``classic''  class of  battles \footnote{This class is also called \emph{standard} in this document (text and proofs).
-The \emph{replication factor} of the hydra is exactly $i$ at the $i$-th round of the battle (see Sect~\vref{sect:replication-def}).}
-and the Wainer-Hardy hierarchy of ``rapidly growing functions'' $H_\alpha$~\cite{Wainer1970}.
-The considered class of battles, which we call \emph{standard}  is the most considered one in the scientific  litterature(including popularization).
+\item In the second part, we are interested in computing $x^n$ with the least number of multiplications as possible.
+We use the notion of \emph{addition chains}~\cite{brauer1939,DBLP:journals/ipl/BerstelB87}, to generate efficient certified exponentiation functions.
 \end{itemize}
 
-Simply put, this document tries to combines the scientific interest of two articles~\cite{KP82, KS81} and a book~\cite{schutte} with the playful activity of truly proving theorems.
-We hope  that such a work, besides exploring a nice piece of discrete maths, 
-will show how \coq{} and its standard library are well fitted to help us to understand some non-trivial mathematical developments, and also to experiment the constructive parts of  the proof through functional programming.
-
- We also hope to provide a little clarification on infinity (both potential and actual) through the notions of function, computation, limit,
- types and proofs.
-
-\section{Remarks}
-
-\subsection{Difference from Kirby and Paris's Work}
-In~\cite{KP82}, Kirby and Paris showed  that there is no proof of termination of all hydra battles in Peano Arithmetic (PA).
-Since we are used to writing proofs in higher order logic, the restriction to PA was quite unnatural for us.
-So we chosed to prove another statement without any reference to PA, by considering a class of proofs indexed by ordinal numbers upto $\epsilon_0$.
-
-\subsection{Trust in our Proofs?}
+%%%%% ICI
+
+\subsection{Trust in our Proofs}
 \label{sect:trust-in-proofs}
@@
 
-\subsection{Main References}
-
-In our development, we adapt the definitions and prove many theorems which
-we found in the following articles. 
-\begin{itemize}
-\item ``Accessible independence results for Peano arithmetic''  by Laurie Kirby and Jeff Paris~\cite{KP82}
-\item ''Rapidly growing Ramsey Functions'' by Jussi Ketonen and Robert Solovay~\cite{KS81}
-\item ``The Termite and the Tower'', by Will Sladek~\cite{Sladek07thetermite}
-\item Chapter V of ``Proof Theory'' by Kurt Schütte~\cite{schutte}
-\end{itemize}
-
 \paragraph*{Warning:}
@@
 
-\subsection{State of the Development}
-The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on  notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
-We added new material : axiomatic definition of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, equations, etc.
-
-We are now working in order to make clumsy proofs more readable, simplify definitions, and ``factorize'' proofs as much as possible. 
-Many possible improvements are suggested as ``todo''s or ``projects'' in this text.
-
 \subsection{Contributions}
-Many thanks to \'Evelyne Contejean,  Florian Hatat and Pascal Manoury  for their contribution.
+Many thanks to Yves Bertot, 
+\'Evelyne Contejean,  Florian Hatat and Pascal Manoury  for their contribution. 
+Yves made nice optimizations  to algorithms presented in Chapter~\ref{chapter-powers}.
 \'Evelyne contributed libraries on the recursive path ordering (\emph{rpo}) for proving the well-foundedness of our representation of $\epsilon_0$ and $\Gamma_0$.
@@
 \label{sec:orgheadline5}
-    Many thanks to David Ilcinkas, Sylvain Salvati, Alan Schmitt and Théo Zimmerman for their help on the elaboration of this document, and to the
+    Many thanks to Yves Bertot, David Ilcinkas, Sylvain Salvati, Alan Schmitt and Théo Zimmerman for their help on the elaboration of this document, and to the
  members of the \emph{Formal Methods} team at laBRI for their helpful comments 
@@
 \begin{itemize}
-\item The present distribution has been checked with version 8.11.0 of the Coq proof assistant, with the plug-ins \texttt{coq-paramcoq} and \texttt{coq-equations}.
-
-\item Just go into the root directory, and type "\texttt{make}"
+\item The present distribution has been checked with version 8.12.0 of the Coq proof assistant, with the plug-ins \texttt{coq-paramcoq}, \texttt{coq-equations}
+and \texttt{coq-mathcomp-algebra}.
+
+\item Just go into the root directory, and type "\texttt{make}" ("\texttt{make -k}"  in case of an incomplete installation).
 \end{itemize}
@@
 
-\section{Future Work (projects)}
-\index{Projects}
-
-This document and the proof scripts are far from being complete.
-
-First, there must be a lot of typos to correct, references and index items to add. Many proofs are too complex and should be simplified, etc.
-
-The following extensions are planned, but help is needed:
-
-\begin{itemize}
-\item Semi automatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
-\item Extension to $\Gamma_0$ (in Veblen normal form)
-\item More lemmas about hierarchies of rapidly growing functions, and their relationship 
-    with primitive recursive functions and provability in Peano arithmetic 
-(following~\cite{KS81, KP82}).
-\item From \coq's point of view, this development could be used as an illustration of the evolution of the software, every time new libraries and sets of tactics could help to simplify the proofs.
-\end{itemize}
-
-%-----------------------------------------------------------
-
-\chapter{Hydras and Hydra Games}
-\label{sec:orgheadline91}
-
-This chapter is dedicated to the representation of hydras and rules of the hydra game in \coq's specification language:~\gallina.
-
-Technically, a \emph{hydra} is just a finite ordered tree, each node of which 
-has any number of sons. Note that, contrary to the computer science tradition, we will show the hydras 
-with the heads up and the foot (i.e., the root of the tree) down.
-Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the 
-module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}).
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {$\bullet$};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H4) at (4,2) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N3) to [bend left= 10]  (H1) ;
-\draw (N3) to [bend right= 16] (H2);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (foot) to [bend right = 10] (H4) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-\caption{The hydra Hy \label{fig:Hy}}
-\end{figure}
-
-We use a specific vocabulary for talking about hydras. Table~\ref{tab:hyd2tree} shows the correspondance between our terminology and the usual vocabulary for trees in computer science.
-
-\begin{figure}[h]
-  \centering
-  \begin{tabular}{ll}
-Hydras & Finite rooted trees\\
-\hline
-foot & root\\
-head & leaf\\
-node & node\\
-segment  & (directed) edge \\
-sub-hydra & subtree\\
-daughter & immediate subtree\\
-\end{tabular}
-  \caption{Translation from hydras to trees}
-  \label{tab:hyd2tree}
-\end{figure}
-
-The hydra \texttt{Hy} has a \emph{foot} (below), five \emph{heads}, and eight \emph{segments}. 
-We leave it to the reader to define various parameters such as the height, the size, the highest arity (number of sons of a node) of a hydra.
-In our example, these parameters have the respective values : $4$, $9$ and $3$.
-
-\subsection{The Rules of the Game}
-
-\label{sec:orgheadline44}
-\label{sect:replication-def}
-
-A \emph{hydra battle} is a fight between Hercules and the Hydra. 
-More formally, a  battle is a sequence of \emph{rounds}.
-At each round:
-\begin{itemize}
-\item If the hydra is composed of just one head, the battle is finished
-and  Hercules is the winner.
-\item Otherwise, Hercules chops off \emph{one} head of the hydra,
-
-\begin{itemize}
-\item If the head is at distance 1 from the foot, the head is just lost by the hydra, with no more reaction.
-\item Otherwise, let us denote by \(r\) the node that was at distance \(2\) from 
-the removed head in the direction of the foot,  and consider the  sub-hydra \(h'\) of \(h\), whose  root is \(r\) \footnote{$h'$ will be called ``the wounded part of the hydra'' in the subsequent text.
-In Figures~\vref{fig:Hy2} and ~\vref{fig:Hy4}, this sub-hydra  is displayed in red.}.
-Let $n$ be some natural number.
-Then $h'$ is replaced by  $n+1$ of copies of \(h'\) which share the same root $r$.
- The \emph{replication factor} $n$ may be different (and generally is)   at each round of the fight.
-It may be chosen by the hydra, according to its strategy, or imposed by some 
-particular rule.
-In many presentations of hydra battles, this number is increased by $1$ at each round.
-In the following presentation, we will also consider battles where the hydra is free to choose its ~replication factor at each round of the battle\footnote{Let us recall that, if the chopped-off head was at distance 1 from the foot, the replication factor is meaningless.}.
-\end{itemize}
-\end{itemize}
-
-Note that the description given in~\cite{KP82} of the replication process in hydra battles is also  semi-formal.
-
-\label{original-rules}
-
-\begin{quote}
-  ``From the node that used to be attached to the head which was just chopped off, traverse one 
-segment towards the root until the next node is reached. From this node sprout $n$ replicas of 
-that part of the hydra (after decapitation) which is ``above'' the segment just traversed, i.e., those 
-nodes and segments from which, in order to reach the root, this segment would have to be 
-traversed. If the head just chopped off had the root of its nodes, no new head is grown. ''
-\end{quote}
-
-Moreover, we note that this description is in \emph{imperative} terms.
-In order to build a formal  study of the properties of hydra battles, we prefer to use a mathematical vocabulary, i.e., graphs, relations, functions, etc.
-Thus, the replication process will be represented as a binary relation on a data type \texttt{Hydra},
-linking the state of the hydra \emph{before} and \emph{after} the transformation.
-A battle will thus be represented as a sequence of terms of type \texttt{Hydra}, respecting the rules of the game.
-
-\subsection{Example}
-Let us start a battle between Hercules and the hydra \texttt{Hy} of Fig.~\ref{fig:Hy}.
-
-At the first round, Hercules choses to chop off the rightmost head of \texttt{Hy}.
-Since this head is near the floor, the hydra loses this head. Let us call 
- \texttt{Hy'} the resulting state of the hydra, represented in Fig.~\vref{fig:Hy-prime}.
-
-Next, assume Hercules choses to chop off one of the two highest heads of \texttt{Hy'}, for instance the rightmost one.
-Fig.~\vref{fig:Hy2} represents the rotten neck in dashed lines, and the part that will be replicated in red.
-Assume also that the hydra decides to add 4 copies of the red part\footnote{In other words, the replication factor at this round is equal to $4$.}.
-We obtain a new state \texttt{Hy''} depicted in Fig.~\ref{fig:Hy3}.
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {$\bullet$};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-%\node (H4) at (6,0) {$\Xey[2][lightgray]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3) ;
-\draw (N3) to [bend left= 10]  (H1) ;
-\draw (N3) to [bend right= 16] (H2);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-
-\caption{Hy': the state  of Hy after one round \label{fig:Hy-prime}}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4)  {{\color{lightred}$\bullet$}};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Sey[2][lightred]$};
-%\node (H2) at (5,0) {$\Xey[2][lightgray]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\node (ex) at (5,8) {};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw  (N2) -- (N3)[draw=lightred];
-\draw (N3) to   [bend left= 10](H1) [draw=lightred];
-\draw [dashed] (N3) to [bend left= 10](ex);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-\caption{A second beheading}
-\label{fig:Hy2}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H11) at (2,8) {$\Smiley[2][vertfluo]$};
-\node (H12) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H13) at (6,8) {$\Smiley[2][vertfluo]$};
-\node (H14) at (8,8) {$\Smiley[2][vertfluo]$};
-
-\node (N3) at (1,6) {$\bullet$};
-\node (N31) at (2,6) {$\bullet$};
-\node (N32) at (3,6) {$\bullet$};
-\node (N33) at (4,6) {$\bullet$};
-\node (N34) at (5,6) {$\bullet$};
-
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N2) -- (N31);
-\draw (N2) -- (N32);
-\draw (N2) -- (N33);
-\draw (N2) -- (N34);
-\draw (N3) to   [bend left= 10](H1) ;
-\draw (N31) to   [bend left= 10](H11) ;
-\draw (N32) to   [bend left= 10](H12) ;
-\draw (N33) to   [bend left= 10](H13) ;
-\draw (N34) to   [bend left= 10](H14) ;
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend left= 10]  (H5) ;
-\end{tikzpicture}
-\caption{Hy'', the state of Hy after two rounds \label{fig:Hy3}}
-\end{figure}
-
-Figs.~\ref{fig:Hy4} and~\vref{fig:Hy5} represent a possible third round of the battle, with a replication factor equal to $2$. Let us call \texttt{Hy'''} the state of the hydra after that third round.
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0)  {{\color{lightred}$\bullet$}};
-\node (N1) at (2,2) {{\color{lightred}$\bullet$}};
-\node (N2) at (2,4) {{\color{lightred}$\bullet$}};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (exN4) at (4,4) {};
-\node (H1) at (0,8) {$\Sey[2][lightred]$};
-\node (H11) at (2,8) {$\Sey[2][lightred]$};
-\node (H12) at (4,8) {$\Sey[2][lightred]$};
-\node (H13) at (6,8) {$\Sey[2][lightred]$};
-\node (H14) at (8,8) {$\Sey[2][lightred]$};
-
-\node (N3) at (1,6) {{\color{lightred}$\bullet$}};
-\node (N31) at (2,6) {{\color{lightred}$\bullet$}};
-\node (N32) at (3,6) {{\color{lightred}$\bullet$}};
-\node (N33) at (4,6) {{\color{lightred}$\bullet$}};
-\node (N34) at (5,6) {{\color{lightred}$\bullet$}};
-
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-%\node (H5) at (4,0) {$\Xey[2][lightgray]$};
-\draw (foot) -- (N1)[very thick,draw=lightred] ;
-\draw (N1) -- (N2)[draw=lightred];
-\draw (N2) -- (N3)[draw=lightred];
-\draw (N2) -- (N31)[draw=lightred];
-\draw (N2) -- (N32)[draw=lightred];
-\draw (N2) -- (N33)[draw=lightred];
-\draw (N2) -- (N34)[draw=lightred];
-\draw (N3) to   [bend left= 10](H1) [draw=lightred];
-\draw (N31) to   [bend left= 10](H11) [draw=lightred];
-\draw (N32) to   [bend left= 10](H12) [draw=lightred];
-\draw (N33) to   [bend left= 10](H13) [draw=lightred];
-\draw (N34) to   [bend left= 10](H14) [draw=lightred];
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw [dashed] (N1) to  [bend left= 10](exN4);
-\end{tikzpicture}
-\caption{A third beheading (wounded part in red) \label{fig:Hy4}}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-
-\node (foot) at (10,0) {$\bullet$};
-
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H1) at (0,8) {$\Smiley[1][vertfluo]$};
-\node (H11) at (2,8) {$\Smiley[1][vertfluo]$};
-\node (H12) at (4,8) {$\Smiley[1][vertfluo]$};
-\node (H13) at (6,8) {$\Smiley[1][vertfluo]$};
-\node (H14) at (8,8) {$\Smiley[1][vertfluo]$};
-
-\node (N3) at (1,6) {$\bullet$};
-\node (N31) at (2,6) {$\bullet$};
-\node (N32) at (3,6) {$\bullet$};
-\node (N33) at (4,6) {$\bullet$};
-\node (N34) at (5,6) {$\bullet$};
-
-\node (H0) at (-3,3) {$\Smiley[1][vertfluo]$};
-
-\draw (foot) to [bend left=10] (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N2) -- (N31);
-\draw (N2) -- (N32);
-\draw (N2) -- (N33);
-\draw (N2) -- (N34);
-\draw (N3) to   [bend left= 10](H1) ;
-\draw (N31) to   [bend left= 10](H11) ;
-\draw (N32) to   [bend left= 10](H12) ;
-\draw (N33) to   [bend left= 10](H13) ;
-\draw (N34) to   [bend left= 10](H14) ;
-\draw (foot) to [bend left = 15]  (H0) ;
-
-% second copy 
-\node (N01) at (12,2) {$\bullet$};
-\node (N02) at (12,4) {$\bullet$};
-\node (N03) at (12,6) {{\color{lightred}$\bullet$}};
-\node (H001) at (10,8) {$\Smiley[1][vertfluo]$};
-\node (H0011) at (12,8) {$\Smiley[1][vertfluo]$};
-\node (H0012) at (14,8) {$\Smiley[1][vertfluo]$};
-\node (H0013) at (16,8) {$\Smiley[1][vertfluo]$};
-\node (H0014) at (18,8) {$\Smiley[1][vertfluo]$};
-
-\node (N03) at (11,6) {$\bullet$};
-\node (N031) at (12,6) {$\bullet$};
-\node (N032) at (13,6) {$\bullet$};
-\node (N033) at (14,6) {$\bullet$};
-\node (N034) at (15,6) {$\bullet$};
-
-\draw (foot) -- (N01)[very thick] ;
-\draw (N01) -- (N02);
-\draw (N02) -- (N03);
-\draw (N02) -- (N031);
-\draw (N02) -- (N032);
-\draw (N02) -- (N033);
-\draw (N02) -- (N034);
-\draw (N03) to   [bend left= 10](H001) ;
-\draw (N031) to   [bend left= 10](H0011) ;
-\draw (N032) to   [bend left= 10](H0012) ;
-\draw (N033) to   [bend left= 10](H0013) ;
-\draw (N034) to   [bend left= 10](H0014) ;
-
-% third copy 
-\node (N001) at (22,2) {$\bullet$};
-\node (N002) at (22,4) {$\bullet$};
-\node (N003) at (22,6) {{\color{lightred}$\bullet$}};
-\node (H001) at (20,8) {$\Smiley[1][vertfluo]$};
-\node (H0011) at (22,8) {$\Smiley[1][vertfluo]$};
-\node (H0012) at (24,8) {$\Smiley[1][vertfluo]$};
-\node (H0013) at (26,8) {$\Smiley[1][vertfluo]$};
-\node (H0014) at (28,8) {$\Smiley[1][vertfluo]$};
-
-\node (N003) at (21,6) {$\bullet$};
-\node (N0031) at (22,6) {$\bullet$};
-\node (N0032) at (23,6) {$\bullet$};
-\node (N0033) at (24,6) {$\bullet$};
-\node (N0034) at (25,6) {$\bullet$};
-
-\draw (foot) -- (N001)[very thick] ;
-\draw (N001) -- (N002);
-\draw (N002) -- (N003);
-\draw (N002) -- (N0031);
-\draw (N002) -- (N0032);
-\draw (N002) -- (N0033);
-\draw (N002) -- (N0034);
-\draw (N003) to   [bend left= 10](H001) ;
-\draw (N0031) to   [bend left= 10](H0011) ;
-\draw (N0032) to   [bend left= 10](H0012) ;
-\draw (N0033) to   [bend left= 10](H0013) ;
-\draw (N0034) to   [bend left= 10](H0014) ;
-\end{tikzpicture}
-\caption{The configuration Hy''' of Hy \label{fig:Hy5}}
-\end{figure}
-\FloatBarrier
-
-We leave it to the reader  to guess the following  rounds of the battle \dots
- % Please keep in mind that, in this 
-% the hydra is free to chose any number of replications at each time, whereas
-% Hercules chops only one head per round.
-
-% Let us precise that, in this game, Hercules wins if the hydra is eventually reduced 
-% to a single head. 
-% We know from~\cite{KP82} that, whichever the initial configuration of the
-% hydra, and the strategies of both players, Hercules eventually wins. The 
-% aforementionned paper shows also that there do not exist any \emph{simple} proof of this result.
-
-\section{Hydras and their Representation in \emph{Coq}}
-\label{sec:orgheadline48}
-
-In order to describe trees where each node can have an arbitrary (but finite) number of sons, it usual to define a type where each node carries a \emph{forest}, \emph{i.e} a list of trees
-(see for instance Chapter 14, pages 400-406 of \cite{BC04}).
-
-For this purpose, we define two mutual \emph{ad-hoc}  inductive types, where \texttt{Hydra} is the main type, and \texttt{Hydrae} is a helper for describing finite sequences of hydra.
-\label{types:Hydra}
-\label{types:Hydrae}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hydra}{Hydra.Hydra\_Definitions}}
-
-\begin{Coqsrc}
-Inductive Hydra : Set :=
-| node :  Hydrae -> Hydra
-with Hydrae : Set :=
-| hnil : Hydrae
-| hcons : Hydra -> Hydrae -> Hydrae.
-\end{Coqsrc}
-
-%\index{To do}
-\index{Projects}
-
-\begin{project}
-Look for an existing library on trees with nodes of arbitrary arity, in order to replace  this ad-hoc type with something more generic.
-\end{project}
-
-\index{Projects}
-
-\begin{project}
-
- Another very similar representation could use the \texttt{list} type family instead of the specific 
-type \texttt{Hydrae}:
-
-\begin{Coqalt}
-Module Alt.
-
-Inductive Hydra: Set :=
-  hnode (daughters : list Hydra).
-
-End Alt.
-\end{Coqalt}
-
-Using this representation, re-define all the constructions of this chapter.
-You will probably have to use patterns described for instance in~\cite{BC04} or the archives of the Coq-club~\cite{Coq}.
-
-\end{project}
-
-\index{Projects}
-
-\begin{project}
-The type \texttt{Hydra} above describes hydras as \emph{plane trees}, i.e., as drawn on a sheet of paper or computer screen. Thus, hydras are \emph{oriented},
-and it is appropriate to consider a \emph{leftmost} or \emph{rightmost} head of
-the beast. It could be interesting to consider another representation, in which
-every non-leaf node has a \emph{multi-set} -- not an ordered list -- of daughters.
-\end{project}
-
-\subsubsection{Abbreviations}
-
-We provide several notations for hydra patterns  which occur often in our developments.
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#head}{Hydra.Hydra\_Definitions}}
-
-\begin{Coqsrc}
-(** heads *)
-Notation head := (node hnil).
-
-(** nodes  with 1, 2 or 3 daughters *)
-Notation hyd1 h := (node (hcons h hnil)).
-Notation hyd2 h h' := (node (hcons h (hcons h' hnil))).
-Notation hyd3 h h' h'' := 
-                   (node (hcons h (hcons h' (hcons h'' hnil)))).
-\end{Coqsrc}
-
-For instance, the hydra \texttt{Hy}  of Figure~\vref{fig:Hy} is defined in \emph{Gallina} as follows:
-
-\vspace{4mm}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html\#Hy}{Hydra.Hydra\_Examples}}
-
-\begin{Coqsrc}
-Example Hy := hyd3 head
-                   (hyd2
-                      (hyd1 
-                         (hyd2 head head))
-                      head) 
-                   head.
-\end{Coqsrc}
-
-Hydras quite frequently contain  multiple  copies of the same pattern. The following functions
-will help us to describe and reason about replications in hydra battles.
-
-\vspace{4pt}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#hcons_mult}{Hydra.Hydra\_Definitions}}
-
-\begin{Coqsrc}
-Fixpoint hcons_mult (h:Hydra)(n:nat)(s:Hydrae):Hydrae :=
-  match n with 
-  | O => s
-  | S p => hcons h (hcons_mult h p s)
-  end.
-
-(** hydra with n copies of the same daughter *)
-
-Definition hyd_mult h n :=
-  node (hcons_mult h n hnil).
-\end{Coqsrc}
-
-\vspace{4mm}
-
-For instance, the hydra $Hy''$ of Fig~\vref{fig:Hy3}  can be defined in \coq{} as follows:
-
-\vspace{4pt}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\begin{Coqsrc}
-Example Hy'' := 
-     hyd2 head
-          (hyd2 (hyd_mult (hyd1 head) 5)
-                head).
-\end{Coqsrc}
-
-\subsubsection{Recursive Functions on type \texttt{Hydra}}
-\label{sec:orgheadline41}
-\label{sec:hsize-def}
-
-When defining a recursive function over the type \texttt{Hydra}, one has to consider the three constructors 
-\texttt{node}, \texttt{hnil} and \texttt{hcons} of the mutually inductive types \texttt{Hydra} and \texttt{Hydrae}. 
-Let us define for instance the function which  computes the number of nodes of any hydra:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\begin{Coqsrc}
-Fixpoint hsize (h:Hydra) : nat :=
-  match h with node l => S (lhsize l)
-  end
-with lhsize l : nat :=
-  match l with hnil => 0
-            | hcons h hs => hsize h + lhsize hs 
-  end.
-
- Compute hsize Hy.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = 9
-     : nat 
-\end{Coqanswer}
-
-Likewise, the \emph{height} (maximum distance between the foot and a head) 
-is defined by mutual recursion:
-
-\begin{Coqsrc}
-Fixpoint height  (h:Hydra) : nat :=
-  match h with node l => lheight l
-  end
-with lheight l : nat :=
-  match l with 
-  | hnil => 0
-  | hcons h hs => Max.max (S (height h)) (lheight hs)
-  end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Compute height Hy.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = 4
-     : nat  
-\end{Coqanswer}
-
-\index{Exercises}
-
-\begin{exercise}
-Define a function \texttt{max\_degree: Hydra $\arrow$ nat} which  returns the highest degree of a node in any hydra.
-For instance, the evaluation of the term \texttt{(max\_degree Hy)} should return $3$.
-\end{exercise}
-
-\subsection{Induction Principles for Hydras}
-\label{sec:orgheadline42}
-
-In this section, we show how induction principles are used to prove properties on the type 
-\texttt{Hydra}. Let us consider for instance the following statement:
-\begin{quote}
-  `` The height of any hydra is strictly less than its size. ''
-\end{quote}
-
-\subsubsection{A failed Attempt}
-
-One may try to use the default tactic of proof by induction, which corresponds to an application of the automatically  generated  induction principle for  type \texttt{Hydra}:
-
-\begin{Coqanswer}
-Hydra_ind :
-forall P : Hydra -> Prop,
-(forall h : Hydrae, P (node h)) -> forall h : Hydra, P h
-\end{Coqanswer}
-
-Ler us start a simple proof by induction.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\begin{Coqbad}
-Module Bad.
-
-Lemma height_lt_size (h:Hydra) :
-  height h <= hsize h.
-Proof.
-  induction h as [s].
-\end{Coqbad}
-
-\begin{Coqanswer}
-1 subgoal, subgoal 1 (ID 11)
-
-  s : Hydrae
-  ============================
-   height (node s) <= hsize (node s)
-\end{Coqanswer}
-
-We might be tempted to do an induction on the sequence \texttt{s}:
-
-% \begin{Coqbad}
-% induction  s as [ | h s'].
-%   -   cbn; auto with arith.
-%   - 
-% \end{Coqbad}
-
-\begin{Coqanswer}
- 1 focused subgoal
-(unfocused: 0), subgoal 1 (ID 19)
-
-  h : Hydra
-  s' : Hydrae
-  IHs' : height (node s') <= hsize (node s')
-  ============================
-   height (node (hcons h s')) <= hsize (node (hcons h s'))
-
-\end{Coqanswer}
-
-Note that the displayed subgoal does not contain any assumption on \texttt{h}, thus there is no way to 
-infer any property about the height and size of the hydra \texttt{(hcons h t)}.
-
-\begin{Coqbad}
-Abort.
-
-End Bad.
-\end{Coqbad}
-
-\subsubsection{A Principle of Mutual Induction}
-In order to get an appropriate induction scheme for the types 
-\texttt{Hydra} and \texttt{Hydrae}, we can use  \coq{}'s  command \texttt{Scheme}.
-
-\index{Coq!Techniques!Mutually inductive types}
-\index{Coq!Commands!Scheme}
-
-\begin{Coqsrc}
-Scheme Hydra_rect2 := Induction for Hydra Sort Type
-with Hydrae_rect2 := Induction for Hydrae Sort Type.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Check Hydra_rect2.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-Hydra_rect2
- : forall (P : Hydra -> Type) (P0 : Hydrae -> Type),
-   (forall h : Hydrae, P0 h -> P (node h)) ->
-    P0 hnil ->
-    (forall h : Hydra, P h -> 
-             forall h0 : Hydrae, P0 h0 -> P0 (hcons h h0)) ->
-    forall h : Hydra, P h
-\end{Coqanswer}
-
-\subsubsection{A Correct Proof}
-
-Let us now use \texttt{Hydra\_rect2} for proving that the height of any hydra is strictly less than its size.
-Using this scheme requires an auxiliary predicate, called \texttt{P0} in \texttt{Hydra\_rect2}'s statement. 
-Let us begin by defining an ad-hoc version of \texttt{List.Forall}.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\begin{Coqsrc}
-(** All elements of s satisfy P *)
-
-Fixpoint h_forall (P: Hydra -> Prop) (s: Hydrae) :=
-  match s with
-      hnil => True
-    | hcons h s' => P h /\ h_forall P s'
-  end.
- \end{Coqsrc}
-
- \begin{Coqsrc}
-Lemma height_lt_size (h:Hydra) :
- height h < hsize h.
-Proof.
-  induction h using Hydra_rect2  with 
-  (P0 :=  h_forall (fun h =>  height h < hsize h)).
- \end{Coqsrc}
-
-\begin{enumerate}
-\item The first subgoal is as follows:
-
-\begin{Coqanswer}
-
-  h: Hydrae
-  IHh : h_forall (fun h : Hydra => height h < hsize h) h
-  ============================
-   height (node s) < hsize (node s)
-
-\end{Coqanswer}
-
-This goal is easily solvable, using some arithmetic. We let the reader look at the source.
-
-\item The second subgoal is trivial:
-
-\begin{Coqanswer}
-
-  ============================
-    h_forall (fun h : Hydra => height h < hsize h) hnil
-
-\end{Coqanswer}
-
-\begin{Coqsrc}
-  reflexivity.
-\end{Coqsrc}
-
-\item Finally, the last subgoal is also easy to solve:
-
-\begin{Coqanswer}
-  h : Hydra
-  h0 : Hydrae
-  IHh : height h < hsize h
-  IHh0 : h_forall (fun h : Hydra => height h < hsize h) h0
-  ============================
-   h_forall (fun h1 : Hydra => height h1 < hsize h1) 
-                 (hcons h h0)
-\end{Coqanswer}
-
-\begin{Coqsrc}
- split;auto. 
-Qed.   
-\end{Coqsrc}
-
-\end{enumerate}
-
-\index{Exercises}
-
-\begin{exercise}
-It happens very often that, in the proof of  a proposition of the form 
-(\texttt{$\forall\,$ h:Hydra, $P$ h}), the predicate \texttt{P0}
-is  (\texttt{h\_forall $P$}).
-Design a tactic for induction on hydras that frees the user from binding explicitly \texttt{P0},  and solves trivial subgoals.
-Apply it for writing  a shorter proof of \texttt{height\_lt\_size}.
-\end{exercise}
-
-\section{Relational Description of Hydra Battles}
-
-In this section, we represent the rules of hydra battles as a binary relation associated with
-a \emph{round}, i.e., an interaction composed of the two following actions:
-\begin{enumerate}
-\item Hercules chops off one head of the hydra
-\item Then, the  hydra replicates the wounded part (if the head is at distance $\geq 2$ from the foot).
-\end{enumerate}
-The relation associated with each round of the battle is parameterized  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot,
-but present for consistency's sake).
-
-In our description,  we will apply the following naming convention: if $h$ represents the configuration of the hydra before a round, then the configuration of $h$ after this round will be called $h'$.
- Thus, we are going to define a proposition  (\texttt{round\_n $n\;h\;h'$})  whose intended meaning will be `` the hydra $h$  is transformed into $h'$  in a single round of a battle, with the expected replication factor $n$ ''.
-
-Since the replication of parts of the hydra depends on the distance of the chopped head from  the foot, we  decompose our description into two main  cases, under the form of a bunch of [mutually] inductive predicates over the types \texttt{Hydra} and \texttt{Hydrae}.
-
-The mutually exclusive cases we consider are the following:
-\begin{itemize}
-\item \textbf{R1}: The chopped off head was at distance 1 from the foot.
-\item \textbf{R2}: The chopped off head was at a distance greater than or equal to  $2$ from the foot.
-\end{itemize}
-
-\subsection{Chopping off a Head at Distance 1 from the Foot (Relation  R1)}
-
-If Hercules chops off a head near the floor, there is no replication at all. We use an auxiliary 
-predicate \texttt{S0}, associated with the removing of one head from a sequence of hydras.
-
-\vspace{4pt}\emph{From Module\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\begin{Coqsrc}
-Inductive S0 :  relation Hydrae :=
-| S0_first : forall s, S0  (hcons head s) s
-| S0_rest : forall  h s s',  S0  s s' ->
-                             S0  (hcons h s) (hcons h s').
-
-Inductive R1  :  Hydra -> Hydra -> Prop :=
-| R1_intro : forall s s', S0 s s' -> R1 (node s) (node s').
-\end{Coqsrc}
-
-\subsubsection{Example}
-\label{sec:orgheadline45}
-
-Let us represent in \coq{}   the transformation of the hydra of Fig.~\vref{fig:Hy} into
-the configuration represented in Fig.~\ref{fig:Hy-prime}.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\begin{Coqsrc}
-Example Hy_1 : R1 Hy Hy'.
-Proof. 
-  split; right; right; left.
-Qed.
-\end{Coqsrc}
-
-\subsection{Chopping off a Head at Distance \texorpdfstring{$\geq 2$}{>= 2} from the Foot (relation R2) }
-
-Let us now consider beheadings  where the chopped-off head is at distance greater than or equal to $2$ from the foot. All the following relations are parameterized by the replication factor  $n$.
-
- Let $s$ be a sequence of hydras. 
-The proposition (\texttt{S1 n s s'}) holds if $s'$ is obtained by replacing some element $h$ of $s$ by 
-$n+1$ copies of $h'$, where  the proposition (\texttt{R1 h h'}) holds, in other words, $h'$ is just $h$, without the chopped-off  head.
-\texttt{S1} is an inductive relation with two constructors that allow us to choose the position in $s'$ of the wounded sub-hydra $h$.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#S1}{Hydra.Hydra\_Definitions}}
-
-\begin{Coqsrc}
-Inductive S1 (n:nat)  : Hydrae -> Hydrae -> Prop :=
-| S1_first : forall s h h' ,   R1 h h' -> 
-                  S1 n (hcons h s) (hcons_mult h' (S n) s)
-| S1_next : forall h s s',  S1 n s s' ->
-                   S1 n (hcons h s) (hcons h s').
-\end{Coqsrc}
-
-The rest of the definition is structured as two mutually inductive relations on hydras and sequences of hydras.
-The first constructor of \texttt{R2} describes the case where the chopped head is exactly at height $2$.
-The others constructors allow us to consider beheadings at height strictly greater than $2$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions}}
-
-\begin{Coqsrc}
-Inductive R2 (n:nat)  :  Hydra -> Hydra -> Prop :=
-| R2_intro : forall s s', S1 n s s' -> R2 n (node s) (node s')
-| R2_intro_2 : forall s s', S2 n s s' -> R2 n (node s) (node s')
-
-with S2 (n:nat) :  Hydrae -> Hydrae -> Prop :=
-|  S2_first : forall h h' s ,
-               R2 n  h h'  -> 
-               S2  n (hcons h s) (hcons h'  s)
-|  S2_next  : forall h   r r',
-               S2 n   r r' ->
-               S2 n (hcons h r) (hcons h r').                  
-\end{Coqsrc}
-
-\subsubsection{Example}
-Let us prove the transformation of \texttt{Hy'} into \texttt{Hy''} (see Fig.~\vref{fig:Hy3}). We use an experimental set of tactics for specifying the place where the 
-interaction between Hercules and the hydra holds.
-
-\vspace{4pt}\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}.
-
-\begin{Coqsrc}
-Example R2_example:  R2 4 Hy' Hy''.
-Proof.
-    (** move to 2nd sub-hydra (0-based indices) *) R2_up 1. 
-    (** move to first sub-hydra *)  R2_up 0.
-    (** we're at distance 2 from the to-be-chopped-off head 
-        let's go to the first daughter, 
-        then chop-off the leftmost head *)  r2_d2  0 0. 
-Qed.
-\end{Coqsrc}
-
-The reader is encouraged to look at all the successive subgoals of this example.
-\emph{Please consider also exercise~\vref{exo:interactive-battle}.}
-
-\subsection{Relation Associated with a Round}
-
-We combine the two cases above into a single relation.
-First,  we define the  relation \texttt{(round\_n n h h')} where \texttt{n} is the expected number of  replications (irrelevant in the case of an \texttt{R1}-transformation).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#round_n}{Hydra.Hydra\_Definitions}}
-
-\index{Predicates!round\_n}
-
-\begin{Coqsrc}
-Definition round_n n h h' := R1 h h' \/ R2 n h h'.  
-\end{Coqsrc}
-
-By abstraction over \texttt{n}, we define a \emph{round} (small step) of a battle:
-
-\index{Predicates!round}
-\label{sect:infix-round}
-\begin{Coqsrc}
-Definition round h h' := exists n,  round_n n h h'.
-
-Infix "-1->" := round (at level 60).
-\end{Coqsrc}
-
-\index{Projects}
-
-\begin{project}
-Give a direct translation of Kirby and Paris's description of hydra battles (quoted on page~\pageref{original-rules}) and prove that our relational description is consistent with theirs.
-\end{project}
-
-\subsection{Rounds and Battles}
-
-Using library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}, we define \texttt{round\_plus},  the transitive closure of \texttt{round}, and \texttt{round\_star},  the reflexive and transitive closure of \texttt{round}.
-
-\label{sect:infix-rounds}
-
-\begin{Coqsrc}
-Definition round_plus := clos_trans_1n Hydra round.
-Infix "-+->" := rounds (at level 60).
-
-Definition round_star h h' := h = h' \/ round_plus h h'.
-Infix "-*->" := round_star (at level 60).
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
-Prove the following lemma:
-
-\begin{Coqsrc}
-Lemma rounds_height : forall h h', 
-   h -+-> h' -> height h' <= height h.  
-\end{Coqsrc}
-
-\end{exercise}
-
-\begin{remark}
-\label{remark:transitive-closure}
-\coq's library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Coq.Relations.Relation\_Operators} 
-contains three logically equivalent definitions of the transitive closure of a binary relation. This equivalence is proved in 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} .
-
-Why three definitions for a single mathematical concept?
-Each definition generates an associated induction principle. 
- According to the form of statement one would like to prove, there is a ``best choice'':
-
-\begin{itemize}
-\item For proving $\forall y, x\,R^+\,y \;\arrow\; P\,y$, prefer 
-\texttt{clos\_trans\_n1}
-\item For proving $\forall x,\,x\,R^+\,y \;\arrow\; P\,x$, prefer \texttt{clos\_trans\_1n}
-\item For proving $\forall x\,y, \,x\,R^+\,y \;\arrow\;P\,x\,y$,  
-prefer \texttt{clos\_trans},
-\end{itemize}
-But there is no ``wrong choice'' at all: the equivalence lemmas in \linebreak 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} 
- allow the user
-to convert any one of the three closures into another one before applying the corresponding elimination tactic.
-The same remark also holds for reflexive and transitive closures. 
-\end{remark}
-
-\index{Exercises}
-
-\begin{exercise}
-Define a restriction of \coqsimple{round},  where Hercules always chops off
-the leftmost among the lowest heads.
-
-Prove that, if $h$ is not a simple head, then there exists a unique $h'$ such that \texttt{h}  is transformed into \texttt{ h'} in one round, according to this restriction.
-
-\end{exercise}
-
-\index{Exercises}
-
-\begin{exercise}[Interactive battles]
-\label{exo:interactive-battle}
-Given a hydra \texttt{h}, the specification of a hydra battle for \texttt{h} is the type 
-\Verb@{h':Hydra | h -*-> h'}@. In order to avoid long sequences of \texttt{split}, \texttt{left}, and 
-\texttt{right}, design a set of dedicated tactics for the interactive building of a battle.
-Your tactics will have the following functionalities:
-\begin{itemize}
-\item  Chose to stop a battle, or continue
-\item Chose an expected number of replications
-\item Navigate in a hydra, looking for a head to chop off.
-\end{itemize}
-
-Use your tactics for simulating a small part of a hydra battle, for instance the rounds which lead from
-\texttt{Hy} to \texttt{Hy'''}  (Fig.~\vref{fig:Hy5}).
-
-\textbf{Hints:} 
-\begin{itemize}
-
-\item Please keep in mind that the last  configuration of your interactively built battle is known only at the end of the battle.
-Thus, you will have to create and solve subgoals with existential variables.
-For that purpose, the tactic \texttt{eexists}, applied to the
-goal \Verb@{h':Hydra | h -*-> h'}@ generates the subgoal \Verb|h -*-> ?h'|.
-\item You may use Gérard Huet's \emph{zipper} data structure~\cite{zipper} for writing tactics associated with Hercule's  interactive search of a head to chop off.
-\end{itemize}
-
-\end{exercise}
-
-\subsection{Classes of Battles}
-\label{sect:battle-classes}
-
-In some presentations of hydra battles, e.g.~\cite{KP82, bauer2008}, the transformation associated with the $i$-th round may depend on $i$.
-For instance, in these articles, the replication factor at the $i$-th round is equal to $i$.
-In other examples, one can allow the hydra to apply any replication factor at any time.
-In order to be the most general as possible, we define the type of predicates which relate the state of the hydra before and after the $i$-th round of a battle.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-\label{types:Battle}
-
-\begin{Coqsrc}
-Definition dep_round_t := nat -> Hydra -> Hydra -> Prop.
-
-Class Battle :=  {battle_r : dep_round_t;
-                  battle_inclusion : forall i h h',
-                      battle_r i h h' -> round h h'}.
-
-\end{Coqsrc}
-
-The most general class of battles is \texttt{free}, which allows the hydra to chose any replication factor at every step:
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#free}{Hydra.Hydra\_Definitions}}
-
-\begin{Coqsrc}
-Program Instance free : Battle :=
-  (Build_Battle ( fun _  h h' => round h h') _).
-\end{Coqsrc}
-
-We chosed to call \emph{standard} the kind of battles which appear  most often in the litterature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#standard}{Hydra.Hydra\_Definitions}}
-
-\begin{Coqsrc}
-Program Instance standard : Battle := (Build_Battle round_n _).
-Next Obligation.
-  now exists i.  
-Defined.
-\end{Coqsrc}
-
-\subsection{Big Steps}
-
-Let $B$ be any instance of class \texttt{Battle}. It is easy to define inductively the relation between the $i$-th and the $j$-th steps of a battle of type $B$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}
-
-\begin{Coqsrc}
-Inductive battle (B:Battle) : nat -> Hydra -> nat -> Hydra -> Prop :=
-
-| battle_1 : forall i h  h', battle_r   B i  h h' -> 
-                            battle B i h (S i) h'
-| battle_n : forall i h  j h' h'',  battle_r  B i h h''  ->
-                                   battle B (S i) h'' j h'  ->
-                                   battle B i h j h'.
-\end{Coqsrc}
-
-% \begin{remark}
-%  The class \texttt{free} is strongly related with the transitive closure  \texttt{round\_plus}, as expressed by the following lemmas.
-
-% \vspace{4pt}
-% \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-
-%  \begin{Coqsrc}
-%  Lemma battle_free_equiv1 : forall i j h h',  
-%              battle free i h j h' ->   h -+-> h'.
-
-%  Lemma battle_free_equiv2 : forall h h',
-%      h -+-> h' ->
-%     forall i, exists j,  battle free i h j h'.
-%  \end{Coqsrc}
-
-% \end{remark}
-
-\section{A Long Battle}
-\label{sect:big-battle}
-
-In this section we consider a simple example of battle, starting with a small hydra,
-shown on figure~\vref{fig:hinit}, with a simple strategy for both players:
-
-\begin{itemize}
-\item At each round, Hercules chops off the rightmost head of the hydra.
-\item The battle is standard: at the round number $i$, the expected replication is $i$.
-\end{itemize}
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (6,0) {$\bullet$};
-\node (n1) at  (3,3) {$\bullet$};
-\node (h1) at  (1,6) {$\Smiley[1][green]$};
-\node (h2) at  (3,6) {$\Smiley[1][green]$};
-\node (h3) at  (6,6) {$\Smiley[1][green]$};
-\node (h4) at  (6,6) {$\Smiley[1][green]$};
-\node (h5) at  (6,3) {$\Smiley[1][green]$};
-\node (h6) at  (9,3) {$\Smiley[1][green]$};
-\draw (foot) -- (n1);
-\draw (n1) to   [bend left=20] (h1);
-\draw (n1) to   (h2);
-\draw (n1) to   [bend right=20] (h3);
-\draw (foot) -- (h5);
-\draw (foot) to  [bend right=20] (h6);
-\end{tikzpicture}
-
-  \caption{The hydra hinit}
-  \label{fig:hinit}
-\end{figure}
-
-\begin{Coqsrc}
-Definition hinit := hyd3 (hyd_mult head 3)  head head.  
-\end{Coqsrc}
-
-The lemma we would like to prove is ``The considered battle lasts exactly $N$ rounds'',
-with $N$ being a natural number we gave to guess.
-
-But the  battle is so long that no \emph{test} can give us an estimation of its length, and we do need the expressive power of logic to compute this length.
-However, in order to  guess this length, we made some experiments, computing with \gallina{}, \coq{}'s  functional programming language.
-Thus, we can consider this development as a collaboration of proof with computation.
-
-In the following lines, we try to show faithfully how we found the value of the number $N$.
-
-The complete proof is in file \url{../theories/html/hydras.Hydra.BigBattle.html}.
-
-\subsection{The beginning of Hostilities}
-During the two first rounds, our hydra loses its two rightmost heads. Thus just before the third round, it looks like in figure~\vref{fig:hinit-plus2}.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (3,0) {$\bullet$};
-\node (n1) at  (3,3) {$\bullet$};
-\node (h1) at  (1,6) {$\Smiley[1][green]$};
-\node (h2) at  (3,6) {$\Smiley[1][green]$};
-\node (h3) at  (6,6) {$\Smiley[1][green]$};
-\node (h4) at  (6,6) {$\Smiley[1][green]$};
-\draw (foot) -- (n1);
-\draw (n1) to   [bend left=20] (h1);
-\draw (n1) to   (h2);
-\draw (n1) to   [bend right=20] (h3);
-\end{tikzpicture}
-
-  \caption{The hydra (hyd1 h3)}
-  \label{fig:hinit-plus2}
-\end{figure}
-
-The following lemma  is a formal description of these first rounds, in terms of the
-\texttt{battle} predicate.
-
-\begin{Coqsrc}
-Lemma L_0_2 : battle standard 0 hinit 2 (hyd1 h3).   
-\end{Coqsrc}
-
-\subsection{Looking for Regularities}
-
-A first study with pencil and paper suggested us that, after three rounds, the hydra always looks like in figure~\vref{fig:hinit-plusn} (with a variable number of 
-subtrees of height 1 or 0).
-Thus, we introduce handy notations.
-
-\begin{Coqsrc}
-Notation h3 := (hyd_mult head 3).
-Notation h2 := (hyd_mult head 2).
-Notation h1 := (hyd1 head).
-
-Definition hyd a b c := 
-  node (hcons_mult h2  a
-             (hcons_mult h1  b
-                         (hcons_mult head c hnil))).
-\end{Coqsrc}
-
-For instance Fig~\vref{fig:hinit-plusn} shows the hydra (\texttt{hyd 3 4 2}). The hydra (\texttt{hyd 0 0 0})  is the ``final'' hydra of any terminating battle, {i.e.},
-a tree whith exactly one node and no edge.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (15,0) {$\bullet$};
-\node (a) at  (3,4) {$\bullet$};
-\node (b) at  (6,4) {$\bullet$};
-\node (c) at  (9,4) {$\bullet$};
-\node (d) at  (13,4) {$\bullet$};
-\node (e) at  (16,4) {$\bullet$};
-\node (f) at  (19,4) {$\bullet$};
-\node (g) at  (22,4) {$\bullet$};
-\node (h) at  (25,4) {$\Smiley[1][green]$};
-\node (i) at  (28,4) {$\Smiley[1][green]$};
-\node (aa) at  (2.5,8) {$\Smiley[1][green]$};
-\node (ab) at  (3.5,8) {$\Smiley[1][green]$};
-\node (ba) at  (5.5,8) {$\Smiley[1][green]$};
-\node (bb) at  (6.5,8) {$\Smiley[1][green]$};
-\node (ca) at  (8.5,8) {$\Smiley[1][green]$};
-\node (cb) at  (9.5,8) {$\Smiley[1][green]$};
-\node (da) at  (13,8) {$\Smiley[1][green]$};
-\node (ea) at  (16,8) {$\Smiley[1][green]$};
-\node (fa) at  (19,8) {$\Smiley[1][green]$};
-\node (ga) at  (22,8) {$\Smiley[1][green]$};
-\draw (foot) -- (a);
-\draw (foot) -- (b);
-\draw (foot) -- (c);
-\draw (foot) -- (d);
-\draw (foot) -- (e);
-\draw (foot) -- (f);
-\draw (foot) -- (g);
-\draw (foot) -- (h);
-\draw (foot) -- (i);
-\draw (a) -- (aa);
-\draw (a) -- (ab);
-\draw (b) -- (ba);
-\draw (b) -- (bb);
-\draw (c) -- (ca);
-\draw (c) -- (cb);
-\draw (d) -- (da);
-\draw (e) -- (ea);
-\draw (f) -- (fa);
-\draw (g) -- (ga);
-% \node (a) at  (3,4) {$\bullet$};
-% \node (h1) at  (1,6) 
-% \node (h2) at  (3,6) {$\Smiley[1][green]$};
-% \node (h3) at  (6,6) {$\Smiley[1][green]$};
-% \node (h4) at  (6,6) {$\Smiley[1][green]$};
-% \draw (foot) -- (n1);
-% \draw (n1) to   [bend left=20] (h1);
-% \draw (n1) to   (h2);
-% \draw (n1) to   [bend right=20] (h3);
-\end{tikzpicture}
-
-  \caption{The hydra (hyd 3 4 2)}
-  \label{fig:hinit-plusn}
-\end{figure}
-
-With these notations, we get a formal description of the first three rounds.
-
-\begin{Coqsrc}
-Lemma L_2_3 : battle standard 2 (hyd1 h3)  3 (hyd 3 0 0).
-
-Lemma L_0_3 : battle standard 0 hinit 3 (hyd 3 0 0).
-\end{Coqsrc}
-
-\subsection{Computing \dots}
-In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple datatype for representing the states as tuples composed of
-the round number, and the respective number of daughters  \texttt{h2}, \texttt{h1}, and heads
-of the current hydra.
-
-\begin{Coqsrc}
- Record state : Type :=
-    mks {round: nat ; n2 : nat ; n1 : nat ; nh : nat}.
-\end{Coqsrc}
-
-The following function returns the next configurarion of the game. 
-Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle will only start with the lemma
-\texttt{lemma:step-battle}, page~\pageref{lemma:step-battle}.
-
-\begin{Coqsrc}
-Definition next (s : state) :=
-  match s with
-  | mks round a b (S c) => mks (S round) a b c
-  | mks round a (S b) 0 => mks (S round) a b (S round)
-  | mks round (S a) 0 0 => mks (S round) a (S round) 0
-  | _ => s
-  end.
-\end{Coqsrc}
-
-We can make bigger steps through iterations of \texttt{next}.
-The functional \texttt{iterate}, similar to Standard Library's \texttt{Nat.iter},
-is defined and studied in~\href{../theories/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
-
-\label{Functions:iterate}
-
-\begin{Coqsrc}
-Fixpoint iterate {A:Type}(f : A -> A) (n: nat)(x:A) :=
-  match n with
-  | 0 => x
-  | S p => f (iterate  f p x)
-  end.
-\end{Coqsrc}
-
-The following function computes the state of the battle at the $n$-th round.
-
-\begin{Coqsrc}
-Definition test n := iterate next (n-3) (mks 3 3 0 0).
-
-Compute test 3.
-   (**
-     = {| round := 3; n2 := 3; n1 := 0; nh := 0 |}
-     : state
-    *)
-
-Compute test 4.
- (*
-  = {| round := 4; n2 := 2; n1 := 4; nh := 0 |}
-     : state
- *)
-
-Compute test 5.
- (*
-   = {| round := 5; n2 := 2; n1 := 3; nh := 5 |}
-     : state
- *)
-
-Compute test 2000.
-(*
-  = {| round := 2000; n2 := 1; n1 := 90; nh := 1102 |}
-     : state
-*)
-\end{Coqsrc}
-
-The battle we are studying seems to be awfully long. Let us concentrate our
-tests on some particular events : the states where $\texttt{nh}=0$.
-From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} is equal to zero.
-
- \begin{Coqsrc}
- Compute test 10.
-(*
-
-    = {| round := 10; n2 := 2; n1 := 3; nh := 0 |}
-     : state
-*)
- \end{Coqsrc}
-
-Thus, $ (1 + 11)$ rounds later, the \texttt{n1} field is equal to $2$, and 
-\texttt{nh}   to $0$.
-
-\begin{Coqsrc}
-Compute test 22.
-(*
-
- = {| round := 22; n2 := 2; n1 := 2; nh := 0 |}
-     : state
-*)
-\end{Coqsrc}
-
-\begin{Coqsrc}
- Compute test 46.
-(*
-
- = {| round := 46; n2 := 2; n1 := 1; nh := 0 |}
-     : state
-
-*)
-
-Compute test 94.
-
-(*
-
- = {| round := 94; n2 := 2; n1 := 0; nh := 0 |}
-     : state
-
-*) 
-\end{Coqsrc}
-
-Next round, we decrement \texttt{n2} and set \texttt{n1} to $95$.
-
-\begin{Coqsrc}
- Compute test 95.
-
-(*
-
-  = {| round := 95; n2 := 1; n1 := 95; nh := 0 |}
-     : state
-
-*)
-\end{Coqsrc}
-
-We now have some intuition of the sequence.
-It looks like the next ``\texttt{nh}=0'' event will happen at the $192=2(95+1)$-th round, then at the $2(192+1)$-th round, etc.
-
-\begin{Coqsrc}
-Definition doubleS (n : nat) := 2 * (S n).
-
-Compute test (doubleS 95).
-
-(**
- = {| round := 192; n2 := 1; n1 := 94; nh := 0 |}
-     : state
- *)
-
-Compute test (iterate doubleS 2 95).
-
-(*
-  = {| round := 386; n2 := 1; n1 := 93; nh := 0 |}
-     : state
-*)
-\end{Coqsrc}
-
-\subsection{Proving \dots}
-We are now able to reason about the sequence of transitions defined by our hydra battle. Instead of using the data-type \texttt{state} we study the relationship
-between different configurations of the battle.
-
-Let us define a binary relation associated with every round of the battle.
-In the following definition \texttt{i} is associated with the round number (or date, if we consider a discrete time), and \texttt{a}, \texttt{b}, \texttt{c} respectively associated with the number of \texttt{h2}, \texttt{h1} and heads connected to the hydra's foot.
-
-\begin{Coqsrc}
-Inductive one_step (i: nat) :
-  nat -> nat -> nat -> nat -> nat -> nat -> Prop :=
-| step1: forall a b c, one_step i a b (S c) a b c
-| step2:  forall a b, one_step i a (S b) 0 a b (S i)
-| step3: forall a, one_step i (S a) 0 0 a (S i) 0.
-\end{Coqsrc}
-
-The relation between \texttt{one\_step} and the rules of hydra battle is asserted by the following lemma.
-
-\label{lemma:step-battle}
-
-\begin{Coqsrc}
-Lemma step_battle : forall i a b c a' b' c', 
-   one_step i a b c a' b' c' ->
-   battle standard i (hyd  a b c)  (S i) (hyd a' b' c').
-\end{Coqsrc}
-
-Next, we define ``big steps'' as the transitive closure of \texttt{one\_step},
-and reachability (from the initial configuration of figure~\ref{fig:hinit} at time $0$).
-
-\begin{Coqsrc}
- Inductive steps : nat -> nat -> nat -> nat ->
-                  nat -> nat -> nat -> nat -> Prop :=
-| steps1 : forall i a b c a' b' c',
-    one_step i a b c a' b' c' -> steps i a b c (S i) a' b' c'
-| steps_S : forall i a b c j a' b' c' k a'' b'' c'',
-    steps i a b c j a' b' c' ->
-    steps j a' b' c' k a'' b'' c'' ->
-    steps i a b c k  a'' b'' c''.
-
-Definition reachable (i a b c : nat) : Prop :=
-  steps 3 3 0 0 i a b c.
-\end{Coqsrc}
-
-The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{battle}.
-
-\begin{Coqsrc}
- Lemma steps_battle : forall i a b c j a' b' c', 
-   steps i a b c j a' b' c' ->
-   battle standard i (hyd  a b c)   j  (hyd a' b' c').
-\end{Coqsrc}
-
-Thus, any result about \texttt{steps} will be applicable to standard battles.
-Using the predicate \texttt{steps} our study of the length of the considered battle
-can  be decomposed into three parts:
-
-\begin{enumerate}
-\item  Characterization of regularities of some events
-\item Study of the beginning of the battle
-\item Computing the exact length of the battle.
-\end{enumerate}
-
-First, we prove that, if at round $i$ the hydra is equal to
-(\texttt{hyd a (S b) 0}), then it will be equal to (\texttt{hyd a b 0}) at the $2(i+1)$-th round.
-
-\begin{Coqsrc}
-Lemma LS : forall c a b i,  steps i a b (S c) (i + S c) a b 0.
-Proof.
-  induction c.
- -   intros;  replace (i + 1) with (S i).
-     + repeat constructor.
-     + ring.
- -  intros; eapply  steps_S.
-   +   eleft;   apply rule1.
-   +   replace (i + S (S c)) with (S i + S c) by ring;  apply IHc.
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma doubleS_law : forall  a b i, steps i a (S b) 0 (doubleS i) a b 0.
-Proof.
-  intros;  eapply steps_S.
-  +   eleft;   apply step2.
-  +   unfold doubleS; replace (2 * S i) with (S i + S i) by ring; 
-        apply LS.
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma reachable_S  : forall i a b, reachable i a (S b) 0 ->
-                                   reachable (doubleS i) a b 0.
-Proof.
-  intros; right with  (1 := H); apply doubleS_law.
-Qed.
-\end{Coqsrc}
-
-From now on, the lemma \texttt{reachable\_S} allows us to watch larger steps of 
-the battle.
-
-\begin{Coqsrc}
- Lemma L4 : reachable 4 2 4 0.
-Proof.
-  left; constructor.
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma L10 : reachable 10 2 3 0.
-Proof.
-  change 10 with (doubleS 4).
-  apply reachable_S, L4.
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma L22 : reachable 22 2 2 0.
-Proof.
-  change 22 with (doubleS 10).
-  apply reachable_S, L10.
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma L46 : reachable 46 2 1 0.
-Proof.
-  change 46 with (doubleS 22); apply  reachable_S, L22.
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma L94 : reachable 94 2 0 0.
-Proof.
-  change 94 with (doubleS 46); apply reachable_S, L46.
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma L95 : reachable 95 1 95 0.
-Proof.
-  eapply steps_S.
-  -  eexact L94.
-  -  repeat constructor.
-Qed.
-\end{Coqsrc}
-
-\subsection{Giant Steps}
-
-We are now able to make bigger steps in the simulation of the battle.
-First, we iterate the lemma \texttt{reachable\_S}.
-
-\begin{Coqsrc}
-Lemma Bigstep : forall b i a , reachable i a b 0 ->
-                               reachable (iterate doubleS b i) a 0 0.
- Proof.
-  induction b.
-  -  trivial.
-  -  intros;  simpl;   apply reachable_S in H.
-     rewrite <- iterate_comm; now apply IHb.
- Qed.
-\end{Coqsrc}
-
-Applying lemmas \texttt{BigStep} and \texttt{L95} we make a first jump.
-
-\begin{Coqsrc}
- Definition M := (iterate doubleS 95 95).
-
-Lemma L2_95 : reachable M 1 0 0.
-Proof.
-  apply Bigstep,  L95.
-Qed.
-\end{Coqsrc}
-
-Figure~\ref{fig:HM}  represents the hydra at the $M$-th round.
-At the $(M+1)$-th round, it will look like in fig~\ref{fig:HM-plus1}.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:HM}}
-The state of the hydra after $M$ rounds.
-% The hydra \texttt{h} of the proof that \(\omega^2\) is too small for proving Hercules' victory
-
-\end{figure}
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (0,5) {$\bullet$};
-\node (N12) at (0,8) {$\Smiley[2][green]$};
-\node (N2) at (2,5) {$\bullet$};
-\node (N22) at (2,8) {$\Smiley[2][green]$};
-\node (N3) at (4,5) {$\bullet$};
-\node (N32) at (4,8) {$\Smiley[2][green]$};
-\node (N4) at (6,5) {$\bullet$};
-\node (N42) at (6,8) {$\Smiley[2][green]$};
-
-\node (Ndots) at (12,8) {\Huge $\dots$};
-\node (Ndots2) at (12,5) {\Huge $\dots$};
-
-\node (N8) at (18,5) {$\bullet$};
-\node (N82) at (18,8) {$\Smiley[2][green]$};
-\node (N9) at (20,5) {$\bullet$};
-\node (N92) at (20,8) {$\Smiley[2][green]$};
-
-\draw (foot) -- (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N8);
-\draw (foot) -- (N9);
-\draw (N1) to  (N12);
-\draw (N2) to  (N22);
-\draw (N3) to  (N32);
-\draw (N4) to  (N42);
-\draw (N8) to  (N82);
-\draw (N9) to  (N92);
-\end{tikzpicture}
-\caption{\label{fig:HM-plus1}}
-The state of the hydra after $M+1$ rounds (with $M+1$ heads).
-
-\end{figure}
-
-\begin{Coqsrc}
-Lemma L2_95_S : reachable (S M) 0 (S M) 0.
-Proof.
-  eright.
-  - apply L2_95.
-  -  left; constructor 3.
-Qed.
-\end{Coqsrc}
-
-Then, applying once more the lemma \texttt{BigStep}, we get the exact time when
-Hercules wins!
-
-\begin{Coqsrc}
-Definition N :=   iterate doubleS (S M) (S M).
-
-Theorem   SuperbigStep : reachable N  0 0 0 .
-Proof.
-  apply Bigstep, L2_95_S.
-Qed.
-\end{Coqsrc}
-
-We are now able to prove formally that the considered battle is 
-composed of $N$ steps.
-
-\begin{Coqsrc}
-Lemma Almost_done :
-  battle standard 3 (hyd 3 0 0) N (hyd 0 0 0).
-Proof. 
-  apply steps_battle, SuperbigStep.
-Qed.
-
-Theorem Done :
-  battle standard 0 hinit N head.
-Proof.
-  eapply battle_trans.
-  -   apply Almost_done.
-  -  apply L_0_3.
-Qed.
-\end{Coqsrc}
-
-\subsection{A Minoration Lemma}
-
-Now, we would like to get an intuition of  how big the number $N$ is.
-For that purpose, we use a minoration of the function \texttt{doubleS} by the
-function (\texttt{fun n => 2 * n}).
-
-\begin{Coqsrc}
-Definition exp2 n := iterate (fun n => 2 * n) n 1.
-\end{Coqsrc}
-Using some facts (proven in 
-\href{../theories/html/hydras.Hydra.BigBattle.html}{hydras.Hydra.BigBattle}),we get several  minorations.
-
-\begin{Coqsrc}
-Lemma minoration_0 : forall n,  2 * n <= doubleS n.
-
-Lemma minoration_1 : forall n x, exp2 n * x <= iterate doubleS n x.
-
-Lemma minoration_2 : exp2 95 * 95 <= M.
-
-Lemma minoration_3 : exp2 (S M) * S M <= N.
-
-Lemma minoration : exp2 (exp2 95 * 95) <= N.
-\end{Coqsrc}
-
-The number $N$ is greater than or  equal to $2^{2^{95}\times 95}.$ If we wrote $N$ in base $10$, $N$ would require at least $10^{30}$ digits!
-
-\section{Generic Properties}
-
-The example we just studied shows that the termination of any battle may take a very long time. If we want to study hydra battles in general, we have to consider 
-any hydra and any strategy, both for Hercules and the hydra itself. So, we first  give some definitions, generally borrowed from transition systems vocabulary (see~\cite{tel_2000} for instance).
-
-\subsection{Liveliness}
-
-Let $B$ be an instance of \texttt{Battle}. We say that $B$ is \emph{alive} if
-for any configuration $(i,h)$, where $h$ is not a head, there exists a further step in class $B$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
-
-\begin{Coqsrc}
-Definition Alive (B : Battle) :=
-  forall i h, 
-     h <> head -> {h' : Hydra |  B i h h'}.
-\end{Coqsrc}
-
-The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
-\url{../theories/html/hydras.Hydra.Hydra_Theorems.html} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
-
-\begin{Coqsrc}
-Theorem Alive_free: Alive free.
-
-Theorem Alive_standard: Alive standard.  
-\end{Coqsrc}
-
-Both theorems are proved with the help of the  following strongly specified function:
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
-
-\begin{Coqsrc}
-Definition  next_round_dec n :
- forall h , (h = head) + {h' : Hydra & {R1 h h'} + {R2 n h  h'}}.
-\end{Coqsrc}
-
-\subsection{Termination}
-
-The termination of all battles is naturally expressed by the predicate \texttt{well\_founded} defined in the module \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf} 
- of the Standard Library.
-
-\index{Predicates!Termination}
-
-\begin{Coqsrc}
-Definition Termination :=  well_founded (transp _ round).
-\end{Coqsrc}
-
-Let $B$ be an instance of class \texttt{Battle}. A \emph{variant} for $B$ consists
-in a well-founded relation $<$  on some type \texttt{A}, and a function
-(also called a \emph{measure}) \texttt{m:Hydra->A} such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $B$, the inequality $m(h')<m(h)$ holds.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
-
-\label{sect:hvariant-def}
-
-\begin{Coqsrc}
-Class Hvariant {A:Type}{Lt:relation A}(Wf: well_founded Lt)(B : Battle)
-  (m: Hydra -> A):   Prop :=
-  {variant_decr :forall i h h',
-      h <> head ->
-      battle_r  B i  h h' -> Lt (m h') (m h)}.
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
- Prove that, if there is an instance of (\texttt{Hvariant Lt wf\_Lt $B$ $m$}), then there exists no infinite battle in  $B$.
-\end{exercise}
-
-\subsection{A  Small Proof of Impossibility}
-\index{Maths!Proofs of impossibility}
-
-\label{omega-case}
-
-When one wants to prove a termination theorem with the help of a variant, 
-one has to consider first a well-founded set $(A,<)$, then a strictly decreasing measure on this set.
-The following two lemmas show that if  the order structure $(A,<)$ is too simple, it is useless to look for a convenient measure, which simply no exists.
-Such kind of result is useful, because it saves you time and effort.
-
-The best known well-founded order is the natural order on the set $\mathbb{N}$ of natural numbers (the type \texttt{nat} of Standard library).
-It would be interesting to look for some measure $m:\texttt{nat}\arrow\texttt{nat}$ and prove it is a variant.
-
-Unfortunately, we can prove that 
-\emph{no} instance of class (\texttt{WfVariant round Peano.lt $m$}) can be built, where
-$m$ is \emph{any} function of type \texttt{Hydra $\arrow$ nat}.
-
-Let us present the main steps of that proof, the script of which  is in the module ~\href{../theories/html/hydras.Hydra.Omega_Small.html}{Hydra/Omega\_Small.v} \footnote{ The name of this file means ``the ordinal $\omega$ is too small for proving the termination of [free] hydra battles ''.
-In effect, the elements of $\omega$, considered as a set, are just the natural numbers (see next chapter for more details)}.
-
-%\subsubsection{Preliminaries}
-
-Let us assume there exists some variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
-    the  termination of all hydra battles.
-
-\begin{Coqsrc}
-Section Impossibility_Proof.
- Variable m : Hydra -> nat.
- Hypothesis Hvar : Hvariant lt_wf free m.
-\end{Coqsrc}
-
-We define an injection $\iota$ from the type \texttt{nat} into \texttt{Hydra}.
-For any natural number $i$, $\iota(i)$ is the hydra composed of a foot and
-$i+1$ heads at height $1$. For instance, Fig.~\ref{fig:flower} represents the hydra $\iota(3)$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (4,0) {$\bullet$};
-\node (N1) at (2,2) {$\Smiley[2][green]$};
-\node (N2) at (4,2) {$\Smiley[2][green]$};
-\node (N3) at (6,2) {$\Smiley[2][green]$};
-\node (N4) at (8,2) {$\Smiley[2][green]$};
-\draw (foot) to [bend left =25] (N1);
-\draw (foot) to [bend left =15] (N2);
-\draw (foot) to [bend right =15] (N3);
-\draw (foot) to [bend right =25] (N4);
-\end{tikzpicture}
-\caption{\label{fig:flower}
-The hydra $\iota(3)$}
-\end{figure}
-
-  \begin{Coqsrc}
-  Let iota (i: nat) := hyd_mult head (S i).    
-  \end{Coqsrc}
-
-Let us consider now some hydra \texttt{big\_h} out of the range of the injection $\iota$ (see Fig.~\vref{fig:h-omega-omega}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to  (N2);
-\end{tikzpicture}
-\caption{\label{fig:h-omega-omega}}
- The hydra \texttt{big\_h}.
-\end{figure}
-
- \begin{Coqsrc}
-  Let big_h := hyd1 (hyd1 head).
- \end{Coqsrc}
-
- Using the functions $m$ and $\iota$, we define a second hydra \texttt{small\_h}, and show
- there is a one-round battle that transforms \texttt{big\_h} into \texttt{small\_h}. Please note that,
-due to the hypothesis \texttt{Hvar}, we are interested in the termination of \emph{free} battles. 
-There is no problem to consider a round with (\texttt{m big\_h}) as the replication factor.
-
-  \begin{Coqsrc}
- Let small_h := iota (m big_h).
-
- Fact big_to_small : big_h -1-> small_h.
- Proof.
-      exists (m big_h); right; repeat constructor.     
- Qed.
-    \end{Coqsrc}
-
-But, by hypothesis, $m$ is a variant. Hence, we infer the following inequality.
-
- \begin{Coqsrc}
-Lemma m_lt : m small_h < m big_h.
- \end{Coqsrc}
-
-In order to get a contradiction, it suffices to  prove the inequality
-\texttt{m big\_h <= m small\_h}, i.e.,  \texttt{m big\_h <= m (iota (m big\_h))}.
-
-More generally, we prove the following lemma:
-
-\begin{Coqsrc}
-Lemma m_ge : forall i:nat, i <= m (iota i).
-\end{Coqsrc}
-
-Intuitively, it means that, from any hydra of the form (\texttt{iota $i$}), the battle will 
-take (at least) $i$ rounds. Thus the associated measure cannot be less than $i$.
-Technically, we prove this lemma by Peano induction on $i$.
-
-\begin{itemize}
-\item The base case $i=0$ is trivial
-\item Otherwise, let $i$ be any natural number and assume  the inequality
-  $i \leq m(\iota(i))$.
-  \begin{enumerate}
-  \item  But the hydra $\iota(S(i))$ can be transformed in one round into
-    $\iota(i)$ (by losing its righmost head, for instance)
-  \item Since $m$ is a variant, we have $m(\iota(i)) < m(\iota(S(i)))$,
-    hence  $i< m(\iota(S(i)))$, which implies  $S(i)\leq  m(\iota(S(i)))$.
-  \end{enumerate}
-\end{itemize}
-
- Then our proof is almost finished.
-
-   \begin{Coqsrc}
-Theorem Contradiction : False.
-Proof.
- apply (Nat.lt_irrefl (m big_h));
-   apply  Lt.le_lt_trans with (m small_h).
-  - apply m_ge.
-  - apply m_lt.
-Qed.
-
-End Impossibility_Proof.
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\subsubsection{Conclusion}
-
-In order to build a variant for proving the termination of all hydra battles, we need to consider order structures more complex than the usual order on type \texttt{nat}. 
-The notion of \emph{ordinal number} provides a catalogue of well-founded order types.
-For a reasonably large bunch of ordinal numbers, \emph{ordinal notations} are data-types which allow the \coq{} user to define functions, to compute and prove some properties, for instance by reflection.
-
-The next chapter is dedicated to a generic implementation of ordinal notations, and chapter~\ref{chap:T1} to a proof of termination of all hydra battles with the help of an ordinal notation for the interval $[0,\epsilon_0)$.
-
-%--------------------------------------------------------------
-
-\chapter{Introduction to Ordinal Numbers and Ordinal Notations}
-
-The proof of termination of all hydra battles presented in~\cite{KP82} is based
-on \emph{ordinal numbers}.
-From a mathematical point of view, an ordinal is a representant of an equivalence class for isomorphims of  totally ordered well-founded sets.
-
-More intuitively, let us have a look at Figure~\ref{fig:ordinal-sequence}. It presents a small sequence of ordinal numbers, which extends the sequence of natural numbers.
-
-\begin{figure}[h]
-  \centering
-\fbox{\Large
-  \begin{minipage}{1.0\linewidth}
-  \begin{align*}
-     &\textcolor{blue}{0},\,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,\ldots\\
-&\textcolor{red}{\omega},\,\omega+1,\omega+2,\omega+3,\ldots\\
-&\textcolor{red}{\omega\times 2},\,\omega\times 2+1,\ldots, \textcolor{red}{\omega\times 3},\,\omega\times 3+1,\ldots, \textcolor{red}{\omega\times 4},\ldots,\\
-&\textcolor{red}{\omega^2},\ldots, \textcolor{red}{\omega^2\times 42},\ldots,\textcolor{red}{\omega^3},\ldots, \textcolor{red}{\omega^4},\omega^4+1,\ldots,\\
-&\textcolor{red}{\omega^\omega},\ldots, \textcolor{red}{\omega^\omega+\omega^7\times 8},\ldots,\textcolor{red}{\omega^\omega\times 2},\omega^\omega\times 2+1, \ldots,\\
-&\textcolor{red}{\omega^{\omega^\omega}},\ldots, \textcolor{red}{\omega^{\omega^\omega}+\omega^\omega\times 42+ \omega^{55}+\omega}, \ldots, \textcolor{red}{\omega^{\omega^{\omega+1}}}, \omega^{\omega^{\omega+1}}+1,\dots\\
-& \textcolor{red}{\epsilon_0 (= \omega^{\epsilon_0)}}, \epsilon_0+1, \epsilon_0+2, \epsilon_0+3, \ldots,\\
-& \textcolor{red}{\epsilon_1}, \ldots, \textcolor{red}{\epsilon_2}, \ldots, \textcolor{red}{\epsilon_\omega},\ldots \\
-& \textcolor{red}{\Gamma_0}, \Gamma_0+1, \Gamma_0+2, \Gamma_0+3,\ldots, \textcolor{red}{\Gamma_0+\omega}, \ldots,\\
-&\ldots
-  \end{align*}   
-  \end{minipage}}
-
-  \caption{A short overview of the sequence of ordinal numbers}
-  \label{fig:ordinal-sequence}
-\end{figure}
-
-Let us comment some features of this figure:
-
-\begin{itemize}
-\item The ordinals are listed in a strictly increasing order. 
-\item Dots : ``$\ldots$'' stand for an infinite sequence of ordinals, not shown for lack of space.
-For instance, the ordinal $42$ is not shown in the first line, but it exists, between $17$ and $\omega$.
-\item Each ordinal printed in black is the immediate successor of another ordinal.
-We call it a \emph{successor} ordinal.
-For instance, $12$ is the successor of $11$, and $\omega^4+1$ the successor of $\omega^4$.
-\item Ordinals (displayed in red)  that  follow immediately dots are called \emph{limit ordinals}.
-With respect to the order induced by this sequence, any limit ordinal $\alpha$ is the least upper bound of  the set $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
-\item
-For instance $\omega$ is the least upper bound of the set of all finite ordinals (in the first line). It is also the first limit ordinal, and the first infinite ordinal, in the sense that 
-the set $\mathbb{O}_\omega$ is infinite.
-\item The ordinal $\epsilon_0$ is the first number which is equal to its own exponential of base $\omega$.
-It plays an important role in proof theory, and is particularly studied in chapters~\ref{chap:T1} to \ref{chap:alpha-large}.
-\item Any ordinal is  either the ordinal \textcolor{blue}{$0$},
-a successor ordinal, or a \textcolor{red}{limit ordinal}.
-\end{itemize}
-
-\section{The Mathematical Point of View}
-
-\subsection{Well-ordered Sets}
-Let us start with some definitions.
-A  \emph{well-ordered set} is a set provided with a binary relation $<$ which has the following properties.
-\begin{description}
-\item[irreflexivity] : $\forall x\in A, x\not< x$
-\item[transitivity] : $\forall x\,y\,z\in A, x<y \Rightarrow y<z \Rightarrow x<z$
-\item[trichotomy]: $\forall x\,y\in A, x<y \vee x = y \vee y < x$
-\item[well foundedness]: $<$ is well-founded (every element of $A$ is accessible)\footnote{In classical mathematics, we would say that there is no infinite sequence $a_1>a_2> \dots a_n> a_{n+1}\dots$ in $A$.
-Please refer to any documenation on \coq{} for having more details on well-foundedness and accessibility.}.
-
-\end{description}
-
-The best known examples of well-ordered sets are the set $\mathbb{N}$ of natural numbers (with the usual order $<$), as well as any finite segment $[0,i)=\{j\in\mathbb{N}\,|\,j<i\}$.
-The disjoint union of two copies of $\mathbb{N}$, \emph{i.e.} the set $\{0,1\}\times\mathbb{N}$ is also well-ordered,
-with respect to the order below:
-
-\begin{align*}
-(i,j) < (i,k) & \;\textit{if} \; j < k\\
-(0,k) < (1,l) & \;\textit{for\,any}\;k \;\textit{and} \; l
-\end{align*}
-
-\subsection{Ordinal Numbers}
-
-\index{Maths!Ordinal numbers}
-
-Let $(A,<_A)$ and $(B,<_B)$ two well-ordered sets. $A$ and $B$ are said to have \emph{the same order type} if 
-there exists a strictly monotonous bijection $b$ from $A$ to $B$, \emph{i.e.} which verifies the proposition
-$\forall x\,y\in A, x <_A y \Rightarrow b(x) <_B  b(y)$.
-
-Having the same order type is an equivalence relation between well-ordered sets. Ordinal numbers (in short \emph{ordinals}) are descriptions (\emph{names}) of the equivalence classes.
-For instance, the order type of $(\mathbb{N},<)$ is associated with the ordinal called  $\omega$, and the order we considered on 
-the disjoint union of $\mathbb{N}$ and itself is named $\omega+\omega$.
-
-In a set-theoretic framework, one can consider any ordinal $\alpha$ as a well-ordered set, whose  elements are just the ordinals strictly less than $\alpha$, \emph{i.e.} the \emph{segment} $\mathbb{O}_\alpha=[0, \alpha)$.
-So, one can speak about \emph{finite}, \emph{infinite}, \emph{countable}, etc., ordinals.
-Nevertheless, since we work within type theory,
-we do not identify ordinals as sets of ordinals, but the correspondance between ordinals and sets of ordinals is the function that maps $\alpha$ to $\mathbb{O}_\alpha$.
-For instance $\mathbb{O}_\omega=\mathbb{N}$, and $\mathbb{O}_7=\{0,1,2,3,4,5,6\}$.
-
-We cannot cite all the litterature published on ordinals since Cantor's book 
-\cite{cantorbook}, and 
-leave it to the reader to explore the bibliography.
-
-\section{Ordinal Numbers in Coq}
-
-Two kinds of representation of ordinals are defined herein.
-
-\begin{itemize}
-\item A ``mathematical'' representation of the set of countable ordinal numbers, afer Kurt Schütte~\cite{schutte}.
-This representation uses several (hopefully harmless) axioms.
-We use it as a reference for proving the correctness of ordinal notations.
-\item A family of \emph{ordinal notations}, \emph{i.e.} data types used to represent segments $[0,\mu)$, where $\mu$ is some countable ordinal.
-Each ordinal notation is defined inside the Calculus of Inductive Constructions (without axioms).
-Many functions are defined, allowing proofs by computation.
-Note that proofs of
-correctness of a given ordinal notation with respect to Schütte's model obviously use axioms.
-Please execute the \texttt{Print Assumptions} command in case of doubt.
-\end{itemize}
-
-\section{Countable Ordinals}
-
-Chapter~\ref{chap:schutte} of this document presents an adaptation to \coq{} of an axiomatization in classical logic of the set of countable ordinals by K. Schütte~\cite{schutte}. 
-That formalization is quite complex, technical and unshamedly non-constructive,  so we put its description  in the last chapter of this document.
-
-Please note that Schütte considers the (uncountable) set $\mathbb{O}$ of all countable ordinals.
-This set is well ordered (which is one of Schütte's axioms), and associates to any ordinal $\alpha$ the segment $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
-
-In our adaptation to \coq{}, we declare a type \texttt{Ord}, a binary relation \texttt{lt} (with infix notation \texttt{"\_<\_"}, and assume Schütte's axiom. In Chapter~\ref{chap:schutte},
-we derive some interesting properties of countable ordinals from these axioms.
-It is interesting to compare proofs of a given property (for instance the associativity of addition) in the computatuinal framework of some ordinal notation, and in the axiomatic model of Schütte.
-
-\section{Ordinal Notations}
-
-Fortunately, the ordinals we need for  studying hydra battles are much simpler than Schütte's, and can be represented as quite simple data types in \gallina.
-So, we will use \emph{ordinal notations} (also called \emph{[ordinal] notation systems}).
-
-Let $\alpha$ be some (countable) ordinal; 
-in \coq{} terms, we call ordinal notation for $\alpha$ a structure composed 
-of:
-\begin{itemize}
-\item A data type $A$ for representing all ordinals strictly below $\alpha$,
-\item A well founded order $<$ on $A$, 
-\item A correct function for comparing two ordinals. Note  that the reflexive closure of $<$ is thus a \emph{total order}.
-\end{itemize}
-
-Such a structure can be proved correct relatively to another ordinal notation or
-to Schütte's model.
-
-\subsubsection*{Ordered Types}
-
-The library ~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Classes.RelationClasses.html}{%
-Coq.Classes.RelationClasses} contains some definitions and facts about binary relations, among them strict orders.
-
-\begin{Coqsrc}
-Variable A: Type.
-
-  Class StrictOrder (R : relation A) : Prop := {
-    StrictOrder_Irreflexive :> Irreflexive R ;
-    StrictOrder_Transitive :> Transitive R }.
-\end{Coqsrc}
-
-\subsection{A Class for Ordinal Notations}
-
-The following class definition, parameterized with a type $A$, a binary relation \texttt{lt} on $A$, specifies that \texttt{lt} is a well-founded strict order.
-The reflexive closure of \texttt{lt}, (called \texttt{le}, for ``less  or equal than'') is a total decidable order, implemented through a comparison function \texttt{compare}.
-The correctness of this function is expressed through Stdlib's type
-\texttt{Datatypes.CompareSpec}.
-
-\begin{Coqsrc}
-Inductive CompareSpec (Peq Plt Pgt : Prop) : comparison -> Prop :=
-    CompEq : Peq -> CompareSpec Peq Plt Pgt Eq
-  | CompLt : Plt -> CompareSpec Peq Plt Pgt Lt
-  | CompGt : Pgt -> CompareSpec Peq Plt Pgt Gt
-\end{Coqsrc}
-
-\vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}}
-
-\label{types:ON}
-
-\begin{Coqsrc}
-Class ON {A:Type}(lt: relation A)
-      (compare : A -> A -> comparison)  :=
-  {
-  sto :> StrictOrder lt;
-  wf : well_founded lt;
-  compare_correct :
-    forall alpha beta:A,
-      CompareSpec (alpha=beta) (lt alpha beta) (lt beta alpha)
-                  (compare alpha beta);
-  }.
-\end{Coqsrc}
-
-The following definitions allow us to make implicit several guessable arguments.
-\label{sect:on-lt-notation}
-\label{sect:on-le-notation}
-\begin{Coqsrc}
-Definition on_t  {A:Type}{lt: relation A}
-            {compare : A -> A -> comparison}
-            {on : ON lt compare} := A.
-
-Definition ON_compare {A:Type}{lt: relation A}
-            {compare : A -> A -> comparison}
-            {on : ON lt compare} := compare.
-
-Definition ON_lt {A:Type}{lt: relation A}
-           {compare : A -> A -> comparison}
-           {on : ON lt compare} := lt.
-Infix "o<" := ON_lt : ON_scope.
-
-Definition ON_le  {A:Type}{lt: relation A}
-           {compare : A -> A -> comparison}
-           {on : ON lt compare} :=
-  clos_refl _ ON_lt.
-
-Infix "o<=" := ON_le : ON_scope.
-\end{Coqsrc}
-
-\begin{remark}
-The infix notations \texttt{o<} and \texttt{o<=} were defined in order to make apparent the distinction between the various notation scopes that may co-exist in a same statement.
-So the infix \texttt{<} and \texttt{<=} are reserved to the natural numbers.
-In the mathematical formulas, we still use $<$ and $\leq$ for comparing ordinals.
-\end{remark}
-
-\subsection{Ordinal Notations and Measures for Proving Termination}
-\label{sect:measure-ON}
-
-The following lemma (together with the type class mechnism) allows us to use simply  measures towards an ordinal notation.
-It is just an application of  the libraries \texttt{Coq.Wellfounded.Inverse\_Image}
-and  \texttt{Coq.Wellfounded.Inclusion}.
-
-\begin{Coqsrc}
-Definition measure_lt {A:Type}{lt: relation A}
-            {compare : A -> A -> comparison}
-            {on : ON lt compare}
-            {B : Type} (m : B -> A) : relation B :=
-             fun x y => on_lt (m x) (m y).
-
-Lemma wf_measure  {A:Type}(lt: relation A)
-            {compare : A -> A -> comparison}
-            {on : ON lt compare}
-            {B : Type}
-            (m : B -> A):  well_founded (measure_lt m). 
-\end{Coqsrc}
-
-A simple example of application is given in Sect.~\vref{sect:merge-example}.
-
-\section{Examples of Ordinal Notations}
-
-\subsection{The Ordinal \texorpdfstring{$\omega$}{omega}}
-
-The simplest example of ordinal notation is built over the type \texttt{nat} of \coq's standard library. We have only to apply already proven lemmas about Peano numbers.
-
-\vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations/ON_Omega.html}{OrdinalNotations.ON\_Omega}}
-
-\begin{Coqsrc}
-Global Instance Omega : ON  Peano.lt Nat.compare.
-Proof.
- split.
- - apply Nat.lt_strorder.
- - apply Wf_nat.lt_wf.
- - apply Nat.compare_spec.
-Qed.
-\end{Coqsrc}
-
-\subsection{Sum of two Ordinal Notations}
-
-Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective types \texttt{A} and \texttt{B}.
-
- We consider a new strict order
-on the disjoint sum of the associated types, by putting all elements of \texttt{A} before the elements of \texttt{B} (thanks to Standard Library's relation operator \texttt{le\_AsB}).
-
-\emph{From Library~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}}.
-
-\begin{Coqanswer}
-Inductive
-le_AsB (A B : Type) (leA : A -> A -> Prop) (leB : B -> B -> Prop)
-  : A + B -> A + B -> Prop :=
-| le_aa : forall x y : A, leA x y -> le_AsB A B leA leB (inl x) (inl y)
-| le_ab : forall (x : A) (y : B), le_AsB A B leA leB (inl x) (inr y)
-| le_bb : forall x y : B, leB x y -> le_AsB A B leA leB (inr x) (inr y)
-\end{Coqanswer}
-
-\vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations/ON_plus.html}{OrdinalNotations.ON\_plus}}
-
-\begin{Coqsrc}
-Section Defs.
-
-  Context `(ltA: relation A)
-          (compareA : A -> A -> comparison)
-          (NA: ON ltA compareA).
-  Context `(ltB: relation B)
-          (compareB : B -> B -> comparison)
-          (NB: ON ltB compareB).
-
-Definition t := (A + B)%type.
-Arguments inl  {A B} _.
-Arguments inr  {A B} _.
-
-Definition lt : relation t := le_AsB _ _ ltA ltB.
-\end{Coqsrc}
-
-Before building an instance of \texttt{ON}, we have to define a comparison function.
-
-\begin{Coqsrc}
-Definition compare (alpha beta: t) : comparison :=
-   match alpha, beta with
-     inl _, inr _ => Lt
-   | inl a, inl a' => compareA a a'
-   | inr b, inr b' => compareB b b'
-   | inr _, inl _ => Gt
-  end.
-
-Lemma compare_correct alpha beta :
-    CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
-                            (compare alpha beta).
-\end{Coqsrc}
-
-The Lemma \texttt{Wellfounded.Disjoint\_Union.wf\_disjoint\_sum} of Standard Library
-helps us to prove that our order \texttt{lt} is well-founded.
-
-\begin{Coqsrc}
-Global Instance ON_plus : ON lt compare.
-Proof.
-  split.
-  - apply lt_strorder.
-  -  apply lt_wf.
-  - apply compare_correct.
-Qed.
-\end{Coqsrc}
-
-\subsection{The Ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
-
-The ordinal $\omega+\omega$ (also known as $\omega\times 2$) may be represented as the concatenation 
-of two copies of $\omega$ (Figure~\ref{fig:omega-plus-omega}).
-
-\begin{figure}[h]
-   \centering
-   \begin{tikzpicture}[very thick, scale=0.5]
-\begin{scope}[color=blue]
-\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
-\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
-\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
-\node (Adots) at (6,0) {$\ldots$};
-\node(An) at (8,0)[label=below:$n$]{$\bullet$};
-\node(A2) at (10,0)[label=below:$n+1$]{$\bullet$};
-\node (Adots1) at (12,0) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red]
-\node(B0) at (14,0)[label=below:$0$,label=above:\textcolor{red}{$\omega$}]{$\bullet$};
-\node(B1) at (16,0)[label=below:$1$, label=above:$\omega+1$]{$\bullet$};
-\node(B2) at (18,0)[label=below:$2$,label=above:$\omega+2$]{$\bullet$};
-\node (Bdots) at (20,0) {$\ldots$};
-\node (Bn) at (22,0) [label=below:$p$, label=above:$\omega+p$]{$\bullet$};
-\node (Bdots2) at (24,0) {$\ldots$};
-\end{scope}
-\end{tikzpicture}
-   \caption{\textcolor{blue}{$\omega+{\color{red}\omega}$}}
-   \label{fig:omega-plus-omega}
- \end{figure}
-
-We can define this notation in \coq{} as an instance of \texttt{ON\_plus}.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\begin{Coqsrc}
-Definition Omega_plus_Omega := ON_plus Omega Omega.
-
-Existing Instance Omega_plus_Omega.
-Definition t := @ON_plus.t nat nat.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Example ex1 : inl 7 o< inr 0.
-Proof. constructor. Qed.
-\end{Coqsrc}
-
-We can now define abbreviations.
-For instance, the finite ordinals are represented by terms built with  the constructor \texttt{inl}, and the first infinite ordinal $\omega$ by the term \texttt{(inr 0)}.
-
-\begin{Coqsrc}
-Definition fin (i:nat) : t := inl i.
-Coercion fin : nat >-> t.
-
-Notation "'omega'" := (inr  0:t).
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Example ex1' : fin 7 o< omega.
-Proof. constructor. Qed.
-
-Lemma lt_omega alpha : 
-     alpha o< omega <-> exists n:nat,  alpha = fin n.
-(* ... *)
-\end{Coqsrc}
-
-% \label{warning:coercions}
-% \index{Coq!Techniques!Coercions} 
-% \begin{remark}
-% Beware of coercions and notation scopes!
-% Let us consider the following goal:
-
-% \begin{Coqsrc}
-%  Goal (6 o< 8).
-%  auto with arith.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% 1 subgoal (ID 9)
-
-%   ============================
-%   6 o< 8
-% \end{Coqanswer}
-
-% Please keep in mind that the current notation scope interprets the infix \texttt{``<''} as the predicate \texttt{Omega\_plus\_omega.lt} and not \texttt{Nat.lt}.
-More,  the coercion mechanism converts the terms \texttt{6:nat} [resp.
-\texttt{8:nat} ]
-% into \texttt{inl 6} [resp. \texttt{inl 8}].  So, the initial goal is correctly interpreted by \coq{}, but not as an inequality between two natural numbers.
-
-% \begin{Coqsrc}
-% Set Printing All.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% 1 subgoal (ID 337)
-
-%   ============================
-%   @on_lt nat Peano.lt Nat.compare Omega (S (S (S (S (S (S O))))))
-%     (S (S (S (S (S (S (S (S O))))))))
-% \end{Coqanswer}
-
-% Anyway, the initial goal is provable, using \texttt{le\_AsB}'s first constructor.
-
-% \begin{Coqsrc}
-%   constructor; auto with arith.
-% Qed.
-% \end{Coqsrc}
-
-% \end{remark}
-%
-
-\section{Limits and Successors}
-
-Let us look again at our implementation of $\omega+\omega$. We can distinguish between the three kinds of ordinals seen in Fig~\ref{fig:ordinal-sequence}:
-
-\begin{itemize}
-\item The least ordinal, \texttt{(inl 0)}, also written \texttt{(fin 0)}.
-\item The limit ordinal $\omega$.
-\item The successor ordinals, either of the form \texttt{(inl (S $i$))} or \texttt{(inr (S $i$))}
-\end{itemize}
-
-\subsection{Definitions}
-It would be interesting to specify at the most generic level, what is a zero, a successor or a limit ordinal. Let $<$ be a strict order on a type $A$.
-
-\begin{itemize}
-\item A \emph{least} element is a minorant (in the large sense) of the full set  on $A$,
-\item $y$ is a \emph{successor} of $x$ if $x<y$ and there is no element between $x$ and $y$. We will also say that $x$ is a \emph{predecessor} of $y$.
-\item $x$ is a \emph{limit} if $x$ is not a least element, and for any $y$ such that $yo<x$,
- there exists some $z$ such that $y<z<x$.
-\end{itemize}
-
-The following definitions are in Library \href{../theories/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
-
-\begin{Coqsrc}
-Section A_given.
-  Variables (A : Type)  (lt: relation A).
-
-Local Infix "<" := lt.
-Local Infix "<=" := (clos_refl _ lt).
-
-Definition Least {sto : StrictOrder lt} (x : A):=
-  forall y,  x <= y.
-
-Definition Successor {sto : StrictOrder lt} (y x : A):=
-  x < y /\ (forall z,  x < z ->  z <  y -> False).
-
-Definition Limit {sto : StrictOrder lt}  (x:A)  :=
-  (exists w:A,  w < x) /\
-  (forall y:A, y < x -> exists z:A, y < z /\ z < x).
-\end{Coqsrc}
-
-\index{Exercises}
-\begin{exercise}
-Prove, that, in any ordinal notation system, that every ordinal has at most one predecessor, and at most one successor. 
-\end{exercise}
-
-\index{Exercises}
-\begin{exercise}
-Prove, that, in any ordinal notation system, that if $\beta$ is a successor of $\alpha$,
-then for any $\gamma$, $\gamma<\beta$ implies 
-$\gamma\leq\alpha$.
-\end{exercise}
-
-\subsection{Limits and Successors in \texorpdfstring{$\omega+\omega$}{omega+omega}}
-
-Using the definitions above, we can prove the following lemma:
-
-\vspace{4pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\begin{Coqsrc}
-Lemma limit_iff (alpha : t) : Limit alpha <-> alpha = omega.
-\end{Coqsrc}
-
-Regarding successors, let us introduce the following definition:
-
-\begin{Coqsrc}
-Definition succ (alpha : t) :=
-  match alpha with
-    inl n => inl (S n)
-  | inr n => inr (S n)
-  end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma Successor_correct alpha beta : Successor beta alpha <->
-                                     beta = succ alpha.
-\end{Coqsrc}
-
-We can also check whether an ordinal is a successor by a simple pattern matching:
-
-\begin{Coqsrc}
-Definition succb (alpha: t) : bool := match alpha with
-                                 | inr (S  _) | inl (S _) => true
-                                 | _ => false
-                                 end.
-
-Lemma succb_correct (alpha : t) :
-    succb alpha <->  exists beta: t, alpha = succ beta.
-\end{Coqsrc}
-
-Finally, the nature of any ordinal is decidable(inside this notation system) :
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}}
-\begin{Coqsrc}
-  Definition ZeroLimitSucc_dec {A:Type}{lt: relation A}
-           {compare : A -> A -> comparison}
-           {on : ON lt compare} :=
-  forall alpha,
-    {Least alpha} +
-    {Limit alpha} +
-    {beta: A | Successor alpha beta}.
-(* ... *)
-\end{Coqsrc}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\begin{Coqsrc}
-Definition Zero_limit_succ_dec : ZeroLimitSucc_dec.
-\end{Coqsrc}
-
-\section{The Ordinal \texorpdfstring{$\omega^2$}{omega^2}}
-
-The ordinal $\omega^2$ (also called $\phi_0(2)$, see Chap.~\ref{chap:schutte}), is represented by the lexicographically ordered cartesian product $\mathbb{N}\times\mathbb{N}$.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{ON\_Omega2}}
-
-\begin{Coqsrc}
-Definition t := (nat * nat)%type.
-
-Definition lt : relation t := lexico Peano.lt Peano.lt.
-Definition le := clos_refl _ lt.
-
-Infix "o<" := lt : o2_scope.
-Infix "o<=" := le : o2_scope.
-\end{Coqsrc}
-
-It is easy to represent  finite ordinals, successors and limits.
-
-\begin{Coqsrc}
-Definition zero: t := (0,0).
-
-Definition fin (n:nat) : t := (0, n).
-
-Notation "'F'" := fin : o2_scope.
-
-Coercion fin : nat >-> t.
-
-Definition succ (alpha : t) := (fst alpha, S (snd alpha)).
-
-Notation "'omega'" := (1,0) : o2_scope.
-
-Definition limitb (alpha : t) := 
- match alpha with
- |  (S _, 0) => true
- | _ => false
-end.
-\end{Coqsrc}
-
-In order to build an ordinal notation out of the type \texttt{t}, we have to define a comparison function, and prove that the order \texttt{lt} is well-founded.
-
-\begin{Coqsrc}
-Instance lt_strorder : StrictOrder lt.
-(* ... *)
-
-Definition compare (alpha beta: t) : comparison :=
-  match Nat.compare (fst alpha) (fst beta) with
-    Eq => Nat.compare (snd alpha) (snd beta)
-  | c => c
-  end.
-
-Lemma compare_correct alpha beta :
-    CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
-                (compare alpha beta).
-
-Lemma lt_wf : well_founded lt.
-
-Instance Omega2: ON lt compare.
-Proof.
-  split.
-  - apply lt_strorder.
-  - apply lt_wf.
-  - apply compare_correct.
-Qed.
-\end{Coqsrc}
-
-\subsection{Arithmetic of \texorpdfstring{$\omega^2$}{omega^2}}
-
-\subsubsection{Addition}
-
-We can define on \texttt{ON\_omega2} an addition which extends the addition on \texttt{nat}.
-
-\begin{Coqsrc}
-Definition  plus (alpha beta : t) : t :=
-  match alpha,beta with
-  | (0, b), (0, b') => (0, b + b')
-  | (0,0), y  => y
-  | x, (0,0)  => x
-  | (0, b), (S n', b') => (S n', b')
-  | (S n, b), (S n', b') => (S n + S n', b')
-  | (S n, b), (0, b') => (S n, b + b')
-   end.
-
-Infix "+" := plus : o2_scope.
-\end{Coqsrc}
-
-Please note that this operation is not commutative:
-
-\begin{Coqsrc}
-Example non_commutativity_of_plus :  omega + 3 <> 3 + omega.
-Proof.
-  cbn.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-1 subgoal (ID 237)
-
-  ============================
-(1, 3) <> omega.
-\end{Coqanswer}
-
-\begin{Coqsrc}
- discriminate.
-Qed.
-\end{Coqsrc}
-
-\subsubsection{Multiplication}
-
-The restriction of ordinal multiplication to the segment $[0,\omega^2)$ is not a total function.
-For instance $\omega\times\omega= \omega^2$ is outside the set of represented values.
-Nevertheless, we can define two operations mixing natural numbers and ordinals.
-
-\begin{Coqsrc}
-(** multiplication of an ordinal by a natural number *)
-
-Definition mult_fin_r  (alpha : t) (p : nat): t :=
-  match alpha, p with
- |  (0,0), _  => zero
- |  _, 0 => zero
- |  (0, n), p => (0, n * p)
- |  ( n, b),  n' => ( n *  n', b)
-               end.
-Infix "*" := mult_fin_r : o2_scope.
-
-(** multiplication of  a natural number by an ordinal *)
-
-Definition mult_fin_l (n:nat)(alpha : t) : t :=
-  match n, alpha with
- |  0, _  => zero
- |  _, (0,0) => zero
- |   n , (0,n') => (0, (n*n')%nat)
- |  n, (n',p') => (n', (n * p')%nat)
- end.
-
-Example e1 : (omega * 7 + 15) * 3 = omega * 21 + 15.
-Proof. reflexivity. Qed.
-
-Example e2 :  mult_fin_l 3 (omega * 7 + 15) = omega * 7 + 45.
-Proof. reflexivity. Qed.
-\end{Coqsrc}
-
-Multiplication with a finite ordinal and addition are related through the following lemma:
-
-\begin{Coqsrc}
-Lemma unique_decomposition alpha : 
-    exists! i j: nat,  alpha = omega * i + j.
-\end{Coqsrc}
-
-\subsection{A Proof of Termination using \texorpdfstring{$\omega^2$}{omega^2}} 
-\label{sect:merge-example}.
-
-Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily a total function which merges two lists.
-
-\index{Coq!Commands!Function}
-
-\begin{Coqsrc}
-(* adapted from Pascal Manoury et al. *)
-
-Require Import Coq.Program.Wf List.
-Require Import FunInd Recdef.
-
-Section Merge.
-
-  Variable A: Type.
-
-  Local Definition m (p : list A * list A) :=
-    omega * length (fst p) + length (snd p).
-
-  Function  merge  (ltb: A -> A -> bool)
-          (xys: list A * list A)
-          {wf (measure_lt m) xys} :
-    list A :=
-    match xys with
-      (nil, ys) => ys
-    | (xs, nil) => xs
-    | (x :: xs, y :: ys) =>
-      if ltb x y then x :: merge  ltb (xs, (y :: ys))
-      else y :: merge  ltb ((x :: xs), ys)
-    end.
-
-  - intros; unfold m, measure_lt; cbn; destruct xs0; simpl; left; lia.
-  - intros; unfold m, measure_lt; cbn; destruct ys0; simpl; right; lia.
-   - auto.
-  Defined.
-
-End Merge.
-
-Goal forall l, merge nat Nat.leb (nil, l) = l.
-  intro; now rewrite merge_equation.
-Qed.
-\end{Coqsrc}
-
-\subsection{Yet Another  Proof of Impossibility}
-\label{omega2-case}
-
-In Sect.~\vref{omega-case}, we proved that there exists no variant towards \texttt{nat}
-(\emph{i.e.} the ordinal $\omega$) for proving the termination of all hydra battles.
-We  prove now that  the ordinal $\omega^2$ is also insufficient for this purpose.
-
-The proof we are going to develop has exactly the same structure as in Section~\ref{omega-case}.
- Nevertheless, the proof of technical  lemmas is a little more complex, due to 
- the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
-Consider for instance that there exists an infinite number of ordinals  between
-$\omega$ and $\omega\times 2$.
-
-The detailed  proof script is in the file \url{../theories/html/hydras.Hydra.Omega2_Small.html}.
-
-\subsubsection{Preliminaries}
-Let us assume there is a variant from \texttt{Hydra} into $\omega^2$  for proving the   termination of all hydra battles.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
-
-\begin{Coqsrc}
-  Section Impossibility_Proof.
-
- Variable m : Hydra -> ON_Omega2.t.
-
- Context (Hvar : Hvariant lt_wf free m).
-\end{Coqsrc}
-
-Let us follow the same pattern as in Sect.~\ref{omega-case}.
-First, we define an injection $\iota$ from type \texttt{t} into \texttt{Hydra}, by
- associating to  each ordinal $\omega\times i+ j = (i,j)$ the hydra with $i$ branches of length $2$ and
-$j$ branches of length $1$.
-
-%% revenir ici
-
-\vspace{4pt}
-\emph{From Module ~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
-
-\begin{Coqsrc}
-Let iota (p: ON_Omega2.t) := 
-    node (hcons_mult (hyd1 head) (fst p)
-                     (hcons_mult head (snd p) hnil)).
-  \end{Coqsrc}
-
-For instance, Figure~\vref{fig:essai2} shows the hydra associated to the ordinal 
-$(3,5)$, a.k.a. $\omega\times 3 + 5$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-\node (foot) at (6,0) {$\bullet$};
-\node (N1) at (1,3) {$\bullet$};
-\node (N2) at (3,3) {$\bullet$};
-\node (N3) at (5,3) {$\bullet$};
-\node (N4) at (8,3) {$\Smiley[2][green]$};
-\node (N5) at (11,3) {$\Smiley[2][green]$};
-\node (N6) at (14,3) {$\Smiley[2][green]$};
-\node (N7) at (17,3){$\Smiley[2][green]$};
-\node (N8) at (20,3){$\Smiley[2][green]$};
-\node  (N9) at (0,5) {$\Smiley[2][green]$};
-\node (N10) at (2,5) {$\Smiley[2][green]$};
-\node (N11) at (4,5) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N5);
-\draw (foot) -- (N6);
-\draw (foot) to [bend right=10] (N7);
-\draw (foot) to [bend right=15] (N8);
-\draw (N1) to [bend left=10] (N9);
-\draw (N2) -- (N10);
-\draw (N3) -- (N11);
-\end{tikzpicture}
-\caption{\label{fig:essai2}
-The hydra $\iota(\omega\times 3+5)$}
-\end{figure}
-
-Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:h-omega2-small}}
- The hydra \texttt{big\_h}.
-\end{figure}
-
-\begin{Coqsrc}
-   Let big_h := hyd1 (hyd2 head head).  
-\end{Coqsrc}
-
- In a second step, we build a ``smaller'' hydra.
-
-\begin{Coqsrc}
-   Let small_h := iota (m big_h).
-\end{Coqsrc}
-
-Like in Sect.~\ref{omega-case}, we prove the double inequality \texttt{m big\_h o<= m small\_h o< m big\_h}, which is impossible.
-
-\subsubsection{Proof of the Inequality \texttt{m small\_h o< m big\_h}}
-
-In order to prove the inequality  \texttt{m\_lt: m small\_h o< m big\_h}, it suffices to
-build a battle transforming \texttt{big\_h} into \texttt{small\_h}.
-
-First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h} as $(i,j)$.
-If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
-If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the following result.
-
-\begin{Coqsrc}
-Lemma big_to_small: big_h -+-> small_h.
-\end{Coqsrc}
-
-Since $m$ is a variant, we infer the following inequality:
-
-\begin{Coqsrc}
-Corollary m_lt : m small_h o< m big_h.
-\end{Coqsrc}
-
-\subsubsection{Proof of the Inequality \texttt{m big\_h o<= m small\_h} }
-
-The proof of the inequality \texttt{m big\_h o<= m small\_h} is quite more complex than in Sect~\ref{omega-case}.
-If we consider any ordinal $\alpha=(i,j)$, where $i>0$, there exists an infinite number of
-ordinals stricly less than $\alpha$, and there exists an infinite number of battles that start from
-$\iota(\alpha)$.
-Indeed, at any configuration $\iota(k,0)$, where $k>0$, the hydra can freely choose any replication number.
-Intuitively, the measure of such a hydra must be large enough for taking into account
-all the possible battles issued from that hydra.
-Let us now give more technical details.
-
-\begin{itemize}
-\item The proof of the lemma \texttt{m\_ge : m big\_h  o<= m small\_h} uses well-founded induction on \texttt{big\_h}.
-
-\item For any pair $p$, we have to distinguish between three cases, according to the value of $p$'s components.
-  \begin{itemize}
-  \item $p=(0,0)$
-  \item $p=(i,0)$, where $i>0$\,: $p$ corresponds to a limit ordinal
-  \item $p=(i,j)$, where $j>0$\,: $p$ is the successor of $(i,j-1)$.
-  \end{itemize}
-\end{itemize}
-
-Let us define the notion of elementary ``step'' of decreasing sequences in
-\texttt{t}
-
-\begin{Coqsrc}
-Inductive step : t -> t -> Prop :=
-| succ_step : forall i j,  step (i, S j) (i, j)
-| limit_step : forall i j, step (S i, 0) (i, j).
-\end{Coqsrc}
-
-The following lemma establishes a correspondance between the relation
-\texttt{step} and hydra battles.
-
-\begin{Coqsrc}
-Lemma step_to_battle : forall p q, step p q -> iota p -+-> iota q.
-\end{Coqsrc}
-
-\index{Maths!Transfinite induction}
-
-Thus, starting from any inequality $q < p$ on type \texttt{t}, we can build 
-by \emph{transfinite induction} (\emph{i.e.} well-founded) over \texttt{p} a battle 
-that transforms the hydra $\iota(p)$ into $\iota(q)$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}
-
-\begin{Coqsrc}
-Lemma m_ge : forall p : t,   p o<= m (iota p).
-Proof.
-  unfold small_h; pattern (m big_h) .   
-   apply  well_founded_induction with (R := lt) (1:= lt_wf).
-  intro p ; pattern p;
-    apply  well_founded_induction with 
-               (R := lt2) (1:= wf_lexico lt_wf lt_wf);
-     intros (i,j) IHij. 
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  i, j : nat
-  IHij : forall y : t, y o< (i, j) -> y o<= m (iota y)
-  ============================
-  (i, j) o<= m (iota (i, j)) 
-\end{Coqanswer}
-
-Then we have  three cases to consider, according to the value of $p$.
-\begin{itemize}
-\item If $p=(0,0)$ then obviously, $\iota(p)\geq p = (0,0)$
-\item If  $p=(i+1,0)$ for some $i\in\mathbb{N}$, we
- remark  that $p$ is strictly greater than any pair $ (i, j)$, where $j$ 
-is any natural number.
-
-Applying the battle rules, for any $j$, we have $\iota(i+1,j)  {\round} \iota(i, j) $, thus $m(\iota(p)) > m(\iota(i,j)$ since  $m$ is assumed to be a variant.
-
-Applying the induction hypothesis, we get the inequality
- $ m(\iota(i,j)) \geq (i,j)$ for any $j$.
-
-Thus, $m(\iota(p)) > (i,j)$ for any $j$.
-Applying the lemma \texttt{limit\_is\_lub}, we get  the inequality
-$m(\iota(i+1,0))\geq (i+1,0)$
-
-\item If $p=(i,j+1)$ with $j\in\mathbb{N}$, we have  $\iota(p)  {\round} \iota(i, j) $,
-hence $m(\iota(p))> m(\iota(i,j)) \geq (i,j)$, thus $m(\iota(p))\geq (i,j+1)=p$
-
-\end{itemize}
-
-\begin{Coqsrc}
-  (* ... *)
-Qed.
-\end{Coqsrc}
-
-\subsubsection{End of the Proof}
-From \texttt{m\_ge}, we get \texttt{m big\_h o<= m small\_h = m (iota (m big\_h)) }. 
-Since $<$ is a strict order (irreflexive  and transitive), this inequality is incompatible with the strict inequality  \texttt{m small\_h o< m big\_h} (lemma \texttt{m\_lt}).
-
-\vspace{4pt}
-\noindent
-In \coq (Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#Impossible}{Hydra.Omega2\_Small}):
-
-\begin{Coqsrc}
-Theorem Impossible: False.
-Proof.
-  destruct (StrictOrder_Irreflexive (m big_h)).
-  apply le2_lt2_trans with (m small_h).
-  -  unfold small_h; apply m_ge.
-  -  apply m_lt. 
-Qed. 
-End Impossibility_Proof.
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into $\omega^2$ for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\begin{remark}
-In Chapter~\ref{ks-chapter}, we  prove a generalization of the impossibility lemmas of
-Sect.~\ref{omega-case} and this section, with the same proof structure, but with much more 
-complex technical details.
- \end{remark}
-
-\index{Exercises}
-\begin{exercise}
-
-\label{sec:orgheadline63}
-Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.
-
-\begin{itemize}
-\item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
-  lexicographic ordering
-\item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.
-
-For instance, the following inequality holds:
-\[\langle 4,3,3,3,3,3,3,2,2,2 \rangle\,<\,\langle 4,4,2 \rangle\]
-\end{itemize}
-
-\end{exercise}
-
-\section{A Notation for Finite Ordinals}
-
-Let $n$ be some natural number. The segment associated with $n$ is the interval 
-$[0,n)\,=\,\{0,1,\dots,n-1\}$. 
-One may represent the ordinal $n$ by a sigma type.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
-
-\label{def: Finite-ord-type}
-\begin{Coqsrc}
-Coercion is_true: bool >-> Sortclass.
-
-Definition t (n:nat) := {i:nat | Nat.ltb i n}.
-\end{Coqsrc}
-
-The order on type (\texttt{t $n$}) is defined through the projection on \texttt{nat}.
-
-\begin{Coqsrc}
-Definition lt {n:nat} : relation (t n) :=
-  fun alpha beta => Nat.ltb ( proj1_sig alpha) (proj1_sig beta).
-\end{Coqsrc}
-
-For instance, let us build two elements of the segment $[0, 7)$, \emph{i.e.} two
-inhabitants of   type (\texttt{t 7}), and prove a simple  inequality (see Fig.~\ref{fig:O7}).
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (N0) at (0,0) {$\bullet$};
-\node (i0) at (0,1) {$0$};
-\node (N1) at (2,0) {$\bullet$};
-\node (i1) at (2,1) {$1$};
-\node (N2) at (4,0) {$\bullet$};
-\node (i2) at (4,1) {$2$};
-\node (N3) at (6,0) {$\bullet$};
-\node (i3) at (6,1) {$3$};
-\node (N4) at (8,0) {$\bullet$};
-\node (i4) at (8,1) {$4$};
-\node (N5) at (10,0) {$\bullet$};
-\node (i5) at (10,1) {$5$};
-\node (N6) at (12,0) {$\bullet$};
-\node (i6) at (12,1) {$6$};
-\node(alpha1) at (4,-1) {$\alpha_1$};
-\node(alpha2) at (10,-1) {$\beta_1$};
-\end{tikzpicture}
-
-\caption{The segment $\mathbb{O}_7$\label{fig:O7}}
-\end{figure}
-
-\index{Coq!Commands!Program}
-
-\begin{Coqsrc}
-Program Example alpha1 : t 7 := 2.
-
-Program Example beta1 : t 7 := 5.
-
-Example i1 : lt  alpha1 beta1.
-Proof. now compute. Qed.
-\end{Coqsrc}
-
-Note that the type (\texttt{t 0}) is empty, and that, for any natural number
- $n$, $n$ does not belong to (\texttt{t $n$}).
-
-\begin{Coqsrc}
-Lemma t0_empty (alpha: t 0): False.
-Proof.
-  destruct alpha.
-  destruct x; cbn in i; discriminate.
-Qed.
-
-Program Definition bad : t 10 := 10.
-Next Obligation.
-  compute.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-1 subgoal (ID 162)
-
-  ============================
-  false = true
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Abort.
-\end{Coqsrc}
-
-Note also that attempting to compare a term  of type (\texttt{t $n$}) with a term of
-type (\texttt{t $p$})  leads to an error if $n$ and $p$ are not convertible.
-
-\begin{Coqsrc}
-
-Program Example gamma1 : t 8 := 7.
-
-Fail Goal lt alpha1 gamma1.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- The command has indeed failed with message:
-The term "gamma1" has type "t 8" while it is expected to have type "t 7".
-\end{Coqanswer}
-
-In order to build an instance of \texttt{OrdinalNotation}, we define a comparison function, by delegation to standard library's  \texttt{Nat.compare}, and prove its correction.
-
-\begin{Coqsrc}
-Definition compare {n:nat} (alpha beta : t n) :=
-  Nat.compare (proj1_sig alpha) (proj1_sig beta).
-
-Lemma compare_correct {n} (alpha beta : t n) :
-  CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
-              (compare alpha beta).
-\end{Coqsrc}
-
-\begin{remark}
- The proof of \texttt{compare\_correct} uses a well-known pattern of \coq{}.
-Let us consider  the following subgoal.
-
-\begin{Coqanswer}
- 1 subgoal (ID 110)
-
-  n, x0 : nat
-  i, i0 : x0 <? S n
-  ============================
-  exist (fun i1 : nat => i1 <=? n) x0 i =
-  exist (fun i1 : nat => i1 <=? n) x0 i0
-\end{Coqanswer}
-
-Applying the tactic \texttt{f\_equal} generates a simpler subgoal.
-
-\begin{Coqanswer}
-1 subgoal (ID 112)
-
-  n, x0 : nat
-  i, i0 : x0 <? S n
-  ============================
-  i = i0
-\end{Coqanswer}
-
-We have now to prove that there exists at most one  proof of (\texttt{Nat.ltb x0 (S n)}). This is not obvious, but  a consequence of the following lemma of library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Eqdep_dec.html}{Coq.Logic.Eqdep\_dec}.
-
-\index{Coq!Techniques!Unicity of equality proofs}
-\label{sect:eq-proof-unicity}
-
-\begin{Coqanswer}
-eq_proofs_unicity_on :
-forall (A : Type) (x : A),
-(forall y : A, x = y \/ x <> y) -> 
-forall (y : A) (p1 p2 : x = y), p1 = p2
-\end{Coqanswer}
-
-Thus unicity of proofs of \texttt{Nat.ltb x0 (S n)}  comes from the decidability of
-equality on type \texttt{bool}.
-This is why we used the boolean function \texttt{Nat.ltb} instead of the inductive predicate \texttt{Nat.lt} in the definition of type \texttt{t $n$} (see page~\pageref{def: Finite-ord-type}).
-For more information about this pattern, please look at the numerous mailing lists and 
-FAQs on \coq{}).
-
-\end{remark}
-
-Applying lemmas of the libraries \texttt{Coq.Wellfounded.Inverse\_Image}, \linebreak
- \texttt{Coq.Wellfounded.Inclusion}, and \texttt{Coq.Arith.Wf\_nat}, we prove that our
-relation \texttt{lt} is well founded.
-
-\begin{Coqsrc}
-Lemma lt_wf (n:nat) : well_founded (@lt n).
-\end{Coqsrc}
-
-Now we can build our instance of \texttt{OrdinalNotation}.
-
-\begin{Coqsrc}
-Global Instance sto n : StrictOrder (@lt n).
-
-Global Instance FinOrd (n:nat) : OrdinalNotation (sto n) compare.
-Proof.
-  split.
-  - apply compare_correct.
-  - apply lt_wf.
-Qed.
-\end{Coqsrc}
-
-\begin{remark}
-It is important to keep in mind  that the integer $n$ is not an ``element'' of \texttt{FinOrd $n$}.
-In set-theoretic presentations of ordinals, the set associated with the ordinal $n$ is $\{0,1,\dots,n-1\}$.
-In our formalization, the interpretation of an ordinal as a set is realized by the following definition
-(in ~\href{../theories/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}).
-
-\begin{Coqsrc}
-Definition bigO `{nA : @OrdinalNotation A ltA stoA compareA}
-           (a: A) : Ensemble A :=
-  fun x: A => ltA x a.
-\end{Coqsrc}
-\end{remark}
-
-\begin{remark}
- There is no interesting arihmetic on finite ordinals, since functions like successor, addition, etc.,  cannot be represented in \coq{} as \emph{total} functions.
-\end{remark}
-
-\begin{remark}
-Finite ordinals are also formalized in MathComp~\cite{SSR}.  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.  
-\end{remark}
-
-%%%
-
-\section{Comparing two Ordinal Notations}
-
-It is sometimes useful to compare two ordinal notations with respect to expressive power
-(the segment of ordinals  they represent).
-
-The following class specifies a strict inclusion of segments.
-The notation \texttt{OA} describes a segment $[0,\alpha($, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented in \texttt{OA}).
-We require also  that the comparison functions of the two notation systems are compatible.
-
-If \texttt{OB} is presumed to be correct, the we can consider that \texttt{OA} ``inherits'' its correcteness from the bigger notation system \texttt{OB}.
-
-\label{types:SubON}
-
-\begin{Coqsrc}
-Class  SubON 
-       `(OA : @ON A ltA  compareA)
-       `(OB : @ON B ltB  compareB)
-       (alpha :  B)
-       (iota : A -> B):=
-  {
-  SubON_compare: forall x y : A,  compareB (iota x) (iota y) =
-                                 compareA x y;
-  SubON_incl : forall x, ltB (iota x) alpha;
-  SubON_onto : forall y, ltB y alpha  -> exists x:A, iota x = y}.
-\end{Coqsrc}
-
-For instance, we prove that \texttt{Omega} is a sub-notation of
-\texttt{Omega\_plus\_Omega} (with $\omega$ as the first ``new'' ordinal, and \texttt{fin} as the injection).
-
-\begin{Coqsrc}
-Instance Incl : SubON Omega Omega_plus_Omega omega fin.
-\end{Coqsrc}
-
-We can also show that, if $i<j$, then the segment $[0,i)$ is a ``sub-segment'' of
-$[0,j)$. Since the terms  ($t\;i$) and ($t\;j$) are not convertible, we consider a ``cast'' 
-function $\iota$ from ($t\;i$) into ($t\;j$), and prove that this function is  a monotonous bijection  from ($t\;i$) to
-the segment $[0,i)$ of ($t\;j$).
-
- \begin{figure}[h]
-   \centering
-   \begin{tikzpicture}[very thick, scale=0.6]
-\begin{scope}[color=blue]
-\node (A) at (0,0) {$A$};
-\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
-\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
-\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
-\node (Adots) at (6,0) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red]
-\node (B) at (0,2) {$B$};
-\node(B0) at (2,2)[label=above:$0$]{$\bullet$};
-\node(B1) at (3,2)[label=above:$1$]{$\bullet$};
-\node(B2) at (4,2)[label=above:$2$]{$\bullet$};
-\node (Bdots) at (6,2) {$\ldots$};
-\node (b) at (8,2) [label=above:$b$]{$\bullet$};
-\node (bsucc) at (9,2) [label=above:$b+1$]{$\bullet$};
-\node (Bdots2) at (10,2) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red!50!blue]
-\draw [->,thin] (A0) -- node [auto] {$\iota$} (B0);
-\draw [->,thin] (A1) -- node [auto] {$\iota$} (B1);
-\draw [->,thin] (A2) -- node [auto] {$\iota$} (B2);
-\draw [->,thin] (Adots) -- node [auto] {$\iota$} (Bdots);
-\end{scope}
-\end{tikzpicture}
-   \caption{\textcolor{blue}{$A$} is a sub-segment  of \textcolor{red}{$B$}}
-   \label{fig:subsegment}
- \end{figure}
-
-\index{Coq!Commands!Program}
-
-We are now able to build an instance of \texttt{SubON}.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
-
-\begin{Coqsrc}
-Section Inclusion_ij.
-
-  Variables i j : nat.
-  Hypothesis Hij : (i < j)%nat.
-
-  Remark Ltb_ij : Nat.ltb i j.
-
-  Program Definition iota_ij  (alpha: t i) : t j :=  alpha.
-
-   Let b : t j := exist _ i Ltb_ij.
-
-   Global Instance F_incl_ij  : SubON  (FinOrd i) (FinOrd j) b iota_ij.
-  (* ... *)
-
-  End Inclusion_ij.
-\end{Coqsrc}
-
-\index{Exercises}
-\begin{exercise}
-Prove that \texttt{Omega\_plus\_Omega} cannot be a sub-notation of \texttt{Omega}.
-\end{exercise}
-
-\index{Projects}
-\begin{project}
-Adapt the definition of \texttt{Hvariant} (Sect.~\ref{sect:hvariant-def}) in order to
-have an ordinal notation as argument. Prove that if $O_A$ is a sub-notation of $O_B$, then any variant defined on  $O_A$ can be automatically transformed into 
-a variant on $O_B$.
-\end{project}
-
-\section{Comparing an Ordinal Notation with Schütte's Model}
-
-Finally, it may be interesting to compare an ordinal notation with the more theoretical model from Schütte (well, at least with our formalization of that model).
-This would be a relative proof of correctenss of the considered  ordinal  notation.
-
-The following class specifies that a notation \texttt{OA} describes a segment $[0,\alpha)$,
-where $\alpha$ is a countable ordinal \emph{à la}  Schütte.
-
-\label{types:ON-for}
-
-\begin{Coqsrc}
-Class ON_correct `(alpha : Schutte_basics.Ord)
-     `(OA : @ON A ltA  compareA)
-      (iota : A -> Schutte_basics.Ord) :=
-  { ON_correct_inj : forall a, Schutte_basics.lt (iota a) alpha;
-    ON_correct_onto : forall beta, Schutte_basics.lt beta alpha ->
-                                exists b, iota b = beta;
-    On_compare_spec : forall a b:A,
-        match compareA a b with
-          Datatypes.Lt => Schutte_basics.lt (iota a) (iota b)
-        | Datatypes.Eq => iota a = iota b
-        | Datatypes.Gt => Schutte_basics.lt (iota b) (iota a)
-        end}.
-\end{Coqsrc}
-
-For instance, the following theorem tells that \texttt{Epsilon0}, our notation system for the segment $[0,\epsilon0)$ is a correct implementation of the theoretically defined  ordinal $\epsilon_0$
-(see chapter~\ref{chap:schutte} for more details).
-
-\begin{Coqsrc}
-Instance Epsilon0_correct :
-  ON_correct epsilon0 Epsilon0  (fun alpha => inject (cnf alpha)).
-\end{Coqsrc}
-
-\index{Projects}
-
-\begin{project}
-  When you have read Chapter~\ref{chap:schutte}, prove that the sum of two ordinal notations \texttt{ON\_plus} implements the addition of ordinals.
-\end{project}
-
-\section{Isomorphism of Ordinal Notations}
-
-In some cases we want to show that two notation systems describe the same segment (for instance $[0,3+\omega)$ and $[0,\omega)$\;).
-For this purpose, one may prove that the two notation systems are order-isomorphic.
-
-\label{types:ON-iso} 
-\begin{Coqsrc}
-Class  ON_Iso 
-       `(OA : @ON A ltA compareA)
-       `(OB : @ON B ltB  compareB)
-       (f : A -> B)
-       (g : B -> A):=
-  {
-  iso_compare: forall x y : A, 
-      compareB (f x) (f y) = compareA x y;
-  iso_inv1 : forall a, g (f a)= a;
-  iso_inv2 : forall b, f (g b) = b}.
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
-Let $i$ be some natural number. Prove that the notation systems 
-\texttt{Omega} and (\texttt{ON\_plus (OrdFin $i$) Omega}) are isomorphic.
-
-{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$, that we prove also in larger notation systems, as well as in Schütte's model.}
-
-This exercise is partially solved for $i=3$ (in ~\href{../theories/html/hydras.OrdinalNotations.Example_3PlusOmega.html}{OrdinalNotations.Example\_3PlusOmega}).
-
-\end{exercise}
-
-\index{Projects}
-\label{exo:ON-mult}
-\begin{project}
-Define in \coq{} the product of two ordinal notations $N_A$ and $N_B$.
-If $A$ [resp. $B$] is the underlying type of $N_A$ [resp. $N_B$], the
-product \texttt{ON\_mult $N_A$ $N_B$} is implemented over the cartesian product $B\times A$ (with the lexicographic ordering).
-
-For instance, the
-elements of the product (\texttt{ON\_mult Omega (FinOrd 3)}) are ordered as follows.
-\[(0,0),(0,1),(0,2),(0,3),(0,4),\dots,{\color{red}(1,0),} (1,1),(1,2),\dots, {\color{red}(2,0)},(2,1),(2,2),\dots\]
-
-Note that the elements of  (\texttt{ON\_mult (FinOrd 3) Omega}) are differently ordered (without limit ordinals):
-\[(0,0),(1,0),(2,0),(0,1),(1,1),(2,1),(0,2),(1,2),(2,2),(0,3),\dots\]
-
-Prove that \texttt{ON\_mult (FinOrd $i$) Omega} is isomorphic to
-\texttt{Omega}  whilst
-\texttt{Omega}  is a sub-notation of \texttt{ON\_mult Omega (FinOrd $i$)},
-for any strictly positive $i$.
-\end{project}
-
-\index{Projects}
-\begin{project}
-Consider two isomorphic ordinal notations \texttt{OA} and \texttt{OB}.
-Prove that, if \texttt{OA} [resp. \texttt{OB}] is a correct implementation 
-of $\alpha$ [resp. $\beta$], then $\alpha=\beta$.
-\end{project}
-
-\index{Projects}
-\begin{project}
-\label{project:succ-limit-dec}
-Add to the class \texttt{ON} the requirement that for any $\alpha$ it is decidable whether $\alpha$ is $0$, a successor or a limit ordinal.
-
-\textbf{Hint:}   Beware of the instances associated with sum and product of notations!
-  You may consider additional fields 
-to make the sum and product of notations ``compositional''.
-
-\end{project}
-
-\index{Projects}
-\begin{project}
-\label{project:on-setoid}
-Reconsider the  class \texttt{ON}, with an equivalence instead of Leibniz equality.
-\end{project}
-
-%%%% ICI ICI
-
-\section{Other Ordinal Notations}
-
-\index{Projects}
-
-\begin{project}
-The directory \texttt{theories/OmegaOmega} contains an ad-hoc formalization of $\omega^\omega$, contributed by Pascal Manoury.
-Every ordinal $\alpha$ is represented by a list $l$ whose elements are the coefficients of $\omega$ in  the Cantor normal form of $\alpha$ (in reverse order).
-For instance, the ordinal
-$\omega^{8}\times 5 + \omega^{6}\times 8 + \omega^2\times 10 + \omega + 7$ is represented by the list \texttt{[5;0;8;0;0.0;10,1,7]}.
-
- Develop this representation and compare it with the other ordinal notations.
-
-\end{project}
-
-\index{Projects}
-
-\begin{project}
-Let $N_A$ be a notation system for ordinals strictly less than $\alpha$, 
-with the strict order $(A,<_A)$. Please build the notation system
-\texttt{ON\_Expl $N_A$}, on the type of multisets of elements of $A$
-(or, if preferred, the type of non-increasing finite sequences on $A$,
-provided with the lexicographic ordering on lists).
-
-For instance, let us take $N_A=\texttt{Omega}$, and take $\alpha=\langle 4,4,2,1,0\rangle$,
- $\beta=\langle 4,3,3,3,3,3,2\rangle$, and $\gamma=\langle 5\rangle$. Then $\beta<\alpha<\gamma$.
-
-In contrast the list $\langle5,6,3,3\rangle$ is not non-increasing (\emph{i.e.} sorted w.r.t. $\geq$), so it is not to be considered.
-
-Note that if the notation $N_A$ implements the ordinal 
-$\alpha$,  the new notation $\omega^{N_A}$ must implement the ordinal $\phi_0(\alpha)$, a.k.a. $\omega^\alpha$ (see chapter~\ref{chap:schutte})
-
-\end{project}
-
-\begin{remark}
- The set of ordinal terms in Cantor normal form (see Chap.~\ref{chap:T1}) and 
-in Veblen normal form (see 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}{Gamma0.Gamma0}) are shown to be ordinal notation systems, but there is a lot of work to be done in order to unify ad-hoc  definitions and proofs which were written before the definition of the \texttt{ON} type class.
-\end{remark}
-
-%------------------------------------------------------------------------
-
-\chapter[A proof of termination, using epsilon0]{A Proof of Termination, using Ordinals below \texorpdfstring{$\epsilon_0$}{Epsilon0}}
-
-\label{cnf-math-def}
-\label{chap:T1}
-
-In this chapter, we adapt to \coq{} the well-known~\cite{KP82}  proof that Hercules eventually wins every battle, whichever the strategy  of each player.
-In other words, we present  a formal and self contained proof of termination  of all [free] hydra battles.
-First, we take from Manolios and Vroon~\cite{Manolios2005} a representation of the ordinal $\epsilon_0$ as terms in Cantor normal form.
-Then, we define a variant for hydra battles as a measure that maps any hydra to some ordinal strictly less than $\epsilon_0$.
-
-\section{The Ordinal \texorpdfstring{\(\epsilon_0\)}{epsilon0}}
-\label{sec:epsilon0-intro}
-
-\subsection{Cantor Normal Form}
-\index{Maths!Cantor normal form}
-
-The ordinal \(\epsilon_0\) is the least ordinal number that satisfies 
-the equation \(\alpha = \omega^\alpha\), where \(\omega\) is 
-the least infinite ordinal. Thus, we can consider \(\epsilon_0\) as an
-\emph{infinite} \(\omega\)-tower.
-Nevertheless, 
-any ordinal strictly less that \(\epsilon_0\) 
-can be finitely represented by a unique  \emph{Cantor normal form}, 
-that is, an expression  which is either  the ordinal \(0\) or 
-a sum  \(\omega^{\alpha_1} \times n_1 + \omega^{\alpha_2} \times n_2 + 
-  \dots + \omega^{\alpha_p} \times n_p\) where all the \(\alpha_i\) 
-are ordinals in Cantor  normal form, \(\alpha_1 > \alpha_2 > \alpha_p\), 
-and all the \(n_i\) are positive integers.
-
-An example of Cantor normal form is displayed in Fig \ref{fig:cnf-example}:
-Note that  any ordinal of
-the form \(\omega^0 \times i + 0\) is just written \(i\).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
-\node[color=blue]{$\omega^{(\omega^\omega\,+\, \omega^2 \times 8 \,+\, \omega)}+ \omega^\omega + \omega^4+ 6$};
-\end{tikzpicture}
-\caption{\label{fig:cnf-example}
-An ordinal in Cantor normal form}
-\end{figure}
-
-In the rest of this section, we define an inductive type for representing in \texttt{Coq}
-all the ordinals strictly  less than  \(\epsilon_0\), then extend some arithmetic operations
-to this type, and finally prove that our representation fits well with 
-the expected mathematical properties: the order we define is a well order, 
-and the decomposition into Cantor normal form  is consistent 
-with the implementation of the arithmetic operations of exponentiation of base \(\omega\) 
-and addition.
-
-\paragraph*{Remark}
-\label{sec:orgheadline65}
-Unless explicitly mentionned, the term ``ordinal" will be used instead of
-``ordinal strictly less than \(\epsilon_0\)" (except in Chapter~\ref{chap:schutte} where it stands for ``countable ordinal'').
-
-\subsection{A Data Type for  Ordinals in Cantor Normal Form}
-\label{sec:orgheadline72}
-\label{sec:T1-inductive-def}
-
-% Our user contribution~\cite{CantorContrib} represents 
-% the set of ordinals strictly less than $\epsilon_0$ in Cantor normal form as in~\cite{Manolios2005}, and also the set
-% of ordinals strictly  less than $\Gamma_0$ in Veblen normal form.
-
-    Let us define an inductive type whose 
-constructors are respectively associated
-with the ways to build Cantor normal forms:
-
-\begin{itemize}
-\item the ordinal \(0\)
-\item the construction \((\alpha,\, n,\,\beta)  \mapsto \omega^\alpha \times (n + 1)+ \beta \quad (n\in\mathbb{N})\)
-\end{itemize}
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#T1}{Epsilon0.T1}}
-
-\label{types:T1}
-%\index{Constants!zero:T1}
-
-\begin{Coqsrc}
-Inductive T1 : Set  :=
-| zero : T1
-| ocons : T1 -> nat -> T1 -> T1.
-\end{Coqsrc}
-
-\subsubsection*{Remark}
-\label{sec:orgheadline66}
-
-The name \texttt{T1} we gave to this data-type  is proper to this development and refers
-to a hierarchy of ordinal notations.
-For instance, in~\cite{CantorContrib}, the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{schutte}).
-
-\emph{Please look also at the library
-\href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2.html}}.
-
-\begin{Coqsrc}
-Inductive T2 : Set :=
-  zero : T2
-| gcons : T2 -> T2 -> nat -> T2 -> T2.
-\end{Coqsrc}
-
-\subsubsection{Example}
-
-\label{alpha0-def}
-For instance, the ordinal  $\omega^\omega+\omega^3\times 5+2$ is represented by the following term:
-
-\begin{Coqsrc}
-Example alpha_0 : T1 :=
-  ocons (ocons (ocons zero 0 zero)
-               0
-               zero)
-        0
-       (ocons (ocons zero 2 zero)
-              4
-              (ocons zero 1 zero)).
-\end{Coqsrc}
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5, level 1/.style={sibling distance=6cm},
-level 2/.style={sibling distance=35mm},  
-level 3/.style={sibling distance=17mm}]
-\node  {ocons}
-  child {  node {ocons}
-            child { node {ocons} child {node {zero}} child {node{0}} child{node{zero}}}
-         child {node {0}}
-         child {node {zero}}}
-    child {node {0}}
-   child {node {ocons} 
- child { node {ocons} child {node {zero}} child {node{2}} child{node{zero}}}
-  child {node {4}}
-         child {node {ocons} child {node {zero}} child {node{1}} child{node{zero}}}};
-
-\end{tikzpicture}
-
-\caption{The tree-like representation of the ordinal $\omega^\omega+\omega^3\times 5 +2$\label{fig:cnf-tree}}
-
-\end{figure}
-
-\paragraph{Remark}
-For simplicity's sake, we chosed to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the contruction (\texttt{ocons $\alpha$ $n$ $\beta$}) is intented to represent the
-ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
-In a future version, we should replace  the type \texttt{nat} with \texttt{positive} in \texttt{T1}'s 
-definition. But this replacement would take a lot of time \dots{}
-
-\subsection{Abbreviations}
-
-Some abbreviations may help to write more concisely complex ordinal terms.
-
-\subsubsection{Finite Ordinals}
-\label{sec:orgheadline67}
-
-For representing finite ordinals, \emph{i.e.} natural numbers, we first introduce a notation for terms of the form $n+1$, then define a coercion from type \texttt{nat} into \texttt{T1}.
-\label{sect:notation-FS}
-
-\begin{Coqsrc}
-Notation "'FS' n" :=
-     (ocons zero n zero) (at level 10) : t1_scope.
-\end{Coqsrc}
-
-\label{sect:notation-F}
-
-\begin{Coqsrc}
-Definition fin (n:nat) : T1 := 
-    match n with 0 => zero | S p => FS p end.
-
-Coercion fin  : nat >-> T1.
-
-Example ten : T1 := 10.   
-\end{Coqsrc}
-
-% \index{Coq!Techniques!Coercions}
-% \index{Functions!Coercions@Coercions (from nat to ordinal types)}
-% \begin{remark}
-% Please refer to the remark~\pageref{warning:coercions} about the use of coercions.
-% % The use of coercions like \texttt{fin} allow us to be close to the mathematical tradition where natural numbers are ordinals too.
-% % Nevertheless, it may happen that a goal like \texttt{3 < 5} could be 
-% % interpreted as \texttt{(lt (fin 3) (fin 5))},  depending on the current notation scope.  
-% % When this misinterpretation happens, tactics like \texttt{auto with arith}, \texttt{lia} do not work!
-% % Thus, it is useful to write \texttt{(3 < 5)\%nat}  an inequality between two natural numbers. 
-% \end{remark}
-
-\subsubsection{The Ordinal \(\omega\)}
-\label{sec:orgheadline68}
-
-  Since \(\omega\)'s Cantor normal form is
-i.e. \(\omega^{\omega^0}\times 1+ 0\), we can define the following abbreviation:
-
-\label{sect:omega-notation2}
-\begin{Coqsrc}
-Notation omega := (ocons (ocons zero 0 zero) 0 zero): t1_scope.
-\end{Coqsrc}
-
-Note that \texttt{omega} is not an identifier, thus any tactic like \texttt{unfold omega} would fail.
-
-\subsubsection{The Ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
-\label{sec:orgheadline70}
-
-We provide also a notation for ordinals of the form $\omega^\alpha$.
-
-\label{sect:notation-phi0}
-
-\index{Notations!phi0@phi0 (exponential of base omega)}
-
-\begin{Coqsrc}
-Notation "'phi0' alpha" := (ocons alpha 0 zero) (at level 29) : t1_scope.
-\end{Coqsrc}
-
-\index{Maths!Additive principal ordinals}
-
-\begin{remark}
-\label{sec:orgheadline69}
-The name \(\phi_0\)
-   comes from ordinal numbers theory. In~\cite{schutte}, Schütte defines 
-$\phi_0$  as the ordering (\emph{i.e.} enumerating) function of the set  of \emph{additive principal ordinals} \emph{i.e.} strictly positive ordinals $\alpha$ that verify $\forall \beta<\alpha, \beta+\alpha=\alpha$.
-For Schütte,  $\omega^\alpha$ is just a notation for $\phi_0(\alpha)$.
-See also Chapter~\ref{chap:schutte} of this document.
-\end{remark}
-
-\subsubsection{The Hierarchy of \(\omega\)-towers:}
-\label{sec:orgheadline71}
-
-The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximed by the sequence of $\omega$-towers (see also Sect~\vref{sect:epsilon0-as-limit} ).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}
-
-\begin{Coqsrc}
-Fixpoint omega_tower (height:nat) : T1 := 
- match height with 
- | 0 =>  1 
- | S h => phi0 (omega_tower h)
- end.
-\end{Coqsrc}
-
-For instance, Figure~\ref{fig:tower7} represents  the ordinal returned by the
- evaluation of the term \texttt{omega\_tower 7}.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
-\node[color=blue]{$\omega^{{{\omega}^{{{\omega}}^{{{\omega}}^{{\omega^{{\omega}^{\omega}}}}}}}}$};
-\end{tikzpicture}
-\caption{\label{fig:tower7}
-The $\omega$-tower of height 7}
-\end{figure}
-
-\subsection{Comparison between Ordinal Terms}
-\label{sec:orgheadline73}
-
-% Our formalisation of Cantor Normal Form will take two steps:
-% 1 Definition of a strict order \texttt{o<} on the type \texttt{T1}, 
-% 2 Using \texttt{o<} for characterizing terms in normal form.
-
-In order to compare two terms of type \texttt{T1}, we define a recursive function \texttt{compare} that maps two ordinals $\alpha$ and $\beta$ to a value of type \texttt{comparison}.
-This type is defined in \coq's standard library
-\texttt{Init.Datatypes} and
-contains three constructors:  \texttt{Lt} (less than), \texttt{Eq} (equal), and
-\texttt{Gt} (greater than).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare}{Epsilon0.T1}}
-
-\begin{Coqsrc}
-Fixpoint compare (alpha alpha':T1):comparison :=
-  match alpha, alpha' with
-    zero, zero => Eq
-  | zero, ocons a' n' b' => Lt
-  | _   , zero => Gt
-  | (ocons a n b),(ocons a' n' b') =>
-      (match compare a a' with 
-          | Lt => Lt
-          | Gt => Gt
-          | Eq => (match lt_eq_lt_dec n n'
-                   with
-                       inleft  (left _) => Lt
-                     | inright _ => Gt
-                     |   _ => compare b b'
-                   end)
-       end)
-  end.
-\end{Coqsrc}
-
-It is now easy to define the boolean predicate \texttt{lt\_b $\alpha$ $\beta$}: 
-`` $\alpha$ is strictly less than $\beta$ ''. By coercion to sort \texttt{Prop} we define also the predicate \texttt{lt}.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}
-
-\begin{Coqsrc}
-Definition lt_b alpha beta : bool :=
-  match compare alpha beta with
-      Lt => true
-    | _ => false
-  end.
-
-Definition lt alpha beta : Prop := lt_b alpha beta.
-\end{Coqsrc}
-
-\label{Predicates:lt-T1}
-Please note that this definition of \texttt{lt} makes it easy to write proofs by reflection, as shown by the following exampgles.
-
-\begin{Coqsrc}
-Example E1 : lt (ocons omega 56 zero) (tower 3).
-Proof. reflexivity. Qed.
-
-Example E2 : ~ lt (tower 3) (tower 3).
-Proof.  discriminate.  Qed.
-\end{Coqsrc}
-
-The following lemmas establish relations between \texttt{compare}, 
-the predicate \texttt{lt} and Leibniz equality \texttt{eq}.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare_refl}{Epsilon0.T1}}
-
-\begin{Coqsrc}
-Lemma compare_refl : forall alpha, compare alpha alpha =  Eq.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma compare_reflect : forall alpha beta,
-    match compare alpha beta with
-    |   Lt => lt alpha  beta
-    |   Eq => alpha = beta
-    |   Gt => lt beta  alpha
-    end.
-\end{Coqsrc}
-
-We prove also that the relation \texttt{lt} is a strict total order.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#lt_irrefl}{Epsilon0.T1}}
-
-\begin{Coqsrc}
-Theorem lt_irrefl (alpha: T1):  ~ lt alpha alpha.
-
-Theorem lt_trans (alpha beta gamma : T1) :
-  lt alpha  beta -> lt beta gamma -> lt alpha gamma.
-
-Definition lt_eq_lt_dec  :
-   forall alpha beta : T1, 
-          {lt alpha  beta} + {alpha = beta} + {lt beta alpha}.
-\end{Coqsrc}
-
-Note that the order \texttt{lt} is not reflected 
-in the structure (size and/or height) of the terms of \texttt{T1}. 
-For instance the ordinal of Fig \ref{fig:cnf-example} is strictly less
-than the structurally simpler \(\omega^{\omega^\omega}\times 2\).
-
-\subsubsection{A Predicate for Characterizing Normal Forms}
-\label{sect:t1-nf}
-
-\label{sec:orgheadline74}
-\label{sec:orgheadline75}
-Our data-type \texttt{T1} allows us to write expressions that
-are not properly in Cantor normal form as specified in Section \ref{sec:epsilon0-intro}.
-For instance, consider the following term of type  \texttt{T1}.
-
-\begin{Coqbad}
-Example bad_term  : T1 := ocons 1 1 (ocons omega 2 zero).
-\end{Coqbad}
-
-This term would have been written \(\omega^1\times 2 + \omega^\omega \times 3\) in the usual mathematical notation.
-We note that the exponents of $\omega$ are not in the right (strictly decreasing) order.
-
-With the help of the order \texttt{lt} on \texttt{T1}, we are now able to characterize
-the set of all well-formed ordinal terms:
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#nf_b}{Epsilon0.T1}}
-
-\label{Predicates:nf-T1}
-
-\begin{Coqsrc}
-Fixpoint nf_b (alpha : T1) : bool :=
-  match alpha with
-    | zero => true
-    | ocons a n zero => nf_b a
-    | ocons a n ((ocons a' n' b') as b) =>
-      (nf_b a && nf_b b && lt_b a' a)%bool
-  end.
-
-Definition nf alpha: Prop := nf_b alpha.
-\end{Coqsrc}
-
-\begin{Coqsrc}
- Compute nf_b alpha_0.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-   = true 
-     : bool
-\end{Coqanswer}
-
-\begin{Coqsrc}
- Compute nf_b bad_term.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-   = false 
-     : bool
-\end{Coqanswer}
-
-\subsection{Making Normality Implicit}
-  We would like to get rid of terms of type \texttt{T1} which are not in Cantor normal form.
-A simple way to do this is to consider statements of the form 
-\texttt{forall alpha: T1, nf alpha -> $P$ alpha}, where $P$ is a predicate over type \texttt{T1}, like in the following lemma \footnote{Ordinal addition is formally defined a little later (page~\ref{sect:infix-plus-T1})}.
-
-\begin{Coqsrc}
-Lemma plus_is_zero alpha beta :
-  nf alpha -> nf beta ->
-  alpha + beta  = zero -> alpha = zero /\  beta = zero.
-\end{Coqsrc}
-
-But this style leads to clumsy statements, and generates too many sub-goals in interactive proofs (although often solved with \texttt{auto} or \texttt{eauto}).
-
-One may encapsulate conditions of the form \texttt{(nf $\alpha$)} in
-the most used predicates. For instance, we introduce the restriction of \texttt{lt} to terms in normal form, and provide a handy notation for this restriction.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Prelude.Restriction.html}{Ordinals.Prelude.Restriction}}
-
-\begin{Coqsrc}
-Definition restrict {A:Type}(E: Ensemble A)(R: relation A) :=
-    fun a b => E a /\ R a b /\ E b.
- \end{Coqsrc}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#LT}{Epsilon0.T1}}
-
-\begin{Coqsrc}
-Definition LT := restrict nf lt.
-Infix "t1<" := LT : t1_scope.
-
-Definition LE := restrict nf le.
-Infix "t1<=" := LE : t1_scope.
-\end{Coqsrc}
-
-\label{Predicates:LT-T1}
-
-For instance, in the following lemma, the condition that $\alpha$ is in normal form is included in the condition $\alpha< 1$.
-
-\begin{Coqsrc}
-Lemma LT_one : forall alpha, alpha t1< one -> alpha = zero.
-\end{Coqsrc}
-
-\subsubsection{A Sigma Type for \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-As we noticed in Sect.~\ref{sect:t1-nf}, the type \texttt{T1} is not a correct ordinal notation, since it contains terms that are not in Cantor normal form.
-In certain contexts (for instance in Sections~\ref{sect:L-equations}, \ref{sect:hardy},
-and \ref{sect:wainer}),  we need to define total recursive functions on well-formed ordinal terms less  than $\epsilon_0$, using the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777}.
- In order to define a type whose inhabitants represent just ordinals, we build a type gathering a term of type \texttt{T1} and a proof that this term is in normal form.
-
-\label{sect:E0-def}
-\label{types:E0}
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
-
-\begin{Coqsrc}
-Class E0 : Type := t1_2o {cnf : T1; cnf_ok : nf cnf}.
-\end{Coqsrc}
-
-Many constructs : types, predicates, functions, notations, etc., on type \texttt{T1} are adapted to \texttt{E0}.
-
-First, we declare a notation scope for \texttt{E0}.
-
-\begin{Coqsrc}
-Declare Scope E0_scope.
-Delimit Scope E0_scope with e0.
-Open Scope E0_scope.
-\end{Coqsrc}
-
-Then we redefine the predicates of comparison.
-
-\label{Predicates:Lt-E0}
-
-\begin{Coqsrc}
-Definition Lt (alpha beta : E0) := T1.LT (@cnf alpha) (@cnf beta).
-Definition Le (alpha beta : E0) := T1.LE (@cnf alpha) (@cnf beta).
-
-Infix "o<" := Lt : E0_scope.
-Infix "o<=" := Le : E0_scope.
-\end{Coqsrc}
-
-Equality in \texttt{E0} is just Leibniz equality. Note that, since \texttt{nf} is
-defined by a Boolean function, for  any term $\alpha:\texttt{T1}$, there exists at most one proof of \texttt{nf $\alpha$}, thus two ordinals of type \texttt{E0} are
-equal if and only iff their projection to \texttt{T1} are equal (see also Sect.~\vref{sect:eq-proof-unicity}).
-
-\index{Coq!Techniques!Unicity of equality proofs}
-
-\begin{Coqsrc}
-Require Import Logic.Eqdep_dec.
-
-Lemma nf_proof_unicity :
-  forall (alpha:T1) (H H': nf alpha), H = H'.
-
-Lemma E0_eq_iff alpha beta : alpha = beta <-> cnf alpha = cnf beta.
-\end{Coqsrc}
-
-\index{Exercises}
-\begin{exercise}
-In earlier versions of this development, the predicate \texttt{nf} was defined  inductively, with various constructors describing all possible cases.
-\begin{enumerate}
-\item Please give such a definition, in a dedicated module.
-\item Prove the logical equivalence between your definition and ours.
-\item Define a variant of the type \texttt{E0} (with your definition of \texttt{nf}).
-\item Can you still prove a lemma like \texttt{E0\_eq\_iff} ? With the help of an axiom from some module of the standard library ?
-\end{enumerate}
-\end{exercise}
-For upgrading constants and fonctions of \texttt{T1}, we have to prove that 
-the term they build is in normal form.
-For instance, let us represent the ordinals $0$ and $\omega$ as instances of the class \texttt{E0}.
-
-\label{sect:omega-T1}
-%\index{Constants!zero:T1}
-
-\begin{Coqsrc}
-Instance Zero : E0.
-Proof.
-  now exists T1.zero.
-Defined.
-
-Instance _Omega : E0.
-Proof.  now exists omega%t1. 
-Defined.
-
-Notation "'omega'"  := _Omega : E0_scope.
-\end{Coqsrc}
-
-\subsection{Syntactic Definition of Limit and Successor Ordinals}
-
-Pattern matching and structural recursion allow us to define the notions of successor and limit ordinal with the help of  boolean functions on type \texttt{T1}.
-
- \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}
-
-\begin{Coqsrc}
-  Fixpoint succb alpha :=
-  match alpha with
-      zero => false
-    | ocons zero _ _ => true
-    | ocons alpha n beta => succb beta
-  end.
-
-Fixpoint limitb alpha :=
-  match alpha with
-      zero => false
-    | ocons zero _ _ => false
-    | ocons alpha n zero => true
-    | ocons alpha n beta => limitb beta
-  end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-  Compute limitb omega.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = true
-     : bool
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Compute succb 42.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = true
-     : bool
-   \end{Coqanswer}
-
-The correctness of these definitions with respect to the mathematical notions of
-limit and successor ordinals is established through several lemmas. For instance,
-Lemma \texttt{canonS\_limit}, page~\pageref{lemma:canonS-limit}, shows that
-if $\alpha$ is (syntactically) a limit ordinal, then it is the least upper bound of
-a strictly increasing sequence of ordinals.
-
-   The following function is very useful in constructions by cases (proofs and function definitions).
-
-\begin{Coqsrc}
-Definition zero_succ_limit (alpha: T1) :
-    {succb alpha} + {limitb alpha} +  {alpha=zero}.
-    (* ... *)
-Defined.
-\end{Coqsrc}
-
-\subsection{Arithmetic on \texorpdfstring{$\epsilon_0$}{epsilon0}}
-\subsubsection{Successor}
-
-The successor of any ordinal $\alpha< \epsilon_0$ is defined by structural 
-recursion on its Cantor normal form.
-
-\label{Functions:succ-T1}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}
-
-\begin{Coqsrc}
-Fixpoint succ (alpha:T1) : T1 :=
-  match alpha with 
-   | zero => 1
-   | ocons zero n _ => ocons zero (S n) zero
-   | ocons beta n gamma => ocons beta n (succ gamma)
- end.
-\end{Coqsrc}
-
-The following lemma establishes the connection between the  function
-\texttt{succ} and the Boolean predicate \texttt{succb}.
-
-\begin{Coqsrc}
- Lemma succb_iff alpha (Halpha : nf alpha) :
-  succb alpha <-> exists beta : T1, nf beta /\ alpha = succ  beta.
-\end{Coqsrc}
-
-\index{Exercises}
- \begin{exercise}
-Prove in \coq{} that for any ordinal $\alpha<\epsilon_0$, $\alpha$ is a limit if 
-and only if for all $\beta<\alpha$, the interval $[\beta,\alpha)$ is infinite.
- \end{exercise}
-
-\subsubsection{Addition and Multiplication}
-
-Ordinal addition and multiplication are also defined by structural recursion over the type \texttt{T1}. Please note that they use the \texttt{compare} function on some subterms of their arguments.
-
-\label{sect:infix-plus-T1}
-
-\begin{Coqsrc}
-Fixpoint plus (alpha beta : T1) : T1 :=
-  match alpha,beta with
- |  zero, y  => y
- |  x, zero  => x
- |  ocons a n b, ocons a' n' b' =>
-    (match compare a a' with
-     | Lt => ocons a' n' b'
-     | Gt => (ocons a n (plus b (ocons a' n' b')))
-     | Eq  => (ocons a (S(n+n')) b')
-     end)
-  end
-where "alpha + beta" := (plus alpha beta) : t1_scope.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Fixpoint mult (alpha beta : T1) :T1 :=
-  match alpha,beta with
- |  zero, y  => zero
- |  x, zero => zero
- |  ocons zero n _, ocons zero n' _ => 
-                 ocons zero (Peano.pred((S n) * (S n'))) zero
- |  ocons a n b, ocons zero n' b' =>  
-                 ocons a (Peano.pred((S n) * (S n'))) b
- |  ocons a n b, ocons a' n' b' =>
-     ocons (a + a') n' ((ocons a n b) * b')
- end
-where  "alpha * beta" := (mult alpha beta) : t1_scope.
-\end{Coqsrc}
-
-\subsubsection{Examples}
-
-The following examples are instances of \emph{proofs by computation}. Please note that  addition and multiplication on \texttt{T1}
-are not commutative. Moreover,  both operations fail to be strictly monotonous in their first argument.
-
-\begin{Coqsrc}
-Example e2 : 6 + omega = omega.
-Proof. reflexivity. Qed.
-
-Example e'2 : omega t1< omega + 6.
-Proof. now compute. Qed.
-
-Example e''2 : 6 * omega = omega.
-Proof. reflexivity. Qed.
-
-Example e'''2 : omega t1< omega * 6.
-Proof. now compute. Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma plus_not_monotonous_l : exists alpha beta gamma : T1,
-    alpha t1< beta /\ alpha + gamma = beta + gamma.
-Proof.
-  exists 3, 5, omega;  now  compute.
-Qed.
-
-Lemma mult_not_monotonous :  exists alpha beta gamma : T1,
-      alpha t1< beta /\ alpha * gamma = beta * gamma.
-Proof.
-  exists 3, 5, omega; now compute.
-Qed.
-\end{Coqsrc}
-
-\subsection{Pretty Printing Ordinals in Cantor Normal Form}
-\label{sect:ppT1}
-
-Let us consider again the ordinal $\alpha_0$ defined in section~\vref{alpha0-def}
-If we ask \coq{} to print its  normal form, we get a hardly readable term of type \texttt{T1}.
-
-\begin{Coqsrc}
-Compute alpha_0.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = ocons omega 0 (ocons (FS 2) 4 (FS 1))
-     : T1
-\end{Coqanswer}
-
-The following data type defines an abstract syntax for more readable ordinals terms in Cantor normal form:
-
-\label{types:ppT1}
-\index{Functions!pp@ pp (pretty printing terms in Cantor normal form)}
-
-\begin{Coqsrc}
-Inductive ppT1 : Set :=
-    P_fin : nat -> ppT1
-  | P_add : ppT1 -> ppT1 -> ppT1
-  | P_mult : ppT1 -> nat -> ppT1
-  | P_exp : ppT1 -> ppT1 -> ppT1
-  | P_omega : ppT1
-\end{Coqsrc}
-
-The function \texttt{pp: T1 -> ppT1} converts any closed term of type \texttt{T1} into a human-readable expression. For instance, let us convert the term \texttt{alpha\_0}.
-
-\begin{Coqsrc}
-Compute pp alpha_0.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-     = (omega ^ omega + omega ^ 3 * 5 + 2)%pT1
-     : ppT1
-\end{Coqanswer}
-
-\index{Projects}
-\begin{project}
-Design  (in \ocaml?) a set of tools for systematically pretty printing ordinal terms in Cantor normal form.
-\end{project}
-
-\subsubsection{Arithmetic on Type \texttt{E0}}
-
- We define an addition in type \texttt{E0}, using the function{T1.plus}
-and its properties.
-
-\begin{Coqsrc}
-Search   (nf (_ + _)%t1).
-\end{Coqsrc}
-
-\begin{Coqanswer}
-plus_nf: forall a : T1, nf a -> forall b : T1, nf b -> nf (a + b)
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Instance plus (alpha beta : E0) : E0.
-Proof.
-  refine (@mkord (T1.plus (@cnf alpha) (@cnf beta))_ );
-    apply plus_nf; apply cnf_ok.
-Defined.
-
-Infix "+" := plus : E0_scope.
-
-Check omega + omega.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-omega + omega
-     : E0
-\end{Coqanswer}
-
-\index{Exercises}
- \begin{exercise}
-Let $\alpha$ be an ordinal. We say that $\alpha$ is \emph{infinite} iff the segment $[0,\alpha)$ is an infinite set. 
-\begin{enumerate}
-\item Adapt this definition to the type \texttt{E0}.
-\item Prove that \emph{being infinite} is decidable on type \texttt{E0}.
-\item Prove that $\alpha$ is infinite if and only if for all finite ordinal $i$,
-$i+\alpha=\alpha$.
-\end{enumerate}
- \end{exercise}
-
-\section{Well-foundedness and Transfinite Induction}
-
-\index{Maths!Transfinite induction}
-
-\subsection{About  Well-foundedness}
-\label{sec:orgheadline82}
-   In order to use \texttt{T1} for proving termination results,
-we need to prove that  our order \texttt{<} is well-founded. Then we will get \emph{transfinite induction} for free.
-
-The proof of well-foundedness of the strict order $<$ on Cantor normal forms is already 
-available in the Cantor contribution by Castéran and Contejean~\cite{CantorContrib}. That proof relies on a library on recursive path orderings written by
-E. Contejean. We present here  a direct proof of the same result, which does not require any knowledge on r.p.o.s.
-
-\index{Exercises}
-
-\begin{exercise}
-Prove that the \emph{total} order \texttt{lt} on \texttt{T1} is not well-founded. 
-\textbf{Hint:}  You will have to build a counter-example with terms of type \texttt{T1}
-which are not in Cantor normal form.
-\end{exercise}
-
-% \subsubsection{The total order \texttt{lt} on \texttt{T1} is \emph{not} well-founded}
-% \label{sec:orgheadline76}
-
-% Let us recall that the data type \texttt{T1} contains too many inhabitants, including
-% terms which are not in Cantor normal form. Thus, the following result is not 
-% very surprising.
-
-% \begin{Coqsrc}
-% Section lt_not_well_founded.
-
-%   (* let us build the sequence of terms :
-%         omega + omega + .... + omega ^ 2   *)
-%   Let f := (fix f (i:nat): T1 :=
-%             match i with 0 => phi0 2
-%                        | S i => ocons 1 1 (f  i)
-%             end).
-
-%  Lemma  f_decreases : forall i, f (S i) <  f i.
-%  Proof.
-%   induction i; compute; auto with T1.
-%  Qed.
-
-%  Theorem lt_not_wf : ~  well_founded lt.
-%  Proof. 
-%    intro wf; case (not_decreasing _ lt);auto.
-%    exists f; apply f_decreases.
-%  Qed.
-
-% End lt_not_well_founded.
-% \end{Coqsrc}
-
-% Thus, we have
-
-\subsubsection{A First Attempt}
-\label{sec:orgheadline77}
-\index{Coq!Techniques!Well-founded induction}
-
-It is natural to try to prove by structural induction over \texttt{T1} 
-that every term in normal form is accessible through \texttt{LT}.
-
-Unfortunately, it won't work. Let us consider some well-formed term
- $\alpha=\texttt{ocons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are accessible
- through \texttt{LT}. For proving the accessibility of $\alpha$, we have to consider
-any well formed term \(\delta\) such that \(\delta<\alpha\). 
-But nothing guarantees that \(\delta\)  is strictly  less than \(\beta\) nor \(\gamma\), and we cannot use the induction hypotheses on   \(\beta\) nor \(\gamma\).
-
-\begin{Coqbad}
-Section First_attempt.
-
- Lemma wf_LT : forall alpha,  nf alpha -> Acc LT alpha. 
- Proof.
-  induction alpha as [| beta IHbeta n gamma IHgamma].
-  - split.
-    inversion 1.
-    destruct H2 as [H3 _];not_neg H3.
-  -  split; intros delta Hdelta.
-\end{Coqbad}
-
-\begin{Coqanswer}
-1 subgoal (ID 560)
-
-  beta : T1
-  n : nat
-  gamma : T1
-  IHbeta : nf beta -> Acc LT beta
-  IHgamma : nf gamma -> Acc LT gamma
-  H : nf (ocons beta n gamma)
-  delta : T1
-  Hdelta : delta t1< ocons beta n gamma
-  ============================
-  Acc LT delta
- \end{Coqanswer}
-
-\begin{Coqbad}
-  Abort.
-\end{Coqbad}
-
-The problem comes from the hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
-\(\gamma\);
-for instance \(\delta\) may be of the form
-\texttt{ocons $\beta'$ $p'$  $\gamma'$},
-where  \(\beta' \leq  \beta\) and  \(p' < n\).
-Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.
-
-\subsubsection{Using a Stronger Inductive Predicate.}
-\label{sec:orgheadline78}
-  Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
-through \texttt{LT}, we propose to show first that any well formed 
-term of the form \(\omega^\alpha\times(n+1)+\beta\) is accessible (which is a stronger result).
-
-\begin{Coqsrc}
- Let Acc_strong (alpha:T1) :=
-      forall n beta, 
-        nf (ocons alpha n beta) -> Acc LT (ocons alpha  n beta).
-\end{Coqsrc}
-
-The following lemma is an application of the strict inequality 
-\showmath{\alpha < \omega ^\alpha}. If \showmath{\omega^\alpha} is accessible,
-then \showmath{\alpha} is \emph{a fortiori} accessible.
-
-\begin{Coqsrc}
- Lemma Acc_strong_stronger : forall alpha, 
-     nf alpha -> Acc_strong  alpha -> Acc LT  alpha.
- Proof.
-  intros alpha H H0; apply acc_imp with (phi0 alpha).
-  - repeat split; trivial.
-    + now apply lt_a_phi0_a.
-  -  apply H0;  now apply single_nf.
-Qed.
-\end{Coqsrc}
-
-Thus, it remains to prove that every ordinal strictly less than \showmath{\epsilon_0} 
-is strongly accessible.
-
-% \subsubsection{Structure of the proof of well-foundedness of \texttt{LT}}
-
-\label{sec:orgheadline81}
-\label{proof-wf-epsilon0}
-\paragraph{A helper}
-\label{sec:orgheadline79}
-
-First, we prove that, for  any \texttt{LT}-accessible term \showmath{\alpha}, \showmath{\alpha} is 
-strongly accessible too (\emph{i.e.} any well formed
-term (\texttt{ocons $\alpha$ $n$ $\beta$})  is accessible).
-
-\begin{Coqsrc}
-Lemma Acc_implies_Acc_strong : 
-   forall alpha, Acc LT  alpha -> Acc_strong alpha.
-\end{Coqsrc}
-
-The proof is structured as an induction on \showmath{\alpha}'s accessibility. Let us consider
-an accessible term $\alpha$.
-
-\begin{Coqanswer}
-  subgoal 1
-
-  alpha : T1
-  Aalpha : forall y : T1,  y t1< alpha -> Acc LT y
-  IHalpha : forall y : T1,
-       LT y alpha ->
-       forall (n : nat) (beta : T1),
-       nf (ocons y n beta) -> Acc LT (ocons y n beta)
-  ============================
-   forall (n : nat) (beta : T1),
-   nf (ocons alpha n beta) -> Acc LT (ocons alpha n beta)
-\end{Coqanswer}
-
-Let \texttt{n:nat} and \texttt{beta:T1} such that \texttt{ocons alpha n beta} is in normal form. 
-We prove first that \texttt{beta} is accessible,  which allows us to by well-founded induction on \texttt{beta}, 
-and natural induction on \texttt{n}, that (\texttt{ocons alpha n beta}) is accessible.
-The proof, quite long, can be consulted in \url{../theories/html/hydras.Epsilon0.T1.html}
-
-\paragraph{Accessibility of any well-formed ordinal term}
-\label{sec:orgheadline80}
-
-Our goal is still to prove accessibility of any well formed ordinal term.
-Thanks to our previous lemmas, we are almost done.
-
-\begin{Coqsrc}
-(* A (last) structural induction *)
-
-Theorem nf_Acc : forall alpha, nf alpha -> Acc LT  alpha.
-Proof.
- induction alpha.
--  intro; apply Acc_zero.
- -  intros; eapply Acc_implies_Acc_strong;auto.
-    apply IHalpha1;eauto.
-    apply nf_inv1 in H; auto. 
-Defined.
-
-Corollary T1_wf : well_founded LT.
-\end{Coqsrc}
-
-\index{Maths!Transfinite induction}
-
-\begin{Coqsrc}
-
-Definition transfinite_recursor :
- forall (P:T1 -> Type),
-   (forall x:T1, 
-     (forall y:T1, nf x -> nf y ->  lt y  x -> P y) -> P x) ->
-    forall alpha:T1, P alpha.
-Proof.
- intros; apply well_founded_induction_type with LT.
- -  exact T1_wf;auto.
- - intros. apply X. intros; apply X0. repeat split;auto. 
-Defined.
-\end{Coqsrc}
-
-The following tactic starts a proof by  transfinite induction on any ordinal \mathcolor{$\alpha<\epsilon_0$}.
-
-\begin{Coqsrc}
-Ltac transfinite_induction alpha :=
-  pattern alpha; apply transfinite_recursor;[ | try assumption].
-\end{Coqsrc}
-
-\begin{remark}
-\label{remark:a3pat}
-The alternate proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
- \end{remark}
-
-\subsection{An Ordinal Notation for  \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-We build an instance of \texttt{ON}, and prove its correction w.r.t. Schutte's model.
-
-\label{instance-epsilon0}
-\begin{Coqsrc}
-Instance Epsilon0 : ON Lt compare.  
-(* ... *)
-\end{Coqsrc}
-
-\emph{From Module~\href{../theories/html/hydras.Schutte.Schutte.Correctness_E0.html}{Schutte.Schutte.Correctness\_E0}}
-
-\begin{Coqsrc}
-Instance Epsilon0_correct :
-  ON_correct epsilon0 Epsilon0 (fun alpha => inject (cnf alpha)).
-\end{Coqsrc}
-
-\index{Projects}
-\begin{project}
- \emph{This exercise is a continuation of Project~\vref{exo:ON-mult}.}
-Use \texttt{ON\_mult} to define an ordinal notation \texttt{Omega2} for $\omega^2=\omega\times\omega$.
-
-Prove that \texttt{Omega2} is a sub-notation of \texttt{Epsilon0}.
-
-Define on \texttt{Omega2} an addition compatible with the addition on \texttt{Epsilon0}.
-
-\textbf{Hint}. You may use the following definition (in 
-\href{../theories/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}).
-
-\begin{Coqsrc}
-Definition SubON_same_op  `{OA : @ON A ltA  compareA}
-       `{OB : @ON B ltB  compareB}
-       {iota : A -> B} 
-       {alpha: B}
-       {_ : SubON OA OB alpha iota}
-       (f : A -> A -> A)
-       (g : B -> B -> B)
-  :=
-  forall x y,  iota (f x y) = g (iota x) (iota y).
-\end{Coqsrc}
-
-\end{project}
-
-\index{Projects}
-\begin{project}
-The class \texttt{ON} of ordinal notations has been defined long after this 
-chapter, and is not used yet in the development of the type \texttt{E0}. 
-A better integration of both notions should simplify the development on ordinals in Cantor normal form. This integration is planned for the future versions.
-
-\end{project}
-
-\section{A Variant for Hydra Battles}
-
-\subsection{Natural Sum (a.k.a. Hessenberg's  Sum)}
-\label{sec:orgheadline87}
-\label{hydra-variant}
-
-Natural sum (Hessenberg's  sum) is a commutative and monotonous version of
-addition. It is used as an auxiliary operation  for defining variants
-for hydra battles, where Hercules is allowed to chop off any  head of the hydra.
-
-In the litterature, the natural sum of ordinals \(\alpha\) and \(\beta\) 
-is often denoted by \(\alpha \# \beta\)  or  \(\alpha \oplus  \beta\).
-Thus we called \texttt{oplus} the associated \emph{Coq} function.
-
-\subsubsection{Definition of \texttt{oplus}}
-\label{sec:orgheadline84}
-%\index{Functions!oplus @ oplus (Hessenberg commutative sum)}
-
-The definition of \texttt{oplus} is recursive in both of its 
-arguments, which makes a structural recursive definition a little 
-complex.
-We used the same pattern as for the \texttt{merge} function on lists of library
-\texttt{Coq.Sorting.Mergesort}.
-
-\begin{enumerate}
-\item Define a nested recursive function, using the \texttt{Fix} 
-    construct
-
-\item Build a principle of induction dedicated to \texttt{oplus}
-
-\item Establish equations associated to each case of the definition.
-\end{enumerate}
-
-\paragraph{Nested recursive definition}
-\label{sec:orgheadline83}
-
-The following definition is composed of 
-\begin{itemize}
-\item A main function \texttt{oplus}, structurally recursive in its 
-first argument \texttt{alpha}
-\item An auxiliary function \texttt{oplus\_aux} within the scope of \texttt{alpha},
-structurally recursive in its argument \texttt{beta};  \texttt{oplus\_aux beta} 
-   is supposed to compute  \texttt{oplus alpha beta}.
-\end{itemize}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}
-
-\label{sect:infix-oplus}
-
-\begin{Coqsrc}
-Fixpoint oplus (alpha beta : T1) : T1 :=
-  let fix oplus_aux beta {struct beta} :=
-      match alpha, beta with
-        | zero, _ => beta
-        | _,  zero => alpha
-        | ocons a1 n1 b1, ocons a2 n2 b2 =>
-          match compare a1 a2 with
-            |  Gt => ocons a1 n1 (oplus b1 beta)
-            |  Lt => ocons a2 n2 (oplus_aux b2)
-            |  Eq => ocons a1 (S (n1 + n2)%nat) (oplus b1 b2)
-          end
-      end
-  in oplus_aux beta.
-
-Infix "o+" := oplus  (at level 50, left associativity).
-\end{Coqsrc}
-
-The reader will note that each recursive call of the functions
-\texttt{oplus} and \texttt{oplus\_aux} satisfies \emph{Coq}'s constraint
-on recursive definitions. The function \texttt{oplus} is recursively called on a sub-term of its first argument,
-and \texttt{oplus\_aux} on a sub-term of its unique argument.
-Thus, \texttt{oplus}'s definition is accepted by \coq{} as a structurally recursive function.
-
-\subsubsection{Rewriting Lemmas}
-\label{sec:orgheadline86}
-
-\emph{Coq}'s constraints on recursive definitions result in 
-the quite  complex form of \texttt{oplus}'s definition.
-Proofs of properties of this function can be simpler if we
- \emph{derive} rewriting lemmas that will help to simplify 
-expressions of the form (\texttt{oplus $a$ $ b$}).
-
-A first set of lemmas correspond to the various cases of \texttt{oplus}'s 
-definition. They can be proved almost immediately, using \texttt{cbn} 
-and \texttt{reflexivity} tactics.
-
-\begin{Coqsrc}
-Lemma oplus_alpha_0 (alpha : T1) : alpha o+ zero = alpha.
-Proof.
-  destruct alpha; reflexivity.
-Qed.
-
-Lemma oplus_0_beta (beta : T1): zero o+ beta = beta.
-Proof.
-  destruct beta; reflexivity.
-Qed.
-\end{Coqsrc}
-
-% \subsubsection{A hand-made induction principle}
-% \label{sec:orgheadline85}
-
-% \index{Coq!Commands!Functional Scheme}
-
-% \emph{Coq} contains a command  \texttt{Functional Scheme} that 
-% generates induction principles which correspond to recursive functions.
-% Unfortunately, the current version ( \texttt{8.11.0} ) doesn't work on \texttt{oplus},
-% probably because of the inner \texttt{Fix}.
-
-% \begin{Coqsrc}
-% Functional Scheme oplus_ind := Induction for oplus Sort Prop.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% Error: Anomaly "todo." Please report at http://coq.inria.fr/bugs/.
-% \end{Coqanswer}
-
-% Fortunately, it's a good exercise for a semi-experienced user, to write
-% her/him-self induction principles similar to the ones returned by
-% \texttt{Functional Scheme}.
-
-% \begin{itemize}
-% \item First, we choose to write a version for sort \texttt{Type}, since versions
-% for sorts \texttt{Prop} and \texttt{Set} can be easily derived from
-% the former one. According to \emph{Coq}'s naming politics, we will call our 
-% principle \texttt{oplus\_rect}
-
-% \item The conclusion of \texttt{oplus\_rect} will be (\texttt{$P$ a b (oplus a b)}),
-% where $P$ is an arbitrary function of type 
-% \texttt{T1 -> T1 -> T1 -> Type}
-
-% \item The premises of \texttt{oplus\_rect} will describe how to build an induction 
-% on the graph of \texttt{oplus}.
-% \end{itemize}
-
-% We are now ready to state and prove \texttt{oplus\_rect}, and the reader
-% will note that the statement is longer than the proof script itself,
-% which is a standard proof by induction, simplification and case-analysis 
-% that follows  \texttt{oplus}'s definition.
-
-% We associate also a tactic to the application of \texttt{oplus\_rect}.
-
-% \begin{Coqsrc}
-%  Lemma oplus_rect:
-%       forall P: T1 -> T1 -> T1 -> Type, 
-%         (forall a:T1, P zero a a) ->
-%         (forall a: T1, P a zero a) ->
-%         (forall a1 n1 b1 a2 n2 b2 o,
-%            compare a1 a2 = Gt ->
-%            P b1 (ocons a2 n2 b2) o ->
-%            P (ocons a1 n1 b1) (ocons a2 n2 b2)
-%              (ocons a1 n1 o)) ->
-%         (forall a1 n1 b1 a2 n2 b2 o,
-%            compare a1 a2 = Lt ->
-%            P (ocons a1 n1 b1) b2 o ->
-%            P (ocons a1 n1 b1) (ocons a2 n2 b2) 
-%            (ocons a2 n2 o)) ->
-%         (forall a1 n1 b1 a2 n2 b2 o,
-%            compare a1 a2 = Eq ->
-%            P b1 b2 o ->
-%           P (ocons a1 n1 b1) (ocons a2 n2 b2)
-%             (ocons a1 (S (n1 + n2)%nat) o)) ->
-%          forall a b, P a b (oplus a b).
-% Proof with auto.
-%    induction a.
-%    -    intro; simpl; destruct b;auto.
-%    -   induction b.
-%        + apply X0.
-%        + case_eq (compare a1 b1).
-%          * intro Comp; unfold oplus; rewrite Comp.
-%            cbn; apply X3 ...
-%          * intro Comp; cbn; rewrite Comp; apply X2...
-%          * intro Comp; cbn; rewrite Comp ...
-%  Defined.
-
-% Ltac oplus_induction a b:= pattern (oplus a b); apply oplus_rect.
-% \end{Coqsrc}
-
-% \index{Exercises}
-
-% \begin{exercise}
-% The induction principle \texttt{oplus\_rect} is still unused in our development. 
-% Please build some nice examples of application.
-% \end{exercise}
-
-\index{Projects}
-\begin{project}
-Compare \texttt{oplus}'s definition (with inner fixpoint) with other possibilities
-(\texttt{coq-equations}, \texttt{Function}, etc.).
-\end{project}
-\subsection{More Theorems on Hessenberg's Sum}
-
-We need to prove some properties of $\oplus$, particularly about 
-its relation with the order $<$ on \texttt{T1}.
-
-\subsubsection{Boundedness}
-If $\alpha$ and $\beta$ are both strictly  less than  $\omega^\gamma$, then so is their natural sum
-$\alpha \oplus \beta$. This result can be proved by structural induction on $\gamma$.
-
-\begin{Coqsrc}
-Lemma lt_phi0_oplus : forall gamma alpha beta,
-                        lt_phi0 alpha gamma ->
-                        lt_phi0 beta gamma ->
-                        lt_phi0 (alpha o+ beta) gamma.
-
-Proof with auto.
-  induction gamma; destruct alpha, beta.  
-(* ... *)
-\end{Coqsrc}
-
-\subsubsection{Commutativity, Associativity}
-
-We prove  the commutativity of $\oplus$ in two steps.
-
-First, we prove by transfinite induction on $\alpha$ that the restriction of $\oplus$ to the
-interval $[0..\alpha)$ is commutative.
-
-\index{Maths!Transfinite induction}
-
-\begin{Coqsrc}
-Lemma oplus_comm_0 : forall alpha, nf alpha ->
-     forall a b,  nf a -> nf b ->
-                  lt a alpha ->
-                  lt b alpha ->
-                  a o+ b = b o+ a.
- Proof with eauto with T1.
-    intros alpha Halph; transfinite_induction alpha.
-(* rest of proof omitted *)  
-\end{Coqsrc}
-
-Then, we infer  $\oplus$'s commutativity for any pair of ordinals:
-Let $\alpha$ and $\beta$ be two ordinals strictly less than $\epsilon_0$. Both ordinals $\alpha$ and $\beta$ are
-strictly less than $\textrm{max}(\alpha,\beta)+1$.
-    Thus, we have just to apply the lemma \coqsimple{oplus\_comm\_0}.
-
-\begin{Coqsrc}
-  Lemma oplus_comm : forall alpha beta, 
-      nf alpha -> nf beta ->
-      alpha o+ beta =  beta o+ alpha.
-  Proof with eauto with T1.
-    intros alpha beta Halpha Hbeta;
-    apply oplus_comm_0 with (succ (max alpha beta)) ...  
-  (* ... *)
-\end{Coqsrc}
-
-The associativity of Hessenberg's sum is proved the same way.
-
-\begin{Coqsrc}
- Lemma oplus_assoc_0 :
-    forall alpha,
-      nf alpha ->
-      forall a b c,  nf a -> nf b -> nf c ->
-                      lt a alpha ->
-                      lt b alpha -> lt c alpha ->
-                      a o+ (b o+ c) = (a o+ b) o+ c.
-  Proof with eauto with T1.
-    intros alpha Halpha.
-    transfinite_induction alpha.
-    (* ... *)
-\end{Coqsrc}
-
-\begin{Coqsrc}
- Lemma oplus_assoc : forall alpha beta gamma,
-                        nf alpha -> nf beta -> nf gamma ->
-                                    alpha o+ (beta o+ gamma) =
-                                    alpha o+ beta o+ gamma.
- Proof with eauto with T1.
-    intros;
-    apply oplus_assoc_0 with (succ (max alpha (max beta gamma))) ...
-    (* ... *)   
-\end{Coqsrc}
-
-\subsubsection{Monotonicity}
-
-At last, we prove that $\oplus$ is strictly monotonous in both of its arguments.
-
-\begin{Coqsrc}
-Lemma oplus_strict_mono_LT_l (alpha beta gamma : T1) :
-  nf gamma   -> alpha  t1< beta ->
-  alpha o+ gamma  t1< beta o+ gamma.
-
-Lemma oplus_strict_mono_LT_r (alpha beta gamma : T1) :
-  nf alpha -> beta t1< gamma ->
-  alpha o+ beta t1< alpha o+ gamma.
-\end{Coqsrc}
-
-\index{Projects}
-
-\begin{project}
-The library \texttt{Hessenberg} looks too long (proof scripts and compilation).
-Please try to make it simpler and more efficient!
-Thanks!
-\end{project}
-
-\subsection{A Measure for Hydra-battle Termination}
-
-\label{sec:hydra-measure}
-
-Let us define a measure from type \texttt{Hydra} into \texttt{T1}.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Termination.html\#m}{Hydra.Hydra\_Termination}}
-
-\begin{Coqsrc}
-Fixpoint m (h:Hydra) : T1 :=
-  match h with head => zero
-             | node hs => ms hs
-end 
-with ms (s:Hydrae) :  T1 :=
-  match s with  hnil => zero
-              | hcons h s' => phi0 (m h) o+  ms s'
- end.  
-\end{Coqsrc}
-
-First, we prove that the measure $m(h)$  of any hydra $h$ is a well-formed ordinal term of type \texttt{T1}.
-
-\begin{Coqsrc}
-Lemma m_nf : forall h, nf (m h).
-Proof.
- intro h; elim h using Hydra_rect2 
-            with (P0 := fun s =>  nf (ms s)).
- (* ... *)
-
-Lemma ms_nf : forall s, nf (ms s).
-Proof with auto with T1.
-(* ... *)
-\end{Coqsrc}
-
-For proving the termination of all hydra battles, we have to prove that
-\texttt{m} is a variant. First, a few technical lemmas follow the decomposition of \texttt{round} into several relations. Then the lemma \texttt{round\_decr} gathers all the cases.
-
-\begin{Coqsrc}
-Lemma S0_decr :
-  forall s s', S0  s s' -> ms s' t1< ms s.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma R1_decr : forall h h',
-                  R1 h h' -> m h' t1< m h.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma S1_decr n:
-  forall s s', S1 n s s' -> ms s' t1<  ms s.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma R2_decr n : forall h h', R2 n h h' -> m h'  t1< m h.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma round_decr : forall h h', h -1-> h' -> m h' t1< m h.
-Proof.
-   destruct 1 as [n [H | H]].
-   -  now apply R1_decr.
-   -  now apply R2_decr with n.
-Qed.
-\end{Coqsrc}
-
-Finally, we prove the termination of all (free) battles.
-
-\label{thm:every-battle-terminates}
-
-\begin{Coqsrc}
-Global Instance HVariant : Hvariant lt_wf free var.
-Proof.
- split; intros; eapply round_decr; eauto.
-Qed.
-
-Theorem every_battle_terminates: Termination.
-Proof. 
-  red; apply Inclusion.wf_incl with 
-         (R2 := fun h h' =>  m h t1< m h').
-   red; intros;  now apply round_decr.
-   apply Inverse_Image.wf_inverse_image, T1_wf.
-Qed.
-\end{Coqsrc}
-
-\section*{Conclusion}
-
-Let us recall three results we have proved so far.
-\begin{itemize}
-\item There exists a strictly decreasing variant which maps \texttt{Hydra} into 
-the segment $[0,\epsilon_0)$ for proving the termination of any hydra battle
-\item There exists \emph{no} such variant from \texttt{Hydra} into 
-$[0,\omega^2)$, \emph{a fortiori} into $[0,\omega)$.
-\end{itemize}
-
-So, a  natural question is `` Does there exist any strictly decreasing variant mapping
-type \texttt{Hydra} into some interval $[0,\alpha[$ (where $\alpha <\epsilon_0$) for proving the termination of all hydra battles''.
-The next chapter is dedicated to a formal proof that there exists no such $\alpha$, even if we consider a restriction to the set of ``standard'' battles.
-
-%\include{epsilon0}
-
-%\include{impossibility-proofs}
-
-%-------------------------------------------------------------------
-
-\chapter[The Ketonen-Solovay machinery]{Strolling inside \texorpdfstring{$\epsilon_0$}{Epsilon0}: The Ketonen-Solovay Machinery\label{ks-chapter}}
-\label{chap:ketonen}
-\index{Maths!Ketonen-Solovay machinery}
-
-\section{Introduction}
-The reader may think that our proof of termination in the previous  chapter requires a lot of mathematical tools and may be too  complex. So, the question is ``is there  any  simpler proof'' ?
-
-In their article~\cite{KP82}, Kirby and Paris show that this result cannot be proved in Peano arithmetic.
-Their proof uses some knowledge about model theory and non-standard models of Peano arithmetic.
-In this chapter, we focus on a specific class of proofs of termination of hydra battles: construction of some variant mapping the type \texttt{Hydra} into a given initial  segment of ordinals.
-Our proof relies only on the Calculus of Inductive Constructions and is a natural complement of the results proven in the previous chapter.
-
-\begin{itemize}
-\item There is no variant mapping the type \texttt{Hydra} into the interval $[0,\omega^2)$ (section ~\vref{omega2-case}), and a fortiori 
-$[0,\omega)$ (section ~\vref{omega-case}).
-
-\item There exists a variant which maps the type \texttt{Hydra} into the
-interval $[0,\epsilon_0)$ (theorem \texttt{every\_battle\_terminates}, in section~\vref{thm:every-battle-terminates}).
-\end{itemize}
-
-Thus, a very natural question is the following one:
-\begin{quote}
-  `` Is there  any variant from
-\texttt{Hydra} into some interval $[0,\mu)$, where $\mu<\epsilon_0$, for proving the termination of all hydra battles ?''
-\end{quote}
-
-We prove in \coq{} the following result:
-
-\begin{quote}
-There is no variant for proving the termination of all hydra battles
-from \texttt{Hydra} into the interval $[0..\mu)$, where
-$\mu< \epsilon_0$.
-The same impossibility holds even if we consider only standard battles (with the successive replication factors $0,1,2,\dots,t,t+1,\dots$).
-\end{quote}
-
-Our proofs are  constructive and require no axioms: they are  closed terms of the CIC, and are mainly composed on function definitions and proofs of properties of these functions. 
-They  share much theoretical material with Kirby and Paris', although they do not use any knowledge about Peano arithmetic nor model  theory.  The combinatorial arguments we use and implement
-come from 
- an article by J.~Ketonen and R.~Solovay~\cite{KS81}, already  cited in the work
- by L.~Kirby et J.~Paris.% on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
- Section $2$ of this article: ''A hierarchy of probably recursive functions'', contains a systematic study of \emph{canonical sequences}, which are closely related to
-rounds of hydra battles. 
-Nevertheless, they have the same global structure as the simple proofs described in
-sections~\vref{omega-case} and \vref{omega2-case}. 
-We invite the reader to compare the three proofs step by step, lemma by lemma.
-
-\section{Canonical Sequences}
-\label{ketonen-solovay-sect}
-\index{Maths!Canonical sequences}
-
-Canonical sequences are functions that associate an ordinal $\canonseq{\alpha}{i}$ to every ordinal $\alpha<\epsilon_0$ and positive integer $i$. They satisfy several nice properties:
-
-\index{Maths!Transfinite induction}
-\begin{itemize}
-\item If $\alpha\not=0$, then $\canonseq{\alpha}{i}<\alpha$. Thus canonical sequences can be used for proofs by transfinite induction or function definition by transfinite recursion
-\item If $\lambda$ is a limit ordinal, then $\lambda$ is the least upper bound of the set 
-$\{\canonseq{\lambda}{i}\;|\,i\in\mathbb{N}_1\}$
-
-\item If $\beta<\alpha<\epsilon_0$, then there is a ``path'' from $\alpha$ to $\beta$, \emph{i.e.} a
-sequence $\alpha_0=\alpha, \alpha_1, \dots, \alpha_n=\beta$, where for every $k<n$, there exists some $i_k$ such that $\alpha_{k+1}=\canonseq{\alpha_k}{i_k}$
-\item Canonical sequences correspond tightly to rounds of hydra battles: if $\alpha\not=0$,
-then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1})$ in one round with
-the replication factor $i$ (Lemma \href{../theories/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
-\item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) battle from $\iota(\alpha)$ to $\iota(\beta)$.
-\end{itemize}
-
-\begin{remark}
-  In~\cite{KS81}, canonical sequences are defined for any ordinal $\alpha <\epsilon_0$,
-by stating that if $\alpha$ is a successor ordinal $\beta+1$,  the sequence associated with 
-$\alpha$ is simply the constant sequence whose terms are equal to $\beta$.
-Likewise, the canonical sequence of $0$ maps any natural number to $0$.
-
-This convention allows us to make total the function that maps any ordinal $\alpha$ and natural number $i$ to the ordinal $\canonseq{\alpha}{i}$.
-
-\end{remark}
-
-First, let us recall how canonical sequences are defined in~\cite{KS81}.
-For efficiency's sake, we decided not to implement directly K.\&S's definitions, but to define in \gallina{} simply typed structurally recursive functions which share the abstract properties which are used in the mathematical proofs\footnote{With a small difference: the $0$-th term of the canonical sequence is not the same in our development as in~\cite{KS81}.}.
-
-\subsubsection{Mathematical Definition of Canonical Sequences}
-
-In~\cite{KS81} the definition of $\canonseq{\alpha}{i}$ is based on the following remark:
-\begin{quote}
-Any non-zero ordinal $\alpha$ can be decomposed in a unique way as the product
-$\omega^\beta\times (\gamma+1)$.
-\end{quote}
-
-Thus the $\canonseq{\alpha}{i}$\,s are defined in terms of this decomposition:
-\begin{definition}[Canonical sequences: mathematical definition]
-\label{def:canonseq-math}
-
-\end{definition}
-\begin{mathframe}
-  \begin{itemize}
-\item Let $\lambda<\epsilon_0$ be a limit ordinal
-
-\begin{itemize}
-\item If $\lambda=\omega^{\alpha+1}\times (\beta+1)$, then 
-$\canonseq{\lambda}{i}= \omega^{\alpha+1}\times\beta +  \omega^\alpha \times i$
-\item If $\lambda=\omega^{\gamma}\times (\beta+1)$, where $\gamma<\lambda$ is a limit ordinal, then 
-$\canonseq{\lambda}{i}=\omega^{\gamma}\times \beta + \omega^{\canonseq{\gamma}{i}}$
-\end{itemize}
-
-\item For successor ordinals, we have $\canonseq{\alpha+1}{i}= \alpha$
-
-\item Finally, $\canonseq{0}{i}= \alpha$.
-\end{itemize}
-\end{mathframe}
-
-\subsubsection{Canonical Sequences in Coq}
-
-Our definition may look more complex than the mathematical one, but
-uses plain structural recursion over the type \coqsimple{T1}. Thus, tactics like
-\coqsimple{cbn}, \coqsimple{simpl}, \coqsimple{compute}, etc., are applicable. For simplicity's sake, 
-we use an auxiliary function \texttt{canonS} of type \texttt{T1 -> nat  -> T1} such that
-(\texttt{canonS  $\alpha$ $i$}) is equal to $\canonseq{\alpha}{i+1}$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS}{Epsilon0.Canon}}
-
-\label{Functions:canonS}
-\label{Functions:canon}
-\begin{Coqsrc}
-Fixpoint canonS alpha (i:nat) :=
-  match alpha with
-      zero => zero
-    | ocons zero 0 zero => zero
-    | ocons zero (S k) zero => FS k
-    | ocons gamma 0 zero =>
-      match pred gamma with
-          Some gamma' => ocons gamma' i zero
-        | None => ocons (canonS gamma i) 0 zero
-      end
-    | ocons gamma (S n) zero =>
-       match pred gamma with
-           Some gamma' => ocons gamma n (ocons gamma' i zero)
-         | None => ocons gamma n (ocons (canonS gamma i) 0 zero)
-       end
-    | ocons alpha n beta => ocons alpha n (canonS beta i)
-  end.
-\end{Coqsrc}
-
-The following function computes $\canonseq{\alpha}{i}$, except for the case $i=0$, where it simply returns $0$\;\footnote{This restriction did not prevent us from proving all the main theorems of~\cite{KS81, KP82}.
-Nevertheless, in a future version of this development, we may define $\canonseq{\alpha}{0}$ exactly as
-in~\cite{KS81}. But we are afraid this would  be done at the cost of making some proofs much more complex.}.
-
-\begin{Coqsrc}
- Definition canon alpha i := 
-   match i with 0 => zero | S j => canonS alpha j end. 
-\end{Coqsrc}
-
-For instance \coq's computing facilities allow us to verify the equalities\linebreak 
-\mathcolor{$\canonseq{\omega^\omega}{3} = \omega^3$} and
-\mathcolor{$\canonseq{\omega^\omega*3}{42} = \omega^\omega*2 + \omega^{42}$}.
-
-\begin{Coqsrc}
-Compute (canon (omega ^ omega) 3).
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = phi0 (FS 2) : T1
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Example canon3 :  canon (omega ^ omega) 3 = omega ^ 3.
-Proof. reflexivity. Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Compute pp (canon (omega ^ omega * 3) 42).  
-\end{Coqsrc}
-
-\begin{Coqanswer}
-    = (omega ^ omega * 2 + omega ^ 42)%pT1
-     : ppT1
-\end{Coqanswer}
-
-\index{Projects}
-\begin{project}
-Many lemmas presented in this chapter were stated and proved before the introduction of 
-the type class \texttt{ON} of ordinal notations, and in particular its  instance \texttt{Epsilon0}.
-Thus definitions and lemmas refer to the type \texttt{T1} of possibly not well-formed terms.
-This should be fixed in  a future version.
-\end{project}
-
-\subsection{Basic Properties of Canonical Sequences}
-
-We did not  try to prove that our definition really implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
-abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
-~\href{../theories/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}
-
-Proving the equality $\canonseq{\alpha+1}{i}=\alpha$ is not 
-as simple as suggested by the equations of definition~\ref{def:canonseq-math}\,.
-Nevertheless, we can prove it by  plain structural induction on $\alpha$.
-
-\begin{Coqsrc}
-Lemma canonS_succ i alpha :
-  nf alpha ->  canonS (succ alpha) i = alpha.
-Proof.
- induction alpha.
- (* ... *)
-\end{Coqsrc}
-
-\subsubsection{Canonical Sequences and the Order $<$}
-
-\index{Maths!Transfinite induction}
-
-We prove by transfinite induction over $\alpha$ that $\canonseq{\alpha}{i+1}$ is an ordinal strictly less than $\alpha$ (assuming $\alpha\not=0$).
-This property allows us to use the function \texttt{canonS} and its derivates in function definitions by transfinite recursion.
-
-\label{lemma:canonS_LT}
-\begin{Coqsrc}
-Lemma canonS_LT i alpha :
-  nf alpha -> alpha <> zero -> canonS alpha i t1<  alpha.
-\end{Coqsrc}
-
-\subsubsection{Limit Ordinals are Really Limits}
-The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
-is the limit of the sequence \showmath{\canonseq{\lambda}{i}\;(1\le i)}.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}
-
-\begin{Coqsrc}
-Lemma canonS_limit_strong (lambda : T1) : 
-     nf lambda ->
-     limitb lambda  ->
-     forall beta, beta t1< lambda ->
-                  {i:nat | beta t1< canonS lambda i}.
-
-Proof.
-  transfinite_induction_LT lambda.
-  (* ... *)
-Defined.
-\end{Coqsrc}
-
-\label{lemma:canonS-limit}
-
-Note the use of \coq's \texttt{sig} type in the theorem's statement, which
-relates the boolean function \texttt{limitb} defined on the \texttt{T1} data-type with a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.
-We can also state directly that $\lambda$ is a (strict) least upper bound of the elements of its canonical sequence.
-
-\begin{Coqsrc}
-Lemma canonS_limit_lub (lambda : T1) :
-  nf lambda -> limitb lambda  ->
-  strict_lub (fun i => canonS lambda i) lambda.
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}\label{exo:simply-typed-canonseq}
-Instead of using the \texttt{sig} type, define a simply typed function that, given two ordinals $\alpha$ and $\beta$, returns a natural number $i$ such that, if $\alpha$ is a limit ordinal and $\beta<\alpha$, then $\beta< \canonseq{\alpha}{i+1}$.
-Of course, you will have to prove the correctness of your function.
-\end{exercise}
-
-\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0} : Paths}
-\index{Maths!Accessibility inside epsilon0}
-\label{sect:pathes-intro}
-
-Let us consider a kind of accessibility problem inside $\epsilon_0$: given two ordinals $\alpha$ and $\beta$, where $\beta<\alpha<\epsilon_0$, find a \emph{path} consisting of a finite sequence $\gamma_0=\alpha,\dots,\gamma_l=\beta$,
-where, for every $i<l$, $\gamma_i \not= 0$ \footnote{This condition allows us to ignore paths which end by a lot of useless $0$s.} and there exists some strictly positive integer $s_i$
-such that $\gamma_{i+1}=\canonseq{\gamma}{s_i}$,
-
-Let $s$ be the sequence $\langle s_0,s_1,\dots, s_{l-1} \rangle$. We describe the
-existence of such a path with the notation $\alpha\xrightarrow [s]{}\beta$.
-
-For instance, we have $\omega*2 \xrightarrow[2,2,2,4,5]{}3$, through the 
-path $\langle\omega\times 2, \omega+2,\omega+1,\omega,4,3\rangle$.
-
-\begin{remark}
-
-Note that, given $\alpha$ and $\beta$, where $\beta < \alpha$, the sequence $s$ which leads from $\alpha$ to $\beta$ is not unique.
-
-Indeed, if $\alpha$ is a limit ordinal, the first element of $s$ can be any integer $i$ such that $\beta<\canonseq{\alpha}{i}$, and if $\alpha$ is a successor ordinal,
-then the sequence $s$ can start with any positive integer.
-
-For instance, we have also 
-$\omega*2 \xrightarrow[3,4,5,6]{}\omega$. 
-Likewise,
-$\omega*2 \xrightarrow[1,2,1,4]{} 0$ and
-$\omega*2 \xrightarrow[3,3,3,3,3,3,3,3]{} 0$.
-\end{remark}
-
-\subsection{Formal Definition}
-
-\label{path-to-definition}
-
-In \coq{}, the notion of path can be simply defined as an inductive predicate 
-parameterized by the destination $\beta$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\index{Predicates!path\_to}
-\label{sect:path-to-def}
-
-\begin{Coqsrc}
-(Definition transition_S i : relation T1 :=
-  fun alpha beta =>  alpha <> zero /\ beta = canonS alpha i.
-
-Definition transition i : relation T1 :=
-  match i with 0 => fun _ _ => False | S j => transition_S j end.
-
-Inductive path_to (beta: T1) : list nat -> T1 -> Prop :=
-  path_to_1 : forall (i:nat) alpha , 
-    i <> 0 ->
-    transition i alpha beta ->
-    path_to beta (i::nil) alpha
-| path_to_cons : forall i alpha s gamma,
-    i <> 0 ->
-    transition i alpha gamma ->
-    path_to beta  s gamma ->
-    path_to beta  (i::s) alpha.
-\end{Coqsrc}
-
-\begin{remark}
-The definition above is parameterized with the \emph{destination} of the path and indexed by the origin, hence the name \texttt{path\_to}.
-The rationale behind this choice is a personal preference of the developer  for the kind of eliminators generated by \coq{} in this case.
-The symmetric option could have been also considered (see also Remark~\vref{remark:transitive-closure}).
-\end{remark}
-
-\begin{remark}
-In the present version of our library, we use a variant \texttt{path\_toS} of
-\texttt{path\_to}, where the proposition
-(\texttt{path\_toS $\beta$ $s$ $\alpha$}) is equivalent to
-(\texttt{path\_to $\beta$ (shift $s$) $\alpha$}). This variant is scheduled to be deprecated.
-\end{remark}
-
-\index{Exercises}
-
-\begin{exercise}
-Write a tactic for solving goals of the form (\texttt{path\_to $\beta$ $s$ $\alpha$})
-where $\alpha$, $\beta$ and $s$ are closed terms. 
-You should solve automatically the following goals:
-
-\begin{Coqsrc}
- path_to omega (2::2::2::nil) (omega * 2).
-
- path_to omega (3::4::5::6::nil) (omega * 2).
-
- path_to zero (interval 3 14) (omega * 2).
-
- path_to zero (repeat 3 8) (omega * 2).
-\end{Coqsrc}
-
-\end{exercise}
-
-\subsection{Existence of a Path}
-
-\index{Maths!Transfinite induction}
-
-By transfinite induction on $\alpha$, we prove that for any $\beta<\alpha$, 
-one can build a path from $\alpha$ to $\beta$ (in other terms, $\beta$ is accessible from $\alpha$).
-
-\begin{Coqsrc}
-Lemma LT_path_to (alpha beta : T1) :
-  beta t1< alpha -> {s : list nat | path_to beta s alpha}.
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}[continuation of exercise~\vref{exo:simply-typed-canonseq}]
-Define a simply typed function for computing a path from $\alpha$ to $\beta$.
-\end{exercise}
-
-\noindent 
-From the lemma \texttt{canonS\_LT}~\vref{lemma:canonS_LT}, we can convert any path into an inequality on ordinals (by induction on paths).
-
-\begin{Coqsrc}
-Lemma path_to_LT beta s alpha :
-  path_to beta s alpha -> nf alpha -> beta t1< alpha.
-\end{Coqsrc}
-
-\subsection{Paths and Hydra Battles}
-\label{KS-o2h}
-
-In order to apply our knowledge about  ordinal numbers less than $\epsilon_0$ to the study of hydra battles, we define an injection
-from the interval $[0,\epsilon_0)$ into the type \texttt{Hydra}.
-
-\vspace{4pt}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}}
-
-\begin{Coqsrc}
-Fixpoint iota (alpha : T1) : Hydra :=
-  match alpha with
-  | T1.zero => head
-  | ocons gamma n beta => 
-         node (hcons_mult (iota gamma) (S n) (iotas beta))
-  end 
-with iotas (alpha : T1) :  Hydrae :=
-       match alpha with
-       | T1.zero => hnil
-       | ocons alpha0 n beta  => 
-           hcons_mult (iota alpha0) (S n) (iotas beta)
-       end.
-\end{Coqsrc}
-
-For instance Fig.~\ref{fig:iota-example} shows the image by $\iota$ of the ordinal  \textcolor{black}{$\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$}
-
-  \begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.3]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (10,2) {$\bullet$};
-\node (N22) at (7,2) {$\bullet$};
-\node (N3) at (14,2) {$\bullet$};
-\node (N4) at (18,2) {$\Smiley[2][green]$};
-\node (N5) at (0,4) {$\bullet$};
-\node (N6) at (2,5) {$\Smiley[2][green]$};
-\node (N7) at (4,6) {$\Smiley[2][green]$};
-\node (N88) at (7,4) {$\bullet$};
-\node (N8) at (10,4) {$\bullet$};
-\node (N9) at (14,6) {$\Smiley[2][green]$};
-\node (N10) at (0,8) {$\Smiley[2][green]$};
-\node (N11) at (10,7) {$\Smiley[2][green]$};
-\node (N111) at (7,7) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N22);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (N1) to  (N5);
-\draw (N1) to   [bend left=10] (N6);
-\draw (N1) to   [bend right=20] (N7);
-\draw (N2) to  (N8);
-\draw (N22) to  (N88);
-\draw (N8) to  (N11);
-\draw (N88) to  (N111);
-\draw (N3) to  (N9);
-\draw (N5) to  (N10);
-\end{tikzpicture}
-\caption{The hydra $\iota(\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$) \label{fig:iota-example}}
-
-\end{figure}
-
-The following lemma (proved in ~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H.v}) maps  canonical sequences to rounds of hydra battles.
-
-\label{lemma:canonS-iota}
-
-\begin{Coqsrc}
-Lemma canonS_iota i alpha :
-    nf alpha -> alpha <> 0 ->
-    iota alpha -1-> iota (canonS alpha i).
-\end{Coqsrc}
-
-The next step of our development extends this relationship to
-the order $<$ on $[0,\epsilon_0)$ on one side, and hydra battles on the other side.
-
-\begin{Coqsrc}
-Lemma path_to_battle alpha s beta :
-  path_to  beta  s alpha -> nf alpha ->
-  iota alpha -+-> iota beta.
-\end{Coqsrc}
-
-As a corollary, we are now able to transform any inequality $\beta<\alpha<\epsilon_0$ into a (free) battle.
-
-\begin{Coqsrc}
-Lemma LT_to_battle alpha beta :
-    beta t1< alpha ->  iota alpha -+-> iota beta.
-\end{Coqsrc}
-
-\section{A  Proof of Impossibility}
-
-We now have  the tools for proving that  there exists no variant bounded by some $\mu<\epsilon_0$ for proving the termination   of all battles.
-The proof we are going to show is a proof by contradiction.
-It  can
- be considered as a generalization of the
-proofs described in  sections~\vref{omega-case} and \vref{omega2-case}.
-
-In the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
-This part of the development is parameterized by some class $B$ of battles, which will be instantiated later to \texttt{free} or \texttt{standard}.
-
-\begin{Coqsrc}
-Class BoundedVariant (B:Battle) :=
-  {
-    mu:T1 ;
-    m: Hydra -> T1;
-    mu_nf: nf mu;
-    Hvar: Hvariant T1_wf B m;
-    m_bounded: forall h, m h t1< mu
-  }.
-\end{Coqsrc}
-
-Let us assume there exists such a variant:
-
-\begin{Coqsrc}
-Section Bounded.
-  Context (B: Battle) (Hy : BoundedVariant B).
-
-  Hypothesis m_decrease : forall  i h h',
-        round_n i h h' -> m h' t1< m h.
-\end{Coqsrc}
-
-\label{remark:m-decrease}
-\begin{remark}
-  The hypothesis \texttt{m\_decrease} is not provable  in general, but is satisfied by
-the  \texttt{free} and \texttt{standard} kinds of battles. This trick allows to 
-``factorize'' our proofs  of impossibility.
-\end{remark}
-
-\index{Maths!Transfinite induction}
-
-First, we prove that $m(\iota(\alpha))$ is always greater than or equal to $\alpha$, by  transfinite induction over $\alpha$.
-
-\begin{Coqsrc}
-Lemma m_ge_0 alpha:  nf alpha -> alpha t1<= m (iota alpha).
-\end{Coqsrc}
-
-\begin{itemize}
-\item If $\alpha=0$, the inequality trivially holds
-\item If $\alpha$ is the successor of  some ordinal $\beta$, the inequality $\beta \leq m(\iota(\beta))$ holds (by induction hypothesis).
-But the hydra $\iota(\alpha)$ is transformed in one round into
-$\iota(\beta)$, thus $m(\iota(\beta))<m(\iota(\alpha))$. Hence $\beta<m(\iota(\alpha))$, which implies $\alpha \leq m(\iota(\alpha))$
-\item If $\alpha$ is a limit ordinal, then $\alpha$ is the least upper bound of the set
-of all  the $\canonseq{\alpha}{i}$.  Thus, we have just to prove that $\canonseq{\alpha}{i}< m(\iota(\alpha))$ for any $i$. 
-\begin{itemize}
-\item Let $i$ be some natural number.
-By the induction hypothesis, we have $\canonseq{\alpha}{i} \leq m(\iota(\canonseq{\alpha}{i}))$.
-But the hydra $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i})$ in one round, thus $m(\iota(\canonseq{\alpha}{i})) < m(\iota(\alpha))$, by our hypothesis \texttt{m\_decrease}.
-\end{itemize}
-\end{itemize}
-
-Please note that the impossibility proofs of 
-sections~\vref{omega-case} and \vref{omega2-case} contain a similar lemma, also called \texttt{m\_ge}.
-We are now able to build a counter-example.
-
-\begin{Coqsrc}
-  Definition big_h := iota mu.
-  Definition beta_h := m big_h.
-  Definition small_h := iota beta_h.
-\end{Coqsrc}
-
-From Lemma \texttt{m\_ge\_0} we infer the following inequality :
-
-\begin{Coqsrc}
-    Corollary m_ge_generic : m big_h t1<= m small_h.
- \end{Coqsrc}
-
-The (big) rest of the proof is dedicated to prove formally the converse inequality 
-\texttt{m small\_h t1< m big\_h}.
-
-\subsection{The case of Free Battles}
-\label{sec:free-battles-case}
-Let us now consider that $B$ is instantiated to \texttt{free} (which means that we are considering proofs of termination of \emph{all} battles).
-The following lemmas are proved in Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
-The case $B=\texttt{standard}$ is studied in section~\vref{std-case}.
-
-\begin{Coqsrc}
-Section Impossibility_Proof.
-
-  Context (Var : BoundedVariant free ).
-  \end{Coqsrc}
-
-\begin{enumerate}
-\item The following lemma is an application of \texttt{m\_ge\_generic}, since \texttt{free}
-satisfies trivially the hypothesis \texttt{m\_decrease} (see page~\pageref{remark:m-decrease}).
-
-\begin{Coqsrc}
-Lemma m_ge : m big_h t1<= m small_h.
-  Proof.
-    apply m_ge_generic.
-   (* ... *)
-\end{Coqsrc}
-
-\item From the hypothesis \texttt{m\_bounded}, we have \texttt{m big\_h t1< mu}
-\item By Lemma \texttt{LT\_to\_battle}, we get a (free) battle from
-\texttt{big\_h = iota mu} to \texttt{small\_h = iota (m big\_h)}.
-
-\begin{Coqsrc}
-  Lemma  big_to_small : big_h  -+-> small_h.
-\end{Coqsrc}
-\item From the hypotheses on $m$, we infer:
-
-\begin{Coqsrc}
-Lemma m_lt : m small_h t1< m big_h.
-\end{Coqsrc}
-
-\item From lemmas \texttt{m\_ge} and \texttt{m\_lt}, and the irreflexivity of $<$, we get a contradiction.
-
-  \begin{Coqsrc}
-Theorem Impossibility_free : False.
-
-End Impossibility_Proof.
-\end{Coqsrc}
-
-\end{enumerate}
-
-We have now proved there exists no bounded variant for the class of free battles.
-
-\begin{Coqsrc}
-Check Impossibility_free.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  Impossibility_free
-     : BoundedVariant free -> False
-\end{Coqanswer}
-%%% ICI
-
-\section{The Case of Standard Battles}
-\label{sec:standard-intro}\label{std-case}
-One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
-Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}.
-
-\begin{Coqsrc}
-Lemma LT_to_round_plus alpha beta :
-    beta t1< alpha ->  iota alpha -+-> iota beta.
-\end{Coqsrc}
-
-This lemma builds a battle out of any inequality $\beta<\alpha$. 
-It is a straightforward application of \texttt{LT\_path\_to} of
-Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}:
-
-\begin{Coqsrc}
-Lemma LT_path_to (alpha beta : T1) :
-  beta t1< alpha -> {s : list nat | path_to beta s alpha}.
-\end{Coqsrc}
-
-The sequence $s$, used to build the sequence of replication factors of the battle depends on 
-$\beta$, so we cannot be sure that the generated battle is a genuine standard battle.
-
-The solution of this issue comes  once again from Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
-$\alpha_0=\alpha,\alpha_1,\dots,\alpha_k=\beta$ where $\alpha_{j+1}$ is equal
-to $\canonseq{\alpha_j}{i_j}$ where $i_j$ is \emph{any} natural number, 
-we consider various constraints on these sequences.
-In particular, a path is called \emph{standard} if $i_{j+1} = i_j + 1$ for every $j<k$.
-It  corresponds to a ``segment'' of some standard battles. 
-Please note that the vocabulary on paths is ours, but all the concepts come really from~\cite{KS81}.
-
-In \coq{}, standard paths can be defined as follows.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.KS.html}{Epsilon0.KS}}
-
-\begin{Coqsrc}
-(**  standard path from (i, alpha) to (j, beta) *)
-
-Inductive standard_pathR(j:nat)( beta:T1):  nat -> T1 -> Prop :=
-  std_1 : forall i alpha, 
-       beta = canon alpha i -> j = S i ->
-       standard_pathR j beta i  alpha
-| std_S : forall i alpha, 
-      standard_pathR j beta (S i) (canon alpha i)  ->
-      standard_pathR j beta i alpha.
-
-Definition standard_path  i alpha j beta := 
-   standard_pathR j beta i alpha.
-\end{Coqsrc}
-
-In the mathematical text and figures, we shall use the notation 
-$\alpha \xrightarrow[i,j]{}\beta$ for the proposition 
-(\texttt{standard\_path $i$ $\alpha$ $j$ $\beta$}).
-In~\cite{KS81} the notation is
-$\alpha \xrightarrow[i]{*}\beta$
-for 
-the proposition  $\exists j, i<j \wedge \alpha \xrightarrow[i,j]{} \beta$.
-
-Our goal is now  to transform any inequality $\beta<\alpha<\epsilon_0$ into a standard path $\alpha \xrightarrow[i,j]{} \beta$ for some $i$ and $j$, then into a standard battle
-from $\iota(\alpha+i)$ to $\iota(\beta)$. 
-Following~\cite{KS81}, we proceed in two stages:
-\begin{enumerate}
-\item we simulate plain (free) paths from $\alpha$ to $\beta$ with
-paths made of steps $(\gamma,\canonseq{\gamma}{n})$, \emph{with the same $n$ all along the path}
-\item we simulate any such path by a standard path.
-\end{enumerate}
-
-\subsection{Paths with a Constant Index}
-
-First of all, paths with a constant index 
-enjoy nice properties. They are defined as paths where all the $i_j$ are equal to the same natural number $i$, for some $i>0$.
-
-Like in~\cite{KS81}, we shall use the notation $\alpha \xrightarrow[i]{} \beta$ for denoting such a path, also called an $i$-path.
-
-\begin{Coqsrc}
-Definition const_pathS i :=
-    clos_trans_1n T1 (fun alpha beta => beta = canonS alpha i).
-
-Definition const_path i alpha beta :=
-  match i with
-    0 => False
-  | S j => const_pathS j alpha beta
-end.
-\end{Coqsrc}
-
-% Paths with a given index can be effectively computed.
-% Given $i$, $\alpha$ and $l$, the following function returns the ordinal $\beta$ such that there exists a path 
-% $\alpha \xrightarrow [i+1] {} \beta$ of length $l$.
-
-% \begin{Coqsrc}
-% Fixpoint const_funS (i:nat)(alpha : T1)(l:nat):  T1  :=
-%   match l
-%   with
-%   | 0 => alpha
-%   | S m => const_funS i (canonS i alpha) m
-%   end.
-% \end{Coqsrc}
-
-% The following computations show  applications of \texttt{constS\_fun} to the 
-% ordinal $\omega^\omega$, with various values of $i$ and $l$.
-
-% \begin{Coqsrc}
-% Compute  (const_funS 2 (omega ^omega)  55).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = zero
-%      : T1 
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (const_funS 2 (omega ^omega) 15).
-% \end{Coqsrc}
-
-%   \begin{Coqanswer}
-%  = (omega ^ 2 * 2)%pT1
-%      : ppT1   
-%   \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (const_funS 4 (omega^omega)  100).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% = (omega ^ 4 * 4 + omega ^ 3 * 4 + omega ^ 2 + omega * 4 + 4)%pT1
-%      : ppT1
-% \end{Coqanswer}
-
-A most interesting property of $i$-paths is that we can ``upgrade'' their index, as stated by K.\&S.'s Corollary 12.
-
-\index{Maths!Transfinite induction}
-
-\begin{Coqsrc}
-Corollary Cor12 (alpha : T1) :  nf alpha ->
-         forall beta i n, beta  t1< alpha  ->
-                i < n ->
-                 const_pathS i alpha beta ->
-                 const_pathS n alpha beta.
-Proof.
-  transfinite_induction_lt alpha.
-  (* (long) proof skipped *)
-\end{Coqsrc}
-
-We  also use a version of \texttt{Cor12} with large inequalities.
-
-\begin{Coqsrc}
-Corollary Cor12_1 (alpha : T1) :
-  nf alpha ->
-  forall beta i n, beta t1< alpha ->
-                   i <= n ->
-                   const_pathS i alpha beta ->
-                   const_pathS n alpha beta.
-\end{Coqsrc}
-
-\subsubsection{Sketch of Proof of \texttt{Cor12}}
-\index{Maths!Transfinite induction}
-
-We prove this lemma by transfinite induction on $\alpha$.
-Let us consider a path $\alpha \xrightarrow [i]{} \beta$ $(i>0)$. Its first step is
-the pair $(\alpha,\canonseq{\alpha}{i})$, We have $\canonseq{\alpha}{i}<\alpha$ and
-$\canonseq{\alpha}{i} \xrightarrow [i]{} \beta$. 
-Let $n$ be any natural number such that $n>i$.
-By the induction hypothesis, there exists a path $\canonseq{\alpha}{n} \xrightarrow[i]{} \beta$.
-\begin{itemize}
-\item  If $\alpha$ is a successor ordinal $\gamma+1$, then $\canonseq{\alpha}{n} =
-\canonseq{\alpha}{i}=\gamma$. Thus we have a path 
-$\alpha  \xrightarrow [n]{}  \gamma \xrightarrow [n]{} \beta$
-\item If $\alpha$ is a limit ordinal, we apply the following theorem (numbered \texttt{2.4} in Ketonen and Solovay's article).
-
-%   \begin{theorem}
-% Let $\lambda$ be a limit ordinal, then for any pair of indices $0<i<j$, there is a path $\canonseq{\lambda}{j} \xrightarrow[1]{} \canonseq{\lambda}{i}$.    
-%   \end{theorem}
-
-  \begin{Coqsrc}
-Theorem Theorem_2_4 (lambda : T1) :
-   nf lambda ->
-   limitb lambda  ->
-   forall i j, (i < j)%nat ->
-               const_pathS 0 (canonS lambda j)
-                             (canonS lambda i). 
-  \end{Coqsrc}
-
- We build the following paths :
-
- \begin{enumerate}
- \item $\alpha \xrightarrow[n]{} \canonseq{\alpha}{n}$
- \item $\canonseq{\alpha}{n} \xrightarrow[1]{} \canonseq{\alpha}{i}$ (by \texttt{Theorem\_2\_4}),
-\item $\canonseq{\alpha}{n} \xrightarrow[n]{} \canonseq{\alpha}{i}$ (applying the induction hypothesis to the preceding path);
-\item $\canonseq{\alpha}{i} \xrightarrow[n]{} \beta$ (applying the induction hypothesis)\item $\alpha \xrightarrow[n]{} \beta$ (by composition of 1, 3, and 4).
-
- \end{enumerate}
-
-\end{itemize}
-
-\begin{remark}
- \texttt{Cor12} ``casts'' $i$-paths into $n$-paths for any $n>i$.
-But the obtained $n$-path can be much longer than the original $i$-path.
-The following exercise will give an idea of this increase. 
-\end{remark}
-
-\index{Exercises}
-\begin{exercise}
-  Prove that  the length of the $i+1$-path from
-  $\omega^\omega$ to $\omega^i$ is $1 + (i+1)^{(i+1)}$, for any $i$. Note that the $i$-path from
-  $\omega^\omega$ to $\omega^i$ is only one step long.
- \end{exercise}
-
-Why is \texttt{Cor12} so useful? 
-Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
-we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying collorary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
-where $n$ is the maximum of the indices $i$ met in the induction.
-
- Lemma 1, Section 2.6 of~\cite{KS81} is naturally expressed in terms of \coq's
-\verb@sig@ construct.
-
-\label{lemma:L-2_6-1}
-\index{Coq!Techniques!Sigma types}
-
-\begin{Coqsrc}
-Lemma Lemma2_6_1 (alpha : T1) :  
-  nf alpha -> forall beta,  beta t1< alpha  ->
-  {n:nat | const_pathS n alpha beta}.
-Proof.
-  transfinite_induction alpha.
-  (* ... *)
-\end{Coqsrc}
-
-Intuitively, lemma   \texttt{L2\_6\_1}  shows that if $\beta<\alpha<\epsilon_0$, then there exists  a battle from $\iota(\alpha)$ to $\iota(\beta)$ where the replication factor is constant, although large enough.
-
-\subsection{Casting Paths with a Constant Index into a Standard Path}
-
-%%% traduire la v.f.  (voir %%% A traduire %%%% )
-
-The article~\cite{KS81} contains 
-the following lemma, the proof of which is quite complex, which allows to simulate $i$-paths by $[i+1,j]$-paths, where $j$ is large enough.
-
-\begin{Coqsrc}
-(* Lemma 1 page 300 of [KS] *)
-
-Lemma constant_to_standard_path 
-  (alpha beta : T1) (i : nat):
-  nf alpha -> const_pathS i alpha beta -> zero  t1< alpha ->
-  {l:nat | standard_path (S i) alpha j beta}.
-\end{Coqsrc}
-
-\subsubsection{Sketch of Proof of \texttt{constant\_to\_standard\_path}}
-
-Our proof follows the proof by Ketonen and Solovay, including its organization as a sequence of lemma.  Since it is a non-trivial proof, we will comment its main steps below.
-
-\subsubsection*{Préliminaries}
-
-Please note that, given an ordinal $\alpha:\texttt{T1}$, and two natural numbers $i$ and $l$, there exists at most a standard path $\alpha \xrightarrow [i,i+l]{*} \beta$.
-The following function computes $\beta$ from $\alpha$, $i$ and $l$.
-
-\begin{Coqsrc}
-Fixpoint standard_gnaw (i:nat)(alpha : T1)(l:nat):  T1  :=
-  match l with
-  | 0 => alpha
-  | S m => standard_gnaw (S i) (canon alpha i) m
-  end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-  Compute standard_gnaw 2 omega 15.
-(*   = zero
-     : T1 *)
-Compute pp (standard_gnaw 2 (omega^omega)  10).
-(*
-= (omega + 7)%pT1
-     : ppT1
-*)
-Compute pp (standard_gnaw 4 (omega^omega)  100).
-(*
- = (omega ^ 3 * 4 + omega ^ 2 * 5 + omega * 3 + 39)%pT1
-     : ppT1 *)
-\end{Coqsrc}
-
-\index{Maths!Transfinite induction}
-
-By transfinite induction over  $\alpha$, we prove that the ordinal $0$ is reachable from any ordinal $\alpha<\epsilon_0$ by some standard path.
-
-\begin{Coqsrc}
-Lemma standard_path_to_zero :
-  forall  alpha i, nf alpha ->
-                   {j: nat | standard_path (S i) alpha j zero}.
-\end{Coqsrc}
-
-\paragraph*{}
-Noq, let us consider two ordinals  $\beta<\alpha<\epsilon_0$.  Let $p$  be some $(n+1)$-path from $\alpha$ to $\beta$.
-
-\begin{Coqsrc}
- Section Constant_to_standard_Proof.
-
-  Variables (alpha beta: T1) (n : nat).
-  Hypotheses (Halpha: nf alpha) (Hpos : zero t1<  beta)
-             (p : const_pathS n alpha  beta).
-\end{Coqsrc}
-
-Applying \texttt{standard\_path\_to\_zero}, $0$ is reachable from $\alpha$ by some standard path  (see figure~\vref{fig:belle-preuve-1}).
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \draw[->, very thick,blue] (alpha)-- node [below]{$n+1$} node [above] {$+$} (beta);
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-    \node (alpha3) at (20,5) {};
-  \node (alphalast) at (35,5) {};
-  \node (zero) at (45,0) {$0$};
-  \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate=40]{$n+1$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha2)-- node [below]{$n+3$}  (alpha3);
-
-  \node (dots) at (24,5) {$\dots$};
-  \draw [->, dashed, very thick,blue] (alphalast)-- node [below, rotate=-26]{$n+p+1$}  (zero);
-
-\end{tikzpicture}
-\caption{A nice proof (1)}
-  \label{fig:belle-preuve-1}
-\end{figure}
-
-\paragraph*{}
-
-Since comparison on \texttt{T1} is decidable, one can compute the last step $\gamma$ of the standard path from $(\alpha,n+1)$  such that $\beta\leq \gamma$.
-Let $l$ be the length of the path from $\alpha$ to $\gamma$.  
-This step of the proof is illustrated in figure~\vref{fig:belle-preuve-2}.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (17,5) {$\ldots$};
-    \node (alpha3) at (20,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (38,0) {$\delta$};
-    \draw [->, dashed,very thick,blue] (alpha)-- node [below,rotate=35]{$n+1$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below,rotate = -48]{\tiny $n+l$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{$n+l+1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{$n+1$} node [above] {$+$} (beta);
-   \draw[thick] (alpha)--  (gamma);
-   \draw[thick] (gamma)--  node [above] {$\geq$} (beta);
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-\end{tikzpicture}
-
-\caption{A nice proof (2)}
-  \label{fig:belle-preuve-2}
-\end{figure}
-
-\paragraph*{}
-
-\begin{itemize}
-\item If $\beta=\gamma$, its OK! We have got a standard path
-from  
-$\alpha$ to $\beta$ with successive indices  $n+1, n+2, \dots, n+l+1$
-
-\item Otherwise,  $\beta < \gamma$.  Let us consider  $\delta=\canonseq{\gamma}{n+l+1}$.
-By applying several times lemma \texttt{Cor12},  one converts  every path of Fig~\ref{fig:belle-preuve-2} into
- a $n+l+1$-path  (see figure~\ref{fig:belle-preuve-3}).
-
-But $\gamma$ is on the $n+l+1$-path from $\alpha$ to $\beta$.
-As shown by figure~\vref{fig:fin-belle-preuve}, the ordinal $\delta$, reachable from
-$\gamma$ in one single step,  must be greater than or equal to $\beta$, which contradicts our  hypothesis $\beta < \gamma$.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-    \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (17,5) {$\ldots$};
-    \node (alpha3) at (20,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (38,0) {$\delta$};
-     \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -48]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\tiny $n+l+1$} node [above]{\color{red} $1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
-    \draw[thick] (gamma)--   node [above]{\color{red} $>$}(beta);
-   \draw[thick] (alpha)--  (gamma);
-
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-
-\end{tikzpicture}
-
-\caption{A nice proof (3)}
-  \label{fig:belle-preuve-3}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (15,5) {$\ldots$};
-    \node (alpha3) at (18,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (42,0) {$\delta$};
-    \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -36]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\small $n+l+1$} node [above]{\color{red} $1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
-   \draw[thick] (alpha)--  (gamma);
-
-    \draw[thick] (gamma)--  node [below]{\tiny $n+l+1$} node [above]{\color{red} $+$}(beta);
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-
-\end{tikzpicture}
-
-\caption{A nice proof (4)}
-  \label{fig:fin-belle-preuve}
-\end{figure}
-
-\end{itemize}
- The only possible case is  thus $\beta=\gamma$, so we have got a standard path  from $\alpha$ to $\beta$.
-
-\begin{Coqsrc}
- Lemma constant_to_standard_0 : 
-    {l : nat | standard_fun (S n) alpha l = beta}.
- (* ... *)
-
-End Constant_to_standard_Proof.
-\end{Coqsrc}
-
-Here is the full statement of the conversion from constant to standard paths.
-
-\begin{Coqsrc}
-Lemma constant_to_standard_path 
-  (alpha beta : T1) (i : nat):
-  nf alpha -> const_pathS i alpha beta -> zero  t1< alpha ->
-  {j:nat | standard_path (S i) alpha j beta}.
-\end{Coqsrc}
-
-Applying \texttt{Lemma2\_6\_1} and \texttt{constant\_to\_standard\_path}, we get the following corollary.
-
-\begin{Coqsrc}
-Corollary  LT_to_standard_path  (alpha beta : T1) :
-  beta t1< alpha ->
-  {n : nat & {j:nat | standard_path (S n) alpha j beta}}.
-\end{Coqsrc}
-
-\subsection{Back to Hydras}
-\label{sec:standard-battles-cases}
-We are now able to complete our proof that there exists no bounded variant for proving the termination of standard hydra battles. This proof can
-be consulted in the module 
-\url{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}.
-Please note that it has the same global structure as in section\ref{sec:free-battles-case} 
-% ICI !
-Applying the  lemmas  \texttt{Lemma2\_6\_1} of the module 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
-{\texttt{Lemma2\_6\_1}}   and 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
-{\texttt{constant\_to\_standard\_path}},
-we can convert any inequality $\beta<\alpha<\epsilon_0$ into a standard path from
-$\alpha$ to  $\beta$, then into a fragment of a standard battle from 
-$\iota(\alpha)$ to $\iota(\beta)$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html\#LT_to_standard_battle}{Hydra.Epsilon0\_Needed\_Std}}
-
-\begin{Coqsrc}
-Lemma LT_to_standard_battle :
-    forall alpha beta,
-      beta t1< alpha ->
-      exists n i,  battle standard  n (iota alpha) i (iota beta).
-\end{Coqsrc}
-
-Next, please consider the following context:
-
-\begin{Coqsrc}
-Section Impossibility_Proof.
-
-  Context (Var : BoundedVariant standard).
- \end{Coqsrc}
-
-In the same way as for free battles, we import a large inequality 
-from 
-the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}.
-
-\begin{Coqsrc}
- Lemma m_ge : m big_h t1<= m small_h.
-\end{Coqsrc}
-
-\paragraph*{} If remains to prove the following strict inequality, in order to have a contradiction.
-
-\begin{Coqsrc}
-Lemma m_lt : m small_h  t1< m big_h.
-\end{Coqsrc}
-
-\paragraph*{Sketch of proof:} Let us recall that $\texttt{big\_h} = \iota(\mu)$
- and $\texttt{small\_h} = \iota (m (\texttt{big\_h}))$.
-
-Since $m(\texttt{big\_h})< \mu$, there exists a standard path from $\mu$ to
-$m(\texttt{big\_h})$, hence a   standard battle from $\iota(\mu)$  to
-$\iota(m(\texttt{big\_h}))$,  i.e. from \texttt{big\_h} to \texttt{small\_h}.
-
-Since $m$ is assumed to be a variant for standard battles, we get the inequality  $m(\texttt{small\_h}) < m(\texttt{big\_h})$.
-
-\subsection{Remarks}
-
-We are grateful to 
- J. Ketonen and R. Solovay  for the high quality of their explanations and proof details.
-Our proof follows tightly the sequence of lemmas in their article, with a focus on 
-constructive aspects.
-Roughly steaking, our implementation \emph{builds}, out of a hypothetic 
-  variant $m$, bounded by some ordinal $\mu<\epsilon_0$, a hydra \texttt{big\_h} which verifies the impossible inequality  $m(\texttt{big\_h})< m(\texttt{big\_h})$.
-
-On may ask whether the preceding results are not too restrictive, since they 
-refer to a particular data type \texttt{T1}.
-In fact, our representation of ordinals strictly less than 
- $\epsilon_0$ is faithful to their mathematical definition, at least 
-Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
-(please see also \href{../theories/html/hydras.Schutte.Correctness_E0.html}{the module \texttt{Ordinals.Schutte.Correctness\_E0}}).
-
-Thus, we can infer that our theorems can be applied to any well order.
-
-\index{Projects}
-\begin{project}
-Study a possible modification of the definition of a variant  (for  standard battles).
-
-\begin{itemize}
-\item The variant is assumed to be strictly decreasing \emph{on configurations 
-reachable from some initial configuration where the replication factor is equal to $0$}
-\item The variant may depend on the number of the current round.
-\end{itemize}
-
-In other words, its type should be \texttt{nat -> Hydra -> T1}, and it must 
-verify the inequality $m\, (S\,i)\, h' < m\,i\, h$ whenever the configuration 
-$(i,h)$ is reachable from some initial configuration $(0,h_0)$
-and \texttt{h} is transformed into \texttt{h'} in the considered round.
-Can we still prove the theorems of section~\ref{std-case} with this new definition?
-
-\end{project}
-
-%---------------------------------------------------------------------
-\chapter{Large Sets and Rapidly Growing Functions}\label{chap:alpha-large}
-
-%\section{Introduction}
-
-In this chapter, we try to feel how long a standard battle can be.
-To be precise, for any ordinal $\alpha<\epsilon_0$ and any positive integer $k$,
-we give a minoration of the number of steps of a standard battle which
-starts with the hydra $\iota(\alpha)$ and the replication factor $k$.
-
-We express this number in terms of the Hardy hierarchy of fast-growing 
-functions~\cite{BW85, Wainer1970, KS81, Promel2013}.
- From the \coq{} user's point of view, such  functions are  very 
-attractive:  they are defined as functions  in \gallina{}, and we can apply them \emph{in theory}, but they are so complex that you will never be able to look at the result of the computation.
- Thus, our knowledge on these functions must rely on \emph{proofs}. In our development, we use often the rewriting rules generated by \coq's \texttt{Equations} plug-in.
-
-\section{Definitions}
-
-%\subsection{Definition}
-
-\begin{definition}
-Let $0<\alpha<\epsilon_0$ be any ordinal, and $s$ a sequence of strictly positive natural numbers. 
-We say that $s$ is \emph{minimally $\alpha$-large} (in short:
-\emph{$\alpha$-mlarge}) if $s$ if $s$ is $\alpha$-large 
- and every strict prefix of $s$ leads to a non-zero ordinal (\emph{cf} Sect.~\vref{sect:path-to-def}).
-
-\index{Maths!Large sequences}
-\index{Maths!Minimal large sequences}
-
-\end{definition}
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}
-
-\index{Predicates!mlarge @ mlarge (minimal large sequences)}
-\begin{Coqsrc}
-Definition mlarge alpha (s:list nat) := path_to zero s alpha.
-\end{Coqsrc}
-
-\begin{remark}
-  Ketonen and Solovay~\cite{KS81} consider  large finite \emph{sets} of natural numbers,  but they are mainly used as sequences. Thus, we chosed to represent them explicitely as (sorted) lists.
-
-They also consider large (but not minimally large) sets. They can be defined in
-\coq{} as follows:
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}
-
-\begin{Coqsrc}
-Fixpoint gnaw (alpha : T1) (s: list nat) :=
-  match s with
-    | nil => alpha
-    | (0::s') => gnaw  alpha s'
-    | (S i :: s')  =>  gnaw (canonS i alpha) s'
-  end.
-
-Definition large alpha (s:list nat) := gnaw alpha s = zero.
-\end{Coqsrc}
-\end{remark}
-
-Let us consider two integers $k$ and $l$, such that $0<k<l$. In order to check whether the interval $[k,l]$ is minimally large for $\alpha$, it is enough to
-follow from $\alpha$ the path associated with the interval $[k,l($ and verify that the last ordinal we obtain is equal to $1$.
-
-\subsection{Example}
-
-For instance the interval $[6,70]$ leads $\omega^2$ to $\omega\times 2 + 56$. Thus this interval is not $\omega^2$-mlarge.
-
-\begin{Coqsrc}
-Compute pp (gnaw (omega * omega) (interval 6 70)).
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = (omega * 2 + 56)%pT1
-     : ppT1
-\end{Coqanswer}
-
-Let us try another computation.
-
-\begin{Coqsrc}
-Compute (gnaw (omega * omega) (interval 6 700)).
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = zero : T1
-\end{Coqanswer}
-
-We may say that the interval $[6,700]$ is $\omega^2$-large, since it leads to $0$, but nothing assures us that the condition of minimality is satisfied.
-
-The following lemma relates minimal largeness with the function 
-\texttt{gnaw}.
-
-\begin{Coqsrc}
-Lemma mlarge_iff alpha x (s:list nat) :
-  s <> nil -> ~ In 0 (x::s) ->
-  mlarge alpha (x::s) <-> gnaw alpha (but_last x s) = one.
- \end{Coqsrc}
-
-For instance, we can verify that the interval $[6,510]$ is $\omega^2$-mlarge.
-
-\vspace{4pt}
- \noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html}{Epsilon0.Large\_Sets\_Examples}}
-\begin{Coqsrc}
-Example Ex1 : mlarge (omega * omega) (interval 6 510).
-\end{Coqsrc}
-
-\section{The Length of Minimal Large Sequences}
-
-Now, consider any natural number $k>0$. We would like to compute
-a number $l$ such that the interval $[k,l]$ is $\alpha$-mlarge. So, 
-the standard battle starting with $\iota(\alpha)$ and the replication factor $k$ will end after $(l-k+1)$ steps.
-
-First, we notice that this  number $l$ exists, since the segment $[0,\epsilon_0)$ is well-founded and $\canonseq{\alpha}{i}<\alpha$ for any $i$ and $\alpha>0$.
-Moreover, it is unique:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-\begin{Coqsrc}
-Lemma mlarge_unicity alpha k l l' : 
-  mlarge alpha (interval (S k) l) ->
-  mlarge alpha (interval (S k) l') ->
-  l = l'.
-\end{Coqsrc}
-
-Thus, it seems obvious that there must exist a function, parameterized by $\alpha$ which associates to any  strictly positive integer $k$ the number $l$ such that
-the interval $[k,l]$ is $\alpha$-mlarge. It would be fine to write in \gallina{} a definition like this:
-
-\begin{Coqbad}
-Function L (alpha: E0) (i:nat) :  nat := ...
-\end{Coqbad}
-
-But we do not know how to fill the dots yet \dots{}   In the next section, we will 
-use \coq{} to reason  about the \emph{specification} of \texttt{L},
-prove properties of any function which satisfies this specification.
-In Sect.~\ref{sect:L-equations}, we use the \texttt{coq-equations} plug-in
-to define a function \texttt{L\_}, and prove its correctness w.r.t. its specification.
-
-%\subsection{Looking for a function definition}
-
-Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider the functions which associate to each stritly positive integer $k$ the number $l$, where 
-the interval $[k,l)$ is $\alpha$-mlarge.
-
-\begin{remark}
-The upper bound of the considered interval has been chosen to be  $l-1$ and not $l$, in order to simplify some statements and proofs in composition lemmas associated to ordinals of the form $\alpha\times i$ and
-$\omega^\alpha\times i + \beta$.
-In order to consider any ordinal below $\epsilon_0$, we consider a special case for $\alpha=0$.
-\end{remark}
-
-\subsection{Formal Specification}
-
-Our specification of the function \texttt{L} is as follows:
-
-\begin{Coqsrc}
-Inductive L_spec : T1 -> (nat -> nat) -> Prop :=
-| L_spec0 : forall f, (forall k, f k = k) ->  L_spec zero f
-| L_spec1 : forall alpha f,
-    alpha <> zero ->
-    (forall k, mlarge alpha (interval (S k) (Nat.pred (f (S k))))) ->
-    L_spec alpha f.
-\end{Coqsrc}
-
-Note that, for $\alpha\not=0$, the value of $f(0)$ is not specified.
-Nevertheless, the restriction of $f$ to the set of strictly positive integers is unique (up to extensionnality).
-
-\begin{Coqsrc}
-Lemma L_spec_unicity alpha f g :
-  L_spec alpha f -> L_spec alpha g -> forall k, f (S k) = g (S k).
-\end{Coqsrc}
-
-\subsection{Abstract Properties}
-
-Let us now prove properties of any function $f$ (if any) which satisfies 
-\texttt{L\_spec}.
-We are looking for properties which could be used for writing \emph{equations} and prove the correctness of the function generated by the \texttt{coq-equations} plug-in.
-Moreover, they will give us some examples of
-$L_\alpha$ for small values of $\alpha$.
-
-Our exploration of the $L_\alpha$s  follows the usual scheme : transfinite induction, and proof-by-cases : zero, successors and limit ordinals.
-
-\index{Maths!Transfinite induction}
-
-\subsubsection{The  Ordinal Zero}
-\label{sect:L-spec-zero}
-The base case is directly a consequence of the specification.
-
-\begin{Coqsrc}
-Lemma L_zero_inv f : L_spec zero f -> forall k, f (S k) = S k.
-\end{Coqsrc}
-
-\subsubsection{Successor Ordinals}
-\label{sect:L-spec-succ}
-Let $\beta$ be some ordinal, and assume the arithmetic function $f$ satisfies 
-the specification $(\texttt{L\_spec}\;\beta)$.  Let $k$ be any natural number.
-Any path from $\texttt{succ}\,\beta$ to $0$ starting at $k+1$ can be decomposed into a first step from $\texttt{succ}\,\beta$ to $\beta$, then a path from
-$\beta$ at $k+2$ to $0$. 
-By hypothesis the interval $[k+2, f(k+2)-1]$ is $\beta$-mlarge.
-But the interval $[k+1, f(k+2)-1]$ is the concatenation of the singleton
-$\{k+1\}$ and the interval $[k+2, f(k+2)-1]$.
-So, the function $\lambda\,k.\,f(k+1)$ satisfies the specification $\texttt{L\_spec}\,\beta$.
-
-Note that our decomposition of intervals works only if the intervals we consider are not empty.
-In order to ensure this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <<= f}, or \texttt{(fun\_le S f)} (defined
-in~\href{../theories/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}).
-
-\begin{Coqsrc}
-Definition fun_le f g  := forall n:nat,  f n <=  g n.
-\end{Coqsrc}
-
-It looks also natural to show that the functions we consider are strictly monotonous. The section on successor ordinals has thus the following structure.
-
-\begin{Coqsrc}
-Section succ.
-   Variables (beta : T1) (f : nat -> nat).
-
-   Hypotheses (Hbeta : nf beta)
-              (f_mono : strict_mono f)
-              (f_Sle : S <<= f)
-              (f_ok : L_spec beta f).
-
-   Definition L_succ := fun k => f (S k).
-
-   Lemma L_succ_mono : strict_mono L_succ.
-
-   Lemma L_succ_Sle : S <<= L_succ.
-
-   Lemma L_succ_ok : L_spec (succ beta) L_succ.
-
-End succ.
-
-\end{Coqsrc}
-
-\subsubsection{Limit Ordinals}
-\label{sect:L-spec-lim}
-
-Let $\lambda<\epsilon_0$ be any limit ordinal. In a similar way as for successors, we decompose any path from $\lambda$ (at $k$) into a step to
-$\canonseq{\lambda}{k}$, then to $0$. In the following section, we assume that there exists à correct function for $\canonseq{\lambda}{k}$, \emph{for any strictly positive $k$}.
-
-\begin{Coqsrc}
-Section lim.
-  Variables (lambda : T1)
-            (Hnf : nf lambda)
-            (Hlim : limitb lambda)
-            (f : nat -> nat -> nat)
-            (H : forall k, L_spec (canonS lambda k) (f (S k))).
-
-  Let L_lim k := f k (S k).
-
-  Lemma L_lim_ok : L_spec lambda L_lim.
-
-End lim.
-\end{Coqsrc}
-
-\subsection{First Results}
-
-Applying the previous lemmas on successors and limit ordinals, 
-we get several correct implementations of \texttt{(L\_spec $\alpha$)} for small values of $\alpha$.
-
-\subsubsection{Finite Ordinals}
-
-By iterating the functional \texttt{L\_succ}, we get a realization of
-\texttt{(L\_spec (fin $i$))} for any natural number $i$.
-
-\begin{Coqsrc}
-Definition L_fin i := fun k => (i + k)%nat.
-
-Lemma L_fin_ok i : L_spec (fin i) (L_fin i).
-\end{Coqsrc}
-
-\subsubsection{The First Limit Ordinal  \texorpdfstring{$\omega$}{omega}}
-
-The lemmas \texttt{L\_fin\_ok} and \texttt{L\_lim\_ok}   allow us to get 
-by diagonalization a correct implementation for 
-\texttt{L\_spec omega}.
-
-\begin{Coqsrc}
-Definition L_omega k := S (2 * k)%nat.
-
-Lemma L_omega_ok : L_spec omega L_omega.
-\end{Coqsrc}
-
-\subsubsection{Towards  \texorpdfstring{$\omega^2$}{omega*omega}}
-
-We would like to get exact formulas for the ordinal $\omega^2$, a.k.a.
-$\phi_0(2)$. This ordinal is the limit of the sequence $\omega\times i\;(i \in \mathbb{N}$. Thus, we have to study ordinals of this form, then use 
-our lemma on limits.
-
-The following lemma establishes a path from $\omega\times ( i+1)$ to
-$\omega \times i$.
-
-\begin{Coqsrc}
-Lemma path_to_omega_mult (i k:nat) :
-  path_to (omega * i) (interval (S k) (2 * (S k))) (omega * (S i)).
-\end{Coqsrc}
-
-Let us consider a path from  $\omega\times(i+1)$ to $0$ starting at $k+1$.
-A first ``big step'' will lead to $\omega\times i$ at $2(k+1)$. If $i>0$, the
-next jump leads to $\omega\times(i-1)$ at $2(2(k+1))+1$, etc.
-
-The following lemma expresses the length of the mlarge sequences associated with the finite multiples of $\omega$.
-
-\begin{Coqsrc}
-Lemma omega_mult_mlarge_0 i  : forall k,
-    mlarge  (omega * (S i))
-            (interval (S k)
-                      (Nat.pred (iterate (fun p =>  S (2 * p)%nat)
-                                         (S i)
-                                         (S k)))).
-\end{Coqsrc}
-
-Thus, we infer the following result:
-
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#L_omega_mult}{Epsilon0.Large\_Sets}}
-
-\begin{Coqsrc}
-Definition L_omega_mult i (x:nat) :=  iterate L_omega i x.
-
-Lemma L_omega_mult_ok (i: nat) :  L_spec (omega * i) (L_omega_mult i).
-\end{Coqsrc}
-
-For instance, let us consider the ordinal $\omega\times 8$, and a sequence 
-starting at $k=5$.
-
-\begin{Coqsrc}
-Compute L_omega_mult 8 5.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-= 1535
-     : nat
-\end{Coqanswer}
-
-More generally, we prove the equality $L_{\omega\times i}(k)=2^i\times(k+1)-1$.
-
-\begin{Coqsrc}
-Lemma L_omega_mult_eqn (i : nat) :
-  forall (k : nat),  (0 < k)%nat  ->
-                     L_omega_mult i k = (exp2 i * S k - 1)%nat.
-\end{Coqsrc}
-
-By diagonalization, we obtain a simple formula for $L_{\omega^2}$.
-
-\begin{Coqsrc}
-Definition L_omega_square k := iterate (fun z => S (2 * z)%nat)
-                                        k
-                                        (S k).
-
-Lemma L_omega_square_eqn k :
-  (0 < k)%nat ->
-  L_omega_square k = (exp2 k * (k + 2) - 1)%nat.
-
-Lemma L_omega_square_ok: L_spec (omega * omega) 
-          L_omega_square.
-
-Compute L_omega_square 8.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = 2559
-     : nat
-\end{Coqanswer}
-
-%%%% ICI
-
-\subsubsection{Going Further}
-Let us consider a last example, ``computing'' $L_{\omega^3}$.
-Since the canonical sequence associated with this ordinal is composed of the
-$\omega^2\times i\;(i\in\mathbb{N}_1)$, we have to study this sequence.
-
-To this end, we prove a generic lemma, which expresses $L_{\omega^\alpha\times i}$ as an iterate of $L_{\omega^\alpha}$.
-Note that in this lemma, we assume that the fonction associated with $\alpha$ is stritly monotonous and
-greater or equal than the successor function, and prove that $L_{\omega^\alpha\times i}$satisfies  the same properties.
-
-\begin{Coqsrc}
-Section phi0_mult.
- Variables (alpha : T1) (f : nat -> nat).
- Hypotheses (Halpha : nf alpha)
-            (f_mono : strict_mono f)
-            (f_Sle : S <<= f)
-            (f_ok : L_spec (phi0 alpha) f).
-
- Definition L_phi0_mult i := iterate f i.
-
-Lemma L_phi0_mult_ok i: 
-  L_spec (ocons alpha i zero)  (L_phi0_mult (S i)).
-
- Lemma L_phi0_mult_smono i: strict_mono (L_phi0_mult i).
-
- Lemma L_phi0_mult_Sle i: S <<=  L_phi0_mult (S i).
-
-End phi0_mult.
-\end{Coqsrc}
-
-Let us look 
-at the ordinal $\omega^2\times i$, using \texttt{L\_phi0\_mult}
-
-\begin{Coqsrc}
- Definition L_omega_square_times i :=  iterate L_omega_square i.
-
- Lemma L_omega_square_times_ok i : 
-    L_spec (ocons 2 i zero) (L_omega_square_times (S i)).
- Proof.
-  apply L_phi0_mult_ok.
-  -  auto with T1.
-  -  apply L_omega_square_Sle.
-  -  apply L_omega_square_ok.
- Qed.
-\end{Coqsrc}
-
-We are now ready to get an exact formula for $L_{\omega^3}$. 
-\begin{Coqsrc}
-Definition L_omega_cube  := L_lim  L_omega_square_times .
-
-Lemma L_omega_cube_ok : L_spec (phi0 3) L_omega_cube.
-\end{Coqsrc}
-
-The function  $L_{\omega^3}$  is just obtained by diagonalization upon $L_{\omega^2\times i}$.
-
-\begin{Coqsrc}
-Lemma L_omega_cube_eqn i : 
-   L_omega_cube i = L_omega_square_times i (S i).
-Proof. reflexivity. Qed.
-\end{Coqsrc}
-
-Thus, for instance, $L_{\omega^3}(3)=L_{\omega^2\times 4}(3)$.
-Thus, we obtain an exact expression of this number.
-
-\begin{Coqsrc}
-Lemma L_omega_cube_3_eq:
-   let N := exp2 95 in
-   let P := (N * 97 - 1)%nat in
-   L_omega_cube 3  =  (exp2 P * (P + 2) - 1)%nat.
-\end{Coqsrc}
-
-This number is quite big. Using \texttt{Ocaml}'s \texttt{float} arithmetic,
-we can [under-]approximate it by $2^{3.8\times10^{30}}\times 3.8\times{10^{30}}$.
-
-\begin{Coqsrc}
-# let exp2 x = 2.0 ** x;;
-
-val exp2 : float -> float = <fun>
-#   exp2 95.0 *. 97.0 -. 1.0;;
-- : float = 3.84256588194182037e+30
-# let n = exp2 95.0 ;;
-# let p = n *. 97.0 -. 1.0;;
-val p : float = 3.84256588194182037e+30
-
-Estimation :
-2 ** (3.84 e+30) * 3.84 e+30.
-\end{Coqsrc}
-
-\subsection{Using \texttt{Equations}}
-\label{sect:L-equations}
-
-Note that we did not define any function $L_\alpha$ \emph{for any $\alpha<\epsilon_0$} yet.
-We have got no more than a collection of proved realizations of $\texttt{L\_spec}\;\alpha$ for several values of $\alpha$.
-
-\index{Coq!Plug-ins!Equations}
-
-Using the \texttt{coq-equations} plug-in by 
-M. Sozeau~\cite{sozeau:hal-01671777}, we will now define a function \texttt{L\_} which maps any ordinal  $\alpha<\epsilon_0$ to a proven realization of 
-$\texttt{L\_spec}\;\alpha$.   
-\subsection{Definition}
-
-In order to get a total function, we use our type \texttt{E0} of well-formed ordinal terms,(see Sect~\vref{sect:E0-def}).  Our definition is structured along a well-founded recursion and
-a case-study (zero, limit and successor ordinals).
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#L_}{L\_alpha}).}
-
-\label{Functions:L-alpha}
-
-\begin{Coqsrc}
-From Equations Require Import Equations.
-Require Import ArithRing Lia.
-
-Instance Olt : WellFounded Lt := Lt_wf.
-Hint Resolve Olt : E0.
-
-(** Using Coq-Equations for building a function which satisfies 
-    Large_sets.L_spec *)
-
-Equations  L_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
-  L_ alpha  i with E0_eq_dec alpha Zero :=
-    { | left _ =>  i ;
-      | right nonzero
-          with Utils.dec (Limitb alpha) :=
-          { | left _ =>  L_ (Canon alpha i)  (S i) ;
-            | right notlimit =>  L_ (Pred alpha) (S i)}}.
-
-Solve All Obligations with auto with E0.
-\end{Coqsrc}
-
-It is worth looking at the answer from \texttt{Equations} and check (with \texttt{About} ) all the lemmas this plug-in gives you for free. We show here only a part of \coq's anwer.
-
-\begin{Coqanswer}
-L__obligations_obligation_1 is defined
-L__obligations_obligation_2 is defined
-L__obligations is defined
-L__clause_1 is defined
-L__functional is defined
-L_ is defined
-...
-L__equation_1 is defined
-L__graph_mut is defined
-L__graph_rect is recursively defined
-L__graph_correct is defined
-L__elim has type-checked, generating 1 obligation
-L__elim is defined
-FunctionalElimination_L_ is defined
-FunctionalInduction_L_ is defined
-\end{Coqanswer}
-
-Sometimes, these automatically generated statements may look cryptic.
-
-\begin{Coqsrc}
-About L__equation_1.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-L__equation_1 :
-forall (alpha : E0) (i : nat),
-L_ alpha i = L__unfold_clause_1 alpha (E0_eq_dec alpha Zero) i
-\end{Coqanswer}
-
-In most cases, it may be useful to write human-readable  paraphrases of these statements.
-
-\begin{Coqsrc}
-Lemma L_zero_eqn : forall i, L_ Zero i = i.
-Proof. intro i; now rewrite L__equation_1. Qed.
-
-Lemma L_lim_eqn alpha i : Limitb alpha -> L_ alpha i =
-                                        L_ (Canon alpha i) (S i).
-
-Lemma L_succ_eqn alpha i :  L_ (Succ alpha) i = L_  alpha (S i).
-
-Hint Rewrite L_zero_eqn L_succ_eqn : L_rw.
-\end{Coqsrc}
-
-Using these three lemmas as rewrite rules, we can prove more properties of the functions \texttt{L\_$\alpha$}.
-
-\begin{Coqsrc}
-Lemma L_finite : forall i k :nat,  L_ i k = (i+k)%nat.  
-(* Proof by induction on i, using L_zero_eqn and L_succ_eqn *)
-
-Lemma L_omega : forall k, L_ omega%e0 k = S (2 * k)%nat.
-(* Proof using L_finite and L_lim_eqn *)
-\end{Coqsrc}
-
-By  well-founded induction on $\alpha$, we prove the following lemmas:
-
-\begin{Coqsrc}
-Lemma L_ge_S alpha : alpha <> Zero -> S <<= L_ alpha.
-
-Theorem L_correct alpha : L_spec (cnf alpha) (L_ alpha).
-\end{Coqsrc}
-
-Please note that the proof of \texttt{L\_correct} applies the lemmas proven in Sections~\ref{sect:L-spec-zero}, ~\ref{sect:L-spec-succ} and ~\ref{sect:L-spec-lim}.
-Our previous study of \texttt{L\_spec} allowed us to pave the way for the definition by \texttt{Equations} and the correctness proof.
-
-\subsubsection{Back to Hydra Battles}
-
-Theorem \texttt{battle\_length\_std } of
-Module~\href{../theories/html/hydras.Hydra.Hydra_Theorems.html\#battle_length_std}{Hydra.Hydra\_Theorems} relates the length of standard battles with the functions $L_\alpha$.
-
-\begin{Coqsrc}
-Theorem battle_length_std (alpha : E0)  :
-  alpha <> Zero ->
-  forall k, (1 <= k)%nat ->
-            battle_length standard k (iota (cnf alpha))
-                         (L_ alpha (S k) - k).
-\end{Coqsrc}
-
-\index{Projects}
-\begin{project}
-Instead of considering standard paths and battles, consider ``constant'' paths and the corresponding battles.
-Please use \texttt{Equations} in order to define the function that computes the length of the $k$-path which leads  from $\alpha$ to $0$.
-Prove a few  exact formulas and minoration lemmas.
-\end{project}
-
-\section{The Wainer-Hardy Hierarchy (Functions \texorpdfstring{$H_\alpha$}{\texttt{H\_alpha}})}
-
-\label{sect:hardy}
-In order to give an idea of the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the so called \emph{Wainer-Hardy hierarchy} of fast growing functions,
-presented for instance in~\cite{Promel2013}. 
-\index{Maths!Hardy Hierarchy}
-
-For each ordinal $\alpha$ below $\epsilon_0$, $H_\alpha$ is a 
-total arithmetic function, defined  by  transfinite recursion on $\alpha$, according to three cases:
-
-\index{Maths!Transfinite induction}
-
-\begin{itemize}
-\item If $\alpha=0$, then $H_\alpha (k)= k$ for any natural number $k$.
-\item If $\alpha=\textrm{succ}(\beta)$, then 
-$H_\alpha(k)=H_\beta(k+1)$ for any $k \in \mathbb{N}$
-\item If $\alpha$ is a limit ordinal, then 
-$H_\alpha(k) = H_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
-\end{itemize}
-
-\subsection{Hardy Functions in \texttt{Coq}}
-
-We define a function \texttt{H\_} of type \texttt{E0 -> nat -> nat} by transfinite induction over the type \texttt{E0} of the well formed ordinals below $\epsilon_0$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.H_alpha.html\#H_}{Epsilon0.H\_alpha}}
-
-\index{Coq!Plug-ins!Equations}
-\label{Functions:H-alpha}
-
-\begin{Coqsrc}
-Equations H_ (alpha: E0) (i:nat) :  nat  by wf  alpha lt :=
-  H_ alpha  i with E0_eq_dec alpha Zero :=
-    { | left _ =>  i ;
-      | right nonzero
-          with Utils.dec (Limitb alpha) :=
-          { | left _ =>  H_ (Canon alpha (S i))  i ;
-            | right notlimit =>  H_ (Pred alpha) (S i)}}.
-
-Solve All Obligations with auto with E0.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma H_eq1 : forall i, H_ Zero i = i.
-Proof.   intro i; now rewrite H__equation_1.  Qed.
-
-Lemma H_eq2 alpha i : Is_Succ alpha ->
-                      H_ alpha i = H_ (Pred alpha) (S i).
-
-Lemma H_eq3 alpha i : Limitb alpha ->
-                      H_ alpha i =  H_ (Canon alpha (S i)) i.
-
-Lemma H_eq4  alpha i :  H_ (Succ alpha) i = H_ alpha (S i).
-\end{Coqsrc}
-
-\subsection{First  Steps of the Hardy hierarchy}
-Using rewrite rules from \texttt{H\_eq1} to \texttt{H\_eq4}, we can explore the functions $H_\alpha$ for some small values of $\alpha$.
-
-\subsubsection{Finite Ordinals}
-
-By induction on $i$, we prove a simple expression of \texttt{H\_ (Fin i)}, where 
-\texttt{Fin $i$}  is the $i$-th finite ordinal.
-
-\begin{Coqsrc}
-Lemma H_Fin : forall i k: nat,  H_ (Fin i) k = (i+k)%nat.
-Proof with eauto with E0.
-  induction i.
-  - intros; simpl OF; simpl; autorewrite with H_rw E0_rw ... 
-  - intros ;simpl; autorewrite with H_rw E0_rw ... 
-    rewrite IHi; lia. 
-Qed.
-\end{Coqsrc}
-
-\subsubsection{Multiples of \texorpdfstring{$\omega$}{omega}}
-
-Since the canonical sequence of $\omega$ is composed of finite ordinals, 
-it is easy to get the formula associated with $H_\omega$.
-
-\begin{Coqsrc}
-Lemma H_omega : forall k, H_ Omega k = S (2 * k)%nat.
-Proof with auto with E0.
-  intro k; rewrite H_eq3 ...
-  - replace (Canon omega (S k)) with (Fin (S k)).
-    + rewrite H_Fin; lia.
-    +  now autorewrite with E0_rw.
-Qed.
-\end{Coqsrc}
-
-Before going further, we prove a useful rewriting lemma:
-
-\begin{Coqsrc}
-Lemma H_Plus_Fin alpha : forall i k : nat,
-    H_ (alpha + i)%e0 k = H_ alpha (i + k)%nat.
-(* Proof by induction on i *)
-\end{Coqsrc}
-
-Then, we get easily formulas for $H_{\omega+i}$, and $H_{\omega\times i}$ for any natural number $i$.
-
-\begin{Coqsrc}
-Lemma H_omega_double k : H_ (omega * 2)%e0 k =  (4 * k + 3)%nat.
-Proof.
- rewrite H_lim_eqn; simpl Canon.
- - ochange  (CanonS  (omega * 2)%e0 k)  (omega + (S k))%e0.
-  + rewrite H_Plus_Fin, H_omega;  lia.
-  -  now compute.
-Qed.
-
-Lemma H_omega_3 k : H_ (omega * 3)%e0 k = (8 * k + 7)%nat.
-
-Lemma H_omega_4 k : H_ (omega * 4)%e0 k = (16 * k + 15)%nat.
-
-Lemma H_omega_i i  : forall k,
-    H_ (omega * i)%e0 k = (exp2 i * k + Nat.pred (exp2 i))%nat.
-\end{Coqsrc}
-
-Crossing a new limit, we prove the following equality: 
-$$H_{\omega^2} (k) = 2 ^ {k+1} \times (k+1) - 1$$.
-
-\begin{Coqsrc}
-
-Lemma H_omega_sqr : forall k,
-    H_ (Phi0  2)%e0 k = (exp2 (S k ) *  S k - 1)%nat.
-Proof.
-  intro k; 
-   rewrite H_lim_eqn; auto with E0.
-  - ochange (Canon (Phi0 2) (S k)) (omega * (S k))%e0.
-    +  rewrite H_omega_i; simpl (exp2 (S k)).
-       *  rewrite Nat.add_pred_r.
-          -- lia. 
-          --   generalize (exp2_not_zero k);  lia.
-    + cbn; f_equal; lia.
-Qed.
-\end{Coqsrc}
-
-\subsubsection{New Limits}
-
-Our next step would be to prove an exact formula for $H_{\omega^\omega}(k)$.
-Since the canonical sequence of $\omega^\omega$ is composed of all the
-$\omega^i$, we first need to express $H_{\omega^i}$ for any natural number $i$.
-
-Let $i$ and $k$ be two natural numbers. 
-The ordinal $\canonseq{\omega^(i+1)}{k}$ is the product
-$\omega^i \times k$, so we need also to consider ordinals of this form.
-
-\begin{enumerate}
-\item First,  we express $H_{\omega^\alpha \times (i+2)}$ in terms of
-$H_{\omega^\alpha \times (i+1)}$.
-
-\begin{Coqsrc}
-Lemma H_Omega_term_1 : alpha <> Zero -> forall  k,  
-    H_ (Omega_term alpha (S i)) k =
-    H_ (Omega_term alpha i) (H_ (Phi0 alpha) k).
-\end{Coqsrc}
-
-\item
-Then, we prove by induction on $i$ that $H_{\omega^\alpha \times (i+1)}$ is just the
-$(i+1)$-th iterate of $H_{\omega^\alpha}$.
-
-\begin{Coqsrc}
-Lemma H_Omega_term (alpha : E0)  :
-alpha <> Zero -> forall i k, 
-  H_ (Omega_term alpha i) k = iterate  (H_ (Phi0 alpha)) (S i) k.
-\end{Coqsrc}
-
-\item In particular, we have got a formula for $H_{\omega^{i+1}}$.
-
-\begin{Coqsrc}
-Definition H_succ_fun f k := iterate f (S k) k.
-
-Lemma H_Phi0_succ alpha  : alpha <> Zero -> forall k,
-      H_ (Phi0 (Succ alpha)) k = H_succ_fun (H_ (Phi0 alpha)) k.
-
-Lemma H_Phi0_Si : forall i k,
-      H_ (Phi0 (S i)) k = iterate H_succ_fun i (H_ omega) k. 
-\end{Coqsrc}
-
-\end{enumerate}
-We get now a  formula for $H_{\omega^3}$:
-
-\begin{Coqsrc}
-Lemma H_omega_cube : forall k,
-    H_ (Phi0 3)%e0 k = iterate (H_ (Phi0 2))  (S k) k.
-Proof.
-  intro k; rewrite <-FinS_eq, -> Fin_Succ, H_Phi0_succ; auto.
-  compute; injection 1; discriminate.
-Qed.
-\end{Coqsrc}
-
-\subsubsection{A Numerical Example}
-
-It seems hard to capture the complexity of this function by looking only at this
-``exact'' formula. 
-Let us consider a simple example, the number $H_{\omega^3}(3)$.
-
-\begin{Coqsrc}
-Section H_omega_cube_3.
-
-Let f k :=   (exp2 (S k ) * (S k) - 1)%nat.
-
-Remark R0 k :  H_ (Phi0 3)%e0 k = iterate f (S k) k.
-\end{Coqsrc}
-
-Thus, the number $H_{\omega^3}(3)$ can be written as four nested applications of $f$.
-
-\begin{Coqsrc}
-Fact F0 : H_ (Phi0 3) 3 = f (f (f (f 3))).
- rewrite R0; reflexivity. 
-Qed.
-\end{Coqsrc}
-
-In order to make this statement more readable, we can introduce a local définition.
-
-\begin{Coqsrc}
-Let N := (exp2 64 * 64 - 1)%nat.
-\end{Coqsrc}
-
-This number looks quite big; let us compute an approximation in \texttt{Ocaml}:
-
-\begin{Coqsrc}
-# (2.0 ** 64.0 *. 64.0 -. 1.0);; 
-\end{Coqsrc}
-
-\begin{Coqanswer}
-- : float = 1.1805916207174113e+21
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Fact F1 : H_ (Phi0 3) 3 = f (f N).
-Proof.
- rewrite H_omega_cube_0; reflexivity. 
-Qed.
-
-Lemma F1_simpl : H_ (Phi0 3) 3 =
-                 (exp2 (exp2 (S N) * S N) * (exp2 (S N) * S N) - 1)%nat.
-
-\end{Coqsrc}
-
-In a more classical writing, this number is displayed as follows:
-
-{\Large
-$$
-H_{\omega^3}(3) =  2 ^ {(2 ^ {N + 1} \times (N+1) } )  \times  (2 ^ {N+1} \times ( N +1) ) - 1
-$$
-}
-
-We leave as an exercise to determine the best approximation as possible of
- the size of this number (for instance its number of digits).  For instance, if
-we do not take into account the multiplications in the formula above,
-we obtain that, in base $2$, the number $H_{\omega^3}(3)$ has at least
-$2^{10^{21}}$  digits. But it is still an under-approximation !
-
-\begin{Coqsrc}
-End H_omega_cube_3.
-\end{Coqsrc}
-
-Now, we have got at last an exact formula for $H_{\omega^\omega}$.
-
-\begin{Coqsrc}
-Lemma H_Phi0_omega : forall k, H_ (Phi0 omega) k =
-                               iterate H_succ_fun  k (H_ omega) k.
-Proof with auto with E0.
-  intro k; rewrite H_lim_eqn, <- H_Phi0_Si ...
-  -  rewrite CanonS_Canon, CanonS_Phi0_lim;  f_equal ...
-Qed.
-\end{Coqsrc}
-
-Using extensionality of the functional \texttt{iterate}, we can get a closed formula.
-
-\begin{Coqsrc}
-Lemma H_Phi0_omega_closed_formula k :
-  H_ (Phi0 omega) k =
-  iterate (fun (f: nat -> nat) (l : nat) => iterate  f (S l) l)
-               k
-               (fun k : nat => S (2 * k)%nat)
-               k.
-\end{Coqsrc}
-
-Note that this short formula contains two occurences of the functional \texttt{iterate}, the outer one is in fact a second-order iteration (on type \texttt{nat -> nat)}
-and the inner one  first-order (on type \texttt{nat}).
-
-\subsection{Abstract Properties of H-functions}
-~\label{sect:H-alpha-prop}
-
-Since pure computation seems to be useless for dealing with expressions of the form $H_\alpha(k)$, even for small values of $\alpha$ and $k$, we need to prove theorems for comparing $H_\alpha(k)$ and $H_\beta(l)$, in terms of comparison
-between $\alpha$ and $\beta$ on the one hand, $k$ and $l$ on the other hand.
-
-But beware of non-theorems! For instance, one could believe that $H$ is monotonous in its first argument. The following proof shows this is false.
-
-\begin{Coqsrc}
-Remark H_non_mono1 :
-  ~ (forall alpha beta k, (alpha o<= beta)%e0 ->
-                          (H_ alpha k <= H_ beta k)%nat).
-Proof.
- intros H ;specialize (H 42 omega 3).
- assert (H0 :(42 o<= omega)%e0) by (repeat split; auto).  
- apply H in H0; rewrite H_Fin, H_omega  in H0; lia.
-Qed.
-\end{Coqsrc}
-
-On the contrary, the fonctions of the Hardy hierarchy have the following five properties~\cite{KS81}: for any $\alpha < \epsilon_0$,
-\begin{itemize}
-\item the function $H_\alpha$ is strictly monotonous :
-      For all $n,p \in\mathbb{N}, n < p \Rightarrow H_\alpha(n)< H_\alpha(p)$.
-\item If $\alpha \not= 0$, then for every $n$, $n<H_\alpha(n)$.
-\item The function $H_\alpha$ is pointwise less or equal than $H_{\alpha+1}$
-
-\item For any $n\geq 1$, $H_\alpha(n)<H_{\alpha+1}(n)$.
-\emph{We say that $H_{\alpha+1}$ dominates $H_\alpha$ from $1$}.
-\item For any $n$ and $\beta$, if $\alpha \xrightarrow[n]{} \beta$, then
-$H_\beta(n)\leq H_\alpha(n)$.
-\end{itemize}
-
-\index{Maths!Transfinite induction}
-
-In \coq{}, we follow the  proof in~\cite{KS81}. This proof is mainly a single  proof by transfinite induction on $\alpha$ of the conjonction of the five properties.
-For each $\alpha$, the three cases : $\alpha=0$, $\alpha$ is a limit, and 
-$\alpha$ is a successor are considered. Inside each case, the five sub-properties are proved sequentially.
-
-\begin{definition}
- Let $f$ zand $g$ be two arithmetic  functions; $f$ is said to \emph{dominate} $g$ if $f(p)>g(p)$ for any all sufficiently large $p$.
-\end{definition}
-
-\begin{Coqsrc}
-Section Proof_of_Abstract_Properties.
-  Record P (alpha:E0) : Prop :=
-    mkP {
-        PA : strict_mono (H_ alpha);
-        PB : alpha <> Zero -> forall n,  (n < H_ alpha n)%nat;
-        PC : H_ alpha <<= H_ (Succ alpha);
-        PD : dominates_from 1 (H_ (Succ alpha)) (H_ alpha);
-        PE : forall beta n, Canon_plus n alpha beta -> 
-                            (H_ beta n <= H_ alpha n)%nat}.
-
-Theorem P_alpha : forall alpha, P alpha.
-  Proof.
-    intro alpha; apply well_founded_induction with lt.
-   (* rest of proof skipped *)
-
-Section Proof_of_Abstract_Properties.
-\end{Coqsrc}
-
-\subsection{Comparison between \texttt{L\_} and \texttt{H\_} }
-
-By well-founded induction on $\alpha$, we prove that our $L$ hierachy is ``almost'' the Hardy hierarchy (up to a small shift).
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#H_L_}{Epsilon0.L\_alpha}}
-
-\begin{Coqsrc}
- Theorem H_L_ alpha: forall i:nat,  (H_ alpha i <= L_ alpha (S i))%nat.
-\end{Coqsrc}
-
-\subsubsection{Back to Hydras}
-
-The following theorem relates the length of (standard) battles with the Hardy hierarchy.
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0.L\_alpha}}
-
-\begin{Coqsrc}
-Theorem battle_length_std_Hardy (alpha : E0) :
-  alpha <> Zero ->
-  forall k , 1 <= k -> exists l: nat,  
-       H_ alpha k - k <= l /\
-       battle_length standard k (iota (cnf alpha)) l.    
-\end{Coqsrc}
-
-\section{The Wainer Hierarchy (Functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
-\label{sect:wainer}
-
-\index{Maths!Wainer Hierarchy}
-
-The Wainer hierarchy~\cite{BW85, Wainer1970, KS81}, is also a family of fast growing functions, indexed by ordinals below $\epsilon_0$, by the following equations:
-
-\label{F_equations}
-\begin{itemize}
-\item $F_0(i)=i+1$
-\item $F_{\beta+1}(i)= (F_\beta)^{(i+1)}(i)$, where $f^{(i)}$ is the $i$-th iterate of $f$.
-\item $F_\alpha(i) = F_{\canonseq{\alpha}{i}} (i)$ if $\alpha$ is a limit ordinal.
-\end{itemize}
-
-A first attempt is to write a definition of $F_\alpha$ by equations, in the same as for $H\_alpha$.  We use the functional \texttt{iterate} defined in 
-Module~\href{../theories/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
-
-\begin{Coqsrc}
-Fixpoint iterate {A:Type}(f : A -> A) (n: nat)(x:A) :=
-  match n with
-  | 0 => x
-  | S p => f (iterate  f p x)
-  end.
-\end{Coqsrc}
-
-The following code comes from 
- \url{../theories/html/hydras.Epsilon0.F_alpha.html}.
-
-\index{Coq!Plug-ins!Equations}
-\begin{Coqsrc}
-Fail Equations F_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
-  F_ alpha  i with E0_eq_dec alpha Zero :=
-    { | left _ =>  i ;
-      | right nonzero
-          with Utils.dec (Limitb alpha) :=
-          { | left _ =>  F_ (Canon alpha i)  i ;
-            | right notlimit =>  iterate (F_ (Pred alpha))  (S i) i}}.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-The command has indeed failed with message:
-In environment
-alpha : E0
-notlimit : Limitb alpha = false
-nonzero : alpha <> Zero
-i : nat
-F_ : forall x : E0, nat -> x o< alpha -> nat
-The term "F_ (Pred alpha) ?x" has type "Pred alpha o< alpha -> nat"
-while it is expected to have type "Pred alpha o< alpha -> Pred alpha o< alpha"
-(cannot unify "nat" and "Pred alpha o< alpha").
-\end{Coqanswer}
-
-We presume that this error comes from the recursive call of \texttt{F\_} inside
-an application of \texttt{iterate}. The workaround we propose is to define first 
-the iteration of \texttt{F\_}  as an helper $F^*$, then to define the function $F$ as a ``iterating $F^*$ once''.
-
-\texttt{Equations} accepts the following definition, relying on  lexicographic ordering on pairs $(\alpha,n)$.
-
-\label{sect:F-equations}
-
-\index{Coq!Plug-ins!Equations}
-\label{Functions:F-alpha}
-\index{Maths! Fast growing functions}
-
-\begin{Coqsrc}
-Definition call_lt (c c' : E0 * nat) :=
-  lexico Lt (Peano.lt) c c'.
-
-Lemma call_lt_wf : well_founded call_lt.
-  unfold call_lt; apply Inverse_Image.wf_inverse_image,  wf_lexico.
-  -  apply E0.Lt_wf.
-  -  unfold Peano.lt; apply Nat.lt_wf_0. 
-Qed.
-
-Instance WF : WellFounded call_lt := call_lt_wf.
-
-Equations  F_star (c: E0 * nat) (i:nat) :  nat by wf  c call_lt :=
-    F_star (alpha, 0) i := i;
-    F_star (alpha, 1) i
-      with E0_eq_dec alpha Zero :=
-           { | left _ => S i ;
-             | right nonzero
-                 with Utils.dec (Limitb alpha) :=
-                 { | left _ => F_star (Canon alpha i,1) i ;
-                   | right notlimit =>
-                     F_star (Pred alpha, S i)  i}};
-    F_star (alpha,(S (S n))) i :=
-               F_star (alpha, 1) (F_star (alpha, (S n)) i).
-
-(* Finally, F_ alpha is defined as its first iterate ! *)
-
-Definition F_  alpha i := F_star (alpha, 1) i.
-\end{Coqsrc}
-
-It is quite easy to prove that our function \texttt{F\_} satisfies the equations on page~\pageref{sect:F-equations}.
-
-\begin{Coqsrc}
-Lemma F_zero_eqn : forall i, F_ Zero i = S i.
-
-Lemma F_lim_eqn : forall alpha i,  Limitb alpha ->
-                               F_ alpha i = F_ (Canon alpha i) i.
-
-Lemma F_succ_eqn : forall alpha i,
-    F_ (Succ alpha) i = iterate (F_ alpha) (S i) i.
-\end{Coqsrc}
-
-As for the Hardy functions, we can use these equalities as rewrite rules for
-``computing'' some values of $F_\alpha(i)$, for small values of $\alpha$.
-
-\begin{Coqsrc}
-Lemma LF1 : forall n,  F_ 1 n = S (2 * n).
-
-Lemma LF2 : forall i, (exp2 i * i < F_ 2 i)%nat.
-\end{Coqsrc}
-
-Like in Sect~\ref{sect:H-alpha-prop}, we prove by induction the following properties (see~\cite{KS81}).
-
-\begin{Coqsrc}
-Theorem F_alpha_mono alpha : strict_mono (F_ alpha).
-
-Theorem F_alpha_ge_S alpha : forall n, (n < F_ alpha n)%nat.
-
-Theorem F_alpha_Succ_le alpha : F_ alpha <<= F_ (Succ alpha).
-
-Theorem F_alpha_dom alpha : dominates_from 1 (F_ (Succ alpha)) (F_ alpha).
-
-Theorem F_alpha_beta alpha : forall beta n, Canon_plus n alpha beta -> 
-                                        (F_ beta n <= F_ alpha n)%nat.
-\end{Coqsrc}
-
-As a corollary, we prove the following proposition, p. 284 of~\cite{KS81}.
-
-\begin{quote}
-  If $\beta<\alpha$, $F_\alpha$ dominates $F_\beta$.
-\end{quote}
-
-\begin{Coqsrc}
-Lemma Propp284 : forall alpha beta : E0, 
-   beta o< alpha -> dominates (F_ alpha) (F_ beta).
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
-Let us quote a theorem from ~\cite{KS81} (page 297).
-
-\begin{quote}
-\begin{align*}
-  H_{\omega^\alpha}(n+1) &\geq F_{\alpha}(n) \quad (n\geq 1, \alpha<\epsilon_0) \\
- F_{\alpha}(n+1) &\geq H_{\omega^\alpha}(n) \quad (n\geq 1, \alpha<\epsilon_0) 
-\end{align*}
-Thus $H_{\omega^\alpha}$ and $F_{\alpha}$ have essentially the same order of growth.
-
-\end{quote}
-
- But, before trying to prove these facts, look at the definition of function $H$ in Ketonen and Solovay's paper !
-Is it really the same as the definition we quote from Pr{\H o}mel's chapter~\cite{Promel2013},
-whereas \cite{KS81} define $H_\alpha(n)$ as ``the least integer $k$ such that $[n,k]$ is $\alpha$-large''. Thus, it may be useful to adapt the statement above.
-
-\end{exercise}
-
-\index{Exercises}
-
-\begin{exercise}
-Prove the following result~\cite{KS81}(p. 298).
-
-\begin{quote}
-   For $n\geq 2$ and $\alpha \geq 3$, $F_\alpha(n+1)\geq 2^{F_\alpha(n)}$.
-\end{quote}
-\end{exercise}
-
-%  \subsection{Gnawing ordinals}
-
-% \begin{definition}[After~\cite{KS81}]
-%   Let $S$ be a finite set of positive integers, and $\alpha$ be an ordinal strictly less than $\epsilon_0$.  Let us denote by $s=s_1,s_2, \dots, s_N$ the sequence of 
-% the elements of $S$, enumerated in strictly increasing order.
-
-% We consider the sequence of ordinals $\alpha_o=\alpha, \alpha_1=\canonseq{\alpha_0}{s_1},\dots,\alpha_{i+1}=\canonseq{\alpha_i}{s_{i+1}},\dots, 
-% \alpha_{N}=\canonseq{\alpha_{N-1}}{s_N}$.
-% We denote by $\gnaw{s}{\alpha}$ the last ordinal of the sequence, \emph{i.e.}  $\alpha_N$.
-% \end{definition}
-
-% The following function computes $\gnaw{s}{\alpha}$ by recursion on $s$.
-% \vspace{4pt}
-
-% For instance, let us consider the ordinal $\omega^\omega$, and try some sequence of integers.
-
-% \begin{Coqsrc}
-% Compute pp (gnaw (omega ^ omega) (1::nil)).  
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = omega%pT1 : ppT1
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (gnaw (omega ^ omega) (1::2::nil)).  
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% = P_fin 2
-%      : ppT1
-% \end{Coqanswer}
-
-% Likewise, we can verify that $\gnaw{\omega^\omega}{\langle 1,2,3 \rangle}=1$
-% and $\gnaw{\omega^\omega}{\langle 1,2,3,4 \rangle}=0$.
-
-% \begin{Coqsrc}
-% Example omega_omega_1_4 : gnaw (omega ^omega) (interval 1 4) = 0.
-% Proof. trivial. Qed.
-
-% Example omega_omega_1_3 : gnaw (omega ^omega) (interval 1 3) = 1.
-% Proof. trivial. Qed.
-% \end{Coqsrc}
-
-% \begin{remark}
-% \label{remark:gnaw-vs-battles}
-% Let us consider an hydra battle, where Hercules always chops off the rightmost head of the hydra. Let $\alpha<\epsilon_0$ be an ordinal number, and 
-% $s=\langle i_1<i_2<\dots<i_N\rangle$ be  any finite sequence of positive integers.
-% Then the battle initiated by the hydra $\iota(\alpha)$, with $s$ as the sequence of successive replication factors, leads to the hydra $\iota(\gnaw{s}{\alpha})$ as the final state.
-% \end{remark}
-
-% \subsection{Large sequences}
-
-% \begin{remark}
-% In their article~\cite{KS81}, Ketonen and Solovay use the appellation ``large set'' instead of ``large sequence'', but their definitions use an enumeration of the elements in increasing order.
-Thus, we shall use the term ``sequence'' when referring to our implementation in \coq{}, and ``set'' when referring to the statements of~\cite{KS81}.
-% \end{remark}
-
-% \begin{definition}[After\cite{KS81}]
-% The sequence $s$ is said to be \emph{$\alpha$-large} if $\gnaw{\alpha}{s}=0$.  
-% \end{definition}
-
-% \begin{Coqsrc}
-% Definition largeb (alpha : T1) (s: list nat) :=
-%   match gnaw alpha s with
-%     | zero => true
-%     | _ => false
-%   end.
-
-% Definition large (alpha : T1) (s : list nat) : Prop :=
-%   largeb alpha s.
-% \end{Coqsrc}
-
-% For instance, the sequence $\langle 1,2,3,4 \rangle$ is $\omega^\omega$-large but not
-% $\omega^{\omega+1}$-large, since $\gnaw{\omega^{\omega+1}}{\langle 1,2,3,4\rangle}$ is equal to $\omega \times 2 + 4$.
-
-% %\omega^\omega+\omega^2\times 3 + \omega\times 4+ 5$.
-
-% \begin{Coqsrc}
-% Compute largeb (omega^omega) (interval 1 4).  
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% = true : bool
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute largeb (omega^(omega+1)) (interval 1 4).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% = false : bool
-% \end{Coqanswer}
-
-% \begin{remark}
-% A sequence $s$ is $\alpha$-large if (still considering Hercules ``righmost-head'', tactic), it leads to Hercules' victory.
-% \end{remark}
-
-% \subsection{A little game}
-
-% Let $\alpha<\epsilon_0$ be an ordinal, and $i$ a positive integer. We want to guess the least natural number  $j$  such that the interval $[i,i+j]$ is $\alpha$-large.
-% Equivalently, we should have $\gnaw{\alpha}{[i, i+j-1]}=1$.
-
-% The following functions takes three arguments: an ordinal $\alpha$, and two positive natural numbers  $i$ and $j$ (we assume, but not verify that $i$ and $j$ are strictly positive).
-It returns one of the three possible answers:
-
-% \begin{itemize}
-% \item \texttt{Ok} if $j$ is the smallest integer such that the interval $[i,i+j]$ is $\alpha$-large
-% \item \texttt{Too\_far} if $[i,i+j]$ is  $\alpha$-large, but $j$ is not the smallest such positive integer
-% \item \texttt{(Remaining $\beta$)} if $j$ is too small, and gnawing $\alpha$ with
-%   $[i,i+j]$ is still equal to $\beta$, instead of $0$
-% \end{itemize}
-
-% \vspace{4pt}
-% \noindent
-% \emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets_Demo.html}{Ordinals.Epsilon0.Large\_Sets\_Demo}}
-
-% \begin{Coqsrc}
-% Inductive answer : Set := You_won | Too_far | Remaining (rest : ppT1).
-
-% Definition large_set_check alpha i j :=
-%   let beta := gnaw alpha (interval i (Nat.pred j))
-%   in match beta with
-%      | one => Ok
-%      | zero => Too_far
-%      |  _ => Remaining (pp (canonseq j beta))
-%      end.
-% \end{Coqsrc}
-
-% \subsubsection{\texorpdfstring{$\omega$}{omega}-large intervals}
-
-% For instance, let us consider the ordinal $\omega$ and start with $i=1$.
-
-% \begin{Coqsrc}
-% Compute large_set_check omega 1 2.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%      = Ok
-%      : answer
-% \end{Coqanswer}
-
-% Let us give greater values of $i$, still with the ordinal $\omega$.
-
-% \begin{Coqsrc}
-% Compute large_set_check omega 2 3.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = Remaining 1
-%      : answer
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check omega 3 3.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = Remaining 3
-%      : answer
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check omega 2 4.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = Ok
-%      : answer
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check omega 3 6.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = Ok
-%      : answer
-% \end{Coqanswer}
-
-% It looks like every request to compute (\texttt{large\_set\_check omega $i$ $2\times i$})  will succeed. Let us try an example.
-
-% \begin{Coqsrc}
-% Compute large_set_check omega 49 98.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = Ok
-%      : answer
-% \end{Coqanswer}
-
-% \subsubsection{\texorpdfstring{$\omega^2$}{omega\^{}2}-large intervals}
-
-% Still using \texttt{Compute} and \texttt{large\_set\_check}, we obtained the 
-% following values of $j$, the least integer such that the interval $[i,j]$ is $\omega^2$ large.
-
-% $$
-% \begin{array}{|c|c|}
-% \hline
-%   i & j \\
-% \hline 
-% 1 & 4 \\
-% 2 & 14 \\
-% 3 & 38 \\
-% 4 & 94 \\
-% 5 & 222 \\
-% % 6 & 510 
-% \hline
-% \end{array}
-% $$
-
-% \begin{exercise}
-% Please give the 6-th and 7-th line of the array above.
-% \end{exercise}
-
-% \subsubsection{The limits of (pure) computation}
-
-% Let us now play with bigger ordinals, for instance $\alpha=\omega^\omega$ or
-% $\alpha=\omega^{\omega + 1}$. We notice that, even for small values of $i$, it is hard
-% to guess values of $j$ such that $[i,i+j]$ is $\alpha$-large.
-
-% \begin{Coqsrc}
-% Compute large_set_check (omega ^ omega) 1 4.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%  = Ok
-%      : answer
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check (omega ^ omega) 2 38.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%  = Ok
-%      : answer
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check (omega ^ omega) 3 1000.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = Remaining (omega ^ 2 * 2 + omega * 220 + 798)%pT1
-%      : answer
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check (omega ^ omega) 3 1798.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = Remaining (omega ^ 2 * 2 + omega * 220)%pT1
-%      : answer
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check (omega ^ omega) 3 5000.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = Remaining (omega ^ 2 * 2 + omega * 218 + 2198)%pT1
-%      : answer
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check (omega ^ (omega + 1)) 3 5000.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%  = Remaining
-%          (omega ^ omega * 2 + omega ^ 3 * 4 + 
-%           omega ^ 2 * 4 + omega * 1148 +  4222)%pT1
-%      : answer  
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check (omega ^ (omega + 1)) 3 10000.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% Warning: To avoid stack overflow, large numbers in nat are interpreted as
-% applications of Init.Nat.of_uint. [abstract-large-number,numbers]
-
-%     = Remaining
-%          (omega ^ omega * 2 + omega ^ 3 * 4 + 
-%           omega ^ 2 * 4 + omega * 1147 +
-%           8446)%pT1
-%      : answer
-% \end{Coqanswer}
-
-% Since computation is not enough, let us comme back to proofs, or, better, proofs \emph{and} computations.
-
-% \subsection{Proving largeness}
-
-% %%% ICI %%%
-
-% \subsection{$n$-large and $\omega$-large intervals }
-
-% A finite ordinal is just made out of \texttt{zero} and \texttt{succ} constructions.
-% Thus any sequence of strictly positive integers of length greater than  or equal to $n$ will completely gnaw the ordinal $n$.
-
-% Concerning the first limit ordinal $\omega$ the largeness of a sequence $s$ depends only on its first element and the length of the rest of the sequence (please keep in mind that the argument $i$ of $\canonseq{\alpha}{i}$ is meaningful only if $\alpha$ is a limit ordinal).
-% The following proposition is labeled $4.2$ in~\cite{KS81}.
-
-% \begin{proposition}
-%   For $n<\omega$, a set $X$id $n$-large if and only if $|X|\geq n$.
-% A finite set $X$ is $\omega$-large if $|X|>\min{X}$.
-% \end{proposition}
-
-% Rewritten in terms of list of strictly positive integers, we get the following statements:
-
-% \begin{Coqsrc}
-% Lemma large_n_iff : forall s (n:nat) , ~ In 0 s  ->
-%                                   large n s  <-> (n <= List.length s)%nat.
-% \end{Coqsrc}
-
-% \begin{Coqsrc}
-% Lemma large_omega_iff : forall s n,  ~In 0 (n::s) -> 
-%                                             large omega (n::s) <->
-%                                             (n <=  List.length s)%nat.
-% \end{Coqsrc}
-
-% % \begin{Coqsrc}
-% % (* sorted list of natural numbers greater than or equal to n *)
-
-% % Inductive sorted_ge (n: nat) : list nat -> Prop :=
-% % | sorted_ge_nil : sorted_ge n nil
-% % | sorted_ge_one : forall p, n<=p ->
-% %                             sorted_ge n (p::nil)
-% % | sorted_ge_cons: forall p q s,  n<=p -> p<q ->
-% %                                  sorted_ge p (q::s) ->
-% %                                  sorted_ge n (p::q::s).
-% % \end{Coqsrc}
-
-% \subsection{Mimimal large sequences}
-
-% Let us consider \emph{minimal} large sequences, \emph{i.e.} large sequences 
-% the strict prefix of which do not lead to $0$. In other words, only the last ordinal 
-% defined by the sequence is null.  For this purpose, we use the predicate \texttt{path\_to} introduced page~\pageref{path-to-definition}.
-
-% \begin{Coqsrc}
-% Definition mlarge alpha s := path_to zero s alpha.
-% \end{Coqsrc}
-
-%----------------------------------------------------------------------
-\chapter[Countable Ordinals (after Sch\"{u}tte)]{Kurt Schütte's Axiomatic Definition of Countable Ordinals}
-
-\label{chap:schutte} 
-%ON
-
-In the present chapter, we  compare our implementation of the segment $[0,\epsilon_0)$ with a mathematical text in order to ``validate'' our constructions.
-Our reference here is the axiomatic definition of the set of countable ordinals,
-in chapter V of Kurt Schütte's book `` Proof Theory ''~\cite{schutte}.
-
-\begin{remark}
-\emph{In all this chapter, the word ``ordinal'' will be considered as a synonymous of
-``countable ordinal''}  
-\end{remark}
-
-Schütte's definition of countable ordinals relies on the following three axioms:
-
-There  exists a strictly ordered set , such that
-\begin{enumerate}
-\item  $(\mathbb{O},<)$ is well-ordered
-\item Every bounded subset of $\mathbb{O}$  is countable
-\item Every countable subset of $\mathbb{O}$  is bounded.
-\end{enumerate}
-
-Starting with these three axioms, Schütte re-defines the vocabulary about ordinal numbers: the null ordinal $0$, limits and successors, the addition of ordinals, the infinite ordinals $\omega$, $\epsilon_0$, $\Gamma_0$, etc.
-
-This chapter describes an adaptation to \coq{} of Schütte's axiomatization. 
- Unlike the rest of our libraries, our library
-\href{../theories/html/hydras.Schutte.Schutte.html}{Ordinals.Schutte}
-is not constructive, and relies on several axioms.
-
-\begin{itemize}
-\item First, please keep in mind  that the set of countable ordinals is not countable.
-Thus, we cannot hope to represent all countable ordinals as finite terms of an inductive type, which was possible with  the set of ordinals strictly less than $\epsilon_0$ (resp.
-$\Gamma_0$)
-\item We tried to be as close as possible to K.
-Schütte's text, which uses ``classical'' mathematics : excluded middle, Hilbert's $\epsilon$ (choice) and Russel's $\iota$ (definite description) operators.
-Both operators allow us to write definitions close to the natural mathematical language, such as ``$\textrm{succ}$ is \emph{the} least ordinal strictly greater than $\alpha$''
-\item Please note that only the library \href{../theories/html/hydras.Schutte.Schutte.html}{Schutte/*.v} is ``contaminated'' by axioms, and that the rest of our libraries remain constructive.
-\end{itemize}
-
-\section{Declarations and Axioms}
-
-Let us declare a type 
-\texttt{Ord} for representing countable ordinals, and a binary relation
- \texttt{lt}. Note that, in our development, \texttt{Ord} is a type, while the \emph{set} of countable ordinals (called $\mathbb{O}$ by Schütte) 
-is the full set over the type \texttt{Ord}.
-
-\label{types:Ord}
-
-We use Florian Hatat's library on countable sets, written as he was a student of  \emph{\'Ecole Normale Supérieure de Lyon}.
-A set $A$ is countable if there is an injective function from $A$ to $\mathbb{N}$ (see
-Library \href{../theories/html/hydras.Schutte.Countable.html}%
-{\texttt{Schutte.Countable}}).
-
-\vspace{6pt}
-
-\emph{From Module\href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{Schutte.Schutte\_basics}}}
-
-\begin{Coqsrc}
-Parameter Ord : Type.
-Parameter lt : relation Ord.
-Infix "<" := lt : schutte_scope.
-
-Definition ordinal := Full_set Ord.
-\end{Coqsrc}
-
-Schütte's first axiom tells that \texttt{lt} is a well order on the set 
-\texttt{ordinal} (The  class \texttt{WO} is defined in
-Module~\href{../theories/html/hydras.Schutte.Well_Orders.html}%
-{\texttt{Well\_Orders.v}}).
-
-\label{types:WO}
-
-\begin{Coqsrc}
-Variables (M:Type)
-         (Lt : relation M).
-
-Class WO : Type:=
-    {
-      Lt_trans : Transitive  Lt;
-      Lt_irreflexive : forall a:M, ~ (Lt a a);
-      well_order : forall (X:Ensemble M)(a:M),
-          In X a ->
-          exists a0:M, least_member  X a0
-    }.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-  Axiom AX1 : WO lt.
-\end{Coqsrc}
-
-The second and third axioms say that a subset $X$ of $\mathbb{O}$ is
-(strictly) bounded if and only if it is countable.
-
-\begin{Coqsrc}
-Axiom AX2 : forall X: Ensemble Ord, 
-   (exists a,  (forall y, In X y -> y < a)) ->
-   countable X.
-
-Axiom AX3 : forall X : Ensemble Ord,
-              countable X -> 
-              exists a,  forall y, In X y -> y < a.
-\end{Coqsrc}
-
-\texttt{AX2} and \texttt{AX3} could have been replaced by a single axiom (using the \texttt{iff} connector), but we decide to respect as most as possible the structure of Schütte's definitions.
-
-\section{Additional  Axioms}
-
-The adaptation of Schütte's mathematical discourse to \coq{} led us to
-import a few axioms from the standard library. We encourage the reader to consult \coq{}'s FAQ about the safe use of axioms
- \url{https://github.com/coq/coq/wiki/The-Logic-of-Coq#axioms}.
-
-\subsubsection{Classical Logic}
-
-In order to work with classical logic, we import the module
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Classical.html}{Coq.Logic.Classical}  of \coq{}'s standard library, specifially the following axiom:
-
-\begin{Coqsrc}
- Axiom classic : forall P:Prop, P \/ ~P.
-\end{Coqsrc}
-
-\subsubsection{Description Operators}
-
-In order to respect Schütte's style, we imported also the library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Epsilon.html}{\texttt{Coq.Logic.Epsilon}}.  The rest of this section presents a few examples of
-how Hilbert's choice operator and Church's definite description allow us
- to write understandable definitions (close to the mathematical natural language).
-
-\subsubsection{The Definition of zero}
-
-According to the  definition of a well order, every non-empty subset of \texttt{Ord} has a least element. Furthermore, this least element is unique.
-
-\begin{Coqsrc}
-Remark R : exists! z : Ord, least_member lt  ordinal z.
-Proof.
-  destruct inh_Ord as [a]; apply (well_order (WO:=AX1)) with a .
-  split.
-Qed.
-\end{Coqsrc}
-
-Assume we want to call this element  \texttt{zero}.
-
-\begin{Coqsrc}
-Definition zero : Ord.
-Proof.
-  Fail destruct R.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-The command has indeed failed with message:
-Case analysis on sort Type is not allowed for inductive 
-definition ex.
-\end{Coqanswer}
-
-Indeed, the basic logic of  \coq{} does not allow us to eliminate a proof of a proposition 
-$\exists\,x:A,\,P(x)$ for building a term whose type lies in the sort \texttt{Type}. 
-The reasons for this impossibility are explained in many documents~\cite{BC04, chlipalacpdt2011, Coq}.
-
-Let us import the library \texttt{Coq.Logic.Epsilon}, which contains the following axiom and lemmas.
-
-\begin{Coqsrc}
-Axiom epsilon_statement:
-  forall (A : Type) (P : A->Prop), inhabited A ->
-    {x : A | (exists x, P x) -> P x}.
-\end{Coqsrc}
-
-Hilbert's $\epsilon$ \emph{operator} is derived from this  axiom.
-
-\begin{Coqsrc}
-  Definition epsilon (A : Type) (i:inhabited A) (P : A->Prop) : A
-  := proj1_sig (epsilon_statement P i).
-
-Lemma constructive_indefinite_description :
-  forall (A : Type) (P : A->Prop),
-    (exists x, P x) -> { x : A | P x }.
-\end{Coqsrc}
-
-If we consider the \emph{unique existential} quantifier $\exists!$, we obtain
-Church's \emph{definite description operator}.
-
-\begin{Coqsrc}
-Definition iota (A : Type) (i:inhabited A) (P : A->Prop) : A
-  := proj1_sig (iota_statement P i).
-\end{Coqsrc}
-
-\begin{Coqsrc}
- Lemma constructive_definite_description :
-  forall (A : Type) (P : A->Prop),
-    (exists! x, P x) -> { x : A | P x }.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Definition iota_spec (A : Type) (i:inhabited A) (P : A->Prop) :
-  (exists! x:A, P x) -> P (iota i P)
-  := proj2_sig (iota_statement P i).
-\end{Coqsrc}
-
-Indeed, the operators \texttt{epsilon} and \texttt{iota} allowed us to make our definitions 
-quite close to Schütte's text. Our libraries \href{../theories/html/hydras.Schutte.MoreEpsilonIota.html}%
-{\texttt{Schutte.MoreEpsilonIota}}
-and
-\href{../theories/html/hydras.Schutte.PartialFun.html}%
-{\texttt{Schutte.PartialFun}} are extensions of \texttt{Coq.logic.Epsilon} for making easier 
-such definitions. See also an article in french~\cite{PCiota}.
-
-\begin{Coqsrc}
-Class InH (A: Type) : Prop :=
-   InHWit : inhabited A.
-
-Definition some {A:Type} {H : InH A} (P: A -> Prop) := 
-   epsilon (@InHWit A H) P.
-
-Definition the {A:Type} {H : InH A} (P: A -> Prop) := 
-   iota (@InHWit A H) P.
-\end{Coqsrc}
-
-In order to use these tools,  we had to tell \coq{}  that the type \texttt{Ord} is not empty:
-
-\begin{Coqsrc}
-Axiom inh_Ord : inhabited Ord.
-\end{Coqsrc}
-
-We are now able to define \texttt{zero} as the least ordinal. For this purpose,
-we define a function returning the least element of any [non-empty]  subset.
-
-\begin{Coqsrc}
-Definition the_least {M: Type} {Lt}
-           {inh : InH M} {WO: WO Lt} (X: Ensemble M)  : M :=
-  the  (least_member  Lt X ).
-\end{Coqsrc}
-
-\vspace{4pt}
-
-From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{~Schutte.Schutte\_basics}}
-
-\label{Constants:zero:Ord}
-
-\begin{Coqsrc}
-Definition zero: Ord :=the_least ordinal.
-\end{Coqsrc}
-
-We want to prove now that zero is less than or equal to any ordinal number.
-
-\begin{Coqsrc}
-Lemma zero_le (alpha : Ord) :  zero <= alpha.
-Proof.
-  unfold zero, the_least, the; apply iota_ind.
-\end{Coqsrc}
-
-According to the use of the description operator \texttt{iota}, we have to solve  two trivial sub-goals.
-\begin{enumerate}
-\item Prove that there exists a unique least member of \texttt{Ord}
-\item Prove that being a least member of \texttt{Ord} entails the announced inequality 
-\end{enumerate}
-
-\begin{Coqanswer}
-2 subgoals (ID 155)
-
-  alpha : Ord
-  ============================
-  exists ! x : Ord, least_member lt ordinal x
-
-subgoal 2 (ID 156) is:
- forall a : Ord, unique (least_member lt ordinal) a -> 
-                a <= alpha
-\end{Coqanswer}
-
-\begin{Coqsrc}
-  -  apply the_least_unicity, Inh_ord.
-  -  destruct 1 as [[_ H1] _]; apply H1; split. 
-Qed.
-\end{Coqsrc}
-
-\subsubsection{Remarks on \texttt{epsilon} and \texttt{iota}}
-
- What would happen in case of a misuse of \texttt{epsilon} or \texttt{iota} ?
-For instance, one could give a unsatisfiable specification to \texttt{epsilon} or 
-a specification for \texttt{iota} that admits several realizations.
-
-Let us consider an example:
-
-\begin{Coqbad}
-Module Bad.
-
- Definition bottom := the_least (Empty_set Ord).
-\end{Coqbad}
-
-\begin{Coqanswer}
- bottom is defined
-\end{Coqanswer}
-
-Since we won't be able to prove the proposition
-\linebreak \Verb|{exists! a: Ord, least_member (Empty_set Ord) a|, the only properties we would be able to prove about \texttt{bottom} would be \emph{trivial} properties, 
-\emph{i.e.}, satisfied by \emph{any} element of type \texttt{Ord}, like for instance
-\texttt{bottom = bottom}, or \texttt{zero <= bottom}.
-
-\begin{Coqbad}
-Lemma le_zero_bottom : zero <= bottom. 
-Proof. apply zero_le. Qed.
-
-Lemma bottom_eq : bottom = bottom.
-Proof. trivial. Qed.
-
-Lemma le_bottom_zero : bottom <= zero.
-Proof.
-   unfold bottom, the_least, the; apply iota_ind.
-\end{Coqbad}
-
-\begin{Coqanswer}
-2 subgoals (ID 413)
-
-  ============================
-  exists ! x : Ord, least_member lt (Empty_set Ord) x
-
-subgoal 2 (ID 414) is:
- forall a : Ord, unique (least_member lt (Empty_set Ord)) a -> 
-      a <= zero
-\end{Coqanswer}
-
-\begin{Coqbad}
-Abort.
-End Bad.
-\end{Coqbad}
-
-In short, using \texttt{epsilon} and \texttt{iota} in our implementation of countable ordinals after Schütte has two main advantages.
-
-\begin{itemize}
-\item It allows us to give a \emph{name} (using \texttt{Definition}) two witnesses 
-of existential quantifiers (let us recall that, in classical logic, one may consider non-constructive proofs of existential statements)
-\item By separating definitions from proofs of [unique] existence, one may make definitions  more concise and readable. Look for instance at 
-the definitions of  \texttt{zero}, \texttt{succ}, \texttt{plus}, etc. in the rest of this chapter.
-\end{itemize}
-%%%% ICI ICI
-
-\section{The  Successor Function}
-
-The definition of the function \texttt{succ:Ord -> Ord} is very concise. The successor of any ordinal $\alpha$ is the smallest ordinal strictly greater than $\alpha$.
-
-\label{Functions:succ-sch}
-
-\begin{Coqsrc}
-Definition succ (alpha : Ord) := the_least (fun beta => alpha < beta).
-\end{Coqsrc}
-
-Using \texttt{succ}, we define the folloing predicates.
-
-\begin{Coqsrc}
-Definition is_succ (alpha:Ord) := exists beta, alpha = succ beta.
-
-Definition is_limit (alpha:Ord) := alpha <> zero /\ ~ is_succ alpha.
-\end{Coqsrc}
-
-% \begin{remark}
-% Please look at remark~\vref{warning:coercions}.
-% \end{remark}
-
-How do we prove properties of the successor function?
-First, we make its specification explicit.
-
-\begin{Coqsrc}
-Definition succ_spec (alpha:Ord) :=
-  least_member   lt (fun z => alpha < z).
-\end{Coqsrc}
-
-Then, we prove that our function \texttt{succ} meets this specification.
-
-\begin{Coqsrc}
-Lemma succ_ok : forall alpha,  succ_spec alpha  (succ alpha).
-Proof.
-  intros; unfold succ, the_least, the;  apply iota_spec.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-1 subgoal (ID 172)
-
-  alpha : Ord
-  ============================
-  exists ! x : Ord, succ_spec alpha x
-\end{Coqanswer}
-
-We have now to prove that the set of all ordinals strictly greater than $\alpha$ has a unique least element.
-But the singleton set $\{\alpha\}$ is countable, hence  bounded (by the axiom \texttt{AX3}).
-Hence; the set $\{\beta\in\mathbb{O}|\alpha < \beta\}$ is not empty
-and therefore has a unique least element.
-
-The \coq{} proof script is quite short.
-
-\begin{Coqsrc}
-  destruct (@AX3 (Singleton _ alpha)).
-  - apply countable_singleton.
-  -  unfold succ_spec; apply the_least_unicity;  exists x; intuition.
-Qed.     
-\end{Coqsrc}
-
-We can ``uncap'' the description operator for proving properties of the
-\texttt{succ} function.
-
-\begin{Coqsrc}
-Lemma lt_succ (alpha : Ord) :  alpha < succ alpha.
-Proof.
-  destruct  (succ_ok  alpha);  tauto.
-Qed.
-
-Hint Resolve lt_succ : schutte.
-
-Lemma lt_succ_le (alpha beta : Ord):
-  alpha < beta -> succ alpha <= beta.
-Proof with eauto with schutte.
-  intros  H;  pattern (succ alpha); apply the_least_ok ... 
-  exists (succ alpha); red;apply lt_succ ...
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma lt_succ_le_2 (alpha beta : Ord):
-  alpha < succ beta -> alpha <= beta.
-
-Lemma succ_mono (alpha beta : Ord):
-  alpha < beta -> succ alpha < succ beta.
-
-Lemma succ_monoR (alpha beta : Ord) :
- succ alpha < succ beta -> alpha < beta.
-
-Lemma lt_succ_lt (alpha beta : Ord) :
-  is_limit beta ->  alpha < beta -> succ alpha < beta.
-\end{Coqsrc}
-
-\section{Finite Ordinals}
-
-Using \texttt{succ}, it is now easy to define recursively all the finite ordinals.
-
-\label{sect:notation-F-sch}
-
-\begin{Coqsrc}
-Reserved Notation "'F' n" (at level 29) .
-
-Fixpoint finite (i:nat) : Ord :=
-  match i with 
-            | 0 => zero
-            | S i => succ (F i)
-  end
-where "'F' i" := (finite i)  : schutte_scope.
-
-Coercion finite : nat >-> Ord.
-\end{Coqsrc}
-
-\section{The Definition of \texttt{omega}}
-In order to define $\omega$, the first infinite ordinal, we use an operator which
-``returns'' the least upper bound (if it exists) of a subset $X\subseteq \mathbb{O}$.
-For that purpose, we first use a predicate:
-(\texttt{is\_lub $D$ \textit{lt} $X$ $a$}) if $a$ belongs to $D$ and is the least 
-upper bound  of $X$ (with respect to \textit{lt}).
-
-\begin{Coqsrc}
-Definition is_lub (M:Type)
-                  (D : Ensemble M)
-                  (lt : relation M)
-                  (X:Ensemble M)
-                  (a:M) :=
-   In _ D a  /\ upper_bound  D lt X a  /\
-   (forall y, In _ D y -> upper_bound  D lt X y  -> 
-                  y = a \/ lt a y).
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Definition sup_spec X lambda := is_lub ordinal lt X lambda.
-
-Definition sup (X: Ensemble Ord) : Ord  := the  (sup_spec X).
-
-Notation "'|_|' X" := (sup X) (at level 29) : schutte_scope.
-\end{Coqsrc}
-
-Then, we define the function \texttt{omega\_limit} which returns the least upper bound 
-of the  (denumerable) range of any sequence \texttt{s: nat -> Ord}. 
-By \texttt{AX3} this range is bounded, hence the set of its upper bounds is not empty and has a least element.
-
-\begin{Coqsrc}
-Definition omega_limit (s:nat->Ord) : Ord 
-  := |_| (seq_range s).
-\end{Coqsrc}
-
-Then we define \texttt{omega} as the limit of the sequence of finite ordinals.
-
-\label{sect:notation-omega}
-\begin{Coqsrc}
-Definition _omega := omega_limit finite.
-
-Notation "'omega'" := (_omega) : schutte_scope.
-\end{Coqsrc}
-
-Among the numerous properties of the ordinal $\omega$, les us quote the following ones
-(proved in Module 
-\href{../theories/html/hydras.Schutte.Schutte_basics.html\#finite_lt_omega}{\texttt{Schutte.Schutte\_basics}})
-
-\begin{Coqsrc}
-Lemma finite_lt_omega : forall i: nat,  i < omega.
-
-Lemma lt_omega_finite alpha : Ord) : 
-  alpha < omega ->  exists i:nat, alpha =  i.
-
-Lemma is_limit_omega : is_limit omega.
-\end{Coqsrc}
-
-\subsection{Ordering Functions and Ordinal Addition}
-
-After having defined the finite ordinals and the infinite ordinal $\omega$, we  define the sum $\alpha+\beta$ of two countable ordinals.
-Schütte's definition looks like the following one:
-
-\begin{quote}
-``$\alpha+\beta$ is the $\beta$-th ordinal greater than or equal to $\alpha$''
-\end{quote}
-
-The purpose of this section is to give a meaning to the construction
-``the $\alpha$-th element of $X$''  where $X$ is any non-empty subset of $\mathbb{O}$.
-We follow Schütte's approach, by defining the notion of \emph{ordering functions},
-a way to associate a unique ordinal to each element of $X$.
-Complete definitions and proofs can be found in Module
- \href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
-{\texttt{Schutte.Ordering\_Functions}} ).
-
-\subsection{Definitions}
-
-A \emph{segment} is a set $A$ of ordinals such that, whenever  $\alpha\in A$ and
-$\beta<\alpha$, then $\beta\in A$; a segment is  \emph{proper} if it strictly included in $\mathbb{O}$.
-
-\begin{Coqsrc}
- Definition segment (A: Ensemble Ord) :=
-  forall alpha beta, In A alpha -> beta < alpha -> In A  beta.
-
-Definition proper_segment (A: Ensemble Ord) :=
-  segment A /\ ~ Same_set A ordinal.
-\end{Coqsrc}
-
-Let  $A$ be a segment, and $B$ a subset of $\mathbb{O}$ : an \emph{ordering function for $A$ and  $B$} is a strictly increasing bijection from $A$ to $B$.
-The set $B$ is said to be an \emph{ordering segment} of $A$.
-Our definition in \coq{} is a direct translation of the mathematical text of~\cite{schutte}.
-
-\index{Maths!Ordering functions}
-
-\begin{Coqsrc}
-Definition ordering_function (f : Ord -> Ord)(A B : Ensemble Ord) :=
- segment A /\
- (forall a, In A a -> In B (f a)) /\
- (forall b, B b -> exists a, In A a /\ f a = b) /\
- forall a b, In A a -> In A b -> a < b ->  f a < f b.
-
-Definition ordering_segment (A B : Ensemble Ord) :=
-  exists f : Ord -> Ord, ordering_function f A B.
-\end{Coqsrc}
-
-We are now able to associate with any subset $B$ of $\mathbb{O}$ its ordering segment and ordering function.
-
-\begin{Coqsrc}
-Definition the_ordering_segment (B : Ensemble Ord) :=
-  the  (fun x => ordering_segment x B).
-
-Definition ord  (B : Ensemble Ord) := 
-  some (fun f => ordering_function f (the_ordering_segment B) B).
-\end{Coqsrc}
-
-Thus (\texttt{ord $B \;\alpha$}) is the $\alpha$-th element of $B$.
-Please note that the last definition uses the epsilon-based operator \texttt{some} and
-not \texttt{the}. This is due to the fact that we cannot prove the unicity (w.r.t. Leibniz' equality) of the ordering function of a given set. 
-By contrast, we admit the axiom  \texttt{Extensionality\_Ensembles}, from the library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Sets.Ensembles.html}{Coq.Sets.Ensembles}, so we use the operator \texttt{the} in the definition of
-\texttt{the\_ordering\_segment}.
-
-One of the main theorems of
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html\#ordering_function_ex}%
-{\texttt{Ordering\_Functions}} 
-associates a unique segment and a unique (up to extensionality) ordering function to every subset $B$ of $\mathbb{O}$.
-
-\begin{Coqsrc}
-About ordering_function_ex.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-forall B : Ensemble Ord,
- exists ! S : Ensemble Ord, 
-      exists f : Ord -> Ord, ordering_function f S B
-\end{Coqanswer}
-
-\begin{Coqanswer}
-ordering_function_unicity :
-forall (B S1 S2 : Ensemble Ord) (f1 f2 : Ord -> Ord),
-ordering_function f1 S B ->
-ordering_function f2 S2 B -> 
-fun_equiv f1 f2 S1 S2
-\end{Coqanswer}
-
-Thus,  our function \texttt{ord}  which enumerates the elements of $B$ is defined in a non-ambiguous way.
-Let us quote the following theorems (see Library
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
-{\texttt{Schutte.Ordering\_Functions}} for more details).
-
-\begin{Coqsrc}
-Theorem ordering_le : forall f A B,
-    ordering_function f A B ->
-    forall alpha, In A alpha -> alpha <= f alpha.
-
-Th_13_5_2 :
-forall (A B : Ensemble Ord) (f : Ord -> Ord),
-ordering_function f A B -> Closed B -> continuous f A B
-\end{Coqsrc}
-
-\subsection{Ordinal Addition}
-
-We are now ready to define and study addition on the type \texttt{Ord}.
-The following definitions and proofs can be consulted in Module
-\href{../theories/html/hydras.Schutte.Addition.html}%
-{\texttt{Schutte.Addition.v}}.
-
-\begin{Coqsrc}
-Definition plus alpha := ord  (ge alpha).
-Notation "alpha + beta " := (plus alpha beta) : schutte_scope.
-\end{Coqsrc}
-
-In other words,  $\alpha + \beta$ is the  $\beta$-th ordinal greater than or equal to $\alpha$. 
-Thanks to generic properties of ordering functions, we can show the following 
-properties of addition on $\mathbb{O}$. First, we prove a useful lemma:
-
-\begin{Coqsrc}
-Lemma plus_elim (alpha : Ord) :
-  forall P : (Ord->Ord)->Prop,
-    (forall f: Ord->Ord, 
-        ordering_function f ordinal (ge alpha)-> P f) ->
-    P (plus alpha).
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma alpha_plus_zero (alpha: Ord): alpha + zero = alpha.
-Proof.
- pattern  (plus alpha); apply plus_elim; eauto.
- \end{Coqsrc}
-
- \begin{Coqanswer}
- 1 subgoal (ID 24)
-
-  alpha : Ord
-  ============================
-  forall f : Ord -> Ord,
-  ordering_function f ordinal (ge alpha) -> 
-  f zero = alpha
- \end{Coqanswer}
-
- \begin{Coqsrc}
- (* rest of proof skipped *)
- \end{Coqsrc}
-
-The following lemmas are proved the same way.
-
- \begin{Coqsrc}
-
-Lemma zero_plus_alpha (alpha : Ord) : zero + alpha = alpha.
-
-Lemma le_plus_l (alpha beta : Ord) : alpha <= alpha + beta.
-
-Lemma le_plus_r (alpha beta : Ord) :  beta <= alpha + beta.
-
-Lemma plus_mono_r (alpha beta gamma : Ord) : 
-    beta < gamma -> alpha + beta < alpha + gamma.
-
-Lemma plus_of_succ (alpha beta : Ord) :
-    alpha + (succ beta) = succ (alpha + beta).
-
-Theorem plus_assoc (alpha beta gamma : Ord) :
-  alpha + (beta + gamma) = (alpha + beta) + gamma.
-
-Lemma one_plus_omega :  1 + omega = omega.
-
-Lemma finite_plus_infinite (n : nat) (alpha : Ord) :
-  omega <= alpha -> n + alpha = alpha.
-\end{Coqsrc}
-
-It isinteresting to compare the proof of these lemmas with the 
-computational proofs of the corresponding statements in Module
-\href{../theories/html/hydras.Epsilon0.T1.html}%
-{\texttt{Epsilon0.T1}}. 
-For instance, the proof of the lemma 
-\texttt{one\_plus\_omega} uses the continuity of ordering functions (applied to  \texttt{(plus 1)}) and compares the limit of the $\omega$-sequences $i_{(i \in \mathbb{N})}$ and
-$(1+i)i_{(i \in \mathbb{N})}$, whereas in the library  \texttt{Epsilon0/T1}, the equality 
-$1+\omega=\omega$ is just proved with \texttt{reflexivity}!
-
-\subsubsection{Multiplication by a Natural Number}
-
-The multiplication of an ordinal by a natural number is defined in terms of addition.
-This operation is useful for the study of Cantor normal forms.
-
-\begin{Coqsrc}
-Fixpoint mult_Sn (alpha:Ord)(n:nat){struct n}: Ord :=
- match n with 
-            | 0 => alpha
-            | S p => mult_Sn  alpha p + alpha
- end.
-
-Definition mult_fin_r alpha n :=
-  match n with
-      0 => zero
-    | S p => mult_Sn alpha p
-  end.
-
-Notation "alpha * n" := (mult_fin_r alpha n) : schutte_scope.
-\end{Coqsrc}
-
-\section{The Exponential of Basis \texorpdfstring{$\omega$}{omega}}
-
-In this section, we define the function which maps any $\alpha\in\mathbb{O}$ to
-the ordinal  $\omega^\alpha$, also written 
-$\phi_0(\alpha)$. 
-It is an opportunity to apply the definitions and results of the preceding section. 
-Indeed,  Schütte first defines a subset of $\mathbb{O}$: the set of additive principal ordinals, and $\phi_0$  is just defined as the ordering function of this set.
-
-\subsection{Additive Principal Ordinals}
-
-\index{Maths!Additive principal ordinals}
-
-\begin{definition}
-A non-zero ordinal  $\alpha$ is said to be \emph{additive principal} if, for all  $\beta<\alpha$, $\beta+\alpha$ is equal to  $\alpha$.
-We call \texttt{AP} the set of additive principal ordinals.
-
-\end{definition}
-
-\noindent\emph{From Module \href{../theories/html/hydras.Schutte.AP.html}%
-{\texttt{Schutte.AP}}}
-
-\begin{Coqsrc}
-Definition AP : Ensemble Ord :=
-  fun alpha => 
-  zero < alpha /\
-  (forall beta, beta < alpha ->  beta + alpha = alpha).
-\end{Coqsrc}
-
-\subsection{The Function \texttt{phi0}}
-
-Let us call  $\phi_0$ the ordering function of \texttt{AP}.
-In the mathematical text, we shall use indifferently the notations  $\omega^\alpha$ and$\phi_0(\alpha)$.
-
-\begin{Coqsrc}
-Definition phi0 := ord AP.
-
-Notation "'omega^'" := phi0 (only parsing) : schutte_scope.
-\end{Coqsrc}
-
-\subsection{Omega-towers and the Ordinal \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-Using $\phi_0$, we can define recursively the set of finite omega-towers.
-
-\begin{Coqsrc}
-Fixpoint omega_tower (i : nat) : Ord :=
-  match i with
-    0 =>  1
-  | S j => phi0 (omega_tower j)
-  end.
-\end{Coqsrc}
-
-\label{sect:epsilon0-as-limit}
-Then, the ordinal  $\epsilon_0$ is defined as the limit of the sequence of all finite towers (a kind of infinite tower).
-
-\begin{Coqsrc}
-Definition epsilon0 := omega_limit omega_tower.
-\end{Coqsrc}
-
-The rest of our library \texttt{AP} is devoted to the proof of properties of additive principal ordinals, hence of the ordering function  $\phi0$ and the ordinal $\epsilon_0$ (which we could not express within the type \texttt{T1}).
-
-\subsection{Properties of the Set  \texttt{AP}}
-
-The set of additive principal ordinals is not empty: it contains at least the ordinals  $1$ and  $\omega$.
-
-\begin{Coqsrc}
-Lemma AP_one : In AP 1.
-
-Lemma AP_omega : In AP omega.
-\end{Coqsrc}
-
-Moreover, $1$ is the least principal ordinal and $\omega$ is the second element of
-\texttt{AP}.
-
-\begin{Coqsrc}
-Lemma least_AP: least_member  lt AP 1.
-
-Lemma omega_second_AP :
-  least_member   lt 
-                  (fun alpha => 1 < alpha /\ In AP alpha)
-                  omega.
-\end{Coqsrc}
-
-The set  \texttt{AP} is  \emph{closed} under addition, and unbounded.
-
-\begin{Coqsrc}
-Lemma AP_plus_closed (alpha beta gamma : Ord): 
-     In AP alpha -> beta < alpha -> gamma < alpha ->
-     beta + gamma < alpha.
-
-Theorem AP_unbounded : Unbounded AP.
-\end{Coqsrc}
-
-Finally, \texttt{AP} is (topologically) \emph{closed} and ordered by the segment of all countable ordinals.
-
-\index{Predicates!Closed}
-
-\begin{Coqsrc} 
-Definition Closed (B : Ensemble Ord) : Prop := 
-  forall M, Included M B -> Inhabited _ M -> 
-                 countable M -> In B (|_| M).
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Theorem AP_closed : Closed AP.
-
-Lemma AP_o_segment :  the_ordering_segment AP = ordinal.
-\end{Coqsrc}
-
-\subsubsection{Properties of the Function \texorpdfstring{$\phi_0$}{phi0}}
-
-The ordering function $\phi_0$ of the set \texttt{AP} is defined on the full set $\mathbb{O}$ and is continuous (Schütte calls such a  function  \emph{normal}).
-
-\begin{Coqsrc}
-Theorem normal_phi0 : normal phi0 AP.
-\end{Coqsrc}
-
-The following properties come from  the definition of $\phi_0$ as the ordering function of \texttt{AP}.
-It may be interesting to compare these proofs with the computational ones described in Chapter ~\ref{chap:T1}.
-
-\begin{Coqsrc}
-Lemma AP_phi0 (alpha : Ord) : In AP (phi0 alpha).
-
-Lemma phi0_zero : phi0 zero =  1.
-
-Lemma phi0_mono (alpha beta : Ord) :
-  alpha < beta ->  phi0 alpha < phi0 beta.
-
-Lemma phi0_inj (alpha beta : Ord) :
-    phi0 alpha = phi0 beta -> alpha = beta.
-
-Lemma phi0_sup : forall (U: Ensemble Ord),
-   Inhabited _ U ->   countable U ->  phi0 (|_| U) = |_| (image U phi0).
-
-Lemma is_limit_phi0 (alpha : Ord) :
-  zero < alpha ->  is_limit (phi0 alpha).
-
-Lemma omega_eq : omega = phi0 1.
-
-Lemma phi0_le (alpha : Ord) : alpha <= phi0 alpha.
-\end{Coqsrc}
-
-Please note that the lemma \texttt{omega\_eq} above, is consistent with the interpretation of the ordering function $\phi_0$ as the exponential of basis $\omega$.
-Indeed we could have written this lemma with our alternative notation:
-
-\begin{Coqsrc}
- Lemma omega_eq : omega = omega^ 1.
-\end{Coqsrc}
-
-\section{More about \texorpdfstring{$\epsilon_0$}{\texttt{epsilon0}}}
-
-Let us recall that the limit ordinal  $\epsilon_0$ cannot be written within the type \texttt{T1}.
-Since we are now considering the set of all countable ordinals, we can now prove some properties of this ordinal.
-
-We prove the inequality  $\alpha<\omega^\alpha$ whenever $\alpha < \epsilon_0$.
-\emph{Note that this condition was implicit in Module \href{../theories/html/hydras.Epsilon0/T1.html\#lt_phi0}%
-{\texttt{Epsilon0.T1}}.}
-
-\begin{Coqsrc}
-Lemma lt_phi0 (alpha : Ord):
-  alpha < epsilon0 -> alpha < phi0 alpha.
-\end{Coqsrc}
-
-The proof is as follows:
-\begin{enumerate}
-\item Since $\alpha<\epsilon_0$, consider the least $i$ such that $\alpha$ is strictly less than the omega-tower of height $i$.
-\item
-  \begin{itemize}
-  \item If $i=0$, then the result is trivial (because $\alpha=0$)
- \item  Otherwise let $i=j+1$; 
-          $\alpha$ is greater than or equal to the omega-tower of height $j$.
-         By monotonicity,  $\phi_0(\alpha)$ is greater than or equal to 
-        the omega-tower of height $j+1$, thus strictly greater than $\alpha$
-  \end{itemize}
- \end{enumerate}
-
-Moreover,  $\epsilon_0$ is the least ordinal $\alpha$ that verifies the equality 
-$\alpha = \omega^\alpha$, in other words the least fixpoint of the function  $\phi_0$.
-
-\begin{Coqsrc}
-Theorem epsilon0_lfp : least_fixpoint lt phi0 epsilon0.
-\end{Coqsrc}
-
-\section{Critical Ordinals}
-
-\index{Maths!Critical ordinals}
-
-For any  (countable) ordinal $\alpha$, the set $\textit{Cr}(\alpha)$ is inductively defined 
-as follows by Schütte (p.81 of~\cite{schutte}).
-
-\begin{quote}
-  \begin{itemize}
-  \item $\textit{Cr}(0)$ is the set \textit{AP} of additive principal ordinals.
-  \item If $0<\alpha$, then $\textit{Cr}(\alpha)$ is the intersection of all the sets of fixpoints of the $\textit{Cr}(\beta)$ for $\beta<\alpha$.
-  \end{itemize}
-\end{quote}
-
-This definition is translated in \coq{} in 
-Module \href{../theories/html/hydras.Schutte.Critical.html}%
-{\texttt{Schutte.Critical}}, as the least fixpoint of a functional.
-
-\begin{Coqsrc}
-Definition Cr_fun : forall alpha : Ord,
-       (forall beta : Ord, beta < alpha -> Ensemble Ord) ->
-        Ensemble Ord 
-:= 
-   fun (alpha :Ord)
-        (Cr : forall beta, 
-                beta < alpha -> Ensemble Ord) 
-        (x : Ord) => (
-       (alpha = zero /\ AP x) \/
-       (zero < alpha /\
-        forall beta (H:beta < alpha),
-          the_ordering_segment (Cr beta H) x /\ ord (Cr  beta H) x = x)).
-
-Definition Cr (alpha : Ord) : Ensemble Ord := 
-    (Fix  all_ord_acc (fun (_:Ord) => Ensemble Ord) Cr_fun) alpha.
-\end{Coqsrc}
-
-\label{sect:phi-schutte}
-
-\begin{Coqsrc}
-Definition phi (alpha : Ord) : Ord -> Ord 
-    :=  ord (Cr alpha).
-
-Definition A (alpha : Ord) : Ensemble Ord :=
-  the_ordering_segment (Cr alpha).
-\end{Coqsrc}
-
-For instance,  we prove that $\textit{Cr}(0)$ is the set of additive principals and that $\epsilon_0$
-belongs to $\textit{Cr}(1)$.
-
-\begin{Coqsrc}
-Lemma Cr_zero_AP :  Cr 0 = AP
-
-Lemma epsilon0_Cr1 : In (Cr 1) epsilon0.
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
- Prove that $\epsilon_0$ is the least element of $\textit{Cr}(1)$.
-\end{exercise}
-
-\subsection{A flavor of Infinity}
-
-The family of the $\textit{Cr}(\alpha)$s is made of infinitely many unbounded (hence infinite) sets.
-Let us quote Lemma 5, p. 82  of~\cite{schutte}:
-\begin{quote}
-  For all $\alpha$, the set $\textit{Cr}(\alpha)$ is closed (for the least upper bound of non-empty countable sets) and unbounded.
-\end{quote}
-
-We prove this result by transfinite induction on $\alpha$ of both properties.
-
-The proof is still quite long, by transfinite induction over $\alpha$.
-
-\index{Coq!Techniques!Transfinite induction}
-
-\begin{Coqsrc}
-Section Proof_of_Lemma5.
-  Let P (alpha:Ord) := Unbounded (Cr alpha) /\ Closed (Cr alpha).
-
- Lemma Lemma5 : forall alpha, P alpha.
-(* ... *)
- End Proof_of_Lemma5.
-
-Corollary Unbounded_Cr alpha : Unbounded (Cr alpha).
-Proof.
-  now destruct (Lemma5 alpha).
-Qed.
-
-Corollary Closed_Cr alpha : Closed (Cr alpha).
-Proof.
-  now destruct (Lemma5 alpha).
-Qed.
-\end{Coqsrc}
-
-\section{Cantor Normal Form}
-
-The notion of Cantor normal form is defined for all countable ordinals.
-Nevertheless, note that, contrary to the implementation based on type \texttt{T1},
-the Cantor normal form of an ordinal $\alpha$ may contain $\alpha$ as a 
-sub-term\footnote{This would prevent us from trying to represent Cantor normal forms as finite trees (like in Sect.~\ref{sec:T1-inductive-def})}.
-
-We represent  Cantor normal forms as lists of ordinals.
-A  list $l$ is a Cantor normal form of a given ordinal $\alpha$ if it satisfies two conditions:
-
-\begin{itemize}
-\item The list  $l$ is sorted (in decreasing order) w.r.t. the order $\leq$
-\item The sum of all the  $\omega^{\beta_i}$ where the $\beta_i$ are the terms of $l$ (in this order) is equal to $\alpha$.
-\end{itemize}
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Schutte.CNF.html\#cnf_t}%
-{\texttt{Schutte.CNF}}}
-
-\begin{Coqsrc}
- Definition cnf_t := list Ord.
-
-Fixpoint eval (l : cnf_t) : Ord :=
-  match l with nil => zero
-              | beta :: l' => phi0 beta + eval l'
-  end.
-
-Definition sorted (l: cnf_t) :=
-  LocallySorted (fun alpha beta => beta <= alpha) l.
-
-Definition is_cnf_of (alpha : Ord)(l : cnf_t) : Prop :=
-  sorted l /\ alpha = eval l.
-\end{Coqsrc}
-
-\index{Maths!Transfinite induction}
-
-By transfinite induction on $\alpha$, we prove that every countable ordinal $\alpha$ 
- has at least a Cantor normal form.
-
-\begin{Coqsrc}
-Theorem cnf_exists (alpha : Ord) :
-  exists l: cnf_t, is_cnf_of alpha l.
-\end{Coqsrc}
-
-By structural induction on lists, we prove that this normal form is unique.
-
-\begin{Coqsrc}
- Lemma cnf_unicity : forall l alpha, 
-   is_cnf_of alpha l -> 
-   forall l',  is_cnf_of alpha l' -> l=l'.
-Proof.
- induction l.
- (*  ...  *)
-
-Theorem cnf_exists_unique (alpha:Ord) :
-  exists! l: cnf_t, is_cnf_of alpha l.
-\end{Coqsrc}
-
-Finally, the following two lemmas relate  $\epsilon_0$ with Cantor normal forms.
-
-If $\alpha<\epsilon_0$, then the Cantor normal form of $\alpha$ is made of ordinals strictly less than $\alpha$.
-
-\begin{Coqsrc}
-Lemma cnf_lt_epsilon0 : 
- forall l alpha, 
-   is_cnf_of alpha l ->  alpha < epsilon0 ->
-   Forall (fun beta =>  beta < alpha) l.
-\end{Coqsrc}
-
-\index{Exercises}
-
-\begin{exercise}
-Please consider the following statement :
-
-\begin{Coqsrc}
-Lemma cnf_lt_epsilon0_iff : 
- forall l alpha, 
-   is_cnf_of alpha l ->  
-   (alpha < epsilon0 <->  Forall (fun beta =>  beta < alpha) l).
-\end{Coqsrc}
-
-Is it true ?
-\end{exercise}
-
-Finally, the Cantor normal form of $\epsilon_0$ is just $\omega^{\epsilon_0}$.
-
-\begin{Coqsrc}
-Lemma cnf_of_epsilon0 : is_cnf_of epsilon0 (epsilon0 :: nil).
-Proof.
-  split.
-  - constructor.  
-  - simpl;  now rewrite alpha_plus_zero, epsilon0_fxp.
-Qed.
-\end{Coqsrc}
-
-\index{Projects}
-
-\begin{project}
-Implement pages 82 to 85 of~\cite{schutte} (critical, strongly critical, maximal critical ordinals, Feferman's ordinal $\Gamma_0$).
-\end{project}
-
-\begin{remark}
-The sub-directory \href{../theories/html/hydras.Gamma0.html}%
-{\texttt{theories/Gamma0}} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form. 
-\end{remark}
-
-\section{An Embedding of \texttt{T1} into \texttt{Ord}}
-
-Our library 
-\href{../theories/html/hydras.Schutte.correctness_E0.html}%
-{\texttt{Schutte.Correctness\_E0}} establishes the link between two very different modelizations of ordinal numbers. In other words, it ``validates'' a data structure in terms of
-a classical mathematical discourse considered as a model. 
-First, we define a function from \texttt{T1} into  \texttt{Ord} by structural recursion.
-
-\begin{Coqsrc}
-Fixpoint inject (t:T1) : Ord :=
- match t with 
-     | T1.zero => zero
-     | T1.ocons a n b =>  AP.phi0 (inject a) * S n + inject b
- end.  
-\end{Coqsrc}
-
-This function enjoys good commutation properties with respect to the main operations which
-allow us to build Cantor normal form.
-
-\begin{Coqsrc}
-Theorem inject_of_zero : inject T1.zero = zero.
-
-Theorem inject_of_finite (n : nat):
-  inject (T1.fin n) =  n.
-
-Theorem inject_of_phi0 (alpha : T1):
-  inject (phi0 alpha) = AP.phi0 (inject alpha).
-
-Theorem inject_plus (alpha beta : T1): nf alpha -> nf beta ->
-  inject (alpha + beta)%t1 = inject alpha + inject beta.
-
-Theorem inject_mult_fin_r (alpha : T1)  :
-  nf alpha -> forall n:nat , inject (alpha *  n)%t1 =  inject alpha * n.
-
-Theorem inject_mono (beta gamma : T1) :
-  T1.lt  beta gamma -> 
-  T1.nf beta -> T1.nf gamma -> 
-  inject beta < inject gamma.
-
-Theorem inject_injective (beta gamma : T1) : nf beta -> nf gamma ->
-  inject beta = inject gamma -> beta = gamma.
-\end{Coqsrc}
-
-Finally, we prove that \texttt{inject} is a bijection from the set of all terms of \texttt{T1} in normal form to the set 
-\texttt{members epsilon0} of the elements of \texttt{Ord} strictly less than  $\epsilon_0$.
-
-\begin{Coqsrc}
-Theorem inject_lt_epsilon0 (alpha : T1):
-      inject alpha < epsilon0.
-
-Theorem embedding : 
-     fun_bijection (nf: Ensemble T1)  (members epsilon0) inject.
- \end{Coqsrc}
-
-\subsection{Remarks}
-Let us recall that the library \href{../theories/html/hydras.Schutte.Schutte.html}%
-{\texttt{Schutte}} depends on five \emph{axioms} and lies explicitly in the  
-framework of classical logic with a weak version of the axiom of choice
-(please look at the documentation of
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.ChoiceFacts.html}{\texttt{Coq.Logic.ChoiceFacts}}).
-Nevertheless, the other modules:
-\href{../theories/html/hydras.Epsilon0.Epsilon0.html}%
-{\texttt{Epsilon0}},
-\href{../theories/html/hydras.Hydra.Hydra.html}%
-{\texttt{Hydra}}, et 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}%
-{\texttt{Gamma0}}
-do not import any axioms and are really constructive.
-
-\index{Projects}
-\begin{project}
-There is no construction of ordinal multiplication in~\cite{schutte}. 
-It would be interesting to derive this operation from Schütte's axioms,
-and prove its consistence with multiplication in ordinal notations for 
-$\epsilon_0$ and $\Gamma_0$.
-\end{project}
-
-\section{Related Work}
-
-In~\cite{grimm:hal-00911710}, José Grimm establishes the consistency between our ordinal notations (\texttt{T1} and \texttt{T2} (Veblen normal form) and his implementation
-of ordinal numbers after Bourbaki's set theory.
-
-\chapter{The Ordinal \texorpdfstring{$\Gamma_0$}{Gamma0} (first draft)}
-
-\emph{This chapter and the files it presents are still very incomplete, considering the impressive properties of $\Gamma_0$~\cite{Gallier91}.
-We hope to add new material soon, and accept contributions!}
-
-\section{Introduction}
-We present a notation system for the ordinal $\Gamma_0$, following Chapter V, Section 14 of~\cite{schutte}: ``A notation system for the ordinals $<\Gamma_0$''.
-We try to be as close as possible to Schütte's text and usual practices of \coq{} developments.
-
-The ordinal $\Gamma_0$ is defined in Section 13 of ~\cite{schutte} as the least \emph{strongly critical ordinal}. It is widely known as the \emph{Feferman-Schütte ordinal}.
-
-Section V, 13 of~\cite{schutte} defines \emph{strongly critical} and
-\emph{maximal $\alpha$-critical} ordinals:
-
-\begin{itemize}
-\item $\alpha$ is strongly critical if
-$\alpha$ is $\alpha$-critical,
-\item $\gamma$ is maximal $\alpha$-critical if $\gamma$ is $\alpha$-critical, and, for all $\xi>\alpha$, $\gamma$ is not $\xi$-critical.
-
-\end{itemize}
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Schutte.Critical.html\#strongly_critical}%
-{\texttt{Schutte.Critical}}}
-
-\begin{Coqsrc}
-Definition strongly_critical alpha := In (Cr alpha) alpha.
-
-Definition maximal_critical alpha : Ensemble Ord :=
-  fun gamma =>
-    In (Cr alpha) gamma /\
-    forall xi, alpha < xi -> ~ In (Cr xi) gamma.
-
-Definition Gamma0 := the_least strongly_critical.
-\end{Coqsrc}
-
-\index{Projects}
-\begin{project}
-Prove that a (countable)  ordinal $\alpha$ is strongly critical iff 
-$\phi_\alpha(0)=\alpha$ (Theorem 13.13 of~\cite{schutte} ). 
-\end{project}
-
-\index{Projects}
-\begin{project}
-Prove that the set of strongly critical ordinals is unbounded and closed (Theorem 13.14 of~\cite{schutte} ).
-Thus this set is not empty,  hence has a least element.
-Otherwise, the definition of $\Gamma_0$ above would be useless.
-\end{project}
-
-In the present version of this development, we  only study $\Gamma_0$ as a notation system, much more powerful than the ordinal notation for $\epsilon_0$.
-
-%\index{Projects}
-%
-%\begin{project}
-% Schûtte's section 13 of~\cite{schutte} contain several definitions and lemmas which give another view on $\Gamma_0$.  We leave it as a project to implement them in \coq{}.  
-%\end{project}
-
-\section{The Type \texttt{T2} of Ordinal Terms}
-
-The notation system for ordinals less than $\gamma_0$ comes from the following theorem of~\cite{schutte}, where $\psi\,\alpha$ is the ordering function 
-of the set of maximal $\alpha$-critical ordinals.
-
-\begin{quote}
-  Any ordinal $\not= 0$ which is not strongly critical can be expressed in terms of $+$ and $\psi$.
-\end{quote}
-
-\index{Projects}
-\begin{project}
-This theorem is not formally proved in this development yet. It should be!
-\end{project}
-
-Like in Chapter~\ref{chap:T1}, we define an inductive type with two constructors, one for $0$, the other for the construction $\psi(\alpha,\beta)\times(n+1)+\gamma$, adapting a Manolios-Vroon-like notation~\cite{Manolios2005} to
-\emph{Veblen normal forms}.
-\label{types:T2}
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T2}%
-{\texttt{Gamma0.T2}}}
-\begin{Coqsrc}
-(**  [gcons alpha beta n gamma] is : [psi(alpha,beta)*(S n)+ gamma]  *)
-
-Inductive T2 : Set :=
-| zero : T2
-| gcons : T2 -> T2  -> nat -> T2 -> T2.
-
-Notation "[ alpha , beta ]" := (gcons alpha beta 0 zero)
-                                 (at level 0): t2_scope.
-\end{Coqsrc}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=11cm]{epsilon0.jpg}
-  \caption{Veblen normal form}
-  \label{fig:gamma0}
-\end{figure}
-
-Like in chapter~\ref{chap:T1}, we get familiar with the type \texttt{T2} by recognising simple constructs like finite ordinals, $\omega$, etc., as inhabitants of \texttt{T2}.
-
-\begin{Coqsrc}
-Notation  "'one'"  := [zero,zero] : T2_scope.
-
-(** The (n+1)-th finite ordinal *)
-Notation "'FS' n" := (gcons zero zero n zero) (at level 10) : T2_scope.
-
-(** the [n]-th ordinal  *)
-Definition fin (n:nat) := match n with 0 => zero | S p => FS p end.
-
-Notation "'omega'"  := [zero,one] : T2_scope.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Notation "'epsilon0'"  := ([one,zero]) : T2_scope.
-
-Definition epsilon alpha := [one, alpha].
-\end{Coqsrc}
-
-\section{How Big is \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}?}
-
-Let us define a strict order on type \texttt{T2}. The following definition is 
-an adaptation of Schütte's, taking into account the multiplications by a natural number (inspired by~\cite{Manolios2005}, and also present in \texttt{T1}).
-
-\label{sect:t2-lt-def}
-
-\begin{Coqsrc}
-Inductive lt : T2 -> T2 -> Prop :=
-| (* 1 *) 
- lt_1 : forall alpha beta n gamma,  zero t2< gcons alpha beta n gamma
-| (* 2 *)
- lt_2 : forall alpha1 alpha2 beta1 beta2 n1 n2 gamma1 gamma2, 
-                alpha1 t2< alpha2 ->
-                beta1 t2< gcons alpha2 beta2 0 zero ->
-               gcons alpha1 beta1 n1 gamma1 t2<
-               gcons alpha2 beta2 n2 gamma2
-| (* 3 *)
- lt_3 : forall alpha1  beta1 beta2 n1 n2 gamma1 gamma2, 
-               beta1 t2< beta2 ->
-               gcons alpha1 beta1 n1 gamma1 t2<
-               gcons alpha1 beta2 n2 gamma2
-
-| (* 4 *)
- lt_4 : forall alpha1 alpha2 beta1 beta2 n1 n2 gamma1 gamma2, 
-               alpha2 t2< alpha1 ->
-               [alpha1, beta1] t2< beta2 ->
-               gcons alpha1 beta1 n1 gamma1 t2<
-               gcons alpha2 beta2 n2 gamma2
-
-| (* 5 *)
-lt_5 : forall alpha1 alpha2 beta1 n1 n2 gamma1 gamma2, 
-               alpha2 t2< alpha1 ->
-               gcons alpha1 beta1 n1 gamma1 t2<
-               gcons alpha2  [alpha1, beta1] n2 gamma2
-
-| (* 6 *)
-lt_6 : forall alpha1 beta1  n1  n2 gamma1 gamma2,  (n1 < n2)%nat ->
-                                    gcons alpha1 beta1 n1 gamma1 t2< 
-                                    gcons alpha1 beta1 n2 gamma2
-
-| (* 7 *)
-  lt_7 : forall alpha1 beta1 n1   gamma1 gamma2,  gamma1 t2< gamma2 ->
-                                      gcons alpha1 beta1 n1 gamma1 t2<
-                                      gcons alpha1 beta1 n1 gamma2
-where  "o1 t2< o2" := (lt o1 o2): T2_scope.
-
-Hint Constructors lt : T2.
-\end{Coqsrc}
-
-Seven constructors! In order to get accustomed with this definition, let us look at a small set of examples, covering all the constructors of \texttt{lt}.
-
-\subsection{Examples}
-
-\subsubsection*{Proof of $0<\epsilon_0$}
-
-\begin{Coqsrc}
-Example Ex1: 0 t2< epsilon0.
-Proof.  constructor 1. Qed.
-\end{Coqsrc}
-
-\subsubsection*{Proof of $\omega<\epsilon_0$}
-
-\begin{Coqsrc}
-Example Ex2: omega t2< epsilon0.
-Proof. info_auto with T2. (* uses lt_1 and lt_2 *) Qed.
-\end{Coqsrc}
-
-\subsubsection*{Proof of $\psi(\omega,8)\times 13+56 < \psi(\omega,8)\times 13+57 $}
-
-\begin{Coqsrc}
-Example Ex3: gcons omega 8 12 56 t2<  gcons omega 8 12 57.
-Proof.
-  constructor 7; constructor 6; auto with arith.
-Qed.
-\end{Coqsrc}
-
-\subsubsection*{Proof of $\epsilon_0<\psi(2,1)$}
-
-\begin{Coqsrc}
-Example Ex4: epsilon0 t2< [2,1].
-Proof.
-   constructor 2; auto with T2.
-   - constructor 6; auto with arith.
-Qed.
-\end{Coqsrc}
-
-\subsubsection*{Proof of $\psi(2,1)<\psi(2,3)$}
-
-\begin{Coqsrc}
-Example Ex5 : [2,1] t2< [2,3].
-Proof.
-  constructor 3; auto with T2.
-  - constructor 6; auto with arith.
-Qed.
-\end{Coqsrc}
-
-\subsubsection*{Proof of $\psi(1,0)\times 13+ \omega < \psi(0,\psi(2,1))$}
-\label{sect:ex6-first-proof}
-
-\begin{Coqsrc}
-Example Ex6 : gcons 1 0 12 omega t2< [0,[2,1]].
-Proof.
-  constructor 4.
-  - constructor 1.
-  - constructor 2.
-    + constructor 6; auto with arith.
-    + constructor 1.
-Qed.
-\end{Coqsrc}
-
-\subsubsection*{Proof of $\psi(2,1)\times 43 + \epsilon_0 < \psi(1,\psi(2,1))$}
-
-\begin{Coqsrc}
-Example Ex7 : gcons 2 1 42 epsilon0 t2< [1, [2,1]].
-Proof.
- constructor 5.
- constructor 6; auto with arith.
-Qed.
-\end{Coqsrc}
-
-\index{Projects}
-\begin{project}
-Write a tactic that solves automatically goals of the form (\texttt{$\alpha$ t2< $\beta$}), where $\alpha$ and $\beta$ are closed terms of type \texttt{T2}.
-\end{project}
-
-\section{Veblen Normal Forms}
-\begin{definition}
-  A term of the form $\psi(\alpha_1,\beta_1)\times n_1+ \psi(\alpha_2,\beta_2)\times n_2+\dots+\psi(\alpha_k,\beta_k)\times n_k$ is said to be in
- \emph{[Veblen] normal form} if for every $i<n$, $\psi(\alpha_i,\beta_i)<\psi(\alpha_{i+1},\beta_{i+1})$, all the $\alpha_i$ and $\beta_i$ are in normal form, and all the $n_i$ are strictly positive integers.
-\end{definition}
-
-\begin{Coqsrc}
-Inductive nf : T2 -> Prop :=
-| zero_nf : nf zero
-| single_nf : forall a b n, nf a ->  nf b -> nf (gcons a b n zero)
-| gcons_nf : forall a b n a' b' n' c', 
-                      [a', b'] t2< [a, b]  -> 
-                      nf a -> nf b -> 
-                      nf(gcons a' b' n' c')-> 
-                      nf(gcons a b n (gcons a' b' n' c')).
-\end{Coqsrc}
-
-Let us look at some positive examples (we have to prove some inversion lemmas before proving counter-examples).
-
-\begin{Coqsrc}
-Lemma  nf_fin i : nf (fin i).
-Proof.
-  destruct i.
-  - auto with T2.
-  - constructor 2; auto with T2.
-Qed.
-
-Lemma nf_omega : nf omega.
-Proof.  compute; auto with T2. Qed.
-
-Lemma nf_epsilon0 : nf epsilon0.
-Proof. constructor 2; auto with T2. Qed.
-
-Lemma nf_epsilon : forall alpha, nf alpha -> nf (epsilon alpha).
-Proof. compute; auto with T2. Qed.
-
-Example Ex8: nf (gcons 2 1 42 epsilon0).
-Proof.
-  constructor 3; auto with T2.
-  - apply Ex4.
-  - apply nf_fin.
-  - apply nf_fin.
-Qed.
-\end{Coqsrc}
-
-\subsection{Length of a Term}
-
-The notion of \emph{term length} is introduced by Schütte as a helper for proving (at least) the \emph{trichotomy} property and transitivity of the strict order \texttt{lt} on \texttt{T2}.
-These properties are proved by induction on length.
-
-\begin{Coqsrc}
-Fixpoint nbterms (t:T2) : nat :=
-  match t with zero => 0
-             | gcons a b n v => (S n) + nbterms v
-  end.
-
-Fixpoint t2_length (t:T2) : nat :=
-  match t  with 
-    zero => 0
-  | gcons a b n v => 
-       nbterms (gcons a b n v) + 
-      2 * (Max.max (t2_length a)
-                              (Max.max (t2_length b) 
-                                                (t2_length_aux v)))
-  end
-with t2_length_aux (t:T2) : nat :=
- match t with 
- | zero => 0
-  | gcons a b n v =>
-           Max.max (t2_length a) 
-                            (Max.max (t2_length b) (t2_length_aux v))
- end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Compute t2_length (gcons 2 1 42 epsilon0).
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = 48 : nat
-\end{Coqanswer}
-
-\subsection{Trichotomy}
-
-\emph{Trichotomy} is another name for the well-known property of decidable total ordering (like Standard Library's \texttt{Compare\_dec.lt\_eq\_lt\_dec}).
-
-We first prove by induction on $l$ the following lemma:
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0\#tricho_aux}%
-{\texttt{Gamma0.Gamma0}}}
-
-\begin{Coqsrc}
-Lemma tricho_aux (l: nat) : forall t t' :T2,
-      t2_length t + t2_length t' < l  ->
-      {t t2< t'} + {t = t'} + {t' t2<  t}.
-\end{Coqsrc}
-
-Then we get our version of \texttt{lt\_eq\_lt\_dec}, and derive a comparison function;
-
-\begin{Coqsrc}
-Definition lt_eq_lt_dec (t t': T2) : {t t2< t'}+{t = t'}+{t' t2<  t}.
-Proof.
-  eapply tricho_aux.
-  eapply lt_n_Sn.
-Defined.
-
-Definition compare (t1 t2 : T2) : comparison := 
-  match lt_eq_lt_dec t1 t2 with
-  | inleft (left _) => Lt
-  | inleft (right _) => Eq
-  | inright _ => Gt
-  end.
-\end{Coqsrc}
-
-With the help of \texttt{compare}, we get a boolean version of \texttt{nf}
-(being in Veblen normal form).
-
-\begin{Coqsrc}
-Fixpoint nfb (alpha : T2) : bool :=
-  match alpha with
-    zero => true
-  | gcons a b n zero => andb (nfb a) (nfb b)
-  | gcons a b n ((gcons a' b' n' c') as c) =>
-    match compare [a', b'] [a, b] with
-           Lt => andb (nfb a) (andb (nfb b) (nfb c))
-           | _ => false
-           end
-end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Compute compare (gcons 2 1 42 epsilon0) [2,2].
-\end{Coqsrc}
-
-\begin{Coqanswer}
-   = Lt
-     : comparison
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Compute nfb  (gcons 2 1 42 epsilon0).
-\end{Coqsrc}
-
-\begin{Coqanswer}
-   = true
-     : bool
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Compute nfb (gcons 2 1 42 (gcons 2 2 4 epsilon0)).
-\end{Coqsrc}
-
-\begin{Coqanswer}
-   = false
-     : bool
-\end{Coqanswer}
-
-\begin{remark}
-The connexion between the predicate \texttt{nf} and the relation \texttt{lt} on one part, and the functions \texttt{nfb} and \texttt{compare} on the other, is expressed by the following lemmas:
-
-\begin{Coqsrc}
-Lemma nfb_equiv gamma : nfb gamma = true <-> nf gamma.
-
-Lemma compare_correct alpha beta :
-  CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
-              (compare alpha beta).
-\end{Coqsrc}
-
-The function \texttt{compare} helps to make easier proofs of inequalities of
-closed terms of type \texttt{T2}.
-
-First, we prove a lemma:
-
-\begin{Coqsrc}
-Lemma compare_Lt : forall alpha beta, compare alpha beta = Lt -> 
-                                         alpha t2< beta.
-Proof.
-  intros alpha beta; destruct (compare_correct alpha beta);
-    trivial; discriminate. 
-Qed.
-\end{Coqsrc}
-
-Then, we give another version of the proof of Sect.~\vref{sect:ex6-first-proof}.
-
-\begin{Coqsrc}
-Example Ex6 : gcons 1 0 12 omega t2< [0,[2,1]].
-Proof. now apply compare_Lt. Qed.
-\end{Coqsrc}
-
-\end{remark}
-
-\section{Main Functions on \texttt{T2}}
-
-\subsection{Successor}
-The successor function is defined by structural recursion.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#succ}%
-{\texttt{Gamma0.T2}}}
-\begin{Coqsrc}
-Fixpoint succ (a:T2) : T2 :=
- match a with zero => one
-             | gcons zero zero n c => fin (S (S n))
-             | gcons a b n c => gcons a b n (succ c)
- end.
-\end{Coqsrc}
-
-\subsection{Addition}
-
-Like for Cantor normal forms (see Sect.~\ref{sect:infix-plus-T1}),  the definition of addition in \texttt{T2}  requires comparison between ordinal terms.
-
-\begin{Coqsrc}
-Fixpoint plus (t1 t2 : T2) {struct t1}:T2 :=
-  match t1,t2 with
-  |  zero, y  => y
-  |  x, zero => x
-  |  gcons a b n c, gcons a' b' n' c' =>
-     (match compare (gcons a b 0 zero)
-                    (gcons a' b' 0 zero) with
-      | Lt => gcons a' b' n' c'
-      | Gt => gcons a b n (c + gcons a' b' n' c')
-      | Eq => gcons a b (S(n+n')) c'
-      end)
-  end
-where "alpha + beta" := (plus alpha beta): T2_scope.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Example Ex7 : 3 + epsilon0 = epsilon0.
-Proof. trivial. Qed.
-\end{Coqsrc}
-
-\subsection{The Veblen Function \texorpdfstring{$\phi$}{\texttt{phi}}}
-
-The enumeration function of critical ordinals, presented in Sect.~\vref{sect:phi-schutte}, is recursively defined in type \texttt{T2}.
-
-\begin{Coqsrc}
-Definition  phi (alpha beta : T2) : T2 :=
-  match beta with zero => [alpha, beta] 
-             | [b1, b2] => 
-               (match compare alpha b1
-                with Datatypes.Lt => [b1, b2 ]
-                | _ => [alpha,[b1, b2]]
-                end)
-             | gcons b1 b2 0 (gcons zero zero  n zero) => 
-               (match compare alpha b1
-                with  Datatypes.Lt => 
-                      [alpha, (gcons b1 b2 0 (fin n))]
-                | _ =>  [alpha, (gcons b1 b2 0 (fin (S n)))]
-                end)
-             | any_beta => [alpha, any_beta]
-  end.
-\end{Coqsrc}
-
-Despite its complexity, the function \texttt{phi} is well adapted to proofs by simplification or computation.
-\begin{Coqsrc}
-Example Ex8:  phi 1 (succ epsilon0) = [1, [1,0] + 1].
-Proof. reflexivity. Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-(**  All epsilons are fixpoints of phi 0 *)
-
-Theorem epsilon_fxp : forall beta, phi zero (epsilon beta) =
-                                   epsilon beta.
-Proof. reflexivity. Qed.
-
-Theorem epsilon0_fxp : epsilon0 = phi zero epsilon0.
-Proof. apply epsilon_fxp. Qed.
-\end{Coqsrc}
-
-The relation between the constructor $\psi$ and the function $\phi$ is
-studied in~\cite{schutte}, and partially implemented in this development.
-\emph{Please contribute!}
-
-For instance, the following theorem states that, if $\gamma$ is the sum of a limit ordinal $\beta$ and a finite ordinal $n$, and $\beta$ is a fixpoint of
-$\phi(\alpha)$, then $\psi(\alpha,\gamma)=\phi_\alpha(\gamma+1)$.
-
-\begin{Coqanswer}
-phi_psi :
-forall (alpha : T2) [beta gamma : T2] [n : nat],
-nf gamma ->
-limit_plus_fin beta n gamma ->
-phi alpha beta = beta -> [alpha, gamma] = phi alpha (succ gamma)
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Example Ex9 : [zero, epsilon 2 + 4] = phi 0 (epsilon 2 + 5).
-Proof. trivial. Qed.
-\end{Coqsrc}
-
-On the other hand, $\phi$ can be expressed in terms of $\psi$.
-
-\begin{Coqanswer}
-phi_of_psi:
-  forall a b1 b2 : T2,
-  phi a [b1, b2] = (if lt_ge_dec a b1 then [b1, b2] else [a, [b1, b2]])
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Example Ex10 : phi omega [epsilon0, 5] = [epsilon0, 5].
-Proof. reflexivity. Qed.
-\end{Coqsrc}
-
-\index{Projects}
-\begin{project}
-Please study a way to pretty print ordinal terms in Veblen normal form (see Section~\vref{sect:ppT1}).
-\end{project}
-
-\section{An Ordinal Notation for \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}}
-
-In order to consider type \texttt{T2} as an ordinal notation, we have to build an instance of class \texttt{ON} (See Definition page~\pageref{types:ON}).
-
-First, we define a type that contains only terms in Veblen normal form, and redefine \texttt{lt} and \texttt{compare} by delegation (see for comparison the construction of type \texttt{E0} in Sect.~\vref{sect:E0-def}).
-
-\begin{Coqsrc}
-Module G0.
-
-Class G0 := mkg0 {vnf : T2; vnf_ok : nfb vnf}.
-
-Definition lt (alpha beta : G0) := T2.lt (@vnf alpha) (@vnf beta).
-
-Definition compare alpha beta := Gamma0.compare (@vnf alpha) (@vnf beta).
-\end{Coqsrc}
-
-Then, we prove that \texttt{lt} is a well-founded strict order and that the
-function \texttt{compare} is correct.
-
-\begin{Coqsrc}
-Instance lt_sto : StrictOrder lt.
-
-Lemma lt_wf : well_founded lt.
-
-Lemma compare_correct alpha beta :
-  CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
-              (compare alpha beta).
-
-Instance Gamma0: ON lt  compare.
-Proof.
-  split.
-  - apply lt_sto.
-  - apply lt_wf. 
-  - apply compare_correct.
-Qed.
-\end{Coqsrc}
-
-\begin{remark}
-The proof of \texttt{lt\_wf} has been written by \'Evelyne Contejean, using her library on the recursive path ordering (see also remark~\vref{remark:a3pat}).
-\end{remark}
-
-\index{Projects}
-\begin{project}
-Prove that \texttt{Epsilon0} (page~\pageref{instance-epsilon0})
-is a sub-notation system of \texttt{Gamma0}.
-
-Prove that the implemantations of \texttt{succ}, \texttt{+}, $\phi_0$, etc.
-are compatible in both notation systems.
-
-Note that a function \texttt{T1\_inj} from \texttt{T1} to \texttt{T2} has already been defined. It may help to complete the task.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T1_to_T2}%
-{\texttt{Gamma0.T2}}}
-\begin{Coqsrc}
-(* injection from T1 *)
-
-Fixpoint T1_to_T2 (alpha :T1) : T2 :=
-  match alpha  with
-  | T1.zero => zero
-  | T1.ocons a n b => gcons zero (T1_to_T2 a) n (T1_to_T2 b)
-  end.
-\end{Coqsrc}
-
-\end{project}
-
-\begin{project}
-Prove that the notation system \texttt{Gamma0} is a correct implementation 
-of the segment $[0,\Gamma_0)$ of the set of countable ordinals.
-\end{project}
-
-\chapter{Appendices}
+\part{Hydras and Ordinals}
+
+\include{part-hydras}
+
+\part{Some Funny Algorithms}
+
+\include{chapter-powers}
+
+\part{Appendices}
 

### 2020-11-15 doc/chapter-powers.tex
Added addition-chains stuff into a general hydras-and-Cie development
- Not tested yet :
  - opam directives, etc.
     the current states uses coq-equations, coq-param, and
         coq-mathcomp-algebra (for the contribution by Yves)

  - There is an ml file, provisionnaly stored as theories/additions/stub.ml,
    and a shell script: theories/additions/make_fib_tests.txt
    stub.ml uses two files generated by extraction from fib.v : bigfib.ml and bigfig.mli.

@@
 \emph{From Module 
-\href{../src/html/addition.power.FirstSteps.html}{\texttt{Powers.FirstSteps}}}
+\href{../theories/html/hydras.additions.FirstSteps.html}{\texttt{additions.FirstSteps}}}
 \label{sect: power-definitions}
@@
 \emph{From Module
-\href{../src/html/addition.power.FirstSteps.html}{Powers.FirstSteps}}
+\href{../theories/html/hydras.additions.FirstSteps.html}{additions.FirstSteps}}
 
@@
 \end{Coqanswer}
+
+\begin{todo}
+Document the files contributed by Yves
+\begin{itemize}
+\item additions/fib.v (to rename ?)
+\item additions/stub.ml (to keep inside theories/ or move to src/ ?)
+\item theories/additions/make\_fib\_tests.txt (to put in a Makefile?)
+\end{itemize}
+\end{todo}
 
@@
 
-\emph{From Module \href{../src/html/addition.power.Monoid_def.html}{Powers/Monoid\_def.v}.}
+\emph{From Module \href{../theories/html/hydras.additions.Monoid_def.html}{additions/Monoid\_def.v}.}
 
@@
 
-\emph{From Module \href{../src/html/addition.power.Monoid_def.html}{Powers.Monoid\_def}}
+\emph{From Module \href{../theories/html/hydras.additions.Monoid_def.html}{additions.Monoid\_def}}
 
@@
 
-The following class definition, from Module \href{../src/html/addition.power.Monoid_def.html}{Powers.Monoid\_def},
+The following class definition, from Module \href{../theories/html/hydras.additions.Monoid_def.html}{additions.Monoid\_def},
 is parameterized with some type $A$,
@@
 Complete definitions and proofs are given in 
-File~\href{../src/html/addition.power.Monoid_instances.html}{Powers/Monoid\_instances.v}.
+File~\href{../theories/html/hydras.additions.Monoid_instances.html}{additions/Monoid\_instances.v}.
 
@@
 \noindent
-\emph{From Module \href{../src/html/addition.power.Monoid_def.html}{Powers.Monoid\_def}}
+\emph{From Module \href{../theories/html/hydras.additions.Monoid_def.html}{additions.Monoid\_def}}
 
@@
 
-\emph{From Module \href{../src/html/addition.power.Monoid_instances.html}{Powers.Monoid\_instances}}
+\emph{From Module \href{../theories/html/hydras.additions.Monoid_instances.html}{additions.Monoid\_instances}}
 
@@
 
-The  module \href{../src/html/addition.power.Pow.html}{Powers.Pow} defines two functions for exponentiation on any 
+The  module \href{../theories/html/hydras.additions.Pow.html}{additions.Pow} defines two functions for exponentiation on any 
 \texttt{EMonoid}  on carrier $A$.
@@
 
-From Module~\href{../src/html/addition.power.Demo_power.html}{Powers.Demo\_power}
+From Module~\href{../theories/html/hydras.additions.Demo_power.html}{additions.Demo\_power}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../src/html/addition.power.Pow.html}{Powers.Pow}}
+\emph{From Module~\href{../theories/html/hydras.additions.Pow.html}{additions.Pow}}
 
@@
 \vspace{4pt}
-From Module~\href{../src/html/addition.power.Pow.html}{Powers.Pow}
+From Module~\href{../theories/html/hydras.additions.Pow.html}{additions.Pow}
 
@@
 Section \texttt{M\_given} of Module 
-~\href{../src/html/addition.power.Pow.html}{Powers.Pow} is devoted to the proof 
+~\href{../theories/html/hydras.additions.Pow.html}{additions.Pow} is devoted to the proof 
 of properties of the functions above.
@@
 
-From Module~\href{../src/html/addition.power.Chains.html}{Powers.Chains}
+From Module~\href{../theories/html/hydras.additions.Chains.html}{additions.Chains}
 
@@
 This tactic is not adapted to much bigger exponents. In \linebreak
- Module~\href{../src/html/addition.power.Euclidean_Chains.html}{Euclidean\_Chains},
+ Module~\href{../theories/html/hydras.additions.Euclidean_Chains.html}{Euclidean\_Chains},
  for instance, we tried to apply this tactic for proving the correctness 
@@
 All the \coq{} material described in this section is available on 
- Module~\href{../src/html/addition.power.Euclidean_Chains.html}{Powers/Euclidean\_Chains.v}
+ Module~\href{../theories/html/hydras.additions.Euclidean_Chains.html}{additions/Euclidean\_Chains.v}
 
@@
 
-Module ~\href{../src/html/addition.power.Dichotomy.html}{Powers.Dichotomy} contains a definition of the function \texttt{delta}, and proofs that if 
+Module ~\href{../theories/html/hydras.additions.Dichotomy.html}{additions.Dichotomy} contains a definition of the function \texttt{delta}, and proofs that if 
 $n>3$ then $1<\delta(n)<n$.

### 2020-11-16 doc/chapter-powers.tex
minor changes

@@
 Fixpoint power (x:A)(n:nat) : A :=
-  match n with 0%nat => 1
+  match n with 
+   | 0%nat => 1
    | S p =>   x * x ^ p
@@
 
-The number of multiplications needed to compute $x^n$ via this function is linear with respect to $n$.
+An application of this function for  computing $x^n$ needs $n$ multiplications.
  Despite this lack of efficiency, and thanks to its simplicity, we keep it as a specification for more efficient and complex exponentiation algorithms.

### 2020-11-17 doc/hydras.tex
minor changes
minor corrections

--- 2020-11-10 doc/hydras.tex
@@
 presented for instance in~\cite{Promel2013}. 
-\index{Maths!Hardy Hierarchy}
+\index{Maths!Hardy hierarchy}
 
@@
 For instance, the following theorem states that, if $\gamma$ is the sum of a limit ordinal $\beta$ and a finite ordinal $n$, and $\beta$ is a fixpoint of
-$\phi(\alpha)$, then $\psi(\alpha,\gamma)=\phi_\alpha(\gamma+1)$.
+$\phi_\alpha$, then $\psi(\alpha,\gamma)=\phi_\alpha(\gamma+1)$.
 
@@
 Name & Gallina & Math & Description & Page \\ \hline
-canon & canon alpha i & $\canonseq{\alpha}{i}$& Canonical sequence & \pageref{Functions:canon}\\
-canonS & canonS alpha i & $\canonseq{\alpha}{i+1}$& Helper for canon & \pageref{Functions:canonS}\\
+canon & canon alpha i & $\canonseq{\alpha}{i}$& canonical sequence & \pageref{Functions:canon}\\
+canonS & canonS alpha i & $\canonseq{\alpha}{i+1}$& helper for canon & \pageref{Functions:canonS}\\
 F\_ & F\_ alpha n & $F_\alpha(n)$ & Wainer's \texttt{F} fast growing hierarchy & \pageref{Functions:F-alpha} \\
 H\_ & H\_ alpha n & $H_\alpha(n)$ & Hardy's \texttt{H} fast growing hierarchy & \pageref{Functions:H-alpha} \\
-iterate & iterate $f$ $n$ $x$ & $f^{(n)}(x)$ & Functional iteration & \pageref{Functions:iterate} \\
+iterate & iterate $f$ $n$ $x$ & $f^{(n)}(x)$ & functional iteration & \pageref{Functions:iterate} \\
 L\_ & L\_ alpha k & $L_\alpha(k)$ & final step of a standard path & \pageref{Functions:L-alpha} \\
-succ & succ alpha & & Successor & \pageref{Functions:succ-T1}, \pageref{Functions:succ-sch} \dots \\
-zero:Ord & & $0$ & The least ordinal (Schütte's model) & \pageref{Constants:zero:Ord} \\
+succ & succ alpha & & successor & \pageref{Functions:succ-T1}, \pageref{Functions:succ-sch} \dots \\
+zero:Ord & & $0$ & the least ordinal (Schütte's model) & \pageref{Constants:zero:Ord} \\
 

### 2020-11-20 doc/chapter-powers.tex
New library additions/AM.v (to certify)
new project in additions

@@
 Study how  to compile efficiently such data structures.
+
 \end{project}
@@
 
+\index{Projects}
+\begin{project}
+Figures~\vref{fig:F3-as-dag} to \vref{fig:FFK} suggest that any computation following an Euclidean chain can be executed on a kind  of abstract machine with a "register'' and a stack, and only four operations:
+\begin{itemize}
+\item multiply the contents of the register by the top of the stack (and pop that stack),
+\item raising the contents of the register to its square,
+\item push the contents of the register into the stack,
+\item swapping the two elements at the top of the stack.
+\end{itemize}
+
+In \coq{}, we define the instructions as the four constructors of an inductive type.
+
+From Module~\href{../theories/html/hydras.additions.AM.html}{additions.AM}
+
+\begin{Coqsrc}
+(** basic instructions *)
+
+Inductive instr : Set :=
+  | MUL : instr
+  | SQR : instr
+  | PUSH : instr
+  | SWAP : instr.
+
+Definition code := list instr.
+
+(* semantics *)
+(*************)
+
+Section Semantics.
+
+ Variable M : Type.
+ Variable mul : M -> M -> M.
+ Variable one : M.
+
+ Definition stack := list M.
+ Definition config := (M * list M)%type.
+
+ Fixpoint exec (c : code) (x:M) (s: stack) : option config :=
+   match c, s with
+     nil, _ => Some (x,s)
+   | MUL::c, y::s => exec c (mul x y) s
+   | SQR::c, s => exec c (mul x x) s
+   | PUSH::c, s => exec c x (x::s)
+   | SWAP::c, y::z::s => exec c x (z::y::s)
+   | _,_ => None
+   end.
+\end{Coqsrc}
+
+Please define an Euclidean chain generator using this data structure, and prove its correctness (partially solved in ~\href{../theories/html/hydras.additions.AM.html}{additions.AM}).
+\end{project}
+
 %\subsection{Optimality and Relative Efficiency}

### 2020-11-20 doc/chapter-powers.tex
redaction chapter-powers

@@
 
- Variable M : Type.
- Variable mul : M -> M -> M.
- Variable one : M.
-
- Definition stack := list M.
- Definition config := (M * list M)%type.
-
- Fixpoint exec (c : code) (x:M) (s: stack) : option config :=
+ Variable A : Type.
+ Variable mul : A -> A -> A.
+ Variable one : A.
+
+ Definition stack := list A.
+ Definition config := (A * list A)%type.
+
+ Fixpoint exec (c : code) (x:A) (s: stack) : option config :=
    match c, s with
@@
    end.
-\end{Coqsrc}
-
-Please define an Euclidean chain generator using this data structure, and prove its correctness (partially solved in ~\href{../theories/html/hydras.additions.AM.html}{additions.AM}).
+
+(* ... *)
+End Semantics.
+\end{Coqsrc}
+
+For instance the chain of Fig.~ \vref{fig:C87} can be represented with the following code:
+
+\begin{Coqsrc}
+  PUSH :: PUSH :: SQR  :: MUL :: PUSH :: SWAP :: SQR :: MUL :: PUSH
+            :: SWAP :: MUL :: SQR :: SQR :: SQR :: MUL :: nil 
+\end{Coqsrc}
+
+Please define an Euclidean chain generator using this data structure, and prove its correctness (partially done  in ~\href{../theories/html/hydras.additions.AM.html}{additions.AM}).
 \end{project}

### 2020-11-21 doc/chapter-powers.tex
modifying the arguments of correctness predicates

@@
 \begin{Coqsrc}
-Definition chain_correct_nat (c: chain) (n:nat) := 
+Definition chain_correct_nat (n:nat) (c: chain) := 
   n <> 0 /\
@@
 
-Definition chain_correct (c: chain) (p:positive) :=
+Definition chain_correct (p:positive) (c: chain) :=
   chain_correct_nat c (Pos.to_nat p). 
@@
   match goal with 
-      [ |- chain_correct ?c ?p ] =>
+      [ |- chain_correct ?p ?c ] =>
       let A := fresh "A" in
@@
 
-Example C7_ok : chain_correct C7 7.
+Example C7_ok : chain_correct 7 C7.
 Proof.
@@
 \begin{Coqsrc}
-Example C87_ok : chain_correct C87 87.
+Example C87_ok : chain_correct 87 C87.
 Proof.
@@
 match goal with
-[ |- chain_correct ?c ?n ] =>
+[ |- chain_correct ?n ?c ] =>
  split; [try discriminate |
@@
  forall c:chain, parametric c ->  
-               chain_correct_nat c  (the_exponent_nat c).
+               chain_correct_nat (the_exponent_nat c)  c.
 (* Proof omitted *)
@@
    parametric c -> 
-   chain_correct c (the_exponent c). 
+   chain_correct (the_exponent c) c. 
 Proof.
@@
 match goal with 
-[|- chain_correct ?c ?p ] => 
+[|- chain_correct ?p ?c ] => 
 apply param_correctness; parametric_tac
@@
 
-Lemma C87_ok' : chain_correct C87 87.
+Lemma C87_ok' : chain_correct 87 C87.
 Time param_chain_correct.
@@
 Definition correct_generator (g : positive -> chain) :=
- forall p, chain_correct p (g p).
+ forall p, chain_correct (g p) p.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Remark F9_correct :chain_correct (F2C F9) 9.
+Remark F9_correct :chain_correct 9 (F2C F9).
 Proof.
@@
 
-Definition Fchain_correct  (f : Fchain) (n:nat) :=
+Definition Fchain_correct  (n:nat) (f : Fchain) :=
   forall A `(M : @EMonoid A op E_one E_equiv) k (a:A),
@@
 \begin{Coqbad}
-Theorem F3_correct : Fchain_correct F3 3.
+Theorem F3_correct : Fchain_correct 3 F3.
 Proof.  
@@
 \begin{Coqsrc}
-Definition Fchain_correct_nat (f : Fchain) (n:nat) :=
+Definition Fchain_correct_nat (n:nat) (f : Fchain) :=
  forall A `(M : @EMonoid A op E_one E_equiv) k
@@
 
-Definition Fchain_correct (f : Fchain) (p:positive) :=
- Fchain_correct_nat f (Pos.to_nat p).
+Definition Fchain_correct (p:positive) (f : Fchain) :=
+ Fchain_correct_nat (Pos.to_nat p) f.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Lemma F1_correct : Fchain_correct F1 1.
+Lemma F1_correct : Fchain_correct 1 F1.
 Proof.
@@
 \begin{Coqsrc}
-Theorem F3_correct : Fchain_correct F3 3.
+Theorem F3_correct : Fchain_correct 3 F3.
 Proof. 
@@
 \begin{Coqsrc}
-Theorem F2_correct : Fchain_correct F2 2.
+Theorem F2_correct : Fchain_correct 2 F2.
 Proof. 
@@
 Lemma Fcompose_correct_attempt :
-  forall f1 f2 n1 n2, Fchain_correct f1 n1 ->
-                      Fchain_correct f2 n2 ->
-                      Fchain_correct (Fcompose f1 f2)
-                                       (n1 * n2).
+  forall f1 f2 n1 n2, Fchain_correct n1 f1 ->
+                      Fchain_correct n2 f2 ->
+                      Fchain_correct (n1 * n2) (Fcompose f1 f2).
 
@@
 Lemma  Fexp2_nat_correct (n:nat) : 
-         Fchain_correct_nat (Fexp2_of_nat n) (2  ^ n).
+         Fchain_correct_nat (2  ^ n) (Fexp2_of_nat n).
 Proof.
@@
 Lemma  Fexp2_correct (p:positive) : 
-                           Fchain_correct (Fexp2 p) (2  ^ p).
+                           Fchain_correct (2  ^ p) (Fexp2 p).
 (* Proof omitted *)
@@
 
-Example F144:  {f : Fchain | Fchain_correct f 144 /\
+Example F144:  {f : Fchain | Fchain_correct 144 f /\
                                 Fchain_proper f}.
@@
  \begin{Coqsrc}
-Definition Kchain_correct_nat (kc : Kchain) (n p : nat) :=
+Definition Kchain_correct_nat (n p : nat) (kc : Kchain) :=
   forall `(M : @EMonoid A op E_one E_equiv)
@@
 \begin{Coqsrc}
-Definition Kchain_correct (kc : Kchain) (n p : positive) :=
-  Kchain_correct_nat kc (Pos.to_nat n) (Pos.to_nat p).
+Definition Kchain_correct (n p : positive) (kc : Kchain) :=
+  Kchain_correct_nat  (Pos.to_nat n) (Pos.to_nat p) kc.
 
@@
 
-Lemma k7_3_correct : Kchain_correct k7_3 7 3.
+Lemma k7_3_correct : Kchain_correct 7 3 k7_3.
 Proof.
@@
   forall knp n p, Kchain_correct kc n p ->
-                 Fchain_correct (K2F knp) n.
+                 Fchain_correct (K2F n) knp.
 (* Proof omitted *)
@@
   forall (b q r : positive) (kbr : Kchain) (fq : Fchain),
-    Kchain_correct kbr b r ->
-    Fchain_correct fq q ->
+    Kchain_correct b r  kbr ->
+    Fchain_correct q fq ->
     Kchain_proper kbr ->
     Fchain_proper fq -> 
-    Kchain_correct (KFK kbr fq) (b * q + r) b.
+    Kchain_correct  (b * q + r) b (KFK kbr fq).
 (* Proof omitted *)
@@
 forall (b q r : positive) (kbr : Kchain) (fq : Fchain),
-Kchain_correct kbr b r ->
-Fchain_correct fq q ->
+Kchain_correct  b r kbr ->
+Fchain_correct q fq ->
 Kchain_proper kbr ->
-Fchain_proper fq -> Fchain_correct (KFF kbr fq) (b * q + r).
+Fchain_proper fq -> Fchain_correct (b * q + r) (KFF kbr fq).
 (* Proof omitted *)
@@
 Lemma FK_correct : forall (p: positive) (Fp : Fchain),
-                     Fchain_correct  Fp p ->
+                     Fchain_correct  p Fp ->
                      Fchain_proper Fp ->
-                     Kchain_correct (FK Fp) p 1.
+                     Kchain_correct p 1 (FK Fp) .
 (* Proof omitted *)
@@
   forall (p q  : positive) (fp fq : Fchain),
-    Fchain_correct fp p  ->
-    Fchain_correct fq q ->
+    Fchain_correct p fp  ->
+    Fchain_correct q fq ->
     Fchain_proper fp ->
-    Fchain_proper fq -> Kchain_correct (FFK fp fq) (p * q ) p.
+    Fchain_proper fq -> Kchain_correct  (p * q ) p (FFK fp fq).
 (* Proof omitted *)
@@
 
-Lemma OK87 : Fchain_correct F87 87.
+Lemma OK87 : Fchain_correct 87 F87.
 Proof.
@@
 Theorem make_chain_correct : 
-   forall p, chain_correct (make_chain p) p.  
+   forall p, chain_correct p (make_chain p).  
 \end{Coqsrc}
@@
 match s  with
-  | gen_F p => fun ch => Fchain_correct ch p
-  | gen_K p d   => fun ch => Kchain_correct ch (p + d) p
+  | gen_F p => fun ch => Fchain_correct p ch
+  | gen_K p d   => fun ch => Kchain_correct (p + d) p ch
 end.
@@
 Theorem make_chain_correct :
-    forall p, chain_correct (make_chain p) p.
+    forall p, chain_correct p (make_chain p).
 Proof.
@@
 
-%\subsection{Optimality and Relative Efficiency}
-
-% \begin{project}[Proof techniques]
-% \noindent
-
-% \begin{enumerate}
-% \item Improve automated proofs on types \texttt{positive} and \texttt{N}.
-% \item Compare  \texttt{Program Fixpoint} and \texttt{Function} for
-% writing \texttt{make\_chain}. Consider measure \emph{vs} well-founded 
-% relations, mutual recursion, possibility of using sigma-types, etc.
-% \item Chains are always associated with strictly positive exponents. 
-% Thus, many lemmas about chain correctness  can be proved using semi-groups instead of
-% monoids. Define type classes for semi-groups and use them whenever possible.
-% \end{enumerate}  
+\begin{project}[Some improvements]
+ \begin{enumerate}
+\item Improve automated proofs on types \texttt{positive} and \texttt{N}.
+\item Compare  \texttt{Program Fixpoint} and \texttt{Function} for
+writing \texttt{make\_chain}. Consider measure \emph{vs} well-founded 
+relations, mutual recursion, possibility of using sigma-types, etc.
+\item Chains are always associated with strictly positive exponents. 
+Thus, many lemmas about chain correctness  can be proved using semi-groups instead of
+monoids. Define type classes for semi-groups and use them whenever possible.
+\end{enumerate}  
+\end{project}
+
 % \end{project}

### 2020-11-23 doc/chapter-powers.tex
Representation of Euclidean chains as lists of instructions. works, but still to simplify and document the proofs.

@@
 
-Please define an Euclidean chain generator using this data structure, and prove its correctness (partially done  in ~\href{../theories/html/hydras.additions.AM.html}{additions.AM}).
+Please define an Euclidean chain generator using this data structure, and prove its correctness (done  in ~\href{../theories/html/hydras.additions.AM.html}{additions.AM}, but with clumsy fresh proofs).
+The project is now to simplify the proofs.
+perhaps considering chains of type \texttt{code} as a refinement of functional chains. Please note that many proof scripts are copied verbatim from 
+\texttt{Euclidean\_Chains} into \texttt{AM}.
 \end{project}

### 2020-11-24 doc/part-hydras.tex
creation of a theories/ordinals/ directory

--- 2020-11-10 doc/hydras.tex
@@
-\documentclass[a4paper]{book}
-\usepackage{alltt}
-\usepackage{makeidx}
-\usepackage{hyperref}
-\usepackage{url}
-\usepackage{mdframed}
-\usepackage{amsmath}
-\usepackage{varioref}
-\usepackage{synttree, proof, centernot}
-\usepackage[firstpage]{draftwatermark}
-\usepackage{verbatim}
-\usepackage[note]{marginote}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{mathtools}
-\usepackage{dsfont}
-\usepackage{threeparttable}
-\setcounter{tocdepth}{1}
-\input{coqmacros}
-
-\author{Pierre Castéran\\ LaBRI, Univ. Bordeaux, CNRS UMR 5800, Bordeaux-INP\\with contributions by \'Evelyne Contejean,  Florian Hatat and Pascal Manoury}
-\date{\today}
-\title{Hydra Battles in Coq}
-\makeindex
-\begin{document}
-\maketitle
-
-\cleardoublepage
-
-% \mbox{~}
-% \clearpage
-% \newpage
-% \mbox{~}
-\vspace{10cm}
-\begin{quote} 
-{ \Large {\it
-``I start from one point and go as far as possible.'' }
-
-John Coltrane.}
-\end{quote}
-
-% \begin{verse}
-% Hydra hydra hydram \\
-
-% Hydrae hydrae hydra \\
-
-% Hydrae hydrae hydras \\
-
-% Hydrarum hydris hydris\\
-% \end{verse}
-
-\tableofcontents
-
-%-------------------------------------------------------------------
-
-\chapter{Introduction}
-
-\vspace{16pt}
-
-Proof assistants are excellent tools for exploring the structure of mathematical proofs,
-studying  which hypotheses are really needed, and which proof patterns are useful and/or
-necessary. Since the development of a theory is represented as a bunch of computer files,
-everyone is able to read the proofs with an arbitrary level of detail, or to play with the theory by writing alternate proofs or definitions.
-
-Among all the theorems proved with the help of proof assistants like \coq{}, \isabelle{}, \hol{}, etc.,
-several statements and proofs  share some interesting features:
-\begin{itemize}
-\item Their statements are easy to understand, even by non-mathematicians
-\item Their proof requires some non-trivial mathematical tools
-\item Their mechanization on computer presents some methodological interest.
-\end{itemize}
-
-This is obviously the case of the four-color theorem~\cite{fourcolors}  and the Kepler conjecture~\cite{flyspeck2015}.
-We do not mention impressive works like the proof of the odd-order theorem ~\cite{oddorderthm}, since understanding its statement requires a quite good mathematical culture.
-
-Hydra games (a.k.a. \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians:
-L. Kirby and J. Paris~\cite{KP82}: \emph{Accessible Independence Results for Peano Arithmetic}. 
-Although the mathematical contents of this 
-paper are quite advanced, the rules of hydra battles are very easy to understand.
-There are now several sites on Internet where you can find tutorials on hydra games, together with simulators you can play with.
-See, for instance, the page written by Andrej Bauer~\cite{bauer2008}.
-
-Hydra battles, as well as Goodstein Sequences~\cite{goodstein_1944, KP82}
-are a nice way to present complex termination problems.
-The article by Kirby and Paris presents a proof of termination
-based on ordinal numbers, as well as a proof that this termination is not
-provable in Peano arithmetic. In the book dedicated to 
-J.P. ~Jouannaud \cite{HommageJPJ}, N.~Dershowitz and G.~Moser  give a thorough survey on this topic~\cite{Dershowitz2007}.
-
-Here, we present a development  for the \coq{} proof assistant, after the work of Kirby and Paris. This formalization contains the following main parts:
+\section*{Introduction}
+
+In this part, we present a development  for the \coq{} proof assistant, after the work of Kirby and Paris. This formalization contains the following main parts:
 
@@
 
-\section{Remarks}
-
-\subsection{Difference from Kirby and Paris's Work}
+%\section{Remarks}
+
+\subsection*{Difference from Kirby and Paris's Work}
 In~\cite{KP82}, Kirby and Paris showed  that there is no proof of termination of all hydra battles in Peano Arithmetic (PA).
@@
 
-\subsection{Trust in our Proofs?}
-\label{sect:trust-in-proofs}
-
-Unlike mathematical literature, where definitions and proofs are spread over many articles and books,
-the whole proof is now inside your computer. It is composed of the \texttt{.v} files you downloaded and 
-parts of \coq's standard library.
-Thus, there is no ambiguity in our definitions and the premises of the theorems.
-Furthermore, you will be able to navigate through the development, using your favourite text editor or IDE, and some commands like \texttt{Search}, \texttt{Locate},  etc.
-
-\subsection{Assumed Redundancy}
-
-It may often happen that several definitions of a given concept, or several proofs of a given theorem are possible.
-If all the versions present some interest, we will make them available, since each one may be of some methodological
-interest (by illustrating some tactic of proof pattern, for instance).
-We may use \coq's module system to make several proofs of a given theorem co-exist in our libraries (see also Sect~\vref{sect:alt-proofs}).
-After some discussions of the pros and cons of each solution, we develop only one of them, leaving the others  as exercises or projects (i.e., big or difficult exercises).
-In order to discuss which assumptions are really needed for proving a theorem, we will also present 
-several aborted proofs.
-Of course, do not hesitate to contribute nice proofs or alternative definitions !
-
-It may also happen that some proof looks to be useless, because the proven theorem is a trivial consequence of another (proven too) result.
-For instance, let us consider the three following statements:
-\begin{enumerate}
-\item There is no measure into $\mathbb{N}$ for proving the termination of all hydra battles (Sect~\vref{omega-case}).
-\item There is no measure into $[0,\omega^2)$ for proving the termination of all hydra battle (Sect~\vref{omega2-case}).
-\item There is no measure into $[0,\mu)$ for proving the termination of all hydra battles, for any $\mu<\epsilon_0$ (Sect\vref{sec:free-battles-case}).
-\end{enumerate}
-
-Obviously, the third theorem implies the second one, which implies the first one.
-So, theoretically, a library would contain only a proof of $(3)$ and remarks for $(2)$ and $(1)$.
-But we found it interesting to make all the three proofs available, allowing the reader to compare their common structure and notice their technical differences.
-In particular, the proof of $(3)$ uses several non-trivial combinatorial properties of ordinal numbers up to $\epsilon_0$~\cite{KS81}, whilst $(1)$ and $(2)$ use simple properties of $\mathbb{N}$ and $\mathbb{N}^2$.
-
-\subsection{About Logic}
-
-Most of the proofs we present are \emph{constructive}.
-Whenever possible, we provide the user with an associated function, which she or he can apply in \gallina{} or \ocaml{} in order to get a ``concrete'' feeling of the meaning of the considered theorem.
-For instance, in Chapter~\vref{chap:ketonen}, the notion of \emph{limit ordinal} is
-made more ``concrete'' thanks to a function \texttt{canon} which computes every item of a sequence which converges on a given limit ordinal $\alpha$.
-This simply typed function allows the user/reader to make her/his own experimentations.
-For instance, one can very easily compute the $42$-nd item of a sequence which converges towards $\omega^{\omega^\omega}$.
-
-Except in the \texttt{Schutte} library, dedicated to an axiomatic presentation of the set of countable ordinal numbers, all our development is axiom-free, and respects the rules of intuitionistic logic.
-Note that we also use the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777} in the definitition of  several rapidly growing hierarchy of functions, in Chap.~\ref{chap:alpha-large}.
-This plug-in imports several known-as-harmless  axioms.
-
-% \begin{Coqsrc}
-% FunctionalExtensionality.functional_extensionality_dep : 
-% forall (A : Type) (B : A -> Type) (f g : forall x : A, B x),
-% (forall x : A, f x = g x) -> f = g
-% \end{Coqsrc}
-
-\index{Coq!Commands!Print Assumptions}
-
-At any place of our development, you may use the  \texttt{Print Assumptions {\it ident}} command in order to verify on which axiom the theorem {\it ident} may depend.
-The following example is extracted from
-Library~\href{../theories/html/hydras.Epsilon0.F_alpha.html}{hydras.Epsilon0.F\_alpha}, where we use the \texttt{coq-equations} plug-in (see Sect.~\vref{sect:wainer}).
-
-\begin{Coqsrc}
-About F_zero_eqn.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-F_zero_eqn : forall i : nat, F_ Zero i = S i
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Print Assumptions F_zero_eqn. 
-\end{Coqsrc}
-
-\begin{Coqanswer}
-Axioms:
-FunctionalExtensionality.functional_extensionality_dep
-  : forall (A : Type) (B : A -> Type) (f g : forall x : A, B x),
-    (forall x : A, f x = g x) -> f = g
-Eqdep.Eq_rect_eq.eq_rect_eq
-  : forall (U : Type) (p : U) (Q : U -> Type) (x : Q p) (h : p = p),
-    x = eq_rect p Q x p h
-\end{Coqanswer}
-
-\subsection{Main References}
+\subsection*{State of the Development}
+The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on  notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
+We added new material : axiomatic definition of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, equations, etc.
+
+We are now working in order to make clumsy proofs more readable, simplify definitions, and ``factorize'' proofs as much as possible. 
+Many possible improvements are suggested as ``todo''s or ``projects'' in this text.
+
+\section*{Future Work (projects)}
+\index{Projects}
+
+This document and the proof scripts are far from being complete.
+
+First, there must be a lot of typos to correct, references and index items to add. Many proofs are too complex and should be simplified, etc.
+
+The following extensions are planned, but help is needed:
+
+\begin{itemize}
+\item Semi automatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
+\item Extension to $\Gamma_0$ (in Veblen normal form)
+\item More lemmas about hierarchies of rapidly growing functions, and their relationship 
+    with primitive recursive functions and provability in Peano arithmetic 
+(following~\cite{KS81, KP82}).
+\item From \coq's point of view, this development could be used as an illustration of the evolution of the software, every time new libraries and sets of tactics could help to simplify the proofs.
+\end{itemize}
+
+\subsection*{Main References}
 
@@
 
-\paragraph*{Warning:}
-
-This document is \emph{not} an introductory text for \coq{}, and there are many aspects of this proof assistant that are not covered. 
- The reader should already have some basic experience with the \coq{} system.
-The Reference Manual and several tutorials are available on \coq{} page~\cite{Coq}.
-First chapters of textbooks like \emph{Interactive Theorem Proving and Program Development}~\cite{BC04}, \emph{Software Foundations}~\cite{SF} or  \emph{Certified Programming with Dependent Types} ~\cite{chlipalacpdt2011} will give you the right background.
-
-\subsection{State of the Development}
-The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on  notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
-We added new material : axiomatic definition of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, equations, etc.
-
-We are now working in order to make clumsy proofs more readable, simplify definitions, and ``factorize'' proofs as much as possible. 
-Many possible improvements are suggested as ``todo''s or ``projects'' in this text.
-
-\subsection{Contributions}
-Many thanks to \'Evelyne Contejean,  Florian Hatat and Pascal Manoury  for their contribution.
-\'Evelyne contributed libraries on the recursive path ordering (\emph{rpo}) for proving the well-foundedness of our representation of $\epsilon_0$ and $\Gamma_0$.
-Florian Hatat proved many useful lemmas on countable sets, which we used in our adaptation of Schütte's formalization of countable ordinals.
-Pascal Manoury is integrating the ordinal $\omega^\omega$ into our hierarchy of ordinal notations.
-
-\label{sec:orgheadline2}
-
-Any form of contribution  is welcome: correction of errors (typos and more serious mistakes), improvement of
-Coq scripts, proposition of inclusion of new chapters, and generally any
-comment or proposition that would help us. The text contains several \emph{projects} which, when completed, may improve the present work.
-Please do not hesitate to bring your contribution!
-
-\subsection{Typographical Conventions}
-
-Quotations from our  \coq{} source are displayed as follows:
-
-  \begin{Coqsrc}
- Require Import Arith.
-
- Definition square (n:nat) := n * n.
-
- Lemma square_double : exists n:nat, n + n = square n.
- Proof.
-    exists 2. 
-  \end{Coqsrc}
-
-Answers from \coq{} (including sub-goals, error messages, etc.) are displayed in slanted style
-with a different background color.
-
- \begin{Coqanswer}
- 1 subgoal, subgoal 1 (ID 5)
-
-  ============================
-   2 + 2 = square 2
-
- \end{Coqanswer}
-
- \begin{Coqsrc}
-   reflexivity.
-Qed.
- \end{Coqsrc}
-
-\subsection{Active Links}
-The  links which appear in this pdf  document lead are of three possible kinds of destination:
-\begin{itemize}
-\item Local links to the document itself,
-\item External links, mainly to \coq's page,
-\item Local links to pages generated by \texttt{coqdoc}. According to the current makefile (through the commands \texttt{make html} and \texttt{make pdf}), 
-  we assume that the page generated from a library \texttt{XXX/YYY.v} is stored as
-the relative address \texttt{../theories/html/hydras.XXX.YYY.html} (from the location of the pdf)
-Thus,  active links towards our \coq{} modules may be incorrect if you got this \texttt{pdf} document otherwise than by compiling the distribution available in
-\url{https://github.com/coq-community/hydra-battles}.
-
-\end{itemize}
-
-\subsection{Alternative or Bad Definitions}
-\label{sect:alt-proofs}
-Finally, we decided to include definitions or lemma statements, as well as tactics,  that lead to
-dead-ends or too complex developments, with the following coloring.
-Bad definitions 
- and encapsulation in modules called \texttt{Bad}, \texttt{Bad1}, etc.
-
-\begin{Coqbad}
-Require Import Lia.
-
-Module Bad.
-
-Definition double (n:nat)  := n + 2.
-
-Lemma lt_double : forall n:nat, n < double  n.
-Proof.
-   unfold double; lia.
-Qed.
-
-End Bad.
-\end{Coqbad}
-
-Likewise, alternative, but still unexplored definitions will be presented in modules
-\texttt{Alt}, \texttt{Alt1}, etc. Using these definitions is left as an implicit exercise.
-
-\begin{Coqalt}
-Require Import Arith Lia Iterates.
-Module Alt.
-  Definition double (n : nat) := iterate S n n.
-End Alt.
-\end{Coqalt}
-
-\begin{Coqsrc}
-Lemma alt_double_ok n : Nat.double n = Alt.double n.
-Proof.
-  unfold Alt.double, Nat.double; induction n; cbn.
-  -  trivial.
-  -  rewrite <- iterate_rw, iterate_S_eqn, <- IHn; lia.
-Qed.
-\end{Coqsrc}
-
-\section{Acknowledgements}
-\label{sec:orgheadline5}
-    Many thanks to David Ilcinkas, Sylvain Salvati, Alan Schmitt and Théo Zimmerman for their help on the elaboration of this document, and to the
- members of the \emph{Formal Methods} team at laBRI for their helpful comments 
-on an oral presentation of this work.
-
-Many thanks also to the Coq development team, Yves Bertot, and members of the \emph{Coq Club} for interesting discussions about the \coq{} system and the Calculus of Inductive Constructions.
-
-The author of the present document wishes to express his gratitude to the late Patrick Dehornoy, whose talk  was determinant for our desire to work on this topic.
-
-I owe my interest in discrete mathematics and their relation to formal proofs and functional programming  to Srecko Brlek.
-Equally, there is W.
-H.
-Burge's book ``\emph{Recursive Programming Techniques}'' ~\cite{burge} which was a great  source of inspiration.
-
-\section{How to Install the Libraries}
-\label{sec:orgheadline4}
-\begin{itemize}
-\item The present distribution has been checked with version 8.11.0 of the Coq proof assistant, with the plug-ins \texttt{coq-paramcoq} and \texttt{coq-equations}.
-
-\item Just go into the root directory, and type "\texttt{make}"
-\end{itemize}
-
-\section{Comments on Exercises and Projects}
-
-Although we do not plan to include complete solutions to the exercises, 
-we think it would be useful to include comments and hints, and questions/answers from the users. In constrast, ``projects'' are supposed, once completed, to be included in the repository.
-
-% \subsubsection{Links to the Coq source}
-
-\section{Future Work (projects)}
-\index{Projects}
-
-This document and the proof scripts are far from being complete.
-
-First, there must be a lot of typos to correct, references and index items to add. Many proofs are too complex and should be simplified, etc.
-
-The following extensions are planned, but help is needed:
-
-\begin{itemize}
-\item Semi automatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
-\item Extension to $\Gamma_0$ (in Veblen normal form)
-\item More lemmas about hierarchies of rapidly growing functions, and their relationship 
-    with primitive recursive functions and provability in Peano arithmetic 
-(following~\cite{KS81, KP82}).
-\item From \coq's point of view, this development could be used as an illustration of the evolution of the software, every time new libraries and sets of tactics could help to simplify the proofs.
-\end{itemize}
-
-%-----------------------------------------------------------
-
 \chapter{Hydras and Hydra Games}
+
 \label{sec:orgheadline91}
@@
 Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the 
-module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}).
+module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}).
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hydra}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Definitions.html\#Hydra}{Hydra.Hydra\_Definitions}}
 
@@
 
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#head}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Definitions.html\#head}{Hydra.Hydra\_Definitions}}
 
@@
 \vspace{4mm}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html\#Hy}{Hydra.Hydra\_Examples}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Examples.html\#Hy}{Hydra.Hydra\_Examples}}
 
@@
 
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#hcons_mult}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Definitions.html\#hcons_mult}{Hydra.Hydra\_Definitions}}
 
@@
 
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
 
@@
 \noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
 
@@
 \noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
 
@@
 
-\vspace{4pt}\emph{From Module\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
+\vspace{4pt}\emph{From Module\href{../theories/html/hydras.ordinals.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
 
@@
 \noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#S1}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Definitions.html\#S1}{Hydra.Hydra\_Definitions}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions}}
 
@@
 
-\vspace{4pt}\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}.
+\vspace{4pt}\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}.
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#round_n}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Definitions.html\#round_n}{Hydra.Hydra\_Definitions}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
 \label{types:Battle}
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#free}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Definitions.html\#free}{Hydra.Hydra\_Definitions}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#standard}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Definitions.html\#standard}{Hydra.Hydra\_Definitions}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}
 
@@
 % \vspace{4pt}
-% \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
+% \emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
 
@@
 
-The complete proof is in file \url{../theories/html/hydras.Hydra.BigBattle.html}.
+The complete proof is in file \url{../theories/html/hydras.ordinals.Hydra.BigBattle.html}.
 
@@
 The functional \texttt{iterate}, similar to Standard Library's \texttt{Nat.iter},
-is defined and studied in~\href{../theories/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
+is defined and studied in~\href{../theories/html/hydras.ordinals.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
 
@@
 Using some facts (proven in 
-\href{../theories/html/hydras.Hydra.BigBattle.html}{hydras.Hydra.BigBattle}),we get several  minorations.
+\href{../theories/html/hydras.ordinals.Hydra.BigBattle.html}{hydras.Hydra.BigBattle}),we get several  minorations.
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
 
@@
 The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
-\url{../theories/html/hydras.Hydra.Hydra_Theorems.html} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
+\url{../theories/html/hydras.ordinals.Hydra.Hydra_Theorems.html} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
 
@@
 
-Let us present the main steps of that proof, the script of which  is in the module ~\href{../theories/html/hydras.Hydra.Omega_Small.html}{Hydra/Omega\_Small.v} \footnote{ The name of this file means ``the ordinal $\omega$ is too small for proving the termination of [free] hydra battles ''.
+Let us present the main steps of that proof, the script of which  is in the module ~\href{../theories/html/hydras.ordinals.Hydra.Omega_Small.html}{Hydra/Omega\_Small.v} \footnote{ The name of this file means ``the ordinal $\omega$ is too small for proving the termination of [free] hydra battles ''.
 In effect, the elements of $\omega$, considered as a set, are just the natural numbers (see next chapter for more details)}.
@@
 
-More intuitively, let us have a look at Figure~\ref{fig:ordinal-sequence}. It presents a small sequence of ordinal numbers, which extends the sequence of natural numbers.
+For the computer scientist, ordinals are tools for proving the totality of a given recursive function, or termination of a transition system. \emph{Ordinal arithmetic} 
+provides a set of functions whose properties, like \emph{monotony}, allow to define \emph{variants}, \emph{i.e.} strictly decreasing measures used in proofs of termination.
+
+Let us have a look at Figure~\ref{fig:ordinal-sequence}. It presents a small sequence of ordinal numbers, which extends the sequence of natural numbers.
 
@@
 \item The ordinals are listed in a strictly increasing order. 
-\item Dots : ``$\ldots$'' stand for an infinite sequence of ordinals, not shown for lack of space.
+\item Dots : ``$\ldots$'' stand for  infinite sequences of ordinals, not shown for lack of space.
 For instance, the ordinal $42$ is not shown in the first line, but it exists, between $17$ and $\omega$.
@@
 \item[well foundedness]: $<$ is well-founded (every element of $A$ is accessible)\footnote{In classical mathematics, we would say that there is no infinite sequence $a_1>a_2> \dots a_n> a_{n+1}\dots$ in $A$.
-Please refer to any documenation on \coq{} for having more details on well-foundedness and accessibility.}.
-
+In contrast, \coq's standard library contains
+an inductive definition of a predicate \texttt{Acc} which allows us to write 
+constructive proofs of accessibility (See \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}).}.
 \end{description}
@@
 \begin{align*}
-(i,j) < (i,k) & \;\textit{if} \; j < k\\
-(0,k) < (1,l) & \;\textit{for\,any}\;k \;\textit{and} \; l
+(i,j) < (i,k) & \;\textit{\textbf{if} }\; j < k\\
+(0,k) < (1,l) & \;\textit{\textbf{for\,any}}\;k \;\textit{\textbf{and}} \; l
 \end{align*}
@@
 we derive some interesting properties of countable ordinals from these axioms.
-It is interesting to compare proofs of a given property (for instance the associativity of addition) in the computatuinal framework of some ordinal notation, and in the axiomatic model of Schütte.
+It is interesting to compare proofs of a given property (for instance the associativity of addition) in the computational framework of some ordinal notation, and in the axiomatic model of Schütte.
 
@@
 \vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}}
+\noindent\emph{From Library~\href{../theories/html/hydras.ordinals.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}}
 
@@
 \vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations/ON_Omega.html}{OrdinalNotations.ON\_Omega}}
+\noindent\emph{From Library~\href{../theories/html/hydras.ordinals.OrdinalNotations/ON_Omega.html}{OrdinalNotations.ON\_Omega}}
 
@@
 \vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations/ON_plus.html}{OrdinalNotations.ON\_plus}}
+\noindent\emph{From Library~\href{../theories/html/hydras.ordinals.OrdinalNotations/ON_plus.html}{OrdinalNotations.ON\_plus}}
 
@@
 \vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
+\noindent\emph{From Module~\href{../theories/html/hydras.ordinals.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
 
@@
 
-The following definitions are in Library \href{../theories/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
+The following definitions are in Library \href{../theories/html/hydras.ordinals.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
 
@@
 
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
+\noindent\emph{From Module~\href{../theories/html/hydras.ordinals.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
 
@@
 
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}}
+\noindent\emph{From Module~\href{../theories/html/hydras.ordinals.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}}
 \begin{Coqsrc}
@@
 \noindent
-\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_mult.html}{OrdinalNotations.ON\_mult}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.OrdinalNotations.ON_mult.html}{OrdinalNotations.ON\_mult}}
 
@@
 \vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
+\noindent\emph{From Module~\href{../theories/html/hydras.ordinals.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
 
@@
 
-The detailed  proof script is in the file \url{../theories/html/hydras.Hydra.Omega2_Small.html}.
+The detailed  proof script is in the file \url{../theories/html/hydras.ordinals.Hydra.Omega2_Small.html}.
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
 
@@
 \vspace{4pt}
-\emph{From Module ~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
+\emph{From Module ~\href{../theories/html/hydras.ordinals.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}
 
@@
 \noindent
-In \coq (Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#Impossible}{Hydra.Omega2\_Small}):
+In \coq (Module~\href{../theories/html/hydras.ordinals.Hydra.Omega2_Small.html\#Impossible}{Hydra.Omega2\_Small}):
 
@@
 \vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
+\noindent\emph{From Module~\href{../theories/html/hydras.ordinals.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
 
@@
 In our formalization, the interpretation of an ordinal as a set is realized by the following definition
-(in ~\href{../theories/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}).
+(in ~\href{../theories/html/hydras.ordinals.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}).
 
@@
 
-This exercise is partially solved for $i=3$ (in ~\href{../theories/html/hydras.OrdinalNotations.Example_3PlusOmega.html}{OrdinalNotations.Example\_3PlusOmega}).
+This exercise is partially solved for $i=3$ (in ~\href{../theories/html/hydras.ordinals.OrdinalNotations.Example_3PlusOmega.html}{OrdinalNotations.Example\_3PlusOmega}).
 
@@
 in Veblen normal form (see 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}{Gamma0.Gamma0}) are shown to be ordinal notation systems, but there is a lot of work to be done in order to unify ad-hoc  definitions and proofs which were written before the definition of the \texttt{ON} type class.
+\href{../theories/html/hydras.ordinals.Gamma0.Gamma0.html}{Gamma0.Gamma0}) are shown to be ordinal notation systems, but there is a lot of work to be done in order to unify ad-hoc  definitions and proofs which were written before the definition of the \texttt{ON} type class.
 \end{remark}
@@
 \vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#T1}{Epsilon0.T1}}
+\noindent\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.T1.html\#T1}{Epsilon0.T1}}
 
@@
 \emph{Please look also at the library
-\href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2.html}}.
+\href{../theories/html/hydras.ordinals.Gamma0.T2.html}{Gamma0.T2.html}}.
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.T1.html}{Epsilon0.T1}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare}{Epsilon0.T1}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.T1.html\#compare}{Epsilon0.T1}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare_refl}{Epsilon0.T1}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.T1.html\#compare_refl}{Epsilon0.T1}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#lt_irrefl}{Epsilon0.T1}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.T1.html\#lt_irrefl}{Epsilon0.T1}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#nf_b}{Epsilon0.T1}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.T1.html\#nf_b}{Epsilon0.T1}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Prelude.Restriction.html}{Ordinals.Prelude.Restriction}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Prelude.Restriction.html}{Ordinals.Prelude.Restriction}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#LT}{Epsilon0.T1}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.T1.html\#LT}{Epsilon0.T1}}
 
@@
 
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.E0.html}{Epsilon0.E0}}
 
@@
  \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.T1.html\#succb}{Epsilon0.T1}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.T1.html\#succ}{Epsilon0.T1}}
 
@@
 and natural induction on \texttt{n}, that (\texttt{ocons alpha n beta}) is accessible.
-The proof, quite long, can be consulted in \url{../theories/html/hydras.Epsilon0.T1.html}
+The proof, quite long, can be consulted in \url{../theories/html/hydras.ordinals.Epsilon0.T1.html}
 
@@
 \label{remark:a3pat}
-The alternate proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
+The alternate proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the library \href{../theories/html/hydras.ordinals.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
  \end{remark}
@@
 
-\emph{From Module~\href{../theories/html/hydras.Schutte.Schutte.Correctness_E0.html}{Schutte.Schutte.Correctness\_E0}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Schutte.Schutte.Correctness_E0.html}{Schutte.Schutte.Correctness\_E0}}
 
@@
 \textbf{Hint}. You may use the following definition (in 
-\href{../theories/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}).
+\href{../theories/html/hydras.ordinals.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}).
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Termination.html\#m}{Hydra.Hydra\_Termination}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Termination.html\#m}{Hydra.Hydra\_Termination}}
 
@@
 then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1})$ in one round with
-the replication factor $i$ (Lemma \href{../theories/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
+the replication factor $i$ (Lemma \href{../theories/html/hydras.ordinals.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
 \item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) battle from $\iota(\alpha)$ to $\iota(\beta)$.
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS}{Epsilon0.Canon}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.Canon.html\#canonS}{Epsilon0.Canon}}
 
@@
 abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
-~\href{../theories/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}
+~\href{../theories/html/hydras.ordinals.Epsilon0.Canon.html}{Epsilon0.Canon}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.Paths.html}{Epsilon0.Paths}}
 
@@
 
-\emph{From Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.O2H.html}{Hydra.O2H}}
 
@@
 
-The following lemma (proved in ~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H.v}) maps  canonical sequences to rounds of hydra battles.
+The following lemma (proved in ~\href{../theories/html/hydras.ordinals.Hydra.O2H.html}{Hydra.O2H.v}) maps  canonical sequences to rounds of hydra battles.
 
@@
 
-In the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
+In the module \href{../theories/html/hydras.ordinals.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
 This part of the development is parameterized by some class $B$ of battles, which will be instantiated later to \texttt{free} or \texttt{standard}.
@@
 Let us now consider that $B$ is instantiated to \texttt{free} (which means that we are considering proofs of termination of \emph{all} battles).
-The following lemmas are proved in Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
+The following lemmas are proved in Module~\href{../theories/html/hydras.ordinals.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
 The case $B=\texttt{standard}$ is studied in section~\vref{std-case}.
@@
 One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
-Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}.
+Module~\href{../theories/html/hydras.ordinals.Hydra.O2H.html}{Hydra.O2H}.
 
@@
 It is a straightforward application of \texttt{LT\_path\_to} of
-Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}:
+Module~\href{../theories/html/hydras.ordinals.Epsilon0.Paths.html}{Epsilon0.Paths}:
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.KS.html}{Epsilon0.KS}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.KS.html}{Epsilon0.KS}}
 
@@
 be consulted in the module 
-\url{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}.
+\url{../theories/html/hydras.ordinals.Hydra.Epsilon0_Needed_Std.html}.
 Please note that it has the same global structure as in section\ref{sec:free-battles-case} 
@@
 Applying the  lemmas  \texttt{Lemma2\_6\_1} of the module 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
+\href{../theories/html/hydras.ordinals.Epsilon0.Paths.html\#Lemma2_6_1}%
 {\texttt{Lemma2\_6\_1}}   and 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
+\href{../theories/html/hydras.ordinals.Epsilon0.Paths.html\#constant_to_standard_path}%
 {\texttt{constant\_to\_standard\_path}},
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html\#LT_to_standard_battle}{Hydra.Epsilon0\_Needed\_Std}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Hydra.Epsilon0_Needed_Std.html\#LT_to_standard_battle}{Hydra.Epsilon0\_Needed\_Std}}
 
@@
 from 
-the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}.
+the module \href{../theories/html/hydras.ordinals.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}.
 
@@
 Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
-(please see also \href{../theories/html/hydras.Schutte.Correctness_E0.html}{the module \texttt{Ordinals.Schutte.Correctness\_E0}}).
+(please see also \href{../theories/html/hydras.ordinals.Schutte.Correctness_E0.html}{the module \texttt{Ordinals.Schutte.Correctness\_E0}}).
 
@@
 \noindent
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}
+\emph{From Module~ \href{../theories/html/hydras.ordinals.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}
 
@@
 \noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}
 
@@
  \noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html}{Epsilon0.Large\_Sets\_Examples}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.Large_Sets_Examples.html}{Epsilon0.Large\_Sets\_Examples}}
 \begin{Coqsrc}
@@
 \noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
 \begin{Coqsrc}
@@
 In order to ensure this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <<= f}, or \texttt{(fun\_le S f)} (defined
-in~\href{../theories/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}).
+in~\href{../theories/html/hydras.ordinals.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}).
 
@@
 
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#L_omega_mult}{Epsilon0.Large\_Sets}}
+\emph{From Module~ \href{../theories/html/hydras.ordinals.Epsilon0.Large_Sets.html\#L_omega_mult}{Epsilon0.Large\_Sets}}
 
@@
 \noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#L_}{L\_alpha}).}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.L_alpha.html\#L_}{L\_alpha}).}
 
@@
 Theorem \texttt{battle\_length\_std } of
-Module~\href{../theories/html/hydras.Hydra.Hydra_Theorems.html\#battle_length_std}{Hydra.Hydra\_Theorems} relates the length of standard battles with the functions $L_\alpha$.
+Module~\href{../theories/html/hydras.ordinals.Hydra.Hydra_Theorems.html\#battle_length_std}{Hydra.Hydra\_Theorems} relates the length of standard battles with the functions $L_\alpha$.
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.H_alpha.html\#H_}{Epsilon0.H\_alpha}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.H_alpha.html\#H_}{Epsilon0.H\_alpha}}
 
@@
 
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#H_L_}{Epsilon0.L\_alpha}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.L_alpha.html\#H_L_}{Epsilon0.L\_alpha}}
 
@@
 
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0.L\_alpha}}
+\emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.L_alpha.html}{Epsilon0.L\_alpha}}
 
@@
 A first attempt is to write a definition of $F_\alpha$ by equations, in the same as for $H\_alpha$.  We use the functional \texttt{iterate} defined in 
-Module~\href{../theories/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
+Module~\href{../theories/html/hydras.ordinals.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
 
@@
 The following code comes from 
- \url{../theories/html/hydras.Epsilon0.F_alpha.html}.
+ \url{../theories/html/hydras.ordinals.Epsilon0.F_alpha.html}.
 
@@
 % \noindent
-% \emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets_Demo.html}{Ordinals.Epsilon0.Large\_Sets\_Demo}}
+% \emph{From Module~\href{../theories/html/hydras.ordinals.Epsilon0.Large_Sets_Demo.html}{Ordinals.Epsilon0.Large\_Sets\_Demo}}
 
@@
  Unlike the rest of our libraries, our library
-\href{../theories/html/hydras.Schutte.Schutte.html}{Ordinals.Schutte}
+\href{../theories/html/hydras.ordinals.Schutte.Schutte.html}{Ordinals.Schutte}
 is not constructive, and relies on several axioms.
@@
 Both operators allow us to write definitions close to the natural mathematical language, such as ``$\textrm{succ}$ is \emph{the} least ordinal strictly greater than $\alpha$''
-\item Please note that only the library \href{../theories/html/hydras.Schutte.Schutte.html}{Schutte/*.v} is ``contaminated'' by axioms, and that the rest of our libraries remain constructive.
+\item Please note that only the library \href{../theories/html/hydras.ordinals.Schutte.Schutte.html}{Schutte/*.v} is ``contaminated'' by axioms, and that the rest of our libraries remain constructive.
 \end{itemize}
@@
 A set $A$ is countable if there is an injective function from $A$ to $\mathbb{N}$ (see
-Library \href{../theories/html/hydras.Schutte.Countable.html}%
+Library \href{../theories/html/hydras.ordinals.Schutte.Countable.html}%
 {\texttt{Schutte.Countable}}).
@@
 
-\emph{From Module\href{../theories/html/hydras.Schutte.Schutte_basics.html}%
+\emph{From Module\href{../theories/html/hydras.ordinals.Schutte.Schutte_basics.html}%
 {\texttt{Schutte.Schutte\_basics}}}
@@
 \texttt{ordinal} (The  class \texttt{WO} is defined in
-Module~\href{../theories/html/hydras.Schutte.Well_Orders.html}%
+Module~\href{../theories/html/hydras.ordinals.Schutte.Well_Orders.html}%
 {\texttt{Well\_Orders.v}}).
@@
 Indeed, the operators \texttt{epsilon} and \texttt{iota} allowed us to make our definitions 
-quite close to Schütte's text. Our libraries \href{../theories/html/hydras.Schutte.MoreEpsilonIota.html}%
+quite close to Schütte's text. Our libraries \href{../theories/html/hydras.ordinals.Schutte.MoreEpsilonIota.html}%
 {\texttt{Schutte.MoreEpsilonIota}}
 and
-\href{../theories/html/hydras.Schutte.PartialFun.html}%
+\href{../theories/html/hydras.ordinals.Schutte.PartialFun.html}%
 {\texttt{Schutte.PartialFun}} are extensions of \texttt{Coq.logic.Epsilon} for making easier 
@@
 
-From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
+From Module \href{../theories/html/hydras.ordinals.Schutte.Schutte_basics.html}%
 {\texttt{~Schutte.Schutte\_basics}}
@@
 (proved in Module 
-\href{../theories/html/hydras.Schutte.Schutte_basics.html\#finite_lt_omega}{\texttt{Schutte.Schutte\_basics}})
+\href{../theories/html/hydras.ordinals.Schutte.Schutte_basics.html\#finite_lt_omega}{\texttt{Schutte.Schutte\_basics}})
 
@@
 Complete definitions and proofs can be found in Module
- \href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
+ \href{../theories/html/hydras.ordinals.Schutte.Ordering_Functions.html}%
 {\texttt{Schutte.Ordering\_Functions}} ).
@@
 One of the main theorems of
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html\#ordering_function_ex}%
+\href{../theories/html/hydras.ordinals.Schutte.Ordering_Functions.html\#ordering_function_ex}%
 {\texttt{Ordering\_Functions}} 
@@
 Let us quote the following theorems (see Library
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
+\href{../theories/html/hydras.ordinals.Schutte.Ordering_Functions.html}%
 {\texttt{Schutte.Ordering\_Functions}} for more details).
@@
 The following definitions and proofs can be consulted in Module
-\href{../theories/html/hydras.Schutte.Addition.html}%
+\href{../theories/html/hydras.ordinals.Schutte.Addition.html}%
 {\texttt{Schutte.Addition.v}}.
@@
 computational proofs of the corresponding statements in Module
-\href{../theories/html/hydras.Epsilon0.T1.html}%
+\href{../theories/html/hydras.ordinals.Epsilon0.T1.html}%
 {\texttt{Epsilon0.T1}}. 
@@
 
-\noindent\emph{From Module \href{../theories/html/hydras.Schutte.AP.html}%
+\noindent\emph{From Module \href{../theories/html/hydras.ordinals.Schutte.AP.html}%
 {\texttt{Schutte.AP}}}
@@
 We prove the inequality  $\alpha<\omega^\alpha$ whenever $\alpha < \epsilon_0$.
-\emph{Note that this condition was implicit in Module \href{../theories/html/hydras.Epsilon0/T1.html\#lt_phi0}%
+\emph{Note that this condition was implicit in Module \href{../theories/html/hydras.ordinals.Epsilon0/T1.html\#lt_phi0}%
 {\texttt{Epsilon0.T1}}.}
@@
 This definition is translated in \coq{} in 
-Module \href{../theories/html/hydras.Schutte.Critical.html}%
+Module \href{../theories/html/hydras.ordinals.Schutte.Critical.html}%
 {\texttt{Schutte.Critical}}, as the least fixpoint of a functional.
@@
 
-\noindent\emph{From \href{../theories/html/hydras.Schutte.CNF.html\#cnf_t}%
+\noindent\emph{From \href{../theories/html/hydras.ordinals.Schutte.CNF.html\#cnf_t}%
 {\texttt{Schutte.CNF}}}
@@
 \begin{remark}
-The sub-directory \href{../theories/html/hydras.Gamma0.html}%
+The sub-directory \href{../theories/html/hydras.ordinals.Gamma0.html}%
 {\texttt{theories/Gamma0}} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form. 
@@
 Our library 
-\href{../theories/html/hydras.Schutte.correctness_E0.html}%
+\href{../theories/html/hydras.ordinals.Schutte.correctness_E0.html}%
 {\texttt{Schutte.Correctness\_E0}} establishes the link between two very different modelizations of ordinal numbers. In other words, it ``validates'' a data structure in terms of
@@
 \subsection{Remarks}
-Let us recall that the library \href{../theories/html/hydras.Schutte.Schutte.html}%
+Let us recall that the library \href{../theories/html/hydras.ordinals.Schutte.Schutte.html}%
 {\texttt{Schutte}} depends on five \emph{axioms} and lies explicitly in the  
@@
 Nevertheless, the other modules:
-\href{../theories/html/hydras.Epsilon0.Epsilon0.html}%
+\href{../theories/html/hydras.ordinals.Epsilon0.Epsilon0.html}%
 {\texttt{Epsilon0}},
-\href{../theories/html/hydras.Hydra.Hydra.html}%
+\href{../theories/html/hydras.ordinals.Hydra.Hydra.html}%
 {\texttt{Hydra}}, et 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}%
+\href{../theories/html/hydras.ordinals.Gamma0.Gamma0.html}%
 {\texttt{Gamma0}}
@@
 
-\noindent\emph{From \href{../theories/html/hydras.Schutte.Critical.html\#strongly_critical}%
+\noindent\emph{From \href{../theories/html/hydras.ordinals.Schutte.Critical.html\#strongly_critical}%
 {\texttt{Schutte.Critical}}}
@@
 
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T2}%
+\noindent\emph{From \href{../theories/html/hydras.ordinals.Gamma0.T2.html\#T2}%
 {\texttt{Gamma0.T2}}}
@@
 
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0\#tricho_aux}%
+\noindent\emph{From \href{../theories/html/hydras.ordinals.Gamma0.Gamma0\#tricho_aux}%
 {\texttt{Gamma0.Gamma0}}}
@@
 
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#succ}%
+\noindent\emph{From \href{../theories/html/hydras.ordinals.Gamma0.T2.html\#succ}%
 {\texttt{Gamma0.T2}}}
@@
 
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T1_to_T2}%
+\noindent\emph{From \href{../theories/html/hydras.ordinals.Gamma0.T2.html\#T1_to_T2}%
 {\texttt{Gamma0.T2}}}
@@
 \end{project}
-
-\chapter{Appendices}
-
-\bibliographystyle{alpha}
-\bibliography{thebib}
-
-\section{Index}
-
-\begin{table}[h]
-  \centering
-  \begin{threeparttable}
-    \caption{Main types and type classes}
-\begin{tabular}{|c|c|c|l|}
-\hline
-Name & Description& Page \\\hline
-Battle & type class of battle types & \pageref{types:Battle} \\
-E0 & (well formed) ordinals below $\epsilon_0$ & \pageref{types:E0}\\
-Hvariant & variants for proving termination of hydra battles & \pageref{sect:hvariant-def} \\
-Hydra & tree-like representation of hydras & \pageref{types:Hydra} \\
-Hydrae & finite sequences of hydras & \pageref{types:Hydrae} \\
-ON & ordinal  notations & \pageref{types:ON} \\
-ON\_for & comparison of an ordinal notation with Schütte's model &  \pageref{types:ON-for}\\
-ON\_iso & isomorphism of ordinal notations  & \pageref{types:ON-iso}\\
-
-Ord & countable ordinals (after Schütte) & \pageref{types:Ord} \\
-ppT1 & pretty printed version of T1 & \pageref{types:ppT1} \\
-SubON & comparison of ordinal notations & \pageref{types:SubON} \\
-T1 & ordinal terms below $\epsilon_0$ & \pageref{types:T1} \\
-T2 & ordinal terms below $\Gamma_0$ & \pageref{types:T2} \\
-WO & well orders (Schütte's definition) & \pageref{types:WO} \\
-\hline
-\end{tabular}
-\end{threeparttable}
-\end{table}
-
-\begin{table}[h]
-  \centering
-  \begin{threeparttable}
-    \caption{Main functions and constants}
-\begin{tabular}{|r | c|c|c|c|l|}
-\hline
-Name & Gallina & Math & Description & Page \\ \hline
-canon & canon alpha i & $\canonseq{\alpha}{i}$& Canonical sequence & \pageref{Functions:canon}\\
-canonS & canonS alpha i & $\canonseq{\alpha}{i+1}$& Helper for canon & \pageref{Functions:canonS}\\
-F\_ & F\_ alpha n & $F_\alpha(n)$ & Wainer's \texttt{F} fast growing hierarchy & \pageref{Functions:F-alpha} \\
-H\_ & H\_ alpha n & $H_\alpha(n)$ & Hardy's \texttt{H} fast growing hierarchy & \pageref{Functions:H-alpha} \\
-iterate & iterate $f$ $n$ $x$ & $f^{(n)}(x)$ & Functional iteration & \pageref{Functions:iterate} \\
-L\_ & L\_ alpha k & $L_\alpha(k)$ & final step of a standard path & \pageref{Functions:L-alpha} \\
-succ & succ alpha & & Successor & \pageref{Functions:succ-T1}, \pageref{Functions:succ-sch} \dots \\
-zero:Ord & & $0$ & The least ordinal (Schütte's model) & \pageref{Constants:zero:Ord} \\
-
-\hline
-\end{tabular}
-\end{threeparttable}
-\end{table}
-
-\begin{table}[h]
-  \centering
-  \begin{threeparttable}
-    \caption{Main predicates}
-\begin{tabular}{|r | c|c|c|c|l|}
-\hline
-Name & Gallina&Math& Description& Page \\\hline
-\texttt{lt : T1->T1->Prop}& lt alpha beta & $\alpha < \beta$& strict order on type \texttt{T1} \tnote{1} & \pageref{Predicates:lt-T1}\\
-\texttt{LT: T1->T1->Prop}& alpha o< beta & $\alpha < \beta$& strict order on type \texttt{T1}   \tnote{2} & \pageref{Predicates:LT-T1}\\
-\texttt{Lt : E0->E0->Prop} & alpha o< beta & $\alpha < \beta$& strict order on type \texttt{E0} \tnote{3} & \pageref{Predicates:Lt-E0} \\
-\texttt{nf: T1->Prop} & \texttt{nf alpha} && alpha is in Cantor normal form & \pageref{Predicates:nf-T1}\\
-\hline
-\end{tabular}
-\begin{tablenotes}
-  \item[1] This order is total, but not well-founded, because of not well formed terms.
-\item[2] Restriction of \texttt{lt} to terms in normal form; this order is partial, but well-founded.
-\item[3] This order is total \emph{and} well-founded.
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\vspace{4pt}
-\begin{table}[h]
-  \centering
-  \begin{threeparttable}
-    \caption{Infix notations}
-\begin{tabular}{|c|c|c|c|l|}
-\hline
-Name & Gallina&Math& Description& Page \\\hline
- \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality \tnote{1} & \pageref{sect:on-lt-notation}\\
- \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
-\texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & ordinal addition & \pageref{sect:infix-plus-T1}, \dots\\
-\texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\
-\texttt{round} & \texttt{h -1-> h'} & & one round of a battle & \pageref{sect:infix-round} \\
-\texttt{rounds} & \texttt{h -+-> h'} & & one or more  rounds of a battle & \pageref{sect:infix-rounds} \\
-\texttt{round\_star} & \texttt{h -*-> h'} & & any number of rounds of a battle & \pageref{sect:infix-rounds} \\
-
-\hline
-\end{tabular}
-
-\begin{tablenotes}
- \item [1]
-Some notations may belong to several scopes. For instance, ``\texttt{o<}'' is
-bound in \texttt{ON\_scope}, \texttt{E0\_scope}, \texttt{t1\_scope}, etc., and locally in several libraries.
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\begin{table}[h]
-  \centering
-  \begin{threeparttable}
-    \caption{Other notations}
-\begin{tabular}{|c|c|c|c|l|}
-\hline
-Name & Gallina&Math& Description& Page \\\hline
-F & \texttt{F $n$} & $n$ & The $n$-th finite ordinal &  
-\pageref{sect:notation-F}, \pageref{sect:notation-F-sch}\\ 
-FS & \texttt{FS $n$} & $n+1$ & The $n+1$-th finite ordinal  \tnote{2} &  
-\pageref{sect:notation-FS}\\ 
-omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega}, \pageref{sect:omega-T1}, \pageref{sect:omega-notation2}, \dots\\
-phi0     & \texttt{phi0 alpha} & $\phi_0(\alpha),\; \omega^\alpha$&exponential of base $\omega$ & \pageref{sect:notation-phi0}\\
-
-\hline
-\end{tabular}
-
-\begin{tablenotes}
-  \item [2] Note that there exist also various coercions from \texttt{nat} to types of ordinal.
-Depending on the current scope and  \coq's syntactic analysis algorithm, \texttt{F} may be left implicit.
-\end{tablenotes}
-  \end{threeparttable}
-
-\end{table}
-
-\printindex
-
-% \section{Main notations}
-
-% \subsubsection*{Intervals}
-
-% Let $\leq$ be a total order on a type $A$. The intervals $\{k\in A| i\leq k\leq j\}$ is written $[i,j]$. Likewise, $\{k\in A| i\leq k<j\}$ is written
-% $[i,j[$.
-
-\end{document}

### 2020-11-24 doc/part-hydras.tex
simplifying the namespace

--- 2020-11-17 doc/hydras.tex
@@
 presented for instance in~\cite{Promel2013}. 
-\index{Maths!Hardy hierarchy}
+\index{Maths!Hardy Hierarchy}
 
@@
 For instance, the following theorem states that, if $\gamma$ is the sum of a limit ordinal $\beta$ and a finite ordinal $n$, and $\beta$ is a fixpoint of
-$\phi_\alpha$, then $\psi(\alpha,\gamma)=\phi_\alpha(\gamma+1)$.
+$\phi(\alpha)$, then $\psi(\alpha,\gamma)=\phi_\alpha(\gamma+1)$.
 
@@
 \end{project}
-
-\chapter{Appendices}
-
-\bibliographystyle{alpha}
-\bibliography{thebib}
-
-\section{Index}
-
-\begin{table}[h]
-  \centering
-  \begin{threeparttable}
-    \caption{Main types and type classes}
-\begin{tabular}{|c|c|c|l|}
-\hline
-Name & Description& Page \\\hline
-Battle & type class of battle types & \pageref{types:Battle} \\
-E0 & (well formed) ordinals below $\epsilon_0$ & \pageref{types:E0}\\
-Hvariant & variants for proving termination of hydra battles & \pageref{sect:hvariant-def} \\
-Hydra & tree-like representation of hydras & \pageref{types:Hydra} \\
-Hydrae & finite sequences of hydras & \pageref{types:Hydrae} \\
-ON & ordinal  notations & \pageref{types:ON} \\
-ON\_for & comparison of an ordinal notation with Schütte's model &  \pageref{types:ON-for}\\
-ON\_iso & isomorphism of ordinal notations  & \pageref{types:ON-iso}\\
-
-Ord & countable ordinals (after Schütte) & \pageref{types:Ord} \\
-ppT1 & pretty printed version of T1 & \pageref{types:ppT1} \\
-SubON & comparison of ordinal notations & \pageref{types:SubON} \\
-T1 & ordinal terms below $\epsilon_0$ & \pageref{types:T1} \\
-T2 & ordinal terms below $\Gamma_0$ & \pageref{types:T2} \\
-WO & well orders (Schütte's definition) & \pageref{types:WO} \\
-\hline
-\end{tabular}
-\end{threeparttable}
-\end{table}
-
-\begin{table}[h]
-  \centering
-  \begin{threeparttable}
-    \caption{Main functions and constants}
-\begin{tabular}{|r | c|c|c|c|l|}
-\hline
-Name & Gallina & Math & Description & Page \\ \hline
-canon & canon alpha i & $\canonseq{\alpha}{i}$& canonical sequence & \pageref{Functions:canon}\\
-canonS & canonS alpha i & $\canonseq{\alpha}{i+1}$& helper for canon & \pageref{Functions:canonS}\\
-F\_ & F\_ alpha n & $F_\alpha(n)$ & Wainer's \texttt{F} fast growing hierarchy & \pageref{Functions:F-alpha} \\
-H\_ & H\_ alpha n & $H_\alpha(n)$ & Hardy's \texttt{H} fast growing hierarchy & \pageref{Functions:H-alpha} \\
-iterate & iterate $f$ $n$ $x$ & $f^{(n)}(x)$ & functional iteration & \pageref{Functions:iterate} \\
-L\_ & L\_ alpha k & $L_\alpha(k)$ & final step of a standard path & \pageref{Functions:L-alpha} \\
-succ & succ alpha & & successor & \pageref{Functions:succ-T1}, \pageref{Functions:succ-sch} \dots \\
-zero:Ord & & $0$ & the least ordinal (Schütte's model) & \pageref{Constants:zero:Ord} \\
-
-\hline
-\end{tabular}
-\end{threeparttable}
-\end{table}
-
-\begin{table}[h]
-  \centering
-  \begin{threeparttable}
-    \caption{Main predicates}
-\begin{tabular}{|r | c|c|c|c|l|}
-\hline
-Name & Gallina&Math& Description& Page \\\hline
-\texttt{lt : T1->T1->Prop}& lt alpha beta & $\alpha < \beta$& strict order on type \texttt{T1} \tnote{1} & \pageref{Predicates:lt-T1}\\
-\texttt{LT: T1->T1->Prop}& alpha o< beta & $\alpha < \beta$& strict order on type \texttt{T1}   \tnote{2} & \pageref{Predicates:LT-T1}\\
-\texttt{Lt : E0->E0->Prop} & alpha o< beta & $\alpha < \beta$& strict order on type \texttt{E0} \tnote{3} & \pageref{Predicates:Lt-E0} \\
-\texttt{nf: T1->Prop} & \texttt{nf alpha} && alpha is in Cantor normal form & \pageref{Predicates:nf-T1}\\
-\hline
-\end{tabular}
-\begin{tablenotes}
-  \item[1] This order is total, but not well-founded, because of not well formed terms.
-\item[2] Restriction of \texttt{lt} to terms in normal form; this order is partial, but well-founded.
-\item[3] This order is total \emph{and} well-founded.
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\vspace{4pt}
-\begin{table}[h]
-  \centering
-  \begin{threeparttable}
-    \caption{Infix notations}
-\begin{tabular}{|c|c|c|c|l|}
-\hline
-Name & Gallina&Math& Description& Page \\\hline
- \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality \tnote{1} & \pageref{sect:on-lt-notation}\\
- \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
-\texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & ordinal addition & \pageref{sect:infix-plus-T1}, \dots\\
-\texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\
-\texttt{round} & \texttt{h -1-> h'} & & one round of a battle & \pageref{sect:infix-round} \\
-\texttt{rounds} & \texttt{h -+-> h'} & & one or more  rounds of a battle & \pageref{sect:infix-rounds} \\
-\texttt{round\_star} & \texttt{h -*-> h'} & & any number of rounds of a battle & \pageref{sect:infix-rounds} \\
-
-\hline
-\end{tabular}
-
-\begin{tablenotes}
- \item [1]
-Some notations may belong to several scopes. For instance, ``\texttt{o<}'' is
-bound in \texttt{ON\_scope}, \texttt{E0\_scope}, \texttt{t1\_scope}, etc., and locally in several libraries.
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\begin{table}[h]
-  \centering
-  \begin{threeparttable}
-    \caption{Other notations}
-\begin{tabular}{|c|c|c|c|l|}
-\hline
-Name & Gallina&Math& Description& Page \\\hline
-F & \texttt{F $n$} & $n$ & The $n$-th finite ordinal &  
-\pageref{sect:notation-F}, \pageref{sect:notation-F-sch}\\ 
-FS & \texttt{FS $n$} & $n+1$ & The $n+1$-th finite ordinal  \tnote{2} &  
-\pageref{sect:notation-FS}\\ 
-omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega}, \pageref{sect:omega-T1}, \pageref{sect:omega-notation2}, \dots\\
-phi0     & \texttt{phi0 alpha} & $\phi_0(\alpha),\; \omega^\alpha$&exponential of base $\omega$ & \pageref{sect:notation-phi0}\\
-
-\hline
-\end{tabular}
-
-\begin{tablenotes}
-  \item [2] Note that there exist also various coercions from \texttt{nat} to types of ordinal.
-Depending on the current scope and  \coq's syntactic analysis algorithm, \texttt{F} may be left implicit.
-\end{tablenotes}
-  \end{threeparttable}
-
-\end{table}
-
-\printindex
-
-% \section{Main notations}
-
-% \subsubsection*{Intervals}
-
-% Let $\leq$ be a total order on a type $A$. The intervals $\{k\in A| i\leq k\leq j\}$ is written $[i,j]$. Likewise, $\{k\in A| i\leq k<j\}$ is written
-% $[i,j[$.
-
-\end{document}

### 2020-11-25 doc/hydras.tex
modifications in doc and README

--- 2020-11-15 doc/hydras.tex
@@
 \usepackage{dsfont}
+\usepackage{float}
 \usepackage{threeparttable}
@@
 \label{sec:orgheadline5}
-    Many thanks to Yves Bertot, David Ilcinkas, Sylvain Salvati, Alan Schmitt and Théo Zimmerman for their help on the elaboration of this document, and to the
+    Many thanks to Yves Bertot, David Ilcinkas, Karl Palmskog, Sylvain Salvati, Alan Schmitt and Théo Zimmerman for their help on the elaboration of this document, and to the
  members of the \emph{Formal Methods} team at laBRI for their helpful comments 
@@
 
-\section{Index}
+\chapter*{Index and tables}
 
-\begin{table}[h]
+\begin{todo}
+ Still very incomplete!
+\end{todo}
+
+\printindex
+
+\section{Hydras and Ordinals}
+
+\begin{table}[H]
   \centering
@@
 
-\begin{table}[h]
+\begin{table}[H]
   \centering
@@
 \vspace{4pt}
-\begin{table}[h]
+\begin{table}[H]
   \centering
@@
 
-\printindex
+\section{Addition Chains}
 
-% \section{Main notations}
-
-% \subsubsection*{Intervals}
-
-% Let $\leq$ be a total order on a type $A$. The intervals $\{k\in A| i\leq k\leq j\}$ is written $[i,j]$. Likewise, $\{k\in A| i\leq k<j\}$ is written
-% $[i,j[$.
+\begin{todo}
+\end{todo}
 

### 2020-12-03 doc/chapter-powers.tex
work on addition chains chapter
add example on Fibonacci

@@
 
-Matrix exponentiation is a well-known method for computing Fibonacci numbers:
+\subsection{Computing Fibonacci Numbers}
+
+The sequence of Fibonacci numbers is defined by the following equations:
+
+\begin{align}
+F_0 & = 1 \\
+F_1 & = 1 \\
+F_n & = F_{n-1} + F_{n-2} \quad (n \geq 2)
+\end{align}
+
+In \coq{}, one can define this function by simple recursion.
+
+\emph{From Library
+\href{../theories/html/hydras.additions.Fib2.html}{additions.Fib2}}
+
+\begin{Coqsrc}
+Fixpoint fib (n:nat) : N :=
+  match n with
+    0%nat | 1%nat => 1%N
+  | S (S p as q) => fib p + fib q
+  end.
+
+Compute fib 20.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+= 10946 : N
+\end{Coqanswer}
+
+In~\cite{BC04}, several exercices~\footnote{Exercises 9.8 (page 270), 9.10 (page 271), 9.15 (page 276), 9.17 (page 284), and 15.8 (page 418).}  
+present ways to compute Fibonacci numbers, with the less number of recursive calls  as possible.
+Please note that these optimizations and the formal proof of their correctness are \emph{ad-hoc}, \emph{i.e.} exclusively written for the
+Fibonacci numbers.
+In contrast, the optimizations we present in this document apply, in their vast majority, \emph{generic} techniques of efficient computation of powers in a monoid. 
+This example of Fibonacci numbers has been developed with Yves Bertot, who wrote a first version with \texttt{SSreflect/Mathcomp}~\cite{MCB, SSR}.
+
+\subsubsection{Using 2x2 integer matrices}
 \index{Fibonacci numbers!Matrix exponentiation}
 
-\begin{Coqsrc}
-
-Import M2.
-
-Arguments M2_mult {A} plus mult  _  _.
-Arguments mat {A} _ _ _ _.
-Arguments Id2 {A}  _ _.
-
-Definition fibonacci (n:N) :=
- c00 N  (N_bpow  (M2_mult Nplus Nmult) 
-                 (Id2  0%N 1%N)
-                 (mat  1 1 1 0)%N 
-                 n).
-
-Compute fibonacci 20.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-= 10946%N
-     : N  
-\end{Coqanswer}
+The following properties are well known. They are left as an exercise, since they are not part of our development.
+
+\index{Exercises}
+
+\begin{exercise}
+  \begin{enumerate}
+  \item
+
+Prove in \coq{} the following equality (for any $n\geq 2$). \label{fibmat-eq1}
+
+\[
+\left(
+  \begin{array}{cc}
+    1 & 1 \\
+    1 & 0 
+  \end{array}
+\right)
+\left(
+  \begin{array}{cc}
+    F_{n}& F_{n-1} \\
+    F_{n-1} & F_{n-2}
+  \end{array}
+\right)
+=
+\left(
+  \begin{array}{cc}
+    F_{n+1}& F_{n} \\
+    F_{n} & F_{n-1} 
+  \end{array}
+\right)
+\]
+
+\item Infer (still in \coq{}) the following equality (still for $n\geq 2$).
+
+\[
+\left(
+  \begin{array}{cc}
+    F_{n}& F_{n-1} \\
+    F_{n-1} & F_{n-2} 
+  \end{array}
+\right)
+= 
+\left(
+  \begin{array}{cc}
+    1 & 1 \\
+    1 & 0 
+  \end{array}
+\right)^n
+\]
+
+\item Write a function using the previous equality for computing the $n$-th Fibonacci number, and prove its equivalence with \texttt{fib}.
+
+\end{enumerate}
+\end{exercise}
+
+\subsubsection{Removing duplicate computations}
+
+Yves Bertot's optimization relies on the observation that all the powers of
+\(  \left(
+  \begin{array}{cc}
+    1 & 1 \\
+    1 & 0 
+  \end{array}
+\right) \) have the form 
+\(  \left(
+  \begin{array}{cc}
+    a+b  & a \\
+    a & b
+  \end{array}
+\right) \) where $a$ and $b$ are natural numbers.
+
+Thus, it is possible to remove duplicate data and computations by reflecting matrix multiplication and identity into $\mathbb{N}\times\mathbb{N}$.
+
+If we define $\varphi(a,b)  \left(
+  \begin{array}{cc}
+    a+b  & a \\
+    a & b
+  \end{array}
+\right)$, then $\varphi(a,b)\times \varphi(c,d)=\varphi(ac + ad + bc, ac + bd)$, and
+$\varphi(0,1)=  \left(
+  \begin{array}{cc}
+    1 & 0 \\
+    0 & 1 
+  \end{array}
+\right) $.
+
+\index{Exercises}
+\begin{exercise}
+  Prove formally these properties. \emph{Please note that their proof is not needed in our development, they just help to understand the following optimization.}
+\end{exercise}
+
+So, let us define a binary operation, which makes $\mathbb{N}\times\mathbb{N}$ a monoid (with $(0,1)$ as neutral element).
+
+\emph{From Library
+\href{../theories/html/hydras.additions.Fib2.html}{additions.Fib2}}
+
+\begin{Coqsrc}
+Definition mul2 (p q : N * N) :=
+  match p, q with (a, b),(c,d) => (a*c + a*d + b*c, a*c + b*d) end.
+
+Instance Mul2 : Monoid  mul2 (0,1).
+(* Proof omitted *)
+\end{Coqsrc}
+
+The following lemma is a simplification of the equality~\vref{fibmat-eq1}.
+
+\begin{Coqsrc}
+Lemma next_fib (n:nat) : mul2 (1,0) (fib (S n), fib n) =
+                         (fib (S (S n)), fib (S n)).  
+\end{Coqsrc}
+
+By induction\footnote{Simple induction, since $n$ has type \texttt{nat}.}
+ over $n$, we obtain the following equality.
+
+ \begin{Coqsrc}
+Lemma fib_power n : fib n = let (a,b) := power (M:=Mul2) (1,0) n
+                            in (a+b).   
+\end{Coqsrc}
+
+Thus, any function able to compute more or less efficiently powers in a monoid will
+give an algorithm for computing Fibonacci numbers. Unlike the \emph{ad-hoc} aforementionned proofs of~\cite{BC04}, the correctness of such an algorithm is a direct consequence 
+of the correctness of the used powering function.
+Several examples will be presented in the rest of this document.
+
+\begin{todo}
+Put links when done.
+\end{todo}
+
+% \begin{Coqsrc}
+
+% Import M2.
+
+% Arguments M2_mult {A} plus mult  _  _.
+% Arguments mat {A} _ _ _ _.
+% Arguments Id2 {A}  _ _.
+
+% Definition fibonacci (n:N) :=
+%  c00 N  (N_bpow  (M2_mult Nplus Nmult) 
+%                  (Id2  0%N 1%N)
+%                  (mat  1 1 1 0)%N 
+%                  n).
+
+% Compute fibonacci 20.
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+% = 10946%N
+%      : N  
+% \end{Coqanswer}
 
@@
 
-\subsubsection{Remark}
-\label{sect:faster}
-
-Our function \texttt{N\_bpow} is really logarithmic. Let us make a comparative 
-test with Standard Library's exponentiation function on type \texttt{N} (see section~\vref{sect:slow-computation}).
-
-\begin{Coqsrc}
-Time Compute (N_bpow N.mul 1 1 56666667)%N.  
-\end{Coqsrc}
-
-\begin{Coqanswer}
-Finished transaction in 0. secs (0.u,0.s) (successful)  
-\end{Coqanswer}
+% \subsubsection{Remark}
+% \label{sect:faster}
+
+% Our function \texttt{N\_bpow} is really logarithmic. Let us make a comparative 
+% test with Standard Library's exponentiation function on type \texttt{N} (see section~\vref{sect:slow-computation}).
+
+% \begin{Coqsrc}
+% Time Compute (N_bpow N.mul 1 1 56666667)%N.  
+% \end{Coqsrc}
+
+% \begin{Coqanswer}
+% Finished transaction in 0. secs (0.u,0.s) (successful)  
+% \end{Coqanswer}
 

### 2020-12-03 doc/chapter-powers.tex
minor revisions on addition chains part (.tex, .v)

@@
 \subsubsection{Removing duplicate computations}
+\label{sect:fibonacci-mul2}
 
@@
 By induction\footnote{Simple induction, since $n$ has type \texttt{nat}.}
- over $n$, we obtain the following equality.
-
- \begin{Coqsrc}
-Lemma fib_power n : fib n = let (a,b) := power (M:=Mul2) (1,0) n
-                            in (a+b).   
+ over $n$, we obtain a the following equalities.
+
+\begin{Coqsrc}
+Definition fib_mul2 n := let (a,b) := power (M:=Mul2) (1,0) n   in (a+b).
+
+Lemma fib_mul2_OK_0 (n:nat) : power (M:=Mul2) (1,0) (S (S n)) =
+                              (fib (S n), fib n).
+
+Lemma fib_mul2_OK n : fib n = fib_mul2 n.
 \end{Coqsrc}
@@
 of the correctness of the used powering function.
-Several examples will be presented in the rest of this document.
+Several examples will be presented in the rest of this document
+(in sections~\vref{sect:fibonacci-pos-bpow}).
 
@@
 \end{Coqsrc}
+
+\subsection{Fibonacci, once again}
+\label{sect:fibonacci-pos-bpow}
+
+We can use the function \texttt{Pos\_bpow} for computing Fibonacci numbers
+(see Section~\vref{sect:fibonacci-mul2}).
+
+\begin{Coqsrc}
+Definition fib_pos n :=
+  let (a,b) := Pos_bpow (M:= Mul2) (1,0) n in
+  (a+b).
+
+Time Compute fib_pos 153%positive.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+  68330027629092351019822533679447
+     : N 
+Finished transaction in 0.002 secs (0.002u,0.s) (successful)
+\end{Coqanswer}
+
+Fibonacci will come back in Sect.~\vref{sect:fibonacci-euclidean}.
 
@@
 Definition chain_correct (p:positive) (c: chain) :=
-  chain_correct_nat c (Pos.to_nat p). 
+  chain_correct_nat c (Pos.
+to_nat p). 
 \end{Coqsrc}
@@
 a * a * a == a * (a * (a * E_one))}
+
 \end{Coqanswer}
@@
 
+\begin{Coqsrc}
+Function dicho_aux (p:positive) {struct p} : positive :=
+ match p with
+   | 1%positive   =>  xH
+   | 2%positive |   3%positive  => 2 
+   | xO (xO q) | xO (xI q) | xI (xO q) | xI (xI q) =>
+                                         xO (dicho_aux q)
+ end.
+
+Definition dicho  (p:positive) : positive :=
+  N2pos (N.div (Npos p) (Npos (dicho_aux p))).
+
+Instance Dicho_strat : Strategy dicho.
+\end{Coqsrc}
+
+\subsection{Other strategies}
+For comparison's sake, we define two other strategies, much simpler but statically less efficient than the dichotomic strategy.
+
+\emph{From Module~\href{../theories/html/hydras.additions.BinaryStrat.html}{additions.BinaryStrat}.}
+
+\begin{Coqsrc}
+Definition half (p:positive) :=
+  match p with xH => xH
+          |    xI q | xO q =>  q
+  end.
+
+Definition two (p:positive) := 2%positive.
+
+Instance Binary_strat : Strategy half.
+Proof.
+  split; destruct p; unfold half; try lia.
+Qed.
+
+Instance Two_strat : Strategy two.
+Proof.
+  split;unfold two; lia.
+Qed.
+\end{Coqsrc}
+
+In Sect.~\ref{sect:test-strat}, we compare the three strategies with respect to the lenght of the built chains.
+
 \subsection{Main chain generation function}
@@
 
+\label{sect:test-strat}
+
+The following tests show various examples of chains for the same exponent, using different strategies. The dichotomic strategy seems clearly to be the winner (at least on this sample).
+
+\begin{Coqsrc}
+Compute chain_length (make_chain two 56789).
+\end{Coqsrc}
+
+\begin{Coqanswer}
+= 25%nat : nat  
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Compute chain_length (make_chain half 56789).
+\end{Coqsrc}
+
+\begin{Coqanswer}
+ = 25%nat : nat
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Compute chain_length (make_chain dicho 56789).
+\end{Coqsrc}
+
+\begin{Coqanswer}
+= 21%nat : nat 
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Compute chain_length (make_chain two 3456789).
+\end{Coqsrc}
+
+\begin{Coqanswer}
+= 33%nat : nat
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Compute chain_length (make_chain half 3456789).
+\end{Coqsrc}
+
+\begin{Coqanswer}
+(= 33%nat : nat
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Compute chain_length (make_chain dicho 3456789).
+\end{Coqsrc}
+
+\begin{Coqanswer}
+= 29%nat : nat
+\end{Coqanswer}
+
 \subsubsection{Correctness of the Euclidean chain generator}
@@
 
+\subsection{Fibonacci, \emph{le retour}}
+\label{sect:fibonacci-euclidean}
+
+It is now possible to use euclidean addition chains for computing Fibonacci numbers
+(see Sections~\vref{sect:fibonacci-mul2} and~\vref{sect:fibonacci-pos-bpow}).
+
+The following function is parameterized by a startegy $\gamma$.
+
+\begin{Coqsrc}
+Definition fib_eucl gamma `{Hgamma: Strategy gamma} n :=
+  let c := make_chain gamma  n
+  in let r := chain_apply c (M:=Mul2) (1,0) in
+       fst r + snd r.
+
+Compute fib_eucl dicho 153.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+    = 68330027629092351019822533679447
+     : N
+Finished transaction in 0.002 secs (0.002u,0.s) (successful)
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Compute fib_eucl two 153.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+    = 68330027629092351019822533679447
+     : N
+Finished transaction in 0.003 secs (0.003u,0.s) (successful)
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Compute fib_eucl half 153.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+    = 68330027629092351019822533679447
+     : N
+Finished transaction in 0.003 secs (0.003u,0.s) (successful)
+\end{Coqanswer}
+
 \section{Projects}

### 2020-12-04 doc/chapter-powers.tex
Added traces of addition chains Small modification of the documentation

@@
 
-\section{Comparing Exponentiation Algorithms with respect to Efficiency}
+\section{Comparing Exponentiation Algorithms with Respect to Efficiency}
 
@@
 
-\subsection{Addition chains}
-An \emph{addition chain} (In short : \emph{a chain})~\cite{brauer1939} is a representation of a sequence of
+\section{Addition chains}
+An \emph{addition chain} (In short : \emph{a chain})~\cite{brauer1939, additionchainspage} is a representation of a sequence of
 intermediate steps that lead to the evaluation of some $x^n$, under the 
@@
 
-From Module~\href{../theories/html/hydras.additions.Chains.html}{additions.Chains}
+From Module~\href{../theories/html/hydras.additions.Addition_Chains.html}{additions.Addition\_Chains}
 
@@
 
+\index{Projects}
+\begin{project}
+Define a function which returns the sequence of operations defined by a chain.
+For instance, the chain \texttt{C87} of Figure \ref{fig:C87} can be represented as a 
+list containing terms of the form \texttt{($i$, Add $j$ $k$)} whenever the associated computation contains the operation $x^i=x^j\times x^k$.
+
+\begin{Coqsrc}
+Compute chain_trace C87.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+(1, Init)
+       :: (2, Add 1 1)
+          :: (3, Add 2 1)
+             :: (6, Add 3 3)
+                :: (7, Add 6 1)
+                   :: (10, Add 7 3)
+                      :: (20, Add 10 10)
+                         :: (40, Add 20 20)
+                            :: (80, Add 40 40) :: (87, Add 80 7) :: nil
+     : list (positive * info)
+\end{Coqanswer}
+
+\textbf{Note} A first solution (in ~\href{../theories/html/hydras.additions.Trace_exercise.html}{additions.Trace\_exercise}) consists in the definition of 
+a (non-associative) multiplication over a type of trace, and apply the function
+\texttt{chain\_execute} as if it were computing a power of \texttt{(1,Init)}.
+
+\end{project}
+
 \subsubsection{Chain Correctness and Optimality}
@@
 
-\section{Proving a chain's correctness}
+\section{Proving a Chain's Correctness}
 \label{chain-correctness-sect}
@@
 
-\subsection{reflection tactic}
+\subsection{Reflection tactic}
 The tactic \texttt{monoid\_eq\_A} converts a goal of the form 

### 2020-12-07 doc/chapter-powers.tex
corrections of typos

@@
 
-\subsection{The dichotomic strategy}
+\subsection{Generation of chains using Euclidean Dibision}
 
@@
 In \cite{DBLP:journals/ita/BrlekCHM95}, the function that maps $n$ into $b$
-is called a \emph{strategy}. In this chapter, we concentrate
-on the so-called \emph{dichotomic strategy}.
-
-$\delta(n) = n \div {2^k}$ where $k=\floor{(\log_2{n})/2}$.
+is called a \emph{strategy}.
+
+\vspace{4pt}
+\noindent
+From ~\href{../theories/html/hydras.additions.Dichotomy.html}{additions.Dichotomy}.
+\begin{Coqsrc}
+Class Strategy (gamma : positive -> positive):=
+{
+gamma_lt :forall p:positive, 3 < p -> gamma  p < p;
+gamma_gt : forall p:positive, 3 < p -> 1 < gamma  p
+}.
+
+\end{Coqsrc}
+\subsection{The dichotomic strategy}
+
+In this chapter, we concentrate
+on the so-called \emph{dichotomic strategy}, defined as follows:
+
+$$n \mapsto  n \div {2^k} \,\textbf{where}\, k=\floor{(\log_2{n})/2}$$
 
@@
 
-Module ~\href{../theories/html/hydras.additions.Dichotomy.html}{additions.Dichotomy} contains a definition of the function \texttt{delta}, and proofs that if 
-$n>3$ then $1<\delta(n)<n$.
+This strategy is defined in Module ~\href{../theories/html/hydras.additions.Dichotomy.html}{additions.Dichotomy}.
 
@@
 
-In Sect.~\ref{sect:test-strat}, we compare the three strategies with respect to the lenght of the built chains.
+Page.~\pageref{sect:test-strat}, we compare the three strategies with respect to the length of the built chains.
 
@@
 the recursive calls. They are solved with the help of Standard Library's lemmas 
-on arithemtic of \texttt{positive} numebers end Euclidean division.
+on arithmetic of \texttt{positive} numbers and Euclidean division.
 
@@
 
+\subsubsection{A few tests}
 \label{sect:test-strat}
@@
 the composition operators on chains: \texttt{Fcompose}, \texttt{KFK}, etc.
-Unfortunately, a lot of interaction is still needed or proving properties of
+Unfortunately, a lot of interaction is still needed for proving properties of
 Euclidean division and binary logarithm.
@@
 
-It is now possible to use euclidean addition chains for computing Fibonacci numbers
+It is now possible to use Euclidean addition chains for computing Fibonacci numbers
 (see Sections~\vref{sect:fibonacci-mul2} and~\vref{sect:fibonacci-pos-bpow}).
 
-The following function is parameterized by a startegy $\gamma$.
+The following function is parameterized by any strategy $\gamma$.
 
@@
 
-\index{Projects}
-\begin{project}
+\subsection{A data structure for Euclidean chains}
+
 Figures~\vref{fig:F3-as-dag} to \vref{fig:FFK} suggest that any computation following an Euclidean chain can be executed on a kind  of abstract machine with a "register'' and a stack, and only four operations:
@@
 
-For instance the chain of Fig.~ \vref{fig:C87} can be represented with the following code:
+For instance the chain of Fig.~\vref{fig:C87} can be represented with the following code:
 
@@
 
-Please define an Euclidean chain generator using this data structure, and prove its correctness (done  in ~\href{../theories/html/hydras.additions.AM.html}{additions.AM}, but with clumsy fresh proofs).
-The project is now to simplify the proofs.
-perhaps considering chains of type \texttt{code} as a refinement of functional chains. Please note that many proof scripts are copied verbatim from 
-\texttt{Euclidean\_Chains} into \texttt{AM}.
-\end{project}
+In the library~\href{../theories/html/hydras.additions.AM.html}{additions.AM},
+we define a chain generator for this data structure. 
+Please note that many proof scripts are copied verbatim from 
+\texttt{Euclidean\_Chains} into \texttt{AM}. Removing such redundancies is left as a project.
 

### 2020-12-10 doc/chapter-powers.tex
minor corrections

@@
 
-Lemma FFK_correct :
-  forall (p q  : positive) (fp fq : Fchain),
+Lemma FFK_correct  (p q  : positive) (fp fq : Fchain):
     Fchain_correct p fp  ->
@@
 
-Instance FFK_proper 
-   (fp: Fchain) (fq : Fchain)
+Instance FFK_proper  (fp fq : Fchain)
     (_ :   Fchain_proper fp)

### 2020-12-10 doc/hydras.tex
Added traces of addition chains Small modification of the documentation
minor corrections

@@
 
+%\SetWatermarkLightness{0.8}
+\definecolor{lightorange}{rgb}{1,0.7,0.5}
+\SetWatermarkScale{0.35}
+\SetWatermarkText{Work in Progress}
+\SetWatermarkColor{lightorange}
+
 \author{Pierre Castéran\\ LaBRI, Univ. Bordeaux, CNRS UMR 5800, Bordeaux-INP\\with contributions by Yves Bertot, \'Evelyne Contejean,  Florian Hatat and Pascal Manoury}
@@
 
-\cleardoublepage
+% \cleardoublepage
 
-% \mbox{~}
-% \clearpage
-% \newpage
-% \mbox{~}
-\vspace{10cm}
-\begin{quote} 
-{ \Large {\it
-``I start from one point and go as far as possible.'' }
+% % \mbox{~}
+% % \clearpage
+% % \newpage
+% % \mbox{~}
+% \vspace{10cm}
+% \begin{quote} 
+% { \Large {\it
+% ``I start from one point and go as far as possible.'' }
 
-John Coltrane.}
-\end{quote}
+% John Coltrane.}
+% \end{quote}
 
@@
 
-Among all the theorems proved with the help of proof assistants like \coq{}, \isabelle{}~\cite{isabelle}, \hol{}~\cite{HOL}, etc.,
+Among all the theorems proved with the help of proof assistants like \coq{}~\cite{Coq,BC04}, \hol{}~\cite{HOL}, \isabelle{}~\cite{isabelle},  etc.,
 several statements and proofs  share some interesting features:
@@
 comment or proposition that would help us. The text contains several \emph{projects} which, when completed, may improve the present work.
-Please do not hesitate to bring your contribution!
+Please do not hesitate to bring your contribution, for instance with Github's proof requests and issues. Thank you in advance!
 
@@
   \begin{Coqsrc}
- Require Import Arith.
-
  Definition square (n:nat) := n * n.
@@
 \begin{Coqbad}
-Require Import Lia.
-
 Module Bad.
@@
 \begin{Coqalt}
-Require Import Arith Lia Iterates.
+From hydras Require Import Iterates.
 Module Alt.
@@
  members of the \emph{Formal Methods} team at laBRI for their helpful comments 
-on an oral presentation of this work.
+on  oral presentations of this work.
 

### 2020-12-10 doc/part-hydras.tex
minor corrections

@@
 
-In order to describe trees where each node can have an arbitrary (but finite) number of sons, it usual to define a type where each node carries a \emph{forest}, \emph{i.e} a list of trees
+In order to describe trees where each node can have an arbitrary (but finite) number of sons, it is usual to define a type where each node carries a \emph{forest}, \emph{i.e} a list of trees
 (see for instance Chapter 14, pages 400-406 of \cite{BC04}).
@@
 \vspace{4pt}
+\noindent
 \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hydra}{Hydra.Hydra\_Definitions}}
@@
 
+\vspace{4pt}
+\noindent
 \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#head}{Hydra.Hydra\_Definitions}}
@@
 \vspace{4mm}
+\noindent
 \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html\#Hy}{Hydra.Hydra\_Examples}}
@@
 \vspace{4pt}
-
+\noindent
 \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#hcons_mult}{Hydra.Hydra\_Definitions}}
@@
 
-(** hydra with n copies of the same daughter *)
+(** A node with n clones of the same daughter *)
 
@@
 \vspace{4pt}
-
+\noindent
 \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
@@
 
-When defining a recursive function over the type \texttt{Hydra}, one has to consider the three constructors 
+In order to  define a recursive function over the type \texttt{Hydra}, one has to consider the three constructors 
 \texttt{node}, \texttt{hnil} and \texttt{hcons} of the mutually inductive types \texttt{Hydra} and \texttt{Hydrae}. 
@@
 Using this scheme requires an auxiliary predicate, called \texttt{P0} in \texttt{Hydra\_rect2}'s statement.
-Let us begin by defining an ad-hoc version of \texttt{List.Forall}.
 
@@
 
-The rest of the definition is structured as two mutually inductive relations on hydras and sequences of hydras.
+The rest of the definition is composed of two mutually inductive relations on hydras and sequences of hydras.
 The first constructor of \texttt{R2} describes the case where the chopped head is exactly at height $2$.
@@
 
- We define an addition in type \texttt{E0}, using the function{T1.plus}
-and its properties.
-
-\begin{Coqsrc}
-Search   (nf (_ + _)%t1).
-\end{Coqsrc}
-
-\begin{Coqanswer}
-plus_nf: forall a : T1, nf a -> forall b : T1, nf b -> nf (a + b)
-\end{Coqanswer}
+ We define an addition in type \texttt{E0}, since the sum of two terms in normal form is in normal form too.
 

### 2020-12-11 doc/part-hydras.tex
minor corrections

@@
 Inductive battle (B:Battle) : nat -> Hydra -> nat -> Hydra -> Prop :=
-
-| battle_1 : forall i h  h', battle_r   B i  h h' -> 
-                            battle B i h (S i) h'
-| battle_n : forall i h  j h' h'',  battle_r  B i h h''  ->
+| battle_1 : forall i h  h', 
+      battle_r   B i  h h' -> battle B i h (S i) h'
+| battle_n : forall i h  j h' h'',  
+      battle_r  B i h h''  ->
       battle B (S i) h'' j h'  ->
@@
 \item At each round, Hercules chops off the rightmost head of the hydra.
-\item The battle is standard: at the round number $i$, the expected replication is $i$.
+\item The battle is called \emph{standard}\footnote{This appellation is ours. If there is a better one, we will change it.}: at the round number $i$, the expected replication is $i$.
 \end{itemize}
@@
 
-But the  battle is so long that no \emph{test} can give us an estimation of its length, and we do need the expressive power of logic to compute this length.
+But the  battle is so long that no \emph{test} can give us any estimation of its length, and we do need the expressive power of logic to compute this length.
 However, in order to  guess this length, we made some experiments, computing with \gallina{}, \coq{}'s  functional programming language.
 Thus, we can consider this development as a collaboration of proof with computation.
-
-In the following lines, we try to show faithfully how we found the value of the number $N$.
-
+In the following lines, we show how we found experimentally the value of the number $N$.
 The complete proof is in file \url{../theories/html/hydras.Hydra.BigBattle.html}.
 
-\subsection{The beginning of Hostilities}
-During the two first rounds, our hydra loses its two rightmost heads. Thus just before the third round, it looks like in figure~\vref{fig:hinit-plus2}.
+\subsection{The Beginning of Hostilities}
+During the two first rounds, our hydra loses its two rightmost heads.  Figure~\vref{fig:hinit-plus2} shows the state of the hydra   just before the third round.
 
@@
 subtrees of height 1 or 0).
-Thus, we introduce handy notations.
+Thus, we introduce a few handy notations.
 
@@
 
-\subsection{Computing \dots}
+\subsection{Testing  \dots}
 In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple datatype for representing the states as tuples composed of
@@
 The following function returns the next configurarion of the game. 
-Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle will only start with the lemma
-\texttt{lemma:step-battle}, page~\pageref{lemma:step-battle}.
+Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle only start with the lemma
+\texttt{step\_battle}, page~\pageref{lemma:step-battle}.
 
@@
 Compute test 3.
-   (**
+\end{Coqsrc}
+
+\begin{Coqanswer}
       = {| round := 3; n2 := 3; n1 := 0; nh := 0 |}
      : state 
-    *)
-
+\end{Coqanswer}
+
+\begin{Coqsrc}
  Compute test 4.
- (*
+\end{Coqsrc}
+
+\begin{Coqanswer}
   = {| round := 4; n2 := 2; n1 := 4; nh := 0 |}
      : state
- *)
-
+\end{Coqanswer}
+
+\begin{Coqsrc}
 Compute test 5.
- (*
+\end{Coqsrc}
+
+\begin{Coqanswer}
   = {| round := 5; n2 := 2; n1 := 3; nh := 5 |}
      : state
- *)
-
+\end{Coqanswer}
+
+\begin{Coqsrc}
 Compute test 2000.
-(*
+\end{Coqsrc}
+
+\begin{Coqanswer}
   = {| round := 2000; n2 := 1; n1 := 90; nh := 1102 |}
      : state
-*)
-\end{Coqsrc}
+\end{Coqanswer}
 
@@
  Compute test 10.
-(*
-
+\end{Coqsrc}
+
+\begin{Coqanswer}
     = {| round := 10; n2 := 2; n1 := 3; nh := 0 |}
      : state
-*)
- \end{Coqsrc}
+ \end{Coqanswer}
 
@@
 Compute test 22.
-(*
-
+\end{Coqsrc}
+
+\begin{Coqanswer}
  = {| round := 22; n2 := 2; n1 := 2; nh := 0 |}
      : state
-*)
-\end{Coqsrc}
+\end{Coqanswer}
 
@@
  Compute test 46.
-(*
-
- = {| round := 46; n2 := 2; n1 := 1; nh := 0 |}
+\end{Coqsrc}
+
+\begin{Coqanswer}
+( = {| round := 46; n2 := 2; n1 := 1; nh := 0 |}
      : state
-
-*)
-
+\end{Coqanswer}
+
+\begin{Coqsrc}
 Compute test 94.
-
-(*
-
+\end{Coqsrc}
+
+\begin{Coqanswer}
  = {| round := 94; n2 := 2; n1 := 0; nh := 0 |}
      : state
-
-*) 
-\end{Coqsrc}
+\end{Coqanswer}
 
@@
  Compute test 95.
-
-(*
-
+\end{Coqsrc}
+
+\begin{Coqanswer}
   = {| round := 95; n2 := 1; n1 := 95; nh := 0 |}
      : state
-
-*)
-\end{Coqsrc}
+\end{Coqanswer}
 
@@
 Compute test (doubleS 95).
-
-(**
+\end{Coqsrc}
+
+\begin{Coqanswer}
  = {| round := 192; n2 := 1; n1 := 94; nh := 0 |}
      : state
- *)
-
+\end{Coqanswer}
+
+\begin{Coqsrc}
 Compute test (iterate doubleS 2 95).
-
-(*
+\end{Coqsrc}
+
+\begin{Coqanswer}
   = {| round := 386; n2 := 1; n1 := 93; nh := 0 |}
      : state
-*)
-\end{Coqsrc}
+\end{Coqanswer}
 

### 2020-12-11 doc/chapter-powers.tex
correction of errors in tex files and bibliography

@@
 \section{Addition chains}
-An \emph{addition chain} (In short : \emph{a chain})~\cite{brauer1939, additionchainspage} is a representation of a sequence of
+An \emph{addition chain} (In short : \emph{a chain})~\cite{brauer1939} is a representation of a sequence of
 intermediate steps that lead to the evaluation of some $x^n$, under the 

### 2020-12-14 doc/hydras.tex
correction of typos

@@
 
+\section{Generalities}
+
 Proof assistants are excellent tools for exploring the structure of mathematical proofs,
@@
 \end{itemize}
+
+\paragraph*{Warning:}
+
+This document is \emph{not} an introductory text for \coq{}, and there are many aspects of this proof assistant that are not covered. 
+ The reader should already have some basic experience with the \coq{} system.
+The Reference Manual and several tutorials are available on \coq{} page~\cite{Coq}.
+First chapters of textbooks like \emph{Interactive Theorem Proving and Program Development}~\cite{BC04}, \emph{Software Foundations}~\cite{SF} or  \emph{Certified Programming with Dependent Types} ~\cite{chlipalacpdt2011} will give you the right background.
 
@@
 
-\paragraph*{Warning:}
-
-This document is \emph{not} an introductory text for \coq{}, and there are many aspects of this proof assistant that are not covered. 
- The reader should already have some basic experience with the \coq{} system.
-The Reference Manual and several tutorials are available on \coq{} page~\cite{Coq}.
-First chapters of textbooks like \emph{Interactive Theorem Proving and Program Development}~\cite{BC04}, \emph{Software Foundations}~\cite{SF} or  \emph{Certified Programming with Dependent Types} ~\cite{chlipalacpdt2011} will give you the right background.
-
-\subsection{Contributions}
-Many thanks to Yves Bertot, 
-\'Evelyne Contejean,  Florian Hatat and Pascal Manoury  for their contribution. 
-Yves made nice optimizations  to algorithms presented in Chapter~\ref{chapter-powers}.
-\'Evelyne contributed libraries on the recursive path ordering (\emph{rpo}) for proving the well-foundedness of our representation of $\epsilon_0$ and $\Gamma_0$.
-Florian Hatat proved many useful lemmas on countable sets, which we used in our adaptation of Schütte's formalization of countable ordinals.
-Pascal Manoury is integrating the ordinal $\omega^\omega$ into our hierarchy of ordinal notations.
-
-\label{sec:orgheadline2}
-
-Any form of contribution  is welcome: correction of errors (typos and more serious mistakes), improvement of
-Coq scripts, proposition of inclusion of new chapters, and generally any
-comment or proposition that would help us. The text contains several \emph{projects} which, when completed, may improve the present work.
-Please do not hesitate to bring your contribution, for instance with Github's proof requests and issues. Thank you in advance!
-
 \subsection{Typographical Conventions}
@@
  \end{Coqsrc}
+
+In general, we do not include full proof scripts in this document.
+The only exceptions are very short proofs (\emph{e.g.} proofs by computation, or by application of automatic tactics).
+Likewise, we may display only the important steps on a long interactive proof, for instance, in the following lemma (\vref{lemma:L-2_6-1}):
+
+\begin{Coqsrc}
+Lemma Lemma2_6_1 (alpha : T1) :  
+  nf alpha -> forall beta,  beta t1< alpha  ->
+  {n:nat | const_pathS n alpha beta}.
+Proof.
+  transfinite_induction alpha.
+  (* ... *)
+\end{Coqsrc}
+
+The reader may consult the full proof scripts with Proof General or CoqIDE, for instance.
 
@@
 
-\section{Acknowledgements}
-\label{sec:orgheadline5}
-    Many thanks to Yves Bertot, David Ilcinkas, Karl Palmskog, Sylvain Salvati, Alan Schmitt and Théo Zimmerman for their help on the elaboration of this document, and to the
- members of the \emph{Formal Methods} team at laBRI for their helpful comments 
-on  oral presentations of this work.
-
-Many thanks also to the Coq development team, Yves Bertot, and members of the \emph{Coq Club} for interesting discussions about the \coq{} system and the Calculus of Inductive Constructions.
-
-The author of the present document wishes to express his gratitude to the late Patrick Dehornoy, whose talk  was determinant for our desire to work on this topic.
-
-I owe my interest in discrete mathematics and their relation to formal proofs and functional programming  to Srecko Brlek.
-Equally, there is W.
-H.
-Burge's book ``\emph{Recursive Programming Techniques}'' ~\cite{burge} which was a great  source of inspiration.
-
 \section{How to Install the Libraries}
@@
 we think it would be useful to include comments and hints, and questions/answers from the users. In constrast, ``projects'' are supposed, once completed, to be included in the repository.
+
+\section{Acknowledgements}
+\label{sec:orgheadline5}
+    Many thanks to Yves Bertot, \'Evelyne Contejean,  Florian Hatat,  David Ilcinkas, 
+Pascal Manoury,  Karl Palmskog, Sylvain Salvati, Alan Schmitt and Théo Zimmerman for their help on the elaboration of this document, and to the
+ members of the \emph{Formal Methods} team at laBRI for their helpful comments 
+on  oral presentations of this work.
+
+Many thanks also to the Coq development team, Yves Bertot, and the members of the \emph{Coq Club} for interesting discussions about the \coq{} system and the Calculus of Inductive Constructions.
+
+The author of the present document wishes to express his gratitude to the late Patrick Dehornoy, whose talk  was determinant for our desire to work on this topic.
+I owe my interest in discrete mathematics and their relation to formal proofs and functional programming  to Srecko Brlek.
+Equally, there is W.
+H.
+Burge's book ``\emph{Recursive Programming Techniques}'' ~\cite{burge} which was a great  source of inspiration.
+
+\subsection{Contributions}
+
+Yves Bertot made nice optimizations  to algorithms presented in Chapter~\ref{chapter-powers}.
+\'Evelyne Contejean contributed libraries on the recursive path ordering (\emph{rpo}) for proving the well-foundedness of our representation of $\epsilon_0$ and $\Gamma_0$.
+Florian Hatat proved many useful lemmas on countable sets, which we used in our adaptation of Schütte's formalization of countable ordinals.
+Pascal Manoury is integrating the ordinal $\omega^\omega$ into our hierarchy of ordinal notations.
+
+\label{sec:orgheadline2}
+
+Any form of contribution  is welcome: correction of errors (typos and more serious mistakes), improvement of
+Coq scripts, proposition of inclusion of new chapters, and generally any
+comment or proposition that would help us. The text contains several \emph{projects} which, when completed, may improve the present work.
+Please do not hesitate to bring your contribution, for instance with Github's proof requests and issues. Thank you in advance!
 

### 2020-12-14 doc/part-hydras.tex
correction of typos

@@
       battle B i h j h'.
+\end{Coqsrc}
+
+The following property allows us to build battles out of smaller ones.
+
+\begin{Coqsrc}
+Lemma battle_trans {b:Battle} :
+  forall i h j h', battle b i h j h' ->
+        forall k h0,  battle b k h0 i h -> battle b k h0 j h'.
 \end{Coqsrc}
@@
 Thus, any result about \texttt{steps} will be applicable to standard battles.
-Using the predicate \texttt{steps} our study of the length of the considered battle
+Using the predicate \texttt{steps},  our study of the length of the considered battle
 can  be decomposed into three parts:
@@
 The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
-\url{../theories/html/hydras.Hydra.Hydra_Theorems.html} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
+\href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
 
@@
 Definition  next_round_dec n :
- forall h , (h = head) + {h' : Hydra & {R1 h h'} + {R2 n h  h'}}.
+ forall h, (h = head) + {h' : Hydra & {R1 h h'} + {R2 n h  h'}}.
 \end{Coqsrc}
@@
 In order to get a contradiction, it suffices to  prove the inequality
-\texttt{m big\_h <= m small\_h}, i.e.,  \texttt{m big\_h <= m (iota (m big\_h))}.
+$m(\texttt{big\_h}) \leq m(\texttt{small\_h})$ i.e.,  $m(\texttt{big\_h}\leq m(\iota (m(\texttt{big\_h})))$.
 
@@
 
-Let us have a look at Figure~\ref{fig:ordinal-sequence}. It presents a small sequence of ordinal numbers, which extends the sequence of natural numbers.
+Let us have a look at Figure~\ref{fig:ordinal-sequence}. It presents a few items of a  sequence of ordinal numbers, which extends the sequence of natural numbers.
 
@@
 &\textcolor{red}{\omega},\,\omega+1,\omega+2,\omega+3,\ldots\\
-&\textcolor{red}{\omega\times 2},\,\omega\times 2+1,\ldots, \textcolor{red}{\omega\times 3},\,\omega\times 3+1,\ldots, \textcolor{red}{\omega\times 4},\ldots,\\
-&\textcolor{red}{\omega^2},\ldots, \textcolor{red}{\omega^2\times 42},\ldots,\textcolor{red}{\omega^3},\ldots, \textcolor{red}{\omega^4},\omega^4+1,\ldots,\\
-&\textcolor{red}{\omega^\omega},\ldots, \textcolor{red}{\omega^\omega+\omega^7\times 8},\ldots,\textcolor{red}{\omega^\omega\times 2},\omega^\omega\times 2+1, \ldots,\\
+&\textcolor{red}{\omega\times 2},\,\omega\times 2+1,\ldots, \textcolor{red}{\omega\times 3},\,\omega\times 3+1,\ldots, \textcolor{red}{\omega\times 4},\ldots\\
+&\textcolor{red}{\omega^2},\ldots, \textcolor{red}{\omega^2\times 42},\ldots,\textcolor{red}{\omega^3},\ldots, \textcolor{red}{\omega^4},\omega^4+1,\ldots\\
+&\textcolor{red}{\omega^\omega},\ldots, \textcolor{red}{\omega^\omega+\omega^7\times 8},\ldots,\textcolor{red}{\omega^\omega\times 2},\omega^\omega\times 2+1, \ldots\\
 &\textcolor{red}{\omega^{\omega^\omega}},\ldots, \textcolor{red}{\omega^{\omega^\omega}+\omega^\omega\times 42+ \omega^{55}+\omega}, \ldots, \textcolor{red}{\omega^{\omega^{\omega+1}}}, \omega^{\omega^{\omega+1}}+1,\dots\\
-& \textcolor{red}{\epsilon_0 (= \omega^{\epsilon_0)}}, \epsilon_0+1, \epsilon_0+2, \epsilon_0+3, \ldots,\\
+& \textcolor{red}{\epsilon_0 (= \omega^{\epsilon_0)}}, \epsilon_0+1, \epsilon_0+2, \epsilon_0+3, \ldots\\
 & \textcolor{red}{\epsilon_1}, \ldots, \textcolor{red}{\epsilon_2}, \ldots, \textcolor{red}{\epsilon_\omega},\ldots \\
-& \textcolor{red}{\Gamma_0}, \Gamma_0+1, \Gamma_0+2, \Gamma_0+3,\ldots, \textcolor{red}{\Gamma_0+\omega}, \ldots,\\
+& \textcolor{red}{\Gamma_0}, \Gamma_0+1, \Gamma_0+2, \Gamma_0+3,\ldots, \textcolor{red}{\Gamma_0+\omega}, \ldots\\
 &\ldots
@@
 \item Dots : ``$\ldots$'' stand for  infinite sequences of ordinals, not shown for lack of space.
-For instance, the ordinal $42$ is not shown in the first line, but it exists, between $17$ and $\omega$.
+For instance, the ordinal $42$ is not shown in the first line, but it exists, somewhere between $17$ and $\omega$.
 \item Each ordinal printed in black is the immediate successor of another ordinal.
@@
 
-Having the same order type is an equivalence relation between well-ordered sets. Ordinal numbers (in short \emph{ordinals}) are descriptions (\emph{names}) of the equivalence classes.
+Having the same order type is an equivalence relation between well-ordered sets. Ordinal numbers (in short: \emph{ordinals})  are descriptions (\emph{names}) of the equivalence classes.
 For instance, the order type of $(\mathbb{N},<)$ is associated with the ordinal called  $\omega$, and the order we considered on 
@@
 So the infix \texttt{<} and \texttt{<=} are reserved to the natural numbers.
-In the mathematical formulas, we still use $<$ and $\leq$ for comparing ordinals.
+In the mathematical formulas, however, we still use $<$ and $\leq$ for comparing ordinals.
 \end{remark}
@@
 
-\section{Example: The Ordinal \texorpdfstring{$\omega$}{omega}}
+\section{Example: the Ordinal \texorpdfstring{$\omega$}{omega}}
 
@@
 
-\section{Sum of  Ordinal Notations}
+\section{Sum of  two Ordinal Notations}
 
@@
 
+\vspace{4pt}
 \noindent
@@
 
-Let us look again at our implementation of $\omega+\omega$. We can distinguish between the three kinds of ordinals seen in Fig~\ref{fig:ordinal-sequence}:
+Let us look again at our implementation of $\omega+\omega$. We can classify its elements into three categories:
 
 \begin{itemize}
-\item The least ordinal, \texttt{(inl 0)}, also written \texttt{(fin 0)}.
+\item The least ordinal, \texttt{(inl 0)}, also known as  \texttt{(fin 0)}.
 \item The limit ordinal $\omega$.
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Prelude.Restriction.html}{Ordinals.Prelude.Restriction}}
+\emph{From Module~\href{../theories/html/hydras.Prelude.Restriction.html}{hydras.Prelude.Restriction}}
 
@@
  Proof with eauto with T1.
-    intros alpha Halph; transfinite_induction alpha.
+    intros alpha Halpha; transfinite_induction alpha.
 (* rest of proof omitted *)  
@@
 Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
-(please see also \href{../theories/html/hydras.Schutte.Correctness_E0.html}{the module \texttt{Ordinals.Schutte.Correctness\_E0}}).
+(please see also \href{../theories/html/hydras.Schutte.Correctness_E0.html}{the module \texttt{hydras.Schutte.Correctness\_E0}}).
 
@@
  Unlike the rest of our libraries, our library
-\href{../theories/html/hydras.Schutte.Schutte.html}{Ordinals.Schutte}
+\href{../theories/html/hydras.Schutte.Schutte.html}{hydras.Schutte}
 is not constructive, and relies on several axioms.

### 2020-12-18 doc/part-hydras.tex
correction of typos

@@
 there exists a strictly monotonous bijection $b$ from $A$ to $B$, \emph{i.e.} which verifies the proposition
-$\forall x\,y\in A, x <_A y \Rightarrow b(x) <_B  b(y)$.
+$\forall x\,y\in A,\, x <_A y \Rightarrow b(x) <_B  b(y)$.
 
@@
 Nevertheless, since we work within type theory,
-we do not identify ordinals as sets of ordinals, but the correspondance between ordinals and sets of ordinals is the function that maps $\alpha$ to $\mathbb{O}_\alpha$.
+we do not identify ordinals as sets of ordinals, but consider the correspondance between ordinals and sets of ordinals as the function that maps $\alpha$ to $\mathbb{O}_\alpha$.
 For instance $\mathbb{O}_\omega=\mathbb{N}$, and $\mathbb{O}_7=\{0,1,2,3,4,5,6\}$.
@@
 
-Two kinds of representation of ordinals are defined herein.
+Two kinds of representation of ordinals are defined in our development.
 
@@
 
-\section{Countable Ordinals}
-
-Chapter~\ref{chap:schutte} of this document presents an adaptation to \coq{} of an axiomatization in classical logic of the set of countable ordinals by K. Schütte~\cite{schutte}. 
-That formalization is quite complex, technical and unshamedly non-constructive,  so we put its description  in the last chapter of this document.
-
-Please note that Schütte considers the (uncountable) set $\mathbb{O}$ of all countable ordinals.
+% \section{Countable Ordinals}
+
+% Chapter~\ref{chap:schutte} of this document presents an adaptation to \coq{} of an axiomatization in classical logic of the set of countable ordinals by K. Schütte~\cite{schutte}. 
+% That formalization is quite complex, technical and unshamedly non-constructive,  so we put its description  in the last chapter of this document.
+
+% Please note that Schütte considers the (uncountable) set $\mathbb{O}$ of all countable ordinals.
 This set is well ordered (which is one of Schütte's axioms), and associates to any ordinal $\alpha$ the segment $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
 
-In our adaptation to \coq{}, we declare a type \texttt{Ord}, a binary relation \texttt{lt} (with infix notation \texttt{"\_<\_"}, and assume Schütte's axiom. In Chapter~\ref{chap:schutte},
-we derive some interesting properties of countable ordinals from these axioms.
-It is interesting to compare proofs of a given property (for instance the associativity of addition) in the computational framework of some ordinal notation, and in the axiomatic model of Schütte.
+% In our adaptation to \coq{}, we declare a type \texttt{Ord}, a binary relation \texttt{lt} (with infix notation \texttt{"\_<\_"}, and assume Schütte's axiom. In Chapter~\ref{chap:schutte},
+% we derive some interesting properties of countable ordinals from these axioms.
+
+It is interesting to compare proofs of a given property (for instance the associativity of addition) both in the computational framework of some ordinal notation, and in the axiomatic model of Schütte.
 
@@
 Let $\alpha$ be some (countable) ordinal; 
-in \coq{} terms, we call ordinal notation for $\alpha$ a structure composed 
+in \coq{} terms, we call \emph{ordinal notation for $\alpha$} a structure composed 
 of:
@@
 
-Such a structure can be proved correct relatively to another ordinal notation or
+Such a structure can be proved correct relatively to a bigger ordinal notation or
 to Schütte's model.
-
-\subsubsection*{Ordered Types}
-
-The library ~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Classes.RelationClasses.html}{%
-Coq.Classes.RelationClasses} contains some definitions and facts about binary relations, among them strict orders.
-
-\begin{Coqsrc}
-Variable A: Type.
-
-  Class StrictOrder (R : relation A) : Prop := {
-    StrictOrder_Irreflexive :> Irreflexive R ;
-    StrictOrder_Transitive :> Transitive R }.
-\end{Coqsrc}
 
@@
 \texttt{Datatypes.CompareSpec}.
+
+First, we import the definition of strict orders from
+the library ~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Classes.RelationClasses.html}{%
+Classes.RelationClasses}, 
+and the inductive specification of a comparison from
+\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Datatypes.html}{%
+Init.Datatypes}.
+
+\begin{Coqsrc}
+Variable A: Type.
+
+  Class StrictOrder (R : relation A) : Prop := {
+    StrictOrder_Irreflexive :> Irreflexive R ;
+    StrictOrder_Transitive :> Transitive R }.
+\end{Coqsrc}
 
@@
 
-The following definitions allow us to make implicit several guessable arguments.
+%The following definitions allow us to make implicit several guessable arguments.
 \label{sect:on-lt-notation}
@@
            {on : ON lt compare} := lt.
+
 Infix "o<" := ON_lt : ON_scope.
@@
 \begin{remark}
-The infix notations \texttt{o<} and \texttt{o<=} were defined in order to make apparent the distinction between the various notation scopes that may co-exist in a same statement.
+The infix notations \texttt{o<} and \texttt{o<=} are defined in order to make apparent the distinction between the various notation scopes that may co-exist in a same statement.
 So the infix \texttt{<} and \texttt{<=} are reserved to the natural numbers.
@@
 
-The following lemma (together with the type class mechnism) allows us to use simply  measures towards an ordinal notation.
+The following lemma (together with the type class mechnism) allows us to define termination measures over any ordinal notation.
 It is just an application of  the libraries \texttt{Coq.Wellfounded.Inverse\_Image}
@@
 
-The simplest example of ordinal notation is built over the type \texttt{nat} of \coq's standard library. We have only to apply already proven lemmas about Peano numbers.
+The simplest example of ordinal notation is built over the type \texttt{nat} of \coq's standard library. We just have to apply already proven lemmas about Peano numbers.
 
@@
 helps us to prove that our order \texttt{lt} is well-founded.
+
+\begin{Coqsrc}
+Lemma lt_wf : well_founded lt.
+Proof. 
+  destruct NA, NB; apply wf_disjoint_sum; [apply wf | apply wf0].
+Qed.
+\end{Coqsrc}
+
+Then, we register an instance of \texttt{ON}:
 
@@
 
-Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily a total function which merges two lists.
+Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily a total function which merges two lists (example contributed by Pascal Manoury).
 
@@
 
-\begin{Coqsrc}
-(* adapted from Pascal Manoury et al. *)
-
-Require Import Coq.Program.Wf List.
-Require Import FunInd Recdef.
-
+\vspace{4pt}
+\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
+
+\begin{Coqsrc}
 Section Merge.
@@
   - intros; unfold m, measure_lt; cbn; destruct ys0; simpl; right; lia.
-   - auto.
+  - apply wf_measure.
   Defined.
@@
 
-In Sect.~\vref{omega-case}, we proved that there exists no variant towards \texttt{nat}
+In Sect.~\vref{omega-case}, we proved that there exists no variant from \texttt{Hydra} to \texttt{(nat,$<$)}
 (\emph{i.e.} the ordinal $\omega$) for proving the termination of all hydra battles.
@@
 
-The proof we are going to develop has exactly the same structure as in Section~\ref{omega-case}.
+The proof we are going to comment has exactly the same structure as in Section~\ref{omega-case}.
  Nevertheless, the proof of technical  lemmas is a little more complex, due to 
@@
 
-Let us follow the same pattern as in Sect.~\ref{omega-case}.
+We  follow the same pattern as in Sect.~\ref{omega-case}.
 First, we define an injection $\iota$ from type \texttt{t} into \texttt{Hydra}, by
@@
 
- In a second step, we build a ``smaller'' hydra.
+ In a second step, we build a ``smaller'' hydra\footnote{With respect to the measure $m$.}.
 
@@
 
-If \texttt{OB} is presumed to be correct, the we can consider that \texttt{OA} ``inherits'' its correcteness from the bigger notation system \texttt{OB}.
-
-\label{types:SubON}
-
-\begin{Coqsrc}
-Class  SubON 
-       `(OA : @ON A ltA  compareA)
-       `(OB : @ON B ltB  compareB)
-       (alpha :  B)
-       (iota : A -> B):=
-  {
-  SubON_compare: forall x y : A,  compareB (iota x) (iota y) =
-                                 compareA x y;
-  SubON_incl : forall x, ltB (iota x) alpha;
-  SubON_onto : forall y, ltB y alpha  -> exists x:A, iota x = y}.
-\end{Coqsrc}
-
-For instance, we prove that \texttt{Omega} is a sub-notation of
-\texttt{Omega\_plus\_Omega} (with $\omega$ as the first ``new'' ordinal, and \texttt{fin} as the injection).
-
-\begin{Coqsrc}
-Instance Incl : SubON Omega Omega_plus_Omega omega fin.
-\end{Coqsrc}
-
-We can also show that, if $i<j$, then the segment $[0,i)$ is a ``sub-segment'' of
-$[0,j)$. Since the terms  ($t\;i$) and ($t\;j$) are not convertible, we consider a ``cast'' 
-function $\iota$ from ($t\;i$) into ($t\;j$), and prove that this function is  a monotonous bijection  from ($t\;i$) to
-the segment $[0,i)$ of ($t\;j$).
-
 \begin{figure}[h]
@@
 \node (Bdots) at (6,2) {$\ldots$};
-\node (b) at (8,2) [label=above:$b$]{$\bullet$};
-\node (bsucc) at (9,2) [label=above:$b+1$]{$\bullet$};
+\node (b) at (8,2) [label=above:$\alpha$]{$\bullet$};
+\node (bsucc) at (9,2) [label=above:$\alpha+1$]{$\bullet$};
 \node (Bdots2) at (10,2) {$\ldots$};
@@
 
+If \texttt{OB} is presumed to be correct, the we can consider that \texttt{OA} ``inherits'' its correcteness from the bigger notation system \texttt{OB}.
+
+\label{types:SubON}
+
+\begin{Coqsrc}
+Class  SubON 
+       `(OA : @ON A ltA  compareA)
+       `(OB : @ON B ltB  compareB)
+       (alpha :  B)
+       (iota : A -> B):=
+  {
+  SubON_compare: forall x y : A,  compareB (iota x) (iota y) =
+                                 compareA x y;
+  SubON_incl : forall x, ltB (iota x) alpha;
+  SubON_onto : forall y, ltB y alpha  -> exists x:A, iota x = y}.
+\end{Coqsrc}
+
+For instance, we prove that \texttt{Omega} is a sub-notation of
+\texttt{Omega\_plus\_Omega} (with $\omega$ as the first ``new'' ordinal, and \texttt{fin} as the injection).
+
+\begin{Coqsrc}
+Instance Incl : SubON Omega Omega_plus_Omega omega fin.
+\end{Coqsrc}
+
+We can also show that, if $i<j$, then the segment $[0,i)$ is a ``sub-segment'' of
+$[0,j)$. Since the terms  ($t\;i$) and ($t\;j$) are not convertible, we consider a ``cast'' 
+function $\iota$ from ($t\;i$) into ($t\;j$), and prove that this function is  a monotonous bijection  from ($t\;i$) to
+the segment $[0,i)$ of ($t\;j$).
+
 \index{Coq!Commands!Program}
@@
 
-\subsubsection*{Remark}
-\label{sec:orgheadline66}
-
+\paragraph{Remark}
 The name \texttt{T1} we gave to this data-type  is proper to this development and refers
 to a hierarchy of ordinal notations.
-For instance, in~\cite{CantorContrib}, the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{schutte}).
-
-\emph{Please look also at the library
-\href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2.html}}.
-
+For instance, in Library \href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2},  the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{schutte}).
+\noindent
 \begin{Coqsrc}
@@
 \subsubsection{The Ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
-\label{sec:orgheadline70}
-
+\label{sect:notation-phi0}
 We provide also a notation for ordinals of the form $\omega^\alpha$.
-
-\label{sect:notation-phi0}
 
@@
 
-The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximed by the sequence of $\omega$-towers (see also Sect~\vref{sect:epsilon0-as-limit} ).
+The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximated by the sequence of $\omega$-towers (see also Sect~\vref{sect:epsilon0-as-limit} ).
 
@@
 \end{figure}
+
+\subsection{Pretty Printing Ordinals in Cantor Normal Form}
+\label{sect:ppT1}
+
+Let us consider again the ordinal $\alpha_0$ defined in section~\vref{alpha0-def}
+If we ask \coq{} to print its  normal form, we get a hardly readable term of type \texttt{T1}.
+
+\begin{Coqsrc}
+Compute alpha_0.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+  = ocons omega 0 (ocons (FS 2) 4 (FS 1))
+     : T1
+\end{Coqanswer}
+
+The following data type defines an abstract syntax for more readable ordinals terms in Cantor normal form:
+
+\label{types:ppT1}
+\index{Functions!pp@ pp (pretty printing terms in Cantor normal form)}
+
+\begin{Coqsrc}
+Inductive ppT1 : Set :=
+    P_fin : nat -> ppT1
+  | P_add : ppT1 -> ppT1 -> ppT1
+  | P_mult : ppT1 -> nat -> ppT1
+  | P_exp : ppT1 -> ppT1 -> ppT1
+  | P_omega : ppT1
+\end{Coqsrc}
+
+The function \texttt{pp: T1 -> ppT1} converts any closed term of type \texttt{T1} into a human-readable expression. For instance, let us convert the term \texttt{alpha\_0}.
+
+\begin{Coqsrc}
+Compute pp alpha_0.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+     = (omega ^ omega + omega ^ 3 * 5 + 2)%pT1
+     : ppT1
+\end{Coqanswer}
+
+\index{Projects}
+\begin{project}
+Design  (in \ocaml?) a set of tools for systematically pretty printing ordinal terms in Cantor normal form.
+\end{project}
 
@@
 \label{Predicates:lt-T1}
-Please note that this definition of \texttt{lt} makes it easy to write proofs by reflection, as shown by the following exampgles.
+Please note that this definition of \texttt{lt} makes it easy to write proofs by computation, as shown by the following examples.
 
@@
 in the structure (size and/or height) of the terms of \texttt{T1}.
-For instance the ordinal of Fig \ref{fig:cnf-example} is strictly less
-than the structurally simpler \(\omega^{\omega^\omega}\times 2\).
+
+\begin{Coqsrc}
+Example Ex0:
+  lt (ocons (phi0 (phi0 omega)) 2
+            (ocons (phi0 10) 33
+                   (ocons (phi0 9) 63 zero)))
+     (ocons  (phi0 (phi0 omega)) 2 (phi0 (phi0 11))).
+Proof.
+  reflexivity. 
+Qed.
+\end{Coqsrc}
 
@@
 We note that the exponents of $\omega$ are not in the right (strictly decreasing) order.
-
-With the help of the order \texttt{lt} on \texttt{T1}, we are now able to characterize
+Nevertheless, with the help of the order \texttt{lt} on \texttt{T1}, we are now able to characterize
 the set of all well-formed ordinal terms:
@@
 \vspace{4pt}
+\noindent
 \emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#nf_b}{Epsilon0.T1}}
@@
 \end{exercise}
-For upgrading constants and fonctions of \texttt{T1}, we have to prove that 
+
+In order to  upgrade constants and fonctions from type \texttt{T1} to \texttt{E0}, we have to prove that 
 the term they build is in normal form.
@@
 
-Pattern matching and structural recursion allow us to define the notions of successor and limit ordinal with the help of  boolean functions on type \texttt{T1}.
+Pattern matching and structural recursion allow us to define boolean characterizations  of successor and limit ordinals.
 
 \vspace{4pt}
+\noindent
 \emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}
@@
 \begin{Coqsrc}
-  Fixpoint succb alpha :=
+Fixpoint succb (alpha:T1) : bool :=
   match alpha with
-      zero => false
+    | zero => false
     | ocons zero _ _ => true
@@
 
-Fixpoint limitb alpha :=
+Fixpoint limitb (alpha : T1) : bool :=
   match alpha with
-      zero => false
+    | zero => false
     | ocons zero _ _ => false
@@
 
-\subsection{Pretty Printing Ordinals in Cantor Normal Form}
-\label{sect:ppT1}
-
-Let us consider again the ordinal $\alpha_0$ defined in section~\vref{alpha0-def}
-If we ask \coq{} to print its  normal form, we get a hardly readable term of type \texttt{T1}.
-
-\begin{Coqsrc}
-Compute alpha_0.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = ocons omega 0 (ocons (FS 2) 4 (FS 1))
-     : T1
-\end{Coqanswer}
-
-The following data type defines an abstract syntax for more readable ordinals terms in Cantor normal form:
-
-\label{types:ppT1}
-\index{Functions!pp@ pp (pretty printing terms in Cantor normal form)}
-
-\begin{Coqsrc}
-Inductive ppT1 : Set :=
-    P_fin : nat -> ppT1
-  | P_add : ppT1 -> ppT1 -> ppT1
-  | P_mult : ppT1 -> nat -> ppT1
-  | P_exp : ppT1 -> ppT1 -> ppT1
-  | P_omega : ppT1
-\end{Coqsrc}
-
-The function \texttt{pp: T1 -> ppT1} converts any closed term of type \texttt{T1} into a human-readable expression. For instance, let us convert the term \texttt{alpha\_0}.
-
-\begin{Coqsrc}
-Compute pp alpha_0.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-     = (omega ^ omega + omega ^ 3 * 5 + 2)%pT1
-     : ppT1
-\end{Coqanswer}
-
-\index{Projects}
-\begin{project}
-Design  (in \ocaml?) a set of tools for systematically pretty printing ordinal terms in Cantor normal form.
-\end{project}
+The function \texttt{succ} is related with addition through the following lemma:
+
+\begin{Coqsrc}
+Lemma succ_is_plus_one (alpha : T1) :  succ alpha = alpha + 1.
+Proof.
+  induction alpha as [|a IHa n b IHb].
+  (* ... *)
+\end{Coqsrc}
 
@@
 \begin{Coqsrc}
+Lemma plus_nf : forall a,  nf a -> forall b, nf b -> nf (a + b).
+Proof.
+ (* ... *)
+
 Instance plus (alpha beta : E0) : E0.
@@
 \end{Coqanswer}
+
+\begin{remark}
+In all this development, two representations of ordinals co-exist: ordinal terms (type \texttt{T1}, notation scope \texttt{t1\_scope}, for reasoning on the tree-structure of Cantor normal forms), and ordinal terms \emph{known to be in normal form} (type \texttt{E0}, notation scope \texttt{E0\_scope}).
+Looking at the contexts displayed by \coq{} prevents you from any risk of confusion.
+\end{remark}
 
@@
 \item Prove that \emph{being infinite} is decidable on type \texttt{E0}.
-\item Prove that $\alpha$ is infinite if and only if for all finite ordinal $i$,
+\item Prove that $\alpha$ is infinite if and only if for any finite ordinal $i$,
 $i+\alpha=\alpha$.
@@
 Let \texttt{n:nat} and \texttt{beta:T1} such that \texttt{ocons alpha n beta} is in normal form. 
-We prove first that \texttt{beta} is accessible,  which allows us to by well-founded induction on \texttt{beta}, 
+We prove first that \texttt{beta} is accessible,  which allows us to prove by well-founded induction on \texttt{beta}, 
 and natural induction on \texttt{n}, that (\texttt{ocons alpha n beta}) is accessible.

### 2020-12-21 doc/part-hydras.tex
deprecating the predicate lt_phi0

@@
 The following lemma is an application of the strict inequality 
-\showmath{\alpha < \omega ^\alpha}. If \showmath{\omega^\alpha} is accessible,
-then \showmath{\alpha} is \emph{a fortiori} accessible.
+\showmath{\alpha < \omega ^\alpha}. If \showmath{\alpha} is strongly accessible, then, by definition, \showmath{\omega^\alpha} is accessible, thus \showmath{\alpha} is \emph{a fortiori} accessible.
 
@@
 The proof is structured as an induction on \showmath{\alpha}'s accessibility. Let us consider
-an accessible term $\alpha$.
+any  accessible term $\alpha$.
 
@@
 
-Let \texttt{n:nat} and \texttt{beta:T1} such that \texttt{ocons alpha n beta} is in normal form. 
+Let \texttt{n:nat} and \texttt{beta:T1} such that (\texttt{ocons alpha n beta}) is in normal form. 
 We prove first that \texttt{beta} is accessible,  which allows us to prove by well-founded induction on \texttt{beta}, 
 and natural induction on \texttt{n}, that (\texttt{ocons alpha n beta}) is accessible.
-The proof, quite long, can be consulted in \url{../theories/html/hydras.Epsilon0.T1.html}
+The proof, quite long, can be consulted in \href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}.
 
@@
 
+In order to prove the termination of any hydra battle, we try to define a variant mapping hydras to ordinals strictly less than $\epsilon_0$.
+In order to make such a variant easy to define (for instance by a structural recursion), we introduce a variant of addition, which, contrary to
+$+$, is commutative and strictly monotonous in both of its arguments. This last property makes it possible to prove that our function is 
+truly a variant for hydra battles (in Sect.~\vref{sect:variant-decr}).
+
 \subsection{Natural Sum (a.k.a. Hessenberg's  Sum)}
@@
 The definition of \texttt{oplus} is recursive in both of its 
-arguments, which makes a structural recursive definition a little 
-complex.
-We used the same pattern as for the \texttt{merge} function on lists of library
+arguments and uses  the same pattern as for the \texttt{merge} function on lists of library
 \texttt{Coq.Sorting.Mergesort}.
@@
 Proofs of properties of this function can be simpler if we
- \emph{derive} rewriting lemmas that will help to simplify 
-expressions of the form (\texttt{oplus $a$ $ b$}).
+ derive a few  rewriting lemmas that will help to simplify 
+expressions of the form (\texttt{oplus $\alpha$ $\beta$}).
 
 A first set of lemmas correspond to the various cases of \texttt{oplus}'s 
-definition. They can be proved almost immediately, using \texttt{cbn} 
-and \texttt{reflexivity} tactics.
+definition. They can be proved almost immediately.
 
@@
 \begin{Coqsrc}
-Lemma lt_phi0_oplus : forall gamma alpha beta,
-                        lt_phi0 alpha gamma ->
-                        lt_phi0 beta gamma ->
-                        lt_phi0 (alpha o+ beta) gamma.
-
-Proof with auto.
-  induction gamma; destruct alpha, beta.  
-(* ... *)
+Lemma oplus_bounded_phi0 alpha beta gamma :
+  nf alpha -> nf beta -> nf gamma ->
+  lt alpha (phi0 gamma) ->
+  lt beta (phi0 gamma) ->
+  lt (alpha o+ beta) (phi0 gamma).
 \end{Coqsrc}
@@
 \texttt{m} is a variant. First, a few technical lemmas follow the decomposition of \texttt{round} into several relations. Then the lemma \texttt{round\_decr} gathers all the cases.
+
+\label{sect:variant-decr}
 

### 2020-12-25 doc/part-hydras.tex
minor modifications

@@
 
+This lemma helps us
+
 \subsubsection{Commutativity, Associativity}
@@
 
-\chapter[The Ketonen-Solovay machinery]{Strolling inside \texorpdfstring{$\epsilon_0$}{Epsilon0}: The Ketonen-Solovay Machinery\label{ks-chapter}}
+\chapter[The Ketonen-Solovay machinery]{Accessibility inside \texorpdfstring{$\epsilon_0$}{Epsilon0}: The Ketonen-Solovay Machinery\label{ks-chapter}}
 \label{chap:ketonen}
@@
 In this chapter, we focus on a specific class of proofs of termination of hydra battles: construction of some variant mapping the type \texttt{Hydra} into a given initial  segment of ordinals.
-Our proof relies only on the Calculus of Inductive Constructions and is a natural complement of the results proven in the previous chapter.
+Our proof relies only on the Calculus of Inductive Constructions and is a natural complement of the results proven in the previous chapters.
 
@@
 
-We did not  try to prove that our definition really implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
+We did not  try to prove that our definition truly implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
 abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
@@
 
-\subsubsection{Limit Ordinals are Really Limits}
+\subsubsection{Limit Ordinals are truly Limits}
 The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
@@
   nf lambda -> limitb lambda  ->
-  strict_lub (fun i => canonS lambda i) lambda.
+  strict_lub (canonS lambda) lambda.
 \end{Coqsrc}
@@
 Of course, you will have to prove the correctness of your function.
+
+\textbf{Hint:} You may add to your function a third argument usually called \texttt{fuel} for allowing you to give a structurally 
+recursive function (\emph{cf} the post of Guillaume Melquiond on Coq-club (Dec 21, 2020)
+\url{https://sympa.inria.fr/sympa/arc/coq-club/2020-12/msg00069.html}).
+The type \texttt{fuel}, an alternative 
+to \texttt{nat} is available on \href{../theories/html/hydras.Prelude.Fuel.html}{Prelude.Fuel}).
+\index{Coq!Techniques!Giving fuel to a long computation}
+
 \end{exercise}
@@
   Lemma m_lt : m small_h  t1< m big_h.
+
+  Theorem Impossibility_std: False.
+
+End Impossibility_Proof.
 \end{Coqsrc}

### 2020-12-26 doc/part-hydras.tex
better definition of BoundedVariant (phase 2)

@@
 \begin{Coqsrc}
-Class BoundedVariant (B:Battle) :=
+Class BoundedVariant {A:Type}{Lt:relation A}
+      {Wf: well_founded Lt}{B : Battle}
+      {m: Hydra -> A} (Var: Hvariant  Wf  B m)(mu:A):=
   {
-    mu:T1 ;
-    m: Hydra -> T1;
-    mu_nf: nf mu;
-    Hvar: Hvariant T1_wf B m;
-    m_bounded: forall h, m h t1< mu
+  m_bounded: forall h, Lt (m h ) mu
   }.
@@
 \begin{Coqsrc}
-Section Bounded.
-  Context (B: Battle) (Hy : BoundedVariant B).
+Context (B: Battle)
+          (mu: T1)
+          (Hmu: nf mu)
+          (m : Hydra -> T1)
+          (Var : Hvariant  T1_wf B m)
+          (Hy : BoundedVariant  Var mu).
 
@@
 
-  Context (Var : BoundedVariant free ).
+  Context (mu: T1)
+          (Hmu: nf mu)
+          (m : Hydra -> T1)
+          (Var : Hvariant  T1_wf free m)
+          (Hy : BoundedVariant Var mu).
   \end{Coqsrc}
@@
 Impossibility_free
-     : BoundedVariant free -> False
+     : forall (mu : T1) (m : Hydra -> T1) (Var : Hvariant T1_wf free m),
+       BoundedVariant Var mu -> False
 \end{Coqanswer}
@@
 \href{../theories/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
-{\texttt{Lemma2\_6\_1}}   and 
+{\texttt{Epsilon0.Paths}}   and 
 \href{../theories/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
@@
 $\alpha$ to  $\beta$, then into a fragment of a standard battle from 
-$\iota(\alpha)$ to $\iota(\beta)$.
+$\iota(\alpha)$ to $\iota(\beta)$, hence the inequality $m(\iota(\beta))<m(\iota(\alpha))$.
 
@@
 
-  Context (Var : BoundedVariant standard).
+Context (mu: T1)
+          (Hmu: nf mu)
+          (m : Hydra -> T1)
+          (Var : Hvariant  T1_wf standard m)
+          (Hy : BoundedVariant Var mu).
  \end{Coqsrc}

### 2020-12-28 doc/part-hydras.tex
minor modifications

@@
 
-The next chapter is dedicated to a generic implementation of ordinal notations, and chapter~\ref{chap:T1} to a proof of termination of all hydra battles with the help of an ordinal notation for the interval $[0,\epsilon_0)$.
+The next chapter is dedicated to a generic formalization of ordinal notations, and chapter~\ref{chap:T1} to a proof of termination of all hydra battles with the help of an ordinal notation for the interval $[0,\epsilon_0)$\,\footnote{We use the mathematical notation $[a,b)$ for the interval $\{x|a\leq x < b\}$.}.
+\index{Maths!Notations!Interval}
 
@@
 
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}}
+\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{ON\_Generic}}
 \begin{Coqsrc}
@@
 
- Context (Hvar : Hvariant (Generic.wf (ON:=Omega2))  free m).
+ Context (Hvar : Hvariant (ON_Generic.wf (ON:=Omega2))  free m).
 \end{Coqsrc}
@@
 In our formalization, the interpretation of an ordinal as a set is realized by the following definition
-(in ~\href{../theories/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}).
+(in ~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{ON\_Generic}).
 
@@
 
-In the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
+In the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
 This part of the development is parameterized by some class $B$ of battles, which will be instantiated later to \texttt{free} or \texttt{standard}.
@@
 from 
-the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}.
+the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Epsilon0\_Needed\_Generic}.
 
@@
 
-On may ask whether the preceding results are not too restrictive, since they 
+On may ask whether the preceding results are too restrictive, since they 
 refer to a particular data type \texttt{T1}.

### 2020-12-31 doc/part-hydras.tex
minor modifications

@@
 The following class specifies a strict inclusion of segments.
-The notation \texttt{OA} describes a segment $[0,\alpha($, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented in \texttt{OA}).
+The notation \texttt{OA} describes a segment $[0,\alpha)$, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented in \texttt{OA}).
 We require also  that the comparison functions of the two notation systems are compatible.
@@
         | Datatypes.Gt => Schutte_basics.lt (iota b) (iota a)
-        end}.
+        end
+}.
 \end{Coqsrc}
@@
   iso_inv1 : forall a, g (f a)= a;
-  iso_inv2 : forall b, f (g b) = b}.
+  iso_inv2 : forall b, f (g b) = b
+}.
 \end{Coqsrc}
@@
 \begin{exercise}
+\label{exo:i-plus-omega}
 Let $i$ be some natural number. Prove that the notation systems 
@@
 {\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$, that we prove also in larger notation systems, as well as in Schütte's model.}
-
 This exercise is partially solved for $i=3$ (in ~\href{../theories/html/hydras.OrdinalNotations.Example_3PlusOmega.html}{OrdinalNotations.Example\_3PlusOmega}).
@@
 for any strictly positive $i$.
+
+\textbf{Note:} Like Exercise~\ref{exo:i-plus-omega}, this project corresponds to the [in]equalities $i+\omega=\omega<\omega+i$, for any natural number $i$.
 \end{project}

### 2021-01-14 doc/chapter-primrec.tex
add a new chapter on primitive recursive functions

\chapter{Primitive Recursive Functions}

\section{Introduction}
The definition of primitive recursive functions we use come
from Russel O'Connors formalization in \coq{} of
G\"odel's incompleteness theorems~\cite{OConnor05}, now hosted in
the \texttt{theories/ordinals/Ackermann} directory.

This chapter contains some comments on Russel's library, as well as a few extensions.

\section{First look at the Ackermann library}

O'Connor's library on Gödel's incompleteness theorems contains a little more 
than 45K lines of scripts. The part dedicated to primitive recursive functions and Peano arithmetics is 32K lines long and is originally structured in 38 modules.

Thus, we propose a slow exploration of this library, through examples and exercises.

\subsection{Basic Libraries}

The formal definition of primitive recursive functions lies in the library
\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec},
with preliminary definitions in 
\href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat}
and
\href{../theories/html/hydras.Ackermann.misc.html}{Ackermann.misc}.

\subsubsection{Functions of arbitrary arity}

The library allows us to consider primitive functions on \texttt{nat}, with any number of arguments, in 
curried form. This is made possible in 
\href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat} by the following definition:

\begin{Coqsrc}
Fixpoint naryFunc (n : nat) : Set :=
  match n with
  | O => nat
  | S n => nat -> naryFunc n
  end.
\end{Coqsrc}

For instance (\texttt{naryFunc 1}) is convertible to \texttt{nat -> nat} and (\texttt{naryFunc 3})
to \texttt{nat -> nat -> nat -> nat}.

\vspace{4pt}
\noindent
\emph{From\href{../theories/html/hydras.Ackermann.PrimRecExamples.html}{Ackermann.PrimRecExamples}}.
\begin{Coqsrc}
Require Import primRec.
Import extEqualNat.

Compute naryFunc 3.
\end{Coqsrc}

\begin{Coqanswer}
= nat -> nat -> nat -> nat
  : Set  
\end{Coqanswer}

Thus, the function \texttt{plus} has also type (\texttt{naryFunc 2}).

Likewise, arbitrary boolean predicates may have an arbitrary number of arguments. The dependent type
(\texttt{naryRel $n$}), defined in the same way as \texttt{naryFunc}, is the type of $n$-ary functions from
\texttt{nat} into \texttt{bool}.

\begin{Coqsrc}
Compute naryRel 2.
\end{Coqsrc}

\begin{Coqanswer}
 = nat -> nat -> bool
     : Set
\end{Coqanswer}

The magic of dependent types makes it possible to define recursively extensionnal equality between functions of the same arity.

\vspace{4pt}
\noindent
\emph{From \href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat}}
\begin{Coqsrc}
Fixpoint  extEqual (n : nat) : forall  (a b : naryFunc n), Prop :=
  match n with
    0 => fun a b => a = b
  | S p => fun a b => forall c, extEqual p (a c) (b c)
  end.
\end{Coqsrc}

\begin{Coqsrc}
Compute extEqual 2.
\end{Coqsrc}

\begin{Coqanswer}
     = fun a b : naryFunc 2 => forall x x0 : nat, a x x0 = b x x0
     : naryFunc 2 -> naryFunc 2 -> Prop
 \end{Coqanswer}

\begin{Coqsrc}
Example extEqual_ex1 : extEqual 2 mult (fun x y =>  y * x + x -x) .
Proof.
  intros x y.
\end{Coqsrc}

\begin{Coqanswer}
  x, y : nat
  ============================
  extEqual 0 (x * y) (y * x)
\end{Coqanswer}

\begin{Coqsrc}
  cbn.
\end{Coqsrc}

\begin{Coqanswer}
1 subgoal (ID 10)

  x, y : nat
  ============================
  x * y = y * x + x - x
\end{Coqanswer}

\begin{Coqsrc}
  rewrite <- Nat.add_sub_assoc, Nat.sub_diag.
  - ring.
  - apply le_n.  
Qed.
\end{Coqsrc}

### 2021-01-14 doc/hydras.tex
minor modifications
add a new chapter on primitive recursive functions

@@
 
-\author{Pierre Castéran\\ LaBRI, Univ. Bordeaux, CNRS UMR 5800, Bordeaux-INP\\with contributions by Yves Bertot, \'Evelyne Contejean,  Florian Hatat and Pascal Manoury}
+\author{Pierre Castéran\\ LaBRI, Univ.
+Bordeaux, CNRS UMR 5800, Bordeaux-INP\\with contributions by Yves Bertot, \'Evelyne Contejean,  Florian Hatat and Pascal Manoury\\ \emph{The formalization of primitive recursive functions was originally authored by Russel O'Connor}}
 \date{\today}
@@
 
+The formalization of primitive recursive functions was originally a part of  Russel O'Connor's work on G\"odel's incompleteness theorems~\cite{OConnor05}.
+
 \label{sec:orgheadline2}
@@
 \include{part-hydras}
-
+\include{chapter-primrec}
 \part{Some Funny Algorithms}
@@
 
-\chapter*{Index and tables}
+\chapter{Index and tables}
 

### 2021-01-15 doc/chapter-powers.tex
comments and examples about the Ackermann part (to continue)

@@
 \label{op-classes}
-\index{Type Classes!Operational Type Classes}
+\index{Coq!Type Classes!Operational Type Classes}
 \emph{Operational type classes}~\cite{BS2011}
@@
 \subsection{The Monoid Type Class}
-\index{Type Classes}
+\index{Coq!Type Classes}
 We are now ready to  give a definition of the \texttt{Monoid} class, using
@@
 \subsection{Monoids and Equivalence Relations}
-\index{Generalized rewriting}
-\index{Equivalence relations}
+\index{Coq!Generalized rewriting}
+\index{Coq!Equivalence relations}
 
@@
 
-\index{Type Classes!Proper}
+\index{Coq!Type Classes}
+\index{Coq!Proper}
 \label{EMonoid-def}
@@
 
-\index{Coercions}
+\index{Coq!Coercions}
 
@@
 \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Classes.Morphisms.html}{Coq.Classes.Morphisms}
-\index{Type Classes!Proper}
-
+\index{Coq!Type Classes}
+\index{Coq!Proper}
 \begin{Coqsrc}
@@
 \label{reflection-section}
-\index{Advanced proof techniques!Proof by reflection}
+\index{Coq!Proof by reflection}
 Instead of letting the tactic \texttt{rewrite} look for contexts in which
@@
 \subsubsection{About parametricity}
-\index{Advanced proof techniques!Parametricity}
+\index{Coq!Parametricity}
 Let us now present another tactic for proving chain correctness,
@@
 \subsubsection{Definition:} 
-\index{Type Classes!Proper}
-
+\index{Coq!Type Classes}
+\index{Coq!Proper}
 A continuation \texttt{k:Fkont A} is \emph{proper}
@@
 any correct chain generator.
-\index{Dependently Typed Functions}
+\index{Coq!Dependently Typed Functions}
 

### 2021-01-15 doc/chapter-primrec.tex
comments and examples about the Ackermann part (to continue)

@@
 
-\subsection{Basic Libraries}
+\section{Basic Libraries}
 
@@
 
-\subsubsection{Functions of arbitrary arity}
+\subsection{Functions of arbitrary arity}
 
@@
 
-The magic of dependent types makes it possible to define recursively extensionnal equality between functions of the same arity.
+The magic of dependent types makes it possible to define recursively extensional equality between functions of the same arity.
 
+\index{Coq!Dependent types}
+\index{Coq!Dependently typed functions}
 \vspace{4pt}
@@
 \end{Coqsrc}
+
+\subsection{A Data-type for Primitive Recursive Functions}
+
+The traditional definition of the set of primitive recursive functions is structured as an inductive definition 
+in five rules: three base cases, and two construction rules. Primitive recursive functions are total functions from $\mathbb{N}^n$ to
+$\mathbb{N}$, for some $n\in\mathbb{N}$.
+
+\begin{description}
+  \item[zero] the constant function of value $0$ is primitive recursive.
+\item[S] The successor function $S:\mathbb{N}\rightarrow\mathbb{N}$ is primitive recursive.
+ \item[projections] For any pair $0\leq i<n$, the projection $\pi_{n,i}: \mathbb{N}^n\rightarrow\mathbb{N}$, defined by $\pi_{n,i}(x_0,x_1,\dots,x_{n-1})=x_i$, is primitive recursive.
+\item[composition] For any $n$ and $m$, if $h: \mathbb{N}^m\rightarrow\mathbb{N}$, and
+$g_0,\dots, g_{m-1}$ are primitive recursive of $n$ arguments, then the function which maps any
+tuple $(x_0,\dots,x_{n-1})$ to $h(g_0(x0,\dots,x_{n-1}),\dots, g_{m-1}(x0,\dots,x_{n-1}))$ is primitive recursive.
+\item[primitive recursion]
+If $g: \mathbb{N}^n\rightarrow\mathbb{N}$ and $h: \mathbb{N}^{n+2}\rightarrow\mathbb{N}$ are primitive recursive, then the function from $\mathbb{N}^{n+1}$ into $\mathbb{N}$ defined by
+\begin{align}
+f(0,x_1,\dots,x_n)&=g(x_1,\dots,x_n)\\
+f(S(p),x_1,\dots,x_n)&=h(p,f(p, x_1,\dots,x_n),  x_1,\dots,x_n)
+\end{align} 
+is primitive recursive.
+\end{description}
+
+O'Connor's formalization of primitive recursive functions takes the form of two mutually inductive dependent data types, each constructor of which is associated with one of these  rules.
+These two types are (\texttt{PrimRec $n$}) (primitive recursive functions of $n$ arguments), and
+(\texttt{PrimRecs $n$ $m$}) ($m$-tuples of primitive recursive functions of $n$ arguments).
+
+\index{Coq!Dependent types}
+\index{Coq!Mutually inductive types}
+
+\vspace{4pt}
+\noindent
+\emph{From \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}}
+\begin{Coqsrc}
+Inductive PrimRec : nat -> Set :=
+  | succFunc : PrimRec 1
+  | zeroFunc : PrimRec 0
+  | projFunc : forall n m : nat, m < n -> PrimRec n
+  | composeFunc :
+      forall (n m : nat) (g : PrimRecs n m) (h : PrimRec m), PrimRec n
+  | primRecFunc :
+      forall (n : nat) (g : PrimRec n) (h : PrimRec (S (S n))), 
+      PrimRec (S n)
+with PrimRecs : nat -> nat -> Set :=
+  | PRnil : forall n : nat, PrimRecs n 0
+  | PRcons : forall n m : nat, PrimRec n -> PrimRecs n m -> 
+                   PrimRecs n (S m).
+\end{Coqsrc}
+
+\subsection{A little bit of Semantics} 
+Please note that inhabitants of type (\texttt{PrimRec $n$}) are not \coq{} functions like \texttt{Nat.mul}, or factorial, etc.
+The data-type (\texttt{PrimRec $n$}) is indeed an abstract syntax for the language of primitive recursive functions.
+The bridge between this language and the word of usual functions
+is an interpretation function (\texttt{evalprimRec $n$})  from  (\texttt{PrimRec $n$}) to  (\texttt{naryFunc $n$}) (with the helper (\texttt{evalprimRecS $n$ $m$})
+from  (\texttt{PrimRecs $n$ $m$}) to  (\texttt{Vector.t (naryFunc $n$) $m$}).
+
+Both functions are mutually defined through dependent pattern matching. We advise the readers who 
+would feel uneasy with dependent types to consult Adam Chlipala's \emph{cpdt}  book~\cite{chlipalacpdt2011}. We leave it to the reader  to look at the helper functions in
+\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.
+
+\begin{Coqsrc}
+Fixpoint evalPrimRec (n : nat) (f : PrimRec n) {struct f} : 
+ naryFunc n :=
+  match f in (PrimRec n) return (naryFunc n) with
+  | succFunc => S
+  | zeroFunc => 0
+  | projFunc n m pf => evalProjFunc n m pf
+  | composeFunc n m l f =>
+      evalComposeFunc n m (evalPrimRecs _ _ l) (evalPrimRec _ f)
+  | primRecFunc n g h =>
+      evalPrimRecFunc n (evalPrimRec _ g) (evalPrimRec _ h)
+  end
+
+ with evalPrimRecs (n m : nat) (fs : PrimRecs n m) {struct fs} :
+ Vector.t (naryFunc n) m :=
+  match fs in (PrimRecs n m) return (Vector.t (naryFunc n) m) with
+  | PRnil a => Vector.nil  (naryFunc a)
+  | PRcons a b g gs =>
+       Vector.cons _ (evalPrimRec _ g) _  (evalPrimRecs _ _ gs)
+  end.
+\end{Coqsrc}
+Let us take as example the following term\footnote{Of course, we never typed this term \emph{verbatim}, but obtained as the result of some computation we leave it the reader to guess and reproduce.}:
+
+\begin{Coqsrc}
+Example weirdPR : PrimRec 1 :=
+primRecFunc 0
+  (composeFunc 0 1 (PRcons 0 0 zeroFunc (PRnil 0)) succFunc)
+  (composeFunc 2 2
+    (PRcons 2 1
+      (composeFunc 2 1
+         (PRcons 2 0 (projFunc 2 1 (le_n 2))
+                 (PRnil 2))
+         succFunc)
+      (PRcons 2 0
+        (composeFunc 2 1
+          (PRcons 2 0
+             (projFunc 2 0
+                       (le_S 1 1 (le_n 1)))
+             (PRnil 2))
+          (projFunc 1 0 (le_n 1))) (PRnil 2)))
+    (primRecFunc 1 (composeFunc 1 0 (PRnil 1) zeroFunc)
+       (composeFunc 3 2
+         (PRcons 3 1
+            (projFunc 3 1 (le_S 2 2 (le_n 2)))
+            (PRcons 3 0 (projFunc 3 0
+                          (le_S 1 2
+                                (le_S 1 1 (le_n 1))))
+                    (PRnil 3)))
+         (primRecFunc 1 (projFunc 1 0 (le_n 1))
+                      (composeFunc 3 1
+                          (PRcons 3 0
+                                  (projFunc 3 1 (le_S 2 2 (le_n 2)))
+                                  (PRnil 3))
+                          succFunc))))). 
+\end{Coqsrc}
+
+Let us now interpret this term as an arithmetic function.
+
+\begin{Coqsrc}
+Example  mystery_fun : nat _> nat := evalPrimRec 1 weirdPR.
+
+Compute List.map mystery_fun (0::1::2::3::4::5::6::nil) : list nat.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+ = 1 :: 1 :: 2 :: 6 :: 24 :: 120 :: 720 :: nil
+     : list nat
+\end{Coqanswer}
+
+Ok, the term \texttt{weirdPR} looks to be a primitive recursive definition of the factorial function, although we haven't proved this fact yet.
+Fortunately, we will see in the following sections much simpler ways to prove that a given function is primitive recursive, whithout looking at an unreadable term.

### 2021-01-18 doc/chapter-primrec.tex
redaction of Ackermann chapter (continued)

@@
 \item[S] The successor function $S:\mathbb{N}\rightarrow\mathbb{N}$ is primitive recursive.
- \item[projections] For any pair $0\leq i<n$, the projection $\pi_{n,i}: \mathbb{N}^n\rightarrow\mathbb{N}$, defined by $\pi_{n,i}(x_0,x_1,\dots,x_{n-1})=x_i$, is primitive recursive.
+ \item[projections] For any pair $0< i\leq n$, the projection $\pi_{i,n}: \mathbb{N}^n\rightarrow\mathbb{N}$, defined by $\pi_{i,n}(x_1,x_2,\dots,x_{n})=x_i$, is primitive recursive.
 \item[composition] For any $n$ and $m$, if $h: \mathbb{N}^m\rightarrow\mathbb{N}$, and
@@
 \noindent
-\emph{From \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}}
+\emph{From \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.}
 \begin{Coqsrc}
@@
 
+\begin{remark}
+Beware of the conventions used in the \texttt{primRec} library!
+The constructor (\texttt{projFunc $n$ $m$})  is associated with the projection $\pi_{n-m,n}$ and \emph{not}
+$\pi_{n, m}$.
+For instance, the projection $\pi_{2,5}$ defined by $\pi_{2,5}(a,b,c,d,e)=b$ corresponds to the term
+(\texttt{projFunc 5 3 H}), where \texttt{H} is a proof of $3<5$.
+ This fact is reported in the comments of \texttt{primRec.v}.
+We presume that this convention makes it easier to define the evaluation function \texttt{evalProjFunc $n$} (see next sub-section) .
+Trying the other conention is left as an exercise.
+\end{remark}
+
 \subsection{A little bit of Semantics} 
@@
 The bridge between this language and the word of usual functions
-is an interpretation function (\texttt{evalprimRec $n$})  from  (\texttt{PrimRec $n$}) to  (\texttt{naryFunc $n$}) (with the helper (\texttt{evalprimRecS $n$ $m$})
-from  (\texttt{PrimRecs $n$ $m$}) to  (\texttt{Vector.t (naryFunc $n$) $m$}).
+is an interpretation function (\texttt{evalprimRec $n$})  of type
+$\texttt{PrimRec}\,n \rightarrow  \texttt{naryFunc}\,n$ (and the helper function 
+(\texttt{evalprimRecS $n$ $m$}) of type 
+$\texttt{PrimRecs}\,n\,m \rightarrow  \texttt{Vector.t}\,(\texttt{naryFunc}\,n)\,m$.
 
 Both functions are mutually defined through dependent pattern matching. We advise the readers who 
-would feel uneasy with dependent types to consult Adam Chlipala's \emph{cpdt}  book~\cite{chlipalacpdt2011}. We leave it to the reader  to look at the helper functions in
+would feel uneasy with dependent types to consult Adam Chlipala's \emph{cpdt}  book~\cite{chlipalacpdt2011}. We leave it to the reader  to look also at the helper functions in
 \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.
@@
 \end{Coqsrc}
-Let us take as example the following term\footnote{Of course, we never typed this term \emph{verbatim}, but obtained as the result of some computation we leave it the reader to guess and reproduce.}:
+Let us take as example the following term\footnote{Of course, we never typed this term \emph{verbatim}; we obtained it as the result of some computation we leave it the reader to guess and reproduce.}:
 
+\label{sect:weirdfac}
 \begin{Coqsrc}
@@
 Fortunately, we will see in the following sections much simpler ways to prove that a given function is primitive recursive, whithout looking at an unreadable term.
+
+\section{Proving that a given arithmetic function is primitive recursive}
+
+The example in the preceding section clearly shows that, in order to prove that a given arithmetic function
+(defined in \gallina{} as usual) is primitive recursive, trying to give  by hand a term  of type \texttt{PrimRec $n$} is not a good method, since such terms may be huge and complex, even for simple arithmetic functions.
+The method proposed in Library \texttt{primRec} is the following one:
+
+\begin{enumerate}
+\item Define a type corresponding to the statement "the function \texttt{$f$:naryFunc $n$} is primitive recursive ''.
+\item Prove handy lemmas which may help to prove that a given function is primitive recursive.
+\end{enumerate}
+
+Thus, the proof that a function, like \texttt{factorial}, is primitive recursive may be interactive, whithout having to type complex terms at any step of the development.
+
+\subsection{The predicate \texttt{isPR}}
+
+Let $f$ be an arithmetic function of arity $n$. We say that $f$ is primitive recursive if $f$ is \textbf{extensionnaly}
+equal to the interpretation of some term of type \texttt{PrimRec $n$}.
+
+\vspace{4pt}
+\noindent
+\emph{From \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.}
+\begin{Coqsrc}
+Definition isPR (n : nat) (f : naryFunc n) : Set :=
+  {p : PrimRec n | extEqual n (evalPrimRec _ p) f}.  
+\end{Coqsrc}
+
+The library \texttt{primRec} contains a large catalogue of lemmas allowing to prove statements 
+of the form (\texttt{isPR $n$ $f$}). We won't list all these lemmas here, but give a few examples of
+how they may be applied.
+
+\begin{remark}
+In the library \texttt{primRec}, all these lemmas are opaque (registered with \texttt{Qed}.
+Thus they do not allow the user to look at the witness of a proof of a \texttt{isPR} statement.
+Our example of page\pageref{sect:weirdfac} was built using a  copy of \texttt{primRec.v} where many \texttt{Qed}s have been replaced with
+\texttt{Defined}s.
+\end{remark}

### 2021-01-18 doc/chapter-primrec.tex
redaction of Ackermann chapter (continued)

@@
 \begin{remark}
+\label{projFunc-order-of-args}
 Beware of the conventions used in the \texttt{primRec} library!
@@
 \end{remark}
+
+\subsubsection{Elementary Proofs of \texttt{isPR} statements}
+
+The constructors \texttt{zeroFunc}, \texttt{succFunc},  and \texttt{projFunc} of type
+\texttt{PrimRec} allows us to write trivial proofs of primitive recursivity. 
+Although  the following lemmas are alreday proven in 
+\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec.v},
+we wrote alternate proofs in 
+\href{../theories/html/hydras.Ackermann.PrimRecExamples.html}%
+{Ackermann.PrimRecExamples.v}, in order to illustrate the main proof patterns.
+
+\begin{Coqsrc}
+Module Alt.
+
+Lemma zeroIsPR : isPR 0 0.
+Proof.
+  exists zeroFunc.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+1 subgoal (ID 90)
+
+  ============================
+  extEqual 0 (evalPrimRec 0 zeroFunc) 0
+\end{Coqanswer}
+
+\begin{Coqsrc}
+  cbn.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+1 subgoal (ID 91)
+
+  ============================
+  0 = 0
+\end{Coqanswer}
+
+\begin{Coqsrc}
+  reflexivity.
+Qed.
+\end{Coqsrc}
+
+Likewise, we prove that the successor function on \texttt{nat} is primitive recursive too.
+
+\begin{Coqsrc}
+Lemma SuccIsPR : isPR 1 S.
+Proof.
+  exists succFunc; cbn; reflexivity.
+Qed.
+\end{Coqsrc}
+
+Projections are proved primitive recursive, case by case (many examples in 
+\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec.v}).
+\emph{Please notice again that the name of the projection follows the mathematical tradition, 
+whilst the arguments of  \texttt{projFunc} use another convention (\emph{cf} remark~\vref{projFunc-order-of-args}).}
+
+\begin{Coqsrc}
+Lemma pi2_5IsPR : isPR 5 (fun a b c d e => b).
+Proof.
+ assert (H: 3 < 5) by auto.
+ exists (projFunc 5 3 H).
+ cbn; reflexivity.
+Qed.
+\end{Coqsrc}
+
+\subsubsection{Using Function Composition}
+
+Let us look at the proof that any constant $n$ of type \texttt{nat} has type (\texttt{PR 0})
+(lemma  \texttt{const1\_NIsPR} of \texttt{primRec}). We carry out a proof by induction on $n$, the base case of which is already proven.
+Now, let us assume $n$ is \texttt{PR $n$}, with $x:\texttt{PrimRec}\,0$ as a ``realizer''.
+Thus we would like to compose this constant function with the unary successor function.
+
+This is exactly the role of the instance \texttt{composeFunc 0 1} of the dependently typed
+function \texttt{composeFunc}, as shown by the following lemma.
+
+\begin{Coqsrc}
+Fact compose_01 :
+    forall (x:PrimRec 0) (t : PrimRec 1),
+    let c := evalPrimRec 0 x in
+    let f := evalPrimRec 1 t in
+    evalPrimRec 0 (composeFunc 0 1
+                               (PRcons 0 0 x (PRnil 0))
+                               t)  =
+     f c.
+Proof. reflexivity. Qed.
+\end{Coqsrc}
+
+Thus, we get a quite simple proof of \texttt{const1\_NIsPR}.
+
+\vspace{4pt}
+\noindent
+\emph{From\href{../theories/html/hydras.Ackermann.PrimRecExamples.html}{Ackermann.PrimRecExamples}}.
+\begin{Coqsrc}
+Lemma  const1_NIsPR n : isPR 0 n. 
+Proof.
+  induction n.
+ - apply zeroIsPR.
+ - destruct IHn as [x Hx].
+   exists (composeFunc 0 1 (PRcons 0 0 x (PRnil 0)) succFunc). 
+   cbn in *; intros; now rewrite Hx.
+Qed.
+\end{Coqsrc}

### 2021-01-18 doc/chapter-primrec.tex
redaction of Ackermann chapter (continued)

@@
 
+Please note that the projection $\pi_{1,1}$ is just the identity on \texttt{nat}, and is realized by 
+(\texttt{projFunc 1 0}).
+
+\vspace{4pt}
+\noindent
+\emph{From \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.}
+
+\begin{Coqsrc}
+Lemma idIsPR : isPR 1 (fun x : nat => x).
+Proof.
+  assert (H: 0 < 1) by auto.
+  exists (projFunc 1 0 H); cbn; auto.
+Qed.
+\end{Coqsrc}
+
 \subsubsection{Using Function Composition}
@@
 \end{Coqsrc}
+
+\begin{todo}
+Comment more examples from   \href{../theories/html/hydras.Ackermann.PrimRecExamples.html}{Ackermann.PrimRecExamples}.
+\end{todo}
+
+\section{Two challenges}
+
+\begin{itemize}
+\item Write a simple and readable proof that the Fibonacci function is primitive recursive.
+
+We give the tail-recursive definition from~\cite{BC04}, chapter 9. See also
+\href{https://coq-community.org/coq-art/ch9_function_specification/SRC/fib_tail.v}{this file}.
+\begin{Coqsrc}
+
+Fixpoint general_fib (a0 a1 n:nat) {struct n} : nat :=
+  match n with
+  | O => a0
+  | S p => general_fib a1 (a0 + a1) p
+  end.
+
+Definition fib_tail (n:nat) := general_fib 1 1 n.
+\end{Coqsrc}
+
+\item Prove formally that the Ackermann function is \emph{not} primitive recursive.
+
+\end{itemize}

### 2021-01-19 doc/chapter-primrec.tex
redaction of Ackermann chapter (continued)

@@
 \noindent
-\emph{From\href{../theories/html/hydras.Ackermann.PrimRecExamples.html}{Ackermann.PrimRecExamples}}.
+\emph{From \href{../theories/html/hydras.Ackermann.PrimRecExamples.html}{Ackermann.PrimRecExamples}}.
 \begin{Coqsrc}
@@
 
+\subsubsection{Proving that \texttt{plus} is primitive recursive}
+
+The lemma \texttt{plusIsPR} is already proven in \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}. We present in 
+\href{../theories/html/hydras.Ackermann.PrimRecExamples.html}{Ackermann.PrimRecExamples}
+a commented version of this proof,
+
+First, we look for lemmas which may help to prove that a given function obtained with the recursor \texttt{nat\_rec} is primitive recursive.
+
+\begin{Coqsrc}
+Search (is_PR 2 (fun _ _ => nat_rec _ _ _ _)).
+\end{Coqsrc}
+
+\begin{Coqanswer}
+ind1ParamIsPR:
+  forall f : nat -> nat -> nat -> nat,
+  isPR 3 f ->
+  forall g : nat -> nat,
+  isPR 1 g ->
+  isPR 2
+    (fun a b : nat =>
+     nat_rec (fun _ : nat => nat)
+                 (g b) (fun x y : nat => f x y b) a)
+\end{Coqanswer}
+
+We prove that the library function \texttt{plus} is extensionally equal to a function defined with
+\texttt{nat\_rec}.
+
+\begin{Coqsrc}
+Definition plus_alt x y  :=
+              nat_rec  (fun n : nat => nat)
+                       y
+                       (fun z t =>  S t)
+                       x.
+
+Lemma plus_alt_ok:
+  extEqual 2 plus_alt plus.
+Proof.
+  intro x; induction x; cbn; auto.
+  intros y; cbn; now rewrite <- (IHx y).
+Qed.
+\end{Coqsrc}
+
+A last lemma before the proof:
+
+\begin{Coqsrc}
+Lemma isPR_extEqual_trans n : forall f g, isPR n f ->
+                                    extEqual n f g ->
+                                    isPR n g.
+Proof.
+ intros f g [x Hx]; exists x.
+ apply extEqualTrans with f; auto.
+Qed.
+\end{Coqsrc}
+
+Let us start now.
+
+\begin{Coqsrc}
+Lemma plusIsPR : isPR 2 plus.
+Proof.
+  apply isPR_extEqual_trans with plus_alt.
+  - unfold plus_alt; apply ind1ParamIsPR.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+2 subgoals (ID 126)
+
+  ============================
+  isPR 3 (fun _ y _ : nat => S y)
+
+subgoal 2 (ID 127) is:
+ isPR 1 (fun b : nat => b)
+\end{Coqanswer}
+
+We alreday proved that \texttt{S} is \texttt{PR 1}, but we need to consider a function of three arguments, which ignores its first and third arguments.
+Fortunately, the library \texttt{primRec} already contains lemmas adapted to this kind of situation.
+
+\begin{Coqanswer}
+filter010IsPR :
+forall g : nat -> nat, isPR 1 g -> isPR 3 (fun _ b _ : nat => g b)
+\end{Coqanswer}
+
+Thus, our first subgoal is solved easily and the proof ends, applying alreday proven lemmas.
+
+\begin{Coqsrc}
+ - unfold plus_alt; apply ind1ParamIsPR.
+    + apply filter010IsPR, succIsPR.
+    + apply idIsPR.
+  - apply plus_alt_ok. 
+Qed.
+\end{Coqsrc}
+
 \begin{todo}
@@
 
+\begin{exercise}
+There is a lot of lemmas similar to \texttt{filter010IsPR} in the \texttt{primRec} library, useful to control the arity of functions.
+Thus, the reader may look at them, and invent simple examples of application for each one.
+\end{exercise}
+
+\begin{exercise}
+Multiplication of natural number is already proven in the \texttt{primRec} library. Write a proof of your own, then compare to the library's version.
+\end{exercise}
+
+\subsubsection{More examples}
+
+The following proof decomposes the \texttt{double} function as the composition of 
+multiplication,with the pair composed of the identity and the constant function wihich returns $2$.
+\emph{Note that the lemma \texttt{const1\_NIsPR} considers this function as an unary function (unlike \texttt{const0\_NIsPR} }.
+
+\begin{Coqsrc}
+Lemma doubleIsPR : isPR 1 double.
+Proof.
+  unfold double; apply compose1_2IsPR.
+  - apply idIsPR.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+subgoal 1 (ID 110) is:
+ isPR 1 (fun _ : nat => 2)
+subgoal 2 (ID 111) is:
+ isPR 2 Init.Nat.mul
+\end{Coqanswer}
+
+ \begin{Coqsrc}
+  - apply const1_NIsPR.
+  - apply multIsPR.
+Qed.
+\end{Coqsrc}
+
+\begin{exercise}
+Consider the following functions:
+
+\begin{Coqsrc}
+Fixpoint fac n :=
+  match n with 
+          | 0 => 1
+          | S p  => n * fac p
+  end.
+
+Fixpoint exp n p :=
+  match p with
+  | 0 => 1
+  | S m =>  exp n m * n
+  end.
+
+Fixpoint tower2 n :=
+  match n with
+  | 0 => 1
+  | S p => exp 2 (tower2 p)
+  end.
+\end{Coqsrc}
+
+Prove they are primitive recursive.
+
+\textbf{Hint:} You may have to look again at the lemmas of the library
+\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec} if you meet some difficulty.
+\end{exercise}
+
 \section{Two challenges}
 
-\begin{itemize}
-\item Write a simple and readable proof that the Fibonacci function is primitive recursive.
+The following projects may require to look at other libraries, and or prove new lemmas.
+Please send us your solutions (even if they are partial).
+
+\index{Projects}
+
+\begin{project}
+Write a simple and readable proof that the Fibonacci function is primitive recursive.
 
@@
 \end{Coqsrc}
+\end{project}
 
-\item Prove formally that the Ackermann function is \emph{not} primitive recursive.
+\begin{project}
+Prove formally that the Ackermann function is \emph{not} primitive recursive.
 
-\end{itemize}
+You may find a definition and some useful lemmas in
+ \href{../theories/ordinals/Prelude/Iterates.html}%
+{ordinals.Prelude.Iterates}.
+
+\end{project}

### 2021-01-20 doc/chapter-primrec.tex
redaction of Ackermann chapter (continued)

@@
 This chapter contains some comments on Russel's library, as well as a few extensions.
+Contributions (under the form of comments, new examples or exercises) are welcomed!.
 
@@
 
-\section{Two challenges}
+\index{Exercises}
 
-The following projects may require to look at other libraries, and or prove new lemmas.
-Please send us your solutions (even if they are partial).
+\begin{exercise}
+Write a simple and readable proof that the Fibonacci function is primitive recursive.
+
+\begin{Coqsrc}
+Fixpoint fib (n:nat) : nat :=
+  match n with
+  | 0 => 1
+  | 1 => 1
+  | S ((S p) as q) => fib q + fib p
+  end.
+\end{Coqsrc}
+
+\textbf{Hint:}  You may use as a helper the function which computes the pair 
+$(\texttt{fib}(n+1),\texttt{fib}(n))$. 
+Library \href{../theories/html/hydras.Ackermann.cPair.html}{Ackermann.cPair} contains
+the definition of the encoding of $\mathbb{N}^2$ into $\mathbb{N}$, and the proofs that 
+the associated constructor and projections are primitive recursive.
+
+\end{exercise}
 
 \index{Projects}
-
-\begin{project}
-Write a simple and readable proof that the Fibonacci function is primitive recursive.
-
-We give the tail-recursive definition from~\cite{BC04}, chapter 9. See also
-\href{https://coq-community.org/coq-art/ch9_function_specification/SRC/fib_tail.v}{this file}.
-\begin{Coqsrc}
-
-Fixpoint general_fib (a0 a1 n:nat) {struct n} : nat :=
-  match n with
-  | O => a0
-  | S p => general_fib a1 (a0 + a1) p
-  end.
-
-Definition fib_tail (n:nat) := general_fib 1 1 n.
-\end{Coqsrc}
-\end{project}
-
 \begin{project}
 Prove formally that the Ackermann function is \emph{not} primitive recursive.
-
-You may find a definition and some useful lemmas in
+You may find a definition of the Ackermann function and a few  lemmas in
  \href{../theories/ordinals/Prelude/Iterates.html}%

### 2021-01-20 doc/chapter-primrec.tex
corrections to the Ackermann doc

@@
 
+\index{Exercises}
+\begin{exercise}
+Show that the function \texttt{min: naryFunc\,2} is primitive recursive.
+\end{exercise}
+
 \index{Projects}
@@
 Prove formally that the Ackermann function is \emph{not} primitive recursive.
-You may find a definition of the Ackermann function and a few  lemmas in
+
+\begin{align}
+A\,0\,m&=m+1\\
+A\,(n+1)\,0 &= A\,n\,1\\
+A\,(n+1)\,(m+1)&= A\,n\,(A\,(n+1)\,m)
+\end{align}
+
+You may find also  definition of the Ackermann function and a few  lemmas in
  \href{../theories/ordinals/Prelude/Iterates.html}%
@@
 
+\textbf{Hint:} the reader may follow the plan suggested by Bruno Salvy (examen given at the French \'Ecole Polytechnique) (available at
+\href{http://www.enseignement.polytechnique.fr/informatique/INF412/uploads/Main/pc-primrec-sujet2014.pdf}{this site}).
+
+\begin{enumerate}
+\item Show that, for any $n$, the function $A\,n$ is primitive recursive.
+\item Show that for all $n$ and $m$, $A\,n\,m > m$ and  that the functions $A\,n$ are strictly monotonous.
+Show also that, for any $n$ and $m$, $A\,n\,m \leq A\,(n+1)\,m$.
+         Finally, show that for any $n$, $m$ and $k$, 
+$A\,k\,(A\,m\,n)\leq A\,(2+\textrm{max}(k,m))\,n$.
+\item Show that for any primitive recursive function $f$ of $k$ arguments,
+for any $x_1,\dots\,x_k$, $f(x_1,\dots,x_k)\leq A\,n\,(x_1+\dots+x_k)$.
+\item Show that the Ackermann function is not primitive recursive.
+
+\end{enumerate}
+
+\begin{remark}
+Note that the Ackermann function is a counter-example to the (false) statement:
+\begin{quote}
+  ``Let $f$ be a function of type \texttt{naryFunc\,2}. If, for any $n$, the fonction $f(n)$ is primitive recursive , then f is primitive recursive.''
+\end{quote}
+\end{remark}
 \end{project}

### 2021-01-20 doc/chapter-primrec.tex
restructuration of the Ackermann directory

@@
 \noindent
-\emph{From\href{../theories/html/hydras.Ackermann.PrimRecExamples.html}{Ackermann.PrimRecExamples}}.
+\emph{From\href{../theories/html/hydras.Ackermann.MoreAck.PrimRecExamples.html}{Ackermann.MoreAck.PrimRecExamples}}.
 \begin{Coqsrc}
@@
 we wrote alternate proofs in 
-\href{../theories/html/hydras.Ackermann.PrimRecExamples.html}%
-{Ackermann.PrimRecExamples.v}, in order to illustrate the main proof patterns.
+\href{../theories/html/hydras.Ackermann.MoreAck.PrimRecExamples.html}%
+{Ackermann.MoreAck.PrimRecExamples.v}, in order to illustrate the main proof patterns.
 
@@
 \noindent
-\emph{From \href{../theories/html/hydras.Ackermann.PrimRecExamples.html}{Ackermann.PrimRecExamples}}.
+\emph{From \href{../theories/html/hydras.Ackermann.MoreAck.PrimRecExamples.html}{Ackermann.MoreAck.PrimRecExamples}}.
 \begin{Coqsrc}
@@
 The lemma \texttt{plusIsPR} is already proven in \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}. We present in 
-\href{../theories/html/hydras.Ackermann.PrimRecExamples.html}{Ackermann.PrimRecExamples}
+\href{../theories/html/hydras.Ackermann.MoreAck.PrimRecExamples.html}{Ackermann.MoreAck.PrimRecExamples}
 a commented version of this proof,
@@
 \begin{todo}
-Comment more examples from   \href{../theories/html/hydras.Ackermann.PrimRecExamples.html}{Ackermann.PrimRecExamples}.
+Comment more examples from   \href{../theories/html/hydras.Ackermann.MoreAck.PrimRecExamples.html}{Ackermann.MoreAck.PrimRecExamples}.
 \end{todo}

### 2021-01-21 doc/chapter-primrec.tex
add an exercises/ directory

@@
 \begin{quote}
-  ``Let $f$ be a function of type \texttt{naryFunc\,2}. If, for any $n$, the fonction $f(n)$ is primitive recursive , then f is primitive recursive.''
+{\color{red}
+  ``Let $f$ be a function of type \texttt{naryFunc\,2}. If, for any $n$, the fonction $f(n)$ is primitive recursive, then f is primitive recursive.''}
 \end{quote}

### 2021-01-22 doc/part-hydras.tex
comments and examples about the Ackermann part (to continue)
add an exercise

@@
 
-\index{Coq!Techniques!Mutually inductive types}
+\index{Coq!Mutually inductive types}
 \index{Coq!Commands!Scheme}
@@
 % \label{warning:coercions}
-% \index{Coq!Techniques!Coercions} 
+% \index{Coq!Coercions} 
 % \begin{remark}
@@
 \begin{exercise}
-Prove, that, in any ordinal notation system, that every ordinal has at most one predecessor, and at most one successor. 
+Prove, that, in any ordinal notation system, every ordinal has at most one predecessor, and at most one successor.
+
+\emph{You may start this exercise with the file
+\url{../exercises/ordinals/predSuccUnicity.v}.}
+
 \end{exercise}
@@
 \begin{exercise}
-Prove, that, in any ordinal notation system, that if $\beta$ is a successor of $\alpha$,
+Prove, that, in any ordinal notation system, if $\beta$ is a successor of $\alpha$,
 then for any $\gamma$, $\gamma<\beta$ implies 
@@
 
-\index{Coq!Techniques!Unicity of equality proofs}
+\index{Coq!Unicity of equality proofs}
 \label{sect:eq-proof-unicity}
@@
 
-% \index{Coq!Techniques!Coercions}
+% \index{Coq!Coercions}
 % \index{Functions!Coercions@Coercions (from nat to ordinal types)}
@@
 
-\index{Coq!Techniques!Unicity of equality proofs}
+\index{Coq!Unicity of equality proofs}
 
@@
 which are not in Cantor normal form.
+
+\emph{You may start this exercise with the file
+\url{../exercises/ordinals/T1_ltNotWf.v}.}
 \end{exercise}
@@
 \label{sec:orgheadline77}
-\index{Coq!Techniques!Well-founded induction}
+\index{Coq!Well-founded induction}
 
@@
 to \texttt{nat} is available on \href{../theories/html/hydras.Prelude.Fuel.html}{Prelude.Fuel}).
-\index{Coq!Techniques!Giving fuel to a long computation}
+\index{Coq!Giving fuel to a long computation}
 
@@
 \label{lemma:L-2_6-1}
-\index{Coq!Techniques!Sigma types}
+\index{Coq!Sigma types}
 
@@
 
-\index{Coq!Techniques!Transfinite induction}
+\index{Maths!Transfinite induction}
 

### 2021-01-24 doc/part-hydras.tex
add an exercise
add partial solutions to exercises

@@
 $\gamma\leq\alpha$.
+
+\emph{You may start this exercise with the file
+\url{../exercises/ordinals/lt_succ_le.v}.}
 \end{exercise}
@@
  \begin{exercise}
-Prove in \coq{} that for any ordinal $\alpha<\epsilon_0$, $\alpha$ is a limit if 
-and only if for all $\beta<\alpha$, the interval $[\beta,\alpha)$ is infinite.
+Prove in \coq{} that for any ordinal $0< \alpha<\epsilon_0$, $\alpha$ is a limit if 
+and only if for all $\beta<\alpha$, the interval $[\beta,\alpha)$ is
+infinite.
+
+\emph{You may start this exercise with the file
+\url{../exercises/ordinals/Limit_Infinity.v}.}
  \end{exercise}

### 2021-01-25 doc/part-hydras.tex
add an exercise

@@
 
+%% To simplify !
 \index{Exercises}
 \begin{exercise}
-Let $\alpha$ be an ordinal. We say that $\alpha$ is \emph{infinite} iff the segment $[0,\alpha)$ is an infinite set. 
-\begin{enumerate}
-\item Adapt this definition to the type \texttt{E0}.
-\item Prove that \emph{being infinite} is decidable on type \texttt{E0}.
-\item Prove that $\alpha$ is infinite if and only if for any finite ordinal $i$,
+Prove that for any ordinal $\alpha:\texttt{E0}$, 
+$\omega\leq \alpha$ if and only if, for any natural number $i$,
 $i+\alpha=\alpha$.
-\end{enumerate}
+
+\emph{You may start this exercise with the file
+\url{../exercises/ordinals/ge_omega_iff.v}.}
 \end{exercise}

### 2021-01-27 doc/chapter-powers.tex
remove an exercise

@@
 
-If we define $\varphi(a,b)  \left(
+If we pose $\varphi(a,b) =\left(
   \begin{array}{cc}
@@
 
-Since \texttt{computation} is an inductive family of types, 
-it could be possible to define a composition operator that 
-works like list appending (replacing the \texttt{Return y} of the first
-computation with the second computation).
-\emph{This approach is left as an exercise.}  The solution we present is based on functional programming and the concept of continuation.
-
-\index{Exercises}
-\begin{exercise}
-Develop the approach suggested in the previous paragraph.  
-\end{exercise}
+  The solution we present is based on functional programming and the concept of continuation.
 

### 2021-01-28 doc/chapter-primrec.tex
add an exercise
add a new exercise
additions to an exercise on Ackermann function

@@
 \begin{Coqsrc}
-Fixpoint fac n :=
+Fixpoint fact n :=
   match n with 
           | 0 => 1
-          | S p  => n * fac p
+          | S p  => n * fact p
   end.
@@
 \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec} if you meet some difficulty.
+
+\emph{You may start this exercise with the file
+\url{../exercises/primrec/MorePRExamples.v}.}
 \end{exercise}
@@
 
+\emph{You may start this exercise with
+\url{../exercises/primrec/FibonacciPR.v}.}
+
 \end{exercise}
@@
 \begin{exercise}
-Show that the function \texttt{min: naryFunc\,2} is primitive recursive.
+Show that the function \texttt{min: naryFunc\,2} is primitive
+recursive.
+
+\emph{You may start this exercise with
+\url{../exercises/primrec/MinPR.v}.}
+
 \end{exercise}
@@
 \item Show that, for any $n$, the function $A\,n$ is primitive recursive.
+\emph{\textbf{Done.}}
+
 \item Show that for all $n$ and $m$, $A\,n\,m > m$ and  that the functions $A\,n$ are strictly monotonous.
-Show also that, for any $n$ and $m$, $A\,n\,m \leq A\,(n+1)\,m$.
+\emph{\textbf{To do.}}
+
+Show also that, for any $n$ and $m$, $A\,n\,m \leq A\,(n+1)\,m$. \emph{\textbf{To do.}}
          Finally, show that for any $n$, $m$ and $k$, 
 $A\,k\,(A\,m\,n)\leq A\,(2+\textrm{max}(k,m))\,n$.
+\emph{\textbf{To do.}}
 \item Show that for any primitive recursive function $f$ of $k$ arguments,
-for any $x_1,\dots\,x_k$, $f(x_1,\dots,x_k)\leq A\,n\,(x_1+\dots+x_k)$.
-\item Show that the Ackermann function is not primitive recursive.
+for any $x_1,\dots\,x_k$, $f(x_1,\dots,x_k)\leq A\,n\,(x_1+\dots+x_k)$. \emph{\textbf{To do.}}
+\item Show that the Ackermann function is not primitive recursive. \emph{\textbf{To do.}}
 
 \end{enumerate}
+
+\emph{You may start this exercise with
+\url{../exercises/primrec/AckermannNotPR.v}.}
 
@@
 \end{remark}
+
 \end{project}

### 2021-01-29 doc/chapter-primrec.tex
new lemmas on Ack

@@
 You may find also  definition of the Ackermann function and a few  lemmas in
- \href{../theories/ordinals/Prelude/Iterates.html}%
-{ordinals.Prelude.Iterates}.
+ \href{../theories/ordinals/Prelude/Ack.html}%
+{ordinals.Prelude.Ack}.
 
@@
 \href{http://www.enseignement.polytechnique.fr/informatique/INF412/uploads/Main/pc-primrec-sujet2014.pdf}{this site}).
+
+The lemmas which are marked as \textbf{Done} are proved
+in the library \texttt{ordinals.Prelude.Ack}.
 
@@
 \item Show that for all $n$ and $m$, $A\,n\,m > m$ and  that the functions $A\,n$ are strictly monotonous.
-\emph{\textbf{To do.}}
 
-Show also that, for any $n$ and $m$, $A\,n\,m \leq A\,(n+1)\,m$. \emph{\textbf{To do.}}
+Show also that, for any $n$ and $m$, $A\,n\,m \leq A\,(n+1)\,m$. 
          Finally, show that for any $n$, $m$ and $k$, 
 $A\,k\,(A\,m\,n)\leq A\,(2+\textrm{max}(k,m))\,n$.
-\emph{\textbf{To do.}}
+\emph{\textbf{Done.}}
 \item Show that for any primitive recursive function $f$ of $k$ arguments,
-for any $x_1,\dots\,x_k$, $f(x_1,\dots,x_k)\leq A\,n\,(x_1+\dots+x_k)$. \emph{\textbf{To do.}}
-\item Show that the Ackermann function is not primitive recursive. \emph{\textbf{To do.}}
+for any $x_1,\dots\,x_k$, $f(x_1,\dots,x_k)\leq A\,n\,(x_1+\dots+x_k)$. \emph{\textbf{Still to do.}}
+\item Show that the Ackermann function is not primitive recursive. \emph{\textbf{Still to do.}}
 
@@
 
-\emph{You may start this exercise with
+\emph{You may start this exercise with the file
 \url{../exercises/primrec/AckermannNotPR.v}.}

### 2021-02-07 doc/chapter-primrec.tex
start redaction doc on Ackermann function

--- 2021-01-28 doc/chapter-primrec.tex
@@
 
+\label{def:Primrec}
 \vspace{4pt}
@@
 
-\index{Projects}
-\begin{project}
-Prove formally that the Ackermann function is \emph{not} primitive recursive.
+\section{Proving that a given function is \emph{not} primitive recursive}
 
-\begin{align}
-A\,0\,m&=m+1\\
-A\,(n+1)\,0 &= A\,n\,1\\
-A\,(n+1)\,(m+1)&= A\,n\,(A\,(n+1)\,m)
-\end{align}
+The best known of total function which is not primitive recursive is the Ackermann function. We show how to adapt the classic proof (see for instance~\cite{planetmath}) to the constraints of \gallina.
 
-You may find also  definition of the Ackermann function and a few  lemmas in
- \href{../theories/ordinals/Prelude/Iterates.html}%
-{ordinals.Prelude.Iterates}.
+\subsection{The Ackermann function}
 
-\textbf{Hint:} the reader may follow the plan suggested by Bruno Salvy (examen given at the French \'Ecole Polytechnique) (available at
-\href{http://www.enseignement.polytechnique.fr/informatique/INF412/uploads/Main/pc-primrec-sujet2014.pdf}{this site}).
+\vspace{4pt}
+\noindent
+\emph{From \href{../theories/html/hydras.Ackermann.MoreAck.Ack.html}{Ackermann.MoreAck.Ack}}.
+
+\begin{Coqsrc}
+Fixpoint Ack (m:nat) : nat -> nat :=
+  match m with
+  | 0 => S
+  |   S n => fun k =>  iterate (Ack n) (S k) 1
+  end.
+\end{Coqsrc}
+
+This function is also known under the form of three equations, which we derive immediately from our definition.
+
+\begin{Coqsrc}
+Lemma Ack_0 : Ack 0 = S.
+Proof refl_equal.
+
+Lemma Ack_S_0 m : Ack (S m) 0 = Ack m 1. 
+Proof.  now cbn. Qed.
+
+Lemma Ack_S_S : forall m p,
+    Ack (S m) (S p) = Ack m (Ack (S m) p).
+Proof.  now cbn. Qed.
+\end{Coqsrc}
+
+The order of growth of the Ackermann function w.r.t. its first argument is illustrated by the following equalities.
+
+\begin{Coqsrc}
+Lemma Ack_1_n n : Ack 1 n = S (S n).
+
+Lemma Ack_2_n n: Ack 2  n = 2 * n + 3.
+
+Lemma Ack_3_n n: Ack 3 n = exp2 (S (S (S n))) - 3.
+
+Lemma Ack_4_n n: Ack 4 n = hyper_exp2 (S (S (S n))) - 3.
+\end{Coqsrc}
+
+Please note that for any given $n$, the unary function
+(\texttt{Ack\,$n$} is primitive recursive.
+
+\emph{From \href{../theories/html/hydras.Ackermann.MoreAck.AckNotPR.html}{Ackermann.MoreAck.AckNotPR}}.
+\begin{Coqsrc}
+Theorem Ackn_IsPR (n: nat) : isPR 1 (Ack n).
+\end{Coqsrc}
+
+Ìn order to prove that \texttt{Ack} (considered as a function of two arguments) is not primitive recursive, the usual method consists in two steps:
 
 \begin{enumerate}
-\item Show that, for any $n$, the function $A\,n$ is primitive recursive.
-\emph{\textbf{Done.}}
-
-\item Show that for all $n$ and $m$, $A\,n\,m > m$ and  that the functions $A\,n$ are strictly monotonous.
-\emph{\textbf{To do.}}
-
-Show also that, for any $n$ and $m$, $A\,n\,m \leq A\,(n+1)\,m$. \emph{\textbf{To do.}}
-         Finally, show that for any $n$, $m$ and $k$, 
-$A\,k\,(A\,m\,n)\leq A\,(2+\textrm{max}(k,m))\,n$.
-\emph{\textbf{To do.}}
-\item Show that for any primitive recursive function $f$ of $k$ arguments,
-for any $x_1,\dots\,x_k$, $f(x_1,\dots,x_k)\leq A\,n\,(x_1+\dots+x_k)$. \emph{\textbf{To do.}}
-\item Show that the Ackermann function is not primitive recursive. \emph{\textbf{To do.}}
-
+\item Prove that for any primitive recursive function $f:\mathbb{N}\rightarrow\mathbb{N}\rightarrow\mathbb{N}$, there exists some natural number $n$ depending on $f$, such that, for any $x$ and $y$, 
+$f\,x\,y \leq \texttt{Ack}\,n\,(\textrm{max}\,x\,y)$. Let us say that $f$ is \emph{majorized} [by some \texttt{Ack\,$n$}].
+\item Show that \texttt{Ack} fails to satisfy this property.
 \end{enumerate}
 
-\emph{You may start this exercise with
-\url{../exercises/primrec/AckermannNotPR.v}.}
+The second part is the easiest one, let us look at it first, and admit the following lemma.
+
+\begin{Coqsrc}
+Lemma PR_2_majorized (f: naryFunc 2)(Hf : isPR 2 f)
+  : exists (n:nat), forall x y,  f x y <= Ack n (x + y).
+\end{Coqsrc}
+
+Then, assume texttt{Ack} is majorized, and find a contradiction.
+
+\begin{Coqsrc}
+Section Impossibility_Proof.
+
+  Hypothesis HAck : isPR 2 Ack.
+
+  Fact Ack_majorized : 
+      exists (n:nat), forall x y,  Ack x y <= Ack n (x + y).
+
+  Section Contrad.
+    Variable n: nat.
+    Hypothesis  Hn :  forall x y,  Ack x y <= Ack n (x + y) .
+
+    Remark R01 : n <> 0.
+
+    Remark R02: n <> 1.
+
+    Remark R03 : max 2 n = n.
+
+    Remark R04 : Ack n (2 * n + 3) = Ack n (Ack 2 n).
+
+    Remark R05 : Ack n (2 * n + 3) <= Ack (n + 2) n.
+
+    Remark R06 : Ack (n + 2) n <= Ack n (2 * n + 2).
+
+    Remark R07 : Ack n (2 * n + 3) <= Ack n (2 * n + 2).
+
+    Remark R08 :  Ack n (2 * n + 2) < Ack n (2 * n + 3).
+
+    Lemma contrad : False.
+    Proof.
+      generalize R07, R08; intros; lia.
+    Qed.
+
+  End Contrad.
+
+  Lemma Ack_not_PR : False.
+
+End Impossibility_Proof.
+\end{Coqsrc}
+
+Now it remains to prove Lemma \texttt{PR\_2\_majorized}.
+If we look at the inductive definition of primitive recursive functions,\pageref{def:Primrec}, it is obvious that a proof by induction on the construction of primitive recursive functions must consider functions of any arity.
+
+\vspace{4pt}
+\noindent
+\emph{From \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.}
+
+\begin{Coqsrc}
+Scheme PrimRec_PrimRecs_ind := Induction for PrimRec
+  Sort Prop
+  with PrimRecs_PrimRec_ind := Induction for PrimRecs 
+  Sort Prop.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+PrimRec_PrimRecs_ind :
+forall (P : forall n : nat, PrimRec n -> Prop)
+  (P0 : forall n n0 : nat, PrimRecs n n0 -> Prop),
+(* successor *)
+P 1 succFunc ->
+
+(* zero *)
+P 0 zeroFunc ->
+
+* projections *)
+(forall (n m : nat) (l : m < n), P n (projFunc n m l)) ->
+
+(* composition *) 
+(forall (n m : nat) (g : PrimRecs n m),
+      P0 n m g -> forall h : PrimRec m, P m h -> 
+      P n (composeFunc n m g h)) ->
+
+(* primitive recursion *)
+(forall (n : nat) (g : PrimRec n),
+ P n g ->
+    forall h : PrimRec (S (S n)), P (S (S n)) h -> 
+     P (S n) (primRecFunc n g h)) ->
+
+(* empty list of functions *)
+(forall n : nat, P0 n 0 (PRnil n)) ->
+
+(* add a function to a list *)
+(forall (n m : nat) (p : PrimRec n),
+   P n p -> 
+   forall p0 : PrimRecs n m, P0 n m p0 -> 
+   P0 n (S m) (PRcons n m p p0)) ->
+
+(* conclusion ! *)
+forall (n : nat) (p : PrimRec n), P n p
+\end{Coqanswer}
+
+Thus the lemma we will have to prove is, for any primitive recursive function of abritrary arity $n$ is:
+
+\[
+   \exists q\in\mathbb{N},\, \forall x_1,\dots,x_n, 
+      f(x_1,\dots,\,x_n)\leq\textrm{Ack}(q,\textrm{max}(x_1,\dots,x_n))
+\]
+
+But dots don't belong to \gallina's syntax! So, we may use \coq's vectors for denoting arbitrary tuples.
+
+%%% ICI
+
+% \item Show that for all $n$ and $m$, $A\,n\,m > m$ and  that the functions $A\,n$ are strictly monotonous.
+
+% Show also that, for any $n$ and $m$, $A\,n\,m \leq A\,(n+1)\,m$. 
+%          Finally, show that for any $n$, $m$ and $k$, 
+% $A\,k\,(A\,m\,n)\leq A\,(2+\textrm{max}(k,m))\,n$.
+% \emph{\textbf{Done.}}
+% \item Show that for any primitive recursive function $f$ of $k$ arguments,
+% for any $x_1,\dots\,x_k$, $f(x_1,\dots,x_k)\leq A\,n\,(x_1+\dots+x_k)$. \emph{\textbf{Still to do.}}
+% \item Show that the Ackermann function is not primitive recursive. \emph{\textbf{Still to do.}}
+
+% \end{enumerate}
 
@@
 \end{remark}
-
-\end{project}

### 2021-02-08 doc/chapter-primrec.tex
redaction on PrimRec and Ackermann function

@@
 \section{Introduction}
-The definition of primitive recursive functions we use come
+The definition of primitive recursive functions we use comes
 from Russel O'Connors formalization in \coq{} of
@@
 than 45K lines of scripts. The part dedicated to primitive recursive functions and Peano arithmetics is 32K lines long and is originally structured in 38 modules.
-
-Thus, we propose a slow exploration of this library, through examples and exercises.
+Thus, we propose a slow exploration of this library, through examples and exercises. Our additions to the original library 
+are stored in a sub-directory \texttt{Ackermann/MoreAck}.
 
@@
 \begin{Coqsrc}
-Example extEqual_ex1 : extEqual 2 mult (fun x y =>  y * x + x -x) .
+Example extEqual_ex1 : extEqual 2 mult (fun x y =>  y * x + x - x) .
 Proof.
@@
 We presume that this convention makes it easier to define the evaluation function \texttt{evalProjFunc $n$} (see next sub-section) .
-Trying the other conention is left as an exercise.
+Trying the other convention is left as an exercise.
 \end{remark}
@@
 
+\subsection{A proof by induction on all primitive recursive functions}
+
 Ìn order to prove that \texttt{Ack} (considered as a function of two arguments) is not primitive recursive, the usual method consists in two steps:
@@
 \item Prove that for any primitive recursive function $f:\mathbb{N}\rightarrow\mathbb{N}\rightarrow\mathbb{N}$, there exists some natural number $n$ depending on $f$, such that, for any $x$ and $y$, 
-$f\,x\,y \leq \texttt{Ack}\,n\,(\textrm{max}\,x\,y)$. Let us say that $f$ is \emph{majorized} [by some \texttt{Ack\,$n$}].
+$f\,x\,y \leq \texttt{Ack}\,n\,(\textrm{max}\,x\,y)$ (we say that $f$ is \emph{majorized}  by \texttt{Ack}).
 \item Show that \texttt{Ack} fails to satisfy this property.
@@
 
-The second part is the easiest one, let us look at it first, and admit the following lemma.
-
-\begin{Coqsrc}
-Lemma PR_2_majorized (f: naryFunc 2)(Hf : isPR 2 f)
-  : exists (n:nat), forall x y,  f x y <= Ack n (x + y).
-\end{Coqsrc}
-
-Then, assume texttt{Ack} is majorized, and find a contradiction.
-
-\begin{Coqsrc}
-Section Impossibility_Proof.
-
-  Hypothesis HAck : isPR 2 Ack.
-
-  Fact Ack_majorized : 
-      exists (n:nat), forall x y,  Ack x y <= Ack n (x + y).
-
-  Section Contrad.
-    Variable n: nat.
-    Hypothesis  Hn :  forall x y,  Ack x y <= Ack n (x + y) .
-
-    Remark R01 : n <> 0.
-
-    Remark R02: n <> 1.
-
-    Remark R03 : max 2 n = n.
-
-    Remark R04 : Ack n (2 * n + 3) = Ack n (Ack 2 n).
-
-    Remark R05 : Ack n (2 * n + 3) <= Ack (n + 2) n.
-
-    Remark R06 : Ack (n + 2) n <= Ack n (2 * n + 2).
-
-    Remark R07 : Ack n (2 * n + 3) <= Ack n (2 * n + 2).
-
-    Remark R08 :  Ack n (2 * n + 2) < Ack n (2 * n + 3).
-
-    Lemma contrad : False.
-    Proof.
-      generalize R07, R08; intros; lia.
-    Qed.
-
-  End Contrad.
-
-  Lemma Ack_not_PR : False.
-
-End Impossibility_Proof.
-\end{Coqsrc}
-
-Now it remains to prove Lemma \texttt{PR\_2\_majorized}.
+First, we prove that any primitive function of two arguments is majorized.
 If we look at the inductive definition of primitive recursive functions,\pageref{def:Primrec}, it is obvious that a proof by induction on the construction of primitive recursive functions must consider functions of any arity.
@@
 
-Thus the lemma we will have to prove is, for any primitive recursive function of abritrary arity $n$ is:
+For instance, proving a property shared by any primitive recursive functions of arity 2 leads to consider the case where that function is obtained by composition with a function of any 
+arity $m$. The same problem happens with primitive 
+recursion, where a function of arity $n$ is built out of a function of arity $n+1$ and a function of arity $n-1$.
 
+Thus the lemma we will have to prove is the following one:
+
+  \begin{quote}
+    For any $n$, and any primitive recursive function $f$ of  arity $n$, there exists some natural number $q$ such that the following inequality holds:
  \[
-   \exists q\in\mathbb{N},\, \forall x_1,\dots,x_n, 
+  \forall x_1,\dots,x_n, 
       f(x_1,\dots,\,x_n)\leq\textrm{Ack}(q,\textrm{max}(x_1,\dots,x_n))
 \]
+ \end{quote}
 
 But dots don't belong to \gallina's syntax! So, we may use \coq's vectors for denoting arbitrary tuples.
+
+First, we extend \texttt{max} to vectors of natural numbers (using the notations of module \texttt{VectorNotations} and some more definitions from 
+\href{../theories/html/hydras.Prelude.MoreVectors.html}{Prelude.MoreVectors}).
+So, \texttt{t\,$A$\,$n$} is the type of vectors of $n$ elements of type $A$, and the constants \texttt{cons}, \texttt{nil}, \texttt{map}, etc., refer to vectors and not to lists.
+Likewise, the notation \texttt{x::v} is an abbreviation for
+\texttt{VectorDef.cons x ?n v}.
+
+\begin{Coqsrc}
+Fixpoint max_v {n:nat} : forall (v: Vector.t nat n) , nat :=
+  match n as n0 return (Vector.t nat n0 -> nat)
+  with
+    0 => fun v => 0
+  | S p => fun (v : Vector.t nat (S p)) =>
+             (max (Vector.hd v) (max_v  (Vector.tl v)))
+  end.
+
+Lemma max_v_2 : forall x y,  max_v (x::y::nil) = max x y.
+
+Lemma max_v_lub : forall n (v: t nat n) y,
+    (Forall (fun x =>  x <= y) v) -> max_v v <= y.
+
+Lemma max_v_ge : forall n (v: t nat n) y,  In  y  v -> y <= max_v v.
+\end{Coqsrc}
+
+We have also to convert any application
+$(f\,x_1\,x_2\,\dots\,x_n)$ into an application of a function 
+to a single argument: the vector of all the $x_i$s.
+This is already done in 
+Library~\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.
+
+\begin{Coqsrc}
+Fixpoint evalList (m : nat) (l : Vector.t nat m) {struct l} :
+ naryFunc m -> nat :=
+  match l in (Vector.t _ m) return (naryFunc m -> nat) with
+  | Vector.nil => fun x : naryFunc 0 => x
+  | Vector.cons a n l' => fun x : naryFunc (S n) => evalList n l' (x a)
+  end.
+\end{Coqsrc}
+
+Indeed, (\texttt{evalList $m$ $v$ $f$}) is the application to the vector $v$ of
+an uncurried version of $f$.
+
+In Library\href{../theories/html/hydras.Ackermann.MoreAck.AckNotPR.html}{Ackermann.MoreAck.AckNotPR}, we introduce a lighter notation.
+
+\begin{Coqsrc}
+(**  uncurried apply:
+
+[v_apply f (x1::x2:: ... ::xn::nil)]  is [f x1 x2 ... xn] 
+ *)
+
+Notation "'v_apply' f v" := (evalList _ v f) ( at level 10, f at level 9).
+
+Example Ex2 : forall (f: naryFunc 2) x y,
+    v_apply f (x::y::nil) = f x y.
+Proof.   intros; now cbn. Qed.
+
+Example Ex4 : forall (f: naryFunc 4) x y z t,
+    v_apply f (x::y::z::t::nil) = f x y z t.
+Proof.  intros; now cbn. Qed.
+\end{Coqsrc}
+
+We are now able to translate in \gallina{} the notion of majoration:
+
+\begin{Coqsrc}
+Definition majorized {n} (f: naryFunc n) (A: naryFunc 2) : Prop :=
+  exists (q:nat), forall (v: t nat n),
+      v_apply f v <= A q  (max_v v).
+
+Definition majorizedPR {n} (x: PrimRec n) A := 
+           majorized (evalPrimRec n x) A.
+
+(** For vectors of functions *)
+
+Definition majorizedS {n m} (fs : Vector.t (naryFunc n) m)
+           (A : naryFunc 2):=
+  exists N, forall (v: t nat n),
+      max_v (map (fun f => v_apply f v) fs) <= A N (max_v v).
+
+Definition majorizedSPR {n m} (x : PrimRecs n m) :=
+  majorizedS (evalPrimRecs _ _ x).
+\end{Coqsrc}
+
+Now, it remains to prove that any primitive function is majorized by \texttt{Ack}.
+The three base cases of construction of primitive recursive function are as follows:
+
+\begin{Coqsrc}
+Lemma majorSucc : majorizedPR  succFunc Ack.
+
+Lemma majorZero : majorizedPR  zeroFunc Ack.
+
+Lemma majorProjection (n m:nat)(H: m < n):
+  majorizedPR (projFunc n m H) Ack.
+\end{Coqsrc}
+
+The rest of the cases are proved within a mutual  induction.
+
+\begin{Coqsrc}
+Lemma majorAnyPR:  forall n (x: PrimRec n),  majorizedPR  x Ack.
+Proof.
+  intros n x; induction x using PrimRec_PrimRecs_ind with
+                  (P0 := fun n m y => majorizedSPR  y Ack).
+  - apply majorSucc.
+  - apply majorZero.
+  - apply majorProjection. 
+  \end{Coqsrc}
+
+  \begin{Coqsrc}
+  - (** function composition *)
+\end{Coqsrc}
+\begin{Coqanswer}
+1 subgoal (ID 265)
+
+  n, m : nat
+  g : PrimRecs n m
+  x : PrimRec m
+  IHx : majorizedSPR g Ack
+  IHx0 : majorizedPR x Ack
+  ============================
+  majorizedPR (composeFunc n m g x) Ack
+\end{Coqanswer}
+
+\begin{Coqsrc}
+  ...
+ - (** primitive recursion *)
+\end{Coqsrc}
+
+\begin{Coqanswer}
+1 subgoal (ID 265)
+
+  n : nat
+  x1 : PrimRec n
+  x2 : PrimRec (S (S n))
+  IHx1 : majorizedPR x1 Ack
+  IHx2 : majorizedPR x2 Ack
+  ============================
+  majorizedPR (primRecFunc n x1 x2) Ack
+\end{Coqanswer}
+
+\begin{Coqsrc}
+ ...
+-  (** empty list of functions *)
+\end{Coqsrc}
+
+\begin{Coqanswer}
+1 subgoal (ID 266)
+
+  n : nat
+  ============================
+  majorizedSPR (PRnil n) Ack
+\end{Coqanswer}
+
+\begin{Coqsrc}
+  ...
+- (** non-empty list of functions *)
+\end{Coqsrc}
+
+\begin{Coqanswer}
+1 subgoal (ID 273)
+
+  n, m : nat
+  x : PrimRec n
+  p : PrimRecs n m
+  IHx : majorizedPR x Ack
+  IHx0 : majorizedSPR p Ack
+  ============================
+  majorizedSPR (PRcons n m x p) Ack
+\end{Coqanswer}
+
+\begin{Coqsrc}
+ ...
+Qed.
+\end{Coqsrc}
+
+\subsection{Looking for a contradiction}
+
+The following lemma is just a specialization of \texttt{majorAnyPR} to
+binary functions (forgetting vectors, coming back to usual notations).
+
+\begin{Coqsrc}
+Lemma majorPR2 (f: naryFunc 2)(Hf : isPR 2 f)
+  : exists (n:nat), forall x y,  f x y <= Ack n (max x  y).
+\end{Coqsrc}
+
+Then, our impossibility proof is just a sequence of easy small steps.
+
+\begin{Coqsrc}
+Section Impossibility_Proof.
+
+  Hypothesis HAck : isPR 2 Ack.
+
+  Fact Ack_majorized : majorized (Ack: naryFunc 2)  Ack.
+
+  Section Contrad.
+    Variable n: nat.
+    Hypothesis  Hn :  forall x y,  Ack x y <= Ack n (max x  y) .
+
+    Remark R01 : n <> 0.
+
+    Remark R02: n <> 1.
+
+    Remark R03 : max 2 n = n.
+
+    Remark R04 : Ack n (2 * n + 3) = Ack n (Ack 2 n).
+
+    Remark R05 : Ack n (2 * n + 3) <= Ack (n + 2) n.
+
+    Remark R06 : Ack (n + 2) n <= Ack n (2 * n + 2).
+
+    Remark R07 : Ack n (2 * n + 3) <= Ack n (2 * n + 2).
+
+    Remark R08 :  Ack n (2 * n + 2) < Ack n (2 * n + 3).
+
+    Lemma contrad : False.
+
+  End Contrad.
+
+  Lemma Ack_not_PR : False.
+
+End Impossibility_Proof.
+\end{Coqsrc}
 

### 2021-02-09 doc/chapter-powers.tex
proof simplification, typos in doc

@@
 \label{op-classes}
-\index{Coq!Type Classes!Operational Type Classes}
+\index{Coq!Type classes!Operational type classes}
 \emph{Operational type classes}~\cite{BS2011}
@@
 \subsection{The Monoid Type Class}
-\index{Coq!Type Classes}
+\index{Coq!Type classes}
 We are now ready to  give a definition of the \texttt{Monoid} class, using
@@
 
-\index{Coq!Type Classes}
+\index{Coq!Type classes}
 \index{Coq!Proper}
@@
 \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Classes.Morphisms.html}{Coq.Classes.Morphisms}
-\index{Coq!Type Classes}
+\index{Coq!Type classes}
 \index{Coq!Proper}
@@
 \begin{itemize}
-\item Continuation Passing Style (CPS) \index{Programming Styles!Continuation Passing Style (CPS)} \cite{reynolds93}
+\item Continuation Passing Style (CPS) \index{Programming styles!Continuation Passing Style (CPS)} \cite{reynolds93}
 is a way to make explicit the 
@@
 
+\index{Coq!Plug-ins!paramcoq}
+
 Marc Lasson's \textbf{paramcoq} plug-in~(available as  \texttt{opam} package 
@@
 \subsubsection{Definition:} 
-\index{Coq!Type Classes}
+\index{Coq!Type classes}
 \index{Coq!Proper}
@@
 
-\index{Programming Styles!Continuation Passing Style (CPS)}
+\index{Programming styles!Continuation Passing Style (CPS)}
 
@@
 any correct chain generator.
-\index{Coq!Dependently Typed Functions}
+\index{Coq!Dependently typed functions}
 

### 2021-02-09 doc/chapter-primrec.tex
proof simplification, typos in doc

@@
 This chapter contains some comments on Russel's library, as well as a few extensions.
-Contributions (under the form of comments, new examples or exercises) are welcomed!.
-
-\section{First look at the Ackermann library}
+Contributions (under the form of comments, new examples or exercises) are welcome!.
+
+\section{First look at the Ackermann Library}
 
@@
 
-\section{Basic Libraries}
+\section{Basic Definitions}
+\index{Maths!Primitive recursive functions}
 
@@
 
-The library allows us to consider primitive functions on \texttt{nat}, with any number of arguments, in 
+Library \texttt{primRec} allows us to consider primitive functions on \texttt{nat}, with any number of arguments, in 
 curried form. This is made possible in 
@@
 is an interpretation function (\texttt{evalprimRec $n$})  of type
-$\texttt{PrimRec}\,n \rightarrow  \texttt{naryFunc}\,n$ (and the helper function 
+$\texttt{PrimRec}\,n \rightarrow  \texttt{naryFunc}\,n$ , together with the helper function 
 (\texttt{evalprimRecS $n$ $m$}) of type 
@@
 
+\index{Coq!Dependent pattern matching}
 Both functions are mutually defined through dependent pattern matching. We advise the readers who 
@@
 \end{Coqsrc}
-Let us take as example the following term\footnote{Of course, we never typed this term \emph{verbatim}; we obtained it as the result of some computation we leave it the reader to guess and reproduce.}:
-
-\label{sect:weirdfac}
-\begin{Coqsrc}
-Example weirdPR : PrimRec 1 :=
+
+Looks complicated? The following examples show that, when
+the arity is fixed, these definitions behave well w.r.t. 
+\coq's reduction rules. Moreover, they make the interpretation functions more ``concrete''.
+
+\begin{Coqsrc}
+Example Ex1 : evalPrimRec 0 zeroFunc = 0.
+Proof. reflexivity. Qed.
+
+Example Ex2 a : evalPrimRec 1 succFunc a = S a.
+Proof. reflexivity. Qed.
+
+Example Ex3 a b c d e f: forall (H: 2 < 6),
+    evalPrimRec 6
+                (projFunc 6 2 H) a b c d e f = d.
+Proof. reflexivity. Qed.
+
+Example Ex4 (x y z : PrimRec 2) (t: PrimRec 3):
+  let u := composeFunc 2 3
+                       (PRcons 2 2 x
+                               (PRcons 2 1 y
+                                       (PRcons 2 0 z
+                                               (PRnil 2))))
+                       t in
+  let f := evalPrimRec 2 x in
+  let g := evalPrimRec 2 y in
+  let h := evalPrimRec 2 z in
+  let i := evalPrimRec 3 t in
+  let j := evalPrimRec 2 u in
+  forall a b, j a b = i (f a b) (g a b) (h a b).
+Proof. reflexivity. Qed.
+
+Example Ex5 (x : PrimRec 2)(y: PrimRec 4):
+  let g := evalPrimRec _ x in
+  let h := evalPrimRec _ y in
+  let f := evalPrimRec _ (primRecFunc _ x y) in
+  forall a b,  f 0 a b = g a b.
+Proof. reflexivity.   Qed.
+
+Example Ex6 (x : PrimRec 2)(y: PrimRec 4):
+  let g := evalPrimRec _ x in
+  let h := evalPrimRec _ y in
+  let f := evalPrimRec _ (primRecFunc _ x y) in
+  forall n a b,  f (S n) a b = h n (f n a b) a b.
+Proof. reflexivity.   Qed.                          
+\end{Coqsrc}
+
+Another example?
+Let us consider the following term\footnote{Of course, we never typed this term \emph{verbatim}; we obtained it as the result of some computation we leave it the reader to guess and reproduce.}:
+
+\label{sect:bigfac}
+\begin{Coqsrc}
+Example bigPR : PrimRec 1 :=
 primRecFunc 0
@@
 \begin{Coqsrc}
-Example  mystery_fun : nat _> nat := evalPrimRec 1 weirdPR.
-
-Compute List.map mystery_fun (0::1::2::3::4::5::6::nil) : list nat.
+Example  mystery_fun : nat -> nat := evalPrimRec 1 bigPR.
+
+Compute map mystery_fun[0;1;2;3;4;5;6] : t nat _.
 \end{Coqsrc}
@@
  = 1 :: 1 :: 2 :: 6 :: 24 :: 120 :: 720 :: nil
-     : list nat
-\end{Coqanswer}
-
-Ok, the term \texttt{weirdPR} looks to be a primitive recursive definition of the factorial function, although we haven't proved this fact yet.
+     : t nat 7
+\end{Coqanswer}
+
+Ok, the term \texttt{bigPR} looks to be a primitive recursive definition of the factorial function, although we haven't proved this fact yet.
 Fortunately, we will see in the following sections much simpler ways to prove that a given function is primitive recursive, whithout looking at an unreadable term.
@@
 Thus they do not allow the user to look at the witness of a proof of a \texttt{isPR} statement.
-Our example of page\pageref{sect:weirdfac} was built using a  copy of \texttt{primRec.v} where many \texttt{Qed}s have been replaced with
+Our example of page\pageref{sect:bigfac} was built using a  copy of \texttt{primRec.v} where many \texttt{Qed}s have been replaced with
 \texttt{Defined}s.
@@
 
-The best known of total function which is not primitive recursive is the Ackermann function. We show how to adapt the classic proof (see for instance~\cite{planetmath}) to the constraints of \gallina.
+The best known example of a total recursive function which is not primitive recursive is the Ackermann function.
+We show how to adapt the classic proof (see for instance~\cite{planetmath}) to the constraints of \gallina.
+We hope this formal proof
+ is a nice opportunity to explore
+the treatment of primitive recursive functions by R. O'Connor,
+and to play with dependent types.
 
 \subsection{The Ackermann function}
+
+Here is a possible definition, where (\texttt{iterate\,$f$\,$n$})
+is the $n$-th iteration of $f$.
+
+\vspace{4pt}
+\noindent
+\emph{From \href{../theories/html/hydras.Prelude.Iterates.html}{Prelude.Iterates}}.
+
+\begin{Coqsrc}
+Fixpoint iterate {A:Type}(f : A -> A) (n: nat)(x:A) :=
+  match n with
+  | 0 => x
+  | S p => f (iterate  f p x)
+  end.
+\end{Coqsrc}
 
@@
 \emph{From \href{../theories/html/hydras.Ackermann.MoreAck.Ack.html}{Ackermann.MoreAck.Ack}}.
+\index{Maths!Ackermann function}
 
@@
 \begin{Coqsrc}
-Lemma Ack_1_n n : Ack 1 n = S (S n).
+Lemma Ack_1_n n: Ack 1 n = S  (S n).
 
@@
 Please note that for any given $n$, the unary function
-(\texttt{Ack\,$n$} is primitive recursive.
+(\texttt{Ack\,$n$}) is primitive recursive.
 
@@
 \item Prove that for any primitive recursive function $f:\mathbb{N}\rightarrow\mathbb{N}\rightarrow\mathbb{N}$, there exists some natural number $n$ depending on $f$, such that, for any $x$ and $y$, 
-$f\,x\,y \leq \texttt{Ack}\,n\,(\textrm{max}\,x\,y)$ (we say that $f$ is \emph{majorized}  by \texttt{Ack}).
+$f\,x\,y \leq \texttt{Ack}\,n\,(\textrm{max}\,x\,y)$ (we say that $f$ is \emph{``majorized''}  by \texttt{Ack}).
 \item Show that \texttt{Ack} fails to satisfy this property.
@@
 
-We are now able to translate in \gallina{} the notion of majoration:
+We are now able to translate in \gallina{} the notion of ``majorization'':
 
@@
 \begin{Coqsrc}
+ assert (L1 : forall i (v: t nat n) ,
+               v_apply f (i::v)  <= Ack q (i + max_v v)).
+    { induction i.
+      ...
+    }
     ...
@@
     Variable n: nat.
-    Hypothesis  Hn :  forall x y,  Ack x y <= Ack n (max x  y) .
-
-    Remark R01 : n <> 0.
+    Hypothesis  Hn: forall x y,  Ack x y <= Ack n (max x  y).
+
+    Remark R01: n <> 0.
 
@@
 
-    Remark R03 : max 2 n = n.
-
-    Remark R04 : Ack n (2 * n + 3) = Ack n (Ack 2 n).
-
-    Remark R05 : Ack n (2 * n + 3) <= Ack (n + 2) n.
-
-    Remark R06 : Ack (n + 2) n <= Ack n (2 * n + 2).
-
-    Remark R07 : Ack n (2 * n + 3) <= Ack n (2 * n + 2).
-
-    Remark R08 :  Ack n (2 * n + 2) < Ack n (2 * n + 3).
+    Remark R03: max 2 n = n.
+
+    Remark R04: Ack n (2 * n + 3) = Ack n (Ack 2 n).
+
+    Remark R05: Ack n (2 * n + 3) <= Ack (n + 2) n.
+
+    Remark R06: Ack (n + 2) n <= Ack n (n + 2).
+
+    Remark R07: Ack n (n + 2) < Ack n (2 * n + 3).
 

### 2021-02-10 doc/chapter-powers.tex
creation of a multiple index

--- 2021-01-27 doc/chapter-powers.tex
@@
 \subsubsection{Using 2x2 integer matrices}
-\index{Fibonacci numbers!Matrix exponentiation}
+\index{maths}{Fibonacci numbers!Matrix exponentiation}
 
@@
 
-\index{Exercises}
+\index{additions}{Exercises}
 
@@
 
-\index{Exercises}
+\index{additions}{Exercises}
 \begin{exercise}
@@
 \label{op-classes}
-\index{Coq!Type Classes!Operational Type Classes}
+\index{coq}{Type classes!Operational type classes}
+≈Ò
 \emph{Operational type classes}~\cite{BS2011}
@@
 \subsection{The Monoid Type Class}
-\index{Coq!Type Classes}
+\index{coq}{Type classes}
 We are now ready to  give a definition of the \texttt{Monoid} class, using
@@
 \subsection{Monoids and Equivalence Relations}
-\index{Coq!Generalized rewriting}
-\index{Coq!Equivalence relations}
+\index{coq}{Generalized rewriting}
+\index{coq}{Type classes!Equivalence relations}
 
@@
 
-\index{Coq!Type Classes}
-\index{Coq!Proper}
+\index{coq}{Type classes}
+\index{Coq}{Type classes!Proper class}
 \label{EMonoid-def}
@@
 
-\index{Coq!Coercions}
+\index{coq}{Coercions}
 
@@
 \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Classes.Morphisms.html}{Coq.Classes.Morphisms}
-\index{Coq!Type Classes}
-\index{Coq!Proper}
+\index{coq}{Type classes}
+\index{coq}{Type classes!Proper class}
 \begin{Coqsrc}
@@
 \begin{itemize}
-\item Continuation Passing Style (CPS) \index{Programming Styles!Continuation Passing Style (CPS)} \cite{reynolds93}
+\item Continuation Passing Style (CPS) 
+\index{coq}{Continuation Passing Style (CPS)} \cite{reynolds93}
 is a way to make explicit the 
@@
 
-\item Like in Parametric Higher Order Abstract Syntax (PHOAS)~\cite{PHOAS}, \index{Parametric Higher-Order Abstract Syntax (PHOAS)}
+\item Like in Parametric Higher Order Abstract Syntax (PHOAS)~\cite{PHOAS}, \index{coq}{Parametric Higher-Order Abstract Syntax (PHOAS)}
      the local variables associated to intermediate results are
@@
 
-\index{Projects}
+\index{additions}{Projects}
 \begin{project}
@@
 \label{reflection-section}
-\index{Coq!Proof by reflection}
+\index{coq}{Proof by reflection}
 Instead of letting the tactic \texttt{rewrite} look for contexts in which
@@
 \subsubsection{About parametricity}
-\index{Coq!Parametricity}
+\index{coq}{Parametricity}
 Let us now present another tactic for proving chain correctness,
@@
 In effect, there is no other way to build terms of type \texttt{A} in the considered context.
+
+\index{coq}{Plug-ins!paramcoq}
 
@@
 \subsubsection{Definition:} 
-\index{Coq!Type Classes}
-\index{Coq!Proper}
+\index{coq}{Type classes}
+\index{coq}{Type classes!Proper class}
 A continuation \texttt{k:Fkont A} is \emph{proper}
@@
 
-\index{Programming Styles!Continuation Passing Style (CPS)}
+\index{coq}{Continuation Passing Style (CPS)}
 
@@
 any correct chain generator.
-\index{Coq!Dependently Typed Functions}
+\index{coq}{Dependently typed functions}
 
@@
 
-\index{Projects}
+\index{additions}{Projects}
 \begin{project}[Optimality and relative efficiency]

### 2021-02-10 doc/part-hydras.tex
add a new exercise
fix errors
creation of a multiple index

@@
 \section*{Future Work (projects)}
-\index{Projects}
+\index{hydras}{Projects}
 
@@
 %\index{To do}
-\index{Projects}
+\index{hydras}{Projects}
 
@@
 
-\index{Projects}
+\index{hydras}{Projects}
 
@@
 
-\index{Exercises}
+\index{hydras}{Exercises}
 
@@
 
-\index{Coq!Mutually inductive types}
-\index{Coq!Commands!Scheme}
+\index{coq}{Mutually inductive types}
+\index{coq}{Commands!Scheme}
 
@@
 
-\index{Predicates!round\_n}
+\index{hydras}{Predicates!round\_n}
 
@@
 
-\index{Predicates!round}
+\index{hydras}{Predicates!round}
 \label{sect:infix-round}
@@
 
-\index{Predicates!Termination}
+\index{hydras}{Predicates!Termination}
 
@@
 \subsection{A  Small Proof of Impossibility}
-\index{Maths!Proofs of impossibility}
+\index{coq}{Proofs of impossibility}
 
@@
 The next chapter is dedicated to a generic formalization of ordinal notations, and chapter~\ref{chap:T1} to a proof of termination of all hydra battles with the help of an ordinal notation for the interval $[0,\epsilon_0)$\,\footnote{We use the mathematical notation $[a,b)$ for the interval $\{x|a\leq x < b\}$.}.
-\index{Maths!Notations!Interval}
+\index{maths}{Notations!Interval}
 
@@
 
-\index{Maths!Ordinal numbers}
+\index{maths}{Ordinal numbers}
 
@@
 
-\index{Exercises}
+\index{hydras}{Exercises}
 \begin{exercise}
@@
 
-\section{The Ordinal \texorpdfstring{$\omega^2$}{omega^2}}
+\section{The Ordinal \texorpdfstring{$\omega^2$}{omega2}}
 
@@
 
-\index{Coq!Commands!Function}
+\index{coq}{Commands!Function}
 
@@
 
-\index{Maths!Transfinite induction}
+\index{maths}{Transfinite induction}
 
@@
 
-\index{Coq!Commands!Program}
+\index{coq}{Commands!Program}
 
@@
 
-\index{Projects}
+\index{hydras}{Projects}
 \begin{project}
@@
 
-\index{Projects}
+\index{hydras}{Projects}
 \label{exo:ON-mult}
@@
 \subsection{Cantor Normal Form}
-\index{Maths!Cantor normal form}
+\index{maths}{Cantor normal form}
 
@@
 
-\index{Notations!phi0@phi0 (exponential of base omega)}
+\index{hydras}{Notations!phi0@phi0 (exponential of base omega)}
 
@@
 
-\index{Maths!Additive principal ordinals}
+\index{maths}{Additive principal ordinals}
 
@@
 \label{types:ppT1}
-\index{Functions!pp@ pp (pretty printing terms in Cantor normal form)}
+\index{hydras}{Functions!pp@ pp (pretty printing terms in Cantor normal form)}
 
@@
 
-\index{Coq!Unicity of equality proofs}
+\index{coq}{Unicity of equality proofs}
 
@@
 
-\index{Exercises}
+\index{hydras}{Exercises}
  \begin{exercise}
@@
 %% To simplify !
-\index{Exercises}
+\index{hydras}{Exercises} 
 \begin{exercise}
@@
 \label{sec:orgheadline77}
-\index{Coq!Well-founded induction}
+\index{coq}{Well-founded induction}
 
@@
 \label{chap:ketonen}
-\index{Maths!Ketonen-Solovay machinery}
+\index{maths}{Ordinal numbers!Ketonen-Solovay machinery}
 
@@
 \label{ketonen-solovay-sect}
-\index{Maths!Canonical sequences}
+\index{maths}{Ordinal numbers!Canonical sequences}
 
@@
 
-\index{Maths!Transfinite induction}
+\index{maths}{Transfinite induction}
 \begin{itemize}
@@
 to \texttt{nat} is available on \href{../theories/html/hydras.Prelude.Fuel.html}{Prelude.Fuel}).
-\index{Coq!Giving fuel to a long computation}
+\index{coq}{Giving fuel to a long computation}
 
@@
 \section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0} : Paths}
-\index{Maths!Accessibility inside epsilon0}
+\index{maths}{Ordinal numbers!Accessibility inside epsilon0}
 \label{sect:pathes-intro}
@@
 
-\index{Predicates!path\_to}
+\index{hydras}{Predicates!path\_to}
 \label{sect:path-to-def}
@@
 \subsubsection{Sketch of Proof of \texttt{Cor12}}
-\index{Maths!Transfinite induction}
+\index{maths}{Transfinite induction}
 
@@
 \label{lemma:L-2_6-1}
-\index{Coq!Sigma types}
+\index{coq}{Sigma types}
 
@@
 
-\index{Maths!Large sequences}
-\index{Maths!Minimal large sequences}
+\index{maths}{Ordinal numbers!Large sets}
+\index{maths}{Ordinal numbers!Minimal large sets}
 
@@
 
-\index{Predicates!mlarge @ mlarge (minimal large sequences)}
+\index{hydras}{Predicates!mlarge @ mlarge (minimal large sequences)}
 \begin{Coqsrc}
@@
 
-\index{Coq!Plug-ins!Equations}
+\index{coq}{Plug-ins!Equations}
 
@@
 presented for instance in~\cite{Promel2013}. 
-\index{Maths!Hardy Hierarchy}
+\index{maths}{Rapidly growing functions!Hardy Hierarchy}
 
@@
 
-\index{Coq!Plug-ins!Equations}
+\index{coq}{Plug-ins!Equations}
 \label{Functions:H-alpha}
@@
 
-\index{Maths!Wainer Hierarchy}
+\index{maths}{Rapidly growing functions!Wainer Hierarchy}
 
@@
 
-\index{Coq!Plug-ins!Equations}
+\index{coq}{Plug-ins!Equations}
 \begin{Coqsrc}
@@
 
-\index{Coq!Plug-ins!Equations}
+\index{coq}{Plug-ins!Equations}
 \label{Functions:F-alpha}
-\index{Maths! Fast growing functions}
+\index{maths}{Rapidly growing functions}
 
@@
 
-\index{Maths!Ordering functions}
+\index{maths}{Ordinal numbers!Ordering functions}
 
@@
 
-\index{Maths!Additive principal ordinals}
+\index{maths}{Ordinal numbers!Additive principal ordinals}
 
@@
 
-\index{Predicates!Closed}
+\index{hydras}{Predicates!Closed}
 
@@
 
-\index{Maths!Critical ordinals}
+\index{maths}{Ordinal numbers!Critical ordinals}
 
@@
 Is it true ?
+
+\emph{You may start this exercise with the file
+\url{../exercises/ordinals/schutte_cnf_counter_example.v}.}
 \end{exercise}

### 2021-02-10 doc/hydras.tex
comments and examples about the Ackermann part (to continue)
add an exercise
creation of a multiple index
work on the index

@@
 \usepackage{alltt}
-\usepackage{makeidx}
+\usepackage{multind}
 \usepackage{hyperref}
@@
 \title{Hydra Battles and Cie}
-\makeindex
+\makeindex{coq}
+\makeindex{maths}
+\makeindex{hydras}
+\makeindex{additions}
+\makeindex{primrec}
 \begin{document}
@@
 
-\index{Coq!Commands!Print Assumptions}
+\index{coq}{Commands!Print Assumptions}
 
@@
 
+Please consult the sub-directory \texttt{exercises/} of the
+ project (in construction).
+
 \section{Acknowledgements}
@@
 
-\begin{todo}
- Still very incomplete!
-\end{todo}
+% \begin{todo}
+%  Still very incomplete!
+% \end{todo}
 
-\printindex
+\printindex{coq}{Coq, plug-ins and standard library}
+\printindex{maths}{Mathematical notions and algorithmics}
+\printindex{hydras}{Library hydras: Ordinals and hydra battles}
+\printindex{primrec}{Library hydras.Ackermann: Primitive recursive functions}
+\printindex{additions}{Library additions: Addition chains}
 
@@
 
-Ord & countable ordinals (after Schütte) & \pageref{types:Ord} \\
 ppT1 & pretty printed version of T1 & \pageref{types:ppT1} \\
@@
 T2 & ordinal terms below $\Gamma_0$ & \pageref{types:T2} \\
-WO & well orders (Schütte's definition) & \pageref{types:WO} \\
 \hline
@@
 H\_ & H\_ alpha n & $H_\alpha(n)$ & Hardy's \texttt{H} fast growing hierarchy & \pageref{Functions:H-alpha} \\
-iterate & iterate $f$ $n$ $x$ & $f^{(n)}(x)$ & Functional iteration & \pageref{Functions:iterate} \\
 L\_ & L\_ alpha k & $L_\alpha(k)$ & final step of a standard path & \pageref{Functions:L-alpha} \\
 succ & succ alpha & & Successor & \pageref{Functions:succ-T1}, \pageref{Functions:succ-sch} \dots \\
-zero:Ord & & $0$ & The least ordinal (Schütte's model) & \pageref{Constants:zero:Ord} \\
-
 \hline

### 2021-02-10 doc/part-hydras.tex
work on the index

@@
 is defined and studied in~\href{../theories/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
+\index{hydras}{Library Prelude!iterate}
 
@@
 
-\index{hydras}{Notations!phi0@phi0 (exponential of base omega)}
+\index{hydras}{Library Epsilon0!Notations!phi0@phi0 (exponential of base omega)}
 
@@
 
+\index{hydras}{Library Prelude!iterate}
 \begin{Coqsrc}
@@
 It is quite easy to prove that our function \texttt{F\_} satisfies the equations on page~\pageref{sect:F-equations}.
+\index{hydras}{Library Prelude!iterate}
 
@@
 
+\index{hydras}{Library Schutte!Types!Ord}
+
 \begin{Coqsrc}
@@
 {\texttt{Well\_Orders.v}}).
+
+\index{hydras}{Library Schutte!Type classes!WO@ WO (well order)}
 
@@
 \label{Constants:zero:Ord}
+\index{hydras}{Library Schutte!Constants!zero}
 
@@
 \label{Functions:succ-sch}
+\index{hydras}{Library Schutte!Functions!succ}
 
@@
 \index{maths}{Ordinal numbers!Ordering functions}
+\index{hydras}{Library Schutte!Predicates!ordering function@ordering\_function}
 
@@
 
+\index{hydras}{Library Schutte!Functions!plus}
 \begin{Coqsrc}
@@
 \index{maths}{Ordinal numbers!Additive principal ordinals}
+\index{hydras}{Library Schutte!AP@AP (additive principal ordinals)}
 
@@
 
+\index{hydras}{Library Schutte!Functions!phi0}
 \begin{Coqsrc}
@@
 
-\index{hydras}{Predicates!Closed}
+\index{hydras}{Library Schutte!Predicates!Closed}
 
@@
 \index{maths}{Ordinal numbers!Critical ordinals}
+\index{hydras}{Library Scutte!Cr@Cr (critical ordinals)}
 
@@
 sub-term\footnote{This would prevent us from trying to represent Cantor normal forms as finite trees (like in Sect.~\ref{sec:T1-inductive-def})}.
+
+\index{maths}{Ordinal numbers!Cantor normal form}
+\index{hydras}{Library Schutte!Predicates!is\_cnf\_of@is\_cnf\_o (to be a Cantor normal form}
 

### 2021-02-11 doc/chapter-powers.tex
Mult index (#28)
* creation of a multiple index

* work on the index

* work on the index

* reorganization of the index (in progress)

@@
 \label{reflection-section}
-\index{coq}{Proof by reflection}
+\index{coq}{Proofs by reflection}
 Instead of letting the tactic \texttt{rewrite} look for contexts in which

### 2021-02-11 doc/hydras.tex
work on the index
Mult index (#28)
* creation of a multiple index

* work on the index

* work on the index

* reorganization of the index (in progress)

@@
 %\SetWatermarkLightness{0.8}
-\definecolor{lightorange}{rgb}{1,0.7,0.5}
-\SetWatermarkScale{0.35}
-\SetWatermarkText{Work in Progress}
-\SetWatermarkColor{lightorange}
+% \definecolor{lightorange}{rgb}{1,0.7,0.5}
+% \SetWatermarkScale{0.35}
+% \SetWatermarkText{Work in Progress}
+% \SetWatermarkColor{lightorange}
 
@@
 % \end{todo}
+{\Large \textbf{In progress} This index is currently under reorganization for a few days. We aplologize for its incompleteness! }
 
@@
 Name & Description& Page \\\hline
-Battle & type class of battle types & \pageref{types:Battle} \\
-E0 & (well formed) ordinals below $\epsilon_0$ & \pageref{types:E0}\\
-Hvariant & variants for proving termination of hydra battles & \pageref{sect:hvariant-def} \\
-Hydra & tree-like representation of hydras & \pageref{types:Hydra} \\
-Hydrae & finite sequences of hydras & \pageref{types:Hydrae} \\
-ON & ordinal  notations & \pageref{types:ON} \\
-ON\_for & comparison of an ordinal notation with Schütte's model &  \pageref{types:ON-for}\\
-ON\_iso & isomorphism of ordinal notations  & \pageref{types:ON-iso}\\
-
 ppT1 & pretty printed version of T1 & \pageref{types:ppT1} \\
-SubON & comparison of ordinal notations & \pageref{types:SubON} \\
 T1 & ordinal terms below $\epsilon_0$ & \pageref{types:T1} \\

### 2021-02-11 doc/part-hydras.tex
Mult index (#28)
* creation of a multiple index

* work on the index

* work on the index

* reorganization of the index (in progress)

@@
 
+\index{hydras}{Library Hydra!Types!Hydra}
+\index{hydras}{Library Hydra!Types!Hydrae}
+
 In order to describe trees where each node can have an arbitrary (but finite) number of sons, it is usual to define a type where each node carries a \emph{forest}, \emph{i.e} a list of trees
@@
 
-\index{hydras}{Predicates!round\_n}
-
+\index{hydras}{Library Hydra!Predicates!round\_n}
 \begin{Coqsrc}
@@
 
-\index{hydras}{Predicates!round}
+\index{hydras}{Library Hydra!Predicates!round}
 \label{sect:infix-round}
@@
 \label{types:Battle}
+\index{hydras}{Library Hydra!Type classes!Battle}
 
@@
 
-\index{hydras}{Predicates!Termination}
+\index{hydras}{Library Hydra!Predicates!Termination}
 
@@
 
+\index{hydras}{Library Hydra!Type classes!Hvariant}
 \begin{Coqsrc}
@@
 \label{types:ON}
+\index{hydras}{Library OrdinalNotations!Type classes!ON}
 
@@
 
-If \texttt{OB} is presumed to be correct, the we can consider that \texttt{OA} ``inherits'' its correcteness from the bigger notation system \texttt{OB}.
+If \texttt{OB} is presumed to be correct, the we can consider that \texttt{OA} ``inherits'' its correctness from the bigger notation system \texttt{OB}.
 
 \label{types:SubON}
+\index{hydras}{Library OrdinalNotations!Type classes!SunON}
 
@@
 \label{types:ON-for}
+\index{hydras}{Library OrdinalNotations}{Type classes!ON\_correct}
 
@@
 For this purpose, one may prove that the two notation systems are order-isomorphic.
+
+\index{hydras}{Library OrdinalNotations!Type classes!ON\_Iso}
 
@@
 \label{types:ppT1}
-\index{hydras}{Functions!pp@ pp (pretty printing terms in Cantor normal form)}
+\index{hydras}{Library Epsilon0!Functions!pp@ pp (pretty printing terms in Cantor normal form)}
 
@@
 \label{types:E0}
+\index{hydras}{Library Epsilon0!Types!E0}
 
@@
 
-\index{hydras}{Predicates!path\_to}
+\index{hydras}{Library Epsilon0!Predicates!path\_to}
 \label{sect:path-to-def}
@@
 
-\index{hydras}{Predicates!mlarge @ mlarge (minimal large sequences)}
+\index{hydras}{Library Epsilon0!Predicates!mlarge@mlarge (minimal large sequences)}
 \begin{Coqsrc}
@@
 \index{maths}{Ordinal numbers!Additive principal ordinals}
-\index{hydras}{Library Schutte!AP@AP (additive principal ordinals)}
+\index{hydras}{Library Schutte!Predicates!AP@AP (additive principal ordinals)}
 
@@
 \index{maths}{Ordinal numbers!Critical ordinals}
-\index{hydras}{Library Scutte!Cr@Cr (critical ordinals)}
+\index{hydras}{Library Schutte!Predicates!Cr@Cr (critical ordinals)}
 
@@
 \index{maths}{Ordinal numbers!Cantor normal form}
-\index{hydras}{Library Schutte!Predicates!is\_cnf\_of@is\_cnf\_o (to be a Cantor normal form}
+\index{hydras}{Library Schutte!Predicates!is\_cnf\_of@is\_cnf\_of (to be a Cantor normal form of}
 

### 2021-02-12 doc/hydras.tex
reorganization of the index (in progress)

@@
 
-%\SetWatermarkLightness{0.8}
-% \definecolor{lightorange}{rgb}{1,0.7,0.5}
-% \SetWatermarkScale{0.35}
-% \SetWatermarkText{Work in Progress}
-% \SetWatermarkColor{lightorange}
+\SetWatermarkLightness{0.8}
+\definecolor{lightorange}{rgb}{1,0.7,0.5}
+\SetWatermarkScale{0.35}
+\SetWatermarkText{Work in Progress}
+\SetWatermarkColor{lightorange}
 
 \author{Pierre Castéran\\ LaBRI, Univ.
-Bordeaux, CNRS UMR 5800, Bordeaux-INP\\with contributions by Yves Bertot, \'Evelyne Contejean,  Florian Hatat and Pascal Manoury\\ \emph{The formalization of primitive recursive functions was originally authored by Russel O'Connor}}
+Bordeaux, CNRS UMR 5800, \\with contributions by Yves Bertot, \'Evelyne Contejean,  Florian Hatat and Pascal Manoury\\ \emph{The formalization of primitive recursive functions was originally authored by Russel O'Connor}}
 \date{\today}
@@
 \begin{itemize}
-\item The present distribution has been checked with version 8.12.0 of the Coq proof assistant, with the plug-ins \texttt{coq-paramcoq}, \texttt{coq-equations}
+\item The present distribution has been checked with version 8.13.0 of the Coq proof assistant, with the plug-ins \texttt{coq-paramcoq}, \texttt{coq-equations}
 and \texttt{coq-mathcomp-algebra}.
@@
 Pascal Manoury,  Karl Palmskog, Sylvain Salvati, Alan Schmitt and Théo Zimmerman for their help on the elaboration of this document, and to the
- members of the \emph{Formal Methods} team at laBRI for their helpful comments 
+ members of the \emph{Formal Methods} team and the \emph{Coq working group} at laBRI for their helpful comments 
 on  oral presentations of this work.
@@
 \section{Hydras and Ordinals}
-
-\begin{table}[H]
-  \centering
-  \begin{threeparttable}
-    \caption{Main types and type classes}
-\begin{tabular}{|c|c|c|l|}
-\hline
-Name & Description& Page \\\hline
-ppT1 & pretty printed version of T1 & \pageref{types:ppT1} \\
-T1 & ordinal terms below $\epsilon_0$ & \pageref{types:T1} \\
-T2 & ordinal terms below $\Gamma_0$ & \pageref{types:T2} \\
-\hline
-\end{tabular}
-\end{threeparttable}
-\end{table}
-
-\begin{table}[H]
-  \centering
-  \begin{threeparttable}
-    \caption{Main functions and constants}
-\begin{tabular}{|r | c|c|c|c|l|}
-\hline
-Name & Gallina & Math & Description & Page \\ \hline
-canon & canon alpha i & $\canonseq{\alpha}{i}$& Canonical sequence & \pageref{Functions:canon}\\
-canonS & canonS alpha i & $\canonseq{\alpha}{i+1}$& Helper for canon & \pageref{Functions:canonS}\\
-F\_ & F\_ alpha n & $F_\alpha(n)$ & Wainer's \texttt{F} fast growing hierarchy & \pageref{Functions:F-alpha} \\
-H\_ & H\_ alpha n & $H_\alpha(n)$ & Hardy's \texttt{H} fast growing hierarchy & \pageref{Functions:H-alpha} \\
-L\_ & L\_ alpha k & $L_\alpha(k)$ & final step of a standard path & \pageref{Functions:L-alpha} \\
-succ & succ alpha & & Successor & \pageref{Functions:succ-T1}, \pageref{Functions:succ-sch} \dots \\
-\hline
-\end{tabular}
-\end{threeparttable}
-\end{table}
 

### 2021-02-12 doc/part-hydras.tex
reorganization of the index (in progress)

@@
 module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}). 
+\emph{For a less technical description of hydras, please see 
+\url{https://www.smbc-comics.com/comic/hydra}.}
 
@@
 
-If \texttt{OB} is presumed to be correct, the we can consider that \texttt{OA} ``inherits'' its correctness from the bigger notation system \texttt{OB}.
+If \texttt{OB} is presumed to be correct, then we may consider that \texttt{OA} ``inherits'' its correctness from the bigger notation system \texttt{OB}.
 
 \label{types:SubON}
-\index{hydras}{Library OrdinalNotations!Type classes!SunON}
+\index{hydras}{Library OrdinalNotations!Type classes!SubON}
 
@@
 \label{types:ON-for}
-\index{hydras}{Library OrdinalNotations}{Type classes!ON\_correct}
+\index{hydras}{Library OrdinalNotations!Type classes!ON\_correct}
 
@@
 %\index{Constants!zero:T1}
+\index{hydras}{Library Epsilon0!Types!T1}
 
@@
 \label{sect:ppT1}
+\index{hydras}{Library Epsilon0!Types!ppT1}
 
@@
 \subsubsection{Successor}
+
+\index{hydras}{Library Epsilon0!Functions!succ}
 
@@
 \subsubsection{Canonical Sequences in Coq}
+\index{hydras}{Library Epsilon0!Functions!canon}
+\index{hydras}{Library Epsilon0!Functions!canonS}
 
@@
 \label{Functions:L-alpha}
-
+ \index{hydras}{Library Epsilon0!Functions!L\_@L\_ (final step of a minimal path}
 \begin{Coqsrc}
@@
 
+\index{hydras}{Library Epsilon0!Functions!H\_@H\_ (Hardy hierarchy)}
 \index{coq}{Plug-ins!Equations}
@@
 \index{maths}{Rapidly growing functions}
+\index{hydras}{Library Epsilon0!Functions!F\_@F\_ (Wainer hierarchy)}
 
@@
 of the set of maximal $\alpha$-critical ordinals.
+
+\index{hydras}{Library Gamma0!Types!T2}
 

### 2021-02-15 doc/chapter-powers.tex
moving sub-library MoreAck to theories/ordinals

@@
 
+\emph{The \texttt{Monoid} type class is defined 
+page~\pageref{sect:monoid-def}.}
+
 \begin{Coqsrc}
@@
 (in sections~\vref{sect:fibonacci-pos-bpow}).
-
-\begin{todo}
-Put links when done.
-\end{todo}
 
@@
 
+\label{sect:monoid-def}
+
+\index{additions}{Type classes!Monoid}
+
 \begin{Coqsrc}
@@
 %\todo{link to Proper in stdlib : Coq.Classes.Morphisms and Coq.Classes.CMorphisms}
+
+\index{additions}{Type classes!EMonoid}
 
@@
 \section{Addition chains}
+\index{maths}{Addition chains}
 An \emph{addition chain} (In short : \emph{a chain})~\cite{brauer1939} is a representation of a sequence of
-intermediate steps that lead to the evaluation of some $x^n$, under the 
+intermediate steps that lead to the evaluation of  $x^n$, under the 
 assumption that each of these steps is a computation of  a power $x^i$, with 
@@
 
+\index{additions}{Types!computation}
+
 The following \emph{monadic} 
@@
 Please look at \emph{CPDT}~\cite{chlipalacpdt2011} for more details on monadic notations in \coq.
+\label{monadic-mult}
 
@@
 on \texttt{A}:
+
+\index{additions}{Types!chain@chain (addition chains)}
 
@@
 \label{euclide-sect}
-
+\index{maths}{Euclidean addition chains}
 In this section, we present an efficient chain generator. The chains built by this generator

### 2021-02-15 doc/chapter-primrec.tex
creation of a multiple index
work on the index
moving sub-library MoreAck to theories/ordinals

@@
 Thus, we propose a slow exploration of this library, through examples and exercises. Our additions to the original library 
-are stored in a sub-directory \texttt{Ackermann/MoreAck}.
+are stored in the directory \texttt{theories/ordinals/MoreAck}.
 
 \section{Basic Definitions}
-\index{Maths!Primitive recursive functions}
+\index{maths}{Primitive recursive functions}
 
@@
 
+\index{primrec}{Types!naryFunc}
 \begin{Coqsrc}
@@
 \noindent
-\emph{From\href{../theories/html/hydras.Ackermann.MoreAck.PrimRecExamples.html}{Ackermann.MoreAck.PrimRecExamples}}.
+\emph{From\href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}}.
 \begin{Coqsrc}
@@
 
-\index{Coq!Dependent types}
-\index{Coq!Dependently typed functions}
+\index{coq}{Dependent types}
+\index{coq}{Dependently typed functions}
 \vspace{4pt}
@@
 \emph{From \href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat}}
+
+\index{primrec}{Predicates!extEqual}
+
 \begin{Coqsrc}
@@
 
-\index{Coq!Dependent types}
-\index{Coq!Mutually inductive types}
-
+\index{coq}{Dependent types}
+\index{coq}{Mutually inductive types}
+
+\index{primrec}{Types!PrimRec}
+\index{primrec}{Types!PrimRecs}
 \label{def:Primrec}
@@
 
-\index{Coq!Dependent pattern matching}
+\index{primrec}{Functions!evalPrimRec}
+\index{primrec}{Functions!evalPrimRecs}
+
+\index{coq}{Dependent pattern matching}
 Both functions are mutually defined through dependent pattern matching. We advise the readers who 
@@
 
+\index{primrec}{Predicates!isPR}
+\index{coq}{Extensionnaly equal functions}
+
 Let $f$ be an arithmetic function of arity $n$. We say that $f$ is primitive recursive if $f$ is \textbf{extensionnaly}
@@
 we wrote alternate proofs in 
-\href{../theories/html/hydras.Ackermann.MoreAck.PrimRecExamples.html}%
+\href{../theories/html/hydras.MoreAck.PrimRecExamples.html}%
 {Ackermann.MoreAck.PrimRecExamples.v}, in order to illustrate the main proof patterns.
@@
 Projections are proved primitive recursive, case by case (many examples in 
-\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec.v}).
+\href{../theories/html/hydras.primRec.html}{Ackermann.primRec.v}).
 \emph{Please notice again that the name of the projection follows the mathematical tradition, 
@@
 \noindent
-\emph{From \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.}
+\emph{From \href{../theories/html/hydras.primRec.html}{Ackermann.primRec}.}
 
@@
 \noindent
-\emph{From \href{../theories/html/hydras.Ackermann.MoreAck.PrimRecExamples.html}{Ackermann.MoreAck.PrimRecExamples}}.
+\emph{From \href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}}.
 \begin{Coqsrc}
@@
 
-The lemma \texttt{plusIsPR} is already proven in \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}. We present in 
-\href{../theories/html/hydras.Ackermann.MoreAck.PrimRecExamples.html}{Ackermann.MoreAck.PrimRecExamples}
+The lemma \texttt{plusIsPR} is already proven in \href{../theories/html/hydras.primRec.html}{Ackermann.primRec}. We present in 
+\href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}
 a commented version of this proof,
@@
 \begin{todo}
-Comment more examples from   \href{../theories/html/hydras.Ackermann.MoreAck.PrimRecExamples.html}{Ackermann.MoreAck.PrimRecExamples}.
+Comment more examples from   \href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}.
 \end{todo}
 
+\index{primrec}{Exercises}
 \begin{exercise}
@@
 
+\index{primrec}{Exercises}
 \begin{exercise}
@@
 
-\index{Exercises}
+\index{primrec}{Exercises}
 
@@
 
-\index{Exercises}
+\index{primrec}{Exercises}
 \begin{exercise}
@@
 \emph{From \href{../theories/html/hydras.Prelude.Iterates.html}{Prelude.Iterates}}.
-
+\index{hydras}{Library Prelude!iterate}
 \begin{Coqsrc}
@@
 \noindent
-\emph{From \href{../theories/html/hydras.Ackermann.MoreAck.Ack.html}{Ackermann.MoreAck.Ack}}.
-\index{Maths!Ackermann function}
+\emph{From \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack}}.
+\index{maths}{Ackermann function}
+\index{primrec}{Ackermann function}
 
@@
 
-\emph{From \href{../theories/html/hydras.Ackermann.MoreAck.AckNotPR.html}{Ackermann.MoreAck.AckNotPR}}.
+\emph{From \href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR}}.
 \begin{Coqsrc}
@@
 
+\index{coq}{Commands!Scheme}
 \begin{Coqsrc}
@@
 
+\index{coq}{Dependently typed functions}
 \begin{Coqsrc}
@@
 
-In Library\href{../theories/html/hydras.Ackermann.MoreAck.AckNotPR.html}{Ackermann.MoreAck.AckNotPR}, we introduce a lighter notation.
-
+In Library\href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR}, we introduce a lighter notation.
+
+\index{coq}{Dependently typed functions}
 \begin{Coqsrc}
@@
 The rest of the cases are proved within a mutual  induction.
+
+\index{coq}{Mutual induction}
 

### 2021-02-15 doc/hydras.tex
moving sub-library MoreAck to theories/ordinals

@@
 \SetWatermarkLightness{0.8}
-\definecolor{lightorange}{rgb}{1,0.7,0.5}
+\definecolor{lightorange}{rgb}{1,0.9,0.9}
 \SetWatermarkScale{0.35}
@@
 \author{Pierre Castéran\\ LaBRI, Univ.
-Bordeaux, CNRS UMR 5800, \\with contributions by Yves Bertot, \'Evelyne Contejean,  Florian Hatat and Pascal Manoury\\ \emph{The formalization of primitive recursive functions was originally authored by Russel O'Connor}}
+Bordeaux, CNRS UMR 5800, \thanks{With contributions by Yves Bertot, \'Evelyne Contejean,  Florian Hatat and Pascal Manoury.
+\emph{The formalization of primitive recursive functions was originally authored by
+Russel O'Connor\cite{OConnor05}.}}}
 \date{\today}
@@
 \include{chapter-primrec}
-\part{Some Funny Algorithms}
+\part{Some  Certified Algorithms}
 
@@
 
-\section{Hydras and Ordinals}
+\section{Main notations}
 
@@
   \begin{threeparttable}
-    \caption{Main predicates}
+    \caption{Ordinals and ordinal notations}
 \begin{tabular}{|r | c|c|c|c|l|}
@@
 \texttt{nf: T1->Prop} & \texttt{nf alpha} && alpha is in Cantor normal form & \pageref{Predicates:nf-T1}\\
+ \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality \tnote{4} & \pageref{sect:on-lt-notation}\\
+ \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
+\texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & ordinal addition & \pageref{sect:infix-plus-T1}, \dots\\
+\texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\
+
+F & \texttt{F $n$} & $n$ & The $n$-th finite ordinal &  
+\pageref{sect:notation-F}, \pageref{sect:notation-F-sch}\\ 
+FS & \texttt{FS $n$} & $n+1$ & The $n+1$-th finite ordinal  \tnote{5} &  
+\pageref{sect:notation-FS}\\ 
+omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega}, \pageref{sect:omega-T1}, \pageref{sect:omega-notation2}, \dots\\
+phi0     & \texttt{phi0 alpha} & $\phi_0(\alpha),\; \omega^\alpha$&exponential of base $\omega$ & \pageref{sect:notation-phi0}\\
+
 \hline
@@
 \item[3] This order is total \emph{and} well-founded.
+\item [4]
+Some notations may belong to several scopes. For instance, ``\texttt{o<}'' is
+bound in \texttt{ON\_scope}, \texttt{E0\_scope}, \texttt{t1\_scope}, etc., and locally in several libraries.
+  \item [5] Note that there exist also various coercions from \texttt{nat} to types of ordinal.
+Depending on the current scope and  \coq's syntactic analysis algorithm, \texttt{F} may be left implicit.
 \end{tablenotes}
@@
 \vspace{4pt}
+
 \begin{table}[H]
-  \centering
+
   \begin{threeparttable}
-    \caption{Infix notations}
+    \caption{hydra battles}
 \begin{tabular}{|c|c|c|c|l|}
@@
 Name & Gallina&Math& Description& Page \\\hline
- \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality \tnote{1} & \pageref{sect:on-lt-notation}\\
- \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
-\texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & ordinal addition & \pageref{sect:infix-plus-T1}, \dots\\
-\texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\
 \texttt{round} & \texttt{h -1-> h'} & & one round of a battle & \pageref{sect:infix-round} \\
@@
 \texttt{round\_star} & \texttt{h -*-> h'} & & any number of rounds of a battle & \pageref{sect:infix-rounds} \\
-
 \hline
@@
 
-\begin{tablenotes}
- \item [1]
-Some notations may belong to several scopes. For instance, ``\texttt{o<}'' is
-bound in \texttt{ON\_scope}, \texttt{E0\_scope}, \texttt{t1\_scope}, etc., and locally in several libraries.
-\end{tablenotes}
+%\begin{tablenotes}
+%\end{tablenotes}
+
   \end{threeparttable}
+
 \end{table}
@@
   \begin{threeparttable}
-    \caption{Other notations}
+    \caption{Addition chains}
 \begin{tabular}{|c|c|c|c|l|}
@@
 Name & Gallina&Math& Description& Page \\\hline
-F & \texttt{F $n$} & $n$ & The $n$-th finite ordinal &  
-\pageref{sect:notation-F}, \pageref{sect:notation-F-sch}\\ 
-FS & \texttt{FS $n$} & $n+1$ & The $n+1$-th finite ordinal  \tnote{2} &  
-\pageref{sect:notation-FS}\\ 
-omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega}, \pageref{sect:omega-T1}, \pageref{sect:omega-notation2}, \dots\\
-phi0     & \texttt{phi0 alpha} & $\phi_0(\alpha),\; \omega^\alpha$&exponential of base $\omega$ & \pageref{sect:notation-phi0}\\
-
+\texttt{Mult} & \texttt{$z$ <--- $x$ times $y$} & & monadic notation & \pageref{monadic-mult} \\
 \hline
@@
 
-\begin{tablenotes}
-  \item [2] Note that there exist also various coercions from \texttt{nat} to types of ordinal.
-Depending on the current scope and  \coq's syntactic analysis algorithm, \texttt{F} may be left implicit.
-\end{tablenotes}
+%\begin{tablenotes}
+%\end{tablenotes}
+
   \end{threeparttable}
@@
 
-\section{Addition Chains}
-
-\begin{todo}
-\end{todo}
-
 \end{document}

### 2021-02-17 doc/chapter-primrec.tex
minor corrections

@@
 \index{primrec}{Exercises}
+\begin{exercise}
+Show that the function \texttt{min: naryFunc\,2} is primitive
+recursive.
+
+\emph{You may start this exercise with
+\url{../exercises/primrec/MinPR.v}.}
+
+\end{exercise}
+
+\index{primrec}{Exercises}
 
@@
 
-\emph{You may start this exercise with
+\emph{You may start this exercise with the file
 \url{../exercises/primrec/FibonacciPR.v}.}
-
-\end{exercise}
-
-\index{primrec}{Exercises}
-\begin{exercise}
-Show that the function \texttt{min: naryFunc\,2} is primitive
-recursive.
-
-\emph{You may start this exercise with
-\url{../exercises/primrec/MinPR.v}.}
 
@@
 
-\subsection{The Ackermann function}
+\subsection{Ackermann function}
 
 Here is a possible definition, where (\texttt{iterate\,$f$\,$n$})
-is the $n$-th iteration of $f$.
+is the $n$-th iteration of $f$\,\footnote{Please not confuse with \texttt{primRec.iterate}, which is monomorphic and does not share the same order of arguments.}
 
@@
 
+Another way of defining this function is 
+using an inner \texttt{fix} (see for instance the discussion in \url{https://stackoverflow.com/questions/10292421/error-in-defining-ackermann-in-coq}).
+
+We preferred to use the former definition, in order to
+take advantage from generic lemmas about \texttt{iterate}.
+
+\begin{Coqsrc}
+Module Alt.
+
+Fixpoint Ack (n m : nat) : nat :=
+  match n with
+  | O => S m
+  | S p => let fix Ackn (m : nat) :=
+               match m with
+               | O => Ack p 1
+               | S q => Ack p (Ackn q)
+               end
+           in Ackn m
+  end.
+
+End Alt.
+\end{Coqsrc}
+
 This function is also known under the form of three equations, which we derive immediately from our definition.
@@
 
-Please note that for any given $n$, the unary function
+An important property of the Ackermann function helps us 
+to overcome the difficulty raised by nested recursion, by climbing up the hierarchy $\texttt{Ack}\,n\,\_\;(n\in\mathbb{N})$.
+
+\begin{Coqsrc}
+Lemma Ack_Ack_le : forall k m n, 
+    Ack k (Ack m n) <= Ack (2 + max k m) n.
+\end{Coqsrc}
+
+Please note also that for any given $n$, the unary function
 (\texttt{Ack\,$n$}) is primitive recursive.
@@
 
-Then, our impossibility proof is just a sequence of easy small steps.
+Please note that, if the Ackermann function were primitive recursive, then there would exist some natural number $n$, such that, for all $x$ and $y$, the inequality 
+$\texttt{Ack}\,x\,y\leq \texttt{Ack}\,n\,(\texttt{max}\,x\,y)$ holds.
+Thus, our impossibility proof is just a sequence of easy small steps.
 
@@
   Fact Ack_majorized : majorized (Ack: naryFunc 2)  Ack.
-
+ \end{Coqsrc}
+
+ In order to share the variable \texttt{n}, let us open a section.
+
+\begin{Coqsrc}
   Section Contrad.
@@
     Hypothesis  Hn: forall x y,  Ack x y <= Ack n (max x  y).
-
+\end{Coqsrc}
+
+Firs, we get rid of the cases $n=0$ and $n=1$, thus we can rewrite
+(\texttt{max 2 n}) into \texttt{n}.
+
+  \begin{Coqsrc}
     Remark R01: n <> 0.
@@
     Remark R03: max 2 n = n.
-
+\end{Coqsrc}
+
+Please keep in mind that \texttt{(2 * n + 3)} can be safely replaced with 
+\texttt{(Ack 2 n)}.
+
+\begin{Coqsrc}
     Remark R04: Ack n (2 * n + 3) = Ack n (Ack 2 n).
@@
     Remark R05: Ack n (2 * n + 3) <= Ack (n + 2) n.
-
+\end{Coqsrc}
+
+ By hypothesis \texttt{Hn}, the following inequality holds.
+
+ \begin{Coqsrc}
     Remark R06: Ack (n + 2) n <= Ack n (n + 2).
-
+ \end{Coqsrc}
+
+By monotony of the Ackerman function, we get:
+
+\begin{Coqsrc}
     Remark R07: Ack n (n + 2) < Ack n (2 * n + 3).
-
+\end{Coqsrc}
+
+which contradicts \texttt{R05}.
+
+\begin{Coqsrc}
     Lemma contrad : False.

### 2021-02-19 doc/chapter-primrec.tex
new exercise on Pr functions
correction of typos

@@
 
+\index{primrec}{Exercises}
+\begin{exercise}
+Prove the following lemmas (which may help to solve thenext  exercise).
+
+\begin{Coqsrc}
+Lemma boundedSearch3 (P: naryRel 2) (b  : nat), 
+    boundedSearch P b <= b.
+
+Lemma boundedSearch4 (P: naryRel 2) (b  : nat):
+  P b b = true -> P b (boundedSearch P b) = true.
+\end{Coqsrc}
+\end{exercise}
+
+\index{primrec}{Exercises}
+
+\begin{exercise}
+Prove that the function which returns the  integer square root of any natural number  is primitive recursive.
+
+\emph{You may start this exercise with the file
+\url{../exercises/primrec/isqrt.v}.}
+
+\end{exercise}
+
 \section{Proving that a given function is \emph{not} primitive recursive}

### 2021-02-24 doc/part-hydras.tex
restructuring some .v files
minor corrections

@@
 attractive:  they are defined as functions  in \gallina{}, and we can apply them \emph{in theory}, but they are so complex that you will never be able to look at the result of the computation.
- Thus, our knowledge on these functions must rely on \emph{proofs}. In our development, we use often the rewriting rules generated by \coq's \texttt{Equations} plug-in.
+ Thus, our knowledge on these functions must rely on \emph{proofs}, not tests. In our development, we use often the rewriting rules generated by \coq's \texttt{Equations} plug-in.
 
@@
 \begin{definition}
-Let $0<\alpha<\epsilon_0$ be any ordinal, and $s$ a sequence of strictly positive natural numbers. 
+Let $0<\alpha<\epsilon_0$ be any ordinal, and $s$ a finite sequence $\langle s_1, s_2, \dots, s_N\rangle$ of strictly positive natural numbers.
+
+We say that $s$ is \emph{$\alpha$-large} if the sequence $\langle \alpha_0=\alpha,\dots,\alpha_{i+1}=\canonseq{\alpha_i}{i+1},\dots \rangle$ leads to $0$.
+
 We say that $s$ is \emph{minimally $\alpha$-large} (in short:
-\emph{$\alpha$-mlarge}) if $s$ if $s$ is $\alpha$-large 
+\emph{$\alpha$-mlarge}) if $s$ is $\alpha$-large 
  and every strict prefix of $s$ leads to a non-zero ordinal (\emph{cf} Sect.~\vref{sect:path-to-def}).
@@
 
-\vspace{4pt}
-
-\noindent
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}
-
-\index{hydras}{Library Epsilon0!Predicates!mlarge@mlarge (minimal large sequences)}
-\begin{Coqsrc}
-Definition mlarge alpha (s:list nat) := path_to zero s alpha.
-\end{Coqsrc}
-
 \begin{remark}
   Ketonen and Solovay~\cite{KS81} consider  large finite \emph{sets} of natural numbers,  but they are mainly used as sequences. Thus, we chosed to represent them explicitely as (sorted) lists.
-
-They also consider large (but not minimally large) sets. They can be defined in
-\coq{} as follows:
 
@@
 \end{Coqsrc}
+
+\vspace{4pt}
+
+\noindent
+\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}
+
+\index{hydras}{Library Epsilon0!Predicates!mlarge@mlarge (minimal large sequences)}
+\begin{Coqsrc}
+Definition mlarge alpha (s:list nat) := path_to zero s alpha.
+\end{Coqsrc}
+
 \end{remark}
@@
 \begin{Coqsrc}
-Lemma LF1 : forall n,  F_ 1 n = S (2 * n).
+Lemma LF1 : forall i,  F_ 1 i = S (2 * i).
 
@@
 \index{hydras}{Exercises}
+
+\begin{exercise}
+Prove the following property:
+
+\begin{Coqsrc}
+Lemma LF3 : dominates_from  2 (F_ 3) (fun  n => iterate exp2 n n).
+\end{Coqsrc}
+\end{exercise}
 

### 2021-02-25 doc/part-hydras.tex
removed useless notation scopes

@@
 
-Lemma LF2 : forall i, (exp2 i * i < F_ 2 i)%nat.
+Lemma LF2 : forall i, exp2 i * i < F_ 2 i.
 \end{Coqsrc}
@@
 
-Theorem F_alpha_ge_S alpha : forall n, (n < F_ alpha n)%nat.
+Theorem F_alpha_ge_S alpha : forall n, n < F_ alpha n.
 
@@
 
-Theorem F_alpha_beta alpha : forall beta n, Canon_plus n alpha beta -> 
-                                        (F_ beta n <= F_ alpha n)%nat.
+Theorem F_alpha_beta alpha : 
+forall beta n, Canon_plus n alpha beta -> 
+                                        F_ beta n <= F_ alpha n.
 \end{Coqsrc}

### 2021-03-02 doc/part-hydras.tex
new exercise
new exercise; minor corrections
small corrections

@@
 module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}). 
-\emph{For a less technical description of hydras, please see 
+\emph{For a less formal description of hydras, please see 
 \url{https://www.smbc-comics.com/comic/hydra}.}
@@
 At the first round, Hercules choses to chop off the rightmost head of \texttt{Hy}.
-Since this head is near the floor, the hydra loses this head. Let us call 
+Since this head is near the floor, the hydra simply loses this head. Let us call 
  \texttt{Hy'} the resulting state of the hydra, represented in Fig.~\vref{fig:Hy-prime}.
@@
 Next, assume Hercules choses to chop off one of the two highest heads of \texttt{Hy'}, for instance the rightmost one.
-Fig.~\vref{fig:Hy2} represents the rotten neck in dashed lines, and the part that will be replicated in red.
+Fig.~\vref{fig:Hy2} represents the broken segment in dashed lines, and the part that will be replicated in red.
 Assume also that the hydra decides to add 4 copies of the red part\footnote{In other words, the replication factor at this round is equal to $4$.}.
@@
 \begin{remark}
-The upper bound of the considered interval has been chosen to be  $l-1$ and not $l$, in order to simplify some statements and proofs in composition lemmas associated to ordinals of the form $\alpha\times i$ and
+The upper bound of the considered interval has been chosen to be  $l-1$ and not $l$, in order to simplify some statements and proofs in composition lemmas associated with the ordinals of the form $\alpha\times i$ and
 $\omega^\alpha\times i + \beta$.
@@
 
+These properties are defined in \coq{} in the library
+~\href{../theories/html/hydras.Prelude.Iterates.html}{Prelude.Iterates}.
+
+\index{maths}{Abstract properties of arithmetic functions}
+\index{hydras}{Abstract properties of arithmetic functions}
+
+\begin{Coqsrc}
+(** ** Abstract properties of arithmetic functions *)
+
+Definition strict_mono f := forall n p,  n < p -> f n < f p.
+
+Definition strict_mono1 f := forall n p,  0 < n < p -> f n < f p.
+
+Definition dominates_from n g f  := forall p, n <= p -> f p < g p.
+
+Definition dominates_strong g f  := {i : nat | dominates_from i g f}.
+
+Definition dominates g f := exists i : nat, dominates_from i g f .
+
+Definition fun_le f g  := forall n:nat,  f n <=  g n.
+
+Infix ">>" := dominates (at level 60).
+
+Infix ">>s" := dominates_strong (at level 60).
+
+Infix "<<=" := fun_le (at level 60).
+\end{Coqsrc}
+
 \index{maths}{Transfinite induction}
@@
 $\alpha$ is a successor are considered. Inside each case, the five sub-properties are proved sequentially.
-
-\begin{definition}
- Let $f$ zand $g$ be two arithmetic  functions; $f$ is said to \emph{dominate} $g$ if $f(p)>g(p)$ for any all sufficiently large $p$.
-\end{definition}
 
@@
 The term "F_ (Pred alpha) ?x" has type "Pred alpha o< alpha -> nat"
-while it is expected to have type "Pred alpha o< alpha -> Pred alpha o< alpha"
+while it is expected to have type 
+"Pred alpha o< alpha -> Pred alpha o< alpha"
 (cannot unify "nat" and "Pred alpha o< alpha").
@@
 
-Theorem F_alpha_dom alpha : dominates_from 1 (F_ (Succ alpha)) (F_ alpha).
+Theorem F_alpha_dom alpha : 
+     dominates_from 1 (F_ (Succ alpha)) (F_ alpha).
 
@@
 \end{Coqsrc}
+
+\emph{You may start this exercise with the file
+\url{../exercises/ordinals/F_3.v}.}
+\end{exercise}
+
+\index{hydras}{Exercises}
+
+\begin{exercise}
+Prove that, for any $\alpha\geq 3$ and $n\geq 2$,
+$F_\alpha(1+n)\geq 2^{F_\alpha(n)}$.
+
+\emph{You may start this exercise with the file
+\url{../exercises/ordinals/F_3.v}.}
+\end{exercise}
+
+%ici
+\index{hydras}{Exercises}
+
+\begin{exercise}
+It is tempting to prove a simple property of monotony 
+of the function \texttt{F\_}.
+
+\begin{quote}
+   Let $\alpha\leq\beta<\epsilon_0$. For any $n\geq 2$,
+$F_\alpha(n)\leq F_\beta(n)$. 
+\end{quote}
+Prove or disprove this statement.
+
+\emph{You may start this exercise with the file
+\url{../exercises/ordinals/is_F_monotonous.v}.}
 \end{exercise}
@@
 
-\end{exercise}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove the following result~\cite{KS81}(p. 298).
-
-\begin{quote}
-   For $n\geq 2$ and $\alpha \geq 3$, $F_\alpha(n+1)\geq 2^{F_\alpha(n)}$.
-\end{quote}
 \end{exercise}

### 2021-03-05 doc/part-hydras.tex
add new exercise
new exercise
improved an exercise

@@
 
+\index{hydras}{Exercises}
+\begin{exercise}
+
+Prove that for any $n\geq 2$, $\textrm{Ack}\,\,n\,n\leq  F_\omega(n)$, where \textrm{Ack} is the Ackermann function.
+Next, prove that $F_\alpha$ is not primitive recursive, for any $\alpha\geq\omega$  (please see Sect.~\vref{sect:ack-not-PR}).
+On the other hand, please show that for any natural number $n$, the function $F_n$ is primitive recursive.
+
+\emph{You may start this exercise with the file
+\url{../exercises/ordinals/F_omega.v}. Properties of the Ackermann function are studied in 
+\url{../theories/ordinals/MoreAck/Ack.v} and
+\url{../theories/ordinals/MoreAck/AcknotPR.v}.}
+\end{exercise}
+
+\index{hydras}{Exercises}
 \begin{exercise}

### 2021-03-06 doc/chapter-primrec.tex
new exercise
minor corrections

@@
 \begin{exercise}
-Prove the following lemmas (which may help to solve thenext  exercise).
+Prove the following lemmas (which may help to solve the next  exercise).
 
@@
 \section{Proving that a given function is \emph{not} primitive recursive}
+\label{sect:ack-not-PR}
 
@@
 
-We preferred to use the former definition, in order to
+We prefered to use the former definition, in order to
 take advantage from generic lemmas about \texttt{iterate}.
@@
 
+\begin{remark}
+It is easy to prove that any unary function which dominatates \texttt{fun n => Ack n n} fails to be primitive recursive. We use an instance on \texttt{majorAnyPR} for unary functions.
+\vspace{4pt}
+\noindent
+
+\emph{From \href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR}}.
+\begin{Coqsrc}
+Lemma majorPR1  (f: naryFunc 1)(Hf : isPR 1 f)
+  : exists (n:nat), forall x ,  f x  <= Ack n x.
+(* ... *)
+\end{Coqsrc}
+
+Then, we write  a short proof by contradiction.
+
+\begin{Coqsrc}
+Section dom_AckNotPR.
+
+  Variable f : nat -> nat.
+  Hypothesis Hf : dominates f (fun n => Ack n n).
+
+ Lemma dom_AckNotPR: isPR 1 f -> False.
+  Proof.
+    intros H;  destruct Hf as [N HN].
+    destruct  (majorPR1 _ H) as [M HM].
+    pose (X := Max.max N M).
+    specialize (HN X  (Max.le_max_l N M)).
+    specialize (HM X);
+      assert (Ack M X <= Ack X X) by (apply Ack_mono_l; subst; lia).
+    lia.
+  Qed.
+
+End dom_AckNotPR.
+\end{Coqsrc}
+\end{remark}
+
 %%% ICI

### 2021-03-07 doc/chapter-primrec.tex
renaming of a lemma

@@
 \begin{remark}
-It is easy to prove that any unary function which dominatates \texttt{fun n => Ack n n} fails to be primitive recursive. We use an instance on \texttt{majorAnyPR} for unary functions.
+It is easy to prove that any unary function which dominatates \texttt{fun n => Ack n n} fails to be primitive recursive. We use an instance of \texttt{majorAnyPR} for unary functions.
 \vspace{4pt}
@@
 Lemma majorPR1  (f: naryFunc 1)(Hf : isPR 1 f)
-  : exists (n:nat), forall x ,  f x  <= Ack n x.
+  : exists (n:nat), forall x, f x <= Ack n x.
 (* ... *)
@@
 
-%%% ICI
-
-% \item Show that for all $n$ and $m$, $A\,n\,m > m$ and  that the functions $A\,n$ are strictly monotonous.
-
-% Show also that, for any $n$ and $m$, $A\,n\,m \leq A\,(n+1)\,m$. 
-%          Finally, show that for any $n$, $m$ and $k$, 
-% $A\,k\,(A\,m\,n)\leq A\,(2+\textrm{max}(k,m))\,n$.
-% \emph{\textbf{Done.}}
-% \item Show that for any primitive recursive function $f$ of $k$ arguments,
-% for any $x_1,\dots\,x_k$, $f(x_1,\dots,x_k)\leq A\,n\,(x_1+\dots+x_k)$. \emph{\textbf{Still to do.}}
-% \item Show that the Ackermann function is not primitive recursive. \emph{\textbf{Still to do.}}
-
-% \end{enumerate}
-
 \begin{remark}

### 2021-03-07 doc/chapter-primrec.tex
minor corrections
simplification of a proof

@@
 
+Since The Ackermann function is strictly monotonous in its first argument\footnote{almost always\dots}, we may replace the large inequality with a strict one.
+
+\vspace{4pt}
+\noindent
+\emph{From \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack}.}
+
+\begin{Coqsrc}
+Lemma Ack_strict_mono_l : forall n m p, n < m ->
+                                        Ack n (S p) < Ack m (S p).
+\end{Coqsrc}
+
+\vspace{4pt}
+\noindent
+\emph{From \href{../theories/html/hydras.MoreAck.AcknotPR.html}{MoreAck.AckNotPR}.}
+
+\begin{Coqsrc}
+Lemma majorPR2_strict (f: naryFunc 2)(Hf : isPR 2 f):
+    exists (n:nat),
+    forall x y, 2 <= x -> 2 <= y -> f x y < Ack n (max x  y).
+\end{Coqsrc}
+
 Please note that, if the Ackermann function were primitive recursive, then there would exist some natural number $n$, such that, for all $x$ and $y$, the inequality 
@@
 
-  Fact Ack_majorized : majorized (Ack: naryFunc 2)  Ack.
- \end{Coqsrc}
-
- In order to share the variable \texttt{n}, let us open a section.
-
-\begin{Coqsrc}
-  Section Contrad.
-    Variable n: nat.
-    Hypothesis  Hn: forall x y,  Ack x y <= Ack n (max x  y).
-\end{Coqsrc}
-
-Firs, we get rid of the cases $n=0$ and $n=1$, thus we can rewrite
-(\texttt{max 2 n}) into \texttt{n}.
-
-  \begin{Coqsrc}
-    Remark R01: n <> 0.
-
-    Remark R02: n <> 1.
-
-    Remark R03: max 2 n = n.
-\end{Coqsrc}
-
-Please keep in mind that \texttt{(2 * n + 3)} can be safely replaced with 
-\texttt{(Ack 2 n)}.
-
-\begin{Coqsrc}
-    Remark R04: Ack n (2 * n + 3) = Ack n (Ack 2 n).
-
-    Remark R05: Ack n (2 * n + 3) <= Ack (n + 2) n.
-\end{Coqsrc}
-
- By hypothesis \texttt{Hn}, the following inequality holds.
-
- \begin{Coqsrc}
-    Remark R06: Ack (n + 2) n <= Ack n (n + 2).
- \end{Coqsrc}
-
-By monotony of the Ackerman function, we get:
-
-\begin{Coqsrc}
-    Remark R07: Ack n (n + 2) < Ack n (2 * n + 3).
-\end{Coqsrc}
-
-which contradicts \texttt{R05}.
-
-\begin{Coqsrc}
-    Lemma contrad : False.
-
-  End Contrad.
-
   Lemma Ack_not_PR : False.
+  Proof.
+    destruct (majorPR2_strict Ack HAck) as [m Hm];
+    pose (X := max 2 m); specialize (Hm X X).
+    rewrite max_idempotent in Hm; 
+    assert (Ack m X <= Ack X X) by (apply Ack_mono_l; lia).
+    lia.
+  Qed.
 

### 2021-03-08 doc/chapter-primrec.tex
correction of typos

@@
 
+\noindent
 \emph{From \href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR}}.
@@
 an uncurried version of $f$.
-
 In Library\href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR}, we introduce a lighter notation.
@@
 
-Since The Ackermann function is strictly monotonous in its first argument\footnote{almost always\dots}, we may replace the large inequality with a strict one.
-
-\vspace{4pt}
-\noindent
-\emph{From \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack}.}
+We prove also a strict version of this lemma, thanks to the following property (proved in Library\href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack}~).
 
@@
 
-Please note that, if the Ackermann function were primitive recursive, then there would exist some natural number $n$, such that, for all $x$ and $y$, the inequality 
+If the Ackermann function were primitive recursive, then there would exist some natural number $n$, such that, for all $x$ and $y$, the inequality 
 $\texttt{Ack}\,x\,y\leq \texttt{Ack}\,n\,(\texttt{max}\,x\,y)$ holds.

### 2021-03-08 doc/part-hydras.tex
renaming of a lemma
correction of typos

@@
 
-Hydras quite frequently contain  multiple  copies of the same pattern. The following functions
+Hydras quite frequently contain  multiple adjacent  copies of the same subtree. The following functions
 will help us to describe and reason about replications in hydra battles.
@@
 where, for every $i<l$, $\gamma_i \not= 0$ \footnote{This condition allows us to ignore paths which end by a lot of useless $0$s.} and there exists some strictly positive integer $s_i$
-such that $\gamma_{i+1}=\canonseq{\gamma}{s_i}$,
+such that $\gamma_{i+1}=\canonseq{\gamma}{s_i}$.
 
@@
 existence of such a path with the notation $\alpha\xrightarrow [s]{}\beta$.
+
+We say also that the considered path from $\alpha$ to $\beta$ \emph{starts at [index] $s_0$ and ends at $s_l$}.
 
@@
   Ketonen and Solovay~\cite{KS81} consider  large finite \emph{sets} of natural numbers,  but they are mainly used as sequences. Thus, we chosed to represent them explicitely as (sorted) lists. 
-
+\end{remark}
+
+The following function ``gnaws'' an ordinal $\alpha$, folowing a sequence of indices (ignoring the $0$s).
 \noindent
@@
 
+Minimal large sequences can be directly defined in terms of the
+predicate \texttt{path\_to} (\vref{sect:path-to-def}) which already prohibits paths containing non-final \texttt{zero}s.
+
 \vspace{4pt}
@@
 \end{Coqsrc}
-
-\end{remark}
 
@@
 
-\section{The Length of Minimal Large Sequences}
+\section{Length of Minimal Large Sequences}
 
@@
 
-%\subsection{Looking for a function definition}
-
-Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider the functions which associate to each stritly positive integer $k$ the number $l$, where 
+\subsection{Formal Specification}
+
+Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider any  function which associates maps  any strictly positive integer $k$ to the number $l$, where 
 the interval $[k,l)$ is $\alpha$-mlarge.
@@
 \end{remark}
-
-\subsection{Formal Specification}
 
@@
 Definition fun_le f g  := forall n:nat,  f n <=  g n.
+Infix "<<=" := fun_le (at level 60).
 \end{Coqsrc}
@@
 
-Let $\lambda<\epsilon_0$ be any limit ordinal. In a similar way as for successors, we decompose any path from $\lambda$ (at $k$) into a step to
-$\canonseq{\lambda}{k}$, then to $0$. In the following section, we assume that there exists à correct function for $\canonseq{\lambda}{k}$, \emph{for any strictly positive $k$}.
+Let $\lambda<\epsilon_0$ be any limit ordinal. In a similar way as for successors, we decompose any path from $\lambda$  into a first step to
+$\canonseq{\lambda}{k}$, followed by a path to $0$. In the following section, we assume that there exists à correct function for $\canonseq{\lambda}{k}$, \emph{for any strictly positive $k$}.
 
@@
 Applying the previous lemmas on successors and limit ordinals, 
-we get several correct implementations of \texttt{(L\_spec $\alpha$)} for small values of $\alpha$.
+we obtain a few  correct implementations of \texttt{(L\_spec $\alpha$)} for small values of $\alpha$.
 
@@
 
-As a corollary, we prove the following proposition, p. 284 of~\cite{KS81}.
+As a corollary, we prove the following proposition, p. 284 of~\cite{KS81}, which states that the function \texttt{F\_} is ``monotonous'' w.r.t. its first argument and ``domination''.
 
@@
 \begin{Coqsrc}
-Lemma Propp284 : forall alpha beta : E0, 
+Lemma F_mono_l: forall alpha beta : E0, 
    beta o< alpha -> dominates (F_ alpha) (F_ beta).

### 2021-03-09 doc/part-hydras.tex
minor improvements

@@
 Note that we did not define any function $L_\alpha$ \emph{for any $\alpha<\epsilon_0$} yet.
-We have got no more than a collection of proved realizations of $\texttt{L\_spec}\;\alpha$ for several values of $\alpha$.
+We have got no more than a collection of proved realizations of $\texttt{L\_spec}\;\alpha$ for a few values of $\alpha$.
 
@@
 $\texttt{L\_spec}\;\alpha$.   
-\subsection{Definition}
-
-In order to get a total function, we use our type \texttt{E0} of well-formed ordinal terms,(see Sect~\vref{sect:E0-def}).  Our definition is structured along a well-founded recursion and
-a case-study (zero, limit and successor ordinals).
+To this end, we represent ordinals as inhabitants of the type 
+\texttt{E0} of well-formed ordinal terms (see Sect~\vref{sect:E0-def}). So, we define a total function \texttt{L\_} of type
+\texttt{E0 -> nat -> nat}, by transfinite recursion, considering the usual three cases : $\alpha=0$, $\alpha$ is a successor, $\alpha$ is a limit ordinal.
+
+\subsubsection{Definition}
 
@@
 \label{sect:hardy}
-In order to give an idea of the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the so called \emph{Wainer-Hardy hierarchy} of fast growing functions,
+In order to give a feeling on  the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the so called \emph{Wainer-Hardy hierarchy} of fast growing functions,
 presented for instance in~\cite{Promel2013}. 
@@
 
-\subsection{Abstract Properties of H-functions}
+\subsection{Abstract Properties of Hardy functions}
 ~\label{sect:H-alpha-prop}
@@
 On the other hand, please show that for any natural number $n$, the function $F_n$ is primitive recursive.
+Thus $F\_alpha$ is primitive recursive if and only if $\alpha$ is finite.
 

### 2021-03-12 doc/chapter-powers.tex
consistent use of capitalization in headers

@@
 
-\subsection{Computing Fibonacci Numbers}
+\subsection{Computing Fibonacci numbers}
 
@@
 
-\section{Representing Monoids in \coq \label{monoid-class-def}}
+\section{Representing monoids in \coq \label{monoid-class-def}}
 
@@
 
-\subsubsection{Multiplication on Peano Numbers}
+\subsubsection{Multiplication on Peano numbers}
 
@@
 
-\subsubsection{String Concatenation}
+\subsubsection{String concatenation}
 We can use the notation \texttt{"\_ * \_ "} for other types than numbers.
@@
 
-\subsubsection{Solving Ambiguities}
+\subsubsection{Solving ambiguities}
 Let $A$ be some type, and let us assume there are several instances of
@@
 
-\subsection{The Monoid Type Class}
+\subsection{The Monoid type class}
 \index{coq}{Type classes}
@@
 
-\subsection{Building Instances of \texttt{Monoid}}
+\subsection{Building instances of \texttt{Monoid}}
 Let \texttt{$A$} be some type, \texttt{{\it op}} an instance of 
@@
 
-\subsection{Monoids and Equivalence Relations}
+\subsection{Monoids and equivalence relations}
 \index{coq}{Generalized rewriting}
@@
 
-\section{Computing Powers in any EMonoid}
+\section{Computing powers in any EMonoid}
 
@@
 
-\subsection{The naive (linear) Algorithm}
+\subsection{The naive (linear) algorithm}
 The new version of the linear exponentiation function is as follows:
@@
 
-\subsection{The Binary Exponentiation Algorithm}
+\subsection{The binary exponentiation algorithm}
 
@@
 
-\subsection{Refinement and Correctness}
+\subsection{Refinement and correctness}
 We have got two functions for computing powers in any monoid. 
@@
 
-\section{Comparing Exponentiation Algorithms with Respect to Efficiency}
+\section{Comparing exponentiation algorithms with respect to efficiency}
 
@@
 \end{Coqsrc}
-\subsubsection{Monadic Notation}
+\subsubsection{Monadic notation}
 
@@
 
-\subsubsection{Chain Correctness and Optimality}
+\subsubsection{Chain correctness and optimality}
 
@@
 
-\section{Proving a Chain's Correctness}
+\section{Proving a chain's correctness}
 \label{chain-correctness-sect}
@@
 
-\subsection{Correctness Proofs by Reflection}
+\subsection{Correctness proofs by reflection}
 \label{reflection-section}
@@
 
-\section{Certified Chain Generators}
+\section{Certified chain generators}
 \label{chain-generation}
@@
 
-\subsection{Chains and Continuations : f-chains}
+\subsection{Chains and continuations : f-chains}
 

### 2021-03-12 doc/chapter-primrec.tex
consistent use of capitalization in headers

@@
 
-\subsection{A little bit of Semantics} 
+\subsection{A little bit of semantics} 
 Please note that inhabitants of type (\texttt{PrimRec $n$}) are not \coq{} functions like \texttt{Nat.mul}, or factorial, etc.
@@
 
-\subsubsection{Elementary Proofs of \texttt{isPR} statements}
+\subsubsection{Elementary proofs of \texttt{isPR} statements}
 
@@
 
-\subsubsection{Using Function Composition}
+\subsubsection{Using function composition}
 

### 2021-03-12 doc/hydras.tex
renaming of a lemma
consistent use of capitalization in headers

@@
 Bordeaux, CNRS UMR 5800, \thanks{With contributions by Yves Bertot, \'Evelyne Contejean,  Florian Hatat and Pascal Manoury.
-\emph{The formalization of primitive recursive functions was originally authored by
-Russel O'Connor\cite{OConnor05}.}}}
+The formalization of primitive recursive functions was originally authored by
+Russel O'Connor\cite{OConnor05}.}}
 \date{\today}
@@
 
+Let us underline the analogy between hydra battles and interactive theorem proving.
+Hercules is the user (you!), and hydra's heads are the sub-goals: you may think that applying a tactic would solve a sub-goal, but it results often in the multiplication of such tasks.
+
 \item In the second part, we are interested in computing $x^n$ with the least number of multiplications as possible.
@@
 
-\subsection{Trust in our Proofs}
+\subsection{Trust in our proofs}
 \label{sect:trust-in-proofs}
@@
 
-\subsection{Assumed Redundancy}
+\subsection{Assumed redundancy}
 
@@
 interest (by illustrating some tactic of proof pattern, for instance).
-We may use \coq's module system to make several proofs of a given theorem co-exist in our libraries (see also Sect~\vref{sect:alt-proofs}).
+We use \coq's module system to make several proofs of a given theorem co-exist in our libraries (see also Sect~\vref{sect:alt-proofs}).
 After some discussions of the pros and cons of each solution, we develop only one of them, leaving the others  as exercises or projects (i.e., big or difficult exercises).
@@
 several aborted proofs.
-Of course, do not hesitate to contribute nice proofs or alternative definitions !
+Of course, do not hesitate to contribute nice proofs or alternative definitions!
 
@@
 \item There is no measure into $\mathbb{N}$ for proving the termination of all hydra battles (Sect~\vref{omega-case}).
-\item There is no measure into $[0,\omega^2)$ for proving the termination of all hydra battle (Sect~\vref{omega2-case}).
+\item There is no measure into the interval\footnote{We use the notation $[a,b)$ for denoting the set of ordinals greater or equal than $a$ and strictly less than $b$.} interval $\{x\$[0,\omega^2)$ for proving the termination of all hydra battles (Sect~\vref{omega2-case}).
 \item There is no measure into $[0,\mu)$ for proving the termination of all hydra battles, for any $\mu<\epsilon_0$ (Sect\vref{sec:free-battles-case}).
@@
 
-\subsection{About Logic}
+\subsection{About logic}
 
@@
 
-\subsection{Alternative or Bad Definitions}
+\subsection{Alternative or bad definitions}
 \label{sect:alt-proofs}
@@
 
-\section{How to Install the Libraries}
+\section{How to install the libraries}
 \label{sec:orgheadline4}
@@
 
-\item Just go into the root directory, and type "\texttt{make}" ("\texttt{make -k}"  in case of an incomplete installation).
+\item Please refer to \href{https://github.com/coq-community/hydra-battles#readme}{the README file of the project}
 \end{itemize}
 
-\section{Comments on Exercises and Projects}
+\section{Comments on exercises and projects}
 
@@
 \include{chapter-primrec}
-\part{Some  Certified Algorithms}
+\part{A Few  Certified Algorithms}
 

### 2021-03-12 doc/part-hydras.tex
consistent use of capitalization in headers

@@
 
-\subsection*{State of the Development}
+\subsection*{State of the development}
 The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on  notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
@@
 
-\section*{Future Work (projects)}
+\section*{Future work (projects)}
 \index{hydras}{Projects}
@@
 
-\subsection*{Main References}
+\subsection*{Main references}
 
@@
 
-\subsection{The Rules of the Game}
+\subsection{The rules of the game}
 
@@
 
-\section{Hydras and their Representation in \emph{Coq}}
+\section{Hydras and their representation in \emph{Coq}}
 \label{sec:orgheadline48}
@@
 
-\subsubsection{Recursive Functions on type \texttt{Hydra}}
+\subsubsection{Recursive functions on type \texttt{Hydra}}
 \label{sec:orgheadline41}
@@
 
-\subsection{Induction Principles for Hydras}
+\subsection{Induction principles for hydras}
 \label{sec:orgheadline42}
@@
 
-\subsubsection{A failed Attempt}
+\subsubsection{A failed attempt}
 
@@
 
-\subsubsection{A Principle of Mutual Induction}
+\subsubsection{A Principle of mutual induction}
 In order to get an appropriate induction scheme for the types 
@@
 
-\subsubsection{A Correct Proof}
+\subsubsection{A Correct proof}
 
@@
 
-\section{Relational Description of Hydra Battles}
+\section{Relational description of hydra battles}
 
@@
 
-\subsection{Chopping off a Head at Distance 1 from the Foot (Relation  R1)}
+\subsection{Chopping off a head at distance 1 from the foot (relation  R1)}
 
@@
 
-\subsection{Chopping off a Head at Distance \texorpdfstring{$\geq 2$}{>= 2} from the Foot (relation R2) }
+\subsection{Chopping off a head at distance \texorpdfstring{$\geq 2$}{>= 2} from the foot (relation R2) }
 
@@
 
-\subsection{Relation Associated with a Round}
+\subsection{Binary relation associated with a round}
 
@@
 
-\subsection{Rounds and Battles}
+\subsection{Rounds and battles}
 
@@
 
-\subsection{Classes of Battles}
+\subsection{Classes of battles}
 \label{sect:battle-classes}
@@
 
-\subsection{Big Steps}
+\subsection{Big steps}
 
@@
 
-\section{A Long Battle}
+\section{A Long battle}
 \label{sect:big-battle}
@@
 
-\subsection{The Beginning of Hostilities}
+\subsection{The beginning of hostilities}
 During the two first rounds, our hydra loses its two rightmost heads.  Figure~\vref{fig:hinit-plus2} shows the state of the hydra   just before the third round.
@@
 
-\subsection{Looking for Regularities}
+\subsection{Looking for regularities}
 
@@
 
-\subsection{Giant Steps}
+\subsection{Giant steps}
 
@@
 
-\subsection{A Minoration Lemma}
+\subsection{A minoration lemma}
 
@@
 
-\section{Generic Properties}
+\section{Generic properties}
 
@@
 
-\subsection{A  Small Proof of Impossibility}
+\subsection{A  small proof of impossibility}
 \index{coq}{Proofs of impossibility}
@@
 
-\section{The Mathematical Point of View}
-
-\subsection{Well-ordered Sets}
+\section{The mathematical point of view}
+
+\subsection{Well-ordered sets}
 Let us start with some definitions.
@@
 
-\subsection{Ordinal Numbers}
+\subsection{Ordinal numbers}
 
@@
 
-\section{Ordinal Numbers in Coq}
+\section{Ordinal numbers in Coq}
 
@@
 
-% \section{Countable Ordinals}
+% \section{Countable ordinals}
 
@@
 
-\subsection{A Class for Ordinal Notations}
+\subsection{A class for ordinal notations}
 
@@
 
-\subsection{Ordinal Notations and Measures for Proving Termination}
+\subsection{Ordinal notations and  termination measures}
 \label{sect:measure-ON}
@@
 
-\section{Example: the Ordinal \texorpdfstring{$\omega$}{omega}}
+\section{Example: the ordinal \texorpdfstring{$\omega$}{omega}}
 
@@
 
-\section{Sum of  two Ordinal Notations}
+\section{Sum of  two ordinal notations}
 
@@
 
-\subsection{The Ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
+\subsection{The ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
 
@@
 
-\section{Limits and Successors}
+\section{Limits and successors}
 
@@
 
-\subsection{Limits and Successors in \texorpdfstring{$\omega+\omega$}{omega+omega}}
+\subsection{Limits and successors in \texorpdfstring{$\omega+\omega$}{omega+omega}}
 
@@
 
-\section{Product of Ordinal Notations}
+\section{Product of ordinal notations}
 
@@
 
-\section{The Ordinal \texorpdfstring{$\omega^2$}{omega2}}
+\section{The ordinal \texorpdfstring{$\omega^2$}{omega2}}
 
@@
 
-\subsection{A Proof of Termination using \texorpdfstring{$\omega^2$}{omega^2}} 
+\subsection{A proof of termination using \texorpdfstring{$\omega^2$}{omega^2}} 
 \label{sect:merge-example}.
@@
 
-\subsection{Yet Another  Proof of Impossibility}
+\subsection{Yet another  proof of impossibility}
 \label{omega2-case}
@@
 
-\subsubsection{Proof of the Inequality \texttt{m small\_h o< m big\_h}}
+\subsubsection{Proof of the inequality \texttt{m small\_h o< m big\_h}}
 
@@
 
-\subsubsection{Proof of the Inequality \texttt{m big\_h o<= m small\_h} }
+\subsubsection{Proof of the inequality \texttt{m big\_h o<= m small\_h} }
 
@@
 
-\subsubsection{End of the Proof}
+\subsubsection{End of the proof}
 From \texttt{m\_ge}, we get \texttt{m big\_h o<= m small\_h = m (iota (m big\_h)) }. 
@@
 
-\section{A Notation for Finite Ordinals}
+\section{A notation for finite ordinals}
 
@@
 
-\section{Comparing two Ordinal Notations}
+\section{Comparing two ordinal notations}
 
@@
 
-\section{Comparing an Ordinal Notation with Schütte's Model}
+\section{Comparing an ordinal notation with Schütte's model}
 
 Finally, it may be interesting to compare an ordinal notation with the more theoretical model from Schütte (well, at least with our formalization of that model).
-This would be a relative proof of correctenss of the considered  ordinal  notation.
+This would be a relative proof of correctness of the considered  ordinal  notation.
 
@@
 
-\section{Isomorphism of Ordinal Notations}
+\section{Isomorphism of ordinal notations}
 
@@
 
-\section{Other Ordinal Notations}
+\section{Other ordinal notations}
 
@@
 
-\section{The Ordinal \texorpdfstring{\(\epsilon_0\)}{epsilon0}}
+\section{The ordinal \texorpdfstring{\(\epsilon_0\)}{epsilon0}}
 \label{sec:epsilon0-intro}
 
-\subsection{Cantor Normal Form}
+\subsection{Cantor normal form}
 \index{maths}{Cantor normal form}
@@
 
-\subsection{A Data Type for  Ordinals in Cantor Normal Form}
+\subsection{A data type for  ordinals in Cantor normal form}
 \label{sec:orgheadline72}
@@
 
-\subsubsection{Finite Ordinals}
+\subsubsection{Finite ordinals}
 \label{sec:orgheadline67}
@@
 
-\subsubsection{The Ordinal \(\omega\)}
+\subsubsection{The ordinal \(\omega\)}
 \label{sec:orgheadline68}
@@
 
-\subsubsection{The Ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
+\subsubsection{The ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
 \label{sect:notation-phi0}
@@
 
-\subsubsection{The Hierarchy of \(\omega\)-towers:}
+\subsubsection{The hierarchy of \(\omega\)-towers:}
 \label{sec:orgheadline71}
@@
 
-\subsection{Pretty Printing Ordinals in Cantor Normal Form}
+\subsection{Pretty-printing ordinals in Cantor normal form}
 \label{sect:ppT1}
@@
 
-\subsection{Comparison between Ordinal Terms}
+\subsection{Comparison between ordinal terms}
 \label{sec:orgheadline73}
@@
 
-\subsection{Making Normality Implicit}
+\subsection{Making normality implicit}
   We would like to get rid of terms of type \texttt{T1} which are not in Cantor normal form.
@@
 
-\subsubsection{A Sigma Type for \texorpdfstring{$\epsilon_0$}{epsilon0}}
+\subsubsection{A sigma-type for \texorpdfstring{$\epsilon_0$}{epsilon0}}
 
@@
 
-\subsection{Syntactic Definition of Limit and Successor Ordinals}
+\subsection{Syntactic definition of limit and successor ordinals}
 
@@
 
-\subsubsection{Addition and Multiplication}
+\subsubsection{Addition and multiplication}
 
@@
 
-\section{Well-foundedness and Transfinite Induction}
+\section{Well-foundedness and transfinite induction}
 
@@
 
-\subsection{About  Well-foundedness}
+\subsection{About  well-foundedness}
 \label{sec:orgheadline82}
@@
 
-\subsubsection{A First Attempt}
+\subsubsection{A first attempt}
 \label{sec:orgheadline77}
@@
 
-\subsubsection{Using a Stronger Inductive Predicate.}
+\subsubsection{Using a stronger inductive predicate.}
 \label{sec:orgheadline78}
@@
 
-\subsection{An Ordinal Notation for  \texorpdfstring{$\epsilon_0$}{epsilon0}}
+\subsection{An ordinal notation for  \texorpdfstring{$\epsilon_0$}{epsilon0}}
 
@@
 
-\section{A Variant for Hydra Battles}
+\section{A Variant for hydra battles}
 
@@
 
-\subsection{Natural Sum (a.k.a. Hessenberg's  Sum)}
+\subsection{Natural sum (a.k.a. Hessenberg's  sum)}
 \label{sec:orgheadline87}
@@
 
-\subsubsection{Rewriting Lemmas}
+\subsubsection{Rewriting lemmas}
 \label{sec:orgheadline86}
@@
 \end{project}
-\subsection{More Theorems on Hessenberg's Sum}
+\subsection{More theorems on Hessenberg's sum}
 
@@
 
-\subsubsection{Commutativity, Associativity}
+\subsubsection{Commutativity, associativity}
 
@@
 
-\subsection{A Measure for Hydra-battle Termination}
+\subsection{A termination measure for hydra battles }
 
@@
 
-\subsubsection{Mathematical Definition of Canonical Sequences}
+\subsubsection{Mathematical definition of canonical sequences}
 
@@
 
-\subsubsection{Canonical Sequences in Coq}
+\subsubsection{Canonical sequences in Coq}
 \index{hydras}{Library Epsilon0!Functions!canon}
@@
 
-\subsection{Basic Properties of Canonical Sequences}
+\subsection{Basic properties of canonical sequences}
 
@@
 
-\subsubsection{Canonical Sequences and the Order $<$}
+\subsubsection{Canonical sequences and the order $<$}
 
@@
 
-\subsubsection{Limit Ordinals are truly Limits}
+\subsubsection{Limit ordinals are truly limits}
 The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
@@
 
-\subsection{Formal Definition}
+\subsection{Formal definition}
 
@@
 
-\subsection{Existence of a Path}
+\subsection{Existence of a path}
 
@@
 
-\subsection{Paths and Hydra Battles}
+\subsection{Paths and hydra battles}
 \label{KS-o2h}
@@
 
-\section{A  Proof of Impossibility}
+\section{A  proof of impossibility}
 
@@
 
-\subsection{The case of Free Battles}
+\subsection{The case of free battles}
 \label{sec:free-battles-case}
@@
 
-\section{The Case of Standard Battles}
+\section{The case of standard battles}
 \label{sec:standard-intro}\label{std-case}
@@
 
-\subsection{Paths with a Constant Index}
+\subsection{Paths with a constant index}
 
@@
 
-\subsubsection{Sketch of Proof of \texttt{Cor12}}
+\subsubsection{Sketch of proof of \texttt{Cor12}}
 \index{maths}{Transfinite induction}
@@
 
-\subsection{Casting Paths with a Constant Index into a Standard Path}
+\subsection{Casting paths with a constant index into a standard path}
 
@@
 
-\subsubsection{Sketch of Proof of \texttt{constant\_to\_standard\_path}}
+\subsubsection{Sketch of proof of \texttt{constant\_to\_standard\_path}}
 
@@
 
-\subsubsection*{Préliminaries}
+\subsubsection*{Preliminaries}
 
@@
 
-\subsection{Back to Hydras}
+\subsection{Back to hydras}
 \label{sec:standard-battles-cases}
@@
 
-\section{Length of Minimal Large Sequences}
+\section{Length of minimal large sequences}
 
@@
 
-\subsection{Formal Specification}
+\subsection{Formal specification}
 
@@
 
-\subsection{Abstract Properties}
+\subsection{Abstract properties}
 
@@
 
-\subsubsection{The  Ordinal Zero}
+\subsubsection{The  ordinal zero}
 \label{sect:L-spec-zero}
@@
 
-\subsubsection{Successor Ordinals}
+\subsubsection{Successor ordinals}
 \label{sect:L-spec-succ}
@@
 
-\subsubsection{Limit Ordinals}
+\subsubsection{Limit ordinals}
 \label{sect:L-spec-lim}
@@
 
-\subsection{First Results}
+\subsection{First results}
 
@@
 
-\subsubsection{Finite Ordinals}
+\subsubsection{Finite ordinals}
 
@@
 
-\subsubsection{The First Limit Ordinal  \texorpdfstring{$\omega$}{omega}}
+\subsubsection{The first limit ordinal  \texorpdfstring{$\omega$}{omega}}
 
@@
 
-\subsubsection{Going Further}
+\subsubsection{Going further}
 Let us consider a last example, ``computing'' $L_{\omega^3}$.
@@
 
-\subsubsection{Back to Hydra Battles}
+\subsubsection{Back to hydra battles}
 
@@
 
-\section{The Wainer-Hardy Hierarchy (Functions \texorpdfstring{$H_\alpha$}{\texttt{H\_alpha}})}
+\section{The Wainer-Hardy hierarchy (functions \texorpdfstring{$H_\alpha$}{\texttt{H\_alpha}})}
 
@@
 
-\subsection{Hardy Functions in \texttt{Coq}}
+\subsection{Hardy functions in \texttt{Coq}}
 
@@
 
-\subsection{First  Steps of the Hardy hierarchy}
+\subsection{First  steps of the Hardy hierarchy}
 Using rewrite rules from \texttt{H\_eq1} to \texttt{H\_eq4}, we can explore the functions $H_\alpha$ for some small values of $\alpha$.
 
-\subsubsection{Finite Ordinals}
+\subsubsection{Finite ordinals}
 
@@
 
-\subsubsection{New Limits}
+\subsubsection{New limits}
 
@@
 
-\subsubsection{A Numerical Example}
+\subsubsection{A numerical example}
 
@@
 
-\subsection{Abstract Properties of Hardy functions}
+\subsection{Abstract properties of Hardy functions}
 ~\label{sect:H-alpha-prop}
@@
 
-\subsubsection{Back to Hydras}
+\subsubsection{Back to hydras}
 
@@
 
-\section{The Wainer Hierarchy (Functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
+\section{The wainer hierarchy (functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
 \label{sect:wainer}
@@
 
-\section{Declarations and Axioms}
+\section{Declarations and axioms}
 
@@
 
-\section{Additional  Axioms}
+\section{Additional  axioms}
 
@@
 
-\subsubsection{Classical Logic}
+\subsubsection{Classical logic}
 
@@
 
-\subsubsection{Description Operators}
+\subsubsection{Description operators}
 
@@
 
-\subsubsection{The Definition of zero}
+\subsubsection{The definition of zero}
 
@@
 
-\section{The  Successor Function}
+\section{The  successor function}
 
@@
 
-\section{Finite Ordinals}
+\section{Finite ordinals}
 
@@
 
-\section{The Definition of \texttt{omega}}
+\section{The definition of \texttt{omega}}
 In order to define $\omega$, the first infinite ordinal, we use an operator which
@@
 
-\subsection{Ordering Functions and Ordinal Addition}
+\subsection{Ordering functions and ordinal addition}
 
@@
 
-\subsection{Ordinal Addition}
+\subsection{Ordinal addition}
 
@@
 
-\subsubsection{Multiplication by a Natural Number}
+\subsubsection{Multiplication by a natural number}
 
@@
 
-\section{The Exponential of Basis \texorpdfstring{$\omega$}{omega}}
+\section{The exponential of basis \texorpdfstring{$\omega$}{omega}}
 
@@
 
-\subsection{Additive Principal Ordinals}
+\subsection{Additive principal ordinals}
 
@@
 
-\subsection{The Function \texttt{phi0}}
+\subsection{The function \texttt{phi0}}
 
@@
 
-\subsection{Omega-towers and the Ordinal \texorpdfstring{$\epsilon_0$}{epsilon0}}
+\subsection{Omega-towers and the ordinal \texorpdfstring{$\epsilon_0$}{epsilon0}}
 
@@
 
-\subsection{Properties of the Set  \texttt{AP}}
+\subsection{Properties of the set  \texttt{AP}}
 
@@
 
-\subsubsection{Properties of the Function \texorpdfstring{$\phi_0$}{phi0}}
+\subsubsection{Properties of the function \texorpdfstring{$\phi_0$}{phi0}}
 
@@
 
-\section{Critical Ordinals}
+\section{Critical ordinals}
 
@@
 
-\subsection{A flavor of Infinity}
+\subsection{A flavor of infinity}
 
@@
 
-\section{Cantor Normal Form}
+\section{Cantor normal form}
 
@@
 
-\section{An Embedding of \texttt{T1} into \texttt{Ord}}
+\section{An embedding of \texttt{T1} into \texttt{Ord}}
 
@@
 
-\section{Related Work}
+\section{Related work}
 
@@
 
-\section{The Type \texttt{T2} of Ordinal Terms}
+\section{The type \texttt{T2} of ordinal terms}
 
@@
 
-\section{How Big is \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}?}
+\section{How big is \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}?}
 
@@
 
-\section{Veblen Normal Forms}
+\section{Veblen normal form}
 \begin{definition}
@@
 
-\subsection{Length of a Term}
+\subsection{Length of a term}
 
@@
 
-\section{Main Functions on \texttt{T2}}
+\section{Main functions on \texttt{T2}}
 
@@
 
-\subsection{The Veblen Function \texorpdfstring{$\phi$}{\texttt{phi}}}
+\subsection{The Veblen function \texorpdfstring{$\phi$}{\texttt{phi}}}
 
@@
 
-\section{An Ordinal Notation for \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}}
+\section{An ordinal notation for \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}}
 

### 2021-03-12 doc/part-hydras.tex
consistent use of capitalization in headers

@@
 
-\chapter{Hydras and Hydra Games}
+\chapter{Hydras and hydra games}
 
@@
 
-\section{A Long battle}
+\section{A long battle}
 \label{sect:big-battle}
@@
 
-\chapter{Introduction to Ordinal Numbers and Ordinal Notations}
+\chapter{Introduction to ordinal numbers and ordinal notations}
 
@@
 
-\chapter[A proof of termination, using epsilon0]{A Proof of Termination, using Ordinals below \texorpdfstring{$\epsilon_0$}{Epsilon0}}
+\chapter[A proof of termination, using epsilon0]{A proof of termination, using ordinals below \texorpdfstring{$\epsilon_0$}{Epsilon0}}
 
@@
 %---------------------------------------------------------------------
-\chapter{Large Sets and Rapidly Growing Functions}\label{chap:alpha-large}
+\chapter{Large sets and rapidly growing functions}\label{chap:alpha-large}
 
@@
 %----------------------------------------------------------------------
-\chapter[Countable Ordinals (after Sch\"{u}tte)]{Kurt Schütte's Axiomatic Definition of Countable Ordinals}
+\chapter[Countable ordinals (after Sch\"{u}tte)]{Kurt Schütte's axiomatic definition of countable ordinals}
 

### 2021-03-18 doc/chapter-primrec.tex
consistent use of capitalization in headers
add variants of Ackermann function + a few typos

@@
-\chapter{Primitive Recursive Functions}
+\chapter{Primitive recursive functions}
 
@@
 is an interpretation function (\texttt{evalprimRec $n$})  of type
-$\texttt{PrimRec}\,n \rightarrow  \texttt{naryFunc}\,n$ , together with the helper function 
+$\texttt{PrimRec}\,n \rightarrow  \texttt{naryFunc}\,n$ , together with the  function 
 (\texttt{evalprimRecS $n$ $m$}) of type 
@@
 The following proof decomposes the \texttt{double} function as the composition of 
-multiplication,with the pair composed of the identity and the constant function wihich returns $2$.
-\emph{Note that the lemma \texttt{const1\_NIsPR} considers this function as an unary function (unlike \texttt{const0\_NIsPR} }.
+multiplication with the identity and the constant function wihich returns $2$.
+\emph{Note that the lemma \texttt{const1\_NIsPR} considers this function as an unary function (unlike \texttt{const0\_NIsPR})}.
 
@@
 \begin{exercise}
-Consider the following functions:
+Prove that the following functions are primitive recursive.
 
@@
 
-Prove they are primitive recursive.
-
 \textbf{Hint:} You may have to look again at the lemmas of the library
 \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec} if you meet some difficulty.
-
-\emph{You may start this exercise with the file
-\url{../exercises/primrec/MorePRExamples.v}.}
+You may start this exercise with the file
+\url{../exercises/primrec/MorePRExamples.v}.
 \end{exercise}
@@
 
+\index{hydras}{Exercises}
+
+\begin{exercise}
+The file \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack} presents two other definitions of the Ackermann functions based on the lexicographic ordering on $\mathbb{N}\times\mathbb{N}$.
+Prove that all the four functions are extensionnally equal.
+\end{exercise}
+
 This function is also known under the form of three equations, which we derive immediately from our definition.

### 2021-03-18 doc/part-hydras.tex
add variants of Ackermann function + a few typos

@@
 
-Simply put, this document tries to combines the scientific interest of two articles~\cite{KP82, KS81} and a book~\cite{schutte} with the playful activity of truly proving theorems.
+Simply put, this document tries to combine the scientific interest of two articles~\cite{KP82, KS81} and a book~\cite{schutte} with the playful activity of truly proving theorems.
 We hope  that such a work, besides exploring a nice piece of discrete maths, 
@@
  We also hope to provide a little clarification on infinity (both potential and actual) through the notions of function, computation, limit,
- types and proofs.
+ type and proof.
 
@@
 \subsection*{Difference from Kirby and Paris's Work}
-In~\cite{KP82}, Kirby and Paris showed  that there is no proof of termination of all hydra battles in Peano Arithmetic (PA).
+In~\cite{KP82}, Kirby and Paris show  that there is no proof of termination of all hydra battles in Peano Arithmetic (PA).
 Since we are used to writing proofs in higher order logic, the restriction to PA was quite unnatural for us.
-So we chosed to prove another statement without any reference to PA, by considering a class of proofs indexed by ordinal numbers upto $\epsilon_0$.
+So we chosed to prove another statement without any reference to PA, by considering a class of proofs indexed by ordinal numbers up to $\epsilon_0$.
 
@@
 Technically, a \emph{hydra} is just a finite ordered tree, each node of which 
-has any number of sons. Note that, contrary to the computer science tradition, we will show the hydras 
+has any number of sons. Contrary to the computer science tradition, we display the hydras 
 with the heads up and the foot (i.e., the root of the tree) down.
@@
 goal \Verb@{h':Hydra | h -*-> h'}@ generates the subgoal \Verb|h -*-> ?h'|.
-\item You may use Gérard Huet's \emph{zipper} data structure~\cite{zipper} for writing tactics associated with Hercule's  interactive search of a head to chop off.
+\item You may use Gérard Huet's \emph{zipper} data structure~\cite{zipper} for writing tactics associated with Hercule's  interactive search for a head to chop off.
 \end{itemize}
@@
 
-Class Battle :=  {battle_r : dep_round_t;
-                  battle_inclusion : forall i h h',
-                      battle_r i h h' -> round h h'}.
-
+Class Battle :=  
+    { battle_rel : dep_round_t;
+      battle_inclusion : forall i h h', battle_rel i h h' -> round h h'}.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Program Instance free : Battle :=
-  (Build_Battle ( fun _  h h' => round h h') _).
+Program Instance free : Battle := Build_Battle (fun _  h h' => round h h') _.
 \end{Coqsrc}
@@
 | battle_1 : forall i h  h', 
-      battle_r   B i  h h' -> battle B i h (S i) h'
+      battle_rel   B i  h h' -> battle B i h (S i) h'
 | battle_n : forall i h  j h' h'',  
-      battle_r  B i h h''  ->
+      battle_rel  B i h h''  ->
       battle B (S i) h'' j h'  ->
@@
   (m: Hydra -> A):   Prop :=
-  {variant_decr :forall i h h',
+  {variant_decr: forall i h h',
       h <> head ->
-      battle_r  B i  h h' -> Lt (m h') (m h)}.
+      battle_rel  B i  h h' -> Lt (m h') (m h)}.
 \end{Coqsrc}

### 2021-03-18 doc/chapter-primrec.tex
get around an issue with equations and modules

@@
 \begin{exercise}
-The file \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack} presents two other definitions of the Ackermann functions based on the lexicographic ordering on $\mathbb{N}\times\mathbb{N}$.
+The file \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack} presents two other definitions\footnote{post on \href{https://stackoverflow.com/questions/10292421/error-in-defining-ackermann-in-coq} by Anton Trunov.} of the Ackermann functions based on the lexicographic ordering on $\mathbb{N}\times\mathbb{N}$.
 Prove that all the four functions are extensionnally equal.

### 2021-03-25 doc/chapter-primrec.tex
minor corrections

@@
 are stored in the directory \texttt{theories/ordinals/MoreAck}.
+In particular, the library~\href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR} contains the well-known  proof that the Ackermann function is not primitive recursive (see Section~\vref{sect:ack-not-PR}).
 
@@
 
-Here is a possible definition, where (\texttt{iterate\,$f$\,$n$})
-is the $n$-th iteration of $f$\,\footnote{Please not confuse with \texttt{primRec.iterate}, which is monomorphic and does not share the same order of arguments.}
+Ackermann function is traditionally defined as a function from 
+$\mathbb{N}\times \mathbb{N}$ into $\mathbb{N}$, through
+three equations:
+
+\begin{align}
+A(0,n)&=n+1\\
+A(m+1,0)&=A(m,1)\\
+A(m+1,n+1)&=A(m,A(m+1,n))
+\end{align}
+
+Let us try to define this function in \coq{} (in curried form).
+
+\begin{Coqsrc}
+Fail
+  Fixpoint Ack (m n : nat) : nat :=
+  match m, n with
+  | 0, n => S n
+  | S m, 0 => Ack m 1
+  | S m0, S p => Ack m0 (Ack m p)
+  end.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+The command has indeed failed with message:
+Cannot guess decreasing argument of fix.
+\end{Coqanswer}
+
+A possible workaround is to make \\texttt{m} be the 
+decreasing argument, and define --- within \texttt{m}'s scope --- a local helper function which computes \texttt{Ack m n} for any \texttt{n}.
+This way, both functions \texttt{Ack} and \texttt{Ackm} have a (structurally) strictly decreasing argument.
+
+\begin{Coqsrc}
+Module Alt.
+
+   Fixpoint Ack (m n : nat) : nat :=
+     match m with
+     | O => S n
+     | S p => let fix Ackm (n : nat) :=
+                  match n with
+                  | O => Ack p 1
+                  | S q => Ack p (Ackm q)
+                  end
+              in Ackm n
+     end.
+
+    Compute Ack 3 2.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+  = 29 : nat
+\end{Coqanswer}
+
+\begin{Coqsrc}
+End Alt.
+\end{Coqsrc}
+
+We prefered to define a variant which uses explicitely 
+ the functional \texttt{iterate},
+where (\texttt{iterate\,$f$\,$n$})
+is the $n$-th iteration of $f$\,\footnote{Please not confuse with \texttt{primRec.iterate}, which is monomorphic and does not share the same order of arguments.}.
+It makes it possible to apply a few lemmas proved in
+\href{../theories/html/hydras.Prelude.Iterates.html}{Prelude.Iterates}, for instance about the monotony of the $n$-th iterate of a given function.
 
@@
 
+\begin{Coqsrc}
+Lemma iterate_le_n_Sn:
+forall f : nat -> nat,
+(forall x : nat, x <= f x) ->
+forall n x : nat, iterate f n x <= iterate f (S n) x.
+\end{Coqsrc}
+
+Thus, our definition of the Ackermann function is as follows:
+
 \vspace{4pt}
@@
   end.
-\end{Coqsrc}
-
-Another way of defining this function is 
-using an inner \texttt{fix} (see for instance the discussion in \url{https://stackoverflow.com/questions/10292421/error-in-defining-ackermann-in-coq}).
-
-We prefered to use the former definition, in order to
-take advantage from generic lemmas about \texttt{iterate}.
-
-\begin{Coqsrc}
-Module Alt.
-
-Fixpoint Ack (n m : nat) : nat :=
-  match n with
-  | O => S m
-  | S p => let fix Ackn (m : nat) :=
-               match m with
-               | O => Ack p 1
-               | S q => Ack p (Ackn q)
-               end
-           in Ackn m
-  end.
-
-End Alt.
-\end{Coqsrc}
+
+Compute Ack 3 2.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+  = 29 : nat
+\end{Coqanswer}
 
@@
 
-This function is also known under the form of three equations, which we derive immediately from our definition.
+\subsubsection{First properties of the Ackermann function}
+
+The three firts lemmas make us sure that our function 
+\texttt{Ack} satifies the ``usual'' equations.
 
@@
 \begin{Coqsrc}
-Lemma Ack_Ack_le : forall k m n, 
+Lemma nested_Ack_bound : forall k m n, 
     Ack k (Ack m n) <= Ack (2 + max k m) n.
@@
 First, we prove that any primitive function of two arguments is majorized.
-If we look at the inductive definition of primitive recursive functions,\pageref{def:Primrec}, it is obvious that a proof by induction on the construction of primitive recursive functions must consider functions of any arity.
+If we look at the inductive definition of primitive recursive functions, page~\pageref{def:Primrec}, it is obvious that a proof by induction on the construction of primitive recursive functions must consider functions of any arity.
 
@@
 
-For instance, proving a property shared by any primitive recursive functions of arity 2 leads to consider the case where that function is obtained by composition with a function of any 
+For instance, proving a property shared by any primitive recursive function of arity 2 leads to consider the case where that function is obtained by composition with a function of any 
 arity $m$. The same problem happens with primitive 
@@
 Likewise, the notation \texttt{x::v} is an abbreviation for
-\texttt{VectorDef.cons x ?n v}.
+\texttt{VectorDef.cons x \_ v}.
 

### 2021-03-27 doc/chapter-primrec.tex
correction to some coqdoc comments

@@
 
-Thus, the function \texttt{plus} has also type (\texttt{naryFunc 2}).
+\begin{Coqsrc}
+Check plus: naryFunc 2.
+
+Check 42: naryFunc 0.
+
+Check (fun n p q : nat =>  n * p + q): naryFunc 3.
+\end{Coqsrc}
 
@@
 \texttt{PrimRec} allows us to write trivial proofs of primitive recursivity. 
-Although  the following lemmas are alreday proven in 
+Although  the following lemmas are already proven in 
 \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec.v},
@@
 
-We alreday proved that \texttt{S} is \texttt{PR 1}, but we need to consider a function of three arguments, which ignores its first and third arguments.
+We already proved that \texttt{S} is \texttt{PR 1}, but we need to consider a function of three arguments, which ignores its first and third arguments.
 Fortunately, the library \texttt{primRec} already contains lemmas adapted to this kind of situation.
@@
 
-Thus, our first subgoal is solved easily and the proof ends, applying alreday proven lemmas.
+Thus, our first subgoal is solved easily and the proof ends, applying already proven lemmas.
 

### 2021-03-29 doc/chapter-primrec.tex
work on coqdoc comments

@@
 
-\section{First look at the Ackermann Library}
+\section{First look at the Ackermann library}
 
@@
 
-\section{Basic Definitions}
+\section{Basic definitions}
 \index{maths}{Primitive recursive functions}
@@
  This fact is reported in the comments of \texttt{primRec.v}.
-We presume that this convention makes it easier to define the evaluation function \texttt{evalProjFunc $n$} (see next sub-section) .
+We presume that this convention makes it easier to define the evaluation function \texttt{evalProjFunc $n$} (see the next sub-section).
 Trying the other convention is left as an exercise.
@@
   end
-
   with evalPrimRecs (n m : nat) (fs : PrimRecs n m) {struct fs} :
@@
 
+\vspace{4pt}
+\noindent
+\emph{From \href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}.}
 \begin{Coqsrc}
@@
   let u := composeFunc 2 3
-                       (PRcons 2 2 x
-                               (PRcons 2 1 y
-                                       (PRcons 2 0 z
+                       (PRcons 2 _ x
+                               (PRcons 2 _ y
+                                       (PRcons 2 _ z
                                                (PRnil 2))))

### 2021-04-13 doc/chapter-primrec.tex
small changes

@@
 \section{Introduction}
-The definition of primitive recursive functions we use comes
-from Russel O'Connors formalization in \coq{} of
-G\"odel's incompleteness theorems~\cite{OConnor05}, now hosted in
-the \texttt{theories/ordinals/Ackermann} directory.
-
-This chapter contains some comments on Russel's library, as well as a few extensions.
-Contributions (under the form of comments, new examples or exercises) are welcome!.
-
-\section{First look at the Ackermann library}
-
-O'Connor's library on Gödel's incompleteness theorems contains a little more 
-than 45K lines of scripts. The part dedicated to primitive recursive functions and Peano arithmetics is 32K lines long and is originally structured in 38 modules.
-Thus, we propose a slow exploration of this library, through examples and exercises. Our additions to the original library 
-are stored in the directory \texttt{theories/ordinals/MoreAck}.
-In particular, the library~\href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR} contains the well-known  proof that the Ackermann function is not primitive recursive (see Section~\vref{sect:ack-not-PR}).
-
-\section{Basic definitions}
-\index{maths}{Primitive recursive functions}
-
-The formal definition of primitive recursive functions lies in the library
-\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec},
-with preliminary definitions in 
-\href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat}
-and
-\href{../theories/html/hydras.Ackermann.misc.html}{Ackermann.misc}.
-
-\subsection{Functions of arbitrary arity}
-
-Library \texttt{primRec} allows us to consider primitive functions on \texttt{nat}, with any number of arguments, in 
-curried form. This is made possible in 
-\href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat} by the following definition:
-
-\index{primrec}{Types!naryFunc}
-\begin{Coqsrc}
-Fixpoint naryFunc (n : nat) : Set :=
-  match n with
-  | O => nat
-  | S n => nat -> naryFunc n
-  end.
-\end{Coqsrc}
-
-For instance (\texttt{naryFunc 1}) is convertible to \texttt{nat -> nat} and (\texttt{naryFunc 3})
-to \texttt{nat -> nat -> nat -> nat}.
-
-\vspace{4pt}
-\noindent
-\emph{From\href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}}.
-\begin{Coqsrc}
-Require Import primRec.
-Import extEqualNat.
-
-Compute naryFunc 3.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-= nat -> nat -> nat -> nat
-  : Set  
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Check plus: naryFunc 2.
-
-Check 42: naryFunc 0.
-
-Check (fun n p q : nat =>  n * p + q): naryFunc 3.
-\end{Coqsrc}
-
-Likewise, arbitrary boolean predicates may have an arbitrary number of arguments. The dependent type
-(\texttt{naryRel $n$}), defined in the same way as \texttt{naryFunc}, is the type of $n$-ary functions from
-\texttt{nat} into \texttt{bool}.
-
-\begin{Coqsrc}
-Compute naryRel 2.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = nat -> nat -> bool
-     : Set
-\end{Coqanswer}
-
-The magic of dependent types makes it possible to define recursively extensional equality between functions of the same arity.
-
-\index{coq}{Dependent types}
-\index{coq}{Dependently typed functions}
-\vspace{4pt}
-\noindent
-\emph{From \href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat}}
-
-\index{primrec}{Predicates!extEqual}
-
-\begin{Coqsrc}
-Fixpoint  extEqual (n : nat) : forall  (a b : naryFunc n), Prop :=
-  match n with
-    0 => fun a b => a = b
-  | S p => fun a b => forall c, extEqual p (a c) (b c)
-  end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Compute extEqual 2.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-     = fun a b : naryFunc 2 => forall x x0 : nat, a x x0 = b x x0
-     : naryFunc 2 -> naryFunc 2 -> Prop
- \end{Coqanswer}
-
-\begin{Coqsrc}
-Example extEqual_ex1 : extEqual 2 mult (fun x y =>  y * x + x - x) .
-Proof.
-  intros x y.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  x, y : nat
-  ============================
-  extEqual 0 (x * y) (y * x)
-\end{Coqanswer}
-
-\begin{Coqsrc}
-  cbn.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-1 subgoal (ID 10)
-
-  x, y : nat
-  ============================
-  x * y = y * x + x - x
-\end{Coqanswer}
-
-\begin{Coqsrc}
-  rewrite <- Nat.add_sub_assoc, Nat.sub_diag.
-  - ring.
-  - apply le_n.  
-Qed.
-\end{Coqsrc}
-
-\subsection{A Data-type for Primitive Recursive Functions}
+Primitive recursive functions are a small class of total functions, corresponding to the expressive power of a simple imperative programming language without \textbf{while} loops, in which every program execution terminates.
+
+Primitive recursive functions are total functions from $\mathbb{N}^n$ to
+$\mathbb{N}$, for some $n\in\mathbb{N}$. Note that not all 
+total $n$-ary recursive functions are primitive recursive
+(see for instance Sect.~\vref{sect:ack-not-PR}).
 
 The traditional definition of the set of primitive recursive functions is structured as an inductive definition 
-in five rules: three base cases, and two construction rules. Primitive recursive functions are total functions from $\mathbb{N}^n$ to
-$\mathbb{N}$, for some $n\in\mathbb{N}$.
+in five rules: three base cases, and two recursive construction rules.
 
@@
 
+Please note the use of dots: $\ldots$ in the definition above. 
+Dots are not part of \gallina's syntax. Thus, the formal definition of the set of primitive recursive function will have to overcome this representation problem.
+
+  We present in this chapter a formalization of  primitive recursive functions, taken from  Russel O'Connor's formalization in \coq{} of
+G\"odel's incompleteness theorems~\cite{OConnor05}.
+
+\begin{remark}
+ The theory of primitive recursive function is now hosted in
+the \texttt{theories/ordinals/Ackermann} directory.
+The specific part on G\"odel's theorem,  is also on
+coq-community (\url{https://github.com/coq-community/goedel}) and requires the 
+\href{https://github.com/coq-community/pocklington}{Pocklington library} for lemmas on primality.
+\end{remark}
+
+This chapter contains some comments on Russel's library, as well as a few extensions.
+Contributions (under the form of comments, new examples or exercises) are welcome!.
+
+\section{First look at the Ackermann library}
+
+O'Connor's library on Gödel's incompleteness theorems contains a little more 
+than 45K lines of scripts. The part dedicated to primitive recursive functions and Peano arithmetics is 32K lines long and is originally structured in 38 modules.
+Thus, we propose a slow exploration of this library, through examples and exercises. Our additions to the original library 
+are stored in the directory \texttt{theories/ordinals/MoreAck}.
+In particular, the library~\href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR} contains the well-known  proof that the Ackermann function is not primitive recursive (see Section~\vref{sect:ack-not-PR}).
+
+\section{Basic definitions}
+\index{maths}{Primitive recursive functions}
+
+The formal definition of primitive recursive functions lies in the library
+\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec},
+with preliminary definitions in 
+\href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat}
+and
+\href{../theories/html/hydras.Ackermann.misc.html}{Ackermann.misc}.
+
+\subsection{Functions of arbitrary arity}
+
+Library \texttt{primRec} allows us to consider primitive functions on \texttt{nat}, with any number of arguments, in 
+curried form. This is made possible in 
+\href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat} by the following definition:
+
+\index{primrec}{Types!naryFunc}
+\begin{Coqsrc}
+Fixpoint naryFunc (n : nat) : Set :=
+  match n with
+  | O => nat
+  | S n => nat -> naryFunc n
+  end.
+\end{Coqsrc}
+
+For instance (\texttt{naryFunc 1}) is convertible to \texttt{nat -> nat} and (\texttt{naryFunc 3})
+to \texttt{nat -> nat -> nat -> nat}.
+
+\vspace{4pt}
+\noindent
+\emph{From\href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}}.
+\begin{Coqsrc}
+Require Import primRec.
+Import extEqualNat.
+
+Compute naryFunc 3.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+= nat -> nat -> nat -> nat
+  : Set  
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Check plus: naryFunc 2.
+
+Check 42: naryFunc 0.
+
+Check (fun n p q : nat =>  n * p + q): naryFunc 3.
+\end{Coqsrc}
+
+Likewise, arbitrary boolean predicates may have an arbitrary number of arguments. The dependent type
+(\texttt{naryRel $n$}), defined in the same way as \texttt{naryFunc}, is the type of $n$-ary functions from
+\texttt{nat} into \texttt{bool}.
+
+\begin{Coqsrc}
+Compute naryRel 2.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+ = nat -> nat -> bool
+     : Set
+\end{Coqanswer}
+
+The magic of dependent types makes it possible to define recursively extensional equality between functions of the same arity.
+
+\index{coq}{Dependent types}
+\index{coq}{Dependently typed functions}
+\vspace{4pt}
+\noindent
+\emph{From \href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat}}
+
+\index{primrec}{Predicates!extEqual}
+
+\begin{Coqsrc}
+Fixpoint  extEqual (n : nat) : forall  (a b : naryFunc n), Prop :=
+  match n with
+    0 => fun a b => a = b
+  | S p => fun a b => forall c, extEqual p (a c) (b c)
+  end.
+\end{Coqsrc}
+
+\begin{Coqsrc}
+Compute extEqual 2.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+     = fun a b : naryFunc 2 => forall x x0 : nat, a x x0 = b x x0
+     : naryFunc 2 -> naryFunc 2 -> Prop
+ \end{Coqanswer}
+
+\begin{Coqsrc}
+Example extEqual_ex1 : extEqual 2 mult (fun x y =>  y * x + x - x) .
+Proof.
+  intros x y.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+  x, y : nat
+  ============================
+  extEqual 0 (x * y) (y * x)
+\end{Coqanswer}
+
+\begin{Coqsrc}
+  cbn.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+1 subgoal (ID 10)
+
+  x, y : nat
+  ============================
+  x * y = y * x + x - x
+\end{Coqanswer}
+
+\begin{Coqsrc}
+  rewrite <- Nat.add_sub_assoc, Nat.sub_diag.
+  - ring.
+  - apply le_n.  
+Qed.
+\end{Coqsrc}
+
+\subsection{A Data-type for Primitive Recursive Functions}
+
 O'Connor's formalization of primitive recursive functions takes the form of two mutually inductive dependent data types, each constructor of which is associated with one of these  rules.
@@
 is an interpretation function (\texttt{evalprimRec $n$})  of type
-$\texttt{PrimRec}\,n \rightarrow  \texttt{naryFunc}\,n$ , together with the  function 
+$\texttt{PrimRec}\,n \rightarrow  \texttt{naryFunc}\,n$.
+This function is definied by mutual recursion,  together with the  function 
 (\texttt{evalprimRecS $n$ $m$}) of type 
@@
 
-Ok, the term \texttt{bigPR} looks to be a primitive recursive definition of the factorial function, although we haven't proved this fact yet.
-Fortunately, we will see in the following sections much simpler ways to prove that a given function is primitive recursive, whithout looking at an unreadable term.
+After this test, the term \texttt{bigPR} looks to be a primitive recursive definition of the factorial function, although we haven't proved this fact yet.
+Fortunately, we will see in the following sections simple ways to prove that a given function is primitive recursive, whithout building such an unreadable term.
 
@@
 \texttt{Defined}s.
+
+If it does not cause compatibility problems (with 
+\href{https://github.com/coq-community/goedel}{goedel library} fr instance), we plan to make all theses lemmas transparent.
 \end{remark}
@@
 
-Thus, our first subgoal is solved easily and the proof ends, applying already proven lemmas.
+Thus, our first subgoal is solved easily. The rest of the proof 
+is just an application of already proven lemmas.
 
@@
 There is a lot of lemmas similar to \texttt{filter010IsPR} in the \texttt{primRec} library, useful to control the arity of functions.
-Thus, the reader may look at them, and invent simple examples of application for each one.
+Thus, the reader may look at them, and invent simple examples of application for each lemma.
 \end{exercise}
@@
 The following proof decomposes the \texttt{double} function as the composition of 
-multiplication with the identity and the constant function wihich returns $2$.
+multiplication with the identity and the constant function which returns $2$.
 \emph{Note that the lemma \texttt{const1\_NIsPR} considers this function as an unary function (unlike \texttt{const0\_NIsPR})}.
@@
 
-The three firts lemmas make us sure that our function 
+The three first lemmas make us sure that our function 
 \texttt{Ack} satifies the ``usual'' equations.
@@
   | S p => fun (v : Vector.t nat (S p)) =>
-             (max (Vector.hd v) (max_v  (Vector.tl v)))
+             max (Vector.hd v) (max_v  (Vector.tl v))
   end.

### 2021-04-30 sphinx-alectryon/source/conf.py
first steps with alacryon and sphinx

@@ -1,0 +1,69 @@
+# Configuration file for the Sphinx documentation builder.
+#
+# This file only contains a selection of the most common options. For a full
+# list see the documentation:
+# https://www.sphinx-doc.org/en/master/usage/configuration.html
+
+# -- Path setup --------------------------------------------------------------
+
+# If extensions (or modules to document with autodoc) are in another directory,
+# add these directories to sys.path here. If the directory is relative to the
+# documentation root, use os.path.abspath to make it absolute, like shown here.
+#
+# import os
+# import sys
+# sys.path.insert(0, os.path.abspath('.'))
+
+
+# -- Project information -----------------------------------------------------
+
+project = 'Hydra battles'
+copyright = '2021, Pierre Castéran'
+author = 'Pierre Castéran'
+
+# The full version, including alpha/beta/rc tags
+release = '0.0'
+
+
+# -- General configuration ---------------------------------------------------
+
+# Add any Sphinx extension module names here, as strings. They can be
+# extensions coming with Sphinx (named 'sphinx.ext.*') or your custom
+# ones.
+extensions = [
+]
+
+# Add any paths that contain templates here, relative to this directory.
+templates_path = ['_templates']
+
+# List of patterns, relative to source directory, that match files and
+# directories to ignore when looking for source files.
+# This pattern also affects html_static_path and html_extra_path.
+exclude_patterns = []
+
+
+# -- Options for HTML output -------------------------------------------------
+
+# The theme to use for HTML and HTML Help pages.  See the documentation for
+# a list of builtin themes.
+#
+#html_theme = 'alabaster'
+
+# Add any paths that contain custom static files (such as style sheets) here,
+# relative to this directory. They are copied after the builtin static files,
+# so a file named "default.css" will overwrite the builtin "default.css".
+html_static_path = ['_static']
+
+
+import sphinx_rtd_theme
+
+html_theme = 'sphinx_rtd_theme'
+
+html_theme_path = [sphinx_rtd_theme.get_html_theme_path()]
+
+
+import os
+import sys
+sys.path.insert(0, os.path.abspath('../../../alectryon/'))
+
+extensions = ["alectryon.sphinx"]

### 2021-05-09 doc/cutting-table/montage1.py
chantier

@@ -1,0 +1,55 @@
+#!/usr/bin/env python3
+import json
+import sys
+from os.path import join, dirname
+from pprint import pprint
+
+root = join(dirname(__file__), "..")
+sys.path.insert(0, root)
+
+sys.path.insert(0, '/Users/casteranpierre/alectryon')
+
+# import alectryon.core
+# alectryon.core.DEBUG = True
+
+def api_annotate():
+    from alectryon.core import annotate
+    annotated = annotate(["Check 1."], ("-Q", "{},logical_name".format(root)))
+    pprint(annotated)
+
+def annotated_to_json():
+    from alectryon.core import annotate
+    from alectryon.json import json_of_annotated
+    annotated = annotate([r"Goal True /\ True. split. ", "all: eauto."],
+                         ("-Q", "{},logical_name".format(root)))
+    print(json.dumps(json_of_annotated(annotated)))
+
+JS = """
+[[{
+            "_type": "sentence",
+            "contents": "Inductive Hydra : Set :=\n|  node :  Hydrae -> Hydra\nwith Hydrae : Set :=\n| hnil : Hydrae\n| hcons : Hydra -> Hydrae -> Hydrae.",
+            "messages": [],
+            "goals": []
+        }]]
+"""
+
+
+print(JS) # type of JS ?
+
+
+def latex_of_movie():
+    from alectryon.json import annotated_of_json
+    from alectryon.latex import LatexGenerator
+    from alectryon.pygments import highlight_latex
+    annotated = annotated_of_json(json.loads(JS, strict=False))
+    for ltx in LatexGenerator(highlight_latex).gen(annotated):
+      #  print("<latex>")
+        print(ltx)
+      #  print("</latex>")
+
+def main():
+    api_annotate()
+    latex_of_movie()
+
+if __name__ == '__main__':
+    main()

### 2021-05-09 sphinx-alectryon/source/conf.py

@@ -1,69 +1,0 @@
-# Configuration file for the Sphinx documentation builder.
-#
-# This file only contains a selection of the most common options. For a full
-# list see the documentation:
-# https://www.sphinx-doc.org/en/master/usage/configuration.html
-
-# -- Path setup --------------------------------------------------------------
-
-# If extensions (or modules to document with autodoc) are in another directory,
-# add these directories to sys.path here. If the directory is relative to the
-# documentation root, use os.path.abspath to make it absolute, like shown here.
-#
-# import os
-# import sys
-# sys.path.insert(0, os.path.abspath('.'))
-
-
-# -- Project information -----------------------------------------------------
-
-project = 'Hydra battles'
-copyright = '2021, Pierre Castéran'
-author = 'Pierre Castéran'
-
-# The full version, including alpha/beta/rc tags
-release = '0.0'
-
-
-# -- General configuration ---------------------------------------------------
-
-# Add any Sphinx extension module names here, as strings. They can be
-# extensions coming with Sphinx (named 'sphinx.ext.*') or your custom
-# ones.
-extensions = [
-]
-
-# Add any paths that contain templates here, relative to this directory.
-templates_path = ['_templates']
-
-# List of patterns, relative to source directory, that match files and
-# directories to ignore when looking for source files.
-# This pattern also affects html_static_path and html_extra_path.
-exclude_patterns = []
-
-
-# -- Options for HTML output -------------------------------------------------
-
-# The theme to use for HTML and HTML Help pages.  See the documentation for
-# a list of builtin themes.
-#
-#html_theme = 'alabaster'
-
-# Add any paths that contain custom static files (such as style sheets) here,
-# relative to this directory. They are copied after the builtin static files,
-# so a file named "default.css" will overwrite the builtin "default.css".
-html_static_path = ['_static']
-
-
-import sphinx_rtd_theme
-
-html_theme = 'sphinx_rtd_theme'
-
-html_theme_path = [sphinx_rtd_theme.get_html_theme_path()]
-
-
-import os
-import sys
-sys.path.insert(0, os.path.abspath('../../../alectryon/'))
-
-extensions = ["alectryon.sphinx"]

### 2021-05-09 doc/cutting-table/montage1.py
chantier

@@ -6,7 +6,6 @@ from pprint import pprint
 
 root = join(dirname(__file__), "..")
 sys.path.insert(0, root)
-
 sys.path.insert(0, '/Users/casteranpierre/alectryon')
 
 # import alectryon.core
@@ -24,6 +23,8 @@ def annotated_to_json():
                          ("-Q", "{},logical_name".format(root)))
     print(json.dumps(json_of_annotated(annotated)))
 
+#Todo  get JS  as an argument on the command line
+
 JS = """
 [[{
             "_type": "sentence",
@@ -34,7 +35,7 @@ JS = """
 """
 
 
-print(JS) # type of JS ?
+# print(JS) # type of JS ?
 
 
 def latex_of_movie():
@@ -48,7 +49,7 @@ def latex_of_movie():
       #  print("</latex>")
 
 def main():
-    api_annotate()
+#    api_annotate()
     latex_of_movie()
 
 if __name__ == '__main__':

### 2021-05-10 doc/cutting-table/jsonToLatex.py
issue in latex generation

@@ -1,0 +1,31 @@
+#!/usr/bin/env python3
+import json
+import sys
+from os.path import join, dirname
+from pprint import pprint
+movie_file = sys.argv[1]
+
+root = join(dirname(__file__), "..")
+sys.path.insert(0, root)
+sys.path.insert(0, '/Users/casteranpierre/alectryon')
+
+with open(movie_file, "r") as read_file:
+    movie =json.load(read_file, strict=False)
+
+# data =  movie[0] 
+
+def latex_of_movie():
+    from alectryon.json import annotated_of_json
+    from alectryon.latex import LatexGenerator
+    from alectryon.pygments import highlight_latex
+    annotated = annotated_of_json(movie)
+    for ltx in LatexGenerator(highlight_latex).gen(annotated):
+      #   print("<latex>")
+        print(ltx)
+      #   print("</latex>")
+
+def main():
+    latex_of_movie()
+
+if __name__ == '__main__':
+    main()       

### 2021-05-10 doc/hydras.tex
consistent use of capitalization in headers
work on coqdoc comments
chantier
corrections of typos

@@
 \usepackage{url}
-\usepackage{mdframed}
+\usepackage{xcolor,fancyvrb, mdframed}
 \usepackage{amsmath}
@@
 \usepackage{threeparttable}
+\usepackage{fontspec}
+\usepackage{mathtools}
+\usepackage{xcolor}
+\usepackage{caption}
+\usepackage{placeins}
+\usepackage{tikz}
+\usepackage{tikzsymbols}
+\usetikzlibrary{arrows}
+\usepackage{graphicx}
+\usepackage{amsmath,mathdots,dsfont}
+\usepackage{amssymb}
+\usepackage{newunicodechar}
+\usepackage{theorem}
+
 \setcounter{tocdepth}{1}
@@
 \item There is no measure into $\mathbb{N}$ for proving the termination of all hydra battles (Sect~\vref{omega-case}).
-\item There is no measure into the interval\footnote{We use the notation $[a,b)$ for denoting the set of ordinals greater or equal than $a$ and strictly less than $b$.} interval $\{x\$[0,\omega^2)$ for proving the termination of all hydra battles (Sect~\vref{omega2-case}).
-\item There is no measure into $[0,\mu)$ for proving the termination of all hydra battles, for any $\mu<\epsilon_0$ (Sect\vref{sec:free-battles-case}).
+\item There is no measure into the interval\footnote{We use the notation $[a,b)$ for denoting the set of ordinals greater or equal than $a$ and strictly less than $b$.}  $[0,\omega^2)$ for proving the termination of all hydra battles (Sect.~\vref{omega2-case}).
+\item There is no measure into $[0,\mu)$ for proving the termination of all hydra battles, for any $\mu<\epsilon_0$ (Sect.\vref{sec:free-battles-case}).
 \end{enumerate}
@@
 Bad definitions 
- and encapsulation in modules called \texttt{Bad}, \texttt{Bad1}, etc.
+ are ''masked '' inside modules called \texttt{Bad}, \texttt{Bad1}, etc.
 
@@
 
-\part{Hydras and Ordinals}
+\part{Hydras and ordinals}
 
@@
 \include{chapter-primrec}
-\part{A Few  Certified Algorithms}
+\part{A few  certified algorithms}
 

### 2021-05-10 doc/part-hydras.tex
minor corrections
small changes
correction of typos
chantier
corrections of typos

@@
 
-\item We prove a relation between the length of a ``classic''  class of  battles \footnote{This class is also called \emph{standard} in this document (text and proofs).
+\item We prove a relation between the length of a ``classic''  kind of  battles \footnote{This class is also called \emph{standard} in this document (text and proofs).
 The \emph{replication factor} of the hydra is exactly $i$ at the $i$-th round of the battle (see Sect~\vref{sect:replication-def}).}
@@
 
-The following function ``gnaws'' an ordinal $\alpha$, folowing a sequence of indices (ignoring the $0$s).
+The following function ``gnaws'' an ordinal $\alpha$, following a sequence of indices (ignoring the $0$s).
 \noindent
@@
 $$
-H_{\omega^3}(3) =  2 ^ {(2 ^ {N + 1} \times (N+1) } )  \times  (2 ^ {N+1} \times ( N +1) ) - 1
+H_{\omega^3}(3) =  2 ^ {(2 ^ {N + 1} \, (N+1) )}   \,  (2 ^ {N+1} \, ( N +1) ) - 1
 $$
@@
 
-\section{The wainer hierarchy (functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
+\section{The Wainer hierarchy (functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
 \label{sect:wainer}

### 2021-05-11 doc/chapter-primrec.tex
chantier
corrections of typos
minor changes

@@
 
+\begin{remark}
+ The statements above can be rewritten in a more uniform way:
+
+ \begin{quote}
+   For $m\in 1..4$, $\texttt{Ack}\,m\,n = f_m\,(n+3)-3$, where 
+   \begin{align*}
+   f_1(n)=&\,n+2 \\
+   f_2(n)=&\,n\times 2\\
+   f_3(n)=&\,2^n\\
+   f_4(n)=&\,2^{2^{\dots^2}}\quad(n\;\textit{levels})
+   \end{align*}
+ \end{quote}
+\end{remark}
+
 An important property of the Ackermann function helps us 
@@
 (\texttt{Ack\,$n$}) is primitive recursive.
+
+\vspace{4pt}
 
@@
 \vspace{4pt}
+
 \noindent
@@
 
-Notation "'v_apply' f v" := (evalList _ v f) ( at level 10, f at level 9).
-
-Example Ex2 : forall (f: naryFunc 2) x y,
+Notation "'v_apply' f v" := (evalList _ v f) 
+     (at level 10, f at level 9).
+
+Example Ex2: forall (f: naryFunc 2) x y,
     v_apply f (x::y::nil) = f x y.
@@
 
-Example Ex4 : forall (f: naryFunc 4) x y z t,
+Example Ex4: forall (f: naryFunc 4) x y z t,
     v_apply f (x::y::z::t::nil) = f x y z t.
@@
 It is easy to prove that any unary function which dominatates \texttt{fun n => Ack n n} fails to be primitive recursive. We use an instance of \texttt{majorAnyPR} for unary functions.
+
 \vspace{4pt}

### 2021-05-14 doc/chapter-primrec.tex
proof that the length  of hydra battles is not primitive recursive

@@
 \end{remark}
+
+\subsection{The length of standard hydra battles}
+
+The module \href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra\_Theorems} contains a proof that the function which computes the lenght of standard hydra battles is not primitive recursive.
+This proof is  a little more complex than the preceding one.
+
+\subsubsection{Definitions}
+
+The function we consider is defined and proven correct in
+Module~\href{../theories/html/hydras.Hydra.Battle_length.html}{Hydra.Battle\_length}.
+
+\begin{Coqsrc}
+Definition l_std alpha k := (L_ alpha (S k) - k)%nat.
+
+Lemma l_std_ok : forall alpha : E0,
+    alpha <> Zero ->
+    forall k : nat,
+      1 <= k -> battle_length standard k (iota (cnf alpha))
+                              (l_std alpha k).
+(* ... *)
+\end{Coqsrc}
+
+\subsubsection{Proof steps}
+
+Now, let us assume that the function \texttt{l\_std} is primitive recursive.
+
+\emph{From \href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems}}.
+
+\begin{Coqsrc}
+Section battle_lenght_notPR.
+
+  (** We assume that the function with computes the length 
+      of standard battles is primitive recursive *)
+
+  Hypothesis H: forall alpha, isPR 1 (l_std alpha). 
+\end{Coqsrc}
+
+Let us consider the hydra represented by the ordinal $\omega^\omega$.
+
+\begin{Coqsrc}
+Let alpha := Phi0 omega%e0.
+Let h := iota (cnf alpha).
+\end{Coqsrc}
+
+In order to get rid of the substraction in the definition of \texttt{l\_std}, we work with a helper function.
+
+\begin{Coqsrc}
+Let m k := L_ alpha (S k).
+
+Remark m_eqn : forall k, m k = (l_std alpha k + k)%nat.
+(* ... *)
+\end{Coqsrc}
+
+Under the hypothesis \texttt{H}, $m$ is also primitive recursive.
+
+\begin{Coqsrc}
+Remark mIsPR : isPR 1 m.
+(* ... *)
+\end{Coqsrc}
+
+Now, let us recall the following lemmas :
+
+\begin{enumerate}
+\item 
+\emph{From \href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}}
+
+\begin{Coqsrc}
+H_F : forall (alpha : E0) (n : nat), 
+     F_ alpha (S n) <= H_ (Phi0 alpha) (S n) 
+\end{Coqsrc}
+
+\emph{Please note that our proof of this lemma is not trivial at all. It uses some lemmas of Chapters~\vref{chap:ketonen} and \vref{chap:alpha-large}.}
+
+\item
+\noindent
+\emph{From \href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0.L\_alpha}}
+
+\begin{Coqsrc}
+H_L_ : forall (alpha : E0) (i : nat), H_ alpha i <= L_ alpha (S i)
+\end{Coqsrc}
+
+\item
+
+\emph{From \href{../theories/html/hydras.Epsilon0.F_omega.html}{Epsilon0.F\_omega}}
+
+\begin{Coqsrc}
+F_vs_Ack : forall n : nat, 2 <= n -> Ack n n <= F_ omega n
+\end{Coqsrc}
+
+\end{enumerate}
+
+By transitivity, we get the inequality
+$F_\omega(k+1)\leq m(k+1)$, for any $k$.
+
+\begin{Coqsrc}
+Remark m_ge_F_omega : forall k,  F_ omega (S k) <= m (S k).
+\end{Coqsrc}
+
+We finish the proof by noting that the function $m$ (composed with \texttt{S}) dominates the Ackermann function, which leads to a contradiction.
+
+\begin{Coqsrc}
+ Remark m_dominates_Ack : 
+     dominates (fun n =>  S (m n)) (fun n => Ack.Ack n n).
+(* ... *)
+
+ Remark SmNotPR : isPR 1 (fun n => S (m n)) -> False.
+
+ Theorem LNotPR : False.
+  Proof.
+    apply SmNotPR,  compose1_1IsPR.
+    - apply mIsPR.
+    - apply succIsPR.
+  Qed.
+
+End battle_lenght_notPR.
+\end{Coqsrc}

### 2021-05-15 doc/part-hydras.tex
minor changes
proof that the length  of hydra battles is not primitive recursive
Fix references to fast growing hierarchies

@@
 
-\section{The Wainer hierarchy (functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
+\section{A variant of the Wainer hierarchy (functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
 \label{sect:wainer}
@@
 
-The Wainer hierarchy~\cite{BW85, Wainer1970, KS81}, is also a family of fast growing functions, indexed by ordinals below $\epsilon_0$, by the following equations:
+Ketonen and Solovay introduce in~\cite{KS81} a ``trivial'' variant of the Wainer hierarchy~\cite{BW85, Wainer1970} of fast growing functions, indexed by ordinals below $\epsilon_0$.
 
@@
 
-A first attempt is to write a definition of $F_\alpha$ by equations, in the same as for $H\_alpha$.  We use the functional \texttt{iterate} defined in 
+\begin{remark}
+The difference with the ``classic'' Wainer hierarchy 
+$f\_alpha\;(\alpha<\epsilon_0)$ lies in the second equation:
+$f_{\beta+1}(i) = (f_\beta)^{(i)}(i)$ and not
+$f_{\beta+1}(i) = (f_\beta)^{(i+1)}(i)$. A module about 
+the classic Wainer hierarchy is in preparation.
+\end{remark}
+
+A first attempt is to write a definition of $F_\alpha$ by equations, in the same way as for $H\_alpha$.  We use the functional \texttt{iterate} defined in 
 Module~\href{../theories/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
@@
 
-Theorem F_alpha_beta alpha : 
+Theorem F_restricted_mono_l alpha : 
    forall beta n, Canon_plus n alpha beta -> 

### 2021-05-17 doc/part-hydras.tex
small change in doc and comments

@@
 
+\begin{remark}
+Some notations (mainly names of fast-growing functions) of our development may differ slightly from the litterature. Although this fact does not affect our proofs, we are preparing a future 
+version where the names $F\_alpha$, $f\_alpha$, $H\_alpha$, etc., are fully consistent with the cited articles.
+This change will be provisionally made in a branch \texttt{Hardy} (in reference to the Hardy hierarchy).
+\end{remark}
 %\section{Introduction}

### 2021-05-17 doc/part-hydras.tex
minor changes

@@
 version where the names $F\_alpha$, $f\_alpha$, $H\_alpha$, etc., are fully consistent with the cited articles.
-This change will be provisionally made in a branch \texttt{Hardy} (in reference to the Hardy hierarchy).
+This change will be provisionally made in a branch called \texttt{Hardy} (in reference to the Hardy hierarchy).
 \end{remark}
@@
 \begin{remark}
-The upper bound of the considered interval has been chosen to be  $l-1$ and not $l$, in order to simplify some statements and proofs in composition lemmas associated with the ordinals of the form $\alpha\times i$ and
+In~\cite{KS81} Ketonen and Solovay consider the least natural number $l$ where the interval $[k,l]$ ($l$ included) is $\alpha$-large, and call $H_\alpha$ the function which maps $k$ to $l$.
+We chose to consider intervals $[l,k)$ instead of $[l,k]$
+in order to simplify  some statements and proofs in composition lemmas associated with the ordinals of the form $\alpha\times i$ and 
 $\omega^\alpha\times i + \beta$.
-In order to consider any ordinal below $\epsilon_0$, we consider a special case for $\alpha=0$.
+Clearly, both approachs are related through the equality
+$L_\alpha(k)=H_\alpha(k)+1$, for any non-null $\alpha$ and $k$.
 \end{remark}
@@
 
-It is worth looking at the answer from \texttt{Equations} and check (with \texttt{About} ) all the lemmas this plug-in gives you for free. We show here only a part of \coq's anwer.
+It is worth looking at the answer from \texttt{Equations} and at all the lemmas this plug-in gives you for free. We show here only a part of \coq's anwer.
 
@@
 \label{sect:hardy}
+
+\begin{remark}
+  Indeed, the functions presented in this section are a \emph{variant} of the Hardy hierarchy of functions.
+In the future versions of this development, we will correct the references to the litterature.
+\end{remark}
+
 In order to give a feeling on  the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the so called \emph{Wainer-Hardy hierarchy} of fast growing functions,
@@
 
-For each ordinal $\alpha$ below $\epsilon_0$, $H_\alpha$ is a 
-total arithmetic function, defined  by  transfinite recursion on $\alpha$, according to three cases:
+For each ordinal $\alpha$ below $\epsilon_0$, $H_\alpha$ is 
+a total arithmetic function, defined  by  transfinite recursion on $\alpha$, according to three cases:
 

### 2021-05-17 doc/part-hydras.tex
Corrected references to the Hardy hierarchy
The variant of the hardy hierarchi is now called H'_ alpha, in order to avoid
confusion with the litterature

@@
 version where the names $F\_alpha$, $f\_alpha$, $H\_alpha$, etc., are fully consistent with the cited articles. 
-This change will be provisionally made in a branch called \texttt{Hardy} (in reference to the Hardy hierarchy).
 \end{remark}
@@
 
-\section{The Wainer-Hardy hierarchy (functions \texorpdfstring{$H_\alpha$}{\texttt{H\_alpha}})}
+\section{A variant of the Wainer-Hardy hierarchy}
 
 \label{sect:hardy}
+
+In order to give a feeling on  the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the so called \emph{Wainer-Hardy hierarchy} of fast growing functions,
+presented for instance in~\cite{Promel2013}. 
+\index{maths}{Rapidly growing functions!Hardy Hierarchy}
 
@@
 In the future versions of this development, we will correct the references to the litterature.
+For the time being, we call our functions $H'\_alpha$ in order to underline the difference from ``classic'' Hardy functions.
 \end{remark}
 
-In order to give a feeling on  the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the so called \emph{Wainer-Hardy hierarchy} of fast growing functions,
-presented for instance in~\cite{Promel2013}. 
-\index{maths}{Rapidly growing functions!Hardy Hierarchy}
-
-For each ordinal $\alpha$ below $\epsilon_0$, $H_\alpha$ is 
+For each ordinal $\alpha$ below $\epsilon_0$, $H'_\alpha$ is 
 a total arithmetic function, defined  by  transfinite recursion on $\alpha$, according to three cases:
@@
 \index{maths}{Transfinite induction}
+
+\begin{itemize}
+\item If $\alpha=0$, then $H'_\alpha (k)= k$ for any natural number $k$.
+\item If $\alpha=\textrm{succ}(\beta)$, then 
+$H'_\alpha(k)=H'_\beta(k+1)$ for any $k \in \mathbb{N}$
+\item If $\alpha$ is a limit ordinal, then 
+$H'_\alpha(k) = H'_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
+\end{itemize}
+
+\begin{remark}
+ The ``classic'' definition of the Wainer-Hardy hierarchy differs in the third equation.
 
@@
 \item If $\alpha$ is a limit ordinal, then 
-$H_\alpha(k) = H_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
+$H_\alpha(k) = H_{(\canonseq{\alpha}{k})}(k)$ for any $k\in \mathbb{N}$.
 \end{itemize}
 
-\subsection{Hardy functions in \texttt{Coq}}
-
-We define a function \texttt{H\_} of type \texttt{E0 -> nat -> nat} by transfinite induction over the type \texttt{E0} of the well formed ordinals below $\epsilon_0$.
+\end{remark}
+
+\subsection{Definition in \texttt{Coq}}
+
+We define a function \texttt{H'\_} of type \texttt{E0 -> nat -> nat} by transfinite induction over the type \texttt{E0} of the well formed ordinals below $\epsilon_0$.
 
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.H_alpha.html\#H_}{Epsilon0.H\_alpha}}
-
-\index{hydras}{Library Epsilon0!Functions!H\_@H\_ (Hardy hierarchy)}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}}
+
+\index{hydras}{Library Epsilon0!Functions!H\_@H\_ (Hardy hierarchy (variant))}
 \index{coq}{Plug-ins!Equations}
-\label{Functions:H-alpha}
-
-\begin{Coqsrc}
-Equations H_ (alpha: E0) (i:nat) :  nat  by wf  alpha lt :=
-  H_ alpha  i with E0_eq_dec alpha Zero :=
+\label{Functions:Hprime-alpha}
+
+\begin{Coqsrc}
+Equations H'_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
+  H'_ alpha  i with E0_eq_dec alpha Zero :=
     { | left _ =>  i ;
@@
           with Utils.dec (Limitb alpha) :=
-          { | left _ =>  H_ (Canon alpha (S i))  i ;
-            | right notlimit =>  H_ (Pred alpha) (S i)}}.
+          { | left _ =>  H'_ (Canon alpha (S i))  i ;
+            | right notlimit =>  H'_ (Pred alpha) (S i)}}.
 
@@
 \begin{Coqsrc}
-Lemma H_eq1 : forall i, H_ Zero i = i.
-Proof.   intro i; now rewrite H__equation_1.  Qed.
-
-Lemma H_eq2 alpha i : Is_Succ alpha ->
-                      H_ alpha i = H_ (Pred alpha) (S i).
-
-Lemma H_eq3 alpha i : Limitb alpha ->
-                      H_ alpha i =  H_ (Canon alpha (S i)) i.
-
-Lemma H_eq4  alpha i :  H_ (Succ alpha) i = H_ alpha (S i).
-\end{Coqsrc}
-
-\subsection{First  steps of the Hardy hierarchy}
-Using rewrite rules from \texttt{H\_eq1} to \texttt{H\_eq4}, we can explore the functions $H_\alpha$ for some small values of $\alpha$.
+Lemma H'_eq1 : forall i, H'_ Zero i = i.
+Proof.
+  intro i; now rewrite H'__equation_1. 
+Qed.
+
+Lemma H'_eq2 alpha i : Succb alpha ->
+                       H'_ alpha i = H'_ (Pred alpha) (S i).
+
+Lemma H'_eq3 alpha i : Limitb alpha ->
+                      H'_ alpha i =  H'_ (Canon alpha (S i)) i.
+
+Lemma H'_succ_eqn  alpha i :  
+            H'_ (Succ alpha) i = H'_ alpha (S i).
+\end{Coqsrc}
+
+\subsection{First  steps of the H' hierarchy}
+Using rewrite rules from \texttt{H'\_eq1} to \texttt{H'\_succ\_eqn}, we can explore the functions $H'_\alpha$ for  small values of $\alpha$.
+
+%ICI
 
@@
 
-By induction on $i$, we prove a simple expression of \texttt{H\_ (Fin i)}, where 
+By induction on $i$, we prove a simple expression of \texttt{H'\_ (Fin i)}, where 
 \texttt{Fin $i$}  is the $i$-th finite ordinal.
@@
 \begin{Coqsrc}
-Lemma H_Fin : forall i k: nat,  H_ (Fin i) k = (i+k)%nat.
+Lemma H'_Fin : forall i k : nat,  H'_  (Fin i) k = (i+k)%nat.
 Proof with eauto with E0.
   induction i.
-  - intros; simpl OF; simpl; autorewrite with H_rw E0_rw ... 
-  - intros ;simpl; autorewrite with H_rw E0_rw ... 
-    rewrite IHi; lia. 
+  - intros; simpl Fin; simpl; autorewrite with H'_rw E0_rw ... 
+  - intros ;simpl; autorewrite with H'_rw E0_rw ... 
+    rewrite IHi; abstract lia. 
 Qed.
@@
 \begin{Coqsrc}
-Lemma H_omega : forall k, H_ Omega k = S (2 * k)%nat.
+Lemma H'_omega : forall k, H'_ omega k = S (2 * k)%nat.
 Proof with auto with E0.
-  intro k; rewrite H_eq3 ...
+  intro k; rewrite H'_eq3 ...
   - replace (Canon omega (S k)) with (Fin (S k)).
-    + rewrite H_Fin; lia.
+    + rewrite H'_Fin; abstract lia.
     + now autorewrite with E0_rw.
@@
 \begin{Coqsrc}
-Lemma H_Plus_Fin alpha : forall i k : nat,
-    H_ (alpha + i)%e0 k = H_ alpha (i + k)%nat.
+Lemma H'_Plus_Fin alpha : forall i k : nat,
+    H'_ (alpha + i)%e0 k = H'_ alpha (i + k)%nat.
 (* Proof by induction on i *)
@@
 
-Then, we get easily formulas for $H_{\omega+i}$, and $H_{\omega\times i}$ for any natural number $i$.
-
-\begin{Coqsrc}
-Lemma H_omega_double k : H_ (omega * 2)%e0 k =  (4 * k + 3)%nat.
+Then, we get easily formulas for $H'_{\omega+i}$, and $H'_{\omega\times i}$ for any natural number $i$.
+
+\begin{Coqsrc}
+Lemma H'_omega_double k : H'_ (omega * 2)%e0 k =  (4 * k + 3)%nat.
 Proof.
- rewrite H_lim_eqn; simpl Canon.
- - ochange  (CanonS  (omega * 2)%e0 k)  (omega + (S k))%e0.
-  + rewrite H_Plus_Fin, H_omega;  lia.
+  rewrite H'_eq3; simpl Canon.
+ -   ochange  (CanonS  (omega * FinS 1)%e0 k)  (omega + (S k))%e0.
+  + rewrite H'_Plus_Fin, H'_omega;  abstract lia.
   -  now compute.
@@
 
-Lemma H_omega_3 k : H_ (omega * 3)%e0 k = (8 * k + 7)%nat.
-
-Lemma H_omega_4 k : H_ (omega * 4)%e0 k = (16 * k + 15)%nat.
-
-Lemma H_omega_i i  : forall k,
-    H_ (omega * i)%e0 k = (exp2 i * k + Nat.pred (exp2 i))%nat.
+Lemma H'_omega_3 k : H'_ (omega * 3)%e0 k = (8 * k + 7)%nat.
+
+Lemma H'_omega_4 k : H'_ (omega * 4)%e0 k = (16 * k + 15)%nat.
+
+Lemma H'_omega_i i  : forall k,
+    H'_ (omega * i)%e0 k = (exp2 i * k + Nat.pred (exp2 i))%nat.
 \end{Coqsrc}
@@
 Crossing a new limit, we prove the following equality: 
-$$H_{\omega^2} (k) = 2 ^ {k+1} \times (k+1) - 1$$.
-
-\begin{Coqsrc}
-
-Lemma H_omega_sqr : forall k,
-    H_ (Phi0  2)%e0 k = (exp2 (S k ) *  S k - 1)%nat.
+$$H'_{\omega^2} (k) = 2 ^ {k+1} \times (k+1) - 1$$.
+
+\begin{Coqsrc}
+Lemma H'_omega_sqr : forall k,
+    H'_ (Phi0  2)%e0 k = (exp2 (S k ) * (S k) - 1)%nat.
 Proof.
-  intro k; 
-   rewrite H_lim_eqn; auto with E0.
+  intro k; rewrite H'_eq3; auto with E0.
   - ochange (Canon (Phi0 2) (S k)) (omega * (S k))%e0.
-    +  rewrite H_omega_i; simpl (exp2 (S k)).
+    +  rewrite H'_omega_i; simpl (exp2 (S k)).
        *  rewrite Nat.add_pred_r.
-          -- lia. 
-          --   generalize (exp2_not_zero k);  lia.
-    + cbn; f_equal; lia.
+          -- abstract lia. 
+          --   generalize (exp2_not_zero k);  abstract lia.
+    + cbn; f_equal; abstract lia.
 Qed.
@@
 
-Our next step would be to prove an exact formula for $H_{\omega^\omega}(k)$.
+Our next step would be to prove an exact formula for $H'_{\omega^\omega}(k)$.
 Since the canonical sequence of $\omega^\omega$ is composed of all the
-$\omega^i$, we first need to express $H_{\omega^i}$ for any natural number $i$.
+$\omega^i$, we first need to express $H'_{\omega^i}$ for any natural number $i$.
 
@@
 \begin{enumerate}
-\item First,  we express $H_{\omega^\alpha \times (i+2)}$ in terms of
-$H_{\omega^\alpha \times (i+1)}$.
-
-\begin{Coqsrc}
-Lemma H_Omega_term_1 : alpha <> Zero -> forall  k,  
-    H_ (Omega_term alpha (S i)) k =
-    H_ (Omega_term alpha i) (H_ (Phi0 alpha) k).
+\item First,  we express $H'_{\omega^\alpha \times (i+2)}$ in terms of
+$H'_{\omega^\alpha \times (i+1)}$.
+
+\begin{Coqsrc}
+Lemma H'_Omega_term_1 : alpha <> Zero -> forall  k,  
+    H'_ (Omega_term alpha (S i)) k =
+    H'_ (Omega_term alpha i) (H'_ (Phi0 alpha) k).
 \end{Coqsrc}
@@
 \item
-Then, we prove by induction on $i$ that $H_{\omega^\alpha \times (i+1)}$ is just the
-$(i+1)$-th iterate of $H_{\omega^\alpha}$.
-
-\begin{Coqsrc}
-Lemma H_Omega_term (alpha : E0)  :
+Then, we prove by induction on $i$ that $H'_{\omega^\alpha \times (i+1)}$ is just the
+$(i+1)$-th iterate of $H'_{\omega^\alpha}$.
+
+\begin{Coqsrc}
+Lemma H'_Omega_term (alpha : E0)  :
 alpha <> Zero -> forall i k, 
-  H_ (Omega_term alpha i) k = iterate  (H_ (Phi0 alpha)) (S i) k.
-\end{Coqsrc}
-
-\item In particular, we have got a formula for $H_{\omega^{i+1}}$.
-
-\begin{Coqsrc}
-Definition H_succ_fun f k := iterate f (S k) k.
-
-Lemma H_Phi0_succ alpha  : alpha <> Zero -> forall k,
-      H_ (Phi0 (Succ alpha)) k = H_succ_fun (H_ (Phi0 alpha)) k.
-
-Lemma H_Phi0_Si : forall i k,
-      H_ (Phi0 (S i)) k = iterate H_succ_fun i (H_ omega) k. 
+  H'_ (Omega_term alpha i) k = iterate  (H'_ (Phi0 alpha)) (S i) k.
+\end{Coqsrc}
+
+\item In particular, we have got a formula for $H'_{\omega^{i+1}}$.
+
+\begin{Coqsrc}
+Definition H'_succ_fun f k := iterate f (S k) k.
+
+Lemma H'_Phi0_succ alpha  : alpha <> Zero -> forall k,
+      H'_ (Phi0 (Succ alpha)) k = H'_succ_fun (H'_ (Phi0 alpha)) k.
+
+Lemma H'_Phi0_Si : forall i k,
+      H'_ (Phi0 (S i)) k = iterate H'_succ_fun i (H'_ omega) k. 
 \end{Coqsrc}
@@
 \end{enumerate}
-We get now a  formula for $H_{\omega^3}$:
-
-\begin{Coqsrc}
-Lemma H_omega_cube : forall k,
-    H_ (Phi0 3)%e0 k = iterate (H_ (Phi0 2))  (S k) k.
+We get now a  formula for $H'_{\omega^3}$:
+
+\begin{Coqsrc}
+Lemma H'_omega_cube : forall k,
+    H'_ (Phi0 3)%e0 k = iterate (H'_ (Phi0 2))  (S k) k.
 Proof.
-  intro k; rewrite <-FinS_eq, -> Fin_Succ, H_Phi0_succ; auto.
+  intro k; rewrite <-FinS_eq, -> Fin_Succ, H'_Phi0_succ; auto.
   compute; injection 1; discriminate.
@@
 ``exact'' formula. 
-Let us consider a simple example, the number $H_{\omega^3}(3)$.
-
-\begin{Coqsrc}
-Section H_omega_cube_3.
+Let us consider a simple example: the number $H'_{\omega^3}(3)$.
+
+\begin{Coqsrc}
+Section H'_omega_cube_3.
 
@@
 
-Remark R0 k :  H_ (Phi0 3)%e0 k = iterate f (S k) k.
+Remark R0 k :  H'_ (Phi0 3)%e0 k = iterate f (S k) k.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Fact F0 : H_ (Phi0 3) 3 = f (f (f (f 3))).
+Fact F0 : H'_ (Phi0 3) 3 = f (f (f (f 3))).
  rewrite R0; reflexivity. 
@@
 \begin{Coqsrc}
-Fact F1 : H_ (Phi0 3) 3 = f (f N).
+Fact F1 : H'_ (Phi0 3) 3 = f (f N).
 Proof.
- rewrite H_omega_cube_0; reflexivity. 
+ rewrite F0; reflexivity. 
 Qed.
 
-Lemma F1_simpl : H_ (Phi0 3) 3 =
+Lemma F1_simpl : H'_ (Phi0 3) 3 =
                  (exp2 (exp2 (S N) * S N) * (exp2 (S N) * S N) - 1)%nat.
@@
 $$
-H_{\omega^3}(3) =  2 ^ {(2 ^ {N + 1} \, (N+1) )}   \,  (2 ^ {N+1} \, ( N +1) ) - 1
+H'_{\omega^3}(3) =  2 ^ {(2 ^ {N + 1} \, (N+1) )}   \,  (2 ^ {N+1} \, ( N +1) ) - 1
 $$
@@
 we do not take into account the multiplications in the formula above,
-we obtain that, in base $2$, the number $H_{\omega^3}(3)$ has at least
+we obtain that, in base $2$, the number $H'_{\omega^3}(3)$ has at least
 $2^{10^{21}}$  digits. But it is still an under-approximation !
@@
 \begin{Coqsrc}
-End H_omega_cube_3.
-\end{Coqsrc}
-
-Now, we have got at last an exact formula for $H_{\omega^\omega}$.
-
-\begin{Coqsrc}
-Lemma H_Phi0_omega : forall k, H_ (Phi0 omega) k =
-                               iterate H_succ_fun  k (H_ omega) k.
+End H'_omega_cube_3.
+\end{Coqsrc}
+
+Now, we have got at last an exact formula for $H'_{\omega^\omega}$.
+
+\begin{Coqsrc}
+Lemma H'_Phi0_omega : forall k, H'_ (Phi0 omega) k =
+                               iterate H'_succ_fun  k (H'_ omega) k.
 Proof with auto with E0.
-  intro k; rewrite H_lim_eqn, <- H_Phi0_Si ...
+  intro k; rewrite H'_eq3, <- H'_Phi0_Si ...
   -  rewrite CanonS_Canon, CanonS_Phi0_lim;  f_equal ...
@@
 \begin{Coqsrc}
-Lemma H_Phi0_omega_closed_formula k :
-  H_ (Phi0 omega) k =
+Lemma H'_Phi0_omega_closed_formula k :
+  H'_ (Phi0 omega) k =
   iterate (fun (f: nat -> nat) (l : nat) => iterate  f (S l) l)
@@
 
-\subsection{Abstract properties of Hardy functions}
+\subsection{Abstract properties of  
+\texorpdfstring{$H'_\alpha$}{H'}}
 ~\label{sect:H-alpha-prop}
 
-Since pure computation seems to be useless for dealing with expressions of the form $H_\alpha(k)$, even for small values of $\alpha$ and $k$, we need to prove theorems for comparing $H_\alpha(k)$ and $H_\beta(l)$, in terms of comparison
+Since pure computation seems to be useless for dealing with expressions of the form $H'_\alpha(k)$, even for small values of $\alpha$ and $k$, we need to prove theorems for comparing $H'_\alpha(k)$ and $H'_\beta(l)$, in terms of comparison
 between $\alpha$ and $\beta$ on the one hand, $k$ and $l$ on the other hand.
 
-But beware of non-theorems! For instance, one could believe that $H$ is monotonous in its first argument. The following proof shows this is false.
-
-\begin{Coqsrc}
-Remark H_non_mono1 :
+But beware of non-theorems! For instance, one could believe that $H'$ is monotonous in its first argument. The following proof shows this is false.
+
+\begin{Coqsrc}
+Remark H'_non_mono1 :
   ~ (forall alpha beta k, (alpha o<= beta)%e0 ->
-                          (H_ alpha k <= H_ beta k)%nat).
+                          (H'_ alpha k <= H'_ beta k)%nat).
 Proof.
@@
  assert (H0 :(42 o<= omega)%e0) by (repeat split; auto).  
- apply H in H0; rewrite H_Fin, H_omega  in H0; lia.
+ apply H in H0; rewrite H'_Fin, H'_omega  in H0; lia.
 Qed.
@@
 
-On the contrary, the fonctions of the Hardy hierarchy have the following five properties~\cite{KS81}: for any $\alpha < \epsilon_0$,
+On the contrary, the fonctions of the $H'$ hierarchy have the following five properties~\cite{KS81}: for any $\alpha < \epsilon_0$,
 \begin{itemize}
-\item the function $H_\alpha$ is strictly monotonous :
-      For all $n,p \in\mathbb{N}, n < p \Rightarrow H_\alpha(n)< H_\alpha(p)$.
-\item If $\alpha \not= 0$, then for every $n$, $n<H_\alpha(n)$.
-\item The function $H_\alpha$ is pointwise less or equal than $H_{\alpha+1}$
-
-\item For any $n\geq 1$, $H_\alpha(n)<H_{\alpha+1}(n)$.
-\emph{We say that $H_{\alpha+1}$ dominates $H_\alpha$ from $1$}.
+\item the function $H'_\alpha$ is strictly monotonous :
+      For all $n,p \in\mathbb{N}, n < p \Rightarrow H'_\alpha(n)< H'_\alpha(p)$.
+\item If $\alpha \not= 0$, then for every $n$, $n<H'_\alpha(n)$.
+\item The function $H'_\alpha$ is pointwise less or equal than $H'_{\alpha+1}$
+
+\item For any $n\geq 1$, $H'_\alpha(n)<H'_{\alpha+1}(n)$.
+\emph{We say that $H'_{\alpha+1}$ dominates $H'_\alpha$ from $1$}.
 \item For any $n$ and $\beta$, if $\alpha \xrightarrow[n]{} \beta$, then
-$H_\beta(n)\leq H_\alpha(n)$.
+$H'_\beta(n)\leq H'_\alpha(n)$.
 \end{itemize}
@@
     mkP {
-        PA : strict_mono (H_ alpha);
-        PB : alpha <> Zero -> forall n,  (n < H_ alpha n)%nat;
-        PC : H_ alpha <<= H_ (Succ alpha);
-        PD : dominates_from 1 (H_ (Succ alpha)) (H_ alpha);
+        PA : strict_mono (H'_ alpha);
+        PB : alpha <> Zero -> forall n,  (n < H'_ alpha n)%nat;
+        PC : H'_ alpha <<= H'_ (Succ alpha);
+        PD : dominates_from 1 (H'_ (Succ alpha)) (H'_ alpha);
         PE : forall beta n, Canon_plus n alpha beta -> 
-                            (H_ beta n <= H_ alpha n)%nat}.
+                            (H'_ beta n <= H'_ alpha n)%nat}.
 
@@
 
-\subsection{Comparison between \texttt{L\_} and \texttt{H\_} }
-
-By well-founded induction on $\alpha$, we prove that our $L$ hierachy is ``almost'' the Hardy hierarchy (up to a small shift).
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#H_L_}{Epsilon0.L\_alpha}}
-
-\begin{Coqsrc}
- Theorem H_L_ alpha: forall i:nat,  (H_ alpha i <= L_ alpha (S i))%nat.
+\subsection{Comparison between \texttt{L\_} and \texttt{H'\_} }
+
+By well-founded induction on $\alpha$, we prove a simple relation between $L_\alpha$ and $H'\_alpha$.
+
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#H'_L_}{Epsilon0.L\_alpha}}
+
+\begin{Coqsrc}
+ Theorem H'_L_ alpha: forall i:nat,  (H'_ alpha i <= L_ alpha (S i))%nat.
 \end{Coqsrc}
@@
 
-The following theorem relates the length of (standard) battles with the Hardy hierarchy.
+The following theorem relates the length of (standard) battles with the the $H'$ family of fast growing functions.
 
@@
   forall k , 1 <= k -> exists l: nat,  
-       H_ alpha k - k <= l /\
+       H'_ alpha k - k <= l /\
        battle_length standard k (iota (cnf alpha)) l.    
@@
 
-\index{hydras}{Exercises}
-\begin{exercise}
-Let us quote a theorem from ~\cite{KS81} (page 297).
-
-\begin{quote}
-\begin{align*}
-  H_{\omega^\alpha}(n+1) &\geq F_{\alpha}(n) \quad (n\geq 1, \alpha<\epsilon_0) \\
- F_{\alpha}(n+1) &\geq H_{\omega^\alpha}(n) \quad (n\geq 1, \alpha<\epsilon_0) 
-\end{align*}
-Thus $H_{\omega^\alpha}$ and $F_{\alpha}$ have essentially the same order of growth.
-
-\end{quote}
-
- But, before trying to prove these facts, look at the definition of function $H$ in Ketonen and Solovay's paper !
+%%% TODO : Check the statement !
+%
+% \index{hydras}{Exercises}
+% \begin{exercise}
+% Let us quote a theorem from ~\cite{KS81} (page 297).
+
+% \begin{quote}
+% \begin{align*}
+%   H'_{\omega^\alpha}(n+1) &\geq F_{\alpha}(n) \quad (n\geq 1, \alpha<\epsilon_0) \\
+%  F_{\alpha}(n+1) &\geq H'_{\omega^\alpha}(n) \quad (n\geq 1, \alpha<\epsilon_0) 
+% \end{align*}
+% Thus $H'_{\omega^\alpha}$ and $F_{\alpha}$ have essentially the same order of growth.
+
+% \end{quote}
+
+%  But, before trying to prove these facts, look at the definition of function $H$ in Ketonen and Solovay's paper !
 Is it really the same as the definition we quote from Pr{\H o}mel's chapter~\cite{Promel2013},
-whereas \cite{KS81} define $H_\alpha(n)$ as ``the least integer $k$ such that $[n,k]$ is $\alpha$-large''. Thus, it may be useful to adapt the statement above.
-
-\end{exercise}
+% whereas \cite{KS81} define $H_\alpha(n)$ as ``the least integer $k$ such that $[n,k]$ is $\alpha$-large''. Thus, it may be useful to adapt the statement above.
+
+% \end{exercise}
 

### 2021-05-20 doc/chapter-primrec.tex
more transparent proofs of predicate isPR

@@
 
-\section{Proving that a given function is \emph{not} primitive recursive}
+\section{Ackermann: a non  primitive recursive function}
 \label{sect:ack-not-PR}
@@
 
-\subsection{The length of standard hydra battles}
+\section{The length of standard hydra battles}
 
@@
 
-\subsubsection{Definitions}
+\subsection{Definitions}
 
@@
 
-\subsubsection{Proof steps}
+\subsection{Main proof steps}
 
@@
 \begin{Coqsrc}
-Section battle_lenght_notPR.
+Section battle_length_notPR.
 
@@
 \begin{Coqsrc}
-Let alpha := Phi0 omega%e0.
-Let h := iota (cnf alpha).
+Let delta := Phi0 omega%e0.
+Let h := iota (cnf delta).
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Let m k := L_ alpha (S k).
-
-Remark m_eqn : forall k, m k = (l_std alpha k + k)%nat.
+Let m k := L_ delta (S k).
+
+Remark m_eqn : forall k, m k = (l_std delta k + k)%nat.
 (* ... *)
@@
 
-\emph{Please note that our proof of this lemma is not trivial at all. It uses some lemmas of Chapters~\vref{chap:ketonen} and \vref{chap:alpha-large}.}
+\emph{Please note that our proof of this lemma is not trivial at all. It uses some lemmas of Chapters~\vref{chap:ketonen} and \vref{chap:alpha-large}. Mainly:}
+
+\begin{Coqanswer}
+H'_restricted_mono_l:
+  forall (alpha beta : E0) (n : nat),
+  Canon_plus n alpha beta -> H'_ beta n <= H'_ alpha n
+\end{Coqanswer}
+
+This lemma is proven in 
+\href{../theories/html/hydras.Epsilon0.Hprime.html}{Epsilon0.Hprime}, with the help of the \emph{Ketonen-Solovay machinery}.
 
@@
 \begin{Coqsrc}
-H_L_ : forall (alpha : E0) (i : nat), H_ alpha i <= L_ alpha (S i)
+H'_L_: forall (alpha : E0) (i : nat), 
+      H'_ alpha i <= L_ alpha (S i)
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-F_vs_Ack : forall n : nat, 2 <= n -> Ack n n <= F_ omega n
+F_vs_Ack: forall n : nat, 2 <= n -> Ack n n <= F_ omega n
 \end{Coqsrc}
@@
 
-End battle_lenght_notPR.
-\end{Coqsrc}
+End battle_length_notPR.
+\end{Coqsrc}

### 2021-05-21 doc/chapter-primrec.tex
corrections to the pdf

--- 2021-05-14 doc/chapter-primrec.tex
@@
 \item[composition] For any $n$ and $m$, if $h: \mathbb{N}^m\rightarrow\mathbb{N}$, and
-$g_0,\dots, g_{m-1}$ are primitive recursive of $n$ arguments, then the function which maps any
-tuple $(x_0,\dots,x_{n-1})$ to $h(g_0(x0,\dots,x_{n-1}),\dots, g_{m-1}(x0,\dots,x_{n-1}))$ is primitive recursive.
+$g_0,\dots, g_{m-1}: \mathbb{N}^n\rightarrow\mathbb{N}$ are primitive recursive of $n$ arguments, then the function which maps any
+tuple $(x_0,\dots,x_{n-1})$ to $h(g_0(x0,\dots,x_{n-1}),\dots, g_{m-1}(x0,\dots,x_{n-1})): \mathbb{N}^n\rightarrow\mathbb{N}$ is primitive recursive.
 \item[primitive recursion]
@@
 than 45K lines of scripts. The part dedicated to primitive recursive functions and Peano arithmetics is 32K lines long and is originally structured in 38 modules.
-Thus, we propose a slow exploration of this library, through examples and exercises. Our additions to the original library 
+Thus, we propose a partial exploration of this library, through examples and exercises. Our additions to the original library --- mainly examples and counter-examples ---,
 are stored in the directory \texttt{theories/ordinals/MoreAck}.
-In particular, the library~\href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR} contains the well-known  proof that the Ackermann function is not primitive recursive (see Section~\vref{sect:ack-not-PR}).
+
+In particular, the library \href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR} contains the well-known  proof that the Ackermann function is not primitive recursive (see Section~\vref{sect:ack-not-PR}).
+MoreOver, the library \href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems} contains 
+a proof that the length of an hydra battle (according to the initial replication factor) is not primitive recursive in general.
 
@@
 
-Library \texttt{primRec} allows us to consider primitive functions on \texttt{nat}, with any number of arguments, in 
+The  \texttt{primRec} library allows us to consider primitive functions on \texttt{nat}, with any number of arguments, in 
 curried form. This is made possible in 
@@
 \noindent
-\emph{From\href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}}.
+\emph{From \href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}}.
 \begin{Coqsrc}
@@
  This fact is reported in the comments of \texttt{primRec.v}.
-We presume that this convention makes it easier to define the evaluation function \texttt{evalProjFunc $n$} (see the next sub-section).
+We presume that this convention makes it easier to define the evaluation function (\texttt{evalProjFunc $n$}) (see the next sub-section).
 Trying the other convention is left as an exercise.
@@
 $\texttt{PrimRec}\,n \rightarrow  \texttt{naryFunc}\,n$.
-This function is definied by mutual recursion,  together with the  function 
+This function is defined by mutual recursion,  together with the  function 
 (\texttt{evalprimRecS $n$ $m$}) of type 
@@
 Another example?
-Let us consider the following term\footnote{Of course, we never typed this term \emph{verbatim}; we obtained it as the result of some computation we leave it the reader to guess and reproduce.}:
+Let us consider the following term\footnote{Of course, we never typed this term \emph{verbatim}; we obtained it by an interactive proof the reader will be able to make after 
+reading Sect.\vref{sect:proofs-of-isPR}.}:
 
@@
 \section{Proving that a given arithmetic function is primitive recursive}
+\label{sect:proofs-of-isPR}
 
 The example in the preceding section clearly shows that, in order to prove that a given arithmetic function
-(defined in \gallina{} as usual) is primitive recursive, trying to give  by hand a term  of type \texttt{PrimRec $n$} is not a good method, since such terms may be huge and complex, even for simple arithmetic functions.
+(defined in \gallina{} as usual) is primitive recursive, trying to give  by hand a term  of type (\texttt{PrimRec $n$}) is not a good method, since such terms may be huge and complex, even for simple arithmetic functions.
 The method proposed in Library \texttt{primRec} is the following one:
@@
 If it does not cause compatibility problems (with 
-\href{https://github.com/coq-community/goedel}{goedel library} fr instance), we plan to make all theses lemmas transparent.
+\href{https://github.com/coq-community/goedel}{goedel library} for instance), we plan to make all theses lemmas transparent.
 \end{remark}
@@
 
-\textbf{Hint:}  You may use as a helper the function which computes the pair 
+\textbf{Hint:}  You may use as a helper the function which computes the pair \linebreak
 $(\texttt{fib}(n+1),\texttt{fib}(n))$. 
@@
 the associated constructor and projections are primitive recursive.
-
 \emph{You may start this exercise with the file
@@
 
-A possible workaround is to make \\texttt{m} be the 
-decreasing argument, and define --- within \texttt{m}'s scope --- a local helper function which computes \texttt{Ack m n} for any \texttt{n}.
+A possible workaround is to make \texttt{m} be the 
+decreasing argument, and define --- within \texttt{m}'s scope --- a local helper function which computes (\texttt{Ack m n}) for any \texttt{n}.
 This way, both functions \texttt{Ack} and \texttt{Ackm} have a (structurally) strictly decreasing argument.
@@
 The file \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack} presents two other definitions\footnote{post on \href{https://stackoverflow.com/questions/10292421/error-in-defining-ackermann-in-coq} by Anton Trunov.} of the Ackermann functions based on the lexicographic ordering on $\mathbb{N}\times\mathbb{N}$.
-Prove that all the four functions are extensionnally equal.
+Prove that the four functions are extensionnally equal.
 \end{exercise}
@@
 
+\noindent
+\emph{From \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack}}.
 \begin{Coqsrc}
@@
 
-First, we prove that any primitive function of two arguments is majorized.
+First, we prove that any primitive function of two arguments is majorized by \texttt{Ack}.
 If we look at the inductive definition of primitive recursive functions, page~\pageref{def:Primrec}, it is obvious that a proof by induction on the construction of primitive recursive functions must consider functions of any arity.
 
+The following scheme allows us to write proofs by induction on the class of primitive recursive functions.
+
 \vspace{4pt}
-
 \noindent
@@
 
-* projections *)
+(* projections *)
 (forall (n m : nat) (l : m < n), P n (projFunc n m l)) ->
@@
 \href{../theories/html/hydras.Prelude.MoreVectors.html}{Prelude.MoreVectors}).
-So, \texttt{t\,$A$\,$n$} is the type of vectors of $n$ elements of type $A$, and the constants \texttt{cons}, \texttt{nil}, \texttt{map}, etc., refer to vectors and not to lists.
+So, (\texttt{t\,$A$\,$n$}) is the type of vectors of $n$ elements of type $A$, and the constants \texttt{cons}, \texttt{nil}, \texttt{map}, etc., refer to vectors and not to lists.
 Likewise, the notation \texttt{x::v} is an abbreviation for
@@
 $(f\,x_1\,x_2\,\dots\,x_n)$ into an application of a function 
-to a single argument: the vector of all the $x_i$s.
-This is already done in 
+to a single argument: the vector of all the $x_i$\,s.
+This is already defined in 
 Library~\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.
@@
 Now, it remains to prove that any primitive function is majorized by \texttt{Ack}.
-The three base cases of construction of primitive recursive function are as follows:
+The three base cases  are as follows:
 
@@
 
-\subsection{The length of standard hydra battles}
-
-The module \href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra\_Theorems} contains a proof that the function which computes the lenght of standard hydra battles is not primitive recursive.
+\section{The length of standard hydra battles}
+
+The module \href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra\_Theorems} contains a proof that the function which computes the length of standard hydra battles is not primitive recursive.
+More precisely, we consider, for a given hydra $h=\iota(\alpha)$, the length of a standard battle which starts with the replication factor $k$ (see Sect~\vref{def:L-alpha}).
+
 This proof is  a little more complex than the preceding one.
 
-\subsubsection{Definitions}
+\subsection{Definitions}
 
@@
 
-\subsubsection{Proof steps}
+\subsection{Proof steps}
 
@@
 
-Now, let us recall the following lemmas :
-
-\begin{enumerate}
-\item 
-\emph{From \href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}}
-
-\begin{Coqsrc}
-H_F : forall (alpha : E0) (n : nat), 
-     F_ alpha (S n) <= H_ (Phi0 alpha) (S n) 
-\end{Coqsrc}
-
-\emph{Please note that our proof of this lemma is not trivial at all. It uses some lemmas of Chapters~\vref{chap:ketonen} and \vref{chap:alpha-large}.}
-
-\item
-\noindent
+\subsubsection{Comparison between $F$ and $H'$}
+
+In \href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}, we prove a relation between the $F$ and $H'$ functional. For any $\alpha$ and $k>0$,
+$H'_{\omega^\alpha}(k)\geq F_\alpha(k)$.
+
+\begin{Coqsrc}
+H'_F : forall (alpha : E0) (n : nat), 
+     F_ alpha (S n) <= H'_ (Phi0 alpha) (S n) 
+\end{Coqsrc}
+
+Our proof of this lemma is not trivial at all. One of it sub-goals is the following one:
+
+\begin{Coqsrc}
+  alpha : E0
+  IHalpha : forall beta : E0, beta o< alpha -> P beta
+  Halpha : Limitb alpha
+  n : nat
+  ============================
+  H'_ (Phi0 (CanonS alpha n)) (S n) <= 
+  H'_ (Phi0 (CanonS alpha (S n))) (S n)
+\end{Coqsrc}
+
+In mathematical notation: $H'_{\omega^{\canonseq{\alpha}{n}}}(n+1) \leq
+H'_{\omega^{\canonseq{\alpha}{n+1}}}(n+1)$.
+
+\vspace{4pt}
+
+But there exists no lemma saying that, if 
+$\beta\leq \alpha$, then $H'_\beta(k)\leq H'_\alpha(k)$, for any $\alpha$ and $\beta$. For instance, 
+$H'_{42}(3)=45> H'_\omega(3)=7$.
+
+Looking for lemmas of the form $H'_\beta(k)\leq H'_\alpha(k)$, we find this one (from our library
+\href{../theories/html/hydras.Epsilon0.Hprimehtml}{Epsilon0.Hprime}):
+
+\begin{Coqanswer}
+H'_restricted_mono_l : 
+    forall (alpha beta : E0) (n : nat), 
+      Canon_plus n alpha beta -> 
+      H'_ beta n <= H'_ alpha n.
+\end{Coqanswer}
+
+Thus, it remains to prove that 
+there exists a path from ${\omega^{\canonseq{\alpha}{n+1}}}$
+to ${\omega^{\canonseq{\alpha}{n}}}$ composed of 
+$n+1$-steps.
+
+Fortunately, the Ketonen-Solovay machinery contains three lemmas which help us to build such a path.
+
+\begin{Coqanswer}
+KS_thm_2_4_lemma5 :
+  forall [i : nat] [alpha beta : T1],
+  const_pathS i alpha beta ->
+  nf alpha -> alpha <> zero -> 
+  const_pathS i (phi0 alpha) (phi0 beta)
+
+KS_thm_2_4 :
+  forall [lambda : T1], nf lambda ->limitb lambda ->
+  forall i j : nat, i < j -> 
+   const_pathS 0 (canonS lambda j) (canonS lambda i)
+
+Cor12_1 :
+forall [alpha : T1], nf alpha ->
+      forall (beta : T1) (i n : nat),
+      beta t1< alpha ->
+     i <= n -> const_pathS i alpha beta -> 
+     const_pathS n alpha beta
+\end{Coqanswer}
+
+\subsubsection{End of the proof}
+
+We finish the proof by comparing several fast growing functions.
+
 \emph{From \href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0.L\_alpha}}
@@
 \begin{Coqsrc}
-H_L_ : forall (alpha : E0) (i : nat), H_ alpha i <= L_ alpha (S i)
-\end{Coqsrc}
-
-\item
+H'_L_ : forall (alpha : E0) (i : nat), H'_ alpha i <= L_ alpha (S i)
+\end{Coqsrc}
 
@@
 \end{Coqsrc}
-
-\end{enumerate}
 

### 2021-05-25 doc/part-hydras.tex
minor changes
corrections to the pdf
correction of typos

@@
 \begin{remark}
-Some notations (mainly names of fast-growing functions) of our development may differ slightly from the litterature. Although this fact does not affect our proofs, we are preparing a future 
-version where the names $F\_alpha$, $f\_alpha$, $H\_alpha$, etc., are fully consistent with the cited articles. 
+Some notations (mainly names of fast-growing functions) of our development may differ slightly from the litterature.
+Although this fact does not affect our proofs, we are preparing a future version where the names $F\_alpha$, $f\_alpha$, $H\_alpha$, etc., are fully consistent with the cited articles.
+
 \end{remark}
@@
 \begin{definition}
-Let $0<\alpha<\epsilon_0$ be any ordinal, and $s$ a finite sequence $\langle s_1, s_2, \dots, s_N\rangle$ of strictly positive natural numbers.
+Let $0<\alpha<\epsilon_0$ be any ordinal, and $s=\langle s_1, s_2, \dots, s_N\rangle$ a finite sequence of strictly positive natural numbers.
 
 We say that $s$ is \emph{$\alpha$-large} if the sequence $\langle \alpha_0=\alpha,\dots,\alpha_{i+1}=\canonseq{\alpha_i}{i+1},\dots \rangle$ leads to $0$. 
-
-We say that $s$ is \emph{minimally $\alpha$-large} (in short:
+We say also that $s$ is \emph{minimally $\alpha$-large} (in short:
 \emph{$\alpha$-mlarge}) if $s$ is $\alpha$-large 
@@
 The following function ``gnaws'' an ordinal $\alpha$, following a sequence of indices (ignoring the $0$s).
+
+\vspace{4pt}
+
 \noindent
@@
 Let us consider two integers $k$ and $l$, such that $0<k<l$. In order to check whether the interval $[k,l]$ is minimally large for $\alpha$, it is enough to
-follow from $\alpha$ the path associated with the interval $[k,l($ and verify that the last ordinal we obtain is equal to $1$.
+follow from $\alpha$ the path associated with the interval $[k,l)$ and verify that the last ordinal we obtain is equal to $1$.
 
@@
 
-Now, consider any natural number $k>0$. We would like to compute
+Now, consider any natural number $k>0$ any ordinal $0<\alpha<\epsilon_0$.  We would like to compute
 a number $l$ such that the interval $[k,l]$ is $\alpha$-mlarge. So, 
@@
 \begin{Coqbad}
-Function L (alpha: E0) (i:nat) :  nat := ...
+Function L_ (alpha: E0) (i:nat) :  nat := ...
 \end{Coqbad}
@@
 
-Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider any  function which associates maps  any strictly positive integer $k$ to the number $l$, where 
+Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider any  function which  maps  any strictly positive integer $k$ to the number $l$, where 
 the interval $[k,l)$ is $\alpha$-mlarge.
@@
 
+\begin{todo}
+ Check if the functions $L_\alpha$ are the same as
+\cite{KS81}' functions $f\_alpha$ (p. 297).
+\end{todo}
+
 Note that, for $\alpha\not=0$, the value of $f(0)$ is not specified.
@@
 We are looking for properties which could be used for writing \emph{equations} and prove the correctness of the function generated by the \texttt{coq-equations} plug-in.
-Moreover, they will give us some examples of
-$L_\alpha$ for small values of $\alpha$.
-
-Our exploration of the $L_\alpha$s  follows the usual scheme : transfinite induction, and proof-by-cases : zero, successors and limit ordinals.
+Moreover, they will give us some examples (for small values of $\alpha$).
+
+Our exploration of the $L_\alpha$\,s  follows the usual scheme : transfinite induction, and proof-by-cases : zero, successors and limit ordinals.
 
@@
 Let $\lambda<\epsilon_0$ be any limit ordinal. In a similar way as for successors, we decompose any path from $\lambda$  into a first step to
-$\canonseq{\lambda}{k}$, followed by a path to $0$. In the following section, we assume that there exists à correct function for $\canonseq{\lambda}{k}$, \emph{for any strictly positive $k$}.
+$\canonseq{\lambda}{k}$, followed by a path to $0$. In the following section, we assume that there exists a correct function computing  $L_{\canonseq{\lambda}{k}}$ for any strictly positive $k$.
 
@@
 We would like to get exact formulas for the ordinal $\omega^2$, a.k.a.
-$\phi_0(2)$. This ordinal is the limit of the sequence $\omega\times i\;(i \in \mathbb{N}$. Thus, we have to study ordinals of this form, then use 
+$\phi_0(2)$. This ordinal is the limit of the sequence $\omega\times i\;(i \in \mathbb{N})$. Thus, we have to study ordinals of this form, then use 
 our lemma on limits.
@@
 Note that in this lemma, we assume that the fonction associated with $\alpha$ is stritly monotonous and
-greater or equal than the successor function, and prove that $L_{\omega^\alpha\times i}$satisfies  the same properties.
+greater or equal than the successor function, and prove that $L_{\omega^\alpha\times i}$ satisfies  the same properties.
 
@@
 \subsubsection{Back to hydra battles}
+\label{def:L-alpha}
 
@@
 \begin{Coqsrc}
-Lemma H'_Omega_term (alpha : E0)  :
+Lemma H'_Omega_term (alpha : E0):
 alpha <> Zero -> forall i k, 
@@
 \end{exercise}
+
+\section{Conclusion}
+
+In Sect.~\vref{sect:battle-length-notPR}, we prove that the length of hydra-battles (for a given hydra, according to the initial replication factor) is not primitive recursive in general. 
+This proof uses properties of the Ackermann function, and the $H'_\alpha$, $F_\alpha$, $L_\alpha$ families of functions.
 

### 2021-05-26 doc/part-hydras.tex
corrections of typos

@@
 Since the canonical sequence of $\omega$ is composed of finite ordinals, 
-it is easy to get the formula associated with $H_\omega$.
+it is easy to get the formula associated with $H'_\omega$.
 
@@
 
-\item In particular, we have got a formula for $H'_{\omega^{i+1}}$.
+\item In particular, we derive a formula for $H'_{\omega^{i+1}}$.
 
@@
 Fact F0 : H'_ (Phi0 3) 3 = f (f (f (f 3))).
- rewrite R0; reflexivity. 
-Qed.
+proof. rewrite R0; reflexivity. Qed.
 \end{Coqsrc}
@@
 
-This number looks quite big; let us compute an approximation in \texttt{Ocaml}:
+This number looks quite big; let us compute an approximation with \texttt{Ocaml}:
 
@@
 
-But beware of non-theorems! For instance, one could believe that $H'$ is monotonous in its first argument. The following proof shows this is false.
+But beware of fake theorems! For instance, one could believe that $H'$ is monotonous in its first argument. The following proof shows this is false.
 
@@
 
+\vspace{4pt}
+
+\noindent
 \emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0.L\_alpha}}
@@
 The difference with the ``classic'' Wainer hierarchy 
-$f\_alpha\;(\alpha<\epsilon_0)$ lies in the second equation:
+$f\_\alpha\;(\alpha<\epsilon_0)$ lies in the second equation:
 $f_{\beta+1}(i) = (f_\beta)^{(i)}(i)$ and not

### 2021-06-01 doc/part-hydras.tex
correction of typos

@@
 Thus, for instance, $L_{\omega^3}(3)=L_{\omega^2\times 4}(3)$.
-Thus, we obtain an exact expression of this number.
 
@@
 This number is quite big. Using \texttt{Ocaml}'s \texttt{float} arithmetic,
-we can [under-]approximate it by $2^{3.8\times10^{30}}\times 3.8\times{10^{30}}$.
+we can under-approximate it by $2^{3.8\times10^{30}}\times 3.8\times{10^{30}}$.
 
@@
 
-In order to give a feeling on  the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the so called \emph{Wainer-Hardy hierarchy} of fast growing functions,
+In order to give a feeling on  the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the  \emph{Wainer-Hardy hierarchy} of fast growing functions,
 presented for instance in~\cite{Promel2013}. 
@@
 In the future versions of this development, we will correct the references to the litterature.
-For the time being, we call our functions $H'\_alpha$ in order to underline the difference from ``classic'' Hardy functions.
+For the time being, we call our functions $H'_\alpha$ in order to underline the difference from ``classic'' Hardy functions.
 \end{remark}
@@
 Section Proof_of_Abstract_Properties.
+\end{Coqsrc}
+
+By elimination, we get a catalogue of properties of the functions $L'_\alpha$:
+
+\begin{Coqsrc}
+Section Abstract_Properties.
+ Variable alpha : E0.
+
+ Theorem H'_alpha_mono : strict_mono (H'_ alpha).
+
+ Theorem H'_alpha_gt : alpha <> Zero ->
+        forall n, n < H'_ alpha n.
+
+ Theorem H'_alpha_Succ_le :
+        H'_ alpha <<= H'_ (Succ alpha).
+
+ Theorem H'_alpha_dom :
+    dominates_from 1 (H'_ (Succ alpha)) (H'_ alpha).
+
+ Theorem H'_restricted_mono_l :
+    forall beta n, Canon_plus n alpha beta -> 
+                                 H'_ beta n <= H'_ alpha n.
+
+  Lemma H'_alpha_ge_id : id <<= H'_ alpha.
+
+  Lemma H'_alpha_mono_weak : forall k l, k <= l ->
+                                        H'_ alpha k <= H'_ alpha l.
+
+End Abstract_Properties.
+\end{Coqsrc}
+
+Using a few lemmas \emph{à la} Ketonen-Solovay, we prove that
+if $\alpha<\beta$, then $H'_\beta$ eventually dominates
+$H'_\alpha$.
+We let the reader look at its proof (Section \texttt{Proof\_of\_H'\_mono\_l} of \href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}).
+
+\noindent
+\begin{Coqsrc}
+  H'_dom : forall alpha beta : E0, alpha o< beta ->
+  dominates_strong (H'_ beta) (H'_ alpha).
 \end{Coqsrc}
@@
 % \end{Coqsrc}
+
+%------------------------------------------------------
+\section{A certified catalogue of rapidly growing functions}
+
+In this section, we try to present an abstract of the properties of the main variants of fast-growing hierarchies of functions
+we fauund in the litterature or we had to define as helpers in our proofs.
+
+\subsection{Ketonen-Solovay's \texttt{F\_alpha}}
+
+  This hierarchy is presented p.280 of~\cite{KS81} as `` a trivial variant of one introduced by Wainer ''.
+  Let us recall the equations shown in Sect.~\vref{F_equations}.
+
+\begin{itemize}
+\item $F_0(i)=i+1$
+\item $F_{\beta+1}(i)= (F_\beta)^{(i+1)}(i)$, where $f^{(i)}$ is the $i$-th iterate of $f$.
+\item $F_\alpha(i) = F_{\canonseq{\alpha}{i}} (i)$ if $\alpha$ is a limit ordinal.
+\end{itemize}
+
+Note that \cite{KS81} defines also $F_{\epsilon_0}$ (by the third equation).
+Since $\epsilon_0$ is not representable in type \texttt{E0}, our translation in \coq{} does not take $F_{\epsilon_0}$ into account.
+
+Several properties of $F_\alpha$ are already presented in Sect.~\ref{F_equations}.
 

### 2021-06-04 doc/part-hydras.tex
update Definition to On_Generic

@@
 
+\vspace{4pt}
+
 Let us have a look at Figure~\ref{fig:ordinal-sequence}. It presents a few items of a  sequence of ordinal numbers, which extends the sequence of natural numbers.
@@
 We use it as a reference for proving the correctness of ordinal notations.
-\item A family of \emph{ordinal notations}, \emph{i.e.} data types used to represent segments $[0,\mu)$, where $\mu$ is some countable ordinal.
+\item A family of \emph{ordinal notations} (also called \emph{[ordinal] notation systems}), \emph{i.e.} data types used to represent segments $[0,\mu)$, where $\mu$ is some countable ordinal.
 Each ordinal notation is defined inside the Calculus of Inductive Constructions (without axioms).
@@
 Fortunately, the ordinals we need for  studying hydra battles are much simpler than Schütte's, and can be represented as quite simple data types in \gallina.
-So, we will use \emph{ordinal notations} (also called \emph{[ordinal] notation systems}).
 
@@
 \vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}}
+\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}}
 
@@
 
-Notation "'omega'" := (inr  0:t).
+Notation omega := (inr  0:t).
 \end{Coqsrc}
@@
 
-Notation "'omega'" := (1,0) : ON_scope.
+Notation omega := (1,0).
 
@@
 \node (Bdots) at (6,2) {$\ldots$};
-\node (b) at (8,2) [label=above:$\alpha$]{$\bullet$};
-\node (bsucc) at (9,2) [label=above:$\alpha+1$]{$\bullet$};
+\node (b) at (8,2) [label=above:$\omega$]{$\bullet$};
+\node (bsucc) at (9,2) [label=above:$\omega+1$]{$\bullet$};
 \node (Bdots2) at (10,2) {$\ldots$};
@@
 \begin{Coqsrc}
-Notation "'FS' n" :=
-     (ocons zero n zero) (at level 10) : t1_scope.
+Notation one := (ocons zero 0 zero).
+
+Notation FS n := (ocons zero n zero).
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Notation omega := (ocons (ocons zero 0 zero) 0 zero): t1_scope.
+Notation omega := (ocons (ocons zero 0 zero) 0 zero).
 \end{Coqsrc}
@@
   | P_omega : ppT1
+
+Notation _omega := P_omega.
+
 \end{Coqsrc}
@@
 \begin{Coqanswer}
-     = (omega ^ omega + omega ^ 3 * 5 + 2)%pT1
+     = (_omega ^ _omega + _omega ^ 3 * 5 + 2)%pT1
      : ppT1
@@
 \begin{Coqsrc}
-Fixpoint compare (alpha alpha':T1):comparison :=
-  match alpha, alpha' with
+Fixpoint compare (alpha beta:T1):comparison :=
+  match alpha, beta with
     zero, zero => Eq
@@
        | Gt => Gt
-          | Eq => (match lt_eq_lt_dec n n'
-                   with
-                       inleft  (left _) => Lt
-                     | inright _ => Gt
-                     |   _ => compare b b'
+       | Eq => (match Nat.compare n n' with
+                | Eq => compare b b'
+                | comp => comp
                 end)
@@
 
-Notation "'omega'"  := _Omega : E0_scope.
+Notation omega  := _Omega.
 \end{Coqsrc}
@@
 
+\index{hydras}{Exercises} 
+\begin{exercise}
+  Prove that that the ordinal addition on \texttt{E0} is associative.
+
+  \textbf{Warning:}  Please do not import theorems from
+  \texttt{ordinals/Schutte/}, since that library uses axioms!
+\end{exercise}
+
 \section{Well-foundedness and transfinite induction}
@@
 The class \texttt{ON} of ordinal notations has been defined long after this 
-chapter, and is not used yet in the development of the type \texttt{E0}. 
+chapter, and is not used in the development of the type \texttt{E0} yet.
 A better integration of both notions should simplify the development on ordinals in Cantor normal form. This integration is planned for the future versions.
@@
 
-\section{A Variant for hydra battles}
+\index{hydras}{Projects}
+\begin{project}
+  Prove that ordinal multiplication is left-distributive over ordinal addition; that is:
+  $$\forall \alpha \, \beta \, \gamma < \epsilon_0, \alpha\times(\beta+\gamma) = \alpha\times\beta+ \alpha\times \gamma$$
+
+\end{project}
+\section{A variant for hydra battles}
 
@@
 uses plain structural recursion over the type \coqsimple{T1}. Thus, tactics like
-\coqsimple{cbn}, \coqsimple{simpl}, \coqsimple{compute}, etc., are applicable. For simplicity's sake, 
-we use an auxiliary function \texttt{canonS} of type \texttt{T1 -> nat  -> T1} such that
-(\texttt{canonS  $\alpha$ $i$}) is equal to $\canonseq{\alpha}{i+1}$.
+\coqsimple{cbn}, \coqsimple{simpl}, \coqsimple{compute}, etc., are applicable.
 
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS}{Epsilon0.Canon}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canon}{Epsilon0.Canon}}
 
@@
 \begin{Coqsrc}
-Fixpoint canonS alpha (i:nat) :=
+  Fixpoint canon alpha (i:nat) : T1 :=
   match alpha with
@@
   | ocons zero (S k) zero  => FS k
-    | ocons gamma 0 zero =>
-      match pred gamma with
-          Some gamma' => ocons gamma' i zero
-        | None => ocons (canonS gamma i) 0 zero
+  | ocons gamma 0 zero => (match T1.pred gamma with
+                            Some gamma' =>
+                            match i with 0 => zero
+                                    |   S j => ocons gamma' j zero
                             end
+                          | None =>
+                            ocons (canon gamma i) 0 zero
+                           end)
+
   |  ocons gamma (S n) zero =>
-       match pred gamma with
-           Some gamma' => ocons gamma n (ocons gamma' i zero)
-         | None => ocons gamma n (ocons (canonS gamma i) 0 zero)
-       end
-    | ocons alpha n beta => ocons alpha n (canonS beta i)
+     (match T1.pred gamma with
+       Some gamma' =>
+       (match i with
+         0 =>  ocons gamma n zero
+       | S j => ocons gamma n (ocons gamma' j zero)
+       end)
+       | None =>
+            ocons gamma n (ocons (canon gamma  i) 0 zero)
+      end)
+   |  ocons alpha n beta => ocons alpha n (canon beta i)  
 end.
-\end{Coqsrc}
-
-The following function computes $\canonseq{\alpha}{i}$, except for the case $i=0$, where it simply returns $0$\;\footnote{This restriction did not prevent us from proving all the main theorems of~\cite{KS81, KP82}.
-Nevertheless, in a future version of this development, we may define $\canonseq{\alpha}{0}$ exactly as
-in~\cite{KS81}. But we are afraid this would  be done at the cost of making some proofs much more complex.}.
-
-\begin{Coqsrc}
- Definition canon alpha i := 
-   match i with 0 => zero | S j => canonS alpha j end. 
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Compute (canon (omega ^ omega) 3).
+Compute pp (canon (T1.omega ^ T1.omega) 3). 
 \end{Coqsrc}
@@
 \begin{Coqanswer}
-  = phi0 (FS 2) : T1
+  = (_omega ^ 3)%pT1
+     : ppT1
    \end{Coqanswer}
@@
 \begin{Coqsrc}
-Example canon3 :  canon (omega ^ omega) 3 = omega ^ 3.
-Proof. reflexivity. Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Compute pp (canon (omega ^ omega * 3) 42).  
+Compute pp (canon (T1.omega ^ T1.omega * 3) 42).  
 \end{Coqsrc}
@@
 \begin{Coqanswer}
-    = (omega ^ omega * 2 + omega ^ 42)%pT1
+    = (_omega ^ _omega * 2 + _omega ^ 42)%pT1
      : ppT1
@@
 \begin{Coqsrc}
-Lemma canonS_succ i alpha :
-  nf alpha ->  canonS (succ alpha) i = alpha.
+  Lemma canon_succ i alpha :
+  nf alpha -> canon (succ alpha) i = alpha.
 Proof.
- induction alpha.
+ revert i; induction alpha.
  (* ... *)
@@
 Lemma canonS_LT i alpha :
-  nf alpha -> alpha <> zero -> canonS alpha i t1<  alpha.
+  nf alpha -> alpha <> zero ->
+  canon alpha (S i) t1<  alpha.
 \end{Coqsrc}
@@
      forall beta, beta t1< lambda ->
-                  {i:nat | beta t1< canonS lambda i}.
+                  {i:nat | beta t1< canon lambda (S i)}.
 
@@
 
-\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0} : Paths}
+\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0} : paths}
 \index{maths}{Ordinal numbers!Accessibility inside epsilon0}
@@
 \begin{Coqsrc}
-(Definition transition_S i : relation T1 :=
-  fun alpha beta =>  alpha <> zero /\ beta = canonS alpha i.
+Definition transition_S i : relation T1 :=
+fun alpha beta =>
+    alpha <> zero /\ beta = canonS alpha (S i).
 
@@
     nf alpha -> alpha <> 0 ->
-    iota alpha -1-> iota (canonS alpha i).
+    iota alpha -1-> iota (canon alpha (S i)).
 \end{Coqsrc}
@@
 Definition const_pathS i :=
-    clos_trans_1n T1 (fun alpha beta => beta = canonS alpha i).
+  clos_trans_1n T1 (fun alpha beta => alpha <> zero /\
+                                      beta = canon alpha (S i)).
 
@@
 \begin{Coqsrc}
-Theorem Theorem_2_4 (lambda : T1) :
+Theorem KS_thm_2_4 (lambda : T1) :
    nf lambda ->
@@
    forall i j, (i < j)%nat ->
-               const_pathS 0 (canonS lambda j)
-                             (canonS lambda i). 
+               const_pathS 0 (canon lambda (S j))
+                             (canon lambda (S i)).
 \end{Coqsrc}
@@
     | (0::s') => gnaw  alpha s'
-    | (S i :: s')  =>  gnaw (canonS i alpha) s'
+    | (S i :: s')  =>  gnaw (canon alpha (S i)) s'
   end.
@@
 \begin{Coqsrc}
-Compute pp (gnaw (omega * omega) (interval 6 70)).
+Compute pp (gnaw (T1.omega * T1.omega) (interval 6 70)).
 \end{Coqsrc}
@@
 \begin{Coqanswer}
- = (omega * 2 + 56)%pT1
+ = (_omega * 2 + 56)%pT1
      : ppT1
@@
 \begin{Coqsrc}
-Compute (gnaw (omega * omega) (interval 6 700)).
+Compute (gnaw (T1.omega * T1.omega) (interval 6 700)).
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Example Ex1 : mlarge (omega * omega) (interval 6 510).
+Example Ex1 : mlarge (T1.omega * T1.omega) (interval 6 510).
 \end{Coqsrc}
@@
             (f : nat -> nat -> nat)
-            (H : forall k, L_spec (canonS lambda k) (f (S k))).
+            (H : forall k, L_spec (canon lambda (S k)) (f (S k))).
 
@@
 
-Lemma L_omega_ok : L_spec omega L_omega.
+Lemma L_omega_ok : L_spec T1.omega L_omega.
 \end{Coqsrc}
@@
 Lemma path_to_omega_mult (i k:nat) :
-  path_to (omega * i) (interval (S k) (2 * (S k))) (omega * (S i)).
+path_to (T1.omega * i) (interval (S k) (2 * (S k)))
+            (T1.omega * (S i)).
 \end{Coqsrc}
@@
 Lemma omega_mult_mlarge_0 i  : forall k,
-    mlarge  (omega * (S i))
+    mlarge  (T1.omega * (S i))
             (interval (S k)
@@
 
-Lemma L_omega_mult_ok (i: nat) :  L_spec (omega * i) (L_omega_mult i).
+Lemma L_omega_mult_ok (i: nat) :  L_spec (T1.omega * i) (L_omega_mult i).
 \end{Coqsrc}
@@
 
-Lemma L_omega_square_ok: L_spec (omega * omega) 
+Lemma L_omega_square_ok: L_spec (T1.omega * T1.omega) 
           L_omega_square.
@@
 \begin{Coqsrc}
-Notation  "'one'"  := [zero,zero] : T2_scope.
+Notation  one  := [zero,zero].
 
 (** The (n+1)-th finite ordinal *)
-Notation "'FS' n" := (gcons zero zero n zero) (at level 10) : T2_scope.
+Notation FS n := (gcons zero zero n zero).
 

### 2021-06-04 doc/part-hydras.tex
update on Omega link
fix link On_plus in pdf

@@
 \vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations/ON_Omega.html}{OrdinalNotations.ON\_Omega}}
+\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_Omega.html}{OrdinalNotations.ON\_Omega}}
 
@@
 \vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations/ON_plus.html}{OrdinalNotations.ON\_plus}}
+\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_plus.html}{OrdinalNotations.ON\_plus}}
 

### 2021-06-09 doc/cutting-table/src/__init__.py
add cli to convert coq file to latex

@@ -1,0 +1,5 @@
+from .cli import parse_arg
+
+__all__ = [
+    "parse_arg"
+]

### 2021-06-09 doc/cutting-table/src/cli.py

@@ -1,0 +1,28 @@
+import argparse
+
+
+def parse_arg():
+    DESCRIPTION = ""  # TODO
+    parser = argparse.ArgumentParser(DESCRIPTION)
+
+    INPUT_FILES_HELP = "coq files '.v'"
+    parser.add_argument("files", nargs='+',
+                        help=INPUT_FILES_HELP)
+
+    OUPUT_DIR_HELP = "output directory"
+    parser.add_argument("-o", "--output-dir",
+                        default=".",
+                        help=OUPUT_DIR_HELP)
+
+    REBUILD_HELP = "rebuild files (WIP)"  # FIXME
+    parser.add_argument("-r", "--rebuild",
+                        default="store_true",
+                        help=REBUILD_HELP)
+
+    THREAD_HELP = "set number of maximum workers (WIP)"  # FIXME
+    parser.add_argument("-t", "--threads",
+                        default=None,
+                        type=int,
+                        help=THREAD_HELP)
+
+    return parser.parse_args()

### 2021-06-09 doc/part-hydras.tex
new exercises

--- 2021-06-01 doc/part-hydras.tex
@@
 
+\index{hydras}{Projects}
+\begin{project}
+  Prove that ordinal multiplication is left-distributive over ordinal addition; that is:
+  $$\forall \alpha \, \beta \, \gamma < \epsilon_0, \alpha\times(\beta+\gamma) = \alpha\times\beta+ \alpha\times \gamma$$
+
+\end{project}
 \section{A Variant for hydra battles}

### 2021-06-09 doc/cutting-table/jsonToLatex.py
add .v to json and read json

@@ -1,31 +1,23 @@
 #!/usr/bin/env python3
-import json
-import sys
-from os.path import join, dirname
-from pprint import pprint
-movie_file = sys.argv[1]
+from src import parse_args
+from src.io_json import coq_to_json, get_annotate_json
 
-root = join(dirname(__file__), "..")
-sys.path.insert(0, root)
-sys.path.insert(0, '/Users/casteranpierre/alectryon')
 
-with open(movie_file, "r") as read_file:
-    movie =json.load(read_file, strict=False)
-
-# data =  movie[0] 
-
-def latex_of_movie():
+def latex_of_movie(movie):
     from alectryon.json import annotated_of_json
     from alectryon.latex import LatexGenerator
     from alectryon.pygments import highlight_latex
     annotated = annotated_of_json(movie)
     for ltx in LatexGenerator(highlight_latex).gen(annotated):
-      #   print("<latex>")
         print(ltx)
-      #   print("</latex>")
 
-def main():
-    latex_of_movie()
 
 if __name__ == '__main__':
-    main()       
+    args = parse_args()
+    for coq_file in args.inputs:
+        json_path = coq_to_json(args.output_dir, coq_file)
+        json = get_annotate_json(json_path)
+        print("------")
+        print(json)
+        print("------")
+    # main()

### 2021-06-09 doc/cutting-table/src/__init__.py

@@ -1,5 +1,5 @@
-from .cli import parse_arg
+from .cli import parse_args
 
 __all__ = [
-    "parse_arg"
+    "parse_args"
 ]

### 2021-06-09 doc/cutting-table/src/cli.py

@@ -1,17 +1,24 @@
 import argparse
+from pathlib import Path
 
+def build_parser() -> argparse.ArgumentParser:
+    """
+    build parser
 
-def parse_arg():
+    :return: Argument parser
+    """
     DESCRIPTION = ""  # TODO
     parser = argparse.ArgumentParser(DESCRIPTION)
 
     INPUT_FILES_HELP = "coq files '.v'"
-    parser.add_argument("files", nargs='+',
+    parser.add_argument("inputs", nargs='+',
+                        type=Path,
                         help=INPUT_FILES_HELP)
 
     OUPUT_DIR_HELP = "output directory"
     parser.add_argument("-o", "--output-dir",
-                        default=".",
+                        default=Path("."),
+                        type=Path,
                         help=OUPUT_DIR_HELP)
 
     REBUILD_HELP = "rebuild files (WIP)"  # FIXME
@@ -25,4 +32,13 @@ def parse_arg():
                         type=int,
                         help=THREAD_HELP)
 
-    return parser.parse_args()
+    return parser
+
+
+def parse_args() -> argparse.Namespace:
+    """
+    parse argument
+
+    :return: Namespace structure
+    """
+    return build_parser().parse_args()

### 2021-06-09 doc/cutting-table/src/io_json.py

@@ -1,0 +1,50 @@
+from pathlib import Path
+
+COQ_EXTENSION = ".v"
+JSON_EXTENSION = ".io.json"
+
+
+def get_annotate_json(json_path: Path) -> dict:
+    """
+    get annotate json with alectryon
+
+    :param json_path: path of json file to read
+    :return: json annotate json
+    """
+    from alectryon.json import annotated_of_json
+    import json
+
+    with open(json_path, "r") as read_file:
+        return annotated_of_json(json.load(read_file, strict=False))
+
+
+def _alectryon_wrapper_json(input_file: Path, output_file: Path):
+    """
+    call alectryon script
+
+    :param output_file: output file name
+    :return:
+    """
+    from alectryon.cli import build_parser, \
+        post_process_arguments, \
+        process_pipelines
+
+    options = f"{input_file} --frontend coq --backend json -o {output_file}"
+
+    parser = build_parser()
+    args = post_process_arguments(parser,
+                                  parser.parse_args(options.split()))
+
+    process_pipelines(args)
+
+
+def coq_to_json(output_dir: Path, coq_file_path: Path) -> Path:
+    name = coq_file_path.stem
+    path_dir = output_dir / name
+    path_dir.mkdir(exist_ok=True)
+
+    json_file_path = path_dir / (name + JSON_EXTENSION)
+
+    _alectryon_wrapper_json(coq_file_path, json_file_path)
+
+    return json_file_path

### 2021-06-09 doc/cutting-table/jsonToLatex.py
cut coq file with snippet

@@ -1,23 +1,12 @@
 #!/usr/bin/env python3
-from src import parse_args
-from src.io_json import coq_to_json, get_annotate_json
-
-
-def latex_of_movie(movie):
-    from alectryon.json import annotated_of_json
-    from alectryon.latex import LatexGenerator
-    from alectryon.pygments import highlight_latex
-    annotated = annotated_of_json(movie)
-    for ltx in LatexGenerator(highlight_latex).gen(annotated):
-        print(ltx)
-
+from src.cli import parse_args
+from src.json import coq_to_json
+from src.snippet import Snippet
 
 if __name__ == '__main__':
     args = parse_args()
     for coq_file in args.inputs:
         json_path = coq_to_json(args.output_dir, coq_file)
-        json = get_annotate_json(json_path)
-        print("------")
-        print(json)
-        print("------")
+        snippets = Snippet(json_path).snippets
+        print(snippets)
     # main()

### 2021-06-09 doc/cutting-table/src/__init__.py

@@ -1,5 +1,0 @@
-from .cli import parse_args
-
-__all__ = [
-    "parse_args"
-]

### 2021-06-09 doc/cutting-table/src/json/__init__.py

@@ -1,0 +1,8 @@
+from .coq_to_json import coq_to_json
+from .read import get_annotate_json
+
+__all__ = \
+    [
+        "coq_to_json",
+        "get_annotate_json"
+    ]

### 2021-06-09 doc/cutting-table/src/json/coq_to_json.py

@@ -1,35 +1,24 @@
 from pathlib import Path
 
+from alectryon.cli import build_parser, \
+    post_process_arguments, \
+    process_pipelines
+
 COQ_EXTENSION = ".v"
 JSON_EXTENSION = ".io.json"
 
-
-def get_annotate_json(json_path: Path) -> dict:
-    """
-    get annotate json with alectryon
-
-    :param json_path: path of json file to read
-    :return: json annotate json
-    """
-    from alectryon.json import annotated_of_json
-    import json
-
-    with open(json_path, "r") as read_file:
-        return annotated_of_json(json.load(read_file, strict=False))
+R_FLAG = "-R ../../theories/ordinals hydras"  # FIXME
 
 
-def _alectryon_wrapper_json(input_file: Path, output_file: Path):
+def _alectryon_wrapper_json(input_file: Path, output_file: Path) -> None:
     """
     call alectryon script
 
     :param output_file: output file name
-    :return:
     """
-    from alectryon.cli import build_parser, \
-        post_process_arguments, \
-        process_pipelines
 
-    options = f"{input_file} --frontend coq --backend json -o {output_file}"
+    options = f"{input_file} --frontend coq --backend json -o {output_file} " \
+              f"{R_FLAG}"
 
     parser = build_parser()
     args = post_process_arguments(parser,

### 2021-06-09 doc/cutting-table/src/json/json_to_latex.py

@@ -1,0 +1,17 @@
+from pathlib import Path
+
+from alectryon.latex import LatexGenerator
+from alectryon.pygments import highlight_latex
+
+LATEX_GENERATOR = LatexGenerator(highlight_latex)
+
+
+def json_to_latex(annotated_json: dict, path: Path) -> None:
+    """
+    transform a json to latex
+
+    :param annotated_json: json to transform
+    """
+    with open(path, 'w') as file:
+        for ltx in LATEX_GENERATOR.gen(annotated_json):
+            file.write(ltx)

### 2021-06-09 doc/cutting-table/src/json/read.py

@@ -1,0 +1,15 @@
+from pathlib import Path
+from alectryon.json import annotated_of_json
+import json
+
+
+def get_annotate_json(json_path: Path):
+    """
+    get annotate json with alectryon
+
+    :param json_path: path of json file to read
+    :return: json annotate json
+    """
+
+    with open(json_path, "r") as read_file:
+        return annotated_of_json(json.load(read_file, strict=False))

### 2021-06-09 doc/cutting-table/src/snippet/Snippet.py

@@ -1,0 +1,133 @@
+from pathlib import Path
+from typing import Dict, Optional, Tuple
+from ..json import get_annotate_json
+from alectryon.core import Text
+import re
+
+from .snippetExceptions import SnippetNotExistException, \
+    SnippetAlreadyCloseException, \
+    SnippetAlreadyExistException, \
+    SnippetNotEndException
+
+
+class Snippet:
+    """
+    find snippet and make annotate structure for all snippet find.
+
+    Use run to compute snippet
+    and snippets to get new structures
+    """
+    BEGIN_STATE = "begin"
+    END_STATE = "end"
+    PATTERN = re.compile(rf"\(\*\s*({BEGIN_STATE}|{END_STATE})\s+snippet\s+(\w+)\s*\*\)")
+
+    def __init__(self, json_path: Path):
+        self.annotate = get_annotate_json(json_path)
+        self.next = None
+        self.__in_process = {}
+        self.__finish = None
+
+    def run(self) -> None:
+        """
+        find all snippet and make dictionary
+        """
+        if self.__finish is not None:
+            return
+        self.__finish = {}
+        for elm in self.annotate:
+            self.__update(elm)
+
+        snippet_not_close = list(self.__in_process.keys())
+        if snippet_not_close:
+            raise SnippetNotEndException(snippet_not_close)
+
+    def __update(self, element) -> None:
+        """
+        add element to al snippets
+        update all snippets
+
+        :param element: element to add
+        """
+        print(element, type)
+        if isinstance(element, Text):
+            self.__update_with_text(element.content)
+            return
+        for key in self.__in_process.keys():
+            self.__in_process[key].append(element)
+
+    def __update_with_text(self, text: Text):
+        dict_match, string = self.list_pattern(text.contents)
+
+        names_already_update = set()
+        for name, (begin, end) in dict_match.items():
+            if (name in list(self.__finish)) or (begin is not None and name in list(self.__in_process)):
+                raise SnippetAlreadyExistException(name)
+            elif name in list(self.__finish) and end is not None:
+                raise SnippetAlreadyCloseException(name)
+            elif begin is None and end is not None and not name in list(self.__in_process):
+                raise SnippetNotExistException(name)
+
+            start_i = (begin, 0)[begin is None]
+            end_i = (end, -1)[end is None]
+
+            self.__in_process[name].append(Text(string[start_i:end_i]))
+
+            if end is not None:
+                self.__finish[name] = self.__in_process.pop(name)
+            else:  # construct list of name already have snippet
[…]

### 2021-06-09 doc/cutting-table/src/snippet/__init__.py

@@ -1,0 +1,15 @@
+from .Snippet import Snippet
+from .snippetExceptions import \
+    SnippetAlreadyCloseException, \
+    SnippetAlreadyExistException, \
+    SnippetException, \
+    SnippetNotExistException
+
+__all__ = \
+    [
+        "Snippet",
+        "SnippetAlreadyCloseException",
+        "SnippetAlreadyExistException",
+        "SnippetException",
+        "SnippetNotExistException",
+    ]

### 2021-06-09 doc/cutting-table/src/snippet/snippetExceptions.py

@@ -1,0 +1,42 @@
+from pathlib import Path
+from typing import List
+
+
+class SnippetException(Exception):
+    pass
+
+
+class SnippetAlreadyExistException(SnippetException):
+    """
+    Raise in snippet if 2 snippet with same name in file.
+    """
+
+    def __init__(self, name_snippet: str):
+        super().__init__(f"snippet \"{name_snippet}\" already exist")
+
+
+class SnippetAlreadyCloseException(SnippetException):
+    """
+    Raise if snippet is already close.
+    """
+
+    def __init__(self, name_snippet: str):
+        super().__init__(f"snippet \"{name_snippet}\" already close")
+
+
+class SnippetNotExistException(SnippetException):
+    """
+    Raise if snippet not exist.
+    """
+
+    def __init__(self, name_snippet: str):
+        super().__init__(f"snippet \"{name_snippet}\" not exist (cannot be close).")
+
+
+class SnippetNotEndException(SnippetException):
+    """
+    Raise if snippet is have not end.
+    """
+
+    def __init__(self, names: List[str]):
+        super().__init__(f"snippet \"{','.join(names)}\" never close")

### 2021-06-09 doc/cutting-table/jsonToLatex.py
first approch to make latex

@@ -1,12 +1,18 @@
 #!/usr/bin/env python3
 from src.cli import parse_args
-from src.json import coq_to_json
+from src.json import coq_to_json, json_to_latex, get_annotate_json
 from src.snippet import Snippet
+
+LATEX_EXTENSION = ".tex"
 
 if __name__ == '__main__':
     args = parse_args()
     for coq_file in args.inputs:
         json_path = coq_to_json(args.output_dir, coq_file)
-        snippets = Snippet(json_path).snippets
-        print(snippets)
-    # main()
+        directory = json_path.parent
+        annotate_json = get_annotate_json(json_path)
+        snippets = Snippet(annotate_json).snippets
+        for name, annotate_json in snippets.items():
+            path = directory / (name + LATEX_EXTENSION)
+            json_to_latex([annotate_json], path)
+            print(f"create latex file: {path}")

### 2021-06-09 doc/cutting-table/src/json/__init__.py

@@ -1,8 +1,10 @@
 from .coq_to_json import coq_to_json
 from .read import get_annotate_json
+from .json_to_latex import json_to_latex
 
 __all__ = \
     [
         "coq_to_json",
-        "get_annotate_json"
+        "get_annotate_json",
+        "json_to_latex"
     ]

### 2021-06-09 doc/cutting-table/src/json/coq_to_json.py

@@ -21,13 +21,19 @@ def _alectryon_wrapper_json(input_file: Path, output_file: Path) -> None:
               f"{R_FLAG}"
 
     parser = build_parser()
-    args = post_process_arguments(parser,
-                                  parser.parse_args(options.split()))
+    args = post_process_arguments(parser, parser.parse_args(options.split()))
 
     process_pipelines(args)
 
 
 def coq_to_json(output_dir: Path, coq_file_path: Path) -> Path:
+    """
+    convert coq '.v' file to 'io.json' file.
+
+    :param output_dir: output directory to make new directory with name of coq file
+    :param coq_file_path: coq file to convert
+    :return: path of json file
+    """
     name = coq_file_path.stem
     path_dir = output_dir / name
     path_dir.mkdir(exist_ok=True)

### 2021-06-09 doc/cutting-table/src/json/json_to_latex.py

@@ -14,4 +14,4 @@ def json_to_latex(annotated_json: dict, path: Path) -> None:
     """
     with open(path, 'w') as file:
         for ltx in LATEX_GENERATOR.gen(annotated_json):
-            file.write(ltx)
+            file.write(str(ltx))

### 2021-06-09 doc/cutting-table/src/snippet/Snippet.py

@@ -1,6 +1,4 @@
-from pathlib import Path
 from typing import Dict, Optional, Tuple
-from ..json import get_annotate_json
 from alectryon.core import Text
 import re
 
@@ -21,8 +19,8 @@ class Snippet:
     END_STATE = "end"
     PATTERN = re.compile(rf"\(\*\s*({BEGIN_STATE}|{END_STATE})\s+snippet\s+(\w+)\s*\*\)")
 
-    def __init__(self, json_path: Path):
-        self.annotate = get_annotate_json(json_path)
+    def __init__(self, annotate_json):
+        self.annotate = annotate_json
         self.next = None
         self.__in_process = {}
         self.__finish = None
@@ -34,7 +32,9 @@ class Snippet:
         if self.__finish is not None:
             return
         self.__finish = {}
-        for elm in self.annotate:
+        print(self.annotate)
+
+        for elm in self.annotate[0]: # FIXME an recursive approch can fix this problem / But check how json to latex work (and do testing)
             self.__update(elm)
 
         snippet_not_close = list(self.__in_process.keys())
@@ -48,16 +48,17 @@ class Snippet:
 
         :param element: element to add
         """
-        print(element, type)
         if isinstance(element, Text):
-            self.__update_with_text(element.content)
+            self.__update_with_text(element)
             return
-        for key in self.__in_process.keys():
-            self.__in_process[key].append(element)
+
+        else:
+            for key in self.__in_process.keys():
+                self.__in_process[key].append(element)
+
 
     def __update_with_text(self, text: Text):
         dict_match, string = self.list_pattern(text.contents)
-
         names_already_update = set()
         for name, (begin, end) in dict_match.items():
             if (name in list(self.__finish)) or (begin is not None and name in list(self.__in_process)):
@@ -66,11 +67,14 @@ class Snippet:
                 raise SnippetAlreadyCloseException(name)
             elif begin is None and end is not None and not name in list(self.__in_process):
                 raise SnippetNotExistException(name)
+            elif begin is not None:
+                self.__in_process[name] = []
 
             start_i = (begin, 0)[begin is None]
             end_i = (end, -1)[end is None]
 
-            self.__in_process[name].append(Text(string[start_i:end_i]))
+            text = Text(string[start_i:end_i])
+            self.__in_process[name].append(text)
 
             if end is not None:
                 self.__finish[name] = self.__in_process.pop(name)
@@ -93,11 +97,10 @@ class Snippet:
         """
 
         dict_name = {}
[…]

### 2021-06-10 doc/cutting-table/coqToLatex.py
update part making

@@ -14,5 +14,5 @@ if __name__ == '__main__':
         snippets = Snippet(annotate_json).snippets
         for name, annotate_json in snippets.items():
             path = directory / (name + LATEX_EXTENSION)
-            json_to_latex([annotate_json], path)
+            json_to_latex(annotate_json, path)
             print(f"create latex file: {path}")

### 2021-06-10 doc/cutting-table/src/json/coq_to_json.py

@@ -20,6 +20,7 @@ def _alectryon_wrapper_json(input_file: Path, output_file: Path) -> None:
     options = f"{input_file} --frontend coq --backend json -o {output_file} " \
               f"{R_FLAG}"
 
+    print(f"alectryon {options}")
     parser = build_parser()
     args = post_process_arguments(parser, parser.parse_args(options.split()))
 

### 2021-06-10 doc/cutting-table/src/snippet/Snippet.py

@@ -32,7 +32,6 @@ class Snippet:
         if self.__finish is not None:
             return
         self.__finish = {}
-        print(self.annotate)
 
         for elm in self.annotate[0]: # FIXME an recursive approch can fix this problem / But check how json to latex work (and do testing)
             self.__update(elm)
@@ -77,7 +76,7 @@ class Snippet:
             self.__in_process[name].append(text)
 
             if end is not None:
-                self.__finish[name] = self.__in_process.pop(name)
+                self.__finish[name] = [self.__in_process.pop(name)]
             else:  # construct list of name already have snippet
                 names_already_update.add(name)
 

### 2021-06-11 doc/cutting-table/src/json/coq_to_json.py
add paramater backend and frontend

@@ -8,6 +8,8 @@ COQ_EXTENSION = ".v"
 JSON_EXTENSION = ".io.json"
 
 R_FLAG = "-R ../../theories/ordinals hydras"  # FIXME
+FRONTEND = "coq"
+BACKEND = "json"
 
 
 def _alectryon_wrapper_json(input_file: Path, output_file: Path) -> None:
@@ -17,8 +19,8 @@ def _alectryon_wrapper_json(input_file: Path, output_file: Path) -> None:
     :param output_file: output file name
     """
 
-    options = f"{input_file} --frontend coq --backend json -o {output_file} " \
-              f"{R_FLAG}"
+    options = f"{input_file} --frontend {FRONTEND} --backend {BACKEND} " \
+              f"-o {output_file} {R_FLAG}"
 
     print(f"alectryon {options}")
     parser = build_parser()

### 2021-06-11 doc/cutting-table/coqToLatex.py
remove json creation and used annotate

@@ -1,18 +1,6 @@
 #!/usr/bin/env python3
-from src.cli import parse_args
-from src.json import coq_to_json, json_to_latex, get_annotate_json
-from src.snippet import Snippet
+from src.cli import main
 
-LATEX_EXTENSION = ".tex"
 
 if __name__ == '__main__':
-    args = parse_args()
-    for coq_file in args.inputs:
-        json_path = coq_to_json(args.output_dir, coq_file)
-        directory = json_path.parent
-        annotate_json = get_annotate_json(json_path)
-        snippets = Snippet(annotate_json).snippets
-        for name, annotate_json in snippets.items():
-            path = directory / (name + LATEX_EXTENSION)
-            json_to_latex(annotate_json, path)
-            print(f"create latex file: {path}")
+    main()

### 2021-06-11 doc/cutting-table/src/annotate.py

@@ -1,0 +1,45 @@
+from typing import List
+
+from alectryon.core import annotate
+from pathlib import Path
+
+annotated_t = List[List]
+
+
+def coq_file_to_annotated(path_coq_file: Path, sertop_args=()) -> annotated_t:
+    """
+    annotate file coq file
+
+    :param path_coq_file: path of coq file to use ('.v' file)
+    :return: annotate content
+    """
+    with open(path_coq_file, 'r') as file:
+        return annotate([file.read()], sertop_args)
+
+
+LATEX_GENERATOR = None
+
+
+def _get_latex_generator():
+    """
+    getter latex generator (lazy)
+    :return: latex generator
+    """
+    global LATEX_GENERATOR
+    if LATEX_GENERATOR is None:
+        from alectryon.latex import LatexGenerator
+        from alectryon.pygments import highlight_latex
+        LATEX_GENERATOR = LatexGenerator(highlight_latex)
+    return LATEX_GENERATOR
+
+
+def annotate_to_latex(annotated_content: annotated_t, latex_file: Path):
+    """
+    transform annotated content to latex file
+
+    :param annotated_content: annotated content to convert
+    :param latex_file: path of latex file to product
+    """
+    with open(latex_file, 'w') as file:
+        for ltx in _get_latex_generator().gen(annotated_content):
+            file.write(str(ltx))

### 2021-06-11 doc/cutting-table/src/cli.py

@@ -32,7 +32,52 @@ def build_parser() -> argparse.ArgumentParser:
                         type=int,
                         help=THREAD_HELP)
 
+    # serapi (copy from alectryon.cli)
+    SUBP_HELP = "Pass arguments to the SerAPI process"
+    subp = parser.add_argument_group("Subprocess arguments", SUBP_HELP)
+
+    SERTOP_ARGS_HELP = "Pass a single argument to SerAPI (e.g. -Q dir,lib)."
+    subp.add_argument("--sertop-arg", dest="sertop_args",
+                      action="append", default=[],
+                      metavar="SERAPI_ARG",
+                      help=SERTOP_ARGS_HELP)
+
+    I_HELP = "Pass -I DIR to the SerAPI subprocess."
+    subp.add_argument("-I", "--ml-include-path", dest="coq_args_I",
+                      metavar="DIR", nargs=1, action="append",
+                      default=[], help=I_HELP)
+
+    Q_HELP = "Pass -Q DIR COQDIR to the SerAPI subprocess."
+    subp.add_argument("-Q", "--load-path", dest="coq_args_Q",
+                      metavar=("DIR", "COQDIR"), nargs=2, action="append",
+                      default=[], help=Q_HELP)
+
+    R_HELP = "Pass -R DIR COQDIR to the SerAPI subprocess."
+    subp.add_argument("-R", "--rec-load-path", dest="coq_args_R",
+                      metavar=("DIR", "COQDIR"), nargs=2, action="append",
+                      default=[], help=R_HELP)
+
+    EXPECT_UNEXPECTED_HELP = "Ignore unexpected output from SerAPI"
+    parser.add_argument("--expect-unexpected", action="store_true",
+                        default=False, help=EXPECT_UNEXPECTED_HELP)
     return parser
+
+
+def post_process_arguments(args: argparse.Namespace) -> argparse.Namespace:
+    """
+    post process sertop arguments
+
+    :param args: args parsed
+    :return: new name space
+    """
+    for dirpath in args.coq_args_I:
+        args.sertop_args.extend(("-I", dirpath))
+    for pair in args.coq_args_R:
+        args.sertop_args.extend(("-R", ",".join(pair)))
+    for pair in args.coq_args_Q:
+        args.sertop_args.extend(("-Q", ",".join(pair)))
+
+    return args
 
 
 def parse_args() -> argparse.Namespace:
@@ -41,4 +86,23 @@ def parse_args() -> argparse.Namespace:
 
[…]

### 2021-06-11 doc/cutting-table/src/json/__init__.py

@@ -1,10 +1,0 @@
-from .coq_to_json import coq_to_json
-from .read import get_annotate_json
-from .json_to_latex import json_to_latex
-
-__all__ = \
-    [
-        "coq_to_json",
-        "get_annotate_json",
-        "json_to_latex"
-    ]

### 2021-06-11 doc/cutting-table/src/json/coq_to_json.py

@@ -1,48 +1,0 @@
-from pathlib import Path
-
-from alectryon.cli import build_parser, \
-    post_process_arguments, \
-    process_pipelines
-
-COQ_EXTENSION = ".v"
-JSON_EXTENSION = ".io.json"
-
-R_FLAG = "-R ../../theories/ordinals hydras"  # FIXME
-FRONTEND = "coq"
-BACKEND = "json"
-
-
-def _alectryon_wrapper_json(input_file: Path, output_file: Path) -> None:
-    """
-    call alectryon script
-
-    :param output_file: output file name
-    """
-
-    options = f"{input_file} --frontend {FRONTEND} --backend {BACKEND} " \
-              f"-o {output_file} {R_FLAG}"
-
-    print(f"alectryon {options}")
-    parser = build_parser()
-    args = post_process_arguments(parser, parser.parse_args(options.split()))
-
-    process_pipelines(args)
-
-
-def coq_to_json(output_dir: Path, coq_file_path: Path) -> Path:
-    """
-    convert coq '.v' file to 'io.json' file.
-
-    :param output_dir: output directory to make new directory with name of coq file
-    :param coq_file_path: coq file to convert
-    :return: path of json file
-    """
-    name = coq_file_path.stem
-    path_dir = output_dir / name
-    path_dir.mkdir(exist_ok=True)
-
-    json_file_path = path_dir / (name + JSON_EXTENSION)
-
-    _alectryon_wrapper_json(coq_file_path, json_file_path)
-
-    return json_file_path

### 2021-06-11 doc/cutting-table/src/json/json_to_latex.py

@@ -1,17 +1,0 @@
-from pathlib import Path
-
-from alectryon.latex import LatexGenerator
-from alectryon.pygments import highlight_latex
-
-LATEX_GENERATOR = LatexGenerator(highlight_latex)
-
-
-def json_to_latex(annotated_json: dict, path: Path) -> None:
-    """
-    transform a json to latex
-
-    :param annotated_json: json to transform
-    """
-    with open(path, 'w') as file:
-        for ltx in LATEX_GENERATOR.gen(annotated_json):
-            file.write(str(ltx))

### 2021-06-11 doc/cutting-table/src/json/read.py

@@ -1,15 +1,0 @@
-from pathlib import Path
-from alectryon.json import annotated_of_json
-import json
-
-
-def get_annotate_json(json_path: Path):
-    """
-    get annotate json with alectryon
-
-    :param json_path: path of json file to read
-    :return: json annotate json
-    """
-
-    with open(json_path, "r") as read_file:
-        return annotated_of_json(json.load(read_file, strict=False))

### 2021-06-11 doc/cutting-table/src/snippet/Snippet.py

@@ -33,11 +33,10 @@ class Snippet:
             return
         self.__finish = {}
 
-        for elm in self.annotate[0]: # FIXME an recursive approch can fix this problem / But check how json to latex work (and do testing)
+        for elm in sum(self.annotate, []):  # flatten list
             self.__update(elm)
 
-        snippet_not_close = list(self.__in_process.keys())
-        if snippet_not_close:
+        if snippet_not_close := list(self.__in_process.keys()):
             raise SnippetNotEndException(snippet_not_close)
 
     def __update(self, element) -> None:
@@ -54,7 +53,6 @@ class Snippet:
         else:
             for key in self.__in_process.keys():
                 self.__in_process[key].append(element)
-
 
     def __update_with_text(self, text: Text):
         dict_match, string = self.list_pattern(text.contents)

### 2021-06-16 doc/part-hydras.tex
update the .tex file

@@
 \begin{Coqsrc}
-Notation  "'one'"  := [zero,zero] : T2_scope.
+Notation  one  := [zero,zero] : T2_scope.
 

### 2021-06-21 doc/part-hydras.tex
Update doc/part-hydras.tex
Co-authored-by: Théo Zimmermann <theo.zimmermann@inria.fr>
Update documentation, and remove useless definitions

@@
 \begin{Coqsrc}
-Notation  one  := [zero,zero] : T2_scope.
+Notation  one  := [zero,zero].
 
 (** The (n+1)-th finite ordinal *)
-Notation "'FS' n" := (gcons zero zero n zero) (at level 10) : T2_scope.
+Notation FS n := (gcons zero zero n zero).
 

### 2021-06-28 doc/hydras.tex

@@
 \author{Pierre Castéran\\ LaBRI, Univ.
-Bordeaux, CNRS UMR 5800, \thanks{With contributions by Yves Bertot, \'Evelyne Contejean,  Florian Hatat and Pascal Manoury.
+Bordeaux, CNRS UMR 5800, \thanks{With contributions by Yves Bertot, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury and Théo Zimmermann.
 The formalization of primitive recursive functions was originally authored by
@@
     Many thanks to Yves Bertot, \'Evelyne Contejean,  Florian Hatat,  David Ilcinkas, 
-Pascal Manoury,  Karl Palmskog, Sylvain Salvati, Alan Schmitt and Théo Zimmerman for their help on the elaboration of this document, and to the
+Pascal Manoury,  Karl Palmskog, Sylvain Salvati, Alan Schmitt and Théo Zimmermann for their help on the elaboration of this document, and to the
  members of the \emph{Formal Methods} team and the \emph{Coq working group} at laBRI for their helpful comments 

### 2021-06-29 doc/chapter-powers.tex
consistent use of capitalization in headers
fix link in pdf

@@
-\chapter{Smart Computation of \texorpdfstring{$x^n$}{Powers}}
+\chapter{Smart computation of \texorpdfstring{$x^n$}{Powers}}
 \label{chapter-powers}
@@
 \emph{From Module 
-\href{../theories/html/hydras.additions.FirstSteps.html}{\texttt{additions.FirstSteps}}}
+\href{../theories/html/additions.FirstSteps.html}{\texttt{additions.FirstSteps}}}
 \label{sect: power-definitions}
@@
 \emph{From Module
-\href{../theories/html/hydras.additions.FirstSteps.html}{additions.FirstSteps}}
+\href{../theories/html/FirstSteps.html}{additions.FirstSteps}}
 
@@
 \emph{From Library
-\href{../theories/html/hydras.additions.Fib2.html}{additions.Fib2}}
+\href{../theories/html/additions.Fib2.html}{additions.Fib2}}
 
@@
 
-\emph{From Module \href{../theories/html/hydras.additions.Monoid_def.html}{additions/Monoid\_def.v}.}
+\emph{From Module \href{../theories/html/additions.Monoid_def.html}{additions/Monoid\_def.v}.}
 
@@
 
-\emph{From Module \href{../theories/html/hydras.additions.Monoid_def.html}{additions.Monoid\_def}}
+\emph{From Module \href{../theories/html/additions.Monoid_def.html}{additions.Monoid\_def}}
 
@@
 
-The following class definition, from Module \href{../theories/html/hydras.additions.Monoid_def.html}{additions.Monoid\_def},
+The following class definition, from Module \href{../theories/html/additions.Monoid_def.html}{additions.Monoid\_def},
 is parameterized with some type $A$,
@@
 Complete definitions and proofs are given in 
-File~\href{../theories/html/hydras.additions.Monoid_instances.html}{additions/Monoid\_instances.v}.
+File~\href{../theories/html/additions.Monoid_instances.html}{additions/Monoid\_instances.v}.
 
@@
 \noindent
-\emph{From Module \href{../theories/html/hydras.additions.Monoid_def.html}{additions.Monoid\_def}}
+\emph{From Module \href{../theories/html/additions.Monoid_def.html}{additions.Monoid\_def}}
 
@@
 
-\emph{From Module \href{../theories/html/hydras.additions.Monoid_instances.html}{additions.Monoid\_instances}}
+\emph{From Module \href{../theories/html/additions.Monoid_instances.html}{additions.Monoid\_instances}}
 
@@
 
-The  module \href{../theories/html/hydras.additions.Pow.html}{additions.Pow} defines two functions for exponentiation on any 
+The  module \href{../theories/html/additions.Pow.html}{additions.Pow} defines two functions for exponentiation on any 
 \texttt{EMonoid}  on carrier $A$.
@@
 
-From Module~\href{../theories/html/hydras.additions.Demo_power.html}{additions.Demo\_power}
+From Module~\href{../theories/html/additions.Demo_power.html}{additions.Demo\_power}
 
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.additions.Pow.html}{additions.Pow}}
+\emph{From Module~\href{../theories/html/additions.Pow.html}{additions.Pow}}
 
@@
 \vspace{4pt}
-From Module~\href{../theories/html/hydras.additions.Pow.html}{additions.Pow}
+From Module~\href{../theories/html/additions.Pow.html}{additions.Pow}
 
@@
 Section \texttt{M\_given} of Module 
-~\href{../theories/html/hydras.additions.Pow.html}{additions.Pow} is devoted to the proof 
+~\href{../theories/html/additions.Pow.html}{additions.Pow} is devoted to the proof 
 of properties of the functions above.
@@
 
-From Module~\href{../theories/html/hydras.additions.Addition_Chains.html}{additions.Addition\_Chains}
+From Module~\href{../theories/html/additions.Addition_Chains.html}{additions.Addition\_Chains}
 
@@
 
-\textbf{Note} A first solution (in ~\href{../theories/html/hydras.additions.Trace_exercise.html}{additions.Trace\_exercise}) consists in the definition of 
+\textbf{Note} A first solution (in ~\href{../theories/html/additions.Trace_exercise.html}{additions.Trace\_exercise}) consists in the definition of 
 a (non-associative) multiplication over a type of trace, and apply the function
@@
 This tactic is not adapted to much bigger exponents. In \linebreak
- Module~\href{../theories/html/hydras.additions.Euclidean_Chains.html}{Euclidean\_Chains},
+ Module~\href{../theories/html/additions.Euclidean_Chains.html}{Euclidean\_Chains},
  for instance, we tried to apply this tactic for proving the correctness 
@@
 All the \coq{} material described in this section is available on 
- Module~\href{../theories/html/hydras.additions.Euclidean_Chains.html}{additions/Euclidean\_Chains.v}
+ Module~\href{../theories/html/additions.Euclidean_Chains.html}{additions/Euclidean\_Chains.v}
 
@@
 \noindent
-From ~\href{../theories/html/hydras.additions.Dichotomy.html}{additions.Dichotomy}.
+From ~\href{../theories/html/additions.Dichotomy.html}{additions.Dichotomy}.
 \begin{Coqsrc}
@@
 
-This strategy is defined in Module ~\href{../theories/html/hydras.additions.Dichotomy.html}{additions.Dichotomy}.
+This strategy is defined in Module ~\href{../theories/html/additions.Dichotomy.html}{additions.Dichotomy}.
 
@@
 
-\emph{From Module~\href{../theories/html/hydras.additions.BinaryStrat.html}{additions.BinaryStrat}.}
+\emph{From Module~\href{../theories/html/additions.BinaryStrat.html}{additions.BinaryStrat}.}
 
@@
 
-From Module~\href{../theories/html/hydras.additions.AM.html}{additions.AM}
+From Module~\href{../theories/html/additions.AM.html}{additions.AM}
 
@@
 
-In the library~\href{../theories/html/hydras.additions.AM.html}{additions.AM},
+In the library~\href{../theories/html/additions.AM.html}{additions.AM},
 we define a chain generator for this data structure. 

### 2021-06-29 doc/part-hydras.tex
fix link in pdf

@@
 \vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}}
+\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_mult.html}{OrdinalNotations.ON\_Generic}}
 
@@
 \emph{You may start this exercise with the file
-\url{../exercises/ordinals/predSuccUnicity.v}.}
+\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/predSuccUnicity.v}{exercises/ordinals/predSuccUnicity.v}.}
 
@@
 \emph{You may start this exercise with the file
-\url{../exercises/ordinals/lt_succ_le.v}.}
+\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/lt_succ_le.v}{exercises/ordinals/lt\_succ\_le.v}.}
 \end{exercise}
@@
 
-The detailed  proof script is in the file \url{../theories/html/hydras.Hydra.Omega2_Small.html}.
+The detailed  proof script is in the file
+\href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/Hydra/Omega2_Small.v}{theories/ordinals/Hydra/Omega2\_Small.v}.
 
@@
 \emph{You may start this exercise with the file
-\url{../exercises/ordinals/Limit_Infinity.v}.}
+     \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/Limit_Infinity.v}{exercises/ordinals/Limit\_Infinity.v}.}
  \end{exercise}
@@
 \emph{You may start this exercise with the file
-\url{../exercises/ordinals/ge_omega_iff.v}.}
+    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/ge_omega_iff.v}{exercises/ordinals/ge\_omega\_iff.v}.}
 \end{exercise}
@@
 \emph{You may start this exercise with the file
-\url{../exercises/ordinals/T1_ltNotWf.v}.}
+    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/T1_ltNotWf.v}{exercises/ordinals/T1\_ltNotWf.v}.}
 \end{exercise}
@@
 The following lemma is an application of the strict inequality 
-\showmath{\alpha < \omega ^\alpha}. If \showmath{\alpha} is strongly accessible, then, by definition, \showmath{\omega^\alpha} is accessible, thus \showmath{\alpha} is \emph{a fortiori} accessible.
+$\alpha < \omega ^\alpha.$ If \alpha is strongly accessible, then, by definition,
+$\omega^\alpha$ is accessible, thus \alpha is \emph{a fortiori} accessible.
 
@@
 
-Thus, it remains to prove that every ordinal strictly less than \showmath{\epsilon_0} 
+Thus, it remains to prove that every ordinal strictly less than $\epsilon_0$
 is strongly accessible.
@@
 
-First, we prove that, for  any \texttt{LT}-accessible term \showmath{\alpha}, \showmath{\alpha} is 
+First, we prove that, for  any \texttt{LT}-accessible term $\alpha$, $\alpha$ is 
 strongly accessible too (\emph{i.e.} any well formed
@@
 
-The proof is structured as an induction on \showmath{\alpha}'s accessibility. Let us consider
+The proof is structured as an induction on $\alpha'$s accessibility. Let us consider
 any  accessible term $\alpha$.
@@
 \label{remark:a3pat}
-The alternate proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
+The alternate proof of well-foundedness using \'Evelyne Contejean's work on
+    recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the
+    library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0.Epsilon0rpo}.
  \end{remark}
@@
 
-\emph{From Module~\href{../theories/html/hydras.Schutte.Schutte.Correctness_E0.html}{Schutte.Schutte.Correctness\_E0}}
+\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
 
@@
 \textbf{Hint}. You may use the following definition (in 
-\href{../theories/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}).
+    \href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}).
 
@@
     \end{project}
-\section{A Variant for hydra battles}
+    \section{A variant for hydra battles}
 
@@
 
-In the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
+In the module
+\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
 This part of the development is parameterized by some class $B$ of battles, which will be instantiated later to \texttt{free} or \texttt{standard}.
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.KS.html}{Epsilon0.KS}}
+\emph{From
+Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
 
@@
 be consulted in the module 
-\url{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}.
+\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}{Hydra.Epsilon0\_Needed\_Std}.
 Please note that it has the same global structure as in section\ref{sec:free-battles-case} 
@@
 \href{../theories/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
-{\texttt{Epsilon0.Paths}}   and 
+{Epsilon0.pathS}   and 
 \href{../theories/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
@@
 Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
-(please see also \href{../theories/html/hydras.Schutte.Correctness_E0.html}{the module \texttt{hydras.Schutte.Correctness\_E0}}).
+(please see also the module
+\href{../theories/html/hydras.Schutte.Correctness_E0.html}{hydras.Schutte.Correctness\_E0}).
 
@@
 The following code comes from 
- \url{../theories/html/hydras.Epsilon0.F_alpha.html}.
+ \href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}.
 
@@
 \emph{You may start this exercise with the file
-\url{../exercises/ordinals/F_3.v}.}
+\href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
 \end{exercise}
@@
 \emph{You may start this exercise with the file
-\url{../exercises/ordinals/F_3.v}.}
+    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
 \end{exercise}
@@
 \emph{You may start this exercise with the file
-\url{../exercises/ordinals/is_F_monotonous.v}.}
+    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/is_F_monotonous.v}{exercises/ordinals/is\_F\_monotonous.v}.}
 \end{exercise}
@@
 \emph{You may start this exercise with the file
-\url{../exercises/ordinals/F_omega.v}. Properties of the Ackermann function are studied in 
-\url{../theories/ordinals/MoreAck/Ack.v} and
-\url{../theories/ordinals/MoreAck/AcknotPR.v}.}
+    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_omega.v}{exercises/ordinals/F\_Omega.v}.
+Properties of the Ackermann function are studied in
+    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/Ack.v}{theories/ordinals/MoreAck/Ack.v} and
+    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/AckNotPR.v}{theories/ordinals/MoreAck/AckNotPR.v}
+.}
 \end{exercise}
@@
 \texttt{ordinal} (The  class \texttt{WO} is defined in
-Module~\href{../theories/html/hydras.Schutte.Well_Orders.html}%
-{\texttt{Well\_Orders.v}}).
+Module~\href{../theories/html/hydras.Schutte.Well_Orders.html}{Schutte.Well\_Orders.v}).
 
@@
 We prove the inequality  $\alpha<\omega^\alpha$ whenever $\alpha < \epsilon_0$.
-\emph{Note that this condition was implicit in Module \href{../theories/html/hydras.Epsilon0/T1.html\#lt_phi0}%
-{\texttt{Epsilon0.T1}}.}
+\emph{Note that this condition was implicit in Module
+\href{../theories/html/hydras.Epsilon0.T1.html\#lt_phi0}{Epsilon0.T1}.}
 
@@
 \emph{You may start this exercise with the file
-\url{../exercises/ordinals/schutte_cnf_counter_example.v}.}
+    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/schutte_cnf_counter_example.v}{exercises/ordinals/schutte\_cnf\_counter\_example.v}.}
 \end{exercise}
@@
 \begin{remark}
-The sub-directory \href{../theories/html/hydras.Gamma0.html}%
-{\texttt{theories/Gamma0}} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form. 
+The sub-directory
+    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/Gamma0}{theories/oridinals/Gamma0} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form.
 \end{remark}
@@
 Our library 
-\href{../theories/html/hydras.Schutte.correctness_E0.html}%
+\href{../theories/html/hydras.Schutte.Correctness_E0.html}%
 {\texttt{Schutte.Correctness\_E0}} establishes the link between two very different modelizations of ordinal numbers. In other words, it ``validates'' a data structure in terms of

### 2021-06-29 doc/cutting-table/src/Snippet.py
try to use separate snippets block in latex file

@@ -1,0 +1,42 @@
+from pathlib import Path
+
+
+class Snippet:
+    LATEX_EXTENSION = ".tex"
+
+    def __init__(self, name: str):
+        self.__name = name
+        self.content = ""
+
+    @property
+    def name(self) -> str:
+        """
+        name of snippet
+        :return: str
+        """
+        return self.__name
+
+    def add_content(self, content_to_add: str) -> None:
+        """
+        add new line to content
+
+        :param content_to_add: content added
+        :return: None
+        """
+        self.content += content_to_add
+
+    def __str__(self) -> str:
+        return self.content
+
+    def get_path(self, directory: Path = Path('.')) -> Path:
+        """
+        get path of file with directory
+
+        :param directory: destination directory
+        :return: Path of file
+        """
+        return directory / (self.name + self.LATEX_EXTENSION)
+
+    def write(self, directory: Path):
+        with open(self.get_path(directory), 'w') as file:
+            file.write(self.content)

### 2021-06-29 doc/cutting-table/src/annotate.py

@@ -1,45 +1,0 @@
-from typing import List
-
-from alectryon.core import annotate
-from pathlib import Path
-
-annotated_t = List[List]
-
-
-def coq_file_to_annotated(path_coq_file: Path, sertop_args=()) -> annotated_t:
-    """
-    annotate file coq file
-
-    :param path_coq_file: path of coq file to use ('.v' file)
-    :return: annotate content
-    """
-    with open(path_coq_file, 'r') as file:
-        return annotate([file.read()], sertop_args)
-
-
-LATEX_GENERATOR = None
-
-
-def _get_latex_generator():
-    """
-    getter latex generator (lazy)
-    :return: latex generator
-    """
-    global LATEX_GENERATOR
-    if LATEX_GENERATOR is None:
-        from alectryon.latex import LatexGenerator
-        from alectryon.pygments import highlight_latex
-        LATEX_GENERATOR = LatexGenerator(highlight_latex)
-    return LATEX_GENERATOR
-
-
-def annotate_to_latex(annotated_content: annotated_t, latex_file: Path):
-    """
-    transform annotated content to latex file
-
-    :param annotated_content: annotated content to convert
-    :param latex_file: path of latex file to product
-    """
-    with open(latex_file, 'w') as file:
-        for ltx in _get_latex_generator().gen(annotated_content):
-            file.write(str(ltx))

### 2021-06-29 doc/cutting-table/src/cli.py

@@ -1,17 +1,18 @@
 import argparse
 from pathlib import Path
 
+
 def build_parser() -> argparse.ArgumentParser:
     """
     build parser
 
     :return: Argument parser
     """
-    DESCRIPTION = ""  # TODO
+    DESCRIPTION = "Extract snippet in latex file generate by alectryon."
     parser = argparse.ArgumentParser(DESCRIPTION)
 
-    INPUT_FILES_HELP = "coq files '.v'"
-    parser.add_argument("inputs", nargs='+',
+    INPUT_FILES_HELP = "latex file '.tex'"
+    parser.add_argument("input",
                         type=Path,
                         help=INPUT_FILES_HELP)
 
@@ -20,64 +21,7 @@ def build_parser() -> argparse.ArgumentParser:
                         default=Path("."),
                         type=Path,
                         help=OUPUT_DIR_HELP)
-
-    REBUILD_HELP = "rebuild files (WIP)"  # FIXME
-    parser.add_argument("-r", "--rebuild",
-                        default="store_true",
-                        help=REBUILD_HELP)
-
-    THREAD_HELP = "set number of maximum workers (WIP)"  # FIXME
-    parser.add_argument("-t", "--threads",
-                        default=None,
-                        type=int,
-                        help=THREAD_HELP)
-
-    # serapi (copy from alectryon.cli)
-    SUBP_HELP = "Pass arguments to the SerAPI process"
-    subp = parser.add_argument_group("Subprocess arguments", SUBP_HELP)
-
-    SERTOP_ARGS_HELP = "Pass a single argument to SerAPI (e.g. -Q dir,lib)."
-    subp.add_argument("--sertop-arg", dest="sertop_args",
-                      action="append", default=[],
-                      metavar="SERAPI_ARG",
-                      help=SERTOP_ARGS_HELP)
-
-    I_HELP = "Pass -I DIR to the SerAPI subprocess."
-    subp.add_argument("-I", "--ml-include-path", dest="coq_args_I",
-                      metavar="DIR", nargs=1, action="append",
-                      default=[], help=I_HELP)
-
-    Q_HELP = "Pass -Q DIR COQDIR to the SerAPI subprocess."
-    subp.add_argument("-Q", "--load-path", dest="coq_args_Q",
-                      metavar=("DIR", "COQDIR"), nargs=2, action="append",
-                      default=[], help=Q_HELP)
-
-    R_HELP = "Pass -R DIR COQDIR to the SerAPI subprocess."
[…]

### 2021-06-29 doc/cutting-table/src/extract.py

@@ -1,0 +1,103 @@
+from pathlib import Path
+from typing import Dict, List
+import re
+from .Snippet import Snippet
+from .snippetExceptions import \
+    SnippetNotBeginException, \
+    SnippetAlreadyCloseException, \
+    SnippetAlreadyExistException, \
+    SnippetNotEndException
+
+
+def get_reader(path: Path) -> str:
+    """
+    reader who yield lines
+
+    :param path: path of file to read
+    :return: line read
+    """
+    with open(path, 'r') as file:
+        yield from enumerate(file, 1)
+
+
+def add_to_snippets(content: str, snippets: Dict[str, Snippet]) -> None:
+    """
+    add content to all snippet
+
+    :param content: content to add
+    :param snippets: snippet used
+    """
+    for snippet in snippets.values():
+        snippet.add_content(content)
+
+
+# text pattern
+BEGIN_STATE = "begin"
+END_STATE = "end"
+TEXT_PATTERN = re.compile(rf"\s*\\({BEGIN_STATE}|{END_STATE})\{{txt}}")
+
+# snippet pattern
+SNIPPET_PATTERN = re.compile(rf".*\\PY{{c}}{{\(\*\~+({BEGIN_STATE}|{END_STATE})\~+snippet\~+(\w+)\~+\*\)}}")
+
+BEGIN_ALECTRYON = "\\begin{alectryon}\n"
+END_ALECTRYON = "\\end{alectryon}\n"
+
+def extract_snippets_text(reader, begin_line: str,
+                          open_snippets: Dict[str, Snippet], close_snippets: List[Snippet]) -> str:
+    """
+    This function is call by extract snippet if we are in text part.
+    This function open new snippets found, or close it.
+    This function end when text part is closed.
+
+    :param reader: reader used
+    :param begin_line: begin line used
+    :param open_snippets: snippet not end
+    :param close_snippets: snippet already end
+    :return: content to add
+    """
+    content = BEGIN_ALECTRYON + begin_line  # TODO found other method
+    for num, line in reader:
+        # match open or close snippet
+        if (match := SNIPPET_PATTERN.match(line)) is not None:
+            state, name = match.groups()
+            if state == BEGIN_STATE:
+                SnippetAlreadyExistException.check(name, num, open_snippets, close_snippets)
+                open_snippets[name] = Snippet(name)
+            else:
+                SnippetAlreadyCloseException.check(name, num, close_snippets)
[…]

### 2021-06-29 doc/cutting-table/src/snippet/Snippet.py

@@ -1,133 +1,0 @@
-from typing import Dict, Optional, Tuple
-from alectryon.core import Text
-import re
-
-from .snippetExceptions import SnippetNotExistException, \
-    SnippetAlreadyCloseException, \
-    SnippetAlreadyExistException, \
-    SnippetNotEndException
-
-
-class Snippet:
-    """
-    find snippet and make annotate structure for all snippet find.
-
-    Use run to compute snippet
-    and snippets to get new structures
-    """
-    BEGIN_STATE = "begin"
-    END_STATE = "end"
-    PATTERN = re.compile(rf"\(\*\s*({BEGIN_STATE}|{END_STATE})\s+snippet\s+(\w+)\s*\*\)")
-
-    def __init__(self, annotate_json):
-        self.annotate = annotate_json
-        self.next = None
-        self.__in_process = {}
-        self.__finish = None
-
-    def run(self) -> None:
-        """
-        find all snippet and make dictionary
-        """
-        if self.__finish is not None:
-            return
-        self.__finish = {}
-
-        for elm in sum(self.annotate, []):  # flatten list
-            self.__update(elm)
-
-        if snippet_not_close := list(self.__in_process.keys()):
-            raise SnippetNotEndException(snippet_not_close)
-
-    def __update(self, element) -> None:
-        """
-        add element to al snippets
-        update all snippets
-
-        :param element: element to add
-        """
-        if isinstance(element, Text):
-            self.__update_with_text(element)
-            return
-
-        else:
-            for key in self.__in_process.keys():
-                self.__in_process[key].append(element)
-
-    def __update_with_text(self, text: Text):
-        dict_match, string = self.list_pattern(text.contents)
-        names_already_update = set()
-        for name, (begin, end) in dict_match.items():
-            if (name in list(self.__finish)) or (begin is not None and name in list(self.__in_process)):
-                raise SnippetAlreadyExistException(name)
-            elif name in list(self.__finish) and end is not None:
-                raise SnippetAlreadyCloseException(name)
-            elif begin is None and end is not None and not name in list(self.__in_process):
-                raise SnippetNotExistException(name)
-            elif begin is not None:
-                self.__in_process[name] = []
-
-            start_i = (begin, 0)[begin is None]
-            end_i = (end, -1)[end is None]
-
-            text = Text(string[start_i:end_i])
-            self.__in_process[name].append(text)
-
-            if end is not None:
-                self.__finish[name] = [self.__in_process.pop(name)]
-            else:  # construct list of name already have snippet
[…]

### 2021-06-29 doc/cutting-table/src/snippet/__init__.py

@@ -1,15 +1,0 @@
-from .Snippet import Snippet
-from .snippetExceptions import \
-    SnippetAlreadyCloseException, \
-    SnippetAlreadyExistException, \
-    SnippetException, \
-    SnippetNotExistException
-
-__all__ = \
-    [
-        "Snippet",
-        "SnippetAlreadyCloseException",
-        "SnippetAlreadyExistException",
-        "SnippetException",
-        "SnippetNotExistException",
-    ]

### 2021-06-29 doc/cutting-table/src/snippet/snippetExceptions.py

@@ -1,42 +1,0 @@
-from pathlib import Path
-from typing import List
-
-
-class SnippetException(Exception):
-    pass
-
-
-class SnippetAlreadyExistException(SnippetException):
-    """
-    Raise in snippet if 2 snippet with same name in file.
-    """
-
-    def __init__(self, name_snippet: str):
-        super().__init__(f"snippet \"{name_snippet}\" already exist")
-
-
-class SnippetAlreadyCloseException(SnippetException):
-    """
-    Raise if snippet is already close.
-    """
-
-    def __init__(self, name_snippet: str):
-        super().__init__(f"snippet \"{name_snippet}\" already close")
-
-
-class SnippetNotExistException(SnippetException):
-    """
-    Raise if snippet not exist.
-    """
-
-    def __init__(self, name_snippet: str):
-        super().__init__(f"snippet \"{name_snippet}\" not exist (cannot be close).")
-
-
-class SnippetNotEndException(SnippetException):
-    """
-    Raise if snippet is have not end.
-    """
-
-    def __init__(self, names: List[str]):
-        super().__init__(f"snippet \"{','.join(names)}\" never close")

### 2021-06-29 doc/cutting-table/src/snippetExceptions.py

@@ -1,0 +1,90 @@
+from typing import Dict, List
+from .Snippet import Snippet
+
+
+class SnippetException(Exception):
+    def __init__(self, prelude: str, line_number: int = None):
+        if line_number is None:
+            super().__init__(prelude)
+        else:
+            super().__init__(f"{prelude}, at line {line_number}.")
+
+
+class SnippetAlreadyExistException(SnippetException):
+    """
+    Raise in snippet if 2 snippet with same name in file.
+    """
+
+    def __init__(self, name_snippet: str, line_number: int):
+        super().__init__(f"Snippet \"{name_snippet}\" already exist", line_number)
+
+    @staticmethod
+    def check(name: str, line_number: int, open_snippets: Dict[str, Snippet], close_snippets: List[Snippet]) -> None:
+        """
+        check if new snippet not already exist
+
+        :param name: name to check
+        :param line_number: line number found this name
+        :param open_snippets: snippets already open
+        :param close_snippets: snippet already closed
+        :raise: SnippetAlreadyExistException if already open or close
+        """
+        list_names = set(open_snippets.keys()) | set(map(lambda x: x.name, close_snippets))
+        if name in list_names:
+            raise SnippetAlreadyExistException(name, line_number)
+
+
+class SnippetAlreadyCloseException(SnippetException):
+    """
+    Raise if snippet is already close.
+    """
+
+    def __init__(self, name_snippet: str, line_number: int):
+        super().__init__(f"Snippet \"{name_snippet}\" already close", line_number)
+
+    @staticmethod
+    def check(name: str, line_number: int, close_snippets: List[Snippet]) -> None:
+        """
+        check if snippet not already closed
+
+        :param name: name to check
+        :param line_number: line number found this name
+        :param close_snippets: snippet already closed
+        :raise: SnippetAlreadyCloseException if exist in close
+        """
+        list_names = set(map(lambda x: x.name, close_snippets))
+        if name in list_names:
+            raise SnippetAlreadyCloseException(name, line_number)
+
+
+class SnippetNotBeginException(SnippetException):
+    """
+    Raise if snippet not exist.
+    """
+
+    def __init__(self, name_snippet: str, line_number: int):
+        super().__init__(f"Snippet \"{name_snippet}\" not begin (cannot be closed)", line_number)
+
+    @staticmethod
+    def check(name: str, line_number: int, open_snippets: Dict[str, Snippet]) -> None:
+        """
+        check if snippet close in open snippets
+
[…]

### 2021-06-30 doc/part-hydras.tex
resolve mistakes

@@
 \vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_mult.html}{OrdinalNotations.ON\_Generic}}
+\noindent\emph{From
+Library~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}}
 
@@
 The following lemma is an application of the strict inequality 
-$\alpha < \omega ^\alpha.$ If \alpha is strongly accessible, then, by definition,
-$\omega^\alpha$ is accessible, thus \alpha is \emph{a fortiori} accessible.
+$\alpha < \omega ^\alpha$. If $\alpha$ is strongly accessible, then, by definition,
+$\omega^\alpha$ is accessible, thus $\alpha$ is \emph{a fortiori} accessible.
 

### 2021-06-30 doc/chapter-primrec.tex
correction of typos
-  Corrections in latex files -  Export distributivity and associativity properties to type E0 -  minor changes (order of lemmas)
Fix pdf links (#41)

@@
 Although  the following lemmas are already proven in 
-\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec.v},
+\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec},
 we wrote alternate proofs in 
@@
 Projections are proved primitive recursive, case by case (many examples in 
-\href{../theories/html/hydras.primRec.html}{Ackermann.primRec.v}).
+\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}).
 \emph{Please notice again that the name of the projection follows the mathematical tradition, 
@@
 \noindent
-\emph{From \href{../theories/html/hydras.primRec.html}{Ackermann.primRec}.}
+\emph{From \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.}
 
@@
 
-The lemma \texttt{plusIsPR} is already proven in \href{../theories/html/hydras.primRec.html}{Ackermann.primRec}. We present in 
+The lemma \texttt{plusIsPR} is already proven in \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}. We present in 
 \href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}
@@
 You may start this exercise with the file
-\url{../exercises/primrec/MorePRExamples.v}.
+    \href{https://https://github.com/coq-community/hydra-battles/blob/master/exercises/primrec/MorePRExamples.v}{exercises/primrec/MorePRExamples.v}.
 \end{exercise}
@@
 \emph{You may start this exercise with
-\url{../exercises/primrec/MinPR.v}.}
+    \href{https://https://github.com/coq-community/hydra-battles/blob/master/exercises/primrec/MinPR.v}{exercises/primrec/MinPR.v}.}
 
@@
 \emph{You may start this exercise with the file
-\url{../exercises/primrec/FibonacciPR.v}.}
+    \href{https://github.com/coq-community/hydra-battles/blob/master/exercises/primrec/FibonacciPR.v}{exercises/primrec/FibonacciPR.v}.}
 
@@
 \emph{You may start this exercise with the file
-\url{../exercises/primrec/isqrt.v}.}
+    \href{https://github.com/coq-community/hydra-battles/blob/master/exercises/primrec/isqrt.v}{exercises/primrec/isqrt.v}.}
 
@@
 Lemma Ack_S_0 m : Ack (S m) 0 = Ack m 1. 
-Proof. now cbn. Qed.
+Proof. reflexivity. Qed.
 
@@
     Ack (S m) (S p) = Ack m (Ack (S m) p).
-Proof. now cbn. Qed.
+Proof. reflexivity. Qed.
 \end{Coqsrc}
@@
 
-We prove also a strict version of this lemma, thanks to the following property (proved in Library\href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack}~).
+We prove also a strict version of this lemma, thanks to the following property (proved in Library
+\href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack}~).
 
@@
 \noindent
-\emph{From \href{../theories/html/hydras.MoreAck.AcknotPR.html}{MoreAck.AckNotPR}.}
+\emph{From \href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR}.}
 
@@
 \section{The length of standard hydra battles}
+\label{sect:battle-length-notPR}
 
@@
 Looking for lemmas of the form $H'_\beta(k)\leq H'_\alpha(k)$, we find this one (from our library
-\href{../theories/html/hydras.Epsilon0.Hprimehtml}{Epsilon0.Hprime}):
+\href{../theories/html/hydras.Epsilon0.Hprime.html}{Epsilon0.Hprime}):
 
@@
 Remark SmNotPR : isPR 1 (fun n => S (m n)) -> False.
+(* ... *)
 

### 2021-06-30 doc/cutting-table/src/Snippet.py
parse coq+rst and extract snippet to latex

@@ -27,16 +27,3 @@ class Snippet:
 
     def __str__(self) -> str:
         return self.content
-
-    def get_path(self, directory: Path = Path('.')) -> Path:
-        """
-        get path of file with directory
-
-        :param directory: destination directory
-        :return: Path of file
-        """
-        return directory / (self.name + self.LATEX_EXTENSION)
-
-    def write(self, directory: Path):
-        with open(self.get_path(directory), 'w') as file:
-            file.write(self.content)

### 2021-06-30 doc/cutting-table/src/cli.py

@@ -11,7 +11,7 @@ def build_parser() -> argparse.ArgumentParser:
     DESCRIPTION = "Extract snippet in latex file generate by alectryon."
     parser = argparse.ArgumentParser(DESCRIPTION)
 
-    INPUT_FILES_HELP = "latex file '.tex'"
+    INPUT_FILES_HELP = "coq file '.v'"
     parser.add_argument("input",
                         type=Path,
                         help=INPUT_FILES_HELP)
@@ -21,27 +21,83 @@ def build_parser() -> argparse.ArgumentParser:
                         default=Path("."),
                         type=Path,
                         help=OUPUT_DIR_HELP)
+
+    # serapi (copy from alectryon.cli)
+    SUBP_HELP = "Pass arguments to the SerAPI process"
+    subp = parser.add_argument_group("Subprocess arguments", SUBP_HELP)
+
+    SERTOP_ARGS_HELP = "Pass a single argument to SerAPI (e.g. -Q dir,lib)."
+    subp.add_argument("--sertop-arg", dest="sertop_args",
+                      action="append", default=[],
+                      metavar="SERAPI_ARG",
+                      help=SERTOP_ARGS_HELP)
+
+    I_HELP = "Pass -I DIR to the SerAPI subprocess."
+    subp.add_argument("-I", "--ml-include-path", dest="coq_args_I",
+                      metavar="DIR", nargs=1, action="append",
+                      default=[], help=I_HELP)
+
+    Q_HELP = "Pass -Q DIR COQDIR to the SerAPI subprocess."
+    subp.add_argument("-Q", "--load-path", dest="coq_args_Q",
+                      metavar=("DIR", "COQDIR"), nargs=2, action="append",
+                      default=[], help=Q_HELP)
+
+    R_HELP = "Pass -R DIR COQDIR to the SerAPI subprocess."
+    subp.add_argument("-R", "--rec-load-path", dest="coq_args_R",
+                      metavar=("DIR", "COQDIR"), nargs=2, action="append",
+                      default=[], help=R_HELP)
+
+    EXPECT_UNEXPECTED_HELP = "Ignore unexpected output from SerAPI"
+    parser.add_argument("--expect-unexpected", action="store_true",
+                        default=False, help=EXPECT_UNEXPECTED_HELP)
+
     return parser
 
 
+def post_process_arguments(args: argparse.Namespace) -> argparse.Namespace:
+    """
+    post process sertop arguments
+
+    :param args: args parsed
+    :return: new name space
+    """
[…]

### 2021-06-30 doc/cutting-table/src/docutils.py

@@ -1,0 +1,57 @@
+import re
+
+
+def register_docutils(sertop_args):
+    """
+    setup docutils
+    :param sertop_args: sertop arguments to used
+    :return:
+    """
+    from alectryon.docutils import setup, AlectryonTransform
+    AlectryonTransform.SERTOP_ARGS = sertop_args
+    setup()
+
+
+def coq_rst_to_latex(source: str) -> str:
+    """
+    convert coq content to latex (please setup doctutils with register_docutils before)
+    (function inspired by `_gen_docutils` in alectryon cli)
+    :param source: source to conver
+    :return:
+    """
+    from alectryon.docutils import LatexWriter
+    from docutils.core import publish_string
+    from alectryon.docutils import \
+        RSTCoqParser as Parser, \
+        RSTCoqStandaloneReader as Reader
+
+    settings_overrides = {
+        'traceback': True,
+        'embed_stylesheet': False,
+        'stylesheet_path': None,
+        'stylesheet_dirs': [],
+        'alectryon_banner': False,
+        'alectryon_vernums': False,
+        'input_encoding': 'utf-8',
+        'output_encoding': 'utf-8',
+    }
+
+    parser = Parser()
+    return publish_string(
+        source=source.encode("utf-8"),
+        reader=Reader(parser), reader_name=None,
+        parser=parser, parser_name=None,
+        writer=LatexWriter(), writer_name=None,
+        settings=None, settings_spec=None,
+        settings_overrides=settings_overrides, config_section=None,
+        enable_exit_status=True).decode("utf-8")
+
+
+PATTERN_BODY = re.compile(r"\\begin\{document\}(.+)\\end\{document\}")
+BEGIN_DOC = r"\begin{document}"
+END_DOC = r"\end{document}"
+def coq_rst_to_latex_snippet(source: str):
+    latex_source = coq_rst_to_latex(source)
+    start_index = latex_source.find(BEGIN_DOC) + len(BEGIN_DOC)
+    end_index = latex_source.find(END_DOC, start_index)
+    return latex_source[start_index:end_index]

### 2021-06-30 doc/cutting-table/src/extract.py

@@ -20,7 +20,7 @@ def get_reader(path: Path) -> str:
         yield from enumerate(file, 1)
 
 
-def add_to_snippets(content: str, snippets: Dict[str, Snippet]) -> None:
+def add_all_snippets(content: str, snippets: Dict[str, Snippet]) -> None:
     """
     add content to all snippet
 
@@ -31,53 +31,10 @@ def add_to_snippets(content: str, snippets: Dict[str, Snippet]) -> None:
         snippet.add_content(content)
 
 
-# text pattern
+# snippet pattern
 BEGIN_STATE = "begin"
 END_STATE = "end"
-TEXT_PATTERN = re.compile(rf"\s*\\({BEGIN_STATE}|{END_STATE})\{{txt}}")
-
-# snippet pattern
-SNIPPET_PATTERN = re.compile(rf".*\\PY{{c}}{{\(\*\~+({BEGIN_STATE}|{END_STATE})\~+snippet\~+(\w+)\~+\*\)}}")
-
-BEGIN_ALECTRYON = "\\begin{alectryon}\n"
-END_ALECTRYON = "\\end{alectryon}\n"
-
-def extract_snippets_text(reader, begin_line: str,
-                          open_snippets: Dict[str, Snippet], close_snippets: List[Snippet]) -> str:
-    """
-    This function is call by extract snippet if we are in text part.
-    This function open new snippets found, or close it.
-    This function end when text part is closed.
-
-    :param reader: reader used
-    :param begin_line: begin line used
-    :param open_snippets: snippet not end
-    :param close_snippets: snippet already end
-    :return: content to add
-    """
-    content = BEGIN_ALECTRYON + begin_line  # TODO found other method
-    for num, line in reader:
-        # match open or close snippet
-        if (match := SNIPPET_PATTERN.match(line)) is not None:
-            state, name = match.groups()
-            if state == BEGIN_STATE:
-                SnippetAlreadyExistException.check(name, num, open_snippets, close_snippets)
-                open_snippets[name] = Snippet(name)
-            else:
-                SnippetAlreadyCloseException.check(name, num, close_snippets)
-                SnippetNotBeginException.check(name, num, open_snippets)
-                open_snippets[name].add_content(END_ALECTRYON)
-                close_snippets.append(open_snippets.pop(name))
-            continue
-
-        # add line to content
-        content += line
-
-        # match end of text
-        if (match := TEXT_PATTERN.match(line)) is not None \
-            and match.group(1) == END_STATE:
[…]

### 2021-06-30 doc/cutting-table/src/snippetExceptions.py

@@ -1,4 +1,4 @@
-from typing import Dict, List
+from typing import Dict, Iterable, List
 from .Snippet import Snippet
 
 
@@ -85,6 +85,6 @@ class SnippetNotEndException(SnippetException):
     Raise if snippet is have not end.
     """
 
-    def __init__(self, snippets: List[Snippet]):
+    def __init__(self, *snippets: Iterable[Snippet]):
         names = ",".join(snippet.name for snippet in snippets)
         super().__init__(f"Snippets \"{names}\" never close")

### 2021-06-30 doc/cutting-table/montage1.py
add Readme

@@ -1,56 +1,0 @@
-#!/usr/bin/env python3
-import json
-import sys
-from os.path import join, dirname
-from pprint import pprint
-
-root = join(dirname(__file__), "..")
-sys.path.insert(0, root)
-sys.path.insert(0, '/Users/casteranpierre/alectryon')
-
-# import alectryon.core
-# alectryon.core.DEBUG = True
-
-def api_annotate():
-    from alectryon.core import annotate
-    annotated = annotate(["Check 1."], ("-Q", "{},logical_name".format(root)))
-    pprint(annotated)
-
-def annotated_to_json():
-    from alectryon.core import annotate
-    from alectryon.json import json_of_annotated
-    annotated = annotate([r"Goal True /\ True. split. ", "all: eauto."],
-                         ("-Q", "{},logical_name".format(root)))
-    print(json.dumps(json_of_annotated(annotated)))
-
-#Todo  get JS  as an argument on the command line
-
-JS = """
-[[{
-            "_type": "sentence",
-            "contents": "Inductive Hydra : Set :=\n|  node :  Hydrae -> Hydra\nwith Hydrae : Set :=\n| hnil : Hydrae\n| hcons : Hydra -> Hydrae -> Hydrae.",
-            "messages": [],
-            "goals": []
-        }]]
-"""
-
-
-# print(JS) # type of JS ?
-
-
-def latex_of_movie():
-    from alectryon.json import annotated_of_json
-    from alectryon.latex import LatexGenerator
-    from alectryon.pygments import highlight_latex
-    annotated = annotated_of_json(json.loads(JS, strict=False))
-    for ltx in LatexGenerator(highlight_latex).gen(annotated):
-      #  print("<latex>")
-        print(ltx)
-      #  print("</latex>")
-
-def main():
-#    api_annotate()
-    latex_of_movie()
-
-if __name__ == '__main__':
-    main()

### 2021-06-30 doc/cutting-table/src/__init__.py

@@ -1,0 +1,11 @@
+from .Snippet import Snippet
+from .docutils import coq_rst_to_latex_snippet, register_docutils
+from .extract import extract_snippets
+
+__all__ = \
+    [
+        "Snippet",
+        "coq_rst_to_latex_snippet",
+        "extract_snippets",
+        "register_docutils",
+    ]

### 2021-06-30 doc/cutting-table/src/cli.py

@@ -8,7 +8,7 @@ def build_parser() -> argparse.ArgumentParser:
 
     :return: Argument parser
     """
-    DESCRIPTION = "Extract snippet in latex file generate by alectryon."
+    DESCRIPTION = "Extract snippet in coq file (.v) generate by alectryon."
     parser = argparse.ArgumentParser(DESCRIPTION)
 
     INPUT_FILES_HELP = "coq file '.v'"

### 2021-07-01 doc/cutting-table/src/Extractor.py
Make a version snippet extractor with latex conversion
Use latex format to extract snippets

@@ -52,7 +52,7 @@ class SnippetExtractor:
         """
         # check
         if snippet_name in self._list_names(self.__snippet_close) or \
-            snippet_name in self._list_names(self.__snippet_open):
+            snippet_name in self._list_names(self.__snippet_open.values()):
             raise SnippetAlreadyExistException(snippet_name, line_number)
 
         # add snippet in open dict

### 2021-07-01 doc/cutting-table/src/Snippet.py

@@ -35,12 +35,13 @@ class Snippet:
         self.content += content_to_add
 
     def _add_tags(self, template: str, tags: List[str], reverse: bool = False):
-        bias = 0
+        indent_func = lambda i: self.INDENT * i
         if reverse:
-            bias = len(tags) - 1
+            max_i = len(tags) - 1
+            indent_func = lambda i: self.INDENT * (max_i - i)
             tags = reversed(tags)
-        self.content += "\n".join(self.INDENT * (bias - i) + template.format(tag=tag)
-                                  for i, tag in enumerate(tags)) + "\n"
+
+        self.content += "\n".join(indent_func(i) + template.format(tag=tag) for i, tag in enumerate(tags)) + "\n"
 
     def close(self, tags: List[str]) -> None:
         """

### 2021-07-01 doc/cutting-table/src/Extractor.py
Make a version snippet extractor with latex conversion
Use latex format to extract snippets

@@ -1,0 +1,126 @@
+import re
+from typing import Iterable, List
+
+from .Reader import Reader
+from .Snippet import Snippet
+from .snippetExceptions import \
+    SnippetAlreadyCloseException, \
+    SnippetAlreadyExistException, \
+    SnippetNotBeginException, \
+    SnippetNotEndWarning
+
+BEGIN_STATE = "begin"
+END_STATE = "end"
+STATES_GROUP = fr"({BEGIN_STATE}|{END_STATE})"
+NAME_GROUP = r"(\w+)"
+
+LATEX_TAG_PATTERN = re.compile(rf"\s*\\{STATES_GROUP}{{{NAME_GROUP}}}")
+SNIPPET_PATTERN = re.compile(rf".*\\PY{{c}}{{\(\*\~+{STATES_GROUP}\~+snippet\~+{NAME_GROUP}\~+\*\)}}")
+
+
+class SnippetExtractor:
+    def __init__(self, reader: Reader):
+        self.reader = reader
+        self.__snippet_open = {}
+        self.__snippet_close = []
+        self.stack_tag = []
+
+    @staticmethod
+    def _list_names(snippets: Iterable[Snippet]) -> Iterable[str]:
+        """
+        get list of name sinppets
+        :param snippets: list of snipepts
+        :return: itterable
+        """
+        return map(lambda snippet: snippet.name, snippets)
+
+    def _match_tag(self, statement: str, tag: str, line_num: int) -> None:
+        if statement == BEGIN_STATE:
+            self.stack_tag.append(tag)
+
+        elif (last_tag := self.stack_tag.pop()) != tag:
+            raise Exception(f"tag '{tag}' close before '{last_tag}' at line {line_num}.")
+
+    def _open_snippet(self, snippet_name: str, line_number: int) -> None:
+        """
+        check if new snippet not already exist
+        and open snippet
+
+        :param snippet_name: name to check
+        :param line_number: line number found this name
+        :raise: SnippetAlreadyExistException if already open or close
+        """
+        # check
+        if snippet_name in self._list_names(self.__snippet_close) or \
+            snippet_name in self._list_names(self.__snippet_open):
+            raise SnippetAlreadyExistException(snippet_name, line_number)
+
+        # add snippet in open dict
+        self.__snippet_open[snippet_name] = Snippet(snippet_name, self.stack_tag)
+
+    def _close_snippet(self, snippet_name: str, line_number: int) -> None:
+        """
+        check if snippet not already closed or
[…]

### 2021-07-01 doc/cutting-table/src/Snippet.py

@@ -1,12 +1,19 @@
-from pathlib import Path
+from typing import List
+
+from .docutils import LATEX_TAG_BEGIN_FORMAT, LATEX_TAG_END_FORMAT
+from .snippetExceptions import SnippetException
 
 
 class Snippet:
-    LATEX_EXTENSION = ".tex"
+    BEGIN_TEMPLATE = LATEX_TAG_BEGIN_FORMAT
+    END_TEMPLATE = LATEX_TAG_END_FORMAT
+    INDENT = "  "
 
-    def __init__(self, name: str):
+    def __init__(self, name: str, tags=[]):
         self.__name = name
         self.content = ""
+        self._add_tags(self.BEGIN_TEMPLATE, tags)
+        self.open = True
 
     @property
     def name(self) -> str:
@@ -23,7 +30,31 @@ class Snippet:
         :param content_to_add: content added
         :return: None
         """
+        if self.is_close():
+            raise SnippetException(f"Snippet {self.name} already closed.")
         self.content += content_to_add
+
+    def _add_tags(self, template: str, tags: List[str], reverse: bool = False):
+        bias = 0
+        if reverse:
+            bias = len(tags) - 1
+            tags = reversed(tags)
+        self.content += "\n".join(self.INDENT * (bias - i) + template.format(tag=tag)
+                                  for i, tag in enumerate(tags)) + "\n"
+
+    def close(self, tags: List[str]) -> None:
+        """
+        close snippet
+        """
+        self._add_tags(self.END_TEMPLATE, tags, reverse=True)
+        self.open = False
+
+    def is_close(self) -> bool:
+        """
+        True if Snippet is closed
+        :return: bool
+        """
+        return not self.open
 
     def __str__(self) -> str:
         return self.content

### 2021-07-01 doc/cutting-table/src/__init__.py

@@ -1,11 +1,0 @@
-from .Snippet import Snippet
-from .docutils import coq_rst_to_latex_snippet, register_docutils
-from .extract import extract_snippets
-
-__all__ = \
-    [
-        "Snippet",
-        "coq_rst_to_latex_snippet",
-        "extract_snippets",
-        "register_docutils",
-    ]

### 2021-07-01 doc/cutting-table/src/cli.py

@@ -22,6 +22,11 @@ def build_parser() -> argparse.ArgumentParser:
                         type=Path,
                         help=OUPUT_DIR_HELP)
 
+    DUMP_LATEX_HELP = "dump latex"
+    parser.add_argument("-d", "--dump-complete-latex",
+                        action="store_true",
+                        help=DUMP_LATEX_HELP)
+
     # serapi (copy from alectryon.cli)
     SUBP_HELP = "Pass arguments to the SerAPI process"
     subp = parser.add_argument_group("Subprocess arguments", SUBP_HELP)
@@ -69,6 +74,8 @@ def post_process_arguments(args: argparse.Namespace) -> argparse.Namespace:
         args.sertop_args.extend(("-Q", ",".join(pair)))
 
     return args
+
+
 def parse_args() -> argparse.Namespace:
     """
     parse argument
@@ -82,22 +89,36 @@ def parse_args() -> argparse.Namespace:
 EXTENSION_LATEX = ".tex"
 
 
+def get_path_latex(output_dir: Path, name: str):
+    return output_dir / (name + EXTENSION_LATEX)
+
+
+def make_latex_file(content: str, path: Path):
+    path.parent.mkdir(parents=True, exist_ok=True)
+    with open(path, 'w') as file:
+        file.write(content)
+
+
 def main():
-    from .extract import extract_snippets
-    from .docutils import coq_rst_to_latex_snippet, register_docutils
+    from .Extractor import SnippetExtractor
+    from .docutils import CoqToLatexReader, register_docutils
     args = parse_args()
-    list_snippets = extract_snippets(args.input)
+    input_name = args.input.stem
+    output_dir = args.output_dir / input_name
 
-    output_dir = args.output_dir / args.input.stem
-    if list_snippets and not output_dir.exists():
-        print(f"Directory {output_dir} not exist, create it.")
-        output_dir.mkdir(parents=True)
-
+    # make latex
     register_docutils(args.sertop_args)
-    for snippet in list_snippets:
-        path = output_dir / (snippet.name + EXTENSION_LATEX)
+    reader = CoqToLatexReader(args.input)
+    snippet_extractor = SnippetExtractor(reader)
+
+    # dump latex to debug
+    if args.dump_complete_latex:
+        path_latex = get_path_latex(output_dir, input_name)
+        print(f"Make latex file {path_latex}")
+        make_latex_file(reader.content_latex, path_latex)
+
+    # write snippets files
[…]

### 2021-07-01 doc/cutting-table/src/docutils.py

@@ -1,5 +1,10 @@
-import re
+from pathlib import Path
 
+from .Reader import Reader
+
+
+LATEX_TAG_BEGIN_FORMAT = "\\begin{{{tag}}}"
+LATEX_TAG_END_FORMAT = "\\end{{{tag}}}"
 
 def register_docutils(sertop_args):
     """
@@ -47,11 +52,21 @@ def coq_rst_to_latex(source: str) -> str:
         enable_exit_status=True).decode("utf-8")
 
 
-PATTERN_BODY = re.compile(r"\\begin\{document\}(.+)\\end\{document\}")
-BEGIN_DOC = r"\begin{document}"
-END_DOC = r"\end{document}"
-def coq_rst_to_latex_snippet(source: str):
-    latex_source = coq_rst_to_latex(source)
-    start_index = latex_source.find(BEGIN_DOC) + len(BEGIN_DOC)
-    end_index = latex_source.find(END_DOC, start_index)
-    return latex_source[start_index:end_index]
+class CoqToLatexReader(Reader):
+    DOCUMENT_BEGIN = LATEX_TAG_BEGIN_FORMAT.format(tag="document")
+    DOCUMENT_END = LATEX_TAG_END_FORMAT.format(tag="document")
+
+    def __init__(self, path_coq: Path):
+        with open(path_coq, 'r') as file:
+            self.content_coq = file.read()
+        self.content_latex = coq_rst_to_latex(self.content_coq)
+
+    def __iter__(self):
+        start = False
+        for num, line in enumerate(self.content_latex.split('\n'), 1):
+            if line == self.DOCUMENT_BEGIN:
+                start = True
+            elif line == self.DOCUMENT_END:
+                return
+            elif start:
+                yield num, line + "\n"

### 2021-07-01 doc/cutting-table/src/extract.py

@@ -1,65 +1,0 @@
-from pathlib import Path
-from typing import Dict, List
-import re
-from .Snippet import Snippet
-from .snippetExceptions import \
-    SnippetNotBeginException, \
-    SnippetAlreadyCloseException, \
-    SnippetAlreadyExistException, \
-    SnippetNotEndException
-
-
-def get_reader(path: Path) -> str:
-    """
-    reader who yield lines
-
-    :param path: path of file to read
-    :return: line read
-    """
-    with open(path, 'r') as file:
-        yield from enumerate(file, 1)
-
-
-def add_all_snippets(content: str, snippets: Dict[str, Snippet]) -> None:
-    """
-    add content to all snippet
-
-    :param content: content to add
-    :param snippets: snippet used
-    """
-    for snippet in snippets.values():
-        snippet.add_content(content)
-
-
-# snippet pattern
-BEGIN_STATE = "begin"
-END_STATE = "end"
-SNIPPET_PATTERN = re.compile(rf"\s*\(\*\s*({BEGIN_STATE}|{END_STATE})\s+snippet\s+(\w+)\s*\*\)")
-
-
-def extract_snippets(path: Path) -> List[Snippet]:
-    """
-    parse latex file and return a list of snippet
-    :param path: path of file to extract snippets
-    :return: list of snippets
-    """
-    reader = get_reader(path)
-    open_snippets = {}
-    close_snippets = []
-    for num, line in reader:
-        if (match := SNIPPET_PATTERN.match(line)) is not None:
-            state, name = match.groups()
-            if state == BEGIN_STATE:
-                SnippetAlreadyExistException.check(name, num, open_snippets, close_snippets)
-                open_snippets[name] = Snippet(name)
-            else:  # end state
-                SnippetAlreadyCloseException.check(name, num, close_snippets)
-                SnippetNotBeginException.check(name, num, open_snippets)
-                close_snippets.append(open_snippets.pop(name))
-            continue
-        add_all_snippets(line, open_snippets)
-
-    if open_snippets:
-        raise SnippetNotEndException(open_snippets.values())
-
-    return close_snippets

### 2021-07-01 doc/cutting-table/src/snippetExceptions.py

@@ -1,5 +1,5 @@
-from typing import Dict, Iterable, List
-from .Snippet import Snippet
+from typing import Iterable
+
 
 
 class SnippetException(Exception):
@@ -18,21 +18,6 @@ class SnippetAlreadyExistException(SnippetException):
     def __init__(self, name_snippet: str, line_number: int):
         super().__init__(f"Snippet \"{name_snippet}\" already exist", line_number)
 
-    @staticmethod
-    def check(name: str, line_number: int, open_snippets: Dict[str, Snippet], close_snippets: List[Snippet]) -> None:
-        """
-        check if new snippet not already exist
-
-        :param name: name to check
-        :param line_number: line number found this name
-        :param open_snippets: snippets already open
-        :param close_snippets: snippet already closed
-        :raise: SnippetAlreadyExistException if already open or close
-        """
-        list_names = set(open_snippets.keys()) | set(map(lambda x: x.name, close_snippets))
-        if name in list_names:
-            raise SnippetAlreadyExistException(name, line_number)
-
 
 class SnippetAlreadyCloseException(SnippetException):
     """
@@ -41,20 +26,6 @@ class SnippetAlreadyCloseException(SnippetException):
 
     def __init__(self, name_snippet: str, line_number: int):
         super().__init__(f"Snippet \"{name_snippet}\" already close", line_number)
-
-    @staticmethod
-    def check(name: str, line_number: int, close_snippets: List[Snippet]) -> None:
-        """
-        check if snippet not already closed
-
-        :param name: name to check
-        :param line_number: line number found this name
-        :param close_snippets: snippet already closed
-        :raise: SnippetAlreadyCloseException if exist in close
-        """
-        list_names = set(map(lambda x: x.name, close_snippets))
-        if name in list_names:
-            raise SnippetAlreadyCloseException(name, line_number)
 
 
 class SnippetNotBeginException(SnippetException):
@@ -65,26 +36,12 @@ class SnippetNotBeginException(SnippetException):
     def __init__(self, name_snippet: str, line_number: int):
         super().__init__(f"Snippet \"{name_snippet}\" not begin (cannot be closed)", line_number)
 
-    @staticmethod
-    def check(name: str, line_number: int, open_snippets: Dict[str, Snippet]) -> None:
-        """
-        check if snippet close in open snippets
 
-        :param name: name to check
-        :param line_number: line number found this name
-        :param open_snippets: snippets already open
[…]

### 2021-07-01 doc/cutting-table/src/Extractor.py
fix regex patter

@@ -14,8 +14,8 @@ END_STATE = "end"
 STATES_GROUP = fr"({BEGIN_STATE}|{END_STATE})"
 NAME_GROUP = r"(\w+)"
 
-LATEX_TAG_PATTERN = re.compile(rf"\s*\\{STATES_GROUP}{{{NAME_GROUP}}}")
-SNIPPET_PATTERN = re.compile(rf".*\\PY{{c}}{{\(\*\~+{STATES_GROUP}\~+snippet\~+{NAME_GROUP}\~+\*\)}}")
+LATEX_TAG_PATTERN = re.compile(rf"\s*\\{STATES_GROUP}{{{NAME_GROUP}}}(.*)") #"\s*\\(begin|end){(\w+)}"
+SNIPPET_PATTERN = re.compile(rf".*\\PY{{c}}{{\(\*\~*{STATES_GROUP}\~+snippet\~+{NAME_GROUP}\~*\*\)}}")
 
 
 class SnippetExtractor:
@@ -34,9 +34,11 @@ class SnippetExtractor:
         """
         return map(lambda snippet: snippet.name, snippets)
 
-    def _match_tag(self, statement: str, tag: str, line_num: int) -> None:
+    def _match_tag(self, statement: str, tag: str, after: str, line_num: int) -> None:
         if statement == BEGIN_STATE:
             self.stack_tag.append(tag)
+            if (match:=LATEX_TAG_PATTERN.match(after)) is not None:
+                self._match_tag(*match.groups(), line_num)
 
         elif (last_tag := self.stack_tag.pop()) != tag:
             raise Exception(f"tag '{tag}' close before '{last_tag}' at line {line_num}.")

### 2021-07-01 doc/hydras.tex
minor corrections

@@
 \label{sec:orgheadline5}
-    Many thanks to Yves Bertot, \'Evelyne Contejean,  Florian Hatat,  David Ilcinkas, 
-Pascal Manoury,  Karl Palmskog, Sylvain Salvati, Alan Schmitt and Théo Zimmermann for their help on the elaboration of this document, and to the
+    Many thanks to Yves Bertot, \'Evelyne Contejean, Jéremy Damour,   Florian Hatat,  David Ilcinkas, 
+Pascal Manoury,  Karl Palmskog, Sylvain Salvati, Alan Schmitt and Théo Zimmermann for their help on the elaboration of thislibrary and  document, and to the
  members of the \emph{Formal Methods} team and the \emph{Coq working group} at laBRI for their helpful comments 

### 2021-07-02 doc/movies/extract_snippets.py
move cutting table in move and create edit makefile

@@ -1,5 +1,5 @@
 #!/usr/bin/env python3
-from src.cli import main
+from extract_snippets.cli import main
 
 
 if __name__ == '__main__':

### 2021-07-02 doc/movies/extract_snippets/Extractor.py

@@ -1,5 +1,5 @@
 import re
-from typing import Iterable, List
+from typing import AnyStr, Iterable, List, Match
 
 from .Reader import Reader
 from .Snippet import Snippet
@@ -14,7 +14,7 @@ END_STATE = "end"
 STATES_GROUP = fr"({BEGIN_STATE}|{END_STATE})"
 NAME_GROUP = r"(\w+)"
 
-LATEX_TAG_PATTERN = re.compile(rf"\s*\\{STATES_GROUP}{{{NAME_GROUP}}}(.*)") #"\s*\\(begin|end){(\w+)}"
+LATEX_TAG_PATTERN = re.compile(rf"\s*\\{STATES_GROUP}(\[.*\])?{{{NAME_GROUP}}}(.*)")  # "\s*(\[.*])\\(begin|end){(\w+)}"
 SNIPPET_PATTERN = re.compile(rf".*\\PY{{c}}{{\(\*\~*{STATES_GROUP}\~+snippet\~+{NAME_GROUP}\~*\*\)}}")
 
 
@@ -34,11 +34,12 @@ class SnippetExtractor:
         """
         return map(lambda snippet: snippet.name, snippets)
 
-    def _match_tag(self, statement: str, tag: str, after: str, line_num: int) -> None:
+    def _match_tag(self, match: Match[AnyStr], line_num: int) -> None:
+        statement, _, tag, after = match.groups()
         if statement == BEGIN_STATE:
             self.stack_tag.append(tag)
-            if (match:=LATEX_TAG_PATTERN.match(after)) is not None:
-                self._match_tag(*match.groups(), line_num)
+            if (match := LATEX_TAG_PATTERN.match(after)) is not None:
+                self._match_tag(match, line_num)
 
         elif (last_tag := self.stack_tag.pop()) != tag:
             raise Exception(f"tag '{tag}' close before '{last_tag}' at line {line_num}.")
@@ -83,7 +84,7 @@ class SnippetExtractor:
         snippet.close(self.stack_tag)
         self.__snippet_close.append(snippet)
 
-    def _match_snippet(self, statement: str, snippet_name: str, line_num: int):
+    def _match_snippet(self, match: Match[AnyStr], line_num: int):
         """
         call if match snippet, to open or close snippet
 
@@ -92,6 +93,7 @@ class SnippetExtractor:
         :param line_num:
         :return:
         """
+        statement, snippet_name = match.groups()
         if statement == BEGIN_STATE:
[…]

### 2021-07-02 doc/movies/extract_snippets/cli.py

@@ -22,10 +22,17 @@ def build_parser() -> argparse.ArgumentParser:
                         type=Path,
                         help=OUPUT_DIR_HELP)
 
-    DUMP_LATEX_HELP = "dump latex"
+    DUMP_LATEX_HELP = "dump latex complete file (usefull to debug)"
     parser.add_argument("-d", "--dump-complete-latex",
                         action="store_true",
                         help=DUMP_LATEX_HELP)
+
+    DUMP_LATEX_DIR_HELP = "directory to dump complete latex" \
+                          "file (please active flag '--dump-complete-latex')"
+    parser.add_argument("-D", "--dump-complete-latex-dir",
+                        type=Path,
+                        default=Path("."),
+                        help=DUMP_LATEX_DIR_HELP)
 
     # serapi (copy from alectryon.cli)
     SUBP_HELP = "Pass arguments to the SerAPI process"
@@ -107,14 +114,17 @@ def main():
     output_dir = args.output_dir / input_name
 
     # make latex
+    print(f"Convert '{args.input}' to latex.")
     register_docutils(args.sertop_args)
     reader = CoqToLatexReader(args.input)
+    print("Extract snippets.")
     snippet_extractor = SnippetExtractor(reader)
 
     # dump latex to debug
     if args.dump_complete_latex:
-        path_latex = get_path_latex(output_dir, input_name)
-        print(f"Make latex file {path_latex}")
+        path_latex = get_path_latex(args.dump_complete_latex_dir,
+                                    input_name)
+        print(f"Make complete latex file {path_latex}.")
         make_latex_file(reader.content_latex, path_latex)
 
     # write snippets files

### 2021-07-07 doc/part-hydras.tex
Simplification of some definitions

@@
 
-The following class definition, parameterized with a type $A$, a binary relation \texttt{lt} on $A$, specifies that \texttt{lt} is a well-founded strict order.
-The reflexive closure of \texttt{lt}, (called \texttt{le}, for ``less  or equal than'') is a total decidable order, implemented through a comparison function \texttt{compare}.
-The correctness of this function is expressed through Stdlib's type
+From the \coq{} user's point of view, an ordinal notation is
+a structure allowing to compare two ordinals by computation, and proving by well-founded induction.
+
+\subsubsection{The \texttt{Comparable class}}
+
+The following class, contributed by Jérémy Damour and Théo Zimmermann, allows us to apply generic lemmas and tactics
+about decidable strict orders.
+The correctness of the comparison function is expressed through Stdlib's type 
 \texttt{Datatypes.CompareSpec}.
-
-First, we import the definition of strict orders from
-the library ~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Classes.RelationClasses.html}{%
-Classes.RelationClasses}, 
-and the inductive specification of a comparison from
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Datatypes.html}{%
-Init.Datatypes}.
-
-\begin{Coqsrc}
-Variable A: Type.
-
-  Class StrictOrder (R : relation A) : Prop := {
-    StrictOrder_Irreflexive :> Irreflexive R ;
-    StrictOrder_Transitive :> Transitive R }.
-\end{Coqsrc}
 
@@
 
+\begin{Coqsrc}
+Class Comparable {A:Type} 
+  (lt: relation A)
+  (compare : A -> A -> comparison) :=
+  {
+  sto :> StrictOrder lt;
+  compare_correct: forall a b,
+      CompareSpec (a = b) (lt a b) (lt b a) (compare a b);
+  }.
+\end{Coqsrc}
+
+\subsubsection{The \texttt{ON} class}
+
+The following class definition, parameterized with a type $A$, a binary relation \texttt{lt} on $A$, specifies that \texttt{lt} is a well-founded strict order, provided with a correct comparison function.
+
 \vspace{4pt}
@@
   {
-  sto :> StrictOrder lt;
+  comp :> Comparable lt compare;
   wf : well_founded lt;
-  compare_correct :
-    forall alpha beta:A,
-      CompareSpec (alpha=beta) (lt alpha beta) (lt beta alpha)
-                  (compare alpha beta);
   }.

### 2021-07-07 doc/part-hydras.tex
Simplification of some definitions

@@
 \vspace{4pt}
-\noindent\emph{From
-Library~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}}
+\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}}
 
@@
 \vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_Omega.html}{OrdinalNotations.ON\_Omega}}
+\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations/ON_Omega.html}{OrdinalNotations.ON\_Omega}}
 
@@
 \vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_plus.html}{OrdinalNotations.ON\_plus}}
+\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations/ON_plus.html}{OrdinalNotations.ON\_plus}}
 
@@
 \emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/predSuccUnicity.v}{exercises/ordinals/predSuccUnicity.v}.}
+\url{../exercises/ordinals/predSuccUnicity.v}.}
 
@@
 \emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/lt_succ_le.v}{exercises/ordinals/lt\_succ\_le.v}.}
+\url{../exercises/ordinals/lt_succ_le.v}.}
 \end{exercise}
@@
 
-The detailed  proof script is in the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/Hydra/Omega2_Small.v}{theories/ordinals/Hydra/Omega2\_Small.v}.
+The detailed  proof script is in the file \url{../theories/html/hydras.Hydra.Omega2_Small.html}.
 
@@
 \emph{You may start this exercise with the file
-     \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/Limit_Infinity.v}{exercises/ordinals/Limit\_Infinity.v}.}
+\url{../exercises/ordinals/Limit_Infinity.v}.}
  \end{exercise}
@@
 \emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/ge_omega_iff.v}{exercises/ordinals/ge\_omega\_iff.v}.}
+\url{../exercises/ordinals/ge_omega_iff.v}.}
 \end{exercise}
@@
 \emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/T1_ltNotWf.v}{exercises/ordinals/T1\_ltNotWf.v}.}
+\url{../exercises/ordinals/T1_ltNotWf.v}.}
 \end{exercise}
@@
 The following lemma is an application of the strict inequality 
-$\alpha < \omega ^\alpha$. If $\alpha$ is strongly accessible, then, by definition,
-$\omega^\alpha$ is accessible, thus $\alpha$ is \emph{a fortiori} accessible.
+\showmath{\alpha < \omega ^\alpha}. If \showmath{\alpha} is strongly accessible, then, by definition, \showmath{\omega^\alpha} is accessible, thus \showmath{\alpha} is \emph{a fortiori} accessible.
 
@@
 
-Thus, it remains to prove that every ordinal strictly less than $\epsilon_0$
+Thus, it remains to prove that every ordinal strictly less than \showmath{\epsilon_0} 
 is strongly accessible.
@@
 
-First, we prove that, for  any \texttt{LT}-accessible term $\alpha$, $\alpha$ is 
+First, we prove that, for  any \texttt{LT}-accessible term \showmath{\alpha}, \showmath{\alpha} is 
 strongly accessible too (\emph{i.e.} any well formed
@@
 
-The proof is structured as an induction on $\alpha'$s accessibility. Let us consider
+The proof is structured as an induction on \showmath{\alpha}'s accessibility. Let us consider
 any  accessible term $\alpha$.
@@
 \label{remark:a3pat}
-The alternate proof of well-foundedness using \'Evelyne Contejean's work on
-    recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the
-    library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0.Epsilon0rpo}.
+The alternate proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
  \end{remark}
@@
 
-\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
+\emph{From Module~\href{../theories/html/hydras.Schutte.Schutte.Correctness_E0.html}{Schutte.Schutte.Correctness\_E0}}
 
@@
 \textbf{Hint}. You may use the following definition (in 
-    \href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}).
+\href{../theories/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}).
 
@@
 
-In the module
-\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
+In the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
 This part of the development is parameterized by some class $B$ of battles, which will be instantiated later to \texttt{free} or \texttt{standard}.
@@
 \vspace{4pt}
-\emph{From
-Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.KS.html}{Epsilon0.KS}}
 
@@
 be consulted in the module 
-\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}{Hydra.Epsilon0\_Needed\_Std}.
+\url{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}.
 Please note that it has the same global structure as in section\ref{sec:free-battles-case} 
@@
 \href{../theories/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
-{Epsilon0.pathS}   and 
+{\texttt{Epsilon0.Paths}}   and 
 \href{../theories/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
@@
 Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
-(please see also the module
-\href{../theories/html/hydras.Schutte.Correctness_E0.html}{hydras.Schutte.Correctness\_E0}).
+(please see also \href{../theories/html/hydras.Schutte.Correctness_E0.html}{the module \texttt{hydras.Schutte.Correctness\_E0}}).
 
@@
 The following code comes from 
- \href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}.
+ \url{../theories/html/hydras.Epsilon0.F_alpha.html}.
 
@@
 \emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
+\url{../exercises/ordinals/F_3.v}.}
 \end{exercise}
@@
 \emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
+\url{../exercises/ordinals/F_3.v}.}
 \end{exercise}
@@
 \emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/is_F_monotonous.v}{exercises/ordinals/is\_F\_monotonous.v}.}
+\url{../exercises/ordinals/is_F_monotonous.v}.}
 \end{exercise}
@@
 \emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_omega.v}{exercises/ordinals/F\_Omega.v}.
-Properties of the Ackermann function are studied in
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/Ack.v}{theories/ordinals/MoreAck/Ack.v} and
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/AckNotPR.v}{theories/ordinals/MoreAck/AckNotPR.v}
-.}
+\url{../exercises/ordinals/F_omega.v}. Properties of the Ackermann function are studied in 
+\url{../theories/ordinals/MoreAck/Ack.v} and
+\url{../theories/ordinals/MoreAck/AcknotPR.v}.}
 \end{exercise}
@@
 \texttt{ordinal} (The  class \texttt{WO} is defined in
-Module~\href{../theories/html/hydras.Schutte.Well_Orders.html}{Schutte.Well\_Orders.v}).
+Module~\href{../theories/html/hydras.Schutte.Well_Orders.html}%
+{\texttt{Well\_Orders.v}}).
 
@@
 We prove the inequality  $\alpha<\omega^\alpha$ whenever $\alpha < \epsilon_0$.
-\emph{Note that this condition was implicit in Module
-\href{../theories/html/hydras.Epsilon0.T1.html\#lt_phi0}{Epsilon0.T1}.}
+\emph{Note that this condition was implicit in Module \href{../theories/html/hydras.Epsilon0/T1.html\#lt_phi0}%
+{\texttt{Epsilon0.T1}}.}
 
@@
 \emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/schutte_cnf_counter_example.v}{exercises/ordinals/schutte\_cnf\_counter\_example.v}.}
+\url{../exercises/ordinals/schutte_cnf_counter_example.v}.}
 \end{exercise}
@@
 \begin{remark}
-The sub-directory
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/Gamma0}{theories/oridinals/Gamma0} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form.
+The sub-directory \href{../theories/html/hydras.Gamma0.html}%
+{\texttt{theories/Gamma0}} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form. 
 \end{remark}
@@
 Our library 
-\href{../theories/html/hydras.Schutte.Correctness_E0.html}%
+\href{../theories/html/hydras.Schutte.correctness_E0.html}%
 {\texttt{Schutte.Correctness\_E0}} establishes the link between two very different modelizations of ordinal numbers. In other words, it ``validates'' a data structure in terms of

### 2021-07-07 doc/hydras.tex
apply suggested changes

--- 2021-06-28 doc/hydras.tex
@@
 \label{sec:orgheadline5}
-    Many thanks to Yves Bertot, \'Evelyne Contejean,  Florian Hatat,  David Ilcinkas, 
-Pascal Manoury,  Karl Palmskog, Sylvain Salvati, Alan Schmitt and Théo Zimmermann for their help on the elaboration of this document, and to the
+    Many thanks to Yves Bertot, \'Evelyne Contejean, Jéremy Damour,   Florian Hatat,  David Ilcinkas, 
+Pascal Manoury,  Karl Palmskog, Sylvain Salvati, Alan Schmitt and Théo Zimmermann for their help on the elaboration of this library and  document, and to the
  members of the \emph{Formal Methods} team and the \emph{Coq working group} at laBRI for their helpful comments 

### 2021-07-07 doc/part-hydras.tex
apply suggested changes

@@
 
-%The following definitions allow us to make implicit several guessable arguments.
+We give  also a few handy definitions and lemmas for any ordinal notation.
+
 \label{sect:on-lt-notation}
@@
 \begin{Coqsrc}
-Definition on_t  {A:Type}{lt: relation A}
+Section Definitions.
+
+  Context  {A:Type}{lt : relation A}
            {compare : A -> A -> comparison}
-            {on : ON lt compare} := A.
-
-Definition ON_compare {A:Type}{lt: relation A}
-            {compare : A -> A -> comparison}
-            {on : ON lt compare} := compare.
-
-Definition ON_lt {A:Type}{lt: relation A}
-           {compare : A -> A -> comparison}
-           {on : ON lt compare} := lt.
+           {on : ON lt compare}.
+
+  #[using="All"]
+   Definition ON_t := A.
+
+  #[using="All"]
+   Definition ON_compare := compare.
+
+  #[using="All"]
+   Definition ON_lt := lt.
+
+  #[using="All"]
+   Definition ON_le:  relation A := leq lt.
+
+  #[using="All"]
+   Definition measure_lt {B : Type} (m : B -> A) : relation B :=
+    fun x y =>  ON_lt (m x) (m y).
+
+  #[using="All"]
+   Lemma wf_measure {B : Type} (m : B -> A) :
+    well_founded (measure_lt m).
+
+  #[using="All"]
+   Definition ZeroLimitSucc_dec :=
+    forall alpha,
+      {Least alpha} +
+      {Limit alpha} +
+      {beta: A | Successor alpha beta}.
+
+  (** The segment called [O alpha] in Schutte's book *)
+
+  #[using="All"]
+   Definition bigO 
+   (a: A) : Ensemble A := fun x: A => lt x a.
+
+End Definitions.
 
 Infix "o<" := ON_lt : ON_scope.
-
-Definition ON_le  {A:Type}{lt: relation A}
-           {compare : A -> A -> comparison}
-           {on : ON lt compare} :=
-  clos_refl _ ON_lt.
-
 Infix "o<=" := ON_le : ON_scope.
@@
 
-\subsection{Ordinal notations and  termination measures}
-\label{sect:measure-ON}
-
-The following lemma (together with the type class mechnism) allows us to define termination measures over any ordinal notation.
+% \subsection{Ordinal notations and  termination measures}
+% \label{sect:measure-ON}
+
+% The following lemma (together with the type class mechnism) allows us to define termination measures over any ordinal notation.
 It is just an application of  the libraries \texttt{Coq.Wellfounded.Inverse\_Image}
-and  \texttt{Coq.Wellfounded.Inclusion}.
-
-\begin{Coqsrc}
-Definition measure_lt {A:Type}{lt: relation A}
-            {compare : A -> A -> comparison}
-            {on : ON lt compare}
-            {B : Type} (m : B -> A) : relation B :=
-             fun x y => on_lt (m x) (m y).
-
-Lemma wf_measure  {A:Type}(lt: relation A)
-            {compare : A -> A -> comparison}
-            {on : ON lt compare}
-            {B : Type}
-            (m : B -> A):  well_founded (measure_lt m). 
-\end{Coqsrc}
-
-A simple example of application is given in Sect.~\vref{sect:merge-example}.
+% and  \texttt{Coq.Wellfounded.Inclusion}.
+
+% \begin{Coqsrc}
+% Definition measure_lt {A:Type}{lt: relation A}
+%             {compare : A -> A -> comparison}
+%             {on : ON lt compare}
+%             {B : Type} (m : B -> A) : relation B :=
+%              fun x y => on_lt (m x) (m y).
+
+% Lemma wf_measure  {A:Type}(lt: relation A)
+%             {compare : A -> A -> comparison}
+%             {on : ON lt compare}
+%             {B : Type}
+%             (m : B -> A):  well_founded (measure_lt m). 
+% \end{Coqsrc}
+
+% A simple example of application is given in Sect.~\vref{sect:merge-example}.
 

### 2021-07-08 doc/part-hydras.tex
update the doc on OrdinalNotations, minor changes

@@
 \begin{Coqsrc}
-Global Instance Omega : ON  Peano.lt Nat.compare.
+#[global]
+ Instance Omega_comp : Comparable Peano.lt  Nat.compare.
 Proof.
@@
   - apply Nat.lt_strorder.
+  - apply Nat.compare_spec.
+Qed.
+
+#[global]
+ Instance Omega : ON  Peano.lt  Nat.compare.
+Proof.
+ split.
+ - apply Omega_comp.
  - apply Wf_nat.lt_wf.
- - apply Nat.compare_spec.
 Qed.
@@
 
-Before building an instance of \texttt{ON}, we have to define a comparison function.
+In order to build an instance of \texttt{ON}, we have to define a comparison function.
 
@@
     (compare alpha beta).
+
+#[global] Instance plus_comp : Comparable lt compare.
+Proof.
+  split.
+  - apply lt_strorder.
+  - apply compare_correct. 
+Qed.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Global Instance ON_plus : ON lt compare.
+#[global] Instance ON_plus : ON lt compare.
 Proof.
   split.
-  - apply lt_strorder.
+  - apply plus_comp.
   - apply lt_wf.
-  - apply compare_correct.
 Qed.
@@
 \begin{Coqsrc}
-Definition Omega_plus_Omega := ON_plus Omega Omega.
-
-Existing Instance Omega_plus_Omega.
-Definition t := @ON_plus.t nat nat.
+Instance Omega_plus_Omega : ON _ _ := ON_plus Omega Omega.
+
+Definition t := ON_t.
 \end{Coqsrc}
@@
 
-Local Infix "<" := lt.
-Local Infix "<=" := (clos_refl _ lt).
+  #[local] Infix "<" := lt.
+  #[local] Infix "<=" := (leq lt).
 
@@
 
-Regarding successors, let us introduce the following definition:
+Regarding successors, let us define the following function:
 
@@
 
-Finally, the nature of any ordinal is decidable(inside this notation system) :
+Finally, the nature of any ordinal is decidable (inside this notation system) :
 

### 2021-07-09 doc/part-hydras.tex
update the doc on ON
minor corrections

@@
 \label{sect:on-le-notation}
+\label{sect:measure-ON}
 \begin{Coqsrc}
@@
 Infix "o<=" := ON_le : ON_scope.
+Infix "o?=" := ON_compare (at level 70) : ON_scope.
 \end{Coqsrc}
@@
 
-Compute ON_compare 6 9.
+#[local] Open Scope ON_scope.
+Compute 6 o?= 9.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Example ex1' : fin 7 o< omega.
-Proof. constructor. Qed.
-
+Compute  fin 8 o?= omega.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+     = Lt
+     : comparison
+\end{Coqanswer}
+
+\begin{Coqsrc}
 Lemma lt_omega alpha : 
@@
 \begin{Coqsrc}
-Definition succb (alpha: t) : bool := match alpha with
+  Definition succb (alpha: t) : bool :=
+  match alpha with
   | inr (S  _) | inl (S _) => true
@@
 Definition Zero_limit_succ_dec : ZeroLimitSucc_dec.
+Proof.
+(* ... *)
+Defined.
 \end{Coqsrc}
@@
 Definition lt : relation t := lexico ltB ltA.
-Definition le := clos_refl _ lt.
 
@@
 
-Lemma compare_reflect alpha beta :
-  match (compare alpha beta)
-  with
-    Lt => lt alpha  beta
-  | Eq => alpha = beta
-  | Gt => lt beta  alpha
-  end.
-
-Global Instance ON_mult : ON lt compare.
+#[global] Instance ON_mult: ON lt compare.
 
@@
 
-The ordinal $\omega^2$ (also called $\phi_0(2)$, see Chap.~\ref{chap:schutte}), is an instance of the multiplication presented in the last section.
+The ordinal $\omega^2$ (also called $\phi_0(2)$, see Chap.~\ref{chap:schutte}), is an instance of the multiplication presented in the preceding section.
 
@@
 \begin{Coqsrc}
-Definition Omega2 := ON_mult Omega Omega.
-Existing Instance Omega2.
+Instance Omega2: ON _ _ := ON_mult Omega Omega.
+
 Definition t := ON_mult.t nat nat.
@@
 
-Definition limitb (alpha : t) := 
+Definition limitb (alpha : t) : bool := 
  match alpha with
@@
  end.
+
+ Definition succb (alpha: t) : bool :=
+ match alpha with
+  (_, S _) => true
+  | _ => false
+end.
 \end{Coqsrc}
@@
 The successor of any ordinal is defined by a simple pattern-matching.
+
+\begin{Coqsrc}
+Definition succ (alpha : t) := (fst alpha, S (snd alpha)).
+\end{Coqsrc}
+
 This function is proved to be correct w.r.t. the \texttt{Successor} predicate.
@@
 \begin{Coqsrc}
-Definition succ (alpha : t) := (fst alpha, S (snd alpha)).
+Lemma succ_ok alpha beta : Successor beta alpha <-> beta = succ alpha.
 \end{Coqsrc}
@@
 
-Lemma succ_ok alpha beta : Successor beta alpha <-> beta = succ alpha.
 \end{Coqsrc}
@@
 Proof.
-  destruct alpha.
-  destruct x; cbn in i; discriminate.
+  destruct alpha; discriminate.
 Qed.
@@
 \begin{Coqsrc}
-
 Program Example gamma1 : t 8 := 7.
@@
 
-In order to build an instance of \texttt{OrdinalNotation}, we define a comparison function, by delegation to standard library's  \texttt{Nat.compare}, and prove its correction.
+In order to build an instance of \texttt{ON}, we define a comparison function,  and prove its correction.
 
@@
 \begin{Coqsrc}
-Global Instance sto n : StrictOrder (@lt n).
-
-Global Instance FinOrd (n:nat) : OrdinalNotation (sto n) compare.
+#[global] Instance sto n : StrictOrder (@lt n).
+
+#[global] Instance comp n: Comparable (@lt n) compare.
+
+#[global] Instance FinOrd n : ON (@lt n) compare. 
 Proof.
   split.
-  - apply compare_correct.
+  - exact (comp n).
   - apply lt_wf.
@@
 \begin{Coqsrc}
-Class ON_correct `(alpha : Schutte_basics.Ord)
+Class ON_correct `(alpha : Ord)
      `(OA : @ON A ltA compareA)
-      (iota : A -> Schutte_basics.Ord) :=
-  { ON_correct_inj : forall a, Schutte_basics.lt (iota a) alpha;
-    ON_correct_onto : forall beta, Schutte_basics.lt beta alpha ->
+      (iota : A -> Ord) :=
+  { ON_correct_inj : forall a, lt (iota a) alpha;
+    ON_correct_onto : forall beta, lt beta alpha ->
                                 exists b, iota b = beta;
@@
         match compareA a b with
-          Datatypes.Lt => Schutte_basics.lt (iota a) (iota b)
+          Datatypes.Lt => lt (iota a) (iota b)
         | Datatypes.Eq => iota a = iota b
-        | Datatypes.Gt => Schutte_basics.lt (iota b) (iota a)
+        | Datatypes.Gt => lt (iota b) (iota a)
         end

### 2021-07-11 doc/part-hydras.tex
fix a naming inconsistency

@@
 Lemma H'_omega_sqr : forall k,
-    H'_ (Phi0  2)%e0 k = (exp2 (S k ) * (S k) - 1)%nat.
+    H'_ (phi0  2)%e0 k = (exp2 (S k ) * (S k) - 1)%nat.
 Proof.
   intro k; rewrite H'_eq3; auto with E0.
-  - ochange (Canon (Phi0 2) (S k)) (omega * (S k))%e0.
+  - ochange (Canon (phi0 2) (S k)) (omega * (S k))%e0.
     +  rewrite H'_omega_i; simpl (exp2 (S k)).
@@
     H'_ (Omega_term alpha (S i)) k =
-    H'_ (Omega_term alpha i) (H'_ (Phi0 alpha) k).
+    H'_ (Omega_term alpha i) (H'_ (phi0 alpha) k).
 \end{Coqsrc}
@@
 alpha <> Zero -> forall i k, 
-  H'_ (Omega_term alpha i) k = iterate (H'_ (Phi0 alpha)) (S i) k.
+  H'_ (Omega_term alpha i) k = iterate (H'_ (phi0 alpha)) (S i) k.
 \end{Coqsrc}
@@
 
-Lemma H'_Phi0_succ alpha  : alpha <> Zero -> forall k,
- H'_ (Phi0 (Succ alpha)) k = H'_succ_fun (H'_ (Phi0 alpha)) k.
-
-Lemma H'_Phi0_Si : forall i k,
- H'_ (Phi0 (S i)) k = iterate H'_succ_fun i (H'_ omega) k. 
+Lemma H'_phi0_succ alpha  : alpha <> Zero -> forall k,
+ H'_ (phi0 (Succ alpha)) k = H'_succ_fun (H'_ (phi0 alpha)) k.
+
+Lemma H'_phi0_Si : forall i k,
+ H'_ (phi0 (S i)) k = iterate H'_succ_fun i (H'_ omega) k. 
 \end{Coqsrc}
@@
 Lemma H'_omega_cube : forall k,
-    H'_ (Phi0 3)%e0 k = iterate (H'_ (Phi0 2))  (S k) k.
+    H'_ (phi0 3)%e0 k = iterate (H'_ (phi0 2))  (S k) k.
 Proof.
-  intro k; rewrite <-FinS_eq, -> Fin_Succ, H'_Phi0_succ; auto.
+  intro k; rewrite <-FinS_eq, -> Fin_Succ, H'_phi0_succ; auto.
   compute; injection 1; discriminate.
@@
 
-Remark R0 k :  H'_ (Phi0 3)%e0 k = iterate f (S k) k.
+Remark R0 k :  H'_ (phi0 3)%e0 k = iterate f (S k) k.
 \end{Coqsrc}
@@
 \begin{Coqsrc}
-Fact F0 : H'_ (Phi0 3) 3 = f (f (f (f 3))).
+Fact F0 : H'_ (phi0 3) 3 = f (f (f (f 3))).
 proof. rewrite R0; reflexivity. Qed.
@@
 \begin{Coqsrc}
-Fact F1 : H'_ (Phi0 3) 3 = f (f N).
+Fact F1 : H'_ (phi0 3) 3 = f (f N).
 Proof.
@@
 
-Lemma F1_simpl : H'_ (Phi0 3) 3 =
+Lemma F1_simpl : H'_ (phi0 3) 3 =
                  (exp2 (exp2 (S N) * S N) * (exp2 (S N) * S N) - 1)%nat.
@@
 \begin{Coqsrc}
-Lemma H'_Phi0_omega : forall k, H'_ (Phi0 omega) k =
+Lemma H'_phi0_omega : forall k, H'_ (phi0 omega) k =
                                iterate H'_succ_fun  k (H'_ omega) k.
 Proof with auto with E0.
-  intro k; rewrite H'_eq3, <- H'_Phi0_Si ...
-  -  rewrite CanonS_Canon, CanonS_Phi0_lim;  f_equal ...
+  intro k; rewrite H'_eq3, <- H'_phi0_Si ...
+  -  rewrite CanonS_Canon, CanonS_phi0_lim;  f_equal ...
 Qed.
@@
 \begin{Coqsrc}
-Lemma H'_Phi0_omega_closed_formula k :
-  H'_ (Phi0 omega) k =
+Lemma H'_phi0_omega_closed_formula k :
+  H'_ (phi0 omega) k =
   iterate (fun (f: nat -> nat) (l : nat) => iterate  f (S l) l)

### 2021-07-12 doc/part-hydras.tex
Remove a buggy subdirectory
The implementation of multiplication in ordinals/OmegaOmega is buggy.
We save the files in branch FixOmegaOmega and temporarily remove them from master.

@@
 \section{Other ordinal notations}
-
-\index{hydras}{Projects}
-
-\begin{project}
-The directory \texttt{theories/OmegaOmega} contains an ad-hoc formalization of $\omega^\omega$, contributed by Pascal Manoury.
+%%% TODO : Fix the multiplication function in branch FixOmegaOmega
+
+% \index{hydras}{Projects}
+
+% \begin{project}
+% The directory \texttt{theories/OmegaOmega} contains an ad-hoc formalization of $\omega^\omega$, contributed by Pascal Manoury.
 Every ordinal $\alpha$ is represented by a list $l$ whose elements are the coefficients of $\omega$ in  the Cantor normal form of $\alpha$ (in reverse order).
 For instance, the ordinal
-$\omega^{8}\times 5 + \omega^{6}\times 8 + \omega^2\times 10 + \omega + 7$ is represented by the list \texttt{[5;0;8;0;0.0;10,1,7]}.
-
- Develop this representation and compare it with the other ordinal notations.
-
-\end{project}
+% $\omega^{8}\times 5 + \omega^{6}\times 8 + \omega^2\times 10 + \omega + 7$ is represented by the list \texttt{[5;0;8;0;0.0;10,1,7]}.
+
+%  Develop this representation and compare it with the other ordinal notations.
+
+% \end{project}
 

### 2021-07-15 doc/part-hydras.tex
more proofs by refinement

@@
     \end{project}
+
+    \section{A refinement of \texttt{E0} : an ordinal notation for \texorpdfstring{$\omega^\omega$}{omega\^omega}}
+
+    In Module   \href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/OrdinalNotations/OmegaOmega.v}{theories/ordinals/OrdinalNotations/OmegaOmega.v},
+we represents ordinals below $\omega^\omega$ by list of pairs of natural numbers.
+we establish this representation as a \emph{refinement} of the data types we used to represent ordinals less than $\epsilon_0$.
+Thus, many properties like well-foundedness of $<$ and associativity of $+$,  of this ordinal notations have very short proofs.
+
+%ici
     \section{A variant for hydra battles}

### 2021-07-16 doc/part-hydras.tex
minor changes

--- 2021-07-12 doc/part-hydras.tex
@@
 
-    \index{hydras}{Projects}
-    \begin{project}
-      Prove that ordinal multiplication is left-distributive over ordinal addition; that is:
-      $$\forall \alpha \, \beta \, \gamma < \epsilon_0, \alpha\times(\beta+\gamma) = \alpha\times\beta+ \alpha\times \gamma$$
-
-    \end{project}
+    \section{A refinement of \texttt{E0} : an ordinal notation for \texorpdfstring{$\omega^\omega$}{omega\^omega}}
+
+    In Module   \href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/OrdinalNotations/OmegaOmega.v}{theories/ordinals/OrdinalNotations/OmegaOmega.v},
+    we represent ordinals below $\omega^\omega$ by list of pairs of natural numbers (with the same coefficient shift as in \texttt{T1}).
+    For instance, the ordinal $\omega^4\times 10 + \omega^3 + \omega+ 5$ is represented by the list \texttt{(4,9)::(3,0)::(1,0)::(0,4)::nil}.
+
+  The usual operations : \texttt{succ}, \texttt{+}, \texttt{*} are simple variants of the same operations in \texttt{T1}.
+
+    We establish this representation as a \emph{refinement} of the data types we used to represent ordinals less than $\epsilon_0$.
+Thus, many properties like well-foundedness of $<$ and associativity of $+$,  of this ordinal notations have very short proofs.
+
+%ici
     \section{A variant for hydra battles}

### 2021-07-19 doc/movies/extract_snippets.py
Use LaTeX snippets extracted with Alectryon.
- add cli to convert coq file to latex
- cut coq file with snippet
- first approch to make latex
- add paramater backend and frontend
- parse coq+rst and extract snippet to latex
- add Readme
- Make a version snippet extractor with latex conversion
- Use latex format to extract snippets
- move cutting table in move and create edit makefile
- try to make snipet in Ackerman to replace in latex file

@@ -1,0 +1,6 @@
+#!/usr/bin/env python3
+from extract_snippets.cli import main
+
+
+if __name__ == '__main__':
+    main()

### 2021-07-19 doc/movies/extract_snippets/snippetExceptions.py

@@ -1,0 +1,47 @@
+from typing import Iterable
+
+
+
+class SnippetException(Exception):
+    def __init__(self, prelude: str, line_number: int = None):
+        if line_number is None:
+            super().__init__(prelude)
+        else:
+            super().__init__(f"{prelude}, at line {line_number}.")
+
+
+class SnippetAlreadyExistException(SnippetException):
+    """
+    Raise in snippet if 2 snippet with same name in file.
+    """
+
+    def __init__(self, name_snippet: str, line_number: int):
+        super().__init__(f"Snippet \"{name_snippet}\" already exist", line_number)
+
+
+class SnippetAlreadyCloseException(SnippetException):
+    """
+    Raise if snippet is already close.
+    """
+
+    def __init__(self, name_snippet: str, line_number: int):
+        super().__init__(f"Snippet \"{name_snippet}\" already close", line_number)
+
+
+class SnippetNotBeginException(SnippetException):
+    """
+    Raise if snippet not exist.
+    """
+
+    def __init__(self, name_snippet: str, line_number: int):
+        super().__init__(f"Snippet \"{name_snippet}\" not begin (cannot be closed)", line_number)
+
+
+class SnippetNotEndWarning(Warning):
+    """
+    Raise if snippet is have not end.
+    """
+
+    def __init__(self, *snippets: Iterable['Snippet']):
+        names = ",".join(snippet.name for snippet in snippets)
+        super().__init__(f"Snippets \"{names}\" never close")

### 2021-07-19 doc/movies/extract_snippets/Extractor.py
Do not use Python 3.8 features.

@@ -38,11 +38,14 @@ class SnippetExtractor:
         statement, _, tag, after = match.groups()
         if statement == BEGIN_STATE:
             self.stack_tag.append(tag)
-            if (match := LATEX_TAG_PATTERN.match(after)) is not None:
+            match = LATEX_TAG_PATTERN.match(after)
+            if match is not None:
                 self._match_tag(match, line_num)
 
-        elif (last_tag := self.stack_tag.pop()) != tag:
-            raise Exception(f"tag '{tag}' close before '{last_tag}' at line {line_num}.")
+        else:
+            last_tag = self.stack_tag.pop()
+            if last_tag != tag:
+                raise Exception(f"tag '{tag}' close before '{last_tag}' at line {line_num}.")
 
     def _open_snippet(self, snippet_name: str, line_number: int) -> None:
         """
@@ -110,12 +113,15 @@ class SnippetExtractor:
         """
 
         for num, line in self.reader:
-            if (match := LATEX_TAG_PATTERN.match(line)) is not None:
+            match = LATEX_TAG_PATTERN.match(line)
+            if match is not None:
                 self._match_tag(match, line_num=num)
 
-            elif (match := SNIPPET_PATTERN.match(line)) is not None:
-                self._match_snippet(match, line_num=num)
-                continue
+            else:
+                match = SNIPPET_PATTERN.match(line)
+                if match is not None:
+                    self._match_snippet(match, line_num=num)
+                    continue
 
             self._add_snippets_open(line)
 

### 2021-07-19 doc/movies/extract_snippets/cli.py
Copy sty alectryon file and more
- copy alectryon latex assets
- Add rule to clear latex generation with make clean
- Add requirements.txt file and change section in Readme.md
- remove compilation of "Issue_Program.v" in latex
- remove pre-compile to get alectryon assets in CI

@@ -106,19 +106,36 @@ def make_latex_file(content: str, path: Path):
         file.write(content)
 
 
-def main():
+def make_latex(input_file: Path, sertop_args):
     from .Extractor import SnippetExtractor
     from .docutils import CoqToLatexReader, register_docutils
+    print(f"Convert '{input_file}' to latex.")
+    register_docutils(sertop_args)
+    reader = CoqToLatexReader(input_file)
+    print("Extract snippets.")
+    return SnippetExtractor(reader), reader
+
+
+def copy_asset(output_dir: Path, dir_name='assets'):
+    from alectryon.cli import copy_assets
+    from alectryon.latex import ASSETS
+    from shutil import copy
+
+    STY = ASSETS.ALECTRYON_STY + ASSETS.PYGMENTS_STY
+
+    output_dir = output_dir / dir_name
+    output_dir.mkdir(exist_ok=True)
+
+    copy_assets(0, STY, copy, output_dir)
+    print(f"copy assets {STY} in {output_dir}")
+
+
+def main():
     args = parse_args()
     input_name = args.input.stem
     output_dir = args.output_dir / input_name
 
-    # make latex
-    print(f"Convert '{args.input}' to latex.")
-    register_docutils(args.sertop_args)
-    reader = CoqToLatexReader(args.input)
-    print("Extract snippets.")
-    snippet_extractor = SnippetExtractor(reader)
+    snippet_extractor, reader = make_latex(args.input, args.sertop_args)
 
     # dump latex to debug
     if args.dump_complete_latex:
@@ -128,7 +145,11 @@ def main():
         make_latex_file(reader.content_latex, path_latex)
 
     # write snippets files
-    for snippet in snippet_extractor.extract():
+    snippets = snippet_extractor.extract()
+    for snippet in snippets:
         path = get_path_latex(output_dir, snippet.name)
         print(f"extract snippet '{snippet.name}', dump file {path}.")
         make_latex_file(str(snippet), path)
+
+    if snippets:
+        copy_asset(args.output_dir)

### 2021-07-20 doc/chapter-primrec.tex

@@
 
-\begin{Coqsrc}
-Fail
-  Fixpoint Ack (m n : nat) : nat :=
-  match m, n with
-  | 0, n => S n
-  | S m, 0 => Ack m 1
-  | S m0, S p => Ack m0 (Ack m p)
-  end.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-The command has indeed failed with message:
-Cannot guess decreasing argument of fix.
-\end{Coqanswer}
+\input{movies/snippets/Ack/AckFixpointFail.tex}
 
@@
 
-\begin{Coqsrc}
-Module Alt.
-
-   Fixpoint Ack (m n : nat) : nat :=
-     match m with
-     | O => S n
-     | S p => let fix Ackm (n : nat) :=
-                  match n with
-                  | O => Ack p 1
-                  | S q => Ack p (Ackm q)
-                  end
-              in Ackm n
-     end.
-
-    Compute Ack 3 2.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = 29 : nat
-\end{Coqanswer}
-
-\begin{Coqsrc}
-End Alt.
-\end{Coqsrc}
+\input{movies/snippets/Ack/AckFixpointAlt.tex}
 
@@
 
-\begin{Coqsrc}
-Fixpoint Ack (m:nat) : nat -> nat :=
-  match m with
-  | 0 => S
-  | S n => fun k =>  iterate (Ack n) (S k) 1
-  end.
-
-Compute Ack 3 2.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = 29 : nat
-\end{Coqanswer}
+\input{movies/snippets/Ack/AckFixpointIterate.tex}
 

### 2021-07-20 doc/chapter-primrec.tex
Insert some new snippets extracted from the code.
more snippets

@@
 \index{primrec}{Types!naryFunc}
-\begin{Coqsrc}
-Fixpoint naryFunc (n : nat) : Set :=
-  match n with
-  | O => nat
-  | S n => nat -> naryFunc n
-  end.
-\end{Coqsrc}
+\input{movies/snippets/extEqualNat/naryFunc.tex}
 
@@
 \emph{From \href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}}.
-\begin{Coqsrc}
-Require Import primRec.
-Import extEqualNat.
-
-Compute naryFunc 3.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-= nat -> nat -> nat -> nat
-  : Set  
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Check plus: naryFunc 2.
-
-Check 42: naryFunc 0.
-
-Check (fun n p q : nat =>  n * p + q): naryFunc 3.
-\end{Coqsrc}
+
+\input{movies/snippets/PrimRecExamples/naryFunc3}
+\input{movies/snippets/PrimRecExamples/checknaryFunc}
 
@@
 
-\begin{Coqsrc}
-Compute naryRel 2.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = nat -> nat -> bool
-     : Set
-\end{Coqanswer}
+\input{movies/snippets/PrimRecExamples/naryRel2}
 
@@
 
-\begin{Coqsrc}
-Compute extEqual 2.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-     = fun a b : naryFunc 2 => forall x x0 : nat, a x x0 = b x x0
-     : naryFunc 2 -> naryFunc 2 -> Prop
- \end{Coqanswer}
-
-\begin{Coqsrc}
-Example extEqual_ex1 : extEqual 2 mult (fun x y =>  y * x + x - x) .
-Proof.
-  intros x y.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  x, y : nat
-  ============================
-  extEqual 0 (x * y) (y * x)
-\end{Coqanswer}
-
-\begin{Coqsrc}
-  cbn.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-1 subgoal (ID 10)
-
-  x, y : nat
-  ============================
-  x * y = y * x + x - x
-\end{Coqanswer}
-
-\begin{Coqsrc}
-  rewrite <- Nat.add_sub_assoc, Nat.sub_diag.
-  - ring.
-  - apply le_n.  
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/PrimRecExamples/extEqual2a}
+
+  Getting rid of the term \texttt{x-x}, we generate two easy-to-solve sub-goals.
+
+\vspace{6pt}
+\noindent
+\input{movies/snippets/PrimRecExamples/extEqual2b}
 
@@
 
-\begin{Coqsrc}
-Search (is_PR 2 (fun _ _ => nat_rec _ _ _ _)).
-\end{Coqsrc}
-
-\begin{Coqanswer}
-ind1ParamIsPR:
-  forall f : nat -> nat -> nat -> nat,
-  isPR 3 f ->
-  forall g : nat -> nat,
-  isPR 1 g ->
-  isPR 2
-    (fun a b : nat =>
-     nat_rec (fun _ : nat => nat)
-                 (g b) (fun x y : nat => f x y b) a)
-\end{Coqanswer}
+\input{movies/snippets/PrimRecExamples/PrimRecExamplesSearch}
 

### 2021-07-21 doc/chapter-primrec.tex
more snippets

@@
 \emph{From \href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}.}
-\begin{Coqsrc}
-Example Ex1 : evalPrimRec 0 zeroFunc = 0.
-Proof. reflexivity. Qed.
-
-Example Ex2 a : evalPrimRec 1 succFunc a = S a.
-Proof. reflexivity. Qed.
-
-Example Ex3 a b c d e f: forall (H: 2 < 6),
-    evalPrimRec 6
-                (projFunc 6 2 H) a b c d e f = d.
-Proof. reflexivity. Qed.
-
-Example Ex4 (x y z : PrimRec 2) (t: PrimRec 3):
-  let u := composeFunc 2 3
-                       (PRcons 2 _ x
-                               (PRcons 2 _ y
-                                       (PRcons 2 _ z
-                                               (PRnil 2))))
-                       t in
-  let f := evalPrimRec 2 x in
-  let g := evalPrimRec 2 y in
-  let h := evalPrimRec 2 z in
-  let i := evalPrimRec 3 t in
-  let j := evalPrimRec 2 u in
-  forall a b, j a b = i (f a b) (g a b) (h a b).
-Proof. reflexivity. Qed.
-
-Example Ex5 (x : PrimRec 2)(y: PrimRec 4):
-  let g := evalPrimRec _ x in
-  let h := evalPrimRec _ y in
-  let f := evalPrimRec _ (primRecFunc _ x y) in
-  forall a b,  f 0 a b = g a b.
-Proof. reflexivity.   Qed.
-
-Example Ex6 (x : PrimRec 2)(y: PrimRec 4):
-  let g := evalPrimRec _ x in
-  let h := evalPrimRec _ y in
-  let f := evalPrimRec _ (primRecFunc _ x y) in
-  forall n a b,  f (S n) a b = h n (f n a b) a b.
-Proof. reflexivity.   Qed.                          
-\end{Coqsrc}
+
+\input{movies/snippets/PrimRecExamples/evalPrimRecEx}
+
+\vspace{4pt}
+\noindent
 
@@
 \label{sect:bigfac}
-\begin{Coqsrc}
-Example bigPR : PrimRec 1 :=
-primRecFunc 0
-  (composeFunc 0 1 (PRcons 0 0 zeroFunc (PRnil 0)) succFunc)
-  (composeFunc 2 2
-    (PRcons 2 1
-      (composeFunc 2 1
-         (PRcons 2 0 (projFunc 2 1 (le_n 2))
-                 (PRnil 2))
-         succFunc)
-      (PRcons 2 0
-        (composeFunc 2 1
-          (PRcons 2 0
-             (projFunc 2 0
-                       (le_S 1 1 (le_n 1)))
-             (PRnil 2))
-          (projFunc 1 0 (le_n 1))) (PRnil 2)))
-    (primRecFunc 1 (composeFunc 1 0 (PRnil 1) zeroFunc)
-       (composeFunc 3 2
-         (PRcons 3 1
-            (projFunc 3 1 (le_S 2 2 (le_n 2)))
-            (PRcons 3 0 (projFunc 3 0
-                          (le_S 1 2
-                                (le_S 1 1 (le_n 1))))
-                    (PRnil 3)))
-         (primRecFunc 1 (projFunc 1 0 (le_n 1))
-                      (composeFunc 3 1
-                          (PRcons 3 0
-                                  (projFunc 3 1 (le_S 2 2 (le_n 2)))
-                                  (PRnil 3))
-                          succFunc))))). 
-\end{Coqsrc}
+
+\input{movies/snippets/PrimRecExamples/bigPRa}
 
@@
 
-\begin{Coqsrc}
-Example  mystery_fun : nat -> nat := evalPrimRec 1 bigPR.
-
-Compute map mystery_fun[0;1;2;3;4;5;6] : t nat _.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = 1 :: 1 :: 2 :: 6 :: 24 :: 120 :: 720 :: nil
-     : t nat 7
-\end{Coqanswer}
+\vspace{4pt}
+\noindent
+
+\input{movies/snippets/PrimRecExamples/bigPRb}
 

### 2021-07-22 doc/chapter-primrec.tex
add a few snippets to chapter 9

@@
 
-\begin{Coqsrc}
-Fixpoint  extEqual (n : nat) : forall  (a b : naryFunc n), Prop :=
-  match n with
-    0 => fun a b => a = b
-  | S p => fun a b => forall c, extEqual p (a c) (b c)
-  end.
-\end{Coqsrc}
+\input{movies/snippets/extEqualNat/extEqualDef}
 
@@
 \emph{From \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.}
-\begin{Coqsrc}
-Inductive PrimRec : nat -> Set :=
-  | succFunc : PrimRec 1
-  | zeroFunc : PrimRec 0
-  | projFunc : forall n m : nat, m < n -> PrimRec n
-  | composeFunc :
-      forall (n m : nat) (g : PrimRecs n m) (h : PrimRec m), PrimRec n
-  | primRecFunc :
-      forall (n : nat) (g : PrimRec n) (h : PrimRec (S (S n))), 
-      PrimRec (S n)
-with PrimRecs : nat -> nat -> Set :=
-  | PRnil : forall n : nat, PrimRecs n 0
-  | PRcons : forall n m : nat, PrimRec n -> PrimRecs n m -> 
-                   PrimRecs n (S m).
-\end{Coqsrc}
+
+\input{movies/snippets/primRec/PrimRecDef}
 
@@
 
-\begin{Coqsrc}
-Fixpoint evalPrimRec (n : nat) (f : PrimRec n) {struct f} : 
- naryFunc n :=
-  match f in (PrimRec n) return (naryFunc n) with
-  | succFunc => S
-  | zeroFunc => 0
-  | projFunc n m pf => evalProjFunc n m pf
-  | composeFunc n m l f =>
-      evalComposeFunc n m (evalPrimRecs _ _ l) (evalPrimRec _ f)
-  | primRecFunc n g h =>
-      evalPrimRecFunc n (evalPrimRec _ g) (evalPrimRec _ h)
-  end
-with evalPrimRecs (n m : nat) (fs : PrimRecs n m) {struct fs} :
- Vector.t (naryFunc n) m :=
-  match fs in (PrimRecs n m) return (Vector.t (naryFunc n) m) with
-  | PRnil a => Vector.nil  (naryFunc a)
-  | PRcons a b g gs =>
-       Vector.cons _ (evalPrimRec _ g) _  (evalPrimRecs _ _ gs)
-  end.
-\end{Coqsrc}
+\vspace{4pt}
+
+\input{movies/snippets/primRec/evalPrimRecDef}
+
+\vspace{4pt}
 
@@
 \emph{From \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.}
-\begin{Coqsrc}
-Definition isPR (n : nat) (f : naryFunc n) : Set :=
-  {p : PrimRec n | extEqual n (evalPrimRec _ p) f}.  
-\end{Coqsrc}
+
+\input{movies/snippets/primRec/isPRDef}
 
@@
 
-\begin{Coqsrc}
-Module Alt.
-
-Lemma zeroIsPR : isPR 0 0.
-Proof.
-  exists zeroFunc.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-1 subgoal (ID 90)
-
-  ============================
-  extEqual 0 (evalPrimRec 0 zeroFunc) 0
-\end{Coqanswer}
-
-\begin{Coqsrc}
-  cbn.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-1 subgoal (ID 91)
-
-  ============================
-  0 = 0
-\end{Coqanswer}
-
-\begin{Coqsrc}
-  reflexivity.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/PrimRecExamples/zeroIsPR}
 
@@
 
-\begin{Coqsrc}
-Lemma SuccIsPR : isPR 1 S.
-Proof.
-  exists succFunc; cbn; reflexivity.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/PrimRecExamples/SuccIsPR}
 
@@
 
-\begin{Coqsrc}
-Lemma pi2_5IsPR : isPR 5 (fun a b c d e => b).
-Proof.
- assert (H: 3 < 5) by auto.
- exists (projFunc 5 3 H).
- cbn; reflexivity.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/PrimRecExamples/pi25IsPR}
 
@@
 
-\begin{Coqsrc}
-Lemma idIsPR : isPR 1 (fun x : nat => x).
-Proof.
-  assert (H: 0 < 1) by auto.
-  exists (projFunc 1 0 H); cbn; auto.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/primRec/idIsPR}
 

### 2021-07-23 doc/chapter-primrec.tex
more snippets

@@
 
-\begin{Coqsrc}
-Fact compose_01 :
-    forall (x:PrimRec 0) (t : PrimRec 1),
-    let c := evalPrimRec 0 x in
-    let f := evalPrimRec 1 t in
-    evalPrimRec 0 (composeFunc 0 1
-                               (PRcons 0 0 x (PRnil 0))
-                               t)  =
-     f c.
-Proof. reflexivity. Qed.
-\end{Coqsrc}
-
+\vspace{4pt}
+\input{movies/snippets/PrimRecExamples/compose01}
+
+\vspace{4pt}
 Thus, we get a quite simple proof of \texttt{const1\_NIsPR}.
@@
 \emph{From \href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}}.
-\begin{Coqsrc}
-Lemma  const1_NIsPR n : isPR 0 n. 
-Proof.
-  induction n.
-  - apply zeroIsPR.
-  - destruct IHn as [x Hx].
-   exists (composeFunc 0 1 (PRcons 0 0 x (PRnil 0)) succFunc). 
-   cbn in *; intros; now rewrite Hx.
-Qed.
-\end{Coqsrc}
+
+\input{movies/snippets/PrimRecExamples/const0NIsPR}
 
@@
 
-We prove that the library function \texttt{plus} is extensionally equal to a function defined with
+The following lemma shows it suffices to prove that
+Standard library's function \texttt{plus} is extensionally equal to a function defined with
 \texttt{nat\_rec}.
 
-\begin{Coqsrc}
-Definition plus_alt x y  :=
-              nat_rec  (fun n : nat => nat)
-                       y
-                       (fun z t =>  S t)
-                       x.
-
-Lemma plus_alt_ok:
-  extEqual 2 plus_alt plus.
-Proof.
-  intro x; induction x; cbn; auto.
-  intros y; cbn; now rewrite <- (IHx y).
-Qed.
-\end{Coqsrc}
-
-A last lemma before the proof:
-
-\begin{Coqsrc}
-Lemma isPR_extEqual_trans n : forall f g, isPR n f ->
-                                    extEqual n f g ->
-                                    isPR n g.
-Proof.
- intros f g [x Hx]; exists x.
- apply extEqualTrans with f; auto.
-Qed.
-\end{Coqsrc}
-
-Let us start now.
+\input{movies/snippets/PrimRecExamples/isPRExtEqualTrans}
+
+Thus, let us define an helper and prove its equivalence with \texttt{plus}.
+
+\input{movies/snippets/PrimRecExamples/plusAlt}
+
+\vspace{4pt}
+
+We are now able to complete the proof.
+
+\input{movies/snippets/PrimRecExamples/plusIsPRa}
 
@@
 
-\begin{Coqanswer}
-filter010IsPR :
-forall g : nat -> nat, isPR 1 g -> isPR 3 (fun _ b _ : nat => g b)
-\end{Coqanswer}
+\vspace{4pt}
+\input{movies/snippets/PrimRecExamples/checkFilter0101IsPR}
+\vspace{4pt}
 
@@
 
-\begin{Coqsrc}
- - unfold plus_alt; apply ind1ParamIsPR.
-    + apply filter010IsPR, succIsPR.
-    + apply idIsPR.
-  - apply plus_alt_ok. 
-Qed.
-\end{Coqsrc}
+\vspace{4pt}
+
+\input{movies/snippets/PrimRecExamples/plusIsPRb}
 
@@
 \emph{Note that the lemma \texttt{const1\_NIsPR} considers this function as an unary function (unlike \texttt{const0\_NIsPR})}. 
-
-\begin{Coqsrc}
-Lemma doubleIsPR : isPR 1 double.
-Proof.
-  unfold double; apply compose1_2IsPR.
-  - apply idIsPR.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-subgoal 1 (ID 110) is:
- isPR 1 (fun _ : nat => 2)
-subgoal 2 (ID 111) is:
- isPR 2 Init.Nat.mul
-\end{Coqanswer}
-
- \begin{Coqsrc}
-  - apply const1_NIsPR.
-  - apply multIsPR.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/PrimRecExamples/doubleIsPRa}
+\input{movies/snippets/PrimRecExamples/doubleIsPRb}
 

### 2021-07-24 doc/chapter-primrec.tex
more snippets

@@
 
-\begin{Coqsrc}
-Lemma plusIsPR : isPR 2 plus.
-Proof.
-  apply isPR_extEqual_trans with plus_alt.
-  - unfold plus_alt; apply ind1ParamIsPR.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-2 subgoals (ID 126)
-
-  ============================
-  isPR 3 (fun _ y _ : nat => S y)
-
-subgoal 2 (ID 127) is:
- isPR 1 (fun b : nat => b)
-\end{Coqanswer}
-
 We already proved that \texttt{S} is \texttt{PR 1}, but we need to consider a function of three arguments, which ignores its first and third arguments.
@@
 
+%\input{movies/snippets/PrimRecExamples/justATest}
+
 \subsubsection{More examples}
@@
 \input{movies/snippets/PrimRecExamples/doubleIsPRa}
-\input{movies/snippets/PrimRecExamples/doubleIsPRb}
 
@@
 
-\begin{Coqsrc}
-Fixpoint fact n :=
-  match n with 
-          | 0 => 1
-          | S p  => n * fact p
-  end.
-
-Fixpoint exp n p :=
-  match p with
-  | 0 => 1
-  | S m =>  exp n m * n
-  end.
-
-Fixpoint tower2 n :=
-  match n with
-  | 0 => 1
-  | S p => exp 2 (tower2 p)
-  end.
-\end{Coqsrc}
+\input{movies/snippets/MorePRExamples/factDef}
+
+\input{movies/snippets/MorePRExamples/expDef}
+
+\input{movies/snippets/MorePRExamples/tower2Def}
 
@@
 
-\begin{Coqsrc}
-Fixpoint fib (n:nat) : nat :=
-  match n with
-  | 0 => 1
-  | 1 => 1
-  | S ((S p) as q) => fib q + fib p
-  end.
-\end{Coqsrc}
+\input{movies/snippets/FibonacciPR/fibDef}
 
@@
 \begin{exercise}
+
 Prove the following lemmas (which may help to solve the next  exercise).
 
-\begin{Coqsrc}
-Lemma boundedSearch3 (P: naryRel 2) (b  : nat), 
-    boundedSearch P b <= b.
-
-Lemma boundedSearch4 (P: naryRel 2) (b  : nat):
-  P b b = true -> P b (boundedSearch P b) = true.
-\end{Coqsrc}
+\input{movies/snippets/isqrt/boundedSearch3}
+\input{movies/snippets/isqrt/boundedSearch4}
+
+\textbf{Note:}  The function \texttt{boundedSearch} is defined
+in Library \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}. 
 \end{exercise}

### 2021-07-26 doc/chapter-primrec.tex
more snippets

@@
 \index{hydras}{Library Prelude!iterate}
-\begin{Coqsrc}
-Fixpoint iterate {A:Type}(f : A -> A) (n: nat)(x:A) :=
-  match n with
-  | 0 => x
-  | S p => f (iterate  f p x)
-  end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma iterate_le_n_Sn:
-forall f : nat -> nat,
-(forall x : nat, x <= f x) ->
-forall n x : nat, iterate f n x <= iterate f (S n) x.
-\end{Coqsrc}
+
+\input{movies/snippets/Iterates/iterateDef}
+
+\input{movies/snippets/Iterates/iterateLeNSN}
 
@@
 \begin{exercise}
-The file \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack} presents two other definitions\footnote{post on \href{https://stackoverflow.com/questions/10292421/error-in-defining-ackermann-in-coq} by Anton Trunov.} of the Ackermann functions based on the lexicographic ordering on $\mathbb{N}\times\mathbb{N}$.
+The file \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack} presents two other definitions of the Ackermann functions based on the lexicographic ordering on $\mathbb{N}\times\mathbb{N}$.
 Prove that the four functions are extensionnally equal.
@@
 
-\begin{Coqsrc}
-Lemma Ack_0 : Ack 0 = S.
-Proof refl_equal.
-
-Lemma Ack_S_0 m : Ack (S m) 0 = Ack m 1. 
-Proof. reflexivity. Qed.
-
-Lemma Ack_S_S : forall m p,
-    Ack (S m) (S p) = Ack m (Ack (S m) p).
-Proof. reflexivity. Qed.
-\end{Coqsrc}
+\input{movies/snippets/Ack/AckRewrite}
+
+\vspace{4pt}
 
@@
 
-\begin{Coqsrc}
-Lemma Ack_1_n n: Ack 1 n = S  (S n).
-
-Lemma Ack_2_n n: Ack 2 n = 2 * n + 3.
-
-Lemma Ack_3_n n: Ack 3 n = exp2 (S (S (S n))) - 3.
-
-Lemma Ack_4_n n: Ack 4 n = hyper_exp2 (S (S (S n))) - 3.
-\end{Coqsrc}
+\input{movies/snippets/Ack/Ack1N}
+\input{movies/snippets/Ack/Ack2N}
+\input{movies/snippets/Ack/Ack3N}
+\input{movies/snippets/Ack/Ack4N}
 
@@
 \emph{From \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack}}.
-\begin{Coqsrc}
-Lemma nested_Ack_bound : forall k m n, 
-    Ack k (Ack m n) <= Ack (2 + max k m) n.
-\end{Coqsrc}
+
+\input{movies/snippets/Ack/nestedAckBound}
 
@@
 \noindent
+
 \emph{From \href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR}}.
-\begin{Coqsrc}
-Theorem Ackn_IsPR (n: nat) : isPR 1 (Ack n).
-\end{Coqsrc}
+
+\input{movies/snippets/AckNotPR/AckNIsPR}
 
@@
 \index{coq}{Commands!Scheme}
-\begin{Coqsrc}
-Scheme PrimRec_PrimRecs_ind := Induction for PrimRec
-  Sort Prop
-  with PrimRecs_PrimRec_ind := Induction for PrimRecs 
-  Sort Prop.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-PrimRec_PrimRecs_ind :
-forall (P : forall n : nat, PrimRec n -> Prop)
-  (P0 : forall n n0 : nat, PrimRecs n n0 -> Prop),
-(* successor *)
-P 1 succFunc ->
-
-(* zero *)
-P 0 zeroFunc ->
-
-(* projections *)
-(forall (n m : nat) (l : m < n), P n (projFunc n m l)) ->
-
-(* composition *) 
-(forall (n m : nat) (g : PrimRecs n m),
-      P0 n m g -> forall h : PrimRec m, P m h -> 
-      P n (composeFunc n m g h)) ->
-
-(* primitive recursion *)
-(forall (n : nat) (g : PrimRec n),
- P n g ->
-    forall h : PrimRec (S (S n)), P (S (S n)) h -> 
-     P (S n) (primRecFunc n g h)) ->
-
-(* empty list of functions *)
-(forall n : nat, P0 n 0 (PRnil n)) ->
-
-(* add a function to a list *)
-(forall (n m : nat) (p : PrimRec n),
-   P n p -> 
-   forall p0 : PrimRecs n m, P0 n m p0 -> 
-   P0 n (S m) (PRcons n m p p0)) ->
-
-(* conclusion ! *)
-forall (n : nat) (p : PrimRec n), P n p
-\end{Coqanswer}
-
-For instance, proving a property shared by any primitive recursive function of arity 2 leads to consider the case where that function is obtained by composition with a function of any 
-arity $m$. The same problem happens with primitive 
-recursion, where a function of arity $n$ is built out of a function of arity $n+1$ and a function of arity $n-1$.
+
+\input{movies/snippets/primRec/SchemePrimRecInd}
+
+Please note that, in order to prove a property shared by any primitive recursive function of, say, arity 2, this induction scheme  leads you to consider an expension of the considered property to primitive recursive function of any arity.
 
@@
 \index{coq}{Dependently typed functions}
-\begin{Coqsrc}
-Fixpoint max_v {n:nat} : forall (v: Vector.t nat n) , nat :=
-  match n as n0 return (Vector.t nat n0 -> nat)
-  with
-    0 => fun v => 0
-  | S p => fun (v : Vector.t nat (S p)) =>
-             max (Vector.hd v) (max_v  (Vector.tl v))
-  end.
-
-Lemma max_v_2 : forall x y,  max_v (x::y::nil) = max x y.
-
-Lemma max_v_lub : forall n (v: t nat n) y,
-    (Forall (fun x =>  x <= y) v) -> max_v v <= y.
-
-Lemma max_v_ge : forall n (v: t nat n) y,  In  y  v -> y <= max_v v.
-\end{Coqsrc}
+
+\input{movies/snippets/MoreVectors/maxvDef}
+
+\input{movies/snippets/MoreVectors/maxvLemmasa}
+
+\input{movies/snippets/MoreVectors/maxvLemmasb}
+
+\input{movies/snippets/MoreVectors/maxvLemmasc}
 
@@
 
-\begin{Coqsrc}
-Fixpoint evalList (m : nat) (l : Vector.t nat m) {struct l} :
- naryFunc m -> nat :=
-  match l in (Vector.t _ m) return (naryFunc m -> nat) with
-  | Vector.nil => fun x : naryFunc 0 => x
-  | Vector.cons a n l' => fun x : naryFunc (S n) => evalList n l' (x a)
-  end.
-\end{Coqsrc}
+\input{movies/snippets/primRec/evalListDef}
 
@@
 \index{coq}{Dependently typed functions}
-\begin{Coqsrc}
-(**  uncurried apply:
-
-[v_apply f (x1::x2:: ... ::xn::nil)]  is [f x1 x2 ... xn] 
- *)
-
-Notation "'v_apply' f v" := (evalList _ v f) 
-     (at level 10, f at level 9).
-
-Example Ex2: forall (f: naryFunc 2) x y,
-    v_apply f (x::y::nil) = f x y.
-Proof.   intros; now cbn. Qed.
-
-Example Ex4: forall (f: naryFunc 4) x y z t,
-    v_apply f (x::y::z::t::nil) = f x y z t.
-Proof.  intros; now cbn. Qed.
-\end{Coqsrc}
+
+\input{movies/snippets/AckNotPR/vApply}
 
@@
 \index{coq}{Dependently typed functions}
-\begin{Coqsrc}
-Definition majorized {n} (f: naryFunc n) (A: naryFunc 2) : Prop :=
-  exists (q:nat), forall (v: t nat n),
-      v_apply f v <= A q  (max_v v).
-
-Definition majorizedPR {n} (x: PrimRec n) A := 
-           majorized (evalPrimRec n x) A.
-
-(** For vectors of functions *)
-
-Definition majorizedS {n m} (fs : Vector.t (naryFunc n) m)
-           (A : naryFunc 2):=
-  exists N, forall (v: t nat n),
-      max_v (map (fun f => v_apply f v) fs) <= A N (max_v v).
-
-Definition majorizedSPR {n m} (x : PrimRecs n m) :=
-  majorizedS (evalPrimRecs _ _ x).
-\end{Coqsrc}
+
+\input{movies/snippets/AckNotPR/majorizedDefs}
 

### 2021-07-26 doc/chapter-primrec.tex
chapter 9 with alectryon: first draft

@@
 
-\begin{Coqsrc}
-Lemma majorSucc : majorizedPR  succFunc Ack.
+\input{movies/snippets/AckNotPR/majorSucc}
 
-Lemma majorZero : majorizedPR  zeroFunc Ack.
+\input{movies/snippets/AckNotPR/majorZero}
 
-Lemma majorProjection (n m:nat)(H: m < n):
-  majorizedPR (projFunc n m H) Ack.
-\end{Coqsrc}
+\input{movies/snippets/AckNotPR/majorProjection}
 
-The rest of the cases are proved within a mutual  induction.
+The remaining cases are proved within a mutual  induction.
 
@@
 
-\begin{Coqsrc}
-Lemma majorAnyPR:  forall n (x: PrimRec n),  majorizedPR  x Ack.
-Proof.
-  intros n x; induction x using PrimRec_PrimRecs_ind with
-                  (P0 := fun n m y => majorizedSPR  y Ack).
-  - apply majorSucc.
-  - apply majorZero.
-  - apply majorProjection. 
-  \end{Coqsrc}
-
-  \begin{Coqsrc}
-  - (** function composition *)
-\end{Coqsrc}
-\begin{Coqanswer}
-1 subgoal (ID 265)
-
-  n, m : nat
-  g : PrimRecs n m
-  x : PrimRec m
-  IHx : majorizedSPR g Ack
-  IHx0 : majorizedPR x Ack
-  ============================
-  majorizedPR (composeFunc n m g x) Ack
-\end{Coqanswer}
-
-\begin{Coqsrc}
-  ...
- - (** primitive recursion *)
-\end{Coqsrc}
-
-\begin{Coqanswer}
-1 subgoal (ID 265)
-
-  n : nat
-  x1 : PrimRec n
-  x2 : PrimRec (S (S n))
-  IHx1 : majorizedPR x1 Ack
-  IHx2 : majorizedPR x2 Ack
-  ============================
-  majorizedPR (primRecFunc n x1 x2) Ack
-\end{Coqanswer}
-
-\begin{Coqsrc}
- assert (L1 : forall i (v: t nat n) ,
-               v_apply f (i::v)  <= Ack q (i + max_v v)).
-    { induction i.
-      ...
-    }
-    ...
--  (** empty list of functions *)
-\end{Coqsrc}
-
-\begin{Coqanswer}
-1 subgoal (ID 266)
-
-  n : nat
-  ============================
-  majorizedSPR (PRnil n) Ack
-\end{Coqanswer}
-
-\begin{Coqsrc}
-  ...
-- (** non-empty list of functions *)
-\end{Coqsrc}
-
-\begin{Coqanswer}
-1 subgoal (ID 273)
-
-  n, m : nat
-  x : PrimRec n
-  p : PrimRecs n m
-  IHx : majorizedPR x Ack
-  IHx0 : majorizedSPR p Ack
-  ============================
-  majorizedSPR (PRcons n m x p) Ack
-\end{Coqanswer}
-
-\begin{Coqsrc}
- ...
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/AckNotPR/majorAnyPR}
 
@@
 
-\begin{Coqsrc}
-Lemma majorPR2 (f: naryFunc 2)(Hf : isPR 2 f)
-  : exists (n:nat), forall x y,  f x y <= Ack n (max x  y).
-\end{Coqsrc}
+\input{movies/snippets/AckNotPR/majorPR2}
 
@@
 
-\begin{Coqsrc}
-Lemma Ack_strict_mono_l : forall n m p, n < m ->
-                                        Ack n (S p) < Ack m (S p).
-\end{Coqsrc}
+\input{movies/snippets/Ack/AckStrictMonoL}
 
@@
 
-\begin{Coqsrc}
-Lemma majorPR2_strict (f: naryFunc 2)(Hf : isPR 2 f):
-    exists (n:nat),
-    forall x y, 2 <= x -> 2 <= y -> f x y < Ack n (max x  y).
-\end{Coqsrc}
+\input{movies/snippets/AckNotPR/majorPR2Strict}
 
@@
 
-\begin{Coqsrc}
-Section Impossibility_Proof.
-
-  Hypothesis HAck : isPR 2 Ack.
-
-  Lemma Ack_not_PR : False.
-  Proof.
-    destruct (majorPR2_strict Ack HAck) as [m Hm];
-    pose (X := max 2 m); specialize (Hm X X).
-    rewrite max_idempotent in Hm; 
-    assert (Ack m X <= Ack X X) by (apply Ack_mono_l; lia).
-    lia.
-  Qed.
-
-End Impossibility_Proof.
-\end{Coqsrc}
+\input{movies/snippets/AckNotPR/AckNotPR}
 
 \begin{remark}
-It is easy to prove that any unary function which dominatates \texttt{fun n => Ack n n} fails to be primitive recursive. We use an instance of \texttt{majorAnyPR} for unary functions.
+It is easy to prove that any unary function which dominatates \texttt{fun n => Ack n n} fails to be primitive recursive.
+To this end, we use an instance of \texttt{majorAnyPR} dealing with unary functions.
 
@@
 \emph{From \href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR}}.
-\begin{Coqsrc}
-Lemma majorPR1  (f: naryFunc 1)(Hf : isPR 1 f)
-  : exists (n:nat), forall x, f x <= Ack n x.
-(* ... *)
-\end{Coqsrc}
+
+\input{movies/snippets/AckNotPR/majorPR1}
 
@@
 
-\begin{Coqsrc}
-Section dom_AckNotPR.
+\input{movies/snippets/AckNotPR/domAckNotPR}
 
-  Variable f : nat -> nat.
-  Hypothesis Hf : dominates f (fun n => Ack n n).
-
- Lemma dom_AckNotPR: isPR 1 f -> False.
-  Proof.
-    intros H;  destruct Hf as [N HN].
-    destruct  (majorPR1 _ H) as [M HM].
-    pose (X := Max.max N M).
-    specialize (HN X  (Max.le_max_l N M)).
-    specialize (HM X);
-      assert (Ack M X <= Ack X X) by (apply Ack_mono_l; subst; lia).
-    lia.
-  Qed.
-
-End dom_AckNotPR.
-\end{Coqsrc}
 \end{remark}
@@
 
-\begin{Coqsrc}
-Definition l_std alpha k := (L_ alpha (S k) - k)%nat.
-
-Lemma l_std_ok : forall alpha : E0,
-    alpha <> Zero ->
-    forall k : nat,
-      1 <= k -> battle_length standard k (iota (cnf alpha))
-                              (l_std alpha k).
-(* ... *)
-\end{Coqsrc}
+\input{movies/snippets/Battle_length/BattleLength}
 
@@
 
-\begin{Coqsrc}
-Section battle_lenght_notPR.
-
-  (** We assume that the function with computes the length 
-      of standard battles is primitive recursive *)
-
-  Hypothesis H: forall alpha, isPR 1 (l_std alpha). 
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Theorems/battleLengthNotPRa}
 
@@
 
-\begin{Coqsrc}
-Let alpha := Phi0 omega%e0.
-Let h := iota (cnf alpha).
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Theorems/battleLengthNotPRb}
 
@@
 
-\begin{Coqsrc}
-Let m k := L_ alpha (S k).
-
-Remark m_eqn : forall k, m k = (l_std alpha k + k)%nat.
-(* ... *)
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Theorems/battleLengthNotPRc}
 
@@
 
-\begin{Coqsrc}
-Remark mIsPR : isPR 1 m.
-(* ... *)
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Theorems/battleLengthNotPRd}
 
@@
 
-In \href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}, we prove a relation between the $F$ and $H'$ functional. For any $\alpha$ and $k>0$,
+In \href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}, we prove a relation between the $F$ and $H'$ functionals. For any $\alpha$ and $k>0$,
 $H'_{\omega^\alpha}(k)\geq F_\alpha(k)$.
 
-\begin{Coqsrc}
-H'_F : forall (alpha : E0) (n : nat), 
-     F_ alpha (S n) <= H'_ (Phi0 alpha) (S n) 
-\end{Coqsrc}
+\input{movies/snippets/F_alpha/HprimeF}
 
-Our proof of this lemma is not trivial at all. One of it sub-goals is the following one:
+Our proof of this lemma is not trivial at all, it uses some properties of the Ketonen-Solovay's toolkit. We advise the reader to explore this proof, with the help of an ide or softwre like alectryon.
+%%
+%%% To move the path chapter
+%%%%
 
-\begin{Coqsrc}
-  alpha : E0
-  IHalpha : forall beta : E0, beta o< alpha -> P beta
-  Halpha : Limitb alpha
-  n : nat
-  ============================
-  H'_ (Phi0 (CanonS alpha n)) (S n) <= 
-  H'_ (Phi0 (CanonS alpha (S n))) (S n)
-\end{Coqsrc}
+% \begin{Coqsrc}
+%   alpha : E0
+%   IHalpha : forall beta : E0, beta o< alpha -> P beta
+%   Halpha : Limitb alpha
+%   n : nat
+%   ============================
+%   H'_ (Phi0 (CanonS alpha n)) (S n) <= 
+%   H'_ (Phi0 (CanonS alpha (S n))) (S n)
+% \end{Coqsrc}
 
-In mathematical notation: $H'_{\omega^{\canonseq{\alpha}{n}}}(n+1) \leq
-H'_{\omega^{\canonseq{\alpha}{n+1}}}(n+1)$.
+% In mathematical notation: $H'_{\omega^{\canonseq{\alpha}{n}}}(n+1) \leq
+% H'_{\omega^{\canonseq{\alpha}{n+1}}}(n+1)$.
 
-\vspace{4pt}
+% \vspace{4pt}
 
-But there exists no lemma saying that, if 
-$\beta\leq \alpha$, then $H'_\beta(k)\leq H'_\alpha(k)$, for any $\alpha$ and $\beta$. For instance, 
-$H'_{42}(3)=45> H'_\omega(3)=7$.
+% But there exists no lemma saying that, if 
+% $\beta\leq \alpha$, then $H'_\beta(k)\leq H'_\alpha(k)$, for any $\alpha$ and $\beta$. For instance, 
+% $H'_{42}(3)=45> H'_\omega(3)=7$.
 
-Looking for lemmas of the form $H'_\beta(k)\leq H'_\alpha(k)$, we find this one (from our library
-\href{../theories/html/hydras.Epsilon0.Hprime.html}{Epsilon0.Hprime}):
+% Looking for lemmas of the form $H'_\beta(k)\leq H'_\alpha(k)$, we find this one (from our library
+% \href{../theories/html/hydras.Epsilon0.Hprime.html}{Epsilon0.Hprime}):
 
-\begin{Coqanswer}
-H'_restricted_mono_l : 
-    forall (alpha beta : E0) (n : nat), 
-      Canon_plus n alpha beta -> 
-      H'_ beta n <= H'_ alpha n.
-\end{Coqanswer}
+% \begin{Coqanswer}
+% H'_restricted_mono_l : 
+%     forall (alpha beta : E0) (n : nat), 
+%       Canon_plus n alpha beta -> 
+%       H'_ beta n <= H'_ alpha n.
+% \end{Coqanswer}
 
-Thus, it remains to prove that 
-there exists a path from ${\omega^{\canonseq{\alpha}{n+1}}}$
-to ${\omega^{\canonseq{\alpha}{n}}}$ composed of 
-$n+1$-steps.
+% Thus, it remains to prove that 
+% there exists a path from ${\omega^{\canonseq{\alpha}{n+1}}}$
+% to ${\omega^{\canonseq{\alpha}{n}}}$ composed of 
+% $n+1$-steps.
 
-Fortunately, the Ketonen-Solovay machinery contains three lemmas which help us to build such a path.
+% Fortunately, the Ketonen-Solovay machinery contains three lemmas which help us to build such a path.
 
-\begin{Coqanswer}
-KS_thm_2_4_lemma5 :
-  forall [i : nat] [alpha beta : T1],
-  const_pathS i alpha beta ->
-  nf alpha -> alpha <> zero -> 
-  const_pathS i (phi0 alpha) (phi0 beta)
+% \begin{Coqanswer}
+% KS_thm_2_4_lemma5 :
+%   forall [i : nat] [alpha beta : T1],
+%   const_pathS i alpha beta ->
+%   nf alpha -> alpha <> zero -> 
+%   const_pathS i (phi0 alpha) (phi0 beta)
 
-KS_thm_2_4 :
-  forall [lambda : T1], nf lambda ->limitb lambda ->
-  forall i j : nat, i < j -> 
-   const_pathS 0 (canonS lambda j) (canonS lambda i)
+% KS_thm_2_4 :
+%   forall [lambda : T1], nf lambda ->limitb lambda ->
+%   forall i j : nat, i < j -> 
+%    const_pathS 0 (canonS lambda j) (canonS lambda i)
 
-Cor12_1 :
-forall [alpha : T1], nf alpha ->
-      forall (beta : T1) (i n : nat),
-      beta t1< alpha ->
-     i <= n -> const_pathS i alpha beta -> 
-     const_pathS n alpha beta
-\end{Coqanswer}
+% Cor12_1 :
+% forall [alpha : T1], nf alpha ->
+%       forall (beta : T1) (i n : nat),
+%       beta t1< alpha ->
+%      i <= n -> const_pathS i alpha beta -> 
+%      const_pathS n alpha beta
+% \end{Coqanswer}
 
@@
 
-\begin{Coqsrc}
-H'_L_ : forall (alpha : E0) (i : nat), H'_ alpha i <= L_ alpha (S i)
-\end{Coqsrc}
+\input{movies/snippets/L_alpha/HprimeL}
 
@@
 
-\begin{Coqsrc}
-F_vs_Ack : forall n : nat, 2 <= n -> Ack n n <= F_ omega n
-\end{Coqsrc}
+\input{movies/snippets/F_omega/FVsAck}
 
@@
 
-\begin{Coqsrc}
-Remark m_ge_F_omega : forall k,  F_ omega (S k) <= m (S k).
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Theorems/mGeFOmega}
 
@@
 
-\begin{Coqsrc}
-Remark m_dominates_Ack : 
-     dominates (fun n =>  S (m n)) (fun n => Ack.Ack n n).
-(* ... *)
+\input{movies/snippets/Hydra_Theorems/mDominatesAck}
 
-Remark SmNotPR : isPR 1 (fun n => S (m n)) -> False.
-(* ... *)
+\input{movies/snippets/Hydra_Theorems/SmNotPR}
 
-Theorem LNotPR : False.
-  Proof.
-    apply SmNotPR,  compose1_1IsPR.
-    - apply mIsPR.
-    - apply succIsPR.
-  Qed.
-
-End battle_lenght_notPR.
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Theorems/LNotPR}

### 2021-07-26 doc/chapter-primrec.tex
draft version of chapter 9 (hybdrid version)
Contains a few hand-made Coqanswer blocks

@@
 
+\begin{Coqanswer}
+3 subgoals (ID 184)
+
+  ============================
+  isPR 1 (fun _ : nat => 2)
+
+subgoal 2 (ID 185) is:
+ isPR 1 (fun x : nat => x)
+subgoal 3 (ID 186) is:
+ isPR 2 Init.Nat.mul
+\end{Coqanswer}
+
+\input{movies/snippets/PrimRecExamples/doubleIsPRb}
+
 \index{primrec}{Exercises}
@@
 
-\input{movies/snippets/AckNotPR/majorAnyPR}
+\input{movies/snippets/AckNotPR/majorAnyPRa}
+
+\begin{Coqanswer}
+ x : PrimRec m
+ x0 : nat
+ H : forall v : t nat n,
+   max_v (map (fun f : naryFunc n => v_apply f v)
+           (evalPrimRecs n m g)) <=
+   Ack x0 (max_v v)
+ x1 : nat
+ H0 : forall v : t nat m, v_apply (evalPrimRec m x) v <=
+                                  Ack x1 (max_v v)
+  ============================
+  forall v : t nat n,
+  v_apply (evalPrimRec n (composeFunc n m g x)) v <=
+  Ack (2 + Nat.max x0 x1) (max_v v)
+
+\end{Coqanswer}
+
+\input{movies/snippets/AckNotPR/majorAnyPRb}
+
+\begin{Coqanswer}
+  n : nat
+  x1 : PrimRec n
+  x2 : PrimRec (S (S n))
+  r : nat
+  Hg : forall v : t nat n,
+     v_apply (evalPrimRec n x1) v <= Ack r (max_v v)
+  s : nat
+  Hh : forall v : t nat (S (S n)),
+       v_apply (evalPrimRec (S (S n)) x2) v <= Ack s (max_v v)
+  ============================
+  majorizedPR (primRecFunc n x1 x2) Ack
+\end{Coqanswer}
+
+The las two goals deal with vectors of functions.
+
+\begin{Coqanswer}
+1 subgoal (ID 289)
+
+  n : nat
+  ============================
+  majorizedSPR (PRnil n) Ack
+
+\end{Coqanswer}
+
+\begin{Coqanswer}
+1 subgoal (ID 296)
+
+  n, m : nat
+  x : PrimRec n
+  p : PrimRecs n m
+  IHx : majorizedPR x Ack
+  IHx0 : majorizedSPR p Ack
+  ============================
+  majorizedSPR (PRcons n m x p) Ack
+\end{Coqanswer}
+
+\begin{Coqsrc}
+Qed.
+\end{Coqsrc}
 
@@
 
-Our proof of this lemma is not trivial at all, it uses some properties of the Ketonen-Solovay's toolkit. We advise the reader to explore this proof, with the help of an ide or softwre like alectryon.
+Our proof of this lemma is not trivial at all, it uses some properties of the Ketonen-Solovay's toolkit. We advise the reader to explore this proof, with the help of an ide or software like alectryon.
 %%
@@
 
+\vspace{4pt}
+
 \emph{From \href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0.L\_alpha}}
@@
 
+\vspace{4pt}
+
 \emph{From \href{../theories/html/hydras.Epsilon0.F_omega.html}{Epsilon0.F\_omega}}
+\input{movies/snippets/F_omega/FVsAck}
 
-\input{movies/snippets/F_omega/FVsAck}
+\vspace{4pt}
 
@@
 
+\vspace{4pt}
+
 \input{movies/snippets/Hydra_Theorems/LNotPR}

### 2021-07-27 doc/chapter-primrec.tex
start to add snippets to chapter 2 (hydras)

@@
 
-In particular, the library \href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR} contains the well-known  proof that the Ackermann function is not primitive recursive (see Section~\vref{sect:ack-not-PR}).
-MoreOver, the library \href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems} contains 
+In particular, the library \href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/MoreAck/AckNotPR.v}{MoreAck.AckNotPR} contains the well-known  proof that the Ackermann function is not primitive recursive (see Section~\vref{sect:ack-not-PR}).
+Moreover, the library \href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/Hydra/Hydra_Theorems.v}{Hydra.Hydra\_Theorems} contains 
 a proof that the length of an hydra battle (according to the initial replication factor) is not primitive recursive in general.
@@
 The formal definition of primitive recursive functions lies in the library
-\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec},
+\href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/Ackermann/primRec.v}{Ackermann.primRec},
 with preliminary definitions in 
-\href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat}
+\href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/Ackermann/extEqualNat.v}{Ackermann.extEqualNat}
 and
-\href{../theories/html/hydras.Ackermann.misc.html}{Ackermann.misc}.
+\href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/Ackermann/misc.v}{Ackermann.misc}.
 
@@
 You may start this exercise with the file
-    \href{https://https://github.com/coq-community/hydra-battles/blob/master/exercises/primrec/MorePRExamples.v}{exercises/primrec/MorePRExamples.v}.
+    \href{https://github.com/coq-community/hydra-battles/blob/master/exercises/primrec/MorePRExamples.v}{exercises/primrec/MorePRExamples.v}.
 \end{exercise}
@@
 \emph{You may start this exercise with
-    \href{https://https://github.com/coq-community/hydra-battles/blob/master/exercises/primrec/MinPR.v}{exercises/primrec/MinPR.v}.}
+    \href{https://github.com/coq-community/hydra-battles/blob/master/exercises/primrec/MinPR.v}{exercises/primrec/MinPR.v}.}
 

### 2021-07-27 doc/part-hydras.tex
start to add snippets to chapter 2 (hydras)

@@
 
-\begin{Coqsrc}
-Inductive Hydra : Set :=
-| node :  Hydrae -> Hydra
-with Hydrae : Set :=
-| hnil : Hydrae
-| hcons : Hydra -> Hydrae -> Hydrae.
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Definitions/HydraDef}
 
@@
 
-\begin{Coqalt}
-Module Alt.
-
-Inductive Hydra: Set :=
-  hnode (daughters : list Hydra).
-
-End Alt.
-\end{Coqalt}
+\input{movies/snippets/Hydra_Definitions/HydraAlt}
 
@@
 
-\begin{Coqsrc}
-(** heads *)
-Notation head := (node hnil).
-
-(** nodes  with 1, 2 or 3 daughters *)
-Notation hyd1 h := (node (hcons h hnil)).
-Notation hyd2 h h' := (node (hcons h (hcons h' hnil))).
-Notation hyd3 h h' h'' := 
-                   (node (hcons h (hcons h' (hcons h'' hnil)))).
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Definitions/headsEtc}
 
@@
 
-\begin{Coqsrc}
-Example Hy := hyd3 head
-                   (hyd2
-                      (hyd1 
-                         (hyd2 head head))
-                      head) 
-                   head.
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Examples/Hy}
 
@@
 
-\begin{Coqsrc}
-Fixpoint hcons_mult (h:Hydra)(n:nat)(s:Hydrae):Hydrae :=
-  match n with 
-  | O => s
-  | S p => hcons h (hcons_mult h p s)
-  end.
-
-(** A node with n clones of the same daughter *)
-
-Definition hyd_mult h n :=
-  node (hcons_mult h n hnil).
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Definitions/hconsMult}
 
@@
 
-\begin{Coqsrc}
-Example Hy'' := 
-     hyd2 head
-          (hyd2 (hyd_mult (hyd1 head) 5)
-                head).
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Examples/HySecond}
 
@@
 
-\begin{Coqsrc}
-Fixpoint hsize (h:Hydra) : nat :=
-  match h with node l => S (lhsize l)
-  end
-with lhsize l : nat :=
-  match l with hnil => 0
-            | hcons h hs => hsize h + lhsize hs 
-  end.
-
- Compute hsize Hy.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = 9
-     : nat 
-\end{Coqanswer}
+\input{movies/snippets/Hydra_Definitions/hsize}
+
+\input{movies/snippets/Hydra_Examples/HySize}
 
@@
 
-\begin{Coqsrc}
-Fixpoint height  (h:Hydra) : nat :=
-  match h with node l => lheight l
-  end
-with lheight l : nat :=
-  match l with 
-  | hnil => 0
-  | hcons h hs => Max.max (S (height h)) (lheight hs)
-  end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Compute height Hy.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = 4
-     : nat  
-\end{Coqanswer}
+\input{movies/snippets/Hydra_Definitions/height}
+
+\input{movies/snippets/Hydra_Examples/HyHeight}
 
@@
 
-\begin{Coqanswer}
-Hydra_ind :
-forall P : Hydra -> Prop,
-(forall h : Hydrae, P (node h)) -> forall h : Hydra, P h
-\end{Coqanswer}
-
-Ler us start a simple proof by induction.
+\input{movies/snippets/Hydra_Examples/HydraInd}
+
+Let us start a simple proof by induction.
 
@@
 
-\begin{Coqbad}
-Module Bad.
-
-Lemma height_lt_size (h:Hydra) :
-  height h <= hsize h.
-Proof.
-  induction h as [s].
-\end{Coqbad}
-
-\begin{Coqanswer}
-1 subgoal, subgoal 1 (ID 11)
-
-  s : Hydrae
-  ============================
-   height (node s) <= hsize (node s)
-\end{Coqanswer}
+\input{movies/snippets/Hydra_Examples/BadInductiona}
 
@@
 
-% \begin{Coqbad}
-% induction  s as [ | h s'].
-%   -   cbn; auto with arith.
-%   - 
-% \end{Coqbad}
-
-\begin{Coqanswer}
- 1 focused subgoal
-(unfocused: 0), subgoal 1 (ID 19)
-
-  h : Hydra
-  s' : Hydrae
-  IHs' : height (node s') <= hsize (node s')
-  ============================
-   height (node (hcons h s')) <= hsize (node (hcons h s'))
-
-\end{Coqanswer}
-
-Note that the displayed subgoal does not contain any assumption on \texttt{h}, thus there is no way to 
-infer any property about the height and size of the hydra \texttt{(hcons h t)}.
-
-\begin{Coqbad}
-Abort.
-
-End Bad.
-\end{Coqbad}
+\input{movies/snippets/Hydra_Examples/BadInductionb}
+
+The first subgoal is trivial.
+
+\input{movies/snippets/Hydra_Examples/BadInductionc}
+
+Let us look at the second sub-goal of the induction.
+
+\input{movies/snippets/Hydra_Examples/BadInductiond}
+
+We notice that this subgoal does not contain any hypothesis
+on the height and size of the hydra \texttt{h}. So, it looks hard to prove the conclusion. Let's stop.
+
+\input{movies/snippets/Hydra_Examples/BadInductione}
 

### 2021-07-29 doc/hydras.tex
Use LaTeX snippets extracted with Alectryon.
- add cli to convert coq file to latex
- cut coq file with snippet
- first approch to make latex
- add paramater backend and frontend
- parse coq+rst and extract snippet to latex
- add Readme
- Make a version snippet extractor with latex conversion
- Use latex format to extract snippets
- move cutting table in move and create edit makefile
- try to make snipet in Ackerman to replace in latex file
Copy sty alectryon file and more
- copy alectryon latex assets
- Add rule to clear latex generation with make clean
- Add requirements.txt file and change

@@
 \usepackage{theorem}
+
+\usepackage{texments}
+%%% for movies by alectryon
+\usepackage{movies/snippets/assets/alectryon}
+\usepackage{movies/snippets/assets/tango_subtle}
 
@@
 \end{itemize}
+
+\subsection{Documenting theories with alectryon}
+
+Quotations of \coq{} source and answers are progressively replaced from copy-pasted \emph{verbatim} to automatically generated \emph{LaTeX} blocks, using Clément Pit-Claudel's \emph{Alectryon} tool~\cite{alectryonpaper, alectryongithub}.
+Many thanks to Jérémy Damour and Théo Zimmermann who designed tools for maintaining consistency between the always evolving \coq{} modules and their documentation written in \emph{LaTeX}.
+
+Besides the guarantee of consistency between theories and documentation, we hope to give a corpus for experimenting new ways of documenting the implementation of non-trivial mathematics on a proof assistant.
+
+At present, this document is a hybrid version: Chapter~\vref{chapter:primrec} is fully adapted to \emph{Alectryon}, and we just started to re-write Chapter~\vref{chapter:hydras}.
 

### 2021-07-29 doc/hydras.tex
moved a paragraph

@@
 
-\subsection{Documenting theories with alectryon}
-
-Quotations of \coq{} source and answers are progressively replaced from copy-pasted \emph{verbatim} to automatically generated \emph{LaTeX} blocks, using Clément Pit-Claudel's \emph{Alectryon} tool~\cite{alectryonpaper, alectryongithub}.
-Many thanks to Jérémy Damour and Théo Zimmermann who designed tools for maintaining consistency between the always evolving \coq{} modules and their documentation written in \emph{LaTeX}.
-
-Besides the guarantee of consistency between theories and documentation, we hope to give a corpus for experimenting new ways of documenting the implementation of non-trivial mathematics on a proof assistant.
-
-At present, this document is a hybrid version: Chapter~\vref{chapter:primrec} is fully adapted to \emph{Alectryon}, and we just started to re-write Chapter~\vref{chapter:hydras}.
-
 \paragraph*{Warning:}
@@
 %%%%% ICI
+
+\subsection{Documenting theories with alectryon}
+
+Quotations of \coq{} source and answers are progressively replaced from copy-pasted \emph{verbatim} to automatically generated \emph{LaTeX} blocks, using Clément Pit-Claudel's \emph{Alectryon} tool~\cite{alectryonpaper, alectryongithub}.
+Many thanks to Jérémy Damour and Théo Zimmermann who designed tools for maintaining consistency between the always evolving \coq{} modules and documentation written in \emph{LaTeX}.
+
+Besides the guarantee of consistency between theories and documentation, we hope to give a corpus for experimenting new ways of documenting the implementation of non-trivial mathematics on a proof assistant.
+
+At present, this document is a hybrid version: Chapter~\vref{chapter:primrec} is fully adapted to \emph{Alectryon}, and we just started to re-write Chapter~\vref{chapter:hydras}.
 

### 2021-07-29 doc/chapter-primrec.tex

--- 2021-07-26 doc/chapter-primrec.tex
@@
 \chapter{Primitive recursive functions}
+\label{chapter:primrec}
 
@@
 
-In particular, the library \href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR} contains the well-known  proof that the Ackermann function is not primitive recursive (see Section~\vref{sect:ack-not-PR}).
-MoreOver, the library \href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems} contains 
+In particular, the library \href{../theories/html/hydras.MoreAck.html}{MoreAck.AckNotPR} contains the well-known  proof that the Ackermann function is not primitive recursive (see Section~\vref{sect:ack-not-PR}).
+Moreover, the library \href{../theories/html/hydras.Hydra_Theorems.html}{Hydra.Hydra\_Theorems} contains 
 a proof that the length of an hydra battle (according to the initial replication factor) is not primitive recursive in general.

### 2021-07-29 doc/part-hydras.tex

@@
 \label{sec:orgheadline91}
+\label{chapter:hydras}
 
@@
 
-\begin{Coqsrc}
-Scheme Hydra_rect2 := Induction for Hydra Sort Type
-with Hydrae_rect2 := Induction for Hydrae Sort Type.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Check Hydra_rect2.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-Hydra_rect2
- : forall (P : Hydra -> Type) (P0 : Hydrae -> Type),
-   (forall h : Hydrae, P0 h -> P (node h)) ->
-    P0 hnil ->
-    (forall h : Hydra, P h -> 
-             forall h0 : Hydrae, P0 h0 -> P0 (hcons h h0)) ->
-    forall h : Hydra, P h
-\end{Coqanswer}
+\input{movies/snippets/Hydra_Definitions/HydraRect2}
+
+\input{movies/snippets/Hydra_Examples/HydraRect2Check}
 
@@
 \noindent
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
+
+\input{movies/snippets/Hydra_Definitions/hForall}
+
 \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
 
-\begin{Coqsrc}
-(** All elements of s satisfy P *)
-
-Fixpoint h_forall (P: Hydra -> Prop) (s: Hydrae) :=
-  match s with
-      hnil => True
-    | hcons h s' => P h /\ h_forall P s'
-  end.
- \end{Coqsrc}
-
- \begin{Coqsrc}
-Lemma height_lt_size (h:Hydra) :
- height h < hsize h.
-Proof.
-  induction h using Hydra_rect2  with 
-  (P0 :=  h_forall (fun h =>  height h < hsize h)).
- \end{Coqsrc}
+\input{movies/snippets/Hydra_Examples/heightLtSizea}
 

### 2021-07-29 doc/part-hydras.tex
converting chapter 2 to alectryon (continued)

@@
 
-\begin{Coqsrc}
-Inductive S0 :  relation Hydrae :=
-| S0_first : forall s, S0  (hcons head s) s
-| S0_rest : forall  h s s',  S0  s s' ->
-                             S0  (hcons h s) (hcons h s').
-
-Inductive R1  :  Hydra -> Hydra -> Prop :=
-| R1_intro : forall s s', S0 s s' -> R1 (node s) (node s').
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Definitions/S0Def}
+
+\input{movies/snippets/Hydra_Definitions/R1Def}
 
@@
 
-\begin{Coqsrc}
-Example Hy_1 : R1 Hy Hy'.
-Proof. 
-  split; right; right; left.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Examples/Hy1}
 
@@
 
-\begin{Coqsrc}
-Inductive S1 (n:nat)  : Hydrae -> Hydrae -> Prop :=
-| S1_first : forall s h h' ,   R1 h h' -> 
-                  S1 n (hcons h s) (hcons_mult h' (S n) s)
-| S1_next : forall h s s',  S1 n s s' ->
-                   S1 n (hcons h s) (hcons h s').
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Definitions/S1Def}
 
@@
 
-\begin{Coqsrc}
-Inductive R2 (n:nat)  :  Hydra -> Hydra -> Prop :=
-| R2_intro : forall s s', S1 n s s' -> R2 n (node s) (node s')
-| R2_intro_2 : forall s s', S2 n s s' -> R2 n (node s) (node s')
-
-with S2 (n:nat) :  Hydrae -> Hydrae -> Prop :=
-|  S2_first : forall h h' s ,
-               R2 n  h h'  -> 
-               S2  n (hcons h s) (hcons h'  s)
-|  S2_next  : forall h   r r',
-               S2 n   r r' ->
-               S2 n (hcons h r) (hcons h r').                  
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Definitions/R2Def}
 
@@
 
-\begin{Coqsrc}
-Example R2_example:  R2 4 Hy' Hy''.
-Proof.
-    (** move to 2nd sub-hydra (0-based indices) *) R2_up 1. 
-    (** move to first sub-hydra *)  R2_up 0.
-    (** we're at distance 2 from the to-be-chopped-off head 
-        let's go to the first daughter, 
-        then chop-off the leftmost head *)  r2_d2  0 0. 
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Examples/R2Example}
 

### 2021-07-30 doc/part-hydras.tex
converting chapter 2 to alectryon (continued)

@@
 \index{hydras}{Library Hydra!Predicates!round\_n}
-\begin{Coqsrc}
-Definition round_n n h h' := R1 h h' \/ R2 n h h'.  
-\end{Coqsrc}
+
+\input{movies/snippets/Hydra_Definitions/roundNDef}
 
@@
 \label{sect:infix-round}
-\begin{Coqsrc}
-Definition round h h' := exists n,  round_n n h h'.
-
-Infix "-1->" := round (at level 60).
-\end{Coqsrc}
+
+\input{movies/snippets/Hydra_Definitions/roundDef}
 
@@
 
-\begin{Coqsrc}
-Definition round_plus := clos_trans_1n Hydra round.
-Infix "-+->" := rounds (at level 60).
-
-Definition round_star h h' := h = h' \/ round_plus h h'.
-Infix "-*->" := round_star (at level 60).
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Definitions/roundPlus}
 
@@
 \begin{exercise}
-Prove the following lemma:
-
-\begin{Coqsrc}
-Lemma rounds_height : forall h h', 
-   h -+-> h' -> height h' <= height h.  
-\end{Coqsrc}
+  Prove that if \texttt{$h$ -+-> $h'$}, then
+  the height of $h'$ is less or equal than the height of $h$.
 
@@
 
-\begin{Coqsrc}
-Definition dep_round_t := nat -> Hydra -> Hydra -> Prop.
-
-Class Battle :=  
-    { battle_rel : dep_round_t;
-      battle_inclusion : forall i h h', battle_rel i h h' -> round h h'}.
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Definitions/BattleDef}
 
@@
 
-\begin{Coqsrc}
-Program Instance free : Battle := Build_Battle (fun _  h h' => round h h') _.
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Definitions/freeDef}
 
@@
 
-\begin{Coqsrc}
-Program Instance standard : Battle := (Build_Battle round_n _).
-Next Obligation.
-  now exists i.  
-Defined.
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Definitions/standardDef}
 
@@
 
-\begin{Coqsrc}
-Inductive battle (B:Battle) : nat -> Hydra -> nat -> Hydra -> Prop :=
-| battle_1 : forall i h  h', 
-      battle_rel   B i  h h' -> battle B i h (S i) h'
-| battle_n : forall i h  j h' h'',  
-      battle_rel  B i h h''  ->
-      battle B (S i) h'' j h'  ->
-      battle B i h j h'.
-\end{Coqsrc}
-
-The following property allows us to build battles out of smaller ones.
-
-\begin{Coqsrc}
-Lemma battle_trans {b:Battle} :
-  forall i h j h', battle b i h j h' ->
-        forall k h0,  battle b k h0 i h -> battle b k h0 j h'.
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Definitions/battleRelDef}
+
+The following property allows us to build battles by composition of smaller ones.
+
+%% TODO display sub-goals when fixed
+
+\input{movies/snippets/Hydra_Lemmas/battleTrans}
 
@@
 
-\begin{Coqsrc}
-Definition hinit := hyd3 (hyd_mult head 3)  head head.  
-\end{Coqsrc}
+\input{movies/snippets/BigBattle/hinitDef}
 
@@
 
-\begin{Coqsrc}
-Lemma L_0_2 : battle standard 0 hinit 2 (hyd1 h3).   
-\end{Coqsrc}
+\input{movies/snippets/BigBattle/L02}
 
@@
 subtrees of height 1 or 0).
-Thus, we introduce a few handy notations.
-
-\begin{Coqsrc}
-Notation h3 := (hyd_mult head 3).
-Notation h2 := (hyd_mult head 2).
-Notation h1 := (hyd1 head).
-
-Definition hyd a b c := 
-  node (hcons_mult h2  a
-             (hcons_mult h1  b
-                         (hcons_mult head c hnil))).
-\end{Coqsrc}
+Thus, we introduce a few handy abbreviations.
+
+\input{movies/snippets/BigBattle/Notations}
 
@@
 
-\begin{Coqsrc}
-Lemma L_2_3 : battle standard 2 (hyd1 h3)  3 (hyd 3 0 0).
-
-Lemma L_0_3 : battle standard 0 hinit 3 (hyd 3 0 0).
-\end{Coqsrc}
+\input{movies/snippets/BigBattle/L23L03}
 
@@
 
-\begin{Coqsrc}
- Record state : Type :=
-    mks {round: nat ; n2 : nat ; n1 : nat ; nh : nat}.
-\end{Coqsrc}
+\input{movies/snippets/BigBattle/stateDef}
 
@@
 
-\begin{Coqsrc}
-Definition next (s : state) :=
-  match s with
-  | mks round a b (S c) => mks (S round) a b c
-  | mks round a (S b) 0 => mks (S round) a b (S round)
-  | mks round (S a) 0 0 => mks (S round) a (S round) 0
-  | _ => s
-  end.
-\end{Coqsrc}
+\input{movies/snippets/BigBattle/nextDef}
 
@@
 
-\begin{Coqsrc}
-Fixpoint iterate {A:Type}(f : A -> A) (n: nat)(x:A) :=
-  match n with
-  | 0 => x
-  | S p => f (iterate  f p x)
-  end.
-\end{Coqsrc}
+\input{movies/snippets/Iterates/iterateDef}
 
@@
 
-\begin{Coqsrc}
-Definition test n := iterate next (n-3) (mks 3 3 0 0).
-
-Compute test 3.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-      = {| round := 3; n2 := 3; n1 := 0; nh := 0 |}
-     : state 
-\end{Coqanswer}
-
-\begin{Coqsrc}
- Compute test 4.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = {| round := 4; n2 := 2; n1 := 4; nh := 0 |}
-     : state
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Compute test 5.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = {| round := 5; n2 := 2; n1 := 3; nh := 5 |}
-     : state
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Compute test 2000.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = {| round := 2000; n2 := 1; n1 := 90; nh := 1102 |}
-     : state
-\end{Coqanswer}
-
-The battle we are studying seems to be awfully long. Let us concentrate our
+\input{movies/snippets/BigBattle/testDefTests}
+
+The battle we are studying looks to be awfully long. Let us concentrate our
 tests on some particular events : the states where $\texttt{nh}=0$.
@@
 
- \begin{Coqsrc}
- Compute test 10.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-    = {| round := 10; n2 := 2; n1 := 3; nh := 0 |}
-     : state
- \end{Coqanswer}
+\input{movies/snippets/BigBattle/smartTest}
 
@@
 
-\begin{Coqsrc}
-Compute test 22.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = {| round := 22; n2 := 2; n1 := 2; nh := 0 |}
-     : state
-\end{Coqanswer}
-
-\begin{Coqsrc}
- Compute test 46.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-( = {| round := 46; n2 := 2; n1 := 1; nh := 0 |}
-     : state
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Compute test 94.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = {| round := 94; n2 := 2; n1 := 0; nh := 0 |}
-     : state
-\end{Coqanswer}
+\input{movies/snippets/BigBattle/smartTestb}
 
@@
 
-\begin{Coqsrc}
- Compute test 95.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = {| round := 95; n2 := 1; n1 := 95; nh := 0 |}
-     : state
-\end{Coqanswer}
+\input{movies/snippets/BigBattle/smartTestc}
 
@@
 
-\begin{Coqsrc}
-Definition doubleS (n : nat) := 2 * (S n).
-
-Compute test (doubleS 95).
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = {| round := 192; n2 := 1; n1 := 94; nh := 0 |}
-     : state
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Compute test (iterate doubleS 2 95).
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = {| round := 386; n2 := 1; n1 := 93; nh := 0 |}
-     : state
-\end{Coqanswer}
+\input{movies/snippets/BigBattle/doubleS}
 

### 2021-07-31 doc/part-hydras.tex
converting chapter 2 to alectryon (continued)

@@
 
-Thus, $ (1 + 11)$ rounds later, the \texttt{n1} field is equal to $2$, and 
+Thus, $(1 + 11)$ rounds later, the \texttt{n1} field is equal to $2$, and 
 \texttt{nh}   to $0$.
@@
 
-\begin{Coqsrc}
-Inductive one_step (i: nat) :
-  nat -> nat -> nat -> nat -> nat -> nat -> Prop :=
-| step1: forall a b c, one_step i a b (S c) a b c
-| step2:  forall a b, one_step i a (S b) 0 a b (S i)
-| step3: forall a, one_step i (S a) 0 0 a (S i) 0.
-\end{Coqsrc}
+\input{movies/snippets/BigBattle/oneStep}
 
@@
 
-\begin{Coqsrc}
-Lemma step_battle : forall i a b c a' b' c', 
-   one_step i a b c a' b' c' ->
-   battle standard i (hyd  a b c)  (S i) (hyd a' b' c').
-\end{Coqsrc}
+\input{movies/snippets/BigBattle/stepBattle}
+
+\vspace{4pt}
 
@@
 
-\begin{Coqsrc}
- Inductive steps : nat -> nat -> nat -> nat ->
-                  nat -> nat -> nat -> nat -> Prop :=
-| steps1 : forall i a b c a' b' c',
-    one_step i a b c a' b' c' -> steps i a b c (S i) a' b' c'
-| steps_S : forall i a b c j a' b' c' k a'' b'' c'',
-    steps i a b c j a' b' c' ->
-    steps j a' b' c' k a'' b'' c'' ->
-    steps i a b c k  a'' b'' c''.
-
-Definition reachable (i a b c : nat) : Prop :=
-  steps 3 3 0 0 i a b c.
-\end{Coqsrc}
+\input{movies/snippets/BigBattle/steps}
 
@@
 
-\begin{Coqsrc}
- Lemma steps_battle : forall i a b c j a' b' c', 
-   steps i a b c j a' b' c' ->
-   battle standard i (hyd  a b c)   j  (hyd a' b' c').
-\end{Coqsrc}
+\input{movies/snippets/BigBattle/stepsBattle}
+
+\vspace{4pt}
 
@@
 
-\begin{Coqsrc}
-Lemma LS : forall c a b i,  steps i a b (S c) (i + S c) a b 0.
-Proof.
-  induction c.
- -   intros;  replace (i + 1) with (S i).
-     + repeat constructor.
-     + ring.
- -  intros; eapply  steps_S.
-   +   eleft;   apply rule1.
-   +   replace (i + S (S c)) with (S i + S c) by ring;  apply IHc.
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma doubleS_law : forall  a b i, steps i a (S b) 0 (doubleS i) a b 0.
-Proof.
-  intros;  eapply steps_S.
-  +   eleft;   apply step2.
-  +   unfold doubleS; replace (2 * S i) with (S i + S i) by ring; 
-        apply LS.
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma reachable_S  : forall i a b, reachable i a (S b) 0 ->
-                                   reachable (doubleS i) a b 0.
-Proof.
-  intros; right with  (1 := H); apply doubleS_law.
-Qed.
-\end{Coqsrc}
-
-From now on, the lemma \texttt{reachable\_S} allows us to watch larger steps of 
+\vspace{4pt}
+
+\input{movies/snippets/BigBattle/LS}
+
+From now on, the lemma \texttt{reachable\_S} allows us to watch larger and larger steps of 
 the battle.
 
-\begin{Coqsrc}
- Lemma L4 : reachable 4 2 4 0.
-Proof.
-  left; constructor.
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma L10 : reachable 10 2 3 0.
-Proof.
-  change 10 with (doubleS 4).
-  apply reachable_S, L4.
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma L22 : reachable 22 2 2 0.
-Proof.
-  change 22 with (doubleS 10).
-  apply reachable_S, L10.
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma L46 : reachable 46 2 1 0.
-Proof.
-  change 46 with (doubleS 22); apply  reachable_S, L22.
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma L94 : reachable 94 2 0 0.
-Proof.
-  change 94 with (doubleS 46); apply reachable_S, L46.
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma L95 : reachable 95 1 95 0.
-Proof.
-  eapply steps_S.
-  -  eexact L94.
-  -  repeat constructor.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/BigBattle/L4}
+
+\input{movies/snippets/BigBattle/L10To95}
 
@@
 
-\begin{Coqsrc}
-Lemma Bigstep : forall b i a , reachable i a b 0 ->
-                               reachable (iterate doubleS b i) a 0 0.
- Proof.
-  induction b.
-  -  trivial.
-  -  intros;  simpl;   apply reachable_S in H.
-     rewrite <- iterate_comm; now apply IHb.
- Qed.
-\end{Coqsrc}
+\vspace{4pt}
+
+\input{movies/snippets/BigBattle/Bigstep}
+
+\vspace{4pt}
 
@@
 
-\begin{Coqsrc}
- Definition M := (iterate doubleS 95 95).
-
-Lemma L2_95 : reachable M 1 0 0.
-Proof.
-  apply Bigstep,  L95.
-Qed.
-\end{Coqsrc}
+\vspace{4pt}
+
+\input{movies/snippets/BigBattle/MDef}
 
@@
 
-\begin{Coqsrc}
-Lemma L2_95_S : reachable (S M) 0 (S M) 0.
-Proof.
-  eright.
-  - apply L2_95.
-  -  left; constructor 3.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/BigBattle/L295S}
+
+\vspace{4pt}
 
@@
 
-\begin{Coqsrc}
-Definition N :=   iterate doubleS (S M) (S M).
-
-Theorem   SuperbigStep : reachable N  0 0 0 .
-Proof.
-  apply Bigstep, L2_95_S.
-Qed.
-\end{Coqsrc}
+\vspace{4pt}
+
+\input{movies/snippets/BigBattle/NDef}
+
+\vspace{4pt}
 
@@
 
-\begin{Coqsrc}
-Lemma Almost_done :
-  battle standard 3 (hyd 3 0 0) N (hyd 0 0 0).
-Proof. 
-  apply steps_battle, SuperbigStep.
-Qed.
-
-Theorem Done :
-  battle standard 0 hinit N head.
-Proof.
-  eapply battle_trans.
-  -   apply Almost_done.
-  -  apply L_0_3.
-Qed.
-\end{Coqsrc}
+\vspace{4pt}
+
+\input{movies/snippets/BigBattle/Done}
 
@@
 
-\begin{Coqsrc}
-Definition exp2 n := iterate (fun n => 2 * n) n 1.
-\end{Coqsrc}
-Using some facts (proven in 
+\vspace{4pt}
+
+\input{movies/snippets/Exp2/exp2Def}
+
+\vspace{4pt}
+
+Using a few facts (proven in 
 \href{../theories/html/hydras.Hydra.BigBattle.html}{hydras.Hydra.BigBattle}),we get several  minorations.
 
-\begin{Coqsrc}
-Lemma minoration_0 : forall n,  2 * n <= doubleS n.
-
-Lemma minoration_1 : forall n x, exp2 n * x <= iterate doubleS n x.
-
-Lemma minoration_2 : exp2 95 * 95 <= M.
-
-Lemma minoration_3 : exp2 (S M) * S M <= N.
-
-Lemma minoration : exp2 (exp2 95 * 95) <= N.
-\end{Coqsrc}
+\input{movies/snippets/BigBattle/minorationLemmas}
+
+\vspace{4pt}
 

### 2021-08-02 doc/part-hydras.tex
chapter 2 (with alectryon) completed
Note: there remain a few Coqnswer blocks (due to a bad display of sub-goals)

@@
 
-\begin{Coqsrc}
-Definition Alive (B : Battle) :=
-  forall i h, 
-     h <> head -> {h' : Hydra |  B i h h'}.
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Definitions/AliveDef}
 
@@
 
-\begin{Coqsrc}
-Theorem Alive_free: Alive free.
-
-Theorem Alive_standard: Alive standard.  
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Theorems/AliveThms}
 
@@
 
-\begin{Coqsrc}
-Definition  next_round_dec n :
- forall h, (h = head) + {h' : Hydra & {R1 h h'} + {R2 n h  h'}}.
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Lemmas/nextRoundDec}
 
@@
 
-The termination of all battles is naturally expressed by the predicate \texttt{well\_founded} defined in the module \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf} 
+The termination of \emph{any}  battle is naturally expressed by the predicate \texttt{well\_founded} defined in the module \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}
  of the Standard Library.
@@
 
-\begin{Coqsrc}
-Definition Termination :=  well_founded (transp _ round).
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Definitions/TerminationDef}
 
@@
 \index{hydras}{Library Hydra!Type classes!Hvariant}
-\begin{Coqsrc}
-Class Hvariant {A:Type}{Lt:relation A}(Wf: well_founded Lt)(B : Battle)
-  (m: Hydra -> A):   Prop :=
-  {variant_decr: forall i h h',
-      h <> head ->
-      battle_rel  B i  h h' -> Lt (m h') (m h)}.
-\end{Coqsrc}
+
+\input{movies/snippets/Hydra_Definitions/HvariantDef}
 
@@
 \begin{exercise}
- Prove that, if there is an instance of (\texttt{Hvariant Lt wf\_Lt $B$ $m$}), then there exists no infinite battle in  $B$.
+ Prove that, if there exists some  instance of (\texttt{Hvariant Lt wf\_Lt $B$ $m$}), then there exists no infinite battle in  $B$.
 \end{exercise}
@@
 
-\begin{Coqsrc}
-Section Impossibility_Proof.
- Variable m : Hydra -> nat.
- Hypothesis Hvar : Hvariant lt_wf free m.
-\end{Coqsrc}
+\input{movies/snippets/Omega_Small/omegaSmalla}
 
@@
 
-  \begin{Coqsrc}
-  Let iota (i: nat) := hyd_mult head (S i).    
-  \end{Coqsrc}
+\input{movies/snippets/Omega_Small/iotaDef}
 
@@
 
- \begin{Coqsrc}
-  Let big_h := hyd1 (hyd1 head).
- \end{Coqsrc}
+\input{movies/snippets/Omega_Small/bigHDef}
 
@@
 
-  \begin{Coqsrc}
- Let small_h := iota (m big_h).
-
- Fact big_to_small : big_h -1-> small_h.
- Proof.
-      exists (m big_h); right; repeat constructor.     
- Qed.
-    \end{Coqsrc}
+\input{movies/snippets/Omega_Small/smallHDef}
 
@@
 
- \begin{Coqsrc}
-Lemma m_lt : m small_h < m big_h.
- \end{Coqsrc}
+\vspace{4pt}
+
+\input{movies/snippets/Omega_Small/mLt}
 
@@
 
-\begin{Coqsrc}
-Lemma m_ge : forall i:nat, i <= m (iota i).
-\end{Coqsrc}
+\input{movies/snippets/Omega_Small/mGea}
+
+\begin{Coqanswer}
+1 subgoal (ID 23)
+  m : Hydra -> nat
+  Hvar : Hvariant lt_wf free m
+
+  small_h := iota (m big_h) : Hydra
+  ============================
+  forall i : nat, i <= m (iota i)
+\end{Coqanswer}
 
@@
 
- Then our proof is almost finished.
-
-   \begin{Coqsrc}
-Theorem Contradiction : False.
-Proof.
- apply (Nat.lt_irrefl (m big_h));
-   apply  Lt.le_lt_trans with (m small_h).
-  - apply m_ge.
-  - apply m_lt.
-Qed.
-
-End Impossibility_Proof.
-\end{Coqsrc}
+We are now ready to complete our impossibility proof.
+
+\vspace{4pt}
+
+\input{movies/snippets/Omega_Small/mGeb}
+
+\input{movies/snippets/Omega_Small/omegaSmallz}
 

### 2021-08-03 doc/part-hydras.tex
converting chapter 3 to alectryon (continued)

@@
   | CompLt : Plt -> CompareSpec Peq Plt Pgt Lt
-  | CompGt : Pgt -> CompareSpec Peq Plt Pgt Gt
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Class Comparable {A:Type} 
-  (lt: relation A)
-  (compare : A -> A -> comparison) :=
-  {
-  sto :> StrictOrder lt;
-  compare_correct: forall a b,
-      CompareSpec (a = b) (lt a b) (lt b a) (compare a b);
-  }.
-\end{Coqsrc}
+  | CompGt : Pgt -> CompareSpec Peq Plt Pgt Gt.
+\end{Coqsrc}
+
+\emph{From Module~\href{../theories/html/hydras.Prelude/mparable.html\#Hvariant}{Prelude.Comparable}}
+
+\input{movies/snippets/Comparable/ComparableDef}
 
@@
 
-\begin{Coqsrc}
-Class ON {A:Type}(lt: relation A)
-      (compare : A -> A -> comparison)  :=
-  {
-  comp :> Comparable lt compare;
-  wf : well_founded lt;
-  }.
-\end{Coqsrc}
+\input{movies/snippets/ON_Generic/ONDef}
 
@@
 \label{sect:measure-ON}
-\begin{Coqsrc}
-Section Definitions.
-
-  Context  {A:Type}{lt : relation A}
-           {compare : A -> A -> comparison}
-           {on : ON lt compare}.
-
-  #[using="All"]
-   Definition ON_t := A.
-
-  #[using="All"]
-   Definition ON_compare := compare.
-
-  #[using="All"]
-   Definition ON_lt := lt.
-
-  #[using="All"]
-   Definition ON_le:  relation A := leq lt.
-
-  #[using="All"]
-   Definition measure_lt {B : Type} (m : B -> A) : relation B :=
-    fun x y =>  ON_lt (m x) (m y).
-
-  #[using="All"]
-   Lemma wf_measure {B : Type} (m : B -> A) :
-    well_founded (measure_lt m).
-
-  #[using="All"]
-   Definition ZeroLimitSucc_dec :=
-    forall alpha,
-      {Least alpha} +
-      {Limit alpha} +
-      {beta: A | Successor alpha beta}.
-
-  (** The segment called [O alpha] in Schutte's book *)
-
-  #[using="All"]
-   Definition bigO 
-   (a: A) : Ensemble A := fun x: A => lt x a.
-
-End Definitions.
-
-Infix "o<" := ON_lt : ON_scope.
-Infix "o<=" := ON_le : ON_scope.
-Infix "o?=" := ON_compare (at level 70) : ON_scope.
-\end{Coqsrc}
+
+\input{movies/snippets/ON_Generic/ONDefs}
 
@@
 
-\begin{Coqsrc}
-#[global]
- Instance Omega_comp : Comparable Peano.lt  Nat.compare.
-Proof.
-  split.
-  - apply Nat.lt_strorder.
-  - apply Nat.compare_spec.
-Qed.
-
-#[global]
- Instance Omega : ON  Peano.lt  Nat.compare.
-Proof.
- split.
- - apply Omega_comp.
- - apply Wf_nat.lt_wf.
-Qed.
-
-#[local] Open Scope ON_scope.
-Compute 6 o?= 9.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-     = Lt
-     : comparison
-\end{Coqanswer}
+\input{movies/snippets/ON_Omega/OmegaDef}
 
@@
 
-\begin{Coqsrc}
-Section Defs.
-
-  Context `(ltA: relation A)
-          (compareA : A -> A -> comparison)
-          (NA: ON ltA compareA).
-  Context `(ltB: relation B)
-          (compareB : B -> B -> comparison)
-          (NB: ON ltB compareB).
-
-Definition t := (A + B)%type.
-Arguments inl  {A B} _.
-Arguments inr  {A B} _.
-
-Definition lt : relation t := le_AsB _ _ ltA ltB.
-\end{Coqsrc}
-
-In order to build an instance of \texttt{ON}, we have to define a comparison function.
-
-\begin{Coqsrc}
-Definition compare (alpha beta: t) : comparison :=
-   match alpha, beta with
-     inl _, inr _ => Lt
-   | inl a, inl a' => compareA a a'
-   | inr b, inr b' => compareB b b'
-   | inr _, inl _ => Gt
-  end.
-
-Lemma compare_correct alpha beta :
-    CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
-    (compare alpha beta).
-
-#[global] Instance plus_comp : Comparable lt compare.
-Proof.
-  split.
-  - apply lt_strorder.
-  - apply compare_correct. 
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/ON_plus/Defs}
+
+In order to build an instance of \texttt{Comparable}, we have to define a correct comparison function.
+
+\input{movies/snippets/ON_plus/compareDef}
+
+\vspace{4pt}
+
+\input{movies/snippets/ON_plus/compareCorrect}
+
+\vspace{4pt}
+
+\input{movies/snippets/ON_plus/plusComp}
+
+\vspace{4pt}
 
@@
 
-\begin{Coqsrc}
-Lemma lt_wf : well_founded lt.
-Proof. 
-  destruct NA, NB; apply wf_disjoint_sum; [apply wf | apply wf0].
-Qed.
-\end{Coqsrc}
-
-Then, we register an instance of \texttt{ON}:
-
-\begin{Coqsrc}
-#[global] Instance ON_plus : ON lt compare.
-Proof.
-  split.
-  - apply plus_comp.
-  - apply lt_wf.
-Qed.
-\end{Coqsrc}
+\vspace{4pt}
+
+\input{movies/snippets/ON_plus/ltWf}
+
+\vspace{4pt}
+
+Then, we can build an instance of \texttt{ON}:
+
+\vspace{4pt}
+
+\input{movies/snippets/ON_plus/OnPlus}
 
@@
 of two copies of $\omega$ (Figure~\ref{fig:omega-plus-omega}).
+It is also represented by the two first lines of Figure~\ref{fig:ordinal-sequence}.
 
@@
 
-\begin{Coqsrc}
-Instance Omega_plus_Omega : ON _ _ := ON_plus Omega Omega.
-
-Definition t := ON_t.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Example ex1 : inl 7 o< inr 0.
-Proof. constructor. Qed.
-\end{Coqsrc}
+\input{movies/snippets/ON_Omega_plus_omega/OmegaPlusOmegaDef}
+
+\vspace{4pt}
 
@@
 
-\begin{Coqsrc}
-Definition fin (i:nat) : t := inl i.
-Coercion fin : nat >-> t.
-
-Notation omega := (inr  0:t).
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Compute  fin 8 o?= omega.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-     = Lt
-     : comparison
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Lemma lt_omega alpha : 
-     alpha o< omega <-> exists n:nat,  alpha = fin n.
-(* ... *)
-\end{Coqsrc}
+\vspace{4pt}
+
+\input{movies/snippets/ON_Omega_plus_omega/finiteOmega}
+
+\vspace{4pt}
+
+\input{movies/snippets/ON_Omega_plus_omega/ltOmega}
 
@@
 \item The least ordinal, \texttt{(inl 0)}, also known as  \texttt{(fin 0)}.
-\item The limit ordinal $\omega$.
-\item The successor ordinals, either of the form \texttt{(inl (S $i$))} or \texttt{(inr (S $i$))}
+\item The first infinite ordinal $\omega$.
+\item The remaining ordinals, either of the form \texttt{(inl (S $i$))} or \texttt{(inr (S $i$))} (in black on Figure~\ref{fig:ordinal-sequence}), called \emph{successor ordinals}.
 \end{itemize}
@@
 
-\begin{Coqsrc}
-Section A_given.
-  Variables (A : Type)  (lt: relation A).
-
-  #[local] Infix "<" := lt.
-  #[local] Infix "<=" := (leq lt).
-
- Definition Least {sto : StrictOrder lt} (x : A):=
-  forall y,  x <= y.
-
- Definition Successor {sto : StrictOrder lt} (y x : A):=
-  x < y /\ (forall z,  x < z ->  z <  y -> False).
-
- Definition Limit {sto : StrictOrder lt}  (x:A)  :=
-   (exists w:A,  w < x) /\
-   (forall y:A, y < x -> exists z:A, y < z /\ z < x).
-\end{Coqsrc}
+\input{movies/snippets/MoreOrders/Defs}
 

### 2021-08-04 doc/part-hydras.tex
converting chapter 3 to alectryon (continued)

@@
 
-\begin{Coqsrc}
-Lemma limit_iff (alpha : t) : Limit alpha <-> alpha = omega.
-\end{Coqsrc}
-
-Regarding successors, let us define the following function:
-
-\begin{Coqsrc}
-Definition succ (alpha : t) :=
-  match alpha with
-    inl n => inl (S n)
-  | inr n => inr (S n)
-  end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma Successor_correct alpha beta : Successor beta alpha <->
-                                     beta = succ alpha.
-\end{Coqsrc}
-
-We can also check whether an ordinal is a successor by a simple pattern matching:
-
-\begin{Coqsrc}
-  Definition succb (alpha: t) : bool :=
-  match alpha with
-  | inr (S  _) | inl (S _) => true
-  | _ => false
-  end.
-
-Lemma succb_correct (alpha : t) :
-    succb alpha <->  exists beta: t, alpha = succ beta.
-\end{Coqsrc}
+\input{movies/snippets/ON_Omega_plus_omega/limitIff}
+
+\vspace{4pt}
+
+Regarding successors, let us define the following function and prove its correctness:
+
+\input{movies/snippets/ON_Omega_plus_omega/succDef}
+
+\input{movies/snippets/ON_Omega_plus_omega/succCorrect}
+
+\vspace{4pt}
+
+We can also check whether an ordinal is a successor by a simple computation:
+
+\input{movies/snippets/ON_Omega_plus_omega/succb}
+
+\vspace{4pt}
 
@@
 
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{ON\_Generic}}
-\begin{Coqsrc}
-  Definition ZeroLimitSucc_dec {A:Type}{lt: relation A}
-           {compare : A -> A -> comparison}
-           {on : ON lt compare} :=
-  forall alpha,
-    {Least alpha} +
-    {Limit alpha} +
-    {beta: A | Successor alpha beta}.
-(* ... *)
-\end{Coqsrc}
-
 \noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
 
-\begin{Coqsrc}
-Definition Zero_limit_succ_dec : ZeroLimitSucc_dec.
-Proof.
-(* ... *)
-Defined.
-\end{Coqsrc}
+\input{movies/snippets/ON_Omega_plus_omega/ZeroLimitSuccDec}
 
@@
 
+In \coq{}, we build an instance of class \texttt{ON} through a sequence of steps as for the sum of ordinal notations.
+
 \vspace{4pt}
@@
 
-\begin{Coqsrc}
-Section Defs.
-
-  Context `(ltA: relation A)
-          (compareA : A -> A -> comparison)
-          (NA: ON ltA compareA).
-  Context `(ltB: relation B)
-          (compareB : B -> B -> comparison)
-          (NB: ON ltB compareB).
-
-Definition t := (B * A)%type.
-Definition lt : relation t := lexico ltB ltA.
-
-Definition compare (alpha beta: t) : comparison :=
-  match compareB (fst alpha) (fst beta) with
-  | Eq => compareA (snd alpha) (snd beta)
-  | c => c
-  end.
-
-#[global] Instance ON_mult: ON lt compare.
-
-End Defs.
-\end{Coqsrc}
+\input{movies/snippets/ON_mult/Defs}
 
@@
 
-\begin{Coqsrc}
-Instance Omega2: ON _ _ := ON_mult Omega Omega.
-
-Definition t := ON_mult.t nat nat.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Definition zero: t := (0,0).
-
-Definition fin (n:nat) : t := (0, n).
-
-Coercion fin : nat >-> t.
-
-Notation omega := (1,0).
-
-Definition limitb (alpha : t) : bool := 
- match alpha with
- |  (S _, 0) => true
- | _ => false
- end.
-
- Definition succb (alpha: t) : bool :=
- match alpha with
-  (_, S _) => true
-  | _ => false
-end.
-\end{Coqsrc}
+\input{movies/snippets/ON_Omega2/Omega2Def}
+
+\input{movies/snippets/ON_Omega2/Defs}
 
@@
 
-\begin{Coqsrc}
-Definition succ (alpha : t) := (fst alpha, S (snd alpha)).
-\end{Coqsrc}
+\input{movies/snippets/ON_Omega2/Defs}
 
@@
 
-\begin{Coqsrc}
-Lemma succ_ok alpha beta : Successor beta alpha <-> beta = succ alpha.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma lt_succ_le alpha beta : alpha o< beta <-> succ alpha o<= beta.
-
-Lemma lt_succ alpha : alpha o< succ alpha.
-
-\end{Coqsrc}
+\input{movies/snippets/ON_Omega2/succOK}
+
+\input{movies/snippets/ON_Omega2/succLemmas}
 

### 2021-08-04 doc/part-hydras.tex
converting chapter 3 to alectryon (continued)

@@
 
-We can define on \texttt{Omega2} an addition which extends the addition on \texttt{nat}.
-
-\begin{Coqsrc}
-Definition  plus (alpha beta : t) : t :=
-  match alpha,beta with
-  | (0, b), (0, b') => (0, b + b')
-  | (0,0), y  => y
-  | x, (0,0)  => x
-  | (0, b), (S n', b') => (S n', b')
-  | (S n, b), (S n', b') => (S n + S n', b')
-  | (S n, b), (0, b') => (S n, b + b')
-   end.
-
-Infix "+" := plus : o2_scope.
-\end{Coqsrc}
-
-Please note that this operation is not commutative:
-
-\begin{Coqsrc}
-Example non_commutativity_of_plus :  omega + 3 <> 3 + omega.
-Proof.
-  cbn.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-1 subgoal (ID 237)
-
-  ============================
-(1, 3) <> omega.
-\end{Coqanswer}
-
-\begin{Coqsrc}
- discriminate.
-Qed.
-\end{Coqsrc}
+We can define on \texttt{Omega2} an addition which extends the addition on \texttt{nat}. Please note that this operation is not commutative:
+
+\input{movies/snippets/ON_Omega2/plusDef}
 

### 2021-08-05 doc/part-hydras.tex
converting chapter 3 to alectryon (continued)

@@
 
-\begin{Coqsrc}
-(** multiplication of an ordinal by a natural number *)
-
-Definition mult_fin_r  (alpha : t) (p : nat): t :=
-  match alpha, p with
- |  (0,0), _  => zero
- |  _, 0 => zero
- |  (0, n), p => (0, n * p)
- |  ( n, b),  n' => ( n *  n', b)
- end.
-Infix "*" := mult_fin_r : o2_scope.
-
-(** multiplication of  a natural number by an ordinal *)
-
-Definition mult_fin_l (n:nat)(alpha : t) : t :=
-  match n, alpha with
- |  0, _  => zero
- |  _, (0,0) => zero
- |   n , (0,n') => (0, (n*n')%nat)
- |  n, (n',p') => (n', (n * p')%nat)
- end.
-
-Example e1 : (omega * 7 + 15) * 3 = omega * 21 + 15.
-Proof. reflexivity. Qed.
-
-Example e2 :  mult_fin_l 3 (omega * 7 + 15) = omega * 7 + 45.
-Proof. reflexivity. Qed.
-\end{Coqsrc}
+\input{movies/snippets/ON_Omega2/multFinDef}
+
+\vspace{4pt}
 
@@
 
-\begin{Coqsrc}
-Lemma unique_decomposition alpha : 
-    exists! i j: nat,  alpha = omega * i + j.
-\end{Coqsrc}
+\input{movies/snippets/ON_Omega2/uniqueDecomposition}
 
@@
 
-\begin{Coqsrc}
-Section Merge.
-
-  Variable A: Type.
-
-  Local Definition m (p : list A * list A) :=
-    omega * length (fst p) + length (snd p).
-
-  Function  merge  (ltb: A -> A -> bool)
-          (xys: list A * list A)
-          {wf (measure_lt m) xys} :
-    list A :=
-    match xys with
-      (nil, ys) => ys
-    | (xs, nil) => xs
-    | (x :: xs, y :: ys) =>
-      if ltb x y then x :: merge  ltb (xs, (y :: ys))
-      else y :: merge  ltb ((x :: xs), ys)
-    end.
-
-  - intros; unfold m, measure_lt; cbn; destruct xs0; simpl; left; lia.
-  - intros; unfold m, measure_lt; cbn; destruct ys0; simpl; right; lia.
-  - apply wf_measure.
-  Defined.
-
-End Merge.
-
-Goal forall l, merge nat Nat.leb (nil, l) = l.
-  intro; now rewrite merge_equation.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/ON_Omega2/Merge}
 
@@
 
-\begin{Coqsrc}
-Section Impossibility_Proof.
-
- Variable m : Hydra -> ON_Omega2.t.
-
- Context (Hvar : Hvariant (ON_Generic.wf (ON:=Omega2))  free m).
-\end{Coqsrc}
+\input{movies/snippets/Omega2_Small/Impossibility}
+\input{movies/snippets/Omega2_Small/Impossibilitya}
 
@@
 
-\begin{Coqsrc}
- Let iota (p: ON_Omega2.t) := 
-     node (hcons_mult (hyd1 head) (fst p)
-                    (hcons_mult head (snd p) hnil)).
-\end{Coqsrc}
+\input{movies/snippets/Omega2_Small/Impossibilityc}
 
@@
 
-\begin{Coqsrc}
-   Let big_h := hyd1 (hyd2 head head).  
-\end{Coqsrc}
+\input{movies/snippets/Omega2_Small/Impossibilityb}
 
@@
 
-\begin{Coqsrc}
-   Let small_h := iota (m big_h).
-\end{Coqsrc}
+\input{movies/snippets/Omega2_Small/Impossibilityd}
+
+\vspace{4pt}
 
@@
 
-\begin{Coqsrc}
-  Lemma big_to_small: big_h -+-> small_h.
-\end{Coqsrc}
+\input{movies/snippets/Omega2_Small/bigToSmall}
 
@@
 
-\begin{Coqsrc}
-   Corollary m_lt : m small_h o< m big_h.
-\end{Coqsrc}
+\input{movies/snippets/Omega2_Small/mLt}
 
@@
 
-\begin{itemize}
-\item The proof of the lemma \texttt{m\_ge : m big\_h  o<= m small\_h} uses well-founded induction on \texttt{big\_h}.
-
-\item For any pair $p$, we have to distinguish between three cases, according to the value of $p$'s components.
-  \begin{itemize}
-  \item $p=(0,0)$
-  \item $p=(i,0)$, where $i>0$\,: $p$ corresponds to a limit ordinal
-  \item $p=(i,j)$, where $j>0$\,: $p$ is the successor of $(i,j-1)$.
-  \end{itemize}
-\end{itemize}
-
-Let us define the notion of elementary ``step'' of decreasing sequences in
-\texttt{t}
-
-\begin{Coqsrc}
-  Inductive step : t -> t -> Prop :=
-  | succ_step : forall i j,  step (i, S j) (i, j)
-  | limit_step : forall i j, step (S i, 0) (i, j).
-\end{Coqsrc}
-
-The following lemma establishes a correspondance between the relation
-\texttt{step} and hydra battles.
-
-\begin{Coqsrc}
-  Lemma step_to_battle : forall p q, step p q -> iota p -+-> iota q.
-\end{Coqsrc}
-
-\index{maths}{Transfinite induction}
-
-Thus, starting from any inequality $q < p$ on type \texttt{t}, we can build 
-by \emph{transfinite induction} (\emph{i.e.} well-founded) over \texttt{p} a battle 
-that transforms the hydra $\iota(p)$ into $\iota(q)$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}
-
-\begin{Coqsrc}
-  Lemma m_ge : forall p : t,   p o<= m (iota p).
-  Proof.
-    unfold small_h; pattern (m big_h) .   
-     apply  well_founded_induction with (R := lt) (1:= lt_wf).
-     intro p ; pattern p;
-     apply  well_founded_induction with 
-               (R := lt2) (1:= wf_lexico lt_wf lt_wf);
-     intros (i,j) IHij. 
-\end{Coqsrc}
+The first steps of our proof prepare a well-founded indsction on $\omega^2$.
+
+\input{movies/snippets/Omega2_Small/mGe}
 
@@
   (i, j) o<= m (iota (i, j))
+
 \end{Coqanswer}
 
-Then we have  three cases to consider, according to the value of $p$.
+Then a case analysis on $i$ and $j$ allows us to
+consider three cases :
+
 \begin{itemize}
-\item If $p=(0,0)$ then obviously, $\iota(p)\geq p = (0,0)$
-\item If  $p=(i+1,0)$ for some $i\in\mathbb{N}$, we
- remark  that $p$ is strictly greater than any pair $ (i, j)$, where $j$ 
-is any natural number.
-
-Applying the battle rules, for any $j$, we have $\iota(i+1,j)  {\round} \iota(i, j) $, thus $m(\iota(p)) > m(\iota(i,j)$ since  $m$ is assumed to be a variant.
-
-Applying the induction hypothesis, we get the inequality
- $ m(\iota(i,j)) \geq (i,j)$ for any $j$.
-
-Thus, $m(\iota(p)) > (i,j)$ for any $j$.
-Applying the lemma \texttt{limit\_is\_lub}, we get  the inequality
-$m(\iota(i+1,0))\geq (i+1,0)$
-
-\item If $p=(i,j+1)$ with $j\in\mathbb{N}$, we have  $\iota(p)  {\round} \iota(i, j) $,
-hence $m(\iota(p))> m(\iota(i,j)) \geq (i,j)$, thus $m(\iota(p))\geq (i,j+1)=p$
-
+\item $i=j=0$: the inequality is trivial.
+\item $i=1+l, j=0$ ($(i,j)$ is a limit ordinal): By the induction hypothesis \texttt{IHij},
+  $(l,k)\leq m(\iota(l,k))$ for any $k$. But (by the rules of the hydra game), $\iota(i,0)$ is transformed into any $\iota(l,k)$ in one round. Thus $m(\iota(l,k)) < m(\iota(i,0))$ for any $k$.
+  Therefore, $(l,k) <  m(\iota(i,0))$ for any $k$, thus
+  $(i,0) \leq m(\iota(i,0))$.
+ \item $j= l+1$  ($(i,j)$ is a successor).  a similar, but simpler case: we apply the induction hypothesis to the pair $(i,l)$.
  \end{itemize}
 
-\begin{Coqsrc}
-  (* ... *)
-Qed.
-\end{Coqsrc}
+ Please look at the proof script for more details.
+
+ \input{movies/snippets/Omega2_Small/mGeb}
 
@@
 \noindent
-In \coq (Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#Impossible}{Hydra.Omega2\_Small}):
-
-\begin{Coqsrc}
-  Theorem Impossible: False.
-  Proof.
-    destruct (StrictOrder_Irreflexive (m big_h)).
-    apply le2_lt2_trans with (m small_h).
-    -  unfold small_h; apply m_ge.
-    -  apply m_lt. 
-Qed. 
-End Impossibility_Proof.
-\end{Coqsrc}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#Impossible}{Hydra.Omega2\_Small}}
+
+  \input{movies/snippets/Omega2_Small/Impossible}
 

### 2021-08-06 doc/part-hydras.tex
converting chapter 3 to alectryon (continued)
converting chapter 3 to alectryon (draft, completed)

@@
 \label{sect:measure-ON}
+\label{sect:bigO-ON}
 
@@
 \label{def: Finite-ord-type}
-\begin{Coqsrc}
-Coercion is_true: bool >-> Sortclass.
-
-Definition t (n:nat) := {i:nat | Nat.ltb i n}.
-\end{Coqsrc}
-
-The order on type (\texttt{t $n$}) is defined through the projection on \texttt{nat}.
-
-\begin{Coqsrc}
-Definition lt {n:nat} : relation (t n) :=
-  fun alpha beta => Nat.ltb ( proj1_sig alpha) (proj1_sig beta).
-\end{Coqsrc}
+
+\input{movies/snippets/ON_Finite/Defs}
 
@@
 
-\begin{Coqsrc}
-Program Example alpha1 : t 7 := 2.
-
-Program Example beta1 : t 7 := 5.
-
-Example i1 : lt  alpha1 beta1.
-Proof. now compute. Qed.
-\end{Coqsrc}
+\input{movies/snippets/ON_Finite/Example1}
 
@@
 
-\begin{Coqsrc}
-Lemma t0_empty (alpha: t 0): False.
-Proof.
-  destruct alpha; discriminate.
-Qed.
-
+ \input{movies/snippets/ON_Finite/t0Empty}
+
+\begin{Coqsrc}
 Program Definition bad : t 10 := 10.
 Next Obligation.
+\end{Coqsrc}
+
+\begin{Coqanswer}
+1 subgoal (ID 118)
+
+  ============================
+  10 <? 10 
+\end{Coqanswer}
+
+\begin{Coqsrc}
   compute.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-1 subgoal (ID 162)
-
-  ============================
-  false = true
-\end{Coqanswer}
-
-\begin{Coqsrc}
 Abort.
@@
 
-Note also that attempting to compare a term  of type (\texttt{t $n$}) with a term of
-type (\texttt{t $p$})  leads to an error if $n$ and $p$ are not convertible.
-
-\begin{Coqsrc}
-Program Example gamma1 : t 8 := 7.
-
-Fail Goal lt alpha1 gamma1.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- The command has indeed failed with message:
-The term "gamma1" has type "t 8" while it is expected to have type "t 7".
-\end{Coqanswer}
+%\input{movies/snippets/ON_Finite/bad}
 
@@
 
-\begin{Coqsrc}
-Definition compare {n:nat} (alpha beta : t n) :=
-  Nat.compare (proj1_sig alpha) (proj1_sig beta).
-
-Lemma compare_correct {n} (alpha beta : t n) :
-  CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
-              (compare alpha beta).
-\end{Coqsrc}
+\input{movies/snippets/ON_Finite/compareDef}
 
@@
 
+Please note  that attempting to compare a term  of type (\texttt{t $n$}) with a term of
+type (\texttt{t $p$})  leads to an error if $n$ and $p$ are not convertible.
+
+\input{movies/snippets/ON_Finite/Example2}
+
 Applying lemmas of the libraries \texttt{Coq.Wellfounded.Inverse\_Image}, \linebreak
@@
 
-\begin{Coqsrc}
-Lemma lt_wf (n:nat) : well_founded (@lt n).
-\end{Coqsrc}
+\input{movies/snippets/ON_Finite/ltWf}
 
@@
 
-\begin{Coqsrc}
-#[global] Instance sto n : StrictOrder (@lt n).
-
-#[global] Instance comp n: Comparable (@lt n) compare.
-
-#[global] Instance FinOrd n : ON (@lt n) compare. 
-Proof.
-  split.
-  - exact (comp n).
-  - apply lt_wf.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/ON_Finite/ONInstance}
 
@@
 In set-theoretic presentations of ordinals, the set associated with the ordinal $n$ is $\{0,1,\dots,n-1\}$.
-In our formalization, the interpretation of an ordinal as a set is realized by the following definition
-(in ~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{ON\_Generic}).
-
-\begin{Coqsrc}
-Definition bigO `{nA : @OrdinalNotation A ltA stoA compareA}
-           (a: A) : Ensemble A :=
-  fun x: A => ltA x a.
-\end{Coqsrc}
+In our formalization, the interpretation of an ordinal as a set is realized by the function \texttt{bigO} (see Section\vref{sect:bigO-ON}).
 \end{remark}
@@
 
-\begin{Coqsrc}
-Class  SubON 
-       `(OA : @ON A ltA  compareA)
-       `(OB : @ON B ltB  compareB)
-       (alpha :  B)
-       (iota : A -> B):=
-  {
-  SubON_compare: forall x y : A,  compareB (iota x) (iota y) =
-                                 compareA x y;
-  SubON_incl : forall x, ltB (iota x) alpha;
-  SubON_onto : forall y, ltB y alpha  -> exists x:A, iota x = y}.
-\end{Coqsrc}
+following definition
+(in ~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{ON\_Generic}).
+
+\input{movies/snippets/ON_Generic/SubONDef}
 
@@
 
-\begin{Coqsrc}
-Instance Incl : SubON Omega Omega_plus_Omega omega fin.
-\end{Coqsrc}
+\vspace{4pt}
+\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
+
+\input{movies/snippets/ON_Omega_plus_omega/Incl}
 
@@
 
-We are now able to build an instance of \texttt{SubON}.
-
 \vspace{4pt}
@@
 
-\begin{Coqsrc}
-Section Inclusion_ij.
-
-  Variables i j : nat.
-  Hypothesis Hij : (i < j)%nat.
-
-  Remark Ltb_ij : Nat.ltb i j.
-
-  Program Definition iota_ij  (alpha: t i) : t j :=  alpha.
-
-   Let b : t j := exist _ i Ltb_ij.
-
-   Global Instance F_incl_ij  : SubON  (FinOrd i) (FinOrd j) b iota_ij.
-  (* ... *)
-
-  End Inclusion_ij.
-\end{Coqsrc}
+\input{movies/snippets/ON_Finite/InclIJ}
+\input{movies/snippets/ON_Finite/InclIJa}
+\input{movies/snippets/ON_Finite/InclIJb}
+\input{movies/snippets/ON_Finite/InclIJc}
+\input{movies/snippets/ON_Finite/InclIJd}
 
@@
 
-\begin{Coqsrc}
-Class ON_correct `(alpha : Ord)
-     `(OA : @ON A ltA compareA)
-      (iota : A -> Ord) :=
-  { ON_correct_inj : forall a, lt (iota a) alpha;
-    ON_correct_onto : forall beta, lt beta alpha ->
-                                exists b, iota b = beta;
-    On_compare_spec : forall a b:A,
-        match compareA a b with
-          Datatypes.Lt => lt (iota a) (iota b)
-        | Datatypes.Eq => iota a = iota b
-        | Datatypes.Gt => lt (iota b) (iota a)
-        end
-  }.
-\end{Coqsrc}
+\input{movies/snippets/ON_Generic/ONCorrect}
 
@@
 
-\begin{Coqsrc}
-Instance Epsilon0_correct :
-  ON_correct epsilon0 Epsilon0  (fun alpha => inject (cnf alpha)).
-\end{Coqsrc}
+\vspace{4pt}
+\noindent\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
+
+\input{movies/snippets/Correctness_E0/Epsilon0Correct}
 
@@
 \label{types:ON-iso}
-\begin{Coqsrc}
-Class  ON_Iso 
-       `(OA : @ON A ltA compareA)
-       `(OB : @ON B ltB  compareB)
-       (f : A -> B)
-       (g : B -> A):=
-  {
-  iso_compare: forall x y : A, 
-      compareB (f x) (f y) = compareA x y;
-  iso_inv1 : forall a, g (f a)= a;
-  iso_inv2 : forall b, f (g b) = b
-}.
-\end{Coqsrc}
+
+\input{movies/snippets/ON_Generic/ONIso}
 

### 2021-08-07 doc/chapter-powers.tex
resolve mistakes

@@
 \emph{From Module
-\href{../theories/html/FirstSteps.html}{additions.FirstSteps}}
+\href{../theories/html/additions.FirstSteps.html}{additions.FirstSteps}}
 
@@
 In contrast, the optimizations we present in this document apply, in their vast majority, \emph{generic} techniques of efficient computation of powers in a monoid. 
-This example of Fibonacci numbers has been developed with Yves Bertot, who wrote a first version with \texttt{SSreflect/Mathcomp}~\cite{MCB, SSR}.
+This example of Fibonacci numbers has been developed with Yves Bertot, who wrote a first version with \texttt{SSreflect/Mathcomp}~\cite{MCB}.
 
@@
 that the two representations of natural numbers and the function \texttt{N.to\_nat}
-form a kind of  \emph{data refinement} \cite{Abrial:2010:MES:1855020, cohen:hal-01113453}.
+form a kind of  \emph{data refinement} \cite{Abrial:2010:MES:1855020, Cohen2013}.
 

### 2021-08-07 doc/part-hydras.tex
remove duplicate references
converting chapter 3 to alectryon (draft, completed)

@@
 
-In order to build an instance of \texttt{ON}, we define a comparison function,  and prove its correction.
+In order to build an instance of \texttt{ON}, we define a comparison function,  and prove its correctiness.
+
+\vspace{4pt}
 
@@
 \begin{remark}
-Finite ordinals are also formalized in MathComp~\cite{SSR}.  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.  
+Finite ordinals are also formalized in MathComp~\cite{MCB}.  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.  
 \end{remark}

### 2021-08-08 doc/part-hydras.tex
converting chapter 4 to Alectryon (start)

@@
 
-\begin{Coqsrc}
-Inductive T1 : Set  :=
-| zero : T1
-| ocons : T1 -> nat -> T1 -> T1.
-\end{Coqsrc}
+\input{movies/snippets/T1/T1Def}
 
@@
 \noindent
-\begin{Coqsrc}
-Inductive T2 : Set :=
-  zero : T2
-| gcons : T2 -> T2 -> nat -> T2 -> T2.
-\end{Coqsrc}
+
+\input{movies/snippets/T2/T2Def}
 
@@
 
-\begin{Coqsrc}
-Example alpha_0 : T1 :=
-  ocons (ocons (ocons zero 0 zero)
-               0
-               zero)
-        0
-       (ocons (ocons zero 2 zero)
-              4
-              (ocons zero 1 zero)).
-\end{Coqsrc}
+\input{movies/snippets/T1/alpha0}
 
@@
 
-\begin{Coqsrc}
-Notation one := (ocons zero 0 zero).
-
-Notation FS n := (ocons zero n zero).
-\end{Coqsrc}
-
 \label{sect:notation-F}
 
-\begin{Coqsrc}
-Definition fin (n:nat) : T1 := 
-    match n with 0 => zero | S p => FS p end.
-
-Coercion fin  : nat >-> T1.
-
-Example ten : T1 := 10.   
-\end{Coqsrc}
+\input{movies/snippets/T1/finiteOrds}
 
@@
 \label{sect:omega-notation2}
-\begin{Coqsrc}
-Notation omega := (ocons (ocons zero 0 zero) 0 zero).
-\end{Coqsrc}
+
+\input{movies/snippets/T1/omegaDef}
 
@@
 
-\begin{Coqsrc}
-Notation "'phi0' alpha" := (ocons alpha 0 zero) (at level 29) : t1_scope.
-\end{Coqsrc}
+\input{movies/snippets/T1/phi0Def}
 
@@
 
-\begin{Coqsrc}
-Fixpoint omega_tower (height:nat) : T1 := 
- match height with 
- | 0 =>  1 
- | S h => phi0 (omega_tower h)
- end.
-\end{Coqsrc}
+\input{movies/snippets/T1/towerDef}
 
@@
 \end{Coqsrc}
+
+% Uncomment for testing 
+%\textbf{Todo: Fix subgoal display }
+%\input{movies/snippets/T1/compareRev}
 

### 2021-08-08 doc/hydras.tex
small note about alectryon output
pass over hydras.tex to fix typos and some grammar

@@
 paper are quite advanced, the rules of hydra battles are very easy to understand.
-There are now several sites on Internet where you can find tutorials on hydra games, together with simulators you can play with.
-See, for instance, the page written by Andrej Bauer~\cite{bauer2008}.
+There are now several sites on the Internet where you can find tutorials on hydra games, together with simulators you can play with.
+See, for instance, the blogpost and source code written by Andrej Bauer~\cite{bauer2008,BauerHydra}.
 
-Hydra battles, as well as Goodstein Sequences~\cite{goodstein_1944, KP82}
+Hydra battles, as well as Goodstein sequences~\cite{goodstein_1944, KP82}
 are a nice way to present complex termination problems.
@@
  The reader should already have some basic experience with the \coq{} system.
-The Reference Manual and several tutorials are available on \coq{} page~\cite{Coq}.
-First chapters of textbooks like \emph{Interactive Theorem Proving and Program Development}~\cite{BC04}, \emph{Software Foundations}~\cite{SF} or  \emph{Certified Programming with Dependent Types} ~\cite{chlipalacpdt2011} will give you the right background.
+The Reference Manual and several tutorials are available on the \coq{} website~\cite{Coq}.
+The first chapters of textbooks like \emph{Interactive Theorem Proving and Program Development}~\cite{BC04}, \emph{Software Foundations}~\cite{SF} or  \emph{Certified Programming with Dependent Types} ~\cite{chlipalacpdt2011} will give you the right background.
 
@@
 
-\subsection{Documenting theories with alectryon}
+\subsection{Documenting theories with Alectryon}
 
@@
 
-Unlike mathematical literature, where definitions and proofs are spread over many articles and books,
-the whole proof is now inside your computer. It is composed of the \texttt{.v} files you downloaded and 
+Unlike mathematical literature, where definitions and proofs are spread out over many articles and books,
+the whole proof is now inside your computer. It is composed from the \texttt{.v} files you downloaded and
 parts of \coq's standard library.
 Thus, there is no ambiguity in our definitions and the premises of the theorems.
-Furthermore, you will be able to navigate through the development, using your favourite text editor or IDE, and some commands like \texttt{Search}, \texttt{Locate},  etc.
+Furthermore, you will be able to navigate through the development, using your favorite text editor or IDE, and some commands like \texttt{Search}, \texttt{Locate},  etc.
 
@@
 
-It may also happen that some proof looks to be useless, because the proven theorem is a trivial consequence of another (proven too) result.
+It may also happen that some proof looks to be useless, because the proven theorem is a trivial consequence of another (also proven) result.
 For instance, let us consider the three following statements:
@@
 
-Except in the \texttt{Schutte} library, dedicated to an axiomatic presentation of the set of countable ordinal numbers, all our development is axiom-free, and respects the rules of intuitionistic logic.
-Note that we also use the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777} in the definitition of  several rapidly growing hierarchy of functions, in Chap.~\ref{chap:alpha-large}.
+Except in the \texttt{Schutte} library, dedicated to an axiomatic presentation of the set of countable ordinal numbers, all of our development is axiom-free, and respects the rules of intuitionistic logic.
+Note that we also use the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777} in the definition of  several rapidly growing hierarchy of functions, in Chap.~\ref{chap:alpha-large}.
 This plug-in imports several known-as-harmless  axioms.
@@
 
+\subsubsection{Verbatim quotations}
+
 Quotations from our  \coq{} source are displayed as follows:
@@
 In general, we do not include full proof scripts in this document.
-The only exceptions are very short proofs (\emph{e.g.} proofs by computation, or by application of automatic tactics).
+The only exceptions are very short proofs (\emph{e.g.}, proofs by computation, or by application of automatic tactics).
 Likewise, we may display only the important steps on a long interactive proof, for instance, in the following lemma (\vref{lemma:L-2_6-1}):
@@
 
-\subsection{Active Links}
+\subsubsection{Using Alectryon}
+
+Alectryon's \emph{LaTeX} mode cannot be mistaken for our hand-made quotations, particularly if you read a color-print.
+Here is an example from Chapter~\ref{chapter:primrec}.
+
+\vspace{4pt}
+
+\input{movies/snippets/Ack/AckFixpointIterate.tex}
+\subsection{Active links}
 The  links which appear in this pdf  document lead are of three possible kinds of destination:
@@
 \item Local links to the document itself,
-\item External links, mainly to \coq's page,
+\item External links, mainly to \coq's website,
 \item Local links to pages generated by \texttt{coqdoc}. According to the current makefile (through the commands \texttt{make html} and \texttt{make pdf}), 
@@
 the relative address \texttt{../theories/html/hydras.XXX.YYY.html} (from the location of the pdf)
-Thus,  active links towards our \coq{} modules may be incorrect if you got this \texttt{pdf} document otherwise than by compiling the distribution available in
+Thus,  active links to our \coq{} modules may be incorrect if you did not get this \texttt{pdf} document by compiling the distribution available at
 \url{https://github.com/coq-community/hydra-battles}.
@@
 \begin{itemize}
-\item The present distribution has been checked with version 8.13.0 of the Coq proof assistant, with the plug-ins \texttt{coq-paramcoq}, \texttt{coq-equations}
+\item The present distribution has been checked with version 8.13.2 of the Coq proof assistant, with the plug-ins \texttt{coq-paramcoq}, \texttt{coq-equations}
 and \texttt{coq-mathcomp-algebra}.
@@
 Although we do not plan to include complete solutions to the exercises, 
-we think it would be useful to include comments and hints, and questions/answers from the users. In constrast, ``projects'' are supposed, once completed, to be included in the repository.
+we think it would be useful to include comments and hints, and questions/answers from the users. In contrast, ``projects'' are supposed, once completed, to be included in the repository.
 
@@
 comment or proposition that would help us. The text contains several \emph{projects} which, when completed, may improve the present work.
-Please do not hesitate to bring your contribution, for instance with Github's proof requests and issues. Thank you in advance!
+Please do not hesitate to share your contributions, for instance using pull requests and issues on GitHub. Thank you in advance!
 
@@
 % \end{todo}
-{\Large \textbf{In progress} This index is currently under reorganization for a few days. We aplologize for its incompleteness! }
+{\Large \textbf{In progress} This index is currently under reorganization for a few days. We apologize for its incompleteness! }
 

### 2021-08-08 doc/part-hydras.tex
fix typos and some notation and grammar in part-hydras.tex

@@
 \begin{itemize}
-\item Representation in \coq{} of hydras and hydra battles
+\item Representation in \coq{} of hydras and hydra battles.
 \item A proof that every battle is finite and won by Hercules.
@@
 
-\item Using a combinatorial toolkit designed by J.~Ketonen and R.~Solovay~\cite{KS81}, we prove that, for any ordinal $\mu<\epsilon_0$, there exists no such variant mapping any hydra to an ordinal stricly less than $\mu$.
+\item Using a combinatorial toolkit designed by J.~Ketonen and R.~Solovay~\cite{KS81}, we prove that, for any ordinal $\mu<\epsilon_0$, there exists no such variant mapping any hydra to an ordinal strictly less than $\mu$.
 Thus, the complexity of $\epsilon_0$ is really needed in the previous proof.
@@
 and the Wainer-Hardy hierarchy of ``rapidly growing functions'' $H_\alpha$~\cite{Wainer1970}.
-The considered class of battles, which we call \emph{standard}  is the most considered one in the scientific  litterature(including popularization).
+The considered class of battles, which we call \emph{standard},  is the most considered one in the scientific  literature (including popularization).
 \end{itemize}
@@
 Simply put, this document tries to combine the scientific interest of two articles~\cite{KP82, KS81} and a book~\cite{schutte} with the playful activity of truly proving theorems.
-We hope  that such a work, besides exploring a nice piece of discrete maths, 
+We hope  that such a work, besides exploring a nice piece of discrete mathematics,
 will show how \coq{} and its standard library are well fitted to help us to understand some non-trivial mathematical developments, and also to experiment the constructive parts of  the proof through functional programming.
@@
 
-\subsection*{Difference from Kirby and Paris's Work}
+\subsection*{Difference from Kirby and Paris's work}
 In~\cite{KP82}, Kirby and Paris show  that there is no proof of termination of all hydra battles in Peano Arithmetic (PA).
 Since we are used to writing proofs in higher order logic, the restriction to PA was quite unnatural for us.
-So we chosed to prove another statement without any reference to PA, by considering a class of proofs indexed by ordinal numbers up to $\epsilon_0$.
+So we chose to prove another statement without any reference to PA, by considering a class of proofs indexed by ordinal numbers up to $\epsilon_0$.
 
@@
 The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on  notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
-We added new material : axiomatic definition of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, equations, etc.
+We added new material: axiomatic definitions of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, equations, etc.
 
@@
 \end{tikzpicture}
-\caption{The hydra Hy \label{fig:Hy}}
+\caption{The hydra \texttt{Hy} \label{fig:Hy}}
 \end{figure}
 
-We use a specific vocabulary for talking about hydras. Table~\ref{tab:hyd2tree} shows the correspondance between our terminology and the usual vocabulary for trees in computer science.
+We use a specific vocabulary for talking about hydras. Table~\ref{tab:hyd2tree} shows the correspondence between our terminology and the usual vocabulary for trees in computer science.
 
@@
 We leave it to the reader to define various parameters such as the height, the size, the highest arity (number of sons of a node) of a hydra.
-In our example, these parameters have the respective values : $4$, $9$ and $3$.
+In our example, these parameters have the respective values $4$, $9$ and $3$.
 
@@
 Moreover, we note that this description is in \emph{imperative} terms.
-In order to build a formal  study of the properties of hydra battles, we prefer to use a mathematical vocabulary, i.e., graphs, relations, functions, etc.
+In order to formally study the properties of hydra battles, we prefer to use a mathematical vocabulary, i.e., graphs, relations, functions, etc.
 Thus, the replication process will be represented as a binary relation on a data type \texttt{Hydra},
@@
 
-At the first round, Hercules choses to chop off the rightmost head of \texttt{Hy}.
+At the first round, Hercules chooses to chop off the rightmost head of \texttt{Hy}.
 Since this head is near the floor, the hydra simply loses this head. Let us call 
@@
 
-Next, assume Hercules choses to chop off one of the two highest heads of \texttt{Hy'}, for instance the rightmost one.
+Next, assume Hercules chooses to chop off one of the two highest heads of \texttt{Hy'}, for instance the rightmost one.
 Fig.~\vref{fig:Hy2} represents the broken segment in dashed lines, and the part that will be replicated in red.
@@
 
-\caption{Hy': the state  of Hy after one round \label{fig:Hy-prime}}
+\caption{\texttt{Hy'}: the state  of \texttt{Hy} after one round \label{fig:Hy-prime}}
 \end{figure}
@@
 \end{tikzpicture}
-\caption{Hy'', the state of Hy after two rounds \label{fig:Hy3}}
+\caption{\texttt{Hy''}: the state of \texttt{Hy} after two rounds \label{fig:Hy3}}
 \end{figure}
@@
 \end{tikzpicture}
-\caption{The configuration Hy''' of Hy \label{fig:Hy5}}
+\caption{The configuration \texttt{Hy'''} of \texttt{Hy} \label{fig:Hy5}}
 \end{figure}
@@
 
-For instance, the hydra $Hy''$ of Fig~\vref{fig:Hy3}  can be defined in \coq{} as follows:
+For instance, the hydra \texttt{Hy''} of Fig~\vref{fig:Hy3}  can be defined in \coq{} as follows:
 
@@
 \begin{enumerate}
-\item Hercules chops off one head of the hydra
+\item Hercules chops off one head of the hydra.
 \item Then, the  hydra replicates the wounded part (if the head is at distance $\geq 2$ from the foot).
@@
 goal \Verb@{h':Hydra | h -*-> h'}@ generates the subgoal \Verb|h -*-> ?h'|.
-\item You may use Gérard Huet's \emph{zipper} data structure~\cite{zipper} for writing tactics associated with Hercule's  interactive search for a head to chop off.
+\item You may use Gérard Huet's \emph{zipper} data structure~\cite{zipper} for writing tactics associated with Hercules's  interactive search for a head to chop off.
 \end{itemize}
@@
 
-We chosed to call \emph{standard} the kind of battles which appear  most often in the litterature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$
+We chose to call \emph{standard} the kind of battles which appear  most often in the literature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$
 
@@
 
-  \caption{The hydra hinit}
+  \caption{The hydra \texttt{hinit}}
   \label{fig:hinit}
@@
 
-  \caption{The hydra (hyd1 h3)}
+  \caption{The hydra (\texttt{hyd1 h3})}
   \label{fig:hinit-plus2}
@@
 
-  \caption{The hydra (hyd 3 4 2)}
+  \caption{The hydra (\texttt{hyd 3 4 2})}
   \label{fig:hinit-plusn}
@@
 \subsection{Testing  \dots}
-In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple datatype for representing the states as tuples composed of
+In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple data type for representing the states as tuples composed of
 the round number, and the respective number of daughters  \texttt{h2}, \texttt{h1}, and heads
@@
 
-The following function returns the next configurarion of the game. 
+The following function returns the next configuration of the game.
 Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle only start with the lemma
@@
   \item  But the hydra $\iota(S(i))$ can be transformed in one round into
-    $\iota(i)$ (by losing its righmost head, for instance)
+    $\iota(i)$ (by losing its rightmost head, for instance)
   \item Since $m$ is a variant, we have $m(\iota(i)) < m(\iota(S(i)))$,
@@
 on \emph{ordinal numbers}.
-From a mathematical point of view, an ordinal is a representant of an equivalence class for isomorphims of  totally ordered well-founded sets.
+From a mathematical point of view, an ordinal is a representative of an equivalence class for isomorphisms of  totally ordered well-founded sets.
 
@@
 Nevertheless, since we work within type theory,
-we do not identify ordinals as sets of ordinals, but consider the correspondance between ordinals and sets of ordinals as the function that maps $\alpha$ to $\mathbb{O}_\alpha$.
+we do not identify ordinals as sets of ordinals, but consider the correspondence between ordinals and sets of ordinals as the function that maps $\alpha$ to $\mathbb{O}_\alpha$.
 For instance $\mathbb{O}_\omega=\mathbb{N}$, and $\mathbb{O}_7=\{0,1,2,3,4,5,6\}$.
 
-We cannot cite all the litterature published on ordinals since Cantor's book 
+We cannot cite all the literature published on ordinals since Cantor's book
 \cite{cantorbook}, and 
@@
 \begin{itemize}
-\item A ``mathematical'' representation of the set of countable ordinal numbers, afer Kurt Schütte~\cite{schutte}.
+\item A ``mathematical'' representation of the set of countable ordinal numbers, after Kurt Schütte~\cite{schutte}.
 This representation uses several (hopefully harmless) axioms.
@@
 If we consider any ordinal $\alpha=(i,j)$, where $i>0$, there exists an infinite number of
-ordinals stricly less than $\alpha$, and there exists an infinite number of battles that start from
+ordinals strictly less than $\alpha$, and there exists an infinite number of battles that start from
 $\iota(\alpha)$.
@@
 
-The first steps of our proof prepare a well-founded indsction on $\omega^2$.
+The first steps of our proof prepare a well-founded induction on $\omega^2$.
 
@@
 \begin{remark}
- There is no interesting arihmetic on finite ordinals, since functions like successor, addition, etc.,  cannot be represented in \coq{} as \emph{total} functions.
+ There is no interesting arithmetic on finite ordinals, since functions like successor, addition, etc.,  cannot be represented in \coq{} as \emph{total} functions.
 \end{remark}
@@
 \label{sec:orgheadline65}
-Unless explicitly mentionned, the term ``ordinal" will be used instead of
+Unless explicitly mentioned, the term ``ordinal" will be used instead of
 ``ordinal strictly less than \(\epsilon_0\)" (except in Chapter~\ref{chap:schutte} where it stands for ``countable ordinal'').
@@
 \paragraph{Remark}
-For simplicity's sake, we chosed to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the contruction (\texttt{ocons $\alpha$ $n$ $\beta$}) is intented to represent the
+For simplicity's sake, we chose to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the construction (\texttt{ocons $\alpha$ $n$ $\beta$}) is intended to represent the
 ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
@@
 
-In order to  upgrade constants and fonctions from type \texttt{T1} to \texttt{E0}, we have to prove that 
+In order to  upgrade constants and functions from type \texttt{T1} to \texttt{E0}, we have to prove that
 the term they build is in normal form.
@@
 
-    In the litterature, the natural sum of ordinals \(\alpha\) and \(\beta\) 
+    In the literature, the natural sum of ordinals \(\alpha\) and \(\beta\)
     is often denoted by \(\alpha \# \beta\)  or  \(\alpha \oplus  \beta\).
@@
  an article by J.~Ketonen and R.~Solovay~\cite{KS81}, already  cited in the work
- by L.~Kirby et J.~Paris.% on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
+ by L.~Kirby and J.~Paris.% on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
  Section $2$ of this article: ''A hierarchy of probably recursive functions'', contains a systematic study of \emph{canonical sequences}, which are closely related to
@@
 We prove by transfinite induction over $\alpha$ that $\canonseq{\alpha}{i+1}$ is an ordinal strictly less than $\alpha$ (assuming $\alpha\not=0$).
-This property allows us to use the function \texttt{canonS} and its derivates in function definitions by transfinite recursion.
+This property allows us to use the function \texttt{canonS} and its derivatives in function definitions by transfinite recursion.
 
@@
 Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
-we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying collorary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
+we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying corollary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
 where $n$ is the maximum of the indices $i$ met in the induction.
@@
 constructive aspects.
-Roughly steaking, our implementation \emph{builds}, out of a hypothetic 
+Roughly speaking, our implementation \emph{builds}, out of a hypothetical
   variant $m$, bounded by some ordinal $\mu<\epsilon_0$, a hydra \texttt{big\_h} which verifies the impossible inequality  $m(\texttt{big\_h})< m(\texttt{big\_h})$.
@@
 \begin{remark}
-Some notations (mainly names of fast-growing functions) of our development may differ slightly from the litterature.
+Some notations (mainly names of fast-growing functions) of our development may differ slightly from the literature.
 Although this fact does not affect our proofs, we are preparing a future version where the names $F\_alpha$, $f\_alpha$, $H\_alpha$, etc., are fully consistent with the cited articles.
@@
 \begin{remark}
-  Ketonen and Solovay~\cite{KS81} consider  large finite \emph{sets} of natural numbers,  but they are mainly used as sequences. Thus, we chosed to represent them explicitely as (sorted) lists. 
+  Ketonen and Solovay~\cite{KS81} consider  large finite \emph{sets} of natural numbers,  but they are mainly used as sequences. Thus, we chose to represent them explicitely as (sorted) lists. 
 \end{remark}
@@
 $\omega^\alpha\times i + \beta$.
-Clearly, both approachs are related through the equality
+Clearly, both approaches are related through the equality
 $L_\alpha(k)=H_\alpha(k)+1$, for any non-null $\alpha$ and $k$.
@@
 Note that, for $\alpha\not=0$, the value of $f(0)$ is not specified.
-Nevertheless, the restriction of $f$ to the set of strictly positive integers is unique (up to extensionnality).
+Nevertheless, the restriction of $f$ to the set of strictly positive integers is unique (up to extensionality).
 
@@
 To this end, we prove a generic lemma, which expresses $L_{\omega^\alpha\times i}$ as an iterate of $L_{\omega^\alpha}$.
-Note that in this lemma, we assume that the fonction associated with $\alpha$ is stritly monotonous and
+Note that in this lemma, we assume that the function associated with $\alpha$ is strictly monotonous and
 greater or equal than the successor function, and prove that $L_{\omega^\alpha\times i}$ satisfies  the same properties.
@@
 
-It is worth looking at the answer from \texttt{Equations} and at all the lemmas this plug-in gives you for free. We show here only a part of \coq's anwer.
+It is worth looking at the answer from \texttt{Equations} and at all the lemmas this plug-in gives you for free. We show here only a part of \coq's answer.
 
@@
   Indeed, the functions presented in this section are a \emph{variant} of the Hardy hierarchy of functions.
-In the future versions of this development, we will correct the references to the litterature.
+In the future versions of this development, we will correct the references to the literature.
 For the time being, we call our functions $H'_\alpha$ in order to underline the difference from ``classic'' Hardy functions.
@@
 
-Note that this short formula contains two occurences of the functional \texttt{iterate}, the outer one is in fact a second-order iteration (on type \texttt{nat -> nat)}
+Note that this short formula contains two occurrences of the functional \texttt{iterate}, the outer one is in fact a second-order iteration (on type \texttt{nat -> nat)}
 and the inner one  first-order (on type \texttt{nat}).
@@
 
-On the contrary, the fonctions of the $H'$ hierarchy have the following five properties~\cite{KS81}: for any $\alpha < \epsilon_0$,
+On the contrary, the functions of the $H'$ hierarchy have the following five properties~\cite{KS81}: for any $\alpha < \epsilon_0$,
 \begin{itemize}
@@
 
-In \coq{}, we follow the  proof in~\cite{KS81}. This proof is mainly a single  proof by transfinite induction on $\alpha$ of the conjonction of the five properties.
+In \coq{}, we follow the  proof in~\cite{KS81}. This proof is mainly a single  proof by transfinite induction on $\alpha$ of the conjunction of the five properties.
 For each $\alpha$, the three cases : $\alpha=0$, $\alpha$ is a limit, and 
@@
 In this section, we try to present an abstract of the properties of the main variants of fast-growing hierarchies of functions
-we fauund in the litterature or we had to define as helpers in our proofs.
+we found in the literature or we had to define as helpers in our proofs.
 
@@
 In order to work with classical logic, we import the module
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Classical.html}{Coq.Logic.Classical}  of \coq{}'s standard library, specifially the following axiom:
+\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Classical.html}{Coq.Logic.Classical}  of \coq{}'s standard library, specifically the following axiom:
 
@@
 
-Among the numerous properties of the ordinal $\omega$, les us quote the following ones
+Among the numerous properties of the ordinal $\omega$, let us quote the following ones
 (proved in Module 
@@
 
-It isinteresting to compare the proof of these lemmas with the 
+It is interesting to compare the proof of these lemmas with the
 computational proofs of the corresponding statements in Module
@@
 
-In~\cite{grimm:hal-00911710}, José Grimm establishes the consistency between our ordinal notations (\texttt{T1} and \texttt{T2} (Veblen normal form) and his implementation
+In~\cite{grimm:hal-00911710}, José Grimm establishes the consistency between our ordinal notations \texttt{T1} and \texttt{T2} (Veblen normal form) and his implementation
 of ordinal numbers after Bourbaki's set theory.
@@
 
-Like in chapter~\ref{chap:T1}, we get familiar with the type \texttt{T2} by recognising simple constructs like finite ordinals, $\omega$, etc., as inhabitants of \texttt{T2}.
+Like in chapter~\ref{chap:T1}, we get familiar with the type \texttt{T2} by recognizing simple constructs like finite ordinals, $\omega$, etc., as inhabitants of \texttt{T2}.
 
@@
 \begin{remark}
-The connexion between the predicate \texttt{nf} and the relation \texttt{lt} on one part, and the functions \texttt{nfb} and \texttt{compare} on the other, is expressed by the following lemmas:
+The connection between the predicate \texttt{nf} and the relation \texttt{lt} on one part, and the functions \texttt{nfb} and \texttt{compare} on the other, is expressed by the following lemmas:
 

### 2021-08-08 doc/part-hydras.tex
fix typos and some notation and grammar in part-hydras.tex

@@
 
-In order to build an instance of \texttt{ON}, we define a comparison function,  and prove its correctiness.
-
-\vspace{4pt}
+In order to build an instance of \texttt{ON}, we define a comparison function,  and prove its correction.
 
@@
 
-\input{movies/snippets/T1/T1Def}
+\begin{Coqsrc}
+Inductive T1 : Set  :=
+| zero : T1
+| ocons : T1 -> nat -> T1 -> T1.
+\end{Coqsrc}
 
@@
 \noindent
-
-\input{movies/snippets/T2/T2Def}
+\begin{Coqsrc}
+Inductive T2 : Set :=
+  zero : T2
+| gcons : T2 -> T2 -> nat -> T2 -> T2.
+\end{Coqsrc}
 
@@
 
-\input{movies/snippets/T1/alpha0}
+\begin{Coqsrc}
+Example alpha_0 : T1 :=
+  ocons (ocons (ocons zero 0 zero)
+               0
+               zero)
+        0
+       (ocons (ocons zero 2 zero)
+              4
+              (ocons zero 1 zero)).
+\end{Coqsrc}
 
@@
 
+\begin{Coqsrc}
+Notation one := (ocons zero 0 zero).
+
+Notation FS n := (ocons zero n zero).
+\end{Coqsrc}
+
 \label{sect:notation-F}
 
-\input{movies/snippets/T1/finiteOrds}
+\begin{Coqsrc}
+Definition fin (n:nat) : T1 := 
+    match n with 0 => zero | S p => FS p end.
+
+Coercion fin  : nat >-> T1.
+
+Example ten : T1 := 10.   
+\end{Coqsrc}
 
@@
 \label{sect:omega-notation2}
-
-\input{movies/snippets/T1/omegaDef}
+\begin{Coqsrc}
+Notation omega := (ocons (ocons zero 0 zero) 0 zero).
+\end{Coqsrc}
 
@@
 
-\input{movies/snippets/T1/phi0Def}
+\begin{Coqsrc}
+Notation "'phi0' alpha" := (ocons alpha 0 zero) (at level 29) : t1_scope.
+\end{Coqsrc}
 
@@
 
-\input{movies/snippets/T1/towerDef}
+\begin{Coqsrc}
+Fixpoint omega_tower (height:nat) : T1 := 
+ match height with 
+ | 0 =>  1 
+ | S h => phi0 (omega_tower h)
+ end.
+\end{Coqsrc}
 
@@
 \end{Coqsrc}
-
-% Uncomment for testing 
-%\textbf{Todo: Fix subgoal display }
-%\input{movies/snippets/T1/compareRev}
 

### 2021-08-09 doc/part-hydras.tex
converting chapter 4 to alectryon; add a link to Gaia

--- 2021-08-08 doc/part-hydras.tex
@@
 
+\label{sect:comparable-def}
+
 \input{movies/snippets/Comparable/ComparableDef}
@@
 
-\begin{Coqsrc}
-Compute alpha_0.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = ocons omega 0 (ocons (FS 2) 4 (FS 1))
-     : T1
-\end{Coqanswer}
+\input{movies/snippets/T1/alpha0Compute}
 
@@
 
-\begin{Coqsrc}
-Inductive ppT1 : Set :=
-    P_fin : nat -> ppT1
-  | P_add : ppT1 -> ppT1 -> ppT1
-  | P_mult : ppT1 -> nat -> ppT1
-  | P_exp : ppT1 -> ppT1 -> ppT1
-  | P_omega : ppT1
-
-Notation _omega := P_omega.
-
-\end{Coqsrc}
+\input{movies/snippets/T1/ppT1Def}
 
@@
 
-\begin{Coqsrc}
-Compute pp alpha_0.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-     = (_omega ^ _omega + _omega ^ 3 * 5 + 2)%pT1
-     : ppT1
-\end{Coqanswer}
+\input{movies/snippets/T1/ppAlpha0}
 
@@
 
-In order to compare two terms of type \texttt{T1}, we define a recursive function \texttt{compare} that maps two ordinals $\alpha$ and $\beta$ to a value of type \texttt{comparison}.
+In order to compare two terms of type \texttt{T1}, we define a recursive function \texttt{compare} that maps two ordinal tems $\alpha$ and $\beta$ to a value of type \texttt{comparison}.
 This type is defined in \coq's standard library
@@
 
-\begin{Coqsrc}
-Fixpoint compare (alpha beta:T1):comparison :=
-  match alpha, beta with
-    zero, zero => Eq
-  | zero, ocons a' n' b' => Lt
-  | _   , zero => Gt
-  | (ocons a n b),(ocons a' n' b') =>
-      (match compare a a' with 
-       | Lt => Lt
-       | Gt => Gt
-       | Eq => (match Nat.compare n n' with
-                | Eq => compare b b'
-                | comp => comp
-                end)
-       end)
-  end.
-\end{Coqsrc}
-
-It is now easy to define the boolean predicate \texttt{lt\_b $\alpha$ $\beta$}: 
-`` $\alpha$ is strictly less than $\beta$ ''. By coercion to sort \texttt{Prop} we define also the predicate \texttt{lt}.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}
-
-\begin{Coqsrc}
-Definition lt_b alpha beta : bool :=
-  match compare alpha beta with
-      Lt => true
-    | _ => false
-  end.
-
-Definition lt alpha beta : Prop := lt_b alpha beta.
-\end{Coqsrc}
+\input{movies/snippets/T1/compareDef}
 
@@
 
-\begin{Coqsrc}
-Example E1 : lt (ocons omega 56 zero) (tower 3).
-Proof. reflexivity. Qed.
-
-Example E2 : ~ lt (tower 3) (tower 3).
-Proof.  discriminate.  Qed.
-\end{Coqsrc}
-
-The following lemmas establish relations between \texttt{compare}, 
-the predicate \texttt{lt} and Leibniz equality \texttt{eq}.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare_refl}{Epsilon0.T1}}
-
-\begin{Coqsrc}
-Lemma compare_refl : forall alpha, compare alpha alpha =  Eq.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma compare_reflect : forall alpha beta,
-    match compare alpha beta with
-    |   Lt => lt alpha  beta
-    |   Eq => alpha = beta
-    |   Gt => lt beta  alpha
-    end.
-\end{Coqsrc}
-
-We prove also that the relation \texttt{lt} is a strict total order.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#lt_irrefl}{Epsilon0.T1}}
-
-\begin{Coqsrc}
-Theorem lt_irrefl (alpha: T1):  ~ lt alpha alpha.
-
-Theorem lt_trans (alpha beta gamma : T1) :
-  lt alpha  beta -> lt beta gamma -> lt alpha gamma.
-
-Definition lt_eq_lt_dec  :
-   forall alpha beta : T1, 
-          {lt alpha  beta} + {alpha = beta} + {lt beta alpha}.
-\end{Coqsrc}
-
-Note that the order \texttt{lt} is not reflected 
-in the structure (size and/or height) of the terms of \texttt{T1}.
-
-\begin{Coqsrc}
-Example Ex0:
-  lt (ocons (phi0 (phi0 omega)) 2
-            (ocons (phi0 10) 33
-                   (ocons (phi0 9) 63 zero)))
-     (ocons  (phi0 (phi0 omega)) 2 (phi0 (phi0 11))).
-Proof.
-  reflexivity. 
-Qed.
-\end{Coqsrc}
+\vspace{4pt}
+
+\input{movies/snippets/T1/ltExamples}
+
+%%% TODO : link to Comparable and StrictOrder classes
+
+Links between the function \texttt{compare} and the relations
+\texttt{lt} and \texttt{eq} are established through the following lemmas (~\vref{sect:comparable-def}).
+\vspace{4pt}
+
+\input{movies/snippets/T1/Instances}
+
+% \vspace{4pt}
+% \emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare_refl}{Epsilon0.T1}}
+
+% \begin{Coqsrc}
+% Lemma compare_refl : forall alpha, compare alpha alpha =  Eq.
+% \end{Coqsrc}
+
+% \begin{Coqsrc}
+% Lemma compare_reflect : forall alpha beta,
+%     match compare alpha beta with
+%     |   Lt => lt alpha  beta
+%     |   Eq => alpha = beta
+%     |   Gt => lt beta  alpha
+%     end.
+% \end{Coqsrc}
+
+% We prove also that the relation \texttt{lt} is a strict total order.
+
+% \vspace{4pt}
+% \emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#lt_irrefl}{Epsilon0.T1}}
+
+% \begin{Coqsrc}
+% Theorem lt_irrefl (alpha: T1):  ~ lt alpha alpha.
+
+% Theorem lt_trans (alpha beta gamma : T1) :
+%   lt alpha  beta -> lt beta gamma -> lt alpha gamma.
+
+% Definition lt_eq_lt_dec  :
+%    forall alpha beta : T1, 
+%           {lt alpha  beta} + {alpha = beta} + {lt beta alpha}.
+% \end{Coqsrc}
+
+% \vspace{4pt}
+
+% Please note that the order \texttt{lt} is not reflected 
+% in the structure (size and/or height) of the terms of \texttt{T1}.
+
+% \begin{Coqsrc}
+% Example Ex0:
+%   lt (ocons (phi0 (phi0 omega)) 2
+%             (ocons (phi0 10) 33
+%                    (ocons (phi0 9) 63 zero)))
+%      (ocons  (phi0 (phi0 omega)) 2 (phi0 (phi0 11))).
+% Proof.
+%   reflexivity. 
+% Qed.
+% \end{Coqsrc}
 
@@
 
-\begin{Coqbad}
-Example bad_term  : T1 := ocons 1 1 (ocons omega 2 zero).
-\end{Coqbad}
+\input{movies/snippets/T1/badTerm}
 
@@
 
-\begin{Coqsrc}
-Fixpoint nf_b (alpha : T1) : bool :=
-  match alpha with
-    | zero => true
-    | ocons a n zero => nf_b a
-    | ocons a n ((ocons a' n' b') as b) =>
-      (nf_b a && nf_b b && lt_b a' a)%bool
-  end.
-
-Definition nf alpha: Prop := nf_b alpha.
-\end{Coqsrc}
-
-\begin{Coqsrc}
- Compute nf_b alpha_0.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-   = true 
-     : bool
-\end{Coqanswer}
-
-\begin{Coqsrc}
- Compute nf_b bad_term.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-   = false 
-     : bool
-\end{Coqanswer}
+\input{movies/snippets/T1/nfDef}
+
+\input{movies/snippets/T1/nfAlpha0}
+
+\input{movies/snippets/T1/nfBadTerm}
 
@@
 
-\begin{Coqsrc}
-Lemma plus_is_zero alpha beta :
-  nf alpha -> nf beta ->
-  alpha + beta  = zero -> alpha = zero /\  beta = zero.
-\end{Coqsrc}
+\input{movies/snippets/T1/plusIsZero}
+
+\vspace{4pt}
 
@@
 
-\begin{Coqsrc}
-Definition restrict {A:Type}(E: Ensemble A)(R: relation A) :=
-    fun a b => E a /\ R a b /\ E b.
- \end{Coqsrc}
+\input{movies/snippets/Restriction/restrictionDef}
 
@@
 
-\begin{Coqsrc}
-Definition LT := restrict nf lt.
-Infix "t1<" := LT : t1_scope.
-
-Definition LE := restrict nf le.
-Infix "t1<=" := LE : t1_scope.
-\end{Coqsrc}
+\input{movies/snippets/T1/LTDef}
 
@@
 
-\begin{Coqsrc}
-Lemma LT_one : forall alpha, alpha t1< one -> alpha = zero.
-\end{Coqsrc}
+\input{movies/snippets/T1/LTOnea}
+
+\begin{Coqanswer}
+1 subgoal (ID 738)
+
+  alpha1 : T1
+  n : nat
+  alpha2 : T1
+  H1 : nf (ocons alpha1 n alpha2)
+  H2 : lt (ocons alpha1 n alpha2) one
+  ============================
+  ocons alpha1 n alpha2 = zero
+\end{Coqanswer}
+
+\input{movies/snippets/T1/LTOneb}
 
@@
 
+The Gaia project ~\url{https://github.com/coq-community/gaia} maintains Grimm's  theory of ordinals as part of coq-community on GitHub. Integration
+of the present ordinal theory with Gaia, i.e., relating the different notions of ordinals
+and transferring relevant results, would be an interesting project.
+
 \chapter{The Ordinal \texorpdfstring{$\Gamma_0$}{Gamma0} (first draft)}

### 2021-08-09 doc/chapter-primrec.tex
fix typos in chapter-powers.tex and chapter-primrec.tex

@@
 O'Connor's library on Gödel's incompleteness theorems contains a little more 
-than 45K lines of scripts. The part dedicated to primitive recursive functions and Peano arithmetics is 32K lines long and is originally structured in 38 modules.
+than 45K lines of scripts. The part dedicated to primitive recursive functions and Peano arithmetic is 32K lines long and is originally structured in 38 modules.
 Thus, we propose a partial exploration of this library, through examples and exercises. Our additions to the original library --- mainly examples and counter-examples ---,
@@
 After this test, the term \texttt{bigPR} looks to be a primitive recursive definition of the factorial function, although we haven't proved this fact yet.
-Fortunately, we will see in the following sections simple ways to prove that a given function is primitive recursive, whithout building such an unreadable term.
+Fortunately, we will see in the following sections simple ways to prove that a given function is primitive recursive, without building such an unreadable term.
 
@@
 
-Thus, the proof that a function, like \texttt{factorial}, is primitive recursive may be interactive, whithout having to type complex terms at any step of the development.
+Thus, the proof that a function, like \texttt{factorial}, is primitive recursive may be interactive, without having to type complex terms at any step of the development.
 
@@
 \index{primrec}{Predicates!isPR}
-\index{coq}{Extensionnaly equal functions}
+\index{coq}{Extensionally equal functions}
 
-Let $f$ be an arithmetic function of arity $n$. We say that $f$ is primitive recursive if $f$ is \textbf{extensionnaly}
+Let $f$ be an arithmetic function of arity $n$. We say that $f$ is primitive recursive if $f$ is \textbf{extensionally}
 equal to the interpretation of some term of type \texttt{PrimRec $n$}.
@@
 
-We prefered to define a variant which uses explicitely 
+We preferred to define a variant which uses explicitly
  the functional \texttt{iterate},
@@
 The file \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack} presents two other definitions of the Ackermann functions based on the lexicographic ordering on $\mathbb{N}\times\mathbb{N}$.
-Prove that the four functions are extensionnally equal.
+Prove that the four functions are extensionally equal.
 \end{exercise}
@@
 The three first lemmas make us sure that our function 
-\texttt{Ack} satifies the ``usual'' equations.
+\texttt{Ack} satisfies the ``usual'' equations.
 
@@
 
-Ìn order to prove that \texttt{Ack} (considered as a function of two arguments) is not primitive recursive, the usual method consists in two steps:
+In order to prove that \texttt{Ack} (considered as a function of two arguments) is not primitive recursive, the usual method consists in two steps:
 
@@
 
-Please note that, in order to prove a property shared by any primitive recursive function of, say, arity 2, this induction scheme  leads you to consider an expension of the considered property to primitive recursive function of any arity.
+Please note that, in order to prove a property shared by any primitive recursive function of, say, arity 2, this induction scheme  leads you to consider an extension of the considered property to primitive recursive function of any arity.
 
@@
 \begin{remark}
-It is easy to prove that any unary function which dominatates \texttt{fun n => Ack n n} fails to be primitive recursive.
+It is easy to prove that any unary function which dominates \texttt{fun n => Ack n n} fails to be primitive recursive.
 To this end, we use an instance of \texttt{majorAnyPR} dealing with unary functions.
@@
 {\color{red}
-  ``Let $f$ be a function of type \texttt{naryFunc\,2}. If, for any $n$, the fonction $f(n)$ is primitive recursive, then f is primitive recursive.''}
+  ``Let $f$ be a function of type \texttt{naryFunc\,2}. If, for any $n$, the function $f(n)$ is primitive recursive, then f is primitive recursive.''}
 \end{quote}
@@
 
-In order to get rid of the substraction in the definition of \texttt{l\_std}, we work with a helper function.
+In order to get rid of the subtraction in the definition of \texttt{l\_std}, we work with a helper function.
 
@@
 
-Our proof of this lemma is not trivial at all, it uses some properties of the Ketonen-Solovay's toolkit. We advise the reader to explore this proof, with the help of an ide or software like alectryon.
+Our proof of this lemma is not trivial at all, it uses some properties of the Ketonen-Solovay's toolkit. We advise the reader to explore this proof, with the help of an IDE or software like Alectryon.
 %%

### 2021-08-09 doc/chapter-powers.tex
fix typos in chapter-powers.tex and chapter-primrec.tex

@@
 
-In~\cite{BC04}, several exercices~\footnote{Exercises 9.8 (page 270), 9.10 (page 271), 9.15 (page 276), 9.17 (page 284), and 15.8 (page 418).}  
+In~\cite{BC04}, several exercises~\footnote{Exercises 9.8 (page 270), 9.10 (page 271), 9.15 (page 276), 9.17 (page 284), and 15.8 (page 418).}
 present ways to compute Fibonacci numbers, with the less number of recursive calls  as possible.
@@
 Thus, any function able to compute more or less efficiently powers in a monoid will
-give an algorithm for computing Fibonacci numbers. Unlike the \emph{ad-hoc} aforementionned proofs of~\cite{BC04}, the correctness of such an algorithm is a direct consequence 
+give an algorithm for computing Fibonacci numbers. Unlike the \emph{ad-hoc} aforementioned proofs of~\cite{BC04}, the correctness of such an algorithm is a direct consequence
 of the correctness of the used powering function.
@@
   In the following inductive type definition, the intended meaning 
-  of the contruct (\texttt{Mult $x$ $y$ $k$})  is \emph{``multiply \texttt{x} with 
+  of the construct (\texttt{Mult $x$ $y$ $k$})  is \emph{``multiply \texttt{x} with
 \texttt{y}, then send  the result of this multiplication to 
@@
 notation makes terms of type \texttt{computation} look like
-expressions of a small programming language dedicated to sequences of mutiplications.
+expressions of a small programming language dedicated to sequences of multiplications.
 Please look at \emph{CPDT}~\cite{chlipalacpdt2011} for more details on monadic notations in \coq.
@@
 obtaining a ``canonical'' form for terms generated from a variable \texttt{x}
-by contructors associated with monoid multiplication and neutral element.
+by constructors associated with monoid multiplication and neutral element.
 
@@
 \subsubsection{Definition}
-A chain $c$ is \emph{parametric} if it has the same behaviour for any
+A chain $c$ is \emph{parametric} if it has the same behavior for any
 pair of types $A$  and $B$, any relation $R$
@@
 Unfortunately, our construct is still very inefficient, since it results in 
-duplications of computations, as shown by the normal form of \texttt{F23}.
+duplication of computations, as shown by the normal form of \texttt{F23}.
 
@@
 We are now ready to define various operators on f- and k-chains, and prove these
-operators preserve correcness and properness. We will also show that 
+operators preserve correctness and properness. We will also show that
 these operators allow to generate easily correct chains for any positive 
@@
 define our chain generator by structural recursion. 
-For simplicity's sake, we chosed to avoid dependent elimination
+For simplicity's sake, we chose to avoid dependent elimination
  and used \texttt{Function}  with a decreasing measure.
@@
 \begin{itemize}
-\item The generation of a k-chain for $n$ and $p$ is meaningfull only if $p < n$. Thus, in order to avoid a clumsy  dependent pattern-matching, we chosed to represent 
+\item The generation of a k-chain for $n$ and $p$ is meaningful only if $p < n$. Thus, in order to avoid a clumsy  dependent pattern-matching, we chose to represent
      a pair $(n,p)$ where $0<p<n$ by a pair of positive numbers $(p,d)$ where 
@@
 
-\subsection{Generation of chains using Euclidean Dibision}
+\subsection{Generation of chains using Euclidean Division}
 
@@
 \item Prove that our implementation of the dichotomic strategy describes
- the same function as in the litterature (for instance ~\cite{DBLP:journals/ita/BrlekCHM95}.)
+ the same function as in the literature (for instance ~\cite{DBLP:journals/ita/BrlekCHM95}.)
 This is important if we want to follow the complexity analyses in this and similar articles.

### 2021-08-09 doc/chapter-powers.tex
fix some small non-typo issues in chapter-powers.tex

@@
 
-Nothing looks simpler than writing some function for computing $x^n$.
-On the contrary, this simple programming exercise allows us to address
+Nothing looks simpler than writing a function for computing $x^n$.
+But on the contrary, this simple programming exercise allows us to address
 advanced programming techniques such as:
@@
 \item type classes, and generalized rewriting
-\item proof engineering, in particular proof re-using
+\item proof engineering, in particular proof reuse
 \item proof by reflection
@@
 present ways to compute Fibonacci numbers, with the less number of recursive calls  as possible.
-Please note that these optimizations and the formal proof of their correctness are \emph{ad-hoc}, \emph{i.e.} exclusively written for the
+Please note that these optimizations and the formal proof of their correctness are \emph{ad-hoc}, \emph{i.e.}, exclusively written for the
 Fibonacci numbers.
@@
 \index{coq}{Type classes!Operational type classes}
-≈Ò
+
 \emph{Operational type classes}~\cite{BS2011}
@@
 These properties are easily expressed within the type class framework, by defining a \emph{family} of monoids.
-It suffices to define an instance of \texttt{Monoid} within the scope of an hypothesis
-of type \texttt{semi\_ring\_theory}
+It suffices to define an instance of \texttt{Monoid} within the scope of a hypothesis
+of type \texttt{semi\_ring\_theory}.
 
@@
 \index{maths}{Addition chains}
-An \emph{addition chain} (In short : \emph{a chain})~\cite{brauer1939} is a representation of a sequence of
+An \emph{addition chain} (in short, a \emph{chain})~\cite{brauer1939} is a representation of a sequence of
 intermediate steps that lead to the evaluation of  $x^n$, under the 
@@
 In articles from the combinatorist  community, 
-\emph{e.g.}~\cite{brauer1939,DBLP:journals/ipl/BerstelB87},  addition chains 
+\emph{e.g.},~\cite{brauer1939,DBLP:journals/ipl/BerstelB87},  addition chains
 are represented as sequences of positive integers, each member of which 
@@
  $c_{87}$  and $c'_{87}$. 
-Please note that some chains may be represented by various different dags.
+Please note that some chains may be represented by various different dags (directed acyclic graphs).
 For instance, we can associate four different dags to the chain $(1,2,3,4,6,9,13)$.
@@
 
-Thus, the main steps of a correctness proof of a given chain, \emph{e.g.}
+Thus, the main steps of a correctness proof of a given chain, \emph{e.g.},
 \texttt{C87} will be the following ones:

### 2021-08-10 doc/part-hydras.tex
initialize the branch Bridge

@@
 
-\index{Coq!Unicity of equality proofs}
+\index{Coq}{Unicity of equality proofs}
+
 \label{sect:eq-proof-unicity}
@@
 
-\begin{Coqsrc}
-Class E0 : Type := t1_2o {cnf : T1; cnf_ok : nf cnf}.
-\end{Coqsrc}
+\input{movies/snippets/E0/E0Def}
 
@@
 
-\begin{Coqsrc}
-Declare Scope E0_scope.
-Delimit Scope E0_scope with e0.
-Open Scope E0_scope.
-\end{Coqsrc}
+\input{movies/snippets/E0/E0Scope}
 
@@
 
-\begin{Coqsrc}
-Definition Lt (alpha beta : E0) := T1.LT (@cnf alpha) (@cnf beta).
-Definition Le (alpha beta : E0) := T1.LE (@cnf alpha) (@cnf beta).
-
-Infix "o<" := Lt : E0_scope.
-Infix "o<=" := Le : E0_scope.
-\end{Coqsrc}
+\input{movies/snippets/E0/LtDef}
 
@@
 
+\vspace{4pt}
+
 \index{coq}{Unicity of equality proofs}
 
-\begin{Coqsrc}
-Require Import Logic.Eqdep_dec.
-
-Lemma nf_proof_unicity :
-  forall (alpha:T1) (H H': nf alpha), H = H'.
-
-Lemma E0_eq_iff alpha beta : alpha = beta <-> cnf alpha = cnf beta.
-\end{Coqsrc}
+\input{movies/snippets/E0/nfProofUnicity}
+
+\input{movies/snippets/E0/E0EqIff}
 
@@
 
+\vspace{4pt}
+
 \label{sect:omega-T1}
+\input{movies/snippets/E0/ZeroOmega}
+
 %\index{Constants!zero:T1}
-
-\begin{Coqsrc}
-Instance Zero : E0.
-Proof.
-  now exists T1.zero.
-Defined.
-
-Instance _Omega : E0.
-Proof.  now exists omega%t1. 
-Defined.
-
-Notation omega  := _Omega.
-\end{Coqsrc}
 
@@
 
-\begin{Coqsrc}
-Fixpoint succb (alpha:T1) : bool :=
-  match alpha with
-    | zero => false
-    | ocons zero _ _ => true
-    | ocons alpha n beta => succb beta
-  end.
-
-Fixpoint limitb (alpha : T1) : bool :=
-  match alpha with
-    | zero => false
-    | ocons zero _ _ => false
-    | ocons alpha n zero => true
-    | ocons alpha n beta => limitb beta
-  end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-  Compute limitb omega.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = true
-     : bool
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Compute succb 42.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = true
-     : bool
-\end{Coqanswer}
+\input{movies/snippets/T1/succbLimitb}
 
@@
 limit and successor ordinals is established through several lemmas. For instance,
-Lemma \texttt{canonS\_limit}, page~\pageref{lemma:canonS-limit}, shows that
+Lemma \texttt{canonS\_limit\_lub}, page~\pageref{lemma:canonS-limit}, shows that
 if $\alpha$ is (syntactically) a limit ordinal, then it is the least upper bound of
 a strictly increasing sequence of ordinals.
-
-   The following function is very useful in constructions by cases (proofs and function definitions).
-
-\begin{Coqsrc}
-Definition zero_succ_limit (alpha: T1) :
-    {succb alpha} + {limitb alpha} +  {alpha=zero}.
-    (* ... *)
-Defined.
-\end{Coqsrc}
 
@@
 \end{Coqsrc}
+%
+% TO do : insert succ in E0
+%
 
@@
 
-\index{hydras}{Exercises} 
-\begin{exercise}
-  Prove that that the ordinal addition on \texttt{E0} is associative.
-
-  \textbf{Warning:}  Please do not import theorems from
-  \texttt{ordinals/Schutte/}, since that library uses axioms!
-\end{exercise}
-
 \section{Well-foundedness and transfinite induction}
@@
 \end{exercise}
-
-% \subsubsection{The total order \texttt{lt} on \texttt{T1} is \emph{not} well-founded}
-% \label{sec:orgheadline76}
-
-% Let us recall that the data type \texttt{T1} contains too many inhabitants, including
-% terms which are not in Cantor normal form. Thus, the following result is not 
-% very surprising.
-
-% \begin{Coqsrc}
-% Section lt_not_well_founded.
-
-%   (* let us build the sequence of terms :
-%         omega + omega + .... + omega ^ 2   *)
-%   Let f := (fix f (i:nat): T1 :=
-%             match i with 0 => phi0 2
-%                        | S i => ocons 1 1 (f  i)
-%             end).
-
-%  Lemma  f_decreases : forall i, f (S i) <  f i.
-%  Proof.
-%   induction i; compute; auto with T1.
-%  Qed.
-
-%  Theorem lt_not_wf : ~  well_founded lt.
-%  Proof. 
-%    intro wf; case (not_decreasing _ lt);auto.
-%    exists f; apply f_decreases.
-%  Qed.
-
-% End lt_not_well_founded.
-% \end{Coqsrc}
-
-% Thus, we have
 
@@
 
-\begin{Coqbad}
-Section First_attempt.
-
- Lemma wf_LT : forall alpha,  nf alpha -> Acc LT alpha. 
- Proof.
-  induction alpha as [| beta IHbeta n gamma IHgamma].
-  - split.
-    inversion 1.
-    destruct H2 as [H3 _];not_neg H3.
-  -  split; intros delta Hdelta.
-\end{Coqbad}
+\input{movies/snippets/T1/wfLTBada}
 
@@
 
-\begin{Coqbad}
-  Abort.
-\end{Coqbad}
-
 The problem comes from the hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
@@
 Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.
+
+\input{movies/snippets/T1/wfLTBadz}
 

### 2021-08-12 doc/part-hydras.tex
converting chapter 4 to Alectryon (start)
converting chapter 4 to Alectryon

--- 2021-08-09 doc/part-hydras.tex
@@
 
-First, we declare a notation scope for \texttt{E0}.
-
-\begin{Coqsrc}
-Declare Scope E0_scope.
-Delimit Scope E0_scope with e0.
-Open Scope E0_scope.
-\end{Coqsrc}
-
-Then we redefine the predicates of comparison.
-
-\label{Predicates:Lt-E0}
-
-\begin{Coqsrc}
-Definition Lt (alpha beta : E0) := T1.LT (@cnf alpha) (@cnf beta).
-Definition Le (alpha beta : E0) := T1.LE (@cnf alpha) (@cnf beta).
-
-Infix "o<" := Lt : E0_scope.
-Infix "o<=" := Le : E0_scope.
-\end{Coqsrc}
+\input{movies/snippets/E0/LtLeDef}
 
@@
 
-\begin{Coqsrc}
-Require Import Logic.Eqdep_dec.
-
-Lemma nf_proof_unicity :
-  forall (alpha:T1) (H H': nf alpha), H = H'.
-
-Lemma E0_eq_iff alpha beta : alpha = beta <-> cnf alpha = cnf beta.
-\end{Coqsrc}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-In earlier versions of this development, the predicate \texttt{nf} was defined  inductively, with various constructors describing all possible cases.
-\begin{enumerate}
-\item Please give such a definition, in a dedicated module.
-\item Prove the logical equivalence between your definition and ours.
-\item Define a variant of the type \texttt{E0} (with your definition of \texttt{nf}).
-\item Can you still prove a lemma like \texttt{E0\_eq\_iff} ? With the help of an axiom from some module of the standard library ?
-\end{enumerate}
-\end{exercise}
+\input{movies/snippets/E0/nfProofUnicity}
+
+\vspace{4pt}
+
+\input{movies/snippets/E0/E0EqIff}
+
+\vspace{4pt}
 
@@
 
+\vspace{4pt}
+
 \label{sect:omega-T1}
-%\index{Constants!zero:T1}
-
-\begin{Coqsrc}
-Instance Zero : E0.
-Proof.
-  now exists T1.zero.
-Defined.
-
-Instance _Omega : E0.
-Proof.  now exists omega%t1. 
-Defined.
-
-Notation omega  := _Omega.
-\end{Coqsrc}
+
+\input{movies/snippets/E0/ZeroOmegaDef}
 
@@
 
-   The following function is very useful in constructions by cases (proofs and function definitions).
-
-\begin{Coqsrc}
-Definition zero_succ_limit (alpha: T1) :
-    {succb alpha} + {limitb alpha} +  {alpha=zero}.
-    (* ... *)
-Defined.
-\end{Coqsrc}
-
 \subsection{Arithmetic on \texorpdfstring{$\epsilon_0$}{epsilon0}}
@@
 
-\begin{Coqsrc}
-Fixpoint succ (alpha:T1) : T1 :=
-  match alpha with 
-   | zero => 1
-   | ocons zero n _ => ocons zero (S n) zero
-   | ocons beta n gamma => ocons beta n (succ gamma)
- end.
-\end{Coqsrc}
+\input{movies/snippets/T1/succDef}
 
@@
 
-\begin{Coqsrc}
- Lemma succb_iff alpha (Halpha : nf alpha) :
-  succb alpha <-> exists beta : T1, nf beta /\ alpha = succ  beta.
-\end{Coqsrc}
+\vspace{4pt}
+
+\input{movies/snippets/T1/succbIff}
 
@@
 
-\begin{Coqsrc}
-Fixpoint plus (alpha beta : T1) : T1 :=
-  match alpha,beta with
- |  zero, y  => y
- |  x, zero  => x
- |  ocons a n b, ocons a' n' b' =>
-    (match compare a a' with
-     | Lt => ocons a' n' b'
-     | Gt => (ocons a n (plus b (ocons a' n' b')))
-     | Eq  => (ocons a (S(n+n')) b')
-     end)
-  end
-where "alpha + beta" := (plus alpha beta) : t1_scope.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Fixpoint mult (alpha beta : T1) :T1 :=
-  match alpha,beta with
- |  zero, y  => zero
- |  x, zero => zero
- |  ocons zero n _, ocons zero n' _ => 
-                 ocons zero (Peano.pred((S n) * (S n'))) zero
- |  ocons a n b, ocons zero n' b' =>  
-                 ocons a (Peano.pred((S n) * (S n'))) b
- |  ocons a n b, ocons a' n' b' =>
-     ocons (a + a') n' ((ocons a n b) * b')
- end
-where  "alpha * beta" := (mult alpha beta) : t1_scope.
-\end{Coqsrc}
+\input{movies/snippets/T1/plusDef}
+
+\input{movies/snippets/T1/multDef}
 
@@
 
-\begin{Coqsrc}
-Example e2 : 6 + omega = omega.
-Proof. reflexivity. Qed.
-
-Example e'2 : omega t1< omega + 6.
-Proof. now compute. Qed.
-
-Example e''2 : 6 * omega = omega.
-Proof. reflexivity. Qed.
-
-Example e'''2 : omega t1< omega * 6.
-Proof. now compute. Qed.
-\end{Coqsrc}
+\input{movies/snippets/T1/plusMultExamples}
 

### 2021-08-14 doc/part-hydras.tex
Update doc/part-hydras.tex
Co-authored-by: Karl Palmskog <palmskog@gmail.com>

--- 2021-08-10 doc/part-hydras.tex
@@
 \index{Coq}{Unicity of equality proofs}
-
 \label{sect:eq-proof-unicity}
@@
 
-\input{movies/snippets/E0/LtDef}
+\input{movies/snippets/E0/LtLeDef}
 
@@
 
+\vspace{4pt}
+
 \input{movies/snippets/E0/E0EqIff}
 
-\index{hydras}{Exercises}
-\begin{exercise}
-In earlier versions of this development, the predicate \texttt{nf} was defined  inductively, with various constructors describing all possible cases.
-\begin{enumerate}
-\item Please give such a definition, in a dedicated module.
-\item Prove the logical equivalence between your definition and ours.
-\item Define a variant of the type \texttt{E0} (with your definition of \texttt{nf}).
-\item Can you still prove a lemma like \texttt{E0\_eq\_iff} ? With the help of an axiom from some module of the standard library ?
-\end{enumerate}
-\end{exercise}
+\vspace{4pt}
 
@@
 \label{sect:omega-T1}
+
 \input{movies/snippets/E0/ZeroOmega}
@@
 \end{Coqsrc}
-%
-% TO do : insert succ in E0
-%
 

### 2021-08-15 doc/movies/extract_snippets/docutils.py
[movies] Remove reference to RSTCoqStandaloneReader
(It was removed it Alectryon 1.3.)

@@ -26,9 +26,7 @@ def coq_rst_to_latex(source: str) -> str:
     """
     from alectryon.docutils import LatexWriter
     from docutils.core import publish_string
-    from alectryon.docutils import \
-        RSTCoqParser as Parser, \
-        RSTCoqStandaloneReader as Reader
+    from alectryon.docutils import RSTCoqParser as Parser
 
     settings_overrides = {
         'traceback': True,
@@ -44,7 +42,6 @@ def coq_rst_to_latex(source: str) -> str:
     parser = Parser()
     return publish_string(
         source=source.encode("utf-8"),
-        reader=Reader(parser), reader_name=None,
         parser=parser, parser_name=None,
         writer=LatexWriter(), writer_name=None,
         settings=None, settings_spec=None,

### 2021-08-15 doc/movies/extract_snippets/cli.py
[movies] Change call to copy_assets for compatibility with alectryon 1.3
This is annoying oversight, sorry.  The function used to take a list of file
names and look them up in a hardcoded directory.  I generalized it to take lists
of (directory, name) pairs, but I didn't realize it had external callers.

@@ -122,11 +122,12 @@ def copy_asset(output_dir: Path, dir_name='assets'):
     from shutil import copy
 
     STY = ASSETS.ALECTRYON_STY + ASSETS.PYGMENTS_STY
+    assets = [(ASSETS.PATH, asset) for asset in STY]
 
     output_dir = output_dir / dir_name
     output_dir.mkdir(exist_ok=True)
 
-    copy_assets(0, STY, copy, output_dir)
+    copy_assets(None, assets, copy, output_dir)
     print(f"copy assets {STY} in {output_dir}")
 
 

### 2021-08-15 doc/movies/extract_snippets/docutils.py
[movies] Use new LuaLaTeX backend

@@ -24,7 +24,7 @@ def coq_rst_to_latex(source: str) -> str:
     :param source: source to conver
     :return:
     """
-    from alectryon.docutils import LatexWriter
+    from alectryon.docutils import LuaLatexWriter
     from docutils.core import publish_string
     from alectryon.docutils import RSTCoqParser as Parser
 
@@ -43,7 +43,7 @@ def coq_rst_to_latex(source: str) -> str:
     return publish_string(
         source=source.encode("utf-8"),
         parser=parser, parser_name=None,
-        writer=LatexWriter(), writer_name=None,
+        writer=LuaLatexWriter(), writer_name=None,
         settings=None, settings_spec=None,
         settings_overrides=settings_overrides, config_section=None,
         enable_exit_status=True).decode("utf-8")

### 2021-08-17 doc/movies/extract_snippets/docutils.py
[movies] Change long-line threshold to 88 to reduce warnings

@@ -12,9 +12,10 @@ def register_docutils(sertop_args):
     :param sertop_args: sertop arguments to used
     :return:
     """
-    from alectryon.docutils import setup, AlectryonTransform
-    AlectryonTransform.SERTOP_ARGS = sertop_args
-    setup()
+    import alectryon.docutils
+    alectryon.docutils.AlectryonTransform.SERTOP_ARGS = sertop_args
+    alectryon.docutils.LONG_LINE_THRESHOLD = 88
+    alectryon.docutils.setup()
 
 
 def coq_rst_to_latex(source: str) -> str:

### 2021-08-17 doc/hydras.tex
use Alectryon 1.3.1
update the doc; minor change to T1Bridge.v

@@
 \input{coqmacros}
+
+%%% One hypothesis per line 
+\makeatletter
+\renewcommand{\alectryon@hyps@sep}{\alectryon@nl}
+\makeatother
 
@@
     Many thanks to Yves Bertot, \'Evelyne Contejean, Jéremy Damour,   Florian Hatat,  David Ilcinkas, 
-Pascal Manoury,  Karl Palmskog, Sylvain Salvati, Alan Schmitt and Théo Zimmermann for their help on the elaboration of this library and  document, and to the
+Pascal Manoury,  Karl Palmskog, Cl\'ement Pit-Claudel, Sylvain Salvati, Alan Schmitt and Théo Zimmermann for their help on the elaboration of this library and  document, and to the
  members of the \emph{Formal Methods} team and the \emph{Coq working group} at laBRI for their helpful comments 

### 2021-08-17 doc/part-hydras.tex
use Alectryon 1.3.1
update the doc; minor change to T1Bridge.v

@@
 
-\input{movies/snippets/T1/LTOnea}
-
-\begin{Coqanswer}
-1 subgoal (ID 738)
-
-  alpha1 : T1
-  n : nat
-  alpha2 : T1
-  H1 : nf (ocons alpha1 n alpha2)
-  H2 : lt (ocons alpha1 n alpha2) one
-  ============================
-  ocons alpha1 n alpha2 = zero
-\end{Coqanswer}
-
-\input{movies/snippets/T1/LTOneb}
+\input{movies/snippets/T1/LTOne}
 
@@
 of the present ordinal theory with Gaia, i.e., relating the different notions of ordinals
-and transferring relevant results, would be an interesting project.
+and transferring relevant results, is an interesting project.
+First experiments in that direction are developped in
+\href{https://github.com/coq-community/hydra-battles/blob/master/theories/gaia/T1Bridge.v}{theories/gaia/T1Bridge.v}.
 

### 2021-08-17 doc/part-hydras.tex
conversion to Alectryon (to be continued)

@@
 
-\begin{enumerate}
-\item The first subgoal is as follows:
-
-\begin{Coqanswer}
-
-  h: Hydrae
-  IHh : h_forall (fun h : Hydra => height h < hsize h) h
-  ============================
-   height (node s) < hsize (node s)
-
-\end{Coqanswer}
-
-This goal is easily solvable, using some arithmetic. We let the reader look at the source.
-
-\item The second subgoal is trivial:
-
-\begin{Coqanswer}
-
-  ============================
-    h_forall (fun h : Hydra => height h < hsize h) hnil
-
-\end{Coqanswer}
-
-\begin{Coqsrc}
-  reflexivity.
-\end{Coqsrc}
-
-\item Finally, the last subgoal is also easy to solve:
-
-\begin{Coqanswer}
-  h : Hydra
-  h0 : Hydrae
-  IHh : height h < hsize h
-  IHh0 : h_forall (fun h : Hydra => height h < hsize h) h0
-  ============================
-   h_forall (fun h1 : Hydra => height h1 < hsize h1) 
-                 (hcons h h0)
-\end{Coqanswer}
-
-\begin{Coqsrc}
- split;auto. 
-Qed.   
-\end{Coqsrc}
-
-\end{enumerate}
+The first subgoal is easily solvable, using some arithmetic. We let the reader look at the source. The second and third are
+almost trivial.
 
@@
 
-More generally, we prove the following lemma:
-
 \input{movies/snippets/Omega_Small/mGea}
-
-\begin{Coqanswer}
-1 subgoal (ID 23)
-  m : Hydra -> nat
-  Hvar : Hvariant lt_wf free m
-
-  small_h := iota (m big_h) : Hydra
-  ============================
-  forall i : nat, i <= m (iota i)
-\end{Coqanswer}
 
@@
 
-\begin{Coqanswer}
-  i, j : nat
-  IHij : forall y : t, y o< (i, j) -> y o<= m (iota y)
-  ============================
-  (i, j) o<= m (iota (i, j))
-
-\end{Coqanswer}
-
 Then a case analysis on $i$ and $j$ allows us to
@@
 
-\begin{Coqsrc}
-Lemma plus_not_monotonous_l : exists alpha beta gamma : T1,
-    alpha t1< beta /\ alpha + gamma = beta + gamma.
-Proof.
-  exists 3, 5, omega;  now  compute.
-Qed.
-
-Lemma mult_not_monotonous :  exists alpha beta gamma : T1,
-      alpha t1< beta /\ alpha * gamma = beta * gamma.
-Proof.
-  exists 3, 5, omega; now compute.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/T1/notMono}
 
@@
 
-\begin{Coqsrc}
-Lemma succ_is_plus_one (alpha : T1) :  succ alpha = alpha + 1.
-Proof.
-  induction alpha as [|a IHa n b IHb].
-  (* ... *)
-\end{Coqsrc}
-
-\subsubsection{Arithmetic on Type \texttt{E0}}
+\input{movies/snippets/T1/succIsPlusOne}
+
+\subsubsection{Arithmetic on type \texttt{E0}}
 

### 2021-08-19 doc/part-hydras.tex
conversion to Alectryon (to be continued)
conversion to Alectryon (chapter 4, to be continued)

@@
 
-\begin{Coqsrc}
-Lemma plus_nf : forall a,  nf a -> forall b, nf b -> nf (a + b).
-Proof.
- (* ... *)
-
-Instance plus (alpha beta : E0) : E0.
-Proof.
-  refine (@mkord (T1.plus (@cnf alpha) (@cnf beta))_ );
-    apply plus_nf; apply cnf_ok.
-Defined.
-
-Infix "+" := plus : E0_scope.
-
-Check omega + omega.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-omega + omega
-     : E0
-\end{Coqanswer}
+\input{movies/snippets/T1/plusNf}
+
+\input{movies/snippets/E0/plusE0}
+\input{movies/snippets/E0/CheckPlus}
 
@@
 
-\begin{Coqanswer}
-1 subgoal (ID 560)
-
-  beta : T1
-  n : nat
-  gamma : T1
-  IHbeta : nf beta -> Acc LT beta
-  IHgamma : nf gamma -> Acc LT gamma
-  H : nf (ocons beta n gamma)
-  delta : T1
-  Hdelta : delta t1< ocons beta n gamma
-  ============================
-  Acc LT delta
- \end{Coqanswer}
-
 The problem comes from the hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
@@
 for instance \(\delta\) may be of the form
-\texttt{ocons $\beta'$ $p'$  $\gamma'$},
-where  \(\beta' \leq  \beta\) and  \(p' < n\).
+\texttt{ocons $\beta$ $p'$  $\gamma'$},
+where    \(p' < n\).
 Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.
@@
 
-\begin{Coqsrc}
- Let Acc_strong (alpha:T1) :=
-      forall n beta, 
-        nf (ocons alpha n beta) -> Acc LT (ocons alpha  n beta).
-\end{Coqsrc}
+\input{movies/snippets/T1/AccStrongDef}
 
@@
 
-\begin{Coqsrc}
- Lemma Acc_strong_stronger : forall alpha, 
-     nf alpha -> Acc_strong  alpha -> Acc LT  alpha.
- Proof.
-  intros alpha H H0; apply acc_imp with (phi0 alpha).
-  - repeat split; trivial.
-    + now apply lt_a_phi0_a.
-  -  apply H0;  now apply single_nf.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/T1/AccStrongStronger}
 
@@
 
-\begin{Coqsrc}
-Lemma Acc_implies_Acc_strong : 
-   forall alpha, Acc LT  alpha -> Acc_strong alpha.
-\end{Coqsrc}
-
 The proof is structured as an induction on $\alpha'$s accessibility. Let us consider
@@
 
-\begin{Coqanswer}
-  subgoal 1
-
-  alpha : T1
-  Aalpha : forall y : T1,  y t1< alpha -> Acc LT y
-  IHalpha : forall y : T1,
-       LT y alpha ->
-       forall (n : nat) (beta : T1),
-       nf (ocons y n beta) -> Acc LT (ocons y n beta)
-  ============================
-   forall (n : nat) (beta : T1),
-   nf (ocons alpha n beta) -> Acc LT (ocons alpha n beta)
-\end{Coqanswer}
+\input{movies/snippets/T1/AccImpAccStrong}
 
@@
 Our goal is still to prove accessibility of any well formed ordinal term.
-Thanks to our previous lemmas, we are almost done.
-
-\begin{Coqsrc}
-(* A (last) structural induction *)
-
-Theorem nf_Acc : forall alpha, nf alpha -> Acc LT  alpha.
-Proof.
- induction alpha.
--  intro; apply Acc_zero.
- -  intros; eapply Acc_implies_Acc_strong;auto.
-    apply IHalpha1;eauto.
-    apply nf_inv1 in H; auto. 
-Defined.
-
-Corollary T1_wf : well_founded LT.
-\end{Coqsrc}
+Thanks to our previous lemmas, we are almost done (by a simple structural induction!).
+
+\input{movies/snippets/T1/nfAcc}
+
+\input{movies/snippets/T1/T1Wf}
 
@@
 
-\begin{Coqsrc}
-
-Definition transfinite_recursor :
- forall (P:T1 -> Type),
-   (forall x:T1, 
-     (forall y:T1, nf x -> nf y ->  lt y  x -> P y) -> P x) ->
-    forall alpha:T1, P alpha.
-Proof.
- intros; apply well_founded_induction_type with LT.
- -  exact T1_wf;auto.
- - intros. apply X. intros; apply X0. repeat split;auto. 
-Defined.
-\end{Coqsrc}
+\input{movies/snippets/T1/transfiniteRecursor}
 
@@
 
-\begin{Coqsrc}
-Ltac transfinite_induction alpha :=
-  pattern alpha; apply transfinite_recursor;[ | try assumption].
-\end{Coqsrc}
+\input{movies/snippets/T1/transfiniteInduction}
 

### 2021-08-19 doc/part-hydras.tex
conversion to Alectryon (chapter 4, to be continued)

@@
 
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
+
+\input{movies/snippets/E0/InstanceEpsilon0}
+
 \label{instance-epsilon0}
-\begin{Coqsrc}
-Instance Epsilon0 : ON Lt compare.  
-(* ... *)
-\end{Coqsrc}
 
@@
 
-\begin{Coqsrc}
-Instance Epsilon0_correct :
-  ON_correct epsilon0 Epsilon0 (fun alpha => inject (cnf alpha)).
-\end{Coqsrc}
+\input{movies/snippets/Correctness_E0/injectDef}
+
+\input{movies/snippets/Correctness_E0/Epsilon0Correct}
 
@@
 
-    \begin{Coqsrc}
-    Definition SubON_same_op  `{OA : @ON A ltA  compareA}
-           `{OB : @ON B ltB  compareB}
-           {iota : A -> B} 
-           {alpha: B}
-           {_ : SubON OA OB alpha iota}
-           (f : A -> A -> A)
-           (g : B -> B -> B)
-      :=
-      forall x y,  iota (f x y) = g (iota x) (iota y).
-    \end{Coqsrc}
+\input{movies/snippets/ON_Generic/SubONSameOp}
 
@@
 
-    \begin{Coqsrc}
-    Fixpoint oplus (alpha beta : T1) : T1 :=
-      let fix oplus_aux beta {struct beta} :=
-          match alpha, beta with
-            | zero, _ => beta
-            | _,  zero => alpha
-            | ocons a1 n1 b1, ocons a2 n2 b2 =>
-              match compare a1 a2 with
-                |  Gt => ocons a1 n1 (oplus b1 beta)
-                |  Lt => ocons a2 n2 (oplus_aux b2)
-                |  Eq => ocons a1 (S (n1 + n2)%nat) (oplus b1 b2)
-              end
-          end
-      in oplus_aux beta.
-
-    Infix "o+" := oplus  (at level 50, left associativity).
-    \end{Coqsrc}
+    \input{movies/snippets/Hessenberg/oplusDef}
 
@@
 
-    \begin{Coqsrc}
-    Lemma oplus_alpha_0 (alpha : T1) : alpha o+ zero = alpha.
-    Proof.
-      destruct alpha; reflexivity.
-    Qed.
-
-    Lemma oplus_0_beta (beta : T1): zero o+ beta = beta.
-    Proof.
-      destruct beta; reflexivity.
-    Qed.
-    \end{Coqsrc}
+    \input{movies/snippets/Hessenberg/oplusNeutral}
 

### 2021-08-20 doc/part-hydras.tex
conversion to Alectryon (chapter 4, completed, starting chap. 5)

@@
 
-    \subsubsection{Boundedness}
-    If $\alpha$ and $\beta$ are both strictly  less than  $\omega^\gamma$, then so is their natural sum
-    $\alpha \oplus \beta$. This result can be proved by structural induction on $\gamma$.
-
-    \begin{Coqsrc}
-    Lemma oplus_bounded_phi0 alpha beta gamma :
-      nf alpha -> nf beta -> nf gamma ->
-      lt alpha (phi0 gamma) ->
-      lt beta (phi0 gamma) ->
-      lt (alpha o+ beta) (phi0 gamma).
-    \end{Coqsrc}
-
-    This lemma helps us
-
     \subsubsection{Commutativity, associativity}
@@
 
-    \begin{Coqsrc}
-    Lemma oplus_comm_0 : forall alpha, nf alpha ->
-         forall a b,  nf a -> nf b ->
-                      lt a alpha ->
-                      lt b alpha ->
-                      a o+ b = b o+ a.
-     Proof with eauto with T1.
-        intros alpha Halpha; transfinite_induction alpha.
-    (* rest of proof omitted *)  
-    \end{Coqsrc}
+    \input{movies/snippets/Hessenberg/oplusComm0}
 
@@
 
-    \begin{Coqsrc}
-      Lemma oplus_comm : forall alpha beta, 
-          nf alpha -> nf beta ->
-          alpha o+ beta =  beta o+ alpha.
-      Proof with eauto with T1.
-        intros alpha beta Halpha Hbeta;
-        apply oplus_comm_0 with (succ (max alpha beta)) ...  
-      (* ... *)
-    \end{Coqsrc}
-
-    The associativity of Hessenberg's sum is proved the same way.
-
-    \begin{Coqsrc}
-     Lemma oplus_assoc_0 :
-        forall alpha,
-          nf alpha ->
-          forall a b c,  nf a -> nf b -> nf c ->
-                          lt a alpha ->
-                          lt b alpha -> lt c alpha ->
-                          a o+ (b o+ c) = (a o+ b) o+ c.
-      Proof with eauto with T1.
-        intros alpha Halpha.
-        transfinite_induction alpha.
-        (* ... *)
-    \end{Coqsrc}
-
-    \begin{Coqsrc}
-     Lemma oplus_assoc : forall alpha beta gamma,
-                            nf alpha -> nf beta -> nf gamma ->
-                                        alpha o+ (beta o+ gamma) =
-                                        alpha o+ beta o+ gamma.
-     Proof with eauto with T1.
-        intros;
-        apply oplus_assoc_0 with (succ (max alpha (max beta gamma))) ...
-        (* ... *)   
-    \end{Coqsrc}
+\input{movies/snippets/Hessenberg/oplusComm}
+
+    Associativity of Hessenberg's sum is proved the same way.
+
+    \input{movies/snippets/Hessenberg/oplusAssoc0}
+    \input{movies/snippets/Hessenberg/oplusAssoc}
 
@@
 
-    \begin{Coqsrc}
-    Lemma oplus_strict_mono_LT_l (alpha beta gamma : T1) :
-      nf gamma   -> alpha  t1< beta ->
-      alpha o+ gamma  t1< beta o+ gamma.
-
-    Lemma oplus_strict_mono_LT_r (alpha beta gamma : T1) :
-      nf alpha -> beta t1< gamma ->
-      alpha o+ beta t1< alpha o+ gamma.
-    \end{Coqsrc}
+    \input{movies/snippets/Hessenberg/oplusMono}
 
@@
 
-    \begin{Coqsrc}
-    Fixpoint m (h:Hydra) : T1 :=
-      match h with head => zero
-                 | node hs => ms hs
-    end 
-    with ms (s:Hydrae) :  T1 :=
-      match s with  hnil => zero
-                  | hcons h s' => phi0 (m h) o+  ms s'
-     end.  
-    \end{Coqsrc}
+   \input{movies/snippets/Hydra_Termination/mDef}
 
@@
 
-    \begin{Coqsrc}
-    Lemma m_nf : forall h, nf (m h).
-    Proof.
-     intro h; elim h using Hydra_rect2 
-                with (P0 := fun s =>  nf (ms s)).
-     (* ... *)
-
-    Lemma ms_nf : forall s, nf (ms s).
-    Proof with auto with T1.
-    (* ... *)
-    \end{Coqsrc}
+    \input{movies/snippets/Hydra_Termination/mNf}
 
@@
 
-    \begin{Coqsrc}
-    Lemma S0_decr :
-      forall s s', S0  s s' -> ms s' t1< ms s.
-    \end{Coqsrc}
-
-    \begin{Coqsrc}
-    Lemma R1_decr : forall h h',
-                      R1 h h' -> m h' t1< m h.
-    \end{Coqsrc}
-
-    \begin{Coqsrc}
-    Lemma S1_decr n:
-      forall s s', S1 n s s' -> ms s' t1<  ms s.
-    \end{Coqsrc}
-
-    \begin{Coqsrc}
-    Lemma R2_decr n : forall h h', R2 n h h' -> m h'  t1< m h.
-    \end{Coqsrc}
-
-    \begin{Coqsrc}
-    Lemma round_decr : forall h h', h -1-> h' -> m h' t1< m h.
-    Proof.
-       destruct 1 as [n [H | H]].
-       -  now apply R1_decr.
-       -  now apply R2_decr with n.
-    Qed.
-    \end{Coqsrc}
-
-    Finally, we prove the termination of all (free) battles.
+    \input{movies/snippets/Hydra_Termination/S0Decr}
+    \input{movies/snippets/Hydra_Termination/R1Decr}
+    \input{movies/snippets/Hydra_Termination/S1Decr}
+    \input{movies/snippets/Hydra_Termination/R2Decr}
+     \input{movies/snippets/Hydra_Termination/RoundDecr}
+
+      Finally, we prove termination of all (free) battles.
 
@@
 
-    \begin{Coqsrc}
-    Global Instance HVariant : Hvariant lt_wf free var.
-    Proof.
-     split; intros; eapply round_decr; eauto.
-    Qed.
-
-    Theorem every_battle_terminates: Termination.
-    Proof. 
-      red; apply Inclusion.wf_incl with 
-             (R2 := fun h h' =>  m h t1< m h').
-       red; intros;  now apply round_decr.
-       apply Inverse_Image.wf_inverse_image, T1_wf.
-    Qed.
-    \end{Coqsrc}
+    \input{movies/snippets/Hydra_Termination/FinalThm}
 
@@
 \label{Functions:canon}
-\begin{Coqsrc}
-  Fixpoint canon alpha (i:nat) : T1 :=
-  match alpha with
-    zero => zero
-  | ocons zero 0 zero  => zero
-  | ocons zero (S k) zero  => FS k
-  | ocons gamma 0 zero => (match T1.pred gamma with
-                            Some gamma' =>
-                            match i with 0 => zero
-                                    |   S j => ocons gamma' j zero
-                            end
-                          | None =>
-                            ocons (canon gamma i) 0 zero
-                           end)
-
-  |  ocons gamma (S n) zero =>
-     (match T1.pred gamma with
-       Some gamma' =>
-       (match i with
-         0 =>  ocons gamma n zero
-       | S j => ocons gamma n (ocons gamma' j zero)
-       end)
-       | None =>
-            ocons gamma n (ocons (canon gamma  i) 0 zero)
-      end)
-   |  ocons alpha n beta => ocons alpha n (canon beta i)  
-end.
-\end{Coqsrc}
+
+\input{movies/snippets/Canon/canonDef}
 
@@
 
-\begin{Coqsrc}
-Compute pp (canon (T1.omega ^ T1.omega) 3). 
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = (_omega ^ 3)%pT1
-     : ppT1
-   \end{Coqanswer}
-
-\begin{Coqsrc}
-Compute pp (canon (T1.omega ^ T1.omega * 3) 42).  
-\end{Coqsrc}
-
-\begin{Coqanswer}
-    = (_omega ^ _omega * 2 + _omega ^ 42)%pT1
-     : ppT1
-\end{Coqanswer}
+\input{movies/snippets/Canon/canonExamples}
 
@@
 
-Proving the equality $\canonseq{\alpha+1}{i}=\alpha$ is not 
-as simple as suggested by the equations of definition~\ref{def:canonseq-math}\,.
-Nevertheless, we can prove it by  plain structural induction on $\alpha$.
-
-\begin{Coqsrc}
-  Lemma canon_succ i alpha :
-  nf alpha -> canon (succ alpha) i = alpha.
-Proof.
- revert i; induction alpha.
- (* ... *)
-\end{Coqsrc}
+For instance, the equality $\canonseq{\alpha+1}{i}=\alpha$  can be  proved by  structural induction on $\alpha$.
+
+\input{movies/snippets/Canon/canonSucc}
 

### 2021-08-21 doc/part-hydras.tex
conversion to Alectryon (chapter 5, to be continued)

@@
     So, a  natural question is `` Does there exist any strictly decreasing variant mapping
-    type \texttt{Hydra} into some interval $[0,\alpha[$ (where $\alpha <\epsilon_0$) for proving the termination of all hydra battles''.
+    type \texttt{Hydra} into some interval $[0,\alpha)$ (where $\alpha <\epsilon_0$) for proving the termination of all hydra battles''.
 The next chapter is dedicated to a formal proof that there exists no such $\alpha$, even if we consider a restriction to the set of ``standard'' battles.
@@
 
+In the present state of this library, the following specializations of \texttt{canon} are still used in some proofs or lemma statements.
+
+\input{movies/snippets/Canon/CanonS0}
+
 For instance \coq's computing facilities allow us to verify the equalities\linebreak 
@@
 
-\index{hydras}{Projects}
-\begin{project}
-Many lemmas presented in this chapter were stated and proved before the introduction of 
-the type class \texttt{ON} of ordinal notations, and in particular its  instance \texttt{Epsilon0}.
-Thus definitions and lemmas refer to the type \texttt{T1} of possibly not well-formed terms.
-This should be fixed in  a future version.
-\end{project}
+% \index{hydras}{Projects}
+% \begin{project}
+% Many lemmas presented in this chapter were stated and proved before the introduction of 
+% the type class \texttt{ON} of ordinal notations, and in particular its  instance \texttt{Epsilon0}.
+% Thus definitions and lemmas refer to the type \texttt{T1} of possibly not well-formed terms.
+% This should be fixed in  a future version.
+% \end{project}
 
@@
 \label{lemma:canonS_LT}
-\begin{Coqsrc}
-Lemma canonS_LT i alpha :
-  nf alpha -> alpha <> zero ->
-  canon alpha (S i) t1<  alpha.
-\end{Coqsrc}
+
+\input{movies/snippets/Canon/canonSLT}
 
@@
 \vspace{4pt}
+
 \emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}
 
-\begin{Coqsrc}
-Lemma canonS_limit_strong (lambda : T1) : 
-     nf lambda ->
-     limitb lambda  ->
-     forall beta, beta t1< lambda ->
-                  {i:nat | beta t1< canon lambda (S i)}.
-
-Proof.
-  transfinite_induction_LT lambda.
-  (* ... *)
-Defined.
-\end{Coqsrc}
+\input{movies/snippets/Canon/canonSLimitStrong}
 
@@
 
-\begin{Coqsrc}
-Lemma canonS_limit_lub (lambda : T1) :
-  nf lambda -> limitb lambda  ->
-  strict_lub (canonS lambda) lambda.
-\end{Coqsrc}
+\input{movies/snippets/Canon/canonSLimitLub}
 
@@
 
-\begin{Coqsrc}
-Definition transition_S i : relation T1 :=
-fun alpha beta =>
-    alpha <> zero /\ beta = canonS alpha (S i).
-
-Definition transition i : relation T1 :=
-  match i with 0 => fun _ _ => False | S j => transition_S j end.
-
-Inductive path_to (beta: T1) : list nat -> T1 -> Prop :=
-  path_to_1 : forall (i:nat) alpha , 
-    i <> 0 ->
-    transition i alpha beta ->
-    path_to beta (i::nil) alpha
-| path_to_cons : forall i alpha s gamma,
-    i <> 0 ->
-    transition i alpha gamma ->
-    path_to beta  s gamma ->
-    path_to beta  (i::s) alpha.
-\end{Coqsrc}
-
-\begin{remark}
-The definition above is parameterized with the \emph{destination} of the path and indexed by the origin, hence the name \texttt{path\_to}.
-The rationale behind this choice is a personal preference of the developer  for the kind of eliminators generated by \coq{} in this case.
-The symmetric option could have been also considered (see also Remark~\vref{remark:transitive-closure}).
-\end{remark}
+\input{movies/snippets/Paths/transitionDefs}
+
+\input{movies/snippets/Paths/pathDef}
 
@@
 (\texttt{path\_toS $\beta$ $s$ $\alpha$}) is equivalent to
-(\texttt{path\_to $\beta$ (shift $s$) $\alpha$}). This variant is scheduled to be deprecated.
+(\texttt{path\_to $\beta$ (List.map S $s$) $\alpha$}). This variant is scheduled to be deprecated.
 \end{remark}
@@
 
-\begin{Coqsrc}
-Lemma LT_path_to (alpha beta : T1) :
-  beta t1< alpha -> {s : list nat | path_to beta s alpha}.
-\end{Coqsrc}
+\input{movies/snippets/Paths/LTPathTo}
+
+\noindent 
+From the lemma \texttt{canonS\_LT}~\vref{lemma:canonS_LT}, we can convert any path into an inequality on ordinals (by induction on paths).
+
+\input{movies/snippets/Paths/pathToLT}
 
@@
 \end{exercise}
-
-\noindent 
-From the lemma \texttt{canonS\_LT}~\vref{lemma:canonS_LT}, we can convert any path into an inequality on ordinals (by induction on paths).
-
-\begin{Coqsrc}
-Lemma path_to_LT beta s alpha :
-  path_to beta s alpha -> nf alpha -> beta t1< alpha.
-\end{Coqsrc}
 
@@
 First experiments in that direction are developped in
-\href{https://github.com/coq-community/hydra-battles/blob/master/theories/gaia/T1Bridge.v}{theories/gaia/T1Bridge.v}.
+the \href{https://github.com/coq-community/hydra-battles/blob/master/theories/gaia/}{theories/gaia/} directory.
 

### 2021-08-21 doc/part-hydras.tex
conversion to Alectryon (chapter 5, half done)

@@
 
-\begin{Coqsrc}
-Fixpoint iota (alpha : T1) : Hydra :=
-  match alpha with
-  | T1.zero => head
-  | ocons gamma n beta => 
-         node (hcons_mult (iota gamma) (S n) (iotas beta))
-  end 
-with iotas (alpha : T1) :  Hydrae :=
-       match alpha with
-       | T1.zero => hnil
-       | ocons alpha0 n beta  => 
-           hcons_mult (iota alpha0) (S n) (iotas beta)
-       end.
-\end{Coqsrc}
+\input{movies/snippets/O2H/iotaDef}
 
@@
 
-\begin{Coqsrc}
-Lemma canonS_iota i alpha :
-    nf alpha -> alpha <> 0 ->
-    iota alpha -1-> iota (canon alpha (S i)).
-\end{Coqsrc}
+\input{movies/snippets/O2H/canonSIota}
 
@@
 
-\begin{Coqsrc}
-Lemma path_to_battle alpha s beta :
-  path_to  beta  s alpha -> nf alpha ->
-  iota alpha -+-> iota beta.
-\end{Coqsrc}
+\input{movies/snippets/O2H/pathToRoundPlus}
 
@@
 
-\begin{Coqsrc}
-Lemma LT_to_battle alpha beta :
-    beta t1< alpha ->  iota alpha -+-> iota beta.
-\end{Coqsrc}
+\input{movies/snippets/O2H/LTToRoundPlus}
 
@@
 
-\begin{Coqsrc}
-Class BoundedVariant {A:Type}{Lt:relation A}
-      {Wf: well_founded Lt}{B : Battle}
-      {m: Hydra -> A} (Var: Hvariant  Wf  B m)(mu:A):=
-  {
-  m_bounded: forall h, Lt (m h ) mu
-  }.
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Definitions/BoundedVariant}
 
@@
 
-\begin{Coqsrc}
-Context (B: Battle)
-          (mu: T1)
-          (Hmu: nf mu)
-          (m : Hydra -> T1)
-          (Var : Hvariant  T1_wf B m)
-          (Hy : BoundedVariant  Var mu).
-
-  Hypothesis m_decrease : forall  i h h',
-      round_n i h h'   -> m h' t1< m h.
-\end{Coqsrc}
+\input{movies/snippets/Epsilon0_Needed_Generic/theContext}
 

### 2021-08-22 doc/part-hydras.tex
conversion to Alectryon (chapter 5, more than half done)

@@
 
-\begin{Coqsrc}
-Lemma m_ge_0 alpha:  nf alpha -> alpha t1<= m (iota alpha).
-\end{Coqsrc}
+\input{movies/snippets/Epsilon0_Needed_Generic/mGe0}
 
@@
 
-\begin{Coqsrc}
-  Definition big_h := iota mu.
-  Definition beta_h := m big_h.
-  Definition small_h := iota beta_h.
-\end{Coqsrc}
-
-From Lemma \texttt{m\_ge\_0} we infer the following inequality :
-
-\begin{Coqsrc}
-    Corollary m_ge_generic : m big_h t1<= m small_h.
- \end{Coqsrc}
+\input{movies/snippets/Epsilon0_Needed_Generic/mGeGeneric}
 
@@
 
-\begin{Coqsrc}
-Section Impossibility_Proof.
-
-  Context (mu: T1)
-          (Hmu: nf mu)
-          (m : Hydra -> T1)
-          (Var : Hvariant  T1_wf free m)
-          (Hy : BoundedVariant Var mu).
-  \end{Coqsrc}
+\input{movies/snippets/Epsilon0_Needed_Free/theContext}
 
@@
 
-\begin{Coqsrc}
-Lemma m_ge : m big_h t1<= m small_h.
-  Proof.
-    apply m_ge_generic.
-   (* ... *)
-\end{Coqsrc}
-
-\item From the hypothesis \texttt{m\_bounded}, we have \texttt{m big\_h t1< mu}
-\item By Lemma \texttt{LT\_to\_battle}, we get a (free) battle from
+\input{movies/snippets/Epsilon0_Needed_Free/mGe}
+
+\item From the hypothesis \texttt{Hy}, we have \texttt{m big\_h t1< mu}
+\item By Lemma \texttt{LT\_to\_round\_plus}, we get a (free) battle from
 \texttt{big\_h = iota mu} to \texttt{small\_h = iota (m big\_h)}.
 
-\begin{Coqsrc}
-  Lemma  big_to_small : big_h  -+-> small_h.
-\end{Coqsrc}
+\input{movies/snippets/Epsilon0_Needed_Free/bigToSmall}
+
 \item From the hypotheses on $m$, we infer:
-
-\begin{Coqsrc}
-Lemma m_lt : m small_h t1< m big_h.
-\end{Coqsrc}
+\input{movies/snippets/Epsilon0_Needed_Free/mLt}
 
 \item From lemmas \texttt{m\_ge} and \texttt{m\_lt}, and the irreflexivity of $<$, we get a contradiction. 
-
-  \begin{Coqsrc}
-Theorem Impossibility_free : False.
-
-End Impossibility_Proof.
-\end{Coqsrc}
+\input{movies/snippets/Epsilon0_Needed_Free/ImpossibilityFree}
 
@@
 
-\begin{Coqsrc}
-Check Impossibility_free.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-Impossibility_free
-     : forall (mu : T1) (m : Hydra -> T1) (Var : Hvariant T1_wf free m),
-       BoundedVariant Var mu -> False
-\end{Coqanswer}
-%%% ICI
+\input{movies/snippets/Epsilon0_Needed_Free/CheckDemo}
 

### 2021-08-22 doc/part-hydras.tex
conversion to Alectryon (chapter 5, more than half done)

@@
 
-\begin{Coqsrc}
-Lemma LT_to_round_plus alpha beta :
-    beta t1< alpha ->  iota alpha -+-> iota beta.
-\end{Coqsrc}
+\input{movies/snippets/O2H/LTToRoundPlus}
 
@@
 
-\begin{Coqsrc}
-Lemma LT_path_to (alpha beta : T1) :
-  beta t1< alpha -> {s : list nat | path_to beta s alpha}.
-\end{Coqsrc}
-
-The sequence $s$, used to build the sequence of replication factors of the battle depends on 
+\input{movies/snippets/Paths/LTPathTo}
+
+The sequence $s$, used to build the sequence of replication factors of the battle, depends on 
 $\beta$, so we cannot be sure that the generated battle is a genuine standard battle.
@@
 
-\begin{Coqsrc}
-(**  standard path from (i, alpha) to (j, beta) *)
-
-Inductive standard_pathR(j:nat)( beta:T1):  nat -> T1 -> Prop :=
-  std_1 : forall i alpha, 
-       beta = canon alpha i -> j = S i ->
-       standard_pathR j beta i  alpha
-| std_S : forall i alpha, 
-      standard_pathR j beta (S i) (canon alpha i)  ->
-      standard_pathR j beta i alpha.
-
-Definition standard_path  i alpha j beta := 
-   standard_pathR j beta i alpha.
-\end{Coqsrc}
+\input{movies/snippets/Paths/standardPathR}
 
@@
 
-\begin{Coqsrc}
-Definition const_pathS i :=
-  clos_trans_1n T1 (fun alpha beta => alpha <> zero /\
-                                      beta = canon alpha (S i)).
-
-Definition const_path i alpha beta :=
-  match i with
-    0 => False
-  | S j => const_pathS j alpha beta
-end.
-\end{Coqsrc}
+\input{movies/snippets/Paths/constPathDef}
 

### 2021-08-23 doc/movies/extract_snippets.py
[movies] Propagate error code returned by Alectryon

@@ -1,6 +1,6 @@
 #!/usr/bin/env python3
 from extract_snippets.cli import main
-
+import sys
 
 if __name__ == '__main__':
-    main()
+    sys.exit(main())

### 2021-08-23 doc/movies/extract_snippets/docutils.py

@@ -22,12 +22,13 @@ def coq_rst_to_latex(source: str) -> str:
     """
     convert coq content to latex (please setup doctutils with register_docutils before)
     (function inspired by `_gen_docutils` in alectryon cli)
-    :param source: source to conver
-    :return:
+    :param source: source to convert
+    :return: Exit code (int) and output (str).
     """
-    from alectryon.docutils import LuaLatexWriter
     from docutils.core import publish_string
-    from alectryon.docutils import RSTCoqParser as Parser
+    from docutils.readers.standalone import Reader
+    from alectryon.cli import _gen_docutils
+    from alectryon.docutils import LuaLatexWriter, RSTCoqParser as Parser
 
     settings_overrides = {
         'traceback': True,
@@ -38,17 +39,15 @@ def coq_rst_to_latex(source: str) -> str:
         'alectryon_vernums': False,
         'input_encoding': 'utf-8',
         'output_encoding': 'utf-8',
+        'exit_status_level': 3,
     }
 
-    parser = Parser()
-    return publish_string(
-        source=source.encode("utf-8"),
-        parser=parser, parser_name=None,
-        writer=LuaLatexWriter(), writer_name=None,
-        settings=None, settings_spec=None,
-        settings_overrides=settings_overrides, config_section=None,
-        enable_exit_status=True).decode("utf-8")
+    output, _pub, exit_code = \
+        _gen_docutils(source, None,
+                      Parser, Reader, LuaLatexWriter,
+                      settings_overrides)
 
+    return exit_code, output
 
 class CoqToLatexReader(Reader):
     DOCUMENT_BEGIN = LATEX_TAG_BEGIN_FORMAT.format(tag="document")
@@ -57,7 +56,7 @@ class CoqToLatexReader(Reader):
     def __init__(self, path_coq: Path):
         with open(path_coq, 'r') as file:
             self.content_coq = file.read()
-        self.content_latex = coq_rst_to_latex(self.content_coq)
+        self.exit_code, self.content_latex = coq_rst_to_latex(self.content_coq)
 
     def __iter__(self):
         start = False

### 2021-08-23 doc/part-hydras.tex
conversion to Alectryon (chapter 5, more than half done)

@@
 
-\begin{Coqsrc}
-Corollary Cor12 (alpha : T1) :  nf alpha ->
-         forall beta i n, beta  t1< alpha  ->
-                i < n ->
-                 const_pathS i alpha beta ->
-                 const_pathS n alpha beta.
-Proof.
-  transfinite_induction_lt alpha.
-  (* (long) proof skipped *)
-\end{Coqsrc}
+\emph{From
+  Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
+
+\input{movies/snippets/Paths/Cor12}
 
@@
 
-\begin{Coqsrc}
-Corollary Cor12_1 (alpha : T1) :
-  nf alpha ->
-  forall beta i n, beta t1< alpha ->
-                   i <= n ->
-                   const_pathS i alpha beta ->
-                   const_pathS n alpha beta.
-\end{Coqsrc}
+\input{movies/snippets/Paths/Cor121}
 

### 2021-08-23 doc/part-hydras.tex
moving to Alectryon: chapter 5 completed

@@
 
-\begin{Coqsrc}
-Theorem KS_thm_2_4 (lambda : T1) :
-   nf lambda ->
-   limitb lambda  ->
-   forall i j, (i < j)%nat ->
-               const_pathS 0 (canon lambda (S j))
-                             (canon lambda (S i)).
-\end{Coqsrc}
+\input{movies/snippets/Paths/KSThm24}
 
@@
 
-\begin{Coqsrc}
-Lemma Lemma2_6_1 (alpha : T1) :  
-  nf alpha -> forall beta,  beta t1< alpha  ->
-  {n:nat | const_pathS n alpha beta}.
-Proof.
-  transfinite_induction alpha.
-  (* ... *)
-\end{Coqsrc}
-
-Intuitively, lemma   \texttt{L2\_6\_1}  shows that if $\beta<\alpha<\epsilon_0$, then there exists  a battle from $\iota(\alpha)$ to $\iota(\beta)$ where the replication factor is constant, although large enough.
+\input{movies/snippets/Paths/Lemma261}
+
+Intuitively, lemma   \texttt{Lemma2\_6\_1}  shows that if $\beta<\alpha<\epsilon_0$, then there exists  a battle from $\iota(\alpha)$ to $\iota(\beta)$ where the replication factor is constant, although large enough.
 
 \subsection{Casting paths with a constant index into a standard path}
-
-%%% traduire la v.f.  (voir %%% A traduire %%%% )
 
@@
 
-\begin{Coqsrc}
-(* Lemma 1 page 300 of [KS] *)
-
-Lemma constant_to_standard_path 
-  (alpha beta : T1) (i : nat):
-  nf alpha -> const_pathS i alpha beta -> zero  t1< alpha ->
-  {l:nat | standard_path (S i) alpha j beta}.
-\end{Coqsrc}
+\input{movies/snippets/Paths/constantToStandard}
 
@@
 
-\begin{Coqsrc}
-Fixpoint standard_gnaw (i:nat)(alpha : T1)(l:nat):  T1  :=
-  match l with
-  | 0 => alpha
-  | S m => standard_gnaw (S i) (canon alpha i) m
-  end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-  Compute standard_gnaw 2 omega 15.
-(*   = zero
-     : T1 *)
-Compute pp (standard_gnaw 2 (omega^omega)  10).
-(*
-= (omega + 7)%pT1
-     : ppT1
-*)
-Compute pp (standard_gnaw 4 (omega^omega)  100).
-(*
- = (omega ^ 3 * 4 + omega ^ 2 * 5 + omega * 3 + 39)%pT1
-     : ppT1 *)
-\end{Coqsrc}
+\input{movies/snippets/Paths/standardGnaw}
 
@@
 
-\begin{Coqsrc}
-Lemma standard_path_to_zero :
-  forall  alpha i, nf alpha ->
-                   {j: nat | standard_path (S i) alpha j zero}.
-\end{Coqsrc}
+\input{movies/snippets/Paths/standardPathToZero}
 
 \paragraph*{}
-Noq, let us consider two ordinals  $\beta<\alpha<\epsilon_0$.  Let $p$  be some $(n+1)$-path from $\alpha$ to $\beta$.
-
-\begin{Coqsrc}
- Section Constant_to_standard_Proof.
-
-  Variables (alpha beta: T1) (n : nat).
-  Hypotheses (Halpha: nf alpha) (Hpos : zero t1<  beta)
-             (p : const_pathS n alpha  beta).
-\end{Coqsrc}
+Now, let us consider two ordinals  $\beta<\alpha<\epsilon_0$.  Let $p$  be some $(n+1)$-path from $\alpha$ to $\beta$.
+
+\input{movies/snippets/Paths/ConstantToStandardProof}
 
@@
 
-\begin{Coqsrc}
- Lemma constant_to_standard_0 : 
-    {l : nat | standard_fun (S n) alpha l = beta}.
- (* ... *)
-
-End Constant_to_standard_Proof.
-\end{Coqsrc}
+\input{movies/snippets/Paths/constantToStandard0}
 
@@
 
-\begin{Coqsrc}
-Lemma constant_to_standard_path 
-  (alpha beta : T1) (i : nat):
-  nf alpha -> const_pathS i alpha beta -> zero  t1< alpha ->
-  {j:nat | standard_path (S i) alpha j beta}.
-\end{Coqsrc}
+\input{movies/snippets/Paths/constantToStandardPath}
 
@@
 
-\begin{Coqsrc}
-Corollary  LT_to_standard_path  (alpha beta : T1) :
-  beta t1< alpha ->
-  {n : nat & {j:nat | standard_path (S n) alpha j beta}}.
-\end{Coqsrc}
+\input{movies/snippets/Paths/LTToStandardPath}
 
@@
 
-\begin{Coqsrc}
-Lemma LT_to_standard_battle :
-    forall alpha beta,
-      beta t1< alpha ->
-      exists n i,  battle standard  n (iota alpha) i (iota beta).
-\end{Coqsrc}
+\input{movies/snippets/Epsilon0_Needed_Std/LTToStandardBattle}
 
@@
 
-\begin{Coqsrc}
-Section Impossibility_Proof.
-
-Context (mu: T1)
-          (Hmu: nf mu)
-          (m : Hydra -> T1)
-          (Var : Hvariant  T1_wf standard m)
-          (Hy : BoundedVariant Var mu).
- \end{Coqsrc}
+\input{movies/snippets/Epsilon0_Needed_Std/theContext}
 
@@
 
-\begin{Coqsrc}
- Lemma m_ge : m big_h t1<= m small_h.
-\end{Coqsrc}
+\input{movies/snippets/Epsilon0_Needed_Std/mGe}
 
@@
 
-\begin{Coqsrc}
-  Lemma m_lt : m small_h  t1< m big_h.
-
-  Theorem Impossibility_std: False.
-
-End Impossibility_Proof.
-\end{Coqsrc}
+\input{movies/snippets/Epsilon0_Needed_Std/mLt}
 
@@
 Since $m$ is assumed to be a variant for standard battles, we get the inequality  $m(\texttt{small\_h}) < m(\texttt{big\_h})$.
+
+\input{movies/snippets/Epsilon0_Needed_Std/endOfProof}
 

### 2021-08-24 doc/part-hydras.tex
Alectryon: 50% of chapter6

@@
 
-\begin{Coqsrc}
-Fixpoint gnaw (alpha : T1) (s: list nat) :=
-  match s with
-    | nil => alpha
-    | (0::s') => gnaw  alpha s'
-    | (S i :: s')  =>  gnaw (canon alpha (S i)) s'
-  end.
-
-Definition large alpha (s:list nat) := gnaw alpha s = zero.
-\end{Coqsrc}
+\input{movies/snippets/Paths/gnawDef}
+
+\noindent
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_SetsPaths.html\#gnaw}{Epsilon0.Large\_Sets}}
+
+\input{movies/snippets/Large_Sets/largeDef}
 
@@
 \index{hydras}{Library Epsilon0!Predicates!mlarge@mlarge (minimal large sequences)}
-\begin{Coqsrc}
-Definition mlarge alpha (s:list nat) := path_to zero s alpha.
-\end{Coqsrc}
+
+\input{movies/snippets/Large_Sets/largeDef}
 
@@
 
-\begin{Coqsrc}
-Compute pp (gnaw (T1.omega * T1.omega) (interval 6 70)).
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = (_omega * 2 + 56)%pT1
-     : ppT1
-\end{Coqanswer}
+\noindent
+\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html\#mlarge}{Epsilon0.Large\_Sets\_Examples}}
+
+\input{movies/snippets/Large_Sets_Examples/gnawEx1}
 
 Let us try another computation.
-
-\begin{Coqsrc}
-Compute (gnaw (T1.omega * T1.omega) (interval 6 700)).
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = zero : T1
-\end{Coqanswer}
+\input{movies/snippets/Large_Sets_Examples/gnawEx2}
 
@@
 
-The following lemma relates minimal largeness with the function 
+The following lemma relates minimal largeness with the
+function 
 \texttt{gnaw}.
 
-\begin{Coqsrc}
-Lemma mlarge_iff alpha x (s:list nat) :
-  s <> nil -> ~ In 0 (x::s) ->
-  mlarge alpha (x::s) <-> gnaw alpha (but_last x s) = one.
- \end{Coqsrc}
+\input{movies/snippets/Large_Sets/mlargeIff}
 
@@
 \emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html}{Epsilon0.Large\_Sets\_Examples}}
-\begin{Coqsrc}
-Example Ex1 : mlarge (T1.omega * T1.omega) (interval 6 510).
-\end{Coqsrc}
+
+\input{movies/snippets/Large_Sets_Examples/Ex1Lemma}
 
@@
 \emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-\begin{Coqsrc}
-Lemma mlarge_unicity alpha k l l' : 
-  mlarge alpha (interval (S k) l) ->
-  mlarge alpha (interval (S k) l') ->
-  l = l'.
-\end{Coqsrc}
-
-Thus, it seems obvious that there must exist a function, parameterized by $\alpha$ which associates to any  strictly positive integer $k$ the number $l$ such that
+
+\input{movies/snippets/Large_Sets/mlargeUnicity}
+
+Thus, we would like to define a function, parameterized by $\alpha$ which associates to any  strictly positive integer $k$ the number $l$ such that
 the interval $[k,l]$ is $\alpha$-mlarge. It would be fine to write in \gallina{} a definition like this:
@@
 
-\begin{Coqsrc}
-Inductive L_spec : T1 -> (nat -> nat) -> Prop :=
-| L_spec0 : forall f, (forall k, f k = k) ->  L_spec zero f
-| L_spec1 : forall alpha f,
-    alpha <> zero ->
-    (forall k, mlarge alpha (interval (S k) (Nat.pred (f (S k))))) ->
-    L_spec alpha f.
-\end{Coqsrc}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
+
+\input{movies/snippets/Large_Sets/LSpecDef}
 
@@
 
-\begin{Coqsrc}
-Lemma L_spec_unicity alpha f g :
-  L_spec alpha f -> L_spec alpha g -> forall k, f (S k) = g (S k).
-\end{Coqsrc}
+\input{movies/snippets/Large_Sets/LSpecUnicity}
 
@@
 
-\begin{Coqsrc}
-Lemma L_zero_inv f : L_spec zero f -> forall k, f (S k) = S k.
-\end{Coqsrc}
+\input{movies/snippets/Large_Sets/LZeroInv}
 
@@
 Note that our decomposition of intervals works only if the intervals we consider are not empty.
-In order to ensure this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <<= f}, or \texttt{(fun\_le S f)} (defined
-in~\href{../theories/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}).
-
-\begin{Coqsrc}
-Definition fun_le f g  := forall n:nat, f n <=  g n.
-Infix "<<=" := fun_le (at level 60).
-\end{Coqsrc}
-
-It looks also natural to show that the functions we consider are strictly monotonous. The section on successor ordinals has thus the following structure.
-
-\begin{Coqsrc}
-Section succ.
-   Variables (beta : T1) (f : nat -> nat).
-
-   Hypotheses (Hbeta : nf beta)
-              (f_mono : strict_mono f)
-              (f_Sle : S <<= f)
-              (f_ok : L_spec beta f).
-
-   Definition L_succ := fun k => f (S k).
-
-   Lemma L_succ_mono : strict_mono L_succ.
-
-   Lemma L_succ_Sle : S <<= L_succ.
-
-   Lemma L_succ_ok : L_spec (succ beta) L_succ.
-
-End succ.
-
-\end{Coqsrc}
+In order to ensure this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <<= f}, or \texttt{(fun\_le S f)}.
+
+Useful abstract properties of arithmetic functions are
+defined 
+in~\href{../theories/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}.
+
+\label{sect:abstract-arith-prop}
+
+\input{movies/snippets/Iterates/funLeDef}
+
+We prove also that the functions we consider are strictly monotonous. The section on successor ordinals has the following structure.
+
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
+
+\input{movies/snippets/Large_Sets/SectionSucc}
 
@@
 
-\begin{Coqsrc}
-Section lim.
-  Variables (lambda : T1)
-            (Hnf : nf lambda)
-            (Hlim : limitb lambda)
-            (f : nat -> nat -> nat)
-            (H : forall k, L_spec (canon lambda (S k)) (f (S k))).
-
-  Let L_lim k := f k (S k).
-
-  Lemma L_lim_ok : L_spec lambda L_lim.
-
-End lim.
-\end{Coqsrc}
+\input{movies/snippets/Large_Sets/SectionLim}
 
@@
 
-\begin{Coqsrc}
-Definition L_fin i := fun k => (i + k)%nat.
-
-Lemma L_fin_ok i : L_spec (fin i) (L_fin i).
-\end{Coqsrc}
+\input{movies/snippets/Large_Sets/LFinDef}
+\vspace{-16pt}
+\input{movies/snippets/Large_Sets/LFinOk}
 
@@
 
-\begin{Coqsrc}
-Definition L_omega k := S (2 * k)%nat.
-
-Lemma L_omega_ok : L_spec T1.omega L_omega.
-\end{Coqsrc}
+\input{movies/snippets/Large_Sets/LOmegaDef}
+\vspace{-16pt}
+\input{movies/snippets/Large_Sets/LOmegaOk}
 
@@
 
-\begin{Coqsrc}
-Lemma path_to_omega_mult (i k:nat) :
-path_to (T1.omega * i) (interval (S k) (2 * (S k)))
-            (T1.omega * (S i)).
-\end{Coqsrc}
+\input{movies/snippets/Large_Sets/pathToOmegaMult}
 
@@
 
-\begin{Coqsrc}
-Lemma omega_mult_mlarge_0 i  : forall k,
-    mlarge  (T1.omega * (S i))
-            (interval (S k)
-                      (Nat.pred (iterate (fun p =>  S (2 * p)%nat)
-                                         (S i)
-                                         (S k)))).
-\end{Coqsrc}
-
-Thus, we infer the following result:
+\input{movies/snippets/Large_Sets/omegaMultMlarge0}
 
@@
 
-\begin{Coqsrc}
-Definition L_omega_mult i (x:nat) :=  iterate L_omega i x.
-
-Lemma L_omega_mult_ok (i: nat) :  L_spec (T1.omega * i) (L_omega_mult i).
-\end{Coqsrc}
-
-For instance, let us consider the ordinal $\omega\times 8$, and a sequence 
-starting at $k=5$.
-
-\begin{Coqsrc}
-Compute L_omega_mult 8 5.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-= 1535
-     : nat
-\end{Coqanswer}
+\input{movies/snippets/Large_Sets/LOmegaMultDef}
 
@@
 
-\begin{Coqsrc}
-Lemma L_omega_mult_eqn (i : nat) :
-  forall (k : nat),  (0 < k)%nat  ->
-                     L_omega_mult i k = (exp2 i * S k - 1)%nat.
-\end{Coqsrc}
+\input{movies/snippets/Large_Sets/LOmegaMultEqn}
+
+Correctness of the function \texttt{L\_omega\_mult} is asserted through the following lemma.
+
+\input{movies/snippets/Large_Sets/LOmegaMultOk}
 
@@
 
-These properties are defined in \coq{} in the library
-~\href{../theories/html/hydras.Prelude.Iterates.html}{Prelude.Iterates}.
-
 \index{maths}{Abstract properties of arithmetic functions}
@@
 
-\begin{Coqsrc}
-(** ** Abstract properties of arithmetic functions *)
-
-Definition strict_mono f := forall n p,  n < p -> f n < f p.
-
-Definition strict_mono1 f := forall n p,  0 < n < p -> f n < f p.
-
-Definition dominates_from n g f  := forall p, n <= p -> f p < g p.
-
-Definition dominates_strong g f  := {i : nat | dominates_from i g f}.
-
-Definition dominates g f := exists i : nat, dominates_from i g f .
-
-Definition fun_le f g  := forall n:nat,  f n <= g n.
-
-Infix ">>" := dominates (at level 60).
-
-Infix ">>s" := dominates_strong (at level 60).
-
-Infix "<<=" := fun_le (at level 60).
-\end{Coqsrc}
-
 \index{maths}{Transfinite induction}
@@
 For each $\alpha$, the three cases : $\alpha=0$, $\alpha$ is a limit, and 
-$\alpha$ is a successor are considered. Inside each case, the five sub-properties are proved sequentially.
+$\alpha$ is a successor are considered. Inside each case, the five sub-properties are proved sequentially, using the abstract properties defined in Sect~\vref{sect:abstract-arith-prop}
 

### 2021-08-24 doc/part-hydras.tex
Alectryon: 75% of chapter6

@@
 
-\begin{Coqsrc}
-Definition L_omega_square k := iterate (fun z => S (2 * z)%nat)
-                                        k
-                                        (S k).
-
-Lemma L_omega_square_eqn k :
-  (0 < k)%nat ->
-  L_omega_square k = (exp2 k * (k + 2) - 1)%nat.
-
-Lemma L_omega_square_ok: L_spec (T1.omega * T1.omega) 
-          L_omega_square.
-
-Compute L_omega_square 8.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = 2559
-     : nat
-\end{Coqanswer}
+\input{movies/snippets/Large_Sets/LOmegaSquare}
+\input{movies/snippets/Large_Sets/LOmegaSquareEqn}
+\input{movies/snippets/Large_Sets/LOmegaSquareOk}
 
@@
 
-\begin{Coqsrc}
-Section phi0_mult.
- Variables (alpha : T1) (f : nat -> nat).
- Hypotheses (Halpha : nf alpha)
-            (f_mono : strict_mono f)
-            (f_Sle : S <<= f)
-            (f_ok : L_spec (phi0 alpha) f).
-
- Definition L_phi0_mult i := iterate f i.
-
-Lemma L_phi0_mult_ok i: 
-  L_spec (ocons alpha i zero)  (L_phi0_mult (S i)).
-
- Lemma L_phi0_mult_smono i: strict_mono (L_phi0_mult i).
-
- Lemma L_phi0_mult_Sle i: S <<=  L_phi0_mult (S i).
-
-End phi0_mult.
-\end{Coqsrc}
-
-Let us look 
-at the ordinal $\omega^2\times i$, using \texttt{L\_phi0\_mult}
-
-\begin{Coqsrc}
- Definition L_omega_square_times i :=  iterate L_omega_square i.
-
- Lemma L_omega_square_times_ok i : 
-    L_spec (ocons 2 i zero) (L_omega_square_times (S i)).
- Proof.
-  apply L_phi0_mult_ok.
-  -  auto with T1.
-  -  apply L_omega_square_Sle.
-  -  apply L_omega_square_ok.
- Qed.
-\end{Coqsrc}
-
-We are now ready to get an exact formula for $L_{\omega^3}$. 
-\begin{Coqsrc}
-Definition L_omega_cube  := L_lim  L_omega_square_times .
-
-Lemma L_omega_cube_ok : L_spec (phi0 3) L_omega_cube.
-\end{Coqsrc}
-
-The function  $L_{\omega^3}$  is just obtained by diagonalization upon $L_{\omega^2\times i}$.
-
-\begin{Coqsrc}
-Lemma L_omega_cube_eqn i : 
-   L_omega_cube i = L_omega_square_times i (S i).
-Proof. reflexivity. Qed.
-\end{Coqsrc}
+\input{movies/snippets/Large_Sets/phi0Mult}
+
+Let us look now
+at the ordinal $\omega^2\times i$, using \texttt{L\_phi0\_mult}.
+
+\input{movies/snippets/Large_Sets/LOmegaSquareTimes}
+
+We are now ready to get an exact formula for $L_{\omega^3}$, by diagonalization upon $L_{\omega^2\times i}$.
+
+\input{movies/snippets/Large_Sets/LOmegaCube}
 
@@
 
-\begin{Coqsrc}
-Lemma L_omega_cube_3_eq:
-   let N := exp2 95 in
-   let P := (N * 97 - 1)%nat in
-   L_omega_cube 3  =  (exp2 P * (P + 2) - 1)%nat.
-\end{Coqsrc}
+\input{movies/snippets/Large_Sets/LOmegaCube3Eq}
 
@@
  \index{hydras}{Library Epsilon0!Functions!L\_@L\_ (final step of a minimal path}
-\begin{Coqsrc}
-From Equations Require Import Equations.
-Require Import ArithRing Lia.
-
-Instance Olt : WellFounded Lt := Lt_wf.
-Hint Resolve Olt : E0.
-
-(** Using Coq-Equations for building a function which satisfies 
-    Large_sets.L_spec *)
-
-Equations  L_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
-  L_ alpha  i with E0_eq_dec alpha Zero :=
-    { | left _ =>  i ;
-      | right nonzero
-          with Utils.dec (Limitb alpha) :=
-          { | left _ =>  L_ (Canon alpha i)  (S i) ;
-            | right notlimit =>  L_ (Pred alpha) (S i)}}.
-
-Solve All Obligations with auto with E0.
-\end{Coqsrc}
-
-It is worth looking at the answer from \texttt{Equations} and at all the lemmas this plug-in gives you for free. We show here only a part of \coq's answer.
-
-\begin{Coqanswer}
-L__obligations_obligation_1 is defined
-L__obligations_obligation_2 is defined
-L__obligations is defined
-L__clause_1 is defined
-L__functional is defined
-L_ is defined
-...
-L__equation_1 is defined
-L__graph_mut is defined
-L__graph_rect is recursively defined
-L__graph_correct is defined
-L__elim has type-checked, generating 1 obligation
-L__elim is defined
-FunctionalElimination_L_ is defined
-FunctionalInduction_L_ is defined
-\end{Coqanswer}
-
-Sometimes, these automatically generated statements may look cryptic.
-
-\begin{Coqsrc}
-About L__equation_1.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-L__equation_1 :
-forall (alpha : E0) (i : nat),
-L_ alpha i = L__unfold_clause_1 alpha (E0_eq_dec alpha Zero) i
-\end{Coqanswer}
+
+\input{movies/snippets/L_alpha/LDef}
+
+This definition results in a bunch of automatically generated lemmas. For instance:
+
+\input{movies/snippets/L_alpha/AboutLEquation1}
 
@@
 
-\begin{Coqsrc}
-Lemma L_zero_eqn : forall i, L_ Zero i = i.
-Proof. intro i; now rewrite L__equation_1. Qed.
-
-Lemma L_lim_eqn alpha i : Limitb alpha -> L_ alpha i =
-                                        L_ (Canon alpha i) (S i).
-
-Lemma L_succ_eqn alpha i :  L_ (Succ alpha) i = L_  alpha (S i).
-
-Hint Rewrite L_zero_eqn L_succ_eqn : L_rw.
-\end{Coqsrc}
+\input{movies/snippets/L_alpha/Paraphrases}
 

### 2021-08-25 doc/part-hydras.tex
Alectryon: 75% of chapter6
Alectryon: 85% of chapter6

@@
 
-\begin{Coqsrc}
-Lemma L_finite : forall i k :nat,  L_ i k = (i+k)%nat.  
-(* Proof by induction on i, using L_zero_eqn and L_succ_eqn *)
-
-Lemma L_omega : forall k, L_ omega%e0 k = S (2 * k)%nat.
-(* Proof using L_finite and L_lim_eqn *)
-\end{Coqsrc}
+\input{movies/snippets/L_alpha/LFiniteOmega}
 
@@
 
-\begin{Coqsrc}
-Lemma L_ge_S alpha : alpha <> Zero -> S <<= L_ alpha.
-
-Theorem L_correct alpha : L_spec (cnf alpha) (L_ alpha).
-\end{Coqsrc}
+\input{movies/snippets/L_alpha/LGeS}
+
+\input{movies/snippets/L_alpha/LCorrect}
 
@@
 
-Theorem \texttt{battle\_length\_std } of
-Module~\href{../theories/html/hydras.Hydra.Hydra_Theorems.html\#battle_length_std}{Hydra.Hydra\_Theorems} relates the length of standard battles with the functions $L_\alpha$.
-
-\begin{Coqsrc}
-Theorem battle_length_std (alpha : E0)  :
-  alpha <> Zero ->
-  forall k, (1 <= k)%nat ->
-            battle_length standard k (iota (cnf alpha))
-                         (L_ alpha (S k) - k).
-\end{Coqsrc}
+Lemma \texttt{battle\_length\_std } of
+Module~\href{../theories/html/hydras.Hydra.Battle_length.html}{Hydra.Battle\_length} relates the length of standard battles with the functions $L_\alpha$.
+
+\input{movies/snippets/Battle_length/battleLengthStd}
 
@@
 
-\begin{Coqsrc}
-Equations H'_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
-  H'_ alpha  i with E0_eq_dec alpha Zero :=
-    { | left _ =>  i ;
-      | right nonzero
-          with Utils.dec (Limitb alpha) :=
-          { | left _ =>  H'_ (Canon alpha (S i))  i ;
-            | right notlimit =>  H'_ (Pred alpha) (S i)}}.
-
-Solve All Obligations with auto with E0.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma H'_eq1 : forall i, H'_ Zero i = i.
-Proof.
-  intro i; now rewrite H'__equation_1. 
-Qed.
-
-Lemma H'_eq2 alpha i : Succb alpha ->
-                       H'_ alpha i = H'_ (Pred alpha) (S i).
-
-Lemma H'_eq3 alpha i : Limitb alpha ->
-                      H'_ alpha i =  H'_ (Canon alpha (S i)) i.
-
-Lemma H'_succ_eqn  alpha i :  
-            H'_ (Succ alpha) i = H'_ alpha (S i).
-\end{Coqsrc}
+\input{movies/snippets/Hprime/HprimeDef}
+
+\input{movies/snippets/Hprime/paraphrases}
 
@@
 
-\begin{Coqsrc}
-Lemma H'_Fin : forall i k : nat,  H'_  (Fin i) k = (i+k)%nat.
-Proof with eauto with E0.
-  induction i.
-  - intros; simpl Fin; simpl; autorewrite with H'_rw E0_rw ... 
-  - intros ;simpl; autorewrite with H'_rw E0_rw ... 
-    rewrite IHi; abstract lia. 
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/Hprime/HprimeFin}
 
@@
 
-\begin{Coqsrc}
-Lemma H'_omega : forall k, H'_ omega k = S (2 * k)%nat.
-Proof with auto with E0.
-  intro k; rewrite H'_eq3 ...
-  - replace (Canon omega (S k)) with (Fin (S k)).
-    + rewrite H'_Fin; abstract lia.
-    + now autorewrite with E0_rw.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/Hprime/HprimeOmega}
 
@@
 
-\begin{Coqsrc}
-Lemma H'_Plus_Fin alpha : forall i k : nat,
-    H'_ (alpha + i)%e0 k = H'_ alpha (i + k)%nat.
-(* Proof by induction on i *)
-\end{Coqsrc}
+\input{movies/snippets/Hprime/HprimePlusFin}
 
@@
 
-\begin{Coqsrc}
-Lemma H'_omega_double k : H'_ (omega * 2)%e0 k = (4 * k + 3)%nat.
-Proof.
-  rewrite H'_eq3; simpl Canon.
- -   ochange  (CanonS  (omega * FinS 1)%e0 k) (omega + (S k))%e0.
-  + rewrite H'_Plus_Fin, H'_omega;  abstract lia.
-  -  now compute.
-Qed.
-
-Lemma H'_omega_3 k : H'_ (omega * 3)%e0 k = (8 * k + 7)%nat.
-
-Lemma H'_omega_4 k : H'_ (omega * 4)%e0 k = (16 * k + 15)%nat.
-
-Lemma H'_omega_i i  : forall k,
-    H'_ (omega * i)%e0 k = (exp2 i * k + Nat.pred (exp2 i))%nat.
-\end{Coqsrc}
+\input{movies/snippets/Hprime/HprimeExamples}
 
@@
 
-\begin{Coqsrc}
-Lemma H'_omega_sqr : forall k,
-    H'_ (phi0  2)%e0 k = (exp2 (S k ) * (S k) - 1)%nat.
-Proof.
-  intro k; rewrite H'_eq3; auto with E0.
-  - ochange (Canon (phi0 2) (S k)) (omega * (S k))%e0.
-    +  rewrite H'_omega_i; simpl (exp2 (S k)).
-       *  rewrite Nat.add_pred_r.
-          -- abstract lia. 
-          -- generalize (exp2_not_zero k); abstract lia.
-    + cbn; f_equal; abstract lia.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/Hprime/HprimeOmegaSqr}
 
@@
 
-\begin{Coqsrc}
-Lemma H'_Omega_term_1 : alpha <> Zero -> forall k,  
-    H'_ (Omega_term alpha (S i)) k =
-    H'_ (Omega_term alpha i) (H'_ (phi0 alpha) k).
-\end{Coqsrc}
+\input{movies/snippets/Hprime/HprimeOmegaTerm1}
 
@@
 
-\begin{Coqsrc}
-Lemma H'_Omega_term (alpha : E0):
-alpha <> Zero -> forall i k, 
-  H'_ (Omega_term alpha i) k = iterate (H'_ (phi0 alpha)) (S i) k.
-\end{Coqsrc}
+\input{movies/snippets/Hprime/HprimeOmegaTerm}
 
@@
 
-\begin{Coqsrc}
-Definition H'_succ_fun f k := iterate f (S k) k.
-
-Lemma H'_phi0_succ alpha  : alpha <> Zero -> forall k,
- H'_ (phi0 (Succ alpha)) k = H'_succ_fun (H'_ (phi0 alpha)) k.
-
-Lemma H'_phi0_Si : forall i k,
- H'_ (phi0 (S i)) k = iterate H'_succ_fun i (H'_ omega) k. 
-\end{Coqsrc}
-
+\input{movies/snippets/Hprime/HprimeSuccFun}
+\input{movies/snippets/Hprime/HprimePhi0SI}
+
+\item We get now a  formula for $H'_{\omega^3}$:
+
+\input{movies/snippets/Hprime/HprimeOmegaCube}
 \end{enumerate}
-We get now a  formula for $H'_{\omega^3}$:
-
-\begin{Coqsrc}
-Lemma H'_omega_cube : forall k,
-    H'_ (phi0 3)%e0 k = iterate (H'_ (phi0 2))  (S k) k.
-Proof.
-  intro k; rewrite <-FinS_eq, -> Fin_Succ, H'_phi0_succ; auto.
-  compute; injection 1; discriminate.
-Qed.
-\end{Coqsrc}
 
@@
 
-It seems hard to capture the complexity of this function by looking only at this
+It looks hard to capture the complexity of this function by looking only at this
 ``exact'' formula. 
@@
 
-\begin{Coqsrc}
-Section H'_omega_cube_3.
-
-Let f k :=   (exp2 (S k ) * (S k) - 1)%nat.
-
-Remark R0 k :  H'_ (phi0 3)%e0 k = iterate f (S k) k.
-\end{Coqsrc}
+\input{movies/snippets/Hprime/HprimeOmegaCube3a}
 
@@
 
-\begin{Coqsrc}
-Fact F0 : H'_ (phi0 3) 3 = f (f (f (f 3))).
-proof. rewrite R0; reflexivity. Qed.
-\end{Coqsrc}
+\input{movies/snippets/Hprime/HprimeOmegaCube3b}
 
@@
 
-\begin{Coqsrc}
-Let N := (exp2 64 * 64 - 1)%nat.
-\end{Coqsrc}
+\input{movies/snippets/Hprime/HprimeOmegaCube3c}
 
@@
 
-\begin{Coqsrc}
-Fact F1 : H'_ (phi0 3) 3 = f (f N).
-Proof.
- rewrite F0; reflexivity. 
-Qed.
-
-Lemma F1_simpl : H'_ (phi0 3) 3 =
-                 (exp2 (exp2 (S N) * S N) * (exp2 (S N) * S N) - 1)%nat.
-
-\end{Coqsrc}
+\input{movies/snippets/Hprime/HprimeOmegaCube3d}
 
@@
 $2^{10^{21}}$  digits. But it is still an under-approximation !
-
-\begin{Coqsrc}
-End H'_omega_cube_3.
-\end{Coqsrc}
 

### 2021-08-25 doc/part-hydras.tex
Alectryon:  chapter 6 (completed, draft)

@@
 
-\begin{Coqsrc}
-Lemma H'_phi0_omega : forall k, H'_ (phi0 omega) k =
-                               iterate H'_succ_fun  k (H'_ omega) k.
-Proof with auto with E0.
-  intro k; rewrite H'_eq3, <- H'_phi0_Si ...
-  -  rewrite CanonS_Canon, CanonS_phi0_lim;  f_equal ...
-Qed.
-\end{Coqsrc}
-
-Using extensionality of the functional \texttt{iterate}, we can get a closed formula.
-
-\begin{Coqsrc}
-Lemma H'_phi0_omega_closed_formula k :
-  H'_ (phi0 omega) k =
-  iterate (fun (f: nat -> nat) (l : nat) => iterate  f (S l) l)
-               k
-               (fun k : nat => S (2 * k)%nat)
-               k.
-\end{Coqsrc}
-
-Note that this short formula contains two occurrences of the functional \texttt{iterate}, the outer one is in fact a second-order iteration (on type \texttt{nat -> nat)}
+\input{movies/snippets/Hprime/HprimePhi0Omega}
+
+Using extensionality of the functional \texttt{iterate}, we also get a closed formula.
+
+\input{movies/snippets/Hprime/HprimePhi0OmegaClosed}
+
+Note that this formula contains two occurrences of the functional \texttt{iterate}, the outer one is in fact a second-order iteration (on type \texttt{nat -> nat)}
 and the inner one  first-order (on type \texttt{nat}).
@@
 
-\begin{Coqsrc}
-Remark H'_non_mono1 :
-  ~ (forall alpha beta k, (alpha o<= beta)%e0 ->
-                          (H'_ alpha k <= H'_ beta k)%nat).
-Proof.
- intros H ;specialize (H 42 omega 3).
- assert (H0 :(42 o<= omega)%e0) by (repeat split; auto).  
- apply H in H0; rewrite H'_Fin, H'_omega  in H0; lia.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/Hprime/HprimeNonMono1}
 
@@
 
-\begin{Coqsrc}
-Section Proof_of_Abstract_Properties.
-  Record P (alpha:E0) : Prop :=
-    mkP {
-        PA : strict_mono (H'_ alpha);
-        PB : alpha <> Zero -> forall n,  (n < H'_ alpha n)%nat;
-        PC : H'_ alpha <<= H'_ (Succ alpha);
-        PD : dominates_from 1 (H'_ (Succ alpha)) (H'_ alpha);
-        PE : forall beta n, Canon_plus n alpha beta -> 
-                            (H'_ beta n <= H'_ alpha n)%nat}.
-
-Theorem P_alpha : forall alpha, P alpha.
-  Proof.
-    intro alpha; apply well_founded_induction with lt.
-   (* rest of proof skipped *)
-
-Section Proof_of_Abstract_Properties.
-\end{Coqsrc}
-
-By elimination, we get a catalogue of properties of the functions $L'_\alpha$:
-
-\begin{Coqsrc}
-Section Abstract_Properties.
- Variable alpha : E0.
-
- Theorem H'_alpha_mono : strict_mono (H'_ alpha).
-
- Theorem H'_alpha_gt : alpha <> Zero ->
-        forall n, n < H'_ alpha n.
-
- Theorem H'_alpha_Succ_le :
-        H'_ alpha <<= H'_ (Succ alpha).
-
- Theorem H'_alpha_dom :
-    dominates_from 1 (H'_ (Succ alpha)) (H'_ alpha).
-
- Theorem H'_restricted_mono_l :
-    forall beta n, Canon_plus n alpha beta -> 
-                                 H'_ beta n <= H'_ alpha n.
-
-  Lemma H'_alpha_ge_id : id <<= H'_ alpha.
-
-  Lemma H'_alpha_mono_weak : forall k l, k <= l ->
-                                        H'_ alpha k <= H'_ alpha l.
-
-End Abstract_Properties.
-\end{Coqsrc}
+\input{movies/snippets/Hprime/PAP}
+
+% By elimination, we get a catalogue of properties of the functions $L'_\alpha$:
+
+% \begin{Coqsrc}
+% Section Abstract_Properties.
+%  Variable alpha : E0.
+
+%  Theorem H'_alpha_mono : strict_mono (H'_ alpha).
+
+%  Theorem H'_alpha_gt : alpha <> Zero ->
+%         forall n, n < H'_ alpha n.
+
+%  Theorem H'_alpha_Succ_le :
+%         H'_ alpha <<= H'_ (Succ alpha).
+
+%  Theorem H'_alpha_dom :
+%     dominates_from 1 (H'_ (Succ alpha)) (H'_ alpha).
+
+%  Theorem H'_restricted_mono_l :
+%     forall beta n, Canon_plus n alpha beta -> 
+%                                  H'_ beta n <= H'_ alpha n.
+
+%   Lemma H'_alpha_ge_id : id <<= H'_ alpha.
+
+%   Lemma H'_alpha_mono_weak : forall k l, k <= l ->
+%                                         H'_ alpha k <= H'_ alpha l.
+
+% End Abstract_Properties.
+% \end{Coqsrc}
 
@@
 \noindent
-\begin{Coqsrc}
-  H'_dom : forall alpha beta : E0, alpha o< beta ->
-  dominates_strong (H'_ beta) (H'_ alpha).
-\end{Coqsrc}
+\input{movies/snippets/Hprime/HprimeDom}
 
@@
 
-\begin{Coqsrc}
- Theorem H'_L_ alpha: forall i:nat,  (H'_ alpha i <= L_ alpha (S i))%nat.
-\end{Coqsrc}
+\input{movies/snippets/L_alpha/HprimeL}
 
@@
 \noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0.L\_alpha}}
-
-\begin{Coqsrc}
-Theorem battle_length_std_Hardy (alpha : E0) :
-  alpha <> Zero ->
-  forall k , 1 <= k -> exists l: nat,  
-       H'_ alpha k - k <= l /\
-       battle_length standard k (iota (cnf alpha)) l.    
-\end{Coqsrc}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems}}
+
+\input{movies/snippets/Hydra_Theorems/battleLengthStdHardy}
 
@@
 \index{hydras}{Library Prelude!iterate}
-\begin{Coqsrc}
-Fixpoint iterate {A:Type}(f : A -> A) (n: nat)(x:A) :=
-  match n with
-  | 0 => x
-  | S p => f (iterate  f p x)
-  end.
-\end{Coqsrc}
+
+\input{movies/snippets/Iterates/iterateDef}
 
@@
  \index{coq}{Plug-ins!Equations}
-\begin{Coqsrc}
-Fail Equations F_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
-  F_ alpha  i with E0_eq_dec alpha Zero :=
-    { | left _ =>  i ;
-      | right nonzero
-          with Utils.dec (Limitb alpha) :=
-          { | left _ =>  F_ (Canon alpha i)  i ;
-            | right notlimit =>  iterate (F_ (Pred alpha))  (S i) i}}.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-The command has indeed failed with message:
-In environment
-alpha : E0
-notlimit : Limitb alpha = false
-nonzero : alpha <> Zero
-i : nat
-F_ : forall x : E0, nat -> x o< alpha -> nat
-The term "F_ (Pred alpha) ?x" has type "Pred alpha o< alpha -> nat"
-while it is expected to have type 
-"Pred alpha o< alpha -> Pred alpha o< alpha"
-(cannot unify "nat" and "Pred alpha o< alpha").
-\end{Coqanswer}
+
+ \input{movies/snippets/F_alpha/FailDemo}
 
@@
 
-\begin{Coqsrc}
-Definition call_lt (c c' : E0 * nat) :=
-  lexico Lt (Peano.lt) c c'.
-
-Lemma call_lt_wf : well_founded call_lt.
-  unfold call_lt; apply Inverse_Image.wf_inverse_image,  wf_lexico.
-  -  apply E0.Lt_wf.
-  -  unfold Peano.lt; apply Nat.lt_wf_0. 
-Qed.
-
-Instance WF : WellFounded call_lt := call_lt_wf.
-
-Equations  F_star (c: E0 * nat) (i:nat) :  nat by wf  c call_lt :=
-    F_star (alpha, 0) i := i;
-    F_star (alpha, 1) i
-      with E0_eq_dec alpha Zero :=
-           { | left _ => S i ;
-             | right nonzero
-                 with Utils.dec (Limitb alpha) :=
-                 { | left _ => F_star (Canon alpha i,1) i ;
-                   | right notlimit =>
-                     F_star (Pred alpha, S i)  i}};
-    F_star (alpha,(S (S n))) i :=
-               F_star (alpha, 1) (F_star (alpha, (S n)) i).
-
-(* Finally, F_ alpha is defined as its first iterate ! *)
-
-Definition F_  alpha i := F_star (alpha, 1) i.
-\end{Coqsrc}
+\input{movies/snippets/F_alpha/goodDef}
 
@@
 
-\begin{Coqsrc}
-Lemma F_zero_eqn : forall i, F_ Zero i = S i.
-
-Lemma F_lim_eqn : forall alpha i,  Limitb alpha ->
-                               F_ alpha i = F_ (Canon alpha i) i.
-
-Lemma F_succ_eqn : forall alpha i,
-    F_ (Succ alpha) i = iterate (F_ alpha) (S i) i.
-\end{Coqsrc}
+\input{movies/snippets/F_alpha/FEquations}
 
@@
 
-\begin{Coqsrc}
-Lemma LF1 : forall i,  F_ 1 i = S (2 * i).
-
-Lemma LF2 : forall i, exp2 i * i < F_ 2 i.
-\end{Coqsrc}
+\input{movies/snippets/F_alpha/FirstValues}
 
@@
 
-\begin{Coqsrc}
-Theorem F_alpha_mono alpha : strict_mono (F_ alpha).
-
-Theorem F_alpha_ge_S alpha : forall n, n < F_ alpha n.
-
-Theorem F_alpha_Succ_le alpha : F_ alpha <<= F_ (Succ alpha).
-
-Theorem F_alpha_dom alpha : 
-     dominates_from 1 (F_ (Succ alpha)) (F_ alpha).
-
-Theorem F_restricted_mono_l alpha : 
-   forall beta n, Canon_plus n alpha beta -> 
-                                        F_ beta n <= F_ alpha n.
-\end{Coqsrc}
-
-As a corollary, we prove the following proposition, p. 284 of~\cite{KS81}, which states that the function \texttt{F\_} is ``monotonous'' w.r.t. its first argument and ``domination''.
-
+\input{movies/snippets/F_alpha/FalphaThms}
+
+As a corollary, we prove the following proposition, p. 284 of~\cite{KS81}.
 \begin{quote}
@@
 
-\begin{Coqsrc}
-Lemma F_mono_l: forall alpha beta : E0, 
-   beta o< alpha -> dominates (F_ alpha) (F_ beta).
-\end{Coqsrc}
+\input{movies/snippets/F_alpha/FDomContext}
+\input{movies/snippets/F_alpha/FDom}
 

### 2021-08-25 doc/part-hydras.tex
Alectryon:  chapter 7 (start)

@@
 
-\begin{Coqsrc}
-Parameter Ord : Type.
-Parameter lt : relation Ord.
-Infix "<" := lt : schutte_scope.
-
-Definition ordinal := Full_set Ord.
-\end{Coqsrc}
+\input{movies/snippets/Schutte_basics/OrdDecl}
 
@@
 
-\begin{Coqsrc}
-Variables (M:Type)
-         (Lt : relation M).
-
-Class WO : Type:=
-    {
-      Lt_trans : Transitive  Lt;
-      Lt_irreflexive : forall a:M, ~ (Lt a a);
-      well_order : forall (X:Ensemble M)(a:M),
-          In X a ->
-          exists a0:M, least_member  X a0
-    }.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-  Axiom AX1 : WO lt.
-\end{Coqsrc}
+\input{movies/snippets/Well_Orders/Mdecl}
+
+\input{movies/snippets/Well_Orders/WODef}
+
+\input{movies/snippets/Schutte_basics/AX1}
 
@@
 
-\begin{Coqsrc}
-Axiom AX2 : forall X: Ensemble Ord, 
-   (exists a,  (forall y, In X y -> y < a)) ->
-   countable X.
-
-Axiom AX3 : forall X : Ensemble Ord,
-              countable X -> 
-              exists a,  forall y, In X y -> y < a.
-\end{Coqsrc}
-
-\texttt{AX2} and \texttt{AX3} could have been replaced by a single axiom (using the \texttt{iff} connector), but we decide to respect as most as possible the structure of Schütte's definitions.
+\input{movies/snippets/Schutte_basics/AX23}
+
+\texttt{AX2} and \texttt{AX3} could have been replaced by a single axiom (using the \texttt{iff} connector), but we decided to respect as most as possible the structure of Schütte's definitions.
 
@@
 
-According to the  definition of a well order, every non-empty subset of \texttt{Ord} has a least element. Furthermore, this least element is unique.
-
-\begin{Coqsrc}
-Remark R : exists! z : Ord, least_member lt  ordinal z.
-Proof.
-  destruct inh_Ord as [a]; apply (well_order (WO:=AX1)) with a .
-  split.
-Qed.
-\end{Coqsrc}
-
-Assume we want to call this element  \texttt{zero}.
-
-\begin{Coqsrc}
-Definition zero : Ord.
-Proof.
-  Fail destruct R.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-The command has indeed failed with message:
-Case analysis on sort Type is not allowed for inductive 
-definition ex.
-\end{Coqanswer}
+According to the  definition of a well order, every non-empty subset of \texttt{Ord} has a least element. Furthermore, this least element is unique. We would like to call this element  \texttt{zero}.
+
+\input{movies/snippets/Schutte_basics/iotaDemo}
 
 Indeed, the basic logic of  \coq{} does not allow us to eliminate a proof of a proposition 
-$\exists\,x:A,\,P(x)$ for building a term whose type lies in the sort \texttt{Type}. 
+$\exists!\,x:A,\,P(x)$ for building a term whose type lies in the sort \texttt{Type}. 
 The reasons for this impossibility are explained in many documents~\cite{BC04, chlipalacpdt2011, Coq}.
@@
 
-\begin{Coqsrc}
-Class InH (A: Type) : Prop :=
-   InHWit : inhabited A.
-
-Definition some {A:Type} {H : InH A} (P: A -> Prop) := 
-   epsilon (@InHWit A H) P.
-
-Definition the {A:Type} {H : InH A} (P: A -> Prop) := 
-   iota (@InHWit A H) P.
-\end{Coqsrc}
-
-In order to use these tools,  we had to tell \coq{}  that the type \texttt{Ord} is not empty:
-
-\begin{Coqsrc}
-Axiom inh_Ord : inhabited Ord.
-\end{Coqsrc}
+\input{movies/snippets/MoreEpsilonIota/Defs}
+
+In order to use these tools,  we have to tell \coq{}  that the declared type \texttt{Ord} is not empty:
+
+\input{movies/snippets/Schutte_basics/inhOrd}
 
@@
 
-\begin{Coqsrc}
-Definition the_least {M: Type} {Lt}
-           {inh : InH M} {WO: WO Lt} (X: Ensemble M)  : M :=
-  the  (least_member  Lt X ).
-\end{Coqsrc}
+\emph{From Module\href{../theories/html/hydras.Schutte.Well_Orders.html}%
+  {\texttt{Schutte.Well\_Orders}}}
+
+\input{movies/snippets/Well_Orders/leastMemberDef}
+\input{movies/snippets/Well_Orders/theLeast}
 
@@
 
-\begin{Coqsrc}
-Definition zero: Ord :=the_least ordinal.
-\end{Coqsrc}
+\input{movies/snippets/Schutte_basics/zeroDef}
 
@@
 
-\begin{Coqsrc}
-Lemma zero_le (alpha : Ord) :  zero <= alpha.
-Proof.
-  unfold zero, the_least, the; apply iota_ind.
-\end{Coqsrc}
-
-According to the use of the description operator \texttt{iota}, we have to solve  two trivial sub-goals.
-\begin{enumerate}
-\item Prove that there exists a unique least member of \texttt{Ord}
-\item Prove that being a least member of \texttt{Ord} entails the announced inequality 
-\end{enumerate}
-
-\begin{Coqanswer}
-2 subgoals (ID 155)
-
-  alpha : Ord
-  ============================
-  exists ! x : Ord, least_member lt ordinal x
-
-subgoal 2 (ID 156) is:
- forall a : Ord, unique (least_member lt ordinal) a -> 
-                a <= alpha
-\end{Coqanswer}
-
-\begin{Coqsrc}
-  -  apply the_least_unicity, Inh_ord.
-  -  destruct 1 as [[_ H1] _]; apply H1; split. 
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/Schutte_basics/zeroLe}
 
@@
 
-\begin{Coqbad}
-Module Bad.
-
- Definition bottom := the_least (Empty_set Ord).
-\end{Coqbad}
-
-\begin{Coqanswer}
- bottom is defined
-\end{Coqanswer}
+\input{movies/snippets/Schutte_basics/BadBottoma}
 
@@
 
-\begin{Coqbad}
-Lemma le_zero_bottom : zero <= bottom. 
-Proof. apply zero_le. Qed.
-
-Lemma bottom_eq : bottom = bottom.
-Proof. trivial. Qed.
-
-Lemma le_bottom_zero : bottom <= zero.
-Proof.
-   unfold bottom, the_least, the; apply iota_ind.
-\end{Coqbad}
-
-\begin{Coqanswer}
-2 subgoals (ID 413)
-
-============================
-exists ! x : Ord, least_member lt (Empty_set Ord) x
-
-subgoal 2 (ID 414) is:
- forall a : Ord, unique (least_member lt (Empty_set Ord)) a -> 
-      a <= zero
-\end{Coqanswer}
-
-\begin{Coqbad}
-Abort.
-End Bad.
-\end{Coqbad}
+\input{movies/snippets/Schutte_basics/trivialProps}
+
+On the other hand, the following attempt fails, because of the unprovable first sub-goal (please notice that the second sub-goal is easy to solve !).
+
+\input{movies/snippets/Schutte_basics/Failure}
 
@@
 
-\begin{Coqsrc}
-Definition succ (alpha : Ord) := the_least (fun beta => alpha < beta).
-\end{Coqsrc}
-
-Using \texttt{succ}, we define the folloing predicates.
-
-\begin{Coqsrc}
-Definition is_succ (alpha:Ord) := exists beta, alpha = succ beta.
-
-Definition is_limit (alpha:Ord) := alpha <> zero /\ ~ is_succ alpha.
-\end{Coqsrc}
+\input{movies/snippets/Schutte_basics/succDef}
+
+Using \texttt{succ}, we define the following predicates.
+
+\input{movies/snippets/Schutte_basics/isSuccIsLimit}
 
@@
 % Please look at remark~\vref{warning:coercions}.
-% \end{remark}
 
@@
 
-\begin{Coqsrc}
-Definition succ_spec (alpha:Ord) :=
-  least_member   lt (fun z => alpha < z).
-\end{Coqsrc}
+\input{movies/snippets/Schutte_basics/succSpec}
 
@@
 
-\begin{Coqsrc}
-Lemma succ_ok : forall alpha,  succ_spec alpha  (succ alpha).
-Proof.
-  intros; unfold succ, the_least, the;  apply iota_spec.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-1 subgoal (ID 172)
-
-  alpha : Ord
-  ============================
-  exists ! x : Ord, succ_spec alpha x
-\end{Coqanswer}
+\input{movies/snippets/Schutte_basics/succOka}
 
@@
 
-The \coq{} proof script is quite short.
-
-\begin{Coqsrc}
-  destruct (@AX3 (Singleton _ alpha)).
-  - apply countable_singleton.
-  -  unfold succ_spec; apply the_least_unicity;  exists x; intuition.
-Qed.     
-\end{Coqsrc}
+The rest of the \coq{} proof script is quite short.
+
+\input{movies/snippets/Schutte_basics/succOkb}
 
@@
 
-\begin{Coqsrc}
-Lemma lt_succ (alpha : Ord) :  alpha < succ alpha.
-Proof.
-  destruct  (succ_ok  alpha);  tauto.
-Qed.
-
-Hint Resolve lt_succ : schutte.
-
-Lemma lt_succ_le (alpha beta : Ord):
-  alpha < beta -> succ alpha <= beta.
-Proof with eauto with schutte.
-  intros  H;  pattern (succ alpha); apply the_least_ok ... 
-  exists (succ alpha); red;apply lt_succ ...
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma lt_succ_le_2 (alpha beta : Ord):
-  alpha < succ beta -> alpha <= beta.
-
-Lemma succ_mono (alpha beta : Ord):
-  alpha < beta -> succ alpha < succ beta.
-
-Lemma succ_monoR (alpha beta : Ord) :
- succ alpha < succ beta -> alpha < beta.
-
-Lemma lt_succ_lt (alpha beta : Ord) :
-  is_limit beta ->  alpha < beta -> succ alpha < beta.
-\end{Coqsrc}
+\input{movies/snippets/Schutte_basics/succProps}
 

### 2021-08-25 doc/part-hydras.tex
Alectryon:  chapter 7 (30%)

@@
 
-\begin{Coqsrc}
-Reserved Notation "'F' n" (at level 29) .
-
-Fixpoint finite (i:nat) : Ord :=
-  match i with 
-            | 0 => zero
-            | S i => succ (F i)
-  end
-where "'F' i" := (finite i)  : schutte_scope.
-
-Coercion finite : nat >-> Ord.
-\end{Coqsrc}
+\input{movies/snippets/Schutte_basics/finiteDef}
 
@@
 
-\begin{Coqsrc}
-Definition is_lub (M:Type)
-                  (D : Ensemble M)
-                  (lt : relation M)
-                  (X:Ensemble M)
-                  (a:M) :=
-   In _ D a  /\ upper_bound  D lt X a  /\
-   (forall y, In _ D y -> upper_bound  D lt X y  -> 
-                  y = a \/ lt a y).
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Definition sup_spec X lambda := is_lub ordinal lt X lambda.
-
-Definition sup (X: Ensemble Ord) : Ord  := the  (sup_spec X).
-
-Notation "'|_|' X" := (sup X) (at level 29) : schutte_scope.
-\end{Coqsrc}
+\input{movies/snippets/Lub/isLubDef}
+
+\input{movies/snippets/Schutte_basics/supDef}
 
@@
 By \texttt{AX3} this range is bounded, hence the set of its upper bounds is not empty and has a least element.
-
-\begin{Coqsrc}
-Definition omega_limit (s:nat->Ord) : Ord 
-  := |_| (seq_range s).
-\end{Coqsrc}
-
 Then we define \texttt{omega} as the limit of the sequence of finite ordinals.
 
+\input{movies/snippets/Schutte_basics/omegaDef}
+
 \label{sect:notation-omega}
-\begin{Coqsrc}
-Definition _omega := omega_limit finite.
-
-Notation "'omega'" := (_omega) : schutte_scope.
-\end{Coqsrc}
 
@@
 
-\begin{Coqsrc}
-Lemma finite_lt_omega : forall i: nat,  i < omega.
-
-Lemma lt_omega_finite alpha : Ord) : 
-  alpha < omega ->  exists i:nat, alpha =  i.
-
-Lemma is_limit_omega : is_limit omega.
-\end{Coqsrc}
+\input{movies/snippets/Schutte_basics/omegaProps}
 

### 2021-08-26 doc/part-hydras.tex
Alectryon:  chapter 7 (100%)

@@
 
-\begin{Coqsrc}
- Definition segment (A: Ensemble Ord) :=
-  forall alpha beta, In A alpha -> beta < alpha -> In A  beta.
-
-Definition proper_segment (A: Ensemble Ord) :=
-  segment A /\ ~ Same_set A ordinal.
-\end{Coqsrc}
+\input{movies/snippets/Ordering_Functions/segmentDef}
 
@@
 
-\begin{Coqsrc}
-Definition ordering_function (f : Ord -> Ord)(A B : Ensemble Ord) :=
- segment A /\
- (forall a, In A a -> In B (f a)) /\
- (forall b, B b -> exists a, In A a /\ f a = b) /\
- forall a b, In A a -> In A b -> a < b ->  f a < f b.
-
-Definition ordering_segment (A B : Ensemble Ord) :=
-  exists f : Ord -> Ord, ordering_function f A B.
-\end{Coqsrc}
+\input{movies/snippets/Ordering_Functions/orderingFunctionDef}
 
@@
 
-\begin{Coqsrc}
-Definition the_ordering_segment (B : Ensemble Ord) :=
-  the  (fun x => ordering_segment x B).
-
-Definition ord  (B : Ensemble Ord) := 
-  some (fun f => ordering_function f (the_ordering_segment B) B).
-\end{Coqsrc}
+\input{movies/snippets/Ordering_Functions/ordDef}
 
@@
 
-\begin{Coqsrc}
-About ordering_function_ex.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-forall B : Ensemble Ord,
- exists ! S : Ensemble Ord, 
-      exists f : Ord -> Ord, ordering_function f S B
-\end{Coqanswer}
-
-\begin{Coqanswer}
-ordering_function_unicity :
-forall (B S1 S2 : Ensemble Ord) (f1 f2 : Ord -> Ord),
-ordering_function f1 S B ->
-ordering_function f2 S2 B -> 
-fun_equiv f1 f2 S1 S2
-\end{Coqanswer}
+\input{movies/snippets/Ordering_Functions/orderingFunctionEx}
 
@@
 
-\begin{Coqsrc}
-Theorem ordering_le : forall f A B,
-    ordering_function f A B ->
-    forall alpha, In A alpha -> alpha <= f alpha.
-
-Th_13_5_2 :
-forall (A B : Ensemble Ord) (f : Ord -> Ord),
-ordering_function f A B -> Closed B -> continuous f A B
-\end{Coqsrc}
+\input{movies/snippets/Ordering_Functions/orderingLe}
+\input{movies/snippets/Ordering_Functions/Th1352}
 
@@
 \index{hydras}{Library Schutte!Functions!plus}
-\begin{Coqsrc}
-Definition plus alpha := ord  (ge alpha).
-Notation "alpha + beta " := (plus alpha beta) : schutte_scope.
-\end{Coqsrc}
+
+\input{movies/snippets/Addition/additionDef}
 
@@
 
-\begin{Coqsrc}
-Lemma alpha_plus_zero (alpha: Ord): alpha + zero = alpha.
-Proof.
- pattern  (plus alpha); apply plus_elim; eauto.
- \end{Coqsrc}
-
- \begin{Coqanswer}
- 1 subgoal (ID 24)
-
-  alpha : Ord
-  ============================
-  forall f : Ord -> Ord,
-  ordering_function f ordinal (ge alpha) -> 
-  f zero = alpha
- \end{Coqanswer}
-
- \begin{Coqsrc}
- (* rest of proof skipped *)
- \end{Coqsrc}
+As a use-case, let us prove that $0$ is a right neutral element of $+$.
+
+\input{movies/snippets/Addition/alphaPlusZero}
 
@@
 
- \begin{Coqsrc}
-
-Lemma zero_plus_alpha (alpha : Ord) : zero + alpha = alpha.
-
-Lemma le_plus_l (alpha beta : Ord) : alpha <= alpha + beta.
-
-Lemma le_plus_r (alpha beta : Ord) :  beta <= alpha + beta.
-
-Lemma plus_mono_r (alpha beta gamma : Ord) : 
-    beta < gamma -> alpha + beta < alpha + gamma.
-
-Lemma plus_of_succ (alpha beta : Ord) :
-    alpha + (succ beta) = succ (alpha + beta).
-
-Theorem plus_assoc (alpha beta gamma : Ord) :
-  alpha + (beta + gamma) = (alpha + beta) + gamma.
-
-Lemma one_plus_omega :  1 + omega = omega.
-
-Lemma finite_plus_infinite (n : nat) (alpha : Ord) :
-  omega <= alpha -> n + alpha = alpha.
-\end{Coqsrc}
+\input{movies/snippets/Addition/bunchOfLemmas}
+
+The following lemmas are not direct applications of \texttt{plus\_elim}.
+
+\input{movies/snippets/Addition/plusAssoc}
+\input{movies/snippets/Addition/finitePlusInfinite}
 
@@
 
-\begin{Coqsrc}
-Fixpoint mult_Sn (alpha:Ord)(n:nat){struct n}: Ord :=
- match n with 
-            | 0 => alpha
-            | S p => mult_Sn  alpha p + alpha
- end.
-
-Definition mult_fin_r alpha n :=
-  match n with
-      0 => zero
-    | S p => mult_Sn alpha p
-  end.
-
-Notation "alpha * n" := (mult_fin_r alpha n) : schutte_scope.
-\end{Coqsrc}
+\input{movies/snippets/Addition/multFin}
 
@@
 
-\begin{Coqsrc}
-Definition AP : Ensemble Ord :=
-  fun alpha => 
-  zero < alpha /\
-  (forall beta, beta < alpha ->  beta + alpha = alpha).
-\end{Coqsrc}
+\input{movies/snippets/AP/APDef}
 
@@
 \index{hydras}{Library Schutte!Functions!phi0}
-\begin{Coqsrc}
-Definition phi0 := ord AP.
-
-Notation "'omega^'" := phi0 (only parsing) : schutte_scope.
-\end{Coqsrc}
+
+\input{movies/snippets/AP/phi0Def}
 
@@
 
-\begin{Coqsrc}
-Fixpoint omega_tower (i : nat) : Ord :=
-  match i with
-    0 =>  1
-  | S j => phi0 (omega_tower j)
-  end.
-\end{Coqsrc}
+\input{movies/snippets/AP/omegaTower}
 
@@
 
-\begin{Coqsrc}
-Definition epsilon0 := omega_limit omega_tower.
-\end{Coqsrc}
+\input{movies/snippets/AP/epsilon0Def}
 
@@
 
-\begin{Coqsrc}
-Lemma AP_one : In AP 1.
-
-Lemma AP_omega : In AP omega.
-\end{Coqsrc}
-
-Moreover, $1$ is the least principal ordinal and $\omega$ is the second element of
-\texttt{AP}.
-
-\begin{Coqsrc}
-Lemma least_AP: least_member  lt AP 1.
-
-Lemma omega_second_AP :
-  least_member   lt 
-                  (fun alpha => 1 < alpha /\ In AP alpha)
-                  omega.
-\end{Coqsrc}
+\input{movies/snippets/AP/APOne}
 
@@
 
-\begin{Coqsrc}
-Lemma AP_plus_closed (alpha beta gamma : Ord): 
-     In AP alpha -> beta < alpha -> gamma < alpha ->
-     beta + gamma < alpha.
-
-Theorem AP_unbounded : Unbounded AP.
-\end{Coqsrc}
+\input{movies/snippets/AP/APPlusClosed}
+
+\input{movies/snippets/AP/APUnbounded}
 
@@
 
-\begin{Coqsrc} 
-Definition Closed (B : Ensemble Ord) : Prop := 
-  forall M, Included M B -> Inhabited _ M -> 
-                 countable M -> In B (|_| M).
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Theorem AP_closed : Closed AP.
-
-Lemma AP_o_segment :  the_ordering_segment AP = ordinal.
-\end{Coqsrc}
+From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
+{\texttt{~Schutte.Schutte\_basics}}
+
+\input{movies/snippets/Schutte_basics/ClosedDef}
+\input{movies/snippets/AP/APClosed}
 
@@
 
-\begin{Coqsrc}
-Lemma AP_phi0 (alpha : Ord) : In AP (phi0 alpha).
-
-Lemma phi0_zero : phi0 zero =  1.
-
-Lemma phi0_mono (alpha beta : Ord) :
-  alpha < beta ->  phi0 alpha < phi0 beta.
-
-Lemma phi0_inj (alpha beta : Ord) :
-    phi0 alpha = phi0 beta -> alpha = beta.
-
-Lemma phi0_sup : forall (U: Ensemble Ord),
-   Inhabited _ U ->   countable U ->  phi0 (|_| U) = |_| (image U phi0).
-
-Lemma is_limit_phi0 (alpha : Ord) :
-  zero < alpha ->  is_limit (phi0 alpha).
-
-Lemma omega_eq : omega = phi0 1.
-
-Lemma phi0_le (alpha : Ord) : alpha <= phi0 alpha.
-\end{Coqsrc}
-
-Please note that the lemma \texttt{omega\_eq} above, is consistent with the interpretation of the ordering function $\phi_0$ as the exponential of basis $\omega$.
-Indeed we could have written this lemma with our alternative notation:
-
-\begin{Coqsrc}
- Lemma omega_eq : omega = omega^ 1.
-\end{Coqsrc}
+\input{movies/snippets/AP/APPhi0}
 
@@
 
-\begin{Coqsrc}
-Lemma lt_phi0 (alpha : Ord):
-  alpha < epsilon0 -> alpha < phi0 alpha.
-\end{Coqsrc}
+\input{movies/snippets/AP/ltPhi0}
 
@@
 Moreover,  $\epsilon_0$ is the least ordinal $\alpha$ that verifies the equality 
-$\alpha = \omega^\alpha$, in other words the least fixpoint of the function  $\phi_0$.
-
-\begin{Coqsrc}
-Theorem epsilon0_lfp : least_fixpoint lt phi0 epsilon0.
-\end{Coqsrc}
+$\alpha = \omega^\alpha$, in other words, the least fixpoint of the function  $\phi_0$.
+
+\input{movies/snippets/AP/epsilon0Lfp}
 
@@
 
-\begin{Coqsrc}
-Definition Cr_fun : forall alpha : Ord,
-       (forall beta : Ord, beta < alpha -> Ensemble Ord) ->
-        Ensemble Ord 
-:= 
-   fun (alpha :Ord)
-        (Cr : forall beta, 
-                beta < alpha -> Ensemble Ord) 
-        (x : Ord) => (
-       (alpha = zero /\ AP x) \/
-       (zero < alpha /\
-        forall beta (H:beta < alpha),
-          the_ordering_segment (Cr beta H) x /\ ord (Cr  beta H) x = x)).
-
-Definition Cr (alpha : Ord) : Ensemble Ord := 
-    (Fix  all_ord_acc (fun (_:Ord) => Ensemble Ord) Cr_fun) alpha.
-\end{Coqsrc}
+\input{movies/snippets/Critical/CrDef}
+
+Lets us denote by $\varphi_\alpha$ the ordering function of the set $\textit{Cr}(\alpha)$ and by $A_\alpha$ its ordering segment.
+
+\input{movies/snippets/Critical/phiDef}
 
 \label{sect:phi-schutte}
-
-\begin{Coqsrc}
-Definition phi (alpha : Ord) : Ord -> Ord 
-    :=  ord (Cr alpha).
-
-Definition A (alpha : Ord) : Ensemble Ord :=
-  the_ordering_segment (Cr alpha).
-\end{Coqsrc}
 
@@
 
-\begin{Coqsrc}
-Lemma Cr_zero_AP :  Cr 0 = AP
-
-Lemma epsilon0_Cr1 : In (Cr 1) epsilon0.
-\end{Coqsrc}
+\input{movies/snippets/Critical/CrZeroAP}
+\input{movies/snippets/Critical/epsilon0Cr1}
 
@@
 
-We prove this result by transfinite induction on $\alpha$ of both properties.
-
-The proof is still quite long, by transfinite induction over $\alpha$.
+We prove this result by a transfinite induction on $\alpha$ of the conjunction of  both properties.
 
@@
 
-\begin{Coqsrc}
-Section Proof_of_Lemma5.
-  Let P (alpha:Ord) := Unbounded (Cr alpha) /\ Closed (Cr alpha).
-
- Lemma Lemma5 : forall alpha, P alpha.
-(* ... *)
- End Proof_of_Lemma5.
-
-Corollary Unbounded_Cr alpha : Unbounded (Cr alpha).
-Proof.
-  now destruct (Lemma5 alpha).
-Qed.
-
-Corollary Closed_Cr alpha : Closed (Cr alpha).
-Proof.
-  now destruct (Lemma5 alpha).
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/Critical/Lemma5}
 
@@
 
-\begin{Coqsrc}
- Definition cnf_t := list Ord.
-
-Fixpoint eval (l : cnf_t) : Ord :=
-  match l with nil => zero
-              | beta :: l' => phi0 beta + eval l'
-  end.
-
-Definition sorted (l: cnf_t) :=
-  LocallySorted (fun alpha beta => beta <= alpha) l.
-
-Definition is_cnf_of (alpha : Ord)(l : cnf_t) : Prop :=
-  sorted l /\ alpha = eval l.
-\end{Coqsrc}
+\input{movies/snippets/CNF/Defs}
 
@@
 
-\begin{Coqsrc}
-Theorem cnf_exists (alpha : Ord) :
-  exists l: cnf_t, is_cnf_of alpha l.
-\end{Coqsrc}
+\input{movies/snippets/CNF/cnfExists}
 
@@
 
-\begin{Coqsrc}
- Lemma cnf_unicity : forall l alpha, 
-   is_cnf_of alpha l -> 
-   forall l',  is_cnf_of alpha l' -> l=l'.
-Proof.
- induction l.
- (*  ...  *)
-
-Theorem cnf_exists_unique (alpha:Ord) :
-  exists! l: cnf_t, is_cnf_of alpha l.
-\end{Coqsrc}
+\input{movies/snippets/CNF/cnfUnicity}
+
+\input{movies/snippets/CNF/cnfExUnique}
 
@@
 
-\begin{Coqsrc}
-Lemma cnf_lt_epsilon0 : 
- forall l alpha, 
-   is_cnf_of alpha l ->  alpha < epsilon0 ->
-   Forall (fun beta =>  beta < alpha) l.
-\end{Coqsrc}
+\input{movies/snippets/CNF/cnfLtEpsilon0}
 
@@
 
-\begin{Coqsrc}
-Lemma cnf_of_epsilon0 : is_cnf_of epsilon0 (epsilon0 :: nil).
-Proof.
-  split.
-  - constructor.  
-  - simpl;  now rewrite alpha_plus_zero, epsilon0_fxp.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/CNF/cnfOfEpsilon0}
 
@@
 
-\begin{Coqsrc}
-Fixpoint inject (t:T1) : Ord :=
- match t with 
-     | T1.zero => zero
-     | T1.ocons a n b =>  AP.phi0 (inject a) * S n + inject b
- end.  
-\end{Coqsrc}
+\input{movies/snippets/Correctness_E0/injectDef}
 
 This function enjoys good commutation properties with respect to the main operations which
-allow us to build Cantor normal form.
-
-\begin{Coqsrc}
-Theorem inject_of_zero : inject T1.zero = zero.
-
-Theorem inject_of_finite (n : nat):
-  inject (T1.fin n) =  n.
-
-Theorem inject_of_phi0 (alpha : T1):
-  inject (phi0 alpha) = AP.phi0 (inject alpha).
-
-Theorem inject_plus (alpha beta : T1): nf alpha -> nf beta ->
-  inject (alpha + beta)%t1 = inject alpha + inject beta.
-
-Theorem inject_mult_fin_r (alpha : T1)  :
-  nf alpha -> forall n:nat , inject (alpha *  n)%t1 =  inject alpha * n.
-
-Theorem inject_mono (beta gamma : T1) :
-  T1.lt  beta gamma -> 
-  T1.nf beta -> T1.nf gamma -> 
-  inject beta < inject gamma.
-
-Theorem inject_injective (beta gamma : T1) : nf beta -> nf gamma ->
-  inject beta = inject gamma -> beta = gamma.
-\end{Coqsrc}
+allow us to build Cantor normal forms.
+
+\input{movies/snippets/Correctness_E0/commutationLemmas}
+\input{movies/snippets/Correctness_E0/injectPlus}
+\input{movies/snippets/Correctness_E0/injectMultFinR}
+
+% \begin{Coqsrc}
+% Theorem inject_mono (beta gamma : T1) :
+%   T1.lt  beta gamma -> 
+%   T1.nf beta -> T1.nf gamma -> 
+%   inject beta < inject gamma.
+
+% Theorem inject_injective (beta gamma : T1) : nf beta -> nf gamma ->
+%   inject beta = inject gamma -> beta = gamma.
+% \end{Coqsrc}
 
@@
 
-\begin{Coqsrc}
-Theorem inject_lt_epsilon0 (alpha : T1):
-      inject alpha < epsilon0.
-
-Theorem embedding : 
-     fun_bijection (nf: Ensemble T1)  (members epsilon0) inject.
- \end{Coqsrc}
+\input{movies/snippets/Correctness_E0/injectLtEpsilon0}
+\input{movies/snippets/Correctness_E0/embedding}
 

### 2021-08-26 doc/part-hydras.tex
Alectryon:  first complete draft

@@
 
-\begin{Coqsrc}
-Definition strongly_critical alpha := In (Cr alpha) alpha.
-
-Definition maximal_critical alpha : Ensemble Ord :=
-  fun gamma =>
-    In (Cr alpha) gamma /\
-    forall xi, alpha < xi -> ~ In (Cr xi) gamma.
-
-Definition Gamma0 := the_least strongly_critical.
-\end{Coqsrc}
+\input{movies/snippets/Critical/Gamma0Def}
 
@@
 {\texttt{Gamma0.T2}}}
-\begin{Coqsrc}
-(**  [gcons alpha beta n gamma] is : [psi(alpha,beta)*(S n)+ gamma]  *)
-
-Inductive T2 : Set :=
-| zero : T2
-| gcons : T2 -> T2  -> nat -> T2 -> T2.
-
-Notation "[ alpha , beta ]" := (gcons alpha beta 0 zero)
-                                 (at level 0): t2_scope.
-\end{Coqsrc}
+
+\input{movies/snippets/T2/T2Def}
 
@@
 
-\begin{Coqsrc}
-Notation  one  := [zero,zero].
-
-(** The (n+1)-th finite ordinal *)
-Notation FS n := (gcons zero zero n zero).
-
-(** the [n]-th ordinal  *)
-Definition fin (n:nat) := match n with 0 => zero | S p => FS p end.
-
-Notation "'omega'"  := [zero,one] : T2_scope.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Notation "'epsilon0'"  := ([one,zero]) : T2_scope.
-
-Definition epsilon alpha := [one, alpha].
-\end{Coqsrc}
+\input{movies/snippets/T2/Notations}
 
@@
 
-\begin{Coqsrc}
-Inductive lt : T2 -> T2 -> Prop :=
-| (* 1 *) 
- lt_1 : forall alpha beta n gamma,  zero t2< gcons alpha beta n gamma
-| (* 2 *)
- lt_2 : forall alpha1 alpha2 beta1 beta2 n1 n2 gamma1 gamma2, 
-                alpha1 t2< alpha2 ->
-                beta1 t2< gcons alpha2 beta2 0 zero ->
-               gcons alpha1 beta1 n1 gamma1 t2<
-               gcons alpha2 beta2 n2 gamma2
-| (* 3 *)
- lt_3 : forall alpha1  beta1 beta2 n1 n2 gamma1 gamma2, 
-               beta1 t2< beta2 ->
-               gcons alpha1 beta1 n1 gamma1 t2<
-               gcons alpha1 beta2 n2 gamma2
-
-| (* 4 *)
- lt_4 : forall alpha1 alpha2 beta1 beta2 n1 n2 gamma1 gamma2, 
-               alpha2 t2< alpha1 ->
-               [alpha1, beta1] t2< beta2 ->
-               gcons alpha1 beta1 n1 gamma1 t2<
-               gcons alpha2 beta2 n2 gamma2
-
-| (* 5 *)
- lt_5 : forall alpha1 alpha2 beta1 n1 n2 gamma1 gamma2, 
-               alpha2 t2< alpha1 ->
-               gcons alpha1 beta1 n1 gamma1 t2<
-               gcons alpha2  [alpha1, beta1] n2 gamma2
-
-| (* 6 *)
- lt_6 : forall alpha1 beta1  n1  n2 gamma1 gamma2,  (n1 < n2)%nat ->
-                                    gcons alpha1 beta1 n1 gamma1 t2< 
-                                    gcons alpha1 beta1 n2 gamma2
-
-| (* 7 *)
- lt_7 : forall alpha1 beta1 n1   gamma1 gamma2,  gamma1 t2< gamma2 ->
-                                      gcons alpha1 beta1 n1 gamma1 t2<
-                                      gcons alpha1 beta1 n1 gamma2
-where  "o1 t2< o2" := (lt o1 o2): T2_scope.
-
-Hint Constructors lt : T2.
-\end{Coqsrc}
+\input{movies/snippets/T2/ltDef}
 
@@
 
-\subsection{Examples}
-
-\subsubsection*{Proof of $0<\epsilon_0$}
-
-\begin{Coqsrc}
-Example Ex1: 0 t2< epsilon0.
-Proof.  constructor 1. Qed.
-\end{Coqsrc}
-
-\subsubsection*{Proof of $\omega<\epsilon_0$}
-
-\begin{Coqsrc}
-Example Ex2: omega t2< epsilon0.
-Proof. info_auto with T2. (* uses lt_1 and lt_2 *) Qed.
-\end{Coqsrc}
-
-\subsubsection*{Proof of $\psi(\omega,8)\times 13+56 < \psi(\omega,8)\times 13+57 $}
-
-\begin{Coqsrc}
-Example Ex3: gcons omega 8 12 56 t2<  gcons omega 8 12 57.
-Proof.
-  constructor 7; constructor 6; auto with arith.
-Qed.
-\end{Coqsrc}
-
-\subsubsection*{Proof of $\epsilon_0<\psi(2,1)$}
-
-\begin{Coqsrc}
-Example Ex4: epsilon0 t2< [2,1].
-Proof.
-   constructor 2; auto with T2.
-   - constructor 6; auto with arith.
-Qed.
-\end{Coqsrc}
-
-\subsubsection*{Proof of $\psi(2,1)<\psi(2,3)$}
-
-\begin{Coqsrc}
-Example Ex5 : [2,1] t2< [2,3].
-Proof.
-  constructor 3; auto with T2.
-  - constructor 6; auto with arith.
-Qed.
-\end{Coqsrc}
-
-\subsubsection*{Proof of $\psi(1,0)\times 13+ \omega < \psi(0,\psi(2,1))$}
-\label{sect:ex6-first-proof}
-
-\begin{Coqsrc}
-Example Ex6 : gcons 1 0 12 omega t2< [0,[2,1]].
-Proof.
-  constructor 4.
-  - constructor 1.
-  - constructor 2.
-    + constructor 6; auto with arith.
-    + constructor 1.
-Qed.
-\end{Coqsrc}
-
-\subsubsection*{Proof of $\psi(2,1)\times 43 + \epsilon_0 < \psi(1,\psi(2,1))$}
-
-\begin{Coqsrc}
-Example Ex7 : gcons 2 1 42 epsilon0 t2< [1, [2,1]].
-Proof.
- constructor 5.
- constructor 6; auto with arith.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/T2/ltExamples}
 
@@
 
-\begin{Coqsrc}
-Inductive nf : T2 -> Prop :=
-| zero_nf : nf zero
-| single_nf : forall a b n, nf a ->  nf b -> nf (gcons a b n zero)
-| gcons_nf : forall a b n a' b' n' c', 
-                      [a', b'] t2< [a, b]  -> 
-                      nf a -> nf b -> 
-                      nf(gcons a' b' n' c')-> 
-                      nf(gcons a b n (gcons a' b' n' c')).
-\end{Coqsrc}
+\input{movies/snippets/T2/nfDef}
 
@@
 
-\begin{Coqsrc}
-Lemma  nf_fin i : nf (fin i).
-Proof.
-  destruct i.
-  - auto with T2.
-  - constructor 2; auto with T2.
-Qed.
-
-Lemma nf_omega : nf omega.
-Proof.  compute; auto with T2. Qed.
-
-Lemma nf_epsilon0 : nf epsilon0.
-Proof. constructor 2; auto with T2. Qed.
-
-Lemma nf_epsilon : forall alpha, nf alpha -> nf (epsilon alpha).
-Proof. compute; auto with T2. Qed.
-
-Example Ex8: nf (gcons 2 1 42 epsilon0).
-Proof.
-  constructor 3; auto with T2.
-  - apply Ex4.
-  - apply nf_fin.
-  - apply nf_fin.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/T2/nfExamples}
 
@@
 
-\begin{Coqsrc}
-Fixpoint nbterms (t:T2) : nat :=
-  match t with zero => 0
-             | gcons a b n v => (S n) + nbterms v
-  end.
-
-Fixpoint t2_length (t:T2) : nat :=
-  match t  with 
-    zero => 0
-  | gcons a b n v => 
-       nbterms (gcons a b n v) + 
-      2 * (Max.max (t2_length a)
-                              (Max.max (t2_length b) 
-                                                (t2_length_aux v)))
-  end
-with t2_length_aux (t:T2) : nat :=
- match t with 
- | zero => 0
-  | gcons a b n v =>
-           Max.max (t2_length a) 
-                            (Max.max (t2_length b) (t2_length_aux v))
- end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Compute t2_length (gcons 2 1 42 epsilon0).
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = 48 : nat
-\end{Coqanswer}
-
 \subsection{Trichotomy}
@@
 
-\begin{Coqsrc}
-Lemma tricho_aux (l: nat) : forall t t' :T2,
-      t2_length t + t2_length t' < l  ->
-      {t t2< t'} + {t = t'} + {t' t2<  t}.
-\end{Coqsrc}
-
-Then we get our version of \texttt{lt\_eq\_lt\_dec}, and derive a comparison function;
-
-\begin{Coqsrc}
-Definition lt_eq_lt_dec (t t': T2) : {t t2< t'}+{t = t'}+{t' t2<  t}.
-Proof.
-  eapply tricho_aux.
-  eapply lt_n_Sn.
-Defined.
-
-Definition compare (t1 t2 : T2) : comparison := 
-  match lt_eq_lt_dec t1 t2 with
-  | inleft (left _) => Lt
-  | inleft (right _) => Eq
-  | inright _ => Gt
-  end.
-\end{Coqsrc}
+\input{movies/snippets/Gamma0/tricho}
+
+\input{movies/snippets/Gamma0/compare}
 
@@
 
-\begin{Coqsrc}
-Fixpoint nfb (alpha : T2) : bool :=
-  match alpha with
-    zero => true
-  | gcons a b n zero => andb (nfb a) (nfb b)
-  | gcons a b n ((gcons a' b' n' c') as c) =>
-    match compare [a', b'] [a, b] with
-           Lt => andb (nfb a) (andb (nfb b) (nfb c))
-           | _ => false
-           end
-end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Compute compare (gcons 2 1 42 epsilon0) [2,2].
-\end{Coqsrc}
-
-\begin{Coqanswer}
-   = Lt
-     : comparison
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Compute nfb  (gcons 2 1 42 epsilon0).
-\end{Coqsrc}
-
-\begin{Coqanswer}
-   = true
-     : bool
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Compute nfb (gcons 2 1 42 (gcons 2 2 4 epsilon0)).
-\end{Coqsrc}
-
-\begin{Coqanswer}
-   = false
-     : bool
-\end{Coqanswer}
-
-\begin{remark}
-The connection between the predicate \texttt{nf} and the relation \texttt{lt} on one part, and the functions \texttt{nfb} and \texttt{compare} on the other, is expressed by the following lemmas:
-
-\begin{Coqsrc}
-Lemma nfb_equiv gamma : nfb gamma = true <-> nf gamma.
-
-Lemma compare_correct alpha beta :
-  CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
-              (compare alpha beta).
-\end{Coqsrc}
-
-The function \texttt{compare} helps to make easier proofs of inequalities of
-closed terms of type \texttt{T2}.
-
-First, we prove a lemma:
-
-\begin{Coqsrc}
-Lemma compare_Lt : forall alpha beta, compare alpha beta = Lt -> 
-                                         alpha t2< beta.
-Proof.
-  intros alpha beta; destruct (compare_correct alpha beta);
-    trivial; discriminate. 
-Qed.
-\end{Coqsrc}
-
-Then, we give another version of the proof of Sect.~\vref{sect:ex6-first-proof}.
-
-\begin{Coqsrc}
-Example Ex6 : gcons 1 0 12 omega t2< [0,[2,1]].
-Proof. now apply compare_Lt. Qed.
-\end{Coqsrc}
-
-\end{remark}
+\input{movies/snippets/Gamma0/nfb}
 
@@
 {\texttt{Gamma0.T2}}}
-\begin{Coqsrc}
-Fixpoint succ (a:T2) : T2 :=
- match a with zero => one
-             | gcons zero zero n c => fin (S (S n))
-             | gcons a b n c => gcons a b n (succ c)
- end.
-\end{Coqsrc}
+
+\input{movies/snippets/T2/succDef}
 
@@
 
-\begin{Coqsrc}
-Fixpoint plus (t1 t2 : T2) {struct t1}:T2 :=
-  match t1,t2 with
-  |  zero, y  => y
-  |  x, zero => x
-  |  gcons a b n c, gcons a' b' n' c' =>
-     (match compare (gcons a b 0 zero)
-                    (gcons a' b' 0 zero) with
-      | Lt => gcons a' b' n' c'
-      | Gt => gcons a b n (c + gcons a' b' n' c')
-      | Eq => gcons a b (S(n+n')) c'
-      end)
-  end
-where "alpha + beta" := (plus alpha beta): T2_scope.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Example Ex7 : 3 + epsilon0 = epsilon0.
-Proof. trivial. Qed.
-\end{Coqsrc}
+\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0T2.html\#succ}%
+  {\texttt{Gamma0.Gamma0}}}
+
+\input{movies/snippets/Gamma0/plusDef}
 
@@
 
-\begin{Coqsrc}
-Definition  phi (alpha beta : T2) : T2 :=
-  match beta with zero => [alpha, beta] 
-             | [b1, b2] => 
-               (match compare alpha b1
-                with Datatypes.Lt => [b1, b2 ]
-                | _ => [alpha,[b1, b2]]
-                end)
-             | gcons b1 b2 0 (gcons zero zero  n zero) => 
-               (match compare alpha b1
-                with  Datatypes.Lt => 
-                      [alpha, (gcons b1 b2 0 (fin n))]
-                | _ =>  [alpha, (gcons b1 b2 0 (fin (S n)))]
-                end)
-             | any_beta => [alpha, any_beta]
-  end.
-\end{Coqsrc}
+\input{movies/snippets/Gamma0/phiDef}
 
 Despite its complexity, the function \texttt{phi} is well adapted to proofs by simplification or computation.
-\begin{Coqsrc}
-Example Ex8:  phi 1 (succ epsilon0) = [1, [1,0] + 1].
-Proof. reflexivity. Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-(**  All epsilons are fixpoints of phi 0 *)
-
-Theorem epsilon_fxp : forall beta, phi zero (epsilon beta) =
-                                   epsilon beta.
-Proof. reflexivity. Qed.
-
-Theorem epsilon0_fxp : epsilon0 = phi zero epsilon0.
-Proof. apply epsilon_fxp. Qed.
-\end{Coqsrc}
 
@@
 
-\begin{Coqanswer}
-phi_psi :
-forall (alpha : T2) [beta gamma : T2] [n : nat],
-nf gamma ->
-limit_plus_fin beta n gamma ->
-phi alpha beta = beta -> [alpha, gamma] = phi alpha (succ gamma)
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Example Ex9 : [zero, epsilon 2 + 4] = phi 0 (epsilon 2 + 5).
-Proof. trivial. Qed.
-\end{Coqsrc}
+\input{movies/snippets/Gamma0/phiPsi}
+\input{movies/snippets/Gamma0/Ex9}
 
@@
 
-\begin{Coqanswer}
-phi_of_psi:
-  forall a b1 b2 : T2,
-  phi a [b1, b2] = (if lt_ge_dec a b1 then [b1, b2] else [a, [b1, b2]])
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Example Ex10 : phi omega [epsilon0, 5] = [epsilon0, 5].
-Proof. reflexivity. Qed.
-\end{Coqsrc}
+\input{movies/snippets/Gamma0/phiOfPsi}
+\input{movies/snippets/Gamma0/Ex10}
 
@@
 
-\begin{Coqsrc}
-Module G0.
-
-Class G0 := mkg0 {vnf : T2; vnf_ok : nfb vnf}.
-
-Definition lt (alpha beta : G0) := T2.lt (@vnf alpha) (@vnf beta).
-
-Definition compare alpha beta := Gamma0.compare (@vnf alpha) (@vnf beta).
-\end{Coqsrc}
-
-Then, we prove that \texttt{lt} is a well-founded strict order and that the
+\input{movies/snippets/Gamma0/G0a}
+\input{movies/snippets/Gamma0/G0b}
+
+Then, we buils an instance of class \texttt{ON}. 
 function \texttt{compare} is correct.
 
-\begin{Coqsrc}
-Instance lt_sto : StrictOrder lt.
-
-Lemma lt_wf : well_founded lt.
-
-Lemma compare_correct alpha beta :
-  CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
-              (compare alpha beta).
-
-Instance Gamma0: ON lt  compare.
-Proof.
-  split.
-  - apply lt_sto.
-  - apply lt_wf. 
-  - apply compare_correct.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/Gamma0/ltSto}
+\input{movies/snippets/Gamma0/ltWf}
+\input{movies/snippets/Gamma0/ONGamma0}
 
@@
 
-Prove that the implemantations of \texttt{succ}, \texttt{+}, $\phi_0$, etc.
+Prove that the implementations of \texttt{succ}, \texttt{+}, $\phi_0$, etc.
 are compatible in both notation systems.
@@
 {\texttt{Gamma0.T2}}}
-\begin{Coqsrc}
-(* injection from T1 *)
-
-Fixpoint T1_to_T2 (alpha :T1) : T2 :=
-  match alpha  with
-  | T1.zero => zero
-  | T1.ocons a n b => gcons zero (T1_to_T2 a) n (T1_to_T2 b)
-  end.
-\end{Coqsrc}
+
+\input{movies/snippets/T2/T1ToT2}
 

### 2021-08-27 doc/hydras.tex
Alectryon:  chapter 7 (30%)
Alectryon:  chapter 7 (100%)
[movies] Adjust to renaming from tango_subtle.sty to pygments.sty

@@
 \usepackage{movies/snippets/assets/alectryon}
-\usepackage{movies/snippets/assets/tango_subtle}
+\usepackage{movies/snippets/assets/pygments}
 
@@
 \date{\today}
+%\date{August 2021}
+
 \title{Hydra Battles and Cie}
@@
 Let us underline the analogy between hydra battles and interactive theorem proving.
-Hercules is the user (you!), and hydra's heads are the sub-goals: you may think that applying a tactic would solve a sub-goal, but it results often in the multiplication of such tasks.
+Hercules is the user (you!), and hydra's heads are the subgoals: you may think that applying a tactic would solve a subgoal, but it results often in the multiplication of such tasks.
 
@@
 
-At present, this document is a hybrid version: Chapter~\vref{chapter:primrec} is fully adapted to \emph{Alectryon}, and we just started to re-write Chapter~\vref{chapter:hydras}.
+At present, this document is a hybrid version: Chapter~\vref{chapter:hydras} to \vref{chapter:primrec} are fully adapted to \emph{Alectryon}.
+The rest of the document (Chapter \vref{chapter-powers} is still to be rewritten.
 
@@
 
-\begin{Coqsrc}
-About F_zero_eqn.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-F_zero_eqn : forall i : nat, F_ Zero i = S i
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Print Assumptions F_zero_eqn. 
-\end{Coqsrc}
-
-\begin{Coqanswer}
-Axioms:
-FunctionalExtensionality.functional_extensionality_dep
-  : forall (A : Type) (B : A -> Type) (f g : forall x : A, B x),
-    (forall x : A, f x = g x) -> f = g
-Eqdep.Eq_rect_eq.eq_rect_eq
-  : forall (U : Type) (p : U) (Q : U -> Type) (x : Q p) (h : p = p),
-    x = eq_rect p Q x p h
-\end{Coqanswer}
+\input{movies/snippets/F_alpha/DemoAssumptions}
 
@@
 
+\subsubsection{Using Alectryon}
+
+Alectryon's \emph{LaTeX} mode cannot be mistaken for our hand-made quotations, particularly if you read a color-print.
+Here is an example from Chapter~\ref{chapter:primrec}.
+
+\vspace{4pt}
+
+\input{movies/snippets/Ack/AckFixpointIterate.tex}
+
 \subsubsection{Verbatim quotations}
 
-Quotations from our  \coq{} source are displayed as follows:
+Quotations of \coq{} source and answers from Chapter~\ref{chapter-powers}, as well as definitions from Standard Library and
+other contributions are displayed as follows.
 
@@
 
-Answers from \coq{} (including sub-goals, error messages, etc.) are displayed in slanted style
+Answers from \coq{} (including subgoals, error messages, etc.) are displayed in slanted style
 with a different background color.
@@
 
+ \subsection{Remark}
+
 In general, we do not include full proof scripts in this document.
@@
 
-\begin{Coqsrc}
-Lemma Lemma2_6_1 (alpha : T1) :  
-  nf alpha -> forall beta,  beta t1< alpha  ->
-  {n:nat | const_pathS n alpha beta}.
-Proof.
-  transfinite_induction alpha.
-  (* ... *)
-\end{Coqsrc}
+\input{movies/snippets/Paths/Lemma261}
 
@@
 
-\subsubsection{Using Alectryon}
-
-Alectryon's \emph{LaTeX} mode cannot be mistaken for our hand-made quotations, particularly if you read a color-print.
-Here is an example from Chapter~\ref{chapter:primrec}.
-
-\vspace{4pt}
-
-\input{movies/snippets/Ack/AckFixpointIterate.tex}
 \subsection{Active links}
@@
 
-\begin{Coqbad}
-Module Bad.
-
-Definition double (n:nat)  := n + 2.
-
-Lemma lt_double : forall n:nat, n < double  n.
-Proof.
-   unfold double; lia.
-Qed.
-
-End Bad.
-\end{Coqbad}
+\input{movies/snippets/Schutte_basics/BadBottoma}
+\input{movies/snippets/Schutte_basics/trivialProps}
+\input{movies/snippets/Schutte_basics/Failure}
 
@@
 
-\begin{Coqalt}
-From hydras Require Import Iterates.
-Module Alt.
-  Definition double (n : nat) := iterate S n n.
-End Alt.
-\end{Coqalt}
-
-\begin{Coqsrc}
-Lemma alt_double_ok n : Nat.double n = Alt.double n.
-Proof.
-  unfold Alt.double, Nat.double; induction n; cbn.
-  -  trivial.
-  -  rewrite <- iterate_rw, iterate_S_eqn, <- IHn; lia.
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/Hydra_Definitions/HydraAlt}
 

### 2021-08-27 doc/hydras.tex
re-reding chapter 1 (introduction)

@@
 \usepackage{movies/snippets/assets/alectryon}
-\usepackage{movies/snippets/assets/pygments}
+\usepackage{movies/snippets/assets/tango_subtle}
 
@@
 Let us underline the analogy between hydra battles and interactive theorem proving.
-Hercules is the user (you!), and hydra's heads are the subgoals: you may think that applying a tactic would solve a subgoal, but it results often in the multiplication of such tasks.
+Hercules is the user (you!), and hydra's heads are the sub-goals: you may think that applying a tactic would solve a sub-goal, but it results often in the multiplication of such tasks.
 
@@
 
-Answers from \coq{} (including subgoals, error messages, etc.) are displayed in slanted style
+Answers from \coq{} (including sub-goals, error messages, etc.) are displayed in slanted style
 with a different background color.

### 2021-08-27 doc/part-hydras.tex
re-reading chapter 2 (partially)

@@
 
-Let us look at the second sub-goal of the induction.
+Let us look at the second subgoal of the induction.
 
@@
 \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
 \input{movies/snippets/Hydra_Examples/heightLtSizea}
 
-The first subgoal is easily solvable, using some arithmetic. We let the reader look at the source. The second and third are
-almost trivial.
+The first subgoal is easily solvable, using some arithmetic.
+The second and third ones are
+almost trivial. We let the reader look at the source. 
+\input{movies/snippets/Hydra_Examples/heightLtSizez}
 
@@
 \input{movies/snippets/Hydra_Definitions/S0Def}
-
 \input{movies/snippets/Hydra_Definitions/R1Def}
@@
 
-%% TODO display sub-goals when fixed
+%% TODO display subgoals when fixed
 
@@
 
-But this style leads to clumsy statements, and generates too many sub-goals in interactive proofs (although often solved with \texttt{auto} or \texttt{eauto}).
+But this style leads to clumsy statements, and generates too many subgoals in interactive proofs (although often solved with \texttt{auto} or \texttt{eauto}).
 
@@
 
-On the other hand, the following attempt fails, because of the unprovable first sub-goal (please notice that the second sub-goal is easy to solve !).
+On the other hand, the following attempt fails, because of the unprovable first subgoal (please notice that the second subgoal is easy to solve !).
 

### 2021-08-27 doc/movies/gen_targets.py
[movies] Implement a simplified driver

@@ -1,0 +1,21 @@
+#!/usr/bin/env python3
+from os import getenv, sep
+from pathlib import Path
+
+root, snippets = Path(getenv("coq_root")), Path(getenv("snippets_root"))
+
+RULE = """
+{}: {}
+	$(alectryon) "$<"
+	@touch "$@"
+"""
+
+targets = []
+
+for coqfile in root.glob('**/*.v'):
+    pth = coqfile.relative_to(root)
+    target = str(snippets / pth.stem) + sep
+    targets.append(target)
+    print(RULE.format(target, coqfile))
+
+print("targets := {}".format(" ".join(str(t) for t in targets)))

### 2021-08-27 doc/movies/driver.py
[movies] Implement a simplified driver

@@ -1,0 +1,118 @@
+#!/usr/bin/env python3
+import re
+from collections import namedtuple
+from itertools import zip_longest, groupby
+from pathlib import Path
+
+from alectryon import cli, transforms
+from alectryon.latex import ASSETS
+
+def pair(*iterable):
+    iters = [iter(iterable)] * 2
+    return zip_longest(*iters, fillvalue=None)
+
+Snippet = namedtuple("Snippet", "name io_annots s")
+
+SNIPPET_DELIM_RE = re.compile(r"""
+  [(][*]\s*
+    (?P<kind>begin|end)\s+ snippet\s+ (?P<name>[^ ]+)
+    (?:[:][:]\s+ (?P<io_flags>.*?))?
+  \s*[*][)]
+""", re.VERBOSE)
+
+SNIPPET_IO_FLAGS_RE = re.compile(r"""
+  (?:\n\s*)? # As if flags were at end of last non-empty line
+  [(][*][|]\s*
+    (?:..\s+ coq::\s+ (.*?))?
+  \s*[|]?[*][)]
+""", re.VERBOSE)
+
+class NoSnippets(Exception):
+    pass
+
+def check_snippet_boundaries(before, after):
+    assert before
+    if not after:
+        MSG = "Missing final ``(* end snippet {} *)`` delimiter"
+        raise ValueError(MSG.format(before.group("before_name")))
+    if before.group("kind") != "begin":
+        raise ValueError("Unexpected delimiter ``{}``".format(before.group()))
+    if after.group("kind") != "end":
+        raise ValueError("Unexpected delimiter ``{}``".format(before.group()))
+    if before.group("name") != after.group("name"):
+        MSG = "Mismatched delimiters: (* begin snippet {} *) and (* end snippet {} *)"
+        raise ValueError(MSG.format(before.group("name"), after.group("name")))
+
+def split_snippet(name, io_flags, contents):
+    for substr, subsnippet in pair(io_flags, *SNIPPET_IO_FLAGS_RE.split(contents)):
+        io_annots = transforms.read_all_io_flags(substr or "")
+        yield Snippet(name, io_annots, subsnippet)
+
+def parse_coq_snippets(coq, ctx):
+    end, chunks = 0, []
+    for before, after in pair(*SNIPPET_DELIM_RE.finditer(coq)):
+        check_snippet_boundaries(before, after)
+        if before.start() > end:
+            chunks.append(coq[end:before.start()])
+        name, io_flags = before.group("name", "io_flags")
[…]

### 2021-08-28 doc/part-hydras.tex
improving the pdf (up to p. 35)

@@
 
-We chose to call \emph{standard} the kind of battles which appear  most often in the literature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$
+We chose to call \emph{standard}\footnote{This appellation is ours.
+If there is a better one, we will change it.} the kind of battles which appear  most often in the literature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$
 
@@
 
+\vspace{4pt}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
+
+\noindent
 \input{movies/snippets/Hydra_Lemmas/battleTrans}
@@
 \item At each round, Hercules chops off the rightmost head of the hydra.
-\item The battle is called \emph{standard}\footnote{This appellation is ours. If there is a better one, we will change it.}: at the round number $i$, the expected replication is $i$.
+\item The battle is {standard}: at the round number $i$, the expected replication is $i$.
 \end{itemize}
@@
 
+\emph{From Module~\href{../theories/html/hydras.Hydra.BigBattle.html}{Hydra.BigBattle}}
+
 \input{movies/snippets/BigBattle/hinitDef}
@@
 
-For instance Fig~\vref{fig:hinit-plusn} shows the hydra (\texttt{hyd 3 4 2}). The hydra (\texttt{hyd 0 0 0})  is the ``final'' hydra of any terminating battle, {i.e.},
+For instance, the hydra shown in  Fig~\ref{fig:hinit-plusn} is  (\texttt{hyd 3 4 2}). The hydra (\texttt{hyd 0 0 0})  is the ``final'' hydra of any terminating battle, {i.e.},
 a tree whith exactly one node and no edge.
@@
 \draw (g) -- (ga);
-% \node (a) at  (3,4) {$\bullet$};
-% \node (h1) at  (1,6) 
-% \node (h2) at  (3,6) {$\Smiley[1][green]$};
-% \node (h3) at  (6,6) {$\Smiley[1][green]$};
-% \node (h4) at  (6,6) {$\Smiley[1][green]$};
-% \draw (foot) -- (n1);
-% \draw (n1) to   [bend left=20] (h1);
-% \draw (n1) to   (h2);
-% \draw (n1) to   [bend right=20] (h3);
 \end{tikzpicture}

### 2021-08-29 doc/hydras.tex
re-reading chapter 2 (partially)
[latex] Prevent line breaks within syntactic units in code
https://github.com/coq-community/hydra-battles/issues/70#issue-980536040

--- 2021-08-27 doc/hydras.tex
@@
 
+% Prevent breaks in the middle of syntactic units
+\let\OldPY\PY
+\def\PY#1#2{\OldPY{#1}{\mbox{#2}}}
+
 \setcounter{tocdepth}{1}
@@
 \renewcommand{\alectryon@hyps@sep}{\alectryon@nl}
+\makeatother
+
+%%% \snippets{A,B,C,…} inputs a series of snippets as one block (with \itemsep
+%%% between them).  A, B, C should be paths to files in movies/snippets/.
+\usepackage{etoolbox}
+\makeatletter
+\newcommand{\inputsnippets}[1]
+  {{\setlength{\itemsep}{1pt}\setlength{\parsep}{0pt}% Adjust spacing
+    \alectryon@copymacros\begin{io}
+      \forcsvlist{\item\@inputsnippet}{#1}
+    \end{io}}}
+\let\input@old\input % Save definition of \input
+\newcommand{\@inputsnippet}[1]
+  {{\renewenvironment{alectryon}{}{}% Skip \begin{alectryon} included in snippet
+    \input@old{movies/snippets/#1}}}
 \makeatother
@@
 \author{Pierre Castéran\\ LaBRI, Univ.
-Bordeaux, CNRS UMR 5800, \thanks{With contributions by Yves Bertot, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury and Théo Zimmermann.
+Bordeaux, CNRS UMR 5800, \thanks{With contributions by Yves Bertot, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Karl Palmskog, Clément Pit-Claudel, and Théo Zimmermann.
 The formalization of primitive recursive functions was originally authored by
@@
 
-% % \mbox{~}
-% % \clearpage
-% % \newpage
-% % \mbox{~}
-% \vspace{10cm}
-% \begin{quote} 
-% { \Large {\it
-% ``I start from one point and go as far as possible.'' }
+\tableofcontents
 
-% John Coltrane.}
-% \end{quote}
+% \mbox{~}
+\clearpage
+% \newpage
+% \mbox{~}
+\begin{quote}
+ (Aprieta el bast\'on con las dos manos, se yergue un tanto,
+casi con entusiasmo) ¡Caramba! Claro... los n\'umeros
+transfinitos, Kantor...
 
-% \begin{verse}
-% Hydra hydra hydram \\
+[ Jorge Luis Borges]
+\end{quote}
+\vspace{2cm}
+\begin{quote}
+  Nessun senso percepisce l'infinito. Nessun senso permette di concludere ch'esso esista. L'infinito, in effetti, non puo' essere l'oggetto dei sensi.
 
-% Hydrae hydrae hydra \\
+  [Giordano Bruno] \emph{De l'infinito, universo e mondi}
+\end{quote}
 
-% Hydrae hydrae hydras \\
+\vspace{2cm}
+\begin{quote} 
+I start from one point and go as far as possible.
 
-% Hydrarum hydris hydris\\
-% \end{verse}
-
-\tableofcontents
+[John Coltrane]
+\end{quote}
 
@@
 Quotations of \coq{} source and answers are progressively replaced from copy-pasted \emph{verbatim} to automatically generated \emph{LaTeX} blocks, using Clément Pit-Claudel's \emph{Alectryon} tool~\cite{alectryonpaper, alectryongithub}.
-Many thanks to Jérémy Damour and Théo Zimmermann who designed tools for maintaining consistency between the always evolving \coq{} modules and documentation written in \emph{LaTeX}.
+Many thanks to Jérémy Damour, Clément Pit-Claudel  and Théo Zimmermann who designed tools for maintaining consistency between the always evolving \coq{} modules and documentation written in \emph{LaTeX}.
 
 Besides the guarantee of consistency between theories and documentation, we hope to give a corpus for experimenting new ways of documenting the implementation of non-trivial mathematics on a proof assistant.
-
-At present, this document is a hybrid version: Chapter~\vref{chapter:hydras} to \vref{chapter:primrec} are fully adapted to \emph{Alectryon}.
-The rest of the document (Chapter \vref{chapter-powers} is still to be rewritten.
 
@@
 
-Alectryon's \emph{LaTeX} mode cannot be mistaken for our hand-made quotations, particularly if you read a color-print.
 Here is an example from Chapter~\ref{chapter:primrec}.
@@
 
-Quotations of \coq{} source and answers from Chapter~\ref{chapter-powers}, as well as definitions from Standard Library and
-other contributions are displayed as follows.
+Quotations of \coq{} source  from Standard Library and code which lead to compilation error or too long computations are displayed as follows.
 
 \begin{Coqsrc}
- Definition square (n:nat) := n * n.
+Definition fib_eucl gamma `{Hgamma: Strategy gamma} n :=
+  let c := make_chain gamma  n
+  in let r := chain_apply c (M:=Mul2) (1,0) in
+       fst r + snd r.
 
- Lemma square_double : exists n:nat, n + n = square n.
- Proof.
-    exists 2. 
+Compute fib_eucl dicho 153.
 \end{Coqsrc}
 
-Answers from \coq{} (including subgoals, error messages, etc.) are displayed in slanted style
-with a different background color.
-
 \begin{Coqanswer}
- 1 subgoal, subgoal 1 (ID 5)
-
-  ============================
-   2 + 2 = square 2
-
+    = 68330027629092351019822533679447
+     : N
 \end{Coqanswer}
-
- \begin{Coqsrc}
-   reflexivity.
-Qed.
- \end{Coqsrc}
 

### 2021-08-30 doc/hydras.tex
[latex] Prevent line breaks within syntactic units in code
https://github.com/coq-community/hydra-battles/issues/70#issue-980536040
[doc] Add an \inputsnippets command and demo it in Schutte/AP.v
update list of contributors

--- 2021-08-27 doc/hydras.tex
@@
 \author{Pierre Castéran\\ LaBRI, Univ.
-Bordeaux, CNRS UMR 5800, \thanks{With contributions by Yves Bertot, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury and Théo Zimmermann.
+Bordeaux, CNRS UMR 5800, \thanks{With contributions by Yves Bertot, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Clément Pit-Claudel, and Théo Zimmermann.
 The formalization of primitive recursive functions was originally authored by

### 2021-08-30 doc/part-hydras.tex
small issue in snippet Omega_Small/mGeb
[doc] Add an \inputsnippets command and demo it in Schutte/AP.v

@@
 
-The relation between \texttt{one\_step} and the rules of hydra battle is asserted by the following lemma.
+The relation between \texttt{one\_step} and the rules of hydra battles is asserted by the following lemma.
 
@@
 \vspace{4pt}
+\noindent
+
 \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
@@
 
-\input{movies/snippets/Omega_Small/mGeb}
-
-\input{movies/snippets/Omega_Small/omegaSmallz}
+\inputsnippets{Omega_Small/mGeb, omega_Small/mGez,
+  Omega_Small/omegaSmallz}
 
@@
 
-\input{movies/snippets/ON_Generic/ONDefs}
+\inputsnippets{ON_Generic/ONDefsa, ON_Generic/ONDefsb}
 
@@
 
-\input{movies/snippets/ON_Omega/OmegaDef}
+\inputsnippets{ON_Omega/OmegaDefa, ON_Omega/OmegaDefb}
 
@@
 \end{Coqanswer}
+
+\pagebreak
 
@@
 
-\input{movies/snippets/AP/APOne}
+\inputsnippets{AP/APOne,AP/APOne_least_AP,AP/APOne_AP_omega,AP/APOne_omega_second_AP}
 

### 2021-08-30 doc/part-hydras.tex
using inputsnippets (chapter 3, up to page 59 of the pdf)

@@
 
-\input{movies/snippets/ON_plus/compareDef}
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_plus/compareCorrect}
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_plus/plusComp}
-
-\vspace{4pt}
+\inputsnippets{ON_plus/compareDef,
+  ON_plus/compareCorrect, ON_plus/plusComp}
 
 The Lemma \texttt{Wellfounded.Disjoint\_Union.wf\_disjoint\_sum} of Standard Library
-helps us to prove that our order \texttt{lt} is well-founded.
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_plus/ltWf}
-
-\vspace{4pt}
-
-Then, we can build an instance of \texttt{ON}:
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_plus/OnPlus}
+helps us to prove that our order \texttt{lt} is well-founded, thenwe can build an instance of \texttt{ON}:
+
+\inputsnippets{ON_plus/ltWf, ON_plus/OnPlus}
 
@@
 
-\input{movies/snippets/ON_mult/Defs}
+\inputsnippets{ON_mult/Defs}
+\inputsnippets{ON_mult/multComp,  ON_mult/ONMult}
+\inputsnippets{ON_mult/endDefs}
 
@@
 
-\input{movies/snippets/ON_Omega2/Omega2Def}
-
-\input{movies/snippets/ON_Omega2/Defs}
+\inputsnippets{ON_Omega2/Omega2Def, ON_Omega2/Defs}
 
@@
 
-\input{movies/snippets/ON_Omega2/Defs}
+\input{movies/snippets/ON_Omega2/succ}
 
@@
 
-\input{movies/snippets/ON_Omega2/plusDef}
+\inputsnippets{ON_Omega2/plusDef, ON_Omega2/plusExamples}
 
@@
 \input{movies/snippets/ON_Omega2/multFinDef}
-
-\vspace{4pt}
+\input{movies/snippets/ON_Omega2/multFinExamples}
 
@@
 
-\input{movies/snippets/ON_Omega2/uniqueDecomposition}
+\inputsnippets{ON_Omega2/uniqueDecompositiona,
+  ON_Omega2/uniqueDecompositionb,
+  ON_Omega2/uniqueDecompositionz}
 

### 2021-08-31 doc/part-hydras.tex
use of inputsnippets in Chap3
correct a typo in part-hydras.tex

@@
 
-\inputsnippets{Omega_Small/mGeb, omega_Small/mGez,
+\inputsnippets{Omega_Small/mGeb, Omega_Small/mGez,
   Omega_Small/omegaSmallz}
@@
 
-\input{movies/snippets/ON_Omega2/Merge}
+\inputsnippets{ON_Omega2/Merge, ON_Omega2/Mergeb}
 
@@
 
+\subsubsection{Examples}
+
 For instance, let us build two elements of the segment $[0, 7)$, \emph{i.e.} two
@@
 
- \input{movies/snippets/ON_Finite/t0Empty}
-
-\begin{Coqsrc}
-Program Definition bad : t 10 := 10.
-Next Obligation.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-1 subgoal (ID 118)
-
-  ============================
-  10 <? 10 
-\end{Coqanswer}
-
-\begin{Coqsrc}
-  compute.
-Abort.
-\end{Coqsrc}
-
-%\input{movies/snippets/ON_Finite/bad}
+ \inputsnippets{ON_Finite/t0Empty, ON_Finite/bad10}
+
+\subsubsection{Comparison function}
 
@@
 
-\begin{Coqanswer}
- 1 subgoal (ID 110)
-
-  n, x0 : nat
-  i, i0 : x0 <? S n
-  ============================
-  exist (fun i1 : nat => i1 <=? n) x0 i =
-  exist (fun i1 : nat => i1 <=? n) x0 i0
-\end{Coqanswer}
+\input{movies/snippets/ON_Finite/compareCorrectb}
 
@@
 
-\begin{Coqanswer}
-1 subgoal (ID 112)
-
-  n, x0 : nat
-  i, i0 : x0 <? S n
-  ============================
-  i = i0
-\end{Coqanswer}
+\input{movies/snippets/ON_Finite/compareCorrectc}
 
@@
 
+\inputsnippets{ON_Finite/compareCorrectd}
+
 \end{remark}
+
+\begin{remark}
 
@@
 \input{movies/snippets/ON_Finite/Example2}
+
+\end{remark}
+
+\subsubsection{Building an instance of \texttt{ON}}
 
@@
 
-\vspace{4pt}
+\vspace{10pt}
+
 \noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
@@
 
-\vspace{4pt}
+\vspace{10pt}
+
 \noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
 
-\input{movies/snippets/ON_Finite/InclIJ}
-\input{movies/snippets/ON_Finite/InclIJa}
-\input{movies/snippets/ON_Finite/InclIJb}
-\input{movies/snippets/ON_Finite/InclIJc}
-\input{movies/snippets/ON_Finite/InclIJd}
+\inputsnippets{ON_Finite/InclIJ, ON_Finite/InclIJa,
+  ON_Finite/InclIJb,ON_Finite/InclIJc, ON_Finite/InclIJd,
+ON_Finite/InclIJz}
 
@@
 
-%%% TODO : link to Comparable and StrictOrder classes
-
 Links between the function \texttt{compare} and the relations
@@
 \input{movies/snippets/T1/Instances}
-
-% \vspace{4pt}
-% \emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare_refl}{Epsilon0.T1}}
-
-% \begin{Coqsrc}
-% Lemma compare_refl : forall alpha, compare alpha alpha =  Eq.
-% \end{Coqsrc}
-
-% \begin{Coqsrc}
-% Lemma compare_reflect : forall alpha beta,
-%     match compare alpha beta with
-%     |   Lt => lt alpha  beta
-%     |   Eq => alpha = beta
-%     |   Gt => lt beta  alpha
-%     end.
-% \end{Coqsrc}
-
-% We prove also that the relation \texttt{lt} is a strict total order.
-
-% \vspace{4pt}
-% \emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#lt_irrefl}{Epsilon0.T1}}
-
-% \begin{Coqsrc}
-% Theorem lt_irrefl (alpha: T1):  ~ lt alpha alpha.
-
-% Theorem lt_trans (alpha beta gamma : T1) :
-%   lt alpha  beta -> lt beta gamma -> lt alpha gamma.
-
-% Definition lt_eq_lt_dec  :
-%    forall alpha beta : T1, 
-%           {lt alpha  beta} + {alpha = beta} + {lt beta alpha}.
-% \end{Coqsrc}
-
-% \vspace{4pt}
-
-% Please note that the order \texttt{lt} is not reflected 
-% in the structure (size and/or height) of the terms of \texttt{T1}.
-
-% \begin{Coqsrc}
-% Example Ex0:
-%   lt (ocons (phi0 (phi0 omega)) 2
-%             (ocons (phi0 10) 33
-%                    (ocons (phi0 9) 63 zero)))
-%      (ocons  (phi0 (phi0 omega)) 2 (phi0 (phi0 11))).
-% Proof.
-%   reflexivity. 
-% Qed.
-% \end{Coqsrc}
-
-% Uncomment for testing 
-%\textbf{Todo: Fix subgoal display }
-%\input{movies/snippets/T1/compareRev}
 

### 2021-09-01 doc/part-hydras.tex
revision chapter 4 (up to page 91)

@@
 \label{alpha0-def}
-For instance, the ordinal  $\omega^\omega+\omega^3\times 5+2$ is represented by the following term:
+For instance, the ordinal  $\omega^\omega+\omega^3\times 5+2$ is represented by the following term of type \texttt{T1}:
 
@@
 defined by a Boolean function, for  any term $\alpha:\texttt{T1}$, there exists at most one proof of \texttt{nf $\alpha$}, thus two ordinals of type \texttt{E0} are
-equal if and only iff their projection to \texttt{T1} are equal (see also Sect.~\vref{sect:eq-proof-unicity}).
+equal if and only if their projection to \texttt{T1} are equal (see also Sect.~\vref{sect:eq-proof-unicity}).
 
@@
 
-\input{movies/snippets/E0/nfProofUnicity}
-
-\vspace{4pt}
-
-\input{movies/snippets/E0/E0EqIff}
-
-\vspace{4pt}
+\inputsnippets{E0/nfProofUnicity}
+\inputsnippets{E0/E0EqIff}
 
@@
 
+The function \texttt{succ} on \texttt{T1} is extended to \texttt{E0} the following way:
+
+ \emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
+
+ \inputsnippets{E0/SuccOnE0}
+
 \index{hydras}{Exercises}
@@
 
-\input{movies/snippets/T1/succIsPlusOne}
+\inputsnippets{T1/succIsPlusOne, T1/succIsPlusOnez}
 
@@
 
-\input{movies/snippets/T1/transfiniteRecursor}
-
-The following tactic starts a proof by  transfinite induction on any ordinal \mathcolor{$\alpha<\epsilon_0$}.
-
-\input{movies/snippets/T1/transfiniteInduction}
+Traditionnally, well-founded induction on ordinals is called \emph{transfinite induction}.
+
+\inputsnippets{T1/transfiniteRecursor, T1/transfiniteInduction}
+
+As a corollary, the order \texttt{Lt} on type \texttt{Ord} is well-founded too.
+
+\inputsnippets{E0/E0LtWf}.
 
@@
 
-We build an instance of \texttt{ON}, and prove its correction w.r.t. Schutte's model.
+We build an instance of \texttt{ON}, and prove its correctness w.r.t. Schutte's model.
 
@@
     In Module   \href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/OrdinalNotations/OmegaOmega.v}{theories/ordinals/OrdinalNotations/OmegaOmega.v},
-    we represent ordinals below $\omega^\omega$ by list of pairs of natural numbers (with the same coefficient shift as in \texttt{T1}).
+    we represent ordinals below $\omega^\omega$ by lists of pairs of natural numbers (with the same coefficient shift as in \texttt{T1}).
     For instance, the ordinal $\omega^4\times 10 + \omega^3 + \omega+ 5$ is represented by the list \texttt{(4,9)::(3,0)::(1,0)::(0,4)::nil}.
 
+\inputsnippets{OmegaOmega/LODef}
+
   The usual operations : \texttt{succ}, \texttt{+}, \texttt{*} are simple variants of the same operations in \texttt{T1}.
+
+  \inputsnippets{OmegaOmega/succPlusMult}
 
@@
 Thus, many properties like well-foundedness of $<$ and associativity of $+$,  of this ordinal notations have very short proofs.
+
+    \inputsnippets{OmegaOmega/refineDef, OmegaOmega/phi0Ref,
+      OmegaOmega/succRef, OmegaOmega/plusRef,
+      OmegaOmega/multRef}
+
+    In order to make an ordinal notation for $\omega^\omega$, we follow the same steps as for $\epsilon_0$:
+    \begin{enumerate}
+    \item Define an order \texttt{lt}, which refines the order
+      \texttt{lt} on \texttt{T1}.
+      \inputsnippets{OmegaOmega/compareDef}
+      \inputsnippets{ OmegaOmega/ltDef, OmegaOmega/ltRef}
+
+    \item Define the predicate ``to be in normal form''.
+      \inputsnippets{OmegaOmega/nfDef}
+    \item Define a class \texttt{OO} of terms in normal form, and an embedding from \texttt{E0} into \texttt{OO}.
+      \inputsnippets{OmegaOmega/OODef}
+       \inputsnippets{OmegaOmega/embedDef}
+      \item Infer well-foundedness of the order on \texttt{OO}.
+    \inputsnippets{OmegaOmega/ltWf}
+
+    \end{enumerate}
+
+    \inputsnippets{OmegaOmega/ONOO, OmegaOmega/OOz}
+
+    Let us show a few examples.
+
+    \inputsnippets{OmegaOmega/OODemo}
 

### 2021-09-02 doc/part-hydras.tex
Corrections in chapter 4
Simplification of old latex snippets

@@
 
+  \pagebreak 
   \vspace{4pt}
@@
     A first set of lemmas correspond to the various cases of \texttt{oplus}'s 
-    definition. They can be proved almost immediately.
-
-    \input{movies/snippets/Hessenberg/oplusNeutral}
+    definition. They can be proved almost immediately. Here are a few examples.
+
+    \inputsnippets{Hessenberg/oplusNeutral,
+      Hessenberg/oplusCompareLt}
 
@@
     We prove  the commutativity of $\oplus$ in two steps. 
-
     First, we prove by transfinite induction on $\alpha$ that the restriction of $\oplus$ to the
@@
     \input{movies/snippets/Hydra_Termination/mNf}
+      \input{movies/snippets/Hydra_Termination/msNf}
 
@@
 The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
-is the limit of the sequence \showmath{\canonseq{\lambda}{i}\;(1\le i)}.
+is the limit of the sequence $\canonseq{\lambda}{i}\;(1\le i)$.
 
@@
 
-\begin{Coqsrc}
- path_to omega (2::2::2::nil) (omega * 2).
-
- path_to omega (3::4::5::6::nil) (omega * 2).
-
- path_to zero (interval 3 14) (omega * 2).
-
- path_to zero (repeat 3 8) (omega * 2).
-\end{Coqsrc}
-
+\inputsnippets{Paths/exPathTac} 
 \end{exercise}

### 2021-09-03 doc/hydras.tex
correct a typo in part-hydras.tex
revision chapter 4 (up to page 91)
Corrections Chapter 5

@@
 
+%%% \snippets{A,B,C,…} inputs a series of snippets as one block (with \itemsep
+%%% between them).  A, B, C should be paths to files in movies/snippets/.
+\usepackage{etoolbox}
+\makeatletter
+\newcommand{\inputsnippets}[1]
+  {{\setlength{\itemsep}{1pt}\setlength{\parsep}{0pt}% Adjust spacing
+    \alectryon@copymacros\begin{io}
+      \forcsvlist{\item\@inputsnippet}{#1}
+    \end{io}}}
+\let\input@old\input % Save definition of \input
+\newcommand{\@inputsnippet}[1]
+  {{\renewenvironment{alectryon}{}{}% Skip \begin{alectryon} included in snippet
+    \input@old{movies/snippets/#1}}}
+\makeatother
+
 \SetWatermarkLightness{0.8}

### 2021-09-03 doc/part-hydras.tex
Corrections Chapter 5

@@
 
-\input{movies/snippets/Epsilon0_Needed_Free/mGe}
+\pagebreak
+
+\inputsnippets{Epsilon0_Needed_Free/mGe}
 
@@
 \vspace{4pt}
+
 \emph{From
@@
 
- We build the following paths :
+The proof of the limit case, is decomposed into a sequence
+of path constructions leading to $\alpha \xrightarrow[n]{} \beta$.
 
  \begin{enumerate}
- \item $\alpha \xrightarrow[n]{} \canonseq{\alpha}{n}$
+ \item $\alpha \xrightarrow[n]{} \canonseq{\alpha}{n}$ (single step path)
  \item $\canonseq{\alpha}{n} \xrightarrow[1]{} \canonseq{\alpha}{i}$ (by \texttt{Theorem\_2\_4}),
 \item $\canonseq{\alpha}{n} \xrightarrow[n]{} \canonseq{\alpha}{i}$ (applying the induction hypothesis to the preceding path);
-\item $\canonseq{\alpha}{i} \xrightarrow[n]{} \beta$ (applying the induction hypothesis)\item $\alpha \xrightarrow[n]{} \beta$ (by composition of 1, 3, and 4).
+\item $\canonseq{\alpha}{i} \xrightarrow[n]{} \beta$ (applying the induction hypothesis)
+\item $\alpha \xrightarrow[n]{} \beta$ (by composition of 1, 3, and 4).
 
@@
 \end{itemize}
+
+Why is \texttt{Cor12} so useful? 
+Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
+we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying corollary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
+where $n$ is the maximum of the indices $i$ met in the induction.
+
+ Lemma 1, Section 2.6 of~\cite{KS81} is naturally expressed in terms of \coq's
+\verb@sig@ construct.
+
+\label{lemma:L-2_6-1}
+\index{coq}{Sigma types}
+
+\input{movies/snippets/Paths/Lemma261}
+
+Intuitively, lemma   \texttt{Lemma2\_6\_1}  shows that if $\beta<\alpha<\epsilon_0$, then there exists  a battle from $\iota(\alpha)$ to $\iota(\beta)$ where the replication factor is constant, although large enough.
 
@@
 
-Why is \texttt{Cor12} so useful? 
-Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
-we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying corollary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
-where $n$ is the maximum of the indices $i$ met in the induction.
-
- Lemma 1, Section 2.6 of~\cite{KS81} is naturally expressed in terms of \coq's
-\verb@sig@ construct.
-
-\label{lemma:L-2_6-1}
-\index{coq}{Sigma types}
-
-\input{movies/snippets/Paths/Lemma261}
-
-Intuitively, lemma   \texttt{Lemma2\_6\_1}  shows that if $\beta<\alpha<\epsilon_0$, then there exists  a battle from $\iota(\alpha)$ to $\iota(\beta)$ where the replication factor is constant, although large enough.
-
 \subsection{Casting paths with a constant index into a standard path}
@@
 The article~\cite{KS81} contains 
-the following lemma, the proof of which is quite complex, which allows to simulate $i$-paths by $[i+1,j]$-paths, where $j$ is large enough.
+the following lemma, which allows us to simulate $i$-paths by $[i+1,j]$-paths, where $j$ is large enough.
 
@@
 \begin{itemize}
-\item If $\beta=\gamma$, its OK! We have got a standard path
+\item If $\beta=\gamma$, it's OK! We have got a standard path
 from  
@@
  \input{movies/snippets/Paths/constantToStandard0}
+ \input{movies/snippets/Paths/constantToStandardz}
 
@@
 
-Next, please consider the following context:
+Now, please consider the following context:
 

### 2021-09-04 doc/part-hydras.tex
corrections chap6

@@
 
-\subsection{Example}
-
-For instance the interval $[6,70]$ leads $\omega^2$ to $\omega\times 2 + 56$. Thus this interval is not $\omega^2$-mlarge.
+\subsection{Examples}
+
+For instance the interval $[6,70]$ leads $\omega^2$ to $\omega\times 2 + 56$. Thus this interval is not $\omega^2$-large.
 
@@
 
-Let us try another computation.
+The interval $[6,700]$ is $\omega^2$-large, but not
+$\omega^2$-mlarge, since $[6,699]$ is also $\omega^2$-large.
+
 \input{movies/snippets/Large_Sets_Examples/gnawEx2}
@@
 
-For instance, we can verify that the interval $[6,510]$ is $\omega^2$-mlarge.
+For instance, we prove that the interval $[6,510]$ is $\omega^2$-mlarge.
 
@@
  Check if the functions $L_\alpha$ are the same as
-\cite{KS81}' functions $f\_alpha$ (p. 297).
+\cite{KS81}' functions $f_\alpha$ (p. 297).
 \end{todo}
@@
 
-Our exploration of the $L_\alpha$\,s  follows the usual scheme : transfinite induction, and proof-by-cases : zero, successors and limit ordinals.
+The properties we consider are defined in \href{../theories/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}.
+
+\label{sect:abstract-arith-prop}
+\input{movies/snippets/Iterates/funLeDef}
+
+Our exploration of the $L_\alpha$\,s  considers the usual cases of a proof by transfinite induction: zero, successors and limit ordinals.
+The lemmas we are going to proved will be applied in a big proof by induction in Sect~\vref{sect:L-correct-proof}.
 
@@
 
-Useful abstract properties of arithmetic functions are
-defined 
-in~\href{../theories/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}.
-
-\label{sect:abstract-arith-prop}
-
-\input{movies/snippets/Iterates/funLeDef}
-
-We prove also that the functions we consider are strictly monotonous. The section on successor ordinals has the following structure.
-
 \emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
@@
 
-\input{movies/snippets/Large_Sets/phi0Mult}
+\inputsnippets{Large_Sets/phi0Mult, Large_Sets/phi0MultOk,
+Large_Sets/Phi0MultSLe}
 
@@
 
-\input{movies/snippets/L_alpha/Paraphrases}
+\inputsnippets{L_alpha/Paraphrasesa}
+\inputsnippets{L_alpha/Paraphrasesb}
+\inputsnippets{L_alpha/Paraphrasesc}
 
@@
 
-By  well-founded induction on $\alpha$, we prove the following lemmas:
+By  well-founded induction on $\alpha$, we prove the following properties:
 
@@
 \input{movies/snippets/L_alpha/LCorrect}
+
+\label{sect:L-correct-proof}
 
@@
 
-\index{hydras}{Projects}
-\begin{project}
-Instead of considering standard paths and battles, consider ``constant'' paths and the corresponding battles.
+\index{hydras}{Exercises}
+\begin{exercise}
+Instead of considering standard paths and battles, consider battles where the replication factor is a constant $k$.
 Please use \texttt{Equations} in order to define the function that computes the length of the $k$-path which leads  from $\alpha$ to $0$.
 Prove a few  exact formulas and minoration lemmas.
-\end{project}
+\end{exercise}
 
@@
 
-\input{movies/snippets/Hprime/paraphrases}
+\input{movies/snippets/Hprime/paraphrasesa}
+\input{movies/snippets/Hprime/paraphrasesb}
+\input{movies/snippets/Hprime/paraphrasesc}
+\input{movies/snippets/Hprime/paraphrasesd}
 
@@
 
-\input{movies/snippets/Hprime/HprimeExamples}
+\input{movies/snippets/Hprime/HprimeExamplesa}
+\input{movies/snippets/Hprime/HprimeExamplesb}
+\input{movies/snippets/Hprime/HprimeExamplesc}
+\input{movies/snippets/Hprime/HprimeExamplesd}
 
@@
 
-It looks hard to capture the complexity of this function by looking only at this
+It is hard to capture the complexity of this function by looking only at this
 ``exact'' formula. 
@@
 
-Now, we have got at last an exact formula for $H'_{\omega^\omega}$.
+\subsubsection{A formula for \texorpdfstring{$H'_{\omega^\omega}$}{\texttt{H'(phi0(omega))}}}
+
+Now, we can get at last an exact formula for $H'_{\omega^\omega}$.
 
@@
 
-In \coq{}, we follow the  proof in~\cite{KS81}. This proof is mainly a single  proof by transfinite induction on $\alpha$ of the conjunction of the five properties.
+In \coq{}, we follow the  proof of~\cite{KS81}. This proof is mainly a single  proof by transfinite induction on $\alpha$ of the conjunction of the five properties.
 For each $\alpha$, the three cases : $\alpha=0$, $\alpha$ is a limit, and 
@@
 \input{movies/snippets/Hprime/PAP}
-
-% By elimination, we get a catalogue of properties of the functions $L'_\alpha$:
-
-% \begin{Coqsrc}
-% Section Abstract_Properties.
-%  Variable alpha : E0.
-
-%  Theorem H'_alpha_mono : strict_mono (H'_ alpha).
-
-%  Theorem H'_alpha_gt : alpha <> Zero ->
-%         forall n, n < H'_ alpha n.
-
-%  Theorem H'_alpha_Succ_le :
-%         H'_ alpha <<= H'_ (Succ alpha).
-
-%  Theorem H'_alpha_dom :
-%     dominates_from 1 (H'_ (Succ alpha)) (H'_ alpha).
-
-%  Theorem H'_restricted_mono_l :
-%     forall beta n, Canon_plus n alpha beta -> 
-%                                  H'_ beta n <= H'_ alpha n.
-
-%   Lemma H'_alpha_ge_id : id <<= H'_ alpha.
-
-%   Lemma H'_alpha_mono_weak : forall k l, k <= l ->
-%                                         H'_ alpha k <= H'_ alpha l.
-
-% End Abstract_Properties.
-% \end{Coqsrc}
 
@@
 $H'_\alpha$.
-We let the reader look at its proof (Section \texttt{Proof\_of\_H'\_mono\_l} of \href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}).
+We let the reader look at the proof (Section \texttt{Proof\_of\_H'\_mono\_l} of \href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}).
 
@@
 Ketonen and Solovay introduce in~\cite{KS81} a ``trivial'' variant of the Wainer hierarchy~\cite{BW85, Wainer1970} of fast growing functions, indexed by ordinals below $\epsilon_0$.
+The functions $F_\alpha$ are defined by the following equations.
 
@@
 $f_{\beta+1}(i) = (f_\beta)^{(i)}(i)$ and not
-$f_{\beta+1}(i) = (f_\beta)^{(i+1)}(i)$. A module about 
+$f_{\beta+1}(i) = (f_\beta)^{(i+1)}(i)$.
+
+A module about 
 the classic Wainer hierarchy is in preparation.
+
+Note also that \cite{KS81} defines also $F_{\epsilon_0}$ (by the third equation).
+Since $\epsilon_0$ is not representable in type \texttt{E0}, our implementation in \coq{} does not take $F_{\epsilon_0}$ into account.
+
 \end{remark}
 
-A first attempt is to write a definition of $F_\alpha$ by equations, in the same way as for $H\_alpha$.  We use the functional \texttt{iterate} defined in 
-Module~\href{../theories/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
+A first attempt is to write a definition of $F_\alpha$ by equations, in the same way as for $H\_alpha$ (the functional \texttt{iterate} has already been used in Sect.\vref{Functions:iterate}).
 
@@
 
-The following code comes from 
- \href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}.
+\emph{From\href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}~}.
 
@@
 
-\input{movies/snippets/F_alpha/goodDef}
-
-It is quite easy to prove that our function \texttt{F\_} satisfies the equations on page~\pageref{sect:F-equations}.
+\input{movies/snippets/F_alpha/goodDefa}
+\input{movies/snippets/F_alpha/goodDefb}
+
+It is quite easy to prove that our functional \texttt{F\_} satisfies the equations on page~\pageref{sect:F-equations}.
 \index{hydras}{Library Prelude!iterate}
@@
 
-As a corollary, we prove the following proposition, p. 284 of~\cite{KS81}.
-\begin{quote}
-  If $\beta<\alpha$, $F_\alpha$ dominates $F_\beta$.
-\end{quote}
+As a corollary, we prove that, if $\beta<\alpha$, then
+$F_\alpha$ dominates $F_\beta$ (p. 284 of~\cite{KS81}).
 
@@
 
-\section{Conclusion}
+\section{More about rapidly growing functions}
 
@@
 This proof uses properties of the Ackermann function, and the $H'_\alpha$, $F_\alpha$, $L_\alpha$ families of functions.
-
-%%% TODO : Check the statement !
-%
-% \index{hydras}{Exercises}
-% \begin{exercise}
-% Let us quote a theorem from ~\cite{KS81} (page 297).
-
-% \begin{quote}
-% \begin{align*}
-%   H'_{\omega^\alpha}(n+1) &\geq F_{\alpha}(n) \quad (n\geq 1, \alpha<\epsilon_0) \\
-%  F_{\alpha}(n+1) &\geq H'_{\omega^\alpha}(n) \quad (n\geq 1, \alpha<\epsilon_0) 
-% \end{align*}
-% Thus $H'_{\omega^\alpha}$ and $F_{\alpha}$ have essentially the same order of growth.
-
-% \end{quote}
-
-%  But, before trying to prove these facts, look at the definition of function $H$ in Ketonen and Solovay's paper !
-Is it really the same as the definition we quote from Pr{\H o}mel's chapter~\cite{Promel2013},
-% whereas \cite{KS81} define $H_\alpha(n)$ as ``the least integer $k$ such that $[n,k]$ is $\alpha$-large''. Thus, it may be useful to adapt the statement above.
-
-% \end{exercise}
-
-%  \subsection{Gnawing ordinals}
-
-% \begin{definition}[After~\cite{KS81}]
-%   Let $S$ be a finite set of positive integers, and $\alpha$ be an ordinal strictly less than $\epsilon_0$.  Let us denote by $s=s_1,s_2, \dots, s_N$ the sequence of 
-% the elements of $S$, enumerated in strictly increasing order.
-
-% We consider the sequence of ordinals $\alpha_o=\alpha, \alpha_1=\canonseq{\alpha_0}{s_1},\dots,\alpha_{i+1}=\canonseq{\alpha_i}{s_{i+1}},\dots, 
-% \alpha_{N}=\canonseq{\alpha_{N-1}}{s_N}$.
-% We denote by $\gnaw{s}{\alpha}$ the last ordinal of the sequence, \emph{i.e.}  $\alpha_N$.
-% \end{definition}
-
-% The following function computes $\gnaw{s}{\alpha}$ by recursion on $s$.
-% \vspace{4pt}
-
-% For instance, let us consider the ordinal $\omega^\omega$, and try some sequence of integers.
-
-% \begin{Coqsrc}
-% Compute pp (gnaw (omega ^ omega) (1::nil)).  
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = omega%pT1 : ppT1
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (gnaw (omega ^ omega) (1::2::nil)).  
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% = P_fin 2
-%      : ppT1
-% \end{Coqanswer}
-
-% Likewise, we can verify that $\gnaw{\omega^\omega}{\langle 1,2,3 \rangle}=1$
-% and $\gnaw{\omega^\omega}{\langle 1,2,3,4 \rangle}=0$.
-
-% \begin{Coqsrc}
-% Example omega_omega_1_4 : gnaw (omega ^omega) (interval 1 4) = 0.
-% Proof. trivial. Qed.
-
-% Example omega_omega_1_3 : gnaw (omega ^omega) (interval 1 3) = 1.
-% Proof. trivial. Qed.
-% \end{Coqsrc}
-
-% \begin{remark}
-% \label{remark:gnaw-vs-battles}
-% Let us consider an hydra battle, where Hercules always chops off the rightmost head of the hydra. Let $\alpha<\epsilon_0$ be an ordinal number, and 
-% $s=\langle i_1<i_2<\dots<i_N\rangle$ be  any finite sequence of positive integers.
-% Then the battle initiated by the hydra $\iota(\alpha)$, with $s$ as the sequence of successive replication factors, leads to the hydra $\iota(\gnaw{s}{\alpha})$ as the final state.
-% \end{remark}
-
-% \subsection{Large sequences}
-
-% \begin{remark}
-% In their article~\cite{KS81}, Ketonen and Solovay use the appellation ``large set'' instead of ``large sequence'', but their definitions use an enumeration of the elements in increasing order.
-Thus, we shall use the term ``sequence'' when referring to our implementation in \coq{}, and ``set'' when referring to the statements of~\cite{KS81}.
-% \end{remark}
-
-% \begin{definition}[After\cite{KS81}]
-% The sequence $s$ is said to be \emph{$\alpha$-large} if $\gnaw{\alpha}{s}=0$.  
-% \end{definition}
-
-% \begin{Coqsrc}
-% Definition largeb (alpha : T1) (s: list nat) :=
-%   match gnaw alpha s with
-%     | zero => true
-%     | _ => false
-%   end.
-
-% Definition large (alpha : T1) (s : list nat) : Prop :=
-%   largeb alpha s.
-% \end{Coqsrc}
-
-% For instance, the sequence $\langle 1,2,3,4 \rangle$ is $\omega^\omega$-large but not
-% $\omega^{\omega+1}$-large, since $\gnaw{\omega^{\omega+1}}{\langle 1,2,3,4\rangle}$ is equal to $\omega \times 2 + 4$.
-
-% %\omega^\omega+\omega^2\times 3 + \omega\times 4+ 5$.
-
-% \begin{Coqsrc}
-% Compute largeb (omega^omega) (interval 1 4).  
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% = true : bool
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute largeb (omega^(omega+1)) (interval 1 4).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% = false : bool
-% \end{Coqanswer}
-
-% \begin{remark}
-% A sequence $s$ is $\alpha$-large if (still considering Hercules ``righmost-head'', tactic), it leads to Hercules' victory.
-% \end{remark}
-
-% \subsection{A little game}
-
-% Let $\alpha<\epsilon_0$ be an ordinal, and $i$ a positive integer. We want to guess the least natural number  $j$  such that the interval $[i,i+j]$ is $\alpha$-large.
-% Equivalently, we should have $\gnaw{\alpha}{[i, i+j-1]}=1$.
-
-% The following functions takes three arguments: an ordinal $\alpha$, and two positive natural numbers  $i$ and $j$ (we assume, but not verify that $i$ and $j$ are strictly positive).
-It returns one of the three possible answers:
-
-% \begin{itemize}
-% \item \texttt{Ok} if $j$ is the smallest integer such that the interval $[i,i+j]$ is $\alpha$-large
-% \item \texttt{Too\_far} if $[i,i+j]$ is  $\alpha$-large, but $j$ is not the smallest such positive integer
-% \item \texttt{(Remaining $\beta$)} if $j$ is too small, and gnawing $\alpha$ with
-%   $[i,i+j]$ is still equal to $\beta$, instead of $0$
-% \end{itemize}
-
-% \vspace{4pt}
-% \noindent
-% \emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets_Demo.html}{Ordinals.Epsilon0.Large\_Sets\_Demo}}
-
-% \begin{Coqsrc}
-% Inductive answer : Set := You_won | Too_far | Remaining (rest : ppT1).
-
-% Definition large_set_check alpha i j :=
-%   let beta := gnaw alpha (interval i (Nat.pred j))
-%   in match beta with
-%      | one => Ok
-%      | zero => Too_far
-%      |  _ => Remaining (pp (canonseq j beta))
-%      end.
-% \end{Coqsrc}
-
-% \subsubsection{\texorpdfstring{$\omega$}{omega}-large intervals}
-
-% For instance, let us consider the ordinal $\omega$ and start with $i=1$.
-
-% \begin{Coqsrc}
-% Compute large_set_check omega 1 2.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%      = Ok
-%      : answer
-% \end{Coqanswer}
-
-% Let us give greater values of $i$, still with the ordinal $\omega$.
-
-% \begin{Coqsrc}
-% Compute large_set_check omega 2 3.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = Remaining 1
-%      : answer
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check omega 3 3.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = Remaining 3
-%      : answer
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check omega 2 4.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = Ok
-%      : answer
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check omega 3 6.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = Ok
-%      : answer
-% \end{Coqanswer}
-
-% It looks like every request to compute (\texttt{large\_set\_check omega $i$ $2\times i$})  will succeed. Let us try an example.
-
-% \begin{Coqsrc}
-% Compute large_set_check omega 49 98.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = Ok
-%      : answer
-% \end{Coqanswer}
-
-% \subsubsection{\texorpdfstring{$\omega^2$}{omega\^{}2}-large intervals}
-
-% Still using \texttt{Compute} and \texttt{large\_set\_check}, we obtained the 
-% following values of $j$, the least integer such that the interval $[i,j]$ is $\omega^2$ large.
-
-% $$
-% \begin{array}{|c|c|}
-% \hline
-%   i & j \\
-% \hline 
-% 1 & 4 \\
-% 2 & 14 \\
-% 3 & 38 \\
-% 4 & 94 \\
-% 5 & 222 \\
-% % 6 & 510 
-% \hline
-% \end{array}
-% $$
-
-% \begin{exercise}
-% Please give the 6-th and 7-th line of the array above.
-% \end{exercise}
-
-% \subsubsection{The limits of (pure) computation}
-
-% Let us now play with bigger ordinals, for instance $\alpha=\omega^\omega$ or
-% $\alpha=\omega^{\omega + 1}$. We notice that, even for small values of $i$, it is hard
-% to guess values of $j$ such that $[i,i+j]$ is $\alpha$-large.
-
-% \begin{Coqsrc}
-% Compute large_set_check (omega ^ omega) 1 4.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%  = Ok
-%      : answer
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check (omega ^ omega) 2 38.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%  = Ok
-%      : answer
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check (omega ^ omega) 3 1000.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = Remaining (omega ^ 2 * 2 + omega * 220 + 798)%pT1
-%      : answer
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check (omega ^ omega) 3 1798.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = Remaining (omega ^ 2 * 2 + omega * 220)%pT1
-%      : answer
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check (omega ^ omega) 3 5000.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = Remaining (omega ^ 2 * 2 + omega * 218 + 2198)%pT1
-%      : answer
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check (omega ^ (omega + 1)) 3 5000.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%  = Remaining
-%          (omega ^ omega * 2 + omega ^ 3 * 4 + 
-%           omega ^ 2 * 4 + omega * 1148 +  4222)%pT1
-%      : answer  
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute large_set_check (omega ^ (omega + 1)) 3 10000.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% Warning: To avoid stack overflow, large numbers in nat are interpreted as
-% applications of Init.Nat.of_uint. [abstract-large-number,numbers]
-
-%     = Remaining
-%          (omega ^ omega * 2 + omega ^ 3 * 4 + 
-%           omega ^ 2 * 4 + omega * 1147 +
-%           8446)%pT1
-%      : answer
-% \end{Coqanswer}
-
-% Since computation is not enough, let us comme back to proofs, or, better, proofs \emph{and} computations.
-
-% \subsection{Proving largeness}
-
-% %%% ICI %%%
-
-% \subsection{$n$-large and $\omega$-large intervals }
-
-% A finite ordinal is just made out of \texttt{zero} and \texttt{succ} constructions.
-% Thus any sequence of strictly positive integers of length greater than  or equal to $n$ will completely gnaw the ordinal $n$.
-
-% Concerning the first limit ordinal $\omega$ the largeness of a sequence $s$ depends only on its first element and the length of the rest of the sequence (please keep in mind that the argument $i$ of $\canonseq{\alpha}{i}$ is meaningful only if $\alpha$ is a limit ordinal).
-% The following proposition is labeled $4.2$ in~\cite{KS81}.
-
-% \begin{proposition}
-%   For $n<\omega$, a set $X$id $n$-large if and only if $|X|\geq n$.
-% A finite set $X$ is $\omega$-large if $|X|>\min{X}$.
-% \end{proposition}
-
-% Rewritten in terms of list of strictly positive integers, we get the following statements:
-
-% \begin{Coqsrc}
-% Lemma large_n_iff : forall s (n:nat) , ~ In 0 s  ->
-%                                   large n s  <-> (n <= List.length s)%nat.
-% \end{Coqsrc}
-
-% \begin{Coqsrc}
-% Lemma large_omega_iff : forall s n,  ~In 0 (n::s) -> 
-%                                             large omega (n::s) <->
-%                                             (n <=  List.length s)%nat.
-% \end{Coqsrc}
-
-% % \begin{Coqsrc}
-% % (* sorted list of natural numbers greater than or equal to n *)
-
-% % Inductive sorted_ge (n: nat) : list nat -> Prop :=
-% % | sorted_ge_nil : sorted_ge n nil
-% % | sorted_ge_one : forall p, n<=p ->
-% %                             sorted_ge n (p::nil)
-% % | sorted_ge_cons: forall p q s,  n<=p -> p<q ->
-% %                                  sorted_ge p (q::s) ->
-% %                                  sorted_ge n (p::q::s).
-% % \end{Coqsrc}
-
-% \subsection{Mimimal large sequences}
-
-% Let us consider \emph{minimal} large sequences, \emph{i.e.} large sequences 
-% the strict prefix of which do not lead to $0$. In other words, only the last ordinal 
-% defined by the sequence is null.  For this purpose, we use the predicate \texttt{path\_to} introduced page~\pageref{path-to-definition}.
-
-% \begin{Coqsrc}
-% Definition mlarge alpha s := path_to zero s alpha.
-% \end{Coqsrc}
-
-%------------------------------------------------------
-\section{A certified catalogue of rapidly growing functions}
-
-In this section, we try to present an abstract of the properties of the main variants of fast-growing hierarchies of functions
-we found in the literature or we had to define as helpers in our proofs.
-
-\subsection{Ketonen-Solovay's \texttt{F\_alpha}}
-
-  This hierarchy is presented p.280 of~\cite{KS81} as `` a trivial variant of one introduced by Wainer ''.
-  Let us recall the equations shown in Sect.~\vref{F_equations}.
-
-\begin{itemize}
-\item $F_0(i)=i+1$
-\item $F_{\beta+1}(i)= (F_\beta)^{(i+1)}(i)$, where $f^{(i)}$ is the $i$-th iterate of $f$.
-\item $F_\alpha(i) = F_{\canonseq{\alpha}{i}} (i)$ if $\alpha$ is a limit ordinal.
-\end{itemize}
-
-Note that \cite{KS81} defines also $F_{\epsilon_0}$ (by the third equation).
-Since $\epsilon_0$ is not representable in type \texttt{E0}, our translation in \coq{} does not take $F_{\epsilon_0}$ into account.
-
-Several properties of $F_\alpha$ are already presented in Sect.~\ref{F_equations}.
 

### 2021-09-06 doc/part-hydras.tex
try to make Ordering_Functions a type class

@@
 
-\subsubsection{The definition of zero}
+\subsection{The definition of zero}
 
@@
 
-\input{movies/snippets/Schutte_basics/iotaDemo}
-
-Indeed, the basic logic of  \coq{} does not allow us to eliminate a proof of a proposition 
+\input{movies/snippets/Schutte_basics/iotaDemoa}
+\input{movies/snippets/Schutte_basics/iotaDemob}
+
+Indeed, the logic of  \coq{} does not allow us to eliminate a proof of a proposition 
 $\exists!\,x:A,\,P(x)$ for building a term whose type lies in the sort \texttt{Type}. 
@@
 
-\begin{Coqsrc}
-Axiom epsilon_statement:
-  forall (A : Type) (P : A->Prop), inhabited A ->
-    {x : A | (exists x, P x) -> P x}.
-\end{Coqsrc}
-
-Hilbert's $\epsilon$ \emph{operator} is derived from this  axiom.
-
-\begin{Coqsrc}
-Definition epsilon (A : Type) (i:inhabited A) (P : A->Prop) : A
-  := proj1_sig (epsilon_statement P i).
-
-Lemma constructive_indefinite_description :
-  forall (A : Type) (P : A->Prop),
-    (exists x, P x) -> { x : A | P x }.
-\end{Coqsrc}
+\inputsnippets{MoreEpsilonIota/EpsilonStatement}
+
+\emph{Hilbert's $\epsilon$ operator} is derived from this  axiom.
+
+\inputsnippets{MoreEpsilonIota/EpsilonDef}
 
@@
 
-\begin{Coqsrc}
-Definition iota (A : Type) (i:inhabited A) (P : A->Prop) : A
-  := proj1_sig (iota_statement P i).
-\end{Coqsrc}
-
-\begin{Coqsrc}
- Lemma constructive_definite_description :
-  forall (A : Type) (P : A->Prop),
-    (exists! x, P x) -> { x : A | P x }.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Definition iota_spec (A : Type) (i:inhabited A) (P : A->Prop) :
-  (exists! x:A, P x) -> P (iota i P)
-  := proj2_sig (iota_statement P i).
-\end{Coqsrc}
-
-Indeed, the operators \texttt{epsilon} and \texttt{iota} allowed us to make our definitions 
-quite close to Schütte's text. Our libraries \href{../theories/html/hydras.Schutte.MoreEpsilonIota.html}%
+\inputsnippets{MoreEpsilonIota/IotaDef}
+
+Indeed, the operators \texttt{epsilon} and \texttt{iota} allow us to write formal definitions close to Schütte's text. Our libraries \href{../theories/html/hydras.Schutte.MoreEpsilonIota.html}%
 {\texttt{Schutte.MoreEpsilonIota}}
@@
 
-\input{movies/snippets/Well_Orders/leastMemberDef}
 \input{movies/snippets/Well_Orders/theLeast}
@@
 Since we won't be able to prove the proposition
-\linebreak \Verb|{exists! a: Ord, least_member (Empty_set Ord) a|, the only properties we would be able to prove about \texttt{bottom} would be \emph{trivial} properties, 
+\linebreak \Verb|exists! a: Ord, least_member (Empty_set Ord) a|, the only properties we would be able to prove about \texttt{bottom} would be \emph{trivial} properties, 
 \emph{i.e.}, satisfied by \emph{any} element of type \texttt{Ord}, like for instance
@@
 
-Using \texttt{succ}, it is now easy to define recursively all the finite ordinals.
+Using \texttt{succ}, it is now easy to define recursively the set of the  finite ordinals.
 
@@
 
-\input{movies/snippets/Schutte_basics/omegaProps}
+\input{movies/snippets/Schutte_basics/omegaPropsa}
+\input{movies/snippets/Schutte_basics/omegaPropsb}
+\input{movies/snippets/Schutte_basics/omegaPropsc}
+\input{movies/snippets/Schutte_basics/omegaPropsd}
 
@@
 \index{hydras}{Library Schutte!Predicates!ordering function@ordering\_function}
+
+\begin{todo}
+Define \texttt{ordering\_function} as a type class !
+\end{todo}
 

### 2021-09-06 doc/chapter-primrec.tex
corrections chapters 8,9

@@
 
-\begin{Coqanswer}
-3 subgoals (ID 184)
-
-  ============================
-  isPR 1 (fun _ : nat => 2)
-
-subgoal 2 (ID 185) is:
- isPR 1 (fun x : nat => x)
-subgoal 3 (ID 186) is:
- isPR 2 Init.Nat.mul
-\end{Coqanswer}
-
 \input{movies/snippets/PrimRecExamples/doubleIsPRb}
@@
 
-\input{movies/snippets/isqrt/boundedSearch3}
-\input{movies/snippets/isqrt/boundedSearch4}
+\inputsnippets{isqrt/boundedSearch3, isqrt/boundedSearch4}
 
@@
 
-\begin{Coqanswer}
- x : PrimRec m
- x0 : nat
- H : forall v : t nat n,
-   max_v (map (fun f : naryFunc n => v_apply f v)
-           (evalPrimRecs n m g)) <=
-   Ack x0 (max_v v)
- x1 : nat
- H0 : forall v : t nat m, v_apply (evalPrimRec m x) v <=
-                                  Ack x1 (max_v v)
-  ============================
-  forall v : t nat n,
-  v_apply (evalPrimRec n (composeFunc n m g x)) v <=
-  Ack (2 + Nat.max x0 x1) (max_v v)
-
-\end{Coqanswer}
-
 \input{movies/snippets/AckNotPR/majorAnyPRb}
 
-\begin{Coqanswer}
-  n : nat
-  x1 : PrimRec n
-  x2 : PrimRec (S (S n))
-  r : nat
-  Hg : forall v : t nat n,
-     v_apply (evalPrimRec n x1) v <= Ack r (max_v v)
-  s : nat
-  Hh : forall v : t nat (S (S n)),
-       v_apply (evalPrimRec (S (S n)) x2) v <= Ack s (max_v v)
-  ============================
-  majorizedPR (primRecFunc n x1 x2) Ack
-\end{Coqanswer}
+The last two goals deal with vectors of functions.
 
-The las two goals deal with vectors of functions.
-
-\begin{Coqanswer}
-1 subgoal (ID 289)
-
-  n : nat
-  ============================
-  majorizedSPR (PRnil n) Ack
-
-\end{Coqanswer}
-
-\begin{Coqanswer}
-1 subgoal (ID 296)
-
-  n, m : nat
-  x : PrimRec n
-  p : PrimRecs n m
-  IHx : majorizedPR x Ack
-  IHx0 : majorizedSPR p Ack
-  ============================
-  majorizedSPR (PRcons n m x p) Ack
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Qed.
-\end{Coqsrc}
+\input{movies/snippets/AckNotPR/majorAnyPRVec}
 

### 2021-09-06 doc/part-hydras.tex
corrections chapter 7
Ordering_Functions no a class type (draft)
corrections chapters 8,9

--- 2021-09-04 doc/part-hydras.tex
@@
 
-\input{movies/snippets/Schutte_basics/iotaDemo}
+\input{movies/snippets/Schutte_basics/iotaDemoa}
+\input{movies/snippets/Schutte_basics/iotaDemob}
 
@@
 
-\begin{Coqsrc}
-Axiom epsilon_statement:
-  forall (A : Type) (P : A->Prop), inhabited A ->
-    {x : A | (exists x, P x) -> P x}.
-\end{Coqsrc}
+\input{movies/snippets/MoreEpsilonIota/EpsilonStatement}
 
@@
 
-\begin{Coqsrc}
-Definition epsilon (A : Type) (i:inhabited A) (P : A->Prop) : A
-  := proj1_sig (epsilon_statement P i).
-
-Lemma constructive_indefinite_description :
-  forall (A : Type) (P : A->Prop),
-    (exists x, P x) -> { x : A | P x }.
-\end{Coqsrc}
+\input{movies/snippets/MoreEpsilonIota/EpsilonDef}
 
@@
 
-\begin{Coqsrc}
-Definition iota (A : Type) (i:inhabited A) (P : A->Prop) : A
-  := proj1_sig (iota_statement P i).
-\end{Coqsrc}
-
-\begin{Coqsrc}
- Lemma constructive_definite_description :
-  forall (A : Type) (P : A->Prop),
-    (exists! x, P x) -> { x : A | P x }.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Definition iota_spec (A : Type) (i:inhabited A) (P : A->Prop) :
-  (exists! x:A, P x) -> P (iota i P)
-  := proj2_sig (iota_statement P i).
-\end{Coqsrc}
+\input{movies/snippets/MoreEpsilonIota/iotaDef}
 
@@
 
-\begin{Coqsrc}
-Lemma plus_elim (alpha : Ord) :
-  forall P : (Ord->Ord)->Prop,
-    (forall f: Ord->Ord, 
-        ordering_function f ordinal (ge alpha)-> P f) ->
-    P (plus alpha).
-\end{Coqsrc}
+\input{movies/snippets/Addition/plusElim}
 
@@
 
-\input{movies/snippets/Critical/CrZeroAP}
-\input{movies/snippets/Critical/epsilon0Cr1}
+\inputsnippets{Critical/CrZeroAP,
+  Critical/epsilon0Cr1}
 
@@
 
-\input{movies/snippets/CNF/cnfUnicity}
-
-\input{movies/snippets/CNF/cnfExUnique}
+\inputsnippets{CNF/cnfUnicity,CNF/cnfExUnique}
 
@@
 
-\section{How big is \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}?}
+\section{A strict order on T2}
 
@@
 \input{movies/snippets/Gamma0/ONGamma0}
+\input{movies/snippets/Gamma0/G0z}
 

### 2021-09-07 doc/chapter-powers.tex
re-reading chapter 2 (partially)
first try
modifications of doc/movies/Makefile (includes additions and gaia)
Fir snippets in theories/additions

@@
 
-\begin{Coqsrc}
-Section Definitions.
-
-Variables (A: Type)
-           (mult: A -> A -> A)
-           (one: A).
-Local Infix "*" := mult.
-Local Notation "1" := one.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Fixpoint power (x:A)(n:nat) : A :=
-  match n with 
-   | 0%nat => 1
-   | S p =>   x * x ^ p
-  end
-where "x ^ n" := (power x n).
-
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Compute power Z.mul 1%Z 2%Z 10.  
-\end{Coqsrc}
-\begin{Coqanswer}
-   = 1024%Z
-     : Z
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Open Scope string_scope.
-Compute power append  "" "ab"  12.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = "abababababababababababab"
-     : string
-\end{Coqanswer}
+\inputsnippets{FirstSteps/Defs}
 
@@
 
-\begin{Coqsrc}
-Compute N_bpow Z.mul 1%Z 2%Z 10. 
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = 1024%Z
-     : Z 
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Require Import String.
-Open Scope string_scope.
-
-Compute N_bpow append "" "ab"  12.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = "abababababababababababab"
-     : string 
-\end{Coqanswer}
+\inputsnippets{FirstSteps/PowerCompute}
 
@@
 
-\begin{Coqsrc}
-Module M2.
-Section Definitions.
-
-  Variables (A: Type) (zero one : A)  (plus mult  : A -> A -> A).
-
-  Variable rt : semi_ring_theory  zero one plus mult   (@eq A).
-  Add Ring Aring : rt.
-
-  Notation "0" := zero.  
-  Notation "1" := one.
-  Notation "x + y" := (plus x y).  
-  Notation "x * y " := (mult x y).
-
-  Structure t : Type := mat{c00 : A;  c01 : A;  c10 : A;  c11 : A}.
-\end{Coqsrc}
+\inputsnippets{FirstSteps/M2a}
 
@@
 
-\begin{Coqsrc}
-Definition M2_mult (M M':t) : t := mat
-  (c00 M * c00 M' + c01 M * c10 M') (c00 M * c01 M' + c01 M * c11 M')
-  (c10 M * c00 M' + c11 M * c10 M') (c10 M * c01 M' + c11 M * c11 M').
-\end{Coqsrc}
-
-The neutral element for \texttt{M2\_mult} is the identity matrix.
-
-\begin{Coqsrc}
-Definition Id2 : t := mat 1 0 0 1.
-
-End M2_Definitions.
-End M2.
-\end{Coqsrc}
+\inputsnippets{FirstSteps/M2Mult}
 
@@
 
-\begin{Coqsrc}
-Fixpoint fib (n:nat) : N :=
-  match n with
-    0%nat | 1%nat => 1%N
-  | S (S p as q) => fib p + fib q
-  end.
-
-Compute fib 20.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-= 10946 : N
-\end{Coqanswer}
+\inputsnippets{Fib2/FibDef}
 
@@
 \begin{exercise}
+  \label{exercise:fibmat}
   \begin{enumerate}
@@
 
-\begin{Coqsrc}
-Definition mul2 (p q : N * N) :=
-  match p, q with (a, b),(c,d) => (a*c + a*d + b*c, a*c + b*d) end.
-
-Instance Mul2 : Monoid  mul2 (0,1).
-(* Proof omitted *)
-\end{Coqsrc}
-
-The following lemma is a simplification of the equality~\vref{fibmat-eq1}.
-
-\begin{Coqsrc}
-Lemma next_fib (n:nat) : mul2 (1,0) (fib (S n), fib n) =
-                         (fib (S (S n)), fib (S n)).  
-\end{Coqsrc}
-
-By induction\footnote{Simple induction, since $n$ has type \texttt{nat}.}
- over $n$, we obtain a the following equalities.
-
-\begin{Coqsrc}
-Definition fib_mul2 n := let (a,b) := power (M:=Mul2) (1,0) n   in (a+b).
-
-Lemma fib_mul2_OK_0 (n:nat) : power (M:=Mul2) (1,0) (S (S n)) =
-                              (fib (S n), fib n).
-
-Lemma fib_mul2_OK n : fib n = fib_mul2 n.
-\end{Coqsrc}
+\inputsnippets{Fib2/mul2Def}
+\inputsnippets{Fib2/mul2Monoid}
+
+The following lemma is a simplification of the equality of Exercise~\ref{exercise:fibmat}.
+
+\inputsnippets{Fib2/nextFib}
+
+Let us consider a new definition of the Fibonacci function.
+
+\inputsnippets{Fib2/fibMul2Def}
+\inputsnippets{Fib2/fibMul2OK0}
+\inputsnippets{Fib2/fibMul2OK}
+\inputsnippets{Fib2/TimeFibMul2}
 
@@
 
-\begin{Coqsrc}
-Definition power_t := forall (A:Type)
-                             (mult : A -> A -> A)
-                             (one:A)
-                             (x:A)
-                             (n:N), A.
-\end{Coqsrc}
+\inputsnippets{FirstSteps/powerTDef}
 
@@
 
-\begin{Coqbad}
-Module Bad.
-
- Definition correct_expt_function (f : power_t) : Prop :=
-  forall A (mult : A -> A -> A) (one:A)
-            (x:A) (n:N), 
-            power mult one x (N.to_nat n) = f A mult one x n.
-\end{Coqbad}
+\inputsnippets{FirstSteps/Bada}
 
@@
 
-\begin{Coqbad}
-Section CounterExample.
-    Let mul (n p : nat) := n + 2 * p.
-    Let one := 0.
-
-    Remark mul_not_associative :
-      exists  n p q,  mul n (mul p q) <> mul (mul n p) q.
-    Proof. 
-        exists 1, 1, 1; discriminate. 
-    Qed.
-
-    Remark one_not_neutral  :
-      exists n : nat, mul one n <> n.
-    Proof.
-      exists 1; discriminate.
-    Qed.
-
-    Lemma correct_expt_too_strong : 
-          ~ correct_expt_function (@N_bpow).
-    Proof.
-      intro H; specialize (H _ mul one 1  7%N).
-      discriminate H.    
-    Qed.
-
-End CounterExample.
-End Bad.
- \end{Coqbad}
+\inputsnippets{FirstSteps/Badb}
 
@@
 
-\begin{Coqsrc}
-Class Mult_op (A:Type) := mult_op : A -> A -> A.  
-\end{Coqsrc}
+\inputsnippets{Monoid_def/MultOpClass}
 
@@
 
+\inputsnippets{Monoid_def/MultOpEq}
+
 We are now ready to define a new notation scope, in which the notation
@@
 
-\begin{Coqsrc}
-Delimit Scope M_scope with M.
-Infix "*" := mult_op : M_scope.
-Open Scope M_scope.  
-\end{Coqsrc}
+\inputsnippets{Monoid_def/MultOpInfix}
 
@@
 
-\begin{Coqsrc}
-Instance nat_mult_op : Mult_op nat  := Nat.mul.
-\end{Coqsrc}
+\inputsnippets{Monoid_def/DemoNatMulta}
 
@@
 
-\begin{Coqsrc}
-Set Printing All.
-Check  3 * 4. 
-\end{Coqsrc}
-\begin{Coqanswer}
-@mult_op nat nat_mult_op (S (S (S O))) (S (S (S (S O))))
-     : nat  
-\end{Coqanswer}
-\begin{Coqsrc}
-Unset Printing All.
-Compute 3 * 4.   
-\end{Coqsrc}
-\begin{Coqanswer}
- = 12  : nat  
-\end{Coqanswer}
+\inputsnippets{Monoid_def/DemoNatMultb}
 
@@
 
-\begin{Coqsrc}
-Require Import String.
-
-Instance string_op : Mult_op string := append.
-Open Scope string_scope.
-
-Example ex_string : "ab" * "cde" = "abcde".
-Proof. reflexivity. Qed.
-\end{Coqsrc}
+\inputsnippets{Monoid_def/DemoStringMult}
 
@@
 
-\begin{Coqsrc}
-Class Monoid {A:Type}(op : Mult_op A)(one : A) : Prop :=
-{
-    op_assoc : forall x y z:A, x * (y * z) = x * y * z;
-    one_left : forall x, one * x = x;
-    one_right : forall x, x * one = x
-}.
-\end{Coqsrc}
+\inputsnippets{Monoid_def/MonoidClass}
 
@@
 
-\begin{Coqsrc}
-Instance Z_mult_op : Mult_op Z := Z.mul.
-
-Instance ZMult : Monoid  Z_mult_op 1.
-Proof. 
-  split.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-3 subgoals, subgoal 1 (ID 8)
-
-  ============================
-   forall x y z : Z, (x * (y * z))%M = (x * y * z)%M
-
-subgoal 2 (ID 9) is:
- forall x : Z, (1 * x)%M = x
-subgoal 3 (ID 10) is:
- forall x : Z, (x * 1)%M = x}
-\end{Coqanswer}
-
-\begin{Coqsrc}
-    all: unfold Z_mult_op, mult_op;intros;ring.
-Qed.
-\end{Coqsrc}
+\inputsnippets{Monoid_instances/ZMultDef}
 
@@
 
-The following function definition generates 9 sub-goals,
+The following function definition generates 9 subgoals,
 for proving that the measure on signatures is strictly decreasing along

### 2021-09-07 doc/part-hydras.tex
Fix compilation issues
minor changes

@@
 
+\subsection{A proof by computation}
+\label{Ex42-E0}
+
+It is interesting to compare the following proof of the equality
+$\omega+42+\omega^2$ with the more theoretical proof in Sect~\vref{Ex42-schutte}.
+
+\inputsnippets{E0/Ex42}
+
 \section{Well-foundedness and transfinite induction}
@@
 
-\input{movies/snippets/Well_Orders/leastMemberDef}
+\input{movies/snippets/Well_Orders/MDecl}
 \input{movies/snippets/Well_Orders/theLeast}
@@
 The set  \texttt{AP} is  \emph{closed} under addition, and unbounded.
+\label{lemma:AP-plus-closed}
 
@@
 \input{movies/snippets/AP/APPhi0}
+
+\subsection{A last example}
+\label{Ex42-schutte}
+
+Let us prove again the equality $\omega+42+\omega^2= \omega^2$.Let us recall that $\omega^2$ is an abbreviation of $\phi_0(2)$,
+\emph{i.e} the third  additive principal ordinal.
+
+\inputsnippets{Schutte/Ex42a}
+
+The proof is very different from the computational proof of Sect~\vref{Ex42-E0}.
+By definition of additive principal ordinals, 
+it suffices to prove the inequality $\omega+42< \phi_0(2)$ to get the equality we want to prove.
+
+\inputsnippets{Schutte/Ex42b}
+
+Since the set \texttt{AP} of additive principals  is closed under addition
+(by Lemma \texttt{AP\_plus\_closed}, page~\pageref{lemma:AP-plus-closed}) , it suffices to prove the inequalities $\omega<\phi_0(2)$ and $42<\phi_0(2)$.
+
+\input{movies/snippets/Schutte/Ex42d}
 
@@
 
-Lets us denote by $\varphi_\alpha$ the ordering function of the set $\textit{Cr}(\alpha)$ and by $A_\alpha$ its ordering segment.
+Lets us denote by $\phi_\alpha$ the ordering function of the set $\textit{Cr}(\alpha)$ and by $A_\alpha$ its ordering segment.
 

### 2021-09-08 doc/chapter-powers.tex
adapt chapter-powers to Alectryon (partial)

@@
 
-\begin{Coqsrc}
-Fixpoint binary_power_mult (x a:A)(p:positive) : A 
-  :=
-  match p with
-    | xH =>  a * x
-    | xO q => binary_power_mult  (x * x) a q
-    | xI q =>  binary_power_mult  (x * x) (a * x) q
-  end.
-
-Fixpoint Pos_bpow (x:A)(p:positive) :=
- match p with
-  | xH => x
-  | xO q => Pos_bpow  (x * x) q
-  | xI q => binary_power_mult   (x * x) x q
-end.
-
-Definition N_bpow x (n:N) := 
-  match n with 
-  | 0%N => 1
-  | Npos p => Pos_bpow x p
-  end.
-
-End Definitions.
-\end{Coqsrc}
-
+\inputsnippets{FirstSteps/bpowDef}
 Let us  close the section \texttt{Definitions} and mark the argument \texttt{A} as implicit.
-
-\begin{Coqsrc}
-End Definitions.
-
-Arguments N_bpow {A}.
-Arguments power {A}.
-\end{Coqsrc}
+\inputsnippets{FirstSteps/EndDefs}
 
@@
 
-\subsubsection{Exponentiation on $2\times 2$ matrices}
-\label{naive-matrix}
-Our second example is a definition of $M^n$ where $M$ is a $2\times 2$ matrix
-over any ``scalar''  type $A$, assuming one can provide $A$ with a semi-ring structure~\cite{Coq}.
-
-%\subsubsection{Representation of $2\times 2$ matrices}
-
-A $2\times 2$ matrix will be simply represented by a structure with four fields;
-each field \texttt{c$ij$} is associated with the $i$-th line and $j$-th column of the considered matrix.
-
-\inputsnippets{FirstSteps/M2a}
+% \subsubsection{Exponentiation on $2\times 2$ matrices}
+% \label{naive-matrix}
+% Our second example is a definition of $M^n$ where $M$ is a $2\times 2$ matrix
+% over any ``scalar''  type $A$, assuming one can provide $A$ with a semi-ring structure~\cite{Coq}.
+
+% %\subsubsection{Representation of $2\times 2$ matrices}
+
+% A $2\times 2$ matrix will be simply represented by a structure with four fields;
+% each field \texttt{c$ij$} is associated with the $i$-th line and $j$-th column of the considered matrix.
 
 %\subsubsection{Matrix Multiplication}
-
-The structure type \texttt{M2.t} allows us to define the product
-of two matrices.
-
-\inputsnippets{FirstSteps/M2Mult}
 
@@
 \item The ``natural'' monoid $(\mathbb{N},\times, 1)$ :
-
-  \begin{Coqsrc}
-Instance nat_mult_op : Mult_op nat | 5 := Nat.mul.
-
-Instance  Natmult : Monoid nat_mult_op  1%nat | 5
-Proof.
-   split;unfold nat_mult_op, mult_op; intros; ring.
-Qed.
-\end{Coqsrc}
+\inputsnippets{Monoid_instances/natMult}
 
@@
 
-\begin{Coqsrc}
-Instance nat_plus_op : Mult_op nat | 12 := Nat.add.
-
-Instance Natplus : Monoid nat_plus_op  0%nat | 12.
-(* Proof omitted *)
-\end{Coqsrc}
+\inputsnippets{Monoid_instances/natPlus}
 \end{itemize}
@@
 
-\begin{Coqsrc}
-Instance int31_mult_op : Mult_op int31 := mul31.
-
-Instance  Int31mult : Monoid int31_mult_op  1.
-Proof.
-   split;unfold int31_mult_op, mult_op; intros; ring.
-Qed.
-\end{Coqsrc}
+\inputsnippets{Monoid_instances/int31}
 
@@
 
-\begin{Coqbad}
-Module Bad.
-
-Fixpoint int31_from_nat (n:nat) :=
-  match n with
-  | O => 1
-  | S p => 1 + int31_from_nat p
-  end.
-
-Coercion int31_from_nat : nat >-> int31.
-
-Fixpoint fact (n:nat) := 
-  match n with
-   | O => 1
-   | S p => n * fact p
-  end.
-
-Example fact_zero : exists n:nat, fact n = 0.
-Proof.  now exists 40%nat.  Qed.
-
-End Bad.
-\end{Coqbad}
+\inputsnippets{Monoid_instances/BadFact}
 
@@
 
-In Sect.\vref{naive-matrix}, we defined a function for computing powers of any $2\times 2$ 
-matrix over any semi-ring. For proving a simple property of matrix exponentiation, we had
- to prove that matrix multiplication is associative and admits the identity matrix as a neutral element.
-These properties are easily expressed within the type class framework, by defining a \emph{family} of monoids.
+Let $(A,+,\times)$ be a semi-ring. We define a multiplicative monoid on the set of \emph{e.g.} $2\times 2$-) square matrices over $A$.
 It suffices to define an instance of \texttt{Monoid} within the scope of a hypothesis
@@
 
-\begin{Coqsrc}
-Section M2_def.
-Variables (A:Type)
-           (zero one : A) 
-           (plus mult  : A -> A -> A).
-
- Variable rt : semi_ring_theory  zero one plus mult  (@eq A).
- Add  Ring Aring : rt.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Structure M2 : Type := {c00 : A;  c01 : A;
-                        c10 : A;  c11 : A}.
-
-Definition Id2 : M2 := Build_M2 1 0 0 1.
-
-Definition M2_mult (m m':M2) : M2 :=
- Build_M2 
-          (c00 m * c00 m' + c01 m * c10 m')
-          (c00 m * c01 m' + c01 m * c11 m')
-          (c10 m * c00 m' + c11 m * c10 m')
-          (c10 m * c01 m' + c11 m * c11 m').
-
-Global Instance M2_op : Mult_op M2 := M2_mult.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Global Instance M2_Monoid : Monoid   M2_op Id2.
-(* Proof omitted *)
-
-End M2_def.
-
-Arguments M2_Monoid {A zero one plus mult} rt.
-\end{Coqsrc}
+\inputsnippets{Monoid_instances/M2Defsa, Monoid_instances/M2Defsb}
 
@@
 
-\begin{Coqsrc}
-Class Equiv A := equiv : relation A.
-
-Infix "==" := equiv (at level 70) : type_scope.
-\end{Coqsrc}
+\inputsnippets{Monoid_def/EquivDef}
 
@@
 
-\begin{Coqsrc}
-Class EMonoid (A:Type)(E_op : Mult_op A)(E_one : A) 
-      (E_eq: Equiv A): Prop :=
-  {
-    Eq_equiv :> Equivalence equiv;
-    Eop_proper :> Proper (equiv ==> equiv ==> equiv) E_op;
-    Eop_assoc : forall x y z, x * (y * z) == x * y * z;
-    Eone_left : forall x,  E_one * x == x;
-    Eone_right : forall x,  x * E_one ==  x
-  }.
-\end{Coqsrc}
+\inputsnippets{Monoid_def/EMonoidDef}
 
@@
 
-\begin{Coqsrc}
-Global Instance eq_equiv {A} : Equiv A := eq.
-
-Global Instance Monoid_EMonoid `(M:@Monoid A op one) :
-        EMonoid  op one eq_equiv.
-Proof.
-split; unfold eq_equiv, equiv in *.
- - apply eq_equivalence.
- - intros x y H z t H0; now subst.
- - intros; now rewrite (op_assoc).
- - intro; now rewrite one_left.
- - intro; now rewrite one_right.
-Defined.
-\end{Coqsrc}
+\inputsnippets{Monoid_def/Coerciona}
 
@@
 
-\begin{Coqsrc}
-Coercion Monoid_EMonoid : Monoid >-> EMonoid.
-\end{Coqsrc}
-
-%ici%
+\inputsnippets{Monoid_def/Coercionb}
 
@@
 
-\begin{Coqsrc}
-Check NMult : EMonoid  N.mul 1%N eq.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  NMult:EMonoid N.mul 1%N eq
-     : EMonoid N.mul 1%N eq
-\end{Coqanswer}
+\inputsnippets{Monoid_instances/CheckCoercion}
 

### 2021-09-08 doc/chapter-powers.tex
adapt chapter-powers to Alectryon (partial)

@@
 
-\begin{Coqsrc}
-Section Nmodulo.
-  Variable m : N.
-  Hypothesis m_gt_1 : 1 < m.
-
-  Definition mult_mod ( x y : N) := (x * y) mod m.
-  Definition mod_eq ( x y: N) := x mod m = y mod m.
-
-  Global Instance mod_equiv : Equiv N := mod_eq.
-
-  Global Instance mod_op : Mult_op N := mult_mod.
-
-  Global Instance mod_Equiv : Equivalence mod_equiv.
-  (* Proof omitted *)
-
-  Global Instance mult_mod_proper : 
-  Proper (mod_equiv ==> mod_equiv ==> mod_equiv)  mod_op.
-  (* Proof omitted *)
-
-  Local Open Scope M_Scope.
-
-  Lemma mult_mod_associative :  
-  forall x y z,  x * (y * z) = x * y * z.
-  (* Proof omitted *)
-
-  Lemma one_mod_neutral_l  : forall x, 1 * x ==  x.
-  (* Proof omitted *)
-
-  Lemma one_mod_neutral_r  : forall x, x * 1 == x.
-  (* Proof omitted *)
-
-  Global Instance Nmod_Monoid : EMonoid  mod_op 1 mod_equiv.
-  (* Proof omitted *)
-
-End Nmodulo.
-
-\end{Coqsrc}
-
-\paragraph{Example}
-In the following interaction, we show how to instantiate the parameter \texttt{$m$} to a 
-concrete value, for instance \texttt{$256$}.
-
-\begin{Coqsrc}
-Section S256.
-Let mod256 :=  mod_op 256.
-Local Existing Instance mod256 | 1.
-
-Compute (211 * 67)
-\end{Coqsrc}
-
-\begin{Coqanswer}
-= 57 : N  
-\end{Coqanswer}
-
-\begin{Coqsrc}
-End S256.
-\end{Coqsrc}
+\inputsnippets{Monoid_instances/Nmoduloa,
+  Monoid_instances/Nmodulob, Monoid_instances/Nmoduloc,
+  Monoid_instances/Nmodulod,
+  Monoid_instances/Nmoduloe,
+  Monoid_instances/Nmodulof,
+  Monoid_instances/Nmodulog}
 
@@
 
-\begin{Coqsrc}
-Compute (211 * 67)%M.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-= 14137 : N   
-\end{Coqanswer}
+\inputsnippets{Monoid_instances/Nmoduloh}
 

### 2021-09-08 doc/hydras.tex
adapt chapter-powers to Alectryon (partial)

@@
 \author{Pierre Castéran\\ LaBRI, Univ.
-Bordeaux, CNRS UMR 5800, \thanks{With contributions by Yves Bertot, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Clément Pit-Claudel, and Théo Zimmermann.
+Bordeaux, CNRS UMR 5800, \thanks{With contributions by Yves Bertot, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Karl Palmskog, Clément Pit-Claudel, and Théo Zimmermann.
 The formalization of primitive recursive functions was originally authored by

### 2021-09-08 doc/JFLA2022/paper.tex
branch for JFLA paper

% easychair.tex,v 3.5 2017/03/15

\documentclass{easychair}
%\documentclass[EPiC]{easychair}
%\documentclass[EPiCempty]{easychair}
%\documentclass[debug]{easychair}
%\documentclass[verbose]{easychair}
%\documentclass[notimes]{easychair}
%\documentclass[withtimes]{easychair}
%\documentclass[a4paper]{easychair}
%\documentclass[letterpaper]{easychair}

\usepackage{doc}

%%%% For Alectryon

\usepackage{texments}
%%% for movies by alectryon
\usepackage{../movies/snippets/assets/alectryon}
\usepackage{../movies/snippets/assets/pygments}
%%% One hypothesis per line 
\makeatletter
\renewcommand{\alectryon@hyps@sep}{\alectryon@nl}
\makeatother

%%% \snippets{A,B,C,…} inputs a series of snippets as one block (with \itemsep
%%% between them).  A, B, C should be paths to files in movies/snippets/.
\usepackage{etoolbox}
\makeatletter
\newcommand{\inputsnippets}[1]
  {{\setlength{\itemsep}{1pt}\setlength{\parsep}{0pt}% Adjust spacing
    \alectryon@copymacros\begin{io}
      \forcsvlist{\item\@inputsnippet}{#1}
    \end{io}}}
\let\input@old\input % Save definition of \input
\newcommand{\@inputsnippet}[1]
  {{\renewenvironment{alectryon}{}{}% Skip \begin{alectryon} included in snippet
    \input@old{../movies/snippets/#1}}}
\makeatother

% use this if you have a long article and want to create an index
% \usepackage{makeidx}

% In order to save space or manage large tables or figures in a
% landcape-like text, you can use the rotating and pdflscape
% packages. Uncomment the desired from the below.
%
% \usepackage{rotating}
% \usepackage{pdflscape}

% Some of our commands for this guide.
%
\newcommand{\easychair}{\textsf{easychair}}
\newcommand{\miktex}{MiK{\TeX}}
\newcommand{\texniccenter}{{\TeX}nicCenter}
\newcommand{\makefile}{\texttt{Makefile}}
\newcommand{\latexeditor}{LEd}

%\makeindex

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Hydras and Cie \\
  Variations on Kirby \& Paris' hydra battles and other entertaining math in Coq}

\author{
Pierre Castéran \inst{1}
\and
    Jérémy Damour \inst{2}
\and
Karl Palmskog \inst{3}
\and Clément Pit-Claudel \inst{4}
\and Théo Zimmermann \inst{5}
}

\institute{
 LaBRI, Université Bordeaux \\
  \email{pierre.casteran@labri.fr}
\and
   University of Miami,
   Miami, Florida, U.S.A.\\
   \email{geoff@cs.miami.edu}\\
\and
   University of Manchester,
   Manchester, U.K.\\
   \email{andrei@voronkov.com}\\
\and
   Chalmers University of Technology,
   Gothenburg, Sweden
\and
   EasyChair
 }

\authorrunning{Castéran, Damour, Palmskog, Pit-Claudel and Zimmermann}

% \titlerunning{} has to be set to either the main title or its shorter
% version for the running heads. When processed by
% EasyChair, this command is mandatory: a document without \titlerunning
% will be rejected by EasyChair
\titlerunning{Hydras and Cie}

\begin{document}

\maketitle

\begin{abstract}
We present the current state of our library \dots
\end{abstract}

\setcounter{tocdepth}{2}
{\small
\tableofcontents}

%\section{To mention}
%
%Processing in EasyChair - number of pages.
%
%Examples of how EasyChair processes papers. Caveats (replacement of EC
%class, errors).

%------------------------------------------------------------------------------
\section{Introduction}
\label{sect:introduction}

\section{Documentation with Alectryon}

We present two ~examples of the documentation generated with Alectryon: A simple proof by computation (Fig.~\ref{fig:Ex42E0}), and a longer proof where only the main steps are displayed, interleaved with plain text.

  \begin{figure}[h]
    \centering
    \fbox{
      \begin{minipage}[h]{1.0\linewidth}
        \inputsnippets{E0/Ex42}
      \end{minipage}}
    \caption{A simple proof by computation}
    \label{fig:Ex42E0}
  \end{figure}
\label{Ex42-E0}.

\begin{figure}[h]
  \centering

\label{Ex42-schutte}
\fbox{\begin{minipage}[h]{1.0\linewidth}
  Let us prove again the equality $\omega+42+\omega^2= \omega^2$.Let us recall that $\omega^2$ is an abbreviation of $\phi_0(2)$,
\emph{i.e} the third  additive principal ordinal.

\inputsnippets{Schutte/Ex42a}

Our proof is very different from the computational proof of Fig\ref{fig:Ex42E0}.
By definition of additive principal ordinals, 
it suffices to prove the inequality $\omega+42< \phi_0(2)$ to get the equality we want to prove.

\inputsnippets{Schutte/Ex42b}

Since the set \texttt{AP} of additive principals  is closed under addition
(by Lemma \texttt{AP\_plus\_closed}, page~\pageref{lemma:AP-plus-closed}) , it suffices to prove the inequalities $\omega<\phi_0(2)$ and $42<\phi_0(2)$.

\inputsnippets{Schutte/Ex42c, Schutte/Ex42d}

\end{minipage}
}
\caption{A proof interleaved with text (from the book)}
  \label{fig:Ex42b}
\end{figure}

\section{Acknowledgments}
\label{sect:acks}

\label{sect:bib}
\bibliographystyle{plain}
%\bibliographystyle{alpha}
%\bibliographystyle{unsrt}
%\bibliographystyle{abbrv}
\bibliography{easychair}

\end{document}

### 2021-09-09 doc/JFLA2022/paper.tex
draft paper

@@
 \and
-    Jérémy Damour \inst{2}
+    Jérémy Damour %\inst{2}
 \and
-Karl Palmskog \inst{3}
-\and Clément Pit-Claudel \inst{4}
-\and Théo Zimmermann \inst{5}
+Karl Palmskog %\inst{3}
+\and Clément Pit-Claudel %\inst{4}
+\and Théo Zimmermann %\inst{5}
 }
@@
 
-% \titlerunning{} has to be set to either the main title or its shorter
-% version for the running heads. When processed by
-% EasyChair, this command is mandatory: a document without \titlerunning
-% will be rejected by EasyChair
 \titlerunning{Hydras and Cie}
@@
 \begin{abstract}
-We present the current state of our library \dots
+  We present the evolution of our library \cite{HydraBattles}
+since \cite{JFLA2018paper}
 \end{abstract}
@@
 
-%\section{To mention}
-%
-%Processing in EasyChair - number of pages.
-%
-%Examples of how EasyChair processes papers. Caveats (replacement of EC
-%class, errors).
-
 %------------------------------------------------------------------------------
@@
 \label{sect:introduction}
+Hydra games (a.k.a. \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians:
+L. Kirby and J. Paris~\cite{KP82}: \emph{Accessible Independence Results for Peano Arithmetic}. 
+Although the mathematical contents of this 
+paper are quite advanced, the rules of hydra battles are very easy to understand.
+There are now several sites on the Internet where you can find tutorials on hydra games, together with simulators you can play with.
+See, for instance, the blogpost and source code written by Andrej Bauer~\cite{bauer2008,BauerHydra}.
+
+Hydra battles have already been presented in
+JFLA2018~\cite{JFLA2018paper}. We present the recent evolution of the library: new results, interaction within the Coq-community project, and documentation generated with Alectryon.
+
+\section{New contents}
+Length of usually considered battles.
+They correspond to the notions of canonical sequences and large sets in \cite{KS81}.
 
@@
 
-We present two ~examples of the documentation generated with Alectryon: A simple proof by computation (Fig.~\ref{fig:Ex42E0}), and a longer proof where only the main steps are displayed, interleaved with plain text.
+We present two ~examples of the documentation generated with Alectryon~\cite{alectryonpaper}: A simple proof by computation (Fig.~\ref{fig:Ex42E0}), and a longer proof where only the main steps are displayed, interleaved with plain text.
 
@@
 
-\begin{figure}[h]
+\begin{figure}[ph]
   \centering
@@
 Since the set \texttt{AP} of additive principals  is closed under addition
-(by Lemma \texttt{AP\_plus\_closed}, page~\pageref{lemma:AP-plus-closed}) , it suffices to prove the inequalities $\omega<\phi_0(2)$ and $42<\phi_0(2)$.
+(by Lemma \texttt{AP\_plus\_closed}, it suffices to prove the inequalities $\omega<\phi_0(2)$ and $42<\phi_0(2)$.
 
@@
 
+\section{Perspectives}
+The Gaia project~\cite{Gaia}, also maintained on \texttt{Coq-community} contains a development by José Grimm~\cite{grimm:hal-00911710}.
+This library -- dedicated  to the implementaion in \textit{Coq (SSreflect)} of books from  Bourbaki's Elements of Mathematics --- contains a generalization of two types of ordinals (\texttt{T1} for the ordinal $\epsilon_0$ and
+\texttt{T2} for $\Gamma_0$) to a larger ordinal defined by Ackermann, and relates these data structures with Bourbaki's set theory.
+It is interesting to note that both libraries \texttt{Gaia} and \texttt{hydra-battles} contain their own (isomorphic) definitions of
+\texttt{T1} and \texttt{T2}, originated in~\cite{CantorContrib}.
+The types --- defined in different modules --- cannot be considered as equal. Even more, some operations like addition, multiplication, etc. have now different definitions.
+
+It would be interesting to build a library which could be used as a gate giving access to both the combinatorial results of \texttt{hydra-battles} and the set-theoretic content of \texttt{Gaia}.
+For instance, the associativity of multiplication of ordinals below $\epsilon_0$, already proven in \texttt{Gaia} could be tranferred ``for free'' to the data structures of \texttt{hydra-battles}, thanks to the isomorphism between both implementations.
+
 \section{Acknowledgments}
@@
 %\bibliographystyle{abbrv}
-\bibliography{easychair}
+\bibliography{../thebib}
 

### 2021-09-09 doc/JFLA2022/paper.tex
draft paper

@@
 \newcommand{\latexeditor}{LEd}
+
+\newtheorem{todo}{To do}
 
@@
 
+\begin{todo}
+  Relation with Russel's library (maintained in Coq-community).
+\end{todo}
+
 \section{Documentation with Alectryon}
@@
 
-Since the set \texttt{AP} of additive principals  is closed under addition
-(by Lemma \texttt{AP\_plus\_closed}, it suffices to prove the inequalities $\omega<\phi_0(2)$ and $42<\phi_0(2)$.
+Since the set \textit{AP} of additive principals  is closed under addition
+(by Lemma \textit{AP\_plus\_closed}, it suffices to prove the inequalities $\omega<\phi_0(2)$ and $42<\phi_0(2)$.
 
@@
 \section{Perspectives}
-The Gaia project~\cite{Gaia}, also maintained on \texttt{Coq-community} contains a development by José Grimm~\cite{grimm:hal-00911710}.
+The Gaia project~\cite{Gaia}, also maintained on \textit{Coq-community} contains a development by José Grimm~\cite{grimm:hal-00911710}.
 This library -- dedicated  to the implementaion in \textit{Coq (SSreflect)} of books from  Bourbaki's Elements of Mathematics --- contains a generalization of two types of ordinals (\texttt{T1} for the ordinal $\epsilon_0$ and
 \texttt{T2} for $\Gamma_0$) to a larger ordinal defined by Ackermann, and relates these data structures with Bourbaki's set theory.
-It is interesting to note that both libraries \texttt{Gaia} and \texttt{hydra-battles} contain their own (isomorphic) definitions of
+It is interesting to note that both libraries \textit{Gaia} and \textit{hydra-battles} contain their own (isomorphic) definitions of
 \texttt{T1} and \texttt{T2}, originated in~\cite{CantorContrib}.
@@
 
-It would be interesting to build a library which could be used as a gate giving access to both the combinatorial results of \texttt{hydra-battles} and the set-theoretic content of \texttt{Gaia}.
-For instance, the associativity of multiplication of ordinals below $\epsilon_0$, already proven in \texttt{Gaia} could be tranferred ``for free'' to the data structures of \texttt{hydra-battles}, thanks to the isomorphism between both implementations.
+It would be interesting to build a library which could be used as a gate giving access to both the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}.
+For instance, the associativity of multiplication of ordinals below $\epsilon_0$, already proven in \textit{Gaia} could be tranferred ``for free'' to the data structures of \textit{hydra-battles}, thanks to the isomorphism between both implementations.
 
-\section{Acknowledgments}
-\label{sect:acks}
+%\section{Acknowledgments}
+%\label{sect:acks}
 

### 2021-09-10 doc/JFLA2022/paper.tex
draft paper

@@
 \newcommand{\latexeditor}{LEd}
+\newcommand{\canonseq}[2]{\mbox{$\{#1\}(#2)$}}
 
@@
 
+Our project is to illustrate  this work by a library written in \emph{Coq} and an electronic book containing explanations of the mathematical content and its possible mechanized formalisations.
+
 Hydra battles have already been presented in
@@
 
+\begin{todo}
+  Comment the multiple views of a same mathematical object:
+  \begin{itemize}
+  \item Ordinals notations: for $\omega$, $\omega^\omega$,
+    $\epsilon_0$, $\Gamma_0$. Proofs by computation whenever possible (see Fig.~\ref{Ex42-E0}.
+  \item (Axiomatic) model by Schütte. Less computations.
+    application of previously proven lemmas (see Fig~\ref{fig:Ex42b}).
+  \item Bridge lemmas (relative correctness of different representations).
+    \item The user/reader may compare  different proofs of the same result (For instance the equality $\omega+42+\omega^2= \omega^2$).
+  \end{itemize}
+\end{todo}
+
 \section{New contents}
-Length of usually considered battles.
-They correspond to the notions of canonical sequences and large sets in \cite{KS81}.
+We are interested in a usual variant of hydra-battles, where the hydra grows $n$ copies at step $n$ of the game (see the examples presented in ~\cite{KP82, BauerHydra}).
+In terms of ordinals, this variant is described as a (finite) sequence $s_n\,(n\geq i)$ of ordinals strictly below $\epsilon_0$.
+\begin{itemize}
+\item If $s_n=0$, then the sequence is empty,
+\item if $s_n=\beta+1$, then $s_{n+1}=\beta$,
+\item Otherwise, $s_n$ is a \emph{limit ordinal} $\lambda$; then $s_{n+1}$ is $\canonseq{\lambda}{n+1}$, \emph{i.e.} the
+    $n+1$-th element of the \emph{canonical sequence} associated with $\lambda$~\cite{KS81}.
+\end{itemize}
 
-\begin{todo}
-  Relation with Russel's library (maintained in Coq-community).
-\end{todo}
+The length of such a sequence is $l-i$, where $s_l=0$.
+It depends both on the initial index $i$ and the initial term $\alpha=s_i$.
+Adapting to \emph{Coq} a few lemmas from~\cite{KS81}, we prove that this length is minored by a fast growing function.
+More precisely, the length  of the sequence associated with
+$\omega^\alpha$ and initial index $i\geq 2$ is greater or equal than
+$F_{\omega^\alpha}(i-1)-(i-1)$, where $F_\alpha$ is a variant of the Wainer hierarchy of rapidly growing functions, defined and studied in~\cite{KS81}.
+From this minoration, we infer that the function which associates to $i$ the length of the considered sequence (for a given $\alpha$) is not primitive recursive in general (our counter-example is for $\alpha=\omega^\omega$).
+
+It is important to notice that, for proving this last result, we decided to host and maintain within \emph{Coq-Community} the formalization of primitive recursive function, part of Russel O'Connor work on G\"{o}del's first incompleteness theorem~\cite{OConnor05}.
+
+Defining and studying families of rapidly growing arithmetic  functions indexed by ordinals below $\epsilon_0$ has been made possible by the plug-in \emph{coq-equations}~\cite{sozeau:hal-01671777}.
+
+\section{The Coq-community ecosystem}
 
@@
 \fbox{\begin{minipage}[h]{1.0\linewidth}
-  Let us prove again the equality $\omega+42+\omega^2= \omega^2$.Let us recall that $\omega^2$ is an abbreviation of $\phi_0(2)$,
+  Let us prove again the equality $\omega+42+\omega^2= \omega^2$. Let us recall that $\omega^2$ is an abbreviation of $\phi_0(2)$,
 \emph{i.e} the third  additive principal ordinal.
@@
 
-\section{Perspectives}
+\section{Conclusion and perspectives}
+
+\begin{todo}
+  \begin{itemize}
+  \item  Soundness (correction, maintenance) of the Alectryon generated doc.
+  \item Communications between co-maintained libraries
+          (Ackermann, Gaia, etc.)
+  \item Work destined both to Coq-users and people interested in discrete math ?
+  \end{itemize}
+\end{todo}
+
+\subsection{Perspectives}
+
 The Gaia project~\cite{Gaia}, also maintained on \textit{Coq-community} contains a development by José Grimm~\cite{grimm:hal-00911710}.

### 2021-09-10 doc/JFLA2022/paper.tex
draft paper

@@
 
+% -------------------------------
 %%%% For Alectryon
@@
 
-% use this if you have a long article and want to create an index
-% \usepackage{makeidx}
+%---------------------------- 
+\newcommand{\canonseq}[2]{\mbox{$\{#1\}(#2)$}}
+\newtheorem{todo}{To do}
+\usepackage{amsfonts}
 
-% In order to save space or manage large tables or figures in a
-% landcape-like text, you can use the rotating and pdflscape
-% packages. Uncomment the desired from the below.
-%
-% \usepackage{rotating}
-% \usepackage{pdflscape}
-
-% Some of our commands for this guide.
-%
 \newcommand{\easychair}{\textsf{easychair}}
@@
 \newcommand{\latexeditor}{LEd}
-\newcommand{\canonseq}[2]{\mbox{$\{#1\}(#2)$}}
-
-\newtheorem{todo}{To do}
 
@@
 \title{Hydras and Cie \\
-  Variations on Kirby \& Paris' hydra battles and other entertaining math in Coq}
+  Variations on Kirby \& Paris' hydra battles and ordinal numbers in Coq}
 
@@
 \begin{abstract}
-  We present the evolution of our library \cite{HydraBattles}
+  \begin{todo}
+  We present the current state and evolution of a  library \cite{HydraBattles}
 since \cite{JFLA2018paper}
+  \end{todo}
 \end{abstract}
 
-\setcounter{tocdepth}{2}
-{\small
-\tableofcontents}
+% \setcounter{tocdepth}{2}
+% {\small
+% \tableofcontents}
 
@@
 In terms of ordinals, this variant is described as a (finite) sequence $s_n\,(n\geq i)$ of ordinals strictly below $\epsilon_0$.
+This sequence is parameterized by an initial index $i\in\mathbb{N}$ and an ordinal $\alpha<\epsilon_0$.
+
 \begin{itemize}
+  \item $s_i=\alpha$
 \item If $s_n=0$, then the sequence is empty,
@@
 
-The length of such a sequence is $l-i$, where $s_l=0$.
-It depends both on the initial index $i$ and the initial term $\alpha=s_i$.
-Adapting to \emph{Coq} a few lemmas from~\cite{KS81}, we prove that this length is minored by a fast growing function.
-More precisely, the length  of the sequence associated with
+The length of such a sequence is $l-i$, where $l$ is the first --- and unique --- natural number such that $s_l=0$.
+For a given $\alpha$, this length is a function of 
+the initial index $i$.
+
+Adapting to \emph{Coq} a few lemmas from~\cite{KS81}, we prove that the length  of the sequence associated with
 $\omega^\alpha$ and initial index $i\geq 2$ is greater or equal than
 $F_{\omega^\alpha}(i-1)-(i-1)$, where $F_\alpha$ is a variant of the Wainer hierarchy of rapidly growing functions, defined and studied in~\cite{KS81}.
+
 From this minoration, we infer that the function which associates to $i$ the length of the considered sequence (for a given $\alpha$) is not primitive recursive in general (our counter-example is for $\alpha=\omega^\omega$).

### 2021-09-10 doc/JFLA2022/paper.tex
draft paper

@@
 Since the set \textit{AP} of additive principals  is closed under addition
-(by Lemma \textit{AP\_plus\_closed}, it suffices to prove the inequalities $\omega<\phi_0(2)$ and $42<\phi_0(2)$.
+(by Lemma \textit{AP\_plus\_closed}), it suffices to prove the inequalities $\omega<\phi_0(2)$ and $42<\phi_0(2)$.
 
-\inputsnippets{Schutte/Ex42c, Schutte/Ex42d}
+\inputsnippets{Schutte/Ex42c, Schutte/Ex42d, Schutte/Ex42e}
 

### 2021-09-11 doc/JFLA2022/paper.tex
minor corrections

@@
 \newtheorem{todo}{To do}
-\usepackage{amsfonts}
+\usepackage{amsfonts, afterpage}
 
@@
   \begin{todo}
-  We present the current state and evolution of a  library \cite{HydraBattles}
-since \cite{JFLA2018paper}
+ \dots
   \end{todo}
@@
 L. Kirby and J. Paris~\cite{KP82}: \emph{Accessible Independence Results for Peano Arithmetic}.
+This article presents a gameful interpretation of accessibility problem in the set of ordinals below $\epsilon_0$, and proves two theorems:
+\begin{quote}
+\begin{itemize}
+\item In the Hydra game, every strategy is a winning strategy for Hercules.
+  \item The theorem above cannot be proved in Peano Arithmettic.
+\end{itemize}
+\end{quote}
+
+Like Goodstein sequences, hydra battles may be considered as a simplified version of G\"{o}del-like theorems.
 Although the mathematical contents of this 
@@
 
-Our project is to illustrate  this work by a library written in \emph{Coq} and an electronic book containing explanations of the mathematical content and its possible mechanized formalisations.
-
-Hydra battles have already been presented in
-JFLA2018~\cite{JFLA2018paper}. We present the recent evolution of the library: new results, interaction within the Coq-community project, and documentation generated with Alectryon.
+Our project is to illustrate  this work by a library written in \emph{Coq}~\cite{HydraBattles} and an electronic book~\cite{HydraBook}  containing explanations of the mathematical content and its possible mechanized formalisations.
+This project has already been presented in
+JFLA2018~\cite{JFLA2018paper}.
+We present the recent evolution of the library: new results, interaction within the Coq-Community project~\cite{CoqCommunity}, and documentation generated with Alectryon~\cite{alectryonpaper, alectryongithub}.
 
@@
 
-\begin{figure}[ph]
+%\afterpage{\clearpage}
+
+\begin{figure}[th]
   \centering

### 2021-09-12 doc/JFLA2022/paper.tex
minor corrections

@@
 
+\usepackage[firstpageonly=false, color={[gray]{0.5}},
+   scale=2.0, text=DRAFT]{draftwatermark}
 % -------------------------------
@@
 %
-\title{Hydras and Cie \\
+\title{Hydras and Cie (Draft) \\
   Variations on Kirby \& Paris' hydra battles and ordinal numbers in Coq}
@@
 \and
-    Jérémy Damour %\inst{2}
+    Jérémy Damour \inst{2}
 \and
-Karl Palmskog %\inst{3}
-\and Clément Pit-Claudel %\inst{4}
+Karl Palmskog \inst{2}
+\and Clément Pit-Claudel \inst{2}
 \and Théo Zimmermann %\inst{5}
@@
 \and
-   University of Miami,
-   Miami, Florida, U.S.A.\\
-   \email{geoff@cs.miami.edu}\\
-\and
-   University of Manchester,
-   Manchester, U.K.\\
-   \email{andrei@voronkov.com}\\
-\and
-   Chalmers University of Technology,
-   Gothenburg, Sweden
-\and
-   EasyChair
+   Ask for agreement !
  }
@@
 
-It is important to notice that, for proving this last result, we decided to host and maintain within \emph{Coq-Community} the formalization of primitive recursive function, part of Russel O'Connor work on G\"{o}del's first incompleteness theorem~\cite{OConnor05}.
+It is important to notice that, for proving this last result, we decided to host and maintain within \emph{Coq-Community} the formalization of primitive recursive function, part of Russel O'Connor's work on G\"{o}del's first incompleteness theorem~\cite{OConnor05}.
 

### 2021-09-12 doc/chapter-powers.tex
minor corrections

@@
 
-\begin{Coqsrc}
-Fixpoint power`{M: @EMonoid A  E_op E_one E_eq} 
-               (x:A) (n:nat) :=
-match n with 
-| 0%nat => E_one
-| S p =>   x * x ^ p
-end
-where "x ^ n" := (power x n) : M_scope.
-\end{Coqsrc}
+\inputsnippets{Pow/powerDef}
 
@@
 
-\begin{Coqsrc}
-Lemma power_eq1 {A:Type} `{M: @EMonoid A  E_op E_one E_eq} 
-               (x:A) :  x ^ 0 = E_one.
-Proof. reflexivity. Qed.
-
-Lemma power_eq2 {A:Type} `{M: @EMonoid A  E_op E_one E_eq}
-                (x:A) (n:nat) :
-                x ^ (S n)  = x * x ^ n.
-Proof. reflexivity. Qed.
-
-Lemma power_eq3 {A:Type} `{M: @EMonoid A  E_op E_one E_eq}
-                (x:A) : x ^ 1 == x.
-Proof. cbn; rewrite Eone_right; reflexivity. Qed.
-\end{Coqsrc}
+\inputsnippets{Pow/powerEqns}
 
@@
 
-\begin{Coqsrc}
-Open Scope M_scope.
-
-Compute 22%Z ^ 20.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-= 705429498686404044207947776%Z
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Import Int31.
-Coercion phi_inv : Z >-> int31.
-
-Compute (22%int31 ^ 20).
-\end{Coqsrc}
-
-\begin{Coqanswer}
-   = 2131755008%int31
-     : int31
-\end{Coqanswer}
+\inputsnippets{Demo_power/Demopower}
 
@@
 
-\vspace{4pt}
+\pagebreak
+
 \emph{From Module~\href{../theories/html/additions.Pow.html}{additions.Pow}}
 
-\begin{Coqsrc}
-Fixpoint binary_power_mult `{M: @EMonoid A E_op E_one E_eq}
-             (x a:A)(p:positive) : A 
-  :=
-  match p with
-    | xH =>    a * x
-    | xO q => binary_power_mult  (x * x) a q
-    | xI q =>   binary_power_mult (x * x) (a * x) q
-  end.
-
-Fixpoint Pos_bpow  `{M: @EMonoid A E_op E_one E_eq} 
-         (x:A)(p:positive) :=
- match p with
-  | xH => x
-  | xO q => Pos_bpow  (x * x) q
-  | xI q => binary_power_mult (x * x) x q
-end.
-\end{Coqsrc}
+\inputsnippets{Pow/binaryPowerMult, Pow/PosBpow}
 
@@
 From Module~\href{../theories/html/additions.Pow.html}{additions.Pow}
-
-\begin{Coqsrc}
-Definition N_bpow {A} `{M: @EMonoid A E_op E_one E_eq} x (n:N) := 
-  match n with 
-  | 0%N => E_one
-  | Npos p => Pos_bpow x p
-  end.
-
-Infix "^b" := N_bpow (at level 30, right associativity): M_scope.
-\end{Coqsrc}
 
@@
 
-\begin{Coqsrc}
-Lemma N_bpow_ok : 
-forall (x:A) (n:N),   x ^b n  == x ^ N.to_nat n.
-(* Proof omitted *)
-\end{Coqsrc}
-
+\inputsnippets{Pow/NBpowOk}
 \end{itemize}

### 2021-09-12 doc/chapter-powers.tex
more snippets

@@
 
-\begin{Coqsrc}
-Section M_given.
- Variables (A:Type) (E_op : Mult_op A)(E_one:A) (E_eq : Equiv A).
- Context (M:EMonoid  E_op E_one E_eq).
-\end{Coqsrc}
+\inputsnippets{Pow/MGiven}
 
@@
 
-\begin{Coqsrc}
-Ltac monoid_rw :=
-    rewrite Eone_left  ||
-    rewrite Eone_right  || 
-    rewrite Eop_assoc .
-
-Ltac monoid_simpl := repeat monoid_rw.
-
-Section About_power.
-\end{Coqsrc}
+\inputsnippets{Pow/MonoidRw}
 
@@
 \index{coq}{Type classes!Proper class}
-\begin{Coqsrc}
-Global Instance power_proper :
-     Proper (equiv ==> eq ==> equiv) power.
-(* Proof omitted *)
-\end{Coqsrc}
+
+\inputsnippets{Pow/powerProper}
 
@@
 
-\begin{Coqsrc}
-Lemma power_of_plus :   forall x n p, x ^ (n + p) ==  x ^ n *  x ^ p.
-(* Proof omitted *)
-
-Ltac power_simpl := 
-    repeat (monoid_rw || rewrite <- power_x_plus).
-\end{Coqsrc}
+\inputsnippets{Pow/powerOfPlus, Pow/powerSimpl}
 

### 2021-09-13 doc/JFLA2022/paper.tex
small corrections

@@
 By definition of additive principal ordinals, 
-it suffices to prove the inequality $\omega+42< \phi_0(2)$ to get the equality we want to prove.
+it suffices to prove the inequality $\omega+42< \phi_0(2)$ to  fprove our lemma.
 
@@
 
-\inputsnippets{Schutte/Ex42c, Schutte/Ex42d, Schutte/Ex42e}
+\inputsnippets{Schutte/Ex42d, Schutte/Ex42c, Schutte/Ex42e}
 

### 2021-09-13 doc/part-hydras.tex
minor changes
draft paper
small changes
small corrections

@@
 
-The proof is very different from the computational proof of Sect~\vref{Ex42-E0}.
+Our proof is very different from the computational proof of Sect~\vref{Ex42-E0}.
 By definition of additive principal ordinals, 
-it suffices to prove the inequality $\omega+42< \phi_0(2)$ to get the equality we want to prove.
+it suffices to prove the inequality $\omega+42< \phi_0(2)$.
 
@@
 
-\input{movies/snippets/Schutte/Ex42d}
+\inputsnippets{Schutte/Ex42d, Schutte/Ex42c,  Schutte/Ex42e}
 

### 2021-09-13 doc/JFLA2022/paper.tex
small corrections

@@
 \begin{quote}
-\begin{itemize}
-\item In the Hydra game, every strategy is a winning strategy for Hercules.
-  \item The theorem above cannot be proved in Peano Arithmettic.
-\end{itemize}
+\begin{description}
+\item[$(1)$] In the Hydra game, every strategy is a winning strategy for Hercules.
+  \item[$(2)$] The theorem $(1)$ cannot be proved in Peano Arithmetic.
+\end{description}
 \end{quote}
@@
 By definition of additive principal ordinals, 
-it suffices to prove the inequality $\omega+42< \phi_0(2)$ to  fprove our lemma.
+it suffices to prove the inequality $\omega+42< \phi_0(2)$.
 

### 2021-09-13 doc/chapter-powers.tex
fix errors in tex files
more snippets
Alectryon: chapter on addition chains 40% done

@@
 
+\inputsnippets{Pow/NBpow}
+
 \subsection{Refinement and correctness}
@@
 
-  Please note that the following two lemmas \emph{do not require} 
+  Please note that the following lemmas \emph{do not require} 
 the operation~\texttt{*} to be commutative.
 
-\begin{Coqsrc}
-Lemma power_commute : 
-    forall x n p, x ^ n * x ^ p ==  x ^ p * x ^ n. 
-(* Proof omitted *)
-
-Lemma power_commute_with_x : 
-    forall x n,  x * x ^ n == x ^ n * x.
-(* Proof omitted *)
-
-Lemma power_of_power : 
-   forall x n p,  (x ^ n) ^ p == x ^ (p * n).
-(* Proof omitted *)
-
-\end{Coqsrc}
+\inputsnippets{Pow/powerCommute}
 
@@
 
-\begin{Coqsrc}
-Lemma sqr_def : forall x, x ^ 2 ==  x * x.
-(* Proof omitted *)
-
-Lemma power_of_square : 
-  forall x n, (x * x) ^ n ==  x ^ n * x ^ n.
-(* Proof omitted *) 
-\end{Coqsrc}
+\inputsnippets{Pow/sqrEqn, Pow/powerOfSquare}
 
@@
 
-\begin{Coqsrc}
-Lemma binary_power_mult_ok :
-  forall p a x,   binary_power_mult  x a p  ==  
-                  a * x ^ Pos.to_nat p.
-Proof.
-  induction p as [q IHq | q IHq| ].
- (* Rest of proof omitted *)
- \end{Coqsrc}
-
- \begin{Coqsrc}
-Lemma Pos_bpow_ok : 
-  forall p x, Pos_bpow x p == x ^ Pos.to_nat p.
-(* Proof omitted *)
-
-Lemma N_bpow_ok : 
-  forall n x,  x ^b n  == x ^ N.to_nat n.
-(* Proof omitted *)
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma N_bpow_ok_R : 
-  forall n x, x ^b (N.of_nat n)   ==  x ^  n.
-(* Proof omitted *)
-
-Lemma Pos_bpow_ok_R : 
-   forall p x, p <> 0 ->
-                      Pos_bpow x  (Pos.of_nat p)   ==  x ^  p.
-(* Proof omitted *)
-
-End About_power.  
-\end{Coqsrc}
+\inputsnippets{Pow/binaryPowerMultOk}
+
+\inputsnippets{Pow/PosBpowOk, Pow/PosBpowOkR,
+  Pow/NBpowOk,
+  Pow/NBpowOkR}
 
@@
 
-\begin{Coqsrc}
-Lemma N_bpow_commute : forall x n p,  
-                        x ^b n *  x ^b p ==  
-                        x ^b p *  x ^b n.
-Proof.
- intros x n p; repeat rewrite N_bpow_ok.
- rewrite power_commute; reflexivity.
-Qed.  
-\end{Coqsrc}
+\inputsnippets{Pow/NBpowCommute}
 
@@
 
-\begin{Coqsrc}
-Definition fib_pos n :=
-  let (a,b) := Pos_bpow (M:= Mul2) (1,0) n in
-  (a+b).
-
-Time Compute fib_pos 153%positive.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  68330027629092351019822533679447
-     : N 
-Finished transaction in 0.002 secs (0.002u,0.s) (successful)
-\end{Coqanswer}
+\emph{From Module~\href{../theories/html/additions.Fib2.html}{additions.Fib2}}
+
+\inputsnippets{Fib2/fibPos}
 
@@
 
-\begin{Coqsrc}
-Eval simpl in   fun (x:A) => x ^b 17.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = fun x : A =>
-       x *
-       (x * x * (x * x) * (x * x * (x * x)) *
-        (x * x * (x * x) * (x * x * (x * x))))
-     : A -> A  
-\end{Coqanswer}
+\inputsnippets{Pow/bpow17}
 
@@
 
-\begin{Coqsrc}
-Eval simpl in   fun x =>  x ^ 17.  
-\end{Coqsrc}
-
-\begin{Coqanswer}
-= fun x : A =>
-   x * (x *  (x *  (x *  (x *   (x *  (x *  (x *
-    (x * (x * (x * (x * (x * (x * (x * (x * (x * one)))))
-  )))))))))))  
-\end{Coqanswer}
+\inputsnippets{Pow/naivePow17}
 
 From these tests, we may infer that  representing exponentiation algorithms as \coq{}  functions hides
-information about the real structure of the computations, particularly the sharing on intermediate computations.
+information about the real structure of the computations, particularly about  sharing intermediate computations.
 
@@
 
-\begin{Coqsrc}
-Definition pow_17  (x:A) :=
-  let x2 := x * x in
-  let x4 := x2 * x2 in
-  let x8 := x4 * x4 in
-  let x16 := x8 * x8 in
-  x16 * x.
-\end{Coqsrc}
+\inputsnippets{Pow/pow17LetIn}
+
 \label{pow-17-let-in}
@@
 
-\begin{Coqsrc}
-Eval cbv  zeta beta delta [pow_17]  in  pow_17.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = fun x : A =>
-       x * x * (x * x) * (x * x * (x * x)) *
-       (x * x * (x * x) * (x * x * (x * x))) * x
-     : A -> A 
-\end{Coqanswer}
-
-In the next section, we propose to use a \emph{data structure} for representing 
+\inputsnippets{Pow/evalPow17LetIn}
+
+In the next section, we define a \emph{data structure} for representing 
 the computations that lead to the evaluation of some power $x^n$, where
@@
 
-\begin{Coqsrc}
-Inductive computation {A:Type}  : Type :=
-| Return (a : A)
-| Mult (x y : A) (k : A -> computation).    
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/computationDef}
+
 \subsubsection{Monadic notation}
@@
 
-\begin{Coqsrc}
-Notation "z '<---'  x 'times' y ';' e2 " :=
-  (Mult x y  (fun z => e2))
-    (right associativity, at level 60).
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/monadicComputation}
 
@@
 
-\begin{Coqsrc}
-Example comp7 : computation  :=
-  x <--- 2 times 2;
-  y <--- x times 2;
-  z <--- y times y ;
-  t <--- 2 times z ;
-  Return t.  
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/comp128}
 
@@
 
-\begin{Coqsrc}
-Definition chain := forall A:Type, A -> @computation A.   
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/chainDef}
 
@@
   \centering
-  \begin{Coqsrc}
-Example  C87 : chain :=
- fun A (x : A) =>
-  x2 <--- x times x ;
-  x3 <--- x2 times x ;
-  x6 <--- x3 times x3 ;
-  x7 <--- x6 times x ;
-  x10 <--- x7 times x3 ;
-  x20 <--- x10 times x10 ;
-  x40 <--- x20 times x20 ;
-  x80 <--- x40 times x40 ;  
-  x87 <--- x80 times x7 ;
-  Return x87.
- \end{Coqsrc}
+  \inputsnippets{Addition_Chains/C87}
    \caption{A chain for raising x to its $87$-th power}
@@
 \label{C87-length}
-\begin{Coqsrc}
-Fixpoint computation_length {A} (a:A)(m : @computation A) 
-  : nat :=
-match m with
-  | Mult _ _ k => S (computation_length a (k a))
-  | _ => 0%nat
-end.
-
-Definition chain_length (c:chain) 
-   := computation_length tt (c _ tt).
-
-Compute chain_length C87.
-\end{Coqsrc}
-\begin{Coqanswer}
- = 9 : nat  
-\end{Coqanswer}
+
+\inputsnippets{Addition_Chains/chainLength}
 
@@
 
-\begin{Coqsrc}
-Fixpoint computation_execute  {A:Type} (op: Mult_op A) 
-                              (c : computation) :=
-match c with 
-| Return x => x 
-| Mult x y k => computation_execute op (k (x * y))
-end.
-
-Definition chain_execute (c:chain) {A} op  (a:A) :=
-  computation_execute op (c A a).
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Definition computation_eval `{M:@EMonoid A E_op E_one E_eq}
-           (c : computation) : A := computation_execute E_op c.
-
-Definition chain_apply (c:chain) 
-    {M:@EMonoid A E_op E_one E_eq} a : A :=
-    computation_eval (c A a).
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/chainExecute}
+
+\subsubsection*{Examples:} 
+The following interactions show how to apply the chain \texttt{C87} 
+for exponentiation within two different monoids:
+
+\inputsnippets{Addition_Chains/C87Apply}
 
@@
 \end{project}
-
-\subsubsection*{Examples:} 
-The following interactions show how to apply the chain \texttt{C87} 
-for exponentiation within two different monoids:
-
-\begin{Coqsrc}
-Compute  chain_apply C87 3%Z.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- =  323257909929174534292273980721360271853387%Z
-     : Z
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Compute chain_apply C87 (M:=M2N) (Build_M2 1 1 1 0)%N.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-  = {|
-      c00 := 1100087778366101931%N;
-      c01 := 679891637638612258%N;
-      c10 := 679891637638612258%N;
-      c11 := 420196140727489673%N |}
-     : M2 N
-\end{Coqanswer}
 
@@
 
-\begin{Coqsrc}
-Compute chain_trace C87.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-(1, Init)
-       :: (2, Add 1 1)
-          :: (3, Add 2 1)
-             :: (6, Add 3 3)
-                :: (7, Add 6 1)
-                   :: (10, Add 7 3)
-                      :: (20, Add 10 10)
-                         :: (40, Add 20 20)
-                            :: (80, Add 40 40) :: (87, Add 80 7) :: nil
-     : list (positive * info)
-\end{Coqanswer}
+\inputsnippets{Trace_exercise/traceC87}
 
@@
 \label{chain-correct-def}
-\begin{Coqsrc}
-Definition chain_correct_nat (n:nat) (c: chain) := 
-  n <> 0 /\
-  forall `(M:@EMonoid  A E_op E_one E_eq) (x:A), 
-      chain_apply c x ==   x ^ n.
-
-Definition chain_correct (p:positive) (c: chain) :=
-  chain_correct_nat c (Pos.
-to_nat p). 
-\end{Coqsrc}
+
+\inputsnippets{Addition_Chains/chainCorrect}
 
@@
 
- \begin{Coqsrc}
-Definition optimal (p:positive) (c : chain) :=
- forall c', chain_correct p c' -> 
-            (chain_length c <= chain_length c')%nat.
- \end{Coqsrc}
+\inputsnippets{Addition_Chains/optimalDef}
 
@@
 
-\begin{Coqsrc}
-Ltac slow_chain_correct_tac :=
-  match goal with 
-      [ |- chain_correct ?p ?c ] =>
-      let A := fresh "A" in
-      let op := fresh "op" in
-      let one := fresh "one" in
-      let eqv := fresh "eqv" in
-      let M := fresh "M" in
-      let x := fresh "x"
-      in  split;
-        [discriminate | 
-         unfold c, chain_apply, computation_eval; simpl;
-         intros A op one eq M x; monoid_simpl M; reflexivity]
-  end.
-
-Example C7_ok : chain_correct 7 C7.
-Proof.
-   slow_chain_correct_tac.
-Qed.
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/slowTac}
 
@@
 
-\begin{Coqsrc}
-Example C87_ok : chain_correct 87 C87.
-Proof.
- Time  slow_chain_correct_tac. 
-\end{Coqsrc}
-
-\begin{Coqanswer}
- Finished transaction in 62.808 secs (62.677u,0.085s) (successful)   
- \end{Coqanswer}
-
-\begin{Coqsrc}
-Qed.
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/slowC87Correct}
 
@@
 
-\begin{Coqanswer}
-1 subgoal, subgoal 1 (ID 219)
-
-  A : Type
-  E_op : Mult_op A
-  E_one : A
-  E_eq : Equiv A
-  M : EMonoid E_op E_one E_eq
-  x : A
-  ============================
-   x * x * x * (x * x * x) * x * (x * x * x) *
-   (x * x * x * (x * x * x) * x * (x * x * x)) *
-   (x * x * x * (x * x * x) * x * (x * x * x) *
-    (x * x * x * (x * x * x) * x * (x * x * x))) *
-   (x * x * x * (x * x * x) * x * (x * x * x) *
-    (x * x * x * (x * x * x) * x * (x * x * x)) *
-    (x * x * x * (x * x * x) * x * (x * x * x) *
-     (x * x * x * (x * x * x) * x * (x * x * x)))) *
-   (x * x * x * (x * x * x) * x) ==
-   x *
-   (x *
-    (x *
-     (x *
-      (x *
-       (x *
-        (x *
-         (x *
-          (x *
-           (x *
-            (x *
-             (x *
-              (x *
-               (x * (x * (x * (x * (x * (x * (x * (x * (x * (x * 
-                ..))))))))))))))))))))))
-\end{Coqanswer}
-\label{fig:big-goal}
-
-This goal is solved by the following tactic composition:
-
-\begin{Coqsrc}
-monoid_simpl M; reflexivity. 
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/WhySoSlow}
 
@@
 
-\begin{Coqsrc}
-(** Binary trees of multiplications over A *)
-
-Inductive Monoid_Exp (A:Type) : Type :=
- Mul_node (t t' : Monoid_Exp A) | One_node | A_node (a:A).
-
-Arguments Mul_node {A} _ _.
-Arguments One_node {A} .
-Arguments A_node {A} _ .
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/MonoidExp}
 
@@
 
-\begin{Coqsrc}
-
-Fixpoint flatten_aux {A:Type} (t fin : Monoid_Exp A) 
-  : Monoid_Exp A :=
-match t with Mul_node  t t' =>
-              flatten_aux t (flatten_aux t' fin)
-           | One_node  => fin
-           |  x => Mul_node  x fin
-end.
-
-Fixpoint flatten {A:Type} (t: Monoid_Exp A) : Monoid_Exp A :=
-match t with
-| Mul_node t t' => flatten_aux t (flatten t')
-| One_node => One_node
-| X => Mul_node X One_node
-end.
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/flattenDef}
 
@@
 
-\begin{Coqsrc}
-Function eval {A:Type} {op one eqv}
-         (M: @EMonoid A op one eqv)
-         (t: Monoid_Exp A) : A :=
- match t with 
-            | Mul_node t1 t2 => (eval M t1 * eval M t2)%M
-            | One_node => one
-            | A_node a => a
-end.
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/evalDef}
 
@@
 
-\begin{Coqsrc}
-Lemma flatten_valid {A} `(M: @EMonoid A op one eqv):
-forall t , eval M t == eval M (flatten t).
-(* Proof omitted *)
-
-Lemma flatten_valid_2 {A} `(M: @EMonoid A op one eqv):
-forall t t' , eval  M (flatten t) == eval M (flatten t')  ->
-     eval M t == eval M t'.
-(* Proof omitted *)
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/flattenValid}
+\inputsnippets{Addition_Chains/flattenValid2}
 
 \subsubsection{Transforming a multiplication into a tree}
-Let us now build a tool for building terms of type  \texttt{Monoid\_Exp $A$} out
+Let us now build a tool for building terms of type  (\texttt{Monoid\_Exp $A$}) out
 of terms of type \texttt{A} containing multiplications of the form 
@@
 
-\begin{Coqsrc}
-(** "Quote" tactic *)
-
-Ltac model A  op one v :=
-match v with 
-| (?x  * ?y)%M => let r1 := model A op one x
-                  with r2 := model A op one y 
-                  in  constr:(@Mul_node A r1 r2)
-| one => constr:(@One_node A)
-| ?x => constr:(@A_node A x)
-end.
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/modelTac}
 

### 2021-09-13 doc/chapter-powers.tex
Alectryon: chapter on addition chains 46% done

@@
 
-\begin{Coqsrc}
-Ltac monoid_eq_A A op one E_eq M  :=
-match goal with 
-| [ |- E_eq  ?X ?Y ] =>
-  let tX := model A op one X with
-      tY := model A op one Y in
-      (change (E_eq (eval M tX) (eval M tY)))
-end.
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/monoidEqTac}
 
@@
 
-\begin{Coqsrc}
-Ltac reflection_correct_tac :=
-match goal with
-[ |- chain_correct ?n ?c ] =>
- split; [try discriminate |
-         let A := fresh "A"
-         in let op := fresh "op"
-         in let one := fresh "one" 
-         in let E_eq := fresh "eq" 
-         in let M := fresh "M"
-         in let x := fresh "x" 
-         in  (try unfold c); unfold chain_apply;
-           simpl; red; intros  A op one E_eq M x;
-           unfold computation_eval;simpl;
-           monoid_eq_A A op one E_eq M;
-           apply flatten_valid_2;try reflexivity
-        ]
-end. 
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/reflectionCorrectTac}
 
@@
 
-\begin{Coqsrc}
-Example C87_ok : chain_correct 87 C87.
-Proof.
-  Time reflection_correct_tac.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- Finished transaction in 0.038 secs (0.038u,0.s) (successful)
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Qed. 
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/reflectionDemo}
 
@@
 
-\begin{Coqsrc}
-Inductive computation {A:Type}  : Type :=
-| Return (a : A)
-| Mult (x y : A) (k : A -> computation).
-
-Definition chain := forall A:Type, A -> @computation A.
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/computationDef,
+  Addition_Chains/chainDef}
 
@@
 
-\begin{Coqanswer}
-Inductive
-computation_R (A B : Type) (R : A -> B -> Type)
-  : computation -> computation -> Type :=
- |    computation_R_Return_R : 
-       forall (a1 : A) (a2 : B),  R a1 a2 ->
-         computation_R A B R (Return a1) (Return a2)
-  | computation_R_Mult_R : forall (x1 : A) (x2 : B),
-         R x1 x2 ->
-          forall (y1 : A) (y2 : B),
-            R y1 y2 ->
-            forall (k1 : A -> computation)
-                   (k2 : B -> computation),
-                      (forall (H : A) (H0 : B),
-                          R H H0 -> 
-                          computation_R A B R (k1 H) (k2 H0)) ->
-                       computation_R A  B R 
-                          (z <--- x1 times y1; k1 z)
-                          (z <--- x2 times y2; k2 z)
-\end{Coqanswer}
+\pagebreak
+
+\inputsnippets{Addition_Chains/paramDemo}
 
@@
 
-\begin{Coqsrc}
-Definition parametric (c:chain) :=
-  forall A B (R: A -> B -> Type) (a:A) (b:B),
-   R a b -> computation_R  R (c A a) (c B b).
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/parametricDef}
 
@@
 
-\begin{Coqsrc}
-Definition the_exponent_nat (c:chain) : nat :=
- chain_apply c (M:=Natplus) 1%nat.
-
-Definition the_exponent (c:chain) : positive :=
-  chain_execute c Pos.add  1%positive.
-
-Compute the_exponent C87.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = 87%positive
-     : positive  
-\end{Coqanswer}
+\inputsnippets{Addition_Chains/theExponent}
 
@@
 
-\begin{Coqsrc}
-Definition power_R  (a:A) :=
-  fun (x:A)(n:nat) => n <> 0 /\ x == a ^ n.  
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/powerR}
 
@@
 
-\begin{Coqsrc}
-Lemma  power_R_is_a_refinemnt (a:A) :
-  forall(gamma : @computation A)
-        (gamma_nat : @computation nat),
-    computation_R  (power_R a) gamma gamma_nat -> 
-     power_R a (computation_eval gamma)
-               (computation_eval (M:= Natplus) gamma_nat).    
-(* Proof omitted *)
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/powerRRef}
 
@@
 
-\begin{Coqsrc}
-Lemma param_correctness_nat :
- forall c:chain, parametric c ->  
-               chain_correct_nat (the_exponent_nat c)  c.
-(* Proof omitted *)
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/paramCorrectnessNat}
 

### 2021-09-13 doc/chapter-powers.tex
Alectryon: chapter on addition chains half done

@@
 
-\begin{Coqsrc}
-Lemma exponent_pos2nat : forall c: chain,  parametric c -> 
-  the_exponent_nat c = Pos.to_nat (the_exponent c).
-
-Lemma exponent_pos_of_nat : forall c: chain,  parametric c -> 
-  the_exponent c = Pos.of_nat (the_exponent_nat c).
-
-Lemma param_correctness (c:chain) :
-   parametric c -> 
-   chain_correct (the_exponent c) c. 
-Proof.
-  intros; rewrite  exponent_pos_of_nat; auto.
-  red;  rewrite  exponent_pos2nat;auto.
-  rewrite Pos2Nat.id,  <- exponent_pos2nat;auto.
-  apply param_correctness_nat; auto.
-Qed.
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/exponentPosToNat}
+\inputsnippets{Addition_Chains/exponentPosOfNat}
+\inputsnippets{Addition_Chains/paramCorrectness}
 
@@
 
-\begin{Coqsrc}
-Ltac parametric_tac  := 
-match goal with [ |- parametric ?c] =>
-   red ; intros;
-   repeat (right;[assumption | assumption | ]);  
-   left; assumption
-end.
-
-Example P87 : parametric C87.
-Proof. Time parametric_tac. 
-\end{Coqsrc}
-
-\begin{Coqanswer}
-Finished transaction in 0.005 secs (0.005u,0.s) (successful)
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Qed. 
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/parametricTac}
 
@@
 
-\begin{Coqsrc}
-Ltac param_chain_correct :=
-match goal with 
-[|- chain_correct ?p ?c ] => 
-apply param_correctness; parametric_tac
-end.
-
-Lemma C87_ok' : chain_correct 87 C87.
-Time param_chain_correct.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-Finished transaction in 0.005 secs (0.005u,0.s) (successful)
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Qed.
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/paramChainCorrect}
 
@@
 
-\begin{Coqbad}
-Definition any_chain_parametric : Type :=
- forall c:chain, parametric c.
-
-Goal any_chain_parametric.
-Proof.
-intros c A B R a b ; induction c.
-\end{Coqbad}
-
-\begin{Coqanswer}
- 2 subgoals, subgoal 1 (ID 556)
-
-  c : chain
-  A : Type
-  B : Type
-  R : A -> B -> Type
-  a : A
-  b : B
-  a0 : A
-  ============================
-   R a b -> computation_R R (Return a0) (c B b)
-
-...
-\end{Coqanswer}
-
-\begin{Coqbad}
-Abort.
-\end{Coqbad}
+\inputsnippets{Addition_Chains/VeryBad,Addition_Chains/VeryBad2}
 

### 2021-09-14 doc/chapter-powers.tex
full book with Alectryon (Draft)

@@
 any positive integer and returns a chain.
-
-\begin{Coqsrc}
-Definition chain_generator := positive -> chain.  
-\end{Coqsrc}
-
 A generator $g$  is \emph{correct} it it returns a correct chain
-for any exponent:
-
-\begin{Coqsrc}
-Definition correct_generator (g : positive -> chain) :=
- forall p, chain_correct (g p) p.
-\end{Coqsrc}
+for any exponent.
+
+\inputsnippets{Addition_Chains/generatorDef}
 
@@
 
-\begin{Coqsrc}
-Definition cpower_pos (g : chain_generator)  p
-           `{M:@EMonoid A E_op E_one E_eq} a :=
-  chain_apply (g p) (M:=M) a.
-
-Definition cpower (g : chain_generator)  n
-           `{M:@EMonoid A E_op E_one E_eq} a :=
-  match n with 0%N => E_one 
-             | Npos p => cpower_pos  g p a
-  end.  
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/cpowerDef}
 
@@
 
-\begin{Coqsrc}
-
-Definition optimal_generator (g : positive -> chain) :=
- forall p:positive, optimal p (g p).
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/optimalGenerator}
 
@@
 product of an accumulator $a$ with an arbitrary power of some value $x$.
-Then, the main function builds a computation for any positive exponent:
-
-\begin{Coqsrc}
-Fixpoint axp_scheme  {A} p : A -> A -> @computation A   :=
- match p with
-   | xH =>  (fun a x => y <--- a  times x ; Return y)
-   | xO q => (fun a x => x2 <--- x times  x ; axp_scheme q a x2)
-   | xI q => (fun a x => ax <--- a times x ;
-                         x2 <--- x times x ;
-                         axp_scheme q ax x2)
-end.
-
-Fixpoint  bin_pow_scheme {A} (p:positive)  
-: A -> @computation A:=
-  match p with 
-  |  xH => fun x => Return x
-  | xI q  => fun x => x2 <--- x times x; axp_scheme q x x2
-  | xO q => fun x => x2 <--- x times x ; bin_pow_scheme q x2
-  end.
-
-\end{Coqsrc}
-
-The following function associates  a chain to any positive exponent:
-
-\begin{Coqsrc}
-Definition binary_chain (p:positive) : chain :=
-  fun A => bin_pow_scheme p.
-
-Compute binary_chain 87.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-    = fun (A : Type) (x : A) =>
-       x0 <--- x times x;
-       x1 <--- x times x0;
-       x2 <--- x0 times x0;
-       x3 <--- x1 times x2;
-       x4 <--- x2 times x2;
-       x5 <--- x4 times x4;
-       x6 <--- x3 times x5;
-       x7 <--- x5 times x5;
-       x8 <--- x7 times x7; 
-       x9 <--- x6 times x8; 
-       Return x9
-     : chain
-\end{Coqanswer}
+
+\inputsnippets{Addition_Chains/binaryChain}
 
@@
 
-\begin{Coqsrc}
-Section binary_power_proof.
-
-Variables (A: Type)
-         (E_op : Mult_op A)
-         (E_one : A)
-         (E_eq: Equiv A).
-
-Context (M : EMonoid  E_op E_one E_eq).
-
-Existing Instance Eop_proper.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma axp_correct : forall p a x,
- computation_eval (axp_scheme p a x) == a *  x ^ (Pos.to_nat p).
-(*  Proof by induction on p *)
-
-Lemma binary_correct : 
-  forall p x,
-        computation_eval  (bin_pow_scheme p (A:=A) x) ==
-        x ^ (Pos.to_nat p).
-(* Proof by induction on p *)
-
-End  binary_power_proof.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma binary_generator_correct : correct_generator binary_chain.
-Proof.
-  red;unfold chain_correct,  binary_chain, chain_apply;
-  split; [auto| intros A op one Eq M x; apply binary_correct].
-Qed.
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/binaryPowerProofa}
+\inputsnippets{Addition_Chains/binaryPowerProofb}
+\inputsnippets{Addition_Chains/binaryPowerProofc}
+\inputsnippets{Addition_Chains/binaryPowerProofd}
+\inputsnippets{Addition_Chains/binaryGeneratorOk}
 
@@
 for computing powers.
-    First, let us  assume that \texttt{binary\_chain} is optimal:
-
-\begin{Coqsrc}
-Section non_optimality_proof.
-
- Hypothesis binary_opt : optimal binary_chain.
-\end{Coqsrc}
-
-Then, let us consider for instance the binary chain generated for the exponent 87.
-
-\begin{Coqsrc}
-Compute chain_length (binary_chain 87).
-\end{Coqsrc}
-
-\begin{Coqanswer}
-   = 10 : nat 
-\end{Coqanswer}
-
-Let us recall that \texttt{C87}'s length has been evaluated to $9$ (Sect~\ref{C87-length},
-and that this chain is correct (Sect~\vref{C87-param-ok}). Thus, it is very easy
-to finish our proof:
-
-\begin{Coqsrc}
-Lemma binary_generator_not_optimal : False.
-Proof.
-  generalize (binary_opt gen _  _ C87_ok); 
-  compute; omega.
-Qed.
-
-End non_optimality_proof.
-\end{Coqsrc}
+
+\inputsnippets{Addition_Chains/nonOpt}
 
@@
 
-\begin{Coqsrc}
-Example C3 : chain :=
- fun A (x:A) =>
-  x2 <--- x times x;
-  x3 <--- x2 times x ;
-  Return x3.
-\end{Coqsrc}
+\inputsnippets{Addition_Chains/C3Example}
 
@@
 build  a computation for raising {$x$} to some  given exponent.
-
-\begin{Coqsrc}
-Definition Fkont (A:Type) := A -> @computation A.
-\end{Coqsrc}
-
 An \texttt{f-chain} is just a polymorphic function that combines  a 
-continuation and en element into a computation:
-
-\begin{Coqsrc}
-Definition Fchain := forall A, Fkont A -> A -> @computation A.
-\end{Coqsrc}
+continuation and an element into a computation.
+
+\inputsnippets{Euclidean_Chains/FchainDef}
 
@@
 
-\begin{Coqsrc}
-Definition F3 : Fchain := 
- fun  A k  (x:A) =>
-  y <--- x times x ;
-  z <--- y times x ;
-  k  z.
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/F3Def}
 
@@
 
-\begin{Coqsrc}
-Definition F2C (f : Fchain) : chain :=
- fun (A:Type) => f A Return.
-
-Compute the_exponent (F2C F3).
-\end{Coqsrc}
-
-\begin{Coqanswer}
-= 3%nat  
-\end{Coqanswer}
+\inputsnippets{Euclidean_Chains/F2C}
 
@@
 
-\begin{Coqsrc}
-Definition F1 : Fchain := 
- fun A k (x:A) => k x.
-
-Definition F2 : Fchain := 
-fun  A k  (x:A) =>
-  y <--- x times x ;
-  k  y.
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/F1F2}
 
@@
 
-\begin{Coqsrc}
-Definition Fapply (f : Fchain) (c: chain) : chain  :=
- fun A  x =>  f  A (fun y => c A  y) x.
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/Fapply}
 
@@
 
-\begin{Coqsrc}
-Definition Fcompose (f1 f2: Fchain) : Fchain  :=
- fun   A k x =>  f1  A (fun y => f2 A k y) x.
-
-Lemma F1_neutral_l : forall f, Fcompose F1 f = f.
-Proof. reflexivity. Qed.
-
-Lemma F1_neutral_r : forall f, Fcompose f F1 = f.
-Proof. reflexivity. Qed.
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/Fcompose}
+
+\inputsnippets{Euclidean_Chains/F1Neutral}
 
@@
 
-\begin{Coqsrc}
-Example F9 := Fcompose F3 F3.
-
-Compute F9.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-= fun (A : Type) (x : Fkont A) (x0 : A) =>
-       x1 <--- x0 times x0;
-       x2 <--- x1 times x0; x3 <--- x2 times x2; 
-       x4 <--- x3 times x2; 
-       x x4
-     : Fchain  
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Remark F9_correct :chain_correct 9 (F2C F9).
-Proof.
-  apply param_correctness_pos;  lazy;  parametric_tac.
-Qed.
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/F9Def}
+\inputsnippets{Euclidean_Chains/F9Ok}
 
@@
 
-\begin{Coqsrc}
- Fixpoint  Fexp2_of_nat (n:nat) : Fchain :=
- match n with O => F1
-            | S p => Fcompose F2 (Fexp2_of_nat p)
- end.
-
-Definition Fexp2 (p:positive) : Fchain :=
-  Fexp2_of_nat (Pos.to_nat p).
-
-Compute Fexp2  4.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- = fun (A : Type) (x : Fkont A) (x0 : A) =>
-       x1 <--- x0 times x0;
-       x2 <--- x1 times x1; x3 <--- x2 times x2; 
-       x4 <--- x3 times x3; x x4
-     : Fchain
-\end{Coqanswer}
-%%% ICI
+\inputsnippets{Euclidean_Chains/Fexp2}
 
@@
 
-\begin{Coqbad}
-Theorem F3_correct : Fchain_correct 3 F3.
-Proof.  
-  intros    A op E_one E_equiv M k  a ; simpl.
-  monoid_simpl M.  
-\end{Coqbad}
-
-\begin{Coqanswer}
-  A : Type
-  op : Mult_op A
-  E_one : A
-  E_equiv : Equiv A
-  M : EMonoid op E_one E_equiv
-  k : Fkont A
-  a : A
-  H : Proper (equiv ==> equiv ==> equiv) op
-  ============================
-   computation_execute op (k (a * a * a)) ==
-   computation_execute op (k (a * (a * (a * E_one))))
-\end{Coqanswer}
-
-\begin{Coqbad}
-Abort.  
-End Bad.
-\end{Coqbad}
+\inputsnippets{Euclidean_Chains/BadDefa}
+\inputsnippets{Euclidean_Chains/BadDefb}
 
@@
 
-\begin{Coqsrc}
-Class Fkont_proper
-      `(M : @EMonoid A op E_one E_equiv) (k: Fkont A )  :=
-  Fkont_proper_prf:
-    Proper (equiv ==> computation_equiv op E_equiv) k.
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/FkontProper}
 
@@
 
-\begin{Coqsrc}
-Definition Fchain_correct_nat (n:nat) (f : Fchain) :=
- forall A `(M : @EMonoid A op E_one E_equiv) k
-        (Hk :Fkont_proper M k)
-        (a : A) ,
- computation_execute op (f A k  a) ==
- computation_execute op (k  (a ^ n)).
-
-Definition Fchain_correct (p:positive) (f : Fchain) :=
- Fchain_correct_nat (Pos.to_nat p) f.
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/GoodFchainCorrect}
 
@@
 
-Let us show some manual correctness proofs for small f-chains:
-
-\begin{Coqsrc}
-Lemma F1_correct : Fchain_correct 1 F1.
-Proof.
-  intros until M ; intros k Hk a ; unfold F1; simpl.
-  apply Hk; monoid_simpl M; reflexivity.
-Qed.
-\end{Coqsrc}
+Let us show some manual correctness proofs of  small f-chains:
+
+\inputsnippets{Euclidean_Chains/F1Ok}
 
@@
 
-\begin{Coqsrc}
-Theorem F3_correct : Fchain_correct 3 F3.
-Proof. 
-  intros until M; intros k Hk a; simpl.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- A : Type
-  op : Mult_op A
-  E_one : A
-  E_equiv : Equiv A
-  M : EMonoid op E_one E_equiv
-  k : Fkont A
-  Hk : Fkont_proper M k
-  a : A
-  ============================
-   computation_execute op (k (a * a * a)) ==
-   computation_execute op (k (a * (a * (a * E_one))))}
-\end{Coqanswer}
-
-\begin{Coqsrc}
-apply Hk. 
-\end{Coqsrc}
-
-\begin{Coqanswer}
-...
-============================
-a * a * a == a * (a * (a * E_one))}
-
-\end{Coqanswer}
-
-\begin{Coqsrc}
-  monoid_simpl M; reflexivity.
-Qed.
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/F3Ok}
 
@@
 
-\begin{Coqsrc}
-Theorem F2_correct : Fchain_correct 2 F2.
-Proof. 
-  intros until M; intros k Hk a; simpl;
-  apply Hk;  monoid_simpl M;  reflexivity.
-Qed.
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/F2Ok}
 
@@
 $n$ and a correct f-chain for $p$.
-
 Let us try to certify this construction:
 
-\begin{Coqbad}
-Module Bad2.
-
-Lemma Fcompose_correct_attempt :
-  forall f1 f2 n1 n2, Fchain_correct n1 f1 ->
-                      Fchain_correct n2 f2 ->
-                      Fchain_correct (n1 * n2) (Fcompose f1 f2).
-
-(* Beginning of proof omitted *)
-\end{Coqbad}
-
-\begin{Coqanswer}
-  Hk : Fkont_proper M k
-  a, x, y : A
-  Hxy : x == y
-  ============================
-   computation_execute op (f2 A k x) == 
-   computation_execute op (f2 A k y) 
-\end{Coqanswer}
+\inputsnippets{Euclidean_Chains/Bad2a, Euclidean_Chains/Bad2b}
 
@@
 
-\begin{Coqbad}
-Abort.
-\end{Coqbad}
+\inputsnippets{Euclidean_Chains/Bad2c}
 
@@
 
-\begin{Coqbad}
-Module Bad3.
-
-Class Fchain_proper_ (fc : Fchain) := Fchain_proper_prf : 
- forall  `(M : @EMonoid A op E_one E_equiv) k  ,
-    Fkont_proper M k 
-    forall x y, x == y ->
-        @computation_equiv _ op E_equiv (fc A k x) (fc A k y).
-\end{Coqbad}
+\inputsnippets{Euclidean_Chains/Bad3}
 
@@
 
-\begin{Coqbad}
-Instance Fcompose_proper_ (f1 f2 : Fchain)
-                             (_ : Fchain_proper_simple f1)
-                             (_ : Fchain_proper_simple f2) :
- Fchain_proper_ (Fcompose f1 f2).
-Proof. 
- intros until M;intros k Hk x y Hxy; unfold Fcompose;cbn. 
- apply (H _ _ _ _ M); auto.
- intros u v Huv;apply (H0 _ _ _ _ M);auto.
-Qed.
-\end{Coqbad}
+\inputsnippets{Euclidean_Chains/Bad3b}
 
@@
 
-\begin{Coqbad}
-End Bad3.
-\end{Coqbad}
+\inputsnippets{Euclidean_Chains/Bad3c}
 
@@
 
-\begin{Coqsrc}
-Definition Fkont_equiv  `(M : @EMonoid A op E_one E_equiv)
- (k k': Fkont A )  := 
- forall x y : A, x == y ->
-                 computation_equiv op E_equiv  (k x)  (k' y).
-
-Class Fchain_proper (fc : Fchain) := Fchain_proper_prf : 
- forall  `(M : @EMonoid A op E_one E_equiv) k k' ,
-    Fkont_proper M k -> Fkont_proper M k' ->    
-    Fkont_equiv M k k' ->
-    forall x y, x == y ->
-        @computation_equiv _ op E_equiv
-                           (fc A k x)
-                           (fc A k' y).
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/correctProper}
 
@@
 
-\begin{Coqsrc}
-Instance F1_proper : Fchain_proper F1.
-Proof.
-  intros until M ; intros k k' Hk Hk' H a b H0; unfold F1; cbn;
-  now apply H.  
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Ltac add_op_proper M H := 
- let h := fresh H in
-   generalize (@Eop_proper _ _ _ _ M); intro h.
-
-Instance F3_proper : Fchain_proper F3.
-Proof.
-  intros  A op one equiv M  k k' Hk Hk'  Hkk' x y Hxy;  
-  apply Hkk'; add_op_proper M H; repeat rewrite Hxy;
-  reflexivity.
-Qed.
-\end{Coqsrc}
-
-We are now able to prove  \texttt{Fexp2 $n$}'s correctness by induction 
-on $n$:
-
-\begin{Coqsrc}
-Instance Fexp2_nat_proper (n:nat) : 
-                           Fchain_proper (Fexp2_of_nat n).
-Proof.
- induction n; cbn.
-   - apply F1_proper.
-   - apply Fcompose_proper ; [apply F2_proper | apply IHn].
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma  Fexp2_nat_correct (n:nat) : 
-         Fchain_correct_nat (2  ^ n) (Fexp2_of_nat n).
-Proof.
-  induction n; cbn.
- - apply F1_correct.
- -  rewrite Nat.add_0_r;
-   replace (2 ^ n + 2 ^ n)%nat with (2 * 2 ^n)%nat by  omega;
-   apply Fcompose_correct_nat;auto.
-   +  apply F2_correct.
-   +  apply  Fexp2_nat_proper.
-Qed.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Lemma  Fexp2_correct (p:positive) : 
-                           Fchain_correct (2  ^ p) (Fexp2 p).
-(* Proof omitted *)
-
-Instance  Fexp2_proper (p:positive) : Fchain_proper (Fexp2 p).
-(* Proof omitted *)
-
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/F1proper}
+\inputsnippets{Euclidean_Chains/F2proper}
+\inputsnippets{Euclidean_Chains/F3proper}
+
+\subsection{Correctness of chain composition}
+
+The \texttt{Fcompose} operator respects chain correctness and properness.
+
+\inputsnippets{Euclidean_Chains/FcomposeCorrect}
+\inputsnippets{Euclidean_Chains/FcomposeProper}
+
+Using chain composition, we get a correct and proper chain for any exponent of the form $2^n$.
+
+\inputsnippets{Euclidean_Chains/Fexp2Correct}
+\inputsnippets{Euclidean_Chains/Fexp2Proper}
 
@@
 
-\begin{Coqsrc}
-Hint Resolve F1_correct F1_proper
-     F3_correct F3_proper Fcompose_correct Fcompose_proper
-     Fexp2_correct Fexp2_proper .
-
-Example F144:  {f : Fchain | Fchain_correct 144 f /\
-                                Fchain_proper f}.
-Proof.
- change 144 with ( (3 * 3) * (2 ^ 4))%positive.
- exists (Fcompose (Fcompose F3 F3) (Fexp2 4)); auto.
-Defined.
-
-Compute proj1_sig F144.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-= fun (A : Type) (x : Fkont A) (x0 : A) =>
-       x1 <--- x0 times x0;
-       x2 <--- x1 times x0;
-       x3 <--- x2 times x2;
-       x4 <--- x3 times x2;
-       x5 <--- x4 times x4;
-       x6 <--- x5 times x5; 
-       x7 <--- x6 times x6; 
-       x8 <--- x7 times x7; 
-       x x8
-     : Fchain  
-\end{Coqanswer}
+\inputsnippets{Euclidean_Chains/F144}
 
@@
 
-\begin{Coqbad}
-Definition Fplus (f1 f2 : Fchain) : Fchain :=
-  fun A k x => 
-   f1 A (fun y => 
-            f2 A (fun z => t <--- z times y; k t) x) 
-            x.
-\end{Coqbad}
-
-For instance, we can define a chain for $23$:
-
-\begin{Coqbad}
-Let F23 := Fplus F3 (Fplus (Fexp2 4) (Fexp2 2)).  
-\end{Coqbad}
+\inputsnippets{Euclidean_Chains/Fplus}
 
@@
 
-\begin{Coqbad}
-Compute F23
-\end{Coqbad}
-
-\begin{Coqanswer}
- = fun (A : Type) (k : Fkont A) (x0 : A) =>
-       x1 <--- x0 times x0;
-       x2 <--- x1 times x0;
-       x3 <--- x0 times x0;
-       x4 <--- x3 times x3;
-       x5 <--- x4 times x4;
-       x6 <--- x5 times x5;
-       x7 <--- x0 times x0;
-       x8 <--- x7 times x7; 
-       x9 <--- x8 times x6; 
-       x10 <--- x9 times x2; 
-       k x10
-\end{Coqanswer}
+\pagebreak
+\inputsnippets{Euclidean_Chains/Fplusb}
 
@@
 
-\begin{Coqsrc}
-Definition Kkont A:=  A -> A -> @computation A.
-
-Definition Kchain :=  forall A, Kkont A -> A -> @computation A.
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/KchainDef}
 
@@
 
-\begin{Coqsrc}
-Example k3_1 : Kchain := fun A (k:Kkont A) (x:A) =>
-  x2 <--- x times x ;
-  x3 <--- x2 times x ;
-  k x3 x.
-
-Example k7_3 : Kchain := fun A (k:Kkont A)  (x:A) =>
-  x2 <--- x times x;
-  x3 <--- x2 times x ;
-  x6 <--- x3 times x3 ;
-  x7 <--- x6 times x ;
-  k  x7 x3.
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/K31}
+\inputsnippets{Euclidean_Chains/K73}
 
@@
 First, we have to adapt to k-chains our definitions of correctness and properness.
-\begin{Coqsrc}
-Definition Kkont_proper `(M : @EMonoid A op E_one E_equiv)
-           (k : Kkont A) :=
- Proper (equiv ==> equiv ==> computation_equiv op E_equiv) k .
-
-Definition Kkont_equiv  `(M : @EMonoid A op E_one E_equiv)
-           (k k': Kkont A )  := 
- forall x y : A, x == y -> forall z t, z == t -> 
-         computation_equiv op E_equiv   (k  x z) (k' y t).
- \end{Coqsrc}
+
+\inputsnippets{Euclidean_Chains/KkontDefs}
 
@@
 
- \begin{Coqsrc}
-Definition Kchain_correct_nat (n p : nat) (kc : Kchain) :=
-  forall `(M : @EMonoid A op E_one E_equiv)
-        (k : Kkont A), 
-        Kkont_proper M k ->
-        forall  (x : A) ,
-              computation_execute op (kc  A k  x) ==
-              computation_execute op (k  (x ^ n) (x ^ p)).
-
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Definition Kchain_correct (n p : positive) (kc : Kchain) :=
-  Kchain_correct_nat  (Pos.to_nat n) (Pos.to_nat p) kc.
-
-Class Kchain_proper (kc : Kchain) :=
-Kchain_proper_prf : 
- forall `(M : @EMonoid A op E_one E_equiv) k k' x y ,
-   Kkont_proper M k ->
-   Kkont_proper M k' -> 
-   Kkont_equiv M k k' ->
-   E_equiv x y ->
-   computation_equiv op E_equiv (kc A k x) (kc A k' y).
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/KchainCorrectDef}
 
@@
 
-\begin{Coqsrc}
-Instance k7_3_proper : Kchain_proper k7_3.
-Proof.
-  intros until M; intros; red; unfold k7_3; cbn;
-  add_op_proper M H3; apply H1;  rewrite H2;   reflexivity. 
-Qed.
-
-Lemma k7_3_correct : Kchain_correct 7 3 k7_3.
-Proof.
-  intros until M; intros; red; unfold k7_3; simpl.
-  apply H; monoid_simpl M;  reflexivity.
-Qed. 
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/K73Ok}
 
@@
 
-\begin{Coqsrc}
-Definition K2F (knp : Kchain) : Fchain :=
-  fun A (k:Fkont A) => kc A (fun  y _ => k y).
-
-Lemma K2F_correct :
-  forall knp n p, Kchain_correct kc n p ->
-                 Fchain_correct (K2F n) knp.
-(* Proof omitted *)
-
-Instance K2F_proper (kc : Kchain)(_ : Kchain_proper kc) :
-                                  Fchain_proper (K2F kc).
-
-(* Proof omitted} *)
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/K2FDef}
+\inputsnippets{Euclidean_Chains/K2FCorrect}
+\inputsnippets{Euclidean_Chains/K2FProper}
 
@@
 
-\begin{Coqsrc}
-Definition KFK (kbr : Kchain) (fq : Fchain) : Kchain  :=
-  fun A k a =>
-    kbr A  (fun xb xr =>
-              fq A (fun y =>
-                      z <--- y times xr; k z xb) xb) a.
-
-Lemma KFK_correct :
-  forall (b q r : positive) (kbr : Kchain) (fq : Fchain),
-    Kchain_correct b r  kbr ->
-    Fchain_correct q fq ->
-    Kchain_proper kbr ->
-    Fchain_proper fq -> 
-    Kchain_correct  (b * q + r) b (KFK kbr fq).
-(* Proof omitted *)
-
-Instance KFK_proper :
-forall (kbr : Kchain) (fq : Fchain),
-  Kchain_proper kbr -> 
-  Fchain_proper fq -> 
-  Kchain_proper (KFK kbr fq)
-(* Proof omitted *)
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/KFKDef}
+\inputsnippets{Euclidean_Chains/KFKCorrect}
+\inputsnippets{Euclidean_Chains/KFKProper}
 
@@
 
-\begin{Coqsrc}
-Definition KFF (kbr : Kchain) (fq : Fchain) : Fchain :=
-  K2F (KFK kbr fq).
-
-Lemma KFF_correct :
-forall (b q r : positive) (kbr : Kchain) (fq : Fchain),
-Kchain_correct  b r kbr ->
-Fchain_correct q fq ->
-Kchain_proper kbr ->
-Fchain_proper fq -> Fchain_correct (b * q + r) (KFF kbr fq).
-(* Proof omitted *)
-
-Instance KFF_proper :
-forall (kbr : Kchain) (fq : Fchain),
-Kchain_proper kbr -> Fchain_proper fq -> Fchain_proper (KFF kbr fq).
-(* Proof omitted *)
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/KFFDef}
+\inputsnippets{Euclidean_Chains/KFFCorrect}
+\inputsnippets{Euclidean_Chains/KFFProper}
 
@@
 
-\begin{Coqsrc}
-
-Definition FK (f : Fchain) : Kchain :=
-  fun (A : Type) (k : Kkont A) (a : A) =>
-    f A (fun y => k y a) a.
-
-Lemma FK_correct : forall (p: positive) (Fp : Fchain),
-                     Fchain_correct  p Fp ->
-                     Fchain_proper Fp ->
-                     Kchain_correct p 1 (FK Fp) .
-(* Proof omitted *)
-
-Instance  FK_proper  (Fp : Fchain) (_ : Fchain_proper Fp):
-  Kchain_proper (FK Fp).
-(* Proof omitted *)
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/FKDef}
+
+Like our other combinators, \texttt{FK} respects chain correctness and properness.
 
@@
 
-\begin{Coqsrc}
-Definition FFK (fp fq : Fchain) : Kchain :=
-  fun A k a =>  fp A (fun xb  => fq A (fun y => k y xb) xb) a.
-
-Lemma FFK_correct  (p q  : positive) (fp fq : Fchain):
-    Fchain_correct p fp  ->
-    Fchain_correct q fq ->
-    Fchain_proper fp ->
-    Fchain_proper fq -> Kchain_correct  (p * q ) p (FFK fp fq).
-(* Proof omitted *)
-
-Instance FFK_proper  (fp fq : Fchain)
-    (_ :   Fchain_proper fp)
-    (_ :  Fchain_proper fq) :  Kchain_proper (FFK fp fq) .
-(* Proof omitted *)
-\end{Coqsrc}
+Our last combinator composes a chain for computing $x^p$ with a chain for computing $x^q$ to build a chain for
+computing $x^p$ and $x^{pq}$.
+
+\inputsnippets{Euclidean_Chains/FFKDef}
+\inputsnippets{Euclidean_Chains/FFKCorrect}
+\inputsnippets{Euclidean_Chains/FFKProper}
 
@@
 
-\begin{Coqsrc}
-Hint Resolve KFF_correct KFF_proper KFK_correct KFK_proper.
-
-Definition F87 :=
- let k7_3 :=  KFK k3_1 (Fexp2 1) in
- let k10_7 := KFK k7_3 F1 in
- KFF k10_7 (Fexp2 3).
-
-Lemma OK87 : Fchain_correct 87 F87.
-Proof.
- unfold F87; change 87 with (10 * (2 ^ 3) + 7)%positive.
- apply KFF_correct;auto.
- change 10 with (7 * 1 + 3); apply KFK_correct;auto.
- change 7 with (3 * 2 ^ 1 + 1)%positive;  apply KFK_correct;auto.
-Qed.
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/HintKchains}
+\inputsnippets{Euclidean_Chains/F87}
+\inputsnippets{Euclidean_Chains/F87Correct}
 
@@
 f-chains, associated with a single exponent, and k-chains, associated with two exponents.
-We would expect that the function \texttt{make\_chain} we want to build and certify is structured as a pair of mutually recursive functions.
+We would expect that the function \texttt{make\_chain} we want to define and certify is structured as a pair of mutually recursive functions.
  In \coq{} , various ways of building such functions are available:
@@
 
-\begin{Coqsrc}
-Inductive signature : Type :=
-|  (** Fchain for the exponent n *)
-     gen_F (n:positive) 
-| (** Kchain for the exponents p+d  and p *) 
-     gen_K (p d: positive).
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/signature}
 
@@
 
-\begin{Coqsrc}
-(** 
-  exponent associated with a signature:
-*)
-Definition signature_exponent (s:signature) : positive :=
- match s with 
-| gen_F n => n 
-| gen_K p d  =>  p + d
-end.
-\end{Coqsrc}
-
-\begin{Coqsrc}
-(**
-Type of the associated continuation 
-*)
-
-Definition kont_type (s: signature)(A:Type) : Type :=
-match s with 
-| gen_F _  => Fkont A 
-| gen_K _ _   => Kkont A
-end.
-
-Definition chain_type (s: signature) : Type :=
- match s with 
-| gen_F _   => Fchain
-|  gen_K _ _  => Kchain
-end.
-
-\end{Coqsrc}
-
-\begin{Coqsrc}
-Definition correctness_statement (s: signature) : 
-chain_type s -> Prop :=
-match s  with
-  | gen_F p => fun ch => Fchain_correct p ch
-  | gen_K p d   => fun ch => Kchain_correct (p + d) p ch
-end.
-
-Definition proper_statement (s: signature) : 
-chain_type s -> Prop :=
-match s  with
-  | gen_F p => fun ch => Fchain_proper ch 
-  | gen_K p d   => fun ch => Kchain_proper ch 
-end.
-
-(**  Full correctness *)
-
-Definition  OK (s: signature) 
-  := fun c: chain_type s => 
-       correctness_statement s c /\
-       proper_statement s c.
-
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/dependentlyTypedFuns}
 
@@
 \noindent
-From ~\href{../theories/html/additions.Dichotomy.html}{additions.Dichotomy}.
-\begin{Coqsrc}
-Class Strategy (gamma : positive -> positive):=
-{
-gamma_lt :forall p:positive, 3 < p -> gamma  p < p;
-gamma_gt : forall p:positive, 3 < p -> 1 < gamma  p
-}.
-
-\end{Coqsrc}
+From ~\href{../theories/html/additions.Strategies.html}{additions.Strategies}.
+
+\inputsnippets{Strategies/StrategyDef}
+
 \subsection{The dichotomic strategy}
@@
 
-\begin{Coqsrc}
-Function dicho_aux (p:positive) {struct p} : positive :=
- match p with
-   | 1%positive   =>  xH
-   | 2%positive |   3%positive  => 2 
-   | xO (xO q) | xO (xI q) | xI (xO q) | xI (xI q) =>
-                                         xO (dicho_aux q)
- end.
-
-Definition dicho  (p:positive) : positive :=
-  N2pos (N.div (Npos p) (Npos (dicho_aux p))).
-
-Instance Dicho_strat : Strategy dicho.
-\end{Coqsrc}
+\inputsnippets{Dichotomy/dichotomy}
+\inputsnippets{Dichotomy/DichoStrat}
 
@@
 
-\begin{Coqsrc}
-Definition half (p:positive) :=
-  match p with xH => xH
-          |    xI q | xO q =>  q
-  end.
-
-Definition two (p:positive) := 2%positive.
-
-Instance Binary_strat : Strategy half.
-Proof.
-  split; destruct p; unfold half; try lia.
-Qed.
-
-Instance Two_strat : Strategy two.
-Proof.
-  split;unfold two; lia.
-Qed.
-\end{Coqsrc}
+\inputsnippets{BinaryStrat/BinaryStrats}
 
@@
 
-\begin{Coqsrc}
-Definition signature_measure (s : signature) : nat :=
-match s with
-  | gen_F n => 2 * Pos.to_nat n 
-  | gen_K p d => 2 * Pos.to_nat (p + d) +1
-end.
-\end{Coqsrc}
-
-The following function definition generates 9 subgoals,
+\inputsnippets{Euclidean_Chains/GammaContext}
+
+The following function definition generates 9 proof obligations subgoals,
 for proving that the measure on signatures is strictly decreasing along
@@
 
-%\todo{Is Function (from Recdef) obsolete ? Use Equations instead ? }
-
-\begin{Coqsrc}
-Function chain_gen  (s:signature) {measure signature_measure}
-:  chain_type s :=
-  match s  return chain_type s with
-    | gen_F i =>
-      if pos_eq_dec i 1 then F1 else
-        if pos_eq_dec i 3
-        then F3
-        else 
-          match exact_log2 i with
-              Some p => Fexp2 p
-            | _ =>
-              match N.pos_div_eucl i (Npos (dicho i))
-              with
-                | (q, 0%N) => 
-                  Fcompose  (chain_gen (gen_F (dicho i)))
-                            (chain_gen (gen_F (N2Pos q)))
-                | (q,r)  => KFF (chain_gen
-                                   (gen_K (N2Pos r)
-                                          (dicho i - N2Pos r)))
-                                (chain_gen (gen_F (N2Pos q)))
-
-              end end
-\end{Coqsrc}
-
-\begin{Coqsrc}
-    | gen_K p d =>
-      if pos_eq_dec p 1 then FK (chain_gen (gen_F (1 + d)))
-      else
-        match N.pos_div_eucl (p + d)  (Npos p) with
-          | (q, 0%N) => FFK   (chain_gen (gen_F p))
-                              (chain_gen (gen_F (N2Pos q)))
-          | (q,r)  => KFK (chain_gen (gen_K (N2Pos r)
-                                            (p - N2Pos r)))
-                          (chain_gen (gen_F (N2Pos q)))
-        end
-  end.
-(* A lot of arithmetic proofs omitted *)
-Defined.
-
-Definition make_chain (n:positive) : chain :=
- F2C (chain_gen (gen_F n)).
-
-\end{Coqsrc}
-
-Thanks to the \texttt{Recdef} package, we are now able to get automatically 
-built chains using the dichotomic strategy.
-
-\begin{Coqsrc}
-Compute make_chain 87.
-\end{Coqsrc}
-
-\begin{Coqanswer}
- =  fun (A : Type) (x : A) =>
-       x0 <--- x times x;
-       x1 <--- x0 times x;
-       x2 <--- x1 times x1;
-       x3 <--- x2 times x;
-       x4 <--- x3 times x1;
-       x5 <--- x4 times x4;
-       x6 <--- x5 times x5;
-       x7 <--- x6 times x6; 
-       x8 <--- x7 times x3; 
-       Return x8
-     : chain 
-\end{Coqanswer}
+\pagebreak
+\inputsnippets{Euclidean_Chains/chainGen}
+\inputsnippets{Euclidean_Chains/makeChain}
+\inputsnippets{Euclidean_Chains/C87Dicho}
 
@@
 
-The following tests show various examples of chains for the same exponent, using different strategies. The dichotomic strategy seems clearly to be the winner (at least on this sample).
+The following tests show various examples of chains for the same exponent, using different strategies.
+The dichotomic strategy seems clearly to be the winner (at least on this sample)\footnote{For efficiency's sake, we commented out some (very) long computations.
+You may uncomment them freely in your own copy.
+For the same reason, we put a verbatim trace instead of \textit{Alectryon} output}.
 
@@
 
-\begin{Coqsrc}
-Lemma chain_gen_OK : forall s:signature, OK  s (chain_gen  s).
-intro s; functional induction chain_gen s.
-Proof.
-(* A lot of arithmetic proofs omitted *)
-
-Theorem make_chain_correct :
-    forall p, chain_correct p (make_chain p).
-Proof.
- intro p; destruct (chain_gen_OK (gen_F p)).
- unfold make_chain; apply F2C_correct; apply H.
-Qed.
-\end{Coqsrc}
+\inputsnippets{Euclidean_Chains/chainGenOK}
 
@@
 
-\begin{Coqsrc}
-(** basic instructions *)
-
-Inductive instr : Set :=
-  | MUL : instr
-  | SQR : instr
-  | PUSH : instr
-  | SWAP : instr.
-
-Definition code := list instr.
-
-(* semantics *)
-(*************)
-
-Section Semantics.
-
- Variable A : Type.
- Variable mul : A -> A -> A.
- Variable one : A.
-
- Definition stack := list A.
- Definition config := (A * list A)%type.
-
- Fixpoint exec (c : code) (x:A) (s: stack) : option config :=
-   match c, s with
-     nil, _ => Some (x,s)
-   | MUL::c, y::s => exec c (mul x y) s
-   | SQR::c, s => exec c (mul x x) s
-   | PUSH::c, s => exec c x (x::s)
-   | SWAP::c, y::z::s => exec c x (z::y::s)
-   | _,_ => None
-   end.
-
-(* ... *)
-End Semantics.
-\end{Coqsrc}
+\inputsnippets{AM/AMDef}
+\inputsnippets{AM/AMSem}
+\inputsnippets{AM/AMSemb}
 
@@
 
-\begin{Coqsrc}
-  PUSH :: PUSH :: SQR  :: MUL :: PUSH :: SWAP :: SQR :: MUL :: PUSH
-            :: SWAP :: MUL :: SQR :: SQR :: SQR :: MUL :: nil 
-\end{Coqsrc}
+\inputsnippets{AM/F87}
 

### 2021-09-15 doc/chapter-powers.tex
corrections Chapter 10

@@
 Several examples will be presented in the rest of this document
-(in sections~\vref{sect:fibonacci-pos-bpow}).
+(in Section~\vref{sect:fibonacci-pos-bpow}).
 
@@
 
-We are now ready to define a new notation scope, in which the notation
+We are now ready to define a new notation scope, in which the notation \linebreak 
 \texttt{x * y} will be interpreted as an application of the function
@@
 and consider ``equivalence modulo \texttt{$m$}'' instead of equality.
-Note that the law of associativity has been stated as Leibniz' equality.
 
@@
 Therefore, we note that the term (\Verb|fun (x:A) =>x ^b 17|)  is
-convertible, --- \emph{thus logically indistinguishable} ---, with a function that performs 16 multiplications.
+convertible --- \emph{thus logically indistinguishable} --- with a function that performs 16 multiplications.
 
@@
 
-From these tests, we may infer that  representing exponentiation algorithms as \coq{}  functions hides
+From these tests, we may infer that  representing exponentiation algorithms as plain arithmetic  functions hides
 information about the real structure of the computations, particularly about  sharing intermediate computations.
@@
 twice in the  computation of $x^{17}$ with the binary algorithm. This information should 
-appear explicitly in the data structure chosen for representing exponentiation 
+appear explicitly in the data structure chosen for representing
+and comparing exponentiation 
 algorithms.
@@
 \begin{quote}
-For any item $i$ of $c$, there exists $j$ and $k$ in $c$, where
+For any item $i$ of $c$ (but the first one), there exists $j$ and $k$ in $c$, where
 $i=j+k$, and $x^j$ and $x^k$ are already computed.
@@
 subgoal of the previous proof generated after a sequence of unfoldings and simplifications. This goal is presented below.
-
+\begin{figure}[h]
+  \centering
 \inputsnippets{Addition_Chains/WhySoSlow}  
+  \caption{A big goal}
+  \label{fig:big-goal}
+\end{figure}
 
@@
 \subsubsection{How does reflection work}
-Let us consider again the subgoal on page~\pageref{fig:big-goal}, the conclusion of which has the form \texttt{|$a_1\,==\,a_2$|}, where \texttt{|$a_1$|} and
+Let us consider again the subgoal on Fig.~\vref{fig:big-goal}, the conclusion of which has the form \texttt{|$a_1\,==\,a_2$|}, where \texttt{|$a_1$|} and
 \texttt{|$a_2$|} are terms of  type \texttt{A}.
@@
 
+\pagebreak
 \inputsnippets{Addition_Chains/MonoidExp}
@@
 \begin{enumerate}
-\item generate a subgoal as in page~\pageref{fig:big-goal},
+\item generate a subgoal as in Fig.~\vref{fig:big-goal},
 \item express each term of the equivalence as the image of a term
@@
 with executions on \texttt{NatPlus}.
-Let us consider the following mathematical relation 
-$$\{(x,n)\in M\times\mathbb{N}\,|\, 0<n \wedge x=a^n\}$$
+Let us consider the  mathematical relation 
+$\{(x,n)\in M\times\mathbb{N}\,|\, 0<n \wedge x=a^n\}$.
 
@@
 \texttt{(a * a * a) == (a * (a * (a * E\_one)))} \linebreak that 
-\texttt{k (a * a * a)} and \texttt{k (a * (a * (a * E\_one)))} are 
+(\texttt{k (a * a * a)}) and (\texttt{k (a * (a * (a * E\_one)))}) are 
 equivalent computations.
@@
 A continuation \texttt{k:Fkont A} is \emph{proper}
-if, whenever \linebreak[3] \texttt{x == y} holds, the computations \texttt{k x} and 
-\texttt{k y} are equivalent.
+if, whenever \linebreak[3] \texttt{x == y} holds, the computations (\texttt{k x}) and 
+(\texttt{k y}) are equivalent.
 
@@
 
-Let us show some manual correctness proofs of  small f-chains:
+Let us show manual correctness proofs of some small f-chains:
 
@@
 
+\pagebreak
 \inputsnippets{Euclidean_Chains/Fplus}
@@
 
-\pagebreak
 \inputsnippets{Euclidean_Chains/Fplusb}
@@
 
-\pagebreak
 \inputsnippets{Euclidean_Chains/chainGen}
 \inputsnippets{Euclidean_Chains/makeChain}
+\inputsnippets{Euclidean_Chains/makeChainCorrect}
 \inputsnippets{Euclidean_Chains/C87Dicho}

### 2021-09-15 doc/JFLA2022/paper.tex
small modifications

@@
 \label{sect:introduction}
-Hydra games (a.k.a. \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians:
-L. Kirby and J. Paris~\cite{KP82}: \emph{Accessible Independence Results for Peano Arithmetic}.
-This article presents a gameful interpretation of accessibility problem in the set of ordinals below $\epsilon_0$, and proves two theorems:
+Hydra games (also know as \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians, 
+Laurie Kirby and Jeff Paris~\cite{KP82}: \emph{Accessible Independence Results for Peano Arithmetic}.
+This article describes a game between two players: Hercules and a hydra, and 
+and proves two theorems:
 \begin{quote}
@@
 
-Like Goodstein sequences, hydra battles may be considered as a simplified version of G\"{o}del-like theorems.
+Indeed, like Goodstein sequences~\cite{goodstein_1944, Sladek07, KP82}, hydra battles are just 
+ a gameful presentation of accessibility problems in the set of ordinals below $\epsilon_0$.
+
 Although the mathematical contents of this 
@@
 
+\section{Main features}
+\subsection{Multiple views}
+
+Whenever possible, we try to present several implementations of the same object, several proofs of the same statement.
+Let us take an example. Ordinal numbers (strictly below $\Gamma_0$) are represented with the help of \emph{ordinal notations}, \emph{i.e.} data types:
+\begin{itemize}
+\item natural numbers (ordinals below $\omega$),
+\item list of natural numbers (ordinals below $\omega^\omega$),
+\item ternary trees (ordinals below $\epsilon_0$ in Cantor normal form),
+\item quaternary trees (ordinals below $\Gamma_0$ in Veblen normal form).
+\end{itemize}
+Comparison and arithmetic operations are implemented by
+structurally recursive functions.
+Figure~\ref{fig:Ex42E0} shows a trivial (computational) proof of the equality \ref{eq:1}.
+
+\begin{eqnarray}
+  \label{eq:1}
+  \omega+42+\omega^2&=&\omega^2
+\end{eqnarray}
+
+We also implemented in \emph{Coq} the axiomatic definition of the set $\mathbb{O}$ of countable ordinals by Schütte~\cite{schutte, PCiota}.
+
+The proof of $\omega+42+\omega^2=\omega^2$ is no more computational, but use the following definitions and properties.
+
+\begin{itemize}
+\item Any set of countable ordinals can be enumerated by a function defined on a segment $[0, \alpha)$ of countable ordinals.
+\item Let $\alpha$ and $\beta$ be (countable) ordinals. Then $\alpha+\beta$ is the $\beta$-th element of the enumeration of the set of countable ordinals greater or equal than $\alpha$.
+\item Let $\phi_0$ be the enumeration of the set of \emph{additive principal ordinals}
+  $\{\alpha\in\mathbb{O}| \forall \beta\in \mathbb{O}, \beta+\alpha=\alpha\}$. Then $\omega^\alpha$ is just a notation for
+  $\phi_0(\alpha)$.
+\end{itemize}
+
+By definition, $\omega^2=\phi_0(2)$ is additive principal, thus
+satisfies the equality. Figure~\ref{fig:Ex42b} shows the mians steps of the proof.
+
 \begin{todo}
@@
   \begin{itemize}
-  \item Ordinals notations: for $\omega$, $\omega^\omega$,
-    $\epsilon_0$, $\Gamma_0$. Proofs by computation whenever possible (see Fig.~\ref{Ex42-E0}.
-  \item (Axiomatic) model by Schütte. Less computations.
-    application of previously proven lemmas (see Fig~\ref{fig:Ex42b}).
+
   \item Bridge lemmas (relative correctness of different representations).
@@
   \end{figure}
-\label{Ex42-E0}.
 

### 2021-09-16 doc/JFLA2022/paper.tex
draft paper

--- 2021-09-13 doc/JFLA2022/paper.tex
@@
 \usepackage{doc}
+\usepackage{tikz}
+\usepackage{tikzsymbols}
+\usetikzlibrary{arrows}
 
@@
 \newcommand{\canonseq}[2]{\mbox{$\{#1\}(#2)$}}
+
 \newtheorem{todo}{To do}
@@
 \label{sect:introduction}
-Hydra games (a.k.a. \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians:
-L. Kirby and J. Paris~\cite{KP82}: \emph{Accessible Independence Results for Peano Arithmetic}.
-This article presents a gameful interpretation of accessibility problem in the set of ordinals below $\epsilon_0$, and proves two theorems:
+Hydra games (also known as \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians, 
+Laurie Kirby and Jeff Paris~\cite{KP82}: \emph{Accessible Independence Results for Peano Arithmetic}.
+This article describes a game between two players: Hercules and a hydra.
+A short description  can be found in Andrej Bauer's blogpost~\cite{bauer2008} as well as a simulator~\cite{BauerHydra}.
+
+In a few words:
+\begin{itemize}
+\item A hydra is a finite tree, traditionnaly presented with the root at the bottom, the leaves of which are called \emph{heads}.
+\item At every round, Hercules chops off one head of the hydra. If the head is at a distance greater or equal than 2 from the root,
+  then some sub-tree $h$ of the hydra is copied a certain amount $i$ of times. The number $i$ of copies and the copied sub-hydra $h$ may depend of the considered variant of the game
+  or the time elapsed since the beginning of the fight.
+\end{itemize}
+
+Kirby and Paris prove the following theorems, applying
+combinatorial results about ordinal numbers by Jussi Ketonen and Robert Solovay~\cite{KS81}.
+
 \begin{quote}
@@
 
-Like Goodstein sequences, hydra battles may be considered as a simplified version of G\"{o}del-like theorems.
-Although the mathematical contents of this 
-paper are quite advanced, the rules of hydra battles are very easy to understand.
-There are now several sites on the Internet where you can find tutorials on hydra games, together with simulators you can play with.
-See, for instance, the blogpost and source code written by Andrej Bauer~\cite{bauer2008,BauerHydra}.
+The contrast between the simplicity of the statements above and the complexity of the mathematical tools used to prove these theorems: ordinal numbers, rapidly growing hierarchies of fast growing functions, etc., convinced us that it is a good thema for a commented library of formal proofs.
 
-Our project is to illustrate  this work by a library written in \emph{Coq}~\cite{HydraBattles} and an electronic book~\cite{HydraBook}  containing explanations of the mathematical content and its possible mechanized formalisations.
-This project has already been presented in
-JFLA2018~\cite{JFLA2018paper}.
-We present the recent evolution of the library: new results, interaction within the Coq-Community project~\cite{CoqCommunity}, and documentation generated with Alectryon~\cite{alectryonpaper, alectryongithub}.
+The benefits we would expect from this development are multiple.
+\begin{itemize}
+\item We present complete proof scripts which can be consulted and replayed by the reader.
+\item Complex formalisations and proofs are explained in an
+  electronic book~\cite{HydraBook}.
+Whenever various reasonable choices exist, we try to present and compare the alternatives.
+Reciprocally, we provide ``concrete'' examples of several formaisation end proof techniques: operational type classes, functions defined by  equations, etc.
+\item It may be used as a library on ordinal numbers, for instance for proving termination properties.
+  \end{itemize}
+
+ Prior stages of this project have already been presented at
+ JFLA~\cite{JFLA2018paper}.
+We present the recent evolutions of the library: new results, interaction within the Coq-Community project~\cite{CoqCommunity}, and documentation generated with Alectryon~\cite{alectryonpaper, alectryongithub}.
+
+\section{New results}
+In order to make evident how long a hydra battle can be,
+one may consider a deterministic variant of the game,
+``compute'' the number of steps of the battles, and thus obtain a minoration of the general case.
+
+The variant we consider is the same as 
+in~\cite{KP82, bauer2008, BauerHydra}.
+
+ \begin{itemize}
+ \item The game starts at an arbitrary initial step $i\in\mathbb{N}$.
+
+   \item The hydra grows $n$ copies at step $n$ of the game.
+ \item  The initial hydra is the graphical representation of an ordinal  $\alpha<\epsilon_0$  in Cantor normal form (See for instance Fig~\ref{fig:iota-example}). 
+ \item Hercules always chops off the rightmost head of the hydra (which is always one of the lowest heads).
+ \end{itemize}
+ It mathematical terms, if at step $n$ the hydra is associated with the ordinal $\alpha$, at step $n+1$ it is associated with
+ $\canonseq{\alpha}{n+1}$, the $(n+1)$th element of the canonical sequence of $\alpha$~\cite{KS81}.
+
+   \begin{figure}[htb]
+\centering
+\begin{tikzpicture}[very thick, scale=0.3]
+\node (foot) at (10,0) {$\bullet$};
+\node (N1) at (2,2) {$\bullet$};
+\node (N2) at (10,2) {$\bullet$};
+\node (N22) at (7,2) {$\bullet$};
+\node (N3) at (14,2) {$\bullet$};
+\node (N4) at (18,2) {$\Smiley[2][green]$};
+\node (N5) at (0,4) {$\bullet$};
+\node (N6) at (2,5) {$\Smiley[2][green]$};
+\node (N7) at (4,6) {$\Smiley[2][green]$};
+\node (N88) at (7,4) {$\bullet$};
+\node (N8) at (10,4) {$\bullet$};
+\node (N9) at (14,6) {$\Smiley[2][green]$};
+\node (N10) at (0,8) {$\Smiley[2][green]$};
+\node (N11) at (10,7) {$\Smiley[2][green]$};
+\node (N111) at (7,7) {$\Smiley[2][green]$};
+\draw (foot) to [bend left=10] (N1);
+\draw (foot) -- (N2);
+\draw (foot) -- (N22);
+\draw (foot) -- (N3);
+\draw (foot) -- (N4);
+\draw (N1) to  (N5);
+\draw (N1) to   [bend left=10] (N6);
+\draw (N1) to   [bend right=20] (N7);
+\draw (N2) to  (N8);
+\draw (N22) to  (N88);
+\draw (N8) to  (N11);
+\draw (N88) to  (N111);
+\draw (N3) to  (N9);
+\draw (N5) to  (N10);
+\end{tikzpicture}
+\caption{The hydra associated with $\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$ \label{fig:iota-example}}
+
+\end{figure}
+
+Let $\alpha<\epsilon_0$ be an ordinal.
+Using the \texttt{coq-equations} plug-in~\cite{sozeau:hal-01671777}, we define the function which returns the number of steps of the battle starting at configuration $(i,\alpha)$ for  any $i\in\mathbb{N}$.
+We prove that this number is greater or equal than
+$H'_\alpha(i)-i$, where $H'$ is a slight variant of the Wainer Hardy hierarchy of rapidly growing functions~\cite{BW85, KS81, Promel2013, Wainer1970}, defined in Fig.~\ref{fig:Hprime} and
+\ref{fig:HprimeCoq}.
+
+\begin{figure}[h]
+  \centering
+  \begin{itemize}
+\item If $\alpha=0$, then $H'_\alpha (k)= k$ for any natural number $k$.
+\item If $\alpha=\beta+1$, then 
+$H'_\alpha(k)=H'_\beta(k+1)$ for any $k \in \mathbb{N}$
+\item If $\alpha$ is a limit ordinal, then 
+$H'_\alpha(k) = H'_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
+\end{itemize}
+\caption{The rapidly growing hierarchy $H'$}
+\label{Hprime}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \inputsnippets{Hprime/HprimeDef}
+  \caption{$H'$: definition in \emph{Coq}}
+  \label{fig:HprimeCoq}
+\end{figure}
+
+From this result, we infer that, for a given $\alpha$, the considered function is not primitive recursive.
+
+\section{Main features}
+
+% \begin{eqnarray}
+%   \label{eq:1}
+%   \omega+42+\omega^2&=&\omega^2
+% \end{eqnarray}
+
+% We also implemented in \emph{Coq} the axiomatic definition of the set $\mathbb{O}$ of countable ordinals by Schütte~\cite{schutte, PCiota}.
+
+% The proof of $\omega+42+\omega^2=\omega^2$ is no more computational, but use the following definitions and properties.
+
+% \begin{itemize}
+% \item Any set of countable ordinals can be enumerated by a function defined on a segment $[0, \alpha)$ of countable ordinals.
+% \item Let $\alpha$ and $\beta$ be (countable) ordinals. Then $\alpha+\beta$ is the $\beta$-th element of the enumeration of the set of countable ordinals greater or equal than $\alpha$.
+% \item Let $\phi_0$ be the enumeration of the set of \emph{additive principal ordinals}
+%   $\{\alpha\in\mathbb{O}| \forall \beta\in \mathbb{O}, \beta+\alpha=\alpha\}$. Then $\omega^\alpha$ is just a notation for
+%   $\phi_0(\alpha)$.
+% \end{itemize}
 
@@
   \begin{itemize}
-  \item Ordinals notations: for $\omega$, $\omega^\omega$,
-    $\epsilon_0$, $\Gamma_0$. Proofs by computation whenever possible (see Fig.~\ref{Ex42-E0}.
-  \item (Axiomatic) model by Schütte. Less computations.
-    application of previously proven lemmas (see Fig~\ref{fig:Ex42b}).
+
   \item Bridge lemmas (relative correctness of different representations).
     \item The user/reader may compare  different proofs of the same result (For instance the equality $\omega+42+\omega^2= \omega^2$).
+\emph{Example in Figures~\ref{fig:Ex42E0} and \ref{fig:Ex42-schutte}}.
   \end{itemize}
 \end{todo}
-
-\section{New contents}
-We are interested in a usual variant of hydra-battles, where the hydra grows $n$ copies at step $n$ of the game (see the examples presented in ~\cite{KP82, BauerHydra}).
-In terms of ordinals, this variant is described as a (finite) sequence $s_n\,(n\geq i)$ of ordinals strictly below $\epsilon_0$.
-This sequence is parameterized by an initial index $i\in\mathbb{N}$ and an ordinal $\alpha<\epsilon_0$.
-
-\begin{itemize}
-  \item $s_i=\alpha$
-\item If $s_n=0$, then the sequence is empty,
-\item if $s_n=\beta+1$, then $s_{n+1}=\beta$,
-\item Otherwise, $s_n$ is a \emph{limit ordinal} $\lambda$; then $s_{n+1}$ is $\canonseq{\lambda}{n+1}$, \emph{i.e.} the
-    $n+1$-th element of the \emph{canonical sequence} associated with $\lambda$~\cite{KS81}.
-\end{itemize}
-
-The length of such a sequence is $l-i$, where $l$ is the first --- and unique --- natural number such that $s_l=0$.
-For a given $\alpha$, this length is a function of 
-the initial index $i$.
-
-Adapting to \emph{Coq} a few lemmas from~\cite{KS81}, we prove that the length  of the sequence associated with
-$\omega^\alpha$ and initial index $i\geq 2$ is greater or equal than
-$F_{\omega^\alpha}(i-1)-(i-1)$, where $F_\alpha$ is a variant of the Wainer hierarchy of rapidly growing functions, defined and studied in~\cite{KS81}.
-
-From this minoration, we infer that the function which associates to $i$ the length of the considered sequence (for a given $\alpha$) is not primitive recursive in general (our counter-example is for $\alpha=\omega^\omega$).
-
-It is important to notice that, for proving this last result, we decided to host and maintain within \emph{Coq-Community} the formalization of primitive recursive function, part of Russel O'Connor's work on G\"{o}del's first incompleteness theorem~\cite{OConnor05}.
-
-Defining and studying families of rapidly growing arithmetic  functions indexed by ordinals below $\epsilon_0$ has been made possible by the plug-in \emph{coq-equations}~\cite{sozeau:hal-01671777}.
 
@@
 
-\label{Ex42-schutte}
+\label{fig:Ex42-schutte}
 \fbox{\begin{minipage}[h]{1.0\linewidth}

### 2021-09-17 doc/JFLA2022/paper.tex
minor corrections

@@
 
+\usepackage{varioref}
 \newtheorem{todo}{To do}
@@
 \begin{abstract}
-  \begin{todo}
- \dots
-  \end{todo}
+  We present the current state of a \textit{Coq} library on ordinal numbers.
+
 \end{abstract}
@@
 A short description  can be found in Andrej Bauer's blogpost~\cite{bauer2008} as well as a simulator~\cite{BauerHydra}.
-
 In a few words:
 \begin{itemize}
-\item A hydra is a finite tree, traditionnaly presented with the root at the bottom, the leaves of which are called \emph{heads}.
+\item A hydra is a finite tree, traditionnaly presented with the root at the bottom, the leaves of which are called \emph{heads}
+  (Fig~\vref{fig:iota-example}).
 \item At every round, Hercules chops off one head of the hydra. If the head is at a distance greater or equal than 2 from the root,
@@
 
-The contrast between the simplicity of the statements above and the complexity of the mathematical tools used to prove these theorems: ordinal numbers, rapidly growing hierarchies of fast growing functions, etc., convinced us that it is a good thema for a commented library of formal proofs.
+The contrast between the simplicity of the statements above and the complexity of their proofs convinced us that it is a good thema for a commented library of formal proofs.
 
-The benefits we would expect from this development are multiple.
 \begin{itemize}
-\item We present complete proof scripts which can be consulted and replayed by the reader.
+\item We present complete proof scripts for the \textit{Coq} proof assistant~\cite{Coq}. These scripts may consulted and replayed by the reader.
 \item Complex formalisations and proofs are explained in an
-  electronic book~\cite{HydraBook}.
-Whenever various reasonable choices exist, we try to present and compare the alternatives.
-Reciprocally, we provide ``concrete'' examples of several formaisation end proof techniques: operational type classes, functions defined by  equations, etc.
+  electronic book~\cite{HydraBook}. Whenever various reasonable choices exist, we try to present and compare the alternatives.
+  For instance, Figures~\vref{fig:Ex42E0} and \vref{fig:Ex42-schutte} show two radically different proofs of the equality
+  $\omega+42+\omega^2=\omega^2$, the first one is a simple proof by computation, the second one shows how this equality
+  is a consequence of the axioms of the set-theoretic model  by Kurt Schütte~\cite{schutte}. 
+\item
+  Reciprocally, we provide ``concrete'' examples of several formalisation and proof techniques: operational type classes, functions defined by  equations, dependently typed functions, etc.
 \item It may be used as a library on ordinal numbers, for instance for proving termination properties.
@@
  Prior stages of this project have already been presented at
- JFLA~\cite{JFLA2018paper}.
+ JFLA~\cite{PCiota, JFLA2018paper}.
 We present the recent evolutions of the library: new results, interaction within the Coq-Community project~\cite{CoqCommunity}, and documentation generated with Alectryon~\cite{alectryonpaper, alectryongithub}.
@@
    \item  The initial hydra is the graphical representation of an ordinal  $\alpha<\epsilon_0$  in Cantor normal form (See for instance Fig~\ref{fig:iota-example}).
+This property is an invariant of the game.
  \item Hercules always chops off the rightmost head of the hydra (which is always one of the lowest heads).
@@
 Let $\alpha<\epsilon_0$ be an ordinal.
-Using the \texttt{coq-equations} plug-in~\cite{sozeau:hal-01671777}, we define the function which returns the number of steps of the battle starting at configuration $(i,\alpha)$ for  any $i\in\mathbb{N}$.
+Using the \texttt{coq-equations} plug-in~\cite{sozeau:hal-01671777}, we define the function which returns the number of steps of the battle starting with $\alpha$ at step $i$ for  any $i\in\mathbb{N}$.
 We prove that this number is greater or equal than
-$H'_\alpha(i)-i$, where $H'$ is a slight variant of the Wainer Hardy hierarchy of rapidly growing functions~\cite{BW85, KS81, Promel2013, Wainer1970}, defined in Fig.~\ref{fig:Hprime} and
-\ref{fig:HprimeCoq}.
+$H'_\alpha(i)-i$, where $H'$ is a slight variant of the Wainer Hardy hierarchy of rapidly growing functions~\cite{BW85, KS81, Promel2013, Wainer1970}, defined by transfinite recursion in Fig.~\ref{fig:Hprime}.
 
@@
   \begin{itemize}
-\item If $\alpha=0$, then $H'_\alpha (k)= k$ for any natural number $k$.
+\item If $\alpha=0$, then $H'_\alpha (i)= i$ for any natural number $i$.
 \item If $\alpha=\beta+1$, then 
-$H'_\alpha(k)=H'_\beta(k+1)$ for any $k \in \mathbb{N}$
+$H'_\alpha(i)=H'_\beta(i+1)$ for any $i \in \mathbb{N}$
 \item If $\alpha$ is a limit ordinal, then 
-$H'_\alpha(k) = H'_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
+$H'_\alpha(i) = H'_{(\canonseq{\alpha}{i+1})}(k)$ for any $i\in \mathbb{N}$.
 \end{itemize}
+\vspace{4pt}
+\inputsnippets{Hprime/HprimeDef}
 \caption{The rapidly growing hierarchy $H'$}
-\label{Hprime}
+\label{fig:Hprime}
 \end{figure}
 
-\begin{figure}[h]
-  \centering
-  \inputsnippets{Hprime/HprimeDef}
-  \caption{$H'$: definition in \emph{Coq}}
-  \label{fig:HprimeCoq}
-\end{figure}
-
-From this result, we infer that, for a given $\alpha$, the considered function is not primitive recursive.
+\label{sect:not-pr}
+From this result, we infer that the considered function is not primitive recursive in general (for $\alpha\geq\omega^\omega$).
 
@@
 
-% \begin{eqnarray}
-%   \label{eq:1}
-%   \omega+42+\omega^2&=&\omega^2
-% \end{eqnarray}
+\subsection{The Coq-community ecosystem}
+\begin{todo}
+Small presentation.
+\end{todo}
 
-% We also implemented in \emph{Coq} the axiomatic definition of the set $\mathbb{O}$ of countable ordinals by Schütte~\cite{schutte, PCiota}.
+In order to prove formally that the length of the considered
+kind of battles is not given by a primitive recursive function, weused the formalisation of primitive recursive functions, part
+of Russell O'Connor's contribution on G\"{o}del first incompleteness theorem \cite{OConnor05}.
+We chose to host and maintain this work in Coq-community.
+The pedagogical interest of this formalisation led us to add new examples and exercises on primtive recursive functions.
 
-% The proof of $\omega+42+\omega^2=\omega^2$ is no more computational, but use the following definitions and properties.
+\subsection{Documentation with Alectryon}
+\begin{todo}
+  Interest of Alectryon:
+  \begin{itemize}
+  \item Despite the frequent changes (improvements) in the Coq scripts, code inclusion (and \textit{Coq}'s anwers are up to date.
 
-% \begin{itemize}
-% \item Any set of countable ordinals can be enumerated by a function defined on a segment $[0, \alpha)$ of countable ordinals.
-% \item Let $\alpha$ and $\beta$ be (countable) ordinals. Then $\alpha+\beta$ is the $\beta$-th element of the enumeration of the set of countable ordinals greater or equal than $\alpha$.
-% \item Let $\phi_0$ be the enumeration of the set of \emph{additive principal ordinals}
-%   $\{\alpha\in\mathbb{O}| \forall \beta\in \mathbb{O}, \beta+\alpha=\alpha\}$. Then $\omega^\alpha$ is just a notation for
-%   $\phi_0(\alpha)$.
-% \end{itemize}
+  \item Presentation of the main steps of proofs, possibly interleaved with text (Figures \vref{fig:Ex42E0} and \vref{fig:Ex42-schutte},).
 
-\begin{todo}
-  Comment the multiple views of a same mathematical object:
-  \begin{itemize}
-
-  \item Bridge lemmas (relative correctness of different representations).
-    \item The user/reader may compare  different proofs of the same result (For instance the equality $\omega+42+\omega^2= \omega^2$).
-\emph{Example in Figures~\ref{fig:Ex42E0} and \ref{fig:Ex42-schutte}}.
   \end{itemize}
 \end{todo}
-
-\section{The Coq-community ecosystem}
-
-\section{Documentation with Alectryon}
-
-We present two ~examples of the documentation generated with Alectryon~\cite{alectryonpaper}: A simple proof by computation (Fig.~\ref{fig:Ex42E0}), and a longer proof where only the main steps are displayed, interleaved with plain text.
 
@@
 
-\label{fig:Ex42-schutte}
 \fbox{\begin{minipage}[h]{1.0\linewidth}
@@
 \caption{A proof interleaved with text (from the book)}
-  \label{fig:Ex42b}
+  \label{fig:Ex42-schutte}
 \end{figure}
@@
 \begin{todo}
-  \begin{itemize}
-  \item  Soundness (correction, maintenance) of the Alectryon generated doc.
-  \item Communications between co-maintained libraries
-          (Ackermann, Gaia, etc.)
-  \item Work destined both to Coq-users and people interested in discrete math ?
-  \end{itemize}
+  Work destined both to Coq-users and people interested in discrete math ?
  \end{todo}
@@
 It would be interesting to build a library which could be used as a gate giving access to both the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}.
-For instance, the associativity of multiplication of ordinals below $\epsilon_0$, already proven in \textit{Gaia} could be tranferred ``for free'' to the data structures of \textit{hydra-battles}, thanks to the isomorphism between both implementations.
+For instance, the associativity of multiplication of ordinals below $\epsilon_0$, already proven in \textit{Gaia} was tranferred ``for free'' to the data structures of \textit{hydra-battles}, thanks to the isomorphism between both implementations.
 

### 2021-09-18 doc/JFLA2022/paper.tex
small changes

@@
 \inputsnippets{Hprime/HprimeDef}
-\caption{The rapidly growing hierarchy $H'$}
+\caption{The rapidly growing hierarchy $H'$ of arithmetic functions}
 \label{fig:Hprime}
@@
 
-\subsection{The Coq-community ecosystem}
+\subsection{The Coq-Community ecosystem}
 \begin{todo}
@@
 of Russell O'Connor's contribution on G\"{o}del first incompleteness theorem \cite{OConnor05}.
-We chose to host and maintain this work in Coq-community.
-The pedagogical interest of this formalisation led us to add new examples and exercises on primtive recursive functions.
+Computability is a key topic in Computer Science teaching.
+Moreover O'Conors library is a nice illustation on dependently typed functions, so we chose to devote a full chapter to this formalisation, with comments on the definitions and proofs, [counter]-examples and exercises.
 

### 2021-09-18 doc/part-hydras.tex
small changes

@@
 
-In this part, we present a development  for the \coq{} proof assistant, after the work of Kirby and Paris. This formalization contains the following main parts:
+In this part, we present a development  for the \coq{} proof assistant, after the work of Kirby and Paris~\cite{KP82}. This formalization contains the following main parts:
 
@@
 The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on  notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
-We added new material: axiomatic definitions of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, equations, etc.
+We added new material: axiomatic definitions of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, function derinition by equations, etc.
 
@@
 \begin{Coqsrc}
-Inductive CompareSpec (Peq Plt Pgt : Prop) : comparison -> Prop :=
+  Inductive CompareSpec (Peq Plt Pgt : Prop) :
+  comparison -> Prop :=
     CompEq : Peq -> CompareSpec Peq Plt Pgt Eq

### 2021-09-18 doc/part-hydras.tex
small changes

@@
 \item Extension to $\Gamma_0$ (in Veblen normal form)
-\item More lemmas about hierarchies of rapidly growing functions, and their relationship 
+\item More lemmas about hierarchies of rapidly growing functioit ns, and their relationship 
     with primitive recursive functions and provability in Peano arithmetic 

### 2021-09-18 doc/hydras.tex

@@
 \begin{quote}
- (Aprieta el bast\'on con las dos manos, se yergue un tanto,
-casi con entusiasmo) ¡Caramba! Claro... los n\'umeros
-transfinitos, Kantor...
+ \emph{(Aprieta el bast\'on con las dos manos, se yergue un tanto,
+casi con entusiasmo)} ¡Caramba! Claro \dots{} los n\'umeros
+transfinitos, Kantor \dots{}
 
@@
 Quotations of \coq{} source  from Standard Library and code which lead to compilation error or too long computations are displayed as follows.
+
+\begin{Coqsrc}
+  Inductive CompareSpec (Peq Plt Pgt : Prop) :
+  comparison -> Prop :=
+    CompEq : Peq -> CompareSpec Peq Plt Pgt Eq
+  | CompLt : Plt -> CompareSpec Peq Plt Pgt Lt
+  | CompGt : Pgt -> CompareSpec Peq Plt Pgt Gt.
+\end{Coqsrc}
 

### 2021-09-18 doc/JFLA2022/paper.tex
minor corrections

@@
 \item If $\alpha=\beta+1$, then 
-$H'_\alpha(i)=H'_\beta(i+1)$ for any $i \in \mathbb{N}$
+$H'_\alpha(i)=H'_\beta(i+1)$ for any $i \in \mathbb{N}$.
 \item If $\alpha$ is a limit ordinal, then 
@@
 kind of battles is not given by a primitive recursive function, weused the formalisation of primitive recursive functions, part
-of Russell O'Connor's contribution on G\"{o}del first incompleteness theorem \cite{OConnor05}.
+of Russell O'Connor's contribution on G\"{o}del first incompleteness theorem \cite{OConnor05, Goedel}.
+For this purpose, and above all by consideration of the scientific interest of this contribution, we chose to host and maintain this work within \textit{Coq-Community}.
+
 Computability is a key topic in Computer Science teaching.
-Moreover O'Conors library is a nice illustation on dependently typed functions, so we chose to devote a full chapter to this formalisation, with comments on the definitions and proofs, [counter]-examples and exercises.
+Moreover O'Conor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter to this formalisation, with comments on the definitions and proofs, [counter]-examples and exercises.
 
@@
   \begin{itemize}
-  \item Despite the frequent changes (improvements) in the Coq scripts, code inclusion (and \textit{Coq}'s anwers are up to date.
+  \item Despite the frequent changes (improvements) in the Coq scripts, code inclusion and \textit{Coq}'s answers are up to date.
 
-  \item Presentation of the main steps of proofs, possibly interleaved with text (Figures \vref{fig:Ex42E0} and \vref{fig:Ex42-schutte},).
+  \item Presentation of the main steps of proofs, possibly interleaved with text (Figures \vref{fig:Ex42E0} and \vref{fig:Ex42-schutte}).
 

### 2021-09-19 doc/JFLA2022/paper.tex
minor corrections

@@
 Hydra games (also known as \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians, 
-Laurie Kirby and Jeff Paris~\cite{KP82}: \emph{Accessible Independence Results for Peano Arithmetic}.
+Laurie Kirby and Jeff Paris: \emph{Accessible Independence Results for Peano Arithmetic}~\cite{KP82}.
 This article describes a game between two players: Hercules and a hydra.
@@
 
-The Gaia project~\cite{Gaia}, also maintained on \textit{Coq-community} contains a development by José Grimm~\cite{grimm:hal-00911710}.
+The Gaia project~\cite{Gaia}, also maintained on \textit{Coq-Community} contains a development by José Grimm~\cite{grimm:hal-00911710}.
 This library -- dedicated  to the implementaion in \textit{Coq (SSreflect)} of books from  Bourbaki's Elements of Mathematics --- contains a generalization of two types of ordinals (\texttt{T1} for the ordinal $\epsilon_0$ and
@@
 
-It would be interesting to build a library which could be used as a gate giving access to both the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}.
+It would be interesting to build a ``gate'' giving access to both the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}.
 For instance, the associativity of multiplication of ordinals below $\epsilon_0$, already proven in \textit{Gaia} was tranferred ``for free'' to the data structures of \textit{hydra-battles}, thanks to the isomorphism between both implementations.
+
+Another direction would be to write in \textit{Coq} a formal proof of the statement of the second theorem of Kirby and Paris, using the formalization of Peano Arithmetic by O'Connor~\cite{Goedel}.
 

### 2021-09-19 doc/JFLA2022/paper.tex
minor corrections

@@
 \begin{abstract}
-  We present the current state of a \textit{Coq} library on ordinal numbers.
+  \textbf{To do!} We present the current state of a \textit{Coq} library on ordinal numbers.
 
@@
 \end{itemize}
+   \begin{figure}[htb]
+\centering
+\begin{tikzpicture}[very thick, scale=0.3]
+\node (foot) at (10,0) {$\bullet$};
+\node (N1) at (2,2) {$\bullet$};
+\node (N2) at (10,2) {$\bullet$};
+\node (N22) at (7,2) {$\bullet$};
+\node (N3) at (14,2) {$\bullet$};
+\node (N4) at (18,2) {$\Smiley[2][green]$};
+\node (N5) at (0,4) {$\bullet$};
+\node (N6) at (2,5) {$\Smiley[2][green]$};
+\node (N7) at (4,6) {$\Smiley[2][green]$};
+\node (N88) at (7,4) {$\bullet$};
+\node (N8) at (10,4) {$\bullet$};
+\node (N9) at (14,6) {$\Smiley[2][green]$};
+\node (N10) at (0,8) {$\Smiley[2][green]$};
+\node (N11) at (10,7) {$\Smiley[2][green]$};
+\node (N111) at (7,7) {$\Smiley[2][green]$};
+\draw (foot) to [bend left=10] (N1);
+\draw (foot) -- (N2);
+\draw (foot) -- (N22);
+\draw (foot) -- (N3);
+\draw (foot) -- (N4);
+\draw (N1) to  (N5);
+\draw (N1) to   [bend left=10] (N6);
+\draw (N1) to   [bend right=20] (N7);
+\draw (N2) to  (N8);
+\draw (N22) to  (N88);
+\draw (N8) to  (N11);
+\draw (N88) to  (N111);
+\draw (N3) to  (N9);
+\draw (N5) to  (N10);
+\end{tikzpicture}
+\caption{The hydra associated with $\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$ \label{fig:iota-example}}
+
+\end{figure}
 
@@
  $\canonseq{\alpha}{n+1}$, the $(n+1)$th element of the canonical sequence of $\alpha$~\cite{KS81}.
-
-   \begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.3]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (10,2) {$\bullet$};
-\node (N22) at (7,2) {$\bullet$};
-\node (N3) at (14,2) {$\bullet$};
-\node (N4) at (18,2) {$\Smiley[2][green]$};
-\node (N5) at (0,4) {$\bullet$};
-\node (N6) at (2,5) {$\Smiley[2][green]$};
-\node (N7) at (4,6) {$\Smiley[2][green]$};
-\node (N88) at (7,4) {$\bullet$};
-\node (N8) at (10,4) {$\bullet$};
-\node (N9) at (14,6) {$\Smiley[2][green]$};
-\node (N10) at (0,8) {$\Smiley[2][green]$};
-\node (N11) at (10,7) {$\Smiley[2][green]$};
-\node (N111) at (7,7) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N22);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (N1) to  (N5);
-\draw (N1) to   [bend left=10] (N6);
-\draw (N1) to   [bend right=20] (N7);
-\draw (N2) to  (N8);
-\draw (N22) to  (N88);
-\draw (N8) to  (N11);
-\draw (N88) to  (N111);
-\draw (N3) to  (N9);
-\draw (N5) to  (N10);
-\end{tikzpicture}
-\caption{The hydra associated with $\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$ \label{fig:iota-example}}
-
-\end{figure}
 
@@
 The Gaia project~\cite{Gaia}, also maintained on \textit{Coq-Community} contains a development by José Grimm~\cite{grimm:hal-00911710}.
-This library -- dedicated  to the implementaion in \textit{Coq (SSreflect)} of books from  Bourbaki's Elements of Mathematics --- contains a generalization of two types of ordinals (\texttt{T1} for the ordinal $\epsilon_0$ and
+This library -- dedicated  to the implementation in \textit{Coq (SSreflect)} of books from  Bourbaki's Elements of Mathematics --- contains a generalization of two types of ordinals (\texttt{T1} for the ordinal $\epsilon_0$ and
 \texttt{T2} for $\Gamma_0$) to a larger ordinal defined by Ackermann, and relates these data structures with Bourbaki's set theory.
-It is interesting to note that both libraries \textit{Gaia} and \textit{hydra-battles} contain their own (isomorphic) definitions of
-\texttt{T1} and \texttt{T2}, originated in~\cite{CantorContrib}.
-The types --- defined in different modules --- cannot be considered as equal. Even more, some operations like addition, multiplication, etc. have now different definitions.
+It is interesting to note that both libraries \textit{Gaia} and \textit{hydra-battles} contain their own (isomorphic but not identical) definitions of types \texttt{T1} and \texttt{T2}, both originated in~\cite{CantorContrib}.
+Our plan is to build a  ``gate'' giving access to both the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}, and make possible the transfer of theorems from one of the two libraries to the other one.
 
-It would be interesting to build a ``gate'' giving access to both the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}.
-For instance, the associativity of multiplication of ordinals below $\epsilon_0$, already proven in \textit{Gaia} was tranferred ``for free'' to the data structures of \textit{hydra-battles}, thanks to the isomorphism between both implementations.
-
-Another direction would be to write in \textit{Coq} a formal proof of the statement of the second theorem of Kirby and Paris, using the formalization of Peano Arithmetic by O'Connor~\cite{Goedel}.
+Another direction would be to write in \textit{Coq} a formal proof of the original statement of the second theorem of Kirby and Paris.
+For this purpose, we plan to import  the formalization of Peano Arithmetic by O'Connor~\cite{Goedel}.
 

### 2021-09-19 doc/part-hydras.tex
adapt hydras text about Comparable typeclass

--- 2021-09-13 doc/part-hydras.tex
@@
 
-\subsection{A class for ordinal notations}
+\subsection{Classes for ordinal notations}
 
@@
 
-\subsubsection{The \texttt{Comparable class}}
-
-The following class, contributed by Jérémy Damour and Théo Zimmermann, allows us to apply generic lemmas and tactics
-about decidable strict orders.
+\subsubsection{The \texttt{Comparison} and \texttt{Comparable} classes}
+
+We use the operational class \texttt{Comparison} of comparison functions to define the \texttt{Comparable} class, contributed by Jérémy Damour and Théo Zimmermann, which allows us to apply generic lemmas and tactics about decidable strict orders.
 The correctness of the comparison function is expressed through Stdlib's type 
-\texttt{Datatypes.CompareSpec}.
+\texttt{Datatypes.CompareSpec} as specialized by \texttt{Datatypes.CompSpec}.
 
@@
   | CompGt : Pgt -> CompareSpec Peq Plt Pgt Gt.
+
+Definition CompSpec {A} (eq lt : A->A->Prop)(x y:A) : comparison -> Prop :=
+ CompareSpec (eq x y) (lt x y) (lt y x).
 \end{Coqsrc}

### 2021-09-20 doc/JFLA2022/paper.tex
minor corrections

@@
 
-\begin{todo}
-  Work destined both to Coq-users and people interested in discrete math ?
- \end{todo}
-
-\subsection{Perspectives}
+%\subsection{Perspectives}
 
@@
 For this purpose, we plan to import  the formalization of Peano Arithmetic by O'Connor~\cite{Goedel}.
+
+The \emph{Hydra-battles} project is not limited to the study of ordinal numbers and applications, but aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
+For instance, we are also developing
+a module about efficient exponentiation algorithms, and hope
+to extend our priject to new topics.
 

### 2021-09-20 doc/JFLA2022/paper.tex
small corrections; add a README.md
Two typos and one addition.

@@
 %
-\title{Hydras and Cie (Draft) \\
+\title{Hydras and Cie  \\
   Variations on Kirby \& Paris' hydra battles and ordinal numbers in Coq}
@@
 
-The contrast between the simplicity of the statements above and the complexity of their proofs convinced us that it is a good thema for a commented library of formal proofs.
+The contrast between the simplicity of the statements above and the complexity of their proofs convinced us that it is a good theme for a commented library of formal proofs.
 
@@
 \item Complex formalisations and proofs are explained in an
-  electronic book~\cite{HydraBook}. Whenever various reasonable choices exist, we try to present and compare the alternatives.
+  electronic book~\cite{HydraBook} (PDF document of over 280 pages). Whenever various reasonable choices exist, we try to present and compare the alternatives.
   For instance, Figures~\vref{fig:Ex42E0} and \vref{fig:Ex42-schutte} show two radically different proofs of the equality
@@
 In order to prove formally that the length of the considered
-kind of battles is not given by a primitive recursive function, weused the formalisation of primitive recursive functions, part
+kind of battles is not given by a primitive recursive function, we used the formalisation of primitive recursive functions, part
 of Russell O'Connor's contribution on G\"{o}del first incompleteness theorem \cite{OConnor05, Goedel}.

### 2021-09-20 doc/JFLA2022/Ex42b.tex
small corrections

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      ~~\PY{n+nb}{assert}~\PY{o}{(}\PY{n}{HAP}\PY{o}{:}~\PY{n}{In}~\PY{n}{AP}~\PY{o}{(}\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{))}~\PY{n+nb+bp}{by}~\PY{n+nb}{apply}~\PY{n}{AP\PYZus{}phi0}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~\PY{n+nb}{elim}~~\PY{n}{HAP}\PY{o}{;}~\PY{n+nb}{intros}~~\PY{n}{\PYZus{}}~\PY{n}{H0}\PY{o}{;}~\PY{n+nb}{apply}~\PY{n}{H0}\PY{o}{;}~\PY{n+nb}{clear}~\PY{n}{H0}\PY{o}{.}
    \end{input}
    \sep
    \begin{output}
      \begin{goals}
        \begin{goal}
          \begin{hyps}
            \hyp{HAP}{\PY{n}{In}~\PY{n}{AP}~\PY{o}{(}\PY{n}{phi0}~\PY{o}{(}\PY{n}{F}~\PY{l+m+mi}{2}\PY{o}{))}}
          \end{hyps}
          \sep
          \infrule{}
          \sep
          \begin{conclusion}
            \PY{n+nb+bp}{omega}~\PY{o}{+}~\PY{n}{F}~\PY{l+m+mi}{42}~\PY{o}{\PYZlt{}}~\PY{n}{phi0}~\PY{o}{(}\PY{n}{F}~\PY{l+m+mi}{2}\PY{o}{)}
          \end{conclusion}
        \end{goal}
      \end{goals}
    \end{output}
  \end{sentence}
\end{alectryon}

### 2021-09-20 doc/JFLA2022/Ex42c.tex
small corrections

--- 2021-09-20 doc/JFLA2022/Ex42b.tex
@@
     \begin{input}
-      ~~\PY{n+nb}{assert}~\PY{o}{(}\PY{n}{HAP}\PY{o}{:}~\PY{n}{In}~\PY{n}{AP}~\PY{o}{(}\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{))}~\PY{n+nb+bp}{by}~\PY{n+nb}{apply}~\PY{n}{AP\PYZus{}phi0}\PY{o}{.}\nl
-    \end{input}
-  \end{sentence}
-  \sep
-  \begin{sentence}
-    \begin{input}
-      ~~\PY{n+nb}{elim}~~\PY{n}{HAP}\PY{o}{;}~\PY{n+nb}{intros}~~\PY{n}{\PYZus{}}~\PY{n}{H0}\PY{o}{;}~\PY{n+nb}{apply}~\PY{n}{H0}\PY{o}{;}~\PY{n+nb}{clear}~\PY{n}{H0}\PY{o}{.}
+      ~~\PY{n+nb}{assert}~\PY{o}{(}\PY{n}{Hlt}\PY{o}{:}~\PY{n+nb+bp}{omega}~\PY{o}{\PYZlt{}}~\PY{n}{phi0}~\PY{o}{(}\PY{n}{F}~\PY{l+m+mi}{2}\PY{o}{))}~\PY{n+nb+bp}{by}~\nl
+      ~~~~~~\PY{o}{(}~\PY{n+nb}{rewrite}~\PY{n}{omega\PYZus{}eqn}\PY{o}{;}~\PY{n+nb}{apply}~\PY{n}{phi0\PYZus{}mono}\PY{o}{,}~\PY{n}{finite\PYZus{}mono}\PY{o}{;}\nl
+      ~~~~~~~~\PY{n+nb}{auto~with}~\PY{n}{arith}\PY{o}{).}
     \end{input}
@@
             \hyp{HAP}{\PY{n}{In}~\PY{n}{AP}~\PY{o}{(}\PY{n}{phi0}~\PY{o}{(}\PY{n}{F}~\PY{l+m+mi}{2}\PY{o}{))}}
+            \sep
+            \hyp{Hlt}{\PY{n+nb+bp}{omega}~\PY{o}{\PYZlt{}}~\PY{n}{phi0}~\PY{o}{(}\PY{n}{F}~\PY{l+m+mi}{2}\PY{o}{)}}
           \end{hyps}

### 2021-09-20 doc/JFLA2022/HprimeDef.tex
small corrections

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      \PY{n}{Equations}~\PY{n}{H\PYZsq{}\PYZus{}}~\PY{o}{(}\PY{n}{alpha}\PY{o}{:}~\PY{n}{E0}\PY{o}{)}~\PY{o}{(}\PY{n}{i}\PY{o}{:}\PY{n}{nat}\PY{o}{)}~\PY{o}{:}~~\PY{n}{nat}~~\PY{n+nb+bp}{by}~\PY{k+kn}{wf}~~\PY{n}{alpha}~\PY{n}{Lt}~\PY{o}{:=}\nl
      ~~\PY{n}{H\PYZsq{}\PYZus{}}~\PY{n}{alpha}~~\PY{n}{i}~\PY{k+kr}{with}~\PY{n}{E0\PYZus{}eq\PYZus{}dec}~\PY{n}{alpha}~\PY{n}{Zero}~\PY{o}{:=}\nl
      ~~~~\PY{o}{\PYZob{}}~\PY{o}{|}~\PY{n+nb}{left}~\PY{n}{\PYZus{}zero}~\PY{o}{=\PYZgt{}}~~\PY{n}{i}~\PY{o}{;}\nl
      ~~~~~~\PY{o}{|}~\PY{n+nb}{right}~\PY{n}{\PYZus{}nonzero}\nl
      ~~~~~~~~~~\PY{k+kr}{with}~\PY{n}{Utils}\PY{o}{.}\PY{n}{dec}~\PY{o}{(}\PY{n}{Limitb}~\PY{n}{alpha}\PY{o}{)}~\PY{o}{:=}\nl
      ~~~~~~~~~~\PY{o}{\PYZob{}}~\PY{o}{|}~\PY{n+nb}{left}~\PY{n}{\PYZus{}limit}~\PY{o}{=\PYZgt{}}~~\PY{n}{H\PYZsq{}\PYZus{}}~\PY{o}{(}\PY{n}{Canon}~\PY{n}{alpha}~\PY{o}{(}\PY{n}{S}~\PY{n}{i}\PY{o}{))}~~\PY{n}{i}~\PY{o}{;}\nl
      ~~~~~~~~~~~~\PY{o}{|}~\PY{n+nb}{right}~\PY{n}{\PYZus{}successor}~\PY{o}{=\PYZgt{}}~~\PY{n}{H\PYZsq{}\PYZus{}}~\PY{o}{(}\PY{n}{Pred}~\PY{n}{alpha}\PY{o}{)}~\PY{o}{(}\PY{n}{S}~\PY{n}{i}\PY{o}{)\PYZcb{}\PYZcb{}.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{txt}
    \nl
  \end{txt}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Solve~All~Obligations~with}~\PY{n+nb}{auto~with}~\PY{n}{E0}\PY{o}{.}
    \end{input}
  \end{sentence}
\end{alectryon}

### 2021-09-20 doc/JFLA2022/paper.tex
change the Makefile
small corrections

--- 2021-09-19 doc/JFLA2022/paper.tex
@@
   {{\renewenvironment{alectryon}{}{}% Skip \begin{alectryon} included in snippet
-    \input@old{../movies/snippets/#1}}}
+    \input@old{#1}}}
 \makeatother
@@
 \begin{abstract}
-  \textbf{To do!} We present the current state of a \textit{Coq} library on ordinal numbers.
-
+  \textbf{To do!} 
 \end{abstract}
@@
 \label{sect:introduction}
+The \emph{Hydra-battles} project, part of \emph{Coq-Community}~\cite{CoqCommunity}  aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
+
 Hydra games (also known as \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians, 
@@
 \vspace{4pt}
-\inputsnippets{Hprime/HprimeDef}
+\inputsnippets{HprimeDef}
 \caption{The rapidly growing hierarchy $H'$ of arithmetic functions}
@@
       \begin{minipage}[h]{1.0\linewidth}
-        \inputsnippets{E0/Ex42}
+        \inputsnippets{Ex42}
       \end{minipage}}
@@
 
-\inputsnippets{Schutte/Ex42a}
+\inputsnippets{Ex42a}
 
@@
 
-\inputsnippets{Schutte/Ex42b}
+\inputsnippets{Ex42b}
 
@@
 
-\inputsnippets{Schutte/Ex42d, Schutte/Ex42c, Schutte/Ex42e}
+\inputsnippets{Ex42d, Ex42c, Ex42e}
 
@@
 For this purpose, we plan to import  the formalization of Peano Arithmetic by O'Connor~\cite{Goedel}.
+
+ \emph{Hydra-battles} is not limited to the study of ordinal numbers and applications.
+ For instance, we are also developing 
+a module about efficient exponentiation algorithms, and hope
+to extend our project to new topics.
 

### 2021-09-22 doc/JFLA2022/paper.tex
small corrections
comparison with the earlier (2018) paper

@@
 \usepackage{tikzsymbols}
+\newtheorem{theorem}{Theorem}
 \usetikzlibrary{arrows}
@@
 \institute{
- LaBRI, Université Bordeaux \\
+Univ. Bordeaux, CNRS, Bordeaux INP, LaBRI, UMR 5800, F-33400 Talence, France \\
   \email{pierre.casteran@labri.fr}
@@
 \item A hydra is a finite tree, traditionnaly presented with the root at the bottom, the leaves of which are called \emph{heads}
-  (Fig~\vref{fig:iota-example}).
+  (Fig~\ref{fig:iota-example}).
 \item At every round, Hercules chops off one head of the hydra. If the head is at a distance greater or equal than 2 from the root,
-  then some sub-tree $h$ of the hydra is copied a certain amount $i$ of times. The number $i$ of copies and the copied sub-hydra $h$ may depend of the considered variant of the game
+  then some sub-tree $h$ of the hydra is copied a certain amount $n$ of times. The number $n$ of copies and the copied sub-hydra $h$ may depend of the considered variant of the game
   or the time elapsed since the beginning of the fight.
@@
 
-\begin{quote}
-\begin{description}
-\item[$(1)$] In the Hydra game, every strategy is a winning strategy for Hercules.
-  \item[$(2)$] The theorem $(1)$ cannot be proved in Peano Arithmetic.
-\end{description}
-\end{quote}
+\begin{theorem}
+  In the Hydra game, every strategy is a winning strategy for Hercules. \label{kp:thm1}
+\end{theorem}
+
+\begin{theorem}
+  Theorem~\ref{kp:thm1} cannot be proved in Peano Arithmetic. \label{kp:thm2}
+\end{theorem}
 
@@
 
-\begin{itemize}
-\item We present complete proof scripts for the \textit{Coq} proof assistant~\cite{Coq}. These scripts may consulted and replayed by the reader.
-\item Complex formalisations and proofs are explained in an
+We present complete proof scripts for the \textit{Coq} proof assistant~\cite{Coq}. These scripts may consulted and replayed by the reader.
+
+Complex formalisations and proofs are explained in an
   electronic book~\cite{HydraBook}. Whenever various reasonable choices exist, we try to present and compare the alternatives.
-  For instance, Figures~\vref{fig:Ex42E0} and \vref{fig:Ex42-schutte} show two radically different proofs of the equality
+  For instance, Figures~\ref{fig:Ex42E0} and \ref{fig:Ex42-schutte} show two radically different proofs of the equality
   $\omega+42+\omega^2=\omega^2$, the first one is a simple proof by computation, the second one shows how this equality
   is a consequence of the axioms of the set-theoretic model  by Kurt Schütte~\cite{schutte}.
-\item
+
   Reciprocally, we provide ``concrete'' examples of several formalisation and proof techniques: operational type classes, functions defined by  equations, dependently typed functions, etc.
-\item It may be used as a library on ordinal numbers, for instance for proving termination properties.
-  \end{itemize}
+
+Our development may be used as a library on ordinal numbers, for instance for proving termination properties.
 
@@
 
-\section{New results}
-In order to make evident how long a hydra battle can be,
-one may consider a deterministic variant of the game,
-``compute'' the number of steps of the battles, and thus obtain a minoration of the general case.
+\section{Changes in the library}
+The 2018 article contains the proof of  a variant of Theorem~\ref{kp:thm2}:
 
-The variant we consider is the same as 
-in~\cite{KP82, bauer2008, BauerHydra}.
+\begin{theorem}
+  There is no termination measure mapping hydras to a segment $[0,\mu)$ where $\mu<\epsilon_0$ for proving the termination of  all hydra battles.\label{thm3}
+\end{theorem}
+
+Indeed, the consideration of \emph{all} battles allowed us to
+consider battles where the hydra can make an arbitrary number of copies at any time, which make easier our proof by absurd.
+Unfortunately, the examples  most shown in the litterature
+(see for instance \cite{KP82, bauer2008, BauerHydra}) 
+assume that the hydra grows $n$ copies at step $n$ of the game, which is incompatible with our proof:
+
+\vspace{6pt}
+
+We prove now that Theorem~\ref{thm3} still holds with this ``usual'' kind of battles, by borrowing new combinatorial results from~\cite{KS81}.
+
+ Without loss of generality, we could consider the following restrictions:
 
  \begin{itemize}
- \item The game starts at an arbitrary initial step $i\in\mathbb{N}$.
+   \item The game starts at an initial step $i\in\mathbb{N}$ (not necessariliy $0$).
+   \item  The initial hydra is the graphical representation of an ordinal  $\alpha<\epsilon_0$  in Cantor normal form (See for instance Fig~\ref{fig:iota-example}). 
+ \item Hercules always chops off the rightmost head of the hydra.
+ \end{itemize}
 
-   \item The hydra grows $n$ copies at step $n$ of the game.
-   \item  The initial hydra is the graphical representation of an ordinal  $\alpha<\epsilon_0$  in Cantor normal form (See for instance Fig~\ref{fig:iota-example}).
-This property is an invariant of the game.
- \item Hercules always chops off the rightmost head of the hydra (which is always one of the lowest heads).
- \end{itemize}
  It mathematical terms, if at step $n$ the hydra is associated with the ordinal $\alpha$, at step $n+1$ it is associated with
@@
 
-  \item Presentation of the main steps of proofs, possibly interleaved with text (Figures \vref{fig:Ex42E0} and \vref{fig:Ex42-schutte}).
+  \item Presentation of the main steps of proofs, possibly interleaved with text (Figures \ref{fig:Ex42E0} and \ref{fig:Ex42-schutte}).
 
@@
 
-Another direction would be to write in \textit{Coq} a formal proof of the original statement of the second theorem of Kirby and Paris.
+Another direction would be to write in \textit{Coq} a formal proof of the original statement of Theorem~\ref{kp:thm2}.
 For this purpose, we plan to import  the formalization of Peano Arithmetic by O'Connor~\cite{Goedel}.

### 2021-09-22 doc/hydras.tex
comparison with the earlier (2018) paper

@@
 
-\author{Pierre Castéran\\ LaBRI, Univ.
-Bordeaux, CNRS UMR 5800, \thanks{With contributions by Yves Bertot, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Karl Palmskog, Clément Pit-Claudel, and Théo Zimmermann.
+\author{Pierre Castéran\\ Univ.
+Bordeaux, CNRS, Bordeaux INP, LaBRI, UMR 5800, F-33400 Talence, France \thanks{With contributions by Yves Bertot, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Karl Palmskog, Clément Pit-Claudel, and Théo Zimmermann.
 The formalization of primitive recursive functions was originally authored by

### 2021-09-22 doc/JFLA2022/paper.tex
comparison with the earlier (2018) paper

@@
 (see for instance \cite{KP82, bauer2008, BauerHydra}) 
-assume that the hydra grows $n$ copies at step $n$ of the game, which is incompatible with our proof:
+assume that the hydra grows $n$ copies at step $n$ of the game, which is incompatible with our proof.
 

### 2021-09-22 doc/JFLA2022/paper.tex
small corrections

@@
 \and
-   Ask for agreement !
+   to fill!
  }
@@
 \section{Changes in the library}
-The 2018 article contains the proof of  a variant of Theorem~\ref{kp:thm2}:
+The 2018 article contains a formal proof of  a variant of Theorem~\ref{kp:thm2}:
 
@@
 We prove now that Theorem~\ref{thm3} still holds with this ``usual'' kind of battles, by borrowing new combinatorial results from~\cite{KS81}.
-
 Without loss of generality, we could consider the following restrictions:
@@
 $H'_\alpha(i)-i$, where $H'$ is a slight variant of the Wainer Hardy hierarchy of rapidly growing functions~\cite{BW85, KS81, Promel2013, Wainer1970}, defined by transfinite recursion in Fig.~\ref{fig:Hprime}.
+From this result, we infer that the considered function is not primitive recursive in general (for $\alpha\geq\omega^\omega$).
 
@@
 \label{sect:not-pr}
-From this result, we infer that the considered function is not primitive recursive in general (for $\alpha\geq\omega^\omega$).
 

### 2021-09-22 doc/JFLA2022/paper.tex
Some simplifications in the description of the recent changes (since 2018)

@@
 \and Clément Pit-Claudel \inst{2}
-\and Théo Zimmermann %\inst{5}
+\and Théo Zimmermann \inst{2}
 }
@@
 \begin{abstract}
-  \textbf{To do!} 
+ To do \dots
 \end{abstract}
@@
 This article describes a game between two players: Hercules and a hydra.
-A short description  can be found in Andrej Bauer's blogpost~\cite{bauer2008} as well as a simulator~\cite{BauerHydra}.
+A short description of the game  can be found in~\cite{bauer2008, KP82, JFLA2018paper}. One can also play with
+Andrej Bauer's simulator~\cite{BauerHydra}.
 In a few words:
@@
 \item At every round, Hercules chops off one head of the hydra. If the head is at a distance greater or equal than 2 from the root,
-  then some sub-tree $h$ of the hydra is copied a certain amount $n$ of times. The number $n$ of copies and the copied sub-hydra $h$ may depend of the considered variant of the game
+  then some sub-tree $h$ of the hydra is copied a certain amount $n$ of times. The number $n$ of copies and the sub-tree $h$ may depend of the considered variant of the game
   or the time elapsed since the beginning of the fight. 
@@
 \end{tikzpicture}
-\caption{The hydra associated with $\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$ \label{fig:iota-example}}
+\caption{The hydra associated with the ordinal $\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$ \label{fig:iota-example}}
 
@@
 
-The contrast between the simplicity of the statements above and the complexity of their proofs convinced us that it is a good thema for a commented library of formal proofs.
-
-We present complete proof scripts for the \textit{Coq} proof assistant~\cite{Coq}. These scripts may consulted and replayed by the reader.
+The contrast between the simplicity of the statements above and the complexity of their proofs convinced us that it is a good thema for a commented library~\cite{HydraBattles} of formal proofs written for the \textit{Coq} proof assistant~\cite{Coq}.
 
@@
 
-  Reciprocally, we provide ``concrete'' examples of several formalisation and proof techniques: operational type classes, functions defined by  equations, dependently typed functions, etc.
-
-Our development may be used as a library on ordinal numbers, for instance for proving termination properties.
+This work is also an opportunity to 
+ provide ``concrete'' examples of several formalisation and proof techniques: operational type classes, functions defined by  equations, dependently typed functions, etc.
+It may be also used as a library on ordinal numbers, for instance for proving termination properties.
 
@@
 \section{Changes in the library}
-The 2018 article contains a formal proof of  a variant of Theorem~\ref{kp:thm2}:
+The 2018 article~\cite{JFLA2018paper} contains a formal proof of  a variant of Theorem~\ref{kp:thm2}:
 

### 2021-09-23 doc/JFLA2022/paper.tex
minor corrections

@@
 
-\usepackage[firstpageonly=false, color={[gray]{0.5}},
-   scale=2.0, text=DRAFT]{draftwatermark}
+%\usepackage[firstpageonly=false, color={[gray]{0.5}},
+%   scale=2.0, text=DRAFT]{draftwatermark}
 % -------------------------------
@@
 \and
-Karl Palmskog \inst{2}
-\and Clément Pit-Claudel \inst{2}
-\and Théo Zimmermann \inst{2}
+Karl Palmskog \inst{3}
+\and Clément Pit-Claudel \inst{4}
+\and Théo Zimmermann \inst{5}
 }
@@
   \email{pierre.casteran@labri.fr}
+\and
+to fill!
+\and
+to fill!
+\and
+to fill!
 \and
@@
 \begin{abstract}
- To do \dots
+  \emph{Hydra-battles} is a collaborative library of commented proofs, developped under the \emph{Coq-Community} project.
+  We present its evolution since a former presentation~\cite{JFLA2018paper} and how its structure is determined by other libraies and the \emph{Alectryon} literate proving tool.
 \end{abstract}
@@
 \begin{itemize}
-\item A hydra is a finite tree, traditionnaly presented with the root at the bottom, the leaves of which are called \emph{heads}
-  (Fig~\ref{fig:iota-example}).
+\item A hydra is a finite tree, traditionally presented with the root at the bottom, the leaves of which are called \emph{heads}
+  (Fig~\ref{fig:round}).
 \item At every round, Hercules chops off one head of the hydra. If the head is at a distance greater or equal than 2 from the root,
@@
   or the time elapsed since the beginning of the fight.
+  Figure~\ref{fig:round} shows an example of such a transformation with $n=2$.
 \end{itemize}
-   \begin{figure}[htb]
+
+\begin{figure}[h]
   \centering
   \begin{tikzpicture}[very thick, scale=0.3]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (10,2) {$\bullet$};
-\node (N22) at (7,2) {$\bullet$};
-\node (N3) at (14,2) {$\bullet$};
-\node (N4) at (18,2) {$\Smiley[2][green]$};
-\node (N5) at (0,4) {$\bullet$};
-\node (N6) at (2,5) {$\Smiley[2][green]$};
-\node (N7) at (4,6) {$\Smiley[2][green]$};
-\node (N88) at (7,4) {$\bullet$};
-\node (N8) at (10,4) {$\bullet$};
-\node (N9) at (14,6) {$\Smiley[2][green]$};
-\node (N10) at (0,8) {$\Smiley[2][green]$};
-\node (N11) at (10,7) {$\Smiley[2][green]$};
-\node (N111) at (7,7) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N22);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (N1) to  (N5);
-\draw (N1) to   [bend left=10] (N6);
-\draw (N1) to   [bend right=20] (N7);
-\draw (N2) to  (N8);
-\draw (N22) to  (N88);
-\draw (N8) to  (N11);
-\draw (N88) to  (N111);
-\draw (N3) to  (N9);
-\draw (N5) to  (N10);
+  \node (h1) at (4,0){$\bullet$};
+  \node (h2) at (4,2){$\bullet$};
+  \node (h3) at (2,4){$\bullet$};
+  \node (h4) at (0,6){$\bullet$};
+  \node (h5) at (4,6){$\bullet$};
+  \node (h6) at (6,4){$\bullet$};
+  \draw (h1) -- (h2) ;
+  \draw (h2) -- (h3) ;
+  \draw (h2) -- (h6);
+  \draw (h3) -- (h4) ;
+  \draw (h3) -- (h5) ;
+
+\node (hn1) at (14,0){$\bullet$};
+\node (hn2) at (12,2) {$\bullet$};
+\node (hn3) at (12,4) {$\bullet$};
+\node (hn4) at (10,6){$\bullet$};
+  \node (hn5) at (12,6){$\bullet$};
+  \draw (hn1) -- (hn2) ;
+  \draw (hn2) -- (hn3) ;
+  \draw (hn3) -- (hn4) ;
+  \draw (hn3) -- (hn5) ;
+\node (hn2b) at (14,2) {$\bullet$};
+\node (hn3b) at (14,4) {$\bullet$};
+\node (hn4b) at (13.4,6){$\bullet$};
+  \node (hn5b) at (15,6){$\bullet$};
+  \draw (hn1) -- (hn2b) ;
+  \draw (hn2b) -- (hn3b) ;
+  \draw (hn3b) -- (hn4b) ;
+  \draw (hn3b) -- (hn5b) ;
+  \node (hn2c) at (16,2) {$\bullet$};
+\node (hn3c) at (16,4) {$\bullet$};
+\node (hn4c) at (17,6){$\bullet$};
+  \node (hn5c) at (18.5,6){$\bullet$};
+  \draw (hn1) -- (hn2c) ;
+  \draw (hn2c) -- (hn3c) ;
+  \draw (hn3c) -- (hn4c) ;
+  \draw (hn3c) -- (hn5c) ;
 \end{tikzpicture}
-\caption{The hydra associated with the ordinal $\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$ \label{fig:iota-example}}
 
+  \caption{Two successive states of a hydra in a battle}
+  \label{fig:round}
 \end{figure}
+
+ %  \begin{figure}[htb]
+% \centering
+% \begin{tikzpicture}[very thick, scale=0.3]
+% \node (foot) at (10,0) {$\bullet$};
+% \node (N1) at (2,2) {$\bullet$};
+% \node (N2) at (10,2) {$\bullet$};
+% \node (N22) at (7,2) {$\bullet$};
+% \node (N3) at (14,2) {$\bullet$};
+% \node (N4) at (18,2) {$\Smiley[2][green]$};
+% \node (N5) at (0,4) {$\bullet$};
+% \node (N6) at (2,5) {$\Smiley[2][green]$};
+% \node (N7) at (4,6) {$\Smiley[2][green]$};
+% \node (N88) at (7,4) {$\bullet$};
+% \node (N8) at (10,4) {$\bullet$};
+% \node (N9) at (14,6) {$\Smiley[2][green]$};
+% \node (N10) at (0,8) {$\Smiley[2][green]$};
+% \node (N11) at (10,7) {$\Smiley[2][green]$};
+% \node (N111) at (7,7) {$\Smiley[2][green]$};
+% \draw (foot) to [bend left=10] (N1);
+% \draw (foot) -- (N2);
+% \draw (foot) -- (N22);
+% \draw (foot) -- (N3);
+% \draw (foot) -- (N4);
+% \draw (N1) to  (N5);
+% \draw (N1) to   [bend left=10] (N6);
+% \draw (N1) to   [bend right=20] (N7);
+% \draw (N2) to  (N8);
+% \draw (N22) to  (N88);
+% \draw (N8) to  (N11);
+% \draw (N88) to  (N111);
+% \draw (N3) to  (N9);
+% \draw (N5) to  (N10);
+% \end{tikzpicture}
+% \caption{The hydra associated with the ordinal $\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$ \label{fig:iota-example}}
+
+% \end{figure}
 
@@
 This work is also an opportunity to 
- provide ``concrete'' examples of several formalisation and proof techniques: operational type classes, functions defined by  equations, dependently typed functions, etc.
+ provide ``concrete'' examples of formalization and proof techniques: operational type classes, functions defined by  equations, dependently typed functions, etc.
 It may be also used as a library on ordinal numbers, for instance for proving termination properties.
@@
 Indeed, the consideration of \emph{all} battles allowed us to
-consider battles where the hydra can make an arbitrary number of copies at any time, which make easier our proof by absurd.
+consider battles where the hydra can make an arbitrary number of copies at any time, which made artificially easier our proof by absurd.
 Unfortunately, the examples  most shown in the litterature
@@
  \begin{itemize}
-   \item The game starts at an initial step $i\in\mathbb{N}$ (not necessariliy $0$).
-   \item  The initial hydra is the graphical representation of an ordinal  $\alpha<\epsilon_0$  in Cantor normal form (See for instance Fig~\ref{fig:iota-example}). 
+   \item The game starts at an initial step $i\in\mathbb{N}$ (not necessarily $0$).
+   \item  The hydra is always the the graphical representation of some ordinal strictly below $\epsilon_0$ in Cantor normal
+     form. For instance, Fig~\ref{fig:round}) shows the hydras respectively associated with  $\omega^{\omega^2+1}$ and $\omega^{\omega^2}\times 3$.
+
  \item Hercules always chops off the rightmost head of the hydra.
@@
 In order to prove formally that the length of the considered
-kind of battles is not given by a primitive recursive function, weused the formalisation of primitive recursive functions, part
+kind of battles is not given by a primitive recursive function, weused the formalization of primitive recursive functions, part
 of Russell O'Connor's contribution on G\"{o}del first incompleteness theorem \cite{OConnor05, Goedel}.
@@
 
-  \item Presentation of the main steps of proofs, possibly interleaved with text (Figures \ref{fig:Ex42E0} and \ref{fig:Ex42-schutte}).
+  \item Presentation of the main steps of proofs, possibly interleaved with text.
 
   \end{itemize}
+
+   Figures~\ref{fig:Hprime} (code snippet), \ref{fig:Ex42E0} and \ref{fig:Ex42-schutte} have been generated with Alectryon.
 \end{todo}

### 2021-09-23 doc/JFLA2022/paper.tex
fix typos

@@
   \emph{Hydra-battles} is a collaborative library of commented proofs, developped under the \emph{Coq-Community} project.
-  We present its evolution since a former presentation~\cite{JFLA2018paper} and how its structure is determined by other libraies and the \emph{Alectryon} literate proving tool.
+  We present its evolution since a former presentation~\cite{JFLA2018paper} and how its structure is determined by other
+  libraries and the \emph{Alectryon} literate proving tool.
 \end{abstract}
@@
   or the time elapsed since the beginning of the fight.
-  Figure~\ref{fig:round} shows an example of such a transformation with $n=2$.
+  Figure~\ref{fig:round} shows an example with $n=2$.
 \end{itemize}

### 2021-09-23 doc/JFLA2022/paper.tex
fix typos
fix more typos

@@
    \item  The hydra is always the the graphical representation of some ordinal strictly below $\epsilon_0$ in Cantor normal
-     form. For instance, Fig~\ref{fig:round}) shows the hydras respectively associated with  $\omega^{\omega^2+1}$ and $\omega^{\omega^2}\times 3$.
+     form. For instance, Fig~\ref{fig:round} shows the hydras respectively associated with  $\omega^{\omega^2+1}$ and $\omega^{\omega^2}\times 3$.
 
@@
 \begin{figure}[h]
-  \centering
+ \noindent
+ For any $i\in\mathbb{N}$,
    \begin{itemize}
-\item If $\alpha=0$, then $H'_\alpha (i)= i$ for any natural number $i$.
+\item If $\alpha=0$, then $H'_\alpha (i)= i$.
 \item If $\alpha=\beta+1$, then 
-$H'_\alpha(i)=H'_\beta(i+1)$ for any $i \in \mathbb{N}$.
+$H'_\alpha(i)=H'_\beta(i+1)$.
 \item If $\alpha$ is a limit ordinal, then 
-$H'_\alpha(i) = H'_{(\canonseq{\alpha}{i+1})}(k)$ for any $i\in \mathbb{N}$.
+$H'_\alpha(i) = H'_{(\canonseq{\alpha}{i+1})}(i)$.
 \end{itemize}

### 2021-09-23 doc/JFLA2022/paper.tex
fix more typos

@@
    \item The game starts at an initial step $i\in\mathbb{N}$ (not necessarily $0$).
-   \item  The hydra is always the the graphical representation of some ordinal strictly below $\epsilon_0$ in Cantor normal
+   \item  The hydra is always the  representation as a tree of some ordinal strictly below $\epsilon_0$ in Cantor normal
      form. For instance, Fig~\ref{fig:round} shows the hydras respectively associated with  $\omega^{\omega^2+1}$ and $\omega^{\omega^2}\times 3$.
@@
 
+ Our new proof of Theorem~\ref{thm3} is based on a systematic study of strictly decreasing sequences of ordinals below $\epsilon_0$, borrowed from~\cite{KS81}.
+
+We also study the number of steps of a battle:
 Let $\alpha<\epsilon_0$ be an ordinal.
-Using the \texttt{coq-equations} plug-in~\cite{sozeau:hal-01671777}, we define the function which returns the number of steps of the battle starting with $\alpha$ at step $i$ for  any $i\in\mathbb{N}$.
+Using the \texttt{coq-equations} plug-in~\cite{sozeau:hal-01671777}, we define the function which returns the number of steps of the battle starting with $\alpha$ at step $i$, for  any $i$ in $\mathbb{N}$.
 We prove that this number is greater or equal than
@@
  \noindent
- For any $i\in\mathbb{N}$,
+ For any $\alpha$ in $[0,\epsilon_0)$ and any $i$ in $\mathbb{N}$,
    \begin{itemize}

### 2021-09-23 doc/JFLA2022/paper.tex
Add institution
Clarify legend of figure 1 (I think?)
Make hydra symmetric
Clarify what happens in figure 1

@@
 \usepackage{tikzsymbols}
+\usepackage{pifont}
 \newtheorem{theorem}{Theorem}
@@
 \and
-to fill!
+MIT CSAIL, Cambridge, Massachusetts, USA
 \and
@@
   \node (h1) at (4,0){$\bullet$};
-  \node (h2) at (4,2){$\bullet$};
-  \node (h3) at (2,4){$\bullet$};
-  \node (h4) at (0,6){$\bullet$};
-  \node (h5) at (4,6){$\bullet$};
-  \node (h6) at (6,4){$\bullet$};
+  \node[blue] (h2) at (4,2){$\bullet$};
+  \node[blue] (h3) at (2,4){$\bullet$};
+  \node[blue] (h4) at (0,6){$\bullet$};
+  \node[blue] (h5) at (4,6){$\bullet$};
+  \node[red!90!black] (h6) at (6,4){$\bullet$};
   \draw (h1) -- (h2) ;
   \draw (h2) -- (h3) ;
-  \draw (h2) -- (h6);
+  \draw (h2) -- node[red!90!black,font=\small,sloped,shift={(0.01,-0.075)},rotate=90]{\textbf{\ding{34}}} (h6);
   \draw (h3) -- (h4) ;
@@
 \node (hn1) at (14,0){$\bullet$};
-\node (hn2) at (12,2) {$\bullet$};
-\node (hn3) at (12,4) {$\bullet$};
-\node (hn4) at (10,6){$\bullet$};
-  \node (hn5) at (12,6){$\bullet$};
+\node[blue] (hn2) at (12,2) {$\bullet$};
+\node[blue] (hn3) at (12,4) {$\bullet$};
+\node[blue] (hn4) at (10,6){$\bullet$};
+  \node[blue] (hn5) at (12,6){$\bullet$};
   \draw (hn1) -- (hn2) ;
@@
 \node (hn4b) at (13.4,6){$\bullet$};
-  \node (hn5b) at (15,6){$\bullet$};
+  \node (hn5b) at (14.6,6){$\bullet$};
   \draw (hn1) -- (hn2b) ;
@@
 \node (hn3c) at (16,4) {$\bullet$};
-\node (hn4c) at (17,6){$\bullet$};
-  \node (hn5c) at (18.5,6){$\bullet$};
+\node (hn4c) at (16,6){$\bullet$};
+  \node (hn5c) at (18,6){$\bullet$};
   \draw (hn1) -- (hn2c) ;
@@
 
-  \caption{Two successive states of a hydra in a battle}
+  \caption{Two successive states of a hydra in a battle.  Hercules chopped off the rightmost head of the hydra (red), and the whole left tree except the root node (blue) was copied twice.}
   \label{fig:round}

### 2021-09-23 doc/JFLA2022/paper.tex
Minor copyediting

@@
 
+\newcommand{\TODO}[2][]{[\textcolor{red}{TODO (#1):} \emph{#2}]}
+
 \begin{document}
@@
   (Fig~\ref{fig:round}).
-\item At every round, Hercules chops off one head of the hydra. If the head is at a distance greater or equal than 2 from the root,
+\item At every round, Hercules chops off one head of the hydra. If the head is at a distance greater than 1 from the root,
   then some sub-tree $h$ of the hydra is copied a certain amount $n$ of times. The number $n$ of copies and the sub-tree $h$ may depend of the considered variant of the game
@@
 
+\TODO[Clément]{We haven't defined winning strategies}
 \begin{theorem}
@@
   For instance, Figures~\ref{fig:Ex42E0} and \ref{fig:Ex42-schutte} show two radically different proofs of the equality
-  $\omega+42+\omega^2=\omega^2$, the first one is a simple proof by computation, the second one shows how this equality
+  $\omega+42+\omega^2=\omega^2$. The first one is a simple proof by computation, the second one shows how this equality
   is a consequence of the axioms of the set-theoretic model  by Kurt Schütte~\cite{schutte}.
@@
  JFLA~\cite{PCiota, JFLA2018paper}.
-We present the recent evolutions of the library: new results, interaction within the Coq-Community project~\cite{CoqCommunity}, and documentation generated with Alectryon~\cite{alectryonpaper, alectryongithub}.
+We present recent evolutions of the library: new results, interaction within the Coq-Community project~\cite{CoqCommunity}, and documentation generated with Alectryon~\cite{alectryonpaper, alectryongithub}.
 
@@
 
-Indeed, the consideration of \emph{all} battles allowed us to
-consider battles where the hydra can make an arbitrary number of copies at any time, which made artificially easier our proof by absurd.
-Unfortunately, the examples  most shown in the litterature
+\TODO[Clément]{Is there a word missing above?}
+Considering measures applicable to \emph{all} battles allowed us to
+focus on battles where the hydra can make an arbitrary number of copies at any time, which made our proof by contradiction artificially easier.
+Unfortunately, the examples  most commonly shown in the litterature
 (see for instance \cite{KP82, bauer2008, BauerHydra}) 
@@
 
-We prove now that Theorem~\ref{thm3} still holds with this ``usual'' kind of battles, by borrowing new combinatorial results from~\cite{KS81}.
-Without loss of generality, we could consider the following restrictions:
+We now prove that Theorem~\ref{thm3} still holds with these typical battles by borrowing new combinatorial results from~\cite{KS81}.
+Without loss of generality, we assume the following restrictions:
 
@@
 $H'_\alpha(i)-i$, where $H'$ is a slight variant of the Wainer Hardy hierarchy of rapidly growing functions~\cite{BW85, KS81, Promel2013, Wainer1970}, defined by transfinite recursion in Fig.~\ref{fig:Hprime}.
+\TODO[Clément]{It seems that the term ``Wainer Hardy'' hierarchy only appears in his paper and the Hydra-battles book}
 From this result, we infer that the considered function is not primitive recursive in general (for $\alpha\geq\omega^\omega$).
@@
 
+\TODO[Clément]{Highlight `Equations' and remove `omega' highlighting; change theme to something less garish}
+
 \section{Main features}
@@
 In order to prove formally that the length of the considered
-kind of battles is not given by a primitive recursive function, weused the formalization of primitive recursive functions, part
+kind of battles is not given by a primitive recursive function, we used the formalization of primitive recursive functions, part
 of Russell O'Connor's contribution on G\"{o}del first incompleteness theorem \cite{OConnor05, Goedel}.
@@
   \item Presentation of the main steps of proofs, possibly interleaved with text.
+  \item \TODO[Clément]{A good part of the LaTeX support in Alectryon was developed for hydra-battles, and this was the first real application with a custom driver, too.}
+  \item \TODO[Clément]{If you kept a trace of them, I would love to see a discussion of any errors that were caught by including the output systematically instead of copy-pasting it.}
 
@@
 It is interesting to note that both libraries \textit{Gaia} and \textit{hydra-battles} contain their own (isomorphic but not identical) definitions of types \texttt{T1} and \texttt{T2}, both originated in~\cite{CantorContrib}.
-Our plan is to build a  ``gate'' giving access to both the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}, and make possible the transfer of theorems from one of the two libraries to the other one.
+Our plan is to build a  ``gate'' \TODO[Clément]{Bridge?} giving access to both the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}, and make possible the transfer of theorems from one of the two libraries to the other one.
 

### 2021-09-24 doc/JFLA2022/paper.tex
Add institutions.

@@
 \and
-to fill!
+Univ. de Paris, F-75013 Paris, France
 \and
@@
 \and
-to fill!
+Inria, Univ. de Paris, CNRS, IRIF, UMR 8243, F-75013 Paris, France
 }
@@
 
-\TODO[Clément]{We haven't defined winning strategies}
 \begin{theorem}
-  In the Hydra game, every strategy is a winning strategy for Hercules. \label{kp:thm1}
+  In the Hydra game, Hercules eventually wins, whichever the strategy of both players :
+  choice of a head to chop off, choice of the number of copies. 
+ \label{kp:thm1}
 \end{theorem}
@@
 
- It mathematical terms, if at step $n$ the hydra is associated with the ordinal $\alpha$, at step $n+1$ it is associated with
+ In mathematical terms, if at step $n$ the hydra is associated with the ordinal $\alpha$, at step $n+1$ it is associated with
  $\canonseq{\alpha}{n+1}$, the $(n+1)$th element of the canonical sequence of $\alpha$~\cite{KS81}.
@@
 We prove that this number is greater or equal than
-$H'_\alpha(i)-i$, where $H'$ is a slight variant of the Wainer Hardy hierarchy of rapidly growing functions~\cite{BW85, KS81, Promel2013, Wainer1970}, defined by transfinite recursion in Fig.~\ref{fig:Hprime}.
-\TODO[Clément]{It seems that the term ``Wainer Hardy'' hierarchy only appears in his paper and the Hydra-battles book}
+$H'_\alpha(i)-i$, where $H'$ is a slight variant of the Hardy hierarchy of rapidly growing functions~\cite{BW85, KS81, Promel2013, Wainer1970}, defined by transfinite recursion in Fig.~\ref{fig:Hprime}.
 From this result, we infer that the considered function is not primitive recursive in general (for $\alpha\geq\omega^\omega$).
@@
 Computability is a key topic in Computer Science teaching.
-Moreover O'Conor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter to this formalisation, with comments on the definitions and proofs, [counter]-examples and exercises.
+Moreover O'Connor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter to this formalisation, with comments on the definitions and proofs, [counter]-examples and exercises.
 
@@
   \item \TODO[Clément]{If you kept a trace of them, I would love to see a discussion of any errors that were caught by including the output systematically instead of copy-pasting it.}
+  \item \TODO[Pierre]{the order of appearance of snippets in the book is independent from the module structure}
+   \item \TODO[Pierre]{Slicing our proofs into snippets allowed to show only the main parts of our proofs (Figure~\ref{fig:Ex42-schutte}).}
 
@@
 It is interesting to note that both libraries \textit{Gaia} and \textit{hydra-battles} contain their own (isomorphic but not identical) definitions of types \texttt{T1} and \texttt{T2}, both originated in~\cite{CantorContrib}.
-Our plan is to build a  ``gate'' \TODO[Clément]{Bridge?} giving access to both the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}, and make possible the transfer of theorems from one of the two libraries to the other one.
+Our plan is to build a  bridge between  the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}, and make possible the transfer of theorems from one of the two libraries to the other one.
 

### 2021-09-24 doc/JFLA2022/paper.tex
Some changes that were accidentally reverted.

--- 2021-09-23 doc/JFLA2022/paper.tex
@@
 
-The contrast between the simplicity of the statements above and the complexity of their proofs convinced us that it is a good thema for a commented library~\cite{HydraBattles} of formal proofs written for the \textit{Coq} proof assistant~\cite{Coq}.
+The contrast between the simplicity of the statements above and the complexity of their proofs convinced us that it is a good theme for a commented library~\cite{HydraBattles} of formal proofs written for the \textit{Coq} proof assistant~\cite{Coq}.
 
 Complex formalisations and proofs are explained in an
-  electronic book~\cite{HydraBook}. Whenever various reasonable choices exist, we try to present and compare the alternatives.
+  electronic book~\cite{HydraBook} (PDF document of over 280 pages). Whenever various reasonable choices exist, we try to present and compare the alternatives.
   For instance, Figures~\ref{fig:Ex42E0} and \ref{fig:Ex42-schutte} show two radically different proofs of the equality

### 2021-09-24 doc/JFLA2022/paper.tex
Clement's corrections

--- 2021-09-24 doc/JFLA2022/paper.tex
@@
 \and
-Univ. de Paris, F-75013 Paris, France
+to fill!
 \and
@@
 \and
-Inria, Univ. de Paris, CNRS, IRIF, UMR 8243, F-75013 Paris, France
+to fill!
  }
@@
 
-\TODO[Clément]{We haven't defined winning strategies}
 \begin{theorem}
-  In the Hydra game, every strategy is a winning strategy for Hercules. \label{kp:thm1}
+  n the Hydra game, Hercules eventually wins, whichever the strategy of both players :
+  choice of a head to chop off, choice of the number of copies. 
+ \label{kp:thm1}
 \end{theorem}
@@
 
-The contrast between the simplicity of the statements above and the complexity of their proofs convinced us that it is a good theme for a commented library~\cite{HydraBattles} of formal proofs written for the \textit{Coq} proof assistant~\cite{Coq}.
+The contrast between the simplicity of the statements above and the complexity of their proofs convinced us that it is a good thema for a commented library~\cite{HydraBattles} of formal proofs written for the \textit{Coq} proof assistant~\cite{Coq}.
 
 Complex formalisations and proofs are explained in an
-  electronic book~\cite{HydraBook} (PDF document of over 280 pages). Whenever various reasonable choices exist, we try to present and compare the alternatives.
+  electronic book~\cite{HydraBook}. Whenever various reasonable choices exist, we try to present and compare the alternatives.
   For instance, Figures~\ref{fig:Ex42E0} and \ref{fig:Ex42-schutte} show two radically different proofs of the equality

### 2021-09-24 doc/JFLA2022/paper.tex
fix a few typos

@@
 \begin{theorem}
-  n the Hydra game, Hercules eventually wins, whichever the strategy of both players :
+  In the Hydra game, Hercules eventually wins, whichever the strategy of both players :
   choice of a head to chop off, choice of the number of copies. 

### 2021-09-24 doc/JFLA2022/paper.tex
Add introducting subsection on the vision of the project.

@@
 \label{sect:introduction}
+
+\subsection{Vision}
 The \emph{Hydra-battles} project, part of \emph{Coq-Community}~\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
+\emph{Coq-Community} is a community organization that we have founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting Coq packages, and working collaboratively on documentation projects.
+The \emph{Hydra-battles} project demonstrates that these two goals are not independent: interesting Coq packages can become the basis for new documentation.
+%
+This umbrella project now includes evolved versions of former \emph{contribs} Cantor and Additions (under the new names of Hydra-battles and Addition-chains), the Pocklington library, extracted from Russel O'Connor's Goedel \emph{contrib} and a bridge to the Gaia library (\emph{contrib} of José Grimm).
+%
+By following this approach of commenting interesting Coq packages, we provide new documentation content, that contributes to the diversity of the thriving Coq documentation ecosystem.
 
+We call on the Coq users in the JFLA community and beyond to come and join us in this effort, by bringing new interesting projects which are worth presenting to Coq learners and guiding them in their exploration.
+%
+We also always have project ideas to extend further our explorations and anyone is welcome to join the team by sending small or larger contributions through pull requests.
+%
+The current state of the project is already the result of such evolutions after several of us contributed project solutions and new proposals to the initial version of the first author.
+
+Futhermore, contrary to traditionally published books, the ``book'' that forms part of this project is intended to be forever evolving.
+As new Coq formalization patterns and proof techniques appear, the book can be adapted to demonstrate their use (in case they fit well with our applications).
+%
+By using modern maintenance techniques such as continuous integration and deployment, we can ensure that this documentation stays up to date with the latest Coq releases.
+With Alectryon, we ensure that code and documentation are always in sync.
+
+\subsection{Hydra games}
 Hydra games (also known as \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians, 

### 2021-09-24 doc/JFLA2022/paper.tex
Add introducting subsection on the vision of the project.

@@
 
+\TODO[Clément]{We haven't defined winning strategies}
 \begin{theorem}
-  In the Hydra game, Hercules eventually wins, whichever the strategy of both players :
-  choice of a head to chop off, choice of the number of copies. 
- \label{kp:thm1}
+  In the Hydra game, every strategy is a winning strategy for Hercules. \label{kp:thm1}
 \end{theorem}
@@
 
- In mathematical terms, if at step $n$ the hydra is associated with the ordinal $\alpha$, at step $n+1$ it is associated with
+ It mathematical terms, if at step $n$ the hydra is associated with the ordinal $\alpha$, at step $n+1$ it is associated with
  $\canonseq{\alpha}{n+1}$, the $(n+1)$th element of the canonical sequence of $\alpha$~\cite{KS81}.
@@
 We prove that this number is greater or equal than
-$H'_\alpha(i)-i$, where $H'$ is a slight variant of the Hardy hierarchy of rapidly growing functions~\cite{BW85, KS81, Promel2013, Wainer1970}, defined by transfinite recursion in Fig.~\ref{fig:Hprime}.
+$H'_\alpha(i)-i$, where $H'$ is a slight variant of the Wainer Hardy hierarchy of rapidly growing functions~\cite{BW85, KS81, Promel2013, Wainer1970}, defined by transfinite recursion in Fig.~\ref{fig:Hprime}.
+\TODO[Clément]{It seems that the term ``Wainer Hardy'' hierarchy only appears in his paper and the Hydra-battles book}
 From this result, we infer that the considered function is not primitive recursive in general (for $\alpha\geq\omega^\omega$).
@@
 Computability is a key topic in Computer Science teaching.
-Moreover O'Connor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter to this formalisation, with comments on the definitions and proofs, [counter]-examples and exercises.
+Moreover O'Conor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter to this formalisation, with comments on the definitions and proofs, [counter]-examples and exercises.
 
@@
   \item \TODO[Clément]{If you kept a trace of them, I would love to see a discussion of any errors that were caught by including the output systematically instead of copy-pasting it.}
-  \item \TODO[Pierre]{the order of appearance of snippets in the book is independent from the module structure}
-   \item \TODO[Pierre]{Slicing our proofs into snippets allowed to show only the main parts of our proofs (Figure~\ref{fig:Ex42-schutte}).}
 
@@
 It is interesting to note that both libraries \textit{Gaia} and \textit{hydra-battles} contain their own (isomorphic but not identical) definitions of types \texttt{T1} and \texttt{T2}, both originated in~\cite{CantorContrib}.
-Our plan is to build a  bridge between  the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}, and make possible the transfer of theorems from one of the two libraries to the other one.
+Our plan is to build a  ``gate'' \TODO[Clément]{Bridge?} giving access to both the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}, and make possible the transfer of theorems from one of the two libraries to the other one.
 

### 2021-09-24 doc/JFLA2022/paper.tex
fix a few typos
Add citations.

--- 2021-09-24 doc/JFLA2022/paper.tex
@@
 %
-This umbrella project now includes evolved versions of former \emph{contribs} Cantor and Additions (under the new names of Hydra-battles and Addition-chains), the Pocklington library, extracted from Russel O'Connor's Goedel \emph{contrib} and a bridge to the Gaia library (\emph{contrib} of José Grimm).
+This umbrella project now includes evolved versions of former \emph{contribs} Cantor and Additions (under the new names of Hydra-battles and Addition-chains), the Pocklington library, extracted from Russel O'Connor's Goedel \emph{contrib}\cite{OConnor05, Goedel} and a bridge to the Gaia library (\emph{contrib} of José Grimm~\cite{Gaia,grimm:hal-00911710}).
 %
@@
 By using modern maintenance techniques such as continuous integration and deployment, we can ensure that this documentation stays up to date with the latest Coq releases.
-With Alectryon, we ensure that code and documentation are always in sync.
+With Alectryon~\cite{alectryonpaper, alectryongithub}, we ensure that code and documentation are always in sync.
 

### 2021-09-24 doc/JFLA2022/paper.tex
Coq users are forever learners.

--- 2021-09-24 doc/JFLA2022/paper.tex
@@
 
-We call on the Coq users in the JFLA community and beyond to come and join us in this effort, by bringing new interesting projects which are worth presenting to Coq learners and guiding them in their exploration.
+We call on the Coq users in the JFLA community and beyond to come and join us in this effort, by bringing new interesting projects which are worth presenting to Coq learners, \emph{a.k.a.} Coq users, and guiding them in their exploration.
 %

### 2021-09-24 doc/JFLA2022/paper.tex
small corrections
Proposal of a new structure.

@@
 %
-\title{Hydras and Cie  \\
-  Variations on Kirby \& Paris' hydra battles and ordinal numbers in Coq}
+\title{Hydras, Ordinals \&  Co.  \\
+  A library in Coq of entertaining formal mathematics}
 
@@
 \and
-to fill!
+KTH Royal Institute of Technology, Stockholm, Sweden
 \and
@@
 
-\titlerunning{Hydras and Cie}
+\titlerunning{Hydras, Ordinals \& Co.}
 
@@
 
+\subsection{Background}
+
+\begin{itemize}
+\item formalizations of mathematics in proof assistants, such as Coq, continually grow in size and scope
+\begin{itemize}
+\item the Mathematical Components library and the four-color theorem and the odd order theorem in Coq
+\item the Mathlib library for Lean
+\end{itemize}
+\item to continue to be useful as examples and reusable libraries, formalizations must be well \emph{documented} and \emph{maintained} to ensure they work with new proof assistant releases \TODO[Pierre]{add evolution of styles of proof and formalization (not limited to Coq litterature)?
+}
+\item Lean and Mathematical Components have accompanying books, which are written mainly to onboard new library users
+\item book authors faced with several design choices: literate documents, using tools to keep the book and the proof assistant synchronized
+\end{itemize}
+
 \subsection{Vision}
@@
 %
-This umbrella project now includes evolved versions of former \emph{contribs} Cantor and Additions (under the new names of Hydra-battles and Addition-chains), the Pocklington library, extracted from Russel O'Connor's Goedel \emph{contrib}\cite{OConnor05, Goedel} and a bridge to the Gaia library (\emph{contrib} of José Grimm~\cite{Gaia,grimm:hal-00911710}).
+This umbrella project now includes evolved versions of former \emph{contribs} Cantor and Additions (under the new names of Hydra-battles and Addition-chains), the Pocklington library, extracted from Russel O'Connor's Goedel \emph{contrib}~\cite{OConnor05, Goedel} and a bridge to the Gaia library (\emph{contrib} of José Grimm~\cite{Gaia,grimm:hal-00911710}).
 %
@@
 \begin{theorem}
-  There is no termination measure mapping hydras to a segment $[0,\mu)$ where $\mu<\epsilon_0$ for proving the termination of  all hydra battles.\label{thm3}
+  Let $\mu$ be any ordinal strictly less than $\epsilon_0$.
+  There is no function mapping hydras to the segment $[0,\mu)$, which could be used as a measure for  proving the termination of  all hydra battles.\label{thm3}
 \end{theorem}
@@
 Unfortunately, the examples  most commonly shown in the litterature
-(see for instance \cite{KP82, bauer2008, BauerHydra}) 
+(see for instance~\cite{KP82, bauer2008, BauerHydra}) 
 assume that the hydra grows $n$ copies at step $n$ of the game, which is incompatible with our proof.
@@
 
-We now prove that Theorem~\ref{thm3} still holds with these typical battles by borrowing new combinatorial results from~\cite{KS81}.
+We prove now that Theorem~\ref{thm3} still holds with these typical battles by borrowing new combinatorial results from~\cite{KS81}.
 Without loss of generality, we assume the following restrictions:
@@
 
+Using $H'$s equations as rewrite rules, we can study a concrete example of hydra battle. Let us start with the hydra of figure~\ref{fig:start} and $0$ as initial step.
+By a sequence of rewritings and induction proof steps, we prove that the number of steps of the considered battle is greater or equal than $2^{2^N}$, where $N=2^{70}-1$.
+
+\begin{figure}[h]
+  \centering
+  \begin{tikzpicture}[very thick, scale=0.3]
+  \node (h1) at (4,0){$\bullet$};
+  \node (h2) at (2,2){$\bullet$};
+  \node (h3) at (0,4){$\bullet$};
+  \node (h4) at (1,4){$\bullet$};
+  \node (h5) at (2,4){$\bullet$};
+  \node (h6) at (4.5,2){$\bullet$};
+  \node (h7) at (6.5,2){$\bullet$};
+  \node (h8) at (8.5,2){$\bullet$};
+    \draw (h1) -- (h2) ;
+    \draw (h2) -- (h3) ;
+    \draw (h2) -- (h4) ;
+    \draw (h2) -- (h5) ;
+    \draw (h1) -- (h6) ;
+    \draw (h1) -- (h7) ;
+    \draw (h1) -- (h8);
+ \end{tikzpicture}
+
+  \caption{The hydra associated with the ordinal $\omega^3+3$}
+  \label{fig:start}
+\end{figure}
+
 \begin{figure}[h]
@@
 \item If $\alpha=0$, then $H'_\alpha (i)= i$.
+\item If $\alpha$ is a limit ordinal, then 
+  $H'_\alpha(i) = H'_{(\canonseq{\alpha}{i+1})}(i)$.
   \item If $\alpha=\beta+1$, then 
 $H'_\alpha(i)=H'_\beta(i+1)$.
-\item If $\alpha$ is a limit ordinal, then 
-$H'_\alpha(i) = H'_{(\canonseq{\alpha}{i+1})}(i)$.
 \end{itemize}
@@
 
-\section{Main features}
+\section{Integration within Coq-Community}
 
-\subsection{The Coq-Community ecosystem}
-\begin{todo}
-Small presentation.
-\end{todo}
+\subsection{Presentation}
+
+\subsection{Interactions with other Coq-Community projects}
 
@@
 kind of battles is not given by a primitive recursive function, we used the formalization of primitive recursive functions, part
-of Russell O'Connor's contribution on G\"{o}del first incompleteness theorem \cite{OConnor05, Goedel}.
+of Russell O'Connor's contribution on G\"{o}del first incompleteness theorem~\cite{OConnor05, Goedel}.
 For this purpose, and above all by consideration of the scientific interest of this contribution, we chose to host and maintain this work within \textit{Coq-Community}.
@@
 
+% We could already talk a bit of the monorepo structure here, and
+% defer some explanations on the tooling to the next section.
+
+\section{Modernizing the infrastructure}
 \subsection{Documentation with Alectryon}
@@
 
-Our proof is very different from the computational proof of Fig\ref{fig:Ex42E0}.
+Our proof is very different from the computational proof of
+Figure~\ref{fig:Ex42E0}.
 By definition of additive principal ordinals, 
@@
 
+\subsection{Continuous integration and deployment, opam packaging, etc.}
 \section{Conclusion and perspectives}

### 2021-09-24 doc/JFLA2022/paper.tex
and Cie -> and Co. (article title)

@@
 %
-\title{Hydras and Cie  \\
+\title{Hydras and Co.  \\
   Variations on Kirby \& Paris' hydra battles and ordinal numbers in Coq}

### 2021-09-25 doc/JFLA2022/paper.tex
and -> ampersand (atricle)
new title
fill in affiliation
small change

@@
 %
-\title{Hydras and Co.  \\
-  Variations on Kirby \& Paris' hydra battles and ordinal numbers in Coq}
+\title{Hydras, Ordinals \&  Co.  \\
+  A library in Coq of entertaining formal mathematics}
 
@@
 
+ \TODO[Pierre]{Using equations for $H'$ as rewrite rules, we could prove nice minorations: for instance, the number $H'_{\omega^3}(3)$ has at least $2^{10^{21}}$ binary digits.}
 \section{Main features}

### 2021-09-25 doc/JFLA2022/paper.tex
paper background outline

@@
 \label{sect:introduction}
+
+\subsection{Background}
+
+\begin{itemize}
+\item formalizations of mathematics in proof assistants, such as Coq, continually grow in size and scope
+\begin{itemize}
+\item the Mathematical Components library and the four-color theorem and the odd order theorem in Coq
+\item the Mathlib library for Lean
+\end{itemize}
+\item to continue to be useful as examples and reusable libraries, formalizations must be well \emph{documented} and \emph{maintained} to ensure they work with new proof assistant releases
+\item Lean and Mathematical Components have accompanying books, which are written mainly to onboard new library users
+\item book authors faced with several design choices: literate documents, using tools to keep the book and the proof assistant synchronized
+\end{itemize}
 

### 2021-09-25 doc/part-hydras.tex
Wainer-Hardy -> Hardy

@@
 
-Definition CompSpec {A} (eq lt : A->A->Prop)(x y:A) : comparison -> Prop :=
+Definition CompSpec {A} (eq lt : A -> A -> Prop) (x y : A) :=
  CompareSpec (eq x y) (lt x y) (lt y x).
@@
 
-\section{A variant of the Wainer-Hardy hierarchy}
+\section{A variant of the Hardy hierarchy}
 
@@
 
-In order to give a feeling on  the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the  \emph{Wainer-Hardy hierarchy} of fast growing functions,
+In order to give a feeling on  the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the  \emph{Hardy hierarchy} of fast growing functions,
 presented for instance in~\cite{Promel2013}. 
@@
 \begin{remark}
- The ``classic'' definition of the Wainer-Hardy hierarchy differs in the third equation.
+ The ``classic'' definition of the Hardy hierarchy differs in the third equation.
 

### 2021-09-25 doc/JFLA2022/paper.tex
add a remark about the number of steps of a battle

@@
 
- \TODO[Pierre]{Using equations for $H'$ as rewrite rules, we could prove nice minorations: for instance, the number $H'_{\omega^3}(3)$ has at least $2^{10^{21}}$ binary digits.}
+\TODO[Pierre]{
+Using the equations for $H'$ as rewrite rules, we could ``compute'' the number of rounds of the battle starting with the hydra of Figure~\ref{fig:start} and $0$ as initial step. this number has
+at least $2^{10^{21}}$ binary digits. Obtaining this result is an interesting example of relationship between proof and computation.}
+
+\begin{figure}[h]
+  \centering
+  \begin{tikzpicture}[very thick, scale=0.3]
+  \node (h1) at (4,0){$\bullet$};
+  \node (h2) at (2,2){$\bullet$};
+  \node (h3) at (0,4){$\bullet$};
+  \node (h4) at (1,4){$\bullet$};
+  \node (h5) at (2,4){$\bullet$};
+  \node (h6) at (4.5,2){$\bullet$};
+  \node (h7) at (6.5,2){$\bullet$};
+  \node (h8) at (8.5,2){$\bullet$};
+    \draw (h1) -- (h2) ;
+    \draw (h2) -- (h3) ;
+    \draw (h2) -- (h4) ;
+    \draw (h2) -- (h5) ;
+    \draw (h1) -- (h6) ;
+    \draw (h1) -- (h7) ;
+    \draw (h1) -- (h8);
+ \end{tikzpicture}
+
+  \caption{The hydra corresponding to the ordinal $\omega^3+3$}
+  \label{fig:start}
+\end{figure}
+
 \section{Main features}

### 2021-09-25 doc/JFLA2022/paper.tex
fix typos
small correction

@@
 \TODO[Pierre]{
-Using the equations for $H'$ as rewrite rules, we could ``compute'' the number of rounds of the battle starting with the hydra of Figure~\ref{fig:start} and $0$ as initial step. this number has
+  Using the equations of $H'$ as rewrite rules, we studied the
+  number of steps  of the battle starting with the hydra of Figure~\ref{fig:start} and $0$ as initial step. This number has
 at least $2^{10^{21}}$ binary digits. Obtaining this result is an interesting example of relationship between proof and computation.}

### 2021-09-26 doc/part-hydras.tex
add a small remark
minor change in text

@@
 \subsubsection{A numerical example}
+\label{sect:bignum}
 
@@
 \input{movies/snippets/Hydra_Theorems/battleLengthStdHardy}
+
+Since $H'_{\omega^3}(3)= H'_{\omega^3+3}(0)$, the big number shown in Section~\vref{sect:bignum} is less or equal than 
+the number of steps of the battle with the initial hydra of
+Figure~\ref{fig:start}, and  $0$ as initial replication factor.
+
+\begin{figure}[h]
+  \centering
+  \begin{tikzpicture}[very thick, scale=0.3]
+  \node (h1) at (4,0){$\bullet$};
+  \node (h2) at (2,2){$\bullet$};
+  \node (h3) at (0,4){$\bullet$};
+  \node (h4) at (1,4){$\bullet$};
+  \node (h5) at (2,4){$\bullet$};
+  \node (h6) at (4.5,2){$\bullet$};
+  \node (h7) at (6.5,2){$\bullet$};
+  \node (h8) at (8.5,2){$\bullet$};
+    \draw (h1) -- (h2) ;
+    \draw (h2) -- (h3) ;
+    \draw (h2) -- (h4) ;
+    \draw (h2) -- (h5) ;
+    \draw (h1) -- (h6) ;
+    \draw (h1) -- (h7) ;
+    \draw (h1) -- (h8);
+ \end{tikzpicture}
+
+  \caption{The hydra corresponding to the ordinal $\omega^3+3$}
+  \label{fig:start}
+\end{figure}
 

### 2021-09-26 doc/JFLA2022/paper.tex
changes in the concrete example (end of Section 2)

@@
 
-\TODO[Pierre]{
-  Using the equations of $H'$ as rewrite rules, we studied the
-  number of steps  of the battle starting with the hydra of Figure~\ref{fig:start} and $0$ as initial step. This number has
-at least $2^{10^{21}}$ binary digits. Obtaining this result is an interesting example of relationship between proof and computation.}
+Using $H'$ equations as rewrite rules, we can study a concrete example of hydra battle. Let us start with the hydra of figure~\ref{fig:start} and $0$ as initial step.
+By a combination of rewriting and induction, we prove that the number of steps of the considered battle has at least $2^{10^{21}}$ binary digits.
 

### 2021-09-27 doc/JFLA2022/paper.tex
minor corrections; add author field for the HOC book
add a TODO in the Background section

@@
 \end{itemize}
-\item to continue to be useful as examples and reusable libraries, formalizations must be well \emph{documented} and \emph{maintained} to ensure they work with new proof assistant releases
+\item to continue to be useful as examples and reusable libraries, formalizations must be well \emph{documented} and \emph{maintained} to ensure they work with new proof assistant releases \TODO[Pierre]{add evolution of styles of proof and formalization (not limited to Coq litterature)?
+}
 \item Lean and Mathematical Components have accompanying books, which are written mainly to onboard new library users
@@
 
-\begin{figure}[h]
- \noindent
- For any $\alpha$ in $[0,\epsilon_0)$ and any $i$ in $\mathbb{N}$,
-   \begin{itemize}
-\item If $\alpha=0$, then $H'_\alpha (i)= i$.
-\item If $\alpha=\beta+1$, then 
-$H'_\alpha(i)=H'_\beta(i+1)$.
-\item If $\alpha$ is a limit ordinal, then 
-$H'_\alpha(i) = H'_{(\canonseq{\alpha}{i+1})}(i)$.
-\end{itemize}
-\vspace{4pt}
-\inputsnippets{HprimeDef}
-\caption{The rapidly growing hierarchy $H'$ of arithmetic functions}
-\label{fig:Hprime}
-\end{figure}
-
-\label{sect:not-pr}
-
-\TODO[Clément]{Highlight `Equations' and remove `omega' highlighting; change theme to something less garish}
-
-Using $H'$ equations as rewrite rules, we can study a concrete example of hydra battle. Let us start with the hydra of figure~\ref{fig:start} and $0$ as initial step.
+Using $H'$s equations as rewrite rules, we can study a concrete example of hydra battle. Let us start with the hydra of figure~\ref{fig:start} and $0$ as initial step.
 By a combination of rewriting and induction, we prove that the number of steps of the considered battle has at least $2^{10^{21}}$ binary digits.
@@
 
-  \caption{The hydra corresponding to the ordinal $\omega^3+3$}
+  \caption{The hydra associated with the ordinal $\omega^3+3$}
   \label{fig:start}
 \end{figure}
+
+\begin{figure}[h]
+ \noindent
+ For any $\alpha$ in $[0,\epsilon_0)$ and any $i$ in $\mathbb{N}$,
+   \begin{itemize}
+\item If $\alpha=0$, then $H'_\alpha (i)= i$.
+\item If $\alpha$ is a limit ordinal, then 
+  $H'_\alpha(i) = H'_{(\canonseq{\alpha}{i+1})}(i)$.
+  \item If $\alpha=\beta+1$, then 
+$H'_\alpha(i)=H'_\beta(i+1)$.
+\end{itemize}
+\vspace{4pt}
+\inputsnippets{HprimeDef}
+\caption{The rapidly growing hierarchy $H'$ of arithmetic functions}
+\label{fig:Hprime}
+\end{figure}
+
+\label{sect:not-pr}
+
+\TODO[Clément]{Highlight `Equations' and remove `omega' highlighting; change theme to something less garish}
 

### 2021-09-27 doc/JFLA2022/paper.tex
more explanations (minoration of the length of a battle)

@@
 Using $H'$s equations as rewrite rules, we can study a concrete example of hydra battle. Let us start with the hydra of figure~\ref{fig:start} and $0$ as initial step.
-By a combination of rewriting and induction, we prove that the number of steps of the considered battle has at least $2^{10^{21}}$ binary digits.
+By a sequence of rewritings and induction proof steps, we prove that the number of steps of the considered battle is greater or equal than $2^{2^N}$, where $N=2^{70}-1$.
 

### 2021-09-27 doc/JFLA2022/paper.tex
improve (?) Section 2

--- 2021-09-27 doc/JFLA2022/paper.tex
@@
   \emph{Hydra-battles} is a collaborative library of commented proofs, developped under the \emph{Coq-Community} project.
-  We present its evolution since a former presentation~\cite{JFLA2018paper} and how its structure is determined by other
+  We present its evolution since a former presentation and how its structure is determined by other
   libraries and the \emph{Alectryon} literate proving tool.
@@
 Let $\alpha<\epsilon_0$ be an ordinal. 
-Using the \texttt{coq-equations} plug-in~\cite{sozeau:hal-01671777}, we define the function which returns the number of steps of the battle starting with $\alpha$ at step $i$, for  any $i$ in $\mathbb{N}$.
-We prove that this number is greater or equal than
-$H'_\alpha(i)-i$, where $H'$ is a slight variant of the Hardy hierarchy of rapidly growing functions~\cite{BW85, KS81, Promel2013, Wainer1970}, defined by transfinite recursion in Fig.~\ref{fig:Hprime}.
-From this result, we infer that the considered function is not primitive recursive in general (for $\alpha\geq\omega^\omega$).
+We prove that  the number of steps of the battle starting with
+$\alpha$ at step $i$ is greater or equal than
+$H'_\alpha(i)-i$, where $H'$ is a slight variant of the Hardy hierarchy of rapidly growing functions~\cite{BW85, KS81, Promel2013, Wainer1970}.
+The function $H'_\alpha$ is defined by transfinite recursion over $\alpha$:
 
-Using $H'$s equations as rewrite rules, we can study a concrete example of hydra battle. Let us start with the hydra of figure~\ref{fig:start} and $0$ as initial step.
-By a combination of rewriting and induction, we prove that the number of steps of the considered battle has at least $2^{10^{21}}$ binary digits.
+\begin{align}
+  H'_0(i) & = i\\
+  H'_\alpha(i) &= H'_{(\canonseq{\alpha}{i+1})}(i)  \quad\textit{if $\alpha$ is a limit ordinal}\\
+  H'_{\alpha}(i) &=H'_\beta(i+1) \quad\textit{if $\alpha=\beta+1$}
+\end{align}
+
+Figure~\ref{fig:Hprime} 
+ shows the  definition of $H'$ with the \texttt{coq-equations}
+ plug-in~\cite{sozeau:hal-01671777}.
+
+ \begin{figure}[h]
+ \inputsnippets{HprimeDef}
+\caption{Coq Equations for $H'$}
+\label{fig:Hprime}
+\end{figure}
+
+Using $H'$s equations as rewrite rules, we can study a realistic example. We take the hydra of figure~\ref{fig:start} and $i=0$ as initial configuration. 
+By a sequence of rewritings and inductions, we prove that the number of steps of the considered battle is greater or equal than $2^{2^N}$, where $N=2^{70}-1$.
 
@@
 
-\begin{figure}[h]
- \noindent
- For any $\alpha$ in $[0,\epsilon_0)$ and any $i$ in $\mathbb{N}$,
-   \begin{itemize}
-\item If $\alpha=0$, then $H'_\alpha (i)= i$.
-\item If $\alpha$ is a limit ordinal, then 
-  $H'_\alpha(i) = H'_{(\canonseq{\alpha}{i+1})}(i)$.
-  \item If $\alpha=\beta+1$, then 
-$H'_\alpha(i)=H'_\beta(i+1)$.
-\end{itemize}
-\vspace{4pt}
-\inputsnippets{HprimeDef}
-\caption{The rapidly growing hierarchy $H'$ of arithmetic functions}
-\label{fig:Hprime}
-\end{figure}
-
 \label{sect:not-pr}
@@
 \TODO[Clément]{Highlight `Equations' and remove `omega' highlighting; change theme to something less garish}
+
+More generally, we prove that, for $\alpha\geq\omega^\omega$
+the function computing the length of the battle starting with the configuration $(\alpha,i)$ is not primitive recursive.
 
@@
 Computability is a key topic in Computer Science teaching.
-Moreover O'Connor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter to this formalisation, with comments on the definitions and proofs, [counter]-examples and exercises.
+Moreover O'Connor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter to this formalization, with comments on the definitions and proofs, [counter]-examples and exercises.
 

### 2021-09-27 doc/JFLA2022/paper.tex
less verbose paragraph on Gaia Bridge

@@
 
+\subsection{Continuous integration and deployment, opam packaging, etc.}
 \section{Conclusion and perspectives}
+
+\TODO{General Conclusion}
 
@@
 
+\subsection{Further extensions}
+
+We plan to extend our libraies in two main directions.
 The Gaia project~\cite{Gaia}, also maintained on \textit{Coq-Community} contains a development by José Grimm~\cite{grimm:hal-00911710}.
-This library -- dedicated  to the implementation in \textit{Coq (SSreflect)} of books from  Bourbaki's Elements of Mathematics --- contains a generalization of two types of ordinals (\texttt{T1} for the ordinal $\epsilon_0$ and
-\texttt{T2} for $\Gamma_0$) to a larger ordinal defined by Ackermann, and relates these data structures with Bourbaki's set theory.
-It is interesting to note that both libraries \textit{Gaia} and \textit{hydra-battles} contain their own (isomorphic but not identical) definitions of types \texttt{T1} and \texttt{T2}, both originated in~\cite{CantorContrib}.
-Our plan is to build a  bridge between  the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}, and make possible the transfer of theorems from one of the two libraries to the other one.
-
-Another direction would be to write in \textit{Coq} a formal proof of the original statement of Theorem~\ref{kp:thm2}.
-For this purpose, we plan to import  the formalization of Peano Arithmetic by O'Connor~\cite{Goedel}.
+This library is dedicated  to the implementation in \textit{Coq (SSreflect)} of books from  Bourbaki's Elements of Mathematics.
+It contains data structures which are compatible with our ordinal notations.
+Our plan is to build a  bridge between  the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}, and make possible the transfer of theorems between both libraries.
+Another direction would be to write in \textit{Coq} a formal proof of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
 

### 2021-09-28 doc/JFLA2022/paper.tex
text simplification in Section 2

@@
 $H'_\alpha(i)-i$, where $H'$ is a slight variant of the Hardy hierarchy of rapidly growing functions~\cite{BW85, KS81, Promel2013, Wainer1970}.
-The function $H'_\alpha$ is defined by transfinite recursion over $\alpha$:
+The function $H'_\alpha$ is defined by transfinite recursion over $\alpha$ on Figures~\ref{fig:hardy-math}
+and~\ref{fig:Hprime}.
 
+\begin{figure}[h]
 \begin{align}
@@
 \end{align}  
-
-Figure~\ref{fig:Hprime} 
- shows the  definition of $H'$ with the \texttt{coq-equations}
- plug-in~\cite{sozeau:hal-01671777}.
+  \caption{The $H'$ rapidly growing hierarchy of arithmetical functions}
+  \label{fig:hardy-math}
+\end{figure}
 
@@
  \inputsnippets{HprimeDef}
-\caption{Coq Equations for $H'$}
+\caption{$H'$ definition with the \texttt{coq-equations}
+ plug-in~\cite{sozeau:hal-01671777}}
 \label{fig:Hprime}
@@
 
-We plan to extend our libraies in two main directions.
+We plan to extend our libraries in two main directions.
 The Gaia project~\cite{Gaia}, also maintained on \textit{Coq-Community} contains a development by José Grimm~\cite{grimm:hal-00911710}.

### 2021-09-28 doc/JFLA2022/paper.tex
minor change

@@
   \item Despite the frequent changes (improvements) in the Coq scripts, code inclusion and \textit{Coq}'s answers are up to date.
-
-  \item Presentation of the main steps of proofs, possibly interleaved with text.
   \item \TODO[Clément]{A good part of the LaTeX support in Alectryon was developed for hydra-battles, and this was the first real application with a custom driver, too.}
@@
   \item \TODO[Pierre]{the order of appearance of snippets in the book is independent from the module structure}
-   \item \TODO[Pierre]{Slicing our proofs into snippets allowed to show only the main parts of our proofs (Figure~\ref{fig:Ex42-schutte}).}
+   \item \TODO[Pierre]{Slicing our proofs into snippets allowed to show only the main parts of our proofsleaved with text (Figure~\ref{fig:Ex42-schutte}).}
 

### 2021-09-29 doc/JFLA2022/paper.tex
changes to the abstract

@@
 \begin{abstract}
-  \emph{Hydra-battles} is a collaborative library of commented proofs, developped under the \emph{Coq-Community} project.
-  We present its evolution since a former presentation and how its structure is determined by other
-  libraries and the \emph{Alectryon} literate proving tool.
+  \emph{Hydra-battles} is a collaborative library on discrete math, written for the Coq proof assistant, and developped under the \emph{Coq-Community} project. The proof scripts are
+  accompanied with an electronic book, made with the help of the \emph{Alectryon} litterate proving tool.
+  We present the evolution of the mathematical contents since
+  former presentations at JFLA meetings.
+  Then, we show how the structure of the project is determined   by two  requirements which must continuously satisfied:
+   maintenance of the library with respect to several Coq-Community projects it requires and evolutions of Coq, and 
+  consistency of the book with respect to the library. 
 \end{abstract}

### 2021-09-30 doc/JFLA2022/paper.tex
Typo fix and uniformize Coq-community casing.

@@
 \begin{abstract}
-  \emph{Hydra-battles} is a collaborative library on discrete math, written for the Coq proof assistant, and developped under the \emph{Coq-Community} project. The proof scripts are
+  \emph{Hydra-battles} is a collaborative library on discrete math, written for the Coq proof assistant, and developed under the \emph{Coq-community} project. The proof scripts are
   accompanied with an electronic book, made with the help of the \emph{Alectryon} litterate proving tool.
@@
   Then, we show how the structure of the project is determined   by two  requirements which must continuously satisfied:
-   maintenance of the library with respect to several Coq-Community projects it requires and evolutions of Coq, and 
+   maintenance of the library with respect to several Coq-community projects it requires and evolutions of Coq, and 
   consistency of the book with respect to the library. 
@@
 \subsection{Vision}
-The \emph{Hydra-battles} project, part of \emph{Coq-Community}~\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
-\emph{Coq-Community} is a community organization that we have founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting Coq packages, and working collaboratively on documentation projects.
+The \emph{Hydra-battles} project, part of \emph{Coq-community}~\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
+\emph{Coq-community} is a community organization that we have founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting Coq packages, and working collaboratively on documentation projects.
 The \emph{Hydra-battles} project demonstrates that these two goals are not independent: interesting Coq packages can become the basis for new documentation.
@@
  JFLA~\cite{PCiota, JFLA2018paper}.
-We present recent evolutions of the library: new results, interaction within the Coq-Community project~\cite{CoqCommunity}, and documentation generated with Alectryon~\cite{alectryonpaper, alectryongithub}.
+We present recent evolutions of the library: new results, interaction within the Coq-community project~\cite{CoqCommunity}, and documentation generated with Alectryon~\cite{alectryonpaper, alectryongithub}.
 
@@
 
-\section{Integration within Coq-Community}
+\section{Integration within Coq-community}
 
@@
 
-\subsection{Interactions with other Coq-Community projects}
+\subsection{Interactions with other Coq-community projects}
 
@@
 of Russell O'Connor's contribution on G\"{o}del first incompleteness theorem~\cite{OConnor05, Goedel}.
-For this purpose, and above all by consideration of the scientific interest of this contribution, we chose to host and maintain this work within \textit{Coq-Community}.
+For this purpose, and above all by consideration of the scientific interest of this contribution, we chose to host and maintain this work within \textit{Coq-community}.
 
@@
 We plan to extend our libraries in two main directions.
-The Gaia project~\cite{Gaia}, also maintained on \textit{Coq-Community} contains a development by José Grimm~\cite{grimm:hal-00911710}.
+The Gaia project~\cite{Gaia}, also maintained on \textit{Coq-community} contains a development by José Grimm~\cite{grimm:hal-00911710}.
 This library is dedicated  to the implementation in \textit{Coq (SSreflect)} of books from  Bourbaki's Elements of Mathematics.

### 2021-09-30 doc/JFLA2022/paper.tex
First stab at presenting Coq-community.

@@
 
+\emph{Coq-community} is an informal organization run by volunteer Coq users with objectives to maintain interesting Coq projects on the long-term and help Coq users to collaborate on documentation, tooling, etc.
+%
+It was created in 2018, inspired by the \emph{Elm Community} organization~\cite{zimmermann:tel-02451322}, and in fact such ``Community Package Maintenance Organizations'' exist in many ecosystems~\cite{zimmermann2021grounded}.
+%
+Today, it hosts over 50 projects maintained by over 30 maintainers.
+
 \subsection{Interactions with other Coq-community projects}

### 2021-10-01 doc/JFLA2022/paper.tex
Remove the word contrib.

@@
 %
-This umbrella project now includes evolved versions of former \emph{contribs} Cantor and Additions (under the new names of Hydra-battles and Addition-chains), the Pocklington library, extracted from Russel O'Connor's Goedel \emph{contrib}~\cite{OConnor05, Goedel} and a bridge to the Gaia library (\emph{contrib} of José Grimm~\cite{Gaia,grimm:hal-00911710}).
+This umbrella project now includes evolved versions of the former Cantor and Additions libraries (under the new names of Hydra-battles and Addition-chains), the Pocklington library, extracted from Russel O'Connor's Goedel library~\cite{OConnor05, Goedel} and a bridge to the Gaia library (by José Grimm~\cite{Gaia,grimm:hal-00911710}).
 %

### 2021-10-01 doc/JFLA2022/paper.tex
Improve / expand presentation of Coq-community.

@@
 
-\emph{Coq-community} is an informal organization run by volunteer Coq users with objectives to maintain interesting Coq projects on the long-term and help Coq users to collaborate on documentation, tooling, etc.
+\emph{Coq-community} is an informal organization run by volunteer Coq users with objectives to maintain interesting Coq projects in the long-term and help Coq users to collaborate on documentation, tooling, etc.
+
+It was created in 2018, inspired by the \emph{Elm Community} organization~\cite{zimmermann:tel-02451322}.
 %
-It was created in 2018, inspired by the \emph{Elm Community} organization~\cite{zimmermann:tel-02451322}, and in fact such ``Community Package Maintenance Organizations'' exist in many ecosystems~\cite{zimmermann2021grounded}.
+Such ``Community Package Maintenance Organizations'' actually exist in many ecosystems as they avoid the very common problem of an important package becoming unmaintained as its author has moved on or disappeared~\cite{zimmermann2021grounded}.
+
+In the case of Coq, this problem may be even more prevalent as many packages are created by PhD students, or by researchers for a specific paper, but their authors do not intend to maintain the package in the long-term.
 %
-Today, it hosts over 50 projects maintained by over 30 maintainers.
+The authors are however generally very open to having someone else who expresses interest in their work continue the maintenance.
+%
+\emph{Coq-community} makes that easy by providing a process for transferring or forking an unmaintained package, tooling for setting up good maintenance practices (such as continuous integration) and by making it possible for someone to take over a package without committing in the long-term (as maintainers who drop out can easily be replaced by some other volunteers).
+%
+Today, \emph{Coq-community} hosts over 50 projects maintained by over 30 maintainers.
 

### 2021-10-01 doc/JFLA2022/paper.tex
More on the origin of the projects.

@@
 \emph{Coq-community} makes that easy by providing a process for transferring or forking an unmaintained package, tooling for setting up good maintenance practices (such as continuous integration) and by making it possible for someone to take over a package without committing in the long-term (as maintainers who drop out can easily be replaced by some other volunteers).
+
+Today, \emph{Coq-community} hosts over 50 projects maintained by over 30 maintainers.
 %
-Today, \emph{Coq-community} hosts over 50 projects maintained by over 30 maintainers.
+The hosted projects come from a variety of origins.
+%
+Some had been maintained in the past by the Coq development team on behalf of the authors, but this meant that only minimal changes required to make the project build with new Coq versions were conducted.
+%
+Some were still maintained by their authors but they wanted help to keep them up-to-date.
+%
+Others were simply left unmaintained and were revived by their adoption in \emph{Coq-community}.
 
 \subsection{Interactions with other Coq-community projects}
+
+Given the objectives of \emph{Coq-community}, it makes sense to propose a transfer anytime we encounter an interesting project that is insufficiently maintained.
+%
+After the project is adopted, large changes and refactorings are explicitly allowed.
+%
+This means, for instance, that we can consolidate libraries together or split them into several packages if it makes sense.
 

### 2021-10-01 doc/JFLA2022/paper.tex
Fix description of part that was extracted from Goedel library.

@@
 %
-This umbrella project now includes evolved versions of the former Cantor and Additions libraries (under the new names of Hydra-battles and Addition-chains), the Pocklington library, extracted from Russel O'Connor's Goedel library~\cite{OConnor05, Goedel} and a bridge to the Gaia library (by José Grimm~\cite{Gaia,grimm:hal-00911710}).
+This umbrella project now includes evolved versions of the former Cantor and Additions libraries (under the new names of Hydra-battles and Addition-chains), the Ackerman sub-library, extracted from Russel O'Connor's Goedel library~\cite{OConnor05, Goedel} and a bridge to the Gaia library (by José Grimm~\cite{Gaia,grimm:hal-00911710}).
 %
@@
 
+We have extracted part of the G\"{o}del library into a new sub-library of the Hydra-battles library: Ackermann.
+
 % We could already talk a bit of the monorepo structure here, and

### 2021-10-01 doc/JFLA2022/paper.tex
Typo.

@@
 %
-This umbrella project now includes evolved versions of the former Cantor and Additions libraries (under the new names of Hydra-battles and Addition-chains), the Ackerman sub-library, extracted from Russel O'Connor's Goedel library~\cite{OConnor05, Goedel} and a bridge to the Gaia library (by José Grimm~\cite{Gaia,grimm:hal-00911710}).
+This umbrella project now includes evolved versions of the former Cantor and Additions libraries (under the new names of Hydra-battles and Addition-chains), the Ackermann sub-library, extracted from Russel O'Connor's Goedel library~\cite{OConnor05, Goedel} and a bridge to the Gaia library (by José Grimm~\cite{Gaia,grimm:hal-00911710}).
 %

### 2021-10-02 doc/JFLA2022/paper.tex
Correctly spell the name of the Goedel library.
work on Conclusion section

@@
 
-We have extracted part of the G\"{o}del library into a new sub-library of the Hydra-battles library: Ackermann.
+We have extracted part of the Goedel library into a new sub-library of the Hydra-battles library: Ackermann.
 
@@
 
-\TODO{General Conclusion}
+\emph{Hydra-battles} wants to be a bridge between scientific litterature (\emph{e.g.}~\cite{KP82, KS81, schutte} and proof assistant technology.
+For the mathematician, it may give a concrete view of the mathematical content, not only through full proofs, but also through significative computations: examples, functions associated with constructive proofs, etc.
+For the Coq user, it provides a consistent set of examples,
+allowing to present and compare various formalization and proving techniques. It may be also used as a library, through opam packages.
+
+It is also a medium sized library (more than 50K LOC), dependent on various tools and libraries of the Coq ecosystem.
+It may be also used to experiment new techniques of continuous maintenace of the code and its documentation.
+
+%\TODO{General Conclusion}
 
@@
 
-\subsection{Further extensions}
+%\subsection{Further extensions}
 

### 2021-10-02 doc/JFLA2022/paper.tex
jfla: Move snippets to a separate directory
jfla: Build snippets separately
jfla: Rebuild snippets
jfla: Minor tweaks

@@
 %%% for movies by alectryon
-\usepackage{../movies/snippets/assets/alectryon}
-\usepackage{../movies/snippets/assets/pygments}
+\usepackage{./assets/alectryon}
+\usepackage{./assets/pygments}
 %%% One hypothesis per line 
@@
 %%% \snippets{A,B,C,…} inputs a series of snippets as one block (with \itemsep
-%%% between them).  A, B, C should be paths to files in movies/snippets/.
+%%% between them).  A, B, C should be paths to files in snippets/.
 \usepackage{etoolbox}
@@
   {{\renewenvironment{alectryon}{}{}% Skip \begin{alectryon} included in snippet
-    \input@old{#1}}}
+    \input@old{snippets/#1}}}
 \makeatother
@@
   Let $\mu$ be any ordinal strictly less than $\epsilon_0$.
-  There is no function mapping hydras to the segment $[0,\mu)$, which could be used as a measure for  proving the termination of  all hydra battles.\label{thm3}
+  There is no function mapping hydras to the segment $[0,\mu)$ that could be used as a measure for  proving the termination of  all hydra battles.\label{thm3}
 \end{theorem}
 
-\TODO[Clément]{Is there a word missing above?}
 Considering measures applicable to \emph{all} battles allowed us to
@@
  \begin{figure}[h]
- \inputsnippets{HprimeDef}
+ \inputsnippets{Hprime/HprimeDef}
 \caption{$H'$ definition with the \texttt{coq-equations}
@@
 
-\TODO[Clément]{Highlight `Equations' and remove `omega' highlighting; change theme to something less garish}
+\TODO[Clément]{Could we use a font with ligatures?}
 
@@
       \begin{minipage}[h]{1.0\linewidth}
-        \inputsnippets{Ex42}
+        \inputsnippets{E0/Ex42}
       \end{minipage}}
@@
 
-\inputsnippets{Ex42a}
+\inputsnippets{Schutte/Ex42a}
 
@@
 
-\inputsnippets{Ex42b}
+\inputsnippets{Schutte/Ex42b}
 
@@
 
-\inputsnippets{Ex42d, Ex42c, Ex42e}
+\inputsnippets{Schutte/Ex42d, Schutte/Ex42c, Schutte/Ex42e}
 

### 2021-10-04 doc/JFLA2022/paper.tex

@@
 \section{Modernizing the infrastructure}
+
 \subsection{Documentation with Alectryon}
+
+The hydra-battles book is written in LaTeX, but it makes very frequent references (about once per page, 274 snippets over 281 pages) to parts of the Coq development, showing definitions (Fig.~\ref{fig:Hprime}), lemmas (Fig.~\ref{fig:Ex42E0}), proof scripts (Fig.~\ref{fig:Ex42-schutte}), and computation results (Fig.~\ref{fig:Pow-evalPow17LetIn}).
+
+Originally, we copied snippets from Coq sources into LaTeX manually, and recorded and inserted the corresponding outputs manually as well.
+This approach is common, but brittle: changes to Coq definitions or lemmas had to be reflected in the book's sources, and we found multiple instances where the book and the Coq development had diverged (see FIXME cite alectryon for a discussion of various approaches to documentation of Coq developments).
+
+We solved this maintenance issue by moving to Alectryon, a tool to automatically record Coq proofs.
+Instead of copy-pasting fragments into LaTeX, we now embed small LaTeX files automatically generated by Alectryon from our Coq development.
+Our build system guarantees that these LaTeX snippets are always up-to-date and consistent with the code (and, by comparing these snippets across releases of Coq or versions of hydra-battles, we can easily spot unexpected changes).
+
+The transition happened gradually, over a few months: for each snippet of Coq code that was in the book, we had to take the following steps:
+
+\begin{enumerate}
+\item Mark the snippet in the Coq sources (we use special comments \texttt{(* begin snippet \textit{name} *)} … \texttt{(* end snippet \textit{name} *)})
+\item Configure output display, using special Alectryon annotations to configure what should be shown (only the inputs, inputs and outputs, some steps of the proof but not all, some proof states at key moments in a proof, etc.)
+\item Replace the copy-pasted inputs and output in LaTeX with an \texttt{input} command.
+\end{enumerate}
+
+Importing snippets into a LaTeX document was not one of the original use cases of Alectryon, so we had to extend it: the original Alectryon did not have support for exporting to LaTeX, and it was towards documenting individual source files (a style where code and prose are interleaved within the same file and code is documented in the same order as it is compiled).
+We extended Alectryon to support our needs by implementing a LaTeX backend, and by programming it to generate individual snippet files, one per \texttt{snippet} comment block.
+The later part was straightforward: all it took was to build a custom Alectryon \textit{driver}, a small (about 100 lines) Python program that leverages most of the Alectryon toolchain but defines a custom frontend that understands our snippet annotations (and otherwise exposes the exact same command line as Alectryon).
+
 \begin{todo}
@@
   \begin{itemize}
-  \item Despite the frequent changes (improvements) in the Coq scripts, code inclusion and \textit{Coq}'s answers are up to date.
-  \item \TODO[Clément]{A good part of the LaTeX support in Alectryon was developed for hydra-battles, and this was the first real application with a custom driver, too.}
-  \item \TODO[Clément]{If you kept a trace of them, I would love to see a discussion of any errors that were caught by including the output systematically instead of copy-pasting it.}
   \item \TODO[Pierre]{the order of appearance of snippets in the book is independent from the module structure}
@@
 
-   Figures~\ref{fig:Hprime} (code snippet), \ref{fig:Ex42E0} and \ref{fig:Ex42-schutte} have been generated with Alectryon.
 \end{todo}
@@
   \label{fig:Ex42-schutte}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \inputsnippets{Pow/evalPow17LetIn}
+  \caption{\label{fig:Pow-evalPow17LetIn}Automatically capturing the output of computations}
 \end{figure}

### 2021-10-04 doc/JFLA2022/snippets/E0/Ex42.tex

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Example}~\PY{n+nf}{Ex42}\PY{o}{:}~\PY{o}{(}\PY{n}{omega}~\PY{o}{+}~\PY{l+m+mi}{42}~\PY{o}{+}~\PY{n}{omega}\PY{o}{\PYZca{}}~\PY{l+m+mi}{2}~\PY{o}{=}~\PY{n}{omega}\PY{o}{\PYZca{}}~\PY{l+m+mi}{2}\PY{o}{)\PYZpc{}}\PY{n}{e0}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Proof}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~\PY{n+nb}{rewrite}~\PY{o}{\PYZlt{}\PYZhy{}}~~\PY{n}{Comparable}\PY{o}{.}\PY{n}{compare\PYZus{}eq\PYZus{}iff}\PY{o}{.}
    \end{input}
    \sep
    \begin{output}
      \begin{goals}
        \begin{goal}
          \begin{hyps}\end{hyps}
          \sep
          \infrule{}
          \sep
          \begin{conclusion}
            \PY{n+nb}{compare}~\PY{o}{(}\PY{n}{omega}~\PY{o}{+}~\PY{l+m+mi}{42}~\PY{o}{+}~\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{)}~\PY{o}{(}\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{)}~\PY{o}{=}~\PY{n}{Eq}
          \end{conclusion}
        \end{goal}
      \end{goals}
    \end{output}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~\PY{n+nb+bp}{reflexivity}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Qed}\PY{o}{.}
    \end{input}
  \end{sentence}
\end{alectryon}

### 2021-10-04 doc/JFLA2022/snippets/Schutte/Ex42c.tex

--- 2021-09-20 doc/JFLA2022/Ex42b.tex
@@
     \begin{input}
-      ~~\PY{n+nb}{assert}~\PY{o}{(}\PY{n}{HAP}\PY{o}{:}~\PY{n}{In}~\PY{n}{AP}~\PY{o}{(}\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{))}~\PY{n+nb+bp}{by}~\PY{n+nb}{apply}~\PY{n}{AP\PYZus{}phi0}\PY{o}{.}\nl
-    \end{input}
-  \end{sentence}
-  \sep
-  \begin{sentence}
-    \begin{input}
-      ~~\PY{n+nb}{elim}~~\PY{n}{HAP}\PY{o}{;}~\PY{n+nb}{intros}~~\PY{n}{\PYZus{}}~\PY{n}{H0}\PY{o}{;}~\PY{n+nb}{apply}~\PY{n}{H0}\PY{o}{;}~\PY{n+nb}{clear}~\PY{n}{H0}\PY{o}{.}
+      ~~\PY{n+nb}{assert}~\PY{o}{(}\PY{n}{Hlt}\PY{o}{:}~\PY{n}{omega}~\PY{o}{\PYZlt{}}~\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{)}~\PY{n+nb+bp}{by}~\nl
+      ~~~~~~\PY{o}{(}\PY{n+nb}{rewrite}~\PY{n}{omega\PYZus{}eqn}\PY{o}{;}~\PY{n+nb}{apply}~\PY{n}{phi0\PYZus{}mono}\PY{o}{,}~\PY{n}{finite\PYZus{}mono}\PY{o}{;}\nl
+      ~~~~~~~~\PY{n+nb}{auto~with}~\PY{n}{arith}\PY{o}{).}
     \end{input}
@@
             \hyp{HAP}{\PY{n}{In}~\PY{n}{AP}~\PY{o}{(}\PY{n}{phi0}~\PY{o}{(}\PY{n}{F}~\PY{l+m+mi}{2}\PY{o}{))}}
+            \sep
+            \hyp{Hlt}{\PY{n}{omega}~\PY{o}{\PYZlt{}}~\PY{n}{phi0}~\PY{o}{(}\PY{n}{F}~\PY{l+m+mi}{2}\PY{o}{)}}
           \end{hyps}
@@
           \begin{conclusion}
-            \PY{n+nb+bp}{omega}~\PY{o}{+}~\PY{n}{F}~\PY{l+m+mi}{42}~\PY{o}{\PYZlt{}}~\PY{n}{phi0}~\PY{o}{(}\PY{n}{F}~\PY{l+m+mi}{2}\PY{o}{)}
+            \PY{n}{omega}~\PY{o}{+}~\PY{n}{F}~\PY{l+m+mi}{42}~\PY{o}{\PYZlt{}}~\PY{n}{phi0}~\PY{o}{(}\PY{n}{F}~\PY{l+m+mi}{2}\PY{o}{)}
           \end{conclusion}

### 2021-10-04 doc/JFLA2022/paper.tex
remove a few TODOs[pierre]

@@
 \end{itemize}
-\item to continue to be useful as examples and reusable libraries, formalizations must be well \emph{documented} and \emph{maintained} to ensure they work with new proof assistant releases \TODO[Pierre]{add evolution of styles of proof and formalization (not limited to Coq litterature)?
+\item to continue to be useful as examples and reusable libraries, formalizations must be well \emph{documented} and \emph{maintained} to ensure they work with new proof assistant releases.
+%\TODO[Pierre]{add evolution of styles of proof and formalization (not limited to Coq litterature)?
 }
@@
 Originally, we copied snippets from Coq sources into LaTeX manually, and recorded and inserted the corresponding outputs manually as well.
-This approach is common, but brittle: changes to Coq definitions or lemmas had to be reflected in the book's sources, and we found multiple instances where the book and the Coq development had diverged (see FIXME cite alectryon for a discussion of various approaches to documentation of Coq developments).
+This approach is common, but brittle: changes to Coq definitions or lemmas had to be reflected in the book's sources, and we found multiple instances where the book and the Coq development had diverged.
+\TODO[Clément]{see FIXME cite alectryon for a discussion of various approaches to documentation of Coq developments}
 
@@
 The later part was straightforward: all it took was to build a custom Alectryon \textit{driver}, a small (about 100 lines) Python program that leverages most of the Alectryon toolchain but defines a custom frontend that understands our snippet annotations (and otherwise exposes the exact same command line as Alectryon).
-
-\begin{todo}
-  Interest of Alectryon:
-  \begin{itemize}
-  \item \TODO[Pierre]{the order of appearance of snippets in the book is independent from the module structure}
-   \item \TODO[Pierre]{Slicing our proofs into snippets allowed to show only the main parts of our proofsleaved with text (Figure~\ref{fig:Ex42-schutte}).}
-
-  \end{itemize}
-
-\end{todo}
 

### 2021-10-04 doc/JFLA2022/paper.tex
add a TODO (alectryon section)

@@
 Computability is a key topic in Computer Science teaching.
-Moreover O'Connor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter to this formalization, with comments on the definitions and proofs, [counter]-examples and exercises.
-
-We have extracted part of the Goedel library into a new sub-library of the Hydra-battles library: Ackermann.
+Moreover O'Connor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter to this formalization, with comments on the definitions and proofs, [counter]-examples and exercises, and  extract part of the Goedel library into a new sub-library of the Hydra-battles library: Ackermann.
 
@@
 The hydra-battles book is written in LaTeX, but it makes very frequent references (about once per page, 274 snippets over 281 pages) to parts of the Coq development, showing definitions (Fig.~\ref{fig:Hprime}), lemmas (Fig.~\ref{fig:Ex42E0}), proof scripts (Fig.~\ref{fig:Ex42-schutte}), and computation results (Fig.~\ref{fig:Pow-evalPow17LetIn}).
+The order in which these references appear in the book is independent of the structure of our libraries.
+Thus, we chose not to include our textbook as comments in the libraries (solution \emph{à la} Coqdoc).
+\TODO[Pierre]{is the preceding sentence OK?}
 
@@
 
-The transition happened gradually, over a few months: for each snippet of Coq code that was in the book, we had to take the following steps:
+The transition happened gradually, over a few weeks: for each snippet of Coq code that was in the book, we had to take the following steps:
 

### 2021-10-04 doc/JFLA2022/paper.tex
Macros for Hydras & co and Coq-community.

@@
 \newcommand{\latexeditor}{LEd}
+\newcommand{\Hydras}{Hydras \& Co. }
+\newcommand{\Coqcommunity}{\emph{Coq-community} }
 
@@
 \begin{abstract}
-  \emph{Hydra-battles} is a collaborative library on discrete math, written for the Coq proof assistant, and developed under the \emph{Coq-community} project. The proof scripts are
+  \Hydras is a collaborative library on discrete math, written for the Coq proof assistant, and developed under the \Coqcommunity project. The proof scripts are
   accompanied with an electronic book, made with the help of the \emph{Alectryon} litterate proving tool.
@@
   Then, we show how the structure of the project is determined   by two  requirements which must continuously satisfied:
-   maintenance of the library with respect to several Coq-community projects it requires and evolutions of Coq, and 
+   maintenance of the library with respect to several \Coqcommunity projects it requires and evolutions of Coq, and 
   consistency of the book with respect to the library. 
@@
 \subsection{Vision}
-The \emph{Hydra-battles} project, part of \emph{Coq-community}~\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
-\emph{Coq-community} is a community organization that we have founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting Coq packages, and working collaboratively on documentation projects.
-The \emph{Hydra-battles} project demonstrates that these two goals are not independent: interesting Coq packages can become the basis for new documentation.
+The \Hydras project, part of \Coqcommunity~\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
+\Coqcommunity is a community organization that we have founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting Coq packages, and working collaboratively on documentation projects.
+The \Hydras project demonstrates that these two goals are not independent: interesting Coq packages can become the basis for new documentation.
 %
@@
  JFLA~\cite{PCiota, JFLA2018paper}.
-We present recent evolutions of the library: new results, interaction within the Coq-community project~\cite{CoqCommunity}, and documentation generated with Alectryon~\cite{alectryonpaper, alectryongithub}.
+We present recent evolutions of the library: new results, interaction within the \Coqcommunity project~\cite{CoqCommunity}, and documentation generated with Alectryon~\cite{alectryonpaper, alectryongithub}.
 
@@
 
-\section{Integration within Coq-community}
+\section{Integration within \Coqcommunity}
 
@@
 
-\emph{Coq-community} is an informal organization run by volunteer Coq users with objectives to maintain interesting Coq projects in the long-term and help Coq users to collaborate on documentation, tooling, etc.
+\Coqcommunity is an informal organization run by volunteer Coq users with objectives to maintain interesting Coq projects in the long-term and help Coq users to collaborate on documentation, tooling, etc.
 
@@
 %
-\emph{Coq-community} makes that easy by providing a process for transferring or forking an unmaintained package, tooling for setting up good maintenance practices (such as continuous integration) and by making it possible for someone to take over a package without committing in the long-term (as maintainers who drop out can easily be replaced by some other volunteers).
+\Coqcommunity makes that easy by providing a process for transferring or forking an unmaintained package, tooling for setting up good maintenance practices (such as continuous integration) and by making it possible for someone to take over a package without committing in the long-term (as maintainers who drop out can easily be replaced by some other volunteers).
 
-Today, \emph{Coq-community} hosts over 50 projects maintained by over 30 maintainers.
+Today, \Coqcommunity hosts over 50 projects maintained by over 30 maintainers.
 %
@@
 %
-Others were simply left unmaintained and were revived by their adoption in \emph{Coq-community}.
+Others were simply left unmaintained and were revived by their adoption in \Coqcommunity.
 
-\subsection{Interactions with other Coq-community projects}
+\subsection{Interactions with other \Coqcommunity projects}
 
-Given the objectives of \emph{Coq-community}, it makes sense to propose a transfer anytime we encounter an interesting project that is insufficiently maintained.
+Given the objectives of \Coqcommunity, it makes sense to propose a transfer anytime we encounter an interesting project that is insufficiently maintained.
 %
@@
 of Russell O'Connor's contribution on G\"{o}del first incompleteness theorem~\cite{OConnor05, Goedel}.
-For this purpose, and above all by consideration of the scientific interest of this contribution, we chose to host and maintain this work within \textit{Coq-community}.
+For this purpose, and above all by consideration of the scientific interest of this contribution, we chose to host and maintain this work within \Coqcommunity.
 
@@
 
-The hydra-battles book is written in LaTeX, but it makes very frequent references (about once per page, 274 snippets over 281 pages) to parts of the Coq development, showing definitions (Fig.~\ref{fig:Hprime}), lemmas (Fig.~\ref{fig:Ex42E0}), proof scripts (Fig.~\ref{fig:Ex42-schutte}), and computation results (Fig.~\ref{fig:Pow-evalPow17LetIn}).
+The \Hydras book is written in LaTeX, but it makes very frequent references (about once per page, 274 snippets over 281 pages) to parts of the Coq development, showing definitions (Fig.~\ref{fig:Hprime}), lemmas (Fig.~\ref{fig:Ex42E0}), proof scripts (Fig.~\ref{fig:Ex42-schutte}), and computation results (Fig.~\ref{fig:Pow-evalPow17LetIn}).
 The order in which these references appear in the book is independent of the structure of our libraries.
@@
 Instead of copy-pasting fragments into LaTeX, we now embed small LaTeX files automatically generated by Alectryon from our Coq development.
-Our build system guarantees that these LaTeX snippets are always up-to-date and consistent with the code (and, by comparing these snippets across releases of Coq or versions of hydra-battles, we can easily spot unexpected changes).
+Our build system guarantees that these LaTeX snippets are always up-to-date and consistent with the code (and, by comparing these snippets across releases of Coq or versions of \Hydras, we can easily spot unexpected changes).
 
@@
 
-Importing snippets into a LaTeX document was not one of the original use cases of Alectryon, so we had to extend it: the original Alectryon did not have support for exporting to LaTeX, and it was towards documenting individual source files (a style where code and prose are interleaved within the same file and code is documented in the same order as it is compiled).
+Importing snippets into a LaTeX document was not one of the original use cases of Alectryon, so we had to extend it: the original Alectryon did not have support for exporting to LaTeX, and it was geared towards documenting individual source files (a style where code and prose are interleaved within the same file and code is documented in the same order as it is compiled).
 We extended Alectryon to support our needs by implementing a LaTeX backend, and by programming it to generate individual snippet files, one per \texttt{snippet} comment block.
@@
 
-\emph{Hydra-battles} wants to be a bridge between scientific litterature (\emph{e.g.}~\cite{KP82, KS81, schutte} and proof assistant technology.
+\Hydras wants to be a bridge between scientific litterature (\emph{e.g.}~\cite{KP82, KS81, schutte} and proof assistant technology.
 For the mathematician, it may give a concrete view of the mathematical content, not only through full proofs, but also through significative computations: examples, functions associated with constructive proofs, etc.
@@
 We plan to extend our libraries in two main directions.
-The Gaia project~\cite{Gaia}, also maintained on \textit{Coq-community} contains a development by José Grimm~\cite{grimm:hal-00911710}.
+The Gaia project~\cite{Gaia}, also maintained on \Coqcommunity contains a development by José Grimm~\cite{grimm:hal-00911710}.
 This library is dedicated  to the implementation in \textit{Coq (SSreflect)} of books from  Bourbaki's Elements of Mathematics.
@@
 
- \emph{Hydra-battles} is not limited to the study of ordinal numbers and applications.
+ \Hydras is not limited to the study of ordinal numbers and applications.
  For instance, we are also developing 

### 2021-10-04 doc/JFLA2022/paper.tex
macro for Coq-community

--- 2021-10-04 doc/JFLA2022/paper.tex
@@
 \usepackage{amsfonts, afterpage}
+\usepackage{xspace}
 
@@
 \newcommand{\TODO}[2][]{[\textcolor{red}{TODO (#1):} \emph{#2}]}
+\newcommand{\coq}{Coq}
+\newcommand{\community}{Coq-community\xspace}
+\newcommand{\gaia}{Gaia}
+\newcommand{\alectr}{Alectryon}
+\newcommand{\equations}{Equations}
 
@@
 \begin{abstract}
-  \emph{Hydra-battles} is a collaborative library on discrete math, written for the Coq proof assistant, and developed under the \emph{Coq-community} project. The proof scripts are
-  accompanied with an electronic book, made with the help of the \emph{Alectryon} litterate proving tool.
+  \emph{Hydra-battles} is a collaborative library on discrete math, written for the {\coq}  proof assistant, and developed under the \community project. The proof scripts are
+  accompanied with an electronic book, made with the help of the {\alectr} litterate proving tool.
   We present the evolution of the mathematical contents since
@@
   Then, we show how the structure of the project is determined   by two  requirements which must continuously satisfied:
-   maintenance of the library with respect to several Coq-community projects it requires and evolutions of Coq, and 
+   maintenance of the library with respect to several \community projects it requires and evolutions of {\coq}, and 
   consistency of the book with respect to the library. 
@@
 \begin{itemize}
-\item formalizations of mathematics in proof assistants, such as Coq, continually grow in size and scope
+\item formalizations of mathematics in proof assistants, such as {\coq}, continually grow in size and scope
 \begin{itemize}
-\item the Mathematical Components library and the four-color theorem and the odd order theorem in Coq
+\item the Mathematical Components library and the four-color theorem and the odd order theorem in {\coq}
 \item the Mathlib library for Lean
@@
 \subsection{Vision}
-The \emph{Hydra-battles} project, part of \emph{Coq-community}~\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
-\emph{Coq-community} is a community organization that we have founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting Coq packages, and working collaboratively on documentation projects.
-The \emph{Hydra-battles} project demonstrates that these two goals are not independent: interesting Coq packages can become the basis for new documentation.
+The \emph{Hydra-battles} project, part of \community~\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
+\community is a community organization that we have founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting {\coq} packages, and working collaboratively on documentation projects.
+The \emph{Hydra-battles} project demonstrates that these two goals are not independent: interesting {\coq} packages can become the basis for new documentation.
 %
@@
 %
-By following this approach of commenting interesting Coq packages, we provide new documentation content, that contributes to the diversity of the thriving Coq documentation ecosystem.
+By following this approach of commenting interesting {\coq} packages, we provide new documentation content, that contributes to the diversity of the thriving {\coq} documentation ecosystem.
 
-We call on the Coq users in the JFLA community and beyond to come and join us in this effort, by bringing new interesting projects which are worth presenting to Coq learners, \emph{a.k.a.} Coq users, and guiding them in their exploration.
+We call on the {\coq} users in the JFLA community and beyond to come and join us in this effort, by bringing new interesting projects which are worth presenting to {\coq} learners, \emph{a.k.a.} {\coq} users, and guiding them in their exploration.
 %
@@
 Futhermore, contrary to traditionally published books, the ``book'' that forms part of this project is intended to be forever evolving.
-As new Coq formalization patterns and proof techniques appear, the book can be adapted to demonstrate their use (in case they fit well with our applications).
+As new {\coq} formalization patterns and proof techniques appear, the book can be adapted to demonstrate their use (in case they fit well with our applications).
 %
-By using modern maintenance techniques such as continuous integration and deployment, we can ensure that this documentation stays up to date with the latest Coq releases.
-With Alectryon~\cite{alectryonpaper, alectryongithub}, we ensure that code and documentation are always in sync.
+By using modern maintenance techniques such as continuous integration and deployment, we can ensure that this documentation stays up to date with the latest {\coq} releases.
+With {\alectr}~\cite{alectryonpaper, alectryongithub}, we ensure that code and documentation are always in sync.
 
@@
 
-The contrast between the simplicity of the statements above and the complexity of their proofs convinced us that it is a good theme for a commented library~\cite{HydraBattles} of formal proofs written for the \textit{Coq} proof assistant~\cite{Coq}.
+The contrast between the simplicity of the statements above and the complexity of their proofs convinced us that it is a good theme for a commented library~\cite{HydraBattles} of formal proofs written for the {\coq} proof assistant~\cite{Coq}.
 
@@
  JFLA~\cite{PCiota, JFLA2018paper}.
-We present recent evolutions of the library: new results, interaction within the Coq-community project~\cite{CoqCommunity}, and documentation generated with Alectryon~\cite{alectryonpaper, alectryongithub}.
+We present recent evolutions of the library: new results, interaction within the \community project~\cite{CoqCommunity}, and documentation generated with {\alectr}~\cite{alectryonpaper, alectryongithub}.
 
@@
 
-\section{Integration within Coq-community}
+\section{Integration within \community}
 
@@
 
-\emph{Coq-community} is an informal organization run by volunteer Coq users with objectives to maintain interesting Coq projects in the long-term and help Coq users to collaborate on documentation, tooling, etc.
+\community is an informal organization run by volunteer {\coq} users with objectives to maintain interesting {\coq} projects in the long-term and help {\coq} users to collaborate on documentation, tooling, etc.
 
-It was created in 2018, inspired by the \emph{Elm Community} organization~\cite{zimmermann:tel-02451322}.
+It was created in 2018, inspired by the {Elm Community} organization~\cite{zimmermann:tel-02451322}.
 %
@@
 
-In the case of Coq, this problem may be even more prevalent as many packages are created by PhD students, or by researchers for a specific paper, but their authors do not intend to maintain the package in the long-term.
+In the case of {\coq}, this problem may be even more prevalent as many packages are created by PhD students, or by researchers for a specific paper, but their authors do not intend to maintain the package in the long-term.
 %
@@
 %
-\emph{Coq-community} makes that easy by providing a process for transferring or forking an unmaintained package, tooling for setting up good maintenance practices (such as continuous integration) and by making it possible for someone to take over a package without committing in the long-term (as maintainers who drop out can easily be replaced by some other volunteers).
+\community makes that easy by providing a process for transferring or forking an unmaintained package, tooling for setting up good maintenance practices (such as continuous integration) and by making it possible for someone to take over a package without committing in the long-term (as maintainers who drop out can easily be replaced by some other volunteers).
 
-Today, \emph{Coq-community} hosts over 50 projects maintained by over 30 maintainers.
+Today, \community hosts over 50 projects maintained by over 30 maintainers.
 %
@@
 %
-Some had been maintained in the past by the Coq development team on behalf of the authors, but this meant that only minimal changes required to make the project build with new Coq versions were conducted.
+Some had been maintained in the past by the {\coq} development team on behalf of the authors, but this meant that only minimal changes required to make the project build with new {\coq} versions were conducted.
 %
@@
 %
-Others were simply left unmaintained and were revived by their adoption in \emph{Coq-community}.
+Others were simply left unmaintained and were revived by their adoption in \community.
 
-\subsection{Interactions with other Coq-community projects}
+\subsection{Interactions with other \community projects}
 
-Given the objectives of \emph{Coq-community}, it makes sense to propose a transfer anytime we encounter an interesting project that is insufficiently maintained.
+Given the objectives of \community, it makes sense to propose a transfer anytime we encounter an interesting project that is insufficiently maintained.
 %
@@
 of Russell O'Connor's contribution on G\"{o}del first incompleteness theorem~\cite{OConnor05, Goedel}.
-For this purpose, and above all by consideration of the scientific interest of this contribution, we chose to host and maintain this work within \textit{Coq-community}.
+For this purpose, and above all by consideration of the scientific interest of this contribution, we chose to host and maintain this work within \community.
 
@@
 
-\subsection{Documentation with Alectryon}
+\subsection{Documentation with {\alectr}}
 
-The hydra-battles book is written in LaTeX, but it makes very frequent references (about once per page, 274 snippets over 281 pages) to parts of the Coq development, showing definitions (Fig.~\ref{fig:Hprime}), lemmas (Fig.~\ref{fig:Ex42E0}), proof scripts (Fig.~\ref{fig:Ex42-schutte}), and computation results (Fig.~\ref{fig:Pow-evalPow17LetIn}).
+The hydra-battles book is written in LaTeX, but it makes very frequent references (about once per page, 274 snippets over 281 pages) to parts of the {\coq} development, showing definitions (Fig.~\ref{fig:Hprime}), lemmas (Fig.~\ref{fig:Ex42E0}), proof scripts (Fig.~\ref{fig:Ex42-schutte}), and computation results (Fig.~\ref{fig:Pow-evalPow17LetIn}).
 The order in which these references appear in the book is independent of the structure of our libraries.
@@
 
-Originally, we copied snippets from Coq sources into LaTeX manually, and recorded and inserted the corresponding outputs manually as well.
-This approach is common, but brittle: changes to Coq definitions or lemmas had to be reflected in the book's sources, and we found multiple instances where the book and the Coq development had diverged.
-\TODO[Clément]{see FIXME cite alectryon for a discussion of various approaches to documentation of Coq developments}
+Originally, we copied snippets from {\coq} sources into LaTeX manually, and recorded and inserted the corresponding outputs manually as well.
+This approach is common, but brittle: changes to {\coq} definitions or lemmas had to be reflected in the book's sources, and we found multiple instances where the book and the {\coq} development had diverged.
+\TODO[Clément]{see FIXME cite alectryon for a discussion of various approaches to documentation of {\coq} developments}
 
-We solved this maintenance issue by moving to Alectryon, a tool to automatically record Coq proofs.
-Instead of copy-pasting fragments into LaTeX, we now embed small LaTeX files automatically generated by Alectryon from our Coq development.
-Our build system guarantees that these LaTeX snippets are always up-to-date and consistent with the code (and, by comparing these snippets across releases of Coq or versions of hydra-battles, we can easily spot unexpected changes).
+We solved this maintenance issue by moving to {\alectr}, a tool to automatically record {\coq} proofs.
+Instead of copy-pasting fragments into LaTeX, we now embed small LaTeX files automatically generated by {\alectr} from our {\coq} development.
+Our build system guarantees that these LaTeX snippets are always up-to-date and consistent with the code (and, by comparing these snippets across releases of {\coq} or versions of hydra-battles, we can easily spot unexpected changes).
 
-The transition happened gradually, over a few weeks: for each snippet of Coq code that was in the book, we had to take the following steps:
+The transition happened gradually, over a few weeks: for each snippet of {\coq} code that was in the book, we had to take the following steps:
 
 \begin{enumerate}
-\item Mark the snippet in the Coq sources (we use special comments \texttt{(* begin snippet \textit{name} *)} … \texttt{(* end snippet \textit{name} *)})
-\item Configure output display, using special Alectryon annotations to configure what should be shown (only the inputs, inputs and outputs, some steps of the proof but not all, some proof states at key moments in a proof, etc.)
+\item Mark the snippet in the {\coq} sources (we use special comments \texttt{(* begin snippet \textit{name} *)} … \texttt{(* end snippet \textit{name} *)})
+\item Configure output display, using special {\alectr} annotations to configure what should be shown (only the inputs, inputs and outputs, some steps of the proof but not all, some proof states at key moments in a proof, etc.)
 \item Replace the copy-pasted inputs and output in LaTeX with an \texttt{input} command.
@@
 
-Importing snippets into a LaTeX document was not one of the original use cases of Alectryon, so we had to extend it: the original Alectryon did not have support for exporting to LaTeX, and it was towards documenting individual source files (a style where code and prose are interleaved within the same file and code is documented in the same order as it is compiled).
-We extended Alectryon to support our needs by implementing a LaTeX backend, and by programming it to generate individual snippet files, one per \texttt{snippet} comment block.
-The later part was straightforward: all it took was to build a custom Alectryon \textit{driver}, a small (about 100 lines) Python program that leverages most of the Alectryon toolchain but defines a custom frontend that understands our snippet annotations (and otherwise exposes the exact same command line as Alectryon).
+Importing snippets into a LaTeX document was not one of the original use cases of {\alectr}, so we had to extend it: the original {\alectr} did not have support for exporting to LaTeX, and it was towards documenting individual source files (a style where code and prose are interleaved within the same file and code is documented in the same order as it is compiled).
+We extended {\alectr} to support our needs by implementing a LaTeX backend, and by programming it to generate individual snippet files, one per \texttt{snippet} comment block.
+The later part was straightforward: all it took was to build a custom {\alectr} \textit{driver}, a small (about 100 lines) Python program that leverages most of the {\alectr} toolchain but defines a custom frontend that understands our snippet annotations (and otherwise exposes the exact same command line as {\alectr}).
 
@@
 For the mathematician, it may give a concrete view of the mathematical content, not only through full proofs, but also through significative computations: examples, functions associated with constructive proofs, etc.
-For the Coq user, it provides a consistent set of examples,
+For the {\coq} user, it provides a consistent set of examples,
 allowing to present and compare various formalization and proving techniques. It may be also used as a library, through opam packages.
 
-It is also a medium sized library (more than 50K LOC), dependent on various tools and libraries of the Coq ecosystem.
-It may be also used to experiment new techniques of continuous maintenace of the code and its documentation.
+It is also a medium sized library (more than 50K LOC), dependent on various tools and libraries of the {\coq} ecosystem.
+It may be also used to experiment new techniques of continuous maintenance of the code and its documentation.
 
@@
 We plan to extend our libraries in two main directions.
-The Gaia project~\cite{Gaia}, also maintained on \textit{Coq-community} contains a development by José Grimm~\cite{grimm:hal-00911710}.
-This library is dedicated  to the implementation in \textit{Coq (SSreflect)} of books from  Bourbaki's Elements of Mathematics.
+The Gaia project~\cite{Gaia}, also maintained on \community, contains a development by José Grimm~\cite{grimm:hal-00911710}.
+This library is dedicated  to the implementation in {\coq} (SSreflect) of books from  Bourbaki's Elements of Mathematics.
 It contains data structures which are compatible with our ordinal notations.
 Our plan is to build a  bridge between  the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}, and make possible the transfer of theorems between both libraries.
-Another direction would be to write in \textit{Coq} a formal proof of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
+Another direction would be to write in {\coq} a formal proof of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
 

### 2021-10-04 doc/JFLA2022/paper.tex
macro for Coq

@@
 \newcommand{\TODO}[2][]{[\textcolor{red}{TODO (#1):} \emph{#2}]}
-\newcommand{\coq}{Coq}
+\newcommand{\coq}{Coq\xspace}
 \newcommand{\community}{Coq-community\xspace}
@@
 \begin{abstract}
-  \emph{Hydra-battles} is a collaborative library on discrete math, written for the {\coq}  proof assistant, and developed under the \community project. The proof scripts are
+  \emph{Hydra-battles} is a collaborative library on discrete math, written for the \coq  proof assistant, and developed under the \community project. The proof scripts are
   accompanied with an electronic book, made with the help of the {\alectr} litterate proving tool.
@@
   Then, we show how the structure of the project is determined   by two  requirements which must continuously satisfied:
-   maintenance of the library with respect to several \community projects it requires and evolutions of {\coq}, and 
+   maintenance of the library with respect to several \community projects it requires and evolutions of \coq, and 
   consistency of the book with respect to the library. 
@@
 \begin{itemize}
-\item formalizations of mathematics in proof assistants, such as {\coq}, continually grow in size and scope
+\item formalizations of mathematics in proof assistants, such as \coq, continually grow in size and scope
 \begin{itemize}
-\item the Mathematical Components library and the four-color theorem and the odd order theorem in {\coq}
+\item the Mathematical Components library and the four-color theorem and the odd order theorem in \coq
 \item the Mathlib library for Lean
@@
 The \emph{Hydra-battles} project, part of \community~\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
-\community is a community organization that we have founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting {\coq} packages, and working collaboratively on documentation projects.
-The \emph{Hydra-battles} project demonstrates that these two goals are not independent: interesting {\coq} packages can become the basis for new documentation.
+\community is a community organization that we have founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting \coq packages, and working collaboratively on documentation projects.
+The \emph{Hydra-battles} project demonstrates that these two goals are not independent: interesting \coq packages can become the basis for new documentation.
 %
@@
 %
-By following this approach of commenting interesting {\coq} packages, we provide new documentation content, that contributes to the diversity of the thriving {\coq} documentation ecosystem.
+By following this approach of commenting interesting \coq packages, we provide new documentation content, that contributes to the diversity of the thriving \coq documentation ecosystem.
 
-We call on the {\coq} users in the JFLA community and beyond to come and join us in this effort, by bringing new interesting projects which are worth presenting to {\coq} learners, \emph{a.k.a.} {\coq} users, and guiding them in their exploration.
+We call on the \coq users in the JFLA community and beyond to come and join us in this effort, by bringing new interesting projects which are worth presenting to \coq learners, \emph{a.k.a.} \coq users, and guiding them in their exploration.
 %
@@
 Futhermore, contrary to traditionally published books, the ``book'' that forms part of this project is intended to be forever evolving.
-As new {\coq} formalization patterns and proof techniques appear, the book can be adapted to demonstrate their use (in case they fit well with our applications).
+As new \coq formalization patterns and proof techniques appear, the book can be adapted to demonstrate their use (in case they fit well with our applications).
 %
-By using modern maintenance techniques such as continuous integration and deployment, we can ensure that this documentation stays up to date with the latest {\coq} releases.
+By using modern maintenance techniques such as continuous integration and deployment, we can ensure that this documentation stays up to date with the latest \coq releases.
 With {\alectr}~\cite{alectryonpaper, alectryongithub}, we ensure that code and documentation are always in sync.
@@
 
-The contrast between the simplicity of the statements above and the complexity of their proofs convinced us that it is a good theme for a commented library~\cite{HydraBattles} of formal proofs written for the {\coq} proof assistant~\cite{Coq}.
+The contrast between the simplicity of the statements above and the complexity of their proofs convinced us that it is a good theme for a commented library~\cite{HydraBattles} of formal proofs written for the \coq proof assistant~\cite{Coq}.
 
@@
 
-\community is an informal organization run by volunteer {\coq} users with objectives to maintain interesting {\coq} projects in the long-term and help {\coq} users to collaborate on documentation, tooling, etc.
+\community is an informal organization run by volunteer \coq users with objectives to maintain interesting \coq projects in the long-term and help \coq users to collaborate on documentation, tooling, etc.
 
@@
 
-In the case of {\coq}, this problem may be even more prevalent as many packages are created by PhD students, or by researchers for a specific paper, but their authors do not intend to maintain the package in the long-term.
+In the case of \coq, this problem may be even more prevalent as many packages are created by PhD students, or by researchers for a specific paper, but their authors do not intend to maintain the package in the long-term.
 %
@@
 %
-Some had been maintained in the past by the {\coq} development team on behalf of the authors, but this meant that only minimal changes required to make the project build with new {\coq} versions were conducted.
+Some had been maintained in the past by the \coq development team on behalf of the authors, but this meant that only minimal changes required to make the project build with new \coq versions were conducted.
 %
@@
 
-The hydra-battles book is written in LaTeX, but it makes very frequent references (about once per page, 274 snippets over 281 pages) to parts of the {\coq} development, showing definitions (Fig.~\ref{fig:Hprime}), lemmas (Fig.~\ref{fig:Ex42E0}), proof scripts (Fig.~\ref{fig:Ex42-schutte}), and computation results (Fig.~\ref{fig:Pow-evalPow17LetIn}).
+The hydra-battles book is written in LaTeX, but it makes very frequent references (about once per page, 274 snippets over 281 pages) to parts of the \coq development, showing definitions (Fig.~\ref{fig:Hprime}), lemmas (Fig.~\ref{fig:Ex42E0}), proof scripts (Fig.~\ref{fig:Ex42-schutte}), and computation results (Fig.~\ref{fig:Pow-evalPow17LetIn}).
 The order in which these references appear in the book is independent of the structure of our libraries.
@@
 
-Originally, we copied snippets from {\coq} sources into LaTeX manually, and recorded and inserted the corresponding outputs manually as well.
-This approach is common, but brittle: changes to {\coq} definitions or lemmas had to be reflected in the book's sources, and we found multiple instances where the book and the {\coq} development had diverged.
-\TODO[Clément]{see FIXME cite alectryon for a discussion of various approaches to documentation of {\coq} developments}
+Originally, we copied snippets from \coq sources into LaTeX manually, and recorded and inserted the corresponding outputs manually as well.
+This approach is common, but brittle: changes to \coq definitions or lemmas had to be reflected in the book's sources, and we found multiple instances where the book and the \coq development had diverged.
+\TODO[Clément]{see FIXME cite alectryon for a discussion of various approaches to documentation of \coq developments}
 
-We solved this maintenance issue by moving to {\alectr}, a tool to automatically record {\coq} proofs.
-Instead of copy-pasting fragments into LaTeX, we now embed small LaTeX files automatically generated by {\alectr} from our {\coq} development.
-Our build system guarantees that these LaTeX snippets are always up-to-date and consistent with the code (and, by comparing these snippets across releases of {\coq} or versions of hydra-battles, we can easily spot unexpected changes).
+We solved this maintenance issue by moving to {\alectr}, a tool to automatically record \coq proofs.
+Instead of copy-pasting fragments into LaTeX, we now embed small LaTeX files automatically generated by {\alectr} from our \coq development.
+Our build system guarantees that these LaTeX snippets are always up-to-date and consistent with the code (and, by comparing these snippets across releases of \coq or versions of hydra-battles, we can easily spot unexpected changes).
 
-The transition happened gradually, over a few weeks: for each snippet of {\coq} code that was in the book, we had to take the following steps:
+The transition happened gradually, over a few weeks: for each snippet of \coq code that was in the book, we had to take the following steps:
 
 \begin{enumerate}
-\item Mark the snippet in the {\coq} sources (we use special comments \texttt{(* begin snippet \textit{name} *)} … \texttt{(* end snippet \textit{name} *)})
+\item Mark the snippet in the \coq sources (we use special comments \texttt{(* begin snippet \textit{name} *)} … \texttt{(* end snippet \textit{name} *)})
 \item Configure output display, using special {\alectr} annotations to configure what should be shown (only the inputs, inputs and outputs, some steps of the proof but not all, some proof states at key moments in a proof, etc.)
@@
 For the mathematician, it may give a concrete view of the mathematical content, not only through full proofs, but also through significative computations: examples, functions associated with constructive proofs, etc.
-For the {\coq} user, it provides a consistent set of examples,
+For the \coq user, it provides a consistent set of examples,
 allowing to present and compare various formalization and proving techniques. It may be also used as a library, through opam packages.
 
-It is also a medium sized library (more than 50K LOC), dependent on various tools and libraries of the {\coq} ecosystem.
+It is also a medium sized library (more than 50K LOC), dependent on various tools and libraries of the \coq ecosystem.
 It may be also used to experiment new techniques of continuous maintenance of the code and its documentation.
@@
 The Gaia project~\cite{Gaia}, also maintained on \community, contains a development by José Grimm~\cite{grimm:hal-00911710}.
-This library is dedicated  to the implementation in {\coq} (SSreflect) of books from  Bourbaki's Elements of Mathematics.
+This library is dedicated  to the implementation in \coq (SSreflect) of books from  Bourbaki's Elements of Mathematics.
 It contains data structures which are compatible with our ordinal notations.
 Our plan is to build a  bridge between  the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}, and make possible the transfer of theorems between both libraries.
-Another direction would be to write in {\coq} a formal proof of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
+Another direction would be to write in \coq a formal proof of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
 

### 2021-10-04 doc/JFLA2022/paper.tex

--- 2021-10-04 doc/JFLA2022/paper.tex
@@
 \newcommand{\latexeditor}{LEd}
-\newcommand{\Hydras}{Hydras \& Co. }
-\newcommand{\Coqcommunity}{\emph{Coq-community} }
+\newcommand{\Hydras}{Hydras \& Co.\xspace}
+\newcommand{\Coqcommunity}{\emph{Coq-community\xspace} }
 
@@
 \newcommand{\TODO}[2][]{[\textcolor{red}{TODO (#1):} \emph{#2}]}
+\newcommand{\coq}{Coq\xspace}
+\newcommand{\community}{Coq-community\xspace}
+\newcommand{\gaia}{Gaia\xspace}
+\newcommand{\alectr}{Alectryon\xspace}
+\newcommand{\equations}{Equations\xspace}
 
@@
   Then, we show how the structure of the project is determined   by two  requirements which must continuously satisfied:
-   maintenance of the library with respect to several \Coqcommunity projects it requires and evolutions of Coq, and 
-  consistency of the book with respect to the library. 
+   maintenance of the library with respect to several \Coqcommunity projects it requires and evolutions of Coq, and  consistency of the book with respect to the library. 
 \end{abstract}
@@
 \begin{itemize}
-\item formalizations of mathematics in proof assistants, such as Coq, continually grow in size and scope
+\item formalizations of mathematics in proof assistants, such as \coq, continually grow in size and scope
 \begin{itemize}
-\item the Mathematical Components library and the four-color theorem and the odd order theorem in Coq
+\item the Mathematical Components library and the four-color theorem and the odd order theorem in \coq
 \item the Mathlib library for Lean
@@
 \subsection{Vision}
-The \Hydras project, part of \Coqcommunity~\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
+The \Hydras project, part of \Coqcommunity\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
 \Coqcommunity is a community organization that we have founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting Coq packages, and working collaboratively on documentation projects.
 The \Hydras project demonstrates that these two goals are not independent: interesting Coq packages can become the basis for new documentation.
+
 %
@@
 %
-By following this approach of commenting interesting Coq packages, we provide new documentation content, that contributes to the diversity of the thriving Coq documentation ecosystem.
+By following this approach of commenting interesting \coq packages, we provide new documentation content, that contributes to the diversity of the thriving \coq documentation ecosystem.
 
-We call on the Coq users in the JFLA community and beyond to come and join us in this effort, by bringing new interesting projects which are worth presenting to Coq learners, \emph{a.k.a.} Coq users, and guiding them in their exploration.
+We call on the \coq users in the JFLA community and beyond to come and join us in this effort, by bringing new interesting projects which are worth presenting to \coq learners, \emph{a.k.a.} \coq users, and guiding them in their exploration.
 %
@@
 Futhermore, contrary to traditionally published books, the ``book'' that forms part of this project is intended to be forever evolving.
-As new Coq formalization patterns and proof techniques appear, the book can be adapted to demonstrate their use (in case they fit well with our applications).
+As new \coq formalization patterns and proof techniques appear, the book can be adapted to demonstrate their use (in case they fit well with our applications).
 %
-By using modern maintenance techniques such as continuous integration and deployment, we can ensure that this documentation stays up to date with the latest Coq releases.
-With Alectryon~\cite{alectryonpaper, alectryongithub}, we ensure that code and documentation are always in sync.
+By using modern maintenance techniques such as continuous integration and deployment, we can ensure that this documentation stays up to date with the latest \coq releases.
+With {\alectr}~\cite{alectryonpaper, alectryongithub}, we ensure that code and documentation are always in sync.
 
@@
 
-The contrast between the simplicity of the statements above and the complexity of their proofs convinced us that it is a good theme for a commented library~\cite{HydraBattles} of formal proofs written for the \textit{Coq} proof assistant~\cite{Coq}.
+The contrast between the simplicity of the statements above and the complexity of their proofs convinced us that it is a good theme for a commented library~\cite{HydraBattles} of formal proofs written for the \coq proof assistant~\cite{Coq}.
 
@@
 
-It was created in 2018, inspired by the \emph{Elm Community} organization~\cite{zimmermann:tel-02451322}.
+It was created in 2018, inspired by the {Elm Community} organization~\cite{zimmermann:tel-02451322}.
 %
@@
 
-In the case of Coq, this problem may be even more prevalent as many packages are created by PhD students, or by researchers for a specific paper, but their authors do not intend to maintain the package in the long-term.
+In the case of \coq, this problem may be even more prevalent as many packages are created by PhD students, or by researchers for a specific paper, but their authors do not intend to maintain the package in the long-term.
 %
@@
 %
+
 \Coqcommunity makes that easy by providing a process for transferring or forking an unmaintained package, tooling for setting up good maintenance practices (such as continuous integration) and by making it possible for someone to take over a package without committing in the long-term (as maintainers who drop out can easily be replaced by some other volunteers).
@@
 %
-Some had been maintained in the past by the Coq development team on behalf of the authors, but this meant that only minimal changes required to make the project build with new Coq versions were conducted.
+Some had been maintained in the past by the \coq development team on behalf of the authors, but this meant that only minimal changes required to make the project build with new \coq versions were conducted.
 %
@@
 
-\subsection{Documentation with Alectryon}
+\subsection{Documentation with {\alectr}}
 
@@
 
-Originally, we copied snippets from Coq sources into LaTeX manually, and recorded and inserted the corresponding outputs manually as well.
-This approach is common, but brittle: changes to Coq definitions or lemmas had to be reflected in the book's sources, and we found multiple instances where the book and the Coq development had diverged.
-\TODO[Clément]{see FIXME cite alectryon for a discussion of various approaches to documentation of Coq developments}
+Originally, we copied snippets from \coq sources into LaTeX manually, and recorded and inserted the corresponding outputs manually as well.
+This approach is common, but brittle: changes to \coq definitions or lemmas had to be reflected in the book's sources, and we found multiple instances where the book and the \coq development had diverged.
+\TODO[Clément]{see FIXME cite alectryon for a discussion of various approaches to documentation of \coq developments}
 
@@
 
-The transition happened gradually, over a few weeks: for each snippet of Coq code that was in the book, we had to take the following steps:
+The transition happened gradually, over a few weeks: for each snippet of \coq code that was in the book, we had to take the following steps:
 
 \begin{enumerate}
-\item Mark the snippet in the Coq sources (we use special comments \texttt{(* begin snippet \textit{name} *)} … \texttt{(* end snippet \textit{name} *)})
-\item Configure output display, using special Alectryon annotations to configure what should be shown (only the inputs, inputs and outputs, some steps of the proof but not all, some proof states at key moments in a proof, etc.)
+\item Mark the snippet in the \coq sources (we use special comments \texttt{(* begin snippet \textit{name} *)} … \texttt{(* end snippet \textit{name} *)})
+\item Configure output display, using special {\alectr} annotations to configure what should be shown (only the inputs, inputs and outputs, some steps of the proof but not all, some proof states at key moments in a proof, etc.)
 \item Replace the copy-pasted inputs and output in LaTeX with an \texttt{input} command.
@@
 We extended Alectryon to support our needs by implementing a LaTeX backend, and by programming it to generate individual snippet files, one per \texttt{snippet} comment block.
-The later part was straightforward: all it took was to build a custom Alectryon \textit{driver}, a small (about 100 lines) Python program that leverages most of the Alectryon toolchain but defines a custom frontend that understands our snippet annotations (and otherwise exposes the exact same command line as Alectryon).
+The later part was straightforward: all it took was to build a custom Alectryon \textit{driver}, a small (about 100 lines) Python program that leverages most of the Alectryon toolchain but defines a custom frontend that understands our snippet annotations (and otherwise exposes the exact same command line as \alectr).
 
@@
 
-\Hydras wants to be a bridge between scientific litterature (\emph{e.g.}~\cite{KP82, KS81, schutte} and proof assistant technology.
+\Hydras wants to be a bridge between scientific litterature (\emph{e.g.}~\cite{KP82, KS81, schutte}) and proof assistant technology.
 For the mathematician, it may give a concrete view of the mathematical content, not only through full proofs, but also through significative computations: examples, functions associated with constructive proofs, etc.
@@
 
-It is also a medium sized library (more than 50K LOC), dependent on various tools and libraries of the Coq ecosystem.
-It may be also used to experiment new techniques of continuous maintenace of the code and its documentation.
+It is also a medium sized library (more than 50K LOC), dependent on various tools and libraries of the \coq ecosystem.
+It may be also used to experiment new techniques of continuous maintenance of the code and its documentation.
 
@@
 We plan to extend our libraries in two main directions.
+
 The Gaia project~\cite{Gaia}, also maintained on \Coqcommunity contains a development by José Grimm~\cite{grimm:hal-00911710}.
@@
 It contains data structures which are compatible with our ordinal notations.
+
 Our plan is to build a  bridge between  the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}, and make possible the transfer of theorems between both libraries.
-Another direction would be to write in \textit{Coq} a formal proof of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
+Another direction would be to write in \coq a formal proof of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
 

### 2021-10-05 doc/JFLA2022/paper.tex
Update doc/JFLA2022/paper.tex
Co-authored-by: Karl Palmskog <palmskog@gmail.com>

@@
 \newcommand{\Hydras}{Hydras \& Co.\xspace}
-\newcommand{\Coqcommunity}{\emph{Coq-community\xspace} }
+\newcommand{\Coqcommunity}{Coq-community\xspace}
 
@@
 \subsection{Vision}
-The \Hydras project, part of \Coqcommunity\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
+The \Hydras project, part of \Coqcommunity~\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
 \Coqcommunity is a community organization that we have founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting Coq packages, and working collaboratively on documentation projects.

### 2021-10-05 doc/JFLA2022/snippets/Hprime/HprimeDef.tex

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      \PY{n}{Equations}~\PY{n}{H\PYZsq{}\PYZus{}}~\PY{o}{(}\PY{n}{alpha}\PY{o}{:}~\PY{n}{E0}\PY{o}{)}~\PY{o}{(}\PY{n}{i}\PY{o}{:}\PY{n}{nat}\PY{o}{)}~\PY{o}{:}~~\PY{n}{nat}~~\PY{n+nb+bp}{by}~\PY{k+kn}{wf}~~\PY{n}{alpha}~\PY{n}{Lt}~\PY{o}{:=}\nl
      ~~\PY{n}{H\PYZsq{}\PYZus{}}~\PY{n}{alpha}~~\PY{n}{i}~\PY{k+kr}{with}~\PY{n}{E0\PYZus{}eq\PYZus{}dec}~\PY{n}{alpha}~\PY{n}{Zero}~\PY{o}{:=}\nl
      ~~~~\PY{o}{\PYZob{}}~\PY{o}{|}~\PY{n+nb}{left}~\PY{n}{\PYZus{}zero}~\PY{o}{=\PYZgt{}}~~\PY{n}{i}~\PY{o}{;}\nl
      ~~~~~~\PY{o}{|}~\PY{n+nb}{right}~\PY{n}{\PYZus{}nonzero}\nl
      ~~~~~~~~~~\PY{k+kr}{with}~\PY{n}{Utils}\PY{o}{.}\PY{n}{dec}~\PY{o}{(}\PY{n}{Limitb}~\PY{n}{alpha}\PY{o}{)}~\PY{o}{:=}\nl
      ~~~~~~~~~~\PY{o}{\PYZob{}}~\PY{o}{|}~\PY{n+nb}{left}~\PY{n}{\PYZus{}limit}~\PY{o}{=\PYZgt{}}~~\PY{n}{H\PYZsq{}\PYZus{}}~\PY{o}{(}\PY{n}{Canon}~\PY{n}{alpha}~\PY{o}{(}\PY{n}{S}~\PY{n}{i}\PY{o}{))}~~\PY{n}{i}~\PY{o}{;}\nl
      ~~~~~~~~~~~~\PY{o}{|}~\PY{n+nb}{right}~\PY{n}{\PYZus{}successor}~\PY{o}{=\PYZgt{}}~~\PY{n}{H\PYZsq{}\PYZus{}}~\PY{o}{(}\PY{n}{Pred}~\PY{n}{alpha}\PY{o}{)}~\PY{o}{(}\PY{n}{S}~\PY{n}{i}\PY{o}{)\PYZcb{}\PYZcb{}.}
    \end{input}
  \end{sentence}
\end{alectryon}

### 2021-10-05 doc/JFLA2022/paper.tex
use macros for citing libraries

--- 2021-10-04 doc/JFLA2022/paper.tex
@@
 \newcommand{\community}{Coq-community\xspace}
-\newcommand{\gaia}{Gaia}
-\newcommand{\alectr}{Alectryon}
-\newcommand{\equations}{Equations}
+\newcommand{\gaia}{Gaia\xspace}
+\newcommand{\alectr}{Alectryon\xspace}
+\newcommand{\equations}{Equations\xspace}
+\newcommand{\Hydras}{Hydras \& Co.\xspace}
 
@@
 \begin{abstract}
-  \emph{Hydra-battles} is a collaborative library on discrete math, written for the \coq  proof assistant, and developed under the \community project. The proof scripts are
-  accompanied with an electronic book, made with the help of the {\alectr} litterate proving tool.
+  \Hydras is a collaborative library on discrete math, written for the \coq  proof assistant, and developed under the \community project. The proof scripts are
+  accompanied with an electronic book, made with the help of the \alectr litterate proving tool.
   We present the evolution of the mathematical contents since
@@
 \subsection{Vision}
-The \emph{Hydra-battles} project, part of \community~\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
+The \Hydras project, part of \community~\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
 \community is a community organization that we have founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting \coq packages, and working collaboratively on documentation projects.
-The \emph{Hydra-battles} project demonstrates that these two goals are not independent: interesting \coq packages can become the basis for new documentation.
+The \Hydras project demonstrates that these two goals are not independent: interesting \coq packages can become the basis for new documentation.
 %
-This umbrella project now includes evolved versions of the former Cantor and Additions libraries (under the new names of Hydra-battles and Addition-chains), the Ackermann sub-library, extracted from Russel O'Connor's Goedel library~\cite{OConnor05, Goedel} and a bridge to the Gaia library (by José Grimm~\cite{Gaia,grimm:hal-00911710}).
+This umbrella project now includes evolved versions of the former Cantor and Additions libraries (under the new names of Hydra-battles and Addition-chains), the Ackermann sub-library, extracted from Russel O'Connor's Goedel library~\cite{OConnor05, Goedel} and a bridge to the \gaia library (by José Grimm~\cite{Gaia,grimm:hal-00911710}).
 %
@@
 By using modern maintenance techniques such as continuous integration and deployment, we can ensure that this documentation stays up to date with the latest \coq releases.
-With {\alectr}~\cite{alectryonpaper, alectryongithub}, we ensure that code and documentation are always in sync.
+With \alectr~\cite{alectryonpaper, alectryongithub}, we ensure that code and documentation are always in sync.
 
@@
  JFLA~\cite{PCiota, JFLA2018paper}.
-We present recent evolutions of the library: new results, interaction within the \community project~\cite{CoqCommunity}, and documentation generated with {\alectr}~\cite{alectryonpaper, alectryongithub}.
+We present recent evolutions of the library: new results, interaction within the \community project~\cite{CoqCommunity}, and documentation generated with \alectr~\cite{alectryonpaper, alectryongithub}.
 
@@
 Computability is a key topic in Computer Science teaching.
-Moreover O'Connor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter to this formalization, with comments on the definitions and proofs, [counter]-examples and exercises, and  extract part of the Goedel library into a new sub-library of the Hydra-battles library: Ackermann.
+Moreover O'Connor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter to this formalization, with comments on the definitions and proofs, [counter]-examples and exercises, and  extract part of the Goedel library into a new sub-library of the \Hydras library: Ackermann.
 
@@
 
-\subsection{Documentation with {\alectr}}
+\subsection{Documentation with \alectr}
 
-The hydra-battles book is written in LaTeX, but it makes very frequent references (about once per page, 274 snippets over 281 pages) to parts of the \coq development, showing definitions (Fig.~\ref{fig:Hprime}), lemmas (Fig.~\ref{fig:Ex42E0}), proof scripts (Fig.~\ref{fig:Ex42-schutte}), and computation results (Fig.~\ref{fig:Pow-evalPow17LetIn}).
+The \Hydras book is written in LaTeX, but it makes very frequent references (about once per page, 274 snippets over 281 pages) to parts of the \coq development, showing definitions (Fig.~\ref{fig:Hprime}), lemmas (Fig.~\ref{fig:Ex42E0}), proof scripts (Fig.~\ref{fig:Ex42-schutte}), and computation results (Fig.~\ref{fig:Pow-evalPow17LetIn}).
 The order in which these references appear in the book is independent of the structure of our libraries.
@@
 
-We solved this maintenance issue by moving to {\alectr}, a tool to automatically record \coq proofs.
-Instead of copy-pasting fragments into LaTeX, we now embed small LaTeX files automatically generated by {\alectr} from our \coq development.
-Our build system guarantees that these LaTeX snippets are always up-to-date and consistent with the code (and, by comparing these snippets across releases of \coq or versions of hydra-battles, we can easily spot unexpected changes).
+We solved this maintenance issue by moving to \alectr, a tool to automatically record \coq proofs.
+Instead of copy-pasting fragments into LaTeX, we now embed small LaTeX files automatically generated by \alectr from our \coq development.
+Our build system guarantees that these LaTeX snippets are always up-to-date and consistent with the code (and, by comparing these snippets across releases of \coq or versions of \Hydras, we can easily spot unexpected changes).
 
@@
 \item Mark the snippet in the \coq sources (we use special comments \texttt{(* begin snippet \textit{name} *)} … \texttt{(* end snippet \textit{name} *)})
-\item Configure output display, using special {\alectr} annotations to configure what should be shown (only the inputs, inputs and outputs, some steps of the proof but not all, some proof states at key moments in a proof, etc.)
+\item Configure output display, using special \alectr annotations to configure what should be shown (only the inputs, inputs and outputs, some steps of the proof but not all, some proof states at key moments in a proof, etc.)
 \item Replace the copy-pasted inputs and output in LaTeX with an \texttt{input} command.
@@
 
-Importing snippets into a LaTeX document was not one of the original use cases of {\alectr}, so we had to extend it: the original {\alectr} did not have support for exporting to LaTeX, and it was towards documenting individual source files (a style where code and prose are interleaved within the same file and code is documented in the same order as it is compiled).
-We extended {\alectr} to support our needs by implementing a LaTeX backend, and by programming it to generate individual snippet files, one per \texttt{snippet} comment block.
-The later part was straightforward: all it took was to build a custom {\alectr} \textit{driver}, a small (about 100 lines) Python program that leverages most of the {\alectr} toolchain but defines a custom frontend that understands our snippet annotations (and otherwise exposes the exact same command line as {\alectr}).
+Importing snippets into a LaTeX document was not one of the original use cases of \alectr, so we had to extend it: the original \alectr did not have support for exporting to LaTeX, and it was towards documenting individual source files (a style where code and prose are interleaved within the same file and code is documented in the same order as it is compiled).
+We extended {alectr to support our needs by implementing a LaTeX backend, and by programming it to generate individual snippet files, one per \texttt{snippet} comment block.
+The later part was straightforward: all it took was to build a custom \alectr \textit{driver}, a small (about 100 lines) Python program that leverages most of the \alectr toolchain but defines a custom frontend that understands our snippet annotations (and otherwise exposes the exact same command line as \alectr).
 
@@
 We plan to extend our libraries in two main directions.
-The Gaia project~\cite{Gaia}, also maintained on \community, contains a development by José Grimm~\cite{grimm:hal-00911710}.
+The \gaia project~\cite{Gaia}, also maintained on \community, contains a development by José Grimm~\cite{grimm:hal-00911710}.
 This library is dedicated  to the implementation in \coq (SSreflect) of books from  Bourbaki's Elements of Mathematics.
 It contains data structures which are compatible with our ordinal notations.
-Our plan is to build a  bridge between  the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}, and make possible the transfer of theorems between both libraries.
+Our plan is to build a  bridge between  the combinatorial results of \Hydras and the set-theoretic content of \gaia, and make possible the transfer of theorems between both libraries.
 Another direction would be to write in \coq a formal proof of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
 
- \emph{Hydra-battles} is not limited to the study of ordinal numbers and applications.
+ \Hydras is not limited to the study of ordinal numbers and applications.
  For instance, we are also developing 

### 2021-10-05 doc/JFLA2022/paper.tex
fix a typo in the abstract Put latex comments on several remaining TODOs. Move a paragraph in Section 2

@@
   former presentations at JFLA meetings.
-  Then, we show how the structure of the project is determined   by two  requirements which must continuously satisfied:
+  Then, we show how the structure of the project is determined   by two  requirements which must be continuously satisfied:
    maintenance of the library with respect to several \community projects it requires and evolutions of \coq, and 
@@
 
+More generally, we prove that, for $\alpha\geq\omega^\omega$
+the function computing the length of the battle starting with the configuration $(\alpha,i)$ is not primitive recursive.
+
 \begin{figure}[h]
@@
 
-\TODO[Clément]{Could we use a font with ligatures?}
-
-More generally, we prove that, for $\alpha\geq\omega^\omega$
-the function computing the length of the battle starting with the configuration $(\alpha,i)$ is not primitive recursive.
+%\TODO[Clément]{Could we use a font with ligatures?}
 
@@
 Thus, we chose not to include our textbook as comments in the libraries (solution \emph{à la} Coqdoc).
-\TODO[Pierre]{is the preceding sentence OK?}
+%\TODO[Pierre]{is the preceding sentence OK?}
 
@@
 This approach is common, but brittle: changes to \coq definitions or lemmas had to be reflected in the book's sources, and we found multiple instances where the book and the \coq development had diverged.
-\TODO[Clément]{see FIXME cite alectryon for a discussion of various approaches to documentation of \coq developments}
+%\TODO[Clément]{see FIXME cite alectryon for a discussion of various approaches to documentation of \coq developments}
 

### 2021-10-05 doc/JFLA2022/paper.tex

@@
 \newcommand{\latexeditor}{LEd}
+\newcommand{\Hydras}{Hydras \& Co.\xspace}
+\newcommand{\Coqcommunity}{Coq-community\xspace}
 
@@
  JFLA~\cite{PCiota, JFLA2018paper}.
+
 We present recent evolutions of the library: new results, interaction within the \community project~\cite{CoqCommunity}, and documentation generated with \alectr~\cite{alectryonpaper, alectryongithub}.
@@
 
-\section{Integration within \community}
+\section{Integration within \Coqcommunity}
 
@@
 
-\community is an informal organization run by volunteer \coq users with objectives to maintain interesting \coq projects in the long-term and help \coq users to collaborate on documentation, tooling, etc.
+\Coqcommunity is an informal organization run by volunteer Coq users with objectives to maintain interesting Coq projects in the long-term and help Coq users to collaborate on documentation, tooling, etc.
 
@@
 %
-\community makes that easy by providing a process for transferring or forking an unmaintained package, tooling for setting up good maintenance practices (such as continuous integration) and by making it possible for someone to take over a package without committing in the long-term (as maintainers who drop out can easily be replaced by some other volunteers).
 
-Today, \community hosts over 50 projects maintained by over 30 maintainers.
+\Coqcommunity makes that easy by providing a process for transferring or forking an unmaintained package, tooling for setting up good maintenance practices (such as continuous integration) and by making it possible for someone to take over a package without committing in the long-term (as maintainers who drop out can easily be replaced by some other volunteers).
+
+Today, \Coqcommunity hosts over 50 projects maintained by over 30 maintainers.
 %
@@
 %
-Others were simply left unmaintained and were revived by their adoption in \community.
+Others were simply left unmaintained and were revived by their adoption in \Coqcommunity.
 
-\subsection{Interactions with other \community projects}
+\subsection{Interactions with other \Coqcommunity projects}
 
-Given the objectives of \community, it makes sense to propose a transfer anytime we encounter an interesting project that is insufficiently maintained.
+Given the objectives of \Coqcommunity, it makes sense to propose a transfer anytime we encounter an interesting project that is insufficiently maintained.
 %
@@
 of Russell O'Connor's contribution on G\"{o}del first incompleteness theorem~\cite{OConnor05, Goedel}.
-For this purpose, and above all by consideration of the scientific interest of this contribution, we chose to host and maintain this work within \community.
+For this purpose, and above all by consideration of the scientific interest of this contribution, we chose to host and maintain this work within \Coqcommunity.
 
@@
 We plan to extend our libraries in two main directions.
+
 The \gaia project~\cite{Gaia}, also maintained on \community, contains a development by José Grimm~\cite{grimm:hal-00911710}.
@@
 Our plan is to build a  bridge between  the combinatorial results of \Hydras and the set-theoretic content of \gaia, and make possible the transfer of theorems between both libraries.
+
 Another direction would be to write in \coq a formal proof of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.

### 2021-10-05 doc/JFLA2022/paper.tex
two macros for Hydras

--- 2021-10-05 doc/JFLA2022/paper.tex
@@
 %
-\title{Hydras, Ordinals \&  Co.  \\
+\title{\HydrasLib  \\
   A library in Coq of entertaining formal mathematics}
@@
 
-\titlerunning{Hydras, Ordinals \& Co.}
+\titlerunning{\HydrasLib}
 
@@
 \newcommand{\equations}{Equations\xspace}
-\newcommand{\Hydras}{Hydras \& Co.\xspace}
+\newcommand{\HydrasBook}{Hydras \& Co.\xspace}
+\newcommand{\HydrasLib}{Hydra-battles\xspace}
 
@@
 \begin{abstract}
-  \Hydras is a collaborative library on discrete math, written for the \coq  proof assistant, and developed under the \community project. The proof scripts are
+  \HydrasLib is a collaborative library on discrete math, written for the \coq  proof assistant, and developed under the \community project. The proof scripts are
   accompanied with an electronic book, made with the help of the \alectr litterate proving tool.
@@
 \subsection{Vision}
-The \Hydras project, part of \community~\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
+The \HydrasLib project, part of \community~\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
 \community is a community organization that we have founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting \coq packages, and working collaboratively on documentation projects.
-The \Hydras project demonstrates that these two goals are not independent: interesting \coq packages can become the basis for new documentation.
+The \HydrasLib project demonstrates that these two goals are not independent: interesting \coq packages can become the basis for new documentation.
 %
@@
 
-\community is an informal organization run by volunteer \coq users with objectives to maintain interesting \coq projects in the long-term and help \coq users to collaborate on documentation, tooling, etc.
+\community is an informal organization run by volunteer Coq users with objectives to maintain interesting Coq projects in the long-term and help Coq users to collaborate on documentation, tooling, etc.
 
@@
 %
+
 \community makes that easy by providing a process for transferring or forking an unmaintained package, tooling for setting up good maintenance practices (such as continuous integration) and by making it possible for someone to take over a package without committing in the long-term (as maintainers who drop out can easily be replaced by some other volunteers).
@@
 Computability is a key topic in Computer Science teaching.
-Moreover O'Connor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter to this formalization, with comments on the definitions and proofs, [counter]-examples and exercises, and  extract part of the Goedel library into a new sub-library of the \Hydras library: Ackermann.
+Moreover O'Connor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter to this formalization, with comments on the definitions and proofs, [counter]-examples and exercises, and  extract part of the Goedel library into a new sub-library of the \HydrasLib library: Ackermann.
 
@@
 
-The \Hydras book is written in LaTeX, but it makes very frequent references (about once per page, 274 snippets over 281 pages) to parts of the \coq development, showing definitions (Fig.~\ref{fig:Hprime}), lemmas (Fig.~\ref{fig:Ex42E0}), proof scripts (Fig.~\ref{fig:Ex42-schutte}), and computation results (Fig.~\ref{fig:Pow-evalPow17LetIn}).
+The \HydrasBook book is written in LaTeX, but it makes very frequent references (about once per page, 274 snippets over 281 pages) to parts of the \coq development, showing definitions (Fig.~\ref{fig:Hprime}), lemmas (Fig.~\ref{fig:Ex42E0}), proof scripts (Fig.~\ref{fig:Ex42-schutte}), and computation results (Fig.~\ref{fig:Pow-evalPow17LetIn}).
 The order in which these references appear in the book is independent of the structure of our libraries.
@@
 Instead of copy-pasting fragments into LaTeX, we now embed small LaTeX files automatically generated by \alectr from our \coq development.
-Our build system guarantees that these LaTeX snippets are always up-to-date and consistent with the code (and, by comparing these snippets across releases of \coq or versions of \Hydras, we can easily spot unexpected changes).
+Our build system guarantees that these LaTeX snippets are always up-to-date and consistent with the code (and, by comparing these snippets across releases of \coq or versions of \HydrasLib, we can easily spot unexpected changes).
 
@@
 
-\Hydras wants to be a bridge between scientific litterature (\emph{e.g.}~\cite{KP82, KS81, schutte} and proof assistant technology.
+\HydrasLib wants to be a bridge between scientific litterature (\emph{e.g.}~\cite{KP82, KS81, schutte} and proof assistant technology.
 For the mathematician, it may give a concrete view of the mathematical content, not only through full proofs, but also through significative computations: examples, functions associated with constructive proofs, etc.
@@
 It contains data structures which are compatible with our ordinal notations.
-Our plan is to build a  bridge between  the combinatorial results of \Hydras and the set-theoretic content of \gaia, and make possible the transfer of theorems between both libraries.
+Our plan is to build a  bridge between  the combinatorial results of \HydrasLib and the set-theoretic content of \gaia, and make possible the transfer of theorems between both libraries.
+
 Another direction would be to write in \coq a formal proof of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
 
- \Hydras is not limited to the study of ordinal numbers and applications.
+ \HydrasLib is not limited to the study of ordinal numbers and applications.
  For instance, we are also developing 

### 2021-10-05 doc/JFLA2022/paper.tex
fix the names of project and book

--- 2021-10-05 doc/JFLA2022/paper.tex
@@
 
-\titlerunning{Hydras, Ordinals \& Co.}
+\titlerunning{\Hydras}
 
@@
 \newcommand{\Hydras}{Hydras \& Co.\xspace}
+%\newcommand{\Hydras}{Hydra-battles\xspace}
 

### 2021-10-06 doc/JFLA2022/paper.tex

@@
 
-\subsection{Continuous integration and deployment, opam packaging, etc.}
+\subsection{Technologies supporting the monorepo structure}
+
+Monorepos (shorthand for monolithic repositories) are single repositories containing multiple independent or related packages.
+%
+They have gained increasing popularity following experiences in large companies such as Google, but are also used at a lesser scale for managing open source projects.
+%
+They are known to simplify the management of dependencies, making cross-packages changes and refactorings~\cite{brito2018monorepos}.
+
+In the context of Coq projects, they did not use to be well-supported.
+%
+One of the only monorepo in the Coq ecosystem was the math-comp repository, and it used a custom build infrastructure to make this possible.
+%
+However, recent improvements in tooling have made their use possible, and therefore we have used them more and more, for instance within \community.
+
+In the context of \Hydras, we rely on the following tools.
+
+\paragraph{The Dune build system}
+It allows building packages contained in a single source tree separately, which is essential to be able to publish the different libraries contained in a single repository as separate (opam) packages to the Coq package index.
+
+In practice, we do still support building the whole project with coq\_makefile, and we actually rely on this build system in our documentation generation pipeline.
+
+\paragraph{The Coq Nix Toolbox~\cite{cohen:hal-03366644}}
+We rely on this toolbox mainly because it can generate a Continuous Integration setup (relying on GitHub Actions) to test the various libraries contained in the project, but also their dependencies.
+%
+This has allowed us, for instance, to test the Goedel library with the current version of the Hydra-battles library when we have moved a subpart of Goedel to Hydra-battles (the Ackermann sub-library).
+
+The build artifacts are stored in the \community binary cache on Cachix~\cite{cachix}.
+%
+This means that a given version never has to be built twice and Continuous Integration can be almost instantaneous when no change has been made (e.g., after merging a pull request).
+
 \section{Conclusion and perspectives}

### 2021-10-06 doc/JFLA2022/paper.tex
minor changes
add an active link to monorepos article; small changes
minor changes in conclusion

@@
 \institute{
-Univ. Bordeaux, CNRS, Bordeaux INP, LaBRI, UMR 5800, F-33400 Talence, France \\
-  \email{pierre.casteran@labri.fr}
+Univ. Bordeaux, CNRS, Bordeaux INP, LaBRI, UMR 5800, F-33400 Talence, France  %\\
+ % \email{pierre.casteran@labri.fr}
 \and
@@
 allowing to present and compare various formalization and proving techniques. It may be also used as a library, through opam packages.
-
 It is also a medium sized library (more than 50K LOC), dependent on various tools and libraries of the \coq ecosystem.
@@
 We plan to extend our libraries in two main directions.
-
-The \gaia project~\cite{Gaia}, also maintained on \community, contains a development by José Grimm~\cite{grimm:hal-00911710}.
-This library is dedicated  to the implementation in \coq (SSreflect) of books from  Bourbaki's Elements of Mathematics.
+The \gaia project~\cite{Gaia}, also maintained on \community, contains a development by José Grimm~\cite{grimm:hal-00911710},  dedicated  to the implementation in \coq (SSreflect) of books from  Bourbaki's Elements of Mathematics.
 It contains data structures which are compatible with our ordinal notations.
 Our plan is to build a  bridge between  the combinatorial results of \Hydras and the set-theoretic content of \gaia, and make possible the transfer of theorems between both libraries.
-
-Another direction would be to write in \coq a formal proof of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
+We aim also to write in \coq a formal proof of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
 

### 2021-10-06 doc/JFLA2022/paper.tex
Changing some  figure positions

@@
  JFLA~\cite{PCiota, JFLA2018paper}.
-
 We present recent evolutions of the library: new results, interaction within the \community project~\cite{CoqCommunity}, and documentation generated with \alectr~\cite{alectryonpaper, alectryongithub}.
@@
 \section{Changes in the library}
-The 2018 article~\cite{JFLA2018paper} contains a formal proof of  a variant of Theorem~\ref{kp:thm2}:
+The 2018 article~\cite{JFLA2018paper} presented a formal proof of  a variant of Theorem~\ref{kp:thm2}:
 
@@
 assume that the hydra grows $n$ copies at step $n$ of the game, which is incompatible with our proof.
-
-\vspace{6pt}
-
 We prove now that Theorem~\ref{thm3} still holds with these typical battles by borrowing new combinatorial results from~\cite{KS81}.
@@
 By a sequence of rewritings and inductions, we prove that the number of steps of the considered battle is greater or equal than $2^{2^N}$, where $N=2^{70}-1$.
-
 More generally, we prove that, for $\alpha\geq\omega^\omega$
@@
 
-The \Hydras book is written in LaTeX, but it makes very frequent references (about once per page, 274 snippets over 281 pages) to parts of the \coq development, showing definitions (Fig.~\ref{fig:Hprime}), lemmas (Fig.~\ref{fig:Ex42E0}), proof scripts (Fig.~\ref{fig:Ex42-schutte}), and computation results (Fig.~\ref{fig:Pow-evalPow17LetIn}).
+The \Hydras book is written in LaTeX, but it makes very frequent references (about once per page, 274 snippets over 281 pages) to parts of the \coq development, showing definitions (Fig.~\ref{fig:Hprime}), computation results (Fig.~\ref{fig:Pow-evalPow17LetIn}), lemmas (Fig.~\ref{fig:Ex42E0}) and parts of  proof scripts (Fig.~\ref{fig:Ex42-schutte}).
 The order in which these references appear in the book is independent of the structure of our libraries.
@@
 The later part was straightforward: all it took was to build a custom \alectr \textit{driver}, a small (about 100 lines) Python program that leverages most of the \alectr toolchain but defines a custom frontend that understands our snippet annotations (and otherwise exposes the exact same command line as \alectr).
+
+   \begin{figure}[h]
+  \centering
+  \inputsnippets{Pow/evalPow17LetIn}
+  \caption{\label{fig:Pow-evalPow17LetIn}Automatically capturing the output of computations}
+\end{figure}
 
@@
 
-\begin{figure}
-  \centering
-  \inputsnippets{Pow/evalPow17LetIn}
-  \caption{\label{fig:Pow-evalPow17LetIn}Automatically capturing the output of computations}
-\end{figure}
-
 \subsection{Technologies supporting the monorepo structure}
@@
 
-The build artifacts are stored in the \community binary cache on Cachix~\cite{cachix}.
+The built artifacts are stored in the \community binary cache on Cachix~\cite{cachix}.
 %

### 2021-10-06 doc/JFLA2022/snippets/E0/Ex42.tex
consistent use of fboxes around snippets

@@
     \begin{input}
-      \PY{k+kn}{Example}~\PY{n+nf}{Ex42}\PY{o}{:}~\PY{o}{(}\PY{n}{omega}~\PY{o}{+}~\PY{l+m+mi}{42}~\PY{o}{+}~\PY{n}{omega}\PY{o}{\PYZca{}}~\PY{l+m+mi}{2}~\PY{o}{=}~\PY{n}{omega}\PY{o}{\PYZca{}}~\PY{l+m+mi}{2}\PY{o}{)\PYZpc{}}\PY{n}{e0}\PY{o}{.}\nl
+      \PY{k+kn}{Example}~\PY{n+nf}{Ex42}\PY{o}{:}~\PY{o}{(}\PY{n+nb+bp}{omega}~\PY{o}{+}~\PY{l+m+mi}{42}~\PY{o}{+}~\PY{n+nb+bp}{omega}\PY{o}{\PYZca{}}~\PY{l+m+mi}{2}~\PY{o}{=}~\PY{n+nb+bp}{omega}\PY{o}{\PYZca{}}~\PY{l+m+mi}{2}\PY{o}{)\PYZpc{}}\PY{n}{e0}\PY{o}{.}\nl
     \end{input}
@@
           \begin{conclusion}
-            \PY{n+nb}{compare}~\PY{o}{(}\PY{n}{omega}~\PY{o}{+}~\PY{l+m+mi}{42}~\PY{o}{+}~\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{)}~\PY{o}{(}\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{)}~\PY{o}{=}~\PY{n}{Eq}
+            \PY{n+nb}{compare}~\PY{o}{(}\PY{n+nb+bp}{omega}~\PY{o}{+}~\PY{l+m+mi}{42}~\PY{o}{+}~\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{)}~\PY{o}{(}\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{)}~\PY{o}{=}~\PY{n}{Eq}
           \end{conclusion}

### 2021-10-06 doc/JFLA2022/snippets/Schutte/Ex42c.tex
consistent use of fboxes around snippets

--- 2021-09-20 doc/JFLA2022/Ex42b.tex
@@
     \begin{input}
-      ~~\PY{n+nb}{assert}~\PY{o}{(}\PY{n}{HAP}\PY{o}{:}~\PY{n}{In}~\PY{n}{AP}~\PY{o}{(}\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{))}~\PY{n+nb+bp}{by}~\PY{n+nb}{apply}~\PY{n}{AP\PYZus{}phi0}\PY{o}{.}\nl
-    \end{input}
-  \end{sentence}
-  \sep
-  \begin{sentence}
-    \begin{input}
-      ~~\PY{n+nb}{elim}~~\PY{n}{HAP}\PY{o}{;}~\PY{n+nb}{intros}~~\PY{n}{\PYZus{}}~\PY{n}{H0}\PY{o}{;}~\PY{n+nb}{apply}~\PY{n}{H0}\PY{o}{;}~\PY{n+nb}{clear}~\PY{n}{H0}\PY{o}{.}
+      ~~\PY{n+nb}{assert}~\PY{o}{(}\PY{n}{Hlt}\PY{o}{:}~\PY{n+nb+bp}{omega}~\PY{o}{\PYZlt{}}~\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{)}~\PY{n+nb+bp}{by}~\nl
+      ~~~~~~\PY{o}{(}\PY{n+nb}{rewrite}~\PY{n}{omega\PYZus{}eqn}\PY{o}{;}~\PY{n+nb}{apply}~\PY{n}{phi0\PYZus{}mono}\PY{o}{,}~\PY{n}{finite\PYZus{}mono}\PY{o}{;}\nl
+      ~~~~~~~~\PY{n+nb}{auto~with}~\PY{n}{arith}\PY{o}{).}
     \end{input}

### 2021-10-06 doc/JFLA2022/paper.tex
consistent use of fboxes around snippets
add a remark in Section 2

@@
 
-\section{Changes in the library}
+\section{Recent developments}
 The 2018 article~\cite{JFLA2018paper} presented a formal proof of  a variant of Theorem~\ref{kp:thm2}:
@@
    \item  The hydra is always the  representation as a tree of some ordinal strictly below $\epsilon_0$ in Cantor normal
-     form. For instance, Fig~\ref{fig:round} shows the hydras respectively associated with  $\omega^{\omega^2+1}$ and $\omega^{\omega^2}\times 3$.
+     form (thus, the rightmost branch is also one of the shortest).
+     For instance, Fig~\ref{fig:round} shows the hydras respectively associated with  $\omega^{\omega^2+1}$ and $\omega^{\omega^2}\times 3$.
 
@@
  In mathematical terms, if at step $n$ the hydra is associated with the ordinal $\alpha$, at step $n+1$ it is associated with
- $\canonseq{\alpha}{n+1}$, the $(n+1)$th element of the canonical sequence of $\alpha$~\cite{KS81}.
+ $\canonseq{\alpha}{n+1}$, \emph{i.e.} the $(n+1)$th element of the canonical sequence of $\alpha$~\cite{KS81}.
 
@@
 \begin{figure}[h]
+  \centering
+  \fbox{
+    \begin{minipage}[h]{1.0\linewidth}
  \inputsnippets{Hprime/HprimeDef}     
+    \end{minipage}}
  \caption{$H'$ definition with the \texttt{coq-equations}
@@
 By a sequence of rewritings and inductions, we prove that the number of steps of the considered battle is greater or equal than $2^{2^N}$, where $N=2^{70}-1$.
-More generally, we prove that, for $\alpha\geq\omega^\omega$
+More generally, we prove that, for $\alpha\geq\omega^\omega$, 
 the function computing the length of the battle starting with the configuration $(\alpha,i)$ is not primitive recursive.
@@
      \centering
+     \fbox{
+       \begin{minipage}[h]{1.0\linewidth}
          \inputsnippets{Pow/evalPow17LetIn}
+       \end{minipage}}
   \caption{\label{fig:Pow-evalPow17LetIn}Automatically capturing the output of computations}

### 2021-10-06 doc/JFLA2022/snippets/E0/Ex42.tex
improvement of two snippets

@@
     \begin{input}
-      \PY{k+kn}{Example}~\PY{n+nf}{Ex42}\PY{o}{:}~\PY{o}{(}\PY{n+nb+bp}{omega}~\PY{o}{+}~\PY{l+m+mi}{42}~\PY{o}{+}~\PY{n+nb+bp}{omega}\PY{o}{\PYZca{}}~\PY{l+m+mi}{2}~\PY{o}{=}~\PY{n+nb+bp}{omega}\PY{o}{\PYZca{}}~\PY{l+m+mi}{2}\PY{o}{)\PYZpc{}}\PY{n}{e0}\PY{o}{.}\nl
+      \PY{k+kn}{Example}~\PY{n+nf}{Ex42}\PY{o}{:}~\PY{n+nb+bp}{omega}~\PY{o}{+}~\PY{l+m+mi}{42}~\PY{o}{+}~\PY{n+nb+bp}{omega}\PY{o}{\PYZca{}}\PY{l+m+mi}{2}~\PY{o}{=}~\PY{n+nb+bp}{omega}\PY{o}{\PYZca{}}\PY{l+m+mi}{2}\PY{o}{.}\nl
     \end{input}

### 2021-10-06 doc/JFLA2022/snippets/Schutte/Ex42c.tex
improvement of two snippets

@@
     \begin{input}
-      ~~\PY{n+nb}{assert}~\PY{o}{(}\PY{n}{Hlt}\PY{o}{:}~\PY{n+nb+bp}{omega}~\PY{o}{\PYZlt{}}~\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{)}~\PY{n+nb+bp}{by}~\nl
+      ~~\PY{n+nb}{assert}~\PY{o}{(}\PY{n}{Hlt}\PY{o}{:}~\PY{n+nb+bp}{omega}~\PY{o}{\PYZlt{}}~\PY{n+nb+bp}{omega}\PY{o}{\PYZca{}}\PY{l+m+mi}{2}\PY{o}{)}~\PY{n+nb+bp}{by}~\nl
       ~~~~~~\PY{o}{(}\PY{n+nb}{rewrite}~\PY{n}{omega\PYZus{}eqn}\PY{o}{;}~\PY{n+nb}{apply}~\PY{n}{phi0\PYZus{}mono}\PY{o}{,}~\PY{n}{finite\PYZus{}mono}\PY{o}{;}\nl

### 2021-10-07 doc/JFLA2022/snippets/Schutte/Ex42c.tex
jfla: Recompile snippets

--- 2021-10-04 doc/JFLA2022/snippets/Schutte/Ex42c.tex
@@
     \begin{input}
-      ~~\PY{n+nb}{assert}~\PY{o}{(}\PY{n}{Hlt}\PY{o}{:}~\PY{n}{omega}~\PY{o}{\PYZlt{}}~\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{)}~\PY{n+nb+bp}{by}~\nl
+      ~~\PY{n+nb}{assert}~\PY{o}{(}\PY{n}{Hlt}\PY{o}{:}~\PY{n}{omega}~\PY{o}{\PYZlt{}}~\PY{n}{omega}\PY{o}{\PYZca{}}\PY{l+m+mi}{2}\PY{o}{)}~\PY{n+nb+bp}{by}~\nl
       ~~~~~~\PY{o}{(}\PY{n+nb}{rewrite}~\PY{n}{omega\PYZus{}eqn}\PY{o}{;}~\PY{n+nb}{apply}~\PY{n}{phi0\PYZus{}mono}\PY{o}{,}~\PY{n}{finite\PYZus{}mono}\PY{o}{;}\nl

### 2021-10-07 doc/JFLA2022/paper.tex
Tweak a sentence about Alectryon

@@
 The \Hydras book is written in LaTeX, but it makes very frequent references (about once per page, 274 snippets over 281 pages) to parts of the \coq development, showing definitions (Fig.~\ref{fig:Hprime}), computation results (Fig.~\ref{fig:Pow-evalPow17LetIn}), lemmas (Fig.~\ref{fig:Ex42E0}) and parts of  proof scripts (Fig.~\ref{fig:Ex42-schutte}).
-The order in which these references appear in the book is independent of the structure of our libraries.
-Thus, we chose not to include our textbook as comments in the libraries (solution \emph{à la} Coqdoc).
-%\TODO[Pierre]{is the preceding sentence OK?}
+The order in which these references appear in the book is independent of the structure of our libraries, so we chose to maintain the book as a standalone document, separate from the \coq source code.
+This in contrast with the Coqdoc approach, where explanatory prose is embedded within \coq source files (a detailed discussion of various approaches to the documentation of \coq developments is in \cite{alectryonpaper}).
 
@@
 This approach is common, but brittle: changes to \coq definitions or lemmas had to be reflected in the book's sources, and we found multiple instances where the book and the \coq development had diverged.
-%\TODO[Clément]{see FIXME cite alectryon for a discussion of various approaches to documentation of \coq developments}
 

### 2021-10-07 doc/JFLA2022/paper.tex
improve (?) introduction of Section 2

--- 2021-10-06 doc/JFLA2022/paper.tex
@@
 This work is also an opportunity to 
- provide ``concrete'' examples of formalization and proof techniques: operational type classes, functions defined by  equations, dependently typed functions, etc.
+ provide concrete examples of formalization and proof techniques: operational type classes, functions defined by  equations, dependently typed functions, etc.
 It may be also used as a library on ordinal numbers, for instance for proving termination properties.
@@
   Let $\mu$ be any ordinal strictly less than $\epsilon_0$.
-  There is no function mapping hydras to the segment $[0,\mu)$ that could be used as a measure for  proving the termination of  all hydra battles.\label{thm3}
+  There is no function mapping hydras to the segment $[0,\mu)$ that could be used as a measure for  proving the termination of  \emph{any} hydra battle.\label{thm3}
 \end{theorem}
 
-Considering measures applicable to \emph{all} battles allowed us to
-focus on battles where the hydra can make an arbitrary number of copies at any time, which made our proof by contradiction artificially easier.
+Our proof was based on the construction of a battle, where the
+number of copies at each round is given by the elimination of an existential quantifier. So, it was mandatory to consider the class of \emph{all} hydra battles.
+
 Unfortunately, the examples  most commonly shown in the litterature
@@
 We prove now that Theorem~\ref{thm3} still holds with these typical battles by borrowing new combinatorial results from~\cite{KS81}.
-Without loss of generality, we assume the following restrictions:
+Since we are looking  for a minoration of the length of such battles,
+we can work with the following hypotheses and invariants, 
+without any loss of generality.
 
@@
 
-Using $H'$s equations as rewrite rules, we can study a realistic example. We take the hydra of figure~\ref{fig:start} and $i=0$ as initial configuration. 
+Using $H'$s equations as rewrite rules, we can study a concrete example. We take the hydra of figure~\ref{fig:start} and $i=0$ as initial configuration. 
 By a sequence of rewritings and inductions, we prove that the number of steps of the considered battle is greater or equal than $2^{2^N}$, where $N=2^{70}-1$.

### 2021-10-07 doc/JFLA2022/paper.tex
improve (?) introduction of Section 2

@@
 
- In mathematical terms, if at step $n$ the hydra is associated with the ordinal $\alpha$, at step $n+1$ it is associated with
+ In mathematical terms, if at step $n\geq i$, the hydra is associated with the ordinal $\alpha$, at step $n+1$ it is associated with
  $\canonseq{\alpha}{n+1}$, \emph{i.e.} the $(n+1)$th element of the canonical sequence of $\alpha$~\cite{KS81}.

### 2021-10-07 doc/JFLA2022/paper.tex
add references to two chapters of the book

@@
 assume that the hydra grows $n$ copies at step $n$ of the game, which is incompatible with our proof.
-We prove now that Theorem~\ref{thm3} still holds with these typical battles by borrowing new combinatorial results from~\cite{KS81}.
+We prove now that Theorem~\ref{thm3} still holds with these typical battles.
+%by borrowing new combinatorial results from~\cite{KS81}.
 Since we are looking  for a minoration of the length of such battles,
@@
 
- Our new proof of Theorem~\ref{thm3} is based on a systematic study of strictly decreasing sequences of ordinals below $\epsilon_0$, borrowed from~\cite{KS81}.
+ Our new proof of Theorem~\ref{thm3} is based on a systematic study of strictly decreasing sequences of ordinals below $\epsilon_0$, borrowed from Ketonen and Solovay~\cite{KS81}, and the formalization of which is described  in chapter 5 of~\cite{HydraBook}.
 
@@
 Computability is a key topic in Computer Science teaching.
-Moreover O'Connor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter to this formalization, with comments on the definitions and proofs, [counter]-examples and exercises, and  extract part of the Goedel library into a new sub-library of the \Hydras library: Ackermann.
+Moreover O'Connor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter (chapter 9 of~\cite{HydraBook}) to this formalization, with comments on the definitions and proofs, [counter]-examples and exercises, and  extract part of the Goedel library into a new sub-library of the \Hydras library: Ackermann.
 

### 2021-10-08 doc/JFLA2022/paper.tex
add a TODO (in the abstract)
minor correction

@@
    consistency of the book with respect to the library.
+   \TODO[Pierre]{Find a better/simpler sentence about the project's structure?}
 \end{abstract}
@@
 
- Our new proof of Theorem~\ref{thm3} is based on a systematic study of strictly decreasing sequences of ordinals below $\epsilon_0$, borrowed from Ketonen and Solovay~\cite{KS81}, and the formalization of which is described  in chapter 5 of~\cite{HydraBook}.
+ Our new proof of Theorem~\ref{thm3} is based on a systematic study of strictly decreasing sequences of ordinals below $\epsilon_0$, borrowed from Ketonen and Solovay~\cite{KS81}, and the formalization of which is described  in chapters 5 and 6 of~\cite{HydraBook}.
 

### 2021-10-11 doc/JFLA2022/paper.tex
fix a few typos

@@
 Importing snippets into a LaTeX document was not one of the original use cases of \alectr, so we had to extend it: the original \alectr did not have support for exporting to LaTeX, and it was towards documenting individual source files (a style where code and prose are interleaved within the same file and code is documented in the same order as it is compiled).
-We extended {alectr to support our needs by implementing a LaTeX backend, and by programming it to generate individual snippet files, one per \texttt{snippet} comment block.
+We extended \alectr to support our needs by implementing a LaTeX backend, and by programming it to generate individual snippet files, one per \texttt{snippet} comment block.
 The later part was straightforward: all it took was to build a custom \alectr \textit{driver}, a small (about 100 lines) Python program that leverages most of the \alectr toolchain but defines a custom frontend that understands our snippet annotations (and otherwise exposes the exact same command line as \alectr).
@@
 
-\inputsnippets{Schutte/Ex42d, Schutte/Ex42c, Schutte/Ex42e}
+\inputsnippets{%Schutte/Ex42d,
+  Schutte/Ex42c, Schutte/Ex42e}
 
@@
 For the mathematician, it may give a concrete view of the mathematical content, not only through full proofs, but also through significative computations: examples, functions associated with constructive proofs, etc.
-For the \coq user, it provides a consistent set of examples,
-allowing to present and compare various formalization and proving techniques. It may be also used as a library, through opam packages.
-It is also a medium sized library (more than 50K LOC), dependent on various tools and libraries of the \coq ecosystem.
-It may be also used to experiment new techniques of continuous maintenance of the code and its documentation.
+For the \coq learner, it provides a consistent set of examples,
+allowing to present and compare various formalization and proving techniques. %It may be also used as a library, through opam packages.
+It is also a medium sized library (more than 50K LOC), dependent on various tools and libraries of the \coq ecosystem, which  may be also used to experiment new techniques of continuous maintenance of the code and its documentation.
 

### 2021-10-11 doc/JFLA2022/paper.tex
fix a few typos

@@
 Since the set \textit{AP} of additive principals  is closed under addition
-(by Lemma \textit{AP\_plus\_closed}), it suffices to prove the inequalities $\omega<\phi_0(2)$ and $42<\phi_0(2)$.
+(by Lemma \textit{AP\_plus\_closed}), it suffices to prove the inequalities $\omega<\omega^2$ and $42<\omega^2$.
 

### 2021-10-12 doc/JFLA2022/paper.tex
make background flowing text

@@
 
-\begin{itemize}
-\item formalizations of mathematics in proof assistants, such as \coq, continually grow in size and scope
-\begin{itemize}
-\item the Mathematical Components library and the four-color theorem and the odd order theorem in \coq
-\item the Mathlib library for Lean
-\end{itemize}
-\item to continue to be useful as examples and reusable libraries, formalizations must be well \emph{documented} and \emph{maintained} to ensure they work with new proof assistant releases.
-%\TODO[Pierre]{add evolution of styles of proof and formalization (not limited to Coq litterature)?
-}
-\item Lean and Mathematical Components have accompanying books, which are written mainly to onboard new library users
-\item book authors faced with several design choices: literate documents, using tools to keep the book and the proof assistant synchronized
-\end{itemize}
+Libraries of formalized mathematics based on proof assistants, such as \coq, Lean, and Isabelle/HOL, are continually growing in size and scope.
+For example, the core projects in the Mathematical Components family for Coq amount to more than 160,000 lines of code and 11,000 lemmas~\cite{Nie2020a}, while Lean's Mathlib consists of more than 140,000 lines of code and 34,000 declarations~\cite{Mathlib}.
+However, after a key mathematical definition or result is added to a library, it needs be well \emph{documented} and \emph{maintained}.
+
+Maintenance includes not only adaptation to changes in new proof assistant versions, but also reorganization to ease future maintenance effort and accommodate new contributions.
+Documentation is usually two-fold: \emph{source code comments} that describe specific definitions and results in-line and \emph{books} that carefully introduce the library and its idioms to beginners~\cite{MCB}.
+The former tends to be terse and dense, while the latter is more long-winded and exhaustive.
+
+As a library changes and expands, its documentation needs to be made consistent and complete.
+Authors use a wide range of techniques, including literate programming and custom tools, to check code snippets and generate output from the proof assistant.
 

### 2021-10-12 doc/JFLA2022/paper.tex
pass over intro

@@
 For example, the core projects in the Mathematical Components family for Coq amount to more than 160,000 lines of code and 11,000 lemmas~\cite{Nie2020a}, while Lean's Mathlib consists of more than 140,000 lines of code and 34,000 declarations~\cite{Mathlib}.
-However, after a key mathematical definition or result is added to a library, it needs be well \emph{documented} and \emph{maintained}.
+However, after a key mathematical definition or result is added to a library, it needs be \emph{documented} and \emph{maintained}.
 
 Maintenance includes not only adaptation to changes in new proof assistant versions, but also reorganization to ease future maintenance effort and accommodate new contributions.
-Documentation is usually two-fold: \emph{source code comments} that describe specific definitions and results in-line and \emph{books} that carefully introduce the library and its idioms to beginners~\cite{MCB}.
-The former tends to be terse and dense, while the latter is more long-winded and exhaustive.
+Documentation is usually two-fold: \emph{source code comments} that describe specific definitions and results in-line and \emph{books} that carefully introduce the library and its idioms~\cite{MCB}.
+The former tends to be terse and dense and serves experienced users, while the latter is more long-winded and exhaustive and serves beginners.
+As a library changes and expands, all its documentation needs to be made consistent and complete.
+Authors use a wide range of techniques, including literate programming~\cite{Knuth1984} and custom tools, to pretty-print and check source code snippets and generate proof assistant output.
 
-As a library changes and expands, its documentation needs to be made consistent and complete.
-Authors use a wide range of techniques, including literate programming and custom tools, to check code snippets and generate output from the proof assistant.
+While books that document proof assistant libraries are valuable to beginners, the purpose of most current books is mainly instrumental, i.e., to teach a certain topic or technique.
+We believe that such books can become \emph{ends-in-themselves}---not just sources of learning, but also of inspiration and entertainment.
 

### 2021-10-12 doc/JFLA2022/paper.tex
more refs and expansion of gaia lineage

@@
 As a library changes and expands, all its documentation needs to be made consistent and complete.
-Authors use a wide range of techniques, including literate programming~\cite{Knuth1984} and custom tools, to pretty-print and check source code snippets and generate proof assistant output.
+Authors use a wide range of techniques, including literate programming~\cite{Knuth1984} and custom tools~\cite{vanDoorn2020}, to pretty-print and check source code snippets and generate proof assistant output.
 
@@
 \subsection{Vision}
-The \Hydras project, part of \community~\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
+The \Hydras project, part of \community~\cite{CoqCommunity}, aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
 \community is a community organization that we have founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting \coq packages, and working collaboratively on documentation projects.
@@
 We plan to extend our libraries in two main directions.
-The \gaia project~\cite{Gaia}, also maintained on \community, contains a development by José Grimm~\cite{grimm:hal-00911710},  dedicated  to the implementation in \coq (SSreflect) of books from  Bourbaki's Elements of Mathematics.
-It contains data structures which are compatible with our ordinal notations.
+
+The \gaia project~\cite{Gaia}, also maintained on \community, contains a development by José Grimm~\cite{grimm:hal-00911710},  dedicated  to the implementation in \coq (SSReflect) of books from  Bourbaki's Elements of Mathematics.
+The set theory underpinning \gaia is derived from an earlier development by Carlos Simpson~\cite{Simpson2004,CatsZFCContrib}.
+\gaia contains data structures which are compatible with our ordinal notations.
 Our plan is to build a  bridge between  the combinatorial results of \Hydras and the set-theoretic content of \gaia, and make possible the transfer of theorems between both libraries.

### 2021-10-12 doc/JFLA2022/paper.tex
small pass over monorepo

@@
 %
-One of the only monorepo in the Coq ecosystem was the math-comp repository, and it used a custom build infrastructure to make this possible.
+One of the few early monorepos in the Coq ecosystem was the Mathematical Components (math-comp) repository, which used a custom build infrastructure to check multiple packages on each commit.
 %
-However, recent improvements in tooling have made their use possible, and therefore we have used them more and more, for instance within \community.
+However, recent improvements in tooling have made their use easier, and therefore we have used them more and more, for instance within \community.
 

### 2021-10-12 doc/JFLA2022/paper.tex
Proposal of an improved abstract.

@@
   former presentations at JFLA meetings.
-  Then, we show how the structure of the project is determined   by two  requirements which must be continuously satisfied:
-   maintenance of the library with respect to several \community projects it requires and evolutions of \coq, and 
-   consistency of the book with respect to the library.
-   \TODO[Pierre]{Find a better/simpler sentence about the project's structure?}
+  Then, we show how the structure of the project is determined   by two  requirements which must be continuously satisfied.
+   First, the Coq code needs to be continuously maintained for compatibility with its dependencies (the Coq proof assistant and several Coq packages from inside and outside \community) and reverse dependencies (\community projects that depend on it).
+   Second, the book needs to be maintained consistent with respect to the evolutions of the Coq code.
+   %\TODO[Pierre]{Find a better/simpler sentence about the project's structure?}
 \end{abstract}

### 2021-10-12 doc/JFLA2022/snippets/Schutte/Ex42b.tex
jfla: Rebuild snippets
Proposal of an improved abstract.

--- 2021-09-20 doc/JFLA2022/Ex42b.tex
@@
     \begin{input}
-      ~~\PY{n+nb}{assert}~\PY{o}{(}\PY{n}{HAP}\PY{o}{:}~\PY{n}{In}~\PY{n}{AP}~\PY{o}{(}\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{))}~\PY{n+nb+bp}{by}~\PY{n+nb}{apply}~\PY{n}{AP\PYZus{}phi0}\PY{o}{.}\nl
+      ~~\PY{n+nb}{assert}~\PY{o}{(}\PY{n}{HAP}\PY{o}{:=}~\PY{n}{AP\PYZus{}phi0}~\PY{l+m+mi}{2}\PY{o}{).}\nl
     \end{input}
@@
     \begin{input}
-      ~~\PY{n+nb}{elim}~~\PY{n}{HAP}\PY{o}{;}~\PY{n+nb}{intros}~~\PY{n}{\PYZus{}}~\PY{n}{H0}\PY{o}{;}~\PY{n+nb}{apply}~\PY{n}{H0}\PY{o}{;}~\PY{n+nb}{clear}~\PY{n}{H0}\PY{o}{.}
+      ~~\PY{n+nb}{elim}~~\PY{n}{HAP}\PY{o}{;}~\PY{n+nb}{intros}~\PY{n}{\PYZus{}}~\PY{n}{H0}\PY{o}{;}~\PY{n+nb}{apply}~\PY{n}{H0}\PY{o}{;}~\PY{n+nb}{clear}~\PY{n}{H0}\PY{o}{.}
     \end{input}

### 2021-10-12 doc/JFLA2022/paper.tex
Remove redundant continously.

@@
   Then, we show how the structure of the project is determined   by two  requirements which must be continuously satisfied.
-   First, the Coq code needs to be continuously maintained for compatibility with its dependencies (the Coq proof assistant and several Coq packages from inside and outside \community) and reverse dependencies (\community projects that depend on it).
+   First, the Coq code needs to be maintained for compatibility with its dependencies (the Coq proof assistant and several Coq packages from inside and outside \community) and reverse dependencies (\community projects that depend on it).
    Second, the book needs to be maintained consistent with respect to the evolutions of the Coq code.

### 2021-10-12 doc/JFLA2022/paper.tex
Fix confusion between project and library.

@@
 Computability is a key topic in Computer Science teaching.
-Moreover O'Connor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter (chapter 9 of~\cite{HydraBook}) to this formalization, with comments on the definitions and proofs, [counter]-examples and exercises, and  extract part of the Goedel library into a new sub-library of the \Hydras library: Ackermann.
+Moreover O'Connor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter (chapter 9 of~\cite{HydraBook}) to this formalization, with comments on the definitions and proofs, [counter]-examples and exercises, and  extract part of the Goedel library into a new sub-library of the Hydra-battles library: Ackermann.
 

### 2021-10-12 doc/JFLA2022/paper.tex
Add missing word.

@@
 
-Importing snippets into a LaTeX document was not one of the original use cases of \alectr, so we had to extend it: the original \alectr did not have support for exporting to LaTeX, and it was towards documenting individual source files (a style where code and prose are interleaved within the same file and code is documented in the same order as it is compiled).
+Importing snippets into a LaTeX document was not one of the original use cases of \alectr, so we had to extend it: the original \alectr did not have support for exporting to LaTeX, and it was geared towards documenting individual source files (a style where code and prose are interleaved within the same file and code is documented in the same order as it is compiled).
 We extended \alectr to support our needs by implementing a LaTeX backend, and by programming it to generate individual snippet files, one per \texttt{snippet} comment block.

### 2021-10-12 doc/JFLA2022/paper.tex
Invert two paragraphs in Alectryon section as suggested by Jérémy.

@@
 
-The transition happened gradually, over a few weeks: for each snippet of \coq code that was in the book, we had to take the following steps:
+Importing snippets into a LaTeX document was not one of the original use cases of \alectr, so we had to extend it: the original \alectr did not have support for exporting to LaTeX, and it was geared towards documenting individual source files (a style where code and prose are interleaved within the same file and code is documented in the same order as it is compiled).
+We extended \alectr to support our needs by implementing a LaTeX backend, and by programming it to generate individual snippet files, one per \texttt{snippet} comment block.
+The later part was straightforward: all it took was to build a custom \alectr \textit{driver}, a small (about 100 lines) Python program that leverages most of the \alectr toolchain but defines a custom frontend that understands our snippet annotations (and otherwise exposes the exact same command line as \alectr).
+
+Once the infrastructure was in place, the transition happened gradually, over a few weeks: for each snippet of \coq code that was in the book, we had to take the following steps:
 
@@
 \end{enumerate}
-
-Importing snippets into a LaTeX document was not one of the original use cases of \alectr, so we had to extend it: the original \alectr did not have support for exporting to LaTeX, and it was geared towards documenting individual source files (a style where code and prose are interleaved within the same file and code is documented in the same order as it is compiled).
-We extended \alectr to support our needs by implementing a LaTeX backend, and by programming it to generate individual snippet files, one per \texttt{snippet} comment block.
-The later part was straightforward: all it took was to build a custom \alectr \textit{driver}, a small (about 100 lines) Python program that leverages most of the \alectr toolchain but defines a custom frontend that understands our snippet annotations (and otherwise exposes the exact same command line as \alectr).
 

### 2021-10-12 doc/JFLA2022/paper.tex
Include a paragraph on Docker-Coq-Action, expand comparison with Nix-based CI.

@@
 
-\paragraph{The Dune build system}
-It allows building packages contained in a single source tree separately, which is essential to be able to publish the different libraries contained in a single repository as separate (opam) packages to the Coq package index.
+\paragraph{The Dune build system~\cite{dune}}
+This build system, originally written for OCaml, but that was extended to support Coq projects, allows building packages contained in a single source tree separately, which is essential to be able to publish the different libraries contained in a single repository as separate (opam) packages to the Coq package index.
 
@@
 
+\paragraph{The Docker-Coq-Action~\cite{martindorel2020dockercoq}}
+This GitHub Action provides a very simple way of setting up Continuous Integration (CI) for a Coq project with an opam file.
+%
+At the current time, we rely on it, together with the mathcomp Docker images~\cite{mathcompdocker}, to test our two main libraries, Hydra-battles and Addition-chains, with multiple versions of Coq.
+
 \paragraph{The Coq Nix Toolbox~\cite{cohen:hal-03366644}}
-We rely on this toolbox mainly because it can generate a Continuous Integration setup (relying on GitHub Actions) to test the various libraries contained in the project, but also their dependencies.
+This toolbox, based on the Nix package manager, provides an alternative way of setting up CI for a Coq project (also relying on GitHub Actions).
 %
-This has allowed us, for instance, to test the Goedel library with the current version of the Hydra-battles library when we have moved a subpart of Goedel to Hydra-battles (the Ackermann sub-library).
+We use this Nix-based CI for several things.
 
-The built artifacts are stored in the \community binary cache on Cachix~\cite{cachix}.
+\begin{enumerate}
+\item To test the build of the project (as a single unit) with coq\_makefile (whereas the Docker-based CI relies on the opam packages, which use Dune to build the libraries contained in the project).
+
+\item To generate the documentation of the project (book in PDF format and coqdoc HTML pages) by relying on the output of the coq\_makefile build.
+
+\item To test the bridge to the Gaia library, because the Nix-based CI supports out-of-the-box caching of build dependencies.
+%
+Given that Gaia takes more than 5 minutes to build, this build is only done once, then reused at each new run of the CI.
+
+\item To test compatibility with the Goedel library, which was made to depend on the Hydra-battles library since the Ackermann sub-library was moved from the former to the latter.
+%
+For this, we rely on the fact that the Coq Nix Toolbox has native support for generating a CI configuration that includes reverse dependency compatibility testing.
+\end{enumerate}
+
+The artifacts of the Nix builds are stored in the \community binary cache on Cachix~\cite{cachix}.
 %
 This means that a given version never has to be built twice and Continuous Integration can be almost instantaneous when no change has been made (e.g., after merging a pull request).
+
+By relying on two different technologies for CI, we are able to fit a larger range of use cases, while also providing more assurance that the project does build correctly in a variety of configurations.
+%
+Given that the two technologies are well maintained (within \community), relying on both does not incur a significant cost, compared to the benefits they provide.
 

### 2021-10-12 doc/JFLA2022/paper.tex
pass over abstract

@@
 \begin{abstract}
-  \Hydras is a collaborative library on discrete math, written for the \coq  proof assistant, and developed under the \community project. The proof scripts are
-  accompanied with an electronic book, made with the help of the \alectr litterate proving tool.
-  We present the evolution of the mathematical contents since
-  former presentations at JFLA meetings.
-  Then, we show how the structure of the project is determined   by two  requirements which must be continuously satisfied.
-   First, the Coq code needs to be maintained for compatibility with its dependencies (the Coq proof assistant and several Coq packages from inside and outside \community) and reverse dependencies (\community projects that depend on it).
-   Second, the book needs to be maintained consistent with respect to the evolutions of the Coq code.
-   %\TODO[Pierre]{Find a better/simpler sentence about the project's structure?}
+  \Hydras is a collaborative library of discrete mathematics,
+  written for the \coq proof assistant and developed as part
+  of the \community project on GitHub. The Coq code is accompanied
+  by an electronic book, generated with the help of the \alectr
+  literate proving tool. We present the evolution of the mathematical
+  contents of the library since former presentations at JFLA meetings.
+  Then, we describe how the structure of the project is determined
+  by two requirements which must be continuously satisfied.
+  First, the Coq code needs to be compatible with its ever-evolving
+  dependencies (the Coq proof assistant and several Coq packages both from
+  inside and outside \community) and its reverse dependencies (\community
+  projects that depend on it). Second, the book needs to be consistent
+  with the library Coq code, which undergoes frequent changes to improve
+  structure and include new definitions and results.
 \end{abstract}
@@
 For example, the core projects in the Mathematical Components family for Coq amount to more than 160,000 lines of code and 11,000 lemmas~\cite{Nie2020a}, while Lean's Mathlib consists of more than 140,000 lines of code and 34,000 declarations~\cite{Mathlib}.
-However, after a key mathematical definition or result is added to a library, it needs be \emph{documented} and \emph{maintained}.
+However, after a key mathematical definition or result is added to a library, it needs be \emph{documented} and \emph{maintained}~\cite{vanDoorn2020}.
 
@@
 As a library changes and expands, all its documentation needs to be made consistent and complete.
-Authors use a wide range of techniques, including literate programming~\cite{Knuth1984} and custom tools~\cite{vanDoorn2020}, to pretty-print and check source code snippets and generate proof assistant output.
+Authors use a wide range of techniques, including literate programming~\cite{Knuth1984} and custom tools, to pretty-print and check source code snippets and generate proof assistant output.
 
-While books that document proof assistant libraries are valuable to beginners, the purpose of most current books is mainly instrumental, i.e., to teach a certain topic or technique.
-We believe that such books can become \emph{ends-in-themselves}---not just sources of learning, but also of inspiration and entertainment.
+While books that document proof assistant libraries are valuable to beginners, the purpose of current books is mainly instrumental, i.e., to teach a certain topic or technique.
+We believe that such books can become \emph{ends in themselves}---not just sources of learning, but also sources of inspiration and entertainment.
 

### 2021-10-12 doc/JFLA2022/paper.tex
one more pass over abstract, use trick to avoid period spacing with project name

@@
 \newcommand{\equations}{Equations\xspace}
-\newcommand{\Hydras}{Hydras \& Co.\xspace}
+\newcommand{\Hydras}{Hydras \& Co$\text.$\xspace}
 %\newcommand{\Hydras}{Hydra-battles\xspace}
@@
 \begin{abstract}
-  \Hydras is a collaborative library of discrete mathematics,
-  written for the \coq proof assistant and developed as part
-  of the \community project on GitHub. The Coq code is accompanied
+  \Hydras is a collaborative library of discrete mathematics for the
+  \coq proof assistant, developed as part of the \community project
+  on GitHub. The Coq code is accompanied
   by an electronic book, generated with the help of the \alectr
@@
   projects that depend on it). Second, the book needs to be consistent
-  with the library Coq code, which undergoes frequent changes to improve
-  structure and include new definitions and results.
+  with the Coq code, which undergoes frequent changes to improve
+  structure and include new definitions and results. We believe \Hydras demonstrates
+  that books on formalized mathematics are not limited to
+  providing exposition of theories and reasoning techniques---they
+  can also provide inspiration and entertainment that transcends educational goals.
 \end{abstract}
@@
 
-While books that document proof assistant libraries are valuable to beginners, the purpose of current books is mainly instrumental, i.e., to teach a certain topic or technique.
-We believe that such books can become \emph{ends in themselves}---not just sources of learning, but also sources of inspiration and entertainment.
+While books that document proof assistant libraries are valuable to beginners, the purpose of recently published books is mainly instrumental, i.e., to teach a certain topic or technique.
+We believe that books can instead become \emph{ends in themselves}---not just sources of learning, but also sources of inspiration and entertainment.
 

### 2021-10-12 doc/JFLA2022/paper.tex
one more abstract pass

@@
   dependencies (the Coq proof assistant and several Coq packages both from
-  inside and outside \community) and its reverse dependencies (\community
+  inside and outside \community) and reverse dependencies (\community
   projects that depend on it). Second, the book needs to be consistent
   with the Coq code, which undergoes frequent changes to improve
-  structure and include new definitions and results. We believe \Hydras demonstrates
+  structure and include new material. We believe \Hydras demonstrates
   that books on formalized mathematics are not limited to

### 2021-10-12 doc/JFLA2022/paper.tex
make abstract and vision more consistent

@@
   \Hydras is a collaborative library of discrete mathematics for the
-  \coq proof assistant, developed as part of the \community project
+  \coq proof assistant, developed as part of the \community organization
   on GitHub. The Coq code is accompanied
@@
 For example, the core projects in the Mathematical Components family for Coq amount to more than 160,000 lines of code and 11,000 lemmas~\cite{Nie2020a}, while Lean's Mathlib consists of more than 140,000 lines of code and 34,000 declarations~\cite{Mathlib}.
-However, after a key mathematical definition or result is added to a library, it needs be \emph{documented} and \emph{maintained}~\cite{vanDoorn2020}.
+However, after a key mathematical definition or result is added to a library, it needs be \emph{documented}~\cite{alectryonpaper} and \emph{maintained}~\cite{vanDoorn2020}.
 
@@
 \subsection{Vision}
-The \Hydras project, part of \community~\cite{CoqCommunity}, aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.
-\community is a community organization that we have founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting \coq packages, and working collaboratively on documentation projects.
+The \Hydras project, part of \community on GitHub~\cite{CoqCommunity}, aims to be an experimental platform for the collaborative development of documented libraries of formal proofs.
+\community is a community organization that we founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting \coq packages, and working collaboratively on documentation projects.
 The \Hydras project demonstrates that these two goals are not independent: interesting \coq packages can become the basis for new documentation.
 %
-This umbrella project now includes evolved versions of the former Cantor and Additions libraries (under the new names of Hydra-battles and Addition-chains), the Ackermann sub-library, extracted from Russel O'Connor's Goedel library~\cite{OConnor05, Goedel} and a bridge to the \gaia library (by José Grimm~\cite{Gaia,grimm:hal-00911710}).
+This umbrella project now includes evolved versions of the former Cantor~\cite{CantorContrib} and Additions~\cite{AdditionsContrib} libraries (under the new names of Hydra-battles and Addition-chains), the Ackermann sub-library which was extracted from Russel O'Connor's Goedel library~\cite{OConnor05, Goedel}, and a bridge to the \gaia library by José Grimm~\cite{Gaia,grimm:hal-00911710}.
 %
-By following this approach of commenting interesting \coq packages, we provide new documentation content, that contributes to the diversity of the thriving \coq documentation ecosystem.
+By following this approach of commenting interesting \coq packages, we provide new documentation content that contributes to the diversity of the thriving \coq documentation ecosystem.
 

### 2021-10-12 doc/JFLA2022/paper.tex
one more pass over vision

@@
 %
-This umbrella project now includes evolved versions of the former Cantor~\cite{CantorContrib} and Additions~\cite{AdditionsContrib} libraries (under the new names of Hydra-battles and Addition-chains), the Ackermann sub-library which was extracted from Russel O'Connor's Goedel library~\cite{OConnor05, Goedel}, and a bridge to the \gaia library by José Grimm~\cite{Gaia,grimm:hal-00911710}.
+This umbrella project now includes evolved versions of the former Cantor and Additions libraries~\cite{CantorContrib,AdditionsContrib} (under the new names of Hydra-battles and Addition-chains), the Ackermann sub-library extracted from Russel O'Connor's Goedel library~\cite{OConnor05, Goedel}, and a bridge to the \gaia library by José Grimm~\cite{Gaia,grimm:hal-00911710}.
 %
-By following this approach of commenting interesting \coq packages, we provide new documentation content that contributes to the diversity of the thriving \coq documentation ecosystem.
+By following this approach of commenting interesting \coq packages, we provide new documentation content that contributes to the diversity of the thriving \coq ecosystem.
 
@@
 
-We solved this maintenance issue by moving to \alectr, a tool to automatically record \coq proofs.
+We solved this maintenance issue by moving to \alectr, a tool that automatically records \coq proofs.
 Instead of copy-pasting fragments into LaTeX, we now embed small LaTeX files automatically generated by \alectr from our \coq development.

### 2021-10-12 doc/JFLA2022/paper.tex
add an introduction to hydras

--- 2021-10-12 doc/JFLA2022/paper.tex
@@
 \subsection{Hydra games}
+
+We chose to build our library and book on two simple themes which allow many variations: computing powers in a monoid, and Kirby and Paris' hydra battles, formerly user contributions of the Coq system.
+In the interest of space, we shall present only the second one in this article.
+
 Hydra games (also known as \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians, 

### 2021-10-12 doc/JFLA2022/paper.tex
add an introduction to Section 1.3

@@
 
-We chose to build our library and book on two simple themes which allow many variations: computing powers in a monoid, and Kirby and Paris' hydra battles, formerly user contributions of the Coq system.
+We chose to build our library and book on two simple themes which allow many variations: computing powers in a monoid, and Kirby and Paris' hydra battles.
 In the interest of space, we shall present only the second one in this article.

### 2021-10-12 doc/JFLA2022/paper.tex
complete pass over section 3

@@
 \subsection{Hydra games}
-
-We chose to build our library and book on two simple themes which allow many variations: computing powers in a monoid, and Kirby and Paris' hydra battles.
-In the interest of space, we shall present only the second one in this article.
-
 Hydra games (also known as \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians, 
@@
 
-\section{Integration within \community}
+\section{Integration with \community}
 
-\subsection{Presentation}
+\subsection{Background}
 
-\community is an informal organization run by volunteer Coq users with objectives to maintain interesting Coq projects in the long-term and help Coq users to collaborate on documentation, tooling, etc.
+\community is an informal organization run by volunteer Coq users on GitHub that aims to maintain interesting Coq projects for the long term and facilitate collaboration among Coq users on documentation, tooling, etc.
+\community was created in 2018, inspired by the {Elm Community} organization~\cite{zimmermann:tel-02451322}.
+%
+Such ``Community Package Maintenance Organizations'' exist in many software ecosystems, as they avoid the common problem of an important package becoming unmaintained after its author has moved on to other projects or has disappeared~\cite{zimmermann2021grounded}.
 
-It was created in 2018, inspired by the {Elm Community} organization~\cite{zimmermann:tel-02451322}.
+In the case of \coq, this problem is likely even more prevalent than in other ecosystems, as many packages are created by graduate students or researchers for a specific paper and not planned to be maintained for the long term by authors.
 %
-Such ``Community Package Maintenance Organizations'' actually exist in many ecosystems as they avoid the very common problem of an important package becoming unmaintained as its author has moved on or disappeared~\cite{zimmermann2021grounded}.
-
-In the case of \coq, this problem may be even more prevalent as many packages are created by PhD students, or by researchers for a specific paper, but their authors do not intend to maintain the package in the long-term.
-%
-The authors are however generally very open to having someone else who expresses interest in their work continue the maintenance.
+However, authors are generally open to having someone else who expresses interest in their work take over package maintenance.
 %
 
-\community makes that easy by providing a process for transferring or forking an unmaintained package, tooling for setting up good maintenance practices (such as continuous integration) and by making it possible for someone to take over a package without committing in the long-term (as maintainers who drop out can easily be replaced by some other volunteers).
+\community makes it easy to change maintainers by defining a process for transferring or forking an unmaintained package, tooling for setting up good maintenance practices (such as continuous integration), and by making it possible for someone to take over a package without making a long-term commitment (as maintainers who drop out can easily be replaced by some other volunteers).
 
-Today, \community hosts over 50 projects maintained by over 30 maintainers.
+As of 2021, \community hosts over 50 projects maintained by over 30 volunteers.
 %
@@
 %
-Some had been maintained in the past by the \coq development team on behalf of the authors, but this meant that only minimal changes required to make the project build with new \coq versions were conducted.
+Some had been maintained in the past by the \coq development team on behalf of the authors, but this meant that only minimal changes required to make the project build with new \coq versions were performed.
 %
-Some were still maintained by their authors but they wanted help to keep them up-to-date.
+Some were still maintained by their original authors, but were transferred to enable other users to help out with maintenance and facilitate adoption of best practices on, e.g., continuous integration.
 %
-Others were simply left unmaintained and were revived by their adoption in \community.
+Others were simply unmaintained and were revived after their transfer to \community.
 
@@
 
-Given the objectives of \community, it makes sense to propose a transfer anytime we encounter an interesting project that is insufficiently maintained.
+Given the objectives of \community, we believe it makes sense to propose a transfer anytime we encounter an interesting Coq project that is insufficiently maintained.
 %
-After the project is adopted, large changes and refactorings are explicitly allowed.
+After a transfer, the \community maintainers are explicitly allowed to perform large changes and refactorings.
 %
-This means, for instance, that we can consolidate libraries together or split them into several packages if it makes sense.
+This means, for instance, that we can consolidate libraries by merging them, or split them up into several packages.
 
-In order to prove formally that the length of the considered
-kind of battles is not given by a primitive recursive function, we used the formalization of primitive recursive functions, part
-of Russell O'Connor's contribution on G\"{o}del first incompleteness theorem~\cite{OConnor05, Goedel}.
-For this purpose, and above all by consideration of the scientific interest of this contribution, we chose to host and maintain this work within \community.
+In order to prove formally that the length of the
+kind of hydra battles we consider is not given by a primitive recursive function, we chose to use a formalization of the primitive recursive functions that was originally part of Russell O'Connor's statement and proof of G\"{o}del's first incompleteness theorem~\cite{OConnor05, Goedel}.
+For this purpose, and above all in consideration of the scientific interest of this contribution, we decided to host and maintain O'Connor's work in \community.
 
-Computability is a key topic in Computer Science teaching.
-Moreover O'Connor's library is a nice illustration of dependently typed programming, so we chose to devote a full chapter (chapter 9 of~\cite{HydraBook}) to this formalization, with comments on the definitions and proofs, [counter]-examples and exercises, and  extract part of the Goedel library into a new sub-library of the Hydra-battles library: Ackermann.
+Since computability is a key topic in computer science teaching and O'Connor's library is a nice illustration of dependently typed programming, we decided to devote a full chapter (chapter 9 of~\cite{HydraBook}) to the formalization of primitive recursive functions, with comments on the definitions and proofs and [counter-]examples and exercises.
+As part of the writing process, we made the formalization into a new sub-library of the Hydra-battles library, dubbed Ackermann.
 

### 2021-10-12 doc/JFLA2022/paper.tex

@@
 \subsection{Hydra games}
+
+We chose to build our library and book on two simple themes which allow many variations: computing powers in a monoid, and Kirby and Paris' hydra battles.
+In the interest of space, we shall present only the second one in this article.
+
 Hydra games (also known as \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians, 

### 2021-10-12 doc/JFLA2022/paper.tex
add acknowledgment section

--- 2021-10-12 doc/JFLA2022/paper.tex
@@
 
-%\section{Acknowledgments}
-%\label{sect:acks}
+\section{Acknowledgments}
+\label{sect:acks}
+We are grateful to the original authors and current maintainers of the Coq packages we use and depend on: José Grimm (Gaia), Russel O'Connor(Goedel), Matthieu Sozeau (Equations), the Mathematical Components team (ssreflect, algebra), Marc Lasson and Chantal Keller (Paramcoq), and the authors and maintainers
+of \coq and its associated tools.
 

### 2021-10-13 doc/JFLA2022/paper.tex
small reorganization of section 3
add history of Gaia

@@
 
-\subsection{Interactions with other \community projects}
-
 Given the objectives of \community, we believe it makes sense to propose a transfer anytime we encounter an interesting Coq project that is insufficiently maintained.
@@
 This means, for instance, that we can consolidate libraries by merging them, or split them up into several packages.
+
+\subsection{Integration of primitive recursive functions}
 
@@
 As part of the writing process, we made the formalization into a new sub-library of the Hydra-battles library, dubbed Ackermann.
+
+\subsection{Bridge to ordinals in Gaia}
+
+The Gaia project by Jos\'e Grimm aimed to formalize mathematics in the style of by N.
+Bourbaki in Coq.
+The formalization of the first book by Bourbaki, on the theory of sets, was initially described in a technical report in July 2009~\cite{Grimm2009a}.
+Grimm then wrote (and continually updated) technical reports describing the formalization of Bourbaki's two subsequent books~\cite{Grimm2009b,Grimm2016} and additional topics in number theory~\cite{grimm:hal-00911710,Grimm2014}, before he passed away in 2019.
+
+In 2020, members of \community transferred the Gaia source code to GitHub and adapted it for recent releases of the Mathematical Components library, which Gaia heavily relies on.
+A volunteer finished the only in-progress proof left by Grimm.
+At around 165,000 lines of code, Gaia is currently one of the largest maintained open source Coq projects~\cite{Gaia}.
+
+\subsection{\Hydras Genealogy and Dependencies}
 

### 2021-10-13 doc/JFLA2022/paper.tex
beginning of genealogy section

@@
 
+\begin{center}
+{\footnotesize
+\begin{tikzpicture}[>=stealth',semithick]
+
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm] (cantor) { Cantor };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3cm] (additions) { Additions };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=3cm] (additions) { Goedel };
+
+\end{tikzpicture}
+}
+\end{center}
+
 % We could already talk a bit of the monorepo structure here, and
@@
 
+\section{Comparison of \Hydras with other Coq books}
+
 \section{Conclusion and perspectives}

### 2021-10-13 doc/JFLA2022/paper.tex
small pass over infrastructure

@@
 
-\paragraph{The Dune build system~\cite{dune}}
-This build system, originally written for OCaml, but that was extended to support Coq projects, allows building packages contained in a single source tree separately, which is essential to be able to publish the different libraries contained in a single repository as separate (opam) packages to the Coq package index.
+\paragraph{The Dune build system~\cite{dune}.}
+Dune was originally designed to build OCaml projects, but was recently extended to support Coq.
+Dune allows building packages contained in a single source tree separately, which is essential to be able to publish the different libraries contained in a single repository as separate (opam) packages to the Coq package index.
 
@@
 
-\paragraph{The Docker-Coq-Action~\cite{martindorel2020dockercoq}}
+\paragraph{The Docker-Coq-Action~\cite{martindorel2020dockercoq}.}
 This GitHub Action provides a very simple way of setting up Continuous Integration (CI) for a Coq project with an opam file.
@@
 
-\paragraph{The Coq Nix Toolbox~\cite{cohen:hal-03366644}}
+\paragraph{The Coq Nix Toolbox~\cite{cohen:hal-03366644}.}
 This toolbox, based on the Nix package manager, provides an alternative way of setting up CI for a Coq project (also relying on GitHub Actions).

### 2021-10-13 doc/JFLA2022/paper.tex
basic dependency elements

@@
 In 2020, members of \community transferred the Gaia source code to GitHub and adapted it for recent releases of the Mathematical Components library, which Gaia heavily relies on.
-A volunteer finished the only in-progress proof left by Grimm.
+Volunteers (``collaborators of Nicolas Bourbaki'') then finished the only in-progress proof left by Grimm.
 At around 165,000 lines of code, Gaia is currently one of the largest maintained open source Coq projects~\cite{Gaia}.
@@
 
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm] (cantor) { Cantor };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3cm] (additions) { Additions };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=3cm] (additions) { Goedel };
+\begin{scope}[xshift=-6cm]
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (additions) { Additions };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (goedel) { Goedel };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2cm] (cantor) { Cantor };
+\end{scope}
+
+\begin{scope}[yshift=-0.7cm,xshift=0.25cm]
+\node[rectangle, draw=black,dashed,minimum height=0.5cm,minimum width=1.5cm] (equations) { Equations };
+\node[rectangle, draw=black,dashed,minimum height=0.5cm,minimum width=1.5cm,right of=equations, node distance=2cm] (mathcomp) { MathComp };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,right of=mathcomp, node distance=2cm] (paramcoq) { Paramcoq };
+\end{scope}
+
+\begin{scope}[yshift=-1.8cm,xshift=-4cm]
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=hydras, node distance=2.5cm] (chains) { Addition-chains };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=chains, node distance=2.2cm] (gaia) { Gaia };
+\end{scope}
+
+\begin{scope}[yshift=-2.8cm,xshift=-6cm]
+ \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
+ \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=2.2cm] (gaiahydras) { Gaia-hydras };
+\end{scope}
 

### 2021-10-13 doc/JFLA2022/paper.tex
Fix typo.
more deps in graph

@@
 
-\begin{center}
+\begin{figure}[h]
+\centering
 {\footnotesize
-\begin{tikzpicture}[>=stealth',semithick]
+\begin{tikzpicture}[thick]
 
 \begin{scope}[xshift=-6cm]
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (additions) { Additions };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (goedel) { Goedel };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2cm] (cantor) { Cantor };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel };    
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2cm] (additions) { Additions };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (cantor) { Cantor };
 \end{scope}
@@
 
+\draw[->,dotted] (additions.south) -- (chains.north west) ;
+\draw[->] (goedel) -- (goedelcc) ;
+\draw[->] (goedel) -- (hydras.north west) ;
+\draw[->] (cantor) -- (hydras.north) ;
+\draw[->] (cantor.south) -- (gaia.north west) ;
+\draw[->] (hydras) -- (gaiahydras) ;
+\draw[->] (gaia.south west) -- (gaiahydras) ;
 \end{tikzpicture}
 }
-\end{center}
+\caption{Genealogy and dependencies for \Hydras packages.}
+  \label{fig:round}
+\end{figure}
+
+In mainstream software, it is common for packages to continually depend on a large number of packages.
+In contrast, projects in the Coq ecosystem are often subject to ``dependency aversion'', where maintainers eschew depending on useful packages because it may lead to work in adapting to upstream changes.
+With \Hydras, we hope to demonstrate that projects with complex dependencies are feasible to manage using recent advances in build management and infrastructure.
 
@@
 
-\section{Modernizing the infrastructure}
+\section{Modernizing the building infrastructure}
 
@@
 \label{sect:acks}
-We are grateful to the original authors and current maintainers of the Coq packages we use and depend on: José Grimm (Gaia), Russel O'Connor(Goedel), Matthieu Sozeau (Equations), the Mathematical Components team (ssreflect, algebra), Marc Lasson and Chantal Keller (Paramcoq), and the authors and maintainers
-of \coq and its associated tools.
+We are grateful to the original authors and current maintainers of the Coq packages we use and depend on: José Grimm (Gaia), Russell O'Connor (Goedel), Matthieu Sozeau (Equations), the Mathematical Components team, Marc Lasson and Chantal Keller (Paramcoq), and the authors and maintainers of \coq and its associated tools.
 

### 2021-10-13 doc/JFLA2022/paper.tex
clarify that gaia fix was anonymous

@@
 
-The Gaia project by Jos\'e Grimm aimed to formalize mathematics in the style of by N.
+The Gaia project by Jos\'e Grimm aimed to formalize mathematics in the style of N.
 Bourbaki in Coq.
@@
 In 2020, members of \community transferred the Gaia source code to GitHub and adapted it for recent releases of the Mathematical Components library, which Gaia heavily relies on.
-Volunteers (``collaborators of Nicolas Bourbaki'') then finished the only in-progress proof left by Grimm.
+Anonymous volunteers (``collaborators of Nicolas Bourbaki'') then finished the only in-progress proof left by Grimm.
 At around 165,000 lines of code, Gaia is currently one of the largest maintained open source Coq projects~\cite{Gaia}.

### 2021-10-13 doc/JFLA2022/paper.tex
small corrections (Section 1.3 and 2)

--- 2021-10-13 doc/JFLA2022/paper.tex
@@
 \item At every round, Hercules chops off one head of the hydra. If the head is at a distance greater than 1 from the root,
-  then some sub-tree $h$ of the hydra is copied a certain amount $n$ of times. The number $n$ of copies and the sub-tree $h$ may depend of the considered variant of the game
-  or the time elapsed since the beginning of the fight.
+  then some sub-tree $h$ of the hydra is copied a certain amount $n$ of times. The number $n$ of copies and the sub-tree $h$ may depend on the considered variant of the game
+  and the time elapsed since the beginning of the fight.
   Figure~\ref{fig:round} shows an example with $n=2$.
@@
 Our proof was based on the construction of a battle, where the
-number of copies at each round is given by the elimination of an existential quantifier. So, it was mandatory to consider the class of \emph{all} hydra battles.
+number of copies at each round was given by the elimination of an existential quantifier.
+So, it was mandatory to consider the class of \emph{all} hydra battles, or, equivalently, the class of battles where the hydra choses arbitrarily  its number $n$ of copies at every round of the battle.
 
@@
 (see for instance~\cite{KP82, bauer2008, BauerHydra}) 
-assume that the hydra grows $n$ copies at step $n$ of the game, which is incompatible with our proof.
+assume that the hydra grows $k$ copies at step $k$ of the game, which is incompatible with our proof.
 We prove now that Theorem~\ref{thm3} still holds with these typical battles.
@@
  \begin{itemize}
-   \item The game starts at an initial step $i\in\mathbb{N}$ (not necessarily $0$).
+   \item The game starts at an initial step $k=i$ where $i$ is any natural number.
    \item  The hydra is always the  representation as a tree of some ordinal strictly below $\epsilon_0$ in Cantor normal
@@
      For instance, Fig~\ref{fig:round} shows the hydras respectively associated with  $\omega^{\omega^2+1}$ and $\omega^{\omega^2}\times 3$.
+     In the following,  the expression  ``the hydra $\alpha$'' is an abbreviation of ``the hydra associated with the ordinal $\alpha$''.
 
@@
 
- In mathematical terms, if at step $n\geq i$, the hydra is associated with the ordinal $\alpha$, at step $n+1$ it is associated with
- $\canonseq{\alpha}{n+1}$, \emph{i.e.} the $(n+1)$th element of the canonical sequence of $\alpha$~\cite{KS81}.
+ In mathematical terms, if at step $k\geq i$, the hydra is associated with the ordinal $\alpha$, at step $k+1$ it is associated with
+ $\canonseq{\alpha}{k+1}$, \emph{i.e.} the $(k+1)$th element of the canonical sequence of $\alpha$~\cite{KS81}.
 
@@
 By a sequence of rewritings and inductions, we prove that the number of steps of the considered battle is greater or equal than $2^{2^N}$, where $N=2^{70}-1$.
-More generally, we prove that, for $\alpha\geq\omega^\omega$, 
-the function computing the length of the battle starting with the configuration $(\alpha,i)$ is not primitive recursive.
+By comparison with the diagonalized Ackermann function $\lambda\,i.\,A(i,i)$, we prove also that, for $\alpha\geq\omega^\omega$, 
+the function computing the length of the battle starting with the hydra $\alpha$ and the initial step $i$ is not primitive recursive.
 
@@
 
-\subsection{Bridge to ordinals in Gaia}
-
-The Gaia project by Jos\'e Grimm aimed to formalize mathematics in the style of by N.
-Bourbaki in Coq.
-The formalization of the first book by Bourbaki, on the theory of sets, was initially described in a technical report in July 2009~\cite{Grimm2009a}.
-Grimm then wrote (and continually updated) technical reports describing the formalization of Bourbaki's two subsequent books~\cite{Grimm2009b,Grimm2016} and additional topics in number theory~\cite{grimm:hal-00911710,Grimm2014}, before he passed away in 2019.
-
-In 2020, members of \community transferred the Gaia source code to GitHub and adapted it for recent releases of the Mathematical Components library, which Gaia heavily relies on.
-Volunteers (``collaborators of Nicolas Bourbaki'') then finished the only in-progress proof left by Grimm.
-At around 165,000 lines of code, Gaia is currently one of the largest maintained open source Coq projects~\cite{Gaia}.
-
-\subsection{\Hydras Genealogy and Dependencies}
-
-\begin{center}
-{\footnotesize
-\begin{tikzpicture}[>=stealth',semithick]
-
-\begin{scope}[xshift=-6cm]
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (additions) { Additions };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (goedel) { Goedel };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2cm] (cantor) { Cantor };
-\end{scope}
-
-\begin{scope}[yshift=-0.7cm,xshift=0.25cm]
-\node[rectangle, draw=black,dashed,minimum height=0.5cm,minimum width=1.5cm] (equations) { Equations };
-\node[rectangle, draw=black,dashed,minimum height=0.5cm,minimum width=1.5cm,right of=equations, node distance=2cm] (mathcomp) { MathComp };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,right of=mathcomp, node distance=2cm] (paramcoq) { Paramcoq };
-\end{scope}
-
-\begin{scope}[yshift=-1.8cm,xshift=-4cm]
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=hydras, node distance=2.5cm] (chains) { Addition-chains };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=chains, node distance=2.2cm] (gaia) { Gaia };
-\end{scope}
-
-\begin{scope}[yshift=-2.8cm,xshift=-6cm]
- \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
- \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=2.2cm] (gaiahydras) { Gaia-hydras };
-\end{scope}
-
-\end{tikzpicture}
-}
-\end{center}
+\subsection{Bridge to alternative ordinal formalization}
 
@@
 
-\inputsnippets{%Schutte/Ex42d,
+\inputsnippets{Schutte/Ex42d, %to remove if short paper
   Schutte/Ex42c, Schutte/Ex42e}
@@
 
-\paragraph{The Dune build system~\cite{dune}.}
-Dune was originally designed to build OCaml projects, but was recently extended to support Coq.
-Dune allows building packages contained in a single source tree separately, which is essential to be able to publish the different libraries contained in a single repository as separate (opam) packages to the Coq package index.
+\paragraph{The Dune build system~\cite{dune}}
+This build system, originally written for OCaml, but that was extended to support Coq projects, allows building packages contained in a single source tree separately, which is essential to be able to publish the different libraries contained in a single repository as separate (opam) packages to the Coq package index.
 
@@
 
-\paragraph{The Docker-Coq-Action~\cite{martindorel2020dockercoq}.}
+\paragraph{The Docker-Coq-Action~\cite{martindorel2020dockercoq}}
 This GitHub Action provides a very simple way of setting up Continuous Integration (CI) for a Coq project with an opam file.
@@
 
-\paragraph{The Coq Nix Toolbox~\cite{cohen:hal-03366644}.}
+\paragraph{The Coq Nix Toolbox~\cite{cohen:hal-03366644}}
 This toolbox, based on the Nix package manager, provides an alternative way of setting up CI for a Coq project (also relying on GitHub Actions).
@@
 Given that the two technologies are well maintained (within \community), relying on both does not incur a significant cost, compared to the benefits they provide.
-
-\section{Comparison of \Hydras with other Coq books}
 

### 2021-10-13 doc/JFLA2022/paper.tex
Section 1.3: an electronic book -> the book (already mentionned)

--- 2021-10-13 doc/JFLA2022/paper.tex
@@
 
-Complex formalisations and proofs are explained in an
-  electronic book~\cite{HydraBook} (PDF document of over 280 pages). Whenever various reasonable choices exist, we try to present and compare the alternatives.
+Complex formalisations and proofs are explained in the book.
+% an  electronic book~\cite{HydraBook} (PDF document of over 280 pages).
+Whenever various reasonable choices exist, we try to present and compare the alternatives.
   For instance, Figures~\ref{fig:Ex42E0} and \ref{fig:Ex42-schutte} show two radically different proofs of the equality

### 2021-10-13 doc/JFLA2022/paper.tex
content in book comparison

--- 2021-10-13 doc/JFLA2022/paper.tex
@@
 
+We believe that current Coq books can broadly be divided into two categories:
+\begin{itemize}
+\item \emph{Executable books} that are in effect large, well-commented Coq programs from which other representations of the material (HTML, PDF) are derived using tools.
+\item \emph{Traditional books} generated from documentation languages such as LaTeX that are accompanied by standalone Coq formalizations.
+\end{itemize}
+
 \section{Conclusion and perspectives}

### 2021-10-13 doc/JFLA2022/paper.tex
fix typos

@@
 
-Unfortunately, the examples  most commonly shown in the litterature
+Unfortunately, the examples  most commonly shown in the literature
 (see for instance~\cite{KP82, bauer2008, BauerHydra}) 
@@
 
-\Hydras wants to be a bridge between scientific litterature (\emph{e.g.}~\cite{KP82, KS81, schutte} and proof assistant technology.
+\Hydras wants to be a bridge between scientific literature (\emph{e.g.}~\cite{KP82, KS81, schutte} and proof assistant technology.
 For the mathematician, it may give a concrete view of the mathematical content, not only through full proofs, but also through significative computations: examples, functions associated with constructive proofs, etc.

### 2021-10-13 doc/JFLA2022/paper.tex
another pass on dependency graph

@@
 
-\subsection{\Hydras Genealogy and Dependencies}
+\subsection{\Hydras Genealogy, Dependencies, and Organization}
 
@@
 \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel };    
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2cm] (additions) { Additions };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (additions) { Additions };
 \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (cantor) { Cantor };
@@
 \begin{scope}[yshift=-0.7cm,xshift=0.25cm]
-\node[rectangle, draw=black,dashed,minimum height=0.5cm,minimum width=1.5cm] (equations) { Equations };
-\node[rectangle, draw=black,dashed,minimum height=0.5cm,minimum width=1.5cm,right of=equations, node distance=2cm] (mathcomp) { MathComp };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,right of=mathcomp, node distance=2cm] (paramcoq) { Paramcoq };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm] (equations) { Equations };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=equations, node distance=2cm] (mathcomp) { MathComp };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=equations, node distance=4.7cm] (paramcoq) { Paramcoq };
 \end{scope}
@@
 
-\draw[->,dotted] (additions.south) -- (chains.north west) ;
-\draw[->] (goedel) -- (goedelcc) ;
-\draw[->] (goedel) -- (hydras.north west) ;
-\draw[->] (cantor) -- (hydras.north) ;
-\draw[->] (cantor.south) -- (gaia.north west) ;
+\draw[->,dotted] (additions.south) -- (chains.north) ;
+\draw[->,dotted] (goedel) -- (goedelcc) ;
+\draw[->,dotted] (goedel) -- (hydras.north west) ;
+\draw[->,dotted] (cantor) -- (hydras.north) ;
+\draw[->,dotted] (cantor.south) -- (gaia.north west) ;
 \draw[->] (hydras) -- (gaiahydras) ;
-\draw[->] (gaia.south west) -- (gaiahydras) ;
+\draw[->] (gaia.south west) -- (gaiahydras.east) ;
+
+\draw[->] (equations) -- (hydras) ;
+\draw[->] (mathcomp) -- (gaia) ;
+\draw[->] (mathcomp) -- (chains) ;
+\draw[->] (equations.south) -- (chains) ;
+\draw[->] (paramcoq.south) -- (hydras) ;
 \end{tikzpicture}
@@
 
-In mainstream software, it is common for packages to continually depend on a large number of packages.
+In traditional software development, it is common for packages to continually depend on a large number of packages.
 In contrast, projects in the Coq ecosystem are often subject to ``dependency aversion'', where maintainers eschew depending on useful packages because it may lead to work in adapting to upstream changes.

### 2021-10-13 doc/JFLA2022/paper.tex
pass over dep graph

@@
 \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington };
 \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (additions) { Additions };
@@
 \begin{scope}[yshift=-0.7cm,xshift=0.25cm]
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm] (equations) { Equations };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=equations, node distance=2cm] (mathcomp) { MathComp };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=equations, node distance=4.7cm] (paramcoq) { Paramcoq };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=2cm] (mathcomp) { MathComp };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=paramcoq, node distance=4.7cm] (equations) { Equations };
 \end{scope}
@@
 \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=hydras, node distance=3.5cm] (pockcc) { Pocklington };
 \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=hydras, node distance=2.5cm] (chains) { Addition-chains };
@@
  \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
- \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=2.2cm] (gaiahydras) { Gaia-hydras };
+ \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=3cm] (gaiahydras) { Gaia-hydras };
 \end{scope}
@@
 \draw[->] (hydras) -- (gaiahydras) ;
+\draw[->] (hydras.south west) -- (goedelcc) ;
 \draw[->] (gaia.south west) -- (gaiahydras.east) ;
@@
 \draw[->] (mathcomp) -- (chains) ;
-\draw[->] (equations.south) -- (chains) ;
-\draw[->] (paramcoq.south) -- (hydras) ;
+\draw[->] (paramcoq.south) -- (chains) ;
+\draw[->] (pockcc) -- (goedelcc) ;
+\draw[->] (pock) -- (pockcc) ;
 \end{tikzpicture}
@@
 
-\section{Modernizing the building infrastructure}
+\section{Modernizing the build infrastructure}
 
@@
 We believe that current Coq books can broadly be divided into two categories:
-\begin{itemize}
-\item \emph{Executable books} that are in effect large, well-commented Coq programs from which other representations of the material (HTML, PDF) are derived using tools.
-\item \emph{Traditional books} generated from documentation languages such as LaTeX that are accompanied by standalone Coq formalizations.
-\end{itemize}
+\begin{enumerate}
+\item[a)] \emph{Executable textbooks} that are in effect large, well-commented Coq programs from which other representations of the material (HTML, PDF) are derived using tools.
+\item[b)] \emph{Traditional textbooks} generated from documentation languages such as LaTeX that are accompanied by standalone Coq formalizations and/or code snippets.
+\end{enumerate}
 

### 2021-10-13 doc/JFLA2022/paper.tex
use grayscale to convey placement

@@
 
-\subsection{\Hydras Genealogy, Dependencies, and Organization}
+\subsection{Project Genealogy, Dependencies, and Organization}
 
@@
 \begin{scope}[yshift=-0.7cm,xshift=0.25cm]
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
+  \node[rectangle, draw=black,fill=gray!30,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
 \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=2cm] (mathcomp) { MathComp };
@@
 \begin{scope}[yshift=-1.8cm,xshift=-4cm]
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=hydras, node distance=3.5cm] (pockcc) { Pocklington };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=hydras, node distance=2.5cm] (chains) { Addition-chains };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=chains, node distance=2.2cm] (gaia) { Gaia };
+\node[rectangle, draw=black,fill=gray!80,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
+\node[rectangle, draw=black,fill=gray!30,minimum height=0.5cm,minimum width=1.5cm,left of=hydras, node distance=3.5cm] (pockcc) { Pocklington };
+\node[rectangle, draw=black,fill=gray!80,minimum height=0.5cm,minimum width=1.5cm,right of=hydras, node distance=2.5cm] (chains) { Addition-chains };
+\node[rectangle, draw=black,fill=gray!30,minimum height=0.5cm,minimum width=1.5cm,right of=chains, node distance=2.2cm] (gaia) { Gaia };
 \end{scope}
@@
 \begin{scope}[yshift=-2.8cm,xshift=-6cm]
- \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
- \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=3cm] (gaiahydras) { Gaia-hydras };
+ \node[rectangle, draw=black,fill=gray!30,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
+ \node[rectangle, draw=black,fill=gray!80,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=3cm] (gaiahydras) { Gaia-hydras };
 \end{scope}
@@
 \draw[->,dotted] (cantor.south) -- (gaia.north west) ;
+\draw[->,dotted] (pock) -- (pockcc) ;
+
 \draw[->] (hydras) -- (gaiahydras) ;
@@
 \draw[->] (pockcc) -- (goedelcc) ;
-\draw[->] (pock) -- (pockcc) ;
 \end{tikzpicture}
 }
-\caption{Genealogy and dependencies for \Hydras packages.}
+\caption{Genealogy and dependencies for packages related to \Hydras.}
   \label{fig:round}

### 2021-10-13 doc/JFLA2022/paper.tex
caption for dependency graph

@@
 
-\begin{scope}[yshift=-0.7cm,xshift=0.25cm]
+\begin{scope}[yshift=-1cm,xshift=0.45cm]
   \node[rectangle, draw=black,fill=gray!30,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
 \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=2cm] (mathcomp) { MathComp };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=paramcoq, node distance=4.7cm] (equations) { Equations };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=paramcoq, node distance=4.5cm] (equations) { Equations };
 \end{scope}
 
-\begin{scope}[yshift=-1.8cm,xshift=-4cm]
+\begin{scope}[yshift=-2cm,xshift=-4cm]
 \node[rectangle, draw=black,fill=gray!80,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
@@
 
-\begin{scope}[yshift=-2.8cm,xshift=-6cm]
+\begin{scope}[yshift=-3cm,xshift=-6cm]
  \node[rectangle, draw=black,fill=gray!30,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
@@
 \draw[->,dotted] (goedel) -- (hydras.north west) ;
-\draw[->,dotted] (cantor) -- (hydras.north) ;
+\draw[->,dotted] (cantor) -- (hydras.north east) ;
 \draw[->,dotted] (cantor.south) -- (gaia.north west) ;
@@
 }
-\caption{Genealogy and dependencies for packages related to \Hydras.}
+\caption{Genealogy and dependencies for \Hydras packages.
+Dotted boxes represent historical Coq contributions, while straight-line boxes represent maintained Coq packages.
+Dark grey packages are maintained in the \Hydras GitHub repository, while light grey packages are maintained in other \community repositories.}
   \label{fig:round}

### 2021-10-13 doc/JFLA2022/paper.tex
add an explanation in Figure 8
outline for table with stats on packages

@@
 Dark grey packages are maintained in the \Hydras GitHub repository, while light grey packages are maintained in other \community repositories.}
-  \label{fig:round}
+  \label{fig:genealogy}
 \end{figure}
@@
 With \Hydras, we hope to demonstrate that projects with complex dependencies are feasible to manage using recent advances in build management and infrastructure.
+
+\begin{table}
+\centering
+\footnotesize
+\begin{tabular}{|l|l|l|l|}
+\hline
+\textbf{Package} & \textbf{Repository} & \textbf{Spec LOC} & \textbf{Proof LOC}\\
+\hline
+Pocklington & \texttt{pocklington} & & \\
+\hline
+Goedel & \texttt{goedel} & & \\
+\hline
+Gaia & \texttt{gaia} & & \\
+\hline
+Hydra-battles & \texttt{hydra-battles} & & \\
+\hline
+Addition-chains & \texttt{hydra-battles} & & \\
+\hline
+Gaia-hydras & \texttt{hydra-battles} & & \\
+\hline
+\end{tabular}
+\caption{Current numbers of lines of code for \Hydras packages.}
+\label{tbl:loc}
+\end{table}
 
@@
 \fbox{\begin{minipage}[h]{1.0\linewidth}
-  Let us prove again the equality $\omega+42+\omega^2= \omega^2$. Let us recall that $\omega^2$ is an abbreviation of $\phi_0(2)$,
-\emph{i.e} the third  additive principal ordinal.
+  Let us prove again the equality $\omega+42+\omega^2= \omega^2$. We recall that $\omega^2$ is an abbreviation of $\phi_0(2)$,
+\emph{i.e} the third  additive principal ordinal, and that \texttt{F} is a notation for the coercion which maps natural numbers to ordinals.
 

### 2021-10-13 doc/JFLA2022/paper.tex
more on Gaia

@@
 In order to prove formally that the length of the
-kind of hydra battles we consider is not given by a primitive recursive function, we chose to use a formalization of the primitive recursive functions that was originally part of Russell O'Connor's statement and proof of G\"{o}del's first incompleteness theorem~\cite{OConnor05, Goedel}.
+kind of hydra battles we consider is not given by a primitive recursive function, we chose to use a formalization of the primitive recursive functions that was originally part of Russell O'Connor's formalization of G\"{o}del's first incompleteness theorem~\cite{OConnor05, Goedel}.
 For this purpose, and above all in consideration of the scientific interest of this contribution, we decided to host and maintain O'Connor's work in \community.
@@
 At around 165,000 lines of code, Gaia is currently one of the largest maintained open source Coq projects~\cite{Gaia}.
+
+The data types for ordinals in Cantor and Veblen form are essentially defined the same way in Gaia and \Hydras, but they are not identical inside Coq, e.g., due to placement in different modules.
+Moreover, there are small differences in the implementation of ordinal arithmetic due to different evolutionary paths taken from the ancestral Cantor library.
+
+Nevertheless, we were able to establish a correspondence between both implementations of ordinals through rewriting lemmas.
+For instance, we proved that multiplication of ordinals in Cantor normal form in \Hydras refine Gaia's corresponding multiplicate, and could import Gaia's proof of associativity of this multiplication almost for free.
 

### 2021-10-13 doc/JFLA2022/paper.tex
intro fix

@@
 For example, the core projects in the Mathematical Components family for Coq amount to more than 160,000 lines of code and 11,000 lemmas~\cite{Nie2020a}, while Lean's Mathlib consists of more than 140,000 lines of code and 34,000 declarations~\cite{Mathlib}.
-However, after a key mathematical definition or result is added to a library, it needs be \emph{documented}~\cite{alectryonpaper} and \emph{maintained}~\cite{vanDoorn2020}.
+However, after a key mathematical definition or result is added to a library, it must be \emph{documented and maintained}~\cite{vanDoorn2020,alectryonpaper}.
 

### 2021-10-13 doc/JFLA2022/paper.tex
small reformulation in section 2

@@
 
-Our proof was based on the construction of a battle, where the
+Our proof was based on the construction of a battle between Herculus and the hydra where the
 number of copies at each round was given by the elimination of an existential quantifier.

### 2021-10-13 doc/JFLA2022/paper.tex
misc additions and fixes

@@
 Libraries of formalized mathematics based on proof assistants, such as \coq, Lean, and Isabelle/HOL, are continually growing in size and scope.
-For example, the core projects in the Mathematical Components family for Coq amount to more than 160,000 lines of code and 11,000 lemmas~\cite{Nie2020a}, while Lean's Mathlib consists of more than 140,000 lines of code and 34,000 declarations~\cite{Mathlib}.
+For example, the core projects in the Mathematical Components family for Coq amount to more than 160,000 lines of code (LOC) and 11,000 lemmas~\cite{Nie2020a}, while Lean's Mathlib consists of more than 140,000 LOC and 34,000 declarations~\cite{Mathlib}.
 However, after a key mathematical definition or result is added to a library, it must be \emph{documented and maintained}~\cite{vanDoorn2020,alectryonpaper}.
@@
 The data types for ordinals in Cantor and Veblen form are essentially defined the same way in Gaia and \Hydras, but they are not identical inside Coq, e.g., due to placement in different modules.
-Moreover, there are small differences in the implementation of ordinal arithmetic due to different evolutionary paths taken from the ancestral Cantor library.
+Moreover, there are small differences in the implementation of ordinal arithmetic due to different evolutionary paths taken since divergence from the ancestral Cantor library.
 
@@
 \subsection{Project Genealogy, Dependencies, and Organization}
+
+Figure~\ref{fig:genealogy} illustrates the relationships between \Hydras packages.
 
@@
 
+Table~\ref{tbl:loc} lists the size of each \Hydras package.
+
 \begin{table}
@@
 \footnotesize
-\begin{tabular}{|l|l|l|l|}
+\begin{tabular}{|l|l|l|l|l|}
 \hline
-\textbf{Package} & \textbf{Repository} & \textbf{Spec LOC} & \textbf{Proof LOC}\\
+\textbf{Package} & \textbf{Repository} & \textbf{Commit} & \textbf{Spec LOC} & \textbf{Proof LOC}\\
 \hline
-Pocklington & \texttt{pocklington} & & \\
+Pocklington & \texttt{pocklington} & & & \\
 \hline
-Goedel & \texttt{goedel} & & \\
+Goedel & \texttt{goedel} & & & \\
 \hline
-Gaia & \texttt{gaia} & & \\
+Gaia & \texttt{gaia} & & & \\
 \hline
-Hydra-battles & \texttt{hydra-battles} & & \\
+Hydra-battles & \texttt{hydra-battles} & & & \\
 \hline
-Addition-chains & \texttt{hydra-battles} & & \\
+Addition-chains & \texttt{hydra-battles} & & & \\
 \hline
-Gaia-hydras & \texttt{hydra-battles} & & \\
+Gaia-hydras & \texttt{hydra-battles} & & & \\
 \hline
@@
 
-\Hydras wants to be a bridge between scientific literature (\emph{e.g.}~\cite{KP82, KS81, schutte} and proof assistant technology.
+\Hydras wants to be a bridge between scientific literature (\emph{e.g.},~\cite{KP82, KS81, schutte}) and proof assistant technology.
 For the mathematician, it may give a concrete view of the mathematical content, not only through full proofs, but also through significative computations: examples, functions associated with constructive proofs, etc.

### 2021-10-13 doc/JFLA2022/paper.tex
fix typo reported by Pierre

@@
 
-Our proof was based on the construction of a battle between Herculus and the hydra where the
+Our proof was based on the construction of a battle between Hercules and the hydra where the
 number of copies at each round was given by the elimination of an existential quantifier.

### 2021-10-13 doc/JFLA2022/paper.tex
work on book table
more details in dep graph caption

@@
 Dotted boxes represent historical Coq contributions, while straight-line boxes represent maintained Coq packages.
-Dark grey packages are maintained in the \Hydras GitHub repository, while light grey packages are maintained in other \community repositories.}
+Dark grey packages are maintained in the \Hydras GitHub repository, while light grey packages are maintained in other \community repositories.
+Dotted lines represent Coq code ancestry, while regular lines represent direct code dependencies.}
   \label{fig:genealogy}
@@
 
+\begin{table}
+  \centering
+  \footnotesize
+  \begin{tabular}{lll}
+  \hline
+  \textbf{Title} & \textbf{Year} & \textbf{Category}\\
+  \hline
+  Coq'Art~\cite{BC04} & 2004 & Traditional\\
+  Formal Reasoning About Programs~\cite{Chlipala2017} & 2017 & Traditional\\
+  Software Foundations~\cite{Pierce2007} & 2007 & Executable\\
+  Program Logics for Certified Compilers~\cite{Appel2014} & 2014 & Traditional\\
+  \end{tabular}
+  \caption{Coq book comparison.}
+  \label{tbl:loc}
+\end{table}
+
 \section{Conclusion and perspectives}

### 2021-10-13 doc/JFLA2022/paper.tex
more on dep graph
fill in fairly complete book list

@@
 
-Figure~\ref{fig:genealogy} illustrates the relationships between \Hydras packages.
+Figure~\ref{fig:genealogy} illustrates the relationships between \Hydras packages and show their ancestry from historical Coq contributions.
 
@@
   Coq'Art~\cite{BC04} & 2004 & Traditional\\
+  Software Foundations~\cite{Pierce2007} & 2007 & Executable\\
+  Certified Programming with Dependent Types~\cite{chlipalacpdt2011} & 2011 &  Executable\\
+  Program Logics for Certified Compilers~\cite{Appel2014} & 2014 & Traditional\\
+  Programs and Proofs~\cite{Sergey:PnP} & 2014 & Executable\\
   Formal Reasoning About Programs~\cite{Chlipala2017} & 2017 & Traditional\\
-  Software Foundations~\cite{Pierce2007} & 2007 & Executable\\
-  Program Logics for Certified Compilers~\cite{Appel2014} & 2014 & Traditional\\
+  Computer Arithmetic and Formal Proofs~\cite{Boldo2017} & 2017 & Traditional\\
+  Mathematical Components~\cite{MCB} & 2018 & Traditional\\
+  \hline
   \end{tabular}

### 2021-10-14 doc/JFLA2022/paper.tex
fix two syntax errors

@@
 In order to prove formally that the length of the
-kind of hydra battles we consider is not given by a primitive recursive function, we chose to use a formalization of the primitive recursive functions that was originally part of Russell O'Connor's formalization of G\"{o}del's first incompleteness theorem~\cite{OConnor05, Goedel}.
+kind of hydra battles we consider is not given by any primitive recursive function, we chose to use a formalization of primitive recursive functions that was originally part of Russell O'Connor's formalization of G\"{o}del's first incompleteness theorem~\cite{OConnor05, Goedel}.
 For this purpose, and above all in consideration of the scientific interest of this contribution, we decided to host and maintain O'Connor's work in \community.

### 2021-10-14 doc/JFLA2022/paper.tex
Fix the multiply defined label warning (tbl:loc -> tbl:books)

@@
 In order to prove formally that the length of the
-kind of hydra battles we consider is not given by any primitive recursive function, we chose to use a formalization of primitive recursive functions that was originally part of Russell O'Connor's formalization of G\"{o}del's first incompleteness theorem~\cite{OConnor05, Goedel}.
+kind of hydra battles we consider is not given by any primitive recursive function, we chose to use a formalization of primitive recursive functions that was originally part of Russell O'Connor's contribution on G\"{o}del's first incompleteness theorem~\cite{OConnor05, Goedel}.
 For this purpose, and above all in consideration of the scientific interest of this contribution, we decided to host and maintain O'Connor's work in \community.
@@
   \caption{Coq book comparison.}
-  \label{tbl:loc}
+  \label{tbl:books}
 \end{table}

### 2021-10-14 doc/JFLA2022/paper.tex
avoid two "formalization" in the same sentence

@@
 In order to prove formally that the length of the
-kind of hydra battles we consider is not given by any primitive recursive function, we chose to use a formalization of primitive recursive functions that was originally part of Russell O'Connor's contribution on G\"{o}del's first incompleteness theorem~\cite{OConnor05, Goedel}.
+kind of hydra battles we consider is not given by any primitive recursive function, we chose to use a formalization of primitive recursive functions that was originally part of Russell O'Connor's formal proof of G\"{o}del's first incompleteness theorem~\cite{OConnor05, Goedel}.
 For this purpose, and above all in consideration of the scientific interest of this contribution, we decided to host and maintain O'Connor's work in \community.

### 2021-10-14 doc/JFLA2022/paper.tex
refine caption in dependency graph

--- 2021-10-14 doc/JFLA2022/paper.tex
@@
 In order to prove formally that the length of the
-kind of hydra battles we consider is not given by any primitive recursive function, we chose to use a formalization of primitive recursive functions that was originally part of Russell O'Connor's formalization of G\"{o}del's first incompleteness theorem~\cite{OConnor05, Goedel}.
+kind of hydra battles we consider is not given by a primitive recursive function, we chose to use a formalization of the primitive recursive functions that was originally part of Russell O'Connor's formalization of G\"{o}del's first incompleteness theorem~\cite{OConnor05, Goedel}.
 For this purpose, and above all in consideration of the scientific interest of this contribution, we decided to host and maintain O'Connor's work in \community.
@@
 \caption{Genealogy and dependencies for \Hydras packages.
-Dotted boxes represent historical Coq contributions, while straight-line boxes represent maintained Coq packages.
+Dotted boxes represent historical Coq contributions, while regular boxes represent maintained Coq packages.
 Dark grey packages are maintained in the \Hydras GitHub repository, while light grey packages are maintained in other \community repositories.

### 2021-10-14 doc/JFLA2022/paper.tex
more on book comparison

--- 2021-10-14 doc/JFLA2022/paper.tex
@@
 
-We believe that current Coq books can broadly be divided into two categories:
+We believe that Coq books can broadly be divided into two categories:
 \begin{enumerate}
@@
 \end{enumerate}
+
+Table~\ref{tbl:books} shows a categorization of the Coq books in English we are aware of.
 
@@
 \end{table}
+
+Several traditional books, such as Coq'Art~\cite{BC04}, used custom tools during their writing process to keep code snippets synchronized with Coq.
+However, once a book is published, the accompanying code generally takes on a life of its own~\cite{CoqArtCommunity} and may come to substantially diverge from the book.
+
+We believe the \Hydras electronic book finds good tradeoffs between the advantages of executable books and traditional books.
 

### 2021-10-14 doc/JFLA2022/paper.tex
another pass over comparison

@@
 
-The Gaia project by Jos\'e Grimm aimed to formalize mathematics in the style of N.
-Bourbaki in Coq.
-The formalization of the first book by Bourbaki, on the theory of sets, was initially described in a technical report in July 2009~\cite{Grimm2009a}.
+The Gaia project by Jos\'e Grimm aimed to formalize mathematics in Coq in the style of Nicolas Bourbaki.
+The formalization of the first book in the Elements of Mathematics series by Bourbaki, on the theory of sets, was initially described in a technical report in July 2009~\cite{Grimm2009a}.
 Grimm then wrote (and continually updated) technical reports describing the formalization of Bourbaki's two subsequent books~\cite{Grimm2009b,Grimm2016} and additional topics in number theory~\cite{grimm:hal-00911710,Grimm2014}, before he passed away in 2019.
@@
 We believe that Coq books can broadly be divided into two categories:
-\begin{enumerate}
-\item[a)] \emph{Executable textbooks} that are in effect large, well-commented Coq programs from which other representations of the material (HTML, PDF) are derived using tools.
-\item[b)] \emph{Traditional textbooks} generated from documentation languages such as LaTeX that are accompanied by standalone Coq formalizations and/or code snippets.
-\end{enumerate}
-
-Table~\ref{tbl:books} shows a categorization of the Coq books in English we are aware of.
+\begin{itemize}
+\item \emph{Executable textbooks} that are in effect large, well-commented Coq programs from which other representations of the material (HTML, PDF) are derived using tools.
+\item \emph{Traditional textbooks} generated from documentation languages such as LaTeX that are accompanied by standalone Coq formalizations and/or code snippets.
+\end{itemize}
+Table~\ref{tbl:books} shows a category breakdown for what we believe are the most prominent Coq books in English.
 
@@
 
-We believe the \Hydras electronic book finds good tradeoffs between the advantages of executable books and traditional books.
+We believe the \Hydras electronic book finds good tradeoffs between the properties of executable books and traditional books.
+Specifically, in an executable book, the proof assistant language places limits on the structure.
+For example, Coq may not permit repeating a definition from another module verbatim, since the definition uses section variables that are not present in the current context.
+Thanks to our \alectr based toolchain, the \Hydras book can include ``live'' code fragments at any point in the text.
+And in contrast to traditional textbooks, we validate the consistency of code fragments during every build of the electronic book.
+We follow executable books in providing ``continuous delivery'' of new revisions at a high pace, but we also aim to make regular timestamped versions compatible with specific versions of Coq.
 

### 2021-10-14 doc/JFLA2022/paper.tex
one more comparison pass

@@
 Table~\ref{tbl:books} shows a category breakdown for what we believe are the most prominent Coq books in English.
+Note that several traditional books, such as Coq'Art~\cite{BC04}, used custom tools during their writing process to keep code snippets synchronized with Coq.
+However, once a book is published, the accompanying code generally takes on a life of its own~\cite{CoqArtCommunity} and may come to substantially diverge from the book.
 
@@
 
-Several traditional books, such as Coq'Art~\cite{BC04}, used custom tools during their writing process to keep code snippets synchronized with Coq.
-However, once a book is published, the accompanying code generally takes on a life of its own~\cite{CoqArtCommunity} and may come to substantially diverge from the book.
-
 We believe the \Hydras electronic book finds good tradeoffs between the properties of executable books and traditional books.
@@
 And in contrast to traditional textbooks, we validate the consistency of code fragments during every build of the electronic book.
-We follow executable books in providing ``continuous delivery'' of new revisions at a high pace, but we also aim to make regular timestamped versions compatible with specific versions of Coq.
+On one hand, we follow executable books in providing ``continuous delivery'' of new revisions at a high pace.
+But on the other hand, we also aim to make regular timestamped versions compatible with specific versions of Coq, akin to new editions of traditional books.
 

### 2021-10-14 doc/JFLA2022/paper.tex
add Simpson's set theory to ancestry graph

@@
 \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (cantor) { Cantor };
+
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3.2cm] (cats) { Categories in ZFC };
 \end{scope}
@@
   \node[rectangle, draw=black,fill=gray!30,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=2cm] (mathcomp) { MathComp };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=2.5cm] (mathcomp) { MathComp };
 \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=paramcoq, node distance=4.5cm] (equations) { Equations };
@@
 \draw[->,dotted] (pock) -- (pockcc) ;
+\draw[->,dotted] ([xshift=2.1cm]cats) -- (gaia) ;
 
@@
 Dotted boxes represent historical Coq contributions, while regular boxes represent maintained Coq packages.
-Dark grey packages are maintained in the \Hydras GitHub repository, while light grey packages are maintained in other \community repositories.
+Dark gray packages are maintained in the \Hydras GitHub repository, while light gray packages are maintained in other \community repositories.
 Dotted lines represent Coq code ancestry, while regular lines represent direct code dependencies.}

### 2021-10-14 doc/JFLA2022/paper.tex
pass over conclusion to place Gaia material in correct place

@@
 The formalization of the first book in the Elements of Mathematics series by Bourbaki, on the theory of sets, was initially described in a technical report in July 2009~\cite{Grimm2009a}.
+The set-theoretic axioms and basic definitions in \gaia were derived from an earlier development by Carlos Simpson~\cite{Simpson2004,CatsZFCContrib}.
 Grimm then wrote (and continually updated) technical reports describing the formalization of Bourbaki's two subsequent books~\cite{Grimm2009b,Grimm2016} and additional topics in number theory~\cite{grimm:hal-00911710,Grimm2014}, before he passed away in 2019.
@@
 
-We plan to extend our libraries in two main directions.
-
-The \gaia project~\cite{Gaia}, also maintained on \community, contains a development by José Grimm~\cite{grimm:hal-00911710},  dedicated  to the implementation in \coq (SSReflect) of books from  Bourbaki's Elements of Mathematics.
-The set theory underpinning \gaia is derived from an earlier development by Carlos Simpson~\cite{Simpson2004,CatsZFCContrib}.
-\gaia contains data structures which are compatible with our ordinal notations.
-Our plan is to build a  bridge between  the combinatorial results of \Hydras and the set-theoretic content of \gaia, and make possible the transfer of theorems between both libraries.
-We aim also to write in \coq a formal proof of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
+We plan to extend \Hydras in two main directions.
+Firstly, we plan to build a bridge between the combinatorial results of \Hydras and the set-theoretic content of \gaia, and make possible the transfer of many interesting theorems between the two packages.
+Secondly, we aim to write a formal proof in \coq of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
 
 \Hydras is not limited to the study of ordinal numbers and applications.
- For instance, we are also developing 
-a module about efficient exponentiation algorithms, and hope
-to extend our project to new topics.
+For instance, we are also developing  a module about efficient exponentiation
+algorithms, and hope to extend our project to new topics.
 

### 2021-10-14 doc/JFLA2022/paper.tex
pass over integration section

@@
 
-Since computability is a key topic in computer science teaching and O'Connor's library is a nice illustration of dependently typed programming, we decided to devote a full chapter (chapter 9 of~\cite{HydraBook}) to the formalization of primitive recursive functions, with comments on the definitions and proofs and [counter-]examples and exercises.
+Since computability is a key topic in computer science teaching and O'Connor's library is a nice illustration of dependently typed programming, we decided to devote a full chapter (chapter 9 of~\cite{HydraBook}) to the formalization of primitive recursive functions, with comments on the definitions and proofs and (counter-)examples and exercises.
 As part of the writing process, we made the formalization into a new sub-library of the Hydra-battles library, dubbed Ackermann.
 
-\subsection{Bridge to ordinals in Gaia}
+\subsection{Bridge to ordinals in \gaia}
 
-The Gaia project by Jos\'e Grimm aimed to formalize mathematics in Coq in the style of Nicolas Bourbaki.
+The \gaia project by Jos\'e Grimm aimed to formalize mathematics in \coq in the style of Nicolas Bourbaki.
 The formalization of the first book in the Elements of Mathematics series by Bourbaki, on the theory of sets, was initially described in a technical report in July 2009~\cite{Grimm2009a}.
@@
 
-In 2020, members of \community transferred the Gaia source code to GitHub and adapted it for recent releases of the Mathematical Components library, which Gaia heavily relies on.
+In 2020, members of \community transferred the \gaia source code to GitHub and adapted it for recent releases of the Mathematical Components library, which \gaia heavily relies on.
 Anonymous volunteers (``collaborators of Nicolas Bourbaki'') then finished the only in-progress proof left by Grimm.
-At around 165,000 lines of code, Gaia is currently one of the largest maintained open source Coq projects~\cite{Gaia}.
+At around 165,000 lines of code, \gaia is currently one of the largest maintained open source \coq projects~\cite{Gaia}.
 
-The data types for ordinals in Cantor and Veblen form are essentially defined the same way in Gaia and \Hydras, but they are not identical inside Coq, e.g., due to placement in different modules.
+The data types for ordinals in Cantor and Veblen form are essentially defined the same way in Gaia and \Hydras, but they are not identical inside \coq, e.g., due to placement in different modules.
 Moreover, there are small differences in the implementation of ordinal arithmetic due to different evolutionary paths taken since divergence from the ancestral Cantor library.

### 2021-10-14 doc/JFLA2022/paper.tex
pass over Gaia bridge text

@@
 
-The data types for ordinals in Cantor and Veblen form are essentially defined the same way in Gaia and \Hydras, but they are not identical inside \coq, e.g., due to placement in different modules.
-Moreover, there are small differences in the implementation of ordinal arithmetic due to different evolutionary paths taken since divergence from the ancestral Cantor library.
-
-Nevertheless, we were able to establish a correspondence between both implementations of ordinals through rewriting lemmas.
-For instance, we proved that multiplication of ordinals in Cantor normal form in \Hydras refine Gaia's corresponding multiplicate, and could import Gaia's proof of associativity of this multiplication almost for free.
+\gaia contains definitions of ordinals in Cantor and Veblen normal form~\cite{grimm:hal-00911710}, adapted from the historical Cantor contribution~\cite{CantorContrib}.
+The data types for ordinals are essentially defined the same way in Gaia and \Hydras, but they are not identical inside \coq, e.g., due to placement in different modules.
+Moreover, there are small differences in the implementation of ordinal arithmetic due to different evolutionary paths taken since divergence from the ancestral library.
+%
+Nevertheless, we were able to establish an initial correspondence between both implementations of ordinals through rewriting lemmas.
+For instance, we proved that multiplication of ordinals in Cantor normal form in \Hydras refine Gaia's corresponding multiplication operation, and were able to import Gaia's proof of associativity of the multiplication almost for free.
 

### 2021-10-14 doc/JFLA2022/paper.tex
cite Lean 4 book in intro

@@
 Maintenance includes not only adaptation to changes in new proof assistant versions, but also reorganization to ease future maintenance effort and accommodate new contributions.
-Documentation is usually two-fold: \emph{source code comments} that describe specific definitions and results in-line and \emph{books} that carefully introduce the library and its idioms~\cite{MCB}.
+Documentation is usually two-fold: \emph{source code comments} that describe specific definitions and results in-line and \emph{books} that carefully introduce the library and its idioms~\cite{MCB,TheoremProvingLean4}.
 The former tends to be terse and dense and serves experienced users, while the latter is more long-winded and exhaustive and serves beginners.

### 2021-10-14 doc/JFLA2022/paper.tex
starting monorepo pass, fix bridge subsection title

@@
 
-\subsection{Bridge to ordinals in \gaia}
+\subsection{First steps towards a bridge to \gaia}
 
@@
 
-We solved this maintenance issue by moving to \alectr, a tool that automatically records \coq proofs.
+We solved this maintenance issue by moving to \alectr, a tool that automatically records \coq proofs~\cite{alectryongithub}.
 Instead of copy-pasting fragments into LaTeX, we now embed small LaTeX files automatically generated by \alectr from our \coq development.
@@
 
-Monorepos (shorthand for monolithic repositories) are single repositories containing multiple independent or related packages.
+A monorepo (shorthand for monolithic repository) is a version control repository containing multiple independent or related packages.
 %
-They have gained increasing popularity following experiences in large companies such as Google, but are also used at a lesser scale for managing open source projects.
+Monorepos have gained increasing popularity following experiences in large companies such as Google, but are also used at a smaller scale for managing open source projects.
 %
-They are known to simplify the management of dependencies, making cross-packages changes and refactorings~\cite{brito2018monorepos}.
+They are known to simplify the management of dependencies, making cross-packages changes, and refactorings~\cite{brito2018monorepos}.
 
-In the context of Coq projects, they did not use to be well-supported.
 %

### 2021-10-14 doc/JFLA2022/paper.tex
pass over monorepo

@@
 %
-One of the few early monorepos in the Coq ecosystem was the Mathematical Components (math-comp) repository, which used a custom build infrastructure to check multiple packages on each commit.
+One of the few early uses of monorepos in the Coq ecosystem was for the Mathematical Components library~\cite{MCB}, whose maintainers developed a custom build infrastructure to check multiple packages on each commit.
 %
-However, recent improvements in tooling have made their use easier, and therefore we have used them more and more, for instance within \community.
+In light of that recent tooling improvements have made monorepo use easier, we chose to use an explicit monorepo structure for \Hydras in \community.
+Other \community projects have since adopted a similar structure.
 
@@
 
-We believe the \Hydras electronic book finds good tradeoffs between the properties of executable books and traditional books.
+We believe the \Hydras electronic book has found an attractive set of tradeoffs between the properties of executable books and traditional books.
 Specifically, in an executable book, the proof assistant language places limits on the structure.
@@
 And in contrast to traditional textbooks, we validate the consistency of code fragments during every build of the electronic book.
-On one hand, we follow executable books in providing ``continuous delivery'' of new revisions at a high pace.
+On the one hand, we follow executable books in providing ``continuous delivery'' of new revisions at a high pace.
 But on the other hand, we also aim to make regular timestamped versions compatible with specific versions of Coq, akin to new editions of traditional books.

### 2021-10-14 doc/JFLA2022/paper.tex
reserve 'bridge' for gaia-hydras
more book comparison stuff

@@
 %
-This means, for instance, that we can consolidate libraries by merging them, or split them up into several packages.
+This means, for instance, that we can consolidate packages by merging them, or split up a single package into several packages.
 
@@
   Coq'Art~\cite{BC04} & 2004 & Traditional\\
-  Software Foundations~\cite{Pierce2007} & 2007 & Executable\\
+  Software Foundations~\cite{SF} & 2007 & Executable\\
   Certified Programming with Dependent Types~\cite{chlipalacpdt2011} & 2011 &  Executable\\
@@
 
+Due to our reliance on LaTeX for writing the book, obtaining a book representation in HTML is more cumbersome than for executable books such as Software Foundations~\cite{SF} and Certified Programming with Dependent Types~\cite{chlipalacpdt2011}.
+Currently, we generate the Coq code documentation as HTML separately from the book.
+
 \section{Conclusion and perspectives}
 
-\Hydras wants to be a bridge between scientific literature (\emph{e.g.},~\cite{KP82, KS81, schutte}) and proof assistant technology.
+\Hydras wants to be provide a connection between scientific literature (\emph{e.g.},~\cite{KP82, KS81, schutte}) and proof assistant technology.
 For the mathematician, it may give a concrete view of the mathematical content, not only through full proofs, but also through significative computations: examples, functions associated with constructive proofs, etc.

### 2021-10-14 doc/JFLA2022/paper.tex
simplification of Alectryon use case description from Pierre

@@
 
-Importing snippets into a LaTeX document was not one of the original use cases of \alectr, so we had to extend it: the original \alectr did not have support for exporting to LaTeX, and it was geared towards documenting individual source files (a style where code and prose are interleaved within the same file and code is documented in the same order as it is compiled).
+Importing snippets into a LaTeX document was not one of the original use cases of Alectryon.
+Firstly, the original Alectryon did not have support for exporting to LaTeX.
+Secondly, it was geared towards documenting individual source files, where code and prose are interleaved within the same file.
+In this style, the code is documented in the same order as it is compiled.
 We extended \alectr to support our needs by implementing a LaTeX backend, and by programming it to generate individual snippet files, one per \texttt{snippet} comment block.

### 2021-10-15 doc/JFLA2022/paper.tex
new title
pass over genealogy section

@@
 %
-\title{Hydras,  Ordinals \& Co.   \\
-  A library in Coq of entertaining formal mathematics}
+% \title{Hydras,  Ordinals \& Co.   \\
+%   A library in Coq of entertaining formal mathematics}
+
+\title{Hydras \& Co.: Formalized mathematics in Coq\\
+ for inspiration and entertainment
+}
 
@@
 We chose to build our library and book on two simple themes which allow many variations: computing powers in a monoid, and Kirby and Paris' hydra battles.
-In the interest of space, we shall present only the second one in this article.
+In the interest of space, we will only present the second theme in this paper.
 
@@
 \gaia contains definitions of ordinals in Cantor and Veblen normal form~\cite{grimm:hal-00911710}, adapted from the historical Cantor contribution~\cite{CantorContrib}.
-The data types for ordinals are essentially defined the same way in Gaia and \Hydras, but they are not identical inside \coq, e.g., due to placement in different modules.
+The data types for ordinals are essentially defined the same way in Gaia and \Hydras, but they are not identical inside \coq, e.g., due to residing in different modules.
 Moreover, there are small differences in the implementation of ordinal arithmetic due to different evolutionary paths taken since divergence from the ancestral library.
@@
 
-\subsection{Project Genealogy, Dependencies, and Organization}
+\subsection{Package Genealogy, Dependencies, and Organization}
 
+Both due to its prior stages~\cite{PCiota,JFLA2018paper} and the recent integrations described just above, the current \Hydras Coq code has a complex inheritance.
 Figure~\ref{fig:genealogy} illustrates the relationships between \Hydras packages and show their ancestry from historical Coq contributions.
+To indicate the scope and relative sizes of packages, Table~\ref{tbl:loc} breaks down lines of code for each \Hydras package, as reported by the \texttt{coqwc} tool.
+We also name and link to the repository in \community for each package.
 
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
@@
 
-Table~\ref{tbl:loc} lists the size of each \Hydras package.
-
-\begin{table}
+\begin{table}[ht]
 \centering
 \footnotesize
-\begin{tabular}{|l|l|l|l|l|}
+\begin{tabular}{|l|l|l|l|l|l|}
 \hline
-\textbf{Package} & \textbf{Repository} & \textbf{Commit} & \textbf{Spec LOC} & \textbf{Proof LOC}\\
+\textbf{Package name} & \textbf{Repository} & \textbf{Package on opam} & \textbf{Version} & \textbf{Spec LOC} & \textbf{Proof LOC}\\
 \hline
-Pocklington & \texttt{pocklington} & & & \\
+Pocklington & \href{https://github.com/coq-community/pocklington}{\texttt{pocklington}} & \texttt{coq-pocklington} & \texttt{8.12.0} & & \\
 \hline
-Goedel & \texttt{goedel} & & & \\
+Goedel & \href{https://github.com/coq-community/goedel}{\texttt{goedel}} & \texttt{coq-goedel} & \texttt{8.13.0} & & \\
 \hline
-Gaia & \texttt{gaia} & & & \\
+Gaia & \href{https://github.com/coq-community/gaia}{\texttt{gaia}} & \texttt{coq-gaia} & \texttt{1.12} & & \\
 \hline
-Hydra-battles & \texttt{hydra-battles} & & & \\
+Hydra-battles & \href{https://github.com/coq-community/hydra-battles}{\texttt{hydra-battles}} & \texttt{coq-hydra-battles} & \texttt{0.5} & & \\
 \hline
-Addition-chains & \texttt{hydra-battles} & & & \\
+Addition-chains & \href{https://github.com/coq-community/hydra-battles}{\texttt{hydra-battles}} & \texttt{coq-addition-chains} & \texttt{0.5} & & \\
 \hline
-Gaia-hydras & \texttt{hydra-battles} & & & \\
+Gaia-hydras & \href{https://github.com/coq-community/hydra-battles}{\texttt{hydra-battles}} & \texttt{coq-gaia-hydras} & \texttt{0.5} & & \\
 \hline

### 2021-10-15 doc/JFLA2022/paper.tex
pass over intro

@@
 
-Maintenance includes not only adaptation to changes in new proof assistant versions, but also reorganization to ease future maintenance effort and accommodate new contributions.
+Maintenance includes not only adaptation to changes in new proof assistant versions, but also reorganization to accommodate new contributions.
 Documentation is usually two-fold: \emph{source code comments} that describe specific definitions and results in-line and \emph{books} that carefully introduce the library and its idioms~\cite{MCB,TheoremProvingLean4}.
@@
 As a library changes and expands, all its documentation needs to be made consistent and complete.
-Authors use a wide range of techniques, including literate programming~\cite{Knuth1984} and custom tools, to pretty-print and check source code snippets and generate proof assistant output.
+Authors use many techniques, including literate programming~\cite{Knuth1984} and custom tools, to pretty-print and check source code snippets and generate proof assistant output.
 
 While books that document proof assistant libraries are valuable to beginners, the purpose of recently published books is mainly instrumental, i.e., to teach a certain topic or technique.
-We believe that books can instead become \emph{ends in themselves}---not just sources of learning, but also sources of inspiration and entertainment.
+We believe that books and libraries can instead become \emph{ends in themselves}---not just sources of exposition and learning, but also sources of inspiration and entertainment.
 

### 2021-10-15 doc/JFLA2022/paper.tex
small intro pass

@@
 \item A hydra is a finite tree, traditionally presented with the root at the bottom, the leaves of which are called \emph{heads}
-  (Fig~\ref{fig:round}).
+  (Figure~\ref{fig:round}).
 \item At every round, Hercules chops off one head of the hydra. If the head is at a distance greater than 1 from the root,
@@
 
-Kirby and Paris prove the following theorems, applying
+Kirby and Paris proved the following theorems, applying
 combinatorial results about ordinal numbers by Jussi Ketonen and Robert Solovay~\cite{KS81}.
@@
 
-Complex formalisations and proofs are explained in the book.
+Complex formalizations and proofs are explained in the book.
 % an  electronic book~\cite{HydraBook} (PDF document of over 280 pages).
@@
 It may be also used as a library on ordinal numbers, for instance for proving termination properties.
-
+%
  Prior stages of this project have already been presented at
  JFLA~\cite{PCiota, JFLA2018paper}.
-We present recent evolutions of the library: new results, interaction within the \community project~\cite{CoqCommunity}, and documentation generated with \alectr~\cite{alectryonpaper, alectryongithub}.
+In this paper, we present recent evolutions of the library: new results, interaction with the \community project~\cite{CoqCommunity}, and documentation generated with \alectr~\cite{alectryonpaper, alectryongithub}.
 

### 2021-10-15 doc/JFLA2022/paper.tex
reformulations in section 2 for space

@@
 \begin{theorem}
-  Let $\mu$ be any ordinal strictly less than $\epsilon_0$.
-  There is no function mapping hydras to the segment $[0,\mu)$ that could be used as a measure for  proving the termination of  \emph{any} hydra battle.\label{thm3}
+  Let $\mu$ be an ordinal strictly less than $\epsilon_0$.
+  There is no function mapping hydras to the segment $[0,\mu)$ that could be a measure used in proving the termination of  \emph{any} hydra battle.\label{thm3}
 \end{theorem}
@@
  \begin{itemize}
-   \item The game starts at an initial step $k=i$ where $i$ is any natural number.
-   \item  The hydra is always the  representation as a tree of some ordinal strictly below $\epsilon_0$ in Cantor normal
+   \item The game starts at an initial step $k=i$, where $i$ is any natural number.
+   \item Hercules always chops off the rightmost head of the hydra.
+   \item  The hydra is always the tree representation of some ordinal strictly below $\epsilon_0$ in Cantor normal
      form (thus, the rightmost branch is also one of the shortest).
-     For instance, Fig~\ref{fig:round} shows the hydras respectively associated with  $\omega^{\omega^2+1}$ and $\omega^{\omega^2}\times 3$.
-     In the following,  the expression  ``the hydra $\alpha$'' is an abbreviation of ``the hydra associated with the ordinal $\alpha$''.
-
- \item Hercules always chops off the rightmost head of the hydra.
+     For instance, Figure~\ref{fig:round} shows the hydras respectively associated with  $\omega^{\omega^2+1}$ and $\omega^{\omega^2}\times 3$.
  \end{itemize}
@@
  $\canonseq{\alpha}{k+1}$, \emph{i.e.} the $(k+1)$th element of the canonical sequence of $\alpha$~\cite{KS81}.
+In the following,  the expression ``the hydra $\alpha$'' is an abbreviation of ``the hydra associated with the ordinal $\alpha$''.
 

### 2021-10-15 doc/JFLA2022/paper.tex
small section 2 reformulation

@@
 \begin{theorem}
-  Let $\mu$ be an ordinal strictly less than $\epsilon_0$.
-  There is no function mapping hydras to the segment $[0,\mu)$ that could be a measure used in proving the termination of  \emph{any} hydra battle.\label{thm3}
+  Let $\mu$ be any ordinal strictly less than $\epsilon_0$.
+  There is no function mapping hydras to the segment $[0,\mu)$ that could be used as a measure in proving termination for \emph{any} hydra battle.\label{thm3}
 \end{theorem}

### 2021-10-15 doc/JFLA2022/paper.tex
more on package organization

@@
 Figure~\ref{fig:genealogy} illustrates the relationships between \Hydras packages and show their ancestry from historical Coq contributions.
-To indicate the scope and relative sizes of packages, Table~\ref{tbl:loc} breaks down lines of code for each \Hydras package, as reported by the \texttt{coqwc} tool.
-We also name and link to the repository in \community for each package.
+To indicate the scope and relative sizes of packages, Table~\ref{tbl:loc} breaks down lines of code for each \Hydras package, as reported by the \texttt{coqwc} tool, and lists their repository names in \community and identifiers in the Coq opam package index~\cite{CoqPackageIndex}.
 
@@
 \hline
-\textbf{Package name} & \textbf{Repository} & \textbf{Package on opam} & \textbf{Version} & \textbf{Spec LOC} & \textbf{Proof LOC}\\
+\textbf{Package name} & \textbf{Repository} & \textbf{Package index identifier} & \textbf{Version} & \textbf{Spec LOC} & \textbf{Proof LOC}\\
 \hline

### 2021-10-15 doc/JFLA2022/paper.tex
fill in proof sizes where possible

--- 2021-10-15 doc/JFLA2022/paper.tex
@@
 Anonymous volunteers (``collaborators of Nicolas Bourbaki'') then finished the only in-progress proof left by Grimm.
-At around 165,000 lines of code, \gaia is currently one of the largest maintained open source \coq projects~\cite{Gaia}.
+At around 155,000 lines of code, \gaia is currently one of the largest maintained open source \coq projects~\cite{Gaia}.
 
 \gaia contains definitions of ordinals in Cantor and Veblen normal form~\cite{grimm:hal-00911710}, adapted from the historical Cantor contribution~\cite{CantorContrib}.
-The data types for ordinals are essentially defined the same way in Gaia and \Hydras, but they are not identical inside \coq, e.g., due to residing in different modules.
+The data types for ordinals are essentially defined the same way in \gaia and \Hydras, but they are not identical inside \coq, e.g., due to residing in different modules.
 Moreover, there are small differences in the implementation of ordinal arithmetic due to different evolutionary paths taken since divergence from the ancestral library.
@@
 Nevertheless, we were able to establish an initial correspondence between both implementations of ordinals through rewriting lemmas.
-For instance, we proved that multiplication of ordinals in Cantor normal form in \Hydras refine Gaia's corresponding multiplication operation, and were able to import Gaia's proof of associativity of the multiplication almost for free.
+For instance, we proved that multiplication of ordinals in Cantor normal form in \Hydras refine \gaia's corresponding multiplication operation, and were able to import \gaia's proof of associativity of the multiplication almost for free.
 
@@
 
-Both due to its prior stages~\cite{PCiota,JFLA2018paper} and the recent integrations described just above, the current \Hydras Coq code has a complex inheritance.
+Both due to its prior stages~\cite{PCiota,JFLA2018paper} and the recent integrations described just above, the current \Hydras \coq code has a complex inheritance.
 Figure~\ref{fig:genealogy} illustrates the relationships between \Hydras packages and show their ancestry from historical Coq contributions.
-To indicate the scope and relative sizes of packages, Table~\ref{tbl:loc} breaks down lines of code for each \Hydras package, as reported by the \texttt{coqwc} tool.
-We also name and link to the repository in \community for each package.
+To indicate the scope and relative sizes of packages, Table~\ref{tbl:loc} breaks down lines of code for each \Hydras package, as reported by the \texttt{coqwc} tool, and lists their repository names in \community and identifiers in the \coq opam package index~\cite{CoqPackageIndex}.
 
@@
 \caption{Genealogy and dependencies for \Hydras packages.
-Dotted boxes represent historical Coq contributions, while regular boxes represent maintained Coq packages.
+Dotted boxes represent historical Coq contributions, while regular boxes represent maintained \coq packages.
 Dark gray packages are maintained in the \Hydras GitHub repository, while light gray packages are maintained in other \community repositories.
-Dotted lines represent Coq code ancestry, while regular lines represent direct code dependencies.}
+Dotted lines represent \coq code ancestry, while regular lines represent direct code dependencies.}
   \label{fig:genealogy}
@@
 \hline
-\textbf{Package name} & \textbf{Repository} & \textbf{Package on opam} & \textbf{Version} & \textbf{Spec LOC} & \textbf{Proof LOC}\\
+\textbf{Package name} & \textbf{Repository} & \textbf{Index identifier} & \textbf{Version} & \textbf{Spec LOC} & \textbf{Proof LOC}\\
 \hline
-Pocklington & \href{https://github.com/coq-community/pocklington}{\texttt{pocklington}} & \texttt{coq-pocklington} & \texttt{8.12.0} & & \\
+Pocklington & \href{https://github.com/coq-community/pocklington}{\texttt{pocklington}} & \texttt{coq-pocklington} & \texttt{8.12.0} & 825 & 3,798 \\
 \hline
-Goedel & \href{https://github.com/coq-community/goedel}{\texttt{goedel}} & \texttt{coq-goedel} & \texttt{8.13.0} & & \\
+Goedel & \href{https://github.com/coq-community/goedel}{\texttt{goedel}} & \texttt{coq-goedel} & \texttt{8.13.0} & 2,799 & 10,762 \\
 \hline
-Gaia & \href{https://github.com/coq-community/gaia}{\texttt{gaia}} & \texttt{coq-gaia} & \texttt{1.12} & & \\
+Gaia & \href{https://github.com/coq-community/gaia}{\texttt{gaia}} & \texttt{coq-gaia} & \texttt{1.12} & 28,850 & 124,839 \\
 \hline
@@
 allowing to present and compare various formalization and proving techniques. %It may be also used as a library, through opam packages.
-It is also a medium sized library (more than 50K LOC), dependent on various tools and libraries of the \coq ecosystem, which  may be also used to experiment new techniques of continuous maintenance of the code and its documentation.
+It is also a medium sized library (more than 50,000 LOC), dependent on various tools and libraries of the \coq ecosystem, which  may be also used to experiment new techniques of continuous maintenance of the code and its documentation.
 

### 2021-10-15 doc/JFLA2022/paper.tex
make table layouts consistent

@@
 \footnotesize
-\begin{tabular}{|l|l|l|l|l|l|}
+\begin{tabular}{llllll}
 \hline
@@
 Pocklington & \href{https://github.com/coq-community/pocklington}{\texttt{pocklington}} & \texttt{coq-pocklington} & \texttt{8.12.0} & 825 & 3,798 \\
-\hline
 Goedel & \href{https://github.com/coq-community/goedel}{\texttt{goedel}} & \texttt{coq-goedel} & \texttt{8.13.0} & 2,799 & 10,762 \\
-\hline
 Gaia & \href{https://github.com/coq-community/gaia}{\texttt{gaia}} & \texttt{coq-gaia} & \texttt{1.12} & 28,850 & 124,839 \\
-\hline
 Hydra-battles & \href{https://github.com/coq-community/hydra-battles}{\texttt{hydra-battles}} & \texttt{coq-hydra-battles} & \texttt{0.5} & & \\
-\hline
 Addition-chains & \href{https://github.com/coq-community/hydra-battles}{\texttt{hydra-battles}} & \texttt{coq-addition-chains} & \texttt{0.5} & & \\
-\hline
 Gaia-hydras & \href{https://github.com/coq-community/hydra-battles}{\texttt{hydra-battles}} & \texttt{coq-gaia-hydras} & \texttt{0.5} & & \\

### 2021-10-15 doc/JFLA2022/paper.tex
fix typo in conclusion

@@
 
-\Hydras wants to be provide a connection between scientific literature (\emph{e.g.},~\cite{KP82, KS81, schutte}) and proof assistant technology.
-For the mathematician, it may give a concrete view of the mathematical content, not only through full proofs, but also through significative computations: examples, functions associated with constructive proofs, etc.
+\Hydras wants to provide a connection between scientific literature (\emph{e.g.},~\cite{KP82, KS81, schutte}) and proof assistant technology.
+For the mathematician, it can give a concrete view of the mathematical content, not only through full proofs, but also through illuminating computations: examples, functions associated with constructive proofs, etc.
 For the \coq learner, it provides a consistent set of examples,

### 2021-10-15 doc/JFLA2022/paper.tex
final pass over section 3.4 text

@@
 
-In traditional software development, it is common for packages to continually depend on a large number of packages.
-In contrast, projects in the Coq ecosystem are often subject to ``dependency aversion'', where maintainers eschew depending on useful packages because it may lead to work in adapting to upstream changes.
-With \Hydras, we hope to demonstrate that projects with complex dependencies are feasible to manage using recent advances in build management and infrastructure.
+Before the Ackermann sublibrary was moved from the Goedel package repository to the \Hydras repository, the Goedel package had around 7,000 specification LOC and 38,000 proof LOC.
+Figure~\ref{fig:genealogy} and Table~\ref{tbl:loc} thus reveal that the seemingly large Coq formalization of G\"{o}del's incompleteness theorem can be considered as a modestly-sized proof on top of general libraries for, on the one hand, first-order logic, primitive recursive functions, and Peano arithmetic (Ackermann), and on the other hand, prime numbers and their properties (Pocklington).
 
@@
 \end{table}
+
+In traditional software development, it is common for packages to continually depend on a large number of packages.
+In contrast, projects in the Coq ecosystem are often subject to ``dependency aversion'', where maintainers eschew depending on useful packages because it may lead to work in adapting to upstream changes.
+With \Hydras, we hope to demonstrate that projects with complex dependencies are feasible to manage using recent advances in build management and infrastructure.
 

### 2021-10-15 doc/JFLA2022/paper.tex
small space adjustment
small fixes
add sentence about new collaborators

@@
   \end{tabular}
-  \caption{Coq book comparison.}
+  \caption{Coq book categorization.}
   \label{tbl:books}
@@
 Due to our reliance on LaTeX for writing the book, obtaining a book representation in HTML is more cumbersome than for executable books such as Software Foundations~\cite{SF} and Certified Programming with Dependent Types~\cite{chlipalacpdt2011}.
-Currently, we generate the Coq code documentation as HTML separately from the book.
+Currently, we generate the Coq code documentation in HTML form separately from the book.
 
@@
 We plan to extend \Hydras in two main directions.
-Firstly, we plan to build a bridge between the combinatorial results of \Hydras and the set-theoretic content of \gaia, and make possible the transfer of many interesting theorems between the two packages.
+Firstly, we plan to bridge the combinatorial results of \Hydras and the set-theoretic content of \gaia, enabling the transfer of many interesting theorems between the two packages.
 Secondly, we aim to write a formal proof in \coq of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
-
-\Hydras is not limited to the study of ordinal numbers and applications.
-For instance, we are also developing  a module about efficient exponentiation
-algorithms, and hope to extend our project to new topics.
+%
+Moreover, \Hydras is not limited to the study of ordinal numbers and applications.
+We are also developing a package about efficient exponentiation
+algorithms, and aim to eventually include new topics. We invite new collaborators to join us in our efforts.
 

### 2021-10-15 doc/JFLA2022/paper.tex
consistently use Peano Arithmetic

@@
 Before the Ackermann sublibrary was moved from the Goedel package repository to the \Hydras repository, the Goedel package had around 7,000 specification LOC and 38,000 proof LOC.
-Figure~\ref{fig:genealogy} and Table~\ref{tbl:loc} thus reveal that the seemingly large Coq formalization of G\"{o}del's incompleteness theorem can be considered as a modestly-sized proof on top of general libraries for, on the one hand, first-order logic, primitive recursive functions, and Peano arithmetic (Ackermann), and on the other hand, prime numbers and their properties (Pocklington).
+Figure~\ref{fig:genealogy} and Table~\ref{tbl:loc} thus reveal that the seemingly large Coq formalization of G\"{o}del's incompleteness theorem can be considered as a modestly-sized proof on top of general libraries for, on the one hand, first-order logic, primitive recursive functions, and Peano Arithmetic (Ackermann), and on the other hand, prime numbers and their properties (Pocklington).
 

### 2021-10-15 doc/JFLA2022/paper.tex
loc counts right-aligned in table
acks section not numbered
Hydras & Co. galaxy.

@@
 Figure~\ref{fig:genealogy} illustrates the relationships between \Hydras packages and show their ancestry from historical Coq contributions.
-To indicate the scope and relative sizes of packages, Table~\ref{tbl:loc} breaks down lines of code for each \Hydras package, as reported by the \texttt{coqwc} tool, and lists their repository names in \community and identifiers in the \coq opam package index~\cite{CoqPackageIndex}.
+To indicate the scope and relative sizes of packages, Table~\ref{tbl:loc} breaks down lines of code for each package in the \Hydras galaxy, as reported by the \texttt{coqwc} tool, and lists their repository names in \community and identifiers in the \coq opam package index~\cite{CoqPackageIndex}.
 
@@
 \footnotesize
-\begin{tabular}{llllll}
+\begin{tabular}{llllrr}
 \hline
@@
 Gaia & \href{https://github.com/coq-community/gaia}{\texttt{gaia}} & \texttt{coq-gaia} & \texttt{1.12} & 28,850 & 124,839 \\
-Hydra-battles & \href{https://github.com/coq-community/hydra-battles}{\texttt{hydra-battles}} & \texttt{coq-hydra-battles} & \texttt{0.5} & & \\
-Addition-chains & \href{https://github.com/coq-community/hydra-battles}{\texttt{hydra-battles}} & \texttt{coq-addition-chains} & \texttt{0.5} & & \\
-Gaia-hydras & \href{https://github.com/coq-community/hydra-battles}{\texttt{hydra-battles}} & \texttt{coq-gaia-hydras} & \texttt{0.5} & & \\
+Hydra-battles & \href{https://github.com/coq-community/hydra-battles}{\texttt{hydra-battles}} & \texttt{coq-hydra-battles} & \texttt{0.5} & 14,368 & 48,385 \\
+Addition-chains & \href{https://github.com/coq-community/hydra-battles}{\texttt{hydra-battles}} & \texttt{coq-addition-chains} & \texttt{0.5} & 1,961 & 2,210 \\
+Gaia-hydras & \href{https://github.com/coq-community/hydra-battles}{\texttt{hydra-battles}} & \texttt{coq-gaia-hydras} & \texttt{0.5} & 81 & 177 \\
 \hline
 \end{tabular}
-\caption{Current numbers of lines of code for \Hydras packages.}
+\caption{Current numbers of lines of code for packages in the \Hydras galaxy.}
 \label{tbl:loc}
@@
 
-\section{Acknowledgments}
+\section*{Acknowledgments}
 \label{sect:acks}

### 2021-10-15 doc/JFLA2022/paper.tex
loc counts for release

@@
 Figure~\ref{fig:genealogy} illustrates the relationships between \Hydras packages and show their ancestry from historical Coq contributions.
-To indicate the scope and relative sizes of packages, Table~\ref{tbl:loc} breaks down lines of code for each package in the \Hydras galaxy, as reported by the \texttt{coqwc} tool, and lists their repository names in \community and identifiers in the \coq opam package index~\cite{CoqPackageIndex}.
+To indicate the scope and relative sizes of packages, Table~\ref{tbl:loc} breaks down lines of code for each \Hydras package, as reported by the \texttt{coqwc} tool, and lists their repository names in \community and identifiers in the \coq opam package index~\cite{CoqPackageIndex}.
 
@@
 \end{tabular}
-\caption{Current numbers of lines of code for packages in the \Hydras galaxy.}
+\caption{Current numbers of lines of code for \Hydras packages.}
 \label{tbl:loc}

### 2021-10-20 doc/gaia-chapter.tex
add a new chapter (Gaia)

\chapter{Gaia and the hydras (draft)}

\section{Introduction}

\section{Data refinement}

\inputsnippets{T1Bridge/iotaPiDef}

### 2021-10-20 doc/gaia-chapter.tex
add a new chapter (Gaia)
minor changes

@@
 
+\emph{The following definitions and lemmas are in
+\href{../theories/html/gaia_hydras.T1Bridge.html}%
+{\texttt{gaia\_hydras.T1Bridge}}}.
+
+Since both libraries \texttt{gaia} and \texttt{hydra-battles} have their own type \texttt{T1}, we will work with the following notations.
+
+\inputsnippets{T1Bridge/hT1gT1}
+
+The bridge between both libraries is made of two functions.
+
 \inputsnippets{T1Bridge/iotaPiDef}
+
+\inputsnippets{T1Bridge/iotaPiRw, T1Bridge/piIotaRw}
+
+The functions \texttt{iota} and \texttt{pi} allow us to define
+a notion of refinement for constants, functions, predicates and relations.
+
+\inputsnippets{T1Bridge/refineDefs}

### 2021-10-21 doc/gaia-chapter.tex
more snippets in gaia-chapter

@@
 \inputsnippets{T1Bridge/refineDefs}
+
+\subsection{Examples of refinements}
+Both libraries define constants like $0$, $1$, $\omega$, and arithmetic functions: successor, addition, multiplication, and exponential of base $\omega$ (function $\phi_0$).
+We prove consistency of these definitions.
+Finally, we prove that the boolean predicates `` to be in normal form'' are equivalent.
+
+\subsubsection{A few constants}
+
+\inputsnippets{T1Bridge/constantRefs}

### 2021-10-22 doc/chapter-powers.tex
fix typos
add a link to Yves' file fib.v

@@
 In contrast, the optimizations we present in this document apply, in their vast majority, \emph{generic} techniques of efficient computation of powers in a monoid. 
-This example of Fibonacci numbers has been developed with Yves Bertot, who wrote a first version with \texttt{SSreflect/Mathcomp}~\cite{MCB}.
+This example of Fibonacci numbers has been developed with Yves Bertot, who wrote a first version with \texttt{SSreflect/Mathcomp}~\cite{MCB}. This original version is available on
+\href{../theories/html/additions.fib.html}{additions.fib}.
 
@@
 
-Page.~\pageref{sect:test-strat}, we compare the three strategies with respect to the length of the built chains.
+Page~\pageref{sect:test-strat}, we compare the three strategies with respect to the length of the built chains.
 

### 2021-10-22 doc/gaia-chapter.tex
Add symmetric notations for hydras/gaia. Fix a few typos
more snippets in gaia-chapter

@@
 \inputsnippets{T1Bridge/hT1gT1}
+
+Most used constants and functions will also get standardized
+non-ambiguous names.
+
+\inputsnippets{T1Bridge/MoreNotations}
 
@@
 \subsubsection{A few constants}
+For each constant: $0$, $1$, \dots, $n$ and $\omega$ we prove
+that the \texttt{hydras}' definition refines \texttt{gaia}'s.
+\inputsnippets{T1Bridge/constantRefs}
 
-\inputsnippets{T1Bridge/constantRefs}
+\subsubsection{Unary functions}
+
+\inputsnippets{T1Bridge/unaryRefs}
+
+\subsubsection{Binary operations}
+The binary operations $+$ and $\times$ are defined the same way in both libraries. Please note that they require comparison of ordinal terms. So, the proof of the following lemmas applies
+compatibility lemmas like \texttt{compare\_ref} (see Section~\vref{sect:gaia-compare-ref}).
+
+\inputsnippets{T1Bridge/plusRef, T1Bridge/multRef}
+
+\subsection{Order and comparison}
+
+\label{sect:gaia-compare-ref}
+\inputsnippets{T1Bridge/compareRef}
+
+\inputsnippets{T1Bridge/ltRef}

### 2021-10-25 doc/gaia-chapter.tex
more snippets in gaia-chapter

@@
 
+\subsubsection{Porting a theorem}
+The \texttt{Gaia} library containsa proof that the multiplication of ordinals less than $\epsilon_0$ is associative.
+
+\emph{From~\href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}}
+
+\begin{Coqsrc}
+Lemma mulA: associative T1mul.
+\end{Coqsrc}
+
+This lemma was missing from \texttt{hydra-battles}. Nevertheless, we could adapt this lemma to \texttt{hydra-battles}' ordinals, by a small sequence of rewritings.
+
+\inputsnippets{T1Bridge/multA}
+
+The module~\href{../theories/html/gaia_hydras.T1Bridge_use.html}%
+{\texttt{gaia\_hydras.T1Bridge\_use}} shows an example of
+application of \texttt{multA}.
+
+\inputsnippets{T1Bridge_use/T1BridgeUse}
+
 \subsection{Order and comparison}
+Leibniz equality and strict order on both libraries are
+compatible:
+\inputsnippets{T1Bridge/eqRef}
 
@@
 \inputsnippets{T1Bridge/ltRef}
+
+\inputsnippets{T1Bridge/decideHLtRw}
+
+Hence, both definitions of ``being in Cantor normal form'' are
+compatible.
+
+\inputsnippets{T1Bridge/nfRef}

### 2021-10-26 doc/gaia-chapter.tex
more snippets

@@
 The module~\href{../theories/html/gaia_hydras.T1Bridge_use.html}%
-{\texttt{gaia\_hydras.T1Bridge\_use}} shows an example of
-application of \texttt{multA}.
+{\texttt{gaia\_hydras.T1Bridge\_use}} shows a small
+example of importation of \texttt{multA} into \texttt{hydra-battles}' world.
 
@@
 \inputsnippets{T1Bridge/nfRef}
+
+\section{Gaia's proof of well-foundedness}
+The library \texttt{gaia-hydras} already contains two proofs
+of well-foundedness of the strict order on ordinal terms in
+Cantor normal form: a direct proof~\vref{sec:T1Wf} and a proof based on the recursive path ordering~\vref{remark:a3pat}, by
+\'Evelyne Contejean. We found it interesting to comment Gaia's proof, which uses  tools already present in~\cite{CantorContrib}.
+
+\subsection{Restricted recursion}
+As remarked in~\vref{sec:T1Wf}, the order \texttt{T1lt} on \texttt{T1} is not well-founded, because of terms not in normal form.
+But the \emph{restriction} of \texttt{T1lt} to terms in normal form \emph{is} well-founded.
+The following section introduces a  vocabulary in order to reason about statements of the form  ``the restriction of the relation $R$ to a subset $P$ of $A$ is well founded''.
+
+\inputsnippets{nfwfgaia/restrictedRecursion}
+
+The following induction principle is just a variant of \texttt{well\_founded\_induction\_type} for relations defined with \texttt{restrict}.
+
+\inputsnippets{nfwfgaia/restrictedRecursiona}
+
+With $P=\texttt{T1nf}$, this principle will allow us to write proofs by transfinite induction.
+
+\subsection{Main proof steps}
+
+\emph{Note: the abreviation \texttt{LT}, and the \texttt{fold LT} and \texttt{unfold LT} tactics,are not in \texttt{gaia}, but  have been added to make the goal display more readable.}
+
+\inputsnippets{nfwfgaia/nfWfProofa}
+
+The following elimination on \texttt{a} pushes a new hypothesis \texttt{Hb} which --- in the same terms as in Section~\vref{sec:strongly-accessible} --- tells that any ordinal less than $a$ is ``strongly accessible''.
+
+\inputsnippets{nfwfgaia/nfWfProofb, nfwfgaia/nfWfProofc}
+
+Now, let us consider any ordinal term which is (by \texttt{LT}) less than \texttt{(cons a n b)}.
+
+There are three cases to consider:
+\begin{itemize}
+\item It is of the form \texttt{(cons a' n' b')} where \texttt{a'} is strictly less than \texttt{a}. Then, it is accessible (by hypothesis \texttt{Hb}).
+  \inputsnippets{nfwfgaia/nfWfProofd}
+
+\end{itemize}

### 2021-10-26 doc/part-hydras.tex
small change in a snippet
fix typos
more snippets

@@
 \subsection{About  well-foundedness}
+\label{sec:T1Wf}
 \label{sec:orgheadline82}
@@
 \label{sec:orgheadline78}
+\label{sec:strongly-accessible}
+
   Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
@@
 
-\item Finally, $\canonseq{0}{i}= \alpha$.
+\item Finally, $\canonseq{0}{i}= 0$.
 \end{itemize}
@@
 \input{movies/snippets/Hprime/HprimeDef}
+\input{movies/snippets/Hprime/HprimeDefb}
 
@@
 \input{movies/snippets/Hprime/HprimeOmegaCube3b}
-
-In order to make this statement more readable, we can introduce a local définition.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3c}
-
-This number looks quite big; let us compute an approximation with \texttt{Ocaml}:
-
-\begin{Coqsrc}
-# (2.0 ** 64.0 *. 64.0 -. 1.0);; 
-\end{Coqsrc}
-
-\begin{Coqanswer}
-- : float = 1.1805916207174113e+21
-\end{Coqanswer}
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3d}
 
@@
 
+In order to make this statement more readable, we introduce a local definition.
+
+\input{movies/snippets/Hprime/HprimeOmegaCube3c}
+
+This number looks quite big; let us compute an approximation with \texttt{Ocaml}:
+
+\begin{Coqsrc}
+# (2.0 ** 64.0 *. 64.0 -. 1.0);; 
+\end{Coqsrc}
+
+\begin{Coqanswer}
+- : float = 1.1805916207174113e+21
+\end{Coqanswer}
+
+\input{movies/snippets/Hprime/HprimeOmegaCube3d}
+
 We leave as an exercise to determine the best approximation as possible of
@@
 $2^{10^{21}}$  digits. But it is still an under-approximation !
+
+\input{movies/snippets/Hprime/HprimeOmegaCube3e}
 

### 2021-10-29 doc/gaia-chapter.tex
more snippets in gaia-chapter
fix typos
untangling gaia's proof of well-foundedness of the order on T1

@@
 
+This proof, as it appears in \href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}, is quite dense and qualified as ``a bit tricky''.
+We present here its main steps with the help of the alectryon tool.
+
 \subsection{Restricted recursion}
@@
 \inputsnippets{nfwfgaia/nfWfProofa}
+We start with classic structural induction on \texttt{a}.
 
-The following elimination on \texttt{a} pushes a new hypothesis \texttt{Hb} which --- in the same terms as in Section~\vref{sec:strongly-accessible} --- tells that any ordinal less than $a$ is ``strongly accessible''.
+\inputsnippets{nfwfgaia/nfWfProofaa}
+The following \texttt{ssreflect} tactic  pushes a new hypothesis \texttt{Hb} which --- in the same terms as in Section~\vref{sec:strongly-accessible} --- tells that any ordinal less than $a$ is ``strongly accessible''.
 
-\inputsnippets{nfwfgaia/nfWfProofb, nfwfgaia/nfWfProofc}
+\inputsnippets{nfwfgaia/nfWfProofb}
 
@@
 
-There are three cases to consider:
-\begin{itemize}
-\item It is of the form \texttt{(cons a' n' b')} where \texttt{a'} is strictly less than \texttt{a}. Then, it is accessible (by hypothesis \texttt{Hb}).
+There are three cases to consider.
+\begin{enumerate}
+\item It is of the form \texttt{(cons a' n' b')} where \texttt{a'} is strictly less than \texttt{a}.
   \inputsnippets{nfwfgaia/nfWfProofd}
 
-\end{itemize}
+  Then, it is accessible (by hypothesis \texttt{Hb}).
+
+  \item The considered term  is equal to
+    \texttt{(cons a n b'')}, where $b''<\omega^a$. 
+    \inputsnippets{nfwfgaia/nfWfProofg}
+
+    Then we can replace \texttt{a''} with \texttt{a}, \texttt{n''} with \texttt{n}, and apply the hypotheses \texttt{Hd} and \texttt{He}.
+
+\item
+  The term is equal to
+  \texttt{(cons a n'' b'')}, where $n''<n$ and $b''<\omega^a$.
+  This term is less than the accessible \texttt{(cons a n b'')}.
+
+  \inputsnippets{nfwfgaia/nfWfProofe}
+
+ Then we can apply the hypothesis \texttt{He}.
+
+\end{enumerate}
+
+In every case, the considered ordinal is accessible.

### 2021-10-29 doc/chapter-powers.tex
Add symmetric notations for hydras/gaia. Fix a few typos
more snippets in gaia-chapter
simplifying latex macros

@@
 In contrast, the optimizations we present in this document apply, in their vast majority, \emph{generic} techniques of efficient computation of powers in a monoid. 
-This example of Fibonacci numbers has been developed with Yves Bertot, who wrote a first version with \texttt{SSreflect/Mathcomp}~\cite{MCB}. This original version is available on
-\href{../theories/html/additions.fib.html}{additions.fib}.
+This example of Fibonacci numbers has been developed with Yves Bertot, who wrote a first version with \texttt{SSreflect/Mathcomp}~\cite{MCB}.
+This original version is available on hydra-battles in the module \href{../theories/html/additions.fib.html}{additions.fib}.
 
@@
 You may uncomment them freely in your own copy.
-For the same reason, we put a verbatim trace instead of \textit{Alectryon} output}.
+For the same reason, we put a verbatim trace instead of \alectr  output}.
 

### 2021-10-29 doc/chapter-primrec.tex
fix typos
simplifying latex macros

@@
 Thus they do not allow the user to look at the witness of a proof of a \texttt{isPR} statement.
-Our example of page\pageref{sect:bigfac} was built using a  copy of \texttt{primRec.v} where many \texttt{Qed}s have been replaced with
+Our example of page~\pageref{sect:bigfac} was built using a  copy of \texttt{primRec.v} where many \texttt{Qed}s have been replaced with
 \texttt{Defined}s.
@@
 
-Our proof of this lemma is not trivial at all, it uses some properties of the Ketonen-Solovay's toolkit. We advise the reader to explore this proof, with the help of an IDE or software like Alectryon.
+Our proof of this lemma is not trivial at all, it uses some properties of the Ketonen-Solovay's toolkit. We advise the reader to explore this proof, with the help of an IDE or software like \alectr.
 %%

### 2021-10-29 doc/gaia-chapter.tex
simplifying latex macros

@@
 
-This proof, as it appears in \href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}, is quite dense and qualified as ``a bit tricky''.
-We present here its main steps with the help of the alectryon tool.
+This proof, as it appears inw, is quite dense and qualified as ``a bit tricky''. We present here its main steps with the help of the \alectr tool.
 
@@
   \texttt{(cons a n'' b'')}, where $n''<n$ and $b''<\omega^a$.
-  This term is less than the accessible \texttt{(cons a n b'')}.
 
@@
 
- Then we can apply the hypothesis \texttt{He}.
+  By \texttt{Hd} and \texttt{He}, and a few technical lemmas about normal forms,   the term  \texttt{(cons a n b'')} is accessible, and so is \texttt{(cons a n'' b'')}.
 
@@
 In every case, the considered ordinal is accessible.
+By the definition of accessibility, the ordinal \texttt{(cons a n b)} is accessible, which ends the proof.
+The reader is invited to replay the full proof in  \href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}, and to compare with
+the proof in Section~\ref{sec:strongly-accessible}.

### 2021-10-29 doc/hydras.tex
and Cie -> and Co. (book title)
and -> ampersand (book)
title change
change title and subtitle
add a new chapter (Gaia)
simplifying latex macros

--- 2021-09-18 doc/hydras.tex
@@
-\documentclass[a4paper]{book}
+\documentclass[twoside,a4paper]{book}
 \usepackage{alltt}
@@
 \usepackage{url}
-\usepackage{xcolor,fancyvrb, mdframed}
+\usepackage{xcolor, fancyvrb, mdframed}
 \usepackage{amsmath}
+%\usepackage{fancyhdr}
 \usepackage{varioref}
@@
 \usepackage{theorem}
-
+\usepackage{xspace}
 \usepackage{texments}
@@
 \usepackage{movies/snippets/assets/pygments}
+
+% Prevent breaks in the middle of syntactic units
+\let\OldPY\PY
+\def\PY#1#2{\OldPY{#1}{\mbox{#2}}}
 
@@
 
-\title{Hydra Battles and Cie}
+\title{Hydras, Ordinals \& Co.\\ A library in Coq of entertaining formal mathematics}
+%\pagestyle{fancy}
+%\fancyhead{Hydras, Ordinals, \& Co.}
 \makeindex{coq}
@@
 
-This document is \emph{not} an introductory text for \coq{}, and there are many aspects of this proof assistant that are not covered. 
- The reader should already have some basic experience with the \coq{} system.
-The Reference Manual and several tutorials are available on the \coq{} website~\cite{Coq}.
+This document is \emph{not} an introductory text for \coq, and there are many aspects of this proof assistant that are not covered. 
+ The reader should already have some basic experience with the \coq system.
+The Reference Manual and several tutorials are available on the \coq website~\cite{Coq}.
 The first chapters of textbooks like \emph{Interactive Theorem Proving and Program Development}~\cite{BC04}, \emph{Software Foundations}~\cite{SF} or  \emph{Certified Programming with Dependent Types} ~\cite{chlipalacpdt2011} will give you the right background.
@@
 
-\subsection{Documenting theories with Alectryon}
+\subsection{Documenting theories with \alectr}
 
-Quotations of \coq{} source and answers are progressively replaced from copy-pasted \emph{verbatim} to automatically generated \emph{LaTeX} blocks, using Clément Pit-Claudel's \emph{Alectryon} tool~\cite{alectryonpaper, alectryongithub}.
+Quotations of \coq source and answers are progressively replaced from copy-pasted \emph{verbatim} to automatically generated \emph{LaTeX} blocks, using Clément Pit-Claudel's \alectr  tool~\cite{alectryonpaper, alectryongithub}.
 Many thanks to Jérémy Damour, Clément Pit-Claudel  and Théo Zimmermann who designed tools for maintaining consistency between the always evolving \coq{} modules and documentation written in \emph{LaTeX}.
@@
 
-\subsubsection{Using Alectryon}
+\subsubsection{Using \alectr}
 
@@
 \include{chapter-primrec}
+\include{gaia-chapter}
 \part{A few  certified algorithms}

### 2021-10-29 doc/gaia-chapter.tex
simplifying latex macros

@@
 
-\emph{Note: the abreviation \texttt{LT}, and the \texttt{fold LT} and \texttt{unfold LT} tactics,are not in \texttt{gaia}, but  have been added to make the goal display more readable.}
+\emph{Notes: the abbreviation \texttt{LT}, and the \texttt{fold LT} and \texttt{unfold LT} tactics,are not in \texttt{gaia}, but  have been added to make the goal display more readable.
+The infix notation \texttt{`` \_ < \_''}  is bound to the order \texttt{T1lt} in \gaia, while,  in \Hydras,  it is bound to its restriction \texttt{LT} to terms in normal form. }
 
 \inputsnippets{nfwfgaia/nfWfProofa}
-We start with classic structural induction on \texttt{a}.
+We start with a classic structural induction on the type \texttt{T1}.
 

### 2021-10-30 doc/gaia-chapter.tex
changes in gaia-chapter

@@
 
-This proof, as it appears inw, is quite dense and qualified as ``a bit tricky''. We present here its main steps with the help of the \alectr tool.
+This proof is quite dense and qualified as ``a bit tricky''. We just present here its main steps with the help of the \alectr tool.
+
+The reader is invited to replay the full proof in  \href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}, and to compare with
+the proof in Section~\ref{sec:strongly-accessible}.
 
@@
 
+The following sentence gives us the possibility of induct over \texttt{a}'s accessibility.
+
+\inputsnippets{nfwfgaia/nfWfProofbb}
+
+%\inputsnippets{nfwfgaia/haveHc}
+
+\inputsnippets{nfwfgaia/haveHd}
+
+\inputsnippets{nfwfgaia/hdProved}
+
 Now, let us consider any ordinal term which is (by \texttt{LT}) less than \texttt{(cons a n b)}.
@@
 
-In every case, the considered ordinal is accessible.
-By the definition of accessibility, the ordinal \texttt{(cons a n b)} is accessible, which ends the proof.
-The reader is invited to replay the full proof in  \href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}, and to compare with
-the proof in Section~\ref{sec:strongly-accessible}.
+In every case, the considered ordinal is accessible. By the definition of accessibility, the ordinal \texttt{(cons a n b)} is accessible, which ends the proof.

### 2021-11-04 doc/JFLA2022/paper.tex
Rephrase lines of code stats in intro

--- 2021-10-15 doc/JFLA2022/paper.tex
@@
 Libraries of formalized mathematics based on proof assistants, such as \coq, Lean, and Isabelle/HOL, are continually growing in size and scope.
-For example, the core projects in the Mathematical Components family for Coq amount to more than 160,000 lines of code (LOC) and 11,000 lemmas~\cite{Nie2020a}, while Lean's Mathlib consists of more than 140,000 LOC and 34,000 declarations~\cite{Mathlib}.
+For example, the core projects in the Mathematical Components family for Coq and in Lean's Mathlib amount to hundreds of thousands of lines and tens of thousands of definitions and theorems~\cite{Nie2020a, Mathlib}.
 However, after a key mathematical definition or result is added to a library, it must be \emph{documented and maintained}~\cite{vanDoorn2020,alectryonpaper}.

### 2021-11-04 doc/JFLA2022/paper.tex
Resurrect the definition of "LOC"

@@
 Libraries of formalized mathematics based on proof assistants, such as \coq, Lean, and Isabelle/HOL, are continually growing in size and scope.
-For example, the core projects in the Mathematical Components family for Coq and in Lean's Mathlib amount to hundreds of thousands of lines and tens of thousands of definitions and theorems~\cite{Nie2020a, Mathlib}.
+For example, the core projects in the Mathematical Components family for Coq and in Lean's Mathlib amount to hundreds of thousands of lines of code (LOC) and tens of thousands of definitions and theorems~\cite{Nie2020a, Mathlib}.
 However, after a key mathematical definition or result is added to a library, it must be \emph{documented and maintained}~\cite{vanDoorn2020,alectryonpaper}.

### 2021-11-04 doc/JFLA2022/paper.tex
change phrasing slightly on MathComp and Mathlib in intro

--- 2021-11-04 doc/JFLA2022/paper.tex
@@
 Libraries of formalized mathematics based on proof assistants, such as \coq, Lean, and Isabelle/HOL, are continually growing in size and scope.
-For example, the core projects in the Mathematical Components family for Coq and in Lean's Mathlib amount to hundreds of thousands of lines and tens of thousands of definitions and theorems~\cite{Nie2020a, Mathlib}.
+For example, Mathlib for Lean and the core set of projects in the Mathematical Components family for Coq both amount to hundreds of thousands of lines of code (LOC) and tens of thousands of definitions and theorems~\cite{Nie2020a, Mathlib}.
 However, after a key mathematical definition or result is added to a library, it must be \emph{documented and maintained}~\cite{vanDoorn2020,alectryonpaper}.

### 2021-11-07 doc/gaia-chapter.tex
minor changes
add a lemma; small change in gaia-chapter.tex

@@
 \section{Gaia's proof of well-foundedness}
-The library \texttt{gaia-hydras} already contains two proofs
+The library \Hydras already contains two proofs
 of well-foundedness of the strict order on ordinal terms in
@@
 
-%\inputsnippets{nfwfgaia/haveHc}
+By the following two forward steps, we show that $b$ is accessible, and will be able to prove properties by induction on $b$'s accessibility.
+
+\inputsnippets{nfwfgaia/haveHc}
 

### 2021-11-12 doc/hydras.tex
small changes

@@
 
-\index{coq}{Commands!Print Assumptions}
+%\index{coq}{Commands!Print Assumptions}
 
@@
 % \end{todo}
-{\Large \textbf{In progress} This index is currently under reorganization for a few days. We apologize for its incompleteness! }
+{\Large \textbf{In progress} This index is currently under reorganization.  We apologize for its incompleteness! }
 
@@
 \printindex{maths}{Mathematical notions and algorithmics}
-\printindex{hydras}{Library hydras: Ordinals and hydra battles}
-\printindex{primrec}{Library hydras.Ackermann: Primitive recursive functions}
-\printindex{additions}{Library additions: Addition chains}
+%\printindex{hydras}{Library hydras: Ordinals and hydra battles}
+%\printindex{primrec}{Library hydras.Ackermann: Primitive recursive functions}
+%\printindex{additions}{Library additions: Addition chains}
 
-\section{Main notations}
+% \section{Main notations}
 
-\begin{table}[H]
-  \centering
-  \begin{threeparttable}
-    \caption{Ordinals and ordinal notations}
-\begin{tabular}{|r | c|c|c|c|l|}
-\hline
-Name & Gallina&Math& Description& Page \\\hline
-\texttt{lt : T1->T1->Prop}& lt alpha beta & $\alpha < \beta$& strict order on type \texttt{T1} \tnote{1} & \pageref{Predicates:lt-T1}\\
-\texttt{LT: T1->T1->Prop}& alpha o< beta & $\alpha < \beta$& strict order on type \texttt{T1}   \tnote{2} & \pageref{Predicates:LT-T1}\\
-\texttt{Lt : E0->E0->Prop} & alpha o< beta & $\alpha < \beta$& strict order on type \texttt{E0} \tnote{3} & \pageref{Predicates:Lt-E0} \\
-\texttt{nf: T1->Prop} & \texttt{nf alpha} && alpha is in Cantor normal form & \pageref{Predicates:nf-T1}\\
- \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality \tnote{4} & \pageref{sect:on-lt-notation}\\
- \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
-\texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & ordinal addition & \pageref{sect:infix-plus-T1}, \dots\\
-\texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\
+% \begin{table}[H]
+%   \centering
+%   \begin{threeparttable}
+%     \caption{Ordinals and ordinal notations}
+% \begin{tabular}{|r | c|c|c|c|l|}
+% \hline
+% Name & Gallina&Math& Description& Page \\\hline
+% \texttt{lt : T1->T1->Prop}& lt alpha beta & $\alpha < \beta$& strict order on type \texttt{T1} \tnote{1} & \pageref{Predicates:lt-T1}\\
+% \texttt{LT: T1->T1->Prop}& alpha o< beta & $\alpha < \beta$& strict order on type \texttt{T1}   \tnote{2} & \pageref{Predicates:LT-T1}\\
+% \texttt{Lt : E0->E0->Prop} & alpha o< beta & $\alpha < \beta$& strict order on type \texttt{E0} \tnote{3} & \pageref{Predicates:Lt-E0} \\
+% \texttt{nf: T1->Prop} & \texttt{nf alpha} && alpha is in Cantor normal form & \pageref{Predicates:nf-T1}\\
+%  \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality \tnote{4} & \pageref{sect:on-lt-notation}\\
+%  \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
+% \texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & ordinal addition & \pageref{sect:infix-plus-T1}, \dots\\
+% \texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\
 
-F & \texttt{F $n$} & $n$ & The $n$-th finite ordinal &  
-\pageref{sect:notation-F}, \pageref{sect:notation-F-sch}\\ 
-FS & \texttt{FS $n$} & $n+1$ & The $n+1$-th finite ordinal  \tnote{5} &  
-\pageref{sect:notation-FS}\\ 
-omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega}, \pageref{sect:omega-T1}, \pageref{sect:omega-notation2}, \dots\\
-phi0     & \texttt{phi0 alpha} & $\phi_0(\alpha),\; \omega^\alpha$&exponential of base $\omega$ & \pageref{sect:notation-phi0}\\
+% F & \texttt{F $n$} & $n$ & The $n$-th finite ordinal &  
+% \pageref{sect:notation-F}, \pageref{sect:notation-F-sch}\\ 
+% FS & \texttt{FS $n$} & $n+1$ & The $n+1$-th finite ordinal  \tnote{5} &  
+% \pageref{sect:notation-FS}\\ 
+% omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega}, \pageref{sect:omega-T1}, \pageref{sect:omega-notation2}, \dots\\
+% phi0     & \texttt{phi0 alpha} & $\phi_0(\alpha),\; \omega^\alpha$&exponential of base $\omega$ & \pageref{sect:notation-phi0}\\
 
-\hline
-\end{tabular}
-\begin{tablenotes}
-  \item[1] This order is total, but not well-founded, because of not well formed terms.
-\item[2] Restriction of \texttt{lt} to terms in normal form; this order is partial, but well-founded.
-\item[3] This order is total \emph{and} well-founded.
-\item [4]
-Some notations may belong to several scopes. For instance, ``\texttt{o<}'' is
-bound in \texttt{ON\_scope}, \texttt{E0\_scope}, \texttt{t1\_scope}, etc., and locally in several libraries.
-  \item [5] Note that there exist also various coercions from \texttt{nat} to types of ordinal.
+% \hline
+% \end{tabular}
+% \begin{tablenotes}
+%   \item[1] This order is total, but not well-founded, because of not well formed terms.
+% \item[2] Restriction of \texttt{lt} to terms in normal form; this order is partial, but well-founded.
+% \item[3] This order is total \emph{and} well-founded.
+% \item [4]
+% Some notations may belong to several scopes. For instance, ``\texttt{o<}'' is
+% bound in \texttt{ON\_scope}, \texttt{E0\_scope}, \texttt{t1\_scope}, etc., and locally in several libraries.
+%   \item [5] Note that there exist also various coercions from \texttt{nat} to types of ordinal.
 Depending on the current scope and  \coq's syntactic analysis algorithm, \texttt{F} may be left implicit.
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
+% \end{tablenotes}
+% \end{threeparttable}
+% \end{table}
 
-\vspace{4pt}
+% \vspace{4pt}
 
-\begin{table}[H]
+% \begin{table}[H]
 
-  \begin{threeparttable}
-    \caption{hydra battles}
-\begin{tabular}{|c|c|c|c|l|}
-\hline
-Name & Gallina&Math& Description& Page \\\hline
-\texttt{round} & \texttt{h -1-> h'} & & one round of a battle & \pageref{sect:infix-round} \\
-\texttt{rounds} & \texttt{h -+-> h'} & & one or more  rounds of a battle & \pageref{sect:infix-rounds} \\
-\texttt{round\_star} & \texttt{h -*-> h'} & & any number of rounds of a battle & \pageref{sect:infix-rounds} \\
-\hline
-\end{tabular}
+%   \begin{threeparttable}
+%     \caption{hydra battles}
+% \begin{tabular}{|c|c|c|c|l|}
+% \hline
+% Name & Gallina&Math& Description& Page \\\hline
+% \texttt{round} & \texttt{h -1-> h'} & & one round of a battle & \pageref{sect:infix-round} \\
+% \texttt{rounds} & \texttt{h -+-> h'} & & one or more  rounds of a battle & \pageref{sect:infix-rounds} \\
+% \texttt{round\_star} & \texttt{h -*-> h'} & & any number of rounds of a battle & \pageref{sect:infix-rounds} \\
+% \hline
+% \end{tabular}
 
-%\begin{tablenotes}
-%\end{tablenotes}
+% %\begin{tablenotes}
+% %\end{tablenotes}
 
-  \end{threeparttable}
+%   \end{threeparttable}
 
-\end{table}
+% \end{table}
 
-\begin{table}[H]
-  \centering
-  \begin{threeparttable}
-    \caption{Addition chains}
-\begin{tabular}{|c|c|c|c|l|}
-\hline
-Name & Gallina&Math& Description& Page \\\hline
-\texttt{Mult} & \texttt{$z$ <--- $x$ times $y$} & & monadic notation & \pageref{monadic-mult} \\
-\hline
-\end{tabular}
+% \begin{table}[H]
+%   \centering
+%   \begin{threeparttable}
+%     \caption{Addition chains}
+% \begin{tabular}{|c|c|c|c|l|}
+% \hline
+% Name & Gallina&Math& Description& Page \\\hline
+% \texttt{Mult} & \texttt{$z$ <--- $x$ times $y$} & & monadic notation & \pageref{monadic-mult} \\
+% \hline
+% \end{tabular}
 
-%\begin{tablenotes}
-%\end{tablenotes}
+% %\begin{tablenotes}
+% %\end{tablenotes}
 
-  \end{threeparttable}
+%   \end{threeparttable}
 
-\end{table}
+% \end{table}
 

### 2021-11-12 doc/part-hydras.tex
minor changes
small changes

@@
 \index{coq}{Mutually inductive types}
-\index{coq}{Commands!Scheme}
+%\index{coq}{Commands!Scheme}
 
@@
 \subsection{A  small proof of impossibility}
-\index{coq}{Proofs of impossibility}
+%\index{coq}{Proofs of impossibility}
 
@@
 
-\index{coq}{Commands!Function}
+%\index{coq}{Commands!Function}
 
@@
 
-\index{coq}{Commands!Program}
+%\index{coq}{Commands!Program}
 
@@
 \label{sec:orgheadline77}
-\index{coq}{Well-founded induction}
+%\index{coq}{Well-founded induction}
 
@@
 
+    \index{hydras}{Exercises}
+
+\begin{exercise}[not trivial]
+  It may be interesting to write a \emph{direct} proof  of well-foundedness of the order in $\omega^\omega$ (\emph{i.e.}
+  without using properties of $\epsilon_0$). This exercise may
+  help to understand better the proof structure of Sect.~\vref{sec:strongly-accessible}.
+    \end{exercise}
 %ici
@@
 to \texttt{nat} is available on \href{../theories/html/hydras.Prelude.Fuel.html}{Prelude.Fuel}).
-\index{coq}{Giving fuel to a long computation}
+%\index{coq}{Giving fuel to a long computation}
 
@@
 \label{lemma:L-2_6-1}
-\index{coq}{Sigma types}
+%\index{coq}{Sigma types}
 

### 2021-11-12 doc/JFLA2022/paper.tex
input paper.bbl in paper.tex
add snippets in the JFLA200 directory
fix a few typos

@@
 %%% for movies by alectryon
-\usepackage{./assets/alectryon}
-\usepackage{./assets/pygments}
+\usepackage{alectryon}
+\usepackage{pygments}
 %%% One hypothesis per line 
@@
   {{\renewenvironment{alectryon}{}{}% Skip \begin{alectryon} included in snippet
-    \input@old{snippets/#1}}}
+    \input@old{#1}}}
 \makeatother
@@
 \begin{theorem}
-  In the Hydra game, Hercules eventually wins, whichever the strategy of both players :
+  In the Hydra game, Hercules eventually wins, whichever the strategy of both players:
   choice of a head to chop off, choice of the number of copies. 
@@
 \section{Recent developments}
-The 2018 article~\cite{JFLA2018paper} presented a formal proof of  a variant of Theorem~\ref{kp:thm2}:
+The 2018 article~\cite{JFLA2018paper} presented a formal proof of the following theorem:
 
@@
     \begin{minipage}[h]{1.0\linewidth}
- \inputsnippets{Hprime/HprimeDef}     
+ \inputsnippets{Hprime_HprimeDef}     
     \end{minipage}}
@@
 The order in which these references appear in the book is independent of the structure of our libraries, so we chose to maintain the book as a standalone document, separate from the \coq source code.
-This in contrast with the Coqdoc approach, where explanatory prose is embedded within \coq source files (a detailed discussion of various approaches to the documentation of \coq developments is in \cite{alectryonpaper}).
+This  contrasts with the Coqdoc approach, where explanatory prose is embedded within \coq source files (a detailed discussion of various approaches to the documentation of \coq developments can be found  in \cite{alectryonpaper}).
 
@@
        \begin{minipage}[h]{1.0\linewidth}
-         \inputsnippets{Pow/evalPow17LetIn}
+         \inputsnippets{Pow_evalPow17LetIn}
        \end{minipage}}
@@
       \begin{minipage}[h]{1.0\linewidth}
-        \inputsnippets{E0/Ex42}
+        \inputsnippets{E0_Ex42}
       \end{minipage}}
@@
 
-\inputsnippets{Schutte/Ex42a}
+\inputsnippets{Schutte_Ex42a}
 
@@
 
-\inputsnippets{Schutte/Ex42b}
+\inputsnippets{Schutte_Ex42b}
 
@@
 
-\inputsnippets{Schutte/Ex42d, %to remove if short paper
-  Schutte/Ex42c, Schutte/Ex42e}
+\inputsnippets{Schutte_Ex42d, %to remove if short paper
+  Schutte_Ex42c, Schutte_Ex42e}
 
@@
 %\bibliographystyle{abbrv}
-\bibliography{../thebib}
+\input{paper.bbl}
+
+%\bibliography{../thebib}
 

### 2021-11-14 doc/JFLA2022/paper.tex
minor changes

@@
 
- Our new proof of Theorem~\ref{thm3} is based on a systematic study of strictly decreasing sequences of ordinals below $\epsilon_0$, borrowed from Ketonen and Solovay~\cite{KS81}, and the formalization of which is described  in chapters 5 and 6 of~\cite{HydraBook}.
+ Our new proof of Theorem~\ref{thm3} is based on a systematic study of strictly decreasing sequences of ordinals below $\epsilon_0$, borrowed from Ketonen and Solovay~\cite{KS81}, and the formalization of which is described in detail  in chapters 5 and 6 of~\cite{HydraBook}.
 
-We also study the number of steps of a battle:
+ Besides  Theorem~\ref{thm3}, we study also the number of steps of a battle:
 Let $\alpha<\epsilon_0$ be an ordinal. 

### 2021-11-14 doc/JFLA2022/paper.tex

@@
   providing exposition of theories and reasoning techniques---they
-  can also provide inspiration and entertainment that transcends educational goals.
+  can also provide inspiration and entertainment that transcend educational goals.
 \end{abstract}

### 2021-11-14 doc/JFLA2022/paper.tex
fix two formulations in final sections

@@
 Specifically, in an executable book, the proof assistant language places limits on the structure.
-For example, Coq may not permit repeating a definition from another module verbatim, since the definition uses section variables that are not present in the current context.
+For example, Coq may not permit repeating a definition from another file verbatim, since the definition uses variables that are not present in the current context.
 Thanks to our \alectr based toolchain, the \Hydras book can include ``live'' code fragments at any point in the text.
-And in contrast to traditional textbooks, we validate the consistency of code fragments during every build of the electronic book.
+And in contrast to traditional books, we validate the consistency of code fragments during every build of the electronic book.
 On the one hand, we follow executable books in providing ``continuous delivery'' of new revisions at a high pace.
@@
 allowing to present and compare various formalization and proving techniques. %It may be also used as a library, through opam packages.
-It is also a medium sized library (more than 50,000 LOC), dependent on various tools and libraries of the \coq ecosystem, which  may be also used to experiment new techniques of continuous maintenance of the code and its documentation.
+It is also a medium sized library (more than 50,000 LOC), dependent on various tools and libraries of the \coq ecosystem, which may be also used to experiment with new maintenance techniques of the code and its documentation.
 

### 2021-11-14 doc/JFLA2022/paper.tex
formulation fix

@@
 Before the Ackermann sublibrary was moved from the Goedel package repository to the \Hydras repository, the Goedel package had around 7,000 specification LOC and 38,000 proof LOC.
-Figure~\ref{fig:genealogy} and Table~\ref{tbl:loc} thus reveal that the seemingly large Coq formalization of G\"{o}del's incompleteness theorem can be considered as a modestly-sized proof on top of general libraries for, on the one hand, first-order logic, primitive recursive functions, and Peano Arithmetic (Ackermann), and on the other hand, prime numbers and their properties (Pocklington).
+Figure~\ref{fig:genealogy} and Table~\ref{tbl:loc} thus reveal that the seemingly large Coq formalization of G\"{o}del's incompleteness theorem can be viewed as a modestly-sized proof on top of general libraries for, on the one hand, first-order logic, primitive recursive functions, and Peano Arithmetic (Ackermann), and on the other hand, prime numbers and their properties (Pocklington).
 

### 2021-11-16 doc/hydras.tex
add a link to JFLA's paper
add a link to JFLA; still have to write a small text
Update doc/hydras.tex
Co-authored-by: Karl Palmskog <palmskog@gmail.com>

@@
 \SetWatermarkScale{0.35}
-\SetWatermarkText{Work in Progress}
+\SetWatermarkText{Work Continuously in Progress}
 \SetWatermarkColor{lightorange}
@@
 
-\title{Hydras, Ordinals \& Co.\\ A library in Coq of entertaining formal mathematics}
+\title{Hydras \& Co.\\ Formalized mathematics in \coq for inspiration and entertainment}
 %\pagestyle{fancy}
@@
 
-%%%%% ICI
+ \subsection{Developing and maintaining the \coq code and documentation}
 
-\subsection{Documenting theories with \alectr}
+ \textbf{TODO} introductory text and link to \cite{jfla2022}
+
+ \subsubsection{Continuous Integration and Continuous Delivery}
+
+ \textbf{TODO}
+
+\subsubsection{Documenting theories with \alectr}
 

### 2021-11-18 doc/JFLA2022/paper.tex
be precise about coq-community founders

@@
 The \Hydras project, part of \community on GitHub~\cite{CoqCommunity}, aims to be an experimental platform for the collaborative development of documented libraries of formal proofs.
-\community is a community organization that we founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting \coq packages, and working collaboratively on documentation projects.
+\community is a community organization that the first and last authors founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting \coq packages, and working collaboratively on documentation projects.
 The \Hydras project demonstrates that these two goals are not independent: interesting \coq packages can become the basis for new documentation.

### 2021-11-18 doc/JFLA2022/paper.tex
consistent notation for coqdoc, slight reformulation on documentation

@@
 \newcommand{\coq}{Coq\xspace}
+\newcommand{\coqdoc}{Coqdoc\xspace}
 \newcommand{\community}{Coq-community\xspace}
@@
 The order in which these references appear in the book is independent of the structure of our libraries, so we chose to maintain the book as a standalone document, separate from the \coq source code.
-This  contrasts with the Coqdoc approach, where explanatory prose is embedded within \coq source files (a detailed discussion of various approaches to the documentation of \coq developments can be found  in \cite{alectryonpaper}).
+This contrasts with the \coqdoc approach, where explanatory prose is embedded within \coq source files (a detailed discussion of different approaches to documenting \coq developments can be found in \cite{alectryonpaper}).
 
@@
 
-\item To generate the documentation of the project (book in PDF format and coqdoc HTML pages) by relying on the output of the coq\_makefile build.
+\item To generate the documentation of the project (book in PDF format and \coqdoc HTML pages) by relying on the output of the coq\_makefile build.
 

### 2021-11-18 doc/JFLA2022/paper.tex
macro for coq_makefile

@@
 \newcommand{\coqdoc}{Coqdoc\xspace}
+\newcommand{\coqmakefile}{coq\_makefile\xspace}
 \newcommand{\community}{Coq-community\xspace}
@@
 
-In practice, we do still support building the whole project with coq\_makefile, and we actually rely on this build system in our documentation generation pipeline.
+In practice, we do still support building the whole project with \coqmakefile, and we actually rely on this build system in our documentation generation pipeline.
 
@@
 \begin{enumerate}
-\item To test the build of the project (as a single unit) with coq\_makefile (whereas the Docker-based CI relies on the opam packages, which use Dune to build the libraries contained in the project).
+\item To test the build of the project (as a single unit) with \coqmakefile (whereas the Docker-based CI relies on the opam packages, which use Dune to build the libraries contained in the project).
 
-\item To generate the documentation of the project (book in PDF format and \coqdoc HTML pages) by relying on the output of the coq\_makefile build.
+\item To generate the documentation of the project (book in PDF format and \coqdoc HTML pages) by relying on the output of the \coqmakefile build.
 

### 2021-11-18 doc/JFLA2022/paper.tex
attempt at explaining why we use coq_makefile

@@
 \newcommand{\Hydras}{Hydras \& Co$\text.$\xspace}
+\newcommand{\make}{\texttt{make}\xspace}
 %\newcommand{\Hydras}{Hydra-battles\xspace}
@@
 
-In practice, we do still support building the whole project with \coqmakefile, and we actually rely on this build system in our documentation generation pipeline.
+In practice, we still support building the whole project using \make via the \coqmakefile tool bundled with \coq, and we actually rely on this build system in our documentation generation pipeline.
+The main reason is that using \coqmakefile is currently more convenient than Dune for local IDE-based Coq code development, e.g., due to Dune placing generated files in a separate directory.
+We expect to migrate fully to Dune in the near future.
 

### 2021-11-18 doc/JFLA2022/paper.tex
use texttt for all Coq-bundled tools
mention the MIT license and why
Uniformize writing of sub-library.

@@
 \newcommand{\coqdoc}{Coqdoc\xspace}
-\newcommand{\coqmakefile}{coq\_makefile\xspace}
+\newcommand{\coqmakefile}{\texttt{coq\_makefile}\xspace}
 \newcommand{\community}{Coq-community\xspace}
@@
 
-Before the Ackermann sublibrary was moved from the Goedel package repository to the \Hydras repository, the Goedel package had around 7,000 specification LOC and 38,000 proof LOC.
+Before the Ackermann sub-library was moved from the Goedel package repository to the \Hydras repository, the Goedel package had around 7,000 specification LOC and 38,000 proof LOC.
 Figure~\ref{fig:genealogy} and Table~\ref{tbl:loc} thus reveal that the seemingly large Coq formalization of G\"{o}del's incompleteness theorem can be viewed as a modestly-sized proof on top of general libraries for, on the one hand, first-order logic, primitive recursive functions, and Peano Arithmetic (Ackermann), and on the other hand, prime numbers and their properties (Pocklington).
@@
 It is also a medium sized library (more than 50,000 LOC), dependent on various tools and libraries of the \coq ecosystem, which may be also used to experiment with new maintenance techniques of the code and its documentation.
+To facilitate easy dissemination and reuse, all code and documentation is available under the permissive MIT license.
 

### 2021-11-18 doc/JFLA2022/paper.tex
qualify coq-community to be about open source packages/projects

--- 2021-11-18 doc/JFLA2022/paper.tex
@@
 The \Hydras project, part of \community on GitHub~\cite{CoqCommunity}, aims to be an experimental platform for the collaborative development of documented libraries of formal proofs.
-\community is a community organization that the first and last authors founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting \coq packages, and working collaboratively on documentation projects.
+\community is a community organization that the first and last authors founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting open source \coq packages, and working collaboratively on documentation projects.
 The \Hydras project demonstrates that these two goals are not independent: interesting \coq packages can become the basis for new documentation.
@@
 
-\community is an informal organization run by volunteer Coq users on GitHub that aims to maintain interesting Coq projects for the long term and facilitate collaboration among Coq users on documentation, tooling, etc.
+\community is an informal organization run by volunteer Coq users on GitHub that aims to maintain interesting open source Coq projects for the long term and facilitate collaboration among Coq users on documentation, tooling, etc.
 \community was created in 2018, inspired by the {Elm Community} organization~\cite{zimmermann:tel-02451322}.

### 2021-11-18 doc/JFLA2022/paper.tex
thank the anonymous reviewers under acknowledgements

@@
 \label{sect:acks}
-We are grateful to the original authors and current maintainers of the Coq packages we use and depend on: José Grimm (Gaia), Russell O'Connor (Goedel), Matthieu Sozeau (Equations), the Mathematical Components team, Marc Lasson and Chantal Keller (Paramcoq), and the authors and maintainers of \coq and its associated tools.
+We thank the anonymous reviewers for their comments, and we are grateful to the original authors and current maintainers of the Coq packages we use and depend on: José Grimm (Gaia), Russell O'Connor (Goedel), Matthieu Sozeau (Equations), the Mathematical Components team, Marc Lasson and Chantal Keller (Paramcoq), and the authors and maintainers of \coq and its associated tools.
 

### 2021-11-18 doc/JFLA2022/paper.tex
Why we cannot migrate to Dune just yet.

@@
 
-In practice, we still support building the whole project using \make via the \coqmakefile tool bundled with \coq, and we actually rely on this build system in our documentation generation pipeline.
-The main reason is that using \coqmakefile is currently more convenient than Dune for local IDE-based Coq code development, e.g., due to Dune placing generated files in a separate directory.
-We expect to migrate fully to Dune in the near future.
+However, Dune is still inconvenient for local IDE-based Coq code development and does not yet support building Coqdoc files.
+%
+Therefore, we still support building the whole project using \make via the \coqmakefile tool bundled with \coq, and we rely on this build system in our documentation generation pipeline.
+%
+We expect to migrate fully to Dune as soon as these limitations are lifted.
 

### 2021-11-18 doc/JFLA2022/paper.tex
pass over Gaia bridge section, mentioning some proof engineering concerns

@@
 
-\subsection{First steps towards a bridge to \gaia}
+\subsection{Towards a bridge to \gaia}
 
@@
 Anonymous volunteers (``collaborators of Nicolas Bourbaki'') then finished the only in-progress proof left by Grimm.
-At around 155,000 lines of code, \gaia is currently one of the largest maintained open source \coq projects~\cite{Gaia}.
+At around 155,000 LOC, \gaia is currently one of the largest maintained open source \coq projects~\cite{Gaia}.
 
 \gaia contains definitions of ordinals in Cantor and Veblen normal form~\cite{grimm:hal-00911710}, adapted from the historical Cantor contribution~\cite{CantorContrib}.
-The data types for ordinals are essentially defined the same way in \gaia and \Hydras, but they are not identical inside \coq, e.g., due to residing in different modules.
-Moreover, there are small differences in the implementation of ordinal arithmetic due to different evolutionary paths taken since divergence from the ancestral library.
-%
-Nevertheless, we were able to establish an initial correspondence between both implementations of ordinals through rewriting lemmas.
-For instance, we proved that multiplication of ordinals in Cantor normal form in \Hydras refine \gaia's corresponding multiplication operation, and were able to import \gaia's proof of associativity of the multiplication almost for free.
+The data types for ordinals are essentially defined the same way as in \Hydras, but they are not identical inside \coq, e.g., due to residing in different modules.
+There are also minor differences in how ordinal arithmetic is implemented, due to the different evolutionary paths taken since divergence from the ancestor library.
+
+As an initial step in bridging \gaia and \Hydras, we were able to establish a correspondence between both implementations of ordinals through rewriting lemmas.
+For instance, we proved that multiplication of ordinals in Cantor normal form in \Hydras refine \gaia's corresponding multiplication operation, and then imported \gaia's proof of associativity of the multiplication almost for free.
+
+The initial bridge code uses the SSReflect proof language and idioms from the Mathematical Components library.
+We made this design decision since we believe it is less challenging to reason about Hydra-battles code using SSReflect and MathComp than to reason about Gaia without SSReflect.
 
@@
 
-However, Dune is still inconvenient for local IDE-based Coq code development and does not yet support building Coqdoc files.
+However, Dune is currently inconvenient for local IDE-based Coq code development and does not yet support building \coqdoc files.
 %

### 2021-11-19 doc/JFLA2022/paper.tex
qualify the bridge code more
paragraph on portability of Alectryon snippet generation

@@
 
-The initial bridge code uses the SSReflect proof language and idioms from the Mathematical Components library.
+The initial draft bridge code mostly uses the SSReflect proof language and idioms from the Mathematical Components library.
 We made this design decision since we believe it is less challenging to reason about Hydra-battles code using SSReflect and MathComp than to reason about Gaia without SSReflect.
@@
 
+\paragraph{Applications in other proof assistants.}
+The Alectryon toolchain is intended to eventually support the languages of many proof assistants, and already has preliminary support for Lean~3.
+Our Alectryon extensions for snippet generation are in principle applicable to any language supported by Alectryon.
+However, snippet delimiters are language specific, and our tooling currently assumes Coq-compatible delimiters; we expect to lift this restriction in the near future.
+We view \Hydras as an incubator for experimental Alectryon features that, when deemed stable and useful, can be disseminated to a wider audience including users of other proof assistants.
+
 \subsection{Technologies supporting the monorepo structure}
@@
 
-\paragraph{The Docker-Coq-Action~\cite{martindorel2020dockercoq}.}
+\paragraph{Docker-Coq-Action~\cite{martindorel2020dockercoq}.}
 This GitHub Action provides a very simple way of setting up Continuous Integration (CI) for a Coq project with an opam file.

### 2021-11-19 doc/JFLA2022/paper.tex
add paragraph summarizing proof engineering improvements

@@
 %
-This umbrella project now includes evolved versions of the former Cantor and Additions libraries~\cite{CantorContrib,AdditionsContrib} (under the new names of Hydra-battles and Addition-chains), the Ackermann sub-library extracted from Russel O'Connor's Goedel library~\cite{OConnor05, Goedel}, and a bridge to the \gaia library by José Grimm~\cite{Gaia,grimm:hal-00911710}.
+This umbrella project now includes evolved versions of the former Cantor and Additions libraries~\cite{CantorContrib,AdditionsContrib} (under the new names of Hydra-battles and Addition-chains), the Ackermann sub-library extracted from Russel O'Connor's Goedel library~\cite{OConnor05, Goedel}, and a preliminary bridge to the \gaia library by José Grimm~\cite{Gaia,grimm:hal-00911710}.
 %
@@
     \end{minipage}}
- \caption{$H'$ definition with the \texttt{coq-equations}
+ \caption{$H'$ definition with the \equations
  plug-in~\cite{sozeau:hal-01671777}}
@@
 \label{sect:not-pr}
+
+\paragraph{Proof engineering improvements.}
+%
+The original release of the Cantor library preceded support for type classes in \coq~\cite{MS08}.
+Changes over time before the move of \Hydras to \community have since then introduced type classes for some concepts in the Hydra-battles library related to ordinals.
+Most type classes initially used a \emph{semi bundled} approach~\cite{Mathlib}, but we have recently \emph{unbundled} some classes to enable more sharing during resolution~\cite{BS2011}.
+Due to the modest depth of our class hierarchy, we believe this will not lead to scalability issues.
+Orthogonally to type classes, we introduced additional automation for definitions via the \equations and Paramcoq plugins.
 

### 2021-11-20 doc/tuto.tex
small tuto on Alectryon (draft)

\documentclass{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% For Alectryon

\usepackage{texments}
%%% for movies by alectryon
\usepackage{./movies/snippets/assets/alectryon}
\usepackage{./movies/snippets/assets/pygments}

% Prevent breaks in the middle of syntactic units
\let\OldPY\PY
\def\PY#1#2{\OldPY{#1}{\mbox{#2}}}

%%% One hypothesis per line 
\makeatletter
\renewcommand{\alectryon@hyps@sep}{\alectryon@nl}
\makeatother

%%% \snippets{A,B,C,…} inputs a series of snippets as one block (with \itemsep
%%% between them).  A, B, C should be paths to files in snippets/.
\usepackage{etoolbox}
\makeatletter

\newcommand{\pathtomovies}{./movies}

\newcommand{\inputsnippets}[1]
  {{\setlength{\itemsep}{1pt}\setlength{\parsep}{0pt}% Adjust spacing
    \alectryon@copymacros\begin{io}
      \forcsvlist{\item\@inputsnippet}{#1}
    \end{io}}}
\let\input@old\input % Save definition of \input
\newcommand{\@inputsnippet}[1]
  {{\renewenvironment{alectryon}{}{}% Skip \begin{alectryon} included in snippet
    \input@old{{\pathtomovies}/snippets/#1}}}
\makeatother

% End of Alectryon stuff
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Lexicographic exponentiation: An example of proof documentation with Alectryon}

\author{Labri Coq working group}

\begin{document}
\maketitle

\section{Introduction}
We show a small but non-trivial example of using Alectryon
for writing a documentation (in pdf) of a proof written in Coq.

\section{Preliminaries}
We prove that, if a type $A$ is well founded for a relation \textit{<},
then the set of sorted (w.r.t. the relation $\geq$ ) sequences on $A$  is well-founded too.
Such sequences are represented in a compact way as multisets over $A$. 
For instance, $\langle a,a,a,b,c,c,c,c,c \rangle $ where $a>b>c$ is represented by the list \texttt{(a,2)::(b,0)::(c,4)::nil}.

Note that a factor $(a,n)$ is interpreted as $n+1$ copies of $a$.

\inputsnippets{MultisetWf/tDef}

Let us consider a relation over some type $A$. We define a lexicographic order on ($\texttt{t}\;A$).

\inputsnippets{MultisetWf/AGiven, MultisetWf/lexpowerDef}

\section{\texttt{lexpower}  is not well-founded in general}

Let us build an infinite strictly decreasing (w.r.t. \texttt{lexpower lt}) sequence.

\inputsnippets{MultisetWf/notWfa}

Then, we prove that any element of the sequence \texttt{seq} is non-accessible.

\inputsnippets{MultisetWf/notWfb}

It's clearly a contradiction, since (\texttt{seq 0}) is an element of the sequence, and accessible (because of the hypothesis \texttt{Hwf}).

\inputsnippets{MultisetWf/notWfc}
\section{Lists in normal form}

We say that a list $(a_1,n_1)::(a_2,n_2)::\dots::(a_p,n_p)::nil$  in  (\texttt{t A}) is in \emph{normal form} if  for any $1<=i<n$,
we have $\texttt{ltA}\; a_{i+1}\; a_i$.

\inputsnippets{MultisetWf/lexnfDef}

The \emph{lexicographic exponentiation} of \texttt{ltA} is the restriction
of \texttt{lexpower ltA} to the set of lists in normal form.
In the next section, we prove that, if \texttt{ltA} is well-founded, then its lexicographic power is well-founded too.
\inputsnippets{MultisetWf/lexltDef}

\paragraph{Note:} The restriction of a binary relation to a set is defined in the file \texttt{Restriction.v}.

\inputsnippets{Restriction/restrictionDef}

\section{Proof of well-foundedness}
It's time to prove that the lexicographic power of a well-founded relation is well-founded too. Let us open a work section.

\inputsnippets{MultisetWf/bigProofa}

We want to prove the following statement:

\inputsnippets{MultisetWf/theStatement}

\subsection{A first attempt}
We may try to prove \texttt{lexwf} by (structural) induction on \texttt{l}.

\inputsnippets{MultisetWf/BadProof,
  MultisetWf/BadProofb}

By the hypothesis \texttt{H0}, the list \texttt{y} may be
equal to \texttt{(a,n)::l'}, where \texttt{LT l' l}. But the induction hypothesis \texttt{IHl} is useless for solving the current goal.
\inputsnippets{MultisetWf/BadProofc}

This is a classical issue. In a naive proof by induction, some variables are fixed too early: here, the variable \texttt{l} in \texttt{IHl}.

\subsection{Introducing a new predicate}

We can express the well-foundedness of \texttt{LT} in the following way :
\begin{quote}
  \begin{enumerate}
    \item  ``The empty list is accessible''
  \item   `` For any $a$ in $A$ and $n$ in \texttt{nat},  any well-formed list starting with $(a,n)$ is accessible. ''

  \end{enumerate}
 \end{quote}

The first property is easy to prove by inversion.

\inputsnippets{MultisetWf/AccNil}

For the second one, we define a new predicate over $A$.

\inputsnippets{MultisetWf/AccsDef}

The new goal is to prove that any element of $A$ satisfies
the predicate \texttt{Accs}.

\subsection{Inversion lemmas}

Before starting the proof, we  prove three useful inversion lemmas:

\inputsnippets{MultisetWf/NFInv1}
\inputsnippets{MultisetWf/NFInv2}
\inputsnippets{MultisetWf/LTInv}

\subsection{The proof by induction}
Le us prove, by induction on $a:A$ that every element of $A$ satisfies \texttt{Accs}.

\inputsnippets{MultisetWf/LAccsa}

Let us note that, for any list \texttt{(a,n)::l} in normal form,
the tail \texttt{l} is either empty or starts with some pair \texttt{(b,p)} where \texttt{ltA b a}, hence is accessible (thanks to \texttt{IHa}).

\inputsnippets{MultisetWf/LAccsb}
\inputsnippets{MultisetWf/LAccsg}

The current goal is universally quantified over \texttt{n}.
Thus we may prove it by well founded induction over \texttt{nat}.

\inputsnippets{MultisetWf/LAccsc, MultisetWf/LAccsd}

Thanks to \texttt{Hl}, we know that \texttt{l} is accessible.
Thus we can build an induction over \texttt{l}'s accessibility.

\inputsnippets{MultisetWf/LAccse}

In order to prove the accessibility of the list \texttt{((a, n) :: x0)}, we have to prove that any predecessor $y$ of this list is accessible, and consider every case (thanks to \texttt{LT\_inv}).

\inputsnippets{MultisetWf/LAccsf}

\subsubsection{First case}
If \texttt{y=nil}, then \texttt{y} is clearly accessible.

\inputsnippets{MultisetWf/case1}

\subsubsection{Second case}
Let us consider a well formed list \texttt{y} of the form \texttt{(b,p)::l'}, where \texttt{ltA b a}. By \texttt{IHa}, this list is accessible.

\inputsnippets{MultisetWf/case2}

\subsubsection{Third case}
The list \texttt{y} may be of the form  \texttt{(a,n)::l'}, where \texttt{LT l' l}. By \texttt{H2},  \texttt{y} is accessible.

\inputsnippets{MultisetWf/case3}

\subsubsection{Last case}
The list \texttt{y} may be of the form  \texttt{(a,p)::l'}, where \texttt{p<n}. Thus, we may apply \texttt{Hn}.

\inputsnippets{MultisetWf/case4}

\subsection{At last!}

\inputsnippets{MultisetWf/NFAcc}

\inputsnippets{MultisetWf/lexwf}

% \section{Examples}

% \inputsnippets{MultisetWf/Examples}

\section{Impossibility proofs}
There is no function of type \texttt{t nat ->  nat} which could serve as a measure for proving \texttt{lexlt lt}'s well-foundedness.

\inputsnippets{MultisetWf/Impossibility1}
\inputsnippets{MultisetWf/Impossibility1a}
\inputsnippets{MultisetWf/Impossibility1b}

\emph{Same exercise with \texttt{nat*nat} (with the lexicographic square of \texttt{lt})}.

\end{document}

### 2021-11-20 doc/tuto.tex
fix  typos

@@
 \subsubsection{Second case}
-Let us consider a well formed list \texttt{y} of the form \texttt{(b,p)::l'}, where \texttt{ltA b a}. By \texttt{IHa}, this list is accessible.
+Let us consider a well formed list \texttt{y} of the form \texttt{(b,p)::l''}, where \texttt{ltA b a}. By \texttt{IHa}, this list is accessible.
 
@@
 \subsubsection{Third case}
-The list \texttt{y} may be of the form  \texttt{(a,n)::l'}, where \texttt{LT l' l}. By \texttt{H2},  \texttt{y} is accessible.
+The list \texttt{y} may be of the form  \texttt{(a,n)::l''}, where \texttt{LT l'' l}. By \texttt{H2},  \texttt{y} is accessible.
 
@@
 \subsubsection{Last case}
-The list \texttt{y} may be of the form  \texttt{(a,p)::l'}, where \texttt{p<n}. Thus, we may apply \texttt{Hn}.
+The list \texttt{y} may be of the form  \texttt{(a,p)::l''}, where \texttt{p<n}. Thus, we may apply \texttt{Hn}.
 

### 2021-11-22 doc/JFLA2022/paper.tex
more concrete wording on proof engineering improvements

@@
 Changes over time before the move of \Hydras to \community have since then introduced type classes for some concepts in the Hydra-battles library related to ordinals.
-Most type classes initially used a \emph{semi bundled} approach~\cite{Mathlib}, but we have recently \emph{unbundled} some classes to enable more sharing during resolution~\cite{BS2011}.
-Due to the modest depth of our class hierarchy, we believe this will not lead to scalability issues.
-Orthogonally to type classes, we introduced additional automation for definitions via the \equations and Paramcoq plugins.
+Most type classes initially used a \emph{semi bundled} approach~\cite{Mathlib}, but we have recently \emph{unbundled} classes to enable more sharing during resolution~\cite{BS2011}.
+For example, we use the (single-field) \texttt{Decision} type class~\cite{BS2011}, that describes deciding a proposition, in parameters to other classes rather than in fields.
+Due to the modest depth of our class hierarchy, we believe unbundling will not lead to scalability issues.
+Orthogonally to type classes, we introduced automation for definitions via the \equations~\cite{sozeau:hal-01671777} and Paramcoq~\cite{Keller2012} plug-ins.
 

### 2021-11-22 doc/JFLA2022/paper.tex
elaborate based on final comment by second reviewer

@@
 Table~\ref{tbl:books} shows a category breakdown for what we believe are the most prominent Coq books in English.
-Note that several traditional books, such as Coq'Art~\cite{BC04}, used custom tools during their writing process to keep code snippets synchronized with Coq.
+Note that several traditional books, such as Coq'Art, used custom tools during their writing process to keep code snippets synchronized with Coq.
 However, once a book is published, the accompanying code generally takes on a life of its own~\cite{CoqArtCommunity} and may come to substantially diverge from the book.
@@
 And in contrast to traditional books, we validate the consistency of code fragments during every build of the electronic book.
-On the one hand, we follow executable books in providing ``continuous delivery'' of new revisions at a high pace.
+On the one hand, we follow executable books in providing continuous delivery of new revisions at a high pace.
 But on the other hand, we also aim to make regular timestamped versions compatible with specific versions of Coq, akin to new editions of traditional books.
 
-Due to our reliance on LaTeX for writing the book, obtaining a book representation in HTML is more cumbersome than for executable books such as Software Foundations~\cite{SF} and Certified Programming with Dependent Types~\cite{chlipalacpdt2011}.
-Currently, we generate the Coq code documentation in HTML form separately from the book.
+The flip side of our approach is that we separate the Coq sources from the LaTeX sources that describe the source code, and our use of snippet delimiters in the Coq source code files may also lower source readability.
+For books and libraries with small scope and specialized audiences, distributing knowledge representations between sources and documentation formats the way we do in \Hydras may introduce unnecessary overhead without providing additional benefits.
+%
+More specifically, we currently distribute the book as a PDF generated by LaTeX with links to corresponding HTML documentation of the Coq sources generated by \coqdoc.
+This means that in contrast to Software Foundations, we cannot easily obtain a uniform single-format representation of all documentation.
 

### 2021-11-22 doc/JFLA2022/paper.tex
small reformulation

@@
 
-The flip side of our approach is that we separate the Coq sources from the LaTeX sources that describe the source code, and our use of snippet delimiters in the Coq source code files may also lower source readability.
-For books and libraries with small scope and specialized audiences, distributing knowledge representations between sources and documentation formats the way we do in \Hydras may introduce unnecessary overhead without providing additional benefits.
+The flip side of our approach is that we separate the Coq sources from the LaTeX sources that describe the code, and our use of snippet delimiters in the Coq source files may also lower source readability.
+For books and libraries with small scope and specialized audiences, distributing knowledge representations between sources and documentation formats the way we do in \Hydras may introduce overhead without providing complementary benefits.
 %

### 2021-11-22 doc/JFLA2022/paper.tex
uniform typesetting of e.g., formulation improvements

@@
     \end{minipage}}
- \caption{$H'$ definition with the \equations
- plug-in~\cite{sozeau:hal-01671777}}
+ \caption{$H'$ definition using the \equations plug-in.}
 \label{fig:Hprime}
@@
 The original release of the Cantor library preceded support for type classes in \coq~\cite{MS08}.
-Changes over time before the move of \Hydras to \community have since then introduced type classes for some concepts in the Hydra-battles library related to ordinals.
-Most type classes initially used a \emph{semi bundled} approach~\cite{Mathlib}, but we have recently \emph{unbundled} classes to enable more sharing during resolution~\cite{BS2011}.
-For example, we use the (single-field) \texttt{Decision} type class~\cite{BS2011}, that describes deciding a proposition, in parameters to other classes rather than in fields.
+%
+Changes over time before the move of \Hydras to \community have since then introduced type classes for some concepts in the Hydra-battles library related to ordinals, e.g., to provide a uniform way to compare ordinals within different representation systems.
+%
+Most type classes initially used a \emph{semi bundled} approach~\cite{Mathlib}, but we have recently \emph{unbundled} classes to enable more sharing during resolution.
+For instance, we use the (single-field) \texttt{Decision} type class~\cite{BS2011}, that describes deciding a proposition, in parameters to other classes rather than in their fields.
 Due to the modest depth of our class hierarchy, we believe unbundling will not lead to scalability issues.
-Orthogonally to type classes, we introduced automation for definitions via the \equations~\cite{sozeau:hal-01671777} and Paramcoq~\cite{Keller2012} plug-ins.
+Orthogonally to type classes, we introduced automation for definitions using the Paramcoq~\cite{Keller2012} and \equations~\cite{sozeau:hal-01671777} plug-ins, as illustrated in Figure~\ref{fig:Hprime}.
 
@@
 
-\Hydras wants to provide a connection between scientific literature (\emph{e.g.},~\cite{KP82, KS81, schutte}) and proof assistant technology.
+\Hydras wants to provide a connection between scientific literature (e.g.,~\cite{KP82, KS81, schutte}) and proof assistant technology.
 For the mathematician, it can give a concrete view of the mathematical content, not only through full proofs, but also through illuminating computations: examples, functions associated with constructive proofs, etc.

### 2021-11-22 doc/JFLA2022/paper.tex
reformulation in comparison section

@@
 
-The flip side of our approach is that we separate the Coq sources from the LaTeX sources that describe the code, and our use of snippet delimiters in the Coq source files may also lower source readability.
+The flip side of our approach is that we separate the Coq sources from the LaTeX sources that describe the code, and our use of snippet delimiters in the Coq source files may lower source readability.
 For books and libraries with small scope and specialized audiences, distributing knowledge representations between sources and documentation formats the way we do in \Hydras may introduce overhead without providing complementary benefits.
 %
-More specifically, we currently distribute the book as a PDF generated by LaTeX with links to corresponding HTML documentation of the Coq sources generated by \coqdoc.
-This means that in contrast to Software Foundations, we cannot easily obtain a uniform single-format representation of all documentation.
+More specifically, we currently distribute the book as a PDF generated by LaTeX that links to corresponding HTML documentation of the Coq sources generated by \coqdoc.
+In contrast, executable books such as Software Foundations can provide a single uniform HTML representation of both book contents and code.
 

### 2021-11-23 doc/JFLA2022/paper.tex
Add reference to latest release in footnote.

@@
 On the one hand, we follow executable books in providing continuous delivery of new revisions at a high pace.
-But on the other hand, we also aim to make regular timestamped versions compatible with specific versions of Coq, akin to new editions of traditional books.
+But on the other hand, we also aim to make regular timestamped versions compatible with specific versions of Coq, akin to new editions of traditional books.\footnote{The latest Hydras \& Co.
+release is 0.5, available at \url{https://github.com/coq-community/hydra-battles/releases/tag/v0.5}.}
 
@@
 In contrast, executable books such as Software Foundations can provide a single uniform HTML representation of both book contents and code.
+We hope that future toolchain improvements can address all these issues.
 
@@
 %
-Moreover, \Hydras is not limited to the study of ordinal numbers and applications.
+Moreover, \Hydras is not limited to the study of ordinal numbers and their applications.
 We are also developing a package about efficient exponentiation

### 2021-11-24 doc/tuto-beamer.tex
experiments with beamer

--- 2021-11-20 doc/tuto.tex
@@
-\documentclass{article}
+\documentclass[10pt]{beamer}
 
@@
 \begin{document}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+  \maketitle
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+  \frametitle{Introduction}
+  \begin{block}{}
+We show a small but non-trivial example of using Alectryon
+for writing a documentation (in pdf) of a proof written in Coq.    
+  \end{block}
+
+  \begin{block}{}
+    \begin{quote}
+
+     `` If a   relation $<$ on a type $A$ is well founded,
+    then the lexicographic exponentiation of $<$  is well-founded too.''
+    \end{quote}
+
+ \end{block}
+\end{frame}
+
+\begin{frame}[fragile]
+
+  \inputsnippets{MultisetWf/tDef}
+
+Let us consider a relation over some type $A$. We define a lexicographic order on ($\texttt{t}\;A$).
+
+\begin{scriptsize}
+\inputsnippets{MultisetWf/AGiven, MultisetWf/lexpowerDef}  
+\end{scriptsize}
+
+\begin{footnotesize}
+\inputsnippets{MultisetWf/AGiven, MultisetWf/lexpowerDef}  
+\end{footnotesize}
+
+\end{frame}
+
+\begin{frame}[fragile]
+     The current goal is universally quantified over \texttt{n}.
+Thus we may prove it by well founded induction over \texttt{nat}.
+
+\begin{footnotesize}
+\inputsnippets{MultisetWf/LAccsc}
+  \end{footnotesize}
+\begin{footnotesize}
+  \inputsnippets{MultisetWf/LAccsd}
+\end{footnotesize}
+
+  Thanks to \texttt{Hl}, we know that \texttt{l} is accessible.
+Thus we can build an induction over \texttt{l}'s accessibility.
+
+\end{frame}
+\end{document}
 \maketitle
 
-\section{Introduction}
-We show a small but non-trivial example of using Alectryon
-for writing a documentation (in pdf) of a proof written in Coq.
-
-\section{Preliminaries}
-We prove that, if a type $A$ is well founded for a relation \textit{<},
-then the set of sorted (w.r.t. the relation $\geq$ ) sequences on $A$  is well-founded too.
 Such sequences are represented in a compact way as multisets over $A$. 

### 2021-12-08 doc/JFLA2022/Slides/Slides.tex
initial slides

--- 2021-11-24 doc/tuto-beamer.tex
@@
 \documentclass[10pt]{beamer}
+
+\definecolor{plugincolor}{rgb}{0.0,0.0,0.0}
+\definecolor{mathcolor}{rgb}{1.0,0.3,0.0}
+\definecolor{coqstylecolor}{rgb}{0.0,0.0,01.0}
+
+\definecolor{metavarcolor}{rgb}{0.5,0.0,1.0}
+\definecolor{darkgreen}{rgb}{0.1,0.7,0.1}
+\definecolor{answercolor}{rgb}{.08,.15,.8}
+\definecolor{normalcolor}{rgb}{0.0,0.0,0.0}
+\definecolor{exbluecolor}{rgb}{0.1,0.1,0.9}
+\definecolor{dontlookcolor}{rgb}{0.5,0.5,0.5}
+\definecolor{termcolor}{rgb}{0.0,0.1,0.9}
+\definecolor{lookcolor}{rgb}{0.9,0.1,0.0}
+\definecolor{prooftermcolor}{rgb}{0.3,0.1,1.0}
+\definecolor{typecolor}{rgb}{1.0,0.6,0.0}
+\definecolor{taccolor}{rgb}{0.70,0.10,0.0}
+\definecolor{pink}{rgb}{0.8,0.6,0.6}
+\definecolor{darkmagenta}{rgb}{0.4,0.0,0.6}
+\definecolor{darkblue}{rgb}{0.0,0.0,0.6}
 
@@
 %%% for movies by alectryon
-\usepackage{./movies/snippets/assets/alectryon}
-\usepackage{./movies/snippets/assets/pygments}
+\usepackage{./alectryon}
+\usepackage{../pygments}
 
@@
 
-\newcommand{\pathtomovies}{./movies}
+\newcommand{\pathtomovies}{..}%/movies}
 
@@
   {{\renewenvironment{alectryon}{}{}% Skip \begin{alectryon} included in snippet
-    \input@old{{\pathtomovies}/snippets/#1}}}
+    \input@old{{\pathtomovies}/#1}}}
 \makeatother
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title{Hydras \& Co.: Formalized mathematics in Coq\\
+ for inspiration and entertainment
+}
+\date{JFLA, February 2022}
+\author{
+Pierre Castéran %\inst{1}
+\and
+    Jérémy Damour %\inst{2}
+\and
+Karl Palmskog %\inst{3}
+\and Clément Pit-Claudel %\inst{4}
+\and Théo Zimmermann %\inst{5}
+}
 
-\title{Lexicographic exponentiation: An example of proof documentation with Alectryon}
-
-\author{Labri Coq working group}
+% \institute{
+% Univ. Bordeaux, CNRS, Bordeaux INP, LaBRI, UMR 5800, F-33400 Talence, France  %\\
+%  % \email{pierre.casteran@labri.fr}
+% \and
+% Univ. de Paris, F-75013 Paris, France
+% \and
+% KTH Royal Institute of Technology, Stockholm, Sweden
+% \and
+% MIT CSAIL, Cambridge, Massachusetts, USA
+% \and
+% Inria, Univ. de Paris, CNRS, IRIF, UMR 8243, F-75013 Paris, France
+% }
 
@@
 \end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+  \frametitle{Coq-community}
+  \begin{block}{Objectives}
+    \begin{itemize}
+    \item
+    \item
+      \item Documentation on libraries
+    \end{itemize}
+  \end{block}
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+  \frametitle{What hydras are talking about?}
+  \begin{block}{Hydra battles}
+    \begin{itemize}
+\item  \textcolor{mathcolor}{ordinals},
+    \textcolor{mathcolor}{rapidly growing functions},
+    \textcolor{mathcolor}{primitive recursive functions}, etc.
+    \item \textcolor{coqstylecolor}{dependently typed functions},
+      \textcolor{coqstylecolor}{proofs of well-foundedness},
+         \textcolor{coqstylecolor}{impossibility proofs},
+    \textcolor{coqstylecolor}{uniqueness of identity proofs (UIP)},
+      \textcolor{coqstylecolor}{indefinite description (Hilbert's epsilon operator)},
+    etc.
+    \item \textcolor{plugincolor}{ackermann (Russel O'Connor)},
+      \textcolor{plugincolor}{gaia (Jos\'e Grimm)},
+    \end{itemize}
+
+  \end{block}
+
+  \begin{block}{Addition chains}
+    \begin{itemize}
+    \item    \textcolor{mathcolor}{algorithmics},
+       \textcolor{mathcolor}{arithmetic},
+  \item 
+    \textcolor{coqstylecolor}{monadic notations},
+    \textcolor{coqstylecolor}{PHOAS},
+  \textcolor{coqstylecolor}{parametricity}, etc.
+    \item
+      \textcolor{plugincolor}{paramcoq (Chantal Keller, Marc Lasson)} 
+          \end{itemize}
+  \end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}[fragile]
+  \frametitle{Tests with Beamer(1)}
+  \begin{footnotesize}
+      \inputsnippets{Schutte_Ex42a}   
+      \inputsnippets{Schutte_Ex42b, Schutte_Ex42c}   
+    \end{footnotesize}
+\end{frame}
+  \begin{frame}[fragile]
+    \frametitle{Tests with Beamer(2)}
+
+  \begin{footnotesize}
+  \inputsnippets{Schutte_Ex42d, Schutte_Ex42e}   
+  \end{footnotesize}
+
+\end{frame}
+\end{document}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{footnotesize}
+\begin{small}
 \inputsnippets{MultisetWf/LAccsc}
-  \end{footnotesize}
+ \end{small}
 \begin{footnotesize}

### 2021-12-08 doc/JFLA2022/Slides/Slides.tex
add coq-equations
a few more slides

@@
 \documentclass[10pt]{beamer}
+
+\usepackage{xspace}
+
+\newcommand{\TODO}[2][]{[\textcolor{red}{TODO (#1):} \emph{#2}]}
+\newcommand{\coq}{Coq\xspace}
+\newcommand{\coqdoc}{Coqdoc\xspace}
+\newcommand{\coqmakefile}{\texttt{coq\_makefile}\xspace}
+\newcommand{\community}{Coq-community\xspace}
+\newcommand{\gaia}{Gaia\xspace}
+\newcommand{\alectr}{Alectryon\xspace}
+\newcommand{\equations}{Equations\xspace}
+\newcommand{\Hydras}{Hydras \& Co$\text.$\xspace}
+\newcommand{\make}{\texttt{make}\xspace}
 
@@
 \definecolor{darkblue}{rgb}{0.0,0.0,0.6}
+
+\usepackage{ifpdf}
+\ifpdf
+\usepackage{graphicx}
+%\usepackage{aeguill}
+\else
+\usepackage[dvips]{graphicx}
+\fi
 
@@
     \begin{itemize}
-    \item
-    \item
+    \item \TODO{}{}
+    \item\TODO{}{}
       \item Documentation on libraries
@@
 \end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+  \frametitle{Yet another book on \coq?}
+  \begin{block}{Kinds of books}
+    \TODO{}{}
+  \end{block}
+
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+  \frametitle{Example-driven books}
+  \begin{block}{Tsukuba Coq User's Group}
+    Regular languages, lambda-calculus, programming language semantics, \dots
+
+   \includegraphics[height=35mm]{tsukubabooks.jpg}
+   %\includegraphics[height=45mm]{leaflet.jpg}
+     \end{block}
+     \begin{block}{\Hydras}
+       Kirby \& Paris'  hydra battles, exponentiation algorithms, \dots?
+     \end{block}
+
+\end{frame}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \frametitle{What hydras are talking about?}
+  \begin{block}{}
+    \Hydras presents a consistent set of examples, which allow us to formalize 
+    \textcolor{mathcolor}{Discrete math}, discuss
+    \textcolor{coqstylecolor}{\coq style, specification and proof patterns}, and use  \textcolor{plugincolor}{a few plug-ins}.
+  \end{block}
   \begin{block}{Hydra battles}
@@
     etc.
-    \item \textcolor{plugincolor}{ackermann (Russel O'Connor)},
-      \textcolor{plugincolor}{gaia (Jos\'e Grimm)},
+  \item
+ \textcolor{plugincolor}{equations (Matthieu Sozeau)},
+    \textcolor{plugincolor}{ackermann (Russel O'Connor)},
+      \textcolor{plugincolor}{gaia (Jos\'e Grimm)}
     \end{itemize}
-
   \end{block}

### 2021-12-08 doc/JFLA2022/Slides/Slides.tex
a few more slides

@@
 
-\definecolor{plugincolor}{rgb}{0.0,0.0,0.0}
+\definecolor{plugincolor}{rgb}{0.6,0.0,0.6}
 \definecolor{mathcolor}{rgb}{1.0,0.3,0.0}
@@
 \author{
-Pierre Castéran %\inst{1}
+Pierre Castéran \inst{1}
 \and
-    Jérémy Damour %\inst{2}
+    Jérémy Damour \inst{2}
 \and
-Karl Palmskog %\inst{3}
-\and Clément Pit-Claudel %\inst{4}
-\and Théo Zimmermann %\inst{5}
+Karl Palmskog \inst{3}
+\and Clément Pit-Claudel \inst{4}
+\and Théo Zimmermann \inst{5}
 }
 
-% \institute{
-% Univ. Bordeaux, CNRS, Bordeaux INP, LaBRI, UMR 5800, F-33400 Talence, France  %\\
-%  % \email{pierre.casteran@labri.fr}
-% \and
-% Univ. de Paris, F-75013 Paris, France
-% \and
-% KTH Royal Institute of Technology, Stockholm, Sweden
-% \and
-% MIT CSAIL, Cambridge, Massachusetts, USA
-% \and
-% Inria, Univ. de Paris, CNRS, IRIF, UMR 8243, F-75013 Paris, France
-% }
+\institute{
+Univ. Bordeaux, CNRS, Bordeaux INP, LaBRI, UMR 5800, F-33400 Talence, France  %\\
+ % \email{pierre.casteran@labri.fr}
+\and
+Univ. de Paris, F-75013 Paris, France
+\and
+KTH Royal Institute of Technology, Stockholm, Sweden
+\and
+MIT CSAIL, Cambridge, Massachusetts, USA
+\and
+Inria, Univ. de Paris, CNRS, IRIF, UMR 8243, F-75013 Paris, France
+}
 
@@
   \end{block}
-
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Example-driven books}
+  \frametitle{Example-driven books (1)}
+  \begin{block}{}
+    \begin{itemize}
+    \item \TODO{}{Definition}
+    \item No need to be complete (there is already a reference manual)
+      \item No need to introduce \coq (there are already a lot of initiation books and tutorials)
+    \item \coq patterns (formalization and proof) are presented  when they are needed.
+    \item Allow to compare several different formalizations of a given notion, and several proofs of the same theorem.
+    \item By definition, the authors choose topics they like to teach.
+      \item \TODO{}{topics adapted to the potential readers?}
+    \end{itemize}
+  \end{block}
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+  \frametitle{Example-driven books (2)}
   \begin{block}{Tsukuba Coq User's Group}
@@
      \begin{block}{\Hydras}
-       Kirby \& Paris'  hydra battles, exponentiation algorithms, \dots?
+       Kirby \& Paris'  hydra battles, exponentiation algorithms, \dots
      \end{block}
@@
     \Hydras presents a consistent set of examples, which allow us to formalize 
-    \textcolor{mathcolor}{Discrete math}, discuss
-    \textcolor{coqstylecolor}{\coq style, specification and proof patterns}, and use  \textcolor{plugincolor}{a few plug-ins}.
+    some \textcolor{mathcolor}{discrete math}, discuss
+    \textcolor{coqstylecolor}{\coq style, specification and proof patterns}, and use\footnote{and depend on!} \textcolor{plugincolor}{a few plug-ins}.
   \end{block}
@@
 \begin{frame}[fragile]
-  \frametitle{Tests with Beamer(1)}
+  \frametitle{Tests with Alectryon(1)}
   \begin{footnotesize}
@@
   \end{frame}
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   \begin{frame}[fragile]
-    \frametitle{Tests with Beamer(2)}
+    \frametitle{Tests with Alectryon(2)}
 

### 2021-12-08 doc/JFLA2022/Slides/Slides.tex
a few more slides

@@
 \definecolor{darkblue}{rgb}{0.0,0.0,0.6}
+
+\usepackage{tikz}
+\usepackage{tikzsymbols}
+\usepackage{pifont}
+\usetikzlibrary{arrows}
 
@@
 \end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+  \TODO{}{Chose an example: for instance the length of hydra-battles: how we use \texttt{equations}, \texttt{ackermann}.
+    Connexion with gaia (or potentially Minki's ordinals)
+    (several slides).}
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
 
+  \begin{figure}[h]
+  \centering
+  \begin{tikzpicture}[very thick, scale=0.3]
+  \node (h1) at (4,0){$\bullet$};
+  \node[blue] (h2) at (4,2){$\bullet$};
+  \node[blue] (h3) at (2,4){$\bullet$};
+  \node[blue] (h4) at (0,6){$\bullet$};
+  \node[red!90!black] (h7) at (6,2){$\bullet$};
+  \node[blue] (h5) at (4,6){$\bullet$};
+  \node[blue] (h6) at (6,4){$\bullet$};
+  \draw (h1) -- (h2) ;
+  \draw (h2) -- (h3) ;
+  \draw (h1) -- node[red!90!black,font=\small,sloped,shift={(0.01,-0.075)},rotate=90]{\textbf{\ding{34}}} (h7);
+  \draw (h3) -- (h4) ;
+  \draw (h3) -- (h5) ;
+  \draw (h2) -- (h6);
+
+  \node (h11) at (14,0){$\bullet$};
+  \node[blue] (h12) at (14,2){$\bullet$};
+  \node[blue] (h13) at (12,4){$\bullet$};
+  \node[blue] (h14) at (10,6){$\bullet$};
+  \node[blue] (h15) at (14,6){$\bullet$};
+  \node[red!90!black] (h16) at (16,4){$\bullet$};
+  \draw (h11) -- (h12) ;
+  \draw (h12) -- (h13) ;
+  \draw (h12) -- node[red!90!black,font=\small,sloped,shift={(0.01,-0.075)},rotate=90]{\textbf{\ding{34}}} (h16);
+  \draw (h13) -- (h14) ;
+  \draw (h13) -- (h15) ;
+
+\node (hn1) at (24,0){$\bullet$};
+\node[blue] (hn2) at (22,2) {$\bullet$};
+\node[blue] (hn3) at (22,4) {$\bullet$};
+\node[blue] (hn4) at (20,6){$\bullet$};
+  \node[blue] (hn5) at (22,6){$\bullet$};
+  \draw (hn1) -- (hn2) ;
+  \draw (hn2) -- (hn3) ;
+  \draw (hn3) -- (hn4) ;
+  \draw (hn3) -- (hn5) ;
+\node [blue](hn2b) at (24,2) {$\bullet$};
+\node[blue] (hn3b) at (24,4) {$\bullet$};
+\node [blue](hn4b) at (23.4,6){$\bullet$};
+  \node [blue](hn5b) at (24.6,6){$\bullet$};
+  \draw (hn1) -- (hn2b) ;
+  \draw (hn2b) -- (hn3b) ;
+  \draw (hn3b) -- (hn4b) ;
+  \draw (hn3b) -- (hn5b) ;
+  \node [blue](hn2c) at (26,2) {$\bullet$};
+\node [blue](hn3c) at (26,4) {$\bullet$};
+\node [blue](hn4c) at (26,6){$\bullet$};
+  \node [red!90!black] (hn5c) at (28,6){$\bullet$};
+  \draw (hn1) -- (hn2c) ;
+  \draw (hn2c) -- (hn3c) ;
+  \draw (hn3c) -- (hn4c) ;
+  \draw (hn3c) -- node[red!90!black,font=\small,sloped,shift={(0.01,-0.075)},rotate=90]{\textbf{\ding{34}}} (hn5c) ;
+\end{tikzpicture}
+
+\caption{Three successive states of a hydra in a battle,
+at time $t=0,1,2$}
+  \label{fig:round}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+
+\begin{tikzpicture}[very thick, scale=0.3]
+\node (hn1) at (4,0){$\bullet$};
+\node[blue] (hn2) at (2,2) {$\bullet$};
+\node[blue] (hn3) at (2,4) {$\bullet$};
+\node[blue] (hn4) at (0,6){$\bullet$};
+  \node[blue] (hn5) at (2,6){$\bullet$};
+  \draw (hn1) -- (hn2) ;
+  \draw (hn2) -- (hn3) ;
+  \draw (hn3) -- (hn4) ;
+  \draw (hn3) -- (hn5) ;
+\node [blue](hn2b) at (4,2) {$\bullet$};
+\node [blue](hn3b) at (4,4) {$\bullet$};
+\node [blue](hn4b) at (3.4,6){$\bullet$};
+  \node [blue] (hn5b) at (4.6,6){$\bullet$};
+  \draw (hn1) -- (hn2b) ;
+  \draw (hn2b) -- (hn3b) ;
+  \draw (hn3b) -- (hn4b) ;
+  \draw (hn3b) -- (hn5b) ;
+  \node [blue](hn2c) at (6,2) {$\bullet$};
+\node [blue](hn3c) at (6,4) {$\bullet$};
+\node [blue](hn4c) at (6,6){$\bullet$};
+ % \node (hn5c) at (8,6){$\bullet$};
+  \draw (hn1) -- (hn2c) ;
+  \draw (hn2c) -- (hn3c) ;
+  \draw (hn3c) -- (hn4c) ;
+  % \draw (hn3c) -- (hn5c) ;
+  \node [blue](hn3d) at (8,4) {$\bullet$};
+  \draw (hn2c) -- (hn3d) ;
+  \node [blue](hn4d) at (8,6) {$\bullet$};
+  \draw (hn3d) -- (hn4d) ;
+
+ \node [blue](hn3e) at (10,4) {$\bullet$};
+  \draw (hn2c) -- (hn3e) ;
+  \node [blue](hn4e) at (10,6) {$\bullet$};
+  \draw (hn3e) -- (hn4e) ;
+
+  \node [blue](hn3f) at (12,4) {$\bullet$};
+  \draw (hn2c) -- (hn3f) ;
+  \node [red!90!black](hn4f) at (12,6) {$\bullet$};
+  \draw (hn3f) -- node[red!90!black,font=\small,sloped,shift={(0.01,-0.075)},rotate=90]{\textbf{\ding{34}}} (hn4f) ;
+
+  \node at (16,0) {$\dots$} ;
+
+ \node [black](hn4h) at (21,0) {$\bullet$};
+\node [red!90!black](hn4g) at (21,2) {$\bullet$};
+   \draw (hn4h) -- node[red!90!black,font=\small,sloped,shift={(0.01,-0.075)},rotate=90]{\textbf{\ding{34}}} (hn4g) ;
+
+  \node [black](hn4f) at (24,0) {$\bullet$};
+\end{tikzpicture}
+ \caption{At $t=3,\dots,x^?-1,x^?$}
+\end{figure}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+  \TODO{}{maintenance, CI/CD,  etc.   (several slides).}
+\end{frame}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+  \TODO{}{Alectryon (several slides)}
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Old stuff
+
 \begin{frame}

### 2021-12-12 doc/JFLA2022/Slides/Slides.tex
a few more slides
draft

@@
 \documentclass[10pt]{beamer}
+\usetheme{Montpellier}
 
-\usepackage{xspace}
+\usepackage{xspace, alltt}
 
@@
 \definecolor{plugincolor}{rgb}{0.6,0.0,0.6}
-\definecolor{mathcolor}{rgb}{1.0,0.3,0.0}
+\definecolor{mathcolor}{rgb}{0.0,0.0,0.6}
 \definecolor{coqstylecolor}{rgb}{0.0,0.0,01.0}
-
+\definecolor{lightblue}{rgb}{0.2,0.2,1.0}
 \definecolor{metavarcolor}{rgb}{0.5,0.0,1.0}
@@
 %%% between them).  A, B, C should be paths to files in snippets/.
+
 \usepackage{etoolbox}
@@
 
-\newcommand{\pathtomovies}{..}%/movies}
+\newcommand{\pathtomovies}{.}%/movies}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%% Specific macros
+
+\newcommand{\canonseq}[2]{\mbox{$\{#1\}(#2)$}}
 \title{Hydras \& Co.: Formalized mathematics in Coq\\
@@
 }
+
+\usepackage{draftwatermark} 
+\setbeamercolor{background canvas}{bg=}%transparent canvas
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-  \frametitle{Example-driven books (2)}
-  \begin{block}{Tsukuba Coq User's Group}
-    Regular languages, lambda-calculus, programming language semantics, \dots
+% \begin{frame}
+%   \frametitle{Example-driven books (2)}
+%   \begin{block}{Tsukuba Coq User's Group}
+%     Regular languages, lambda-calculus, programming language semantics, \dots
 
-   \includegraphics[height=35mm]{tsukubabooks.jpg}
-   %\includegraphics[height=45mm]{leaflet.jpg}
-     \end{block}
-     \begin{block}{\Hydras}
-       Kirby \& Paris'  hydra battles, exponentiation algorithms, \dots
-     \end{block}
+%    \includegraphics[height=35mm]{tsukubabooks.jpg}
+%    %\includegraphics[height=45mm]{leaflet.jpg}
+%      \end{block}
+%      \begin{block}{\Hydras}
+%        Kirby \& Paris'  hydra battles, exponentiation algorithms, \dots
+%      \end{block}
 
-\end{frame}
+% \end{frame}
 
@@
   \end{block}
-  \begin{block}{Hydra battles}
-    \begin{itemize}
-\item  \textcolor{mathcolor}{ordinals},
+  \begin{block}{}
+    \begin{description}
+    \item[hydra battles:]
+
+  \textcolor{mathcolor}{ordinals},
     \textcolor{mathcolor}{rapidly growing functions},
-    \textcolor{mathcolor}{primitive recursive functions}, etc.
-    \item \textcolor{coqstylecolor}{dependently typed functions},
+    \textcolor{mathcolor}{primitive recursive functions}, 
+   \textcolor{coqstylecolor}{dependently typed functions},
+    \textcolor{coqstylecolor}{operational type classes},
       \textcolor{coqstylecolor}{proofs of well-foundedness},
@@
       \textcolor{coqstylecolor}{indefinite description (Hilbert's epsilon operator)}, 
-    etc.
-  \item
- \textcolor{plugincolor}{equations (Matthieu Sozeau)},
-    \textcolor{plugincolor}{ackermann (Russel O'Connor)},
-      \textcolor{plugincolor}{gaia (Jos\'e Grimm)}
-    \end{itemize}
-  \end{block}
+     \textcolor{plugincolor}{equations},
+    \textcolor{plugincolor}{ackermann (goedel)},
+      \textcolor{plugincolor}{gaia}
 
-  \begin{block}{Addition chains}
-    \begin{itemize}
-    \item    \textcolor{mathcolor}{algorithmics},
+ \item[addition chains (exponentiation algorithms):]
+      \textcolor{mathcolor}{algorithmics},
        \textcolor{mathcolor}{arithmetic},
-  \item 
+            \textcolor{coqstylecolor}{operational type classes},
     \textcolor{coqstylecolor}{monadic notations},
     \textcolor{coqstylecolor}{PHOAS},
-  \textcolor{coqstylecolor}{parametricity}, etc.
-    \item
-      \textcolor{plugincolor}{paramcoq (Chantal Keller, Marc Lasson)} 
-          \end{itemize}
+  \textcolor{coqstylecolor}{parametricity}, 
+        \textcolor{plugincolor}{paramcoq}
+        \end{description}
   \end{block}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Example: Kirby \& Paris' hydras}
 \begin{frame}
-  \TODO{}{Chose an example: for instance the length of hydra-battles: how we use \texttt{equations}, \texttt{ackermann}.
-    Connexion with gaia (or potentially Minki's ordinals)
-    (several slides).}
-\end{frame}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
+  \frametitle{Example: Kirby \& Paris' hydras}
+ %  \begin{block}{}
+
+%   \begin{enumerate}
+%   \item Intuitive presentation
+%     \item From discrete math to \coq
+%   \end{enumerate}
+%   \end{block}
+
+% \end{frame}
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \begin{frame}
+%   \frametitle{The life of $\omega^{\omega^2+1}+1$}
 
@@
   \centering
-  \begin{tikzpicture}[very thick, scale=0.3]
+  \begin{tikzpicture}[very thick, scale=0.25]
   \node (h1) at (4,0){$\bullet$};
@@
 \caption{Three successive states of a hydra in a battle,
-at time $t=0,1,2$}
+at time $t=0,1,2$\label{life1}}
   \label{fig:round}
@@
 
-\begin{tikzpicture}[very thick, scale=0.3]
+\begin{tikzpicture}[very thick, scale=0.25]
 \node (hn1) at (4,0){$\bullet$};
@@
 \end{tikzpicture}
- \caption{At $t=3,\dots,x^?-1,x^?$}
+\caption{At $t=3,\dots,\textcolor{red}{x}-1,\textcolor{red}{x=?}$
+\label{life2}}
 \end{figure}
+\end{frame}
 
+\begin{frame}
+  \frametitle{Mathematical formalization of hydra battles}
+   \begin{block}{}
+
+    \begin{itemize}
+    \item   We associate a hydra to each ordinal \textcolor{mathcolor}{$\alpha<\epsilon_0$} (in Cantor normal form).
+For instance, the first hydra of our example is the image of \textcolor{mathcolor}{$omega^{\omega^2+1}+1$}.
+    \item A round (at time $t$) of a battle transforms an hydra associated with \textcolor{mathcolor}{$\alpha$} into the hydra associated with \textcolor{mathcolor}{$\canonseq{\alpha}{t+1}$}, the $t+1$-st element of the \emph{canonical sequence} of \textcolor{mathcolor}{$\alpha$}.
+    \item We want to ``compute'' the length of sequences of the form 
+\textcolor{mathcolor}{$\langle \alpha_k=\alpha,\dots,\alpha_{i+1}=\canonseq{\alpha_i}{i+1},\dots, 1,0\rangle$}, where \textcolor{mathcolor}{$k\in\mathbb{N}$}.
+
+  \end{itemize}
+\end{block}
 \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+
+ \begin{block}{}
+    Ordinals less than $\epsilon_0$ are represented as \emph{ordinal notations} which allow the user to write proofs by computation (comparison, canonical sequences, etc.)
+    \begin{footnotesize}
+      \inputsnippets{../../movies/snippets/ON_Generic/ONDef}
+
+       \inputsnippets{../E0_Ex42}
+    \end{footnotesize}
+
+  \end{block}
+\end{frame}
+
+\begin{frame}
+  \begin{block}{}
+    \begin{itemize}
+    \item We borrow the systematic study of the sequencesfrom Ketonen \& Solovay's article
+      \emph{Rapidly Growing {R}amsey Functions (1981)}.
+
+    \item Let \textcolor{mathcolor}{$0<\alpha<\epsilon_0$}, and \textcolor{mathcolor}{$i\in\mathbb{N}$}.
+      The least index \textcolor{mathcolor}{$k$} such that \textcolor{mathcolor}{$\alpha_k=0$} is  \textcolor{mathcolor}{$H'_\alpha(i)$} where \textcolor{mathcolor}{$H'_\alpha$} is defined by transfinite induction over \textcolor{mathcolor}{$\alpha$}.
+
+   %    \begin{align}
+%   H'_0(i) & = i\\
+%   H'_\alpha(i) &= H'_{(\canonseq{\alpha}{i+1})}(i)  \quad\textit{if $\alpha$ is a limit ordinal}\\
+%   H'_{\alpha}(i) &=H'_\beta(i+1) \quad\textit{if $\alpha=\beta+1$}
+% \end{align}
+\vspace{5pt}
+      \begin{footnotesize}
+        \inputsnippets{../Hprime_HprimeDef}   
+      \end{footnotesize}
+
+    \end{itemize}
+
+  \end{block}
+
+  \begin{block}{}
+    The \textcolor{mathcolor}{$H'_\alpha$} family is an example of \emph{hierarchy of rapidly growing functions}.
+For instance, \textcolor{mathcolor}{$H'_{\omega^3+3}(0)\geq2^{2^N}$}, where \textcolor{mathcolor}{$N=2^{70}-1$}.
+  \end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+  \begin{block}{}
+
+    Properties of the \textcolor{mathcolor}{$H'_\alpha$}s are proved with the help of the lemmas automatically generated by \textcolor{plugincolor}{equations}.
+For instance, the \textcolor{mathcolor}{$H'_\alpha$} are strictly monotonous,
+and if \textcolor{mathcolor}{$\beta<\alpha$}, then \textcolor{mathcolor}{$H'_\alpha(i)>H'_\beta(i)$} for all but a finite number of $i$.
+  \end{block}
+
+  \begin{block}{}
+    We prove also that, for any   \textcolor{mathcolor}{$\alpha\geq\omega^\omega$} the function \textcolor{mathcolor}{$H'_\alpha$} is not primitive recursive.
+
+    \begin{itemize}
+    \item \emph{In order to prove this theorem, we chose to maintain under \community Russel O'Connor's library \textcolor{plugincolor}{goedel}, which contains a formalization of primitive recursive functions.}
+    \item \emph{Proving formally that a given function is not primitive recursive is a nice example of application of a complex (mutually recursive, dependently typed) induction principle.
+You have to take into account the set of PR functions of any arity.}
+
+        \item    \emph{The interest of O'Connor's contribution led us to write a chapter on primitive recursive functions, with examples and exercices.}
+    \end{itemize}
+  \end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+  \frametitle{Some features of the \Hydras project}
+  \begin{block}{}
+    \begin{itemize}
+    \item The order in which \coq code is presented in the book is determined by the difficulty of the mathematical examples, or the assumed level in \coq of the reader, not by the structure of the libraries.
+    \item We often present and compare several different definitions and proofs of the same mathematical notions, including ``bad'' ones.
+      \item We would be happy to include new proofs, if they improve old ones, or they illustrate new proof patterns. \emph{Please send PRs or issues!}.
+    \end{itemize}
+  \end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \begin{frame}
-  \TODO{}{Alectryon (several slides)}
+  \frametitle{Presenting proofs with \alectr}
+  \begin{small}
+
+ %\begin{footnotesize}
+  \inputsnippets{../Schutte_Ex42a}   
+ %\end{footnotesize}
+
+By definition of additive principal ordinals, 
+it suffices to prove the inequality $\omega+42< \phi_0(2)$.
+%\begin{footnotesize}
+\inputsnippets{../Schutte_Ex42b}
+By Lemma \textit{AP\_plus\_closed}, it suffices to prove  $\omega<\omega^2$ and $42<\omega^2$.
+%\end{footnotesize}
+\end{small}
+
+\end{frame}
+
+\begin{frame}
+  %\frametitle{Presenting proofs with \alectr (2)}
+\begin{small}
+%\begin{footnotesize}
+\inputsnippets{
+  ../Schutte_Ex42c, ../Schutte_Ex42d, ../Schutte_Ex42e}
+%\end{footnotesize}
+  \end{small}
+
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}[fragile]
+ \begin{alltt}
+{\color{lightblue}{(* begin snippet Ex42a:: no-out *)}}
+Example Ex42: omega + 42 + omega^2 = omega^2. 
+{\color{lightblue}{(* end snippet Ex42a *)}}
+Proof.
+  {\color{lightblue}{(* begin snippet Ex42b *)}}
+  assert (HAP:= AP_phi0 2). {\color{lightblue}{(* .no-out *)}}
+  elim  HAP; intros _ H0; apply H0; clear H0. 
+  {\color{lightblue}{(* end snippet Ex42b *)}}
+  {\color{lightblue}{(* begin snippet Ex42d *)}}
+  Check AP_plus_closed. {\color{lightblue}{(* .unfold .no-goals *)}}
+  {\color{lightblue}{(* end snippet Ex42d *)}}
+   {\color{lightblue}{(* begin snippet Ex42c *)}}
+  assert (Hlt: omega < omega^2) by 
+      (rewrite omega_eqn; apply phi0_mono, finite_mono;
+        auto with arith).
+ {\color{lightblue}{ (* end  snippet Ex42c *)}}
+\end{alltt}
 \end{frame}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}[fragile]
-  \frametitle{Tests with Alectryon(1)}
-  \begin{footnotesize}
-      \inputsnippets{Schutte_Ex42a}   
-      \inputsnippets{Schutte_Ex42b, Schutte_Ex42c}   
-    \end{footnotesize}
-  \end{frame}
-  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \begin{frame}[fragile]
-    \frametitle{Tests with Alectryon(2)}
-
-  \begin{footnotesize}
-  \inputsnippets{Schutte_Ex42d, Schutte_Ex42e}   
-  \end{footnotesize}
-
-\end{frame}
-\end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Old stuff
-
 \begin{frame}
-  \frametitle{Introduction}
-  \begin{block}{}
-We show a small but non-trivial example of using Alectryon
-for writing a documentation (in pdf) of a proof written in Coq.    
-  \end{block}
-
-  \begin{block}{}
-    \begin{quote}
-
-     `` If a   relation $<$ on a type $A$ is well founded,
-    then the lexicographic exponentiation of $<$  is well-founded too.''
-    \end{quote}
-
+  \frametitle{Perspectives}
+  \begin{block}{Library on ordinals}
+    \begin{itemize}
+    \item \TODO{}{Gaia}
+      \item \TODO{}{Aczel, Minki}
+    \end{itemize}
   \end{block}
@@
 
-\begin{frame}[fragile]
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+  \frametitle{Perspectives(2)}
+  \begin{block}{Documentation with Alectryon}
+    \begin{itemize}
+    \item \TODO{}{}
+    \end{itemize}
+  \end{block}
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+  \frametitle{Perspectives(3)}
+  \begin{block}{CI/CD}
+     \TODO{}{}
+  \end{block}
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-  \inputsnippets{MultisetWf/tDef}
-
-Let us consider a relation over some type $A$. We define a lexicographic order on ($\texttt{t}\;A$).
-
-\begin{scriptsize}
-\inputsnippets{MultisetWf/AGiven, MultisetWf/lexpowerDef}  
-\end{scriptsize}
-
-\begin{footnotesize}
-\inputsnippets{MultisetWf/AGiven, MultisetWf/lexpowerDef}  
-\end{footnotesize}
+\begin{frame}
+  \frametitle{Conclusion}
+  \begin{block}{}
+    \begin{itemize}
+    \item \TODO{}{}
+    \item \Hydras and \community are on github. Please contribute!
+      \begin{itemize}
+    \item New chapters
+    \item Any improvement (text and/or proofs)
+    \end{itemize} 
+    \end{itemize}
+ \end{block}
 
@@
 
-\begin{frame}[fragile]
-     The current goal is universally quantified over \texttt{n}.
-Thus we may prove it by well founded induction over \texttt{nat}.
-
-\begin{small}
-\inputsnippets{MultisetWf/LAccsc}
- \end{small}
-\begin{footnotesize}
-  \inputsnippets{MultisetWf/LAccsd}
-\end{footnotesize}
-
-  Thanks to \texttt{Hl}, we know that \texttt{l} is accessible.
-Thus we can build an induction over \texttt{l}'s accessibility.
-
-\end{frame}
 \end{document}
-\maketitle
-
-Such sequences are represented in a compact way as multisets over $A$. 
-For instance, $\langle a,a,a,b,c,c,c,c,c \rangle $ where $a>b>c$ is represented by the list \texttt{(a,2)::(b,0)::(c,4)::nil}.
-
-Note that a factor $(a,n)$ is interpreted as $n+1$ copies of $a$.
-
-\inputsnippets{MultisetWf/tDef}
-
-Let us consider a relation over some type $A$. We define a lexicographic order on ($\texttt{t}\;A$).
-
-\inputsnippets{MultisetWf/AGiven, MultisetWf/lexpowerDef}
-
-\section{\texttt{lexpower}  is not well-founded in general}
-
-Let us build an infinite strictly decreasing (w.r.t. \texttt{lexpower lt}) sequence.
-
-\inputsnippets{MultisetWf/notWfa}
-
-Then, we prove that any element of the sequence \texttt{seq} is non-accessible.
-
-\inputsnippets{MultisetWf/notWfb}
-
-It's clearly a contradiction, since (\texttt{seq 0}) is an element of the sequence, and accessible (because of the hypothesis \texttt{Hwf}).
-
-\inputsnippets{MultisetWf/notWfc}
-\section{Lists in normal form}
-
-We say that a list $(a_1,n_1)::(a_2,n_2)::\dots::(a_p,n_p)::nil$  in  (\texttt{t A}) is in \emph{normal form} if  for any $1<=i<n$,
-we have $\texttt{ltA}\; a_{i+1}\; a_i$.
-
-\inputsnippets{MultisetWf/lexnfDef}
-
-The \emph{lexicographic exponentiation} of \texttt{ltA} is the restriction
-of \texttt{lexpower ltA} to the set of lists in normal form.
-In the next section, we prove that, if \texttt{ltA} is well-founded, then its lexicographic power is well-founded too.
-\inputsnippets{MultisetWf/lexltDef}
-
-\paragraph{Note:} The restriction of a binary relation to a set is defined in the file \texttt{Restriction.v}.
-
-\inputsnippets{Restriction/restrictionDef}
-
-\section{Proof of well-foundedness}
-It's time to prove that the lexicographic power of a well-founded relation is well-founded too. Let us open a work section.
-
-\inputsnippets{MultisetWf/bigProofa}
-
-We want to prove the following statement:
-
-\inputsnippets{MultisetWf/theStatement}
-
-\subsection{A first attempt}
-We may try to prove \texttt{lexwf} by (structural) induction on \texttt{l}.
-
-\inputsnippets{MultisetWf/BadProof,
-  MultisetWf/BadProofb}
-
-By the hypothesis \texttt{H0}, the list \texttt{y} may be
-equal to \texttt{(a,n)::l'}, where \texttt{LT l' l}. But the induction hypothesis \texttt{IHl} is useless for solving the current goal.
-\inputsnippets{MultisetWf/BadProofc}
-
-This is a classical issue. In a naive proof by induction, some variables are fixed too early: here, the variable \texttt{l} in \texttt{IHl}.
-
-\subsection{Introducing a new predicate}
-
-We can express the well-foundedness of \texttt{LT} in the following way :
-\begin{quote}
-  \begin{enumerate}
-    \item  ``The empty list is accessible''
-  \item   `` For any $a$ in $A$ and $n$ in \texttt{nat},  any well-formed list starting with $(a,n)$ is accessible. ''
-
-  \end{enumerate}
- \end{quote}
-
-The first property is easy to prove by inversion.
-
-\inputsnippets{MultisetWf/AccNil}
-
-For the second one, we define a new predicate over $A$.
-
-\inputsnippets{MultisetWf/AccsDef}
-
-The new goal is to prove that any element of $A$ satisfies
-the predicate \texttt{Accs}.
-
-\subsection{Inversion lemmas}
-
-Before starting the proof, we  prove three useful inversion lemmas:
-
-\inputsnippets{MultisetWf/NFInv1}
-\inputsnippets{MultisetWf/NFInv2}
-\inputsnippets{MultisetWf/LTInv}
-
-\subsection{The proof by induction}
-Le us prove, by induction on $a:A$ that every element of $A$ satisfies \texttt{Accs}.
-
-\inputsnippets{MultisetWf/LAccsa}
-
-Let us note that, for any list \texttt{(a,n)::l} in normal form,
-the tail \texttt{l} is either empty or starts with some pair \texttt{(b,p)} where \texttt{ltA b a}, hence is accessible (thanks to \texttt{IHa}).
-
-\inputsnippets{MultisetWf/LAccsb}
-\inputsnippets{MultisetWf/LAccsg}
-
-The current goal is universally quantified over \texttt{n}.
-Thus we may prove it by well founded induction over \texttt{nat}.
-
-\inputsnippets{MultisetWf/LAccsc, MultisetWf/LAccsd}
-
-Thanks to \texttt{Hl}, we know that \texttt{l} is accessible.
-Thus we can build an induction over \texttt{l}'s accessibility.
-
-\inputsnippets{MultisetWf/LAccse}
-
-In order to prove the accessibility of the list \texttt{((a, n) :: x0)}, we have to prove that any predecessor $y$ of this list is accessible, and consider every case (thanks to \texttt{LT\_inv}).
-
-\inputsnippets{MultisetWf/LAccsf}
-
-\subsubsection{First case}
-If \texttt{y=nil}, then \texttt{y} is clearly accessible.
-
-\inputsnippets{MultisetWf/case1}
-
-\subsubsection{Second case}
-Let us consider a well formed list \texttt{y} of the form \texttt{(b,p)::l''}, where \texttt{ltA b a}. By \texttt{IHa}, this list is accessible.
-
-\inputsnippets{MultisetWf/case2}
-
-\subsubsection{Third case}
-The list \texttt{y} may be of the form  \texttt{(a,n)::l''}, where \texttt{LT l'' l}. By \texttt{H2},  \texttt{y} is accessible.
-
-\inputsnippets{MultisetWf/case3}
-
-\subsubsection{Last case}
-The list \texttt{y} may be of the form  \texttt{(a,p)::l''}, where \texttt{p<n}. Thus, we may apply \texttt{Hn}.
-
-\inputsnippets{MultisetWf/case4}
-
-\subsection{At last!}
-
-\inputsnippets{MultisetWf/NFAcc}
-
-\inputsnippets{MultisetWf/lexwf}
-
-% \section{Examples}
-
-% \inputsnippets{MultisetWf/Examples}
-
-\section{Impossibility proofs}
-There is no function of type \texttt{t nat ->  nat} which could serve as a measure for proving \texttt{lexlt lt}'s well-foundedness.
-
-\inputsnippets{MultisetWf/Impossibility1}
-\inputsnippets{MultisetWf/Impossibility1a}
-\inputsnippets{MultisetWf/Impossibility1b}
-
-\emph{Same exercise with \texttt{nat*nat} (with the lexicographic square of \texttt{lt})}.
-
-\end{document}

### 2021-12-21 doc/JFLA2022/Slides/Slides.tex
JFLA Slides (draft)

@@
 \definecolor{termcolor}{rgb}{0.0,0.1,0.9}
-\definecolor{lookcolor}{rgb}{0.9,0.1,0.0}
+\definecolor{lookcolor}{rgb}{0.5,0.1,0.0}
 \definecolor{prooftermcolor}{rgb}{0.3,0.1,1.0}
@@
 \usepackage{graphicx}
-%\usepackage{aeguill}
 \else
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%% For Alectryon
@@
 \usepackage{texments}
-%%% for movies by alectryon
 \usepackage{./alectryon}
@@
 
-\newcommand{\pathtomovies}{.}%/movies}
+\newcommand{\pathtomovies}{.}
 
@@
 % End of Alectryon stuff
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \newcommand{\canonseq}[2]{\mbox{$\{#1\}(#2)$}}
+
 \title{Hydras \& Co.: Formalized mathematics in Coq\\
@@
 
-\usepackage{draftwatermark} 
+\usepackage[firstpage]{draftwatermark} 
 \setbeamercolor{background canvas}{bg=}%transparent canvas
@@
 \begin{document}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frame}
@@
 \end{frame}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Introduction}
+
 \begin{frame}
@@
     \item\TODO{}{}
-      \item Documentation on libraries
     \end{itemize}
@@
 \end{frame}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frame}
   \frametitle{Yet another book on \coq?}
-  \begin{block}{Kinds of books}
+  \begin{block}{Kinds of books (existing)}
     \TODO{}{}
@@
 \end{frame}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frame}
-  \frametitle{Example-driven books (1)}
+  \frametitle{Example-driven books}
   \begin{block}{}
     \begin{itemize}
-    \item \TODO{}{Definition}
+    \item The structure is determined by the presented topics (discrete math, algorithmics, etc.). \coq technology (formalization and proof patterns) is presented when needed.
     \item No need to be complete (there is already a reference manual)
     \item No need to introduce \coq (there are already a lot of initiation books and tutorials)
-    \item \coq patterns (formalization and proof) are presented  when they are needed.
     \item Allow to compare several different formalizations of a given notion, and several proofs of the same theorem.
-    \item By definition, the authors choose topics they like to teach.
-      \item \TODO{}{topics adapted to the potential readers?}
+    \end{itemize}
+  \end{block}
+  \begin{block}{Examples}
+    \begin{itemize}
+    \item Tsukuba Coq Users's Group: regular languages, $\lambda$-calculus, arithmetic, ... (in Japanese + ssreflect)
+    \item Software Foundations (B. Pierce \emph{et al.})
     \end{itemize}
@@
 \end{frame}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \begin{frame}
-%   \frametitle{Example-driven books (2)}
-%   \begin{block}{Tsukuba Coq User's Group}
-%     Regular languages, lambda-calculus, programming language semantics, \dots
 
-%    \includegraphics[height=35mm]{tsukubabooks.jpg}
-%    %\includegraphics[height=45mm]{leaflet.jpg}
-%      \end{block}
-%      \begin{block}{\Hydras}
-%        Kirby \& Paris'  hydra battles, exponentiation algorithms, \dots
-%      \end{block}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% \end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}
@@
 \end{frame}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Example: Kirby \& Paris' hydras}
+
 \begin{frame}
   \frametitle{Example: Kirby \& Paris' hydras}
- %  \begin{block}{}
-
-%   \begin{enumerate}
-%   \item Intuitive presentation
-%     \item From discrete math to \coq
-%   \end{enumerate}
-%   \end{block}
-
-% \end{frame}
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \begin{frame}
-%   \frametitle{The life of $\omega^{\omega^2+1}+1$}
-
   \begin{figure}[h]
@@
       \draw (h2) -- (h3) ;
-  \draw (h1) -- node[red!90!black,font=\small,sloped,shift={(0.01,-0.075)},rotate=90]{\textbf{\ding{34}}} (h7);
+      \draw (h1) -- node[red!90!black,font=\small,sloped,
+      shift={(0.01,-0.075)},rotate=90]{\textbf{\ding{34}}} (h7);
       \draw (h3) -- (h4) ;
@@
       \draw (h12) -- (h13) ;
-  \draw (h12) -- node[red!90!black,font=\small,sloped,shift={(0.01,-0.075)},rotate=90]{\textbf{\ding{34}}} (h16);
+      \draw (h12) -- node[red!90!black,
+      font=\small,sloped,shift={(0.01,-0.075)},
+      rotate=90]{\textbf{\ding{34}}} (h16);
       \draw (h13) -- (h14) ;
@@
       \draw (hn3c) -- (hn4c) ;
-  \draw (hn3c) -- node[red!90!black,font=\small,sloped,shift={(0.01,-0.075)},rotate=90]{\textbf{\ding{34}}} (hn5c) ;
+      \draw (hn3c) -- node[red!90!black,
+      font=\small,sloped,shift={(0.01,-0.075)},
+      rotate=90]{\textbf{\ding{34}}} (hn5c) ;
     \end{tikzpicture}
@@
     \centering
-
     \begin{tikzpicture}[very thick, scale=0.25]
@@
       \node [blue](hn4c) at (6,6){$\bullet$};
- % \node (hn5c) at (8,6){$\bullet$};
       \draw (hn1) -- (hn2c) ;
@@
       \draw (hn3c) -- (hn4c) ;
-  % \draw (hn3c) -- (hn5c) ;
       \node [blue](hn3d) at (8,4) {$\bullet$};
@@
       \node [red!90!black](hn4f) at (12,6) {$\bullet$};
-  \draw (hn3f) -- node[red!90!black,font=\small,sloped,shift={(0.01,-0.075)},rotate=90]{\textbf{\ding{34}}} (hn4f) ;
+      \draw (hn3f) -- node[red!90!black,
+      font=\small,sloped,shift={(0.01,-0.075)},
+      rotate=90]{\textbf{\ding{34}}} (hn4f) ;
 
@@
       \node [red!90!black](hn4g) at (21,2) {$\bullet$};
-   \draw (hn4h) -- node[red!90!black,font=\small,sloped,shift={(0.01,-0.075)},rotate=90]{\textbf{\ding{34}}} (hn4g) ;
+      \draw (hn4h) -- node[red!90!black,
+      font=\small,sloped,shift={(0.01,-0.075)},
+      rotate=90]{\textbf{\ding{34}}} (hn4g) ;
 
       \node [black](hn4f) at (24,0) {$\bullet$};
+
     \end{tikzpicture}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frame}
-  \frametitle{Mathematical formalization of hydra battles}
-   \begin{block}{}
+    \begin{block}{Mathematical formalization}
 
@@
     \item   We associate a hydra to each ordinal \textcolor{mathcolor}{$\alpha<\epsilon_0$} (in Cantor normal form).
-For instance, the first hydra of our example is the image of \textcolor{mathcolor}{$omega^{\omega^2+1}+1$}.
-    \item A round (at time $t$) of a battle transforms an hydra associated with \textcolor{mathcolor}{$\alpha$} into the hydra associated with \textcolor{mathcolor}{$\canonseq{\alpha}{t+1}$}, the $t+1$-st element of the \emph{canonical sequence} of \textcolor{mathcolor}{$\alpha$}.
+For instance, the first hydra of our example is the image of \textcolor{mathcolor}{$\omega^{\omega^2+1}+1$}.
+    \item A round (at time $t\in\mathbb{N}$) of a battle transforms an hydra associated with \textcolor{mathcolor}{$\alpha$} into the hydra associated with \textcolor{mathcolor}{$\canonseq{\alpha}{t+1}$}, the $t+1$-st element of the \emph{canonical sequence} of \textcolor{mathcolor}{$\alpha$}.
     \item We want to ``compute'' the length of sequences of the form 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frame}
+  \begin{block}{}
+    \begin{itemize}
+    \item Sequences associated with hydra battles are studied
+      as \emph{$\alpha$-large sets of natural numbers}  in Ketonen \& Solovay's article
+      \emph{Rapidly Growing {R}amsey Functions (1981)}.
+
+    \item Let \textcolor{mathcolor}{$0<\alpha<\epsilon_0$}, and \textcolor{mathcolor}{$i\in\mathbb{N}$}.
+      The least index \textcolor{mathcolor}{$k$} such that \textcolor{mathcolor}{$\alpha_k=0$} is  \textcolor{mathcolor}{$H'_\alpha(i)$} where \textcolor{mathcolor}{$H'_\alpha$} is defined by transfinite induction over \textcolor{mathcolor}{$\alpha$}.
+
+      \begin{align}
+        H'_0(i) & = i\\
+        H'_\alpha(i) &= H'_{(\canonseq{\alpha}{i+1})}(i)  \quad\textit{if $\alpha$ is a limit ordinal}\\
+        H'_{\alpha}(i) &=H'_\beta(i+1) \quad\textit{if $\alpha=\beta+1$}
+      \end{align}
+
+    \end{itemize}
+  \end{block}
 
   \begin{block}{}
-    Ordinals less than $\epsilon_0$ are represented as \emph{ordinal notations} which allow the user to write proofs by computation (comparison, canonical sequences, etc.)
+    The \textcolor{mathcolor}{$H'_\alpha$} family,  is a variant of the  \emph{Hardy hierarchy of rapidly growing functions}.
+For instance, \textcolor{mathcolor}{$H'_{\omega^3+3}(0)\geq2^{2^N}$}, where \textcolor{mathcolor}{$N=2^{70}-1$}.
+  \end{block}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+  \frametitle{Ordinals in \coq}
+
+  \begin{block}{}
+    Ordinals less than $\epsilon_0$ are represented as \emph{ordinal notations}: comparison, canonical sequences, partition limit/successors/zero ordinals are implemented by certified functions, allowing the user to write proofs by computation.
+
     \begin{footnotesize}
-      \inputsnippets{../../movies/snippets/ON_Generic/ONDef}
+      \inputsnippets{./ONDef}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frame}
   \begin{block}{}
-    \begin{itemize}
-    \item We borrow the systematic study of the sequencesfrom Ketonen \& Solovay's article
-      \emph{Rapidly Growing {R}amsey Functions (1981)}.
-
-    \item Let \textcolor{mathcolor}{$0<\alpha<\epsilon_0$}, and \textcolor{mathcolor}{$i\in\mathbb{N}$}.
-      The least index \textcolor{mathcolor}{$k$} such that \textcolor{mathcolor}{$\alpha_k=0$} is  \textcolor{mathcolor}{$H'_\alpha(i)$} where \textcolor{mathcolor}{$H'_\alpha$} is defined by transfinite induction over \textcolor{mathcolor}{$\alpha$}.
-
-   %    \begin{align}
-%   H'_0(i) & = i\\
-%   H'_\alpha(i) &= H'_{(\canonseq{\alpha}{i+1})}(i)  \quad\textit{if $\alpha$ is a limit ordinal}\\
-%   H'_{\alpha}(i) &=H'_\beta(i+1) \quad\textit{if $\alpha=\beta+1$}
-% \end{align}
+    The functions $H'_\alpha$ are ``implemented'' with \textcolor{plugincolor}{coq-equations}.
     \vspace{5pt}
@@
 
-    \end{itemize}
-
   \end{block}
@@
   \begin{block}{}
-    The \textcolor{mathcolor}{$H'_\alpha$} family is an example of \emph{hierarchy of rapidly growing functions}.
-For instance, \textcolor{mathcolor}{$H'_{\omega^3+3}(0)\geq2^{2^N}$}, where \textcolor{mathcolor}{$N=2^{70}-1$}.
+    Using combinatorial lemmas by  [K. \& S.], we prove (by transfinite induction on \textcolor{mathcolor}{$\alpha$}):
+    \begin{quote}
+      \begin{itemize}
+      \item \textcolor{mathcolor}{$H'_\alpha$} is strictly monotonous
+      \item \textcolor{mathcolor}{$\forall i, i \leq H'_\alpha(i)$}
+      \item   If \textcolor{mathcolor}{$\beta<\alpha$}, then \textcolor{mathcolor}{$H'_\alpha(i)>H'_\beta(i)$} for all but a finite number of $i$ (we say that \textcolor{mathcolor}{$H'_\alpha$}
+        \emph{dominates} \textcolor{mathcolor}{$H_\beta$}).
+      \end{itemize}
+    \end{quote}
+
   \end{block}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frame}
-  \begin{block}{}
-
-    Properties of the \textcolor{mathcolor}{$H'_\alpha$}s are proved with the help of the lemmas automatically generated by \textcolor{plugincolor}{equations}.
-For instance, the \textcolor{mathcolor}{$H'_\alpha$} are strictly monotonous,
-and if \textcolor{mathcolor}{$\beta<\alpha$}, then \textcolor{mathcolor}{$H'_\alpha(i)>H'_\beta(i)$} for all but a finite number of $i$.
+  \begin{block}{Theorem}
+    For any   \textcolor{mathcolor}{$\alpha\geq\omega^\omega$} the function \textcolor{mathcolor}{$H'_\alpha$} is not primitive recursive.
   \end{block}
@@
   \begin{block}{}
-    We prove also that, for any   \textcolor{mathcolor}{$\alpha\geq\omega^\omega$} the function \textcolor{mathcolor}{$H'_\alpha$} is not primitive recursive.
+    {\color{coqstylecolor}
+      \begin{itemize}
+      \item \emph{In order to prove this theorem, we chose to host and  maintain under \community Russel O'Connor's library \textcolor{plugincolor}{goedel}, which contains a formalization of primitive recursive functions.}
 
-    \begin{itemize}
-    \item \emph{In order to prove this theorem, we chose to maintain under \community Russel O'Connor's library \textcolor{plugincolor}{goedel}, which contains a formalization of primitive recursive functions.}
+      \item    \emph{The interest of O'Connor's contribution led us to write a chapter on primitive recursive functions, with [counter-]examples and exercices.}
+
       \item \emph{Proving formally that a given function is not primitive recursive is a nice example of application of a complex (mutually recursive, dependently typed) induction principle.
@@
 
-        \item    \emph{The interest of O'Connor's contribution led us to write a chapter on primitive recursive functions, with examples and exercices.}
       \end{itemize}
+    }
   \end{block}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frame}
-  \frametitle{Some features of the \Hydras project}
+
+  \begin{scriptsize}
+    \inputsnippets{./SchemePrimRecInda}
+  \end{scriptsize}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+  \begin{block}{}
+    Using  \texttt{PrimRec\_PrimRecs\_ind}, we prove the following lemma:
+
+    \begin{quote}
+      For any \textcolor{mathcolor}{$n$}, and any primitive recursive function \textcolor{mathcolor}{$f$} of  arity \textcolor{mathcolor}{$n$}, there exists some natural number \textcolor{mathcolor}{$q$} such that the following inequality holds:
+      {\color{mathcolor}
+        \[
+          \forall x_1,\dots,x_n, 
+          f(x_1,\dots,\,x_n)\leq\textrm{Ack}(q,\textrm{max}(x_1,\dots,x_n))
+        \]}
+    \end{quote}
+  \end{block}
+
+  \begin{block}{}
+
+    Instantiating with \textcolor{mathcolor}{$n=1$}, we obtain:
+
+    \begin{quote}
+      If \textcolor{mathcolor}{$H'_{\omega^\omega}$} were primitive recursive, there would exists some \textcolor{mathcolor}{$q$} such that, for any \textcolor{mathcolor}{$k$}, \textcolor{mathcolor}{$H'_{\omega^\omega}(k) \leq\textrm{Ack}(q,k)$}, which is false.
+    \end{quote}
+
+  \end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+\section{Presenting proofs with \alectr}
+\begin{frame}
+  \frametitle{Presenting proofs with \alectr}
   \begin{block}{}
     \begin{itemize}
-    \item The order in which \coq code is presented in the book is determined by the difficulty of the mathematical examples, or the assumed level in \coq of the reader, not by the structure of the libraries.
-    \item We often present and compare several different definitions and proofs of the same mathematical notions, including ``bad'' ones.
-      \item We would be happy to include new proofs, if they improve old ones, or they illustrate new proof patterns. \emph{Please send PRs or issues!}.
+    \item The order in which \coq code; goals and answers are presented in the book is determined by the difficulty of the mathematical examples, or the assumed level in \coq of the reader, not by the structure of the libraries.
+    \item
+      \emph{\coq libraries contain \emph{snippet declarations} which are converted during the compilation of the book into \LaTeX\xspace blocks which are inserted into the book.}
+
+    \item {\color{lookcolor} \emph{Note that our build system guarantees that the accompanying book is always up-to-date and consistent with the \coq code.}}
     \end{itemize}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-  \TODO{}{maintenance, CI/CD,  etc.   (several slides).}
-\end{frame}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-  \frametitle{Presenting proofs with \alectr}
-  \begin{small}
-
- %\begin{footnotesize}
-  \inputsnippets{../Schutte_Ex42a}   
- %\end{footnotesize}
-
-By definition of additive principal ordinals, 
-it suffices to prove the inequality $\omega+42< \phi_0(2)$.
-%\begin{footnotesize}
-\inputsnippets{../Schutte_Ex42b}
-By Lemma \textit{AP\_plus\_closed}, it suffices to prove  $\omega<\omega^2$ and $42<\omega^2$.
-%\end{footnotesize}
-\end{small}
-
-\end{frame}
-
-\begin{frame}
-  %\frametitle{Presenting proofs with \alectr (2)}
-\begin{small}
-%\begin{footnotesize}
-\inputsnippets{
-  ../Schutte_Ex42c, ../Schutte_Ex42d, ../Schutte_Ex42e}
-%\end{footnotesize}
-  \end{small}
-
-\end{frame}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]
@@
 \end{frame}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frame}
-  \frametitle{Perspectives}
-  \begin{block}{Library on ordinals}
+  \begin{small}
+
+     \inputsnippets{../Schutte_Ex42a}
+
+     By definition of additive principal ordinals, 
+    it suffices to prove the inequality $\omega+42< \phi_0(2)$.
+
+    \inputsnippets{../Schutte_Ex42b}
+    By Lemma \textit{AP\_plus\_closed}, it suffices to prove  $\omega<\omega^2$ and $42<\omega^2$.
+  \end{small}
+
+\end{frame}
+
+\begin{frame}
+  \begin{small}
+    \inputsnippets{
+      ../Schutte_Ex42c, ../Schutte_Ex42d, ../Schutte_Ex42e}
+  \end{small}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+
+  \begin{block}{Remarks}
     \begin{itemize}
-    \item \TODO{}{Gaia}
-      \item \TODO{}{Aczel, Minki}
+    \item We could write the book, because we are also the maintainers of the cited \coq source, and could insert the snippet instructions in the \texttt{.v} files.
+\textcolor{lookcolor}{But what about commenting interesting code from other people (\emph{e.g}
+        \gaia) ?}
+    \item The same  piece of \coq code may be used in several documents, with different styles (book, beamer), and thus
+      different filters (hiding or showing subgoals, hypotheses, etc.).
+
+    \end{itemize}
+  \end{block}
+
+  \begin{block}{To do ?}
+    Define a language for
+    \begin{itemize}
+    \item Searching sentences in any \coq file,
+    \item Applying filters in order to generate as many snippets as needed.
     \end{itemize}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frame}
-  \frametitle{Perspectives(2)}
-  \begin{block}{Documentation with Alectryon}
-    \begin{itemize}
-    \item \TODO{}{}
-    \end{itemize}
-  \end{block}
+  \TODO{}{maintenance, CI/CD,  etc.   (several slides).}
 \end{frame}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-  \frametitle{Perspectives(3)}
-  \begin{block}{CI/CD}
-     \TODO{}{}
-  \end{block}
-\end{frame}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
     \begin{itemize}
-    \item \TODO{}{}
+    \item Changes in definitions and proofs already happen
+      (\emph{e.g.} type class structure), without making the book's maintenance difficult.
     \item \Hydras and \community are on github. Please contribute!
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[fragile]
+  \frametitle{Perspectives}
+  \begin{block}{Library on ordinals}
+    There exist already \coq libraries which formalize ordinal numbers.
+    \begin{itemize}
+    \item \gaia (José Grimm) : Formalization of Bourbaki's Elements of Mathematics.
+{\color{lookcolor}\emph{This library uses an ordinal representation compatible with ours.
+It is  written in SSreflect/Mathcomp, and is also maintained on \community}.}
+    \item A formalization by Minki Cho, using a different representation.
+\begin{verbatim}
+Inductive Ord.t :=
+| Ord.build: 
+    forall (A: Type), (A -> Ord.t) -> Ord.t.
+\end{verbatim}
+
+    \end{itemize}
+
+    Both libraries contain set-theoretical results about larger ordinals than ours.
+Having a library containing both our combinatorial results (large sets and rapidly growing functions) and mathematical definitions of ordinals would be an interesting project.
+  \end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+  \frametitle{Perspectives(2)}
+  \begin{block}{CI/CD}
+    \TODO{}{}
+  \end{block}
+  \begin{block}{Alectryon}
+     \TODO{}{}
+  \end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \end{document}

### 2021-12-26 doc/JFLA2022/Slides/Slides.tex
small changes

@@
 
+\definecolor{orange}{rgb}{1.0,0.6,0.0}
 \definecolor{plugincolor}{rgb}{0.6,0.0,0.6}
@@
 \section{Introduction}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}
-  \frametitle{Coq-community}
-  \begin{block}{Objectives}
-    \begin{itemize}
-    \item \TODO{}{}
-    \item\TODO{}{}
-    \end{itemize}
-  \end{block}
+  \frametitle{Introduction}
+  \TODO{}{Develop the abstract and background of the paper}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Yet another book on \coq?}
-  \begin{block}{Kinds of books (existing)}
-    \TODO{}{}
-  \end{block}
+  %\frametitle{Yet another book on \coq?}
+  \begin{table}
+  \centering
+  \footnotesize
+  \begin{tabular}{lll}
+  \hline
+  \textbf{Title} & \textbf{Year} & \textbf{Category}\\
+    \hline
+   \coq reference manual &  $\geq 1989$ & \\
+  Coq'Art & 2004 & Traditional\\
+  Software Foundations & 2007 & Executable\\
+  Certified Programming with Dependent Types & 2011 &  Executable\\
+  Program Logics for Certified Compilers & 2014 & Traditional\\
+  Programs and Proofs & 2014 & Executable\\
+  Formal Reasoning About Programs& 2017 & Traditional\\
+  Computer Arithmetic and Formal Proofs & 2017 & Traditional\\
+    Mathematical Components & 2018 & Traditional\\
+  \hline
+  \end{tabular}
+  \caption{Coq book categorization.}
+  \label{tbl:books}
+\end{table}
+
 \end{frame}
@@
       \textcolor{mathcolor}{$\langle \alpha_k=\alpha,\dots,\alpha_{i+1}=\canonseq{\alpha_i}{i+1},\dots, 1,0\rangle$}, where \textcolor{mathcolor}{$k\in\mathbb{N}$}.
-
+      \item Such sequences are studied
+      as \emph{$\alpha$-large sets of natural numbers}  in Ketonen \& Solovay's article
+      \emph{Rapidly Growing {R}amsey Functions (1981)}.
     \end{itemize}
@@
     \begin{itemize}
-    \item Sequences associated with hydra battles are studied
-      as \emph{$\alpha$-large sets of natural numbers}  in Ketonen \& Solovay's article
-      \emph{Rapidly Growing {R}amsey Functions (1981)}.
-
     \item Let \textcolor{mathcolor}{$0<\alpha<\epsilon_0$}, and \textcolor{mathcolor}{$i\in\mathbb{N}$}.
@@
   \begin{block}{}
-    The functions $H'_\alpha$ are ``implemented'' with \textcolor{plugincolor}{coq-equations}.
+    The functions \textcolor{mathcolor}{$H'_\alpha$} are implemented with \textcolor{plugincolor}{coq-equations}.
     \vspace{5pt}
@@
   \begin{block}{}
-    Using combinatorial lemmas by  [K. \& S.], we prove (by transfinite induction on \textcolor{mathcolor}{$\alpha$}):
+    Using combinatorial lemmas by  K. \& S., we prove (by mutual  transfinite induction on \textcolor{mathcolor}{$\alpha$}):
     \begin{quote}
@@
 
+  \end{block}
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Project Organization}
+  \begin{frame}
+    \frametitle{Project  Organization}
+    \begin{itemize}
+    \item Documentation with \alectr
+     \item Integration with \community
+     \item Package dependencies
+       \item Continuous integration
+    \end{itemize}
+  \end{frame}
+
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  \begin{frame}
+    \frametitle{\community}
+    \begin{block}{}
+      \community is an informal organization that aims to maintain interesting open source \coq projects  andfacilitate collaboration among Coq users on documentation, tooling, etc.
+      \begin{itemize}
+      \item Created in 2018,  run by volunteer Coq users on GitHub
+      \item At present, over than 50 projects, maintained by over than 30 volunteers
+      \item Some packages were unmaintained before their integration to \community. According to their scientific or historical interest, we decided to revive them.
+        \item After their transfer, packages may be the object of large changes  and refactorings.
+      \end{itemize}
     \end{block}
@@
 \end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+\begin{frame}
+  \frametitle{Package organization and dependencies}
+
+\begin{figure}[ht]
+\centering
+\scalebox{0.7}{
+\begin{tikzpicture}[thick]
+
+\begin{scope}[xshift=-6cm]
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (additions) { Additions };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (cantor) { Cantor };
+
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3.2cm] (cats) { Categories in ZFC };
+\end{scope}
+
+\begin{scope}[yshift=-1cm,xshift=0.45cm]
+  \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=2.5cm] (mathcomp) { MathComp };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=paramcoq, node distance=4.5cm] (equations) { Equations };
+\end{scope}
+
+\begin{scope}[yshift=-2cm,xshift=-4cm]
+\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
+\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,left of=hydras, node distance=3.5cm] (pockcc) { Pocklington };
+\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=hydras, node distance=2.5cm] (chains) { Addition-chains };
+\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,right of=chains, node distance=2.2cm] (gaia) { Gaia };
+\end{scope}
+
+\begin{scope}[yshift=-3cm,xshift=-6cm]
+ \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
+ \node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=3cm] (gaiahydras) { Gaia-hydras };
+\end{scope}
+
+\draw[lookcolor,->,dashed] (additions.south) -- (chains.north) ;
+\draw[lookcolor,->,dashed] (goedel) -- (goedelcc) ;
+\draw[lookcolor,->,dashed] (goedel) -- (hydras.north west) ;
+\draw[lookcolor,->,dashed] (cantor) -- (hydras.north east) ;
+\draw[lookcolor,->,dashed] (cantor.south) -- (gaia.north west) ;
+\draw[lookcolor,->,dashed] (pock) -- (pockcc) ;
+\draw[lookcolor,->,dashed] ([xshift=2.1cm]cats) -- (gaia) ;
+
+\draw[->] (hydras) -- (gaiahydras) ;
+\draw[->] (hydras.south west) -- (goedelcc) ;
+\draw[->] (gaia.south west) -- (gaiahydras.east) ;
+
+\draw[->] (equations) -- (hydras) ;
+\draw[->] (mathcomp) -- (gaia) ;
+\draw[->] (mathcomp) -- (chains) ;
+\draw[->] (paramcoq.south) -- (chains) ;
+\draw[->] (pockcc) -- (goedelcc) ;
+\end{tikzpicture}
+}
+\caption{\textcolor{lookcolor}{Genealogy} and dependencies for \Hydras packages.}
+ %Dotted boxes represent historical Coq contributions, while regular boxes represent maintained \coq packages.
+Orange packages are maintained in the \Hydras GitHub repository, while blue packages are maintained in other \community repositories.
+Dotted lines represent \coq code ancestry, while regular lines represent direct code dependencies.}
+  \label{fig:genealogy}
+\end{figure}
+
+%%%% TODO : Resize  !
+{\scriptsize
+\begin{table}[ht]
+\centering
+\scriptsize
+\begin{tabular}{llllrr}
+\hline
+\textbf{Package name} &  \textbf{Version} & \textbf{Spec LOC} & \textbf{Proof LOC}\\
+\hline
+Pocklington & \texttt{8.12.0} & 825 & 3,798 \\
+Goedel  & \texttt{8.13.0} & 2,799 & 10,762 \\
+Gaia &  \texttt{1.12} & 28,850 & 124,839 \\
+Hydra-battles & \texttt{0.5} & 14,368 & 48,385 \\
+Addition-chains & \texttt{0.5} & 1,961 & 2,210 \\
+Gaia-hydras &  \texttt{0.5} & 81 & 177 \\
+\hline
+\end{tabular}
+\caption{\scriptsize Current numbers of lines of code for packages in the \Hydras galaxy.}
+\label{tbl:loc}
+\end{table}}
+\end{frame}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-12-27 doc/JFLA2022/Slides/Slides.tex
new draft version of JFLA Slides

@@
 \definecolor{orange}{rgb}{1.0,0.6,0.0}
+\definecolor{darkorange}{rgb}{0.7,0.25,0.0}
 \definecolor{plugincolor}{rgb}{0.6,0.0,0.6}
@@
 \begin{frame}
-  \frametitle{Introduction}
-  \TODO{}{Develop the abstract and background of the paper}
+  \frametitle{The \Hydras project}
+  \begin{block}{}
+    \begin{itemize}
+    \item Experimental platform for the  development of \textcolor{lookcolor}{maintained} and \textcolor{lookcolor}{documented} libraries of formal proofs.
+    \item A project of \community, developped on Github
+      \begin{itemize}
+      \item CI/CD
+        \item Tools for collaboration (issues, PR, etc.)
+      \end{itemize}
+    \end{itemize}
+  \end{block}
+  \begin{block}{Contents}
+    \begin{itemize}
+    \item Evolved version  of two contributions: \textcolor{plugincolor}{Cantor} and \textcolor{plugincolor}{Additions}
+    \item Maintenance of \textcolor{plugincolor}{Ackermann}
+      (part of Russel O'Connor's \textcolor{plugincolor}{Goedel})
+    \item Preliminary bridge to \textcolor{plugincolor}{Gaia} (José Grimm)
+      \item Documentation in pdf (280p), generated  with \textcolor{lookcolor}{\alectr} and \LaTeX, continuously updated and deployed
+    \end{itemize}
+  \end{block}
+
 \end{frame}
@@
 \begin{frame}
-  %\frametitle{Yet another book on \coq?}
+  \frametitle{Yet another book on \coq?}
   \begin{table}
@@
 \begin{frame}
-  \frametitle{Example-driven books}
-  \begin{block}{}
+   \begin{block}{Example-driven books}
     \begin{itemize}
@@
 For instance, \textcolor{mathcolor}{$H'_{\omega^3+3}(0)\geq2^{2^N}$}, where \textcolor{mathcolor}{$N=2^{70}-1$}.
+
+    {\color{lookcolor} Such functions defy common programming intuition: no test is possible, only proofs!}
   \end{block}
@@
   \begin{block}{}
-    Ordinals less than $\epsilon_0$ are represented as \emph{ordinal notations}: comparison, canonical sequences, partition limit/successors/zero ordinals are implemented by certified functions, allowing the user to write proofs by computation.
+    Ordinals less than {\color{mathcolor}$\omega^\omega, \epsilon_0,  \Gamma_0$} are represented as \emph{ordinal notations}: comparison, canonical sequences, partition limit/successors/zero ordinals are implemented by certified functions, allowing the user to write proofs by computation.
 
@@
   \begin{block}{}
-    The functions \textcolor{mathcolor}{$H'_\alpha$} are implemented with \textcolor{plugincolor}{coq-equations}.
+    The functions \textcolor{mathcolor}{$H'_\alpha\;(\alpha< \epsilon_0)$} are implemented with \textcolor{plugincolor}{coq-equations}.
     \vspace{5pt}
@@
   \begin{itemize}
+  \item Integration with \community
   \item Documentation with \alectr
-     \item Integration with \community
   \item Package dependencies
@@
 
-\begin{frame}
-  \TODO{}{maintenance, CI/CD,  etc.   (several slides).}
-\end{frame}
+% \begin{frame}
+%   \TODO{}{maintenance, CI/CD,  etc.   (several slides).}
+% \end{frame}
 
@@
 \begin{frame}
-  \frametitle{Conclusion}
+  \frametitle{Conclusion and perspectives}
   \begin{block}{}
     \begin{itemize}
-    \item Changes in definitions and proofs already happen
-      (\emph{e.g.} type class structure), without making the book's maintenance difficult.
-    \item \Hydras and \community are on github. Please contribute!
+    \item \Hydras wants to be a connection between scientific litterature and proof assistant technology:
       \begin{itemize}
-      \item New chapters
-      \item Any improvement (text and/or proofs)
+      \item For the mathematician, it gives a concrete view of the mathematical contents: full proofs, examples, computable functions whenever possible, etc.
+        \item  For the \coq learner, it provides a consistent set of examples,
+allowing to present and compare various formalization and proving techniques.
       \end{itemize}
@@
   \end{block}
-
   \end{frame}
@@
     \item A formalization by Minki Cho, using a different representation.
+{\color{darkorange}
 \begin{verbatim}
@@
 \end{verbatim}
-
+}  
     \end{itemize}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \begin{frame}
-  \frametitle{Perspectives(2)}
-  \begin{block}{CI/CD}
-    \TODO{}{}
+  \frametitle{More TODOs}
+  \begin{block}{}
+    \begin{itemize}
+    \item More results about hyper-complexity and fast growing functions
+    \item New toolchain for \alectr
+    \item More proof and specification patterns
+    \item etc.
+    \item \Hydras and \community are on github. Please contribute!
+      \begin{itemize}
+      \item New chapters
+      \item Any correction or improvement (text and/or proofs)
+
+      \end{itemize}
+    \end{itemize}
+
   \end{block}
-  \begin{block}{Alectryon}
-     \TODO{}{}
-  \end{block}
+
 \end{frame}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \end{document}

### 2021-12-27 doc/JFLA2022/Slides/Slides.tex
Minor fixes.

@@
     \begin{block}{}
-      \community is an informal organization that aims to maintain interesting open source \coq projects  andfacilitate collaboration among Coq users on documentation, tooling, etc.
+      \community is an informal organization that aims to maintain interesting open source \coq projects and facilitate collaboration among Coq users on documentation, tooling, etc.
       \begin{itemize}
@@
     \begin{itemize}
-    \item The order in which \coq code; goals and answers are presented in the book is determined by the difficulty of the mathematical examples, or the assumed level in \coq of the reader, not by the structure of the libraries.
+    \item The order in which \coq code, goals and answers are presented in the book is determined by the difficulty of the mathematical examples, or the assumed level in \coq of the reader, not by the structure of the libraries.
     \item
@@
     \item We could write the book, because we are also the maintainers of the cited \coq source, and could insert the snippet instructions in the \texttt{.v} files.
-\textcolor{lookcolor}{But what about commenting interesting code from other people (\emph{e.g}
+\textcolor{lookcolor}{But what about commenting interesting code from other people (\emph{e.g.},
         \gaia) ?}
@@
 {\color{lookcolor}\emph{This library uses an ordinal representation compatible with ours.
-It is  written in SSreflect/Mathcomp, and is also maintained on \community}.}
+It is  written in SSReflect/Mathcomp, and is also maintained in \community}.}
     \item A formalization by Minki Cho, using a different representation.
@@
     \item etc.
-    \item \Hydras and \community are on github. Please contribute!
+    \item \Hydras and \community are on GitHub. Please contribute!
       \begin{itemize}

### 2021-12-27 doc/JFLA2022/Slides/Slides.tex
Merge changes
add a slide on CI/CD

@@
 
-% \begin{frame}
-%   \TODO{}{maintenance, CI/CD,  etc.   (several slides).}
-% \end{frame}
+\begin{frame}
+  \frametitle{Maintenance, CI/CD}
+  \begin{block}{CD}
+    The continuously evolving version of the book is continuously deployed from the master branch to the \community website.
+  \end{block}
+
+  \begin{block}{Reverse Dependencies Compatibility Testing}
+    Since we have extracted part of {\color{plugincolor}goedel}, now {\color{plugincolor}goedel} depends on {\color{plugincolor}\Hydras}, and thus we are checking in CI that we do not break it (thanks to the \textcolor{lookcolor}{Coq Nix Toolbox},
+    which can generate such CI workflows)
+  \end{block}
+
+  \begin{block}{}
+  {\color{lookcolor}  With \Hydras, we hope to demonstrate that projects with complex dependencies are feasible to manage using recent advances in build management and infrastructure.}
+  \end{block}
+
+\end{frame}
 
@@
     \begin{itemize}
-    \item \Hydras wants to be a connection between scientific litterature and proof assistant technology:
+    \item \Hydras wants to provide  a connection between scientific litterature and proof assistant technology:
       \begin{itemize}
@@
     \item etc.
-    \item \Hydras and \community are on GitHub. Please contribute!
+    \end{itemize}
+  \end{block}
+
+  \begin{block}{Call to contributors }
+  {\color{plugincolor}\Hydras} and  {\color{plugincolor}\community} are on GitHub. Please contribute!
       \begin{itemize}
@@
     \end{itemize}
-    \end{itemize}
-
      \end{block}

### 2021-12-28 doc/hydras.tex
comment out the watermark text
remove the book's subtitle
new title page

@@
 \usepackage{synttree, proof, centernot}
-\usepackage[firstpage]{draftwatermark}
+%\usepackage[firstpage]{draftwatermark}
 \usepackage{verbatim}
@@
 
-\SetWatermarkLightness{0.8}
-\definecolor{lightorange}{rgb}{1,0.9,0.9}
-\SetWatermarkScale{0.35}
-\SetWatermarkText{Work Continuously in Progress}
-\SetWatermarkColor{lightorange}
+% \SetWatermarkLightness{0.8}
+\definecolor{darkblue}{rgb}{0,0.1,0.7}
+% \SetWatermarkScale{0.35}
+% \SetWatermarkText{Work Continuously in Progress}
+% \SetWatermarkColor{lightorange}
 
@@
 
-\title{Hydras \& Co.\\ Formalized mathematics in \coq for inspiration and entertainment}
+\title{Hydras \& Co.}
 %\pagestyle{fancy}
@@
 \begin{document}
-\maketitle
+\begin{titlepage}
+   \begin{center}
+       \vspace*{1cm}
 
-% \cleardoublepage
+       \textbf{\LARGE Hydras \& Co.}
 
-\tableofcontents
+       \vspace{0.5cm}
+        Formalized mathematics in \coq for inspiration and entertainement
 
-% \mbox{~}
+       \vspace{1.5cm}
+
+       \emph{Pierre Castéran, LaBRI, Univ.
+Bordeaux, CNRS UMR 5800, \\ With contributions by Yves Bertot, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Karl Palmskog, Clément Pit-Claudel, and Théo Zimmermann.
+\\ The formalization of primitive recursive functions was originally authored by Russel O'Connor\cite{OConnor05}.}
+
+ \vfill
+        \today
+
+      \end{center}
+\vfill
+      \centering
+      \includegraphics[width=13cm]{epsilon0.jpg}
+      \centerline{\emph{\small {\color{darkblue}Ordinal numbers (less than $\Gamma_0$)}}}
+\end{titlepage}
+
 \clearpage
-% \newpage
+\newpage
 % \mbox{~}
@@
 \end{quote}
+
+ \tableofcontents
 

### 2021-12-28 doc/hydras.tex
new title page

@@
 
-       \textbf{\LARGE Hydras \& Co.}
+       \textbf{\Huge Hydras \& Co.}
 
        \vspace{0.5cm}
-        Formalized mathematics in \coq for inspiration and entertainement
+        {\emph{\Large Formalized mathematics in \coq for inspiration and entertainement}}
 
@@
 Bordeaux, CNRS UMR 5800, \\ With contributions by Yves Bertot, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Karl Palmskog, Clément Pit-Claudel, and Théo Zimmermann.
-\\ The formalization of primitive recursive functions was originally authored by Russel O'Connor\cite{OConnor05}.}
+\\ The formalization of primitive recursive functions was originally authored by Russel O'Connor.}%\cite{OConnor05}.}
 
@@
       \includegraphics[width=13cm]{epsilon0.jpg}
-      \centerline{\emph{\small {\color{darkblue}Ordinal numbers (less than $\Gamma_0$)}}}
+      \centerline{\emph{\small {\color{darkblue}Ordinal numbers in Veblen normal form}}}
 \end{titlepage}
@@
 \begin{itemize}
-\item The present distribution has been checked with version 8.13.2 of the Coq proof assistant, with the plug-ins \texttt{coq-paramcoq}, \texttt{coq-equations}
+\item The present distribution has been checked with version 8.14.1 of the Coq proof assistant, with the plug-ins \texttt{coq-paramcoq}, \texttt{coq-equations}
 and \texttt{coq-mathcomp-algebra}.
@@
 
-Many thanks also to the Coq development team, Yves Bertot, and the members of the \emph{Coq Club} for interesting discussions about the \coq{} system and the Calculus of Inductive Constructions.
+Many thanks also to the Coq development team and the members of the \emph{Coq Club} for interesting discussions about the \coq{} system and the Calculus of Inductive Constructions.
 

### 2021-12-29 doc/JFLA2022/snippets/Hprime/HprimeDef.tex

@@
     \begin{input}
-      \PY{n}{Equations}~\PY{n}{H\PYZsq{}\PYZus{}}~\PY{o}{(}\PY{n}{alpha}\PY{o}{:}~\PY{n}{E0}\PY{o}{)}~\PY{o}{(}\PY{n}{i}\PY{o}{:}\PY{n}{nat}\PY{o}{)}~\PY{o}{:}~~\PY{n}{nat}~~\PY{n+nb+bp}{by}~\PY{k+kn}{wf}~~\PY{n}{alpha}~\PY{n}{Lt}~\PY{o}{:=}\nl
+      \PY{k+kn}{Equations}~\PY{n+nf}{H\PYZsq{}\PYZus{}}~\PY{o}{(}\PY{n+nv}{alpha}\PY{o}{:}~\PY{n}{E0}\PY{o}{)}~\PY{o}{(}\PY{n+nv}{i}\PY{o}{:}\PY{n}{nat}\PY{o}{)}~\PY{o}{:}~~\PY{n}{nat}~~\PY{n+nb+bp}{by}~\PY{k+kn}{wf}~~\PY{n}{alpha}~\PY{n}{Lt}~\PY{o}{:=}\nl
       ~~\PY{n}{H\PYZsq{}\PYZus{}}~\PY{n}{alpha}~~\PY{n}{i}~\PY{k+kr}{with}~\PY{n}{E0\PYZus{}eq\PYZus{}dec}~\PY{n}{alpha}~\PY{n}{Zero}~\PY{o}{:=}\nl

### 2021-12-29 doc/hydras.tex

@@
 \makeatletter
+\newcounter{snippets}
 \newcommand{\inputsnippets}[1]
@@
   {{\renewenvironment{alectryon}{}{}% Skip \begin{alectryon} included in snippet
+    \refstepcounter{snippets}%
     \input@old{movies/snippets/#1}}}
+\AtEndDocument{\message{Snippets: \thesnippets}}
 \makeatother

### 2022-01-04 doc/JFLA2022/snippets/E0/Ex42.tex

--- 2021-10-04 doc/JFLA2022/snippets/E0/Ex42.tex
@@
     \begin{input}
-      \PY{k+kn}{Example}~\PY{n+nf}{Ex42}\PY{o}{:}~\PY{o}{(}\PY{n}{omega}~\PY{o}{+}~\PY{l+m+mi}{42}~\PY{o}{+}~\PY{n}{omega}\PY{o}{\PYZca{}}~\PY{l+m+mi}{2}~\PY{o}{=}~\PY{n}{omega}\PY{o}{\PYZca{}}~\PY{l+m+mi}{2}\PY{o}{)\PYZpc{}}\PY{n}{e0}\PY{o}{.}\nl
+      \PY{k+kn}{Example}~\PY{n+nf}{Ex42}\PY{o}{:}~\PY{n}{omega}~\PY{o}{+}~\PY{l+m+mi}{42}~\PY{o}{+}~\PY{n}{omega}\PY{o}{\PYZca{}}\PY{l+m+mi}{2}~\PY{o}{=}~\PY{n}{omega}\PY{o}{\PYZca{}}\PY{l+m+mi}{2}\PY{o}{.}\nl
     \end{input}

### 2022-01-04 doc/gaia-chapter.tex
new title page

@@
 \chapter{Gaia and the hydras (draft)}
+\index{coq}{Plug-ins!Gaia}
 
 \section{Introduction}
+The \gaia project by Jos\'e Grimm aimed to formalize mathematics in \coq in the style of Nicolas Bourbaki.
+The formalization of the first book in the Elements of Mathematics series by Bourbaki, on the theory of sets, was initially described in a technical report in July 2009~\cite{Grimm2009a}.
+The set-theoretic axioms and basic definitions in \gaia were derived from an earlier development by Carlos Simpson~\cite{Simpson2004,CatsZFCContrib}.
+Grimm then wrote (and continually updated) technical reports describing the formalization of Bourbaki's two subsequent books~\cite{Grimm2009b,Grimm2016} and additional topics in number theory~\cite{grimm:hal-00911710,Grimm2014}, before he passed away in 2019.
+
+In 2020, members of \community transferred the \gaia source code to GitHub and adapted it for recent releases of the Mathematical Components library, which \gaia heavily relies on.
+Anonymous volunteers (``collaborators of Nicolas Bourbaki'') then finished the only in-progress proof left by Grimm.
+At around 155,000 LOC, \gaia is currently one of the largest maintained open source \coq projects~\cite{Gaia}.
+
+\gaia contains definitions of ordinals in Cantor and Veblen normal form~\cite{grimm:hal-00911710}, adapted from the historical Cantor contribution~\cite{CantorContrib}.
+The data types for ordinals are essentially defined the same way as in \Hydras, but they are not identical inside \coq, e.g., due to residing in different modules.
+There are also minor differences in how ordinal arithmetic is implemented, due to the different evolutionary paths taken since divergence from the ancestor library.
+
+As an initial step in bridging \gaia and \Hydras, we were able to establish a correspondence between both implementations of ordinals through rewriting lemmas.
+For instance, we proved that multiplication of ordinals in Cantor normal form in \Hydras refine \gaia's corresponding multiplication operation, and then imported \gaia's proof of associativity of the multiplication almost for free.
+
+The initial draft bridge code mostly uses the SSReflect proof language and idioms from the Mathematical Components library.
+We made this design decision since we believe it is less challenging to reason about Hydra-battles code using SSReflect and MathComp than to reason about Gaia without SSReflect.
 

### 2022-01-05 doc/part-hydras.tex
new title page
fix typos in chapter 2 and bibliography

@@
 \begin{itemize}
-\item Semi automatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
+\item Semiautomatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
 \item Extension to $\Gamma_0$ (in Veblen normal form)
-\item More lemmas about hierarchies of rapidly growing functioit ns, and their relationship 
+\item More lemmas about hierarchies of rapidly growing
+  functions, and their relationship 
     with primitive recursive functions and provability in Peano arithmetic 
@@
 In order to describe trees where each node can have an arbitrary (but finite) number of sons, it is usual to define a type where each node carries a \emph{forest}, \emph{i.e} a list of trees
-(see for instance Chapter 14, pages 400-406 of \cite{BC04}).
-
-For this purpose, we define two mutual \emph{ad-hoc}  inductive types, where \texttt{Hydra} is the main type, and \texttt{Hydrae} is a helper for describing finite sequences of hydra.
+(see for instance Chapter 14, pages 400-406 of \cite{BC04,BC04ch14}).
+
+For this purpose, we define two mutual \emph{ad-hoc}  inductive types, where \texttt{Hydra} is the main type, and \texttt{Hydrae} is a helper for describing finite sequences of hydras.
 \label{types:Hydra}
@@
 Using this representation, re-define all the constructions of this chapter.
-You will probably have to use patterns described for instance in~\cite{BC04} or the archives of the Coq-club~\cite{Coq}.
+You will probably have to use patterns described for instance in~\cite{BC04} or the archives of the \coq communication channels (please consult~\url{https://coq.inria.fr/community.html}).
 
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=11cm]{epsilon0.jpg}
-  \caption{Veblen normal form}
-  \label{fig:gamma0}
-\end{figure}
-
 Like in chapter~\ref{chap:T1}, we get familiar with the type \texttt{T2} by recognizing simple constructs like finite ordinals, $\omega$, etc., as inhabitants of \texttt{T2}.

### 2022-01-05 doc/JFLA2022/Slides/ONDef.tex

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Class}~\PY{n+nf}{ON}~\PY{o}{\PYZob{}}\PY{n+nv}{A}\PY{o}{:}\PY{k+kt}{Type}\PY{o}{\PYZcb{}}~\PY{o}{(}\PY{n+nv}{lt}\PY{o}{:}~\PY{n}{relation}~\PY{n}{A}\PY{o}{)}~\PY{o}{(}\PY{n+nv}{cmp}\PY{o}{:}~\PY{n}{Compare}~\PY{n}{A}\PY{o}{)}~\PY{o}{:=}\nl
      ~~\PY{o}{\PYZob{}}\nl
      ~~\PY{n}{ON\PYZus{}comp}~\PY{o}{:\PYZgt{}}~\PY{n}{Comparable}~\PY{n}{lt}~\PY{n}{cmp}\PY{o}{;}\nl
      ~~\PY{n}{ON\PYZus{}wf}~\PY{o}{:}~\PY{n}{well\PYZus{}founded}~\PY{n}{lt}\PY{o}{;}\nl
      ~~\PY{o}{\PYZcb{}.}
    \end{input}
  \end{sentence}
\end{alectryon}

### 2022-01-05 doc/JFLA2022/Slides/SchemePrimRecInda.tex

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Check}~\PY{n}{PrimRec\PYZus{}PrimRecs\PYZus{}ind}\PY{o}{.}
    \end{input}
    \sep
    \begin{output}
      \begin{messages}
        \begin{message}
          \PY{n}{PrimRec\PYZus{}PrimRecs\PYZus{}ind}\nl
          ~~~~~\PY{o}{:}~\PY{k+kr}{forall}~\PY{o}{(}\PY{n+nv}{P}~\PY{o}{:}~\PY{k+kr}{forall}~\PY{n+nv}{n}~\PY{o}{:}~\PY{n}{nat}\PY{o}{,}~\PY{n}{PrimRec}~\PY{n}{n}~\PY{o}{\PYZhy{}\PYZgt{}}~\PY{k+kt}{Prop}\PY{o}{)}\nl
          ~~~~~~~~~\PY{o}{(}\PY{n+nv}{P0}~\PY{o}{:}~\PY{k+kr}{forall}~\PY{n+nv}{n}~\PY{n+nv}{n0}~\PY{o}{:}~\PY{n}{nat}\PY{o}{,}\nl
          ~~~~~~~~~~~~~~~\PY{n}{PrimRecs}~\PY{n}{n}~\PY{n}{n0}~\PY{o}{\PYZhy{}\PYZgt{}}~\PY{k+kt}{Prop}\PY{o}{),}\nl
          ~~~~~~~\PY{n}{P}~\PY{l+m+mi}{1}~\PY{n}{succFunc}~\PY{o}{\PYZhy{}\PYZgt{}}\nl
          ~~~~~~~\PY{n}{P}~\PY{l+m+mi}{0}~\PY{n}{zeroFunc}~\PY{o}{\PYZhy{}\PYZgt{}}\nl
          ~~~~~~~\PY{o}{(}\PY{k+kr}{forall}~\PY{o}{(}\PY{n+nv}{n}~\PY{n+nv}{m}~\PY{o}{:}~\PY{n}{nat}\PY{o}{)}~\PY{o}{(}\PY{n+nv}{l}~\PY{o}{:}~\PY{n}{m}~\PY{o}{\PYZlt{}}~\PY{n}{n}\PY{o}{),}\nl
          ~~~~~~~~\PY{n}{P}~\PY{n}{n}~\PY{o}{(}\PY{n}{projFunc}~\PY{n}{n}~\PY{n}{m}~\PY{n}{l}\PY{o}{))}~\PY{o}{\PYZhy{}\PYZgt{}}\nl
          ~~~~~~~\PY{o}{(}\PY{k+kr}{forall}~\PY{o}{(}\PY{n+nv}{n}~\PY{n+nv}{m}~\PY{o}{:}~\PY{n}{nat}\PY{o}{)}~\PY{o}{(}\PY{n+nv}{g}~\PY{o}{:}~\PY{n}{PrimRecs}~\PY{n}{n}~\PY{n}{m}\PY{o}{),}\nl
          ~~~~~~~~\PY{n}{P0}~\PY{n}{n}~\PY{n}{m}~\PY{n}{g}~\PY{o}{\PYZhy{}\PYZgt{}}\nl
          ~~~~~~~~\PY{k+kr}{forall}~\PY{n+nv}{h}~\PY{o}{:}~\PY{n}{PrimRec}~\PY{n}{m}\PY{o}{,}\nl
          ~~~~~~~~\PY{n}{P}~\PY{n}{m}~\PY{n}{h}~\PY{o}{\PYZhy{}\PYZgt{}}~\PY{n}{P}~\PY{n}{n}~\PY{o}{(}\PY{n}{composeFunc}~\PY{n}{n}~\PY{n}{m}~\PY{n}{g}~\PY{n}{h}\PY{o}{))}~\PY{o}{\PYZhy{}\PYZgt{}}\nl
          ~~~~~~~\PY{o}{(}\PY{k+kr}{forall}~\PY{o}{(}\PY{n+nv}{n}~\PY{o}{:}~\PY{n}{nat}\PY{o}{)}~\PY{o}{(}\PY{n+nv}{g}~\PY{o}{:}~\PY{n}{PrimRec}~\PY{n}{n}\PY{o}{),}\nl
          ~~~~~~~~\PY{n}{P}~\PY{n}{n}~\PY{n}{g}~\PY{o}{\PYZhy{}\PYZgt{}}\nl
          ~~~~~~~~\PY{k+kr}{forall}~\PY{n+nv}{h}~\PY{o}{:}~\PY{n}{PrimRec}~\PY{o}{(}\PY{n}{S}~\PY{o}{(}\PY{n}{S}~\PY{n}{n}\PY{o}{)),}\nl
          ~~~~~~~~\PY{n}{P}~\PY{o}{(}\PY{n}{S}~\PY{o}{(}\PY{n}{S}~\PY{n}{n}\PY{o}{))}~\PY{n}{h}~\PY{o}{\PYZhy{}\PYZgt{}}~\PY{n}{P}~\PY{o}{(}\PY{n}{S}~\PY{n}{n}\PY{o}{)}~\PY{o}{(}\PY{n}{primRecFunc}~\PY{n}{n}~\PY{n}{g}~\PY{n}{h}\PY{o}{))}~\PY{o}{\PYZhy{}\PYZgt{}}\nl
          ~~~~~~~\PY{o}{(}\PY{k+kr}{forall}~\PY{n+nv}{n}~\PY{o}{:}~\PY{n}{nat}\PY{o}{,}~\PY{n}{P0}~\PY{n}{n}~\PY{l+m+mi}{0}~\PY{o}{(}\PY{n}{PRnil}~\PY{n}{n}\PY{o}{))}~\PY{o}{\PYZhy{}\PYZgt{}}\nl
          ~~~~~~~\PY{o}{(}\PY{k+kr}{forall}~\PY{o}{(}\PY{n+nv}{n}~\PY{n+nv}{m}~\PY{o}{:}~\PY{n}{nat}\PY{o}{)}~\PY{o}{(}\PY{n+nv}{p}~\PY{o}{:}~\PY{n}{PrimRec}~\PY{n}{n}\PY{o}{),}\nl
          ~~~~~~~~\PY{n}{P}~\PY{n}{n}~\PY{n}{p}~\PY{o}{\PYZhy{}\PYZgt{}}\nl
          ~~~~~~~~\PY{k+kr}{forall}~\PY{n+nv}{p0}~\PY{o}{:}~\PY{n}{PrimRecs}~\PY{n}{n}~\PY{n}{m}\PY{o}{,}\nl
          ~~~~~~~~\PY{n}{P0}~\PY{n}{n}~\PY{n}{m}~\PY{n}{p0}~\PY{o}{\PYZhy{}\PYZgt{}}~\PY{n}{P0}~\PY{n}{n}~\PY{o}{(}\PY{n}{S}~\PY{n}{m}\PY{o}{)}~\PY{o}{(}\PY{n}{PRcons}~\PY{n}{n}~\PY{n}{m}~\PY{n}{p}~\PY{n}{p0}\PY{o}{))}~\PY{o}{\PYZhy{}\PYZgt{}}\nl
          ~~~~~~~\PY{k+kr}{forall}~\PY{o}{(}\PY{n+nv}{n}~\PY{o}{:}~\PY{n}{nat}\PY{o}{)}~\PY{o}{(}\PY{n+nv}{p}~\PY{o}{:}~\PY{n}{PrimRec}~\PY{n}{n}\PY{o}{),}~\PY{n}{P}~\PY{n}{n}~\PY{n}{p}
        \end{message}
      \end{messages}
    \end{output}
  \end{sentence}
\end{alectryon}

### 2022-01-06 doc/JFLA2022/Slides/Slides.tex
small correction in the slides

@@
 \definecolor{darkblue}{rgb}{0.0,0.0,0.6}
-
+\definecolor{cyan}{rgb}{0.0,0.4,0.8}
 \usepackage{tikz}
@@
 
-\usepackage[firstpage]{draftwatermark} 
+%\usepackage[firstpage]{draftwatermark} 
 \setbeamercolor{background canvas}{bg=}%transparent canvas
@@
       \item \textcolor{mathcolor}{$\forall i, i \leq H'_\alpha(i)$}
-      \item   If \textcolor{mathcolor}{$\beta<\alpha$}, then \textcolor{mathcolor}{$H'_\alpha(i)>H'_\beta(i)$} for all but a finite number of $i$ (we say that \textcolor{mathcolor}{$H'_\alpha$}
+      \item   If \textcolor{mathcolor}{$\beta<\alpha$}, then \textcolor{mathcolor}{$H'_\beta(i)<H'_\alpha(i)$} for all but a finite number of $i$ (we say that \textcolor{mathcolor}{$H'_\alpha$}
         \emph{dominates} \textcolor{mathcolor}{$H_\beta$}).
@@
     \item \gaia (José Grimm) : Formalization of Bourbaki's Elements of Mathematics.
-{\color{lookcolor}\emph{This library uses an ordinal representation compatible with ours.
+{\color{lookcolor}\emph{This library uses an ordinal representation compatible with ours (from the \textcolor{plugincolor}{Cantor} contrib) .
 It is  written in SSReflect/Mathcomp, and is also maintained in \community}.}
@@
   \begin{block}{Call to contributors }
-  {\color{plugincolor}\Hydras} and  {\color{plugincolor}\community} are on GitHub. Please contribute!
+  {\color{plugincolor}\Hydras} and  {\color{plugincolor}\community} are on GitHub. {\color{cyan}Please contribute, reuse!}
       \begin{itemize}

### 2022-01-10 doc/gaia-chapter.tex
experiments in bridge building

@@
 
+\section{Well formed ordinals}
+For sake of compatibility, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}.
+
+\inputsnippets{T1Bridge/gE0Def}
+\inputsnippets{T1Bridge/E0EqP}
+
+In order to import definitions and lemmas
+from~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}, we define a pair of translations.
+
+\inputsnippets{T1Bridge/E0iotaPi}
+
+These tools allow us to import \texttt{E0.lt}'s  well-foundedness.
+
+\inputsnippets{T1Bridge/gE0LtWf}
+
 \section{Gaia's proof of well-foundedness}
@@
 the proof in Section~\ref{sec:strongly-accessible}.
+
+\begin{remark}
+  The following snippets have been generated, not from
+  \href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}, but on a provisional copy:
+\url{../theories/gaia/nfgaia.v} we made in order to insert comments for \alectr.
+This copy is planned to be removed when we are able to add \alectr directives to a \coq source we are not allowed to modify.
+\end{remark}
 

### 2022-01-14 doc/gaia-chapter.tex
fix a typo in gaia-chapter.tex
first try
utf8 omega accepted

@@
 \subsubsection{A few constants}
-For each constant: $0$, $1$, \dots, $n$ and $\omega$ we prove
+For each constant: $0$, $1$, \dots, $n$ and ω, ω again, we prove
 that the \texttt{hydras}' definition refines \texttt{gaia}'s.
 \inputsnippets{T1Bridge/constantRefs}
+
+\inputsnippets{T1Bridge/utf8try}
 
@@
 
-\inputsnippets{T1Bridge/gE0Def}
+\inputsnippets{T1Bridge/E0Def}
 \inputsnippets{T1Bridge/E0EqP}
@@
 
-\inputsnippets{T1Bridge/E0iotaPi}
+\inputsnippets{T1Bridge/gE0iotaPi}
 

### 2022-01-15 doc/gaia-chapter.tex
try to use DejaVu Sans Mono

@@
 \subsubsection{A few constants}
-For each constant: $0$, $1$, \dots, $n$ and ω, ω again, we prove
-that the \texttt{hydras}' definition refines \texttt{gaia}'s.
+For each constant: $0$, $1$, \dots, $n$ and $\omega$, we prove
+that \texttt{hydras}' definition refines \texttt{gaia}'s.
 \inputsnippets{T1Bridge/constantRefs}

### 2022-01-29 doc/JFLA2022/Slides/multA.tex
update slides about Gaia

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Lemma}~\PY{n+nf}{mult\PYZus{}refR}~\PY{o}{(}\PY{n+nv}{a}~\PY{n+nv}{b}~\PY{o}{:}~\PY{n}{T1}\PY{o}{)}~\PY{o}{:}~\PY{n}{hmult}~\PY{o}{(}\PY{n}{g2h}~\PY{n}{a}\PY{o}{)}~\PY{o}{(}\PY{n}{g2h}~\PY{n}{b}\PY{o}{)}~\PY{o}{=}~\PY{n}{g2h}~\PY{o}{(}\PY{n}{a}~\PY{o}{*}~\PY{n}{b}\PY{o}{).}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Proof}\PY{o}{.}~
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{n+nb}{apply}~\PY{n}{refines2\PYZus{}R}\PY{o}{,}~~\PY{n}{mult\PYZus{}ref}\PY{o}{.}~
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Qed}\PY{o}{.}~\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{txt}
    \nl
  \end{txt}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Lemma}~\PY{n+nf}{hmultA}~\PY{o}{:}~\PY{n}{associative}~\PY{n}{hmult}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Proof}\PY{o}{.}~\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~\PY{n+nb}{move}~\PY{o}{=\PYZgt{}}~\PY{n}{a}~\PY{n}{b}~\PY{n}{c}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~\PY{n+nb+bp}{by}~\PY{n+nb}{rewrite}~\PY{o}{\PYZhy{}(}\PY{n}{g2h\PYZus{}h2g}~\PY{n}{a}\PY{o}{)}~\PY{o}{\PYZhy{}(}\PY{n}{g2h\PYZus{}h2g}~\PY{n}{b}\PY{o}{)}~\PY{o}{\PYZhy{}(}\PY{n}{g2h\PYZus{}h2g}~\PY{n}{c}\PY{o}{)}~\PY{o}{!}\PY{n}{mult\PYZus{}refR}~\PY{n}{mulA}\PY{o}{.}~\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Qed}\PY{o}{.}
    \end{input}
  \end{sentence}
\end{alectryon}

### 2022-01-30 doc/gaia-chapter.tex
add a few lemmas
fix a typo
Import lemmas from theories/ordinal/Epsilon0/Canon (clumsy, partial) (#115)
* Import lemmas from theories/ordinal/Epsilon0/Canon (clumsy, partial)

* minor changes (SSreflect style)

* simplification of rewritings from hydras to gaia (init)

* fix an error in gaia-chapter.tex

* Simplify some ports in Bridge

* minor changes
new snippets for gaia-hydras

@@
 \section{Introduction}
-The \gaia project by Jos\'e Grimm aimed to formalize mathematics in \coq in the style of Nicolas Bourbaki.
-The formalization of the first book in the Elements of Mathematics series by Bourbaki, on the theory of sets, was initially described in a technical report in July 2009~\cite{Grimm2009a}.
-The set-theoretic axioms and basic definitions in \gaia were derived from an earlier development by Carlos Simpson~\cite{Simpson2004,CatsZFCContrib}.
-Grimm then wrote (and continually updated) technical reports describing the formalization of Bourbaki's two subsequent books~\cite{Grimm2009b,Grimm2016} and additional topics in number theory~\cite{grimm:hal-00911710,Grimm2014}, before he passed away in 2019.
+The \gaia project by Jos\'e Grimm aimed to formalize mathematics in \coq in
+the style of Nicolas Bourbaki. The formalization of the first book in the
+Elements of Mathematics series by Bourbaki, on the theory of sets, was
+initially described in a technical report in July 2009~\cite{Grimm2009a}.
+The set-theoretic axioms and basic definitions in \gaia were derived
+from an earlier development by Carlos Simpson~\cite{Simpson2004,CatsZFCContrib}.
+Grimm then wrote (and continually updated) technical reports describing the
+formalization of Bourbaki's two subsequent books~\cite{Grimm2009b,Grimm2016}
+and additional topics in number theory~\cite{grimm:hal-00911710,Grimm2014},
+before he passed away in 2019.
 
-In 2020, members of \community transferred the \gaia source code to GitHub and adapted it for recent releases of the Mathematical Components library, which \gaia heavily relies on.
-Anonymous volunteers (``collaborators of Nicolas Bourbaki'') then finished the only in-progress proof left by Grimm.
-At around 155,000 LOC, \gaia is currently one of the largest maintained open source \coq projects~\cite{Gaia}.
+In 2020, members of \community transferred the \gaia source code to
+GitHub and adapted it for recent releases of the Mathematical Components
+library, which \gaia heavily relies on.
+Anonymous volunteers (``collaborators of Nicolas Bourbaki'') then finished
+the only in-progress proof left by Grimm. At around 155,000 LOC, \gaia is currently one of the largest maintained open source \coq projects~\cite{Gaia}.
 
@@
 
-As an initial step in bridging \gaia and \Hydras, we were able to establish a correspondence between both implementations of ordinals through rewriting lemmas.
-For instance, we proved that multiplication of ordinals in Cantor normal form in \Hydras refine \gaia's corresponding multiplication operation, and then imported \gaia's proof of associativity of the multiplication almost for free.
+As an initial step in bridging \gaia and \Hydras, we were able to establish a
+correspondence between both implementations of ordinals through rewriting
+lemmas. For instance, we proved that multiplication of ordinals in Cantor
+normal form in \Hydras refine \gaia's corresponding multiplication operation,
+and then imported \gaia's proof of associativity of the multiplication
+almost for free.
 
-The initial draft bridge code mostly uses the SSReflect proof language and idioms from the Mathematical Components library.
-We made this design decision since we believe it is less challenging to reason about Hydra-battles code using SSReflect and MathComp than to reason about Gaia without SSReflect.
+The initial draft bridge code mostly uses the SSReflect proof language
+and idioms from the Mathematical Components library. We made this design
+decision since we believe it is less challenging to reason about
+Hydra-battles code using SSReflect and MathComp than to reason about
+Gaia without SSReflect.
 
@@
 
-\inputsnippets{T1Bridge/iotaPiDef}
+\inputsnippets{T1Bridge/h2gG2hDef}
 
-\inputsnippets{T1Bridge/iotaPiRw, T1Bridge/piIotaRw}
+\inputsnippets{T1Bridge/h2gG2hRw, T1Bridge/g2hH2gRw}
 
@@
 
-\inputsnippets{T1Bridge/utf8try}
+%\inputsnippets{T1Bridge/utf8try}
 
@@
 \inputsnippets{T1Bridge/plusRef, T1Bridge/multRef}
-
-\subsubsection{Porting a theorem}
-The \texttt{Gaia} library containsa proof that the multiplication of ordinals less than $\epsilon_0$ is associative.
-
-\emph{From~\href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}}
-
-\begin{Coqsrc}
-Lemma mulA: associative T1mul.
-\end{Coqsrc}
-
-This lemma was missing from \texttt{hydra-battles}. Nevertheless, we could adapt this lemma to \texttt{hydra-battles}' ordinals, by a small sequence of rewritings.
-
-\inputsnippets{T1Bridge/multA}
-
-The module~\href{../theories/html/gaia_hydras.T1Bridge_use.html}%
-{\texttt{gaia\_hydras.T1Bridge\_use}} shows a small
-example of importation of \texttt{multA} into \texttt{hydra-battles}' world.
-
-\inputsnippets{T1Bridge_use/T1BridgeUse}
 
@@
 
-\inputsnippets{T1Bridge/gE0iotaPi}
+\inputsnippets{T1Bridge/gE0h2gG2h}
 
@@
 \inputsnippets{T1Bridge/gE0LtWf}
+
+\section{Porting theorems without big effort}
+
+\subsection{Porting a theorem from \gaia}
+The \texttt{Gaia} library contains a proof that the multiplication of ordinals less than $\epsilon_0$ is associative.
+
+\emph{From~\href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}}
+
+\begin{Coqsrc}
+Lemma mulA: associative T1mul.
+\end{Coqsrc}
+
+This lemma was missing from \texttt{hydra-battles}. Nevertheless, we could adapt this lemma to \texttt{hydra-battles}' ordinals, by a small sequence of rewritings.
+
+\inputsnippets{T1Bridge/multA}
+
+The module~\href{../theories/html/gaia_hydras.T1Bridge_use.html}%
+{\texttt{gaia\_hydras.T1Bridge\_use}} shows a small
+example of importation of \texttt{multA} into \texttt{hydra-battles}' world.
+
+\inputsnippets{T1Bridge_use/T1BridgeUse}
+
+\subsection{Porting theorems from \Hydras}
+
+\subsubsection{Canonical sequences}
+
+In~\href{../theories/html/gaia_hydras.GCanon.html}%
+{\texttt{gaia\_hydras.GCanon}}, canonical sequences are just imported from the library \linebreak
+\href{../theories/html/hydras.Epsilon0.Canon.html}%
+{\texttt{hydras.Epsilon0.Canon}}.
+
+First, we rename \Hydras's \texttt{canon} function into
+\texttt{hcanon}. Then we define a new function \emph{via} the
+translations \texttt{g2h} and \texttt{h2g}.
+
+\inputsnippets{GCanon/canonDef}
+
+\begin{todo}
+Explain the proof !
+\end{todo}
+
+\inputsnippets{GCanon/gcanonLimitOf}
 

### 2022-01-30 doc/JFLA2022/Slides/canonDef.tex
Update slides (#118)
* update slides about Gaia

* new snippets for gaia-hydras

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Definition}~\PY{n+nf}{canon}~\PY{o}{(}\PY{n+nv}{a}\PY{o}{:}~\PY{n}{T1}\PY{o}{)}~\PY{o}{(}\PY{n+nv}{i}\PY{o}{:}\PY{n}{nat}\PY{o}{)}~\PY{o}{:}~\PY{n}{T1}~\PY{o}{:=}\nl
      ~~\PY{n}{h2g}~\PY{o}{(}\PY{n}{hcanon}~\PY{o}{(}\PY{n}{g2h}~\PY{n}{a}\PY{o}{)}~\PY{n}{i}\PY{o}{).}
    \end{input}
  \end{sentence}
\end{alectryon}

### 2022-01-30 doc/JFLA2022/Slides/gcanonLimitOf.tex
Update slides (#118)
* update slides about Gaia

* new snippets for gaia-hydras

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      ~\PY{k+kn}{Lemma}~\PY{n+nf}{gcanon\PYZus{}limit\PYZus{}of}~\PY{n+nv}{lambda}~\PY{o}{(}\PY{n+nv}{Hnf}~\PY{o}{:}~\PY{n}{T1nf}~\PY{n}{lambda}\PY{o}{)}~\PY{o}{(}\PY{n+nv}{Hlim}~\PY{o}{:}~\PY{n}{T1limit}~\PY{n}{lambda}\PY{o}{)}~\PY{o}{:}\nl
      ~~~\PY{n}{limit\PYZus{}of}~\PY{o}{(}\PY{n}{canon}~\PY{n}{lambda}\PY{o}{)}~\PY{n}{lambda}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~\PY{k+kn}{Proof}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~~\PY{n+nb}{split}~\PY{o}{=\PYZgt{}}~\PY{o}{//.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~~\PY{n+nb}{move}~\PY{o}{=\PYZgt{}}~\PY{n}{n}~\PY{n}{m}~\PY{n}{Hnm}\PY{o}{;}~\PY{n+nb}{apply}~\PY{n}{gcanon\PYZus{}limit\PYZus{}mono}~\PY{o}{=\PYZgt{}}~\PY{o}{//.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~~\PY{n+nb}{apply}~\PY{n}{gcanon\PYZus{}limit\PYZus{}v2}~\PY{o}{=\PYZgt{}}~\PY{o}{//.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~\PY{k+kn}{Qed}\PY{o}{.}~~~~
    \end{input}
  \end{sentence}
\end{alectryon}

### 2022-01-31 doc/gaia-chapter.tex
add more snippets
add stuff in gaia-chapter

@@
 
-\section{Well formed ordinals}
-For sake of compatibility, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}.
-
-\inputsnippets{T1Bridge/E0Def}
-\inputsnippets{T1Bridge/E0EqP}
-
-In order to import definitions and lemmas
-from~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}, we define a pair of translations.
-
-\inputsnippets{T1Bridge/gE0h2gG2h}
-
-These tools allow us to import \texttt{E0.lt}'s  well-foundedness.
-
-\inputsnippets{T1Bridge/gE0LtWf}
-
-\section{Porting theorems without big effort}
-
 \subsection{Porting a theorem from \gaia}
-The \texttt{Gaia} library contains a proof that the multiplication of ordinals less than $\epsilon_0$ is associative.
-
+The \texttt{Gaia} library already contains a proof that the multiplication of ordinals less than $\epsilon_0$ is associative.
 \emph{From~\href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}}
@@
 
+Many lemmas of the library \texttt{ordinals/Epsilon0/} are statements about inhabitants of type \texttt{hT1}. In order to get
+similar statements on \gaia's type \texttt{T1}, we prove and
+apply rewriting rules of the following forms:
+
+\vspace{4pt}
+
+\noindent\emph{From Module~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge}}
+
+\inputsnippets{T1Bridge/T1leIff}
+\inputsnippets{T1Bridge/rewritingRules}
+\inputsnippets{T1Bridge/limitbRef}
+
+Please note that in this module, we favour \gaia's notations over hydra's.
+A snippet like \texttt{alpha < succ beta} should be interpreted as \texttt{T1lt alpha beta}, where \texttt{T1lt} is defined in \texttt{gaia.ssete9.v}.
+The corresponding statement is
+written \texttt{hlt (g2h alpha) (g2h beta)}.
+
 \subsubsection{Canonical sequences}
@@
 
-\begin{todo}
-Explain the proof !
-\end{todo}
+The following lemmas are proved by rewriting from corresponding statements proved in \Hydras.
 
+\inputsnippets{GCanon/g2hCanonRw}
+\inputsnippets{GCanon/gcanonLt}
+\inputsnippets{GCanon/glimitCanonSNotZero}
+\inputsnippets{GCanon/gcanonLimitV2}
+\inputsnippets{GCanon/gcanonLimitMono}
 \inputsnippets{GCanon/gcanonLimitOf}
@@
 In every case, the considered ordinal is accessible. By the definition of accessibility, the ordinal \texttt{(cons a n b)} is accessible, which ends the proof.
+
+\section{Well formed ordinals}
+For sake of compatibility, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}.
+
+\inputsnippets{T1Bridge/E0Def}
+\inputsnippets{T1Bridge/E0EqP}
+
+In order to import definitions and lemmas
+from~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}, we define a pair of translations.
+
+\inputsnippets{T1Bridge/gE0h2gG2h}
+
+These tools allow us to import \texttt{E0.lt}'s  well-foundedness.
+
+\inputsnippets{T1Bridge/gE0LtWf}

### 2022-02-02 doc/gaia-chapter.tex
minor changes

@@
 
+From now on, names like  \texttt{T1}, \texttt{omega}, etc. are  bound to \Hydras's meaning.
+
+ \inputsnippets{ T1Bridge_use/LocateT1, T1Bridge_use/T1BridgeUseb}
+
 \subsection{Porting theorems from \Hydras}
@@
 Please note that in this module, we favour \gaia's notations over hydra's.
-A snippet like \texttt{alpha < succ beta} should be interpreted as \texttt{T1lt alpha beta}, where \texttt{T1lt} is defined in \texttt{gaia.ssete9.v}.
-The corresponding statement is
+A snippet like \texttt{alpha < succ beta} should be interpreted as \texttt{T1lt alpha beta}, where \texttt{T1lt} is defined in ~\href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}.
+
+\texttt{gaia.ssete9.v}. The corresponding statement is
 written \texttt{hlt (g2h alpha) (g2h beta)}.
@@
 The following lemmas are proved by rewriting from corresponding statements proved in \Hydras.
+\emph{The author of these proof scripts is still a beginner in Ssreflect. Please forgive the clumsiness of the scripts below.}
 

### 2022-02-09 doc/gaia-chapter.tex
re-redaction of chapter 10 (start)

@@
-\chapter{Gaia and the hydras (draft)}
+\chapter{Gaia and the hydras (experimental)}
 \index{coq}{Plug-ins!Gaia}
@@
 
-As an initial step in bridging \gaia and \Hydras, we were able to establish a
-correspondence between both implementations of ordinals through rewriting
-lemmas. For instance, we proved that multiplication of ordinals in Cantor
-normal form in \Hydras refine \gaia's corresponding multiplication operation,
-and then imported \gaia's proof of associativity of the multiplication
-almost for free.
+The bridge we aim to build and cross is a two-way bridge:
 
-The initial draft bridge code mostly uses the SSReflect proof language
+\begin{itemize}
+\item From \gaia, we import the proof of associativity of multiplication of ordinals strictly less than $\epsilon_0$, which was missing from \Hydras.
+  \item The other way, we started to import definitions and properties of canonical sequences and rapidly growing functions from \Hydras (see Chapter~\ref{chap:ketonen} and\ref{chap:alpha-large}~of this book).
+\end{itemize}
+
+Please note that by ``import'' we do not mean ``duplicate the proofs''.  Whenever possible, we proceed by rewriting steps over the statements and not the proofs.
+
+This initial draft bridge code mostly uses the SSReflect proof language
 and idioms from the Mathematical Components library. We made this design
@@
 
-\section{Data refinement}
+\section{Library structure}
+The \gaiaHydras library is designed for use  by users of
+\gaia and/or  \HydrasLib. Whenever possible, we managed to respect the notations and naming conventions of these libraries.
 
-\emph{The following definitions and lemmas are in
+In case of name clash, we often added a one-letter prefix to names : \texttt{g} or \texttt{G} for \gaia, \texttt{h} or \texttt{H}
+  for \HydrasLib.
+
+For instance, both \gaia and \HydrasLib have a type called \texttt{T1} (data type for Cantor normal forms).
+\begin{itemize}
+\item In
+  \href{../theories/html/gaia_hydras.T1Bridge_use.html}%
+  {\texttt{gaia\_hydras.T1Bridge\_use}}, we adopt \HydrasLib's point of view, and the name \texttt{T1} is by default bound to
+\HydrasLib' data structures.
+
+\inputsnippets{T1Bridge_use/T1BridgeUse,
+  T1Bridge_use/LocateT1}
+
+Associativity of multiplication over \HydrasLib' type \texttt{T1},
+although imported from \gaia, is expressed with \HydrasLib' notations.
+
+\inputsnippets{T1Bridge_use/T1BridgeUseb}
+
+\item The other way, the module \href{../theories/html/gaia_hydras.GCanon.html}%
+  {\texttt{gaia\_hydras.GCanon}} is an adaptation to \gaia of
+  \href{../theories/html/hydras.Epsilon0/Canon.html}%
+  {\texttt{hydras.Epsilon0/Canon}}.
+  In this module, the name \texttt{T1} is bound by default to \gaia's meaning. \HydrasLib' name \texttt{T1} is renamed into \texttt{hT1}.
+
+  The following lemmas are expressed in \gaia's vocabulary.
+  \inputsnippets{GCanon/gcanonLimitMono,
+    GCanon/gcanonLimitOf}
+
+\end{itemize}
+
+The \gaiaHydras entry point is the file
 \href{../theories/html/gaia_hydras.T1Bridge.html}%
-{\texttt{gaia\_hydras.T1Bridge}}}.
+{\texttt{theories/gaia.T1Bridge.v}}, which requires modules both from \HydrasLib and \gaia. By default, we give priority to \gaia's notation and vocabulary.
 
-Since both libraries \texttt{gaia} and \texttt{hydra-battles} have their own type \texttt{T1}, we will work with the following notations.
+\inputsnippets{T1Bridge/Requirements}
 
 \inputsnippets{T1Bridge/hT1gT1}
+
+\subsection{Data refinement}
 
@@
 
-From now on, names like  \texttt{T1}, \texttt{omega}, etc. are  bound to \Hydras's meaning.
+From now on, names like  \texttt{T1}, \texttt{omega}, etc. are  bound to \HydrasLib' meaning.
 
@@
 
-\subsection{Porting theorems from \Hydras}
+\subsection{Porting theorems from \HydrasLib}
 
@@
 
-First, we rename \Hydras's \texttt{canon} function into
+First, we rename \HydrasLib' \texttt{canon} function into
 \texttt{hcanon}. Then we define a new function \emph{via} the
@@
 
-The following lemmas are proved by rewriting from corresponding statements proved in \Hydras.
+The following lemmas are proved by rewriting from corresponding statements proved in \HydrasLib.
 \emph{The author of these proof scripts is still a beginner in Ssreflect. Please forgive the clumsiness of the scripts below.}
@@
 \section{Gaia's proof of well-foundedness}
-The library \Hydras already contains two proofs
+The library \HydrasLib already contains two proofs
 of well-foundedness of the strict order on ordinal terms in
@@
 \emph{Notes: the abbreviation \texttt{LT}, and the \texttt{fold LT} and \texttt{unfold LT} tactics,are not in \texttt{gaia}, but  have been added to make the goal display more readable.
-The infix notation \texttt{`` \_ < \_''}  is bound to the order \texttt{T1lt} in \gaia, while,  in \Hydras,  it is bound to its restriction \texttt{LT} to terms in normal form. }
+The infix notation \texttt{`` \_ < \_''}  is bound to the order \texttt{T1lt} in \gaia, while,  in \HydrasLib,  it is bound to its restriction \texttt{LT} to terms in normal form. }
 

### 2022-02-10 doc/gaia-chapter.tex
revision of chapter 10 (continued)

@@
 
-Most used constants and functions will also get standardized
-non-ambiguous names.
+\subsubsection{Renamings}
+
+Most used constants and functions are also unambiguously renamed.
 
 \inputsnippets{T1Bridge/MoreNotations}
+
+Please note that in this module, we favour \gaia's notations over hydra's.
+A snippet like \texttt{alpha < succ beta} should be interpreted as \texttt{T1lt alpha beta}, where \texttt{T1lt} is defined in ~\href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{\texttt{gaia.ssete9.v}}.
+
+\subsubsection{Translation functions and data refinement}
 
@@
 
+The following cancel lemmas will be often used in order to simplify sub-terms of the form \texttt{g2h (h2g {\it t})} and \texttt{h2g (g2h {\it t})} which appear in many proofs by rewriting.
+
 \inputsnippets{T1Bridge/h2gG2hRw, T1Bridge/g2hH2gRw}
 
-The functions \texttt{iota} and \texttt{pi} allow us to define
-a notion of refinement for constants, functions, predicates and relations.
+\inputsnippets{T1Bridge/h2gEqIff, T1Bridge/g2hEqIff}
+
+Functions \texttt{h2g} and \texttt{g2h} allow us to define
+a notion of ``data-refinement''  for constants, functions, predicates and relations. The following definitions express that some
+constant, function, relation defined in \HydrasLib ``implements'' the same concept of \gaia.
 
 \inputsnippets{T1Bridge/refineDefs}
+
+%%% ICI
 
@@
 
-\subsubsection{Binary operations}
-The binary operations $+$ and $\times$ are defined the same way in both libraries. Please note that they require comparison of ordinal terms. So, the proof of the following lemmas applies
-compatibility lemmas like \texttt{compare\_ref} (see Section~\vref{sect:gaia-compare-ref}).
-
-\inputsnippets{T1Bridge/plusRef, T1Bridge/multRef}
-
-\subsection{Order and comparison}
+\subsubsection{Order and comparison}
 Leibniz equality and strict order on both libraries are
 compatible:
+
 \inputsnippets{T1Bridge/eqRef}
@@
 
-\inputsnippets{T1Bridge/ltRef}
+\inputsnippets{T1Bridge/ltRef, T1Bridge/leRef}
 
 \inputsnippets{T1Bridge/decideHLtRw}
+
+\subsubsection{Cantor normal form}
 
@@
 
-\subsection{Porting a theorem from \gaia}
+\subsubsection{Addition and multiplication}
+
+The binary operations $+$ and $\times$ are defined the same way in both libraries. Please note that they require comparison of ordinal terms. So, the proof of the following lemmas applies
+compatibility lemmas like \texttt{compare\_ref} (see Section~\vref{sect:gaia-compare-ref}).
+
+\inputsnippets{T1Bridge/plusRef, T1Bridge/multRef}
+
+\subsection{Importing a theorem from \gaia}
 The \texttt{Gaia} library already contains a proof that the multiplication of ordinals less than $\epsilon_0$ is associative.
@@
 
-From now on, names like  \texttt{T1}, \texttt{omega}, etc. are  bound to \HydrasLib' meaning.
+In the rest of this module, names like  \texttt{T1}, \texttt{omega}, etc. are  bound to \HydrasLib' meaning.
 
@@
 
-\subsection{Porting theorems from \HydrasLib}
+\subsection{Importing theorems from \HydrasLib}
 
@@
 \inputsnippets{T1Bridge/rewritingRules}
-\inputsnippets{T1Bridge/limitbRef}
+\inputsnippets{T1Bridge/T1ltIff}
 
-Please note that in this module, we favour \gaia's notations over hydra's.
-A snippet like \texttt{alpha < succ beta} should be interpreted as \texttt{T1lt alpha beta}, where \texttt{T1lt} is defined in ~\href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}.
-
-\texttt{gaia.ssete9.v}. The corresponding statement is
-written \texttt{hlt (g2h alpha) (g2h beta)}.
+\inputsnippets{T1Bridge/limitbRef, T1Bridge/isSuccRef}
 
@@
 
-\section{Gaia's proof of well-foundedness}
-The library \HydrasLib already contains two proofs
-of well-foundedness of the strict order on ordinal terms in
-Cantor normal form: a direct proof~\vref{sec:T1Wf} and a proof based on the recursive path ordering~\vref{remark:a3pat}, by
-\'Evelyne Contejean. We found it interesting to comment Gaia's proof, which uses tools already present in~\cite{CantorContrib}.
+% \section{Gaia's proof of well-foundedness}
+% The library \HydrasLib already contains two proofs
+% of well-foundedness of the strict order on ordinal terms in
+% Cantor normal form: a direct proof~\vref{sec:T1Wf} and a proof based on the recursive path ordering~\vref{remark:a3pat}, by
+% \'Evelyne Contejean. We found it interesting to comment Gaia's proof, which uses tools already present in~\cite{CantorContrib}.
 
-This proof is quite dense and qualified as ``a bit tricky''. We just present here its main steps with the help of the \alectr tool.
+% This proof is quite dense and qualified as ``a bit tricky''. We just present here its main steps with the help of the \alectr tool.
 
-The reader is invited to replay the full proof in  \href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}, and to compare with
-the proof in Section~\ref{sec:strongly-accessible}.
+% The reader is invited to replay the full proof in  \href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}, and to compare with
+% the proof in Section~\ref{sec:strongly-accessible}.
 
-\begin{remark}
-  The following snippets have been generated, not from
-  \href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}, but on a provisional copy:
-\url{../theories/gaia/nfgaia.v} we made in order to insert comments for \alectr.
+% \begin{remark}
+%   The following snippets have been generated, not from
+%   \href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}, but on a provisional copy:
+% \url{../theories/gaia/nfgaia.v} we made in order to insert comments for \alectr.
 This copy is planned to be removed when we are able to add \alectr directives to a \coq source we are not allowed to modify.
-\end{remark}
+% \end{remark}
 
-\subsection{Restricted recursion}
-As remarked in~\vref{sec:T1Wf}, the order \texttt{T1lt} on \texttt{T1} is not well-founded, because of terms not in normal form.
-But the \emph{restriction} of \texttt{T1lt} to terms in normal form \emph{is} well-founded.
+% \subsection{Restricted recursion}
+% As remarked in~\vref{sec:T1Wf}, the order \texttt{T1lt} on \texttt{T1} is not well-founded, because of terms not in normal form.
+% But the \emph{restriction} of \texttt{T1lt} to terms in normal form \emph{is} well-founded.
 The following section introduces a  vocabulary in order to reason about statements of the form  ``the restriction of the relation $R$ to a subset $P$ of $A$ is well founded''.
 
-\inputsnippets{nfwfgaia/restrictedRecursion}
+% \inputsnippets{nfwfgaia/restrictedRecursion}
 
-The following induction principle is just a variant of \texttt{well\_founded\_induction\_type} for relations defined with \texttt{restrict}.
+% The following induction principle is just a variant of \texttt{well\_founded\_induction\_type} for relations defined with \texttt{restrict}.
 
-\inputsnippets{nfwfgaia/restrictedRecursiona}
+% \inputsnippets{nfwfgaia/restrictedRecursiona}
 
-With $P=\texttt{T1nf}$, this principle will allow us to write proofs by transfinite induction.
+% With $P=\texttt{T1nf}$, this principle will allow us to write proofs by transfinite induction.
 
-\subsection{Main proof steps}
+% \subsection{Main proof steps}
 
-\emph{Notes: the abbreviation \texttt{LT}, and the \texttt{fold LT} and \texttt{unfold LT} tactics,are not in \texttt{gaia}, but  have been added to make the goal display more readable.
-The infix notation \texttt{`` \_ < \_''}  is bound to the order \texttt{T1lt} in \gaia, while,  in \HydrasLib,  it is bound to its restriction \texttt{LT} to terms in normal form. }
+% \emph{Notes: the abbreviation \texttt{LT}, and the \texttt{fold LT} and \texttt{unfold LT} tactics,are not in \texttt{gaia}, but  have been added to make the goal display more readable.
+% The infix notation \texttt{`` \_ < \_''}  is bound to the order \texttt{T1lt} in \gaia, while,  in \HydrasLib,  it is bound to its restriction \texttt{LT} to terms in normal form. }
 
-\inputsnippets{nfwfgaia/nfWfProofa}
-We start with a classic structural induction on the type \texttt{T1}.
+% \inputsnippets{nfwfgaia/nfWfProofa}
+% We start with a classic structural induction on the type \texttt{T1}.
 
-\inputsnippets{nfwfgaia/nfWfProofaa}
-The following \texttt{ssreflect} tactic  pushes a new hypothesis \texttt{Hb} which --- in the same terms as in Section~\vref{sec:strongly-accessible} --- tells that any ordinal less than $a$ is ``strongly accessible''.
+% \inputsnippets{nfwfgaia/nfWfProofaa}
+% The following \texttt{ssreflect} tactic  pushes a new hypothesis \texttt{Hb} which --- in the same terms as in Section~\vref{sec:strongly-accessible} --- tells that any ordinal less than $a$ is ``strongly accessible''.
 
-\inputsnippets{nfwfgaia/nfWfProofb}
+% \inputsnippets{nfwfgaia/nfWfProofb}
 
-The following sentence gives us the possibility of induct over \texttt{a}'s accessibility.
+% The following sentence gives us the possibility of induct over \texttt{a}'s accessibility.
 
-\inputsnippets{nfwfgaia/nfWfProofbb}
+% \inputsnippets{nfwfgaia/nfWfProofbb}
 
-By the following two forward steps, we show that $b$ is accessible, and will be able to prove properties by induction on $b$'s accessibility.
+% By the following two forward steps, we show that $b$ is accessible, and will be able to prove properties by induction on $b$'s accessibility.
 
-\inputsnippets{nfwfgaia/haveHc}
+% \inputsnippets{nfwfgaia/haveHc}
 
-\inputsnippets{nfwfgaia/haveHd}
+% \inputsnippets{nfwfgaia/haveHd}
 
-\inputsnippets{nfwfgaia/hdProved}
+% \inputsnippets{nfwfgaia/hdProved}
 
-Now, let us consider any ordinal term which is (by \texttt{LT}) less than \texttt{(cons a n b)}.
+% Now, let us consider any ordinal term which is (by \texttt{LT}) less than \texttt{(cons a n b)}.
 
-There are three cases to consider.
-\begin{enumerate}
-\item It is of the form \texttt{(cons a' n' b')} where \texttt{a'} is strictly less than \texttt{a}.
-  \inputsnippets{nfwfgaia/nfWfProofd}
+% There are three cases to consider.
+% \begin{enumerate}
+% \item It is of the form \texttt{(cons a' n' b')} where \texttt{a'} is strictly less than \texttt{a}.
+%   \inputsnippets{nfwfgaia/nfWfProofd}
 
-  Then, it is accessible (by hypothesis \texttt{Hb}).
+%   Then, it is accessible (by hypothesis \texttt{Hb}).
 
-  \item The considered term  is equal to
-    \texttt{(cons a n b'')}, where $b''<\omega^a$. 
-    \inputsnippets{nfwfgaia/nfWfProofg}
+%   \item The considered term  is equal to
+%     \texttt{(cons a n b'')}, where $b''<\omega^a$. 
+%     \inputsnippets{nfwfgaia/nfWfProofg}
 
-    Then we can replace \texttt{a''} with \texttt{a}, \texttt{n''} with \texttt{n}, and apply the hypotheses \texttt{Hd} and \texttt{He}.
+%     Then we can replace \texttt{a''} with \texttt{a}, \texttt{n''} with \texttt{n}, and apply the hypotheses \texttt{Hd} and \texttt{He}.
 
-\item
-  The term is equal to
-  \texttt{(cons a n'' b'')}, where $n''<n$ and $b''<\omega^a$.
+% \item
+%   The term is equal to
+%   \texttt{(cons a n'' b'')}, where $n''<n$ and $b''<\omega^a$.
 
-  \inputsnippets{nfwfgaia/nfWfProofe}
+%   \inputsnippets{nfwfgaia/nfWfProofe}
 
-  By \texttt{Hd} and \texttt{He}, and a few technical lemmas about normal forms,   the term  \texttt{(cons a n b'')} is accessible, and so is \texttt{(cons a n'' b'')}.
+%   By \texttt{Hd} and \texttt{He}, and a few technical lemmas about normal forms,   the term  \texttt{(cons a n b'')} is accessible, and so is \texttt{(cons a n'' b'')}.
 
-\end{enumerate}
+% \end{enumerate}
 
-In every case, the considered ordinal is accessible. By the definition of accessibility, the ordinal \texttt{(cons a n b)} is accessible, which ends the proof.
+% In every case, the considered ordinal is accessible. By the definition of accessibility, the ordinal \texttt{(cons a n b)} is accessible, which ends the proof.
 

### 2022-02-15 doc/gaia-chapter.tex
initialization

@@
 \section{Introduction}
-The \gaia project by Jos\'e Grimm aimed to formalize mathematics in \coq in
+The \gaia project~\cite{Gaia} by Jos\'e Grimm aimed to formalize mathematics in \coq  in
 the style of Nicolas Bourbaki. The formalization of the first book in the
@@
 Anonymous volunteers (``collaborators of Nicolas Bourbaki'') then finished
-the only in-progress proof left by Grimm. At around 155,000 LOC, \gaia is currently one of the largest maintained open source \coq projects~\cite{Gaia}.
+the only in-progress proof left by Grimm. At around 155,000 LOC, \gaia is currently one of the largest maintained open source \coq projects.
 
@@
 \begin{itemize}
-\item From \gaia, we import the proof of associativity of multiplication of ordinals strictly less than $\epsilon_0$, which was missing from \Hydras.
-  \item The other way, we started to import definitions and properties of canonical sequences and rapidly growing functions from \Hydras (see Chapter~\ref{chap:ketonen} and\ref{chap:alpha-large}~of this book).
+\item \gaia contains lemmas on ordinal arithmetic (\emph{e.g.} associativity of multiplication of ordinals strictly less than $\epsilon_0$), which were missing from \Hydras. 
+  \item The other way, \Hydras contains formal definitions and proofs of properties of canonical sequences and rapidly growing functions  (see Chapters~\ref{chap:ketonen} and~\ref{chap:alpha-large}~of this book).
   \end{itemize}
 
+  We plan to make accessible the first kind of lemmas to \HydrasLib' users and the second kind to \gaia's.
+For this purpose, we write modules dedicated to the importation of definitions and lemmas from each of both libraries into the other one.
 Please note that by ``import'' we do not mean ``duplicate the proofs''.  Whenever possible, we proceed by rewriting steps over the statements and not the proofs.
@@
 Hydra-battles code using SSReflect and MathComp than to reason about
-Gaia without SSReflect.
+Gaia without SSReflect \footnote{The author of these proof scripts is still a beginner in Ssreflect. Please forgive the clumsiness of the current proof scripts.}.
 
@@
 The \gaiaHydras library is designed for use  by users of
-\gaia and/or  \HydrasLib. Whenever possible, we managed to respect the notations and naming conventions of these libraries.
+\gaia and/or  \HydrasLib. Whenever possible, we managed to respect the notations and naming conventions of both libraries.
 
-In case of name clash, we often added a one-letter prefix to names : \texttt{g} or \texttt{G} for \gaia, \texttt{h} or \texttt{H}
+In case of name clash (for instance the type \texttt{T1} of ordinal terms below $\epsilon_0$), we often added a one-letter prefix to names : \texttt{g} or \texttt{G} for \gaia, \texttt{h} or \texttt{H}
 for \HydrasLib.
 
-For instance, both \gaia and \HydrasLib have a type called \texttt{T1} (data type for Cantor normal forms).
+In the following two examples, the bare name \texttt{T1} is given by default either  to \HydrasLib' world or \gaia's.
+
 \begin{itemize}
@@
 Please note that in this module, we favour \gaia's notations over hydra's.
-A snippet like \texttt{alpha < succ beta} should be interpreted as \texttt{T1lt alpha beta}, where \texttt{T1lt} is defined in ~\href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{\texttt{gaia.ssete9.v}}.
+A snippet like \texttt{(alpha < succ beta)} should be interpreted as \texttt{(T1lt alpha beta)}, where \texttt{T1lt} is defined in ~\href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{\texttt{gaia.ssete9.v}}.
 
@@
 \subsection{Examples of refinements}
-Both libraries define constants like $0$, $1$, $\omega$, and arithmetic functions: successor, addition, multiplication, and exponential of base $\omega$ (function $\phi_0$).
+Both of our libraries define constants like $0$, $1$, $\omega$, and arithmetic functions: successor, addition, multiplication, and exponential of base $\omega$ (function $\phi_0$).
 We prove consistency of these definitions.
@@
 
-\subsection{Importing a theorem from \gaia}
+\section{Importing a theorem from \gaia}
 The \texttt{Gaia} library already contains a proof that the multiplication of ordinals less than $\epsilon_0$ is associative.
@@
 
-\subsection{Importing theorems from \HydrasLib}
+\section{Importing theorems from \HydrasLib}
 
@@
 
-\subsubsection{Canonical sequences}
+\subsection{Canonical sequences}
 
@@
 {\texttt{hydras.Epsilon0.Canon}}.
-
 First, we rename \HydrasLib' \texttt{canon} function into
@@
 The following lemmas are proved by rewriting from corresponding statements proved in \HydrasLib.
-\emph{The author of these proof scripts is still a beginner in Ssreflect. Please forgive the clumsiness of the scripts below.}
 
@@
 \inputsnippets{GCanon/gcanonLimitOf}
+
+\subsection{A family of rapidly growing functions}
+
+The functions $F_\alpha$ are described in Section~\vref{sect:wainer}. The symbol \texttt{F\_} has type \texttt{E0 -> nat -> nat},
+where \texttt{E0} is the type of well formed ordinal terms less than $\epsilon_0$.
+\subsubsection{Well-formed ordinal terms}
+
+For sake of compatibility, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}.
+
+\inputsnippets{T1Bridge/E0Def}
+\inputsnippets{T1Bridge/E0EqP}
+
+In order to import definitions and lemmas
+from~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}, we define a pair of translations.
+
+\inputsnippets{T1Bridge/gE0h2gG2h}
+
+These tools allow us to import \texttt{E0.lt}'s  well-foundedness.
+
+\inputsnippets{T1Bridge/gE0LtWf}
+
+\subsubsection{Importing the \texttt{F\_alpha} family}
+
+We are now able to import \HydrasLib' $F_\alpha$ functions.
+
+\inputsnippets{GF_alpha/FAlphaDef}
+
+Mainly by rewriting, we import a few properties of the $F_\alpha$s.
+
+\inputsnippets{T1Bridge/MonoDef, GF_alpha/FAlphaProps}.
 
@@
 % In every case, the considered ordinal is accessible. By the definition of accessibility, the ordinal \texttt{(cons a n b)} is accessible, which ends the proof.
-
-\section{Well formed ordinals}
-For sake of compatibility, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}.
-
-\inputsnippets{T1Bridge/E0Def}
-\inputsnippets{T1Bridge/E0EqP}
-
-In order to import definitions and lemmas
-from~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}, we define a pair of translations.
-
-\inputsnippets{T1Bridge/gE0h2gG2h}
-
-These tools allow us to import \texttt{E0.lt}'s  well-foundedness.
-
-\inputsnippets{T1Bridge/gE0LtWf}

### 2022-02-15 doc/gaia-chapter.tex
small module restructuration

@@
 \item In
-  \href{../theories/html/gaia_hydras.T1Bridge_use.html}%
-  {\texttt{gaia\_hydras.T1Bridge\_use}}, we adopt \HydrasLib's point of view, and the name \texttt{T1} is by default bound to
+  \href{../theories/html/gaia_hydras.GaiaToHydra.html}%
+  {\texttt{gaia\_hydras.GaiaToHydra}}, we adopt \HydrasLib's point of view, and the name \texttt{T1} is by default bound to
 \HydrasLib' data structures.
 
-\inputsnippets{T1Bridge_use/T1BridgeUse,
-  T1Bridge_use/LocateT1}
+\inputsnippets{GaiaToHydra/T1BridgeUse,
+  GaiaToHydra/LocateT1}
 
@@
 
-\inputsnippets{T1Bridge_use/T1BridgeUseb}
+\inputsnippets{GaiaToHydra/T1BridgeUseb}
 
@@
 
-The module~\href{../theories/html/gaia_hydras.T1Bridge_use.html}%
-{\texttt{gaia\_hydras.T1Bridge\_use}} shows a small
+The module~\href{../theories/html/gaia_hydras.GaiaToHydra.html}%
+{\texttt{gaia\_hydras.GaiaToHydra}} shows a small
 example of importation of \texttt{multA} into \texttt{hydra-battles}' world.
 
-\inputsnippets{T1Bridge_use/T1BridgeUse}
+\inputsnippets{GaiaToHydra/T1BridgeUse}
 
@@
 
- \inputsnippets{ T1Bridge_use/LocateT1, T1Bridge_use/T1BridgeUseb}
+ \inputsnippets{GaiaToHydra/LocateT1, GaiaToHydra/T1BridgeUseb}
 

### 2022-02-16 doc/gaia-chapter.tex
first version of chapter 10

@@
 \index{coq}{Plug-ins!Gaia}
+\label{gaia-chapter}
 
@@
 \section{Importing a theorem from \gaia}
-The \texttt{Gaia} library already contains a proof that the multiplication of ordinals less than $\epsilon_0$ is associative.
+The \texttt{Gaia} library already contains many lemmas about
+ordinal arithmetic. In this section, we give two examples of
+porting such a lemma to \HydrasLib' vocabulary.
+
+\subsection{Associativity of ordinal multiplication (below \texorpdfstring{$\epsilon_0$}{epsilon\_0})}
+\gaia already contains a proof that the multiplication of ordinals less than $\epsilon_0$ is associative.
 \emph{From~\href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}}
@@
 
+ \subsection{Right Distributivity, etc.}
+ Likewise, we prove almost for free that ordinal multiplication is right distributive over addition (with \HydrasLib' definitions).
+
+ \inputsnippets{GaiaToHydra/distributivity}.
+
+We plan to automatize the most as possible this transfer of lemmas in this direction.
+
 \section{Importing theorems from \HydrasLib}
@@
 \inputsnippets{T1Bridge/limitbRef, T1Bridge/isSuccRef}
+
+\subsection{A type for well-formed ordinal terms}
+
+For sake of compatibility, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}.
+
+\inputsnippets{T1Bridge/E0Def}
+\inputsnippets{T1Bridge/E0EqP}
+
+In order to import definitions and lemmas
+from~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}, we define a pair of translations.
+
+\inputsnippets{T1Bridge/gE0h2gG2h}
+
+These tools allow us to import \texttt{E0.lt}'s  well-foundedness.
+
+\inputsnippets{T1Bridge/gE0LtWf}
 
@@
 \inputsnippets{GCanon/glimitCanonSNotZero}
+\inputsnippets{GCanon/gcanonLimitMono}
+
+Let us recall \gaia's definition of $\omega$-limit.
+
+\inputsnippets{nfwfgaia/LimitV2Def, nfwfgaia/LimitOFDef}
+
+We adapt theorems from \href{../theories/html/hydras.Epsilon0.Canon.html}%
+{\texttt{hydras.Epsilon0.Canon}} to \gaia's vocabulary.
+
+\inputsnippets{GCanon/gcanonLimitStrong}
 \inputsnippets{GCanon/gcanonLimitV2}
-\inputsnippets{GCanon/gcanonLimitMono}
 \inputsnippets{GCanon/gcanonLimitOf}
@@
 where \texttt{E0} is the type of well formed ordinal terms less than $\epsilon_0$.
-\subsubsection{Well-formed ordinal terms}
-
-For sake of compatibility, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}.
-
-\inputsnippets{T1Bridge/E0Def}
-\inputsnippets{T1Bridge/E0EqP}
-
-In order to import definitions and lemmas
-from~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}, we define a pair of translations.
-
-\inputsnippets{T1Bridge/gE0h2gG2h}
-
-These tools allow us to import \texttt{E0.lt}'s  well-foundedness.
-
-\inputsnippets{T1Bridge/gE0LtWf}
 
@@
 
-Mainly by rewriting, we import a few properties of the $F_\alpha$s.
+The following definitions adapt \HydrasLib' to \ssreflect's 
+definitions of $<$ and $\leq$ on \texttt{nat}.
 
-\inputsnippets{T1Bridge/MonoDef, GF_alpha/FAlphaProps}.
+\inputsnippets{T1Bridge/MonoDef}
+
+By rewriting, we import a few lemmas from
+~\href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}.
+
+\inputsnippets{GF_alpha/FAlphaProps}
+
+With the same technique, we prove that if
+$\alpha\geq\omega$, then $F_\alpha$ is not primitive recursive.
+
+\inputsnippets{GF_alpha/FAlphaNotPR}
 

### 2022-02-16 doc/part-hydras.tex
first version of chapter 10

@@
  type and proof.
+
+\subsection*{Compatibility with \gaia\cite{Gaia} (in progress)}
+The \gaia project~\cite{Gaia} by Jos\'e Grimm aimed to formalize mathematics in \coq  in
+the style of Nicolas Bourbaki.
+It contains many definitions and results about ordinal numbers.
+In Chapter~\ref{gaia-chapter}, we present some modules which allow \HydrasLib' users to apply lemmas proven in \gaia, and vice versa.
 

### 2022-02-16 doc/gaia-chapter.tex
Prep for release (#121)
* initialization

* small module restructuration

* small module restructuration

* first version of chapter 10

* first version of chapter 10

--- 2022-02-15 doc/gaia-chapter.tex
@@
 \begin{itemize}
-\item \gaia contains lemmas on ordinal arithmetic (\emph{e.g.} associativity of multiplication of ordinals strictly less than $\epsilon_0$), which were missing from \Hydras. 
-  \item The other way, \Hydras contains formal definitions and proofs of properties of canonical sequences and rapidly growing functions  (see Chapters~\ref{chap:ketonen} and~\ref{chap:alpha-large}~of this book).
+\item From \gaia, we import the proof of associativity of multiplication of ordinals strictly less than $\epsilon_0$, which was missing from \Hydras.
+  \item The other way, we started to import definitions and properties of canonical sequences and rapidly growing functions from \Hydras (see Chapter~\ref{chap:ketonen} and\ref{chap:alpha-large}~of this book).
 \end{itemize}
 
-  We plan to make accessible the first kind of lemmas to \HydrasLib' users and the second kind to \gaia's.
-For this purpose, we write modules dedicated to the importation of definitions and lemmas from each of both libraries into the other one.
 Please note that by ``import'' we do not mean ``duplicate the proofs''.  Whenever possible, we proceed by rewriting steps over the statements and not the proofs.
@@
 Hydra-battles code using SSReflect and MathComp than to reason about
-Gaia without SSReflect \footnote{The author of these proof scripts is still a beginner in Ssreflect. Please forgive the clumsiness of the current proof scripts.}.
+Gaia without SSReflect \footnote{The author of these proof scripts is still a beginner in Ssreflect. Please forgive the clumsiness of the current proof scripts.}
 

### 2022-02-17 doc/gaia-chapter.tex
small corrections

--- 2022-02-16 doc/gaia-chapter.tex
@@
 
-The following cancel lemmas will be often used in order to simplify sub-terms of the form \texttt{g2h (h2g {\it t})} and \texttt{h2g (g2h {\it t})} which appear in many proofs by rewriting.
+The following cancel lemmas will be often used in order to simplify sub-terms of the form (\texttt{g2h (h2g {\it t})}) and (\texttt{h2g (g2h {\it t})}) which appear in many proofs by rewriting.
 
@@
 
-%%% ICI
-
-\subsection{Examples of refinements}
+\subsection{Examples of refinement}
 Both of our libraries define constants like $0$, $1$, $\omega$, and arithmetic functions: successor, addition, multiplication, and exponential of base $\omega$ (function $\phi_0$).
-We prove consistency of these definitions.
+We prove that these definitions are mutually consistent.
 Finally, we prove that the boolean predicates `` to be in normal form'' are equivalent.
@@
 
+\subsection{Looking for a lemma}
+\coq's command \texttt{Search} and notation scopes allow you to explore both libraries.
+
+\inputsnippets{T1Bridge/SearchDemo}
+
 \section{Importing a theorem from \gaia}
@@
 
- \inputsnippets{GaiaToHydra/distributivity}.
+ \inputsnippets{GaiaToHydra/distributivity}
 
-We plan to automatize the most as possible this transfer of lemmas in this direction.
+We plan to automatize as soon as possible this kind of transfer of lemmas from \gaia to \HydrasLib.
 

### 2022-02-18 doc/part-hydras.tex
making gaia's T1 an instance of ON (doc)

@@
 \chapter{Introduction to ordinal numbers and ordinal notations}
+\label{chap:ON}
 
@@
 \end{remark}
+
+\section{Ordinal Notations and \gaia's ordinals}
+Module~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge} contains an instance of class \texttt{ON E0lt compare}, where \texttt{E0lt} is the order on the well-formed ordinal terms below $\epsilon_0$ defined in \gaia library (please see Chapter~\ref{gaia-chapter}).
 

### 2022-02-19 doc/gaia-chapter.tex
making gaia's T1 an instance of ON (doc)
extension of T1add and T1mul to type E0
add a new lemma to GCanon.v

@@
 
+\inputsnippets{T1Bridge/refines1R, T1Bridge/refines2R}
+
 \subsection{Examples of refinement}
@@
 \inputsnippets{T1Bridge/E0Def}
+\inputsnippets{T1Bridge/E0plusMultDef}
+
 \inputsnippets{T1Bridge/E0EqP}
@@
 \inputsnippets{T1Bridge/gE0LtWf}
+
+\subsection{E0 as an ordinal notation}
+The module \href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge} defines an instance of class \texttt{ON E0lt compare}
+(see Chapter~\ref{chap:ON}).
+
+\inputsnippets{T1Bridge/T1compareDef}
+\inputsnippets{ T1Bridge/T1compareCorrect}
+\inputsnippets{T1Bridge/E0compare}
+\inputsnippets{T1Bridge/E0Sto}
+\inputsnippets{T1Bridge/gEpsilon0}
+
+\inputsnippets{T1Bridge/ExampleComp}
 
@@
 \inputsnippets{GCanon/canonDef}
+
+\begin{todo}
+Give equations for \texttt{canon} (corresponding to the structural recursive definition in \HydrasLib.
+\end{todo}
 
@@
 \inputsnippets{GF_alpha/FAlphaDef}
+
+\begin{todo}
+  Give equations for \texttt{F\_} (corresponding to the equation
+  in \HydrasLib.
+\end{todo}
 

### 2022-02-21 doc/gaia-chapter.tex
add snippets to GCanon.v
Small changes in gaia-chapter; add an ON instance for Mathcomp finite ordinals

@@
 
-Associativity of multiplication over \HydrasLib' type \texttt{T1},
-although imported from \gaia, is expressed with \HydrasLib' notations.
-
-\inputsnippets{GaiaToHydra/T1BridgeUseb}
+Section~\ref{sect:gaia2hydra} shows how lemmas on arithmetic properties of ordinals strictly less than $\epsilon_0$ are  imported from \gaia , with \HydrasLib's notations.
 
@@
 \section{Importing a theorem from \gaia}
+\label{sect:gaia2hydra}
+
 The \texttt{Gaia} library already contains many lemmas about
@@
 The following lemmas are proved by rewriting from corresponding statements proved in \HydrasLib.
+\inputsnippets{GCanon/g2hCanonRw}
 
-\inputsnippets{GCanon/g2hCanonRw}
+\inputsnippets{GCanon/gcanonZero}
+\inputsnippets{GCanon/gcanonSucc}
+\inputsnippets{GCanon/gcanonSSn}
+\inputsnippets{GCanon/gcanonLim1}
+\inputsnippets{GCanon/gcanonLim2}
+\inputsnippets{GCanon/gcanonLim3}
+
 \inputsnippets{GCanon/gcanonLt}
@@
 \inputsnippets{GCanon/gcanonLimitMono}
+\inputsnippets{GCanon/glimitCanonSNotZero}
 

### 2022-02-21 doc/gaia-chapter.tex
Add MathComp's finite ordinals to ON instances

@@
 \section{Library structure}
-The \gaiaHydras library is designed for use  by users of
+The \gaiaHydras library is designed for users of
 \gaia and/or  \HydrasLib. Whenever possible, we managed to respect the notations and naming conventions of both libraries.
@@
 
-\begin{todo}
-Give equations for \texttt{canon} (corresponding to the structural recursive definition in \HydrasLib.
-\end{todo}
+\inputsnippets{GCanon/g2hCanonRw}
 
 The following lemmas are proved by rewriting from corresponding statements proved in \HydrasLib.
-\inputsnippets{GCanon/g2hCanonRw}
+The first six lemmas correspond to rewriting rules derived from the body of the definition of \texttt{Epsilon0.Canon.canon}.
 

### 2022-02-24 doc/gaia-chapter.tex
More uses of 'decide equality' tactic
add stuff in gaia-chapter.tex

@@
 
+\subsubsection{Canonical sequences and limits}
+
 Let us recall \gaia's definition of $\omega$-limit.
@@
 \inputsnippets{GCanon/gcanonLimitOf}
+
+\subsection{Accessibility in \texorpdfstring{$\epsilon_0$}{epsilon\_0}}
+
+The library \href{../theories/html/gaia.GPaths.html}%
+{\texttt{gaia.GPaths}} imports definitions and lemmas from
+\href{../theories/html/hydras.Epsilon0.Paths.html}%
+{\texttt{hydras.Epsilon0.Paths}}.
+
+\subsubsection{Remark}
+\label{sect:gaia-masking}
+
+First, we import some definitions. When we import some constant \emph{Id} from {\texttt{hydras.Epsilon0.Paths}} the name
+\emph{Id} is given to {\texttt{gaia.GPaths}}. In order to get
+{\texttt{hydras.Epsilon0.Paths}}'s sense, please refer it as
+\emph{Paths.Id}.
+Sometimes, we rename \emph{Paths.Id} into \emph{{\bf h}Id}.
+
+\subsubsection{Transitions}
+
+\inputsnippets{GPaths/importationsa}
+
+\subsubsection{Paths}
+
+\inputsnippets{GPaths/pathsDefs}
+
+\subsubsection{Main theorems about paths}
+
+If $\beta$ is (strictly) accessible from $\alpha$, then
+$\beta<\alpha$.
+
+\inputsnippets{GPaths/pathToLT}
+
+Constant index paths can be simulated with constant paths with larger index.
+
+\inputsnippets{GPaths/Cor12}
+
+If $\beta<\alpha$, then $\beta$ is accessible from $\alpha$
+through some constant-index path.
+
+\inputsnippets{GPaths/Lemma261}
+
+Any constant-index path can be simulated by a ``standard'' path (with index incremented by 1 at each step).
+
+\inputsnippets{GPaths/constantToStandard}
+
+As a corollary, we relate ordinal inequality to standard paths.
+
+\inputsnippets{GPaths/LTToStandard}
 

### 2022-02-26 doc/gaia-chapter.tex
improve redaction of gaia-chapter.tex

@@
 
-In case of name clash (for instance the type \texttt{T1} of ordinal terms below $\epsilon_0$), we often added a one-letter prefix to names : \texttt{g} or \texttt{G} for \gaia, \texttt{h} or \texttt{H}
-for \HydrasLib.
+The modules are located in {\texttt{theories/gaia/*.v}}.
+The main entry point is
+\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge}, which requires both Gaia's \texttt{ssete9} and a few modules of \texttt{theories/ordinals/Epsilon0}.
+
+Other modules are dedicated to the adaptation of several notions of \HydrasLib to \mathcomp's and \gaia's vocabulary and structures.
+By default, we give priority to \gaia's notation and vocabulary.
+
+\subsection{The \texttt{T1Bridge} Module}
+
+\texttt{T1Bridge} starts with a few requirements.
+\inputsnippets{T1Bridge/Requirements}
+
+Because of the common origin of \HydrasLib and \gaia, both libraries share the common name \texttt{T1} (data type of ordinal terms below $\epsilon_0$).
+As mentionned before, the bare name \texttt{T1} is given by priority to \gaia, and we prefix this name with an \texttt{h} (for ``hydras'') to refer to
+\href{../theories/html/Epsilon0.T1.html}{hydras.Epsilon0.T1.T1}.
+
+\inputsnippets{T1Bridge/hT1gT1}
+
+Many constants are common to \HydrasLib and \gaia. The following notations allow us to get rid of possible ambiguities, still
+keeping with the prefix convention: ``h'' for \HydrasLib, ``g'' for \gaia.
+
+\inputsnippets{T1Bridge/MoreNotations}
+
+\subsection{Translation functions and data refinement}
+
+The bridge between both libraries is made of two straightforward bijections.
+
+\inputsnippets{T1Bridge/h2gG2hDef}
+\inputsnippets{HydraGaia_Examples/ExamplesG2H}
+
+The following cancel lemmas will be often used in order to simplify sub-terms of the form (\texttt{g2h (h2g {\it t})}) and (\texttt{h2g (g2h {\it t})}) which appear in many proofs by rewriting.
+
+\inputsnippets{T1Bridge/h2gG2hRw, T1Bridge/g2hH2gRw}
+
+\inputsnippets{T1Bridge/h2gEqIff, T1Bridge/g2hEqIff}
+
+The function \texttt{g2h} is also used to provide a ``pretty'' printing of Cantor normal forms (through \texttt{Epsilon0.T1.pp}).
+
+\inputsnippets{T1Bridge/T1ppDef}
+
+\inputsnippets{HydraGaia_Examples/T1pp}
+
+Functions \texttt{h2g} and \texttt{g2h} allow us to define
+a notion of ``data-refinement''  for constants, functions, predicates and relations. The following definitions express that some
+constant, function, relation defined in \HydrasLib ``implements'' the same concept of \gaia.
+
+\section{Importing Material from \Hydras}
 
@@
 
-The \gaiaHydras entry point is the file
-\href{../theories/html/gaia_hydras.T1Bridge.html}%
-{\texttt{theories/gaia.T1Bridge.v}}, which requires modules both from \HydrasLib and \gaia. By default, we give priority to \gaia's notation and vocabulary.
-
-\inputsnippets{T1Bridge/Requirements}
-
-\inputsnippets{T1Bridge/hT1gT1}
-
 \subsection{Data refinement}
-
-\subsubsection{Renamings}
-
-Most used constants and functions are also unambiguously renamed.
-
-\inputsnippets{T1Bridge/MoreNotations}
 
@@
 A snippet like \texttt{(alpha < succ beta)} should be interpreted as \texttt{(T1lt alpha beta)}, where \texttt{T1lt} is defined in ~\href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{\texttt{gaia.ssete9.v}}.
-
-\subsubsection{Translation functions and data refinement}
-
-The bridge between both libraries is made of two functions.
-
-\inputsnippets{T1Bridge/h2gG2hDef}
-
-The following cancel lemmas will be often used in order to simplify sub-terms of the form (\texttt{g2h (h2g {\it t})}) and (\texttt{h2g (g2h {\it t})}) which appear in many proofs by rewriting.
-
-\inputsnippets{T1Bridge/h2gG2hRw, T1Bridge/g2hH2gRw}
-
-\inputsnippets{T1Bridge/h2gEqIff, T1Bridge/g2hEqIff}
-
-Functions \texttt{h2g} and \texttt{g2h} allow us to define
-a notion of ``data-refinement''  for constants, functions, predicates and relations. The following definitions express that some
-constant, function, relation defined in \HydrasLib ``implements'' the same concept of \gaia.
 

### 2022-02-27 doc/gaia-chapter.tex
corrections in gaia-chapter

@@
 \href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge}, which requires both Gaia's \texttt{ssete9} and a few modules of \texttt{theories/ordinals/Epsilon0}.
-
 Other modules are dedicated to the adaptation of several notions of \HydrasLib to \mathcomp's and \gaia's vocabulary and structures.
@@
 
+Please keep in mind that, in this module, we favour \gaia's notations over hydra's.
+A snippet like \texttt{(alpha < succ beta)} should be interpreted as \texttt{(T1lt alpha beta)}, where \texttt{T1lt} is defined in ~\href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{\texttt{gaia.ssete9.v}}.
+
 \subsection{Translation functions and data refinement}
@@
 
-The following cancel lemmas will be often used in order to simplify sub-terms of the form (\texttt{g2h (h2g {\it t})}) and (\texttt{h2g (g2h {\it t})}) which appear in many proofs by rewriting.
+The following cancel lemmas will be often applied in order to simplify sub-terms of the form (\texttt{g2h (h2g {\it t})}) and (\texttt{h2g (g2h {\it t})}), which appear in many proofs by rewriting.
 
@@
 
-The function \texttt{g2h} is also used to provide a ``pretty'' printing of Cantor normal forms (through \texttt{Epsilon0.T1.pp}).
+\subsubsection{Pretty printing Cantor normal forms}
+
+The function \texttt{g2h} is also used to provide a ``pretty'' printing of Cantor normal forms (please see Sect.~\vref{sect:ppT1}).
 
@@
 
+\subsubsection{Kind of refinement}
+
 Functions \texttt{h2g} and \texttt{g2h} allow us to define
@@
 constant, function, relation defined in \HydrasLib ``implements'' the same concept of \gaia.
-
-\section{Importing Material from \Hydras}
-
-In the following two examples, the bare name \texttt{T1} is given by default either  to \HydrasLib' world or \gaia's.
-
-\begin{itemize}
-\item In
-  \href{../theories/html/gaia_hydras.GaiaToHydra.html}%
-  {\texttt{gaia\_hydras.GaiaToHydra}}, we adopt \HydrasLib's point of view, and the name \texttt{T1} is by default bound to
-\HydrasLib' data structures.
-
-\inputsnippets{GaiaToHydra/T1BridgeUse,
-  GaiaToHydra/LocateT1}
-
-Section~\ref{sect:gaia2hydra} shows how lemmas on arithmetic properties of ordinals strictly less than $\epsilon_0$ are  imported from \gaia , with \HydrasLib's notations.
-
-\item The other way, the module \href{../theories/html/gaia_hydras.GCanon.html}%
-  {\texttt{gaia\_hydras.GCanon}} is an adaptation to \gaia of
-  \href{../theories/html/hydras.Epsilon0/Canon.html}%
-  {\texttt{hydras.Epsilon0/Canon}}.
-  In this module, the name \texttt{T1} is bound by default to \gaia's meaning. \HydrasLib' name \texttt{T1} is renamed into \texttt{hT1}.
-
-  The following lemmas are expressed in \gaia's vocabulary.
-  \inputsnippets{GCanon/gcanonLimitMono,
-    GCanon/gcanonLimitOf}
-
-\end{itemize}
-
-\subsection{Data refinement}
-
-Please note that in this module, we favour \gaia's notations over hydra's.
-A snippet like \texttt{(alpha < succ beta)} should be interpreted as \texttt{(T1lt alpha beta)}, where \texttt{T1lt} is defined in ~\href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{\texttt{gaia.ssete9.v}}.
 
@@
 \coq's command \texttt{Search} and notation scopes allow you to explore both libraries.
+For instance, let us look for lemmas whose conclusion is
+$\alpha \times \beta=\beta$.
 
 \inputsnippets{T1Bridge/SearchDemo}
-
-\section{Importing a theorem from \gaia}
-\label{sect:gaia2hydra}
-
-The \texttt{Gaia} library already contains many lemmas about
-ordinal arithmetic. In this section, we give two examples of
-porting such a lemma to \HydrasLib' vocabulary.
-
-\subsection{Associativity of ordinal multiplication (below \texorpdfstring{$\epsilon_0$}{epsilon\_0})}
-\gaia already contains a proof that the multiplication of ordinals less than $\epsilon_0$ is associative.
-\emph{From~\href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}}
-
-\begin{Coqsrc}
-Lemma mulA: associative T1mul.
-\end{Coqsrc}
-
-This lemma was missing from \texttt{hydra-battles}. Nevertheless, we could adapt this lemma to \texttt{hydra-battles}' ordinals, by a small sequence of rewritings.
-
-\inputsnippets{T1Bridge/multA}
-
-The module~\href{../theories/html/gaia_hydras.GaiaToHydra.html}%
-{\texttt{gaia\_hydras.GaiaToHydra}} shows a small
-example of importation of \texttt{multA} into \texttt{hydra-battles}' world.
-
-\inputsnippets{GaiaToHydra/T1BridgeUse}
-
-In the rest of this module, names like  \texttt{T1}, \texttt{omega}, etc. are  bound to \HydrasLib' meaning.
-
- \inputsnippets{GaiaToHydra/LocateT1, GaiaToHydra/T1BridgeUseb}
-
- \subsection{Right Distributivity, etc.}
- Likewise, we prove almost for free that ordinal multiplication is right distributive over addition (with \HydrasLib' definitions).
-
- \inputsnippets{GaiaToHydra/distributivity}
-
-We plan to automatize as soon as possible this kind of transfer of lemmas from \gaia to \HydrasLib.
 
@@
 
+\section{Importing a theorem from \gaia}
+\label{sect:gaia2hydra}
+
+The \texttt{Gaia} library already contains many lemmas about
+ordinal arithmetic. In this section, we give two examples of
+porting such a lemma to \HydrasLib' vocabulary.
+
+\subsection{Associativity of ordinal multiplication (below \texorpdfstring{$\epsilon_0$}{epsilon\_0})}
+\gaia already contains a proof that the multiplication of ordinals less than $\epsilon_0$ is associative.
+\emph{From~\href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}}
+
+\begin{Coqsrc}
+Lemma mulA: associative T1mul.
+\end{Coqsrc}
+
+This lemma was missing from \texttt{hydra-battles}. Nevertheless, we could adapt this lemma to \texttt{hydra-battles}' ordinals, by a small sequence of rewritings.
+
+\inputsnippets{T1Bridge/multA}
+
+The module~\href{../theories/html/gaia_hydras.GaiaToHydra.html}%
+{\texttt{gaia\_hydras.GaiaToHydra}} shows a small
+example of importation of \texttt{multA} into \texttt{hydra-battles}' world.
+
+\inputsnippets{GaiaToHydra/T1BridgeUse}
+
+In the rest of this module, names like  \texttt{T1}, \texttt{omega}, etc. are  bound to \HydrasLib' meaning.
+
+ \inputsnippets{GaiaToHydra/LocateT1, GaiaToHydra/T1BridgeUseb}
+
+ \subsection{Right Distributivity, etc.}
+ Likewise, we prove almost for free that ordinal multiplication is right distributive over addition (with \HydrasLib' definitions).
+
+ \inputsnippets{GaiaToHydra/distributivity}
+
+We plan to automatize as soon as possible this kind of transfer of lemmas from \gaia to \HydrasLib.
+
 % \section{Gaia's proof of well-foundedness}

### 2022-02-28 doc/gaia-chapter.tex
gaia-hydras first draft

@@
-\chapter{Gaia and the hydras (experimental)}
+\chapter{Gaia and the hydras (draft)}
 \index{coq}{Plug-ins!Gaia}
@@
 
+Refinement lemmas can be easily ``reversed''.
+
 \inputsnippets{T1Bridge/refines1R, T1Bridge/refines2R}
@@
 
-\subsubsection{Cantor normal form}
+\subsubsection{More rewriting lemmas}
 
-Hence, both definitions of ``being in Cantor normal form'' are
-compatible.
+\inputsnippets{T1Bridge/T1ltIff}
+\inputsnippets{T1Bridge/T1leIff}
 
-\inputsnippets{T1Bridge/nfRef}
+\inputsnippets{T1Bridge/rewritingRules}
+
+\inputsnippets{T1Bridge/limitbRef, T1Bridge/isSuccRef}
 
@@
 
+\subsubsection{Cantor normal form}
+
+Both definitions of ``being in Cantor normal form'' are
+compatible.
+
+\inputsnippets{T1Bridge/nfRef}
+
 \subsection{Looking for a lemma}
@@
 \inputsnippets{T1Bridge/SearchDemo}
-
-\section{Importing theorems from \HydrasLib}
-
-Many lemmas of the library \texttt{ordinals/Epsilon0/} are statements about inhabitants of type \texttt{hT1}. In order to get
-similar statements on \gaia's type \texttt{T1}, we prove and
-apply rewriting rules of the following forms:
-
-\vspace{4pt}
-
-\noindent\emph{From Module~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge}}
-
-\inputsnippets{T1Bridge/T1leIff}
-\inputsnippets{T1Bridge/rewritingRules}
-\inputsnippets{T1Bridge/T1ltIff}
-
-\inputsnippets{T1Bridge/limitbRef, T1Bridge/isSuccRef}
 
@@
 
-These tools allow us to import \texttt{E0.lt}'s  well-foundedness.
+These tools allow us, for instance,  to import \texttt{E0.lt}'s  well-foundedness.
 
@@
 
-\subsection{E0 as an ordinal notation}
-The module \href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge} defines an instance of class \texttt{ON E0lt compare}
+\section{E0 as an ordinal notation}
+The module \href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines an instance of class \texttt{ON E0lt compare}
 (see Chapter~\ref{chap:ON}).
@@
 \inputsnippets{T1Bridge/ExampleComp}
+
+\section{Importing Definitions and theorems from \HydrasLib}
+
+Some constructions of \HydrasLib were (to our knowledge) not implemented in \gaia yet.
+For instance, commutative (``Hessenberg'') addition, canonical sequences and Ketonen-Solovay machinery (Chapters~\ref{chap:T1}, \ref{chap:ketonen} and \ref{chap:alpha-large}).
+In order not to copy long proofs, we chose to
+\emph{derive} several constructions from \HydrasLib into \gaia's world, and rewrite \HydrasLib statements in order to make them \gaia-compatible.
+
+\subsubsection{Remark}
+\label{sect:gaia-masking}
+We try to respect the following ``protocol''.
+Let us assume we adapt some definitions and properties
+from a given module \texttt{hydras.Epsilon0.{\it XXX}}.
+\begin{enumerate}
+\item We create a module \texttt{gaia\_hydras.{\it GXXX}},
+\item When we import some constant \emph{Id} from
+ \texttt{hydras.Epsilon0.{\it XXX}}, the name 
+\emph{Id} is given to \texttt{gaia\_hydras.{\it GXXX.Id}}.
+In order to get
+\texttt{hydras.Epsilon0.{\it XXX}}'s sense, please refer it as
+  \emph{XXX.Id} or \emph{Epsilon0.XXX.Id}.
+  \item Sometimes, we rename \emph{XXX.Id} into \emph{{\bf h}Id} (for readability's sake).
+\end{enumerate}
+
+\subsection{Hessenberg sum}
+
+Natural sum  (a.k.a. Hessenberg's  sum) is defined in
+\href{../theories/html/hydras.Epsilon0.Hessenberg.html}{hydras.Epsilon0.Hessenberg} and is used for proving termination of all hydra battles (See Sect.~\ref{sect:hessenberg-def}).
+
+We define this function by delegation to \HydrasLib'.
+\vspace{4pt} 
+\noindent.
+
+From\href{../theories/html/gaia\_hydras.GHessenberg.html}{gaia\_hydras.GHessenberg}
+
+\inputsnippets{GHessenberg/oplusDef}
+
+By simple rewriting, we import the following lemmas.
+
+\inputsnippets{GHessenberg/oplusEquations}
+
+Hessenberg sum is associative, commutative and strictly monotonous (on ordinal terms in normal form).
+
+\inputsnippets{GHessenberg/oplusLemmasa}
+\inputsnippets{GHessenberg/oplusLemmasb}
+\inputsnippets{GHessenberg/oplusLemmasc}
+\inputsnippets{GHessenberg/oplusLemmasd}
+\inputsnippets{GHessenberg/oplusLemmase}
 
@@
 
-The library \href{../theories/html/gaia.GPaths.html}%
-{\texttt{gaia.GPaths}} imports definitions and lemmas from
+The library \href{../theories/html/gaia_hydras.GPaths.html}%
+{\texttt{gaia\_hydras.GPaths}} imports definitions and lemmas from
 \href{../theories/html/hydras.Epsilon0.Paths.html}%
 {\texttt{hydras.Epsilon0.Paths}}.
-
-\subsubsection{Remark}
-\label{sect:gaia-masking}
-
-First, we import some definitions. When we import some constant \emph{Id} from {\texttt{hydras.Epsilon0.Paths}} the name
-\emph{Id} is given to {\texttt{gaia.GPaths}}. In order to get
-{\texttt{hydras.Epsilon0.Paths}}'s sense, please refer it as
-\emph{Paths.Id}.
-Sometimes, we rename \emph{Paths.Id} into \emph{{\bf h}Id}.
 

### 2022-02-28 doc/gaia-chapter.tex
fix minor typos

@@
 \subsection{Canonical sequences}
-
-In~\href{../theories/html/gaia_hydras.GCanon.html}%
-{\texttt{gaia\_hydras.GCanon}}, canonical sequences are just imported from the library \linebreak
+Canonical sequences are described in Chapter~\vref{chap:ketonen}, and implemented in
 \href{../theories/html/hydras.Epsilon0.Canon.html}%
 {\texttt{hydras.Epsilon0.Canon}}.
+Module~\href{../theories/html/gaia_hydras.GCanon.html}%
+{\texttt{gaia\_hydras.GCanon}}, imports definitions and results from that library.
+
 First, we rename \HydrasLib' \texttt{canon} function into
@@
 {\texttt{hydras.Epsilon0.Paths}}.
+
+Please refer to Sect.~\ref{sect:pathes-intro} for a descrition of that library.
 

### 2022-03-02 doc/part-hydras.tex
Small changes in gaia-chapter; add an ON instance for Mathcomp finite ordinals
corrections in gaia-chapter
gaia-hydras first draft
Remove ocons (#124)
* minor script simplifications

* minor script simplifications

* make T1's constructors gaia-compatible

@@
 
+The introduction of Jos\'e Grimm's report~\cite{grimm:hal-00911710} contains a nice presentation of the main properties of ordinals.
+
 \section{Ordinal numbers in Coq}
@@
 level 3/.style={sibling distance=17mm}]
-\node  {ocons}
-  child {  node {ocons}
-            child { node {ocons} child {node {zero}} child {node{0}} child{node{zero}}}
+\node  {cons}
+  child {  node {cons}
+            child { node {cons} child {node {zero}} child {node{0}} child{node{zero}}}
          child {node {0}}
@@
     child {node {0}}
-   child {node {ocons} 
- child { node {ocons} child {node {zero}} child {node{2}} child{node{zero}}}
+   child {node {cons} 
+ child { node {cons} child {node {zero}} child {node{2}} child{node{zero}}}
   child {node {4}}
-         child {node {ocons} child {node {zero}} child {node{1}} child{node{zero}}}};
+         child {node {cons} child {node {zero}} child {node{1}} child{node{zero}}}};
 
@@
 \paragraph{Remark}
-For simplicity's sake, we chose to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the construction (\texttt{ocons $\alpha$ $n$ $\beta$}) is intended to represent the
+For simplicity's sake, we chose to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the construction (\texttt{cons $\alpha$ $n$ $\beta$}) is intended to represent the
 ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
@@
 Unfortunately, it won't work. Let us consider some well-formed term
- $\alpha=\texttt{ocons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are accessible
+ $\alpha=\texttt{cons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are accessible
  through \texttt{LT}. For proving the accessibility of $\alpha$, we have to consider
@@
 for instance \(\delta\) may be of the form
-\texttt{ocons $\beta$ $p'$  $\gamma'$},
+\texttt{cons $\beta$ $p'$  $\gamma'$},
 where    \(p' < n\).
@@
 strongly accessible too (\emph{i.e.} any well formed
-term (\texttt{ocons $\alpha$ $n$ $\beta$})  is accessible).
+term (\texttt{cons $\alpha$ $n$ $\beta$})  is accessible).
 
@@
 
-Let \texttt{n:nat} and \texttt{beta:T1} such that (\texttt{ocons alpha n beta}) is in normal form. 
+Let \texttt{n:nat} and \texttt{beta:T1} such that (\texttt{cons alpha n beta}) is in normal form. 
 We prove first that \texttt{beta} is accessible,  which allows us to prove by well-founded induction on \texttt{beta}, 
-and natural induction on \texttt{n}, that (\texttt{ocons alpha n beta}) is accessible.
+and natural induction on \texttt{n}, that (\texttt{cons alpha n beta}) is accessible.
 The proof, quite long, can be consulted in \href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}.
@@
     \label{hydra-variant}
+    \label{sect:hessenberg-def}
 
@@
     %            compare a1 a2 = Gt ->
-    %            P b1 (ocons a2 n2 b2) o ->
-    %            P (ocons a1 n1 b1) (ocons a2 n2 b2)
-    %              (ocons a1 n1 o)) ->
+    %            P b1 (cons a2 n2 b2) o ->
+    %            P (cons a1 n1 b1) (cons a2 n2 b2)
+    %              (cons a1 n1 o)) ->
     %         (forall a1 n1 b1 a2 n2 b2 o,
     %            compare a1 a2 = Lt ->
-    %            P (ocons a1 n1 b1) b2 o ->
-    %            P (ocons a1 n1 b1) (ocons a2 n2 b2) 
-    %            (ocons a2 n2 o)) ->
+    %            P (cons a1 n1 b1) b2 o ->
+    %            P (cons a1 n1 b1) (cons a2 n2 b2) 
+    %            (cons a2 n2 o)) ->
     %         (forall a1 n1 b1 a2 n2 b2 o,
@@
     %            P b1 b2 o ->
-    %           P (ocons a1 n1 b1) (ocons a2 n2 b2)
-    %             (ocons a1 (S (n1 + n2)%nat) o)) ->
+    %           P (cons a1 n1 b1) (cons a2 n2 b2)
+    %             (cons a1 (S (n1 + n2)%nat) o)) ->
     %          forall a b, P a b (oplus a b).
@@
 
+% ---------------------------------------------------
+\input{gaia-chapter}
+
 %----------------------------------------------------------------------

### 2022-03-02 doc/part-hydras.tex
minor corrections

@@
 
-\paragraph{Remark}
-The name \texttt{T1} we gave to this data-type  is proper to this development and refers
-to a hierarchy of ordinal notations.
-For instance, in Library \href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2},  the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{schutte}).
-\noindent
-
-\input{movies/snippets/T2/T2Def}
-
 \subsubsection{Example}
@@
 
+\paragraph{Remark}
+The name \texttt{T1} we gave to this data-type  is proper to this development and refers
+to a hierarchy of ordinal notations.
+For instance, in Library \href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2},  the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{schutte}).
+\noindent
+
+\input{movies/snippets/T2/T2Def}
+
+\subsection{About Gaia}
+Chapter~\vref{gaia-chapter} describes the present state of a
+project of making compatible the libraries \HydrasLib and \gaia. At present, both libraries contain a version of the inductive type \texttt{T1}, with the same base name for the constructors.
+
+\inputsnippets{nfwfgaia/T1def}
+
+Thus, many examples of this and following chapters can be run
+in \gaia's context. We will signal any difference (notation, name) which may appear.
+ A closer integration (same data type and functions) is still in project.
+
 \subsection{Abbreviations}
@@
 \input{movies/snippets/T1/finiteOrds}
+
+\begin{gaiacompat}
+
+  In \texttt{gaia.ssete9}, the $n$-th finite ordinal is written
+  \texttt{$\backslash${F}\,$n$}. There is no coercion in \gaia from \texttt{nat} to finite ordinals.
+
+  \vspace{4pt}
+
+  From~\href{../theories/html/gaia_hydras.HydraGaia_Examples.html}{gaia\_hydras.HydraGaia\_Examples}.
+
+  \inputsnippets{HydraGaia_Examples/Demoa}
+
+\end{gaiacompat}
 
@@
 
-\input{movies/snippets/T2/T2Def}
+\inputsnippets{T2/T2Def, T2/psiDef}
 

### 2022-03-03 doc/gaia-chapter.tex
add references to gaia in chapter 4

@@
 \subsubsection{Pretty printing Cantor normal forms}
+\label{sect:gaia-ppT1}
 
@@
 \subsubsection{Order and comparison}
+\label{sect:lt-compat-gaia}
 Leibniz equality and strict order on both libraries are
@@
 \subsubsection{Cantor normal form}
-
+\label{nf-gaia-compat}
 Both definitions of ``being in Cantor normal form'' are

### 2022-03-03 doc/part-hydras.tex
minor corrections
add references to gaia in chapter 4

@@
 
-\begin{gaiacompat}
+\paragraph{\gaiacompat}
 
@@
   \inputsnippets{HydraGaia_Examples/Demoa}
-
-\end{gaiacompat}
 
@@
 
+\paragraph{\gaiacompat}
+In \texttt{gaia.ssete9}, the ordinal $\omega$ is bound to
+a \emph{constant} called \texttt{T1omega}.
+
 \subsubsection{The ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
@@
 
+\paragraph{\gaiacompat}
+In \texttt{gaia.ssete9} the identifier\texttt{phi0} is bound to a constant (not a notation).
+
 \subsubsection{The hierarchy of \(\omega\)-towers:}
@@
 
+\paragraph{\gaiacompat}
+In ~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge},
+we define a variant of \texttt{pp} for pretty-printing terms of type \texttt{ssete9.CantorOrdinal.T1} (see Sect.~\vref{sect:gaia-ppT1}).
+
 \index{hydras}{Projects}
@@
 
+\paragraph{\gaiacompat}
+In \gaia, the strict order \texttt{T1lt} on ordinal terms is directly defined as a boolean function (in \ssreflect's style).
+
+\inputsnippets{nfwfgaia/T1ltDef}
+
+In section~\vref{sect:lt-compat-gaia}, we show that both definitions are mutually equivalent.
+
 \subsubsection{A Predicate for Characterizing Normal Forms}
@@
 \input{movies/snippets/T1/nfBadTerm}
+
+\paragraph{\gaiacompat}
+In \gaia, the boolean function which characterizes ordinal terms in normal form is defined as follows:
+
+\inputsnippets{nfwfgaia/T1nfDef}
+
+In Sect.~\vref{nf-gaia-compat}, we show that \gaia's \texttt{T1nf} is extensionnaly equivalent with \HydrasLib'  \texttt{nf}.
 

### 2022-03-03 doc/part-hydras.tex
add compatibility notes (Gaia) (up to page 85)
add compatibility notes (Gaia)

@@
 
+\paragraph{\gaiacompat}
+In \gaia, the boolean functions associated with limit and successor ordinals are called \texttt{T1limit} and \texttt{T1is\_succ}.
+
+ From~\href{../theories/html/gaia_hydras.HydraGaia_Examples.html}{gaia\_hydras.HydraGaia\_Examples}.
+
+\inputsnippets{HydraGaia_Examples/Demob}
+
 \subsection{Arithmetic on \texorpdfstring{$\epsilon_0$}{epsilon0}}
@@
 \input{movies/snippets/T1/succbIff}
+
+\begin{exercise}[\gaia]
+  Look for the \gaia-theorem which corresponds to \texttt{succb\_iff}.
+\end{exercise}
+
+\subsubsection{Successor function on \texttt{E0}}
 
@@
 \input{movies/snippets/T1/multDef}
+
+\paragraph{\gaiacompat}
+In \gaia, addition and multiplication of ordinal terms below $\epsilon_0$ are respectively called \texttt{T1add} and \texttt{T1mul}.
+Please refer to Sect.~\ref{sect:plus-mult-gaia-hydras} about compatibility of both arithmetics.
 

### 2022-03-04 doc/part-hydras.tex
same name for T1add

--- 2022-03-02 doc/part-hydras.tex
@@
 
+\begin{remark}
+  \label{remark:hydra-gaia-renaming}
+  \textcolor{lookcolor}{In future releases, we plan to make some
+    \HydrasLib identifiers progressively  deprecated, in favour of \gaia's names.}
+\end{remark}
+
 \subsection{Abbreviations}
@@
 
-\begin{gaiacompat}
+\paragraph{\gaiacompat (cf. Remark ~\ref{remark:hydra-gaia-renaming})}
 
@@
 
+\paragraph{\gaiacompat (cf. Remark ~\vref{remark:hydra-gaia-renaming})}
+In \texttt{gaia.ssete9}, the ordinal $\omega$ is bound to
+a \emph{constant} called \texttt{T1omega}.
+
 \subsubsection{The ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
@@
 
+\paragraph{\gaiacompat (cf. Remark ~\vref{remark:hydra-gaia-renaming})}
+In \texttt{gaia.ssete9} the identifier\texttt{phi0} is bound to a constant (not a notation).
+
 \subsubsection{The hierarchy of \(\omega\)-towers:}
@@
 
+\paragraph{\gaiacompat (cf. Remark ~\ref{remark:hydra-gaia-renaming})}
+In ~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge},
+we define a variant of \texttt{pp} for pretty-printing terms of type \texttt{ssete9.CantorOrdinal.T1} (see Sect.~\vref{sect:gaia-ppT1}).
+
 \index{hydras}{Projects}
@@
 
+\paragraph{\gaiacompat (cf. Remark ~\ref{remark:hydra-gaia-renaming})}
+In \gaia, the strict order \texttt{T1lt} on ordinal terms is directly defined as a boolean function (in \ssreflect's style).
+
+\inputsnippets{nfwfgaia/T1ltDef}
+
+In section~\vref{sect:lt-compat-gaia}, we show that both definitions are mutually equivalent.
+
 \subsubsection{A Predicate for Characterizing Normal Forms}
@@
 
+\paragraph{\gaiacompat (cf. Remark ~\ref{remark:hydra-gaia-renaming})}
+In \gaia, the boolean function which characterizes ordinal terms in normal form is defined as follows:
+
+\inputsnippets{nfwfgaia/T1nfDef}
+
+In Sect.~\vref{nf-gaia-compat}, we show that \gaia's \texttt{T1nf} is extensionnaly equivalent with \HydrasLib'  \texttt{nf}.
+
 \subsection{Making normality implicit}
@@
 
+\paragraph{\gaiacompat (cf. Remark ~\ref{remark:hydra-gaia-renaming})}
+In \gaia, the boolean functions associated with limit and successor ordinals are called \texttt{T1limit} and \texttt{T1is\_succ}.
+
+ From~\href{../theories/html/gaia_hydras.HydraGaia_Examples.html}{gaia\_hydras.HydraGaia\_Examples}.
+
+\inputsnippets{HydraGaia_Examples/Demob}
+
 \subsection{Arithmetic on \texorpdfstring{$\epsilon_0$}{epsilon0}}
@@
 \input{movies/snippets/T1/multDef}
+
+\paragraph{\gaiacompat (cf. Remark ~\ref{remark:hydra-gaia-renaming})}
+In \gaia, addition and multiplication of ordinal terms below $\epsilon_0$ are respectively called \texttt{T1add} and \texttt{T1mul}.
+Please refer to Sect.~\ref{sect:plus-mult-gaia-hydras} about compatibility of both arithmetics.
 

### 2022-03-04 doc/part-hydras.tex
unification of T1add and T1mul names

@@
 \paragraph{\gaiacompat (cf. Remark ~\ref{remark:hydra-gaia-renaming})}
-In \gaia, addition and multiplication of ordinal terms below $\epsilon_0$ are respectively called \texttt{T1add} and \texttt{T1mul}.
+We keep \gaia's base names for
+addition and multiplication of ordinal terms below $\epsilon_0$.
 Please refer to Sect.~\ref{sect:plus-mult-gaia-hydras} about compatibility of both arithmetics.

### 2022-03-09 doc/gaia-chapter.tex
add compatibility notes (Gaia)
Minor corrections; add a bibitem (Talia Ringer et al.)

@@
 For this purpose, we write modules dedicated to the importation of definitions and lemmas from each of both libraries into the other one.
-Please note that by ``import'' we do not mean ``duplicate the proofs''.  Whenever possible, we proceed by rewriting steps over the statements and not the proofs.
+  Please note that by ``import'' we do not mean ``duplicate''
+  nor ``repairing'' \HydrasLib' proofs.
+Whenever possible, we proceed by rewriting steps over the statements and not the proofs.
+In a future version, it would be nice to have our proofs in the same context, after porting \HydrasLib' proofs under \gaia's
+  definition, following  proof repair  techniques~\cite{ringer2021}.
 
@@
 Hydra-battles code using SSReflect and MathComp than to reason about
-Gaia without SSReflect \footnote{The author of these proof scripts is still a beginner in Ssreflect. Please forgive the clumsiness of the current proof scripts.}.
+Gaia without SSReflect\footnote{The author of these proof scripts is still a beginner in Ssreflect. Please forgive the clumsiness of the current proof scripts.}.
 
@@
 
-The following cancel lemmas will be often applied in order to simplify sub-terms of the form (\texttt{g2h (h2g {\it t})}) and (\texttt{h2g (g2h {\it t})}), which appear in many proofs by rewriting.
-
-\inputsnippets{T1Bridge/h2gG2hRw, T1Bridge/g2hH2gRw}
-
-\inputsnippets{T1Bridge/h2gEqIff, T1Bridge/g2hEqIff}
-
 \subsubsection{Pretty printing Cantor normal forms}
@@
 
-\subsubsection{Kind of refinement}
+\subsubsection{Cancel lemmas}
+
+The following cancel lemmas will be often applied in order to simplify sub-terms of the form (\texttt{g2h (h2g {\it t})}) and (\texttt{h2g (g2h {\it t})}), which appear in many proofs by rewriting.
+
+\inputsnippets{T1Bridge/h2gG2hRw, T1Bridge/g2hH2gRw}
+
+\inputsnippets{T1Bridge/h2gEqIff, T1Bridge/g2hEqIff}
+
+\subsubsection{Refinements: Definitions}
 
@@
 \subsubsection{Addition and multiplication}
+\label{sect:plus-mult-gaia-hydras}
 
@@
 Let us assume we adapt some definitions and properties
-from a given module \texttt{hydras.Epsilon0.{\it XXX}}.
+from a given module \texttt{hydras.Epsilon0.Foo}, and a constant \texttt{bar} of this module, whose absolute name is
+thus \texttt{hydras.Epsilon0.Foo.bar}.
 \begin{enumerate}
-\item We create a module \texttt{gaia\_hydras.{\it GXXX}},
-\item When we import some constant \emph{Id} from
- \texttt{hydras.Epsilon0.{\it XXX}}, the name 
-\emph{Id} is given to \texttt{gaia\_hydras.{\it GXXX.Id}}.
-In order to get
-\texttt{hydras.Epsilon0.{\it XXX}}'s sense, please refer it as
-  \emph{XXX.Id} or \emph{Epsilon0.XXX.Id}.
-  \item Sometimes, we rename \emph{XXX.Id} into \emph{{\bf h}Id} (for readability's sake).
+\item We create a module \texttt{gaia\_hydras.GFoo}
+  (``G'' for \gaia).
+\item Inside \texttt{GFoo}, the identifier \texttt{bar}
+  is reserved in priority to \gaia's version, and the notation
+  \texttt{hbar} is used for \texttt{hydras.Epsilon0.Foo.bar}.
 \end{enumerate}

### 2022-03-09 doc/gaia-chapter.tex
Minor corrections

@@
 \href{../theories/html/hydras.Epsilon0.Paths.html}%
-{\texttt{hydras.Epsilon0.Paths}}.
-
-Please refer to Sect.~\ref{sect:pathes-intro} for a descrition of that library.
+{\texttt{hydras.Epsilon0.Paths}} (described in Sect.~\ref{sect:pathes-intro}).
 
@@
 
-\subsection{A family of rapidly growing functions}
+\subsection{A family of rapidly growing functions
+(Library \texttt{GF\_alpha})}
 
@@
 We are now able to import \HydrasLib' $F_\alpha$ functions.
+
+From~\href{../theories/html/gaia_hydras.G~F_alpha.html}%
+{\texttt{gaia\_hydras.GF\_alpha}}.
 

### 2022-03-10 doc/chapter-powers.tex
small changes
new title page
Fix a few (too long to generate) snippets

@@
 much as possible within the more generic framework of \texttt{EMonoid}s.
-
-\index{coq}{Coercions}
 
@@
 
-\inputsnippets{Addition_Chains/slowC87Correct}
+\begin{Coqbad}
+Example C87_ok_slow : chain_correct 87 C87.
+Proof.
+Time slow_chain_correct_tac.
+\end{Coqbad}
+
+\begin{Coqanswer}
+Finished transaction in 49.927 secs (49.445u,0.079s) (successful)
+\end{Coqanswer}
+
+\begin{Coqbad}
+Qed.
+\end{Coqbad}
+
+%\inputsnippets{Addition_Chains/slowC87Correct}
 
@@
 
-\index{coq}{Plug-ins!paramcoq}
+\index{coq}{Plug-ins!Paramcoq}
 
@@
 
-\begin{Coqbad}
-Module Bad.
-
-Definition Fchain_correct  (n:nat) (f : Fchain) :=
-  forall A `(M : @EMonoid A op E_one E_equiv) k (a:A),
-    computation_execute op (f A k  a)==
-    computation_execute op (k  (a ^ n)).
-\end{Coqbad}
+\inputsnippets{Euclidean_Chains/BadDefa}
 
@@
 
-\inputsnippets{Euclidean_Chains/BadDefa}
 \inputsnippets{Euclidean_Chains/BadDefb}
@@
 
-\begin{Coqsrc}
-Definition fib_eucl gamma `{Hgamma: Strategy gamma} n :=
-  let c := make_chain gamma  n
-  in let r := chain_apply c (M:=Mul2) (1,0) in
-       fst r + snd r.
-
-Compute fib_eucl dicho 153.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-    = 68330027629092351019822533679447
-     : N
-Finished transaction in 0.002 secs (0.002u,0.s) (successful)
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Compute fib_eucl two 153.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-    = 68330027629092351019822533679447
-     : N
-Finished transaction in 0.003 secs (0.003u,0.s) (successful)
-\end{Coqanswer}
-
-\begin{Coqsrc}
-Compute fib_eucl half 153.
-\end{Coqsrc}
-
-\begin{Coqanswer}
-    = 68330027629092351019822533679447
-     : N
-Finished transaction in 0.003 secs (0.003u,0.s) (successful)
-\end{Coqanswer}
+\inputsnippets{Fib2/fibEuclDemo}
 

### 2022-03-10 doc/hydras.tex
first try
failed attempt to generate snippets with utf8 characters
utf8 omega accepted
try to use DejaVu Sans Mono
add a few lemmas
Bridge dev (#112)
* more lemmas (draft)

* add a type E0 (in Bridge)

* experiments in bridge building

* experiments in bridge building

* experiments in bridge building

* fix a typo in gaia-chapter.tex

* fix a typo in gaia-chapter.tex

* first try

* failed attempt to generate snippets with utf8 characters

* additions to GCanon.V

* utf8 omega accepted

* Canonical sequences in Bridge (initial)

* try to use DejaVu Sans Mono

* add a few lemmas

* fix a typo

@@
 \documentclass[twoside,a4paper]{book}
+
+%\usepackage{fontenc}
+\usepackage[utf8x]{inputenc}
+
+\usepackage{fontspec}
+%\setmainfont{DejaVu Serif}
+
+\setmonofont{DejaVu Sans Mono}[Scale=MatchLowercase]
 \usepackage{alltt}
@@
 \newpage
-% \mbox{~}
+\thispagestyle{empty}
 \begin{quote}
@@
 \end{quote}
-\vspace{2cm}
+\vspace{17mm}
 \begin{quote}
@@
 
-\vspace{2cm}
+\vspace{17mm}
 \begin{quote} 
@@
 \end{quote}
+\mbox{}
+\newpage
+\newpage
+% \mbox{~}
+\thispagestyle{empty}
+\label{watercolor}
+ {\centering
+      \includegraphics[width=13cm]{blue-hydra.jpg}
+      \centerline{\emph{``The blue hydra'': Watercolor  by Pierrette Cassou-Noguès}}
+}
 
@@
 \subsubsection{Using \alectr}
+Whenever possible, we use \alectr to display \coq code (definition, proof scripts) and answers.
 
-Here is an example from Chapter~\ref{chapter:primrec}.
+Here are two  examples from Chapters~\ref{chapter:primrec}
+and~\ref{chapter-powers}.
 
@@
 \input{movies/snippets/Ack/AckFixpointIterate.tex}
+
+\inputsnippets{Fib2/fibEuclDemo}
 
@@
 
-\begin{Coqsrc}
-Definition fib_eucl gamma `{Hgamma: Strategy gamma} n :=
-  let c := make_chain gamma  n
-  in let r := chain_apply c (M:=Mul2) (1,0) in
-       fst r + snd r.
-
-Compute fib_eucl dicho 153.
-\end{Coqsrc}
+\begin{Coqbad}
+Example C87_ok_slow : chain_correct 87 C87.
+Proof.
+Time slow_chain_correct_tac.
+\end{Coqbad}
 
 \begin{Coqanswer}
-    = 68330027629092351019822533679447
-     : N
+Finished transaction in 49.927 secs (49.445u,0.079s) (successful)
 \end{Coqanswer}
+
+\begin{Coqbad}
+Qed.
+\end{Coqbad}
 
@@
 
+Last but not least, many thanks to Pierrette Cassou-Noguès for the watercolor on page~\pageref{watercolor}.
+
 \subsection{Contributions}
@@
 \include{chapter-primrec}
-\include{gaia-chapter}
 \part{A few  certified algorithms}

### 2022-03-10 doc/part-hydras.tex
more gaia-compatible names
small changes (book)

@@
 In Chapter~\ref{gaia-chapter}, we present some modules which allow \HydrasLib' users to apply lemmas proven in \gaia, and vice versa.
+Remarks about compatibility with \gaia signalled with the picture \gaiasign.
 
@@
 
-\paragraph{\gaiacompat (cf. Remark ~\ref{remark:hydra-gaia-renaming})}
-
+\paragraph*{\gaiasign}
   In \texttt{gaia.ssete9}, the $n$-th finite ordinal is written
@@
 
-\paragraph{\gaiacompat (cf. Remark ~\vref{remark:hydra-gaia-renaming})}
-In \texttt{gaia.ssete9}, the ordinal $\omega$ is bound to
-a \emph{constant} called \texttt{T1omega}.
+\paragraph*{\gaiasign}
+In \texttt{gaia.ssete9}, the ordinal $\omega$ is also named as \emph{constant}  \texttt{T1omega}.
 
@@
 
-\paragraph{\gaiacompat (cf. Remark ~\vref{remark:hydra-gaia-renaming})}
+\paragraph*{\gaiasign}
 In \texttt{gaia.ssete9} the identifier\texttt{phi0} is bound to a constant (not a notation).
@@
 
-\paragraph{\gaiacompat (cf. Remark ~\ref{remark:hydra-gaia-renaming})}
+\paragraph*{\gaiasign}
 In ~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge},
@@
 
-\paragraph{\gaiacompat (cf. Remark ~\ref{remark:hydra-gaia-renaming})}
+\paragraph*{\gaiasign}
 In \gaia, the strict order \texttt{T1lt} on ordinal terms is directly defined as a boolean function (in \ssreflect's style).
@@
 
-\paragraph{\gaiacompat (cf. Remark ~\ref{remark:hydra-gaia-renaming})}
+\paragraph*{\gaiasign}
 In \gaia, the boolean function which characterizes ordinal terms in normal form is defined as follows:
@@
 
-\paragraph{\gaiacompat (cf. Remark ~\ref{remark:hydra-gaia-renaming})}
+\paragraph*{\gaiasign}
 In \gaia, the boolean functions associated with limit and successor ordinals are called \texttt{T1limit} and \texttt{T1is\_succ}.
@@
 
-\paragraph{\gaiacompat (cf. Remark ~\ref{remark:hydra-gaia-renaming})}
+\paragraph*{\gaiasign}
 We keep \gaia's base names for

### 2022-03-10 doc/chapter-powers.tex
fix a bug (snippet)

@@
 In order to understand this poor performance, let us consider an intermediate
-subgoal of the previous proof generated after a sequence of unfoldings and simplifications. This goal is presented below.
+subgoal of the previous proof generated after a sequence of unfoldings and simplifications. This goal is presented in Figure~\vref{fig:big-goal}.
+
 \begin{figure}[h]
   \centering
-\inputsnippets{Addition_Chains/WhySoSlow}  
+  \begin{Coqanswer}
+    1 goal (ID 506)
+
+  A : Type
+  E_op : Mult_op A
+  E_one : A
+  E_eq : Equiv A
+  M : EMonoid E_op E_one E_eq
+  x : A
+  ============================
+  E_eq
+    (x * x * x * (x * x * x) * x * (x * x * x) *
+     (x * x * x * (x * x * x) * x * (x * x * x)) *
+     (x * x * x * (x * x * x) * x * (x * x * x) *
+      (x * x * x * (x * x * x) * x * (x * x * x))) *
+     (x * x * x * (x * x * x) * x * (x * x * x) *
+      (x * x * x * (x * x * x) * x * (x * x * x)) *
+      (x * x * x * (x * x * x) * x * (x * x * x) *
+       (x * x * x * (x * x * x) * x * (x * x * x)))) *
+     (x * x * x * (x * x * x) * x))
+    (x *
+     (x *
+      (x *
+       (x *
+        (x *
+         (x *
+          (x *
+           (x *
+            (x *
+             (x *
+              (x *
+               (x *
+                (x * (x * (x *  ...)))))))))))))))
+
+  \end{Coqanswer}
+%\inputsnippets{Addition_Chains/WhySoSlow}  
   \caption{A big goal}

### 2022-03-11 doc/part-hydras.tex
minor corrections

@@
 \section{Ordinal Notations and \gaia's ordinals}
-Module~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge} contains an instance of class \texttt{ON E0lt compare}, where \texttt{E0lt} is the order on the well-formed ordinal terms below $\epsilon_0$ defined in \gaia library (please see Chapter~\ref{gaia-chapter}).
+\gaiasign Module~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge} contains an instance of class \texttt{ON E0lt compare}, where \texttt{E0lt} is the order on the well-formed ordinal terms below $\epsilon_0$ defined in \gaiaHydras library (please see Chapter~\ref{gaia-chapter}).
 
@@
 \subsection{About Gaia}
-Chapter~\vref{gaia-chapter} describes the present state of a
+\gaiasign Chapter~\vref{gaia-chapter} describes the present state of a
 project of making compatible the libraries \HydrasLib and \gaia. At present, both libraries contain a version of the inductive type \texttt{T1}, with the same base name for the constructors.
@@
 
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
+\emph{From Module~w}
 
@@
 %\index{Constants!zero:T1}
+
+\paragraph*{\gaiasign}
+ The library~\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines a \gaia-compatible type \texttt{E0}.
 
@@
     \input{movies/snippets/Hydra_Termination/FinalThm}
+
+    \paragraph*{\gaiasign}
+    The module~\href{../theories/html/gaia_hydras.GHydra.html}{gaia\_hydras.GHydra} contains a proof of \gaiaHydras version of termination of all battles.
+
+    \inputsnippets{GHydra/mDef, GHydra/mVariant, GHydra/Termination}
 

### 2022-03-11 doc/gaia-chapter.tex
minor corrections

@@
-\chapter{Gaia and the hydras (draft)}
+\chapter{Gaia and the hydra (draft)}
 \index{coq}{Plug-ins!Gaia}
@@
 
-The modules are located in {\texttt{theories/gaia/*.v}}.
+The modules are named {\texttt{theories/gaia/*.v}} and coqdocumented in {\texttt{theories/html/gaia\_hydras.*.html}}
 The main entry point is
@@
 
-The function \texttt{g2h} is also used to provide a ``pretty'' printing of Cantor normal forms (please see Sect.~\vref{sect:ppT1}).
+The following function provides us with a more readable  printing of Cantor normal forms (please see Sect.~\vref{sect:ppT1}).
 
@@
 
-For sake of compatibility, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}.
+For compatibility's sake, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}.
 
@@
 \inputsnippets{GCanon/gcanonLt}
-\inputsnippets{GCanon/glimitCanonSNotZero}
 \inputsnippets{GCanon/gcanonLimitMono}
@@
   Give equations for \texttt{F\_} (corresponding to the equation
-  in \HydrasLib.
+  in \HydrasLib).
 \end{todo}

### 2022-03-12 doc/hydras.tex
small changes (book)
minor corrections
book introduction: minor corrections

@@
 \usepackage{texments}
+
 %%% for movies by alectryon
@@
 \input{coqmacros}
+
+\newcommand{\gaiacompat}{\textcolor{blue}{Gaia}}
+\newcommand{\gaiasign}{\texorpdfstring{\mycircled[cyan]{{\mbox{\color{blue}{\scriptsize{G}}}}}}{Gaia-Hydra}}
+\newtheorem{gaiahydra}{\gaiasign}
 
@@
 
+Let us underline the analogy between hydra battles and interactive theorem proving.
+Hercules is the user (you!), and hydra's heads are the subgoals: you may think that applying a tactic would solve a subgoal, but it results often in the multiplication of such tasks.
+
 Hydra battles, as well as Goodstein sequences~\cite{goodstein_1944, KP82}
@@
 
-Let us underline the analogy between hydra battles and interactive theorem proving.
-Hercules is the user (you!), and hydra's heads are the subgoals: you may think that applying a tactic would solve a subgoal, but it results often in the multiplication of such tasks.
+We present a (still partial, under continuous development) implementation in \coq of the various techniques shown in
+Kirby \& Paris' and Ketonen \& Solovay's~\cite{KS81} article.
+
+Our library \gaiaHydras is dedicated to make compatible our lemmas with José Grimm's \gaia project (designed for \ssreflect/\mathcomp).
 
@@
 
- \subsection{Developing and maintaining the \coq code and documentation}
+ \subsection{Structure of \Hydras}
 
- \textbf{TODO} introductory text and link to \cite{jfla2022}
+ \Hydras is made of three main packages: \HydrasLib, \gaiaHydras, and \additions. Figure~\ref{fig:genealogy}  illustrates
+ the complex relationships: inheritance from historical contributions to \coq, and dependency with other \coq packages.
+ Many thanks to  Karl Palmskog and Théo Zimmermann for the CI/CD design of \Hydras and the automation of documentation maintenance.
+ Please look for a more detailed description in~\cite{jfla2022}.
 
- \subsubsection{Continuous Integration and Continuous Delivery}
+ \begin{figure}[ht]
+\centering
+{\footnotesize
+\begin{tikzpicture}[thick]
 
- \textbf{TODO}
+\begin{scope}[xshift=-6cm]
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (additions) { Additions };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (cantor) { Cantor };
 
-\subsubsection{Documenting theories with \alectr}
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3.2cm] (cats) { Categories in ZFC };
+\end{scope}
+
+\begin{scope}[yshift=-1cm,xshift=0.25cm]
+  \node[rectangle, draw=black,fill=blue!20,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=1.8cm] (mathcomp) { MathComp };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=1.5cm,left of=paramcoq, node distance=4.5cm] (equations) { Equations };
+\end{scope}
+
+\begin{scope}[yshift=-2cm,xshift=-4cm]
+\node[rectangle, draw=black,fill=orange!30,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
+\node[rectangle, draw=black,fill=blue!20,minimum height=0.5cm,minimum width=1.5cm,left of=hydras, node distance=3.5cm] (pockcc) { Pocklington };
+\node[rectangle, draw=black,fill=orange!30,minimum height=0.5cm,minimum width=1.5cm,right of=hydras, node distance=2.5cm] (chains) { Addition-chains };
+\node[rectangle, draw=black,fill=blue!20,minimum height=0.5cm,minimum width=1.5cm,right of=chains, node distance=2.2cm] (gaia) { Gaia };
+\end{scope}
+
+\begin{scope}[yshift=-3cm,xshift=-6cm]
+ \node[rectangle, draw=black,fill=blue!20,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
+ \node[rectangle, draw=black,fill=orange!30,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=3cm] (gaiahydras) { Gaia-hydras };
+\end{scope}
+
+\draw[->,dotted] (additions.south) -- (chains.north) ;
+\draw[->,dotted] (goedel) -- (goedelcc) ;
+\draw[->,dotted] (goedel) -- (hydras.north west) ;
+\draw[->,dotted] (cantor) -- (hydras.north east) ;
+\draw[->,dotted] (cantor.south) -- (gaia.north west) ;
+\draw[->,dotted] (pock) -- (pockcc) ;
+\draw[->,dotted] ([xshift=2.1cm]cats) -- (gaia) ;
+
+\draw[->] (hydras) -- (gaiahydras) ;
+\draw[->] (hydras.south west) -- (goedelcc) ;
+\draw[->] (gaia.south west) -- (gaiahydras.east) ;
+
+\draw[->] (equations) -- (hydras) ;
+\draw[->] (mathcomp) -- (gaia) ;
+\draw[->] (mathcomp) -- (chains) ;
+\draw[->] (paramcoq.south) -- (chains) ;
+\draw[->] (pockcc) -- (goedelcc) ;
+\end{tikzpicture}
+}
+\caption{Genealogy and dependencies for \Hydras packages.
+Dotted boxes represent historical Coq contributions, while regular boxes represent maintained \coq packages.
+Orange packages are maintained in the \Hydras GitHub repository, while light blue packages are maintained in other \community repositories.
+Dotted lines represent \coq code ancestry, while regular lines represent direct code dependencies.}
+  \label{fig:genealogy}
+\end{figure}
+
+\subsection{Documenting theories with \alectr}
 
@@
 \item There is no measure into $\mathbb{N}$ for proving the termination of all hydra battles (Sect~\vref{omega-case}).
-\item There is no measure into the interval\footnote{We use the notation $[a,b)$ for denoting the set of ordinals greater or equal than $a$ and strictly less than $b$.}  $[0,\omega^2)$ for proving the termination of all hydra battles (Sect.~\vref{omega2-case}).
+\item There is no measure into the interval\,\footnote{We use the notation $[a,b)$ for denoting the set of ordinals greater or equal than $a$ and strictly less than $b$.}  $[0,\omega^2)$ for proving the termination of all hydra battles (Sect.~\vref{omega2-case}).
 \item There is no measure into $[0,\mu)$ for proving the termination of all hydra battles, for any $\mu<\epsilon_0$ (Sect.\vref{sec:free-battles-case}).

### 2022-03-12 doc/hydras.tex
re-reading pages 9-21

@@
 \newcommand{\gaiacompat}{\textcolor{blue}{Gaia}}
-\newcommand{\gaiasign}{\texorpdfstring{\mycircled[cyan]{{\mbox{\color{blue}{\scriptsize{G}}}}}}{Gaia-Hydra}}
+\newcommand{\gaiasign}{\texorpdfstring{\mycircled[cyan!40]{{\mbox{\color{blue}{\scriptsize{G}}}}}}{Gaia-Hydra}}
 \newtheorem{gaiahydra}{\gaiasign}
@@
 Although the mathematical contents of this 
-paper are quite advanced, the rules of hydra battles are very easy to understand.
+paper are quite advanced, the rules of hydra battles are very easy to understand\footnote{Let us underline the analogy between hydra battles and interactive theorem proving.
+Hercules is the user (you!), and hydra's heads are the subgoals: you may think that applying a tactic would solve a subgoal, but it results often in the multiplication of such tasks.}.
 There are now several sites on the Internet where you can find tutorials on hydra games, together with simulators you can play with.
 See, for instance, the blogpost and source code written by Andrej Bauer~\cite{bauer2008,BauerHydra}.
-
-Let us underline the analogy between hydra battles and interactive theorem proving.
-Hercules is the user (you!), and hydra's heads are the subgoals: you may think that applying a tactic would solve a subgoal, but it results often in the multiplication of such tasks.
 
@@
 
-\begin{scope}[xshift=-6cm]
+\begin{scope}[xshift=-6.5cm]
 \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (additions) { Additions };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (cantor) { Cantor };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=1.9cm] (pock) { Pocklington };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=1.8cm] (additions) { Additions };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=1.8cm] (cantor) { Cantor };
 
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3.2cm] (cats) { Categories in ZFC };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=2.2cm] (cats) { Categories in ZFC };
 \end{scope}
 
-\begin{scope}[yshift=-1cm,xshift=0.25cm]
+\begin{scope}[yshift=-1cm,xshift=-4.25cm]
   \node[rectangle, draw=black,fill=blue!20,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=1.8cm] (mathcomp) { MathComp };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=5.8cm] (mathcomp) { MathComp };
 \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=1.5cm,left of=paramcoq, node distance=4.5cm] (equations) { Equations };
@@
 
-Quotations of \coq{} source  from Standard Library and code which lead to compilation error or too long computations are displayed as follows.
+Quotations of \coq{} source  from others libraries (\coq's standard library, borrowed plug-ins)  are displayed as follows.
 
@@
 \end{Coqsrc}
+
+We use also verbatim code inclusions when the examples lead to very long computations.
 

### 2022-03-12 doc/part-hydras.tex
re-reading pages 9-21

@@
 
-\subsection*{Compatibility with \gaia\cite{Gaia} (in progress)}
-The \gaia project~\cite{Gaia} by Jos\'e Grimm aimed to formalize mathematics in \coq  in
+\subsection*{Compatibility with \gaia (in progress)}
+The \gaia project~\cite{Gaia} by Jos\'e Grimm, Alban Quadrat,  and Carlos Simpson,
+aims  to formalize mathematics in \coq  in
 the style of Nicolas Bourbaki.
@@
 In Chapter~\ref{gaia-chapter}, we present some modules which allow \HydrasLib' users to apply lemmas proven in \gaia, and vice versa.
-Remarks about compatibility with \gaia signalled with the picture \gaiasign.
+Remarks about compatibility with \gaia are signalled with the picture \gaiasign.
 
@@
 \item Semiautomatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
-\item Extension to $\Gamma_0$ (in Veblen normal form)
+%\item Extension to $\Gamma_0$ (in Veblen normal form)
 \item More lemmas about hierarchies of rapidly growing

### 2022-03-12 doc/hydras.tex
re-reading pages 9-21

@@
 \item Local links to pages generated by \texttt{coqdoc}. According to the current makefile (through the commands \texttt{make html} and \texttt{make pdf}), 
-  we assume that the page generated from a library \texttt{XXX/YYY.v} is stored as
-the relative address \texttt{../theories/html/hydras.XXX.YYY.html} (from the location of the pdf)
+  the pages generated by \texttt{coqdoc} are stored at 
+the relative address \texttt{../theories/html/*.html} (from the location of the pdf).
 Thus,  active links to our \coq{} modules may be incorrect if you did not get this \texttt{pdf} document by compiling the distribution available at

### 2022-03-14 doc/part-hydras.tex
corrections chapter 2

@@
 In order to describe trees where each node can have an arbitrary (but finite) number of sons, it is usual to define a type where each node carries a \emph{forest}, \emph{i.e} a list of trees
-(see for instance Chapter 14, pages 400-406 of \cite{BC04,BC04ch14}).
+(see for instance Chapter 14, pages 400-406 of \cite{BC04}, also available as~\cite{BC04ch14}).
 
@@
 \subsubsection{Example}
-Let us prove the transformation of \texttt{Hy'} into \texttt{Hy''} (see Fig.~\vref{fig:Hy3}). We use an experimental set of tactics for specifying the place where the 
+Let us prove the transformation of \texttt{Hy'} into \texttt{Hy''} (see Fig.~\vref{fig:Hy3}).
+We use an experimental set of tactics\footnote{See the \texttt{Ltac} definitions in~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions.}}  for specifying the place where the
 interaction between Hercules and the hydra holds.
@@
 
-Prove that, if $h$ is not a simple head, then there exists a unique $h'$ such that \texttt{h}  is transformed into \texttt{ h'} in one round, according to this restriction.
+Prove that, if $h$ is not a simple head, then there exists a unique $h'$ such that $h$  is transformed into $h'$ in one round, according to this restriction.
 
@@
 
-\vspace{4pt}
+\vspace{6pt}
+\noindent
 \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
@@
 
-\vspace{4pt}
+\vspace{6pt}
+\noindent
 \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#free}{Hydra.Hydra\_Definitions}}
@@
 
-\vspace{4pt}
+\vspace{6pt}
+\noindent
 \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#standard}{Hydra.Hydra\_Definitions}}
-
 \input{movies/snippets/Hydra_Definitions/standardDef}
@@
 
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}
+\vspace{6pt}
+\noindent\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}
 
@@
 
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-
+\vspace{6pt}
+\noindent\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
 \noindent
@@
 
-\subsection{The beginning of hostilities}
+\subsection{First rounds}
 During the two first rounds, our hydra loses its two rightmost heads.  Figure~\vref{fig:hinit-plus2} shows the state of the hydra   just before the third round.
@@
 \subsection{Testing  \dots}
+\label{sect:testing}
 In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple data type for representing the states as tuples composed of
@@
 Let us define a binary relation associated with every round of the battle.
-In the following definition \texttt{i} is associated with the round number (or date, if we consider a discrete time), and \texttt{a}, \texttt{b}, \texttt{c} respectively associated with the number of \texttt{h2}, \texttt{h1} and heads connected to the hydra's foot.
+In the following definition \texttt{i} is associated with the round number (or date, if we consider a discrete time), and \texttt{a}, \texttt{b}, \texttt{c} respectively associated with the number of occurrences of \texttt{h2}, \texttt{h1} and heads connected to the hydra's foot.
 
@@
 \begin{enumerate}
-\item  Characterization of regularities of some events
+\item  Characterization and proofs of regularities of some events (inspired by our experiments of Sect.~\ref{sect:testing}).
 \item Study of the beginning of the battle
@@
 in a well-founded relation $<$  on some type \texttt{A}, and a function
-(also called a \emph{measure}) \texttt{m:Hydra->A} such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $B$, the inequality $m(h')<m(h)$ holds.
+(also called a \emph{measure}) $m:\texttt{Hydra}\rightarrow A$ such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $B$, the inequality $m(h')<m(h)$ holds.
 
@@
 
-\inputsnippets{ON_Generic/ONDefsa, ON_Generic/ONDefsb}
+\inputsnippets{ON_Generic/ONDefsa}
+\inputsnippets{ON_Generic/ONDefsb}
 

### 2022-03-15 doc/part-hydras.tex
corrections chapter 3 (up to p. 63)

@@
 We use it as a reference for proving the correctness of ordinal notations.
+  We are also progressively importing \gaia's~\cite{Gaia} definitions and theorems, based on Bourbaki's set theory~\cite{Grimm2016}.
+Chapter~\ref{gaia-chapter} is dedicated to the connexion between libraries \HydrasLib and \gaia.
+
 \item A family of \emph{ordinal notations} (also called \emph{[ordinal] notation systems}), \emph{i.e.} data types used to represent segments $[0,\mu)$, where $\mu$ is some countable ordinal.
@@
 
-It is interesting to compare proofs of a given property (for instance the associativity of addition) both in the computational framework of some ordinal notation, and in the axiomatic model of Schütte.
+It is interesting to compare proofs of a given property (for instance the associativity of addition) both in the computational framework of some ordinal notation, and in mathematical  models
+of Schütte or \gaia.
 
@@
 
-Such a structure can be proved correct relatively to a bigger ordinal notation or
-to Schütte's model.
+Such a structure may be proved correct relatively to a bigger ordinal notation or to Schütte's or \gaia's model.
 
@@
 
-\pagebreak
+%\pagebreak
 
@@
 
-\inputsnippets{ON_plus/compareDef,
-  ON_plus/compareCorrect, ON_plus/plusComp}
-
-The Lemma \texttt{Wellfounded.Disjoint\_Union.wf\_disjoint\_sum} of Standard Library
-helps us to prove that our order \texttt{lt} is well-founded, thenwe can build an instance of \texttt{ON}:
+\inputsnippets{ON_plus/compareDef, ON_plus/compareCorrect}
+\inputsnippets{ON_plus/plusComp}
+
+Standard library's lemma \texttt{Wellfounded.Disjoint\_Union.wf\_disjoint\_sum} 
+is applied  to prove that our order \texttt{lt} is well-founded, allowing us to build an instance of \texttt{ON}:
 
@@
 \inputsnippets{ON_Omega2/Omega2Def, ON_Omega2/Defs}
+\inputsnippets{ON_Omega2/Ex1}
 
@@
 \subsection{A proof of termination using \texorpdfstring{$\omega^2$}{omega^2}} 
-\label{sect:merge-example}.
+\label{sect:merge-example}
 
@@
 The type \texttt{fuel}, an alternative 
-to \texttt{nat} is available on \href{../theories/html/hydras.Prelude.Fuel.html}{Prelude.Fuel}).
+to \texttt{nat} is available on \href{../theories/html/hydras.Prelude.Fuel.html}{Prelude.Fuel})
+.
 %\index{coq}{Giving fuel to a long computation}

### 2022-03-16 doc/part-hydras.tex
corrections chapter 3

@@
 We  follow the same pattern as in Sect.~\ref{omega-case}.
-First, we define an injection $\iota$ from type \texttt{t} into \texttt{Hydra}, by
+First, we define an injection $\iota$ from type \texttt{ON\_Omega2.t} into \texttt{Hydra}, by
  associating to  each ordinal $\omega\times i+ j = (i,j)$ the hydra with $i$ branches of length $2$ and
@@
 
-Like in Sect.~\ref{omega-case}, we prove the double inequality \texttt{m big\_h o<= m small\_h o< m big\_h}, which is impossible.
+Like in Sect.~\ref{omega-case}, we prove the inequality $m\;\texttt{big\_h} \,\texttt{o<=}\, m\;\texttt{small\_h} \,\texttt{o<=}\, m\;\texttt{big\_h}$, which is obviously false.
 
@@
 
-The first steps of our proof prepare a well-founded induction on $\omega^2$.
+The first steps of our proof start a well-founded induction over $\omega^2$.
 
@@
 
-Then a case analysis on $i$ and $j$ allows us to
+A case analysis on $i$ and $j$ allows us to
 consider three cases :
@@
 \item $i=1+l, j=0$ ($(i,j)$ is a limit ordinal): By the induction hypothesis \texttt{IHij},
-  $(l,k)\leq m(\iota(l,k))$ for any $k$. But (by the rules of the hydra game), $\iota(i,0)$ is transformed into any $\iota(l,k)$ in one round. Thus $m(\iota(l,k)) < m(\iota(i,0))$ for any $k$.
+  $(l,k)\texttt{o<=} m(\iota(l,k))$ for any $k$. But (by the rules of the hydra game), $\iota(i,0)$ is transformed into any $\iota(l,k)$ in one round. Thus $m(\iota(l,k)) < m(\iota(i,0))$ for any $k$.
   Therefore, $(l,k) <  m(\iota(i,0))$ for any $k$, thus
-  $(i,0) \leq m(\iota(i,0))$.
- \item $j= l+1$  ($(i,j)$ is a successor).  a similar, but simpler case: we apply the induction hypothesis to the pair $(i,l)$.
+  $(i,0)\,\texttt{o<=}\,m(\iota(i,0))$.
+ \item $j= l+1$  ($(i,j)$ is a successor). We apply the induction hypothesis to the pair $(i,l)$.
  \end{itemize}
@@
 \subsubsection{End of the proof}
-From \texttt{m\_ge}, we get \texttt{m big\_h o<= m small\_h = m (iota (m big\_h)) }. 
+From \texttt{m\_ge}, we get \texttt{$m$\, big\_h o<= $m$\, small\_h = $m$\,(iota (m big\_h)) }. 
 Since $<$ is a strict order (irreflexive  and transitive), this inequality is incompatible with the strict inequality  \texttt{m small\_h o< m big\_h} (lemma \texttt{m\_lt}).
@@
 \begin{remark}
-In Chapter~\ref{ks-chapter}, we  prove a generalization of the impossibility lemmas of
+ We  prove in Chapter~\ref{ks-chapter} a generalization of the impossibility lemmas of
 Sect.~\ref{omega-case} and this section, with the same proof structure, but with much more 
@@
 
-\begin{remark}
-Finite ordinals are also formalized in MathComp~\cite{MCB}.  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.  
-\end{remark}
+\section{See also \dots}
+
+Finite ordinals are also formalized in \ssreflect/\mathcomp~\cite{MCB}.
+In Module~\href{../theories/html/gaia_hydras.ON_gfinite.html}{gaia\_hydras.ON\_gfinite}, we build an instance of class \texttt{ON}.
+
+\inputsnippets{ON_gfinite/finordLtDef, ON_gfinite/finordON}
+
+\inputsnippets{ON_gfinite/Examples, ON_gfinite/Examplesb}
+
+  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.
 
@@
 \begin{project}
+Same work, but replace Schütte's model with \gaia's.
+\end{project}
+
+\begin{project}
   When you have read Chapter~\ref{chap:schutte}, prove that the sum of two ordinal notations \texttt{ON\_plus} implements the addition of ordinals.
@@
 
-Prove formally  that \texttt{ON\_mult (FinOrd $i$) Omega} is isomorphic to
+Prove formally  that (\texttt{ON\_mult (FinOrd $i$) Omega}) is isomorphic to
 \texttt{Omega}  whilst
-\texttt{Omega}  is a sub-notation of \texttt{ON\_mult Omega (FinOrd $i$)},
+\texttt{Omega}  is a sub-notation of (\texttt{ON\_mult Omega (FinOrd $i$)}),
 for any strictly positive $i$.

### 2022-03-16 doc/part-hydras.tex
corrections chapter 4 (up to page 80)

@@
 
+The ordinal \(\epsilon_0\) is the least ordinal number that satisfies 
+the equation \(\alpha = \omega^\alpha\), where \(\omega\) is 
+the least infinite ordinal.
+Thus, we can intuitively consider \(\epsilon_0\) as an
+\emph{infinite} \(\omega\)-tower.
+
 \subsection{Cantor normal form}
@@
 
-The ordinal \(\epsilon_0\) is the least ordinal number that satisfies 
-the equation \(\alpha = \omega^\alpha\), where \(\omega\) is 
-the least infinite ordinal. Thus, we can consider \(\epsilon_0\) as an
-\emph{infinite} \(\omega\)-tower.
-Nevertheless, 
-any ordinal strictly less that \(\epsilon_0\) 
+Any ordinal strictly less that \(\epsilon_0\) 
 can be finitely represented by a unique  \emph{Cantor normal form}, 
@@
 to a hierarchy of ordinal notations.
-For instance, in Library \href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2},  the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{schutte}).
+For instance, in Library \href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2},  the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{Gaia, schutte}).
 \noindent
@@
 \gaiasign Chapter~\vref{gaia-chapter} describes the present state of a
-project of making compatible the libraries \HydrasLib and \gaia. At present, both libraries contain a version of the inductive type \texttt{T1}, with the same base name for the constructors.
+project of making compatible the libraries \HydrasLib and \gaia. At present, both libraries contain their own version of the inductive type \texttt{T1}, with the same base name for the constructors.
 
@@
 \paragraph*{\gaiasign}
-  In \texttt{gaia.ssete9}, the $n$-th finite ordinal is written
+  In \texttt{gaia.ssete9}, the $n$-th finite ordinal is also written
   \texttt{$\backslash${F}\,$n$}. There is no coercion in \gaia from \texttt{nat} to finite ordinals.
@@
 
-Note that \texttt{omega} is not an identifier, thus any tactic like \texttt{unfold omega} would fail.
+Note that \texttt{omega} is not an identifier in \HydrasLib, thus any tactic like \texttt{unfold omega} would fail.
 
 \paragraph*{\gaiasign}
-In \texttt{gaia.ssete9}, the ordinal $\omega$ is also named as \emph{constant}  \texttt{T1omega}.
+In \texttt{gaia.ssete9}, the ordinal $\omega$ is bound to the \emph{constant}  \texttt{T1omega}.
 
@@
 \paragraph*{\gaiasign}
-In \texttt{gaia.ssete9} the identifier\texttt{phi0} is bound to a constant (not a notation).
+In \texttt{gaia.ssete9} the identifier \texttt{phi0} is bound to a plain constant (not a notation).
 
@@
 
-Let us consider again the ordinal $\alpha_0$ defined in section~\vref{alpha0-def}
+Let us consider again the ordinal $\alpha_0$ defined in section~\vref{alpha0-def}. 
 If we ask \coq{} to print its  normal form, we get a hardly readable term of type \texttt{T1}.
@@
 
-The following data type defines an abstract syntax for more readable ordinals terms in Cantor normal form:
+The following data type defines a more readable abstract syntax for  ordinals terms in Cantor normal form:
 
@@
 
-In Sect.~\vref{nf-gaia-compat}, we show that \gaia's \texttt{T1nf} is extensionnaly equivalent with \HydrasLib'  \texttt{nf}.
+In Sect.~\vref{nf-gaia-compat}, we show that \gaia's \texttt{T1nf} is extensionally equivalent with \HydrasLib'  \texttt{nf}.
 

### 2022-03-17 doc/part-hydras.tex
corrections chapter 4 (up to page 86)

@@
 
-\subsubsection{A sigma-type for \texorpdfstring{$\epsilon_0$}{epsilon0}}
+\subsubsection{\texttt{E0}: a sigma-type for \texorpdfstring{$\epsilon_0$}{epsilon0}}
 
@@
 
-\emph{From Module~w}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html\#E0}{Epsilon0.E0}}
 
@@
 
-First, we declare a notation scope for \texttt{E0}.
+First, we declare a notation scope for \texttt{E0}, then we redefine the predicates of comparison.
 
 \input{movies/snippets/E0/E0Scope}
-
-Then we redefine the predicates of comparison.
-
 \label{Predicates:Lt-E0}
-
 \input{movies/snippets/E0/LtLeDef}
@@
 
-\begin{exercise}[\gaia]
+\begin{exercise}[\gaiasign]
   Look for the \gaia-theorem which corresponds to \texttt{succb\_iff}.

### 2022-03-17 doc/part-hydras.tex
corrections chapter 4

@@
 It is natural to try to prove by structural induction over \texttt{T1} 
-that every term in normal form is accessible through \texttt{LT}.
+that every term in normal form is \texttt{LT}-accessible.
 
 Unfortunately, it won't work. Let us consider some well-formed term
- $\alpha=\texttt{cons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are accessible
- through \texttt{LT}. For proving the accessibility of $\alpha$, we have to consider
+ $\alpha=\texttt{cons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are \texttt{LT}-accessible.
+In order to prove the accessibility of $\alpha$, we have to consider
 any well formed term \(\delta\) such that \(\delta<\alpha\). 
@@
 
-The problem comes from the hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
+The problem comes from the too weak hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
 \(\gamma\);
@@
   Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
-through \texttt{LT}, we propose to show first that any well formed 
-term of the form \(\omega^\alpha\times(n+1)+\beta\) is accessible (which is a stronger result).
+through \texttt{LT}, we propose to consider the following (stronger) predicate:
 
@@
 
-As a corollary, the order \texttt{Lt} on type \texttt{Ord} is well-founded too.
-
-\inputsnippets{E0/E0LtWf}.
+As a corollary, the order \texttt{Lt} on type \texttt{E0} is well-founded too.
+
+\inputsnippets{E0/E0LtWf}
 
@@
 
-We build an instance of \texttt{ON}, and prove its correctness w.r.t. Schutte's model.
+We build an instance of \texttt{ON}, and prove its correctness w.r.t. Schutte's model (see Chapter~\ref{chap:schutte}).
 
@@
     \index{hydras}{Exercises}
+
+    \subsection{Related work}
+    The article~\cite{Manoury2016}  defines another representation of ordinals below $\omega^\omega$ based on lists of natural numbers.
 
@@
 
-  \pagebreak 
+%  \pagebreak 
   \vspace{4pt}
@@
 
-    \subsubsection{Monotonicity}
+    \subsubsection{Monotony}
 
@@
     \end{project}
+
+    \paragraph*{\gaiasign} The module
+      \href{../theories/html/gaia_hydras.~GHessenberg.html}{gaia\_hydras.GHessenberg} defines a version of Hessenberg's sum
+      compatible with \gaia's type \texttt{T1}.
 
@@
           $\alpha$ is greater than or equal to the omega-tower of height $j$.
-         By monotonicity,  $\phi_0(\alpha)$ is greater than or equal to 
+         By monotony,  $\phi_0(\alpha)$ is greater than or equal to 
         the omega-tower of height $j+1$, thus strictly greater than $\alpha$

### 2022-03-17 doc/part-hydras.tex
corrections chapter 5 (up to p. 106)

@@
 \begin{itemize}
-\item There is no variant mapping the type \texttt{Hydra} into the interval $[0,\omega^2)$ (section ~\vref{omega2-case}), and a fortiori 
-$[0,\omega)$ (section ~\vref{omega-case}).
+\item There is no variant mapping the type \texttt{Hydra} into the interval $[0,\omega^2)$ (section ~\vref{omega2-case}), and \emph{a fortiori} $[0,\omega)$ (section ~\vref{omega-case}).
 
@@
  an article by J.~Ketonen and R.~Solovay~\cite{KS81}, already  cited in the work
- by L.~Kirby and J.~Paris.% on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
+ by L.~Kirby and J.~Paris. % on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
  Section $2$ of this article: ''A hierarchy of probably recursive functions'', contains a systematic study of \emph{canonical sequences}, which are closely related to
@@
 
-\subsubsection{Canonical sequences in Coq}
+\subsubsection{Canonical sequences in \coq}
 \index{hydras}{Library Epsilon0!Functions!canon}
@@
 \vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canon}{Epsilon0.Canon}}
+\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canon}{Epsilon0.Canon}}
 
@@
 \input{movies/snippets/Canon/canonDef}
+
+\paragraph*{\gaiasign} Our library~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon} transposes the  function \texttt{canon} into \gaia's type \texttt{T1}.
 
@@
 abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
-~\href{../theories/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}
-
+~\href{../theories/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}.
 For instance, the equality $\canonseq{\alpha+1}{i}=\alpha$  can be  proved by  structural induction on $\alpha$.
@@
 
-We prove by transfinite induction over $\alpha$ that $\canonseq{\alpha}{i+1}$ is an ordinal strictly less than $\alpha$ (assuming $\alpha\not=0$).
+We prove by transfinite induction over $\alpha$ that $\canonseq{\alpha}{i}$ is an ordinal strictly less than $\alpha$ (assuming $\alpha\not=0$).
 This property allows us to use the function \texttt{canonS} and its derivatives in function definitions by transfinite recursion.
 
-\label{lemma:canonS_LT}
-
-\input{movies/snippets/Canon/canonSLT}
+\label{lemma:canon_LT}
+
+\inputsnippets{Canon/canonLT}
+
+\paragraph*{\gaiasign} This lemma is also available in
+Library ~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon}:
+
+\inputsnippets{GCanon/gcanonLt}
 
@@
 \input{movies/snippets/Canon/canonSLimitLub}
+
+\paragraph*{\gaiasign}  In \gaiaHydras, the statement use a slightly different vocabulary:
+\inputsnippets{GCanon/gcanonLimitOf}
 
@@
 \noindent 
-From the lemma \texttt{canonS\_LT}~\vref{lemma:canonS_LT}, we can convert any path into an inequality on ordinals (by induction on paths).
+From the lemma \texttt{canon\_LT}~\vref{lemma:canon_LT}, we can convert any path into an inequality on ordinals (by induction on paths).
 

### 2022-03-18 doc/gaia-chapter.tex
corrections chapter 5 (completed)

@@
 \subsection{Canonical sequences}
+\label{sect:gcanon}
 Canonical sequences are described in Chapter~\vref{chap:ketonen}, and implemented in
@@
 \subsection{Accessibility in \texorpdfstring{$\epsilon_0$}{epsilon\_0}}
+\label{sect:gpath}
 
@@
 
-If $\beta$ is (strictly) accessible from $\alpha$, then
+$\beta$ is (strictly) accessible from $\alpha$, if and only iff
 $\beta<\alpha$.
@@
 \inputsnippets{GPaths/pathToLT}
+\inputsnippets{GPaths/LTPathTo}
 

### 2022-03-18 doc/part-hydras.tex
corrections chapter 5 (completed)

@@
 
-\paragraph*{\gaiasign} Our library~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon} transposes the  function \texttt{canon} into \gaia's type \texttt{T1}.
-
-In the present state of this library, the following specializations of \texttt{canon} are still used in some proofs or lemma statements.
+\paragraph*{\gaiasign} Our library~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon} transposes the  function \texttt{canon} into \gaia's type \texttt{T1}
+(please see Sect.\~ref{sect:gcanon}).
+
+\paragraph*{Remark}
+In the present state of this library, the following specializations of \texttt{canon} are still used in some proofs or lemma statements. They are planned to be deprecated.
 
@@
 
-\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0} : paths}
+\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0}\,: paths}
 \index{maths}{Ordinal numbers!Accessibility inside epsilon0}
@@
 existence of such a path with the notation $\alpha\xrightarrow [s]{}\beta$.
-
 We say also that the considered path from $\alpha$ to $\beta$ \emph{starts at [index] $s_0$ and ends at $s_l$}.
@@
 Note that, given $\alpha$ and $\beta$, where $\beta < \alpha$, the sequence $s$ which leads from $\alpha$ to $\beta$ is not unique.
-
 Indeed, if $\alpha$ is a limit ordinal, the first element of $s$ can be any integer $i$ such that $\beta<\canonseq{\alpha}{i}$, and if $\alpha$ is a successor ordinal,
@@
 \end{remark}
+
+\paragraph*{\gaiasign}
+The library~\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} transposes the  notion of path into \gaia's type \texttt{T1}
+(please see Sect.\~ref{sect:gpath}).
 
@@
 
+\vspace{4pt}
+\noindent
+From~\href{../theories/html/hydras.Hydra.~Hydra_Definitions.html}{Hydra.Hydra\_Definitions}:
+
 \input{movies/snippets/Hydra_Definitions/BoundedVariant}
@@
 
-\pagebreak
-
 \inputsnippets{Epsilon0_Needed_Free/mGe}
@@
 The sequence $s$, used to build the sequence of replication factors of the battle, depends on 
-$\beta$, so we cannot be sure that the generated battle is a genuine standard battle.
-
-The solution of this issue comes  once again from Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
+$\alpha$ and $\beta$, so we cannot be sure that the generated battle is a genuine standard battle.
+
+The solution to this issue comes  once again from Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
 $\alpha_0=\alpha,\alpha_1,\dots,\alpha_k=\beta$ where $\alpha_{j+1}$ is equal
@@
 
+\vspace{4pt}
+\noindent
 \emph{From
@@
 
+ \paragraph*{\gaiasign}
+ Corollary \texttt{Cor12} and Lemma \texttt{Lemma261} are also available in
+\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} 
+(please see Sect.\~ref{sect:gpath}).
+
 \subsection{Casting paths with a constant index into a standard path}
@@
 
-Our proof follows the proof by Ketonen and Solovay, including its organization as a sequence of lemma.  Since it is a non-trivial proof, we will comment its main steps below.
+Our proof follows the proof by Ketonen and Solovay, including its organization as a sequence of lemma.
+Since it is a non-trivial proof, we will comment its main steps below   (see Figure~\vref{fig:belle-preuve-1} to Figure~\vref{fig:fin-belle-preuve}).
 
@@
 
-Applying \texttt{standard\_path\_to\_zero}, $0$ is reachable from $\alpha$ by some standard path  (see figure~\vref{fig:belle-preuve-1}).
+Applying \texttt{standard\_path\_to\_zero}, $0$ is reachable from $\alpha$ by some standard path.
 
@@
 \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}{Hydra.Epsilon0\_Needed\_Std}.
-Please note that it has the same global structure as in section\ref{sec:free-battles-case} 
+Please note that it has the same global structure as in section~\ref{sec:free-battles-case} 
 % ICI !

### 2022-03-18 doc/part-hydras.tex
corrections chapter 6 (up to page 130)

@@
 
-\input{movies/snippets/Large_Sets/largeDef}
+\input{movies/snippets/Large_Sets/mlargeDef}
 
@@
 \input{movies/snippets/Large_Sets_Examples/gnawEx2}
-
-We may say that the interval $[6,700]$ is $\omega^2$-large, since it leads to $0$, but nothing assures us that the condition of minimality is satisfied.
 
@@
 \input{movies/snippets/Large_Sets/mlargeIff}
-
-For instance, we prove that the interval $[6,510]$ is $\omega^2$-mlarge.
 
@@
 Our exploration of the $L_\alpha$\,s  considers the usual cases of a proof by transfinite induction: zero, successors and limit ordinals.
-The lemmas we are going to proved will be applied in a big proof by induction in Sect~\vref{sect:L-correct-proof}.
+The lemmas we are going to prove will be applied in a big proof by induction in Sect~\vref{sect:L-correct-proof}.
 
@@
 By iterating the functional \texttt{L\_succ}, we get a realization of
-\texttt{(L\_spec (fin $i$))} for any natural number $i$.
+\texttt{(L\_spec $i$)} for any natural number $i$.
 
@@
 
-The following lemma expresses the length of the mlarge sequences associated with the finite multiples of $\omega$.
+The following lemma expresses the length of the mlarge sequences associated with any finite multiple of $\omega$.
 
@@
 Let us consider a last example, ``computing'' $L_{\omega^3}$.
-Since the canonical sequence associated with this ordinal is composed of the
-$\omega^2\times i\;(i\in\mathbb{N}_1)$, we have to study this sequence.
+Let us study the canonical sequence associated with $\omega^3$, the elements of which are 
+$\omega^2\times i\;(i\in\mathbb{N}_1)$.
 
@@
 
-\inputsnippets{Large_Sets/phi0Mult, Large_Sets/phi0MultOk,
+\inputsnippets{Large_Sets/phi0Mult}
+
+\inputsnippets{Large_Sets/phi0MultOk,
 Large_Sets/Phi0MultSLe}

### 2022-03-18 doc/part-hydras.tex
corrections chapter 6 (up to page 137)

@@
 
-We are now ready to get an exact formula for $L_{\omega^3}$, by diagonalization upon $L_{\omega^2\times i}$.
+We are now ready to get an exact formula for $L_{\omega^3}$, by diagonalization over $L_{\omega^2\times i}$.
 
@@
 
-Crossing a new limit, we prove the following equality: 
-$$H'_{\omega^2} (k) = 2 ^ {k+1} \times (k+1) - 1$$.
+Since $\omega^2$ is a limit ordinal, we prove the following equality: 
+\[H'_{\omega^2} (k) = 2 ^ {k+1} \times (k+1) - 1\]
 
@@
 
-By well-founded induction on $\alpha$, we prove a simple relation between $L_\alpha$ and $H'\_alpha$.
-
+By well-founded induction on $\alpha$, we prove that
+ $H'_\alpha(i) \leq L_\alpha(i+1)$ for any $i$.
+
+\noindent
 \emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#H'_L_}{Epsilon0.L\_alpha}}

### 2022-03-19 doc/part-hydras.tex
corrections chapter 6 (completed)

@@
 The difference with the ``classic'' Wainer hierarchy 
-$f\_\alpha\;(\alpha<\epsilon_0)$ lies in the second equation:
+$f_\alpha\;(\alpha<\epsilon_0)$ lies in the second equation:
 $f_{\beta+1}(i) = (f_\beta)^{(i)}(i)$ and not
@@
 
-A first attempt is to write a definition of $F_\alpha$ by equations, in the same way as for $H\_alpha$ (the functional \texttt{iterate} has already been used in Sect.\vref{Functions:iterate}).
+A first attempt is to write a definition of $F_\alpha$ by equations, in the same way as for $H_\alpha$ (the functional \texttt{iterate} has already been used in Sect.\vref{Functions:iterate}).
 
@@
 On the other hand, please show that for any natural number $n$, the function $F_n$ is primitive recursive.
-Thus $F\_alpha$ is primitive recursive if and only if $\alpha$ is finite.
+Thus $F_\alpha$ is primitive recursive if and only if $\alpha$ is finite.
 
@@
 \end{exercise}
+
+\paragraph*{\gaiasign} Library~\href{../theories/html/gaia_hydras.GF_alpha.html}{gaia\_hydras.GF\_alpha}  defines the functions $F_\alpha$ for \gaia's world.
 

### 2022-03-19 doc/part-hydras.tex
corrections chapter 7 (completed)

@@
 Since we won't be able to prove the proposition
-\linebreak \Verb|{exists! a: Ord, least_member (Empty_set Ord) a|, the only properties we would be able to prove about \texttt{bottom} would be \emph{trivial} properties, 
+\linebreak \Verb|(exists! a: Ord, least_member (Empty_set Ord) a)|, the only properties we would be able to prove about \texttt{bottom} would be \emph{trivial} properties, 
 \emph{i.e.}, satisfied by \emph{any} element of type \texttt{Ord}, like for instance
@@
 
+\begin{todo}
+  Build a T2Bridge! (note: Gamma0's wellfoundedness is missing in \gaia (?).
+\end{todo}
+
 \section{Introduction}

### 2022-03-25 doc/part-hydras.tex
minor improvements

@@
 
+First, let us prove the following lemma (numbered \texttt{2.4} in Ketonen and Solovay's article).
+
+\input{movies/snippets/Paths/KSThm24}
+
 We prove this lemma by transfinite induction on $\alpha$.
@@
 $\alpha  \xrightarrow [n]{}  \gamma \xrightarrow [n]{} \beta$
-\item If $\alpha$ is a limit ordinal, we apply the following theorem (numbered \texttt{2.4} in Ketonen and Solovay's article).
+\item If $\alpha$ is a limit ordinal, we apply Theorem \texttt{KS\_thm\_2\_4} (see above).
 
@@
 %   \end{theorem}
-
-\input{movies/snippets/Paths/KSThm24}
 
@@
 
-Why is \texttt{Cor12} so useful? 
-Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
-we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying corollary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
-where $n$ is the maximum of the indices $i$ met in the induction.
-
- Lemma 1, Section 2.6 of~\cite{KS81} is naturally expressed in terms of \coq's
-\verb@sig@ construct.
-
-\label{lemma:L-2_6-1}
-%\index{coq}{Sigma types}
-
-\input{movies/snippets/Paths/Lemma261}
-
-Intuitively, lemma   \texttt{Lemma2\_6\_1}  shows that if $\beta<\alpha<\epsilon_0$, then there exists  a battle from $\iota(\alpha)$ to $\iota(\beta)$ where the replication factor is constant, although large enough.
-
 \begin{remark}
@@
 
+Why is \texttt{Cor12} so useful? 
+Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
+we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying corollary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
+where $n$ is the maximum of the indices $i$ met in the induction.
+
+ Lemma 1, Section 2.6 of~\cite{KS81} is naturally expressed in terms of \coq's
+\verb@sig@ construct.
+
+\label{lemma:L-2_6-1}
+%\index{coq}{Sigma types}
+
+\input{movies/snippets/Paths/Lemma261}
+
+Intuitively, Lemma~\texttt{2\_6\_1}  shows that if $\beta<\alpha<\epsilon_0$, then there exists  a battle from $\iota(\alpha)$ to $\iota(\beta)$ where the replication factor is constant, although large enough.
+
  \paragraph*{\gaiasign}
- Corollary \texttt{Cor12} and Lemma \texttt{Lemma261} are also available in
+ Corollary \texttt{Cor12} and Lemma \texttt{Lemma2\_6\_1} are also available in
 \href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} 
-(please see Sect.\~ref{sect:gpath}).
+(please see Sect.~\vref{sect:gpath}).
 

### 2022-04-06 doc/gaia-chapter.tex
corrections chapter 7 (completed)
make E0 an eqtype
simplification of notations

@@
 
-The modules are named {\texttt{theories/gaia/*.v}} and coqdocumented in {\texttt{theories/html/gaia\_hydras.*.html}}
+The modules are named {\texttt{theories/gaia/*.v}} and coqdocumented in \linebreak {\texttt{theories/html/gaia\_hydras.*.html}}
 The main entry point is
@@
 
-\inputsnippets{T1Bridge/limitbRef, T1Bridge/isSuccRef}
+\inputsnippets{T1Bridge/limitbRef}
+\inputsnippets{T1Bridge/isSuccRef}
 
@@
 \subsection{A type for well-formed ordinal terms}
-
+\label{sect:E0-gaia-hydras}
 For compatibility's sake, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}.
@@
 \inputsnippets{T1Bridge/E0plusMultDef}
-
-\inputsnippets{T1Bridge/E0EqP}
 
@@
 
+\section{Already ported definitions and lemmas}
+
+\textbf{These tables are still in construction.}
+
+\begin{table}[H]
+  \centering
+  \begin{threeparttable}
+    \caption{Types for ordinals less than $\epsilon_0$}
+\begin{tabular}{|r | c|c|c|c|l|}
+\hline
+Hydras & page & Gaia-Hydras& page& Description \\\hline
+  \texttt{T1.T1} 
+       &  \pageref{types:T1}
+              & \texttt{CantorOrdinal.T1}
+                           &  & ordinal terms \tnote{1} \\
+
+  \texttt{E0.E0}
+       & \pageref{types:E0}
+              & \texttt{T1Bridge.E0}
+                           & \pageref{sect:E0-gaia-hydras}
+                                 & ordinal terms in Cantor normal form \\
+
+  \texttt{ON\_Finite.t}
+       & \pageref{def:Finite-ord-type}
+              & \ssreflect's finite ordinals \tnote{2}&  & ordinals below $\omega$ \\
+\hline
+\end{tabular}
+\begin{tablenotes}
+  \item[1] From \gaia library.
+   \item[2] From \ssreflect/\mathcomp libraries.
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
 \section{Importing a theorem from \gaia}

### 2022-04-06 doc/hydras.tex
corrections chapter 2
corrections chapter 5 (up to p. 106)
simplification of notations

@@
 \newcommand{\gaiacompat}{\textcolor{blue}{Gaia}}
-\newcommand{\gaiasign}{\texorpdfstring{\mycircled[cyan!40]{{\mbox{\color{blue}{\scriptsize{G}}}}}}{Gaia-Hydra}}
+\newcommand{\gaiasign}{\texorpdfstring{\mycircled[orange!40]{{\mbox{\color{blue!65}{\scriptsize{G}}}}}}{Gaia-Hydra}}
 \newtheorem{gaiahydra}{\gaiasign}
@@
 
-Our library \gaiaHydras is dedicated to make compatible our lemmas with José Grimm's \gaia project (designed for \ssreflect/\mathcomp).
+Our library \gaiaHydras is dedicated to make compatible our lemmas with José Grimm's \gaia project (designed for \ssreflect/\mathcomp) (please look at Sect.~\vref{sect:gaia-first-intro} and the paragraphs signalled with {\gaiasign}).
 
@@
 Bad definitions 
- are ''masked '' inside modules called \texttt{Bad}, \texttt{Bad1}, etc.
+ are ''masked'' inside modules called \texttt{Bad}, \texttt{Bad1}, etc.
 

### 2022-04-06 doc/part-hydras.tex
simplification of notations

@@
  \subsection*{Compatibility with \gaia (in progress)}
+ \label{sect:gaia-first-intro}
 The \gaia project~\cite{Gaia} by Jos\'e Grimm, Alban Quadrat,  and Carlos Simpson,
@@
 
-\label{def: Finite-ord-type}
+\label{def:Finite-ord-type}
 
@@
 
+\paragraph*{\gaiasign}
+The module
+\href{../theories/html/gaia_hydras.GL_alpha.html}{gaia\_hydras.GL\_alpha} contains an adaptation of
+\href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0/L\_alpha}.
+
 \subsubsection{Back to hydra battles}
@@
 
-%ICI
+\paragraph*{\gaiasign}
+The module
+\href{../theories/html/gaia_hydras.GHprime.html}{gaia\_hydras.GHprime} contains an adaptation of
+\href{../theories/html/hydras.Epsilon0.Hprime.html}{Epsilon0.Hprime}.
 
@@
 
+%ICI
 \subsection{Comparison between \texttt{L\_} and \texttt{H'\_} }

### 2022-04-06 doc/gaia-chapter.tex
simplification of notations in theories/gaia

@@
 
-Many constants are common to \HydrasLib and \gaia. The following notations allow us to get rid of possible ambiguities, still
-keeping with the prefix convention: ``h'' for \HydrasLib, ``g'' for \gaia.
+Many constants are common to \HydrasLib and \gaia. For instance, the name \texttt{zero} may refer to \gaia's
+\texttt{gaia.ssete9.CantorOrdinal.zero}  and
+\HydrasLib' \texttt{hydras.Epsilon0.T1.zero}. By default, the former will be simply called \texttt{zero} and the latter \texttt{T1.zero}.
+
+There are a few exceptions to this convention.
 
@@
 
+\begin{table}[H]
+  \centering
+  \begin{threeparttable}
+    \caption{Main constants}
+\begin{tabular}{|r | c|c|c|c|l|}
+\hline
+Hydras & page & Gaia-Hydras& page& Description \\\hline
+  \texttt{T1.zero}
+       &  \pageref{types:T1}
+              & \texttt{CantorOrdinal.zero \tnote{1}}
+                           &  & The ordinal $0$  \\
+
+\hline
+\end{tabular}
+\begin{tablenotes}
+  \item[1] From \gaia library.
+   \item[2] From \ssreflect/\mathcomp libraries.
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
 \section{Importing a theorem from \gaia}

### 2022-04-06 doc/gaia-chapter.tex
simplification of notations in theories/gaia

@@
 
-\section{Already ported definitions and lemmas}
+% \section{Already ported definitions and lemmas}
 
-\textbf{These tables are still in construction.}
+% \textbf{These tables are still in construction.}
 
-\begin{table}[H]
-  \centering
-  \begin{threeparttable}
-    \caption{Types for ordinals less than $\epsilon_0$}
-\begin{tabular}{|r | c|c|c|c|l|}
-\hline
-Hydras & page & Gaia-Hydras& page& Description \\\hline
-  \texttt{T1.T1} 
-       &  \pageref{types:T1}
-              & \texttt{CantorOrdinal.T1}
-                           &  & ordinal terms \tnote{1} \\
+% \begin{table}[H]
+%   \centering
+%   \begin{threeparttable}
+%     \caption{Types for ordinals less than $\epsilon_0$}
+% \begin{tabular}{|r | c|c|c|c|l|}
+% \hline
+% Hydras & page & Gaia-Hydras& page& Description \\\hline
+%   \texttt{T1.T1} 
+%        &  \pageref{types:T1}
+%               & \texttt{CantorOrdinal.T1}
+%                            &  & ordinal terms \tnote{1} \\
 
-  \texttt{E0.E0}
-       & \pageref{types:E0}
-              & \texttt{T1Bridge.E0}
-                           & \pageref{sect:E0-gaia-hydras}
-                                 & ordinal terms in Cantor normal form \\
+%   \texttt{E0.E0}
+%        & \pageref{types:E0}
+%               & \texttt{T1Bridge.E0}
+%                            & \pageref{sect:E0-gaia-hydras}
+%                                  & ordinal terms in Cantor normal form \\
 
-  \texttt{ON\_Finite.t}
-       & \pageref{def:Finite-ord-type}
-              & \ssreflect's finite ordinals \tnote{2}&  & ordinals below $\omega$ \\
-\hline
-\end{tabular}
-\begin{tablenotes}
-  \item[1] From \gaia library.
-   \item[2] From \ssreflect/\mathcomp libraries.
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
+%   \texttt{ON\_Finite.t}
+%        & \pageref{def:Finite-ord-type}
+%               & \ssreflect's finite ordinals \tnote{2}&  & ordinals below $\omega$ \\
+% \hline
+% \end{tabular}
+% \begin{tablenotes}
+%   \item[1] From \gaia library.
+%    \item[2] From \ssreflect/\mathcomp libraries.
+% \end{tablenotes}
+% \end{threeparttable}
+% \end{table}
 
-\begin{table}[H]
-  \centering
-  \begin{threeparttable}
-    \caption{Main constants}
-\begin{tabular}{|r | c|c|c|c|l|}
-\hline
-Hydras & page & Gaia-Hydras& page& Description \\\hline
-  \texttt{T1.zero}
-       &  \pageref{types:T1}
-              & \texttt{CantorOrdinal.zero \tnote{1}}
-                           &  & The ordinal $0$  \\
+% \begin{table}[H]
+%   \centering
+%   \begin{threeparttable}
+%     \caption{Main constants}
+% \begin{tabular}{|r | c|c|c|c|l|}
+% \hline
+% Hydras & page & Gaia-Hydras& page& Description \\\hline
+%   \texttt{T1.zero}
+%        &  \pageref{types:T1}
+%               & \texttt{CantorOrdinal.zero \tnote{1}}
+%                            &  & The ordinal $0$  \\
 
-\hline
-\end{tabular}
-\begin{tablenotes}
-  \item[1] From \gaia library.
-   \item[2] From \ssreflect/\mathcomp libraries.
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
+% \hline
+% \end{tabular}
+% \begin{tablenotes}
+%   \item[1] From \gaia library.
+%    \item[2] From \ssreflect/\mathcomp libraries.
+% \end{tablenotes}
+% \end{threeparttable}
+% \end{table}
 

### 2022-04-11 doc/gaia-chapter.tex
small snippet change
corrections in gaia chapter

@@
 In a future version, it would be nice to have our proofs in the same context, after porting \HydrasLib' proofs under \gaia's
-  definition, following  proof repair  techniques~\cite{ringer2021}.
+  definition, following  proof re-use or repair  techniques~\cite{Barthe2001, Magaud03a, ringer2021}.
 
@@
 \href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge}, which requires both Gaia's \texttt{ssete9} and a few modules of \texttt{theories/ordinals/Epsilon0}.
-Other modules are dedicated to the adaptation of several notions of \HydrasLib to \mathcomp's and \gaia's vocabulary and structures.
+Other modules are dedicated to the adaptation of several notions developped in \HydrasLib to \mathcomp's and \gaia's vocabulary and structures.
 By default, we give priority to \gaia's notation and vocabulary.
@@
 Because of the common origin of \HydrasLib and \gaia, both libraries share the common name \texttt{T1} (data type of ordinal terms below $\epsilon_0$).
-As mentionned before, the bare name \texttt{T1} is given by priority to \gaia, and we prefix this name with an \texttt{h} (for ``hydras'') to refer to
-\href{../theories/html/Epsilon0.T1.html}{hydras.Epsilon0.T1.T1}.
+As mentionned before, the bare name \texttt{T1} is given by priority to \gaia.
+\HydrasLib' type \texttt{T1} may be called \texttt{hT1}
+(\texttt{h} for ``hydras'') or simply \texttt{T1.T1}, \texttt{Epsilon0.T1.T1}, etc.
 
@@
 
-There are a few exceptions to this convention.
+The following notations are not defined in \gaia; we add them to \texttt{T1Bridge} for simplicity's sake.
 
@@
 
+\begin{remark}
 Please keep in mind that, in this module, we favour \gaia's notations over hydra's.
-A snippet like \texttt{(alpha < succ beta)} should be interpreted as \texttt{(T1lt alpha beta)}, where \texttt{T1lt} is defined in ~\href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{\texttt{gaia.ssete9.v}}.
+For instance, the mathematical inequality $42<\omega$ is 
+expressed in \gaia's terms as follows:
+
+\inputsnippets{HydraGaia_Examples/F42Omega}
+
+The same inequality, with \HydrasLib vocabulary:
+
+\inputsnippets{HydraGaia_Examples/F42Omegab}
+\end{remark}
 
@@
 \inputsnippets{HydraGaia_Examples/ExamplesG2H}
+
+The following cancel lemmas will be often applied in order to simplify sub-terms of the form (\texttt{g2h (h2g {\it t})}) and (\texttt{h2g (g2h {\it t})}), which appear in many proofs by rewriting.
+
+\inputsnippets{T1Bridge/h2gG2hRw, T1Bridge/g2hH2gRw}
+
+\inputsnippets{T1Bridge/h2gEqIff, T1Bridge/g2hEqIff}
 
@@
 \inputsnippets{HydraGaia_Examples/T1pp}
-
-\subsubsection{Cancel lemmas}
-
-The following cancel lemmas will be often applied in order to simplify sub-terms of the form (\texttt{g2h (h2g {\it t})}) and (\texttt{h2g (g2h {\it t})}), which appear in many proofs by rewriting.
-
-\inputsnippets{T1Bridge/h2gG2hRw, T1Bridge/g2hH2gRw}
-
-\inputsnippets{T1Bridge/h2gEqIff, T1Bridge/g2hEqIff}
 
@@
 \inputsnippets{T1Bridge/eqRef}
+Note that \HydrasLib' comparison on ordinal terms uses a
+\stdpp-like trichotomy, hence the first two lemmas below:
 
@@
 \inputsnippets{T1Bridge/compareRef}
-
+\inputsnippets{T1Bridge/decideHLtRw}
 \inputsnippets{T1Bridge/ltRef, T1Bridge/leRef}
-
-\inputsnippets{T1Bridge/decideHLtRw}
 
@@
 
-\inputsnippets{T1Bridge/SearchDemo}
+The following command lists us `gaia's lemmas (thanks to
+\gaia's  \texttt{cantor\_scope}.
+\inputsnippets{T1Bridge/SearchDemoa}
 
-\subsection{A type for well-formed ordinal terms}
-\label{sect:E0-gaia-hydras}
-For compatibility's sake, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}.
-
-\inputsnippets{T1Bridge/E0Def}
-\inputsnippets{T1Bridge/E0plusMultDef}
-
-In order to import definitions and lemmas
-from~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}, we define a pair of translations.
-
-\inputsnippets{T1Bridge/gE0h2gG2h}
-
-These tools allow us, for instance,  to import \texttt{E0.lt}'s  well-foundedness.
-
-\inputsnippets{T1Bridge/gE0LtWf}
-
-\section{E0 as an ordinal notation}
-The module \href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines an instance of class \texttt{ON E0lt compare}
-(see Chapter~\ref{chap:ON}).
-
-\inputsnippets{T1Bridge/T1compareDef}
-\inputsnippets{ T1Bridge/T1compareCorrect}
-\inputsnippets{T1Bridge/E0compare}
-\inputsnippets{T1Bridge/E0Sto}
-\inputsnippets{T1Bridge/gEpsilon0}
-
-\inputsnippets{T1Bridge/ExampleComp}
+With \texttt{t1\_scope}:
+\inputsnippets{T1Bridge/SearchDemob}
 
@@
 
-We define this function by delegation to \HydrasLib'.
+We define this function in \href{../theories/html/gaia\_hydras.GHessenberg.html}{gaia\_hydras.GHessenberg} by delegation to \HydrasLib'.
+
 \vspace{4pt} 
 \noindent.
-
-From\href{../theories/html/gaia\_hydras.GHessenberg.html}{gaia\_hydras.GHessenberg}
 
@@
 \inputsnippets{GHessenberg/oplusEquations}
+\inputsnippets{GHessenberg/oplusEquationsb}
+\inputsnippets{GHessenberg/oplusEquationsc}
+
+\inputsnippets{GHessenberg/oplusNf}
 
@@
 
+\inputsnippets{GHessenberg/oplusCoplusA}
 \inputsnippets{GHessenberg/oplusLemmasa}
@@
 
-\subsubsection{Transitions}
+\subsubsection{Transitions and paths}
+Let us consider a kind of transition system, the states of which are ordinals below $\epsilon_0$.
+The elementary transitions (``small steps'') are jumps from an ordinal $\alpha$ to $\canonseq{\alpha}{i}$, for some $i\not=0$.
 
@@
 
-\subsubsection{Paths}
+Paths (sequences of transitions) are defined by delegation to
+\href{../theories/html/hydras.Epsilon0.Paths.html}%
+{\texttt{hydras.Epsilon0.Paths}}.
 
@@
 
+\inputsnippets{GPaths/KSThm24}
 \inputsnippets{GPaths/Cor12}
@@
 \inputsnippets{GPaths/LTToStandard}
+
+\section{Rapidly growing arithmetic functions}
+In this section, we consider a few families of arithmetic functions, indexed by ordinal numbers below $\epsilon_0$.
+In \HydrasLib, these functions are defined by transfinite induction, using the \texttt{coq-equations} plug-in~\cite{sozeau:hal-01671777}.
+For that purpose, we defined a sigma-type
+\texttt{hydras.Epsilon0.E0.E0} of ordinal-terms in Cantor normal form.
+
+These families share some monotony properties, also re-defined in
+\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge}:
+
+\inputsnippets{T1Bridge/MonoDef}
+\subsection{A type for well-formed ordinal terms}
+\label{sect:E0-gaia-hydras}
+For compatibility's sake, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}.
+
+\inputsnippets{T1Bridge/E0Def}
+\inputsnippets{T1Bridge/E0plusMultDef}
+
+In order to import definitions and lemmas
+from~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}, we define a pair of translations.
+
+\inputsnippets{T1Bridge/gE0h2gG2h}
+
+These tools allow us, for instance,  to import \texttt{E0.lt}'s  well-foundedness.
+
+\inputsnippets{T1Bridge/gE0LtWf}
+
+\subsection{E0 as an ordinal notation}
+The module \href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines an instance of class \texttt{ON E0lt compare}
+(see Chapter~\ref{chap:ON}).
+
+\inputsnippets{T1Bridge/T1compareDef}
+\inputsnippets{ T1Bridge/T1compareCorrect}
+\inputsnippets{T1Bridge/E0compare}
+\inputsnippets{T1Bridge/E0Sto}
+\inputsnippets{T1Bridge/gEpsilon0}
+
+\inputsnippets{T1Bridge/ExampleComp}
 
@@
 \end{todo}
-
-The following definitions adapt \HydrasLib' to \ssreflect's 
-definitions of $<$ and $\leq$ on \texttt{nat}.
-
-\inputsnippets{T1Bridge/MonoDef}
 

### 2022-04-11 doc/part-hydras.tex
minor corrections
corrections in gaia chapter

@@
   In \texttt{gaia.ssete9}, the $n$-th finite ordinal is also written
-  \texttt{$\backslash${F}\,$n$}. There is no coercion in \gaia from \texttt{nat} to finite ordinals.
+  \texttt{$\backslash${F}\,$n$}. There is no coercion in \gaia from \texttt{nat} to \texttt{T1}.
 
@@
 
-    In Module   \href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/OrdinalNotations/OmegaOmega.v}{theories/ordinals/OrdinalNotations/OmegaOmega.v},
+    In Module   \href{https://github.com/coq-community/hydra-battles/blob/master/theories/
+      ordinals/OrdinalNotations/OmegaOmega.v}{theories/ordinals/OrdinalNotations/OmegaOmega.v},
     we represent ordinals below $\omega^\omega$ by lists of pairs of natural numbers (with the same coefficient shift as in \texttt{T1}).

### 2022-04-12 doc/gaia-chapter.tex
corrections in gaia chapter

@@
 \item Inside \texttt{GFoo}, the identifier \texttt{bar}
-  is reserved in priority to \gaia's version, and the notation
-  \texttt{hbar} is used for \texttt{hydras.Epsilon0.Foo.bar}.
+  is reserved in priority to \gaia's version.
+\HydrasLib' constructions are available through qualified names, like \texttt{Epsilon0.Foo.bar}, \texttt{Foo.bar}, or even a short notation like \texttt{hbar}  (``h'' for \HydrasLib).
 \end{enumerate}
@@
 \href{../theories/html/hydras.Epsilon0.Hessenberg.html}{hydras.Epsilon0.Hessenberg} and is used for proving termination of all hydra battles (See Sect.~\ref{sect:hessenberg-def}).
-
 We define this function in \href{../theories/html/gaia\_hydras.GHessenberg.html}{gaia\_hydras.GHessenberg} by delegation to \HydrasLib'.
@@
 \vspace{4pt} 
-\noindent.
+\noindent
 
@@
 
+\inputsnippets{GHessenberg/oplusNf}
 \inputsnippets{GHessenberg/oplusEquations}
@@
 \inputsnippets{GHessenberg/oplusEquationsc}
-
-\inputsnippets{GHessenberg/oplusNf}
 
@@
 
-First, we rename \HydrasLib' \texttt{canon} function into
-\texttt{hcanon}. Then we define a new function \emph{via} the
+Like for Hessenberg's sum, we define a function \texttt{canon}
+\emph{via} the
 translations \texttt{g2h} and \texttt{h2g}.
@@
 \inputsnippets{GCanon/canonDef}
-
 \inputsnippets{GCanon/g2hCanonRw}
@@
 
+\subsubsection{Canonical sequences and the order \texttt{T1lt}}
+
 \inputsnippets{GCanon/gcanonLt}
 \inputsnippets{GCanon/gcanonLimitMono}
-\inputsnippets{GCanon/glimitCanonSNotZero}
+%\inputsnippets{GCanon/glimitCanonSNotZero}
+\subsubsection{Canonical sequences and limits}
 
-\subsubsection{Canonical sequences and limits}
+The following lemma (borrowed from\href{../theories/html/hydras.Epsilon0.Canon.html}%
+{\texttt{hydras.Epsilon0.Canon}}) links limit ordinals to canonical sequences.
+
+\inputsnippets{GCanon/gcanonLimitStrong}
 
@@
 
-We adapt theorems from \href{../theories/html/hydras.Epsilon0.Canon.html}%
-{\texttt{hydras.Epsilon0.Canon}} to \gaia's vocabulary.
-
-\inputsnippets{GCanon/gcanonLimitStrong}
 \inputsnippets{GCanon/gcanonLimitV2}
@@
 \inputsnippets{GPaths/pathsDefs}
+
+ Paths with constant index and paths whose index is incremented by 1 at each step play an important role in the so-called Ketonen-Solovay machinery~\cite{KP81}.
+
+ \inputsnippets{GPaths/pathsDefsb}
 
@@
 In this section, we consider a few families of arithmetic functions, indexed by ordinal numbers below $\epsilon_0$.
-In \HydrasLib, these functions are defined by transfinite induction, using the \texttt{coq-equations} plug-in~\cite{sozeau:hal-01671777}.
-For that purpose, we defined a sigma-type
+The type of such families is \texttt{E0 -> nat -> nat} where
+\texttt{E0} is the type of ordinal terms in Cantor normal form, strictly below $\epsilon_0$.
+In order to compare such functions, we define a few abstract properties in
+\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} (adapted from 
+\href{../theories/html/hydras.Prelude.Iterates.html}{hydras.Prelude.Iterates}).
+
+\inputsnippets{T1Bridge/MonoDef}
+
+In \HydrasLib, these functions are defined by transfinite induction, using the \texttt{coq-equations} plug-in~\cite{sozeau:hal-01671777}. For that purpose, we defined a sigma-type
 \texttt{hydras.Epsilon0.E0.E0} of ordinal-terms in Cantor normal form.
 
-These families share some monotony properties, also re-defined in
-\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge}:
-
-\inputsnippets{T1Bridge/MonoDef}
 \subsection{A type for well-formed ordinal terms}
@@
 \inputsnippets{T1Bridge/E0Def}
+\inputsnippets{T1Bridge/E0EqMixin, T1Bridge/E0EqType}
+
+Many functions on type \texttt{T1} are ``lifted'' to \texttt{E0}.
+
+\inputsnippets{T1Bridge/E0Defb}
 \inputsnippets{T1Bridge/E0plusMultDef}
+\inputsnippets{GCanon/E0CanonDef, GCanon/E0CanonLt}
 
@@
 
-\subsection{E0 as an ordinal notation}
+\subsubsection{E0 as an ordinal notation}
 The module \href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines an instance of class \texttt{ON E0lt compare}
@@
 
-\subsection{A family of rapidly growing functions
-(Library \texttt{GF\_alpha})}
+\subsection{The \texorpdfstring{$H'_\alpha$}{H'\_alpha} family}
 
-The functions $F_\alpha$ are described in Section~\vref{sect:wainer}. The symbol \texttt{F\_} has type \texttt{E0 -> nat -> nat},
-where \texttt{E0} is the type of well formed ordinal terms less than $\epsilon_0$.
+This variant of the Hardy hierarchy is described in Sect.~\vref{sect:hardy}, and adapted to \gaia in
+Module \href{../theories/html/gaia_hydras.GHprime.html}{gaia\_hydras.GHprime}.
 
-\subsubsection{Importing the \texttt{F\_alpha} family}
+\inputsnippets{GHprime/Requirements, GHprime/Hprimedef}
 
-We are now able to import \HydrasLib' $F_\alpha$ functions.
+Here is a list of lemmas proved by rewriting from \HydrasLib version ( ~\href{../theories/html/hydras.Epsilon0.Hprime.html}{Epsilon0.Hprime} ).
 
-From~\href{../theories/html/gaia_hydras.G~F_alpha.html}%
+\subsubsection{Equations for $H'$}
+In \HydrasLib', $H'_\alpha(i)$  is defined by transfinite induction over $\alpha$, using \texttt{coq-equations}. In the present module, equations for $H'$ are just imported as lemmas.
+
+\inputsnippets{GHprime/HprimeEquations}
+
+\subsubsection{Examples}
+
+\inputsnippets{GHprime/Examples}
+\inputsnippets{GHprime/Examplesb}
+
+\subsubsection{Order and monotony properties}
+
+\inputsnippets{GHprime/abstract}
+\inputsnippets{GHprime/abstractb}
+\inputsnippets{GHprime/abstractc}
+
+\subsection{The \texorpdfstring{$F_\alpha$}{\texttt{F\_alpha}} hierarchy (Library \texttt{GF\_alpha})}
+
+The functions $F_\alpha$ are described in Section~\vref{sect:wainer}.
+
+%\subsubsection{Importing the \texttt{F\_alpha} family}
+
+From~\href{../theories/html/gaia_hydras.GF_alpha.html}%
 {\texttt{gaia\_hydras.GF\_alpha}}.
@@
 
-\begin{todo}
-  Give equations for \texttt{F\_} (corresponding to the equation
-  in \HydrasLib).
-\end{todo}
+The following equalities correspond to the equations of
+~\href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}
+
+\inputsnippets{GF_alpha/FZeroEqn}
+\inputsnippets{GF_alpha/FAlphaZeroEqn}
+\inputsnippets{GF_alpha/FSuccEqn}
+\inputsnippets{GF_alpha/FLimEqn}
 
@@
 \inputsnippets{GF_alpha/FAlphaProps}
+
+The $F_\alpha$ and $H'_\alpha$  hierarchies satisfy the inequality
+$F_\alpha(n)\leq H'_{\omega^\alpha}(n)$ for any $\alpha$ and $n>0$.
+
+\inputsnippets{GF_alpha/HprimeF}
 

### 2022-04-13 doc/gaia-chapter.tex
additions to gaia chapter; remove several proof-based definitions

@@
 \subsubsection{E0 as an ordinal notation}
-The module \href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines an instance of class \texttt{ON E0lt compare}
-(see Chapter~\ref{chap:ON}).
+
+The notion of ordinal notation is defined in Chapter~\ref{chap:ON}.
+In \href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge}, we define an instance of class \texttt{ON E0lt compare}.
+
+First, we define an instance of (\texttt{Compare T1}):
 
@@
 \inputsnippets{ T1Bridge/T1compareCorrect}
+
+Then, we build an instance of (\texttt{ON E0lt compare}).
+
 \inputsnippets{T1Bridge/E0compare}
 \inputsnippets{T1Bridge/E0Sto}
+\inputsnippets{T1Bridge/compEpsilon0}
+\inputsnippets{T1Bridge/ExampleComp}
 \inputsnippets{T1Bridge/gEpsilon0}
-
-\inputsnippets{T1Bridge/ExampleComp}
 
@@
 
+\vspace{4pt}
+
+\noindent
 From~\href{../theories/html/gaia_hydras.GF_alpha.html}%
@@
 \inputsnippets{GF_alpha/FAlphaNotPR}
+
+\subsection{The \texorpdfstring{$L_\alpha$}{\texttt{L\_alpha}} hierarchy (Library \texttt{GL\_alpha})}
+
+The functions $L_\alpha$, described in Sect.~\vref{sect:lalpha-section}  allow us to ``compute'' the length of standard hydra battles.
+In the same way as for the hierarchies $F_\alpha$ and $H'_\alpha$, we import  definitions and lemmas from the module
+~\href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0.L\_alpha}.
+
+\vspace{4pt}
+\noindent
+From\href{../theories/html/gaia_hydras.GF_alpha.html}%
+{\texttt{gaia\_hydras.GF\_alpha}}.
+
+\inputsnippets{GL_alpha/Ldef}
+\inputsnippets{GL_alpha/Leq2}
+\inputsnippets{GL_alpha/Leq3}
+\inputsnippets{GL_alpha/Leq4}
+
+The following theorem compares the $L_\alpha$ and $H'_\alpha$ hierarchies.
+
+\inputsnippets{GL_alpha/HprimeLAlpha}
 

### 2022-04-13 doc/gaia-chapter.tex
fix a typo

@@
 
- Paths with constant index and paths whose index is incremented by 1 at each step play an important role in the so-called Ketonen-Solovay machinery~\cite{KP81}.
+ Paths with constant index and paths whose index is incremented by 1 at each step play an important role in the so-called Ketonen-Solovay machinery~\cite{KS81}.
 

### 2022-04-13 doc/gaia-chapter.tex
add a new subsection

@@
 
+\subsection{Back to hydras}
+
+In \href{../theories/html/gaia_hydras.GHydra.html}%
+{\texttt{gaia\_hydras.GHydra}}, we import a few theorem abouthydra battles (see
+~\href{../theories/html/theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems}).
+
+\subsubsection{Termination of all battles}
+
+The termination measure is re-defined with \gaia's ordinals.
+
+\inputsnippets{GHydra/mDef}.
+
+\inputsnippets{GHydra/mVariant, GHydra/Termination}.
+
+\subsubsection{Length of standard battles}
+
+Finally, we relate the lenght of standard-battles with the $L_\alpha$ fast growing functions.
+
+\inputsnippets{GHydra/lDef, GHydra/lStdOk}.
+
 % \section{Already ported definitions and lemmas}

### 2022-04-13 doc/gaia-chapter.tex
fix typos

@@
 
-Here is a list of lemmas proved by rewriting from \HydrasLib version ( ~\href{../theories/html/hydras.Epsilon0.Hprime.html}{Epsilon0.Hprime} ).
+In the rest of this section, we prove a few lemmas  by rewriting from \HydrasLib version ( ~\href{../theories/html/hydras.Epsilon0.Hprime.html}{Epsilon0.Hprime} ).
 
@@
 In \href{../theories/html/gaia_hydras.GHydra.html}%
-{\texttt{gaia\_hydras.GHydra}}, we import a few theorem abouthydra battles (see
+{\texttt{gaia\_hydras.GHydra}}, we import a few theorem about hydra battles (see
 ~\href{../theories/html/theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems}).

### 2022-04-17 doc/gaia-chapter.tex
minor corrections
update gaia-chapter

@@
 \inputsnippets{T1Bridge/E0plusMultDef}
-\inputsnippets{GCanon/E0CanonDef, GCanon/E0CanonLt}
+
+Canonical sequences are lifted to type \texttt{E0} in
+\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon}.
+
+\inputsnippets{GCanon/E0CanonDef}
+\inputsnippets{GCanon/E0CanonLt}
 
@@
 
-\inputsnippets{GHydra/mDef}.
+\inputsnippets{GHydra/mDef}
 
-\inputsnippets{GHydra/mVariant, GHydra/Termination}.
+\inputsnippets{GHydra/mVariant, GHydra/Termination}
+
+\subsection{Impossibility Theorems}
+\label{sect:impossibility-gaia-version}
+The theorems described in Sect.~\vref{sect:impossibility-free}
+and~\vref{sec:standard-battles-cases} are adapted to \gaia's world.
+
+\inputsnippets{GHydra/impossibilityThms}
 
@@
 
-Finally, we relate the lenght of standard-battles with the $L_\alpha$ fast growing functions.
+Finally, we relate the length of standard-battles with the $L_\alpha$ fast growing functions.
 

### 2022-04-18 doc/gaia-chapter.tex
add an index for gaia bridge

@@
 \index{coq}{Plug-ins!Gaia}
+\index{gaiabridge}{Introduction}
 \label{gaia-chapter}
@@
 \subsection{Hessenberg sum}
-
-Natural sum  (a.k.a. Hessenberg's  sum) is defined in
+\index{gaiabridge}{Hessenberg sum}
+Natural sum  (a.k.a. Hessenberg  sum) is defined in
 \href{../theories/html/hydras.Epsilon0.Hessenberg.html}{hydras.Epsilon0.Hessenberg} and is used for proving termination of all hydra battles (See Sect.~\ref{sect:hessenberg-def}).
@@
 \label{sect:gcanon}
+\index{gaiabridge}{Canonical sequences}
+
 Canonical sequences are described in Chapter~\vref{chap:ketonen}, and implemented in
@@
 
-Like for Hessenberg's sum, we define a function \texttt{canon}
+Like for Hessenberg sum, we define a function \texttt{canon}
 \emph{via} the
@@
 \label{sect:gpath}
+\index{gaiabridge}{Accessibility and paths inside $\epsilon_0$}
 
@@
 \section{Rapidly growing arithmetic functions}
+\index{gaiabridge}{Rapidly growing functions}
+
 In this section, we consider a few families of arithmetic functions, indexed by ordinal numbers below $\epsilon_0$.

### 2022-04-18 doc/hydras.tex
add an index for gaia bridge

@@
 %\fancyhead{Hydras, Ordinals, \& Co.}
+\makeindex{gaiabridge}
 \makeindex{coq}
@@
 
+\index{gaiabridge}{Introduction}
 Our library \gaiaHydras is dedicated to make compatible our lemmas with José Grimm's \gaia project (designed for \ssreflect/\mathcomp) (please look at Sect.~\vref{sect:gaia-first-intro} and the paragraphs signalled with {\gaiasign}).
@@
 
+\printindex{gaiabridge}{Links to \gaia Library}
+\label{gaia-index}
 \printindex{coq}{Coq, plug-ins and standard library}

### 2022-04-18 doc/part-hydras.tex
additions to gaia chapter; remove several proof-based definitions
update gaia-chapter
add an index for gaia bridge

@@
 aims  to formalize mathematics in \coq  in
+\index{gaiabridge}{Introduction}
+
 the style of Nicolas Bourbaki.
@@
 Remarks about compatibility with \gaia are signalled with the picture \gaiasign.
+A special index is in construction (page~\pageref{gaia-index}).
 
@@
 We use it as a reference for proving the correctness of ordinal notations.
+
+  \index{gaiabridge}{Introduction}
+  \gaiasign
   We are also progressively importing \gaia's~\cite{Gaia} definitions and theorems, based on Bourbaki's set theory~\cite{Grimm2016}.
@@
 
-\section{Ordinal Notations and \gaia's ordinals}
-\gaiasign Module~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge} contains an instance of class \texttt{ON E0lt compare}, where \texttt{E0lt} is the order on the well-formed ordinal terms below $\epsilon_0$ defined in \gaiaHydras library (please see Chapter~\ref{gaia-chapter}).
-
 %------------------------------------------------------------------------
@@
 \subsection{About Gaia}
+\index{gaiabridge}{Introduction}
 \gaiasign Chapter~\vref{gaia-chapter} describes the present state of a
@@
 \input{movies/snippets/T1/finiteOrds}
-
+\index{gaiabridge}{Finite ordinals}
 \paragraph*{\gaiasign}
@@
 
+\index{gaiabridge}{The ordinal $\omega$}
 \paragraph*{\gaiasign}
@@
 \paragraph*{\gaiasign}
+\index{gaiabridge}{Exponential of base $\omega$}
 In \texttt{gaia.ssete9} the identifier \texttt{phi0} is bound to a plain constant (not a notation).
@@
 \paragraph*{\gaiasign}
+\index{gaiabridge}{Pretty printing Cantor normal forms}
 In ~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge},
@@
 \paragraph*{\gaiasign}
+\index{gaiabridge}{Strict order on ordinals below $\epsilon_0$}
+
 In \gaia, the strict order \texttt{T1lt} on ordinal terms is directly defined as a boolean function (in \ssreflect's style).
@@
 \paragraph*{\gaiasign}
+\index{gaiabridge}{Ordinal terms in Cantor normal form}
 In \gaia, the boolean function which characterizes ordinal terms in normal form is defined as follows:
@@
 \paragraph*{\gaiasign}
+\index{gaiabridge}{Type of well formed ordinal terms below $\epsilon_0$}
  The library~\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines a \gaia-compatible type \texttt{E0}.
@@
 \paragraph*{\gaiasign}
+\index{gaiabridge}{Limit and successor ordinals}
 In \gaia, the boolean functions associated with limit and successor ordinals are called \texttt{T1limit} and \texttt{T1is\_succ}.
@@
 \input{movies/snippets/Correctness_E0/Epsilon0Correct}
+
+\subsection{An ordinal notation for  \gaia's ordinals}
+\index{gaiabridge}{Ordinal notations}
+
+\gaiasign Module~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge} contains an instance of class \texttt{ON E0lt compare}, where \texttt{E0lt} is the order on the well-formed ordinal terms below $\epsilon_0$ defined in \gaiaHydras library (please see Chapter~\ref{gaia-chapter}).
+
+\inputsnippets{T1Bridge/gEpsilon0}
 
@@
 
-    Natural sum (Hessenberg's  sum) is a commutative and monotonous version of
+    Natural sum (Hessenberg  sum) is a commutative and monotonous version of
     addition. It is used as an auxiliary operation  for defining variants
@@
     \end{project}
-    \subsection{More theorems on Hessenberg's sum}
-
+    \subsection{More theorems on Hessenberg sum}
+\index{gaiabridge}{Hessenberg sum}
     We need to prove some properties of $\oplus$, particularly about 
@@
 
-    Associativity of Hessenberg's sum is proved the same way.
+    Associativity of Hessenberg sum is proved the same way.
 
@@
     \paragraph*{\gaiasign} The module
-      \href{../theories/html/gaia_hydras.~GHessenberg.html}{gaia\_hydras.GHessenberg} defines a version of Hessenberg's sum
+      \href{../theories/html/gaia_hydras.~GHessenberg.html}{gaia\_hydras.GHessenberg} defines a version of Hessenberg sum
       compatible with \gaia's type \texttt{T1}.
-
+      \index{gaiabridge}{Hessenberg sum}
     \subsection{A termination measure for hydra battles }
@@
     \paragraph*{\gaiasign}
+    \index{gaiabridge}{Termination of all hydra battles}
     The module~\href{../theories/html/gaia_hydras.GHydra.html}{gaia\_hydras.GHydra} contains a proof of \gaiaHydras version of termination of all battles.
@@
 (please see Sect.\~ref{sect:gcanon}).
-
+\index{gaiabridge}{Canonical sequences}
 \paragraph*{Remark}
@@
 Library ~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon}:
-
+\index{gaiabridge}{Canonical sequences}
 \inputsnippets{GCanon/gcanonLt}
@@
 \input{movies/snippets/Canon/canonSLimitLub}
+\index{gaiabridge}{Canonical sequences}
 
@@
 \section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0}\,: paths}
+\index{gaiabridge}{Accessibility and paths inside $\epsilon_0$}
 \index{maths}{Ordinal numbers!Accessibility inside epsilon0}
@@
 \paragraph*{\gaiasign}
+
 The library~\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} transposes the  notion of path into \gaia's type \texttt{T1}
@@
 \section{A  proof of impossibility}
+\label{sect:impossibility-free}
 
@@
 
+\paragraph*{\gaiasign} Please look at the \gaia version of this theorem in Sect.~\vref{sect:impossibility-gaia-version}.
+\index{gaiabridge}{Impossibility theorems}
 \section{The case of standard battles}
@@
 
+\paragraph*{\gaiasign} Please look at the \gaia version of this theorem in Sect.~\vref{sect:impossibility-gaia-version}.
+\index{gaiabridge}{Impossibility theorems}
+
 \subsection{Remarks}
@@
 \section{Length of minimal large sequences}
+\label{sect:lalpha-section}
 
@@
 \paragraph*{\gaiasign} Library~\href{../theories/html/gaia_hydras.GF_alpha.html}{gaia\_hydras.GF\_alpha}  defines the functions $F_\alpha$ for \gaia's world.
+\index{gaiabridge}{Rapidly growing functions}
 

### 2022-04-23 doc/JFLA2022/Slides/Slides.tex
update slides about Gaia
changes to JFLA slides

@@
 \usetheme{Montpellier}
+\usepackage{xspace, alltt}
+\usepackage[utf8x]{inputenc}
 
-\usepackage{xspace, alltt}
+\usepackage{fontspec}
+%\setmainfont{DejaVu Serif}
+
+\setmonofont{DejaVu Sans Mono}[Scale=MatchLowercase]
 
@@
 \newcommand{\canonseq}[2]{\mbox{$\{#1\}(#2)$}}
+\newcommand{\rounds}{\mbox{\,\texttt{-+->}\,}}
+\newcommand{\round}{\mbox{\,\texttt{-1->}\,}}
+\newcommand{\rplus}[1]{\mbox{$\,\underset{#1}{\xrightarrow{\textcolor{white}{\,#1\,}}}\,$}}
 
@@
 }
-\date{JFLA, February 2022}
+\date{JFLA, June 2022}
 \author{
@@
         \end{itemize}
+           \item book in pdf (300p), generated  with \textcolor{lookcolor}{\alectr} and \LaTeX, continuously updated and deployed
     \end{itemize}
@@
       (part of Russel O'Connor's \textcolor{plugincolor}{Goedel})
-    \item Preliminary bridge to \textcolor{plugincolor}{Gaia} (José Grimm)
-      \item Documentation in pdf (280p), generated  with \textcolor{lookcolor}{\alectr} and \LaTeX, continuously updated and deployed
+    \item Bridge to \textcolor{plugincolor}{Gaia} (José Grimm) (\emph{in construction})
+
     \end{itemize}
@@
     some \textcolor{mathcolor}{discrete math}, discuss
-    \textcolor{coqstylecolor}{\coq style, specification and proof patterns}, and use\footnote{and depend on!} \textcolor{plugincolor}{a few plug-ins}.
+    \textcolor{coqstylecolor}{\coq style, specification and proof patterns}, and depend on \textcolor{plugincolor}{a few plug-ins}.
   \end{block}
@@
     \caption{Three successive states of a hydra in a battle,
-      at time $t=0,1,2$\label{life1}}
+      at time $t=1,2,3$\label{life1}}
     \label{fig:round}
@@
     \end{tikzpicture}
-    \caption{At $t=3,\dots,\textcolor{red}{x}-1,\textcolor{red}{x=?}$
+    \caption{At $t=4,\dots,\textcolor{red}{x}-1,\textcolor{red}{x=?}$
       \label{life2}}
@@
 \begin{frame}
-    \begin{block}{Mathematical formalization}
+  \frametitle{The Ketonen-Solovay machinery\footnote{Rapidly Growing {R}amsey Functions (1981)}}
+    \begin{block}{Formalization}
 
+      $$\omega^{\omega^2+1}+1 \rplus{1} \omega^{\omega^2+1}
+      \rplus{2} \omega^{\omega^2}\times 3 
+      \rplus{3} \omega^{\omega^2}\times 2 +
+      \omega ^ {\omega \times 4}, \dots,2,1,0$$
+
+      \vspace{6pt}
+
+      $$ \omega^{\omega^2+1}+1 \rplus{1,2,3,4,\dots,x} 0$$
+    \end{block}
+
+      \begin{block}{Small steps}
+        At time $i$, 
+           $\alpha \rplus{i} \canonseq{\alpha}{i}$,
+        where $\canonseq{\alpha}{\_}$ is the \emph{canonical sequence} of $\alpha$.
         \begin{itemize}
-    \item   We associate a hydra to each ordinal \textcolor{mathcolor}{$\alpha<\epsilon_0$} (in Cantor normal form).
+        \item $\canonseq{0}{i}=0$
+        \item $\canonseq{\beta+1}{i}=\beta$
+         \item If $\lambda$ is a limit ordinal, then $\lambda$ is the $\omega$-limit of $\canonseq{\lambda}{\_}$.
+        \end{itemize}
+    \end{block}
+
+      % We associate a hydra to each ordinal \textcolor{mathcolor}{$\alpha<\epsilon_0$} (in Cantor normal form).
 For instance, the first hydra of our example is the image of \textcolor{mathcolor}{$\omega^{\omega^2+1}+1$}.
-    \item A round (at time $t\in\mathbb{N}$) of a battle transforms an hydra associated with \textcolor{mathcolor}{$\alpha$} into the hydra associated with \textcolor{mathcolor}{$\canonseq{\alpha}{t+1}$}, the $t+1$-st element of the \emph{canonical sequence} of \textcolor{mathcolor}{$\alpha$}.
-    \item We want to ``compute'' the length of sequences of the form 
-      \textcolor{mathcolor}{$\langle \alpha_k=\alpha,\dots,\alpha_{i+1}=\canonseq{\alpha_i}{i+1},\dots, 1,0\rangle$}, where \textcolor{mathcolor}{$k\in\mathbb{N}$}.
-      \item Such sequences are studied
-      as \emph{$\alpha$-large sets of natural numbers}  in Ketonen \& Solovay's article
-      \emph{Rapidly Growing {R}amsey Functions (1981)}.
+
+    % \item We want to ``compute'' the length of sequences of the form 
+    %   \textcolor{mathcolor}{$\langle \alpha_k=\alpha,\dots,\alpha_{i+1}=\canonseq{\alpha_i}{i+1},\dots, 1,0\rangle$}, where \textcolor{mathcolor}{$k\in\mathbb{N}$}.
+
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+
+     \begin{block}{}
+       \begin{itemize}
+       \item In Coq, canonical sequences are defined by structural recursion on ordinal terms in Cantor normal form.
+         \input{../../movies/snippets/Canon/canonExamplesb}
+       \item Not the same definition as in K. \& S.'s article, but same abstract properties (monotony, limits, \dots).
        \end{itemize}
@@
 
+  \end{block}
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+  \frametitle{Hydras \& Gaia}
+  \begin{block}{From \gaia to \Hydras}
+    {\footnotesize
+    \inputsnippets{./multA}}
+  \end{block}
+\end{frame}
+
+\begin{frame}
+  \begin{block}{From \Hydras to \gaia}
+    {\footnotesize
+      \inputsnippets{./canonDef}
+      \inputsnippets{./gcanonLimitOf}}
   \end{block}
@@
 Addition-chains & \texttt{0.5} & 1,961 & 2,210 \\
-Gaia-hydras &  \texttt{0.5} & 81 & 177 \\
+Gaia-hydras &  \texttt{0.8} & 677 & 2,023 \\
 \hline

### 2022-04-25 doc/JFLA2022/Slides/Slides.tex
work on JFLA slides

@@
 \usetheme{Montpellier}
-\usepackage{xspace, alltt}
-\usepackage[utf8x]{inputenc}
 
-\usepackage{fontspec}
-%\setmainfont{DejaVu Serif}
-
-\setmonofont{DejaVu Sans Mono}[Scale=MatchLowercase]
-
-\newcommand{\TODO}[2][]{[\textcolor{red}{TODO (#1):} \emph{#2}]}
-\newcommand{\coq}{Coq\xspace}
-\newcommand{\coqdoc}{Coqdoc\xspace}
-\newcommand{\coqmakefile}{\texttt{coq\_makefile}\xspace}
-\newcommand{\community}{Coq-community\xspace}
-\newcommand{\gaia}{Gaia\xspace}
-\newcommand{\alectr}{Alectryon\xspace}
-\newcommand{\equations}{Equations\xspace}
-\newcommand{\Hydras}{Hydras \& Co$\text.$\xspace}
-\newcommand{\make}{\texttt{make}\xspace}
-
-\definecolor{orange}{rgb}{1.0,0.6,0.0}
-\definecolor{darkorange}{rgb}{0.7,0.25,0.0}
-\definecolor{plugincolor}{rgb}{0.6,0.0,0.6}
-\definecolor{mathcolor}{rgb}{0.0,0.0,0.6}
-\definecolor{coqstylecolor}{rgb}{0.0,0.0,01.0}
-\definecolor{lightblue}{rgb}{0.2,0.2,1.0}
-\definecolor{metavarcolor}{rgb}{0.5,0.0,1.0}
-\definecolor{darkgreen}{rgb}{0.1,0.7,0.1}
-\definecolor{answercolor}{rgb}{.08,.15,.8}
-\definecolor{normalcolor}{rgb}{0.0,0.0,0.0}
-\definecolor{exbluecolor}{rgb}{0.1,0.1,0.9}
-\definecolor{dontlookcolor}{rgb}{0.5,0.5,0.5}
-\definecolor{termcolor}{rgb}{0.0,0.1,0.9}
-\definecolor{lookcolor}{rgb}{0.5,0.1,0.0}
-\definecolor{prooftermcolor}{rgb}{0.3,0.1,1.0}
-\definecolor{typecolor}{rgb}{1.0,0.6,0.0}
-\definecolor{taccolor}{rgb}{0.70,0.10,0.0}
-\definecolor{pink}{rgb}{0.8,0.6,0.6}
-\definecolor{darkmagenta}{rgb}{0.4,0.0,0.6}
-\definecolor{darkblue}{rgb}{0.0,0.0,0.6}
-\definecolor{cyan}{rgb}{0.0,0.4,0.8}
-\usepackage{tikz}
-\usepackage{tikzsymbols}
-\usepackage{pifont}
-\usetikzlibrary{arrows}
-
-\usepackage{ifpdf}
-\ifpdf
-\usepackage{graphicx}
-\else
-\usepackage[dvips]{graphicx}
-\fi
-
-%%%% For Alectryon
-
-\usepackage{texments}
-\usepackage{./alectryon}
-\usepackage{../pygments}
-
-% Prevent breaks in the middle of syntactic units
-\let\OldPY\PY
-\def\PY#1#2{\OldPY{#1}{\mbox{#2}}}
-
-%%% One hypothesis per line 
-\makeatletter
-\renewcommand{\alectryon@hyps@sep}{\alectryon@nl}
-\makeatother
-
-%%% \snippets{A,B,C,…} inputs a series of snippets as one block (with \itemsep
-%%% between them).  A, B, C should be paths to files in snippets/.
-
-\usepackage{etoolbox}
-\makeatletter
-
-\newcommand{\pathtomovies}{.}
-
-\newcommand{\inputsnippets}[1]
-{{\setlength{\itemsep}{1pt}\setlength{\parsep}{0pt}% Adjust spacing
-    \alectryon@copymacros\begin{io}
-      \forcsvlist{\item\@inputsnippet}{#1}
-    \end{io}}}
-\let\input@old\input % Save definition of \input
-\newcommand{\@inputsnippet}[1]
-{{\renewenvironment{alectryon}{}{}% Skip \begin{alectryon} included in snippet
-    \input@old{{\pathtomovies}/#1}}}
-\makeatother
-
-% End of Alectryon stuff
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%% Specific macros
-
-\newcommand{\canonseq}[2]{\mbox{$\{#1\}(#2)$}}
-\newcommand{\rounds}{\mbox{\,\texttt{-+->}\,}}
-\newcommand{\round}{\mbox{\,\texttt{-1->}\,}}
-\newcommand{\rplus}[1]{\mbox{$\,\underset{#1}{\xrightarrow{\textcolor{white}{\,#1\,}}}\,$}}
+\input{macros}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+\begin{frame}
+  \tableofcontents
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+    \frametitle{\community}
+    \begin{block}{}
+      \community is an informal organization that aims to maintain interesting open source \coq projects and facilitate collaboration among Coq users on documentation, tooling, etc.
+      \begin{itemize}
+      \item Created in 2018,  run by volunteer Coq users on GitHub
+      \item At present, over than 50 projects, maintained by over than 30 volunteers
+      \item Some packages were unmaintained before their integration to \community. According to their scientific or historical interest, we decided to revive them.
+        \item After their transfer, packages may be the object of large changes  and refactorings.
+      \end{itemize}
+    \end{block}
+  \end{frame}
 
-\section{Introduction}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \item Allow to compare several different formalizations of a given notion, and several proofs of the same theorem.
+      \item Show also interesting dead ends (formalization, tactics, proof structure).
     \end{itemize}
   \end{block}
-  \begin{block}{Examples}
-    \begin{itemize}
-    \item Tsukuba Coq Users's Group: regular languages, $\lambda$-calculus, arithmetic, ... (in Japanese + ssreflect)
-    \item Software Foundations (B. Pierce \emph{et al.})
-    \end{itemize}
-  \end{block}
+  % \begin{block}{Examples}
+  %   \begin{itemize}
+  %   \item Tsukuba Coq Users's Group: regular languages, $\lambda$-calculus, arithmetic, ... (in Japanese + ssreflect)
+  %   \item Software Foundations (B. Pierce \emph{et al.})
+  %   \end{itemize}
+  % \end{block}
 \end{frame}
@@
 
-    % \item We want to ``compute'' the length of sequences of the form 
-    %   \textcolor{mathcolor}{$\langle \alpha_k=\alpha,\dots,\alpha_{i+1}=\canonseq{\alpha_i}{i+1},\dots, 1,0\rangle$}, where \textcolor{mathcolor}{$k\in\mathbb{N}$}.
+  \end{frame}
 
-\end{frame}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  \begin{frame}
+    \frametitle{Paths and accessibility}
+    \begin{block}{Theorem (K. \& S., 1981) }
+      % If $\beta<\alpha<\epsilon_0$, there exists $i$ and $l$ in $\mathbb{N}$ such that $ \alpha \rplus{i, i+1, i+2, \dots,l} \beta$.
+
+      \vspace{6pt}
+
+    \begin{tikzpicture}[very thick, scale=0.25]
+\node (alpha) at (0,0) {$\alpha$};
+    \node (beta) at (40, 0){$\beta$};
+
+  \draw[->, very thick,black] (alpha)--  node [above]{$>$} (beta);
+
+  \node (alpha1) at (5,5) {};
+  \node (alpha2) at (13,5) {};
+    \node (alpha3) at (20,5) {};
+  \node (alpha4) at (25,5) {};
+  \draw [->, dashed,very thick,red] (alpha)-- node [below, rotate=40]{$i$}  (alpha1);
+  \draw [->, dashed,very thick,red] (alpha1)-- node [below]{$i+1$}  (alpha2);
+   \draw [->, dashed,very thick,red] (alpha2)-- node [below]{$i+2$}  (alpha3);
+    \draw [->, dashed,very thick,red] (alpha3)-- node [below]{$i+3$}  (alpha4);
+  \node (dots) at (30,5) {$\dots$};
+  \node (search) at (32,5){};
+ \draw [->, dashed, very thick,red] (search)-- node [below, rotate=-26]{$l$}  (beta);
+\end{tikzpicture}
+
+\pause
+\textbf{Remark:} The proof is not trivial. The built path could miss $\beta$!
+
+\vspace{4pt}
+
+\textbf{Cor:} There is no decreasing measure mapping hydras to  $[0, \mu)$ (for any $\mu<\epsilon_0$)  for proving the termination of the considered kind of hydra battles.
+
+\end{block}
+
+  \end{frame}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \item Let \textcolor{mathcolor}{$0<\alpha<\epsilon_0$}, and \textcolor{mathcolor}{$i\in\mathbb{N}$}.
-      The least index \textcolor{mathcolor}{$k$} such that \textcolor{mathcolor}{$\alpha_k=0$} is  \textcolor{mathcolor}{$H'_\alpha(i)$} where \textcolor{mathcolor}{$H'_\alpha$} is defined by transfinite induction over \textcolor{mathcolor}{$\alpha$}.
+      The least index \textcolor{mathcolor}{$l$} such that \textcolor{mathcolor}{$\alpha_l=0$} is  \textcolor{mathcolor}{$H'_\alpha(i)$} where \textcolor{mathcolor}{$H'_\alpha$} is defined by transfinite induction over \textcolor{mathcolor}{$\alpha$}.
 
@@
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \begin{frame}
-    \frametitle{\community}
-    \begin{block}{}
-      \community is an informal organization that aims to maintain interesting open source \coq projects and facilitate collaboration among Coq users on documentation, tooling, etc.
-      \begin{itemize}
-      \item Created in 2018,  run by volunteer Coq users on GitHub
-      \item At present, over than 50 projects, maintained by over than 30 volunteers
-      \item Some packages were unmaintained before their integration to \community. According to their scientific or historical interest, we decided to revive them.
-        \item After their transfer, packages may be the object of large changes  and refactorings.
-      \end{itemize}
-    \end{block}
-  \end{frame}
 

### 2022-04-26 doc/JFLA2022/Slides/Slides.tex
work on JFLA slides

@@
        \item In Coq, canonical sequences are defined by structural recursion on ordinal terms in Cantor normal form.
+
+   {\color{mathcolor}
+         $$\canonseq{\omega^{\omega^{\omega+1}+1}}{42}=
+         \omega^{\omega^{\omega+1}}\times 42$$
+         }
+
          \input{../../movies/snippets/Canon/canonExamplesb}
@@
 
+          \label{fig:accessibility}
         \begin{tikzpicture}[very thick, scale=0.25]
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+\begin{frame}
+  \frametitle{Fast growing hierarchies of arithmetic functions}
 
-\begin{frame}
 \begin{block}{}
-    \begin{itemize}
-    \item Let \textcolor{mathcolor}{$0<\alpha<\epsilon_0$}, and \textcolor{mathcolor}{$i\in\mathbb{N}$}.
-      The least index \textcolor{mathcolor}{$l$} such that \textcolor{mathcolor}{$\alpha_l=0$} is  \textcolor{mathcolor}{$H'_\alpha(i)$} where \textcolor{mathcolor}{$H'_\alpha$} is defined by transfinite induction over \textcolor{mathcolor}{$\alpha$}.
 
+  {\small
+\textbf{Hardy Hierarchy (variant)}
+
+ Let us define \textcolor{mathcolor}{$H'_\alpha(i)$} as the least $l$ such that \textcolor{mathcolor}{$\alpha_l=0$} (Figure page.~\pageref{fig:accessibility}). 
+ {\color{mathcolor}
 \begin{align}
   H'_0(i) & = i\\
-        H'_\alpha(i) &= H'_{(\canonseq{\alpha}{i+1})}(i)  \quad\textit{if $\alpha$ is a limit ordinal}\\
-        H'_{\alpha}(i) &=H'_\beta(i+1) \quad\textit{if $\alpha=\beta+1$}
-      \end{align}
+  H'_{\beta+1}(i) &=H'_\beta(i+1) \\
+  H'_\lambda(i) &= H'_{(\canonseq{\lambda}{i+1})}(i)  \quad\textit{($\lambda$ limit ordinal)} 
+\end{align}}
 
-    \end{itemize}
-  \end{block}
+\vspace{4pt}
 
-  \begin{block}{}
-    The \textcolor{mathcolor}{$H'_\alpha$} family,  is a variant of the  \emph{Hardy hierarchy of rapidly growing functions}.
-For instance, \textcolor{mathcolor}{$H'_{\omega^3+3}(0)\geq2^{2^N}$}, where \textcolor{mathcolor}{$N=2^{70}-1$}.
+\textbf{Wainer hierarchy (variant)}
 
-    {\color{lookcolor} Such functions defy common programming intuition: no test is possible, only proofs!}
-  \end{block}
+We use also the following family in our proofs.
 
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{frame}
-  \frametitle{Ordinals in \coq}
-
-  \begin{block}{}
-    Ordinals less than {\color{mathcolor}$\omega^\omega, \epsilon_0,  \Gamma_0$} are represented as \emph{ordinal notations}: comparison, canonical sequences, partition limit/successors/zero ordinals are implemented by certified functions, allowing the user to write proofs by computation.
-
-    \begin{footnotesize}
-      \inputsnippets{./ONDef}
-
-      \inputsnippets{../E0_Ex42}
-    \end{footnotesize}
-
+{\color{mathcolor}
+\begin{align}
+F_0(i)& =i+1\\
+F_{\beta+1}(i)&= (F_\beta)^{(i+1)}(i)\\
+F_\lambda(i) &= F_{(\canonseq{\lambda}{i})} (i) \quad \textit{($\lambda$ limit ordinal)}
+\end{align}}}
 \end{block}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}[fragile]
+
       \begin{block}{}
-    The functions \textcolor{mathcolor}{$H'_\alpha\;(\alpha< \epsilon_0)$} are implemented with \textcolor{plugincolor}{coq-equations}.
+    The functions \textcolor{mathcolor}{$H'_\alpha, F_\alpha\;(\alpha< \epsilon_0)$} are implemented with \textcolor{plugincolor}{coq-equations}.
     \vspace{5pt}
@@
 
+\end{frame}
+
+\begin{frame}[fragile]
+  \begin{block}{Examples}
+
+    Using the lemmas generated by \emph{coq-equations}, we prove minoration lemmas or exact formulas for given $\alpha$\,s:
+
+    \begin{itemize}
+    \item \input{../../movies/snippets/Hprime/HprimeOmegaCube3c}
+      \input{../../movies/snippets/Hprime/HprimeOmegaCube3e}
+     \item   \input{../../movies/snippets/Hprime/HprimeHexp2}
+    \item
+      \input{../../movies/snippets/Hprime/HprimePhi0OmegaClosed}
+
+      \end{itemize}
+
+    \end{block}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+
   \begin{block}{}
@@
       \item \textcolor{mathcolor}{$H'_\alpha$} is strictly monotonous
-      \item \textcolor{mathcolor}{$\forall i, i \leq H'_\alpha(i)$}
-      \item   If \textcolor{mathcolor}{$\beta<\alpha$}, then \textcolor{mathcolor}{$H'_\beta(i)<H'_\alpha(i)$} for all but a finite number of $i$ (we say that \textcolor{mathcolor}{$H'_\alpha$}
-        \emph{dominates} \textcolor{mathcolor}{$H_\beta$}).
+      \item If $\alpha>0$ then \textcolor{mathcolor}{$ i < H'_\alpha(i)$} for all $i$
+        \item $H'_\alpha(i) < H'_{\alpha+1}(i) \; (i \not=0)$
+      \item   If \textcolor{mathcolor}{$\alpha \rplus{n,n,\dots,n} \beta$} then \textcolor{mathcolor}{$H'_\beta(n)\leq H'_\alpha(n)$} 
       \end{itemize}
@@
   \end{block}
+  \begin{block}{Theorem}
+    The function which computes the length of hydra battles
+    (according to the initial index) is not primitive recursive
+    (for $\alpha\geq \omega^\omega$).
+    \vspace{4pt}
+
+\emph{We use   Russel O'Connor's formalization of PR functions.}
+  \end{block}
+
 \end{frame}
@@
 
-\begin{frame}
-  \begin{block}{Theorem}
-    For any   \textcolor{mathcolor}{$\alpha\geq\omega^\omega$} the function \textcolor{mathcolor}{$H'_\alpha$} is not primitive recursive.
-  \end{block}
+% \begin{frame}
 
-  \begin{block}{}
-    {\color{coqstylecolor}
-      \begin{itemize}
-      \item \emph{In order to prove this theorem, we chose to host and  maintain under \community Russel O'Connor's library \textcolor{plugincolor}{goedel}, which contains a formalization of primitive recursive functions.}
+%   \begin{block}{}
+%     {\color{coqstylecolor}
+%       \begin{itemize}
+%       \item \emph{In order to prove this theorem, we chose to host and  maintain under \community Russel O'Connor's library \textcolor{plugincolor}{goedel}, which contains a formalization of primitive recursive functions.}
 
-      \item    \emph{The interest of O'Connor's contribution led us to write a chapter on primitive recursive functions, with [counter-]examples and exercices.}
+%       \item    \emph{The interest of O'Connor's contribution led us to write a chapter on primitive recursive functions, with [counter-]examples and exercices.}
 
-      \item \emph{Proving formally that a given function is not primitive recursive is a nice example of application of a complex (mutually recursive, dependently typed) induction principle.
+%       \item \emph{Proving formally that a given function is not primitive recursive is a nice example of application of a complex (mutually recursive, dependently typed) induction principle.
 You have to take into account the set of PR functions of any arity.}
 
-      \end{itemize}
-    }
-  \end{block}
-\end{frame}
+%       \end{itemize}
+%     }
+%   \end{block}
+% \end{frame}
 
@@
     \end{quote}
-  \end{block}
 
-  \begin{block}{}
-
-    Instantiating with \textcolor{mathcolor}{$n=1$}, we obtain:
-
-    \begin{quote}
-      If \textcolor{mathcolor}{$H'_{\omega^\omega}$} were primitive recursive, there would exists some \textcolor{mathcolor}{$q$} such that, for any \textcolor{mathcolor}{$k$}, \textcolor{mathcolor}{$H'_{\omega^\omega}(k) \leq\textrm{Ack}(q,k)$}, which is false.
-    \end{quote}
+       \input{../../movies/snippets/F_omega/lemmas}
 
@@
 \end{alltt}
+
 \end{frame}
@@
 \caption{\textcolor{lookcolor}{Genealogy} and dependencies for \Hydras packages.}
- %Dotted boxes represent historical Coq contributions, while regular boxes represent maintained \coq packages.
+% Dotted boxes represent historical Coq contributions, while regular boxes represent maintained \coq packages.
 Orange packages are maintained in the \Hydras GitHub repository, while blue packages are maintained in other \community repositories.
 Dotted lines represent \coq code ancestry, while regular lines represent direct code dependencies.}
+
   \label{fig:genealogy}
@@
 \end{document}
+
+\begin{frame}
+  \frametitle{Ordinals in \coq}
+
+  \begin{block}{}
+    Ordinals less than {\color{mathcolor}$\omega^\omega, \epsilon_0,  \Gamma_0$} are represented as \emph{ordinal notations}: comparison, canonical sequences, partition limit/successors/zero ordinals are implemented by certified functions, allowing the user to write proofs by computation.
+
+    \begin{footnotesize}
+      \inputsnippets{./ONDef}
+
+      \inputsnippets{../E0_Ex42}
+    \end{footnotesize}
+
+  \end{block}
+\end{frame}

### 2022-04-26 doc/part-hydras.tex
changes to JFLA slides
work on JFLA slides

@@
 \mathcolor{$\canonseq{\omega^\omega}{3} = \omega^3$} and
-\mathcolor{$\canonseq{\omega^\omega*3}{42} = \omega^\omega*2 + \omega^{42}$}.
+\mathcolor{$\canonseq{\omega^{\omega^{\omega+1}+1}}{42}=
+ \omega^{\omega^{\omega+1}}\times 42$}
 
 \input{movies/snippets/Canon/canonExamples}
+\input{movies/snippets/Canon/canonExamplesb}
 
@@
 
-Note that this formula contains two occurrences of the functional \texttt{iterate}, the outer one is in fact a second-order iteration (on type \texttt{nat -> nat)}
-and the inner one  first-order (on type \texttt{nat}).
+Note that this formula contains two occurrences of the functional \texttt{iterate}, the second  one is in fact a second-order iteration (on type \texttt{nat -> nat)}
+and the first one  first-order (on type \texttt{nat}).
 
@@
 Thus,  our function \texttt{ord}  which enumerates the elements of $B$ is defined in a non-ambiguous way.
-Let us quote the following theorems (see Library
+Let us quote the following theorems (see Library \linebreak
 \href{../theories/html/hydras.Schutte.Ordering_Functions.html}%

### 2022-04-27 doc/JFLA2022/Slides/Slides.tex
work on JFLA slides

@@
-\documentclass[10pt]{beamer}
+\documentclass[10pt, fleqn]{beamer}
 \usetheme{Montpellier}
@@
 
-    Using the lemmas generated by \emph{coq-equations}, we prove minoration lemmas or exact formulas for given $\alpha$\,s:
+      {\color{mathcolor}
 
+    \begin{align*}
+      H'_{\omega^3}(k) \geq\, & \textit{hyper\_exp2}(k+1) \\
+      H'_{\omega^\omega}(k)=\,& \textbf{let}\,F\;f\;i := f^{(i+1)}(i)\,
+                      \textbf{in}\, F^{(k)}(\lambda k.\,2k+1)\,k\\
+    \end{align*}
+       }%mathcolor
+  \end{block}
+
+\begin{block}{Lemmas (after K.\& S.)}
    \begin{itemize} 
-    \item \input{../../movies/snippets/Hprime/HprimeOmegaCube3c}
-      \input{../../movies/snippets/Hprime/HprimeOmegaCube3e}
-     \item   \input{../../movies/snippets/Hprime/HprimeHexp2}
-    \item
-      \input{../../movies/snippets/Hprime/HprimePhi0OmegaClosed}
+    \item \textcolor{mathcolor}{$H'_\alpha$} is strictly monotonous
+
+     \item  If $\alpha>0$ then \textcolor{mathcolor}{$ i < H'_\alpha(i)$} for all $i$
+
+    \item   $H'_\alpha(i) < H'_{\alpha+1}(i) \; (i \not=0)$
+
+       \item     If \textcolor{mathcolor}{$\alpha \rplus{n,n,\dots,n} \beta$} then \textcolor{mathcolor}{$H'_\beta(n)\leq H'_\alpha(n)$}
 
        \end{itemize}
-
+       (Proof by mutual transfinite induction)
  \end{block}
@@
 
-  \begin{block}{}
-    Using combinatorial lemmas by  K. \& S., we prove (by mutual  transfinite induction on \textcolor{mathcolor}{$\alpha$}):
-    \begin{quote}
-      \begin{itemize}
-      \item \textcolor{mathcolor}{$H'_\alpha$} is strictly monotonous
-      \item If $\alpha>0$ then \textcolor{mathcolor}{$ i < H'_\alpha(i)$} for all $i$
-        \item $H'_\alpha(i) < H'_{\alpha+1}(i) \; (i \not=0)$
-      \item   If \textcolor{mathcolor}{$\alpha \rplus{n,n,\dots,n} \beta$} then \textcolor{mathcolor}{$H'_\beta(n)\leq H'_\alpha(n)$} 
-      \end{itemize}
-    \end{quote}
-
-  \end{block}
   \begin{block}{Theorem}
@@
 \begin{frame}
-
+  \frametitle{Proof sketch}
   \begin{scriptsize}
@@
 
-       \input{../../movies/snippets/F_omega/lemmas}
-
+Comparison of the length of hydra-battles (w.r.t to the initial replication factor) with the diagonalized Ackermann function proves the theorem.
   \end{block}
@@
 \begin{frame}
-  \frametitle{Hydras \& Gaia}
-  \begin{block}{From \gaia to \Hydras}
-    {\footnotesize
-    \inputsnippets{./multA}}
-  \end{block}
-\end{frame}
-
-\begin{frame}
+  \frametitle{Compatibility with Gaia (José Grimm)}
   \begin{block}{From \Hydras to \gaia}
@@
   {\color{lightblue}{(* end snippet Ex42b *)}}
-  {\color{lightblue}{(* begin snippet Ex42d *)}}
-  Check AP_plus_closed. {\color{lightblue}{(* .unfold .no-goals *)}}
-  {\color{lightblue}{(* end snippet Ex42d *)}}
-   {\color{lightblue}{(* begin snippet Ex42c *)}}
-  assert (Hlt: omega < omega^2) by 
-      (rewrite omega_eqn; apply phi0_mono, finite_mono;
-        auto with arith).
- {\color{lightblue}{ (* end  snippet Ex42c *)}}
  \end{alltt}
@@
     By Lemma \textit{AP\_plus\_closed}, it suffices to prove  $\omega<\omega^2$ and $42<\omega^2$.
-  \end{small}
-
-\end{frame}
-
-\begin{frame}
-  \begin{small}
-    \inputsnippets{
-      ../Schutte_Ex42c, ../Schutte_Ex42d, ../Schutte_Ex42e}
   \end{small}

### 2022-05-02 doc/JFLA2022/Slides/Slides.tex
minor changes in theories/gaia

@@
     \vspace{4pt}
+  \end{block}
 
-\emph{We use   Russel O'Connor's formalization of PR functions.}
+  \begin{block}{Sketch of proof}
+
+    We prove and apply a famous lemma:
+
+       \begin{quote}
+
+      {\color{mathcolor}
+        \[
+          \exists q, \forall x_1,\dots,x_n, 
+          f(x_1,\dots,\,x_n)\leq\textrm{Ack}(q,\textrm{max}(x_1,\dots,x_n))
+        \]}
+    \end{quote}
+
     \end{block}
@@
 \begin{frame}
-  \frametitle{Proof sketch}
+  \frametitle{}
   \begin{scriptsize}
@@
 
-    \begin{quote}
-      For any \textcolor{mathcolor}{$n$}, and any primitive recursive function \textcolor{mathcolor}{$f$} of  arity \textcolor{mathcolor}{$n$}, there exists some natural number \textcolor{mathcolor}{$q$} such that the following inequality holds:
-      {\color{mathcolor}
-        \[
-          \forall x_1,\dots,x_n, 
-          f(x_1,\dots,\,x_n)\leq\textrm{Ack}(q,\textrm{max}(x_1,\dots,x_n))
-        \]}
-    \end{quote}
-
 Comparison of the length of hydra-battles (w.r.t to the initial replication factor) with the diagonalized Ackermann function proves the theorem.
@@
 \end{frame}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{frame}
-  \frametitle{Compatibility with Gaia (José Grimm)}
-  \begin{block}{From \Hydras to \gaia}
-    {\footnotesize
-      \inputsnippets{./canonDef}
-      \inputsnippets{./gcanonLimitOf}}
-  \end{block}
-\end{frame}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \begin{itemize}
+   \item Compatibility with \gaia
   \item Integration with \community
@@
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+  \begin{figure}[ht]
+\centering
+\scalebox{0.7}{
+\begin{tikzpicture}[thick]
+
+  \begin{scope}[yshift=1.5cm, xshift=-1.5cm]
+\node[ellipse, dotted, draw= black!20,fill=green!20,minimum height=0.5cm,minimum width=1.5cm] (acl2) { [ACL2]\footnote{P. Manolios \&  D. Vroon}};
+\end{scope}
+
+\begin{scope}[xshift=-6cm]
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel\footnote{R. O'Connor} };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (additions) { Additions };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (cantor) { Cantor\footnote{C. \& E. Contejean} };
+
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3.2cm] (cats) { Categories in ZFC };
+\end{scope}
+
+\begin{scope}[yshift=-1cm,xshift=0.45cm]
+  \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=2.5cm] (mathcomp) { MathComp };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=paramcoq, node distance=4.5cm] (equations) { Equations };
+\end{scope}
+
+\begin{scope}[yshift=-2cm,xshift=-4cm]
+\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
+\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,left of=hydras, node distance=3.5cm] (pockcc) { Pocklington };
+\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=hydras, node distance=2.5cm] (chains) { Addition-chains };
+\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,right of=chains, node distance=2.2cm] (gaia) { Gaia\footnote{J. Grimm} };
+\end{scope}
+
+\begin{scope}[yshift=-3cm,xshift=-6cm]
+ \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
+ \node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=3cm] (gaiahydras) { Gaia-hydras };
+\end{scope}
+
+\draw[lookcolor,->,dotted] (acl2.south) -- (cantor.north) ;
+\draw[lookcolor,->,dashed] (additions.south) -- (chains.north) ;
+\draw[lookcolor,->,dashed] (goedel) -- (goedelcc) ;
+\draw[lookcolor,->,dashed] (goedel) -- (hydras.north west) ;
+\draw[lookcolor,->,dashed] (cantor) -- (hydras.north east) ;
+\draw[lookcolor,->,dashed] (cantor.south) -- (gaia.north west) ;
+\draw[lookcolor,->,dashed] (pock) -- (pockcc) ;
+\draw[lookcolor,->,dashed] ([xshift=2.1cm]cats) -- (gaia) ;
+
+\draw[->] (hydras) -- (gaiahydras) ;
+\draw[->] (hydras.south west) -- (goedelcc) ;
+\draw[->] (gaia.south west) -- (gaiahydras.east) ;
+
+\draw[->] (equations) -- (hydras) ;
+\draw[->] (mathcomp) -- (gaia) ;
+\draw[->] (mathcomp) -- (chains) ;
+\draw[->] (paramcoq.south) -- (chains) ;
+\draw[->] (pockcc) -- (goedelcc) ;
+\end{tikzpicture}
+}
+\caption{Genealogy and dependencies for \Hydras packages.}
+% Dotted boxes represent historical Coq contributions, while regular boxes represent maintained \coq packages.
+Orange packages are maintained in the \Hydras GitHub repository, while blue packages are maintained in other \community repositories.
+Dotted lines represent \coq code ancestry, while regular lines represent direct code dependencies.}
+
+  \label{fig:genealogy}
+\end{figure}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+  \frametitle{Compatibility with Gaia}
+  \begin{block}{From \Hydras to \gaia}
+    {\footnotesize
+      \inputsnippets{./canonDef}
+      \inputsnippets{./gcanonLimitOf}}
+  \end{block}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Maintenance, CI/CD}
+
+  \begin{block}{Reverse Dependencies Compatibility Testing}
+    Since we have extracted part of {\color{plugincolor}goedel}, now {\color{plugincolor}goedel} depends on {\color{plugincolor}\Hydras}, and thus we are checking in CI that we do not break it (thanks to the \textcolor{lookcolor}{Coq Nix Toolbox},
+    which can generate such CI workflows)
+  \end{block}
+
+  \begin{block}{}
+  {\color{lookcolor}  With \Hydras, we hope to demonstrate that projects with complex dependencies are feasible to manage using recent advances in build management and infrastructure.}
+  \end{block}
+
+\end{frame}
 
@@
     \begin{itemize}
+ \item     The continuously evolving version of the book is continuously deployed from the master branch to the \community website.
+
+     \item
+      \coq libraries contain \emph{snippet declarations} which are converted during the compilation of the book into \LaTeX\xspace blocks which are inserted into the book.
+
     \item The order in which \coq code, goals and answers are presented in the book is determined by the difficulty of the mathematical examples, or the assumed level in \coq of the reader, not by the structure of the libraries.
-    \item
-      \emph{\coq libraries contain \emph{snippet declarations} which are converted during the compilation of the book into \LaTeX\xspace blocks which are inserted into the book.}
 
-    \item {\color{lookcolor} \emph{Note that our build system guarantees that the accompanying book is always up-to-date and consistent with the \coq code.}}
     \end{itemize}
@@
 
-\begin{frame}
-  \frametitle{Package organization and dependencies}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{figure}[ht]
-\centering
-\scalebox{0.7}{
-\begin{tikzpicture}[thick]
+% \begin{frame}
+%   \frametitle{Package organization and dependencies}
 
-\begin{scope}[xshift=-6cm]
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (additions) { Additions };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (cantor) { Cantor };
+% \begin{figure}[ht]
+% \centering
+% \scalebox{0.7}{
+% \begin{tikzpicture}[thick]
 
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3.2cm] (cats) { Categories in ZFC };
-\end{scope}
+% \begin{scope}[xshift=-6cm]
+% \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel };
+% \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington };
+% \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (additions) { Additions };
+% \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (cantor) { Cantor };
 
-\begin{scope}[yshift=-1cm,xshift=0.45cm]
-  \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=2.5cm] (mathcomp) { MathComp };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=paramcoq, node distance=4.5cm] (equations) { Equations };
-\end{scope}
+% \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3.2cm] (cats) { Categories in ZFC };
+% \end{scope}
 
-\begin{scope}[yshift=-2cm,xshift=-4cm]
-\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
-\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,left of=hydras, node distance=3.5cm] (pockcc) { Pocklington };
-\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=hydras, node distance=2.5cm] (chains) { Addition-chains };
-\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,right of=chains, node distance=2.2cm] (gaia) { Gaia };
-\end{scope}
+% \begin{scope}[yshift=-1cm,xshift=0.45cm]
+%   \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
+% \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=2.5cm] (mathcomp) { MathComp };
+% \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=paramcoq, node distance=4.5cm] (equations) { Equations };
+% \end{scope}
 
-\begin{scope}[yshift=-3cm,xshift=-6cm]
- \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
- \node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=3cm] (gaiahydras) { Gaia-hydras };
-\end{scope}
+% \begin{scope}[yshift=-2cm,xshift=-4cm]
+% \node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
+% \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,left of=hydras, node distance=3.5cm] (pockcc) { Pocklington };
+% \node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=hydras, node distance=2.5cm] (chains) { Addition-chains };
+% \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,right of=chains, node distance=2.2cm] (gaia) { Gaia };
+% \end{scope}
 
-\draw[lookcolor,->,dashed] (additions.south) -- (chains.north) ;
-\draw[lookcolor,->,dashed] (goedel) -- (goedelcc) ;
-\draw[lookcolor,->,dashed] (goedel) -- (hydras.north west) ;
-\draw[lookcolor,->,dashed] (cantor) -- (hydras.north east) ;
-\draw[lookcolor,->,dashed] (cantor.south) -- (gaia.north west) ;
-\draw[lookcolor,->,dashed] (pock) -- (pockcc) ;
-\draw[lookcolor,->,dashed] ([xshift=2.1cm]cats) -- (gaia) ;
+% \begin{scope}[yshift=-3cm,xshift=-6cm]
+%  \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
+%  \node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=3cm] (gaiahydras) { Gaia-hydras };
+% \end{scope}
 
-\draw[->] (hydras) -- (gaiahydras) ;
-\draw[->] (hydras.south west) -- (goedelcc) ;
-\draw[->] (gaia.south west) -- (gaiahydras.east) ;
+% \draw[lookcolor,->,dashed] (additions.south) -- (chains.north) ;
+% \draw[lookcolor,->,dashed] (goedel) -- (goedelcc) ;
+% \draw[lookcolor,->,dashed] (goedel) -- (hydras.north west) ;
+% \draw[lookcolor,->,dashed] (cantor) -- (hydras.north east) ;
+% \draw[lookcolor,->,dashed] (cantor.south) -- (gaia.north west) ;
+% \draw[lookcolor,->,dashed] (pock) -- (pockcc) ;
+% \draw[lookcolor,->,dashed] ([xshift=2.1cm]cats) -- (gaia) ;
 
-\draw[->] (equations) -- (hydras) ;
-\draw[->] (mathcomp) -- (gaia) ;
-\draw[->] (mathcomp) -- (chains) ;
-\draw[->] (paramcoq.south) -- (chains) ;
-\draw[->] (pockcc) -- (goedelcc) ;
-\end{tikzpicture}
-}
-\caption{\textcolor{lookcolor}{Genealogy} and dependencies for \Hydras packages.}
+% \draw[->] (hydras) -- (gaiahydras) ;
+% \draw[->] (hydras.south west) -- (goedelcc) ;
+% \draw[->] (gaia.south west) -- (gaiahydras.east) ;
+
+% \draw[->] (equations) -- (hydras) ;
+% \draw[->] (mathcomp) -- (gaia) ;
+% \draw[->] (mathcomp) -- (chains) ;
+% \draw[->] (paramcoq.south) -- (chains) ;
+% \draw[->] (pockcc) -- (goedelcc) ;
+% \end{tikzpicture}
+% }
+% \caption{\textcolor{lookcolor}{Genealogy} and dependencies for \Hydras packages.}
 % Dotted boxes represent historical Coq contributions, while regular boxes represent maintained \coq packages.
@@
 
-  \label{fig:genealogy}
-\end{figure}
+%   \label{fig:genealogy}
+% \end{figure}
 
-%%%% TODO : Resize  !
-{\scriptsize
-\begin{table}[ht]
-\centering
-\scriptsize
-\begin{tabular}{llllrr}
-\hline
-\textbf{Package name} &  \textbf{Version} & \textbf{Spec LOC} & \textbf{Proof LOC}\\
-\hline
-Pocklington & \texttt{8.12.0} & 825 & 3,798 \\
-Goedel  & \texttt{8.13.0} & 2,799 & 10,762 \\
-Gaia &  \texttt{1.12} & 28,850 & 124,839 \\
-Hydra-battles & \texttt{0.5} & 14,368 & 48,385 \\
-Addition-chains & \texttt{0.5} & 1,961 & 2,210 \\
-Gaia-hydras &  \texttt{0.8} & 677 & 2,023 \\
-\hline
-\end{tabular}
-\caption{\scriptsize Current numbers of lines of code for packages in the \Hydras galaxy.}
-\label{tbl:loc}
-\end{table}}
-\end{frame}
+% %%%% TODO : Resize  !
+% {\scriptsize
+% \begin{table}[ht]
+% \centering
+% \scriptsize
+% \begin{tabular}{llllrr}
+% \hline
+% \textbf{Package name} &  \textbf{Version} & \textbf{Spec LOC} & \textbf{Proof LOC}\\
+% \hline
+% Pocklington & \texttt{8.12.0} & 825 & 3,798 \\
+% Goedel  & \texttt{8.13.0} & 2,799 & 10,762 \\
+% Gaia &  \texttt{1.12} & 28,850 & 124,839 \\
+% Hydra-battles & \texttt{0.5} & 14,368 & 48,385 \\
+% Addition-chains & \texttt{0.5} & 1,961 & 2,210 \\
+% Gaia-hydras &  \texttt{0.8} & 677 & 2,023 \\
+% \hline
+% \end{tabular}
+% \caption{\scriptsize Current numbers of lines of code for packages in the \Hydras galaxy.}
+% \label{tbl:loc}
+% \end{table}}
+%\end{frame}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{frame}
-  \frametitle{Maintenance, CI/CD}
-  \begin{block}{CD}
-    The continuously evolving version of the book is continuously deployed from the master branch to the \community website.
-  \end{block}
-
-  \begin{block}{Reverse Dependencies Compatibility Testing}
-    Since we have extracted part of {\color{plugincolor}goedel}, now {\color{plugincolor}goedel} depends on {\color{plugincolor}\Hydras}, and thus we are checking in CI that we do not break it (thanks to the \textcolor{lookcolor}{Coq Nix Toolbox},
-    which can generate such CI workflows)
-  \end{block}
-
-  \begin{block}{}
-  {\color{lookcolor}  With \Hydras, we hope to demonstrate that projects with complex dependencies are feasible to manage using recent advances in build management and infrastructure.}
-  \end{block}
-
-\end{frame}
 
@@
   \begin{block}{Library on ordinals}
-    There exist already \coq libraries which formalize ordinal numbers.
       \begin{itemize}
-    \item \gaia (José Grimm) : Formalization of Bourbaki's Elements of Mathematics.
-{\color{lookcolor}\emph{This library uses an ordinal representation compatible with ours (from the \textcolor{plugincolor}{Cantor} contrib) .
-It is  written in SSReflect/Mathcomp, and is also maintained in \community}.}
-    \item A formalization by Minki Cho, using a different representation.
+  \item \gaiahydra:
+    \begin{itemize}
+    \item Remove definition duplicates
+     \item types \texttt{T2}, \texttt{T3}, connection between 
+    Schütte's model with \texttt{ssete10} (Bourbaki' set theory).
+    \end{itemize}
+
+    \item Formalization by Minki Cho, using a different representation.
 {\color{darkorange}
@@
 
-    Both libraries contain set-theoretical results about larger ordinals than ours.
-Having a library containing both our combinatorial results (large sets and rapidly growing functions) and mathematical definitions of ordinals would be an interesting project.
   \end{block}
@@
 \begin{frame}
-  \frametitle{More TODOs}
   \begin{block}{}
@@
     \item More proof and specification patterns
+    \item More examples and exercises
     \item etc.
@@
   \end{block}
+\end{frame}
 
-  \begin{block}{Call to contributors }
+  \begin{frame}
+\frametitle{Call to contributors}
+  \begin{block}{}
   {\color{plugincolor}\Hydras} and  {\color{plugincolor}\community} are on GitHub. {\color{cyan}Please contribute, reuse!}

### 2022-05-06 doc/gaia-chapter.tex
minor changes in theories/gaia
minor corrections

@@
 
-The bridge we aim to build and cross is a two-way bridge:
+The main objective of the \texttt{theories/gaia/*.v} files is
+to provide \gaia's users definitions and lemmas on
+some combinatorial aspects of ordinal numbers less than $\epsilon_0$ \cite{KS81}, already proven in \Hydras: canonical sequences, accessibility properties, large sets and rapidly growing arithmetical functions.
+We may also complete \Hydras with some lemmas proven in \gaia.
 
-\begin{itemize}
-\item \gaia contains lemmas on ordinal arithmetic (\emph{e.g.} associativity of multiplication of ordinals strictly less than $\epsilon_0$), which were missing from \Hydras. 
-  \item The other way, \Hydras contains formal definitions and proofs of properties of canonical sequences and rapidly growing functions  (see Chapters~\ref{chap:ketonen} and~\ref{chap:alpha-large}~of this book).
-  \end{itemize}
-
-  We plan to make accessible the first kind of lemmas to \HydrasLib' users and the second kind to \gaia's.
-For this purpose, we write modules dedicated to the importation of definitions and lemmas from each of both libraries into the other one.
+For these purposes, we write modules dedicated to the importation of definitions and lemmas from each of both libraries into the other one.
   Please note that by ``import'' we do not mean ``duplicate''
@@
 
-Many constants are common to \HydrasLib and \gaia. For instance, the name \texttt{zero} may refer to \gaia's
+Likewise, many constants are common to \HydrasLib and \gaia. For instance, the name \texttt{zero} may refer to \gaia's
 \texttt{gaia.ssete9.CantorOrdinal.zero}  and
@@
 \subsubsection{A few constants}
-For each constant: $0$, $1$, \dots, $n$ and $\omega$, we prove
+For each constant: $0$, $1$,\dots, $n$ and $\omega$, we prove
 that \texttt{hydras}' definition refines \texttt{gaia}'s.
@@
 \label{sect:lt-compat-gaia}
-Leibniz equality and strict order on both libraries are
+The  strict order on both types \texttt{T1} are
 compatible:
@@
 
-The binary operations $+$ and $\times$ are defined the same way in both libraries. Please note that they require comparison of ordinal terms. So, the proof of the following lemmas applies
+The definition of the binary operations $+$ and $\times$
+requires the comparison of ordinal terms, which is not implemented the same way in both libraries. Thus, the proof of the following lemmas applies applies
 compatibility lemmas like \texttt{compare\_ref} (see Section~\vref{sect:gaia-compare-ref}).
@@
 
-\subsubsection{Cantor normal form}
+\subsubsection{Well formed ordinal terms (Cantor normal form)}
 \label{nf-gaia-compat}
@@
 \href{../theories/html/hydras.Epsilon0.Hessenberg.html}{hydras.Epsilon0.Hessenberg} and is used for proving termination of all hydra battles (See Sect.~\ref{sect:hessenberg-def}).
-We define this function in \href{../theories/html/gaia\_hydras.GHessenberg.html}{gaia\_hydras.GHessenberg} by delegation to \HydrasLib'.
+
+Instead of defining this operation \emph{ex nihilo} in \texttt{theories/gaia}, we chose to \emph{import} the definition
+from \href{../theories/html/hydras.Epsilon0.Hessenberg.html}{hydras.Epsilon0.Hessenberg} as follows:
 
@@
 
-\inputsnippets{GHessenberg/oplusNf}
+\inputsnippets{GHessenberg/oplusEquationsc}
 \inputsnippets{GHessenberg/oplusEquations}
 \inputsnippets{GHessenberg/oplusEquationsb}
-\inputsnippets{GHessenberg/oplusEquationsc}
+\inputsnippets{GHessenberg/oplusNf}
 
@@
 
-\inputsnippets{GPaths/KSThm24}
+%\inputsnippets{GPaths/KSThm24}
 \inputsnippets{GPaths/Cor12}
@@
 
-We plan to automatize as soon as possible this kind of transfer of lemmas from \gaia to \HydrasLib.
+We plan to automatize as soon as possible the proof of this kind of transfer lemmas from \gaia to \HydrasLib.
 

### 2022-05-06 doc/gaia-chapter.tex
fix an error in latex doc

@@
 \label{sect:lt-compat-gaia}
-The  strict order on both types \texttt{T1} are
-compatible:
 
-\inputsnippets{T1Bridge/eqRef}
+The  strict orders on types \texttt{T1} and \texttt{hT1} are
+compatible.
 Note that \HydrasLib' comparison on ordinal terms uses a

### 2022-05-08 doc/Gamma0-chapter.tex
reorganization of tex files

--- 2022-04-11 doc/part-hydras.tex
@@
-\section*{Introduction}
-
-In this part, we present a development  for the \coq{} proof assistant, after the work of Kirby and Paris~\cite{KP82}. This formalization contains the following main parts:
-
-\begin{itemize}
-\item Representation in \coq{} of hydras and hydra battles.
-\item A proof that every battle is finite and won by Hercules.
-This proof is based on a \emph{variant} which maps any hydra to an ordinal strictly less than $\epsilon_0$ and is strictly decreasing along any battle.
-
-\item Using a combinatorial toolkit designed by J.~Ketonen and R.~Solovay~\cite{KS81}, we prove that, for any ordinal $\mu<\epsilon_0$, there exists no such variant mapping any hydra to an ordinal strictly less than $\mu$.
-Thus, the complexity of $\epsilon_0$ is really needed in the previous proof.
-
-\item We prove a relation between the length of a ``classic''  kind of  battles \footnote{This class is also called \emph{standard} in this document (text and proofs).
-The \emph{replication factor} of the hydra is exactly $i$ at the $i$-th round of the battle (see Sect~\vref{sect:replication-def}).}
-and the Wainer-Hardy hierarchy of ``rapidly growing functions'' $H_\alpha$~\cite{Wainer1970}.
-The considered class of battles, which we call \emph{standard},  is the most considered one in the scientific  literature (including popularization).
-\end{itemize}
-
-Simply put, this document tries to combine the scientific interest of two articles~\cite{KP82, KS81} and a book~\cite{schutte} with the playful activity of truly proving theorems.
-We hope  that such a work, besides exploring a nice piece of discrete mathematics,
-will show how \coq{} and its standard library are well fitted to help us to understand some non-trivial mathematical developments, and also to experiment the constructive parts of  the proof through functional programming.
-
- We also hope to provide a little clarification on infinity (both potential and actual) through the notions of function, computation, limit,
- type and proof.
-
- \subsection*{Compatibility with \gaia (in progress)}
- \label{sect:gaia-first-intro}
-The \gaia project~\cite{Gaia} by Jos\'e Grimm, Alban Quadrat,  and Carlos Simpson,
-aims  to formalize mathematics in \coq  in
-the style of Nicolas Bourbaki.
-It contains many definitions and results about ordinal numbers.
-In Chapter~\ref{gaia-chapter}, we present some modules which allow \HydrasLib' users to apply lemmas proven in \gaia, and vice versa.
-Remarks about compatibility with \gaia are signalled with the picture \gaiasign.
-
-%\section{Remarks}
-
-\subsection*{Difference from Kirby and Paris's work}
-In~\cite{KP82}, Kirby and Paris show  that there is no proof of termination of all hydra battles in Peano Arithmetic (PA).
-Since we are used to writing proofs in higher order logic, the restriction to PA was quite unnatural for us.
-So we chose to prove another statement without any reference to PA, by considering a class of proofs indexed by ordinal numbers up to $\epsilon_0$.
-
-\subsection*{State of the development}
-The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on  notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
-We added new material: axiomatic definitions of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, function derinition by equations, etc.
-
-We are now working in order to make clumsy proofs more readable, simplify definitions, and ``factorize'' proofs as much as possible. 
-Many possible improvements are suggested as ``todo''s or ``projects'' in this text.
-
-\section*{Future work (projects)}
-\index{hydras}{Projects}
-
-This document and the proof scripts are far from being complete.
-
-First, there must be a lot of typos to correct, references and index items to add. Many proofs are too complex and should be simplified, etc.
-
-The following extensions are planned, but help is needed:
-
-\begin{itemize}
-\item Semiautomatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
-%\item Extension to $\Gamma_0$ (in Veblen normal form)
-\item More lemmas about hierarchies of rapidly growing
-  functions, and their relationship 
-    with primitive recursive functions and provability in Peano arithmetic 
-(following~\cite{KS81, KP82}).
-\item From \coq's point of view, this development could be used as an illustration of the evolution of the software, every time new libraries and sets of tactics could help to simplify the proofs.
-\end{itemize}
-
-\subsection*{Main references}
-
-In our development, we adapt the definitions and prove many theorems which
-we found in the following articles. 
-\begin{itemize}
-\item ``Accessible independence results for Peano arithmetic''  by Laurie Kirby and Jeff Paris~\cite{KP82}
-\item ''Rapidly growing Ramsey Functions'' by Jussi Ketonen and Robert Solovay~\cite{KS81}
-\item ``The Termite and the Tower'', by Will Sladek~\cite{Sladek07thetermite}
-\item Chapter V of ``Proof Theory'' by Kurt Schütte~\cite{schutte}
-\end{itemize}
-
-\chapter{Hydras and hydra games}
-
-\label{sec:orgheadline91}
-\label{chapter:hydras}
-
-This chapter is dedicated to the representation of hydras and rules of the hydra game in \coq's specification language:~\gallina.
-
-Technically, a \emph{hydra} is just a finite ordered tree, each node of which 
-has any number of sons. Contrary to the computer science tradition, we display the hydras 
-with the heads up and the foot (i.e., the root of the tree) down.
-Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the 
-module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}). 
-\emph{For a less formal description of hydras, please see 
-\url{https://www.smbc-comics.com/comic/hydra}.}
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {$\bullet$};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H4) at (4,2) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N3) to [bend left= 10]  (H1) ;
-\draw (N3) to [bend right= 16] (H2);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (foot) to [bend right = 10] (H4) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-\caption{The hydra \texttt{Hy} \label{fig:Hy}}
-\end{figure}
-
-We use a specific vocabulary for talking about hydras. Table~\ref{tab:hyd2tree} shows the correspondence between our terminology and the usual vocabulary for trees in computer science.
-
-\begin{figure}[h]
-  \centering
-  \begin{tabular}{ll}
-Hydras & Finite rooted trees\\
-\hline
-foot & root\\
-head & leaf\\
-node & node\\
-segment  & (directed) edge \\
-sub-hydra & subtree\\
-daughter & immediate subtree\\
-\end{tabular}
-  \caption{Translation from hydras to trees}
-  \label{tab:hyd2tree}
-\end{figure}
-
-The hydra \texttt{Hy} has a \emph{foot} (below), five \emph{heads}, and eight \emph{segments}. 
-We leave it to the reader to define various parameters such as the height, the size, the highest arity (number of sons of a node) of a hydra.
-In our example, these parameters have the respective values $4$, $9$ and $3$.
-
-\subsection{The rules of the game}
-
-\label{sec:orgheadline44}
-\label{sect:replication-def}
-
-A \emph{hydra battle} is a fight between Hercules and the Hydra. 
-More formally, a  battle is a sequence of \emph{rounds}.
-At each round:
-\begin{itemize}
-\item If the hydra is composed of just one head, the battle is finished
-and  Hercules is the winner.
-\item Otherwise, Hercules chops off \emph{one} head of the hydra,
-
-\begin{itemize}
-\item If the head is at distance 1 from the foot, the head is just lost by the hydra, with no more reaction.
-\item Otherwise, let us denote by \(r\) the node that was at distance \(2\) from 
-the removed head in the direction of the foot,  and consider the  sub-hydra \(h'\) of \(h\), whose  root is \(r\) \footnote{$h'$ will be called ``the wounded part of the hydra'' in the subsequent text.
-In Figures~\vref{fig:Hy2} and ~\vref{fig:Hy4}, this sub-hydra  is displayed in red.}.
-Let $n$ be some natural number.
-Then $h'$ is replaced by  $n+1$ of copies of \(h'\) which share the same root $r$.
- The \emph{replication factor} $n$ may be different (and generally is)   at each round of the fight.
-It may be chosen by the hydra, according to its strategy, or imposed by some 
-particular rule.
-In many presentations of hydra battles, this number is increased by $1$ at each round.
-In the following presentation, we will also consider battles where the hydra is free to choose its ~replication factor at each round of the battle\footnote{Let us recall that, if the chopped-off head was at distance 1 from the foot, the replication factor is meaningless.}.
-\end{itemize}
-\end{itemize}
-
-Note that the description given in~\cite{KP82} of the replication process in hydra battles is also  semi-formal.
-
-\label{original-rules}
-
-\begin{quote}
-  ``From the node that used to be attached to the head which was just chopped off, traverse one 
-segment towards the root until the next node is reached. From this node sprout $n$ replicas of 
-that part of the hydra (after decapitation) which is ``above'' the segment just traversed, i.e., those 
-nodes and segments from which, in order to reach the root, this segment would have to be 
-traversed. If the head just chopped off had the root of its nodes, no new head is grown. ''
-\end{quote}
-
-Moreover, we note that this description is in \emph{imperative} terms.
-In order to formally study the properties of hydra battles, we prefer to use a mathematical vocabulary, i.e., graphs, relations, functions, etc.
-Thus, the replication process will be represented as a binary relation on a data type \texttt{Hydra},
-linking the state of the hydra \emph{before} and \emph{after} the transformation.
-A battle will thus be represented as a sequence of terms of type \texttt{Hydra}, respecting the rules of the game.
-
-\subsection{Example}
-Let us start a battle between Hercules and the hydra \texttt{Hy} of Fig.~\ref{fig:Hy}.
-
-At the first round, Hercules chooses to chop off the rightmost head of \texttt{Hy}.
-Since this head is near the floor, the hydra simply loses this head. Let us call 
- \texttt{Hy'} the resulting state of the hydra, represented in Fig.~\vref{fig:Hy-prime}.
-
-Next, assume Hercules chooses to chop off one of the two highest heads of \texttt{Hy'}, for instance the rightmost one.
-Fig.~\vref{fig:Hy2} represents the broken segment in dashed lines, and the part that will be replicated in red.
-Assume also that the hydra decides to add 4 copies of the red part\footnote{In other words, the replication factor at this round is equal to $4$.}.
-We obtain a new state \texttt{Hy''} depicted in Fig.~\ref{fig:Hy3}.
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {$\bullet$};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-%\node (H4) at (6,0) {$\Xey[2][lightgray]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3) ;
-\draw (N3) to [bend left= 10]  (H1) ;
-\draw (N3) to [bend right= 16] (H2);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-
-\caption{\texttt{Hy'}: the state  of \texttt{Hy} after one round \label{fig:Hy-prime}}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4)  {{\color{lightred}$\bullet$}};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Sey[2][lightred]$};
-%\node (H2) at (5,0) {$\Xey[2][lightgray]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\node (ex) at (5,8) {};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw  (N2) -- (N3)[draw=lightred];
-\draw (N3) to   [bend left= 10](H1) [draw=lightred];
-\draw [dashed] (N3) to [bend left= 10](ex);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-\caption{A second beheading}
-\label{fig:Hy2}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H11) at (2,8) {$\Smiley[2][vertfluo]$};
-\node (H12) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H13) at (6,8) {$\Smiley[2][vertfluo]$};
-\node (H14) at (8,8) {$\Smiley[2][vertfluo]$};
-
-\node (N3) at (1,6) {$\bullet$};
-\node (N31) at (2,6) {$\bullet$};
-\node (N32) at (3,6) {$\bullet$};
-\node (N33) at (4,6) {$\bullet$};
-\node (N34) at (5,6) {$\bullet$};
-
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N2) -- (N31);
-\draw (N2) -- (N32);
-\draw (N2) -- (N33);
-\draw (N2) -- (N34);
-\draw (N3) to   [bend left= 10](H1) ;
-\draw (N31) to   [bend left= 10](H11) ;
-\draw (N32) to   [bend left= 10](H12) ;
-\draw (N33) to   [bend left= 10](H13) ;
-\draw (N34) to   [bend left= 10](H14) ;
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend left= 10]  (H5) ;
-\end{tikzpicture}
-\caption{\texttt{Hy''}: the state of \texttt{Hy} after two rounds \label{fig:Hy3}}
-\end{figure}
-
-Figs.~\ref{fig:Hy4} and~\vref{fig:Hy5} represent a possible third round of the battle, with a replication factor equal to $2$. Let us call \texttt{Hy'''} the state of the hydra after that third round.
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0)  {{\color{lightred}$\bullet$}};
-\node (N1) at (2,2) {{\color{lightred}$\bullet$}};
-\node (N2) at (2,4) {{\color{lightred}$\bullet$}};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (exN4) at (4,4) {};
-\node (H1) at (0,8) {$\Sey[2][lightred]$};
-\node (H11) at (2,8) {$\Sey[2][lightred]$};
-\node (H12) at (4,8) {$\Sey[2][lightred]$};
-\node (H13) at (6,8) {$\Sey[2][lightred]$};
-\node (H14) at (8,8) {$\Sey[2][lightred]$};
-
-\node (N3) at (1,6) {{\color{lightred}$\bullet$}};
-\node (N31) at (2,6) {{\color{lightred}$\bullet$}};
-\node (N32) at (3,6) {{\color{lightred}$\bullet$}};
-\node (N33) at (4,6) {{\color{lightred}$\bullet$}};
-\node (N34) at (5,6) {{\color{lightred}$\bullet$}};
-
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-%\node (H5) at (4,0) {$\Xey[2][lightgray]$};
-\draw (foot) -- (N1)[very thick,draw=lightred] ;
-\draw (N1) -- (N2)[draw=lightred];
-\draw (N2) -- (N3)[draw=lightred];
-\draw (N2) -- (N31)[draw=lightred];
-\draw (N2) -- (N32)[draw=lightred];
-\draw (N2) -- (N33)[draw=lightred];
-\draw (N2) -- (N34)[draw=lightred];
-\draw (N3) to   [bend left= 10](H1) [draw=lightred];
-\draw (N31) to   [bend left= 10](H11) [draw=lightred];
-\draw (N32) to   [bend left= 10](H12) [draw=lightred];
-\draw (N33) to   [bend left= 10](H13) [draw=lightred];
-\draw (N34) to   [bend left= 10](H14) [draw=lightred];
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw [dashed] (N1) to  [bend left= 10](exN4);
-\end{tikzpicture}
-\caption{A third beheading (wounded part in red) \label{fig:Hy4}}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-
-\node (foot) at (10,0) {$\bullet$};
-
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H1) at (0,8) {$\Smiley[1][vertfluo]$};
-\node (H11) at (2,8) {$\Smiley[1][vertfluo]$};
-\node (H12) at (4,8) {$\Smiley[1][vertfluo]$};
-\node (H13) at (6,8) {$\Smiley[1][vertfluo]$};
-\node (H14) at (8,8) {$\Smiley[1][vertfluo]$};
-
-\node (N3) at (1,6) {$\bullet$};
-\node (N31) at (2,6) {$\bullet$};
-\node (N32) at (3,6) {$\bullet$};
-\node (N33) at (4,6) {$\bullet$};
-\node (N34) at (5,6) {$\bullet$};
-
-\node (H0) at (-3,3) {$\Smiley[1][vertfluo]$};
-
-\draw (foot) to [bend left=10] (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N2) -- (N31);
-\draw (N2) -- (N32);
-\draw (N2) -- (N33);
-\draw (N2) -- (N34);
-\draw (N3) to   [bend left= 10](H1) ;
-\draw (N31) to   [bend left= 10](H11) ;
-\draw (N32) to   [bend left= 10](H12) ;
-\draw (N33) to   [bend left= 10](H13) ;
-\draw (N34) to   [bend left= 10](H14) ;
-\draw (foot) to [bend left = 15]  (H0) ;
-
-% second copy 
-\node (N01) at (12,2) {$\bullet$};
-\node (N02) at (12,4) {$\bullet$};
-\node (N03) at (12,6) {{\color{lightred}$\bullet$}};
-\node (H001) at (10,8) {$\Smiley[1][vertfluo]$};
-\node (H0011) at (12,8) {$\Smiley[1][vertfluo]$};
-\node (H0012) at (14,8) {$\Smiley[1][vertfluo]$};
-\node (H0013) at (16,8) {$\Smiley[1][vertfluo]$};
-\node (H0014) at (18,8) {$\Smiley[1][vertfluo]$};
-
-\node (N03) at (11,6) {$\bullet$};
-\node (N031) at (12,6) {$\bullet$};
-\node (N032) at (13,6) {$\bullet$};
-\node (N033) at (14,6) {$\bullet$};
-\node (N034) at (15,6) {$\bullet$};
-
-\draw (foot) -- (N01)[very thick] ;
-\draw (N01) -- (N02);
-\draw (N02) -- (N03);
-\draw (N02) -- (N031);
-\draw (N02) -- (N032);
-\draw (N02) -- (N033);
-\draw (N02) -- (N034);
-\draw (N03) to   [bend left= 10](H001) ;
-\draw (N031) to   [bend left= 10](H0011) ;
-\draw (N032) to   [bend left= 10](H0012) ;
-\draw (N033) to   [bend left= 10](H0013) ;
-\draw (N034) to   [bend left= 10](H0014) ;
-
-% third copy 
-\node (N001) at (22,2) {$\bullet$};
-\node (N002) at (22,4) {$\bullet$};
-\node (N003) at (22,6) {{\color{lightred}$\bullet$}};
-\node (H001) at (20,8) {$\Smiley[1][vertfluo]$};
-\node (H0011) at (22,8) {$\Smiley[1][vertfluo]$};
-\node (H0012) at (24,8) {$\Smiley[1][vertfluo]$};
-\node (H0013) at (26,8) {$\Smiley[1][vertfluo]$};
-\node (H0014) at (28,8) {$\Smiley[1][vertfluo]$};
-
-\node (N003) at (21,6) {$\bullet$};
-\node (N0031) at (22,6) {$\bullet$};
-\node (N0032) at (23,6) {$\bullet$};
-\node (N0033) at (24,6) {$\bullet$};
-\node (N0034) at (25,6) {$\bullet$};
-
-\draw (foot) -- (N001)[very thick] ;
-\draw (N001) -- (N002);
-\draw (N002) -- (N003);
-\draw (N002) -- (N0031);
-\draw (N002) -- (N0032);
-\draw (N002) -- (N0033);
-\draw (N002) -- (N0034);
-\draw (N003) to   [bend left= 10](H001) ;
-\draw (N0031) to   [bend left= 10](H0011) ;
-\draw (N0032) to   [bend left= 10](H0012) ;
-\draw (N0033) to   [bend left= 10](H0013) ;
-\draw (N0034) to   [bend left= 10](H0014) ;
-\end{tikzpicture}
-\caption{The configuration \texttt{Hy'''} of \texttt{Hy} \label{fig:Hy5}}
-\end{figure}
-\FloatBarrier
-
-We leave it to the reader  to guess the following  rounds of the battle \dots
- % Please keep in mind that, in this 
-% the hydra is free to chose any number of replications at each time, whereas
-% Hercules chops only one head per round.
-
-% Let us precise that, in this game, Hercules wins if the hydra is eventually reduced 
-% to a single head. 
-% We know from~\cite{KP82} that, whichever the initial configuration of the
-% hydra, and the strategies of both players, Hercules eventually wins. The 
-% aforementionned paper shows also that there do not exist any \emph{simple} proof of this result.
-
-\section{Hydras and their representation in \emph{Coq}}
-\label{sec:orgheadline48}
-
-\index{hydras}{Library Hydra!Types!Hydra}
-\index{hydras}{Library Hydra!Types!Hydrae}
-
-In order to describe trees where each node can have an arbitrary (but finite) number of sons, it is usual to define a type where each node carries a \emph{forest}, \emph{i.e} a list of trees
-(see for instance Chapter 14, pages 400-406 of \cite{BC04}, also available as~\cite{BC04ch14}).
-
-For this purpose, we define two mutual \emph{ad-hoc}  inductive types, where \texttt{Hydra} is the main type, and \texttt{Hydrae} is a helper for describing finite sequences of hydras.
-\label{types:Hydra}
-\label{types:Hydrae}
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hydra}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/HydraDef}
-
-%\index{To do}
-\index{hydras}{Projects}
-
-\begin{project}
-Look for an existing library on trees with nodes of arbitrary arity, in order to replace  this ad-hoc type with something more generic.
-\end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-
- Another very similar representation could use the \texttt{list} type family instead of the specific 
-type \texttt{Hydrae}:
-
-\input{movies/snippets/Hydra_Definitions/HydraAlt}
-
-Using this representation, re-define all the constructions of this chapter.
-You will probably have to use patterns described for instance in~\cite{BC04} or the archives of the \coq communication channels (please consult~\url{https://coq.inria.fr/community.html}).
-
-\end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-The type \texttt{Hydra} above describes hydras as \emph{plane trees}, i.e., as drawn on a sheet of paper or computer screen. Thus, hydras are \emph{oriented},
-and it is appropriate to consider a \emph{leftmost} or \emph{rightmost} head of
-the beast. It could be interesting to consider another representation, in which
-every non-leaf node has a \emph{multi-set} -- not an ordered list -- of daughters.
-\end{project}
-
-\subsubsection{Abbreviations}
-
-We provide several notations for hydra patterns  which occur often in our developments.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#head}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/headsEtc}
-
-For instance, the hydra \texttt{Hy}  of Figure~\vref{fig:Hy} is defined in \emph{Gallina} as follows:
-
-\vspace{4mm}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html\#Hy}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/Hy}
-
-Hydras quite frequently contain  multiple adjacent  copies of the same subtree. The following functions
-will help us to describe and reason about replications in hydra battles.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#hcons_mult}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hconsMult}
-
-\vspace{4mm}
-
-For instance, the hydra \texttt{Hy''} of Fig~\vref{fig:Hy3}  can be defined in \coq{} as follows:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/HySecond}
-
-\subsubsection{Recursive functions on type \texttt{Hydra}}
-\label{sec:orgheadline41}
-\label{sec:hsize-def}
-
-In order to  define a recursive function over the type \texttt{Hydra}, one has to consider the three constructors 
-\texttt{node}, \texttt{hnil} and \texttt{hcons} of the mutually inductive types \texttt{Hydra} and \texttt{Hydrae}. 
-Let us define for instance the function which  computes the number of nodes of any hydra:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hsize}
-
-\input{movies/snippets/Hydra_Examples/HySize}
-
-Likewise, the \emph{height} (maximum distance between the foot and a head) 
-is defined by mutual recursion:
-
-\input{movies/snippets/Hydra_Definitions/height}
-
-\input{movies/snippets/Hydra_Examples/HyHeight}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Define a function \texttt{max\_degree: Hydra $\arrow$ nat} which  returns the highest degree of a node in any hydra.
-For instance, the evaluation of the term \texttt{(max\_degree Hy)} should return $3$.
-\end{exercise}
-
-\subsection{Induction principles for hydras}
-\label{sec:orgheadline42}
-
-In this section, we show how induction principles are used to prove properties on the type 
-\texttt{Hydra}. Let us consider for instance the following statement:
-\begin{quote}
-  `` The height of any hydra is strictly less than its size. ''
-\end{quote}
-
-\subsubsection{A failed attempt}
-
-One may try to use the default tactic of proof by induction, which corresponds to an application of the automatically  generated  induction principle for  type \texttt{Hydra}:
-
-\input{movies/snippets/Hydra_Examples/HydraInd}
-
-Let us start a simple proof by induction.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/BadInductiona}
-
-We might be tempted to do an induction on the sequence \texttt{s}:
-
-\input{movies/snippets/Hydra_Examples/BadInductionb}
-
-The first subgoal is trivial.
-
-\input{movies/snippets/Hydra_Examples/BadInductionc}
-
-Let us look at the second subgoal of the induction.
-
-\input{movies/snippets/Hydra_Examples/BadInductiond}
-
-We notice that this subgoal does not contain any hypothesis
-on the height and size of the hydra \texttt{h}. So, it looks hard to prove the conclusion. Let's stop.
-
-\input{movies/snippets/Hydra_Examples/BadInductione}
-
-\subsubsection{A Principle of mutual induction}
-In order to get an appropriate induction scheme for the types 
-\texttt{Hydra} and \texttt{Hydrae}, we can use  \coq{}'s  command \texttt{Scheme}.
-
-\index{coq}{Mutually inductive types}
-%\index{coq}{Commands!Scheme}
-
-\input{movies/snippets/Hydra_Definitions/HydraRect2}
-
-\input{movies/snippets/Hydra_Examples/HydraRect2Check}
-
-\subsubsection{A Correct proof}
-
-Let us now use \texttt{Hydra\_rect2} for proving that the height of any hydra is strictly less than its size.
-Using this scheme requires an auxiliary predicate, called \texttt{P0} in \texttt{Hydra\_rect2}'s statement.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hForall}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-\input{movies/snippets/Hydra_Examples/heightLtSizea}
-
-The first subgoal is easily solvable, using some arithmetic.
-The second and third ones are
-almost trivial. We let the reader look at the source. 
-\input{movies/snippets/Hydra_Examples/heightLtSizez}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-It happens very often that, in the proof of  a proposition of the form 
-(\texttt{$\forall\,$ h:Hydra, $P$ h}), the predicate \texttt{P0}
-is  (\texttt{h\_forall $P$}).
-Design a tactic for induction on hydras that frees the user from binding explicitly \texttt{P0},  and solves trivial subgoals.
-Apply it for writing  a shorter proof of \texttt{height\_lt\_size}.
-\end{exercise}
-
-\section{Relational description of hydra battles}
-
-In this section, we represent the rules of hydra battles as a binary relation associated with
-a \emph{round}, i.e., an interaction composed of the two following actions:
-\begin{enumerate}
-\item Hercules chops off one head of the hydra.
-\item Then, the  hydra replicates the wounded part (if the head is at distance $\geq 2$ from the foot).
-\end{enumerate}
-The relation associated with each round of the battle is parameterized  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot,
-but present for consistency's sake).
-
-In our description,  we will apply the following naming convention: if $h$ represents the configuration of the hydra before a round, then the configuration of $h$ after this round will be called $h'$.
- Thus, we are going to define a proposition  (\texttt{round\_n $n\;h\;h'$})  whose intended meaning will be `` the hydra $h$  is transformed into $h'$  in a single round of a battle, with the expected replication factor $n$ ''.
-
-Since the replication of parts of the hydra depends on the distance of the chopped head from  the foot, we  decompose our description into two main  cases, under the form of a bunch of [mutually] inductive predicates over the types \texttt{Hydra} and \texttt{Hydrae}.
-
-The mutually exclusive cases we consider are the following:
-\begin{itemize}
-\item \textbf{R1}: The chopped off head was at distance 1 from the foot.
-\item \textbf{R2}: The chopped off head was at a distance greater than or equal to  $2$ from the foot.
-\end{itemize}
-
-\subsection{Chopping off a head at distance 1 from the foot (relation  R1)}
-
-If Hercules chops off a head near the floor, there is no replication at all. We use an auxiliary 
-predicate \texttt{S0}, associated with the removing of one head from a sequence of hydras.
-
-\vspace{4pt}\emph{From Module\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/S0Def}
-\input{movies/snippets/Hydra_Definitions/R1Def}
-
-\subsubsection{Example}
-\label{sec:orgheadline45}
-
-Let us represent in \coq{}   the transformation of the hydra of Fig.~\vref{fig:Hy} into
-the configuration represented in Fig.~\ref{fig:Hy-prime}.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/Hy1}
-
-\subsection{Chopping off a head at distance \texorpdfstring{$\geq 2$}{>= 2} from the foot (relation R2) }
-
-Let us now consider beheadings  where the chopped-off head is at distance greater than or equal to $2$ from the foot. All the following relations are parameterized by the replication factor  $n$.
-
- Let $s$ be a sequence of hydras. 
-The proposition (\texttt{S1 n s s'}) holds if $s'$ is obtained by replacing some element $h$ of $s$ by 
-$n+1$ copies of $h'$, where  the proposition (\texttt{R1 h h'}) holds, in other words, $h'$ is just $h$, without the chopped-off  head.
-\texttt{S1} is an inductive relation with two constructors that allow us to choose the position in $s'$ of the wounded sub-hydra $h$.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#S1}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/S1Def}
-
-The rest of the definition is composed of two mutually inductive relations on hydras and sequences of hydras.
-The first constructor of \texttt{R2} describes the case where the chopped head is exactly at height $2$.
-The others constructors allow us to consider beheadings at height strictly greater than $2$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/R2Def}
-
-\subsubsection{Example}
-Let us prove the transformation of \texttt{Hy'} into \texttt{Hy''} (see Fig.~\vref{fig:Hy3}).
-We use an experimental set of tactics\footnote{See the \texttt{Ltac} definitions in~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions.}}  for specifying the place where the
-interaction between Hercules and the hydra holds.
-
-\vspace{4pt}\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}.
-
-\input{movies/snippets/Hydra_Examples/R2Example}
-
-The reader is encouraged to look at all the successive subgoals of this example.
-\emph{Please consider also exercise~\vref{exo:interactive-battle}.}
-
-\subsection{Binary relation associated with a round}
-
-We combine the two cases above into a single relation.
-First,  we define the  relation \texttt{(round\_n n h h')} where \texttt{n} is the expected number of  replications (irrelevant in the case of an \texttt{R1}-transformation).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#round_n}{Hydra.Hydra\_Definitions}}
-
-\index{hydras}{Library Hydra!Predicates!round\_n}
-
-\input{movies/snippets/Hydra_Definitions/roundNDef}
-
-By abstraction over \texttt{n}, we define a \emph{round} (small step) of a battle:
-
-\index{hydras}{Library Hydra!Predicates!round}
-\label{sect:infix-round}
-
-\input{movies/snippets/Hydra_Definitions/roundDef}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Give a direct translation of Kirby and Paris's description of hydra battles (quoted on page~\pageref{original-rules}) and prove that our relational description is consistent with theirs.
-\end{project}
-
-\subsection{Rounds and battles}
-
-Using library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}, we define \texttt{round\_plus},  the transitive closure of \texttt{round}, and \texttt{round\_star},  the reflexive and transitive closure of \texttt{round}.
-
-\label{sect:infix-rounds}
-
-\input{movies/snippets/Hydra_Definitions/roundPlus}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-  Prove that if \texttt{$h$ -+-> $h'$}, then
-  the height of $h'$ is less or equal than the height of $h$.
-
-\end{exercise}
-
-\begin{remark}
-\label{remark:transitive-closure}
-\coq's library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Coq.Relations.Relation\_Operators} 
-contains three logically equivalent definitions of the transitive closure of a binary relation. This equivalence is proved in 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} .
-
-Why three definitions for a single mathematical concept?
-Each definition generates an associated induction principle. 
- According to the form of statement one would like to prove, there is a ``best choice'':
-
-\begin{itemize}
-\item For proving $\forall y, x\,R^+\,y \;\arrow\; P\,y$, prefer 
-\texttt{clos\_trans\_n1}
-\item For proving $\forall x,\,x\,R^+\,y \;\arrow\; P\,x$, prefer \texttt{clos\_trans\_1n}
-\item For proving $\forall x\,y, \,x\,R^+\,y \;\arrow\;P\,x\,y$,  
-prefer \texttt{clos\_trans},
-\end{itemize}
-But there is no ``wrong choice'' at all: the equivalence lemmas in \linebreak 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} 
- allow the user
-to convert any one of the three closures into another one before applying the corresponding elimination tactic.
-The same remark also holds for reflexive and transitive closures. 
-\end{remark}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Define a restriction of \coqsimple{round},  where Hercules always chops off
-the leftmost among the lowest heads.
-
-Prove that, if $h$ is not a simple head, then there exists a unique $h'$ such that $h$  is transformed into $h'$ in one round, according to this restriction.
-
-\end{exercise}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}[Interactive battles]
-\label{exo:interactive-battle}
-Given a hydra \texttt{h}, the specification of a hydra battle for \texttt{h} is the type 
-\Verb@{h':Hydra | h -*-> h'}@. In order to avoid long sequences of \texttt{split}, \texttt{left}, and 
-\texttt{right}, design a set of dedicated tactics for the interactive building of a battle.
-Your tactics will have the following functionalities:
-\begin{itemize}
-\item  Chose to stop a battle, or continue
-\item Chose an expected number of replications
-\item Navigate in a hydra, looking for a head to chop off.
-\end{itemize}
-
-Use your tactics for simulating a small part of a hydra battle, for instance the rounds which lead from
-\texttt{Hy} to \texttt{Hy'''}  (Fig.~\vref{fig:Hy5}).
-
-\textbf{Hints:} 
-\begin{itemize}
-
-\item Please keep in mind that the last  configuration of your interactively built battle is known only at the end of the battle.
-Thus, you will have to create and solve subgoals with existential variables.
-For that purpose, the tactic \texttt{eexists}, applied to the
-goal \Verb@{h':Hydra | h -*-> h'}@ generates the subgoal \Verb|h -*-> ?h'|.
-\item You may use Gérard Huet's \emph{zipper} data structure~\cite{zipper} for writing tactics associated with Hercules's  interactive search for a head to chop off.
-\end{itemize}
-
-\end{exercise}
-
-\subsection{Classes of battles}
-\label{sect:battle-classes}
-
-In some presentations of hydra battles, e.g.~\cite{KP82, bauer2008}, the transformation associated with the $i$-th round may depend on $i$.
-For instance, in these articles, the replication factor at the $i$-th round is equal to $i$.
-In other examples, one can allow the hydra to apply any replication factor at any time.
-In order to be the most general as possible, we define the type of predicates which relate the state of the hydra before and after the $i$-th round of a battle.
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-\label{types:Battle}
-\index{hydras}{Library Hydra!Type classes!Battle}
-
-\input{movies/snippets/Hydra_Definitions/BattleDef}
-
-The most general class of battles is \texttt{free}, which allows the hydra to chose any replication factor at every step:
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#free}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/freeDef}
-
-We chose to call \emph{standard}\footnote{This appellation is ours.
-If there is a better one, we will change it.} the kind of battles which appear  most often in the literature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#standard}{Hydra.Hydra\_Definitions}}
-\input{movies/snippets/Hydra_Definitions/standardDef}
-
-\subsection{Big steps}
-
-Let $B$ be any instance of class \texttt{Battle}. It is easy to define inductively the relation between the $i$-th and the $j$-th steps of a battle of type $B$.
-
-\vspace{6pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/battleRelDef}
-
-The following property allows us to build battles by composition of smaller ones.
-
-%% TODO display subgoals when fixed
-
-\vspace{6pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-\noindent
-\input{movies/snippets/Hydra_Lemmas/battleTrans}
-
-% \begin{remark}
-%  The class \texttt{free} is strongly related with the transitive closure  \texttt{round\_plus}, as expressed by the following lemmas.
-
-% \vspace{4pt}
-% \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-
-%  \begin{Coqsrc}
-%  Lemma battle_free_equiv1 : forall i j h h',  
-%              battle free i h j h' ->   h -+-> h'.
-
-%  Lemma battle_free_equiv2 : forall h h',
-%      h -+-> h' ->
-%     forall i, exists j,  battle free i h j h'.
-%  \end{Coqsrc}
-
-% \end{remark}
-
-\section{A long battle}
-\label{sect:big-battle}
-
-In this section we consider a simple example of battle, starting with a small hydra,
-shown on figure~\vref{fig:hinit}, with a simple strategy for both players:
-
-\begin{itemize}
-\item At each round, Hercules chops off the rightmost head of the hydra.
-\item The battle is {standard}: at the round number $i$, the expected replication is $i$.
-\end{itemize}
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (6,0) {$\bullet$};
-\node (n1) at  (3,3) {$\bullet$};
-\node (h1) at  (1,6) {$\Smiley[1][green]$};
-\node (h2) at  (3,6) {$\Smiley[1][green]$};
-\node (h3) at  (6,6) {$\Smiley[1][green]$};
-\node (h4) at  (6,6) {$\Smiley[1][green]$};
-\node (h5) at  (6,3) {$\Smiley[1][green]$};
-\node (h6) at  (9,3) {$\Smiley[1][green]$};
-\draw (foot) -- (n1);
-\draw (n1) to   [bend left=20] (h1);
-\draw (n1) to   (h2);
-\draw (n1) to   [bend right=20] (h3);
-\draw (foot) -- (h5);
-\draw (foot) to  [bend right=20] (h6);
-\end{tikzpicture}
-
-  \caption{The hydra \texttt{hinit}}
-  \label{fig:hinit}
-\end{figure}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.BigBattle.html}{Hydra.BigBattle}}
-
-\input{movies/snippets/BigBattle/hinitDef}
-
-The lemma we would like to prove is ``The considered battle lasts exactly $N$ rounds'',
-with $N$ being a natural number we gave to guess.
-
-But the  battle is so long that no \emph{test} can give us any estimation of its length, and we do need the expressive power of logic to compute this length.
-However, in order to  guess this length, we made some experiments, computing with \gallina{}, \coq{}'s  functional programming language.
-Thus, we can consider this development as a collaboration of proof with computation.
-In the following lines, we show how we found experimentally the value of the number $N$.
-The complete proof is in file \url{../theories/html/hydras.Hydra.BigBattle.html}.
-
-\subsection{First rounds}
-During the two first rounds, our hydra loses its two rightmost heads.  Figure~\vref{fig:hinit-plus2} shows the state of the hydra   just before the third round.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (3,0) {$\bullet$};
-\node (n1) at  (3,3) {$\bullet$};
-\node (h1) at  (1,6) {$\Smiley[1][green]$};
-\node (h2) at  (3,6) {$\Smiley[1][green]$};
-\node (h3) at  (6,6) {$\Smiley[1][green]$};
-\node (h4) at  (6,6) {$\Smiley[1][green]$};
-\draw (foot) -- (n1);
-\draw (n1) to   [bend left=20] (h1);
-\draw (n1) to   (h2);
-\draw (n1) to   [bend right=20] (h3);
-\end{tikzpicture}
-
-  \caption{The hydra (\texttt{hyd1 h3})}
-  \label{fig:hinit-plus2}
-\end{figure}
-
-The following lemma  is a formal description of these first rounds, in terms of the
-\texttt{battle} predicate.
-
-\input{movies/snippets/BigBattle/L02}
-
-\subsection{Looking for regularities}
-
-A first study with pencil and paper suggested us that, after three rounds, the hydra always looks like in figure~\vref{fig:hinit-plusn} (with a variable number of 
-subtrees of height 1 or 0).
-Thus, we introduce a few handy abbreviations.
-
-\input{movies/snippets/BigBattle/Notations}
-
-For instance, the hydra shown in  Fig~\ref{fig:hinit-plusn} is  (\texttt{hyd 3 4 2}). The hydra (\texttt{hyd 0 0 0})  is the ``final'' hydra of any terminating battle, {i.e.},
-a tree whith exactly one node and no edge.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (15,0) {$\bullet$};
-\node (a) at  (3,4) {$\bullet$};
-\node (b) at  (6,4) {$\bullet$};
-\node (c) at  (9,4) {$\bullet$};
-\node (d) at  (13,4) {$\bullet$};
-\node (e) at  (16,4) {$\bullet$};
-\node (f) at  (19,4) {$\bullet$};
-\node (g) at  (22,4) {$\bullet$};
-\node (h) at  (25,4) {$\Smiley[1][green]$};
-\node (i) at  (28,4) {$\Smiley[1][green]$};
-\node (aa) at  (2.5,8) {$\Smiley[1][green]$};
-\node (ab) at  (3.5,8) {$\Smiley[1][green]$};
-\node (ba) at  (5.5,8) {$\Smiley[1][green]$};
-\node (bb) at  (6.5,8) {$\Smiley[1][green]$};
-\node (ca) at  (8.5,8) {$\Smiley[1][green]$};
-\node (cb) at  (9.5,8) {$\Smiley[1][green]$};
-\node (da) at  (13,8) {$\Smiley[1][green]$};
-\node (ea) at  (16,8) {$\Smiley[1][green]$};
-\node (fa) at  (19,8) {$\Smiley[1][green]$};
-\node (ga) at  (22,8) {$\Smiley[1][green]$};
-\draw (foot) -- (a);
-\draw (foot) -- (b);
-\draw (foot) -- (c);
-\draw (foot) -- (d);
-\draw (foot) -- (e);
-\draw (foot) -- (f);
-\draw (foot) -- (g);
-\draw (foot) -- (h);
-\draw (foot) -- (i);
-\draw (a) -- (aa);
-\draw (a) -- (ab);
-\draw (b) -- (ba);
-\draw (b) -- (bb);
-\draw (c) -- (ca);
-\draw (c) -- (cb);
-\draw (d) -- (da);
-\draw (e) -- (ea);
-\draw (f) -- (fa);
-\draw (g) -- (ga);
-\end{tikzpicture}
-
-  \caption{The hydra (\texttt{hyd 3 4 2})}
-  \label{fig:hinit-plusn}
-\end{figure}
-
-With these notations, we get a formal description of the first three rounds.
-
-\input{movies/snippets/BigBattle/L23L03}
-
-\subsection{Testing  \dots}
-\label{sect:testing}
-In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple data type for representing the states as tuples composed of
-the round number, and the respective number of daughters  \texttt{h2}, \texttt{h1}, and heads
-of the current hydra.
-
-\input{movies/snippets/BigBattle/stateDef}
-
-The following function returns the next configuration of the game.
-Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle only start with the lemma
-\texttt{step\_battle}, page~\pageref{lemma:step-battle}.
-
-\input{movies/snippets/BigBattle/nextDef}
-
-We can make bigger steps through iterations of \texttt{next}.
-The functional \texttt{iterate}, similar to Standard Library's \texttt{Nat.iter},
-is defined and studied in~\href{../theories/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
-\index{hydras}{Library Prelude!iterate}
-
-\label{Functions:iterate}
-
-\input{movies/snippets/Iterates/iterateDef}
-
-The following function computes the state of the battle at the $n$-th round.
-
-\input{movies/snippets/BigBattle/testDefTests}
-
-The battle we are studying looks to be awfully long. Let us concentrate our
-tests on some particular events : the states where $\texttt{nh}=0$.
-From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} is equal to zero.
-
-\input{movies/snippets/BigBattle/smartTest}
-
-Thus, $(1 + 11)$ rounds later, the \texttt{n1} field is equal to $2$, and 
-\texttt{nh}   to $0$.
-
-\input{movies/snippets/BigBattle/smartTestb}
-
-Next round, we decrement \texttt{n2} and set \texttt{n1} to $95$.
-
-\input{movies/snippets/BigBattle/smartTestc}
-
-We now have some intuition of the sequence.
-It looks like the next ``\texttt{nh}=0'' event will happen at the $192=2(95+1)$-th round, then at the $2(192+1)$-th round, etc.
-
-\input{movies/snippets/BigBattle/doubleS}
-
-\subsection{Proving \dots}
-We are now able to reason about the sequence of transitions defined by our hydra battle. Instead of using the data-type \texttt{state} we study the relationship
-between different configurations of the battle.
-
-Let us define a binary relation associated with every round of the battle.
-In the following definition \texttt{i} is associated with the round number (or date, if we consider a discrete time), and \texttt{a}, \texttt{b}, \texttt{c} respectively associated with the number of occurrences of \texttt{h2}, \texttt{h1} and heads connected to the hydra's foot.
-
-\input{movies/snippets/BigBattle/oneStep}
-
-The relation between \texttt{one\_step} and the rules of hydra battles is asserted by the following lemma.
-
-\label{lemma:step-battle}
-
-\input{movies/snippets/BigBattle/stepBattle}
-
-\vspace{4pt}
-
-Next, we define ``big steps'' as the transitive closure of \texttt{one\_step},
-and reachability (from the initial configuration of figure~\ref{fig:hinit} at time $0$).
-
-\input{movies/snippets/BigBattle/steps}
-
-The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{battle}.
-
-\input{movies/snippets/BigBattle/stepsBattle}
-
-\vspace{4pt}
-
-Thus, any result about \texttt{steps} will be applicable to standard battles.
-Using the predicate \texttt{steps},  our study of the length of the considered battle
-can  be decomposed into three parts:
-
-\begin{enumerate}
-\item  Characterization and proofs of regularities of some events (inspired by our experiments of Sect.~\ref{sect:testing}).
-\item Study of the beginning of the battle
-\item Computing the exact length of the battle.
-\end{enumerate}
-
-First, we prove that, if at round $i$ the hydra is equal to
-(\texttt{hyd a (S b) 0}), then it will be equal to (\texttt{hyd a b 0}) at the $2(i+1)$-th round.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/LS}
-
-From now on, the lemma \texttt{reachable\_S} allows us to watch larger and larger steps of 
-the battle.
-
-\input{movies/snippets/BigBattle/L4}
-
-\input{movies/snippets/BigBattle/L10To95}
-
-\subsection{Giant steps}
-
-We are now able to make bigger steps in the simulation of the battle.
-First, we iterate the lemma \texttt{reachable\_S}.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/Bigstep}
-
-\vspace{4pt}
-
-Applying lemmas \texttt{BigStep} and \texttt{L95} we make a first jump.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/MDef}
-
-Figure~\ref{fig:HM}  represents the hydra at the $M$-th round.
-At the $(M+1)$-th round, it will look like in fig~\ref{fig:HM-plus1}.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:HM}}
-The state of the hydra after $M$ rounds.
-% The hydra \texttt{h} of the proof that \(\omega^2\) is too small for proving Hercules' victory
-
-\end{figure}
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (0,5) {$\bullet$};
-\node (N12) at (0,8) {$\Smiley[2][green]$};
-\node (N2) at (2,5) {$\bullet$};
-\node (N22) at (2,8) {$\Smiley[2][green]$};
-\node (N3) at (4,5) {$\bullet$};
-\node (N32) at (4,8) {$\Smiley[2][green]$};
-\node (N4) at (6,5) {$\bullet$};
-\node (N42) at (6,8) {$\Smiley[2][green]$};
-
-\node (Ndots) at (12,8) {\Huge $\dots$};
-\node (Ndots2) at (12,5) {\Huge $\dots$};
-
-\node (N8) at (18,5) {$\bullet$};
-\node (N82) at (18,8) {$\Smiley[2][green]$};
-\node (N9) at (20,5) {$\bullet$};
-\node (N92) at (20,8) {$\Smiley[2][green]$};
-
-\draw (foot) -- (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N8);
-\draw (foot) -- (N9);
-\draw (N1) to  (N12);
-\draw (N2) to  (N22);
-\draw (N3) to  (N32);
-\draw (N4) to  (N42);
-\draw (N8) to  (N82);
-\draw (N9) to  (N92);
-\end{tikzpicture}
-\caption{\label{fig:HM-plus1}}
-The state of the hydra after $M+1$ rounds (with $M+1$ heads).
-
-\end{figure}
-
-\input{movies/snippets/BigBattle/L295S}
-
-\vspace{4pt}
-
-Then, applying once more the lemma \texttt{BigStep}, we get the exact time when
-Hercules wins!
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/NDef}
-
-\vspace{4pt}
-
-We are now able to prove formally that the considered battle is 
-composed of $N$ steps.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/Done}
-
-\subsection{A minoration lemma}
-
-Now, we would like to get an intuition of  how big the number $N$ is.
-For that purpose, we use a minoration of the function \texttt{doubleS} by the
-function (\texttt{fun n => 2 * n}).
-
-\vspace{4pt}
-
-\input{movies/snippets/Exp2/exp2Def}
-
-\vspace{4pt}
-
-Using a few facts (proven in 
-\href{../theories/html/hydras.Hydra.BigBattle.html}{hydras.Hydra.BigBattle}),we get several  minorations.
-
-\input{movies/snippets/BigBattle/minorationLemmas}
-
-\vspace{4pt}
-
-The number $N$ is greater than or  equal to $2^{2^{95}\times 95}.$ If we wrote $N$ in base $10$, $N$ would require at least $10^{30}$ digits!
-
-\section{Generic properties}
-
-The example we just studied shows that the termination of any battle may take a very long time. If we want to study hydra battles in general, we have to consider 
-any hydra and any strategy, both for Hercules and the hydra itself. So, we first  give some definitions, generally borrowed from transition systems vocabulary (see~\cite{tel_2000} for instance).
-
-\subsection{Liveliness}
-
-Let $B$ be an instance of \texttt{Battle}. We say that $B$ is \emph{alive} if
-for any configuration $(i,h)$, where $h$ is not a head, there exists a further step in class $B$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/AliveDef}
-
-The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
-\href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
-
-\input{movies/snippets/Hydra_Theorems/AliveThms}
-
-Both theorems are proved with the help of the  following strongly specified function:
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
-
-\input{movies/snippets/Hydra_Lemmas/nextRoundDec}
-
-\subsection{Termination}
-
-The termination of \emph{any}  battle is naturally expressed by the predicate \texttt{well\_founded} defined in the module \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}
- of the Standard Library.
-
-\index{hydras}{Library Hydra!Predicates!Termination}
-
-\input{movies/snippets/Hydra_Definitions/TerminationDef}
-
-Let $B$ be an instance of class \texttt{Battle}. A \emph{variant} for $B$ consists
-in a well-founded relation $<$  on some type \texttt{A}, and a function
-(also called a \emph{measure}) $m:\texttt{Hydra}\rightarrow A$ such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $B$, the inequality $m(h')<m(h)$ holds.
-
-\vspace{4pt}
-\noindent
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
-
-\label{sect:hvariant-def}
-
-\index{hydras}{Library Hydra!Type classes!Hvariant}
-
-\input{movies/snippets/Hydra_Definitions/HvariantDef}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
- Prove that, if there exists some  instance of (\texttt{Hvariant Lt wf\_Lt $B$ $m$}), then there exists no infinite battle in  $B$.
-\end{exercise}
-
-\subsection{A  small proof of impossibility}
-%\index{coq}{Proofs of impossibility}
-
-\label{omega-case}
-
-When one wants to prove a termination theorem with the help of a variant, 
-one has to consider first a well-founded set $(A,<)$, then a strictly decreasing measure on this set.
-The following two lemmas show that if  the order structure $(A,<)$ is too simple, it is useless to look for a convenient measure, which simply no exists.
-Such kind of result is useful, because it saves you time and effort.
-
-The best known well-founded order is the natural order on the set $\mathbb{N}$ of natural numbers (the type \texttt{nat} of Standard library).
-It would be interesting to look for some measure $m:\texttt{nat}\arrow\texttt{nat}$ and prove it is a variant.
-
-Unfortunately, we can prove that 
-\emph{no} instance of class (\texttt{WfVariant round Peano.lt $m$}) can be built, where
-$m$ is \emph{any} function of type \texttt{Hydra $\arrow$ nat}.
-
-Let us present the main steps of that proof, the script of which  is in the module ~\href{../theories/html/hydras.Hydra.Omega_Small.html}{Hydra/Omega\_Small.v} \footnote{ The name of this file means ``the ordinal $\omega$ is too small for proving the termination of [free] hydra battles ''.
-In effect, the elements of $\omega$, considered as a set, are just the natural numbers (see next chapter for more details)}.
-
-%\subsubsection{Preliminaries}
-
-Let us assume there exists some variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
-    the  termination of all hydra battles.
-
-\input{movies/snippets/Omega_Small/omegaSmalla}
-
-We define an injection $\iota$ from the type \texttt{nat} into \texttt{Hydra}.
-For any natural number $i$, $\iota(i)$ is the hydra composed of a foot and
-$i+1$ heads at height $1$. For instance, Fig.~\ref{fig:flower} represents the hydra $\iota(3)$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (4,0) {$\bullet$};
-\node (N1) at (2,2) {$\Smiley[2][green]$};
-\node (N2) at (4,2) {$\Smiley[2][green]$};
-\node (N3) at (6,2) {$\Smiley[2][green]$};
-\node (N4) at (8,2) {$\Smiley[2][green]$};
-\draw (foot) to [bend left =25] (N1);
-\draw (foot) to [bend left =15] (N2);
-\draw (foot) to [bend right =15] (N3);
-\draw (foot) to [bend right =25] (N4);
-\end{tikzpicture}
-\caption{\label{fig:flower}
-The hydra $\iota(3)$}
-\end{figure}
-
-\input{movies/snippets/Omega_Small/iotaDef}
-
-Let us consider now some hydra \texttt{big\_h} out of the range of the injection $\iota$ (see Fig.~\vref{fig:h-omega-omega}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to  (N2);
-\end{tikzpicture}
-\caption{\label{fig:h-omega-omega}}
- The hydra \texttt{big\_h}.
-\end{figure}
-
-\input{movies/snippets/Omega_Small/bigHDef}
-
- Using the functions $m$ and $\iota$, we define a second hydra \texttt{small\_h}, and show
- there is a one-round battle that transforms \texttt{big\_h} into \texttt{small\_h}. Please note that,
-due to the hypothesis \texttt{Hvar}, we are interested in the termination of \emph{free} battles. 
-There is no problem to consider a round with (\texttt{m big\_h}) as the replication factor.
-
-\input{movies/snippets/Omega_Small/smallHDef}
-
-But, by hypothesis, $m$ is a variant. Hence, we infer the following inequality.
-
-\vspace{4pt}
-
-\input{movies/snippets/Omega_Small/mLt}
-
-In order to get a contradiction, it suffices to  prove the inequality
-$m(\texttt{big\_h}) \leq m(\texttt{small\_h})$ i.e.,  $m(\texttt{big\_h}\leq m(\iota (m(\texttt{big\_h})))$.
-
-\input{movies/snippets/Omega_Small/mGea}
-
-Intuitively, it means that, from any hydra of the form (\texttt{iota $i$}), the battle will 
-take (at least) $i$ rounds. Thus the associated measure cannot be less than $i$.
-Technically, we prove this lemma by Peano induction on $i$.
-
-\begin{itemize}
-\item The base case $i=0$ is trivial
-\item Otherwise, let $i$ be any natural number and assume  the inequality
-  $i \leq m(\iota(i))$.
-  \begin{enumerate}
-  \item  But the hydra $\iota(S(i))$ can be transformed in one round into
-    $\iota(i)$ (by losing its rightmost head, for instance)
-  \item Since $m$ is a variant, we have $m(\iota(i)) < m(\iota(S(i)))$,
-    hence  $i< m(\iota(S(i)))$, which implies  $S(i)\leq  m(\iota(S(i)))$.
-  \end{enumerate}
-\end{itemize}
-
-We are now ready to complete our impossibility proof.
-
-\vspace{4pt}
-
-\inputsnippets{Omega_Small/mGeb, Omega_Small/mGez,
-  Omega_Small/omegaSmallz}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\subsubsection{Conclusion}
-
-In order to build a variant for proving the termination of all hydra battles, we need to consider order structures more complex than the usual order on type \texttt{nat}. 
-The notion of \emph{ordinal number} provides a catalogue of well-founded order types.
-For a reasonably large bunch of ordinal numbers, \emph{ordinal notations} are data-types which allow the \coq{} user to define functions, to compute and prove some properties, for instance by reflection.
-
-The next chapter is dedicated to a generic formalization of ordinal notations, and chapter~\ref{chap:T1} to a proof of termination of all hydra battles with the help of an ordinal notation for the interval $[0,\epsilon_0)$\,\footnote{We use the mathematical notation $[a,b)$ for the interval $\{x|a\leq x < b\}$.}.
-\index{maths}{Notations!Interval}
-
-%--------------------------------------------------------------
-
-\chapter{Introduction to ordinal numbers and ordinal notations}
-\label{chap:ON}
-
-The proof of termination of all hydra battles presented in~\cite{KP82} is based
-on \emph{ordinal numbers}.
-From a mathematical point of view, an ordinal is a representative of an equivalence class for isomorphisms of  totally ordered well-founded sets.
-
-For the computer scientist, ordinals are tools for proving the totality of a given recursive function, or termination of a transition system. \emph{Ordinal arithmetic} 
-provides a set of functions whose properties, like \emph{monotony}, allow to define \emph{variants}, \emph{i.e.} strictly decreasing measures used in proofs of termination.
-
-\vspace{4pt}
-
-Let us have a look at Figure~\ref{fig:ordinal-sequence}. It presents a few items of a  sequence of ordinal numbers, which extends the sequence of natural numbers.
-
-\begin{figure}[h]
-  \centering
-\fbox{\Large
-  \begin{minipage}{1.0\linewidth}
-  \begin{align*}
-     &\textcolor{blue}{0},\,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,\ldots\\
-&\textcolor{red}{\omega},\,\omega+1,\omega+2,\omega+3,\ldots\\
-&\textcolor{red}{\omega\times 2},\,\omega\times 2+1,\ldots, \textcolor{red}{\omega\times 3},\,\omega\times 3+1,\ldots, \textcolor{red}{\omega\times 4},\ldots\\
-&\textcolor{red}{\omega^2},\ldots, \textcolor{red}{\omega^2\times 42},\ldots,\textcolor{red}{\omega^3},\ldots, \textcolor{red}{\omega^4},\omega^4+1,\ldots\\
-&\textcolor{red}{\omega^\omega},\ldots, \textcolor{red}{\omega^\omega+\omega^7\times 8},\ldots,\textcolor{red}{\omega^\omega\times 2},\omega^\omega\times 2+1, \ldots\\
-&\textcolor{red}{\omega^{\omega^\omega}},\ldots, \textcolor{red}{\omega^{\omega^\omega}+\omega^\omega\times 42+ \omega^{55}+\omega}, \ldots, \textcolor{red}{\omega^{\omega^{\omega+1}}}, \omega^{\omega^{\omega+1}}+1,\dots\\
-& \textcolor{red}{\epsilon_0 (= \omega^{\epsilon_0)}}, \epsilon_0+1, \epsilon_0+2, \epsilon_0+3, \ldots\\
-& \textcolor{red}{\epsilon_1}, \ldots, \textcolor{red}{\epsilon_2}, \ldots, \textcolor{red}{\epsilon_\omega},\ldots \\
-& \textcolor{red}{\Gamma_0}, \Gamma_0+1, \Gamma_0+2, \Gamma_0+3,\ldots, \textcolor{red}{\Gamma_0+\omega}, \ldots\\
-&\ldots
-  \end{align*}   
-  \end{minipage}}
-
-  \caption{A short overview of the sequence of ordinal numbers}
-  \label{fig:ordinal-sequence}
-\end{figure}
-
-Let us comment some features of this figure:
-
-\begin{itemize}
-\item The ordinals are listed in a strictly increasing order. 
-\item Dots : ``$\ldots$'' stand for  infinite sequences of ordinals, not shown for lack of space.
-For instance, the ordinal $42$ is not shown in the first line, but it exists, somewhere between $17$ and $\omega$.
-\item Each ordinal printed in black is the immediate successor of another ordinal.
-We call it a \emph{successor} ordinal.
-For instance, $12$ is the successor of $11$, and $\omega^4+1$ the successor of $\omega^4$.
-\item Ordinals (displayed in red)  that  follow immediately dots are called \emph{limit ordinals}.
-With respect to the order induced by this sequence, any limit ordinal $\alpha$ is the least upper bound of  the set $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
-\item
-For instance $\omega$ is the least upper bound of the set of all finite ordinals (in the first line). It is also the first limit ordinal, and the first infinite ordinal, in the sense that 
-the set $\mathbb{O}_\omega$ is infinite.
-\item The ordinal $\epsilon_0$ is the first number which is equal to its own exponential of base $\omega$.
-It plays an important role in proof theory, and is particularly studied in chapters~\ref{chap:T1} to \ref{chap:alpha-large}.
-\item Any ordinal is  either the ordinal \textcolor{blue}{$0$},
-a successor ordinal, or a \textcolor{red}{limit ordinal}.
-\end{itemize}
-
-\section{The mathematical point of view}
-
-\subsection{Well-ordered sets}
-Let us start with some definitions.
-A  \emph{well-ordered set} is a set provided with a binary relation $<$ which has the following properties.
-\begin{description}
-\item[irreflexivity] : $\forall x\in A, x\not< x$
-\item[transitivity] : $\forall x\,y\,z\in A, x<y \Rightarrow y<z \Rightarrow x<z$
-\item[trichotomy]: $\forall x\,y\in A, x<y \vee x = y \vee y < x$
-\item[well foundedness]: $<$ is well-founded (every element of $A$ is accessible)\footnote{In classical mathematics, we would say that there is no infinite sequence $a_1>a_2> \dots a_n> a_{n+1}\dots$ in $A$.
-In contrast, \coq's standard library contains
-an inductive definition of a predicate \texttt{Acc} which allows us to write 
-constructive proofs of accessibility (See \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}).}.
-\end{description}
-
-The best known examples of well-ordered sets are the set $\mathbb{N}$ of natural numbers (with the usual order $<$), as well as any finite segment $[0,i)=\{j\in\mathbb{N}\,|\,j<i\}$.
-The disjoint union of two copies of $\mathbb{N}$, \emph{i.e.} the set $\{0,1\}\times\mathbb{N}$ is also well-ordered,
-with respect to the order below:
-
-\begin{align*}
-(i,j) < (i,k) & \;\textit{\textbf{if} }\; j < k\\
-(0,k) < (1,l) & \;\textit{\textbf{for\,any}}\;k \;\textit{\textbf{and}} \; l
-\end{align*}
-
-\subsection{Ordinal numbers}
-
-\index{maths}{Ordinal numbers}
-
-Let $(A,<_A)$ and $(B,<_B)$ two well-ordered sets. $A$ and $B$ are said to have \emph{the same order type} if 
-there exists a strictly monotonous bijection $b$ from $A$ to $B$, \emph{i.e.} which verifies the proposition
-$\forall x\,y\in A,\, x <_A y \Rightarrow b(x) <_B  b(y)$.
-
-Having the same order type is an equivalence relation between well-ordered sets. Ordinal numbers (in short: \emph{ordinals})  are descriptions (\emph{names}) of the equivalence classes.
-For instance, the order type of $(\mathbb{N},<)$ is associated with the ordinal called  $\omega$, and the order we considered on 
-the disjoint union of $\mathbb{N}$ and itself is named $\omega+\omega$.
-
-In a set-theoretic framework, one can consider any ordinal $\alpha$ as a well-ordered set, whose  elements are just the ordinals strictly less than $\alpha$, \emph{i.e.} the \emph{segment} $\mathbb{O}_\alpha=[0, \alpha)$.
-So, one can speak about \emph{finite}, \emph{infinite}, \emph{countable}, etc., ordinals.
-Nevertheless, since we work within type theory,
-we do not identify ordinals as sets of ordinals, but consider the correspondence between ordinals and sets of ordinals as the function that maps $\alpha$ to $\mathbb{O}_\alpha$.
-For instance $\mathbb{O}_\omega=\mathbb{N}$, and $\mathbb{O}_7=\{0,1,2,3,4,5,6\}$.
-
-We cannot cite all the literature published on ordinals since Cantor's book
-\cite{cantorbook}, and 
-leave it to the reader to explore the bibliography.
-
-The introduction of Jos\'e Grimm's report~\cite{grimm:hal-00911710} contains a nice presentation of the main properties of ordinals.
-
-\section{Ordinal numbers in Coq}
-
-Two kinds of representation of ordinals are defined in our development.
-
-\begin{itemize}
-\item A ``mathematical'' representation of the set of countable ordinal numbers, after Kurt Schütte~\cite{schutte}.
-This representation uses several (hopefully harmless) axioms.
-We use it as a reference for proving the correctness of ordinal notations.
-  We are also progressively importing \gaia's~\cite{Gaia} definitions and theorems, based on Bourbaki's set theory~\cite{Grimm2016}.
-Chapter~\ref{gaia-chapter} is dedicated to the connexion between libraries \HydrasLib and \gaia.
-
-\item A family of \emph{ordinal notations} (also called \emph{[ordinal] notation systems}), \emph{i.e.} data types used to represent segments $[0,\mu)$, where $\mu$ is some countable ordinal.
-Each ordinal notation is defined inside the Calculus of Inductive Constructions (without axioms).
-Many functions are defined, allowing proofs by computation.
-Note that proofs of
-correctness of a given ordinal notation with respect to Schütte's model obviously use axioms.
-Please execute the \texttt{Print Assumptions} command in case of doubt.
-\end{itemize}
-
-% \section{Countable ordinals}
-
-% Chapter~\ref{chap:schutte} of this document presents an adaptation to \coq{} of an axiomatization in classical logic of the set of countable ordinals by K. Schütte~\cite{schutte}. 
-% That formalization is quite complex, technical and unshamedly non-constructive,  so we put its description  in the last chapter of this document.
-
-% Please note that Schütte considers the (uncountable) set $\mathbb{O}$ of all countable ordinals.
-This set is well ordered (which is one of Schütte's axioms), and associates to any ordinal $\alpha$ the segment $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
-
-% In our adaptation to \coq{}, we declare a type \texttt{Ord}, a binary relation \texttt{lt} (with infix notation \texttt{"\_<\_"}, and assume Schütte's axiom. In Chapter~\ref{chap:schutte},
-% we derive some interesting properties of countable ordinals from these axioms.
-
-It is interesting to compare proofs of a given property (for instance the associativity of addition) both in the computational framework of some ordinal notation, and in mathematical  models
-of Schütte or \gaia.
-
-\section{Ordinal Notations}
-
-Fortunately, the ordinals we need for  studying hydra battles are much simpler than Schütte's, and can be represented as quite simple data types in \gallina.
-
-Let $\alpha$ be some (countable) ordinal; 
-in \coq{} terms, we call \emph{ordinal notation for $\alpha$} a structure composed 
-of:
-\begin{itemize}
-\item A data type $A$ for representing all ordinals strictly below $\alpha$,
-\item A well founded order $<$ on $A$, 
-\item A correct function for comparing two ordinals. Note  that the reflexive closure of $<$ is thus a \emph{total order}.
-\end{itemize}
-
-Such a structure may be proved correct relatively to a bigger ordinal notation or to Schütte's or \gaia's model.
-
-\subsection{Classes for ordinal notations}
-
-From the \coq{} user's point of view, an ordinal notation is
-a structure allowing to compare two ordinals by computation, and proving by well-founded induction.
-
-\subsubsection{The \texttt{Comparison} and \texttt{Comparable} classes}
-
-We use the operational class \texttt{Comparison} of comparison functions to define the \texttt{Comparable} class, contributed by Jérémy Damour and Théo Zimmermann, which allows us to apply generic lemmas and tactics about decidable strict orders.
-The correctness of the comparison function is expressed through Stdlib's type 
-\texttt{Datatypes.CompareSpec} as specialized by \texttt{Datatypes.CompSpec}.
-
-\begin{Coqsrc}
-  Inductive CompareSpec (Peq Plt Pgt : Prop) :
-  comparison -> Prop :=
-    CompEq : Peq -> CompareSpec Peq Plt Pgt Eq
-  | CompLt : Plt -> CompareSpec Peq Plt Pgt Lt
-  | CompGt : Pgt -> CompareSpec Peq Plt Pgt Gt.
-
-Definition CompSpec {A} (eq lt : A -> A -> Prop) (x y : A) :=
- CompareSpec (eq x y) (lt x y) (lt y x).
-\end{Coqsrc}
-
-\emph{From Module~\href{../theories/html/hydras.Prelude/mparable.html\#Hvariant}{Prelude.Comparable}}
-
-\label{sect:comparable-def}
-
-\input{movies/snippets/Comparable/ComparableDef}
-
-\subsubsection{The \texttt{ON} class}
-
-The following class definition, parameterized with a type $A$, a binary relation \texttt{lt} on $A$, specifies that \texttt{lt} is a well-founded strict order, provided with a correct comparison function.
-
-\vspace{4pt}
-\noindent\emph{From
-Library~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}}
-
-\label{types:ON}
-\index{hydras}{Library OrdinalNotations!Type classes!ON}
-
-\input{movies/snippets/ON_Generic/ONDef}
-
-We give  also a few handy definitions and lemmas for any ordinal notation.
-
-\label{sect:on-lt-notation}
-\label{sect:on-le-notation}
-\label{sect:measure-ON}
-\label{sect:bigO-ON}
-
-\inputsnippets{ON_Generic/ONDefsa}
-\inputsnippets{ON_Generic/ONDefsb}
-
-\begin{remark}
-The infix notations \texttt{o<} and \texttt{o<=} are defined in order to make apparent the distinction between the various notation scopes that may co-exist in a same statement.
-So the infix \texttt{<} and \texttt{<=} are reserved to the natural numbers.
-In the mathematical formulas, however, we still use $<$ and $\leq$ for comparing ordinals.
-\end{remark}
-
-% \subsection{Ordinal notations and  termination measures}
-% \label{sect:measure-ON}
-
-% The following lemma (together with the type class mechnism) allows us to define termination measures over any ordinal notation.
-It is just an application of  the libraries \texttt{Coq.Wellfounded.Inverse\_Image}
-% and  \texttt{Coq.Wellfounded.Inclusion}.
-
-% \begin{Coqsrc}
-% Definition measure_lt {A:Type}{lt: relation A}
-%             {compare : A -> A -> comparison}
-%             {on : ON lt compare}
-%             {B : Type} (m : B -> A) : relation B :=
-%              fun x y => on_lt (m x) (m y).
-
-% Lemma wf_measure  {A:Type}(lt: relation A)
-%             {compare : A -> A -> comparison}
-%             {on : ON lt compare}
-%             {B : Type}
-%             (m : B -> A):  well_founded (measure_lt m). 
-% \end{Coqsrc}
-
-% A simple example of application is given in Sect.~\vref{sect:merge-example}.
-
-\section{Example: the ordinal \texorpdfstring{$\omega$}{omega}}
-
-The simplest example of ordinal notation is built over the type \texttt{nat} of \coq's standard library. We just have to apply already proven lemmas about Peano numbers.
-
-\vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_Omega.html}{OrdinalNotations.ON\_Omega}}
-
-\inputsnippets{ON_Omega/OmegaDefa, ON_Omega/OmegaDefb}
-
-\section{Sum of  two ordinal notations}
-
-Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective types \texttt{A} and \texttt{B}.
-
- We consider a new strict order
-on the disjoint sum of the associated types, by putting all elements of \texttt{A} before the elements of \texttt{B} (thanks to Standard Library's relation operator \texttt{le\_AsB}).
-
-\vspace{4pt}
-\noindent
-\emph{From Library~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}}.
-
-\begin{Coqanswer}
-Inductive
-le_AsB (A B : Type) (leA : A -> A -> Prop) (leB : B -> B -> Prop)
-  : A + B -> A + B -> Prop :=
-| le_aa : forall x y : A, leA x y -> le_AsB A B leA leB (inl x) (inl y)
-| le_ab : forall (x : A) (y : B), le_AsB A B leA leB (inl x) (inr y)
-| le_bb : forall x y : B, leB x y -> le_AsB A B leA leB (inr x) (inr y)
-\end{Coqanswer}
-
-%\pagebreak
-
-\vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_plus.html}{OrdinalNotations.ON\_plus}}
-
-\input{movies/snippets/ON_plus/Defs}
-
-In order to build an instance of \texttt{Comparable}, we have to define a correct comparison function.
-
-\inputsnippets{ON_plus/compareDef, ON_plus/compareCorrect}
-\inputsnippets{ON_plus/plusComp}
-
-Standard library's lemma \texttt{Wellfounded.Disjoint\_Union.wf\_disjoint\_sum} 
-is applied  to prove that our order \texttt{lt} is well-founded, allowing us to build an instance of \texttt{ON}:
-
-\inputsnippets{ON_plus/ltWf, ON_plus/OnPlus}
-
-\subsection{The ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
-
-The ordinal $\omega+\omega$ (also known as $\omega\times 2$) may be represented as the concatenation 
-of two copies of $\omega$ (Figure~\ref{fig:omega-plus-omega}).
-It is also represented by the two first lines of Figure~\ref{fig:ordinal-sequence}.
-
-\begin{figure}[h]
-   \centering
-   \begin{tikzpicture}[very thick, scale=0.5]
-\begin{scope}[color=blue]
-\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
-\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
-\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
-\node (Adots) at (6,0) {$\ldots$};
-\node(An) at (8,0)[label=below:$n$]{$\bullet$};
-\node(A2) at (10,0)[label=below:$n+1$]{$\bullet$};
-\node (Adots1) at (12,0) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red]
-\node(B0) at (14,0)[label=below:$0$,label=above:\textcolor{red}{$\omega$}]{$\bullet$};
-\node(B1) at (16,0)[label=below:$1$, label=above:$\omega+1$]{$\bullet$};
-\node(B2) at (18,0)[label=below:$2$,label=above:$\omega+2$]{$\bullet$};
-\node (Bdots) at (20,0) {$\ldots$};
-\node (Bn) at (22,0) [label=below:$p$, label=above:$\omega+p$]{$\bullet$};
-\node (Bdots2) at (24,0) {$\ldots$};
-\end{scope}
-\end{tikzpicture}
-   \caption{\textcolor{blue}{$\omega+{\color{red}\omega}$}}
-   \label{fig:omega-plus-omega}
- \end{figure}
-
-We can define this notation in \coq{} as an instance of \texttt{ON\_plus}.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/OmegaPlusOmegaDef}
-
-\vspace{4pt}
-
-We can now define abbreviations.
-For instance, the finite ordinals are represented by terms built with  the constructor \texttt{inl}, and the first infinite ordinal $\omega$ by the term \texttt{(inr 0)}.
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Omega_plus_omega/finiteOmega}
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Omega_plus_omega/ltOmega}
-
-% \label{warning:coercions}
-% \index{Coq!Coercions} 
-% \begin{remark}
-% Beware of coercions and notation scopes!
-% Let us consider the following goal:
-
-% \begin{Coqsrc}
-%  Goal (6 o< 8).
-%  auto with arith.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% 1 subgoal (ID 9)
-
-%   ============================
-%   6 o< 8
-% \end{Coqanswer}
-
-% Please keep in mind that the current notation scope interprets the infix \texttt{``<''} as the predicate \texttt{Omega\_plus\_omega.lt} and not \texttt{Nat.lt}.
-More,  the coercion mechanism converts the terms \texttt{6:nat} [resp.
-\texttt{8:nat} ]
-% into \texttt{inl 6} [resp. \texttt{inl 8}].  So, the initial goal is correctly interpreted by \coq{}, but not as an inequality between two natural numbers.
-
-% \begin{Coqsrc}
-% Set Printing All.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% 1 subgoal (ID 337)
-
-%   ============================
-%   @on_lt nat Peano.lt Nat.compare Omega (S (S (S (S (S (S O))))))
-%     (S (S (S (S (S (S (S (S O))))))))
-% \end{Coqanswer}
-
-% Anyway, the initial goal is provable, using \texttt{le\_AsB}'s first constructor.
-
-% \begin{Coqsrc}
-%   constructor; auto with arith.
-% Qed.
-% \end{Coqsrc}
-
-% \end{remark}
-%
-
-\section{Limits and successors}
-
-Let us look again at our implementation of $\omega+\omega$. We can classify its elements into three categories:
-
-\begin{itemize}
-\item The least ordinal, \texttt{(inl 0)}, also known as  \texttt{(fin 0)}.
-\item The first infinite ordinal $\omega$.
-\item The remaining ordinals, either of the form \texttt{(inl (S $i$))} or \texttt{(inr (S $i$))} (in black on Figure~\ref{fig:ordinal-sequence}), called \emph{successor ordinals}.
-\end{itemize}
-
-\subsection{Definitions}
-It would be interesting to specify at the most generic level, what is a zero, a successor or a limit ordinal. Let $<$ be a strict order on a type $A$.
-
-\begin{itemize}
-\item A \emph{least} element is a minorant (in the large sense) of the full set  on $A$,
-\item $y$ is a \emph{successor} of $x$ if $x<y$ and there is no element between $x$ and $y$. We will also say that $x$ is a \emph{predecessor} of $y$.
-\item $x$ is a \emph{limit} if $x$ is not a least element, and for any $y$ such that $yo<x$,
- there exists some $z$ such that $y<z<x$.
-\end{itemize}
-
-The following definitions are in Library \href{../theories/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
-
-\input{movies/snippets/MoreOrders/Defs}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove, that, in any ordinal notation system, every ordinal has at most one predecessor, and at most one successor.
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/predSuccUnicity.v}{exercises/ordinals/predSuccUnicity.v}.}
-
-\end{exercise}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove, that, in any ordinal notation system, if $\beta$ is a successor of $\alpha$,
-then for any $\gamma$, $\gamma<\beta$ implies 
-$\gamma\leq\alpha$.
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/lt_succ_le.v}{exercises/ordinals/lt\_succ\_le.v}.}
-\end{exercise}
-
-\subsection{Limits and successors in \texorpdfstring{$\omega+\omega$}{omega+omega}}
-
-Using the definitions above, we can prove the following lemma:
-
-\vspace{4pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/limitIff}
-
-\vspace{4pt}
-
-Regarding successors, let us define the following function and prove its correctness:
-
-\input{movies/snippets/ON_Omega_plus_omega/succDef}
-
-\input{movies/snippets/ON_Omega_plus_omega/succCorrect}
-
-\vspace{4pt}
-
-We can also check whether an ordinal is a successor by a simple computation:
-
-\input{movies/snippets/ON_Omega_plus_omega/succb}
-
-\vspace{4pt}
-
-Finally, the nature of any ordinal is decidable (inside this notation system) :
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/ZeroLimitSuccDec}
-
-\section{Product of ordinal notations}
-
-Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective  ordered types \texttt{A} and \texttt{B}.
-The product of \texttt{NA} and \texttt{NB} is considered as the concatenation of $B$ copies of $A$, ordered by the lexicographic order on $B\times A$.
-
-In \coq{}, we build an instance of class \texttt{ON} through a sequence of steps as for the sum of ordinal notations.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_mult.html}{OrdinalNotations.ON\_mult}}
-
-\inputsnippets{ON_mult/Defs}
-\inputsnippets{ON_mult/multComp,  ON_mult/ONMult}
-\inputsnippets{ON_mult/endDefs}
-
-\section{The ordinal \texorpdfstring{$\omega^2$}{omega2}}
-
-The ordinal $\omega^2$ (also called $\phi_0(2)$, see Chap.~\ref{chap:schutte}), is an instance of the multiplication presented in the preceding section.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
-
-\inputsnippets{ON_Omega2/Omega2Def, ON_Omega2/Defs}
-\inputsnippets{ON_Omega2/Ex1}
-
-\subsection{Arithmetic of \texorpdfstring{$\omega^2$}{omega^2}}
-
-\subsubsection{Successor}
-
-The successor of any ordinal is defined by a simple pattern-matching.
-
-\input{movies/snippets/ON_Omega2/succ}
-
-This function is proved to be correct w.r.t. the \texttt{Successor} predicate.
-
-\input{movies/snippets/ON_Omega2/succOK}
-
-\input{movies/snippets/ON_Omega2/succLemmas}
-
-\subsubsection{Addition}
-
-We can define on \texttt{Omega2} an addition which extends the addition on \texttt{nat}. Please note that this operation is not commutative:
-
-\inputsnippets{ON_Omega2/plusDef, ON_Omega2/plusExamples}
-
-\subsubsection{Multiplication}
-
-The restriction of ordinal multiplication to the segment $[0,\omega^2)$ is not a total function.
-For instance $\omega\times\omega= \omega^2$ is outside the set of represented values.
-Nevertheless, we can define two operations mixing natural numbers and ordinals.
-
-\input{movies/snippets/ON_Omega2/multFinDef}
-\input{movies/snippets/ON_Omega2/multFinExamples}
-
-Multiplication with a finite ordinal and addition are related through the following lemma:
-
-\inputsnippets{ON_Omega2/uniqueDecompositiona,
-  ON_Omega2/uniqueDecompositionb,
-  ON_Omega2/uniqueDecompositionz}
-
-\subsection{A proof of termination using \texorpdfstring{$\omega^2$}{omega^2}} 
-\label{sect:merge-example}
-
-Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily a total function which merges two lists (example contributed by Pascal Manoury).
-
-%\index{coq}{Commands!Function}
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
-
-\inputsnippets{ON_Omega2/Merge, ON_Omega2/Mergeb}
-
-\subsection{Yet another  proof of impossibility}
-\label{omega2-case}
-
-In Sect.~\vref{omega-case}, we proved that there exists no variant from \texttt{Hydra} to \texttt{(nat,$<$)}
-(\emph{i.e.} the ordinal $\omega$) for proving the termination of all hydra battles.
-We  prove now that  the ordinal $\omega^2$ is also insufficient for this purpose.
-
-The proof we are going to comment has exactly the same structure as in Section~\ref{omega-case}.
- Nevertheless, the proof of technical  lemmas is a little more complex, due to 
- the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
-Consider for instance that there exists an infinite number of ordinals  between
-$\omega$ and $\omega\times 2$.
-
-The detailed  proof script is in the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/Hydra/Omega2_Small.v}{theories/ordinals/Hydra/Omega2\_Small.v}.
-
-\subsubsection{Preliminaries}
-Let us assume there is a variant from \texttt{Hydra} into $\omega^2$  for proving the   termination of all hydra battles.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
-
-\input{movies/snippets/Omega2_Small/Impossibility}
-\input{movies/snippets/Omega2_Small/Impossibilitya}
-
-We  follow the same pattern as in Sect.~\ref{omega-case}.
-First, we define an injection $\iota$ from type \texttt{ON\_Omega2.t} into \texttt{Hydra}, by
- associating to  each ordinal $\omega\times i+ j = (i,j)$ the hydra with $i$ branches of length $2$ and
-$j$ branches of length $1$.
-
-%% revenir ici
-
-\vspace{4pt}
-\emph{From Module ~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
-
-\input{movies/snippets/Omega2_Small/Impossibilityc}
-
-For instance, Figure~\vref{fig:essai2} shows the hydra associated to the ordinal 
-$(3,5)$, a.k.a. $\omega\times 3 + 5$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-\node (foot) at (6,0) {$\bullet$};
-\node (N1) at (1,3) {$\bullet$};
-\node (N2) at (3,3) {$\bullet$};
-\node (N3) at (5,3) {$\bullet$};
-\node (N4) at (8,3) {$\Smiley[2][green]$};
-\node (N5) at (11,3) {$\Smiley[2][green]$};
-\node (N6) at (14,3) {$\Smiley[2][green]$};
-\node (N7) at (17,3){$\Smiley[2][green]$};
-\node (N8) at (20,3){$\Smiley[2][green]$};
-\node  (N9) at (0,5) {$\Smiley[2][green]$};
-\node (N10) at (2,5) {$\Smiley[2][green]$};
-\node (N11) at (4,5) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N5);
-\draw (foot) -- (N6);
-\draw (foot) to [bend right=10] (N7);
-\draw (foot) to [bend right=15] (N8);
-\draw (N1) to [bend left=10] (N9);
-\draw (N2) -- (N10);
-\draw (N3) -- (N11);
-\end{tikzpicture}
-\caption{\label{fig:essai2}
-The hydra $\iota(\omega\times 3+5)$}
-\end{figure}
-
-Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:h-omega2-small}}
- The hydra \texttt{big\_h}.
-\end{figure}
-
-\input{movies/snippets/Omega2_Small/Impossibilityb}
-
- In a second step, we build a ``smaller'' hydra\footnote{With respect to the measure $m$.}.
-
-\input{movies/snippets/Omega2_Small/Impossibilityd}
-
-\vspace{4pt}
-
-Like in Sect.~\ref{omega-case}, we prove the inequality $m\;\texttt{big\_h} \,\texttt{o<=}\, m\;\texttt{small\_h} \,\texttt{o<=}\, m\;\texttt{big\_h}$, which is obviously false.
-
-\subsubsection{Proof of the inequality \texttt{m small\_h o< m big\_h}}
-
-In order to prove the inequality  \texttt{m\_lt: m small\_h o< m big\_h}, it suffices to
-build a battle transforming \texttt{big\_h} into \texttt{small\_h}.
-
-First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h} as $(i,j)$.
-If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
-If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the following result.
-
-\input{movies/snippets/Omega2_Small/bigToSmall}
-
-Since $m$ is a variant, we infer the following inequality:
-
-\input{movies/snippets/Omega2_Small/mLt}
-
-\subsubsection{Proof of the inequality \texttt{m big\_h o<= m small\_h} }
-
-The proof of the inequality \texttt{m big\_h o<= m small\_h} is quite more complex than in Sect~\ref{omega-case}.
-If we consider any ordinal $\alpha=(i,j)$, where $i>0$, there exists an infinite number of
-ordinals strictly less than $\alpha$, and there exists an infinite number of battles that start from
-$\iota(\alpha)$.
-Indeed, at any configuration $\iota(k,0)$, where $k>0$, the hydra can freely choose any replication number.
-Intuitively, the measure of such a hydra must be large enough for taking into account
-all the possible battles issued from that hydra.
-Let us now give more technical details.
-
-The first steps of our proof start a well-founded induction over $\omega^2$.
-
-\input{movies/snippets/Omega2_Small/mGe}
-
-A case analysis on $i$ and $j$ allows us to
-consider three cases :
-
-\begin{itemize}
-\item $i=j=0$: the inequality is trivial.
-\item $i=1+l, j=0$ ($(i,j)$ is a limit ordinal): By the induction hypothesis \texttt{IHij},
-  $(l,k)\texttt{o<=} m(\iota(l,k))$ for any $k$. But (by the rules of the hydra game), $\iota(i,0)$ is transformed into any $\iota(l,k)$ in one round. Thus $m(\iota(l,k)) < m(\iota(i,0))$ for any $k$.
-  Therefore, $(l,k) <  m(\iota(i,0))$ for any $k$, thus
-  $(i,0)\,\texttt{o<=}\,m(\iota(i,0))$.
- \item $j= l+1$  ($(i,j)$ is a successor). We apply the induction hypothesis to the pair $(i,l)$.
- \end{itemize}
-
- Please look at the proof script for more details.
-
- \input{movies/snippets/Omega2_Small/mGeb}
-
-\subsubsection{End of the proof}
-From \texttt{m\_ge}, we get \texttt{$m$\, big\_h o<= $m$\, small\_h = $m$\,(iota (m big\_h)) }. 
-Since $<$ is a strict order (irreflexive  and transitive), this inequality is incompatible with the strict inequality  \texttt{m small\_h o< m big\_h} (lemma \texttt{m\_lt}).
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#Impossible}{Hydra.Omega2\_Small}}
-
-  \input{movies/snippets/Omega2_Small/Impossible}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into $\omega^2$ for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\begin{remark}
- We  prove in Chapter~\ref{ks-chapter} a generalization of the impossibility lemmas of
-Sect.~\ref{omega-case} and this section, with the same proof structure, but with much more 
-complex technical details.
- \end{remark}
-
-% \index{Exercises}
-% \begin{exercise}
-
-% \label{sec:orgheadline63}
-% Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.
-
-% \begin{itemize}
-% \item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
-%   lexicographic ordering
-% \item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.
-
-% For instance, the following inequality holds:
-% \[\langle 4,3,3,3,3,3,3,2,2,2 \rangle\,<\,\langle 4,4,2 \rangle\]
-% \end{itemize}
-
-% \end{exercise}
-
-\section{A notation for finite ordinals}
-
-Let $n$ be some natural number. The segment associated with $n$ is the interval 
-$[0,n)\,=\,\{0,1,\dots,n-1\}$. 
-One may represent the ordinal $n$ by a sigma type.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
-
-\label{def:Finite-ord-type}
-
-\input{movies/snippets/ON_Finite/Defs}
-
-\subsubsection{Examples}
-
-For instance, let us build two elements of the segment $[0, 7)$, \emph{i.e.} two
-inhabitants of   type (\texttt{t 7}), and prove a simple  inequality (see Fig.~\ref{fig:O7}).
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (N0) at (0,0) {$\bullet$};
-\node (i0) at (0,1) {$0$};
-\node (N1) at (2,0) {$\bullet$};
-\node (i1) at (2,1) {$1$};
-\node (N2) at (4,0) {$\bullet$};
-\node (i2) at (4,1) {$2$};
-\node (N3) at (6,0) {$\bullet$};
-\node (i3) at (6,1) {$3$};
-\node (N4) at (8,0) {$\bullet$};
-\node (i4) at (8,1) {$4$};
-\node (N5) at (10,0) {$\bullet$};
-\node (i5) at (10,1) {$5$};
-\node (N6) at (12,0) {$\bullet$};
-\node (i6) at (12,1) {$6$};
-\node(alpha1) at (4,-1) {$\alpha_1$};
-\node(alpha2) at (10,-1) {$\beta_1$};
-\end{tikzpicture}
-
-\caption{The segment $\mathbb{O}_7$\label{fig:O7}}
-\end{figure}
-
-%\index{coq}{Commands!Program}
-
-\input{movies/snippets/ON_Finite/Example1}
-
-Note that the type (\texttt{t 0}) is empty, and that, for any natural number
- $n$, $n$ does not belong to (\texttt{t $n$}).
-
- \inputsnippets{ON_Finite/t0Empty, ON_Finite/bad10}
-
-\subsubsection{Comparison function}
-
-In order to build an instance of \texttt{ON}, we define a comparison function,  and prove its correctiness.
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Finite/compareDef}
-
-\begin{remark}
- The proof of \texttt{compare\_correct} uses a well-known pattern of \coq{}.
-Let us consider  the following subgoal.
-
-\input{movies/snippets/ON_Finite/compareCorrectb}
-
-Applying the tactic \texttt{f\_equal} generates a simpler subgoal.
-
-\input{movies/snippets/ON_Finite/compareCorrectc}
-
-We have now to prove that there exists at most one  proof of (\texttt{Nat.ltb x0 (S n)}). This is not obvious, but  a consequence of the following lemma of library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Eqdep_dec.html}{Coq.Logic.Eqdep\_dec}.
-
-\index{Coq}{Unicity of equality proofs}
-\label{sect:eq-proof-unicity}
-
-\begin{Coqanswer}
-eq_proofs_unicity_on :
-forall (A : Type) (x : A),
-(forall y : A, x = y \/ x <> y) -> 
-forall (y : A) (p1 p2 : x = y), p1 = p2
-\end{Coqanswer}
-
-Thus unicity of proofs of \texttt{Nat.ltb x0 (S n)}  comes from the decidability of
-equality on type \texttt{bool}.
-This is why we used the boolean function \texttt{Nat.ltb} instead of the inductive predicate \texttt{Nat.lt} in the definition of type \texttt{t $n$} (see page~\pageref{def: Finite-ord-type}).
-For more information about this pattern, please look at the numerous mailing lists and 
-FAQs on \coq{}).
-
-\inputsnippets{ON_Finite/compareCorrectd}
-
-\end{remark}
-
-\begin{remark}
-
-Please note  that attempting to compare a term  of type (\texttt{t $n$}) with a term of
-type (\texttt{t $p$})  leads to an error if $n$ and $p$ are not convertible.
-
-\input{movies/snippets/ON_Finite/Example2}
-
-\end{remark}
-
-\subsubsection{Building an instance of \texttt{ON}}
-
-Applying lemmas of the libraries \texttt{Coq.Wellfounded.Inverse\_Image}, \linebreak
- \texttt{Coq.Wellfounded.Inclusion}, and \texttt{Coq.Arith.Wf\_nat}, we prove that our
-relation \texttt{lt} is well founded.
-
-\input{movies/snippets/ON_Finite/ltWf}
-
-Now we can build our instance of \texttt{OrdinalNotation}.
-
-\input{movies/snippets/ON_Finite/ONInstance}
-
-\begin{remark}
-It is important to keep in mind  that the integer $n$ is not an ``element'' of \texttt{FinOrd $n$}.
-In set-theoretic presentations of ordinals, the set associated with the ordinal $n$ is $\{0,1,\dots,n-1\}$.
-In our formalization, the interpretation of an ordinal as a set is realized by the function \texttt{bigO} (see Section\vref{sect:bigO-ON}).
-\end{remark}
-
-\begin{remark}
- There is no interesting arithmetic on finite ordinals, since functions like successor, addition, etc.,  cannot be represented in \coq{} as \emph{total} functions.
-\end{remark}
-
-\section{See also \dots}
-
-Finite ordinals are also formalized in \ssreflect/\mathcomp~\cite{MCB}.
-In Module~\href{../theories/html/gaia_hydras.ON_gfinite.html}{gaia\_hydras.ON\_gfinite}, we build an instance of class \texttt{ON}.
-
-\inputsnippets{ON_gfinite/finordLtDef, ON_gfinite/finordON}
-
-\inputsnippets{ON_gfinite/Examples, ON_gfinite/Examplesb}
-
-  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.
-
-%%%
-
-\section{Comparing two ordinal notations}
-
-It is sometimes useful to compare two ordinal notations with respect to expressive power
-(the segment of ordinals  they represent).
-
-The following class specifies a strict inclusion of segments.
-The notation \texttt{OA} describes a segment $[0,\alpha)$, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented in \texttt{OA}).
-We require also  that the comparison functions of the two notation systems are compatible.
-
-\begin{figure}[h]
-   \centering
-   \begin{tikzpicture}[very thick, scale=0.6]
-\begin{scope}[color=blue]
-\node (A) at (0,0) {$A$};
-\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
-\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
-\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
-\node (Adots) at (6,0) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red]
-\node (B) at (0,2) {$B$};
-\node(B0) at (2,2)[label=above:$0$]{$\bullet$};
-\node(B1) at (3,2)[label=above:$1$]{$\bullet$};
-\node(B2) at (4,2)[label=above:$2$]{$\bullet$};
-\node (Bdots) at (6,2) {$\ldots$};
-\node (b) at (8,2) [label=above:$\omega$]{$\bullet$};
-\node (bsucc) at (9,2) [label=above:$\omega+1$]{$\bullet$};
-\node (Bdots2) at (10,2) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red!50!blue]
-\draw [->,thin] (A0) -- node [auto] {$\iota$} (B0);
-\draw [->,thin] (A1) -- node [auto] {$\iota$} (B1);
-\draw [->,thin] (A2) -- node [auto] {$\iota$} (B2);
-\draw [->,thin] (Adots) -- node [auto] {$\iota$} (Bdots);
-\end{scope}
-\end{tikzpicture}
-   \caption{\textcolor{blue}{$A$} is a sub-segment  of \textcolor{red}{$B$}}
-   \label{fig:subsegment}
- \end{figure}
-
-If \texttt{OB} is presumed to be correct, then we may consider that \texttt{OA} ``inherits'' its correctness from the bigger notation system \texttt{OB}.
-
-\label{types:SubON}
-\index{hydras}{Library OrdinalNotations!Type classes!SubON}
-
-following definition
-(in ~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{ON\_Generic}).
-
-\input{movies/snippets/ON_Generic/SubONDef}
-
-For instance, we prove that \texttt{Omega} is a sub-notation of
-\texttt{Omega\_plus\_Omega} (with $\omega$ as the first ``new'' ordinal, and \texttt{fin} as the injection).
-
-\vspace{10pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/Incl}
-
-We can also show that, if $i<j$, then the segment $[0,i)$ is a ``sub-segment'' of
-$[0,j)$. Since the terms  ($t\;i$) and ($t\;j$) are not convertible, we consider a ``cast'' 
-function $\iota$ from ($t\;i$) into ($t\;j$), and prove that this function is  a monotonous bijection  from ($t\;i$) to
-the segment $[0,i)$ of ($t\;j$).
-
-%\index{coq}{Commands!Program}
-
-\vspace{10pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
-
-\inputsnippets{ON_Finite/InclIJ, ON_Finite/InclIJa,
-  ON_Finite/InclIJb,ON_Finite/InclIJc, ON_Finite/InclIJd,
-ON_Finite/InclIJz}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove that \texttt{Omega\_plus\_Omega} cannot be a sub-notation of \texttt{Omega}.
-\end{exercise}
-
-\index{hydras}{Projects}
-\begin{project}
-Adapt the definition of \texttt{Hvariant} (Sect.~\ref{sect:hvariant-def}) in order to
-have an ordinal notation as argument. Prove that if $O_A$ is a sub-notation of $O_B$, then any variant defined on  $O_A$ can be automatically transformed into 
-a variant on $O_B$.
-\end{project}
-
-\section{Comparing an ordinal notation with Schütte's model}
-
-Finally, it may be interesting to compare an ordinal notation with the more theoretical model from Schütte (well, at least with our formalization of that model).
-This would be a relative proof of correctness of the considered  ordinal  notation.
-
-The following class specifies that a notation \texttt{OA} describes a segment $[0,\alpha)$,
-where $\alpha$ is a countable ordinal \emph{à la}  Schütte.
-
-\label{types:ON-for}
-\index{hydras}{Library OrdinalNotations!Type classes!ON\_correct}
-
-\input{movies/snippets/ON_Generic/ONCorrect}
-
-For instance, the following theorem tells that \texttt{Epsilon0}, our notation system for the segment $[0,\epsilon0)$ is a correct implementation of the theoretically defined  ordinal $\epsilon_0$
-(see chapter~\ref{chap:schutte} for more details).
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
-
-\input{movies/snippets/Correctness_E0/Epsilon0Correct}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Same work, but replace Schütte's model with \gaia's.
-\end{project}
-
-\begin{project}
-  When you have read Chapter~\ref{chap:schutte}, prove that the sum of two ordinal notations \texttt{ON\_plus} implements the addition of ordinals.
-\end{project}
-
-\section{Isomorphism of ordinal notations}
-
-In some cases we want to show that two notation systems describe the same segment (for instance $[0,3+\omega)$ and $[0,\omega)$\;).
-For this purpose, one may prove that the two notation systems are order-isomorphic.
-
-\index{hydras}{Library OrdinalNotations!Type classes!ON\_Iso}
-
-\label{types:ON-iso}
-
-\input{movies/snippets/ON_Generic/ONIso}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-\label{exo:i-plus-omega}
-Let $i$ be some natural number. Prove that the notation systems 
-\texttt{Omega} and (\texttt{ON\_plus (OrdFin $i$) Omega}) are isomorphic.
-
-{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$, that we prove also in larger notation systems, as well as in Schütte's model.}
-This exercise is partially solved for $i=3$ (in ~\href{../theories/html/hydras.OrdinalNotations.Example_3PlusOmega.html}{OrdinalNotations.Example\_3PlusOmega}).
-
-\end{exercise}
-
-\index{hydras}{Projects}
-\label{exo:ON-mult}
-\begin{project}
-% Define in \coq{} the product of two ordinal notations $N_A$ and $N_B$.
-% If $A$ [resp. $B$] is the underlying type of $N_A$ [resp. $N_B$], the
-% product \texttt{ON\_mult $N_A$ $N_B$} is implemented over the cartesian product $B\times A$ (with the lexicographic ordering).
-
-This exercise is about the non-commutativity of the multiplication of ordinals, reflected in ordinal notations.
-
-For instance, the
-elements of the product (\texttt{ON\_mult Omega (FinOrd 3)}) are ordered as follows.
-\[(0,0),(0,1),(0,2),(0,3),(0,4),\dots,{\color{red}(1,0),} (1,1),(1,2),\dots, {\color{red}(2,0)},(2,1),(2,2),\dots\]
-
-Note that the elements of  (\texttt{ON\_mult (FinOrd 3) Omega}) are differently ordered (without limit ordinals):
-\[(0,0),(1,0),(2,0),(0,1),(1,1),(2,1),(0,2),(1,2),(2,2),(0,3),\dots\]
-
-Prove formally  that (\texttt{ON\_mult (FinOrd $i$) Omega}) is isomorphic to
-\texttt{Omega}  whilst
-\texttt{Omega}  is a sub-notation of (\texttt{ON\_mult Omega (FinOrd $i$)}),
-for any strictly positive $i$.
-
-\textbf{Note:} Like Exercise~\ref{exo:i-plus-omega}, this project corresponds to the [in]equalities $i+\omega=\omega<\omega+i$, for any natural number $i$.
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-Consider two isomorphic ordinal notations \texttt{OA} and \texttt{OB}.
-Prove that, if \texttt{OA} [resp. \texttt{OB}] is a correct implementation 
-of $\alpha$ [resp. $\beta$], then $\alpha=\beta$.
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-\label{project:succ-limit-dec}
-Add to the class \texttt{ON} the requirement that for any $\alpha$ it is decidable whether $\alpha$ is $0$, a successor or a limit ordinal.
-
-\textbf{Hint:}   Beware of the instances associated with sum and product of notations!
-  You may consider additional fields 
-to make the sum and product of notations ``compositional''.
-
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-\label{project:on-setoid}
-Reconsider the  class \texttt{ON}, with an equivalence instead of Leibniz equality.
-\end{project}
-
-%%%% ICI ICI
-
-\section{Other ordinal notations}
-%%% TODO : Fix the multiplication function in branch FixOmegaOmega
-
-% \index{hydras}{Projects}
-
-% \begin{project}
-% The directory \texttt{theories/OmegaOmega} contains an ad-hoc formalization of $\omega^\omega$, contributed by Pascal Manoury.
-Every ordinal $\alpha$ is represented by a list $l$ whose elements are the coefficients of $\omega$ in  the Cantor normal form of $\alpha$ (in reverse order).
-For instance, the ordinal
-% $\omega^{8}\times 5 + \omega^{6}\times 8 + \omega^2\times 10 + \omega + 7$ is represented by the list \texttt{[5;0;8;0;0.0;10,1,7]}.
-
-%  Develop this representation and compare it with the other ordinal notations.
-
-% \end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Let $N_A$ be a notation system for ordinals strictly less than $\alpha$, 
-with the strict order $(A,<_A)$. Please build the notation system
-\texttt{ON\_Expl $N_A$}, on the type of multisets of elements of $A$
-(or, if preferred, the type of non-increasing finite sequences on $A$,
-provided with the lexicographic ordering on lists).
-
-For instance, let us take $N_A=\texttt{Omega}$, and take $\alpha=\langle 4,4,2,1,0\rangle$,
- $\beta=\langle 4,3,3,3,3,3,2\rangle$, and $\gamma=\langle 5\rangle$. Then $\beta<\alpha<\gamma$.
-
-In contrast the list $\langle5,6,3,3\rangle$ is not non-increasing (\emph{i.e.} sorted w.r.t. $\geq$), so it is not to be considered.
-
-Note that if the notation $N_A$ implements the ordinal 
-$\alpha$,  the new notation $\omega^{N_A}$ must implement the ordinal $\phi_0(\alpha)$, a.k.a. $\omega^\alpha$ (see chapter~\ref{chap:schutte})
-
-\end{project}
-
-\begin{remark}
- The set of ordinal terms in Cantor normal form (see Chap.~\ref{chap:T1}) and 
-in Veblen normal form (see 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}{Gamma0.Gamma0}) are shown to be ordinal notation systems, but there is a lot of work to be done in order to unify ad-hoc  definitions and proofs which were written before the definition of the \texttt{ON} type class.
-\end{remark}
-
-\section{Ordinal Notations and \gaia's ordinals}
-\gaiasign Module~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge} contains an instance of class \texttt{ON E0lt compare}, where \texttt{E0lt} is the order on the well-formed ordinal terms below $\epsilon_0$ defined in \gaiaHydras library (please see Chapter~\ref{gaia-chapter}).
-
-%------------------------------------------------------------------------
-
-\chapter[A proof of termination, using epsilon0]{A proof of termination, using ordinals below \texorpdfstring{$\epsilon_0$}{Epsilon0}}
-
-\label{cnf-math-def}
-\label{chap:T1}
-
-In this chapter, we adapt to \coq{} the well-known~\cite{KP82}  proof that Hercules eventually wins every battle, whichever the strategy  of each player.
-In other words, we present  a formal and self contained proof of termination  of all [free] hydra battles.
-First, we take from Manolios and Vroon~\cite{Manolios2005} a representation of the ordinal $\epsilon_0$ as terms in Cantor normal form.
-Then, we define a variant for hydra battles as a measure that maps any hydra to some ordinal strictly less than $\epsilon_0$.
-
-\section{The ordinal \texorpdfstring{\(\epsilon_0\)}{epsilon0}}
-\label{sec:epsilon0-intro}
-
-The ordinal \(\epsilon_0\) is the least ordinal number that satisfies 
-the equation \(\alpha = \omega^\alpha\), where \(\omega\) is 
-the least infinite ordinal.
-Thus, we can intuitively consider \(\epsilon_0\) as an
-\emph{infinite} \(\omega\)-tower.
-
-\subsection{Cantor normal form}
-\index{maths}{Cantor normal form}
-
-Any ordinal strictly less that \(\epsilon_0\) 
-can be finitely represented by a unique  \emph{Cantor normal form}, 
-that is, an expression  which is either  the ordinal \(0\) or 
-a sum  \(\omega^{\alpha_1} \times n_1 + \omega^{\alpha_2} \times n_2 + 
-  \dots + \omega^{\alpha_p} \times n_p\) where all the \(\alpha_i\) 
-are ordinals in Cantor  normal form, \(\alpha_1 > \alpha_2 > \alpha_p\), 
-and all the \(n_i\) are positive integers.
-
-An example of Cantor normal form is displayed in Fig \ref{fig:cnf-example}:
-Note that  any ordinal of
-the form \(\omega^0 \times i + 0\) is just written \(i\).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
-\node[color=blue]{$\omega^{(\omega^\omega\,+\, \omega^2 \times 8 \,+\, \omega)}+ \omega^\omega + \omega^4+ 6$};
-\end{tikzpicture}
-\caption{\label{fig:cnf-example}
-An ordinal in Cantor normal form}
-\end{figure}
-
-In the rest of this section, we define an inductive type for representing in \texttt{Coq}
-all the ordinals strictly  less than  \(\epsilon_0\), then extend some arithmetic operations
-to this type, and finally prove that our representation fits well with 
-the expected mathematical properties: the order we define is a well order, 
-and the decomposition into Cantor normal form  is consistent 
-with the implementation of the arithmetic operations of exponentiation of base \(\omega\) 
-and addition.
-
-\paragraph*{Remark}
-\label{sec:orgheadline65}
-Unless explicitly mentioned, the term ``ordinal" will be used instead of
-``ordinal strictly less than \(\epsilon_0\)" (except in Chapter~\ref{chap:schutte} where it stands for ``countable ordinal'').
-
-\subsection{A data type for  ordinals in Cantor normal form}
-\label{sec:orgheadline72}
-\label{sec:T1-inductive-def}
-
-% Our user contribution~\cite{CantorContrib} represents 
-% the set of ordinals strictly less than $\epsilon_0$ in Cantor normal form as in~\cite{Manolios2005}, and also the set
-% of ordinals strictly  less than $\Gamma_0$ in Veblen normal form.
-
-    Let us define an inductive type whose 
-constructors are respectively associated
-with the ways to build Cantor normal forms:
-
-\begin{itemize}
-\item the ordinal \(0\)
-\item the construction \((\alpha,\, n,\,\beta)  \mapsto \omega^\alpha \times (n + 1)+ \beta \quad (n\in\mathbb{N})\)
-\end{itemize}
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#T1}{Epsilon0.T1}}
-
-\label{types:T1}
-%\index{Constants!zero:T1}
-\index{hydras}{Library Epsilon0!Types!T1}
-
-\input{movies/snippets/T1/T1Def}
-
-\subsubsection{Example}
-
-\label{alpha0-def}
-For instance, the ordinal  $\omega^\omega+\omega^3\times 5+2$ is represented by the following term of type \texttt{T1}:
-
-\input{movies/snippets/T1/alpha0}
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5, level 1/.style={sibling distance=6cm},
-level 2/.style={sibling distance=35mm},  
-level 3/.style={sibling distance=17mm}]
-\node  {cons}
-  child {  node {cons}
-            child { node {cons} child {node {zero}} child {node{0}} child{node{zero}}}
-         child {node {0}}
-         child {node {zero}}}
-    child {node {0}}
-   child {node {cons} 
- child { node {cons} child {node {zero}} child {node{2}} child{node{zero}}}
-  child {node {4}}
-         child {node {cons} child {node {zero}} child {node{1}} child{node{zero}}}};
-
-\end{tikzpicture}
-
-\caption{The tree-like representation of the ordinal $\omega^\omega+\omega^3\times 5 +2$\label{fig:cnf-tree}}
-
-\end{figure}
-
-\paragraph{Remark}
-For simplicity's sake, we chose to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the construction (\texttt{cons $\alpha$ $n$ $\beta$}) is intended to represent the
-ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
-In a future version, we should replace  the type \texttt{nat} with \texttt{positive} in \texttt{T1}'s 
-definition. But this replacement would take a lot of time \dots{}
-
-\paragraph{Remark}
-The name \texttt{T1} we gave to this data-type  is proper to this development and refers
-to a hierarchy of ordinal notations.
-For instance, in Library \href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2},  the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{Gaia, schutte}).
-\noindent
-
-\input{movies/snippets/T2/T2Def}
-
-\subsection{About Gaia}
-\gaiasign Chapter~\vref{gaia-chapter} describes the present state of a
-project of making compatible the libraries \HydrasLib and \gaia. At present, both libraries contain their own version of the inductive type \texttt{T1}, with the same base name for the constructors.
-
-\inputsnippets{nfwfgaia/T1def}
-
-Thus, many examples of this and following chapters can be run
-in \gaia's context. We will signal any difference (notation, name) which may appear.
-A closer integration (same data type and functions) is still in project.
-
-\begin{remark}
-  \label{remark:hydra-gaia-renaming}
-  \textcolor{lookcolor}{In future releases, we plan to make some
-    \HydrasLib identifiers progressively  deprecated, in favour of \gaia's names.}
-\end{remark}
-
-\subsection{Abbreviations}
-
-Some abbreviations may help to write more concisely complex ordinal terms.
-
-\subsubsection{Finite ordinals}
-\label{sec:orgheadline67}
-
-For representing finite ordinals, \emph{i.e.} natural numbers, we first introduce a notation for terms of the form $n+1$, then define a coercion from type \texttt{nat} into \texttt{T1}.
-\label{sect:notation-FS}
-
-\label{sect:notation-F}
-
-\input{movies/snippets/T1/finiteOrds}
-
-\paragraph*{\gaiasign}
-  In \texttt{gaia.ssete9}, the $n$-th finite ordinal is also written
-  \texttt{$\backslash${F}\,$n$}. There is no coercion in \gaia from \texttt{nat} to \texttt{T1}.
-
-  \vspace{4pt}
-
-  From~\href{../theories/html/gaia_hydras.HydraGaia_Examples.html}{gaia\_hydras.HydraGaia\_Examples}.
-
-  \inputsnippets{HydraGaia_Examples/Demoa}
-
-% \index{Coq!Coercions}
-% \index{Functions!Coercions@Coercions (from nat to ordinal types)}
-% \begin{remark}
-% Please refer to the remark~\pageref{warning:coercions} about the use of coercions.
-% % The use of coercions like \texttt{fin} allow us to be close to the mathematical tradition where natural numbers are ordinals too.
-% % Nevertheless, it may happen that a goal like \texttt{3 < 5} could be 
-% % interpreted as \texttt{(lt (fin 3) (fin 5))},  depending on the current notation scope.  
-% % When this misinterpretation happens, tactics like \texttt{auto with arith}, \texttt{lia} do not work!
-% % Thus, it is useful to write \texttt{(3 < 5)\%nat}  an inequality between two natural numbers. 
-% \end{remark}
-
-\subsubsection{The ordinal \(\omega\)}
-\label{sec:orgheadline68}
-
-  Since \(\omega\)'s Cantor normal form is
-i.e. \(\omega^{\omega^0}\times 1+ 0\), we can define the following abbreviation:
-
-\label{sect:omega-notation2}
-
-\input{movies/snippets/T1/omegaDef}
-
-Note that \texttt{omega} is not an identifier in \HydrasLib, thus any tactic like \texttt{unfold omega} would fail.
-
-\paragraph*{\gaiasign}
-In \texttt{gaia.ssete9}, the ordinal $\omega$ is bound to the \emph{constant}  \texttt{T1omega}.
-
-\subsubsection{The ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
-\label{sect:notation-phi0}
-We provide also a notation for ordinals of the form $\omega^\alpha$.
-
-\index{hydras}{Library Epsilon0!Notations!phi0@phi0 (exponential of base omega)}
-
-\input{movies/snippets/T1/phi0Def}
-
-\index{maths}{Additive principal ordinals}
-
-\begin{remark}
-\label{sec:orgheadline69}
-The name \(\phi_0\)
-   comes from ordinal numbers theory. In~\cite{schutte}, Schütte defines 
-$\phi_0$  as the ordering (\emph{i.e.} enumerating) function of the set  of \emph{additive principal ordinals} \emph{i.e.} strictly positive ordinals $\alpha$ that verify $\forall \beta<\alpha, \beta+\alpha=\alpha$.
-For Schütte,  $\omega^\alpha$ is just a notation for $\phi_0(\alpha)$.
-See also Chapter~\ref{chap:schutte} of this document.
-\end{remark}
-
-\paragraph*{\gaiasign}
-In \texttt{gaia.ssete9} the identifier \texttt{phi0} is bound to a plain constant (not a notation).
-
-\subsubsection{The hierarchy of \(\omega\)-towers:}
-\label{sec:orgheadline71}
-
-The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximated by the sequence of $\omega$-towers (see also Sect~\vref{sect:epsilon0-as-limit} ).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/towerDef}
-
-For instance, Figure~\ref{fig:tower7} represents  the ordinal returned by the
- evaluation of the term \texttt{omega\_tower 7}.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
-\node[color=blue]{$\omega^{{{\omega}^{{{\omega}}^{{{\omega}}^{{\omega^{{\omega}^{\omega}}}}}}}}$};
-\end{tikzpicture}
-\caption{\label{fig:tower7}
-The $\omega$-tower of height 7}
-\end{figure}
-
-\subsection{Pretty-printing ordinals in Cantor normal form}
-\label{sect:ppT1}
-\index{hydras}{Library Epsilon0!Types!ppT1}
-
-Let us consider again the ordinal $\alpha_0$ defined in section~\vref{alpha0-def}. 
-If we ask \coq{} to print its  normal form, we get a hardly readable term of type \texttt{T1}.
-
-\input{movies/snippets/T1/alpha0Compute}
-
-The following data type defines a more readable abstract syntax for  ordinals terms in Cantor normal form:
-
-\label{types:ppT1}
-\index{hydras}{Library Epsilon0!Functions!pp@ pp (pretty printing terms in Cantor normal form)}
-
-\input{movies/snippets/T1/ppT1Def}
-
-The function \texttt{pp: T1 -> ppT1} converts any closed term of type \texttt{T1} into a human-readable expression. For instance, let us convert the term \texttt{alpha\_0}.
-
-\input{movies/snippets/T1/ppAlpha0}
-
-\paragraph*{\gaiasign}
-In ~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge},
-we define a variant of \texttt{pp} for pretty-printing terms of type \texttt{ssete9.CantorOrdinal.T1} (see Sect.~\vref{sect:gaia-ppT1}).
-
-\index{hydras}{Projects}
-\begin{project}
-Design  (in \ocaml?) a set of tools for systematically pretty printing ordinal terms in Cantor normal form.
-\end{project}
-
-\subsection{Comparison between ordinal terms}
-\label{sec:orgheadline73}
-
-% Our formalisation of Cantor Normal Form will take two steps:
-% 1 Definition of a strict order \texttt{o<} on the type \texttt{T1}, 
-% 2 Using \texttt{o<} for characterizing terms in normal form.
-
-In order to compare two terms of type \texttt{T1}, we define a recursive function \texttt{compare} that maps two ordinal tems $\alpha$ and $\beta$ to a value of type \texttt{comparison}.
-This type is defined in \coq's standard library
-\texttt{Init.Datatypes} and
-contains three constructors:  \texttt{Lt} (less than), \texttt{Eq} (equal), and
-\texttt{Gt} (greater than).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/compareDef}
-
-\label{Predicates:lt-T1}
-Please note that this definition of \texttt{lt} makes it easy to write proofs by computation, as shown by the following examples.
-
-\vspace{4pt}
-
-\input{movies/snippets/T1/ltExamples}
-
-Links between the function \texttt{compare} and the relations
-\texttt{lt} and \texttt{eq} are established through the following lemmas (~\vref{sect:comparable-def}).
-\vspace{4pt}
-
-\input{movies/snippets/T1/Instances}
-
-\paragraph*{\gaiasign}
-In \gaia, the strict order \texttt{T1lt} on ordinal terms is directly defined as a boolean function (in \ssreflect's style).
-
-\inputsnippets{nfwfgaia/T1ltDef}
-
-In section~\vref{sect:lt-compat-gaia}, we show that both definitions are mutually equivalent.
-
-\subsubsection{A Predicate for Characterizing Normal Forms}
-\label{sect:t1-nf}
-
-\label{sec:orgheadline74}
-\label{sec:orgheadline75}
-Our data-type \texttt{T1} allows us to write expressions that
-are not properly in Cantor normal form as specified in Section \ref{sec:epsilon0-intro}.
-For instance, consider the following term of type  \texttt{T1}.
-
-\input{movies/snippets/T1/badTerm}
-
-This term would have been written \(\omega^1\times 2 + \omega^\omega \times 3\) in the usual mathematical notation.
-We note that the exponents of $\omega$ are not in the right (strictly decreasing) order.
-Nevertheless, with the help of the order \texttt{lt} on \texttt{T1}, we are now able to characterize
-the set of all well-formed ordinal terms:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#nf_b}{Epsilon0.T1}}
-
-\label{Predicates:nf-T1}
-
-\input{movies/snippets/T1/nfDef}
-
-\input{movies/snippets/T1/nfAlpha0}
-
-\input{movies/snippets/T1/nfBadTerm}
-
-\paragraph*{\gaiasign}
-In \gaia, the boolean function which characterizes ordinal terms in normal form is defined as follows:
-
-\inputsnippets{nfwfgaia/T1nfDef}
-
-In Sect.~\vref{nf-gaia-compat}, we show that \gaia's \texttt{T1nf} is extensionally equivalent with \HydrasLib'  \texttt{nf}.
-
-\subsection{Making normality implicit}
-  We would like to get rid of terms of type \texttt{T1} which are not in Cantor normal form.
-A simple way to do this is to consider statements of the form 
-\texttt{forall alpha: T1, nf alpha -> $P$ alpha}, where $P$ is a predicate over type \texttt{T1}, like in the following lemma \footnote{Ordinal addition is formally defined a little later (page~\ref{sect:infix-plus-T1})}.
-
-\input{movies/snippets/T1/plusIsZero}
-
-\vspace{4pt}
-
-But this style leads to clumsy statements, and generates too many subgoals in interactive proofs (although often solved with \texttt{auto} or \texttt{eauto}).
-
-One may encapsulate conditions of the form \texttt{(nf $\alpha$)} in
-the most used predicates. For instance, we introduce the restriction of \texttt{lt} to terms in normal form, and provide a handy notation for this restriction.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Prelude.Restriction.html}{hydras.Prelude.Restriction}}
-
-\input{movies/snippets/Restriction/restrictionDef}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#LT}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/LTDef}
-
-\label{Predicates:LT-T1}
-
-For instance, in the following lemma, the condition that $\alpha$ is in normal form is included in the condition $\alpha< 1$.
-
-\input{movies/snippets/T1/LTOne}
-
-\subsubsection{\texttt{E0}: a sigma-type for \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-As we noticed in Sect.~\ref{sect:t1-nf}, the type \texttt{T1} is not a correct ordinal notation, since it contains terms that are not in Cantor normal form.
-In certain contexts (for instance in Sections~\ref{sect:L-equations}, \ref{sect:hardy},
-and \ref{sect:wainer}),  we need to define total recursive functions on well-formed ordinal terms less  than $\epsilon_0$, using the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777}.
- In order to define a type whose inhabitants represent just ordinals, we build a type gathering a term of type \texttt{T1} and a proof that this term is in normal form.
-
-\label{sect:E0-def}
-\label{types:E0}
-\index{hydras}{Library Epsilon0!Types!E0}
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html\#E0}{Epsilon0.E0}}
-
-\input{movies/snippets/E0/E0Def}
-
-Many constructs : types, predicates, functions, notations, etc., on type \texttt{T1} are adapted to \texttt{E0}.
-
-First, we declare a notation scope for \texttt{E0}, then we redefine the predicates of comparison.
-
-\input{movies/snippets/E0/E0Scope}
-\label{Predicates:Lt-E0}
-\input{movies/snippets/E0/LtLeDef}
-
-Equality in \texttt{E0} is just Leibniz equality. Note that, since \texttt{nf} is
-defined by a Boolean function, for  any term $\alpha:\texttt{T1}$, there exists at most one proof of \texttt{nf $\alpha$}, thus two ordinals of type \texttt{E0} are
-equal if and only if their projection to \texttt{T1} are equal (see also Sect.~\vref{sect:eq-proof-unicity}).
-
-\vspace{4pt}
-
-\index{coq}{Unicity of equality proofs}
-
-\inputsnippets{E0/nfProofUnicity}
-\inputsnippets{E0/E0EqIff}
-
-In order to  upgrade constants and functions from type \texttt{T1} to \texttt{E0}, we have to prove that
-the term they build is in normal form.
-For instance, let us represent the ordinals $0$ and $\omega$   as instances of the class \texttt{E0}.
-
-\vspace{4pt}
-
-\label{sect:omega-T1}
-
-\input{movies/snippets/E0/ZeroOmega}
-
-%\index{Constants!zero:T1}
-
-\paragraph*{\gaiasign}
- The library~\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines a \gaia-compatible type \texttt{E0}.
-
-\subsection{Syntactic definition of limit and successor ordinals}
-
-Pattern matching and structural recursion allow us to define boolean characterizations  of successor and limit ordinals.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/succbLimitb}
-
-The correctness of these definitions with respect to the mathematical notions of
-limit and successor ordinals is established through several lemmas. For instance,
-Lemma \texttt{canonS\_limit\_lub}, page~\pageref{lemma:canonS-limit}, shows that
-if $\alpha$ is (syntactically) a limit ordinal, then it is the least upper bound of
-a strictly increasing sequence of ordinals.
-
-\paragraph*{\gaiasign}
-In \gaia, the boolean functions associated with limit and successor ordinals are called \texttt{T1limit} and \texttt{T1is\_succ}.
-
- From~\href{../theories/html/gaia_hydras.HydraGaia_Examples.html}{gaia\_hydras.HydraGaia\_Examples}.
-
-\inputsnippets{HydraGaia_Examples/Demob}
-
-\subsection{Arithmetic on \texorpdfstring{$\epsilon_0$}{epsilon0}}
-\subsubsection{Successor}
-
-\index{hydras}{Library Epsilon0!Functions!succ}
-
-The successor of any ordinal $\alpha< \epsilon_0$ is defined by structural 
-recursion on its Cantor normal form.
-
-\label{Functions:succ-T1}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/succDef}
-
-The following lemma establishes the connection between the  function
-\texttt{succ} and the Boolean predicate \texttt{succb}.
-
-\vspace{4pt}
-
-\input{movies/snippets/T1/succbIff}
-
-\begin{exercise}[\gaiasign]
-  Look for the \gaia-theorem which corresponds to \texttt{succb\_iff}.
-\end{exercise}
-
-\subsubsection{Successor function on \texttt{E0}}
-
-The function \texttt{succ} on \texttt{T1} is extended to \texttt{E0} the following way:
-
- \emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
-
- \inputsnippets{E0/SuccOnE0}
-
-\index{hydras}{Exercises}
- \begin{exercise}
-Prove in \coq{} that for any ordinal $0< \alpha<\epsilon_0$, $\alpha$ is a limit if 
-and only if for all $\beta<\alpha$, the interval $[\beta,\alpha)$ is
-infinite.
-
-\emph{You may start this exercise with the file
-     \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/Limit_Infinity.v}{exercises/ordinals/Limit\_Infinity.v}.}
- \end{exercise}
-
-\subsubsection{Addition and multiplication}
-
-Ordinal addition and multiplication are also defined by structural recursion over the type \texttt{T1}. Please note that they use the \texttt{compare} function on some subterms of their arguments.
-
-\label{sect:infix-plus-T1}
-
-\input{movies/snippets/T1/plusDef}
-
-\input{movies/snippets/T1/multDef}
-
-\paragraph*{\gaiasign}
-We keep \gaia's base names for
-addition and multiplication of ordinal terms below $\epsilon_0$.
-Please refer to Sect.~\ref{sect:plus-mult-gaia-hydras} about compatibility of both arithmetics.
-
-\subsubsection{Examples}
-
-The following examples are instances of \emph{proofs by computation}. Please note that  addition and multiplication on \texttt{T1}
-are not commutative. Moreover,  both operations fail to be strictly monotonous in their first argument.
-
-\input{movies/snippets/T1/plusMultExamples}
-
-\input{movies/snippets/T1/notMono}
-
-The function \texttt{succ} is related with addition through the following lemma:
-
-\inputsnippets{T1/succIsPlusOne, T1/succIsPlusOnez}
-
-\subsubsection{Arithmetic on type \texttt{E0}}
-
- We define an addition in type \texttt{E0}, since the sum of two terms in normal form is in normal form too.
-
-\input{movies/snippets/T1/plusNf}
-
-\input{movies/snippets/E0/plusE0}
-\input{movies/snippets/E0/CheckPlus}
-
-\begin{remark}
-In all this development, two representations of ordinals co-exist: ordinal terms (type \texttt{T1}, notation scope \texttt{t1\_scope}, for reasoning on the tree-structure of Cantor normal forms), and ordinal terms \emph{known to be in normal form} (type \texttt{E0}, notation scope \texttt{E0\_scope}).
-Looking at the contexts displayed by \coq{} prevents you from any risk of confusion.
-\end{remark}
-
-%% To simplify !
-\index{hydras}{Exercises} 
-\begin{exercise}
-Prove that for any ordinal $\alpha:\texttt{E0}$, 
-$\omega\leq \alpha$ if and only if, for any natural number $i$,
-$i+\alpha=\alpha$.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/ge_omega_iff.v}{exercises/ordinals/ge\_omega\_iff.v}.}
-\end{exercise}
-
-\subsection{A proof by computation}
-\label{Ex42-E0}
-
-It is interesting to compare the following proof of the equality
-$\omega+42+\omega^2$ with the more theoretical proof in Sect~\vref{Ex42-schutte}.
-
-\inputsnippets{E0/Ex42}
-
-\section{Well-foundedness and transfinite induction}
-
-\index{maths}{Transfinite induction}
-
-\subsection{About  well-foundedness}
-\label{sec:T1Wf}
-\label{sec:orgheadline82}
-   In order to use \texttt{T1} for proving termination results,
-we need to prove that  our order \texttt{<} is well-founded. Then we will get \emph{transfinite induction} for free.
-
-The proof of well-foundedness of the strict order $<$ on Cantor normal forms is already 
-available in the Cantor contribution by Castéran and Contejean~\cite{CantorContrib}. That proof relies on a library on recursive path orderings written by
-E. Contejean. We present here  a direct proof of the same result, which does not require any knowledge on r.p.o.s.
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that the \emph{total} order \texttt{lt} on \texttt{T1} is not well-founded. 
-\textbf{Hint:}  You will have to build a counter-example with terms of type \texttt{T1}
-which are not in Cantor normal form.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/T1_ltNotWf.v}{exercises/ordinals/T1\_ltNotWf.v}.}
-\end{exercise}
-
-\subsubsection{A first attempt}
-\label{sec:orgheadline77}
-%\index{coq}{Well-founded induction}
-
-It is natural to try to prove by structural induction over \texttt{T1} 
-that every term in normal form is \texttt{LT}-accessible.
-
-Unfortunately, it won't work. Let us consider some well-formed term
- $\alpha=\texttt{cons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are \texttt{LT}-accessible.
-In order to prove the accessibility of $\alpha$, we have to consider
-any well formed term \(\delta\) such that \(\delta<\alpha\). 
-But nothing guarantees that \(\delta\)  is strictly  less than \(\beta\) nor \(\gamma\), and we cannot use the induction hypotheses on   \(\beta\) nor \(\gamma\).
-
-\input{movies/snippets/T1/wfLTBada}
-
-The problem comes from the too weak hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
-\(\gamma\);
-for instance \(\delta\) may be of the form
-\texttt{cons $\beta$ $p'$  $\gamma'$},
-where    \(p' < n\).
-Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.
-
-\input{movies/snippets/T1/wfLTBadz}
-
-\subsubsection{Using a stronger inductive predicate.}
-\label{sec:orgheadline78}
-\label{sec:strongly-accessible}
-
-  Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
-through \texttt{LT}, we propose to consider the following (stronger) predicate:
-
-\input{movies/snippets/T1/AccStrongDef}
-
-The following lemma is an application of the strict inequality 
-$\alpha < \omega ^\alpha$. If $\alpha$ is strongly accessible, then, by definition,
-$\omega^\alpha$ is accessible, thus $\alpha$ is \emph{a fortiori} accessible.
-
-\input{movies/snippets/T1/AccStrongStronger}
-
-Thus, it remains to prove that every ordinal strictly less than $\epsilon_0$
-is strongly accessible.
-
-% \subsubsection{Structure of the proof of well-foundedness of \texttt{LT}}
-
-\label{sec:orgheadline81}
-\label{proof-wf-epsilon0}
-\paragraph{A helper}
-\label{sec:orgheadline79}
-
-First, we prove that, for  any \texttt{LT}-accessible term $\alpha$, $\alpha$ is 
-strongly accessible too (\emph{i.e.} any well formed
-term (\texttt{cons $\alpha$ $n$ $\beta$})  is accessible).
-
-The proof is structured as an induction on $\alpha'$s accessibility. Let us consider
-any  accessible term $\alpha$.
-
-\input{movies/snippets/T1/AccImpAccStrong}
-
-Let \texttt{n:nat} and \texttt{beta:T1} such that (\texttt{cons alpha n beta}) is in normal form. 
-We prove first that \texttt{beta} is accessible,  which allows us to prove by well-founded induction on \texttt{beta}, 
-and natural induction on \texttt{n}, that (\texttt{cons alpha n beta}) is accessible.
-The proof, quite long, can be consulted in \href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}.
-
-\paragraph{Accessibility of any well-formed ordinal term}
-\label{sec:orgheadline80}
-
-Our goal is still to prove accessibility of any well formed ordinal term.
-Thanks to our previous lemmas, we are almost done (by a simple structural induction!).
-
-\input{movies/snippets/T1/nfAcc}
-
-\input{movies/snippets/T1/T1Wf}
-
-\index{maths}{Transfinite induction}
-
-Traditionnally, well-founded induction on ordinals is called \emph{transfinite induction}.
-
-\inputsnippets{T1/transfiniteRecursor, T1/transfiniteInduction}
-
-As a corollary, the order \texttt{Lt} on type \texttt{E0} is well-founded too.
-
-\inputsnippets{E0/E0LtWf}
-
-\begin{remark}
-\label{remark:a3pat}
-The alternate proof of well-foundedness using \'Evelyne Contejean's work on
-    recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the
-    library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0.Epsilon0rpo}.
- \end{remark}
-
-\subsection{An ordinal notation for  \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-We build an instance of \texttt{ON}, and prove its correctness w.r.t. Schutte's model (see Chapter~\ref{chap:schutte}).
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
-
-\input{movies/snippets/E0/InstanceEpsilon0}
-
-\label{instance-epsilon0}
-
-\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
-
-\input{movies/snippets/Correctness_E0/injectDef}
-
-\input{movies/snippets/Correctness_E0/Epsilon0Correct}
-
-\index{hydras}{Projects}
-\begin{project}
- \emph{This exercise is a continuation of Project~\vref{exo:ON-mult}.}
-Use \texttt{ON\_mult} to define an ordinal notation \texttt{Omega2} for $\omega^2=\omega\times\omega$.
-
-Prove that \texttt{Omega2} is a sub-notation of \texttt{Epsilon0}.
-
-Define on \texttt{Omega2} an addition compatible with the addition on \texttt{Epsilon0}.
-
-\textbf{Hint}. You may use the following definition (in 
-    \href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}).
-
-\input{movies/snippets/ON_Generic/SubONSameOp}
-
-     \end{project}
-
-    \index{hydras}{Projects}
-    \begin{project}
-    The class \texttt{ON} of ordinal notations has been defined long after this 
-    chapter, and is not used in the development of the type \texttt{E0} yet.
-    A better integration of both notions should simplify the development on ordinals in Cantor normal form. This integration is planned for the future versions.
-
-    \end{project}
-
-    \section{A refinement of \texttt{E0} : an ordinal notation for \texorpdfstring{$\omega^\omega$}{omega\^omega}}
-
-    In Module   \href{https://github.com/coq-community/hydra-battles/blob/master/theories/
-      ordinals/OrdinalNotations/OmegaOmega.v}{theories/ordinals/OrdinalNotations/OmegaOmega.v},
-    we represent ordinals below $\omega^\omega$ by lists of pairs of natural numbers (with the same coefficient shift as in \texttt{T1}).
-    For instance, the ordinal $\omega^4\times 10 + \omega^3 + \omega+ 5$ is represented by the list \texttt{(4,9)::(3,0)::(1,0)::(0,4)::nil}.
-
-\inputsnippets{OmegaOmega/LODef}
-
-  The usual operations : \texttt{succ}, \texttt{+}, \texttt{*} are simple variants of the same operations in \texttt{T1}.
-
-  \inputsnippets{OmegaOmega/succPlusMult}
-
-    We establish this representation as a \emph{refinement} of the data types we used to represent ordinals less than $\epsilon_0$.
-Thus, many properties like well-foundedness of $<$ and associativity of $+$,  of this ordinal notations have very short proofs.
-
-    \inputsnippets{OmegaOmega/refineDef, OmegaOmega/phi0Ref,
-      OmegaOmega/succRef, OmegaOmega/plusRef,
-      OmegaOmega/multRef}
-
-    In order to make an ordinal notation for $\omega^\omega$, we follow the same steps as for $\epsilon_0$:
-    \begin{enumerate}
-    \item Define an order \texttt{lt}, which refines the order
-      \texttt{lt} on \texttt{T1}.
-      \inputsnippets{OmegaOmega/compareDef}
-      \inputsnippets{ OmegaOmega/ltDef, OmegaOmega/ltRef}
-
-    \item Define the predicate ``to be in normal form''.
-      \inputsnippets{OmegaOmega/nfDef}
-    \item Define a class \texttt{OO} of terms in normal form, and an embedding from \texttt{E0} into \texttt{OO}.
-      \inputsnippets{OmegaOmega/OODef}
-       \inputsnippets{OmegaOmega/embedDef}
-      \item Infer well-foundedness of the order on \texttt{OO}.
-    \inputsnippets{OmegaOmega/ltWf}
-
-    \end{enumerate}
-
-    \inputsnippets{OmegaOmega/ONOO, OmegaOmega/OOz}
-
-    Let us show a few examples.
-
-    \inputsnippets{OmegaOmega/OODemo}
-
-    \index{hydras}{Exercises}
-
-    \subsection{Related work}
-    The article~\cite{Manoury2016}  defines another representation of ordinals below $\omega^\omega$ based on lists of natural numbers.
-
-\begin{exercise}[not trivial]
-  It may be interesting to write a \emph{direct} proof  of well-foundedness of the order in $\omega^\omega$ (\emph{i.e.}
-  without using properties of $\epsilon_0$). This exercise may
-  help to understand better the proof structure of Sect.~\vref{sec:strongly-accessible}.
-    \end{exercise}
-%ici
-    \section{A variant for hydra battles}
-
-    In order to prove the termination of any hydra battle, we try to define a variant mapping hydras to ordinals strictly less than $\epsilon_0$.
-    In order to make such a variant easy to define (for instance by a structural recursion), we introduce a variant of addition, which, contrary to
-    $+$, is commutative and strictly monotonous in both of its arguments. This last property makes it possible to prove that our function is 
-    truly a variant for hydra battles (in Sect.~\vref{sect:variant-decr}).
-
-    \subsection{Natural sum (a.k.a. Hessenberg's  sum)}
-    \label{sec:orgheadline87}
-    \label{hydra-variant}
-    \label{sect:hessenberg-def}
-
-    Natural sum (Hessenberg's  sum) is a commutative and monotonous version of
-    addition. It is used as an auxiliary operation  for defining variants
-    for hydra battles, where Hercules is allowed to chop off any  head of the hydra.
-
-    In the literature, the natural sum of ordinals \(\alpha\) and \(\beta\)
-    is often denoted by \(\alpha \# \beta\)  or  \(\alpha \oplus  \beta\).
-    Thus we called \texttt{oplus} the associated \emph{Coq} function.
-
-    \subsubsection{Definition of \texttt{oplus}}
-    \label{sec:orgheadline84}
-    %\index{Functions!oplus @ oplus (Hessenberg commutative sum)}
-
-    The definition of \texttt{oplus} is recursive in both of its 
-    arguments and uses  the same pattern as for the \texttt{merge} function on lists of library
-    \texttt{Coq.Sorting.Mergesort}.
-
-    \begin{enumerate}
-    \item Define a nested recursive function, using the \texttt{Fix} 
-        construct
-
-    \item Build a principle of induction dedicated to \texttt{oplus}
-
-    \item Establish equations associated to each case of the definition.
-    \end{enumerate}
-
-    \paragraph{Nested recursive definition}
-    \label{sec:orgheadline83}
-
-    The following definition is composed of 
-    \begin{itemize}
-    \item A main function \texttt{oplus}, structurally recursive in its 
-    first argument \texttt{alpha}
-    \item An auxiliary function \texttt{oplus\_aux} within the scope of \texttt{alpha},
-    structurally recursive in its argument \texttt{beta};  \texttt{oplus\_aux beta} 
-       is supposed to compute  \texttt{oplus alpha beta}.
-    \end{itemize}
-
-%  \pagebreak 
-  \vspace{4pt}
-    \emph{From Module~\href{../theories/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}
-
-    \label{sect:infix-oplus}
-
-    \input{movies/snippets/Hessenberg/oplusDef}
-
-    The reader will note that each recursive call of the functions
-    \texttt{oplus} and \texttt{oplus\_aux} satisfies \emph{Coq}'s constraint
-    on recursive definitions. The function \texttt{oplus} is recursively called on a sub-term of its first argument,
-    and \texttt{oplus\_aux} on a sub-term of its unique argument.
-    Thus, \texttt{oplus}'s definition is accepted by \coq{} as a structurally recursive function.
-
-    \subsubsection{Rewriting lemmas}
-    \label{sec:orgheadline86}
-
-    \emph{Coq}'s constraints on recursive definitions result in 
-    the quite  complex form of \texttt{oplus}'s definition.
-    Proofs of properties of this function can be simpler if we
-     derive a few  rewriting lemmas that will help to simplify 
-    expressions of the form (\texttt{oplus $\alpha$ $\beta$}).
-
-    A first set of lemmas correspond to the various cases of \texttt{oplus}'s 
-    definition. They can be proved almost immediately. Here are a few examples.
-
-    \inputsnippets{Hessenberg/oplusNeutral,
-      Hessenberg/oplusCompareLt}
-
-    % \subsubsection{A hand-made induction principle}
-    % \label{sec:orgheadline85}
-
-    % \index{Coq!Commands!Functional Scheme}
-
-    % \emph{Coq} contains a command  \texttt{Functional Scheme} that 
-    % generates induction principles which correspond to recursive functions.
-    % Unfortunately, the current version ( \texttt{8.11.0} ) doesn't work on \texttt{oplus},
-    % probably because of the inner \texttt{Fix}.
-
-    % \begin{Coqsrc}
-    % Functional Scheme oplus_ind := Induction for oplus Sort Prop.
-    % \end{Coqsrc}
-
-    % \begin{Coqanswer}
-    % Error: Anomaly "todo." Please report at http://coq.inria.fr/bugs/.
-    % \end{Coqanswer}
-
-    % Fortunately, it's a good exercise for a semi-experienced user, to write
-    % her/him-self induction principles similar to the ones returned by
-    % \texttt{Functional Scheme}.
-
-    % \begin{itemize}
-    % \item First, we choose to write a version for sort \texttt{Type}, since versions
-    % for sorts \texttt{Prop} and \texttt{Set} can be easily derived from
-    % the former one. According to \emph{Coq}'s naming politics, we will call our 
-    % principle \texttt{oplus\_rect}
-
-    % \item The conclusion of \texttt{oplus\_rect} will be (\texttt{$P$ a b (oplus a b)}),
-    % where $P$ is an arbitrary function of type 
-    % \texttt{T1 -> T1 -> T1 -> Type}
-
-    % \item The premises of \texttt{oplus\_rect} will describe how to build an induction 
-    % on the graph of \texttt{oplus}.
-    % \end{itemize}
-
-    % We are now ready to state and prove \texttt{oplus\_rect}, and the reader
-    % will note that the statement is longer than the proof script itself,
-    % which is a standard proof by induction, simplification and case-analysis 
-    % that follows  \texttt{oplus}'s definition.
-
-    % We associate also a tactic to the application of \texttt{oplus\_rect}.
-
-    % \begin{Coqsrc}
-    %  Lemma oplus_rect:
-    %       forall P: T1 -> T1 -> T1 -> Type, 
-    %         (forall a:T1, P zero a a) ->
-    %         (forall a: T1, P a zero a) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Gt ->
-    %            P b1 (cons a2 n2 b2) o ->
-    %            P (cons a1 n1 b1) (cons a2 n2 b2)
-    %              (cons a1 n1 o)) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Lt ->
-    %            P (cons a1 n1 b1) b2 o ->
-    %            P (cons a1 n1 b1) (cons a2 n2 b2) 
-    %            (cons a2 n2 o)) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Eq ->
-    %            P b1 b2 o ->
-    %           P (cons a1 n1 b1) (cons a2 n2 b2)
-    %             (cons a1 (S (n1 + n2)%nat) o)) ->
-    %          forall a b, P a b (oplus a b).
-    % Proof with auto.
-    %    induction a.
-    %    -    intro; simpl; destruct b;auto.
-    %    -   induction b.
-    %        + apply X0.
-    %        + case_eq (compare a1 b1).
-    %          * intro Comp; unfold oplus; rewrite Comp.
-    %            cbn; apply X3 ...
-    %          * intro Comp; cbn; rewrite Comp; apply X2...
-    %          * intro Comp; cbn; rewrite Comp ...
-    %  Defined.
-
-    % Ltac oplus_induction a b:= pattern (oplus a b); apply oplus_rect.
-    % \end{Coqsrc}
-
-    % \index{Exercises}
-
-    % \begin{exercise}
-    % The induction principle \texttt{oplus\_rect} is still unused in our development. 
-    % Please build some nice examples of application.
-    % \end{exercise}
-
-    \index{hydras}{Projects}
-    \begin{project}
-    Compare \texttt{oplus}'s definition (with inner fixpoint) with other possibilities
-    (\texttt{coq-equations}, \texttt{Function}, etc.).
-    \end{project}
-    \subsection{More theorems on Hessenberg's sum}
-
-    We need to prove some properties of $\oplus$, particularly about 
-    its relation with the order $<$ on \texttt{T1}.
-
-    \subsubsection{Commutativity, associativity}
-
-    We prove  the commutativity of $\oplus$ in two steps. 
-    First, we prove by transfinite induction on $\alpha$ that the restriction of $\oplus$ to the
-    interval $[0..\alpha)$ is commutative.
-
-    \index{maths}{Transfinite induction}
-
-    \input{movies/snippets/Hessenberg/oplusComm0}
-
-    Then, we infer  $\oplus$'s commutativity for any pair of ordinals:
-    Let $\alpha$ and $\beta$ be two ordinals strictly less than $\epsilon_0$. Both ordinals $\alpha$ and $\beta$ are
-    strictly less than $\textrm{max}(\alpha,\beta)+1$.
-        Thus, we have just to apply the lemma \coqsimple{oplus\_comm\_0}.
-
-\input{movies/snippets/Hessenberg/oplusComm}
-
-    Associativity of Hessenberg's sum is proved the same way.
-
-    \input{movies/snippets/Hessenberg/oplusAssoc0}
-    \input{movies/snippets/Hessenberg/oplusAssoc}
-
-    \subsubsection{Monotony}
-
-    At last, we prove that $\oplus$ is strictly monotonous in both of its arguments.
-
-    \input{movies/snippets/Hessenberg/oplusMono}
-
-      \index{hydras}{Projects}
-
-    \begin{project}
-    The library \texttt{Hessenberg} looks too long (proof scripts and compilation).
-    Please try to make it simpler and more efficient!
-    Thanks!
-    \end{project}
-
-    \paragraph*{\gaiasign} The module
-      \href{../theories/html/gaia_hydras.~GHessenberg.html}{gaia\_hydras.GHessenberg} defines a version of Hessenberg's sum
-      compatible with \gaia's type \texttt{T1}.
-
-    \subsection{A termination measure for hydra battles }
-
-    \label{sec:hydra-measure}
-
-    Let us define a measure from type \texttt{Hydra} into \texttt{T1}.
-
-    \vspace{4pt}
-    \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Termination.html\#m}{Hydra.Hydra\_Termination}}
-
-   \input{movies/snippets/Hydra_Termination/mDef}
-
-    First, we prove that the measure $m(h)$  of any hydra $h$ is a well-formed ordinal term of type \texttt{T1}.
-
-    \input{movies/snippets/Hydra_Termination/mNf}
-      \input{movies/snippets/Hydra_Termination/msNf}
-
-    For proving the termination of all hydra battles, we have to prove that
-    \texttt{m} is a variant. First, a few technical lemmas follow the decomposition of \texttt{round} into several relations. Then the lemma \texttt{round\_decr} gathers all the cases.
-
-    \label{sect:variant-decr}
-
-    \input{movies/snippets/Hydra_Termination/S0Decr}
-    \input{movies/snippets/Hydra_Termination/R1Decr}
-    \input{movies/snippets/Hydra_Termination/S1Decr}
-    \input{movies/snippets/Hydra_Termination/R2Decr}
-     \input{movies/snippets/Hydra_Termination/RoundDecr}
-
-      Finally, we prove termination of all (free) battles.
-
-    \label{thm:every-battle-terminates}
-
-    \input{movies/snippets/Hydra_Termination/FinalThm}
-
-    \paragraph*{\gaiasign}
-    The module~\href{../theories/html/gaia_hydras.GHydra.html}{gaia\_hydras.GHydra} contains a proof of \gaiaHydras version of termination of all battles.
-
-    \inputsnippets{GHydra/mDef, GHydra/mVariant, GHydra/Termination}
-
-    \section*{Conclusion}
-
-    Let us recall three results we have proved so far.
-    \begin{itemize}
-    \item There exists a strictly decreasing variant which maps \texttt{Hydra} into 
-    the segment $[0,\epsilon_0)$ for proving the termination of any hydra battle
-    \item There exists \emph{no} such variant from \texttt{Hydra} into 
-    $[0,\omega^2)$, \emph{a fortiori} into $[0,\omega)$.
-    \end{itemize}
-
-    So, a  natural question is `` Does there exist any strictly decreasing variant mapping
-    type \texttt{Hydra} into some interval $[0,\alpha)$ (where $\alpha <\epsilon_0$) for proving the termination of all hydra battles''.
-The next chapter is dedicated to a formal proof that there exists no such $\alpha$, even if we consider a restriction to the set of ``standard'' battles.
-
-    %\include{epsilon0}
-
-    %\include{impossibility-proofs}
-
-%-------------------------------------------------------------------
-
-\chapter[The Ketonen-Solovay machinery]{Accessibility inside \texorpdfstring{$\epsilon_0$}{Epsilon0}: The Ketonen-Solovay Machinery\label{ks-chapter}}
-\label{chap:ketonen}
-\index{maths}{Ordinal numbers!Ketonen-Solovay machinery}
-
-\section{Introduction}
-The reader may think that our proof of termination in the previous  chapter requires a lot of mathematical tools and may be too  complex. So, the question is ``is there  any  simpler proof'' ?
-
-In their article~\cite{KP82}, Kirby and Paris show that this result cannot be proved in Peano arithmetic.
-Their proof uses some knowledge about model theory and non-standard models of Peano arithmetic.
-In this chapter, we focus on a specific class of proofs of termination of hydra battles: construction of some variant mapping the type \texttt{Hydra} into a given initial  segment of ordinals.
-Our proof relies only on the Calculus of Inductive Constructions and is a natural complement of the results proven in the previous chapters.
-
-\begin{itemize}
-\item There is no variant mapping the type \texttt{Hydra} into the interval $[0,\omega^2)$ (section ~\vref{omega2-case}), and \emph{a fortiori} $[0,\omega)$ (section ~\vref{omega-case}).
-
-\item There exists a variant which maps the type \texttt{Hydra} into the
-interval $[0,\epsilon_0)$ (theorem \texttt{every\_battle\_terminates}, in section~\vref{thm:every-battle-terminates}).
-\end{itemize}
-
-Thus, a very natural question is the following one:
-\begin{quote}
-  `` Is there  any variant from
-\texttt{Hydra} into some interval $[0,\mu)$, where $\mu<\epsilon_0$, for proving the termination of all hydra battles ?''
-\end{quote}
-
-We prove in \coq{} the following result:
-
-\begin{quote}
-There is no variant for proving the termination of all hydra battles
-from \texttt{Hydra} into the interval $[0..\mu)$, where
-$\mu< \epsilon_0$.
-The same impossibility holds even if we consider only standard battles (with the successive replication factors $0,1,2,\dots,t,t+1,\dots$).
-\end{quote}
-
-Our proofs are  constructive and require no axioms: they are  closed terms of the CIC, and are mainly composed on function definitions and proofs of properties of these functions. 
-They  share much theoretical material with Kirby and Paris', although they do not use any knowledge about Peano arithmetic nor model  theory.  The combinatorial arguments we use and implement
-come from 
- an article by J.~Ketonen and R.~Solovay~\cite{KS81}, already  cited in the work
- by L.~Kirby and J.~Paris. % on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
- Section $2$ of this article: ''A hierarchy of probably recursive functions'', contains a systematic study of \emph{canonical sequences}, which are closely related to
-rounds of hydra battles. 
-Nevertheless, they have the same global structure as the simple proofs described in
-sections~\vref{omega-case} and \vref{omega2-case}. 
-We invite the reader to compare the three proofs step by step, lemma by lemma.
-
-\section{Canonical Sequences}
-\label{ketonen-solovay-sect}
-\index{maths}{Ordinal numbers!Canonical sequences}
-
-Canonical sequences are functions that associate an ordinal $\canonseq{\alpha}{i}$ to every ordinal $\alpha<\epsilon_0$ and positive integer $i$. They satisfy several nice properties:
-
-\index{maths}{Transfinite induction}
-\begin{itemize}
-\item If $\alpha\not=0$, then $\canonseq{\alpha}{i}<\alpha$. Thus canonical sequences can be used for proofs by transfinite induction or function definition by transfinite recursion
-\item If $\lambda$ is a limit ordinal, then $\lambda$ is the least upper bound of the set 
-$\{\canonseq{\lambda}{i}\;|\,i\in\mathbb{N}_1\}$
-
-\item If $\beta<\alpha<\epsilon_0$, then there is a ``path'' from $\alpha$ to $\beta$, \emph{i.e.} a
-sequence $\alpha_0=\alpha, \alpha_1, \dots, \alpha_n=\beta$, where for every $k<n$, there exists some $i_k$ such that $\alpha_{k+1}=\canonseq{\alpha_k}{i_k}$
-\item Canonical sequences correspond tightly to rounds of hydra battles: if $\alpha\not=0$,
-then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1})$ in one round with
-the replication factor $i$ (Lemma \href{../theories/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
-\item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) battle from $\iota(\alpha)$ to $\iota(\beta)$.
-\end{itemize}
-
-\begin{remark}
-  In~\cite{KS81}, canonical sequences are defined for any ordinal $\alpha <\epsilon_0$,
-by stating that if $\alpha$ is a successor ordinal $\beta+1$,  the sequence associated with 
-$\alpha$ is simply the constant sequence whose terms are equal to $\beta$.
-Likewise, the canonical sequence of $0$ maps any natural number to $0$.
-
-This convention allows us to make total the function that maps any ordinal $\alpha$ and natural number $i$ to the ordinal $\canonseq{\alpha}{i}$.
-
-\end{remark}
-
-First, let us recall how canonical sequences are defined in~\cite{KS81}.
-For efficiency's sake, we decided not to implement directly K.\&S's definitions, but to define in \gallina{} simply typed structurally recursive functions which share the abstract properties which are used in the mathematical proofs\footnote{With a small difference: the $0$-th term of the canonical sequence is not the same in our development as in~\cite{KS81}.}.
-
-\subsubsection{Mathematical definition of canonical sequences}
-
-In~\cite{KS81} the definition of $\canonseq{\alpha}{i}$ is based on the following remark:
-\begin{quote}
-Any non-zero ordinal $\alpha$ can be decomposed in a unique way as the product
-$\omega^\beta\times (\gamma+1)$.
-\end{quote}
-
-Thus the $\canonseq{\alpha}{i}$\,s are defined in terms of this decomposition:
-\begin{definition}[Canonical sequences: mathematical definition]
-\label{def:canonseq-math}
-
-\end{definition}
-\begin{mathframe}
-  \begin{itemize}
-\item Let $\lambda<\epsilon_0$ be a limit ordinal
-
-\begin{itemize}
-\item If $\lambda=\omega^{\alpha+1}\times (\beta+1)$, then 
-$\canonseq{\lambda}{i}= \omega^{\alpha+1}\times\beta +  \omega^\alpha \times i$
-\item If $\lambda=\omega^{\gamma}\times (\beta+1)$, where $\gamma<\lambda$ is a limit ordinal, then 
-$\canonseq{\lambda}{i}=\omega^{\gamma}\times \beta + \omega^{\canonseq{\gamma}{i}}$
-\end{itemize}
-
-\item For successor ordinals, we have $\canonseq{\alpha+1}{i}= \alpha$
-
-\item Finally, $\canonseq{0}{i}= 0$.
-\end{itemize}
-\end{mathframe}
-
-\subsubsection{Canonical sequences in \coq}
-\index{hydras}{Library Epsilon0!Functions!canon}
-\index{hydras}{Library Epsilon0!Functions!canonS}
-
-Our definition may look more complex than the mathematical one, but
-uses plain structural recursion over the type \coqsimple{T1}. Thus, tactics like
-\coqsimple{cbn}, \coqsimple{simpl}, \coqsimple{compute}, etc., are applicable.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canon}{Epsilon0.Canon}}
-
-\label{Functions:canonS}
-\label{Functions:canon}
-
-\input{movies/snippets/Canon/canonDef}
-
-\paragraph*{\gaiasign} Our library~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon} transposes the  function \texttt{canon} into \gaia's type \texttt{T1}
-(please see Sect.\~ref{sect:gcanon}).
-
-\paragraph*{Remark}
-In the present state of this library, the following specializations of \texttt{canon} are still used in some proofs or lemma statements. They are planned to be deprecated.
-
-\input{movies/snippets/Canon/CanonS0}
-
-For instance \coq's computing facilities allow us to verify the equalities\linebreak 
-\mathcolor{$\canonseq{\omega^\omega}{3} = \omega^3$} and
-\mathcolor{$\canonseq{\omega^\omega*3}{42} = \omega^\omega*2 + \omega^{42}$}.
-
-\input{movies/snippets/Canon/canonExamples}
-
-% \index{hydras}{Projects}
-% \begin{project}
-% Many lemmas presented in this chapter were stated and proved before the introduction of 
-% the type class \texttt{ON} of ordinal notations, and in particular its  instance \texttt{Epsilon0}.
-% Thus definitions and lemmas refer to the type \texttt{T1} of possibly not well-formed terms.
-% This should be fixed in  a future version.
-% \end{project}
-
-\subsection{Basic properties of canonical sequences}
-
-We did not  try to prove that our definition truly implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
-abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
-~\href{../theories/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}.
-For instance, the equality $\canonseq{\alpha+1}{i}=\alpha$  can be  proved by  structural induction on $\alpha$.
-
-\input{movies/snippets/Canon/canonSucc}
-
-\subsubsection{Canonical sequences and the order $<$}
-
-\index{maths}{Transfinite induction}
-
-We prove by transfinite induction over $\alpha$ that $\canonseq{\alpha}{i}$ is an ordinal strictly less than $\alpha$ (assuming $\alpha\not=0$).
-This property allows us to use the function \texttt{canonS} and its derivatives in function definitions by transfinite recursion.
-
-\label{lemma:canon_LT}
-
-\inputsnippets{Canon/canonLT}
-
-\paragraph*{\gaiasign} This lemma is also available in
-Library ~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon}:
-
-\inputsnippets{GCanon/gcanonLt}
-
-\subsubsection{Limit ordinals are truly limits}
-The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
-is the limit of the sequence $\canonseq{\lambda}{i}\;(1\le i)$.
-
-\vspace{4pt}
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}
-
-\input{movies/snippets/Canon/canonSLimitStrong}
-
-\label{lemma:canonS-limit}
-
-Note the use of \coq's \texttt{sig} type in the theorem's statement, which
-relates the boolean function \texttt{limitb} defined on the \texttt{T1} data-type with a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.
-We can also state directly that $\lambda$ is a (strict) least upper bound of the elements of its canonical sequence.
-
-\input{movies/snippets/Canon/canonSLimitLub}
-
-\paragraph*{\gaiasign}  In \gaiaHydras, the statement use a slightly different vocabulary:
-\inputsnippets{GCanon/gcanonLimitOf}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}\label{exo:simply-typed-canonseq}
-Instead of using the \texttt{sig} type, define a simply typed function that, given two ordinals $\alpha$ and $\beta$, returns a natural number $i$ such that, if $\alpha$ is a limit ordinal and $\beta<\alpha$, then $\beta< \canonseq{\alpha}{i+1}$.
-Of course, you will have to prove the correctness of your function.
-
-\textbf{Hint:} You may add to your function a third argument usually called \texttt{fuel} for allowing you to give a structurally 
-recursive function (\emph{cf} the post of Guillaume Melquiond on Coq-club (Dec 21, 2020)
-\url{https://sympa.inria.fr/sympa/arc/coq-club/2020-12/msg00069.html}).
-The type \texttt{fuel}, an alternative 
-to \texttt{nat} is available on \href{../theories/html/hydras.Prelude.Fuel.html}{Prelude.Fuel})
-.
-%\index{coq}{Giving fuel to a long computation}
-
-\end{exercise}
-
-\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0}\,: paths}
-\index{maths}{Ordinal numbers!Accessibility inside epsilon0}
-\label{sect:pathes-intro}
-
-Let us consider a kind of accessibility problem inside $\epsilon_0$: given two ordinals $\alpha$ and $\beta$, where $\beta<\alpha<\epsilon_0$, find a \emph{path} consisting of a finite sequence $\gamma_0=\alpha,\dots,\gamma_l=\beta$,
-where, for every $i<l$, $\gamma_i \not= 0$ \footnote{This condition allows us to ignore paths which end by a lot of useless $0$s.} and there exists some strictly positive integer $s_i$
-such that $\gamma_{i+1}=\canonseq{\gamma}{s_i}$.
-
-Let $s$ be the sequence $\langle s_0,s_1,\dots, s_{l-1} \rangle$. We describe the
-existence of such a path with the notation $\alpha\xrightarrow [s]{}\beta$.
-We say also that the considered path from $\alpha$ to $\beta$ \emph{starts at [index] $s_0$ and ends at $s_l$}.
-
-For instance, we have $\omega*2 \xrightarrow[2,2,2,4,5]{}3$, through the 
-path $\langle\omega\times 2, \omega+2,\omega+1,\omega,4,3\rangle$.
-
-\begin{remark}
-
-Note that, given $\alpha$ and $\beta$, where $\beta < \alpha$, the sequence $s$ which leads from $\alpha$ to $\beta$ is not unique.
-Indeed, if $\alpha$ is a limit ordinal, the first element of $s$ can be any integer $i$ such that $\beta<\canonseq{\alpha}{i}$, and if $\alpha$ is a successor ordinal,
-then the sequence $s$ can start with any positive integer.
-
-For instance, we have also 
-$\omega*2 \xrightarrow[3,4,5,6]{}\omega$. 
-Likewise,
-$\omega*2 \xrightarrow[1,2,1,4]{} 0$ and
-$\omega*2 \xrightarrow[3,3,3,3,3,3,3,3]{} 0$.
-\end{remark}
-
-\subsection{Formal definition}
-
-\label{path-to-definition}
-
-In \coq{}, the notion of path can be simply defined as an inductive predicate 
-parameterized by the destination $\beta$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\index{hydras}{Library Epsilon0!Predicates!path\_to}
-\label{sect:path-to-def}
-
-\input{movies/snippets/Paths/transitionDefs}
-
-\input{movies/snippets/Paths/pathDef}
-
-\begin{remark}
-In the present version of our library, we use a variant \texttt{path\_toS} of
-\texttt{path\_to}, where the proposition
-(\texttt{path\_toS $\beta$ $s$ $\alpha$}) is equivalent to
-(\texttt{path\_to $\beta$ (List.map S $s$) $\alpha$}). This variant is scheduled to be deprecated.
-\end{remark}
-
-\paragraph*{\gaiasign}
-The library~\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} transposes the  notion of path into \gaia's type \texttt{T1}
-(please see Sect.\~ref{sect:gpath}).
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Write a tactic for solving goals of the form (\texttt{path\_to $\beta$ $s$ $\alpha$})
-where $\alpha$, $\beta$ and $s$ are closed terms. 
-You should solve automatically the following goals:
-
-\inputsnippets{Paths/exPathTac} 
-\end{exercise}
-
-\subsection{Existence of a path}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction on $\alpha$, we prove that for any $\beta<\alpha$, 
-one can build a path from $\alpha$ to $\beta$ (in other terms, $\beta$ is accessible from $\alpha$).
-
-\input{movies/snippets/Paths/LTPathTo}
-
-\noindent 
-From the lemma \texttt{canon\_LT}~\vref{lemma:canon_LT}, we can convert any path into an inequality on ordinals (by induction on paths).
-
-\input{movies/snippets/Paths/pathToLT}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}[continuation of exercise~\vref{exo:simply-typed-canonseq}]
-Define a simply typed function for computing a path from $\alpha$ to $\beta$.
-\end{exercise}
-
-\subsection{Paths and hydra battles}
-\label{KS-o2h}
-
-In order to apply our knowledge about  ordinal numbers less than $\epsilon_0$ to the study of hydra battles, we define an injection
-from the interval $[0,\epsilon_0)$ into the type \texttt{Hydra}.
-
-\vspace{4pt}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}}
-
-\input{movies/snippets/O2H/iotaDef}
-
-For instance Fig.~\ref{fig:iota-example} shows the image by $\iota$ of the ordinal  \textcolor{black}{$\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$}
-
-  \begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.3]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (10,2) {$\bullet$};
-\node (N22) at (7,2) {$\bullet$};
-\node (N3) at (14,2) {$\bullet$};
-\node (N4) at (18,2) {$\Smiley[2][green]$};
-\node (N5) at (0,4) {$\bullet$};
-\node (N6) at (2,5) {$\Smiley[2][green]$};
-\node (N7) at (4,6) {$\Smiley[2][green]$};
-\node (N88) at (7,4) {$\bullet$};
-\node (N8) at (10,4) {$\bullet$};
-\node (N9) at (14,6) {$\Smiley[2][green]$};
-\node (N10) at (0,8) {$\Smiley[2][green]$};
-\node (N11) at (10,7) {$\Smiley[2][green]$};
-\node (N111) at (7,7) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N22);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (N1) to  (N5);
-\draw (N1) to   [bend left=10] (N6);
-\draw (N1) to   [bend right=20] (N7);
-\draw (N2) to  (N8);
-\draw (N22) to  (N88);
-\draw (N8) to  (N11);
-\draw (N88) to  (N111);
-\draw (N3) to  (N9);
-\draw (N5) to  (N10);
-\end{tikzpicture}
-\caption{The hydra $\iota(\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$) \label{fig:iota-example}}
-
-\end{figure}
-
-The following lemma (proved in ~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H.v}) maps  canonical sequences to rounds of hydra battles.
-
-\label{lemma:canonS-iota}
-
-\input{movies/snippets/O2H/canonSIota}
-
-The next step of our development extends this relationship to
-the order $<$ on $[0,\epsilon_0)$ on one side, and hydra battles on the other side.
-
-\input{movies/snippets/O2H/pathToRoundPlus}
-
-As a corollary, we are now able to transform any inequality $\beta<\alpha<\epsilon_0$ into a (free) battle.
-
-\input{movies/snippets/O2H/LTToRoundPlus}
-
-\section{A  proof of impossibility}
-
-We now have  the tools for proving that  there exists no variant bounded by some $\mu<\epsilon_0$ for proving the termination   of all battles.
-The proof we are going to show is a proof by contradiction.
-It  can
- be considered as a generalization of the
-proofs described in  sections~\vref{omega-case} and \vref{omega2-case}.
-
-In the module
-\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
-This part of the development is parameterized by some class $B$ of battles, which will be instantiated later to \texttt{free} or \texttt{standard}.
-
-\vspace{4pt}
-\noindent
-From~\href{../theories/html/hydras.Hydra.~Hydra_Definitions.html}{Hydra.Hydra\_Definitions}:
-
-\input{movies/snippets/Hydra_Definitions/BoundedVariant}
-
-Let us assume there exists such a variant:
-
-\input{movies/snippets/Epsilon0_Needed_Generic/theContext}
-
-\label{remark:m-decrease}
-\begin{remark}
-  The hypothesis \texttt{m\_decrease} is not provable  in general, but is satisfied by
-the  \texttt{free} and \texttt{standard} kinds of battles. This trick allows to 
-``factorize'' our proofs  of impossibility.
-\end{remark}
-
-\index{maths}{Transfinite induction}
-
-First, we prove that $m(\iota(\alpha))$ is always greater than or equal to $\alpha$, by  transfinite induction over $\alpha$.
-
-\input{movies/snippets/Epsilon0_Needed_Generic/mGe0}
-
-\begin{itemize}
-\item If $\alpha=0$, the inequality trivially holds
-\item If $\alpha$ is the successor of  some ordinal $\beta$, the inequality $\beta \leq m(\iota(\beta))$ holds (by induction hypothesis).
-But the hydra $\iota(\alpha)$ is transformed in one round into
-$\iota(\beta)$, thus $m(\iota(\beta))<m(\iota(\alpha))$. Hence $\beta<m(\iota(\alpha))$, which implies $\alpha \leq m(\iota(\alpha))$
-\item If $\alpha$ is a limit ordinal, then $\alpha$ is the least upper bound of the set
-of all  the $\canonseq{\alpha}{i}$.  Thus, we have just to prove that $\canonseq{\alpha}{i}< m(\iota(\alpha))$ for any $i$. 
-\begin{itemize}
-\item Let $i$ be some natural number.
-By the induction hypothesis, we have $\canonseq{\alpha}{i} \leq m(\iota(\canonseq{\alpha}{i}))$.
-But the hydra $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i})$ in one round, thus $m(\iota(\canonseq{\alpha}{i})) < m(\iota(\alpha))$, by our hypothesis \texttt{m\_decrease}.
-\end{itemize}
-\end{itemize}
-
-Please note that the impossibility proofs of 
-sections~\vref{omega-case} and \vref{omega2-case} contain a similar lemma, also called \texttt{m\_ge}.
-We are now able to build a counter-example.
-
-\input{movies/snippets/Epsilon0_Needed_Generic/mGeGeneric}
-
-The (big) rest of the proof is dedicated to prove formally the converse inequality 
-\texttt{m small\_h t1< m big\_h}.
-
-\subsection{The case of free battles}
-\label{sec:free-battles-case}
-Let us now consider that $B$ is instantiated to \texttt{free} (which means that we are considering proofs of termination of \emph{all} battles).
-The following lemmas are proved in Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
-The case $B=\texttt{standard}$ is studied in section~\vref{std-case}.
-
-\input{movies/snippets/Epsilon0_Needed_Free/theContext}
-
-\begin{enumerate}
-\item The following lemma is an application of \texttt{m\_ge\_generic}, since \texttt{free}
-satisfies trivially the hypothesis \texttt{m\_decrease} (see page~\pageref{remark:m-decrease}).
-
-\inputsnippets{Epsilon0_Needed_Free/mGe}
-
-\item From the hypothesis \texttt{Hy}, we have \texttt{m big\_h t1< mu}
-\item By Lemma \texttt{LT\_to\_round\_plus}, we get a (free) battle from
-\texttt{big\_h = iota mu} to \texttt{small\_h = iota (m big\_h)}.
-
-\input{movies/snippets/Epsilon0_Needed_Free/bigToSmall}
-
-\item From the hypotheses on $m$, we infer:
-\input{movies/snippets/Epsilon0_Needed_Free/mLt}
-
-\item From lemmas \texttt{m\_ge} and \texttt{m\_lt}, and the irreflexivity of $<$, we get a contradiction. 
-\input{movies/snippets/Epsilon0_Needed_Free/ImpossibilityFree}
-
-\end{enumerate}
-
-We have now proved there exists no bounded variant for the class of free battles.
-
-\input{movies/snippets/Epsilon0_Needed_Free/CheckDemo}
-
-\section{The case of standard battles}
-\label{sec:standard-intro}\label{std-case}
-One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
-Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}.
-
-\input{movies/snippets/O2H/LTToRoundPlus}
-
-This lemma builds a battle out of any inequality $\beta<\alpha$. 
-It is a straightforward application of \texttt{LT\_path\_to} of
-Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}:
-
-\input{movies/snippets/Paths/LTPathTo}
-
-The sequence $s$, used to build the sequence of replication factors of the battle, depends on 
-$\alpha$ and $\beta$, so we cannot be sure that the generated battle is a genuine standard battle.
-
-The solution to this issue comes  once again from Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
-$\alpha_0=\alpha,\alpha_1,\dots,\alpha_k=\beta$ where $\alpha_{j+1}$ is equal
-to $\canonseq{\alpha_j}{i_j}$ where $i_j$ is \emph{any} natural number, 
-we consider various constraints on these sequences.
-In particular, a path is called \emph{standard} if $i_{j+1} = i_j + 1$ for every $j<k$.
-It  corresponds to a ``segment'' of some standard battles. 
-Please note that the vocabulary on paths is ours, but all the concepts come really from~\cite{KS81}.
-
-In \coq{}, standard paths can be defined as follows.
-
-\vspace{4pt}
-
-\emph{From
-Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/standardPathR}
-
-In the mathematical text and figures, we shall use the notation 
-$\alpha \xrightarrow[i,j]{}\beta$ for the proposition 
-(\texttt{standard\_path $i$ $\alpha$ $j$ $\beta$}).
-In~\cite{KS81} the notation is
-$\alpha \xrightarrow[i]{*}\beta$
-for 
-the proposition  $\exists j, i<j \wedge \alpha \xrightarrow[i,j]{} \beta$.
-
-Our goal is now  to transform any inequality $\beta<\alpha<\epsilon_0$ into a standard path $\alpha \xrightarrow[i,j]{} \beta$ for some $i$ and $j$, then into a standard battle
-from $\iota(\alpha+i)$ to $\iota(\beta)$. 
-Following~\cite{KS81}, we proceed in two stages:
-\begin{enumerate}
-\item we simulate plain (free) paths from $\alpha$ to $\beta$ with
-paths made of steps $(\gamma,\canonseq{\gamma}{n})$, \emph{with the same $n$ all along the path}
-\item we simulate any such path by a standard path.
-\end{enumerate}
-
-\subsection{Paths with a constant index}
-
-First of all, paths with a constant index 
-enjoy nice properties. They are defined as paths where all the $i_j$ are equal to the same natural number $i$, for some $i>0$.
-
-Like in~\cite{KS81}, we shall use the notation $\alpha \xrightarrow[i]{} \beta$ for denoting such a path, also called an $i$-path.
-
-\input{movies/snippets/Paths/constPathDef}
-
-% Paths with a given index can be effectively computed.
-% Given $i$, $\alpha$ and $l$, the following function returns the ordinal $\beta$ such that there exists a path 
-% $\alpha \xrightarrow [i+1] {} \beta$ of length $l$.
-
-% \begin{Coqsrc}
-% Fixpoint const_funS (i:nat)(alpha : T1)(l:nat):  T1  :=
-%   match l
-%   with
-%   | 0 => alpha
-%   | S m => const_funS i (canonS i alpha) m
-%   end.
-% \end{Coqsrc}
-
-% The following computations show  applications of \texttt{constS\_fun} to the 
-% ordinal $\omega^\omega$, with various values of $i$ and $l$.
-
-% \begin{Coqsrc}
-% Compute  (const_funS 2 (omega ^omega)  55).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = zero
-%      : T1 
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (const_funS 2 (omega ^omega) 15).
-% \end{Coqsrc}
-
-%   \begin{Coqanswer}
-%  = (omega ^ 2 * 2)%pT1
-%      : ppT1   
-%   \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (const_funS 4 (omega^omega)  100).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% = (omega ^ 4 * 4 + omega ^ 3 * 4 + omega ^ 2 + omega * 4 + 4)%pT1
-%      : ppT1
-% \end{Coqanswer}
-
-A most interesting property of $i$-paths is that we can ``upgrade'' their index, as stated by K.\&S.'s Corollary 12.
-
-\index{maths}{Transfinite induction}
-
-\vspace{4pt}
-\noindent
-\emph{From
-  Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/Cor12}
-
-We  also use a version of \texttt{Cor12} with large inequalities.
-
-\input{movies/snippets/Paths/Cor121}
-
-\subsubsection{Sketch of proof of \texttt{Cor12}}
-\index{maths}{Transfinite induction}
-
-First, let us prove the following lemma (numbered \texttt{2.4} in Ketonen and Solovay's article).
-
-\input{movies/snippets/Paths/KSThm24}
-
-We prove this lemma by transfinite induction on $\alpha$.
-Let us consider a path $\alpha \xrightarrow [i]{} \beta$ $(i>0)$. Its first step is
-the pair $(\alpha,\canonseq{\alpha}{i})$, We have $\canonseq{\alpha}{i}<\alpha$ and
-$\canonseq{\alpha}{i} \xrightarrow [i]{} \beta$. 
-Let $n$ be any natural number such that $n>i$.
-By the induction hypothesis, there exists a path $\canonseq{\alpha}{n} \xrightarrow[i]{} \beta$.
-\begin{itemize}
-\item  If $\alpha$ is a successor ordinal $\gamma+1$, then $\canonseq{\alpha}{n} =
-\canonseq{\alpha}{i}=\gamma$. Thus we have a path 
-$\alpha  \xrightarrow [n]{}  \gamma \xrightarrow [n]{} \beta$
-\item If $\alpha$ is a limit ordinal, we apply Theorem \texttt{KS\_thm\_2\_4} (see above).
-
-%   \begin{theorem}
-% Let $\lambda$ be a limit ordinal, then for any pair of indices $0<i<j$, there is a path $\canonseq{\lambda}{j} \xrightarrow[1]{} \canonseq{\lambda}{i}$.    
-%   \end{theorem}
-
-The proof of the limit case, is decomposed into a sequence
-of path constructions leading to $\alpha \xrightarrow[n]{} \beta$.
-
- \begin{enumerate}
- \item $\alpha \xrightarrow[n]{} \canonseq{\alpha}{n}$ (single step path)
- \item $\canonseq{\alpha}{n} \xrightarrow[1]{} \canonseq{\alpha}{i}$ (by \texttt{Theorem\_2\_4}),
-\item $\canonseq{\alpha}{n} \xrightarrow[n]{} \canonseq{\alpha}{i}$ (applying the induction hypothesis to the preceding path);
-\item $\canonseq{\alpha}{i} \xrightarrow[n]{} \beta$ (applying the induction hypothesis)
-\item $\alpha \xrightarrow[n]{} \beta$ (by composition of 1, 3, and 4).
-
- \end{enumerate}
-
-\end{itemize}
-
-\begin{remark}
- \texttt{Cor12} ``casts'' $i$-paths into $n$-paths for any $n>i$.
-But the obtained $n$-path can be much longer than the original $i$-path.
-The following exercise will give an idea of this increase. 
-\end{remark}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-  Prove that  the length of the $i+1$-path from
-  $\omega^\omega$ to $\omega^i$ is $1 + (i+1)^{(i+1)}$, for any $i$. Note that the $i$-path from
-  $\omega^\omega$ to $\omega^i$ is only one step long.
- \end{exercise}
-
-Why is \texttt{Cor12} so useful? 
-Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
-we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying corollary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
-where $n$ is the maximum of the indices $i$ met in the induction.
-
- Lemma 1, Section 2.6 of~\cite{KS81} is naturally expressed in terms of \coq's
-\verb@sig@ construct.
-
-\label{lemma:L-2_6-1}
-%\index{coq}{Sigma types}
-
-\input{movies/snippets/Paths/Lemma261}
-
-Intuitively, Lemma~\texttt{2\_6\_1}  shows that if $\beta<\alpha<\epsilon_0$, then there exists  a battle from $\iota(\alpha)$ to $\iota(\beta)$ where the replication factor is constant, although large enough.
-
- \paragraph*{\gaiasign}
- Corollary \texttt{Cor12} and Lemma \texttt{Lemma2\_6\_1} are also available in
-\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} 
-(please see Sect.~\vref{sect:gpath}).
-
-\subsection{Casting paths with a constant index into a standard path}
-
-The article~\cite{KS81} contains 
-the following lemma, which allows us to simulate $i$-paths by $[i+1,j]$-paths, where $j$ is large enough.
-
-\input{movies/snippets/Paths/constantToStandard}
-
-\subsubsection{Sketch of proof of \texttt{constant\_to\_standard\_path}}
-
-Our proof follows the proof by Ketonen and Solovay, including its organization as a sequence of lemma.
-Since it is a non-trivial proof, we will comment its main steps below   (see Figure~\vref{fig:belle-preuve-1} to Figure~\vref{fig:fin-belle-preuve}).
-
-\subsubsection*{Preliminaries}
-
-Please note that, given an ordinal $\alpha:\texttt{T1}$, and two natural numbers $i$ and $l$, there exists at most a standard path $\alpha \xrightarrow [i,i+l]{*} \beta$.
-The following function computes $\beta$ from $\alpha$, $i$ and $l$.
-
-\input{movies/snippets/Paths/standardGnaw}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction over  $\alpha$, we prove that the ordinal $0$ is reachable from any ordinal $\alpha<\epsilon_0$ by some standard path.
-
-\input{movies/snippets/Paths/standardPathToZero}
-
-\paragraph*{}
-Now, let us consider two ordinals  $\beta<\alpha<\epsilon_0$.  Let $p$  be some $(n+1)$-path from $\alpha$ to $\beta$.
-
-\input{movies/snippets/Paths/ConstantToStandardProof}
-
-Applying \texttt{standard\_path\_to\_zero}, $0$ is reachable from $\alpha$ by some standard path.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \draw[->, very thick,blue] (alpha)-- node [below]{$n+1$} node [above] {$+$} (beta);
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-    \node (alpha3) at (20,5) {};
-  \node (alphalast) at (35,5) {};
-  \node (zero) at (45,0) {$0$};
-  \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate=40]{$n+1$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha2)-- node [below]{$n+3$}  (alpha3);
-
-  \node (dots) at (24,5) {$\dots$};
-  \draw [->, dashed, very thick,blue] (alphalast)-- node [below, rotate=-26]{$n+p+1$}  (zero);
-
-\end{tikzpicture}
-\caption{A nice proof (1)}
-  \label{fig:belle-preuve-1}
-\end{figure}
-
-\paragraph*{}
-
-Since comparison on \texttt{T1} is decidable, one can compute the last step $\gamma$ of the standard path from $(\alpha,n+1)$  such that $\beta\leq \gamma$.
-Let $l$ be the length of the path from $\alpha$ to $\gamma$.  
-This step of the proof is illustrated in figure~\vref{fig:belle-preuve-2}.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (17,5) {$\ldots$};
-    \node (alpha3) at (20,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (38,0) {$\delta$};
-    \draw [->, dashed,very thick,blue] (alpha)-- node [below,rotate=35]{$n+1$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below,rotate = -48]{\tiny $n+l$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{$n+l+1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{$n+1$} node [above] {$+$} (beta);
-   \draw[thick] (alpha)--  (gamma);
-   \draw[thick] (gamma)--  node [above] {$\geq$} (beta);
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-\end{tikzpicture}
-
-\caption{A nice proof (2)}
-  \label{fig:belle-preuve-2}
-\end{figure}
-
-\paragraph*{}
-
-\begin{itemize}
-\item If $\beta=\gamma$, it's OK! We have got a standard path
-from  
-$\alpha$ to $\beta$ with successive indices  $n+1, n+2, \dots, n+l+1$
-
-\item Otherwise,  $\beta < \gamma$.  Let us consider  $\delta=\canonseq{\gamma}{n+l+1}$.
-By applying several times lemma \texttt{Cor12},  one converts  every path of Fig~\ref{fig:belle-preuve-2} into
- a $n+l+1$-path  (see figure~\ref{fig:belle-preuve-3}).
-
-But $\gamma$ is on the $n+l+1$-path from $\alpha$ to $\beta$.
-As shown by figure~\vref{fig:fin-belle-preuve}, the ordinal $\delta$, reachable from
-$\gamma$ in one single step,  must be greater than or equal to $\beta$, which contradicts our  hypothesis $\beta < \gamma$.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-    \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (17,5) {$\ldots$};
-    \node (alpha3) at (20,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (38,0) {$\delta$};
-     \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -48]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\tiny $n+l+1$} node [above]{\color{red} $1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
-    \draw[thick] (gamma)--   node [above]{\color{red} $>$}(beta);
-   \draw[thick] (alpha)--  (gamma);
-
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-
-\end{tikzpicture}
-
-\caption{A nice proof (3)}
-  \label{fig:belle-preuve-3}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (15,5) {$\ldots$};
-    \node (alpha3) at (18,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (42,0) {$\delta$};
-    \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -36]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\small $n+l+1$} node [above]{\color{red} $1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
-   \draw[thick] (alpha)--  (gamma);
-
-    \draw[thick] (gamma)--  node [below]{\tiny $n+l+1$} node [above]{\color{red} $+$}(beta);
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-
-\end{tikzpicture}
-
-\caption{A nice proof (4)}
-  \label{fig:fin-belle-preuve}
-\end{figure}
-
-\end{itemize}
- The only possible case is  thus $\beta=\gamma$, so we have got a standard path  from $\alpha$ to $\beta$.
-
- \input{movies/snippets/Paths/constantToStandard0}
- \input{movies/snippets/Paths/constantToStandardz}
-
-Here is the full statement of the conversion from constant to standard paths.
-
-\input{movies/snippets/Paths/constantToStandardPath}
-
-Applying \texttt{Lemma2\_6\_1} and \texttt{constant\_to\_standard\_path}, we get the following corollary.
-
-\input{movies/snippets/Paths/LTToStandardPath}
-
-\subsection{Back to hydras}
-\label{sec:standard-battles-cases}
-We are now able to complete our proof that there exists no bounded variant for proving the termination of standard hydra battles. This proof can
-be consulted in the module 
-\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}{Hydra.Epsilon0\_Needed\_Std}.
-Please note that it has the same global structure as in section~\ref{sec:free-battles-case} 
-% ICI !
-Applying the  lemmas  \texttt{Lemma2\_6\_1} of the module 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
-{Epsilon0.pathS}   and 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
-{\texttt{constant\_to\_standard\_path}},
-we can convert any inequality $\beta<\alpha<\epsilon_0$ into a standard path from
-$\alpha$ to  $\beta$, then into a fragment of a standard battle from 
-$\iota(\alpha)$ to $\iota(\beta)$, hence the inequality $m(\iota(\beta))<m(\iota(\alpha))$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html\#LT_to_standard_battle}{Hydra.Epsilon0\_Needed\_Std}}
-
-\input{movies/snippets/Epsilon0_Needed_Std/LTToStandardBattle}
-
-Now, please consider the following context:
-
-\input{movies/snippets/Epsilon0_Needed_Std/theContext}
-
-In the same way as for free battles, we import a large inequality 
-from 
-the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Epsilon0\_Needed\_Generic}.
-
-\input{movies/snippets/Epsilon0_Needed_Std/mGe}
-
-\paragraph*{} If remains to prove the following strict inequality, in order to have a contradiction.
-
-\input{movies/snippets/Epsilon0_Needed_Std/mLt}
-
-\paragraph*{Sketch of proof:} Let us recall that $\texttt{big\_h} = \iota(\mu)$
- and $\texttt{small\_h} = \iota (m (\texttt{big\_h}))$.
-
-Since $m(\texttt{big\_h})< \mu$, there exists a standard path from $\mu$ to
-$m(\texttt{big\_h})$, hence a   standard battle from $\iota(\mu)$  to
-$\iota(m(\texttt{big\_h}))$,  i.e. from \texttt{big\_h} to \texttt{small\_h}.
-
-Since $m$ is assumed to be a variant for standard battles, we get the inequality  $m(\texttt{small\_h}) < m(\texttt{big\_h})$.
-
-\input{movies/snippets/Epsilon0_Needed_Std/endOfProof}
-
-\subsection{Remarks}
-
-We are grateful to 
- J. Ketonen and R. Solovay  for the high quality of their explanations and proof details.
-Our proof follows tightly the sequence of lemmas in their article, with a focus on 
-constructive aspects.
-Roughly speaking, our implementation \emph{builds}, out of a hypothetical
-  variant $m$, bounded by some ordinal $\mu<\epsilon_0$, a hydra \texttt{big\_h} which verifies the impossible inequality  $m(\texttt{big\_h})< m(\texttt{big\_h})$.
-
-On may ask whether the preceding results are too restrictive, since they 
-refer to a particular data type \texttt{T1}.
-In fact, our representation of ordinals strictly less than 
- $\epsilon_0$ is faithful to their mathematical definition, at least 
-Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
-(please see also the module
-\href{../theories/html/hydras.Schutte.Correctness_E0.html}{hydras.Schutte.Correctness\_E0}).
-
-Thus, we can infer that our theorems can be applied to any well order.
-
-\index{hydras}{Projects}
-\begin{project}
-Study a possible modification of the definition of a variant  (for  standard battles).
-
-\begin{itemize}
-\item The variant is assumed to be strictly decreasing \emph{on configurations 
-reachable from some initial configuration where the replication factor is equal to $0$}
-\item The variant may depend on the number of the current round.
-\end{itemize}
-
-In other words, its type should be \texttt{nat -> Hydra -> T1}, and it must 
-verify the inequality $m\, (S\,i)\, h' < m\,i\, h$ whenever the configuration 
-$(i,h)$ is reachable from some initial configuration $(0,h_0)$
-and \texttt{h} is transformed into \texttt{h'} in the considered round.
-Can we still prove the theorems of section~\ref{std-case} with this new definition?
-
-\end{project}
-
-%---------------------------------------------------------------------
-\chapter{Large sets and rapidly growing functions}\label{chap:alpha-large}
-
-\begin{remark}
-Some notations (mainly names of fast-growing functions) of our development may differ slightly from the literature.
-Although this fact does not affect our proofs, we are preparing a future version where the names $F\_alpha$, $f\_alpha$, $H\_alpha$, etc., are fully consistent with the cited articles.
-
-\end{remark}
-%\section{Introduction}
-
-In this chapter, we try to feel how long a standard battle can be.
-To be precise, for any ordinal $\alpha<\epsilon_0$ and any positive integer $k$,
-we give a minoration of the number of steps of a standard battle which
-starts with the hydra $\iota(\alpha)$ and the replication factor $k$.
-
-We express this number in terms of the Hardy hierarchy of fast-growing 
-functions~\cite{BW85, Wainer1970, KS81, Promel2013}.
- From the \coq{} user's point of view, such  functions are  very 
-attractive:  they are defined as functions  in \gallina{}, and we can apply them \emph{in theory}, but they are so complex that you will never be able to look at the result of the computation.
- Thus, our knowledge on these functions must rely on \emph{proofs}, not tests. In our development, we use often the rewriting rules generated by \coq's \texttt{Equations} plug-in.
-
-\section{Definitions}
-
-%\subsection{Definition}
-
-\begin{definition}
-Let $0<\alpha<\epsilon_0$ be any ordinal, and $s=\langle s_1, s_2, \dots, s_N\rangle$ a finite sequence of strictly positive natural numbers.
-
-We say that $s$ is \emph{$\alpha$-large} if the sequence $\langle \alpha_0=\alpha,\dots,\alpha_{i+1}=\canonseq{\alpha_i}{i+1},\dots \rangle$ leads to $0$. 
-We say also that $s$ is \emph{minimally $\alpha$-large} (in short:
-\emph{$\alpha$-mlarge}) if $s$ is $\alpha$-large 
- and every strict prefix of $s$ leads to a non-zero ordinal (\emph{cf} Sect.~\vref{sect:path-to-def}).
-
-\index{maths}{Ordinal numbers!Large sets}
-\index{maths}{Ordinal numbers!Minimal large sets}
-
-\end{definition}
-
-\begin{remark}
-  Ketonen and Solovay~\cite{KS81} consider  large finite \emph{sets} of natural numbers,  but they are mainly used as sequences. Thus, we chose to represent them explicitely as (sorted) lists. 
-\end{remark}
-
-The following function ``gnaws'' an ordinal $\alpha$, following a sequence of indices (ignoring the $0$s).
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/gnawDef}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_SetsPaths.html\#gnaw}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/largeDef}
-
-Minimal large sequences can be directly defined in terms of the
-predicate \texttt{path\_to} (\vref{sect:path-to-def}) which already prohibits paths containing non-final \texttt{zero}s.
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}
-
-\index{hydras}{Library Epsilon0!Predicates!mlarge@mlarge (minimal large sequences)}
-
-\input{movies/snippets/Large_Sets/mlargeDef}
-
-Let us consider two integers $k$ and $l$, such that $0<k<l$. In order to check whether the interval $[k,l]$ is minimally large for $\alpha$, it is enough to
-follow from $\alpha$ the path associated with the interval $[k,l)$ and verify that the last ordinal we obtain is equal to $1$.
-
-\subsection{Examples}
-
-For instance the interval $[6,70]$ leads $\omega^2$ to $\omega\times 2 + 56$. Thus this interval is not $\omega^2$-large.
-
-\noindent
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html\#mlarge}{Epsilon0.Large\_Sets\_Examples}}
-
-\input{movies/snippets/Large_Sets_Examples/gnawEx1}
-
-The interval $[6,700]$ is $\omega^2$-large, but not
-$\omega^2$-mlarge, since $[6,699]$ is also $\omega^2$-large.
-
-\input{movies/snippets/Large_Sets_Examples/gnawEx2}
-
-The following lemma relates minimal largeness with the
-function 
-\texttt{gnaw}.
-
-\input{movies/snippets/Large_Sets/mlargeIff}
-
-\vspace{4pt}
- \noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html}{Epsilon0.Large\_Sets\_Examples}}
-
-\input{movies/snippets/Large_Sets_Examples/Ex1Lemma}
-
-\section{Length of minimal large sequences}
-
-Now, consider any natural number $k>0$ any ordinal $0<\alpha<\epsilon_0$.  We would like to compute
-a number $l$ such that the interval $[k,l]$ is $\alpha$-mlarge. So, 
-the standard battle starting with $\iota(\alpha)$ and the replication factor $k$ will end after $(l-k+1)$ steps.
-
-First, we notice that this  number $l$ exists, since the segment $[0,\epsilon_0)$ is well-founded and $\canonseq{\alpha}{i}<\alpha$ for any $i$ and $\alpha>0$.
-Moreover, it is unique:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/mlargeUnicity}
-
-Thus, we would like to define a function, parameterized by $\alpha$ which associates to any  strictly positive integer $k$ the number $l$ such that
-the interval $[k,l]$ is $\alpha$-mlarge. It would be fine to write in \gallina{} a definition like this:
-
-\begin{Coqbad}
-Function L_ (alpha: E0) (i:nat) :  nat := ...
-\end{Coqbad}
-
-But we do not know how to fill the dots yet \dots{}   In the next section, we will 
-use \coq{} to reason  about the \emph{specification} of \texttt{L},
-prove properties of any function which satisfies this specification.
-In Sect.~\ref{sect:L-equations}, we use the \texttt{coq-equations} plug-in
-to define a function \texttt{L\_}, and prove its correctness w.r.t. its specification.
-
-\subsection{Formal specification}
-
-Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider any  function which  maps  any strictly positive integer $k$ to the number $l$, where 
-the interval $[k,l)$ is $\alpha$-mlarge.
-
-\begin{remark}
-In~\cite{KS81} Ketonen and Solovay consider the least natural number $l$ where the interval $[k,l]$ ($l$ included) is $\alpha$-large, and call $H_\alpha$ the function which maps $k$ to $l$.
-We chose to consider intervals $[l,k)$ instead of $[l,k]$
-in order to simplify  some statements and proofs in composition lemmas associated with the ordinals of the form $\alpha\times i$ and 
-$\omega^\alpha\times i + \beta$.
-Clearly, both approaches are related through the equality
-$L_\alpha(k)=H_\alpha(k)+1$, for any non-null $\alpha$ and $k$.
-\end{remark}
-
-Our specification of the function \texttt{L} is as follows:
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/LSpecDef}
-
-\begin{todo}
- Check if the functions $L_\alpha$ are the same as
-\cite{KS81}' functions $f_\alpha$ (p. 297).
-\end{todo}
-
-Note that, for $\alpha\not=0$, the value of $f(0)$ is not specified.
-Nevertheless, the restriction of $f$ to the set of strictly positive integers is unique (up to extensionality).
-
-\input{movies/snippets/Large_Sets/LSpecUnicity}
-
-\subsection{Abstract properties}
-
-Let us now prove properties of any function $f$ (if any) which satisfies 
-\texttt{L\_spec}.
-We are looking for properties which could be used for writing \emph{equations} and prove the correctness of the function generated by the \texttt{coq-equations} plug-in.
-Moreover, they will give us some examples (for small values of $\alpha$).
-
-The properties we consider are defined in \href{../theories/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}.
-
-\label{sect:abstract-arith-prop}
-\input{movies/snippets/Iterates/funLeDef}
-
-Our exploration of the $L_\alpha$\,s  considers the usual cases of a proof by transfinite induction: zero, successors and limit ordinals.
-The lemmas we are going to prove will be applied in a big proof by induction in Sect~\vref{sect:L-correct-proof}.
-
-\index{maths}{Transfinite induction}
-
-\subsubsection{The  ordinal zero}
-\label{sect:L-spec-zero}
-The base case is directly a consequence of the specification.
-
-\input{movies/snippets/Large_Sets/LZeroInv}
-
-\subsubsection{Successor ordinals}
-\label{sect:L-spec-succ}
-Let $\beta$ be some ordinal, and assume the arithmetic function $f$ satisfies 
-the specification $(\texttt{L\_spec}\;\beta)$.  Let $k$ be any natural number.
-Any path from $\texttt{succ}\,\beta$ to $0$ starting at $k+1$ can be decomposed into a first step from $\texttt{succ}\,\beta$ to $\beta$, then a path from
-$\beta$ at $k+2$ to $0$. 
-By hypothesis the interval $[k+2, f(k+2)-1]$ is $\beta$-mlarge.
-But the interval $[k+1, f(k+2)-1]$ is the concatenation of the singleton
-$\{k+1\}$ and the interval $[k+2, f(k+2)-1]$.
-So, the function $\lambda\,k.\,f(k+1)$ satisfies the specification $\texttt{L\_spec}\,\beta$.
-
-Note that our decomposition of intervals works only if the intervals we consider are not empty.
-In order to ensure this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <<= f}, or \texttt{(fun\_le S f)}.
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/SectionSucc}
-
-\subsubsection{Limit ordinals}
-\label{sect:L-spec-lim}
-
-Let $\lambda<\epsilon_0$ be any limit ordinal. In a similar way as for successors, we decompose any path from $\lambda$  into a first step to
-$\canonseq{\lambda}{k}$, followed by a path to $0$. In the following section, we assume that there exists a correct function computing  $L_{\canonseq{\lambda}{k}}$ for any strictly positive $k$.
-
-\input{movies/snippets/Large_Sets/SectionLim}
-
-\subsection{First results}
-
-Applying the previous lemmas on successors and limit ordinals, 
-we obtain a few  correct implementations of \texttt{(L\_spec $\alpha$)} for small values of $\alpha$.
-
-\subsubsection{Finite ordinals}
-
-By iterating the functional \texttt{L\_succ}, we get a realization of
-\texttt{(L\_spec $i$)} for any natural number $i$.
-
-\input{movies/snippets/Large_Sets/LFinDef}
-\vspace{-16pt}
-\input{movies/snippets/Large_Sets/LFinOk}
-
-\subsubsection{The first limit ordinal  \texorpdfstring{$\omega$}{omega}}
-
-The lemmas \texttt{L\_fin\_ok} and \texttt{L\_lim\_ok}   allow us to get 
-by diagonalization a correct implementation for 
-\texttt{L\_spec omega}.
-
-\input{movies/snippets/Large_Sets/LOmegaDef}
-\vspace{-16pt}
-\input{movies/snippets/Large_Sets/LOmegaOk}
-
-\subsubsection{Towards  \texorpdfstring{$\omega^2$}{omega*omega}}
-
-We would like to get exact formulas for the ordinal $\omega^2$, a.k.a.
-$\phi_0(2)$. This ordinal is the limit of the sequence $\omega\times i\;(i \in \mathbb{N})$. Thus, we have to study ordinals of this form, then use 
-our lemma on limits.
-
-The following lemma establishes a path from $\omega\times ( i+1)$ to
-$\omega \times i$.
-
-\input{movies/snippets/Large_Sets/pathToOmegaMult}
-
-Let us consider a path from  $\omega\times(i+1)$ to $0$ starting at $k+1$.
-A first ``big step'' will lead to $\omega\times i$ at $2(k+1)$. If $i>0$, the
-next jump leads to $\omega\times(i-1)$ at $2(2(k+1))+1$, etc.
-
-The following lemma expresses the length of the mlarge sequences associated with any finite multiple of $\omega$.
-
-\input{movies/snippets/Large_Sets/omegaMultMlarge0}
-
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#L_omega_mult}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/LOmegaMultDef}
-
-More generally, we prove the equality $L_{\omega\times i}(k)=2^i\times(k+1)-1$.
-
-\input{movies/snippets/Large_Sets/LOmegaMultEqn}
-
-Correctness of the function \texttt{L\_omega\_mult} is asserted through the following lemma.
-
-\input{movies/snippets/Large_Sets/LOmegaMultOk}
-
-By diagonalization, we obtain a simple formula for $L_{\omega^2}$.
-
-\input{movies/snippets/Large_Sets/LOmegaSquare}
-\input{movies/snippets/Large_Sets/LOmegaSquareEqn}
-\input{movies/snippets/Large_Sets/LOmegaSquareOk}
-
-%%%% ICI
-
-\subsubsection{Going further}
-Let us consider a last example, ``computing'' $L_{\omega^3}$.
-Let us study the canonical sequence associated with $\omega^3$, the elements of which are 
-$\omega^2\times i\;(i\in\mathbb{N}_1)$.
-
-To this end, we prove a generic lemma, which expresses $L_{\omega^\alpha\times i}$ as an iterate of $L_{\omega^\alpha}$.
-Note that in this lemma, we assume that the function associated with $\alpha$ is strictly monotonous and
-greater or equal than the successor function, and prove that $L_{\omega^\alpha\times i}$ satisfies  the same properties.
-
-\inputsnippets{Large_Sets/phi0Mult}
-
-\inputsnippets{Large_Sets/phi0MultOk,
-Large_Sets/Phi0MultSLe}
-
-Let us look now
-at the ordinal $\omega^2\times i$, using \texttt{L\_phi0\_mult}.
-
-\input{movies/snippets/Large_Sets/LOmegaSquareTimes}
-
-We are now ready to get an exact formula for $L_{\omega^3}$, by diagonalization over $L_{\omega^2\times i}$.
-
-\input{movies/snippets/Large_Sets/LOmegaCube}
-
-Thus, for instance, $L_{\omega^3}(3)=L_{\omega^2\times 4}(3)$.
-
-\input{movies/snippets/Large_Sets/LOmegaCube3Eq}
-
-This number is quite big. Using \texttt{Ocaml}'s \texttt{float} arithmetic,
-we can under-approximate it by $2^{3.8\times10^{30}}\times 3.8\times{10^{30}}$.
-
-\begin{Coqsrc}
-# let exp2 x = 2.0 ** x;;
-
-val exp2 : float -> float = <fun>
-#   exp2 95.0 *. 97.0 -. 1.0;;
-- : float = 3.84256588194182037e+30
-# let n = exp2 95.0 ;;
-# let p = n *. 97.0 -. 1.0;;
-val p : float = 3.84256588194182037e+30
-
-Estimation :
-2 ** (3.84 e+30) * 3.84 e+30.
-\end{Coqsrc}
-
-\subsection{Using \texttt{Equations}}
-\label{sect:L-equations}
-
-Note that we did not define any function $L_\alpha$ \emph{for any $\alpha<\epsilon_0$} yet.
-We have got no more than a collection of proved realizations of $\texttt{L\_spec}\;\alpha$ for a few values of $\alpha$.
-
-\index{coq}{Plug-ins!Equations}
-
-Using the \texttt{coq-equations} plug-in by 
-M. Sozeau~\cite{sozeau:hal-01671777}, we will now define a function \texttt{L\_} which maps any ordinal  $\alpha<\epsilon_0$ to a proven realization of 
-$\texttt{L\_spec}\;\alpha$.   
-To this end, we represent ordinals as inhabitants of the type 
-\texttt{E0} of well-formed ordinal terms (see Sect~\vref{sect:E0-def}). So, we define a total function \texttt{L\_} of type
-\texttt{E0 -> nat -> nat}, by transfinite recursion, considering the usual three cases : $\alpha=0$, $\alpha$ is a successor, $\alpha$ is a limit ordinal.
-
-\subsubsection{Definition}
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#L_}{L\_alpha}).}
-
-\label{Functions:L-alpha}
- \index{hydras}{Library Epsilon0!Functions!L\_@L\_ (final step of a minimal path}
-
-\input{movies/snippets/L_alpha/LDef}
-
-This definition results in a bunch of automatically generated lemmas. For instance:
-
-\input{movies/snippets/L_alpha/AboutLEquation1}
-
-In most cases, it may be useful to write human-readable  paraphrases of these statements.
-
-\inputsnippets{L_alpha/Paraphrasesa}
-\inputsnippets{L_alpha/Paraphrasesb}
-\inputsnippets{L_alpha/Paraphrasesc}
-
-Using these three lemmas as rewrite rules, we can prove more properties of the functions \texttt{L\_$\alpha$}.
-
-\input{movies/snippets/L_alpha/LFiniteOmega}
-
-By  well-founded induction on $\alpha$, we prove the following properties:
-
-\input{movies/snippets/L_alpha/LGeS}
-
-\input{movies/snippets/L_alpha/LCorrect}
-
-\label{sect:L-correct-proof}
-
-Please note that the proof of \texttt{L\_correct} applies the lemmas proven in Sections~\ref{sect:L-spec-zero}, ~\ref{sect:L-spec-succ} and ~\ref{sect:L-spec-lim}.
-Our previous study of \texttt{L\_spec} allowed us to pave the way for the definition by \texttt{Equations} and the correctness proof.
-
-\paragraph*{\gaiasign}
-The module
-\href{../theories/html/gaia_hydras.GL_alpha.html}{gaia\_hydras.GL\_alpha} contains an adaptation of
-\href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0/L\_alpha}.
-
-\subsubsection{Back to hydra battles}
-\label{def:L-alpha}
-
-Lemma \texttt{battle\_length\_std } of
-Module~\href{../theories/html/hydras.Hydra.Battle_length.html}{Hydra.Battle\_length} relates the length of standard battles with the functions $L_\alpha$.
-
-\input{movies/snippets/Battle_length/battleLengthStd}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Instead of considering standard paths and battles, consider battles where the replication factor is a constant $k$.
-Please use \texttt{Equations} in order to define the function that computes the length of the $k$-path which leads  from $\alpha$ to $0$.
-Prove a few  exact formulas and minoration lemmas.
-\end{exercise}
-
-\section{A variant of the Hardy hierarchy}
-
-\label{sect:hardy}
-
-In order to give a feeling on  the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the  \emph{Hardy hierarchy} of fast growing functions,
-presented for instance in~\cite{Promel2013}. 
-\index{maths}{Rapidly growing functions!Hardy Hierarchy}
-
-\begin{remark}
-  Indeed, the functions presented in this section are a \emph{variant} of the Hardy hierarchy of functions.
-In the future versions of this development, we will correct the references to the literature.
-For the time being, we call our functions $H'_\alpha$ in order to underline the difference from ``classic'' Hardy functions.
-\end{remark}
-
-For each ordinal $\alpha$ below $\epsilon_0$, $H'_\alpha$ is 
-a total arithmetic function, defined  by  transfinite recursion on $\alpha$, according to three cases:
-
-\index{maths}{Transfinite induction}
-
-\begin{itemize}
-\item If $\alpha=0$, then $H'_\alpha (k)= k$ for any natural number $k$.
-\item If $\alpha=\textrm{succ}(\beta)$, then 
-$H'_\alpha(k)=H'_\beta(k+1)$ for any $k \in \mathbb{N}$
-\item If $\alpha$ is a limit ordinal, then 
-$H'_\alpha(k) = H'_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
-\end{itemize}
-
-\begin{remark}
- The ``classic'' definition of the Hardy hierarchy differs in the third equation.
-
-\begin{itemize}
-\item If $\alpha=0$, then $H_\alpha (k)= k$ for any natural number $k$.
-\item If $\alpha=\textrm{succ}(\beta)$, then 
-$H_\alpha(k)=H_\beta(k+1)$ for any $k \in \mathbb{N}$
-\item If $\alpha$ is a limit ordinal, then 
-$H_\alpha(k) = H_{(\canonseq{\alpha}{k})}(k)$ for any $k\in \mathbb{N}$.
-\end{itemize}
-
-\end{remark}
-
-\subsection{Definition in \texttt{Coq}}
-
-We define a function \texttt{H'\_} of type \texttt{E0 -> nat -> nat} by transfinite induction over the type \texttt{E0} of the well formed ordinals below $\epsilon_0$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}}
-
-\index{hydras}{Library Epsilon0!Functions!H\_@H\_ (Hardy hierarchy (variant))}
-\index{coq}{Plug-ins!Equations}
-\label{Functions:Hprime-alpha}
-
-\input{movies/snippets/Hprime/HprimeDef}
-\input{movies/snippets/Hprime/HprimeDefb}
-
-\input{movies/snippets/Hprime/paraphrasesa}
-\input{movies/snippets/Hprime/paraphrasesb}
-\input{movies/snippets/Hprime/paraphrasesc}
-\input{movies/snippets/Hprime/paraphrasesd}
-
-\subsection{First  steps of the H' hierarchy}
-Using rewrite rules from \texttt{H'\_eq1} to \texttt{H'\_succ\_eqn}, we can explore the functions $H'_\alpha$ for  small values of $\alpha$.
-
-\paragraph*{\gaiasign}
-The module
-\href{../theories/html/gaia_hydras.GHprime.html}{gaia\_hydras.GHprime} contains an adaptation of
-\href{../theories/html/hydras.Epsilon0.Hprime.html}{Epsilon0.Hprime}.
-
-\subsubsection{Finite ordinals}
-
-By induction on $i$, we prove a simple expression of \texttt{H'\_ (Fin i)}, where 
-\texttt{Fin $i$}  is the $i$-th finite ordinal.
-
-\input{movies/snippets/Hprime/HprimeFin}
-
-\subsubsection{Multiples of \texorpdfstring{$\omega$}{omega}}
-
-Since the canonical sequence of $\omega$ is composed of finite ordinals, 
-it is easy to get the formula associated with $H'_\omega$.
-
-\input{movies/snippets/Hprime/HprimeOmega}
-
-Before going further, we prove a useful rewriting lemma:
-
-\input{movies/snippets/Hprime/HprimePlusFin}
-
-Then, we get easily formulas for $H'_{\omega+i}$, and $H'_{\omega\times i}$ for any natural number $i$.
-
-\input{movies/snippets/Hprime/HprimeExamplesa}
-\input{movies/snippets/Hprime/HprimeExamplesb}
-\input{movies/snippets/Hprime/HprimeExamplesc}
-\input{movies/snippets/Hprime/HprimeExamplesd}
-
-Since $\omega^2$ is a limit ordinal, we prove the following equality: 
-\[H'_{\omega^2} (k) = 2 ^ {k+1} \times (k+1) - 1\]
-
-\input{movies/snippets/Hprime/HprimeOmegaSqr}
-
-\subsubsection{New limits}
-
-Our next step would be to prove an exact formula for $H'_{\omega^\omega}(k)$.
-Since the canonical sequence of $\omega^\omega$ is composed of all the
-$\omega^i$, we first need to express $H'_{\omega^i}$ for any natural number $i$.
-
-Let $i$ and $k$ be two natural numbers. 
-The ordinal $\canonseq{\omega^(i+1)}{k}$ is the product
-$\omega^i \times k$, so we need also to consider ordinals of this form.
-
-\begin{enumerate}
-\item First,  we express $H'_{\omega^\alpha \times (i+2)}$ in terms of
-$H'_{\omega^\alpha \times (i+1)}$.
-
-\input{movies/snippets/Hprime/HprimeOmegaTerm1}
-
-\item
-Then, we prove by induction on $i$ that $H'_{\omega^\alpha \times (i+1)}$ is just the
-$(i+1)$-th iterate of $H'_{\omega^\alpha}$.
-
-\input{movies/snippets/Hprime/HprimeOmegaTerm}
-
-\item In particular, we derive a formula for $H'_{\omega^{i+1}}$.
-
-\input{movies/snippets/Hprime/HprimeSuccFun}
-\input{movies/snippets/Hprime/HprimePhi0SI}
-
-\item We get now a  formula for $H'_{\omega^3}$:
-
-\input{movies/snippets/Hprime/HprimeOmegaCube}
-\end{enumerate}
-
-\subsubsection{A numerical example}
-\label{sect:bignum}
-
-It is hard to capture the complexity of this function by looking only at this
-``exact'' formula. 
-Let us consider a simple example: the number $H'_{\omega^3}(3)$.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3a}
-
-Thus, the number $H_{\omega^3}(3)$ can be written as four nested applications of $f$.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3b}
-
-In a more classical writing, this number is displayed as follows:
-
-{\Large
-$$
-H'_{\omega^3}(3) =  2 ^ {(2 ^ {N + 1} \, (N+1) )}   \,  (2 ^ {N+1} \, ( N +1) ) - 1
-$$
-}
-
-In order to make this statement more readable, we introduce a local definition.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3c}
-
-This number looks quite big; let us compute an approximation with \texttt{Ocaml}:
-
-\begin{Coqsrc}
-# (2.0 ** 64.0 *. 64.0 -. 1.0);; 
-\end{Coqsrc}
-
-\begin{Coqanswer}
-- : float = 1.1805916207174113e+21
-\end{Coqanswer}
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3d}
-
-We leave as an exercise to determine the best approximation as possible of
- the size of this number (for instance its number of digits).  For instance, if
-we do not take into account the multiplications in the formula above,
-we obtain that, in base $2$, the number $H'_{\omega^3}(3)$ has at least
-$2^{10^{21}}$  digits. But it is still an under-approximation !
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3e}
-
-\subsubsection{A formula for \texorpdfstring{$H'_{\omega^\omega}$}{\texttt{H'(phi0(omega))}}}
-
-Now, we can get at last an exact formula for $H'_{\omega^\omega}$.
-
-\input{movies/snippets/Hprime/HprimePhi0Omega}
-
-Using extensionality of the functional \texttt{iterate}, we also get a closed formula.
-
-\input{movies/snippets/Hprime/HprimePhi0OmegaClosed}
-
-Note that this formula contains two occurrences of the functional \texttt{iterate}, the outer one is in fact a second-order iteration (on type \texttt{nat -> nat)}
-and the inner one  first-order (on type \texttt{nat}).
-
-\subsection{Abstract properties of  
-\texorpdfstring{$H'_\alpha$}{H'}}
-~\label{sect:H-alpha-prop}
-
-Since pure computation seems to be useless for dealing with expressions of the form $H'_\alpha(k)$, even for small values of $\alpha$ and $k$, we need to prove theorems for comparing $H'_\alpha(k)$ and $H'_\beta(l)$, in terms of comparison
-between $\alpha$ and $\beta$ on the one hand, $k$ and $l$ on the other hand.
-
-But beware of fake theorems! For instance, one could believe that $H'$ is monotonous in its first argument. The following proof shows this is false.
-
-\input{movies/snippets/Hprime/HprimeNonMono1}
-
-On the contrary, the functions of the $H'$ hierarchy have the following five properties~\cite{KS81}: for any $\alpha < \epsilon_0$,
-\begin{itemize}
-\item the function $H'_\alpha$ is strictly monotonous :
-      For all $n,p \in\mathbb{N}, n < p \Rightarrow H'_\alpha(n)< H'_\alpha(p)$.
-\item If $\alpha \not= 0$, then for every $n$, $n<H'_\alpha(n)$.
-\item The function $H'_\alpha$ is pointwise less or equal than $H'_{\alpha+1}$
-
-\item For any $n\geq 1$, $H'_\alpha(n)<H'_{\alpha+1}(n)$.
-\emph{We say that $H'_{\alpha+1}$ dominates $H'_\alpha$ from $1$}.
-\item For any $n$ and $\beta$, if $\alpha \xrightarrow[n]{} \beta$, then
-$H'_\beta(n)\leq H'_\alpha(n)$.
-\end{itemize}
-
-\index{maths}{Abstract properties of arithmetic functions}
-\index{hydras}{Abstract properties of arithmetic functions}
-
-\index{maths}{Transfinite induction}
-
-In \coq{}, we follow the  proof of~\cite{KS81}. This proof is mainly a single  proof by transfinite induction on $\alpha$ of the conjunction of the five properties.
-For each $\alpha$, the three cases : $\alpha=0$, $\alpha$ is a limit, and 
-$\alpha$ is a successor are considered. Inside each case, the five sub-properties are proved sequentially, using the abstract properties defined in Sect~\vref{sect:abstract-arith-prop}
-
-\input{movies/snippets/Hprime/PAP}
-
-Using a few lemmas \emph{à la} Ketonen-Solovay, we prove that
-if $\alpha<\beta$, then $H'_\beta$ eventually dominates
-$H'_\alpha$.
-We let the reader look at the proof (Section \texttt{Proof\_of\_H'\_mono\_l} of \href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}).
-
-\noindent
-\input{movies/snippets/Hprime/HprimeDom}
-
-%ICI
-\subsection{Comparison between \texttt{L\_} and \texttt{H'\_} }
-
-By well-founded induction on $\alpha$, we prove that
- $H'_\alpha(i) \leq L_\alpha(i+1)$ for any $i$.
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#H'_L_}{Epsilon0.L\_alpha}}
-
-\input{movies/snippets/L_alpha/HprimeL}
-
-\subsubsection{Back to hydras}
-
-The following theorem relates the length of (standard) battles with the the $H'$ family of fast growing functions.
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems}}
-
-\input{movies/snippets/Hydra_Theorems/battleLengthStdHardy}
-
-Since $H'_{\omega^3}(3)= H'_{\omega^3+3}(0)$, the big number shown in Section~\vref{sect:bignum} is less or equal than 
-the number of steps of the battle with the initial hydra of
-Figure~\ref{fig:start}, and  $0$ as initial replication factor.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[very thick, scale=0.3]
-  \node (h1) at (4,0){$\bullet$};
-  \node (h2) at (2,2){$\bullet$};
-  \node (h3) at (0,4){$\bullet$};
-  \node (h4) at (1,4){$\bullet$};
-  \node (h5) at (2,4){$\bullet$};
-  \node (h6) at (4.5,2){$\bullet$};
-  \node (h7) at (6.5,2){$\bullet$};
-  \node (h8) at (8.5,2){$\bullet$};
-    \draw (h1) -- (h2) ;
-    \draw (h2) -- (h3) ;
-    \draw (h2) -- (h4) ;
-    \draw (h2) -- (h5) ;
-    \draw (h1) -- (h6) ;
-    \draw (h1) -- (h7) ;
-    \draw (h1) -- (h8);
- \end{tikzpicture}
-
-  \caption{The hydra corresponding to the ordinal $\omega^3+3$}
-  \label{fig:start}
-\end{figure}
-
-\section{A variant of the Wainer hierarchy (functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
-\label{sect:wainer}
-
-\index{maths}{Rapidly growing functions!Wainer Hierarchy}
-
-Ketonen and Solovay introduce in~\cite{KS81} a ``trivial'' variant of the Wainer hierarchy~\cite{BW85, Wainer1970} of fast growing functions, indexed by ordinals below $\epsilon_0$.
-The functions $F_\alpha$ are defined by the following equations.
-
-\label{F_equations}
-\begin{itemize}
-\item $F_0(i)=i+1$
-\item $F_{\beta+1}(i)= (F_\beta)^{(i+1)}(i)$, where $f^{(i)}$ is the $i$-th iterate of $f$.
-\item $F_\alpha(i) = F_{\canonseq{\alpha}{i}} (i)$ if $\alpha$ is a limit ordinal.
-\end{itemize}
-
-\begin{remark}
-The difference with the ``classic'' Wainer hierarchy 
-$f_\alpha\;(\alpha<\epsilon_0)$ lies in the second equation:
-$f_{\beta+1}(i) = (f_\beta)^{(i)}(i)$ and not
-$f_{\beta+1}(i) = (f_\beta)^{(i+1)}(i)$.
-
-A module about 
-the classic Wainer hierarchy is in preparation.
-
-Note also that \cite{KS81} defines also $F_{\epsilon_0}$ (by the third equation).
-Since $\epsilon_0$ is not representable in type \texttt{E0}, our implementation in \coq{} does not take $F_{\epsilon_0}$ into account.
-
-\end{remark}
-
-A first attempt is to write a definition of $F_\alpha$ by equations, in the same way as for $H_\alpha$ (the functional \texttt{iterate} has already been used in Sect.\vref{Functions:iterate}).
-
-\index{hydras}{Library Prelude!iterate}
-
-\input{movies/snippets/Iterates/iterateDef}
-
-\emph{From\href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}~}.
-
-\index{coq}{Plug-ins!Equations}
-
- \input{movies/snippets/F_alpha/FailDemo}
-
-We presume that this error comes from the recursive call of \texttt{F\_} inside
-an application of \texttt{iterate}. The workaround we propose is to define first 
-the iteration of \texttt{F\_}  as an helper $F^*$, then to define the function $F$ as a ``iterating $F^*$ once''.
-
-\texttt{Equations} accepts the following definition, relying on  lexicographic ordering on pairs $(\alpha,n)$.
-
-\label{sect:F-equations}
-
-\index{coq}{Plug-ins!Equations}
-\label{Functions:F-alpha}
-\index{maths}{Rapidly growing functions}
-\index{hydras}{Library Epsilon0!Functions!F\_@F\_ (Wainer hierarchy)}
-
-\input{movies/snippets/F_alpha/goodDefa}
-\input{movies/snippets/F_alpha/goodDefb}
-
-It is quite easy to prove that our functional \texttt{F\_} satisfies the equations on page~\pageref{sect:F-equations}.
-\index{hydras}{Library Prelude!iterate}
-
-\input{movies/snippets/F_alpha/FEquations}
-
-As for the Hardy functions, we can use these equalities as rewrite rules for
-``computing'' some values of $F_\alpha(i)$, for small values of $\alpha$.
-
-\input{movies/snippets/F_alpha/FirstValues}
-
-Like in Sect~\ref{sect:H-alpha-prop}, we prove by induction the following properties (see~\cite{KS81}).
-
-\input{movies/snippets/F_alpha/FalphaThms}
-
-As a corollary, we prove that, if $\beta<\alpha$, then
-$F_\alpha$ dominates $F_\beta$ (p. 284 of~\cite{KS81}).
-
-\input{movies/snippets/F_alpha/FDomContext}
-\input{movies/snippets/F_alpha/FDom}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove the following property:
-
-\begin{Coqsrc}
-Lemma LF3 : dominates_from  2 (F_ 3) (fun  n => iterate exp2 n n).
-\end{Coqsrc}
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
-\end{exercise}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that, for any $\alpha\geq 3$ and $n\geq 2$,
-$F_\alpha(1+n)\geq 2^{F_\alpha(n)}$.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
-\end{exercise}
-
-%ici
-\index{hydras}{Exercises}
-
-\begin{exercise}
-It is tempting to prove a simple property of monotony 
-of the function \texttt{F\_}.
-
-\begin{quote}
-   Let $\alpha\leq\beta<\epsilon_0$. For any $n\geq 2$,
-$F_\alpha(n)\leq F_\beta(n)$. 
-\end{quote}
-Prove or disprove this statement.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/is_F_monotonous.v}{exercises/ordinals/is\_F\_monotonous.v}.}
-\end{exercise}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-
-Prove that for any $n\geq 2$, $\textrm{Ack}\,\,n\,n\leq  F_\omega(n)$, where \textrm{Ack} is the Ackermann function.
-Next, prove that $F_\alpha$ is not primitive recursive, for any $\alpha\geq\omega$  (please see Sect.~\vref{sect:ack-not-PR}).
-On the other hand, please show that for any natural number $n$, the function $F_n$ is primitive recursive.
-Thus $F_\alpha$ is primitive recursive if and only if $\alpha$ is finite.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_omega.v}{exercises/ordinals/F\_Omega.v}.
-Properties of the Ackermann function are studied in
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/Ack.v}{theories/ordinals/MoreAck/Ack.v} and
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/AckNotPR.v}{theories/ordinals/MoreAck/AckNotPR.v}
-.}
-\end{exercise}
-
-\paragraph*{\gaiasign} Library~\href{../theories/html/gaia_hydras.GF_alpha.html}{gaia\_hydras.GF\_alpha}  defines the functions $F_\alpha$ for \gaia's world.
-
-\section{More about rapidly growing functions}
-
-In Sect.~\vref{sect:battle-length-notPR}, we prove that the length of hydra-battles (for a given hydra, according to the initial replication factor) is not primitive recursive in general. 
-This proof uses properties of the Ackermann function, and the $H'_\alpha$, $F_\alpha$, $L_\alpha$ families of functions.
-
-% ---------------------------------------------------
-\input{gaia-chapter}
-
-%----------------------------------------------------------------------
-\chapter[Countable ordinals (after Sch\"{u}tte)]{Kurt Schütte's axiomatic definition of countable ordinals}
-
-\label{chap:schutte} 
-%ON
-
-In the present chapter, we  compare our implementation of the segment $[0,\epsilon_0)$ with a mathematical text in order to ``validate'' our constructions.
-Our reference here is the axiomatic definition of the set of countable ordinals,
-in chapter V of Kurt Schütte's book `` Proof Theory ''~\cite{schutte}.
-
-\begin{remark}
-\emph{In all this chapter, the word ``ordinal'' will be considered as a synonymous of
-``countable ordinal''}  
-\end{remark}
-
-Schütte's definition of countable ordinals relies on the following three axioms:
-
-There  exists a strictly ordered set , such that
-\begin{enumerate}
-\item  $(\mathbb{O},<)$ is well-ordered
-\item Every bounded subset of $\mathbb{O}$  is countable
-\item Every countable subset of $\mathbb{O}$  is bounded.
-\end{enumerate}
-
-Starting with these three axioms, Schütte re-defines the vocabulary about ordinal numbers: the null ordinal $0$, limits and successors, the addition of ordinals, the infinite ordinals $\omega$, $\epsilon_0$, $\Gamma_0$, etc.
-
-This chapter describes an adaptation to \coq{} of Schütte's axiomatization. 
- Unlike the rest of our libraries, our library
-\href{../theories/html/hydras.Schutte.Schutte.html}{hydras.Schutte}
-is not constructive, and relies on several axioms.
-
-\begin{itemize}
-\item First, please keep in mind  that the set of countable ordinals is not countable.
-Thus, we cannot hope to represent all countable ordinals as finite terms of an inductive type, which was possible with  the set of ordinals strictly less than $\epsilon_0$ (resp.
-$\Gamma_0$)
-\item We tried to be as close as possible to K.
-Schütte's text, which uses ``classical'' mathematics : excluded middle, Hilbert's $\epsilon$ (choice) and Russel's $\iota$ (definite description) operators.
-Both operators allow us to write definitions close to the natural mathematical language, such as ``$\textrm{succ}$ is \emph{the} least ordinal strictly greater than $\alpha$''
-\item Please note that only the library \href{../theories/html/hydras.Schutte.Schutte.html}{Schutte/*.v} is ``contaminated'' by axioms, and that the rest of our libraries remain constructive.
-\end{itemize}
-
-\section{Declarations and axioms}
-
-Let us declare a type 
-\texttt{Ord} for representing countable ordinals, and a binary relation
- \texttt{lt}. Note that, in our development, \texttt{Ord} is a type, while the \emph{set} of countable ordinals (called $\mathbb{O}$ by Schütte) 
-is the full set over the type \texttt{Ord}.
-
-\label{types:Ord}
-
-We use Florian Hatat's library on countable sets, written as he was a student of  \emph{\'Ecole Normale Supérieure de Lyon}.
-A set $A$ is countable if there is an injective function from $A$ to $\mathbb{N}$ (see
-Library \href{../theories/html/hydras.Schutte.Countable.html}%
-{\texttt{Schutte.Countable}}).
-
-\vspace{6pt}
-
-\emph{From Module\href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{Schutte.Schutte\_basics}}}
-
-\index{hydras}{Library Schutte!Types!Ord}
-
-\input{movies/snippets/Schutte_basics/OrdDecl}
-
-Schütte's first axiom tells that \texttt{lt} is a well order on the set 
-\texttt{ordinal} (The  class \texttt{WO} is defined in
-Module~\href{../theories/html/hydras.Schutte.Well_Orders.html}{Schutte.Well\_Orders.v}).
-
-\index{hydras}{Library Schutte!Type classes!WO@ WO (well order)}
-
-\label{types:WO}
-
-\input{movies/snippets/Well_Orders/Mdecl}
-
-\input{movies/snippets/Well_Orders/WODef}
-
-\input{movies/snippets/Schutte_basics/AX1}
-
-The second and third axioms say that a subset $X$ of $\mathbb{O}$ is
-(strictly) bounded if and only if it is countable.
-
-\input{movies/snippets/Schutte_basics/AX23}
-
-\texttt{AX2} and \texttt{AX3} could have been replaced by a single axiom (using the \texttt{iff} connector), but we decided to respect as most as possible the structure of Schütte's definitions.
-
-\section{Additional  axioms}
-
-The adaptation of Schütte's mathematical discourse to \coq{} led us to
-import a few axioms from the standard library. We encourage the reader to consult \coq{}'s FAQ about the safe use of axioms
- \url{https://github.com/coq/coq/wiki/The-Logic-of-Coq#axioms}.
-
-\subsubsection{Classical logic}
-
-In order to work with classical logic, we import the module
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Classical.html}{Coq.Logic.Classical}  of \coq{}'s standard library, specifically the following axiom:
-
-\begin{Coqsrc}
- Axiom classic : forall P:Prop, P \/ ~P.
-\end{Coqsrc}
-
-\subsubsection{Description operators}
-
-In order to respect Schütte's style, we imported also the library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Epsilon.html}{\texttt{Coq.Logic.Epsilon}}.  The rest of this section presents a few examples of
-how Hilbert's choice operator and Church's definite description allow us
- to write understandable definitions (close to the mathematical natural language).
-
-\subsubsection{The definition of zero}
-
-According to the  definition of a well order, every non-empty subset of \texttt{Ord} has a least element. Furthermore, this least element is unique. We would like to call this element  \texttt{zero}.
-
-\input{movies/snippets/Schutte_basics/iotaDemoa}
-\input{movies/snippets/Schutte_basics/iotaDemob}
-
-Indeed, the basic logic of  \coq{} does not allow us to eliminate a proof of a proposition 
-$\exists!\,x:A,\,P(x)$ for building a term whose type lies in the sort \texttt{Type}. 
-The reasons for this impossibility are explained in many documents~\cite{BC04, chlipalacpdt2011, Coq}.
-
-Let us import the library \texttt{Coq.Logic.Epsilon}, which contains the following axiom and lemmas.
-
-\input{movies/snippets/MoreEpsilonIota/EpsilonStatement}
-
-Hilbert's $\epsilon$ \emph{operator} is derived from this  axiom.
-
-\input{movies/snippets/MoreEpsilonIota/EpsilonDef}
-
-If we consider the \emph{unique existential} quantifier $\exists!$, we obtain
-Church's \emph{definite description operator}.
-
-\input{movies/snippets/MoreEpsilonIota/iotaDef}
-
-Indeed, the operators \texttt{epsilon} and \texttt{iota} allowed us to make our definitions 
-quite close to Schütte's text. Our libraries \href{../theories/html/hydras.Schutte.MoreEpsilonIota.html}%
-{\texttt{Schutte.MoreEpsilonIota}}
-and
-\href{../theories/html/hydras.Schutte.PartialFun.html}%
-{\texttt{Schutte.PartialFun}} are extensions of \texttt{Coq.logic.Epsilon} for making easier 
-such definitions. See also an article in french~\cite{PCiota}.
-
-\input{movies/snippets/MoreEpsilonIota/Defs}
-
-In order to use these tools,  we have to tell \coq{}  that the declared type \texttt{Ord} is not empty:
-
-\input{movies/snippets/Schutte_basics/inhOrd}
-
-We are now able to define \texttt{zero} as the least ordinal. For this purpose,
-we define a function returning the least element of any [non-empty]  subset.
-
-\emph{From Module\href{../theories/html/hydras.Schutte.Well_Orders.html}%
-  {\texttt{Schutte.Well\_Orders}}}
-
-\input{movies/snippets/Well_Orders/MDecl}
-\input{movies/snippets/Well_Orders/theLeast}
-
-\vspace{4pt}
-
-From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{~Schutte.Schutte\_basics}}
-
-\label{Constants:zero:Ord}
-\index{hydras}{Library Schutte!Constants!zero}
-
-\input{movies/snippets/Schutte_basics/zeroDef}
-
-We want to prove now that zero is less than or equal to any ordinal number.
-
-\input{movies/snippets/Schutte_basics/zeroLe}
-
-\subsubsection{Remarks on \texttt{epsilon} and \texttt{iota}}
-
- What would happen in case of a misuse of \texttt{epsilon} or \texttt{iota} ?
-For instance, one could give a unsatisfiable specification to \texttt{epsilon} or 
-a specification for \texttt{iota} that admits several realizations.
-
-Let us consider an example:
-
-\input{movies/snippets/Schutte_basics/BadBottoma}
-
-Since we won't be able to prove the proposition
-\linebreak \Verb|(exists! a: Ord, least_member (Empty_set Ord) a)|, the only properties we would be able to prove about \texttt{bottom} would be \emph{trivial} properties, 
-\emph{i.e.}, satisfied by \emph{any} element of type \texttt{Ord}, like for instance
-\texttt{bottom = bottom}, or \texttt{zero <= bottom}.
-
-\input{movies/snippets/Schutte_basics/trivialProps}
-
-On the other hand, the following attempt fails, because of the unprovable first subgoal (please notice that the second subgoal is easy to solve !).
-
-\input{movies/snippets/Schutte_basics/Failure}
-
-In short, using \texttt{epsilon} and \texttt{iota} in our implementation of countable ordinals after Schütte has two main advantages.
-
-\begin{itemize}
-\item It allows us to give a \emph{name} (using \texttt{Definition}) two witnesses 
-of existential quantifiers (let us recall that, in classical logic, one may consider non-constructive proofs of existential statements)
-\item By separating definitions from proofs of [unique] existence, one may make definitions  more concise and readable. Look for instance at 
-the definitions of  \texttt{zero}, \texttt{succ}, \texttt{plus}, etc. in the rest of this chapter.
-\end{itemize}
-%%%% ICI ICI
-
-\section{The  successor function}
-
-The definition of the function \texttt{succ:Ord -> Ord} is very concise. The successor of any ordinal $\alpha$ is the smallest ordinal strictly greater than $\alpha$.
-
-\label{Functions:succ-sch}
-\index{hydras}{Library Schutte!Functions!succ}
-
-\input{movies/snippets/Schutte_basics/succDef}
-
-Using \texttt{succ}, we define the following predicates.
-
-\input{movies/snippets/Schutte_basics/isSuccIsLimit}
-
-% \begin{remark}
-% Please look at remark~\vref{warning:coercions}.
-
-How do we prove properties of the successor function?
-First, we make its specification explicit.
-
-\input{movies/snippets/Schutte_basics/succSpec}
-
-Then, we prove that our function \texttt{succ} meets this specification.
-
-\input{movies/snippets/Schutte_basics/succOka}
-
-We have now to prove that the set of all ordinals strictly greater than $\alpha$ has a unique least element.
-But the singleton set $\{\alpha\}$ is countable, hence  bounded (by the axiom \texttt{AX3}).
-Hence; the set $\{\beta\in\mathbb{O}|\alpha < \beta\}$ is not empty
-and therefore has a unique least element.
-
-The rest of the \coq{} proof script is quite short.
-
-\input{movies/snippets/Schutte_basics/succOkb}
-
-We can ``uncap'' the description operator for proving properties of the
-\texttt{succ} function.
-
-\input{movies/snippets/Schutte_basics/succProps}
-
-\section{Finite ordinals}
-
-Using \texttt{succ}, it is now easy to define recursively all the finite ordinals.
-
-\label{sect:notation-F-sch}
-
-\input{movies/snippets/Schutte_basics/finiteDef}
-
-\section{The definition of \texttt{omega}}
-In order to define $\omega$, the first infinite ordinal, we use an operator which
-``returns'' the least upper bound (if it exists) of a subset $X\subseteq \mathbb{O}$.
-For that purpose, we first use a predicate:
-(\texttt{is\_lub $D$ \textit{lt} $X$ $a$}) if $a$ belongs to $D$ and is the least 
-upper bound  of $X$ (with respect to \textit{lt}).
-
-\input{movies/snippets/Lub/isLubDef}
-
-\input{movies/snippets/Schutte_basics/supDef}
-
-Then, we define the function \texttt{omega\_limit} which returns the least upper bound 
-of the  (denumerable) range of any sequence \texttt{s: nat -> Ord}. 
-By \texttt{AX3} this range is bounded, hence the set of its upper bounds is not empty and has a least element.
-Then we define \texttt{omega} as the limit of the sequence of finite ordinals.
-
-\input{movies/snippets/Schutte_basics/omegaDef}
-
-\label{sect:notation-omega}
-
-Among the numerous properties of the ordinal $\omega$, let us quote the following ones
-(proved in Module 
-\href{../theories/html/hydras.Schutte.Schutte_basics.html\#finite_lt_omega}{\texttt{Schutte.Schutte\_basics}})
-
-\input{movies/snippets/Schutte_basics/omegaPropsa}
-\input{movies/snippets/Schutte_basics/omegaPropsb}
-\input{movies/snippets/Schutte_basics/omegaPropsc}
-\input{movies/snippets/Schutte_basics/omegaPropsd}
-
-\subsection{Ordering functions and ordinal addition}
-
-After having defined the finite ordinals and the infinite ordinal $\omega$, we  define the sum $\alpha+\beta$ of two countable ordinals.
-Schütte's definition looks like the following one:
-
-\begin{quote}
-``$\alpha+\beta$ is the $\beta$-th ordinal greater than or equal to $\alpha$''
-\end{quote}
-
-The purpose of this section is to give a meaning to the construction
-``the $\alpha$-th element of $X$''  where $X$ is any non-empty subset of $\mathbb{O}$.
-We follow Schütte's approach, by defining the notion of \emph{ordering functions},
-a way to associate a unique ordinal to each element of $X$.
-Complete definitions and proofs can be found in Module
- \href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
-{\texttt{Schutte.Ordering\_Functions}} ).
-
-\subsection{Definitions}
-
-A \emph{segment} is a set $A$ of ordinals such that, whenever  $\alpha\in A$ and
-$\beta<\alpha$, then $\beta\in A$; a segment is  \emph{proper} if it strictly included in $\mathbb{O}$.
-
-\input{movies/snippets/Ordering_Functions/segmentDef}
-
-Let  $A$ be a segment, and $B$ a subset of $\mathbb{O}$ : an \emph{ordering function for $A$ and  $B$} is a strictly increasing bijection from $A$ to $B$.
-The set $B$ is said to be an \emph{ordering segment} of $A$.
-Our definition in \coq{} is a direct translation of the mathematical text of~\cite{schutte}.
-
-\index{maths}{Ordinal numbers!Ordering functions}
-\index{hydras}{Library Schutte!Predicates!ordering function@ordering\_function}
-
-\input{movies/snippets/Ordering_Functions/orderingFunctionDef}
-
-We are now able to associate with any subset $B$ of $\mathbb{O}$ its ordering segment and ordering function.
-
-\input{movies/snippets/Ordering_Functions/ordDef}
-
-Thus (\texttt{ord $B \;\alpha$}) is the $\alpha$-th element of $B$.
-Please note that the last definition uses the epsilon-based operator \texttt{some} and
-not \texttt{the}. This is due to the fact that we cannot prove the unicity (w.r.t. Leibniz' equality) of the ordering function of a given set. 
-By contrast, we admit the axiom  \texttt{Extensionality\_Ensembles}, from the library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Sets.Ensembles.html}{Coq.Sets.Ensembles}, so we use the operator \texttt{the} in the definition of
-\texttt{the\_ordering\_segment}.
-
-One of the main theorems of
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html\#ordering_function_ex}%
-{\texttt{Ordering\_Functions}} 
-associates a unique segment and a unique (up to extensionality) ordering function to every subset $B$ of $\mathbb{O}$.
-
-\input{movies/snippets/Ordering_Functions/orderingFunctionEx}
-
-Thus,  our function \texttt{ord}  which enumerates the elements of $B$ is defined in a non-ambiguous way.
-Let us quote the following theorems (see Library
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
-{\texttt{Schutte.Ordering\_Functions}} for more details).
-
-\input{movies/snippets/Ordering_Functions/orderingLe}
-\input{movies/snippets/Ordering_Functions/Th1352}
-
-\subsection{Ordinal addition}
-
-We are now ready to define and study addition on the type \texttt{Ord}.
-The following definitions and proofs can be consulted in Module
-\href{../theories/html/hydras.Schutte.Addition.html}%
-{\texttt{Schutte.Addition.v}}.
-
-\index{hydras}{Library Schutte!Functions!plus}
-
-\input{movies/snippets/Addition/additionDef}
-
-In other words,  $\alpha + \beta$ is the  $\beta$-th ordinal greater than or equal to $\alpha$. 
-Thanks to generic properties of ordering functions, we can show the following 
-properties of addition on $\mathbb{O}$. First, we prove a useful lemma:
-
-\input{movies/snippets/Addition/plusElim}
-
-As a use-case, let us prove that $0$ is a right neutral element of $+$.
-
-\input{movies/snippets/Addition/alphaPlusZero}
-
-The following lemmas are proved the same way.
-
-\input{movies/snippets/Addition/bunchOfLemmas}
-
-The following lemmas are not direct applications of \texttt{plus\_elim}.
-
-\input{movies/snippets/Addition/plusAssoc}
-\input{movies/snippets/Addition/finitePlusInfinite}
-
-It is interesting to compare the proof of these lemmas with the
-computational proofs of the corresponding statements in Module
-\href{../theories/html/hydras.Epsilon0.T1.html}%
-{\texttt{Epsilon0.T1}}. 
-For instance, the proof of the lemma 
-\texttt{one\_plus\_omega} uses the continuity of ordering functions (applied to  \texttt{(plus 1)}) and compares the limit of the $\omega$-sequences $i_{(i \in \mathbb{N})}$ and
-$(1+i)i_{(i \in \mathbb{N})}$, whereas in the library  \texttt{Epsilon0/T1}, the equality 
-$1+\omega=\omega$ is just proved with \texttt{reflexivity}!
-
-\subsubsection{Multiplication by a natural number}
-
-The multiplication of an ordinal by a natural number is defined in terms of addition.
-This operation is useful for the study of Cantor normal forms.
-
-\input{movies/snippets/Addition/multFin}
-
-\section{The exponential of basis \texorpdfstring{$\omega$}{omega}}
-
-In this section, we define the function which maps any $\alpha\in\mathbb{O}$ to
-the ordinal  $\omega^\alpha$, also written 
-$\phi_0(\alpha)$. 
-It is an opportunity to apply the definitions and results of the preceding section. 
-Indeed,  Schütte first defines a subset of $\mathbb{O}$: the set of additive principal ordinals, and $\phi_0$  is just defined as the ordering function of this set.
-
-\subsection{Additive principal ordinals}
-
-\index{maths}{Ordinal numbers!Additive principal ordinals}
-\index{hydras}{Library Schutte!Predicates!AP@AP (additive principal ordinals)}
-
-\begin{definition}
-A non-zero ordinal  $\alpha$ is said to be \emph{additive principal} if, for all  $\beta<\alpha$, $\beta+\alpha$ is equal to  $\alpha$.
-We call \texttt{AP} the set of additive principal ordinals.
-
-\end{definition}
-
-\noindent\emph{From Module \href{../theories/html/hydras.Schutte.AP.html}%
-{\texttt{Schutte.AP}}}
-
-\input{movies/snippets/AP/APDef}
-
-\subsection{The function \texttt{phi0}}
-
-Let us call  $\phi_0$ the ordering function of \texttt{AP}.
-In the mathematical text, we shall use indifferently the notations  $\omega^\alpha$ and$\phi_0(\alpha)$.
-
-\index{hydras}{Library Schutte!Functions!phi0}
-
-\input{movies/snippets/AP/phi0Def}
-
-\subsection{Omega-towers and the ordinal \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-Using $\phi_0$, we can define recursively the set of finite omega-towers.
-
-\input{movies/snippets/AP/omegaTower}
-
-\label{sect:epsilon0-as-limit}
-Then, the ordinal  $\epsilon_0$ is defined as the limit of the sequence of all finite towers (a kind of infinite tower).
-
-\input{movies/snippets/AP/epsilon0Def}
-
-The rest of our library \texttt{AP} is devoted to the proof of properties of additive principal ordinals, hence of the ordering function  $\phi0$ and the ordinal $\epsilon_0$ (which we could not express within the type \texttt{T1}).
-
-\subsection{Properties of the set  \texttt{AP}}
-
-The set of additive principal ordinals is not empty: it contains at least the ordinals  $1$ and  $\omega$.
-
-\inputsnippets{AP/APOne,AP/APOne_least_AP,AP/APOne_AP_omega,AP/APOne_omega_second_AP}
-
-The set  \texttt{AP} is  \emph{closed} under addition, and unbounded.
-\label{lemma:AP-plus-closed}
-
-\input{movies/snippets/AP/APPlusClosed}
-
-\input{movies/snippets/AP/APUnbounded}
-
-Finally, \texttt{AP} is (topologically) \emph{closed} and ordered by the segment of all countable ordinals.
-
-\index{hydras}{Library Schutte!Predicates!Closed}
-
-From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{~Schutte.Schutte\_basics}}
-
-\input{movies/snippets/Schutte_basics/ClosedDef}
-\input{movies/snippets/AP/APClosed}
-
-\subsubsection{Properties of the function \texorpdfstring{$\phi_0$}{phi0}}
-
-The ordering function $\phi_0$ of the set \texttt{AP} is defined on the full set $\mathbb{O}$ and is continuous (Schütte calls such a  function  \emph{normal}).
-
-\begin{Coqsrc}
-Theorem normal_phi0 : normal phi0 AP.
-\end{Coqsrc}
-
-The following properties come from  the definition of $\phi_0$ as the ordering function of \texttt{AP}.
-It may be interesting to compare these proofs with the computational ones described in Chapter ~\ref{chap:T1}.
-
-\input{movies/snippets/AP/APPhi0}
-
-\subsection{A last example}
-\label{Ex42-schutte}
-
-Let us prove again the equality $\omega+42+\omega^2= \omega^2$.Let us recall that $\omega^2$ is an abbreviation of $\phi_0(2)$,
-\emph{i.e} the third  additive principal ordinal.
-
-\inputsnippets{Schutte/Ex42a}
-
-Our proof is very different from the computational proof of Sect~\vref{Ex42-E0}.
-By definition of additive principal ordinals, 
-it suffices to prove the inequality $\omega+42< \phi_0(2)$.
-
-\inputsnippets{Schutte/Ex42b}
-
-Since the set \texttt{AP} of additive principals  is closed under addition
-(by Lemma \texttt{AP\_plus\_closed}, page~\pageref{lemma:AP-plus-closed}) , it suffices to prove the inequalities $\omega<\phi_0(2)$ and $42<\phi_0(2)$.
-
-\inputsnippets{Schutte/Ex42d, Schutte/Ex42c, Schutte/Ex42e}
-
-\section{More about \texorpdfstring{$\epsilon_0$}{\texttt{epsilon0}}}
-
-Let us recall that the limit ordinal  $\epsilon_0$ cannot be written within the type \texttt{T1}.
-Since we are now considering the set of all countable ordinals, we can now prove some properties of this ordinal.
-
-We prove the inequality  $\alpha<\omega^\alpha$ whenever $\alpha < \epsilon_0$.
-\emph{Note that this condition was implicit in Module
-\href{../theories/html/hydras.Epsilon0.T1.html\#lt_phi0}{Epsilon0.T1}.}
-
-\input{movies/snippets/AP/ltPhi0}
-
-The proof is as follows:
-\begin{enumerate}
-\item Since $\alpha<\epsilon_0$, consider the least $i$ such that $\alpha$ is strictly less than the omega-tower of height $i$.
-\item
-  \begin{itemize}
-  \item If $i=0$, then the result is trivial (because $\alpha=0$)
- \item  Otherwise let $i=j+1$; 
-          $\alpha$ is greater than or equal to the omega-tower of height $j$.
-         By monotony,  $\phi_0(\alpha)$ is greater than or equal to 
-        the omega-tower of height $j+1$, thus strictly greater than $\alpha$
-  \end{itemize}
- \end{enumerate}
-
-Moreover,  $\epsilon_0$ is the least ordinal $\alpha$ that verifies the equality 
-$\alpha = \omega^\alpha$, in other words, the least fixpoint of the function  $\phi_0$.
-
-\input{movies/snippets/AP/epsilon0Lfp}
-
-\section{Critical ordinals}
-
-\index{maths}{Ordinal numbers!Critical ordinals}
-\index{hydras}{Library Schutte!Predicates!Cr@Cr (critical ordinals)}
-
-For any  (countable) ordinal $\alpha$, the set $\textit{Cr}(\alpha)$ is inductively defined 
-as follows by Schütte (p.81 of~\cite{schutte}).
-
-\begin{quote}
-  \begin{itemize}
-  \item $\textit{Cr}(0)$ is the set \textit{AP} of additive principal ordinals.
-  \item If $0<\alpha$, then $\textit{Cr}(\alpha)$ is the intersection of all the sets of fixpoints of the $\textit{Cr}(\beta)$ for $\beta<\alpha$.
-  \end{itemize}
-\end{quote}
-
-This definition is translated in \coq{} in 
-Module \href{../theories/html/hydras.Schutte.Critical.html}%
-{\texttt{Schutte.Critical}}, as the least fixpoint of a functional.
-
-\input{movies/snippets/Critical/CrDef}
-
-Lets us denote by $\phi_\alpha$ the ordering function of the set $\textit{Cr}(\alpha)$ and by $A_\alpha$ its ordering segment.
-
-\input{movies/snippets/Critical/phiDef}
-
-\label{sect:phi-schutte}
-
-For instance,  we prove that $\textit{Cr}(0)$ is the set of additive principals and that $\epsilon_0$
-belongs to $\textit{Cr}(1)$.
-
-\inputsnippets{Critical/CrZeroAP,
-  Critical/epsilon0Cr1}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
- Prove that $\epsilon_0$ is the least element of $\textit{Cr}(1)$.
-\end{exercise}
-
-\subsection{A flavor of infinity}
-
-The family of the $\textit{Cr}(\alpha)$s is made of infinitely many unbounded (hence infinite) sets.
-Let us quote Lemma 5, p. 82  of~\cite{schutte}:
-\begin{quote}
-  For all $\alpha$, the set $\textit{Cr}(\alpha)$ is closed (for the least upper bound of non-empty countable sets) and unbounded.
-\end{quote}
-
-We prove this result by a transfinite induction on $\alpha$ of the conjunction of  both properties.
-
-\index{maths}{Transfinite induction}
-
-\input{movies/snippets/Critical/Lemma5}
-
-\section{Cantor normal form}
-
-The notion of Cantor normal form is defined for all countable ordinals.
-Nevertheless, note that, contrary to the implementation based on type \texttt{T1},
-the Cantor normal form of an ordinal $\alpha$ may contain $\alpha$ as a 
-sub-term\footnote{This would prevent us from trying to represent Cantor normal forms as finite trees (like in Sect.~\ref{sec:T1-inductive-def})}.
-
-\index{maths}{Ordinal numbers!Cantor normal form}
-\index{hydras}{Library Schutte!Predicates!is\_cnf\_of@is\_cnf\_of (to be a Cantor normal form of}
-
-We represent  Cantor normal forms as lists of ordinals.
-A  list $l$ is a Cantor normal form of a given ordinal $\alpha$ if it satisfies two conditions:
-
-\begin{itemize}
-\item The list  $l$ is sorted (in decreasing order) w.r.t. the order $\leq$
-\item The sum of all the  $\omega^{\beta_i}$ where the $\beta_i$ are the terms of $l$ (in this order) is equal to $\alpha$.
-\end{itemize}
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Schutte.CNF.html\#cnf_t}%
-{\texttt{Schutte.CNF}}}
-
-\input{movies/snippets/CNF/Defs}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction on $\alpha$, we prove that every countable ordinal $\alpha$ 
- has at least a Cantor normal form.
-
-\input{movies/snippets/CNF/cnfExists}
-
-By structural induction on lists, we prove that this normal form is unique.
-
-\inputsnippets{CNF/cnfUnicity,CNF/cnfExUnique}
-
-Finally, the following two lemmas relate  $\epsilon_0$ with Cantor normal forms.
-
-If $\alpha<\epsilon_0$, then the Cantor normal form of $\alpha$ is made of ordinals strictly less than $\alpha$.
-
-\input{movies/snippets/CNF/cnfLtEpsilon0}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Please consider the following statement :
-
-\begin{Coqsrc}
-Lemma cnf_lt_epsilon0_iff : 
- forall l alpha, 
-   is_cnf_of alpha l ->  
-   (alpha < epsilon0 <->  Forall (fun beta =>  beta < alpha) l).
-\end{Coqsrc}
-
-Is it true ?
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/schutte_cnf_counter_example.v}{exercises/ordinals/schutte\_cnf\_counter\_example.v}.}
-\end{exercise}
-
-Finally, the Cantor normal form of $\epsilon_0$ is just $\omega^{\epsilon_0}$.
-
-\input{movies/snippets/CNF/cnfOfEpsilon0}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Implement pages 82 to 85 of~\cite{schutte} (critical, strongly critical, maximal critical ordinals, Feferman's ordinal $\Gamma_0$).
-\end{project}
-
-\begin{remark}
-The sub-directory
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/Gamma0}{theories/oridinals/Gamma0} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form.
-\end{remark}
-
-\section{An embedding of \texttt{T1} into \texttt{Ord}}
-
-Our library 
-\href{../theories/html/hydras.Schutte.Correctness_E0.html}%
-{\texttt{Schutte.Correctness\_E0}} establishes the link between two very different modelizations of ordinal numbers. In other words, it ``validates'' a data structure in terms of
-a classical mathematical discourse considered as a model. 
-First, we define a function from \texttt{T1} into  \texttt{Ord} by structural recursion.
-
-\input{movies/snippets/Correctness_E0/injectDef}
-
-This function enjoys good commutation properties with respect to the main operations which
-allow us to build Cantor normal forms.
-
-\input{movies/snippets/Correctness_E0/commutationLemmas}
-\input{movies/snippets/Correctness_E0/injectPlus}
-\input{movies/snippets/Correctness_E0/injectMultFinR}
-
-% \begin{Coqsrc}
-% Theorem inject_mono (beta gamma : T1) :
-%   T1.lt  beta gamma -> 
-%   T1.nf beta -> T1.nf gamma -> 
-%   inject beta < inject gamma.
-
-% Theorem inject_injective (beta gamma : T1) : nf beta -> nf gamma ->
-%   inject beta = inject gamma -> beta = gamma.
-% \end{Coqsrc}
-
-Finally, we prove that \texttt{inject} is a bijection from the set of all terms of \texttt{T1} in normal form to the set 
-\texttt{members epsilon0} of the elements of \texttt{Ord} strictly less than  $\epsilon_0$.
-
-\input{movies/snippets/Correctness_E0/injectLtEpsilon0}
-\input{movies/snippets/Correctness_E0/embedding}
-
-\subsection{Remarks}
-Let us recall that the library \href{../theories/html/hydras.Schutte.Schutte.html}%
-{\texttt{Schutte}} depends on five \emph{axioms} and lies explicitly in the  
-framework of classical logic with a weak version of the axiom of choice
-(please look at the documentation of
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.ChoiceFacts.html}{\texttt{Coq.Logic.ChoiceFacts}}).
-Nevertheless, the other modules:
-\href{../theories/html/hydras.Epsilon0.Epsilon0.html}%
-{\texttt{Epsilon0}},
-\href{../theories/html/hydras.Hydra.Hydra.html}%
-{\texttt{Hydra}}, et 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}%
-{\texttt{Gamma0}}
-do not import any axioms and are really constructive.
-
-\index{hydras}{Projects}
-\begin{project}
-There is no construction of ordinal multiplication in~\cite{schutte}. 
-It would be interesting to derive this operation from Schütte's axioms,
-and prove its consistence with multiplication in ordinal notations for 
-$\epsilon_0$ and $\Gamma_0$.
-\end{project}
-
-\section{Related work}
-
-In~\cite{grimm:hal-00911710}, José Grimm establishes the consistency between our ordinal notations \texttt{T1} and \texttt{T2} (Veblen normal form) and his implementation
-of ordinal numbers after Bourbaki's set theory.
-
-The Gaia project ~\url{https://github.com/coq-community/gaia} maintains Grimm's  theory of ordinals as part of coq-community on GitHub. Integration
-of the present ordinal theory with Gaia, i.e., relating the different notions of ordinals
-and transferring relevant results, is an interesting project.
-First experiments in that direction are developped in
-the \href{https://github.com/coq-community/hydra-battles/blob/master/theories/gaia/}{theories/gaia/} directory.
-
 \chapter{The Ordinal \texorpdfstring{$\Gamma_0$}{Gamma0} (first draft)}

### 2022-05-08 doc/epsilon0-chapter.tex
reorganization of tex files

--- 2022-04-18 doc/part-hydras.tex
@@
-\section*{Introduction}
-
-In this part, we present a development  for the \coq{} proof assistant, after the work of Kirby and Paris~\cite{KP82}. This formalization contains the following main parts:
-
-\begin{itemize}
-\item Representation in \coq{} of hydras and hydra battles.
-\item A proof that every battle is finite and won by Hercules.
-This proof is based on a \emph{variant} which maps any hydra to an ordinal strictly less than $\epsilon_0$ and is strictly decreasing along any battle.
-
-\item Using a combinatorial toolkit designed by J.~Ketonen and R.~Solovay~\cite{KS81}, we prove that, for any ordinal $\mu<\epsilon_0$, there exists no such variant mapping any hydra to an ordinal strictly less than $\mu$.
-Thus, the complexity of $\epsilon_0$ is really needed in the previous proof.
-
-\item We prove a relation between the length of a ``classic''  kind of  battles \footnote{This class is also called \emph{standard} in this document (text and proofs).
-The \emph{replication factor} of the hydra is exactly $i$ at the $i$-th round of the battle (see Sect~\vref{sect:replication-def}).}
-and the Wainer-Hardy hierarchy of ``rapidly growing functions'' $H_\alpha$~\cite{Wainer1970}.
-The considered class of battles, which we call \emph{standard},  is the most considered one in the scientific  literature (including popularization).
-\end{itemize}
-
-Simply put, this document tries to combine the scientific interest of two articles~\cite{KP82, KS81} and a book~\cite{schutte} with the playful activity of truly proving theorems.
-We hope  that such a work, besides exploring a nice piece of discrete mathematics,
-will show how \coq{} and its standard library are well fitted to help us to understand some non-trivial mathematical developments, and also to experiment the constructive parts of  the proof through functional programming.
-
- We also hope to provide a little clarification on infinity (both potential and actual) through the notions of function, computation, limit,
- type and proof.
-
- \subsection*{Compatibility with \gaia (in progress)}
- \label{sect:gaia-first-intro}
-The \gaia project~\cite{Gaia} by Jos\'e Grimm, Alban Quadrat,  and Carlos Simpson,
-aims  to formalize mathematics in \coq  in
-\index{gaiabridge}{Introduction}
-
-the style of Nicolas Bourbaki.
-It contains many definitions and results about ordinal numbers.
-In Chapter~\ref{gaia-chapter}, we present some modules which allow \HydrasLib' users to apply lemmas proven in \gaia, and vice versa.
-Remarks about compatibility with \gaia are signalled with the picture \gaiasign.
-A special index is in construction (page~\pageref{gaia-index}).
-
-%\section{Remarks}
-
-\subsection*{Difference from Kirby and Paris's work}
-In~\cite{KP82}, Kirby and Paris show  that there is no proof of termination of all hydra battles in Peano Arithmetic (PA).
-Since we are used to writing proofs in higher order logic, the restriction to PA was quite unnatural for us.
-So we chose to prove another statement without any reference to PA, by considering a class of proofs indexed by ordinal numbers up to $\epsilon_0$.
-
-\subsection*{State of the development}
-The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on  notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
-We added new material: axiomatic definitions of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, function derinition by equations, etc.
-
-We are now working in order to make clumsy proofs more readable, simplify definitions, and ``factorize'' proofs as much as possible. 
-Many possible improvements are suggested as ``todo''s or ``projects'' in this text.
-
-\section*{Future work (projects)}
-\index{hydras}{Projects}
-
-This document and the proof scripts are far from being complete.
-
-First, there must be a lot of typos to correct, references and index items to add. Many proofs are too complex and should be simplified, etc.
-
-The following extensions are planned, but help is needed:
-
-\begin{itemize}
-\item Semiautomatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
-%\item Extension to $\Gamma_0$ (in Veblen normal form)
-\item More lemmas about hierarchies of rapidly growing
-  functions, and their relationship 
-    with primitive recursive functions and provability in Peano arithmetic 
-(following~\cite{KS81, KP82}).
-\item From \coq's point of view, this development could be used as an illustration of the evolution of the software, every time new libraries and sets of tactics could help to simplify the proofs.
-\end{itemize}
-
-\subsection*{Main references}
-
-In our development, we adapt the definitions and prove many theorems which
-we found in the following articles. 
-\begin{itemize}
-\item ``Accessible independence results for Peano arithmetic''  by Laurie Kirby and Jeff Paris~\cite{KP82}
-\item ''Rapidly growing Ramsey Functions'' by Jussi Ketonen and Robert Solovay~\cite{KS81}
-\item ``The Termite and the Tower'', by Will Sladek~\cite{Sladek07thetermite}
-\item Chapter V of ``Proof Theory'' by Kurt Schütte~\cite{schutte}
-\end{itemize}
-
-\chapter{Hydras and hydra games}
-
-\label{sec:orgheadline91}
-\label{chapter:hydras}
-
-This chapter is dedicated to the representation of hydras and rules of the hydra game in \coq's specification language:~\gallina.
-
-Technically, a \emph{hydra} is just a finite ordered tree, each node of which 
-has any number of sons. Contrary to the computer science tradition, we display the hydras 
-with the heads up and the foot (i.e., the root of the tree) down.
-Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the 
-module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}). 
-\emph{For a less formal description of hydras, please see 
-\url{https://www.smbc-comics.com/comic/hydra}.}
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {$\bullet$};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H4) at (4,2) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N3) to [bend left= 10]  (H1) ;
-\draw (N3) to [bend right= 16] (H2);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (foot) to [bend right = 10] (H4) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-\caption{The hydra \texttt{Hy} \label{fig:Hy}}
-\end{figure}
-
-We use a specific vocabulary for talking about hydras. Table~\ref{tab:hyd2tree} shows the correspondence between our terminology and the usual vocabulary for trees in computer science.
-
-\begin{figure}[h]
-  \centering
-  \begin{tabular}{ll}
-Hydras & Finite rooted trees\\
-\hline
-foot & root\\
-head & leaf\\
-node & node\\
-segment  & (directed) edge \\
-sub-hydra & subtree\\
-daughter & immediate subtree\\
-\end{tabular}
-  \caption{Translation from hydras to trees}
-  \label{tab:hyd2tree}
-\end{figure}
-
-The hydra \texttt{Hy} has a \emph{foot} (below), five \emph{heads}, and eight \emph{segments}. 
-We leave it to the reader to define various parameters such as the height, the size, the highest arity (number of sons of a node) of a hydra.
-In our example, these parameters have the respective values $4$, $9$ and $3$.
-
-\subsection{The rules of the game}
-
-\label{sec:orgheadline44}
-\label{sect:replication-def}
-
-A \emph{hydra battle} is a fight between Hercules and the Hydra. 
-More formally, a  battle is a sequence of \emph{rounds}.
-At each round:
-\begin{itemize}
-\item If the hydra is composed of just one head, the battle is finished
-and  Hercules is the winner.
-\item Otherwise, Hercules chops off \emph{one} head of the hydra,
-
-\begin{itemize}
-\item If the head is at distance 1 from the foot, the head is just lost by the hydra, with no more reaction.
-\item Otherwise, let us denote by \(r\) the node that was at distance \(2\) from 
-the removed head in the direction of the foot,  and consider the  sub-hydra \(h'\) of \(h\), whose  root is \(r\) \footnote{$h'$ will be called ``the wounded part of the hydra'' in the subsequent text.
-In Figures~\vref{fig:Hy2} and ~\vref{fig:Hy4}, this sub-hydra  is displayed in red.}.
-Let $n$ be some natural number.
-Then $h'$ is replaced by  $n+1$ of copies of \(h'\) which share the same root $r$.
- The \emph{replication factor} $n$ may be different (and generally is)   at each round of the fight.
-It may be chosen by the hydra, according to its strategy, or imposed by some 
-particular rule.
-In many presentations of hydra battles, this number is increased by $1$ at each round.
-In the following presentation, we will also consider battles where the hydra is free to choose its ~replication factor at each round of the battle\footnote{Let us recall that, if the chopped-off head was at distance 1 from the foot, the replication factor is meaningless.}.
-\end{itemize}
-\end{itemize}
-
-Note that the description given in~\cite{KP82} of the replication process in hydra battles is also  semi-formal.
-
-\label{original-rules}
-
-\begin{quote}
-  ``From the node that used to be attached to the head which was just chopped off, traverse one 
-segment towards the root until the next node is reached. From this node sprout $n$ replicas of 
-that part of the hydra (after decapitation) which is ``above'' the segment just traversed, i.e., those 
-nodes and segments from which, in order to reach the root, this segment would have to be 
-traversed. If the head just chopped off had the root of its nodes, no new head is grown. ''
-\end{quote}
-
-Moreover, we note that this description is in \emph{imperative} terms.
-In order to formally study the properties of hydra battles, we prefer to use a mathematical vocabulary, i.e., graphs, relations, functions, etc.
-Thus, the replication process will be represented as a binary relation on a data type \texttt{Hydra},
-linking the state of the hydra \emph{before} and \emph{after} the transformation.
-A battle will thus be represented as a sequence of terms of type \texttt{Hydra}, respecting the rules of the game.
-
-\subsection{Example}
-Let us start a battle between Hercules and the hydra \texttt{Hy} of Fig.~\ref{fig:Hy}.
-
-At the first round, Hercules chooses to chop off the rightmost head of \texttt{Hy}.
-Since this head is near the floor, the hydra simply loses this head. Let us call 
- \texttt{Hy'} the resulting state of the hydra, represented in Fig.~\vref{fig:Hy-prime}.
-
-Next, assume Hercules chooses to chop off one of the two highest heads of \texttt{Hy'}, for instance the rightmost one.
-Fig.~\vref{fig:Hy2} represents the broken segment in dashed lines, and the part that will be replicated in red.
-Assume also that the hydra decides to add 4 copies of the red part\footnote{In other words, the replication factor at this round is equal to $4$.}.
-We obtain a new state \texttt{Hy''} depicted in Fig.~\ref{fig:Hy3}.
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {$\bullet$};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-%\node (H4) at (6,0) {$\Xey[2][lightgray]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3) ;
-\draw (N3) to [bend left= 10]  (H1) ;
-\draw (N3) to [bend right= 16] (H2);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-
-\caption{\texttt{Hy'}: the state  of \texttt{Hy} after one round \label{fig:Hy-prime}}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4)  {{\color{lightred}$\bullet$}};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Sey[2][lightred]$};
-%\node (H2) at (5,0) {$\Xey[2][lightgray]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\node (ex) at (5,8) {};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw  (N2) -- (N3)[draw=lightred];
-\draw (N3) to   [bend left= 10](H1) [draw=lightred];
-\draw [dashed] (N3) to [bend left= 10](ex);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-\caption{A second beheading}
-\label{fig:Hy2}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H11) at (2,8) {$\Smiley[2][vertfluo]$};
-\node (H12) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H13) at (6,8) {$\Smiley[2][vertfluo]$};
-\node (H14) at (8,8) {$\Smiley[2][vertfluo]$};
-
-\node (N3) at (1,6) {$\bullet$};
-\node (N31) at (2,6) {$\bullet$};
-\node (N32) at (3,6) {$\bullet$};
-\node (N33) at (4,6) {$\bullet$};
-\node (N34) at (5,6) {$\bullet$};
-
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N2) -- (N31);
-\draw (N2) -- (N32);
-\draw (N2) -- (N33);
-\draw (N2) -- (N34);
-\draw (N3) to   [bend left= 10](H1) ;
-\draw (N31) to   [bend left= 10](H11) ;
-\draw (N32) to   [bend left= 10](H12) ;
-\draw (N33) to   [bend left= 10](H13) ;
-\draw (N34) to   [bend left= 10](H14) ;
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend left= 10]  (H5) ;
-\end{tikzpicture}
-\caption{\texttt{Hy''}: the state of \texttt{Hy} after two rounds \label{fig:Hy3}}
-\end{figure}
-
-Figs.~\ref{fig:Hy4} and~\vref{fig:Hy5} represent a possible third round of the battle, with a replication factor equal to $2$. Let us call \texttt{Hy'''} the state of the hydra after that third round.
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0)  {{\color{lightred}$\bullet$}};
-\node (N1) at (2,2) {{\color{lightred}$\bullet$}};
-\node (N2) at (2,4) {{\color{lightred}$\bullet$}};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (exN4) at (4,4) {};
-\node (H1) at (0,8) {$\Sey[2][lightred]$};
-\node (H11) at (2,8) {$\Sey[2][lightred]$};
-\node (H12) at (4,8) {$\Sey[2][lightred]$};
-\node (H13) at (6,8) {$\Sey[2][lightred]$};
-\node (H14) at (8,8) {$\Sey[2][lightred]$};
-
-\node (N3) at (1,6) {{\color{lightred}$\bullet$}};
-\node (N31) at (2,6) {{\color{lightred}$\bullet$}};
-\node (N32) at (3,6) {{\color{lightred}$\bullet$}};
-\node (N33) at (4,6) {{\color{lightred}$\bullet$}};
-\node (N34) at (5,6) {{\color{lightred}$\bullet$}};
-
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-%\node (H5) at (4,0) {$\Xey[2][lightgray]$};
-\draw (foot) -- (N1)[very thick,draw=lightred] ;
-\draw (N1) -- (N2)[draw=lightred];
-\draw (N2) -- (N3)[draw=lightred];
-\draw (N2) -- (N31)[draw=lightred];
-\draw (N2) -- (N32)[draw=lightred];
-\draw (N2) -- (N33)[draw=lightred];
-\draw (N2) -- (N34)[draw=lightred];
-\draw (N3) to   [bend left= 10](H1) [draw=lightred];
-\draw (N31) to   [bend left= 10](H11) [draw=lightred];
-\draw (N32) to   [bend left= 10](H12) [draw=lightred];
-\draw (N33) to   [bend left= 10](H13) [draw=lightred];
-\draw (N34) to   [bend left= 10](H14) [draw=lightred];
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw [dashed] (N1) to  [bend left= 10](exN4);
-\end{tikzpicture}
-\caption{A third beheading (wounded part in red) \label{fig:Hy4}}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-
-\node (foot) at (10,0) {$\bullet$};
-
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H1) at (0,8) {$\Smiley[1][vertfluo]$};
-\node (H11) at (2,8) {$\Smiley[1][vertfluo]$};
-\node (H12) at (4,8) {$\Smiley[1][vertfluo]$};
-\node (H13) at (6,8) {$\Smiley[1][vertfluo]$};
-\node (H14) at (8,8) {$\Smiley[1][vertfluo]$};
-
-\node (N3) at (1,6) {$\bullet$};
-\node (N31) at (2,6) {$\bullet$};
-\node (N32) at (3,6) {$\bullet$};
-\node (N33) at (4,6) {$\bullet$};
-\node (N34) at (5,6) {$\bullet$};
-
-\node (H0) at (-3,3) {$\Smiley[1][vertfluo]$};
-
-\draw (foot) to [bend left=10] (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N2) -- (N31);
-\draw (N2) -- (N32);
-\draw (N2) -- (N33);
-\draw (N2) -- (N34);
-\draw (N3) to   [bend left= 10](H1) ;
-\draw (N31) to   [bend left= 10](H11) ;
-\draw (N32) to   [bend left= 10](H12) ;
-\draw (N33) to   [bend left= 10](H13) ;
-\draw (N34) to   [bend left= 10](H14) ;
-\draw (foot) to [bend left = 15]  (H0) ;
-
-% second copy 
-\node (N01) at (12,2) {$\bullet$};
-\node (N02) at (12,4) {$\bullet$};
-\node (N03) at (12,6) {{\color{lightred}$\bullet$}};
-\node (H001) at (10,8) {$\Smiley[1][vertfluo]$};
-\node (H0011) at (12,8) {$\Smiley[1][vertfluo]$};
-\node (H0012) at (14,8) {$\Smiley[1][vertfluo]$};
-\node (H0013) at (16,8) {$\Smiley[1][vertfluo]$};
-\node (H0014) at (18,8) {$\Smiley[1][vertfluo]$};
-
-\node (N03) at (11,6) {$\bullet$};
-\node (N031) at (12,6) {$\bullet$};
-\node (N032) at (13,6) {$\bullet$};
-\node (N033) at (14,6) {$\bullet$};
-\node (N034) at (15,6) {$\bullet$};
-
-\draw (foot) -- (N01)[very thick] ;
-\draw (N01) -- (N02);
-\draw (N02) -- (N03);
-\draw (N02) -- (N031);
-\draw (N02) -- (N032);
-\draw (N02) -- (N033);
-\draw (N02) -- (N034);
-\draw (N03) to   [bend left= 10](H001) ;
-\draw (N031) to   [bend left= 10](H0011) ;
-\draw (N032) to   [bend left= 10](H0012) ;
-\draw (N033) to   [bend left= 10](H0013) ;
-\draw (N034) to   [bend left= 10](H0014) ;
-
-% third copy 
-\node (N001) at (22,2) {$\bullet$};
-\node (N002) at (22,4) {$\bullet$};
-\node (N003) at (22,6) {{\color{lightred}$\bullet$}};
-\node (H001) at (20,8) {$\Smiley[1][vertfluo]$};
-\node (H0011) at (22,8) {$\Smiley[1][vertfluo]$};
-\node (H0012) at (24,8) {$\Smiley[1][vertfluo]$};
-\node (H0013) at (26,8) {$\Smiley[1][vertfluo]$};
-\node (H0014) at (28,8) {$\Smiley[1][vertfluo]$};
-
-\node (N003) at (21,6) {$\bullet$};
-\node (N0031) at (22,6) {$\bullet$};
-\node (N0032) at (23,6) {$\bullet$};
-\node (N0033) at (24,6) {$\bullet$};
-\node (N0034) at (25,6) {$\bullet$};
-
-\draw (foot) -- (N001)[very thick] ;
-\draw (N001) -- (N002);
-\draw (N002) -- (N003);
-\draw (N002) -- (N0031);
-\draw (N002) -- (N0032);
-\draw (N002) -- (N0033);
-\draw (N002) -- (N0034);
-\draw (N003) to   [bend left= 10](H001) ;
-\draw (N0031) to   [bend left= 10](H0011) ;
-\draw (N0032) to   [bend left= 10](H0012) ;
-\draw (N0033) to   [bend left= 10](H0013) ;
-\draw (N0034) to   [bend left= 10](H0014) ;
-\end{tikzpicture}
-\caption{The configuration \texttt{Hy'''} of \texttt{Hy} \label{fig:Hy5}}
-\end{figure}
-\FloatBarrier
-
-We leave it to the reader  to guess the following  rounds of the battle \dots
- % Please keep in mind that, in this 
-% the hydra is free to chose any number of replications at each time, whereas
-% Hercules chops only one head per round.
-
-% Let us precise that, in this game, Hercules wins if the hydra is eventually reduced 
-% to a single head. 
-% We know from~\cite{KP82} that, whichever the initial configuration of the
-% hydra, and the strategies of both players, Hercules eventually wins. The 
-% aforementionned paper shows also that there do not exist any \emph{simple} proof of this result.
-
-\section{Hydras and their representation in \emph{Coq}}
-\label{sec:orgheadline48}
-
-\index{hydras}{Library Hydra!Types!Hydra}
-\index{hydras}{Library Hydra!Types!Hydrae}
-
-In order to describe trees where each node can have an arbitrary (but finite) number of sons, it is usual to define a type where each node carries a \emph{forest}, \emph{i.e} a list of trees
-(see for instance Chapter 14, pages 400-406 of \cite{BC04}, also available as~\cite{BC04ch14}).
-
-For this purpose, we define two mutual \emph{ad-hoc}  inductive types, where \texttt{Hydra} is the main type, and \texttt{Hydrae} is a helper for describing finite sequences of hydras.
-\label{types:Hydra}
-\label{types:Hydrae}
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hydra}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/HydraDef}
-
-%\index{To do}
-\index{hydras}{Projects}
-
-\begin{project}
-Look for an existing library on trees with nodes of arbitrary arity, in order to replace  this ad-hoc type with something more generic.
-\end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-
- Another very similar representation could use the \texttt{list} type family instead of the specific 
-type \texttt{Hydrae}:
-
-\input{movies/snippets/Hydra_Definitions/HydraAlt}
-
-Using this representation, re-define all the constructions of this chapter.
-You will probably have to use patterns described for instance in~\cite{BC04} or the archives of the \coq communication channels (please consult~\url{https://coq.inria.fr/community.html}).
-
-\end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-The type \texttt{Hydra} above describes hydras as \emph{plane trees}, i.e., as drawn on a sheet of paper or computer screen. Thus, hydras are \emph{oriented},
-and it is appropriate to consider a \emph{leftmost} or \emph{rightmost} head of
-the beast. It could be interesting to consider another representation, in which
-every non-leaf node has a \emph{multi-set} -- not an ordered list -- of daughters.
-\end{project}
-
-\subsubsection{Abbreviations}
-
-We provide several notations for hydra patterns  which occur often in our developments.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#head}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/headsEtc}
-
-For instance, the hydra \texttt{Hy}  of Figure~\vref{fig:Hy} is defined in \emph{Gallina} as follows:
-
-\vspace{4mm}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html\#Hy}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/Hy}
-
-Hydras quite frequently contain  multiple adjacent  copies of the same subtree. The following functions
-will help us to describe and reason about replications in hydra battles.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#hcons_mult}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hconsMult}
-
-\vspace{4mm}
-
-For instance, the hydra \texttt{Hy''} of Fig~\vref{fig:Hy3}  can be defined in \coq{} as follows:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/HySecond}
-
-\subsubsection{Recursive functions on type \texttt{Hydra}}
-\label{sec:orgheadline41}
-\label{sec:hsize-def}
-
-In order to  define a recursive function over the type \texttt{Hydra}, one has to consider the three constructors 
-\texttt{node}, \texttt{hnil} and \texttt{hcons} of the mutually inductive types \texttt{Hydra} and \texttt{Hydrae}. 
-Let us define for instance the function which  computes the number of nodes of any hydra:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hsize}
-
-\input{movies/snippets/Hydra_Examples/HySize}
-
-Likewise, the \emph{height} (maximum distance between the foot and a head) 
-is defined by mutual recursion:
-
-\input{movies/snippets/Hydra_Definitions/height}
-
-\input{movies/snippets/Hydra_Examples/HyHeight}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Define a function \texttt{max\_degree: Hydra $\arrow$ nat} which  returns the highest degree of a node in any hydra.
-For instance, the evaluation of the term \texttt{(max\_degree Hy)} should return $3$.
-\end{exercise}
-
-\subsection{Induction principles for hydras}
-\label{sec:orgheadline42}
-
-In this section, we show how induction principles are used to prove properties on the type 
-\texttt{Hydra}. Let us consider for instance the following statement:
-\begin{quote}
-  `` The height of any hydra is strictly less than its size. ''
-\end{quote}
-
-\subsubsection{A failed attempt}
-
-One may try to use the default tactic of proof by induction, which corresponds to an application of the automatically  generated  induction principle for  type \texttt{Hydra}:
-
-\input{movies/snippets/Hydra_Examples/HydraInd}
-
-Let us start a simple proof by induction.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/BadInductiona}
-
-We might be tempted to do an induction on the sequence \texttt{s}:
-
-\input{movies/snippets/Hydra_Examples/BadInductionb}
-
-The first subgoal is trivial.
-
-\input{movies/snippets/Hydra_Examples/BadInductionc}
-
-Let us look at the second subgoal of the induction.
-
-\input{movies/snippets/Hydra_Examples/BadInductiond}
-
-We notice that this subgoal does not contain any hypothesis
-on the height and size of the hydra \texttt{h}. So, it looks hard to prove the conclusion. Let's stop.
-
-\input{movies/snippets/Hydra_Examples/BadInductione}
-
-\subsubsection{A Principle of mutual induction}
-In order to get an appropriate induction scheme for the types 
-\texttt{Hydra} and \texttt{Hydrae}, we can use  \coq{}'s  command \texttt{Scheme}.
-
-\index{coq}{Mutually inductive types}
-%\index{coq}{Commands!Scheme}
-
-\input{movies/snippets/Hydra_Definitions/HydraRect2}
-
-\input{movies/snippets/Hydra_Examples/HydraRect2Check}
-
-\subsubsection{A Correct proof}
-
-Let us now use \texttt{Hydra\_rect2} for proving that the height of any hydra is strictly less than its size.
-Using this scheme requires an auxiliary predicate, called \texttt{P0} in \texttt{Hydra\_rect2}'s statement.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hForall}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-\input{movies/snippets/Hydra_Examples/heightLtSizea}
-
-The first subgoal is easily solvable, using some arithmetic.
-The second and third ones are
-almost trivial. We let the reader look at the source. 
-\input{movies/snippets/Hydra_Examples/heightLtSizez}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-It happens very often that, in the proof of  a proposition of the form 
-(\texttt{$\forall\,$ h:Hydra, $P$ h}), the predicate \texttt{P0}
-is  (\texttt{h\_forall $P$}).
-Design a tactic for induction on hydras that frees the user from binding explicitly \texttt{P0},  and solves trivial subgoals.
-Apply it for writing  a shorter proof of \texttt{height\_lt\_size}.
-\end{exercise}
-
-\section{Relational description of hydra battles}
-
-In this section, we represent the rules of hydra battles as a binary relation associated with
-a \emph{round}, i.e., an interaction composed of the two following actions:
-\begin{enumerate}
-\item Hercules chops off one head of the hydra.
-\item Then, the  hydra replicates the wounded part (if the head is at distance $\geq 2$ from the foot).
-\end{enumerate}
-The relation associated with each round of the battle is parameterized  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot,
-but present for consistency's sake).
-
-In our description,  we will apply the following naming convention: if $h$ represents the configuration of the hydra before a round, then the configuration of $h$ after this round will be called $h'$.
- Thus, we are going to define a proposition  (\texttt{round\_n $n\;h\;h'$})  whose intended meaning will be `` the hydra $h$  is transformed into $h'$  in a single round of a battle, with the expected replication factor $n$ ''.
-
-Since the replication of parts of the hydra depends on the distance of the chopped head from  the foot, we  decompose our description into two main  cases, under the form of a bunch of [mutually] inductive predicates over the types \texttt{Hydra} and \texttt{Hydrae}.
-
-The mutually exclusive cases we consider are the following:
-\begin{itemize}
-\item \textbf{R1}: The chopped off head was at distance 1 from the foot.
-\item \textbf{R2}: The chopped off head was at a distance greater than or equal to  $2$ from the foot.
-\end{itemize}
-
-\subsection{Chopping off a head at distance 1 from the foot (relation  R1)}
-
-If Hercules chops off a head near the floor, there is no replication at all. We use an auxiliary 
-predicate \texttt{S0}, associated with the removing of one head from a sequence of hydras.
-
-\vspace{4pt}\emph{From Module\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/S0Def}
-\input{movies/snippets/Hydra_Definitions/R1Def}
-
-\subsubsection{Example}
-\label{sec:orgheadline45}
-
-Let us represent in \coq{}   the transformation of the hydra of Fig.~\vref{fig:Hy} into
-the configuration represented in Fig.~\ref{fig:Hy-prime}.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/Hy1}
-
-\subsection{Chopping off a head at distance \texorpdfstring{$\geq 2$}{>= 2} from the foot (relation R2) }
-
-Let us now consider beheadings  where the chopped-off head is at distance greater than or equal to $2$ from the foot. All the following relations are parameterized by the replication factor  $n$.
-
- Let $s$ be a sequence of hydras. 
-The proposition (\texttt{S1 n s s'}) holds if $s'$ is obtained by replacing some element $h$ of $s$ by 
-$n+1$ copies of $h'$, where  the proposition (\texttt{R1 h h'}) holds, in other words, $h'$ is just $h$, without the chopped-off  head.
-\texttt{S1} is an inductive relation with two constructors that allow us to choose the position in $s'$ of the wounded sub-hydra $h$.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#S1}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/S1Def}
-
-The rest of the definition is composed of two mutually inductive relations on hydras and sequences of hydras.
-The first constructor of \texttt{R2} describes the case where the chopped head is exactly at height $2$.
-The others constructors allow us to consider beheadings at height strictly greater than $2$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/R2Def}
-
-\subsubsection{Example}
-Let us prove the transformation of \texttt{Hy'} into \texttt{Hy''} (see Fig.~\vref{fig:Hy3}).
-We use an experimental set of tactics\footnote{See the \texttt{Ltac} definitions in~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions.}}  for specifying the place where the
-interaction between Hercules and the hydra holds.
-
-\vspace{4pt}\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}.
-
-\input{movies/snippets/Hydra_Examples/R2Example}
-
-The reader is encouraged to look at all the successive subgoals of this example.
-\emph{Please consider also exercise~\vref{exo:interactive-battle}.}
-
-\subsection{Binary relation associated with a round}
-
-We combine the two cases above into a single relation.
-First,  we define the  relation \texttt{(round\_n n h h')} where \texttt{n} is the expected number of  replications (irrelevant in the case of an \texttt{R1}-transformation).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#round_n}{Hydra.Hydra\_Definitions}}
-
-\index{hydras}{Library Hydra!Predicates!round\_n}
-
-\input{movies/snippets/Hydra_Definitions/roundNDef}
-
-By abstraction over \texttt{n}, we define a \emph{round} (small step) of a battle:
-
-\index{hydras}{Library Hydra!Predicates!round}
-\label{sect:infix-round}
-
-\input{movies/snippets/Hydra_Definitions/roundDef}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Give a direct translation of Kirby and Paris's description of hydra battles (quoted on page~\pageref{original-rules}) and prove that our relational description is consistent with theirs.
-\end{project}
-
-\subsection{Rounds and battles}
-
-Using library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}, we define \texttt{round\_plus},  the transitive closure of \texttt{round}, and \texttt{round\_star},  the reflexive and transitive closure of \texttt{round}.
-
-\label{sect:infix-rounds}
-
-\input{movies/snippets/Hydra_Definitions/roundPlus}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-  Prove that if \texttt{$h$ -+-> $h'$}, then
-  the height of $h'$ is less or equal than the height of $h$.
-
-\end{exercise}
-
-\begin{remark}
-\label{remark:transitive-closure}
-\coq's library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Coq.Relations.Relation\_Operators} 
-contains three logically equivalent definitions of the transitive closure of a binary relation. This equivalence is proved in 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} .
-
-Why three definitions for a single mathematical concept?
-Each definition generates an associated induction principle. 
- According to the form of statement one would like to prove, there is a ``best choice'':
-
-\begin{itemize}
-\item For proving $\forall y, x\,R^+\,y \;\arrow\; P\,y$, prefer 
-\texttt{clos\_trans\_n1}
-\item For proving $\forall x,\,x\,R^+\,y \;\arrow\; P\,x$, prefer \texttt{clos\_trans\_1n}
-\item For proving $\forall x\,y, \,x\,R^+\,y \;\arrow\;P\,x\,y$,  
-prefer \texttt{clos\_trans},
-\end{itemize}
-But there is no ``wrong choice'' at all: the equivalence lemmas in \linebreak 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} 
- allow the user
-to convert any one of the three closures into another one before applying the corresponding elimination tactic.
-The same remark also holds for reflexive and transitive closures. 
-\end{remark}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Define a restriction of \coqsimple{round},  where Hercules always chops off
-the leftmost among the lowest heads.
-
-Prove that, if $h$ is not a simple head, then there exists a unique $h'$ such that $h$  is transformed into $h'$ in one round, according to this restriction.
-
-\end{exercise}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}[Interactive battles]
-\label{exo:interactive-battle}
-Given a hydra \texttt{h}, the specification of a hydra battle for \texttt{h} is the type 
-\Verb@{h':Hydra | h -*-> h'}@. In order to avoid long sequences of \texttt{split}, \texttt{left}, and 
-\texttt{right}, design a set of dedicated tactics for the interactive building of a battle.
-Your tactics will have the following functionalities:
-\begin{itemize}
-\item  Chose to stop a battle, or continue
-\item Chose an expected number of replications
-\item Navigate in a hydra, looking for a head to chop off.
-\end{itemize}
-
-Use your tactics for simulating a small part of a hydra battle, for instance the rounds which lead from
-\texttt{Hy} to \texttt{Hy'''}  (Fig.~\vref{fig:Hy5}).
-
-\textbf{Hints:} 
-\begin{itemize}
-
-\item Please keep in mind that the last  configuration of your interactively built battle is known only at the end of the battle.
-Thus, you will have to create and solve subgoals with existential variables.
-For that purpose, the tactic \texttt{eexists}, applied to the
-goal \Verb@{h':Hydra | h -*-> h'}@ generates the subgoal \Verb|h -*-> ?h'|.
-\item You may use Gérard Huet's \emph{zipper} data structure~\cite{zipper} for writing tactics associated with Hercules's  interactive search for a head to chop off.
-\end{itemize}
-
-\end{exercise}
-
-\subsection{Classes of battles}
-\label{sect:battle-classes}
-
-In some presentations of hydra battles, e.g.~\cite{KP82, bauer2008}, the transformation associated with the $i$-th round may depend on $i$.
-For instance, in these articles, the replication factor at the $i$-th round is equal to $i$.
-In other examples, one can allow the hydra to apply any replication factor at any time.
-In order to be the most general as possible, we define the type of predicates which relate the state of the hydra before and after the $i$-th round of a battle.
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-\label{types:Battle}
-\index{hydras}{Library Hydra!Type classes!Battle}
-
-\input{movies/snippets/Hydra_Definitions/BattleDef}
-
-The most general class of battles is \texttt{free}, which allows the hydra to chose any replication factor at every step:
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#free}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/freeDef}
-
-We chose to call \emph{standard}\footnote{This appellation is ours.
-If there is a better one, we will change it.} the kind of battles which appear  most often in the literature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#standard}{Hydra.Hydra\_Definitions}}
-\input{movies/snippets/Hydra_Definitions/standardDef}
-
-\subsection{Big steps}
-
-Let $B$ be any instance of class \texttt{Battle}. It is easy to define inductively the relation between the $i$-th and the $j$-th steps of a battle of type $B$.
-
-\vspace{6pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/battleRelDef}
-
-The following property allows us to build battles by composition of smaller ones.
-
-%% TODO display subgoals when fixed
-
-\vspace{6pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-\noindent
-\input{movies/snippets/Hydra_Lemmas/battleTrans}
-
-% \begin{remark}
-%  The class \texttt{free} is strongly related with the transitive closure  \texttt{round\_plus}, as expressed by the following lemmas.
-
-% \vspace{4pt}
-% \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-
-%  \begin{Coqsrc}
-%  Lemma battle_free_equiv1 : forall i j h h',  
-%              battle free i h j h' ->   h -+-> h'.
-
-%  Lemma battle_free_equiv2 : forall h h',
-%      h -+-> h' ->
-%     forall i, exists j,  battle free i h j h'.
-%  \end{Coqsrc}
-
-% \end{remark}
-
-\section{A long battle}
-\label{sect:big-battle}
-
-In this section we consider a simple example of battle, starting with a small hydra,
-shown on figure~\vref{fig:hinit}, with a simple strategy for both players:
-
-\begin{itemize}
-\item At each round, Hercules chops off the rightmost head of the hydra.
-\item The battle is {standard}: at the round number $i$, the expected replication is $i$.
-\end{itemize}
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (6,0) {$\bullet$};
-\node (n1) at  (3,3) {$\bullet$};
-\node (h1) at  (1,6) {$\Smiley[1][green]$};
-\node (h2) at  (3,6) {$\Smiley[1][green]$};
-\node (h3) at  (6,6) {$\Smiley[1][green]$};
-\node (h4) at  (6,6) {$\Smiley[1][green]$};
-\node (h5) at  (6,3) {$\Smiley[1][green]$};
-\node (h6) at  (9,3) {$\Smiley[1][green]$};
-\draw (foot) -- (n1);
-\draw (n1) to   [bend left=20] (h1);
-\draw (n1) to   (h2);
-\draw (n1) to   [bend right=20] (h3);
-\draw (foot) -- (h5);
-\draw (foot) to  [bend right=20] (h6);
-\end{tikzpicture}
-
-  \caption{The hydra \texttt{hinit}}
-  \label{fig:hinit}
-\end{figure}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.BigBattle.html}{Hydra.BigBattle}}
-
-\input{movies/snippets/BigBattle/hinitDef}
-
-The lemma we would like to prove is ``The considered battle lasts exactly $N$ rounds'',
-with $N$ being a natural number we gave to guess.
-
-But the  battle is so long that no \emph{test} can give us any estimation of its length, and we do need the expressive power of logic to compute this length.
-However, in order to  guess this length, we made some experiments, computing with \gallina{}, \coq{}'s  functional programming language.
-Thus, we can consider this development as a collaboration of proof with computation.
-In the following lines, we show how we found experimentally the value of the number $N$.
-The complete proof is in file \url{../theories/html/hydras.Hydra.BigBattle.html}.
-
-\subsection{First rounds}
-During the two first rounds, our hydra loses its two rightmost heads.  Figure~\vref{fig:hinit-plus2} shows the state of the hydra   just before the third round.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (3,0) {$\bullet$};
-\node (n1) at  (3,3) {$\bullet$};
-\node (h1) at  (1,6) {$\Smiley[1][green]$};
-\node (h2) at  (3,6) {$\Smiley[1][green]$};
-\node (h3) at  (6,6) {$\Smiley[1][green]$};
-\node (h4) at  (6,6) {$\Smiley[1][green]$};
-\draw (foot) -- (n1);
-\draw (n1) to   [bend left=20] (h1);
-\draw (n1) to   (h2);
-\draw (n1) to   [bend right=20] (h3);
-\end{tikzpicture}
-
-  \caption{The hydra (\texttt{hyd1 h3})}
-  \label{fig:hinit-plus2}
-\end{figure}
-
-The following lemma  is a formal description of these first rounds, in terms of the
-\texttt{battle} predicate.
-
-\input{movies/snippets/BigBattle/L02}
-
-\subsection{Looking for regularities}
-
-A first study with pencil and paper suggested us that, after three rounds, the hydra always looks like in figure~\vref{fig:hinit-plusn} (with a variable number of 
-subtrees of height 1 or 0).
-Thus, we introduce a few handy abbreviations.
-
-\input{movies/snippets/BigBattle/Notations}
-
-For instance, the hydra shown in  Fig~\ref{fig:hinit-plusn} is  (\texttt{hyd 3 4 2}). The hydra (\texttt{hyd 0 0 0})  is the ``final'' hydra of any terminating battle, {i.e.},
-a tree whith exactly one node and no edge.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (15,0) {$\bullet$};
-\node (a) at  (3,4) {$\bullet$};
-\node (b) at  (6,4) {$\bullet$};
-\node (c) at  (9,4) {$\bullet$};
-\node (d) at  (13,4) {$\bullet$};
-\node (e) at  (16,4) {$\bullet$};
-\node (f) at  (19,4) {$\bullet$};
-\node (g) at  (22,4) {$\bullet$};
-\node (h) at  (25,4) {$\Smiley[1][green]$};
-\node (i) at  (28,4) {$\Smiley[1][green]$};
-\node (aa) at  (2.5,8) {$\Smiley[1][green]$};
-\node (ab) at  (3.5,8) {$\Smiley[1][green]$};
-\node (ba) at  (5.5,8) {$\Smiley[1][green]$};
-\node (bb) at  (6.5,8) {$\Smiley[1][green]$};
-\node (ca) at  (8.5,8) {$\Smiley[1][green]$};
-\node (cb) at  (9.5,8) {$\Smiley[1][green]$};
-\node (da) at  (13,8) {$\Smiley[1][green]$};
-\node (ea) at  (16,8) {$\Smiley[1][green]$};
-\node (fa) at  (19,8) {$\Smiley[1][green]$};
-\node (ga) at  (22,8) {$\Smiley[1][green]$};
-\draw (foot) -- (a);
-\draw (foot) -- (b);
-\draw (foot) -- (c);
-\draw (foot) -- (d);
-\draw (foot) -- (e);
-\draw (foot) -- (f);
-\draw (foot) -- (g);
-\draw (foot) -- (h);
-\draw (foot) -- (i);
-\draw (a) -- (aa);
-\draw (a) -- (ab);
-\draw (b) -- (ba);
-\draw (b) -- (bb);
-\draw (c) -- (ca);
-\draw (c) -- (cb);
-\draw (d) -- (da);
-\draw (e) -- (ea);
-\draw (f) -- (fa);
-\draw (g) -- (ga);
-\end{tikzpicture}
-
-  \caption{The hydra (\texttt{hyd 3 4 2})}
-  \label{fig:hinit-plusn}
-\end{figure}
-
-With these notations, we get a formal description of the first three rounds.
-
-\input{movies/snippets/BigBattle/L23L03}
-
-\subsection{Testing  \dots}
-\label{sect:testing}
-In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple data type for representing the states as tuples composed of
-the round number, and the respective number of daughters  \texttt{h2}, \texttt{h1}, and heads
-of the current hydra.
-
-\input{movies/snippets/BigBattle/stateDef}
-
-The following function returns the next configuration of the game.
-Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle only start with the lemma
-\texttt{step\_battle}, page~\pageref{lemma:step-battle}.
-
-\input{movies/snippets/BigBattle/nextDef}
-
-We can make bigger steps through iterations of \texttt{next}.
-The functional \texttt{iterate}, similar to Standard Library's \texttt{Nat.iter},
-is defined and studied in~\href{../theories/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
-\index{hydras}{Library Prelude!iterate}
-
-\label{Functions:iterate}
-
-\input{movies/snippets/Iterates/iterateDef}
-
-The following function computes the state of the battle at the $n$-th round.
-
-\input{movies/snippets/BigBattle/testDefTests}
-
-The battle we are studying looks to be awfully long. Let us concentrate our
-tests on some particular events : the states where $\texttt{nh}=0$.
-From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} is equal to zero.
-
-\input{movies/snippets/BigBattle/smartTest}
-
-Thus, $(1 + 11)$ rounds later, the \texttt{n1} field is equal to $2$, and 
-\texttt{nh}   to $0$.
-
-\input{movies/snippets/BigBattle/smartTestb}
-
-Next round, we decrement \texttt{n2} and set \texttt{n1} to $95$.
-
-\input{movies/snippets/BigBattle/smartTestc}
-
-We now have some intuition of the sequence.
-It looks like the next ``\texttt{nh}=0'' event will happen at the $192=2(95+1)$-th round, then at the $2(192+1)$-th round, etc.
-
-\input{movies/snippets/BigBattle/doubleS}
-
-\subsection{Proving \dots}
-We are now able to reason about the sequence of transitions defined by our hydra battle. Instead of using the data-type \texttt{state} we study the relationship
-between different configurations of the battle.
-
-Let us define a binary relation associated with every round of the battle.
-In the following definition \texttt{i} is associated with the round number (or date, if we consider a discrete time), and \texttt{a}, \texttt{b}, \texttt{c} respectively associated with the number of occurrences of \texttt{h2}, \texttt{h1} and heads connected to the hydra's foot.
-
-\input{movies/snippets/BigBattle/oneStep}
-
-The relation between \texttt{one\_step} and the rules of hydra battles is asserted by the following lemma.
-
-\label{lemma:step-battle}
-
-\input{movies/snippets/BigBattle/stepBattle}
-
-\vspace{4pt}
-
-Next, we define ``big steps'' as the transitive closure of \texttt{one\_step},
-and reachability (from the initial configuration of figure~\ref{fig:hinit} at time $0$).
-
-\input{movies/snippets/BigBattle/steps}
-
-The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{battle}.
-
-\input{movies/snippets/BigBattle/stepsBattle}
-
-\vspace{4pt}
-
-Thus, any result about \texttt{steps} will be applicable to standard battles.
-Using the predicate \texttt{steps},  our study of the length of the considered battle
-can  be decomposed into three parts:
-
-\begin{enumerate}
-\item  Characterization and proofs of regularities of some events (inspired by our experiments of Sect.~\ref{sect:testing}).
-\item Study of the beginning of the battle
-\item Computing the exact length of the battle.
-\end{enumerate}
-
-First, we prove that, if at round $i$ the hydra is equal to
-(\texttt{hyd a (S b) 0}), then it will be equal to (\texttt{hyd a b 0}) at the $2(i+1)$-th round.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/LS}
-
-From now on, the lemma \texttt{reachable\_S} allows us to watch larger and larger steps of 
-the battle.
-
-\input{movies/snippets/BigBattle/L4}
-
-\input{movies/snippets/BigBattle/L10To95}
-
-\subsection{Giant steps}
-
-We are now able to make bigger steps in the simulation of the battle.
-First, we iterate the lemma \texttt{reachable\_S}.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/Bigstep}
-
-\vspace{4pt}
-
-Applying lemmas \texttt{BigStep} and \texttt{L95} we make a first jump.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/MDef}
-
-Figure~\ref{fig:HM}  represents the hydra at the $M$-th round.
-At the $(M+1)$-th round, it will look like in fig~\ref{fig:HM-plus1}.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:HM}}
-The state of the hydra after $M$ rounds.
-% The hydra \texttt{h} of the proof that \(\omega^2\) is too small for proving Hercules' victory
-
-\end{figure}
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (0,5) {$\bullet$};
-\node (N12) at (0,8) {$\Smiley[2][green]$};
-\node (N2) at (2,5) {$\bullet$};
-\node (N22) at (2,8) {$\Smiley[2][green]$};
-\node (N3) at (4,5) {$\bullet$};
-\node (N32) at (4,8) {$\Smiley[2][green]$};
-\node (N4) at (6,5) {$\bullet$};
-\node (N42) at (6,8) {$\Smiley[2][green]$};
-
-\node (Ndots) at (12,8) {\Huge $\dots$};
-\node (Ndots2) at (12,5) {\Huge $\dots$};
-
-\node (N8) at (18,5) {$\bullet$};
-\node (N82) at (18,8) {$\Smiley[2][green]$};
-\node (N9) at (20,5) {$\bullet$};
-\node (N92) at (20,8) {$\Smiley[2][green]$};
-
-\draw (foot) -- (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N8);
-\draw (foot) -- (N9);
-\draw (N1) to  (N12);
-\draw (N2) to  (N22);
-\draw (N3) to  (N32);
-\draw (N4) to  (N42);
-\draw (N8) to  (N82);
-\draw (N9) to  (N92);
-\end{tikzpicture}
-\caption{\label{fig:HM-plus1}}
-The state of the hydra after $M+1$ rounds (with $M+1$ heads).
-
-\end{figure}
-
-\input{movies/snippets/BigBattle/L295S}
-
-\vspace{4pt}
-
-Then, applying once more the lemma \texttt{BigStep}, we get the exact time when
-Hercules wins!
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/NDef}
-
-\vspace{4pt}
-
-We are now able to prove formally that the considered battle is 
-composed of $N$ steps.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/Done}
-
-\subsection{A minoration lemma}
-
-Now, we would like to get an intuition of  how big the number $N$ is.
-For that purpose, we use a minoration of the function \texttt{doubleS} by the
-function (\texttt{fun n => 2 * n}).
-
-\vspace{4pt}
-
-\input{movies/snippets/Exp2/exp2Def}
-
-\vspace{4pt}
-
-Using a few facts (proven in 
-\href{../theories/html/hydras.Hydra.BigBattle.html}{hydras.Hydra.BigBattle}),we get several  minorations.
-
-\input{movies/snippets/BigBattle/minorationLemmas}
-
-\vspace{4pt}
-
-The number $N$ is greater than or  equal to $2^{2^{95}\times 95}.$ If we wrote $N$ in base $10$, $N$ would require at least $10^{30}$ digits!
-
-\section{Generic properties}
-
-The example we just studied shows that the termination of any battle may take a very long time. If we want to study hydra battles in general, we have to consider 
-any hydra and any strategy, both for Hercules and the hydra itself. So, we first  give some definitions, generally borrowed from transition systems vocabulary (see~\cite{tel_2000} for instance).
-
-\subsection{Liveliness}
-
-Let $B$ be an instance of \texttt{Battle}. We say that $B$ is \emph{alive} if
-for any configuration $(i,h)$, where $h$ is not a head, there exists a further step in class $B$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/AliveDef}
-
-The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
-\href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
-
-\input{movies/snippets/Hydra_Theorems/AliveThms}
-
-Both theorems are proved with the help of the  following strongly specified function:
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
-
-\input{movies/snippets/Hydra_Lemmas/nextRoundDec}
-
-\subsection{Termination}
-
-The termination of \emph{any}  battle is naturally expressed by the predicate \texttt{well\_founded} defined in the module \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}
- of the Standard Library.
-
-\index{hydras}{Library Hydra!Predicates!Termination}
-
-\input{movies/snippets/Hydra_Definitions/TerminationDef}
-
-Let $B$ be an instance of class \texttt{Battle}. A \emph{variant} for $B$ consists
-in a well-founded relation $<$  on some type \texttt{A}, and a function
-(also called a \emph{measure}) $m:\texttt{Hydra}\rightarrow A$ such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $B$, the inequality $m(h')<m(h)$ holds.
-
-\vspace{4pt}
-\noindent
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
-
-\label{sect:hvariant-def}
-
-\index{hydras}{Library Hydra!Type classes!Hvariant}
-
-\input{movies/snippets/Hydra_Definitions/HvariantDef}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
- Prove that, if there exists some  instance of (\texttt{Hvariant Lt wf\_Lt $B$ $m$}), then there exists no infinite battle in  $B$.
-\end{exercise}
-
-\subsection{A  small proof of impossibility}
-%\index{coq}{Proofs of impossibility}
-
-\label{omega-case}
-
-When one wants to prove a termination theorem with the help of a variant, 
-one has to consider first a well-founded set $(A,<)$, then a strictly decreasing measure on this set.
-The following two lemmas show that if  the order structure $(A,<)$ is too simple, it is useless to look for a convenient measure, which simply no exists.
-Such kind of result is useful, because it saves you time and effort.
-
-The best known well-founded order is the natural order on the set $\mathbb{N}$ of natural numbers (the type \texttt{nat} of Standard library).
-It would be interesting to look for some measure $m:\texttt{nat}\arrow\texttt{nat}$ and prove it is a variant.
-
-Unfortunately, we can prove that 
-\emph{no} instance of class (\texttt{WfVariant round Peano.lt $m$}) can be built, where
-$m$ is \emph{any} function of type \texttt{Hydra $\arrow$ nat}.
-
-Let us present the main steps of that proof, the script of which  is in the module ~\href{../theories/html/hydras.Hydra.Omega_Small.html}{Hydra/Omega\_Small.v} \footnote{ The name of this file means ``the ordinal $\omega$ is too small for proving the termination of [free] hydra battles ''.
-In effect, the elements of $\omega$, considered as a set, are just the natural numbers (see next chapter for more details)}.
-
-%\subsubsection{Preliminaries}
-
-Let us assume there exists some variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
-    the  termination of all hydra battles.
-
-\input{movies/snippets/Omega_Small/omegaSmalla}
-
-We define an injection $\iota$ from the type \texttt{nat} into \texttt{Hydra}.
-For any natural number $i$, $\iota(i)$ is the hydra composed of a foot and
-$i+1$ heads at height $1$. For instance, Fig.~\ref{fig:flower} represents the hydra $\iota(3)$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (4,0) {$\bullet$};
-\node (N1) at (2,2) {$\Smiley[2][green]$};
-\node (N2) at (4,2) {$\Smiley[2][green]$};
-\node (N3) at (6,2) {$\Smiley[2][green]$};
-\node (N4) at (8,2) {$\Smiley[2][green]$};
-\draw (foot) to [bend left =25] (N1);
-\draw (foot) to [bend left =15] (N2);
-\draw (foot) to [bend right =15] (N3);
-\draw (foot) to [bend right =25] (N4);
-\end{tikzpicture}
-\caption{\label{fig:flower}
-The hydra $\iota(3)$}
-\end{figure}
-
-\input{movies/snippets/Omega_Small/iotaDef}
-
-Let us consider now some hydra \texttt{big\_h} out of the range of the injection $\iota$ (see Fig.~\vref{fig:h-omega-omega}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to  (N2);
-\end{tikzpicture}
-\caption{\label{fig:h-omega-omega}}
- The hydra \texttt{big\_h}.
-\end{figure}
-
-\input{movies/snippets/Omega_Small/bigHDef}
-
- Using the functions $m$ and $\iota$, we define a second hydra \texttt{small\_h}, and show
- there is a one-round battle that transforms \texttt{big\_h} into \texttt{small\_h}. Please note that,
-due to the hypothesis \texttt{Hvar}, we are interested in the termination of \emph{free} battles. 
-There is no problem to consider a round with (\texttt{m big\_h}) as the replication factor.
-
-\input{movies/snippets/Omega_Small/smallHDef}
-
-But, by hypothesis, $m$ is a variant. Hence, we infer the following inequality.
-
-\vspace{4pt}
-
-\input{movies/snippets/Omega_Small/mLt}
-
-In order to get a contradiction, it suffices to  prove the inequality
-$m(\texttt{big\_h}) \leq m(\texttt{small\_h})$ i.e.,  $m(\texttt{big\_h}\leq m(\iota (m(\texttt{big\_h})))$.
-
-\input{movies/snippets/Omega_Small/mGea}
-
-Intuitively, it means that, from any hydra of the form (\texttt{iota $i$}), the battle will 
-take (at least) $i$ rounds. Thus the associated measure cannot be less than $i$.
-Technically, we prove this lemma by Peano induction on $i$.
-
-\begin{itemize}
-\item The base case $i=0$ is trivial
-\item Otherwise, let $i$ be any natural number and assume  the inequality
-  $i \leq m(\iota(i))$.
-  \begin{enumerate}
-  \item  But the hydra $\iota(S(i))$ can be transformed in one round into
-    $\iota(i)$ (by losing its rightmost head, for instance)
-  \item Since $m$ is a variant, we have $m(\iota(i)) < m(\iota(S(i)))$,
-    hence  $i< m(\iota(S(i)))$, which implies  $S(i)\leq  m(\iota(S(i)))$.
-  \end{enumerate}
-\end{itemize}
-
-We are now ready to complete our impossibility proof.
-
-\vspace{4pt}
-
-\inputsnippets{Omega_Small/mGeb, Omega_Small/mGez,
-  Omega_Small/omegaSmallz}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\subsubsection{Conclusion}
-
-In order to build a variant for proving the termination of all hydra battles, we need to consider order structures more complex than the usual order on type \texttt{nat}. 
-The notion of \emph{ordinal number} provides a catalogue of well-founded order types.
-For a reasonably large bunch of ordinal numbers, \emph{ordinal notations} are data-types which allow the \coq{} user to define functions, to compute and prove some properties, for instance by reflection.
-
-The next chapter is dedicated to a generic formalization of ordinal notations, and chapter~\ref{chap:T1} to a proof of termination of all hydra battles with the help of an ordinal notation for the interval $[0,\epsilon_0)$\,\footnote{We use the mathematical notation $[a,b)$ for the interval $\{x|a\leq x < b\}$.}.
-\index{maths}{Notations!Interval}
-
-%--------------------------------------------------------------
-
-\chapter{Introduction to ordinal numbers and ordinal notations}
-\label{chap:ON}
-
-The proof of termination of all hydra battles presented in~\cite{KP82} is based
-on \emph{ordinal numbers}.
-From a mathematical point of view, an ordinal is a representative of an equivalence class for isomorphisms of  totally ordered well-founded sets.
-
-For the computer scientist, ordinals are tools for proving the totality of a given recursive function, or termination of a transition system. \emph{Ordinal arithmetic} 
-provides a set of functions whose properties, like \emph{monotony}, allow to define \emph{variants}, \emph{i.e.} strictly decreasing measures used in proofs of termination.
-
-\vspace{4pt}
-
-Let us have a look at Figure~\ref{fig:ordinal-sequence}. It presents a few items of a  sequence of ordinal numbers, which extends the sequence of natural numbers.
-
-\begin{figure}[h]
-  \centering
-\fbox{\Large
-  \begin{minipage}{1.0\linewidth}
-  \begin{align*}
-     &\textcolor{blue}{0},\,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,\ldots\\
-&\textcolor{red}{\omega},\,\omega+1,\omega+2,\omega+3,\ldots\\
-&\textcolor{red}{\omega\times 2},\,\omega\times 2+1,\ldots, \textcolor{red}{\omega\times 3},\,\omega\times 3+1,\ldots, \textcolor{red}{\omega\times 4},\ldots\\
-&\textcolor{red}{\omega^2},\ldots, \textcolor{red}{\omega^2\times 42},\ldots,\textcolor{red}{\omega^3},\ldots, \textcolor{red}{\omega^4},\omega^4+1,\ldots\\
-&\textcolor{red}{\omega^\omega},\ldots, \textcolor{red}{\omega^\omega+\omega^7\times 8},\ldots,\textcolor{red}{\omega^\omega\times 2},\omega^\omega\times 2+1, \ldots\\
-&\textcolor{red}{\omega^{\omega^\omega}},\ldots, \textcolor{red}{\omega^{\omega^\omega}+\omega^\omega\times 42+ \omega^{55}+\omega}, \ldots, \textcolor{red}{\omega^{\omega^{\omega+1}}}, \omega^{\omega^{\omega+1}}+1,\dots\\
-& \textcolor{red}{\epsilon_0 (= \omega^{\epsilon_0)}}, \epsilon_0+1, \epsilon_0+2, \epsilon_0+3, \ldots\\
-& \textcolor{red}{\epsilon_1}, \ldots, \textcolor{red}{\epsilon_2}, \ldots, \textcolor{red}{\epsilon_\omega},\ldots \\
-& \textcolor{red}{\Gamma_0}, \Gamma_0+1, \Gamma_0+2, \Gamma_0+3,\ldots, \textcolor{red}{\Gamma_0+\omega}, \ldots\\
-&\ldots
-  \end{align*}   
-  \end{minipage}}
-
-  \caption{A short overview of the sequence of ordinal numbers}
-  \label{fig:ordinal-sequence}
-\end{figure}
-
-Let us comment some features of this figure:
-
-\begin{itemize}
-\item The ordinals are listed in a strictly increasing order. 
-\item Dots : ``$\ldots$'' stand for  infinite sequences of ordinals, not shown for lack of space.
-For instance, the ordinal $42$ is not shown in the first line, but it exists, somewhere between $17$ and $\omega$.
-\item Each ordinal printed in black is the immediate successor of another ordinal.
-We call it a \emph{successor} ordinal.
-For instance, $12$ is the successor of $11$, and $\omega^4+1$ the successor of $\omega^4$.
-\item Ordinals (displayed in red)  that  follow immediately dots are called \emph{limit ordinals}.
-With respect to the order induced by this sequence, any limit ordinal $\alpha$ is the least upper bound of  the set $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
-\item
-For instance $\omega$ is the least upper bound of the set of all finite ordinals (in the first line). It is also the first limit ordinal, and the first infinite ordinal, in the sense that 
-the set $\mathbb{O}_\omega$ is infinite.
-\item The ordinal $\epsilon_0$ is the first number which is equal to its own exponential of base $\omega$.
-It plays an important role in proof theory, and is particularly studied in chapters~\ref{chap:T1} to \ref{chap:alpha-large}.
-\item Any ordinal is  either the ordinal \textcolor{blue}{$0$},
-a successor ordinal, or a \textcolor{red}{limit ordinal}.
-\end{itemize}
-
-\section{The mathematical point of view}
-
-\subsection{Well-ordered sets}
-Let us start with some definitions.
-A  \emph{well-ordered set} is a set provided with a binary relation $<$ which has the following properties.
-\begin{description}
-\item[irreflexivity] : $\forall x\in A, x\not< x$
-\item[transitivity] : $\forall x\,y\,z\in A, x<y \Rightarrow y<z \Rightarrow x<z$
-\item[trichotomy]: $\forall x\,y\in A, x<y \vee x = y \vee y < x$
-\item[well foundedness]: $<$ is well-founded (every element of $A$ is accessible)\footnote{In classical mathematics, we would say that there is no infinite sequence $a_1>a_2> \dots a_n> a_{n+1}\dots$ in $A$.
-In contrast, \coq's standard library contains
-an inductive definition of a predicate \texttt{Acc} which allows us to write 
-constructive proofs of accessibility (See \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}).}.
-\end{description}
-
-The best known examples of well-ordered sets are the set $\mathbb{N}$ of natural numbers (with the usual order $<$), as well as any finite segment $[0,i)=\{j\in\mathbb{N}\,|\,j<i\}$.
-The disjoint union of two copies of $\mathbb{N}$, \emph{i.e.} the set $\{0,1\}\times\mathbb{N}$ is also well-ordered,
-with respect to the order below:
-
-\begin{align*}
-(i,j) < (i,k) & \;\textit{\textbf{if} }\; j < k\\
-(0,k) < (1,l) & \;\textit{\textbf{for\,any}}\;k \;\textit{\textbf{and}} \; l
-\end{align*}
-
-\subsection{Ordinal numbers}
-
-\index{maths}{Ordinal numbers}
-
-Let $(A,<_A)$ and $(B,<_B)$ two well-ordered sets. $A$ and $B$ are said to have \emph{the same order type} if 
-there exists a strictly monotonous bijection $b$ from $A$ to $B$, \emph{i.e.} which verifies the proposition
-$\forall x\,y\in A,\, x <_A y \Rightarrow b(x) <_B  b(y)$.
-
-Having the same order type is an equivalence relation between well-ordered sets. Ordinal numbers (in short: \emph{ordinals})  are descriptions (\emph{names}) of the equivalence classes.
-For instance, the order type of $(\mathbb{N},<)$ is associated with the ordinal called  $\omega$, and the order we considered on 
-the disjoint union of $\mathbb{N}$ and itself is named $\omega+\omega$.
-
-In a set-theoretic framework, one can consider any ordinal $\alpha$ as a well-ordered set, whose  elements are just the ordinals strictly less than $\alpha$, \emph{i.e.} the \emph{segment} $\mathbb{O}_\alpha=[0, \alpha)$.
-So, one can speak about \emph{finite}, \emph{infinite}, \emph{countable}, etc., ordinals.
-Nevertheless, since we work within type theory,
-we do not identify ordinals as sets of ordinals, but consider the correspondence between ordinals and sets of ordinals as the function that maps $\alpha$ to $\mathbb{O}_\alpha$.
-For instance $\mathbb{O}_\omega=\mathbb{N}$, and $\mathbb{O}_7=\{0,1,2,3,4,5,6\}$.
-
-We cannot cite all the literature published on ordinals since Cantor's book
-\cite{cantorbook}, and 
-leave it to the reader to explore the bibliography.
-
-The introduction of Jos\'e Grimm's report~\cite{grimm:hal-00911710} contains a nice presentation of the main properties of ordinals.
-
-\section{Ordinal numbers in Coq}
-
-Two kinds of representation of ordinals are defined in our development.
-
-\begin{itemize}
-\item A ``mathematical'' representation of the set of countable ordinal numbers, after Kurt Schütte~\cite{schutte}.
-This representation uses several (hopefully harmless) axioms.
-We use it as a reference for proving the correctness of ordinal notations.
-
-  \index{gaiabridge}{Introduction}
-  \gaiasign
-  We are also progressively importing \gaia's~\cite{Gaia} definitions and theorems, based on Bourbaki's set theory~\cite{Grimm2016}.
-Chapter~\ref{gaia-chapter} is dedicated to the connexion between libraries \HydrasLib and \gaia.
-
-\item A family of \emph{ordinal notations} (also called \emph{[ordinal] notation systems}), \emph{i.e.} data types used to represent segments $[0,\mu)$, where $\mu$ is some countable ordinal.
-Each ordinal notation is defined inside the Calculus of Inductive Constructions (without axioms).
-Many functions are defined, allowing proofs by computation.
-Note that proofs of
-correctness of a given ordinal notation with respect to Schütte's model obviously use axioms.
-Please execute the \texttt{Print Assumptions} command in case of doubt.
-\end{itemize}
-
-% \section{Countable ordinals}
-
-% Chapter~\ref{chap:schutte} of this document presents an adaptation to \coq{} of an axiomatization in classical logic of the set of countable ordinals by K. Schütte~\cite{schutte}. 
-% That formalization is quite complex, technical and unshamedly non-constructive,  so we put its description  in the last chapter of this document.
-
-% Please note that Schütte considers the (uncountable) set $\mathbb{O}$ of all countable ordinals.
-This set is well ordered (which is one of Schütte's axioms), and associates to any ordinal $\alpha$ the segment $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
-
-% In our adaptation to \coq{}, we declare a type \texttt{Ord}, a binary relation \texttt{lt} (with infix notation \texttt{"\_<\_"}, and assume Schütte's axiom. In Chapter~\ref{chap:schutte},
-% we derive some interesting properties of countable ordinals from these axioms.
-
-It is interesting to compare proofs of a given property (for instance the associativity of addition) both in the computational framework of some ordinal notation, and in mathematical  models
-of Schütte or \gaia.
-
-\section{Ordinal Notations}
-
-Fortunately, the ordinals we need for  studying hydra battles are much simpler than Schütte's, and can be represented as quite simple data types in \gallina.
-
-Let $\alpha$ be some (countable) ordinal; 
-in \coq{} terms, we call \emph{ordinal notation for $\alpha$} a structure composed 
-of:
-\begin{itemize}
-\item A data type $A$ for representing all ordinals strictly below $\alpha$,
-\item A well founded order $<$ on $A$, 
-\item A correct function for comparing two ordinals. Note  that the reflexive closure of $<$ is thus a \emph{total order}.
-\end{itemize}
-
-Such a structure may be proved correct relatively to a bigger ordinal notation or to Schütte's or \gaia's model.
-
-\subsection{Classes for ordinal notations}
-
-From the \coq{} user's point of view, an ordinal notation is
-a structure allowing to compare two ordinals by computation, and proving by well-founded induction.
-
-\subsubsection{The \texttt{Comparison} and \texttt{Comparable} classes}
-
-We use the operational class \texttt{Comparison} of comparison functions to define the \texttt{Comparable} class, contributed by Jérémy Damour and Théo Zimmermann, which allows us to apply generic lemmas and tactics about decidable strict orders.
-The correctness of the comparison function is expressed through Stdlib's type 
-\texttt{Datatypes.CompareSpec} as specialized by \texttt{Datatypes.CompSpec}.
-
-\begin{Coqsrc}
-  Inductive CompareSpec (Peq Plt Pgt : Prop) :
-  comparison -> Prop :=
-    CompEq : Peq -> CompareSpec Peq Plt Pgt Eq
-  | CompLt : Plt -> CompareSpec Peq Plt Pgt Lt
-  | CompGt : Pgt -> CompareSpec Peq Plt Pgt Gt.
-
-Definition CompSpec {A} (eq lt : A -> A -> Prop) (x y : A) :=
- CompareSpec (eq x y) (lt x y) (lt y x).
-\end{Coqsrc}
-
-\emph{From Module~\href{../theories/html/hydras.Prelude/mparable.html\#Hvariant}{Prelude.Comparable}}
-
-\label{sect:comparable-def}
-
-\input{movies/snippets/Comparable/ComparableDef}
-
-\subsubsection{The \texttt{ON} class}
-
-The following class definition, parameterized with a type $A$, a binary relation \texttt{lt} on $A$, specifies that \texttt{lt} is a well-founded strict order, provided with a correct comparison function.
-
-\vspace{4pt}
-\noindent\emph{From
-Library~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}}
-
-\label{types:ON}
-\index{hydras}{Library OrdinalNotations!Type classes!ON}
-
-\input{movies/snippets/ON_Generic/ONDef}
-
-We give  also a few handy definitions and lemmas for any ordinal notation.
-
-\label{sect:on-lt-notation}
-\label{sect:on-le-notation}
-\label{sect:measure-ON}
-\label{sect:bigO-ON}
-
-\inputsnippets{ON_Generic/ONDefsa}
-\inputsnippets{ON_Generic/ONDefsb}
-
-\begin{remark}
-The infix notations \texttt{o<} and \texttt{o<=} are defined in order to make apparent the distinction between the various notation scopes that may co-exist in a same statement.
-So the infix \texttt{<} and \texttt{<=} are reserved to the natural numbers.
-In the mathematical formulas, however, we still use $<$ and $\leq$ for comparing ordinals.
-\end{remark}
-
-% \subsection{Ordinal notations and  termination measures}
-% \label{sect:measure-ON}
-
-% The following lemma (together with the type class mechnism) allows us to define termination measures over any ordinal notation.
-It is just an application of  the libraries \texttt{Coq.Wellfounded.Inverse\_Image}
-% and  \texttt{Coq.Wellfounded.Inclusion}.
-
-% \begin{Coqsrc}
-% Definition measure_lt {A:Type}{lt: relation A}
-%             {compare : A -> A -> comparison}
-%             {on : ON lt compare}
-%             {B : Type} (m : B -> A) : relation B :=
-%              fun x y => on_lt (m x) (m y).
-
-% Lemma wf_measure  {A:Type}(lt: relation A)
-%             {compare : A -> A -> comparison}
-%             {on : ON lt compare}
-%             {B : Type}
-%             (m : B -> A):  well_founded (measure_lt m). 
-% \end{Coqsrc}
-
-% A simple example of application is given in Sect.~\vref{sect:merge-example}.
-
-\section{Example: the ordinal \texorpdfstring{$\omega$}{omega}}
-
-The simplest example of ordinal notation is built over the type \texttt{nat} of \coq's standard library. We just have to apply already proven lemmas about Peano numbers.
-
-\vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_Omega.html}{OrdinalNotations.ON\_Omega}}
-
-\inputsnippets{ON_Omega/OmegaDefa, ON_Omega/OmegaDefb}
-
-\section{Sum of  two ordinal notations}
-
-Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective types \texttt{A} and \texttt{B}.
-
- We consider a new strict order
-on the disjoint sum of the associated types, by putting all elements of \texttt{A} before the elements of \texttt{B} (thanks to Standard Library's relation operator \texttt{le\_AsB}).
-
-\vspace{4pt}
-\noindent
-\emph{From Library~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}}.
-
-\begin{Coqanswer}
-Inductive
-le_AsB (A B : Type) (leA : A -> A -> Prop) (leB : B -> B -> Prop)
-  : A + B -> A + B -> Prop :=
-| le_aa : forall x y : A, leA x y -> le_AsB A B leA leB (inl x) (inl y)
-| le_ab : forall (x : A) (y : B), le_AsB A B leA leB (inl x) (inr y)
-| le_bb : forall x y : B, leB x y -> le_AsB A B leA leB (inr x) (inr y)
-\end{Coqanswer}
-
-%\pagebreak
-
-\vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_plus.html}{OrdinalNotations.ON\_plus}}
-
-\input{movies/snippets/ON_plus/Defs}
-
-In order to build an instance of \texttt{Comparable}, we have to define a correct comparison function.
-
-\inputsnippets{ON_plus/compareDef, ON_plus/compareCorrect}
-\inputsnippets{ON_plus/plusComp}
-
-Standard library's lemma \texttt{Wellfounded.Disjoint\_Union.wf\_disjoint\_sum} 
-is applied  to prove that our order \texttt{lt} is well-founded, allowing us to build an instance of \texttt{ON}:
-
-\inputsnippets{ON_plus/ltWf, ON_plus/OnPlus}
-
-\subsection{The ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
-
-The ordinal $\omega+\omega$ (also known as $\omega\times 2$) may be represented as the concatenation 
-of two copies of $\omega$ (Figure~\ref{fig:omega-plus-omega}).
-It is also represented by the two first lines of Figure~\ref{fig:ordinal-sequence}.
-
-\begin{figure}[h]
-   \centering
-   \begin{tikzpicture}[very thick, scale=0.5]
-\begin{scope}[color=blue]
-\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
-\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
-\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
-\node (Adots) at (6,0) {$\ldots$};
-\node(An) at (8,0)[label=below:$n$]{$\bullet$};
-\node(A2) at (10,0)[label=below:$n+1$]{$\bullet$};
-\node (Adots1) at (12,0) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red]
-\node(B0) at (14,0)[label=below:$0$,label=above:\textcolor{red}{$\omega$}]{$\bullet$};
-\node(B1) at (16,0)[label=below:$1$, label=above:$\omega+1$]{$\bullet$};
-\node(B2) at (18,0)[label=below:$2$,label=above:$\omega+2$]{$\bullet$};
-\node (Bdots) at (20,0) {$\ldots$};
-\node (Bn) at (22,0) [label=below:$p$, label=above:$\omega+p$]{$\bullet$};
-\node (Bdots2) at (24,0) {$\ldots$};
-\end{scope}
-\end{tikzpicture}
-   \caption{\textcolor{blue}{$\omega+{\color{red}\omega}$}}
-   \label{fig:omega-plus-omega}
- \end{figure}
-
-We can define this notation in \coq{} as an instance of \texttt{ON\_plus}.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/OmegaPlusOmegaDef}
-
-\vspace{4pt}
-
-We can now define abbreviations.
-For instance, the finite ordinals are represented by terms built with  the constructor \texttt{inl}, and the first infinite ordinal $\omega$ by the term \texttt{(inr 0)}.
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Omega_plus_omega/finiteOmega}
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Omega_plus_omega/ltOmega}
-
-% \label{warning:coercions}
-% \index{Coq!Coercions} 
-% \begin{remark}
-% Beware of coercions and notation scopes!
-% Let us consider the following goal:
-
-% \begin{Coqsrc}
-%  Goal (6 o< 8).
-%  auto with arith.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% 1 subgoal (ID 9)
-
-%   ============================
-%   6 o< 8
-% \end{Coqanswer}
-
-% Please keep in mind that the current notation scope interprets the infix \texttt{``<''} as the predicate \texttt{Omega\_plus\_omega.lt} and not \texttt{Nat.lt}.
-More,  the coercion mechanism converts the terms \texttt{6:nat} [resp.
-\texttt{8:nat} ]
-% into \texttt{inl 6} [resp. \texttt{inl 8}].  So, the initial goal is correctly interpreted by \coq{}, but not as an inequality between two natural numbers.
-
-% \begin{Coqsrc}
-% Set Printing All.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% 1 subgoal (ID 337)
-
-%   ============================
-%   @on_lt nat Peano.lt Nat.compare Omega (S (S (S (S (S (S O))))))
-%     (S (S (S (S (S (S (S (S O))))))))
-% \end{Coqanswer}
-
-% Anyway, the initial goal is provable, using \texttt{le\_AsB}'s first constructor.
-
-% \begin{Coqsrc}
-%   constructor; auto with arith.
-% Qed.
-% \end{Coqsrc}
-
-% \end{remark}
-%
-
-\section{Limits and successors}
-
-Let us look again at our implementation of $\omega+\omega$. We can classify its elements into three categories:
-
-\begin{itemize}
-\item The least ordinal, \texttt{(inl 0)}, also known as  \texttt{(fin 0)}.
-\item The first infinite ordinal $\omega$.
-\item The remaining ordinals, either of the form \texttt{(inl (S $i$))} or \texttt{(inr (S $i$))} (in black on Figure~\ref{fig:ordinal-sequence}), called \emph{successor ordinals}.
-\end{itemize}
-
-\subsection{Definitions}
-It would be interesting to specify at the most generic level, what is a zero, a successor or a limit ordinal. Let $<$ be a strict order on a type $A$.
-
-\begin{itemize}
-\item A \emph{least} element is a minorant (in the large sense) of the full set  on $A$,
-\item $y$ is a \emph{successor} of $x$ if $x<y$ and there is no element between $x$ and $y$. We will also say that $x$ is a \emph{predecessor} of $y$.
-\item $x$ is a \emph{limit} if $x$ is not a least element, and for any $y$ such that $yo<x$,
- there exists some $z$ such that $y<z<x$.
-\end{itemize}
-
-The following definitions are in Library \href{../theories/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
-
-\input{movies/snippets/MoreOrders/Defs}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove, that, in any ordinal notation system, every ordinal has at most one predecessor, and at most one successor.
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/predSuccUnicity.v}{exercises/ordinals/predSuccUnicity.v}.}
-
-\end{exercise}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove, that, in any ordinal notation system, if $\beta$ is a successor of $\alpha$,
-then for any $\gamma$, $\gamma<\beta$ implies 
-$\gamma\leq\alpha$.
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/lt_succ_le.v}{exercises/ordinals/lt\_succ\_le.v}.}
-\end{exercise}
-
-\subsection{Limits and successors in \texorpdfstring{$\omega+\omega$}{omega+omega}}
-
-Using the definitions above, we can prove the following lemma:
-
-\vspace{4pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/limitIff}
-
-\vspace{4pt}
-
-Regarding successors, let us define the following function and prove its correctness:
-
-\input{movies/snippets/ON_Omega_plus_omega/succDef}
-
-\input{movies/snippets/ON_Omega_plus_omega/succCorrect}
-
-\vspace{4pt}
-
-We can also check whether an ordinal is a successor by a simple computation:
-
-\input{movies/snippets/ON_Omega_plus_omega/succb}
-
-\vspace{4pt}
-
-Finally, the nature of any ordinal is decidable (inside this notation system) :
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/ZeroLimitSuccDec}
-
-\section{Product of ordinal notations}
-
-Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective  ordered types \texttt{A} and \texttt{B}.
-The product of \texttt{NA} and \texttt{NB} is considered as the concatenation of $B$ copies of $A$, ordered by the lexicographic order on $B\times A$.
-
-In \coq{}, we build an instance of class \texttt{ON} through a sequence of steps as for the sum of ordinal notations.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_mult.html}{OrdinalNotations.ON\_mult}}
-
-\inputsnippets{ON_mult/Defs}
-\inputsnippets{ON_mult/multComp,  ON_mult/ONMult}
-\inputsnippets{ON_mult/endDefs}
-
-\section{The ordinal \texorpdfstring{$\omega^2$}{omega2}}
-
-The ordinal $\omega^2$ (also called $\phi_0(2)$, see Chap.~\ref{chap:schutte}), is an instance of the multiplication presented in the preceding section.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
-
-\inputsnippets{ON_Omega2/Omega2Def, ON_Omega2/Defs}
-\inputsnippets{ON_Omega2/Ex1}
-
-\subsection{Arithmetic of \texorpdfstring{$\omega^2$}{omega^2}}
-
-\subsubsection{Successor}
-
-The successor of any ordinal is defined by a simple pattern-matching.
-
-\input{movies/snippets/ON_Omega2/succ}
-
-This function is proved to be correct w.r.t. the \texttt{Successor} predicate.
-
-\input{movies/snippets/ON_Omega2/succOK}
-
-\input{movies/snippets/ON_Omega2/succLemmas}
-
-\subsubsection{Addition}
-
-We can define on \texttt{Omega2} an addition which extends the addition on \texttt{nat}. Please note that this operation is not commutative:
-
-\inputsnippets{ON_Omega2/plusDef, ON_Omega2/plusExamples}
-
-\subsubsection{Multiplication}
-
-The restriction of ordinal multiplication to the segment $[0,\omega^2)$ is not a total function.
-For instance $\omega\times\omega= \omega^2$ is outside the set of represented values.
-Nevertheless, we can define two operations mixing natural numbers and ordinals.
-
-\input{movies/snippets/ON_Omega2/multFinDef}
-\input{movies/snippets/ON_Omega2/multFinExamples}
-
-Multiplication with a finite ordinal and addition are related through the following lemma:
-
-\inputsnippets{ON_Omega2/uniqueDecompositiona,
-  ON_Omega2/uniqueDecompositionb,
-  ON_Omega2/uniqueDecompositionz}
-
-\subsection{A proof of termination using \texorpdfstring{$\omega^2$}{omega^2}} 
-\label{sect:merge-example}
-
-Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily a total function which merges two lists (example contributed by Pascal Manoury).
-
-%\index{coq}{Commands!Function}
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
-
-\inputsnippets{ON_Omega2/Merge, ON_Omega2/Mergeb}
-
-\subsection{Yet another  proof of impossibility}
-\label{omega2-case}
-
-In Sect.~\vref{omega-case}, we proved that there exists no variant from \texttt{Hydra} to \texttt{(nat,$<$)}
-(\emph{i.e.} the ordinal $\omega$) for proving the termination of all hydra battles.
-We  prove now that  the ordinal $\omega^2$ is also insufficient for this purpose.
-
-The proof we are going to comment has exactly the same structure as in Section~\ref{omega-case}.
- Nevertheless, the proof of technical  lemmas is a little more complex, due to 
- the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
-Consider for instance that there exists an infinite number of ordinals  between
-$\omega$ and $\omega\times 2$.
-
-The detailed  proof script is in the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/Hydra/Omega2_Small.v}{theories/ordinals/Hydra/Omega2\_Small.v}.
-
-\subsubsection{Preliminaries}
-Let us assume there is a variant from \texttt{Hydra} into $\omega^2$  for proving the   termination of all hydra battles.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
-
-\input{movies/snippets/Omega2_Small/Impossibility}
-\input{movies/snippets/Omega2_Small/Impossibilitya}
-
-We  follow the same pattern as in Sect.~\ref{omega-case}.
-First, we define an injection $\iota$ from type \texttt{ON\_Omega2.t} into \texttt{Hydra}, by
- associating to  each ordinal $\omega\times i+ j = (i,j)$ the hydra with $i$ branches of length $2$ and
-$j$ branches of length $1$.
-
-%% revenir ici
-
-\vspace{4pt}
-\emph{From Module ~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
-
-\input{movies/snippets/Omega2_Small/Impossibilityc}
-
-For instance, Figure~\vref{fig:essai2} shows the hydra associated to the ordinal 
-$(3,5)$, a.k.a. $\omega\times 3 + 5$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-\node (foot) at (6,0) {$\bullet$};
-\node (N1) at (1,3) {$\bullet$};
-\node (N2) at (3,3) {$\bullet$};
-\node (N3) at (5,3) {$\bullet$};
-\node (N4) at (8,3) {$\Smiley[2][green]$};
-\node (N5) at (11,3) {$\Smiley[2][green]$};
-\node (N6) at (14,3) {$\Smiley[2][green]$};
-\node (N7) at (17,3){$\Smiley[2][green]$};
-\node (N8) at (20,3){$\Smiley[2][green]$};
-\node  (N9) at (0,5) {$\Smiley[2][green]$};
-\node (N10) at (2,5) {$\Smiley[2][green]$};
-\node (N11) at (4,5) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N5);
-\draw (foot) -- (N6);
-\draw (foot) to [bend right=10] (N7);
-\draw (foot) to [bend right=15] (N8);
-\draw (N1) to [bend left=10] (N9);
-\draw (N2) -- (N10);
-\draw (N3) -- (N11);
-\end{tikzpicture}
-\caption{\label{fig:essai2}
-The hydra $\iota(\omega\times 3+5)$}
-\end{figure}
-
-Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:h-omega2-small}}
- The hydra \texttt{big\_h}.
-\end{figure}
-
-\input{movies/snippets/Omega2_Small/Impossibilityb}
-
- In a second step, we build a ``smaller'' hydra\footnote{With respect to the measure $m$.}.
-
-\input{movies/snippets/Omega2_Small/Impossibilityd}
-
-\vspace{4pt}
-
-Like in Sect.~\ref{omega-case}, we prove the inequality $m\;\texttt{big\_h} \,\texttt{o<=}\, m\;\texttt{small\_h} \,\texttt{o<=}\, m\;\texttt{big\_h}$, which is obviously false.
-
-\subsubsection{Proof of the inequality \texttt{m small\_h o< m big\_h}}
-
-In order to prove the inequality  \texttt{m\_lt: m small\_h o< m big\_h}, it suffices to
-build a battle transforming \texttt{big\_h} into \texttt{small\_h}.
-
-First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h} as $(i,j)$.
-If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
-If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the following result.
-
-\input{movies/snippets/Omega2_Small/bigToSmall}
-
-Since $m$ is a variant, we infer the following inequality:
-
-\input{movies/snippets/Omega2_Small/mLt}
-
-\subsubsection{Proof of the inequality \texttt{m big\_h o<= m small\_h} }
-
-The proof of the inequality \texttt{m big\_h o<= m small\_h} is quite more complex than in Sect~\ref{omega-case}.
-If we consider any ordinal $\alpha=(i,j)$, where $i>0$, there exists an infinite number of
-ordinals strictly less than $\alpha$, and there exists an infinite number of battles that start from
-$\iota(\alpha)$.
-Indeed, at any configuration $\iota(k,0)$, where $k>0$, the hydra can freely choose any replication number.
-Intuitively, the measure of such a hydra must be large enough for taking into account
-all the possible battles issued from that hydra.
-Let us now give more technical details.
-
-The first steps of our proof start a well-founded induction over $\omega^2$.
-
-\input{movies/snippets/Omega2_Small/mGe}
-
-A case analysis on $i$ and $j$ allows us to
-consider three cases :
-
-\begin{itemize}
-\item $i=j=0$: the inequality is trivial.
-\item $i=1+l, j=0$ ($(i,j)$ is a limit ordinal): By the induction hypothesis \texttt{IHij},
-  $(l,k)\texttt{o<=} m(\iota(l,k))$ for any $k$. But (by the rules of the hydra game), $\iota(i,0)$ is transformed into any $\iota(l,k)$ in one round. Thus $m(\iota(l,k)) < m(\iota(i,0))$ for any $k$.
-  Therefore, $(l,k) <  m(\iota(i,0))$ for any $k$, thus
-  $(i,0)\,\texttt{o<=}\,m(\iota(i,0))$.
- \item $j= l+1$  ($(i,j)$ is a successor). We apply the induction hypothesis to the pair $(i,l)$.
- \end{itemize}
-
- Please look at the proof script for more details.
-
- \input{movies/snippets/Omega2_Small/mGeb}
-
-\subsubsection{End of the proof}
-From \texttt{m\_ge}, we get \texttt{$m$\, big\_h o<= $m$\, small\_h = $m$\,(iota (m big\_h)) }. 
-Since $<$ is a strict order (irreflexive  and transitive), this inequality is incompatible with the strict inequality  \texttt{m small\_h o< m big\_h} (lemma \texttt{m\_lt}).
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#Impossible}{Hydra.Omega2\_Small}}
-
-  \input{movies/snippets/Omega2_Small/Impossible}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into $\omega^2$ for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\begin{remark}
- We  prove in Chapter~\ref{ks-chapter} a generalization of the impossibility lemmas of
-Sect.~\ref{omega-case} and this section, with the same proof structure, but with much more 
-complex technical details.
- \end{remark}
-
-% \index{Exercises}
-% \begin{exercise}
-
-% \label{sec:orgheadline63}
-% Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.
-
-% \begin{itemize}
-% \item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
-%   lexicographic ordering
-% \item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.
-
-% For instance, the following inequality holds:
-% \[\langle 4,3,3,3,3,3,3,2,2,2 \rangle\,<\,\langle 4,4,2 \rangle\]
-% \end{itemize}
-
-% \end{exercise}
-
-\section{A notation for finite ordinals}
-
-Let $n$ be some natural number. The segment associated with $n$ is the interval 
-$[0,n)\,=\,\{0,1,\dots,n-1\}$. 
-One may represent the ordinal $n$ by a sigma type.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
-
-\label{def:Finite-ord-type}
-
-\input{movies/snippets/ON_Finite/Defs}
-
-\subsubsection{Examples}
-
-For instance, let us build two elements of the segment $[0, 7)$, \emph{i.e.} two
-inhabitants of   type (\texttt{t 7}), and prove a simple  inequality (see Fig.~\ref{fig:O7}).
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (N0) at (0,0) {$\bullet$};
-\node (i0) at (0,1) {$0$};
-\node (N1) at (2,0) {$\bullet$};
-\node (i1) at (2,1) {$1$};
-\node (N2) at (4,0) {$\bullet$};
-\node (i2) at (4,1) {$2$};
-\node (N3) at (6,0) {$\bullet$};
-\node (i3) at (6,1) {$3$};
-\node (N4) at (8,0) {$\bullet$};
-\node (i4) at (8,1) {$4$};
-\node (N5) at (10,0) {$\bullet$};
-\node (i5) at (10,1) {$5$};
-\node (N6) at (12,0) {$\bullet$};
-\node (i6) at (12,1) {$6$};
-\node(alpha1) at (4,-1) {$\alpha_1$};
-\node(alpha2) at (10,-1) {$\beta_1$};
-\end{tikzpicture}
-
-\caption{The segment $\mathbb{O}_7$\label{fig:O7}}
-\end{figure}
-
-%\index{coq}{Commands!Program}
-
-\input{movies/snippets/ON_Finite/Example1}
-
-Note that the type (\texttt{t 0}) is empty, and that, for any natural number
- $n$, $n$ does not belong to (\texttt{t $n$}).
-
- \inputsnippets{ON_Finite/t0Empty, ON_Finite/bad10}
-
-\subsubsection{Comparison function}
-
-In order to build an instance of \texttt{ON}, we define a comparison function,  and prove its correctiness.
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Finite/compareDef}
-
-\begin{remark}
- The proof of \texttt{compare\_correct} uses a well-known pattern of \coq{}.
-Let us consider  the following subgoal.
-
-\input{movies/snippets/ON_Finite/compareCorrectb}
-
-Applying the tactic \texttt{f\_equal} generates a simpler subgoal.
-
-\input{movies/snippets/ON_Finite/compareCorrectc}
-
-We have now to prove that there exists at most one  proof of (\texttt{Nat.ltb x0 (S n)}). This is not obvious, but  a consequence of the following lemma of library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Eqdep_dec.html}{Coq.Logic.Eqdep\_dec}.
-
-\index{Coq}{Unicity of equality proofs}
-\label{sect:eq-proof-unicity}
-
-\begin{Coqanswer}
-eq_proofs_unicity_on :
-forall (A : Type) (x : A),
-(forall y : A, x = y \/ x <> y) -> 
-forall (y : A) (p1 p2 : x = y), p1 = p2
-\end{Coqanswer}
-
-Thus unicity of proofs of \texttt{Nat.ltb x0 (S n)}  comes from the decidability of
-equality on type \texttt{bool}.
-This is why we used the boolean function \texttt{Nat.ltb} instead of the inductive predicate \texttt{Nat.lt} in the definition of type \texttt{t $n$} (see page~\pageref{def: Finite-ord-type}).
-For more information about this pattern, please look at the numerous mailing lists and 
-FAQs on \coq{}).
-
-\inputsnippets{ON_Finite/compareCorrectd}
-
-\end{remark}
-
-\begin{remark}
-
-Please note  that attempting to compare a term  of type (\texttt{t $n$}) with a term of
-type (\texttt{t $p$})  leads to an error if $n$ and $p$ are not convertible.
-
-\input{movies/snippets/ON_Finite/Example2}
-
-\end{remark}
-
-\subsubsection{Building an instance of \texttt{ON}}
-
-Applying lemmas of the libraries \texttt{Coq.Wellfounded.Inverse\_Image}, \linebreak
- \texttt{Coq.Wellfounded.Inclusion}, and \texttt{Coq.Arith.Wf\_nat}, we prove that our
-relation \texttt{lt} is well founded.
-
-\input{movies/snippets/ON_Finite/ltWf}
-
-Now we can build our instance of \texttt{OrdinalNotation}.
-
-\input{movies/snippets/ON_Finite/ONInstance}
-
-\begin{remark}
-It is important to keep in mind  that the integer $n$ is not an ``element'' of \texttt{FinOrd $n$}.
-In set-theoretic presentations of ordinals, the set associated with the ordinal $n$ is $\{0,1,\dots,n-1\}$.
-In our formalization, the interpretation of an ordinal as a set is realized by the function \texttt{bigO} (see Section\vref{sect:bigO-ON}).
-\end{remark}
-
-\begin{remark}
- There is no interesting arithmetic on finite ordinals, since functions like successor, addition, etc.,  cannot be represented in \coq{} as \emph{total} functions.
-\end{remark}
-
-\section{See also \dots}
-
-Finite ordinals are also formalized in \ssreflect/\mathcomp~\cite{MCB}.
-In Module~\href{../theories/html/gaia_hydras.ON_gfinite.html}{gaia\_hydras.ON\_gfinite}, we build an instance of class \texttt{ON}.
-
-\inputsnippets{ON_gfinite/finordLtDef, ON_gfinite/finordON}
-
-\inputsnippets{ON_gfinite/Examples, ON_gfinite/Examplesb}
-
-  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.
-
-%%%
-
-\section{Comparing two ordinal notations}
-
-It is sometimes useful to compare two ordinal notations with respect to expressive power
-(the segment of ordinals  they represent).
-
-The following class specifies a strict inclusion of segments.
-The notation \texttt{OA} describes a segment $[0,\alpha)$, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented in \texttt{OA}).
-We require also  that the comparison functions of the two notation systems are compatible.
-
-\begin{figure}[h]
-   \centering
-   \begin{tikzpicture}[very thick, scale=0.6]
-\begin{scope}[color=blue]
-\node (A) at (0,0) {$A$};
-\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
-\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
-\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
-\node (Adots) at (6,0) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red]
-\node (B) at (0,2) {$B$};
-\node(B0) at (2,2)[label=above:$0$]{$\bullet$};
-\node(B1) at (3,2)[label=above:$1$]{$\bullet$};
-\node(B2) at (4,2)[label=above:$2$]{$\bullet$};
-\node (Bdots) at (6,2) {$\ldots$};
-\node (b) at (8,2) [label=above:$\omega$]{$\bullet$};
-\node (bsucc) at (9,2) [label=above:$\omega+1$]{$\bullet$};
-\node (Bdots2) at (10,2) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red!50!blue]
-\draw [->,thin] (A0) -- node [auto] {$\iota$} (B0);
-\draw [->,thin] (A1) -- node [auto] {$\iota$} (B1);
-\draw [->,thin] (A2) -- node [auto] {$\iota$} (B2);
-\draw [->,thin] (Adots) -- node [auto] {$\iota$} (Bdots);
-\end{scope}
-\end{tikzpicture}
-   \caption{\textcolor{blue}{$A$} is a sub-segment  of \textcolor{red}{$B$}}
-   \label{fig:subsegment}
- \end{figure}
-
-If \texttt{OB} is presumed to be correct, then we may consider that \texttt{OA} ``inherits'' its correctness from the bigger notation system \texttt{OB}.
-
-\label{types:SubON}
-\index{hydras}{Library OrdinalNotations!Type classes!SubON}
-
-following definition
-(in ~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{ON\_Generic}).
-
-\input{movies/snippets/ON_Generic/SubONDef}
-
-For instance, we prove that \texttt{Omega} is a sub-notation of
-\texttt{Omega\_plus\_Omega} (with $\omega$ as the first ``new'' ordinal, and \texttt{fin} as the injection).
-
-\vspace{10pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/Incl}
-
-We can also show that, if $i<j$, then the segment $[0,i)$ is a ``sub-segment'' of
-$[0,j)$. Since the terms  ($t\;i$) and ($t\;j$) are not convertible, we consider a ``cast'' 
-function $\iota$ from ($t\;i$) into ($t\;j$), and prove that this function is  a monotonous bijection  from ($t\;i$) to
-the segment $[0,i)$ of ($t\;j$).
-
-%\index{coq}{Commands!Program}
-
-\vspace{10pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
-
-\inputsnippets{ON_Finite/InclIJ, ON_Finite/InclIJa,
-  ON_Finite/InclIJb,ON_Finite/InclIJc, ON_Finite/InclIJd,
-ON_Finite/InclIJz}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove that \texttt{Omega\_plus\_Omega} cannot be a sub-notation of \texttt{Omega}.
-\end{exercise}
-
-\index{hydras}{Projects}
-\begin{project}
-Adapt the definition of \texttt{Hvariant} (Sect.~\ref{sect:hvariant-def}) in order to
-have an ordinal notation as argument. Prove that if $O_A$ is a sub-notation of $O_B$, then any variant defined on  $O_A$ can be automatically transformed into 
-a variant on $O_B$.
-\end{project}
-
-\section{Comparing an ordinal notation with Schütte's model}
-
-Finally, it may be interesting to compare an ordinal notation with the more theoretical model from Schütte (well, at least with our formalization of that model).
-This would be a relative proof of correctness of the considered  ordinal  notation.
-
-The following class specifies that a notation \texttt{OA} describes a segment $[0,\alpha)$,
-where $\alpha$ is a countable ordinal \emph{à la}  Schütte.
-
-\label{types:ON-for}
-\index{hydras}{Library OrdinalNotations!Type classes!ON\_correct}
-
-\input{movies/snippets/ON_Generic/ONCorrect}
-
-For instance, the following theorem tells that \texttt{Epsilon0}, our notation system for the segment $[0,\epsilon0)$ is a correct implementation of the theoretically defined  ordinal $\epsilon_0$
-(see chapter~\ref{chap:schutte} for more details).
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
-
-\input{movies/snippets/Correctness_E0/Epsilon0Correct}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Same work, but replace Schütte's model with \gaia's.
-\end{project}
-
-\begin{project}
-  When you have read Chapter~\ref{chap:schutte}, prove that the sum of two ordinal notations \texttt{ON\_plus} implements the addition of ordinals.
-\end{project}
-
-\section{Isomorphism of ordinal notations}
-
-In some cases we want to show that two notation systems describe the same segment (for instance $[0,3+\omega)$ and $[0,\omega)$\;).
-For this purpose, one may prove that the two notation systems are order-isomorphic.
-
-\index{hydras}{Library OrdinalNotations!Type classes!ON\_Iso}
-
-\label{types:ON-iso}
-
-\input{movies/snippets/ON_Generic/ONIso}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-\label{exo:i-plus-omega}
-Let $i$ be some natural number. Prove that the notation systems 
-\texttt{Omega} and (\texttt{ON\_plus (OrdFin $i$) Omega}) are isomorphic.
-
-{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$, that we prove also in larger notation systems, as well as in Schütte's model.}
-This exercise is partially solved for $i=3$ (in ~\href{../theories/html/hydras.OrdinalNotations.Example_3PlusOmega.html}{OrdinalNotations.Example\_3PlusOmega}).
-
-\end{exercise}
-
-\index{hydras}{Projects}
-\label{exo:ON-mult}
-\begin{project}
-% Define in \coq{} the product of two ordinal notations $N_A$ and $N_B$.
-% If $A$ [resp. $B$] is the underlying type of $N_A$ [resp. $N_B$], the
-% product \texttt{ON\_mult $N_A$ $N_B$} is implemented over the cartesian product $B\times A$ (with the lexicographic ordering).
-
-This exercise is about the non-commutativity of the multiplication of ordinals, reflected in ordinal notations.
-
-For instance, the
-elements of the product (\texttt{ON\_mult Omega (FinOrd 3)}) are ordered as follows.
-\[(0,0),(0,1),(0,2),(0,3),(0,4),\dots,{\color{red}(1,0),} (1,1),(1,2),\dots, {\color{red}(2,0)},(2,1),(2,2),\dots\]
-
-Note that the elements of  (\texttt{ON\_mult (FinOrd 3) Omega}) are differently ordered (without limit ordinals):
-\[(0,0),(1,0),(2,0),(0,1),(1,1),(2,1),(0,2),(1,2),(2,2),(0,3),\dots\]
-
-Prove formally  that (\texttt{ON\_mult (FinOrd $i$) Omega}) is isomorphic to
-\texttt{Omega}  whilst
-\texttt{Omega}  is a sub-notation of (\texttt{ON\_mult Omega (FinOrd $i$)}),
-for any strictly positive $i$.
-
-\textbf{Note:} Like Exercise~\ref{exo:i-plus-omega}, this project corresponds to the [in]equalities $i+\omega=\omega<\omega+i$, for any natural number $i$.
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-Consider two isomorphic ordinal notations \texttt{OA} and \texttt{OB}.
-Prove that, if \texttt{OA} [resp. \texttt{OB}] is a correct implementation 
-of $\alpha$ [resp. $\beta$], then $\alpha=\beta$.
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-\label{project:succ-limit-dec}
-Add to the class \texttt{ON} the requirement that for any $\alpha$ it is decidable whether $\alpha$ is $0$, a successor or a limit ordinal.
-
-\textbf{Hint:}   Beware of the instances associated with sum and product of notations!
-  You may consider additional fields 
-to make the sum and product of notations ``compositional''.
-
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-\label{project:on-setoid}
-Reconsider the  class \texttt{ON}, with an equivalence instead of Leibniz equality.
-\end{project}
-
-%%%% ICI ICI
-
-\section{Other ordinal notations}
-%%% TODO : Fix the multiplication function in branch FixOmegaOmega
-
-% \index{hydras}{Projects}
-
-% \begin{project}
-% The directory \texttt{theories/OmegaOmega} contains an ad-hoc formalization of $\omega^\omega$, contributed by Pascal Manoury.
-Every ordinal $\alpha$ is represented by a list $l$ whose elements are the coefficients of $\omega$ in  the Cantor normal form of $\alpha$ (in reverse order).
-For instance, the ordinal
-% $\omega^{8}\times 5 + \omega^{6}\times 8 + \omega^2\times 10 + \omega + 7$ is represented by the list \texttt{[5;0;8;0;0.0;10,1,7]}.
-
-%  Develop this representation and compare it with the other ordinal notations.
-
-% \end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Let $N_A$ be a notation system for ordinals strictly less than $\alpha$, 
-with the strict order $(A,<_A)$. Please build the notation system
-\texttt{ON\_Expl $N_A$}, on the type of multisets of elements of $A$
-(or, if preferred, the type of non-increasing finite sequences on $A$,
-provided with the lexicographic ordering on lists).
-
-For instance, let us take $N_A=\texttt{Omega}$, and take $\alpha=\langle 4,4,2,1,0\rangle$,
- $\beta=\langle 4,3,3,3,3,3,2\rangle$, and $\gamma=\langle 5\rangle$. Then $\beta<\alpha<\gamma$.
-
-In contrast the list $\langle5,6,3,3\rangle$ is not non-increasing (\emph{i.e.} sorted w.r.t. $\geq$), so it is not to be considered.
-
-Note that if the notation $N_A$ implements the ordinal 
-$\alpha$,  the new notation $\omega^{N_A}$ must implement the ordinal $\phi_0(\alpha)$, a.k.a. $\omega^\alpha$ (see chapter~\ref{chap:schutte})
-
-\end{project}
-
-\begin{remark}
- The set of ordinal terms in Cantor normal form (see Chap.~\ref{chap:T1}) and 
-in Veblen normal form (see 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}{Gamma0.Gamma0}) are shown to be ordinal notation systems, but there is a lot of work to be done in order to unify ad-hoc  definitions and proofs which were written before the definition of the \texttt{ON} type class.
-\end{remark}
-
 %------------------------------------------------------------------------
@@
     \label{sec:orgheadline83}
+    \label{sec:oplusdefT1}
 
@@
     %\include{impossibility-proofs}
-
-%-------------------------------------------------------------------
-
-\chapter[The Ketonen-Solovay machinery]{Accessibility inside \texorpdfstring{$\epsilon_0$}{Epsilon0}: The Ketonen-Solovay Machinery\label{ks-chapter}}
-\label{chap:ketonen}
-\index{maths}{Ordinal numbers!Ketonen-Solovay machinery}
-
-\section{Introduction}
-The reader may think that our proof of termination in the previous  chapter requires a lot of mathematical tools and may be too  complex. So, the question is ``is there  any  simpler proof'' ?
-
-In their article~\cite{KP82}, Kirby and Paris show that this result cannot be proved in Peano arithmetic.
-Their proof uses some knowledge about model theory and non-standard models of Peano arithmetic.
-In this chapter, we focus on a specific class of proofs of termination of hydra battles: construction of some variant mapping the type \texttt{Hydra} into a given initial  segment of ordinals.
-Our proof relies only on the Calculus of Inductive Constructions and is a natural complement of the results proven in the previous chapters.
-
-\begin{itemize}
-\item There is no variant mapping the type \texttt{Hydra} into the interval $[0,\omega^2)$ (section ~\vref{omega2-case}), and \emph{a fortiori} $[0,\omega)$ (section ~\vref{omega-case}).
-
-\item There exists a variant which maps the type \texttt{Hydra} into the
-interval $[0,\epsilon_0)$ (theorem \texttt{every\_battle\_terminates}, in section~\vref{thm:every-battle-terminates}).
-\end{itemize}
-
-Thus, a very natural question is the following one:
-\begin{quote}
-  `` Is there  any variant from
-\texttt{Hydra} into some interval $[0,\mu)$, where $\mu<\epsilon_0$, for proving the termination of all hydra battles ?''
-\end{quote}
-
-We prove in \coq{} the following result:
-
-\begin{quote}
-There is no variant for proving the termination of all hydra battles
-from \texttt{Hydra} into the interval $[0..\mu)$, where
-$\mu< \epsilon_0$.
-The same impossibility holds even if we consider only standard battles (with the successive replication factors $0,1,2,\dots,t,t+1,\dots$).
-\end{quote}
-
-Our proofs are  constructive and require no axioms: they are  closed terms of the CIC, and are mainly composed on function definitions and proofs of properties of these functions. 
-They  share much theoretical material with Kirby and Paris', although they do not use any knowledge about Peano arithmetic nor model  theory.  The combinatorial arguments we use and implement
-come from 
- an article by J.~Ketonen and R.~Solovay~\cite{KS81}, already  cited in the work
- by L.~Kirby and J.~Paris. % on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
- Section $2$ of this article: ''A hierarchy of probably recursive functions'', contains a systematic study of \emph{canonical sequences}, which are closely related to
-rounds of hydra battles. 
-Nevertheless, they have the same global structure as the simple proofs described in
-sections~\vref{omega-case} and \vref{omega2-case}. 
-We invite the reader to compare the three proofs step by step, lemma by lemma.
-
-\section{Canonical Sequences}
-\label{ketonen-solovay-sect}
-\index{maths}{Ordinal numbers!Canonical sequences}
-
-Canonical sequences are functions that associate an ordinal $\canonseq{\alpha}{i}$ to every ordinal $\alpha<\epsilon_0$ and positive integer $i$. They satisfy several nice properties:
-
-\index{maths}{Transfinite induction}
-\begin{itemize}
-\item If $\alpha\not=0$, then $\canonseq{\alpha}{i}<\alpha$. Thus canonical sequences can be used for proofs by transfinite induction or function definition by transfinite recursion
-\item If $\lambda$ is a limit ordinal, then $\lambda$ is the least upper bound of the set 
-$\{\canonseq{\lambda}{i}\;|\,i\in\mathbb{N}_1\}$
-
-\item If $\beta<\alpha<\epsilon_0$, then there is a ``path'' from $\alpha$ to $\beta$, \emph{i.e.} a
-sequence $\alpha_0=\alpha, \alpha_1, \dots, \alpha_n=\beta$, where for every $k<n$, there exists some $i_k$ such that $\alpha_{k+1}=\canonseq{\alpha_k}{i_k}$
-\item Canonical sequences correspond tightly to rounds of hydra battles: if $\alpha\not=0$,
-then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1})$ in one round with
-the replication factor $i$ (Lemma \href{../theories/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
-\item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) battle from $\iota(\alpha)$ to $\iota(\beta)$.
-\end{itemize}
-
-\begin{remark}
-  In~\cite{KS81}, canonical sequences are defined for any ordinal $\alpha <\epsilon_0$,
-by stating that if $\alpha$ is a successor ordinal $\beta+1$,  the sequence associated with 
-$\alpha$ is simply the constant sequence whose terms are equal to $\beta$.
-Likewise, the canonical sequence of $0$ maps any natural number to $0$.
-
-This convention allows us to make total the function that maps any ordinal $\alpha$ and natural number $i$ to the ordinal $\canonseq{\alpha}{i}$.
-
-\end{remark}
-
-First, let us recall how canonical sequences are defined in~\cite{KS81}.
-For efficiency's sake, we decided not to implement directly K.\&S's definitions, but to define in \gallina{} simply typed structurally recursive functions which share the abstract properties which are used in the mathematical proofs\footnote{With a small difference: the $0$-th term of the canonical sequence is not the same in our development as in~\cite{KS81}.}.
-
-\subsubsection{Mathematical definition of canonical sequences}
-
-In~\cite{KS81} the definition of $\canonseq{\alpha}{i}$ is based on the following remark:
-\begin{quote}
-Any non-zero ordinal $\alpha$ can be decomposed in a unique way as the product
-$\omega^\beta\times (\gamma+1)$.
-\end{quote}
-
-Thus the $\canonseq{\alpha}{i}$\,s are defined in terms of this decomposition:
-\begin{definition}[Canonical sequences: mathematical definition]
-\label{def:canonseq-math}
-
-\end{definition}
-\begin{mathframe}
-  \begin{itemize}
-\item Let $\lambda<\epsilon_0$ be a limit ordinal
-
-\begin{itemize}
-\item If $\lambda=\omega^{\alpha+1}\times (\beta+1)$, then 
-$\canonseq{\lambda}{i}= \omega^{\alpha+1}\times\beta +  \omega^\alpha \times i$
-\item If $\lambda=\omega^{\gamma}\times (\beta+1)$, where $\gamma<\lambda$ is a limit ordinal, then 
-$\canonseq{\lambda}{i}=\omega^{\gamma}\times \beta + \omega^{\canonseq{\gamma}{i}}$
-\end{itemize}
-
-\item For successor ordinals, we have $\canonseq{\alpha+1}{i}= \alpha$
-
-\item Finally, $\canonseq{0}{i}= 0$.
-\end{itemize}
-\end{mathframe}
-
-\subsubsection{Canonical sequences in \coq}
-\index{hydras}{Library Epsilon0!Functions!canon}
-\index{hydras}{Library Epsilon0!Functions!canonS}
-
-Our definition may look more complex than the mathematical one, but
-uses plain structural recursion over the type \coqsimple{T1}. Thus, tactics like
-\coqsimple{cbn}, \coqsimple{simpl}, \coqsimple{compute}, etc., are applicable.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canon}{Epsilon0.Canon}}
-
-\label{Functions:canonS}
-\label{Functions:canon}
-
-\input{movies/snippets/Canon/canonDef}
-
-\paragraph*{\gaiasign} Our library~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon} transposes the  function \texttt{canon} into \gaia's type \texttt{T1}
-(please see Sect.\~ref{sect:gcanon}).
-\index{gaiabridge}{Canonical sequences}
-\paragraph*{Remark}
-In the present state of this library, the following specializations of \texttt{canon} are still used in some proofs or lemma statements. They are planned to be deprecated.
-
-\input{movies/snippets/Canon/CanonS0}
-
-For instance \coq's computing facilities allow us to verify the equalities\linebreak 
-\mathcolor{$\canonseq{\omega^\omega}{3} = \omega^3$} and
-\mathcolor{$\canonseq{\omega^\omega*3}{42} = \omega^\omega*2 + \omega^{42}$}.
-
-\input{movies/snippets/Canon/canonExamples}
-
-% \index{hydras}{Projects}
-% \begin{project}
-% Many lemmas presented in this chapter were stated and proved before the introduction of 
-% the type class \texttt{ON} of ordinal notations, and in particular its  instance \texttt{Epsilon0}.
-% Thus definitions and lemmas refer to the type \texttt{T1} of possibly not well-formed terms.
-% This should be fixed in  a future version.
-% \end{project}
-
-\subsection{Basic properties of canonical sequences}
-
-We did not  try to prove that our definition truly implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
-abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
-~\href{../theories/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}.
-For instance, the equality $\canonseq{\alpha+1}{i}=\alpha$  can be  proved by  structural induction on $\alpha$.
-
-\input{movies/snippets/Canon/canonSucc}
-
-\subsubsection{Canonical sequences and the order $<$}
-
-\index{maths}{Transfinite induction}
-
-We prove by transfinite induction over $\alpha$ that $\canonseq{\alpha}{i}$ is an ordinal strictly less than $\alpha$ (assuming $\alpha\not=0$).
-This property allows us to use the function \texttt{canonS} and its derivatives in function definitions by transfinite recursion.
-
-\label{lemma:canon_LT}
-
-\inputsnippets{Canon/canonLT}
-
-\paragraph*{\gaiasign} This lemma is also available in
-Library ~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon}:
-\index{gaiabridge}{Canonical sequences}
-\inputsnippets{GCanon/gcanonLt}
-
-\subsubsection{Limit ordinals are truly limits}
-The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
-is the limit of the sequence $\canonseq{\lambda}{i}\;(1\le i)$.
-
-\vspace{4pt}
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}
-
-\input{movies/snippets/Canon/canonSLimitStrong}
-
-\label{lemma:canonS-limit}
-
-Note the use of \coq's \texttt{sig} type in the theorem's statement, which
-relates the boolean function \texttt{limitb} defined on the \texttt{T1} data-type with a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.
-We can also state directly that $\lambda$ is a (strict) least upper bound of the elements of its canonical sequence.
-
-\input{movies/snippets/Canon/canonSLimitLub}
-\index{gaiabridge}{Canonical sequences}
-
-\paragraph*{\gaiasign}  In \gaiaHydras, the statement use a slightly different vocabulary:
-\inputsnippets{GCanon/gcanonLimitOf}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}\label{exo:simply-typed-canonseq}
-Instead of using the \texttt{sig} type, define a simply typed function that, given two ordinals $\alpha$ and $\beta$, returns a natural number $i$ such that, if $\alpha$ is a limit ordinal and $\beta<\alpha$, then $\beta< \canonseq{\alpha}{i+1}$.
-Of course, you will have to prove the correctness of your function.
-
-\textbf{Hint:} You may add to your function a third argument usually called \texttt{fuel} for allowing you to give a structurally 
-recursive function (\emph{cf} the post of Guillaume Melquiond on Coq-club (Dec 21, 2020)
-\url{https://sympa.inria.fr/sympa/arc/coq-club/2020-12/msg00069.html}).
-The type \texttt{fuel}, an alternative 
-to \texttt{nat} is available on \href{../theories/html/hydras.Prelude.Fuel.html}{Prelude.Fuel})
-.
-%\index{coq}{Giving fuel to a long computation}
-
-\end{exercise}
-
-\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0}\,: paths}
-\index{gaiabridge}{Accessibility and paths inside $\epsilon_0$}
-\index{maths}{Ordinal numbers!Accessibility inside epsilon0}
-\label{sect:pathes-intro}
-
-Let us consider a kind of accessibility problem inside $\epsilon_0$: given two ordinals $\alpha$ and $\beta$, where $\beta<\alpha<\epsilon_0$, find a \emph{path} consisting of a finite sequence $\gamma_0=\alpha,\dots,\gamma_l=\beta$,
-where, for every $i<l$, $\gamma_i \not= 0$ \footnote{This condition allows us to ignore paths which end by a lot of useless $0$s.} and there exists some strictly positive integer $s_i$
-such that $\gamma_{i+1}=\canonseq{\gamma}{s_i}$.
-
-Let $s$ be the sequence $\langle s_0,s_1,\dots, s_{l-1} \rangle$. We describe the
-existence of such a path with the notation $\alpha\xrightarrow [s]{}\beta$.
-We say also that the considered path from $\alpha$ to $\beta$ \emph{starts at [index] $s_0$ and ends at $s_l$}.
-
-For instance, we have $\omega*2 \xrightarrow[2,2,2,4,5]{}3$, through the 
-path $\langle\omega\times 2, \omega+2,\omega+1,\omega,4,3\rangle$.
-
-\begin{remark}
-
-Note that, given $\alpha$ and $\beta$, where $\beta < \alpha$, the sequence $s$ which leads from $\alpha$ to $\beta$ is not unique.
-Indeed, if $\alpha$ is a limit ordinal, the first element of $s$ can be any integer $i$ such that $\beta<\canonseq{\alpha}{i}$, and if $\alpha$ is a successor ordinal,
-then the sequence $s$ can start with any positive integer.
-
-For instance, we have also 
-$\omega*2 \xrightarrow[3,4,5,6]{}\omega$. 
-Likewise,
-$\omega*2 \xrightarrow[1,2,1,4]{} 0$ and
-$\omega*2 \xrightarrow[3,3,3,3,3,3,3,3]{} 0$.
-\end{remark}
-
-\subsection{Formal definition}
-
-\label{path-to-definition}
-
-In \coq{}, the notion of path can be simply defined as an inductive predicate 
-parameterized by the destination $\beta$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\index{hydras}{Library Epsilon0!Predicates!path\_to}
-\label{sect:path-to-def}
-
-\input{movies/snippets/Paths/transitionDefs}
-
-\input{movies/snippets/Paths/pathDef}
-
-\begin{remark}
-In the present version of our library, we use a variant \texttt{path\_toS} of
-\texttt{path\_to}, where the proposition
-(\texttt{path\_toS $\beta$ $s$ $\alpha$}) is equivalent to
-(\texttt{path\_to $\beta$ (List.map S $s$) $\alpha$}). This variant is scheduled to be deprecated.
-\end{remark}
-
-\paragraph*{\gaiasign}
-
-The library~\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} transposes the  notion of path into \gaia's type \texttt{T1}
-(please see Sect.\~ref{sect:gpath}).
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Write a tactic for solving goals of the form (\texttt{path\_to $\beta$ $s$ $\alpha$})
-where $\alpha$, $\beta$ and $s$ are closed terms. 
-You should solve automatically the following goals:
-
-\inputsnippets{Paths/exPathTac} 
-\end{exercise}
-
-\subsection{Existence of a path}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction on $\alpha$, we prove that for any $\beta<\alpha$, 
-one can build a path from $\alpha$ to $\beta$ (in other terms, $\beta$ is accessible from $\alpha$).
-
-\input{movies/snippets/Paths/LTPathTo}
-
-\noindent 
-From the lemma \texttt{canon\_LT}~\vref{lemma:canon_LT}, we can convert any path into an inequality on ordinals (by induction on paths).
-
-\input{movies/snippets/Paths/pathToLT}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}[continuation of exercise~\vref{exo:simply-typed-canonseq}]
-Define a simply typed function for computing a path from $\alpha$ to $\beta$.
-\end{exercise}
-
-\subsection{Paths and hydra battles}
-\label{KS-o2h}
-
-In order to apply our knowledge about  ordinal numbers less than $\epsilon_0$ to the study of hydra battles, we define an injection
-from the interval $[0,\epsilon_0)$ into the type \texttt{Hydra}.
-
-\vspace{4pt}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}}
-
-\input{movies/snippets/O2H/iotaDef}
-
-For instance Fig.~\ref{fig:iota-example} shows the image by $\iota$ of the ordinal  \textcolor{black}{$\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$}
-
-  \begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.3]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (10,2) {$\bullet$};
-\node (N22) at (7,2) {$\bullet$};
-\node (N3) at (14,2) {$\bullet$};
-\node (N4) at (18,2) {$\Smiley[2][green]$};
-\node (N5) at (0,4) {$\bullet$};
-\node (N6) at (2,5) {$\Smiley[2][green]$};
-\node (N7) at (4,6) {$\Smiley[2][green]$};
-\node (N88) at (7,4) {$\bullet$};
-\node (N8) at (10,4) {$\bullet$};
-\node (N9) at (14,6) {$\Smiley[2][green]$};
-\node (N10) at (0,8) {$\Smiley[2][green]$};
-\node (N11) at (10,7) {$\Smiley[2][green]$};
-\node (N111) at (7,7) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N22);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (N1) to  (N5);
-\draw (N1) to   [bend left=10] (N6);
-\draw (N1) to   [bend right=20] (N7);
-\draw (N2) to  (N8);
-\draw (N22) to  (N88);
-\draw (N8) to  (N11);
-\draw (N88) to  (N111);
-\draw (N3) to  (N9);
-\draw (N5) to  (N10);
-\end{tikzpicture}
-\caption{The hydra $\iota(\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$) \label{fig:iota-example}}
-
-\end{figure}
-
-The following lemma (proved in ~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H.v}) maps  canonical sequences to rounds of hydra battles.
-
-\label{lemma:canonS-iota}
-
-\input{movies/snippets/O2H/canonSIota}
-
-The next step of our development extends this relationship to
-the order $<$ on $[0,\epsilon_0)$ on one side, and hydra battles on the other side.
-
-\input{movies/snippets/O2H/pathToRoundPlus}
-
-As a corollary, we are now able to transform any inequality $\beta<\alpha<\epsilon_0$ into a (free) battle.
-
-\input{movies/snippets/O2H/LTToRoundPlus}
-
-\section{A  proof of impossibility}
-\label{sect:impossibility-free}
-
-We now have  the tools for proving that  there exists no variant bounded by some $\mu<\epsilon_0$ for proving the termination   of all battles.
-The proof we are going to show is a proof by contradiction.
-It  can
- be considered as a generalization of the
-proofs described in  sections~\vref{omega-case} and \vref{omega2-case}.
-
-In the module
-\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
-This part of the development is parameterized by some class $B$ of battles, which will be instantiated later to \texttt{free} or \texttt{standard}.
-
-\vspace{4pt}
-\noindent
-From~\href{../theories/html/hydras.Hydra.~Hydra_Definitions.html}{Hydra.Hydra\_Definitions}:
-
-\input{movies/snippets/Hydra_Definitions/BoundedVariant}
-
-Let us assume there exists such a variant:
-
-\input{movies/snippets/Epsilon0_Needed_Generic/theContext}
-
-\label{remark:m-decrease}
-\begin{remark}
-  The hypothesis \texttt{m\_decrease} is not provable  in general, but is satisfied by
-the  \texttt{free} and \texttt{standard} kinds of battles. This trick allows to 
-``factorize'' our proofs  of impossibility.
-\end{remark}
-
-\index{maths}{Transfinite induction}
-
-First, we prove that $m(\iota(\alpha))$ is always greater than or equal to $\alpha$, by  transfinite induction over $\alpha$.
-
-\input{movies/snippets/Epsilon0_Needed_Generic/mGe0}
-
-\begin{itemize}
-\item If $\alpha=0$, the inequality trivially holds
-\item If $\alpha$ is the successor of  some ordinal $\beta$, the inequality $\beta \leq m(\iota(\beta))$ holds (by induction hypothesis).
-But the hydra $\iota(\alpha)$ is transformed in one round into
-$\iota(\beta)$, thus $m(\iota(\beta))<m(\iota(\alpha))$. Hence $\beta<m(\iota(\alpha))$, which implies $\alpha \leq m(\iota(\alpha))$
-\item If $\alpha$ is a limit ordinal, then $\alpha$ is the least upper bound of the set
-of all  the $\canonseq{\alpha}{i}$.  Thus, we have just to prove that $\canonseq{\alpha}{i}< m(\iota(\alpha))$ for any $i$. 
-\begin{itemize}
-\item Let $i$ be some natural number.
-By the induction hypothesis, we have $\canonseq{\alpha}{i} \leq m(\iota(\canonseq{\alpha}{i}))$.
-But the hydra $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i})$ in one round, thus $m(\iota(\canonseq{\alpha}{i})) < m(\iota(\alpha))$, by our hypothesis \texttt{m\_decrease}.
-\end{itemize}
-\end{itemize}
-
-Please note that the impossibility proofs of 
-sections~\vref{omega-case} and \vref{omega2-case} contain a similar lemma, also called \texttt{m\_ge}.
-We are now able to build a counter-example.
-
-\input{movies/snippets/Epsilon0_Needed_Generic/mGeGeneric}
-
-The (big) rest of the proof is dedicated to prove formally the converse inequality 
-\texttt{m small\_h t1< m big\_h}.
-
-\subsection{The case of free battles}
-\label{sec:free-battles-case}
-Let us now consider that $B$ is instantiated to \texttt{free} (which means that we are considering proofs of termination of \emph{all} battles).
-The following lemmas are proved in Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
-The case $B=\texttt{standard}$ is studied in section~\vref{std-case}.
-
-\input{movies/snippets/Epsilon0_Needed_Free/theContext}
-
-\begin{enumerate}
-\item The following lemma is an application of \texttt{m\_ge\_generic}, since \texttt{free}
-satisfies trivially the hypothesis \texttt{m\_decrease} (see page~\pageref{remark:m-decrease}).
-
-\inputsnippets{Epsilon0_Needed_Free/mGe}
-
-\item From the hypothesis \texttt{Hy}, we have \texttt{m big\_h t1< mu}
-\item By Lemma \texttt{LT\_to\_round\_plus}, we get a (free) battle from
-\texttt{big\_h = iota mu} to \texttt{small\_h = iota (m big\_h)}.
-
-\input{movies/snippets/Epsilon0_Needed_Free/bigToSmall}
-
-\item From the hypotheses on $m$, we infer:
-\input{movies/snippets/Epsilon0_Needed_Free/mLt}
-
-\item From lemmas \texttt{m\_ge} and \texttt{m\_lt}, and the irreflexivity of $<$, we get a contradiction. 
-\input{movies/snippets/Epsilon0_Needed_Free/ImpossibilityFree}
-
-\end{enumerate}
-
-We have now proved there exists no bounded variant for the class of free battles.
-
-\input{movies/snippets/Epsilon0_Needed_Free/CheckDemo}
-
-\paragraph*{\gaiasign} Please look at the \gaia version of this theorem in Sect.~\vref{sect:impossibility-gaia-version}.
-\index{gaiabridge}{Impossibility theorems}
-\section{The case of standard battles}
-\label{sec:standard-intro}\label{std-case}
-One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
-Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}.
-
-\input{movies/snippets/O2H/LTToRoundPlus}
-
-This lemma builds a battle out of any inequality $\beta<\alpha$. 
-It is a straightforward application of \texttt{LT\_path\_to} of
-Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}:
-
-\input{movies/snippets/Paths/LTPathTo}
-
-The sequence $s$, used to build the sequence of replication factors of the battle, depends on 
-$\alpha$ and $\beta$, so we cannot be sure that the generated battle is a genuine standard battle.
-
-The solution to this issue comes  once again from Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
-$\alpha_0=\alpha,\alpha_1,\dots,\alpha_k=\beta$ where $\alpha_{j+1}$ is equal
-to $\canonseq{\alpha_j}{i_j}$ where $i_j$ is \emph{any} natural number, 
-we consider various constraints on these sequences.
-In particular, a path is called \emph{standard} if $i_{j+1} = i_j + 1$ for every $j<k$.
-It  corresponds to a ``segment'' of some standard battles. 
-Please note that the vocabulary on paths is ours, but all the concepts come really from~\cite{KS81}.
-
-In \coq{}, standard paths can be defined as follows.
-
-\vspace{4pt}
-
-\emph{From
-Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/standardPathR}
-
-In the mathematical text and figures, we shall use the notation 
-$\alpha \xrightarrow[i,j]{}\beta$ for the proposition 
-(\texttt{standard\_path $i$ $\alpha$ $j$ $\beta$}).
-In~\cite{KS81} the notation is
-$\alpha \xrightarrow[i]{*}\beta$
-for 
-the proposition  $\exists j, i<j \wedge \alpha \xrightarrow[i,j]{} \beta$.
-
-Our goal is now  to transform any inequality $\beta<\alpha<\epsilon_0$ into a standard path $\alpha \xrightarrow[i,j]{} \beta$ for some $i$ and $j$, then into a standard battle
-from $\iota(\alpha+i)$ to $\iota(\beta)$. 
-Following~\cite{KS81}, we proceed in two stages:
-\begin{enumerate}
-\item we simulate plain (free) paths from $\alpha$ to $\beta$ with
-paths made of steps $(\gamma,\canonseq{\gamma}{n})$, \emph{with the same $n$ all along the path}
-\item we simulate any such path by a standard path.
-\end{enumerate}
-
-\subsection{Paths with a constant index}
-
-First of all, paths with a constant index 
-enjoy nice properties. They are defined as paths where all the $i_j$ are equal to the same natural number $i$, for some $i>0$.
-
-Like in~\cite{KS81}, we shall use the notation $\alpha \xrightarrow[i]{} \beta$ for denoting such a path, also called an $i$-path.
-
-\input{movies/snippets/Paths/constPathDef}
-
-% Paths with a given index can be effectively computed.
-% Given $i$, $\alpha$ and $l$, the following function returns the ordinal $\beta$ such that there exists a path 
-% $\alpha \xrightarrow [i+1] {} \beta$ of length $l$.
-
-% \begin{Coqsrc}
-% Fixpoint const_funS (i:nat)(alpha : T1)(l:nat):  T1  :=
-%   match l
-%   with
-%   | 0 => alpha
-%   | S m => const_funS i (canonS i alpha) m
-%   end.
-% \end{Coqsrc}
-
-% The following computations show  applications of \texttt{constS\_fun} to the 
-% ordinal $\omega^\omega$, with various values of $i$ and $l$.
-
-% \begin{Coqsrc}
-% Compute  (const_funS 2 (omega ^omega)  55).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = zero
-%      : T1 
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (const_funS 2 (omega ^omega) 15).
-% \end{Coqsrc}
-
-%   \begin{Coqanswer}
-%  = (omega ^ 2 * 2)%pT1
-%      : ppT1   
-%   \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (const_funS 4 (omega^omega)  100).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% = (omega ^ 4 * 4 + omega ^ 3 * 4 + omega ^ 2 + omega * 4 + 4)%pT1
-%      : ppT1
-% \end{Coqanswer}
-
-A most interesting property of $i$-paths is that we can ``upgrade'' their index, as stated by K.\&S.'s Corollary 12.
-
-\index{maths}{Transfinite induction}
-
-\vspace{4pt}
-\noindent
-\emph{From
-  Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/Cor12}
-
-We  also use a version of \texttt{Cor12} with large inequalities.
-
-\input{movies/snippets/Paths/Cor121}
-
-\subsubsection{Sketch of proof of \texttt{Cor12}}
-\index{maths}{Transfinite induction}
-
-First, let us prove the following lemma (numbered \texttt{2.4} in Ketonen and Solovay's article).
-
-\input{movies/snippets/Paths/KSThm24}
-
-We prove this lemma by transfinite induction on $\alpha$.
-Let us consider a path $\alpha \xrightarrow [i]{} \beta$ $(i>0)$. Its first step is
-the pair $(\alpha,\canonseq{\alpha}{i})$, We have $\canonseq{\alpha}{i}<\alpha$ and
-$\canonseq{\alpha}{i} \xrightarrow [i]{} \beta$. 
-Let $n$ be any natural number such that $n>i$.
-By the induction hypothesis, there exists a path $\canonseq{\alpha}{n} \xrightarrow[i]{} \beta$.
-\begin{itemize}
-\item  If $\alpha$ is a successor ordinal $\gamma+1$, then $\canonseq{\alpha}{n} =
-\canonseq{\alpha}{i}=\gamma$. Thus we have a path 
-$\alpha  \xrightarrow [n]{}  \gamma \xrightarrow [n]{} \beta$
-\item If $\alpha$ is a limit ordinal, we apply Theorem \texttt{KS\_thm\_2\_4} (see above).
-
-%   \begin{theorem}
-% Let $\lambda$ be a limit ordinal, then for any pair of indices $0<i<j$, there is a path $\canonseq{\lambda}{j} \xrightarrow[1]{} \canonseq{\lambda}{i}$.    
-%   \end{theorem}
-
-The proof of the limit case, is decomposed into a sequence
-of path constructions leading to $\alpha \xrightarrow[n]{} \beta$.
-
- \begin{enumerate}
- \item $\alpha \xrightarrow[n]{} \canonseq{\alpha}{n}$ (single step path)
- \item $\canonseq{\alpha}{n} \xrightarrow[1]{} \canonseq{\alpha}{i}$ (by \texttt{Theorem\_2\_4}),
-\item $\canonseq{\alpha}{n} \xrightarrow[n]{} \canonseq{\alpha}{i}$ (applying the induction hypothesis to the preceding path);
-\item $\canonseq{\alpha}{i} \xrightarrow[n]{} \beta$ (applying the induction hypothesis)
-\item $\alpha \xrightarrow[n]{} \beta$ (by composition of 1, 3, and 4).
-
- \end{enumerate}
-
-\end{itemize}
-
-\begin{remark}
- \texttt{Cor12} ``casts'' $i$-paths into $n$-paths for any $n>i$.
-But the obtained $n$-path can be much longer than the original $i$-path.
-The following exercise will give an idea of this increase. 
-\end{remark}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-  Prove that  the length of the $i+1$-path from
-  $\omega^\omega$ to $\omega^i$ is $1 + (i+1)^{(i+1)}$, for any $i$. Note that the $i$-path from
-  $\omega^\omega$ to $\omega^i$ is only one step long.
- \end{exercise}
-
-Why is \texttt{Cor12} so useful? 
-Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
-we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying corollary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
-where $n$ is the maximum of the indices $i$ met in the induction.
-
- Lemma 1, Section 2.6 of~\cite{KS81} is naturally expressed in terms of \coq's
-\verb@sig@ construct.
-
-\label{lemma:L-2_6-1}
-%\index{coq}{Sigma types}
-
-\input{movies/snippets/Paths/Lemma261}
-
-Intuitively, Lemma~\texttt{2\_6\_1}  shows that if $\beta<\alpha<\epsilon_0$, then there exists  a battle from $\iota(\alpha)$ to $\iota(\beta)$ where the replication factor is constant, although large enough.
-
- \paragraph*{\gaiasign}
- Corollary \texttt{Cor12} and Lemma \texttt{Lemma2\_6\_1} are also available in
-\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} 
-(please see Sect.~\vref{sect:gpath}).
-
-\subsection{Casting paths with a constant index into a standard path}
-
-The article~\cite{KS81} contains 
-the following lemma, which allows us to simulate $i$-paths by $[i+1,j]$-paths, where $j$ is large enough.
-
-\input{movies/snippets/Paths/constantToStandard}
-
-\subsubsection{Sketch of proof of \texttt{constant\_to\_standard\_path}}
-
-Our proof follows the proof by Ketonen and Solovay, including its organization as a sequence of lemma.
-Since it is a non-trivial proof, we will comment its main steps below   (see Figure~\vref{fig:belle-preuve-1} to Figure~\vref{fig:fin-belle-preuve}).
-
-\subsubsection*{Preliminaries}
-
-Please note that, given an ordinal $\alpha:\texttt{T1}$, and two natural numbers $i$ and $l$, there exists at most a standard path $\alpha \xrightarrow [i,i+l]{*} \beta$.
-The following function computes $\beta$ from $\alpha$, $i$ and $l$.
-
-\input{movies/snippets/Paths/standardGnaw}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction over  $\alpha$, we prove that the ordinal $0$ is reachable from any ordinal $\alpha<\epsilon_0$ by some standard path.
-
-\input{movies/snippets/Paths/standardPathToZero}
-
-\paragraph*{}
-Now, let us consider two ordinals  $\beta<\alpha<\epsilon_0$.  Let $p$  be some $(n+1)$-path from $\alpha$ to $\beta$.
-
-\input{movies/snippets/Paths/ConstantToStandardProof}
-
-Applying \texttt{standard\_path\_to\_zero}, $0$ is reachable from $\alpha$ by some standard path.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \draw[->, very thick,blue] (alpha)-- node [below]{$n+1$} node [above] {$+$} (beta);
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-    \node (alpha3) at (20,5) {};
-  \node (alphalast) at (35,5) {};
-  \node (zero) at (45,0) {$0$};
-  \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate=40]{$n+1$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha2)-- node [below]{$n+3$}  (alpha3);
-
-  \node (dots) at (24,5) {$\dots$};
-  \draw [->, dashed, very thick,blue] (alphalast)-- node [below, rotate=-26]{$n+p+1$}  (zero);
-
-\end{tikzpicture}
-\caption{A nice proof (1)}
-  \label{fig:belle-preuve-1}
-\end{figure}
-
-\paragraph*{}
-
-Since comparison on \texttt{T1} is decidable, one can compute the last step $\gamma$ of the standard path from $(\alpha,n+1)$  such that $\beta\leq \gamma$.
-Let $l$ be the length of the path from $\alpha$ to $\gamma$.  
-This step of the proof is illustrated in figure~\vref{fig:belle-preuve-2}.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (17,5) {$\ldots$};
-    \node (alpha3) at (20,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (38,0) {$\delta$};
-    \draw [->, dashed,very thick,blue] (alpha)-- node [below,rotate=35]{$n+1$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below,rotate = -48]{\tiny $n+l$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{$n+l+1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{$n+1$} node [above] {$+$} (beta);
-   \draw[thick] (alpha)--  (gamma);
-   \draw[thick] (gamma)--  node [above] {$\geq$} (beta);
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-\end{tikzpicture}
-
-\caption{A nice proof (2)}
-  \label{fig:belle-preuve-2}
-\end{figure}
-
-\paragraph*{}
-
-\begin{itemize}
-\item If $\beta=\gamma$, it's OK! We have got a standard path
-from  
-$\alpha$ to $\beta$ with successive indices  $n+1, n+2, \dots, n+l+1$
-
-\item Otherwise,  $\beta < \gamma$.  Let us consider  $\delta=\canonseq{\gamma}{n+l+1}$.
-By applying several times lemma \texttt{Cor12},  one converts  every path of Fig~\ref{fig:belle-preuve-2} into
- a $n+l+1$-path  (see figure~\ref{fig:belle-preuve-3}).
-
-But $\gamma$ is on the $n+l+1$-path from $\alpha$ to $\beta$.
-As shown by figure~\vref{fig:fin-belle-preuve}, the ordinal $\delta$, reachable from
-$\gamma$ in one single step,  must be greater than or equal to $\beta$, which contradicts our  hypothesis $\beta < \gamma$.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-    \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (17,5) {$\ldots$};
-    \node (alpha3) at (20,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (38,0) {$\delta$};
-     \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -48]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\tiny $n+l+1$} node [above]{\color{red} $1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
-    \draw[thick] (gamma)--   node [above]{\color{red} $>$}(beta);
-   \draw[thick] (alpha)--  (gamma);
-
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-
-\end{tikzpicture}
-
-\caption{A nice proof (3)}
-  \label{fig:belle-preuve-3}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (15,5) {$\ldots$};
-    \node (alpha3) at (18,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (42,0) {$\delta$};
-    \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -36]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\small $n+l+1$} node [above]{\color{red} $1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
-   \draw[thick] (alpha)--  (gamma);
-
-    \draw[thick] (gamma)--  node [below]{\tiny $n+l+1$} node [above]{\color{red} $+$}(beta);
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-
-\end{tikzpicture}
-
-\caption{A nice proof (4)}
-  \label{fig:fin-belle-preuve}
-\end{figure}
-
-\end{itemize}
- The only possible case is  thus $\beta=\gamma$, so we have got a standard path  from $\alpha$ to $\beta$.
-
- \input{movies/snippets/Paths/constantToStandard0}
- \input{movies/snippets/Paths/constantToStandardz}
-
-Here is the full statement of the conversion from constant to standard paths.
-
-\input{movies/snippets/Paths/constantToStandardPath}
-
-Applying \texttt{Lemma2\_6\_1} and \texttt{constant\_to\_standard\_path}, we get the following corollary.
-
-\input{movies/snippets/Paths/LTToStandardPath}
-
-\subsection{Back to hydras}
-\label{sec:standard-battles-cases}
-We are now able to complete our proof that there exists no bounded variant for proving the termination of standard hydra battles. This proof can
-be consulted in the module 
-\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}{Hydra.Epsilon0\_Needed\_Std}.
-Please note that it has the same global structure as in section~\ref{sec:free-battles-case} 
-% ICI !
-Applying the  lemmas  \texttt{Lemma2\_6\_1} of the module 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
-{Epsilon0.pathS}   and 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
-{\texttt{constant\_to\_standard\_path}},
-we can convert any inequality $\beta<\alpha<\epsilon_0$ into a standard path from
-$\alpha$ to  $\beta$, then into a fragment of a standard battle from 
-$\iota(\alpha)$ to $\iota(\beta)$, hence the inequality $m(\iota(\beta))<m(\iota(\alpha))$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html\#LT_to_standard_battle}{Hydra.Epsilon0\_Needed\_Std}}
-
-\input{movies/snippets/Epsilon0_Needed_Std/LTToStandardBattle}
-
-Now, please consider the following context:
-
-\input{movies/snippets/Epsilon0_Needed_Std/theContext}
-
-In the same way as for free battles, we import a large inequality 
-from 
-the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Epsilon0\_Needed\_Generic}.
-
-\input{movies/snippets/Epsilon0_Needed_Std/mGe}
-
-\paragraph*{} If remains to prove the following strict inequality, in order to have a contradiction.
-
-\input{movies/snippets/Epsilon0_Needed_Std/mLt}
-
-\paragraph*{Sketch of proof:} Let us recall that $\texttt{big\_h} = \iota(\mu)$
- and $\texttt{small\_h} = \iota (m (\texttt{big\_h}))$.
-
-Since $m(\texttt{big\_h})< \mu$, there exists a standard path from $\mu$ to
-$m(\texttt{big\_h})$, hence a   standard battle from $\iota(\mu)$  to
-$\iota(m(\texttt{big\_h}))$,  i.e. from \texttt{big\_h} to \texttt{small\_h}.
-
-Since $m$ is assumed to be a variant for standard battles, we get the inequality  $m(\texttt{small\_h}) < m(\texttt{big\_h})$.
-
-\input{movies/snippets/Epsilon0_Needed_Std/endOfProof}
-
-\paragraph*{\gaiasign} Please look at the \gaia version of this theorem in Sect.~\vref{sect:impossibility-gaia-version}.
-\index{gaiabridge}{Impossibility theorems}
-
-\subsection{Remarks}
-
-We are grateful to 
- J. Ketonen and R. Solovay  for the high quality of their explanations and proof details.
-Our proof follows tightly the sequence of lemmas in their article, with a focus on 
-constructive aspects.
-Roughly speaking, our implementation \emph{builds}, out of a hypothetical
-  variant $m$, bounded by some ordinal $\mu<\epsilon_0$, a hydra \texttt{big\_h} which verifies the impossible inequality  $m(\texttt{big\_h})< m(\texttt{big\_h})$.
-
-On may ask whether the preceding results are too restrictive, since they 
-refer to a particular data type \texttt{T1}.
-In fact, our representation of ordinals strictly less than 
- $\epsilon_0$ is faithful to their mathematical definition, at least 
-Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
-(please see also the module
-\href{../theories/html/hydras.Schutte.Correctness_E0.html}{hydras.Schutte.Correctness\_E0}).
-
-Thus, we can infer that our theorems can be applied to any well order.
-
-\index{hydras}{Projects}
-\begin{project}
-Study a possible modification of the definition of a variant  (for  standard battles).
-
-\begin{itemize}
-\item The variant is assumed to be strictly decreasing \emph{on configurations 
-reachable from some initial configuration where the replication factor is equal to $0$}
-\item The variant may depend on the number of the current round.
-\end{itemize}
-
-In other words, its type should be \texttt{nat -> Hydra -> T1}, and it must 
-verify the inequality $m\, (S\,i)\, h' < m\,i\, h$ whenever the configuration 
-$(i,h)$ is reachable from some initial configuration $(0,h_0)$
-and \texttt{h} is transformed into \texttt{h'} in the considered round.
-Can we still prove the theorems of section~\ref{std-case} with this new definition?
-
-\end{project}
-
-%---------------------------------------------------------------------
-\chapter{Large sets and rapidly growing functions}\label{chap:alpha-large}
-
-\begin{remark}
-Some notations (mainly names of fast-growing functions) of our development may differ slightly from the literature.
-Although this fact does not affect our proofs, we are preparing a future version where the names $F\_alpha$, $f\_alpha$, $H\_alpha$, etc., are fully consistent with the cited articles.
-
-\end{remark}
-%\section{Introduction}
-
-In this chapter, we try to feel how long a standard battle can be.
-To be precise, for any ordinal $\alpha<\epsilon_0$ and any positive integer $k$,
-we give a minoration of the number of steps of a standard battle which
-starts with the hydra $\iota(\alpha)$ and the replication factor $k$.
-
-We express this number in terms of the Hardy hierarchy of fast-growing 
-functions~\cite{BW85, Wainer1970, KS81, Promel2013}.
- From the \coq{} user's point of view, such  functions are  very 
-attractive:  they are defined as functions  in \gallina{}, and we can apply them \emph{in theory}, but they are so complex that you will never be able to look at the result of the computation.
- Thus, our knowledge on these functions must rely on \emph{proofs}, not tests. In our development, we use often the rewriting rules generated by \coq's \texttt{Equations} plug-in.
-
-\section{Definitions}
-
-%\subsection{Definition}
-
-\begin{definition}
-Let $0<\alpha<\epsilon_0$ be any ordinal, and $s=\langle s_1, s_2, \dots, s_N\rangle$ a finite sequence of strictly positive natural numbers.
-
-We say that $s$ is \emph{$\alpha$-large} if the sequence $\langle \alpha_0=\alpha,\dots,\alpha_{i+1}=\canonseq{\alpha_i}{i+1},\dots \rangle$ leads to $0$. 
-We say also that $s$ is \emph{minimally $\alpha$-large} (in short:
-\emph{$\alpha$-mlarge}) if $s$ is $\alpha$-large 
- and every strict prefix of $s$ leads to a non-zero ordinal (\emph{cf} Sect.~\vref{sect:path-to-def}).
-
-\index{maths}{Ordinal numbers!Large sets}
-\index{maths}{Ordinal numbers!Minimal large sets}
-
-\end{definition}
-
-\begin{remark}
-  Ketonen and Solovay~\cite{KS81} consider  large finite \emph{sets} of natural numbers,  but they are mainly used as sequences. Thus, we chose to represent them explicitely as (sorted) lists. 
-\end{remark}
-
-The following function ``gnaws'' an ordinal $\alpha$, following a sequence of indices (ignoring the $0$s).
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/gnawDef}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_SetsPaths.html\#gnaw}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/largeDef}
-
-Minimal large sequences can be directly defined in terms of the
-predicate \texttt{path\_to} (\vref{sect:path-to-def}) which already prohibits paths containing non-final \texttt{zero}s.
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}
-
-\index{hydras}{Library Epsilon0!Predicates!mlarge@mlarge (minimal large sequences)}
-
-\input{movies/snippets/Large_Sets/mlargeDef}
-
-Let us consider two integers $k$ and $l$, such that $0<k<l$. In order to check whether the interval $[k,l]$ is minimally large for $\alpha$, it is enough to
-follow from $\alpha$ the path associated with the interval $[k,l)$ and verify that the last ordinal we obtain is equal to $1$.
-
-\subsection{Examples}
-
-For instance the interval $[6,70]$ leads $\omega^2$ to $\omega\times 2 + 56$. Thus this interval is not $\omega^2$-large.
-
-\noindent
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html\#mlarge}{Epsilon0.Large\_Sets\_Examples}}
-
-\input{movies/snippets/Large_Sets_Examples/gnawEx1}
-
-The interval $[6,700]$ is $\omega^2$-large, but not
-$\omega^2$-mlarge, since $[6,699]$ is also $\omega^2$-large.
-
-\input{movies/snippets/Large_Sets_Examples/gnawEx2}
-
-The following lemma relates minimal largeness with the
-function 
-\texttt{gnaw}.
-
-\input{movies/snippets/Large_Sets/mlargeIff}
-
-\vspace{4pt}
- \noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html}{Epsilon0.Large\_Sets\_Examples}}
-
-\input{movies/snippets/Large_Sets_Examples/Ex1Lemma}
-
-\section{Length of minimal large sequences}
-\label{sect:lalpha-section}
-
-Now, consider any natural number $k>0$ any ordinal $0<\alpha<\epsilon_0$.  We would like to compute
-a number $l$ such that the interval $[k,l]$ is $\alpha$-mlarge. So, 
-the standard battle starting with $\iota(\alpha)$ and the replication factor $k$ will end after $(l-k+1)$ steps.
-
-First, we notice that this  number $l$ exists, since the segment $[0,\epsilon_0)$ is well-founded and $\canonseq{\alpha}{i}<\alpha$ for any $i$ and $\alpha>0$.
-Moreover, it is unique:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/mlargeUnicity}
-
-Thus, we would like to define a function, parameterized by $\alpha$ which associates to any  strictly positive integer $k$ the number $l$ such that
-the interval $[k,l]$ is $\alpha$-mlarge. It would be fine to write in \gallina{} a definition like this:
-
-\begin{Coqbad}
-Function L_ (alpha: E0) (i:nat) :  nat := ...
-\end{Coqbad}
-
-But we do not know how to fill the dots yet \dots{}   In the next section, we will 
-use \coq{} to reason  about the \emph{specification} of \texttt{L},
-prove properties of any function which satisfies this specification.
-In Sect.~\ref{sect:L-equations}, we use the \texttt{coq-equations} plug-in
-to define a function \texttt{L\_}, and prove its correctness w.r.t. its specification.
-
-\subsection{Formal specification}
-
-Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider any  function which  maps  any strictly positive integer $k$ to the number $l$, where 
-the interval $[k,l)$ is $\alpha$-mlarge.
-
-\begin{remark}
-In~\cite{KS81} Ketonen and Solovay consider the least natural number $l$ where the interval $[k,l]$ ($l$ included) is $\alpha$-large, and call $H_\alpha$ the function which maps $k$ to $l$.
-We chose to consider intervals $[l,k)$ instead of $[l,k]$
-in order to simplify  some statements and proofs in composition lemmas associated with the ordinals of the form $\alpha\times i$ and 
-$\omega^\alpha\times i + \beta$.
-Clearly, both approaches are related through the equality
-$L_\alpha(k)=H_\alpha(k)+1$, for any non-null $\alpha$ and $k$.
-\end{remark}
-
-Our specification of the function \texttt{L} is as follows:
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/LSpecDef}
-
-\begin{todo}
- Check if the functions $L_\alpha$ are the same as
-\cite{KS81}' functions $f_\alpha$ (p. 297).
-\end{todo}
-
-Note that, for $\alpha\not=0$, the value of $f(0)$ is not specified.
-Nevertheless, the restriction of $f$ to the set of strictly positive integers is unique (up to extensionality).
-
-\input{movies/snippets/Large_Sets/LSpecUnicity}
-
-\subsection{Abstract properties}
-
-Let us now prove properties of any function $f$ (if any) which satisfies 
-\texttt{L\_spec}.
-We are looking for properties which could be used for writing \emph{equations} and prove the correctness of the function generated by the \texttt{coq-equations} plug-in.
-Moreover, they will give us some examples (for small values of $\alpha$).
-
-The properties we consider are defined in \href{../theories/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}.
-
-\label{sect:abstract-arith-prop}
-\input{movies/snippets/Iterates/funLeDef}
-
-Our exploration of the $L_\alpha$\,s  considers the usual cases of a proof by transfinite induction: zero, successors and limit ordinals.
-The lemmas we are going to prove will be applied in a big proof by induction in Sect~\vref{sect:L-correct-proof}.
-
-\index{maths}{Transfinite induction}
-
-\subsubsection{The  ordinal zero}
-\label{sect:L-spec-zero}
-The base case is directly a consequence of the specification.
-
-\input{movies/snippets/Large_Sets/LZeroInv}
-
-\subsubsection{Successor ordinals}
-\label{sect:L-spec-succ}
-Let $\beta$ be some ordinal, and assume the arithmetic function $f$ satisfies 
-the specification $(\texttt{L\_spec}\;\beta)$.  Let $k$ be any natural number.
-Any path from $\texttt{succ}\,\beta$ to $0$ starting at $k+1$ can be decomposed into a first step from $\texttt{succ}\,\beta$ to $\beta$, then a path from
-$\beta$ at $k+2$ to $0$. 
-By hypothesis the interval $[k+2, f(k+2)-1]$ is $\beta$-mlarge.
-But the interval $[k+1, f(k+2)-1]$ is the concatenation of the singleton
-$\{k+1\}$ and the interval $[k+2, f(k+2)-1]$.
-So, the function $\lambda\,k.\,f(k+1)$ satisfies the specification $\texttt{L\_spec}\,\beta$.
-
-Note that our decomposition of intervals works only if the intervals we consider are not empty.
-In order to ensure this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <<= f}, or \texttt{(fun\_le S f)}.
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/SectionSucc}
-
-\subsubsection{Limit ordinals}
-\label{sect:L-spec-lim}
-
-Let $\lambda<\epsilon_0$ be any limit ordinal. In a similar way as for successors, we decompose any path from $\lambda$  into a first step to
-$\canonseq{\lambda}{k}$, followed by a path to $0$. In the following section, we assume that there exists a correct function computing  $L_{\canonseq{\lambda}{k}}$ for any strictly positive $k$.
-
-\input{movies/snippets/Large_Sets/SectionLim}
-
-\subsection{First results}
-
-Applying the previous lemmas on successors and limit ordinals, 
-we obtain a few  correct implementations of \texttt{(L\_spec $\alpha$)} for small values of $\alpha$.
-
-\subsubsection{Finite ordinals}
-
-By iterating the functional \texttt{L\_succ}, we get a realization of
-\texttt{(L\_spec $i$)} for any natural number $i$.
-
-\input{movies/snippets/Large_Sets/LFinDef}
-\vspace{-16pt}
-\input{movies/snippets/Large_Sets/LFinOk}
-
-\subsubsection{The first limit ordinal  \texorpdfstring{$\omega$}{omega}}
-
-The lemmas \texttt{L\_fin\_ok} and \texttt{L\_lim\_ok}   allow us to get 
-by diagonalization a correct implementation for 
-\texttt{L\_spec omega}.
-
-\input{movies/snippets/Large_Sets/LOmegaDef}
-\vspace{-16pt}
-\input{movies/snippets/Large_Sets/LOmegaOk}
-
-\subsubsection{Towards  \texorpdfstring{$\omega^2$}{omega*omega}}
-
-We would like to get exact formulas for the ordinal $\omega^2$, a.k.a.
-$\phi_0(2)$. This ordinal is the limit of the sequence $\omega\times i\;(i \in \mathbb{N})$. Thus, we have to study ordinals of this form, then use 
-our lemma on limits.
-
-The following lemma establishes a path from $\omega\times ( i+1)$ to
-$\omega \times i$.
-
-\input{movies/snippets/Large_Sets/pathToOmegaMult}
-
-Let us consider a path from  $\omega\times(i+1)$ to $0$ starting at $k+1$.
-A first ``big step'' will lead to $\omega\times i$ at $2(k+1)$. If $i>0$, the
-next jump leads to $\omega\times(i-1)$ at $2(2(k+1))+1$, etc.
-
-The following lemma expresses the length of the mlarge sequences associated with any finite multiple of $\omega$.
-
-\input{movies/snippets/Large_Sets/omegaMultMlarge0}
-
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#L_omega_mult}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/LOmegaMultDef}
-
-More generally, we prove the equality $L_{\omega\times i}(k)=2^i\times(k+1)-1$.
-
-\input{movies/snippets/Large_Sets/LOmegaMultEqn}
-
-Correctness of the function \texttt{L\_omega\_mult} is asserted through the following lemma.
-
-\input{movies/snippets/Large_Sets/LOmegaMultOk}
-
-By diagonalization, we obtain a simple formula for $L_{\omega^2}$.
-
-\input{movies/snippets/Large_Sets/LOmegaSquare}
-\input{movies/snippets/Large_Sets/LOmegaSquareEqn}
-\input{movies/snippets/Large_Sets/LOmegaSquareOk}
-
-%%%% ICI
-
-\subsubsection{Going further}
-Let us consider a last example, ``computing'' $L_{\omega^3}$.
-Let us study the canonical sequence associated with $\omega^3$, the elements of which are 
-$\omega^2\times i\;(i\in\mathbb{N}_1)$.
-
-To this end, we prove a generic lemma, which expresses $L_{\omega^\alpha\times i}$ as an iterate of $L_{\omega^\alpha}$.
-Note that in this lemma, we assume that the function associated with $\alpha$ is strictly monotonous and
-greater or equal than the successor function, and prove that $L_{\omega^\alpha\times i}$ satisfies  the same properties.
-
-\inputsnippets{Large_Sets/phi0Mult}
-
-\inputsnippets{Large_Sets/phi0MultOk,
-Large_Sets/Phi0MultSLe}
-
-Let us look now
-at the ordinal $\omega^2\times i$, using \texttt{L\_phi0\_mult}.
-
-\input{movies/snippets/Large_Sets/LOmegaSquareTimes}
-
-We are now ready to get an exact formula for $L_{\omega^3}$, by diagonalization over $L_{\omega^2\times i}$.
-
-\input{movies/snippets/Large_Sets/LOmegaCube}
-
-Thus, for instance, $L_{\omega^3}(3)=L_{\omega^2\times 4}(3)$.
-
-\input{movies/snippets/Large_Sets/LOmegaCube3Eq}
-
-This number is quite big. Using \texttt{Ocaml}'s \texttt{float} arithmetic,
-we can under-approximate it by $2^{3.8\times10^{30}}\times 3.8\times{10^{30}}$.
-
-\begin{Coqsrc}
-# let exp2 x = 2.0 ** x;;
-
-val exp2 : float -> float = <fun>
-#   exp2 95.0 *. 97.0 -. 1.0;;
-- : float = 3.84256588194182037e+30
-# let n = exp2 95.0 ;;
-# let p = n *. 97.0 -. 1.0;;
-val p : float = 3.84256588194182037e+30
-
-Estimation :
-2 ** (3.84 e+30) * 3.84 e+30.
-\end{Coqsrc}
-
-\subsection{Using \texttt{Equations}}
-\label{sect:L-equations}
-
-Note that we did not define any function $L_\alpha$ \emph{for any $\alpha<\epsilon_0$} yet.
-We have got no more than a collection of proved realizations of $\texttt{L\_spec}\;\alpha$ for a few values of $\alpha$.
-
-\index{coq}{Plug-ins!Equations}
-
-Using the \texttt{coq-equations} plug-in by 
-M. Sozeau~\cite{sozeau:hal-01671777}, we will now define a function \texttt{L\_} which maps any ordinal  $\alpha<\epsilon_0$ to a proven realization of 
-$\texttt{L\_spec}\;\alpha$.   
-To this end, we represent ordinals as inhabitants of the type 
-\texttt{E0} of well-formed ordinal terms (see Sect~\vref{sect:E0-def}). So, we define a total function \texttt{L\_} of type
-\texttt{E0 -> nat -> nat}, by transfinite recursion, considering the usual three cases : $\alpha=0$, $\alpha$ is a successor, $\alpha$ is a limit ordinal.
-
-\subsubsection{Definition}
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#L_}{L\_alpha}).}
-
-\label{Functions:L-alpha}
- \index{hydras}{Library Epsilon0!Functions!L\_@L\_ (final step of a minimal path}
-
-\input{movies/snippets/L_alpha/LDef}
-
-This definition results in a bunch of automatically generated lemmas. For instance:
-
-\input{movies/snippets/L_alpha/AboutLEquation1}
-
-In most cases, it may be useful to write human-readable  paraphrases of these statements.
-
-\inputsnippets{L_alpha/Paraphrasesa}
-\inputsnippets{L_alpha/Paraphrasesb}
-\inputsnippets{L_alpha/Paraphrasesc}
-
-Using these three lemmas as rewrite rules, we can prove more properties of the functions \texttt{L\_$\alpha$}.
-
-\input{movies/snippets/L_alpha/LFiniteOmega}
-
-By  well-founded induction on $\alpha$, we prove the following properties:
-
-\input{movies/snippets/L_alpha/LGeS}
-
-\input{movies/snippets/L_alpha/LCorrect}
-
-\label{sect:L-correct-proof}
-
-Please note that the proof of \texttt{L\_correct} applies the lemmas proven in Sections~\ref{sect:L-spec-zero}, ~\ref{sect:L-spec-succ} and ~\ref{sect:L-spec-lim}.
-Our previous study of \texttt{L\_spec} allowed us to pave the way for the definition by \texttt{Equations} and the correctness proof.
-
-\paragraph*{\gaiasign}
-The module
-\href{../theories/html/gaia_hydras.GL_alpha.html}{gaia\_hydras.GL\_alpha} contains an adaptation of
-\href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0/L\_alpha}.
-
-\subsubsection{Back to hydra battles}
-\label{def:L-alpha}
-
-Lemma \texttt{battle\_length\_std } of
-Module~\href{../theories/html/hydras.Hydra.Battle_length.html}{Hydra.Battle\_length} relates the length of standard battles with the functions $L_\alpha$.
-
-\input{movies/snippets/Battle_length/battleLengthStd}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Instead of considering standard paths and battles, consider battles where the replication factor is a constant $k$.
-Please use \texttt{Equations} in order to define the function that computes the length of the $k$-path which leads  from $\alpha$ to $0$.
-Prove a few  exact formulas and minoration lemmas.
-\end{exercise}
-
-\section{A variant of the Hardy hierarchy}
-
-\label{sect:hardy}
-
-In order to give a feeling on  the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the  \emph{Hardy hierarchy} of fast growing functions,
-presented for instance in~\cite{Promel2013}. 
-\index{maths}{Rapidly growing functions!Hardy Hierarchy}
-
-\begin{remark}
-  Indeed, the functions presented in this section are a \emph{variant} of the Hardy hierarchy of functions.
-In the future versions of this development, we will correct the references to the literature.
-For the time being, we call our functions $H'_\alpha$ in order to underline the difference from ``classic'' Hardy functions.
-\end{remark}
-
-For each ordinal $\alpha$ below $\epsilon_0$, $H'_\alpha$ is 
-a total arithmetic function, defined  by  transfinite recursion on $\alpha$, according to three cases:
-
-\index{maths}{Transfinite induction}
-
-\begin{itemize}
-\item If $\alpha=0$, then $H'_\alpha (k)= k$ for any natural number $k$.
-\item If $\alpha=\textrm{succ}(\beta)$, then 
-$H'_\alpha(k)=H'_\beta(k+1)$ for any $k \in \mathbb{N}$
-\item If $\alpha$ is a limit ordinal, then 
-$H'_\alpha(k) = H'_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
-\end{itemize}
-
-\begin{remark}
- The ``classic'' definition of the Hardy hierarchy differs in the third equation.
-
-\begin{itemize}
-\item If $\alpha=0$, then $H_\alpha (k)= k$ for any natural number $k$.
-\item If $\alpha=\textrm{succ}(\beta)$, then 
-$H_\alpha(k)=H_\beta(k+1)$ for any $k \in \mathbb{N}$
-\item If $\alpha$ is a limit ordinal, then 
-$H_\alpha(k) = H_{(\canonseq{\alpha}{k})}(k)$ for any $k\in \mathbb{N}$.
-\end{itemize}
-
-\end{remark}
-
-\subsection{Definition in \texttt{Coq}}
-
-We define a function \texttt{H'\_} of type \texttt{E0 -> nat -> nat} by transfinite induction over the type \texttt{E0} of the well formed ordinals below $\epsilon_0$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}}
-
-\index{hydras}{Library Epsilon0!Functions!H\_@H\_ (Hardy hierarchy (variant))}
-\index{coq}{Plug-ins!Equations}
-\label{Functions:Hprime-alpha}
-
-\input{movies/snippets/Hprime/HprimeDef}
-\input{movies/snippets/Hprime/HprimeDefb}
-
-\input{movies/snippets/Hprime/paraphrasesa}
-\input{movies/snippets/Hprime/paraphrasesb}
-\input{movies/snippets/Hprime/paraphrasesc}
-\input{movies/snippets/Hprime/paraphrasesd}
-
-\subsection{First  steps of the H' hierarchy}
-Using rewrite rules from \texttt{H'\_eq1} to \texttt{H'\_succ\_eqn}, we can explore the functions $H'_\alpha$ for  small values of $\alpha$.
-
-\paragraph*{\gaiasign}
-The module
-\href{../theories/html/gaia_hydras.GHprime.html}{gaia\_hydras.GHprime} contains an adaptation of
-\href{../theories/html/hydras.Epsilon0.Hprime.html}{Epsilon0.Hprime}.
-
-\subsubsection{Finite ordinals}
-
-By induction on $i$, we prove a simple expression of \texttt{H'\_ (Fin i)}, where 
-\texttt{Fin $i$}  is the $i$-th finite ordinal.
-
-\input{movies/snippets/Hprime/HprimeFin}
-
-\subsubsection{Multiples of \texorpdfstring{$\omega$}{omega}}
-
-Since the canonical sequence of $\omega$ is composed of finite ordinals, 
-it is easy to get the formula associated with $H'_\omega$.
-
-\input{movies/snippets/Hprime/HprimeOmega}
-
-Before going further, we prove a useful rewriting lemma:
-
-\input{movies/snippets/Hprime/HprimePlusFin}
-
-Then, we get easily formulas for $H'_{\omega+i}$, and $H'_{\omega\times i}$ for any natural number $i$.
-
-\input{movies/snippets/Hprime/HprimeExamplesa}
-\input{movies/snippets/Hprime/HprimeExamplesb}
-\input{movies/snippets/Hprime/HprimeExamplesc}
-\input{movies/snippets/Hprime/HprimeExamplesd}
-
-Since $\omega^2$ is a limit ordinal, we prove the following equality: 
-\[H'_{\omega^2} (k) = 2 ^ {k+1} \times (k+1) - 1\]
-
-\input{movies/snippets/Hprime/HprimeOmegaSqr}
-
-\subsubsection{New limits}
-
-Our next step would be to prove an exact formula for $H'_{\omega^\omega}(k)$.
-Since the canonical sequence of $\omega^\omega$ is composed of all the
-$\omega^i$, we first need to express $H'_{\omega^i}$ for any natural number $i$.
-
-Let $i$ and $k$ be two natural numbers. 
-The ordinal $\canonseq{\omega^(i+1)}{k}$ is the product
-$\omega^i \times k$, so we need also to consider ordinals of this form.
-
-\begin{enumerate}
-\item First,  we express $H'_{\omega^\alpha \times (i+2)}$ in terms of
-$H'_{\omega^\alpha \times (i+1)}$.
-
-\input{movies/snippets/Hprime/HprimeOmegaTerm1}
-
-\item
-Then, we prove by induction on $i$ that $H'_{\omega^\alpha \times (i+1)}$ is just the
-$(i+1)$-th iterate of $H'_{\omega^\alpha}$.
-
-\input{movies/snippets/Hprime/HprimeOmegaTerm}
-
-\item In particular, we derive a formula for $H'_{\omega^{i+1}}$.
-
-\input{movies/snippets/Hprime/HprimeSuccFun}
-\input{movies/snippets/Hprime/HprimePhi0SI}
-
-\item We get now a  formula for $H'_{\omega^3}$:
-
-\input{movies/snippets/Hprime/HprimeOmegaCube}
-\end{enumerate}
-
-\subsubsection{A numerical example}
-\label{sect:bignum}
-
-It is hard to capture the complexity of this function by looking only at this
-``exact'' formula. 
-Let us consider a simple example: the number $H'_{\omega^3}(3)$.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3a}
-
-Thus, the number $H_{\omega^3}(3)$ can be written as four nested applications of $f$.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3b}
-
-In a more classical writing, this number is displayed as follows:
-
-{\Large
-$$
-H'_{\omega^3}(3) =  2 ^ {(2 ^ {N + 1} \, (N+1) )}   \,  (2 ^ {N+1} \, ( N +1) ) - 1
-$$
-}
-
-In order to make this statement more readable, we introduce a local definition.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3c}
-
-This number looks quite big; let us compute an approximation with \texttt{Ocaml}:
-
-\begin{Coqsrc}
-# (2.0 ** 64.0 *. 64.0 -. 1.0);; 
-\end{Coqsrc}
-
-\begin{Coqanswer}
-- : float = 1.1805916207174113e+21
-\end{Coqanswer}
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3d}
-
-We leave as an exercise to determine the best approximation as possible of
- the size of this number (for instance its number of digits).  For instance, if
-we do not take into account the multiplications in the formula above,
-we obtain that, in base $2$, the number $H'_{\omega^3}(3)$ has at least
-$2^{10^{21}}$  digits. But it is still an under-approximation !
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3e}
-
-\subsubsection{A formula for \texorpdfstring{$H'_{\omega^\omega}$}{\texttt{H'(phi0(omega))}}}
-
-Now, we can get at last an exact formula for $H'_{\omega^\omega}$.
-
-\input{movies/snippets/Hprime/HprimePhi0Omega}
-
-Using extensionality of the functional \texttt{iterate}, we also get a closed formula.
-
-\input{movies/snippets/Hprime/HprimePhi0OmegaClosed}
-
-Note that this formula contains two occurrences of the functional \texttt{iterate}, the outer one is in fact a second-order iteration (on type \texttt{nat -> nat)}
-and the inner one  first-order (on type \texttt{nat}).
-
-\subsection{Abstract properties of  
-\texorpdfstring{$H'_\alpha$}{H'}}
-~\label{sect:H-alpha-prop}
-
-Since pure computation seems to be useless for dealing with expressions of the form $H'_\alpha(k)$, even for small values of $\alpha$ and $k$, we need to prove theorems for comparing $H'_\alpha(k)$ and $H'_\beta(l)$, in terms of comparison
-between $\alpha$ and $\beta$ on the one hand, $k$ and $l$ on the other hand.
-
-But beware of fake theorems! For instance, one could believe that $H'$ is monotonous in its first argument. The following proof shows this is false.
-
-\input{movies/snippets/Hprime/HprimeNonMono1}
-
-On the contrary, the functions of the $H'$ hierarchy have the following five properties~\cite{KS81}: for any $\alpha < \epsilon_0$,
-\begin{itemize}
-\item the function $H'_\alpha$ is strictly monotonous :
-      For all $n,p \in\mathbb{N}, n < p \Rightarrow H'_\alpha(n)< H'_\alpha(p)$.
-\item If $\alpha \not= 0$, then for every $n$, $n<H'_\alpha(n)$.
-\item The function $H'_\alpha$ is pointwise less or equal than $H'_{\alpha+1}$
-
-\item For any $n\geq 1$, $H'_\alpha(n)<H'_{\alpha+1}(n)$.
-\emph{We say that $H'_{\alpha+1}$ dominates $H'_\alpha$ from $1$}.
-\item For any $n$ and $\beta$, if $\alpha \xrightarrow[n]{} \beta$, then
-$H'_\beta(n)\leq H'_\alpha(n)$.
-\end{itemize}
-
-\index{maths}{Abstract properties of arithmetic functions}
-\index{hydras}{Abstract properties of arithmetic functions}
-
-\index{maths}{Transfinite induction}
-
-In \coq{}, we follow the  proof of~\cite{KS81}. This proof is mainly a single  proof by transfinite induction on $\alpha$ of the conjunction of the five properties.
-For each $\alpha$, the three cases : $\alpha=0$, $\alpha$ is a limit, and 
-$\alpha$ is a successor are considered. Inside each case, the five sub-properties are proved sequentially, using the abstract properties defined in Sect~\vref{sect:abstract-arith-prop}
-
-\input{movies/snippets/Hprime/PAP}
-
-Using a few lemmas \emph{à la} Ketonen-Solovay, we prove that
-if $\alpha<\beta$, then $H'_\beta$ eventually dominates
-$H'_\alpha$.
-We let the reader look at the proof (Section \texttt{Proof\_of\_H'\_mono\_l} of \href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}).
-
-\noindent
-\input{movies/snippets/Hprime/HprimeDom}
-
-%ICI
-\subsection{Comparison between \texttt{L\_} and \texttt{H'\_} }
-
-By well-founded induction on $\alpha$, we prove that
- $H'_\alpha(i) \leq L_\alpha(i+1)$ for any $i$.
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#H'_L_}{Epsilon0.L\_alpha}}
-
-\input{movies/snippets/L_alpha/HprimeL}
-
-\subsubsection{Back to hydras}
-
-The following theorem relates the length of (standard) battles with the the $H'$ family of fast growing functions.
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems}}
-
-\input{movies/snippets/Hydra_Theorems/battleLengthStdHardy}
-
-Since $H'_{\omega^3}(3)= H'_{\omega^3+3}(0)$, the big number shown in Section~\vref{sect:bignum} is less or equal than 
-the number of steps of the battle with the initial hydra of
-Figure~\ref{fig:start}, and  $0$ as initial replication factor.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[very thick, scale=0.3]
-  \node (h1) at (4,0){$\bullet$};
-  \node (h2) at (2,2){$\bullet$};
-  \node (h3) at (0,4){$\bullet$};
-  \node (h4) at (1,4){$\bullet$};
-  \node (h5) at (2,4){$\bullet$};
-  \node (h6) at (4.5,2){$\bullet$};
-  \node (h7) at (6.5,2){$\bullet$};
-  \node (h8) at (8.5,2){$\bullet$};
-    \draw (h1) -- (h2) ;
-    \draw (h2) -- (h3) ;
-    \draw (h2) -- (h4) ;
-    \draw (h2) -- (h5) ;
-    \draw (h1) -- (h6) ;
-    \draw (h1) -- (h7) ;
-    \draw (h1) -- (h8);
- \end{tikzpicture}
-
-  \caption{The hydra corresponding to the ordinal $\omega^3+3$}
-  \label{fig:start}
-\end{figure}
-
-\section{A variant of the Wainer hierarchy (functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
-\label{sect:wainer}
-
-\index{maths}{Rapidly growing functions!Wainer Hierarchy}
-
-Ketonen and Solovay introduce in~\cite{KS81} a ``trivial'' variant of the Wainer hierarchy~\cite{BW85, Wainer1970} of fast growing functions, indexed by ordinals below $\epsilon_0$.
-The functions $F_\alpha$ are defined by the following equations.
-
-\label{F_equations}
-\begin{itemize}
-\item $F_0(i)=i+1$
-\item $F_{\beta+1}(i)= (F_\beta)^{(i+1)}(i)$, where $f^{(i)}$ is the $i$-th iterate of $f$.
-\item $F_\alpha(i) = F_{\canonseq{\alpha}{i}} (i)$ if $\alpha$ is a limit ordinal.
-\end{itemize}
-
-\begin{remark}
-The difference with the ``classic'' Wainer hierarchy 
-$f_\alpha\;(\alpha<\epsilon_0)$ lies in the second equation:
-$f_{\beta+1}(i) = (f_\beta)^{(i)}(i)$ and not
-$f_{\beta+1}(i) = (f_\beta)^{(i+1)}(i)$.
-
-A module about 
-the classic Wainer hierarchy is in preparation.
-
-Note also that \cite{KS81} defines also $F_{\epsilon_0}$ (by the third equation).
-Since $\epsilon_0$ is not representable in type \texttt{E0}, our implementation in \coq{} does not take $F_{\epsilon_0}$ into account.
-
-\end{remark}
-
-A first attempt is to write a definition of $F_\alpha$ by equations, in the same way as for $H_\alpha$ (the functional \texttt{iterate} has already been used in Sect.\vref{Functions:iterate}).
-
-\index{hydras}{Library Prelude!iterate}
-
-\input{movies/snippets/Iterates/iterateDef}
-
-\emph{From\href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}~}.
-
-\index{coq}{Plug-ins!Equations}
-
- \input{movies/snippets/F_alpha/FailDemo}
-
-We presume that this error comes from the recursive call of \texttt{F\_} inside
-an application of \texttt{iterate}. The workaround we propose is to define first 
-the iteration of \texttt{F\_}  as an helper $F^*$, then to define the function $F$ as a ``iterating $F^*$ once''.
-
-\texttt{Equations} accepts the following definition, relying on  lexicographic ordering on pairs $(\alpha,n)$.
-
-\label{sect:F-equations}
-
-\index{coq}{Plug-ins!Equations}
-\label{Functions:F-alpha}
-\index{maths}{Rapidly growing functions}
-\index{hydras}{Library Epsilon0!Functions!F\_@F\_ (Wainer hierarchy)}
-
-\input{movies/snippets/F_alpha/goodDefa}
-\input{movies/snippets/F_alpha/goodDefb}
-
-It is quite easy to prove that our functional \texttt{F\_} satisfies the equations on page~\pageref{sect:F-equations}.
-\index{hydras}{Library Prelude!iterate}
-
-\input{movies/snippets/F_alpha/FEquations}
-
-As for the Hardy functions, we can use these equalities as rewrite rules for
-``computing'' some values of $F_\alpha(i)$, for small values of $\alpha$.
-
-\input{movies/snippets/F_alpha/FirstValues}
-
-Like in Sect~\ref{sect:H-alpha-prop}, we prove by induction the following properties (see~\cite{KS81}).
-
-\input{movies/snippets/F_alpha/FalphaThms}
-
-As a corollary, we prove that, if $\beta<\alpha$, then
-$F_\alpha$ dominates $F_\beta$ (p. 284 of~\cite{KS81}).
-
-\input{movies/snippets/F_alpha/FDomContext}
-\input{movies/snippets/F_alpha/FDom}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove the following property:
-
-\begin{Coqsrc}
-Lemma LF3 : dominates_from  2 (F_ 3) (fun  n => iterate exp2 n n).
-\end{Coqsrc}
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
-\end{exercise}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that, for any $\alpha\geq 3$ and $n\geq 2$,
-$F_\alpha(1+n)\geq 2^{F_\alpha(n)}$.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
-\end{exercise}
-
-%ici
-\index{hydras}{Exercises}
-
-\begin{exercise}
-It is tempting to prove a simple property of monotony 
-of the function \texttt{F\_}.
-
-\begin{quote}
-   Let $\alpha\leq\beta<\epsilon_0$. For any $n\geq 2$,
-$F_\alpha(n)\leq F_\beta(n)$. 
-\end{quote}
-Prove or disprove this statement.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/is_F_monotonous.v}{exercises/ordinals/is\_F\_monotonous.v}.}
-\end{exercise}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-
-Prove that for any $n\geq 2$, $\textrm{Ack}\,\,n\,n\leq  F_\omega(n)$, where \textrm{Ack} is the Ackermann function.
-Next, prove that $F_\alpha$ is not primitive recursive, for any $\alpha\geq\omega$  (please see Sect.~\vref{sect:ack-not-PR}).
-On the other hand, please show that for any natural number $n$, the function $F_n$ is primitive recursive.
-Thus $F_\alpha$ is primitive recursive if and only if $\alpha$ is finite.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_omega.v}{exercises/ordinals/F\_Omega.v}.
-Properties of the Ackermann function are studied in
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/Ack.v}{theories/ordinals/MoreAck/Ack.v} and
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/AckNotPR.v}{theories/ordinals/MoreAck/AckNotPR.v}
-.}
-\end{exercise}
-
-\paragraph*{\gaiasign} Library~\href{../theories/html/gaia_hydras.GF_alpha.html}{gaia\_hydras.GF\_alpha}  defines the functions $F_\alpha$ for \gaia's world.
-\index{gaiabridge}{Rapidly growing functions}
-
-\section{More about rapidly growing functions}
-
-In Sect.~\vref{sect:battle-length-notPR}, we prove that the length of hydra-battles (for a given hydra, according to the initial replication factor) is not primitive recursive in general. 
-This proof uses properties of the Ackermann function, and the $H'_\alpha$, $F_\alpha$, $L_\alpha$ families of functions.
-
-% ---------------------------------------------------
-\input{gaia-chapter}
-
-%----------------------------------------------------------------------
-\chapter[Countable ordinals (after Sch\"{u}tte)]{Kurt Schütte's axiomatic definition of countable ordinals}
-
-\label{chap:schutte} 
-%ON
-
-In the present chapter, we  compare our implementation of the segment $[0,\epsilon_0)$ with a mathematical text in order to ``validate'' our constructions.
-Our reference here is the axiomatic definition of the set of countable ordinals,
-in chapter V of Kurt Schütte's book `` Proof Theory ''~\cite{schutte}.
-
-\begin{remark}
-\emph{In all this chapter, the word ``ordinal'' will be considered as a synonymous of
-``countable ordinal''}  
-\end{remark}
-
-Schütte's definition of countable ordinals relies on the following three axioms:
-
-There  exists a strictly ordered set , such that
-\begin{enumerate}
-\item  $(\mathbb{O},<)$ is well-ordered
-\item Every bounded subset of $\mathbb{O}$  is countable
-\item Every countable subset of $\mathbb{O}$  is bounded.
-\end{enumerate}
-
-Starting with these three axioms, Schütte re-defines the vocabulary about ordinal numbers: the null ordinal $0$, limits and successors, the addition of ordinals, the infinite ordinals $\omega$, $\epsilon_0$, $\Gamma_0$, etc.
-
-This chapter describes an adaptation to \coq{} of Schütte's axiomatization. 
- Unlike the rest of our libraries, our library
-\href{../theories/html/hydras.Schutte.Schutte.html}{hydras.Schutte}
-is not constructive, and relies on several axioms.
-
-\begin{itemize}
-\item First, please keep in mind  that the set of countable ordinals is not countable.
-Thus, we cannot hope to represent all countable ordinals as finite terms of an inductive type, which was possible with  the set of ordinals strictly less than $\epsilon_0$ (resp.
-$\Gamma_0$)
-\item We tried to be as close as possible to K.
-Schütte's text, which uses ``classical'' mathematics : excluded middle, Hilbert's $\epsilon$ (choice) and Russel's $\iota$ (definite description) operators.
-Both operators allow us to write definitions close to the natural mathematical language, such as ``$\textrm{succ}$ is \emph{the} least ordinal strictly greater than $\alpha$''
-\item Please note that only the library \href{../theories/html/hydras.Schutte.Schutte.html}{Schutte/*.v} is ``contaminated'' by axioms, and that the rest of our libraries remain constructive.
-\end{itemize}
-
-\section{Declarations and axioms}
-
-Let us declare a type 
-\texttt{Ord} for representing countable ordinals, and a binary relation
- \texttt{lt}. Note that, in our development, \texttt{Ord} is a type, while the \emph{set} of countable ordinals (called $\mathbb{O}$ by Schütte) 
-is the full set over the type \texttt{Ord}.
-
-\label{types:Ord}
-
-We use Florian Hatat's library on countable sets, written as he was a student of  \emph{\'Ecole Normale Supérieure de Lyon}.
-A set $A$ is countable if there is an injective function from $A$ to $\mathbb{N}$ (see
-Library \href{../theories/html/hydras.Schutte.Countable.html}%
-{\texttt{Schutte.Countable}}).
-
-\vspace{6pt}
-
-\emph{From Module\href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{Schutte.Schutte\_basics}}}
-
-\index{hydras}{Library Schutte!Types!Ord}
-
-\input{movies/snippets/Schutte_basics/OrdDecl}
-
-Schütte's first axiom tells that \texttt{lt} is a well order on the set 
-\texttt{ordinal} (The  class \texttt{WO} is defined in
-Module~\href{../theories/html/hydras.Schutte.Well_Orders.html}{Schutte.Well\_Orders.v}).
-
-\index{hydras}{Library Schutte!Type classes!WO@ WO (well order)}
-
-\label{types:WO}
-
-\input{movies/snippets/Well_Orders/Mdecl}
-
-\input{movies/snippets/Well_Orders/WODef}
-
-\input{movies/snippets/Schutte_basics/AX1}
-
-The second and third axioms say that a subset $X$ of $\mathbb{O}$ is
-(strictly) bounded if and only if it is countable.
-
-\input{movies/snippets/Schutte_basics/AX23}
-
-\texttt{AX2} and \texttt{AX3} could have been replaced by a single axiom (using the \texttt{iff} connector), but we decided to respect as most as possible the structure of Schütte's definitions.
-
-\section{Additional  axioms}
-
-The adaptation of Schütte's mathematical discourse to \coq{} led us to
-import a few axioms from the standard library. We encourage the reader to consult \coq{}'s FAQ about the safe use of axioms
- \url{https://github.com/coq/coq/wiki/The-Logic-of-Coq#axioms}.
-
-\subsubsection{Classical logic}
-
-In order to work with classical logic, we import the module
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Classical.html}{Coq.Logic.Classical}  of \coq{}'s standard library, specifically the following axiom:
-
-\begin{Coqsrc}
- Axiom classic : forall P:Prop, P \/ ~P.
-\end{Coqsrc}
-
-\subsubsection{Description operators}
-
-In order to respect Schütte's style, we imported also the library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Epsilon.html}{\texttt{Coq.Logic.Epsilon}}.  The rest of this section presents a few examples of
-how Hilbert's choice operator and Church's definite description allow us
- to write understandable definitions (close to the mathematical natural language).
-
-\subsubsection{The definition of zero}
-
-According to the  definition of a well order, every non-empty subset of \texttt{Ord} has a least element. Furthermore, this least element is unique. We would like to call this element  \texttt{zero}.
-
-\input{movies/snippets/Schutte_basics/iotaDemoa}
-\input{movies/snippets/Schutte_basics/iotaDemob}
-
-Indeed, the basic logic of  \coq{} does not allow us to eliminate a proof of a proposition 
-$\exists!\,x:A,\,P(x)$ for building a term whose type lies in the sort \texttt{Type}. 
-The reasons for this impossibility are explained in many documents~\cite{BC04, chlipalacpdt2011, Coq}.
-
-Let us import the library \texttt{Coq.Logic.Epsilon}, which contains the following axiom and lemmas.
-
-\input{movies/snippets/MoreEpsilonIota/EpsilonStatement}
-
-Hilbert's $\epsilon$ \emph{operator} is derived from this  axiom.
-
-\input{movies/snippets/MoreEpsilonIota/EpsilonDef}
-
-If we consider the \emph{unique existential} quantifier $\exists!$, we obtain
-Church's \emph{definite description operator}.
-
-\input{movies/snippets/MoreEpsilonIota/iotaDef}
-
-Indeed, the operators \texttt{epsilon} and \texttt{iota} allowed us to make our definitions 
-quite close to Schütte's text. Our libraries \href{../theories/html/hydras.Schutte.MoreEpsilonIota.html}%
-{\texttt{Schutte.MoreEpsilonIota}}
-and
-\href{../theories/html/hydras.Schutte.PartialFun.html}%
-{\texttt{Schutte.PartialFun}} are extensions of \texttt{Coq.logic.Epsilon} for making easier 
-such definitions. See also an article in french~\cite{PCiota}.
-
-\input{movies/snippets/MoreEpsilonIota/Defs}
-
-In order to use these tools,  we have to tell \coq{}  that the declared type \texttt{Ord} is not empty:
-
-\input{movies/snippets/Schutte_basics/inhOrd}
-
-We are now able to define \texttt{zero} as the least ordinal. For this purpose,
-we define a function returning the least element of any [non-empty]  subset.
-
-\emph{From Module\href{../theories/html/hydras.Schutte.Well_Orders.html}%
-  {\texttt{Schutte.Well\_Orders}}}
-
-\input{movies/snippets/Well_Orders/MDecl}
-\input{movies/snippets/Well_Orders/theLeast}
-
-\vspace{4pt}
-
-From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{~Schutte.Schutte\_basics}}
-
-\label{Constants:zero:Ord}
-\index{hydras}{Library Schutte!Constants!zero}
-
-\input{movies/snippets/Schutte_basics/zeroDef}
-
-We want to prove now that zero is less than or equal to any ordinal number.
-
-\input{movies/snippets/Schutte_basics/zeroLe}
-
-\subsubsection{Remarks on \texttt{epsilon} and \texttt{iota}}
-
- What would happen in case of a misuse of \texttt{epsilon} or \texttt{iota} ?
-For instance, one could give a unsatisfiable specification to \texttt{epsilon} or 
-a specification for \texttt{iota} that admits several realizations.
-
-Let us consider an example:
-
-\input{movies/snippets/Schutte_basics/BadBottoma}
-
-Since we won't be able to prove the proposition
-\linebreak \Verb|(exists! a: Ord, least_member (Empty_set Ord) a)|, the only properties we would be able to prove about \texttt{bottom} would be \emph{trivial} properties, 
-\emph{i.e.}, satisfied by \emph{any} element of type \texttt{Ord}, like for instance
-\texttt{bottom = bottom}, or \texttt{zero <= bottom}.
-
-\input{movies/snippets/Schutte_basics/trivialProps}
-
-On the other hand, the following attempt fails, because of the unprovable first subgoal (please notice that the second subgoal is easy to solve !).
-
-\input{movies/snippets/Schutte_basics/Failure}
-
-In short, using \texttt{epsilon} and \texttt{iota} in our implementation of countable ordinals after Schütte has two main advantages.
-
-\begin{itemize}
-\item It allows us to give a \emph{name} (using \texttt{Definition}) two witnesses 
-of existential quantifiers (let us recall that, in classical logic, one may consider non-constructive proofs of existential statements)
-\item By separating definitions from proofs of [unique] existence, one may make definitions  more concise and readable. Look for instance at 
-the definitions of  \texttt{zero}, \texttt{succ}, \texttt{plus}, etc. in the rest of this chapter.
-\end{itemize}
-%%%% ICI ICI
-
-\section{The  successor function}
-
-The definition of the function \texttt{succ:Ord -> Ord} is very concise. The successor of any ordinal $\alpha$ is the smallest ordinal strictly greater than $\alpha$.
-
-\label{Functions:succ-sch}
-\index{hydras}{Library Schutte!Functions!succ}
-
-\input{movies/snippets/Schutte_basics/succDef}
-
-Using \texttt{succ}, we define the following predicates.
-
-\input{movies/snippets/Schutte_basics/isSuccIsLimit}
-
-% \begin{remark}
-% Please look at remark~\vref{warning:coercions}.
-
-How do we prove properties of the successor function?
-First, we make its specification explicit.
-
-\input{movies/snippets/Schutte_basics/succSpec}
-
-Then, we prove that our function \texttt{succ} meets this specification.
-
-\input{movies/snippets/Schutte_basics/succOka}
-
-We have now to prove that the set of all ordinals strictly greater than $\alpha$ has a unique least element.
-But the singleton set $\{\alpha\}$ is countable, hence  bounded (by the axiom \texttt{AX3}).
-Hence; the set $\{\beta\in\mathbb{O}|\alpha < \beta\}$ is not empty
-and therefore has a unique least element.
-
-The rest of the \coq{} proof script is quite short.
-
-\input{movies/snippets/Schutte_basics/succOkb}
-
-We can ``uncap'' the description operator for proving properties of the
-\texttt{succ} function.
-
-\input{movies/snippets/Schutte_basics/succProps}
-
-\section{Finite ordinals}
-
-Using \texttt{succ}, it is now easy to define recursively all the finite ordinals.
-
-\label{sect:notation-F-sch}
-
-\input{movies/snippets/Schutte_basics/finiteDef}
-
-\section{The definition of \texttt{omega}}
-In order to define $\omega$, the first infinite ordinal, we use an operator which
-``returns'' the least upper bound (if it exists) of a subset $X\subseteq \mathbb{O}$.
-For that purpose, we first use a predicate:
-(\texttt{is\_lub $D$ \textit{lt} $X$ $a$}) if $a$ belongs to $D$ and is the least 
-upper bound  of $X$ (with respect to \textit{lt}).
-
-\input{movies/snippets/Lub/isLubDef}
-
-\input{movies/snippets/Schutte_basics/supDef}
-
-Then, we define the function \texttt{omega\_limit} which returns the least upper bound 
-of the  (denumerable) range of any sequence \texttt{s: nat -> Ord}. 
-By \texttt{AX3} this range is bounded, hence the set of its upper bounds is not empty and has a least element.
-Then we define \texttt{omega} as the limit of the sequence of finite ordinals.
-
-\input{movies/snippets/Schutte_basics/omegaDef}
-
-\label{sect:notation-omega}
-
-Among the numerous properties of the ordinal $\omega$, let us quote the following ones
-(proved in Module 
-\href{../theories/html/hydras.Schutte.Schutte_basics.html\#finite_lt_omega}{\texttt{Schutte.Schutte\_basics}})
-
-\input{movies/snippets/Schutte_basics/omegaPropsa}
-\input{movies/snippets/Schutte_basics/omegaPropsb}
-\input{movies/snippets/Schutte_basics/omegaPropsc}
-\input{movies/snippets/Schutte_basics/omegaPropsd}
-
-\subsection{Ordering functions and ordinal addition}
-
-After having defined the finite ordinals and the infinite ordinal $\omega$, we  define the sum $\alpha+\beta$ of two countable ordinals.
-Schütte's definition looks like the following one:
-
-\begin{quote}
-``$\alpha+\beta$ is the $\beta$-th ordinal greater than or equal to $\alpha$''
-\end{quote}
-
-The purpose of this section is to give a meaning to the construction
-``the $\alpha$-th element of $X$''  where $X$ is any non-empty subset of $\mathbb{O}$.
-We follow Schütte's approach, by defining the notion of \emph{ordering functions},
-a way to associate a unique ordinal to each element of $X$.
-Complete definitions and proofs can be found in Module
- \href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
-{\texttt{Schutte.Ordering\_Functions}} ).
-
-\subsection{Definitions}
-
-A \emph{segment} is a set $A$ of ordinals such that, whenever  $\alpha\in A$ and
-$\beta<\alpha$, then $\beta\in A$; a segment is  \emph{proper} if it strictly included in $\mathbb{O}$.
-
-\input{movies/snippets/Ordering_Functions/segmentDef}
-
-Let  $A$ be a segment, and $B$ a subset of $\mathbb{O}$ : an \emph{ordering function for $A$ and  $B$} is a strictly increasing bijection from $A$ to $B$.
-The set $B$ is said to be an \emph{ordering segment} of $A$.
-Our definition in \coq{} is a direct translation of the mathematical text of~\cite{schutte}.
-
-\index{maths}{Ordinal numbers!Ordering functions}
-\index{hydras}{Library Schutte!Predicates!ordering function@ordering\_function}
-
-\input{movies/snippets/Ordering_Functions/orderingFunctionDef}
-
-We are now able to associate with any subset $B$ of $\mathbb{O}$ its ordering segment and ordering function.
-
-\input{movies/snippets/Ordering_Functions/ordDef}
-
-Thus (\texttt{ord $B \;\alpha$}) is the $\alpha$-th element of $B$.
-Please note that the last definition uses the epsilon-based operator \texttt{some} and
-not \texttt{the}. This is due to the fact that we cannot prove the unicity (w.r.t. Leibniz' equality) of the ordering function of a given set. 
-By contrast, we admit the axiom  \texttt{Extensionality\_Ensembles}, from the library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Sets.Ensembles.html}{Coq.Sets.Ensembles}, so we use the operator \texttt{the} in the definition of
-\texttt{the\_ordering\_segment}.
-
-One of the main theorems of
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html\#ordering_function_ex}%
-{\texttt{Ordering\_Functions}} 
-associates a unique segment and a unique (up to extensionality) ordering function to every subset $B$ of $\mathbb{O}$.
-
-\input{movies/snippets/Ordering_Functions/orderingFunctionEx}
-
-Thus,  our function \texttt{ord}  which enumerates the elements of $B$ is defined in a non-ambiguous way.
-Let us quote the following theorems (see Library
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
-{\texttt{Schutte.Ordering\_Functions}} for more details).
-
-\input{movies/snippets/Ordering_Functions/orderingLe}
-\input{movies/snippets/Ordering_Functions/Th1352}
-
-\subsection{Ordinal addition}
-
-We are now ready to define and study addition on the type \texttt{Ord}.
-The following definitions and proofs can be consulted in Module
-\href{../theories/html/hydras.Schutte.Addition.html}%
-{\texttt{Schutte.Addition.v}}.
-
-\index{hydras}{Library Schutte!Functions!plus}
-
-\input{movies/snippets/Addition/additionDef}
-
-In other words,  $\alpha + \beta$ is the  $\beta$-th ordinal greater than or equal to $\alpha$. 
-Thanks to generic properties of ordering functions, we can show the following 
-properties of addition on $\mathbb{O}$. First, we prove a useful lemma:
-
-\input{movies/snippets/Addition/plusElim}
-
-As a use-case, let us prove that $0$ is a right neutral element of $+$.
-
-\input{movies/snippets/Addition/alphaPlusZero}
-
-The following lemmas are proved the same way.
-
-\input{movies/snippets/Addition/bunchOfLemmas}
-
-The following lemmas are not direct applications of \texttt{plus\_elim}.
-
-\input{movies/snippets/Addition/plusAssoc}
-\input{movies/snippets/Addition/finitePlusInfinite}
-
-It is interesting to compare the proof of these lemmas with the
-computational proofs of the corresponding statements in Module
-\href{../theories/html/hydras.Epsilon0.T1.html}%
-{\texttt{Epsilon0.T1}}. 
-For instance, the proof of the lemma 
-\texttt{one\_plus\_omega} uses the continuity of ordering functions (applied to  \texttt{(plus 1)}) and compares the limit of the $\omega$-sequences $i_{(i \in \mathbb{N})}$ and
-$(1+i)i_{(i \in \mathbb{N})}$, whereas in the library  \texttt{Epsilon0/T1}, the equality 
-$1+\omega=\omega$ is just proved with \texttt{reflexivity}!
-
-\subsubsection{Multiplication by a natural number}
-
-The multiplication of an ordinal by a natural number is defined in terms of addition.
-This operation is useful for the study of Cantor normal forms.
-
-\input{movies/snippets/Addition/multFin}
-
-\section{The exponential of basis \texorpdfstring{$\omega$}{omega}}
-
-In this section, we define the function which maps any $\alpha\in\mathbb{O}$ to
-the ordinal  $\omega^\alpha$, also written 
-$\phi_0(\alpha)$. 
-It is an opportunity to apply the definitions and results of the preceding section. 
-Indeed,  Schütte first defines a subset of $\mathbb{O}$: the set of additive principal ordinals, and $\phi_0$  is just defined as the ordering function of this set.
-
-\subsection{Additive principal ordinals}
-
-\index{maths}{Ordinal numbers!Additive principal ordinals}
-\index{hydras}{Library Schutte!Predicates!AP@AP (additive principal ordinals)}
-
-\begin{definition}
-A non-zero ordinal  $\alpha$ is said to be \emph{additive principal} if, for all  $\beta<\alpha$, $\beta+\alpha$ is equal to  $\alpha$.
-We call \texttt{AP} the set of additive principal ordinals.
-
-\end{definition}
-
-\noindent\emph{From Module \href{../theories/html/hydras.Schutte.AP.html}%
-{\texttt{Schutte.AP}}}
-
-\input{movies/snippets/AP/APDef}
-
-\subsection{The function \texttt{phi0}}
-
-Let us call  $\phi_0$ the ordering function of \texttt{AP}.
-In the mathematical text, we shall use indifferently the notations  $\omega^\alpha$ and$\phi_0(\alpha)$.
-
-\index{hydras}{Library Schutte!Functions!phi0}
-
-\input{movies/snippets/AP/phi0Def}
-
-\subsection{Omega-towers and the ordinal \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-Using $\phi_0$, we can define recursively the set of finite omega-towers.
-
-\input{movies/snippets/AP/omegaTower}
-
-\label{sect:epsilon0-as-limit}
-Then, the ordinal  $\epsilon_0$ is defined as the limit of the sequence of all finite towers (a kind of infinite tower).
-
-\input{movies/snippets/AP/epsilon0Def}
-
-The rest of our library \texttt{AP} is devoted to the proof of properties of additive principal ordinals, hence of the ordering function  $\phi0$ and the ordinal $\epsilon_0$ (which we could not express within the type \texttt{T1}).
-
-\subsection{Properties of the set  \texttt{AP}}
-
-The set of additive principal ordinals is not empty: it contains at least the ordinals  $1$ and  $\omega$.
-
-\inputsnippets{AP/APOne,AP/APOne_least_AP,AP/APOne_AP_omega,AP/APOne_omega_second_AP}
-
-The set  \texttt{AP} is  \emph{closed} under addition, and unbounded.
-\label{lemma:AP-plus-closed}
-
-\input{movies/snippets/AP/APPlusClosed}
-
-\input{movies/snippets/AP/APUnbounded}
-
-Finally, \texttt{AP} is (topologically) \emph{closed} and ordered by the segment of all countable ordinals.
-
-\index{hydras}{Library Schutte!Predicates!Closed}
-
-From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{~Schutte.Schutte\_basics}}
-
-\input{movies/snippets/Schutte_basics/ClosedDef}
-\input{movies/snippets/AP/APClosed}
-
-\subsubsection{Properties of the function \texorpdfstring{$\phi_0$}{phi0}}
-
-The ordering function $\phi_0$ of the set \texttt{AP} is defined on the full set $\mathbb{O}$ and is continuous (Schütte calls such a  function  \emph{normal}).
-
-\begin{Coqsrc}
-Theorem normal_phi0 : normal phi0 AP.
-\end{Coqsrc}
-
-The following properties come from  the definition of $\phi_0$ as the ordering function of \texttt{AP}.
-It may be interesting to compare these proofs with the computational ones described in Chapter ~\ref{chap:T1}.
-
-\input{movies/snippets/AP/APPhi0}
-
-\subsection{A last example}
-\label{Ex42-schutte}
-
-Let us prove again the equality $\omega+42+\omega^2= \omega^2$.Let us recall that $\omega^2$ is an abbreviation of $\phi_0(2)$,
-\emph{i.e} the third  additive principal ordinal.
-
-\inputsnippets{Schutte/Ex42a}
-
-Our proof is very different from the computational proof of Sect~\vref{Ex42-E0}.
-By definition of additive principal ordinals, 
-it suffices to prove the inequality $\omega+42< \phi_0(2)$.
-
-\inputsnippets{Schutte/Ex42b}
-
-Since the set \texttt{AP} of additive principals  is closed under addition
-(by Lemma \texttt{AP\_plus\_closed}, page~\pageref{lemma:AP-plus-closed}) , it suffices to prove the inequalities $\omega<\phi_0(2)$ and $42<\phi_0(2)$.
-
-\inputsnippets{Schutte/Ex42d, Schutte/Ex42c, Schutte/Ex42e}
-
-\section{More about \texorpdfstring{$\epsilon_0$}{\texttt{epsilon0}}}
-
-Let us recall that the limit ordinal  $\epsilon_0$ cannot be written within the type \texttt{T1}.
-Since we are now considering the set of all countable ordinals, we can now prove some properties of this ordinal.
-
-We prove the inequality  $\alpha<\omega^\alpha$ whenever $\alpha < \epsilon_0$.
-\emph{Note that this condition was implicit in Module
-\href{../theories/html/hydras.Epsilon0.T1.html\#lt_phi0}{Epsilon0.T1}.}
-
-\input{movies/snippets/AP/ltPhi0}
-
-The proof is as follows:
-\begin{enumerate}
-\item Since $\alpha<\epsilon_0$, consider the least $i$ such that $\alpha$ is strictly less than the omega-tower of height $i$.
-\item
-  \begin{itemize}
-  \item If $i=0$, then the result is trivial (because $\alpha=0$)
- \item  Otherwise let $i=j+1$; 
-          $\alpha$ is greater than or equal to the omega-tower of height $j$.
-         By monotony,  $\phi_0(\alpha)$ is greater than or equal to 
-        the omega-tower of height $j+1$, thus strictly greater than $\alpha$
-  \end{itemize}
- \end{enumerate}
-
-Moreover,  $\epsilon_0$ is the least ordinal $\alpha$ that verifies the equality 
-$\alpha = \omega^\alpha$, in other words, the least fixpoint of the function  $\phi_0$.
-
-\input{movies/snippets/AP/epsilon0Lfp}
-
-\section{Critical ordinals}
-
-\index{maths}{Ordinal numbers!Critical ordinals}
-\index{hydras}{Library Schutte!Predicates!Cr@Cr (critical ordinals)}
-
-For any  (countable) ordinal $\alpha$, the set $\textit{Cr}(\alpha)$ is inductively defined 
-as follows by Schütte (p.81 of~\cite{schutte}).
-
-\begin{quote}
-  \begin{itemize}
-  \item $\textit{Cr}(0)$ is the set \textit{AP} of additive principal ordinals.
-  \item If $0<\alpha$, then $\textit{Cr}(\alpha)$ is the intersection of all the sets of fixpoints of the $\textit{Cr}(\beta)$ for $\beta<\alpha$.
-  \end{itemize}
-\end{quote}
-
-This definition is translated in \coq{} in 
-Module \href{../theories/html/hydras.Schutte.Critical.html}%
-{\texttt{Schutte.Critical}}, as the least fixpoint of a functional.
-
-\input{movies/snippets/Critical/CrDef}
-
-Lets us denote by $\phi_\alpha$ the ordering function of the set $\textit{Cr}(\alpha)$ and by $A_\alpha$ its ordering segment.
-
-\input{movies/snippets/Critical/phiDef}
-
-\label{sect:phi-schutte}
-
-For instance,  we prove that $\textit{Cr}(0)$ is the set of additive principals and that $\epsilon_0$
-belongs to $\textit{Cr}(1)$.
-
-\inputsnippets{Critical/CrZeroAP,
-  Critical/epsilon0Cr1}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
- Prove that $\epsilon_0$ is the least element of $\textit{Cr}(1)$.
-\end{exercise}
-
-\subsection{A flavor of infinity}
-
-The family of the $\textit{Cr}(\alpha)$s is made of infinitely many unbounded (hence infinite) sets.
-Let us quote Lemma 5, p. 82  of~\cite{schutte}:
-\begin{quote}
-  For all $\alpha$, the set $\textit{Cr}(\alpha)$ is closed (for the least upper bound of non-empty countable sets) and unbounded.
-\end{quote}
-
-We prove this result by a transfinite induction on $\alpha$ of the conjunction of  both properties.
-
-\index{maths}{Transfinite induction}
-
-\input{movies/snippets/Critical/Lemma5}
-
-\section{Cantor normal form}
-
-The notion of Cantor normal form is defined for all countable ordinals.
-Nevertheless, note that, contrary to the implementation based on type \texttt{T1},
-the Cantor normal form of an ordinal $\alpha$ may contain $\alpha$ as a 
-sub-term\footnote{This would prevent us from trying to represent Cantor normal forms as finite trees (like in Sect.~\ref{sec:T1-inductive-def})}.
-
-\index{maths}{Ordinal numbers!Cantor normal form}
-\index{hydras}{Library Schutte!Predicates!is\_cnf\_of@is\_cnf\_of (to be a Cantor normal form of}
-
-We represent  Cantor normal forms as lists of ordinals.
-A  list $l$ is a Cantor normal form of a given ordinal $\alpha$ if it satisfies two conditions:
-
-\begin{itemize}
-\item The list  $l$ is sorted (in decreasing order) w.r.t. the order $\leq$
-\item The sum of all the  $\omega^{\beta_i}$ where the $\beta_i$ are the terms of $l$ (in this order) is equal to $\alpha$.
-\end{itemize}
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Schutte.CNF.html\#cnf_t}%
-{\texttt{Schutte.CNF}}}
-
-\input{movies/snippets/CNF/Defs}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction on $\alpha$, we prove that every countable ordinal $\alpha$ 
- has at least a Cantor normal form.
-
-\input{movies/snippets/CNF/cnfExists}
-
-By structural induction on lists, we prove that this normal form is unique.
-
-\inputsnippets{CNF/cnfUnicity,CNF/cnfExUnique}
-
-Finally, the following two lemmas relate  $\epsilon_0$ with Cantor normal forms.
-
-If $\alpha<\epsilon_0$, then the Cantor normal form of $\alpha$ is made of ordinals strictly less than $\alpha$.
-
-\input{movies/snippets/CNF/cnfLtEpsilon0}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Please consider the following statement :
-
-\begin{Coqsrc}
-Lemma cnf_lt_epsilon0_iff : 
- forall l alpha, 
-   is_cnf_of alpha l ->  
-   (alpha < epsilon0 <->  Forall (fun beta =>  beta < alpha) l).
-\end{Coqsrc}
-
-Is it true ?
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/schutte_cnf_counter_example.v}{exercises/ordinals/schutte\_cnf\_counter\_example.v}.}
-\end{exercise}
-
-Finally, the Cantor normal form of $\epsilon_0$ is just $\omega^{\epsilon_0}$.
-
-\input{movies/snippets/CNF/cnfOfEpsilon0}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Implement pages 82 to 85 of~\cite{schutte} (critical, strongly critical, maximal critical ordinals, Feferman's ordinal $\Gamma_0$).
-\end{project}
-
-\begin{remark}
-The sub-directory
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/Gamma0}{theories/oridinals/Gamma0} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form.
-\end{remark}
-
-\section{An embedding of \texttt{T1} into \texttt{Ord}}
-
-Our library 
-\href{../theories/html/hydras.Schutte.Correctness_E0.html}%
-{\texttt{Schutte.Correctness\_E0}} establishes the link between two very different modelizations of ordinal numbers. In other words, it ``validates'' a data structure in terms of
-a classical mathematical discourse considered as a model. 
-First, we define a function from \texttt{T1} into  \texttt{Ord} by structural recursion.
-
-\input{movies/snippets/Correctness_E0/injectDef}
-
-This function enjoys good commutation properties with respect to the main operations which
-allow us to build Cantor normal forms.
-
-\input{movies/snippets/Correctness_E0/commutationLemmas}
-\input{movies/snippets/Correctness_E0/injectPlus}
-\input{movies/snippets/Correctness_E0/injectMultFinR}
-
-% \begin{Coqsrc}
-% Theorem inject_mono (beta gamma : T1) :
-%   T1.lt  beta gamma -> 
-%   T1.nf beta -> T1.nf gamma -> 
-%   inject beta < inject gamma.
-
-% Theorem inject_injective (beta gamma : T1) : nf beta -> nf gamma ->
-%   inject beta = inject gamma -> beta = gamma.
-% \end{Coqsrc}
-
-Finally, we prove that \texttt{inject} is a bijection from the set of all terms of \texttt{T1} in normal form to the set 
-\texttt{members epsilon0} of the elements of \texttt{Ord} strictly less than  $\epsilon_0$.
-
-\input{movies/snippets/Correctness_E0/injectLtEpsilon0}
-\input{movies/snippets/Correctness_E0/embedding}
-
-\subsection{Remarks}
-Let us recall that the library \href{../theories/html/hydras.Schutte.Schutte.html}%
-{\texttt{Schutte}} depends on five \emph{axioms} and lies explicitly in the  
-framework of classical logic with a weak version of the axiom of choice
-(please look at the documentation of
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.ChoiceFacts.html}{\texttt{Coq.Logic.ChoiceFacts}}).
-Nevertheless, the other modules:
-\href{../theories/html/hydras.Epsilon0.Epsilon0.html}%
-{\texttt{Epsilon0}},
-\href{../theories/html/hydras.Hydra.Hydra.html}%
-{\texttt{Hydra}}, et 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}%
-{\texttt{Gamma0}}
-do not import any axioms and are really constructive.
-
-\index{hydras}{Projects}
-\begin{project}
-There is no construction of ordinal multiplication in~\cite{schutte}. 
-It would be interesting to derive this operation from Schütte's axioms,
-and prove its consistence with multiplication in ordinal notations for 
-$\epsilon_0$ and $\Gamma_0$.
-\end{project}
-
-\section{Related work}
-
-In~\cite{grimm:hal-00911710}, José Grimm establishes the consistency between our ordinal notations \texttt{T1} and \texttt{T2} (Veblen normal form) and his implementation
-of ordinal numbers after Bourbaki's set theory.
-
-The Gaia project ~\url{https://github.com/coq-community/gaia} maintains Grimm's  theory of ordinals as part of coq-community on GitHub. Integration
-of the present ordinal theory with Gaia, i.e., relating the different notions of ordinals
-and transferring relevant results, is an interesting project.
-First experiments in that direction are developped in
-the \href{https://github.com/coq-community/hydra-battles/blob/master/theories/gaia/}{theories/gaia/} directory.
-
-\chapter{The Ordinal \texorpdfstring{$\Gamma_0$}{Gamma0} (first draft)}
-
-\emph{This chapter and the files it presents are still very incomplete, considering the impressive properties of $\Gamma_0$~\cite{Gallier91}.
-We hope to add new material soon, and accept contributions!}
-
-\begin{todo}
-  Build a T2Bridge! (note: Gamma0's wellfoundedness is missing in \gaia (?).
-\end{todo}
-
-\section{Introduction}
-We present a notation system for the ordinal $\Gamma_0$, following Chapter V, Section 14 of~\cite{schutte}: ``A notation system for the ordinals $<\Gamma_0$''.
-We try to be as close as possible to Schütte's text and usual practices of \coq{} developments.
-
-The ordinal $\Gamma_0$ is defined in Section 13 of ~\cite{schutte} as the least \emph{strongly critical ordinal}. It is widely known as the \emph{Feferman-Schütte ordinal}.
-
-Section V, 13 of~\cite{schutte} defines \emph{strongly critical} and
-\emph{maximal $\alpha$-critical} ordinals:
-
-\begin{itemize}
-\item $\alpha$ is strongly critical if
-$\alpha$ is $\alpha$-critical,
-\item $\gamma$ is maximal $\alpha$-critical if $\gamma$ is $\alpha$-critical, and, for all $\xi>\alpha$, $\gamma$ is not $\xi$-critical.
-
-\end{itemize}
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Schutte.Critical.html\#strongly_critical}%
-{\texttt{Schutte.Critical}}}
-
-\input{movies/snippets/Critical/Gamma0Def}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that a (countable)  ordinal $\alpha$ is strongly critical iff 
-$\phi_\alpha(0)=\alpha$ (Theorem 13.13 of~\cite{schutte} ). 
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that the set of strongly critical ordinals is unbounded and closed (Theorem 13.14 of~\cite{schutte} ).
-Thus this set is not empty,  hence has a least element.
-Otherwise, the definition of $\Gamma_0$ above would be useless.
-\end{project}
-
-In the present version of this development, we  only study $\Gamma_0$ as a notation system, much more powerful than the ordinal notation for $\epsilon_0$.
-
-%\index{Projects}
-%
-%\begin{project}
-% Schûtte's section 13 of~\cite{schutte} contain several definitions and lemmas which give another view on $\Gamma_0$.  We leave it as a project to implement them in \coq{}.  
-%\end{project}
-
-\section{The type \texttt{T2} of ordinal terms}
-
-The notation system for ordinals less than $\gamma_0$ comes from the following theorem of~\cite{schutte}, where $\psi\,\alpha$ is the ordering function 
-of the set of maximal $\alpha$-critical ordinals.
-
-\index{hydras}{Library Gamma0!Types!T2}
-
-\begin{quote}
-  Any ordinal $\not= 0$ which is not strongly critical can be expressed in terms of $+$ and $\psi$.
-\end{quote}
-
-\index{hydras}{Projects}
-\begin{project}
-This theorem is not formally proved in this development yet. It should be!
-\end{project}
-
-Like in Chapter~\ref{chap:T1}, we define an inductive type with two constructors, one for $0$, the other for the construction $\psi(\alpha,\beta)\times(n+1)+\gamma$, adapting a Manolios-Vroon-like notation~\cite{Manolios2005} to
-\emph{Veblen normal forms}.
-\label{types:T2}
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T2}%
-{\texttt{Gamma0.T2}}}
-
-\inputsnippets{T2/T2Def, T2/psiDef}
-
-Like in chapter~\ref{chap:T1}, we get familiar with the type \texttt{T2} by recognizing simple constructs like finite ordinals, $\omega$, etc., as inhabitants of \texttt{T2}.
-
-\input{movies/snippets/T2/Notations}
-
-\section{A strict order on T2}
-
-Let us define a strict order on type \texttt{T2}. The following definition is 
-an adaptation of Schütte's, taking into account the multiplications by a natural number (inspired by~\cite{Manolios2005}, and also present in \texttt{T1}).
-
-\label{sect:t2-lt-def}
-
-\input{movies/snippets/T2/ltDef}
-
-Seven constructors! In order to get accustomed with this definition, let us look at a small set of examples, covering all the constructors of \texttt{lt}.
-
-\input{movies/snippets/T2/ltExamples}
-
-\index{hydras}{Projects}
-\begin{project}
-Write a tactic that solves automatically goals of the form (\texttt{$\alpha$ t2< $\beta$}), where $\alpha$ and $\beta$ are closed terms of type \texttt{T2}.
-\end{project}
-
-\section{Veblen normal form}
-\begin{definition}
-  A term of the form $\psi(\alpha_1,\beta_1)\times n_1+ \psi(\alpha_2,\beta_2)\times n_2+\dots+\psi(\alpha_k,\beta_k)\times n_k$ is said to be in
- \emph{[Veblen] normal form} if for every $i<n$, $\psi(\alpha_i,\beta_i)<\psi(\alpha_{i+1},\beta_{i+1})$, all the $\alpha_i$ and $\beta_i$ are in normal form, and all the $n_i$ are strictly positive integers.
-\end{definition}
-
-\input{movies/snippets/T2/nfDef}
-
-Let us look at some positive examples (we have to prove some inversion lemmas before proving counter-examples).
-
-\input{movies/snippets/T2/nfExamples}
-
-\subsection{Length of a term}
-
-The notion of \emph{term length} is introduced by Schütte as a helper for proving (at least) the \emph{trichotomy} property and transitivity of the strict order \texttt{lt} on \texttt{T2}.
-These properties are proved by induction on length.
-
-\subsection{Trichotomy}
-
-\emph{Trichotomy} is another name for the well-known property of decidable total ordering (like Standard Library's \texttt{Compare\_dec.lt\_eq\_lt\_dec}).
-
-We first prove by induction on $l$ the following lemma:
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0\#tricho_aux}%
-{\texttt{Gamma0.Gamma0}}}
-
-\input{movies/snippets/Gamma0/tricho}
-
-\input{movies/snippets/Gamma0/compare}
-
-With the help of \texttt{compare}, we get a boolean version of \texttt{nf}
-(being in Veblen normal form).
-
-\input{movies/snippets/Gamma0/nfb}
-
-\section{Main functions on \texttt{T2}}
-
-\subsection{Successor}
-The successor function is defined by structural recursion.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#succ}%
-{\texttt{Gamma0.T2}}}
-
-\input{movies/snippets/T2/succDef}
-
-\subsection{Addition}
-
-Like for Cantor normal forms (see Sect.~\ref{sect:infix-plus-T1}),  the definition of addition in \texttt{T2}  requires comparison between ordinal terms.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0T2.html\#succ}%
-  {\texttt{Gamma0.Gamma0}}}
-
-\input{movies/snippets/Gamma0/plusDef}
-
-\subsection{The Veblen function \texorpdfstring{$\phi$}{\texttt{phi}}}
-
-The enumeration function of critical ordinals, presented in Sect.~\vref{sect:phi-schutte}, is recursively defined in type \texttt{T2}.
-
-\input{movies/snippets/Gamma0/phiDef}
-
-Despite its complexity, the function \texttt{phi} is well adapted to proofs by simplification or computation.
-
-The relation between the constructor $\psi$ and the function $\phi$ is
-studied in~\cite{schutte}, and partially implemented in this development.
-\emph{Please contribute!}
-
-For instance, the following theorem states that, if $\gamma$ is the sum of a limit ordinal $\beta$ and a finite ordinal $n$, and $\beta$ is a fixpoint of
-$\phi(\alpha)$, then $\psi(\alpha,\gamma)=\phi_\alpha(\gamma+1)$.
-
-\input{movies/snippets/Gamma0/phiPsi}
-\input{movies/snippets/Gamma0/Ex9}
-
-On the other hand, $\phi$ can be expressed in terms of $\psi$.
-
-\input{movies/snippets/Gamma0/phiOfPsi}
-\input{movies/snippets/Gamma0/Ex10}
-
-\index{hydras}{Projects}
-\begin{project}
-Please study a way to pretty print ordinal terms in Veblen normal form (see Section~\vref{sect:ppT1}).
-\end{project}
-
-\section{An ordinal notation for \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}}
-
-In order to consider type \texttt{T2} as an ordinal notation, we have to build an instance of class \texttt{ON} (See Definition page~\pageref{types:ON}).
-
-First, we define a type that contains only terms in Veblen normal form, and redefine \texttt{lt} and \texttt{compare} by delegation (see for comparison the construction of type \texttt{E0} in Sect.~\vref{sect:E0-def}).
-
-\input{movies/snippets/Gamma0/G0a}
-\input{movies/snippets/Gamma0/G0b}
-
-Then, we buils an instance of class \texttt{ON}. 
-function \texttt{compare} is correct.
-
-\input{movies/snippets/Gamma0/ltSto}
-\input{movies/snippets/Gamma0/ltWf}
-\input{movies/snippets/Gamma0/ONGamma0}
-\input{movies/snippets/Gamma0/G0z}
-
-\begin{remark}
-The proof of \texttt{lt\_wf} has been written by \'Evelyne Contejean, using her library on the recursive path ordering (see also remark~\vref{remark:a3pat}).
-\end{remark}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that \texttt{Epsilon0} (page~\pageref{instance-epsilon0})
-is a sub-notation system of \texttt{Gamma0}.
-
-Prove that the implementations of \texttt{succ}, \texttt{+}, $\phi_0$, etc.
-are compatible in both notation systems.
-
-Note that a function \texttt{T1\_inj} from \texttt{T1} to \texttt{T2} has already been defined. It may help to complete the task.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T1_to_T2}%
-{\texttt{Gamma0.T2}}}
-
-\input{movies/snippets/T2/T1ToT2}
-
-\end{project}
-
-\begin{project}
-Prove that the notation system \texttt{Gamma0} is a correct implementation 
-of the segment $[0,\Gamma_0)$ of the set of countable ordinals.
-\end{project}

### 2022-05-08 doc/gaia-chapter.tex
reorganization of tex files

@@
 
-By simple rewriting, we import the following lemmas.
+\index{hydras}{Exercises}
+
+\begin{exercise}
+
+  We could also have defined \texttt{oplus} by a structurally
+  recursive definition, as in 
+  \vref{sec:oplusdefT1}.
+
+\inputsnippets{GHessenberg/FixOplus}
+
+Prove that \texttt{Hessenberg.oplus} refines
+the new function. 
+\end{exercise}
+
+\subsubsection{A few lemmas}
+
+Many properties of Hessenberg's sum can be obtained from
+\href{../theories/html/hydras.Epsilon0.Hessenberg.html}{hydras.Epsilon0.Hessenberg} just by sequences of rewritings.
 
@@
 
+The following lemmas are also borrowed from\href{../theories/html/hydras.Epsilon0.Canon.html}%
+{\texttt{hydras.Epsilon0.Canon}}.
+
 \inputsnippets{GCanon/gcanonLt}
@@
 %\inputsnippets{GCanon/glimitCanonSNotZero}
-\subsubsection{Canonical sequences and limits}
-
-The following lemma (borrowed from\href{../theories/html/hydras.Epsilon0.Canon.html}%
-{\texttt{hydras.Epsilon0.Canon}}) links limit ordinals to canonical sequences.
+%\subsubsection{Canonical sequences and limits}
 
@@
 \inputsnippets{nfwfgaia/LimitV2Def, nfwfgaia/LimitOFDef}
+
+We can prove the following statements, in \gaia's  style.
 
@@
  \inputsnippets{GPaths/pathsDefsb}
+
+The following examples are proved using the \texttt{Epsilon0.Paths.path\_tac} tactic.
+
+\inputsnippets{GPaths/pathExamples}
 

### 2022-05-08 doc/hydra-chapter.tex
reorganization of tex files

--- 2022-04-11 doc/part-hydras.tex
@@
-\section*{Introduction}
-
-In this part, we present a development  for the \coq{} proof assistant, after the work of Kirby and Paris~\cite{KP82}. This formalization contains the following main parts:
-
-\begin{itemize}
-\item Representation in \coq{} of hydras and hydra battles.
-\item A proof that every battle is finite and won by Hercules.
-This proof is based on a \emph{variant} which maps any hydra to an ordinal strictly less than $\epsilon_0$ and is strictly decreasing along any battle.
-
-\item Using a combinatorial toolkit designed by J.~Ketonen and R.~Solovay~\cite{KS81}, we prove that, for any ordinal $\mu<\epsilon_0$, there exists no such variant mapping any hydra to an ordinal strictly less than $\mu$.
-Thus, the complexity of $\epsilon_0$ is really needed in the previous proof.
-
-\item We prove a relation between the length of a ``classic''  kind of  battles \footnote{This class is also called \emph{standard} in this document (text and proofs).
-The \emph{replication factor} of the hydra is exactly $i$ at the $i$-th round of the battle (see Sect~\vref{sect:replication-def}).}
-and the Wainer-Hardy hierarchy of ``rapidly growing functions'' $H_\alpha$~\cite{Wainer1970}.
-The considered class of battles, which we call \emph{standard},  is the most considered one in the scientific  literature (including popularization).
-\end{itemize}
-
-Simply put, this document tries to combine the scientific interest of two articles~\cite{KP82, KS81} and a book~\cite{schutte} with the playful activity of truly proving theorems.
-We hope  that such a work, besides exploring a nice piece of discrete mathematics,
-will show how \coq{} and its standard library are well fitted to help us to understand some non-trivial mathematical developments, and also to experiment the constructive parts of  the proof through functional programming.
-
- We also hope to provide a little clarification on infinity (both potential and actual) through the notions of function, computation, limit,
- type and proof.
-
- \subsection*{Compatibility with \gaia (in progress)}
- \label{sect:gaia-first-intro}
-The \gaia project~\cite{Gaia} by Jos\'e Grimm, Alban Quadrat,  and Carlos Simpson,
-aims  to formalize mathematics in \coq  in
-the style of Nicolas Bourbaki.
-It contains many definitions and results about ordinal numbers.
-In Chapter~\ref{gaia-chapter}, we present some modules which allow \HydrasLib' users to apply lemmas proven in \gaia, and vice versa.
-Remarks about compatibility with \gaia are signalled with the picture \gaiasign.
-
-%\section{Remarks}
-
-\subsection*{Difference from Kirby and Paris's work}
-In~\cite{KP82}, Kirby and Paris show  that there is no proof of termination of all hydra battles in Peano Arithmetic (PA).
-Since we are used to writing proofs in higher order logic, the restriction to PA was quite unnatural for us.
-So we chose to prove another statement without any reference to PA, by considering a class of proofs indexed by ordinal numbers up to $\epsilon_0$.
-
-\subsection*{State of the development}
-The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on  notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
-We added new material: axiomatic definitions of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, function derinition by equations, etc.
-
-We are now working in order to make clumsy proofs more readable, simplify definitions, and ``factorize'' proofs as much as possible. 
-Many possible improvements are suggested as ``todo''s or ``projects'' in this text.
-
-\section*{Future work (projects)}
-\index{hydras}{Projects}
-
-This document and the proof scripts are far from being complete.
-
-First, there must be a lot of typos to correct, references and index items to add. Many proofs are too complex and should be simplified, etc.
-
-The following extensions are planned, but help is needed:
-
-\begin{itemize}
-\item Semiautomatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
-%\item Extension to $\Gamma_0$ (in Veblen normal form)
-\item More lemmas about hierarchies of rapidly growing
-  functions, and their relationship 
-    with primitive recursive functions and provability in Peano arithmetic 
-(following~\cite{KS81, KP82}).
-\item From \coq's point of view, this development could be used as an illustration of the evolution of the software, every time new libraries and sets of tactics could help to simplify the proofs.
-\end{itemize}
-
-\subsection*{Main references}
-
-In our development, we adapt the definitions and prove many theorems which
-we found in the following articles. 
-\begin{itemize}
-\item ``Accessible independence results for Peano arithmetic''  by Laurie Kirby and Jeff Paris~\cite{KP82}
-\item ''Rapidly growing Ramsey Functions'' by Jussi Ketonen and Robert Solovay~\cite{KS81}
-\item ``The Termite and the Tower'', by Will Sladek~\cite{Sladek07thetermite}
-\item Chapter V of ``Proof Theory'' by Kurt Schütte~\cite{schutte}
-\end{itemize}
-
 \chapter{Hydras and hydra games}
@@
 \index{maths}{Notations!Interval}
-
-%--------------------------------------------------------------
-
-\chapter{Introduction to ordinal numbers and ordinal notations}
-\label{chap:ON}
-
-The proof of termination of all hydra battles presented in~\cite{KP82} is based
-on \emph{ordinal numbers}.
-From a mathematical point of view, an ordinal is a representative of an equivalence class for isomorphisms of  totally ordered well-founded sets.
-
-For the computer scientist, ordinals are tools for proving the totality of a given recursive function, or termination of a transition system. \emph{Ordinal arithmetic} 
-provides a set of functions whose properties, like \emph{monotony}, allow to define \emph{variants}, \emph{i.e.} strictly decreasing measures used in proofs of termination.
-
-\vspace{4pt}
-
-Let us have a look at Figure~\ref{fig:ordinal-sequence}. It presents a few items of a  sequence of ordinal numbers, which extends the sequence of natural numbers.
-
-\begin{figure}[h]
-  \centering
-\fbox{\Large
-  \begin{minipage}{1.0\linewidth}
-  \begin{align*}
-     &\textcolor{blue}{0},\,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,\ldots\\
-&\textcolor{red}{\omega},\,\omega+1,\omega+2,\omega+3,\ldots\\
-&\textcolor{red}{\omega\times 2},\,\omega\times 2+1,\ldots, \textcolor{red}{\omega\times 3},\,\omega\times 3+1,\ldots, \textcolor{red}{\omega\times 4},\ldots\\
-&\textcolor{red}{\omega^2},\ldots, \textcolor{red}{\omega^2\times 42},\ldots,\textcolor{red}{\omega^3},\ldots, \textcolor{red}{\omega^4},\omega^4+1,\ldots\\
-&\textcolor{red}{\omega^\omega},\ldots, \textcolor{red}{\omega^\omega+\omega^7\times 8},\ldots,\textcolor{red}{\omega^\omega\times 2},\omega^\omega\times 2+1, \ldots\\
-&\textcolor{red}{\omega^{\omega^\omega}},\ldots, \textcolor{red}{\omega^{\omega^\omega}+\omega^\omega\times 42+ \omega^{55}+\omega}, \ldots, \textcolor{red}{\omega^{\omega^{\omega+1}}}, \omega^{\omega^{\omega+1}}+1,\dots\\
-& \textcolor{red}{\epsilon_0 (= \omega^{\epsilon_0)}}, \epsilon_0+1, \epsilon_0+2, \epsilon_0+3, \ldots\\
-& \textcolor{red}{\epsilon_1}, \ldots, \textcolor{red}{\epsilon_2}, \ldots, \textcolor{red}{\epsilon_\omega},\ldots \\
-& \textcolor{red}{\Gamma_0}, \Gamma_0+1, \Gamma_0+2, \Gamma_0+3,\ldots, \textcolor{red}{\Gamma_0+\omega}, \ldots\\
-&\ldots
-  \end{align*}   
-  \end{minipage}}
-
-  \caption{A short overview of the sequence of ordinal numbers}
-  \label{fig:ordinal-sequence}
-\end{figure}
-
-Let us comment some features of this figure:
-
-\begin{itemize}
-\item The ordinals are listed in a strictly increasing order. 
-\item Dots : ``$\ldots$'' stand for  infinite sequences of ordinals, not shown for lack of space.
-For instance, the ordinal $42$ is not shown in the first line, but it exists, somewhere between $17$ and $\omega$.
-\item Each ordinal printed in black is the immediate successor of another ordinal.
-We call it a \emph{successor} ordinal.
-For instance, $12$ is the successor of $11$, and $\omega^4+1$ the successor of $\omega^4$.
-\item Ordinals (displayed in red)  that  follow immediately dots are called \emph{limit ordinals}.
-With respect to the order induced by this sequence, any limit ordinal $\alpha$ is the least upper bound of  the set $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
-\item
-For instance $\omega$ is the least upper bound of the set of all finite ordinals (in the first line). It is also the first limit ordinal, and the first infinite ordinal, in the sense that 
-the set $\mathbb{O}_\omega$ is infinite.
-\item The ordinal $\epsilon_0$ is the first number which is equal to its own exponential of base $\omega$.
-It plays an important role in proof theory, and is particularly studied in chapters~\ref{chap:T1} to \ref{chap:alpha-large}.
-\item Any ordinal is  either the ordinal \textcolor{blue}{$0$},
-a successor ordinal, or a \textcolor{red}{limit ordinal}.
-\end{itemize}
-
-\section{The mathematical point of view}
-
-\subsection{Well-ordered sets}
-Let us start with some definitions.
-A  \emph{well-ordered set} is a set provided with a binary relation $<$ which has the following properties.
-\begin{description}
-\item[irreflexivity] : $\forall x\in A, x\not< x$
-\item[transitivity] : $\forall x\,y\,z\in A, x<y \Rightarrow y<z \Rightarrow x<z$
-\item[trichotomy]: $\forall x\,y\in A, x<y \vee x = y \vee y < x$
-\item[well foundedness]: $<$ is well-founded (every element of $A$ is accessible)\footnote{In classical mathematics, we would say that there is no infinite sequence $a_1>a_2> \dots a_n> a_{n+1}\dots$ in $A$.
-In contrast, \coq's standard library contains
-an inductive definition of a predicate \texttt{Acc} which allows us to write 
-constructive proofs of accessibility (See \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}).}.
-\end{description}
-
-The best known examples of well-ordered sets are the set $\mathbb{N}$ of natural numbers (with the usual order $<$), as well as any finite segment $[0,i)=\{j\in\mathbb{N}\,|\,j<i\}$.
-The disjoint union of two copies of $\mathbb{N}$, \emph{i.e.} the set $\{0,1\}\times\mathbb{N}$ is also well-ordered,
-with respect to the order below:
-
-\begin{align*}
-(i,j) < (i,k) & \;\textit{\textbf{if} }\; j < k\\
-(0,k) < (1,l) & \;\textit{\textbf{for\,any}}\;k \;\textit{\textbf{and}} \; l
-\end{align*}
-
-\subsection{Ordinal numbers}
-
-\index{maths}{Ordinal numbers}
-
-Let $(A,<_A)$ and $(B,<_B)$ two well-ordered sets. $A$ and $B$ are said to have \emph{the same order type} if 
-there exists a strictly monotonous bijection $b$ from $A$ to $B$, \emph{i.e.} which verifies the proposition
-$\forall x\,y\in A,\, x <_A y \Rightarrow b(x) <_B  b(y)$.
-
-Having the same order type is an equivalence relation between well-ordered sets. Ordinal numbers (in short: \emph{ordinals})  are descriptions (\emph{names}) of the equivalence classes.
-For instance, the order type of $(\mathbb{N},<)$ is associated with the ordinal called  $\omega$, and the order we considered on 
-the disjoint union of $\mathbb{N}$ and itself is named $\omega+\omega$.
-
-In a set-theoretic framework, one can consider any ordinal $\alpha$ as a well-ordered set, whose  elements are just the ordinals strictly less than $\alpha$, \emph{i.e.} the \emph{segment} $\mathbb{O}_\alpha=[0, \alpha)$.
-So, one can speak about \emph{finite}, \emph{infinite}, \emph{countable}, etc., ordinals.
-Nevertheless, since we work within type theory,
-we do not identify ordinals as sets of ordinals, but consider the correspondence between ordinals and sets of ordinals as the function that maps $\alpha$ to $\mathbb{O}_\alpha$.
-For instance $\mathbb{O}_\omega=\mathbb{N}$, and $\mathbb{O}_7=\{0,1,2,3,4,5,6\}$.
-
-We cannot cite all the literature published on ordinals since Cantor's book
-\cite{cantorbook}, and 
-leave it to the reader to explore the bibliography.
-
-The introduction of Jos\'e Grimm's report~\cite{grimm:hal-00911710} contains a nice presentation of the main properties of ordinals.
-
-\section{Ordinal numbers in Coq}
-
-Two kinds of representation of ordinals are defined in our development.
-
-\begin{itemize}
-\item A ``mathematical'' representation of the set of countable ordinal numbers, after Kurt Schütte~\cite{schutte}.
-This representation uses several (hopefully harmless) axioms.
-We use it as a reference for proving the correctness of ordinal notations.
-  We are also progressively importing \gaia's~\cite{Gaia} definitions and theorems, based on Bourbaki's set theory~\cite{Grimm2016}.
-Chapter~\ref{gaia-chapter} is dedicated to the connexion between libraries \HydrasLib and \gaia.
-
-\item A family of \emph{ordinal notations} (also called \emph{[ordinal] notation systems}), \emph{i.e.} data types used to represent segments $[0,\mu)$, where $\mu$ is some countable ordinal.
-Each ordinal notation is defined inside the Calculus of Inductive Constructions (without axioms).
-Many functions are defined, allowing proofs by computation.
-Note that proofs of
-correctness of a given ordinal notation with respect to Schütte's model obviously use axioms.
-Please execute the \texttt{Print Assumptions} command in case of doubt.
-\end{itemize}
-
-% \section{Countable ordinals}
-
-% Chapter~\ref{chap:schutte} of this document presents an adaptation to \coq{} of an axiomatization in classical logic of the set of countable ordinals by K. Schütte~\cite{schutte}. 
-% That formalization is quite complex, technical and unshamedly non-constructive,  so we put its description  in the last chapter of this document.
-
-% Please note that Schütte considers the (uncountable) set $\mathbb{O}$ of all countable ordinals.
-This set is well ordered (which is one of Schütte's axioms), and associates to any ordinal $\alpha$ the segment $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
-
-% In our adaptation to \coq{}, we declare a type \texttt{Ord}, a binary relation \texttt{lt} (with infix notation \texttt{"\_<\_"}, and assume Schütte's axiom. In Chapter~\ref{chap:schutte},
-% we derive some interesting properties of countable ordinals from these axioms.
-
-It is interesting to compare proofs of a given property (for instance the associativity of addition) both in the computational framework of some ordinal notation, and in mathematical  models
-of Schütte or \gaia.
-
-\section{Ordinal Notations}
-
-Fortunately, the ordinals we need for  studying hydra battles are much simpler than Schütte's, and can be represented as quite simple data types in \gallina.
-
-Let $\alpha$ be some (countable) ordinal; 
-in \coq{} terms, we call \emph{ordinal notation for $\alpha$} a structure composed 
-of:
-\begin{itemize}
-\item A data type $A$ for representing all ordinals strictly below $\alpha$,
-\item A well founded order $<$ on $A$, 
-\item A correct function for comparing two ordinals. Note  that the reflexive closure of $<$ is thus a \emph{total order}.
-\end{itemize}
-
-Such a structure may be proved correct relatively to a bigger ordinal notation or to Schütte's or \gaia's model.
-
-\subsection{Classes for ordinal notations}
-
-From the \coq{} user's point of view, an ordinal notation is
-a structure allowing to compare two ordinals by computation, and proving by well-founded induction.
-
-\subsubsection{The \texttt{Comparison} and \texttt{Comparable} classes}
-
-We use the operational class \texttt{Comparison} of comparison functions to define the \texttt{Comparable} class, contributed by Jérémy Damour and Théo Zimmermann, which allows us to apply generic lemmas and tactics about decidable strict orders.
-The correctness of the comparison function is expressed through Stdlib's type 
-\texttt{Datatypes.CompareSpec} as specialized by \texttt{Datatypes.CompSpec}.
-
-\begin{Coqsrc}
-  Inductive CompareSpec (Peq Plt Pgt : Prop) :
-  comparison -> Prop :=
-    CompEq : Peq -> CompareSpec Peq Plt Pgt Eq
-  | CompLt : Plt -> CompareSpec Peq Plt Pgt Lt
-  | CompGt : Pgt -> CompareSpec Peq Plt Pgt Gt.
-
-Definition CompSpec {A} (eq lt : A -> A -> Prop) (x y : A) :=
- CompareSpec (eq x y) (lt x y) (lt y x).
-\end{Coqsrc}
-
-\emph{From Module~\href{../theories/html/hydras.Prelude/mparable.html\#Hvariant}{Prelude.Comparable}}
-
-\label{sect:comparable-def}
-
-\input{movies/snippets/Comparable/ComparableDef}
-
-\subsubsection{The \texttt{ON} class}
-
-The following class definition, parameterized with a type $A$, a binary relation \texttt{lt} on $A$, specifies that \texttt{lt} is a well-founded strict order, provided with a correct comparison function.
-
-\vspace{4pt}
-\noindent\emph{From
-Library~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}}
-
-\label{types:ON}
-\index{hydras}{Library OrdinalNotations!Type classes!ON}
-
-\input{movies/snippets/ON_Generic/ONDef}
-
-We give  also a few handy definitions and lemmas for any ordinal notation.
-
-\label{sect:on-lt-notation}
-\label{sect:on-le-notation}
-\label{sect:measure-ON}
-\label{sect:bigO-ON}
-
-\inputsnippets{ON_Generic/ONDefsa}
-\inputsnippets{ON_Generic/ONDefsb}
-
-\begin{remark}
-The infix notations \texttt{o<} and \texttt{o<=} are defined in order to make apparent the distinction between the various notation scopes that may co-exist in a same statement.
-So the infix \texttt{<} and \texttt{<=} are reserved to the natural numbers.
-In the mathematical formulas, however, we still use $<$ and $\leq$ for comparing ordinals.
-\end{remark}
-
-% \subsection{Ordinal notations and  termination measures}
-% \label{sect:measure-ON}
-
-% The following lemma (together with the type class mechnism) allows us to define termination measures over any ordinal notation.
-It is just an application of  the libraries \texttt{Coq.Wellfounded.Inverse\_Image}
-% and  \texttt{Coq.Wellfounded.Inclusion}.
-
-% \begin{Coqsrc}
-% Definition measure_lt {A:Type}{lt: relation A}
-%             {compare : A -> A -> comparison}
-%             {on : ON lt compare}
-%             {B : Type} (m : B -> A) : relation B :=
-%              fun x y => on_lt (m x) (m y).
-
-% Lemma wf_measure  {A:Type}(lt: relation A)
-%             {compare : A -> A -> comparison}
-%             {on : ON lt compare}
-%             {B : Type}
-%             (m : B -> A):  well_founded (measure_lt m). 
-% \end{Coqsrc}
-
-% A simple example of application is given in Sect.~\vref{sect:merge-example}.
-
-\section{Example: the ordinal \texorpdfstring{$\omega$}{omega}}
-
-The simplest example of ordinal notation is built over the type \texttt{nat} of \coq's standard library. We just have to apply already proven lemmas about Peano numbers.
-
-\vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_Omega.html}{OrdinalNotations.ON\_Omega}}
-
-\inputsnippets{ON_Omega/OmegaDefa, ON_Omega/OmegaDefb}
-
-\section{Sum of  two ordinal notations}
-
-Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective types \texttt{A} and \texttt{B}.
-
- We consider a new strict order
-on the disjoint sum of the associated types, by putting all elements of \texttt{A} before the elements of \texttt{B} (thanks to Standard Library's relation operator \texttt{le\_AsB}).
-
-\vspace{4pt}
-\noindent
-\emph{From Library~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}}.
-
-\begin{Coqanswer}
-Inductive
-le_AsB (A B : Type) (leA : A -> A -> Prop) (leB : B -> B -> Prop)
-  : A + B -> A + B -> Prop :=
-| le_aa : forall x y : A, leA x y -> le_AsB A B leA leB (inl x) (inl y)
-| le_ab : forall (x : A) (y : B), le_AsB A B leA leB (inl x) (inr y)
-| le_bb : forall x y : B, leB x y -> le_AsB A B leA leB (inr x) (inr y)
-\end{Coqanswer}
-
-%\pagebreak
-
-\vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_plus.html}{OrdinalNotations.ON\_plus}}
-
-\input{movies/snippets/ON_plus/Defs}
-
-In order to build an instance of \texttt{Comparable}, we have to define a correct comparison function.
-
-\inputsnippets{ON_plus/compareDef, ON_plus/compareCorrect}
-\inputsnippets{ON_plus/plusComp}
-
-Standard library's lemma \texttt{Wellfounded.Disjoint\_Union.wf\_disjoint\_sum} 
-is applied  to prove that our order \texttt{lt} is well-founded, allowing us to build an instance of \texttt{ON}:
-
-\inputsnippets{ON_plus/ltWf, ON_plus/OnPlus}
-
-\subsection{The ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
-
-The ordinal $\omega+\omega$ (also known as $\omega\times 2$) may be represented as the concatenation 
-of two copies of $\omega$ (Figure~\ref{fig:omega-plus-omega}).
-It is also represented by the two first lines of Figure~\ref{fig:ordinal-sequence}.
-
-\begin{figure}[h]
-   \centering
-   \begin{tikzpicture}[very thick, scale=0.5]
-\begin{scope}[color=blue]
-\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
-\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
-\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
-\node (Adots) at (6,0) {$\ldots$};
-\node(An) at (8,0)[label=below:$n$]{$\bullet$};
-\node(A2) at (10,0)[label=below:$n+1$]{$\bullet$};
-\node (Adots1) at (12,0) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red]
-\node(B0) at (14,0)[label=below:$0$,label=above:\textcolor{red}{$\omega$}]{$\bullet$};
-\node(B1) at (16,0)[label=below:$1$, label=above:$\omega+1$]{$\bullet$};
-\node(B2) at (18,0)[label=below:$2$,label=above:$\omega+2$]{$\bullet$};
-\node (Bdots) at (20,0) {$\ldots$};
-\node (Bn) at (22,0) [label=below:$p$, label=above:$\omega+p$]{$\bullet$};
-\node (Bdots2) at (24,0) {$\ldots$};
-\end{scope}
-\end{tikzpicture}
-   \caption{\textcolor{blue}{$\omega+{\color{red}\omega}$}}
-   \label{fig:omega-plus-omega}
- \end{figure}
-
-We can define this notation in \coq{} as an instance of \texttt{ON\_plus}.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/OmegaPlusOmegaDef}
-
-\vspace{4pt}
-
-We can now define abbreviations.
-For instance, the finite ordinals are represented by terms built with  the constructor \texttt{inl}, and the first infinite ordinal $\omega$ by the term \texttt{(inr 0)}.
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Omega_plus_omega/finiteOmega}
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Omega_plus_omega/ltOmega}
-
-% \label{warning:coercions}
-% \index{Coq!Coercions} 
-% \begin{remark}
-% Beware of coercions and notation scopes!
-% Let us consider the following goal:
-
-% \begin{Coqsrc}
-%  Goal (6 o< 8).
-%  auto with arith.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% 1 subgoal (ID 9)
-
-%   ============================
-%   6 o< 8
-% \end{Coqanswer}
-
-% Please keep in mind that the current notation scope interprets the infix \texttt{``<''} as the predicate \texttt{Omega\_plus\_omega.lt} and not \texttt{Nat.lt}.
-More,  the coercion mechanism converts the terms \texttt{6:nat} [resp.
-\texttt{8:nat} ]
-% into \texttt{inl 6} [resp. \texttt{inl 8}].  So, the initial goal is correctly interpreted by \coq{}, but not as an inequality between two natural numbers.
-
-% \begin{Coqsrc}
-% Set Printing All.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% 1 subgoal (ID 337)
-
-%   ============================
-%   @on_lt nat Peano.lt Nat.compare Omega (S (S (S (S (S (S O))))))
-%     (S (S (S (S (S (S (S (S O))))))))
-% \end{Coqanswer}
-
-% Anyway, the initial goal is provable, using \texttt{le\_AsB}'s first constructor.
-
-% \begin{Coqsrc}
-%   constructor; auto with arith.
-% Qed.
-% \end{Coqsrc}
-
-% \end{remark}
-%
-
-\section{Limits and successors}
-
-Let us look again at our implementation of $\omega+\omega$. We can classify its elements into three categories:
-
-\begin{itemize}
-\item The least ordinal, \texttt{(inl 0)}, also known as  \texttt{(fin 0)}.
-\item The first infinite ordinal $\omega$.
-\item The remaining ordinals, either of the form \texttt{(inl (S $i$))} or \texttt{(inr (S $i$))} (in black on Figure~\ref{fig:ordinal-sequence}), called \emph{successor ordinals}.
-\end{itemize}
-
-\subsection{Definitions}
-It would be interesting to specify at the most generic level, what is a zero, a successor or a limit ordinal. Let $<$ be a strict order on a type $A$.
-
-\begin{itemize}
-\item A \emph{least} element is a minorant (in the large sense) of the full set  on $A$,
-\item $y$ is a \emph{successor} of $x$ if $x<y$ and there is no element between $x$ and $y$. We will also say that $x$ is a \emph{predecessor} of $y$.
-\item $x$ is a \emph{limit} if $x$ is not a least element, and for any $y$ such that $yo<x$,
- there exists some $z$ such that $y<z<x$.
-\end{itemize}
-
-The following definitions are in Library \href{../theories/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
-
-\input{movies/snippets/MoreOrders/Defs}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove, that, in any ordinal notation system, every ordinal has at most one predecessor, and at most one successor.
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/predSuccUnicity.v}{exercises/ordinals/predSuccUnicity.v}.}
-
-\end{exercise}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove, that, in any ordinal notation system, if $\beta$ is a successor of $\alpha$,
-then for any $\gamma$, $\gamma<\beta$ implies 
-$\gamma\leq\alpha$.
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/lt_succ_le.v}{exercises/ordinals/lt\_succ\_le.v}.}
-\end{exercise}
-
-\subsection{Limits and successors in \texorpdfstring{$\omega+\omega$}{omega+omega}}
-
-Using the definitions above, we can prove the following lemma:
-
-\vspace{4pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/limitIff}
-
-\vspace{4pt}
-
-Regarding successors, let us define the following function and prove its correctness:
-
-\input{movies/snippets/ON_Omega_plus_omega/succDef}
-
-\input{movies/snippets/ON_Omega_plus_omega/succCorrect}
-
-\vspace{4pt}
-
-We can also check whether an ordinal is a successor by a simple computation:
-
-\input{movies/snippets/ON_Omega_plus_omega/succb}
-
-\vspace{4pt}
-
-Finally, the nature of any ordinal is decidable (inside this notation system) :
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/ZeroLimitSuccDec}
-
-\section{Product of ordinal notations}
-
-Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective  ordered types \texttt{A} and \texttt{B}.
-The product of \texttt{NA} and \texttt{NB} is considered as the concatenation of $B$ copies of $A$, ordered by the lexicographic order on $B\times A$.
-
-In \coq{}, we build an instance of class \texttt{ON} through a sequence of steps as for the sum of ordinal notations.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_mult.html}{OrdinalNotations.ON\_mult}}
-
-\inputsnippets{ON_mult/Defs}
-\inputsnippets{ON_mult/multComp,  ON_mult/ONMult}
-\inputsnippets{ON_mult/endDefs}
-
-\section{The ordinal \texorpdfstring{$\omega^2$}{omega2}}
-
-The ordinal $\omega^2$ (also called $\phi_0(2)$, see Chap.~\ref{chap:schutte}), is an instance of the multiplication presented in the preceding section.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
-
-\inputsnippets{ON_Omega2/Omega2Def, ON_Omega2/Defs}
-\inputsnippets{ON_Omega2/Ex1}
-
-\subsection{Arithmetic of \texorpdfstring{$\omega^2$}{omega^2}}
-
-\subsubsection{Successor}
-
-The successor of any ordinal is defined by a simple pattern-matching.
-
-\input{movies/snippets/ON_Omega2/succ}
-
-This function is proved to be correct w.r.t. the \texttt{Successor} predicate.
-
-\input{movies/snippets/ON_Omega2/succOK}
-
-\input{movies/snippets/ON_Omega2/succLemmas}
-
-\subsubsection{Addition}
-
-We can define on \texttt{Omega2} an addition which extends the addition on \texttt{nat}. Please note that this operation is not commutative:
-
-\inputsnippets{ON_Omega2/plusDef, ON_Omega2/plusExamples}
-
-\subsubsection{Multiplication}
-
-The restriction of ordinal multiplication to the segment $[0,\omega^2)$ is not a total function.
-For instance $\omega\times\omega= \omega^2$ is outside the set of represented values.
-Nevertheless, we can define two operations mixing natural numbers and ordinals.
-
-\input{movies/snippets/ON_Omega2/multFinDef}
-\input{movies/snippets/ON_Omega2/multFinExamples}
-
-Multiplication with a finite ordinal and addition are related through the following lemma:
-
-\inputsnippets{ON_Omega2/uniqueDecompositiona,
-  ON_Omega2/uniqueDecompositionb,
-  ON_Omega2/uniqueDecompositionz}
-
-\subsection{A proof of termination using \texorpdfstring{$\omega^2$}{omega^2}} 
-\label{sect:merge-example}
-
-Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily a total function which merges two lists (example contributed by Pascal Manoury).
-
-%\index{coq}{Commands!Function}
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
-
-\inputsnippets{ON_Omega2/Merge, ON_Omega2/Mergeb}
-
-\subsection{Yet another  proof of impossibility}
-\label{omega2-case}
-
-In Sect.~\vref{omega-case}, we proved that there exists no variant from \texttt{Hydra} to \texttt{(nat,$<$)}
-(\emph{i.e.} the ordinal $\omega$) for proving the termination of all hydra battles.
-We  prove now that  the ordinal $\omega^2$ is also insufficient for this purpose.
-
-The proof we are going to comment has exactly the same structure as in Section~\ref{omega-case}.
- Nevertheless, the proof of technical  lemmas is a little more complex, due to 
- the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
-Consider for instance that there exists an infinite number of ordinals  between
-$\omega$ and $\omega\times 2$.
-
-The detailed  proof script is in the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/Hydra/Omega2_Small.v}{theories/ordinals/Hydra/Omega2\_Small.v}.
-
-\subsubsection{Preliminaries}
-Let us assume there is a variant from \texttt{Hydra} into $\omega^2$  for proving the   termination of all hydra battles.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
-
-\input{movies/snippets/Omega2_Small/Impossibility}
-\input{movies/snippets/Omega2_Small/Impossibilitya}
-
-We  follow the same pattern as in Sect.~\ref{omega-case}.
-First, we define an injection $\iota$ from type \texttt{ON\_Omega2.t} into \texttt{Hydra}, by
- associating to  each ordinal $\omega\times i+ j = (i,j)$ the hydra with $i$ branches of length $2$ and
-$j$ branches of length $1$.
-
-%% revenir ici
-
-\vspace{4pt}
-\emph{From Module ~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
-
-\input{movies/snippets/Omega2_Small/Impossibilityc}
-
-For instance, Figure~\vref{fig:essai2} shows the hydra associated to the ordinal 
-$(3,5)$, a.k.a. $\omega\times 3 + 5$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-\node (foot) at (6,0) {$\bullet$};
-\node (N1) at (1,3) {$\bullet$};
-\node (N2) at (3,3) {$\bullet$};
-\node (N3) at (5,3) {$\bullet$};
-\node (N4) at (8,3) {$\Smiley[2][green]$};
-\node (N5) at (11,3) {$\Smiley[2][green]$};
-\node (N6) at (14,3) {$\Smiley[2][green]$};
-\node (N7) at (17,3){$\Smiley[2][green]$};
-\node (N8) at (20,3){$\Smiley[2][green]$};
-\node  (N9) at (0,5) {$\Smiley[2][green]$};
-\node (N10) at (2,5) {$\Smiley[2][green]$};
-\node (N11) at (4,5) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N5);
-\draw (foot) -- (N6);
-\draw (foot) to [bend right=10] (N7);
-\draw (foot) to [bend right=15] (N8);
-\draw (N1) to [bend left=10] (N9);
-\draw (N2) -- (N10);
-\draw (N3) -- (N11);
-\end{tikzpicture}
-\caption{\label{fig:essai2}
-The hydra $\iota(\omega\times 3+5)$}
-\end{figure}
-
-Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:h-omega2-small}}
- The hydra \texttt{big\_h}.
-\end{figure}
-
-\input{movies/snippets/Omega2_Small/Impossibilityb}
-
- In a second step, we build a ``smaller'' hydra\footnote{With respect to the measure $m$.}.
-
-\input{movies/snippets/Omega2_Small/Impossibilityd}
-
-\vspace{4pt}
-
-Like in Sect.~\ref{omega-case}, we prove the inequality $m\;\texttt{big\_h} \,\texttt{o<=}\, m\;\texttt{small\_h} \,\texttt{o<=}\, m\;\texttt{big\_h}$, which is obviously false.
-
-\subsubsection{Proof of the inequality \texttt{m small\_h o< m big\_h}}
-
-In order to prove the inequality  \texttt{m\_lt: m small\_h o< m big\_h}, it suffices to
-build a battle transforming \texttt{big\_h} into \texttt{small\_h}.
-
-First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h} as $(i,j)$.
-If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
-If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the following result.
-
-\input{movies/snippets/Omega2_Small/bigToSmall}
-
-Since $m$ is a variant, we infer the following inequality:
-
-\input{movies/snippets/Omega2_Small/mLt}
-
-\subsubsection{Proof of the inequality \texttt{m big\_h o<= m small\_h} }
-
-The proof of the inequality \texttt{m big\_h o<= m small\_h} is quite more complex than in Sect~\ref{omega-case}.
-If we consider any ordinal $\alpha=(i,j)$, where $i>0$, there exists an infinite number of
-ordinals strictly less than $\alpha$, and there exists an infinite number of battles that start from
-$\iota(\alpha)$.
-Indeed, at any configuration $\iota(k,0)$, where $k>0$, the hydra can freely choose any replication number.
-Intuitively, the measure of such a hydra must be large enough for taking into account
-all the possible battles issued from that hydra.
-Let us now give more technical details.
-
-The first steps of our proof start a well-founded induction over $\omega^2$.
-
-\input{movies/snippets/Omega2_Small/mGe}
-
-A case analysis on $i$ and $j$ allows us to
-consider three cases :
-
-\begin{itemize}
-\item $i=j=0$: the inequality is trivial.
-\item $i=1+l, j=0$ ($(i,j)$ is a limit ordinal): By the induction hypothesis \texttt{IHij},
-  $(l,k)\texttt{o<=} m(\iota(l,k))$ for any $k$. But (by the rules of the hydra game), $\iota(i,0)$ is transformed into any $\iota(l,k)$ in one round. Thus $m(\iota(l,k)) < m(\iota(i,0))$ for any $k$.
-  Therefore, $(l,k) <  m(\iota(i,0))$ for any $k$, thus
-  $(i,0)\,\texttt{o<=}\,m(\iota(i,0))$.
- \item $j= l+1$  ($(i,j)$ is a successor). We apply the induction hypothesis to the pair $(i,l)$.
- \end{itemize}
-
- Please look at the proof script for more details.
-
- \input{movies/snippets/Omega2_Small/mGeb}
-
-\subsubsection{End of the proof}
-From \texttt{m\_ge}, we get \texttt{$m$\, big\_h o<= $m$\, small\_h = $m$\,(iota (m big\_h)) }. 
-Since $<$ is a strict order (irreflexive  and transitive), this inequality is incompatible with the strict inequality  \texttt{m small\_h o< m big\_h} (lemma \texttt{m\_lt}).
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#Impossible}{Hydra.Omega2\_Small}}
-
-  \input{movies/snippets/Omega2_Small/Impossible}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into $\omega^2$ for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\begin{remark}
- We  prove in Chapter~\ref{ks-chapter} a generalization of the impossibility lemmas of
-Sect.~\ref{omega-case} and this section, with the same proof structure, but with much more 
-complex technical details.
- \end{remark}
-
-% \index{Exercises}
-% \begin{exercise}
-
-% \label{sec:orgheadline63}
-% Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.
-
-% \begin{itemize}
-% \item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
-%   lexicographic ordering
-% \item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.
-
-% For instance, the following inequality holds:
-% \[\langle 4,3,3,3,3,3,3,2,2,2 \rangle\,<\,\langle 4,4,2 \rangle\]
-% \end{itemize}
-
-% \end{exercise}
-
-\section{A notation for finite ordinals}
-
-Let $n$ be some natural number. The segment associated with $n$ is the interval 
-$[0,n)\,=\,\{0,1,\dots,n-1\}$. 
-One may represent the ordinal $n$ by a sigma type.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
-
-\label{def:Finite-ord-type}
-
-\input{movies/snippets/ON_Finite/Defs}
-
-\subsubsection{Examples}
-
-For instance, let us build two elements of the segment $[0, 7)$, \emph{i.e.} two
-inhabitants of   type (\texttt{t 7}), and prove a simple  inequality (see Fig.~\ref{fig:O7}).
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (N0) at (0,0) {$\bullet$};
-\node (i0) at (0,1) {$0$};
-\node (N1) at (2,0) {$\bullet$};
-\node (i1) at (2,1) {$1$};
-\node (N2) at (4,0) {$\bullet$};
-\node (i2) at (4,1) {$2$};
-\node (N3) at (6,0) {$\bullet$};
-\node (i3) at (6,1) {$3$};
-\node (N4) at (8,0) {$\bullet$};
-\node (i4) at (8,1) {$4$};
-\node (N5) at (10,0) {$\bullet$};
-\node (i5) at (10,1) {$5$};
-\node (N6) at (12,0) {$\bullet$};
-\node (i6) at (12,1) {$6$};
-\node(alpha1) at (4,-1) {$\alpha_1$};
-\node(alpha2) at (10,-1) {$\beta_1$};
-\end{tikzpicture}
-
-\caption{The segment $\mathbb{O}_7$\label{fig:O7}}
-\end{figure}
-
-%\index{coq}{Commands!Program}
-
-\input{movies/snippets/ON_Finite/Example1}
-
-Note that the type (\texttt{t 0}) is empty, and that, for any natural number
- $n$, $n$ does not belong to (\texttt{t $n$}).
-
- \inputsnippets{ON_Finite/t0Empty, ON_Finite/bad10}
-
-\subsubsection{Comparison function}
-
-In order to build an instance of \texttt{ON}, we define a comparison function,  and prove its correctiness.
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Finite/compareDef}
-
-\begin{remark}
- The proof of \texttt{compare\_correct} uses a well-known pattern of \coq{}.
-Let us consider  the following subgoal.
-
-\input{movies/snippets/ON_Finite/compareCorrectb}
-
-Applying the tactic \texttt{f\_equal} generates a simpler subgoal.
-
-\input{movies/snippets/ON_Finite/compareCorrectc}
-
-We have now to prove that there exists at most one  proof of (\texttt{Nat.ltb x0 (S n)}). This is not obvious, but  a consequence of the following lemma of library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Eqdep_dec.html}{Coq.Logic.Eqdep\_dec}.
-
-\index{Coq}{Unicity of equality proofs}
-\label{sect:eq-proof-unicity}
-
-\begin{Coqanswer}
-eq_proofs_unicity_on :
-forall (A : Type) (x : A),
-(forall y : A, x = y \/ x <> y) -> 
-forall (y : A) (p1 p2 : x = y), p1 = p2
-\end{Coqanswer}
-
-Thus unicity of proofs of \texttt{Nat.ltb x0 (S n)}  comes from the decidability of
-equality on type \texttt{bool}.
-This is why we used the boolean function \texttt{Nat.ltb} instead of the inductive predicate \texttt{Nat.lt} in the definition of type \texttt{t $n$} (see page~\pageref{def: Finite-ord-type}).
-For more information about this pattern, please look at the numerous mailing lists and 
-FAQs on \coq{}).
-
-\inputsnippets{ON_Finite/compareCorrectd}
-
-\end{remark}
-
-\begin{remark}
-
-Please note  that attempting to compare a term  of type (\texttt{t $n$}) with a term of
-type (\texttt{t $p$})  leads to an error if $n$ and $p$ are not convertible.
-
-\input{movies/snippets/ON_Finite/Example2}
-
-\end{remark}
-
-\subsubsection{Building an instance of \texttt{ON}}
-
-Applying lemmas of the libraries \texttt{Coq.Wellfounded.Inverse\_Image}, \linebreak
- \texttt{Coq.Wellfounded.Inclusion}, and \texttt{Coq.Arith.Wf\_nat}, we prove that our
-relation \texttt{lt} is well founded.
-
-\input{movies/snippets/ON_Finite/ltWf}
-
-Now we can build our instance of \texttt{OrdinalNotation}.
-
-\input{movies/snippets/ON_Finite/ONInstance}
-
-\begin{remark}
-It is important to keep in mind  that the integer $n$ is not an ``element'' of \texttt{FinOrd $n$}.
-In set-theoretic presentations of ordinals, the set associated with the ordinal $n$ is $\{0,1,\dots,n-1\}$.
-In our formalization, the interpretation of an ordinal as a set is realized by the function \texttt{bigO} (see Section\vref{sect:bigO-ON}).
-\end{remark}
-
-\begin{remark}
- There is no interesting arithmetic on finite ordinals, since functions like successor, addition, etc.,  cannot be represented in \coq{} as \emph{total} functions.
-\end{remark}
-
-\section{See also \dots}
-
-Finite ordinals are also formalized in \ssreflect/\mathcomp~\cite{MCB}.
-In Module~\href{../theories/html/gaia_hydras.ON_gfinite.html}{gaia\_hydras.ON\_gfinite}, we build an instance of class \texttt{ON}.
-
-\inputsnippets{ON_gfinite/finordLtDef, ON_gfinite/finordON}
-
-\inputsnippets{ON_gfinite/Examples, ON_gfinite/Examplesb}
-
-  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.
-
-%%%
-
-\section{Comparing two ordinal notations}
-
-It is sometimes useful to compare two ordinal notations with respect to expressive power
-(the segment of ordinals  they represent).
-
-The following class specifies a strict inclusion of segments.
-The notation \texttt{OA} describes a segment $[0,\alpha)$, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented in \texttt{OA}).
-We require also  that the comparison functions of the two notation systems are compatible.
-
-\begin{figure}[h]
-   \centering
-   \begin{tikzpicture}[very thick, scale=0.6]
-\begin{scope}[color=blue]
-\node (A) at (0,0) {$A$};
-\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
-\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
-\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
-\node (Adots) at (6,0) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red]
-\node (B) at (0,2) {$B$};
-\node(B0) at (2,2)[label=above:$0$]{$\bullet$};
-\node(B1) at (3,2)[label=above:$1$]{$\bullet$};
-\node(B2) at (4,2)[label=above:$2$]{$\bullet$};
-\node (Bdots) at (6,2) {$\ldots$};
-\node (b) at (8,2) [label=above:$\omega$]{$\bullet$};
-\node (bsucc) at (9,2) [label=above:$\omega+1$]{$\bullet$};
-\node (Bdots2) at (10,2) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red!50!blue]
-\draw [->,thin] (A0) -- node [auto] {$\iota$} (B0);
-\draw [->,thin] (A1) -- node [auto] {$\iota$} (B1);
-\draw [->,thin] (A2) -- node [auto] {$\iota$} (B2);
-\draw [->,thin] (Adots) -- node [auto] {$\iota$} (Bdots);
-\end{scope}
-\end{tikzpicture}
-   \caption{\textcolor{blue}{$A$} is a sub-segment  of \textcolor{red}{$B$}}
-   \label{fig:subsegment}
- \end{figure}
-
-If \texttt{OB} is presumed to be correct, then we may consider that \texttt{OA} ``inherits'' its correctness from the bigger notation system \texttt{OB}.
-
-\label{types:SubON}
-\index{hydras}{Library OrdinalNotations!Type classes!SubON}
-
-following definition
-(in ~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{ON\_Generic}).
-
-\input{movies/snippets/ON_Generic/SubONDef}
-
-For instance, we prove that \texttt{Omega} is a sub-notation of
-\texttt{Omega\_plus\_Omega} (with $\omega$ as the first ``new'' ordinal, and \texttt{fin} as the injection).
-
-\vspace{10pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/Incl}
-
-We can also show that, if $i<j$, then the segment $[0,i)$ is a ``sub-segment'' of
-$[0,j)$. Since the terms  ($t\;i$) and ($t\;j$) are not convertible, we consider a ``cast'' 
-function $\iota$ from ($t\;i$) into ($t\;j$), and prove that this function is  a monotonous bijection  from ($t\;i$) to
-the segment $[0,i)$ of ($t\;j$).
-
-%\index{coq}{Commands!Program}
-
-\vspace{10pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
-
-\inputsnippets{ON_Finite/InclIJ, ON_Finite/InclIJa,
-  ON_Finite/InclIJb,ON_Finite/InclIJc, ON_Finite/InclIJd,
-ON_Finite/InclIJz}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove that \texttt{Omega\_plus\_Omega} cannot be a sub-notation of \texttt{Omega}.
-\end{exercise}
-
-\index{hydras}{Projects}
-\begin{project}
-Adapt the definition of \texttt{Hvariant} (Sect.~\ref{sect:hvariant-def}) in order to
-have an ordinal notation as argument. Prove that if $O_A$ is a sub-notation of $O_B$, then any variant defined on  $O_A$ can be automatically transformed into 
-a variant on $O_B$.
-\end{project}
-
-\section{Comparing an ordinal notation with Schütte's model}
-
-Finally, it may be interesting to compare an ordinal notation with the more theoretical model from Schütte (well, at least with our formalization of that model).
-This would be a relative proof of correctness of the considered  ordinal  notation.
-
-The following class specifies that a notation \texttt{OA} describes a segment $[0,\alpha)$,
-where $\alpha$ is a countable ordinal \emph{à la}  Schütte.
-
-\label{types:ON-for}
-\index{hydras}{Library OrdinalNotations!Type classes!ON\_correct}
-
-\input{movies/snippets/ON_Generic/ONCorrect}
-
-For instance, the following theorem tells that \texttt{Epsilon0}, our notation system for the segment $[0,\epsilon0)$ is a correct implementation of the theoretically defined  ordinal $\epsilon_0$
-(see chapter~\ref{chap:schutte} for more details).
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
-
-\input{movies/snippets/Correctness_E0/Epsilon0Correct}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Same work, but replace Schütte's model with \gaia's.
-\end{project}
-
-\begin{project}
-  When you have read Chapter~\ref{chap:schutte}, prove that the sum of two ordinal notations \texttt{ON\_plus} implements the addition of ordinals.
-\end{project}
-
-\section{Isomorphism of ordinal notations}
-
-In some cases we want to show that two notation systems describe the same segment (for instance $[0,3+\omega)$ and $[0,\omega)$\;).
-For this purpose, one may prove that the two notation systems are order-isomorphic.
-
-\index{hydras}{Library OrdinalNotations!Type classes!ON\_Iso}
-
-\label{types:ON-iso}
-
-\input{movies/snippets/ON_Generic/ONIso}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-\label{exo:i-plus-omega}
-Let $i$ be some natural number. Prove that the notation systems 
-\texttt{Omega} and (\texttt{ON\_plus (OrdFin $i$) Omega}) are isomorphic.
-
-{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$, that we prove also in larger notation systems, as well as in Schütte's model.}
-This exercise is partially solved for $i=3$ (in ~\href{../theories/html/hydras.OrdinalNotations.Example_3PlusOmega.html}{OrdinalNotations.Example\_3PlusOmega}).
-
-\end{exercise}
-
-\index{hydras}{Projects}
-\label{exo:ON-mult}
-\begin{project}
-% Define in \coq{} the product of two ordinal notations $N_A$ and $N_B$.
-% If $A$ [resp. $B$] is the underlying type of $N_A$ [resp. $N_B$], the
-% product \texttt{ON\_mult $N_A$ $N_B$} is implemented over the cartesian product $B\times A$ (with the lexicographic ordering).
-
-This exercise is about the non-commutativity of the multiplication of ordinals, reflected in ordinal notations.
-
-For instance, the
-elements of the product (\texttt{ON\_mult Omega (FinOrd 3)}) are ordered as follows.
-\[(0,0),(0,1),(0,2),(0,3),(0,4),\dots,{\color{red}(1,0),} (1,1),(1,2),\dots, {\color{red}(2,0)},(2,1),(2,2),\dots\]
-
-Note that the elements of  (\texttt{ON\_mult (FinOrd 3) Omega}) are differently ordered (without limit ordinals):
-\[(0,0),(1,0),(2,0),(0,1),(1,1),(2,1),(0,2),(1,2),(2,2),(0,3),\dots\]
-
-Prove formally  that (\texttt{ON\_mult (FinOrd $i$) Omega}) is isomorphic to
-\texttt{Omega}  whilst
-\texttt{Omega}  is a sub-notation of (\texttt{ON\_mult Omega (FinOrd $i$)}),
-for any strictly positive $i$.
-
-\textbf{Note:} Like Exercise~\ref{exo:i-plus-omega}, this project corresponds to the [in]equalities $i+\omega=\omega<\omega+i$, for any natural number $i$.
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-Consider two isomorphic ordinal notations \texttt{OA} and \texttt{OB}.
-Prove that, if \texttt{OA} [resp. \texttt{OB}] is a correct implementation 
-of $\alpha$ [resp. $\beta$], then $\alpha=\beta$.
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-\label{project:succ-limit-dec}
-Add to the class \texttt{ON} the requirement that for any $\alpha$ it is decidable whether $\alpha$ is $0$, a successor or a limit ordinal.
-
-\textbf{Hint:}   Beware of the instances associated with sum and product of notations!
-  You may consider additional fields 
-to make the sum and product of notations ``compositional''.
-
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-\label{project:on-setoid}
-Reconsider the  class \texttt{ON}, with an equivalence instead of Leibniz equality.
-\end{project}
-
-%%%% ICI ICI
-
-\section{Other ordinal notations}
-%%% TODO : Fix the multiplication function in branch FixOmegaOmega
-
-% \index{hydras}{Projects}
-
-% \begin{project}
-% The directory \texttt{theories/OmegaOmega} contains an ad-hoc formalization of $\omega^\omega$, contributed by Pascal Manoury.
-Every ordinal $\alpha$ is represented by a list $l$ whose elements are the coefficients of $\omega$ in  the Cantor normal form of $\alpha$ (in reverse order).
-For instance, the ordinal
-% $\omega^{8}\times 5 + \omega^{6}\times 8 + \omega^2\times 10 + \omega + 7$ is represented by the list \texttt{[5;0;8;0;0.0;10,1,7]}.
-
-%  Develop this representation and compare it with the other ordinal notations.
-
-% \end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Let $N_A$ be a notation system for ordinals strictly less than $\alpha$, 
-with the strict order $(A,<_A)$. Please build the notation system
-\texttt{ON\_Expl $N_A$}, on the type of multisets of elements of $A$
-(or, if preferred, the type of non-increasing finite sequences on $A$,
-provided with the lexicographic ordering on lists).
-
-For instance, let us take $N_A=\texttt{Omega}$, and take $\alpha=\langle 4,4,2,1,0\rangle$,
- $\beta=\langle 4,3,3,3,3,3,2\rangle$, and $\gamma=\langle 5\rangle$. Then $\beta<\alpha<\gamma$.
-
-In contrast the list $\langle5,6,3,3\rangle$ is not non-increasing (\emph{i.e.} sorted w.r.t. $\geq$), so it is not to be considered.
-
-Note that if the notation $N_A$ implements the ordinal 
-$\alpha$,  the new notation $\omega^{N_A}$ must implement the ordinal $\phi_0(\alpha)$, a.k.a. $\omega^\alpha$ (see chapter~\ref{chap:schutte})
-
-\end{project}
-
-\begin{remark}
- The set of ordinal terms in Cantor normal form (see Chap.~\ref{chap:T1}) and 
-in Veblen normal form (see 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}{Gamma0.Gamma0}) are shown to be ordinal notation systems, but there is a lot of work to be done in order to unify ad-hoc  definitions and proofs which were written before the definition of the \texttt{ON} type class.
-\end{remark}
-
-\section{Ordinal Notations and \gaia's ordinals}
-\gaiasign Module~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge} contains an instance of class \texttt{ON E0lt compare}, where \texttt{E0lt} is the order on the well-formed ordinal terms below $\epsilon_0$ defined in \gaiaHydras library (please see Chapter~\ref{gaia-chapter}).
-
-%------------------------------------------------------------------------
-
-\chapter[A proof of termination, using epsilon0]{A proof of termination, using ordinals below \texorpdfstring{$\epsilon_0$}{Epsilon0}}
-
-\label{cnf-math-def}
-\label{chap:T1}
-
-In this chapter, we adapt to \coq{} the well-known~\cite{KP82}  proof that Hercules eventually wins every battle, whichever the strategy  of each player.
-In other words, we present  a formal and self contained proof of termination  of all [free] hydra battles.
-First, we take from Manolios and Vroon~\cite{Manolios2005} a representation of the ordinal $\epsilon_0$ as terms in Cantor normal form.
-Then, we define a variant for hydra battles as a measure that maps any hydra to some ordinal strictly less than $\epsilon_0$.
-
-\section{The ordinal \texorpdfstring{\(\epsilon_0\)}{epsilon0}}
-\label{sec:epsilon0-intro}
-
-The ordinal \(\epsilon_0\) is the least ordinal number that satisfies 
-the equation \(\alpha = \omega^\alpha\), where \(\omega\) is 
-the least infinite ordinal.
-Thus, we can intuitively consider \(\epsilon_0\) as an
-\emph{infinite} \(\omega\)-tower.
-
-\subsection{Cantor normal form}
-\index{maths}{Cantor normal form}
-
-Any ordinal strictly less that \(\epsilon_0\) 
-can be finitely represented by a unique  \emph{Cantor normal form}, 
-that is, an expression  which is either  the ordinal \(0\) or 
-a sum  \(\omega^{\alpha_1} \times n_1 + \omega^{\alpha_2} \times n_2 + 
-  \dots + \omega^{\alpha_p} \times n_p\) where all the \(\alpha_i\) 
-are ordinals in Cantor  normal form, \(\alpha_1 > \alpha_2 > \alpha_p\), 
-and all the \(n_i\) are positive integers.
-
-An example of Cantor normal form is displayed in Fig \ref{fig:cnf-example}:
-Note that  any ordinal of
-the form \(\omega^0 \times i + 0\) is just written \(i\).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
-\node[color=blue]{$\omega^{(\omega^\omega\,+\, \omega^2 \times 8 \,+\, \omega)}+ \omega^\omega + \omega^4+ 6$};
-\end{tikzpicture}
-\caption{\label{fig:cnf-example}
-An ordinal in Cantor normal form}
-\end{figure}
-
-In the rest of this section, we define an inductive type for representing in \texttt{Coq}
-all the ordinals strictly  less than  \(\epsilon_0\), then extend some arithmetic operations
-to this type, and finally prove that our representation fits well with 
-the expected mathematical properties: the order we define is a well order, 
-and the decomposition into Cantor normal form  is consistent 
-with the implementation of the arithmetic operations of exponentiation of base \(\omega\) 
-and addition.
-
-\paragraph*{Remark}
-\label{sec:orgheadline65}
-Unless explicitly mentioned, the term ``ordinal" will be used instead of
-``ordinal strictly less than \(\epsilon_0\)" (except in Chapter~\ref{chap:schutte} where it stands for ``countable ordinal'').
-
-\subsection{A data type for  ordinals in Cantor normal form}
-\label{sec:orgheadline72}
-\label{sec:T1-inductive-def}
-
-% Our user contribution~\cite{CantorContrib} represents 
-% the set of ordinals strictly less than $\epsilon_0$ in Cantor normal form as in~\cite{Manolios2005}, and also the set
-% of ordinals strictly  less than $\Gamma_0$ in Veblen normal form.
-
-    Let us define an inductive type whose 
-constructors are respectively associated
-with the ways to build Cantor normal forms:
-
-\begin{itemize}
-\item the ordinal \(0\)
-\item the construction \((\alpha,\, n,\,\beta)  \mapsto \omega^\alpha \times (n + 1)+ \beta \quad (n\in\mathbb{N})\)
-\end{itemize}
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#T1}{Epsilon0.T1}}
-
-\label{types:T1}
-%\index{Constants!zero:T1}
-\index{hydras}{Library Epsilon0!Types!T1}
-
-\input{movies/snippets/T1/T1Def}
-
-\subsubsection{Example}
-
-\label{alpha0-def}
-For instance, the ordinal  $\omega^\omega+\omega^3\times 5+2$ is represented by the following term of type \texttt{T1}:
-
-\input{movies/snippets/T1/alpha0}
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5, level 1/.style={sibling distance=6cm},
-level 2/.style={sibling distance=35mm},  
-level 3/.style={sibling distance=17mm}]
-\node  {cons}
-  child {  node {cons}
-            child { node {cons} child {node {zero}} child {node{0}} child{node{zero}}}
-         child {node {0}}
-         child {node {zero}}}
-    child {node {0}}
-   child {node {cons} 
- child { node {cons} child {node {zero}} child {node{2}} child{node{zero}}}
-  child {node {4}}
-         child {node {cons} child {node {zero}} child {node{1}} child{node{zero}}}};
-
-\end{tikzpicture}
-
-\caption{The tree-like representation of the ordinal $\omega^\omega+\omega^3\times 5 +2$\label{fig:cnf-tree}}
-
-\end{figure}
-
-\paragraph{Remark}
-For simplicity's sake, we chose to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the construction (\texttt{cons $\alpha$ $n$ $\beta$}) is intended to represent the
-ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
-In a future version, we should replace  the type \texttt{nat} with \texttt{positive} in \texttt{T1}'s 
-definition. But this replacement would take a lot of time \dots{}
-
-\paragraph{Remark}
-The name \texttt{T1} we gave to this data-type  is proper to this development and refers
-to a hierarchy of ordinal notations.
-For instance, in Library \href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2},  the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{Gaia, schutte}).
-\noindent
-
-\input{movies/snippets/T2/T2Def}
-
-\subsection{About Gaia}
-\gaiasign Chapter~\vref{gaia-chapter} describes the present state of a
-project of making compatible the libraries \HydrasLib and \gaia. At present, both libraries contain their own version of the inductive type \texttt{T1}, with the same base name for the constructors.
-
-\inputsnippets{nfwfgaia/T1def}
-
-Thus, many examples of this and following chapters can be run
-in \gaia's context. We will signal any difference (notation, name) which may appear.
-A closer integration (same data type and functions) is still in project.
-
-\begin{remark}
-  \label{remark:hydra-gaia-renaming}
-  \textcolor{lookcolor}{In future releases, we plan to make some
-    \HydrasLib identifiers progressively  deprecated, in favour of \gaia's names.}
-\end{remark}
-
-\subsection{Abbreviations}
-
-Some abbreviations may help to write more concisely complex ordinal terms.
-
-\subsubsection{Finite ordinals}
-\label{sec:orgheadline67}
-
-For representing finite ordinals, \emph{i.e.} natural numbers, we first introduce a notation for terms of the form $n+1$, then define a coercion from type \texttt{nat} into \texttt{T1}.
-\label{sect:notation-FS}
-
-\label{sect:notation-F}
-
-\input{movies/snippets/T1/finiteOrds}
-
-\paragraph*{\gaiasign}
-  In \texttt{gaia.ssete9}, the $n$-th finite ordinal is also written
-  \texttt{$\backslash${F}\,$n$}. There is no coercion in \gaia from \texttt{nat} to \texttt{T1}.
-
-  \vspace{4pt}
-
-  From~\href{../theories/html/gaia_hydras.HydraGaia_Examples.html}{gaia\_hydras.HydraGaia\_Examples}.
-
-  \inputsnippets{HydraGaia_Examples/Demoa}
-
-% \index{Coq!Coercions}
-% \index{Functions!Coercions@Coercions (from nat to ordinal types)}
-% \begin{remark}
-% Please refer to the remark~\pageref{warning:coercions} about the use of coercions.
-% % The use of coercions like \texttt{fin} allow us to be close to the mathematical tradition where natural numbers are ordinals too.
-% % Nevertheless, it may happen that a goal like \texttt{3 < 5} could be 
-% % interpreted as \texttt{(lt (fin 3) (fin 5))},  depending on the current notation scope.  
-% % When this misinterpretation happens, tactics like \texttt{auto with arith}, \texttt{lia} do not work!
-% % Thus, it is useful to write \texttt{(3 < 5)\%nat}  an inequality between two natural numbers. 
-% \end{remark}
-
-\subsubsection{The ordinal \(\omega\)}
-\label{sec:orgheadline68}
-
-  Since \(\omega\)'s Cantor normal form is
-i.e. \(\omega^{\omega^0}\times 1+ 0\), we can define the following abbreviation:
-
-\label{sect:omega-notation2}
-
-\input{movies/snippets/T1/omegaDef}
-
-Note that \texttt{omega} is not an identifier in \HydrasLib, thus any tactic like \texttt{unfold omega} would fail.
-
-\paragraph*{\gaiasign}
-In \texttt{gaia.ssete9}, the ordinal $\omega$ is bound to the \emph{constant}  \texttt{T1omega}.
-
-\subsubsection{The ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
-\label{sect:notation-phi0}
-We provide also a notation for ordinals of the form $\omega^\alpha$.
-
-\index{hydras}{Library Epsilon0!Notations!phi0@phi0 (exponential of base omega)}
-
-\input{movies/snippets/T1/phi0Def}
-
-\index{maths}{Additive principal ordinals}
-
-\begin{remark}
-\label{sec:orgheadline69}
-The name \(\phi_0\)
-   comes from ordinal numbers theory. In~\cite{schutte}, Schütte defines 
-$\phi_0$  as the ordering (\emph{i.e.} enumerating) function of the set  of \emph{additive principal ordinals} \emph{i.e.} strictly positive ordinals $\alpha$ that verify $\forall \beta<\alpha, \beta+\alpha=\alpha$.
-For Schütte,  $\omega^\alpha$ is just a notation for $\phi_0(\alpha)$.
-See also Chapter~\ref{chap:schutte} of this document.
-\end{remark}
-
-\paragraph*{\gaiasign}
-In \texttt{gaia.ssete9} the identifier \texttt{phi0} is bound to a plain constant (not a notation).
-
-\subsubsection{The hierarchy of \(\omega\)-towers:}
-\label{sec:orgheadline71}
-
-The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximated by the sequence of $\omega$-towers (see also Sect~\vref{sect:epsilon0-as-limit} ).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/towerDef}
-
-For instance, Figure~\ref{fig:tower7} represents  the ordinal returned by the
- evaluation of the term \texttt{omega\_tower 7}.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
-\node[color=blue]{$\omega^{{{\omega}^{{{\omega}}^{{{\omega}}^{{\omega^{{\omega}^{\omega}}}}}}}}$};
-\end{tikzpicture}
-\caption{\label{fig:tower7}
-The $\omega$-tower of height 7}
-\end{figure}
-
-\subsection{Pretty-printing ordinals in Cantor normal form}
-\label{sect:ppT1}
-\index{hydras}{Library Epsilon0!Types!ppT1}
-
-Let us consider again the ordinal $\alpha_0$ defined in section~\vref{alpha0-def}. 
-If we ask \coq{} to print its  normal form, we get a hardly readable term of type \texttt{T1}.
-
-\input{movies/snippets/T1/alpha0Compute}
-
-The following data type defines a more readable abstract syntax for  ordinals terms in Cantor normal form:
-
-\label{types:ppT1}
-\index{hydras}{Library Epsilon0!Functions!pp@ pp (pretty printing terms in Cantor normal form)}
-
-\input{movies/snippets/T1/ppT1Def}
-
-The function \texttt{pp: T1 -> ppT1} converts any closed term of type \texttt{T1} into a human-readable expression. For instance, let us convert the term \texttt{alpha\_0}.
-
-\input{movies/snippets/T1/ppAlpha0}
-
-\paragraph*{\gaiasign}
-In ~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge},
-we define a variant of \texttt{pp} for pretty-printing terms of type \texttt{ssete9.CantorOrdinal.T1} (see Sect.~\vref{sect:gaia-ppT1}).
-
-\index{hydras}{Projects}
-\begin{project}
-Design  (in \ocaml?) a set of tools for systematically pretty printing ordinal terms in Cantor normal form.
-\end{project}
-
-\subsection{Comparison between ordinal terms}
-\label{sec:orgheadline73}
-
-% Our formalisation of Cantor Normal Form will take two steps:
-% 1 Definition of a strict order \texttt{o<} on the type \texttt{T1}, 
-% 2 Using \texttt{o<} for characterizing terms in normal form.
-
-In order to compare two terms of type \texttt{T1}, we define a recursive function \texttt{compare} that maps two ordinal tems $\alpha$ and $\beta$ to a value of type \texttt{comparison}.
-This type is defined in \coq's standard library
-\texttt{Init.Datatypes} and
-contains three constructors:  \texttt{Lt} (less than), \texttt{Eq} (equal), and
-\texttt{Gt} (greater than).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/compareDef}
-
-\label{Predicates:lt-T1}
-Please note that this definition of \texttt{lt} makes it easy to write proofs by computation, as shown by the following examples.
-
-\vspace{4pt}
-
-\input{movies/snippets/T1/ltExamples}
-
-Links between the function \texttt{compare} and the relations
-\texttt{lt} and \texttt{eq} are established through the following lemmas (~\vref{sect:comparable-def}).
-\vspace{4pt}
-
-\input{movies/snippets/T1/Instances}
-
-\paragraph*{\gaiasign}
-In \gaia, the strict order \texttt{T1lt} on ordinal terms is directly defined as a boolean function (in \ssreflect's style).
-
-\inputsnippets{nfwfgaia/T1ltDef}
-
-In section~\vref{sect:lt-compat-gaia}, we show that both definitions are mutually equivalent.
-
-\subsubsection{A Predicate for Characterizing Normal Forms}
-\label{sect:t1-nf}
-
-\label{sec:orgheadline74}
-\label{sec:orgheadline75}
-Our data-type \texttt{T1} allows us to write expressions that
-are not properly in Cantor normal form as specified in Section \ref{sec:epsilon0-intro}.
-For instance, consider the following term of type  \texttt{T1}.
-
-\input{movies/snippets/T1/badTerm}
-
-This term would have been written \(\omega^1\times 2 + \omega^\omega \times 3\) in the usual mathematical notation.
-We note that the exponents of $\omega$ are not in the right (strictly decreasing) order.
-Nevertheless, with the help of the order \texttt{lt} on \texttt{T1}, we are now able to characterize
-the set of all well-formed ordinal terms:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#nf_b}{Epsilon0.T1}}
-
-\label{Predicates:nf-T1}
-
-\input{movies/snippets/T1/nfDef}
-
-\input{movies/snippets/T1/nfAlpha0}
-
-\input{movies/snippets/T1/nfBadTerm}
-
-\paragraph*{\gaiasign}
-In \gaia, the boolean function which characterizes ordinal terms in normal form is defined as follows:
-
-\inputsnippets{nfwfgaia/T1nfDef}
-
-In Sect.~\vref{nf-gaia-compat}, we show that \gaia's \texttt{T1nf} is extensionally equivalent with \HydrasLib'  \texttt{nf}.
-
-\subsection{Making normality implicit}
-  We would like to get rid of terms of type \texttt{T1} which are not in Cantor normal form.
-A simple way to do this is to consider statements of the form 
-\texttt{forall alpha: T1, nf alpha -> $P$ alpha}, where $P$ is a predicate over type \texttt{T1}, like in the following lemma \footnote{Ordinal addition is formally defined a little later (page~\ref{sect:infix-plus-T1})}.
-
-\input{movies/snippets/T1/plusIsZero}
-
-\vspace{4pt}
-
-But this style leads to clumsy statements, and generates too many subgoals in interactive proofs (although often solved with \texttt{auto} or \texttt{eauto}).
-
-One may encapsulate conditions of the form \texttt{(nf $\alpha$)} in
-the most used predicates. For instance, we introduce the restriction of \texttt{lt} to terms in normal form, and provide a handy notation for this restriction.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Prelude.Restriction.html}{hydras.Prelude.Restriction}}
-
-\input{movies/snippets/Restriction/restrictionDef}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#LT}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/LTDef}
-
-\label{Predicates:LT-T1}
-
-For instance, in the following lemma, the condition that $\alpha$ is in normal form is included in the condition $\alpha< 1$.
-
-\input{movies/snippets/T1/LTOne}
-
-\subsubsection{\texttt{E0}: a sigma-type for \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-As we noticed in Sect.~\ref{sect:t1-nf}, the type \texttt{T1} is not a correct ordinal notation, since it contains terms that are not in Cantor normal form.
-In certain contexts (for instance in Sections~\ref{sect:L-equations}, \ref{sect:hardy},
-and \ref{sect:wainer}),  we need to define total recursive functions on well-formed ordinal terms less  than $\epsilon_0$, using the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777}.
- In order to define a type whose inhabitants represent just ordinals, we build a type gathering a term of type \texttt{T1} and a proof that this term is in normal form.
-
-\label{sect:E0-def}
-\label{types:E0}
-\index{hydras}{Library Epsilon0!Types!E0}
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html\#E0}{Epsilon0.E0}}
-
-\input{movies/snippets/E0/E0Def}
-
-Many constructs : types, predicates, functions, notations, etc., on type \texttt{T1} are adapted to \texttt{E0}.
-
-First, we declare a notation scope for \texttt{E0}, then we redefine the predicates of comparison.
-
-\input{movies/snippets/E0/E0Scope}
-\label{Predicates:Lt-E0}
-\input{movies/snippets/E0/LtLeDef}
-
-Equality in \texttt{E0} is just Leibniz equality. Note that, since \texttt{nf} is
-defined by a Boolean function, for  any term $\alpha:\texttt{T1}$, there exists at most one proof of \texttt{nf $\alpha$}, thus two ordinals of type \texttt{E0} are
-equal if and only if their projection to \texttt{T1} are equal (see also Sect.~\vref{sect:eq-proof-unicity}).
-
-\vspace{4pt}
-
-\index{coq}{Unicity of equality proofs}
-
-\inputsnippets{E0/nfProofUnicity}
-\inputsnippets{E0/E0EqIff}
-
-In order to  upgrade constants and functions from type \texttt{T1} to \texttt{E0}, we have to prove that
-the term they build is in normal form.
-For instance, let us represent the ordinals $0$ and $\omega$   as instances of the class \texttt{E0}.
-
-\vspace{4pt}
-
-\label{sect:omega-T1}
-
-\input{movies/snippets/E0/ZeroOmega}
-
-%\index{Constants!zero:T1}
-
-\paragraph*{\gaiasign}
- The library~\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines a \gaia-compatible type \texttt{E0}.
-
-\subsection{Syntactic definition of limit and successor ordinals}
-
-Pattern matching and structural recursion allow us to define boolean characterizations  of successor and limit ordinals.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/succbLimitb}
-
-The correctness of these definitions with respect to the mathematical notions of
-limit and successor ordinals is established through several lemmas. For instance,
-Lemma \texttt{canonS\_limit\_lub}, page~\pageref{lemma:canonS-limit}, shows that
-if $\alpha$ is (syntactically) a limit ordinal, then it is the least upper bound of
-a strictly increasing sequence of ordinals.
-
-\paragraph*{\gaiasign}
-In \gaia, the boolean functions associated with limit and successor ordinals are called \texttt{T1limit} and \texttt{T1is\_succ}.
-
- From~\href{../theories/html/gaia_hydras.HydraGaia_Examples.html}{gaia\_hydras.HydraGaia\_Examples}.
-
-\inputsnippets{HydraGaia_Examples/Demob}
-
-\subsection{Arithmetic on \texorpdfstring{$\epsilon_0$}{epsilon0}}
-\subsubsection{Successor}
-
-\index{hydras}{Library Epsilon0!Functions!succ}
-
-The successor of any ordinal $\alpha< \epsilon_0$ is defined by structural 
-recursion on its Cantor normal form.
-
-\label{Functions:succ-T1}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/succDef}
-
-The following lemma establishes the connection between the  function
-\texttt{succ} and the Boolean predicate \texttt{succb}.
-
-\vspace{4pt}
-
-\input{movies/snippets/T1/succbIff}
-
-\begin{exercise}[\gaiasign]
-  Look for the \gaia-theorem which corresponds to \texttt{succb\_iff}.
-\end{exercise}
-
-\subsubsection{Successor function on \texttt{E0}}
-
-The function \texttt{succ} on \texttt{T1} is extended to \texttt{E0} the following way:
-
- \emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
-
- \inputsnippets{E0/SuccOnE0}
-
-\index{hydras}{Exercises}
- \begin{exercise}
-Prove in \coq{} that for any ordinal $0< \alpha<\epsilon_0$, $\alpha$ is a limit if 
-and only if for all $\beta<\alpha$, the interval $[\beta,\alpha)$ is
-infinite.
-
-\emph{You may start this exercise with the file
-     \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/Limit_Infinity.v}{exercises/ordinals/Limit\_Infinity.v}.}
- \end{exercise}
-
-\subsubsection{Addition and multiplication}
-
-Ordinal addition and multiplication are also defined by structural recursion over the type \texttt{T1}. Please note that they use the \texttt{compare} function on some subterms of their arguments.
-
-\label{sect:infix-plus-T1}
-
-\input{movies/snippets/T1/plusDef}
-
-\input{movies/snippets/T1/multDef}
-
-\paragraph*{\gaiasign}
-We keep \gaia's base names for
-addition and multiplication of ordinal terms below $\epsilon_0$.
-Please refer to Sect.~\ref{sect:plus-mult-gaia-hydras} about compatibility of both arithmetics.
-
-\subsubsection{Examples}
-
-The following examples are instances of \emph{proofs by computation}. Please note that  addition and multiplication on \texttt{T1}
-are not commutative. Moreover,  both operations fail to be strictly monotonous in their first argument.
-
-\input{movies/snippets/T1/plusMultExamples}
-
-\input{movies/snippets/T1/notMono}
-
-The function \texttt{succ} is related with addition through the following lemma:
-
-\inputsnippets{T1/succIsPlusOne, T1/succIsPlusOnez}
-
-\subsubsection{Arithmetic on type \texttt{E0}}
-
- We define an addition in type \texttt{E0}, since the sum of two terms in normal form is in normal form too.
-
-\input{movies/snippets/T1/plusNf}
-
-\input{movies/snippets/E0/plusE0}
-\input{movies/snippets/E0/CheckPlus}
-
-\begin{remark}
-In all this development, two representations of ordinals co-exist: ordinal terms (type \texttt{T1}, notation scope \texttt{t1\_scope}, for reasoning on the tree-structure of Cantor normal forms), and ordinal terms \emph{known to be in normal form} (type \texttt{E0}, notation scope \texttt{E0\_scope}).
-Looking at the contexts displayed by \coq{} prevents you from any risk of confusion.
-\end{remark}
-
-%% To simplify !
-\index{hydras}{Exercises} 
-\begin{exercise}
-Prove that for any ordinal $\alpha:\texttt{E0}$, 
-$\omega\leq \alpha$ if and only if, for any natural number $i$,
-$i+\alpha=\alpha$.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/ge_omega_iff.v}{exercises/ordinals/ge\_omega\_iff.v}.}
-\end{exercise}
-
-\subsection{A proof by computation}
-\label{Ex42-E0}
-
-It is interesting to compare the following proof of the equality
-$\omega+42+\omega^2$ with the more theoretical proof in Sect~\vref{Ex42-schutte}.
-
-\inputsnippets{E0/Ex42}
-
-\section{Well-foundedness and transfinite induction}
-
-\index{maths}{Transfinite induction}
-
-\subsection{About  well-foundedness}
-\label{sec:T1Wf}
-\label{sec:orgheadline82}
-   In order to use \texttt{T1} for proving termination results,
-we need to prove that  our order \texttt{<} is well-founded. Then we will get \emph{transfinite induction} for free.
-
-The proof of well-foundedness of the strict order $<$ on Cantor normal forms is already 
-available in the Cantor contribution by Castéran and Contejean~\cite{CantorContrib}. That proof relies on a library on recursive path orderings written by
-E. Contejean. We present here  a direct proof of the same result, which does not require any knowledge on r.p.o.s.
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that the \emph{total} order \texttt{lt} on \texttt{T1} is not well-founded. 
-\textbf{Hint:}  You will have to build a counter-example with terms of type \texttt{T1}
-which are not in Cantor normal form.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/T1_ltNotWf.v}{exercises/ordinals/T1\_ltNotWf.v}.}
-\end{exercise}
-
-\subsubsection{A first attempt}
-\label{sec:orgheadline77}
-%\index{coq}{Well-founded induction}
-
-It is natural to try to prove by structural induction over \texttt{T1} 
-that every term in normal form is \texttt{LT}-accessible.
-
-Unfortunately, it won't work. Let us consider some well-formed term
- $\alpha=\texttt{cons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are \texttt{LT}-accessible.
-In order to prove the accessibility of $\alpha$, we have to consider
-any well formed term \(\delta\) such that \(\delta<\alpha\). 
-But nothing guarantees that \(\delta\)  is strictly  less than \(\beta\) nor \(\gamma\), and we cannot use the induction hypotheses on   \(\beta\) nor \(\gamma\).
-
-\input{movies/snippets/T1/wfLTBada}
-
-The problem comes from the too weak hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
-\(\gamma\);
-for instance \(\delta\) may be of the form
-\texttt{cons $\beta$ $p'$  $\gamma'$},
-where    \(p' < n\).
-Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.
-
-\input{movies/snippets/T1/wfLTBadz}
-
-\subsubsection{Using a stronger inductive predicate.}
-\label{sec:orgheadline78}
-\label{sec:strongly-accessible}
-
-  Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
-through \texttt{LT}, we propose to consider the following (stronger) predicate:
-
-\input{movies/snippets/T1/AccStrongDef}
-
-The following lemma is an application of the strict inequality 
-$\alpha < \omega ^\alpha$. If $\alpha$ is strongly accessible, then, by definition,
-$\omega^\alpha$ is accessible, thus $\alpha$ is \emph{a fortiori} accessible.
-
-\input{movies/snippets/T1/AccStrongStronger}
-
-Thus, it remains to prove that every ordinal strictly less than $\epsilon_0$
-is strongly accessible.
-
-% \subsubsection{Structure of the proof of well-foundedness of \texttt{LT}}
-
-\label{sec:orgheadline81}
-\label{proof-wf-epsilon0}
-\paragraph{A helper}
-\label{sec:orgheadline79}
-
-First, we prove that, for  any \texttt{LT}-accessible term $\alpha$, $\alpha$ is 
-strongly accessible too (\emph{i.e.} any well formed
-term (\texttt{cons $\alpha$ $n$ $\beta$})  is accessible).
-
-The proof is structured as an induction on $\alpha'$s accessibility. Let us consider
-any  accessible term $\alpha$.
-
-\input{movies/snippets/T1/AccImpAccStrong}
-
-Let \texttt{n:nat} and \texttt{beta:T1} such that (\texttt{cons alpha n beta}) is in normal form. 
-We prove first that \texttt{beta} is accessible,  which allows us to prove by well-founded induction on \texttt{beta}, 
-and natural induction on \texttt{n}, that (\texttt{cons alpha n beta}) is accessible.
-The proof, quite long, can be consulted in \href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}.
-
-\paragraph{Accessibility of any well-formed ordinal term}
-\label{sec:orgheadline80}
-
-Our goal is still to prove accessibility of any well formed ordinal term.
-Thanks to our previous lemmas, we are almost done (by a simple structural induction!).
-
-\input{movies/snippets/T1/nfAcc}
-
-\input{movies/snippets/T1/T1Wf}
-
-\index{maths}{Transfinite induction}
-
-Traditionnally, well-founded induction on ordinals is called \emph{transfinite induction}.
-
-\inputsnippets{T1/transfiniteRecursor, T1/transfiniteInduction}
-
-As a corollary, the order \texttt{Lt} on type \texttt{E0} is well-founded too.
-
-\inputsnippets{E0/E0LtWf}
-
-\begin{remark}
-\label{remark:a3pat}
-The alternate proof of well-foundedness using \'Evelyne Contejean's work on
-    recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the
-    library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0.Epsilon0rpo}.
- \end{remark}
-
-\subsection{An ordinal notation for  \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-We build an instance of \texttt{ON}, and prove its correctness w.r.t. Schutte's model (see Chapter~\ref{chap:schutte}).
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
-
-\input{movies/snippets/E0/InstanceEpsilon0}
-
-\label{instance-epsilon0}
-
-\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
-
-\input{movies/snippets/Correctness_E0/injectDef}
-
-\input{movies/snippets/Correctness_E0/Epsilon0Correct}
-
-\index{hydras}{Projects}
-\begin{project}
- \emph{This exercise is a continuation of Project~\vref{exo:ON-mult}.}
-Use \texttt{ON\_mult} to define an ordinal notation \texttt{Omega2} for $\omega^2=\omega\times\omega$.
-
-Prove that \texttt{Omega2} is a sub-notation of \texttt{Epsilon0}.
-
-Define on \texttt{Omega2} an addition compatible with the addition on \texttt{Epsilon0}.
-
-\textbf{Hint}. You may use the following definition (in 
-    \href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}).
-
-\input{movies/snippets/ON_Generic/SubONSameOp}
-
-     \end{project}
-
-    \index{hydras}{Projects}
-    \begin{project}
-    The class \texttt{ON} of ordinal notations has been defined long after this 
-    chapter, and is not used in the development of the type \texttt{E0} yet.
-    A better integration of both notions should simplify the development on ordinals in Cantor normal form. This integration is planned for the future versions.
-
-    \end{project}
-
-    \section{A refinement of \texttt{E0} : an ordinal notation for \texorpdfstring{$\omega^\omega$}{omega\^omega}}
-
-    In Module   \href{https://github.com/coq-community/hydra-battles/blob/master/theories/
-      ordinals/OrdinalNotations/OmegaOmega.v}{theories/ordinals/OrdinalNotations/OmegaOmega.v},
-    we represent ordinals below $\omega^\omega$ by lists of pairs of natural numbers (with the same coefficient shift as in \texttt{T1}).
-    For instance, the ordinal $\omega^4\times 10 + \omega^3 + \omega+ 5$ is represented by the list \texttt{(4,9)::(3,0)::(1,0)::(0,4)::nil}.
-
-\inputsnippets{OmegaOmega/LODef}
-
-  The usual operations : \texttt{succ}, \texttt{+}, \texttt{*} are simple variants of the same operations in \texttt{T1}.
-
-  \inputsnippets{OmegaOmega/succPlusMult}
-
-    We establish this representation as a \emph{refinement} of the data types we used to represent ordinals less than $\epsilon_0$.
-Thus, many properties like well-foundedness of $<$ and associativity of $+$,  of this ordinal notations have very short proofs.
-
-    \inputsnippets{OmegaOmega/refineDef, OmegaOmega/phi0Ref,
-      OmegaOmega/succRef, OmegaOmega/plusRef,
-      OmegaOmega/multRef}
-
-    In order to make an ordinal notation for $\omega^\omega$, we follow the same steps as for $\epsilon_0$:
-    \begin{enumerate}
-    \item Define an order \texttt{lt}, which refines the order
-      \texttt{lt} on \texttt{T1}.
-      \inputsnippets{OmegaOmega/compareDef}
-      \inputsnippets{ OmegaOmega/ltDef, OmegaOmega/ltRef}
-
-    \item Define the predicate ``to be in normal form''.
-      \inputsnippets{OmegaOmega/nfDef}
-    \item Define a class \texttt{OO} of terms in normal form, and an embedding from \texttt{E0} into \texttt{OO}.
-      \inputsnippets{OmegaOmega/OODef}
-       \inputsnippets{OmegaOmega/embedDef}
-      \item Infer well-foundedness of the order on \texttt{OO}.
-    \inputsnippets{OmegaOmega/ltWf}
-
-    \end{enumerate}
-
-    \inputsnippets{OmegaOmega/ONOO, OmegaOmega/OOz}
-
-    Let us show a few examples.
-
-    \inputsnippets{OmegaOmega/OODemo}
-
-    \index{hydras}{Exercises}
-
-    \subsection{Related work}
-    The article~\cite{Manoury2016}  defines another representation of ordinals below $\omega^\omega$ based on lists of natural numbers.
-
-\begin{exercise}[not trivial]
-  It may be interesting to write a \emph{direct} proof  of well-foundedness of the order in $\omega^\omega$ (\emph{i.e.}
-  without using properties of $\epsilon_0$). This exercise may
-  help to understand better the proof structure of Sect.~\vref{sec:strongly-accessible}.
-    \end{exercise}
-%ici
-    \section{A variant for hydra battles}
-
-    In order to prove the termination of any hydra battle, we try to define a variant mapping hydras to ordinals strictly less than $\epsilon_0$.
-    In order to make such a variant easy to define (for instance by a structural recursion), we introduce a variant of addition, which, contrary to
-    $+$, is commutative and strictly monotonous in both of its arguments. This last property makes it possible to prove that our function is 
-    truly a variant for hydra battles (in Sect.~\vref{sect:variant-decr}).
-
-    \subsection{Natural sum (a.k.a. Hessenberg's  sum)}
-    \label{sec:orgheadline87}
-    \label{hydra-variant}
-    \label{sect:hessenberg-def}
-
-    Natural sum (Hessenberg's  sum) is a commutative and monotonous version of
-    addition. It is used as an auxiliary operation  for defining variants
-    for hydra battles, where Hercules is allowed to chop off any  head of the hydra.
-
-    In the literature, the natural sum of ordinals \(\alpha\) and \(\beta\)
-    is often denoted by \(\alpha \# \beta\)  or  \(\alpha \oplus  \beta\).
-    Thus we called \texttt{oplus} the associated \emph{Coq} function.
-
-    \subsubsection{Definition of \texttt{oplus}}
-    \label{sec:orgheadline84}
-    %\index{Functions!oplus @ oplus (Hessenberg commutative sum)}
-
-    The definition of \texttt{oplus} is recursive in both of its 
-    arguments and uses  the same pattern as for the \texttt{merge} function on lists of library
-    \texttt{Coq.Sorting.Mergesort}.
-
-    \begin{enumerate}
-    \item Define a nested recursive function, using the \texttt{Fix} 
-        construct
-
-    \item Build a principle of induction dedicated to \texttt{oplus}
-
-    \item Establish equations associated to each case of the definition.
-    \end{enumerate}
-
-    \paragraph{Nested recursive definition}
-    \label{sec:orgheadline83}
-
-    The following definition is composed of 
-    \begin{itemize}
-    \item A main function \texttt{oplus}, structurally recursive in its 
-    first argument \texttt{alpha}
-    \item An auxiliary function \texttt{oplus\_aux} within the scope of \texttt{alpha},
-    structurally recursive in its argument \texttt{beta};  \texttt{oplus\_aux beta} 
-       is supposed to compute  \texttt{oplus alpha beta}.
-    \end{itemize}
-
-%  \pagebreak 
-  \vspace{4pt}
-    \emph{From Module~\href{../theories/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}
-
-    \label{sect:infix-oplus}
-
-    \input{movies/snippets/Hessenberg/oplusDef}
-
-    The reader will note that each recursive call of the functions
-    \texttt{oplus} and \texttt{oplus\_aux} satisfies \emph{Coq}'s constraint
-    on recursive definitions. The function \texttt{oplus} is recursively called on a sub-term of its first argument,
-    and \texttt{oplus\_aux} on a sub-term of its unique argument.
-    Thus, \texttt{oplus}'s definition is accepted by \coq{} as a structurally recursive function.
-
-    \subsubsection{Rewriting lemmas}
-    \label{sec:orgheadline86}
-
-    \emph{Coq}'s constraints on recursive definitions result in 
-    the quite  complex form of \texttt{oplus}'s definition.
-    Proofs of properties of this function can be simpler if we
-     derive a few  rewriting lemmas that will help to simplify 
-    expressions of the form (\texttt{oplus $\alpha$ $\beta$}).
-
-    A first set of lemmas correspond to the various cases of \texttt{oplus}'s 
-    definition. They can be proved almost immediately. Here are a few examples.
-
-    \inputsnippets{Hessenberg/oplusNeutral,
-      Hessenberg/oplusCompareLt}
-
-    % \subsubsection{A hand-made induction principle}
-    % \label{sec:orgheadline85}
-
-    % \index{Coq!Commands!Functional Scheme}
-
-    % \emph{Coq} contains a command  \texttt{Functional Scheme} that 
-    % generates induction principles which correspond to recursive functions.
-    % Unfortunately, the current version ( \texttt{8.11.0} ) doesn't work on \texttt{oplus},
-    % probably because of the inner \texttt{Fix}.
-
-    % \begin{Coqsrc}
-    % Functional Scheme oplus_ind := Induction for oplus Sort Prop.
-    % \end{Coqsrc}
-
-    % \begin{Coqanswer}
-    % Error: Anomaly "todo." Please report at http://coq.inria.fr/bugs/.
-    % \end{Coqanswer}
-
-    % Fortunately, it's a good exercise for a semi-experienced user, to write
-    % her/him-self induction principles similar to the ones returned by
-    % \texttt{Functional Scheme}.
-
-    % \begin{itemize}
-    % \item First, we choose to write a version for sort \texttt{Type}, since versions
-    % for sorts \texttt{Prop} and \texttt{Set} can be easily derived from
-    % the former one. According to \emph{Coq}'s naming politics, we will call our 
-    % principle \texttt{oplus\_rect}
-
-    % \item The conclusion of \texttt{oplus\_rect} will be (\texttt{$P$ a b (oplus a b)}),
-    % where $P$ is an arbitrary function of type 
-    % \texttt{T1 -> T1 -> T1 -> Type}
-
-    % \item The premises of \texttt{oplus\_rect} will describe how to build an induction 
-    % on the graph of \texttt{oplus}.
-    % \end{itemize}
-
-    % We are now ready to state and prove \texttt{oplus\_rect}, and the reader
-    % will note that the statement is longer than the proof script itself,
-    % which is a standard proof by induction, simplification and case-analysis 
-    % that follows  \texttt{oplus}'s definition.
-
-    % We associate also a tactic to the application of \texttt{oplus\_rect}.
-
-    % \begin{Coqsrc}
-    %  Lemma oplus_rect:
-    %       forall P: T1 -> T1 -> T1 -> Type, 
-    %         (forall a:T1, P zero a a) ->
-    %         (forall a: T1, P a zero a) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Gt ->
-    %            P b1 (cons a2 n2 b2) o ->
-    %            P (cons a1 n1 b1) (cons a2 n2 b2)
-    %              (cons a1 n1 o)) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Lt ->
-    %            P (cons a1 n1 b1) b2 o ->
-    %            P (cons a1 n1 b1) (cons a2 n2 b2) 
-    %            (cons a2 n2 o)) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Eq ->
-    %            P b1 b2 o ->
-    %           P (cons a1 n1 b1) (cons a2 n2 b2)
-    %             (cons a1 (S (n1 + n2)%nat) o)) ->
-    %          forall a b, P a b (oplus a b).
-    % Proof with auto.
-    %    induction a.
-    %    -    intro; simpl; destruct b;auto.
-    %    -   induction b.
-    %        + apply X0.
-    %        + case_eq (compare a1 b1).
-    %          * intro Comp; unfold oplus; rewrite Comp.
-    %            cbn; apply X3 ...
-    %          * intro Comp; cbn; rewrite Comp; apply X2...
-    %          * intro Comp; cbn; rewrite Comp ...
-    %  Defined.
-
-    % Ltac oplus_induction a b:= pattern (oplus a b); apply oplus_rect.
-    % \end{Coqsrc}
-
-    % \index{Exercises}
-
-    % \begin{exercise}
-    % The induction principle \texttt{oplus\_rect} is still unused in our development. 
-    % Please build some nice examples of application.
-    % \end{exercise}
-
-    \index{hydras}{Projects}
-    \begin{project}
-    Compare \texttt{oplus}'s definition (with inner fixpoint) with other possibilities
-    (\texttt{coq-equations}, \texttt{Function}, etc.).
-    \end{project}
-    \subsection{More theorems on Hessenberg's sum}
-
-    We need to prove some properties of $\oplus$, particularly about 
-    its relation with the order $<$ on \texttt{T1}.
-
-    \subsubsection{Commutativity, associativity}
-
-    We prove  the commutativity of $\oplus$ in two steps. 
-    First, we prove by transfinite induction on $\alpha$ that the restriction of $\oplus$ to the
-    interval $[0..\alpha)$ is commutative.
-
-    \index{maths}{Transfinite induction}
-
-    \input{movies/snippets/Hessenberg/oplusComm0}
-
-    Then, we infer  $\oplus$'s commutativity for any pair of ordinals:
-    Let $\alpha$ and $\beta$ be two ordinals strictly less than $\epsilon_0$. Both ordinals $\alpha$ and $\beta$ are
-    strictly less than $\textrm{max}(\alpha,\beta)+1$.
-        Thus, we have just to apply the lemma \coqsimple{oplus\_comm\_0}.
-
-\input{movies/snippets/Hessenberg/oplusComm}
-
-    Associativity of Hessenberg's sum is proved the same way.
-
-    \input{movies/snippets/Hessenberg/oplusAssoc0}
-    \input{movies/snippets/Hessenberg/oplusAssoc}
-
-    \subsubsection{Monotony}
-
-    At last, we prove that $\oplus$ is strictly monotonous in both of its arguments.
-
-    \input{movies/snippets/Hessenberg/oplusMono}
-
-      \index{hydras}{Projects}
-
-    \begin{project}
-    The library \texttt{Hessenberg} looks too long (proof scripts and compilation).
-    Please try to make it simpler and more efficient!
-    Thanks!
-    \end{project}
-
-    \paragraph*{\gaiasign} The module
-      \href{../theories/html/gaia_hydras.~GHessenberg.html}{gaia\_hydras.GHessenberg} defines a version of Hessenberg's sum
-      compatible with \gaia's type \texttt{T1}.
-
-    \subsection{A termination measure for hydra battles }
-
-    \label{sec:hydra-measure}
-
-    Let us define a measure from type \texttt{Hydra} into \texttt{T1}.
-
-    \vspace{4pt}
-    \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Termination.html\#m}{Hydra.Hydra\_Termination}}
-
-   \input{movies/snippets/Hydra_Termination/mDef}
-
-    First, we prove that the measure $m(h)$  of any hydra $h$ is a well-formed ordinal term of type \texttt{T1}.
-
-    \input{movies/snippets/Hydra_Termination/mNf}
-      \input{movies/snippets/Hydra_Termination/msNf}
-
-    For proving the termination of all hydra battles, we have to prove that
-    \texttt{m} is a variant. First, a few technical lemmas follow the decomposition of \texttt{round} into several relations. Then the lemma \texttt{round\_decr} gathers all the cases.
-
-    \label{sect:variant-decr}
-
-    \input{movies/snippets/Hydra_Termination/S0Decr}
-    \input{movies/snippets/Hydra_Termination/R1Decr}
-    \input{movies/snippets/Hydra_Termination/S1Decr}
-    \input{movies/snippets/Hydra_Termination/R2Decr}
-     \input{movies/snippets/Hydra_Termination/RoundDecr}
-
-      Finally, we prove termination of all (free) battles.
-
-    \label{thm:every-battle-terminates}
-
-    \input{movies/snippets/Hydra_Termination/FinalThm}
-
-    \paragraph*{\gaiasign}
-    The module~\href{../theories/html/gaia_hydras.GHydra.html}{gaia\_hydras.GHydra} contains a proof of \gaiaHydras version of termination of all battles.
-
-    \inputsnippets{GHydra/mDef, GHydra/mVariant, GHydra/Termination}
-
-    \section*{Conclusion}
-
-    Let us recall three results we have proved so far.
-    \begin{itemize}
-    \item There exists a strictly decreasing variant which maps \texttt{Hydra} into 
-    the segment $[0,\epsilon_0)$ for proving the termination of any hydra battle
-    \item There exists \emph{no} such variant from \texttt{Hydra} into 
-    $[0,\omega^2)$, \emph{a fortiori} into $[0,\omega)$.
-    \end{itemize}
-
-    So, a  natural question is `` Does there exist any strictly decreasing variant mapping
-    type \texttt{Hydra} into some interval $[0,\alpha)$ (where $\alpha <\epsilon_0$) for proving the termination of all hydra battles''.
-The next chapter is dedicated to a formal proof that there exists no such $\alpha$, even if we consider a restriction to the set of ``standard'' battles.
-
-    %\include{epsilon0}
-
-    %\include{impossibility-proofs}
-
-%-------------------------------------------------------------------
-
-\chapter[The Ketonen-Solovay machinery]{Accessibility inside \texorpdfstring{$\epsilon_0$}{Epsilon0}: The Ketonen-Solovay Machinery\label{ks-chapter}}
-\label{chap:ketonen}
-\index{maths}{Ordinal numbers!Ketonen-Solovay machinery}
-
-\section{Introduction}
-The reader may think that our proof of termination in the previous  chapter requires a lot of mathematical tools and may be too  complex. So, the question is ``is there  any  simpler proof'' ?
-
-In their article~\cite{KP82}, Kirby and Paris show that this result cannot be proved in Peano arithmetic.
-Their proof uses some knowledge about model theory and non-standard models of Peano arithmetic.
-In this chapter, we focus on a specific class of proofs of termination of hydra battles: construction of some variant mapping the type \texttt{Hydra} into a given initial  segment of ordinals.
-Our proof relies only on the Calculus of Inductive Constructions and is a natural complement of the results proven in the previous chapters.
-
-\begin{itemize}
-\item There is no variant mapping the type \texttt{Hydra} into the interval $[0,\omega^2)$ (section ~\vref{omega2-case}), and \emph{a fortiori} $[0,\omega)$ (section ~\vref{omega-case}).
-
-\item There exists a variant which maps the type \texttt{Hydra} into the
-interval $[0,\epsilon_0)$ (theorem \texttt{every\_battle\_terminates}, in section~\vref{thm:every-battle-terminates}).
-\end{itemize}
-
-Thus, a very natural question is the following one:
-\begin{quote}
-  `` Is there  any variant from
-\texttt{Hydra} into some interval $[0,\mu)$, where $\mu<\epsilon_0$, for proving the termination of all hydra battles ?''
-\end{quote}
-
-We prove in \coq{} the following result:
-
-\begin{quote}
-There is no variant for proving the termination of all hydra battles
-from \texttt{Hydra} into the interval $[0..\mu)$, where
-$\mu< \epsilon_0$.
-The same impossibility holds even if we consider only standard battles (with the successive replication factors $0,1,2,\dots,t,t+1,\dots$).
-\end{quote}
-
-Our proofs are  constructive and require no axioms: they are  closed terms of the CIC, and are mainly composed on function definitions and proofs of properties of these functions. 
-They  share much theoretical material with Kirby and Paris', although they do not use any knowledge about Peano arithmetic nor model  theory.  The combinatorial arguments we use and implement
-come from 
- an article by J.~Ketonen and R.~Solovay~\cite{KS81}, already  cited in the work
- by L.~Kirby and J.~Paris. % on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
- Section $2$ of this article: ''A hierarchy of probably recursive functions'', contains a systematic study of \emph{canonical sequences}, which are closely related to
-rounds of hydra battles. 
-Nevertheless, they have the same global structure as the simple proofs described in
-sections~\vref{omega-case} and \vref{omega2-case}. 
-We invite the reader to compare the three proofs step by step, lemma by lemma.
-
-\section{Canonical Sequences}
-\label{ketonen-solovay-sect}
-\index{maths}{Ordinal numbers!Canonical sequences}
-
-Canonical sequences are functions that associate an ordinal $\canonseq{\alpha}{i}$ to every ordinal $\alpha<\epsilon_0$ and positive integer $i$. They satisfy several nice properties:
-
-\index{maths}{Transfinite induction}
-\begin{itemize}
-\item If $\alpha\not=0$, then $\canonseq{\alpha}{i}<\alpha$. Thus canonical sequences can be used for proofs by transfinite induction or function definition by transfinite recursion
-\item If $\lambda$ is a limit ordinal, then $\lambda$ is the least upper bound of the set 
-$\{\canonseq{\lambda}{i}\;|\,i\in\mathbb{N}_1\}$
-
-\item If $\beta<\alpha<\epsilon_0$, then there is a ``path'' from $\alpha$ to $\beta$, \emph{i.e.} a
-sequence $\alpha_0=\alpha, \alpha_1, \dots, \alpha_n=\beta$, where for every $k<n$, there exists some $i_k$ such that $\alpha_{k+1}=\canonseq{\alpha_k}{i_k}$
-\item Canonical sequences correspond tightly to rounds of hydra battles: if $\alpha\not=0$,
-then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1})$ in one round with
-the replication factor $i$ (Lemma \href{../theories/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
-\item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) battle from $\iota(\alpha)$ to $\iota(\beta)$.
-\end{itemize}
-
-\begin{remark}
-  In~\cite{KS81}, canonical sequences are defined for any ordinal $\alpha <\epsilon_0$,
-by stating that if $\alpha$ is a successor ordinal $\beta+1$,  the sequence associated with 
-$\alpha$ is simply the constant sequence whose terms are equal to $\beta$.
-Likewise, the canonical sequence of $0$ maps any natural number to $0$.
-
-This convention allows us to make total the function that maps any ordinal $\alpha$ and natural number $i$ to the ordinal $\canonseq{\alpha}{i}$.
-
-\end{remark}
-
-First, let us recall how canonical sequences are defined in~\cite{KS81}.
-For efficiency's sake, we decided not to implement directly K.\&S's definitions, but to define in \gallina{} simply typed structurally recursive functions which share the abstract properties which are used in the mathematical proofs\footnote{With a small difference: the $0$-th term of the canonical sequence is not the same in our development as in~\cite{KS81}.}.
-
-\subsubsection{Mathematical definition of canonical sequences}
-
-In~\cite{KS81} the definition of $\canonseq{\alpha}{i}$ is based on the following remark:
-\begin{quote}
-Any non-zero ordinal $\alpha$ can be decomposed in a unique way as the product
-$\omega^\beta\times (\gamma+1)$.
-\end{quote}
-
-Thus the $\canonseq{\alpha}{i}$\,s are defined in terms of this decomposition:
-\begin{definition}[Canonical sequences: mathematical definition]
-\label{def:canonseq-math}
-
-\end{definition}
-\begin{mathframe}
-  \begin{itemize}
-\item Let $\lambda<\epsilon_0$ be a limit ordinal
-
-\begin{itemize}
-\item If $\lambda=\omega^{\alpha+1}\times (\beta+1)$, then 
-$\canonseq{\lambda}{i}= \omega^{\alpha+1}\times\beta +  \omega^\alpha \times i$
-\item If $\lambda=\omega^{\gamma}\times (\beta+1)$, where $\gamma<\lambda$ is a limit ordinal, then 
-$\canonseq{\lambda}{i}=\omega^{\gamma}\times \beta + \omega^{\canonseq{\gamma}{i}}$
-\end{itemize}
-
-\item For successor ordinals, we have $\canonseq{\alpha+1}{i}= \alpha$
-
-\item Finally, $\canonseq{0}{i}= 0$.
-\end{itemize}
-\end{mathframe}
-
-\subsubsection{Canonical sequences in \coq}
-\index{hydras}{Library Epsilon0!Functions!canon}
-\index{hydras}{Library Epsilon0!Functions!canonS}
-
-Our definition may look more complex than the mathematical one, but
-uses plain structural recursion over the type \coqsimple{T1}. Thus, tactics like
-\coqsimple{cbn}, \coqsimple{simpl}, \coqsimple{compute}, etc., are applicable.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canon}{Epsilon0.Canon}}
-
-\label{Functions:canonS}
-\label{Functions:canon}
-
-\input{movies/snippets/Canon/canonDef}
-
-\paragraph*{\gaiasign} Our library~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon} transposes the  function \texttt{canon} into \gaia's type \texttt{T1}
-(please see Sect.\~ref{sect:gcanon}).
-
-\paragraph*{Remark}
-In the present state of this library, the following specializations of \texttt{canon} are still used in some proofs or lemma statements. They are planned to be deprecated.
-
-\input{movies/snippets/Canon/CanonS0}
-
-For instance \coq's computing facilities allow us to verify the equalities\linebreak 
-\mathcolor{$\canonseq{\omega^\omega}{3} = \omega^3$} and
-\mathcolor{$\canonseq{\omega^\omega*3}{42} = \omega^\omega*2 + \omega^{42}$}.
-
-\input{movies/snippets/Canon/canonExamples}
-
-% \index{hydras}{Projects}
-% \begin{project}
-% Many lemmas presented in this chapter were stated and proved before the introduction of 
-% the type class \texttt{ON} of ordinal notations, and in particular its  instance \texttt{Epsilon0}.
-% Thus definitions and lemmas refer to the type \texttt{T1} of possibly not well-formed terms.
-% This should be fixed in  a future version.
-% \end{project}
-
-\subsection{Basic properties of canonical sequences}
-
-We did not  try to prove that our definition truly implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
-abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
-~\href{../theories/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}.
-For instance, the equality $\canonseq{\alpha+1}{i}=\alpha$  can be  proved by  structural induction on $\alpha$.
-
-\input{movies/snippets/Canon/canonSucc}
-
-\subsubsection{Canonical sequences and the order $<$}
-
-\index{maths}{Transfinite induction}
-
-We prove by transfinite induction over $\alpha$ that $\canonseq{\alpha}{i}$ is an ordinal strictly less than $\alpha$ (assuming $\alpha\not=0$).
-This property allows us to use the function \texttt{canonS} and its derivatives in function definitions by transfinite recursion.
-
-\label{lemma:canon_LT}
-
-\inputsnippets{Canon/canonLT}
-
-\paragraph*{\gaiasign} This lemma is also available in
-Library ~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon}:
-
-\inputsnippets{GCanon/gcanonLt}
-
-\subsubsection{Limit ordinals are truly limits}
-The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
-is the limit of the sequence $\canonseq{\lambda}{i}\;(1\le i)$.
-
-\vspace{4pt}
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}
-
-\input{movies/snippets/Canon/canonSLimitStrong}
-
-\label{lemma:canonS-limit}
-
-Note the use of \coq's \texttt{sig} type in the theorem's statement, which
-relates the boolean function \texttt{limitb} defined on the \texttt{T1} data-type with a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.
-We can also state directly that $\lambda$ is a (strict) least upper bound of the elements of its canonical sequence.
-
-\input{movies/snippets/Canon/canonSLimitLub}
-
-\paragraph*{\gaiasign}  In \gaiaHydras, the statement use a slightly different vocabulary:
-\inputsnippets{GCanon/gcanonLimitOf}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}\label{exo:simply-typed-canonseq}
-Instead of using the \texttt{sig} type, define a simply typed function that, given two ordinals $\alpha$ and $\beta$, returns a natural number $i$ such that, if $\alpha$ is a limit ordinal and $\beta<\alpha$, then $\beta< \canonseq{\alpha}{i+1}$.
-Of course, you will have to prove the correctness of your function.
-
-\textbf{Hint:} You may add to your function a third argument usually called \texttt{fuel} for allowing you to give a structurally 
-recursive function (\emph{cf} the post of Guillaume Melquiond on Coq-club (Dec 21, 2020)
-\url{https://sympa.inria.fr/sympa/arc/coq-club/2020-12/msg00069.html}).
-The type \texttt{fuel}, an alternative 
-to \texttt{nat} is available on \href{../theories/html/hydras.Prelude.Fuel.html}{Prelude.Fuel})
-.
-%\index{coq}{Giving fuel to a long computation}
-
-\end{exercise}
-
-\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0}\,: paths}
-\index{maths}{Ordinal numbers!Accessibility inside epsilon0}
-\label{sect:pathes-intro}
-
-Let us consider a kind of accessibility problem inside $\epsilon_0$: given two ordinals $\alpha$ and $\beta$, where $\beta<\alpha<\epsilon_0$, find a \emph{path} consisting of a finite sequence $\gamma_0=\alpha,\dots,\gamma_l=\beta$,
-where, for every $i<l$, $\gamma_i \not= 0$ \footnote{This condition allows us to ignore paths which end by a lot of useless $0$s.} and there exists some strictly positive integer $s_i$
-such that $\gamma_{i+1}=\canonseq{\gamma}{s_i}$.
-
-Let $s$ be the sequence $\langle s_0,s_1,\dots, s_{l-1} \rangle$. We describe the
-existence of such a path with the notation $\alpha\xrightarrow [s]{}\beta$.
-We say also that the considered path from $\alpha$ to $\beta$ \emph{starts at [index] $s_0$ and ends at $s_l$}.
-
-For instance, we have $\omega*2 \xrightarrow[2,2,2,4,5]{}3$, through the 
-path $\langle\omega\times 2, \omega+2,\omega+1,\omega,4,3\rangle$.
-
-\begin{remark}
-
-Note that, given $\alpha$ and $\beta$, where $\beta < \alpha$, the sequence $s$ which leads from $\alpha$ to $\beta$ is not unique.
-Indeed, if $\alpha$ is a limit ordinal, the first element of $s$ can be any integer $i$ such that $\beta<\canonseq{\alpha}{i}$, and if $\alpha$ is a successor ordinal,
-then the sequence $s$ can start with any positive integer.
-
-For instance, we have also 
-$\omega*2 \xrightarrow[3,4,5,6]{}\omega$. 
-Likewise,
-$\omega*2 \xrightarrow[1,2,1,4]{} 0$ and
-$\omega*2 \xrightarrow[3,3,3,3,3,3,3,3]{} 0$.
-\end{remark}
-
-\subsection{Formal definition}
-
-\label{path-to-definition}
-
-In \coq{}, the notion of path can be simply defined as an inductive predicate 
-parameterized by the destination $\beta$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\index{hydras}{Library Epsilon0!Predicates!path\_to}
-\label{sect:path-to-def}
-
-\input{movies/snippets/Paths/transitionDefs}
-
-\input{movies/snippets/Paths/pathDef}
-
-\begin{remark}
-In the present version of our library, we use a variant \texttt{path\_toS} of
-\texttt{path\_to}, where the proposition
-(\texttt{path\_toS $\beta$ $s$ $\alpha$}) is equivalent to
-(\texttt{path\_to $\beta$ (List.map S $s$) $\alpha$}). This variant is scheduled to be deprecated.
-\end{remark}
-
-\paragraph*{\gaiasign}
-The library~\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} transposes the  notion of path into \gaia's type \texttt{T1}
-(please see Sect.\~ref{sect:gpath}).
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Write a tactic for solving goals of the form (\texttt{path\_to $\beta$ $s$ $\alpha$})
-where $\alpha$, $\beta$ and $s$ are closed terms. 
-You should solve automatically the following goals:
-
-\inputsnippets{Paths/exPathTac} 
-\end{exercise}
-
-\subsection{Existence of a path}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction on $\alpha$, we prove that for any $\beta<\alpha$, 
-one can build a path from $\alpha$ to $\beta$ (in other terms, $\beta$ is accessible from $\alpha$).
-
-\input{movies/snippets/Paths/LTPathTo}
-
-\noindent 
-From the lemma \texttt{canon\_LT}~\vref{lemma:canon_LT}, we can convert any path into an inequality on ordinals (by induction on paths).
-
-\input{movies/snippets/Paths/pathToLT}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}[continuation of exercise~\vref{exo:simply-typed-canonseq}]
-Define a simply typed function for computing a path from $\alpha$ to $\beta$.
-\end{exercise}
-
-\subsection{Paths and hydra battles}
-\label{KS-o2h}
-
-In order to apply our knowledge about  ordinal numbers less than $\epsilon_0$ to the study of hydra battles, we define an injection
-from the interval $[0,\epsilon_0)$ into the type \texttt{Hydra}.
-
-\vspace{4pt}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}}
-
-\input{movies/snippets/O2H/iotaDef}
-
-For instance Fig.~\ref{fig:iota-example} shows the image by $\iota$ of the ordinal  \textcolor{black}{$\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$}
-
-  \begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.3]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (10,2) {$\bullet$};
-\node (N22) at (7,2) {$\bullet$};
-\node (N3) at (14,2) {$\bullet$};
-\node (N4) at (18,2) {$\Smiley[2][green]$};
-\node (N5) at (0,4) {$\bullet$};
-\node (N6) at (2,5) {$\Smiley[2][green]$};
-\node (N7) at (4,6) {$\Smiley[2][green]$};
-\node (N88) at (7,4) {$\bullet$};
-\node (N8) at (10,4) {$\bullet$};
-\node (N9) at (14,6) {$\Smiley[2][green]$};
-\node (N10) at (0,8) {$\Smiley[2][green]$};
-\node (N11) at (10,7) {$\Smiley[2][green]$};
-\node (N111) at (7,7) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N22);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (N1) to  (N5);
-\draw (N1) to   [bend left=10] (N6);
-\draw (N1) to   [bend right=20] (N7);
-\draw (N2) to  (N8);
-\draw (N22) to  (N88);
-\draw (N8) to  (N11);
-\draw (N88) to  (N111);
-\draw (N3) to  (N9);
-\draw (N5) to  (N10);
-\end{tikzpicture}
-\caption{The hydra $\iota(\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$) \label{fig:iota-example}}
-
-\end{figure}
-
-The following lemma (proved in ~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H.v}) maps  canonical sequences to rounds of hydra battles.
-
-\label{lemma:canonS-iota}
-
-\input{movies/snippets/O2H/canonSIota}
-
-The next step of our development extends this relationship to
-the order $<$ on $[0,\epsilon_0)$ on one side, and hydra battles on the other side.
-
-\input{movies/snippets/O2H/pathToRoundPlus}
-
-As a corollary, we are now able to transform any inequality $\beta<\alpha<\epsilon_0$ into a (free) battle.
-
-\input{movies/snippets/O2H/LTToRoundPlus}
-
-\section{A  proof of impossibility}
-
-We now have  the tools for proving that  there exists no variant bounded by some $\mu<\epsilon_0$ for proving the termination   of all battles.
-The proof we are going to show is a proof by contradiction.
-It  can
- be considered as a generalization of the
-proofs described in  sections~\vref{omega-case} and \vref{omega2-case}.
-
-In the module
-\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
-This part of the development is parameterized by some class $B$ of battles, which will be instantiated later to \texttt{free} or \texttt{standard}.
-
-\vspace{4pt}
-\noindent
-From~\href{../theories/html/hydras.Hydra.~Hydra_Definitions.html}{Hydra.Hydra\_Definitions}:
-
-\input{movies/snippets/Hydra_Definitions/BoundedVariant}
-
-Let us assume there exists such a variant:
-
-\input{movies/snippets/Epsilon0_Needed_Generic/theContext}
-
-\label{remark:m-decrease}
-\begin{remark}
-  The hypothesis \texttt{m\_decrease} is not provable  in general, but is satisfied by
-the  \texttt{free} and \texttt{standard} kinds of battles. This trick allows to 
-``factorize'' our proofs  of impossibility.
-\end{remark}
-
-\index{maths}{Transfinite induction}
-
-First, we prove that $m(\iota(\alpha))$ is always greater than or equal to $\alpha$, by  transfinite induction over $\alpha$.
-
-\input{movies/snippets/Epsilon0_Needed_Generic/mGe0}
-
-\begin{itemize}
-\item If $\alpha=0$, the inequality trivially holds
-\item If $\alpha$ is the successor of  some ordinal $\beta$, the inequality $\beta \leq m(\iota(\beta))$ holds (by induction hypothesis).
-But the hydra $\iota(\alpha)$ is transformed in one round into
-$\iota(\beta)$, thus $m(\iota(\beta))<m(\iota(\alpha))$. Hence $\beta<m(\iota(\alpha))$, which implies $\alpha \leq m(\iota(\alpha))$
-\item If $\alpha$ is a limit ordinal, then $\alpha$ is the least upper bound of the set
-of all  the $\canonseq{\alpha}{i}$.  Thus, we have just to prove that $\canonseq{\alpha}{i}< m(\iota(\alpha))$ for any $i$. 
-\begin{itemize}
-\item Let $i$ be some natural number.
-By the induction hypothesis, we have $\canonseq{\alpha}{i} \leq m(\iota(\canonseq{\alpha}{i}))$.
-But the hydra $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i})$ in one round, thus $m(\iota(\canonseq{\alpha}{i})) < m(\iota(\alpha))$, by our hypothesis \texttt{m\_decrease}.
-\end{itemize}
-\end{itemize}
-
-Please note that the impossibility proofs of 
-sections~\vref{omega-case} and \vref{omega2-case} contain a similar lemma, also called \texttt{m\_ge}.
-We are now able to build a counter-example.
-
-\input{movies/snippets/Epsilon0_Needed_Generic/mGeGeneric}
-
-The (big) rest of the proof is dedicated to prove formally the converse inequality 
-\texttt{m small\_h t1< m big\_h}.
-
-\subsection{The case of free battles}
-\label{sec:free-battles-case}
-Let us now consider that $B$ is instantiated to \texttt{free} (which means that we are considering proofs of termination of \emph{all} battles).
-The following lemmas are proved in Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
-The case $B=\texttt{standard}$ is studied in section~\vref{std-case}.
-
-\input{movies/snippets/Epsilon0_Needed_Free/theContext}
-
-\begin{enumerate}
-\item The following lemma is an application of \texttt{m\_ge\_generic}, since \texttt{free}
-satisfies trivially the hypothesis \texttt{m\_decrease} (see page~\pageref{remark:m-decrease}).
-
-\inputsnippets{Epsilon0_Needed_Free/mGe}
-
-\item From the hypothesis \texttt{Hy}, we have \texttt{m big\_h t1< mu}
-\item By Lemma \texttt{LT\_to\_round\_plus}, we get a (free) battle from
-\texttt{big\_h = iota mu} to \texttt{small\_h = iota (m big\_h)}.
-
-\input{movies/snippets/Epsilon0_Needed_Free/bigToSmall}
-
-\item From the hypotheses on $m$, we infer:
-\input{movies/snippets/Epsilon0_Needed_Free/mLt}
-
-\item From lemmas \texttt{m\_ge} and \texttt{m\_lt}, and the irreflexivity of $<$, we get a contradiction. 
-\input{movies/snippets/Epsilon0_Needed_Free/ImpossibilityFree}
-
-\end{enumerate}
-
-We have now proved there exists no bounded variant for the class of free battles.
-
-\input{movies/snippets/Epsilon0_Needed_Free/CheckDemo}
-
-\section{The case of standard battles}
-\label{sec:standard-intro}\label{std-case}
-One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
-Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}.
-
-\input{movies/snippets/O2H/LTToRoundPlus}
-
-This lemma builds a battle out of any inequality $\beta<\alpha$. 
-It is a straightforward application of \texttt{LT\_path\_to} of
-Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}:
-
-\input{movies/snippets/Paths/LTPathTo}
-
-The sequence $s$, used to build the sequence of replication factors of the battle, depends on 
-$\alpha$ and $\beta$, so we cannot be sure that the generated battle is a genuine standard battle.
-
-The solution to this issue comes  once again from Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
-$\alpha_0=\alpha,\alpha_1,\dots,\alpha_k=\beta$ where $\alpha_{j+1}$ is equal
-to $\canonseq{\alpha_j}{i_j}$ where $i_j$ is \emph{any} natural number, 
-we consider various constraints on these sequences.
-In particular, a path is called \emph{standard} if $i_{j+1} = i_j + 1$ for every $j<k$.
-It  corresponds to a ``segment'' of some standard battles. 
-Please note that the vocabulary on paths is ours, but all the concepts come really from~\cite{KS81}.
-
-In \coq{}, standard paths can be defined as follows.
-
-\vspace{4pt}
-
-\emph{From
-Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/standardPathR}
-
-In the mathematical text and figures, we shall use the notation 
-$\alpha \xrightarrow[i,j]{}\beta$ for the proposition 
-(\texttt{standard\_path $i$ $\alpha$ $j$ $\beta$}).
-In~\cite{KS81} the notation is
-$\alpha \xrightarrow[i]{*}\beta$
-for 
-the proposition  $\exists j, i<j \wedge \alpha \xrightarrow[i,j]{} \beta$.
-
-Our goal is now  to transform any inequality $\beta<\alpha<\epsilon_0$ into a standard path $\alpha \xrightarrow[i,j]{} \beta$ for some $i$ and $j$, then into a standard battle
-from $\iota(\alpha+i)$ to $\iota(\beta)$. 
-Following~\cite{KS81}, we proceed in two stages:
-\begin{enumerate}
-\item we simulate plain (free) paths from $\alpha$ to $\beta$ with
-paths made of steps $(\gamma,\canonseq{\gamma}{n})$, \emph{with the same $n$ all along the path}
-\item we simulate any such path by a standard path.
-\end{enumerate}
-
-\subsection{Paths with a constant index}
-
-First of all, paths with a constant index 
-enjoy nice properties. They are defined as paths where all the $i_j$ are equal to the same natural number $i$, for some $i>0$.
-
-Like in~\cite{KS81}, we shall use the notation $\alpha \xrightarrow[i]{} \beta$ for denoting such a path, also called an $i$-path.
-
-\input{movies/snippets/Paths/constPathDef}
-
-% Paths with a given index can be effectively computed.
-% Given $i$, $\alpha$ and $l$, the following function returns the ordinal $\beta$ such that there exists a path 
-% $\alpha \xrightarrow [i+1] {} \beta$ of length $l$.
-
-% \begin{Coqsrc}
-% Fixpoint const_funS (i:nat)(alpha : T1)(l:nat):  T1  :=
-%   match l
-%   with
-%   | 0 => alpha
-%   | S m => const_funS i (canonS i alpha) m
-%   end.
-% \end{Coqsrc}
-
-% The following computations show  applications of \texttt{constS\_fun} to the 
-% ordinal $\omega^\omega$, with various values of $i$ and $l$.
-
-% \begin{Coqsrc}
-% Compute  (const_funS 2 (omega ^omega)  55).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = zero
-%      : T1 
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (const_funS 2 (omega ^omega) 15).
-% \end{Coqsrc}
-
-%   \begin{Coqanswer}
-%  = (omega ^ 2 * 2)%pT1
-%      : ppT1   
-%   \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (const_funS 4 (omega^omega)  100).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% = (omega ^ 4 * 4 + omega ^ 3 * 4 + omega ^ 2 + omega * 4 + 4)%pT1
-%      : ppT1
-% \end{Coqanswer}
-
-A most interesting property of $i$-paths is that we can ``upgrade'' their index, as stated by K.\&S.'s Corollary 12.
-
-\index{maths}{Transfinite induction}
-
-\vspace{4pt}
-\noindent
-\emph{From
-  Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/Cor12}
-
-We  also use a version of \texttt{Cor12} with large inequalities.
-
-\input{movies/snippets/Paths/Cor121}
-
-\subsubsection{Sketch of proof of \texttt{Cor12}}
-\index{maths}{Transfinite induction}
-
-First, let us prove the following lemma (numbered \texttt{2.4} in Ketonen and Solovay's article).
-
-\input{movies/snippets/Paths/KSThm24}
-
-We prove this lemma by transfinite induction on $\alpha$.
-Let us consider a path $\alpha \xrightarrow [i]{} \beta$ $(i>0)$. Its first step is
-the pair $(\alpha,\canonseq{\alpha}{i})$, We have $\canonseq{\alpha}{i}<\alpha$ and
-$\canonseq{\alpha}{i} \xrightarrow [i]{} \beta$. 
-Let $n$ be any natural number such that $n>i$.
-By the induction hypothesis, there exists a path $\canonseq{\alpha}{n} \xrightarrow[i]{} \beta$.
-\begin{itemize}
-\item  If $\alpha$ is a successor ordinal $\gamma+1$, then $\canonseq{\alpha}{n} =
-\canonseq{\alpha}{i}=\gamma$. Thus we have a path 
-$\alpha  \xrightarrow [n]{}  \gamma \xrightarrow [n]{} \beta$
-\item If $\alpha$ is a limit ordinal, we apply Theorem \texttt{KS\_thm\_2\_4} (see above).
-
-%   \begin{theorem}
-% Let $\lambda$ be a limit ordinal, then for any pair of indices $0<i<j$, there is a path $\canonseq{\lambda}{j} \xrightarrow[1]{} \canonseq{\lambda}{i}$.    
-%   \end{theorem}
-
-The proof of the limit case, is decomposed into a sequence
-of path constructions leading to $\alpha \xrightarrow[n]{} \beta$.
-
- \begin{enumerate}
- \item $\alpha \xrightarrow[n]{} \canonseq{\alpha}{n}$ (single step path)
- \item $\canonseq{\alpha}{n} \xrightarrow[1]{} \canonseq{\alpha}{i}$ (by \texttt{Theorem\_2\_4}),
-\item $\canonseq{\alpha}{n} \xrightarrow[n]{} \canonseq{\alpha}{i}$ (applying the induction hypothesis to the preceding path);
-\item $\canonseq{\alpha}{i} \xrightarrow[n]{} \beta$ (applying the induction hypothesis)
-\item $\alpha \xrightarrow[n]{} \beta$ (by composition of 1, 3, and 4).
-
- \end{enumerate}
-
-\end{itemize}
-
-\begin{remark}
- \texttt{Cor12} ``casts'' $i$-paths into $n$-paths for any $n>i$.
-But the obtained $n$-path can be much longer than the original $i$-path.
-The following exercise will give an idea of this increase. 
-\end{remark}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-  Prove that  the length of the $i+1$-path from
-  $\omega^\omega$ to $\omega^i$ is $1 + (i+1)^{(i+1)}$, for any $i$. Note that the $i$-path from
-  $\omega^\omega$ to $\omega^i$ is only one step long.
- \end{exercise}
-
-Why is \texttt{Cor12} so useful? 
-Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
-we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying corollary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
-where $n$ is the maximum of the indices $i$ met in the induction.
-
- Lemma 1, Section 2.6 of~\cite{KS81} is naturally expressed in terms of \coq's
-\verb@sig@ construct.
-
-\label{lemma:L-2_6-1}
-%\index{coq}{Sigma types}
-
-\input{movies/snippets/Paths/Lemma261}
-
-Intuitively, Lemma~\texttt{2\_6\_1}  shows that if $\beta<\alpha<\epsilon_0$, then there exists  a battle from $\iota(\alpha)$ to $\iota(\beta)$ where the replication factor is constant, although large enough.
-
- \paragraph*{\gaiasign}
- Corollary \texttt{Cor12} and Lemma \texttt{Lemma2\_6\_1} are also available in
-\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} 
-(please see Sect.~\vref{sect:gpath}).
-
-\subsection{Casting paths with a constant index into a standard path}
-
-The article~\cite{KS81} contains 
-the following lemma, which allows us to simulate $i$-paths by $[i+1,j]$-paths, where $j$ is large enough.
-
-\input{movies/snippets/Paths/constantToStandard}
-
-\subsubsection{Sketch of proof of \texttt{constant\_to\_standard\_path}}
-
-Our proof follows the proof by Ketonen and Solovay, including its organization as a sequence of lemma.
-Since it is a non-trivial proof, we will comment its main steps below   (see Figure~\vref{fig:belle-preuve-1} to Figure~\vref{fig:fin-belle-preuve}).
-
-\subsubsection*{Preliminaries}
-
-Please note that, given an ordinal $\alpha:\texttt{T1}$, and two natural numbers $i$ and $l$, there exists at most a standard path $\alpha \xrightarrow [i,i+l]{*} \beta$.
-The following function computes $\beta$ from $\alpha$, $i$ and $l$.
-
-\input{movies/snippets/Paths/standardGnaw}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction over  $\alpha$, we prove that the ordinal $0$ is reachable from any ordinal $\alpha<\epsilon_0$ by some standard path.
-
-\input{movies/snippets/Paths/standardPathToZero}
-
-\paragraph*{}
-Now, let us consider two ordinals  $\beta<\alpha<\epsilon_0$.  Let $p$  be some $(n+1)$-path from $\alpha$ to $\beta$.
-
-\input{movies/snippets/Paths/ConstantToStandardProof}
-
-Applying \texttt{standard\_path\_to\_zero}, $0$ is reachable from $\alpha$ by some standard path.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \draw[->, very thick,blue] (alpha)-- node [below]{$n+1$} node [above] {$+$} (beta);
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-    \node (alpha3) at (20,5) {};
-  \node (alphalast) at (35,5) {};
-  \node (zero) at (45,0) {$0$};
-  \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate=40]{$n+1$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha2)-- node [below]{$n+3$}  (alpha3);
-
-  \node (dots) at (24,5) {$\dots$};
-  \draw [->, dashed, very thick,blue] (alphalast)-- node [below, rotate=-26]{$n+p+1$}  (zero);
-
-\end{tikzpicture}
-\caption{A nice proof (1)}
-  \label{fig:belle-preuve-1}
-\end{figure}
-
-\paragraph*{}
-
-Since comparison on \texttt{T1} is decidable, one can compute the last step $\gamma$ of the standard path from $(\alpha,n+1)$  such that $\beta\leq \gamma$.
-Let $l$ be the length of the path from $\alpha$ to $\gamma$.  
-This step of the proof is illustrated in figure~\vref{fig:belle-preuve-2}.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (17,5) {$\ldots$};
-    \node (alpha3) at (20,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (38,0) {$\delta$};
-    \draw [->, dashed,very thick,blue] (alpha)-- node [below,rotate=35]{$n+1$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below,rotate = -48]{\tiny $n+l$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{$n+l+1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{$n+1$} node [above] {$+$} (beta);
-   \draw[thick] (alpha)--  (gamma);
-   \draw[thick] (gamma)--  node [above] {$\geq$} (beta);
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-\end{tikzpicture}
-
-\caption{A nice proof (2)}
-  \label{fig:belle-preuve-2}
-\end{figure}
-
-\paragraph*{}
-
-\begin{itemize}
-\item If $\beta=\gamma$, it's OK! We have got a standard path
-from  
-$\alpha$ to $\beta$ with successive indices  $n+1, n+2, \dots, n+l+1$
-
-\item Otherwise,  $\beta < \gamma$.  Let us consider  $\delta=\canonseq{\gamma}{n+l+1}$.
-By applying several times lemma \texttt{Cor12},  one converts  every path of Fig~\ref{fig:belle-preuve-2} into
- a $n+l+1$-path  (see figure~\ref{fig:belle-preuve-3}).
-
-But $\gamma$ is on the $n+l+1$-path from $\alpha$ to $\beta$.
-As shown by figure~\vref{fig:fin-belle-preuve}, the ordinal $\delta$, reachable from
-$\gamma$ in one single step,  must be greater than or equal to $\beta$, which contradicts our  hypothesis $\beta < \gamma$.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-    \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (17,5) {$\ldots$};
-    \node (alpha3) at (20,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (38,0) {$\delta$};
-     \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -48]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\tiny $n+l+1$} node [above]{\color{red} $1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
-    \draw[thick] (gamma)--   node [above]{\color{red} $>$}(beta);
-   \draw[thick] (alpha)--  (gamma);
-
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-
-\end{tikzpicture}
-
-\caption{A nice proof (3)}
-  \label{fig:belle-preuve-3}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (15,5) {$\ldots$};
-    \node (alpha3) at (18,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (42,0) {$\delta$};
-    \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -36]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\small $n+l+1$} node [above]{\color{red} $1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
-   \draw[thick] (alpha)--  (gamma);
-
-    \draw[thick] (gamma)--  node [below]{\tiny $n+l+1$} node [above]{\color{red} $+$}(beta);
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-
-\end{tikzpicture}
-
-\caption{A nice proof (4)}
-  \label{fig:fin-belle-preuve}
-\end{figure}
-
-\end{itemize}
- The only possible case is  thus $\beta=\gamma$, so we have got a standard path  from $\alpha$ to $\beta$.
-
- \input{movies/snippets/Paths/constantToStandard0}
- \input{movies/snippets/Paths/constantToStandardz}
-
-Here is the full statement of the conversion from constant to standard paths.
-
-\input{movies/snippets/Paths/constantToStandardPath}
-
-Applying \texttt{Lemma2\_6\_1} and \texttt{constant\_to\_standard\_path}, we get the following corollary.
-
-\input{movies/snippets/Paths/LTToStandardPath}
-
-\subsection{Back to hydras}
-\label{sec:standard-battles-cases}
-We are now able to complete our proof that there exists no bounded variant for proving the termination of standard hydra battles. This proof can
-be consulted in the module 
-\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}{Hydra.Epsilon0\_Needed\_Std}.
-Please note that it has the same global structure as in section~\ref{sec:free-battles-case} 
-% ICI !
-Applying the  lemmas  \texttt{Lemma2\_6\_1} of the module 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
-{Epsilon0.pathS}   and 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
-{\texttt{constant\_to\_standard\_path}},
-we can convert any inequality $\beta<\alpha<\epsilon_0$ into a standard path from
-$\alpha$ to  $\beta$, then into a fragment of a standard battle from 
-$\iota(\alpha)$ to $\iota(\beta)$, hence the inequality $m(\iota(\beta))<m(\iota(\alpha))$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html\#LT_to_standard_battle}{Hydra.Epsilon0\_Needed\_Std}}
-
-\input{movies/snippets/Epsilon0_Needed_Std/LTToStandardBattle}
-
-Now, please consider the following context:
-
-\input{movies/snippets/Epsilon0_Needed_Std/theContext}
-
-In the same way as for free battles, we import a large inequality 
-from 
-the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Epsilon0\_Needed\_Generic}.
-
-\input{movies/snippets/Epsilon0_Needed_Std/mGe}
-
-\paragraph*{} If remains to prove the following strict inequality, in order to have a contradiction.
-
-\input{movies/snippets/Epsilon0_Needed_Std/mLt}
-
-\paragraph*{Sketch of proof:} Let us recall that $\texttt{big\_h} = \iota(\mu)$
- and $\texttt{small\_h} = \iota (m (\texttt{big\_h}))$.
-
-Since $m(\texttt{big\_h})< \mu$, there exists a standard path from $\mu$ to
-$m(\texttt{big\_h})$, hence a   standard battle from $\iota(\mu)$  to
-$\iota(m(\texttt{big\_h}))$,  i.e. from \texttt{big\_h} to \texttt{small\_h}.
-
-Since $m$ is assumed to be a variant for standard battles, we get the inequality  $m(\texttt{small\_h}) < m(\texttt{big\_h})$.
-
-\input{movies/snippets/Epsilon0_Needed_Std/endOfProof}
-
-\subsection{Remarks}
-
-We are grateful to 
- J. Ketonen and R. Solovay  for the high quality of their explanations and proof details.
-Our proof follows tightly the sequence of lemmas in their article, with a focus on 
-constructive aspects.
-Roughly speaking, our implementation \emph{builds}, out of a hypothetical
-  variant $m$, bounded by some ordinal $\mu<\epsilon_0$, a hydra \texttt{big\_h} which verifies the impossible inequality  $m(\texttt{big\_h})< m(\texttt{big\_h})$.
-
-On may ask whether the preceding results are too restrictive, since they 
-refer to a particular data type \texttt{T1}.
-In fact, our representation of ordinals strictly less than 
- $\epsilon_0$ is faithful to their mathematical definition, at least 
-Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
-(please see also the module
-\href{../theories/html/hydras.Schutte.Correctness_E0.html}{hydras.Schutte.Correctness\_E0}).
-
-Thus, we can infer that our theorems can be applied to any well order.
-
-\index{hydras}{Projects}
-\begin{project}
-Study a possible modification of the definition of a variant  (for  standard battles).
-
-\begin{itemize}
-\item The variant is assumed to be strictly decreasing \emph{on configurations 
-reachable from some initial configuration where the replication factor is equal to $0$}
-\item The variant may depend on the number of the current round.
-\end{itemize}
-
-In other words, its type should be \texttt{nat -> Hydra -> T1}, and it must 
-verify the inequality $m\, (S\,i)\, h' < m\,i\, h$ whenever the configuration 
-$(i,h)$ is reachable from some initial configuration $(0,h_0)$
-and \texttt{h} is transformed into \texttt{h'} in the considered round.
-Can we still prove the theorems of section~\ref{std-case} with this new definition?
-
-\end{project}
-
-%---------------------------------------------------------------------
-\chapter{Large sets and rapidly growing functions}\label{chap:alpha-large}
-
-\begin{remark}
-Some notations (mainly names of fast-growing functions) of our development may differ slightly from the literature.
-Although this fact does not affect our proofs, we are preparing a future version where the names $F\_alpha$, $f\_alpha$, $H\_alpha$, etc., are fully consistent with the cited articles.
-
-\end{remark}
-%\section{Introduction}
-
-In this chapter, we try to feel how long a standard battle can be.
-To be precise, for any ordinal $\alpha<\epsilon_0$ and any positive integer $k$,
-we give a minoration of the number of steps of a standard battle which
-starts with the hydra $\iota(\alpha)$ and the replication factor $k$.
-
-We express this number in terms of the Hardy hierarchy of fast-growing 
-functions~\cite{BW85, Wainer1970, KS81, Promel2013}.
- From the \coq{} user's point of view, such  functions are  very 
-attractive:  they are defined as functions  in \gallina{}, and we can apply them \emph{in theory}, but they are so complex that you will never be able to look at the result of the computation.
- Thus, our knowledge on these functions must rely on \emph{proofs}, not tests. In our development, we use often the rewriting rules generated by \coq's \texttt{Equations} plug-in.
-
-\section{Definitions}
-
-%\subsection{Definition}
-
-\begin{definition}
-Let $0<\alpha<\epsilon_0$ be any ordinal, and $s=\langle s_1, s_2, \dots, s_N\rangle$ a finite sequence of strictly positive natural numbers.
-
-We say that $s$ is \emph{$\alpha$-large} if the sequence $\langle \alpha_0=\alpha,\dots,\alpha_{i+1}=\canonseq{\alpha_i}{i+1},\dots \rangle$ leads to $0$. 
-We say also that $s$ is \emph{minimally $\alpha$-large} (in short:
-\emph{$\alpha$-mlarge}) if $s$ is $\alpha$-large 
- and every strict prefix of $s$ leads to a non-zero ordinal (\emph{cf} Sect.~\vref{sect:path-to-def}).
-
-\index{maths}{Ordinal numbers!Large sets}
-\index{maths}{Ordinal numbers!Minimal large sets}
-
-\end{definition}
-
-\begin{remark}
-  Ketonen and Solovay~\cite{KS81} consider  large finite \emph{sets} of natural numbers,  but they are mainly used as sequences. Thus, we chose to represent them explicitely as (sorted) lists. 
-\end{remark}
-
-The following function ``gnaws'' an ordinal $\alpha$, following a sequence of indices (ignoring the $0$s).
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/gnawDef}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_SetsPaths.html\#gnaw}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/largeDef}
-
-Minimal large sequences can be directly defined in terms of the
-predicate \texttt{path\_to} (\vref{sect:path-to-def}) which already prohibits paths containing non-final \texttt{zero}s.
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}
-
-\index{hydras}{Library Epsilon0!Predicates!mlarge@mlarge (minimal large sequences)}
-
-\input{movies/snippets/Large_Sets/mlargeDef}
-
-Let us consider two integers $k$ and $l$, such that $0<k<l$. In order to check whether the interval $[k,l]$ is minimally large for $\alpha$, it is enough to
-follow from $\alpha$ the path associated with the interval $[k,l)$ and verify that the last ordinal we obtain is equal to $1$.
-
-\subsection{Examples}
-
-For instance the interval $[6,70]$ leads $\omega^2$ to $\omega\times 2 + 56$. Thus this interval is not $\omega^2$-large.
-
-\noindent
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html\#mlarge}{Epsilon0.Large\_Sets\_Examples}}
-
-\input{movies/snippets/Large_Sets_Examples/gnawEx1}
-
-The interval $[6,700]$ is $\omega^2$-large, but not
-$\omega^2$-mlarge, since $[6,699]$ is also $\omega^2$-large.
-
-\input{movies/snippets/Large_Sets_Examples/gnawEx2}
-
-The following lemma relates minimal largeness with the
-function 
-\texttt{gnaw}.
-
-\input{movies/snippets/Large_Sets/mlargeIff}
-
-\vspace{4pt}
- \noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html}{Epsilon0.Large\_Sets\_Examples}}
-
-\input{movies/snippets/Large_Sets_Examples/Ex1Lemma}
-
-\section{Length of minimal large sequences}
-
-Now, consider any natural number $k>0$ any ordinal $0<\alpha<\epsilon_0$.  We would like to compute
-a number $l$ such that the interval $[k,l]$ is $\alpha$-mlarge. So, 
-the standard battle starting with $\iota(\alpha)$ and the replication factor $k$ will end after $(l-k+1)$ steps.
-
-First, we notice that this  number $l$ exists, since the segment $[0,\epsilon_0)$ is well-founded and $\canonseq{\alpha}{i}<\alpha$ for any $i$ and $\alpha>0$.
-Moreover, it is unique:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/mlargeUnicity}
-
-Thus, we would like to define a function, parameterized by $\alpha$ which associates to any  strictly positive integer $k$ the number $l$ such that
-the interval $[k,l]$ is $\alpha$-mlarge. It would be fine to write in \gallina{} a definition like this:
-
-\begin{Coqbad}
-Function L_ (alpha: E0) (i:nat) :  nat := ...
-\end{Coqbad}
-
-But we do not know how to fill the dots yet \dots{}   In the next section, we will 
-use \coq{} to reason  about the \emph{specification} of \texttt{L},
-prove properties of any function which satisfies this specification.
-In Sect.~\ref{sect:L-equations}, we use the \texttt{coq-equations} plug-in
-to define a function \texttt{L\_}, and prove its correctness w.r.t. its specification.
-
-\subsection{Formal specification}
-
-Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider any  function which  maps  any strictly positive integer $k$ to the number $l$, where 
-the interval $[k,l)$ is $\alpha$-mlarge.
-
-\begin{remark}
-In~\cite{KS81} Ketonen and Solovay consider the least natural number $l$ where the interval $[k,l]$ ($l$ included) is $\alpha$-large, and call $H_\alpha$ the function which maps $k$ to $l$.
-We chose to consider intervals $[l,k)$ instead of $[l,k]$
-in order to simplify  some statements and proofs in composition lemmas associated with the ordinals of the form $\alpha\times i$ and 
-$\omega^\alpha\times i + \beta$.
-Clearly, both approaches are related through the equality
-$L_\alpha(k)=H_\alpha(k)+1$, for any non-null $\alpha$ and $k$.
-\end{remark}
-
-Our specification of the function \texttt{L} is as follows:
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/LSpecDef}
-
-\begin{todo}
- Check if the functions $L_\alpha$ are the same as
-\cite{KS81}' functions $f_\alpha$ (p. 297).
-\end{todo}
-
-Note that, for $\alpha\not=0$, the value of $f(0)$ is not specified.
-Nevertheless, the restriction of $f$ to the set of strictly positive integers is unique (up to extensionality).
-
-\input{movies/snippets/Large_Sets/LSpecUnicity}
-
-\subsection{Abstract properties}
-
-Let us now prove properties of any function $f$ (if any) which satisfies 
-\texttt{L\_spec}.
-We are looking for properties which could be used for writing \emph{equations} and prove the correctness of the function generated by the \texttt{coq-equations} plug-in.
-Moreover, they will give us some examples (for small values of $\alpha$).
-
-The properties we consider are defined in \href{../theories/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}.
-
-\label{sect:abstract-arith-prop}
-\input{movies/snippets/Iterates/funLeDef}
-
-Our exploration of the $L_\alpha$\,s  considers the usual cases of a proof by transfinite induction: zero, successors and limit ordinals.
-The lemmas we are going to prove will be applied in a big proof by induction in Sect~\vref{sect:L-correct-proof}.
-
-\index{maths}{Transfinite induction}
-
-\subsubsection{The  ordinal zero}
-\label{sect:L-spec-zero}
-The base case is directly a consequence of the specification.
-
-\input{movies/snippets/Large_Sets/LZeroInv}
-
-\subsubsection{Successor ordinals}
-\label{sect:L-spec-succ}
-Let $\beta$ be some ordinal, and assume the arithmetic function $f$ satisfies 
-the specification $(\texttt{L\_spec}\;\beta)$.  Let $k$ be any natural number.
-Any path from $\texttt{succ}\,\beta$ to $0$ starting at $k+1$ can be decomposed into a first step from $\texttt{succ}\,\beta$ to $\beta$, then a path from
-$\beta$ at $k+2$ to $0$. 
-By hypothesis the interval $[k+2, f(k+2)-1]$ is $\beta$-mlarge.
-But the interval $[k+1, f(k+2)-1]$ is the concatenation of the singleton
-$\{k+1\}$ and the interval $[k+2, f(k+2)-1]$.
-So, the function $\lambda\,k.\,f(k+1)$ satisfies the specification $\texttt{L\_spec}\,\beta$.
-
-Note that our decomposition of intervals works only if the intervals we consider are not empty.
-In order to ensure this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <<= f}, or \texttt{(fun\_le S f)}.
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/SectionSucc}
-
-\subsubsection{Limit ordinals}
-\label{sect:L-spec-lim}
-
-Let $\lambda<\epsilon_0$ be any limit ordinal. In a similar way as for successors, we decompose any path from $\lambda$  into a first step to
-$\canonseq{\lambda}{k}$, followed by a path to $0$. In the following section, we assume that there exists a correct function computing  $L_{\canonseq{\lambda}{k}}$ for any strictly positive $k$.
-
-\input{movies/snippets/Large_Sets/SectionLim}
-
-\subsection{First results}
-
-Applying the previous lemmas on successors and limit ordinals, 
-we obtain a few  correct implementations of \texttt{(L\_spec $\alpha$)} for small values of $\alpha$.
-
-\subsubsection{Finite ordinals}
-
-By iterating the functional \texttt{L\_succ}, we get a realization of
-\texttt{(L\_spec $i$)} for any natural number $i$.
-
-\input{movies/snippets/Large_Sets/LFinDef}
-\vspace{-16pt}
-\input{movies/snippets/Large_Sets/LFinOk}
-
-\subsubsection{The first limit ordinal  \texorpdfstring{$\omega$}{omega}}
-
-The lemmas \texttt{L\_fin\_ok} and \texttt{L\_lim\_ok}   allow us to get 
-by diagonalization a correct implementation for 
-\texttt{L\_spec omega}.
-
-\input{movies/snippets/Large_Sets/LOmegaDef}
-\vspace{-16pt}
-\input{movies/snippets/Large_Sets/LOmegaOk}
-
-\subsubsection{Towards  \texorpdfstring{$\omega^2$}{omega*omega}}
-
-We would like to get exact formulas for the ordinal $\omega^2$, a.k.a.
-$\phi_0(2)$. This ordinal is the limit of the sequence $\omega\times i\;(i \in \mathbb{N})$. Thus, we have to study ordinals of this form, then use 
-our lemma on limits.
-
-The following lemma establishes a path from $\omega\times ( i+1)$ to
-$\omega \times i$.
-
-\input{movies/snippets/Large_Sets/pathToOmegaMult}
-
-Let us consider a path from  $\omega\times(i+1)$ to $0$ starting at $k+1$.
-A first ``big step'' will lead to $\omega\times i$ at $2(k+1)$. If $i>0$, the
-next jump leads to $\omega\times(i-1)$ at $2(2(k+1))+1$, etc.
-
-The following lemma expresses the length of the mlarge sequences associated with any finite multiple of $\omega$.
-
-\input{movies/snippets/Large_Sets/omegaMultMlarge0}
-
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#L_omega_mult}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/LOmegaMultDef}
-
-More generally, we prove the equality $L_{\omega\times i}(k)=2^i\times(k+1)-1$.
-
-\input{movies/snippets/Large_Sets/LOmegaMultEqn}
-
-Correctness of the function \texttt{L\_omega\_mult} is asserted through the following lemma.
-
-\input{movies/snippets/Large_Sets/LOmegaMultOk}
-
-By diagonalization, we obtain a simple formula for $L_{\omega^2}$.
-
-\input{movies/snippets/Large_Sets/LOmegaSquare}
-\input{movies/snippets/Large_Sets/LOmegaSquareEqn}
-\input{movies/snippets/Large_Sets/LOmegaSquareOk}
-
-%%%% ICI
-
-\subsubsection{Going further}
-Let us consider a last example, ``computing'' $L_{\omega^3}$.
-Let us study the canonical sequence associated with $\omega^3$, the elements of which are 
-$\omega^2\times i\;(i\in\mathbb{N}_1)$.
-
-To this end, we prove a generic lemma, which expresses $L_{\omega^\alpha\times i}$ as an iterate of $L_{\omega^\alpha}$.
-Note that in this lemma, we assume that the function associated with $\alpha$ is strictly monotonous and
-greater or equal than the successor function, and prove that $L_{\omega^\alpha\times i}$ satisfies  the same properties.
-
-\inputsnippets{Large_Sets/phi0Mult}
-
-\inputsnippets{Large_Sets/phi0MultOk,
-Large_Sets/Phi0MultSLe}
-
-Let us look now
-at the ordinal $\omega^2\times i$, using \texttt{L\_phi0\_mult}.
-
-\input{movies/snippets/Large_Sets/LOmegaSquareTimes}
-
-We are now ready to get an exact formula for $L_{\omega^3}$, by diagonalization over $L_{\omega^2\times i}$.
-
-\input{movies/snippets/Large_Sets/LOmegaCube}
-
-Thus, for instance, $L_{\omega^3}(3)=L_{\omega^2\times 4}(3)$.
-
-\input{movies/snippets/Large_Sets/LOmegaCube3Eq}
-
-This number is quite big. Using \texttt{Ocaml}'s \texttt{float} arithmetic,
-we can under-approximate it by $2^{3.8\times10^{30}}\times 3.8\times{10^{30}}$.
-
-\begin{Coqsrc}
-# let exp2 x = 2.0 ** x;;
-
-val exp2 : float -> float = <fun>
-#   exp2 95.0 *. 97.0 -. 1.0;;
-- : float = 3.84256588194182037e+30
-# let n = exp2 95.0 ;;
-# let p = n *. 97.0 -. 1.0;;
-val p : float = 3.84256588194182037e+30
-
-Estimation :
-2 ** (3.84 e+30) * 3.84 e+30.
-\end{Coqsrc}
-
-\subsection{Using \texttt{Equations}}
-\label{sect:L-equations}
-
-Note that we did not define any function $L_\alpha$ \emph{for any $\alpha<\epsilon_0$} yet.
-We have got no more than a collection of proved realizations of $\texttt{L\_spec}\;\alpha$ for a few values of $\alpha$.
-
-\index{coq}{Plug-ins!Equations}
-
-Using the \texttt{coq-equations} plug-in by 
-M. Sozeau~\cite{sozeau:hal-01671777}, we will now define a function \texttt{L\_} which maps any ordinal  $\alpha<\epsilon_0$ to a proven realization of 
-$\texttt{L\_spec}\;\alpha$.   
-To this end, we represent ordinals as inhabitants of the type 
-\texttt{E0} of well-formed ordinal terms (see Sect~\vref{sect:E0-def}). So, we define a total function \texttt{L\_} of type
-\texttt{E0 -> nat -> nat}, by transfinite recursion, considering the usual three cases : $\alpha=0$, $\alpha$ is a successor, $\alpha$ is a limit ordinal.
-
-\subsubsection{Definition}
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#L_}{L\_alpha}).}
-
-\label{Functions:L-alpha}
- \index{hydras}{Library Epsilon0!Functions!L\_@L\_ (final step of a minimal path}
-
-\input{movies/snippets/L_alpha/LDef}
-
-This definition results in a bunch of automatically generated lemmas. For instance:
-
-\input{movies/snippets/L_alpha/AboutLEquation1}
-
-In most cases, it may be useful to write human-readable  paraphrases of these statements.
-
-\inputsnippets{L_alpha/Paraphrasesa}
-\inputsnippets{L_alpha/Paraphrasesb}
-\inputsnippets{L_alpha/Paraphrasesc}
-
-Using these three lemmas as rewrite rules, we can prove more properties of the functions \texttt{L\_$\alpha$}.
-
-\input{movies/snippets/L_alpha/LFiniteOmega}
-
-By  well-founded induction on $\alpha$, we prove the following properties:
-
-\input{movies/snippets/L_alpha/LGeS}
-
-\input{movies/snippets/L_alpha/LCorrect}
-
-\label{sect:L-correct-proof}
-
-Please note that the proof of \texttt{L\_correct} applies the lemmas proven in Sections~\ref{sect:L-spec-zero}, ~\ref{sect:L-spec-succ} and ~\ref{sect:L-spec-lim}.
-Our previous study of \texttt{L\_spec} allowed us to pave the way for the definition by \texttt{Equations} and the correctness proof.
-
-\paragraph*{\gaiasign}
-The module
-\href{../theories/html/gaia_hydras.GL_alpha.html}{gaia\_hydras.GL\_alpha} contains an adaptation of
-\href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0/L\_alpha}.
-
-\subsubsection{Back to hydra battles}
-\label{def:L-alpha}
-
-Lemma \texttt{battle\_length\_std } of
-Module~\href{../theories/html/hydras.Hydra.Battle_length.html}{Hydra.Battle\_length} relates the length of standard battles with the functions $L_\alpha$.
-
-\input{movies/snippets/Battle_length/battleLengthStd}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Instead of considering standard paths and battles, consider battles where the replication factor is a constant $k$.
-Please use \texttt{Equations} in order to define the function that computes the length of the $k$-path which leads  from $\alpha$ to $0$.
-Prove a few  exact formulas and minoration lemmas.
-\end{exercise}
-
-\section{A variant of the Hardy hierarchy}
-
-\label{sect:hardy}
-
-In order to give a feeling on  the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the  \emph{Hardy hierarchy} of fast growing functions,
-presented for instance in~\cite{Promel2013}. 
-\index{maths}{Rapidly growing functions!Hardy Hierarchy}
-
-\begin{remark}
-  Indeed, the functions presented in this section are a \emph{variant} of the Hardy hierarchy of functions.
-In the future versions of this development, we will correct the references to the literature.
-For the time being, we call our functions $H'_\alpha$ in order to underline the difference from ``classic'' Hardy functions.
-\end{remark}
-
-For each ordinal $\alpha$ below $\epsilon_0$, $H'_\alpha$ is 
-a total arithmetic function, defined  by  transfinite recursion on $\alpha$, according to three cases:
-
-\index{maths}{Transfinite induction}
-
-\begin{itemize}
-\item If $\alpha=0$, then $H'_\alpha (k)= k$ for any natural number $k$.
-\item If $\alpha=\textrm{succ}(\beta)$, then 
-$H'_\alpha(k)=H'_\beta(k+1)$ for any $k \in \mathbb{N}$
-\item If $\alpha$ is a limit ordinal, then 
-$H'_\alpha(k) = H'_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
-\end{itemize}
-
-\begin{remark}
- The ``classic'' definition of the Hardy hierarchy differs in the third equation.
-
-\begin{itemize}
-\item If $\alpha=0$, then $H_\alpha (k)= k$ for any natural number $k$.
-\item If $\alpha=\textrm{succ}(\beta)$, then 
-$H_\alpha(k)=H_\beta(k+1)$ for any $k \in \mathbb{N}$
-\item If $\alpha$ is a limit ordinal, then 
-$H_\alpha(k) = H_{(\canonseq{\alpha}{k})}(k)$ for any $k\in \mathbb{N}$.
-\end{itemize}
-
-\end{remark}
-
-\subsection{Definition in \texttt{Coq}}
-
-We define a function \texttt{H'\_} of type \texttt{E0 -> nat -> nat} by transfinite induction over the type \texttt{E0} of the well formed ordinals below $\epsilon_0$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}}
-
-\index{hydras}{Library Epsilon0!Functions!H\_@H\_ (Hardy hierarchy (variant))}
-\index{coq}{Plug-ins!Equations}
-\label{Functions:Hprime-alpha}
-
-\input{movies/snippets/Hprime/HprimeDef}
-\input{movies/snippets/Hprime/HprimeDefb}
-
-\input{movies/snippets/Hprime/paraphrasesa}
-\input{movies/snippets/Hprime/paraphrasesb}
-\input{movies/snippets/Hprime/paraphrasesc}
-\input{movies/snippets/Hprime/paraphrasesd}
-
-\subsection{First  steps of the H' hierarchy}
-Using rewrite rules from \texttt{H'\_eq1} to \texttt{H'\_succ\_eqn}, we can explore the functions $H'_\alpha$ for  small values of $\alpha$.
-
-\paragraph*{\gaiasign}
-The module
-\href{../theories/html/gaia_hydras.GHprime.html}{gaia\_hydras.GHprime} contains an adaptation of
-\href{../theories/html/hydras.Epsilon0.Hprime.html}{Epsilon0.Hprime}.
-
-\subsubsection{Finite ordinals}
-
-By induction on $i$, we prove a simple expression of \texttt{H'\_ (Fin i)}, where 
-\texttt{Fin $i$}  is the $i$-th finite ordinal.
-
-\input{movies/snippets/Hprime/HprimeFin}
-
-\subsubsection{Multiples of \texorpdfstring{$\omega$}{omega}}
-
-Since the canonical sequence of $\omega$ is composed of finite ordinals, 
-it is easy to get the formula associated with $H'_\omega$.
-
-\input{movies/snippets/Hprime/HprimeOmega}
-
-Before going further, we prove a useful rewriting lemma:
-
-\input{movies/snippets/Hprime/HprimePlusFin}
-
-Then, we get easily formulas for $H'_{\omega+i}$, and $H'_{\omega\times i}$ for any natural number $i$.
-
-\input{movies/snippets/Hprime/HprimeExamplesa}
-\input{movies/snippets/Hprime/HprimeExamplesb}
-\input{movies/snippets/Hprime/HprimeExamplesc}
-\input{movies/snippets/Hprime/HprimeExamplesd}
-
-Since $\omega^2$ is a limit ordinal, we prove the following equality: 
-\[H'_{\omega^2} (k) = 2 ^ {k+1} \times (k+1) - 1\]
-
-\input{movies/snippets/Hprime/HprimeOmegaSqr}
-
-\subsubsection{New limits}
-
-Our next step would be to prove an exact formula for $H'_{\omega^\omega}(k)$.
-Since the canonical sequence of $\omega^\omega$ is composed of all the
-$\omega^i$, we first need to express $H'_{\omega^i}$ for any natural number $i$.
-
-Let $i$ and $k$ be two natural numbers. 
-The ordinal $\canonseq{\omega^(i+1)}{k}$ is the product
-$\omega^i \times k$, so we need also to consider ordinals of this form.
-
-\begin{enumerate}
-\item First,  we express $H'_{\omega^\alpha \times (i+2)}$ in terms of
-$H'_{\omega^\alpha \times (i+1)}$.
-
-\input{movies/snippets/Hprime/HprimeOmegaTerm1}
-
-\item
-Then, we prove by induction on $i$ that $H'_{\omega^\alpha \times (i+1)}$ is just the
-$(i+1)$-th iterate of $H'_{\omega^\alpha}$.
-
-\input{movies/snippets/Hprime/HprimeOmegaTerm}
-
-\item In particular, we derive a formula for $H'_{\omega^{i+1}}$.
-
-\input{movies/snippets/Hprime/HprimeSuccFun}
-\input{movies/snippets/Hprime/HprimePhi0SI}
-
-\item We get now a  formula for $H'_{\omega^3}$:
-
-\input{movies/snippets/Hprime/HprimeOmegaCube}
-\end{enumerate}
-
-\subsubsection{A numerical example}
-\label{sect:bignum}
-
-It is hard to capture the complexity of this function by looking only at this
-``exact'' formula. 
-Let us consider a simple example: the number $H'_{\omega^3}(3)$.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3a}
-
-Thus, the number $H_{\omega^3}(3)$ can be written as four nested applications of $f$.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3b}
-
-In a more classical writing, this number is displayed as follows:
-
-{\Large
-$$
-H'_{\omega^3}(3) =  2 ^ {(2 ^ {N + 1} \, (N+1) )}   \,  (2 ^ {N+1} \, ( N +1) ) - 1
-$$
-}
-
-In order to make this statement more readable, we introduce a local definition.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3c}
-
-This number looks quite big; let us compute an approximation with \texttt{Ocaml}:
-
-\begin{Coqsrc}
-# (2.0 ** 64.0 *. 64.0 -. 1.0);; 
-\end{Coqsrc}
-
-\begin{Coqanswer}
-- : float = 1.1805916207174113e+21
-\end{Coqanswer}
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3d}
-
-We leave as an exercise to determine the best approximation as possible of
- the size of this number (for instance its number of digits).  For instance, if
-we do not take into account the multiplications in the formula above,
-we obtain that, in base $2$, the number $H'_{\omega^3}(3)$ has at least
-$2^{10^{21}}$  digits. But it is still an under-approximation !
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3e}
-
-\subsubsection{A formula for \texorpdfstring{$H'_{\omega^\omega}$}{\texttt{H'(phi0(omega))}}}
-
-Now, we can get at last an exact formula for $H'_{\omega^\omega}$.
-
-\input{movies/snippets/Hprime/HprimePhi0Omega}
-
-Using extensionality of the functional \texttt{iterate}, we also get a closed formula.
-
-\input{movies/snippets/Hprime/HprimePhi0OmegaClosed}
-
-Note that this formula contains two occurrences of the functional \texttt{iterate}, the outer one is in fact a second-order iteration (on type \texttt{nat -> nat)}
-and the inner one  first-order (on type \texttt{nat}).
-
-\subsection{Abstract properties of  
-\texorpdfstring{$H'_\alpha$}{H'}}
-~\label{sect:H-alpha-prop}
-
-Since pure computation seems to be useless for dealing with expressions of the form $H'_\alpha(k)$, even for small values of $\alpha$ and $k$, we need to prove theorems for comparing $H'_\alpha(k)$ and $H'_\beta(l)$, in terms of comparison
-between $\alpha$ and $\beta$ on the one hand, $k$ and $l$ on the other hand.
-
-But beware of fake theorems! For instance, one could believe that $H'$ is monotonous in its first argument. The following proof shows this is false.
-
-\input{movies/snippets/Hprime/HprimeNonMono1}
-
-On the contrary, the functions of the $H'$ hierarchy have the following five properties~\cite{KS81}: for any $\alpha < \epsilon_0$,
-\begin{itemize}
-\item the function $H'_\alpha$ is strictly monotonous :
-      For all $n,p \in\mathbb{N}, n < p \Rightarrow H'_\alpha(n)< H'_\alpha(p)$.
-\item If $\alpha \not= 0$, then for every $n$, $n<H'_\alpha(n)$.
-\item The function $H'_\alpha$ is pointwise less or equal than $H'_{\alpha+1}$
-
-\item For any $n\geq 1$, $H'_\alpha(n)<H'_{\alpha+1}(n)$.
-\emph{We say that $H'_{\alpha+1}$ dominates $H'_\alpha$ from $1$}.
-\item For any $n$ and $\beta$, if $\alpha \xrightarrow[n]{} \beta$, then
-$H'_\beta(n)\leq H'_\alpha(n)$.
-\end{itemize}
-
-\index{maths}{Abstract properties of arithmetic functions}
-\index{hydras}{Abstract properties of arithmetic functions}
-
-\index{maths}{Transfinite induction}
-
-In \coq{}, we follow the  proof of~\cite{KS81}. This proof is mainly a single  proof by transfinite induction on $\alpha$ of the conjunction of the five properties.
-For each $\alpha$, the three cases : $\alpha=0$, $\alpha$ is a limit, and 
-$\alpha$ is a successor are considered. Inside each case, the five sub-properties are proved sequentially, using the abstract properties defined in Sect~\vref{sect:abstract-arith-prop}
-
-\input{movies/snippets/Hprime/PAP}
-
-Using a few lemmas \emph{à la} Ketonen-Solovay, we prove that
-if $\alpha<\beta$, then $H'_\beta$ eventually dominates
-$H'_\alpha$.
-We let the reader look at the proof (Section \texttt{Proof\_of\_H'\_mono\_l} of \href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}).
-
-\noindent
-\input{movies/snippets/Hprime/HprimeDom}
-
-%ICI
-\subsection{Comparison between \texttt{L\_} and \texttt{H'\_} }
-
-By well-founded induction on $\alpha$, we prove that
- $H'_\alpha(i) \leq L_\alpha(i+1)$ for any $i$.
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#H'_L_}{Epsilon0.L\_alpha}}
-
-\input{movies/snippets/L_alpha/HprimeL}
-
-\subsubsection{Back to hydras}
-
-The following theorem relates the length of (standard) battles with the the $H'$ family of fast growing functions.
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems}}
-
-\input{movies/snippets/Hydra_Theorems/battleLengthStdHardy}
-
-Since $H'_{\omega^3}(3)= H'_{\omega^3+3}(0)$, the big number shown in Section~\vref{sect:bignum} is less or equal than 
-the number of steps of the battle with the initial hydra of
-Figure~\ref{fig:start}, and  $0$ as initial replication factor.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[very thick, scale=0.3]
-  \node (h1) at (4,0){$\bullet$};
-  \node (h2) at (2,2){$\bullet$};
-  \node (h3) at (0,4){$\bullet$};
-  \node (h4) at (1,4){$\bullet$};
-  \node (h5) at (2,4){$\bullet$};
-  \node (h6) at (4.5,2){$\bullet$};
-  \node (h7) at (6.5,2){$\bullet$};
-  \node (h8) at (8.5,2){$\bullet$};
-    \draw (h1) -- (h2) ;
-    \draw (h2) -- (h3) ;
-    \draw (h2) -- (h4) ;
-    \draw (h2) -- (h5) ;
-    \draw (h1) -- (h6) ;
-    \draw (h1) -- (h7) ;
-    \draw (h1) -- (h8);
- \end{tikzpicture}
-
-  \caption{The hydra corresponding to the ordinal $\omega^3+3$}
-  \label{fig:start}
-\end{figure}
-
-\section{A variant of the Wainer hierarchy (functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
-\label{sect:wainer}
-
-\index{maths}{Rapidly growing functions!Wainer Hierarchy}
-
-Ketonen and Solovay introduce in~\cite{KS81} a ``trivial'' variant of the Wainer hierarchy~\cite{BW85, Wainer1970} of fast growing functions, indexed by ordinals below $\epsilon_0$.
-The functions $F_\alpha$ are defined by the following equations.
-
-\label{F_equations}
-\begin{itemize}
-\item $F_0(i)=i+1$
-\item $F_{\beta+1}(i)= (F_\beta)^{(i+1)}(i)$, where $f^{(i)}$ is the $i$-th iterate of $f$.
-\item $F_\alpha(i) = F_{\canonseq{\alpha}{i}} (i)$ if $\alpha$ is a limit ordinal.
-\end{itemize}
-
-\begin{remark}
-The difference with the ``classic'' Wainer hierarchy 
-$f_\alpha\;(\alpha<\epsilon_0)$ lies in the second equation:
-$f_{\beta+1}(i) = (f_\beta)^{(i)}(i)$ and not
-$f_{\beta+1}(i) = (f_\beta)^{(i+1)}(i)$.
-
-A module about 
-the classic Wainer hierarchy is in preparation.
-
-Note also that \cite{KS81} defines also $F_{\epsilon_0}$ (by the third equation).
-Since $\epsilon_0$ is not representable in type \texttt{E0}, our implementation in \coq{} does not take $F_{\epsilon_0}$ into account.
-
-\end{remark}
-
-A first attempt is to write a definition of $F_\alpha$ by equations, in the same way as for $H_\alpha$ (the functional \texttt{iterate} has already been used in Sect.\vref{Functions:iterate}).
-
-\index{hydras}{Library Prelude!iterate}
-
-\input{movies/snippets/Iterates/iterateDef}
-
-\emph{From\href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}~}.
-
-\index{coq}{Plug-ins!Equations}
-
- \input{movies/snippets/F_alpha/FailDemo}
-
-We presume that this error comes from the recursive call of \texttt{F\_} inside
-an application of \texttt{iterate}. The workaround we propose is to define first 
-the iteration of \texttt{F\_}  as an helper $F^*$, then to define the function $F$ as a ``iterating $F^*$ once''.
-
-\texttt{Equations} accepts the following definition, relying on  lexicographic ordering on pairs $(\alpha,n)$.
-
-\label{sect:F-equations}
-
-\index{coq}{Plug-ins!Equations}
-\label{Functions:F-alpha}
-\index{maths}{Rapidly growing functions}
-\index{hydras}{Library Epsilon0!Functions!F\_@F\_ (Wainer hierarchy)}
-
-\input{movies/snippets/F_alpha/goodDefa}
-\input{movies/snippets/F_alpha/goodDefb}
-
-It is quite easy to prove that our functional \texttt{F\_} satisfies the equations on page~\pageref{sect:F-equations}.
-\index{hydras}{Library Prelude!iterate}
-
-\input{movies/snippets/F_alpha/FEquations}
-
-As for the Hardy functions, we can use these equalities as rewrite rules for
-``computing'' some values of $F_\alpha(i)$, for small values of $\alpha$.
-
-\input{movies/snippets/F_alpha/FirstValues}
-
-Like in Sect~\ref{sect:H-alpha-prop}, we prove by induction the following properties (see~\cite{KS81}).
-
-\input{movies/snippets/F_alpha/FalphaThms}
-
-As a corollary, we prove that, if $\beta<\alpha$, then
-$F_\alpha$ dominates $F_\beta$ (p. 284 of~\cite{KS81}).
-
-\input{movies/snippets/F_alpha/FDomContext}
-\input{movies/snippets/F_alpha/FDom}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove the following property:
-
-\begin{Coqsrc}
-Lemma LF3 : dominates_from  2 (F_ 3) (fun  n => iterate exp2 n n).
-\end{Coqsrc}
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
-\end{exercise}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that, for any $\alpha\geq 3$ and $n\geq 2$,
-$F_\alpha(1+n)\geq 2^{F_\alpha(n)}$.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
-\end{exercise}
-
-%ici
-\index{hydras}{Exercises}
-
-\begin{exercise}
-It is tempting to prove a simple property of monotony 
-of the function \texttt{F\_}.
-
-\begin{quote}
-   Let $\alpha\leq\beta<\epsilon_0$. For any $n\geq 2$,
-$F_\alpha(n)\leq F_\beta(n)$. 
-\end{quote}
-Prove or disprove this statement.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/is_F_monotonous.v}{exercises/ordinals/is\_F\_monotonous.v}.}
-\end{exercise}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-
-Prove that for any $n\geq 2$, $\textrm{Ack}\,\,n\,n\leq  F_\omega(n)$, where \textrm{Ack} is the Ackermann function.
-Next, prove that $F_\alpha$ is not primitive recursive, for any $\alpha\geq\omega$  (please see Sect.~\vref{sect:ack-not-PR}).
-On the other hand, please show that for any natural number $n$, the function $F_n$ is primitive recursive.
-Thus $F_\alpha$ is primitive recursive if and only if $\alpha$ is finite.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_omega.v}{exercises/ordinals/F\_Omega.v}.
-Properties of the Ackermann function are studied in
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/Ack.v}{theories/ordinals/MoreAck/Ack.v} and
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/AckNotPR.v}{theories/ordinals/MoreAck/AckNotPR.v}
-.}
-\end{exercise}
-
-\paragraph*{\gaiasign} Library~\href{../theories/html/gaia_hydras.GF_alpha.html}{gaia\_hydras.GF\_alpha}  defines the functions $F_\alpha$ for \gaia's world.
-
-\section{More about rapidly growing functions}
-
-In Sect.~\vref{sect:battle-length-notPR}, we prove that the length of hydra-battles (for a given hydra, according to the initial replication factor) is not primitive recursive in general. 
-This proof uses properties of the Ackermann function, and the $H'_\alpha$, $F_\alpha$, $L_\alpha$ families of functions.
-
-% ---------------------------------------------------
-\input{gaia-chapter}
-
-%----------------------------------------------------------------------
-\chapter[Countable ordinals (after Sch\"{u}tte)]{Kurt Schütte's axiomatic definition of countable ordinals}
-
-\label{chap:schutte} 
-%ON
-
-In the present chapter, we  compare our implementation of the segment $[0,\epsilon_0)$ with a mathematical text in order to ``validate'' our constructions.
-Our reference here is the axiomatic definition of the set of countable ordinals,
-in chapter V of Kurt Schütte's book `` Proof Theory ''~\cite{schutte}.
-
-\begin{remark}
-\emph{In all this chapter, the word ``ordinal'' will be considered as a synonymous of
-``countable ordinal''}  
-\end{remark}
-
-Schütte's definition of countable ordinals relies on the following three axioms:
-
-There  exists a strictly ordered set , such that
-\begin{enumerate}
-\item  $(\mathbb{O},<)$ is well-ordered
-\item Every bounded subset of $\mathbb{O}$  is countable
-\item Every countable subset of $\mathbb{O}$  is bounded.
-\end{enumerate}
-
-Starting with these three axioms, Schütte re-defines the vocabulary about ordinal numbers: the null ordinal $0$, limits and successors, the addition of ordinals, the infinite ordinals $\omega$, $\epsilon_0$, $\Gamma_0$, etc.
-
-This chapter describes an adaptation to \coq{} of Schütte's axiomatization. 
- Unlike the rest of our libraries, our library
-\href{../theories/html/hydras.Schutte.Schutte.html}{hydras.Schutte}
-is not constructive, and relies on several axioms.
-
-\begin{itemize}
-\item First, please keep in mind  that the set of countable ordinals is not countable.
-Thus, we cannot hope to represent all countable ordinals as finite terms of an inductive type, which was possible with  the set of ordinals strictly less than $\epsilon_0$ (resp.
-$\Gamma_0$)
-\item We tried to be as close as possible to K.
-Schütte's text, which uses ``classical'' mathematics : excluded middle, Hilbert's $\epsilon$ (choice) and Russel's $\iota$ (definite description) operators.
-Both operators allow us to write definitions close to the natural mathematical language, such as ``$\textrm{succ}$ is \emph{the} least ordinal strictly greater than $\alpha$''
-\item Please note that only the library \href{../theories/html/hydras.Schutte.Schutte.html}{Schutte/*.v} is ``contaminated'' by axioms, and that the rest of our libraries remain constructive.
-\end{itemize}
-
-\section{Declarations and axioms}
-
-Let us declare a type 
-\texttt{Ord} for representing countable ordinals, and a binary relation
- \texttt{lt}. Note that, in our development, \texttt{Ord} is a type, while the \emph{set} of countable ordinals (called $\mathbb{O}$ by Schütte) 
-is the full set over the type \texttt{Ord}.
-
-\label{types:Ord}
-
-We use Florian Hatat's library on countable sets, written as he was a student of  \emph{\'Ecole Normale Supérieure de Lyon}.
-A set $A$ is countable if there is an injective function from $A$ to $\mathbb{N}$ (see
-Library \href{../theories/html/hydras.Schutte.Countable.html}%
-{\texttt{Schutte.Countable}}).
-
-\vspace{6pt}
-
-\emph{From Module\href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{Schutte.Schutte\_basics}}}
-
-\index{hydras}{Library Schutte!Types!Ord}
-
-\input{movies/snippets/Schutte_basics/OrdDecl}
-
-Schütte's first axiom tells that \texttt{lt} is a well order on the set 
-\texttt{ordinal} (The  class \texttt{WO} is defined in
-Module~\href{../theories/html/hydras.Schutte.Well_Orders.html}{Schutte.Well\_Orders.v}).
-
-\index{hydras}{Library Schutte!Type classes!WO@ WO (well order)}
-
-\label{types:WO}
-
-\input{movies/snippets/Well_Orders/Mdecl}
-
-\input{movies/snippets/Well_Orders/WODef}
-
-\input{movies/snippets/Schutte_basics/AX1}
-
-The second and third axioms say that a subset $X$ of $\mathbb{O}$ is
-(strictly) bounded if and only if it is countable.
-
-\input{movies/snippets/Schutte_basics/AX23}
-
-\texttt{AX2} and \texttt{AX3} could have been replaced by a single axiom (using the \texttt{iff} connector), but we decided to respect as most as possible the structure of Schütte's definitions.
-
-\section{Additional  axioms}
-
-The adaptation of Schütte's mathematical discourse to \coq{} led us to
-import a few axioms from the standard library. We encourage the reader to consult \coq{}'s FAQ about the safe use of axioms
- \url{https://github.com/coq/coq/wiki/The-Logic-of-Coq#axioms}.
-
-\subsubsection{Classical logic}
-
-In order to work with classical logic, we import the module
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Classical.html}{Coq.Logic.Classical}  of \coq{}'s standard library, specifically the following axiom:
-
-\begin{Coqsrc}
- Axiom classic : forall P:Prop, P \/ ~P.
-\end{Coqsrc}
-
-\subsubsection{Description operators}
-
-In order to respect Schütte's style, we imported also the library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Epsilon.html}{\texttt{Coq.Logic.Epsilon}}.  The rest of this section presents a few examples of
-how Hilbert's choice operator and Church's definite description allow us
- to write understandable definitions (close to the mathematical natural language).
-
-\subsubsection{The definition of zero}
-
-According to the  definition of a well order, every non-empty subset of \texttt{Ord} has a least element. Furthermore, this least element is unique. We would like to call this element  \texttt{zero}.
-
-\input{movies/snippets/Schutte_basics/iotaDemoa}
-\input{movies/snippets/Schutte_basics/iotaDemob}
-
-Indeed, the basic logic of  \coq{} does not allow us to eliminate a proof of a proposition 
-$\exists!\,x:A,\,P(x)$ for building a term whose type lies in the sort \texttt{Type}. 
-The reasons for this impossibility are explained in many documents~\cite{BC04, chlipalacpdt2011, Coq}.
-
-Let us import the library \texttt{Coq.Logic.Epsilon}, which contains the following axiom and lemmas.
-
-\input{movies/snippets/MoreEpsilonIota/EpsilonStatement}
-
-Hilbert's $\epsilon$ \emph{operator} is derived from this  axiom.
-
-\input{movies/snippets/MoreEpsilonIota/EpsilonDef}
-
-If we consider the \emph{unique existential} quantifier $\exists!$, we obtain
-Church's \emph{definite description operator}.
-
-\input{movies/snippets/MoreEpsilonIota/iotaDef}
-
-Indeed, the operators \texttt{epsilon} and \texttt{iota} allowed us to make our definitions 
-quite close to Schütte's text. Our libraries \href{../theories/html/hydras.Schutte.MoreEpsilonIota.html}%
-{\texttt{Schutte.MoreEpsilonIota}}
-and
-\href{../theories/html/hydras.Schutte.PartialFun.html}%
-{\texttt{Schutte.PartialFun}} are extensions of \texttt{Coq.logic.Epsilon} for making easier 
-such definitions. See also an article in french~\cite{PCiota}.
-
-\input{movies/snippets/MoreEpsilonIota/Defs}
-
-In order to use these tools,  we have to tell \coq{}  that the declared type \texttt{Ord} is not empty:
-
-\input{movies/snippets/Schutte_basics/inhOrd}
-
-We are now able to define \texttt{zero} as the least ordinal. For this purpose,
-we define a function returning the least element of any [non-empty]  subset.
-
-\emph{From Module\href{../theories/html/hydras.Schutte.Well_Orders.html}%
-  {\texttt{Schutte.Well\_Orders}}}
-
-\input{movies/snippets/Well_Orders/MDecl}
-\input{movies/snippets/Well_Orders/theLeast}
-
-\vspace{4pt}
-
-From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{~Schutte.Schutte\_basics}}
-
-\label{Constants:zero:Ord}
-\index{hydras}{Library Schutte!Constants!zero}
-
-\input{movies/snippets/Schutte_basics/zeroDef}
-
-We want to prove now that zero is less than or equal to any ordinal number.
-
-\input{movies/snippets/Schutte_basics/zeroLe}
-
-\subsubsection{Remarks on \texttt{epsilon} and \texttt{iota}}
-
- What would happen in case of a misuse of \texttt{epsilon} or \texttt{iota} ?
-For instance, one could give a unsatisfiable specification to \texttt{epsilon} or 
-a specification for \texttt{iota} that admits several realizations.
-
-Let us consider an example:
-
-\input{movies/snippets/Schutte_basics/BadBottoma}
-
-Since we won't be able to prove the proposition
-\linebreak \Verb|(exists! a: Ord, least_member (Empty_set Ord) a)|, the only properties we would be able to prove about \texttt{bottom} would be \emph{trivial} properties, 
-\emph{i.e.}, satisfied by \emph{any} element of type \texttt{Ord}, like for instance
-\texttt{bottom = bottom}, or \texttt{zero <= bottom}.
-
-\input{movies/snippets/Schutte_basics/trivialProps}
-
-On the other hand, the following attempt fails, because of the unprovable first subgoal (please notice that the second subgoal is easy to solve !).
-
-\input{movies/snippets/Schutte_basics/Failure}
-
-In short, using \texttt{epsilon} and \texttt{iota} in our implementation of countable ordinals after Schütte has two main advantages.
-
-\begin{itemize}
-\item It allows us to give a \emph{name} (using \texttt{Definition}) two witnesses 
-of existential quantifiers (let us recall that, in classical logic, one may consider non-constructive proofs of existential statements)
-\item By separating definitions from proofs of [unique] existence, one may make definitions  more concise and readable. Look for instance at 
-the definitions of  \texttt{zero}, \texttt{succ}, \texttt{plus}, etc. in the rest of this chapter.
-\end{itemize}
-%%%% ICI ICI
-
-\section{The  successor function}
-
-The definition of the function \texttt{succ:Ord -> Ord} is very concise. The successor of any ordinal $\alpha$ is the smallest ordinal strictly greater than $\alpha$.
-
-\label{Functions:succ-sch}
-\index{hydras}{Library Schutte!Functions!succ}
-
-\input{movies/snippets/Schutte_basics/succDef}
-
-Using \texttt{succ}, we define the following predicates.
-
-\input{movies/snippets/Schutte_basics/isSuccIsLimit}
-
-% \begin{remark}
-% Please look at remark~\vref{warning:coercions}.
-
-How do we prove properties of the successor function?
-First, we make its specification explicit.
-
-\input{movies/snippets/Schutte_basics/succSpec}
-
-Then, we prove that our function \texttt{succ} meets this specification.
-
-\input{movies/snippets/Schutte_basics/succOka}
-
-We have now to prove that the set of all ordinals strictly greater than $\alpha$ has a unique least element.
-But the singleton set $\{\alpha\}$ is countable, hence  bounded (by the axiom \texttt{AX3}).
-Hence; the set $\{\beta\in\mathbb{O}|\alpha < \beta\}$ is not empty
-and therefore has a unique least element.
-
-The rest of the \coq{} proof script is quite short.
-
-\input{movies/snippets/Schutte_basics/succOkb}
-
-We can ``uncap'' the description operator for proving properties of the
-\texttt{succ} function.
-
-\input{movies/snippets/Schutte_basics/succProps}
-
-\section{Finite ordinals}
-
-Using \texttt{succ}, it is now easy to define recursively all the finite ordinals.
-
-\label{sect:notation-F-sch}
-
-\input{movies/snippets/Schutte_basics/finiteDef}
-
-\section{The definition of \texttt{omega}}
-In order to define $\omega$, the first infinite ordinal, we use an operator which
-``returns'' the least upper bound (if it exists) of a subset $X\subseteq \mathbb{O}$.
-For that purpose, we first use a predicate:
-(\texttt{is\_lub $D$ \textit{lt} $X$ $a$}) if $a$ belongs to $D$ and is the least 
-upper bound  of $X$ (with respect to \textit{lt}).
-
-\input{movies/snippets/Lub/isLubDef}
-
-\input{movies/snippets/Schutte_basics/supDef}
-
-Then, we define the function \texttt{omega\_limit} which returns the least upper bound 
-of the  (denumerable) range of any sequence \texttt{s: nat -> Ord}. 
-By \texttt{AX3} this range is bounded, hence the set of its upper bounds is not empty and has a least element.
-Then we define \texttt{omega} as the limit of the sequence of finite ordinals.
-
-\input{movies/snippets/Schutte_basics/omegaDef}
-
-\label{sect:notation-omega}
-
-Among the numerous properties of the ordinal $\omega$, let us quote the following ones
-(proved in Module 
-\href{../theories/html/hydras.Schutte.Schutte_basics.html\#finite_lt_omega}{\texttt{Schutte.Schutte\_basics}})
-
-\input{movies/snippets/Schutte_basics/omegaPropsa}
-\input{movies/snippets/Schutte_basics/omegaPropsb}
-\input{movies/snippets/Schutte_basics/omegaPropsc}
-\input{movies/snippets/Schutte_basics/omegaPropsd}
-
-\subsection{Ordering functions and ordinal addition}
-
-After having defined the finite ordinals and the infinite ordinal $\omega$, we  define the sum $\alpha+\beta$ of two countable ordinals.
-Schütte's definition looks like the following one:
-
-\begin{quote}
-``$\alpha+\beta$ is the $\beta$-th ordinal greater than or equal to $\alpha$''
-\end{quote}
-
-The purpose of this section is to give a meaning to the construction
-``the $\alpha$-th element of $X$''  where $X$ is any non-empty subset of $\mathbb{O}$.
-We follow Schütte's approach, by defining the notion of \emph{ordering functions},
-a way to associate a unique ordinal to each element of $X$.
-Complete definitions and proofs can be found in Module
- \href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
-{\texttt{Schutte.Ordering\_Functions}} ).
-
-\subsection{Definitions}
-
-A \emph{segment} is a set $A$ of ordinals such that, whenever  $\alpha\in A$ and
-$\beta<\alpha$, then $\beta\in A$; a segment is  \emph{proper} if it strictly included in $\mathbb{O}$.
-
-\input{movies/snippets/Ordering_Functions/segmentDef}
-
-Let  $A$ be a segment, and $B$ a subset of $\mathbb{O}$ : an \emph{ordering function for $A$ and  $B$} is a strictly increasing bijection from $A$ to $B$.
-The set $B$ is said to be an \emph{ordering segment} of $A$.
-Our definition in \coq{} is a direct translation of the mathematical text of~\cite{schutte}.
-
-\index{maths}{Ordinal numbers!Ordering functions}
-\index{hydras}{Library Schutte!Predicates!ordering function@ordering\_function}
-
-\input{movies/snippets/Ordering_Functions/orderingFunctionDef}
-
-We are now able to associate with any subset $B$ of $\mathbb{O}$ its ordering segment and ordering function.
-
-\input{movies/snippets/Ordering_Functions/ordDef}
-
-Thus (\texttt{ord $B \;\alpha$}) is the $\alpha$-th element of $B$.
-Please note that the last definition uses the epsilon-based operator \texttt{some} and
-not \texttt{the}. This is due to the fact that we cannot prove the unicity (w.r.t. Leibniz' equality) of the ordering function of a given set. 
-By contrast, we admit the axiom  \texttt{Extensionality\_Ensembles}, from the library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Sets.Ensembles.html}{Coq.Sets.Ensembles}, so we use the operator \texttt{the} in the definition of
-\texttt{the\_ordering\_segment}.
-
-One of the main theorems of
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html\#ordering_function_ex}%
-{\texttt{Ordering\_Functions}} 
-associates a unique segment and a unique (up to extensionality) ordering function to every subset $B$ of $\mathbb{O}$.
-
-\input{movies/snippets/Ordering_Functions/orderingFunctionEx}
-
-Thus,  our function \texttt{ord}  which enumerates the elements of $B$ is defined in a non-ambiguous way.
-Let us quote the following theorems (see Library
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
-{\texttt{Schutte.Ordering\_Functions}} for more details).
-
-\input{movies/snippets/Ordering_Functions/orderingLe}
-\input{movies/snippets/Ordering_Functions/Th1352}
-
-\subsection{Ordinal addition}
-
-We are now ready to define and study addition on the type \texttt{Ord}.
-The following definitions and proofs can be consulted in Module
-\href{../theories/html/hydras.Schutte.Addition.html}%
-{\texttt{Schutte.Addition.v}}.
-
-\index{hydras}{Library Schutte!Functions!plus}
-
-\input{movies/snippets/Addition/additionDef}
-
-In other words,  $\alpha + \beta$ is the  $\beta$-th ordinal greater than or equal to $\alpha$. 
-Thanks to generic properties of ordering functions, we can show the following 
-properties of addition on $\mathbb{O}$. First, we prove a useful lemma:
-
-\input{movies/snippets/Addition/plusElim}
-
-As a use-case, let us prove that $0$ is a right neutral element of $+$.
-
-\input{movies/snippets/Addition/alphaPlusZero}
-
-The following lemmas are proved the same way.
-
-\input{movies/snippets/Addition/bunchOfLemmas}
-
-The following lemmas are not direct applications of \texttt{plus\_elim}.
-
-\input{movies/snippets/Addition/plusAssoc}
-\input{movies/snippets/Addition/finitePlusInfinite}
-
-It is interesting to compare the proof of these lemmas with the
-computational proofs of the corresponding statements in Module
-\href{../theories/html/hydras.Epsilon0.T1.html}%
-{\texttt{Epsilon0.T1}}. 
-For instance, the proof of the lemma 
-\texttt{one\_plus\_omega} uses the continuity of ordering functions (applied to  \texttt{(plus 1)}) and compares the limit of the $\omega$-sequences $i_{(i \in \mathbb{N})}$ and
-$(1+i)i_{(i \in \mathbb{N})}$, whereas in the library  \texttt{Epsilon0/T1}, the equality 
-$1+\omega=\omega$ is just proved with \texttt{reflexivity}!
-
-\subsubsection{Multiplication by a natural number}
-
-The multiplication of an ordinal by a natural number is defined in terms of addition.
-This operation is useful for the study of Cantor normal forms.
-
-\input{movies/snippets/Addition/multFin}
-
-\section{The exponential of basis \texorpdfstring{$\omega$}{omega}}
-
-In this section, we define the function which maps any $\alpha\in\mathbb{O}$ to
-the ordinal  $\omega^\alpha$, also written 
-$\phi_0(\alpha)$. 
-It is an opportunity to apply the definitions and results of the preceding section. 
-Indeed,  Schütte first defines a subset of $\mathbb{O}$: the set of additive principal ordinals, and $\phi_0$  is just defined as the ordering function of this set.
-
-\subsection{Additive principal ordinals}
-
-\index{maths}{Ordinal numbers!Additive principal ordinals}
-\index{hydras}{Library Schutte!Predicates!AP@AP (additive principal ordinals)}
-
-\begin{definition}
-A non-zero ordinal  $\alpha$ is said to be \emph{additive principal} if, for all  $\beta<\alpha$, $\beta+\alpha$ is equal to  $\alpha$.
-We call \texttt{AP} the set of additive principal ordinals.
-
-\end{definition}
-
-\noindent\emph{From Module \href{../theories/html/hydras.Schutte.AP.html}%
-{\texttt{Schutte.AP}}}
-
-\input{movies/snippets/AP/APDef}
-
-\subsection{The function \texttt{phi0}}
-
-Let us call  $\phi_0$ the ordering function of \texttt{AP}.
-In the mathematical text, we shall use indifferently the notations  $\omega^\alpha$ and$\phi_0(\alpha)$.
-
-\index{hydras}{Library Schutte!Functions!phi0}
-
-\input{movies/snippets/AP/phi0Def}
-
-\subsection{Omega-towers and the ordinal \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-Using $\phi_0$, we can define recursively the set of finite omega-towers.
-
-\input{movies/snippets/AP/omegaTower}
-
-\label{sect:epsilon0-as-limit}
-Then, the ordinal  $\epsilon_0$ is defined as the limit of the sequence of all finite towers (a kind of infinite tower).
-
-\input{movies/snippets/AP/epsilon0Def}
-
-The rest of our library \texttt{AP} is devoted to the proof of properties of additive principal ordinals, hence of the ordering function  $\phi0$ and the ordinal $\epsilon_0$ (which we could not express within the type \texttt{T1}).
-
-\subsection{Properties of the set  \texttt{AP}}
-
-The set of additive principal ordinals is not empty: it contains at least the ordinals  $1$ and  $\omega$.
-
-\inputsnippets{AP/APOne,AP/APOne_least_AP,AP/APOne_AP_omega,AP/APOne_omega_second_AP}
-
-The set  \texttt{AP} is  \emph{closed} under addition, and unbounded.
-\label{lemma:AP-plus-closed}
-
-\input{movies/snippets/AP/APPlusClosed}
-
-\input{movies/snippets/AP/APUnbounded}
-
-Finally, \texttt{AP} is (topologically) \emph{closed} and ordered by the segment of all countable ordinals.
-
-\index{hydras}{Library Schutte!Predicates!Closed}
-
-From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{~Schutte.Schutte\_basics}}
-
-\input{movies/snippets/Schutte_basics/ClosedDef}
-\input{movies/snippets/AP/APClosed}
-
-\subsubsection{Properties of the function \texorpdfstring{$\phi_0$}{phi0}}
-
-The ordering function $\phi_0$ of the set \texttt{AP} is defined on the full set $\mathbb{O}$ and is continuous (Schütte calls such a  function  \emph{normal}).
-
-\begin{Coqsrc}
-Theorem normal_phi0 : normal phi0 AP.
-\end{Coqsrc}
-
-The following properties come from  the definition of $\phi_0$ as the ordering function of \texttt{AP}.
-It may be interesting to compare these proofs with the computational ones described in Chapter ~\ref{chap:T1}.
-
-\input{movies/snippets/AP/APPhi0}
-
-\subsection{A last example}
-\label{Ex42-schutte}
-
-Let us prove again the equality $\omega+42+\omega^2= \omega^2$.Let us recall that $\omega^2$ is an abbreviation of $\phi_0(2)$,
-\emph{i.e} the third  additive principal ordinal.
-
-\inputsnippets{Schutte/Ex42a}
-
-Our proof is very different from the computational proof of Sect~\vref{Ex42-E0}.
-By definition of additive principal ordinals, 
-it suffices to prove the inequality $\omega+42< \phi_0(2)$.
-
-\inputsnippets{Schutte/Ex42b}
-
-Since the set \texttt{AP} of additive principals  is closed under addition
-(by Lemma \texttt{AP\_plus\_closed}, page~\pageref{lemma:AP-plus-closed}) , it suffices to prove the inequalities $\omega<\phi_0(2)$ and $42<\phi_0(2)$.
-
-\inputsnippets{Schutte/Ex42d, Schutte/Ex42c, Schutte/Ex42e}
-
-\section{More about \texorpdfstring{$\epsilon_0$}{\texttt{epsilon0}}}
-
-Let us recall that the limit ordinal  $\epsilon_0$ cannot be written within the type \texttt{T1}.
-Since we are now considering the set of all countable ordinals, we can now prove some properties of this ordinal.
-
-We prove the inequality  $\alpha<\omega^\alpha$ whenever $\alpha < \epsilon_0$.
-\emph{Note that this condition was implicit in Module
-\href{../theories/html/hydras.Epsilon0.T1.html\#lt_phi0}{Epsilon0.T1}.}
-
-\input{movies/snippets/AP/ltPhi0}
-
-The proof is as follows:
-\begin{enumerate}
-\item Since $\alpha<\epsilon_0$, consider the least $i$ such that $\alpha$ is strictly less than the omega-tower of height $i$.
-\item
-  \begin{itemize}
-  \item If $i=0$, then the result is trivial (because $\alpha=0$)
- \item  Otherwise let $i=j+1$; 
-          $\alpha$ is greater than or equal to the omega-tower of height $j$.
-         By monotony,  $\phi_0(\alpha)$ is greater than or equal to 
-        the omega-tower of height $j+1$, thus strictly greater than $\alpha$
-  \end{itemize}
- \end{enumerate}
-
-Moreover,  $\epsilon_0$ is the least ordinal $\alpha$ that verifies the equality 
-$\alpha = \omega^\alpha$, in other words, the least fixpoint of the function  $\phi_0$.
-
-\input{movies/snippets/AP/epsilon0Lfp}
-
-\section{Critical ordinals}
-
-\index{maths}{Ordinal numbers!Critical ordinals}
-\index{hydras}{Library Schutte!Predicates!Cr@Cr (critical ordinals)}
-
-For any  (countable) ordinal $\alpha$, the set $\textit{Cr}(\alpha)$ is inductively defined 
-as follows by Schütte (p.81 of~\cite{schutte}).
-
-\begin{quote}
-  \begin{itemize}
-  \item $\textit{Cr}(0)$ is the set \textit{AP} of additive principal ordinals.
-  \item If $0<\alpha$, then $\textit{Cr}(\alpha)$ is the intersection of all the sets of fixpoints of the $\textit{Cr}(\beta)$ for $\beta<\alpha$.
-  \end{itemize}
-\end{quote}
-
-This definition is translated in \coq{} in 
-Module \href{../theories/html/hydras.Schutte.Critical.html}%
-{\texttt{Schutte.Critical}}, as the least fixpoint of a functional.
-
-\input{movies/snippets/Critical/CrDef}
-
-Lets us denote by $\phi_\alpha$ the ordering function of the set $\textit{Cr}(\alpha)$ and by $A_\alpha$ its ordering segment.
-
-\input{movies/snippets/Critical/phiDef}
-
-\label{sect:phi-schutte}
-
-For instance,  we prove that $\textit{Cr}(0)$ is the set of additive principals and that $\epsilon_0$
-belongs to $\textit{Cr}(1)$.
-
-\inputsnippets{Critical/CrZeroAP,
-  Critical/epsilon0Cr1}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
- Prove that $\epsilon_0$ is the least element of $\textit{Cr}(1)$.
-\end{exercise}
-
-\subsection{A flavor of infinity}
-
-The family of the $\textit{Cr}(\alpha)$s is made of infinitely many unbounded (hence infinite) sets.
-Let us quote Lemma 5, p. 82  of~\cite{schutte}:
-\begin{quote}
-  For all $\alpha$, the set $\textit{Cr}(\alpha)$ is closed (for the least upper bound of non-empty countable sets) and unbounded.
-\end{quote}
-
-We prove this result by a transfinite induction on $\alpha$ of the conjunction of  both properties.
-
-\index{maths}{Transfinite induction}
-
-\input{movies/snippets/Critical/Lemma5}
-
-\section{Cantor normal form}
-
-The notion of Cantor normal form is defined for all countable ordinals.
-Nevertheless, note that, contrary to the implementation based on type \texttt{T1},
-the Cantor normal form of an ordinal $\alpha$ may contain $\alpha$ as a 
-sub-term\footnote{This would prevent us from trying to represent Cantor normal forms as finite trees (like in Sect.~\ref{sec:T1-inductive-def})}.
-
-\index{maths}{Ordinal numbers!Cantor normal form}
-\index{hydras}{Library Schutte!Predicates!is\_cnf\_of@is\_cnf\_of (to be a Cantor normal form of}
-
-We represent  Cantor normal forms as lists of ordinals.
-A  list $l$ is a Cantor normal form of a given ordinal $\alpha$ if it satisfies two conditions:
-
-\begin{itemize}
-\item The list  $l$ is sorted (in decreasing order) w.r.t. the order $\leq$
-\item The sum of all the  $\omega^{\beta_i}$ where the $\beta_i$ are the terms of $l$ (in this order) is equal to $\alpha$.
-\end{itemize}
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Schutte.CNF.html\#cnf_t}%
-{\texttt{Schutte.CNF}}}
-
-\input{movies/snippets/CNF/Defs}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction on $\alpha$, we prove that every countable ordinal $\alpha$ 
- has at least a Cantor normal form.
-
-\input{movies/snippets/CNF/cnfExists}
-
-By structural induction on lists, we prove that this normal form is unique.
-
-\inputsnippets{CNF/cnfUnicity,CNF/cnfExUnique}
-
-Finally, the following two lemmas relate  $\epsilon_0$ with Cantor normal forms.
-
-If $\alpha<\epsilon_0$, then the Cantor normal form of $\alpha$ is made of ordinals strictly less than $\alpha$.
-
-\input{movies/snippets/CNF/cnfLtEpsilon0}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Please consider the following statement :
-
-\begin{Coqsrc}
-Lemma cnf_lt_epsilon0_iff : 
- forall l alpha, 
-   is_cnf_of alpha l ->  
-   (alpha < epsilon0 <->  Forall (fun beta =>  beta < alpha) l).
-\end{Coqsrc}
-
-Is it true ?
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/schutte_cnf_counter_example.v}{exercises/ordinals/schutte\_cnf\_counter\_example.v}.}
-\end{exercise}
-
-Finally, the Cantor normal form of $\epsilon_0$ is just $\omega^{\epsilon_0}$.
-
-\input{movies/snippets/CNF/cnfOfEpsilon0}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Implement pages 82 to 85 of~\cite{schutte} (critical, strongly critical, maximal critical ordinals, Feferman's ordinal $\Gamma_0$).
-\end{project}
-
-\begin{remark}
-The sub-directory
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/Gamma0}{theories/oridinals/Gamma0} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form.
-\end{remark}
-
-\section{An embedding of \texttt{T1} into \texttt{Ord}}
-
-Our library 
-\href{../theories/html/hydras.Schutte.Correctness_E0.html}%
-{\texttt{Schutte.Correctness\_E0}} establishes the link between two very different modelizations of ordinal numbers. In other words, it ``validates'' a data structure in terms of
-a classical mathematical discourse considered as a model. 
-First, we define a function from \texttt{T1} into  \texttt{Ord} by structural recursion.
-
-\input{movies/snippets/Correctness_E0/injectDef}
-
-This function enjoys good commutation properties with respect to the main operations which
-allow us to build Cantor normal forms.
-
-\input{movies/snippets/Correctness_E0/commutationLemmas}
-\input{movies/snippets/Correctness_E0/injectPlus}
-\input{movies/snippets/Correctness_E0/injectMultFinR}
-
-% \begin{Coqsrc}
-% Theorem inject_mono (beta gamma : T1) :
-%   T1.lt  beta gamma -> 
-%   T1.nf beta -> T1.nf gamma -> 
-%   inject beta < inject gamma.
-
-% Theorem inject_injective (beta gamma : T1) : nf beta -> nf gamma ->
-%   inject beta = inject gamma -> beta = gamma.
-% \end{Coqsrc}
-
-Finally, we prove that \texttt{inject} is a bijection from the set of all terms of \texttt{T1} in normal form to the set 
-\texttt{members epsilon0} of the elements of \texttt{Ord} strictly less than  $\epsilon_0$.
-
-\input{movies/snippets/Correctness_E0/injectLtEpsilon0}
-\input{movies/snippets/Correctness_E0/embedding}
-
-\subsection{Remarks}
-Let us recall that the library \href{../theories/html/hydras.Schutte.Schutte.html}%
-{\texttt{Schutte}} depends on five \emph{axioms} and lies explicitly in the  
-framework of classical logic with a weak version of the axiom of choice
-(please look at the documentation of
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.ChoiceFacts.html}{\texttt{Coq.Logic.ChoiceFacts}}).
-Nevertheless, the other modules:
-\href{../theories/html/hydras.Epsilon0.Epsilon0.html}%
-{\texttt{Epsilon0}},
-\href{../theories/html/hydras.Hydra.Hydra.html}%
-{\texttt{Hydra}}, et 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}%
-{\texttt{Gamma0}}
-do not import any axioms and are really constructive.
-
-\index{hydras}{Projects}
-\begin{project}
-There is no construction of ordinal multiplication in~\cite{schutte}. 
-It would be interesting to derive this operation from Schütte's axioms,
-and prove its consistence with multiplication in ordinal notations for 
-$\epsilon_0$ and $\Gamma_0$.
-\end{project}
-
-\section{Related work}
-
-In~\cite{grimm:hal-00911710}, José Grimm establishes the consistency between our ordinal notations \texttt{T1} and \texttt{T2} (Veblen normal form) and his implementation
-of ordinal numbers after Bourbaki's set theory.
-
-The Gaia project ~\url{https://github.com/coq-community/gaia} maintains Grimm's  theory of ordinals as part of coq-community on GitHub. Integration
-of the present ordinal theory with Gaia, i.e., relating the different notions of ordinals
-and transferring relevant results, is an interesting project.
-First experiments in that direction are developped in
-the \href{https://github.com/coq-community/hydra-battles/blob/master/theories/gaia/}{theories/gaia/} directory.
-
-\chapter{The Ordinal \texorpdfstring{$\Gamma_0$}{Gamma0} (first draft)}
-
-\emph{This chapter and the files it presents are still very incomplete, considering the impressive properties of $\Gamma_0$~\cite{Gallier91}.
-We hope to add new material soon, and accept contributions!}
-
-\begin{todo}
-  Build a T2Bridge! (note: Gamma0's wellfoundedness is missing in \gaia (?).
-\end{todo}
-
-\section{Introduction}
-We present a notation system for the ordinal $\Gamma_0$, following Chapter V, Section 14 of~\cite{schutte}: ``A notation system for the ordinals $<\Gamma_0$''.
-We try to be as close as possible to Schütte's text and usual practices of \coq{} developments.
-
-The ordinal $\Gamma_0$ is defined in Section 13 of ~\cite{schutte} as the least \emph{strongly critical ordinal}. It is widely known as the \emph{Feferman-Schütte ordinal}.
-
-Section V, 13 of~\cite{schutte} defines \emph{strongly critical} and
-\emph{maximal $\alpha$-critical} ordinals:
-
-\begin{itemize}
-\item $\alpha$ is strongly critical if
-$\alpha$ is $\alpha$-critical,
-\item $\gamma$ is maximal $\alpha$-critical if $\gamma$ is $\alpha$-critical, and, for all $\xi>\alpha$, $\gamma$ is not $\xi$-critical.
-
-\end{itemize}
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Schutte.Critical.html\#strongly_critical}%
-{\texttt{Schutte.Critical}}}
-
-\input{movies/snippets/Critical/Gamma0Def}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that a (countable)  ordinal $\alpha$ is strongly critical iff 
-$\phi_\alpha(0)=\alpha$ (Theorem 13.13 of~\cite{schutte} ). 
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that the set of strongly critical ordinals is unbounded and closed (Theorem 13.14 of~\cite{schutte} ).
-Thus this set is not empty,  hence has a least element.
-Otherwise, the definition of $\Gamma_0$ above would be useless.
-\end{project}
-
-In the present version of this development, we  only study $\Gamma_0$ as a notation system, much more powerful than the ordinal notation for $\epsilon_0$.
-
-%\index{Projects}
-%
-%\begin{project}
-% Schûtte's section 13 of~\cite{schutte} contain several definitions and lemmas which give another view on $\Gamma_0$.  We leave it as a project to implement them in \coq{}.  
-%\end{project}
-
-\section{The type \texttt{T2} of ordinal terms}
-
-The notation system for ordinals less than $\gamma_0$ comes from the following theorem of~\cite{schutte}, where $\psi\,\alpha$ is the ordering function 
-of the set of maximal $\alpha$-critical ordinals.
-
-\index{hydras}{Library Gamma0!Types!T2}
-
-\begin{quote}
-  Any ordinal $\not= 0$ which is not strongly critical can be expressed in terms of $+$ and $\psi$.
-\end{quote}
-
-\index{hydras}{Projects}
-\begin{project}
-This theorem is not formally proved in this development yet. It should be!
-\end{project}
-
-Like in Chapter~\ref{chap:T1}, we define an inductive type with two constructors, one for $0$, the other for the construction $\psi(\alpha,\beta)\times(n+1)+\gamma$, adapting a Manolios-Vroon-like notation~\cite{Manolios2005} to
-\emph{Veblen normal forms}.
-\label{types:T2}
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T2}%
-{\texttt{Gamma0.T2}}}
-
-\inputsnippets{T2/T2Def, T2/psiDef}
-
-Like in chapter~\ref{chap:T1}, we get familiar with the type \texttt{T2} by recognizing simple constructs like finite ordinals, $\omega$, etc., as inhabitants of \texttt{T2}.
-
-\input{movies/snippets/T2/Notations}
-
-\section{A strict order on T2}
-
-Let us define a strict order on type \texttt{T2}. The following definition is 
-an adaptation of Schütte's, taking into account the multiplications by a natural number (inspired by~\cite{Manolios2005}, and also present in \texttt{T1}).
-
-\label{sect:t2-lt-def}
-
-\input{movies/snippets/T2/ltDef}
-
-Seven constructors! In order to get accustomed with this definition, let us look at a small set of examples, covering all the constructors of \texttt{lt}.
-
-\input{movies/snippets/T2/ltExamples}
-
-\index{hydras}{Projects}
-\begin{project}
-Write a tactic that solves automatically goals of the form (\texttt{$\alpha$ t2< $\beta$}), where $\alpha$ and $\beta$ are closed terms of type \texttt{T2}.
-\end{project}
-
-\section{Veblen normal form}
-\begin{definition}
-  A term of the form $\psi(\alpha_1,\beta_1)\times n_1+ \psi(\alpha_2,\beta_2)\times n_2+\dots+\psi(\alpha_k,\beta_k)\times n_k$ is said to be in
- \emph{[Veblen] normal form} if for every $i<n$, $\psi(\alpha_i,\beta_i)<\psi(\alpha_{i+1},\beta_{i+1})$, all the $\alpha_i$ and $\beta_i$ are in normal form, and all the $n_i$ are strictly positive integers.
-\end{definition}
-
-\input{movies/snippets/T2/nfDef}
-
-Let us look at some positive examples (we have to prove some inversion lemmas before proving counter-examples).
-
-\input{movies/snippets/T2/nfExamples}
-
-\subsection{Length of a term}
-
-The notion of \emph{term length} is introduced by Schütte as a helper for proving (at least) the \emph{trichotomy} property and transitivity of the strict order \texttt{lt} on \texttt{T2}.
-These properties are proved by induction on length.
-
-\subsection{Trichotomy}
-
-\emph{Trichotomy} is another name for the well-known property of decidable total ordering (like Standard Library's \texttt{Compare\_dec.lt\_eq\_lt\_dec}).
-
-We first prove by induction on $l$ the following lemma:
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0\#tricho_aux}%
-{\texttt{Gamma0.Gamma0}}}
-
-\input{movies/snippets/Gamma0/tricho}
-
-\input{movies/snippets/Gamma0/compare}
-
-With the help of \texttt{compare}, we get a boolean version of \texttt{nf}
-(being in Veblen normal form).
-
-\input{movies/snippets/Gamma0/nfb}
-
-\section{Main functions on \texttt{T2}}
-
-\subsection{Successor}
-The successor function is defined by structural recursion.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#succ}%
-{\texttt{Gamma0.T2}}}
-
-\input{movies/snippets/T2/succDef}
-
-\subsection{Addition}
-
-Like for Cantor normal forms (see Sect.~\ref{sect:infix-plus-T1}),  the definition of addition in \texttt{T2}  requires comparison between ordinal terms.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0T2.html\#succ}%
-  {\texttt{Gamma0.Gamma0}}}
-
-\input{movies/snippets/Gamma0/plusDef}
-
-\subsection{The Veblen function \texorpdfstring{$\phi$}{\texttt{phi}}}
-
-The enumeration function of critical ordinals, presented in Sect.~\vref{sect:phi-schutte}, is recursively defined in type \texttt{T2}.
-
-\input{movies/snippets/Gamma0/phiDef}
-
-Despite its complexity, the function \texttt{phi} is well adapted to proofs by simplification or computation.
-
-The relation between the constructor $\psi$ and the function $\phi$ is
-studied in~\cite{schutte}, and partially implemented in this development.
-\emph{Please contribute!}
-
-For instance, the following theorem states that, if $\gamma$ is the sum of a limit ordinal $\beta$ and a finite ordinal $n$, and $\beta$ is a fixpoint of
-$\phi(\alpha)$, then $\psi(\alpha,\gamma)=\phi_\alpha(\gamma+1)$.
-
-\input{movies/snippets/Gamma0/phiPsi}
-\input{movies/snippets/Gamma0/Ex9}
-
-On the other hand, $\phi$ can be expressed in terms of $\psi$.
-
-\input{movies/snippets/Gamma0/phiOfPsi}
-\input{movies/snippets/Gamma0/Ex10}
-
-\index{hydras}{Projects}
-\begin{project}
-Please study a way to pretty print ordinal terms in Veblen normal form (see Section~\vref{sect:ppT1}).
-\end{project}
-
-\section{An ordinal notation for \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}}
-
-In order to consider type \texttt{T2} as an ordinal notation, we have to build an instance of class \texttt{ON} (See Definition page~\pageref{types:ON}).
-
-First, we define a type that contains only terms in Veblen normal form, and redefine \texttt{lt} and \texttt{compare} by delegation (see for comparison the construction of type \texttt{E0} in Sect.~\vref{sect:E0-def}).
-
-\input{movies/snippets/Gamma0/G0a}
-\input{movies/snippets/Gamma0/G0b}
-
-Then, we buils an instance of class \texttt{ON}. 
-function \texttt{compare} is correct.
-
-\input{movies/snippets/Gamma0/ltSto}
-\input{movies/snippets/Gamma0/ltWf}
-\input{movies/snippets/Gamma0/ONGamma0}
-\input{movies/snippets/Gamma0/G0z}
-
-\begin{remark}
-The proof of \texttt{lt\_wf} has been written by \'Evelyne Contejean, using her library on the recursive path ordering (see also remark~\vref{remark:a3pat}).
-\end{remark}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that \texttt{Epsilon0} (page~\pageref{instance-epsilon0})
-is a sub-notation system of \texttt{Gamma0}.
-
-Prove that the implementations of \texttt{succ}, \texttt{+}, $\phi_0$, etc.
-are compatible in both notation systems.
-
-Note that a function \texttt{T1\_inj} from \texttt{T1} to \texttt{T2} has already been defined. It may help to complete the task.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T1_to_T2}%
-{\texttt{Gamma0.T2}}}
-
-\input{movies/snippets/T2/T1ToT2}
-
-\end{project}
-
-\begin{project}
-Prove that the notation system \texttt{Gamma0} is a correct implementation 
-of the segment $[0,\Gamma_0)$ of the set of countable ordinals.
-\end{project}

### 2022-05-08 doc/ks-chapter.tex
reorganization of tex files

--- 2022-04-26 doc/part-hydras.tex
@@
-\section*{Introduction}
-
-In this part, we present a development  for the \coq{} proof assistant, after the work of Kirby and Paris~\cite{KP82}. This formalization contains the following main parts:
-
-\begin{itemize}
-\item Representation in \coq{} of hydras and hydra battles.
-\item A proof that every battle is finite and won by Hercules.
-This proof is based on a \emph{variant} which maps any hydra to an ordinal strictly less than $\epsilon_0$ and is strictly decreasing along any battle.
-
-\item Using a combinatorial toolkit designed by J.~Ketonen and R.~Solovay~\cite{KS81}, we prove that, for any ordinal $\mu<\epsilon_0$, there exists no such variant mapping any hydra to an ordinal strictly less than $\mu$.
-Thus, the complexity of $\epsilon_0$ is really needed in the previous proof.
-
-\item We prove a relation between the length of a ``classic''  kind of  battles \footnote{This class is also called \emph{standard} in this document (text and proofs).
-The \emph{replication factor} of the hydra is exactly $i$ at the $i$-th round of the battle (see Sect~\vref{sect:replication-def}).}
-and the Wainer-Hardy hierarchy of ``rapidly growing functions'' $H_\alpha$~\cite{Wainer1970}.
-The considered class of battles, which we call \emph{standard},  is the most considered one in the scientific  literature (including popularization).
-\end{itemize}
-
-Simply put, this document tries to combine the scientific interest of two articles~\cite{KP82, KS81} and a book~\cite{schutte} with the playful activity of truly proving theorems.
-We hope  that such a work, besides exploring a nice piece of discrete mathematics,
-will show how \coq{} and its standard library are well fitted to help us to understand some non-trivial mathematical developments, and also to experiment the constructive parts of  the proof through functional programming.
-
- We also hope to provide a little clarification on infinity (both potential and actual) through the notions of function, computation, limit,
- type and proof.
-
- \subsection*{Compatibility with \gaia (in progress)}
- \label{sect:gaia-first-intro}
-The \gaia project~\cite{Gaia} by Jos\'e Grimm, Alban Quadrat,  and Carlos Simpson,
-aims  to formalize mathematics in \coq  in
-\index{gaiabridge}{Introduction}
-
-the style of Nicolas Bourbaki.
-It contains many definitions and results about ordinal numbers.
-In Chapter~\ref{gaia-chapter}, we present some modules which allow \HydrasLib' users to apply lemmas proven in \gaia, and vice versa.
-Remarks about compatibility with \gaia are signalled with the picture \gaiasign.
-A special index is in construction (page~\pageref{gaia-index}).
-
-%\section{Remarks}
-
-\subsection*{Difference from Kirby and Paris's work}
-In~\cite{KP82}, Kirby and Paris show  that there is no proof of termination of all hydra battles in Peano Arithmetic (PA).
-Since we are used to writing proofs in higher order logic, the restriction to PA was quite unnatural for us.
-So we chose to prove another statement without any reference to PA, by considering a class of proofs indexed by ordinal numbers up to $\epsilon_0$.
-
-\subsection*{State of the development}
-The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on  notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
-We added new material: axiomatic definitions of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, function derinition by equations, etc.
-
-We are now working in order to make clumsy proofs more readable, simplify definitions, and ``factorize'' proofs as much as possible. 
-Many possible improvements are suggested as ``todo''s or ``projects'' in this text.
-
-\section*{Future work (projects)}
-\index{hydras}{Projects}
-
-This document and the proof scripts are far from being complete.
-
-First, there must be a lot of typos to correct, references and index items to add. Many proofs are too complex and should be simplified, etc.
-
-The following extensions are planned, but help is needed:
-
-\begin{itemize}
-\item Semiautomatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
-%\item Extension to $\Gamma_0$ (in Veblen normal form)
-\item More lemmas about hierarchies of rapidly growing
-  functions, and their relationship 
-    with primitive recursive functions and provability in Peano arithmetic 
-(following~\cite{KS81, KP82}).
-\item From \coq's point of view, this development could be used as an illustration of the evolution of the software, every time new libraries and sets of tactics could help to simplify the proofs.
-\end{itemize}
-
-\subsection*{Main references}
-
-In our development, we adapt the definitions and prove many theorems which
-we found in the following articles. 
-\begin{itemize}
-\item ``Accessible independence results for Peano arithmetic''  by Laurie Kirby and Jeff Paris~\cite{KP82}
-\item ''Rapidly growing Ramsey Functions'' by Jussi Ketonen and Robert Solovay~\cite{KS81}
-\item ``The Termite and the Tower'', by Will Sladek~\cite{Sladek07thetermite}
-\item Chapter V of ``Proof Theory'' by Kurt Schütte~\cite{schutte}
-\end{itemize}
-
-\chapter{Hydras and hydra games}
-
-\label{sec:orgheadline91}
-\label{chapter:hydras}
-
-This chapter is dedicated to the representation of hydras and rules of the hydra game in \coq's specification language:~\gallina.
-
-Technically, a \emph{hydra} is just a finite ordered tree, each node of which 
-has any number of sons. Contrary to the computer science tradition, we display the hydras 
-with the heads up and the foot (i.e., the root of the tree) down.
-Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the 
-module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}). 
-\emph{For a less formal description of hydras, please see 
-\url{https://www.smbc-comics.com/comic/hydra}.}
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {$\bullet$};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H4) at (4,2) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N3) to [bend left= 10]  (H1) ;
-\draw (N3) to [bend right= 16] (H2);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (foot) to [bend right = 10] (H4) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-\caption{The hydra \texttt{Hy} \label{fig:Hy}}
-\end{figure}
-
-We use a specific vocabulary for talking about hydras. Table~\ref{tab:hyd2tree} shows the correspondence between our terminology and the usual vocabulary for trees in computer science.
-
-\begin{figure}[h]
-  \centering
-  \begin{tabular}{ll}
-Hydras & Finite rooted trees\\
-\hline
-foot & root\\
-head & leaf\\
-node & node\\
-segment  & (directed) edge \\
-sub-hydra & subtree\\
-daughter & immediate subtree\\
-\end{tabular}
-  \caption{Translation from hydras to trees}
-  \label{tab:hyd2tree}
-\end{figure}
-
-The hydra \texttt{Hy} has a \emph{foot} (below), five \emph{heads}, and eight \emph{segments}. 
-We leave it to the reader to define various parameters such as the height, the size, the highest arity (number of sons of a node) of a hydra.
-In our example, these parameters have the respective values $4$, $9$ and $3$.
-
-\subsection{The rules of the game}
-
-\label{sec:orgheadline44}
-\label{sect:replication-def}
-
-A \emph{hydra battle} is a fight between Hercules and the Hydra. 
-More formally, a  battle is a sequence of \emph{rounds}.
-At each round:
-\begin{itemize}
-\item If the hydra is composed of just one head, the battle is finished
-and  Hercules is the winner.
-\item Otherwise, Hercules chops off \emph{one} head of the hydra,
-
-\begin{itemize}
-\item If the head is at distance 1 from the foot, the head is just lost by the hydra, with no more reaction.
-\item Otherwise, let us denote by \(r\) the node that was at distance \(2\) from 
-the removed head in the direction of the foot,  and consider the  sub-hydra \(h'\) of \(h\), whose  root is \(r\) \footnote{$h'$ will be called ``the wounded part of the hydra'' in the subsequent text.
-In Figures~\vref{fig:Hy2} and ~\vref{fig:Hy4}, this sub-hydra  is displayed in red.}.
-Let $n$ be some natural number.
-Then $h'$ is replaced by  $n+1$ of copies of \(h'\) which share the same root $r$.
- The \emph{replication factor} $n$ may be different (and generally is)   at each round of the fight.
-It may be chosen by the hydra, according to its strategy, or imposed by some 
-particular rule.
-In many presentations of hydra battles, this number is increased by $1$ at each round.
-In the following presentation, we will also consider battles where the hydra is free to choose its ~replication factor at each round of the battle\footnote{Let us recall that, if the chopped-off head was at distance 1 from the foot, the replication factor is meaningless.}.
-\end{itemize}
-\end{itemize}
-
-Note that the description given in~\cite{KP82} of the replication process in hydra battles is also  semi-formal.
-
-\label{original-rules}
-
-\begin{quote}
-  ``From the node that used to be attached to the head which was just chopped off, traverse one 
-segment towards the root until the next node is reached. From this node sprout $n$ replicas of 
-that part of the hydra (after decapitation) which is ``above'' the segment just traversed, i.e., those 
-nodes and segments from which, in order to reach the root, this segment would have to be 
-traversed. If the head just chopped off had the root of its nodes, no new head is grown. ''
-\end{quote}
-
-Moreover, we note that this description is in \emph{imperative} terms.
-In order to formally study the properties of hydra battles, we prefer to use a mathematical vocabulary, i.e., graphs, relations, functions, etc.
-Thus, the replication process will be represented as a binary relation on a data type \texttt{Hydra},
-linking the state of the hydra \emph{before} and \emph{after} the transformation.
-A battle will thus be represented as a sequence of terms of type \texttt{Hydra}, respecting the rules of the game.
-
-\subsection{Example}
-Let us start a battle between Hercules and the hydra \texttt{Hy} of Fig.~\ref{fig:Hy}.
-
-At the first round, Hercules chooses to chop off the rightmost head of \texttt{Hy}.
-Since this head is near the floor, the hydra simply loses this head. Let us call 
- \texttt{Hy'} the resulting state of the hydra, represented in Fig.~\vref{fig:Hy-prime}.
-
-Next, assume Hercules chooses to chop off one of the two highest heads of \texttt{Hy'}, for instance the rightmost one.
-Fig.~\vref{fig:Hy2} represents the broken segment in dashed lines, and the part that will be replicated in red.
-Assume also that the hydra decides to add 4 copies of the red part\footnote{In other words, the replication factor at this round is equal to $4$.}.
-We obtain a new state \texttt{Hy''} depicted in Fig.~\ref{fig:Hy3}.
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {$\bullet$};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-%\node (H4) at (6,0) {$\Xey[2][lightgray]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3) ;
-\draw (N3) to [bend left= 10]  (H1) ;
-\draw (N3) to [bend right= 16] (H2);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-
-\caption{\texttt{Hy'}: the state  of \texttt{Hy} after one round \label{fig:Hy-prime}}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4)  {{\color{lightred}$\bullet$}};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Sey[2][lightred]$};
-%\node (H2) at (5,0) {$\Xey[2][lightgray]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\node (ex) at (5,8) {};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw  (N2) -- (N3)[draw=lightred];
-\draw (N3) to   [bend left= 10](H1) [draw=lightred];
-\draw [dashed] (N3) to [bend left= 10](ex);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-\caption{A second beheading}
-\label{fig:Hy2}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H11) at (2,8) {$\Smiley[2][vertfluo]$};
-\node (H12) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H13) at (6,8) {$\Smiley[2][vertfluo]$};
-\node (H14) at (8,8) {$\Smiley[2][vertfluo]$};
-
-\node (N3) at (1,6) {$\bullet$};
-\node (N31) at (2,6) {$\bullet$};
-\node (N32) at (3,6) {$\bullet$};
-\node (N33) at (4,6) {$\bullet$};
-\node (N34) at (5,6) {$\bullet$};
-
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N2) -- (N31);
-\draw (N2) -- (N32);
-\draw (N2) -- (N33);
-\draw (N2) -- (N34);
-\draw (N3) to   [bend left= 10](H1) ;
-\draw (N31) to   [bend left= 10](H11) ;
-\draw (N32) to   [bend left= 10](H12) ;
-\draw (N33) to   [bend left= 10](H13) ;
-\draw (N34) to   [bend left= 10](H14) ;
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend left= 10]  (H5) ;
-\end{tikzpicture}
-\caption{\texttt{Hy''}: the state of \texttt{Hy} after two rounds \label{fig:Hy3}}
-\end{figure}
-
-Figs.~\ref{fig:Hy4} and~\vref{fig:Hy5} represent a possible third round of the battle, with a replication factor equal to $2$. Let us call \texttt{Hy'''} the state of the hydra after that third round.
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0)  {{\color{lightred}$\bullet$}};
-\node (N1) at (2,2) {{\color{lightred}$\bullet$}};
-\node (N2) at (2,4) {{\color{lightred}$\bullet$}};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (exN4) at (4,4) {};
-\node (H1) at (0,8) {$\Sey[2][lightred]$};
-\node (H11) at (2,8) {$\Sey[2][lightred]$};
-\node (H12) at (4,8) {$\Sey[2][lightred]$};
-\node (H13) at (6,8) {$\Sey[2][lightred]$};
-\node (H14) at (8,8) {$\Sey[2][lightred]$};
-
-\node (N3) at (1,6) {{\color{lightred}$\bullet$}};
-\node (N31) at (2,6) {{\color{lightred}$\bullet$}};
-\node (N32) at (3,6) {{\color{lightred}$\bullet$}};
-\node (N33) at (4,6) {{\color{lightred}$\bullet$}};
-\node (N34) at (5,6) {{\color{lightred}$\bullet$}};
-
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-%\node (H5) at (4,0) {$\Xey[2][lightgray]$};
-\draw (foot) -- (N1)[very thick,draw=lightred] ;
-\draw (N1) -- (N2)[draw=lightred];
-\draw (N2) -- (N3)[draw=lightred];
-\draw (N2) -- (N31)[draw=lightred];
-\draw (N2) -- (N32)[draw=lightred];
-\draw (N2) -- (N33)[draw=lightred];
-\draw (N2) -- (N34)[draw=lightred];
-\draw (N3) to   [bend left= 10](H1) [draw=lightred];
-\draw (N31) to   [bend left= 10](H11) [draw=lightred];
-\draw (N32) to   [bend left= 10](H12) [draw=lightred];
-\draw (N33) to   [bend left= 10](H13) [draw=lightred];
-\draw (N34) to   [bend left= 10](H14) [draw=lightred];
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw [dashed] (N1) to  [bend left= 10](exN4);
-\end{tikzpicture}
-\caption{A third beheading (wounded part in red) \label{fig:Hy4}}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-
-\node (foot) at (10,0) {$\bullet$};
-
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H1) at (0,8) {$\Smiley[1][vertfluo]$};
-\node (H11) at (2,8) {$\Smiley[1][vertfluo]$};
-\node (H12) at (4,8) {$\Smiley[1][vertfluo]$};
-\node (H13) at (6,8) {$\Smiley[1][vertfluo]$};
-\node (H14) at (8,8) {$\Smiley[1][vertfluo]$};
-
-\node (N3) at (1,6) {$\bullet$};
-\node (N31) at (2,6) {$\bullet$};
-\node (N32) at (3,6) {$\bullet$};
-\node (N33) at (4,6) {$\bullet$};
-\node (N34) at (5,6) {$\bullet$};
-
-\node (H0) at (-3,3) {$\Smiley[1][vertfluo]$};
-
-\draw (foot) to [bend left=10] (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N2) -- (N31);
-\draw (N2) -- (N32);
-\draw (N2) -- (N33);
-\draw (N2) -- (N34);
-\draw (N3) to   [bend left= 10](H1) ;
-\draw (N31) to   [bend left= 10](H11) ;
-\draw (N32) to   [bend left= 10](H12) ;
-\draw (N33) to   [bend left= 10](H13) ;
-\draw (N34) to   [bend left= 10](H14) ;
-\draw (foot) to [bend left = 15]  (H0) ;
-
-% second copy 
-\node (N01) at (12,2) {$\bullet$};
-\node (N02) at (12,4) {$\bullet$};
-\node (N03) at (12,6) {{\color{lightred}$\bullet$}};
-\node (H001) at (10,8) {$\Smiley[1][vertfluo]$};
-\node (H0011) at (12,8) {$\Smiley[1][vertfluo]$};
-\node (H0012) at (14,8) {$\Smiley[1][vertfluo]$};
-\node (H0013) at (16,8) {$\Smiley[1][vertfluo]$};
-\node (H0014) at (18,8) {$\Smiley[1][vertfluo]$};
-
-\node (N03) at (11,6) {$\bullet$};
-\node (N031) at (12,6) {$\bullet$};
-\node (N032) at (13,6) {$\bullet$};
-\node (N033) at (14,6) {$\bullet$};
-\node (N034) at (15,6) {$\bullet$};
-
-\draw (foot) -- (N01)[very thick] ;
-\draw (N01) -- (N02);
-\draw (N02) -- (N03);
-\draw (N02) -- (N031);
-\draw (N02) -- (N032);
-\draw (N02) -- (N033);
-\draw (N02) -- (N034);
-\draw (N03) to   [bend left= 10](H001) ;
-\draw (N031) to   [bend left= 10](H0011) ;
-\draw (N032) to   [bend left= 10](H0012) ;
-\draw (N033) to   [bend left= 10](H0013) ;
-\draw (N034) to   [bend left= 10](H0014) ;
-
-% third copy 
-\node (N001) at (22,2) {$\bullet$};
-\node (N002) at (22,4) {$\bullet$};
-\node (N003) at (22,6) {{\color{lightred}$\bullet$}};
-\node (H001) at (20,8) {$\Smiley[1][vertfluo]$};
-\node (H0011) at (22,8) {$\Smiley[1][vertfluo]$};
-\node (H0012) at (24,8) {$\Smiley[1][vertfluo]$};
-\node (H0013) at (26,8) {$\Smiley[1][vertfluo]$};
-\node (H0014) at (28,8) {$\Smiley[1][vertfluo]$};
-
-\node (N003) at (21,6) {$\bullet$};
-\node (N0031) at (22,6) {$\bullet$};
-\node (N0032) at (23,6) {$\bullet$};
-\node (N0033) at (24,6) {$\bullet$};
-\node (N0034) at (25,6) {$\bullet$};
-
-\draw (foot) -- (N001)[very thick] ;
-\draw (N001) -- (N002);
-\draw (N002) -- (N003);
-\draw (N002) -- (N0031);
-\draw (N002) -- (N0032);
-\draw (N002) -- (N0033);
-\draw (N002) -- (N0034);
-\draw (N003) to   [bend left= 10](H001) ;
-\draw (N0031) to   [bend left= 10](H0011) ;
-\draw (N0032) to   [bend left= 10](H0012) ;
-\draw (N0033) to   [bend left= 10](H0013) ;
-\draw (N0034) to   [bend left= 10](H0014) ;
-\end{tikzpicture}
-\caption{The configuration \texttt{Hy'''} of \texttt{Hy} \label{fig:Hy5}}
-\end{figure}
-\FloatBarrier
-
-We leave it to the reader  to guess the following  rounds of the battle \dots
- % Please keep in mind that, in this 
-% the hydra is free to chose any number of replications at each time, whereas
-% Hercules chops only one head per round.
-
-% Let us precise that, in this game, Hercules wins if the hydra is eventually reduced 
-% to a single head. 
-% We know from~\cite{KP82} that, whichever the initial configuration of the
-% hydra, and the strategies of both players, Hercules eventually wins. The 
-% aforementionned paper shows also that there do not exist any \emph{simple} proof of this result.
-
-\section{Hydras and their representation in \emph{Coq}}
-\label{sec:orgheadline48}
-
-\index{hydras}{Library Hydra!Types!Hydra}
-\index{hydras}{Library Hydra!Types!Hydrae}
-
-In order to describe trees where each node can have an arbitrary (but finite) number of sons, it is usual to define a type where each node carries a \emph{forest}, \emph{i.e} a list of trees
-(see for instance Chapter 14, pages 400-406 of \cite{BC04}, also available as~\cite{BC04ch14}).
-
-For this purpose, we define two mutual \emph{ad-hoc}  inductive types, where \texttt{Hydra} is the main type, and \texttt{Hydrae} is a helper for describing finite sequences of hydras.
-\label{types:Hydra}
-\label{types:Hydrae}
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hydra}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/HydraDef}
-
-%\index{To do}
-\index{hydras}{Projects}
-
-\begin{project}
-Look for an existing library on trees with nodes of arbitrary arity, in order to replace  this ad-hoc type with something more generic.
-\end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-
- Another very similar representation could use the \texttt{list} type family instead of the specific 
-type \texttt{Hydrae}:
-
-\input{movies/snippets/Hydra_Definitions/HydraAlt}
-
-Using this representation, re-define all the constructions of this chapter.
-You will probably have to use patterns described for instance in~\cite{BC04} or the archives of the \coq communication channels (please consult~\url{https://coq.inria.fr/community.html}).
-
-\end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-The type \texttt{Hydra} above describes hydras as \emph{plane trees}, i.e., as drawn on a sheet of paper or computer screen. Thus, hydras are \emph{oriented},
-and it is appropriate to consider a \emph{leftmost} or \emph{rightmost} head of
-the beast. It could be interesting to consider another representation, in which
-every non-leaf node has a \emph{multi-set} -- not an ordered list -- of daughters.
-\end{project}
-
-\subsubsection{Abbreviations}
-
-We provide several notations for hydra patterns  which occur often in our developments.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#head}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/headsEtc}
-
-For instance, the hydra \texttt{Hy}  of Figure~\vref{fig:Hy} is defined in \emph{Gallina} as follows:
-
-\vspace{4mm}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html\#Hy}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/Hy}
-
-Hydras quite frequently contain  multiple adjacent  copies of the same subtree. The following functions
-will help us to describe and reason about replications in hydra battles.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#hcons_mult}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hconsMult}
-
-\vspace{4mm}
-
-For instance, the hydra \texttt{Hy''} of Fig~\vref{fig:Hy3}  can be defined in \coq{} as follows:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/HySecond}
-
-\subsubsection{Recursive functions on type \texttt{Hydra}}
-\label{sec:orgheadline41}
-\label{sec:hsize-def}
-
-In order to  define a recursive function over the type \texttt{Hydra}, one has to consider the three constructors 
-\texttt{node}, \texttt{hnil} and \texttt{hcons} of the mutually inductive types \texttt{Hydra} and \texttt{Hydrae}. 
-Let us define for instance the function which  computes the number of nodes of any hydra:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hsize}
-
-\input{movies/snippets/Hydra_Examples/HySize}
-
-Likewise, the \emph{height} (maximum distance between the foot and a head) 
-is defined by mutual recursion:
-
-\input{movies/snippets/Hydra_Definitions/height}
-
-\input{movies/snippets/Hydra_Examples/HyHeight}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Define a function \texttt{max\_degree: Hydra $\arrow$ nat} which  returns the highest degree of a node in any hydra.
-For instance, the evaluation of the term \texttt{(max\_degree Hy)} should return $3$.
-\end{exercise}
-
-\subsection{Induction principles for hydras}
-\label{sec:orgheadline42}
-
-In this section, we show how induction principles are used to prove properties on the type 
-\texttt{Hydra}. Let us consider for instance the following statement:
-\begin{quote}
-  `` The height of any hydra is strictly less than its size. ''
-\end{quote}
-
-\subsubsection{A failed attempt}
-
-One may try to use the default tactic of proof by induction, which corresponds to an application of the automatically  generated  induction principle for  type \texttt{Hydra}:
-
-\input{movies/snippets/Hydra_Examples/HydraInd}
-
-Let us start a simple proof by induction.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/BadInductiona}
-
-We might be tempted to do an induction on the sequence \texttt{s}:
-
-\input{movies/snippets/Hydra_Examples/BadInductionb}
-
-The first subgoal is trivial.
-
-\input{movies/snippets/Hydra_Examples/BadInductionc}
-
-Let us look at the second subgoal of the induction.
-
-\input{movies/snippets/Hydra_Examples/BadInductiond}
-
-We notice that this subgoal does not contain any hypothesis
-on the height and size of the hydra \texttt{h}. So, it looks hard to prove the conclusion. Let's stop.
-
-\input{movies/snippets/Hydra_Examples/BadInductione}
-
-\subsubsection{A Principle of mutual induction}
-In order to get an appropriate induction scheme for the types 
-\texttt{Hydra} and \texttt{Hydrae}, we can use  \coq{}'s  command \texttt{Scheme}.
-
-\index{coq}{Mutually inductive types}
-%\index{coq}{Commands!Scheme}
-
-\input{movies/snippets/Hydra_Definitions/HydraRect2}
-
-\input{movies/snippets/Hydra_Examples/HydraRect2Check}
-
-\subsubsection{A Correct proof}
-
-Let us now use \texttt{Hydra\_rect2} for proving that the height of any hydra is strictly less than its size.
-Using this scheme requires an auxiliary predicate, called \texttt{P0} in \texttt{Hydra\_rect2}'s statement.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hForall}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-\input{movies/snippets/Hydra_Examples/heightLtSizea}
-
-The first subgoal is easily solvable, using some arithmetic.
-The second and third ones are
-almost trivial. We let the reader look at the source. 
-\input{movies/snippets/Hydra_Examples/heightLtSizez}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-It happens very often that, in the proof of  a proposition of the form 
-(\texttt{$\forall\,$ h:Hydra, $P$ h}), the predicate \texttt{P0}
-is  (\texttt{h\_forall $P$}).
-Design a tactic for induction on hydras that frees the user from binding explicitly \texttt{P0},  and solves trivial subgoals.
-Apply it for writing  a shorter proof of \texttt{height\_lt\_size}.
-\end{exercise}
-
-\section{Relational description of hydra battles}
-
-In this section, we represent the rules of hydra battles as a binary relation associated with
-a \emph{round}, i.e., an interaction composed of the two following actions:
-\begin{enumerate}
-\item Hercules chops off one head of the hydra.
-\item Then, the  hydra replicates the wounded part (if the head is at distance $\geq 2$ from the foot).
-\end{enumerate}
-The relation associated with each round of the battle is parameterized  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot,
-but present for consistency's sake).
-
-In our description,  we will apply the following naming convention: if $h$ represents the configuration of the hydra before a round, then the configuration of $h$ after this round will be called $h'$.
- Thus, we are going to define a proposition  (\texttt{round\_n $n\;h\;h'$})  whose intended meaning will be `` the hydra $h$  is transformed into $h'$  in a single round of a battle, with the expected replication factor $n$ ''.
-
-Since the replication of parts of the hydra depends on the distance of the chopped head from  the foot, we  decompose our description into two main  cases, under the form of a bunch of [mutually] inductive predicates over the types \texttt{Hydra} and \texttt{Hydrae}.
-
-The mutually exclusive cases we consider are the following:
-\begin{itemize}
-\item \textbf{R1}: The chopped off head was at distance 1 from the foot.
-\item \textbf{R2}: The chopped off head was at a distance greater than or equal to  $2$ from the foot.
-\end{itemize}
-
-\subsection{Chopping off a head at distance 1 from the foot (relation  R1)}
-
-If Hercules chops off a head near the floor, there is no replication at all. We use an auxiliary 
-predicate \texttt{S0}, associated with the removing of one head from a sequence of hydras.
-
-\vspace{4pt}\emph{From Module\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/S0Def}
-\input{movies/snippets/Hydra_Definitions/R1Def}
-
-\subsubsection{Example}
-\label{sec:orgheadline45}
-
-Let us represent in \coq{}   the transformation of the hydra of Fig.~\vref{fig:Hy} into
-the configuration represented in Fig.~\ref{fig:Hy-prime}.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/Hy1}
-
-\subsection{Chopping off a head at distance \texorpdfstring{$\geq 2$}{>= 2} from the foot (relation R2) }
-
-Let us now consider beheadings  where the chopped-off head is at distance greater than or equal to $2$ from the foot. All the following relations are parameterized by the replication factor  $n$.
-
- Let $s$ be a sequence of hydras. 
-The proposition (\texttt{S1 n s s'}) holds if $s'$ is obtained by replacing some element $h$ of $s$ by 
-$n+1$ copies of $h'$, where  the proposition (\texttt{R1 h h'}) holds, in other words, $h'$ is just $h$, without the chopped-off  head.
-\texttt{S1} is an inductive relation with two constructors that allow us to choose the position in $s'$ of the wounded sub-hydra $h$.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#S1}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/S1Def}
-
-The rest of the definition is composed of two mutually inductive relations on hydras and sequences of hydras.
-The first constructor of \texttt{R2} describes the case where the chopped head is exactly at height $2$.
-The others constructors allow us to consider beheadings at height strictly greater than $2$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/R2Def}
-
-\subsubsection{Example}
-Let us prove the transformation of \texttt{Hy'} into \texttt{Hy''} (see Fig.~\vref{fig:Hy3}).
-We use an experimental set of tactics\footnote{See the \texttt{Ltac} definitions in~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions.}}  for specifying the place where the
-interaction between Hercules and the hydra holds.
-
-\vspace{4pt}\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}.
-
-\input{movies/snippets/Hydra_Examples/R2Example}
-
-The reader is encouraged to look at all the successive subgoals of this example.
-\emph{Please consider also exercise~\vref{exo:interactive-battle}.}
-
-\subsection{Binary relation associated with a round}
-
-We combine the two cases above into a single relation.
-First,  we define the  relation \texttt{(round\_n n h h')} where \texttt{n} is the expected number of  replications (irrelevant in the case of an \texttt{R1}-transformation).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#round_n}{Hydra.Hydra\_Definitions}}
-
-\index{hydras}{Library Hydra!Predicates!round\_n}
-
-\input{movies/snippets/Hydra_Definitions/roundNDef}
-
-By abstraction over \texttt{n}, we define a \emph{round} (small step) of a battle:
-
-\index{hydras}{Library Hydra!Predicates!round}
-\label{sect:infix-round}
-
-\input{movies/snippets/Hydra_Definitions/roundDef}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Give a direct translation of Kirby and Paris's description of hydra battles (quoted on page~\pageref{original-rules}) and prove that our relational description is consistent with theirs.
-\end{project}
-
-\subsection{Rounds and battles}
-
-Using library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}, we define \texttt{round\_plus},  the transitive closure of \texttt{round}, and \texttt{round\_star},  the reflexive and transitive closure of \texttt{round}.
-
-\label{sect:infix-rounds}
-
-\input{movies/snippets/Hydra_Definitions/roundPlus}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-  Prove that if \texttt{$h$ -+-> $h'$}, then
-  the height of $h'$ is less or equal than the height of $h$.
-
-\end{exercise}
-
-\begin{remark}
-\label{remark:transitive-closure}
-\coq's library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Coq.Relations.Relation\_Operators} 
-contains three logically equivalent definitions of the transitive closure of a binary relation. This equivalence is proved in 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} .
-
-Why three definitions for a single mathematical concept?
-Each definition generates an associated induction principle. 
- According to the form of statement one would like to prove, there is a ``best choice'':
-
-\begin{itemize}
-\item For proving $\forall y, x\,R^+\,y \;\arrow\; P\,y$, prefer 
-\texttt{clos\_trans\_n1}
-\item For proving $\forall x,\,x\,R^+\,y \;\arrow\; P\,x$, prefer \texttt{clos\_trans\_1n}
-\item For proving $\forall x\,y, \,x\,R^+\,y \;\arrow\;P\,x\,y$,  
-prefer \texttt{clos\_trans},
-\end{itemize}
-But there is no ``wrong choice'' at all: the equivalence lemmas in \linebreak 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} 
- allow the user
-to convert any one of the three closures into another one before applying the corresponding elimination tactic.
-The same remark also holds for reflexive and transitive closures. 
-\end{remark}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Define a restriction of \coqsimple{round},  where Hercules always chops off
-the leftmost among the lowest heads.
-
-Prove that, if $h$ is not a simple head, then there exists a unique $h'$ such that $h$  is transformed into $h'$ in one round, according to this restriction.
-
-\end{exercise}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}[Interactive battles]
-\label{exo:interactive-battle}
-Given a hydra \texttt{h}, the specification of a hydra battle for \texttt{h} is the type 
-\Verb@{h':Hydra | h -*-> h'}@. In order to avoid long sequences of \texttt{split}, \texttt{left}, and 
-\texttt{right}, design a set of dedicated tactics for the interactive building of a battle.
-Your tactics will have the following functionalities:
-\begin{itemize}
-\item  Chose to stop a battle, or continue
-\item Chose an expected number of replications
-\item Navigate in a hydra, looking for a head to chop off.
-\end{itemize}
-
-Use your tactics for simulating a small part of a hydra battle, for instance the rounds which lead from
-\texttt{Hy} to \texttt{Hy'''}  (Fig.~\vref{fig:Hy5}).
-
-\textbf{Hints:} 
-\begin{itemize}
-
-\item Please keep in mind that the last  configuration of your interactively built battle is known only at the end of the battle.
-Thus, you will have to create and solve subgoals with existential variables.
-For that purpose, the tactic \texttt{eexists}, applied to the
-goal \Verb@{h':Hydra | h -*-> h'}@ generates the subgoal \Verb|h -*-> ?h'|.
-\item You may use Gérard Huet's \emph{zipper} data structure~\cite{zipper} for writing tactics associated with Hercules's  interactive search for a head to chop off.
-\end{itemize}
-
-\end{exercise}
-
-\subsection{Classes of battles}
-\label{sect:battle-classes}
-
-In some presentations of hydra battles, e.g.~\cite{KP82, bauer2008}, the transformation associated with the $i$-th round may depend on $i$.
-For instance, in these articles, the replication factor at the $i$-th round is equal to $i$.
-In other examples, one can allow the hydra to apply any replication factor at any time.
-In order to be the most general as possible, we define the type of predicates which relate the state of the hydra before and after the $i$-th round of a battle.
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-\label{types:Battle}
-\index{hydras}{Library Hydra!Type classes!Battle}
-
-\input{movies/snippets/Hydra_Definitions/BattleDef}
-
-The most general class of battles is \texttt{free}, which allows the hydra to chose any replication factor at every step:
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#free}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/freeDef}
-
-We chose to call \emph{standard}\footnote{This appellation is ours.
-If there is a better one, we will change it.} the kind of battles which appear  most often in the literature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#standard}{Hydra.Hydra\_Definitions}}
-\input{movies/snippets/Hydra_Definitions/standardDef}
-
-\subsection{Big steps}
-
-Let $B$ be any instance of class \texttt{Battle}. It is easy to define inductively the relation between the $i$-th and the $j$-th steps of a battle of type $B$.
-
-\vspace{6pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/battleRelDef}
-
-The following property allows us to build battles by composition of smaller ones.
-
-%% TODO display subgoals when fixed
-
-\vspace{6pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-\noindent
-\input{movies/snippets/Hydra_Lemmas/battleTrans}
-
-% \begin{remark}
-%  The class \texttt{free} is strongly related with the transitive closure  \texttt{round\_plus}, as expressed by the following lemmas.
-
-% \vspace{4pt}
-% \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-
-%  \begin{Coqsrc}
-%  Lemma battle_free_equiv1 : forall i j h h',  
-%              battle free i h j h' ->   h -+-> h'.
-
-%  Lemma battle_free_equiv2 : forall h h',
-%      h -+-> h' ->
-%     forall i, exists j,  battle free i h j h'.
-%  \end{Coqsrc}
-
-% \end{remark}
-
-\section{A long battle}
-\label{sect:big-battle}
-
-In this section we consider a simple example of battle, starting with a small hydra,
-shown on figure~\vref{fig:hinit}, with a simple strategy for both players:
-
-\begin{itemize}
-\item At each round, Hercules chops off the rightmost head of the hydra.
-\item The battle is {standard}: at the round number $i$, the expected replication is $i$.
-\end{itemize}
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (6,0) {$\bullet$};
-\node (n1) at  (3,3) {$\bullet$};
-\node (h1) at  (1,6) {$\Smiley[1][green]$};
-\node (h2) at  (3,6) {$\Smiley[1][green]$};
-\node (h3) at  (6,6) {$\Smiley[1][green]$};
-\node (h4) at  (6,6) {$\Smiley[1][green]$};
-\node (h5) at  (6,3) {$\Smiley[1][green]$};
-\node (h6) at  (9,3) {$\Smiley[1][green]$};
-\draw (foot) -- (n1);
-\draw (n1) to   [bend left=20] (h1);
-\draw (n1) to   (h2);
-\draw (n1) to   [bend right=20] (h3);
-\draw (foot) -- (h5);
-\draw (foot) to  [bend right=20] (h6);
-\end{tikzpicture}
-
-  \caption{The hydra \texttt{hinit}}
-  \label{fig:hinit}
-\end{figure}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.BigBattle.html}{Hydra.BigBattle}}
-
-\input{movies/snippets/BigBattle/hinitDef}
-
-The lemma we would like to prove is ``The considered battle lasts exactly $N$ rounds'',
-with $N$ being a natural number we gave to guess.
-
-But the  battle is so long that no \emph{test} can give us any estimation of its length, and we do need the expressive power of logic to compute this length.
-However, in order to  guess this length, we made some experiments, computing with \gallina{}, \coq{}'s  functional programming language.
-Thus, we can consider this development as a collaboration of proof with computation.
-In the following lines, we show how we found experimentally the value of the number $N$.
-The complete proof is in file \url{../theories/html/hydras.Hydra.BigBattle.html}.
-
-\subsection{First rounds}
-During the two first rounds, our hydra loses its two rightmost heads.  Figure~\vref{fig:hinit-plus2} shows the state of the hydra   just before the third round.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (3,0) {$\bullet$};
-\node (n1) at  (3,3) {$\bullet$};
-\node (h1) at  (1,6) {$\Smiley[1][green]$};
-\node (h2) at  (3,6) {$\Smiley[1][green]$};
-\node (h3) at  (6,6) {$\Smiley[1][green]$};
-\node (h4) at  (6,6) {$\Smiley[1][green]$};
-\draw (foot) -- (n1);
-\draw (n1) to   [bend left=20] (h1);
-\draw (n1) to   (h2);
-\draw (n1) to   [bend right=20] (h3);
-\end{tikzpicture}
-
-  \caption{The hydra (\texttt{hyd1 h3})}
-  \label{fig:hinit-plus2}
-\end{figure}
-
-The following lemma  is a formal description of these first rounds, in terms of the
-\texttt{battle} predicate.
-
-\input{movies/snippets/BigBattle/L02}
-
-\subsection{Looking for regularities}
-
-A first study with pencil and paper suggested us that, after three rounds, the hydra always looks like in figure~\vref{fig:hinit-plusn} (with a variable number of 
-subtrees of height 1 or 0).
-Thus, we introduce a few handy abbreviations.
-
-\input{movies/snippets/BigBattle/Notations}
-
-For instance, the hydra shown in  Fig~\ref{fig:hinit-plusn} is  (\texttt{hyd 3 4 2}). The hydra (\texttt{hyd 0 0 0})  is the ``final'' hydra of any terminating battle, {i.e.},
-a tree whith exactly one node and no edge.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (15,0) {$\bullet$};
-\node (a) at  (3,4) {$\bullet$};
-\node (b) at  (6,4) {$\bullet$};
-\node (c) at  (9,4) {$\bullet$};
-\node (d) at  (13,4) {$\bullet$};
-\node (e) at  (16,4) {$\bullet$};
-\node (f) at  (19,4) {$\bullet$};
-\node (g) at  (22,4) {$\bullet$};
-\node (h) at  (25,4) {$\Smiley[1][green]$};
-\node (i) at  (28,4) {$\Smiley[1][green]$};
-\node (aa) at  (2.5,8) {$\Smiley[1][green]$};
-\node (ab) at  (3.5,8) {$\Smiley[1][green]$};
-\node (ba) at  (5.5,8) {$\Smiley[1][green]$};
-\node (bb) at  (6.5,8) {$\Smiley[1][green]$};
-\node (ca) at  (8.5,8) {$\Smiley[1][green]$};
-\node (cb) at  (9.5,8) {$\Smiley[1][green]$};
-\node (da) at  (13,8) {$\Smiley[1][green]$};
-\node (ea) at  (16,8) {$\Smiley[1][green]$};
-\node (fa) at  (19,8) {$\Smiley[1][green]$};
-\node (ga) at  (22,8) {$\Smiley[1][green]$};
-\draw (foot) -- (a);
-\draw (foot) -- (b);
-\draw (foot) -- (c);
-\draw (foot) -- (d);
-\draw (foot) -- (e);
-\draw (foot) -- (f);
-\draw (foot) -- (g);
-\draw (foot) -- (h);
-\draw (foot) -- (i);
-\draw (a) -- (aa);
-\draw (a) -- (ab);
-\draw (b) -- (ba);
-\draw (b) -- (bb);
-\draw (c) -- (ca);
-\draw (c) -- (cb);
-\draw (d) -- (da);
-\draw (e) -- (ea);
-\draw (f) -- (fa);
-\draw (g) -- (ga);
-\end{tikzpicture}
-
-  \caption{The hydra (\texttt{hyd 3 4 2})}
-  \label{fig:hinit-plusn}
-\end{figure}
-
-With these notations, we get a formal description of the first three rounds.
-
-\input{movies/snippets/BigBattle/L23L03}
-
-\subsection{Testing  \dots}
-\label{sect:testing}
-In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple data type for representing the states as tuples composed of
-the round number, and the respective number of daughters  \texttt{h2}, \texttt{h1}, and heads
-of the current hydra.
-
-\input{movies/snippets/BigBattle/stateDef}
-
-The following function returns the next configuration of the game.
-Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle only start with the lemma
-\texttt{step\_battle}, page~\pageref{lemma:step-battle}.
-
-\input{movies/snippets/BigBattle/nextDef}
-
-We can make bigger steps through iterations of \texttt{next}.
-The functional \texttt{iterate}, similar to Standard Library's \texttt{Nat.iter},
-is defined and studied in~\href{../theories/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
-\index{hydras}{Library Prelude!iterate}
-
-\label{Functions:iterate}
-
-\input{movies/snippets/Iterates/iterateDef}
-
-The following function computes the state of the battle at the $n$-th round.
-
-\input{movies/snippets/BigBattle/testDefTests}
-
-The battle we are studying looks to be awfully long. Let us concentrate our
-tests on some particular events : the states where $\texttt{nh}=0$.
-From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} is equal to zero.
-
-\input{movies/snippets/BigBattle/smartTest}
-
-Thus, $(1 + 11)$ rounds later, the \texttt{n1} field is equal to $2$, and 
-\texttt{nh}   to $0$.
-
-\input{movies/snippets/BigBattle/smartTestb}
-
-Next round, we decrement \texttt{n2} and set \texttt{n1} to $95$.
-
-\input{movies/snippets/BigBattle/smartTestc}
-
-We now have some intuition of the sequence.
-It looks like the next ``\texttt{nh}=0'' event will happen at the $192=2(95+1)$-th round, then at the $2(192+1)$-th round, etc.
-
-\input{movies/snippets/BigBattle/doubleS}
-
-\subsection{Proving \dots}
-We are now able to reason about the sequence of transitions defined by our hydra battle. Instead of using the data-type \texttt{state} we study the relationship
-between different configurations of the battle.
-
-Let us define a binary relation associated with every round of the battle.
-In the following definition \texttt{i} is associated with the round number (or date, if we consider a discrete time), and \texttt{a}, \texttt{b}, \texttt{c} respectively associated with the number of occurrences of \texttt{h2}, \texttt{h1} and heads connected to the hydra's foot.
-
-\input{movies/snippets/BigBattle/oneStep}
-
-The relation between \texttt{one\_step} and the rules of hydra battles is asserted by the following lemma.
-
-\label{lemma:step-battle}
-
-\input{movies/snippets/BigBattle/stepBattle}
-
-\vspace{4pt}
-
-Next, we define ``big steps'' as the transitive closure of \texttt{one\_step},
-and reachability (from the initial configuration of figure~\ref{fig:hinit} at time $0$).
-
-\input{movies/snippets/BigBattle/steps}
-
-The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{battle}.
-
-\input{movies/snippets/BigBattle/stepsBattle}
-
-\vspace{4pt}
-
-Thus, any result about \texttt{steps} will be applicable to standard battles.
-Using the predicate \texttt{steps},  our study of the length of the considered battle
-can  be decomposed into three parts:
-
-\begin{enumerate}
-\item  Characterization and proofs of regularities of some events (inspired by our experiments of Sect.~\ref{sect:testing}).
-\item Study of the beginning of the battle
-\item Computing the exact length of the battle.
-\end{enumerate}
-
-First, we prove that, if at round $i$ the hydra is equal to
-(\texttt{hyd a (S b) 0}), then it will be equal to (\texttt{hyd a b 0}) at the $2(i+1)$-th round.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/LS}
-
-From now on, the lemma \texttt{reachable\_S} allows us to watch larger and larger steps of 
-the battle.
-
-\input{movies/snippets/BigBattle/L4}
-
-\input{movies/snippets/BigBattle/L10To95}
-
-\subsection{Giant steps}
-
-We are now able to make bigger steps in the simulation of the battle.
-First, we iterate the lemma \texttt{reachable\_S}.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/Bigstep}
-
-\vspace{4pt}
-
-Applying lemmas \texttt{BigStep} and \texttt{L95} we make a first jump.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/MDef}
-
-Figure~\ref{fig:HM}  represents the hydra at the $M$-th round.
-At the $(M+1)$-th round, it will look like in fig~\ref{fig:HM-plus1}.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:HM}}
-The state of the hydra after $M$ rounds.
-% The hydra \texttt{h} of the proof that \(\omega^2\) is too small for proving Hercules' victory
-
-\end{figure}
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (0,5) {$\bullet$};
-\node (N12) at (0,8) {$\Smiley[2][green]$};
-\node (N2) at (2,5) {$\bullet$};
-\node (N22) at (2,8) {$\Smiley[2][green]$};
-\node (N3) at (4,5) {$\bullet$};
-\node (N32) at (4,8) {$\Smiley[2][green]$};
-\node (N4) at (6,5) {$\bullet$};
-\node (N42) at (6,8) {$\Smiley[2][green]$};
-
-\node (Ndots) at (12,8) {\Huge $\dots$};
-\node (Ndots2) at (12,5) {\Huge $\dots$};
-
-\node (N8) at (18,5) {$\bullet$};
-\node (N82) at (18,8) {$\Smiley[2][green]$};
-\node (N9) at (20,5) {$\bullet$};
-\node (N92) at (20,8) {$\Smiley[2][green]$};
-
-\draw (foot) -- (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N8);
-\draw (foot) -- (N9);
-\draw (N1) to  (N12);
-\draw (N2) to  (N22);
-\draw (N3) to  (N32);
-\draw (N4) to  (N42);
-\draw (N8) to  (N82);
-\draw (N9) to  (N92);
-\end{tikzpicture}
-\caption{\label{fig:HM-plus1}}
-The state of the hydra after $M+1$ rounds (with $M+1$ heads).
-
-\end{figure}
-
-\input{movies/snippets/BigBattle/L295S}
-
-\vspace{4pt}
-
-Then, applying once more the lemma \texttt{BigStep}, we get the exact time when
-Hercules wins!
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/NDef}
-
-\vspace{4pt}
-
-We are now able to prove formally that the considered battle is 
-composed of $N$ steps.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/Done}
-
-\subsection{A minoration lemma}
-
-Now, we would like to get an intuition of  how big the number $N$ is.
-For that purpose, we use a minoration of the function \texttt{doubleS} by the
-function (\texttt{fun n => 2 * n}).
-
-\vspace{4pt}
-
-\input{movies/snippets/Exp2/exp2Def}
-
-\vspace{4pt}
-
-Using a few facts (proven in 
-\href{../theories/html/hydras.Hydra.BigBattle.html}{hydras.Hydra.BigBattle}),we get several  minorations.
-
-\input{movies/snippets/BigBattle/minorationLemmas}
-
-\vspace{4pt}
-
-The number $N$ is greater than or  equal to $2^{2^{95}\times 95}.$ If we wrote $N$ in base $10$, $N$ would require at least $10^{30}$ digits!
-
-\section{Generic properties}
-
-The example we just studied shows that the termination of any battle may take a very long time. If we want to study hydra battles in general, we have to consider 
-any hydra and any strategy, both for Hercules and the hydra itself. So, we first  give some definitions, generally borrowed from transition systems vocabulary (see~\cite{tel_2000} for instance).
-
-\subsection{Liveliness}
-
-Let $B$ be an instance of \texttt{Battle}. We say that $B$ is \emph{alive} if
-for any configuration $(i,h)$, where $h$ is not a head, there exists a further step in class $B$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/AliveDef}
-
-The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
-\href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
-
-\input{movies/snippets/Hydra_Theorems/AliveThms}
-
-Both theorems are proved with the help of the  following strongly specified function:
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
-
-\input{movies/snippets/Hydra_Lemmas/nextRoundDec}
-
-\subsection{Termination}
-
-The termination of \emph{any}  battle is naturally expressed by the predicate \texttt{well\_founded} defined in the module \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}
- of the Standard Library.
-
-\index{hydras}{Library Hydra!Predicates!Termination}
-
-\input{movies/snippets/Hydra_Definitions/TerminationDef}
-
-Let $B$ be an instance of class \texttt{Battle}. A \emph{variant} for $B$ consists
-in a well-founded relation $<$  on some type \texttt{A}, and a function
-(also called a \emph{measure}) $m:\texttt{Hydra}\rightarrow A$ such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $B$, the inequality $m(h')<m(h)$ holds.
-
-\vspace{4pt}
-\noindent
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
-
-\label{sect:hvariant-def}
-
-\index{hydras}{Library Hydra!Type classes!Hvariant}
-
-\input{movies/snippets/Hydra_Definitions/HvariantDef}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
- Prove that, if there exists some  instance of (\texttt{Hvariant Lt wf\_Lt $B$ $m$}), then there exists no infinite battle in  $B$.
-\end{exercise}
-
-\subsection{A  small proof of impossibility}
-%\index{coq}{Proofs of impossibility}
-
-\label{omega-case}
-
-When one wants to prove a termination theorem with the help of a variant, 
-one has to consider first a well-founded set $(A,<)$, then a strictly decreasing measure on this set.
-The following two lemmas show that if  the order structure $(A,<)$ is too simple, it is useless to look for a convenient measure, which simply no exists.
-Such kind of result is useful, because it saves you time and effort.
-
-The best known well-founded order is the natural order on the set $\mathbb{N}$ of natural numbers (the type \texttt{nat} of Standard library).
-It would be interesting to look for some measure $m:\texttt{nat}\arrow\texttt{nat}$ and prove it is a variant.
-
-Unfortunately, we can prove that 
-\emph{no} instance of class (\texttt{WfVariant round Peano.lt $m$}) can be built, where
-$m$ is \emph{any} function of type \texttt{Hydra $\arrow$ nat}.
-
-Let us present the main steps of that proof, the script of which  is in the module ~\href{../theories/html/hydras.Hydra.Omega_Small.html}{Hydra/Omega\_Small.v} \footnote{ The name of this file means ``the ordinal $\omega$ is too small for proving the termination of [free] hydra battles ''.
-In effect, the elements of $\omega$, considered as a set, are just the natural numbers (see next chapter for more details)}.
-
-%\subsubsection{Preliminaries}
-
-Let us assume there exists some variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
-    the  termination of all hydra battles.
-
-\input{movies/snippets/Omega_Small/omegaSmalla}
-
-We define an injection $\iota$ from the type \texttt{nat} into \texttt{Hydra}.
-For any natural number $i$, $\iota(i)$ is the hydra composed of a foot and
-$i+1$ heads at height $1$. For instance, Fig.~\ref{fig:flower} represents the hydra $\iota(3)$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (4,0) {$\bullet$};
-\node (N1) at (2,2) {$\Smiley[2][green]$};
-\node (N2) at (4,2) {$\Smiley[2][green]$};
-\node (N3) at (6,2) {$\Smiley[2][green]$};
-\node (N4) at (8,2) {$\Smiley[2][green]$};
-\draw (foot) to [bend left =25] (N1);
-\draw (foot) to [bend left =15] (N2);
-\draw (foot) to [bend right =15] (N3);
-\draw (foot) to [bend right =25] (N4);
-\end{tikzpicture}
-\caption{\label{fig:flower}
-The hydra $\iota(3)$}
-\end{figure}
-
-\input{movies/snippets/Omega_Small/iotaDef}
-
-Let us consider now some hydra \texttt{big\_h} out of the range of the injection $\iota$ (see Fig.~\vref{fig:h-omega-omega}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to  (N2);
-\end{tikzpicture}
-\caption{\label{fig:h-omega-omega}}
- The hydra \texttt{big\_h}.
-\end{figure}
-
-\input{movies/snippets/Omega_Small/bigHDef}
-
- Using the functions $m$ and $\iota$, we define a second hydra \texttt{small\_h}, and show
- there is a one-round battle that transforms \texttt{big\_h} into \texttt{small\_h}. Please note that,
-due to the hypothesis \texttt{Hvar}, we are interested in the termination of \emph{free} battles. 
-There is no problem to consider a round with (\texttt{m big\_h}) as the replication factor.
-
-\input{movies/snippets/Omega_Small/smallHDef}
-
-But, by hypothesis, $m$ is a variant. Hence, we infer the following inequality.
-
-\vspace{4pt}
-
-\input{movies/snippets/Omega_Small/mLt}
-
-In order to get a contradiction, it suffices to  prove the inequality
-$m(\texttt{big\_h}) \leq m(\texttt{small\_h})$ i.e.,  $m(\texttt{big\_h}\leq m(\iota (m(\texttt{big\_h})))$.
-
-\input{movies/snippets/Omega_Small/mGea}
-
-Intuitively, it means that, from any hydra of the form (\texttt{iota $i$}), the battle will 
-take (at least) $i$ rounds. Thus the associated measure cannot be less than $i$.
-Technically, we prove this lemma by Peano induction on $i$.
-
-\begin{itemize}
-\item The base case $i=0$ is trivial
-\item Otherwise, let $i$ be any natural number and assume  the inequality
-  $i \leq m(\iota(i))$.
-  \begin{enumerate}
-  \item  But the hydra $\iota(S(i))$ can be transformed in one round into
-    $\iota(i)$ (by losing its rightmost head, for instance)
-  \item Since $m$ is a variant, we have $m(\iota(i)) < m(\iota(S(i)))$,
-    hence  $i< m(\iota(S(i)))$, which implies  $S(i)\leq  m(\iota(S(i)))$.
-  \end{enumerate}
-\end{itemize}
-
-We are now ready to complete our impossibility proof.
-
-\vspace{4pt}
-
-\inputsnippets{Omega_Small/mGeb, Omega_Small/mGez,
-  Omega_Small/omegaSmallz}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\subsubsection{Conclusion}
-
-In order to build a variant for proving the termination of all hydra battles, we need to consider order structures more complex than the usual order on type \texttt{nat}. 
-The notion of \emph{ordinal number} provides a catalogue of well-founded order types.
-For a reasonably large bunch of ordinal numbers, \emph{ordinal notations} are data-types which allow the \coq{} user to define functions, to compute and prove some properties, for instance by reflection.
-
-The next chapter is dedicated to a generic formalization of ordinal notations, and chapter~\ref{chap:T1} to a proof of termination of all hydra battles with the help of an ordinal notation for the interval $[0,\epsilon_0)$\,\footnote{We use the mathematical notation $[a,b)$ for the interval $\{x|a\leq x < b\}$.}.
-\index{maths}{Notations!Interval}
-
-%--------------------------------------------------------------
-
-\chapter{Introduction to ordinal numbers and ordinal notations}
-\label{chap:ON}
-
-The proof of termination of all hydra battles presented in~\cite{KP82} is based
-on \emph{ordinal numbers}.
-From a mathematical point of view, an ordinal is a representative of an equivalence class for isomorphisms of  totally ordered well-founded sets.
-
-For the computer scientist, ordinals are tools for proving the totality of a given recursive function, or termination of a transition system. \emph{Ordinal arithmetic} 
-provides a set of functions whose properties, like \emph{monotony}, allow to define \emph{variants}, \emph{i.e.} strictly decreasing measures used in proofs of termination.
-
-\vspace{4pt}
-
-Let us have a look at Figure~\ref{fig:ordinal-sequence}. It presents a few items of a  sequence of ordinal numbers, which extends the sequence of natural numbers.
-
-\begin{figure}[h]
-  \centering
-\fbox{\Large
-  \begin{minipage}{1.0\linewidth}
-  \begin{align*}
-     &\textcolor{blue}{0},\,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,\ldots\\
-&\textcolor{red}{\omega},\,\omega+1,\omega+2,\omega+3,\ldots\\
-&\textcolor{red}{\omega\times 2},\,\omega\times 2+1,\ldots, \textcolor{red}{\omega\times 3},\,\omega\times 3+1,\ldots, \textcolor{red}{\omega\times 4},\ldots\\
-&\textcolor{red}{\omega^2},\ldots, \textcolor{red}{\omega^2\times 42},\ldots,\textcolor{red}{\omega^3},\ldots, \textcolor{red}{\omega^4},\omega^4+1,\ldots\\
-&\textcolor{red}{\omega^\omega},\ldots, \textcolor{red}{\omega^\omega+\omega^7\times 8},\ldots,\textcolor{red}{\omega^\omega\times 2},\omega^\omega\times 2+1, \ldots\\
-&\textcolor{red}{\omega^{\omega^\omega}},\ldots, \textcolor{red}{\omega^{\omega^\omega}+\omega^\omega\times 42+ \omega^{55}+\omega}, \ldots, \textcolor{red}{\omega^{\omega^{\omega+1}}}, \omega^{\omega^{\omega+1}}+1,\dots\\
-& \textcolor{red}{\epsilon_0 (= \omega^{\epsilon_0)}}, \epsilon_0+1, \epsilon_0+2, \epsilon_0+3, \ldots\\
-& \textcolor{red}{\epsilon_1}, \ldots, \textcolor{red}{\epsilon_2}, \ldots, \textcolor{red}{\epsilon_\omega},\ldots \\
-& \textcolor{red}{\Gamma_0}, \Gamma_0+1, \Gamma_0+2, \Gamma_0+3,\ldots, \textcolor{red}{\Gamma_0+\omega}, \ldots\\
-&\ldots
-  \end{align*}   
-  \end{minipage}}
-
-  \caption{A short overview of the sequence of ordinal numbers}
-  \label{fig:ordinal-sequence}
-\end{figure}
-
-Let us comment some features of this figure:
-
-\begin{itemize}
-\item The ordinals are listed in a strictly increasing order. 
-\item Dots : ``$\ldots$'' stand for  infinite sequences of ordinals, not shown for lack of space.
-For instance, the ordinal $42$ is not shown in the first line, but it exists, somewhere between $17$ and $\omega$.
-\item Each ordinal printed in black is the immediate successor of another ordinal.
-We call it a \emph{successor} ordinal.
-For instance, $12$ is the successor of $11$, and $\omega^4+1$ the successor of $\omega^4$.
-\item Ordinals (displayed in red)  that  follow immediately dots are called \emph{limit ordinals}.
-With respect to the order induced by this sequence, any limit ordinal $\alpha$ is the least upper bound of  the set $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
-\item
-For instance $\omega$ is the least upper bound of the set of all finite ordinals (in the first line). It is also the first limit ordinal, and the first infinite ordinal, in the sense that 
-the set $\mathbb{O}_\omega$ is infinite.
-\item The ordinal $\epsilon_0$ is the first number which is equal to its own exponential of base $\omega$.
-It plays an important role in proof theory, and is particularly studied in chapters~\ref{chap:T1} to \ref{chap:alpha-large}.
-\item Any ordinal is  either the ordinal \textcolor{blue}{$0$},
-a successor ordinal, or a \textcolor{red}{limit ordinal}.
-\end{itemize}
-
-\section{The mathematical point of view}
-
-\subsection{Well-ordered sets}
-Let us start with some definitions.
-A  \emph{well-ordered set} is a set provided with a binary relation $<$ which has the following properties.
-\begin{description}
-\item[irreflexivity] : $\forall x\in A, x\not< x$
-\item[transitivity] : $\forall x\,y\,z\in A, x<y \Rightarrow y<z \Rightarrow x<z$
-\item[trichotomy]: $\forall x\,y\in A, x<y \vee x = y \vee y < x$
-\item[well foundedness]: $<$ is well-founded (every element of $A$ is accessible)\footnote{In classical mathematics, we would say that there is no infinite sequence $a_1>a_2> \dots a_n> a_{n+1}\dots$ in $A$.
-In contrast, \coq's standard library contains
-an inductive definition of a predicate \texttt{Acc} which allows us to write 
-constructive proofs of accessibility (See \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}).}.
-\end{description}
-
-The best known examples of well-ordered sets are the set $\mathbb{N}$ of natural numbers (with the usual order $<$), as well as any finite segment $[0,i)=\{j\in\mathbb{N}\,|\,j<i\}$.
-The disjoint union of two copies of $\mathbb{N}$, \emph{i.e.} the set $\{0,1\}\times\mathbb{N}$ is also well-ordered,
-with respect to the order below:
-
-\begin{align*}
-(i,j) < (i,k) & \;\textit{\textbf{if} }\; j < k\\
-(0,k) < (1,l) & \;\textit{\textbf{for\,any}}\;k \;\textit{\textbf{and}} \; l
-\end{align*}
-
-\subsection{Ordinal numbers}
-
-\index{maths}{Ordinal numbers}
-
-Let $(A,<_A)$ and $(B,<_B)$ two well-ordered sets. $A$ and $B$ are said to have \emph{the same order type} if 
-there exists a strictly monotonous bijection $b$ from $A$ to $B$, \emph{i.e.} which verifies the proposition
-$\forall x\,y\in A,\, x <_A y \Rightarrow b(x) <_B  b(y)$.
-
-Having the same order type is an equivalence relation between well-ordered sets. Ordinal numbers (in short: \emph{ordinals})  are descriptions (\emph{names}) of the equivalence classes.
-For instance, the order type of $(\mathbb{N},<)$ is associated with the ordinal called  $\omega$, and the order we considered on 
-the disjoint union of $\mathbb{N}$ and itself is named $\omega+\omega$.
-
-In a set-theoretic framework, one can consider any ordinal $\alpha$ as a well-ordered set, whose  elements are just the ordinals strictly less than $\alpha$, \emph{i.e.} the \emph{segment} $\mathbb{O}_\alpha=[0, \alpha)$.
-So, one can speak about \emph{finite}, \emph{infinite}, \emph{countable}, etc., ordinals.
-Nevertheless, since we work within type theory,
-we do not identify ordinals as sets of ordinals, but consider the correspondence between ordinals and sets of ordinals as the function that maps $\alpha$ to $\mathbb{O}_\alpha$.
-For instance $\mathbb{O}_\omega=\mathbb{N}$, and $\mathbb{O}_7=\{0,1,2,3,4,5,6\}$.
-
-We cannot cite all the literature published on ordinals since Cantor's book
-\cite{cantorbook}, and 
-leave it to the reader to explore the bibliography.
-
-The introduction of Jos\'e Grimm's report~\cite{grimm:hal-00911710} contains a nice presentation of the main properties of ordinals.
-
-\section{Ordinal numbers in Coq}
-
-Two kinds of representation of ordinals are defined in our development.
-
-\begin{itemize}
-\item A ``mathematical'' representation of the set of countable ordinal numbers, after Kurt Schütte~\cite{schutte}.
-This representation uses several (hopefully harmless) axioms.
-We use it as a reference for proving the correctness of ordinal notations.
-
-  \index{gaiabridge}{Introduction}
-  \gaiasign
-  We are also progressively importing \gaia's~\cite{Gaia} definitions and theorems, based on Bourbaki's set theory~\cite{Grimm2016}.
-Chapter~\ref{gaia-chapter} is dedicated to the connexion between libraries \HydrasLib and \gaia.
-
-\item A family of \emph{ordinal notations} (also called \emph{[ordinal] notation systems}), \emph{i.e.} data types used to represent segments $[0,\mu)$, where $\mu$ is some countable ordinal.
-Each ordinal notation is defined inside the Calculus of Inductive Constructions (without axioms).
-Many functions are defined, allowing proofs by computation.
-Note that proofs of
-correctness of a given ordinal notation with respect to Schütte's model obviously use axioms.
-Please execute the \texttt{Print Assumptions} command in case of doubt.
-\end{itemize}
-
-% \section{Countable ordinals}
-
-% Chapter~\ref{chap:schutte} of this document presents an adaptation to \coq{} of an axiomatization in classical logic of the set of countable ordinals by K. Schütte~\cite{schutte}. 
-% That formalization is quite complex, technical and unshamedly non-constructive,  so we put its description  in the last chapter of this document.
-
-% Please note that Schütte considers the (uncountable) set $\mathbb{O}$ of all countable ordinals.
-This set is well ordered (which is one of Schütte's axioms), and associates to any ordinal $\alpha$ the segment $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
-
-% In our adaptation to \coq{}, we declare a type \texttt{Ord}, a binary relation \texttt{lt} (with infix notation \texttt{"\_<\_"}, and assume Schütte's axiom. In Chapter~\ref{chap:schutte},
-% we derive some interesting properties of countable ordinals from these axioms.
-
-It is interesting to compare proofs of a given property (for instance the associativity of addition) both in the computational framework of some ordinal notation, and in mathematical  models
-of Schütte or \gaia.
-
-\section{Ordinal Notations}
-
-Fortunately, the ordinals we need for  studying hydra battles are much simpler than Schütte's, and can be represented as quite simple data types in \gallina.
-
-Let $\alpha$ be some (countable) ordinal; 
-in \coq{} terms, we call \emph{ordinal notation for $\alpha$} a structure composed 
-of:
-\begin{itemize}
-\item A data type $A$ for representing all ordinals strictly below $\alpha$,
-\item A well founded order $<$ on $A$, 
-\item A correct function for comparing two ordinals. Note  that the reflexive closure of $<$ is thus a \emph{total order}.
-\end{itemize}
-
-Such a structure may be proved correct relatively to a bigger ordinal notation or to Schütte's or \gaia's model.
-
-\subsection{Classes for ordinal notations}
-
-From the \coq{} user's point of view, an ordinal notation is
-a structure allowing to compare two ordinals by computation, and proving by well-founded induction.
-
-\subsubsection{The \texttt{Comparison} and \texttt{Comparable} classes}
-
-We use the operational class \texttt{Comparison} of comparison functions to define the \texttt{Comparable} class, contributed by Jérémy Damour and Théo Zimmermann, which allows us to apply generic lemmas and tactics about decidable strict orders.
-The correctness of the comparison function is expressed through Stdlib's type 
-\texttt{Datatypes.CompareSpec} as specialized by \texttt{Datatypes.CompSpec}.
-
-\begin{Coqsrc}
-  Inductive CompareSpec (Peq Plt Pgt : Prop) :
-  comparison -> Prop :=
-    CompEq : Peq -> CompareSpec Peq Plt Pgt Eq
-  | CompLt : Plt -> CompareSpec Peq Plt Pgt Lt
-  | CompGt : Pgt -> CompareSpec Peq Plt Pgt Gt.
-
-Definition CompSpec {A} (eq lt : A -> A -> Prop) (x y : A) :=
- CompareSpec (eq x y) (lt x y) (lt y x).
-\end{Coqsrc}
-
-\emph{From Module~\href{../theories/html/hydras.Prelude/mparable.html\#Hvariant}{Prelude.Comparable}}
-
-\label{sect:comparable-def}
-
-\input{movies/snippets/Comparable/ComparableDef}
-
-\subsubsection{The \texttt{ON} class}
-
-The following class definition, parameterized with a type $A$, a binary relation \texttt{lt} on $A$, specifies that \texttt{lt} is a well-founded strict order, provided with a correct comparison function.
-
-\vspace{4pt}
-\noindent\emph{From
-Library~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}}
-
-\label{types:ON}
-\index{hydras}{Library OrdinalNotations!Type classes!ON}
-
-\input{movies/snippets/ON_Generic/ONDef}
-
-We give  also a few handy definitions and lemmas for any ordinal notation.
-
-\label{sect:on-lt-notation}
-\label{sect:on-le-notation}
-\label{sect:measure-ON}
-\label{sect:bigO-ON}
-
-\inputsnippets{ON_Generic/ONDefsa}
-\inputsnippets{ON_Generic/ONDefsb}
-
-\begin{remark}
-The infix notations \texttt{o<} and \texttt{o<=} are defined in order to make apparent the distinction between the various notation scopes that may co-exist in a same statement.
-So the infix \texttt{<} and \texttt{<=} are reserved to the natural numbers.
-In the mathematical formulas, however, we still use $<$ and $\leq$ for comparing ordinals.
-\end{remark}
-
-% \subsection{Ordinal notations and  termination measures}
-% \label{sect:measure-ON}
-
-% The following lemma (together with the type class mechnism) allows us to define termination measures over any ordinal notation.
-It is just an application of  the libraries \texttt{Coq.Wellfounded.Inverse\_Image}
-% and  \texttt{Coq.Wellfounded.Inclusion}.
-
-% \begin{Coqsrc}
-% Definition measure_lt {A:Type}{lt: relation A}
-%             {compare : A -> A -> comparison}
-%             {on : ON lt compare}
-%             {B : Type} (m : B -> A) : relation B :=
-%              fun x y => on_lt (m x) (m y).
-
-% Lemma wf_measure  {A:Type}(lt: relation A)
-%             {compare : A -> A -> comparison}
-%             {on : ON lt compare}
-%             {B : Type}
-%             (m : B -> A):  well_founded (measure_lt m). 
-% \end{Coqsrc}
-
-% A simple example of application is given in Sect.~\vref{sect:merge-example}.
-
-\section{Example: the ordinal \texorpdfstring{$\omega$}{omega}}
-
-The simplest example of ordinal notation is built over the type \texttt{nat} of \coq's standard library. We just have to apply already proven lemmas about Peano numbers.
-
-\vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_Omega.html}{OrdinalNotations.ON\_Omega}}
-
-\inputsnippets{ON_Omega/OmegaDefa, ON_Omega/OmegaDefb}
-
-\section{Sum of  two ordinal notations}
-
-Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective types \texttt{A} and \texttt{B}.
-
- We consider a new strict order
-on the disjoint sum of the associated types, by putting all elements of \texttt{A} before the elements of \texttt{B} (thanks to Standard Library's relation operator \texttt{le\_AsB}).
-
-\vspace{4pt}
-\noindent
-\emph{From Library~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}}.
-
-\begin{Coqanswer}
-Inductive
-le_AsB (A B : Type) (leA : A -> A -> Prop) (leB : B -> B -> Prop)
-  : A + B -> A + B -> Prop :=
-| le_aa : forall x y : A, leA x y -> le_AsB A B leA leB (inl x) (inl y)
-| le_ab : forall (x : A) (y : B), le_AsB A B leA leB (inl x) (inr y)
-| le_bb : forall x y : B, leB x y -> le_AsB A B leA leB (inr x) (inr y)
-\end{Coqanswer}
-
-%\pagebreak
-
-\vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_plus.html}{OrdinalNotations.ON\_plus}}
-
-\input{movies/snippets/ON_plus/Defs}
-
-In order to build an instance of \texttt{Comparable}, we have to define a correct comparison function.
-
-\inputsnippets{ON_plus/compareDef, ON_plus/compareCorrect}
-\inputsnippets{ON_plus/plusComp}
-
-Standard library's lemma \texttt{Wellfounded.Disjoint\_Union.wf\_disjoint\_sum} 
-is applied  to prove that our order \texttt{lt} is well-founded, allowing us to build an instance of \texttt{ON}:
-
-\inputsnippets{ON_plus/ltWf, ON_plus/OnPlus}
-
-\subsection{The ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
-
-The ordinal $\omega+\omega$ (also known as $\omega\times 2$) may be represented as the concatenation 
-of two copies of $\omega$ (Figure~\ref{fig:omega-plus-omega}).
-It is also represented by the two first lines of Figure~\ref{fig:ordinal-sequence}.
-
-\begin{figure}[h]
-   \centering
-   \begin{tikzpicture}[very thick, scale=0.5]
-\begin{scope}[color=blue]
-\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
-\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
-\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
-\node (Adots) at (6,0) {$\ldots$};
-\node(An) at (8,0)[label=below:$n$]{$\bullet$};
-\node(A2) at (10,0)[label=below:$n+1$]{$\bullet$};
-\node (Adots1) at (12,0) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red]
-\node(B0) at (14,0)[label=below:$0$,label=above:\textcolor{red}{$\omega$}]{$\bullet$};
-\node(B1) at (16,0)[label=below:$1$, label=above:$\omega+1$]{$\bullet$};
-\node(B2) at (18,0)[label=below:$2$,label=above:$\omega+2$]{$\bullet$};
-\node (Bdots) at (20,0) {$\ldots$};
-\node (Bn) at (22,0) [label=below:$p$, label=above:$\omega+p$]{$\bullet$};
-\node (Bdots2) at (24,0) {$\ldots$};
-\end{scope}
-\end{tikzpicture}
-   \caption{\textcolor{blue}{$\omega+{\color{red}\omega}$}}
-   \label{fig:omega-plus-omega}
- \end{figure}
-
-We can define this notation in \coq{} as an instance of \texttt{ON\_plus}.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/OmegaPlusOmegaDef}
-
-\vspace{4pt}
-
-We can now define abbreviations.
-For instance, the finite ordinals are represented by terms built with  the constructor \texttt{inl}, and the first infinite ordinal $\omega$ by the term \texttt{(inr 0)}.
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Omega_plus_omega/finiteOmega}
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Omega_plus_omega/ltOmega}
-
-% \label{warning:coercions}
-% \index{Coq!Coercions} 
-% \begin{remark}
-% Beware of coercions and notation scopes!
-% Let us consider the following goal:
-
-% \begin{Coqsrc}
-%  Goal (6 o< 8).
-%  auto with arith.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% 1 subgoal (ID 9)
-
-%   ============================
-%   6 o< 8
-% \end{Coqanswer}
-
-% Please keep in mind that the current notation scope interprets the infix \texttt{``<''} as the predicate \texttt{Omega\_plus\_omega.lt} and not \texttt{Nat.lt}.
-More,  the coercion mechanism converts the terms \texttt{6:nat} [resp.
-\texttt{8:nat} ]
-% into \texttt{inl 6} [resp. \texttt{inl 8}].  So, the initial goal is correctly interpreted by \coq{}, but not as an inequality between two natural numbers.
-
-% \begin{Coqsrc}
-% Set Printing All.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% 1 subgoal (ID 337)
-
-%   ============================
-%   @on_lt nat Peano.lt Nat.compare Omega (S (S (S (S (S (S O))))))
-%     (S (S (S (S (S (S (S (S O))))))))
-% \end{Coqanswer}
-
-% Anyway, the initial goal is provable, using \texttt{le\_AsB}'s first constructor.
-
-% \begin{Coqsrc}
-%   constructor; auto with arith.
-% Qed.
-% \end{Coqsrc}
-
-% \end{remark}
-%
-
-\section{Limits and successors}
-
-Let us look again at our implementation of $\omega+\omega$. We can classify its elements into three categories:
-
-\begin{itemize}
-\item The least ordinal, \texttt{(inl 0)}, also known as  \texttt{(fin 0)}.
-\item The first infinite ordinal $\omega$.
-\item The remaining ordinals, either of the form \texttt{(inl (S $i$))} or \texttt{(inr (S $i$))} (in black on Figure~\ref{fig:ordinal-sequence}), called \emph{successor ordinals}.
-\end{itemize}
-
-\subsection{Definitions}
-It would be interesting to specify at the most generic level, what is a zero, a successor or a limit ordinal. Let $<$ be a strict order on a type $A$.
-
-\begin{itemize}
-\item A \emph{least} element is a minorant (in the large sense) of the full set  on $A$,
-\item $y$ is a \emph{successor} of $x$ if $x<y$ and there is no element between $x$ and $y$. We will also say that $x$ is a \emph{predecessor} of $y$.
-\item $x$ is a \emph{limit} if $x$ is not a least element, and for any $y$ such that $yo<x$,
- there exists some $z$ such that $y<z<x$.
-\end{itemize}
-
-The following definitions are in Library \href{../theories/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
-
-\input{movies/snippets/MoreOrders/Defs}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove, that, in any ordinal notation system, every ordinal has at most one predecessor, and at most one successor.
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/predSuccUnicity.v}{exercises/ordinals/predSuccUnicity.v}.}
-
-\end{exercise}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove, that, in any ordinal notation system, if $\beta$ is a successor of $\alpha$,
-then for any $\gamma$, $\gamma<\beta$ implies 
-$\gamma\leq\alpha$.
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/lt_succ_le.v}{exercises/ordinals/lt\_succ\_le.v}.}
-\end{exercise}
-
-\subsection{Limits and successors in \texorpdfstring{$\omega+\omega$}{omega+omega}}
-
-Using the definitions above, we can prove the following lemma:
-
-\vspace{4pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/limitIff}
-
-\vspace{4pt}
-
-Regarding successors, let us define the following function and prove its correctness:
-
-\input{movies/snippets/ON_Omega_plus_omega/succDef}
-
-\input{movies/snippets/ON_Omega_plus_omega/succCorrect}
-
-\vspace{4pt}
-
-We can also check whether an ordinal is a successor by a simple computation:
-
-\input{movies/snippets/ON_Omega_plus_omega/succb}
-
-\vspace{4pt}
-
-Finally, the nature of any ordinal is decidable (inside this notation system) :
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/ZeroLimitSuccDec}
-
-\section{Product of ordinal notations}
-
-Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective  ordered types \texttt{A} and \texttt{B}.
-The product of \texttt{NA} and \texttt{NB} is considered as the concatenation of $B$ copies of $A$, ordered by the lexicographic order on $B\times A$.
-
-In \coq{}, we build an instance of class \texttt{ON} through a sequence of steps as for the sum of ordinal notations.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_mult.html}{OrdinalNotations.ON\_mult}}
-
-\inputsnippets{ON_mult/Defs}
-\inputsnippets{ON_mult/multComp,  ON_mult/ONMult}
-\inputsnippets{ON_mult/endDefs}
-
-\section{The ordinal \texorpdfstring{$\omega^2$}{omega2}}
-
-The ordinal $\omega^2$ (also called $\phi_0(2)$, see Chap.~\ref{chap:schutte}), is an instance of the multiplication presented in the preceding section.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
-
-\inputsnippets{ON_Omega2/Omega2Def, ON_Omega2/Defs}
-\inputsnippets{ON_Omega2/Ex1}
-
-\subsection{Arithmetic of \texorpdfstring{$\omega^2$}{omega^2}}
-
-\subsubsection{Successor}
-
-The successor of any ordinal is defined by a simple pattern-matching.
-
-\input{movies/snippets/ON_Omega2/succ}
-
-This function is proved to be correct w.r.t. the \texttt{Successor} predicate.
-
-\input{movies/snippets/ON_Omega2/succOK}
-
-\input{movies/snippets/ON_Omega2/succLemmas}
-
-\subsubsection{Addition}
-
-We can define on \texttt{Omega2} an addition which extends the addition on \texttt{nat}. Please note that this operation is not commutative:
-
-\inputsnippets{ON_Omega2/plusDef, ON_Omega2/plusExamples}
-
-\subsubsection{Multiplication}
-
-The restriction of ordinal multiplication to the segment $[0,\omega^2)$ is not a total function.
-For instance $\omega\times\omega= \omega^2$ is outside the set of represented values.
-Nevertheless, we can define two operations mixing natural numbers and ordinals.
-
-\input{movies/snippets/ON_Omega2/multFinDef}
-\input{movies/snippets/ON_Omega2/multFinExamples}
-
-Multiplication with a finite ordinal and addition are related through the following lemma:
-
-\inputsnippets{ON_Omega2/uniqueDecompositiona,
-  ON_Omega2/uniqueDecompositionb,
-  ON_Omega2/uniqueDecompositionz}
-
-\subsection{A proof of termination using \texorpdfstring{$\omega^2$}{omega^2}} 
-\label{sect:merge-example}
-
-Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily a total function which merges two lists (example contributed by Pascal Manoury).
-
-%\index{coq}{Commands!Function}
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
-
-\inputsnippets{ON_Omega2/Merge, ON_Omega2/Mergeb}
-
-\subsection{Yet another  proof of impossibility}
-\label{omega2-case}
-
-In Sect.~\vref{omega-case}, we proved that there exists no variant from \texttt{Hydra} to \texttt{(nat,$<$)}
-(\emph{i.e.} the ordinal $\omega$) for proving the termination of all hydra battles.
-We  prove now that  the ordinal $\omega^2$ is also insufficient for this purpose.
-
-The proof we are going to comment has exactly the same structure as in Section~\ref{omega-case}.
- Nevertheless, the proof of technical  lemmas is a little more complex, due to 
- the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
-Consider for instance that there exists an infinite number of ordinals  between
-$\omega$ and $\omega\times 2$.
-
-The detailed  proof script is in the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/Hydra/Omega2_Small.v}{theories/ordinals/Hydra/Omega2\_Small.v}.
-
-\subsubsection{Preliminaries}
-Let us assume there is a variant from \texttt{Hydra} into $\omega^2$  for proving the   termination of all hydra battles.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
-
-\input{movies/snippets/Omega2_Small/Impossibility}
-\input{movies/snippets/Omega2_Small/Impossibilitya}
-
-We  follow the same pattern as in Sect.~\ref{omega-case}.
-First, we define an injection $\iota$ from type \texttt{ON\_Omega2.t} into \texttt{Hydra}, by
- associating to  each ordinal $\omega\times i+ j = (i,j)$ the hydra with $i$ branches of length $2$ and
-$j$ branches of length $1$.
-
-%% revenir ici
-
-\vspace{4pt}
-\emph{From Module ~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
-
-\input{movies/snippets/Omega2_Small/Impossibilityc}
-
-For instance, Figure~\vref{fig:essai2} shows the hydra associated to the ordinal 
-$(3,5)$, a.k.a. $\omega\times 3 + 5$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-\node (foot) at (6,0) {$\bullet$};
-\node (N1) at (1,3) {$\bullet$};
-\node (N2) at (3,3) {$\bullet$};
-\node (N3) at (5,3) {$\bullet$};
-\node (N4) at (8,3) {$\Smiley[2][green]$};
-\node (N5) at (11,3) {$\Smiley[2][green]$};
-\node (N6) at (14,3) {$\Smiley[2][green]$};
-\node (N7) at (17,3){$\Smiley[2][green]$};
-\node (N8) at (20,3){$\Smiley[2][green]$};
-\node  (N9) at (0,5) {$\Smiley[2][green]$};
-\node (N10) at (2,5) {$\Smiley[2][green]$};
-\node (N11) at (4,5) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N5);
-\draw (foot) -- (N6);
-\draw (foot) to [bend right=10] (N7);
-\draw (foot) to [bend right=15] (N8);
-\draw (N1) to [bend left=10] (N9);
-\draw (N2) -- (N10);
-\draw (N3) -- (N11);
-\end{tikzpicture}
-\caption{\label{fig:essai2}
-The hydra $\iota(\omega\times 3+5)$}
-\end{figure}
-
-Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:h-omega2-small}}
- The hydra \texttt{big\_h}.
-\end{figure}
-
-\input{movies/snippets/Omega2_Small/Impossibilityb}
-
- In a second step, we build a ``smaller'' hydra\footnote{With respect to the measure $m$.}.
-
-\input{movies/snippets/Omega2_Small/Impossibilityd}
-
-\vspace{4pt}
-
-Like in Sect.~\ref{omega-case}, we prove the inequality $m\;\texttt{big\_h} \,\texttt{o<=}\, m\;\texttt{small\_h} \,\texttt{o<=}\, m\;\texttt{big\_h}$, which is obviously false.
-
-\subsubsection{Proof of the inequality \texttt{m small\_h o< m big\_h}}
-
-In order to prove the inequality  \texttt{m\_lt: m small\_h o< m big\_h}, it suffices to
-build a battle transforming \texttt{big\_h} into \texttt{small\_h}.
-
-First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h} as $(i,j)$.
-If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
-If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the following result.
-
-\input{movies/snippets/Omega2_Small/bigToSmall}
-
-Since $m$ is a variant, we infer the following inequality:
-
-\input{movies/snippets/Omega2_Small/mLt}
-
-\subsubsection{Proof of the inequality \texttt{m big\_h o<= m small\_h} }
-
-The proof of the inequality \texttt{m big\_h o<= m small\_h} is quite more complex than in Sect~\ref{omega-case}.
-If we consider any ordinal $\alpha=(i,j)$, where $i>0$, there exists an infinite number of
-ordinals strictly less than $\alpha$, and there exists an infinite number of battles that start from
-$\iota(\alpha)$.
-Indeed, at any configuration $\iota(k,0)$, where $k>0$, the hydra can freely choose any replication number.
-Intuitively, the measure of such a hydra must be large enough for taking into account
-all the possible battles issued from that hydra.
-Let us now give more technical details.
-
-The first steps of our proof start a well-founded induction over $\omega^2$.
-
-\input{movies/snippets/Omega2_Small/mGe}
-
-A case analysis on $i$ and $j$ allows us to
-consider three cases :
-
-\begin{itemize}
-\item $i=j=0$: the inequality is trivial.
-\item $i=1+l, j=0$ ($(i,j)$ is a limit ordinal): By the induction hypothesis \texttt{IHij},
-  $(l,k)\texttt{o<=} m(\iota(l,k))$ for any $k$. But (by the rules of the hydra game), $\iota(i,0)$ is transformed into any $\iota(l,k)$ in one round. Thus $m(\iota(l,k)) < m(\iota(i,0))$ for any $k$.
-  Therefore, $(l,k) <  m(\iota(i,0))$ for any $k$, thus
-  $(i,0)\,\texttt{o<=}\,m(\iota(i,0))$.
- \item $j= l+1$  ($(i,j)$ is a successor). We apply the induction hypothesis to the pair $(i,l)$.
- \end{itemize}
-
- Please look at the proof script for more details.
-
- \input{movies/snippets/Omega2_Small/mGeb}
-
-\subsubsection{End of the proof}
-From \texttt{m\_ge}, we get \texttt{$m$\, big\_h o<= $m$\, small\_h = $m$\,(iota (m big\_h)) }. 
-Since $<$ is a strict order (irreflexive  and transitive), this inequality is incompatible with the strict inequality  \texttt{m small\_h o< m big\_h} (lemma \texttt{m\_lt}).
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#Impossible}{Hydra.Omega2\_Small}}
-
-  \input{movies/snippets/Omega2_Small/Impossible}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into $\omega^2$ for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\begin{remark}
- We  prove in Chapter~\ref{ks-chapter} a generalization of the impossibility lemmas of
-Sect.~\ref{omega-case} and this section, with the same proof structure, but with much more 
-complex technical details.
- \end{remark}
-
-% \index{Exercises}
-% \begin{exercise}
-
-% \label{sec:orgheadline63}
-% Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.
-
-% \begin{itemize}
-% \item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
-%   lexicographic ordering
-% \item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.
-
-% For instance, the following inequality holds:
-% \[\langle 4,3,3,3,3,3,3,2,2,2 \rangle\,<\,\langle 4,4,2 \rangle\]
-% \end{itemize}
-
-% \end{exercise}
-
-\section{A notation for finite ordinals}
-
-Let $n$ be some natural number. The segment associated with $n$ is the interval 
-$[0,n)\,=\,\{0,1,\dots,n-1\}$. 
-One may represent the ordinal $n$ by a sigma type.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
-
-\label{def:Finite-ord-type}
-
-\input{movies/snippets/ON_Finite/Defs}
-
-\subsubsection{Examples}
-
-For instance, let us build two elements of the segment $[0, 7)$, \emph{i.e.} two
-inhabitants of   type (\texttt{t 7}), and prove a simple  inequality (see Fig.~\ref{fig:O7}).
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (N0) at (0,0) {$\bullet$};
-\node (i0) at (0,1) {$0$};
-\node (N1) at (2,0) {$\bullet$};
-\node (i1) at (2,1) {$1$};
-\node (N2) at (4,0) {$\bullet$};
-\node (i2) at (4,1) {$2$};
-\node (N3) at (6,0) {$\bullet$};
-\node (i3) at (6,1) {$3$};
-\node (N4) at (8,0) {$\bullet$};
-\node (i4) at (8,1) {$4$};
-\node (N5) at (10,0) {$\bullet$};
-\node (i5) at (10,1) {$5$};
-\node (N6) at (12,0) {$\bullet$};
-\node (i6) at (12,1) {$6$};
-\node(alpha1) at (4,-1) {$\alpha_1$};
-\node(alpha2) at (10,-1) {$\beta_1$};
-\end{tikzpicture}
-
-\caption{The segment $\mathbb{O}_7$\label{fig:O7}}
-\end{figure}
-
-%\index{coq}{Commands!Program}
-
-\input{movies/snippets/ON_Finite/Example1}
-
-Note that the type (\texttt{t 0}) is empty, and that, for any natural number
- $n$, $n$ does not belong to (\texttt{t $n$}).
-
- \inputsnippets{ON_Finite/t0Empty, ON_Finite/bad10}
-
-\subsubsection{Comparison function}
-
-In order to build an instance of \texttt{ON}, we define a comparison function,  and prove its correctiness.
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Finite/compareDef}
-
-\begin{remark}
- The proof of \texttt{compare\_correct} uses a well-known pattern of \coq{}.
-Let us consider  the following subgoal.
-
-\input{movies/snippets/ON_Finite/compareCorrectb}
-
-Applying the tactic \texttt{f\_equal} generates a simpler subgoal.
-
-\input{movies/snippets/ON_Finite/compareCorrectc}
-
-We have now to prove that there exists at most one  proof of (\texttt{Nat.ltb x0 (S n)}). This is not obvious, but  a consequence of the following lemma of library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Eqdep_dec.html}{Coq.Logic.Eqdep\_dec}.
-
-\index{Coq}{Unicity of equality proofs}
-\label{sect:eq-proof-unicity}
-
-\begin{Coqanswer}
-eq_proofs_unicity_on :
-forall (A : Type) (x : A),
-(forall y : A, x = y \/ x <> y) -> 
-forall (y : A) (p1 p2 : x = y), p1 = p2
-\end{Coqanswer}
-
-Thus unicity of proofs of \texttt{Nat.ltb x0 (S n)}  comes from the decidability of
-equality on type \texttt{bool}.
-This is why we used the boolean function \texttt{Nat.ltb} instead of the inductive predicate \texttt{Nat.lt} in the definition of type \texttt{t $n$} (see page~\pageref{def: Finite-ord-type}).
-For more information about this pattern, please look at the numerous mailing lists and 
-FAQs on \coq{}).
-
-\inputsnippets{ON_Finite/compareCorrectd}
-
-\end{remark}
-
-\begin{remark}
-
-Please note  that attempting to compare a term  of type (\texttt{t $n$}) with a term of
-type (\texttt{t $p$})  leads to an error if $n$ and $p$ are not convertible.
-
-\input{movies/snippets/ON_Finite/Example2}
-
-\end{remark}
-
-\subsubsection{Building an instance of \texttt{ON}}
-
-Applying lemmas of the libraries \texttt{Coq.Wellfounded.Inverse\_Image}, \linebreak
- \texttt{Coq.Wellfounded.Inclusion}, and \texttt{Coq.Arith.Wf\_nat}, we prove that our
-relation \texttt{lt} is well founded.
-
-\input{movies/snippets/ON_Finite/ltWf}
-
-Now we can build our instance of \texttt{OrdinalNotation}.
-
-\input{movies/snippets/ON_Finite/ONInstance}
-
-\begin{remark}
-It is important to keep in mind  that the integer $n$ is not an ``element'' of \texttt{FinOrd $n$}.
-In set-theoretic presentations of ordinals, the set associated with the ordinal $n$ is $\{0,1,\dots,n-1\}$.
-In our formalization, the interpretation of an ordinal as a set is realized by the function \texttt{bigO} (see Section\vref{sect:bigO-ON}).
-\end{remark}
-
-\begin{remark}
- There is no interesting arithmetic on finite ordinals, since functions like successor, addition, etc.,  cannot be represented in \coq{} as \emph{total} functions.
-\end{remark}
-
-\section{See also \dots}
-
-Finite ordinals are also formalized in \ssreflect/\mathcomp~\cite{MCB}.
-In Module~\href{../theories/html/gaia_hydras.ON_gfinite.html}{gaia\_hydras.ON\_gfinite}, we build an instance of class \texttt{ON}.
-
-\inputsnippets{ON_gfinite/finordLtDef, ON_gfinite/finordON}
-
-\inputsnippets{ON_gfinite/Examples, ON_gfinite/Examplesb}
-
-  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.
-
-%%%
-
-\section{Comparing two ordinal notations}
-
-It is sometimes useful to compare two ordinal notations with respect to expressive power
-(the segment of ordinals  they represent).
-
-The following class specifies a strict inclusion of segments.
-The notation \texttt{OA} describes a segment $[0,\alpha)$, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented in \texttt{OA}).
-We require also  that the comparison functions of the two notation systems are compatible.
-
-\begin{figure}[h]
-   \centering
-   \begin{tikzpicture}[very thick, scale=0.6]
-\begin{scope}[color=blue]
-\node (A) at (0,0) {$A$};
-\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
-\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
-\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
-\node (Adots) at (6,0) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red]
-\node (B) at (0,2) {$B$};
-\node(B0) at (2,2)[label=above:$0$]{$\bullet$};
-\node(B1) at (3,2)[label=above:$1$]{$\bullet$};
-\node(B2) at (4,2)[label=above:$2$]{$\bullet$};
-\node (Bdots) at (6,2) {$\ldots$};
-\node (b) at (8,2) [label=above:$\omega$]{$\bullet$};
-\node (bsucc) at (9,2) [label=above:$\omega+1$]{$\bullet$};
-\node (Bdots2) at (10,2) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red!50!blue]
-\draw [->,thin] (A0) -- node [auto] {$\iota$} (B0);
-\draw [->,thin] (A1) -- node [auto] {$\iota$} (B1);
-\draw [->,thin] (A2) -- node [auto] {$\iota$} (B2);
-\draw [->,thin] (Adots) -- node [auto] {$\iota$} (Bdots);
-\end{scope}
-\end{tikzpicture}
-   \caption{\textcolor{blue}{$A$} is a sub-segment  of \textcolor{red}{$B$}}
-   \label{fig:subsegment}
- \end{figure}
-
-If \texttt{OB} is presumed to be correct, then we may consider that \texttt{OA} ``inherits'' its correctness from the bigger notation system \texttt{OB}.
-
-\label{types:SubON}
-\index{hydras}{Library OrdinalNotations!Type classes!SubON}
-
-following definition
-(in ~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{ON\_Generic}).
-
-\input{movies/snippets/ON_Generic/SubONDef}
-
-For instance, we prove that \texttt{Omega} is a sub-notation of
-\texttt{Omega\_plus\_Omega} (with $\omega$ as the first ``new'' ordinal, and \texttt{fin} as the injection).
-
-\vspace{10pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/Incl}
-
-We can also show that, if $i<j$, then the segment $[0,i)$ is a ``sub-segment'' of
-$[0,j)$. Since the terms  ($t\;i$) and ($t\;j$) are not convertible, we consider a ``cast'' 
-function $\iota$ from ($t\;i$) into ($t\;j$), and prove that this function is  a monotonous bijection  from ($t\;i$) to
-the segment $[0,i)$ of ($t\;j$).
-
-%\index{coq}{Commands!Program}
-
-\vspace{10pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
-
-\inputsnippets{ON_Finite/InclIJ, ON_Finite/InclIJa,
-  ON_Finite/InclIJb,ON_Finite/InclIJc, ON_Finite/InclIJd,
-ON_Finite/InclIJz}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove that \texttt{Omega\_plus\_Omega} cannot be a sub-notation of \texttt{Omega}.
-\end{exercise}
-
-\index{hydras}{Projects}
-\begin{project}
-Adapt the definition of \texttt{Hvariant} (Sect.~\ref{sect:hvariant-def}) in order to
-have an ordinal notation as argument. Prove that if $O_A$ is a sub-notation of $O_B$, then any variant defined on  $O_A$ can be automatically transformed into 
-a variant on $O_B$.
-\end{project}
-
-\section{Comparing an ordinal notation with Schütte's model}
-
-Finally, it may be interesting to compare an ordinal notation with the more theoretical model from Schütte (well, at least with our formalization of that model).
-This would be a relative proof of correctness of the considered  ordinal  notation.
-
-The following class specifies that a notation \texttt{OA} describes a segment $[0,\alpha)$,
-where $\alpha$ is a countable ordinal \emph{à la}  Schütte.
-
-\label{types:ON-for}
-\index{hydras}{Library OrdinalNotations!Type classes!ON\_correct}
-
-\input{movies/snippets/ON_Generic/ONCorrect}
-
-For instance, the following theorem tells that \texttt{Epsilon0}, our notation system for the segment $[0,\epsilon0)$ is a correct implementation of the theoretically defined  ordinal $\epsilon_0$
-(see chapter~\ref{chap:schutte} for more details).
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
-
-\input{movies/snippets/Correctness_E0/Epsilon0Correct}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Same work, but replace Schütte's model with \gaia's.
-\end{project}
-
-\begin{project}
-  When you have read Chapter~\ref{chap:schutte}, prove that the sum of two ordinal notations \texttt{ON\_plus} implements the addition of ordinals.
-\end{project}
-
-\section{Isomorphism of ordinal notations}
-
-In some cases we want to show that two notation systems describe the same segment (for instance $[0,3+\omega)$ and $[0,\omega)$\;).
-For this purpose, one may prove that the two notation systems are order-isomorphic.
-
-\index{hydras}{Library OrdinalNotations!Type classes!ON\_Iso}
-
-\label{types:ON-iso}
-
-\input{movies/snippets/ON_Generic/ONIso}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-\label{exo:i-plus-omega}
-Let $i$ be some natural number. Prove that the notation systems 
-\texttt{Omega} and (\texttt{ON\_plus (OrdFin $i$) Omega}) are isomorphic.
-
-{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$, that we prove also in larger notation systems, as well as in Schütte's model.}
-This exercise is partially solved for $i=3$ (in ~\href{../theories/html/hydras.OrdinalNotations.Example_3PlusOmega.html}{OrdinalNotations.Example\_3PlusOmega}).
-
-\end{exercise}
-
-\index{hydras}{Projects}
-\label{exo:ON-mult}
-\begin{project}
-% Define in \coq{} the product of two ordinal notations $N_A$ and $N_B$.
-% If $A$ [resp. $B$] is the underlying type of $N_A$ [resp. $N_B$], the
-% product \texttt{ON\_mult $N_A$ $N_B$} is implemented over the cartesian product $B\times A$ (with the lexicographic ordering).
-
-This exercise is about the non-commutativity of the multiplication of ordinals, reflected in ordinal notations.
-
-For instance, the
-elements of the product (\texttt{ON\_mult Omega (FinOrd 3)}) are ordered as follows.
-\[(0,0),(0,1),(0,2),(0,3),(0,4),\dots,{\color{red}(1,0),} (1,1),(1,2),\dots, {\color{red}(2,0)},(2,1),(2,2),\dots\]
-
-Note that the elements of  (\texttt{ON\_mult (FinOrd 3) Omega}) are differently ordered (without limit ordinals):
-\[(0,0),(1,0),(2,0),(0,1),(1,1),(2,1),(0,2),(1,2),(2,2),(0,3),\dots\]
-
-Prove formally  that (\texttt{ON\_mult (FinOrd $i$) Omega}) is isomorphic to
-\texttt{Omega}  whilst
-\texttt{Omega}  is a sub-notation of (\texttt{ON\_mult Omega (FinOrd $i$)}),
-for any strictly positive $i$.
-
-\textbf{Note:} Like Exercise~\ref{exo:i-plus-omega}, this project corresponds to the [in]equalities $i+\omega=\omega<\omega+i$, for any natural number $i$.
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-Consider two isomorphic ordinal notations \texttt{OA} and \texttt{OB}.
-Prove that, if \texttt{OA} [resp. \texttt{OB}] is a correct implementation 
-of $\alpha$ [resp. $\beta$], then $\alpha=\beta$.
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-\label{project:succ-limit-dec}
-Add to the class \texttt{ON} the requirement that for any $\alpha$ it is decidable whether $\alpha$ is $0$, a successor or a limit ordinal.
-
-\textbf{Hint:}   Beware of the instances associated with sum and product of notations!
-  You may consider additional fields 
-to make the sum and product of notations ``compositional''.
-
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-\label{project:on-setoid}
-Reconsider the  class \texttt{ON}, with an equivalence instead of Leibniz equality.
-\end{project}
-
-%%%% ICI ICI
-
-\section{Other ordinal notations}
-%%% TODO : Fix the multiplication function in branch FixOmegaOmega
-
-% \index{hydras}{Projects}
-
-% \begin{project}
-% The directory \texttt{theories/OmegaOmega} contains an ad-hoc formalization of $\omega^\omega$, contributed by Pascal Manoury.
-Every ordinal $\alpha$ is represented by a list $l$ whose elements are the coefficients of $\omega$ in  the Cantor normal form of $\alpha$ (in reverse order).
-For instance, the ordinal
-% $\omega^{8}\times 5 + \omega^{6}\times 8 + \omega^2\times 10 + \omega + 7$ is represented by the list \texttt{[5;0;8;0;0.0;10,1,7]}.
-
-%  Develop this representation and compare it with the other ordinal notations.
-
-% \end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Let $N_A$ be a notation system for ordinals strictly less than $\alpha$, 
-with the strict order $(A,<_A)$. Please build the notation system
-\texttt{ON\_Expl $N_A$}, on the type of multisets of elements of $A$
-(or, if preferred, the type of non-increasing finite sequences on $A$,
-provided with the lexicographic ordering on lists).
-
-For instance, let us take $N_A=\texttt{Omega}$, and take $\alpha=\langle 4,4,2,1,0\rangle$,
- $\beta=\langle 4,3,3,3,3,3,2\rangle$, and $\gamma=\langle 5\rangle$. Then $\beta<\alpha<\gamma$.
-
-In contrast the list $\langle5,6,3,3\rangle$ is not non-increasing (\emph{i.e.} sorted w.r.t. $\geq$), so it is not to be considered.
-
-Note that if the notation $N_A$ implements the ordinal 
-$\alpha$,  the new notation $\omega^{N_A}$ must implement the ordinal $\phi_0(\alpha)$, a.k.a. $\omega^\alpha$ (see chapter~\ref{chap:schutte})
-
-\end{project}
-
-\begin{remark}
- The set of ordinal terms in Cantor normal form (see Chap.~\ref{chap:T1}) and 
-in Veblen normal form (see 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}{Gamma0.Gamma0}) are shown to be ordinal notation systems, but there is a lot of work to be done in order to unify ad-hoc  definitions and proofs which were written before the definition of the \texttt{ON} type class.
-\end{remark}
-
-%------------------------------------------------------------------------
-
-\chapter[A proof of termination, using epsilon0]{A proof of termination, using ordinals below \texorpdfstring{$\epsilon_0$}{Epsilon0}}
-
-\label{cnf-math-def}
-\label{chap:T1}
-
-In this chapter, we adapt to \coq{} the well-known~\cite{KP82}  proof that Hercules eventually wins every battle, whichever the strategy  of each player.
-In other words, we present  a formal and self contained proof of termination  of all [free] hydra battles.
-First, we take from Manolios and Vroon~\cite{Manolios2005} a representation of the ordinal $\epsilon_0$ as terms in Cantor normal form.
-Then, we define a variant for hydra battles as a measure that maps any hydra to some ordinal strictly less than $\epsilon_0$.
-
-\section{The ordinal \texorpdfstring{\(\epsilon_0\)}{epsilon0}}
-\label{sec:epsilon0-intro}
-
-The ordinal \(\epsilon_0\) is the least ordinal number that satisfies 
-the equation \(\alpha = \omega^\alpha\), where \(\omega\) is 
-the least infinite ordinal.
-Thus, we can intuitively consider \(\epsilon_0\) as an
-\emph{infinite} \(\omega\)-tower.
-
-\subsection{Cantor normal form}
-\index{maths}{Cantor normal form}
-
-Any ordinal strictly less that \(\epsilon_0\) 
-can be finitely represented by a unique  \emph{Cantor normal form}, 
-that is, an expression  which is either  the ordinal \(0\) or 
-a sum  \(\omega^{\alpha_1} \times n_1 + \omega^{\alpha_2} \times n_2 + 
-  \dots + \omega^{\alpha_p} \times n_p\) where all the \(\alpha_i\) 
-are ordinals in Cantor  normal form, \(\alpha_1 > \alpha_2 > \alpha_p\), 
-and all the \(n_i\) are positive integers.
-
-An example of Cantor normal form is displayed in Fig \ref{fig:cnf-example}:
-Note that  any ordinal of
-the form \(\omega^0 \times i + 0\) is just written \(i\).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
-\node[color=blue]{$\omega^{(\omega^\omega\,+\, \omega^2 \times 8 \,+\, \omega)}+ \omega^\omega + \omega^4+ 6$};
-\end{tikzpicture}
-\caption{\label{fig:cnf-example}
-An ordinal in Cantor normal form}
-\end{figure}
-
-In the rest of this section, we define an inductive type for representing in \texttt{Coq}
-all the ordinals strictly  less than  \(\epsilon_0\), then extend some arithmetic operations
-to this type, and finally prove that our representation fits well with 
-the expected mathematical properties: the order we define is a well order, 
-and the decomposition into Cantor normal form  is consistent 
-with the implementation of the arithmetic operations of exponentiation of base \(\omega\) 
-and addition.
-
-\paragraph*{Remark}
-\label{sec:orgheadline65}
-Unless explicitly mentioned, the term ``ordinal" will be used instead of
-``ordinal strictly less than \(\epsilon_0\)" (except in Chapter~\ref{chap:schutte} where it stands for ``countable ordinal'').
-
-\subsection{A data type for  ordinals in Cantor normal form}
-\label{sec:orgheadline72}
-\label{sec:T1-inductive-def}
-
-% Our user contribution~\cite{CantorContrib} represents 
-% the set of ordinals strictly less than $\epsilon_0$ in Cantor normal form as in~\cite{Manolios2005}, and also the set
-% of ordinals strictly  less than $\Gamma_0$ in Veblen normal form.
-
-    Let us define an inductive type whose 
-constructors are respectively associated
-with the ways to build Cantor normal forms:
-
-\begin{itemize}
-\item the ordinal \(0\)
-\item the construction \((\alpha,\, n,\,\beta)  \mapsto \omega^\alpha \times (n + 1)+ \beta \quad (n\in\mathbb{N})\)
-\end{itemize}
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#T1}{Epsilon0.T1}}
-
-\label{types:T1}
-%\index{Constants!zero:T1}
-\index{hydras}{Library Epsilon0!Types!T1}
-
-\input{movies/snippets/T1/T1Def}
-
-\subsubsection{Example}
-
-\label{alpha0-def}
-For instance, the ordinal  $\omega^\omega+\omega^3\times 5+2$ is represented by the following term of type \texttt{T1}:
-
-\input{movies/snippets/T1/alpha0}
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5, level 1/.style={sibling distance=6cm},
-level 2/.style={sibling distance=35mm},  
-level 3/.style={sibling distance=17mm}]
-\node  {cons}
-  child {  node {cons}
-            child { node {cons} child {node {zero}} child {node{0}} child{node{zero}}}
-         child {node {0}}
-         child {node {zero}}}
-    child {node {0}}
-   child {node {cons} 
- child { node {cons} child {node {zero}} child {node{2}} child{node{zero}}}
-  child {node {4}}
-         child {node {cons} child {node {zero}} child {node{1}} child{node{zero}}}};
-
-\end{tikzpicture}
-
-\caption{The tree-like representation of the ordinal $\omega^\omega+\omega^3\times 5 +2$\label{fig:cnf-tree}}
-
-\end{figure}
-
-\paragraph{Remark}
-For simplicity's sake, we chose to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the construction (\texttt{cons $\alpha$ $n$ $\beta$}) is intended to represent the
-ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
-In a future version, we should replace  the type \texttt{nat} with \texttt{positive} in \texttt{T1}'s 
-definition. But this replacement would take a lot of time \dots{}
-
-\paragraph{Remark}
-The name \texttt{T1} we gave to this data-type  is proper to this development and refers
-to a hierarchy of ordinal notations.
-For instance, in Library \href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2},  the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{Gaia, schutte}).
-\noindent
-
-\input{movies/snippets/T2/T2Def}
-
-\subsection{About Gaia}
-\index{gaiabridge}{Introduction}
-\gaiasign Chapter~\vref{gaia-chapter} describes the present state of a
-project of making compatible the libraries \HydrasLib and \gaia. At present, both libraries contain their own version of the inductive type \texttt{T1}, with the same base name for the constructors.
-
-\inputsnippets{nfwfgaia/T1def}
-
-Thus, many examples of this and following chapters can be run
-in \gaia's context. We will signal any difference (notation, name) which may appear.
-A closer integration (same data type and functions) is still in project.
-
-\begin{remark}
-  \label{remark:hydra-gaia-renaming}
-  \textcolor{lookcolor}{In future releases, we plan to make some
-    \HydrasLib identifiers progressively  deprecated, in favour of \gaia's names.}
-\end{remark}
-
-\subsection{Abbreviations}
-
-Some abbreviations may help to write more concisely complex ordinal terms.
-
-\subsubsection{Finite ordinals}
-\label{sec:orgheadline67}
-
-For representing finite ordinals, \emph{i.e.} natural numbers, we first introduce a notation for terms of the form $n+1$, then define a coercion from type \texttt{nat} into \texttt{T1}.
-\label{sect:notation-FS}
-
-\label{sect:notation-F}
-
-\input{movies/snippets/T1/finiteOrds}
-\index{gaiabridge}{Finite ordinals}
-\paragraph*{\gaiasign}
-  In \texttt{gaia.ssete9}, the $n$-th finite ordinal is also written
-  \texttt{$\backslash${F}\,$n$}. There is no coercion in \gaia from \texttt{nat} to \texttt{T1}.
-
-  \vspace{4pt}
-
-  From~\href{../theories/html/gaia_hydras.HydraGaia_Examples.html}{gaia\_hydras.HydraGaia\_Examples}.
-
-  \inputsnippets{HydraGaia_Examples/Demoa}
-
-% \index{Coq!Coercions}
-% \index{Functions!Coercions@Coercions (from nat to ordinal types)}
-% \begin{remark}
-% Please refer to the remark~\pageref{warning:coercions} about the use of coercions.
-% % The use of coercions like \texttt{fin} allow us to be close to the mathematical tradition where natural numbers are ordinals too.
-% % Nevertheless, it may happen that a goal like \texttt{3 < 5} could be 
-% % interpreted as \texttt{(lt (fin 3) (fin 5))},  depending on the current notation scope.  
-% % When this misinterpretation happens, tactics like \texttt{auto with arith}, \texttt{lia} do not work!
-% % Thus, it is useful to write \texttt{(3 < 5)\%nat}  an inequality between two natural numbers. 
-% \end{remark}
-
-\subsubsection{The ordinal \(\omega\)}
-\label{sec:orgheadline68}
-
-  Since \(\omega\)'s Cantor normal form is
-i.e. \(\omega^{\omega^0}\times 1+ 0\), we can define the following abbreviation:
-
-\label{sect:omega-notation2}
-
-\input{movies/snippets/T1/omegaDef}
-
-Note that \texttt{omega} is not an identifier in \HydrasLib, thus any tactic like \texttt{unfold omega} would fail.
-
-\index{gaiabridge}{The ordinal $\omega$}
-\paragraph*{\gaiasign}
-In \texttt{gaia.ssete9}, the ordinal $\omega$ is bound to the \emph{constant}  \texttt{T1omega}.
-
-\subsubsection{The ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
-\label{sect:notation-phi0}
-We provide also a notation for ordinals of the form $\omega^\alpha$.
-
-\index{hydras}{Library Epsilon0!Notations!phi0@phi0 (exponential of base omega)}
-
-\input{movies/snippets/T1/phi0Def}
-
-\index{maths}{Additive principal ordinals}
-
-\begin{remark}
-\label{sec:orgheadline69}
-The name \(\phi_0\)
-   comes from ordinal numbers theory. In~\cite{schutte}, Schütte defines 
-$\phi_0$  as the ordering (\emph{i.e.} enumerating) function of the set  of \emph{additive principal ordinals} \emph{i.e.} strictly positive ordinals $\alpha$ that verify $\forall \beta<\alpha, \beta+\alpha=\alpha$.
-For Schütte,  $\omega^\alpha$ is just a notation for $\phi_0(\alpha)$.
-See also Chapter~\ref{chap:schutte} of this document.
-\end{remark}
-
-\paragraph*{\gaiasign}
-\index{gaiabridge}{Exponential of base $\omega$}
-In \texttt{gaia.ssete9} the identifier \texttt{phi0} is bound to a plain constant (not a notation).
-
-\subsubsection{The hierarchy of \(\omega\)-towers:}
-\label{sec:orgheadline71}
-
-The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximated by the sequence of $\omega$-towers (see also Sect~\vref{sect:epsilon0-as-limit} ).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/towerDef}
-
-For instance, Figure~\ref{fig:tower7} represents  the ordinal returned by the
- evaluation of the term \texttt{omega\_tower 7}.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
-\node[color=blue]{$\omega^{{{\omega}^{{{\omega}}^{{{\omega}}^{{\omega^{{\omega}^{\omega}}}}}}}}$};
-\end{tikzpicture}
-\caption{\label{fig:tower7}
-The $\omega$-tower of height 7}
-\end{figure}
-
-\subsection{Pretty-printing ordinals in Cantor normal form}
-\label{sect:ppT1}
-\index{hydras}{Library Epsilon0!Types!ppT1}
-
-Let us consider again the ordinal $\alpha_0$ defined in section~\vref{alpha0-def}. 
-If we ask \coq{} to print its  normal form, we get a hardly readable term of type \texttt{T1}.
-
-\input{movies/snippets/T1/alpha0Compute}
-
-The following data type defines a more readable abstract syntax for  ordinals terms in Cantor normal form:
-
-\label{types:ppT1}
-\index{hydras}{Library Epsilon0!Functions!pp@ pp (pretty printing terms in Cantor normal form)}
-
-\input{movies/snippets/T1/ppT1Def}
-
-The function \texttt{pp: T1 -> ppT1} converts any closed term of type \texttt{T1} into a human-readable expression. For instance, let us convert the term \texttt{alpha\_0}.
-
-\input{movies/snippets/T1/ppAlpha0}
-
-\paragraph*{\gaiasign}
-\index{gaiabridge}{Pretty printing Cantor normal forms}
-In ~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge},
-we define a variant of \texttt{pp} for pretty-printing terms of type \texttt{ssete9.CantorOrdinal.T1} (see Sect.~\vref{sect:gaia-ppT1}).
-
-\index{hydras}{Projects}
-\begin{project}
-Design  (in \ocaml?) a set of tools for systematically pretty printing ordinal terms in Cantor normal form.
-\end{project}
-
-\subsection{Comparison between ordinal terms}
-\label{sec:orgheadline73}
-
-% Our formalisation of Cantor Normal Form will take two steps:
-% 1 Definition of a strict order \texttt{o<} on the type \texttt{T1}, 
-% 2 Using \texttt{o<} for characterizing terms in normal form.
-
-In order to compare two terms of type \texttt{T1}, we define a recursive function \texttt{compare} that maps two ordinal tems $\alpha$ and $\beta$ to a value of type \texttt{comparison}.
-This type is defined in \coq's standard library
-\texttt{Init.Datatypes} and
-contains three constructors:  \texttt{Lt} (less than), \texttt{Eq} (equal), and
-\texttt{Gt} (greater than).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/compareDef}
-
-\label{Predicates:lt-T1}
-Please note that this definition of \texttt{lt} makes it easy to write proofs by computation, as shown by the following examples.
-
-\vspace{4pt}
-
-\input{movies/snippets/T1/ltExamples}
-
-Links between the function \texttt{compare} and the relations
-\texttt{lt} and \texttt{eq} are established through the following lemmas (~\vref{sect:comparable-def}).
-\vspace{4pt}
-
-\input{movies/snippets/T1/Instances}
-
-\paragraph*{\gaiasign}
-\index{gaiabridge}{Strict order on ordinals below $\epsilon_0$}
-
-In \gaia, the strict order \texttt{T1lt} on ordinal terms is directly defined as a boolean function (in \ssreflect's style).
-
-\inputsnippets{nfwfgaia/T1ltDef}
-
-In section~\vref{sect:lt-compat-gaia}, we show that both definitions are mutually equivalent.
-
-\subsubsection{A Predicate for Characterizing Normal Forms}
-\label{sect:t1-nf}
-
-\label{sec:orgheadline74}
-\label{sec:orgheadline75}
-Our data-type \texttt{T1} allows us to write expressions that
-are not properly in Cantor normal form as specified in Section \ref{sec:epsilon0-intro}.
-For instance, consider the following term of type  \texttt{T1}.
-
-\input{movies/snippets/T1/badTerm}
-
-This term would have been written \(\omega^1\times 2 + \omega^\omega \times 3\) in the usual mathematical notation.
-We note that the exponents of $\omega$ are not in the right (strictly decreasing) order.
-Nevertheless, with the help of the order \texttt{lt} on \texttt{T1}, we are now able to characterize
-the set of all well-formed ordinal terms:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#nf_b}{Epsilon0.T1}}
-
-\label{Predicates:nf-T1}
-
-\input{movies/snippets/T1/nfDef}
-
-\input{movies/snippets/T1/nfAlpha0}
-
-\input{movies/snippets/T1/nfBadTerm}
-
-\paragraph*{\gaiasign}
-\index{gaiabridge}{Ordinal terms in Cantor normal form}
-In \gaia, the boolean function which characterizes ordinal terms in normal form is defined as follows:
-
-\inputsnippets{nfwfgaia/T1nfDef}
-
-In Sect.~\vref{nf-gaia-compat}, we show that \gaia's \texttt{T1nf} is extensionally equivalent with \HydrasLib'  \texttt{nf}.
-
-\subsection{Making normality implicit}
-  We would like to get rid of terms of type \texttt{T1} which are not in Cantor normal form.
-A simple way to do this is to consider statements of the form 
-\texttt{forall alpha: T1, nf alpha -> $P$ alpha}, where $P$ is a predicate over type \texttt{T1}, like in the following lemma \footnote{Ordinal addition is formally defined a little later (page~\ref{sect:infix-plus-T1})}.
-
-\input{movies/snippets/T1/plusIsZero}
-
-\vspace{4pt}
-
-But this style leads to clumsy statements, and generates too many subgoals in interactive proofs (although often solved with \texttt{auto} or \texttt{eauto}).
-
-One may encapsulate conditions of the form \texttt{(nf $\alpha$)} in
-the most used predicates. For instance, we introduce the restriction of \texttt{lt} to terms in normal form, and provide a handy notation for this restriction.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Prelude.Restriction.html}{hydras.Prelude.Restriction}}
-
-\input{movies/snippets/Restriction/restrictionDef}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#LT}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/LTDef}
-
-\label{Predicates:LT-T1}
-
-For instance, in the following lemma, the condition that $\alpha$ is in normal form is included in the condition $\alpha< 1$.
-
-\input{movies/snippets/T1/LTOne}
-
-\subsubsection{\texttt{E0}: a sigma-type for \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-As we noticed in Sect.~\ref{sect:t1-nf}, the type \texttt{T1} is not a correct ordinal notation, since it contains terms that are not in Cantor normal form.
-In certain contexts (for instance in Sections~\ref{sect:L-equations}, \ref{sect:hardy},
-and \ref{sect:wainer}),  we need to define total recursive functions on well-formed ordinal terms less  than $\epsilon_0$, using the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777}.
- In order to define a type whose inhabitants represent just ordinals, we build a type gathering a term of type \texttt{T1} and a proof that this term is in normal form.
-
-\label{sect:E0-def}
-\label{types:E0}
-\index{hydras}{Library Epsilon0!Types!E0}
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html\#E0}{Epsilon0.E0}}
-
-\input{movies/snippets/E0/E0Def}
-
-Many constructs : types, predicates, functions, notations, etc., on type \texttt{T1} are adapted to \texttt{E0}.
-
-First, we declare a notation scope for \texttt{E0}, then we redefine the predicates of comparison.
-
-\input{movies/snippets/E0/E0Scope}
-\label{Predicates:Lt-E0}
-\input{movies/snippets/E0/LtLeDef}
-
-Equality in \texttt{E0} is just Leibniz equality. Note that, since \texttt{nf} is
-defined by a Boolean function, for  any term $\alpha:\texttt{T1}$, there exists at most one proof of \texttt{nf $\alpha$}, thus two ordinals of type \texttt{E0} are
-equal if and only if their projection to \texttt{T1} are equal (see also Sect.~\vref{sect:eq-proof-unicity}).
-
-\vspace{4pt}
-
-\index{coq}{Unicity of equality proofs}
-
-\inputsnippets{E0/nfProofUnicity}
-\inputsnippets{E0/E0EqIff}
-
-In order to  upgrade constants and functions from type \texttt{T1} to \texttt{E0}, we have to prove that
-the term they build is in normal form.
-For instance, let us represent the ordinals $0$ and $\omega$   as instances of the class \texttt{E0}.
-
-\vspace{4pt}
-
-\label{sect:omega-T1}
-
-\input{movies/snippets/E0/ZeroOmega}
-
-%\index{Constants!zero:T1}
-
-\paragraph*{\gaiasign}
-\index{gaiabridge}{Type of well formed ordinal terms below $\epsilon_0$}
- The library~\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines a \gaia-compatible type \texttt{E0}.
-
-\subsection{Syntactic definition of limit and successor ordinals}
-
-Pattern matching and structural recursion allow us to define boolean characterizations  of successor and limit ordinals.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/succbLimitb}
-
-The correctness of these definitions with respect to the mathematical notions of
-limit and successor ordinals is established through several lemmas. For instance,
-Lemma \texttt{canonS\_limit\_lub}, page~\pageref{lemma:canonS-limit}, shows that
-if $\alpha$ is (syntactically) a limit ordinal, then it is the least upper bound of
-a strictly increasing sequence of ordinals.
-
-\paragraph*{\gaiasign}
-\index{gaiabridge}{Limit and successor ordinals}
-In \gaia, the boolean functions associated with limit and successor ordinals are called \texttt{T1limit} and \texttt{T1is\_succ}.
-
- From~\href{../theories/html/gaia_hydras.HydraGaia_Examples.html}{gaia\_hydras.HydraGaia\_Examples}.
-
-\inputsnippets{HydraGaia_Examples/Demob}
-
-\subsection{Arithmetic on \texorpdfstring{$\epsilon_0$}{epsilon0}}
-\subsubsection{Successor}
-
-\index{hydras}{Library Epsilon0!Functions!succ}
-
-The successor of any ordinal $\alpha< \epsilon_0$ is defined by structural 
-recursion on its Cantor normal form.
-
-\label{Functions:succ-T1}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/succDef}
-
-The following lemma establishes the connection between the  function
-\texttt{succ} and the Boolean predicate \texttt{succb}.
-
-\vspace{4pt}
-
-\input{movies/snippets/T1/succbIff}
-
-\begin{exercise}[\gaiasign]
-  Look for the \gaia-theorem which corresponds to \texttt{succb\_iff}.
-\end{exercise}
-
-\subsubsection{Successor function on \texttt{E0}}
-
-The function \texttt{succ} on \texttt{T1} is extended to \texttt{E0} the following way:
-
- \emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
-
- \inputsnippets{E0/SuccOnE0}
-
-\index{hydras}{Exercises}
- \begin{exercise}
-Prove in \coq{} that for any ordinal $0< \alpha<\epsilon_0$, $\alpha$ is a limit if 
-and only if for all $\beta<\alpha$, the interval $[\beta,\alpha)$ is
-infinite.
-
-\emph{You may start this exercise with the file
-     \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/Limit_Infinity.v}{exercises/ordinals/Limit\_Infinity.v}.}
- \end{exercise}
-
-\subsubsection{Addition and multiplication}
-
-Ordinal addition and multiplication are also defined by structural recursion over the type \texttt{T1}. Please note that they use the \texttt{compare} function on some subterms of their arguments.
-
-\label{sect:infix-plus-T1}
-
-\input{movies/snippets/T1/plusDef}
-
-\input{movies/snippets/T1/multDef}
-
-\paragraph*{\gaiasign}
-We keep \gaia's base names for
-addition and multiplication of ordinal terms below $\epsilon_0$.
-Please refer to Sect.~\ref{sect:plus-mult-gaia-hydras} about compatibility of both arithmetics.
-
-\subsubsection{Examples}
-
-The following examples are instances of \emph{proofs by computation}. Please note that  addition and multiplication on \texttt{T1}
-are not commutative. Moreover,  both operations fail to be strictly monotonous in their first argument.
-
-\input{movies/snippets/T1/plusMultExamples}
-
-\input{movies/snippets/T1/notMono}
-
-The function \texttt{succ} is related with addition through the following lemma:
-
-\inputsnippets{T1/succIsPlusOne, T1/succIsPlusOnez}
-
-\subsubsection{Arithmetic on type \texttt{E0}}
-
- We define an addition in type \texttt{E0}, since the sum of two terms in normal form is in normal form too.
-
-\input{movies/snippets/T1/plusNf}
-
-\input{movies/snippets/E0/plusE0}
-\input{movies/snippets/E0/CheckPlus}
-
-\begin{remark}
-In all this development, two representations of ordinals co-exist: ordinal terms (type \texttt{T1}, notation scope \texttt{t1\_scope}, for reasoning on the tree-structure of Cantor normal forms), and ordinal terms \emph{known to be in normal form} (type \texttt{E0}, notation scope \texttt{E0\_scope}).
-Looking at the contexts displayed by \coq{} prevents you from any risk of confusion.
-\end{remark}
-
-%% To simplify !
-\index{hydras}{Exercises} 
-\begin{exercise}
-Prove that for any ordinal $\alpha:\texttt{E0}$, 
-$\omega\leq \alpha$ if and only if, for any natural number $i$,
-$i+\alpha=\alpha$.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/ge_omega_iff.v}{exercises/ordinals/ge\_omega\_iff.v}.}
-\end{exercise}
-
-\subsection{A proof by computation}
-\label{Ex42-E0}
-
-It is interesting to compare the following proof of the equality
-$\omega+42+\omega^2$ with the more theoretical proof in Sect~\vref{Ex42-schutte}.
-
-\inputsnippets{E0/Ex42}
-
-\section{Well-foundedness and transfinite induction}
-
-\index{maths}{Transfinite induction}
-
-\subsection{About  well-foundedness}
-\label{sec:T1Wf}
-\label{sec:orgheadline82}
-   In order to use \texttt{T1} for proving termination results,
-we need to prove that  our order \texttt{<} is well-founded. Then we will get \emph{transfinite induction} for free.
-
-The proof of well-foundedness of the strict order $<$ on Cantor normal forms is already 
-available in the Cantor contribution by Castéran and Contejean~\cite{CantorContrib}. That proof relies on a library on recursive path orderings written by
-E. Contejean. We present here  a direct proof of the same result, which does not require any knowledge on r.p.o.s.
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that the \emph{total} order \texttt{lt} on \texttt{T1} is not well-founded. 
-\textbf{Hint:}  You will have to build a counter-example with terms of type \texttt{T1}
-which are not in Cantor normal form.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/T1_ltNotWf.v}{exercises/ordinals/T1\_ltNotWf.v}.}
-\end{exercise}
-
-\subsubsection{A first attempt}
-\label{sec:orgheadline77}
-%\index{coq}{Well-founded induction}
-
-It is natural to try to prove by structural induction over \texttt{T1} 
-that every term in normal form is \texttt{LT}-accessible.
-
-Unfortunately, it won't work. Let us consider some well-formed term
- $\alpha=\texttt{cons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are \texttt{LT}-accessible.
-In order to prove the accessibility of $\alpha$, we have to consider
-any well formed term \(\delta\) such that \(\delta<\alpha\). 
-But nothing guarantees that \(\delta\)  is strictly  less than \(\beta\) nor \(\gamma\), and we cannot use the induction hypotheses on   \(\beta\) nor \(\gamma\).
-
-\input{movies/snippets/T1/wfLTBada}
-
-The problem comes from the too weak hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
-\(\gamma\);
-for instance \(\delta\) may be of the form
-\texttt{cons $\beta$ $p'$  $\gamma'$},
-where    \(p' < n\).
-Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.
-
-\input{movies/snippets/T1/wfLTBadz}
-
-\subsubsection{Using a stronger inductive predicate.}
-\label{sec:orgheadline78}
-\label{sec:strongly-accessible}
-
-  Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
-through \texttt{LT}, we propose to consider the following (stronger) predicate:
-
-\input{movies/snippets/T1/AccStrongDef}
-
-The following lemma is an application of the strict inequality 
-$\alpha < \omega ^\alpha$. If $\alpha$ is strongly accessible, then, by definition,
-$\omega^\alpha$ is accessible, thus $\alpha$ is \emph{a fortiori} accessible.
-
-\input{movies/snippets/T1/AccStrongStronger}
-
-Thus, it remains to prove that every ordinal strictly less than $\epsilon_0$
-is strongly accessible.
-
-% \subsubsection{Structure of the proof of well-foundedness of \texttt{LT}}
-
-\label{sec:orgheadline81}
-\label{proof-wf-epsilon0}
-\paragraph{A helper}
-\label{sec:orgheadline79}
-
-First, we prove that, for  any \texttt{LT}-accessible term $\alpha$, $\alpha$ is 
-strongly accessible too (\emph{i.e.} any well formed
-term (\texttt{cons $\alpha$ $n$ $\beta$})  is accessible).
-
-The proof is structured as an induction on $\alpha'$s accessibility. Let us consider
-any  accessible term $\alpha$.
-
-\input{movies/snippets/T1/AccImpAccStrong}
-
-Let \texttt{n:nat} and \texttt{beta:T1} such that (\texttt{cons alpha n beta}) is in normal form. 
-We prove first that \texttt{beta} is accessible,  which allows us to prove by well-founded induction on \texttt{beta}, 
-and natural induction on \texttt{n}, that (\texttt{cons alpha n beta}) is accessible.
-The proof, quite long, can be consulted in \href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}.
-
-\paragraph{Accessibility of any well-formed ordinal term}
-\label{sec:orgheadline80}
-
-Our goal is still to prove accessibility of any well formed ordinal term.
-Thanks to our previous lemmas, we are almost done (by a simple structural induction!).
-
-\input{movies/snippets/T1/nfAcc}
-
-\input{movies/snippets/T1/T1Wf}
-
-\index{maths}{Transfinite induction}
-
-Traditionnally, well-founded induction on ordinals is called \emph{transfinite induction}.
-
-\inputsnippets{T1/transfiniteRecursor, T1/transfiniteInduction}
-
-As a corollary, the order \texttt{Lt} on type \texttt{E0} is well-founded too.
-
-\inputsnippets{E0/E0LtWf}
-
-\begin{remark}
-\label{remark:a3pat}
-The alternate proof of well-foundedness using \'Evelyne Contejean's work on
-    recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the
-    library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0.Epsilon0rpo}.
- \end{remark}
-
-\subsection{An ordinal notation for  \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-We build an instance of \texttt{ON}, and prove its correctness w.r.t. Schutte's model (see Chapter~\ref{chap:schutte}).
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
-
-\input{movies/snippets/E0/InstanceEpsilon0}
-
-\label{instance-epsilon0}
-
-\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
-
-\input{movies/snippets/Correctness_E0/injectDef}
-
-\input{movies/snippets/Correctness_E0/Epsilon0Correct}
-
-\subsection{An ordinal notation for  \gaia's ordinals}
-\index{gaiabridge}{Ordinal notations}
-
-\gaiasign Module~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge} contains an instance of class \texttt{ON E0lt compare}, where \texttt{E0lt} is the order on the well-formed ordinal terms below $\epsilon_0$ defined in \gaiaHydras library (please see Chapter~\ref{gaia-chapter}).
-
-\inputsnippets{T1Bridge/gEpsilon0}
-
-\index{hydras}{Projects}
-\begin{project}
- \emph{This exercise is a continuation of Project~\vref{exo:ON-mult}.}
-Use \texttt{ON\_mult} to define an ordinal notation \texttt{Omega2} for $\omega^2=\omega\times\omega$.
-
-Prove that \texttt{Omega2} is a sub-notation of \texttt{Epsilon0}.
-
-Define on \texttt{Omega2} an addition compatible with the addition on \texttt{Epsilon0}.
-
-\textbf{Hint}. You may use the following definition (in 
-    \href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}).
-
-\input{movies/snippets/ON_Generic/SubONSameOp}
-
-     \end{project}
-
-    \index{hydras}{Projects}
-    \begin{project}
-    The class \texttt{ON} of ordinal notations has been defined long after this 
-    chapter, and is not used in the development of the type \texttt{E0} yet.
-    A better integration of both notions should simplify the development on ordinals in Cantor normal form. This integration is planned for the future versions.
-
-    \end{project}
-
-    \section{A refinement of \texttt{E0} : an ordinal notation for \texorpdfstring{$\omega^\omega$}{omega\^omega}}
-
-    In Module   \href{https://github.com/coq-community/hydra-battles/blob/master/theories/
-      ordinals/OrdinalNotations/OmegaOmega.v}{theories/ordinals/OrdinalNotations/OmegaOmega.v},
-    we represent ordinals below $\omega^\omega$ by lists of pairs of natural numbers (with the same coefficient shift as in \texttt{T1}).
-    For instance, the ordinal $\omega^4\times 10 + \omega^3 + \omega+ 5$ is represented by the list \texttt{(4,9)::(3,0)::(1,0)::(0,4)::nil}.
-
-\inputsnippets{OmegaOmega/LODef}
-
-  The usual operations : \texttt{succ}, \texttt{+}, \texttt{*} are simple variants of the same operations in \texttt{T1}.
-
-  \inputsnippets{OmegaOmega/succPlusMult}
-
-    We establish this representation as a \emph{refinement} of the data types we used to represent ordinals less than $\epsilon_0$.
-Thus, many properties like well-foundedness of $<$ and associativity of $+$,  of this ordinal notations have very short proofs.
-
-    \inputsnippets{OmegaOmega/refineDef, OmegaOmega/phi0Ref,
-      OmegaOmega/succRef, OmegaOmega/plusRef,
-      OmegaOmega/multRef}
-
-    In order to make an ordinal notation for $\omega^\omega$, we follow the same steps as for $\epsilon_0$:
-    \begin{enumerate}
-    \item Define an order \texttt{lt}, which refines the order
-      \texttt{lt} on \texttt{T1}.
-      \inputsnippets{OmegaOmega/compareDef}
-      \inputsnippets{ OmegaOmega/ltDef, OmegaOmega/ltRef}
-
-    \item Define the predicate ``to be in normal form''.
-      \inputsnippets{OmegaOmega/nfDef}
-    \item Define a class \texttt{OO} of terms in normal form, and an embedding from \texttt{E0} into \texttt{OO}.
-      \inputsnippets{OmegaOmega/OODef}
-       \inputsnippets{OmegaOmega/embedDef}
-      \item Infer well-foundedness of the order on \texttt{OO}.
-    \inputsnippets{OmegaOmega/ltWf}
-
-    \end{enumerate}
-
-    \inputsnippets{OmegaOmega/ONOO, OmegaOmega/OOz}
-
-    Let us show a few examples.
-
-    \inputsnippets{OmegaOmega/OODemo}
-
-    \index{hydras}{Exercises}
-
-    \subsection{Related work}
-    The article~\cite{Manoury2016}  defines another representation of ordinals below $\omega^\omega$ based on lists of natural numbers.
-
-\begin{exercise}[not trivial]
-  It may be interesting to write a \emph{direct} proof  of well-foundedness of the order in $\omega^\omega$ (\emph{i.e.}
-  without using properties of $\epsilon_0$). This exercise may
-  help to understand better the proof structure of Sect.~\vref{sec:strongly-accessible}.
-    \end{exercise}
-%ici
-    \section{A variant for hydra battles}
-
-    In order to prove the termination of any hydra battle, we try to define a variant mapping hydras to ordinals strictly less than $\epsilon_0$.
-    In order to make such a variant easy to define (for instance by a structural recursion), we introduce a variant of addition, which, contrary to
-    $+$, is commutative and strictly monotonous in both of its arguments. This last property makes it possible to prove that our function is 
-    truly a variant for hydra battles (in Sect.~\vref{sect:variant-decr}).
-
-    \subsection{Natural sum (a.k.a. Hessenberg's  sum)}
-    \label{sec:orgheadline87}
-    \label{hydra-variant}
-    \label{sect:hessenberg-def}
-
-    Natural sum (Hessenberg  sum) is a commutative and monotonous version of
-    addition. It is used as an auxiliary operation  for defining variants
-    for hydra battles, where Hercules is allowed to chop off any  head of the hydra.
-
-    In the literature, the natural sum of ordinals \(\alpha\) and \(\beta\)
-    is often denoted by \(\alpha \# \beta\)  or  \(\alpha \oplus  \beta\).
-    Thus we called \texttt{oplus} the associated \emph{Coq} function.
-
-    \subsubsection{Definition of \texttt{oplus}}
-    \label{sec:orgheadline84}
-    %\index{Functions!oplus @ oplus (Hessenberg commutative sum)}
-
-    The definition of \texttt{oplus} is recursive in both of its 
-    arguments and uses  the same pattern as for the \texttt{merge} function on lists of library
-    \texttt{Coq.Sorting.Mergesort}.
-
-    \begin{enumerate}
-    \item Define a nested recursive function, using the \texttt{Fix} 
-        construct
-
-    \item Build a principle of induction dedicated to \texttt{oplus}
-
-    \item Establish equations associated to each case of the definition.
-    \end{enumerate}
-
-    \paragraph{Nested recursive definition}
-    \label{sec:orgheadline83}
-
-    The following definition is composed of 
-    \begin{itemize}
-    \item A main function \texttt{oplus}, structurally recursive in its 
-    first argument \texttt{alpha}
-    \item An auxiliary function \texttt{oplus\_aux} within the scope of \texttt{alpha},
-    structurally recursive in its argument \texttt{beta};  \texttt{oplus\_aux beta} 
-       is supposed to compute  \texttt{oplus alpha beta}.
-    \end{itemize}
-
-%  \pagebreak 
-  \vspace{4pt}
-    \emph{From Module~\href{../theories/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}
-
-    \label{sect:infix-oplus}
-
-    \input{movies/snippets/Hessenberg/oplusDef}
-
-    The reader will note that each recursive call of the functions
-    \texttt{oplus} and \texttt{oplus\_aux} satisfies \emph{Coq}'s constraint
-    on recursive definitions. The function \texttt{oplus} is recursively called on a sub-term of its first argument,
-    and \texttt{oplus\_aux} on a sub-term of its unique argument.
-    Thus, \texttt{oplus}'s definition is accepted by \coq{} as a structurally recursive function.
-
-    \subsubsection{Rewriting lemmas}
-    \label{sec:orgheadline86}
-
-    \emph{Coq}'s constraints on recursive definitions result in 
-    the quite  complex form of \texttt{oplus}'s definition.
-    Proofs of properties of this function can be simpler if we
-     derive a few  rewriting lemmas that will help to simplify 
-    expressions of the form (\texttt{oplus $\alpha$ $\beta$}).
-
-    A first set of lemmas correspond to the various cases of \texttt{oplus}'s 
-    definition. They can be proved almost immediately. Here are a few examples.
-
-    \inputsnippets{Hessenberg/oplusNeutral,
-      Hessenberg/oplusCompareLt}
-
-    % \subsubsection{A hand-made induction principle}
-    % \label{sec:orgheadline85}
-
-    % \index{Coq!Commands!Functional Scheme}
-
-    % \emph{Coq} contains a command  \texttt{Functional Scheme} that 
-    % generates induction principles which correspond to recursive functions.
-    % Unfortunately, the current version ( \texttt{8.11.0} ) doesn't work on \texttt{oplus},
-    % probably because of the inner \texttt{Fix}.
-
-    % \begin{Coqsrc}
-    % Functional Scheme oplus_ind := Induction for oplus Sort Prop.
-    % \end{Coqsrc}
-
-    % \begin{Coqanswer}
-    % Error: Anomaly "todo." Please report at http://coq.inria.fr/bugs/.
-    % \end{Coqanswer}
-
-    % Fortunately, it's a good exercise for a semi-experienced user, to write
-    % her/him-self induction principles similar to the ones returned by
-    % \texttt{Functional Scheme}.
-
-    % \begin{itemize}
-    % \item First, we choose to write a version for sort \texttt{Type}, since versions
-    % for sorts \texttt{Prop} and \texttt{Set} can be easily derived from
-    % the former one. According to \emph{Coq}'s naming politics, we will call our 
-    % principle \texttt{oplus\_rect}
-
-    % \item The conclusion of \texttt{oplus\_rect} will be (\texttt{$P$ a b (oplus a b)}),
-    % where $P$ is an arbitrary function of type 
-    % \texttt{T1 -> T1 -> T1 -> Type}
-
-    % \item The premises of \texttt{oplus\_rect} will describe how to build an induction 
-    % on the graph of \texttt{oplus}.
-    % \end{itemize}
-
-    % We are now ready to state and prove \texttt{oplus\_rect}, and the reader
-    % will note that the statement is longer than the proof script itself,
-    % which is a standard proof by induction, simplification and case-analysis 
-    % that follows  \texttt{oplus}'s definition.
-
-    % We associate also a tactic to the application of \texttt{oplus\_rect}.
-
-    % \begin{Coqsrc}
-    %  Lemma oplus_rect:
-    %       forall P: T1 -> T1 -> T1 -> Type, 
-    %         (forall a:T1, P zero a a) ->
-    %         (forall a: T1, P a zero a) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Gt ->
-    %            P b1 (cons a2 n2 b2) o ->
-    %            P (cons a1 n1 b1) (cons a2 n2 b2)
-    %              (cons a1 n1 o)) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Lt ->
-    %            P (cons a1 n1 b1) b2 o ->
-    %            P (cons a1 n1 b1) (cons a2 n2 b2) 
-    %            (cons a2 n2 o)) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Eq ->
-    %            P b1 b2 o ->
-    %           P (cons a1 n1 b1) (cons a2 n2 b2)
-    %             (cons a1 (S (n1 + n2)%nat) o)) ->
-    %          forall a b, P a b (oplus a b).
-    % Proof with auto.
-    %    induction a.
-    %    -    intro; simpl; destruct b;auto.
-    %    -   induction b.
-    %        + apply X0.
-    %        + case_eq (compare a1 b1).
-    %          * intro Comp; unfold oplus; rewrite Comp.
-    %            cbn; apply X3 ...
-    %          * intro Comp; cbn; rewrite Comp; apply X2...
-    %          * intro Comp; cbn; rewrite Comp ...
-    %  Defined.
-
-    % Ltac oplus_induction a b:= pattern (oplus a b); apply oplus_rect.
-    % \end{Coqsrc}
-
-    % \index{Exercises}
-
-    % \begin{exercise}
-    % The induction principle \texttt{oplus\_rect} is still unused in our development. 
-    % Please build some nice examples of application.
-    % \end{exercise}
-
-    \index{hydras}{Projects}
-    \begin{project}
-    Compare \texttt{oplus}'s definition (with inner fixpoint) with other possibilities
-    (\texttt{coq-equations}, \texttt{Function}, etc.).
-    \end{project}
-    \subsection{More theorems on Hessenberg sum}
-\index{gaiabridge}{Hessenberg sum}
-    We need to prove some properties of $\oplus$, particularly about 
-    its relation with the order $<$ on \texttt{T1}.
-
-    \subsubsection{Commutativity, associativity}
-
-    We prove  the commutativity of $\oplus$ in two steps. 
-    First, we prove by transfinite induction on $\alpha$ that the restriction of $\oplus$ to the
-    interval $[0..\alpha)$ is commutative.
-
-    \index{maths}{Transfinite induction}
-
-    \input{movies/snippets/Hessenberg/oplusComm0}
-
-    Then, we infer  $\oplus$'s commutativity for any pair of ordinals:
-    Let $\alpha$ and $\beta$ be two ordinals strictly less than $\epsilon_0$. Both ordinals $\alpha$ and $\beta$ are
-    strictly less than $\textrm{max}(\alpha,\beta)+1$.
-        Thus, we have just to apply the lemma \coqsimple{oplus\_comm\_0}.
-
-\input{movies/snippets/Hessenberg/oplusComm}
-
-    Associativity of Hessenberg sum is proved the same way.
-
-    \input{movies/snippets/Hessenberg/oplusAssoc0}
-    \input{movies/snippets/Hessenberg/oplusAssoc}
-
-    \subsubsection{Monotony}
-
-    At last, we prove that $\oplus$ is strictly monotonous in both of its arguments.
-
-    \input{movies/snippets/Hessenberg/oplusMono}
-
-      \index{hydras}{Projects}
-
-    \begin{project}
-    The library \texttt{Hessenberg} looks too long (proof scripts and compilation).
-    Please try to make it simpler and more efficient!
-    Thanks!
-    \end{project}
-
-    \paragraph*{\gaiasign} The module
-      \href{../theories/html/gaia_hydras.~GHessenberg.html}{gaia\_hydras.GHessenberg} defines a version of Hessenberg sum
-      compatible with \gaia's type \texttt{T1}.
-      \index{gaiabridge}{Hessenberg sum}
-    \subsection{A termination measure for hydra battles }
-
-    \label{sec:hydra-measure}
-
-    Let us define a measure from type \texttt{Hydra} into \texttt{T1}.
-
-    \vspace{4pt}
-    \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Termination.html\#m}{Hydra.Hydra\_Termination}}
-
-   \input{movies/snippets/Hydra_Termination/mDef}
-
-    First, we prove that the measure $m(h)$  of any hydra $h$ is a well-formed ordinal term of type \texttt{T1}.
-
-    \input{movies/snippets/Hydra_Termination/mNf}
-      \input{movies/snippets/Hydra_Termination/msNf}
-
-    For proving the termination of all hydra battles, we have to prove that
-    \texttt{m} is a variant. First, a few technical lemmas follow the decomposition of \texttt{round} into several relations. Then the lemma \texttt{round\_decr} gathers all the cases.
-
-    \label{sect:variant-decr}
-
-    \input{movies/snippets/Hydra_Termination/S0Decr}
-    \input{movies/snippets/Hydra_Termination/R1Decr}
-    \input{movies/snippets/Hydra_Termination/S1Decr}
-    \input{movies/snippets/Hydra_Termination/R2Decr}
-     \input{movies/snippets/Hydra_Termination/RoundDecr}
-
-      Finally, we prove termination of all (free) battles.
-
-    \label{thm:every-battle-terminates}
-
-    \input{movies/snippets/Hydra_Termination/FinalThm}
-
-    \paragraph*{\gaiasign}
-    \index{gaiabridge}{Termination of all hydra battles}
-    The module~\href{../theories/html/gaia_hydras.GHydra.html}{gaia\_hydras.GHydra} contains a proof of \gaiaHydras version of termination of all battles.
-
-    \inputsnippets{GHydra/mDef, GHydra/mVariant, GHydra/Termination}
-
-    \section*{Conclusion}
-
-    Let us recall three results we have proved so far.
-    \begin{itemize}
-    \item There exists a strictly decreasing variant which maps \texttt{Hydra} into 
-    the segment $[0,\epsilon_0)$ for proving the termination of any hydra battle
-    \item There exists \emph{no} such variant from \texttt{Hydra} into 
-    $[0,\omega^2)$, \emph{a fortiori} into $[0,\omega)$.
-    \end{itemize}
-
-    So, a  natural question is `` Does there exist any strictly decreasing variant mapping
-    type \texttt{Hydra} into some interval $[0,\alpha)$ (where $\alpha <\epsilon_0$) for proving the termination of all hydra battles''.
-The next chapter is dedicated to a formal proof that there exists no such $\alpha$, even if we consider a restriction to the set of ``standard'' battles.
-
-    %\include{epsilon0}
-
-    %\include{impossibility-proofs}
-
-%-------------------------------------------------------------------
-
 \chapter[The Ketonen-Solovay machinery]{Accessibility inside \texorpdfstring{$\epsilon_0$}{Epsilon0}: The Ketonen-Solovay Machinery\label{ks-chapter}}
@@
 \end{project}
-
-%---------------------------------------------------------------------
-\chapter{Large sets and rapidly growing functions}\label{chap:alpha-large}
-
-\begin{remark}
-Some notations (mainly names of fast-growing functions) of our development may differ slightly from the literature.
-Although this fact does not affect our proofs, we are preparing a future version where the names $F\_alpha$, $f\_alpha$, $H\_alpha$, etc., are fully consistent with the cited articles.
-
-\end{remark}
-%\section{Introduction}
-
-In this chapter, we try to feel how long a standard battle can be.
-To be precise, for any ordinal $\alpha<\epsilon_0$ and any positive integer $k$,
-we give a minoration of the number of steps of a standard battle which
-starts with the hydra $\iota(\alpha)$ and the replication factor $k$.
-
-We express this number in terms of the Hardy hierarchy of fast-growing 
-functions~\cite{BW85, Wainer1970, KS81, Promel2013}.
- From the \coq{} user's point of view, such  functions are  very 
-attractive:  they are defined as functions  in \gallina{}, and we can apply them \emph{in theory}, but they are so complex that you will never be able to look at the result of the computation.
- Thus, our knowledge on these functions must rely on \emph{proofs}, not tests. In our development, we use often the rewriting rules generated by \coq's \texttt{Equations} plug-in.
-
-\section{Definitions}
-
-%\subsection{Definition}
-
-\begin{definition}
-Let $0<\alpha<\epsilon_0$ be any ordinal, and $s=\langle s_1, s_2, \dots, s_N\rangle$ a finite sequence of strictly positive natural numbers.
-
-We say that $s$ is \emph{$\alpha$-large} if the sequence $\langle \alpha_0=\alpha,\dots,\alpha_{i+1}=\canonseq{\alpha_i}{i+1},\dots \rangle$ leads to $0$. 
-We say also that $s$ is \emph{minimally $\alpha$-large} (in short:
-\emph{$\alpha$-mlarge}) if $s$ is $\alpha$-large 
- and every strict prefix of $s$ leads to a non-zero ordinal (\emph{cf} Sect.~\vref{sect:path-to-def}).
-
-\index{maths}{Ordinal numbers!Large sets}
-\index{maths}{Ordinal numbers!Minimal large sets}
-
-\end{definition}
-
-\begin{remark}
-  Ketonen and Solovay~\cite{KS81} consider  large finite \emph{sets} of natural numbers,  but they are mainly used as sequences. Thus, we chose to represent them explicitely as (sorted) lists. 
-\end{remark}
-
-The following function ``gnaws'' an ordinal $\alpha$, following a sequence of indices (ignoring the $0$s).
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/gnawDef}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_SetsPaths.html\#gnaw}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/largeDef}
-
-Minimal large sequences can be directly defined in terms of the
-predicate \texttt{path\_to} (\vref{sect:path-to-def}) which already prohibits paths containing non-final \texttt{zero}s.
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}
-
-\index{hydras}{Library Epsilon0!Predicates!mlarge@mlarge (minimal large sequences)}
-
-\input{movies/snippets/Large_Sets/mlargeDef}
-
-Let us consider two integers $k$ and $l$, such that $0<k<l$. In order to check whether the interval $[k,l]$ is minimally large for $\alpha$, it is enough to
-follow from $\alpha$ the path associated with the interval $[k,l)$ and verify that the last ordinal we obtain is equal to $1$.
-
-\subsection{Examples}
-
-For instance the interval $[6,70]$ leads $\omega^2$ to $\omega\times 2 + 56$. Thus this interval is not $\omega^2$-large.
-
-\noindent
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html\#mlarge}{Epsilon0.Large\_Sets\_Examples}}
-
-\input{movies/snippets/Large_Sets_Examples/gnawEx1}
-
-The interval $[6,700]$ is $\omega^2$-large, but not
-$\omega^2$-mlarge, since $[6,699]$ is also $\omega^2$-large.
-
-\input{movies/snippets/Large_Sets_Examples/gnawEx2}
-
-The following lemma relates minimal largeness with the
-function 
-\texttt{gnaw}.
-
-\input{movies/snippets/Large_Sets/mlargeIff}
-
-\vspace{4pt}
- \noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html}{Epsilon0.Large\_Sets\_Examples}}
-
-\input{movies/snippets/Large_Sets_Examples/Ex1Lemma}
-
-\section{Length of minimal large sequences}
-\label{sect:lalpha-section}
-
-Now, consider any natural number $k>0$ any ordinal $0<\alpha<\epsilon_0$.  We would like to compute
-a number $l$ such that the interval $[k,l]$ is $\alpha$-mlarge. So, 
-the standard battle starting with $\iota(\alpha)$ and the replication factor $k$ will end after $(l-k+1)$ steps.
-
-First, we notice that this  number $l$ exists, since the segment $[0,\epsilon_0)$ is well-founded and $\canonseq{\alpha}{i}<\alpha$ for any $i$ and $\alpha>0$.
-Moreover, it is unique:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/mlargeUnicity}
-
-Thus, we would like to define a function, parameterized by $\alpha$ which associates to any  strictly positive integer $k$ the number $l$ such that
-the interval $[k,l]$ is $\alpha$-mlarge. It would be fine to write in \gallina{} a definition like this:
-
-\begin{Coqbad}
-Function L_ (alpha: E0) (i:nat) :  nat := ...
-\end{Coqbad}
-
-But we do not know how to fill the dots yet \dots{}   In the next section, we will 
-use \coq{} to reason  about the \emph{specification} of \texttt{L},
-prove properties of any function which satisfies this specification.
-In Sect.~\ref{sect:L-equations}, we use the \texttt{coq-equations} plug-in
-to define a function \texttt{L\_}, and prove its correctness w.r.t. its specification.
-
-\subsection{Formal specification}
-
-Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider any  function which  maps  any strictly positive integer $k$ to the number $l$, where 
-the interval $[k,l)$ is $\alpha$-mlarge.
-
-\begin{remark}
-In~\cite{KS81} Ketonen and Solovay consider the least natural number $l$ where the interval $[k,l]$ ($l$ included) is $\alpha$-large, and call $H_\alpha$ the function which maps $k$ to $l$.
-We chose to consider intervals $[l,k)$ instead of $[l,k]$
-in order to simplify  some statements and proofs in composition lemmas associated with the ordinals of the form $\alpha\times i$ and 
-$\omega^\alpha\times i + \beta$.
-Clearly, both approaches are related through the equality
-$L_\alpha(k)=H_\alpha(k)+1$, for any non-null $\alpha$ and $k$.
-\end{remark}
-
-Our specification of the function \texttt{L} is as follows:
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/LSpecDef}
-
-\begin{todo}
- Check if the functions $L_\alpha$ are the same as
-\cite{KS81}' functions $f_\alpha$ (p. 297).
-\end{todo}
-
-Note that, for $\alpha\not=0$, the value of $f(0)$ is not specified.
-Nevertheless, the restriction of $f$ to the set of strictly positive integers is unique (up to extensionality).
-
-\input{movies/snippets/Large_Sets/LSpecUnicity}
-
-\subsection{Abstract properties}
-
-Let us now prove properties of any function $f$ (if any) which satisfies 
-\texttt{L\_spec}.
-We are looking for properties which could be used for writing \emph{equations} and prove the correctness of the function generated by the \texttt{coq-equations} plug-in.
-Moreover, they will give us some examples (for small values of $\alpha$).
-
-The properties we consider are defined in \href{../theories/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}.
-
-\label{sect:abstract-arith-prop}
-\input{movies/snippets/Iterates/funLeDef}
-
-Our exploration of the $L_\alpha$\,s  considers the usual cases of a proof by transfinite induction: zero, successors and limit ordinals.
-The lemmas we are going to prove will be applied in a big proof by induction in Sect~\vref{sect:L-correct-proof}.
-
-\index{maths}{Transfinite induction}
-
-\subsubsection{The  ordinal zero}
-\label{sect:L-spec-zero}
-The base case is directly a consequence of the specification.
-
-\input{movies/snippets/Large_Sets/LZeroInv}
-
-\subsubsection{Successor ordinals}
-\label{sect:L-spec-succ}
-Let $\beta$ be some ordinal, and assume the arithmetic function $f$ satisfies 
-the specification $(\texttt{L\_spec}\;\beta)$.  Let $k$ be any natural number.
-Any path from $\texttt{succ}\,\beta$ to $0$ starting at $k+1$ can be decomposed into a first step from $\texttt{succ}\,\beta$ to $\beta$, then a path from
-$\beta$ at $k+2$ to $0$. 
-By hypothesis the interval $[k+2, f(k+2)-1]$ is $\beta$-mlarge.
-But the interval $[k+1, f(k+2)-1]$ is the concatenation of the singleton
-$\{k+1\}$ and the interval $[k+2, f(k+2)-1]$.
-So, the function $\lambda\,k.\,f(k+1)$ satisfies the specification $\texttt{L\_spec}\,\beta$.
-
-Note that our decomposition of intervals works only if the intervals we consider are not empty.
-In order to ensure this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <<= f}, or \texttt{(fun\_le S f)}.
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/SectionSucc}
-
-\subsubsection{Limit ordinals}
-\label{sect:L-spec-lim}
-
-Let $\lambda<\epsilon_0$ be any limit ordinal. In a similar way as for successors, we decompose any path from $\lambda$  into a first step to
-$\canonseq{\lambda}{k}$, followed by a path to $0$. In the following section, we assume that there exists a correct function computing  $L_{\canonseq{\lambda}{k}}$ for any strictly positive $k$.
-
-\input{movies/snippets/Large_Sets/SectionLim}
-
-\subsection{First results}
-
-Applying the previous lemmas on successors and limit ordinals, 
-we obtain a few  correct implementations of \texttt{(L\_spec $\alpha$)} for small values of $\alpha$.
-
-\subsubsection{Finite ordinals}
-
-By iterating the functional \texttt{L\_succ}, we get a realization of
-\texttt{(L\_spec $i$)} for any natural number $i$.
-
-\input{movies/snippets/Large_Sets/LFinDef}
-\vspace{-16pt}
-\input{movies/snippets/Large_Sets/LFinOk}
-
-\subsubsection{The first limit ordinal  \texorpdfstring{$\omega$}{omega}}
-
-The lemmas \texttt{L\_fin\_ok} and \texttt{L\_lim\_ok}   allow us to get 
-by diagonalization a correct implementation for 
-\texttt{L\_spec omega}.
-
-\input{movies/snippets/Large_Sets/LOmegaDef}
-\vspace{-16pt}
-\input{movies/snippets/Large_Sets/LOmegaOk}
-
-\subsubsection{Towards  \texorpdfstring{$\omega^2$}{omega*omega}}
-
-We would like to get exact formulas for the ordinal $\omega^2$, a.k.a.
-$\phi_0(2)$. This ordinal is the limit of the sequence $\omega\times i\;(i \in \mathbb{N})$. Thus, we have to study ordinals of this form, then use 
-our lemma on limits.
-
-The following lemma establishes a path from $\omega\times ( i+1)$ to
-$\omega \times i$.
-
-\input{movies/snippets/Large_Sets/pathToOmegaMult}
-
-Let us consider a path from  $\omega\times(i+1)$ to $0$ starting at $k+1$.
-A first ``big step'' will lead to $\omega\times i$ at $2(k+1)$. If $i>0$, the
-next jump leads to $\omega\times(i-1)$ at $2(2(k+1))+1$, etc.
-
-The following lemma expresses the length of the mlarge sequences associated with any finite multiple of $\omega$.
-
-\input{movies/snippets/Large_Sets/omegaMultMlarge0}
-
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#L_omega_mult}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/LOmegaMultDef}
-
-More generally, we prove the equality $L_{\omega\times i}(k)=2^i\times(k+1)-1$.
-
-\input{movies/snippets/Large_Sets/LOmegaMultEqn}
-
-Correctness of the function \texttt{L\_omega\_mult} is asserted through the following lemma.
-
-\input{movies/snippets/Large_Sets/LOmegaMultOk}
-
-By diagonalization, we obtain a simple formula for $L_{\omega^2}$.
-
-\input{movies/snippets/Large_Sets/LOmegaSquare}
-\input{movies/snippets/Large_Sets/LOmegaSquareEqn}
-\input{movies/snippets/Large_Sets/LOmegaSquareOk}
-
-%%%% ICI
-
-\subsubsection{Going further}
-Let us consider a last example, ``computing'' $L_{\omega^3}$.
-Let us study the canonical sequence associated with $\omega^3$, the elements of which are 
-$\omega^2\times i\;(i\in\mathbb{N}_1)$.
-
-To this end, we prove a generic lemma, which expresses $L_{\omega^\alpha\times i}$ as an iterate of $L_{\omega^\alpha}$.
-Note that in this lemma, we assume that the function associated with $\alpha$ is strictly monotonous and
-greater or equal than the successor function, and prove that $L_{\omega^\alpha\times i}$ satisfies  the same properties.
-
-\inputsnippets{Large_Sets/phi0Mult}
-
-\inputsnippets{Large_Sets/phi0MultOk,
-Large_Sets/Phi0MultSLe}
-
-Let us look now
-at the ordinal $\omega^2\times i$, using \texttt{L\_phi0\_mult}.
-
-\input{movies/snippets/Large_Sets/LOmegaSquareTimes}
-
-We are now ready to get an exact formula for $L_{\omega^3}$, by diagonalization over $L_{\omega^2\times i}$.
-
-\input{movies/snippets/Large_Sets/LOmegaCube}
-
-Thus, for instance, $L_{\omega^3}(3)=L_{\omega^2\times 4}(3)$.
-
-\input{movies/snippets/Large_Sets/LOmegaCube3Eq}
-
-This number is quite big. Using \texttt{Ocaml}'s \texttt{float} arithmetic,
-we can under-approximate it by $2^{3.8\times10^{30}}\times 3.8\times{10^{30}}$.
-
-\begin{Coqsrc}
-# let exp2 x = 2.0 ** x;;
-
-val exp2 : float -> float = <fun>
-#   exp2 95.0 *. 97.0 -. 1.0;;
-- : float = 3.84256588194182037e+30
-# let n = exp2 95.0 ;;
-# let p = n *. 97.0 -. 1.0;;
-val p : float = 3.84256588194182037e+30
-
-Estimation :
-2 ** (3.84 e+30) * 3.84 e+30.
-\end{Coqsrc}
-
-\subsection{Using \texttt{Equations}}
-\label{sect:L-equations}
-
-Note that we did not define any function $L_\alpha$ \emph{for any $\alpha<\epsilon_0$} yet.
-We have got no more than a collection of proved realizations of $\texttt{L\_spec}\;\alpha$ for a few values of $\alpha$.
-
-\index{coq}{Plug-ins!Equations}
-
-Using the \texttt{coq-equations} plug-in by 
-M. Sozeau~\cite{sozeau:hal-01671777}, we will now define a function \texttt{L\_} which maps any ordinal  $\alpha<\epsilon_0$ to a proven realization of 
-$\texttt{L\_spec}\;\alpha$.   
-To this end, we represent ordinals as inhabitants of the type 
-\texttt{E0} of well-formed ordinal terms (see Sect~\vref{sect:E0-def}). So, we define a total function \texttt{L\_} of type
-\texttt{E0 -> nat -> nat}, by transfinite recursion, considering the usual three cases : $\alpha=0$, $\alpha$ is a successor, $\alpha$ is a limit ordinal.
-
-\subsubsection{Definition}
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#L_}{L\_alpha}).}
-
-\label{Functions:L-alpha}
- \index{hydras}{Library Epsilon0!Functions!L\_@L\_ (final step of a minimal path}
-
-\input{movies/snippets/L_alpha/LDef}
-
-This definition results in a bunch of automatically generated lemmas. For instance:
-
-\input{movies/snippets/L_alpha/AboutLEquation1}
-
-In most cases, it may be useful to write human-readable  paraphrases of these statements.
-
-\inputsnippets{L_alpha/Paraphrasesa}
-\inputsnippets{L_alpha/Paraphrasesb}
-\inputsnippets{L_alpha/Paraphrasesc}
-
-Using these three lemmas as rewrite rules, we can prove more properties of the functions \texttt{L\_$\alpha$}.
-
-\input{movies/snippets/L_alpha/LFiniteOmega}
-
-By  well-founded induction on $\alpha$, we prove the following properties:
-
-\input{movies/snippets/L_alpha/LGeS}
-
-\input{movies/snippets/L_alpha/LCorrect}
-
-\label{sect:L-correct-proof}
-
-Please note that the proof of \texttt{L\_correct} applies the lemmas proven in Sections~\ref{sect:L-spec-zero}, ~\ref{sect:L-spec-succ} and ~\ref{sect:L-spec-lim}.
-Our previous study of \texttt{L\_spec} allowed us to pave the way for the definition by \texttt{Equations} and the correctness proof.
-
-\paragraph*{\gaiasign}
-The module
-\href{../theories/html/gaia_hydras.GL_alpha.html}{gaia\_hydras.GL\_alpha} contains an adaptation of
-\href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0/L\_alpha}.
-
-\subsubsection{Back to hydra battles}
-\label{def:L-alpha}
-
-Lemma \texttt{battle\_length\_std } of
-Module~\href{../theories/html/hydras.Hydra.Battle_length.html}{Hydra.Battle\_length} relates the length of standard battles with the functions $L_\alpha$.
-
-\input{movies/snippets/Battle_length/battleLengthStd}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Instead of considering standard paths and battles, consider battles where the replication factor is a constant $k$.
-Please use \texttt{Equations} in order to define the function that computes the length of the $k$-path which leads  from $\alpha$ to $0$.
-Prove a few  exact formulas and minoration lemmas.
-\end{exercise}
-
-\section{A variant of the Hardy hierarchy}
-
-\label{sect:hardy}
-
-In order to give a feeling on  the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the  \emph{Hardy hierarchy} of fast growing functions,
-presented for instance in~\cite{Promel2013}. 
-\index{maths}{Rapidly growing functions!Hardy Hierarchy}
-
-\begin{remark}
-  Indeed, the functions presented in this section are a \emph{variant} of the Hardy hierarchy of functions.
-In the future versions of this development, we will correct the references to the literature.
-For the time being, we call our functions $H'_\alpha$ in order to underline the difference from ``classic'' Hardy functions.
-\end{remark}
-
-For each ordinal $\alpha$ below $\epsilon_0$, $H'_\alpha$ is 
-a total arithmetic function, defined  by  transfinite recursion on $\alpha$, according to three cases:
-
-\index{maths}{Transfinite induction}
-
-\begin{itemize}
-\item If $\alpha=0$, then $H'_\alpha (k)= k$ for any natural number $k$.
-\item If $\alpha=\textrm{succ}(\beta)$, then 
-$H'_\alpha(k)=H'_\beta(k+1)$ for any $k \in \mathbb{N}$
-\item If $\alpha$ is a limit ordinal, then 
-$H'_\alpha(k) = H'_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
-\end{itemize}
-
-\begin{remark}
- The ``classic'' definition of the Hardy hierarchy differs in the third equation.
-
-\begin{itemize}
-\item If $\alpha=0$, then $H_\alpha (k)= k$ for any natural number $k$.
-\item If $\alpha=\textrm{succ}(\beta)$, then 
-$H_\alpha(k)=H_\beta(k+1)$ for any $k \in \mathbb{N}$
-\item If $\alpha$ is a limit ordinal, then 
-$H_\alpha(k) = H_{(\canonseq{\alpha}{k})}(k)$ for any $k\in \mathbb{N}$.
-\end{itemize}
-
-\end{remark}
-
-\subsection{Definition in \texttt{Coq}}
-
-We define a function \texttt{H'\_} of type \texttt{E0 -> nat -> nat} by transfinite induction over the type \texttt{E0} of the well formed ordinals below $\epsilon_0$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}}
-
-\index{hydras}{Library Epsilon0!Functions!H\_@H\_ (Hardy hierarchy (variant))}
-\index{coq}{Plug-ins!Equations}
-\label{Functions:Hprime-alpha}
-
-\input{movies/snippets/Hprime/HprimeDef}
-\input{movies/snippets/Hprime/HprimeDefb}
-
-\input{movies/snippets/Hprime/paraphrasesa}
-\input{movies/snippets/Hprime/paraphrasesb}
-\input{movies/snippets/Hprime/paraphrasesc}
-\input{movies/snippets/Hprime/paraphrasesd}
-
-\subsection{First  steps of the H' hierarchy}
-Using rewrite rules from \texttt{H'\_eq1} to \texttt{H'\_succ\_eqn}, we can explore the functions $H'_\alpha$ for  small values of $\alpha$.
-
-\paragraph*{\gaiasign}
-The module
-\href{../theories/html/gaia_hydras.GHprime.html}{gaia\_hydras.GHprime} contains an adaptation of
-\href{../theories/html/hydras.Epsilon0.Hprime.html}{Epsilon0.Hprime}.
-
-\subsubsection{Finite ordinals}
-
-By induction on $i$, we prove a simple expression of \texttt{H'\_ (Fin i)}, where 
-\texttt{Fin $i$}  is the $i$-th finite ordinal.
-
-\input{movies/snippets/Hprime/HprimeFin}
-
-\subsubsection{Multiples of \texorpdfstring{$\omega$}{omega}}
-
-Since the canonical sequence of $\omega$ is composed of finite ordinals, 
-it is easy to get the formula associated with $H'_\omega$.
-
-\input{movies/snippets/Hprime/HprimeOmega}
-
-Before going further, we prove a useful rewriting lemma:
-
-\input{movies/snippets/Hprime/HprimePlusFin}
-
-Then, we get easily formulas for $H'_{\omega+i}$, and $H'_{\omega\times i}$ for any natural number $i$.
-
-\input{movies/snippets/Hprime/HprimeExamplesa}
-\input{movies/snippets/Hprime/HprimeExamplesb}
-\input{movies/snippets/Hprime/HprimeExamplesc}
-\input{movies/snippets/Hprime/HprimeExamplesd}
-
-Since $\omega^2$ is a limit ordinal, we prove the following equality: 
-\[H'_{\omega^2} (k) = 2 ^ {k+1} \times (k+1) - 1\]
-
-\input{movies/snippets/Hprime/HprimeOmegaSqr}
-
-\subsubsection{New limits}
-
-Our next step would be to prove an exact formula for $H'_{\omega^\omega}(k)$.
-Since the canonical sequence of $\omega^\omega$ is composed of all the
-$\omega^i$, we first need to express $H'_{\omega^i}$ for any natural number $i$.
-
-Let $i$ and $k$ be two natural numbers. 
-The ordinal $\canonseq{\omega^(i+1)}{k}$ is the product
-$\omega^i \times k$, so we need also to consider ordinals of this form.
-
-\begin{enumerate}
-\item First,  we express $H'_{\omega^\alpha \times (i+2)}$ in terms of
-$H'_{\omega^\alpha \times (i+1)}$.
-
-\input{movies/snippets/Hprime/HprimeOmegaTerm1}
-
-\item
-Then, we prove by induction on $i$ that $H'_{\omega^\alpha \times (i+1)}$ is just the
-$(i+1)$-th iterate of $H'_{\omega^\alpha}$.
-
-\input{movies/snippets/Hprime/HprimeOmegaTerm}
-
-\item In particular, we derive a formula for $H'_{\omega^{i+1}}$.
-
-\input{movies/snippets/Hprime/HprimeSuccFun}
-\input{movies/snippets/Hprime/HprimePhi0SI}
-
-\item We get now a  formula for $H'_{\omega^3}$:
-
-\input{movies/snippets/Hprime/HprimeOmegaCube}
-\end{enumerate}
-
-\subsubsection{A numerical example}
-\label{sect:bignum}
-
-It is hard to capture the complexity of this function by looking only at this
-``exact'' formula. 
-Let us consider a simple example: the number $H'_{\omega^3}(3)$.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3a}
-
-Thus, the number $H_{\omega^3}(3)$ can be written as four nested applications of $f$.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3b}
-
-In a more classical writing, this number is displayed as follows:
-
-{\Large
-$$
-H'_{\omega^3}(3) =  2 ^ {(2 ^ {N + 1} \, (N+1) )}   \,  (2 ^ {N+1} \, ( N +1) ) - 1
-$$
-}
-
-In order to make this statement more readable, we introduce a local definition.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3c}
-
-This number looks quite big; let us compute an approximation with \texttt{Ocaml}:
-
-\begin{Coqsrc}
-# (2.0 ** 64.0 *. 64.0 -. 1.0);; 
-\end{Coqsrc}
-
-\begin{Coqanswer}
-- : float = 1.1805916207174113e+21
-\end{Coqanswer}
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3d}
-
-We leave as an exercise to determine the best approximation as possible of
- the size of this number (for instance its number of digits).  For instance, if
-we do not take into account the multiplications in the formula above,
-we obtain that, in base $2$, the number $H'_{\omega^3}(3)$ has at least
-$2^{10^{21}}$  digits. But it is still an under-approximation !
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3e}
-
-\subsubsection{A formula for \texorpdfstring{$H'_{\omega^\omega}$}{\texttt{H'(phi0(omega))}}}
-
-Now, we can get at last an exact formula for $H'_{\omega^\omega}$.
-
-\input{movies/snippets/Hprime/HprimePhi0Omega}
-
-Using extensionality of the functional \texttt{iterate}, we also get a closed formula.
-
-\input{movies/snippets/Hprime/HprimePhi0OmegaClosed}
-
-Note that this formula contains two occurrences of the functional \texttt{iterate}, the second  one is in fact a second-order iteration (on type \texttt{nat -> nat)}
-and the first one  first-order (on type \texttt{nat}).
-
-\subsection{Abstract properties of  
-\texorpdfstring{$H'_\alpha$}{H'}}
-~\label{sect:H-alpha-prop}
-
-Since pure computation seems to be useless for dealing with expressions of the form $H'_\alpha(k)$, even for small values of $\alpha$ and $k$, we need to prove theorems for comparing $H'_\alpha(k)$ and $H'_\beta(l)$, in terms of comparison
-between $\alpha$ and $\beta$ on the one hand, $k$ and $l$ on the other hand.
-
-But beware of fake theorems! For instance, one could believe that $H'$ is monotonous in its first argument. The following proof shows this is false.
-
-\input{movies/snippets/Hprime/HprimeNonMono1}
-
-On the contrary, the functions of the $H'$ hierarchy have the following five properties~\cite{KS81}: for any $\alpha < \epsilon_0$,
-\begin{itemize}
-\item the function $H'_\alpha$ is strictly monotonous :
-      For all $n,p \in\mathbb{N}, n < p \Rightarrow H'_\alpha(n)< H'_\alpha(p)$.
-\item If $\alpha \not= 0$, then for every $n$, $n<H'_\alpha(n)$.
-\item The function $H'_\alpha$ is pointwise less or equal than $H'_{\alpha+1}$
-
-\item For any $n\geq 1$, $H'_\alpha(n)<H'_{\alpha+1}(n)$.
-\emph{We say that $H'_{\alpha+1}$ dominates $H'_\alpha$ from $1$}.
-\item For any $n$ and $\beta$, if $\alpha \xrightarrow[n]{} \beta$, then
-$H'_\beta(n)\leq H'_\alpha(n)$.
-\end{itemize}
-
-\index{maths}{Abstract properties of arithmetic functions}
-\index{hydras}{Abstract properties of arithmetic functions}
-
-\index{maths}{Transfinite induction}
-
-In \coq{}, we follow the  proof of~\cite{KS81}. This proof is mainly a single  proof by transfinite induction on $\alpha$ of the conjunction of the five properties.
-For each $\alpha$, the three cases : $\alpha=0$, $\alpha$ is a limit, and 
-$\alpha$ is a successor are considered. Inside each case, the five sub-properties are proved sequentially, using the abstract properties defined in Sect~\vref{sect:abstract-arith-prop}
-
-\input{movies/snippets/Hprime/PAP}
-
-Using a few lemmas \emph{à la} Ketonen-Solovay, we prove that
-if $\alpha<\beta$, then $H'_\beta$ eventually dominates
-$H'_\alpha$.
-We let the reader look at the proof (Section \texttt{Proof\_of\_H'\_mono\_l} of \href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}).
-
-\noindent
-\input{movies/snippets/Hprime/HprimeDom}
-
-%ICI
-\subsection{Comparison between \texttt{L\_} and \texttt{H'\_} }
-
-By well-founded induction on $\alpha$, we prove that
- $H'_\alpha(i) \leq L_\alpha(i+1)$ for any $i$.
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#H'_L_}{Epsilon0.L\_alpha}}
-
-\input{movies/snippets/L_alpha/HprimeL}
-
-\subsubsection{Back to hydras}
-
-The following theorem relates the length of (standard) battles with the the $H'$ family of fast growing functions.
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems}}
-
-\input{movies/snippets/Hydra_Theorems/battleLengthStdHardy}
-
-Since $H'_{\omega^3}(3)= H'_{\omega^3+3}(0)$, the big number shown in Section~\vref{sect:bignum} is less or equal than 
-the number of steps of the battle with the initial hydra of
-Figure~\ref{fig:start}, and  $0$ as initial replication factor.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[very thick, scale=0.3]
-  \node (h1) at (4,0){$\bullet$};
-  \node (h2) at (2,2){$\bullet$};
-  \node (h3) at (0,4){$\bullet$};
-  \node (h4) at (1,4){$\bullet$};
-  \node (h5) at (2,4){$\bullet$};
-  \node (h6) at (4.5,2){$\bullet$};
-  \node (h7) at (6.5,2){$\bullet$};
-  \node (h8) at (8.5,2){$\bullet$};
-    \draw (h1) -- (h2) ;
-    \draw (h2) -- (h3) ;
-    \draw (h2) -- (h4) ;
-    \draw (h2) -- (h5) ;
-    \draw (h1) -- (h6) ;
-    \draw (h1) -- (h7) ;
-    \draw (h1) -- (h8);
- \end{tikzpicture}
-
-  \caption{The hydra corresponding to the ordinal $\omega^3+3$}
-  \label{fig:start}
-\end{figure}
-
-\section{A variant of the Wainer hierarchy (functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
-\label{sect:wainer}
-
-\index{maths}{Rapidly growing functions!Wainer Hierarchy}
-
-Ketonen and Solovay introduce in~\cite{KS81} a ``trivial'' variant of the Wainer hierarchy~\cite{BW85, Wainer1970} of fast growing functions, indexed by ordinals below $\epsilon_0$.
-The functions $F_\alpha$ are defined by the following equations.
-
-\label{F_equations}
-\begin{itemize}
-\item $F_0(i)=i+1$
-\item $F_{\beta+1}(i)= (F_\beta)^{(i+1)}(i)$, where $f^{(i)}$ is the $i$-th iterate of $f$.
-\item $F_\alpha(i) = F_{\canonseq{\alpha}{i}} (i)$ if $\alpha$ is a limit ordinal.
-\end{itemize}
-
-\begin{remark}
-The difference with the ``classic'' Wainer hierarchy 
-$f_\alpha\;(\alpha<\epsilon_0)$ lies in the second equation:
-$f_{\beta+1}(i) = (f_\beta)^{(i)}(i)$ and not
-$f_{\beta+1}(i) = (f_\beta)^{(i+1)}(i)$.
-
-A module about 
-the classic Wainer hierarchy is in preparation.
-
-Note also that \cite{KS81} defines also $F_{\epsilon_0}$ (by the third equation).
-Since $\epsilon_0$ is not representable in type \texttt{E0}, our implementation in \coq{} does not take $F_{\epsilon_0}$ into account.
-
-\end{remark}
-
-A first attempt is to write a definition of $F_\alpha$ by equations, in the same way as for $H_\alpha$ (the functional \texttt{iterate} has already been used in Sect.\vref{Functions:iterate}).
-
-\index{hydras}{Library Prelude!iterate}
-
-\input{movies/snippets/Iterates/iterateDef}
-
-\emph{From\href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}~}.
-
-\index{coq}{Plug-ins!Equations}
-
- \input{movies/snippets/F_alpha/FailDemo}
-
-We presume that this error comes from the recursive call of \texttt{F\_} inside
-an application of \texttt{iterate}. The workaround we propose is to define first 
-the iteration of \texttt{F\_}  as an helper $F^*$, then to define the function $F$ as a ``iterating $F^*$ once''.
-
-\texttt{Equations} accepts the following definition, relying on  lexicographic ordering on pairs $(\alpha,n)$.
-
-\label{sect:F-equations}
-
-\index{coq}{Plug-ins!Equations}
-\label{Functions:F-alpha}
-\index{maths}{Rapidly growing functions}
-\index{hydras}{Library Epsilon0!Functions!F\_@F\_ (Wainer hierarchy)}
-
-\input{movies/snippets/F_alpha/goodDefa}
-\input{movies/snippets/F_alpha/goodDefb}
-
-It is quite easy to prove that our functional \texttt{F\_} satisfies the equations on page~\pageref{sect:F-equations}.
-\index{hydras}{Library Prelude!iterate}
-
-\input{movies/snippets/F_alpha/FEquations}
-
-As for the Hardy functions, we can use these equalities as rewrite rules for
-``computing'' some values of $F_\alpha(i)$, for small values of $\alpha$.
-
-\input{movies/snippets/F_alpha/FirstValues}
-
-Like in Sect~\ref{sect:H-alpha-prop}, we prove by induction the following properties (see~\cite{KS81}).
-
-\input{movies/snippets/F_alpha/FalphaThms}
-
-As a corollary, we prove that, if $\beta<\alpha$, then
-$F_\alpha$ dominates $F_\beta$ (p. 284 of~\cite{KS81}).
-
-\input{movies/snippets/F_alpha/FDomContext}
-\input{movies/snippets/F_alpha/FDom}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove the following property:
-
-\begin{Coqsrc}
-Lemma LF3 : dominates_from  2 (F_ 3) (fun  n => iterate exp2 n n).
-\end{Coqsrc}
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
-\end{exercise}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that, for any $\alpha\geq 3$ and $n\geq 2$,
-$F_\alpha(1+n)\geq 2^{F_\alpha(n)}$.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
-\end{exercise}
-
-%ici
-\index{hydras}{Exercises}
-
-\begin{exercise}
-It is tempting to prove a simple property of monotony 
-of the function \texttt{F\_}.
-
-\begin{quote}
-   Let $\alpha\leq\beta<\epsilon_0$. For any $n\geq 2$,
-$F_\alpha(n)\leq F_\beta(n)$. 
-\end{quote}
-Prove or disprove this statement.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/is_F_monotonous.v}{exercises/ordinals/is\_F\_monotonous.v}.}
-\end{exercise}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-
-Prove that for any $n\geq 2$, $\textrm{Ack}\,\,n\,n\leq  F_\omega(n)$, where \textrm{Ack} is the Ackermann function.
-Next, prove that $F_\alpha$ is not primitive recursive, for any $\alpha\geq\omega$  (please see Sect.~\vref{sect:ack-not-PR}).
-On the other hand, please show that for any natural number $n$, the function $F_n$ is primitive recursive.
-Thus $F_\alpha$ is primitive recursive if and only if $\alpha$ is finite.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_omega.v}{exercises/ordinals/F\_Omega.v}.
-Properties of the Ackermann function are studied in
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/Ack.v}{theories/ordinals/MoreAck/Ack.v} and
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/AckNotPR.v}{theories/ordinals/MoreAck/AckNotPR.v}
-.}
-\end{exercise}
-
-\paragraph*{\gaiasign} Library~\href{../theories/html/gaia_hydras.GF_alpha.html}{gaia\_hydras.GF\_alpha}  defines the functions $F_\alpha$ for \gaia's world.
-\index{gaiabridge}{Rapidly growing functions}
-
-\section{More about rapidly growing functions}
-
-In Sect.~\vref{sect:battle-length-notPR}, we prove that the length of hydra-battles (for a given hydra, according to the initial replication factor) is not primitive recursive in general. 
-This proof uses properties of the Ackermann function, and the $H'_\alpha$, $F_\alpha$, $L_\alpha$ families of functions.
-
-% ---------------------------------------------------
-\input{gaia-chapter}
-
-%----------------------------------------------------------------------
-\chapter[Countable ordinals (after Sch\"{u}tte)]{Kurt Schütte's axiomatic definition of countable ordinals}
-
-\label{chap:schutte} 
-%ON
-
-In the present chapter, we  compare our implementation of the segment $[0,\epsilon_0)$ with a mathematical text in order to ``validate'' our constructions.
-Our reference here is the axiomatic definition of the set of countable ordinals,
-in chapter V of Kurt Schütte's book `` Proof Theory ''~\cite{schutte}.
-
-\begin{remark}
-\emph{In all this chapter, the word ``ordinal'' will be considered as a synonymous of
-``countable ordinal''}  
-\end{remark}
-
-Schütte's definition of countable ordinals relies on the following three axioms:
-
-There  exists a strictly ordered set , such that
-\begin{enumerate}
-\item  $(\mathbb{O},<)$ is well-ordered
-\item Every bounded subset of $\mathbb{O}$  is countable
-\item Every countable subset of $\mathbb{O}$  is bounded.
-\end{enumerate}
-
-Starting with these three axioms, Schütte re-defines the vocabulary about ordinal numbers: the null ordinal $0$, limits and successors, the addition of ordinals, the infinite ordinals $\omega$, $\epsilon_0$, $\Gamma_0$, etc.
-
-This chapter describes an adaptation to \coq{} of Schütte's axiomatization. 
- Unlike the rest of our libraries, our library
-\href{../theories/html/hydras.Schutte.Schutte.html}{hydras.Schutte}
-is not constructive, and relies on several axioms.
-
-\begin{itemize}
-\item First, please keep in mind  that the set of countable ordinals is not countable.
-Thus, we cannot hope to represent all countable ordinals as finite terms of an inductive type, which was possible with  the set of ordinals strictly less than $\epsilon_0$ (resp.
-$\Gamma_0$)
-\item We tried to be as close as possible to K.
-Schütte's text, which uses ``classical'' mathematics : excluded middle, Hilbert's $\epsilon$ (choice) and Russel's $\iota$ (definite description) operators.
-Both operators allow us to write definitions close to the natural mathematical language, such as ``$\textrm{succ}$ is \emph{the} least ordinal strictly greater than $\alpha$''
-\item Please note that only the library \href{../theories/html/hydras.Schutte.Schutte.html}{Schutte/*.v} is ``contaminated'' by axioms, and that the rest of our libraries remain constructive.
-\end{itemize}
-
-\section{Declarations and axioms}
-
-Let us declare a type 
-\texttt{Ord} for representing countable ordinals, and a binary relation
- \texttt{lt}. Note that, in our development, \texttt{Ord} is a type, while the \emph{set} of countable ordinals (called $\mathbb{O}$ by Schütte) 
-is the full set over the type \texttt{Ord}.
-
-\label{types:Ord}
-
-We use Florian Hatat's library on countable sets, written as he was a student of  \emph{\'Ecole Normale Supérieure de Lyon}.
-A set $A$ is countable if there is an injective function from $A$ to $\mathbb{N}$ (see
-Library \href{../theories/html/hydras.Schutte.Countable.html}%
-{\texttt{Schutte.Countable}}).
-
-\vspace{6pt}
-
-\emph{From Module\href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{Schutte.Schutte\_basics}}}
-
-\index{hydras}{Library Schutte!Types!Ord}
-
-\input{movies/snippets/Schutte_basics/OrdDecl}
-
-Schütte's first axiom tells that \texttt{lt} is a well order on the set 
-\texttt{ordinal} (The  class \texttt{WO} is defined in
-Module~\href{../theories/html/hydras.Schutte.Well_Orders.html}{Schutte.Well\_Orders.v}).
-
-\index{hydras}{Library Schutte!Type classes!WO@ WO (well order)}
-
-\label{types:WO}
-
-\input{movies/snippets/Well_Orders/Mdecl}
-
-\input{movies/snippets/Well_Orders/WODef}
-
-\input{movies/snippets/Schutte_basics/AX1}
-
-The second and third axioms say that a subset $X$ of $\mathbb{O}$ is
-(strictly) bounded if and only if it is countable.
-
-\input{movies/snippets/Schutte_basics/AX23}
-
-\texttt{AX2} and \texttt{AX3} could have been replaced by a single axiom (using the \texttt{iff} connector), but we decided to respect as most as possible the structure of Schütte's definitions.
-
-\section{Additional  axioms}
-
-The adaptation of Schütte's mathematical discourse to \coq{} led us to
-import a few axioms from the standard library. We encourage the reader to consult \coq{}'s FAQ about the safe use of axioms
- \url{https://github.com/coq/coq/wiki/The-Logic-of-Coq#axioms}.
-
-\subsubsection{Classical logic}
-
-In order to work with classical logic, we import the module
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Classical.html}{Coq.Logic.Classical}  of \coq{}'s standard library, specifically the following axiom:
-
-\begin{Coqsrc}
- Axiom classic : forall P:Prop, P \/ ~P.
-\end{Coqsrc}
-
-\subsubsection{Description operators}
-
-In order to respect Schütte's style, we imported also the library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Epsilon.html}{\texttt{Coq.Logic.Epsilon}}.  The rest of this section presents a few examples of
-how Hilbert's choice operator and Church's definite description allow us
- to write understandable definitions (close to the mathematical natural language).
-
-\subsubsection{The definition of zero}
-
-According to the  definition of a well order, every non-empty subset of \texttt{Ord} has a least element. Furthermore, this least element is unique. We would like to call this element  \texttt{zero}.
-
-\input{movies/snippets/Schutte_basics/iotaDemoa}
-\input{movies/snippets/Schutte_basics/iotaDemob}
-
-Indeed, the basic logic of  \coq{} does not allow us to eliminate a proof of a proposition 
-$\exists!\,x:A,\,P(x)$ for building a term whose type lies in the sort \texttt{Type}. 
-The reasons for this impossibility are explained in many documents~\cite{BC04, chlipalacpdt2011, Coq}.
-
-Let us import the library \texttt{Coq.Logic.Epsilon}, which contains the following axiom and lemmas.
-
-\input{movies/snippets/MoreEpsilonIota/EpsilonStatement}
-
-Hilbert's $\epsilon$ \emph{operator} is derived from this  axiom.
-
-\input{movies/snippets/MoreEpsilonIota/EpsilonDef}
-
-If we consider the \emph{unique existential} quantifier $\exists!$, we obtain
-Church's \emph{definite description operator}.
-
-\input{movies/snippets/MoreEpsilonIota/iotaDef}
-
-Indeed, the operators \texttt{epsilon} and \texttt{iota} allowed us to make our definitions 
-quite close to Schütte's text. Our libraries \href{../theories/html/hydras.Schutte.MoreEpsilonIota.html}%
-{\texttt{Schutte.MoreEpsilonIota}}
-and
-\href{../theories/html/hydras.Schutte.PartialFun.html}%
-{\texttt{Schutte.PartialFun}} are extensions of \texttt{Coq.logic.Epsilon} for making easier 
-such definitions. See also an article in french~\cite{PCiota}.
-
-\input{movies/snippets/MoreEpsilonIota/Defs}
-
-In order to use these tools,  we have to tell \coq{}  that the declared type \texttt{Ord} is not empty:
-
-\input{movies/snippets/Schutte_basics/inhOrd}
-
-We are now able to define \texttt{zero} as the least ordinal. For this purpose,
-we define a function returning the least element of any [non-empty]  subset.
-
-\emph{From Module\href{../theories/html/hydras.Schutte.Well_Orders.html}%
-  {\texttt{Schutte.Well\_Orders}}}
-
-\input{movies/snippets/Well_Orders/MDecl}
-\input{movies/snippets/Well_Orders/theLeast}
-
-\vspace{4pt}
-
-From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{~Schutte.Schutte\_basics}}
-
-\label{Constants:zero:Ord}
-\index{hydras}{Library Schutte!Constants!zero}
-
-\input{movies/snippets/Schutte_basics/zeroDef}
-
-We want to prove now that zero is less than or equal to any ordinal number.
-
-\input{movies/snippets/Schutte_basics/zeroLe}
-
-\subsubsection{Remarks on \texttt{epsilon} and \texttt{iota}}
-
- What would happen in case of a misuse of \texttt{epsilon} or \texttt{iota} ?
-For instance, one could give a unsatisfiable specification to \texttt{epsilon} or 
-a specification for \texttt{iota} that admits several realizations.
-
-Let us consider an example:
-
-\input{movies/snippets/Schutte_basics/BadBottoma}
-
-Since we won't be able to prove the proposition
-\linebreak \Verb|(exists! a: Ord, least_member (Empty_set Ord) a)|, the only properties we would be able to prove about \texttt{bottom} would be \emph{trivial} properties, 
-\emph{i.e.}, satisfied by \emph{any} element of type \texttt{Ord}, like for instance
-\texttt{bottom = bottom}, or \texttt{zero <= bottom}.
-
-\input{movies/snippets/Schutte_basics/trivialProps}
-
-On the other hand, the following attempt fails, because of the unprovable first subgoal (please notice that the second subgoal is easy to solve !).
-
-\input{movies/snippets/Schutte_basics/Failure}
-
-In short, using \texttt{epsilon} and \texttt{iota} in our implementation of countable ordinals after Schütte has two main advantages.
-
-\begin{itemize}
-\item It allows us to give a \emph{name} (using \texttt{Definition}) two witnesses 
-of existential quantifiers (let us recall that, in classical logic, one may consider non-constructive proofs of existential statements)
-\item By separating definitions from proofs of [unique] existence, one may make definitions  more concise and readable. Look for instance at 
-the definitions of  \texttt{zero}, \texttt{succ}, \texttt{plus}, etc. in the rest of this chapter.
-\end{itemize}
-%%%% ICI ICI
-
-\section{The  successor function}
-
-The definition of the function \texttt{succ:Ord -> Ord} is very concise. The successor of any ordinal $\alpha$ is the smallest ordinal strictly greater than $\alpha$.
-
-\label{Functions:succ-sch}
-\index{hydras}{Library Schutte!Functions!succ}
-
-\input{movies/snippets/Schutte_basics/succDef}
-
-Using \texttt{succ}, we define the following predicates.
-
-\input{movies/snippets/Schutte_basics/isSuccIsLimit}
-
-% \begin{remark}
-% Please look at remark~\vref{warning:coercions}.
-
-How do we prove properties of the successor function?
-First, we make its specification explicit.
-
-\input{movies/snippets/Schutte_basics/succSpec}
-
-Then, we prove that our function \texttt{succ} meets this specification.
-
-\input{movies/snippets/Schutte_basics/succOka}
-
-We have now to prove that the set of all ordinals strictly greater than $\alpha$ has a unique least element.
-But the singleton set $\{\alpha\}$ is countable, hence  bounded (by the axiom \texttt{AX3}).
-Hence; the set $\{\beta\in\mathbb{O}|\alpha < \beta\}$ is not empty
-and therefore has a unique least element.
-
-The rest of the \coq{} proof script is quite short.
-
-\input{movies/snippets/Schutte_basics/succOkb}
-
-We can ``uncap'' the description operator for proving properties of the
-\texttt{succ} function.
-
-\input{movies/snippets/Schutte_basics/succProps}
-
-\section{Finite ordinals}
-
-Using \texttt{succ}, it is now easy to define recursively all the finite ordinals.
-
-\label{sect:notation-F-sch}
-
-\input{movies/snippets/Schutte_basics/finiteDef}
-
-\section{The definition of \texttt{omega}}
-In order to define $\omega$, the first infinite ordinal, we use an operator which
-``returns'' the least upper bound (if it exists) of a subset $X\subseteq \mathbb{O}$.
-For that purpose, we first use a predicate:
-(\texttt{is\_lub $D$ \textit{lt} $X$ $a$}) if $a$ belongs to $D$ and is the least 
-upper bound  of $X$ (with respect to \textit{lt}).
-
-\input{movies/snippets/Lub/isLubDef}
-
-\input{movies/snippets/Schutte_basics/supDef}
-
-Then, we define the function \texttt{omega\_limit} which returns the least upper bound 
-of the  (denumerable) range of any sequence \texttt{s: nat -> Ord}. 
-By \texttt{AX3} this range is bounded, hence the set of its upper bounds is not empty and has a least element.
-Then we define \texttt{omega} as the limit of the sequence of finite ordinals.
-
-\input{movies/snippets/Schutte_basics/omegaDef}
-
-\label{sect:notation-omega}
-
-Among the numerous properties of the ordinal $\omega$, let us quote the following ones
-(proved in Module 
-\href{../theories/html/hydras.Schutte.Schutte_basics.html\#finite_lt_omega}{\texttt{Schutte.Schutte\_basics}})
-
-\input{movies/snippets/Schutte_basics/omegaPropsa}
-\input{movies/snippets/Schutte_basics/omegaPropsb}
-\input{movies/snippets/Schutte_basics/omegaPropsc}
-\input{movies/snippets/Schutte_basics/omegaPropsd}
-
-\subsection{Ordering functions and ordinal addition}
-
-After having defined the finite ordinals and the infinite ordinal $\omega$, we  define the sum $\alpha+\beta$ of two countable ordinals.
-Schütte's definition looks like the following one:
-
-\begin{quote}
-``$\alpha+\beta$ is the $\beta$-th ordinal greater than or equal to $\alpha$''
-\end{quote}
-
-The purpose of this section is to give a meaning to the construction
-``the $\alpha$-th element of $X$''  where $X$ is any non-empty subset of $\mathbb{O}$.
-We follow Schütte's approach, by defining the notion of \emph{ordering functions},
-a way to associate a unique ordinal to each element of $X$.
-Complete definitions and proofs can be found in Module
- \href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
-{\texttt{Schutte.Ordering\_Functions}} ).
-
-\subsection{Definitions}
-
-A \emph{segment} is a set $A$ of ordinals such that, whenever  $\alpha\in A$ and
-$\beta<\alpha$, then $\beta\in A$; a segment is  \emph{proper} if it strictly included in $\mathbb{O}$.
-
-\input{movies/snippets/Ordering_Functions/segmentDef}
-
-Let  $A$ be a segment, and $B$ a subset of $\mathbb{O}$ : an \emph{ordering function for $A$ and  $B$} is a strictly increasing bijection from $A$ to $B$.
-The set $B$ is said to be an \emph{ordering segment} of $A$.
-Our definition in \coq{} is a direct translation of the mathematical text of~\cite{schutte}.
-
-\index{maths}{Ordinal numbers!Ordering functions}
-\index{hydras}{Library Schutte!Predicates!ordering function@ordering\_function}
-
-\input{movies/snippets/Ordering_Functions/orderingFunctionDef}
-
-We are now able to associate with any subset $B$ of $\mathbb{O}$ its ordering segment and ordering function.
-
-\input{movies/snippets/Ordering_Functions/ordDef}
-
-Thus (\texttt{ord $B \;\alpha$}) is the $\alpha$-th element of $B$.
-Please note that the last definition uses the epsilon-based operator \texttt{some} and
-not \texttt{the}. This is due to the fact that we cannot prove the unicity (w.r.t. Leibniz' equality) of the ordering function of a given set. 
-By contrast, we admit the axiom  \texttt{Extensionality\_Ensembles}, from the library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Sets.Ensembles.html}{Coq.Sets.Ensembles}, so we use the operator \texttt{the} in the definition of
-\texttt{the\_ordering\_segment}.
-
-One of the main theorems of
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html\#ordering_function_ex}%
-{\texttt{Ordering\_Functions}} 
-associates a unique segment and a unique (up to extensionality) ordering function to every subset $B$ of $\mathbb{O}$.
-
-\input{movies/snippets/Ordering_Functions/orderingFunctionEx}
-
-Thus,  our function \texttt{ord}  which enumerates the elements of $B$ is defined in a non-ambiguous way.
-Let us quote the following theorems (see Library \linebreak
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
-{\texttt{Schutte.Ordering\_Functions}} for more details).
-
-\input{movies/snippets/Ordering_Functions/orderingLe}
-\input{movies/snippets/Ordering_Functions/Th1352}
-
-\subsection{Ordinal addition}
-
-We are now ready to define and study addition on the type \texttt{Ord}.
-The following definitions and proofs can be consulted in Module
-\href{../theories/html/hydras.Schutte.Addition.html}%
-{\texttt{Schutte.Addition.v}}.
-
-\index{hydras}{Library Schutte!Functions!plus}
-
-\input{movies/snippets/Addition/additionDef}
-
-In other words,  $\alpha + \beta$ is the  $\beta$-th ordinal greater than or equal to $\alpha$. 
-Thanks to generic properties of ordering functions, we can show the following 
-properties of addition on $\mathbb{O}$. First, we prove a useful lemma:
-
-\input{movies/snippets/Addition/plusElim}
-
-As a use-case, let us prove that $0$ is a right neutral element of $+$.
-
-\input{movies/snippets/Addition/alphaPlusZero}
-
-The following lemmas are proved the same way.
-
-\input{movies/snippets/Addition/bunchOfLemmas}
-
-The following lemmas are not direct applications of \texttt{plus\_elim}.
-
-\input{movies/snippets/Addition/plusAssoc}
-\input{movies/snippets/Addition/finitePlusInfinite}
-
-It is interesting to compare the proof of these lemmas with the
-computational proofs of the corresponding statements in Module
-\href{../theories/html/hydras.Epsilon0.T1.html}%
-{\texttt{Epsilon0.T1}}. 
-For instance, the proof of the lemma 
-\texttt{one\_plus\_omega} uses the continuity of ordering functions (applied to  \texttt{(plus 1)}) and compares the limit of the $\omega$-sequences $i_{(i \in \mathbb{N})}$ and
-$(1+i)i_{(i \in \mathbb{N})}$, whereas in the library  \texttt{Epsilon0/T1}, the equality 
-$1+\omega=\omega$ is just proved with \texttt{reflexivity}!
-
-\subsubsection{Multiplication by a natural number}
-
-The multiplication of an ordinal by a natural number is defined in terms of addition.
-This operation is useful for the study of Cantor normal forms.
-
-\input{movies/snippets/Addition/multFin}
-
-\section{The exponential of basis \texorpdfstring{$\omega$}{omega}}
-
-In this section, we define the function which maps any $\alpha\in\mathbb{O}$ to
-the ordinal  $\omega^\alpha$, also written 
-$\phi_0(\alpha)$. 
-It is an opportunity to apply the definitions and results of the preceding section. 
-Indeed,  Schütte first defines a subset of $\mathbb{O}$: the set of additive principal ordinals, and $\phi_0$  is just defined as the ordering function of this set.
-
-\subsection{Additive principal ordinals}
-
-\index{maths}{Ordinal numbers!Additive principal ordinals}
-\index{hydras}{Library Schutte!Predicates!AP@AP (additive principal ordinals)}
-
-\begin{definition}
-A non-zero ordinal  $\alpha$ is said to be \emph{additive principal} if, for all  $\beta<\alpha$, $\beta+\alpha$ is equal to  $\alpha$.
-We call \texttt{AP} the set of additive principal ordinals.
-
-\end{definition}
-
-\noindent\emph{From Module \href{../theories/html/hydras.Schutte.AP.html}%
-{\texttt{Schutte.AP}}}
-
-\input{movies/snippets/AP/APDef}
-
-\subsection{The function \texttt{phi0}}
-
-Let us call  $\phi_0$ the ordering function of \texttt{AP}.
-In the mathematical text, we shall use indifferently the notations  $\omega^\alpha$ and$\phi_0(\alpha)$.
-
-\index{hydras}{Library Schutte!Functions!phi0}
-
-\input{movies/snippets/AP/phi0Def}
-
-\subsection{Omega-towers and the ordinal \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-Using $\phi_0$, we can define recursively the set of finite omega-towers.
-
-\input{movies/snippets/AP/omegaTower}
-
-\label{sect:epsilon0-as-limit}
-Then, the ordinal  $\epsilon_0$ is defined as the limit of the sequence of all finite towers (a kind of infinite tower).
-
-\input{movies/snippets/AP/epsilon0Def}
-
-The rest of our library \texttt{AP} is devoted to the proof of properties of additive principal ordinals, hence of the ordering function  $\phi0$ and the ordinal $\epsilon_0$ (which we could not express within the type \texttt{T1}).
-
-\subsection{Properties of the set  \texttt{AP}}
-
-The set of additive principal ordinals is not empty: it contains at least the ordinals  $1$ and  $\omega$.
-
-\inputsnippets{AP/APOne,AP/APOne_least_AP,AP/APOne_AP_omega,AP/APOne_omega_second_AP}
-
-The set  \texttt{AP} is  \emph{closed} under addition, and unbounded.
-\label{lemma:AP-plus-closed}
-
-\input{movies/snippets/AP/APPlusClosed}
-
-\input{movies/snippets/AP/APUnbounded}
-
-Finally, \texttt{AP} is (topologically) \emph{closed} and ordered by the segment of all countable ordinals.
-
-\index{hydras}{Library Schutte!Predicates!Closed}
-
-From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{~Schutte.Schutte\_basics}}
-
-\input{movies/snippets/Schutte_basics/ClosedDef}
-\input{movies/snippets/AP/APClosed}
-
-\subsubsection{Properties of the function \texorpdfstring{$\phi_0$}{phi0}}
-
-The ordering function $\phi_0$ of the set \texttt{AP} is defined on the full set $\mathbb{O}$ and is continuous (Schütte calls such a  function  \emph{normal}).
-
-\begin{Coqsrc}
-Theorem normal_phi0 : normal phi0 AP.
-\end{Coqsrc}
-
-The following properties come from  the definition of $\phi_0$ as the ordering function of \texttt{AP}.
-It may be interesting to compare these proofs with the computational ones described in Chapter ~\ref{chap:T1}.
-
-\input{movies/snippets/AP/APPhi0}
-
-\subsection{A last example}
-\label{Ex42-schutte}
-
-Let us prove again the equality $\omega+42+\omega^2= \omega^2$.Let us recall that $\omega^2$ is an abbreviation of $\phi_0(2)$,
-\emph{i.e} the third  additive principal ordinal.
-
-\inputsnippets{Schutte/Ex42a}
-
-Our proof is very different from the computational proof of Sect~\vref{Ex42-E0}.
-By definition of additive principal ordinals, 
-it suffices to prove the inequality $\omega+42< \phi_0(2)$.
-
-\inputsnippets{Schutte/Ex42b}
-
-Since the set \texttt{AP} of additive principals  is closed under addition
-(by Lemma \texttt{AP\_plus\_closed}, page~\pageref{lemma:AP-plus-closed}) , it suffices to prove the inequalities $\omega<\phi_0(2)$ and $42<\phi_0(2)$.
-
-\inputsnippets{Schutte/Ex42d, Schutte/Ex42c, Schutte/Ex42e}
-
-\section{More about \texorpdfstring{$\epsilon_0$}{\texttt{epsilon0}}}
-
-Let us recall that the limit ordinal  $\epsilon_0$ cannot be written within the type \texttt{T1}.
-Since we are now considering the set of all countable ordinals, we can now prove some properties of this ordinal.
-
-We prove the inequality  $\alpha<\omega^\alpha$ whenever $\alpha < \epsilon_0$.
-\emph{Note that this condition was implicit in Module
-\href{../theories/html/hydras.Epsilon0.T1.html\#lt_phi0}{Epsilon0.T1}.}
-
-\input{movies/snippets/AP/ltPhi0}
-
-The proof is as follows:
-\begin{enumerate}
-\item Since $\alpha<\epsilon_0$, consider the least $i$ such that $\alpha$ is strictly less than the omega-tower of height $i$.
-\item
-  \begin{itemize}
-  \item If $i=0$, then the result is trivial (because $\alpha=0$)
- \item  Otherwise let $i=j+1$; 
-          $\alpha$ is greater than or equal to the omega-tower of height $j$.
-         By monotony,  $\phi_0(\alpha)$ is greater than or equal to 
-        the omega-tower of height $j+1$, thus strictly greater than $\alpha$
-  \end{itemize}
- \end{enumerate}
-
-Moreover,  $\epsilon_0$ is the least ordinal $\alpha$ that verifies the equality 
-$\alpha = \omega^\alpha$, in other words, the least fixpoint of the function  $\phi_0$.
-
-\input{movies/snippets/AP/epsilon0Lfp}
-
-\section{Critical ordinals}
-
-\index{maths}{Ordinal numbers!Critical ordinals}
-\index{hydras}{Library Schutte!Predicates!Cr@Cr (critical ordinals)}
-
-For any  (countable) ordinal $\alpha$, the set $\textit{Cr}(\alpha)$ is inductively defined 
-as follows by Schütte (p.81 of~\cite{schutte}).
-
-\begin{quote}
-  \begin{itemize}
-  \item $\textit{Cr}(0)$ is the set \textit{AP} of additive principal ordinals.
-  \item If $0<\alpha$, then $\textit{Cr}(\alpha)$ is the intersection of all the sets of fixpoints of the $\textit{Cr}(\beta)$ for $\beta<\alpha$.
-  \end{itemize}
-\end{quote}
-
-This definition is translated in \coq{} in 
-Module \href{../theories/html/hydras.Schutte.Critical.html}%
-{\texttt{Schutte.Critical}}, as the least fixpoint of a functional.
-
-\input{movies/snippets/Critical/CrDef}
-
-Lets us denote by $\phi_\alpha$ the ordering function of the set $\textit{Cr}(\alpha)$ and by $A_\alpha$ its ordering segment.
-
-\input{movies/snippets/Critical/phiDef}
-
-\label{sect:phi-schutte}
-
-For instance,  we prove that $\textit{Cr}(0)$ is the set of additive principals and that $\epsilon_0$
-belongs to $\textit{Cr}(1)$.
-
-\inputsnippets{Critical/CrZeroAP,
-  Critical/epsilon0Cr1}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
- Prove that $\epsilon_0$ is the least element of $\textit{Cr}(1)$.
-\end{exercise}
-
-\subsection{A flavor of infinity}
-
-The family of the $\textit{Cr}(\alpha)$s is made of infinitely many unbounded (hence infinite) sets.
-Let us quote Lemma 5, p. 82  of~\cite{schutte}:
-\begin{quote}
-  For all $\alpha$, the set $\textit{Cr}(\alpha)$ is closed (for the least upper bound of non-empty countable sets) and unbounded.
-\end{quote}
-
-We prove this result by a transfinite induction on $\alpha$ of the conjunction of  both properties.
-
-\index{maths}{Transfinite induction}
-
-\input{movies/snippets/Critical/Lemma5}
-
-\section{Cantor normal form}
-
-The notion of Cantor normal form is defined for all countable ordinals.
-Nevertheless, note that, contrary to the implementation based on type \texttt{T1},
-the Cantor normal form of an ordinal $\alpha$ may contain $\alpha$ as a 
-sub-term\footnote{This would prevent us from trying to represent Cantor normal forms as finite trees (like in Sect.~\ref{sec:T1-inductive-def})}.
-
-\index{maths}{Ordinal numbers!Cantor normal form}
-\index{hydras}{Library Schutte!Predicates!is\_cnf\_of@is\_cnf\_of (to be a Cantor normal form of}
-
-We represent  Cantor normal forms as lists of ordinals.
-A  list $l$ is a Cantor normal form of a given ordinal $\alpha$ if it satisfies two conditions:
-
-\begin{itemize}
-\item The list  $l$ is sorted (in decreasing order) w.r.t. the order $\leq$
-\item The sum of all the  $\omega^{\beta_i}$ where the $\beta_i$ are the terms of $l$ (in this order) is equal to $\alpha$.
-\end{itemize}
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Schutte.CNF.html\#cnf_t}%
-{\texttt{Schutte.CNF}}}
-
-\input{movies/snippets/CNF/Defs}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction on $\alpha$, we prove that every countable ordinal $\alpha$ 
- has at least a Cantor normal form.
-
-\input{movies/snippets/CNF/cnfExists}
-
-By structural induction on lists, we prove that this normal form is unique.
-
-\inputsnippets{CNF/cnfUnicity,CNF/cnfExUnique}
-
-Finally, the following two lemmas relate  $\epsilon_0$ with Cantor normal forms.
-
-If $\alpha<\epsilon_0$, then the Cantor normal form of $\alpha$ is made of ordinals strictly less than $\alpha$.
-
-\input{movies/snippets/CNF/cnfLtEpsilon0}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Please consider the following statement :
-
-\begin{Coqsrc}
-Lemma cnf_lt_epsilon0_iff : 
- forall l alpha, 
-   is_cnf_of alpha l ->  
-   (alpha < epsilon0 <->  Forall (fun beta =>  beta < alpha) l).
-\end{Coqsrc}
-
-Is it true ?
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/schutte_cnf_counter_example.v}{exercises/ordinals/schutte\_cnf\_counter\_example.v}.}
-\end{exercise}
-
-Finally, the Cantor normal form of $\epsilon_0$ is just $\omega^{\epsilon_0}$.
-
-\input{movies/snippets/CNF/cnfOfEpsilon0}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Implement pages 82 to 85 of~\cite{schutte} (critical, strongly critical, maximal critical ordinals, Feferman's ordinal $\Gamma_0$).
-\end{project}
-
-\begin{remark}
-The sub-directory
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/Gamma0}{theories/oridinals/Gamma0} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form.
-\end{remark}
-
-\section{An embedding of \texttt{T1} into \texttt{Ord}}
-
-Our library 
-\href{../theories/html/hydras.Schutte.Correctness_E0.html}%
-{\texttt{Schutte.Correctness\_E0}} establishes the link between two very different modelizations of ordinal numbers. In other words, it ``validates'' a data structure in terms of
-a classical mathematical discourse considered as a model. 
-First, we define a function from \texttt{T1} into  \texttt{Ord} by structural recursion.
-
-\input{movies/snippets/Correctness_E0/injectDef}
-
-This function enjoys good commutation properties with respect to the main operations which
-allow us to build Cantor normal forms.
-
-\input{movies/snippets/Correctness_E0/commutationLemmas}
-\input{movies/snippets/Correctness_E0/injectPlus}
-\input{movies/snippets/Correctness_E0/injectMultFinR}
-
-% \begin{Coqsrc}
-% Theorem inject_mono (beta gamma : T1) :
-%   T1.lt  beta gamma -> 
-%   T1.nf beta -> T1.nf gamma -> 
-%   inject beta < inject gamma.
-
-% Theorem inject_injective (beta gamma : T1) : nf beta -> nf gamma ->
-%   inject beta = inject gamma -> beta = gamma.
-% \end{Coqsrc}
-
-Finally, we prove that \texttt{inject} is a bijection from the set of all terms of \texttt{T1} in normal form to the set 
-\texttt{members epsilon0} of the elements of \texttt{Ord} strictly less than  $\epsilon_0$.
-
-\input{movies/snippets/Correctness_E0/injectLtEpsilon0}
-\input{movies/snippets/Correctness_E0/embedding}
-
-\subsection{Remarks}
-Let us recall that the library \href{../theories/html/hydras.Schutte.Schutte.html}%
-{\texttt{Schutte}} depends on five \emph{axioms} and lies explicitly in the  
-framework of classical logic with a weak version of the axiom of choice
-(please look at the documentation of
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.ChoiceFacts.html}{\texttt{Coq.Logic.ChoiceFacts}}).
-Nevertheless, the other modules:
-\href{../theories/html/hydras.Epsilon0.Epsilon0.html}%
-{\texttt{Epsilon0}},
-\href{../theories/html/hydras.Hydra.Hydra.html}%
-{\texttt{Hydra}}, et 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}%
-{\texttt{Gamma0}}
-do not import any axioms and are really constructive.
-
-\index{hydras}{Projects}
-\begin{project}
-There is no construction of ordinal multiplication in~\cite{schutte}. 
-It would be interesting to derive this operation from Schütte's axioms,
-and prove its consistence with multiplication in ordinal notations for 
-$\epsilon_0$ and $\Gamma_0$.
-\end{project}
-
-\section{Related work}
-
-In~\cite{grimm:hal-00911710}, José Grimm establishes the consistency between our ordinal notations \texttt{T1} and \texttt{T2} (Veblen normal form) and his implementation
-of ordinal numbers after Bourbaki's set theory.
-
-The Gaia project ~\url{https://github.com/coq-community/gaia} maintains Grimm's  theory of ordinals as part of coq-community on GitHub. Integration
-of the present ordinal theory with Gaia, i.e., relating the different notions of ordinals
-and transferring relevant results, is an interesting project.
-First experiments in that direction are developped in
-the \href{https://github.com/coq-community/hydra-battles/blob/master/theories/gaia/}{theories/gaia/} directory.
-
-\chapter{The Ordinal \texorpdfstring{$\Gamma_0$}{Gamma0} (first draft)}
-
-\emph{This chapter and the files it presents are still very incomplete, considering the impressive properties of $\Gamma_0$~\cite{Gallier91}.
-We hope to add new material soon, and accept contributions!}
-
-\begin{todo}
-  Build a T2Bridge! (note: Gamma0's wellfoundedness is missing in \gaia (?).
-\end{todo}
-
-\section{Introduction}
-We present a notation system for the ordinal $\Gamma_0$, following Chapter V, Section 14 of~\cite{schutte}: ``A notation system for the ordinals $<\Gamma_0$''.
-We try to be as close as possible to Schütte's text and usual practices of \coq{} developments.
-
-The ordinal $\Gamma_0$ is defined in Section 13 of ~\cite{schutte} as the least \emph{strongly critical ordinal}. It is widely known as the \emph{Feferman-Schütte ordinal}.
-
-Section V, 13 of~\cite{schutte} defines \emph{strongly critical} and
-\emph{maximal $\alpha$-critical} ordinals:
-
-\begin{itemize}
-\item $\alpha$ is strongly critical if
-$\alpha$ is $\alpha$-critical,
-\item $\gamma$ is maximal $\alpha$-critical if $\gamma$ is $\alpha$-critical, and, for all $\xi>\alpha$, $\gamma$ is not $\xi$-critical.
-
-\end{itemize}
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Schutte.Critical.html\#strongly_critical}%
-{\texttt{Schutte.Critical}}}
-
-\input{movies/snippets/Critical/Gamma0Def}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that a (countable)  ordinal $\alpha$ is strongly critical iff 
-$\phi_\alpha(0)=\alpha$ (Theorem 13.13 of~\cite{schutte} ). 
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that the set of strongly critical ordinals is unbounded and closed (Theorem 13.14 of~\cite{schutte} ).
-Thus this set is not empty,  hence has a least element.
-Otherwise, the definition of $\Gamma_0$ above would be useless.
-\end{project}
-
-In the present version of this development, we  only study $\Gamma_0$ as a notation system, much more powerful than the ordinal notation for $\epsilon_0$.
-
-%\index{Projects}
-%
-%\begin{project}
-% Schûtte's section 13 of~\cite{schutte} contain several definitions and lemmas which give another view on $\Gamma_0$.  We leave it as a project to implement them in \coq{}.  
-%\end{project}
-
-\section{The type \texttt{T2} of ordinal terms}
-
-The notation system for ordinals less than $\gamma_0$ comes from the following theorem of~\cite{schutte}, where $\psi\,\alpha$ is the ordering function 
-of the set of maximal $\alpha$-critical ordinals.
-
-\index{hydras}{Library Gamma0!Types!T2}
-
-\begin{quote}
-  Any ordinal $\not= 0$ which is not strongly critical can be expressed in terms of $+$ and $\psi$.
-\end{quote}
-
-\index{hydras}{Projects}
-\begin{project}
-This theorem is not formally proved in this development yet. It should be!
-\end{project}
-
-Like in Chapter~\ref{chap:T1}, we define an inductive type with two constructors, one for $0$, the other for the construction $\psi(\alpha,\beta)\times(n+1)+\gamma$, adapting a Manolios-Vroon-like notation~\cite{Manolios2005} to
-\emph{Veblen normal forms}.
-\label{types:T2}
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T2}%
-{\texttt{Gamma0.T2}}}
-
-\inputsnippets{T2/T2Def, T2/psiDef}
-
-Like in chapter~\ref{chap:T1}, we get familiar with the type \texttt{T2} by recognizing simple constructs like finite ordinals, $\omega$, etc., as inhabitants of \texttt{T2}.
-
-\input{movies/snippets/T2/Notations}
-
-\section{A strict order on T2}
-
-Let us define a strict order on type \texttt{T2}. The following definition is 
-an adaptation of Schütte's, taking into account the multiplications by a natural number (inspired by~\cite{Manolios2005}, and also present in \texttt{T1}).
-
-\label{sect:t2-lt-def}
-
-\input{movies/snippets/T2/ltDef}
-
-Seven constructors! In order to get accustomed with this definition, let us look at a small set of examples, covering all the constructors of \texttt{lt}.
-
-\input{movies/snippets/T2/ltExamples}
-
-\index{hydras}{Projects}
-\begin{project}
-Write a tactic that solves automatically goals of the form (\texttt{$\alpha$ t2< $\beta$}), where $\alpha$ and $\beta$ are closed terms of type \texttt{T2}.
-\end{project}
-
-\section{Veblen normal form}
-\begin{definition}
-  A term of the form $\psi(\alpha_1,\beta_1)\times n_1+ \psi(\alpha_2,\beta_2)\times n_2+\dots+\psi(\alpha_k,\beta_k)\times n_k$ is said to be in
- \emph{[Veblen] normal form} if for every $i<n$, $\psi(\alpha_i,\beta_i)<\psi(\alpha_{i+1},\beta_{i+1})$, all the $\alpha_i$ and $\beta_i$ are in normal form, and all the $n_i$ are strictly positive integers.
-\end{definition}
-
-\input{movies/snippets/T2/nfDef}
-
-Let us look at some positive examples (we have to prove some inversion lemmas before proving counter-examples).
-
-\input{movies/snippets/T2/nfExamples}
-
-\subsection{Length of a term}
-
-The notion of \emph{term length} is introduced by Schütte as a helper for proving (at least) the \emph{trichotomy} property and transitivity of the strict order \texttt{lt} on \texttt{T2}.
-These properties are proved by induction on length.
-
-\subsection{Trichotomy}
-
-\emph{Trichotomy} is another name for the well-known property of decidable total ordering (like Standard Library's \texttt{Compare\_dec.lt\_eq\_lt\_dec}).
-
-We first prove by induction on $l$ the following lemma:
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0\#tricho_aux}%
-{\texttt{Gamma0.Gamma0}}}
-
-\input{movies/snippets/Gamma0/tricho}
-
-\input{movies/snippets/Gamma0/compare}
-
-With the help of \texttt{compare}, we get a boolean version of \texttt{nf}
-(being in Veblen normal form).
-
-\input{movies/snippets/Gamma0/nfb}
-
-\section{Main functions on \texttt{T2}}
-
-\subsection{Successor}
-The successor function is defined by structural recursion.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#succ}%
-{\texttt{Gamma0.T2}}}
-
-\input{movies/snippets/T2/succDef}
-
-\subsection{Addition}
-
-Like for Cantor normal forms (see Sect.~\ref{sect:infix-plus-T1}),  the definition of addition in \texttt{T2}  requires comparison between ordinal terms.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0T2.html\#succ}%
-  {\texttt{Gamma0.Gamma0}}}
-
-\input{movies/snippets/Gamma0/plusDef}
-
-\subsection{The Veblen function \texorpdfstring{$\phi$}{\texttt{phi}}}
-
-The enumeration function of critical ordinals, presented in Sect.~\vref{sect:phi-schutte}, is recursively defined in type \texttt{T2}.
-
-\input{movies/snippets/Gamma0/phiDef}
-
-Despite its complexity, the function \texttt{phi} is well adapted to proofs by simplification or computation.
-
-The relation between the constructor $\psi$ and the function $\phi$ is
-studied in~\cite{schutte}, and partially implemented in this development.
-\emph{Please contribute!}
-
-For instance, the following theorem states that, if $\gamma$ is the sum of a limit ordinal $\beta$ and a finite ordinal $n$, and $\beta$ is a fixpoint of
-$\phi(\alpha)$, then $\psi(\alpha,\gamma)=\phi_\alpha(\gamma+1)$.
-
-\input{movies/snippets/Gamma0/phiPsi}
-\input{movies/snippets/Gamma0/Ex9}
-
-On the other hand, $\phi$ can be expressed in terms of $\psi$.
-
-\input{movies/snippets/Gamma0/phiOfPsi}
-\input{movies/snippets/Gamma0/Ex10}
-
-\index{hydras}{Projects}
-\begin{project}
-Please study a way to pretty print ordinal terms in Veblen normal form (see Section~\vref{sect:ppT1}).
-\end{project}
-
-\section{An ordinal notation for \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}}
-
-In order to consider type \texttt{T2} as an ordinal notation, we have to build an instance of class \texttt{ON} (See Definition page~\pageref{types:ON}).
-
-First, we define a type that contains only terms in Veblen normal form, and redefine \texttt{lt} and \texttt{compare} by delegation (see for comparison the construction of type \texttt{E0} in Sect.~\vref{sect:E0-def}).
-
-\input{movies/snippets/Gamma0/G0a}
-\input{movies/snippets/Gamma0/G0b}
-
-Then, we buils an instance of class \texttt{ON}. 
-function \texttt{compare} is correct.
-
-\input{movies/snippets/Gamma0/ltSto}
-\input{movies/snippets/Gamma0/ltWf}
-\input{movies/snippets/Gamma0/ONGamma0}
-\input{movies/snippets/Gamma0/G0z}
-
-\begin{remark}
-The proof of \texttt{lt\_wf} has been written by \'Evelyne Contejean, using her library on the recursive path ordering (see also remark~\vref{remark:a3pat}).
-\end{remark}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that \texttt{Epsilon0} (page~\pageref{instance-epsilon0})
-is a sub-notation system of \texttt{Gamma0}.
-
-Prove that the implementations of \texttt{succ}, \texttt{+}, $\phi_0$, etc.
-are compatible in both notation systems.
-
-Note that a function \texttt{T1\_inj} from \texttt{T1} to \texttt{T2} has already been defined. It may help to complete the task.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T1_to_T2}%
-{\texttt{Gamma0.T2}}}
-
-\input{movies/snippets/T2/T1ToT2}
-
-\end{project}
-
-\begin{project}
-Prove that the notation system \texttt{Gamma0} is a correct implementation 
-of the segment $[0,\Gamma_0)$ of the set of countable ordinals.
-\end{project}

### 2022-05-08 doc/large-sets-chapter.tex
reorganization of tex files

--- 2022-04-11 doc/part-hydras.tex
@@
-\section*{Introduction}
-
-In this part, we present a development  for the \coq{} proof assistant, after the work of Kirby and Paris~\cite{KP82}. This formalization contains the following main parts:
-
-\begin{itemize}
-\item Representation in \coq{} of hydras and hydra battles.
-\item A proof that every battle is finite and won by Hercules.
-This proof is based on a \emph{variant} which maps any hydra to an ordinal strictly less than $\epsilon_0$ and is strictly decreasing along any battle.
-
-\item Using a combinatorial toolkit designed by J.~Ketonen and R.~Solovay~\cite{KS81}, we prove that, for any ordinal $\mu<\epsilon_0$, there exists no such variant mapping any hydra to an ordinal strictly less than $\mu$.
-Thus, the complexity of $\epsilon_0$ is really needed in the previous proof.
-
-\item We prove a relation between the length of a ``classic''  kind of  battles \footnote{This class is also called \emph{standard} in this document (text and proofs).
-The \emph{replication factor} of the hydra is exactly $i$ at the $i$-th round of the battle (see Sect~\vref{sect:replication-def}).}
-and the Wainer-Hardy hierarchy of ``rapidly growing functions'' $H_\alpha$~\cite{Wainer1970}.
-The considered class of battles, which we call \emph{standard},  is the most considered one in the scientific  literature (including popularization).
-\end{itemize}
-
-Simply put, this document tries to combine the scientific interest of two articles~\cite{KP82, KS81} and a book~\cite{schutte} with the playful activity of truly proving theorems.
-We hope  that such a work, besides exploring a nice piece of discrete mathematics,
-will show how \coq{} and its standard library are well fitted to help us to understand some non-trivial mathematical developments, and also to experiment the constructive parts of  the proof through functional programming.
-
- We also hope to provide a little clarification on infinity (both potential and actual) through the notions of function, computation, limit,
- type and proof.
-
- \subsection*{Compatibility with \gaia (in progress)}
- \label{sect:gaia-first-intro}
-The \gaia project~\cite{Gaia} by Jos\'e Grimm, Alban Quadrat,  and Carlos Simpson,
-aims  to formalize mathematics in \coq  in
-the style of Nicolas Bourbaki.
-It contains many definitions and results about ordinal numbers.
-In Chapter~\ref{gaia-chapter}, we present some modules which allow \HydrasLib' users to apply lemmas proven in \gaia, and vice versa.
-Remarks about compatibility with \gaia are signalled with the picture \gaiasign.
-
-%\section{Remarks}
-
-\subsection*{Difference from Kirby and Paris's work}
-In~\cite{KP82}, Kirby and Paris show  that there is no proof of termination of all hydra battles in Peano Arithmetic (PA).
-Since we are used to writing proofs in higher order logic, the restriction to PA was quite unnatural for us.
-So we chose to prove another statement without any reference to PA, by considering a class of proofs indexed by ordinal numbers up to $\epsilon_0$.
-
-\subsection*{State of the development}
-The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on  notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
-We added new material: axiomatic definitions of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, function derinition by equations, etc.
-
-We are now working in order to make clumsy proofs more readable, simplify definitions, and ``factorize'' proofs as much as possible. 
-Many possible improvements are suggested as ``todo''s or ``projects'' in this text.
-
-\section*{Future work (projects)}
-\index{hydras}{Projects}
-
-This document and the proof scripts are far from being complete.
-
-First, there must be a lot of typos to correct, references and index items to add. Many proofs are too complex and should be simplified, etc.
-
-The following extensions are planned, but help is needed:
-
-\begin{itemize}
-\item Semiautomatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
-%\item Extension to $\Gamma_0$ (in Veblen normal form)
-\item More lemmas about hierarchies of rapidly growing
-  functions, and their relationship 
-    with primitive recursive functions and provability in Peano arithmetic 
-(following~\cite{KS81, KP82}).
-\item From \coq's point of view, this development could be used as an illustration of the evolution of the software, every time new libraries and sets of tactics could help to simplify the proofs.
-\end{itemize}
-
-\subsection*{Main references}
-
-In our development, we adapt the definitions and prove many theorems which
-we found in the following articles. 
-\begin{itemize}
-\item ``Accessible independence results for Peano arithmetic''  by Laurie Kirby and Jeff Paris~\cite{KP82}
-\item ''Rapidly growing Ramsey Functions'' by Jussi Ketonen and Robert Solovay~\cite{KS81}
-\item ``The Termite and the Tower'', by Will Sladek~\cite{Sladek07thetermite}
-\item Chapter V of ``Proof Theory'' by Kurt Schütte~\cite{schutte}
-\end{itemize}
-
-\chapter{Hydras and hydra games}
-
-\label{sec:orgheadline91}
-\label{chapter:hydras}
-
-This chapter is dedicated to the representation of hydras and rules of the hydra game in \coq's specification language:~\gallina.
-
-Technically, a \emph{hydra} is just a finite ordered tree, each node of which 
-has any number of sons. Contrary to the computer science tradition, we display the hydras 
-with the heads up and the foot (i.e., the root of the tree) down.
-Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the 
-module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}). 
-\emph{For a less formal description of hydras, please see 
-\url{https://www.smbc-comics.com/comic/hydra}.}
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {$\bullet$};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H4) at (4,2) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N3) to [bend left= 10]  (H1) ;
-\draw (N3) to [bend right= 16] (H2);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (foot) to [bend right = 10] (H4) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-\caption{The hydra \texttt{Hy} \label{fig:Hy}}
-\end{figure}
-
-We use a specific vocabulary for talking about hydras. Table~\ref{tab:hyd2tree} shows the correspondence between our terminology and the usual vocabulary for trees in computer science.
-
-\begin{figure}[h]
-  \centering
-  \begin{tabular}{ll}
-Hydras & Finite rooted trees\\
-\hline
-foot & root\\
-head & leaf\\
-node & node\\
-segment  & (directed) edge \\
-sub-hydra & subtree\\
-daughter & immediate subtree\\
-\end{tabular}
-  \caption{Translation from hydras to trees}
-  \label{tab:hyd2tree}
-\end{figure}
-
-The hydra \texttt{Hy} has a \emph{foot} (below), five \emph{heads}, and eight \emph{segments}. 
-We leave it to the reader to define various parameters such as the height, the size, the highest arity (number of sons of a node) of a hydra.
-In our example, these parameters have the respective values $4$, $9$ and $3$.
-
-\subsection{The rules of the game}
-
-\label{sec:orgheadline44}
-\label{sect:replication-def}
-
-A \emph{hydra battle} is a fight between Hercules and the Hydra. 
-More formally, a  battle is a sequence of \emph{rounds}.
-At each round:
-\begin{itemize}
-\item If the hydra is composed of just one head, the battle is finished
-and  Hercules is the winner.
-\item Otherwise, Hercules chops off \emph{one} head of the hydra,
-
-\begin{itemize}
-\item If the head is at distance 1 from the foot, the head is just lost by the hydra, with no more reaction.
-\item Otherwise, let us denote by \(r\) the node that was at distance \(2\) from 
-the removed head in the direction of the foot,  and consider the  sub-hydra \(h'\) of \(h\), whose  root is \(r\) \footnote{$h'$ will be called ``the wounded part of the hydra'' in the subsequent text.
-In Figures~\vref{fig:Hy2} and ~\vref{fig:Hy4}, this sub-hydra  is displayed in red.}.
-Let $n$ be some natural number.
-Then $h'$ is replaced by  $n+1$ of copies of \(h'\) which share the same root $r$.
- The \emph{replication factor} $n$ may be different (and generally is)   at each round of the fight.
-It may be chosen by the hydra, according to its strategy, or imposed by some 
-particular rule.
-In many presentations of hydra battles, this number is increased by $1$ at each round.
-In the following presentation, we will also consider battles where the hydra is free to choose its ~replication factor at each round of the battle\footnote{Let us recall that, if the chopped-off head was at distance 1 from the foot, the replication factor is meaningless.}.
-\end{itemize}
-\end{itemize}
-
-Note that the description given in~\cite{KP82} of the replication process in hydra battles is also  semi-formal.
-
-\label{original-rules}
-
-\begin{quote}
-  ``From the node that used to be attached to the head which was just chopped off, traverse one 
-segment towards the root until the next node is reached. From this node sprout $n$ replicas of 
-that part of the hydra (after decapitation) which is ``above'' the segment just traversed, i.e., those 
-nodes and segments from which, in order to reach the root, this segment would have to be 
-traversed. If the head just chopped off had the root of its nodes, no new head is grown. ''
-\end{quote}
-
-Moreover, we note that this description is in \emph{imperative} terms.
-In order to formally study the properties of hydra battles, we prefer to use a mathematical vocabulary, i.e., graphs, relations, functions, etc.
-Thus, the replication process will be represented as a binary relation on a data type \texttt{Hydra},
-linking the state of the hydra \emph{before} and \emph{after} the transformation.
-A battle will thus be represented as a sequence of terms of type \texttt{Hydra}, respecting the rules of the game.
-
-\subsection{Example}
-Let us start a battle between Hercules and the hydra \texttt{Hy} of Fig.~\ref{fig:Hy}.
-
-At the first round, Hercules chooses to chop off the rightmost head of \texttt{Hy}.
-Since this head is near the floor, the hydra simply loses this head. Let us call 
- \texttt{Hy'} the resulting state of the hydra, represented in Fig.~\vref{fig:Hy-prime}.
-
-Next, assume Hercules chooses to chop off one of the two highest heads of \texttt{Hy'}, for instance the rightmost one.
-Fig.~\vref{fig:Hy2} represents the broken segment in dashed lines, and the part that will be replicated in red.
-Assume also that the hydra decides to add 4 copies of the red part\footnote{In other words, the replication factor at this round is equal to $4$.}.
-We obtain a new state \texttt{Hy''} depicted in Fig.~\ref{fig:Hy3}.
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {$\bullet$};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-%\node (H4) at (6,0) {$\Xey[2][lightgray]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3) ;
-\draw (N3) to [bend left= 10]  (H1) ;
-\draw (N3) to [bend right= 16] (H2);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-
-\caption{\texttt{Hy'}: the state  of \texttt{Hy} after one round \label{fig:Hy-prime}}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4)  {{\color{lightred}$\bullet$}};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Sey[2][lightred]$};
-%\node (H2) at (5,0) {$\Xey[2][lightgray]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\node (ex) at (5,8) {};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw  (N2) -- (N3)[draw=lightred];
-\draw (N3) to   [bend left= 10](H1) [draw=lightred];
-\draw [dashed] (N3) to [bend left= 10](ex);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-\caption{A second beheading}
-\label{fig:Hy2}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H11) at (2,8) {$\Smiley[2][vertfluo]$};
-\node (H12) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H13) at (6,8) {$\Smiley[2][vertfluo]$};
-\node (H14) at (8,8) {$\Smiley[2][vertfluo]$};
-
-\node (N3) at (1,6) {$\bullet$};
-\node (N31) at (2,6) {$\bullet$};
-\node (N32) at (3,6) {$\bullet$};
-\node (N33) at (4,6) {$\bullet$};
-\node (N34) at (5,6) {$\bullet$};
-
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N2) -- (N31);
-\draw (N2) -- (N32);
-\draw (N2) -- (N33);
-\draw (N2) -- (N34);
-\draw (N3) to   [bend left= 10](H1) ;
-\draw (N31) to   [bend left= 10](H11) ;
-\draw (N32) to   [bend left= 10](H12) ;
-\draw (N33) to   [bend left= 10](H13) ;
-\draw (N34) to   [bend left= 10](H14) ;
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend left= 10]  (H5) ;
-\end{tikzpicture}
-\caption{\texttt{Hy''}: the state of \texttt{Hy} after two rounds \label{fig:Hy3}}
-\end{figure}
-
-Figs.~\ref{fig:Hy4} and~\vref{fig:Hy5} represent a possible third round of the battle, with a replication factor equal to $2$. Let us call \texttt{Hy'''} the state of the hydra after that third round.
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0)  {{\color{lightred}$\bullet$}};
-\node (N1) at (2,2) {{\color{lightred}$\bullet$}};
-\node (N2) at (2,4) {{\color{lightred}$\bullet$}};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (exN4) at (4,4) {};
-\node (H1) at (0,8) {$\Sey[2][lightred]$};
-\node (H11) at (2,8) {$\Sey[2][lightred]$};
-\node (H12) at (4,8) {$\Sey[2][lightred]$};
-\node (H13) at (6,8) {$\Sey[2][lightred]$};
-\node (H14) at (8,8) {$\Sey[2][lightred]$};
-
-\node (N3) at (1,6) {{\color{lightred}$\bullet$}};
-\node (N31) at (2,6) {{\color{lightred}$\bullet$}};
-\node (N32) at (3,6) {{\color{lightred}$\bullet$}};
-\node (N33) at (4,6) {{\color{lightred}$\bullet$}};
-\node (N34) at (5,6) {{\color{lightred}$\bullet$}};
-
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-%\node (H5) at (4,0) {$\Xey[2][lightgray]$};
-\draw (foot) -- (N1)[very thick,draw=lightred] ;
-\draw (N1) -- (N2)[draw=lightred];
-\draw (N2) -- (N3)[draw=lightred];
-\draw (N2) -- (N31)[draw=lightred];
-\draw (N2) -- (N32)[draw=lightred];
-\draw (N2) -- (N33)[draw=lightred];
-\draw (N2) -- (N34)[draw=lightred];
-\draw (N3) to   [bend left= 10](H1) [draw=lightred];
-\draw (N31) to   [bend left= 10](H11) [draw=lightred];
-\draw (N32) to   [bend left= 10](H12) [draw=lightred];
-\draw (N33) to   [bend left= 10](H13) [draw=lightred];
-\draw (N34) to   [bend left= 10](H14) [draw=lightred];
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw [dashed] (N1) to  [bend left= 10](exN4);
-\end{tikzpicture}
-\caption{A third beheading (wounded part in red) \label{fig:Hy4}}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-
-\node (foot) at (10,0) {$\bullet$};
-
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H1) at (0,8) {$\Smiley[1][vertfluo]$};
-\node (H11) at (2,8) {$\Smiley[1][vertfluo]$};
-\node (H12) at (4,8) {$\Smiley[1][vertfluo]$};
-\node (H13) at (6,8) {$\Smiley[1][vertfluo]$};
-\node (H14) at (8,8) {$\Smiley[1][vertfluo]$};
-
-\node (N3) at (1,6) {$\bullet$};
-\node (N31) at (2,6) {$\bullet$};
-\node (N32) at (3,6) {$\bullet$};
-\node (N33) at (4,6) {$\bullet$};
-\node (N34) at (5,6) {$\bullet$};
-
-\node (H0) at (-3,3) {$\Smiley[1][vertfluo]$};
-
-\draw (foot) to [bend left=10] (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N2) -- (N31);
-\draw (N2) -- (N32);
-\draw (N2) -- (N33);
-\draw (N2) -- (N34);
-\draw (N3) to   [bend left= 10](H1) ;
-\draw (N31) to   [bend left= 10](H11) ;
-\draw (N32) to   [bend left= 10](H12) ;
-\draw (N33) to   [bend left= 10](H13) ;
-\draw (N34) to   [bend left= 10](H14) ;
-\draw (foot) to [bend left = 15]  (H0) ;
-
-% second copy 
-\node (N01) at (12,2) {$\bullet$};
-\node (N02) at (12,4) {$\bullet$};
-\node (N03) at (12,6) {{\color{lightred}$\bullet$}};
-\node (H001) at (10,8) {$\Smiley[1][vertfluo]$};
-\node (H0011) at (12,8) {$\Smiley[1][vertfluo]$};
-\node (H0012) at (14,8) {$\Smiley[1][vertfluo]$};
-\node (H0013) at (16,8) {$\Smiley[1][vertfluo]$};
-\node (H0014) at (18,8) {$\Smiley[1][vertfluo]$};
-
-\node (N03) at (11,6) {$\bullet$};
-\node (N031) at (12,6) {$\bullet$};
-\node (N032) at (13,6) {$\bullet$};
-\node (N033) at (14,6) {$\bullet$};
-\node (N034) at (15,6) {$\bullet$};
-
-\draw (foot) -- (N01)[very thick] ;
-\draw (N01) -- (N02);
-\draw (N02) -- (N03);
-\draw (N02) -- (N031);
-\draw (N02) -- (N032);
-\draw (N02) -- (N033);
-\draw (N02) -- (N034);
-\draw (N03) to   [bend left= 10](H001) ;
-\draw (N031) to   [bend left= 10](H0011) ;
-\draw (N032) to   [bend left= 10](H0012) ;
-\draw (N033) to   [bend left= 10](H0013) ;
-\draw (N034) to   [bend left= 10](H0014) ;
-
-% third copy 
-\node (N001) at (22,2) {$\bullet$};
-\node (N002) at (22,4) {$\bullet$};
-\node (N003) at (22,6) {{\color{lightred}$\bullet$}};
-\node (H001) at (20,8) {$\Smiley[1][vertfluo]$};
-\node (H0011) at (22,8) {$\Smiley[1][vertfluo]$};
-\node (H0012) at (24,8) {$\Smiley[1][vertfluo]$};
-\node (H0013) at (26,8) {$\Smiley[1][vertfluo]$};
-\node (H0014) at (28,8) {$\Smiley[1][vertfluo]$};
-
-\node (N003) at (21,6) {$\bullet$};
-\node (N0031) at (22,6) {$\bullet$};
-\node (N0032) at (23,6) {$\bullet$};
-\node (N0033) at (24,6) {$\bullet$};
-\node (N0034) at (25,6) {$\bullet$};
-
-\draw (foot) -- (N001)[very thick] ;
-\draw (N001) -- (N002);
-\draw (N002) -- (N003);
-\draw (N002) -- (N0031);
-\draw (N002) -- (N0032);
-\draw (N002) -- (N0033);
-\draw (N002) -- (N0034);
-\draw (N003) to   [bend left= 10](H001) ;
-\draw (N0031) to   [bend left= 10](H0011) ;
-\draw (N0032) to   [bend left= 10](H0012) ;
-\draw (N0033) to   [bend left= 10](H0013) ;
-\draw (N0034) to   [bend left= 10](H0014) ;
-\end{tikzpicture}
-\caption{The configuration \texttt{Hy'''} of \texttt{Hy} \label{fig:Hy5}}
-\end{figure}
-\FloatBarrier
-
-We leave it to the reader  to guess the following  rounds of the battle \dots
- % Please keep in mind that, in this 
-% the hydra is free to chose any number of replications at each time, whereas
-% Hercules chops only one head per round.
-
-% Let us precise that, in this game, Hercules wins if the hydra is eventually reduced 
-% to a single head. 
-% We know from~\cite{KP82} that, whichever the initial configuration of the
-% hydra, and the strategies of both players, Hercules eventually wins. The 
-% aforementionned paper shows also that there do not exist any \emph{simple} proof of this result.
-
-\section{Hydras and their representation in \emph{Coq}}
-\label{sec:orgheadline48}
-
-\index{hydras}{Library Hydra!Types!Hydra}
-\index{hydras}{Library Hydra!Types!Hydrae}
-
-In order to describe trees where each node can have an arbitrary (but finite) number of sons, it is usual to define a type where each node carries a \emph{forest}, \emph{i.e} a list of trees
-(see for instance Chapter 14, pages 400-406 of \cite{BC04}, also available as~\cite{BC04ch14}).
-
-For this purpose, we define two mutual \emph{ad-hoc}  inductive types, where \texttt{Hydra} is the main type, and \texttt{Hydrae} is a helper for describing finite sequences of hydras.
-\label{types:Hydra}
-\label{types:Hydrae}
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hydra}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/HydraDef}
-
-%\index{To do}
-\index{hydras}{Projects}
-
-\begin{project}
-Look for an existing library on trees with nodes of arbitrary arity, in order to replace  this ad-hoc type with something more generic.
-\end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-
- Another very similar representation could use the \texttt{list} type family instead of the specific 
-type \texttt{Hydrae}:
-
-\input{movies/snippets/Hydra_Definitions/HydraAlt}
-
-Using this representation, re-define all the constructions of this chapter.
-You will probably have to use patterns described for instance in~\cite{BC04} or the archives of the \coq communication channels (please consult~\url{https://coq.inria.fr/community.html}).
-
-\end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-The type \texttt{Hydra} above describes hydras as \emph{plane trees}, i.e., as drawn on a sheet of paper or computer screen. Thus, hydras are \emph{oriented},
-and it is appropriate to consider a \emph{leftmost} or \emph{rightmost} head of
-the beast. It could be interesting to consider another representation, in which
-every non-leaf node has a \emph{multi-set} -- not an ordered list -- of daughters.
-\end{project}
-
-\subsubsection{Abbreviations}
-
-We provide several notations for hydra patterns  which occur often in our developments.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#head}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/headsEtc}
-
-For instance, the hydra \texttt{Hy}  of Figure~\vref{fig:Hy} is defined in \emph{Gallina} as follows:
-
-\vspace{4mm}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html\#Hy}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/Hy}
-
-Hydras quite frequently contain  multiple adjacent  copies of the same subtree. The following functions
-will help us to describe and reason about replications in hydra battles.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#hcons_mult}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hconsMult}
-
-\vspace{4mm}
-
-For instance, the hydra \texttt{Hy''} of Fig~\vref{fig:Hy3}  can be defined in \coq{} as follows:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/HySecond}
-
-\subsubsection{Recursive functions on type \texttt{Hydra}}
-\label{sec:orgheadline41}
-\label{sec:hsize-def}
-
-In order to  define a recursive function over the type \texttt{Hydra}, one has to consider the three constructors 
-\texttt{node}, \texttt{hnil} and \texttt{hcons} of the mutually inductive types \texttt{Hydra} and \texttt{Hydrae}. 
-Let us define for instance the function which  computes the number of nodes of any hydra:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hsize}
-
-\input{movies/snippets/Hydra_Examples/HySize}
-
-Likewise, the \emph{height} (maximum distance between the foot and a head) 
-is defined by mutual recursion:
-
-\input{movies/snippets/Hydra_Definitions/height}
-
-\input{movies/snippets/Hydra_Examples/HyHeight}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Define a function \texttt{max\_degree: Hydra $\arrow$ nat} which  returns the highest degree of a node in any hydra.
-For instance, the evaluation of the term \texttt{(max\_degree Hy)} should return $3$.
-\end{exercise}
-
-\subsection{Induction principles for hydras}
-\label{sec:orgheadline42}
-
-In this section, we show how induction principles are used to prove properties on the type 
-\texttt{Hydra}. Let us consider for instance the following statement:
-\begin{quote}
-  `` The height of any hydra is strictly less than its size. ''
-\end{quote}
-
-\subsubsection{A failed attempt}
-
-One may try to use the default tactic of proof by induction, which corresponds to an application of the automatically  generated  induction principle for  type \texttt{Hydra}:
-
-\input{movies/snippets/Hydra_Examples/HydraInd}
-
-Let us start a simple proof by induction.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/BadInductiona}
-
-We might be tempted to do an induction on the sequence \texttt{s}:
-
-\input{movies/snippets/Hydra_Examples/BadInductionb}
-
-The first subgoal is trivial.
-
-\input{movies/snippets/Hydra_Examples/BadInductionc}
-
-Let us look at the second subgoal of the induction.
-
-\input{movies/snippets/Hydra_Examples/BadInductiond}
-
-We notice that this subgoal does not contain any hypothesis
-on the height and size of the hydra \texttt{h}. So, it looks hard to prove the conclusion. Let's stop.
-
-\input{movies/snippets/Hydra_Examples/BadInductione}
-
-\subsubsection{A Principle of mutual induction}
-In order to get an appropriate induction scheme for the types 
-\texttt{Hydra} and \texttt{Hydrae}, we can use  \coq{}'s  command \texttt{Scheme}.
-
-\index{coq}{Mutually inductive types}
-%\index{coq}{Commands!Scheme}
-
-\input{movies/snippets/Hydra_Definitions/HydraRect2}
-
-\input{movies/snippets/Hydra_Examples/HydraRect2Check}
-
-\subsubsection{A Correct proof}
-
-Let us now use \texttt{Hydra\_rect2} for proving that the height of any hydra is strictly less than its size.
-Using this scheme requires an auxiliary predicate, called \texttt{P0} in \texttt{Hydra\_rect2}'s statement.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hForall}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-\input{movies/snippets/Hydra_Examples/heightLtSizea}
-
-The first subgoal is easily solvable, using some arithmetic.
-The second and third ones are
-almost trivial. We let the reader look at the source. 
-\input{movies/snippets/Hydra_Examples/heightLtSizez}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-It happens very often that, in the proof of  a proposition of the form 
-(\texttt{$\forall\,$ h:Hydra, $P$ h}), the predicate \texttt{P0}
-is  (\texttt{h\_forall $P$}).
-Design a tactic for induction on hydras that frees the user from binding explicitly \texttt{P0},  and solves trivial subgoals.
-Apply it for writing  a shorter proof of \texttt{height\_lt\_size}.
-\end{exercise}
-
-\section{Relational description of hydra battles}
-
-In this section, we represent the rules of hydra battles as a binary relation associated with
-a \emph{round}, i.e., an interaction composed of the two following actions:
-\begin{enumerate}
-\item Hercules chops off one head of the hydra.
-\item Then, the  hydra replicates the wounded part (if the head is at distance $\geq 2$ from the foot).
-\end{enumerate}
-The relation associated with each round of the battle is parameterized  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot,
-but present for consistency's sake).
-
-In our description,  we will apply the following naming convention: if $h$ represents the configuration of the hydra before a round, then the configuration of $h$ after this round will be called $h'$.
- Thus, we are going to define a proposition  (\texttt{round\_n $n\;h\;h'$})  whose intended meaning will be `` the hydra $h$  is transformed into $h'$  in a single round of a battle, with the expected replication factor $n$ ''.
-
-Since the replication of parts of the hydra depends on the distance of the chopped head from  the foot, we  decompose our description into two main  cases, under the form of a bunch of [mutually] inductive predicates over the types \texttt{Hydra} and \texttt{Hydrae}.
-
-The mutually exclusive cases we consider are the following:
-\begin{itemize}
-\item \textbf{R1}: The chopped off head was at distance 1 from the foot.
-\item \textbf{R2}: The chopped off head was at a distance greater than or equal to  $2$ from the foot.
-\end{itemize}
-
-\subsection{Chopping off a head at distance 1 from the foot (relation  R1)}
-
-If Hercules chops off a head near the floor, there is no replication at all. We use an auxiliary 
-predicate \texttt{S0}, associated with the removing of one head from a sequence of hydras.
-
-\vspace{4pt}\emph{From Module\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/S0Def}
-\input{movies/snippets/Hydra_Definitions/R1Def}
-
-\subsubsection{Example}
-\label{sec:orgheadline45}
-
-Let us represent in \coq{}   the transformation of the hydra of Fig.~\vref{fig:Hy} into
-the configuration represented in Fig.~\ref{fig:Hy-prime}.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/Hy1}
-
-\subsection{Chopping off a head at distance \texorpdfstring{$\geq 2$}{>= 2} from the foot (relation R2) }
-
-Let us now consider beheadings  where the chopped-off head is at distance greater than or equal to $2$ from the foot. All the following relations are parameterized by the replication factor  $n$.
-
- Let $s$ be a sequence of hydras. 
-The proposition (\texttt{S1 n s s'}) holds if $s'$ is obtained by replacing some element $h$ of $s$ by 
-$n+1$ copies of $h'$, where  the proposition (\texttt{R1 h h'}) holds, in other words, $h'$ is just $h$, without the chopped-off  head.
-\texttt{S1} is an inductive relation with two constructors that allow us to choose the position in $s'$ of the wounded sub-hydra $h$.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#S1}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/S1Def}
-
-The rest of the definition is composed of two mutually inductive relations on hydras and sequences of hydras.
-The first constructor of \texttt{R2} describes the case where the chopped head is exactly at height $2$.
-The others constructors allow us to consider beheadings at height strictly greater than $2$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/R2Def}
-
-\subsubsection{Example}
-Let us prove the transformation of \texttt{Hy'} into \texttt{Hy''} (see Fig.~\vref{fig:Hy3}).
-We use an experimental set of tactics\footnote{See the \texttt{Ltac} definitions in~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions.}}  for specifying the place where the
-interaction between Hercules and the hydra holds.
-
-\vspace{4pt}\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}.
-
-\input{movies/snippets/Hydra_Examples/R2Example}
-
-The reader is encouraged to look at all the successive subgoals of this example.
-\emph{Please consider also exercise~\vref{exo:interactive-battle}.}
-
-\subsection{Binary relation associated with a round}
-
-We combine the two cases above into a single relation.
-First,  we define the  relation \texttt{(round\_n n h h')} where \texttt{n} is the expected number of  replications (irrelevant in the case of an \texttt{R1}-transformation).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#round_n}{Hydra.Hydra\_Definitions}}
-
-\index{hydras}{Library Hydra!Predicates!round\_n}
-
-\input{movies/snippets/Hydra_Definitions/roundNDef}
-
-By abstraction over \texttt{n}, we define a \emph{round} (small step) of a battle:
-
-\index{hydras}{Library Hydra!Predicates!round}
-\label{sect:infix-round}
-
-\input{movies/snippets/Hydra_Definitions/roundDef}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Give a direct translation of Kirby and Paris's description of hydra battles (quoted on page~\pageref{original-rules}) and prove that our relational description is consistent with theirs.
-\end{project}
-
-\subsection{Rounds and battles}
-
-Using library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}, we define \texttt{round\_plus},  the transitive closure of \texttt{round}, and \texttt{round\_star},  the reflexive and transitive closure of \texttt{round}.
-
-\label{sect:infix-rounds}
-
-\input{movies/snippets/Hydra_Definitions/roundPlus}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-  Prove that if \texttt{$h$ -+-> $h'$}, then
-  the height of $h'$ is less or equal than the height of $h$.
-
-\end{exercise}
-
-\begin{remark}
-\label{remark:transitive-closure}
-\coq's library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Coq.Relations.Relation\_Operators} 
-contains three logically equivalent definitions of the transitive closure of a binary relation. This equivalence is proved in 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} .
-
-Why three definitions for a single mathematical concept?
-Each definition generates an associated induction principle. 
- According to the form of statement one would like to prove, there is a ``best choice'':
-
-\begin{itemize}
-\item For proving $\forall y, x\,R^+\,y \;\arrow\; P\,y$, prefer 
-\texttt{clos\_trans\_n1}
-\item For proving $\forall x,\,x\,R^+\,y \;\arrow\; P\,x$, prefer \texttt{clos\_trans\_1n}
-\item For proving $\forall x\,y, \,x\,R^+\,y \;\arrow\;P\,x\,y$,  
-prefer \texttt{clos\_trans},
-\end{itemize}
-But there is no ``wrong choice'' at all: the equivalence lemmas in \linebreak 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} 
- allow the user
-to convert any one of the three closures into another one before applying the corresponding elimination tactic.
-The same remark also holds for reflexive and transitive closures. 
-\end{remark}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Define a restriction of \coqsimple{round},  where Hercules always chops off
-the leftmost among the lowest heads.
-
-Prove that, if $h$ is not a simple head, then there exists a unique $h'$ such that $h$  is transformed into $h'$ in one round, according to this restriction.
-
-\end{exercise}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}[Interactive battles]
-\label{exo:interactive-battle}
-Given a hydra \texttt{h}, the specification of a hydra battle for \texttt{h} is the type 
-\Verb@{h':Hydra | h -*-> h'}@. In order to avoid long sequences of \texttt{split}, \texttt{left}, and 
-\texttt{right}, design a set of dedicated tactics for the interactive building of a battle.
-Your tactics will have the following functionalities:
-\begin{itemize}
-\item  Chose to stop a battle, or continue
-\item Chose an expected number of replications
-\item Navigate in a hydra, looking for a head to chop off.
-\end{itemize}
-
-Use your tactics for simulating a small part of a hydra battle, for instance the rounds which lead from
-\texttt{Hy} to \texttt{Hy'''}  (Fig.~\vref{fig:Hy5}).
-
-\textbf{Hints:} 
-\begin{itemize}
-
-\item Please keep in mind that the last  configuration of your interactively built battle is known only at the end of the battle.
-Thus, you will have to create and solve subgoals with existential variables.
-For that purpose, the tactic \texttt{eexists}, applied to the
-goal \Verb@{h':Hydra | h -*-> h'}@ generates the subgoal \Verb|h -*-> ?h'|.
-\item You may use Gérard Huet's \emph{zipper} data structure~\cite{zipper} for writing tactics associated with Hercules's  interactive search for a head to chop off.
-\end{itemize}
-
-\end{exercise}
-
-\subsection{Classes of battles}
-\label{sect:battle-classes}
-
-In some presentations of hydra battles, e.g.~\cite{KP82, bauer2008}, the transformation associated with the $i$-th round may depend on $i$.
-For instance, in these articles, the replication factor at the $i$-th round is equal to $i$.
-In other examples, one can allow the hydra to apply any replication factor at any time.
-In order to be the most general as possible, we define the type of predicates which relate the state of the hydra before and after the $i$-th round of a battle.
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-\label{types:Battle}
-\index{hydras}{Library Hydra!Type classes!Battle}
-
-\input{movies/snippets/Hydra_Definitions/BattleDef}
-
-The most general class of battles is \texttt{free}, which allows the hydra to chose any replication factor at every step:
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#free}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/freeDef}
-
-We chose to call \emph{standard}\footnote{This appellation is ours.
-If there is a better one, we will change it.} the kind of battles which appear  most often in the literature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#standard}{Hydra.Hydra\_Definitions}}
-\input{movies/snippets/Hydra_Definitions/standardDef}
-
-\subsection{Big steps}
-
-Let $B$ be any instance of class \texttt{Battle}. It is easy to define inductively the relation between the $i$-th and the $j$-th steps of a battle of type $B$.
-
-\vspace{6pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/battleRelDef}
-
-The following property allows us to build battles by composition of smaller ones.
-
-%% TODO display subgoals when fixed
-
-\vspace{6pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-\noindent
-\input{movies/snippets/Hydra_Lemmas/battleTrans}
-
-% \begin{remark}
-%  The class \texttt{free} is strongly related with the transitive closure  \texttt{round\_plus}, as expressed by the following lemmas.
-
-% \vspace{4pt}
-% \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-
-%  \begin{Coqsrc}
-%  Lemma battle_free_equiv1 : forall i j h h',  
-%              battle free i h j h' ->   h -+-> h'.
-
-%  Lemma battle_free_equiv2 : forall h h',
-%      h -+-> h' ->
-%     forall i, exists j,  battle free i h j h'.
-%  \end{Coqsrc}
-
-% \end{remark}
-
-\section{A long battle}
-\label{sect:big-battle}
-
-In this section we consider a simple example of battle, starting with a small hydra,
-shown on figure~\vref{fig:hinit}, with a simple strategy for both players:
-
-\begin{itemize}
-\item At each round, Hercules chops off the rightmost head of the hydra.
-\item The battle is {standard}: at the round number $i$, the expected replication is $i$.
-\end{itemize}
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (6,0) {$\bullet$};
-\node (n1) at  (3,3) {$\bullet$};
-\node (h1) at  (1,6) {$\Smiley[1][green]$};
-\node (h2) at  (3,6) {$\Smiley[1][green]$};
-\node (h3) at  (6,6) {$\Smiley[1][green]$};
-\node (h4) at  (6,6) {$\Smiley[1][green]$};
-\node (h5) at  (6,3) {$\Smiley[1][green]$};
-\node (h6) at  (9,3) {$\Smiley[1][green]$};
-\draw (foot) -- (n1);
-\draw (n1) to   [bend left=20] (h1);
-\draw (n1) to   (h2);
-\draw (n1) to   [bend right=20] (h3);
-\draw (foot) -- (h5);
-\draw (foot) to  [bend right=20] (h6);
-\end{tikzpicture}
-
-  \caption{The hydra \texttt{hinit}}
-  \label{fig:hinit}
-\end{figure}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.BigBattle.html}{Hydra.BigBattle}}
-
-\input{movies/snippets/BigBattle/hinitDef}
-
-The lemma we would like to prove is ``The considered battle lasts exactly $N$ rounds'',
-with $N$ being a natural number we gave to guess.
-
-But the  battle is so long that no \emph{test} can give us any estimation of its length, and we do need the expressive power of logic to compute this length.
-However, in order to  guess this length, we made some experiments, computing with \gallina{}, \coq{}'s  functional programming language.
-Thus, we can consider this development as a collaboration of proof with computation.
-In the following lines, we show how we found experimentally the value of the number $N$.
-The complete proof is in file \url{../theories/html/hydras.Hydra.BigBattle.html}.
-
-\subsection{First rounds}
-During the two first rounds, our hydra loses its two rightmost heads.  Figure~\vref{fig:hinit-plus2} shows the state of the hydra   just before the third round.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (3,0) {$\bullet$};
-\node (n1) at  (3,3) {$\bullet$};
-\node (h1) at  (1,6) {$\Smiley[1][green]$};
-\node (h2) at  (3,6) {$\Smiley[1][green]$};
-\node (h3) at  (6,6) {$\Smiley[1][green]$};
-\node (h4) at  (6,6) {$\Smiley[1][green]$};
-\draw (foot) -- (n1);
-\draw (n1) to   [bend left=20] (h1);
-\draw (n1) to   (h2);
-\draw (n1) to   [bend right=20] (h3);
-\end{tikzpicture}
-
-  \caption{The hydra (\texttt{hyd1 h3})}
-  \label{fig:hinit-plus2}
-\end{figure}
-
-The following lemma  is a formal description of these first rounds, in terms of the
-\texttt{battle} predicate.
-
-\input{movies/snippets/BigBattle/L02}
-
-\subsection{Looking for regularities}
-
-A first study with pencil and paper suggested us that, after three rounds, the hydra always looks like in figure~\vref{fig:hinit-plusn} (with a variable number of 
-subtrees of height 1 or 0).
-Thus, we introduce a few handy abbreviations.
-
-\input{movies/snippets/BigBattle/Notations}
-
-For instance, the hydra shown in  Fig~\ref{fig:hinit-plusn} is  (\texttt{hyd 3 4 2}). The hydra (\texttt{hyd 0 0 0})  is the ``final'' hydra of any terminating battle, {i.e.},
-a tree whith exactly one node and no edge.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (15,0) {$\bullet$};
-\node (a) at  (3,4) {$\bullet$};
-\node (b) at  (6,4) {$\bullet$};
-\node (c) at  (9,4) {$\bullet$};
-\node (d) at  (13,4) {$\bullet$};
-\node (e) at  (16,4) {$\bullet$};
-\node (f) at  (19,4) {$\bullet$};
-\node (g) at  (22,4) {$\bullet$};
-\node (h) at  (25,4) {$\Smiley[1][green]$};
-\node (i) at  (28,4) {$\Smiley[1][green]$};
-\node (aa) at  (2.5,8) {$\Smiley[1][green]$};
-\node (ab) at  (3.5,8) {$\Smiley[1][green]$};
-\node (ba) at  (5.5,8) {$\Smiley[1][green]$};
-\node (bb) at  (6.5,8) {$\Smiley[1][green]$};
-\node (ca) at  (8.5,8) {$\Smiley[1][green]$};
-\node (cb) at  (9.5,8) {$\Smiley[1][green]$};
-\node (da) at  (13,8) {$\Smiley[1][green]$};
-\node (ea) at  (16,8) {$\Smiley[1][green]$};
-\node (fa) at  (19,8) {$\Smiley[1][green]$};
-\node (ga) at  (22,8) {$\Smiley[1][green]$};
-\draw (foot) -- (a);
-\draw (foot) -- (b);
-\draw (foot) -- (c);
-\draw (foot) -- (d);
-\draw (foot) -- (e);
-\draw (foot) -- (f);
-\draw (foot) -- (g);
-\draw (foot) -- (h);
-\draw (foot) -- (i);
-\draw (a) -- (aa);
-\draw (a) -- (ab);
-\draw (b) -- (ba);
-\draw (b) -- (bb);
-\draw (c) -- (ca);
-\draw (c) -- (cb);
-\draw (d) -- (da);
-\draw (e) -- (ea);
-\draw (f) -- (fa);
-\draw (g) -- (ga);
-\end{tikzpicture}
-
-  \caption{The hydra (\texttt{hyd 3 4 2})}
-  \label{fig:hinit-plusn}
-\end{figure}
-
-With these notations, we get a formal description of the first three rounds.
-
-\input{movies/snippets/BigBattle/L23L03}
-
-\subsection{Testing  \dots}
-\label{sect:testing}
-In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple data type for representing the states as tuples composed of
-the round number, and the respective number of daughters  \texttt{h2}, \texttt{h1}, and heads
-of the current hydra.
-
-\input{movies/snippets/BigBattle/stateDef}
-
-The following function returns the next configuration of the game.
-Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle only start with the lemma
-\texttt{step\_battle}, page~\pageref{lemma:step-battle}.
-
-\input{movies/snippets/BigBattle/nextDef}
-
-We can make bigger steps through iterations of \texttt{next}.
-The functional \texttt{iterate}, similar to Standard Library's \texttt{Nat.iter},
-is defined and studied in~\href{../theories/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
-\index{hydras}{Library Prelude!iterate}
-
-\label{Functions:iterate}
-
-\input{movies/snippets/Iterates/iterateDef}
-
-The following function computes the state of the battle at the $n$-th round.
-
-\input{movies/snippets/BigBattle/testDefTests}
-
-The battle we are studying looks to be awfully long. Let us concentrate our
-tests on some particular events : the states where $\texttt{nh}=0$.
-From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} is equal to zero.
-
-\input{movies/snippets/BigBattle/smartTest}
-
-Thus, $(1 + 11)$ rounds later, the \texttt{n1} field is equal to $2$, and 
-\texttt{nh}   to $0$.
-
-\input{movies/snippets/BigBattle/smartTestb}
-
-Next round, we decrement \texttt{n2} and set \texttt{n1} to $95$.
-
-\input{movies/snippets/BigBattle/smartTestc}
-
-We now have some intuition of the sequence.
-It looks like the next ``\texttt{nh}=0'' event will happen at the $192=2(95+1)$-th round, then at the $2(192+1)$-th round, etc.
-
-\input{movies/snippets/BigBattle/doubleS}
-
-\subsection{Proving \dots}
-We are now able to reason about the sequence of transitions defined by our hydra battle. Instead of using the data-type \texttt{state} we study the relationship
-between different configurations of the battle.
-
-Let us define a binary relation associated with every round of the battle.
-In the following definition \texttt{i} is associated with the round number (or date, if we consider a discrete time), and \texttt{a}, \texttt{b}, \texttt{c} respectively associated with the number of occurrences of \texttt{h2}, \texttt{h1} and heads connected to the hydra's foot.
-
-\input{movies/snippets/BigBattle/oneStep}
-
-The relation between \texttt{one\_step} and the rules of hydra battles is asserted by the following lemma.
-
-\label{lemma:step-battle}
-
-\input{movies/snippets/BigBattle/stepBattle}
-
-\vspace{4pt}
-
-Next, we define ``big steps'' as the transitive closure of \texttt{one\_step},
-and reachability (from the initial configuration of figure~\ref{fig:hinit} at time $0$).
-
-\input{movies/snippets/BigBattle/steps}
-
-The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{battle}.
-
-\input{movies/snippets/BigBattle/stepsBattle}
-
-\vspace{4pt}
-
-Thus, any result about \texttt{steps} will be applicable to standard battles.
-Using the predicate \texttt{steps},  our study of the length of the considered battle
-can  be decomposed into three parts:
-
-\begin{enumerate}
-\item  Characterization and proofs of regularities of some events (inspired by our experiments of Sect.~\ref{sect:testing}).
-\item Study of the beginning of the battle
-\item Computing the exact length of the battle.
-\end{enumerate}
-
-First, we prove that, if at round $i$ the hydra is equal to
-(\texttt{hyd a (S b) 0}), then it will be equal to (\texttt{hyd a b 0}) at the $2(i+1)$-th round.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/LS}
-
-From now on, the lemma \texttt{reachable\_S} allows us to watch larger and larger steps of 
-the battle.
-
-\input{movies/snippets/BigBattle/L4}
-
-\input{movies/snippets/BigBattle/L10To95}
-
-\subsection{Giant steps}
-
-We are now able to make bigger steps in the simulation of the battle.
-First, we iterate the lemma \texttt{reachable\_S}.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/Bigstep}
-
-\vspace{4pt}
-
-Applying lemmas \texttt{BigStep} and \texttt{L95} we make a first jump.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/MDef}
-
-Figure~\ref{fig:HM}  represents the hydra at the $M$-th round.
-At the $(M+1)$-th round, it will look like in fig~\ref{fig:HM-plus1}.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:HM}}
-The state of the hydra after $M$ rounds.
-% The hydra \texttt{h} of the proof that \(\omega^2\) is too small for proving Hercules' victory
-
-\end{figure}
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (0,5) {$\bullet$};
-\node (N12) at (0,8) {$\Smiley[2][green]$};
-\node (N2) at (2,5) {$\bullet$};
-\node (N22) at (2,8) {$\Smiley[2][green]$};
-\node (N3) at (4,5) {$\bullet$};
-\node (N32) at (4,8) {$\Smiley[2][green]$};
-\node (N4) at (6,5) {$\bullet$};
-\node (N42) at (6,8) {$\Smiley[2][green]$};
-
-\node (Ndots) at (12,8) {\Huge $\dots$};
-\node (Ndots2) at (12,5) {\Huge $\dots$};
-
-\node (N8) at (18,5) {$\bullet$};
-\node (N82) at (18,8) {$\Smiley[2][green]$};
-\node (N9) at (20,5) {$\bullet$};
-\node (N92) at (20,8) {$\Smiley[2][green]$};
-
-\draw (foot) -- (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N8);
-\draw (foot) -- (N9);
-\draw (N1) to  (N12);
-\draw (N2) to  (N22);
-\draw (N3) to  (N32);
-\draw (N4) to  (N42);
-\draw (N8) to  (N82);
-\draw (N9) to  (N92);
-\end{tikzpicture}
-\caption{\label{fig:HM-plus1}}
-The state of the hydra after $M+1$ rounds (with $M+1$ heads).
-
-\end{figure}
-
-\input{movies/snippets/BigBattle/L295S}
-
-\vspace{4pt}
-
-Then, applying once more the lemma \texttt{BigStep}, we get the exact time when
-Hercules wins!
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/NDef}
-
-\vspace{4pt}
-
-We are now able to prove formally that the considered battle is 
-composed of $N$ steps.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/Done}
-
-\subsection{A minoration lemma}
-
-Now, we would like to get an intuition of  how big the number $N$ is.
-For that purpose, we use a minoration of the function \texttt{doubleS} by the
-function (\texttt{fun n => 2 * n}).
-
-\vspace{4pt}
-
-\input{movies/snippets/Exp2/exp2Def}
-
-\vspace{4pt}
-
-Using a few facts (proven in 
-\href{../theories/html/hydras.Hydra.BigBattle.html}{hydras.Hydra.BigBattle}),we get several  minorations.
-
-\input{movies/snippets/BigBattle/minorationLemmas}
-
-\vspace{4pt}
-
-The number $N$ is greater than or  equal to $2^{2^{95}\times 95}.$ If we wrote $N$ in base $10$, $N$ would require at least $10^{30}$ digits!
-
-\section{Generic properties}
-
-The example we just studied shows that the termination of any battle may take a very long time. If we want to study hydra battles in general, we have to consider 
-any hydra and any strategy, both for Hercules and the hydra itself. So, we first  give some definitions, generally borrowed from transition systems vocabulary (see~\cite{tel_2000} for instance).
-
-\subsection{Liveliness}
-
-Let $B$ be an instance of \texttt{Battle}. We say that $B$ is \emph{alive} if
-for any configuration $(i,h)$, where $h$ is not a head, there exists a further step in class $B$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/AliveDef}
-
-The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
-\href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
-
-\input{movies/snippets/Hydra_Theorems/AliveThms}
-
-Both theorems are proved with the help of the  following strongly specified function:
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
-
-\input{movies/snippets/Hydra_Lemmas/nextRoundDec}
-
-\subsection{Termination}
-
-The termination of \emph{any}  battle is naturally expressed by the predicate \texttt{well\_founded} defined in the module \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}
- of the Standard Library.
-
-\index{hydras}{Library Hydra!Predicates!Termination}
-
-\input{movies/snippets/Hydra_Definitions/TerminationDef}
-
-Let $B$ be an instance of class \texttt{Battle}. A \emph{variant} for $B$ consists
-in a well-founded relation $<$  on some type \texttt{A}, and a function
-(also called a \emph{measure}) $m:\texttt{Hydra}\rightarrow A$ such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $B$, the inequality $m(h')<m(h)$ holds.
-
-\vspace{4pt}
-\noindent
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
-
-\label{sect:hvariant-def}
-
-\index{hydras}{Library Hydra!Type classes!Hvariant}
-
-\input{movies/snippets/Hydra_Definitions/HvariantDef}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
- Prove that, if there exists some  instance of (\texttt{Hvariant Lt wf\_Lt $B$ $m$}), then there exists no infinite battle in  $B$.
-\end{exercise}
-
-\subsection{A  small proof of impossibility}
-%\index{coq}{Proofs of impossibility}
-
-\label{omega-case}
-
-When one wants to prove a termination theorem with the help of a variant, 
-one has to consider first a well-founded set $(A,<)$, then a strictly decreasing measure on this set.
-The following two lemmas show that if  the order structure $(A,<)$ is too simple, it is useless to look for a convenient measure, which simply no exists.
-Such kind of result is useful, because it saves you time and effort.
-
-The best known well-founded order is the natural order on the set $\mathbb{N}$ of natural numbers (the type \texttt{nat} of Standard library).
-It would be interesting to look for some measure $m:\texttt{nat}\arrow\texttt{nat}$ and prove it is a variant.
-
-Unfortunately, we can prove that 
-\emph{no} instance of class (\texttt{WfVariant round Peano.lt $m$}) can be built, where
-$m$ is \emph{any} function of type \texttt{Hydra $\arrow$ nat}.
-
-Let us present the main steps of that proof, the script of which  is in the module ~\href{../theories/html/hydras.Hydra.Omega_Small.html}{Hydra/Omega\_Small.v} \footnote{ The name of this file means ``the ordinal $\omega$ is too small for proving the termination of [free] hydra battles ''.
-In effect, the elements of $\omega$, considered as a set, are just the natural numbers (see next chapter for more details)}.
-
-%\subsubsection{Preliminaries}
-
-Let us assume there exists some variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
-    the  termination of all hydra battles.
-
-\input{movies/snippets/Omega_Small/omegaSmalla}
-
-We define an injection $\iota$ from the type \texttt{nat} into \texttt{Hydra}.
-For any natural number $i$, $\iota(i)$ is the hydra composed of a foot and
-$i+1$ heads at height $1$. For instance, Fig.~\ref{fig:flower} represents the hydra $\iota(3)$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (4,0) {$\bullet$};
-\node (N1) at (2,2) {$\Smiley[2][green]$};
-\node (N2) at (4,2) {$\Smiley[2][green]$};
-\node (N3) at (6,2) {$\Smiley[2][green]$};
-\node (N4) at (8,2) {$\Smiley[2][green]$};
-\draw (foot) to [bend left =25] (N1);
-\draw (foot) to [bend left =15] (N2);
-\draw (foot) to [bend right =15] (N3);
-\draw (foot) to [bend right =25] (N4);
-\end{tikzpicture}
-\caption{\label{fig:flower}
-The hydra $\iota(3)$}
-\end{figure}
-
-\input{movies/snippets/Omega_Small/iotaDef}
-
-Let us consider now some hydra \texttt{big\_h} out of the range of the injection $\iota$ (see Fig.~\vref{fig:h-omega-omega}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to  (N2);
-\end{tikzpicture}
-\caption{\label{fig:h-omega-omega}}
- The hydra \texttt{big\_h}.
-\end{figure}
-
-\input{movies/snippets/Omega_Small/bigHDef}
-
- Using the functions $m$ and $\iota$, we define a second hydra \texttt{small\_h}, and show
- there is a one-round battle that transforms \texttt{big\_h} into \texttt{small\_h}. Please note that,
-due to the hypothesis \texttt{Hvar}, we are interested in the termination of \emph{free} battles. 
-There is no problem to consider a round with (\texttt{m big\_h}) as the replication factor.
-
-\input{movies/snippets/Omega_Small/smallHDef}
-
-But, by hypothesis, $m$ is a variant. Hence, we infer the following inequality.
-
-\vspace{4pt}
-
-\input{movies/snippets/Omega_Small/mLt}
-
-In order to get a contradiction, it suffices to  prove the inequality
-$m(\texttt{big\_h}) \leq m(\texttt{small\_h})$ i.e.,  $m(\texttt{big\_h}\leq m(\iota (m(\texttt{big\_h})))$.
-
-\input{movies/snippets/Omega_Small/mGea}
-
-Intuitively, it means that, from any hydra of the form (\texttt{iota $i$}), the battle will 
-take (at least) $i$ rounds. Thus the associated measure cannot be less than $i$.
-Technically, we prove this lemma by Peano induction on $i$.
-
-\begin{itemize}
-\item The base case $i=0$ is trivial
-\item Otherwise, let $i$ be any natural number and assume  the inequality
-  $i \leq m(\iota(i))$.
-  \begin{enumerate}
-  \item  But the hydra $\iota(S(i))$ can be transformed in one round into
-    $\iota(i)$ (by losing its rightmost head, for instance)
-  \item Since $m$ is a variant, we have $m(\iota(i)) < m(\iota(S(i)))$,
-    hence  $i< m(\iota(S(i)))$, which implies  $S(i)\leq  m(\iota(S(i)))$.
-  \end{enumerate}
-\end{itemize}
-
-We are now ready to complete our impossibility proof.
-
-\vspace{4pt}
-
-\inputsnippets{Omega_Small/mGeb, Omega_Small/mGez,
-  Omega_Small/omegaSmallz}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\subsubsection{Conclusion}
-
-In order to build a variant for proving the termination of all hydra battles, we need to consider order structures more complex than the usual order on type \texttt{nat}. 
-The notion of \emph{ordinal number} provides a catalogue of well-founded order types.
-For a reasonably large bunch of ordinal numbers, \emph{ordinal notations} are data-types which allow the \coq{} user to define functions, to compute and prove some properties, for instance by reflection.
-
-The next chapter is dedicated to a generic formalization of ordinal notations, and chapter~\ref{chap:T1} to a proof of termination of all hydra battles with the help of an ordinal notation for the interval $[0,\epsilon_0)$\,\footnote{We use the mathematical notation $[a,b)$ for the interval $\{x|a\leq x < b\}$.}.
-\index{maths}{Notations!Interval}
-
-%--------------------------------------------------------------
-
-\chapter{Introduction to ordinal numbers and ordinal notations}
-\label{chap:ON}
-
-The proof of termination of all hydra battles presented in~\cite{KP82} is based
-on \emph{ordinal numbers}.
-From a mathematical point of view, an ordinal is a representative of an equivalence class for isomorphisms of  totally ordered well-founded sets.
-
-For the computer scientist, ordinals are tools for proving the totality of a given recursive function, or termination of a transition system. \emph{Ordinal arithmetic} 
-provides a set of functions whose properties, like \emph{monotony}, allow to define \emph{variants}, \emph{i.e.} strictly decreasing measures used in proofs of termination.
-
-\vspace{4pt}
-
-Let us have a look at Figure~\ref{fig:ordinal-sequence}. It presents a few items of a  sequence of ordinal numbers, which extends the sequence of natural numbers.
-
-\begin{figure}[h]
-  \centering
-\fbox{\Large
-  \begin{minipage}{1.0\linewidth}
-  \begin{align*}
-     &\textcolor{blue}{0},\,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,\ldots\\
-&\textcolor{red}{\omega},\,\omega+1,\omega+2,\omega+3,\ldots\\
-&\textcolor{red}{\omega\times 2},\,\omega\times 2+1,\ldots, \textcolor{red}{\omega\times 3},\,\omega\times 3+1,\ldots, \textcolor{red}{\omega\times 4},\ldots\\
-&\textcolor{red}{\omega^2},\ldots, \textcolor{red}{\omega^2\times 42},\ldots,\textcolor{red}{\omega^3},\ldots, \textcolor{red}{\omega^4},\omega^4+1,\ldots\\
-&\textcolor{red}{\omega^\omega},\ldots, \textcolor{red}{\omega^\omega+\omega^7\times 8},\ldots,\textcolor{red}{\omega^\omega\times 2},\omega^\omega\times 2+1, \ldots\\
-&\textcolor{red}{\omega^{\omega^\omega}},\ldots, \textcolor{red}{\omega^{\omega^\omega}+\omega^\omega\times 42+ \omega^{55}+\omega}, \ldots, \textcolor{red}{\omega^{\omega^{\omega+1}}}, \omega^{\omega^{\omega+1}}+1,\dots\\
-& \textcolor{red}{\epsilon_0 (= \omega^{\epsilon_0)}}, \epsilon_0+1, \epsilon_0+2, \epsilon_0+3, \ldots\\
-& \textcolor{red}{\epsilon_1}, \ldots, \textcolor{red}{\epsilon_2}, \ldots, \textcolor{red}{\epsilon_\omega},\ldots \\
-& \textcolor{red}{\Gamma_0}, \Gamma_0+1, \Gamma_0+2, \Gamma_0+3,\ldots, \textcolor{red}{\Gamma_0+\omega}, \ldots\\
-&\ldots
-  \end{align*}   
-  \end{minipage}}
-
-  \caption{A short overview of the sequence of ordinal numbers}
-  \label{fig:ordinal-sequence}
-\end{figure}
-
-Let us comment some features of this figure:
-
-\begin{itemize}
-\item The ordinals are listed in a strictly increasing order. 
-\item Dots : ``$\ldots$'' stand for  infinite sequences of ordinals, not shown for lack of space.
-For instance, the ordinal $42$ is not shown in the first line, but it exists, somewhere between $17$ and $\omega$.
-\item Each ordinal printed in black is the immediate successor of another ordinal.
-We call it a \emph{successor} ordinal.
-For instance, $12$ is the successor of $11$, and $\omega^4+1$ the successor of $\omega^4$.
-\item Ordinals (displayed in red)  that  follow immediately dots are called \emph{limit ordinals}.
-With respect to the order induced by this sequence, any limit ordinal $\alpha$ is the least upper bound of  the set $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
-\item
-For instance $\omega$ is the least upper bound of the set of all finite ordinals (in the first line). It is also the first limit ordinal, and the first infinite ordinal, in the sense that 
-the set $\mathbb{O}_\omega$ is infinite.
-\item The ordinal $\epsilon_0$ is the first number which is equal to its own exponential of base $\omega$.
-It plays an important role in proof theory, and is particularly studied in chapters~\ref{chap:T1} to \ref{chap:alpha-large}.
-\item Any ordinal is  either the ordinal \textcolor{blue}{$0$},
-a successor ordinal, or a \textcolor{red}{limit ordinal}.
-\end{itemize}
-
-\section{The mathematical point of view}
-
-\subsection{Well-ordered sets}
-Let us start with some definitions.
-A  \emph{well-ordered set} is a set provided with a binary relation $<$ which has the following properties.
-\begin{description}
-\item[irreflexivity] : $\forall x\in A, x\not< x$
-\item[transitivity] : $\forall x\,y\,z\in A, x<y \Rightarrow y<z \Rightarrow x<z$
-\item[trichotomy]: $\forall x\,y\in A, x<y \vee x = y \vee y < x$
-\item[well foundedness]: $<$ is well-founded (every element of $A$ is accessible)\footnote{In classical mathematics, we would say that there is no infinite sequence $a_1>a_2> \dots a_n> a_{n+1}\dots$ in $A$.
-In contrast, \coq's standard library contains
-an inductive definition of a predicate \texttt{Acc} which allows us to write 
-constructive proofs of accessibility (See \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}).}.
-\end{description}
-
-The best known examples of well-ordered sets are the set $\mathbb{N}$ of natural numbers (with the usual order $<$), as well as any finite segment $[0,i)=\{j\in\mathbb{N}\,|\,j<i\}$.
-The disjoint union of two copies of $\mathbb{N}$, \emph{i.e.} the set $\{0,1\}\times\mathbb{N}$ is also well-ordered,
-with respect to the order below:
-
-\begin{align*}
-(i,j) < (i,k) & \;\textit{\textbf{if} }\; j < k\\
-(0,k) < (1,l) & \;\textit{\textbf{for\,any}}\;k \;\textit{\textbf{and}} \; l
-\end{align*}
-
-\subsection{Ordinal numbers}
-
-\index{maths}{Ordinal numbers}
-
-Let $(A,<_A)$ and $(B,<_B)$ two well-ordered sets. $A$ and $B$ are said to have \emph{the same order type} if 
-there exists a strictly monotonous bijection $b$ from $A$ to $B$, \emph{i.e.} which verifies the proposition
-$\forall x\,y\in A,\, x <_A y \Rightarrow b(x) <_B  b(y)$.
-
-Having the same order type is an equivalence relation between well-ordered sets. Ordinal numbers (in short: \emph{ordinals})  are descriptions (\emph{names}) of the equivalence classes.
-For instance, the order type of $(\mathbb{N},<)$ is associated with the ordinal called  $\omega$, and the order we considered on 
-the disjoint union of $\mathbb{N}$ and itself is named $\omega+\omega$.
-
-In a set-theoretic framework, one can consider any ordinal $\alpha$ as a well-ordered set, whose  elements are just the ordinals strictly less than $\alpha$, \emph{i.e.} the \emph{segment} $\mathbb{O}_\alpha=[0, \alpha)$.
-So, one can speak about \emph{finite}, \emph{infinite}, \emph{countable}, etc., ordinals.
-Nevertheless, since we work within type theory,
-we do not identify ordinals as sets of ordinals, but consider the correspondence between ordinals and sets of ordinals as the function that maps $\alpha$ to $\mathbb{O}_\alpha$.
-For instance $\mathbb{O}_\omega=\mathbb{N}$, and $\mathbb{O}_7=\{0,1,2,3,4,5,6\}$.
-
-We cannot cite all the literature published on ordinals since Cantor's book
-\cite{cantorbook}, and 
-leave it to the reader to explore the bibliography.
-
-The introduction of Jos\'e Grimm's report~\cite{grimm:hal-00911710} contains a nice presentation of the main properties of ordinals.
-
-\section{Ordinal numbers in Coq}
-
-Two kinds of representation of ordinals are defined in our development.
-
-\begin{itemize}
-\item A ``mathematical'' representation of the set of countable ordinal numbers, after Kurt Schütte~\cite{schutte}.
-This representation uses several (hopefully harmless) axioms.
-We use it as a reference for proving the correctness of ordinal notations.
-  We are also progressively importing \gaia's~\cite{Gaia} definitions and theorems, based on Bourbaki's set theory~\cite{Grimm2016}.
-Chapter~\ref{gaia-chapter} is dedicated to the connexion between libraries \HydrasLib and \gaia.
-
-\item A family of \emph{ordinal notations} (also called \emph{[ordinal] notation systems}), \emph{i.e.} data types used to represent segments $[0,\mu)$, where $\mu$ is some countable ordinal.
-Each ordinal notation is defined inside the Calculus of Inductive Constructions (without axioms).
-Many functions are defined, allowing proofs by computation.
-Note that proofs of
-correctness of a given ordinal notation with respect to Schütte's model obviously use axioms.
-Please execute the \texttt{Print Assumptions} command in case of doubt.
-\end{itemize}
-
-% \section{Countable ordinals}
-
-% Chapter~\ref{chap:schutte} of this document presents an adaptation to \coq{} of an axiomatization in classical logic of the set of countable ordinals by K. Schütte~\cite{schutte}. 
-% That formalization is quite complex, technical and unshamedly non-constructive,  so we put its description  in the last chapter of this document.
-
-% Please note that Schütte considers the (uncountable) set $\mathbb{O}$ of all countable ordinals.
-This set is well ordered (which is one of Schütte's axioms), and associates to any ordinal $\alpha$ the segment $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
-
-% In our adaptation to \coq{}, we declare a type \texttt{Ord}, a binary relation \texttt{lt} (with infix notation \texttt{"\_<\_"}, and assume Schütte's axiom. In Chapter~\ref{chap:schutte},
-% we derive some interesting properties of countable ordinals from these axioms.
-
-It is interesting to compare proofs of a given property (for instance the associativity of addition) both in the computational framework of some ordinal notation, and in mathematical  models
-of Schütte or \gaia.
-
-\section{Ordinal Notations}
-
-Fortunately, the ordinals we need for  studying hydra battles are much simpler than Schütte's, and can be represented as quite simple data types in \gallina.
-
-Let $\alpha$ be some (countable) ordinal; 
-in \coq{} terms, we call \emph{ordinal notation for $\alpha$} a structure composed 
-of:
-\begin{itemize}
-\item A data type $A$ for representing all ordinals strictly below $\alpha$,
-\item A well founded order $<$ on $A$, 
-\item A correct function for comparing two ordinals. Note  that the reflexive closure of $<$ is thus a \emph{total order}.
-\end{itemize}
-
-Such a structure may be proved correct relatively to a bigger ordinal notation or to Schütte's or \gaia's model.
-
-\subsection{Classes for ordinal notations}
-
-From the \coq{} user's point of view, an ordinal notation is
-a structure allowing to compare two ordinals by computation, and proving by well-founded induction.
-
-\subsubsection{The \texttt{Comparison} and \texttt{Comparable} classes}
-
-We use the operational class \texttt{Comparison} of comparison functions to define the \texttt{Comparable} class, contributed by Jérémy Damour and Théo Zimmermann, which allows us to apply generic lemmas and tactics about decidable strict orders.
-The correctness of the comparison function is expressed through Stdlib's type 
-\texttt{Datatypes.CompareSpec} as specialized by \texttt{Datatypes.CompSpec}.
-
-\begin{Coqsrc}
-  Inductive CompareSpec (Peq Plt Pgt : Prop) :
-  comparison -> Prop :=
-    CompEq : Peq -> CompareSpec Peq Plt Pgt Eq
-  | CompLt : Plt -> CompareSpec Peq Plt Pgt Lt
-  | CompGt : Pgt -> CompareSpec Peq Plt Pgt Gt.
-
-Definition CompSpec {A} (eq lt : A -> A -> Prop) (x y : A) :=
- CompareSpec (eq x y) (lt x y) (lt y x).
-\end{Coqsrc}
-
-\emph{From Module~\href{../theories/html/hydras.Prelude/mparable.html\#Hvariant}{Prelude.Comparable}}
-
-\label{sect:comparable-def}
-
-\input{movies/snippets/Comparable/ComparableDef}
-
-\subsubsection{The \texttt{ON} class}
-
-The following class definition, parameterized with a type $A$, a binary relation \texttt{lt} on $A$, specifies that \texttt{lt} is a well-founded strict order, provided with a correct comparison function.
-
-\vspace{4pt}
-\noindent\emph{From
-Library~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}}
-
-\label{types:ON}
-\index{hydras}{Library OrdinalNotations!Type classes!ON}
-
-\input{movies/snippets/ON_Generic/ONDef}
-
-We give  also a few handy definitions and lemmas for any ordinal notation.
-
-\label{sect:on-lt-notation}
-\label{sect:on-le-notation}
-\label{sect:measure-ON}
-\label{sect:bigO-ON}
-
-\inputsnippets{ON_Generic/ONDefsa}
-\inputsnippets{ON_Generic/ONDefsb}
-
-\begin{remark}
-The infix notations \texttt{o<} and \texttt{o<=} are defined in order to make apparent the distinction between the various notation scopes that may co-exist in a same statement.
-So the infix \texttt{<} and \texttt{<=} are reserved to the natural numbers.
-In the mathematical formulas, however, we still use $<$ and $\leq$ for comparing ordinals.
-\end{remark}
-
-% \subsection{Ordinal notations and  termination measures}
-% \label{sect:measure-ON}
-
-% The following lemma (together with the type class mechnism) allows us to define termination measures over any ordinal notation.
-It is just an application of  the libraries \texttt{Coq.Wellfounded.Inverse\_Image}
-% and  \texttt{Coq.Wellfounded.Inclusion}.
-
-% \begin{Coqsrc}
-% Definition measure_lt {A:Type}{lt: relation A}
-%             {compare : A -> A -> comparison}
-%             {on : ON lt compare}
-%             {B : Type} (m : B -> A) : relation B :=
-%              fun x y => on_lt (m x) (m y).
-
-% Lemma wf_measure  {A:Type}(lt: relation A)
-%             {compare : A -> A -> comparison}
-%             {on : ON lt compare}
-%             {B : Type}
-%             (m : B -> A):  well_founded (measure_lt m). 
-% \end{Coqsrc}
-
-% A simple example of application is given in Sect.~\vref{sect:merge-example}.
-
-\section{Example: the ordinal \texorpdfstring{$\omega$}{omega}}
-
-The simplest example of ordinal notation is built over the type \texttt{nat} of \coq's standard library. We just have to apply already proven lemmas about Peano numbers.
-
-\vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_Omega.html}{OrdinalNotations.ON\_Omega}}
-
-\inputsnippets{ON_Omega/OmegaDefa, ON_Omega/OmegaDefb}
-
-\section{Sum of  two ordinal notations}
-
-Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective types \texttt{A} and \texttt{B}.
-
- We consider a new strict order
-on the disjoint sum of the associated types, by putting all elements of \texttt{A} before the elements of \texttt{B} (thanks to Standard Library's relation operator \texttt{le\_AsB}).
-
-\vspace{4pt}
-\noindent
-\emph{From Library~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}}.
-
-\begin{Coqanswer}
-Inductive
-le_AsB (A B : Type) (leA : A -> A -> Prop) (leB : B -> B -> Prop)
-  : A + B -> A + B -> Prop :=
-| le_aa : forall x y : A, leA x y -> le_AsB A B leA leB (inl x) (inl y)
-| le_ab : forall (x : A) (y : B), le_AsB A B leA leB (inl x) (inr y)
-| le_bb : forall x y : B, leB x y -> le_AsB A B leA leB (inr x) (inr y)
-\end{Coqanswer}
-
-%\pagebreak
-
-\vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_plus.html}{OrdinalNotations.ON\_plus}}
-
-\input{movies/snippets/ON_plus/Defs}
-
-In order to build an instance of \texttt{Comparable}, we have to define a correct comparison function.
-
-\inputsnippets{ON_plus/compareDef, ON_plus/compareCorrect}
-\inputsnippets{ON_plus/plusComp}
-
-Standard library's lemma \texttt{Wellfounded.Disjoint\_Union.wf\_disjoint\_sum} 
-is applied  to prove that our order \texttt{lt} is well-founded, allowing us to build an instance of \texttt{ON}:
-
-\inputsnippets{ON_plus/ltWf, ON_plus/OnPlus}
-
-\subsection{The ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
-
-The ordinal $\omega+\omega$ (also known as $\omega\times 2$) may be represented as the concatenation 
-of two copies of $\omega$ (Figure~\ref{fig:omega-plus-omega}).
-It is also represented by the two first lines of Figure~\ref{fig:ordinal-sequence}.
-
-\begin{figure}[h]
-   \centering
-   \begin{tikzpicture}[very thick, scale=0.5]
-\begin{scope}[color=blue]
-\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
-\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
-\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
-\node (Adots) at (6,0) {$\ldots$};
-\node(An) at (8,0)[label=below:$n$]{$\bullet$};
-\node(A2) at (10,0)[label=below:$n+1$]{$\bullet$};
-\node (Adots1) at (12,0) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red]
-\node(B0) at (14,0)[label=below:$0$,label=above:\textcolor{red}{$\omega$}]{$\bullet$};
-\node(B1) at (16,0)[label=below:$1$, label=above:$\omega+1$]{$\bullet$};
-\node(B2) at (18,0)[label=below:$2$,label=above:$\omega+2$]{$\bullet$};
-\node (Bdots) at (20,0) {$\ldots$};
-\node (Bn) at (22,0) [label=below:$p$, label=above:$\omega+p$]{$\bullet$};
-\node (Bdots2) at (24,0) {$\ldots$};
-\end{scope}
-\end{tikzpicture}
-   \caption{\textcolor{blue}{$\omega+{\color{red}\omega}$}}
-   \label{fig:omega-plus-omega}
- \end{figure}
-
-We can define this notation in \coq{} as an instance of \texttt{ON\_plus}.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/OmegaPlusOmegaDef}
-
-\vspace{4pt}
-
-We can now define abbreviations.
-For instance, the finite ordinals are represented by terms built with  the constructor \texttt{inl}, and the first infinite ordinal $\omega$ by the term \texttt{(inr 0)}.
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Omega_plus_omega/finiteOmega}
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Omega_plus_omega/ltOmega}
-
-% \label{warning:coercions}
-% \index{Coq!Coercions} 
-% \begin{remark}
-% Beware of coercions and notation scopes!
-% Let us consider the following goal:
-
-% \begin{Coqsrc}
-%  Goal (6 o< 8).
-%  auto with arith.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% 1 subgoal (ID 9)
-
-%   ============================
-%   6 o< 8
-% \end{Coqanswer}
-
-% Please keep in mind that the current notation scope interprets the infix \texttt{``<''} as the predicate \texttt{Omega\_plus\_omega.lt} and not \texttt{Nat.lt}.
-More,  the coercion mechanism converts the terms \texttt{6:nat} [resp.
-\texttt{8:nat} ]
-% into \texttt{inl 6} [resp. \texttt{inl 8}].  So, the initial goal is correctly interpreted by \coq{}, but not as an inequality between two natural numbers.
-
-% \begin{Coqsrc}
-% Set Printing All.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% 1 subgoal (ID 337)
-
-%   ============================
-%   @on_lt nat Peano.lt Nat.compare Omega (S (S (S (S (S (S O))))))
-%     (S (S (S (S (S (S (S (S O))))))))
-% \end{Coqanswer}
-
-% Anyway, the initial goal is provable, using \texttt{le\_AsB}'s first constructor.
-
-% \begin{Coqsrc}
-%   constructor; auto with arith.
-% Qed.
-% \end{Coqsrc}
-
-% \end{remark}
-%
-
-\section{Limits and successors}
-
-Let us look again at our implementation of $\omega+\omega$. We can classify its elements into three categories:
-
-\begin{itemize}
-\item The least ordinal, \texttt{(inl 0)}, also known as  \texttt{(fin 0)}.
-\item The first infinite ordinal $\omega$.
-\item The remaining ordinals, either of the form \texttt{(inl (S $i$))} or \texttt{(inr (S $i$))} (in black on Figure~\ref{fig:ordinal-sequence}), called \emph{successor ordinals}.
-\end{itemize}
-
-\subsection{Definitions}
-It would be interesting to specify at the most generic level, what is a zero, a successor or a limit ordinal. Let $<$ be a strict order on a type $A$.
-
-\begin{itemize}
-\item A \emph{least} element is a minorant (in the large sense) of the full set  on $A$,
-\item $y$ is a \emph{successor} of $x$ if $x<y$ and there is no element between $x$ and $y$. We will also say that $x$ is a \emph{predecessor} of $y$.
-\item $x$ is a \emph{limit} if $x$ is not a least element, and for any $y$ such that $yo<x$,
- there exists some $z$ such that $y<z<x$.
-\end{itemize}
-
-The following definitions are in Library \href{../theories/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
-
-\input{movies/snippets/MoreOrders/Defs}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove, that, in any ordinal notation system, every ordinal has at most one predecessor, and at most one successor.
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/predSuccUnicity.v}{exercises/ordinals/predSuccUnicity.v}.}
-
-\end{exercise}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove, that, in any ordinal notation system, if $\beta$ is a successor of $\alpha$,
-then for any $\gamma$, $\gamma<\beta$ implies 
-$\gamma\leq\alpha$.
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/lt_succ_le.v}{exercises/ordinals/lt\_succ\_le.v}.}
-\end{exercise}
-
-\subsection{Limits and successors in \texorpdfstring{$\omega+\omega$}{omega+omega}}
-
-Using the definitions above, we can prove the following lemma:
-
-\vspace{4pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/limitIff}
-
-\vspace{4pt}
-
-Regarding successors, let us define the following function and prove its correctness:
-
-\input{movies/snippets/ON_Omega_plus_omega/succDef}
-
-\input{movies/snippets/ON_Omega_plus_omega/succCorrect}
-
-\vspace{4pt}
-
-We can also check whether an ordinal is a successor by a simple computation:
-
-\input{movies/snippets/ON_Omega_plus_omega/succb}
-
-\vspace{4pt}
-
-Finally, the nature of any ordinal is decidable (inside this notation system) :
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/ZeroLimitSuccDec}
-
-\section{Product of ordinal notations}
-
-Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective  ordered types \texttt{A} and \texttt{B}.
-The product of \texttt{NA} and \texttt{NB} is considered as the concatenation of $B$ copies of $A$, ordered by the lexicographic order on $B\times A$.
-
-In \coq{}, we build an instance of class \texttt{ON} through a sequence of steps as for the sum of ordinal notations.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_mult.html}{OrdinalNotations.ON\_mult}}
-
-\inputsnippets{ON_mult/Defs}
-\inputsnippets{ON_mult/multComp,  ON_mult/ONMult}
-\inputsnippets{ON_mult/endDefs}
-
-\section{The ordinal \texorpdfstring{$\omega^2$}{omega2}}
-
-The ordinal $\omega^2$ (also called $\phi_0(2)$, see Chap.~\ref{chap:schutte}), is an instance of the multiplication presented in the preceding section.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
-
-\inputsnippets{ON_Omega2/Omega2Def, ON_Omega2/Defs}
-\inputsnippets{ON_Omega2/Ex1}
-
-\subsection{Arithmetic of \texorpdfstring{$\omega^2$}{omega^2}}
-
-\subsubsection{Successor}
-
-The successor of any ordinal is defined by a simple pattern-matching.
-
-\input{movies/snippets/ON_Omega2/succ}
-
-This function is proved to be correct w.r.t. the \texttt{Successor} predicate.
-
-\input{movies/snippets/ON_Omega2/succOK}
-
-\input{movies/snippets/ON_Omega2/succLemmas}
-
-\subsubsection{Addition}
-
-We can define on \texttt{Omega2} an addition which extends the addition on \texttt{nat}. Please note that this operation is not commutative:
-
-\inputsnippets{ON_Omega2/plusDef, ON_Omega2/plusExamples}
-
-\subsubsection{Multiplication}
-
-The restriction of ordinal multiplication to the segment $[0,\omega^2)$ is not a total function.
-For instance $\omega\times\omega= \omega^2$ is outside the set of represented values.
-Nevertheless, we can define two operations mixing natural numbers and ordinals.
-
-\input{movies/snippets/ON_Omega2/multFinDef}
-\input{movies/snippets/ON_Omega2/multFinExamples}
-
-Multiplication with a finite ordinal and addition are related through the following lemma:
-
-\inputsnippets{ON_Omega2/uniqueDecompositiona,
-  ON_Omega2/uniqueDecompositionb,
-  ON_Omega2/uniqueDecompositionz}
-
-\subsection{A proof of termination using \texorpdfstring{$\omega^2$}{omega^2}} 
-\label{sect:merge-example}
-
-Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily a total function which merges two lists (example contributed by Pascal Manoury).
-
-%\index{coq}{Commands!Function}
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
-
-\inputsnippets{ON_Omega2/Merge, ON_Omega2/Mergeb}
-
-\subsection{Yet another  proof of impossibility}
-\label{omega2-case}
-
-In Sect.~\vref{omega-case}, we proved that there exists no variant from \texttt{Hydra} to \texttt{(nat,$<$)}
-(\emph{i.e.} the ordinal $\omega$) for proving the termination of all hydra battles.
-We  prove now that  the ordinal $\omega^2$ is also insufficient for this purpose.
-
-The proof we are going to comment has exactly the same structure as in Section~\ref{omega-case}.
- Nevertheless, the proof of technical  lemmas is a little more complex, due to 
- the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
-Consider for instance that there exists an infinite number of ordinals  between
-$\omega$ and $\omega\times 2$.
-
-The detailed  proof script is in the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/Hydra/Omega2_Small.v}{theories/ordinals/Hydra/Omega2\_Small.v}.
-
-\subsubsection{Preliminaries}
-Let us assume there is a variant from \texttt{Hydra} into $\omega^2$  for proving the   termination of all hydra battles.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
-
-\input{movies/snippets/Omega2_Small/Impossibility}
-\input{movies/snippets/Omega2_Small/Impossibilitya}
-
-We  follow the same pattern as in Sect.~\ref{omega-case}.
-First, we define an injection $\iota$ from type \texttt{ON\_Omega2.t} into \texttt{Hydra}, by
- associating to  each ordinal $\omega\times i+ j = (i,j)$ the hydra with $i$ branches of length $2$ and
-$j$ branches of length $1$.
-
-%% revenir ici
-
-\vspace{4pt}
-\emph{From Module ~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
-
-\input{movies/snippets/Omega2_Small/Impossibilityc}
-
-For instance, Figure~\vref{fig:essai2} shows the hydra associated to the ordinal 
-$(3,5)$, a.k.a. $\omega\times 3 + 5$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-\node (foot) at (6,0) {$\bullet$};
-\node (N1) at (1,3) {$\bullet$};
-\node (N2) at (3,3) {$\bullet$};
-\node (N3) at (5,3) {$\bullet$};
-\node (N4) at (8,3) {$\Smiley[2][green]$};
-\node (N5) at (11,3) {$\Smiley[2][green]$};
-\node (N6) at (14,3) {$\Smiley[2][green]$};
-\node (N7) at (17,3){$\Smiley[2][green]$};
-\node (N8) at (20,3){$\Smiley[2][green]$};
-\node  (N9) at (0,5) {$\Smiley[2][green]$};
-\node (N10) at (2,5) {$\Smiley[2][green]$};
-\node (N11) at (4,5) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N5);
-\draw (foot) -- (N6);
-\draw (foot) to [bend right=10] (N7);
-\draw (foot) to [bend right=15] (N8);
-\draw (N1) to [bend left=10] (N9);
-\draw (N2) -- (N10);
-\draw (N3) -- (N11);
-\end{tikzpicture}
-\caption{\label{fig:essai2}
-The hydra $\iota(\omega\times 3+5)$}
-\end{figure}
-
-Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:h-omega2-small}}
- The hydra \texttt{big\_h}.
-\end{figure}
-
-\input{movies/snippets/Omega2_Small/Impossibilityb}
-
- In a second step, we build a ``smaller'' hydra\footnote{With respect to the measure $m$.}.
-
-\input{movies/snippets/Omega2_Small/Impossibilityd}
-
-\vspace{4pt}
-
-Like in Sect.~\ref{omega-case}, we prove the inequality $m\;\texttt{big\_h} \,\texttt{o<=}\, m\;\texttt{small\_h} \,\texttt{o<=}\, m\;\texttt{big\_h}$, which is obviously false.
-
-\subsubsection{Proof of the inequality \texttt{m small\_h o< m big\_h}}
-
-In order to prove the inequality  \texttt{m\_lt: m small\_h o< m big\_h}, it suffices to
-build a battle transforming \texttt{big\_h} into \texttt{small\_h}.
-
-First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h} as $(i,j)$.
-If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
-If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the following result.
-
-\input{movies/snippets/Omega2_Small/bigToSmall}
-
-Since $m$ is a variant, we infer the following inequality:
-
-\input{movies/snippets/Omega2_Small/mLt}
-
-\subsubsection{Proof of the inequality \texttt{m big\_h o<= m small\_h} }
-
-The proof of the inequality \texttt{m big\_h o<= m small\_h} is quite more complex than in Sect~\ref{omega-case}.
-If we consider any ordinal $\alpha=(i,j)$, where $i>0$, there exists an infinite number of
-ordinals strictly less than $\alpha$, and there exists an infinite number of battles that start from
-$\iota(\alpha)$.
-Indeed, at any configuration $\iota(k,0)$, where $k>0$, the hydra can freely choose any replication number.
-Intuitively, the measure of such a hydra must be large enough for taking into account
-all the possible battles issued from that hydra.
-Let us now give more technical details.
-
-The first steps of our proof start a well-founded induction over $\omega^2$.
-
-\input{movies/snippets/Omega2_Small/mGe}
-
-A case analysis on $i$ and $j$ allows us to
-consider three cases :
-
-\begin{itemize}
-\item $i=j=0$: the inequality is trivial.
-\item $i=1+l, j=0$ ($(i,j)$ is a limit ordinal): By the induction hypothesis \texttt{IHij},
-  $(l,k)\texttt{o<=} m(\iota(l,k))$ for any $k$. But (by the rules of the hydra game), $\iota(i,0)$ is transformed into any $\iota(l,k)$ in one round. Thus $m(\iota(l,k)) < m(\iota(i,0))$ for any $k$.
-  Therefore, $(l,k) <  m(\iota(i,0))$ for any $k$, thus
-  $(i,0)\,\texttt{o<=}\,m(\iota(i,0))$.
- \item $j= l+1$  ($(i,j)$ is a successor). We apply the induction hypothesis to the pair $(i,l)$.
- \end{itemize}
-
- Please look at the proof script for more details.
-
- \input{movies/snippets/Omega2_Small/mGeb}
-
-\subsubsection{End of the proof}
-From \texttt{m\_ge}, we get \texttt{$m$\, big\_h o<= $m$\, small\_h = $m$\,(iota (m big\_h)) }. 
-Since $<$ is a strict order (irreflexive  and transitive), this inequality is incompatible with the strict inequality  \texttt{m small\_h o< m big\_h} (lemma \texttt{m\_lt}).
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#Impossible}{Hydra.Omega2\_Small}}
-
-  \input{movies/snippets/Omega2_Small/Impossible}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into $\omega^2$ for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\begin{remark}
- We  prove in Chapter~\ref{ks-chapter} a generalization of the impossibility lemmas of
-Sect.~\ref{omega-case} and this section, with the same proof structure, but with much more 
-complex technical details.
- \end{remark}
-
-% \index{Exercises}
-% \begin{exercise}
-
-% \label{sec:orgheadline63}
-% Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.
-
-% \begin{itemize}
-% \item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
-%   lexicographic ordering
-% \item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.
-
-% For instance, the following inequality holds:
-% \[\langle 4,3,3,3,3,3,3,2,2,2 \rangle\,<\,\langle 4,4,2 \rangle\]
-% \end{itemize}
-
-% \end{exercise}
-
-\section{A notation for finite ordinals}
-
-Let $n$ be some natural number. The segment associated with $n$ is the interval 
-$[0,n)\,=\,\{0,1,\dots,n-1\}$. 
-One may represent the ordinal $n$ by a sigma type.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
-
-\label{def:Finite-ord-type}
-
-\input{movies/snippets/ON_Finite/Defs}
-
-\subsubsection{Examples}
-
-For instance, let us build two elements of the segment $[0, 7)$, \emph{i.e.} two
-inhabitants of   type (\texttt{t 7}), and prove a simple  inequality (see Fig.~\ref{fig:O7}).
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (N0) at (0,0) {$\bullet$};
-\node (i0) at (0,1) {$0$};
-\node (N1) at (2,0) {$\bullet$};
-\node (i1) at (2,1) {$1$};
-\node (N2) at (4,0) {$\bullet$};
-\node (i2) at (4,1) {$2$};
-\node (N3) at (6,0) {$\bullet$};
-\node (i3) at (6,1) {$3$};
-\node (N4) at (8,0) {$\bullet$};
-\node (i4) at (8,1) {$4$};
-\node (N5) at (10,0) {$\bullet$};
-\node (i5) at (10,1) {$5$};
-\node (N6) at (12,0) {$\bullet$};
-\node (i6) at (12,1) {$6$};
-\node(alpha1) at (4,-1) {$\alpha_1$};
-\node(alpha2) at (10,-1) {$\beta_1$};
-\end{tikzpicture}
-
-\caption{The segment $\mathbb{O}_7$\label{fig:O7}}
-\end{figure}
-
-%\index{coq}{Commands!Program}
-
-\input{movies/snippets/ON_Finite/Example1}
-
-Note that the type (\texttt{t 0}) is empty, and that, for any natural number
- $n$, $n$ does not belong to (\texttt{t $n$}).
-
- \inputsnippets{ON_Finite/t0Empty, ON_Finite/bad10}
-
-\subsubsection{Comparison function}
-
-In order to build an instance of \texttt{ON}, we define a comparison function,  and prove its correctiness.
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Finite/compareDef}
-
-\begin{remark}
- The proof of \texttt{compare\_correct} uses a well-known pattern of \coq{}.
-Let us consider  the following subgoal.
-
-\input{movies/snippets/ON_Finite/compareCorrectb}
-
-Applying the tactic \texttt{f\_equal} generates a simpler subgoal.
-
-\input{movies/snippets/ON_Finite/compareCorrectc}
-
-We have now to prove that there exists at most one  proof of (\texttt{Nat.ltb x0 (S n)}). This is not obvious, but  a consequence of the following lemma of library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Eqdep_dec.html}{Coq.Logic.Eqdep\_dec}.
-
-\index{Coq}{Unicity of equality proofs}
-\label{sect:eq-proof-unicity}
-
-\begin{Coqanswer}
-eq_proofs_unicity_on :
-forall (A : Type) (x : A),
-(forall y : A, x = y \/ x <> y) -> 
-forall (y : A) (p1 p2 : x = y), p1 = p2
-\end{Coqanswer}
-
-Thus unicity of proofs of \texttt{Nat.ltb x0 (S n)}  comes from the decidability of
-equality on type \texttt{bool}.
-This is why we used the boolean function \texttt{Nat.ltb} instead of the inductive predicate \texttt{Nat.lt} in the definition of type \texttt{t $n$} (see page~\pageref{def: Finite-ord-type}).
-For more information about this pattern, please look at the numerous mailing lists and 
-FAQs on \coq{}).
-
-\inputsnippets{ON_Finite/compareCorrectd}
-
-\end{remark}
-
-\begin{remark}
-
-Please note  that attempting to compare a term  of type (\texttt{t $n$}) with a term of
-type (\texttt{t $p$})  leads to an error if $n$ and $p$ are not convertible.
-
-\input{movies/snippets/ON_Finite/Example2}
-
-\end{remark}
-
-\subsubsection{Building an instance of \texttt{ON}}
-
-Applying lemmas of the libraries \texttt{Coq.Wellfounded.Inverse\_Image}, \linebreak
- \texttt{Coq.Wellfounded.Inclusion}, and \texttt{Coq.Arith.Wf\_nat}, we prove that our
-relation \texttt{lt} is well founded.
-
-\input{movies/snippets/ON_Finite/ltWf}
-
-Now we can build our instance of \texttt{OrdinalNotation}.
-
-\input{movies/snippets/ON_Finite/ONInstance}
-
-\begin{remark}
-It is important to keep in mind  that the integer $n$ is not an ``element'' of \texttt{FinOrd $n$}.
-In set-theoretic presentations of ordinals, the set associated with the ordinal $n$ is $\{0,1,\dots,n-1\}$.
-In our formalization, the interpretation of an ordinal as a set is realized by the function \texttt{bigO} (see Section\vref{sect:bigO-ON}).
-\end{remark}
-
-\begin{remark}
- There is no interesting arithmetic on finite ordinals, since functions like successor, addition, etc.,  cannot be represented in \coq{} as \emph{total} functions.
-\end{remark}
-
-\section{See also \dots}
-
-Finite ordinals are also formalized in \ssreflect/\mathcomp~\cite{MCB}.
-In Module~\href{../theories/html/gaia_hydras.ON_gfinite.html}{gaia\_hydras.ON\_gfinite}, we build an instance of class \texttt{ON}.
-
-\inputsnippets{ON_gfinite/finordLtDef, ON_gfinite/finordON}
-
-\inputsnippets{ON_gfinite/Examples, ON_gfinite/Examplesb}
-
-  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.
-
-%%%
-
-\section{Comparing two ordinal notations}
-
-It is sometimes useful to compare two ordinal notations with respect to expressive power
-(the segment of ordinals  they represent).
-
-The following class specifies a strict inclusion of segments.
-The notation \texttt{OA} describes a segment $[0,\alpha)$, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented in \texttt{OA}).
-We require also  that the comparison functions of the two notation systems are compatible.
-
-\begin{figure}[h]
-   \centering
-   \begin{tikzpicture}[very thick, scale=0.6]
-\begin{scope}[color=blue]
-\node (A) at (0,0) {$A$};
-\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
-\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
-\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
-\node (Adots) at (6,0) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red]
-\node (B) at (0,2) {$B$};
-\node(B0) at (2,2)[label=above:$0$]{$\bullet$};
-\node(B1) at (3,2)[label=above:$1$]{$\bullet$};
-\node(B2) at (4,2)[label=above:$2$]{$\bullet$};
-\node (Bdots) at (6,2) {$\ldots$};
-\node (b) at (8,2) [label=above:$\omega$]{$\bullet$};
-\node (bsucc) at (9,2) [label=above:$\omega+1$]{$\bullet$};
-\node (Bdots2) at (10,2) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red!50!blue]
-\draw [->,thin] (A0) -- node [auto] {$\iota$} (B0);
-\draw [->,thin] (A1) -- node [auto] {$\iota$} (B1);
-\draw [->,thin] (A2) -- node [auto] {$\iota$} (B2);
-\draw [->,thin] (Adots) -- node [auto] {$\iota$} (Bdots);
-\end{scope}
-\end{tikzpicture}
-   \caption{\textcolor{blue}{$A$} is a sub-segment  of \textcolor{red}{$B$}}
-   \label{fig:subsegment}
- \end{figure}
-
-If \texttt{OB} is presumed to be correct, then we may consider that \texttt{OA} ``inherits'' its correctness from the bigger notation system \texttt{OB}.
-
-\label{types:SubON}
-\index{hydras}{Library OrdinalNotations!Type classes!SubON}
-
-following definition
-(in ~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{ON\_Generic}).
-
-\input{movies/snippets/ON_Generic/SubONDef}
-
-For instance, we prove that \texttt{Omega} is a sub-notation of
-\texttt{Omega\_plus\_Omega} (with $\omega$ as the first ``new'' ordinal, and \texttt{fin} as the injection).
-
-\vspace{10pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/Incl}
-
-We can also show that, if $i<j$, then the segment $[0,i)$ is a ``sub-segment'' of
-$[0,j)$. Since the terms  ($t\;i$) and ($t\;j$) are not convertible, we consider a ``cast'' 
-function $\iota$ from ($t\;i$) into ($t\;j$), and prove that this function is  a monotonous bijection  from ($t\;i$) to
-the segment $[0,i)$ of ($t\;j$).
-
-%\index{coq}{Commands!Program}
-
-\vspace{10pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
-
-\inputsnippets{ON_Finite/InclIJ, ON_Finite/InclIJa,
-  ON_Finite/InclIJb,ON_Finite/InclIJc, ON_Finite/InclIJd,
-ON_Finite/InclIJz}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove that \texttt{Omega\_plus\_Omega} cannot be a sub-notation of \texttt{Omega}.
-\end{exercise}
-
-\index{hydras}{Projects}
-\begin{project}
-Adapt the definition of \texttt{Hvariant} (Sect.~\ref{sect:hvariant-def}) in order to
-have an ordinal notation as argument. Prove that if $O_A$ is a sub-notation of $O_B$, then any variant defined on  $O_A$ can be automatically transformed into 
-a variant on $O_B$.
-\end{project}
-
-\section{Comparing an ordinal notation with Schütte's model}
-
-Finally, it may be interesting to compare an ordinal notation with the more theoretical model from Schütte (well, at least with our formalization of that model).
-This would be a relative proof of correctness of the considered  ordinal  notation.
-
-The following class specifies that a notation \texttt{OA} describes a segment $[0,\alpha)$,
-where $\alpha$ is a countable ordinal \emph{à la}  Schütte.
-
-\label{types:ON-for}
-\index{hydras}{Library OrdinalNotations!Type classes!ON\_correct}
-
-\input{movies/snippets/ON_Generic/ONCorrect}
-
-For instance, the following theorem tells that \texttt{Epsilon0}, our notation system for the segment $[0,\epsilon0)$ is a correct implementation of the theoretically defined  ordinal $\epsilon_0$
-(see chapter~\ref{chap:schutte} for more details).
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
-
-\input{movies/snippets/Correctness_E0/Epsilon0Correct}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Same work, but replace Schütte's model with \gaia's.
-\end{project}
-
-\begin{project}
-  When you have read Chapter~\ref{chap:schutte}, prove that the sum of two ordinal notations \texttt{ON\_plus} implements the addition of ordinals.
-\end{project}
-
-\section{Isomorphism of ordinal notations}
-
-In some cases we want to show that two notation systems describe the same segment (for instance $[0,3+\omega)$ and $[0,\omega)$\;).
-For this purpose, one may prove that the two notation systems are order-isomorphic.
-
-\index{hydras}{Library OrdinalNotations!Type classes!ON\_Iso}
-
-\label{types:ON-iso}
-
-\input{movies/snippets/ON_Generic/ONIso}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-\label{exo:i-plus-omega}
-Let $i$ be some natural number. Prove that the notation systems 
-\texttt{Omega} and (\texttt{ON\_plus (OrdFin $i$) Omega}) are isomorphic.
-
-{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$, that we prove also in larger notation systems, as well as in Schütte's model.}
-This exercise is partially solved for $i=3$ (in ~\href{../theories/html/hydras.OrdinalNotations.Example_3PlusOmega.html}{OrdinalNotations.Example\_3PlusOmega}).
-
-\end{exercise}
-
-\index{hydras}{Projects}
-\label{exo:ON-mult}
-\begin{project}
-% Define in \coq{} the product of two ordinal notations $N_A$ and $N_B$.
-% If $A$ [resp. $B$] is the underlying type of $N_A$ [resp. $N_B$], the
-% product \texttt{ON\_mult $N_A$ $N_B$} is implemented over the cartesian product $B\times A$ (with the lexicographic ordering).
-
-This exercise is about the non-commutativity of the multiplication of ordinals, reflected in ordinal notations.
-
-For instance, the
-elements of the product (\texttt{ON\_mult Omega (FinOrd 3)}) are ordered as follows.
-\[(0,0),(0,1),(0,2),(0,3),(0,4),\dots,{\color{red}(1,0),} (1,1),(1,2),\dots, {\color{red}(2,0)},(2,1),(2,2),\dots\]
-
-Note that the elements of  (\texttt{ON\_mult (FinOrd 3) Omega}) are differently ordered (without limit ordinals):
-\[(0,0),(1,0),(2,0),(0,1),(1,1),(2,1),(0,2),(1,2),(2,2),(0,3),\dots\]
-
-Prove formally  that (\texttt{ON\_mult (FinOrd $i$) Omega}) is isomorphic to
-\texttt{Omega}  whilst
-\texttt{Omega}  is a sub-notation of (\texttt{ON\_mult Omega (FinOrd $i$)}),
-for any strictly positive $i$.
-
-\textbf{Note:} Like Exercise~\ref{exo:i-plus-omega}, this project corresponds to the [in]equalities $i+\omega=\omega<\omega+i$, for any natural number $i$.
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-Consider two isomorphic ordinal notations \texttt{OA} and \texttt{OB}.
-Prove that, if \texttt{OA} [resp. \texttt{OB}] is a correct implementation 
-of $\alpha$ [resp. $\beta$], then $\alpha=\beta$.
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-\label{project:succ-limit-dec}
-Add to the class \texttt{ON} the requirement that for any $\alpha$ it is decidable whether $\alpha$ is $0$, a successor or a limit ordinal.
-
-\textbf{Hint:}   Beware of the instances associated with sum and product of notations!
-  You may consider additional fields 
-to make the sum and product of notations ``compositional''.
-
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-\label{project:on-setoid}
-Reconsider the  class \texttt{ON}, with an equivalence instead of Leibniz equality.
-\end{project}
-
-%%%% ICI ICI
-
-\section{Other ordinal notations}
-%%% TODO : Fix the multiplication function in branch FixOmegaOmega
-
-% \index{hydras}{Projects}
-
-% \begin{project}
-% The directory \texttt{theories/OmegaOmega} contains an ad-hoc formalization of $\omega^\omega$, contributed by Pascal Manoury.
-Every ordinal $\alpha$ is represented by a list $l$ whose elements are the coefficients of $\omega$ in  the Cantor normal form of $\alpha$ (in reverse order).
-For instance, the ordinal
-% $\omega^{8}\times 5 + \omega^{6}\times 8 + \omega^2\times 10 + \omega + 7$ is represented by the list \texttt{[5;0;8;0;0.0;10,1,7]}.
-
-%  Develop this representation and compare it with the other ordinal notations.
-
-% \end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Let $N_A$ be a notation system for ordinals strictly less than $\alpha$, 
-with the strict order $(A,<_A)$. Please build the notation system
-\texttt{ON\_Expl $N_A$}, on the type of multisets of elements of $A$
-(or, if preferred, the type of non-increasing finite sequences on $A$,
-provided with the lexicographic ordering on lists).
-
-For instance, let us take $N_A=\texttt{Omega}$, and take $\alpha=\langle 4,4,2,1,0\rangle$,
- $\beta=\langle 4,3,3,3,3,3,2\rangle$, and $\gamma=\langle 5\rangle$. Then $\beta<\alpha<\gamma$.
-
-In contrast the list $\langle5,6,3,3\rangle$ is not non-increasing (\emph{i.e.} sorted w.r.t. $\geq$), so it is not to be considered.
-
-Note that if the notation $N_A$ implements the ordinal 
-$\alpha$,  the new notation $\omega^{N_A}$ must implement the ordinal $\phi_0(\alpha)$, a.k.a. $\omega^\alpha$ (see chapter~\ref{chap:schutte})
-
-\end{project}
-
-\begin{remark}
- The set of ordinal terms in Cantor normal form (see Chap.~\ref{chap:T1}) and 
-in Veblen normal form (see 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}{Gamma0.Gamma0}) are shown to be ordinal notation systems, but there is a lot of work to be done in order to unify ad-hoc  definitions and proofs which were written before the definition of the \texttt{ON} type class.
-\end{remark}
-
-\section{Ordinal Notations and \gaia's ordinals}
-\gaiasign Module~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge} contains an instance of class \texttt{ON E0lt compare}, where \texttt{E0lt} is the order on the well-formed ordinal terms below $\epsilon_0$ defined in \gaiaHydras library (please see Chapter~\ref{gaia-chapter}).
-
-%------------------------------------------------------------------------
-
-\chapter[A proof of termination, using epsilon0]{A proof of termination, using ordinals below \texorpdfstring{$\epsilon_0$}{Epsilon0}}
-
-\label{cnf-math-def}
-\label{chap:T1}
-
-In this chapter, we adapt to \coq{} the well-known~\cite{KP82}  proof that Hercules eventually wins every battle, whichever the strategy  of each player.
-In other words, we present  a formal and self contained proof of termination  of all [free] hydra battles.
-First, we take from Manolios and Vroon~\cite{Manolios2005} a representation of the ordinal $\epsilon_0$ as terms in Cantor normal form.
-Then, we define a variant for hydra battles as a measure that maps any hydra to some ordinal strictly less than $\epsilon_0$.
-
-\section{The ordinal \texorpdfstring{\(\epsilon_0\)}{epsilon0}}
-\label{sec:epsilon0-intro}
-
-The ordinal \(\epsilon_0\) is the least ordinal number that satisfies 
-the equation \(\alpha = \omega^\alpha\), where \(\omega\) is 
-the least infinite ordinal.
-Thus, we can intuitively consider \(\epsilon_0\) as an
-\emph{infinite} \(\omega\)-tower.
-
-\subsection{Cantor normal form}
-\index{maths}{Cantor normal form}
-
-Any ordinal strictly less that \(\epsilon_0\) 
-can be finitely represented by a unique  \emph{Cantor normal form}, 
-that is, an expression  which is either  the ordinal \(0\) or 
-a sum  \(\omega^{\alpha_1} \times n_1 + \omega^{\alpha_2} \times n_2 + 
-  \dots + \omega^{\alpha_p} \times n_p\) where all the \(\alpha_i\) 
-are ordinals in Cantor  normal form, \(\alpha_1 > \alpha_2 > \alpha_p\), 
-and all the \(n_i\) are positive integers.
-
-An example of Cantor normal form is displayed in Fig \ref{fig:cnf-example}:
-Note that  any ordinal of
-the form \(\omega^0 \times i + 0\) is just written \(i\).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
-\node[color=blue]{$\omega^{(\omega^\omega\,+\, \omega^2 \times 8 \,+\, \omega)}+ \omega^\omega + \omega^4+ 6$};
-\end{tikzpicture}
-\caption{\label{fig:cnf-example}
-An ordinal in Cantor normal form}
-\end{figure}
-
-In the rest of this section, we define an inductive type for representing in \texttt{Coq}
-all the ordinals strictly  less than  \(\epsilon_0\), then extend some arithmetic operations
-to this type, and finally prove that our representation fits well with 
-the expected mathematical properties: the order we define is a well order, 
-and the decomposition into Cantor normal form  is consistent 
-with the implementation of the arithmetic operations of exponentiation of base \(\omega\) 
-and addition.
-
-\paragraph*{Remark}
-\label{sec:orgheadline65}
-Unless explicitly mentioned, the term ``ordinal" will be used instead of
-``ordinal strictly less than \(\epsilon_0\)" (except in Chapter~\ref{chap:schutte} where it stands for ``countable ordinal'').
-
-\subsection{A data type for  ordinals in Cantor normal form}
-\label{sec:orgheadline72}
-\label{sec:T1-inductive-def}
-
-% Our user contribution~\cite{CantorContrib} represents 
-% the set of ordinals strictly less than $\epsilon_0$ in Cantor normal form as in~\cite{Manolios2005}, and also the set
-% of ordinals strictly  less than $\Gamma_0$ in Veblen normal form.
-
-    Let us define an inductive type whose 
-constructors are respectively associated
-with the ways to build Cantor normal forms:
-
-\begin{itemize}
-\item the ordinal \(0\)
-\item the construction \((\alpha,\, n,\,\beta)  \mapsto \omega^\alpha \times (n + 1)+ \beta \quad (n\in\mathbb{N})\)
-\end{itemize}
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#T1}{Epsilon0.T1}}
-
-\label{types:T1}
-%\index{Constants!zero:T1}
-\index{hydras}{Library Epsilon0!Types!T1}
-
-\input{movies/snippets/T1/T1Def}
-
-\subsubsection{Example}
-
-\label{alpha0-def}
-For instance, the ordinal  $\omega^\omega+\omega^3\times 5+2$ is represented by the following term of type \texttt{T1}:
-
-\input{movies/snippets/T1/alpha0}
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5, level 1/.style={sibling distance=6cm},
-level 2/.style={sibling distance=35mm},  
-level 3/.style={sibling distance=17mm}]
-\node  {cons}
-  child {  node {cons}
-            child { node {cons} child {node {zero}} child {node{0}} child{node{zero}}}
-         child {node {0}}
-         child {node {zero}}}
-    child {node {0}}
-   child {node {cons} 
- child { node {cons} child {node {zero}} child {node{2}} child{node{zero}}}
-  child {node {4}}
-         child {node {cons} child {node {zero}} child {node{1}} child{node{zero}}}};
-
-\end{tikzpicture}
-
-\caption{The tree-like representation of the ordinal $\omega^\omega+\omega^3\times 5 +2$\label{fig:cnf-tree}}
-
-\end{figure}
-
-\paragraph{Remark}
-For simplicity's sake, we chose to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the construction (\texttt{cons $\alpha$ $n$ $\beta$}) is intended to represent the
-ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
-In a future version, we should replace  the type \texttt{nat} with \texttt{positive} in \texttt{T1}'s 
-definition. But this replacement would take a lot of time \dots{}
-
-\paragraph{Remark}
-The name \texttt{T1} we gave to this data-type  is proper to this development and refers
-to a hierarchy of ordinal notations.
-For instance, in Library \href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2},  the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{Gaia, schutte}).
-\noindent
-
-\input{movies/snippets/T2/T2Def}
-
-\subsection{About Gaia}
-\gaiasign Chapter~\vref{gaia-chapter} describes the present state of a
-project of making compatible the libraries \HydrasLib and \gaia. At present, both libraries contain their own version of the inductive type \texttt{T1}, with the same base name for the constructors.
-
-\inputsnippets{nfwfgaia/T1def}
-
-Thus, many examples of this and following chapters can be run
-in \gaia's context. We will signal any difference (notation, name) which may appear.
-A closer integration (same data type and functions) is still in project.
-
-\begin{remark}
-  \label{remark:hydra-gaia-renaming}
-  \textcolor{lookcolor}{In future releases, we plan to make some
-    \HydrasLib identifiers progressively  deprecated, in favour of \gaia's names.}
-\end{remark}
-
-\subsection{Abbreviations}
-
-Some abbreviations may help to write more concisely complex ordinal terms.
-
-\subsubsection{Finite ordinals}
-\label{sec:orgheadline67}
-
-For representing finite ordinals, \emph{i.e.} natural numbers, we first introduce a notation for terms of the form $n+1$, then define a coercion from type \texttt{nat} into \texttt{T1}.
-\label{sect:notation-FS}
-
-\label{sect:notation-F}
-
-\input{movies/snippets/T1/finiteOrds}
-
-\paragraph*{\gaiasign}
-  In \texttt{gaia.ssete9}, the $n$-th finite ordinal is also written
-  \texttt{$\backslash${F}\,$n$}. There is no coercion in \gaia from \texttt{nat} to \texttt{T1}.
-
-  \vspace{4pt}
-
-  From~\href{../theories/html/gaia_hydras.HydraGaia_Examples.html}{gaia\_hydras.HydraGaia\_Examples}.
-
-  \inputsnippets{HydraGaia_Examples/Demoa}
-
-% \index{Coq!Coercions}
-% \index{Functions!Coercions@Coercions (from nat to ordinal types)}
-% \begin{remark}
-% Please refer to the remark~\pageref{warning:coercions} about the use of coercions.
-% % The use of coercions like \texttt{fin} allow us to be close to the mathematical tradition where natural numbers are ordinals too.
-% % Nevertheless, it may happen that a goal like \texttt{3 < 5} could be 
-% % interpreted as \texttt{(lt (fin 3) (fin 5))},  depending on the current notation scope.  
-% % When this misinterpretation happens, tactics like \texttt{auto with arith}, \texttt{lia} do not work!
-% % Thus, it is useful to write \texttt{(3 < 5)\%nat}  an inequality between two natural numbers. 
-% \end{remark}
-
-\subsubsection{The ordinal \(\omega\)}
-\label{sec:orgheadline68}
-
-  Since \(\omega\)'s Cantor normal form is
-i.e. \(\omega^{\omega^0}\times 1+ 0\), we can define the following abbreviation:
-
-\label{sect:omega-notation2}
-
-\input{movies/snippets/T1/omegaDef}
-
-Note that \texttt{omega} is not an identifier in \HydrasLib, thus any tactic like \texttt{unfold omega} would fail.
-
-\paragraph*{\gaiasign}
-In \texttt{gaia.ssete9}, the ordinal $\omega$ is bound to the \emph{constant}  \texttt{T1omega}.
-
-\subsubsection{The ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
-\label{sect:notation-phi0}
-We provide also a notation for ordinals of the form $\omega^\alpha$.
-
-\index{hydras}{Library Epsilon0!Notations!phi0@phi0 (exponential of base omega)}
-
-\input{movies/snippets/T1/phi0Def}
-
-\index{maths}{Additive principal ordinals}
-
-\begin{remark}
-\label{sec:orgheadline69}
-The name \(\phi_0\)
-   comes from ordinal numbers theory. In~\cite{schutte}, Schütte defines 
-$\phi_0$  as the ordering (\emph{i.e.} enumerating) function of the set  of \emph{additive principal ordinals} \emph{i.e.} strictly positive ordinals $\alpha$ that verify $\forall \beta<\alpha, \beta+\alpha=\alpha$.
-For Schütte,  $\omega^\alpha$ is just a notation for $\phi_0(\alpha)$.
-See also Chapter~\ref{chap:schutte} of this document.
-\end{remark}
-
-\paragraph*{\gaiasign}
-In \texttt{gaia.ssete9} the identifier \texttt{phi0} is bound to a plain constant (not a notation).
-
-\subsubsection{The hierarchy of \(\omega\)-towers:}
-\label{sec:orgheadline71}
-
-The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximated by the sequence of $\omega$-towers (see also Sect~\vref{sect:epsilon0-as-limit} ).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/towerDef}
-
-For instance, Figure~\ref{fig:tower7} represents  the ordinal returned by the
- evaluation of the term \texttt{omega\_tower 7}.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
-\node[color=blue]{$\omega^{{{\omega}^{{{\omega}}^{{{\omega}}^{{\omega^{{\omega}^{\omega}}}}}}}}$};
-\end{tikzpicture}
-\caption{\label{fig:tower7}
-The $\omega$-tower of height 7}
-\end{figure}
-
-\subsection{Pretty-printing ordinals in Cantor normal form}
-\label{sect:ppT1}
-\index{hydras}{Library Epsilon0!Types!ppT1}
-
-Let us consider again the ordinal $\alpha_0$ defined in section~\vref{alpha0-def}. 
-If we ask \coq{} to print its  normal form, we get a hardly readable term of type \texttt{T1}.
-
-\input{movies/snippets/T1/alpha0Compute}
-
-The following data type defines a more readable abstract syntax for  ordinals terms in Cantor normal form:
-
-\label{types:ppT1}
-\index{hydras}{Library Epsilon0!Functions!pp@ pp (pretty printing terms in Cantor normal form)}
-
-\input{movies/snippets/T1/ppT1Def}
-
-The function \texttt{pp: T1 -> ppT1} converts any closed term of type \texttt{T1} into a human-readable expression. For instance, let us convert the term \texttt{alpha\_0}.
-
-\input{movies/snippets/T1/ppAlpha0}
-
-\paragraph*{\gaiasign}
-In ~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge},
-we define a variant of \texttt{pp} for pretty-printing terms of type \texttt{ssete9.CantorOrdinal.T1} (see Sect.~\vref{sect:gaia-ppT1}).
-
-\index{hydras}{Projects}
-\begin{project}
-Design  (in \ocaml?) a set of tools for systematically pretty printing ordinal terms in Cantor normal form.
-\end{project}
-
-\subsection{Comparison between ordinal terms}
-\label{sec:orgheadline73}
-
-% Our formalisation of Cantor Normal Form will take two steps:
-% 1 Definition of a strict order \texttt{o<} on the type \texttt{T1}, 
-% 2 Using \texttt{o<} for characterizing terms in normal form.
-
-In order to compare two terms of type \texttt{T1}, we define a recursive function \texttt{compare} that maps two ordinal tems $\alpha$ and $\beta$ to a value of type \texttt{comparison}.
-This type is defined in \coq's standard library
-\texttt{Init.Datatypes} and
-contains three constructors:  \texttt{Lt} (less than), \texttt{Eq} (equal), and
-\texttt{Gt} (greater than).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/compareDef}
-
-\label{Predicates:lt-T1}
-Please note that this definition of \texttt{lt} makes it easy to write proofs by computation, as shown by the following examples.
-
-\vspace{4pt}
-
-\input{movies/snippets/T1/ltExamples}
-
-Links between the function \texttt{compare} and the relations
-\texttt{lt} and \texttt{eq} are established through the following lemmas (~\vref{sect:comparable-def}).
-\vspace{4pt}
-
-\input{movies/snippets/T1/Instances}
-
-\paragraph*{\gaiasign}
-In \gaia, the strict order \texttt{T1lt} on ordinal terms is directly defined as a boolean function (in \ssreflect's style).
-
-\inputsnippets{nfwfgaia/T1ltDef}
-
-In section~\vref{sect:lt-compat-gaia}, we show that both definitions are mutually equivalent.
-
-\subsubsection{A Predicate for Characterizing Normal Forms}
-\label{sect:t1-nf}
-
-\label{sec:orgheadline74}
-\label{sec:orgheadline75}
-Our data-type \texttt{T1} allows us to write expressions that
-are not properly in Cantor normal form as specified in Section \ref{sec:epsilon0-intro}.
-For instance, consider the following term of type  \texttt{T1}.
-
-\input{movies/snippets/T1/badTerm}
-
-This term would have been written \(\omega^1\times 2 + \omega^\omega \times 3\) in the usual mathematical notation.
-We note that the exponents of $\omega$ are not in the right (strictly decreasing) order.
-Nevertheless, with the help of the order \texttt{lt} on \texttt{T1}, we are now able to characterize
-the set of all well-formed ordinal terms:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#nf_b}{Epsilon0.T1}}
-
-\label{Predicates:nf-T1}
-
-\input{movies/snippets/T1/nfDef}
-
-\input{movies/snippets/T1/nfAlpha0}
-
-\input{movies/snippets/T1/nfBadTerm}
-
-\paragraph*{\gaiasign}
-In \gaia, the boolean function which characterizes ordinal terms in normal form is defined as follows:
-
-\inputsnippets{nfwfgaia/T1nfDef}
-
-In Sect.~\vref{nf-gaia-compat}, we show that \gaia's \texttt{T1nf} is extensionally equivalent with \HydrasLib'  \texttt{nf}.
-
-\subsection{Making normality implicit}
-  We would like to get rid of terms of type \texttt{T1} which are not in Cantor normal form.
-A simple way to do this is to consider statements of the form 
-\texttt{forall alpha: T1, nf alpha -> $P$ alpha}, where $P$ is a predicate over type \texttt{T1}, like in the following lemma \footnote{Ordinal addition is formally defined a little later (page~\ref{sect:infix-plus-T1})}.
-
-\input{movies/snippets/T1/plusIsZero}
-
-\vspace{4pt}
-
-But this style leads to clumsy statements, and generates too many subgoals in interactive proofs (although often solved with \texttt{auto} or \texttt{eauto}).
-
-One may encapsulate conditions of the form \texttt{(nf $\alpha$)} in
-the most used predicates. For instance, we introduce the restriction of \texttt{lt} to terms in normal form, and provide a handy notation for this restriction.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Prelude.Restriction.html}{hydras.Prelude.Restriction}}
-
-\input{movies/snippets/Restriction/restrictionDef}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#LT}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/LTDef}
-
-\label{Predicates:LT-T1}
-
-For instance, in the following lemma, the condition that $\alpha$ is in normal form is included in the condition $\alpha< 1$.
-
-\input{movies/snippets/T1/LTOne}
-
-\subsubsection{\texttt{E0}: a sigma-type for \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-As we noticed in Sect.~\ref{sect:t1-nf}, the type \texttt{T1} is not a correct ordinal notation, since it contains terms that are not in Cantor normal form.
-In certain contexts (for instance in Sections~\ref{sect:L-equations}, \ref{sect:hardy},
-and \ref{sect:wainer}),  we need to define total recursive functions on well-formed ordinal terms less  than $\epsilon_0$, using the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777}.
- In order to define a type whose inhabitants represent just ordinals, we build a type gathering a term of type \texttt{T1} and a proof that this term is in normal form.
-
-\label{sect:E0-def}
-\label{types:E0}
-\index{hydras}{Library Epsilon0!Types!E0}
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html\#E0}{Epsilon0.E0}}
-
-\input{movies/snippets/E0/E0Def}
-
-Many constructs : types, predicates, functions, notations, etc., on type \texttt{T1} are adapted to \texttt{E0}.
-
-First, we declare a notation scope for \texttt{E0}, then we redefine the predicates of comparison.
-
-\input{movies/snippets/E0/E0Scope}
-\label{Predicates:Lt-E0}
-\input{movies/snippets/E0/LtLeDef}
-
-Equality in \texttt{E0} is just Leibniz equality. Note that, since \texttt{nf} is
-defined by a Boolean function, for  any term $\alpha:\texttt{T1}$, there exists at most one proof of \texttt{nf $\alpha$}, thus two ordinals of type \texttt{E0} are
-equal if and only if their projection to \texttt{T1} are equal (see also Sect.~\vref{sect:eq-proof-unicity}).
-
-\vspace{4pt}
-
-\index{coq}{Unicity of equality proofs}
-
-\inputsnippets{E0/nfProofUnicity}
-\inputsnippets{E0/E0EqIff}
-
-In order to  upgrade constants and functions from type \texttt{T1} to \texttt{E0}, we have to prove that
-the term they build is in normal form.
-For instance, let us represent the ordinals $0$ and $\omega$   as instances of the class \texttt{E0}.
-
-\vspace{4pt}
-
-\label{sect:omega-T1}
-
-\input{movies/snippets/E0/ZeroOmega}
-
-%\index{Constants!zero:T1}
-
-\paragraph*{\gaiasign}
- The library~\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines a \gaia-compatible type \texttt{E0}.
-
-\subsection{Syntactic definition of limit and successor ordinals}
-
-Pattern matching and structural recursion allow us to define boolean characterizations  of successor and limit ordinals.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/succbLimitb}
-
-The correctness of these definitions with respect to the mathematical notions of
-limit and successor ordinals is established through several lemmas. For instance,
-Lemma \texttt{canonS\_limit\_lub}, page~\pageref{lemma:canonS-limit}, shows that
-if $\alpha$ is (syntactically) a limit ordinal, then it is the least upper bound of
-a strictly increasing sequence of ordinals.
-
-\paragraph*{\gaiasign}
-In \gaia, the boolean functions associated with limit and successor ordinals are called \texttt{T1limit} and \texttt{T1is\_succ}.
-
- From~\href{../theories/html/gaia_hydras.HydraGaia_Examples.html}{gaia\_hydras.HydraGaia\_Examples}.
-
-\inputsnippets{HydraGaia_Examples/Demob}
-
-\subsection{Arithmetic on \texorpdfstring{$\epsilon_0$}{epsilon0}}
-\subsubsection{Successor}
-
-\index{hydras}{Library Epsilon0!Functions!succ}
-
-The successor of any ordinal $\alpha< \epsilon_0$ is defined by structural 
-recursion on its Cantor normal form.
-
-\label{Functions:succ-T1}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/succDef}
-
-The following lemma establishes the connection between the  function
-\texttt{succ} and the Boolean predicate \texttt{succb}.
-
-\vspace{4pt}
-
-\input{movies/snippets/T1/succbIff}
-
-\begin{exercise}[\gaiasign]
-  Look for the \gaia-theorem which corresponds to \texttt{succb\_iff}.
-\end{exercise}
-
-\subsubsection{Successor function on \texttt{E0}}
-
-The function \texttt{succ} on \texttt{T1} is extended to \texttt{E0} the following way:
-
- \emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
-
- \inputsnippets{E0/SuccOnE0}
-
-\index{hydras}{Exercises}
- \begin{exercise}
-Prove in \coq{} that for any ordinal $0< \alpha<\epsilon_0$, $\alpha$ is a limit if 
-and only if for all $\beta<\alpha$, the interval $[\beta,\alpha)$ is
-infinite.
-
-\emph{You may start this exercise with the file
-     \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/Limit_Infinity.v}{exercises/ordinals/Limit\_Infinity.v}.}
- \end{exercise}
-
-\subsubsection{Addition and multiplication}
-
-Ordinal addition and multiplication are also defined by structural recursion over the type \texttt{T1}. Please note that they use the \texttt{compare} function on some subterms of their arguments.
-
-\label{sect:infix-plus-T1}
-
-\input{movies/snippets/T1/plusDef}
-
-\input{movies/snippets/T1/multDef}
-
-\paragraph*{\gaiasign}
-We keep \gaia's base names for
-addition and multiplication of ordinal terms below $\epsilon_0$.
-Please refer to Sect.~\ref{sect:plus-mult-gaia-hydras} about compatibility of both arithmetics.
-
-\subsubsection{Examples}
-
-The following examples are instances of \emph{proofs by computation}. Please note that  addition and multiplication on \texttt{T1}
-are not commutative. Moreover,  both operations fail to be strictly monotonous in their first argument.
-
-\input{movies/snippets/T1/plusMultExamples}
-
-\input{movies/snippets/T1/notMono}
-
-The function \texttt{succ} is related with addition through the following lemma:
-
-\inputsnippets{T1/succIsPlusOne, T1/succIsPlusOnez}
-
-\subsubsection{Arithmetic on type \texttt{E0}}
-
- We define an addition in type \texttt{E0}, since the sum of two terms in normal form is in normal form too.
-
-\input{movies/snippets/T1/plusNf}
-
-\input{movies/snippets/E0/plusE0}
-\input{movies/snippets/E0/CheckPlus}
-
-\begin{remark}
-In all this development, two representations of ordinals co-exist: ordinal terms (type \texttt{T1}, notation scope \texttt{t1\_scope}, for reasoning on the tree-structure of Cantor normal forms), and ordinal terms \emph{known to be in normal form} (type \texttt{E0}, notation scope \texttt{E0\_scope}).
-Looking at the contexts displayed by \coq{} prevents you from any risk of confusion.
-\end{remark}
-
-%% To simplify !
-\index{hydras}{Exercises} 
-\begin{exercise}
-Prove that for any ordinal $\alpha:\texttt{E0}$, 
-$\omega\leq \alpha$ if and only if, for any natural number $i$,
-$i+\alpha=\alpha$.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/ge_omega_iff.v}{exercises/ordinals/ge\_omega\_iff.v}.}
-\end{exercise}
-
-\subsection{A proof by computation}
-\label{Ex42-E0}
-
-It is interesting to compare the following proof of the equality
-$\omega+42+\omega^2$ with the more theoretical proof in Sect~\vref{Ex42-schutte}.
-
-\inputsnippets{E0/Ex42}
-
-\section{Well-foundedness and transfinite induction}
-
-\index{maths}{Transfinite induction}
-
-\subsection{About  well-foundedness}
-\label{sec:T1Wf}
-\label{sec:orgheadline82}
-   In order to use \texttt{T1} for proving termination results,
-we need to prove that  our order \texttt{<} is well-founded. Then we will get \emph{transfinite induction} for free.
-
-The proof of well-foundedness of the strict order $<$ on Cantor normal forms is already 
-available in the Cantor contribution by Castéran and Contejean~\cite{CantorContrib}. That proof relies on a library on recursive path orderings written by
-E. Contejean. We present here  a direct proof of the same result, which does not require any knowledge on r.p.o.s.
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that the \emph{total} order \texttt{lt} on \texttt{T1} is not well-founded. 
-\textbf{Hint:}  You will have to build a counter-example with terms of type \texttt{T1}
-which are not in Cantor normal form.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/T1_ltNotWf.v}{exercises/ordinals/T1\_ltNotWf.v}.}
-\end{exercise}
-
-\subsubsection{A first attempt}
-\label{sec:orgheadline77}
-%\index{coq}{Well-founded induction}
-
-It is natural to try to prove by structural induction over \texttt{T1} 
-that every term in normal form is \texttt{LT}-accessible.
-
-Unfortunately, it won't work. Let us consider some well-formed term
- $\alpha=\texttt{cons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are \texttt{LT}-accessible.
-In order to prove the accessibility of $\alpha$, we have to consider
-any well formed term \(\delta\) such that \(\delta<\alpha\). 
-But nothing guarantees that \(\delta\)  is strictly  less than \(\beta\) nor \(\gamma\), and we cannot use the induction hypotheses on   \(\beta\) nor \(\gamma\).
-
-\input{movies/snippets/T1/wfLTBada}
-
-The problem comes from the too weak hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
-\(\gamma\);
-for instance \(\delta\) may be of the form
-\texttt{cons $\beta$ $p'$  $\gamma'$},
-where    \(p' < n\).
-Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.
-
-\input{movies/snippets/T1/wfLTBadz}
-
-\subsubsection{Using a stronger inductive predicate.}
-\label{sec:orgheadline78}
-\label{sec:strongly-accessible}
-
-  Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
-through \texttt{LT}, we propose to consider the following (stronger) predicate:
-
-\input{movies/snippets/T1/AccStrongDef}
-
-The following lemma is an application of the strict inequality 
-$\alpha < \omega ^\alpha$. If $\alpha$ is strongly accessible, then, by definition,
-$\omega^\alpha$ is accessible, thus $\alpha$ is \emph{a fortiori} accessible.
-
-\input{movies/snippets/T1/AccStrongStronger}
-
-Thus, it remains to prove that every ordinal strictly less than $\epsilon_0$
-is strongly accessible.
-
-% \subsubsection{Structure of the proof of well-foundedness of \texttt{LT}}
-
-\label{sec:orgheadline81}
-\label{proof-wf-epsilon0}
-\paragraph{A helper}
-\label{sec:orgheadline79}
-
-First, we prove that, for  any \texttt{LT}-accessible term $\alpha$, $\alpha$ is 
-strongly accessible too (\emph{i.e.} any well formed
-term (\texttt{cons $\alpha$ $n$ $\beta$})  is accessible).
-
-The proof is structured as an induction on $\alpha'$s accessibility. Let us consider
-any  accessible term $\alpha$.
-
-\input{movies/snippets/T1/AccImpAccStrong}
-
-Let \texttt{n:nat} and \texttt{beta:T1} such that (\texttt{cons alpha n beta}) is in normal form. 
-We prove first that \texttt{beta} is accessible,  which allows us to prove by well-founded induction on \texttt{beta}, 
-and natural induction on \texttt{n}, that (\texttt{cons alpha n beta}) is accessible.
-The proof, quite long, can be consulted in \href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}.
-
-\paragraph{Accessibility of any well-formed ordinal term}
-\label{sec:orgheadline80}
-
-Our goal is still to prove accessibility of any well formed ordinal term.
-Thanks to our previous lemmas, we are almost done (by a simple structural induction!).
-
-\input{movies/snippets/T1/nfAcc}
-
-\input{movies/snippets/T1/T1Wf}
-
-\index{maths}{Transfinite induction}
-
-Traditionnally, well-founded induction on ordinals is called \emph{transfinite induction}.
-
-\inputsnippets{T1/transfiniteRecursor, T1/transfiniteInduction}
-
-As a corollary, the order \texttt{Lt} on type \texttt{E0} is well-founded too.
-
-\inputsnippets{E0/E0LtWf}
-
-\begin{remark}
-\label{remark:a3pat}
-The alternate proof of well-foundedness using \'Evelyne Contejean's work on
-    recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the
-    library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0.Epsilon0rpo}.
- \end{remark}
-
-\subsection{An ordinal notation for  \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-We build an instance of \texttt{ON}, and prove its correctness w.r.t. Schutte's model (see Chapter~\ref{chap:schutte}).
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
-
-\input{movies/snippets/E0/InstanceEpsilon0}
-
-\label{instance-epsilon0}
-
-\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
-
-\input{movies/snippets/Correctness_E0/injectDef}
-
-\input{movies/snippets/Correctness_E0/Epsilon0Correct}
-
-\index{hydras}{Projects}
-\begin{project}
- \emph{This exercise is a continuation of Project~\vref{exo:ON-mult}.}
-Use \texttt{ON\_mult} to define an ordinal notation \texttt{Omega2} for $\omega^2=\omega\times\omega$.
-
-Prove that \texttt{Omega2} is a sub-notation of \texttt{Epsilon0}.
-
-Define on \texttt{Omega2} an addition compatible with the addition on \texttt{Epsilon0}.
-
-\textbf{Hint}. You may use the following definition (in 
-    \href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}).
-
-\input{movies/snippets/ON_Generic/SubONSameOp}
-
-     \end{project}
-
-    \index{hydras}{Projects}
-    \begin{project}
-    The class \texttt{ON} of ordinal notations has been defined long after this 
-    chapter, and is not used in the development of the type \texttt{E0} yet.
-    A better integration of both notions should simplify the development on ordinals in Cantor normal form. This integration is planned for the future versions.
-
-    \end{project}
-
-    \section{A refinement of \texttt{E0} : an ordinal notation for \texorpdfstring{$\omega^\omega$}{omega\^omega}}
-
-    In Module   \href{https://github.com/coq-community/hydra-battles/blob/master/theories/
-      ordinals/OrdinalNotations/OmegaOmega.v}{theories/ordinals/OrdinalNotations/OmegaOmega.v},
-    we represent ordinals below $\omega^\omega$ by lists of pairs of natural numbers (with the same coefficient shift as in \texttt{T1}).
-    For instance, the ordinal $\omega^4\times 10 + \omega^3 + \omega+ 5$ is represented by the list \texttt{(4,9)::(3,0)::(1,0)::(0,4)::nil}.
-
-\inputsnippets{OmegaOmega/LODef}
-
-  The usual operations : \texttt{succ}, \texttt{+}, \texttt{*} are simple variants of the same operations in \texttt{T1}.
-
-  \inputsnippets{OmegaOmega/succPlusMult}
-
-    We establish this representation as a \emph{refinement} of the data types we used to represent ordinals less than $\epsilon_0$.
-Thus, many properties like well-foundedness of $<$ and associativity of $+$,  of this ordinal notations have very short proofs.
-
-    \inputsnippets{OmegaOmega/refineDef, OmegaOmega/phi0Ref,
-      OmegaOmega/succRef, OmegaOmega/plusRef,
-      OmegaOmega/multRef}
-
-    In order to make an ordinal notation for $\omega^\omega$, we follow the same steps as for $\epsilon_0$:
-    \begin{enumerate}
-    \item Define an order \texttt{lt}, which refines the order
-      \texttt{lt} on \texttt{T1}.
-      \inputsnippets{OmegaOmega/compareDef}
-      \inputsnippets{ OmegaOmega/ltDef, OmegaOmega/ltRef}
-
-    \item Define the predicate ``to be in normal form''.
-      \inputsnippets{OmegaOmega/nfDef}
-    \item Define a class \texttt{OO} of terms in normal form, and an embedding from \texttt{E0} into \texttt{OO}.
-      \inputsnippets{OmegaOmega/OODef}
-       \inputsnippets{OmegaOmega/embedDef}
-      \item Infer well-foundedness of the order on \texttt{OO}.
-    \inputsnippets{OmegaOmega/ltWf}
-
-    \end{enumerate}
-
-    \inputsnippets{OmegaOmega/ONOO, OmegaOmega/OOz}
-
-    Let us show a few examples.
-
-    \inputsnippets{OmegaOmega/OODemo}
-
-    \index{hydras}{Exercises}
-
-    \subsection{Related work}
-    The article~\cite{Manoury2016}  defines another representation of ordinals below $\omega^\omega$ based on lists of natural numbers.
-
-\begin{exercise}[not trivial]
-  It may be interesting to write a \emph{direct} proof  of well-foundedness of the order in $\omega^\omega$ (\emph{i.e.}
-  without using properties of $\epsilon_0$). This exercise may
-  help to understand better the proof structure of Sect.~\vref{sec:strongly-accessible}.
-    \end{exercise}
-%ici
-    \section{A variant for hydra battles}
-
-    In order to prove the termination of any hydra battle, we try to define a variant mapping hydras to ordinals strictly less than $\epsilon_0$.
-    In order to make such a variant easy to define (for instance by a structural recursion), we introduce a variant of addition, which, contrary to
-    $+$, is commutative and strictly monotonous in both of its arguments. This last property makes it possible to prove that our function is 
-    truly a variant for hydra battles (in Sect.~\vref{sect:variant-decr}).
-
-    \subsection{Natural sum (a.k.a. Hessenberg's  sum)}
-    \label{sec:orgheadline87}
-    \label{hydra-variant}
-    \label{sect:hessenberg-def}
-
-    Natural sum (Hessenberg's  sum) is a commutative and monotonous version of
-    addition. It is used as an auxiliary operation  for defining variants
-    for hydra battles, where Hercules is allowed to chop off any  head of the hydra.
-
-    In the literature, the natural sum of ordinals \(\alpha\) and \(\beta\)
-    is often denoted by \(\alpha \# \beta\)  or  \(\alpha \oplus  \beta\).
-    Thus we called \texttt{oplus} the associated \emph{Coq} function.
-
-    \subsubsection{Definition of \texttt{oplus}}
-    \label{sec:orgheadline84}
-    %\index{Functions!oplus @ oplus (Hessenberg commutative sum)}
-
-    The definition of \texttt{oplus} is recursive in both of its 
-    arguments and uses  the same pattern as for the \texttt{merge} function on lists of library
-    \texttt{Coq.Sorting.Mergesort}.
-
-    \begin{enumerate}
-    \item Define a nested recursive function, using the \texttt{Fix} 
-        construct
-
-    \item Build a principle of induction dedicated to \texttt{oplus}
-
-    \item Establish equations associated to each case of the definition.
-    \end{enumerate}
-
-    \paragraph{Nested recursive definition}
-    \label{sec:orgheadline83}
-
-    The following definition is composed of 
-    \begin{itemize}
-    \item A main function \texttt{oplus}, structurally recursive in its 
-    first argument \texttt{alpha}
-    \item An auxiliary function \texttt{oplus\_aux} within the scope of \texttt{alpha},
-    structurally recursive in its argument \texttt{beta};  \texttt{oplus\_aux beta} 
-       is supposed to compute  \texttt{oplus alpha beta}.
-    \end{itemize}
-
-%  \pagebreak 
-  \vspace{4pt}
-    \emph{From Module~\href{../theories/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}
-
-    \label{sect:infix-oplus}
-
-    \input{movies/snippets/Hessenberg/oplusDef}
-
-    The reader will note that each recursive call of the functions
-    \texttt{oplus} and \texttt{oplus\_aux} satisfies \emph{Coq}'s constraint
-    on recursive definitions. The function \texttt{oplus} is recursively called on a sub-term of its first argument,
-    and \texttt{oplus\_aux} on a sub-term of its unique argument.
-    Thus, \texttt{oplus}'s definition is accepted by \coq{} as a structurally recursive function.
-
-    \subsubsection{Rewriting lemmas}
-    \label{sec:orgheadline86}
-
-    \emph{Coq}'s constraints on recursive definitions result in 
-    the quite  complex form of \texttt{oplus}'s definition.
-    Proofs of properties of this function can be simpler if we
-     derive a few  rewriting lemmas that will help to simplify 
-    expressions of the form (\texttt{oplus $\alpha$ $\beta$}).
-
-    A first set of lemmas correspond to the various cases of \texttt{oplus}'s 
-    definition. They can be proved almost immediately. Here are a few examples.
-
-    \inputsnippets{Hessenberg/oplusNeutral,
-      Hessenberg/oplusCompareLt}
-
-    % \subsubsection{A hand-made induction principle}
-    % \label{sec:orgheadline85}
-
-    % \index{Coq!Commands!Functional Scheme}
-
-    % \emph{Coq} contains a command  \texttt{Functional Scheme} that 
-    % generates induction principles which correspond to recursive functions.
-    % Unfortunately, the current version ( \texttt{8.11.0} ) doesn't work on \texttt{oplus},
-    % probably because of the inner \texttt{Fix}.
-
-    % \begin{Coqsrc}
-    % Functional Scheme oplus_ind := Induction for oplus Sort Prop.
-    % \end{Coqsrc}
-
-    % \begin{Coqanswer}
-    % Error: Anomaly "todo." Please report at http://coq.inria.fr/bugs/.
-    % \end{Coqanswer}
-
-    % Fortunately, it's a good exercise for a semi-experienced user, to write
-    % her/him-self induction principles similar to the ones returned by
-    % \texttt{Functional Scheme}.
-
-    % \begin{itemize}
-    % \item First, we choose to write a version for sort \texttt{Type}, since versions
-    % for sorts \texttt{Prop} and \texttt{Set} can be easily derived from
-    % the former one. According to \emph{Coq}'s naming politics, we will call our 
-    % principle \texttt{oplus\_rect}
-
-    % \item The conclusion of \texttt{oplus\_rect} will be (\texttt{$P$ a b (oplus a b)}),
-    % where $P$ is an arbitrary function of type 
-    % \texttt{T1 -> T1 -> T1 -> Type}
-
-    % \item The premises of \texttt{oplus\_rect} will describe how to build an induction 
-    % on the graph of \texttt{oplus}.
-    % \end{itemize}
-
-    % We are now ready to state and prove \texttt{oplus\_rect}, and the reader
-    % will note that the statement is longer than the proof script itself,
-    % which is a standard proof by induction, simplification and case-analysis 
-    % that follows  \texttt{oplus}'s definition.
-
-    % We associate also a tactic to the application of \texttt{oplus\_rect}.
-
-    % \begin{Coqsrc}
-    %  Lemma oplus_rect:
-    %       forall P: T1 -> T1 -> T1 -> Type, 
-    %         (forall a:T1, P zero a a) ->
-    %         (forall a: T1, P a zero a) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Gt ->
-    %            P b1 (cons a2 n2 b2) o ->
-    %            P (cons a1 n1 b1) (cons a2 n2 b2)
-    %              (cons a1 n1 o)) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Lt ->
-    %            P (cons a1 n1 b1) b2 o ->
-    %            P (cons a1 n1 b1) (cons a2 n2 b2) 
-    %            (cons a2 n2 o)) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Eq ->
-    %            P b1 b2 o ->
-    %           P (cons a1 n1 b1) (cons a2 n2 b2)
-    %             (cons a1 (S (n1 + n2)%nat) o)) ->
-    %          forall a b, P a b (oplus a b).
-    % Proof with auto.
-    %    induction a.
-    %    -    intro; simpl; destruct b;auto.
-    %    -   induction b.
-    %        + apply X0.
-    %        + case_eq (compare a1 b1).
-    %          * intro Comp; unfold oplus; rewrite Comp.
-    %            cbn; apply X3 ...
-    %          * intro Comp; cbn; rewrite Comp; apply X2...
-    %          * intro Comp; cbn; rewrite Comp ...
-    %  Defined.
-
-    % Ltac oplus_induction a b:= pattern (oplus a b); apply oplus_rect.
-    % \end{Coqsrc}
-
-    % \index{Exercises}
-
-    % \begin{exercise}
-    % The induction principle \texttt{oplus\_rect} is still unused in our development. 
-    % Please build some nice examples of application.
-    % \end{exercise}
-
-    \index{hydras}{Projects}
-    \begin{project}
-    Compare \texttt{oplus}'s definition (with inner fixpoint) with other possibilities
-    (\texttt{coq-equations}, \texttt{Function}, etc.).
-    \end{project}
-    \subsection{More theorems on Hessenberg's sum}
-
-    We need to prove some properties of $\oplus$, particularly about 
-    its relation with the order $<$ on \texttt{T1}.
-
-    \subsubsection{Commutativity, associativity}
-
-    We prove  the commutativity of $\oplus$ in two steps. 
-    First, we prove by transfinite induction on $\alpha$ that the restriction of $\oplus$ to the
-    interval $[0..\alpha)$ is commutative.
-
-    \index{maths}{Transfinite induction}
-
-    \input{movies/snippets/Hessenberg/oplusComm0}
-
-    Then, we infer  $\oplus$'s commutativity for any pair of ordinals:
-    Let $\alpha$ and $\beta$ be two ordinals strictly less than $\epsilon_0$. Both ordinals $\alpha$ and $\beta$ are
-    strictly less than $\textrm{max}(\alpha,\beta)+1$.
-        Thus, we have just to apply the lemma \coqsimple{oplus\_comm\_0}.
-
-\input{movies/snippets/Hessenberg/oplusComm}
-
-    Associativity of Hessenberg's sum is proved the same way.
-
-    \input{movies/snippets/Hessenberg/oplusAssoc0}
-    \input{movies/snippets/Hessenberg/oplusAssoc}
-
-    \subsubsection{Monotony}
-
-    At last, we prove that $\oplus$ is strictly monotonous in both of its arguments.
-
-    \input{movies/snippets/Hessenberg/oplusMono}
-
-      \index{hydras}{Projects}
-
-    \begin{project}
-    The library \texttt{Hessenberg} looks too long (proof scripts and compilation).
-    Please try to make it simpler and more efficient!
-    Thanks!
-    \end{project}
-
-    \paragraph*{\gaiasign} The module
-      \href{../theories/html/gaia_hydras.~GHessenberg.html}{gaia\_hydras.GHessenberg} defines a version of Hessenberg's sum
-      compatible with \gaia's type \texttt{T1}.
-
-    \subsection{A termination measure for hydra battles }
-
-    \label{sec:hydra-measure}
-
-    Let us define a measure from type \texttt{Hydra} into \texttt{T1}.
-
-    \vspace{4pt}
-    \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Termination.html\#m}{Hydra.Hydra\_Termination}}
-
-   \input{movies/snippets/Hydra_Termination/mDef}
-
-    First, we prove that the measure $m(h)$  of any hydra $h$ is a well-formed ordinal term of type \texttt{T1}.
-
-    \input{movies/snippets/Hydra_Termination/mNf}
-      \input{movies/snippets/Hydra_Termination/msNf}
-
-    For proving the termination of all hydra battles, we have to prove that
-    \texttt{m} is a variant. First, a few technical lemmas follow the decomposition of \texttt{round} into several relations. Then the lemma \texttt{round\_decr} gathers all the cases.
-
-    \label{sect:variant-decr}
-
-    \input{movies/snippets/Hydra_Termination/S0Decr}
-    \input{movies/snippets/Hydra_Termination/R1Decr}
-    \input{movies/snippets/Hydra_Termination/S1Decr}
-    \input{movies/snippets/Hydra_Termination/R2Decr}
-     \input{movies/snippets/Hydra_Termination/RoundDecr}
-
-      Finally, we prove termination of all (free) battles.
-
-    \label{thm:every-battle-terminates}
-
-    \input{movies/snippets/Hydra_Termination/FinalThm}
-
-    \paragraph*{\gaiasign}
-    The module~\href{../theories/html/gaia_hydras.GHydra.html}{gaia\_hydras.GHydra} contains a proof of \gaiaHydras version of termination of all battles.
-
-    \inputsnippets{GHydra/mDef, GHydra/mVariant, GHydra/Termination}
-
-    \section*{Conclusion}
-
-    Let us recall three results we have proved so far.
-    \begin{itemize}
-    \item There exists a strictly decreasing variant which maps \texttt{Hydra} into 
-    the segment $[0,\epsilon_0)$ for proving the termination of any hydra battle
-    \item There exists \emph{no} such variant from \texttt{Hydra} into 
-    $[0,\omega^2)$, \emph{a fortiori} into $[0,\omega)$.
-    \end{itemize}
-
-    So, a  natural question is `` Does there exist any strictly decreasing variant mapping
-    type \texttt{Hydra} into some interval $[0,\alpha)$ (where $\alpha <\epsilon_0$) for proving the termination of all hydra battles''.
-The next chapter is dedicated to a formal proof that there exists no such $\alpha$, even if we consider a restriction to the set of ``standard'' battles.
-
-    %\include{epsilon0}
-
-    %\include{impossibility-proofs}
-
-%-------------------------------------------------------------------
-
-\chapter[The Ketonen-Solovay machinery]{Accessibility inside \texorpdfstring{$\epsilon_0$}{Epsilon0}: The Ketonen-Solovay Machinery\label{ks-chapter}}
-\label{chap:ketonen}
-\index{maths}{Ordinal numbers!Ketonen-Solovay machinery}
-
-\section{Introduction}
-The reader may think that our proof of termination in the previous  chapter requires a lot of mathematical tools and may be too  complex. So, the question is ``is there  any  simpler proof'' ?
-
-In their article~\cite{KP82}, Kirby and Paris show that this result cannot be proved in Peano arithmetic.
-Their proof uses some knowledge about model theory and non-standard models of Peano arithmetic.
-In this chapter, we focus on a specific class of proofs of termination of hydra battles: construction of some variant mapping the type \texttt{Hydra} into a given initial  segment of ordinals.
-Our proof relies only on the Calculus of Inductive Constructions and is a natural complement of the results proven in the previous chapters.
-
-\begin{itemize}
-\item There is no variant mapping the type \texttt{Hydra} into the interval $[0,\omega^2)$ (section ~\vref{omega2-case}), and \emph{a fortiori} $[0,\omega)$ (section ~\vref{omega-case}).
-
-\item There exists a variant which maps the type \texttt{Hydra} into the
-interval $[0,\epsilon_0)$ (theorem \texttt{every\_battle\_terminates}, in section~\vref{thm:every-battle-terminates}).
-\end{itemize}
-
-Thus, a very natural question is the following one:
-\begin{quote}
-  `` Is there  any variant from
-\texttt{Hydra} into some interval $[0,\mu)$, where $\mu<\epsilon_0$, for proving the termination of all hydra battles ?''
-\end{quote}
-
-We prove in \coq{} the following result:
-
-\begin{quote}
-There is no variant for proving the termination of all hydra battles
-from \texttt{Hydra} into the interval $[0..\mu)$, where
-$\mu< \epsilon_0$.
-The same impossibility holds even if we consider only standard battles (with the successive replication factors $0,1,2,\dots,t,t+1,\dots$).
-\end{quote}
-
-Our proofs are  constructive and require no axioms: they are  closed terms of the CIC, and are mainly composed on function definitions and proofs of properties of these functions. 
-They  share much theoretical material with Kirby and Paris', although they do not use any knowledge about Peano arithmetic nor model  theory.  The combinatorial arguments we use and implement
-come from 
- an article by J.~Ketonen and R.~Solovay~\cite{KS81}, already  cited in the work
- by L.~Kirby and J.~Paris. % on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
- Section $2$ of this article: ''A hierarchy of probably recursive functions'', contains a systematic study of \emph{canonical sequences}, which are closely related to
-rounds of hydra battles. 
-Nevertheless, they have the same global structure as the simple proofs described in
-sections~\vref{omega-case} and \vref{omega2-case}. 
-We invite the reader to compare the three proofs step by step, lemma by lemma.
-
-\section{Canonical Sequences}
-\label{ketonen-solovay-sect}
-\index{maths}{Ordinal numbers!Canonical sequences}
-
-Canonical sequences are functions that associate an ordinal $\canonseq{\alpha}{i}$ to every ordinal $\alpha<\epsilon_0$ and positive integer $i$. They satisfy several nice properties:
-
-\index{maths}{Transfinite induction}
-\begin{itemize}
-\item If $\alpha\not=0$, then $\canonseq{\alpha}{i}<\alpha$. Thus canonical sequences can be used for proofs by transfinite induction or function definition by transfinite recursion
-\item If $\lambda$ is a limit ordinal, then $\lambda$ is the least upper bound of the set 
-$\{\canonseq{\lambda}{i}\;|\,i\in\mathbb{N}_1\}$
-
-\item If $\beta<\alpha<\epsilon_0$, then there is a ``path'' from $\alpha$ to $\beta$, \emph{i.e.} a
-sequence $\alpha_0=\alpha, \alpha_1, \dots, \alpha_n=\beta$, where for every $k<n$, there exists some $i_k$ such that $\alpha_{k+1}=\canonseq{\alpha_k}{i_k}$
-\item Canonical sequences correspond tightly to rounds of hydra battles: if $\alpha\not=0$,
-then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1})$ in one round with
-the replication factor $i$ (Lemma \href{../theories/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
-\item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) battle from $\iota(\alpha)$ to $\iota(\beta)$.
-\end{itemize}
-
-\begin{remark}
-  In~\cite{KS81}, canonical sequences are defined for any ordinal $\alpha <\epsilon_0$,
-by stating that if $\alpha$ is a successor ordinal $\beta+1$,  the sequence associated with 
-$\alpha$ is simply the constant sequence whose terms are equal to $\beta$.
-Likewise, the canonical sequence of $0$ maps any natural number to $0$.
-
-This convention allows us to make total the function that maps any ordinal $\alpha$ and natural number $i$ to the ordinal $\canonseq{\alpha}{i}$.
-
-\end{remark}
-
-First, let us recall how canonical sequences are defined in~\cite{KS81}.
-For efficiency's sake, we decided not to implement directly K.\&S's definitions, but to define in \gallina{} simply typed structurally recursive functions which share the abstract properties which are used in the mathematical proofs\footnote{With a small difference: the $0$-th term of the canonical sequence is not the same in our development as in~\cite{KS81}.}.
-
-\subsubsection{Mathematical definition of canonical sequences}
-
-In~\cite{KS81} the definition of $\canonseq{\alpha}{i}$ is based on the following remark:
-\begin{quote}
-Any non-zero ordinal $\alpha$ can be decomposed in a unique way as the product
-$\omega^\beta\times (\gamma+1)$.
-\end{quote}
-
-Thus the $\canonseq{\alpha}{i}$\,s are defined in terms of this decomposition:
-\begin{definition}[Canonical sequences: mathematical definition]
-\label{def:canonseq-math}
-
-\end{definition}
-\begin{mathframe}
-  \begin{itemize}
-\item Let $\lambda<\epsilon_0$ be a limit ordinal
-
-\begin{itemize}
-\item If $\lambda=\omega^{\alpha+1}\times (\beta+1)$, then 
-$\canonseq{\lambda}{i}= \omega^{\alpha+1}\times\beta +  \omega^\alpha \times i$
-\item If $\lambda=\omega^{\gamma}\times (\beta+1)$, where $\gamma<\lambda$ is a limit ordinal, then 
-$\canonseq{\lambda}{i}=\omega^{\gamma}\times \beta + \omega^{\canonseq{\gamma}{i}}$
-\end{itemize}
-
-\item For successor ordinals, we have $\canonseq{\alpha+1}{i}= \alpha$
-
-\item Finally, $\canonseq{0}{i}= 0$.
-\end{itemize}
-\end{mathframe}
-
-\subsubsection{Canonical sequences in \coq}
-\index{hydras}{Library Epsilon0!Functions!canon}
-\index{hydras}{Library Epsilon0!Functions!canonS}
-
-Our definition may look more complex than the mathematical one, but
-uses plain structural recursion over the type \coqsimple{T1}. Thus, tactics like
-\coqsimple{cbn}, \coqsimple{simpl}, \coqsimple{compute}, etc., are applicable.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canon}{Epsilon0.Canon}}
-
-\label{Functions:canonS}
-\label{Functions:canon}
-
-\input{movies/snippets/Canon/canonDef}
-
-\paragraph*{\gaiasign} Our library~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon} transposes the  function \texttt{canon} into \gaia's type \texttt{T1}
-(please see Sect.\~ref{sect:gcanon}).
-
-\paragraph*{Remark}
-In the present state of this library, the following specializations of \texttt{canon} are still used in some proofs or lemma statements. They are planned to be deprecated.
-
-\input{movies/snippets/Canon/CanonS0}
-
-For instance \coq's computing facilities allow us to verify the equalities\linebreak 
-\mathcolor{$\canonseq{\omega^\omega}{3} = \omega^3$} and
-\mathcolor{$\canonseq{\omega^\omega*3}{42} = \omega^\omega*2 + \omega^{42}$}.
-
-\input{movies/snippets/Canon/canonExamples}
-
-% \index{hydras}{Projects}
-% \begin{project}
-% Many lemmas presented in this chapter were stated and proved before the introduction of 
-% the type class \texttt{ON} of ordinal notations, and in particular its  instance \texttt{Epsilon0}.
-% Thus definitions and lemmas refer to the type \texttt{T1} of possibly not well-formed terms.
-% This should be fixed in  a future version.
-% \end{project}
-
-\subsection{Basic properties of canonical sequences}
-
-We did not  try to prove that our definition truly implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
-abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
-~\href{../theories/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}.
-For instance, the equality $\canonseq{\alpha+1}{i}=\alpha$  can be  proved by  structural induction on $\alpha$.
-
-\input{movies/snippets/Canon/canonSucc}
-
-\subsubsection{Canonical sequences and the order $<$}
-
-\index{maths}{Transfinite induction}
-
-We prove by transfinite induction over $\alpha$ that $\canonseq{\alpha}{i}$ is an ordinal strictly less than $\alpha$ (assuming $\alpha\not=0$).
-This property allows us to use the function \texttt{canonS} and its derivatives in function definitions by transfinite recursion.
-
-\label{lemma:canon_LT}
-
-\inputsnippets{Canon/canonLT}
-
-\paragraph*{\gaiasign} This lemma is also available in
-Library ~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon}:
-
-\inputsnippets{GCanon/gcanonLt}
-
-\subsubsection{Limit ordinals are truly limits}
-The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
-is the limit of the sequence $\canonseq{\lambda}{i}\;(1\le i)$.
-
-\vspace{4pt}
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}
-
-\input{movies/snippets/Canon/canonSLimitStrong}
-
-\label{lemma:canonS-limit}
-
-Note the use of \coq's \texttt{sig} type in the theorem's statement, which
-relates the boolean function \texttt{limitb} defined on the \texttt{T1} data-type with a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.
-We can also state directly that $\lambda$ is a (strict) least upper bound of the elements of its canonical sequence.
-
-\input{movies/snippets/Canon/canonSLimitLub}
-
-\paragraph*{\gaiasign}  In \gaiaHydras, the statement use a slightly different vocabulary:
-\inputsnippets{GCanon/gcanonLimitOf}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}\label{exo:simply-typed-canonseq}
-Instead of using the \texttt{sig} type, define a simply typed function that, given two ordinals $\alpha$ and $\beta$, returns a natural number $i$ such that, if $\alpha$ is a limit ordinal and $\beta<\alpha$, then $\beta< \canonseq{\alpha}{i+1}$.
-Of course, you will have to prove the correctness of your function.
-
-\textbf{Hint:} You may add to your function a third argument usually called \texttt{fuel} for allowing you to give a structurally 
-recursive function (\emph{cf} the post of Guillaume Melquiond on Coq-club (Dec 21, 2020)
-\url{https://sympa.inria.fr/sympa/arc/coq-club/2020-12/msg00069.html}).
-The type \texttt{fuel}, an alternative 
-to \texttt{nat} is available on \href{../theories/html/hydras.Prelude.Fuel.html}{Prelude.Fuel})
-.
-%\index{coq}{Giving fuel to a long computation}
-
-\end{exercise}
-
-\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0}\,: paths}
-\index{maths}{Ordinal numbers!Accessibility inside epsilon0}
-\label{sect:pathes-intro}
-
-Let us consider a kind of accessibility problem inside $\epsilon_0$: given two ordinals $\alpha$ and $\beta$, where $\beta<\alpha<\epsilon_0$, find a \emph{path} consisting of a finite sequence $\gamma_0=\alpha,\dots,\gamma_l=\beta$,
-where, for every $i<l$, $\gamma_i \not= 0$ \footnote{This condition allows us to ignore paths which end by a lot of useless $0$s.} and there exists some strictly positive integer $s_i$
-such that $\gamma_{i+1}=\canonseq{\gamma}{s_i}$.
-
-Let $s$ be the sequence $\langle s_0,s_1,\dots, s_{l-1} \rangle$. We describe the
-existence of such a path with the notation $\alpha\xrightarrow [s]{}\beta$.
-We say also that the considered path from $\alpha$ to $\beta$ \emph{starts at [index] $s_0$ and ends at $s_l$}.
-
-For instance, we have $\omega*2 \xrightarrow[2,2,2,4,5]{}3$, through the 
-path $\langle\omega\times 2, \omega+2,\omega+1,\omega,4,3\rangle$.
-
-\begin{remark}
-
-Note that, given $\alpha$ and $\beta$, where $\beta < \alpha$, the sequence $s$ which leads from $\alpha$ to $\beta$ is not unique.
-Indeed, if $\alpha$ is a limit ordinal, the first element of $s$ can be any integer $i$ such that $\beta<\canonseq{\alpha}{i}$, and if $\alpha$ is a successor ordinal,
-then the sequence $s$ can start with any positive integer.
-
-For instance, we have also 
-$\omega*2 \xrightarrow[3,4,5,6]{}\omega$. 
-Likewise,
-$\omega*2 \xrightarrow[1,2,1,4]{} 0$ and
-$\omega*2 \xrightarrow[3,3,3,3,3,3,3,3]{} 0$.
-\end{remark}
-
-\subsection{Formal definition}
-
-\label{path-to-definition}
-
-In \coq{}, the notion of path can be simply defined as an inductive predicate 
-parameterized by the destination $\beta$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\index{hydras}{Library Epsilon0!Predicates!path\_to}
-\label{sect:path-to-def}
-
-\input{movies/snippets/Paths/transitionDefs}
-
-\input{movies/snippets/Paths/pathDef}
-
-\begin{remark}
-In the present version of our library, we use a variant \texttt{path\_toS} of
-\texttt{path\_to}, where the proposition
-(\texttt{path\_toS $\beta$ $s$ $\alpha$}) is equivalent to
-(\texttt{path\_to $\beta$ (List.map S $s$) $\alpha$}). This variant is scheduled to be deprecated.
-\end{remark}
-
-\paragraph*{\gaiasign}
-The library~\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} transposes the  notion of path into \gaia's type \texttt{T1}
-(please see Sect.\~ref{sect:gpath}).
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Write a tactic for solving goals of the form (\texttt{path\_to $\beta$ $s$ $\alpha$})
-where $\alpha$, $\beta$ and $s$ are closed terms. 
-You should solve automatically the following goals:
-
-\inputsnippets{Paths/exPathTac} 
-\end{exercise}
-
-\subsection{Existence of a path}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction on $\alpha$, we prove that for any $\beta<\alpha$, 
-one can build a path from $\alpha$ to $\beta$ (in other terms, $\beta$ is accessible from $\alpha$).
-
-\input{movies/snippets/Paths/LTPathTo}
-
-\noindent 
-From the lemma \texttt{canon\_LT}~\vref{lemma:canon_LT}, we can convert any path into an inequality on ordinals (by induction on paths).
-
-\input{movies/snippets/Paths/pathToLT}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}[continuation of exercise~\vref{exo:simply-typed-canonseq}]
-Define a simply typed function for computing a path from $\alpha$ to $\beta$.
-\end{exercise}
-
-\subsection{Paths and hydra battles}
-\label{KS-o2h}
-
-In order to apply our knowledge about  ordinal numbers less than $\epsilon_0$ to the study of hydra battles, we define an injection
-from the interval $[0,\epsilon_0)$ into the type \texttt{Hydra}.
-
-\vspace{4pt}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}}
-
-\input{movies/snippets/O2H/iotaDef}
-
-For instance Fig.~\ref{fig:iota-example} shows the image by $\iota$ of the ordinal  \textcolor{black}{$\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$}
-
-  \begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.3]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (10,2) {$\bullet$};
-\node (N22) at (7,2) {$\bullet$};
-\node (N3) at (14,2) {$\bullet$};
-\node (N4) at (18,2) {$\Smiley[2][green]$};
-\node (N5) at (0,4) {$\bullet$};
-\node (N6) at (2,5) {$\Smiley[2][green]$};
-\node (N7) at (4,6) {$\Smiley[2][green]$};
-\node (N88) at (7,4) {$\bullet$};
-\node (N8) at (10,4) {$\bullet$};
-\node (N9) at (14,6) {$\Smiley[2][green]$};
-\node (N10) at (0,8) {$\Smiley[2][green]$};
-\node (N11) at (10,7) {$\Smiley[2][green]$};
-\node (N111) at (7,7) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N22);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (N1) to  (N5);
-\draw (N1) to   [bend left=10] (N6);
-\draw (N1) to   [bend right=20] (N7);
-\draw (N2) to  (N8);
-\draw (N22) to  (N88);
-\draw (N8) to  (N11);
-\draw (N88) to  (N111);
-\draw (N3) to  (N9);
-\draw (N5) to  (N10);
-\end{tikzpicture}
-\caption{The hydra $\iota(\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$) \label{fig:iota-example}}
-
-\end{figure}
-
-The following lemma (proved in ~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H.v}) maps  canonical sequences to rounds of hydra battles.
-
-\label{lemma:canonS-iota}
-
-\input{movies/snippets/O2H/canonSIota}
-
-The next step of our development extends this relationship to
-the order $<$ on $[0,\epsilon_0)$ on one side, and hydra battles on the other side.
-
-\input{movies/snippets/O2H/pathToRoundPlus}
-
-As a corollary, we are now able to transform any inequality $\beta<\alpha<\epsilon_0$ into a (free) battle.
-
-\input{movies/snippets/O2H/LTToRoundPlus}
-
-\section{A  proof of impossibility}
-
-We now have  the tools for proving that  there exists no variant bounded by some $\mu<\epsilon_0$ for proving the termination   of all battles.
-The proof we are going to show is a proof by contradiction.
-It  can
- be considered as a generalization of the
-proofs described in  sections~\vref{omega-case} and \vref{omega2-case}.
-
-In the module
-\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
-This part of the development is parameterized by some class $B$ of battles, which will be instantiated later to \texttt{free} or \texttt{standard}.
-
-\vspace{4pt}
-\noindent
-From~\href{../theories/html/hydras.Hydra.~Hydra_Definitions.html}{Hydra.Hydra\_Definitions}:
-
-\input{movies/snippets/Hydra_Definitions/BoundedVariant}
-
-Let us assume there exists such a variant:
-
-\input{movies/snippets/Epsilon0_Needed_Generic/theContext}
-
-\label{remark:m-decrease}
-\begin{remark}
-  The hypothesis \texttt{m\_decrease} is not provable  in general, but is satisfied by
-the  \texttt{free} and \texttt{standard} kinds of battles. This trick allows to 
-``factorize'' our proofs  of impossibility.
-\end{remark}
-
-\index{maths}{Transfinite induction}
-
-First, we prove that $m(\iota(\alpha))$ is always greater than or equal to $\alpha$, by  transfinite induction over $\alpha$.
-
-\input{movies/snippets/Epsilon0_Needed_Generic/mGe0}
-
-\begin{itemize}
-\item If $\alpha=0$, the inequality trivially holds
-\item If $\alpha$ is the successor of  some ordinal $\beta$, the inequality $\beta \leq m(\iota(\beta))$ holds (by induction hypothesis).
-But the hydra $\iota(\alpha)$ is transformed in one round into
-$\iota(\beta)$, thus $m(\iota(\beta))<m(\iota(\alpha))$. Hence $\beta<m(\iota(\alpha))$, which implies $\alpha \leq m(\iota(\alpha))$
-\item If $\alpha$ is a limit ordinal, then $\alpha$ is the least upper bound of the set
-of all  the $\canonseq{\alpha}{i}$.  Thus, we have just to prove that $\canonseq{\alpha}{i}< m(\iota(\alpha))$ for any $i$. 
-\begin{itemize}
-\item Let $i$ be some natural number.
-By the induction hypothesis, we have $\canonseq{\alpha}{i} \leq m(\iota(\canonseq{\alpha}{i}))$.
-But the hydra $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i})$ in one round, thus $m(\iota(\canonseq{\alpha}{i})) < m(\iota(\alpha))$, by our hypothesis \texttt{m\_decrease}.
-\end{itemize}
-\end{itemize}
-
-Please note that the impossibility proofs of 
-sections~\vref{omega-case} and \vref{omega2-case} contain a similar lemma, also called \texttt{m\_ge}.
-We are now able to build a counter-example.
-
-\input{movies/snippets/Epsilon0_Needed_Generic/mGeGeneric}
-
-The (big) rest of the proof is dedicated to prove formally the converse inequality 
-\texttt{m small\_h t1< m big\_h}.
-
-\subsection{The case of free battles}
-\label{sec:free-battles-case}
-Let us now consider that $B$ is instantiated to \texttt{free} (which means that we are considering proofs of termination of \emph{all} battles).
-The following lemmas are proved in Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
-The case $B=\texttt{standard}$ is studied in section~\vref{std-case}.
-
-\input{movies/snippets/Epsilon0_Needed_Free/theContext}
-
-\begin{enumerate}
-\item The following lemma is an application of \texttt{m\_ge\_generic}, since \texttt{free}
-satisfies trivially the hypothesis \texttt{m\_decrease} (see page~\pageref{remark:m-decrease}).
-
-\inputsnippets{Epsilon0_Needed_Free/mGe}
-
-\item From the hypothesis \texttt{Hy}, we have \texttt{m big\_h t1< mu}
-\item By Lemma \texttt{LT\_to\_round\_plus}, we get a (free) battle from
-\texttt{big\_h = iota mu} to \texttt{small\_h = iota (m big\_h)}.
-
-\input{movies/snippets/Epsilon0_Needed_Free/bigToSmall}
-
-\item From the hypotheses on $m$, we infer:
-\input{movies/snippets/Epsilon0_Needed_Free/mLt}
-
-\item From lemmas \texttt{m\_ge} and \texttt{m\_lt}, and the irreflexivity of $<$, we get a contradiction. 
-\input{movies/snippets/Epsilon0_Needed_Free/ImpossibilityFree}
-
-\end{enumerate}
-
-We have now proved there exists no bounded variant for the class of free battles.
-
-\input{movies/snippets/Epsilon0_Needed_Free/CheckDemo}
-
-\section{The case of standard battles}
-\label{sec:standard-intro}\label{std-case}
-One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
-Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}.
-
-\input{movies/snippets/O2H/LTToRoundPlus}
-
-This lemma builds a battle out of any inequality $\beta<\alpha$. 
-It is a straightforward application of \texttt{LT\_path\_to} of
-Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}:
-
-\input{movies/snippets/Paths/LTPathTo}
-
-The sequence $s$, used to build the sequence of replication factors of the battle, depends on 
-$\alpha$ and $\beta$, so we cannot be sure that the generated battle is a genuine standard battle.
-
-The solution to this issue comes  once again from Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
-$\alpha_0=\alpha,\alpha_1,\dots,\alpha_k=\beta$ where $\alpha_{j+1}$ is equal
-to $\canonseq{\alpha_j}{i_j}$ where $i_j$ is \emph{any} natural number, 
-we consider various constraints on these sequences.
-In particular, a path is called \emph{standard} if $i_{j+1} = i_j + 1$ for every $j<k$.
-It  corresponds to a ``segment'' of some standard battles. 
-Please note that the vocabulary on paths is ours, but all the concepts come really from~\cite{KS81}.
-
-In \coq{}, standard paths can be defined as follows.
-
-\vspace{4pt}
-
-\emph{From
-Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/standardPathR}
-
-In the mathematical text and figures, we shall use the notation 
-$\alpha \xrightarrow[i,j]{}\beta$ for the proposition 
-(\texttt{standard\_path $i$ $\alpha$ $j$ $\beta$}).
-In~\cite{KS81} the notation is
-$\alpha \xrightarrow[i]{*}\beta$
-for 
-the proposition  $\exists j, i<j \wedge \alpha \xrightarrow[i,j]{} \beta$.
-
-Our goal is now  to transform any inequality $\beta<\alpha<\epsilon_0$ into a standard path $\alpha \xrightarrow[i,j]{} \beta$ for some $i$ and $j$, then into a standard battle
-from $\iota(\alpha+i)$ to $\iota(\beta)$. 
-Following~\cite{KS81}, we proceed in two stages:
-\begin{enumerate}
-\item we simulate plain (free) paths from $\alpha$ to $\beta$ with
-paths made of steps $(\gamma,\canonseq{\gamma}{n})$, \emph{with the same $n$ all along the path}
-\item we simulate any such path by a standard path.
-\end{enumerate}
-
-\subsection{Paths with a constant index}
-
-First of all, paths with a constant index 
-enjoy nice properties. They are defined as paths where all the $i_j$ are equal to the same natural number $i$, for some $i>0$.
-
-Like in~\cite{KS81}, we shall use the notation $\alpha \xrightarrow[i]{} \beta$ for denoting such a path, also called an $i$-path.
-
-\input{movies/snippets/Paths/constPathDef}
-
-% Paths with a given index can be effectively computed.
-% Given $i$, $\alpha$ and $l$, the following function returns the ordinal $\beta$ such that there exists a path 
-% $\alpha \xrightarrow [i+1] {} \beta$ of length $l$.
-
-% \begin{Coqsrc}
-% Fixpoint const_funS (i:nat)(alpha : T1)(l:nat):  T1  :=
-%   match l
-%   with
-%   | 0 => alpha
-%   | S m => const_funS i (canonS i alpha) m
-%   end.
-% \end{Coqsrc}
-
-% The following computations show  applications of \texttt{constS\_fun} to the 
-% ordinal $\omega^\omega$, with various values of $i$ and $l$.
-
-% \begin{Coqsrc}
-% Compute  (const_funS 2 (omega ^omega)  55).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = zero
-%      : T1 
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (const_funS 2 (omega ^omega) 15).
-% \end{Coqsrc}
-
-%   \begin{Coqanswer}
-%  = (omega ^ 2 * 2)%pT1
-%      : ppT1   
-%   \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (const_funS 4 (omega^omega)  100).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% = (omega ^ 4 * 4 + omega ^ 3 * 4 + omega ^ 2 + omega * 4 + 4)%pT1
-%      : ppT1
-% \end{Coqanswer}
-
-A most interesting property of $i$-paths is that we can ``upgrade'' their index, as stated by K.\&S.'s Corollary 12.
-
-\index{maths}{Transfinite induction}
-
-\vspace{4pt}
-\noindent
-\emph{From
-  Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/Cor12}
-
-We  also use a version of \texttt{Cor12} with large inequalities.
-
-\input{movies/snippets/Paths/Cor121}
-
-\subsubsection{Sketch of proof of \texttt{Cor12}}
-\index{maths}{Transfinite induction}
-
-First, let us prove the following lemma (numbered \texttt{2.4} in Ketonen and Solovay's article).
-
-\input{movies/snippets/Paths/KSThm24}
-
-We prove this lemma by transfinite induction on $\alpha$.
-Let us consider a path $\alpha \xrightarrow [i]{} \beta$ $(i>0)$. Its first step is
-the pair $(\alpha,\canonseq{\alpha}{i})$, We have $\canonseq{\alpha}{i}<\alpha$ and
-$\canonseq{\alpha}{i} \xrightarrow [i]{} \beta$. 
-Let $n$ be any natural number such that $n>i$.
-By the induction hypothesis, there exists a path $\canonseq{\alpha}{n} \xrightarrow[i]{} \beta$.
-\begin{itemize}
-\item  If $\alpha$ is a successor ordinal $\gamma+1$, then $\canonseq{\alpha}{n} =
-\canonseq{\alpha}{i}=\gamma$. Thus we have a path 
-$\alpha  \xrightarrow [n]{}  \gamma \xrightarrow [n]{} \beta$
-\item If $\alpha$ is a limit ordinal, we apply Theorem \texttt{KS\_thm\_2\_4} (see above).
-
-%   \begin{theorem}
-% Let $\lambda$ be a limit ordinal, then for any pair of indices $0<i<j$, there is a path $\canonseq{\lambda}{j} \xrightarrow[1]{} \canonseq{\lambda}{i}$.    
-%   \end{theorem}
-
-The proof of the limit case, is decomposed into a sequence
-of path constructions leading to $\alpha \xrightarrow[n]{} \beta$.
-
- \begin{enumerate}
- \item $\alpha \xrightarrow[n]{} \canonseq{\alpha}{n}$ (single step path)
- \item $\canonseq{\alpha}{n} \xrightarrow[1]{} \canonseq{\alpha}{i}$ (by \texttt{Theorem\_2\_4}),
-\item $\canonseq{\alpha}{n} \xrightarrow[n]{} \canonseq{\alpha}{i}$ (applying the induction hypothesis to the preceding path);
-\item $\canonseq{\alpha}{i} \xrightarrow[n]{} \beta$ (applying the induction hypothesis)
-\item $\alpha \xrightarrow[n]{} \beta$ (by composition of 1, 3, and 4).
-
- \end{enumerate}
-
-\end{itemize}
-
-\begin{remark}
- \texttt{Cor12} ``casts'' $i$-paths into $n$-paths for any $n>i$.
-But the obtained $n$-path can be much longer than the original $i$-path.
-The following exercise will give an idea of this increase. 
-\end{remark}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-  Prove that  the length of the $i+1$-path from
-  $\omega^\omega$ to $\omega^i$ is $1 + (i+1)^{(i+1)}$, for any $i$. Note that the $i$-path from
-  $\omega^\omega$ to $\omega^i$ is only one step long.
- \end{exercise}
-
-Why is \texttt{Cor12} so useful? 
-Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
-we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying corollary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
-where $n$ is the maximum of the indices $i$ met in the induction.
-
- Lemma 1, Section 2.6 of~\cite{KS81} is naturally expressed in terms of \coq's
-\verb@sig@ construct.
-
-\label{lemma:L-2_6-1}
-%\index{coq}{Sigma types}
-
-\input{movies/snippets/Paths/Lemma261}
-
-Intuitively, Lemma~\texttt{2\_6\_1}  shows that if $\beta<\alpha<\epsilon_0$, then there exists  a battle from $\iota(\alpha)$ to $\iota(\beta)$ where the replication factor is constant, although large enough.
-
- \paragraph*{\gaiasign}
- Corollary \texttt{Cor12} and Lemma \texttt{Lemma2\_6\_1} are also available in
-\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} 
-(please see Sect.~\vref{sect:gpath}).
-
-\subsection{Casting paths with a constant index into a standard path}
-
-The article~\cite{KS81} contains 
-the following lemma, which allows us to simulate $i$-paths by $[i+1,j]$-paths, where $j$ is large enough.
-
-\input{movies/snippets/Paths/constantToStandard}
-
-\subsubsection{Sketch of proof of \texttt{constant\_to\_standard\_path}}
-
-Our proof follows the proof by Ketonen and Solovay, including its organization as a sequence of lemma.
-Since it is a non-trivial proof, we will comment its main steps below   (see Figure~\vref{fig:belle-preuve-1} to Figure~\vref{fig:fin-belle-preuve}).
-
-\subsubsection*{Preliminaries}
-
-Please note that, given an ordinal $\alpha:\texttt{T1}$, and two natural numbers $i$ and $l$, there exists at most a standard path $\alpha \xrightarrow [i,i+l]{*} \beta$.
-The following function computes $\beta$ from $\alpha$, $i$ and $l$.
-
-\input{movies/snippets/Paths/standardGnaw}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction over  $\alpha$, we prove that the ordinal $0$ is reachable from any ordinal $\alpha<\epsilon_0$ by some standard path.
-
-\input{movies/snippets/Paths/standardPathToZero}
-
-\paragraph*{}
-Now, let us consider two ordinals  $\beta<\alpha<\epsilon_0$.  Let $p$  be some $(n+1)$-path from $\alpha$ to $\beta$.
-
-\input{movies/snippets/Paths/ConstantToStandardProof}
-
-Applying \texttt{standard\_path\_to\_zero}, $0$ is reachable from $\alpha$ by some standard path.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \draw[->, very thick,blue] (alpha)-- node [below]{$n+1$} node [above] {$+$} (beta);
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-    \node (alpha3) at (20,5) {};
-  \node (alphalast) at (35,5) {};
-  \node (zero) at (45,0) {$0$};
-  \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate=40]{$n+1$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha2)-- node [below]{$n+3$}  (alpha3);
-
-  \node (dots) at (24,5) {$\dots$};
-  \draw [->, dashed, very thick,blue] (alphalast)-- node [below, rotate=-26]{$n+p+1$}  (zero);
-
-\end{tikzpicture}
-\caption{A nice proof (1)}
-  \label{fig:belle-preuve-1}
-\end{figure}
-
-\paragraph*{}
-
-Since comparison on \texttt{T1} is decidable, one can compute the last step $\gamma$ of the standard path from $(\alpha,n+1)$  such that $\beta\leq \gamma$.
-Let $l$ be the length of the path from $\alpha$ to $\gamma$.  
-This step of the proof is illustrated in figure~\vref{fig:belle-preuve-2}.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (17,5) {$\ldots$};
-    \node (alpha3) at (20,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (38,0) {$\delta$};
-    \draw [->, dashed,very thick,blue] (alpha)-- node [below,rotate=35]{$n+1$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below,rotate = -48]{\tiny $n+l$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{$n+l+1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{$n+1$} node [above] {$+$} (beta);
-   \draw[thick] (alpha)--  (gamma);
-   \draw[thick] (gamma)--  node [above] {$\geq$} (beta);
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-\end{tikzpicture}
-
-\caption{A nice proof (2)}
-  \label{fig:belle-preuve-2}
-\end{figure}
-
-\paragraph*{}
-
-\begin{itemize}
-\item If $\beta=\gamma$, it's OK! We have got a standard path
-from  
-$\alpha$ to $\beta$ with successive indices  $n+1, n+2, \dots, n+l+1$
-
-\item Otherwise,  $\beta < \gamma$.  Let us consider  $\delta=\canonseq{\gamma}{n+l+1}$.
-By applying several times lemma \texttt{Cor12},  one converts  every path of Fig~\ref{fig:belle-preuve-2} into
- a $n+l+1$-path  (see figure~\ref{fig:belle-preuve-3}).
-
-But $\gamma$ is on the $n+l+1$-path from $\alpha$ to $\beta$.
-As shown by figure~\vref{fig:fin-belle-preuve}, the ordinal $\delta$, reachable from
-$\gamma$ in one single step,  must be greater than or equal to $\beta$, which contradicts our  hypothesis $\beta < \gamma$.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-    \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (17,5) {$\ldots$};
-    \node (alpha3) at (20,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (38,0) {$\delta$};
-     \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -48]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\tiny $n+l+1$} node [above]{\color{red} $1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
-    \draw[thick] (gamma)--   node [above]{\color{red} $>$}(beta);
-   \draw[thick] (alpha)--  (gamma);
-
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-
-\end{tikzpicture}
-
-\caption{A nice proof (3)}
-  \label{fig:belle-preuve-3}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (15,5) {$\ldots$};
-    \node (alpha3) at (18,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (42,0) {$\delta$};
-    \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -36]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\small $n+l+1$} node [above]{\color{red} $1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
-   \draw[thick] (alpha)--  (gamma);
-
-    \draw[thick] (gamma)--  node [below]{\tiny $n+l+1$} node [above]{\color{red} $+$}(beta);
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-
-\end{tikzpicture}
-
-\caption{A nice proof (4)}
-  \label{fig:fin-belle-preuve}
-\end{figure}
-
-\end{itemize}
- The only possible case is  thus $\beta=\gamma$, so we have got a standard path  from $\alpha$ to $\beta$.
-
- \input{movies/snippets/Paths/constantToStandard0}
- \input{movies/snippets/Paths/constantToStandardz}
-
-Here is the full statement of the conversion from constant to standard paths.
-
-\input{movies/snippets/Paths/constantToStandardPath}
-
-Applying \texttt{Lemma2\_6\_1} and \texttt{constant\_to\_standard\_path}, we get the following corollary.
-
-\input{movies/snippets/Paths/LTToStandardPath}
-
-\subsection{Back to hydras}
-\label{sec:standard-battles-cases}
-We are now able to complete our proof that there exists no bounded variant for proving the termination of standard hydra battles. This proof can
-be consulted in the module 
-\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}{Hydra.Epsilon0\_Needed\_Std}.
-Please note that it has the same global structure as in section~\ref{sec:free-battles-case} 
-% ICI !
-Applying the  lemmas  \texttt{Lemma2\_6\_1} of the module 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
-{Epsilon0.pathS}   and 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
-{\texttt{constant\_to\_standard\_path}},
-we can convert any inequality $\beta<\alpha<\epsilon_0$ into a standard path from
-$\alpha$ to  $\beta$, then into a fragment of a standard battle from 
-$\iota(\alpha)$ to $\iota(\beta)$, hence the inequality $m(\iota(\beta))<m(\iota(\alpha))$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html\#LT_to_standard_battle}{Hydra.Epsilon0\_Needed\_Std}}
-
-\input{movies/snippets/Epsilon0_Needed_Std/LTToStandardBattle}
-
-Now, please consider the following context:
-
-\input{movies/snippets/Epsilon0_Needed_Std/theContext}
-
-In the same way as for free battles, we import a large inequality 
-from 
-the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Epsilon0\_Needed\_Generic}.
-
-\input{movies/snippets/Epsilon0_Needed_Std/mGe}
-
-\paragraph*{} If remains to prove the following strict inequality, in order to have a contradiction.
-
-\input{movies/snippets/Epsilon0_Needed_Std/mLt}
-
-\paragraph*{Sketch of proof:} Let us recall that $\texttt{big\_h} = \iota(\mu)$
- and $\texttt{small\_h} = \iota (m (\texttt{big\_h}))$.
-
-Since $m(\texttt{big\_h})< \mu$, there exists a standard path from $\mu$ to
-$m(\texttt{big\_h})$, hence a   standard battle from $\iota(\mu)$  to
-$\iota(m(\texttt{big\_h}))$,  i.e. from \texttt{big\_h} to \texttt{small\_h}.
-
-Since $m$ is assumed to be a variant for standard battles, we get the inequality  $m(\texttt{small\_h}) < m(\texttt{big\_h})$.
-
-\input{movies/snippets/Epsilon0_Needed_Std/endOfProof}
-
-\subsection{Remarks}
-
-We are grateful to 
- J. Ketonen and R. Solovay  for the high quality of their explanations and proof details.
-Our proof follows tightly the sequence of lemmas in their article, with a focus on 
-constructive aspects.
-Roughly speaking, our implementation \emph{builds}, out of a hypothetical
-  variant $m$, bounded by some ordinal $\mu<\epsilon_0$, a hydra \texttt{big\_h} which verifies the impossible inequality  $m(\texttt{big\_h})< m(\texttt{big\_h})$.
-
-On may ask whether the preceding results are too restrictive, since they 
-refer to a particular data type \texttt{T1}.
-In fact, our representation of ordinals strictly less than 
- $\epsilon_0$ is faithful to their mathematical definition, at least 
-Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
-(please see also the module
-\href{../theories/html/hydras.Schutte.Correctness_E0.html}{hydras.Schutte.Correctness\_E0}).
-
-Thus, we can infer that our theorems can be applied to any well order.
-
-\index{hydras}{Projects}
-\begin{project}
-Study a possible modification of the definition of a variant  (for  standard battles).
-
-\begin{itemize}
-\item The variant is assumed to be strictly decreasing \emph{on configurations 
-reachable from some initial configuration where the replication factor is equal to $0$}
-\item The variant may depend on the number of the current round.
-\end{itemize}
-
-In other words, its type should be \texttt{nat -> Hydra -> T1}, and it must 
-verify the inequality $m\, (S\,i)\, h' < m\,i\, h$ whenever the configuration 
-$(i,h)$ is reachable from some initial configuration $(0,h_0)$
-and \texttt{h} is transformed into \texttt{h'} in the considered round.
-Can we still prove the theorems of section~\ref{std-case} with this new definition?
-
-\end{project}
-
-%---------------------------------------------------------------------
 \chapter{Large sets and rapidly growing functions}\label{chap:alpha-large}
@@
 This proof uses properties of the Ackermann function, and the $H'_\alpha$, $F_\alpha$, $L_\alpha$ families of functions.
-
-% ---------------------------------------------------
-\input{gaia-chapter}
-
-%----------------------------------------------------------------------
-\chapter[Countable ordinals (after Sch\"{u}tte)]{Kurt Schütte's axiomatic definition of countable ordinals}
-
-\label{chap:schutte} 
-%ON
-
-In the present chapter, we  compare our implementation of the segment $[0,\epsilon_0)$ with a mathematical text in order to ``validate'' our constructions.
-Our reference here is the axiomatic definition of the set of countable ordinals,
-in chapter V of Kurt Schütte's book `` Proof Theory ''~\cite{schutte}.
-
-\begin{remark}
-\emph{In all this chapter, the word ``ordinal'' will be considered as a synonymous of
-``countable ordinal''}  
-\end{remark}
-
-Schütte's definition of countable ordinals relies on the following three axioms:
-
-There  exists a strictly ordered set , such that
-\begin{enumerate}
-\item  $(\mathbb{O},<)$ is well-ordered
-\item Every bounded subset of $\mathbb{O}$  is countable
-\item Every countable subset of $\mathbb{O}$  is bounded.
-\end{enumerate}
-
-Starting with these three axioms, Schütte re-defines the vocabulary about ordinal numbers: the null ordinal $0$, limits and successors, the addition of ordinals, the infinite ordinals $\omega$, $\epsilon_0$, $\Gamma_0$, etc.
-
-This chapter describes an adaptation to \coq{} of Schütte's axiomatization. 
- Unlike the rest of our libraries, our library
-\href{../theories/html/hydras.Schutte.Schutte.html}{hydras.Schutte}
-is not constructive, and relies on several axioms.
-
-\begin{itemize}
-\item First, please keep in mind  that the set of countable ordinals is not countable.
-Thus, we cannot hope to represent all countable ordinals as finite terms of an inductive type, which was possible with  the set of ordinals strictly less than $\epsilon_0$ (resp.
-$\Gamma_0$)
-\item We tried to be as close as possible to K.
-Schütte's text, which uses ``classical'' mathematics : excluded middle, Hilbert's $\epsilon$ (choice) and Russel's $\iota$ (definite description) operators.
-Both operators allow us to write definitions close to the natural mathematical language, such as ``$\textrm{succ}$ is \emph{the} least ordinal strictly greater than $\alpha$''
-\item Please note that only the library \href{../theories/html/hydras.Schutte.Schutte.html}{Schutte/*.v} is ``contaminated'' by axioms, and that the rest of our libraries remain constructive.
-\end{itemize}
-
-\section{Declarations and axioms}
-
-Let us declare a type 
-\texttt{Ord} for representing countable ordinals, and a binary relation
- \texttt{lt}. Note that, in our development, \texttt{Ord} is a type, while the \emph{set} of countable ordinals (called $\mathbb{O}$ by Schütte) 
-is the full set over the type \texttt{Ord}.
-
-\label{types:Ord}
-
-We use Florian Hatat's library on countable sets, written as he was a student of  \emph{\'Ecole Normale Supérieure de Lyon}.
-A set $A$ is countable if there is an injective function from $A$ to $\mathbb{N}$ (see
-Library \href{../theories/html/hydras.Schutte.Countable.html}%
-{\texttt{Schutte.Countable}}).
-
-\vspace{6pt}
-
-\emph{From Module\href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{Schutte.Schutte\_basics}}}
-
-\index{hydras}{Library Schutte!Types!Ord}
-
-\input{movies/snippets/Schutte_basics/OrdDecl}
-
-Schütte's first axiom tells that \texttt{lt} is a well order on the set 
-\texttt{ordinal} (The  class \texttt{WO} is defined in
-Module~\href{../theories/html/hydras.Schutte.Well_Orders.html}{Schutte.Well\_Orders.v}).
-
-\index{hydras}{Library Schutte!Type classes!WO@ WO (well order)}
-
-\label{types:WO}
-
-\input{movies/snippets/Well_Orders/Mdecl}
-
-\input{movies/snippets/Well_Orders/WODef}
-
-\input{movies/snippets/Schutte_basics/AX1}
-
-The second and third axioms say that a subset $X$ of $\mathbb{O}$ is
-(strictly) bounded if and only if it is countable.
-
-\input{movies/snippets/Schutte_basics/AX23}
-
-\texttt{AX2} and \texttt{AX3} could have been replaced by a single axiom (using the \texttt{iff} connector), but we decided to respect as most as possible the structure of Schütte's definitions.
-
-\section{Additional  axioms}
-
-The adaptation of Schütte's mathematical discourse to \coq{} led us to
-import a few axioms from the standard library. We encourage the reader to consult \coq{}'s FAQ about the safe use of axioms
- \url{https://github.com/coq/coq/wiki/The-Logic-of-Coq#axioms}.
-
-\subsubsection{Classical logic}
-
-In order to work with classical logic, we import the module
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Classical.html}{Coq.Logic.Classical}  of \coq{}'s standard library, specifically the following axiom:
-
-\begin{Coqsrc}
- Axiom classic : forall P:Prop, P \/ ~P.
-\end{Coqsrc}
-
-\subsubsection{Description operators}
-
-In order to respect Schütte's style, we imported also the library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Epsilon.html}{\texttt{Coq.Logic.Epsilon}}.  The rest of this section presents a few examples of
-how Hilbert's choice operator and Church's definite description allow us
- to write understandable definitions (close to the mathematical natural language).
-
-\subsubsection{The definition of zero}
-
-According to the  definition of a well order, every non-empty subset of \texttt{Ord} has a least element. Furthermore, this least element is unique. We would like to call this element  \texttt{zero}.
-
-\input{movies/snippets/Schutte_basics/iotaDemoa}
-\input{movies/snippets/Schutte_basics/iotaDemob}
-
-Indeed, the basic logic of  \coq{} does not allow us to eliminate a proof of a proposition 
-$\exists!\,x:A,\,P(x)$ for building a term whose type lies in the sort \texttt{Type}. 
-The reasons for this impossibility are explained in many documents~\cite{BC04, chlipalacpdt2011, Coq}.
-
-Let us import the library \texttt{Coq.Logic.Epsilon}, which contains the following axiom and lemmas.
-
-\input{movies/snippets/MoreEpsilonIota/EpsilonStatement}
-
-Hilbert's $\epsilon$ \emph{operator} is derived from this  axiom.
-
-\input{movies/snippets/MoreEpsilonIota/EpsilonDef}
-
-If we consider the \emph{unique existential} quantifier $\exists!$, we obtain
-Church's \emph{definite description operator}.
-
-\input{movies/snippets/MoreEpsilonIota/iotaDef}
-
-Indeed, the operators \texttt{epsilon} and \texttt{iota} allowed us to make our definitions 
-quite close to Schütte's text. Our libraries \href{../theories/html/hydras.Schutte.MoreEpsilonIota.html}%
-{\texttt{Schutte.MoreEpsilonIota}}
-and
-\href{../theories/html/hydras.Schutte.PartialFun.html}%
-{\texttt{Schutte.PartialFun}} are extensions of \texttt{Coq.logic.Epsilon} for making easier 
-such definitions. See also an article in french~\cite{PCiota}.
-
-\input{movies/snippets/MoreEpsilonIota/Defs}
-
-In order to use these tools,  we have to tell \coq{}  that the declared type \texttt{Ord} is not empty:
-
-\input{movies/snippets/Schutte_basics/inhOrd}
-
-We are now able to define \texttt{zero} as the least ordinal. For this purpose,
-we define a function returning the least element of any [non-empty]  subset.
-
-\emph{From Module\href{../theories/html/hydras.Schutte.Well_Orders.html}%
-  {\texttt{Schutte.Well\_Orders}}}
-
-\input{movies/snippets/Well_Orders/MDecl}
-\input{movies/snippets/Well_Orders/theLeast}
-
-\vspace{4pt}
-
-From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{~Schutte.Schutte\_basics}}
-
-\label{Constants:zero:Ord}
-\index{hydras}{Library Schutte!Constants!zero}
-
-\input{movies/snippets/Schutte_basics/zeroDef}
-
-We want to prove now that zero is less than or equal to any ordinal number.
-
-\input{movies/snippets/Schutte_basics/zeroLe}
-
-\subsubsection{Remarks on \texttt{epsilon} and \texttt{iota}}
-
- What would happen in case of a misuse of \texttt{epsilon} or \texttt{iota} ?
-For instance, one could give a unsatisfiable specification to \texttt{epsilon} or 
-a specification for \texttt{iota} that admits several realizations.
-
-Let us consider an example:
-
-\input{movies/snippets/Schutte_basics/BadBottoma}
-
-Since we won't be able to prove the proposition
-\linebreak \Verb|(exists! a: Ord, least_member (Empty_set Ord) a)|, the only properties we would be able to prove about \texttt{bottom} would be \emph{trivial} properties, 
-\emph{i.e.}, satisfied by \emph{any} element of type \texttt{Ord}, like for instance
-\texttt{bottom = bottom}, or \texttt{zero <= bottom}.
-
-\input{movies/snippets/Schutte_basics/trivialProps}
-
-On the other hand, the following attempt fails, because of the unprovable first subgoal (please notice that the second subgoal is easy to solve !).
-
-\input{movies/snippets/Schutte_basics/Failure}
-
-In short, using \texttt{epsilon} and \texttt{iota} in our implementation of countable ordinals after Schütte has two main advantages.
-
-\begin{itemize}
-\item It allows us to give a \emph{name} (using \texttt{Definition}) two witnesses 
-of existential quantifiers (let us recall that, in classical logic, one may consider non-constructive proofs of existential statements)
-\item By separating definitions from proofs of [unique] existence, one may make definitions  more concise and readable. Look for instance at 
-the definitions of  \texttt{zero}, \texttt{succ}, \texttt{plus}, etc. in the rest of this chapter.
-\end{itemize}
-%%%% ICI ICI
-
-\section{The  successor function}
-
-The definition of the function \texttt{succ:Ord -> Ord} is very concise. The successor of any ordinal $\alpha$ is the smallest ordinal strictly greater than $\alpha$.
-
-\label{Functions:succ-sch}
-\index{hydras}{Library Schutte!Functions!succ}
-
-\input{movies/snippets/Schutte_basics/succDef}
-
-Using \texttt{succ}, we define the following predicates.
-
-\input{movies/snippets/Schutte_basics/isSuccIsLimit}
-
-% \begin{remark}
-% Please look at remark~\vref{warning:coercions}.
-
-How do we prove properties of the successor function?
-First, we make its specification explicit.
-
-\input{movies/snippets/Schutte_basics/succSpec}
-
-Then, we prove that our function \texttt{succ} meets this specification.
-
-\input{movies/snippets/Schutte_basics/succOka}
-
-We have now to prove that the set of all ordinals strictly greater than $\alpha$ has a unique least element.
-But the singleton set $\{\alpha\}$ is countable, hence  bounded (by the axiom \texttt{AX3}).
-Hence; the set $\{\beta\in\mathbb{O}|\alpha < \beta\}$ is not empty
-and therefore has a unique least element.
-
-The rest of the \coq{} proof script is quite short.
-
-\input{movies/snippets/Schutte_basics/succOkb}
-
-We can ``uncap'' the description operator for proving properties of the
-\texttt{succ} function.
-
-\input{movies/snippets/Schutte_basics/succProps}
-
-\section{Finite ordinals}
-
-Using \texttt{succ}, it is now easy to define recursively all the finite ordinals.
-
-\label{sect:notation-F-sch}
-
-\input{movies/snippets/Schutte_basics/finiteDef}
-
-\section{The definition of \texttt{omega}}
-In order to define $\omega$, the first infinite ordinal, we use an operator which
-``returns'' the least upper bound (if it exists) of a subset $X\subseteq \mathbb{O}$.
-For that purpose, we first use a predicate:
-(\texttt{is\_lub $D$ \textit{lt} $X$ $a$}) if $a$ belongs to $D$ and is the least 
-upper bound  of $X$ (with respect to \textit{lt}).
-
-\input{movies/snippets/Lub/isLubDef}
-
-\input{movies/snippets/Schutte_basics/supDef}
-
-Then, we define the function \texttt{omega\_limit} which returns the least upper bound 
-of the  (denumerable) range of any sequence \texttt{s: nat -> Ord}. 
-By \texttt{AX3} this range is bounded, hence the set of its upper bounds is not empty and has a least element.
-Then we define \texttt{omega} as the limit of the sequence of finite ordinals.
-
-\input{movies/snippets/Schutte_basics/omegaDef}
-
-\label{sect:notation-omega}
-
-Among the numerous properties of the ordinal $\omega$, let us quote the following ones
-(proved in Module 
-\href{../theories/html/hydras.Schutte.Schutte_basics.html\#finite_lt_omega}{\texttt{Schutte.Schutte\_basics}})
-
-\input{movies/snippets/Schutte_basics/omegaPropsa}
-\input{movies/snippets/Schutte_basics/omegaPropsb}
-\input{movies/snippets/Schutte_basics/omegaPropsc}
-\input{movies/snippets/Schutte_basics/omegaPropsd}
-
-\subsection{Ordering functions and ordinal addition}
-
-After having defined the finite ordinals and the infinite ordinal $\omega$, we  define the sum $\alpha+\beta$ of two countable ordinals.
-Schütte's definition looks like the following one:
-
-\begin{quote}
-``$\alpha+\beta$ is the $\beta$-th ordinal greater than or equal to $\alpha$''
-\end{quote}
-
-The purpose of this section is to give a meaning to the construction
-``the $\alpha$-th element of $X$''  where $X$ is any non-empty subset of $\mathbb{O}$.
-We follow Schütte's approach, by defining the notion of \emph{ordering functions},
-a way to associate a unique ordinal to each element of $X$.
-Complete definitions and proofs can be found in Module
- \href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
-{\texttt{Schutte.Ordering\_Functions}} ).
-
-\subsection{Definitions}
-
-A \emph{segment} is a set $A$ of ordinals such that, whenever  $\alpha\in A$ and
-$\beta<\alpha$, then $\beta\in A$; a segment is  \emph{proper} if it strictly included in $\mathbb{O}$.
-
-\input{movies/snippets/Ordering_Functions/segmentDef}
-
-Let  $A$ be a segment, and $B$ a subset of $\mathbb{O}$ : an \emph{ordering function for $A$ and  $B$} is a strictly increasing bijection from $A$ to $B$.
-The set $B$ is said to be an \emph{ordering segment} of $A$.
-Our definition in \coq{} is a direct translation of the mathematical text of~\cite{schutte}.
-
-\index{maths}{Ordinal numbers!Ordering functions}
-\index{hydras}{Library Schutte!Predicates!ordering function@ordering\_function}
-
-\input{movies/snippets/Ordering_Functions/orderingFunctionDef}
-
-We are now able to associate with any subset $B$ of $\mathbb{O}$ its ordering segment and ordering function.
-
-\input{movies/snippets/Ordering_Functions/ordDef}
-
-Thus (\texttt{ord $B \;\alpha$}) is the $\alpha$-th element of $B$.
-Please note that the last definition uses the epsilon-based operator \texttt{some} and
-not \texttt{the}. This is due to the fact that we cannot prove the unicity (w.r.t. Leibniz' equality) of the ordering function of a given set. 
-By contrast, we admit the axiom  \texttt{Extensionality\_Ensembles}, from the library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Sets.Ensembles.html}{Coq.Sets.Ensembles}, so we use the operator \texttt{the} in the definition of
-\texttt{the\_ordering\_segment}.
-
-One of the main theorems of
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html\#ordering_function_ex}%
-{\texttt{Ordering\_Functions}} 
-associates a unique segment and a unique (up to extensionality) ordering function to every subset $B$ of $\mathbb{O}$.
-
-\input{movies/snippets/Ordering_Functions/orderingFunctionEx}
-
-Thus,  our function \texttt{ord}  which enumerates the elements of $B$ is defined in a non-ambiguous way.
-Let us quote the following theorems (see Library
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
-{\texttt{Schutte.Ordering\_Functions}} for more details).
-
-\input{movies/snippets/Ordering_Functions/orderingLe}
-\input{movies/snippets/Ordering_Functions/Th1352}
-
-\subsection{Ordinal addition}
-
-We are now ready to define and study addition on the type \texttt{Ord}.
-The following definitions and proofs can be consulted in Module
-\href{../theories/html/hydras.Schutte.Addition.html}%
-{\texttt{Schutte.Addition.v}}.
-
-\index{hydras}{Library Schutte!Functions!plus}
-
-\input{movies/snippets/Addition/additionDef}
-
-In other words,  $\alpha + \beta$ is the  $\beta$-th ordinal greater than or equal to $\alpha$. 
-Thanks to generic properties of ordering functions, we can show the following 
-properties of addition on $\mathbb{O}$. First, we prove a useful lemma:
-
-\input{movies/snippets/Addition/plusElim}
-
-As a use-case, let us prove that $0$ is a right neutral element of $+$.
-
-\input{movies/snippets/Addition/alphaPlusZero}
-
-The following lemmas are proved the same way.
-
-\input{movies/snippets/Addition/bunchOfLemmas}
-
-The following lemmas are not direct applications of \texttt{plus\_elim}.
-
-\input{movies/snippets/Addition/plusAssoc}
-\input{movies/snippets/Addition/finitePlusInfinite}
-
-It is interesting to compare the proof of these lemmas with the
-computational proofs of the corresponding statements in Module
-\href{../theories/html/hydras.Epsilon0.T1.html}%
-{\texttt{Epsilon0.T1}}. 
-For instance, the proof of the lemma 
-\texttt{one\_plus\_omega} uses the continuity of ordering functions (applied to  \texttt{(plus 1)}) and compares the limit of the $\omega$-sequences $i_{(i \in \mathbb{N})}$ and
-$(1+i)i_{(i \in \mathbb{N})}$, whereas in the library  \texttt{Epsilon0/T1}, the equality 
-$1+\omega=\omega$ is just proved with \texttt{reflexivity}!
-
-\subsubsection{Multiplication by a natural number}
-
-The multiplication of an ordinal by a natural number is defined in terms of addition.
-This operation is useful for the study of Cantor normal forms.
-
-\input{movies/snippets/Addition/multFin}
-
-\section{The exponential of basis \texorpdfstring{$\omega$}{omega}}
-
-In this section, we define the function which maps any $\alpha\in\mathbb{O}$ to
-the ordinal  $\omega^\alpha$, also written 
-$\phi_0(\alpha)$. 
-It is an opportunity to apply the definitions and results of the preceding section. 
-Indeed,  Schütte first defines a subset of $\mathbb{O}$: the set of additive principal ordinals, and $\phi_0$  is just defined as the ordering function of this set.
-
-\subsection{Additive principal ordinals}
-
-\index{maths}{Ordinal numbers!Additive principal ordinals}
-\index{hydras}{Library Schutte!Predicates!AP@AP (additive principal ordinals)}
-
-\begin{definition}
-A non-zero ordinal  $\alpha$ is said to be \emph{additive principal} if, for all  $\beta<\alpha$, $\beta+\alpha$ is equal to  $\alpha$.
-We call \texttt{AP} the set of additive principal ordinals.
-
-\end{definition}
-
-\noindent\emph{From Module \href{../theories/html/hydras.Schutte.AP.html}%
-{\texttt{Schutte.AP}}}
-
-\input{movies/snippets/AP/APDef}
-
-\subsection{The function \texttt{phi0}}
-
-Let us call  $\phi_0$ the ordering function of \texttt{AP}.
-In the mathematical text, we shall use indifferently the notations  $\omega^\alpha$ and$\phi_0(\alpha)$.
-
-\index{hydras}{Library Schutte!Functions!phi0}
-
-\input{movies/snippets/AP/phi0Def}
-
-\subsection{Omega-towers and the ordinal \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-Using $\phi_0$, we can define recursively the set of finite omega-towers.
-
-\input{movies/snippets/AP/omegaTower}
-
-\label{sect:epsilon0-as-limit}
-Then, the ordinal  $\epsilon_0$ is defined as the limit of the sequence of all finite towers (a kind of infinite tower).
-
-\input{movies/snippets/AP/epsilon0Def}
-
-The rest of our library \texttt{AP} is devoted to the proof of properties of additive principal ordinals, hence of the ordering function  $\phi0$ and the ordinal $\epsilon_0$ (which we could not express within the type \texttt{T1}).
-
-\subsection{Properties of the set  \texttt{AP}}
-
-The set of additive principal ordinals is not empty: it contains at least the ordinals  $1$ and  $\omega$.
-
-\inputsnippets{AP/APOne,AP/APOne_least_AP,AP/APOne_AP_omega,AP/APOne_omega_second_AP}
-
-The set  \texttt{AP} is  \emph{closed} under addition, and unbounded.
-\label{lemma:AP-plus-closed}
-
-\input{movies/snippets/AP/APPlusClosed}
-
-\input{movies/snippets/AP/APUnbounded}
-
-Finally, \texttt{AP} is (topologically) \emph{closed} and ordered by the segment of all countable ordinals.
-
-\index{hydras}{Library Schutte!Predicates!Closed}
-
-From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{~Schutte.Schutte\_basics}}
-
-\input{movies/snippets/Schutte_basics/ClosedDef}
-\input{movies/snippets/AP/APClosed}
-
-\subsubsection{Properties of the function \texorpdfstring{$\phi_0$}{phi0}}
-
-The ordering function $\phi_0$ of the set \texttt{AP} is defined on the full set $\mathbb{O}$ and is continuous (Schütte calls such a  function  \emph{normal}).
-
-\begin{Coqsrc}
-Theorem normal_phi0 : normal phi0 AP.
-\end{Coqsrc}
-
-The following properties come from  the definition of $\phi_0$ as the ordering function of \texttt{AP}.
-It may be interesting to compare these proofs with the computational ones described in Chapter ~\ref{chap:T1}.
-
-\input{movies/snippets/AP/APPhi0}
-
-\subsection{A last example}
-\label{Ex42-schutte}
-
-Let us prove again the equality $\omega+42+\omega^2= \omega^2$.Let us recall that $\omega^2$ is an abbreviation of $\phi_0(2)$,
-\emph{i.e} the third  additive principal ordinal.
-
-\inputsnippets{Schutte/Ex42a}
-
-Our proof is very different from the computational proof of Sect~\vref{Ex42-E0}.
-By definition of additive principal ordinals, 
-it suffices to prove the inequality $\omega+42< \phi_0(2)$.
-
-\inputsnippets{Schutte/Ex42b}
-
-Since the set \texttt{AP} of additive principals  is closed under addition
-(by Lemma \texttt{AP\_plus\_closed}, page~\pageref{lemma:AP-plus-closed}) , it suffices to prove the inequalities $\omega<\phi_0(2)$ and $42<\phi_0(2)$.
-
-\inputsnippets{Schutte/Ex42d, Schutte/Ex42c, Schutte/Ex42e}
-
-\section{More about \texorpdfstring{$\epsilon_0$}{\texttt{epsilon0}}}
-
-Let us recall that the limit ordinal  $\epsilon_0$ cannot be written within the type \texttt{T1}.
-Since we are now considering the set of all countable ordinals, we can now prove some properties of this ordinal.
-
-We prove the inequality  $\alpha<\omega^\alpha$ whenever $\alpha < \epsilon_0$.
-\emph{Note that this condition was implicit in Module
-\href{../theories/html/hydras.Epsilon0.T1.html\#lt_phi0}{Epsilon0.T1}.}
-
-\input{movies/snippets/AP/ltPhi0}
-
-The proof is as follows:
-\begin{enumerate}
-\item Since $\alpha<\epsilon_0$, consider the least $i$ such that $\alpha$ is strictly less than the omega-tower of height $i$.
-\item
-  \begin{itemize}
-  \item If $i=0$, then the result is trivial (because $\alpha=0$)
- \item  Otherwise let $i=j+1$; 
-          $\alpha$ is greater than or equal to the omega-tower of height $j$.
-         By monotony,  $\phi_0(\alpha)$ is greater than or equal to 
-        the omega-tower of height $j+1$, thus strictly greater than $\alpha$
-  \end{itemize}
- \end{enumerate}
-
-Moreover,  $\epsilon_0$ is the least ordinal $\alpha$ that verifies the equality 
-$\alpha = \omega^\alpha$, in other words, the least fixpoint of the function  $\phi_0$.
-
-\input{movies/snippets/AP/epsilon0Lfp}
-
-\section{Critical ordinals}
-
-\index{maths}{Ordinal numbers!Critical ordinals}
-\index{hydras}{Library Schutte!Predicates!Cr@Cr (critical ordinals)}
-
-For any  (countable) ordinal $\alpha$, the set $\textit{Cr}(\alpha)$ is inductively defined 
-as follows by Schütte (p.81 of~\cite{schutte}).
-
-\begin{quote}
-  \begin{itemize}
-  \item $\textit{Cr}(0)$ is the set \textit{AP} of additive principal ordinals.
-  \item If $0<\alpha$, then $\textit{Cr}(\alpha)$ is the intersection of all the sets of fixpoints of the $\textit{Cr}(\beta)$ for $\beta<\alpha$.
-  \end{itemize}
-\end{quote}
-
-This definition is translated in \coq{} in 
-Module \href{../theories/html/hydras.Schutte.Critical.html}%
-{\texttt{Schutte.Critical}}, as the least fixpoint of a functional.
-
-\input{movies/snippets/Critical/CrDef}
-
-Lets us denote by $\phi_\alpha$ the ordering function of the set $\textit{Cr}(\alpha)$ and by $A_\alpha$ its ordering segment.
-
-\input{movies/snippets/Critical/phiDef}
-
-\label{sect:phi-schutte}
-
-For instance,  we prove that $\textit{Cr}(0)$ is the set of additive principals and that $\epsilon_0$
-belongs to $\textit{Cr}(1)$.
-
-\inputsnippets{Critical/CrZeroAP,
-  Critical/epsilon0Cr1}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
- Prove that $\epsilon_0$ is the least element of $\textit{Cr}(1)$.
-\end{exercise}
-
-\subsection{A flavor of infinity}
-
-The family of the $\textit{Cr}(\alpha)$s is made of infinitely many unbounded (hence infinite) sets.
-Let us quote Lemma 5, p. 82  of~\cite{schutte}:
-\begin{quote}
-  For all $\alpha$, the set $\textit{Cr}(\alpha)$ is closed (for the least upper bound of non-empty countable sets) and unbounded.
-\end{quote}
-
-We prove this result by a transfinite induction on $\alpha$ of the conjunction of  both properties.
-
-\index{maths}{Transfinite induction}
-
-\input{movies/snippets/Critical/Lemma5}
-
-\section{Cantor normal form}
-
-The notion of Cantor normal form is defined for all countable ordinals.
-Nevertheless, note that, contrary to the implementation based on type \texttt{T1},
-the Cantor normal form of an ordinal $\alpha$ may contain $\alpha$ as a 
-sub-term\footnote{This would prevent us from trying to represent Cantor normal forms as finite trees (like in Sect.~\ref{sec:T1-inductive-def})}.
-
-\index{maths}{Ordinal numbers!Cantor normal form}
-\index{hydras}{Library Schutte!Predicates!is\_cnf\_of@is\_cnf\_of (to be a Cantor normal form of}
-
-We represent  Cantor normal forms as lists of ordinals.
-A  list $l$ is a Cantor normal form of a given ordinal $\alpha$ if it satisfies two conditions:
-
-\begin{itemize}
-\item The list  $l$ is sorted (in decreasing order) w.r.t. the order $\leq$
-\item The sum of all the  $\omega^{\beta_i}$ where the $\beta_i$ are the terms of $l$ (in this order) is equal to $\alpha$.
-\end{itemize}
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Schutte.CNF.html\#cnf_t}%
-{\texttt{Schutte.CNF}}}
-
-\input{movies/snippets/CNF/Defs}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction on $\alpha$, we prove that every countable ordinal $\alpha$ 
- has at least a Cantor normal form.
-
-\input{movies/snippets/CNF/cnfExists}
-
-By structural induction on lists, we prove that this normal form is unique.
-
-\inputsnippets{CNF/cnfUnicity,CNF/cnfExUnique}
-
-Finally, the following two lemmas relate  $\epsilon_0$ with Cantor normal forms.
-
-If $\alpha<\epsilon_0$, then the Cantor normal form of $\alpha$ is made of ordinals strictly less than $\alpha$.
-
-\input{movies/snippets/CNF/cnfLtEpsilon0}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Please consider the following statement :
-
-\begin{Coqsrc}
-Lemma cnf_lt_epsilon0_iff : 
- forall l alpha, 
-   is_cnf_of alpha l ->  
-   (alpha < epsilon0 <->  Forall (fun beta =>  beta < alpha) l).
-\end{Coqsrc}
-
-Is it true ?
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/schutte_cnf_counter_example.v}{exercises/ordinals/schutte\_cnf\_counter\_example.v}.}
-\end{exercise}
-
-Finally, the Cantor normal form of $\epsilon_0$ is just $\omega^{\epsilon_0}$.
-
-\input{movies/snippets/CNF/cnfOfEpsilon0}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Implement pages 82 to 85 of~\cite{schutte} (critical, strongly critical, maximal critical ordinals, Feferman's ordinal $\Gamma_0$).
-\end{project}
-
-\begin{remark}
-The sub-directory
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/Gamma0}{theories/oridinals/Gamma0} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form.
-\end{remark}
-
-\section{An embedding of \texttt{T1} into \texttt{Ord}}
-
-Our library 
-\href{../theories/html/hydras.Schutte.Correctness_E0.html}%
-{\texttt{Schutte.Correctness\_E0}} establishes the link between two very different modelizations of ordinal numbers. In other words, it ``validates'' a data structure in terms of
-a classical mathematical discourse considered as a model. 
-First, we define a function from \texttt{T1} into  \texttt{Ord} by structural recursion.
-
-\input{movies/snippets/Correctness_E0/injectDef}
-
-This function enjoys good commutation properties with respect to the main operations which
-allow us to build Cantor normal forms.
-
-\input{movies/snippets/Correctness_E0/commutationLemmas}
-\input{movies/snippets/Correctness_E0/injectPlus}
-\input{movies/snippets/Correctness_E0/injectMultFinR}
-
-% \begin{Coqsrc}
-% Theorem inject_mono (beta gamma : T1) :
-%   T1.lt  beta gamma -> 
-%   T1.nf beta -> T1.nf gamma -> 
-%   inject beta < inject gamma.
-
-% Theorem inject_injective (beta gamma : T1) : nf beta -> nf gamma ->
-%   inject beta = inject gamma -> beta = gamma.
-% \end{Coqsrc}
-
-Finally, we prove that \texttt{inject} is a bijection from the set of all terms of \texttt{T1} in normal form to the set 
-\texttt{members epsilon0} of the elements of \texttt{Ord} strictly less than  $\epsilon_0$.
-
-\input{movies/snippets/Correctness_E0/injectLtEpsilon0}
-\input{movies/snippets/Correctness_E0/embedding}
-
-\subsection{Remarks}
-Let us recall that the library \href{../theories/html/hydras.Schutte.Schutte.html}%
-{\texttt{Schutte}} depends on five \emph{axioms} and lies explicitly in the  
-framework of classical logic with a weak version of the axiom of choice
-(please look at the documentation of
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.ChoiceFacts.html}{\texttt{Coq.Logic.ChoiceFacts}}).
-Nevertheless, the other modules:
-\href{../theories/html/hydras.Epsilon0.Epsilon0.html}%
-{\texttt{Epsilon0}},
-\href{../theories/html/hydras.Hydra.Hydra.html}%
-{\texttt{Hydra}}, et 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}%
-{\texttt{Gamma0}}
-do not import any axioms and are really constructive.
-
-\index{hydras}{Projects}
-\begin{project}
-There is no construction of ordinal multiplication in~\cite{schutte}. 
-It would be interesting to derive this operation from Schütte's axioms,
-and prove its consistence with multiplication in ordinal notations for 
-$\epsilon_0$ and $\Gamma_0$.
-\end{project}
-
-\section{Related work}
-
-In~\cite{grimm:hal-00911710}, José Grimm establishes the consistency between our ordinal notations \texttt{T1} and \texttt{T2} (Veblen normal form) and his implementation
-of ordinal numbers after Bourbaki's set theory.
-
-The Gaia project ~\url{https://github.com/coq-community/gaia} maintains Grimm's  theory of ordinals as part of coq-community on GitHub. Integration
-of the present ordinal theory with Gaia, i.e., relating the different notions of ordinals
-and transferring relevant results, is an interesting project.
-First experiments in that direction are developped in
-the \href{https://github.com/coq-community/hydra-battles/blob/master/theories/gaia/}{theories/gaia/} directory.
-
-\chapter{The Ordinal \texorpdfstring{$\Gamma_0$}{Gamma0} (first draft)}
-
-\emph{This chapter and the files it presents are still very incomplete, considering the impressive properties of $\Gamma_0$~\cite{Gallier91}.
-We hope to add new material soon, and accept contributions!}
-
-\begin{todo}
-  Build a T2Bridge! (note: Gamma0's wellfoundedness is missing in \gaia (?).
-\end{todo}
-
-\section{Introduction}
-We present a notation system for the ordinal $\Gamma_0$, following Chapter V, Section 14 of~\cite{schutte}: ``A notation system for the ordinals $<\Gamma_0$''.
-We try to be as close as possible to Schütte's text and usual practices of \coq{} developments.
-
-The ordinal $\Gamma_0$ is defined in Section 13 of ~\cite{schutte} as the least \emph{strongly critical ordinal}. It is widely known as the \emph{Feferman-Schütte ordinal}.
-
-Section V, 13 of~\cite{schutte} defines \emph{strongly critical} and
-\emph{maximal $\alpha$-critical} ordinals:
-
-\begin{itemize}
-\item $\alpha$ is strongly critical if
-$\alpha$ is $\alpha$-critical,
-\item $\gamma$ is maximal $\alpha$-critical if $\gamma$ is $\alpha$-critical, and, for all $\xi>\alpha$, $\gamma$ is not $\xi$-critical.
-
-\end{itemize}
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Schutte.Critical.html\#strongly_critical}%
-{\texttt{Schutte.Critical}}}
-
-\input{movies/snippets/Critical/Gamma0Def}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that a (countable)  ordinal $\alpha$ is strongly critical iff 
-$\phi_\alpha(0)=\alpha$ (Theorem 13.13 of~\cite{schutte} ). 
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that the set of strongly critical ordinals is unbounded and closed (Theorem 13.14 of~\cite{schutte} ).
-Thus this set is not empty,  hence has a least element.
-Otherwise, the definition of $\Gamma_0$ above would be useless.
-\end{project}
-
-In the present version of this development, we  only study $\Gamma_0$ as a notation system, much more powerful than the ordinal notation for $\epsilon_0$.
-
-%\index{Projects}
-%
-%\begin{project}
-% Schûtte's section 13 of~\cite{schutte} contain several definitions and lemmas which give another view on $\Gamma_0$.  We leave it as a project to implement them in \coq{}.  
-%\end{project}
-
-\section{The type \texttt{T2} of ordinal terms}
-
-The notation system for ordinals less than $\gamma_0$ comes from the following theorem of~\cite{schutte}, where $\psi\,\alpha$ is the ordering function 
-of the set of maximal $\alpha$-critical ordinals.
-
-\index{hydras}{Library Gamma0!Types!T2}
-
-\begin{quote}
-  Any ordinal $\not= 0$ which is not strongly critical can be expressed in terms of $+$ and $\psi$.
-\end{quote}
-
-\index{hydras}{Projects}
-\begin{project}
-This theorem is not formally proved in this development yet. It should be!
-\end{project}
-
-Like in Chapter~\ref{chap:T1}, we define an inductive type with two constructors, one for $0$, the other for the construction $\psi(\alpha,\beta)\times(n+1)+\gamma$, adapting a Manolios-Vroon-like notation~\cite{Manolios2005} to
-\emph{Veblen normal forms}.
-\label{types:T2}
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T2}%
-{\texttt{Gamma0.T2}}}
-
-\inputsnippets{T2/T2Def, T2/psiDef}
-
-Like in chapter~\ref{chap:T1}, we get familiar with the type \texttt{T2} by recognizing simple constructs like finite ordinals, $\omega$, etc., as inhabitants of \texttt{T2}.
-
-\input{movies/snippets/T2/Notations}
-
-\section{A strict order on T2}
-
-Let us define a strict order on type \texttt{T2}. The following definition is 
-an adaptation of Schütte's, taking into account the multiplications by a natural number (inspired by~\cite{Manolios2005}, and also present in \texttt{T1}).
-
-\label{sect:t2-lt-def}
-
-\input{movies/snippets/T2/ltDef}
-
-Seven constructors! In order to get accustomed with this definition, let us look at a small set of examples, covering all the constructors of \texttt{lt}.
-
-\input{movies/snippets/T2/ltExamples}
-
-\index{hydras}{Projects}
-\begin{project}
-Write a tactic that solves automatically goals of the form (\texttt{$\alpha$ t2< $\beta$}), where $\alpha$ and $\beta$ are closed terms of type \texttt{T2}.
-\end{project}
-
-\section{Veblen normal form}
-\begin{definition}
-  A term of the form $\psi(\alpha_1,\beta_1)\times n_1+ \psi(\alpha_2,\beta_2)\times n_2+\dots+\psi(\alpha_k,\beta_k)\times n_k$ is said to be in
- \emph{[Veblen] normal form} if for every $i<n$, $\psi(\alpha_i,\beta_i)<\psi(\alpha_{i+1},\beta_{i+1})$, all the $\alpha_i$ and $\beta_i$ are in normal form, and all the $n_i$ are strictly positive integers.
-\end{definition}
-
-\input{movies/snippets/T2/nfDef}
-
-Let us look at some positive examples (we have to prove some inversion lemmas before proving counter-examples).
-
-\input{movies/snippets/T2/nfExamples}
-
-\subsection{Length of a term}
-
-The notion of \emph{term length} is introduced by Schütte as a helper for proving (at least) the \emph{trichotomy} property and transitivity of the strict order \texttt{lt} on \texttt{T2}.
-These properties are proved by induction on length.
-
-\subsection{Trichotomy}
-
-\emph{Trichotomy} is another name for the well-known property of decidable total ordering (like Standard Library's \texttt{Compare\_dec.lt\_eq\_lt\_dec}).
-
-We first prove by induction on $l$ the following lemma:
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0\#tricho_aux}%
-{\texttt{Gamma0.Gamma0}}}
-
-\input{movies/snippets/Gamma0/tricho}
-
-\input{movies/snippets/Gamma0/compare}
-
-With the help of \texttt{compare}, we get a boolean version of \texttt{nf}
-(being in Veblen normal form).
-
-\input{movies/snippets/Gamma0/nfb}
-
-\section{Main functions on \texttt{T2}}
-
-\subsection{Successor}
-The successor function is defined by structural recursion.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#succ}%
-{\texttt{Gamma0.T2}}}
-
-\input{movies/snippets/T2/succDef}
-
-\subsection{Addition}
-
-Like for Cantor normal forms (see Sect.~\ref{sect:infix-plus-T1}),  the definition of addition in \texttt{T2}  requires comparison between ordinal terms.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0T2.html\#succ}%
-  {\texttt{Gamma0.Gamma0}}}
-
-\input{movies/snippets/Gamma0/plusDef}
-
-\subsection{The Veblen function \texorpdfstring{$\phi$}{\texttt{phi}}}
-
-The enumeration function of critical ordinals, presented in Sect.~\vref{sect:phi-schutte}, is recursively defined in type \texttt{T2}.
-
-\input{movies/snippets/Gamma0/phiDef}
-
-Despite its complexity, the function \texttt{phi} is well adapted to proofs by simplification or computation.
-
-The relation between the constructor $\psi$ and the function $\phi$ is
-studied in~\cite{schutte}, and partially implemented in this development.
-\emph{Please contribute!}
-
-For instance, the following theorem states that, if $\gamma$ is the sum of a limit ordinal $\beta$ and a finite ordinal $n$, and $\beta$ is a fixpoint of
-$\phi(\alpha)$, then $\psi(\alpha,\gamma)=\phi_\alpha(\gamma+1)$.
-
-\input{movies/snippets/Gamma0/phiPsi}
-\input{movies/snippets/Gamma0/Ex9}
-
-On the other hand, $\phi$ can be expressed in terms of $\psi$.
-
-\input{movies/snippets/Gamma0/phiOfPsi}
-\input{movies/snippets/Gamma0/Ex10}
-
-\index{hydras}{Projects}
-\begin{project}
-Please study a way to pretty print ordinal terms in Veblen normal form (see Section~\vref{sect:ppT1}).
-\end{project}
-
-\section{An ordinal notation for \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}}
-
-In order to consider type \texttt{T2} as an ordinal notation, we have to build an instance of class \texttt{ON} (See Definition page~\pageref{types:ON}).
-
-First, we define a type that contains only terms in Veblen normal form, and redefine \texttt{lt} and \texttt{compare} by delegation (see for comparison the construction of type \texttt{E0} in Sect.~\vref{sect:E0-def}).
-
-\input{movies/snippets/Gamma0/G0a}
-\input{movies/snippets/Gamma0/G0b}
-
-Then, we buils an instance of class \texttt{ON}. 
-function \texttt{compare} is correct.
-
-\input{movies/snippets/Gamma0/ltSto}
-\input{movies/snippets/Gamma0/ltWf}
-\input{movies/snippets/Gamma0/ONGamma0}
-\input{movies/snippets/Gamma0/G0z}
-
-\begin{remark}
-The proof of \texttt{lt\_wf} has been written by \'Evelyne Contejean, using her library on the recursive path ordering (see also remark~\vref{remark:a3pat}).
-\end{remark}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that \texttt{Epsilon0} (page~\pageref{instance-epsilon0})
-is a sub-notation system of \texttt{Gamma0}.
-
-Prove that the implementations of \texttt{succ}, \texttt{+}, $\phi_0$, etc.
-are compatible in both notation systems.
-
-Note that a function \texttt{T1\_inj} from \texttt{T1} to \texttt{T2} has already been defined. It may help to complete the task.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T1_to_T2}%
-{\texttt{Gamma0.T2}}}
-
-\input{movies/snippets/T2/T1ToT2}
-
-\end{project}
-
-\begin{project}
-Prove that the notation system \texttt{Gamma0} is a correct implementation 
-of the segment $[0,\Gamma_0)$ of the set of countable ordinals.
-\end{project}

### 2022-05-08 doc/ordinal-chapter.tex
reorganization of tex files

--- 2022-04-11 doc/part-hydras.tex
@@
-\section*{Introduction}
-
-In this part, we present a development  for the \coq{} proof assistant, after the work of Kirby and Paris~\cite{KP82}. This formalization contains the following main parts:
-
-\begin{itemize}
-\item Representation in \coq{} of hydras and hydra battles.
-\item A proof that every battle is finite and won by Hercules.
-This proof is based on a \emph{variant} which maps any hydra to an ordinal strictly less than $\epsilon_0$ and is strictly decreasing along any battle.
-
-\item Using a combinatorial toolkit designed by J.~Ketonen and R.~Solovay~\cite{KS81}, we prove that, for any ordinal $\mu<\epsilon_0$, there exists no such variant mapping any hydra to an ordinal strictly less than $\mu$.
-Thus, the complexity of $\epsilon_0$ is really needed in the previous proof.
-
-\item We prove a relation between the length of a ``classic''  kind of  battles \footnote{This class is also called \emph{standard} in this document (text and proofs).
-The \emph{replication factor} of the hydra is exactly $i$ at the $i$-th round of the battle (see Sect~\vref{sect:replication-def}).}
-and the Wainer-Hardy hierarchy of ``rapidly growing functions'' $H_\alpha$~\cite{Wainer1970}.
-The considered class of battles, which we call \emph{standard},  is the most considered one in the scientific  literature (including popularization).
-\end{itemize}
-
-Simply put, this document tries to combine the scientific interest of two articles~\cite{KP82, KS81} and a book~\cite{schutte} with the playful activity of truly proving theorems.
-We hope  that such a work, besides exploring a nice piece of discrete mathematics,
-will show how \coq{} and its standard library are well fitted to help us to understand some non-trivial mathematical developments, and also to experiment the constructive parts of  the proof through functional programming.
-
- We also hope to provide a little clarification on infinity (both potential and actual) through the notions of function, computation, limit,
- type and proof.
-
- \subsection*{Compatibility with \gaia (in progress)}
- \label{sect:gaia-first-intro}
-The \gaia project~\cite{Gaia} by Jos\'e Grimm, Alban Quadrat,  and Carlos Simpson,
-aims  to formalize mathematics in \coq  in
-the style of Nicolas Bourbaki.
-It contains many definitions and results about ordinal numbers.
-In Chapter~\ref{gaia-chapter}, we present some modules which allow \HydrasLib' users to apply lemmas proven in \gaia, and vice versa.
-Remarks about compatibility with \gaia are signalled with the picture \gaiasign.
-
-%\section{Remarks}
-
-\subsection*{Difference from Kirby and Paris's work}
-In~\cite{KP82}, Kirby and Paris show  that there is no proof of termination of all hydra battles in Peano Arithmetic (PA).
-Since we are used to writing proofs in higher order logic, the restriction to PA was quite unnatural for us.
-So we chose to prove another statement without any reference to PA, by considering a class of proofs indexed by ordinal numbers up to $\epsilon_0$.
-
-\subsection*{State of the development}
-The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on  notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
-We added new material: axiomatic definitions of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, function derinition by equations, etc.
-
-We are now working in order to make clumsy proofs more readable, simplify definitions, and ``factorize'' proofs as much as possible. 
-Many possible improvements are suggested as ``todo''s or ``projects'' in this text.
-
-\section*{Future work (projects)}
-\index{hydras}{Projects}
-
-This document and the proof scripts are far from being complete.
-
-First, there must be a lot of typos to correct, references and index items to add. Many proofs are too complex and should be simplified, etc.
-
-The following extensions are planned, but help is needed:
-
-\begin{itemize}
-\item Semiautomatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
-%\item Extension to $\Gamma_0$ (in Veblen normal form)
-\item More lemmas about hierarchies of rapidly growing
-  functions, and their relationship 
-    with primitive recursive functions and provability in Peano arithmetic 
-(following~\cite{KS81, KP82}).
-\item From \coq's point of view, this development could be used as an illustration of the evolution of the software, every time new libraries and sets of tactics could help to simplify the proofs.
-\end{itemize}
-
-\subsection*{Main references}
-
-In our development, we adapt the definitions and prove many theorems which
-we found in the following articles. 
-\begin{itemize}
-\item ``Accessible independence results for Peano arithmetic''  by Laurie Kirby and Jeff Paris~\cite{KP82}
-\item ''Rapidly growing Ramsey Functions'' by Jussi Ketonen and Robert Solovay~\cite{KS81}
-\item ``The Termite and the Tower'', by Will Sladek~\cite{Sladek07thetermite}
-\item Chapter V of ``Proof Theory'' by Kurt Schütte~\cite{schutte}
-\end{itemize}
-
-\chapter{Hydras and hydra games}
-
-\label{sec:orgheadline91}
-\label{chapter:hydras}
-
-This chapter is dedicated to the representation of hydras and rules of the hydra game in \coq's specification language:~\gallina.
-
-Technically, a \emph{hydra} is just a finite ordered tree, each node of which 
-has any number of sons. Contrary to the computer science tradition, we display the hydras 
-with the heads up and the foot (i.e., the root of the tree) down.
-Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the 
-module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}). 
-\emph{For a less formal description of hydras, please see 
-\url{https://www.smbc-comics.com/comic/hydra}.}
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {$\bullet$};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H4) at (4,2) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N3) to [bend left= 10]  (H1) ;
-\draw (N3) to [bend right= 16] (H2);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (foot) to [bend right = 10] (H4) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-\caption{The hydra \texttt{Hy} \label{fig:Hy}}
-\end{figure}
-
-We use a specific vocabulary for talking about hydras. Table~\ref{tab:hyd2tree} shows the correspondence between our terminology and the usual vocabulary for trees in computer science.
-
-\begin{figure}[h]
-  \centering
-  \begin{tabular}{ll}
-Hydras & Finite rooted trees\\
-\hline
-foot & root\\
-head & leaf\\
-node & node\\
-segment  & (directed) edge \\
-sub-hydra & subtree\\
-daughter & immediate subtree\\
-\end{tabular}
-  \caption{Translation from hydras to trees}
-  \label{tab:hyd2tree}
-\end{figure}
-
-The hydra \texttt{Hy} has a \emph{foot} (below), five \emph{heads}, and eight \emph{segments}. 
-We leave it to the reader to define various parameters such as the height, the size, the highest arity (number of sons of a node) of a hydra.
-In our example, these parameters have the respective values $4$, $9$ and $3$.
-
-\subsection{The rules of the game}
-
-\label{sec:orgheadline44}
-\label{sect:replication-def}
-
-A \emph{hydra battle} is a fight between Hercules and the Hydra. 
-More formally, a  battle is a sequence of \emph{rounds}.
-At each round:
-\begin{itemize}
-\item If the hydra is composed of just one head, the battle is finished
-and  Hercules is the winner.
-\item Otherwise, Hercules chops off \emph{one} head of the hydra,
-
-\begin{itemize}
-\item If the head is at distance 1 from the foot, the head is just lost by the hydra, with no more reaction.
-\item Otherwise, let us denote by \(r\) the node that was at distance \(2\) from 
-the removed head in the direction of the foot,  and consider the  sub-hydra \(h'\) of \(h\), whose  root is \(r\) \footnote{$h'$ will be called ``the wounded part of the hydra'' in the subsequent text.
-In Figures~\vref{fig:Hy2} and ~\vref{fig:Hy4}, this sub-hydra  is displayed in red.}.
-Let $n$ be some natural number.
-Then $h'$ is replaced by  $n+1$ of copies of \(h'\) which share the same root $r$.
- The \emph{replication factor} $n$ may be different (and generally is)   at each round of the fight.
-It may be chosen by the hydra, according to its strategy, or imposed by some 
-particular rule.
-In many presentations of hydra battles, this number is increased by $1$ at each round.
-In the following presentation, we will also consider battles where the hydra is free to choose its ~replication factor at each round of the battle\footnote{Let us recall that, if the chopped-off head was at distance 1 from the foot, the replication factor is meaningless.}.
-\end{itemize}
-\end{itemize}
-
-Note that the description given in~\cite{KP82} of the replication process in hydra battles is also  semi-formal.
-
-\label{original-rules}
-
-\begin{quote}
-  ``From the node that used to be attached to the head which was just chopped off, traverse one 
-segment towards the root until the next node is reached. From this node sprout $n$ replicas of 
-that part of the hydra (after decapitation) which is ``above'' the segment just traversed, i.e., those 
-nodes and segments from which, in order to reach the root, this segment would have to be 
-traversed. If the head just chopped off had the root of its nodes, no new head is grown. ''
-\end{quote}
-
-Moreover, we note that this description is in \emph{imperative} terms.
-In order to formally study the properties of hydra battles, we prefer to use a mathematical vocabulary, i.e., graphs, relations, functions, etc.
-Thus, the replication process will be represented as a binary relation on a data type \texttt{Hydra},
-linking the state of the hydra \emph{before} and \emph{after} the transformation.
-A battle will thus be represented as a sequence of terms of type \texttt{Hydra}, respecting the rules of the game.
-
-\subsection{Example}
-Let us start a battle between Hercules and the hydra \texttt{Hy} of Fig.~\ref{fig:Hy}.
-
-At the first round, Hercules chooses to chop off the rightmost head of \texttt{Hy}.
-Since this head is near the floor, the hydra simply loses this head. Let us call 
- \texttt{Hy'} the resulting state of the hydra, represented in Fig.~\vref{fig:Hy-prime}.
-
-Next, assume Hercules chooses to chop off one of the two highest heads of \texttt{Hy'}, for instance the rightmost one.
-Fig.~\vref{fig:Hy2} represents the broken segment in dashed lines, and the part that will be replicated in red.
-Assume also that the hydra decides to add 4 copies of the red part\footnote{In other words, the replication factor at this round is equal to $4$.}.
-We obtain a new state \texttt{Hy''} depicted in Fig.~\ref{fig:Hy3}.
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {$\bullet$};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-%\node (H4) at (6,0) {$\Xey[2][lightgray]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3) ;
-\draw (N3) to [bend left= 10]  (H1) ;
-\draw (N3) to [bend right= 16] (H2);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-
-\caption{\texttt{Hy'}: the state  of \texttt{Hy} after one round \label{fig:Hy-prime}}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4)  {{\color{lightred}$\bullet$}};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Sey[2][lightred]$};
-%\node (H2) at (5,0) {$\Xey[2][lightgray]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\node (ex) at (5,8) {};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw  (N2) -- (N3)[draw=lightred];
-\draw (N3) to   [bend left= 10](H1) [draw=lightred];
-\draw [dashed] (N3) to [bend left= 10](ex);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-\caption{A second beheading}
-\label{fig:Hy2}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H11) at (2,8) {$\Smiley[2][vertfluo]$};
-\node (H12) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H13) at (6,8) {$\Smiley[2][vertfluo]$};
-\node (H14) at (8,8) {$\Smiley[2][vertfluo]$};
-
-\node (N3) at (1,6) {$\bullet$};
-\node (N31) at (2,6) {$\bullet$};
-\node (N32) at (3,6) {$\bullet$};
-\node (N33) at (4,6) {$\bullet$};
-\node (N34) at (5,6) {$\bullet$};
-
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N2) -- (N31);
-\draw (N2) -- (N32);
-\draw (N2) -- (N33);
-\draw (N2) -- (N34);
-\draw (N3) to   [bend left= 10](H1) ;
-\draw (N31) to   [bend left= 10](H11) ;
-\draw (N32) to   [bend left= 10](H12) ;
-\draw (N33) to   [bend left= 10](H13) ;
-\draw (N34) to   [bend left= 10](H14) ;
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend left= 10]  (H5) ;
-\end{tikzpicture}
-\caption{\texttt{Hy''}: the state of \texttt{Hy} after two rounds \label{fig:Hy3}}
-\end{figure}
-
-Figs.~\ref{fig:Hy4} and~\vref{fig:Hy5} represent a possible third round of the battle, with a replication factor equal to $2$. Let us call \texttt{Hy'''} the state of the hydra after that third round.
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0)  {{\color{lightred}$\bullet$}};
-\node (N1) at (2,2) {{\color{lightred}$\bullet$}};
-\node (N2) at (2,4) {{\color{lightred}$\bullet$}};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (exN4) at (4,4) {};
-\node (H1) at (0,8) {$\Sey[2][lightred]$};
-\node (H11) at (2,8) {$\Sey[2][lightred]$};
-\node (H12) at (4,8) {$\Sey[2][lightred]$};
-\node (H13) at (6,8) {$\Sey[2][lightred]$};
-\node (H14) at (8,8) {$\Sey[2][lightred]$};
-
-\node (N3) at (1,6) {{\color{lightred}$\bullet$}};
-\node (N31) at (2,6) {{\color{lightred}$\bullet$}};
-\node (N32) at (3,6) {{\color{lightred}$\bullet$}};
-\node (N33) at (4,6) {{\color{lightred}$\bullet$}};
-\node (N34) at (5,6) {{\color{lightred}$\bullet$}};
-
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-%\node (H5) at (4,0) {$\Xey[2][lightgray]$};
-\draw (foot) -- (N1)[very thick,draw=lightred] ;
-\draw (N1) -- (N2)[draw=lightred];
-\draw (N2) -- (N3)[draw=lightred];
-\draw (N2) -- (N31)[draw=lightred];
-\draw (N2) -- (N32)[draw=lightred];
-\draw (N2) -- (N33)[draw=lightred];
-\draw (N2) -- (N34)[draw=lightred];
-\draw (N3) to   [bend left= 10](H1) [draw=lightred];
-\draw (N31) to   [bend left= 10](H11) [draw=lightred];
-\draw (N32) to   [bend left= 10](H12) [draw=lightred];
-\draw (N33) to   [bend left= 10](H13) [draw=lightred];
-\draw (N34) to   [bend left= 10](H14) [draw=lightred];
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw [dashed] (N1) to  [bend left= 10](exN4);
-\end{tikzpicture}
-\caption{A third beheading (wounded part in red) \label{fig:Hy4}}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-
-\node (foot) at (10,0) {$\bullet$};
-
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H1) at (0,8) {$\Smiley[1][vertfluo]$};
-\node (H11) at (2,8) {$\Smiley[1][vertfluo]$};
-\node (H12) at (4,8) {$\Smiley[1][vertfluo]$};
-\node (H13) at (6,8) {$\Smiley[1][vertfluo]$};
-\node (H14) at (8,8) {$\Smiley[1][vertfluo]$};
-
-\node (N3) at (1,6) {$\bullet$};
-\node (N31) at (2,6) {$\bullet$};
-\node (N32) at (3,6) {$\bullet$};
-\node (N33) at (4,6) {$\bullet$};
-\node (N34) at (5,6) {$\bullet$};
-
-\node (H0) at (-3,3) {$\Smiley[1][vertfluo]$};
-
-\draw (foot) to [bend left=10] (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N2) -- (N31);
-\draw (N2) -- (N32);
-\draw (N2) -- (N33);
-\draw (N2) -- (N34);
-\draw (N3) to   [bend left= 10](H1) ;
-\draw (N31) to   [bend left= 10](H11) ;
-\draw (N32) to   [bend left= 10](H12) ;
-\draw (N33) to   [bend left= 10](H13) ;
-\draw (N34) to   [bend left= 10](H14) ;
-\draw (foot) to [bend left = 15]  (H0) ;
-
-% second copy 
-\node (N01) at (12,2) {$\bullet$};
-\node (N02) at (12,4) {$\bullet$};
-\node (N03) at (12,6) {{\color{lightred}$\bullet$}};
-\node (H001) at (10,8) {$\Smiley[1][vertfluo]$};
-\node (H0011) at (12,8) {$\Smiley[1][vertfluo]$};
-\node (H0012) at (14,8) {$\Smiley[1][vertfluo]$};
-\node (H0013) at (16,8) {$\Smiley[1][vertfluo]$};
-\node (H0014) at (18,8) {$\Smiley[1][vertfluo]$};
-
-\node (N03) at (11,6) {$\bullet$};
-\node (N031) at (12,6) {$\bullet$};
-\node (N032) at (13,6) {$\bullet$};
-\node (N033) at (14,6) {$\bullet$};
-\node (N034) at (15,6) {$\bullet$};
-
-\draw (foot) -- (N01)[very thick] ;
-\draw (N01) -- (N02);
-\draw (N02) -- (N03);
-\draw (N02) -- (N031);
-\draw (N02) -- (N032);
-\draw (N02) -- (N033);
-\draw (N02) -- (N034);
-\draw (N03) to   [bend left= 10](H001) ;
-\draw (N031) to   [bend left= 10](H0011) ;
-\draw (N032) to   [bend left= 10](H0012) ;
-\draw (N033) to   [bend left= 10](H0013) ;
-\draw (N034) to   [bend left= 10](H0014) ;
-
-% third copy 
-\node (N001) at (22,2) {$\bullet$};
-\node (N002) at (22,4) {$\bullet$};
-\node (N003) at (22,6) {{\color{lightred}$\bullet$}};
-\node (H001) at (20,8) {$\Smiley[1][vertfluo]$};
-\node (H0011) at (22,8) {$\Smiley[1][vertfluo]$};
-\node (H0012) at (24,8) {$\Smiley[1][vertfluo]$};
-\node (H0013) at (26,8) {$\Smiley[1][vertfluo]$};
-\node (H0014) at (28,8) {$\Smiley[1][vertfluo]$};
-
-\node (N003) at (21,6) {$\bullet$};
-\node (N0031) at (22,6) {$\bullet$};
-\node (N0032) at (23,6) {$\bullet$};
-\node (N0033) at (24,6) {$\bullet$};
-\node (N0034) at (25,6) {$\bullet$};
-
-\draw (foot) -- (N001)[very thick] ;
-\draw (N001) -- (N002);
-\draw (N002) -- (N003);
-\draw (N002) -- (N0031);
-\draw (N002) -- (N0032);
-\draw (N002) -- (N0033);
-\draw (N002) -- (N0034);
-\draw (N003) to   [bend left= 10](H001) ;
-\draw (N0031) to   [bend left= 10](H0011) ;
-\draw (N0032) to   [bend left= 10](H0012) ;
-\draw (N0033) to   [bend left= 10](H0013) ;
-\draw (N0034) to   [bend left= 10](H0014) ;
-\end{tikzpicture}
-\caption{The configuration \texttt{Hy'''} of \texttt{Hy} \label{fig:Hy5}}
-\end{figure}
-\FloatBarrier
-
-We leave it to the reader  to guess the following  rounds of the battle \dots
- % Please keep in mind that, in this 
-% the hydra is free to chose any number of replications at each time, whereas
-% Hercules chops only one head per round.
-
-% Let us precise that, in this game, Hercules wins if the hydra is eventually reduced 
-% to a single head. 
-% We know from~\cite{KP82} that, whichever the initial configuration of the
-% hydra, and the strategies of both players, Hercules eventually wins. The 
-% aforementionned paper shows also that there do not exist any \emph{simple} proof of this result.
-
-\section{Hydras and their representation in \emph{Coq}}
-\label{sec:orgheadline48}
-
-\index{hydras}{Library Hydra!Types!Hydra}
-\index{hydras}{Library Hydra!Types!Hydrae}
-
-In order to describe trees where each node can have an arbitrary (but finite) number of sons, it is usual to define a type where each node carries a \emph{forest}, \emph{i.e} a list of trees
-(see for instance Chapter 14, pages 400-406 of \cite{BC04}, also available as~\cite{BC04ch14}).
-
-For this purpose, we define two mutual \emph{ad-hoc}  inductive types, where \texttt{Hydra} is the main type, and \texttt{Hydrae} is a helper for describing finite sequences of hydras.
-\label{types:Hydra}
-\label{types:Hydrae}
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hydra}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/HydraDef}
-
-%\index{To do}
-\index{hydras}{Projects}
-
-\begin{project}
-Look for an existing library on trees with nodes of arbitrary arity, in order to replace  this ad-hoc type with something more generic.
-\end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-
- Another very similar representation could use the \texttt{list} type family instead of the specific 
-type \texttt{Hydrae}:
-
-\input{movies/snippets/Hydra_Definitions/HydraAlt}
-
-Using this representation, re-define all the constructions of this chapter.
-You will probably have to use patterns described for instance in~\cite{BC04} or the archives of the \coq communication channels (please consult~\url{https://coq.inria.fr/community.html}).
-
-\end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-The type \texttt{Hydra} above describes hydras as \emph{plane trees}, i.e., as drawn on a sheet of paper or computer screen. Thus, hydras are \emph{oriented},
-and it is appropriate to consider a \emph{leftmost} or \emph{rightmost} head of
-the beast. It could be interesting to consider another representation, in which
-every non-leaf node has a \emph{multi-set} -- not an ordered list -- of daughters.
-\end{project}
-
-\subsubsection{Abbreviations}
-
-We provide several notations for hydra patterns  which occur often in our developments.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#head}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/headsEtc}
-
-For instance, the hydra \texttt{Hy}  of Figure~\vref{fig:Hy} is defined in \emph{Gallina} as follows:
-
-\vspace{4mm}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html\#Hy}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/Hy}
-
-Hydras quite frequently contain  multiple adjacent  copies of the same subtree. The following functions
-will help us to describe and reason about replications in hydra battles.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#hcons_mult}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hconsMult}
-
-\vspace{4mm}
-
-For instance, the hydra \texttt{Hy''} of Fig~\vref{fig:Hy3}  can be defined in \coq{} as follows:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/HySecond}
-
-\subsubsection{Recursive functions on type \texttt{Hydra}}
-\label{sec:orgheadline41}
-\label{sec:hsize-def}
-
-In order to  define a recursive function over the type \texttt{Hydra}, one has to consider the three constructors 
-\texttt{node}, \texttt{hnil} and \texttt{hcons} of the mutually inductive types \texttt{Hydra} and \texttt{Hydrae}. 
-Let us define for instance the function which  computes the number of nodes of any hydra:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hsize}
-
-\input{movies/snippets/Hydra_Examples/HySize}
-
-Likewise, the \emph{height} (maximum distance between the foot and a head) 
-is defined by mutual recursion:
-
-\input{movies/snippets/Hydra_Definitions/height}
-
-\input{movies/snippets/Hydra_Examples/HyHeight}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Define a function \texttt{max\_degree: Hydra $\arrow$ nat} which  returns the highest degree of a node in any hydra.
-For instance, the evaluation of the term \texttt{(max\_degree Hy)} should return $3$.
-\end{exercise}
-
-\subsection{Induction principles for hydras}
-\label{sec:orgheadline42}
-
-In this section, we show how induction principles are used to prove properties on the type 
-\texttt{Hydra}. Let us consider for instance the following statement:
-\begin{quote}
-  `` The height of any hydra is strictly less than its size. ''
-\end{quote}
-
-\subsubsection{A failed attempt}
-
-One may try to use the default tactic of proof by induction, which corresponds to an application of the automatically  generated  induction principle for  type \texttt{Hydra}:
-
-\input{movies/snippets/Hydra_Examples/HydraInd}
-
-Let us start a simple proof by induction.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/BadInductiona}
-
-We might be tempted to do an induction on the sequence \texttt{s}:
-
-\input{movies/snippets/Hydra_Examples/BadInductionb}
-
-The first subgoal is trivial.
-
-\input{movies/snippets/Hydra_Examples/BadInductionc}
-
-Let us look at the second subgoal of the induction.
-
-\input{movies/snippets/Hydra_Examples/BadInductiond}
-
-We notice that this subgoal does not contain any hypothesis
-on the height and size of the hydra \texttt{h}. So, it looks hard to prove the conclusion. Let's stop.
-
-\input{movies/snippets/Hydra_Examples/BadInductione}
-
-\subsubsection{A Principle of mutual induction}
-In order to get an appropriate induction scheme for the types 
-\texttt{Hydra} and \texttt{Hydrae}, we can use  \coq{}'s  command \texttt{Scheme}.
-
-\index{coq}{Mutually inductive types}
-%\index{coq}{Commands!Scheme}
-
-\input{movies/snippets/Hydra_Definitions/HydraRect2}
-
-\input{movies/snippets/Hydra_Examples/HydraRect2Check}
-
-\subsubsection{A Correct proof}
-
-Let us now use \texttt{Hydra\_rect2} for proving that the height of any hydra is strictly less than its size.
-Using this scheme requires an auxiliary predicate, called \texttt{P0} in \texttt{Hydra\_rect2}'s statement.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hForall}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-\input{movies/snippets/Hydra_Examples/heightLtSizea}
-
-The first subgoal is easily solvable, using some arithmetic.
-The second and third ones are
-almost trivial. We let the reader look at the source. 
-\input{movies/snippets/Hydra_Examples/heightLtSizez}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-It happens very often that, in the proof of  a proposition of the form 
-(\texttt{$\forall\,$ h:Hydra, $P$ h}), the predicate \texttt{P0}
-is  (\texttt{h\_forall $P$}).
-Design a tactic for induction on hydras that frees the user from binding explicitly \texttt{P0},  and solves trivial subgoals.
-Apply it for writing  a shorter proof of \texttt{height\_lt\_size}.
-\end{exercise}
-
-\section{Relational description of hydra battles}
-
-In this section, we represent the rules of hydra battles as a binary relation associated with
-a \emph{round}, i.e., an interaction composed of the two following actions:
-\begin{enumerate}
-\item Hercules chops off one head of the hydra.
-\item Then, the  hydra replicates the wounded part (if the head is at distance $\geq 2$ from the foot).
-\end{enumerate}
-The relation associated with each round of the battle is parameterized  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot,
-but present for consistency's sake).
-
-In our description,  we will apply the following naming convention: if $h$ represents the configuration of the hydra before a round, then the configuration of $h$ after this round will be called $h'$.
- Thus, we are going to define a proposition  (\texttt{round\_n $n\;h\;h'$})  whose intended meaning will be `` the hydra $h$  is transformed into $h'$  in a single round of a battle, with the expected replication factor $n$ ''.
-
-Since the replication of parts of the hydra depends on the distance of the chopped head from  the foot, we  decompose our description into two main  cases, under the form of a bunch of [mutually] inductive predicates over the types \texttt{Hydra} and \texttt{Hydrae}.
-
-The mutually exclusive cases we consider are the following:
-\begin{itemize}
-\item \textbf{R1}: The chopped off head was at distance 1 from the foot.
-\item \textbf{R2}: The chopped off head was at a distance greater than or equal to  $2$ from the foot.
-\end{itemize}
-
-\subsection{Chopping off a head at distance 1 from the foot (relation  R1)}
-
-If Hercules chops off a head near the floor, there is no replication at all. We use an auxiliary 
-predicate \texttt{S0}, associated with the removing of one head from a sequence of hydras.
-
-\vspace{4pt}\emph{From Module\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/S0Def}
-\input{movies/snippets/Hydra_Definitions/R1Def}
-
-\subsubsection{Example}
-\label{sec:orgheadline45}
-
-Let us represent in \coq{}   the transformation of the hydra of Fig.~\vref{fig:Hy} into
-the configuration represented in Fig.~\ref{fig:Hy-prime}.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/Hy1}
-
-\subsection{Chopping off a head at distance \texorpdfstring{$\geq 2$}{>= 2} from the foot (relation R2) }
-
-Let us now consider beheadings  where the chopped-off head is at distance greater than or equal to $2$ from the foot. All the following relations are parameterized by the replication factor  $n$.
-
- Let $s$ be a sequence of hydras. 
-The proposition (\texttt{S1 n s s'}) holds if $s'$ is obtained by replacing some element $h$ of $s$ by 
-$n+1$ copies of $h'$, where  the proposition (\texttt{R1 h h'}) holds, in other words, $h'$ is just $h$, without the chopped-off  head.
-\texttt{S1} is an inductive relation with two constructors that allow us to choose the position in $s'$ of the wounded sub-hydra $h$.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#S1}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/S1Def}
-
-The rest of the definition is composed of two mutually inductive relations on hydras and sequences of hydras.
-The first constructor of \texttt{R2} describes the case where the chopped head is exactly at height $2$.
-The others constructors allow us to consider beheadings at height strictly greater than $2$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/R2Def}
-
-\subsubsection{Example}
-Let us prove the transformation of \texttt{Hy'} into \texttt{Hy''} (see Fig.~\vref{fig:Hy3}).
-We use an experimental set of tactics\footnote{See the \texttt{Ltac} definitions in~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions.}}  for specifying the place where the
-interaction between Hercules and the hydra holds.
-
-\vspace{4pt}\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}.
-
-\input{movies/snippets/Hydra_Examples/R2Example}
-
-The reader is encouraged to look at all the successive subgoals of this example.
-\emph{Please consider also exercise~\vref{exo:interactive-battle}.}
-
-\subsection{Binary relation associated with a round}
-
-We combine the two cases above into a single relation.
-First,  we define the  relation \texttt{(round\_n n h h')} where \texttt{n} is the expected number of  replications (irrelevant in the case of an \texttt{R1}-transformation).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#round_n}{Hydra.Hydra\_Definitions}}
-
-\index{hydras}{Library Hydra!Predicates!round\_n}
-
-\input{movies/snippets/Hydra_Definitions/roundNDef}
-
-By abstraction over \texttt{n}, we define a \emph{round} (small step) of a battle:
-
-\index{hydras}{Library Hydra!Predicates!round}
-\label{sect:infix-round}
-
-\input{movies/snippets/Hydra_Definitions/roundDef}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Give a direct translation of Kirby and Paris's description of hydra battles (quoted on page~\pageref{original-rules}) and prove that our relational description is consistent with theirs.
-\end{project}
-
-\subsection{Rounds and battles}
-
-Using library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}, we define \texttt{round\_plus},  the transitive closure of \texttt{round}, and \texttt{round\_star},  the reflexive and transitive closure of \texttt{round}.
-
-\label{sect:infix-rounds}
-
-\input{movies/snippets/Hydra_Definitions/roundPlus}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-  Prove that if \texttt{$h$ -+-> $h'$}, then
-  the height of $h'$ is less or equal than the height of $h$.
-
-\end{exercise}
-
-\begin{remark}
-\label{remark:transitive-closure}
-\coq's library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Coq.Relations.Relation\_Operators} 
-contains three logically equivalent definitions of the transitive closure of a binary relation. This equivalence is proved in 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} .
-
-Why three definitions for a single mathematical concept?
-Each definition generates an associated induction principle. 
- According to the form of statement one would like to prove, there is a ``best choice'':
-
-\begin{itemize}
-\item For proving $\forall y, x\,R^+\,y \;\arrow\; P\,y$, prefer 
-\texttt{clos\_trans\_n1}
-\item For proving $\forall x,\,x\,R^+\,y \;\arrow\; P\,x$, prefer \texttt{clos\_trans\_1n}
-\item For proving $\forall x\,y, \,x\,R^+\,y \;\arrow\;P\,x\,y$,  
-prefer \texttt{clos\_trans},
-\end{itemize}
-But there is no ``wrong choice'' at all: the equivalence lemmas in \linebreak 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} 
- allow the user
-to convert any one of the three closures into another one before applying the corresponding elimination tactic.
-The same remark also holds for reflexive and transitive closures. 
-\end{remark}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Define a restriction of \coqsimple{round},  where Hercules always chops off
-the leftmost among the lowest heads.
-
-Prove that, if $h$ is not a simple head, then there exists a unique $h'$ such that $h$  is transformed into $h'$ in one round, according to this restriction.
-
-\end{exercise}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}[Interactive battles]
-\label{exo:interactive-battle}
-Given a hydra \texttt{h}, the specification of a hydra battle for \texttt{h} is the type 
-\Verb@{h':Hydra | h -*-> h'}@. In order to avoid long sequences of \texttt{split}, \texttt{left}, and 
-\texttt{right}, design a set of dedicated tactics for the interactive building of a battle.
-Your tactics will have the following functionalities:
-\begin{itemize}
-\item  Chose to stop a battle, or continue
-\item Chose an expected number of replications
-\item Navigate in a hydra, looking for a head to chop off.
-\end{itemize}
-
-Use your tactics for simulating a small part of a hydra battle, for instance the rounds which lead from
-\texttt{Hy} to \texttt{Hy'''}  (Fig.~\vref{fig:Hy5}).
-
-\textbf{Hints:} 
-\begin{itemize}
-
-\item Please keep in mind that the last  configuration of your interactively built battle is known only at the end of the battle.
-Thus, you will have to create and solve subgoals with existential variables.
-For that purpose, the tactic \texttt{eexists}, applied to the
-goal \Verb@{h':Hydra | h -*-> h'}@ generates the subgoal \Verb|h -*-> ?h'|.
-\item You may use Gérard Huet's \emph{zipper} data structure~\cite{zipper} for writing tactics associated with Hercules's  interactive search for a head to chop off.
-\end{itemize}
-
-\end{exercise}
-
-\subsection{Classes of battles}
-\label{sect:battle-classes}
-
-In some presentations of hydra battles, e.g.~\cite{KP82, bauer2008}, the transformation associated with the $i$-th round may depend on $i$.
-For instance, in these articles, the replication factor at the $i$-th round is equal to $i$.
-In other examples, one can allow the hydra to apply any replication factor at any time.
-In order to be the most general as possible, we define the type of predicates which relate the state of the hydra before and after the $i$-th round of a battle.
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-\label{types:Battle}
-\index{hydras}{Library Hydra!Type classes!Battle}
-
-\input{movies/snippets/Hydra_Definitions/BattleDef}
-
-The most general class of battles is \texttt{free}, which allows the hydra to chose any replication factor at every step:
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#free}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/freeDef}
-
-We chose to call \emph{standard}\footnote{This appellation is ours.
-If there is a better one, we will change it.} the kind of battles which appear  most often in the literature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#standard}{Hydra.Hydra\_Definitions}}
-\input{movies/snippets/Hydra_Definitions/standardDef}
-
-\subsection{Big steps}
-
-Let $B$ be any instance of class \texttt{Battle}. It is easy to define inductively the relation between the $i$-th and the $j$-th steps of a battle of type $B$.
-
-\vspace{6pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/battleRelDef}
-
-The following property allows us to build battles by composition of smaller ones.
-
-%% TODO display subgoals when fixed
-
-\vspace{6pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-\noindent
-\input{movies/snippets/Hydra_Lemmas/battleTrans}
-
-% \begin{remark}
-%  The class \texttt{free} is strongly related with the transitive closure  \texttt{round\_plus}, as expressed by the following lemmas.
-
-% \vspace{4pt}
-% \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-
-%  \begin{Coqsrc}
-%  Lemma battle_free_equiv1 : forall i j h h',  
-%              battle free i h j h' ->   h -+-> h'.
-
-%  Lemma battle_free_equiv2 : forall h h',
-%      h -+-> h' ->
-%     forall i, exists j,  battle free i h j h'.
-%  \end{Coqsrc}
-
-% \end{remark}
-
-\section{A long battle}
-\label{sect:big-battle}
-
-In this section we consider a simple example of battle, starting with a small hydra,
-shown on figure~\vref{fig:hinit}, with a simple strategy for both players:
-
-\begin{itemize}
-\item At each round, Hercules chops off the rightmost head of the hydra.
-\item The battle is {standard}: at the round number $i$, the expected replication is $i$.
-\end{itemize}
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (6,0) {$\bullet$};
-\node (n1) at  (3,3) {$\bullet$};
-\node (h1) at  (1,6) {$\Smiley[1][green]$};
-\node (h2) at  (3,6) {$\Smiley[1][green]$};
-\node (h3) at  (6,6) {$\Smiley[1][green]$};
-\node (h4) at  (6,6) {$\Smiley[1][green]$};
-\node (h5) at  (6,3) {$\Smiley[1][green]$};
-\node (h6) at  (9,3) {$\Smiley[1][green]$};
-\draw (foot) -- (n1);
-\draw (n1) to   [bend left=20] (h1);
-\draw (n1) to   (h2);
-\draw (n1) to   [bend right=20] (h3);
-\draw (foot) -- (h5);
-\draw (foot) to  [bend right=20] (h6);
-\end{tikzpicture}
-
-  \caption{The hydra \texttt{hinit}}
-  \label{fig:hinit}
-\end{figure}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.BigBattle.html}{Hydra.BigBattle}}
-
-\input{movies/snippets/BigBattle/hinitDef}
-
-The lemma we would like to prove is ``The considered battle lasts exactly $N$ rounds'',
-with $N$ being a natural number we gave to guess.
-
-But the  battle is so long that no \emph{test} can give us any estimation of its length, and we do need the expressive power of logic to compute this length.
-However, in order to  guess this length, we made some experiments, computing with \gallina{}, \coq{}'s  functional programming language.
-Thus, we can consider this development as a collaboration of proof with computation.
-In the following lines, we show how we found experimentally the value of the number $N$.
-The complete proof is in file \url{../theories/html/hydras.Hydra.BigBattle.html}.
-
-\subsection{First rounds}
-During the two first rounds, our hydra loses its two rightmost heads.  Figure~\vref{fig:hinit-plus2} shows the state of the hydra   just before the third round.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (3,0) {$\bullet$};
-\node (n1) at  (3,3) {$\bullet$};
-\node (h1) at  (1,6) {$\Smiley[1][green]$};
-\node (h2) at  (3,6) {$\Smiley[1][green]$};
-\node (h3) at  (6,6) {$\Smiley[1][green]$};
-\node (h4) at  (6,6) {$\Smiley[1][green]$};
-\draw (foot) -- (n1);
-\draw (n1) to   [bend left=20] (h1);
-\draw (n1) to   (h2);
-\draw (n1) to   [bend right=20] (h3);
-\end{tikzpicture}
-
-  \caption{The hydra (\texttt{hyd1 h3})}
-  \label{fig:hinit-plus2}
-\end{figure}
-
-The following lemma  is a formal description of these first rounds, in terms of the
-\texttt{battle} predicate.
-
-\input{movies/snippets/BigBattle/L02}
-
-\subsection{Looking for regularities}
-
-A first study with pencil and paper suggested us that, after three rounds, the hydra always looks like in figure~\vref{fig:hinit-plusn} (with a variable number of 
-subtrees of height 1 or 0).
-Thus, we introduce a few handy abbreviations.
-
-\input{movies/snippets/BigBattle/Notations}
-
-For instance, the hydra shown in  Fig~\ref{fig:hinit-plusn} is  (\texttt{hyd 3 4 2}). The hydra (\texttt{hyd 0 0 0})  is the ``final'' hydra of any terminating battle, {i.e.},
-a tree whith exactly one node and no edge.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (15,0) {$\bullet$};
-\node (a) at  (3,4) {$\bullet$};
-\node (b) at  (6,4) {$\bullet$};
-\node (c) at  (9,4) {$\bullet$};
-\node (d) at  (13,4) {$\bullet$};
-\node (e) at  (16,4) {$\bullet$};
-\node (f) at  (19,4) {$\bullet$};
-\node (g) at  (22,4) {$\bullet$};
-\node (h) at  (25,4) {$\Smiley[1][green]$};
-\node (i) at  (28,4) {$\Smiley[1][green]$};
-\node (aa) at  (2.5,8) {$\Smiley[1][green]$};
-\node (ab) at  (3.5,8) {$\Smiley[1][green]$};
-\node (ba) at  (5.5,8) {$\Smiley[1][green]$};
-\node (bb) at  (6.5,8) {$\Smiley[1][green]$};
-\node (ca) at  (8.5,8) {$\Smiley[1][green]$};
-\node (cb) at  (9.5,8) {$\Smiley[1][green]$};
-\node (da) at  (13,8) {$\Smiley[1][green]$};
-\node (ea) at  (16,8) {$\Smiley[1][green]$};
-\node (fa) at  (19,8) {$\Smiley[1][green]$};
-\node (ga) at  (22,8) {$\Smiley[1][green]$};
-\draw (foot) -- (a);
-\draw (foot) -- (b);
-\draw (foot) -- (c);
-\draw (foot) -- (d);
-\draw (foot) -- (e);
-\draw (foot) -- (f);
-\draw (foot) -- (g);
-\draw (foot) -- (h);
-\draw (foot) -- (i);
-\draw (a) -- (aa);
-\draw (a) -- (ab);
-\draw (b) -- (ba);
-\draw (b) -- (bb);
-\draw (c) -- (ca);
-\draw (c) -- (cb);
-\draw (d) -- (da);
-\draw (e) -- (ea);
-\draw (f) -- (fa);
-\draw (g) -- (ga);
-\end{tikzpicture}
-
-  \caption{The hydra (\texttt{hyd 3 4 2})}
-  \label{fig:hinit-plusn}
-\end{figure}
-
-With these notations, we get a formal description of the first three rounds.
-
-\input{movies/snippets/BigBattle/L23L03}
-
-\subsection{Testing  \dots}
-\label{sect:testing}
-In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple data type for representing the states as tuples composed of
-the round number, and the respective number of daughters  \texttt{h2}, \texttt{h1}, and heads
-of the current hydra.
-
-\input{movies/snippets/BigBattle/stateDef}
-
-The following function returns the next configuration of the game.
-Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle only start with the lemma
-\texttt{step\_battle}, page~\pageref{lemma:step-battle}.
-
-\input{movies/snippets/BigBattle/nextDef}
-
-We can make bigger steps through iterations of \texttt{next}.
-The functional \texttt{iterate}, similar to Standard Library's \texttt{Nat.iter},
-is defined and studied in~\href{../theories/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
-\index{hydras}{Library Prelude!iterate}
-
-\label{Functions:iterate}
-
-\input{movies/snippets/Iterates/iterateDef}
-
-The following function computes the state of the battle at the $n$-th round.
-
-\input{movies/snippets/BigBattle/testDefTests}
-
-The battle we are studying looks to be awfully long. Let us concentrate our
-tests on some particular events : the states where $\texttt{nh}=0$.
-From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} is equal to zero.
-
-\input{movies/snippets/BigBattle/smartTest}
-
-Thus, $(1 + 11)$ rounds later, the \texttt{n1} field is equal to $2$, and 
-\texttt{nh}   to $0$.
-
-\input{movies/snippets/BigBattle/smartTestb}
-
-Next round, we decrement \texttt{n2} and set \texttt{n1} to $95$.
-
-\input{movies/snippets/BigBattle/smartTestc}
-
-We now have some intuition of the sequence.
-It looks like the next ``\texttt{nh}=0'' event will happen at the $192=2(95+1)$-th round, then at the $2(192+1)$-th round, etc.
-
-\input{movies/snippets/BigBattle/doubleS}
-
-\subsection{Proving \dots}
-We are now able to reason about the sequence of transitions defined by our hydra battle. Instead of using the data-type \texttt{state} we study the relationship
-between different configurations of the battle.
-
-Let us define a binary relation associated with every round of the battle.
-In the following definition \texttt{i} is associated with the round number (or date, if we consider a discrete time), and \texttt{a}, \texttt{b}, \texttt{c} respectively associated with the number of occurrences of \texttt{h2}, \texttt{h1} and heads connected to the hydra's foot.
-
-\input{movies/snippets/BigBattle/oneStep}
-
-The relation between \texttt{one\_step} and the rules of hydra battles is asserted by the following lemma.
-
-\label{lemma:step-battle}
-
-\input{movies/snippets/BigBattle/stepBattle}
-
-\vspace{4pt}
-
-Next, we define ``big steps'' as the transitive closure of \texttt{one\_step},
-and reachability (from the initial configuration of figure~\ref{fig:hinit} at time $0$).
-
-\input{movies/snippets/BigBattle/steps}
-
-The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{battle}.
-
-\input{movies/snippets/BigBattle/stepsBattle}
-
-\vspace{4pt}
-
-Thus, any result about \texttt{steps} will be applicable to standard battles.
-Using the predicate \texttt{steps},  our study of the length of the considered battle
-can  be decomposed into three parts:
-
-\begin{enumerate}
-\item  Characterization and proofs of regularities of some events (inspired by our experiments of Sect.~\ref{sect:testing}).
-\item Study of the beginning of the battle
-\item Computing the exact length of the battle.
-\end{enumerate}
-
-First, we prove that, if at round $i$ the hydra is equal to
-(\texttt{hyd a (S b) 0}), then it will be equal to (\texttt{hyd a b 0}) at the $2(i+1)$-th round.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/LS}
-
-From now on, the lemma \texttt{reachable\_S} allows us to watch larger and larger steps of 
-the battle.
-
-\input{movies/snippets/BigBattle/L4}
-
-\input{movies/snippets/BigBattle/L10To95}
-
-\subsection{Giant steps}
-
-We are now able to make bigger steps in the simulation of the battle.
-First, we iterate the lemma \texttt{reachable\_S}.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/Bigstep}
-
-\vspace{4pt}
-
-Applying lemmas \texttt{BigStep} and \texttt{L95} we make a first jump.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/MDef}
-
-Figure~\ref{fig:HM}  represents the hydra at the $M$-th round.
-At the $(M+1)$-th round, it will look like in fig~\ref{fig:HM-plus1}.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:HM}}
-The state of the hydra after $M$ rounds.
-% The hydra \texttt{h} of the proof that \(\omega^2\) is too small for proving Hercules' victory
-
-\end{figure}
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (0,5) {$\bullet$};
-\node (N12) at (0,8) {$\Smiley[2][green]$};
-\node (N2) at (2,5) {$\bullet$};
-\node (N22) at (2,8) {$\Smiley[2][green]$};
-\node (N3) at (4,5) {$\bullet$};
-\node (N32) at (4,8) {$\Smiley[2][green]$};
-\node (N4) at (6,5) {$\bullet$};
-\node (N42) at (6,8) {$\Smiley[2][green]$};
-
-\node (Ndots) at (12,8) {\Huge $\dots$};
-\node (Ndots2) at (12,5) {\Huge $\dots$};
-
-\node (N8) at (18,5) {$\bullet$};
-\node (N82) at (18,8) {$\Smiley[2][green]$};
-\node (N9) at (20,5) {$\bullet$};
-\node (N92) at (20,8) {$\Smiley[2][green]$};
-
-\draw (foot) -- (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N8);
-\draw (foot) -- (N9);
-\draw (N1) to  (N12);
-\draw (N2) to  (N22);
-\draw (N3) to  (N32);
-\draw (N4) to  (N42);
-\draw (N8) to  (N82);
-\draw (N9) to  (N92);
-\end{tikzpicture}
-\caption{\label{fig:HM-plus1}}
-The state of the hydra after $M+1$ rounds (with $M+1$ heads).
-
-\end{figure}
-
-\input{movies/snippets/BigBattle/L295S}
-
-\vspace{4pt}
-
-Then, applying once more the lemma \texttt{BigStep}, we get the exact time when
-Hercules wins!
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/NDef}
-
-\vspace{4pt}
-
-We are now able to prove formally that the considered battle is 
-composed of $N$ steps.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/Done}
-
-\subsection{A minoration lemma}
-
-Now, we would like to get an intuition of  how big the number $N$ is.
-For that purpose, we use a minoration of the function \texttt{doubleS} by the
-function (\texttt{fun n => 2 * n}).
-
-\vspace{4pt}
-
-\input{movies/snippets/Exp2/exp2Def}
-
-\vspace{4pt}
-
-Using a few facts (proven in 
-\href{../theories/html/hydras.Hydra.BigBattle.html}{hydras.Hydra.BigBattle}),we get several  minorations.
-
-\input{movies/snippets/BigBattle/minorationLemmas}
-
-\vspace{4pt}
-
-The number $N$ is greater than or  equal to $2^{2^{95}\times 95}.$ If we wrote $N$ in base $10$, $N$ would require at least $10^{30}$ digits!
-
-\section{Generic properties}
-
-The example we just studied shows that the termination of any battle may take a very long time. If we want to study hydra battles in general, we have to consider 
-any hydra and any strategy, both for Hercules and the hydra itself. So, we first  give some definitions, generally borrowed from transition systems vocabulary (see~\cite{tel_2000} for instance).
-
-\subsection{Liveliness}
-
-Let $B$ be an instance of \texttt{Battle}. We say that $B$ is \emph{alive} if
-for any configuration $(i,h)$, where $h$ is not a head, there exists a further step in class $B$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/AliveDef}
-
-The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
-\href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
-
-\input{movies/snippets/Hydra_Theorems/AliveThms}
-
-Both theorems are proved with the help of the  following strongly specified function:
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
-
-\input{movies/snippets/Hydra_Lemmas/nextRoundDec}
-
-\subsection{Termination}
-
-The termination of \emph{any}  battle is naturally expressed by the predicate \texttt{well\_founded} defined in the module \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}
- of the Standard Library.
-
-\index{hydras}{Library Hydra!Predicates!Termination}
-
-\input{movies/snippets/Hydra_Definitions/TerminationDef}
-
-Let $B$ be an instance of class \texttt{Battle}. A \emph{variant} for $B$ consists
-in a well-founded relation $<$  on some type \texttt{A}, and a function
-(also called a \emph{measure}) $m:\texttt{Hydra}\rightarrow A$ such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $B$, the inequality $m(h')<m(h)$ holds.
-
-\vspace{4pt}
-\noindent
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
-
-\label{sect:hvariant-def}
-
-\index{hydras}{Library Hydra!Type classes!Hvariant}
-
-\input{movies/snippets/Hydra_Definitions/HvariantDef}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
- Prove that, if there exists some  instance of (\texttt{Hvariant Lt wf\_Lt $B$ $m$}), then there exists no infinite battle in  $B$.
-\end{exercise}
-
-\subsection{A  small proof of impossibility}
-%\index{coq}{Proofs of impossibility}
-
-\label{omega-case}
-
-When one wants to prove a termination theorem with the help of a variant, 
-one has to consider first a well-founded set $(A,<)$, then a strictly decreasing measure on this set.
-The following two lemmas show that if  the order structure $(A,<)$ is too simple, it is useless to look for a convenient measure, which simply no exists.
-Such kind of result is useful, because it saves you time and effort.
-
-The best known well-founded order is the natural order on the set $\mathbb{N}$ of natural numbers (the type \texttt{nat} of Standard library).
-It would be interesting to look for some measure $m:\texttt{nat}\arrow\texttt{nat}$ and prove it is a variant.
-
-Unfortunately, we can prove that 
-\emph{no} instance of class (\texttt{WfVariant round Peano.lt $m$}) can be built, where
-$m$ is \emph{any} function of type \texttt{Hydra $\arrow$ nat}.
-
-Let us present the main steps of that proof, the script of which  is in the module ~\href{../theories/html/hydras.Hydra.Omega_Small.html}{Hydra/Omega\_Small.v} \footnote{ The name of this file means ``the ordinal $\omega$ is too small for proving the termination of [free] hydra battles ''.
-In effect, the elements of $\omega$, considered as a set, are just the natural numbers (see next chapter for more details)}.
-
-%\subsubsection{Preliminaries}
-
-Let us assume there exists some variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
-    the  termination of all hydra battles.
-
-\input{movies/snippets/Omega_Small/omegaSmalla}
-
-We define an injection $\iota$ from the type \texttt{nat} into \texttt{Hydra}.
-For any natural number $i$, $\iota(i)$ is the hydra composed of a foot and
-$i+1$ heads at height $1$. For instance, Fig.~\ref{fig:flower} represents the hydra $\iota(3)$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (4,0) {$\bullet$};
-\node (N1) at (2,2) {$\Smiley[2][green]$};
-\node (N2) at (4,2) {$\Smiley[2][green]$};
-\node (N3) at (6,2) {$\Smiley[2][green]$};
-\node (N4) at (8,2) {$\Smiley[2][green]$};
-\draw (foot) to [bend left =25] (N1);
-\draw (foot) to [bend left =15] (N2);
-\draw (foot) to [bend right =15] (N3);
-\draw (foot) to [bend right =25] (N4);
-\end{tikzpicture}
-\caption{\label{fig:flower}
-The hydra $\iota(3)$}
-\end{figure}
-
-\input{movies/snippets/Omega_Small/iotaDef}
-
-Let us consider now some hydra \texttt{big\_h} out of the range of the injection $\iota$ (see Fig.~\vref{fig:h-omega-omega}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to  (N2);
-\end{tikzpicture}
-\caption{\label{fig:h-omega-omega}}
- The hydra \texttt{big\_h}.
-\end{figure}
-
-\input{movies/snippets/Omega_Small/bigHDef}
-
- Using the functions $m$ and $\iota$, we define a second hydra \texttt{small\_h}, and show
- there is a one-round battle that transforms \texttt{big\_h} into \texttt{small\_h}. Please note that,
-due to the hypothesis \texttt{Hvar}, we are interested in the termination of \emph{free} battles. 
-There is no problem to consider a round with (\texttt{m big\_h}) as the replication factor.
-
-\input{movies/snippets/Omega_Small/smallHDef}
-
-But, by hypothesis, $m$ is a variant. Hence, we infer the following inequality.
-
-\vspace{4pt}
-
-\input{movies/snippets/Omega_Small/mLt}
-
-In order to get a contradiction, it suffices to  prove the inequality
-$m(\texttt{big\_h}) \leq m(\texttt{small\_h})$ i.e.,  $m(\texttt{big\_h}\leq m(\iota (m(\texttt{big\_h})))$.
-
-\input{movies/snippets/Omega_Small/mGea}
-
-Intuitively, it means that, from any hydra of the form (\texttt{iota $i$}), the battle will 
-take (at least) $i$ rounds. Thus the associated measure cannot be less than $i$.
-Technically, we prove this lemma by Peano induction on $i$.
-
-\begin{itemize}
-\item The base case $i=0$ is trivial
-\item Otherwise, let $i$ be any natural number and assume  the inequality
-  $i \leq m(\iota(i))$.
-  \begin{enumerate}
-  \item  But the hydra $\iota(S(i))$ can be transformed in one round into
-    $\iota(i)$ (by losing its rightmost head, for instance)
-  \item Since $m$ is a variant, we have $m(\iota(i)) < m(\iota(S(i)))$,
-    hence  $i< m(\iota(S(i)))$, which implies  $S(i)\leq  m(\iota(S(i)))$.
-  \end{enumerate}
-\end{itemize}
-
-We are now ready to complete our impossibility proof.
-
-\vspace{4pt}
-
-\inputsnippets{Omega_Small/mGeb, Omega_Small/mGez,
-  Omega_Small/omegaSmallz}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\subsubsection{Conclusion}
-
-In order to build a variant for proving the termination of all hydra battles, we need to consider order structures more complex than the usual order on type \texttt{nat}. 
-The notion of \emph{ordinal number} provides a catalogue of well-founded order types.
-For a reasonably large bunch of ordinal numbers, \emph{ordinal notations} are data-types which allow the \coq{} user to define functions, to compute and prove some properties, for instance by reflection.
-
-The next chapter is dedicated to a generic formalization of ordinal notations, and chapter~\ref{chap:T1} to a proof of termination of all hydra battles with the help of an ordinal notation for the interval $[0,\epsilon_0)$\,\footnote{We use the mathematical notation $[a,b)$ for the interval $\{x|a\leq x < b\}$.}.
-\index{maths}{Notations!Interval}
-
-%--------------------------------------------------------------
-
 \chapter{Introduction to ordinal numbers and ordinal notations}
@@
 \end{remark}
-
-\section{Ordinal Notations and \gaia's ordinals}
-\gaiasign Module~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge} contains an instance of class \texttt{ON E0lt compare}, where \texttt{E0lt} is the order on the well-formed ordinal terms below $\epsilon_0$ defined in \gaiaHydras library (please see Chapter~\ref{gaia-chapter}).
-
-%------------------------------------------------------------------------
-
-\chapter[A proof of termination, using epsilon0]{A proof of termination, using ordinals below \texorpdfstring{$\epsilon_0$}{Epsilon0}}
-
-\label{cnf-math-def}
-\label{chap:T1}
-
-In this chapter, we adapt to \coq{} the well-known~\cite{KP82}  proof that Hercules eventually wins every battle, whichever the strategy  of each player.
-In other words, we present  a formal and self contained proof of termination  of all [free] hydra battles.
-First, we take from Manolios and Vroon~\cite{Manolios2005} a representation of the ordinal $\epsilon_0$ as terms in Cantor normal form.
-Then, we define a variant for hydra battles as a measure that maps any hydra to some ordinal strictly less than $\epsilon_0$.
-
-\section{The ordinal \texorpdfstring{\(\epsilon_0\)}{epsilon0}}
-\label{sec:epsilon0-intro}
-
-The ordinal \(\epsilon_0\) is the least ordinal number that satisfies 
-the equation \(\alpha = \omega^\alpha\), where \(\omega\) is 
-the least infinite ordinal.
-Thus, we can intuitively consider \(\epsilon_0\) as an
-\emph{infinite} \(\omega\)-tower.
-
-\subsection{Cantor normal form}
-\index{maths}{Cantor normal form}
-
-Any ordinal strictly less that \(\epsilon_0\) 
-can be finitely represented by a unique  \emph{Cantor normal form}, 
-that is, an expression  which is either  the ordinal \(0\) or 
-a sum  \(\omega^{\alpha_1} \times n_1 + \omega^{\alpha_2} \times n_2 + 
-  \dots + \omega^{\alpha_p} \times n_p\) where all the \(\alpha_i\) 
-are ordinals in Cantor  normal form, \(\alpha_1 > \alpha_2 > \alpha_p\), 
-and all the \(n_i\) are positive integers.
-
-An example of Cantor normal form is displayed in Fig \ref{fig:cnf-example}:
-Note that  any ordinal of
-the form \(\omega^0 \times i + 0\) is just written \(i\).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
-\node[color=blue]{$\omega^{(\omega^\omega\,+\, \omega^2 \times 8 \,+\, \omega)}+ \omega^\omega + \omega^4+ 6$};
-\end{tikzpicture}
-\caption{\label{fig:cnf-example}
-An ordinal in Cantor normal form}
-\end{figure}
-
-In the rest of this section, we define an inductive type for representing in \texttt{Coq}
-all the ordinals strictly  less than  \(\epsilon_0\), then extend some arithmetic operations
-to this type, and finally prove that our representation fits well with 
-the expected mathematical properties: the order we define is a well order, 
-and the decomposition into Cantor normal form  is consistent 
-with the implementation of the arithmetic operations of exponentiation of base \(\omega\) 
-and addition.
-
-\paragraph*{Remark}
-\label{sec:orgheadline65}
-Unless explicitly mentioned, the term ``ordinal" will be used instead of
-``ordinal strictly less than \(\epsilon_0\)" (except in Chapter~\ref{chap:schutte} where it stands for ``countable ordinal'').
-
-\subsection{A data type for  ordinals in Cantor normal form}
-\label{sec:orgheadline72}
-\label{sec:T1-inductive-def}
-
-% Our user contribution~\cite{CantorContrib} represents 
-% the set of ordinals strictly less than $\epsilon_0$ in Cantor normal form as in~\cite{Manolios2005}, and also the set
-% of ordinals strictly  less than $\Gamma_0$ in Veblen normal form.
-
-    Let us define an inductive type whose 
-constructors are respectively associated
-with the ways to build Cantor normal forms:
-
-\begin{itemize}
-\item the ordinal \(0\)
-\item the construction \((\alpha,\, n,\,\beta)  \mapsto \omega^\alpha \times (n + 1)+ \beta \quad (n\in\mathbb{N})\)
-\end{itemize}
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#T1}{Epsilon0.T1}}
-
-\label{types:T1}
-%\index{Constants!zero:T1}
-\index{hydras}{Library Epsilon0!Types!T1}
-
-\input{movies/snippets/T1/T1Def}
-
-\subsubsection{Example}
-
-\label{alpha0-def}
-For instance, the ordinal  $\omega^\omega+\omega^3\times 5+2$ is represented by the following term of type \texttt{T1}:
-
-\input{movies/snippets/T1/alpha0}
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5, level 1/.style={sibling distance=6cm},
-level 2/.style={sibling distance=35mm},  
-level 3/.style={sibling distance=17mm}]
-\node  {cons}
-  child {  node {cons}
-            child { node {cons} child {node {zero}} child {node{0}} child{node{zero}}}
-         child {node {0}}
-         child {node {zero}}}
-    child {node {0}}
-   child {node {cons} 
- child { node {cons} child {node {zero}} child {node{2}} child{node{zero}}}
-  child {node {4}}
-         child {node {cons} child {node {zero}} child {node{1}} child{node{zero}}}};
-
-\end{tikzpicture}
-
-\caption{The tree-like representation of the ordinal $\omega^\omega+\omega^3\times 5 +2$\label{fig:cnf-tree}}
-
-\end{figure}
-
-\paragraph{Remark}
-For simplicity's sake, we chose to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the construction (\texttt{cons $\alpha$ $n$ $\beta$}) is intended to represent the
-ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
-In a future version, we should replace  the type \texttt{nat} with \texttt{positive} in \texttt{T1}'s 
-definition. But this replacement would take a lot of time \dots{}
-
-\paragraph{Remark}
-The name \texttt{T1} we gave to this data-type  is proper to this development and refers
-to a hierarchy of ordinal notations.
-For instance, in Library \href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2},  the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{Gaia, schutte}).
-\noindent
-
-\input{movies/snippets/T2/T2Def}
-
-\subsection{About Gaia}
-\gaiasign Chapter~\vref{gaia-chapter} describes the present state of a
-project of making compatible the libraries \HydrasLib and \gaia. At present, both libraries contain their own version of the inductive type \texttt{T1}, with the same base name for the constructors.
-
-\inputsnippets{nfwfgaia/T1def}
-
-Thus, many examples of this and following chapters can be run
-in \gaia's context. We will signal any difference (notation, name) which may appear.
-A closer integration (same data type and functions) is still in project.
-
-\begin{remark}
-  \label{remark:hydra-gaia-renaming}
-  \textcolor{lookcolor}{In future releases, we plan to make some
-    \HydrasLib identifiers progressively  deprecated, in favour of \gaia's names.}
-\end{remark}
-
-\subsection{Abbreviations}
-
-Some abbreviations may help to write more concisely complex ordinal terms.
-
-\subsubsection{Finite ordinals}
-\label{sec:orgheadline67}
-
-For representing finite ordinals, \emph{i.e.} natural numbers, we first introduce a notation for terms of the form $n+1$, then define a coercion from type \texttt{nat} into \texttt{T1}.
-\label{sect:notation-FS}
-
-\label{sect:notation-F}
-
-\input{movies/snippets/T1/finiteOrds}
-
-\paragraph*{\gaiasign}
-  In \texttt{gaia.ssete9}, the $n$-th finite ordinal is also written
-  \texttt{$\backslash${F}\,$n$}. There is no coercion in \gaia from \texttt{nat} to \texttt{T1}.
-
-  \vspace{4pt}
-
-  From~\href{../theories/html/gaia_hydras.HydraGaia_Examples.html}{gaia\_hydras.HydraGaia\_Examples}.
-
-  \inputsnippets{HydraGaia_Examples/Demoa}
-
-% \index{Coq!Coercions}
-% \index{Functions!Coercions@Coercions (from nat to ordinal types)}
-% \begin{remark}
-% Please refer to the remark~\pageref{warning:coercions} about the use of coercions.
-% % The use of coercions like \texttt{fin} allow us to be close to the mathematical tradition where natural numbers are ordinals too.
-% % Nevertheless, it may happen that a goal like \texttt{3 < 5} could be 
-% % interpreted as \texttt{(lt (fin 3) (fin 5))},  depending on the current notation scope.  
-% % When this misinterpretation happens, tactics like \texttt{auto with arith}, \texttt{lia} do not work!
-% % Thus, it is useful to write \texttt{(3 < 5)\%nat}  an inequality between two natural numbers. 
-% \end{remark}
-
-\subsubsection{The ordinal \(\omega\)}
-\label{sec:orgheadline68}
-
-  Since \(\omega\)'s Cantor normal form is
-i.e. \(\omega^{\omega^0}\times 1+ 0\), we can define the following abbreviation:
-
-\label{sect:omega-notation2}
-
-\input{movies/snippets/T1/omegaDef}
-
-Note that \texttt{omega} is not an identifier in \HydrasLib, thus any tactic like \texttt{unfold omega} would fail.
-
-\paragraph*{\gaiasign}
-In \texttt{gaia.ssete9}, the ordinal $\omega$ is bound to the \emph{constant}  \texttt{T1omega}.
-
-\subsubsection{The ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
-\label{sect:notation-phi0}
-We provide also a notation for ordinals of the form $\omega^\alpha$.
-
-\index{hydras}{Library Epsilon0!Notations!phi0@phi0 (exponential of base omega)}
-
-\input{movies/snippets/T1/phi0Def}
-
-\index{maths}{Additive principal ordinals}
-
-\begin{remark}
-\label{sec:orgheadline69}
-The name \(\phi_0\)
-   comes from ordinal numbers theory. In~\cite{schutte}, Schütte defines 
-$\phi_0$  as the ordering (\emph{i.e.} enumerating) function of the set  of \emph{additive principal ordinals} \emph{i.e.} strictly positive ordinals $\alpha$ that verify $\forall \beta<\alpha, \beta+\alpha=\alpha$.
-For Schütte,  $\omega^\alpha$ is just a notation for $\phi_0(\alpha)$.
-See also Chapter~\ref{chap:schutte} of this document.
-\end{remark}
-
-\paragraph*{\gaiasign}
-In \texttt{gaia.ssete9} the identifier \texttt{phi0} is bound to a plain constant (not a notation).
-
-\subsubsection{The hierarchy of \(\omega\)-towers:}
-\label{sec:orgheadline71}
-
-The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximated by the sequence of $\omega$-towers (see also Sect~\vref{sect:epsilon0-as-limit} ).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/towerDef}
-
-For instance, Figure~\ref{fig:tower7} represents  the ordinal returned by the
- evaluation of the term \texttt{omega\_tower 7}.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
-\node[color=blue]{$\omega^{{{\omega}^{{{\omega}}^{{{\omega}}^{{\omega^{{\omega}^{\omega}}}}}}}}$};
-\end{tikzpicture}
-\caption{\label{fig:tower7}
-The $\omega$-tower of height 7}
-\end{figure}
-
-\subsection{Pretty-printing ordinals in Cantor normal form}
-\label{sect:ppT1}
-\index{hydras}{Library Epsilon0!Types!ppT1}
-
-Let us consider again the ordinal $\alpha_0$ defined in section~\vref{alpha0-def}. 
-If we ask \coq{} to print its  normal form, we get a hardly readable term of type \texttt{T1}.
-
-\input{movies/snippets/T1/alpha0Compute}
-
-The following data type defines a more readable abstract syntax for  ordinals terms in Cantor normal form:
-
-\label{types:ppT1}
-\index{hydras}{Library Epsilon0!Functions!pp@ pp (pretty printing terms in Cantor normal form)}
-
-\input{movies/snippets/T1/ppT1Def}
-
-The function \texttt{pp: T1 -> ppT1} converts any closed term of type \texttt{T1} into a human-readable expression. For instance, let us convert the term \texttt{alpha\_0}.
-
-\input{movies/snippets/T1/ppAlpha0}
-
-\paragraph*{\gaiasign}
-In ~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge},
-we define a variant of \texttt{pp} for pretty-printing terms of type \texttt{ssete9.CantorOrdinal.T1} (see Sect.~\vref{sect:gaia-ppT1}).
-
-\index{hydras}{Projects}
-\begin{project}
-Design  (in \ocaml?) a set of tools for systematically pretty printing ordinal terms in Cantor normal form.
-\end{project}
-
-\subsection{Comparison between ordinal terms}
-\label{sec:orgheadline73}
-
-% Our formalisation of Cantor Normal Form will take two steps:
-% 1 Definition of a strict order \texttt{o<} on the type \texttt{T1}, 
-% 2 Using \texttt{o<} for characterizing terms in normal form.
-
-In order to compare two terms of type \texttt{T1}, we define a recursive function \texttt{compare} that maps two ordinal tems $\alpha$ and $\beta$ to a value of type \texttt{comparison}.
-This type is defined in \coq's standard library
-\texttt{Init.Datatypes} and
-contains three constructors:  \texttt{Lt} (less than), \texttt{Eq} (equal), and
-\texttt{Gt} (greater than).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/compareDef}
-
-\label{Predicates:lt-T1}
-Please note that this definition of \texttt{lt} makes it easy to write proofs by computation, as shown by the following examples.
-
-\vspace{4pt}
-
-\input{movies/snippets/T1/ltExamples}
-
-Links between the function \texttt{compare} and the relations
-\texttt{lt} and \texttt{eq} are established through the following lemmas (~\vref{sect:comparable-def}).
-\vspace{4pt}
-
-\input{movies/snippets/T1/Instances}
-
-\paragraph*{\gaiasign}
-In \gaia, the strict order \texttt{T1lt} on ordinal terms is directly defined as a boolean function (in \ssreflect's style).
-
-\inputsnippets{nfwfgaia/T1ltDef}
-
-In section~\vref{sect:lt-compat-gaia}, we show that both definitions are mutually equivalent.
-
-\subsubsection{A Predicate for Characterizing Normal Forms}
-\label{sect:t1-nf}
-
-\label{sec:orgheadline74}
-\label{sec:orgheadline75}
-Our data-type \texttt{T1} allows us to write expressions that
-are not properly in Cantor normal form as specified in Section \ref{sec:epsilon0-intro}.
-For instance, consider the following term of type  \texttt{T1}.
-
-\input{movies/snippets/T1/badTerm}
-
-This term would have been written \(\omega^1\times 2 + \omega^\omega \times 3\) in the usual mathematical notation.
-We note that the exponents of $\omega$ are not in the right (strictly decreasing) order.
-Nevertheless, with the help of the order \texttt{lt} on \texttt{T1}, we are now able to characterize
-the set of all well-formed ordinal terms:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#nf_b}{Epsilon0.T1}}
-
-\label{Predicates:nf-T1}
-
-\input{movies/snippets/T1/nfDef}
-
-\input{movies/snippets/T1/nfAlpha0}
-
-\input{movies/snippets/T1/nfBadTerm}
-
-\paragraph*{\gaiasign}
-In \gaia, the boolean function which characterizes ordinal terms in normal form is defined as follows:
-
-\inputsnippets{nfwfgaia/T1nfDef}
-
-In Sect.~\vref{nf-gaia-compat}, we show that \gaia's \texttt{T1nf} is extensionally equivalent with \HydrasLib'  \texttt{nf}.
-
-\subsection{Making normality implicit}
-  We would like to get rid of terms of type \texttt{T1} which are not in Cantor normal form.
-A simple way to do this is to consider statements of the form 
-\texttt{forall alpha: T1, nf alpha -> $P$ alpha}, where $P$ is a predicate over type \texttt{T1}, like in the following lemma \footnote{Ordinal addition is formally defined a little later (page~\ref{sect:infix-plus-T1})}.
-
-\input{movies/snippets/T1/plusIsZero}
-
-\vspace{4pt}
-
-But this style leads to clumsy statements, and generates too many subgoals in interactive proofs (although often solved with \texttt{auto} or \texttt{eauto}).
-
-One may encapsulate conditions of the form \texttt{(nf $\alpha$)} in
-the most used predicates. For instance, we introduce the restriction of \texttt{lt} to terms in normal form, and provide a handy notation for this restriction.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Prelude.Restriction.html}{hydras.Prelude.Restriction}}
-
-\input{movies/snippets/Restriction/restrictionDef}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#LT}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/LTDef}
-
-\label{Predicates:LT-T1}
-
-For instance, in the following lemma, the condition that $\alpha$ is in normal form is included in the condition $\alpha< 1$.
-
-\input{movies/snippets/T1/LTOne}
-
-\subsubsection{\texttt{E0}: a sigma-type for \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-As we noticed in Sect.~\ref{sect:t1-nf}, the type \texttt{T1} is not a correct ordinal notation, since it contains terms that are not in Cantor normal form.
-In certain contexts (for instance in Sections~\ref{sect:L-equations}, \ref{sect:hardy},
-and \ref{sect:wainer}),  we need to define total recursive functions on well-formed ordinal terms less  than $\epsilon_0$, using the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777}.
- In order to define a type whose inhabitants represent just ordinals, we build a type gathering a term of type \texttt{T1} and a proof that this term is in normal form.
-
-\label{sect:E0-def}
-\label{types:E0}
-\index{hydras}{Library Epsilon0!Types!E0}
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html\#E0}{Epsilon0.E0}}
-
-\input{movies/snippets/E0/E0Def}
-
-Many constructs : types, predicates, functions, notations, etc., on type \texttt{T1} are adapted to \texttt{E0}.
-
-First, we declare a notation scope for \texttt{E0}, then we redefine the predicates of comparison.
-
-\input{movies/snippets/E0/E0Scope}
-\label{Predicates:Lt-E0}
-\input{movies/snippets/E0/LtLeDef}
-
-Equality in \texttt{E0} is just Leibniz equality. Note that, since \texttt{nf} is
-defined by a Boolean function, for  any term $\alpha:\texttt{T1}$, there exists at most one proof of \texttt{nf $\alpha$}, thus two ordinals of type \texttt{E0} are
-equal if and only if their projection to \texttt{T1} are equal (see also Sect.~\vref{sect:eq-proof-unicity}).
-
-\vspace{4pt}
-
-\index{coq}{Unicity of equality proofs}
-
-\inputsnippets{E0/nfProofUnicity}
-\inputsnippets{E0/E0EqIff}
-
-In order to  upgrade constants and functions from type \texttt{T1} to \texttt{E0}, we have to prove that
-the term they build is in normal form.
-For instance, let us represent the ordinals $0$ and $\omega$   as instances of the class \texttt{E0}.
-
-\vspace{4pt}
-
-\label{sect:omega-T1}
-
-\input{movies/snippets/E0/ZeroOmega}
-
-%\index{Constants!zero:T1}
-
-\paragraph*{\gaiasign}
- The library~\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines a \gaia-compatible type \texttt{E0}.
-
-\subsection{Syntactic definition of limit and successor ordinals}
-
-Pattern matching and structural recursion allow us to define boolean characterizations  of successor and limit ordinals.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/succbLimitb}
-
-The correctness of these definitions with respect to the mathematical notions of
-limit and successor ordinals is established through several lemmas. For instance,
-Lemma \texttt{canonS\_limit\_lub}, page~\pageref{lemma:canonS-limit}, shows that
-if $\alpha$ is (syntactically) a limit ordinal, then it is the least upper bound of
-a strictly increasing sequence of ordinals.
-
-\paragraph*{\gaiasign}
-In \gaia, the boolean functions associated with limit and successor ordinals are called \texttt{T1limit} and \texttt{T1is\_succ}.
-
- From~\href{../theories/html/gaia_hydras.HydraGaia_Examples.html}{gaia\_hydras.HydraGaia\_Examples}.
-
-\inputsnippets{HydraGaia_Examples/Demob}
-
-\subsection{Arithmetic on \texorpdfstring{$\epsilon_0$}{epsilon0}}
-\subsubsection{Successor}
-
-\index{hydras}{Library Epsilon0!Functions!succ}
-
-The successor of any ordinal $\alpha< \epsilon_0$ is defined by structural 
-recursion on its Cantor normal form.
-
-\label{Functions:succ-T1}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/succDef}
-
-The following lemma establishes the connection between the  function
-\texttt{succ} and the Boolean predicate \texttt{succb}.
-
-\vspace{4pt}
-
-\input{movies/snippets/T1/succbIff}
-
-\begin{exercise}[\gaiasign]
-  Look for the \gaia-theorem which corresponds to \texttt{succb\_iff}.
-\end{exercise}
-
-\subsubsection{Successor function on \texttt{E0}}
-
-The function \texttt{succ} on \texttt{T1} is extended to \texttt{E0} the following way:
-
- \emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
-
- \inputsnippets{E0/SuccOnE0}
-
-\index{hydras}{Exercises}
- \begin{exercise}
-Prove in \coq{} that for any ordinal $0< \alpha<\epsilon_0$, $\alpha$ is a limit if 
-and only if for all $\beta<\alpha$, the interval $[\beta,\alpha)$ is
-infinite.
-
-\emph{You may start this exercise with the file
-     \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/Limit_Infinity.v}{exercises/ordinals/Limit\_Infinity.v}.}
- \end{exercise}
-
-\subsubsection{Addition and multiplication}
-
-Ordinal addition and multiplication are also defined by structural recursion over the type \texttt{T1}. Please note that they use the \texttt{compare} function on some subterms of their arguments.
-
-\label{sect:infix-plus-T1}
-
-\input{movies/snippets/T1/plusDef}
-
-\input{movies/snippets/T1/multDef}
-
-\paragraph*{\gaiasign}
-We keep \gaia's base names for
-addition and multiplication of ordinal terms below $\epsilon_0$.
-Please refer to Sect.~\ref{sect:plus-mult-gaia-hydras} about compatibility of both arithmetics.
-
-\subsubsection{Examples}
-
-The following examples are instances of \emph{proofs by computation}. Please note that  addition and multiplication on \texttt{T1}
-are not commutative. Moreover,  both operations fail to be strictly monotonous in their first argument.
-
-\input{movies/snippets/T1/plusMultExamples}
-
-\input{movies/snippets/T1/notMono}
-
-The function \texttt{succ} is related with addition through the following lemma:
-
-\inputsnippets{T1/succIsPlusOne, T1/succIsPlusOnez}
-
-\subsubsection{Arithmetic on type \texttt{E0}}
-
- We define an addition in type \texttt{E0}, since the sum of two terms in normal form is in normal form too.
-
-\input{movies/snippets/T1/plusNf}
-
-\input{movies/snippets/E0/plusE0}
-\input{movies/snippets/E0/CheckPlus}
-
-\begin{remark}
-In all this development, two representations of ordinals co-exist: ordinal terms (type \texttt{T1}, notation scope \texttt{t1\_scope}, for reasoning on the tree-structure of Cantor normal forms), and ordinal terms \emph{known to be in normal form} (type \texttt{E0}, notation scope \texttt{E0\_scope}).
-Looking at the contexts displayed by \coq{} prevents you from any risk of confusion.
-\end{remark}
-
-%% To simplify !
-\index{hydras}{Exercises} 
-\begin{exercise}
-Prove that for any ordinal $\alpha:\texttt{E0}$, 
-$\omega\leq \alpha$ if and only if, for any natural number $i$,
-$i+\alpha=\alpha$.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/ge_omega_iff.v}{exercises/ordinals/ge\_omega\_iff.v}.}
-\end{exercise}
-
-\subsection{A proof by computation}
-\label{Ex42-E0}
-
-It is interesting to compare the following proof of the equality
-$\omega+42+\omega^2$ with the more theoretical proof in Sect~\vref{Ex42-schutte}.
-
-\inputsnippets{E0/Ex42}
-
-\section{Well-foundedness and transfinite induction}
-
-\index{maths}{Transfinite induction}
-
-\subsection{About  well-foundedness}
-\label{sec:T1Wf}
-\label{sec:orgheadline82}
-   In order to use \texttt{T1} for proving termination results,
-we need to prove that  our order \texttt{<} is well-founded. Then we will get \emph{transfinite induction} for free.
-
-The proof of well-foundedness of the strict order $<$ on Cantor normal forms is already 
-available in the Cantor contribution by Castéran and Contejean~\cite{CantorContrib}. That proof relies on a library on recursive path orderings written by
-E. Contejean. We present here  a direct proof of the same result, which does not require any knowledge on r.p.o.s.
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that the \emph{total} order \texttt{lt} on \texttt{T1} is not well-founded. 
-\textbf{Hint:}  You will have to build a counter-example with terms of type \texttt{T1}
-which are not in Cantor normal form.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/T1_ltNotWf.v}{exercises/ordinals/T1\_ltNotWf.v}.}
-\end{exercise}
-
-\subsubsection{A first attempt}
-\label{sec:orgheadline77}
-%\index{coq}{Well-founded induction}
-
-It is natural to try to prove by structural induction over \texttt{T1} 
-that every term in normal form is \texttt{LT}-accessible.
-
-Unfortunately, it won't work. Let us consider some well-formed term
- $\alpha=\texttt{cons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are \texttt{LT}-accessible.
-In order to prove the accessibility of $\alpha$, we have to consider
-any well formed term \(\delta\) such that \(\delta<\alpha\). 
-But nothing guarantees that \(\delta\)  is strictly  less than \(\beta\) nor \(\gamma\), and we cannot use the induction hypotheses on   \(\beta\) nor \(\gamma\).
-
-\input{movies/snippets/T1/wfLTBada}
-
-The problem comes from the too weak hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
-\(\gamma\);
-for instance \(\delta\) may be of the form
-\texttt{cons $\beta$ $p'$  $\gamma'$},
-where    \(p' < n\).
-Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.
-
-\input{movies/snippets/T1/wfLTBadz}
-
-\subsubsection{Using a stronger inductive predicate.}
-\label{sec:orgheadline78}
-\label{sec:strongly-accessible}
-
-  Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
-through \texttt{LT}, we propose to consider the following (stronger) predicate:
-
-\input{movies/snippets/T1/AccStrongDef}
-
-The following lemma is an application of the strict inequality 
-$\alpha < \omega ^\alpha$. If $\alpha$ is strongly accessible, then, by definition,
-$\omega^\alpha$ is accessible, thus $\alpha$ is \emph{a fortiori} accessible.
-
-\input{movies/snippets/T1/AccStrongStronger}
-
-Thus, it remains to prove that every ordinal strictly less than $\epsilon_0$
-is strongly accessible.
-
-% \subsubsection{Structure of the proof of well-foundedness of \texttt{LT}}
-
-\label{sec:orgheadline81}
-\label{proof-wf-epsilon0}
-\paragraph{A helper}
-\label{sec:orgheadline79}
-
-First, we prove that, for  any \texttt{LT}-accessible term $\alpha$, $\alpha$ is 
-strongly accessible too (\emph{i.e.} any well formed
-term (\texttt{cons $\alpha$ $n$ $\beta$})  is accessible).
-
-The proof is structured as an induction on $\alpha'$s accessibility. Let us consider
-any  accessible term $\alpha$.
-
-\input{movies/snippets/T1/AccImpAccStrong}
-
-Let \texttt{n:nat} and \texttt{beta:T1} such that (\texttt{cons alpha n beta}) is in normal form. 
-We prove first that \texttt{beta} is accessible,  which allows us to prove by well-founded induction on \texttt{beta}, 
-and natural induction on \texttt{n}, that (\texttt{cons alpha n beta}) is accessible.
-The proof, quite long, can be consulted in \href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}.
-
-\paragraph{Accessibility of any well-formed ordinal term}
-\label{sec:orgheadline80}
-
-Our goal is still to prove accessibility of any well formed ordinal term.
-Thanks to our previous lemmas, we are almost done (by a simple structural induction!).
-
-\input{movies/snippets/T1/nfAcc}
-
-\input{movies/snippets/T1/T1Wf}
-
-\index{maths}{Transfinite induction}
-
-Traditionnally, well-founded induction on ordinals is called \emph{transfinite induction}.
-
-\inputsnippets{T1/transfiniteRecursor, T1/transfiniteInduction}
-
-As a corollary, the order \texttt{Lt} on type \texttt{E0} is well-founded too.
-
-\inputsnippets{E0/E0LtWf}
-
-\begin{remark}
-\label{remark:a3pat}
-The alternate proof of well-foundedness using \'Evelyne Contejean's work on
-    recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the
-    library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0.Epsilon0rpo}.
- \end{remark}
-
-\subsection{An ordinal notation for  \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-We build an instance of \texttt{ON}, and prove its correctness w.r.t. Schutte's model (see Chapter~\ref{chap:schutte}).
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
-
-\input{movies/snippets/E0/InstanceEpsilon0}
-
-\label{instance-epsilon0}
-
-\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
-
-\input{movies/snippets/Correctness_E0/injectDef}
-
-\input{movies/snippets/Correctness_E0/Epsilon0Correct}
-
-\index{hydras}{Projects}
-\begin{project}
- \emph{This exercise is a continuation of Project~\vref{exo:ON-mult}.}
-Use \texttt{ON\_mult} to define an ordinal notation \texttt{Omega2} for $\omega^2=\omega\times\omega$.
-
-Prove that \texttt{Omega2} is a sub-notation of \texttt{Epsilon0}.
-
-Define on \texttt{Omega2} an addition compatible with the addition on \texttt{Epsilon0}.
-
-\textbf{Hint}. You may use the following definition (in 
-    \href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}).
-
-\input{movies/snippets/ON_Generic/SubONSameOp}
-
-     \end{project}
-
-    \index{hydras}{Projects}
-    \begin{project}
-    The class \texttt{ON} of ordinal notations has been defined long after this 
-    chapter, and is not used in the development of the type \texttt{E0} yet.
-    A better integration of both notions should simplify the development on ordinals in Cantor normal form. This integration is planned for the future versions.
-
-    \end{project}
-
-    \section{A refinement of \texttt{E0} : an ordinal notation for \texorpdfstring{$\omega^\omega$}{omega\^omega}}
-
-    In Module   \href{https://github.com/coq-community/hydra-battles/blob/master/theories/
-      ordinals/OrdinalNotations/OmegaOmega.v}{theories/ordinals/OrdinalNotations/OmegaOmega.v},
-    we represent ordinals below $\omega^\omega$ by lists of pairs of natural numbers (with the same coefficient shift as in \texttt{T1}).
-    For instance, the ordinal $\omega^4\times 10 + \omega^3 + \omega+ 5$ is represented by the list \texttt{(4,9)::(3,0)::(1,0)::(0,4)::nil}.
-
-\inputsnippets{OmegaOmega/LODef}
-
-  The usual operations : \texttt{succ}, \texttt{+}, \texttt{*} are simple variants of the same operations in \texttt{T1}.
-
-  \inputsnippets{OmegaOmega/succPlusMult}
-
-    We establish this representation as a \emph{refinement} of the data types we used to represent ordinals less than $\epsilon_0$.
-Thus, many properties like well-foundedness of $<$ and associativity of $+$,  of this ordinal notations have very short proofs.
-
-    \inputsnippets{OmegaOmega/refineDef, OmegaOmega/phi0Ref,
-      OmegaOmega/succRef, OmegaOmega/plusRef,
-      OmegaOmega/multRef}
-
-    In order to make an ordinal notation for $\omega^\omega$, we follow the same steps as for $\epsilon_0$:
-    \begin{enumerate}
-    \item Define an order \texttt{lt}, which refines the order
-      \texttt{lt} on \texttt{T1}.
-      \inputsnippets{OmegaOmega/compareDef}
-      \inputsnippets{ OmegaOmega/ltDef, OmegaOmega/ltRef}
-
-    \item Define the predicate ``to be in normal form''.
-      \inputsnippets{OmegaOmega/nfDef}
-    \item Define a class \texttt{OO} of terms in normal form, and an embedding from \texttt{E0} into \texttt{OO}.
-      \inputsnippets{OmegaOmega/OODef}
-       \inputsnippets{OmegaOmega/embedDef}
-      \item Infer well-foundedness of the order on \texttt{OO}.
-    \inputsnippets{OmegaOmega/ltWf}
-
-    \end{enumerate}
-
-    \inputsnippets{OmegaOmega/ONOO, OmegaOmega/OOz}
-
-    Let us show a few examples.
-
-    \inputsnippets{OmegaOmega/OODemo}
-
-    \index{hydras}{Exercises}
-
-    \subsection{Related work}
-    The article~\cite{Manoury2016}  defines another representation of ordinals below $\omega^\omega$ based on lists of natural numbers.
-
-\begin{exercise}[not trivial]
-  It may be interesting to write a \emph{direct} proof  of well-foundedness of the order in $\omega^\omega$ (\emph{i.e.}
-  without using properties of $\epsilon_0$). This exercise may
-  help to understand better the proof structure of Sect.~\vref{sec:strongly-accessible}.
-    \end{exercise}
-%ici
-    \section{A variant for hydra battles}
-
-    In order to prove the termination of any hydra battle, we try to define a variant mapping hydras to ordinals strictly less than $\epsilon_0$.
-    In order to make such a variant easy to define (for instance by a structural recursion), we introduce a variant of addition, which, contrary to
-    $+$, is commutative and strictly monotonous in both of its arguments. This last property makes it possible to prove that our function is 
-    truly a variant for hydra battles (in Sect.~\vref{sect:variant-decr}).
-
-    \subsection{Natural sum (a.k.a. Hessenberg's  sum)}
-    \label{sec:orgheadline87}
-    \label{hydra-variant}
-    \label{sect:hessenberg-def}
-
-    Natural sum (Hessenberg's  sum) is a commutative and monotonous version of
-    addition. It is used as an auxiliary operation  for defining variants
-    for hydra battles, where Hercules is allowed to chop off any  head of the hydra.
-
-    In the literature, the natural sum of ordinals \(\alpha\) and \(\beta\)
-    is often denoted by \(\alpha \# \beta\)  or  \(\alpha \oplus  \beta\).
-    Thus we called \texttt{oplus} the associated \emph{Coq} function.
-
-    \subsubsection{Definition of \texttt{oplus}}
-    \label{sec:orgheadline84}
-    %\index{Functions!oplus @ oplus (Hessenberg commutative sum)}
-
-    The definition of \texttt{oplus} is recursive in both of its 
-    arguments and uses  the same pattern as for the \texttt{merge} function on lists of library
-    \texttt{Coq.Sorting.Mergesort}.
-
-    \begin{enumerate}
-    \item Define a nested recursive function, using the \texttt{Fix} 
-        construct
-
-    \item Build a principle of induction dedicated to \texttt{oplus}
-
-    \item Establish equations associated to each case of the definition.
-    \end{enumerate}
-
-    \paragraph{Nested recursive definition}
-    \label{sec:orgheadline83}
-
-    The following definition is composed of 
-    \begin{itemize}
-    \item A main function \texttt{oplus}, structurally recursive in its 
-    first argument \texttt{alpha}
-    \item An auxiliary function \texttt{oplus\_aux} within the scope of \texttt{alpha},
-    structurally recursive in its argument \texttt{beta};  \texttt{oplus\_aux beta} 
-       is supposed to compute  \texttt{oplus alpha beta}.
-    \end{itemize}
-
-%  \pagebreak 
-  \vspace{4pt}
-    \emph{From Module~\href{../theories/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}
-
-    \label{sect:infix-oplus}
-
-    \input{movies/snippets/Hessenberg/oplusDef}
-
-    The reader will note that each recursive call of the functions
-    \texttt{oplus} and \texttt{oplus\_aux} satisfies \emph{Coq}'s constraint
-    on recursive definitions. The function \texttt{oplus} is recursively called on a sub-term of its first argument,
-    and \texttt{oplus\_aux} on a sub-term of its unique argument.
-    Thus, \texttt{oplus}'s definition is accepted by \coq{} as a structurally recursive function.
-
-    \subsubsection{Rewriting lemmas}
-    \label{sec:orgheadline86}
-
-    \emph{Coq}'s constraints on recursive definitions result in 
-    the quite  complex form of \texttt{oplus}'s definition.
-    Proofs of properties of this function can be simpler if we
-     derive a few  rewriting lemmas that will help to simplify 
-    expressions of the form (\texttt{oplus $\alpha$ $\beta$}).
-
-    A first set of lemmas correspond to the various cases of \texttt{oplus}'s 
-    definition. They can be proved almost immediately. Here are a few examples.
-
-    \inputsnippets{Hessenberg/oplusNeutral,
-      Hessenberg/oplusCompareLt}
-
-    % \subsubsection{A hand-made induction principle}
-    % \label{sec:orgheadline85}
-
-    % \index{Coq!Commands!Functional Scheme}
-
-    % \emph{Coq} contains a command  \texttt{Functional Scheme} that 
-    % generates induction principles which correspond to recursive functions.
-    % Unfortunately, the current version ( \texttt{8.11.0} ) doesn't work on \texttt{oplus},
-    % probably because of the inner \texttt{Fix}.
-
-    % \begin{Coqsrc}
-    % Functional Scheme oplus_ind := Induction for oplus Sort Prop.
-    % \end{Coqsrc}
-
-    % \begin{Coqanswer}
-    % Error: Anomaly "todo." Please report at http://coq.inria.fr/bugs/.
-    % \end{Coqanswer}
-
-    % Fortunately, it's a good exercise for a semi-experienced user, to write
-    % her/him-self induction principles similar to the ones returned by
-    % \texttt{Functional Scheme}.
-
-    % \begin{itemize}
-    % \item First, we choose to write a version for sort \texttt{Type}, since versions
-    % for sorts \texttt{Prop} and \texttt{Set} can be easily derived from
-    % the former one. According to \emph{Coq}'s naming politics, we will call our 
-    % principle \texttt{oplus\_rect}
-
-    % \item The conclusion of \texttt{oplus\_rect} will be (\texttt{$P$ a b (oplus a b)}),
-    % where $P$ is an arbitrary function of type 
-    % \texttt{T1 -> T1 -> T1 -> Type}
-
-    % \item The premises of \texttt{oplus\_rect} will describe how to build an induction 
-    % on the graph of \texttt{oplus}.
-    % \end{itemize}
-
-    % We are now ready to state and prove \texttt{oplus\_rect}, and the reader
-    % will note that the statement is longer than the proof script itself,
-    % which is a standard proof by induction, simplification and case-analysis 
-    % that follows  \texttt{oplus}'s definition.
-
-    % We associate also a tactic to the application of \texttt{oplus\_rect}.
-
-    % \begin{Coqsrc}
-    %  Lemma oplus_rect:
-    %       forall P: T1 -> T1 -> T1 -> Type, 
-    %         (forall a:T1, P zero a a) ->
-    %         (forall a: T1, P a zero a) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Gt ->
-    %            P b1 (cons a2 n2 b2) o ->
-    %            P (cons a1 n1 b1) (cons a2 n2 b2)
-    %              (cons a1 n1 o)) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Lt ->
-    %            P (cons a1 n1 b1) b2 o ->
-    %            P (cons a1 n1 b1) (cons a2 n2 b2) 
-    %            (cons a2 n2 o)) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Eq ->
-    %            P b1 b2 o ->
-    %           P (cons a1 n1 b1) (cons a2 n2 b2)
-    %             (cons a1 (S (n1 + n2)%nat) o)) ->
-    %          forall a b, P a b (oplus a b).
-    % Proof with auto.
-    %    induction a.
-    %    -    intro; simpl; destruct b;auto.
-    %    -   induction b.
-    %        + apply X0.
-    %        + case_eq (compare a1 b1).
-    %          * intro Comp; unfold oplus; rewrite Comp.
-    %            cbn; apply X3 ...
-    %          * intro Comp; cbn; rewrite Comp; apply X2...
-    %          * intro Comp; cbn; rewrite Comp ...
-    %  Defined.
-
-    % Ltac oplus_induction a b:= pattern (oplus a b); apply oplus_rect.
-    % \end{Coqsrc}
-
-    % \index{Exercises}
-
-    % \begin{exercise}
-    % The induction principle \texttt{oplus\_rect} is still unused in our development. 
-    % Please build some nice examples of application.
-    % \end{exercise}
-
-    \index{hydras}{Projects}
-    \begin{project}
-    Compare \texttt{oplus}'s definition (with inner fixpoint) with other possibilities
-    (\texttt{coq-equations}, \texttt{Function}, etc.).
-    \end{project}
-    \subsection{More theorems on Hessenberg's sum}
-
-    We need to prove some properties of $\oplus$, particularly about 
-    its relation with the order $<$ on \texttt{T1}.
-
-    \subsubsection{Commutativity, associativity}
-
-    We prove  the commutativity of $\oplus$ in two steps. 
-    First, we prove by transfinite induction on $\alpha$ that the restriction of $\oplus$ to the
-    interval $[0..\alpha)$ is commutative.
-
-    \index{maths}{Transfinite induction}
-
-    \input{movies/snippets/Hessenberg/oplusComm0}
-
-    Then, we infer  $\oplus$'s commutativity for any pair of ordinals:
-    Let $\alpha$ and $\beta$ be two ordinals strictly less than $\epsilon_0$. Both ordinals $\alpha$ and $\beta$ are
-    strictly less than $\textrm{max}(\alpha,\beta)+1$.
-        Thus, we have just to apply the lemma \coqsimple{oplus\_comm\_0}.
-
-\input{movies/snippets/Hessenberg/oplusComm}
-
-    Associativity of Hessenberg's sum is proved the same way.
-
-    \input{movies/snippets/Hessenberg/oplusAssoc0}
-    \input{movies/snippets/Hessenberg/oplusAssoc}
-
-    \subsubsection{Monotony}
-
-    At last, we prove that $\oplus$ is strictly monotonous in both of its arguments.
-
-    \input{movies/snippets/Hessenberg/oplusMono}
-
-      \index{hydras}{Projects}
-
-    \begin{project}
-    The library \texttt{Hessenberg} looks too long (proof scripts and compilation).
-    Please try to make it simpler and more efficient!
-    Thanks!
-    \end{project}
-
-    \paragraph*{\gaiasign} The module
-      \href{../theories/html/gaia_hydras.~GHessenberg.html}{gaia\_hydras.GHessenberg} defines a version of Hessenberg's sum
-      compatible with \gaia's type \texttt{T1}.
-
-    \subsection{A termination measure for hydra battles }
-
-    \label{sec:hydra-measure}
-
-    Let us define a measure from type \texttt{Hydra} into \texttt{T1}.
-
-    \vspace{4pt}
-    \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Termination.html\#m}{Hydra.Hydra\_Termination}}
-
-   \input{movies/snippets/Hydra_Termination/mDef}
-
-    First, we prove that the measure $m(h)$  of any hydra $h$ is a well-formed ordinal term of type \texttt{T1}.
-
-    \input{movies/snippets/Hydra_Termination/mNf}
-      \input{movies/snippets/Hydra_Termination/msNf}
-
-    For proving the termination of all hydra battles, we have to prove that
-    \texttt{m} is a variant. First, a few technical lemmas follow the decomposition of \texttt{round} into several relations. Then the lemma \texttt{round\_decr} gathers all the cases.
-
-    \label{sect:variant-decr}
-
-    \input{movies/snippets/Hydra_Termination/S0Decr}
-    \input{movies/snippets/Hydra_Termination/R1Decr}
-    \input{movies/snippets/Hydra_Termination/S1Decr}
-    \input{movies/snippets/Hydra_Termination/R2Decr}
-     \input{movies/snippets/Hydra_Termination/RoundDecr}
-
-      Finally, we prove termination of all (free) battles.
-
-    \label{thm:every-battle-terminates}
-
-    \input{movies/snippets/Hydra_Termination/FinalThm}
-
-    \paragraph*{\gaiasign}
-    The module~\href{../theories/html/gaia_hydras.GHydra.html}{gaia\_hydras.GHydra} contains a proof of \gaiaHydras version of termination of all battles.
-
-    \inputsnippets{GHydra/mDef, GHydra/mVariant, GHydra/Termination}
-
-    \section*{Conclusion}
-
-    Let us recall three results we have proved so far.
-    \begin{itemize}
-    \item There exists a strictly decreasing variant which maps \texttt{Hydra} into 
-    the segment $[0,\epsilon_0)$ for proving the termination of any hydra battle
-    \item There exists \emph{no} such variant from \texttt{Hydra} into 
-    $[0,\omega^2)$, \emph{a fortiori} into $[0,\omega)$.
-    \end{itemize}
-
-    So, a  natural question is `` Does there exist any strictly decreasing variant mapping
-    type \texttt{Hydra} into some interval $[0,\alpha)$ (where $\alpha <\epsilon_0$) for proving the termination of all hydra battles''.
-The next chapter is dedicated to a formal proof that there exists no such $\alpha$, even if we consider a restriction to the set of ``standard'' battles.
-
-    %\include{epsilon0}
-
-    %\include{impossibility-proofs}
-
-%-------------------------------------------------------------------
-
-\chapter[The Ketonen-Solovay machinery]{Accessibility inside \texorpdfstring{$\epsilon_0$}{Epsilon0}: The Ketonen-Solovay Machinery\label{ks-chapter}}
-\label{chap:ketonen}
-\index{maths}{Ordinal numbers!Ketonen-Solovay machinery}
-
-\section{Introduction}
-The reader may think that our proof of termination in the previous  chapter requires a lot of mathematical tools and may be too  complex. So, the question is ``is there  any  simpler proof'' ?
-
-In their article~\cite{KP82}, Kirby and Paris show that this result cannot be proved in Peano arithmetic.
-Their proof uses some knowledge about model theory and non-standard models of Peano arithmetic.
-In this chapter, we focus on a specific class of proofs of termination of hydra battles: construction of some variant mapping the type \texttt{Hydra} into a given initial  segment of ordinals.
-Our proof relies only on the Calculus of Inductive Constructions and is a natural complement of the results proven in the previous chapters.
-
-\begin{itemize}
-\item There is no variant mapping the type \texttt{Hydra} into the interval $[0,\omega^2)$ (section ~\vref{omega2-case}), and \emph{a fortiori} $[0,\omega)$ (section ~\vref{omega-case}).
-
-\item There exists a variant which maps the type \texttt{Hydra} into the
-interval $[0,\epsilon_0)$ (theorem \texttt{every\_battle\_terminates}, in section~\vref{thm:every-battle-terminates}).
-\end{itemize}
-
-Thus, a very natural question is the following one:
-\begin{quote}
-  `` Is there  any variant from
-\texttt{Hydra} into some interval $[0,\mu)$, where $\mu<\epsilon_0$, for proving the termination of all hydra battles ?''
-\end{quote}
-
-We prove in \coq{} the following result:
-
-\begin{quote}
-There is no variant for proving the termination of all hydra battles
-from \texttt{Hydra} into the interval $[0..\mu)$, where
-$\mu< \epsilon_0$.
-The same impossibility holds even if we consider only standard battles (with the successive replication factors $0,1,2,\dots,t,t+1,\dots$).
-\end{quote}
-
-Our proofs are  constructive and require no axioms: they are  closed terms of the CIC, and are mainly composed on function definitions and proofs of properties of these functions. 
-They  share much theoretical material with Kirby and Paris', although they do not use any knowledge about Peano arithmetic nor model  theory.  The combinatorial arguments we use and implement
-come from 
- an article by J.~Ketonen and R.~Solovay~\cite{KS81}, already  cited in the work
- by L.~Kirby and J.~Paris. % on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
- Section $2$ of this article: ''A hierarchy of probably recursive functions'', contains a systematic study of \emph{canonical sequences}, which are closely related to
-rounds of hydra battles. 
-Nevertheless, they have the same global structure as the simple proofs described in
-sections~\vref{omega-case} and \vref{omega2-case}. 
-We invite the reader to compare the three proofs step by step, lemma by lemma.
-
-\section{Canonical Sequences}
-\label{ketonen-solovay-sect}
-\index{maths}{Ordinal numbers!Canonical sequences}
-
-Canonical sequences are functions that associate an ordinal $\canonseq{\alpha}{i}$ to every ordinal $\alpha<\epsilon_0$ and positive integer $i$. They satisfy several nice properties:
-
-\index{maths}{Transfinite induction}
-\begin{itemize}
-\item If $\alpha\not=0$, then $\canonseq{\alpha}{i}<\alpha$. Thus canonical sequences can be used for proofs by transfinite induction or function definition by transfinite recursion
-\item If $\lambda$ is a limit ordinal, then $\lambda$ is the least upper bound of the set 
-$\{\canonseq{\lambda}{i}\;|\,i\in\mathbb{N}_1\}$
-
-\item If $\beta<\alpha<\epsilon_0$, then there is a ``path'' from $\alpha$ to $\beta$, \emph{i.e.} a
-sequence $\alpha_0=\alpha, \alpha_1, \dots, \alpha_n=\beta$, where for every $k<n$, there exists some $i_k$ such that $\alpha_{k+1}=\canonseq{\alpha_k}{i_k}$
-\item Canonical sequences correspond tightly to rounds of hydra battles: if $\alpha\not=0$,
-then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1})$ in one round with
-the replication factor $i$ (Lemma \href{../theories/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
-\item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) battle from $\iota(\alpha)$ to $\iota(\beta)$.
-\end{itemize}
-
-\begin{remark}
-  In~\cite{KS81}, canonical sequences are defined for any ordinal $\alpha <\epsilon_0$,
-by stating that if $\alpha$ is a successor ordinal $\beta+1$,  the sequence associated with 
-$\alpha$ is simply the constant sequence whose terms are equal to $\beta$.
-Likewise, the canonical sequence of $0$ maps any natural number to $0$.
-
-This convention allows us to make total the function that maps any ordinal $\alpha$ and natural number $i$ to the ordinal $\canonseq{\alpha}{i}$.
-
-\end{remark}
-
-First, let us recall how canonical sequences are defined in~\cite{KS81}.
-For efficiency's sake, we decided not to implement directly K.\&S's definitions, but to define in \gallina{} simply typed structurally recursive functions which share the abstract properties which are used in the mathematical proofs\footnote{With a small difference: the $0$-th term of the canonical sequence is not the same in our development as in~\cite{KS81}.}.
-
-\subsubsection{Mathematical definition of canonical sequences}
-
-In~\cite{KS81} the definition of $\canonseq{\alpha}{i}$ is based on the following remark:
-\begin{quote}
-Any non-zero ordinal $\alpha$ can be decomposed in a unique way as the product
-$\omega^\beta\times (\gamma+1)$.
-\end{quote}
-
-Thus the $\canonseq{\alpha}{i}$\,s are defined in terms of this decomposition:
-\begin{definition}[Canonical sequences: mathematical definition]
-\label{def:canonseq-math}
-
-\end{definition}
-\begin{mathframe}
-  \begin{itemize}
-\item Let $\lambda<\epsilon_0$ be a limit ordinal
-
-\begin{itemize}
-\item If $\lambda=\omega^{\alpha+1}\times (\beta+1)$, then 
-$\canonseq{\lambda}{i}= \omega^{\alpha+1}\times\beta +  \omega^\alpha \times i$
-\item If $\lambda=\omega^{\gamma}\times (\beta+1)$, where $\gamma<\lambda$ is a limit ordinal, then 
-$\canonseq{\lambda}{i}=\omega^{\gamma}\times \beta + \omega^{\canonseq{\gamma}{i}}$
-\end{itemize}
-
-\item For successor ordinals, we have $\canonseq{\alpha+1}{i}= \alpha$
-
-\item Finally, $\canonseq{0}{i}= 0$.
-\end{itemize}
-\end{mathframe}
-
-\subsubsection{Canonical sequences in \coq}
-\index{hydras}{Library Epsilon0!Functions!canon}
-\index{hydras}{Library Epsilon0!Functions!canonS}
-
-Our definition may look more complex than the mathematical one, but
-uses plain structural recursion over the type \coqsimple{T1}. Thus, tactics like
-\coqsimple{cbn}, \coqsimple{simpl}, \coqsimple{compute}, etc., are applicable.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canon}{Epsilon0.Canon}}
-
-\label{Functions:canonS}
-\label{Functions:canon}
-
-\input{movies/snippets/Canon/canonDef}
-
-\paragraph*{\gaiasign} Our library~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon} transposes the  function \texttt{canon} into \gaia's type \texttt{T1}
-(please see Sect.\~ref{sect:gcanon}).
-
-\paragraph*{Remark}
-In the present state of this library, the following specializations of \texttt{canon} are still used in some proofs or lemma statements. They are planned to be deprecated.
-
-\input{movies/snippets/Canon/CanonS0}
-
-For instance \coq's computing facilities allow us to verify the equalities\linebreak 
-\mathcolor{$\canonseq{\omega^\omega}{3} = \omega^3$} and
-\mathcolor{$\canonseq{\omega^\omega*3}{42} = \omega^\omega*2 + \omega^{42}$}.
-
-\input{movies/snippets/Canon/canonExamples}
-
-% \index{hydras}{Projects}
-% \begin{project}
-% Many lemmas presented in this chapter were stated and proved before the introduction of 
-% the type class \texttt{ON} of ordinal notations, and in particular its  instance \texttt{Epsilon0}.
-% Thus definitions and lemmas refer to the type \texttt{T1} of possibly not well-formed terms.
-% This should be fixed in  a future version.
-% \end{project}
-
-\subsection{Basic properties of canonical sequences}
-
-We did not  try to prove that our definition truly implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
-abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
-~\href{../theories/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}.
-For instance, the equality $\canonseq{\alpha+1}{i}=\alpha$  can be  proved by  structural induction on $\alpha$.
-
-\input{movies/snippets/Canon/canonSucc}
-
-\subsubsection{Canonical sequences and the order $<$}
-
-\index{maths}{Transfinite induction}
-
-We prove by transfinite induction over $\alpha$ that $\canonseq{\alpha}{i}$ is an ordinal strictly less than $\alpha$ (assuming $\alpha\not=0$).
-This property allows us to use the function \texttt{canonS} and its derivatives in function definitions by transfinite recursion.
-
-\label{lemma:canon_LT}
-
-\inputsnippets{Canon/canonLT}
-
-\paragraph*{\gaiasign} This lemma is also available in
-Library ~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon}:
-
-\inputsnippets{GCanon/gcanonLt}
-
-\subsubsection{Limit ordinals are truly limits}
-The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
-is the limit of the sequence $\canonseq{\lambda}{i}\;(1\le i)$.
-
-\vspace{4pt}
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}
-
-\input{movies/snippets/Canon/canonSLimitStrong}
-
-\label{lemma:canonS-limit}
-
-Note the use of \coq's \texttt{sig} type in the theorem's statement, which
-relates the boolean function \texttt{limitb} defined on the \texttt{T1} data-type with a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.
-We can also state directly that $\lambda$ is a (strict) least upper bound of the elements of its canonical sequence.
-
-\input{movies/snippets/Canon/canonSLimitLub}
-
-\paragraph*{\gaiasign}  In \gaiaHydras, the statement use a slightly different vocabulary:
-\inputsnippets{GCanon/gcanonLimitOf}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}\label{exo:simply-typed-canonseq}
-Instead of using the \texttt{sig} type, define a simply typed function that, given two ordinals $\alpha$ and $\beta$, returns a natural number $i$ such that, if $\alpha$ is a limit ordinal and $\beta<\alpha$, then $\beta< \canonseq{\alpha}{i+1}$.
-Of course, you will have to prove the correctness of your function.
-
-\textbf{Hint:} You may add to your function a third argument usually called \texttt{fuel} for allowing you to give a structurally 
-recursive function (\emph{cf} the post of Guillaume Melquiond on Coq-club (Dec 21, 2020)
-\url{https://sympa.inria.fr/sympa/arc/coq-club/2020-12/msg00069.html}).
-The type \texttt{fuel}, an alternative 
-to \texttt{nat} is available on \href{../theories/html/hydras.Prelude.Fuel.html}{Prelude.Fuel})
-.
-%\index{coq}{Giving fuel to a long computation}
-
-\end{exercise}
-
-\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0}\,: paths}
-\index{maths}{Ordinal numbers!Accessibility inside epsilon0}
-\label{sect:pathes-intro}
-
-Let us consider a kind of accessibility problem inside $\epsilon_0$: given two ordinals $\alpha$ and $\beta$, where $\beta<\alpha<\epsilon_0$, find a \emph{path} consisting of a finite sequence $\gamma_0=\alpha,\dots,\gamma_l=\beta$,
-where, for every $i<l$, $\gamma_i \not= 0$ \footnote{This condition allows us to ignore paths which end by a lot of useless $0$s.} and there exists some strictly positive integer $s_i$
-such that $\gamma_{i+1}=\canonseq{\gamma}{s_i}$.
-
-Let $s$ be the sequence $\langle s_0,s_1,\dots, s_{l-1} \rangle$. We describe the
-existence of such a path with the notation $\alpha\xrightarrow [s]{}\beta$.
-We say also that the considered path from $\alpha$ to $\beta$ \emph{starts at [index] $s_0$ and ends at $s_l$}.
-
-For instance, we have $\omega*2 \xrightarrow[2,2,2,4,5]{}3$, through the 
-path $\langle\omega\times 2, \omega+2,\omega+1,\omega,4,3\rangle$.
-
-\begin{remark}
-
-Note that, given $\alpha$ and $\beta$, where $\beta < \alpha$, the sequence $s$ which leads from $\alpha$ to $\beta$ is not unique.
-Indeed, if $\alpha$ is a limit ordinal, the first element of $s$ can be any integer $i$ such that $\beta<\canonseq{\alpha}{i}$, and if $\alpha$ is a successor ordinal,
-then the sequence $s$ can start with any positive integer.
-
-For instance, we have also 
-$\omega*2 \xrightarrow[3,4,5,6]{}\omega$. 
-Likewise,
-$\omega*2 \xrightarrow[1,2,1,4]{} 0$ and
-$\omega*2 \xrightarrow[3,3,3,3,3,3,3,3]{} 0$.
-\end{remark}
-
-\subsection{Formal definition}
-
-\label{path-to-definition}
-
-In \coq{}, the notion of path can be simply defined as an inductive predicate 
-parameterized by the destination $\beta$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\index{hydras}{Library Epsilon0!Predicates!path\_to}
-\label{sect:path-to-def}
-
-\input{movies/snippets/Paths/transitionDefs}
-
-\input{movies/snippets/Paths/pathDef}
-
-\begin{remark}
-In the present version of our library, we use a variant \texttt{path\_toS} of
-\texttt{path\_to}, where the proposition
-(\texttt{path\_toS $\beta$ $s$ $\alpha$}) is equivalent to
-(\texttt{path\_to $\beta$ (List.map S $s$) $\alpha$}). This variant is scheduled to be deprecated.
-\end{remark}
-
-\paragraph*{\gaiasign}
-The library~\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} transposes the  notion of path into \gaia's type \texttt{T1}
-(please see Sect.\~ref{sect:gpath}).
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Write a tactic for solving goals of the form (\texttt{path\_to $\beta$ $s$ $\alpha$})
-where $\alpha$, $\beta$ and $s$ are closed terms. 
-You should solve automatically the following goals:
-
-\inputsnippets{Paths/exPathTac} 
-\end{exercise}
-
-\subsection{Existence of a path}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction on $\alpha$, we prove that for any $\beta<\alpha$, 
-one can build a path from $\alpha$ to $\beta$ (in other terms, $\beta$ is accessible from $\alpha$).
-
-\input{movies/snippets/Paths/LTPathTo}
-
-\noindent 
-From the lemma \texttt{canon\_LT}~\vref{lemma:canon_LT}, we can convert any path into an inequality on ordinals (by induction on paths).
-
-\input{movies/snippets/Paths/pathToLT}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}[continuation of exercise~\vref{exo:simply-typed-canonseq}]
-Define a simply typed function for computing a path from $\alpha$ to $\beta$.
-\end{exercise}
-
-\subsection{Paths and hydra battles}
-\label{KS-o2h}
-
-In order to apply our knowledge about  ordinal numbers less than $\epsilon_0$ to the study of hydra battles, we define an injection
-from the interval $[0,\epsilon_0)$ into the type \texttt{Hydra}.
-
-\vspace{4pt}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}}
-
-\input{movies/snippets/O2H/iotaDef}
-
-For instance Fig.~\ref{fig:iota-example} shows the image by $\iota$ of the ordinal  \textcolor{black}{$\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$}
-
-  \begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.3]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (10,2) {$\bullet$};
-\node (N22) at (7,2) {$\bullet$};
-\node (N3) at (14,2) {$\bullet$};
-\node (N4) at (18,2) {$\Smiley[2][green]$};
-\node (N5) at (0,4) {$\bullet$};
-\node (N6) at (2,5) {$\Smiley[2][green]$};
-\node (N7) at (4,6) {$\Smiley[2][green]$};
-\node (N88) at (7,4) {$\bullet$};
-\node (N8) at (10,4) {$\bullet$};
-\node (N9) at (14,6) {$\Smiley[2][green]$};
-\node (N10) at (0,8) {$\Smiley[2][green]$};
-\node (N11) at (10,7) {$\Smiley[2][green]$};
-\node (N111) at (7,7) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N22);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (N1) to  (N5);
-\draw (N1) to   [bend left=10] (N6);
-\draw (N1) to   [bend right=20] (N7);
-\draw (N2) to  (N8);
-\draw (N22) to  (N88);
-\draw (N8) to  (N11);
-\draw (N88) to  (N111);
-\draw (N3) to  (N9);
-\draw (N5) to  (N10);
-\end{tikzpicture}
-\caption{The hydra $\iota(\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$) \label{fig:iota-example}}
-
-\end{figure}
-
-The following lemma (proved in ~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H.v}) maps  canonical sequences to rounds of hydra battles.
-
-\label{lemma:canonS-iota}
-
-\input{movies/snippets/O2H/canonSIota}
-
-The next step of our development extends this relationship to
-the order $<$ on $[0,\epsilon_0)$ on one side, and hydra battles on the other side.
-
-\input{movies/snippets/O2H/pathToRoundPlus}
-
-As a corollary, we are now able to transform any inequality $\beta<\alpha<\epsilon_0$ into a (free) battle.
-
-\input{movies/snippets/O2H/LTToRoundPlus}
-
-\section{A  proof of impossibility}
-
-We now have  the tools for proving that  there exists no variant bounded by some $\mu<\epsilon_0$ for proving the termination   of all battles.
-The proof we are going to show is a proof by contradiction.
-It  can
- be considered as a generalization of the
-proofs described in  sections~\vref{omega-case} and \vref{omega2-case}.
-
-In the module
-\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
-This part of the development is parameterized by some class $B$ of battles, which will be instantiated later to \texttt{free} or \texttt{standard}.
-
-\vspace{4pt}
-\noindent
-From~\href{../theories/html/hydras.Hydra.~Hydra_Definitions.html}{Hydra.Hydra\_Definitions}:
-
-\input{movies/snippets/Hydra_Definitions/BoundedVariant}
-
-Let us assume there exists such a variant:
-
-\input{movies/snippets/Epsilon0_Needed_Generic/theContext}
-
-\label{remark:m-decrease}
-\begin{remark}
-  The hypothesis \texttt{m\_decrease} is not provable  in general, but is satisfied by
-the  \texttt{free} and \texttt{standard} kinds of battles. This trick allows to 
-``factorize'' our proofs  of impossibility.
-\end{remark}
-
-\index{maths}{Transfinite induction}
-
-First, we prove that $m(\iota(\alpha))$ is always greater than or equal to $\alpha$, by  transfinite induction over $\alpha$.
-
-\input{movies/snippets/Epsilon0_Needed_Generic/mGe0}
-
-\begin{itemize}
-\item If $\alpha=0$, the inequality trivially holds
-\item If $\alpha$ is the successor of  some ordinal $\beta$, the inequality $\beta \leq m(\iota(\beta))$ holds (by induction hypothesis).
-But the hydra $\iota(\alpha)$ is transformed in one round into
-$\iota(\beta)$, thus $m(\iota(\beta))<m(\iota(\alpha))$. Hence $\beta<m(\iota(\alpha))$, which implies $\alpha \leq m(\iota(\alpha))$
-\item If $\alpha$ is a limit ordinal, then $\alpha$ is the least upper bound of the set
-of all  the $\canonseq{\alpha}{i}$.  Thus, we have just to prove that $\canonseq{\alpha}{i}< m(\iota(\alpha))$ for any $i$. 
-\begin{itemize}
-\item Let $i$ be some natural number.
-By the induction hypothesis, we have $\canonseq{\alpha}{i} \leq m(\iota(\canonseq{\alpha}{i}))$.
-But the hydra $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i})$ in one round, thus $m(\iota(\canonseq{\alpha}{i})) < m(\iota(\alpha))$, by our hypothesis \texttt{m\_decrease}.
-\end{itemize}
-\end{itemize}
-
-Please note that the impossibility proofs of 
-sections~\vref{omega-case} and \vref{omega2-case} contain a similar lemma, also called \texttt{m\_ge}.
-We are now able to build a counter-example.
-
-\input{movies/snippets/Epsilon0_Needed_Generic/mGeGeneric}
-
-The (big) rest of the proof is dedicated to prove formally the converse inequality 
-\texttt{m small\_h t1< m big\_h}.
-
-\subsection{The case of free battles}
-\label{sec:free-battles-case}
-Let us now consider that $B$ is instantiated to \texttt{free} (which means that we are considering proofs of termination of \emph{all} battles).
-The following lemmas are proved in Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
-The case $B=\texttt{standard}$ is studied in section~\vref{std-case}.
-
-\input{movies/snippets/Epsilon0_Needed_Free/theContext}
-
-\begin{enumerate}
-\item The following lemma is an application of \texttt{m\_ge\_generic}, since \texttt{free}
-satisfies trivially the hypothesis \texttt{m\_decrease} (see page~\pageref{remark:m-decrease}).
-
-\inputsnippets{Epsilon0_Needed_Free/mGe}
-
-\item From the hypothesis \texttt{Hy}, we have \texttt{m big\_h t1< mu}
-\item By Lemma \texttt{LT\_to\_round\_plus}, we get a (free) battle from
-\texttt{big\_h = iota mu} to \texttt{small\_h = iota (m big\_h)}.
-
-\input{movies/snippets/Epsilon0_Needed_Free/bigToSmall}
-
-\item From the hypotheses on $m$, we infer:
-\input{movies/snippets/Epsilon0_Needed_Free/mLt}
-
-\item From lemmas \texttt{m\_ge} and \texttt{m\_lt}, and the irreflexivity of $<$, we get a contradiction. 
-\input{movies/snippets/Epsilon0_Needed_Free/ImpossibilityFree}
-
-\end{enumerate}
-
-We have now proved there exists no bounded variant for the class of free battles.
-
-\input{movies/snippets/Epsilon0_Needed_Free/CheckDemo}
-
-\section{The case of standard battles}
-\label{sec:standard-intro}\label{std-case}
-One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
-Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}.
-
-\input{movies/snippets/O2H/LTToRoundPlus}
-
-This lemma builds a battle out of any inequality $\beta<\alpha$. 
-It is a straightforward application of \texttt{LT\_path\_to} of
-Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}:
-
-\input{movies/snippets/Paths/LTPathTo}
-
-The sequence $s$, used to build the sequence of replication factors of the battle, depends on 
-$\alpha$ and $\beta$, so we cannot be sure that the generated battle is a genuine standard battle.
-
-The solution to this issue comes  once again from Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
-$\alpha_0=\alpha,\alpha_1,\dots,\alpha_k=\beta$ where $\alpha_{j+1}$ is equal
-to $\canonseq{\alpha_j}{i_j}$ where $i_j$ is \emph{any} natural number, 
-we consider various constraints on these sequences.
-In particular, a path is called \emph{standard} if $i_{j+1} = i_j + 1$ for every $j<k$.
-It  corresponds to a ``segment'' of some standard battles. 
-Please note that the vocabulary on paths is ours, but all the concepts come really from~\cite{KS81}.
-
-In \coq{}, standard paths can be defined as follows.
-
-\vspace{4pt}
-
-\emph{From
-Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/standardPathR}
-
-In the mathematical text and figures, we shall use the notation 
-$\alpha \xrightarrow[i,j]{}\beta$ for the proposition 
-(\texttt{standard\_path $i$ $\alpha$ $j$ $\beta$}).
-In~\cite{KS81} the notation is
-$\alpha \xrightarrow[i]{*}\beta$
-for 
-the proposition  $\exists j, i<j \wedge \alpha \xrightarrow[i,j]{} \beta$.
-
-Our goal is now  to transform any inequality $\beta<\alpha<\epsilon_0$ into a standard path $\alpha \xrightarrow[i,j]{} \beta$ for some $i$ and $j$, then into a standard battle
-from $\iota(\alpha+i)$ to $\iota(\beta)$. 
-Following~\cite{KS81}, we proceed in two stages:
-\begin{enumerate}
-\item we simulate plain (free) paths from $\alpha$ to $\beta$ with
-paths made of steps $(\gamma,\canonseq{\gamma}{n})$, \emph{with the same $n$ all along the path}
-\item we simulate any such path by a standard path.
-\end{enumerate}
-
-\subsection{Paths with a constant index}
-
-First of all, paths with a constant index 
-enjoy nice properties. They are defined as paths where all the $i_j$ are equal to the same natural number $i$, for some $i>0$.
-
-Like in~\cite{KS81}, we shall use the notation $\alpha \xrightarrow[i]{} \beta$ for denoting such a path, also called an $i$-path.
-
-\input{movies/snippets/Paths/constPathDef}
-
-% Paths with a given index can be effectively computed.
-% Given $i$, $\alpha$ and $l$, the following function returns the ordinal $\beta$ such that there exists a path 
-% $\alpha \xrightarrow [i+1] {} \beta$ of length $l$.
-
-% \begin{Coqsrc}
-% Fixpoint const_funS (i:nat)(alpha : T1)(l:nat):  T1  :=
-%   match l
-%   with
-%   | 0 => alpha
-%   | S m => const_funS i (canonS i alpha) m
-%   end.
-% \end{Coqsrc}
-
-% The following computations show  applications of \texttt{constS\_fun} to the 
-% ordinal $\omega^\omega$, with various values of $i$ and $l$.
-
-% \begin{Coqsrc}
-% Compute  (const_funS 2 (omega ^omega)  55).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = zero
-%      : T1 
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (const_funS 2 (omega ^omega) 15).
-% \end{Coqsrc}
-
-%   \begin{Coqanswer}
-%  = (omega ^ 2 * 2)%pT1
-%      : ppT1   
-%   \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (const_funS 4 (omega^omega)  100).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% = (omega ^ 4 * 4 + omega ^ 3 * 4 + omega ^ 2 + omega * 4 + 4)%pT1
-%      : ppT1
-% \end{Coqanswer}
-
-A most interesting property of $i$-paths is that we can ``upgrade'' their index, as stated by K.\&S.'s Corollary 12.
-
-\index{maths}{Transfinite induction}
-
-\vspace{4pt}
-\noindent
-\emph{From
-  Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/Cor12}
-
-We  also use a version of \texttt{Cor12} with large inequalities.
-
-\input{movies/snippets/Paths/Cor121}
-
-\subsubsection{Sketch of proof of \texttt{Cor12}}
-\index{maths}{Transfinite induction}
-
-First, let us prove the following lemma (numbered \texttt{2.4} in Ketonen and Solovay's article).
-
-\input{movies/snippets/Paths/KSThm24}
-
-We prove this lemma by transfinite induction on $\alpha$.
-Let us consider a path $\alpha \xrightarrow [i]{} \beta$ $(i>0)$. Its first step is
-the pair $(\alpha,\canonseq{\alpha}{i})$, We have $\canonseq{\alpha}{i}<\alpha$ and
-$\canonseq{\alpha}{i} \xrightarrow [i]{} \beta$. 
-Let $n$ be any natural number such that $n>i$.
-By the induction hypothesis, there exists a path $\canonseq{\alpha}{n} \xrightarrow[i]{} \beta$.
-\begin{itemize}
-\item  If $\alpha$ is a successor ordinal $\gamma+1$, then $\canonseq{\alpha}{n} =
-\canonseq{\alpha}{i}=\gamma$. Thus we have a path 
-$\alpha  \xrightarrow [n]{}  \gamma \xrightarrow [n]{} \beta$
-\item If $\alpha$ is a limit ordinal, we apply Theorem \texttt{KS\_thm\_2\_4} (see above).
-
-%   \begin{theorem}
-% Let $\lambda$ be a limit ordinal, then for any pair of indices $0<i<j$, there is a path $\canonseq{\lambda}{j} \xrightarrow[1]{} \canonseq{\lambda}{i}$.    
-%   \end{theorem}
-
-The proof of the limit case, is decomposed into a sequence
-of path constructions leading to $\alpha \xrightarrow[n]{} \beta$.
-
- \begin{enumerate}
- \item $\alpha \xrightarrow[n]{} \canonseq{\alpha}{n}$ (single step path)
- \item $\canonseq{\alpha}{n} \xrightarrow[1]{} \canonseq{\alpha}{i}$ (by \texttt{Theorem\_2\_4}),
-\item $\canonseq{\alpha}{n} \xrightarrow[n]{} \canonseq{\alpha}{i}$ (applying the induction hypothesis to the preceding path);
-\item $\canonseq{\alpha}{i} \xrightarrow[n]{} \beta$ (applying the induction hypothesis)
-\item $\alpha \xrightarrow[n]{} \beta$ (by composition of 1, 3, and 4).
-
- \end{enumerate}
-
-\end{itemize}
-
-\begin{remark}
- \texttt{Cor12} ``casts'' $i$-paths into $n$-paths for any $n>i$.
-But the obtained $n$-path can be much longer than the original $i$-path.
-The following exercise will give an idea of this increase. 
-\end{remark}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-  Prove that  the length of the $i+1$-path from
-  $\omega^\omega$ to $\omega^i$ is $1 + (i+1)^{(i+1)}$, for any $i$. Note that the $i$-path from
-  $\omega^\omega$ to $\omega^i$ is only one step long.
- \end{exercise}
-
-Why is \texttt{Cor12} so useful? 
-Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
-we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying corollary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
-where $n$ is the maximum of the indices $i$ met in the induction.
-
- Lemma 1, Section 2.6 of~\cite{KS81} is naturally expressed in terms of \coq's
-\verb@sig@ construct.
-
-\label{lemma:L-2_6-1}
-%\index{coq}{Sigma types}
-
-\input{movies/snippets/Paths/Lemma261}
-
-Intuitively, Lemma~\texttt{2\_6\_1}  shows that if $\beta<\alpha<\epsilon_0$, then there exists  a battle from $\iota(\alpha)$ to $\iota(\beta)$ where the replication factor is constant, although large enough.
-
- \paragraph*{\gaiasign}
- Corollary \texttt{Cor12} and Lemma \texttt{Lemma2\_6\_1} are also available in
-\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} 
-(please see Sect.~\vref{sect:gpath}).
-
-\subsection{Casting paths with a constant index into a standard path}
-
-The article~\cite{KS81} contains 
-the following lemma, which allows us to simulate $i$-paths by $[i+1,j]$-paths, where $j$ is large enough.
-
-\input{movies/snippets/Paths/constantToStandard}
-
-\subsubsection{Sketch of proof of \texttt{constant\_to\_standard\_path}}
-
-Our proof follows the proof by Ketonen and Solovay, including its organization as a sequence of lemma.
-Since it is a non-trivial proof, we will comment its main steps below   (see Figure~\vref{fig:belle-preuve-1} to Figure~\vref{fig:fin-belle-preuve}).
-
-\subsubsection*{Preliminaries}
-
-Please note that, given an ordinal $\alpha:\texttt{T1}$, and two natural numbers $i$ and $l$, there exists at most a standard path $\alpha \xrightarrow [i,i+l]{*} \beta$.
-The following function computes $\beta$ from $\alpha$, $i$ and $l$.
-
-\input{movies/snippets/Paths/standardGnaw}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction over  $\alpha$, we prove that the ordinal $0$ is reachable from any ordinal $\alpha<\epsilon_0$ by some standard path.
-
-\input{movies/snippets/Paths/standardPathToZero}
-
-\paragraph*{}
-Now, let us consider two ordinals  $\beta<\alpha<\epsilon_0$.  Let $p$  be some $(n+1)$-path from $\alpha$ to $\beta$.
-
-\input{movies/snippets/Paths/ConstantToStandardProof}
-
-Applying \texttt{standard\_path\_to\_zero}, $0$ is reachable from $\alpha$ by some standard path.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \draw[->, very thick,blue] (alpha)-- node [below]{$n+1$} node [above] {$+$} (beta);
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-    \node (alpha3) at (20,5) {};
-  \node (alphalast) at (35,5) {};
-  \node (zero) at (45,0) {$0$};
-  \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate=40]{$n+1$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha2)-- node [below]{$n+3$}  (alpha3);
-
-  \node (dots) at (24,5) {$\dots$};
-  \draw [->, dashed, very thick,blue] (alphalast)-- node [below, rotate=-26]{$n+p+1$}  (zero);
-
-\end{tikzpicture}
-\caption{A nice proof (1)}
-  \label{fig:belle-preuve-1}
-\end{figure}
-
-\paragraph*{}
-
-Since comparison on \texttt{T1} is decidable, one can compute the last step $\gamma$ of the standard path from $(\alpha,n+1)$  such that $\beta\leq \gamma$.
-Let $l$ be the length of the path from $\alpha$ to $\gamma$.  
-This step of the proof is illustrated in figure~\vref{fig:belle-preuve-2}.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (17,5) {$\ldots$};
-    \node (alpha3) at (20,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (38,0) {$\delta$};
-    \draw [->, dashed,very thick,blue] (alpha)-- node [below,rotate=35]{$n+1$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below,rotate = -48]{\tiny $n+l$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{$n+l+1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{$n+1$} node [above] {$+$} (beta);
-   \draw[thick] (alpha)--  (gamma);
-   \draw[thick] (gamma)--  node [above] {$\geq$} (beta);
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-\end{tikzpicture}
-
-\caption{A nice proof (2)}
-  \label{fig:belle-preuve-2}
-\end{figure}
-
-\paragraph*{}
-
-\begin{itemize}
-\item If $\beta=\gamma$, it's OK! We have got a standard path
-from  
-$\alpha$ to $\beta$ with successive indices  $n+1, n+2, \dots, n+l+1$
-
-\item Otherwise,  $\beta < \gamma$.  Let us consider  $\delta=\canonseq{\gamma}{n+l+1}$.
-By applying several times lemma \texttt{Cor12},  one converts  every path of Fig~\ref{fig:belle-preuve-2} into
- a $n+l+1$-path  (see figure~\ref{fig:belle-preuve-3}).
-
-But $\gamma$ is on the $n+l+1$-path from $\alpha$ to $\beta$.
-As shown by figure~\vref{fig:fin-belle-preuve}, the ordinal $\delta$, reachable from
-$\gamma$ in one single step,  must be greater than or equal to $\beta$, which contradicts our  hypothesis $\beta < \gamma$.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-    \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (17,5) {$\ldots$};
-    \node (alpha3) at (20,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (38,0) {$\delta$};
-     \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -48]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\tiny $n+l+1$} node [above]{\color{red} $1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
-    \draw[thick] (gamma)--   node [above]{\color{red} $>$}(beta);
-   \draw[thick] (alpha)--  (gamma);
-
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-
-\end{tikzpicture}
-
-\caption{A nice proof (3)}
-  \label{fig:belle-preuve-3}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (15,5) {$\ldots$};
-    \node (alpha3) at (18,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (42,0) {$\delta$};
-    \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -36]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\small $n+l+1$} node [above]{\color{red} $1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
-   \draw[thick] (alpha)--  (gamma);
-
-    \draw[thick] (gamma)--  node [below]{\tiny $n+l+1$} node [above]{\color{red} $+$}(beta);
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-
-\end{tikzpicture}
-
-\caption{A nice proof (4)}
-  \label{fig:fin-belle-preuve}
-\end{figure}
-
-\end{itemize}
- The only possible case is  thus $\beta=\gamma$, so we have got a standard path  from $\alpha$ to $\beta$.
-
- \input{movies/snippets/Paths/constantToStandard0}
- \input{movies/snippets/Paths/constantToStandardz}
-
-Here is the full statement of the conversion from constant to standard paths.
-
-\input{movies/snippets/Paths/constantToStandardPath}
-
-Applying \texttt{Lemma2\_6\_1} and \texttt{constant\_to\_standard\_path}, we get the following corollary.
-
-\input{movies/snippets/Paths/LTToStandardPath}
-
-\subsection{Back to hydras}
-\label{sec:standard-battles-cases}
-We are now able to complete our proof that there exists no bounded variant for proving the termination of standard hydra battles. This proof can
-be consulted in the module 
-\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}{Hydra.Epsilon0\_Needed\_Std}.
-Please note that it has the same global structure as in section~\ref{sec:free-battles-case} 
-% ICI !
-Applying the  lemmas  \texttt{Lemma2\_6\_1} of the module 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
-{Epsilon0.pathS}   and 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
-{\texttt{constant\_to\_standard\_path}},
-we can convert any inequality $\beta<\alpha<\epsilon_0$ into a standard path from
-$\alpha$ to  $\beta$, then into a fragment of a standard battle from 
-$\iota(\alpha)$ to $\iota(\beta)$, hence the inequality $m(\iota(\beta))<m(\iota(\alpha))$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html\#LT_to_standard_battle}{Hydra.Epsilon0\_Needed\_Std}}
-
-\input{movies/snippets/Epsilon0_Needed_Std/LTToStandardBattle}
-
-Now, please consider the following context:
-
-\input{movies/snippets/Epsilon0_Needed_Std/theContext}
-
-In the same way as for free battles, we import a large inequality 
-from 
-the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Epsilon0\_Needed\_Generic}.
-
-\input{movies/snippets/Epsilon0_Needed_Std/mGe}
-
-\paragraph*{} If remains to prove the following strict inequality, in order to have a contradiction.
-
-\input{movies/snippets/Epsilon0_Needed_Std/mLt}
-
-\paragraph*{Sketch of proof:} Let us recall that $\texttt{big\_h} = \iota(\mu)$
- and $\texttt{small\_h} = \iota (m (\texttt{big\_h}))$.
-
-Since $m(\texttt{big\_h})< \mu$, there exists a standard path from $\mu$ to
-$m(\texttt{big\_h})$, hence a   standard battle from $\iota(\mu)$  to
-$\iota(m(\texttt{big\_h}))$,  i.e. from \texttt{big\_h} to \texttt{small\_h}.
-
-Since $m$ is assumed to be a variant for standard battles, we get the inequality  $m(\texttt{small\_h}) < m(\texttt{big\_h})$.
-
-\input{movies/snippets/Epsilon0_Needed_Std/endOfProof}
-
-\subsection{Remarks}
-
-We are grateful to 
- J. Ketonen and R. Solovay  for the high quality of their explanations and proof details.
-Our proof follows tightly the sequence of lemmas in their article, with a focus on 
-constructive aspects.
-Roughly speaking, our implementation \emph{builds}, out of a hypothetical
-  variant $m$, bounded by some ordinal $\mu<\epsilon_0$, a hydra \texttt{big\_h} which verifies the impossible inequality  $m(\texttt{big\_h})< m(\texttt{big\_h})$.
-
-On may ask whether the preceding results are too restrictive, since they 
-refer to a particular data type \texttt{T1}.
-In fact, our representation of ordinals strictly less than 
- $\epsilon_0$ is faithful to their mathematical definition, at least 
-Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
-(please see also the module
-\href{../theories/html/hydras.Schutte.Correctness_E0.html}{hydras.Schutte.Correctness\_E0}).
-
-Thus, we can infer that our theorems can be applied to any well order.
-
-\index{hydras}{Projects}
-\begin{project}
-Study a possible modification of the definition of a variant  (for  standard battles).
-
-\begin{itemize}
-\item The variant is assumed to be strictly decreasing \emph{on configurations 
-reachable from some initial configuration where the replication factor is equal to $0$}
-\item The variant may depend on the number of the current round.
-\end{itemize}
-
-In other words, its type should be \texttt{nat -> Hydra -> T1}, and it must 
-verify the inequality $m\, (S\,i)\, h' < m\,i\, h$ whenever the configuration 
-$(i,h)$ is reachable from some initial configuration $(0,h_0)$
-and \texttt{h} is transformed into \texttt{h'} in the considered round.
-Can we still prove the theorems of section~\ref{std-case} with this new definition?
-
-\end{project}
-
-%---------------------------------------------------------------------
-\chapter{Large sets and rapidly growing functions}\label{chap:alpha-large}
-
-\begin{remark}
-Some notations (mainly names of fast-growing functions) of our development may differ slightly from the literature.
-Although this fact does not affect our proofs, we are preparing a future version where the names $F\_alpha$, $f\_alpha$, $H\_alpha$, etc., are fully consistent with the cited articles.
-
-\end{remark}
-%\section{Introduction}
-
-In this chapter, we try to feel how long a standard battle can be.
-To be precise, for any ordinal $\alpha<\epsilon_0$ and any positive integer $k$,
-we give a minoration of the number of steps of a standard battle which
-starts with the hydra $\iota(\alpha)$ and the replication factor $k$.
-
-We express this number in terms of the Hardy hierarchy of fast-growing 
-functions~\cite{BW85, Wainer1970, KS81, Promel2013}.
- From the \coq{} user's point of view, such  functions are  very 
-attractive:  they are defined as functions  in \gallina{}, and we can apply them \emph{in theory}, but they are so complex that you will never be able to look at the result of the computation.
- Thus, our knowledge on these functions must rely on \emph{proofs}, not tests. In our development, we use often the rewriting rules generated by \coq's \texttt{Equations} plug-in.
-
-\section{Definitions}
-
-%\subsection{Definition}
-
-\begin{definition}
-Let $0<\alpha<\epsilon_0$ be any ordinal, and $s=\langle s_1, s_2, \dots, s_N\rangle$ a finite sequence of strictly positive natural numbers.
-
-We say that $s$ is \emph{$\alpha$-large} if the sequence $\langle \alpha_0=\alpha,\dots,\alpha_{i+1}=\canonseq{\alpha_i}{i+1},\dots \rangle$ leads to $0$. 
-We say also that $s$ is \emph{minimally $\alpha$-large} (in short:
-\emph{$\alpha$-mlarge}) if $s$ is $\alpha$-large 
- and every strict prefix of $s$ leads to a non-zero ordinal (\emph{cf} Sect.~\vref{sect:path-to-def}).
-
-\index{maths}{Ordinal numbers!Large sets}
-\index{maths}{Ordinal numbers!Minimal large sets}
-
-\end{definition}
-
-\begin{remark}
-  Ketonen and Solovay~\cite{KS81} consider  large finite \emph{sets} of natural numbers,  but they are mainly used as sequences. Thus, we chose to represent them explicitely as (sorted) lists. 
-\end{remark}
-
-The following function ``gnaws'' an ordinal $\alpha$, following a sequence of indices (ignoring the $0$s).
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/gnawDef}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_SetsPaths.html\#gnaw}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/largeDef}
-
-Minimal large sequences can be directly defined in terms of the
-predicate \texttt{path\_to} (\vref{sect:path-to-def}) which already prohibits paths containing non-final \texttt{zero}s.
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}
-
-\index{hydras}{Library Epsilon0!Predicates!mlarge@mlarge (minimal large sequences)}
-
-\input{movies/snippets/Large_Sets/mlargeDef}
-
-Let us consider two integers $k$ and $l$, such that $0<k<l$. In order to check whether the interval $[k,l]$ is minimally large for $\alpha$, it is enough to
-follow from $\alpha$ the path associated with the interval $[k,l)$ and verify that the last ordinal we obtain is equal to $1$.
-
-\subsection{Examples}
-
-For instance the interval $[6,70]$ leads $\omega^2$ to $\omega\times 2 + 56$. Thus this interval is not $\omega^2$-large.
-
-\noindent
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html\#mlarge}{Epsilon0.Large\_Sets\_Examples}}
-
-\input{movies/snippets/Large_Sets_Examples/gnawEx1}
-
-The interval $[6,700]$ is $\omega^2$-large, but not
-$\omega^2$-mlarge, since $[6,699]$ is also $\omega^2$-large.
-
-\input{movies/snippets/Large_Sets_Examples/gnawEx2}
-
-The following lemma relates minimal largeness with the
-function 
-\texttt{gnaw}.
-
-\input{movies/snippets/Large_Sets/mlargeIff}
-
-\vspace{4pt}
- \noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html}{Epsilon0.Large\_Sets\_Examples}}
-
-\input{movies/snippets/Large_Sets_Examples/Ex1Lemma}
-
-\section{Length of minimal large sequences}
-
-Now, consider any natural number $k>0$ any ordinal $0<\alpha<\epsilon_0$.  We would like to compute
-a number $l$ such that the interval $[k,l]$ is $\alpha$-mlarge. So, 
-the standard battle starting with $\iota(\alpha)$ and the replication factor $k$ will end after $(l-k+1)$ steps.
-
-First, we notice that this  number $l$ exists, since the segment $[0,\epsilon_0)$ is well-founded and $\canonseq{\alpha}{i}<\alpha$ for any $i$ and $\alpha>0$.
-Moreover, it is unique:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/mlargeUnicity}
-
-Thus, we would like to define a function, parameterized by $\alpha$ which associates to any  strictly positive integer $k$ the number $l$ such that
-the interval $[k,l]$ is $\alpha$-mlarge. It would be fine to write in \gallina{} a definition like this:
-
-\begin{Coqbad}
-Function L_ (alpha: E0) (i:nat) :  nat := ...
-\end{Coqbad}
-
-But we do not know how to fill the dots yet \dots{}   In the next section, we will 
-use \coq{} to reason  about the \emph{specification} of \texttt{L},
-prove properties of any function which satisfies this specification.
-In Sect.~\ref{sect:L-equations}, we use the \texttt{coq-equations} plug-in
-to define a function \texttt{L\_}, and prove its correctness w.r.t. its specification.
-
-\subsection{Formal specification}
-
-Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider any  function which  maps  any strictly positive integer $k$ to the number $l$, where 
-the interval $[k,l)$ is $\alpha$-mlarge.
-
-\begin{remark}
-In~\cite{KS81} Ketonen and Solovay consider the least natural number $l$ where the interval $[k,l]$ ($l$ included) is $\alpha$-large, and call $H_\alpha$ the function which maps $k$ to $l$.
-We chose to consider intervals $[l,k)$ instead of $[l,k]$
-in order to simplify  some statements and proofs in composition lemmas associated with the ordinals of the form $\alpha\times i$ and 
-$\omega^\alpha\times i + \beta$.
-Clearly, both approaches are related through the equality
-$L_\alpha(k)=H_\alpha(k)+1$, for any non-null $\alpha$ and $k$.
-\end{remark}
-
-Our specification of the function \texttt{L} is as follows:
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/LSpecDef}
-
-\begin{todo}
- Check if the functions $L_\alpha$ are the same as
-\cite{KS81}' functions $f_\alpha$ (p. 297).
-\end{todo}
-
-Note that, for $\alpha\not=0$, the value of $f(0)$ is not specified.
-Nevertheless, the restriction of $f$ to the set of strictly positive integers is unique (up to extensionality).
-
-\input{movies/snippets/Large_Sets/LSpecUnicity}
-
-\subsection{Abstract properties}
-
-Let us now prove properties of any function $f$ (if any) which satisfies 
-\texttt{L\_spec}.
-We are looking for properties which could be used for writing \emph{equations} and prove the correctness of the function generated by the \texttt{coq-equations} plug-in.
-Moreover, they will give us some examples (for small values of $\alpha$).
-
-The properties we consider are defined in \href{../theories/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}.
-
-\label{sect:abstract-arith-prop}
-\input{movies/snippets/Iterates/funLeDef}
-
-Our exploration of the $L_\alpha$\,s  considers the usual cases of a proof by transfinite induction: zero, successors and limit ordinals.
-The lemmas we are going to prove will be applied in a big proof by induction in Sect~\vref{sect:L-correct-proof}.
-
-\index{maths}{Transfinite induction}
-
-\subsubsection{The  ordinal zero}
-\label{sect:L-spec-zero}
-The base case is directly a consequence of the specification.
-
-\input{movies/snippets/Large_Sets/LZeroInv}
-
-\subsubsection{Successor ordinals}
-\label{sect:L-spec-succ}
-Let $\beta$ be some ordinal, and assume the arithmetic function $f$ satisfies 
-the specification $(\texttt{L\_spec}\;\beta)$.  Let $k$ be any natural number.
-Any path from $\texttt{succ}\,\beta$ to $0$ starting at $k+1$ can be decomposed into a first step from $\texttt{succ}\,\beta$ to $\beta$, then a path from
-$\beta$ at $k+2$ to $0$. 
-By hypothesis the interval $[k+2, f(k+2)-1]$ is $\beta$-mlarge.
-But the interval $[k+1, f(k+2)-1]$ is the concatenation of the singleton
-$\{k+1\}$ and the interval $[k+2, f(k+2)-1]$.
-So, the function $\lambda\,k.\,f(k+1)$ satisfies the specification $\texttt{L\_spec}\,\beta$.
-
-Note that our decomposition of intervals works only if the intervals we consider are not empty.
-In order to ensure this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <<= f}, or \texttt{(fun\_le S f)}.
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/SectionSucc}
-
-\subsubsection{Limit ordinals}
-\label{sect:L-spec-lim}
-
-Let $\lambda<\epsilon_0$ be any limit ordinal. In a similar way as for successors, we decompose any path from $\lambda$  into a first step to
-$\canonseq{\lambda}{k}$, followed by a path to $0$. In the following section, we assume that there exists a correct function computing  $L_{\canonseq{\lambda}{k}}$ for any strictly positive $k$.
-
-\input{movies/snippets/Large_Sets/SectionLim}
-
-\subsection{First results}
-
-Applying the previous lemmas on successors and limit ordinals, 
-we obtain a few  correct implementations of \texttt{(L\_spec $\alpha$)} for small values of $\alpha$.
-
-\subsubsection{Finite ordinals}
-
-By iterating the functional \texttt{L\_succ}, we get a realization of
-\texttt{(L\_spec $i$)} for any natural number $i$.
-
-\input{movies/snippets/Large_Sets/LFinDef}
-\vspace{-16pt}
-\input{movies/snippets/Large_Sets/LFinOk}
-
-\subsubsection{The first limit ordinal  \texorpdfstring{$\omega$}{omega}}
-
-The lemmas \texttt{L\_fin\_ok} and \texttt{L\_lim\_ok}   allow us to get 
-by diagonalization a correct implementation for 
-\texttt{L\_spec omega}.
-
-\input{movies/snippets/Large_Sets/LOmegaDef}
-\vspace{-16pt}
-\input{movies/snippets/Large_Sets/LOmegaOk}
-
-\subsubsection{Towards  \texorpdfstring{$\omega^2$}{omega*omega}}
-
-We would like to get exact formulas for the ordinal $\omega^2$, a.k.a.
-$\phi_0(2)$. This ordinal is the limit of the sequence $\omega\times i\;(i \in \mathbb{N})$. Thus, we have to study ordinals of this form, then use 
-our lemma on limits.
-
-The following lemma establishes a path from $\omega\times ( i+1)$ to
-$\omega \times i$.
-
-\input{movies/snippets/Large_Sets/pathToOmegaMult}
-
-Let us consider a path from  $\omega\times(i+1)$ to $0$ starting at $k+1$.
-A first ``big step'' will lead to $\omega\times i$ at $2(k+1)$. If $i>0$, the
-next jump leads to $\omega\times(i-1)$ at $2(2(k+1))+1$, etc.
-
-The following lemma expresses the length of the mlarge sequences associated with any finite multiple of $\omega$.
-
-\input{movies/snippets/Large_Sets/omegaMultMlarge0}
-
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#L_omega_mult}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/LOmegaMultDef}
-
-More generally, we prove the equality $L_{\omega\times i}(k)=2^i\times(k+1)-1$.
-
-\input{movies/snippets/Large_Sets/LOmegaMultEqn}
-
-Correctness of the function \texttt{L\_omega\_mult} is asserted through the following lemma.
-
-\input{movies/snippets/Large_Sets/LOmegaMultOk}
-
-By diagonalization, we obtain a simple formula for $L_{\omega^2}$.
-
-\input{movies/snippets/Large_Sets/LOmegaSquare}
-\input{movies/snippets/Large_Sets/LOmegaSquareEqn}
-\input{movies/snippets/Large_Sets/LOmegaSquareOk}
-
-%%%% ICI
-
-\subsubsection{Going further}
-Let us consider a last example, ``computing'' $L_{\omega^3}$.
-Let us study the canonical sequence associated with $\omega^3$, the elements of which are 
-$\omega^2\times i\;(i\in\mathbb{N}_1)$.
-
-To this end, we prove a generic lemma, which expresses $L_{\omega^\alpha\times i}$ as an iterate of $L_{\omega^\alpha}$.
-Note that in this lemma, we assume that the function associated with $\alpha$ is strictly monotonous and
-greater or equal than the successor function, and prove that $L_{\omega^\alpha\times i}$ satisfies  the same properties.
-
-\inputsnippets{Large_Sets/phi0Mult}
-
-\inputsnippets{Large_Sets/phi0MultOk,
-Large_Sets/Phi0MultSLe}
-
-Let us look now
-at the ordinal $\omega^2\times i$, using \texttt{L\_phi0\_mult}.
-
-\input{movies/snippets/Large_Sets/LOmegaSquareTimes}
-
-We are now ready to get an exact formula for $L_{\omega^3}$, by diagonalization over $L_{\omega^2\times i}$.
-
-\input{movies/snippets/Large_Sets/LOmegaCube}
-
-Thus, for instance, $L_{\omega^3}(3)=L_{\omega^2\times 4}(3)$.
-
-\input{movies/snippets/Large_Sets/LOmegaCube3Eq}
-
-This number is quite big. Using \texttt{Ocaml}'s \texttt{float} arithmetic,
-we can under-approximate it by $2^{3.8\times10^{30}}\times 3.8\times{10^{30}}$.
-
-\begin{Coqsrc}
-# let exp2 x = 2.0 ** x;;
-
-val exp2 : float -> float = <fun>
-#   exp2 95.0 *. 97.0 -. 1.0;;
-- : float = 3.84256588194182037e+30
-# let n = exp2 95.0 ;;
-# let p = n *. 97.0 -. 1.0;;
-val p : float = 3.84256588194182037e+30
-
-Estimation :
-2 ** (3.84 e+30) * 3.84 e+30.
-\end{Coqsrc}
-
-\subsection{Using \texttt{Equations}}
-\label{sect:L-equations}
-
-Note that we did not define any function $L_\alpha$ \emph{for any $\alpha<\epsilon_0$} yet.
-We have got no more than a collection of proved realizations of $\texttt{L\_spec}\;\alpha$ for a few values of $\alpha$.
-
-\index{coq}{Plug-ins!Equations}
-
-Using the \texttt{coq-equations} plug-in by 
-M. Sozeau~\cite{sozeau:hal-01671777}, we will now define a function \texttt{L\_} which maps any ordinal  $\alpha<\epsilon_0$ to a proven realization of 
-$\texttt{L\_spec}\;\alpha$.   
-To this end, we represent ordinals as inhabitants of the type 
-\texttt{E0} of well-formed ordinal terms (see Sect~\vref{sect:E0-def}). So, we define a total function \texttt{L\_} of type
-\texttt{E0 -> nat -> nat}, by transfinite recursion, considering the usual three cases : $\alpha=0$, $\alpha$ is a successor, $\alpha$ is a limit ordinal.
-
-\subsubsection{Definition}
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#L_}{L\_alpha}).}
-
-\label{Functions:L-alpha}
- \index{hydras}{Library Epsilon0!Functions!L\_@L\_ (final step of a minimal path}
-
-\input{movies/snippets/L_alpha/LDef}
-
-This definition results in a bunch of automatically generated lemmas. For instance:
-
-\input{movies/snippets/L_alpha/AboutLEquation1}
-
-In most cases, it may be useful to write human-readable  paraphrases of these statements.
-
-\inputsnippets{L_alpha/Paraphrasesa}
-\inputsnippets{L_alpha/Paraphrasesb}
-\inputsnippets{L_alpha/Paraphrasesc}
-
-Using these three lemmas as rewrite rules, we can prove more properties of the functions \texttt{L\_$\alpha$}.
-
-\input{movies/snippets/L_alpha/LFiniteOmega}
-
-By  well-founded induction on $\alpha$, we prove the following properties:
-
-\input{movies/snippets/L_alpha/LGeS}
-
-\input{movies/snippets/L_alpha/LCorrect}
-
-\label{sect:L-correct-proof}
-
-Please note that the proof of \texttt{L\_correct} applies the lemmas proven in Sections~\ref{sect:L-spec-zero}, ~\ref{sect:L-spec-succ} and ~\ref{sect:L-spec-lim}.
-Our previous study of \texttt{L\_spec} allowed us to pave the way for the definition by \texttt{Equations} and the correctness proof.
-
-\paragraph*{\gaiasign}
-The module
-\href{../theories/html/gaia_hydras.GL_alpha.html}{gaia\_hydras.GL\_alpha} contains an adaptation of
-\href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0/L\_alpha}.
-
-\subsubsection{Back to hydra battles}
-\label{def:L-alpha}
-
-Lemma \texttt{battle\_length\_std } of
-Module~\href{../theories/html/hydras.Hydra.Battle_length.html}{Hydra.Battle\_length} relates the length of standard battles with the functions $L_\alpha$.
-
-\input{movies/snippets/Battle_length/battleLengthStd}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Instead of considering standard paths and battles, consider battles where the replication factor is a constant $k$.
-Please use \texttt{Equations} in order to define the function that computes the length of the $k$-path which leads  from $\alpha$ to $0$.
-Prove a few  exact formulas and minoration lemmas.
-\end{exercise}
-
-\section{A variant of the Hardy hierarchy}
-
-\label{sect:hardy}
-
-In order to give a feeling on  the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the  \emph{Hardy hierarchy} of fast growing functions,
-presented for instance in~\cite{Promel2013}. 
-\index{maths}{Rapidly growing functions!Hardy Hierarchy}
-
-\begin{remark}
-  Indeed, the functions presented in this section are a \emph{variant} of the Hardy hierarchy of functions.
-In the future versions of this development, we will correct the references to the literature.
-For the time being, we call our functions $H'_\alpha$ in order to underline the difference from ``classic'' Hardy functions.
-\end{remark}
-
-For each ordinal $\alpha$ below $\epsilon_0$, $H'_\alpha$ is 
-a total arithmetic function, defined  by  transfinite recursion on $\alpha$, according to three cases:
-
-\index{maths}{Transfinite induction}
-
-\begin{itemize}
-\item If $\alpha=0$, then $H'_\alpha (k)= k$ for any natural number $k$.
-\item If $\alpha=\textrm{succ}(\beta)$, then 
-$H'_\alpha(k)=H'_\beta(k+1)$ for any $k \in \mathbb{N}$
-\item If $\alpha$ is a limit ordinal, then 
-$H'_\alpha(k) = H'_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
-\end{itemize}
-
-\begin{remark}
- The ``classic'' definition of the Hardy hierarchy differs in the third equation.
-
-\begin{itemize}
-\item If $\alpha=0$, then $H_\alpha (k)= k$ for any natural number $k$.
-\item If $\alpha=\textrm{succ}(\beta)$, then 
-$H_\alpha(k)=H_\beta(k+1)$ for any $k \in \mathbb{N}$
-\item If $\alpha$ is a limit ordinal, then 
-$H_\alpha(k) = H_{(\canonseq{\alpha}{k})}(k)$ for any $k\in \mathbb{N}$.
-\end{itemize}
-
-\end{remark}
-
-\subsection{Definition in \texttt{Coq}}
-
-We define a function \texttt{H'\_} of type \texttt{E0 -> nat -> nat} by transfinite induction over the type \texttt{E0} of the well formed ordinals below $\epsilon_0$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}}
-
-\index{hydras}{Library Epsilon0!Functions!H\_@H\_ (Hardy hierarchy (variant))}
-\index{coq}{Plug-ins!Equations}
-\label{Functions:Hprime-alpha}
-
-\input{movies/snippets/Hprime/HprimeDef}
-\input{movies/snippets/Hprime/HprimeDefb}
-
-\input{movies/snippets/Hprime/paraphrasesa}
-\input{movies/snippets/Hprime/paraphrasesb}
-\input{movies/snippets/Hprime/paraphrasesc}
-\input{movies/snippets/Hprime/paraphrasesd}
-
-\subsection{First  steps of the H' hierarchy}
-Using rewrite rules from \texttt{H'\_eq1} to \texttt{H'\_succ\_eqn}, we can explore the functions $H'_\alpha$ for  small values of $\alpha$.
-
-\paragraph*{\gaiasign}
-The module
-\href{../theories/html/gaia_hydras.GHprime.html}{gaia\_hydras.GHprime} contains an adaptation of
-\href{../theories/html/hydras.Epsilon0.Hprime.html}{Epsilon0.Hprime}.
-
-\subsubsection{Finite ordinals}
-
-By induction on $i$, we prove a simple expression of \texttt{H'\_ (Fin i)}, where 
-\texttt{Fin $i$}  is the $i$-th finite ordinal.
-
-\input{movies/snippets/Hprime/HprimeFin}
-
-\subsubsection{Multiples of \texorpdfstring{$\omega$}{omega}}
-
-Since the canonical sequence of $\omega$ is composed of finite ordinals, 
-it is easy to get the formula associated with $H'_\omega$.
-
-\input{movies/snippets/Hprime/HprimeOmega}
-
-Before going further, we prove a useful rewriting lemma:
-
-\input{movies/snippets/Hprime/HprimePlusFin}
-
-Then, we get easily formulas for $H'_{\omega+i}$, and $H'_{\omega\times i}$ for any natural number $i$.
-
-\input{movies/snippets/Hprime/HprimeExamplesa}
-\input{movies/snippets/Hprime/HprimeExamplesb}
-\input{movies/snippets/Hprime/HprimeExamplesc}
-\input{movies/snippets/Hprime/HprimeExamplesd}
-
-Since $\omega^2$ is a limit ordinal, we prove the following equality: 
-\[H'_{\omega^2} (k) = 2 ^ {k+1} \times (k+1) - 1\]
-
-\input{movies/snippets/Hprime/HprimeOmegaSqr}
-
-\subsubsection{New limits}
-
-Our next step would be to prove an exact formula for $H'_{\omega^\omega}(k)$.
-Since the canonical sequence of $\omega^\omega$ is composed of all the
-$\omega^i$, we first need to express $H'_{\omega^i}$ for any natural number $i$.
-
-Let $i$ and $k$ be two natural numbers. 
-The ordinal $\canonseq{\omega^(i+1)}{k}$ is the product
-$\omega^i \times k$, so we need also to consider ordinals of this form.
-
-\begin{enumerate}
-\item First,  we express $H'_{\omega^\alpha \times (i+2)}$ in terms of
-$H'_{\omega^\alpha \times (i+1)}$.
-
-\input{movies/snippets/Hprime/HprimeOmegaTerm1}
-
-\item
-Then, we prove by induction on $i$ that $H'_{\omega^\alpha \times (i+1)}$ is just the
-$(i+1)$-th iterate of $H'_{\omega^\alpha}$.
-
-\input{movies/snippets/Hprime/HprimeOmegaTerm}
-
-\item In particular, we derive a formula for $H'_{\omega^{i+1}}$.
-
-\input{movies/snippets/Hprime/HprimeSuccFun}
-\input{movies/snippets/Hprime/HprimePhi0SI}
-
-\item We get now a  formula for $H'_{\omega^3}$:
-
-\input{movies/snippets/Hprime/HprimeOmegaCube}
-\end{enumerate}
-
-\subsubsection{A numerical example}
-\label{sect:bignum}
-
-It is hard to capture the complexity of this function by looking only at this
-``exact'' formula. 
-Let us consider a simple example: the number $H'_{\omega^3}(3)$.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3a}
-
-Thus, the number $H_{\omega^3}(3)$ can be written as four nested applications of $f$.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3b}
-
-In a more classical writing, this number is displayed as follows:
-
-{\Large
-$$
-H'_{\omega^3}(3) =  2 ^ {(2 ^ {N + 1} \, (N+1) )}   \,  (2 ^ {N+1} \, ( N +1) ) - 1
-$$
-}
-
-In order to make this statement more readable, we introduce a local definition.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3c}
-
-This number looks quite big; let us compute an approximation with \texttt{Ocaml}:
-
-\begin{Coqsrc}
-# (2.0 ** 64.0 *. 64.0 -. 1.0);; 
-\end{Coqsrc}
-
-\begin{Coqanswer}
-- : float = 1.1805916207174113e+21
-\end{Coqanswer}
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3d}
-
-We leave as an exercise to determine the best approximation as possible of
- the size of this number (for instance its number of digits).  For instance, if
-we do not take into account the multiplications in the formula above,
-we obtain that, in base $2$, the number $H'_{\omega^3}(3)$ has at least
-$2^{10^{21}}$  digits. But it is still an under-approximation !
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3e}
-
-\subsubsection{A formula for \texorpdfstring{$H'_{\omega^\omega}$}{\texttt{H'(phi0(omega))}}}
-
-Now, we can get at last an exact formula for $H'_{\omega^\omega}$.
-
-\input{movies/snippets/Hprime/HprimePhi0Omega}
-
-Using extensionality of the functional \texttt{iterate}, we also get a closed formula.
-
-\input{movies/snippets/Hprime/HprimePhi0OmegaClosed}
-
-Note that this formula contains two occurrences of the functional \texttt{iterate}, the outer one is in fact a second-order iteration (on type \texttt{nat -> nat)}
-and the inner one  first-order (on type \texttt{nat}).
-
-\subsection{Abstract properties of  
-\texorpdfstring{$H'_\alpha$}{H'}}
-~\label{sect:H-alpha-prop}
-
-Since pure computation seems to be useless for dealing with expressions of the form $H'_\alpha(k)$, even for small values of $\alpha$ and $k$, we need to prove theorems for comparing $H'_\alpha(k)$ and $H'_\beta(l)$, in terms of comparison
-between $\alpha$ and $\beta$ on the one hand, $k$ and $l$ on the other hand.
-
-But beware of fake theorems! For instance, one could believe that $H'$ is monotonous in its first argument. The following proof shows this is false.
-
-\input{movies/snippets/Hprime/HprimeNonMono1}
-
-On the contrary, the functions of the $H'$ hierarchy have the following five properties~\cite{KS81}: for any $\alpha < \epsilon_0$,
-\begin{itemize}
-\item the function $H'_\alpha$ is strictly monotonous :
-      For all $n,p \in\mathbb{N}, n < p \Rightarrow H'_\alpha(n)< H'_\alpha(p)$.
-\item If $\alpha \not= 0$, then for every $n$, $n<H'_\alpha(n)$.
-\item The function $H'_\alpha$ is pointwise less or equal than $H'_{\alpha+1}$
-
-\item For any $n\geq 1$, $H'_\alpha(n)<H'_{\alpha+1}(n)$.
-\emph{We say that $H'_{\alpha+1}$ dominates $H'_\alpha$ from $1$}.
-\item For any $n$ and $\beta$, if $\alpha \xrightarrow[n]{} \beta$, then
-$H'_\beta(n)\leq H'_\alpha(n)$.
-\end{itemize}
-
-\index{maths}{Abstract properties of arithmetic functions}
-\index{hydras}{Abstract properties of arithmetic functions}
-
-\index{maths}{Transfinite induction}
-
-In \coq{}, we follow the  proof of~\cite{KS81}. This proof is mainly a single  proof by transfinite induction on $\alpha$ of the conjunction of the five properties.
-For each $\alpha$, the three cases : $\alpha=0$, $\alpha$ is a limit, and 
-$\alpha$ is a successor are considered. Inside each case, the five sub-properties are proved sequentially, using the abstract properties defined in Sect~\vref{sect:abstract-arith-prop}
-
-\input{movies/snippets/Hprime/PAP}
-
-Using a few lemmas \emph{à la} Ketonen-Solovay, we prove that
-if $\alpha<\beta$, then $H'_\beta$ eventually dominates
-$H'_\alpha$.
-We let the reader look at the proof (Section \texttt{Proof\_of\_H'\_mono\_l} of \href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}).
-
-\noindent
-\input{movies/snippets/Hprime/HprimeDom}
-
-%ICI
-\subsection{Comparison between \texttt{L\_} and \texttt{H'\_} }
-
-By well-founded induction on $\alpha$, we prove that
- $H'_\alpha(i) \leq L_\alpha(i+1)$ for any $i$.
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#H'_L_}{Epsilon0.L\_alpha}}
-
-\input{movies/snippets/L_alpha/HprimeL}
-
-\subsubsection{Back to hydras}
-
-The following theorem relates the length of (standard) battles with the the $H'$ family of fast growing functions.
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems}}
-
-\input{movies/snippets/Hydra_Theorems/battleLengthStdHardy}
-
-Since $H'_{\omega^3}(3)= H'_{\omega^3+3}(0)$, the big number shown in Section~\vref{sect:bignum} is less or equal than 
-the number of steps of the battle with the initial hydra of
-Figure~\ref{fig:start}, and  $0$ as initial replication factor.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[very thick, scale=0.3]
-  \node (h1) at (4,0){$\bullet$};
-  \node (h2) at (2,2){$\bullet$};
-  \node (h3) at (0,4){$\bullet$};
-  \node (h4) at (1,4){$\bullet$};
-  \node (h5) at (2,4){$\bullet$};
-  \node (h6) at (4.5,2){$\bullet$};
-  \node (h7) at (6.5,2){$\bullet$};
-  \node (h8) at (8.5,2){$\bullet$};
-    \draw (h1) -- (h2) ;
-    \draw (h2) -- (h3) ;
-    \draw (h2) -- (h4) ;
-    \draw (h2) -- (h5) ;
-    \draw (h1) -- (h6) ;
-    \draw (h1) -- (h7) ;
-    \draw (h1) -- (h8);
- \end{tikzpicture}
-
-  \caption{The hydra corresponding to the ordinal $\omega^3+3$}
-  \label{fig:start}
-\end{figure}
-
-\section{A variant of the Wainer hierarchy (functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
-\label{sect:wainer}
-
-\index{maths}{Rapidly growing functions!Wainer Hierarchy}
-
-Ketonen and Solovay introduce in~\cite{KS81} a ``trivial'' variant of the Wainer hierarchy~\cite{BW85, Wainer1970} of fast growing functions, indexed by ordinals below $\epsilon_0$.
-The functions $F_\alpha$ are defined by the following equations.
-
-\label{F_equations}
-\begin{itemize}
-\item $F_0(i)=i+1$
-\item $F_{\beta+1}(i)= (F_\beta)^{(i+1)}(i)$, where $f^{(i)}$ is the $i$-th iterate of $f$.
-\item $F_\alpha(i) = F_{\canonseq{\alpha}{i}} (i)$ if $\alpha$ is a limit ordinal.
-\end{itemize}
-
-\begin{remark}
-The difference with the ``classic'' Wainer hierarchy 
-$f_\alpha\;(\alpha<\epsilon_0)$ lies in the second equation:
-$f_{\beta+1}(i) = (f_\beta)^{(i)}(i)$ and not
-$f_{\beta+1}(i) = (f_\beta)^{(i+1)}(i)$.
-
-A module about 
-the classic Wainer hierarchy is in preparation.
-
-Note also that \cite{KS81} defines also $F_{\epsilon_0}$ (by the third equation).
-Since $\epsilon_0$ is not representable in type \texttt{E0}, our implementation in \coq{} does not take $F_{\epsilon_0}$ into account.
-
-\end{remark}
-
-A first attempt is to write a definition of $F_\alpha$ by equations, in the same way as for $H_\alpha$ (the functional \texttt{iterate} has already been used in Sect.\vref{Functions:iterate}).
-
-\index{hydras}{Library Prelude!iterate}
-
-\input{movies/snippets/Iterates/iterateDef}
-
-\emph{From\href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}~}.
-
-\index{coq}{Plug-ins!Equations}
-
- \input{movies/snippets/F_alpha/FailDemo}
-
-We presume that this error comes from the recursive call of \texttt{F\_} inside
-an application of \texttt{iterate}. The workaround we propose is to define first 
-the iteration of \texttt{F\_}  as an helper $F^*$, then to define the function $F$ as a ``iterating $F^*$ once''.
-
-\texttt{Equations} accepts the following definition, relying on  lexicographic ordering on pairs $(\alpha,n)$.
-
-\label{sect:F-equations}
-
-\index{coq}{Plug-ins!Equations}
-\label{Functions:F-alpha}
-\index{maths}{Rapidly growing functions}
-\index{hydras}{Library Epsilon0!Functions!F\_@F\_ (Wainer hierarchy)}
-
-\input{movies/snippets/F_alpha/goodDefa}
-\input{movies/snippets/F_alpha/goodDefb}
-
-It is quite easy to prove that our functional \texttt{F\_} satisfies the equations on page~\pageref{sect:F-equations}.
-\index{hydras}{Library Prelude!iterate}
-
-\input{movies/snippets/F_alpha/FEquations}
-
-As for the Hardy functions, we can use these equalities as rewrite rules for
-``computing'' some values of $F_\alpha(i)$, for small values of $\alpha$.
-
-\input{movies/snippets/F_alpha/FirstValues}
-
-Like in Sect~\ref{sect:H-alpha-prop}, we prove by induction the following properties (see~\cite{KS81}).
-
-\input{movies/snippets/F_alpha/FalphaThms}
-
-As a corollary, we prove that, if $\beta<\alpha$, then
-$F_\alpha$ dominates $F_\beta$ (p. 284 of~\cite{KS81}).
-
-\input{movies/snippets/F_alpha/FDomContext}
-\input{movies/snippets/F_alpha/FDom}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove the following property:
-
-\begin{Coqsrc}
-Lemma LF3 : dominates_from  2 (F_ 3) (fun  n => iterate exp2 n n).
-\end{Coqsrc}
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
-\end{exercise}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that, for any $\alpha\geq 3$ and $n\geq 2$,
-$F_\alpha(1+n)\geq 2^{F_\alpha(n)}$.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
-\end{exercise}
-
-%ici
-\index{hydras}{Exercises}
-
-\begin{exercise}
-It is tempting to prove a simple property of monotony 
-of the function \texttt{F\_}.
-
-\begin{quote}
-   Let $\alpha\leq\beta<\epsilon_0$. For any $n\geq 2$,
-$F_\alpha(n)\leq F_\beta(n)$. 
-\end{quote}
-Prove or disprove this statement.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/is_F_monotonous.v}{exercises/ordinals/is\_F\_monotonous.v}.}
-\end{exercise}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-
-Prove that for any $n\geq 2$, $\textrm{Ack}\,\,n\,n\leq  F_\omega(n)$, where \textrm{Ack} is the Ackermann function.
-Next, prove that $F_\alpha$ is not primitive recursive, for any $\alpha\geq\omega$  (please see Sect.~\vref{sect:ack-not-PR}).
-On the other hand, please show that for any natural number $n$, the function $F_n$ is primitive recursive.
-Thus $F_\alpha$ is primitive recursive if and only if $\alpha$ is finite.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_omega.v}{exercises/ordinals/F\_Omega.v}.
-Properties of the Ackermann function are studied in
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/Ack.v}{theories/ordinals/MoreAck/Ack.v} and
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/AckNotPR.v}{theories/ordinals/MoreAck/AckNotPR.v}
-.}
-\end{exercise}
-
-\paragraph*{\gaiasign} Library~\href{../theories/html/gaia_hydras.GF_alpha.html}{gaia\_hydras.GF\_alpha}  defines the functions $F_\alpha$ for \gaia's world.
-
-\section{More about rapidly growing functions}
-
-In Sect.~\vref{sect:battle-length-notPR}, we prove that the length of hydra-battles (for a given hydra, according to the initial replication factor) is not primitive recursive in general. 
-This proof uses properties of the Ackermann function, and the $H'_\alpha$, $F_\alpha$, $L_\alpha$ families of functions.
-
-% ---------------------------------------------------
-\input{gaia-chapter}
-
-%----------------------------------------------------------------------
-\chapter[Countable ordinals (after Sch\"{u}tte)]{Kurt Schütte's axiomatic definition of countable ordinals}
-
-\label{chap:schutte} 
-%ON
-
-In the present chapter, we  compare our implementation of the segment $[0,\epsilon_0)$ with a mathematical text in order to ``validate'' our constructions.
-Our reference here is the axiomatic definition of the set of countable ordinals,
-in chapter V of Kurt Schütte's book `` Proof Theory ''~\cite{schutte}.
-
-\begin{remark}
-\emph{In all this chapter, the word ``ordinal'' will be considered as a synonymous of
-``countable ordinal''}  
-\end{remark}
-
-Schütte's definition of countable ordinals relies on the following three axioms:
-
-There  exists a strictly ordered set , such that
-\begin{enumerate}
-\item  $(\mathbb{O},<)$ is well-ordered
-\item Every bounded subset of $\mathbb{O}$  is countable
-\item Every countable subset of $\mathbb{O}$  is bounded.
-\end{enumerate}
-
-Starting with these three axioms, Schütte re-defines the vocabulary about ordinal numbers: the null ordinal $0$, limits and successors, the addition of ordinals, the infinite ordinals $\omega$, $\epsilon_0$, $\Gamma_0$, etc.
-
-This chapter describes an adaptation to \coq{} of Schütte's axiomatization. 
- Unlike the rest of our libraries, our library
-\href{../theories/html/hydras.Schutte.Schutte.html}{hydras.Schutte}
-is not constructive, and relies on several axioms.
-
-\begin{itemize}
-\item First, please keep in mind  that the set of countable ordinals is not countable.
-Thus, we cannot hope to represent all countable ordinals as finite terms of an inductive type, which was possible with  the set of ordinals strictly less than $\epsilon_0$ (resp.
-$\Gamma_0$)
-\item We tried to be as close as possible to K.
-Schütte's text, which uses ``classical'' mathematics : excluded middle, Hilbert's $\epsilon$ (choice) and Russel's $\iota$ (definite description) operators.
-Both operators allow us to write definitions close to the natural mathematical language, such as ``$\textrm{succ}$ is \emph{the} least ordinal strictly greater than $\alpha$''
-\item Please note that only the library \href{../theories/html/hydras.Schutte.Schutte.html}{Schutte/*.v} is ``contaminated'' by axioms, and that the rest of our libraries remain constructive.
-\end{itemize}
-
-\section{Declarations and axioms}
-
-Let us declare a type 
-\texttt{Ord} for representing countable ordinals, and a binary relation
- \texttt{lt}. Note that, in our development, \texttt{Ord} is a type, while the \emph{set} of countable ordinals (called $\mathbb{O}$ by Schütte) 
-is the full set over the type \texttt{Ord}.
-
-\label{types:Ord}
-
-We use Florian Hatat's library on countable sets, written as he was a student of  \emph{\'Ecole Normale Supérieure de Lyon}.
-A set $A$ is countable if there is an injective function from $A$ to $\mathbb{N}$ (see
-Library \href{../theories/html/hydras.Schutte.Countable.html}%
-{\texttt{Schutte.Countable}}).
-
-\vspace{6pt}
-
-\emph{From Module\href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{Schutte.Schutte\_basics}}}
-
-\index{hydras}{Library Schutte!Types!Ord}
-
-\input{movies/snippets/Schutte_basics/OrdDecl}
-
-Schütte's first axiom tells that \texttt{lt} is a well order on the set 
-\texttt{ordinal} (The  class \texttt{WO} is defined in
-Module~\href{../theories/html/hydras.Schutte.Well_Orders.html}{Schutte.Well\_Orders.v}).
-
-\index{hydras}{Library Schutte!Type classes!WO@ WO (well order)}
-
-\label{types:WO}
-
-\input{movies/snippets/Well_Orders/Mdecl}
-
-\input{movies/snippets/Well_Orders/WODef}
-
-\input{movies/snippets/Schutte_basics/AX1}
-
-The second and third axioms say that a subset $X$ of $\mathbb{O}$ is
-(strictly) bounded if and only if it is countable.
-
-\input{movies/snippets/Schutte_basics/AX23}
-
-\texttt{AX2} and \texttt{AX3} could have been replaced by a single axiom (using the \texttt{iff} connector), but we decided to respect as most as possible the structure of Schütte's definitions.
-
-\section{Additional  axioms}
-
-The adaptation of Schütte's mathematical discourse to \coq{} led us to
-import a few axioms from the standard library. We encourage the reader to consult \coq{}'s FAQ about the safe use of axioms
- \url{https://github.com/coq/coq/wiki/The-Logic-of-Coq#axioms}.
-
-\subsubsection{Classical logic}
-
-In order to work with classical logic, we import the module
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Classical.html}{Coq.Logic.Classical}  of \coq{}'s standard library, specifically the following axiom:
-
-\begin{Coqsrc}
- Axiom classic : forall P:Prop, P \/ ~P.
-\end{Coqsrc}
-
-\subsubsection{Description operators}
-
-In order to respect Schütte's style, we imported also the library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Epsilon.html}{\texttt{Coq.Logic.Epsilon}}.  The rest of this section presents a few examples of
-how Hilbert's choice operator and Church's definite description allow us
- to write understandable definitions (close to the mathematical natural language).
-
-\subsubsection{The definition of zero}
-
-According to the  definition of a well order, every non-empty subset of \texttt{Ord} has a least element. Furthermore, this least element is unique. We would like to call this element  \texttt{zero}.
-
-\input{movies/snippets/Schutte_basics/iotaDemoa}
-\input{movies/snippets/Schutte_basics/iotaDemob}
-
-Indeed, the basic logic of  \coq{} does not allow us to eliminate a proof of a proposition 
-$\exists!\,x:A,\,P(x)$ for building a term whose type lies in the sort \texttt{Type}. 
-The reasons for this impossibility are explained in many documents~\cite{BC04, chlipalacpdt2011, Coq}.
-
-Let us import the library \texttt{Coq.Logic.Epsilon}, which contains the following axiom and lemmas.
-
-\input{movies/snippets/MoreEpsilonIota/EpsilonStatement}
-
-Hilbert's $\epsilon$ \emph{operator} is derived from this  axiom.
-
-\input{movies/snippets/MoreEpsilonIota/EpsilonDef}
-
-If we consider the \emph{unique existential} quantifier $\exists!$, we obtain
-Church's \emph{definite description operator}.
-
-\input{movies/snippets/MoreEpsilonIota/iotaDef}
-
-Indeed, the operators \texttt{epsilon} and \texttt{iota} allowed us to make our definitions 
-quite close to Schütte's text. Our libraries \href{../theories/html/hydras.Schutte.MoreEpsilonIota.html}%
-{\texttt{Schutte.MoreEpsilonIota}}
-and
-\href{../theories/html/hydras.Schutte.PartialFun.html}%
-{\texttt{Schutte.PartialFun}} are extensions of \texttt{Coq.logic.Epsilon} for making easier 
-such definitions. See also an article in french~\cite{PCiota}.
-
-\input{movies/snippets/MoreEpsilonIota/Defs}
-
-In order to use these tools,  we have to tell \coq{}  that the declared type \texttt{Ord} is not empty:
-
-\input{movies/snippets/Schutte_basics/inhOrd}
-
-We are now able to define \texttt{zero} as the least ordinal. For this purpose,
-we define a function returning the least element of any [non-empty]  subset.
-
-\emph{From Module\href{../theories/html/hydras.Schutte.Well_Orders.html}%
-  {\texttt{Schutte.Well\_Orders}}}
-
-\input{movies/snippets/Well_Orders/MDecl}
-\input{movies/snippets/Well_Orders/theLeast}
-
-\vspace{4pt}
-
-From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{~Schutte.Schutte\_basics}}
-
-\label{Constants:zero:Ord}
-\index{hydras}{Library Schutte!Constants!zero}
-
-\input{movies/snippets/Schutte_basics/zeroDef}
-
-We want to prove now that zero is less than or equal to any ordinal number.
-
-\input{movies/snippets/Schutte_basics/zeroLe}
-
-\subsubsection{Remarks on \texttt{epsilon} and \texttt{iota}}
-
- What would happen in case of a misuse of \texttt{epsilon} or \texttt{iota} ?
-For instance, one could give a unsatisfiable specification to \texttt{epsilon} or 
-a specification for \texttt{iota} that admits several realizations.
-
-Let us consider an example:
-
-\input{movies/snippets/Schutte_basics/BadBottoma}
-
-Since we won't be able to prove the proposition
-\linebreak \Verb|(exists! a: Ord, least_member (Empty_set Ord) a)|, the only properties we would be able to prove about \texttt{bottom} would be \emph{trivial} properties, 
-\emph{i.e.}, satisfied by \emph{any} element of type \texttt{Ord}, like for instance
-\texttt{bottom = bottom}, or \texttt{zero <= bottom}.
-
-\input{movies/snippets/Schutte_basics/trivialProps}
-
-On the other hand, the following attempt fails, because of the unprovable first subgoal (please notice that the second subgoal is easy to solve !).
-
-\input{movies/snippets/Schutte_basics/Failure}
-
-In short, using \texttt{epsilon} and \texttt{iota} in our implementation of countable ordinals after Schütte has two main advantages.
-
-\begin{itemize}
-\item It allows us to give a \emph{name} (using \texttt{Definition}) two witnesses 
-of existential quantifiers (let us recall that, in classical logic, one may consider non-constructive proofs of existential statements)
-\item By separating definitions from proofs of [unique] existence, one may make definitions  more concise and readable. Look for instance at 
-the definitions of  \texttt{zero}, \texttt{succ}, \texttt{plus}, etc. in the rest of this chapter.
-\end{itemize}
-%%%% ICI ICI
-
-\section{The  successor function}
-
-The definition of the function \texttt{succ:Ord -> Ord} is very concise. The successor of any ordinal $\alpha$ is the smallest ordinal strictly greater than $\alpha$.
-
-\label{Functions:succ-sch}
-\index{hydras}{Library Schutte!Functions!succ}
-
-\input{movies/snippets/Schutte_basics/succDef}
-
-Using \texttt{succ}, we define the following predicates.
-
-\input{movies/snippets/Schutte_basics/isSuccIsLimit}
-
-% \begin{remark}
-% Please look at remark~\vref{warning:coercions}.
-
-How do we prove properties of the successor function?
-First, we make its specification explicit.
-
-\input{movies/snippets/Schutte_basics/succSpec}
-
-Then, we prove that our function \texttt{succ} meets this specification.
-
-\input{movies/snippets/Schutte_basics/succOka}
-
-We have now to prove that the set of all ordinals strictly greater than $\alpha$ has a unique least element.
-But the singleton set $\{\alpha\}$ is countable, hence  bounded (by the axiom \texttt{AX3}).
-Hence; the set $\{\beta\in\mathbb{O}|\alpha < \beta\}$ is not empty
-and therefore has a unique least element.
-
-The rest of the \coq{} proof script is quite short.
-
-\input{movies/snippets/Schutte_basics/succOkb}
-
-We can ``uncap'' the description operator for proving properties of the
-\texttt{succ} function.
-
-\input{movies/snippets/Schutte_basics/succProps}
-
-\section{Finite ordinals}
-
-Using \texttt{succ}, it is now easy to define recursively all the finite ordinals.
-
-\label{sect:notation-F-sch}
-
-\input{movies/snippets/Schutte_basics/finiteDef}
-
-\section{The definition of \texttt{omega}}
-In order to define $\omega$, the first infinite ordinal, we use an operator which
-``returns'' the least upper bound (if it exists) of a subset $X\subseteq \mathbb{O}$.
-For that purpose, we first use a predicate:
-(\texttt{is\_lub $D$ \textit{lt} $X$ $a$}) if $a$ belongs to $D$ and is the least 
-upper bound  of $X$ (with respect to \textit{lt}).
-
-\input{movies/snippets/Lub/isLubDef}
-
-\input{movies/snippets/Schutte_basics/supDef}
-
-Then, we define the function \texttt{omega\_limit} which returns the least upper bound 
-of the  (denumerable) range of any sequence \texttt{s: nat -> Ord}. 
-By \texttt{AX3} this range is bounded, hence the set of its upper bounds is not empty and has a least element.
-Then we define \texttt{omega} as the limit of the sequence of finite ordinals.
-
-\input{movies/snippets/Schutte_basics/omegaDef}
-
-\label{sect:notation-omega}
-
-Among the numerous properties of the ordinal $\omega$, let us quote the following ones
-(proved in Module 
-\href{../theories/html/hydras.Schutte.Schutte_basics.html\#finite_lt_omega}{\texttt{Schutte.Schutte\_basics}})
-
-\input{movies/snippets/Schutte_basics/omegaPropsa}
-\input{movies/snippets/Schutte_basics/omegaPropsb}
-\input{movies/snippets/Schutte_basics/omegaPropsc}
-\input{movies/snippets/Schutte_basics/omegaPropsd}
-
-\subsection{Ordering functions and ordinal addition}
-
-After having defined the finite ordinals and the infinite ordinal $\omega$, we  define the sum $\alpha+\beta$ of two countable ordinals.
-Schütte's definition looks like the following one:
-
-\begin{quote}
-``$\alpha+\beta$ is the $\beta$-th ordinal greater than or equal to $\alpha$''
-\end{quote}
-
-The purpose of this section is to give a meaning to the construction
-``the $\alpha$-th element of $X$''  where $X$ is any non-empty subset of $\mathbb{O}$.
-We follow Schütte's approach, by defining the notion of \emph{ordering functions},
-a way to associate a unique ordinal to each element of $X$.
-Complete definitions and proofs can be found in Module
- \href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
-{\texttt{Schutte.Ordering\_Functions}} ).
-
-\subsection{Definitions}
-
-A \emph{segment} is a set $A$ of ordinals such that, whenever  $\alpha\in A$ and
-$\beta<\alpha$, then $\beta\in A$; a segment is  \emph{proper} if it strictly included in $\mathbb{O}$.
-
-\input{movies/snippets/Ordering_Functions/segmentDef}
-
-Let  $A$ be a segment, and $B$ a subset of $\mathbb{O}$ : an \emph{ordering function for $A$ and  $B$} is a strictly increasing bijection from $A$ to $B$.
-The set $B$ is said to be an \emph{ordering segment} of $A$.
-Our definition in \coq{} is a direct translation of the mathematical text of~\cite{schutte}.
-
-\index{maths}{Ordinal numbers!Ordering functions}
-\index{hydras}{Library Schutte!Predicates!ordering function@ordering\_function}
-
-\input{movies/snippets/Ordering_Functions/orderingFunctionDef}
-
-We are now able to associate with any subset $B$ of $\mathbb{O}$ its ordering segment and ordering function.
-
-\input{movies/snippets/Ordering_Functions/ordDef}
-
-Thus (\texttt{ord $B \;\alpha$}) is the $\alpha$-th element of $B$.
-Please note that the last definition uses the epsilon-based operator \texttt{some} and
-not \texttt{the}. This is due to the fact that we cannot prove the unicity (w.r.t. Leibniz' equality) of the ordering function of a given set. 
-By contrast, we admit the axiom  \texttt{Extensionality\_Ensembles}, from the library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Sets.Ensembles.html}{Coq.Sets.Ensembles}, so we use the operator \texttt{the} in the definition of
-\texttt{the\_ordering\_segment}.
-
-One of the main theorems of
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html\#ordering_function_ex}%
-{\texttt{Ordering\_Functions}} 
-associates a unique segment and a unique (up to extensionality) ordering function to every subset $B$ of $\mathbb{O}$.
-
-\input{movies/snippets/Ordering_Functions/orderingFunctionEx}
-
-Thus,  our function \texttt{ord}  which enumerates the elements of $B$ is defined in a non-ambiguous way.
-Let us quote the following theorems (see Library
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
-{\texttt{Schutte.Ordering\_Functions}} for more details).
-
-\input{movies/snippets/Ordering_Functions/orderingLe}
-\input{movies/snippets/Ordering_Functions/Th1352}
-
-\subsection{Ordinal addition}
-
-We are now ready to define and study addition on the type \texttt{Ord}.
-The following definitions and proofs can be consulted in Module
-\href{../theories/html/hydras.Schutte.Addition.html}%
-{\texttt{Schutte.Addition.v}}.
-
-\index{hydras}{Library Schutte!Functions!plus}
-
-\input{movies/snippets/Addition/additionDef}
-
-In other words,  $\alpha + \beta$ is the  $\beta$-th ordinal greater than or equal to $\alpha$. 
-Thanks to generic properties of ordering functions, we can show the following 
-properties of addition on $\mathbb{O}$. First, we prove a useful lemma:
-
-\input{movies/snippets/Addition/plusElim}
-
-As a use-case, let us prove that $0$ is a right neutral element of $+$.
-
-\input{movies/snippets/Addition/alphaPlusZero}
-
-The following lemmas are proved the same way.
-
-\input{movies/snippets/Addition/bunchOfLemmas}
-
-The following lemmas are not direct applications of \texttt{plus\_elim}.
-
-\input{movies/snippets/Addition/plusAssoc}
-\input{movies/snippets/Addition/finitePlusInfinite}
-
-It is interesting to compare the proof of these lemmas with the
-computational proofs of the corresponding statements in Module
-\href{../theories/html/hydras.Epsilon0.T1.html}%
-{\texttt{Epsilon0.T1}}. 
-For instance, the proof of the lemma 
-\texttt{one\_plus\_omega} uses the continuity of ordering functions (applied to  \texttt{(plus 1)}) and compares the limit of the $\omega$-sequences $i_{(i \in \mathbb{N})}$ and
-$(1+i)i_{(i \in \mathbb{N})}$, whereas in the library  \texttt{Epsilon0/T1}, the equality 
-$1+\omega=\omega$ is just proved with \texttt{reflexivity}!
-
-\subsubsection{Multiplication by a natural number}
-
-The multiplication of an ordinal by a natural number is defined in terms of addition.
-This operation is useful for the study of Cantor normal forms.
-
-\input{movies/snippets/Addition/multFin}
-
-\section{The exponential of basis \texorpdfstring{$\omega$}{omega}}
-
-In this section, we define the function which maps any $\alpha\in\mathbb{O}$ to
-the ordinal  $\omega^\alpha$, also written 
-$\phi_0(\alpha)$. 
-It is an opportunity to apply the definitions and results of the preceding section. 
-Indeed,  Schütte first defines a subset of $\mathbb{O}$: the set of additive principal ordinals, and $\phi_0$  is just defined as the ordering function of this set.
-
-\subsection{Additive principal ordinals}
-
-\index{maths}{Ordinal numbers!Additive principal ordinals}
-\index{hydras}{Library Schutte!Predicates!AP@AP (additive principal ordinals)}
-
-\begin{definition}
-A non-zero ordinal  $\alpha$ is said to be \emph{additive principal} if, for all  $\beta<\alpha$, $\beta+\alpha$ is equal to  $\alpha$.
-We call \texttt{AP} the set of additive principal ordinals.
-
-\end{definition}
-
-\noindent\emph{From Module \href{../theories/html/hydras.Schutte.AP.html}%
-{\texttt{Schutte.AP}}}
-
-\input{movies/snippets/AP/APDef}
-
-\subsection{The function \texttt{phi0}}
-
-Let us call  $\phi_0$ the ordering function of \texttt{AP}.
-In the mathematical text, we shall use indifferently the notations  $\omega^\alpha$ and$\phi_0(\alpha)$.
-
-\index{hydras}{Library Schutte!Functions!phi0}
-
-\input{movies/snippets/AP/phi0Def}
-
-\subsection{Omega-towers and the ordinal \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-Using $\phi_0$, we can define recursively the set of finite omega-towers.
-
-\input{movies/snippets/AP/omegaTower}
-
-\label{sect:epsilon0-as-limit}
-Then, the ordinal  $\epsilon_0$ is defined as the limit of the sequence of all finite towers (a kind of infinite tower).
-
-\input{movies/snippets/AP/epsilon0Def}
-
-The rest of our library \texttt{AP} is devoted to the proof of properties of additive principal ordinals, hence of the ordering function  $\phi0$ and the ordinal $\epsilon_0$ (which we could not express within the type \texttt{T1}).
-
-\subsection{Properties of the set  \texttt{AP}}
-
-The set of additive principal ordinals is not empty: it contains at least the ordinals  $1$ and  $\omega$.
-
-\inputsnippets{AP/APOne,AP/APOne_least_AP,AP/APOne_AP_omega,AP/APOne_omega_second_AP}
-
-The set  \texttt{AP} is  \emph{closed} under addition, and unbounded.
-\label{lemma:AP-plus-closed}
-
-\input{movies/snippets/AP/APPlusClosed}
-
-\input{movies/snippets/AP/APUnbounded}
-
-Finally, \texttt{AP} is (topologically) \emph{closed} and ordered by the segment of all countable ordinals.
-
-\index{hydras}{Library Schutte!Predicates!Closed}
-
-From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{~Schutte.Schutte\_basics}}
-
-\input{movies/snippets/Schutte_basics/ClosedDef}
-\input{movies/snippets/AP/APClosed}
-
-\subsubsection{Properties of the function \texorpdfstring{$\phi_0$}{phi0}}
-
-The ordering function $\phi_0$ of the set \texttt{AP} is defined on the full set $\mathbb{O}$ and is continuous (Schütte calls such a  function  \emph{normal}).
-
-\begin{Coqsrc}
-Theorem normal_phi0 : normal phi0 AP.
-\end{Coqsrc}
-
-The following properties come from  the definition of $\phi_0$ as the ordering function of \texttt{AP}.
-It may be interesting to compare these proofs with the computational ones described in Chapter ~\ref{chap:T1}.
-
-\input{movies/snippets/AP/APPhi0}
-
-\subsection{A last example}
-\label{Ex42-schutte}
-
-Let us prove again the equality $\omega+42+\omega^2= \omega^2$.Let us recall that $\omega^2$ is an abbreviation of $\phi_0(2)$,
-\emph{i.e} the third  additive principal ordinal.
-
-\inputsnippets{Schutte/Ex42a}
-
-Our proof is very different from the computational proof of Sect~\vref{Ex42-E0}.
-By definition of additive principal ordinals, 
-it suffices to prove the inequality $\omega+42< \phi_0(2)$.
-
-\inputsnippets{Schutte/Ex42b}
-
-Since the set \texttt{AP} of additive principals  is closed under addition
-(by Lemma \texttt{AP\_plus\_closed}, page~\pageref{lemma:AP-plus-closed}) , it suffices to prove the inequalities $\omega<\phi_0(2)$ and $42<\phi_0(2)$.
-
-\inputsnippets{Schutte/Ex42d, Schutte/Ex42c, Schutte/Ex42e}
-
-\section{More about \texorpdfstring{$\epsilon_0$}{\texttt{epsilon0}}}
-
-Let us recall that the limit ordinal  $\epsilon_0$ cannot be written within the type \texttt{T1}.
-Since we are now considering the set of all countable ordinals, we can now prove some properties of this ordinal.
-
-We prove the inequality  $\alpha<\omega^\alpha$ whenever $\alpha < \epsilon_0$.
-\emph{Note that this condition was implicit in Module
-\href{../theories/html/hydras.Epsilon0.T1.html\#lt_phi0}{Epsilon0.T1}.}
-
-\input{movies/snippets/AP/ltPhi0}
-
-The proof is as follows:
-\begin{enumerate}
-\item Since $\alpha<\epsilon_0$, consider the least $i$ such that $\alpha$ is strictly less than the omega-tower of height $i$.
-\item
-  \begin{itemize}
-  \item If $i=0$, then the result is trivial (because $\alpha=0$)
- \item  Otherwise let $i=j+1$; 
-          $\alpha$ is greater than or equal to the omega-tower of height $j$.
-         By monotony,  $\phi_0(\alpha)$ is greater than or equal to 
-        the omega-tower of height $j+1$, thus strictly greater than $\alpha$
-  \end{itemize}
- \end{enumerate}
-
-Moreover,  $\epsilon_0$ is the least ordinal $\alpha$ that verifies the equality 
-$\alpha = \omega^\alpha$, in other words, the least fixpoint of the function  $\phi_0$.
-
-\input{movies/snippets/AP/epsilon0Lfp}
-
-\section{Critical ordinals}
-
-\index{maths}{Ordinal numbers!Critical ordinals}
-\index{hydras}{Library Schutte!Predicates!Cr@Cr (critical ordinals)}
-
-For any  (countable) ordinal $\alpha$, the set $\textit{Cr}(\alpha)$ is inductively defined 
-as follows by Schütte (p.81 of~\cite{schutte}).
-
-\begin{quote}
-  \begin{itemize}
-  \item $\textit{Cr}(0)$ is the set \textit{AP} of additive principal ordinals.
-  \item If $0<\alpha$, then $\textit{Cr}(\alpha)$ is the intersection of all the sets of fixpoints of the $\textit{Cr}(\beta)$ for $\beta<\alpha$.
-  \end{itemize}
-\end{quote}
-
-This definition is translated in \coq{} in 
-Module \href{../theories/html/hydras.Schutte.Critical.html}%
-{\texttt{Schutte.Critical}}, as the least fixpoint of a functional.
-
-\input{movies/snippets/Critical/CrDef}
-
-Lets us denote by $\phi_\alpha$ the ordering function of the set $\textit{Cr}(\alpha)$ and by $A_\alpha$ its ordering segment.
-
-\input{movies/snippets/Critical/phiDef}
-
-\label{sect:phi-schutte}
-
-For instance,  we prove that $\textit{Cr}(0)$ is the set of additive principals and that $\epsilon_0$
-belongs to $\textit{Cr}(1)$.
-
-\inputsnippets{Critical/CrZeroAP,
-  Critical/epsilon0Cr1}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
- Prove that $\epsilon_0$ is the least element of $\textit{Cr}(1)$.
-\end{exercise}
-
-\subsection{A flavor of infinity}
-
-The family of the $\textit{Cr}(\alpha)$s is made of infinitely many unbounded (hence infinite) sets.
-Let us quote Lemma 5, p. 82  of~\cite{schutte}:
-\begin{quote}
-  For all $\alpha$, the set $\textit{Cr}(\alpha)$ is closed (for the least upper bound of non-empty countable sets) and unbounded.
-\end{quote}
-
-We prove this result by a transfinite induction on $\alpha$ of the conjunction of  both properties.
-
-\index{maths}{Transfinite induction}
-
-\input{movies/snippets/Critical/Lemma5}
-
-\section{Cantor normal form}
-
-The notion of Cantor normal form is defined for all countable ordinals.
-Nevertheless, note that, contrary to the implementation based on type \texttt{T1},
-the Cantor normal form of an ordinal $\alpha$ may contain $\alpha$ as a 
-sub-term\footnote{This would prevent us from trying to represent Cantor normal forms as finite trees (like in Sect.~\ref{sec:T1-inductive-def})}.
-
-\index{maths}{Ordinal numbers!Cantor normal form}
-\index{hydras}{Library Schutte!Predicates!is\_cnf\_of@is\_cnf\_of (to be a Cantor normal form of}
-
-We represent  Cantor normal forms as lists of ordinals.
-A  list $l$ is a Cantor normal form of a given ordinal $\alpha$ if it satisfies two conditions:
-
-\begin{itemize}
-\item The list  $l$ is sorted (in decreasing order) w.r.t. the order $\leq$
-\item The sum of all the  $\omega^{\beta_i}$ where the $\beta_i$ are the terms of $l$ (in this order) is equal to $\alpha$.
-\end{itemize}
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Schutte.CNF.html\#cnf_t}%
-{\texttt{Schutte.CNF}}}
-
-\input{movies/snippets/CNF/Defs}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction on $\alpha$, we prove that every countable ordinal $\alpha$ 
- has at least a Cantor normal form.
-
-\input{movies/snippets/CNF/cnfExists}
-
-By structural induction on lists, we prove that this normal form is unique.
-
-\inputsnippets{CNF/cnfUnicity,CNF/cnfExUnique}
-
-Finally, the following two lemmas relate  $\epsilon_0$ with Cantor normal forms.
-
-If $\alpha<\epsilon_0$, then the Cantor normal form of $\alpha$ is made of ordinals strictly less than $\alpha$.
-
-\input{movies/snippets/CNF/cnfLtEpsilon0}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Please consider the following statement :
-
-\begin{Coqsrc}
-Lemma cnf_lt_epsilon0_iff : 
- forall l alpha, 
-   is_cnf_of alpha l ->  
-   (alpha < epsilon0 <->  Forall (fun beta =>  beta < alpha) l).
-\end{Coqsrc}
-
-Is it true ?
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/schutte_cnf_counter_example.v}{exercises/ordinals/schutte\_cnf\_counter\_example.v}.}
-\end{exercise}
-
-Finally, the Cantor normal form of $\epsilon_0$ is just $\omega^{\epsilon_0}$.
-
-\input{movies/snippets/CNF/cnfOfEpsilon0}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Implement pages 82 to 85 of~\cite{schutte} (critical, strongly critical, maximal critical ordinals, Feferman's ordinal $\Gamma_0$).
-\end{project}
-
-\begin{remark}
-The sub-directory
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/Gamma0}{theories/oridinals/Gamma0} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form.
-\end{remark}
-
-\section{An embedding of \texttt{T1} into \texttt{Ord}}
-
-Our library 
-\href{../theories/html/hydras.Schutte.Correctness_E0.html}%
-{\texttt{Schutte.Correctness\_E0}} establishes the link between two very different modelizations of ordinal numbers. In other words, it ``validates'' a data structure in terms of
-a classical mathematical discourse considered as a model. 
-First, we define a function from \texttt{T1} into  \texttt{Ord} by structural recursion.
-
-\input{movies/snippets/Correctness_E0/injectDef}
-
-This function enjoys good commutation properties with respect to the main operations which
-allow us to build Cantor normal forms.
-
-\input{movies/snippets/Correctness_E0/commutationLemmas}
-\input{movies/snippets/Correctness_E0/injectPlus}
-\input{movies/snippets/Correctness_E0/injectMultFinR}
-
-% \begin{Coqsrc}
-% Theorem inject_mono (beta gamma : T1) :
-%   T1.lt  beta gamma -> 
-%   T1.nf beta -> T1.nf gamma -> 
-%   inject beta < inject gamma.
-
-% Theorem inject_injective (beta gamma : T1) : nf beta -> nf gamma ->
-%   inject beta = inject gamma -> beta = gamma.
-% \end{Coqsrc}
-
-Finally, we prove that \texttt{inject} is a bijection from the set of all terms of \texttt{T1} in normal form to the set 
-\texttt{members epsilon0} of the elements of \texttt{Ord} strictly less than  $\epsilon_0$.
-
-\input{movies/snippets/Correctness_E0/injectLtEpsilon0}
-\input{movies/snippets/Correctness_E0/embedding}
-
-\subsection{Remarks}
-Let us recall that the library \href{../theories/html/hydras.Schutte.Schutte.html}%
-{\texttt{Schutte}} depends on five \emph{axioms} and lies explicitly in the  
-framework of classical logic with a weak version of the axiom of choice
-(please look at the documentation of
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.ChoiceFacts.html}{\texttt{Coq.Logic.ChoiceFacts}}).
-Nevertheless, the other modules:
-\href{../theories/html/hydras.Epsilon0.Epsilon0.html}%
-{\texttt{Epsilon0}},
-\href{../theories/html/hydras.Hydra.Hydra.html}%
-{\texttt{Hydra}}, et 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}%
-{\texttt{Gamma0}}
-do not import any axioms and are really constructive.
-
-\index{hydras}{Projects}
-\begin{project}
-There is no construction of ordinal multiplication in~\cite{schutte}. 
-It would be interesting to derive this operation from Schütte's axioms,
-and prove its consistence with multiplication in ordinal notations for 
-$\epsilon_0$ and $\Gamma_0$.
-\end{project}
-
-\section{Related work}
-
-In~\cite{grimm:hal-00911710}, José Grimm establishes the consistency between our ordinal notations \texttt{T1} and \texttt{T2} (Veblen normal form) and his implementation
-of ordinal numbers after Bourbaki's set theory.
-
-The Gaia project ~\url{https://github.com/coq-community/gaia} maintains Grimm's  theory of ordinals as part of coq-community on GitHub. Integration
-of the present ordinal theory with Gaia, i.e., relating the different notions of ordinals
-and transferring relevant results, is an interesting project.
-First experiments in that direction are developped in
-the \href{https://github.com/coq-community/hydra-battles/blob/master/theories/gaia/}{theories/gaia/} directory.
-
-\chapter{The Ordinal \texorpdfstring{$\Gamma_0$}{Gamma0} (first draft)}
-
-\emph{This chapter and the files it presents are still very incomplete, considering the impressive properties of $\Gamma_0$~\cite{Gallier91}.
-We hope to add new material soon, and accept contributions!}
-
-\begin{todo}
-  Build a T2Bridge! (note: Gamma0's wellfoundedness is missing in \gaia (?).
-\end{todo}
-
-\section{Introduction}
-We present a notation system for the ordinal $\Gamma_0$, following Chapter V, Section 14 of~\cite{schutte}: ``A notation system for the ordinals $<\Gamma_0$''.
-We try to be as close as possible to Schütte's text and usual practices of \coq{} developments.
-
-The ordinal $\Gamma_0$ is defined in Section 13 of ~\cite{schutte} as the least \emph{strongly critical ordinal}. It is widely known as the \emph{Feferman-Schütte ordinal}.
-
-Section V, 13 of~\cite{schutte} defines \emph{strongly critical} and
-\emph{maximal $\alpha$-critical} ordinals:
-
-\begin{itemize}
-\item $\alpha$ is strongly critical if
-$\alpha$ is $\alpha$-critical,
-\item $\gamma$ is maximal $\alpha$-critical if $\gamma$ is $\alpha$-critical, and, for all $\xi>\alpha$, $\gamma$ is not $\xi$-critical.
-
-\end{itemize}
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Schutte.Critical.html\#strongly_critical}%
-{\texttt{Schutte.Critical}}}
-
-\input{movies/snippets/Critical/Gamma0Def}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that a (countable)  ordinal $\alpha$ is strongly critical iff 
-$\phi_\alpha(0)=\alpha$ (Theorem 13.13 of~\cite{schutte} ). 
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that the set of strongly critical ordinals is unbounded and closed (Theorem 13.14 of~\cite{schutte} ).
-Thus this set is not empty,  hence has a least element.
-Otherwise, the definition of $\Gamma_0$ above would be useless.
-\end{project}
-
-In the present version of this development, we  only study $\Gamma_0$ as a notation system, much more powerful than the ordinal notation for $\epsilon_0$.
-
-%\index{Projects}
-%
-%\begin{project}
-% Schûtte's section 13 of~\cite{schutte} contain several definitions and lemmas which give another view on $\Gamma_0$.  We leave it as a project to implement them in \coq{}.  
-%\end{project}
-
-\section{The type \texttt{T2} of ordinal terms}
-
-The notation system for ordinals less than $\gamma_0$ comes from the following theorem of~\cite{schutte}, where $\psi\,\alpha$ is the ordering function 
-of the set of maximal $\alpha$-critical ordinals.
-
-\index{hydras}{Library Gamma0!Types!T2}
-
-\begin{quote}
-  Any ordinal $\not= 0$ which is not strongly critical can be expressed in terms of $+$ and $\psi$.
-\end{quote}
-
-\index{hydras}{Projects}
-\begin{project}
-This theorem is not formally proved in this development yet. It should be!
-\end{project}
-
-Like in Chapter~\ref{chap:T1}, we define an inductive type with two constructors, one for $0$, the other for the construction $\psi(\alpha,\beta)\times(n+1)+\gamma$, adapting a Manolios-Vroon-like notation~\cite{Manolios2005} to
-\emph{Veblen normal forms}.
-\label{types:T2}
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T2}%
-{\texttt{Gamma0.T2}}}
-
-\inputsnippets{T2/T2Def, T2/psiDef}
-
-Like in chapter~\ref{chap:T1}, we get familiar with the type \texttt{T2} by recognizing simple constructs like finite ordinals, $\omega$, etc., as inhabitants of \texttt{T2}.
-
-\input{movies/snippets/T2/Notations}
-
-\section{A strict order on T2}
-
-Let us define a strict order on type \texttt{T2}. The following definition is 
-an adaptation of Schütte's, taking into account the multiplications by a natural number (inspired by~\cite{Manolios2005}, and also present in \texttt{T1}).
-
-\label{sect:t2-lt-def}
-
-\input{movies/snippets/T2/ltDef}
-
-Seven constructors! In order to get accustomed with this definition, let us look at a small set of examples, covering all the constructors of \texttt{lt}.
-
-\input{movies/snippets/T2/ltExamples}
-
-\index{hydras}{Projects}
-\begin{project}
-Write a tactic that solves automatically goals of the form (\texttt{$\alpha$ t2< $\beta$}), where $\alpha$ and $\beta$ are closed terms of type \texttt{T2}.
-\end{project}
-
-\section{Veblen normal form}
-\begin{definition}
-  A term of the form $\psi(\alpha_1,\beta_1)\times n_1+ \psi(\alpha_2,\beta_2)\times n_2+\dots+\psi(\alpha_k,\beta_k)\times n_k$ is said to be in
- \emph{[Veblen] normal form} if for every $i<n$, $\psi(\alpha_i,\beta_i)<\psi(\alpha_{i+1},\beta_{i+1})$, all the $\alpha_i$ and $\beta_i$ are in normal form, and all the $n_i$ are strictly positive integers.
-\end{definition}
-
-\input{movies/snippets/T2/nfDef}
-
-Let us look at some positive examples (we have to prove some inversion lemmas before proving counter-examples).
-
-\input{movies/snippets/T2/nfExamples}
-
-\subsection{Length of a term}
-
-The notion of \emph{term length} is introduced by Schütte as a helper for proving (at least) the \emph{trichotomy} property and transitivity of the strict order \texttt{lt} on \texttt{T2}.
-These properties are proved by induction on length.
-
-\subsection{Trichotomy}
-
-\emph{Trichotomy} is another name for the well-known property of decidable total ordering (like Standard Library's \texttt{Compare\_dec.lt\_eq\_lt\_dec}).
-
-We first prove by induction on $l$ the following lemma:
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0\#tricho_aux}%
-{\texttt{Gamma0.Gamma0}}}
-
-\input{movies/snippets/Gamma0/tricho}
-
-\input{movies/snippets/Gamma0/compare}
-
-With the help of \texttt{compare}, we get a boolean version of \texttt{nf}
-(being in Veblen normal form).
-
-\input{movies/snippets/Gamma0/nfb}
-
-\section{Main functions on \texttt{T2}}
-
-\subsection{Successor}
-The successor function is defined by structural recursion.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#succ}%
-{\texttt{Gamma0.T2}}}
-
-\input{movies/snippets/T2/succDef}
-
-\subsection{Addition}
-
-Like for Cantor normal forms (see Sect.~\ref{sect:infix-plus-T1}),  the definition of addition in \texttt{T2}  requires comparison between ordinal terms.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0T2.html\#succ}%
-  {\texttt{Gamma0.Gamma0}}}
-
-\input{movies/snippets/Gamma0/plusDef}
-
-\subsection{The Veblen function \texorpdfstring{$\phi$}{\texttt{phi}}}
-
-The enumeration function of critical ordinals, presented in Sect.~\vref{sect:phi-schutte}, is recursively defined in type \texttt{T2}.
-
-\input{movies/snippets/Gamma0/phiDef}
-
-Despite its complexity, the function \texttt{phi} is well adapted to proofs by simplification or computation.
-
-The relation between the constructor $\psi$ and the function $\phi$ is
-studied in~\cite{schutte}, and partially implemented in this development.
-\emph{Please contribute!}
-
-For instance, the following theorem states that, if $\gamma$ is the sum of a limit ordinal $\beta$ and a finite ordinal $n$, and $\beta$ is a fixpoint of
-$\phi(\alpha)$, then $\psi(\alpha,\gamma)=\phi_\alpha(\gamma+1)$.
-
-\input{movies/snippets/Gamma0/phiPsi}
-\input{movies/snippets/Gamma0/Ex9}
-
-On the other hand, $\phi$ can be expressed in terms of $\psi$.
-
-\input{movies/snippets/Gamma0/phiOfPsi}
-\input{movies/snippets/Gamma0/Ex10}
-
-\index{hydras}{Projects}
-\begin{project}
-Please study a way to pretty print ordinal terms in Veblen normal form (see Section~\vref{sect:ppT1}).
-\end{project}
-
-\section{An ordinal notation for \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}}
-
-In order to consider type \texttt{T2} as an ordinal notation, we have to build an instance of class \texttt{ON} (See Definition page~\pageref{types:ON}).
-
-First, we define a type that contains only terms in Veblen normal form, and redefine \texttt{lt} and \texttt{compare} by delegation (see for comparison the construction of type \texttt{E0} in Sect.~\vref{sect:E0-def}).
-
-\input{movies/snippets/Gamma0/G0a}
-\input{movies/snippets/Gamma0/G0b}
-
-Then, we buils an instance of class \texttt{ON}. 
-function \texttt{compare} is correct.
-
-\input{movies/snippets/Gamma0/ltSto}
-\input{movies/snippets/Gamma0/ltWf}
-\input{movies/snippets/Gamma0/ONGamma0}
-\input{movies/snippets/Gamma0/G0z}
-
-\begin{remark}
-The proof of \texttt{lt\_wf} has been written by \'Evelyne Contejean, using her library on the recursive path ordering (see also remark~\vref{remark:a3pat}).
-\end{remark}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that \texttt{Epsilon0} (page~\pageref{instance-epsilon0})
-is a sub-notation system of \texttt{Gamma0}.
-
-Prove that the implementations of \texttt{succ}, \texttt{+}, $\phi_0$, etc.
-are compatible in both notation systems.
-
-Note that a function \texttt{T1\_inj} from \texttt{T1} to \texttt{T2} has already been defined. It may help to complete the task.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T1_to_T2}%
-{\texttt{Gamma0.T2}}}
-
-\input{movies/snippets/T2/T1ToT2}
-
-\end{project}
-
-\begin{project}
-Prove that the notation system \texttt{Gamma0} is a correct implementation 
-of the segment $[0,\Gamma_0)$ of the set of countable ordinals.
-\end{project}

### 2022-05-08 doc/part-hydras-intro.tex
reorganization of tex files

--- 2022-04-11 doc/part-hydras.tex
@@
 \end{itemize}
-
-\chapter{Hydras and hydra games}
-
-\label{sec:orgheadline91}
-\label{chapter:hydras}
-
-This chapter is dedicated to the representation of hydras and rules of the hydra game in \coq's specification language:~\gallina.
-
-Technically, a \emph{hydra} is just a finite ordered tree, each node of which 
-has any number of sons. Contrary to the computer science tradition, we display the hydras 
-with the heads up and the foot (i.e., the root of the tree) down.
-Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the 
-module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}). 
-\emph{For a less formal description of hydras, please see 
-\url{https://www.smbc-comics.com/comic/hydra}.}
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {$\bullet$};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H4) at (4,2) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N3) to [bend left= 10]  (H1) ;
-\draw (N3) to [bend right= 16] (H2);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (foot) to [bend right = 10] (H4) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-\caption{The hydra \texttt{Hy} \label{fig:Hy}}
-\end{figure}
-
-We use a specific vocabulary for talking about hydras. Table~\ref{tab:hyd2tree} shows the correspondence between our terminology and the usual vocabulary for trees in computer science.
-
-\begin{figure}[h]
-  \centering
-  \begin{tabular}{ll}
-Hydras & Finite rooted trees\\
-\hline
-foot & root\\
-head & leaf\\
-node & node\\
-segment  & (directed) edge \\
-sub-hydra & subtree\\
-daughter & immediate subtree\\
-\end{tabular}
-  \caption{Translation from hydras to trees}
-  \label{tab:hyd2tree}
-\end{figure}
-
-The hydra \texttt{Hy} has a \emph{foot} (below), five \emph{heads}, and eight \emph{segments}. 
-We leave it to the reader to define various parameters such as the height, the size, the highest arity (number of sons of a node) of a hydra.
-In our example, these parameters have the respective values $4$, $9$ and $3$.
-
-\subsection{The rules of the game}
-
-\label{sec:orgheadline44}
-\label{sect:replication-def}
-
-A \emph{hydra battle} is a fight between Hercules and the Hydra. 
-More formally, a  battle is a sequence of \emph{rounds}.
-At each round:
-\begin{itemize}
-\item If the hydra is composed of just one head, the battle is finished
-and  Hercules is the winner.
-\item Otherwise, Hercules chops off \emph{one} head of the hydra,
-
-\begin{itemize}
-\item If the head is at distance 1 from the foot, the head is just lost by the hydra, with no more reaction.
-\item Otherwise, let us denote by \(r\) the node that was at distance \(2\) from 
-the removed head in the direction of the foot,  and consider the  sub-hydra \(h'\) of \(h\), whose  root is \(r\) \footnote{$h'$ will be called ``the wounded part of the hydra'' in the subsequent text.
-In Figures~\vref{fig:Hy2} and ~\vref{fig:Hy4}, this sub-hydra  is displayed in red.}.
-Let $n$ be some natural number.
-Then $h'$ is replaced by  $n+1$ of copies of \(h'\) which share the same root $r$.
- The \emph{replication factor} $n$ may be different (and generally is)   at each round of the fight.
-It may be chosen by the hydra, according to its strategy, or imposed by some 
-particular rule.
-In many presentations of hydra battles, this number is increased by $1$ at each round.
-In the following presentation, we will also consider battles where the hydra is free to choose its ~replication factor at each round of the battle\footnote{Let us recall that, if the chopped-off head was at distance 1 from the foot, the replication factor is meaningless.}.
-\end{itemize}
-\end{itemize}
-
-Note that the description given in~\cite{KP82} of the replication process in hydra battles is also  semi-formal.
-
-\label{original-rules}
-
-\begin{quote}
-  ``From the node that used to be attached to the head which was just chopped off, traverse one 
-segment towards the root until the next node is reached. From this node sprout $n$ replicas of 
-that part of the hydra (after decapitation) which is ``above'' the segment just traversed, i.e., those 
-nodes and segments from which, in order to reach the root, this segment would have to be 
-traversed. If the head just chopped off had the root of its nodes, no new head is grown. ''
-\end{quote}
-
-Moreover, we note that this description is in \emph{imperative} terms.
-In order to formally study the properties of hydra battles, we prefer to use a mathematical vocabulary, i.e., graphs, relations, functions, etc.
-Thus, the replication process will be represented as a binary relation on a data type \texttt{Hydra},
-linking the state of the hydra \emph{before} and \emph{after} the transformation.
-A battle will thus be represented as a sequence of terms of type \texttt{Hydra}, respecting the rules of the game.
-
-\subsection{Example}
-Let us start a battle between Hercules and the hydra \texttt{Hy} of Fig.~\ref{fig:Hy}.
-
-At the first round, Hercules chooses to chop off the rightmost head of \texttt{Hy}.
-Since this head is near the floor, the hydra simply loses this head. Let us call 
- \texttt{Hy'} the resulting state of the hydra, represented in Fig.~\vref{fig:Hy-prime}.
-
-Next, assume Hercules chooses to chop off one of the two highest heads of \texttt{Hy'}, for instance the rightmost one.
-Fig.~\vref{fig:Hy2} represents the broken segment in dashed lines, and the part that will be replicated in red.
-Assume also that the hydra decides to add 4 copies of the red part\footnote{In other words, the replication factor at this round is equal to $4$.}.
-We obtain a new state \texttt{Hy''} depicted in Fig.~\ref{fig:Hy3}.
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {$\bullet$};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-%\node (H4) at (6,0) {$\Xey[2][lightgray]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3) ;
-\draw (N3) to [bend left= 10]  (H1) ;
-\draw (N3) to [bend right= 16] (H2);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-
-\caption{\texttt{Hy'}: the state  of \texttt{Hy} after one round \label{fig:Hy-prime}}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4)  {{\color{lightred}$\bullet$}};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Sey[2][lightred]$};
-%\node (H2) at (5,0) {$\Xey[2][lightgray]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\node (ex) at (5,8) {};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw  (N2) -- (N3)[draw=lightred];
-\draw (N3) to   [bend left= 10](H1) [draw=lightred];
-\draw [dashed] (N3) to [bend left= 10](ex);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-\caption{A second beheading}
-\label{fig:Hy2}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H11) at (2,8) {$\Smiley[2][vertfluo]$};
-\node (H12) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H13) at (6,8) {$\Smiley[2][vertfluo]$};
-\node (H14) at (8,8) {$\Smiley[2][vertfluo]$};
-
-\node (N3) at (1,6) {$\bullet$};
-\node (N31) at (2,6) {$\bullet$};
-\node (N32) at (3,6) {$\bullet$};
-\node (N33) at (4,6) {$\bullet$};
-\node (N34) at (5,6) {$\bullet$};
-
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N2) -- (N31);
-\draw (N2) -- (N32);
-\draw (N2) -- (N33);
-\draw (N2) -- (N34);
-\draw (N3) to   [bend left= 10](H1) ;
-\draw (N31) to   [bend left= 10](H11) ;
-\draw (N32) to   [bend left= 10](H12) ;
-\draw (N33) to   [bend left= 10](H13) ;
-\draw (N34) to   [bend left= 10](H14) ;
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend left= 10]  (H5) ;
-\end{tikzpicture}
-\caption{\texttt{Hy''}: the state of \texttt{Hy} after two rounds \label{fig:Hy3}}
-\end{figure}
-
-Figs.~\ref{fig:Hy4} and~\vref{fig:Hy5} represent a possible third round of the battle, with a replication factor equal to $2$. Let us call \texttt{Hy'''} the state of the hydra after that third round.
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0)  {{\color{lightred}$\bullet$}};
-\node (N1) at (2,2) {{\color{lightred}$\bullet$}};
-\node (N2) at (2,4) {{\color{lightred}$\bullet$}};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (exN4) at (4,4) {};
-\node (H1) at (0,8) {$\Sey[2][lightred]$};
-\node (H11) at (2,8) {$\Sey[2][lightred]$};
-\node (H12) at (4,8) {$\Sey[2][lightred]$};
-\node (H13) at (6,8) {$\Sey[2][lightred]$};
-\node (H14) at (8,8) {$\Sey[2][lightred]$};
-
-\node (N3) at (1,6) {{\color{lightred}$\bullet$}};
-\node (N31) at (2,6) {{\color{lightred}$\bullet$}};
-\node (N32) at (3,6) {{\color{lightred}$\bullet$}};
-\node (N33) at (4,6) {{\color{lightred}$\bullet$}};
-\node (N34) at (5,6) {{\color{lightred}$\bullet$}};
-
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-%\node (H5) at (4,0) {$\Xey[2][lightgray]$};
-\draw (foot) -- (N1)[very thick,draw=lightred] ;
-\draw (N1) -- (N2)[draw=lightred];
-\draw (N2) -- (N3)[draw=lightred];
-\draw (N2) -- (N31)[draw=lightred];
-\draw (N2) -- (N32)[draw=lightred];
-\draw (N2) -- (N33)[draw=lightred];
-\draw (N2) -- (N34)[draw=lightred];
-\draw (N3) to   [bend left= 10](H1) [draw=lightred];
-\draw (N31) to   [bend left= 10](H11) [draw=lightred];
-\draw (N32) to   [bend left= 10](H12) [draw=lightred];
-\draw (N33) to   [bend left= 10](H13) [draw=lightred];
-\draw (N34) to   [bend left= 10](H14) [draw=lightred];
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw [dashed] (N1) to  [bend left= 10](exN4);
-\end{tikzpicture}
-\caption{A third beheading (wounded part in red) \label{fig:Hy4}}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-
-\node (foot) at (10,0) {$\bullet$};
-
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H1) at (0,8) {$\Smiley[1][vertfluo]$};
-\node (H11) at (2,8) {$\Smiley[1][vertfluo]$};
-\node (H12) at (4,8) {$\Smiley[1][vertfluo]$};
-\node (H13) at (6,8) {$\Smiley[1][vertfluo]$};
-\node (H14) at (8,8) {$\Smiley[1][vertfluo]$};
-
-\node (N3) at (1,6) {$\bullet$};
-\node (N31) at (2,6) {$\bullet$};
-\node (N32) at (3,6) {$\bullet$};
-\node (N33) at (4,6) {$\bullet$};
-\node (N34) at (5,6) {$\bullet$};
-
-\node (H0) at (-3,3) {$\Smiley[1][vertfluo]$};
-
-\draw (foot) to [bend left=10] (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N2) -- (N31);
-\draw (N2) -- (N32);
-\draw (N2) -- (N33);
-\draw (N2) -- (N34);
-\draw (N3) to   [bend left= 10](H1) ;
-\draw (N31) to   [bend left= 10](H11) ;
-\draw (N32) to   [bend left= 10](H12) ;
-\draw (N33) to   [bend left= 10](H13) ;
-\draw (N34) to   [bend left= 10](H14) ;
-\draw (foot) to [bend left = 15]  (H0) ;
-
-% second copy 
-\node (N01) at (12,2) {$\bullet$};
-\node (N02) at (12,4) {$\bullet$};
-\node (N03) at (12,6) {{\color{lightred}$\bullet$}};
-\node (H001) at (10,8) {$\Smiley[1][vertfluo]$};
-\node (H0011) at (12,8) {$\Smiley[1][vertfluo]$};
-\node (H0012) at (14,8) {$\Smiley[1][vertfluo]$};
-\node (H0013) at (16,8) {$\Smiley[1][vertfluo]$};
-\node (H0014) at (18,8) {$\Smiley[1][vertfluo]$};
-
-\node (N03) at (11,6) {$\bullet$};
-\node (N031) at (12,6) {$\bullet$};
-\node (N032) at (13,6) {$\bullet$};
-\node (N033) at (14,6) {$\bullet$};
-\node (N034) at (15,6) {$\bullet$};
-
-\draw (foot) -- (N01)[very thick] ;
-\draw (N01) -- (N02);
-\draw (N02) -- (N03);
-\draw (N02) -- (N031);
-\draw (N02) -- (N032);
-\draw (N02) -- (N033);
-\draw (N02) -- (N034);
-\draw (N03) to   [bend left= 10](H001) ;
-\draw (N031) to   [bend left= 10](H0011) ;
-\draw (N032) to   [bend left= 10](H0012) ;
-\draw (N033) to   [bend left= 10](H0013) ;
-\draw (N034) to   [bend left= 10](H0014) ;
-
-% third copy 
-\node (N001) at (22,2) {$\bullet$};
-\node (N002) at (22,4) {$\bullet$};
-\node (N003) at (22,6) {{\color{lightred}$\bullet$}};
-\node (H001) at (20,8) {$\Smiley[1][vertfluo]$};
-\node (H0011) at (22,8) {$\Smiley[1][vertfluo]$};
-\node (H0012) at (24,8) {$\Smiley[1][vertfluo]$};
-\node (H0013) at (26,8) {$\Smiley[1][vertfluo]$};
-\node (H0014) at (28,8) {$\Smiley[1][vertfluo]$};
-
-\node (N003) at (21,6) {$\bullet$};
-\node (N0031) at (22,6) {$\bullet$};
-\node (N0032) at (23,6) {$\bullet$};
-\node (N0033) at (24,6) {$\bullet$};
-\node (N0034) at (25,6) {$\bullet$};
-
-\draw (foot) -- (N001)[very thick] ;
-\draw (N001) -- (N002);
-\draw (N002) -- (N003);
-\draw (N002) -- (N0031);
-\draw (N002) -- (N0032);
-\draw (N002) -- (N0033);
-\draw (N002) -- (N0034);
-\draw (N003) to   [bend left= 10](H001) ;
-\draw (N0031) to   [bend left= 10](H0011) ;
-\draw (N0032) to   [bend left= 10](H0012) ;
-\draw (N0033) to   [bend left= 10](H0013) ;
-\draw (N0034) to   [bend left= 10](H0014) ;
-\end{tikzpicture}
-\caption{The configuration \texttt{Hy'''} of \texttt{Hy} \label{fig:Hy5}}
-\end{figure}
-\FloatBarrier
-
-We leave it to the reader  to guess the following  rounds of the battle \dots
- % Please keep in mind that, in this 
-% the hydra is free to chose any number of replications at each time, whereas
-% Hercules chops only one head per round.
-
-% Let us precise that, in this game, Hercules wins if the hydra is eventually reduced 
-% to a single head. 
-% We know from~\cite{KP82} that, whichever the initial configuration of the
-% hydra, and the strategies of both players, Hercules eventually wins. The 
-% aforementionned paper shows also that there do not exist any \emph{simple} proof of this result.
-
-\section{Hydras and their representation in \emph{Coq}}
-\label{sec:orgheadline48}
-
-\index{hydras}{Library Hydra!Types!Hydra}
-\index{hydras}{Library Hydra!Types!Hydrae}
-
-In order to describe trees where each node can have an arbitrary (but finite) number of sons, it is usual to define a type where each node carries a \emph{forest}, \emph{i.e} a list of trees
-(see for instance Chapter 14, pages 400-406 of \cite{BC04}, also available as~\cite{BC04ch14}).
-
-For this purpose, we define two mutual \emph{ad-hoc}  inductive types, where \texttt{Hydra} is the main type, and \texttt{Hydrae} is a helper for describing finite sequences of hydras.
-\label{types:Hydra}
-\label{types:Hydrae}
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hydra}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/HydraDef}
-
-%\index{To do}
-\index{hydras}{Projects}
-
-\begin{project}
-Look for an existing library on trees with nodes of arbitrary arity, in order to replace  this ad-hoc type with something more generic.
-\end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-
- Another very similar representation could use the \texttt{list} type family instead of the specific 
-type \texttt{Hydrae}:
-
-\input{movies/snippets/Hydra_Definitions/HydraAlt}
-
-Using this representation, re-define all the constructions of this chapter.
-You will probably have to use patterns described for instance in~\cite{BC04} or the archives of the \coq communication channels (please consult~\url{https://coq.inria.fr/community.html}).
-
-\end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-The type \texttt{Hydra} above describes hydras as \emph{plane trees}, i.e., as drawn on a sheet of paper or computer screen. Thus, hydras are \emph{oriented},
-and it is appropriate to consider a \emph{leftmost} or \emph{rightmost} head of
-the beast. It could be interesting to consider another representation, in which
-every non-leaf node has a \emph{multi-set} -- not an ordered list -- of daughters.
-\end{project}
-
-\subsubsection{Abbreviations}
-
-We provide several notations for hydra patterns  which occur often in our developments.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#head}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/headsEtc}
-
-For instance, the hydra \texttt{Hy}  of Figure~\vref{fig:Hy} is defined in \emph{Gallina} as follows:
-
-\vspace{4mm}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html\#Hy}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/Hy}
-
-Hydras quite frequently contain  multiple adjacent  copies of the same subtree. The following functions
-will help us to describe and reason about replications in hydra battles.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#hcons_mult}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hconsMult}
-
-\vspace{4mm}
-
-For instance, the hydra \texttt{Hy''} of Fig~\vref{fig:Hy3}  can be defined in \coq{} as follows:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/HySecond}
-
-\subsubsection{Recursive functions on type \texttt{Hydra}}
-\label{sec:orgheadline41}
-\label{sec:hsize-def}
-
-In order to  define a recursive function over the type \texttt{Hydra}, one has to consider the three constructors 
-\texttt{node}, \texttt{hnil} and \texttt{hcons} of the mutually inductive types \texttt{Hydra} and \texttt{Hydrae}. 
-Let us define for instance the function which  computes the number of nodes of any hydra:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hsize}
-
-\input{movies/snippets/Hydra_Examples/HySize}
-
-Likewise, the \emph{height} (maximum distance between the foot and a head) 
-is defined by mutual recursion:
-
-\input{movies/snippets/Hydra_Definitions/height}
-
-\input{movies/snippets/Hydra_Examples/HyHeight}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Define a function \texttt{max\_degree: Hydra $\arrow$ nat} which  returns the highest degree of a node in any hydra.
-For instance, the evaluation of the term \texttt{(max\_degree Hy)} should return $3$.
-\end{exercise}
-
-\subsection{Induction principles for hydras}
-\label{sec:orgheadline42}
-
-In this section, we show how induction principles are used to prove properties on the type 
-\texttt{Hydra}. Let us consider for instance the following statement:
-\begin{quote}
-  `` The height of any hydra is strictly less than its size. ''
-\end{quote}
-
-\subsubsection{A failed attempt}
-
-One may try to use the default tactic of proof by induction, which corresponds to an application of the automatically  generated  induction principle for  type \texttt{Hydra}:
-
-\input{movies/snippets/Hydra_Examples/HydraInd}
-
-Let us start a simple proof by induction.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/BadInductiona}
-
-We might be tempted to do an induction on the sequence \texttt{s}:
-
-\input{movies/snippets/Hydra_Examples/BadInductionb}
-
-The first subgoal is trivial.
-
-\input{movies/snippets/Hydra_Examples/BadInductionc}
-
-Let us look at the second subgoal of the induction.
-
-\input{movies/snippets/Hydra_Examples/BadInductiond}
-
-We notice that this subgoal does not contain any hypothesis
-on the height and size of the hydra \texttt{h}. So, it looks hard to prove the conclusion. Let's stop.
-
-\input{movies/snippets/Hydra_Examples/BadInductione}
-
-\subsubsection{A Principle of mutual induction}
-In order to get an appropriate induction scheme for the types 
-\texttt{Hydra} and \texttt{Hydrae}, we can use  \coq{}'s  command \texttt{Scheme}.
-
-\index{coq}{Mutually inductive types}
-%\index{coq}{Commands!Scheme}
-
-\input{movies/snippets/Hydra_Definitions/HydraRect2}
-
-\input{movies/snippets/Hydra_Examples/HydraRect2Check}
-
-\subsubsection{A Correct proof}
-
-Let us now use \texttt{Hydra\_rect2} for proving that the height of any hydra is strictly less than its size.
-Using this scheme requires an auxiliary predicate, called \texttt{P0} in \texttt{Hydra\_rect2}'s statement.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hForall}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-\input{movies/snippets/Hydra_Examples/heightLtSizea}
-
-The first subgoal is easily solvable, using some arithmetic.
-The second and third ones are
-almost trivial. We let the reader look at the source. 
-\input{movies/snippets/Hydra_Examples/heightLtSizez}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-It happens very often that, in the proof of  a proposition of the form 
-(\texttt{$\forall\,$ h:Hydra, $P$ h}), the predicate \texttt{P0}
-is  (\texttt{h\_forall $P$}).
-Design a tactic for induction on hydras that frees the user from binding explicitly \texttt{P0},  and solves trivial subgoals.
-Apply it for writing  a shorter proof of \texttt{height\_lt\_size}.
-\end{exercise}
-
-\section{Relational description of hydra battles}
-
-In this section, we represent the rules of hydra battles as a binary relation associated with
-a \emph{round}, i.e., an interaction composed of the two following actions:
-\begin{enumerate}
-\item Hercules chops off one head of the hydra.
-\item Then, the  hydra replicates the wounded part (if the head is at distance $\geq 2$ from the foot).
-\end{enumerate}
-The relation associated with each round of the battle is parameterized  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot,
-but present for consistency's sake).
-
-In our description,  we will apply the following naming convention: if $h$ represents the configuration of the hydra before a round, then the configuration of $h$ after this round will be called $h'$.
- Thus, we are going to define a proposition  (\texttt{round\_n $n\;h\;h'$})  whose intended meaning will be `` the hydra $h$  is transformed into $h'$  in a single round of a battle, with the expected replication factor $n$ ''.
-
-Since the replication of parts of the hydra depends on the distance of the chopped head from  the foot, we  decompose our description into two main  cases, under the form of a bunch of [mutually] inductive predicates over the types \texttt{Hydra} and \texttt{Hydrae}.
-
-The mutually exclusive cases we consider are the following:
-\begin{itemize}
-\item \textbf{R1}: The chopped off head was at distance 1 from the foot.
-\item \textbf{R2}: The chopped off head was at a distance greater than or equal to  $2$ from the foot.
-\end{itemize}
-
-\subsection{Chopping off a head at distance 1 from the foot (relation  R1)}
-
-If Hercules chops off a head near the floor, there is no replication at all. We use an auxiliary 
-predicate \texttt{S0}, associated with the removing of one head from a sequence of hydras.
-
-\vspace{4pt}\emph{From Module\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/S0Def}
-\input{movies/snippets/Hydra_Definitions/R1Def}
-
-\subsubsection{Example}
-\label{sec:orgheadline45}
-
-Let us represent in \coq{}   the transformation of the hydra of Fig.~\vref{fig:Hy} into
-the configuration represented in Fig.~\ref{fig:Hy-prime}.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/Hy1}
-
-\subsection{Chopping off a head at distance \texorpdfstring{$\geq 2$}{>= 2} from the foot (relation R2) }
-
-Let us now consider beheadings  where the chopped-off head is at distance greater than or equal to $2$ from the foot. All the following relations are parameterized by the replication factor  $n$.
-
- Let $s$ be a sequence of hydras. 
-The proposition (\texttt{S1 n s s'}) holds if $s'$ is obtained by replacing some element $h$ of $s$ by 
-$n+1$ copies of $h'$, where  the proposition (\texttt{R1 h h'}) holds, in other words, $h'$ is just $h$, without the chopped-off  head.
-\texttt{S1} is an inductive relation with two constructors that allow us to choose the position in $s'$ of the wounded sub-hydra $h$.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#S1}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/S1Def}
-
-The rest of the definition is composed of two mutually inductive relations on hydras and sequences of hydras.
-The first constructor of \texttt{R2} describes the case where the chopped head is exactly at height $2$.
-The others constructors allow us to consider beheadings at height strictly greater than $2$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/R2Def}
-
-\subsubsection{Example}
-Let us prove the transformation of \texttt{Hy'} into \texttt{Hy''} (see Fig.~\vref{fig:Hy3}).
-We use an experimental set of tactics\footnote{See the \texttt{Ltac} definitions in~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions.}}  for specifying the place where the
-interaction between Hercules and the hydra holds.
-
-\vspace{4pt}\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}.
-
-\input{movies/snippets/Hydra_Examples/R2Example}
-
-The reader is encouraged to look at all the successive subgoals of this example.
-\emph{Please consider also exercise~\vref{exo:interactive-battle}.}
-
-\subsection{Binary relation associated with a round}
-
-We combine the two cases above into a single relation.
-First,  we define the  relation \texttt{(round\_n n h h')} where \texttt{n} is the expected number of  replications (irrelevant in the case of an \texttt{R1}-transformation).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#round_n}{Hydra.Hydra\_Definitions}}
-
-\index{hydras}{Library Hydra!Predicates!round\_n}
-
-\input{movies/snippets/Hydra_Definitions/roundNDef}
-
-By abstraction over \texttt{n}, we define a \emph{round} (small step) of a battle:
-
-\index{hydras}{Library Hydra!Predicates!round}
-\label{sect:infix-round}
-
-\input{movies/snippets/Hydra_Definitions/roundDef}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Give a direct translation of Kirby and Paris's description of hydra battles (quoted on page~\pageref{original-rules}) and prove that our relational description is consistent with theirs.
-\end{project}
-
-\subsection{Rounds and battles}
-
-Using library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}, we define \texttt{round\_plus},  the transitive closure of \texttt{round}, and \texttt{round\_star},  the reflexive and transitive closure of \texttt{round}.
-
-\label{sect:infix-rounds}
-
-\input{movies/snippets/Hydra_Definitions/roundPlus}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-  Prove that if \texttt{$h$ -+-> $h'$}, then
-  the height of $h'$ is less or equal than the height of $h$.
-
-\end{exercise}
-
-\begin{remark}
-\label{remark:transitive-closure}
-\coq's library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Coq.Relations.Relation\_Operators} 
-contains three logically equivalent definitions of the transitive closure of a binary relation. This equivalence is proved in 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} .
-
-Why three definitions for a single mathematical concept?
-Each definition generates an associated induction principle. 
- According to the form of statement one would like to prove, there is a ``best choice'':
-
-\begin{itemize}
-\item For proving $\forall y, x\,R^+\,y \;\arrow\; P\,y$, prefer 
-\texttt{clos\_trans\_n1}
-\item For proving $\forall x,\,x\,R^+\,y \;\arrow\; P\,x$, prefer \texttt{clos\_trans\_1n}
-\item For proving $\forall x\,y, \,x\,R^+\,y \;\arrow\;P\,x\,y$,  
-prefer \texttt{clos\_trans},
-\end{itemize}
-But there is no ``wrong choice'' at all: the equivalence lemmas in \linebreak 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} 
- allow the user
-to convert any one of the three closures into another one before applying the corresponding elimination tactic.
-The same remark also holds for reflexive and transitive closures. 
-\end{remark}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Define a restriction of \coqsimple{round},  where Hercules always chops off
-the leftmost among the lowest heads.
-
-Prove that, if $h$ is not a simple head, then there exists a unique $h'$ such that $h$  is transformed into $h'$ in one round, according to this restriction.
-
-\end{exercise}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}[Interactive battles]
-\label{exo:interactive-battle}
-Given a hydra \texttt{h}, the specification of a hydra battle for \texttt{h} is the type 
-\Verb@{h':Hydra | h -*-> h'}@. In order to avoid long sequences of \texttt{split}, \texttt{left}, and 
-\texttt{right}, design a set of dedicated tactics for the interactive building of a battle.
-Your tactics will have the following functionalities:
-\begin{itemize}
-\item  Chose to stop a battle, or continue
-\item Chose an expected number of replications
-\item Navigate in a hydra, looking for a head to chop off.
-\end{itemize}
-
-Use your tactics for simulating a small part of a hydra battle, for instance the rounds which lead from
-\texttt{Hy} to \texttt{Hy'''}  (Fig.~\vref{fig:Hy5}).
-
-\textbf{Hints:} 
-\begin{itemize}
-
-\item Please keep in mind that the last  configuration of your interactively built battle is known only at the end of the battle.
-Thus, you will have to create and solve subgoals with existential variables.
-For that purpose, the tactic \texttt{eexists}, applied to the
-goal \Verb@{h':Hydra | h -*-> h'}@ generates the subgoal \Verb|h -*-> ?h'|.
-\item You may use Gérard Huet's \emph{zipper} data structure~\cite{zipper} for writing tactics associated with Hercules's  interactive search for a head to chop off.
-\end{itemize}
-
-\end{exercise}
-
-\subsection{Classes of battles}
-\label{sect:battle-classes}
-
-In some presentations of hydra battles, e.g.~\cite{KP82, bauer2008}, the transformation associated with the $i$-th round may depend on $i$.
-For instance, in these articles, the replication factor at the $i$-th round is equal to $i$.
-In other examples, one can allow the hydra to apply any replication factor at any time.
-In order to be the most general as possible, we define the type of predicates which relate the state of the hydra before and after the $i$-th round of a battle.
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-\label{types:Battle}
-\index{hydras}{Library Hydra!Type classes!Battle}
-
-\input{movies/snippets/Hydra_Definitions/BattleDef}
-
-The most general class of battles is \texttt{free}, which allows the hydra to chose any replication factor at every step:
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#free}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/freeDef}
-
-We chose to call \emph{standard}\footnote{This appellation is ours.
-If there is a better one, we will change it.} the kind of battles which appear  most often in the literature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#standard}{Hydra.Hydra\_Definitions}}
-\input{movies/snippets/Hydra_Definitions/standardDef}
-
-\subsection{Big steps}
-
-Let $B$ be any instance of class \texttt{Battle}. It is easy to define inductively the relation between the $i$-th and the $j$-th steps of a battle of type $B$.
-
-\vspace{6pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/battleRelDef}
-
-The following property allows us to build battles by composition of smaller ones.
-
-%% TODO display subgoals when fixed
-
-\vspace{6pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-\noindent
-\input{movies/snippets/Hydra_Lemmas/battleTrans}
-
-% \begin{remark}
-%  The class \texttt{free} is strongly related with the transitive closure  \texttt{round\_plus}, as expressed by the following lemmas.
-
-% \vspace{4pt}
-% \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-
-%  \begin{Coqsrc}
-%  Lemma battle_free_equiv1 : forall i j h h',  
-%              battle free i h j h' ->   h -+-> h'.
-
-%  Lemma battle_free_equiv2 : forall h h',
-%      h -+-> h' ->
-%     forall i, exists j,  battle free i h j h'.
-%  \end{Coqsrc}
-
-% \end{remark}
-
-\section{A long battle}
-\label{sect:big-battle}
-
-In this section we consider a simple example of battle, starting with a small hydra,
-shown on figure~\vref{fig:hinit}, with a simple strategy for both players:
-
-\begin{itemize}
-\item At each round, Hercules chops off the rightmost head of the hydra.
-\item The battle is {standard}: at the round number $i$, the expected replication is $i$.
-\end{itemize}
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (6,0) {$\bullet$};
-\node (n1) at  (3,3) {$\bullet$};
-\node (h1) at  (1,6) {$\Smiley[1][green]$};
-\node (h2) at  (3,6) {$\Smiley[1][green]$};
-\node (h3) at  (6,6) {$\Smiley[1][green]$};
-\node (h4) at  (6,6) {$\Smiley[1][green]$};
-\node (h5) at  (6,3) {$\Smiley[1][green]$};
-\node (h6) at  (9,3) {$\Smiley[1][green]$};
-\draw (foot) -- (n1);
-\draw (n1) to   [bend left=20] (h1);
-\draw (n1) to   (h2);
-\draw (n1) to   [bend right=20] (h3);
-\draw (foot) -- (h5);
-\draw (foot) to  [bend right=20] (h6);
-\end{tikzpicture}
-
-  \caption{The hydra \texttt{hinit}}
-  \label{fig:hinit}
-\end{figure}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.BigBattle.html}{Hydra.BigBattle}}
-
-\input{movies/snippets/BigBattle/hinitDef}
-
-The lemma we would like to prove is ``The considered battle lasts exactly $N$ rounds'',
-with $N$ being a natural number we gave to guess.
-
-But the  battle is so long that no \emph{test} can give us any estimation of its length, and we do need the expressive power of logic to compute this length.
-However, in order to  guess this length, we made some experiments, computing with \gallina{}, \coq{}'s  functional programming language.
-Thus, we can consider this development as a collaboration of proof with computation.
-In the following lines, we show how we found experimentally the value of the number $N$.
-The complete proof is in file \url{../theories/html/hydras.Hydra.BigBattle.html}.
-
-\subsection{First rounds}
-During the two first rounds, our hydra loses its two rightmost heads.  Figure~\vref{fig:hinit-plus2} shows the state of the hydra   just before the third round.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (3,0) {$\bullet$};
-\node (n1) at  (3,3) {$\bullet$};
-\node (h1) at  (1,6) {$\Smiley[1][green]$};
-\node (h2) at  (3,6) {$\Smiley[1][green]$};
-\node (h3) at  (6,6) {$\Smiley[1][green]$};
-\node (h4) at  (6,6) {$\Smiley[1][green]$};
-\draw (foot) -- (n1);
-\draw (n1) to   [bend left=20] (h1);
-\draw (n1) to   (h2);
-\draw (n1) to   [bend right=20] (h3);
-\end{tikzpicture}
-
-  \caption{The hydra (\texttt{hyd1 h3})}
-  \label{fig:hinit-plus2}
-\end{figure}
-
-The following lemma  is a formal description of these first rounds, in terms of the
-\texttt{battle} predicate.
-
-\input{movies/snippets/BigBattle/L02}
-
-\subsection{Looking for regularities}
-
-A first study with pencil and paper suggested us that, after three rounds, the hydra always looks like in figure~\vref{fig:hinit-plusn} (with a variable number of 
-subtrees of height 1 or 0).
-Thus, we introduce a few handy abbreviations.
-
-\input{movies/snippets/BigBattle/Notations}
-
-For instance, the hydra shown in  Fig~\ref{fig:hinit-plusn} is  (\texttt{hyd 3 4 2}). The hydra (\texttt{hyd 0 0 0})  is the ``final'' hydra of any terminating battle, {i.e.},
-a tree whith exactly one node and no edge.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (15,0) {$\bullet$};
-\node (a) at  (3,4) {$\bullet$};
-\node (b) at  (6,4) {$\bullet$};
-\node (c) at  (9,4) {$\bullet$};
-\node (d) at  (13,4) {$\bullet$};
-\node (e) at  (16,4) {$\bullet$};
-\node (f) at  (19,4) {$\bullet$};
-\node (g) at  (22,4) {$\bullet$};
-\node (h) at  (25,4) {$\Smiley[1][green]$};
-\node (i) at  (28,4) {$\Smiley[1][green]$};
-\node (aa) at  (2.5,8) {$\Smiley[1][green]$};
-\node (ab) at  (3.5,8) {$\Smiley[1][green]$};
-\node (ba) at  (5.5,8) {$\Smiley[1][green]$};
-\node (bb) at  (6.5,8) {$\Smiley[1][green]$};
-\node (ca) at  (8.5,8) {$\Smiley[1][green]$};
-\node (cb) at  (9.5,8) {$\Smiley[1][green]$};
-\node (da) at  (13,8) {$\Smiley[1][green]$};
-\node (ea) at  (16,8) {$\Smiley[1][green]$};
-\node (fa) at  (19,8) {$\Smiley[1][green]$};
-\node (ga) at  (22,8) {$\Smiley[1][green]$};
-\draw (foot) -- (a);
-\draw (foot) -- (b);
-\draw (foot) -- (c);
-\draw (foot) -- (d);
-\draw (foot) -- (e);
-\draw (foot) -- (f);
-\draw (foot) -- (g);
-\draw (foot) -- (h);
-\draw (foot) -- (i);
-\draw (a) -- (aa);
-\draw (a) -- (ab);
-\draw (b) -- (ba);
-\draw (b) -- (bb);
-\draw (c) -- (ca);
-\draw (c) -- (cb);
-\draw (d) -- (da);
-\draw (e) -- (ea);
-\draw (f) -- (fa);
-\draw (g) -- (ga);
-\end{tikzpicture}
-
-  \caption{The hydra (\texttt{hyd 3 4 2})}
-  \label{fig:hinit-plusn}
-\end{figure}
-
-With these notations, we get a formal description of the first three rounds.
-
-\input{movies/snippets/BigBattle/L23L03}
-
-\subsection{Testing  \dots}
-\label{sect:testing}
-In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple data type for representing the states as tuples composed of
-the round number, and the respective number of daughters  \texttt{h2}, \texttt{h1}, and heads
-of the current hydra.
-
-\input{movies/snippets/BigBattle/stateDef}
-
-The following function returns the next configuration of the game.
-Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle only start with the lemma
-\texttt{step\_battle}, page~\pageref{lemma:step-battle}.
-
-\input{movies/snippets/BigBattle/nextDef}
-
-We can make bigger steps through iterations of \texttt{next}.
-The functional \texttt{iterate}, similar to Standard Library's \texttt{Nat.iter},
-is defined and studied in~\href{../theories/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
-\index{hydras}{Library Prelude!iterate}
-
-\label{Functions:iterate}
-
-\input{movies/snippets/Iterates/iterateDef}
-
-The following function computes the state of the battle at the $n$-th round.
-
-\input{movies/snippets/BigBattle/testDefTests}
-
-The battle we are studying looks to be awfully long. Let us concentrate our
-tests on some particular events : the states where $\texttt{nh}=0$.
-From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} is equal to zero.
-
-\input{movies/snippets/BigBattle/smartTest}
-
-Thus, $(1 + 11)$ rounds later, the \texttt{n1} field is equal to $2$, and 
-\texttt{nh}   to $0$.
-
-\input{movies/snippets/BigBattle/smartTestb}
-
-Next round, we decrement \texttt{n2} and set \texttt{n1} to $95$.
-
-\input{movies/snippets/BigBattle/smartTestc}
-
-We now have some intuition of the sequence.
-It looks like the next ``\texttt{nh}=0'' event will happen at the $192=2(95+1)$-th round, then at the $2(192+1)$-th round, etc.
-
-\input{movies/snippets/BigBattle/doubleS}
-
-\subsection{Proving \dots}
-We are now able to reason about the sequence of transitions defined by our hydra battle. Instead of using the data-type \texttt{state} we study the relationship
-between different configurations of the battle.
-
-Let us define a binary relation associated with every round of the battle.
-In the following definition \texttt{i} is associated with the round number (or date, if we consider a discrete time), and \texttt{a}, \texttt{b}, \texttt{c} respectively associated with the number of occurrences of \texttt{h2}, \texttt{h1} and heads connected to the hydra's foot.
-
-\input{movies/snippets/BigBattle/oneStep}
-
-The relation between \texttt{one\_step} and the rules of hydra battles is asserted by the following lemma.
-
-\label{lemma:step-battle}
-
-\input{movies/snippets/BigBattle/stepBattle}
-
-\vspace{4pt}
-
-Next, we define ``big steps'' as the transitive closure of \texttt{one\_step},
-and reachability (from the initial configuration of figure~\ref{fig:hinit} at time $0$).
-
-\input{movies/snippets/BigBattle/steps}
-
-The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{battle}.
-
-\input{movies/snippets/BigBattle/stepsBattle}
-
-\vspace{4pt}
-
-Thus, any result about \texttt{steps} will be applicable to standard battles.
-Using the predicate \texttt{steps},  our study of the length of the considered battle
-can  be decomposed into three parts:
-
-\begin{enumerate}
-\item  Characterization and proofs of regularities of some events (inspired by our experiments of Sect.~\ref{sect:testing}).
-\item Study of the beginning of the battle
-\item Computing the exact length of the battle.
-\end{enumerate}
-
-First, we prove that, if at round $i$ the hydra is equal to
-(\texttt{hyd a (S b) 0}), then it will be equal to (\texttt{hyd a b 0}) at the $2(i+1)$-th round.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/LS}
-
-From now on, the lemma \texttt{reachable\_S} allows us to watch larger and larger steps of 
-the battle.
-
-\input{movies/snippets/BigBattle/L4}
-
-\input{movies/snippets/BigBattle/L10To95}
-
-\subsection{Giant steps}
-
-We are now able to make bigger steps in the simulation of the battle.
-First, we iterate the lemma \texttt{reachable\_S}.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/Bigstep}
-
-\vspace{4pt}
-
-Applying lemmas \texttt{BigStep} and \texttt{L95} we make a first jump.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/MDef}
-
-Figure~\ref{fig:HM}  represents the hydra at the $M$-th round.
-At the $(M+1)$-th round, it will look like in fig~\ref{fig:HM-plus1}.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:HM}}
-The state of the hydra after $M$ rounds.
-% The hydra \texttt{h} of the proof that \(\omega^2\) is too small for proving Hercules' victory
-
-\end{figure}
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (0,5) {$\bullet$};
-\node (N12) at (0,8) {$\Smiley[2][green]$};
-\node (N2) at (2,5) {$\bullet$};
-\node (N22) at (2,8) {$\Smiley[2][green]$};
-\node (N3) at (4,5) {$\bullet$};
-\node (N32) at (4,8) {$\Smiley[2][green]$};
-\node (N4) at (6,5) {$\bullet$};
-\node (N42) at (6,8) {$\Smiley[2][green]$};
-
-\node (Ndots) at (12,8) {\Huge $\dots$};
-\node (Ndots2) at (12,5) {\Huge $\dots$};
-
-\node (N8) at (18,5) {$\bullet$};
-\node (N82) at (18,8) {$\Smiley[2][green]$};
-\node (N9) at (20,5) {$\bullet$};
-\node (N92) at (20,8) {$\Smiley[2][green]$};
-
-\draw (foot) -- (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N8);
-\draw (foot) -- (N9);
-\draw (N1) to  (N12);
-\draw (N2) to  (N22);
-\draw (N3) to  (N32);
-\draw (N4) to  (N42);
-\draw (N8) to  (N82);
-\draw (N9) to  (N92);
-\end{tikzpicture}
-\caption{\label{fig:HM-plus1}}
-The state of the hydra after $M+1$ rounds (with $M+1$ heads).
-
-\end{figure}
-
-\input{movies/snippets/BigBattle/L295S}
-
-\vspace{4pt}
-
-Then, applying once more the lemma \texttt{BigStep}, we get the exact time when
-Hercules wins!
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/NDef}
-
-\vspace{4pt}
-
-We are now able to prove formally that the considered battle is 
-composed of $N$ steps.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/Done}
-
-\subsection{A minoration lemma}
-
-Now, we would like to get an intuition of  how big the number $N$ is.
-For that purpose, we use a minoration of the function \texttt{doubleS} by the
-function (\texttt{fun n => 2 * n}).
-
-\vspace{4pt}
-
-\input{movies/snippets/Exp2/exp2Def}
-
-\vspace{4pt}
-
-Using a few facts (proven in 
-\href{../theories/html/hydras.Hydra.BigBattle.html}{hydras.Hydra.BigBattle}),we get several  minorations.
-
-\input{movies/snippets/BigBattle/minorationLemmas}
-
-\vspace{4pt}
-
-The number $N$ is greater than or  equal to $2^{2^{95}\times 95}.$ If we wrote $N$ in base $10$, $N$ would require at least $10^{30}$ digits!
-
-\section{Generic properties}
-
-The example we just studied shows that the termination of any battle may take a very long time. If we want to study hydra battles in general, we have to consider 
-any hydra and any strategy, both for Hercules and the hydra itself. So, we first  give some definitions, generally borrowed from transition systems vocabulary (see~\cite{tel_2000} for instance).
-
-\subsection{Liveliness}
-
-Let $B$ be an instance of \texttt{Battle}. We say that $B$ is \emph{alive} if
-for any configuration $(i,h)$, where $h$ is not a head, there exists a further step in class $B$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/AliveDef}
-
-The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
-\href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
-
-\input{movies/snippets/Hydra_Theorems/AliveThms}
-
-Both theorems are proved with the help of the  following strongly specified function:
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
-
-\input{movies/snippets/Hydra_Lemmas/nextRoundDec}
-
-\subsection{Termination}
-
-The termination of \emph{any}  battle is naturally expressed by the predicate \texttt{well\_founded} defined in the module \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}
- of the Standard Library.
-
-\index{hydras}{Library Hydra!Predicates!Termination}
-
-\input{movies/snippets/Hydra_Definitions/TerminationDef}
-
-Let $B$ be an instance of class \texttt{Battle}. A \emph{variant} for $B$ consists
-in a well-founded relation $<$  on some type \texttt{A}, and a function
-(also called a \emph{measure}) $m:\texttt{Hydra}\rightarrow A$ such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $B$, the inequality $m(h')<m(h)$ holds.
-
-\vspace{4pt}
-\noindent
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
-
-\label{sect:hvariant-def}
-
-\index{hydras}{Library Hydra!Type classes!Hvariant}
-
-\input{movies/snippets/Hydra_Definitions/HvariantDef}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
- Prove that, if there exists some  instance of (\texttt{Hvariant Lt wf\_Lt $B$ $m$}), then there exists no infinite battle in  $B$.
-\end{exercise}
-
-\subsection{A  small proof of impossibility}
-%\index{coq}{Proofs of impossibility}
-
-\label{omega-case}
-
-When one wants to prove a termination theorem with the help of a variant, 
-one has to consider first a well-founded set $(A,<)$, then a strictly decreasing measure on this set.
-The following two lemmas show that if  the order structure $(A,<)$ is too simple, it is useless to look for a convenient measure, which simply no exists.
-Such kind of result is useful, because it saves you time and effort.
-
-The best known well-founded order is the natural order on the set $\mathbb{N}$ of natural numbers (the type \texttt{nat} of Standard library).
-It would be interesting to look for some measure $m:\texttt{nat}\arrow\texttt{nat}$ and prove it is a variant.
-
-Unfortunately, we can prove that 
-\emph{no} instance of class (\texttt{WfVariant round Peano.lt $m$}) can be built, where
-$m$ is \emph{any} function of type \texttt{Hydra $\arrow$ nat}.
-
-Let us present the main steps of that proof, the script of which  is in the module ~\href{../theories/html/hydras.Hydra.Omega_Small.html}{Hydra/Omega\_Small.v} \footnote{ The name of this file means ``the ordinal $\omega$ is too small for proving the termination of [free] hydra battles ''.
-In effect, the elements of $\omega$, considered as a set, are just the natural numbers (see next chapter for more details)}.
-
-%\subsubsection{Preliminaries}
-
-Let us assume there exists some variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
-    the  termination of all hydra battles.
-
-\input{movies/snippets/Omega_Small/omegaSmalla}
-
-We define an injection $\iota$ from the type \texttt{nat} into \texttt{Hydra}.
-For any natural number $i$, $\iota(i)$ is the hydra composed of a foot and
-$i+1$ heads at height $1$. For instance, Fig.~\ref{fig:flower} represents the hydra $\iota(3)$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (4,0) {$\bullet$};
-\node (N1) at (2,2) {$\Smiley[2][green]$};
-\node (N2) at (4,2) {$\Smiley[2][green]$};
-\node (N3) at (6,2) {$\Smiley[2][green]$};
-\node (N4) at (8,2) {$\Smiley[2][green]$};
-\draw (foot) to [bend left =25] (N1);
-\draw (foot) to [bend left =15] (N2);
-\draw (foot) to [bend right =15] (N3);
-\draw (foot) to [bend right =25] (N4);
-\end{tikzpicture}
-\caption{\label{fig:flower}
-The hydra $\iota(3)$}
-\end{figure}
-
-\input{movies/snippets/Omega_Small/iotaDef}
-
-Let us consider now some hydra \texttt{big\_h} out of the range of the injection $\iota$ (see Fig.~\vref{fig:h-omega-omega}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to  (N2);
-\end{tikzpicture}
-\caption{\label{fig:h-omega-omega}}
- The hydra \texttt{big\_h}.
-\end{figure}
-
-\input{movies/snippets/Omega_Small/bigHDef}
-
- Using the functions $m$ and $\iota$, we define a second hydra \texttt{small\_h}, and show
- there is a one-round battle that transforms \texttt{big\_h} into \texttt{small\_h}. Please note that,
-due to the hypothesis \texttt{Hvar}, we are interested in the termination of \emph{free} battles. 
-There is no problem to consider a round with (\texttt{m big\_h}) as the replication factor.
-
-\input{movies/snippets/Omega_Small/smallHDef}
-
-But, by hypothesis, $m$ is a variant. Hence, we infer the following inequality.
-
-\vspace{4pt}
-
-\input{movies/snippets/Omega_Small/mLt}
-
-In order to get a contradiction, it suffices to  prove the inequality
-$m(\texttt{big\_h}) \leq m(\texttt{small\_h})$ i.e.,  $m(\texttt{big\_h}\leq m(\iota (m(\texttt{big\_h})))$.
-
-\input{movies/snippets/Omega_Small/mGea}
-
-Intuitively, it means that, from any hydra of the form (\texttt{iota $i$}), the battle will 
-take (at least) $i$ rounds. Thus the associated measure cannot be less than $i$.
-Technically, we prove this lemma by Peano induction on $i$.
-
-\begin{itemize}
-\item The base case $i=0$ is trivial
-\item Otherwise, let $i$ be any natural number and assume  the inequality
-  $i \leq m(\iota(i))$.
-  \begin{enumerate}
-  \item  But the hydra $\iota(S(i))$ can be transformed in one round into
-    $\iota(i)$ (by losing its rightmost head, for instance)
-  \item Since $m$ is a variant, we have $m(\iota(i)) < m(\iota(S(i)))$,
-    hence  $i< m(\iota(S(i)))$, which implies  $S(i)\leq  m(\iota(S(i)))$.
-  \end{enumerate}
-\end{itemize}
-
-We are now ready to complete our impossibility proof.
-
-\vspace{4pt}
-
-\inputsnippets{Omega_Small/mGeb, Omega_Small/mGez,
-  Omega_Small/omegaSmallz}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\subsubsection{Conclusion}
-
-In order to build a variant for proving the termination of all hydra battles, we need to consider order structures more complex than the usual order on type \texttt{nat}. 
-The notion of \emph{ordinal number} provides a catalogue of well-founded order types.
-For a reasonably large bunch of ordinal numbers, \emph{ordinal notations} are data-types which allow the \coq{} user to define functions, to compute and prove some properties, for instance by reflection.
-
-The next chapter is dedicated to a generic formalization of ordinal notations, and chapter~\ref{chap:T1} to a proof of termination of all hydra battles with the help of an ordinal notation for the interval $[0,\epsilon_0)$\,\footnote{We use the mathematical notation $[a,b)$ for the interval $\{x|a\leq x < b\}$.}.
-\index{maths}{Notations!Interval}
-
-%--------------------------------------------------------------
-
-\chapter{Introduction to ordinal numbers and ordinal notations}
-\label{chap:ON}
-
-The proof of termination of all hydra battles presented in~\cite{KP82} is based
-on \emph{ordinal numbers}.
-From a mathematical point of view, an ordinal is a representative of an equivalence class for isomorphisms of  totally ordered well-founded sets.
-
-For the computer scientist, ordinals are tools for proving the totality of a given recursive function, or termination of a transition system. \emph{Ordinal arithmetic} 
-provides a set of functions whose properties, like \emph{monotony}, allow to define \emph{variants}, \emph{i.e.} strictly decreasing measures used in proofs of termination.
-
-\vspace{4pt}
-
-Let us have a look at Figure~\ref{fig:ordinal-sequence}. It presents a few items of a  sequence of ordinal numbers, which extends the sequence of natural numbers.
-
-\begin{figure}[h]
-  \centering
-\fbox{\Large
-  \begin{minipage}{1.0\linewidth}
-  \begin{align*}
-     &\textcolor{blue}{0},\,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,\ldots\\
-&\textcolor{red}{\omega},\,\omega+1,\omega+2,\omega+3,\ldots\\
-&\textcolor{red}{\omega\times 2},\,\omega\times 2+1,\ldots, \textcolor{red}{\omega\times 3},\,\omega\times 3+1,\ldots, \textcolor{red}{\omega\times 4},\ldots\\
-&\textcolor{red}{\omega^2},\ldots, \textcolor{red}{\omega^2\times 42},\ldots,\textcolor{red}{\omega^3},\ldots, \textcolor{red}{\omega^4},\omega^4+1,\ldots\\
-&\textcolor{red}{\omega^\omega},\ldots, \textcolor{red}{\omega^\omega+\omega^7\times 8},\ldots,\textcolor{red}{\omega^\omega\times 2},\omega^\omega\times 2+1, \ldots\\
-&\textcolor{red}{\omega^{\omega^\omega}},\ldots, \textcolor{red}{\omega^{\omega^\omega}+\omega^\omega\times 42+ \omega^{55}+\omega}, \ldots, \textcolor{red}{\omega^{\omega^{\omega+1}}}, \omega^{\omega^{\omega+1}}+1,\dots\\
-& \textcolor{red}{\epsilon_0 (= \omega^{\epsilon_0)}}, \epsilon_0+1, \epsilon_0+2, \epsilon_0+3, \ldots\\
-& \textcolor{red}{\epsilon_1}, \ldots, \textcolor{red}{\epsilon_2}, \ldots, \textcolor{red}{\epsilon_\omega},\ldots \\
-& \textcolor{red}{\Gamma_0}, \Gamma_0+1, \Gamma_0+2, \Gamma_0+3,\ldots, \textcolor{red}{\Gamma_0+\omega}, \ldots\\
-&\ldots
-  \end{align*}   
-  \end{minipage}}
-
-  \caption{A short overview of the sequence of ordinal numbers}
-  \label{fig:ordinal-sequence}
-\end{figure}
-
-Let us comment some features of this figure:
-
-\begin{itemize}
-\item The ordinals are listed in a strictly increasing order. 
-\item Dots : ``$\ldots$'' stand for  infinite sequences of ordinals, not shown for lack of space.
-For instance, the ordinal $42$ is not shown in the first line, but it exists, somewhere between $17$ and $\omega$.
-\item Each ordinal printed in black is the immediate successor of another ordinal.
-We call it a \emph{successor} ordinal.
-For instance, $12$ is the successor of $11$, and $\omega^4+1$ the successor of $\omega^4$.
-\item Ordinals (displayed in red)  that  follow immediately dots are called \emph{limit ordinals}.
-With respect to the order induced by this sequence, any limit ordinal $\alpha$ is the least upper bound of  the set $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
-\item
-For instance $\omega$ is the least upper bound of the set of all finite ordinals (in the first line). It is also the first limit ordinal, and the first infinite ordinal, in the sense that 
-the set $\mathbb{O}_\omega$ is infinite.
-\item The ordinal $\epsilon_0$ is the first number which is equal to its own exponential of base $\omega$.
-It plays an important role in proof theory, and is particularly studied in chapters~\ref{chap:T1} to \ref{chap:alpha-large}.
-\item Any ordinal is  either the ordinal \textcolor{blue}{$0$},
-a successor ordinal, or a \textcolor{red}{limit ordinal}.
-\end{itemize}
-
-\section{The mathematical point of view}
-
-\subsection{Well-ordered sets}
-Let us start with some definitions.
-A  \emph{well-ordered set} is a set provided with a binary relation $<$ which has the following properties.
-\begin{description}
-\item[irreflexivity] : $\forall x\in A, x\not< x$
-\item[transitivity] : $\forall x\,y\,z\in A, x<y \Rightarrow y<z \Rightarrow x<z$
-\item[trichotomy]: $\forall x\,y\in A, x<y \vee x = y \vee y < x$
-\item[well foundedness]: $<$ is well-founded (every element of $A$ is accessible)\footnote{In classical mathematics, we would say that there is no infinite sequence $a_1>a_2> \dots a_n> a_{n+1}\dots$ in $A$.
-In contrast, \coq's standard library contains
-an inductive definition of a predicate \texttt{Acc} which allows us to write 
-constructive proofs of accessibility (See \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}).}.
-\end{description}
-
-The best known examples of well-ordered sets are the set $\mathbb{N}$ of natural numbers (with the usual order $<$), as well as any finite segment $[0,i)=\{j\in\mathbb{N}\,|\,j<i\}$.
-The disjoint union of two copies of $\mathbb{N}$, \emph{i.e.} the set $\{0,1\}\times\mathbb{N}$ is also well-ordered,
-with respect to the order below:
-
-\begin{align*}
-(i,j) < (i,k) & \;\textit{\textbf{if} }\; j < k\\
-(0,k) < (1,l) & \;\textit{\textbf{for\,any}}\;k \;\textit{\textbf{and}} \; l
-\end{align*}
-
-\subsection{Ordinal numbers}
-
-\index{maths}{Ordinal numbers}
-
-Let $(A,<_A)$ and $(B,<_B)$ two well-ordered sets. $A$ and $B$ are said to have \emph{the same order type} if 
-there exists a strictly monotonous bijection $b$ from $A$ to $B$, \emph{i.e.} which verifies the proposition
-$\forall x\,y\in A,\, x <_A y \Rightarrow b(x) <_B  b(y)$.
-
-Having the same order type is an equivalence relation between well-ordered sets. Ordinal numbers (in short: \emph{ordinals})  are descriptions (\emph{names}) of the equivalence classes.
-For instance, the order type of $(\mathbb{N},<)$ is associated with the ordinal called  $\omega$, and the order we considered on 
-the disjoint union of $\mathbb{N}$ and itself is named $\omega+\omega$.
-
-In a set-theoretic framework, one can consider any ordinal $\alpha$ as a well-ordered set, whose  elements are just the ordinals strictly less than $\alpha$, \emph{i.e.} the \emph{segment} $\mathbb{O}_\alpha=[0, \alpha)$.
-So, one can speak about \emph{finite}, \emph{infinite}, \emph{countable}, etc., ordinals.
-Nevertheless, since we work within type theory,
-we do not identify ordinals as sets of ordinals, but consider the correspondence between ordinals and sets of ordinals as the function that maps $\alpha$ to $\mathbb{O}_\alpha$.
-For instance $\mathbb{O}_\omega=\mathbb{N}$, and $\mathbb{O}_7=\{0,1,2,3,4,5,6\}$.
-
-We cannot cite all the literature published on ordinals since Cantor's book
-\cite{cantorbook}, and 
-leave it to the reader to explore the bibliography.
-
-The introduction of Jos\'e Grimm's report~\cite{grimm:hal-00911710} contains a nice presentation of the main properties of ordinals.
-
-\section{Ordinal numbers in Coq}
-
-Two kinds of representation of ordinals are defined in our development.
-
-\begin{itemize}
-\item A ``mathematical'' representation of the set of countable ordinal numbers, after Kurt Schütte~\cite{schutte}.
-This representation uses several (hopefully harmless) axioms.
-We use it as a reference for proving the correctness of ordinal notations.
-  We are also progressively importing \gaia's~\cite{Gaia} definitions and theorems, based on Bourbaki's set theory~\cite{Grimm2016}.
-Chapter~\ref{gaia-chapter} is dedicated to the connexion between libraries \HydrasLib and \gaia.
-
-\item A family of \emph{ordinal notations} (also called \emph{[ordinal] notation systems}), \emph{i.e.} data types used to represent segments $[0,\mu)$, where $\mu$ is some countable ordinal.
-Each ordinal notation is defined inside the Calculus of Inductive Constructions (without axioms).
-Many functions are defined, allowing proofs by computation.
-Note that proofs of
-correctness of a given ordinal notation with respect to Schütte's model obviously use axioms.
-Please execute the \texttt{Print Assumptions} command in case of doubt.
-\end{itemize}
-
-% \section{Countable ordinals}
-
-% Chapter~\ref{chap:schutte} of this document presents an adaptation to \coq{} of an axiomatization in classical logic of the set of countable ordinals by K. Schütte~\cite{schutte}. 
-% That formalization is quite complex, technical and unshamedly non-constructive,  so we put its description  in the last chapter of this document.
-
-% Please note that Schütte considers the (uncountable) set $\mathbb{O}$ of all countable ordinals.
-This set is well ordered (which is one of Schütte's axioms), and associates to any ordinal $\alpha$ the segment $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
-
-% In our adaptation to \coq{}, we declare a type \texttt{Ord}, a binary relation \texttt{lt} (with infix notation \texttt{"\_<\_"}, and assume Schütte's axiom. In Chapter~\ref{chap:schutte},
-% we derive some interesting properties of countable ordinals from these axioms.
-
-It is interesting to compare proofs of a given property (for instance the associativity of addition) both in the computational framework of some ordinal notation, and in mathematical  models
-of Schütte or \gaia.
-
-\section{Ordinal Notations}
-
-Fortunately, the ordinals we need for  studying hydra battles are much simpler than Schütte's, and can be represented as quite simple data types in \gallina.
-
-Let $\alpha$ be some (countable) ordinal; 
-in \coq{} terms, we call \emph{ordinal notation for $\alpha$} a structure composed 
-of:
-\begin{itemize}
-\item A data type $A$ for representing all ordinals strictly below $\alpha$,
-\item A well founded order $<$ on $A$, 
-\item A correct function for comparing two ordinals. Note  that the reflexive closure of $<$ is thus a \emph{total order}.
-\end{itemize}
-
-Such a structure may be proved correct relatively to a bigger ordinal notation or to Schütte's or \gaia's model.
-
-\subsection{Classes for ordinal notations}
-
-From the \coq{} user's point of view, an ordinal notation is
-a structure allowing to compare two ordinals by computation, and proving by well-founded induction.
-
-\subsubsection{The \texttt{Comparison} and \texttt{Comparable} classes}
-
-We use the operational class \texttt{Comparison} of comparison functions to define the \texttt{Comparable} class, contributed by Jérémy Damour and Théo Zimmermann, which allows us to apply generic lemmas and tactics about decidable strict orders.
-The correctness of the comparison function is expressed through Stdlib's type 
-\texttt{Datatypes.CompareSpec} as specialized by \texttt{Datatypes.CompSpec}.
-
-\begin{Coqsrc}
-  Inductive CompareSpec (Peq Plt Pgt : Prop) :
-  comparison -> Prop :=
-    CompEq : Peq -> CompareSpec Peq Plt Pgt Eq
-  | CompLt : Plt -> CompareSpec Peq Plt Pgt Lt
-  | CompGt : Pgt -> CompareSpec Peq Plt Pgt Gt.
-
-Definition CompSpec {A} (eq lt : A -> A -> Prop) (x y : A) :=
- CompareSpec (eq x y) (lt x y) (lt y x).
-\end{Coqsrc}
-
-\emph{From Module~\href{../theories/html/hydras.Prelude/mparable.html\#Hvariant}{Prelude.Comparable}}
-
-\label{sect:comparable-def}
-
-\input{movies/snippets/Comparable/ComparableDef}
-
-\subsubsection{The \texttt{ON} class}
-
-The following class definition, parameterized with a type $A$, a binary relation \texttt{lt} on $A$, specifies that \texttt{lt} is a well-founded strict order, provided with a correct comparison function.
-
-\vspace{4pt}
-\noindent\emph{From
-Library~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}}
-
-\label{types:ON}
-\index{hydras}{Library OrdinalNotations!Type classes!ON}
-
-\input{movies/snippets/ON_Generic/ONDef}
-
-We give  also a few handy definitions and lemmas for any ordinal notation.
-
-\label{sect:on-lt-notation}
-\label{sect:on-le-notation}
-\label{sect:measure-ON}
-\label{sect:bigO-ON}
-
-\inputsnippets{ON_Generic/ONDefsa}
-\inputsnippets{ON_Generic/ONDefsb}
-
-\begin{remark}
-The infix notations \texttt{o<} and \texttt{o<=} are defined in order to make apparent the distinction between the various notation scopes that may co-exist in a same statement.
-So the infix \texttt{<} and \texttt{<=} are reserved to the natural numbers.
-In the mathematical formulas, however, we still use $<$ and $\leq$ for comparing ordinals.
-\end{remark}
-
-% \subsection{Ordinal notations and  termination measures}
-% \label{sect:measure-ON}
-
-% The following lemma (together with the type class mechnism) allows us to define termination measures over any ordinal notation.
-It is just an application of  the libraries \texttt{Coq.Wellfounded.Inverse\_Image}
-% and  \texttt{Coq.Wellfounded.Inclusion}.
-
-% \begin{Coqsrc}
-% Definition measure_lt {A:Type}{lt: relation A}
-%             {compare : A -> A -> comparison}
-%             {on : ON lt compare}
-%             {B : Type} (m : B -> A) : relation B :=
-%              fun x y => on_lt (m x) (m y).
-
-% Lemma wf_measure  {A:Type}(lt: relation A)
-%             {compare : A -> A -> comparison}
-%             {on : ON lt compare}
-%             {B : Type}
-%             (m : B -> A):  well_founded (measure_lt m). 
-% \end{Coqsrc}
-
-% A simple example of application is given in Sect.~\vref{sect:merge-example}.
-
-\section{Example: the ordinal \texorpdfstring{$\omega$}{omega}}
-
-The simplest example of ordinal notation is built over the type \texttt{nat} of \coq's standard library. We just have to apply already proven lemmas about Peano numbers.
-
-\vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_Omega.html}{OrdinalNotations.ON\_Omega}}
-
-\inputsnippets{ON_Omega/OmegaDefa, ON_Omega/OmegaDefb}
-
-\section{Sum of  two ordinal notations}
-
-Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective types \texttt{A} and \texttt{B}.
-
- We consider a new strict order
-on the disjoint sum of the associated types, by putting all elements of \texttt{A} before the elements of \texttt{B} (thanks to Standard Library's relation operator \texttt{le\_AsB}).
-
-\vspace{4pt}
-\noindent
-\emph{From Library~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}}.
-
-\begin{Coqanswer}
-Inductive
-le_AsB (A B : Type) (leA : A -> A -> Prop) (leB : B -> B -> Prop)
-  : A + B -> A + B -> Prop :=
-| le_aa : forall x y : A, leA x y -> le_AsB A B leA leB (inl x) (inl y)
-| le_ab : forall (x : A) (y : B), le_AsB A B leA leB (inl x) (inr y)
-| le_bb : forall x y : B, leB x y -> le_AsB A B leA leB (inr x) (inr y)
-\end{Coqanswer}
-
-%\pagebreak
-
-\vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_plus.html}{OrdinalNotations.ON\_plus}}
-
-\input{movies/snippets/ON_plus/Defs}
-
-In order to build an instance of \texttt{Comparable}, we have to define a correct comparison function.
-
-\inputsnippets{ON_plus/compareDef, ON_plus/compareCorrect}
-\inputsnippets{ON_plus/plusComp}
-
-Standard library's lemma \texttt{Wellfounded.Disjoint\_Union.wf\_disjoint\_sum} 
-is applied  to prove that our order \texttt{lt} is well-founded, allowing us to build an instance of \texttt{ON}:
-
-\inputsnippets{ON_plus/ltWf, ON_plus/OnPlus}
-
-\subsection{The ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
-
-The ordinal $\omega+\omega$ (also known as $\omega\times 2$) may be represented as the concatenation 
-of two copies of $\omega$ (Figure~\ref{fig:omega-plus-omega}).
-It is also represented by the two first lines of Figure~\ref{fig:ordinal-sequence}.
-
-\begin{figure}[h]
-   \centering
-   \begin{tikzpicture}[very thick, scale=0.5]
-\begin{scope}[color=blue]
-\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
-\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
-\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
-\node (Adots) at (6,0) {$\ldots$};
-\node(An) at (8,0)[label=below:$n$]{$\bullet$};
-\node(A2) at (10,0)[label=below:$n+1$]{$\bullet$};
-\node (Adots1) at (12,0) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red]
-\node(B0) at (14,0)[label=below:$0$,label=above:\textcolor{red}{$\omega$}]{$\bullet$};
-\node(B1) at (16,0)[label=below:$1$, label=above:$\omega+1$]{$\bullet$};
-\node(B2) at (18,0)[label=below:$2$,label=above:$\omega+2$]{$\bullet$};
-\node (Bdots) at (20,0) {$\ldots$};
-\node (Bn) at (22,0) [label=below:$p$, label=above:$\omega+p$]{$\bullet$};
-\node (Bdots2) at (24,0) {$\ldots$};
-\end{scope}
-\end{tikzpicture}
-   \caption{\textcolor{blue}{$\omega+{\color{red}\omega}$}}
-   \label{fig:omega-plus-omega}
- \end{figure}
-
-We can define this notation in \coq{} as an instance of \texttt{ON\_plus}.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/OmegaPlusOmegaDef}
-
-\vspace{4pt}
-
-We can now define abbreviations.
-For instance, the finite ordinals are represented by terms built with  the constructor \texttt{inl}, and the first infinite ordinal $\omega$ by the term \texttt{(inr 0)}.
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Omega_plus_omega/finiteOmega}
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Omega_plus_omega/ltOmega}
-
-% \label{warning:coercions}
-% \index{Coq!Coercions} 
-% \begin{remark}
-% Beware of coercions and notation scopes!
-% Let us consider the following goal:
-
-% \begin{Coqsrc}
-%  Goal (6 o< 8).
-%  auto with arith.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% 1 subgoal (ID 9)
-
-%   ============================
-%   6 o< 8
-% \end{Coqanswer}
-
-% Please keep in mind that the current notation scope interprets the infix \texttt{``<''} as the predicate \texttt{Omega\_plus\_omega.lt} and not \texttt{Nat.lt}.
-More,  the coercion mechanism converts the terms \texttt{6:nat} [resp.
-\texttt{8:nat} ]
-% into \texttt{inl 6} [resp. \texttt{inl 8}].  So, the initial goal is correctly interpreted by \coq{}, but not as an inequality between two natural numbers.
-
-% \begin{Coqsrc}
-% Set Printing All.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% 1 subgoal (ID 337)
-
-%   ============================
-%   @on_lt nat Peano.lt Nat.compare Omega (S (S (S (S (S (S O))))))
-%     (S (S (S (S (S (S (S (S O))))))))
-% \end{Coqanswer}
-
-% Anyway, the initial goal is provable, using \texttt{le\_AsB}'s first constructor.
-
-% \begin{Coqsrc}
-%   constructor; auto with arith.
-% Qed.
-% \end{Coqsrc}
-
-% \end{remark}
-%
-
-\section{Limits and successors}
-
-Let us look again at our implementation of $\omega+\omega$. We can classify its elements into three categories:
-
-\begin{itemize}
-\item The least ordinal, \texttt{(inl 0)}, also known as  \texttt{(fin 0)}.
-\item The first infinite ordinal $\omega$.
-\item The remaining ordinals, either of the form \texttt{(inl (S $i$))} or \texttt{(inr (S $i$))} (in black on Figure~\ref{fig:ordinal-sequence}), called \emph{successor ordinals}.
-\end{itemize}
-
-\subsection{Definitions}
-It would be interesting to specify at the most generic level, what is a zero, a successor or a limit ordinal. Let $<$ be a strict order on a type $A$.
-
-\begin{itemize}
-\item A \emph{least} element is a minorant (in the large sense) of the full set  on $A$,
-\item $y$ is a \emph{successor} of $x$ if $x<y$ and there is no element between $x$ and $y$. We will also say that $x$ is a \emph{predecessor} of $y$.
-\item $x$ is a \emph{limit} if $x$ is not a least element, and for any $y$ such that $yo<x$,
- there exists some $z$ such that $y<z<x$.
-\end{itemize}
-
-The following definitions are in Library \href{../theories/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
-
-\input{movies/snippets/MoreOrders/Defs}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove, that, in any ordinal notation system, every ordinal has at most one predecessor, and at most one successor.
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/predSuccUnicity.v}{exercises/ordinals/predSuccUnicity.v}.}
-
-\end{exercise}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove, that, in any ordinal notation system, if $\beta$ is a successor of $\alpha$,
-then for any $\gamma$, $\gamma<\beta$ implies 
-$\gamma\leq\alpha$.
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/lt_succ_le.v}{exercises/ordinals/lt\_succ\_le.v}.}
-\end{exercise}
-
-\subsection{Limits and successors in \texorpdfstring{$\omega+\omega$}{omega+omega}}
-
-Using the definitions above, we can prove the following lemma:
-
-\vspace{4pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/limitIff}
-
-\vspace{4pt}
-
-Regarding successors, let us define the following function and prove its correctness:
-
-\input{movies/snippets/ON_Omega_plus_omega/succDef}
-
-\input{movies/snippets/ON_Omega_plus_omega/succCorrect}
-
-\vspace{4pt}
-
-We can also check whether an ordinal is a successor by a simple computation:
-
-\input{movies/snippets/ON_Omega_plus_omega/succb}
-
-\vspace{4pt}
-
-Finally, the nature of any ordinal is decidable (inside this notation system) :
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/ZeroLimitSuccDec}
-
-\section{Product of ordinal notations}
-
-Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective  ordered types \texttt{A} and \texttt{B}.
-The product of \texttt{NA} and \texttt{NB} is considered as the concatenation of $B$ copies of $A$, ordered by the lexicographic order on $B\times A$.
-
-In \coq{}, we build an instance of class \texttt{ON} through a sequence of steps as for the sum of ordinal notations.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_mult.html}{OrdinalNotations.ON\_mult}}
-
-\inputsnippets{ON_mult/Defs}
-\inputsnippets{ON_mult/multComp,  ON_mult/ONMult}
-\inputsnippets{ON_mult/endDefs}
-
-\section{The ordinal \texorpdfstring{$\omega^2$}{omega2}}
-
-The ordinal $\omega^2$ (also called $\phi_0(2)$, see Chap.~\ref{chap:schutte}), is an instance of the multiplication presented in the preceding section.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
-
-\inputsnippets{ON_Omega2/Omega2Def, ON_Omega2/Defs}
-\inputsnippets{ON_Omega2/Ex1}
-
-\subsection{Arithmetic of \texorpdfstring{$\omega^2$}{omega^2}}
-
-\subsubsection{Successor}
-
-The successor of any ordinal is defined by a simple pattern-matching.
-
-\input{movies/snippets/ON_Omega2/succ}
-
-This function is proved to be correct w.r.t. the \texttt{Successor} predicate.
-
-\input{movies/snippets/ON_Omega2/succOK}
-
-\input{movies/snippets/ON_Omega2/succLemmas}
-
-\subsubsection{Addition}
-
-We can define on \texttt{Omega2} an addition which extends the addition on \texttt{nat}. Please note that this operation is not commutative:
-
-\inputsnippets{ON_Omega2/plusDef, ON_Omega2/plusExamples}
-
-\subsubsection{Multiplication}
-
-The restriction of ordinal multiplication to the segment $[0,\omega^2)$ is not a total function.
-For instance $\omega\times\omega= \omega^2$ is outside the set of represented values.
-Nevertheless, we can define two operations mixing natural numbers and ordinals.
-
-\input{movies/snippets/ON_Omega2/multFinDef}
-\input{movies/snippets/ON_Omega2/multFinExamples}
-
-Multiplication with a finite ordinal and addition are related through the following lemma:
-
-\inputsnippets{ON_Omega2/uniqueDecompositiona,
-  ON_Omega2/uniqueDecompositionb,
-  ON_Omega2/uniqueDecompositionz}
-
-\subsection{A proof of termination using \texorpdfstring{$\omega^2$}{omega^2}} 
-\label{sect:merge-example}
-
-Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily a total function which merges two lists (example contributed by Pascal Manoury).
-
-%\index{coq}{Commands!Function}
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
-
-\inputsnippets{ON_Omega2/Merge, ON_Omega2/Mergeb}
-
-\subsection{Yet another  proof of impossibility}
-\label{omega2-case}
-
-In Sect.~\vref{omega-case}, we proved that there exists no variant from \texttt{Hydra} to \texttt{(nat,$<$)}
-(\emph{i.e.} the ordinal $\omega$) for proving the termination of all hydra battles.
-We  prove now that  the ordinal $\omega^2$ is also insufficient for this purpose.
-
-The proof we are going to comment has exactly the same structure as in Section~\ref{omega-case}.
- Nevertheless, the proof of technical  lemmas is a little more complex, due to 
- the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
-Consider for instance that there exists an infinite number of ordinals  between
-$\omega$ and $\omega\times 2$.
-
-The detailed  proof script is in the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/Hydra/Omega2_Small.v}{theories/ordinals/Hydra/Omega2\_Small.v}.
-
-\subsubsection{Preliminaries}
-Let us assume there is a variant from \texttt{Hydra} into $\omega^2$  for proving the   termination of all hydra battles.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
-
-\input{movies/snippets/Omega2_Small/Impossibility}
-\input{movies/snippets/Omega2_Small/Impossibilitya}
-
-We  follow the same pattern as in Sect.~\ref{omega-case}.
-First, we define an injection $\iota$ from type \texttt{ON\_Omega2.t} into \texttt{Hydra}, by
- associating to  each ordinal $\omega\times i+ j = (i,j)$ the hydra with $i$ branches of length $2$ and
-$j$ branches of length $1$.
-
-%% revenir ici
-
-\vspace{4pt}
-\emph{From Module ~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
-
-\input{movies/snippets/Omega2_Small/Impossibilityc}
-
-For instance, Figure~\vref{fig:essai2} shows the hydra associated to the ordinal 
-$(3,5)$, a.k.a. $\omega\times 3 + 5$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-\node (foot) at (6,0) {$\bullet$};
-\node (N1) at (1,3) {$\bullet$};
-\node (N2) at (3,3) {$\bullet$};
-\node (N3) at (5,3) {$\bullet$};
-\node (N4) at (8,3) {$\Smiley[2][green]$};
-\node (N5) at (11,3) {$\Smiley[2][green]$};
-\node (N6) at (14,3) {$\Smiley[2][green]$};
-\node (N7) at (17,3){$\Smiley[2][green]$};
-\node (N8) at (20,3){$\Smiley[2][green]$};
-\node  (N9) at (0,5) {$\Smiley[2][green]$};
-\node (N10) at (2,5) {$\Smiley[2][green]$};
-\node (N11) at (4,5) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N5);
-\draw (foot) -- (N6);
-\draw (foot) to [bend right=10] (N7);
-\draw (foot) to [bend right=15] (N8);
-\draw (N1) to [bend left=10] (N9);
-\draw (N2) -- (N10);
-\draw (N3) -- (N11);
-\end{tikzpicture}
-\caption{\label{fig:essai2}
-The hydra $\iota(\omega\times 3+5)$}
-\end{figure}
-
-Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:h-omega2-small}}
- The hydra \texttt{big\_h}.
-\end{figure}
-
-\input{movies/snippets/Omega2_Small/Impossibilityb}
-
- In a second step, we build a ``smaller'' hydra\footnote{With respect to the measure $m$.}.
-
-\input{movies/snippets/Omega2_Small/Impossibilityd}
-
-\vspace{4pt}
-
-Like in Sect.~\ref{omega-case}, we prove the inequality $m\;\texttt{big\_h} \,\texttt{o<=}\, m\;\texttt{small\_h} \,\texttt{o<=}\, m\;\texttt{big\_h}$, which is obviously false.
-
-\subsubsection{Proof of the inequality \texttt{m small\_h o< m big\_h}}
-
-In order to prove the inequality  \texttt{m\_lt: m small\_h o< m big\_h}, it suffices to
-build a battle transforming \texttt{big\_h} into \texttt{small\_h}.
-
-First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h} as $(i,j)$.
-If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
-If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the following result.
-
-\input{movies/snippets/Omega2_Small/bigToSmall}
-
-Since $m$ is a variant, we infer the following inequality:
-
-\input{movies/snippets/Omega2_Small/mLt}
-
-\subsubsection{Proof of the inequality \texttt{m big\_h o<= m small\_h} }
-
-The proof of the inequality \texttt{m big\_h o<= m small\_h} is quite more complex than in Sect~\ref{omega-case}.
-If we consider any ordinal $\alpha=(i,j)$, where $i>0$, there exists an infinite number of
-ordinals strictly less than $\alpha$, and there exists an infinite number of battles that start from
-$\iota(\alpha)$.
-Indeed, at any configuration $\iota(k,0)$, where $k>0$, the hydra can freely choose any replication number.
-Intuitively, the measure of such a hydra must be large enough for taking into account
-all the possible battles issued from that hydra.
-Let us now give more technical details.
-
-The first steps of our proof start a well-founded induction over $\omega^2$.
-
-\input{movies/snippets/Omega2_Small/mGe}
-
-A case analysis on $i$ and $j$ allows us to
-consider three cases :
-
-\begin{itemize}
-\item $i=j=0$: the inequality is trivial.
-\item $i=1+l, j=0$ ($(i,j)$ is a limit ordinal): By the induction hypothesis \texttt{IHij},
-  $(l,k)\texttt{o<=} m(\iota(l,k))$ for any $k$. But (by the rules of the hydra game), $\iota(i,0)$ is transformed into any $\iota(l,k)$ in one round. Thus $m(\iota(l,k)) < m(\iota(i,0))$ for any $k$.
-  Therefore, $(l,k) <  m(\iota(i,0))$ for any $k$, thus
-  $(i,0)\,\texttt{o<=}\,m(\iota(i,0))$.
- \item $j= l+1$  ($(i,j)$ is a successor). We apply the induction hypothesis to the pair $(i,l)$.
- \end{itemize}
-
- Please look at the proof script for more details.
-
- \input{movies/snippets/Omega2_Small/mGeb}
-
-\subsubsection{End of the proof}
-From \texttt{m\_ge}, we get \texttt{$m$\, big\_h o<= $m$\, small\_h = $m$\,(iota (m big\_h)) }. 
-Since $<$ is a strict order (irreflexive  and transitive), this inequality is incompatible with the strict inequality  \texttt{m small\_h o< m big\_h} (lemma \texttt{m\_lt}).
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#Impossible}{Hydra.Omega2\_Small}}
-
-  \input{movies/snippets/Omega2_Small/Impossible}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into $\omega^2$ for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\begin{remark}
- We  prove in Chapter~\ref{ks-chapter} a generalization of the impossibility lemmas of
-Sect.~\ref{omega-case} and this section, with the same proof structure, but with much more 
-complex technical details.
- \end{remark}
-
-% \index{Exercises}
-% \begin{exercise}
-
-% \label{sec:orgheadline63}
-% Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.
-
-% \begin{itemize}
-% \item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
-%   lexicographic ordering
-% \item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.
-
-% For instance, the following inequality holds:
-% \[\langle 4,3,3,3,3,3,3,2,2,2 \rangle\,<\,\langle 4,4,2 \rangle\]
-% \end{itemize}
-
-% \end{exercise}
-
-\section{A notation for finite ordinals}
-
-Let $n$ be some natural number. The segment associated with $n$ is the interval 
-$[0,n)\,=\,\{0,1,\dots,n-1\}$. 
-One may represent the ordinal $n$ by a sigma type.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
-
-\label{def:Finite-ord-type}
-
-\input{movies/snippets/ON_Finite/Defs}
-
-\subsubsection{Examples}
-
-For instance, let us build two elements of the segment $[0, 7)$, \emph{i.e.} two
-inhabitants of   type (\texttt{t 7}), and prove a simple  inequality (see Fig.~\ref{fig:O7}).
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (N0) at (0,0) {$\bullet$};
-\node (i0) at (0,1) {$0$};
-\node (N1) at (2,0) {$\bullet$};
-\node (i1) at (2,1) {$1$};
-\node (N2) at (4,0) {$\bullet$};
-\node (i2) at (4,1) {$2$};
-\node (N3) at (6,0) {$\bullet$};
-\node (i3) at (6,1) {$3$};
-\node (N4) at (8,0) {$\bullet$};
-\node (i4) at (8,1) {$4$};
-\node (N5) at (10,0) {$\bullet$};
-\node (i5) at (10,1) {$5$};
-\node (N6) at (12,0) {$\bullet$};
-\node (i6) at (12,1) {$6$};
-\node(alpha1) at (4,-1) {$\alpha_1$};
-\node(alpha2) at (10,-1) {$\beta_1$};
-\end{tikzpicture}
-
-\caption{The segment $\mathbb{O}_7$\label{fig:O7}}
-\end{figure}
-
-%\index{coq}{Commands!Program}
-
-\input{movies/snippets/ON_Finite/Example1}
-
-Note that the type (\texttt{t 0}) is empty, and that, for any natural number
- $n$, $n$ does not belong to (\texttt{t $n$}).
-
- \inputsnippets{ON_Finite/t0Empty, ON_Finite/bad10}
-
-\subsubsection{Comparison function}
-
-In order to build an instance of \texttt{ON}, we define a comparison function,  and prove its correctiness.
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Finite/compareDef}
-
-\begin{remark}
- The proof of \texttt{compare\_correct} uses a well-known pattern of \coq{}.
-Let us consider  the following subgoal.
-
-\input{movies/snippets/ON_Finite/compareCorrectb}
-
-Applying the tactic \texttt{f\_equal} generates a simpler subgoal.
-
-\input{movies/snippets/ON_Finite/compareCorrectc}
-
-We have now to prove that there exists at most one  proof of (\texttt{Nat.ltb x0 (S n)}). This is not obvious, but  a consequence of the following lemma of library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Eqdep_dec.html}{Coq.Logic.Eqdep\_dec}.
-
-\index{Coq}{Unicity of equality proofs}
-\label{sect:eq-proof-unicity}
-
-\begin{Coqanswer}
-eq_proofs_unicity_on :
-forall (A : Type) (x : A),
-(forall y : A, x = y \/ x <> y) -> 
-forall (y : A) (p1 p2 : x = y), p1 = p2
-\end{Coqanswer}
-
-Thus unicity of proofs of \texttt{Nat.ltb x0 (S n)}  comes from the decidability of
-equality on type \texttt{bool}.
-This is why we used the boolean function \texttt{Nat.ltb} instead of the inductive predicate \texttt{Nat.lt} in the definition of type \texttt{t $n$} (see page~\pageref{def: Finite-ord-type}).
-For more information about this pattern, please look at the numerous mailing lists and 
-FAQs on \coq{}).
-
-\inputsnippets{ON_Finite/compareCorrectd}
-
-\end{remark}
-
-\begin{remark}
-
-Please note  that attempting to compare a term  of type (\texttt{t $n$}) with a term of
-type (\texttt{t $p$})  leads to an error if $n$ and $p$ are not convertible.
-
-\input{movies/snippets/ON_Finite/Example2}
-
-\end{remark}
-
-\subsubsection{Building an instance of \texttt{ON}}
-
-Applying lemmas of the libraries \texttt{Coq.Wellfounded.Inverse\_Image}, \linebreak
- \texttt{Coq.Wellfounded.Inclusion}, and \texttt{Coq.Arith.Wf\_nat}, we prove that our
-relation \texttt{lt} is well founded.
-
-\input{movies/snippets/ON_Finite/ltWf}
-
-Now we can build our instance of \texttt{OrdinalNotation}.
-
-\input{movies/snippets/ON_Finite/ONInstance}
-
-\begin{remark}
-It is important to keep in mind  that the integer $n$ is not an ``element'' of \texttt{FinOrd $n$}.
-In set-theoretic presentations of ordinals, the set associated with the ordinal $n$ is $\{0,1,\dots,n-1\}$.
-In our formalization, the interpretation of an ordinal as a set is realized by the function \texttt{bigO} (see Section\vref{sect:bigO-ON}).
-\end{remark}
-
-\begin{remark}
- There is no interesting arithmetic on finite ordinals, since functions like successor, addition, etc.,  cannot be represented in \coq{} as \emph{total} functions.
-\end{remark}
-
-\section{See also \dots}
-
-Finite ordinals are also formalized in \ssreflect/\mathcomp~\cite{MCB}.
-In Module~\href{../theories/html/gaia_hydras.ON_gfinite.html}{gaia\_hydras.ON\_gfinite}, we build an instance of class \texttt{ON}.
-
-\inputsnippets{ON_gfinite/finordLtDef, ON_gfinite/finordON}
-
-\inputsnippets{ON_gfinite/Examples, ON_gfinite/Examplesb}
-
-  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.
-
-%%%
-
-\section{Comparing two ordinal notations}
-
-It is sometimes useful to compare two ordinal notations with respect to expressive power
-(the segment of ordinals  they represent).
-
-The following class specifies a strict inclusion of segments.
-The notation \texttt{OA} describes a segment $[0,\alpha)$, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented in \texttt{OA}).
-We require also  that the comparison functions of the two notation systems are compatible.
-
-\begin{figure}[h]
-   \centering
-   \begin{tikzpicture}[very thick, scale=0.6]
-\begin{scope}[color=blue]
-\node (A) at (0,0) {$A$};
-\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
-\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
-\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
-\node (Adots) at (6,0) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red]
-\node (B) at (0,2) {$B$};
-\node(B0) at (2,2)[label=above:$0$]{$\bullet$};
-\node(B1) at (3,2)[label=above:$1$]{$\bullet$};
-\node(B2) at (4,2)[label=above:$2$]{$\bullet$};
-\node (Bdots) at (6,2) {$\ldots$};
-\node (b) at (8,2) [label=above:$\omega$]{$\bullet$};
-\node (bsucc) at (9,2) [label=above:$\omega+1$]{$\bullet$};
-\node (Bdots2) at (10,2) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red!50!blue]
-\draw [->,thin] (A0) -- node [auto] {$\iota$} (B0);
-\draw [->,thin] (A1) -- node [auto] {$\iota$} (B1);
-\draw [->,thin] (A2) -- node [auto] {$\iota$} (B2);
-\draw [->,thin] (Adots) -- node [auto] {$\iota$} (Bdots);
-\end{scope}
-\end{tikzpicture}
-   \caption{\textcolor{blue}{$A$} is a sub-segment  of \textcolor{red}{$B$}}
-   \label{fig:subsegment}
- \end{figure}
-
-If \texttt{OB} is presumed to be correct, then we may consider that \texttt{OA} ``inherits'' its correctness from the bigger notation system \texttt{OB}.
-
-\label{types:SubON}
-\index{hydras}{Library OrdinalNotations!Type classes!SubON}
-
-following definition
-(in ~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{ON\_Generic}).
-
-\input{movies/snippets/ON_Generic/SubONDef}
-
-For instance, we prove that \texttt{Omega} is a sub-notation of
-\texttt{Omega\_plus\_Omega} (with $\omega$ as the first ``new'' ordinal, and \texttt{fin} as the injection).
-
-\vspace{10pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/Incl}
-
-We can also show that, if $i<j$, then the segment $[0,i)$ is a ``sub-segment'' of
-$[0,j)$. Since the terms  ($t\;i$) and ($t\;j$) are not convertible, we consider a ``cast'' 
-function $\iota$ from ($t\;i$) into ($t\;j$), and prove that this function is  a monotonous bijection  from ($t\;i$) to
-the segment $[0,i)$ of ($t\;j$).
-
-%\index{coq}{Commands!Program}
-
-\vspace{10pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
-
-\inputsnippets{ON_Finite/InclIJ, ON_Finite/InclIJa,
-  ON_Finite/InclIJb,ON_Finite/InclIJc, ON_Finite/InclIJd,
-ON_Finite/InclIJz}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove that \texttt{Omega\_plus\_Omega} cannot be a sub-notation of \texttt{Omega}.
-\end{exercise}
-
-\index{hydras}{Projects}
-\begin{project}
-Adapt the definition of \texttt{Hvariant} (Sect.~\ref{sect:hvariant-def}) in order to
-have an ordinal notation as argument. Prove that if $O_A$ is a sub-notation of $O_B$, then any variant defined on  $O_A$ can be automatically transformed into 
-a variant on $O_B$.
-\end{project}
-
-\section{Comparing an ordinal notation with Schütte's model}
-
-Finally, it may be interesting to compare an ordinal notation with the more theoretical model from Schütte (well, at least with our formalization of that model).
-This would be a relative proof of correctness of the considered  ordinal  notation.
-
-The following class specifies that a notation \texttt{OA} describes a segment $[0,\alpha)$,
-where $\alpha$ is a countable ordinal \emph{à la}  Schütte.
-
-\label{types:ON-for}
-\index{hydras}{Library OrdinalNotations!Type classes!ON\_correct}
-
-\input{movies/snippets/ON_Generic/ONCorrect}
-
-For instance, the following theorem tells that \texttt{Epsilon0}, our notation system for the segment $[0,\epsilon0)$ is a correct implementation of the theoretically defined  ordinal $\epsilon_0$
-(see chapter~\ref{chap:schutte} for more details).
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
-
-\input{movies/snippets/Correctness_E0/Epsilon0Correct}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Same work, but replace Schütte's model with \gaia's.
-\end{project}
-
-\begin{project}
-  When you have read Chapter~\ref{chap:schutte}, prove that the sum of two ordinal notations \texttt{ON\_plus} implements the addition of ordinals.
-\end{project}
-
-\section{Isomorphism of ordinal notations}
-
-In some cases we want to show that two notation systems describe the same segment (for instance $[0,3+\omega)$ and $[0,\omega)$\;).
-For this purpose, one may prove that the two notation systems are order-isomorphic.
-
-\index{hydras}{Library OrdinalNotations!Type classes!ON\_Iso}
-
-\label{types:ON-iso}
-
-\input{movies/snippets/ON_Generic/ONIso}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-\label{exo:i-plus-omega}
-Let $i$ be some natural number. Prove that the notation systems 
-\texttt{Omega} and (\texttt{ON\_plus (OrdFin $i$) Omega}) are isomorphic.
-
-{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$, that we prove also in larger notation systems, as well as in Schütte's model.}
-This exercise is partially solved for $i=3$ (in ~\href{../theories/html/hydras.OrdinalNotations.Example_3PlusOmega.html}{OrdinalNotations.Example\_3PlusOmega}).
-
-\end{exercise}
-
-\index{hydras}{Projects}
-\label{exo:ON-mult}
-\begin{project}
-% Define in \coq{} the product of two ordinal notations $N_A$ and $N_B$.
-% If $A$ [resp. $B$] is the underlying type of $N_A$ [resp. $N_B$], the
-% product \texttt{ON\_mult $N_A$ $N_B$} is implemented over the cartesian product $B\times A$ (with the lexicographic ordering).
-
-This exercise is about the non-commutativity of the multiplication of ordinals, reflected in ordinal notations.
-
-For instance, the
-elements of the product (\texttt{ON\_mult Omega (FinOrd 3)}) are ordered as follows.
-\[(0,0),(0,1),(0,2),(0,3),(0,4),\dots,{\color{red}(1,0),} (1,1),(1,2),\dots, {\color{red}(2,0)},(2,1),(2,2),\dots\]
-
-Note that the elements of  (\texttt{ON\_mult (FinOrd 3) Omega}) are differently ordered (without limit ordinals):
-\[(0,0),(1,0),(2,0),(0,1),(1,1),(2,1),(0,2),(1,2),(2,2),(0,3),\dots\]
-
-Prove formally  that (\texttt{ON\_mult (FinOrd $i$) Omega}) is isomorphic to
-\texttt{Omega}  whilst
-\texttt{Omega}  is a sub-notation of (\texttt{ON\_mult Omega (FinOrd $i$)}),
-for any strictly positive $i$.
-
-\textbf{Note:} Like Exercise~\ref{exo:i-plus-omega}, this project corresponds to the [in]equalities $i+\omega=\omega<\omega+i$, for any natural number $i$.
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-Consider two isomorphic ordinal notations \texttt{OA} and \texttt{OB}.
-Prove that, if \texttt{OA} [resp. \texttt{OB}] is a correct implementation 
-of $\alpha$ [resp. $\beta$], then $\alpha=\beta$.
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-\label{project:succ-limit-dec}
-Add to the class \texttt{ON} the requirement that for any $\alpha$ it is decidable whether $\alpha$ is $0$, a successor or a limit ordinal.
-
-\textbf{Hint:}   Beware of the instances associated with sum and product of notations!
-  You may consider additional fields 
-to make the sum and product of notations ``compositional''.
-
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-\label{project:on-setoid}
-Reconsider the  class \texttt{ON}, with an equivalence instead of Leibniz equality.
-\end{project}
-
-%%%% ICI ICI
-
-\section{Other ordinal notations}
-%%% TODO : Fix the multiplication function in branch FixOmegaOmega
-
-% \index{hydras}{Projects}
-
-% \begin{project}
-% The directory \texttt{theories/OmegaOmega} contains an ad-hoc formalization of $\omega^\omega$, contributed by Pascal Manoury.
-Every ordinal $\alpha$ is represented by a list $l$ whose elements are the coefficients of $\omega$ in  the Cantor normal form of $\alpha$ (in reverse order).
-For instance, the ordinal
-% $\omega^{8}\times 5 + \omega^{6}\times 8 + \omega^2\times 10 + \omega + 7$ is represented by the list \texttt{[5;0;8;0;0.0;10,1,7]}.
-
-%  Develop this representation and compare it with the other ordinal notations.
-
-% \end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Let $N_A$ be a notation system for ordinals strictly less than $\alpha$, 
-with the strict order $(A,<_A)$. Please build the notation system
-\texttt{ON\_Expl $N_A$}, on the type of multisets of elements of $A$
-(or, if preferred, the type of non-increasing finite sequences on $A$,
-provided with the lexicographic ordering on lists).
-
-For instance, let us take $N_A=\texttt{Omega}$, and take $\alpha=\langle 4,4,2,1,0\rangle$,
- $\beta=\langle 4,3,3,3,3,3,2\rangle$, and $\gamma=\langle 5\rangle$. Then $\beta<\alpha<\gamma$.
-
-In contrast the list $\langle5,6,3,3\rangle$ is not non-increasing (\emph{i.e.} sorted w.r.t. $\geq$), so it is not to be considered.
-
-Note that if the notation $N_A$ implements the ordinal 
-$\alpha$,  the new notation $\omega^{N_A}$ must implement the ordinal $\phi_0(\alpha)$, a.k.a. $\omega^\alpha$ (see chapter~\ref{chap:schutte})
-
-\end{project}
-
-\begin{remark}
- The set of ordinal terms in Cantor normal form (see Chap.~\ref{chap:T1}) and 
-in Veblen normal form (see 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}{Gamma0.Gamma0}) are shown to be ordinal notation systems, but there is a lot of work to be done in order to unify ad-hoc  definitions and proofs which were written before the definition of the \texttt{ON} type class.
-\end{remark}
-
-\section{Ordinal Notations and \gaia's ordinals}
-\gaiasign Module~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge} contains an instance of class \texttt{ON E0lt compare}, where \texttt{E0lt} is the order on the well-formed ordinal terms below $\epsilon_0$ defined in \gaiaHydras library (please see Chapter~\ref{gaia-chapter}).
-
-%------------------------------------------------------------------------
-
-\chapter[A proof of termination, using epsilon0]{A proof of termination, using ordinals below \texorpdfstring{$\epsilon_0$}{Epsilon0}}
-
-\label{cnf-math-def}
-\label{chap:T1}
-
-In this chapter, we adapt to \coq{} the well-known~\cite{KP82}  proof that Hercules eventually wins every battle, whichever the strategy  of each player.
-In other words, we present  a formal and self contained proof of termination  of all [free] hydra battles.
-First, we take from Manolios and Vroon~\cite{Manolios2005} a representation of the ordinal $\epsilon_0$ as terms in Cantor normal form.
-Then, we define a variant for hydra battles as a measure that maps any hydra to some ordinal strictly less than $\epsilon_0$.
-
-\section{The ordinal \texorpdfstring{\(\epsilon_0\)}{epsilon0}}
-\label{sec:epsilon0-intro}
-
-The ordinal \(\epsilon_0\) is the least ordinal number that satisfies 
-the equation \(\alpha = \omega^\alpha\), where \(\omega\) is 
-the least infinite ordinal.
-Thus, we can intuitively consider \(\epsilon_0\) as an
-\emph{infinite} \(\omega\)-tower.
-
-\subsection{Cantor normal form}
-\index{maths}{Cantor normal form}
-
-Any ordinal strictly less that \(\epsilon_0\) 
-can be finitely represented by a unique  \emph{Cantor normal form}, 
-that is, an expression  which is either  the ordinal \(0\) or 
-a sum  \(\omega^{\alpha_1} \times n_1 + \omega^{\alpha_2} \times n_2 + 
-  \dots + \omega^{\alpha_p} \times n_p\) where all the \(\alpha_i\) 
-are ordinals in Cantor  normal form, \(\alpha_1 > \alpha_2 > \alpha_p\), 
-and all the \(n_i\) are positive integers.
-
-An example of Cantor normal form is displayed in Fig \ref{fig:cnf-example}:
-Note that  any ordinal of
-the form \(\omega^0 \times i + 0\) is just written \(i\).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
-\node[color=blue]{$\omega^{(\omega^\omega\,+\, \omega^2 \times 8 \,+\, \omega)}+ \omega^\omega + \omega^4+ 6$};
-\end{tikzpicture}
-\caption{\label{fig:cnf-example}
-An ordinal in Cantor normal form}
-\end{figure}
-
-In the rest of this section, we define an inductive type for representing in \texttt{Coq}
-all the ordinals strictly  less than  \(\epsilon_0\), then extend some arithmetic operations
-to this type, and finally prove that our representation fits well with 
-the expected mathematical properties: the order we define is a well order, 
-and the decomposition into Cantor normal form  is consistent 
-with the implementation of the arithmetic operations of exponentiation of base \(\omega\) 
-and addition.
-
-\paragraph*{Remark}
-\label{sec:orgheadline65}
-Unless explicitly mentioned, the term ``ordinal" will be used instead of
-``ordinal strictly less than \(\epsilon_0\)" (except in Chapter~\ref{chap:schutte} where it stands for ``countable ordinal'').
-
-\subsection{A data type for  ordinals in Cantor normal form}
-\label{sec:orgheadline72}
-\label{sec:T1-inductive-def}
-
-% Our user contribution~\cite{CantorContrib} represents 
-% the set of ordinals strictly less than $\epsilon_0$ in Cantor normal form as in~\cite{Manolios2005}, and also the set
-% of ordinals strictly  less than $\Gamma_0$ in Veblen normal form.
-
-    Let us define an inductive type whose 
-constructors are respectively associated
-with the ways to build Cantor normal forms:
-
-\begin{itemize}
-\item the ordinal \(0\)
-\item the construction \((\alpha,\, n,\,\beta)  \mapsto \omega^\alpha \times (n + 1)+ \beta \quad (n\in\mathbb{N})\)
-\end{itemize}
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#T1}{Epsilon0.T1}}
-
-\label{types:T1}
-%\index{Constants!zero:T1}
-\index{hydras}{Library Epsilon0!Types!T1}
-
-\input{movies/snippets/T1/T1Def}
-
-\subsubsection{Example}
-
-\label{alpha0-def}
-For instance, the ordinal  $\omega^\omega+\omega^3\times 5+2$ is represented by the following term of type \texttt{T1}:
-
-\input{movies/snippets/T1/alpha0}
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5, level 1/.style={sibling distance=6cm},
-level 2/.style={sibling distance=35mm},  
-level 3/.style={sibling distance=17mm}]
-\node  {cons}
-  child {  node {cons}
-            child { node {cons} child {node {zero}} child {node{0}} child{node{zero}}}
-         child {node {0}}
-         child {node {zero}}}
-    child {node {0}}
-   child {node {cons} 
- child { node {cons} child {node {zero}} child {node{2}} child{node{zero}}}
-  child {node {4}}
-         child {node {cons} child {node {zero}} child {node{1}} child{node{zero}}}};
-
-\end{tikzpicture}
-
-\caption{The tree-like representation of the ordinal $\omega^\omega+\omega^3\times 5 +2$\label{fig:cnf-tree}}
-
-\end{figure}
-
-\paragraph{Remark}
-For simplicity's sake, we chose to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the construction (\texttt{cons $\alpha$ $n$ $\beta$}) is intended to represent the
-ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
-In a future version, we should replace  the type \texttt{nat} with \texttt{positive} in \texttt{T1}'s 
-definition. But this replacement would take a lot of time \dots{}
-
-\paragraph{Remark}
-The name \texttt{T1} we gave to this data-type  is proper to this development and refers
-to a hierarchy of ordinal notations.
-For instance, in Library \href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2},  the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{Gaia, schutte}).
-\noindent
-
-\input{movies/snippets/T2/T2Def}
-
-\subsection{About Gaia}
-\gaiasign Chapter~\vref{gaia-chapter} describes the present state of a
-project of making compatible the libraries \HydrasLib and \gaia. At present, both libraries contain their own version of the inductive type \texttt{T1}, with the same base name for the constructors.
-
-\inputsnippets{nfwfgaia/T1def}
-
-Thus, many examples of this and following chapters can be run
-in \gaia's context. We will signal any difference (notation, name) which may appear.
-A closer integration (same data type and functions) is still in project.
-
-\begin{remark}
-  \label{remark:hydra-gaia-renaming}
-  \textcolor{lookcolor}{In future releases, we plan to make some
-    \HydrasLib identifiers progressively  deprecated, in favour of \gaia's names.}
-\end{remark}
-
-\subsection{Abbreviations}
-
-Some abbreviations may help to write more concisely complex ordinal terms.
-
-\subsubsection{Finite ordinals}
-\label{sec:orgheadline67}
-
-For representing finite ordinals, \emph{i.e.} natural numbers, we first introduce a notation for terms of the form $n+1$, then define a coercion from type \texttt{nat} into \texttt{T1}.
-\label{sect:notation-FS}
-
-\label{sect:notation-F}
-
-\input{movies/snippets/T1/finiteOrds}
-
-\paragraph*{\gaiasign}
-  In \texttt{gaia.ssete9}, the $n$-th finite ordinal is also written
-  \texttt{$\backslash${F}\,$n$}. There is no coercion in \gaia from \texttt{nat} to \texttt{T1}.
-
-  \vspace{4pt}
-
-  From~\href{../theories/html/gaia_hydras.HydraGaia_Examples.html}{gaia\_hydras.HydraGaia\_Examples}.
-
-  \inputsnippets{HydraGaia_Examples/Demoa}
-
-% \index{Coq!Coercions}
-% \index{Functions!Coercions@Coercions (from nat to ordinal types)}
-% \begin{remark}
-% Please refer to the remark~\pageref{warning:coercions} about the use of coercions.
-% % The use of coercions like \texttt{fin} allow us to be close to the mathematical tradition where natural numbers are ordinals too.
-% % Nevertheless, it may happen that a goal like \texttt{3 < 5} could be 
-% % interpreted as \texttt{(lt (fin 3) (fin 5))},  depending on the current notation scope.  
-% % When this misinterpretation happens, tactics like \texttt{auto with arith}, \texttt{lia} do not work!
-% % Thus, it is useful to write \texttt{(3 < 5)\%nat}  an inequality between two natural numbers. 
-% \end{remark}
-
-\subsubsection{The ordinal \(\omega\)}
-\label{sec:orgheadline68}
-
-  Since \(\omega\)'s Cantor normal form is
-i.e. \(\omega^{\omega^0}\times 1+ 0\), we can define the following abbreviation:
-
-\label{sect:omega-notation2}
-
-\input{movies/snippets/T1/omegaDef}
-
-Note that \texttt{omega} is not an identifier in \HydrasLib, thus any tactic like \texttt{unfold omega} would fail.
-
-\paragraph*{\gaiasign}
-In \texttt{gaia.ssete9}, the ordinal $\omega$ is bound to the \emph{constant}  \texttt{T1omega}.
-
-\subsubsection{The ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
-\label{sect:notation-phi0}
-We provide also a notation for ordinals of the form $\omega^\alpha$.
-
-\index{hydras}{Library Epsilon0!Notations!phi0@phi0 (exponential of base omega)}
-
-\input{movies/snippets/T1/phi0Def}
-
-\index{maths}{Additive principal ordinals}
-
-\begin{remark}
-\label{sec:orgheadline69}
-The name \(\phi_0\)
-   comes from ordinal numbers theory. In~\cite{schutte}, Schütte defines 
-$\phi_0$  as the ordering (\emph{i.e.} enumerating) function of the set  of \emph{additive principal ordinals} \emph{i.e.} strictly positive ordinals $\alpha$ that verify $\forall \beta<\alpha, \beta+\alpha=\alpha$.
-For Schütte,  $\omega^\alpha$ is just a notation for $\phi_0(\alpha)$.
-See also Chapter~\ref{chap:schutte} of this document.
-\end{remark}
-
-\paragraph*{\gaiasign}
-In \texttt{gaia.ssete9} the identifier \texttt{phi0} is bound to a plain constant (not a notation).
-
-\subsubsection{The hierarchy of \(\omega\)-towers:}
-\label{sec:orgheadline71}
-
-The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximated by the sequence of $\omega$-towers (see also Sect~\vref{sect:epsilon0-as-limit} ).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/towerDef}
-
-For instance, Figure~\ref{fig:tower7} represents  the ordinal returned by the
- evaluation of the term \texttt{omega\_tower 7}.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
-\node[color=blue]{$\omega^{{{\omega}^{{{\omega}}^{{{\omega}}^{{\omega^{{\omega}^{\omega}}}}}}}}$};
-\end{tikzpicture}
-\caption{\label{fig:tower7}
-The $\omega$-tower of height 7}
-\end{figure}
-
-\subsection{Pretty-printing ordinals in Cantor normal form}
-\label{sect:ppT1}
-\index{hydras}{Library Epsilon0!Types!ppT1}
-
-Let us consider again the ordinal $\alpha_0$ defined in section~\vref{alpha0-def}. 
-If we ask \coq{} to print its  normal form, we get a hardly readable term of type \texttt{T1}.
-
-\input{movies/snippets/T1/alpha0Compute}
-
-The following data type defines a more readable abstract syntax for  ordinals terms in Cantor normal form:
-
-\label{types:ppT1}
-\index{hydras}{Library Epsilon0!Functions!pp@ pp (pretty printing terms in Cantor normal form)}
-
-\input{movies/snippets/T1/ppT1Def}
-
-The function \texttt{pp: T1 -> ppT1} converts any closed term of type \texttt{T1} into a human-readable expression. For instance, let us convert the term \texttt{alpha\_0}.
-
-\input{movies/snippets/T1/ppAlpha0}
-
-\paragraph*{\gaiasign}
-In ~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge},
-we define a variant of \texttt{pp} for pretty-printing terms of type \texttt{ssete9.CantorOrdinal.T1} (see Sect.~\vref{sect:gaia-ppT1}).
-
-\index{hydras}{Projects}
-\begin{project}
-Design  (in \ocaml?) a set of tools for systematically pretty printing ordinal terms in Cantor normal form.
-\end{project}
-
-\subsection{Comparison between ordinal terms}
-\label{sec:orgheadline73}
-
-% Our formalisation of Cantor Normal Form will take two steps:
-% 1 Definition of a strict order \texttt{o<} on the type \texttt{T1}, 
-% 2 Using \texttt{o<} for characterizing terms in normal form.
-
-In order to compare two terms of type \texttt{T1}, we define a recursive function \texttt{compare} that maps two ordinal tems $\alpha$ and $\beta$ to a value of type \texttt{comparison}.
-This type is defined in \coq's standard library
-\texttt{Init.Datatypes} and
-contains three constructors:  \texttt{Lt} (less than), \texttt{Eq} (equal), and
-\texttt{Gt} (greater than).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/compareDef}
-
-\label{Predicates:lt-T1}
-Please note that this definition of \texttt{lt} makes it easy to write proofs by computation, as shown by the following examples.
-
-\vspace{4pt}
-
-\input{movies/snippets/T1/ltExamples}
-
-Links between the function \texttt{compare} and the relations
-\texttt{lt} and \texttt{eq} are established through the following lemmas (~\vref{sect:comparable-def}).
-\vspace{4pt}
-
-\input{movies/snippets/T1/Instances}
-
-\paragraph*{\gaiasign}
-In \gaia, the strict order \texttt{T1lt} on ordinal terms is directly defined as a boolean function (in \ssreflect's style).
-
-\inputsnippets{nfwfgaia/T1ltDef}
-
-In section~\vref{sect:lt-compat-gaia}, we show that both definitions are mutually equivalent.
-
-\subsubsection{A Predicate for Characterizing Normal Forms}
-\label{sect:t1-nf}
-
-\label{sec:orgheadline74}
-\label{sec:orgheadline75}
-Our data-type \texttt{T1} allows us to write expressions that
-are not properly in Cantor normal form as specified in Section \ref{sec:epsilon0-intro}.
-For instance, consider the following term of type  \texttt{T1}.
-
-\input{movies/snippets/T1/badTerm}
-
-This term would have been written \(\omega^1\times 2 + \omega^\omega \times 3\) in the usual mathematical notation.
-We note that the exponents of $\omega$ are not in the right (strictly decreasing) order.
-Nevertheless, with the help of the order \texttt{lt} on \texttt{T1}, we are now able to characterize
-the set of all well-formed ordinal terms:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#nf_b}{Epsilon0.T1}}
-
-\label{Predicates:nf-T1}
-
-\input{movies/snippets/T1/nfDef}
-
-\input{movies/snippets/T1/nfAlpha0}
-
-\input{movies/snippets/T1/nfBadTerm}
-
-\paragraph*{\gaiasign}
-In \gaia, the boolean function which characterizes ordinal terms in normal form is defined as follows:
-
-\inputsnippets{nfwfgaia/T1nfDef}
-
-In Sect.~\vref{nf-gaia-compat}, we show that \gaia's \texttt{T1nf} is extensionally equivalent with \HydrasLib'  \texttt{nf}.
-
-\subsection{Making normality implicit}
-  We would like to get rid of terms of type \texttt{T1} which are not in Cantor normal form.
-A simple way to do this is to consider statements of the form 
-\texttt{forall alpha: T1, nf alpha -> $P$ alpha}, where $P$ is a predicate over type \texttt{T1}, like in the following lemma \footnote{Ordinal addition is formally defined a little later (page~\ref{sect:infix-plus-T1})}.
-
-\input{movies/snippets/T1/plusIsZero}
-
-\vspace{4pt}
-
-But this style leads to clumsy statements, and generates too many subgoals in interactive proofs (although often solved with \texttt{auto} or \texttt{eauto}).
-
-One may encapsulate conditions of the form \texttt{(nf $\alpha$)} in
-the most used predicates. For instance, we introduce the restriction of \texttt{lt} to terms in normal form, and provide a handy notation for this restriction.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Prelude.Restriction.html}{hydras.Prelude.Restriction}}
-
-\input{movies/snippets/Restriction/restrictionDef}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#LT}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/LTDef}
-
-\label{Predicates:LT-T1}
-
-For instance, in the following lemma, the condition that $\alpha$ is in normal form is included in the condition $\alpha< 1$.
-
-\input{movies/snippets/T1/LTOne}
-
-\subsubsection{\texttt{E0}: a sigma-type for \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-As we noticed in Sect.~\ref{sect:t1-nf}, the type \texttt{T1} is not a correct ordinal notation, since it contains terms that are not in Cantor normal form.
-In certain contexts (for instance in Sections~\ref{sect:L-equations}, \ref{sect:hardy},
-and \ref{sect:wainer}),  we need to define total recursive functions on well-formed ordinal terms less  than $\epsilon_0$, using the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777}.
- In order to define a type whose inhabitants represent just ordinals, we build a type gathering a term of type \texttt{T1} and a proof that this term is in normal form.
-
-\label{sect:E0-def}
-\label{types:E0}
-\index{hydras}{Library Epsilon0!Types!E0}
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html\#E0}{Epsilon0.E0}}
-
-\input{movies/snippets/E0/E0Def}
-
-Many constructs : types, predicates, functions, notations, etc., on type \texttt{T1} are adapted to \texttt{E0}.
-
-First, we declare a notation scope for \texttt{E0}, then we redefine the predicates of comparison.
-
-\input{movies/snippets/E0/E0Scope}
-\label{Predicates:Lt-E0}
-\input{movies/snippets/E0/LtLeDef}
-
-Equality in \texttt{E0} is just Leibniz equality. Note that, since \texttt{nf} is
-defined by a Boolean function, for  any term $\alpha:\texttt{T1}$, there exists at most one proof of \texttt{nf $\alpha$}, thus two ordinals of type \texttt{E0} are
-equal if and only if their projection to \texttt{T1} are equal (see also Sect.~\vref{sect:eq-proof-unicity}).
-
-\vspace{4pt}
-
-\index{coq}{Unicity of equality proofs}
-
-\inputsnippets{E0/nfProofUnicity}
-\inputsnippets{E0/E0EqIff}
-
-In order to  upgrade constants and functions from type \texttt{T1} to \texttt{E0}, we have to prove that
-the term they build is in normal form.
-For instance, let us represent the ordinals $0$ and $\omega$   as instances of the class \texttt{E0}.
-
-\vspace{4pt}
-
-\label{sect:omega-T1}
-
-\input{movies/snippets/E0/ZeroOmega}
-
-%\index{Constants!zero:T1}
-
-\paragraph*{\gaiasign}
- The library~\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines a \gaia-compatible type \texttt{E0}.
-
-\subsection{Syntactic definition of limit and successor ordinals}
-
-Pattern matching and structural recursion allow us to define boolean characterizations  of successor and limit ordinals.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/succbLimitb}
-
-The correctness of these definitions with respect to the mathematical notions of
-limit and successor ordinals is established through several lemmas. For instance,
-Lemma \texttt{canonS\_limit\_lub}, page~\pageref{lemma:canonS-limit}, shows that
-if $\alpha$ is (syntactically) a limit ordinal, then it is the least upper bound of
-a strictly increasing sequence of ordinals.
-
-\paragraph*{\gaiasign}
-In \gaia, the boolean functions associated with limit and successor ordinals are called \texttt{T1limit} and \texttt{T1is\_succ}.
-
- From~\href{../theories/html/gaia_hydras.HydraGaia_Examples.html}{gaia\_hydras.HydraGaia\_Examples}.
-
-\inputsnippets{HydraGaia_Examples/Demob}
-
-\subsection{Arithmetic on \texorpdfstring{$\epsilon_0$}{epsilon0}}
-\subsubsection{Successor}
-
-\index{hydras}{Library Epsilon0!Functions!succ}
-
-The successor of any ordinal $\alpha< \epsilon_0$ is defined by structural 
-recursion on its Cantor normal form.
-
-\label{Functions:succ-T1}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/succDef}
-
-The following lemma establishes the connection between the  function
-\texttt{succ} and the Boolean predicate \texttt{succb}.
-
-\vspace{4pt}
-
-\input{movies/snippets/T1/succbIff}
-
-\begin{exercise}[\gaiasign]
-  Look for the \gaia-theorem which corresponds to \texttt{succb\_iff}.
-\end{exercise}
-
-\subsubsection{Successor function on \texttt{E0}}
-
-The function \texttt{succ} on \texttt{T1} is extended to \texttt{E0} the following way:
-
- \emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
-
- \inputsnippets{E0/SuccOnE0}
-
-\index{hydras}{Exercises}
- \begin{exercise}
-Prove in \coq{} that for any ordinal $0< \alpha<\epsilon_0$, $\alpha$ is a limit if 
-and only if for all $\beta<\alpha$, the interval $[\beta,\alpha)$ is
-infinite.
-
-\emph{You may start this exercise with the file
-     \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/Limit_Infinity.v}{exercises/ordinals/Limit\_Infinity.v}.}
- \end{exercise}
-
-\subsubsection{Addition and multiplication}
-
-Ordinal addition and multiplication are also defined by structural recursion over the type \texttt{T1}. Please note that they use the \texttt{compare} function on some subterms of their arguments.
-
-\label{sect:infix-plus-T1}
-
-\input{movies/snippets/T1/plusDef}
-
-\input{movies/snippets/T1/multDef}
-
-\paragraph*{\gaiasign}
-We keep \gaia's base names for
-addition and multiplication of ordinal terms below $\epsilon_0$.
-Please refer to Sect.~\ref{sect:plus-mult-gaia-hydras} about compatibility of both arithmetics.
-
-\subsubsection{Examples}
-
-The following examples are instances of \emph{proofs by computation}. Please note that  addition and multiplication on \texttt{T1}
-are not commutative. Moreover,  both operations fail to be strictly monotonous in their first argument.
-
-\input{movies/snippets/T1/plusMultExamples}
-
-\input{movies/snippets/T1/notMono}
-
-The function \texttt{succ} is related with addition through the following lemma:
-
-\inputsnippets{T1/succIsPlusOne, T1/succIsPlusOnez}
-
-\subsubsection{Arithmetic on type \texttt{E0}}
-
- We define an addition in type \texttt{E0}, since the sum of two terms in normal form is in normal form too.
-
-\input{movies/snippets/T1/plusNf}
-
-\input{movies/snippets/E0/plusE0}
-\input{movies/snippets/E0/CheckPlus}
-
-\begin{remark}
-In all this development, two representations of ordinals co-exist: ordinal terms (type \texttt{T1}, notation scope \texttt{t1\_scope}, for reasoning on the tree-structure of Cantor normal forms), and ordinal terms \emph{known to be in normal form} (type \texttt{E0}, notation scope \texttt{E0\_scope}).
-Looking at the contexts displayed by \coq{} prevents you from any risk of confusion.
-\end{remark}
-
-%% To simplify !
-\index{hydras}{Exercises} 
-\begin{exercise}
-Prove that for any ordinal $\alpha:\texttt{E0}$, 
-$\omega\leq \alpha$ if and only if, for any natural number $i$,
-$i+\alpha=\alpha$.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/ge_omega_iff.v}{exercises/ordinals/ge\_omega\_iff.v}.}
-\end{exercise}
-
-\subsection{A proof by computation}
-\label{Ex42-E0}
-
-It is interesting to compare the following proof of the equality
-$\omega+42+\omega^2$ with the more theoretical proof in Sect~\vref{Ex42-schutte}.
-
-\inputsnippets{E0/Ex42}
-
-\section{Well-foundedness and transfinite induction}
-
-\index{maths}{Transfinite induction}
-
-\subsection{About  well-foundedness}
-\label{sec:T1Wf}
-\label{sec:orgheadline82}
-   In order to use \texttt{T1} for proving termination results,
-we need to prove that  our order \texttt{<} is well-founded. Then we will get \emph{transfinite induction} for free.
-
-The proof of well-foundedness of the strict order $<$ on Cantor normal forms is already 
-available in the Cantor contribution by Castéran and Contejean~\cite{CantorContrib}. That proof relies on a library on recursive path orderings written by
-E. Contejean. We present here  a direct proof of the same result, which does not require any knowledge on r.p.o.s.
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that the \emph{total} order \texttt{lt} on \texttt{T1} is not well-founded. 
-\textbf{Hint:}  You will have to build a counter-example with terms of type \texttt{T1}
-which are not in Cantor normal form.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/T1_ltNotWf.v}{exercises/ordinals/T1\_ltNotWf.v}.}
-\end{exercise}
-
-\subsubsection{A first attempt}
-\label{sec:orgheadline77}
-%\index{coq}{Well-founded induction}
-
-It is natural to try to prove by structural induction over \texttt{T1} 
-that every term in normal form is \texttt{LT}-accessible.
-
-Unfortunately, it won't work. Let us consider some well-formed term
- $\alpha=\texttt{cons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are \texttt{LT}-accessible.
-In order to prove the accessibility of $\alpha$, we have to consider
-any well formed term \(\delta\) such that \(\delta<\alpha\). 
-But nothing guarantees that \(\delta\)  is strictly  less than \(\beta\) nor \(\gamma\), and we cannot use the induction hypotheses on   \(\beta\) nor \(\gamma\).
-
-\input{movies/snippets/T1/wfLTBada}
-
-The problem comes from the too weak hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
-\(\gamma\);
-for instance \(\delta\) may be of the form
-\texttt{cons $\beta$ $p'$  $\gamma'$},
-where    \(p' < n\).
-Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.
-
-\input{movies/snippets/T1/wfLTBadz}
-
-\subsubsection{Using a stronger inductive predicate.}
-\label{sec:orgheadline78}
-\label{sec:strongly-accessible}
-
-  Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
-through \texttt{LT}, we propose to consider the following (stronger) predicate:
-
-\input{movies/snippets/T1/AccStrongDef}
-
-The following lemma is an application of the strict inequality 
-$\alpha < \omega ^\alpha$. If $\alpha$ is strongly accessible, then, by definition,
-$\omega^\alpha$ is accessible, thus $\alpha$ is \emph{a fortiori} accessible.
-
-\input{movies/snippets/T1/AccStrongStronger}
-
-Thus, it remains to prove that every ordinal strictly less than $\epsilon_0$
-is strongly accessible.
-
-% \subsubsection{Structure of the proof of well-foundedness of \texttt{LT}}
-
-\label{sec:orgheadline81}
-\label{proof-wf-epsilon0}
-\paragraph{A helper}
-\label{sec:orgheadline79}
-
-First, we prove that, for  any \texttt{LT}-accessible term $\alpha$, $\alpha$ is 
-strongly accessible too (\emph{i.e.} any well formed
-term (\texttt{cons $\alpha$ $n$ $\beta$})  is accessible).
-
-The proof is structured as an induction on $\alpha'$s accessibility. Let us consider
-any  accessible term $\alpha$.
-
-\input{movies/snippets/T1/AccImpAccStrong}
-
-Let \texttt{n:nat} and \texttt{beta:T1} such that (\texttt{cons alpha n beta}) is in normal form. 
-We prove first that \texttt{beta} is accessible,  which allows us to prove by well-founded induction on \texttt{beta}, 
-and natural induction on \texttt{n}, that (\texttt{cons alpha n beta}) is accessible.
-The proof, quite long, can be consulted in \href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}.
-
-\paragraph{Accessibility of any well-formed ordinal term}
-\label{sec:orgheadline80}
-
-Our goal is still to prove accessibility of any well formed ordinal term.
-Thanks to our previous lemmas, we are almost done (by a simple structural induction!).
-
-\input{movies/snippets/T1/nfAcc}
-
-\input{movies/snippets/T1/T1Wf}
-
-\index{maths}{Transfinite induction}
-
-Traditionnally, well-founded induction on ordinals is called \emph{transfinite induction}.
-
-\inputsnippets{T1/transfiniteRecursor, T1/transfiniteInduction}
-
-As a corollary, the order \texttt{Lt} on type \texttt{E0} is well-founded too.
-
-\inputsnippets{E0/E0LtWf}
-
-\begin{remark}
-\label{remark:a3pat}
-The alternate proof of well-foundedness using \'Evelyne Contejean's work on
-    recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the
-    library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0.Epsilon0rpo}.
- \end{remark}
-
-\subsection{An ordinal notation for  \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-We build an instance of \texttt{ON}, and prove its correctness w.r.t. Schutte's model (see Chapter~\ref{chap:schutte}).
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
-
-\input{movies/snippets/E0/InstanceEpsilon0}
-
-\label{instance-epsilon0}
-
-\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
-
-\input{movies/snippets/Correctness_E0/injectDef}
-
-\input{movies/snippets/Correctness_E0/Epsilon0Correct}
-
-\index{hydras}{Projects}
-\begin{project}
- \emph{This exercise is a continuation of Project~\vref{exo:ON-mult}.}
-Use \texttt{ON\_mult} to define an ordinal notation \texttt{Omega2} for $\omega^2=\omega\times\omega$.
-
-Prove that \texttt{Omega2} is a sub-notation of \texttt{Epsilon0}.
-
-Define on \texttt{Omega2} an addition compatible with the addition on \texttt{Epsilon0}.
-
-\textbf{Hint}. You may use the following definition (in 
-    \href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}).
-
-\input{movies/snippets/ON_Generic/SubONSameOp}
-
-     \end{project}
-
-    \index{hydras}{Projects}
-    \begin{project}
-    The class \texttt{ON} of ordinal notations has been defined long after this 
-    chapter, and is not used in the development of the type \texttt{E0} yet.
-    A better integration of both notions should simplify the development on ordinals in Cantor normal form. This integration is planned for the future versions.
-
-    \end{project}
-
-    \section{A refinement of \texttt{E0} : an ordinal notation for \texorpdfstring{$\omega^\omega$}{omega\^omega}}
-
-    In Module   \href{https://github.com/coq-community/hydra-battles/blob/master/theories/
-      ordinals/OrdinalNotations/OmegaOmega.v}{theories/ordinals/OrdinalNotations/OmegaOmega.v},
-    we represent ordinals below $\omega^\omega$ by lists of pairs of natural numbers (with the same coefficient shift as in \texttt{T1}).
-    For instance, the ordinal $\omega^4\times 10 + \omega^3 + \omega+ 5$ is represented by the list \texttt{(4,9)::(3,0)::(1,0)::(0,4)::nil}.
-
-\inputsnippets{OmegaOmega/LODef}
-
-  The usual operations : \texttt{succ}, \texttt{+}, \texttt{*} are simple variants of the same operations in \texttt{T1}.
-
-  \inputsnippets{OmegaOmega/succPlusMult}
-
-    We establish this representation as a \emph{refinement} of the data types we used to represent ordinals less than $\epsilon_0$.
-Thus, many properties like well-foundedness of $<$ and associativity of $+$,  of this ordinal notations have very short proofs.
-
-    \inputsnippets{OmegaOmega/refineDef, OmegaOmega/phi0Ref,
-      OmegaOmega/succRef, OmegaOmega/plusRef,
-      OmegaOmega/multRef}
-
-    In order to make an ordinal notation for $\omega^\omega$, we follow the same steps as for $\epsilon_0$:
-    \begin{enumerate}
-    \item Define an order \texttt{lt}, which refines the order
-      \texttt{lt} on \texttt{T1}.
-      \inputsnippets{OmegaOmega/compareDef}
-      \inputsnippets{ OmegaOmega/ltDef, OmegaOmega/ltRef}
-
-    \item Define the predicate ``to be in normal form''.
-      \inputsnippets{OmegaOmega/nfDef}
-    \item Define a class \texttt{OO} of terms in normal form, and an embedding from \texttt{E0} into \texttt{OO}.
-      \inputsnippets{OmegaOmega/OODef}
-       \inputsnippets{OmegaOmega/embedDef}
-      \item Infer well-foundedness of the order on \texttt{OO}.
-    \inputsnippets{OmegaOmega/ltWf}
-
-    \end{enumerate}
-
-    \inputsnippets{OmegaOmega/ONOO, OmegaOmega/OOz}
-
-    Let us show a few examples.
-
-    \inputsnippets{OmegaOmega/OODemo}
-
-    \index{hydras}{Exercises}
-
-    \subsection{Related work}
-    The article~\cite{Manoury2016}  defines another representation of ordinals below $\omega^\omega$ based on lists of natural numbers.
-
-\begin{exercise}[not trivial]
-  It may be interesting to write a \emph{direct} proof  of well-foundedness of the order in $\omega^\omega$ (\emph{i.e.}
-  without using properties of $\epsilon_0$). This exercise may
-  help to understand better the proof structure of Sect.~\vref{sec:strongly-accessible}.
-    \end{exercise}
-%ici
-    \section{A variant for hydra battles}
-
-    In order to prove the termination of any hydra battle, we try to define a variant mapping hydras to ordinals strictly less than $\epsilon_0$.
-    In order to make such a variant easy to define (for instance by a structural recursion), we introduce a variant of addition, which, contrary to
-    $+$, is commutative and strictly monotonous in both of its arguments. This last property makes it possible to prove that our function is 
-    truly a variant for hydra battles (in Sect.~\vref{sect:variant-decr}).
-
-    \subsection{Natural sum (a.k.a. Hessenberg's  sum)}
-    \label{sec:orgheadline87}
-    \label{hydra-variant}
-    \label{sect:hessenberg-def}
-
-    Natural sum (Hessenberg's  sum) is a commutative and monotonous version of
-    addition. It is used as an auxiliary operation  for defining variants
-    for hydra battles, where Hercules is allowed to chop off any  head of the hydra.
-
-    In the literature, the natural sum of ordinals \(\alpha\) and \(\beta\)
-    is often denoted by \(\alpha \# \beta\)  or  \(\alpha \oplus  \beta\).
-    Thus we called \texttt{oplus} the associated \emph{Coq} function.
-
-    \subsubsection{Definition of \texttt{oplus}}
-    \label{sec:orgheadline84}
-    %\index{Functions!oplus @ oplus (Hessenberg commutative sum)}
-
-    The definition of \texttt{oplus} is recursive in both of its 
-    arguments and uses  the same pattern as for the \texttt{merge} function on lists of library
-    \texttt{Coq.Sorting.Mergesort}.
-
-    \begin{enumerate}
-    \item Define a nested recursive function, using the \texttt{Fix} 
-        construct
-
-    \item Build a principle of induction dedicated to \texttt{oplus}
-
-    \item Establish equations associated to each case of the definition.
-    \end{enumerate}
-
-    \paragraph{Nested recursive definition}
-    \label{sec:orgheadline83}
-
-    The following definition is composed of 
-    \begin{itemize}
-    \item A main function \texttt{oplus}, structurally recursive in its 
-    first argument \texttt{alpha}
-    \item An auxiliary function \texttt{oplus\_aux} within the scope of \texttt{alpha},
-    structurally recursive in its argument \texttt{beta};  \texttt{oplus\_aux beta} 
-       is supposed to compute  \texttt{oplus alpha beta}.
-    \end{itemize}
-
-%  \pagebreak 
-  \vspace{4pt}
-    \emph{From Module~\href{../theories/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}
-
-    \label{sect:infix-oplus}
-
-    \input{movies/snippets/Hessenberg/oplusDef}
-
-    The reader will note that each recursive call of the functions
-    \texttt{oplus} and \texttt{oplus\_aux} satisfies \emph{Coq}'s constraint
-    on recursive definitions. The function \texttt{oplus} is recursively called on a sub-term of its first argument,
-    and \texttt{oplus\_aux} on a sub-term of its unique argument.
-    Thus, \texttt{oplus}'s definition is accepted by \coq{} as a structurally recursive function.
-
-    \subsubsection{Rewriting lemmas}
-    \label{sec:orgheadline86}
-
-    \emph{Coq}'s constraints on recursive definitions result in 
-    the quite  complex form of \texttt{oplus}'s definition.
-    Proofs of properties of this function can be simpler if we
-     derive a few  rewriting lemmas that will help to simplify 
-    expressions of the form (\texttt{oplus $\alpha$ $\beta$}).
-
-    A first set of lemmas correspond to the various cases of \texttt{oplus}'s 
-    definition. They can be proved almost immediately. Here are a few examples.
-
-    \inputsnippets{Hessenberg/oplusNeutral,
-      Hessenberg/oplusCompareLt}
-
-    % \subsubsection{A hand-made induction principle}
-    % \label{sec:orgheadline85}
-
-    % \index{Coq!Commands!Functional Scheme}
-
-    % \emph{Coq} contains a command  \texttt{Functional Scheme} that 
-    % generates induction principles which correspond to recursive functions.
-    % Unfortunately, the current version ( \texttt{8.11.0} ) doesn't work on \texttt{oplus},
-    % probably because of the inner \texttt{Fix}.
-
-    % \begin{Coqsrc}
-    % Functional Scheme oplus_ind := Induction for oplus Sort Prop.
-    % \end{Coqsrc}
-
-    % \begin{Coqanswer}
-    % Error: Anomaly "todo." Please report at http://coq.inria.fr/bugs/.
-    % \end{Coqanswer}
-
-    % Fortunately, it's a good exercise for a semi-experienced user, to write
-    % her/him-self induction principles similar to the ones returned by
-    % \texttt{Functional Scheme}.
-
-    % \begin{itemize}
-    % \item First, we choose to write a version for sort \texttt{Type}, since versions
-    % for sorts \texttt{Prop} and \texttt{Set} can be easily derived from
-    % the former one. According to \emph{Coq}'s naming politics, we will call our 
-    % principle \texttt{oplus\_rect}
-
-    % \item The conclusion of \texttt{oplus\_rect} will be (\texttt{$P$ a b (oplus a b)}),
-    % where $P$ is an arbitrary function of type 
-    % \texttt{T1 -> T1 -> T1 -> Type}
-
-    % \item The premises of \texttt{oplus\_rect} will describe how to build an induction 
-    % on the graph of \texttt{oplus}.
-    % \end{itemize}
-
-    % We are now ready to state and prove \texttt{oplus\_rect}, and the reader
-    % will note that the statement is longer than the proof script itself,
-    % which is a standard proof by induction, simplification and case-analysis 
-    % that follows  \texttt{oplus}'s definition.
-
-    % We associate also a tactic to the application of \texttt{oplus\_rect}.
-
-    % \begin{Coqsrc}
-    %  Lemma oplus_rect:
-    %       forall P: T1 -> T1 -> T1 -> Type, 
-    %         (forall a:T1, P zero a a) ->
-    %         (forall a: T1, P a zero a) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Gt ->
-    %            P b1 (cons a2 n2 b2) o ->
-    %            P (cons a1 n1 b1) (cons a2 n2 b2)
-    %              (cons a1 n1 o)) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Lt ->
-    %            P (cons a1 n1 b1) b2 o ->
-    %            P (cons a1 n1 b1) (cons a2 n2 b2) 
-    %            (cons a2 n2 o)) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Eq ->
-    %            P b1 b2 o ->
-    %           P (cons a1 n1 b1) (cons a2 n2 b2)
-    %             (cons a1 (S (n1 + n2)%nat) o)) ->
-    %          forall a b, P a b (oplus a b).
-    % Proof with auto.
-    %    induction a.
-    %    -    intro; simpl; destruct b;auto.
-    %    -   induction b.
-    %        + apply X0.
-    %        + case_eq (compare a1 b1).
-    %          * intro Comp; unfold oplus; rewrite Comp.
-    %            cbn; apply X3 ...
-    %          * intro Comp; cbn; rewrite Comp; apply X2...
-    %          * intro Comp; cbn; rewrite Comp ...
-    %  Defined.
-
-    % Ltac oplus_induction a b:= pattern (oplus a b); apply oplus_rect.
-    % \end{Coqsrc}
-
-    % \index{Exercises}
-
-    % \begin{exercise}
-    % The induction principle \texttt{oplus\_rect} is still unused in our development. 
-    % Please build some nice examples of application.
-    % \end{exercise}
-
-    \index{hydras}{Projects}
-    \begin{project}
-    Compare \texttt{oplus}'s definition (with inner fixpoint) with other possibilities
-    (\texttt{coq-equations}, \texttt{Function}, etc.).
-    \end{project}
-    \subsection{More theorems on Hessenberg's sum}
-
-    We need to prove some properties of $\oplus$, particularly about 
-    its relation with the order $<$ on \texttt{T1}.
-
-    \subsubsection{Commutativity, associativity}
-
-    We prove  the commutativity of $\oplus$ in two steps. 
-    First, we prove by transfinite induction on $\alpha$ that the restriction of $\oplus$ to the
-    interval $[0..\alpha)$ is commutative.
-
-    \index{maths}{Transfinite induction}
-
-    \input{movies/snippets/Hessenberg/oplusComm0}
-
-    Then, we infer  $\oplus$'s commutativity for any pair of ordinals:
-    Let $\alpha$ and $\beta$ be two ordinals strictly less than $\epsilon_0$. Both ordinals $\alpha$ and $\beta$ are
-    strictly less than $\textrm{max}(\alpha,\beta)+1$.
-        Thus, we have just to apply the lemma \coqsimple{oplus\_comm\_0}.
-
-\input{movies/snippets/Hessenberg/oplusComm}
-
-    Associativity of Hessenberg's sum is proved the same way.
-
-    \input{movies/snippets/Hessenberg/oplusAssoc0}
-    \input{movies/snippets/Hessenberg/oplusAssoc}
-
-    \subsubsection{Monotony}
-
-    At last, we prove that $\oplus$ is strictly monotonous in both of its arguments.
-
-    \input{movies/snippets/Hessenberg/oplusMono}
-
-      \index{hydras}{Projects}
-
-    \begin{project}
-    The library \texttt{Hessenberg} looks too long (proof scripts and compilation).
-    Please try to make it simpler and more efficient!
-    Thanks!
-    \end{project}
-
-    \paragraph*{\gaiasign} The module
-      \href{../theories/html/gaia_hydras.~GHessenberg.html}{gaia\_hydras.GHessenberg} defines a version of Hessenberg's sum
-      compatible with \gaia's type \texttt{T1}.
-
-    \subsection{A termination measure for hydra battles }
-
-    \label{sec:hydra-measure}
-
-    Let us define a measure from type \texttt{Hydra} into \texttt{T1}.
-
-    \vspace{4pt}
-    \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Termination.html\#m}{Hydra.Hydra\_Termination}}
-
-   \input{movies/snippets/Hydra_Termination/mDef}
-
-    First, we prove that the measure $m(h)$  of any hydra $h$ is a well-formed ordinal term of type \texttt{T1}.
-
-    \input{movies/snippets/Hydra_Termination/mNf}
-      \input{movies/snippets/Hydra_Termination/msNf}
-
-    For proving the termination of all hydra battles, we have to prove that
-    \texttt{m} is a variant. First, a few technical lemmas follow the decomposition of \texttt{round} into several relations. Then the lemma \texttt{round\_decr} gathers all the cases.
-
-    \label{sect:variant-decr}
-
-    \input{movies/snippets/Hydra_Termination/S0Decr}
-    \input{movies/snippets/Hydra_Termination/R1Decr}
-    \input{movies/snippets/Hydra_Termination/S1Decr}
-    \input{movies/snippets/Hydra_Termination/R2Decr}
-     \input{movies/snippets/Hydra_Termination/RoundDecr}
-
-      Finally, we prove termination of all (free) battles.
-
-    \label{thm:every-battle-terminates}
-
-    \input{movies/snippets/Hydra_Termination/FinalThm}
-
-    \paragraph*{\gaiasign}
-    The module~\href{../theories/html/gaia_hydras.GHydra.html}{gaia\_hydras.GHydra} contains a proof of \gaiaHydras version of termination of all battles.
-
-    \inputsnippets{GHydra/mDef, GHydra/mVariant, GHydra/Termination}
-
-    \section*{Conclusion}
-
-    Let us recall three results we have proved so far.
-    \begin{itemize}
-    \item There exists a strictly decreasing variant which maps \texttt{Hydra} into 
-    the segment $[0,\epsilon_0)$ for proving the termination of any hydra battle
-    \item There exists \emph{no} such variant from \texttt{Hydra} into 
-    $[0,\omega^2)$, \emph{a fortiori} into $[0,\omega)$.
-    \end{itemize}
-
-    So, a  natural question is `` Does there exist any strictly decreasing variant mapping
-    type \texttt{Hydra} into some interval $[0,\alpha)$ (where $\alpha <\epsilon_0$) for proving the termination of all hydra battles''.
-The next chapter is dedicated to a formal proof that there exists no such $\alpha$, even if we consider a restriction to the set of ``standard'' battles.
-
-    %\include{epsilon0}
-
-    %\include{impossibility-proofs}
-
-%-------------------------------------------------------------------
-
-\chapter[The Ketonen-Solovay machinery]{Accessibility inside \texorpdfstring{$\epsilon_0$}{Epsilon0}: The Ketonen-Solovay Machinery\label{ks-chapter}}
-\label{chap:ketonen}
-\index{maths}{Ordinal numbers!Ketonen-Solovay machinery}
-
-\section{Introduction}
-The reader may think that our proof of termination in the previous  chapter requires a lot of mathematical tools and may be too  complex. So, the question is ``is there  any  simpler proof'' ?
-
-In their article~\cite{KP82}, Kirby and Paris show that this result cannot be proved in Peano arithmetic.
-Their proof uses some knowledge about model theory and non-standard models of Peano arithmetic.
-In this chapter, we focus on a specific class of proofs of termination of hydra battles: construction of some variant mapping the type \texttt{Hydra} into a given initial  segment of ordinals.
-Our proof relies only on the Calculus of Inductive Constructions and is a natural complement of the results proven in the previous chapters.
-
-\begin{itemize}
-\item There is no variant mapping the type \texttt{Hydra} into the interval $[0,\omega^2)$ (section ~\vref{omega2-case}), and \emph{a fortiori} $[0,\omega)$ (section ~\vref{omega-case}).
-
-\item There exists a variant which maps the type \texttt{Hydra} into the
-interval $[0,\epsilon_0)$ (theorem \texttt{every\_battle\_terminates}, in section~\vref{thm:every-battle-terminates}).
-\end{itemize}
-
-Thus, a very natural question is the following one:
-\begin{quote}
-  `` Is there  any variant from
-\texttt{Hydra} into some interval $[0,\mu)$, where $\mu<\epsilon_0$, for proving the termination of all hydra battles ?''
-\end{quote}
-
-We prove in \coq{} the following result:
-
-\begin{quote}
-There is no variant for proving the termination of all hydra battles
-from \texttt{Hydra} into the interval $[0..\mu)$, where
-$\mu< \epsilon_0$.
-The same impossibility holds even if we consider only standard battles (with the successive replication factors $0,1,2,\dots,t,t+1,\dots$).
-\end{quote}
-
-Our proofs are  constructive and require no axioms: they are  closed terms of the CIC, and are mainly composed on function definitions and proofs of properties of these functions. 
-They  share much theoretical material with Kirby and Paris', although they do not use any knowledge about Peano arithmetic nor model  theory.  The combinatorial arguments we use and implement
-come from 
- an article by J.~Ketonen and R.~Solovay~\cite{KS81}, already  cited in the work
- by L.~Kirby and J.~Paris. % on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
- Section $2$ of this article: ''A hierarchy of probably recursive functions'', contains a systematic study of \emph{canonical sequences}, which are closely related to
-rounds of hydra battles. 
-Nevertheless, they have the same global structure as the simple proofs described in
-sections~\vref{omega-case} and \vref{omega2-case}. 
-We invite the reader to compare the three proofs step by step, lemma by lemma.
-
-\section{Canonical Sequences}
-\label{ketonen-solovay-sect}
-\index{maths}{Ordinal numbers!Canonical sequences}
-
-Canonical sequences are functions that associate an ordinal $\canonseq{\alpha}{i}$ to every ordinal $\alpha<\epsilon_0$ and positive integer $i$. They satisfy several nice properties:
-
-\index{maths}{Transfinite induction}
-\begin{itemize}
-\item If $\alpha\not=0$, then $\canonseq{\alpha}{i}<\alpha$. Thus canonical sequences can be used for proofs by transfinite induction or function definition by transfinite recursion
-\item If $\lambda$ is a limit ordinal, then $\lambda$ is the least upper bound of the set 
-$\{\canonseq{\lambda}{i}\;|\,i\in\mathbb{N}_1\}$
-
-\item If $\beta<\alpha<\epsilon_0$, then there is a ``path'' from $\alpha$ to $\beta$, \emph{i.e.} a
-sequence $\alpha_0=\alpha, \alpha_1, \dots, \alpha_n=\beta$, where for every $k<n$, there exists some $i_k$ such that $\alpha_{k+1}=\canonseq{\alpha_k}{i_k}$
-\item Canonical sequences correspond tightly to rounds of hydra battles: if $\alpha\not=0$,
-then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1})$ in one round with
-the replication factor $i$ (Lemma \href{../theories/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
-\item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) battle from $\iota(\alpha)$ to $\iota(\beta)$.
-\end{itemize}
-
-\begin{remark}
-  In~\cite{KS81}, canonical sequences are defined for any ordinal $\alpha <\epsilon_0$,
-by stating that if $\alpha$ is a successor ordinal $\beta+1$,  the sequence associated with 
-$\alpha$ is simply the constant sequence whose terms are equal to $\beta$.
-Likewise, the canonical sequence of $0$ maps any natural number to $0$.
-
-This convention allows us to make total the function that maps any ordinal $\alpha$ and natural number $i$ to the ordinal $\canonseq{\alpha}{i}$.
-
-\end{remark}
-
-First, let us recall how canonical sequences are defined in~\cite{KS81}.
-For efficiency's sake, we decided not to implement directly K.\&S's definitions, but to define in \gallina{} simply typed structurally recursive functions which share the abstract properties which are used in the mathematical proofs\footnote{With a small difference: the $0$-th term of the canonical sequence is not the same in our development as in~\cite{KS81}.}.
-
-\subsubsection{Mathematical definition of canonical sequences}
-
-In~\cite{KS81} the definition of $\canonseq{\alpha}{i}$ is based on the following remark:
-\begin{quote}
-Any non-zero ordinal $\alpha$ can be decomposed in a unique way as the product
-$\omega^\beta\times (\gamma+1)$.
-\end{quote}
-
-Thus the $\canonseq{\alpha}{i}$\,s are defined in terms of this decomposition:
-\begin{definition}[Canonical sequences: mathematical definition]
-\label{def:canonseq-math}
-
-\end{definition}
-\begin{mathframe}
-  \begin{itemize}
-\item Let $\lambda<\epsilon_0$ be a limit ordinal
-
-\begin{itemize}
-\item If $\lambda=\omega^{\alpha+1}\times (\beta+1)$, then 
-$\canonseq{\lambda}{i}= \omega^{\alpha+1}\times\beta +  \omega^\alpha \times i$
-\item If $\lambda=\omega^{\gamma}\times (\beta+1)$, where $\gamma<\lambda$ is a limit ordinal, then 
-$\canonseq{\lambda}{i}=\omega^{\gamma}\times \beta + \omega^{\canonseq{\gamma}{i}}$
-\end{itemize}
-
-\item For successor ordinals, we have $\canonseq{\alpha+1}{i}= \alpha$
-
-\item Finally, $\canonseq{0}{i}= 0$.
-\end{itemize}
-\end{mathframe}
-
-\subsubsection{Canonical sequences in \coq}
-\index{hydras}{Library Epsilon0!Functions!canon}
-\index{hydras}{Library Epsilon0!Functions!canonS}
-
-Our definition may look more complex than the mathematical one, but
-uses plain structural recursion over the type \coqsimple{T1}. Thus, tactics like
-\coqsimple{cbn}, \coqsimple{simpl}, \coqsimple{compute}, etc., are applicable.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canon}{Epsilon0.Canon}}
-
-\label{Functions:canonS}
-\label{Functions:canon}
-
-\input{movies/snippets/Canon/canonDef}
-
-\paragraph*{\gaiasign} Our library~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon} transposes the  function \texttt{canon} into \gaia's type \texttt{T1}
-(please see Sect.\~ref{sect:gcanon}).
-
-\paragraph*{Remark}
-In the present state of this library, the following specializations of \texttt{canon} are still used in some proofs or lemma statements. They are planned to be deprecated.
-
-\input{movies/snippets/Canon/CanonS0}
-
-For instance \coq's computing facilities allow us to verify the equalities\linebreak 
-\mathcolor{$\canonseq{\omega^\omega}{3} = \omega^3$} and
-\mathcolor{$\canonseq{\omega^\omega*3}{42} = \omega^\omega*2 + \omega^{42}$}.
-
-\input{movies/snippets/Canon/canonExamples}
-
-% \index{hydras}{Projects}
-% \begin{project}
-% Many lemmas presented in this chapter were stated and proved before the introduction of 
-% the type class \texttt{ON} of ordinal notations, and in particular its  instance \texttt{Epsilon0}.
-% Thus definitions and lemmas refer to the type \texttt{T1} of possibly not well-formed terms.
-% This should be fixed in  a future version.
-% \end{project}
-
-\subsection{Basic properties of canonical sequences}
-
-We did not  try to prove that our definition truly implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
-abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
-~\href{../theories/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}.
-For instance, the equality $\canonseq{\alpha+1}{i}=\alpha$  can be  proved by  structural induction on $\alpha$.
-
-\input{movies/snippets/Canon/canonSucc}
-
-\subsubsection{Canonical sequences and the order $<$}
-
-\index{maths}{Transfinite induction}
-
-We prove by transfinite induction over $\alpha$ that $\canonseq{\alpha}{i}$ is an ordinal strictly less than $\alpha$ (assuming $\alpha\not=0$).
-This property allows us to use the function \texttt{canonS} and its derivatives in function definitions by transfinite recursion.
-
-\label{lemma:canon_LT}
-
-\inputsnippets{Canon/canonLT}
-
-\paragraph*{\gaiasign} This lemma is also available in
-Library ~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon}:
-
-\inputsnippets{GCanon/gcanonLt}
-
-\subsubsection{Limit ordinals are truly limits}
-The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
-is the limit of the sequence $\canonseq{\lambda}{i}\;(1\le i)$.
-
-\vspace{4pt}
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}
-
-\input{movies/snippets/Canon/canonSLimitStrong}
-
-\label{lemma:canonS-limit}
-
-Note the use of \coq's \texttt{sig} type in the theorem's statement, which
-relates the boolean function \texttt{limitb} defined on the \texttt{T1} data-type with a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.
-We can also state directly that $\lambda$ is a (strict) least upper bound of the elements of its canonical sequence.
-
-\input{movies/snippets/Canon/canonSLimitLub}
-
-\paragraph*{\gaiasign}  In \gaiaHydras, the statement use a slightly different vocabulary:
-\inputsnippets{GCanon/gcanonLimitOf}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}\label{exo:simply-typed-canonseq}
-Instead of using the \texttt{sig} type, define a simply typed function that, given two ordinals $\alpha$ and $\beta$, returns a natural number $i$ such that, if $\alpha$ is a limit ordinal and $\beta<\alpha$, then $\beta< \canonseq{\alpha}{i+1}$.
-Of course, you will have to prove the correctness of your function.
-
-\textbf{Hint:} You may add to your function a third argument usually called \texttt{fuel} for allowing you to give a structurally 
-recursive function (\emph{cf} the post of Guillaume Melquiond on Coq-club (Dec 21, 2020)
-\url{https://sympa.inria.fr/sympa/arc/coq-club/2020-12/msg00069.html}).
-The type \texttt{fuel}, an alternative 
-to \texttt{nat} is available on \href{../theories/html/hydras.Prelude.Fuel.html}{Prelude.Fuel})
-.
-%\index{coq}{Giving fuel to a long computation}
-
-\end{exercise}
-
-\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0}\,: paths}
-\index{maths}{Ordinal numbers!Accessibility inside epsilon0}
-\label{sect:pathes-intro}
-
-Let us consider a kind of accessibility problem inside $\epsilon_0$: given two ordinals $\alpha$ and $\beta$, where $\beta<\alpha<\epsilon_0$, find a \emph{path} consisting of a finite sequence $\gamma_0=\alpha,\dots,\gamma_l=\beta$,
-where, for every $i<l$, $\gamma_i \not= 0$ \footnote{This condition allows us to ignore paths which end by a lot of useless $0$s.} and there exists some strictly positive integer $s_i$
-such that $\gamma_{i+1}=\canonseq{\gamma}{s_i}$.
-
-Let $s$ be the sequence $\langle s_0,s_1,\dots, s_{l-1} \rangle$. We describe the
-existence of such a path with the notation $\alpha\xrightarrow [s]{}\beta$.
-We say also that the considered path from $\alpha$ to $\beta$ \emph{starts at [index] $s_0$ and ends at $s_l$}.
-
-For instance, we have $\omega*2 \xrightarrow[2,2,2,4,5]{}3$, through the 
-path $\langle\omega\times 2, \omega+2,\omega+1,\omega,4,3\rangle$.
-
-\begin{remark}
-
-Note that, given $\alpha$ and $\beta$, where $\beta < \alpha$, the sequence $s$ which leads from $\alpha$ to $\beta$ is not unique.
-Indeed, if $\alpha$ is a limit ordinal, the first element of $s$ can be any integer $i$ such that $\beta<\canonseq{\alpha}{i}$, and if $\alpha$ is a successor ordinal,
-then the sequence $s$ can start with any positive integer.
-
-For instance, we have also 
-$\omega*2 \xrightarrow[3,4,5,6]{}\omega$. 
-Likewise,
-$\omega*2 \xrightarrow[1,2,1,4]{} 0$ and
-$\omega*2 \xrightarrow[3,3,3,3,3,3,3,3]{} 0$.
-\end{remark}
-
-\subsection{Formal definition}
-
-\label{path-to-definition}
-
-In \coq{}, the notion of path can be simply defined as an inductive predicate 
-parameterized by the destination $\beta$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\index{hydras}{Library Epsilon0!Predicates!path\_to}
-\label{sect:path-to-def}
-
-\input{movies/snippets/Paths/transitionDefs}
-
-\input{movies/snippets/Paths/pathDef}
-
-\begin{remark}
-In the present version of our library, we use a variant \texttt{path\_toS} of
-\texttt{path\_to}, where the proposition
-(\texttt{path\_toS $\beta$ $s$ $\alpha$}) is equivalent to
-(\texttt{path\_to $\beta$ (List.map S $s$) $\alpha$}). This variant is scheduled to be deprecated.
-\end{remark}
-
-\paragraph*{\gaiasign}
-The library~\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} transposes the  notion of path into \gaia's type \texttt{T1}
-(please see Sect.\~ref{sect:gpath}).
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Write a tactic for solving goals of the form (\texttt{path\_to $\beta$ $s$ $\alpha$})
-where $\alpha$, $\beta$ and $s$ are closed terms. 
-You should solve automatically the following goals:
-
-\inputsnippets{Paths/exPathTac} 
-\end{exercise}
-
-\subsection{Existence of a path}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction on $\alpha$, we prove that for any $\beta<\alpha$, 
-one can build a path from $\alpha$ to $\beta$ (in other terms, $\beta$ is accessible from $\alpha$).
-
-\input{movies/snippets/Paths/LTPathTo}
-
-\noindent 
-From the lemma \texttt{canon\_LT}~\vref{lemma:canon_LT}, we can convert any path into an inequality on ordinals (by induction on paths).
-
-\input{movies/snippets/Paths/pathToLT}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}[continuation of exercise~\vref{exo:simply-typed-canonseq}]
-Define a simply typed function for computing a path from $\alpha$ to $\beta$.
-\end{exercise}
-
-\subsection{Paths and hydra battles}
-\label{KS-o2h}
-
-In order to apply our knowledge about  ordinal numbers less than $\epsilon_0$ to the study of hydra battles, we define an injection
-from the interval $[0,\epsilon_0)$ into the type \texttt{Hydra}.
-
-\vspace{4pt}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}}
-
-\input{movies/snippets/O2H/iotaDef}
-
-For instance Fig.~\ref{fig:iota-example} shows the image by $\iota$ of the ordinal  \textcolor{black}{$\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$}
-
-  \begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.3]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (10,2) {$\bullet$};
-\node (N22) at (7,2) {$\bullet$};
-\node (N3) at (14,2) {$\bullet$};
-\node (N4) at (18,2) {$\Smiley[2][green]$};
-\node (N5) at (0,4) {$\bullet$};
-\node (N6) at (2,5) {$\Smiley[2][green]$};
-\node (N7) at (4,6) {$\Smiley[2][green]$};
-\node (N88) at (7,4) {$\bullet$};
-\node (N8) at (10,4) {$\bullet$};
-\node (N9) at (14,6) {$\Smiley[2][green]$};
-\node (N10) at (0,8) {$\Smiley[2][green]$};
-\node (N11) at (10,7) {$\Smiley[2][green]$};
-\node (N111) at (7,7) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N22);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (N1) to  (N5);
-\draw (N1) to   [bend left=10] (N6);
-\draw (N1) to   [bend right=20] (N7);
-\draw (N2) to  (N8);
-\draw (N22) to  (N88);
-\draw (N8) to  (N11);
-\draw (N88) to  (N111);
-\draw (N3) to  (N9);
-\draw (N5) to  (N10);
-\end{tikzpicture}
-\caption{The hydra $\iota(\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$) \label{fig:iota-example}}
-
-\end{figure}
-
-The following lemma (proved in ~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H.v}) maps  canonical sequences to rounds of hydra battles.
-
-\label{lemma:canonS-iota}
-
-\input{movies/snippets/O2H/canonSIota}
-
-The next step of our development extends this relationship to
-the order $<$ on $[0,\epsilon_0)$ on one side, and hydra battles on the other side.
-
-\input{movies/snippets/O2H/pathToRoundPlus}
-
-As a corollary, we are now able to transform any inequality $\beta<\alpha<\epsilon_0$ into a (free) battle.
-
-\input{movies/snippets/O2H/LTToRoundPlus}
-
-\section{A  proof of impossibility}
-
-We now have  the tools for proving that  there exists no variant bounded by some $\mu<\epsilon_0$ for proving the termination   of all battles.
-The proof we are going to show is a proof by contradiction.
-It  can
- be considered as a generalization of the
-proofs described in  sections~\vref{omega-case} and \vref{omega2-case}.
-
-In the module
-\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
-This part of the development is parameterized by some class $B$ of battles, which will be instantiated later to \texttt{free} or \texttt{standard}.
-
-\vspace{4pt}
-\noindent
-From~\href{../theories/html/hydras.Hydra.~Hydra_Definitions.html}{Hydra.Hydra\_Definitions}:
-
-\input{movies/snippets/Hydra_Definitions/BoundedVariant}
-
-Let us assume there exists such a variant:
-
-\input{movies/snippets/Epsilon0_Needed_Generic/theContext}
-
-\label{remark:m-decrease}
-\begin{remark}
-  The hypothesis \texttt{m\_decrease} is not provable  in general, but is satisfied by
-the  \texttt{free} and \texttt{standard} kinds of battles. This trick allows to 
-``factorize'' our proofs  of impossibility.
-\end{remark}
-
-\index{maths}{Transfinite induction}
-
-First, we prove that $m(\iota(\alpha))$ is always greater than or equal to $\alpha$, by  transfinite induction over $\alpha$.
-
-\input{movies/snippets/Epsilon0_Needed_Generic/mGe0}
-
-\begin{itemize}
-\item If $\alpha=0$, the inequality trivially holds
-\item If $\alpha$ is the successor of  some ordinal $\beta$, the inequality $\beta \leq m(\iota(\beta))$ holds (by induction hypothesis).
-But the hydra $\iota(\alpha)$ is transformed in one round into
-$\iota(\beta)$, thus $m(\iota(\beta))<m(\iota(\alpha))$. Hence $\beta<m(\iota(\alpha))$, which implies $\alpha \leq m(\iota(\alpha))$
-\item If $\alpha$ is a limit ordinal, then $\alpha$ is the least upper bound of the set
-of all  the $\canonseq{\alpha}{i}$.  Thus, we have just to prove that $\canonseq{\alpha}{i}< m(\iota(\alpha))$ for any $i$. 
-\begin{itemize}
-\item Let $i$ be some natural number.
-By the induction hypothesis, we have $\canonseq{\alpha}{i} \leq m(\iota(\canonseq{\alpha}{i}))$.
-But the hydra $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i})$ in one round, thus $m(\iota(\canonseq{\alpha}{i})) < m(\iota(\alpha))$, by our hypothesis \texttt{m\_decrease}.
-\end{itemize}
-\end{itemize}
-
-Please note that the impossibility proofs of 
-sections~\vref{omega-case} and \vref{omega2-case} contain a similar lemma, also called \texttt{m\_ge}.
-We are now able to build a counter-example.
-
-\input{movies/snippets/Epsilon0_Needed_Generic/mGeGeneric}
-
-The (big) rest of the proof is dedicated to prove formally the converse inequality 
-\texttt{m small\_h t1< m big\_h}.
-
-\subsection{The case of free battles}
-\label{sec:free-battles-case}
-Let us now consider that $B$ is instantiated to \texttt{free} (which means that we are considering proofs of termination of \emph{all} battles).
-The following lemmas are proved in Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
-The case $B=\texttt{standard}$ is studied in section~\vref{std-case}.
-
-\input{movies/snippets/Epsilon0_Needed_Free/theContext}
-
-\begin{enumerate}
-\item The following lemma is an application of \texttt{m\_ge\_generic}, since \texttt{free}
-satisfies trivially the hypothesis \texttt{m\_decrease} (see page~\pageref{remark:m-decrease}).
-
-\inputsnippets{Epsilon0_Needed_Free/mGe}
-
-\item From the hypothesis \texttt{Hy}, we have \texttt{m big\_h t1< mu}
-\item By Lemma \texttt{LT\_to\_round\_plus}, we get a (free) battle from
-\texttt{big\_h = iota mu} to \texttt{small\_h = iota (m big\_h)}.
-
-\input{movies/snippets/Epsilon0_Needed_Free/bigToSmall}
-
-\item From the hypotheses on $m$, we infer:
-\input{movies/snippets/Epsilon0_Needed_Free/mLt}
-
-\item From lemmas \texttt{m\_ge} and \texttt{m\_lt}, and the irreflexivity of $<$, we get a contradiction. 
-\input{movies/snippets/Epsilon0_Needed_Free/ImpossibilityFree}
-
-\end{enumerate}
-
-We have now proved there exists no bounded variant for the class of free battles.
-
-\input{movies/snippets/Epsilon0_Needed_Free/CheckDemo}
-
-\section{The case of standard battles}
-\label{sec:standard-intro}\label{std-case}
-One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
-Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}.
-
-\input{movies/snippets/O2H/LTToRoundPlus}
-
-This lemma builds a battle out of any inequality $\beta<\alpha$. 
-It is a straightforward application of \texttt{LT\_path\_to} of
-Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}:
-
-\input{movies/snippets/Paths/LTPathTo}
-
-The sequence $s$, used to build the sequence of replication factors of the battle, depends on 
-$\alpha$ and $\beta$, so we cannot be sure that the generated battle is a genuine standard battle.
-
-The solution to this issue comes  once again from Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
-$\alpha_0=\alpha,\alpha_1,\dots,\alpha_k=\beta$ where $\alpha_{j+1}$ is equal
-to $\canonseq{\alpha_j}{i_j}$ where $i_j$ is \emph{any} natural number, 
-we consider various constraints on these sequences.
-In particular, a path is called \emph{standard} if $i_{j+1} = i_j + 1$ for every $j<k$.
-It  corresponds to a ``segment'' of some standard battles. 
-Please note that the vocabulary on paths is ours, but all the concepts come really from~\cite{KS81}.
-
-In \coq{}, standard paths can be defined as follows.
-
-\vspace{4pt}
-
-\emph{From
-Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/standardPathR}
-
-In the mathematical text and figures, we shall use the notation 
-$\alpha \xrightarrow[i,j]{}\beta$ for the proposition 
-(\texttt{standard\_path $i$ $\alpha$ $j$ $\beta$}).
-In~\cite{KS81} the notation is
-$\alpha \xrightarrow[i]{*}\beta$
-for 
-the proposition  $\exists j, i<j \wedge \alpha \xrightarrow[i,j]{} \beta$.
-
-Our goal is now  to transform any inequality $\beta<\alpha<\epsilon_0$ into a standard path $\alpha \xrightarrow[i,j]{} \beta$ for some $i$ and $j$, then into a standard battle
-from $\iota(\alpha+i)$ to $\iota(\beta)$. 
-Following~\cite{KS81}, we proceed in two stages:
-\begin{enumerate}
-\item we simulate plain (free) paths from $\alpha$ to $\beta$ with
-paths made of steps $(\gamma,\canonseq{\gamma}{n})$, \emph{with the same $n$ all along the path}
-\item we simulate any such path by a standard path.
-\end{enumerate}
-
-\subsection{Paths with a constant index}
-
-First of all, paths with a constant index 
-enjoy nice properties. They are defined as paths where all the $i_j$ are equal to the same natural number $i$, for some $i>0$.
-
-Like in~\cite{KS81}, we shall use the notation $\alpha \xrightarrow[i]{} \beta$ for denoting such a path, also called an $i$-path.
-
-\input{movies/snippets/Paths/constPathDef}
-
-% Paths with a given index can be effectively computed.
-% Given $i$, $\alpha$ and $l$, the following function returns the ordinal $\beta$ such that there exists a path 
-% $\alpha \xrightarrow [i+1] {} \beta$ of length $l$.
-
-% \begin{Coqsrc}
-% Fixpoint const_funS (i:nat)(alpha : T1)(l:nat):  T1  :=
-%   match l
-%   with
-%   | 0 => alpha
-%   | S m => const_funS i (canonS i alpha) m
-%   end.
-% \end{Coqsrc}
-
-% The following computations show  applications of \texttt{constS\_fun} to the 
-% ordinal $\omega^\omega$, with various values of $i$ and $l$.
-
-% \begin{Coqsrc}
-% Compute  (const_funS 2 (omega ^omega)  55).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = zero
-%      : T1 
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (const_funS 2 (omega ^omega) 15).
-% \end{Coqsrc}
-
-%   \begin{Coqanswer}
-%  = (omega ^ 2 * 2)%pT1
-%      : ppT1   
-%   \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (const_funS 4 (omega^omega)  100).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% = (omega ^ 4 * 4 + omega ^ 3 * 4 + omega ^ 2 + omega * 4 + 4)%pT1
-%      : ppT1
-% \end{Coqanswer}
-
-A most interesting property of $i$-paths is that we can ``upgrade'' their index, as stated by K.\&S.'s Corollary 12.
-
-\index{maths}{Transfinite induction}
-
-\vspace{4pt}
-\noindent
-\emph{From
-  Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/Cor12}
-
-We  also use a version of \texttt{Cor12} with large inequalities.
-
-\input{movies/snippets/Paths/Cor121}
-
-\subsubsection{Sketch of proof of \texttt{Cor12}}
-\index{maths}{Transfinite induction}
-
-First, let us prove the following lemma (numbered \texttt{2.4} in Ketonen and Solovay's article).
-
-\input{movies/snippets/Paths/KSThm24}
-
-We prove this lemma by transfinite induction on $\alpha$.
-Let us consider a path $\alpha \xrightarrow [i]{} \beta$ $(i>0)$. Its first step is
-the pair $(\alpha,\canonseq{\alpha}{i})$, We have $\canonseq{\alpha}{i}<\alpha$ and
-$\canonseq{\alpha}{i} \xrightarrow [i]{} \beta$. 
-Let $n$ be any natural number such that $n>i$.
-By the induction hypothesis, there exists a path $\canonseq{\alpha}{n} \xrightarrow[i]{} \beta$.
-\begin{itemize}
-\item  If $\alpha$ is a successor ordinal $\gamma+1$, then $\canonseq{\alpha}{n} =
-\canonseq{\alpha}{i}=\gamma$. Thus we have a path 
-$\alpha  \xrightarrow [n]{}  \gamma \xrightarrow [n]{} \beta$
-\item If $\alpha$ is a limit ordinal, we apply Theorem \texttt{KS\_thm\_2\_4} (see above).
-
-%   \begin{theorem}
-% Let $\lambda$ be a limit ordinal, then for any pair of indices $0<i<j$, there is a path $\canonseq{\lambda}{j} \xrightarrow[1]{} \canonseq{\lambda}{i}$.    
-%   \end{theorem}
-
-The proof of the limit case, is decomposed into a sequence
-of path constructions leading to $\alpha \xrightarrow[n]{} \beta$.
-
- \begin{enumerate}
- \item $\alpha \xrightarrow[n]{} \canonseq{\alpha}{n}$ (single step path)
- \item $\canonseq{\alpha}{n} \xrightarrow[1]{} \canonseq{\alpha}{i}$ (by \texttt{Theorem\_2\_4}),
-\item $\canonseq{\alpha}{n} \xrightarrow[n]{} \canonseq{\alpha}{i}$ (applying the induction hypothesis to the preceding path);
-\item $\canonseq{\alpha}{i} \xrightarrow[n]{} \beta$ (applying the induction hypothesis)
-\item $\alpha \xrightarrow[n]{} \beta$ (by composition of 1, 3, and 4).
-
- \end{enumerate}
-
-\end{itemize}
-
-\begin{remark}
- \texttt{Cor12} ``casts'' $i$-paths into $n$-paths for any $n>i$.
-But the obtained $n$-path can be much longer than the original $i$-path.
-The following exercise will give an idea of this increase. 
-\end{remark}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-  Prove that  the length of the $i+1$-path from
-  $\omega^\omega$ to $\omega^i$ is $1 + (i+1)^{(i+1)}$, for any $i$. Note that the $i$-path from
-  $\omega^\omega$ to $\omega^i$ is only one step long.
- \end{exercise}
-
-Why is \texttt{Cor12} so useful? 
-Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
-we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying corollary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
-where $n$ is the maximum of the indices $i$ met in the induction.
-
- Lemma 1, Section 2.6 of~\cite{KS81} is naturally expressed in terms of \coq's
-\verb@sig@ construct.
-
-\label{lemma:L-2_6-1}
-%\index{coq}{Sigma types}
-
-\input{movies/snippets/Paths/Lemma261}
-
-Intuitively, Lemma~\texttt{2\_6\_1}  shows that if $\beta<\alpha<\epsilon_0$, then there exists  a battle from $\iota(\alpha)$ to $\iota(\beta)$ where the replication factor is constant, although large enough.
-
- \paragraph*{\gaiasign}
- Corollary \texttt{Cor12} and Lemma \texttt{Lemma2\_6\_1} are also available in
-\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} 
-(please see Sect.~\vref{sect:gpath}).
-
-\subsection{Casting paths with a constant index into a standard path}
-
-The article~\cite{KS81} contains 
-the following lemma, which allows us to simulate $i$-paths by $[i+1,j]$-paths, where $j$ is large enough.
-
-\input{movies/snippets/Paths/constantToStandard}
-
-\subsubsection{Sketch of proof of \texttt{constant\_to\_standard\_path}}
-
-Our proof follows the proof by Ketonen and Solovay, including its organization as a sequence of lemma.
-Since it is a non-trivial proof, we will comment its main steps below   (see Figure~\vref{fig:belle-preuve-1} to Figure~\vref{fig:fin-belle-preuve}).
-
-\subsubsection*{Preliminaries}
-
-Please note that, given an ordinal $\alpha:\texttt{T1}$, and two natural numbers $i$ and $l$, there exists at most a standard path $\alpha \xrightarrow [i,i+l]{*} \beta$.
-The following function computes $\beta$ from $\alpha$, $i$ and $l$.
-
-\input{movies/snippets/Paths/standardGnaw}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction over  $\alpha$, we prove that the ordinal $0$ is reachable from any ordinal $\alpha<\epsilon_0$ by some standard path.
-
-\input{movies/snippets/Paths/standardPathToZero}
-
-\paragraph*{}
-Now, let us consider two ordinals  $\beta<\alpha<\epsilon_0$.  Let $p$  be some $(n+1)$-path from $\alpha$ to $\beta$.
-
-\input{movies/snippets/Paths/ConstantToStandardProof}
-
-Applying \texttt{standard\_path\_to\_zero}, $0$ is reachable from $\alpha$ by some standard path.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \draw[->, very thick,blue] (alpha)-- node [below]{$n+1$} node [above] {$+$} (beta);
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-    \node (alpha3) at (20,5) {};
-  \node (alphalast) at (35,5) {};
-  \node (zero) at (45,0) {$0$};
-  \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate=40]{$n+1$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha2)-- node [below]{$n+3$}  (alpha3);
-
-  \node (dots) at (24,5) {$\dots$};
-  \draw [->, dashed, very thick,blue] (alphalast)-- node [below, rotate=-26]{$n+p+1$}  (zero);
-
-\end{tikzpicture}
-\caption{A nice proof (1)}
-  \label{fig:belle-preuve-1}
-\end{figure}
-
-\paragraph*{}
-
-Since comparison on \texttt{T1} is decidable, one can compute the last step $\gamma$ of the standard path from $(\alpha,n+1)$  such that $\beta\leq \gamma$.
-Let $l$ be the length of the path from $\alpha$ to $\gamma$.  
-This step of the proof is illustrated in figure~\vref{fig:belle-preuve-2}.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (17,5) {$\ldots$};
-    \node (alpha3) at (20,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (38,0) {$\delta$};
-    \draw [->, dashed,very thick,blue] (alpha)-- node [below,rotate=35]{$n+1$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below,rotate = -48]{\tiny $n+l$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{$n+l+1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{$n+1$} node [above] {$+$} (beta);
-   \draw[thick] (alpha)--  (gamma);
-   \draw[thick] (gamma)--  node [above] {$\geq$} (beta);
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-\end{tikzpicture}
-
-\caption{A nice proof (2)}
-  \label{fig:belle-preuve-2}
-\end{figure}
-
-\paragraph*{}
-
-\begin{itemize}
-\item If $\beta=\gamma$, it's OK! We have got a standard path
-from  
-$\alpha$ to $\beta$ with successive indices  $n+1, n+2, \dots, n+l+1$
-
-\item Otherwise,  $\beta < \gamma$.  Let us consider  $\delta=\canonseq{\gamma}{n+l+1}$.
-By applying several times lemma \texttt{Cor12},  one converts  every path of Fig~\ref{fig:belle-preuve-2} into
- a $n+l+1$-path  (see figure~\ref{fig:belle-preuve-3}).
-
-But $\gamma$ is on the $n+l+1$-path from $\alpha$ to $\beta$.
-As shown by figure~\vref{fig:fin-belle-preuve}, the ordinal $\delta$, reachable from
-$\gamma$ in one single step,  must be greater than or equal to $\beta$, which contradicts our  hypothesis $\beta < \gamma$.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-    \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (17,5) {$\ldots$};
-    \node (alpha3) at (20,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (38,0) {$\delta$};
-     \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -48]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\tiny $n+l+1$} node [above]{\color{red} $1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
-    \draw[thick] (gamma)--   node [above]{\color{red} $>$}(beta);
-   \draw[thick] (alpha)--  (gamma);
-
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-
-\end{tikzpicture}
-
-\caption{A nice proof (3)}
-  \label{fig:belle-preuve-3}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (15,5) {$\ldots$};
-    \node (alpha3) at (18,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (42,0) {$\delta$};
-    \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -36]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\small $n+l+1$} node [above]{\color{red} $1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
-   \draw[thick] (alpha)--  (gamma);
-
-    \draw[thick] (gamma)--  node [below]{\tiny $n+l+1$} node [above]{\color{red} $+$}(beta);
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-
-\end{tikzpicture}
-
-\caption{A nice proof (4)}
-  \label{fig:fin-belle-preuve}
-\end{figure}
-
-\end{itemize}
- The only possible case is  thus $\beta=\gamma$, so we have got a standard path  from $\alpha$ to $\beta$.
-
- \input{movies/snippets/Paths/constantToStandard0}
- \input{movies/snippets/Paths/constantToStandardz}
-
-Here is the full statement of the conversion from constant to standard paths.
-
-\input{movies/snippets/Paths/constantToStandardPath}
-
-Applying \texttt{Lemma2\_6\_1} and \texttt{constant\_to\_standard\_path}, we get the following corollary.
-
-\input{movies/snippets/Paths/LTToStandardPath}
-
-\subsection{Back to hydras}
-\label{sec:standard-battles-cases}
-We are now able to complete our proof that there exists no bounded variant for proving the termination of standard hydra battles. This proof can
-be consulted in the module 
-\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}{Hydra.Epsilon0\_Needed\_Std}.
-Please note that it has the same global structure as in section~\ref{sec:free-battles-case} 
-% ICI !
-Applying the  lemmas  \texttt{Lemma2\_6\_1} of the module 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
-{Epsilon0.pathS}   and 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
-{\texttt{constant\_to\_standard\_path}},
-we can convert any inequality $\beta<\alpha<\epsilon_0$ into a standard path from
-$\alpha$ to  $\beta$, then into a fragment of a standard battle from 
-$\iota(\alpha)$ to $\iota(\beta)$, hence the inequality $m(\iota(\beta))<m(\iota(\alpha))$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html\#LT_to_standard_battle}{Hydra.Epsilon0\_Needed\_Std}}
-
-\input{movies/snippets/Epsilon0_Needed_Std/LTToStandardBattle}
-
-Now, please consider the following context:
-
-\input{movies/snippets/Epsilon0_Needed_Std/theContext}
-
-In the same way as for free battles, we import a large inequality 
-from 
-the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Epsilon0\_Needed\_Generic}.
-
-\input{movies/snippets/Epsilon0_Needed_Std/mGe}
-
-\paragraph*{} If remains to prove the following strict inequality, in order to have a contradiction.
-
-\input{movies/snippets/Epsilon0_Needed_Std/mLt}
-
-\paragraph*{Sketch of proof:} Let us recall that $\texttt{big\_h} = \iota(\mu)$
- and $\texttt{small\_h} = \iota (m (\texttt{big\_h}))$.
-
-Since $m(\texttt{big\_h})< \mu$, there exists a standard path from $\mu$ to
-$m(\texttt{big\_h})$, hence a   standard battle from $\iota(\mu)$  to
-$\iota(m(\texttt{big\_h}))$,  i.e. from \texttt{big\_h} to \texttt{small\_h}.
-
-Since $m$ is assumed to be a variant for standard battles, we get the inequality  $m(\texttt{small\_h}) < m(\texttt{big\_h})$.
-
-\input{movies/snippets/Epsilon0_Needed_Std/endOfProof}
-
-\subsection{Remarks}
-
-We are grateful to 
- J. Ketonen and R. Solovay  for the high quality of their explanations and proof details.
-Our proof follows tightly the sequence of lemmas in their article, with a focus on 
-constructive aspects.
-Roughly speaking, our implementation \emph{builds}, out of a hypothetical
-  variant $m$, bounded by some ordinal $\mu<\epsilon_0$, a hydra \texttt{big\_h} which verifies the impossible inequality  $m(\texttt{big\_h})< m(\texttt{big\_h})$.
-
-On may ask whether the preceding results are too restrictive, since they 
-refer to a particular data type \texttt{T1}.
-In fact, our representation of ordinals strictly less than 
- $\epsilon_0$ is faithful to their mathematical definition, at least 
-Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
-(please see also the module
-\href{../theories/html/hydras.Schutte.Correctness_E0.html}{hydras.Schutte.Correctness\_E0}).
-
-Thus, we can infer that our theorems can be applied to any well order.
-
-\index{hydras}{Projects}
-\begin{project}
-Study a possible modification of the definition of a variant  (for  standard battles).
-
-\begin{itemize}
-\item The variant is assumed to be strictly decreasing \emph{on configurations 
-reachable from some initial configuration where the replication factor is equal to $0$}
-\item The variant may depend on the number of the current round.
-\end{itemize}
-
-In other words, its type should be \texttt{nat -> Hydra -> T1}, and it must 
-verify the inequality $m\, (S\,i)\, h' < m\,i\, h$ whenever the configuration 
-$(i,h)$ is reachable from some initial configuration $(0,h_0)$
-and \texttt{h} is transformed into \texttt{h'} in the considered round.
-Can we still prove the theorems of section~\ref{std-case} with this new definition?
-
-\end{project}
-
-%---------------------------------------------------------------------
-\chapter{Large sets and rapidly growing functions}\label{chap:alpha-large}
-
-\begin{remark}
-Some notations (mainly names of fast-growing functions) of our development may differ slightly from the literature.
-Although this fact does not affect our proofs, we are preparing a future version where the names $F\_alpha$, $f\_alpha$, $H\_alpha$, etc., are fully consistent with the cited articles.
-
-\end{remark}
-%\section{Introduction}
-
-In this chapter, we try to feel how long a standard battle can be.
-To be precise, for any ordinal $\alpha<\epsilon_0$ and any positive integer $k$,
-we give a minoration of the number of steps of a standard battle which
-starts with the hydra $\iota(\alpha)$ and the replication factor $k$.
-
-We express this number in terms of the Hardy hierarchy of fast-growing 
-functions~\cite{BW85, Wainer1970, KS81, Promel2013}.
- From the \coq{} user's point of view, such  functions are  very 
-attractive:  they are defined as functions  in \gallina{}, and we can apply them \emph{in theory}, but they are so complex that you will never be able to look at the result of the computation.
- Thus, our knowledge on these functions must rely on \emph{proofs}, not tests. In our development, we use often the rewriting rules generated by \coq's \texttt{Equations} plug-in.
-
-\section{Definitions}
-
-%\subsection{Definition}
-
-\begin{definition}
-Let $0<\alpha<\epsilon_0$ be any ordinal, and $s=\langle s_1, s_2, \dots, s_N\rangle$ a finite sequence of strictly positive natural numbers.
-
-We say that $s$ is \emph{$\alpha$-large} if the sequence $\langle \alpha_0=\alpha,\dots,\alpha_{i+1}=\canonseq{\alpha_i}{i+1},\dots \rangle$ leads to $0$. 
-We say also that $s$ is \emph{minimally $\alpha$-large} (in short:
-\emph{$\alpha$-mlarge}) if $s$ is $\alpha$-large 
- and every strict prefix of $s$ leads to a non-zero ordinal (\emph{cf} Sect.~\vref{sect:path-to-def}).
-
-\index{maths}{Ordinal numbers!Large sets}
-\index{maths}{Ordinal numbers!Minimal large sets}
-
-\end{definition}
-
-\begin{remark}
-  Ketonen and Solovay~\cite{KS81} consider  large finite \emph{sets} of natural numbers,  but they are mainly used as sequences. Thus, we chose to represent them explicitely as (sorted) lists. 
-\end{remark}
-
-The following function ``gnaws'' an ordinal $\alpha$, following a sequence of indices (ignoring the $0$s).
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/gnawDef}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_SetsPaths.html\#gnaw}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/largeDef}
-
-Minimal large sequences can be directly defined in terms of the
-predicate \texttt{path\_to} (\vref{sect:path-to-def}) which already prohibits paths containing non-final \texttt{zero}s.
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}
-
-\index{hydras}{Library Epsilon0!Predicates!mlarge@mlarge (minimal large sequences)}
-
-\input{movies/snippets/Large_Sets/mlargeDef}
-
-Let us consider two integers $k$ and $l$, such that $0<k<l$. In order to check whether the interval $[k,l]$ is minimally large for $\alpha$, it is enough to
-follow from $\alpha$ the path associated with the interval $[k,l)$ and verify that the last ordinal we obtain is equal to $1$.
-
-\subsection{Examples}
-
-For instance the interval $[6,70]$ leads $\omega^2$ to $\omega\times 2 + 56$. Thus this interval is not $\omega^2$-large.
-
-\noindent
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html\#mlarge}{Epsilon0.Large\_Sets\_Examples}}
-
-\input{movies/snippets/Large_Sets_Examples/gnawEx1}
-
-The interval $[6,700]$ is $\omega^2$-large, but not
-$\omega^2$-mlarge, since $[6,699]$ is also $\omega^2$-large.
-
-\input{movies/snippets/Large_Sets_Examples/gnawEx2}
-
-The following lemma relates minimal largeness with the
-function 
-\texttt{gnaw}.
-
-\input{movies/snippets/Large_Sets/mlargeIff}
-
-\vspace{4pt}
- \noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html}{Epsilon0.Large\_Sets\_Examples}}
-
-\input{movies/snippets/Large_Sets_Examples/Ex1Lemma}
-
-\section{Length of minimal large sequences}
-
-Now, consider any natural number $k>0$ any ordinal $0<\alpha<\epsilon_0$.  We would like to compute
-a number $l$ such that the interval $[k,l]$ is $\alpha$-mlarge. So, 
-the standard battle starting with $\iota(\alpha)$ and the replication factor $k$ will end after $(l-k+1)$ steps.
-
-First, we notice that this  number $l$ exists, since the segment $[0,\epsilon_0)$ is well-founded and $\canonseq{\alpha}{i}<\alpha$ for any $i$ and $\alpha>0$.
-Moreover, it is unique:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/mlargeUnicity}
-
-Thus, we would like to define a function, parameterized by $\alpha$ which associates to any  strictly positive integer $k$ the number $l$ such that
-the interval $[k,l]$ is $\alpha$-mlarge. It would be fine to write in \gallina{} a definition like this:
-
-\begin{Coqbad}
-Function L_ (alpha: E0) (i:nat) :  nat := ...
-\end{Coqbad}
-
-But we do not know how to fill the dots yet \dots{}   In the next section, we will 
-use \coq{} to reason  about the \emph{specification} of \texttt{L},
-prove properties of any function which satisfies this specification.
-In Sect.~\ref{sect:L-equations}, we use the \texttt{coq-equations} plug-in
-to define a function \texttt{L\_}, and prove its correctness w.r.t. its specification.
-
-\subsection{Formal specification}
-
-Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider any  function which  maps  any strictly positive integer $k$ to the number $l$, where 
-the interval $[k,l)$ is $\alpha$-mlarge.
-
-\begin{remark}
-In~\cite{KS81} Ketonen and Solovay consider the least natural number $l$ where the interval $[k,l]$ ($l$ included) is $\alpha$-large, and call $H_\alpha$ the function which maps $k$ to $l$.
-We chose to consider intervals $[l,k)$ instead of $[l,k]$
-in order to simplify  some statements and proofs in composition lemmas associated with the ordinals of the form $\alpha\times i$ and 
-$\omega^\alpha\times i + \beta$.
-Clearly, both approaches are related through the equality
-$L_\alpha(k)=H_\alpha(k)+1$, for any non-null $\alpha$ and $k$.
-\end{remark}
-
-Our specification of the function \texttt{L} is as follows:
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/LSpecDef}
-
-\begin{todo}
- Check if the functions $L_\alpha$ are the same as
-\cite{KS81}' functions $f_\alpha$ (p. 297).
-\end{todo}
-
-Note that, for $\alpha\not=0$, the value of $f(0)$ is not specified.
-Nevertheless, the restriction of $f$ to the set of strictly positive integers is unique (up to extensionality).
-
-\input{movies/snippets/Large_Sets/LSpecUnicity}
-
-\subsection{Abstract properties}
-
-Let us now prove properties of any function $f$ (if any) which satisfies 
-\texttt{L\_spec}.
-We are looking for properties which could be used for writing \emph{equations} and prove the correctness of the function generated by the \texttt{coq-equations} plug-in.
-Moreover, they will give us some examples (for small values of $\alpha$).
-
-The properties we consider are defined in \href{../theories/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}.
-
-\label{sect:abstract-arith-prop}
-\input{movies/snippets/Iterates/funLeDef}
-
-Our exploration of the $L_\alpha$\,s  considers the usual cases of a proof by transfinite induction: zero, successors and limit ordinals.
-The lemmas we are going to prove will be applied in a big proof by induction in Sect~\vref{sect:L-correct-proof}.
-
-\index{maths}{Transfinite induction}
-
-\subsubsection{The  ordinal zero}
-\label{sect:L-spec-zero}
-The base case is directly a consequence of the specification.
-
-\input{movies/snippets/Large_Sets/LZeroInv}
-
-\subsubsection{Successor ordinals}
-\label{sect:L-spec-succ}
-Let $\beta$ be some ordinal, and assume the arithmetic function $f$ satisfies 
-the specification $(\texttt{L\_spec}\;\beta)$.  Let $k$ be any natural number.
-Any path from $\texttt{succ}\,\beta$ to $0$ starting at $k+1$ can be decomposed into a first step from $\texttt{succ}\,\beta$ to $\beta$, then a path from
-$\beta$ at $k+2$ to $0$. 
-By hypothesis the interval $[k+2, f(k+2)-1]$ is $\beta$-mlarge.
-But the interval $[k+1, f(k+2)-1]$ is the concatenation of the singleton
-$\{k+1\}$ and the interval $[k+2, f(k+2)-1]$.
-So, the function $\lambda\,k.\,f(k+1)$ satisfies the specification $\texttt{L\_spec}\,\beta$.
-
-Note that our decomposition of intervals works only if the intervals we consider are not empty.
-In order to ensure this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <<= f}, or \texttt{(fun\_le S f)}.
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/SectionSucc}
-
-\subsubsection{Limit ordinals}
-\label{sect:L-spec-lim}
-
-Let $\lambda<\epsilon_0$ be any limit ordinal. In a similar way as for successors, we decompose any path from $\lambda$  into a first step to
-$\canonseq{\lambda}{k}$, followed by a path to $0$. In the following section, we assume that there exists a correct function computing  $L_{\canonseq{\lambda}{k}}$ for any strictly positive $k$.
-
-\input{movies/snippets/Large_Sets/SectionLim}
-
-\subsection{First results}
-
-Applying the previous lemmas on successors and limit ordinals, 
-we obtain a few  correct implementations of \texttt{(L\_spec $\alpha$)} for small values of $\alpha$.
-
-\subsubsection{Finite ordinals}
-
-By iterating the functional \texttt{L\_succ}, we get a realization of
-\texttt{(L\_spec $i$)} for any natural number $i$.
-
-\input{movies/snippets/Large_Sets/LFinDef}
-\vspace{-16pt}
-\input{movies/snippets/Large_Sets/LFinOk}
-
-\subsubsection{The first limit ordinal  \texorpdfstring{$\omega$}{omega}}
-
-The lemmas \texttt{L\_fin\_ok} and \texttt{L\_lim\_ok}   allow us to get 
-by diagonalization a correct implementation for 
-\texttt{L\_spec omega}.
-
-\input{movies/snippets/Large_Sets/LOmegaDef}
-\vspace{-16pt}
-\input{movies/snippets/Large_Sets/LOmegaOk}
-
-\subsubsection{Towards  \texorpdfstring{$\omega^2$}{omega*omega}}
-
-We would like to get exact formulas for the ordinal $\omega^2$, a.k.a.
-$\phi_0(2)$. This ordinal is the limit of the sequence $\omega\times i\;(i \in \mathbb{N})$. Thus, we have to study ordinals of this form, then use 
-our lemma on limits.
-
-The following lemma establishes a path from $\omega\times ( i+1)$ to
-$\omega \times i$.
-
-\input{movies/snippets/Large_Sets/pathToOmegaMult}
-
-Let us consider a path from  $\omega\times(i+1)$ to $0$ starting at $k+1$.
-A first ``big step'' will lead to $\omega\times i$ at $2(k+1)$. If $i>0$, the
-next jump leads to $\omega\times(i-1)$ at $2(2(k+1))+1$, etc.
-
-The following lemma expresses the length of the mlarge sequences associated with any finite multiple of $\omega$.
-
-\input{movies/snippets/Large_Sets/omegaMultMlarge0}
-
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#L_omega_mult}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/LOmegaMultDef}
-
-More generally, we prove the equality $L_{\omega\times i}(k)=2^i\times(k+1)-1$.
-
-\input{movies/snippets/Large_Sets/LOmegaMultEqn}
-
-Correctness of the function \texttt{L\_omega\_mult} is asserted through the following lemma.
-
-\input{movies/snippets/Large_Sets/LOmegaMultOk}
-
-By diagonalization, we obtain a simple formula for $L_{\omega^2}$.
-
-\input{movies/snippets/Large_Sets/LOmegaSquare}
-\input{movies/snippets/Large_Sets/LOmegaSquareEqn}
-\input{movies/snippets/Large_Sets/LOmegaSquareOk}
-
-%%%% ICI
-
-\subsubsection{Going further}
-Let us consider a last example, ``computing'' $L_{\omega^3}$.
-Let us study the canonical sequence associated with $\omega^3$, the elements of which are 
-$\omega^2\times i\;(i\in\mathbb{N}_1)$.
-
-To this end, we prove a generic lemma, which expresses $L_{\omega^\alpha\times i}$ as an iterate of $L_{\omega^\alpha}$.
-Note that in this lemma, we assume that the function associated with $\alpha$ is strictly monotonous and
-greater or equal than the successor function, and prove that $L_{\omega^\alpha\times i}$ satisfies  the same properties.
-
-\inputsnippets{Large_Sets/phi0Mult}
-
-\inputsnippets{Large_Sets/phi0MultOk,
-Large_Sets/Phi0MultSLe}
-
-Let us look now
-at the ordinal $\omega^2\times i$, using \texttt{L\_phi0\_mult}.
-
-\input{movies/snippets/Large_Sets/LOmegaSquareTimes}
-
-We are now ready to get an exact formula for $L_{\omega^3}$, by diagonalization over $L_{\omega^2\times i}$.
-
-\input{movies/snippets/Large_Sets/LOmegaCube}
-
-Thus, for instance, $L_{\omega^3}(3)=L_{\omega^2\times 4}(3)$.
-
-\input{movies/snippets/Large_Sets/LOmegaCube3Eq}
-
-This number is quite big. Using \texttt{Ocaml}'s \texttt{float} arithmetic,
-we can under-approximate it by $2^{3.8\times10^{30}}\times 3.8\times{10^{30}}$.
-
-\begin{Coqsrc}
-# let exp2 x = 2.0 ** x;;
-
-val exp2 : float -> float = <fun>
-#   exp2 95.0 *. 97.0 -. 1.0;;
-- : float = 3.84256588194182037e+30
-# let n = exp2 95.0 ;;
-# let p = n *. 97.0 -. 1.0;;
-val p : float = 3.84256588194182037e+30
-
-Estimation :
-2 ** (3.84 e+30) * 3.84 e+30.
-\end{Coqsrc}
-
-\subsection{Using \texttt{Equations}}
-\label{sect:L-equations}
-
-Note that we did not define any function $L_\alpha$ \emph{for any $\alpha<\epsilon_0$} yet.
-We have got no more than a collection of proved realizations of $\texttt{L\_spec}\;\alpha$ for a few values of $\alpha$.
-
-\index{coq}{Plug-ins!Equations}
-
-Using the \texttt{coq-equations} plug-in by 
-M. Sozeau~\cite{sozeau:hal-01671777}, we will now define a function \texttt{L\_} which maps any ordinal  $\alpha<\epsilon_0$ to a proven realization of 
-$\texttt{L\_spec}\;\alpha$.   
-To this end, we represent ordinals as inhabitants of the type 
-\texttt{E0} of well-formed ordinal terms (see Sect~\vref{sect:E0-def}). So, we define a total function \texttt{L\_} of type
-\texttt{E0 -> nat -> nat}, by transfinite recursion, considering the usual three cases : $\alpha=0$, $\alpha$ is a successor, $\alpha$ is a limit ordinal.
-
-\subsubsection{Definition}
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#L_}{L\_alpha}).}
-
-\label{Functions:L-alpha}
- \index{hydras}{Library Epsilon0!Functions!L\_@L\_ (final step of a minimal path}
-
-\input{movies/snippets/L_alpha/LDef}
-
-This definition results in a bunch of automatically generated lemmas. For instance:
-
-\input{movies/snippets/L_alpha/AboutLEquation1}
-
-In most cases, it may be useful to write human-readable  paraphrases of these statements.
-
-\inputsnippets{L_alpha/Paraphrasesa}
-\inputsnippets{L_alpha/Paraphrasesb}
-\inputsnippets{L_alpha/Paraphrasesc}
-
-Using these three lemmas as rewrite rules, we can prove more properties of the functions \texttt{L\_$\alpha$}.
-
-\input{movies/snippets/L_alpha/LFiniteOmega}
-
-By  well-founded induction on $\alpha$, we prove the following properties:
-
-\input{movies/snippets/L_alpha/LGeS}
-
-\input{movies/snippets/L_alpha/LCorrect}
-
-\label{sect:L-correct-proof}
-
-Please note that the proof of \texttt{L\_correct} applies the lemmas proven in Sections~\ref{sect:L-spec-zero}, ~\ref{sect:L-spec-succ} and ~\ref{sect:L-spec-lim}.
-Our previous study of \texttt{L\_spec} allowed us to pave the way for the definition by \texttt{Equations} and the correctness proof.
-
-\paragraph*{\gaiasign}
-The module
-\href{../theories/html/gaia_hydras.GL_alpha.html}{gaia\_hydras.GL\_alpha} contains an adaptation of
-\href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0/L\_alpha}.
-
-\subsubsection{Back to hydra battles}
-\label{def:L-alpha}
-
-Lemma \texttt{battle\_length\_std } of
-Module~\href{../theories/html/hydras.Hydra.Battle_length.html}{Hydra.Battle\_length} relates the length of standard battles with the functions $L_\alpha$.
-
-\input{movies/snippets/Battle_length/battleLengthStd}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Instead of considering standard paths and battles, consider battles where the replication factor is a constant $k$.
-Please use \texttt{Equations} in order to define the function that computes the length of the $k$-path which leads  from $\alpha$ to $0$.
-Prove a few  exact formulas and minoration lemmas.
-\end{exercise}
-
-\section{A variant of the Hardy hierarchy}
-
-\label{sect:hardy}
-
-In order to give a feeling on  the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the  \emph{Hardy hierarchy} of fast growing functions,
-presented for instance in~\cite{Promel2013}. 
-\index{maths}{Rapidly growing functions!Hardy Hierarchy}
-
-\begin{remark}
-  Indeed, the functions presented in this section are a \emph{variant} of the Hardy hierarchy of functions.
-In the future versions of this development, we will correct the references to the literature.
-For the time being, we call our functions $H'_\alpha$ in order to underline the difference from ``classic'' Hardy functions.
-\end{remark}
-
-For each ordinal $\alpha$ below $\epsilon_0$, $H'_\alpha$ is 
-a total arithmetic function, defined  by  transfinite recursion on $\alpha$, according to three cases:
-
-\index{maths}{Transfinite induction}
-
-\begin{itemize}
-\item If $\alpha=0$, then $H'_\alpha (k)= k$ for any natural number $k$.
-\item If $\alpha=\textrm{succ}(\beta)$, then 
-$H'_\alpha(k)=H'_\beta(k+1)$ for any $k \in \mathbb{N}$
-\item If $\alpha$ is a limit ordinal, then 
-$H'_\alpha(k) = H'_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
-\end{itemize}
-
-\begin{remark}
- The ``classic'' definition of the Hardy hierarchy differs in the third equation.
-
-\begin{itemize}
-\item If $\alpha=0$, then $H_\alpha (k)= k$ for any natural number $k$.
-\item If $\alpha=\textrm{succ}(\beta)$, then 
-$H_\alpha(k)=H_\beta(k+1)$ for any $k \in \mathbb{N}$
-\item If $\alpha$ is a limit ordinal, then 
-$H_\alpha(k) = H_{(\canonseq{\alpha}{k})}(k)$ for any $k\in \mathbb{N}$.
-\end{itemize}
-
-\end{remark}
-
-\subsection{Definition in \texttt{Coq}}
-
-We define a function \texttt{H'\_} of type \texttt{E0 -> nat -> nat} by transfinite induction over the type \texttt{E0} of the well formed ordinals below $\epsilon_0$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}}
-
-\index{hydras}{Library Epsilon0!Functions!H\_@H\_ (Hardy hierarchy (variant))}
-\index{coq}{Plug-ins!Equations}
-\label{Functions:Hprime-alpha}
-
-\input{movies/snippets/Hprime/HprimeDef}
-\input{movies/snippets/Hprime/HprimeDefb}
-
-\input{movies/snippets/Hprime/paraphrasesa}
-\input{movies/snippets/Hprime/paraphrasesb}
-\input{movies/snippets/Hprime/paraphrasesc}
-\input{movies/snippets/Hprime/paraphrasesd}
-
-\subsection{First  steps of the H' hierarchy}
-Using rewrite rules from \texttt{H'\_eq1} to \texttt{H'\_succ\_eqn}, we can explore the functions $H'_\alpha$ for  small values of $\alpha$.
-
-\paragraph*{\gaiasign}
-The module
-\href{../theories/html/gaia_hydras.GHprime.html}{gaia\_hydras.GHprime} contains an adaptation of
-\href{../theories/html/hydras.Epsilon0.Hprime.html}{Epsilon0.Hprime}.
-
-\subsubsection{Finite ordinals}
-
-By induction on $i$, we prove a simple expression of \texttt{H'\_ (Fin i)}, where 
-\texttt{Fin $i$}  is the $i$-th finite ordinal.
-
-\input{movies/snippets/Hprime/HprimeFin}
-
-\subsubsection{Multiples of \texorpdfstring{$\omega$}{omega}}
-
-Since the canonical sequence of $\omega$ is composed of finite ordinals, 
-it is easy to get the formula associated with $H'_\omega$.
-
-\input{movies/snippets/Hprime/HprimeOmega}
-
-Before going further, we prove a useful rewriting lemma:
-
-\input{movies/snippets/Hprime/HprimePlusFin}
-
-Then, we get easily formulas for $H'_{\omega+i}$, and $H'_{\omega\times i}$ for any natural number $i$.
-
-\input{movies/snippets/Hprime/HprimeExamplesa}
-\input{movies/snippets/Hprime/HprimeExamplesb}
-\input{movies/snippets/Hprime/HprimeExamplesc}
-\input{movies/snippets/Hprime/HprimeExamplesd}
-
-Since $\omega^2$ is a limit ordinal, we prove the following equality: 
-\[H'_{\omega^2} (k) = 2 ^ {k+1} \times (k+1) - 1\]
-
-\input{movies/snippets/Hprime/HprimeOmegaSqr}
-
-\subsubsection{New limits}
-
-Our next step would be to prove an exact formula for $H'_{\omega^\omega}(k)$.
-Since the canonical sequence of $\omega^\omega$ is composed of all the
-$\omega^i$, we first need to express $H'_{\omega^i}$ for any natural number $i$.
-
-Let $i$ and $k$ be two natural numbers. 
-The ordinal $\canonseq{\omega^(i+1)}{k}$ is the product
-$\omega^i \times k$, so we need also to consider ordinals of this form.
-
-\begin{enumerate}
-\item First,  we express $H'_{\omega^\alpha \times (i+2)}$ in terms of
-$H'_{\omega^\alpha \times (i+1)}$.
-
-\input{movies/snippets/Hprime/HprimeOmegaTerm1}
-
-\item
-Then, we prove by induction on $i$ that $H'_{\omega^\alpha \times (i+1)}$ is just the
-$(i+1)$-th iterate of $H'_{\omega^\alpha}$.
-
-\input{movies/snippets/Hprime/HprimeOmegaTerm}
-
-\item In particular, we derive a formula for $H'_{\omega^{i+1}}$.
-
-\input{movies/snippets/Hprime/HprimeSuccFun}
-\input{movies/snippets/Hprime/HprimePhi0SI}
-
-\item We get now a  formula for $H'_{\omega^3}$:
-
-\input{movies/snippets/Hprime/HprimeOmegaCube}
-\end{enumerate}
-
-\subsubsection{A numerical example}
-\label{sect:bignum}
-
-It is hard to capture the complexity of this function by looking only at this
-``exact'' formula. 
-Let us consider a simple example: the number $H'_{\omega^3}(3)$.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3a}
-
-Thus, the number $H_{\omega^3}(3)$ can be written as four nested applications of $f$.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3b}
-
-In a more classical writing, this number is displayed as follows:
-
-{\Large
-$$
-H'_{\omega^3}(3) =  2 ^ {(2 ^ {N + 1} \, (N+1) )}   \,  (2 ^ {N+1} \, ( N +1) ) - 1
-$$
-}
-
-In order to make this statement more readable, we introduce a local definition.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3c}
-
-This number looks quite big; let us compute an approximation with \texttt{Ocaml}:
-
-\begin{Coqsrc}
-# (2.0 ** 64.0 *. 64.0 -. 1.0);; 
-\end{Coqsrc}
-
-\begin{Coqanswer}
-- : float = 1.1805916207174113e+21
-\end{Coqanswer}
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3d}
-
-We leave as an exercise to determine the best approximation as possible of
- the size of this number (for instance its number of digits).  For instance, if
-we do not take into account the multiplications in the formula above,
-we obtain that, in base $2$, the number $H'_{\omega^3}(3)$ has at least
-$2^{10^{21}}$  digits. But it is still an under-approximation !
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3e}
-
-\subsubsection{A formula for \texorpdfstring{$H'_{\omega^\omega}$}{\texttt{H'(phi0(omega))}}}
-
-Now, we can get at last an exact formula for $H'_{\omega^\omega}$.
-
-\input{movies/snippets/Hprime/HprimePhi0Omega}
-
-Using extensionality of the functional \texttt{iterate}, we also get a closed formula.
-
-\input{movies/snippets/Hprime/HprimePhi0OmegaClosed}
-
-Note that this formula contains two occurrences of the functional \texttt{iterate}, the outer one is in fact a second-order iteration (on type \texttt{nat -> nat)}
-and the inner one  first-order (on type \texttt{nat}).
-
-\subsection{Abstract properties of  
-\texorpdfstring{$H'_\alpha$}{H'}}
-~\label{sect:H-alpha-prop}
-
-Since pure computation seems to be useless for dealing with expressions of the form $H'_\alpha(k)$, even for small values of $\alpha$ and $k$, we need to prove theorems for comparing $H'_\alpha(k)$ and $H'_\beta(l)$, in terms of comparison
-between $\alpha$ and $\beta$ on the one hand, $k$ and $l$ on the other hand.
-
-But beware of fake theorems! For instance, one could believe that $H'$ is monotonous in its first argument. The following proof shows this is false.
-
-\input{movies/snippets/Hprime/HprimeNonMono1}
-
-On the contrary, the functions of the $H'$ hierarchy have the following five properties~\cite{KS81}: for any $\alpha < \epsilon_0$,
-\begin{itemize}
-\item the function $H'_\alpha$ is strictly monotonous :
-      For all $n,p \in\mathbb{N}, n < p \Rightarrow H'_\alpha(n)< H'_\alpha(p)$.
-\item If $\alpha \not= 0$, then for every $n$, $n<H'_\alpha(n)$.
-\item The function $H'_\alpha$ is pointwise less or equal than $H'_{\alpha+1}$
-
-\item For any $n\geq 1$, $H'_\alpha(n)<H'_{\alpha+1}(n)$.
-\emph{We say that $H'_{\alpha+1}$ dominates $H'_\alpha$ from $1$}.
-\item For any $n$ and $\beta$, if $\alpha \xrightarrow[n]{} \beta$, then
-$H'_\beta(n)\leq H'_\alpha(n)$.
-\end{itemize}
-
-\index{maths}{Abstract properties of arithmetic functions}
-\index{hydras}{Abstract properties of arithmetic functions}
-
-\index{maths}{Transfinite induction}
-
-In \coq{}, we follow the  proof of~\cite{KS81}. This proof is mainly a single  proof by transfinite induction on $\alpha$ of the conjunction of the five properties.
-For each $\alpha$, the three cases : $\alpha=0$, $\alpha$ is a limit, and 
-$\alpha$ is a successor are considered. Inside each case, the five sub-properties are proved sequentially, using the abstract properties defined in Sect~\vref{sect:abstract-arith-prop}
-
-\input{movies/snippets/Hprime/PAP}
-
-Using a few lemmas \emph{à la} Ketonen-Solovay, we prove that
-if $\alpha<\beta$, then $H'_\beta$ eventually dominates
-$H'_\alpha$.
-We let the reader look at the proof (Section \texttt{Proof\_of\_H'\_mono\_l} of \href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}).
-
-\noindent
-\input{movies/snippets/Hprime/HprimeDom}
-
-%ICI
-\subsection{Comparison between \texttt{L\_} and \texttt{H'\_} }
-
-By well-founded induction on $\alpha$, we prove that
- $H'_\alpha(i) \leq L_\alpha(i+1)$ for any $i$.
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#H'_L_}{Epsilon0.L\_alpha}}
-
-\input{movies/snippets/L_alpha/HprimeL}
-
-\subsubsection{Back to hydras}
-
-The following theorem relates the length of (standard) battles with the the $H'$ family of fast growing functions.
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems}}
-
-\input{movies/snippets/Hydra_Theorems/battleLengthStdHardy}
-
-Since $H'_{\omega^3}(3)= H'_{\omega^3+3}(0)$, the big number shown in Section~\vref{sect:bignum} is less or equal than 
-the number of steps of the battle with the initial hydra of
-Figure~\ref{fig:start}, and  $0$ as initial replication factor.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[very thick, scale=0.3]
-  \node (h1) at (4,0){$\bullet$};
-  \node (h2) at (2,2){$\bullet$};
-  \node (h3) at (0,4){$\bullet$};
-  \node (h4) at (1,4){$\bullet$};
-  \node (h5) at (2,4){$\bullet$};
-  \node (h6) at (4.5,2){$\bullet$};
-  \node (h7) at (6.5,2){$\bullet$};
-  \node (h8) at (8.5,2){$\bullet$};
-    \draw (h1) -- (h2) ;
-    \draw (h2) -- (h3) ;
-    \draw (h2) -- (h4) ;
-    \draw (h2) -- (h5) ;
-    \draw (h1) -- (h6) ;
-    \draw (h1) -- (h7) ;
-    \draw (h1) -- (h8);
- \end{tikzpicture}
-
-  \caption{The hydra corresponding to the ordinal $\omega^3+3$}
-  \label{fig:start}
-\end{figure}
-
-\section{A variant of the Wainer hierarchy (functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
-\label{sect:wainer}
-
-\index{maths}{Rapidly growing functions!Wainer Hierarchy}
-
-Ketonen and Solovay introduce in~\cite{KS81} a ``trivial'' variant of the Wainer hierarchy~\cite{BW85, Wainer1970} of fast growing functions, indexed by ordinals below $\epsilon_0$.
-The functions $F_\alpha$ are defined by the following equations.
-
-\label{F_equations}
-\begin{itemize}
-\item $F_0(i)=i+1$
-\item $F_{\beta+1}(i)= (F_\beta)^{(i+1)}(i)$, where $f^{(i)}$ is the $i$-th iterate of $f$.
-\item $F_\alpha(i) = F_{\canonseq{\alpha}{i}} (i)$ if $\alpha$ is a limit ordinal.
-\end{itemize}
-
-\begin{remark}
-The difference with the ``classic'' Wainer hierarchy 
-$f_\alpha\;(\alpha<\epsilon_0)$ lies in the second equation:
-$f_{\beta+1}(i) = (f_\beta)^{(i)}(i)$ and not
-$f_{\beta+1}(i) = (f_\beta)^{(i+1)}(i)$.
-
-A module about 
-the classic Wainer hierarchy is in preparation.
-
-Note also that \cite{KS81} defines also $F_{\epsilon_0}$ (by the third equation).
-Since $\epsilon_0$ is not representable in type \texttt{E0}, our implementation in \coq{} does not take $F_{\epsilon_0}$ into account.
-
-\end{remark}
-
-A first attempt is to write a definition of $F_\alpha$ by equations, in the same way as for $H_\alpha$ (the functional \texttt{iterate} has already been used in Sect.\vref{Functions:iterate}).
-
-\index{hydras}{Library Prelude!iterate}
-
-\input{movies/snippets/Iterates/iterateDef}
-
-\emph{From\href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}~}.
-
-\index{coq}{Plug-ins!Equations}
-
- \input{movies/snippets/F_alpha/FailDemo}
-
-We presume that this error comes from the recursive call of \texttt{F\_} inside
-an application of \texttt{iterate}. The workaround we propose is to define first 
-the iteration of \texttt{F\_}  as an helper $F^*$, then to define the function $F$ as a ``iterating $F^*$ once''.
-
-\texttt{Equations} accepts the following definition, relying on  lexicographic ordering on pairs $(\alpha,n)$.
-
-\label{sect:F-equations}
-
-\index{coq}{Plug-ins!Equations}
-\label{Functions:F-alpha}
-\index{maths}{Rapidly growing functions}
-\index{hydras}{Library Epsilon0!Functions!F\_@F\_ (Wainer hierarchy)}
-
-\input{movies/snippets/F_alpha/goodDefa}
-\input{movies/snippets/F_alpha/goodDefb}
-
-It is quite easy to prove that our functional \texttt{F\_} satisfies the equations on page~\pageref{sect:F-equations}.
-\index{hydras}{Library Prelude!iterate}
-
-\input{movies/snippets/F_alpha/FEquations}
-
-As for the Hardy functions, we can use these equalities as rewrite rules for
-``computing'' some values of $F_\alpha(i)$, for small values of $\alpha$.
-
-\input{movies/snippets/F_alpha/FirstValues}
-
-Like in Sect~\ref{sect:H-alpha-prop}, we prove by induction the following properties (see~\cite{KS81}).
-
-\input{movies/snippets/F_alpha/FalphaThms}
-
-As a corollary, we prove that, if $\beta<\alpha$, then
-$F_\alpha$ dominates $F_\beta$ (p. 284 of~\cite{KS81}).
-
-\input{movies/snippets/F_alpha/FDomContext}
-\input{movies/snippets/F_alpha/FDom}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove the following property:
-
-\begin{Coqsrc}
-Lemma LF3 : dominates_from  2 (F_ 3) (fun  n => iterate exp2 n n).
-\end{Coqsrc}
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
-\end{exercise}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that, for any $\alpha\geq 3$ and $n\geq 2$,
-$F_\alpha(1+n)\geq 2^{F_\alpha(n)}$.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
-\end{exercise}
-
-%ici
-\index{hydras}{Exercises}
-
-\begin{exercise}
-It is tempting to prove a simple property of monotony 
-of the function \texttt{F\_}.
-
-\begin{quote}
-   Let $\alpha\leq\beta<\epsilon_0$. For any $n\geq 2$,
-$F_\alpha(n)\leq F_\beta(n)$. 
-\end{quote}
-Prove or disprove this statement.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/is_F_monotonous.v}{exercises/ordinals/is\_F\_monotonous.v}.}
-\end{exercise}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-
-Prove that for any $n\geq 2$, $\textrm{Ack}\,\,n\,n\leq  F_\omega(n)$, where \textrm{Ack} is the Ackermann function.
-Next, prove that $F_\alpha$ is not primitive recursive, for any $\alpha\geq\omega$  (please see Sect.~\vref{sect:ack-not-PR}).
-On the other hand, please show that for any natural number $n$, the function $F_n$ is primitive recursive.
-Thus $F_\alpha$ is primitive recursive if and only if $\alpha$ is finite.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_omega.v}{exercises/ordinals/F\_Omega.v}.
-Properties of the Ackermann function are studied in
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/Ack.v}{theories/ordinals/MoreAck/Ack.v} and
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/AckNotPR.v}{theories/ordinals/MoreAck/AckNotPR.v}
-.}
-\end{exercise}
-
-\paragraph*{\gaiasign} Library~\href{../theories/html/gaia_hydras.GF_alpha.html}{gaia\_hydras.GF\_alpha}  defines the functions $F_\alpha$ for \gaia's world.
-
-\section{More about rapidly growing functions}
-
-In Sect.~\vref{sect:battle-length-notPR}, we prove that the length of hydra-battles (for a given hydra, according to the initial replication factor) is not primitive recursive in general. 
-This proof uses properties of the Ackermann function, and the $H'_\alpha$, $F_\alpha$, $L_\alpha$ families of functions.
-
-% ---------------------------------------------------
-\input{gaia-chapter}
-
-%----------------------------------------------------------------------
-\chapter[Countable ordinals (after Sch\"{u}tte)]{Kurt Schütte's axiomatic definition of countable ordinals}
-
-\label{chap:schutte} 
-%ON
-
-In the present chapter, we  compare our implementation of the segment $[0,\epsilon_0)$ with a mathematical text in order to ``validate'' our constructions.
-Our reference here is the axiomatic definition of the set of countable ordinals,
-in chapter V of Kurt Schütte's book `` Proof Theory ''~\cite{schutte}.
-
-\begin{remark}
-\emph{In all this chapter, the word ``ordinal'' will be considered as a synonymous of
-``countable ordinal''}  
-\end{remark}
-
-Schütte's definition of countable ordinals relies on the following three axioms:
-
-There  exists a strictly ordered set , such that
-\begin{enumerate}
-\item  $(\mathbb{O},<)$ is well-ordered
-\item Every bounded subset of $\mathbb{O}$  is countable
-\item Every countable subset of $\mathbb{O}$  is bounded.
-\end{enumerate}
-
-Starting with these three axioms, Schütte re-defines the vocabulary about ordinal numbers: the null ordinal $0$, limits and successors, the addition of ordinals, the infinite ordinals $\omega$, $\epsilon_0$, $\Gamma_0$, etc.
-
-This chapter describes an adaptation to \coq{} of Schütte's axiomatization. 
- Unlike the rest of our libraries, our library
-\href{../theories/html/hydras.Schutte.Schutte.html}{hydras.Schutte}
-is not constructive, and relies on several axioms.
-
-\begin{itemize}
-\item First, please keep in mind  that the set of countable ordinals is not countable.
-Thus, we cannot hope to represent all countable ordinals as finite terms of an inductive type, which was possible with  the set of ordinals strictly less than $\epsilon_0$ (resp.
-$\Gamma_0$)
-\item We tried to be as close as possible to K.
-Schütte's text, which uses ``classical'' mathematics : excluded middle, Hilbert's $\epsilon$ (choice) and Russel's $\iota$ (definite description) operators.
-Both operators allow us to write definitions close to the natural mathematical language, such as ``$\textrm{succ}$ is \emph{the} least ordinal strictly greater than $\alpha$''
-\item Please note that only the library \href{../theories/html/hydras.Schutte.Schutte.html}{Schutte/*.v} is ``contaminated'' by axioms, and that the rest of our libraries remain constructive.
-\end{itemize}
-
-\section{Declarations and axioms}
-
-Let us declare a type 
-\texttt{Ord} for representing countable ordinals, and a binary relation
- \texttt{lt}. Note that, in our development, \texttt{Ord} is a type, while the \emph{set} of countable ordinals (called $\mathbb{O}$ by Schütte) 
-is the full set over the type \texttt{Ord}.
-
-\label{types:Ord}
-
-We use Florian Hatat's library on countable sets, written as he was a student of  \emph{\'Ecole Normale Supérieure de Lyon}.
-A set $A$ is countable if there is an injective function from $A$ to $\mathbb{N}$ (see
-Library \href{../theories/html/hydras.Schutte.Countable.html}%
-{\texttt{Schutte.Countable}}).
-
-\vspace{6pt}
-
-\emph{From Module\href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{Schutte.Schutte\_basics}}}
-
-\index{hydras}{Library Schutte!Types!Ord}
-
-\input{movies/snippets/Schutte_basics/OrdDecl}
-
-Schütte's first axiom tells that \texttt{lt} is a well order on the set 
-\texttt{ordinal} (The  class \texttt{WO} is defined in
-Module~\href{../theories/html/hydras.Schutte.Well_Orders.html}{Schutte.Well\_Orders.v}).
-
-\index{hydras}{Library Schutte!Type classes!WO@ WO (well order)}
-
-\label{types:WO}
-
-\input{movies/snippets/Well_Orders/Mdecl}
-
-\input{movies/snippets/Well_Orders/WODef}
-
-\input{movies/snippets/Schutte_basics/AX1}
-
-The second and third axioms say that a subset $X$ of $\mathbb{O}$ is
-(strictly) bounded if and only if it is countable.
-
-\input{movies/snippets/Schutte_basics/AX23}
-
-\texttt{AX2} and \texttt{AX3} could have been replaced by a single axiom (using the \texttt{iff} connector), but we decided to respect as most as possible the structure of Schütte's definitions.
-
-\section{Additional  axioms}
-
-The adaptation of Schütte's mathematical discourse to \coq{} led us to
-import a few axioms from the standard library. We encourage the reader to consult \coq{}'s FAQ about the safe use of axioms
- \url{https://github.com/coq/coq/wiki/The-Logic-of-Coq#axioms}.
-
-\subsubsection{Classical logic}
-
-In order to work with classical logic, we import the module
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Classical.html}{Coq.Logic.Classical}  of \coq{}'s standard library, specifically the following axiom:
-
-\begin{Coqsrc}
- Axiom classic : forall P:Prop, P \/ ~P.
-\end{Coqsrc}
-
-\subsubsection{Description operators}
-
-In order to respect Schütte's style, we imported also the library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Epsilon.html}{\texttt{Coq.Logic.Epsilon}}.  The rest of this section presents a few examples of
-how Hilbert's choice operator and Church's definite description allow us
- to write understandable definitions (close to the mathematical natural language).
-
-\subsubsection{The definition of zero}
-
-According to the  definition of a well order, every non-empty subset of \texttt{Ord} has a least element. Furthermore, this least element is unique. We would like to call this element  \texttt{zero}.
-
-\input{movies/snippets/Schutte_basics/iotaDemoa}
-\input{movies/snippets/Schutte_basics/iotaDemob}
-
-Indeed, the basic logic of  \coq{} does not allow us to eliminate a proof of a proposition 
-$\exists!\,x:A,\,P(x)$ for building a term whose type lies in the sort \texttt{Type}. 
-The reasons for this impossibility are explained in many documents~\cite{BC04, chlipalacpdt2011, Coq}.
-
-Let us import the library \texttt{Coq.Logic.Epsilon}, which contains the following axiom and lemmas.
-
-\input{movies/snippets/MoreEpsilonIota/EpsilonStatement}
-
-Hilbert's $\epsilon$ \emph{operator} is derived from this  axiom.
-
-\input{movies/snippets/MoreEpsilonIota/EpsilonDef}
-
-If we consider the \emph{unique existential} quantifier $\exists!$, we obtain
-Church's \emph{definite description operator}.
-
-\input{movies/snippets/MoreEpsilonIota/iotaDef}
-
-Indeed, the operators \texttt{epsilon} and \texttt{iota} allowed us to make our definitions 
-quite close to Schütte's text. Our libraries \href{../theories/html/hydras.Schutte.MoreEpsilonIota.html}%
-{\texttt{Schutte.MoreEpsilonIota}}
-and
-\href{../theories/html/hydras.Schutte.PartialFun.html}%
-{\texttt{Schutte.PartialFun}} are extensions of \texttt{Coq.logic.Epsilon} for making easier 
-such definitions. See also an article in french~\cite{PCiota}.
-
-\input{movies/snippets/MoreEpsilonIota/Defs}
-
-In order to use these tools,  we have to tell \coq{}  that the declared type \texttt{Ord} is not empty:
-
-\input{movies/snippets/Schutte_basics/inhOrd}
-
-We are now able to define \texttt{zero} as the least ordinal. For this purpose,
-we define a function returning the least element of any [non-empty]  subset.
-
-\emph{From Module\href{../theories/html/hydras.Schutte.Well_Orders.html}%
-  {\texttt{Schutte.Well\_Orders}}}
-
-\input{movies/snippets/Well_Orders/MDecl}
-\input{movies/snippets/Well_Orders/theLeast}
-
-\vspace{4pt}
-
-From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{~Schutte.Schutte\_basics}}
-
-\label{Constants:zero:Ord}
-\index{hydras}{Library Schutte!Constants!zero}
-
-\input{movies/snippets/Schutte_basics/zeroDef}
-
-We want to prove now that zero is less than or equal to any ordinal number.
-
-\input{movies/snippets/Schutte_basics/zeroLe}
-
-\subsubsection{Remarks on \texttt{epsilon} and \texttt{iota}}
-
- What would happen in case of a misuse of \texttt{epsilon} or \texttt{iota} ?
-For instance, one could give a unsatisfiable specification to \texttt{epsilon} or 
-a specification for \texttt{iota} that admits several realizations.
-
-Let us consider an example:
-
-\input{movies/snippets/Schutte_basics/BadBottoma}
-
-Since we won't be able to prove the proposition
-\linebreak \Verb|(exists! a: Ord, least_member (Empty_set Ord) a)|, the only properties we would be able to prove about \texttt{bottom} would be \emph{trivial} properties, 
-\emph{i.e.}, satisfied by \emph{any} element of type \texttt{Ord}, like for instance
-\texttt{bottom = bottom}, or \texttt{zero <= bottom}.
-
-\input{movies/snippets/Schutte_basics/trivialProps}
-
-On the other hand, the following attempt fails, because of the unprovable first subgoal (please notice that the second subgoal is easy to solve !).
-
-\input{movies/snippets/Schutte_basics/Failure}
-
-In short, using \texttt{epsilon} and \texttt{iota} in our implementation of countable ordinals after Schütte has two main advantages.
-
-\begin{itemize}
-\item It allows us to give a \emph{name} (using \texttt{Definition}) two witnesses 
-of existential quantifiers (let us recall that, in classical logic, one may consider non-constructive proofs of existential statements)
-\item By separating definitions from proofs of [unique] existence, one may make definitions  more concise and readable. Look for instance at 
-the definitions of  \texttt{zero}, \texttt{succ}, \texttt{plus}, etc. in the rest of this chapter.
-\end{itemize}
-%%%% ICI ICI
-
-\section{The  successor function}
-
-The definition of the function \texttt{succ:Ord -> Ord} is very concise. The successor of any ordinal $\alpha$ is the smallest ordinal strictly greater than $\alpha$.
-
-\label{Functions:succ-sch}
-\index{hydras}{Library Schutte!Functions!succ}
-
-\input{movies/snippets/Schutte_basics/succDef}
-
-Using \texttt{succ}, we define the following predicates.
-
-\input{movies/snippets/Schutte_basics/isSuccIsLimit}
-
-% \begin{remark}
-% Please look at remark~\vref{warning:coercions}.
-
-How do we prove properties of the successor function?
-First, we make its specification explicit.
-
-\input{movies/snippets/Schutte_basics/succSpec}
-
-Then, we prove that our function \texttt{succ} meets this specification.
-
-\input{movies/snippets/Schutte_basics/succOka}
-
-We have now to prove that the set of all ordinals strictly greater than $\alpha$ has a unique least element.
-But the singleton set $\{\alpha\}$ is countable, hence  bounded (by the axiom \texttt{AX3}).
-Hence; the set $\{\beta\in\mathbb{O}|\alpha < \beta\}$ is not empty
-and therefore has a unique least element.
-
-The rest of the \coq{} proof script is quite short.
-
-\input{movies/snippets/Schutte_basics/succOkb}
-
-We can ``uncap'' the description operator for proving properties of the
-\texttt{succ} function.
-
-\input{movies/snippets/Schutte_basics/succProps}
-
-\section{Finite ordinals}
-
-Using \texttt{succ}, it is now easy to define recursively all the finite ordinals.
-
-\label{sect:notation-F-sch}
-
-\input{movies/snippets/Schutte_basics/finiteDef}
-
-\section{The definition of \texttt{omega}}
-In order to define $\omega$, the first infinite ordinal, we use an operator which
-``returns'' the least upper bound (if it exists) of a subset $X\subseteq \mathbb{O}$.
-For that purpose, we first use a predicate:
-(\texttt{is\_lub $D$ \textit{lt} $X$ $a$}) if $a$ belongs to $D$ and is the least 
-upper bound  of $X$ (with respect to \textit{lt}).
-
-\input{movies/snippets/Lub/isLubDef}
-
-\input{movies/snippets/Schutte_basics/supDef}
-
-Then, we define the function \texttt{omega\_limit} which returns the least upper bound 
-of the  (denumerable) range of any sequence \texttt{s: nat -> Ord}. 
-By \texttt{AX3} this range is bounded, hence the set of its upper bounds is not empty and has a least element.
-Then we define \texttt{omega} as the limit of the sequence of finite ordinals.
-
-\input{movies/snippets/Schutte_basics/omegaDef}
-
-\label{sect:notation-omega}
-
-Among the numerous properties of the ordinal $\omega$, let us quote the following ones
-(proved in Module 
-\href{../theories/html/hydras.Schutte.Schutte_basics.html\#finite_lt_omega}{\texttt{Schutte.Schutte\_basics}})
-
-\input{movies/snippets/Schutte_basics/omegaPropsa}
-\input{movies/snippets/Schutte_basics/omegaPropsb}
-\input{movies/snippets/Schutte_basics/omegaPropsc}
-\input{movies/snippets/Schutte_basics/omegaPropsd}
-
-\subsection{Ordering functions and ordinal addition}
-
-After having defined the finite ordinals and the infinite ordinal $\omega$, we  define the sum $\alpha+\beta$ of two countable ordinals.
-Schütte's definition looks like the following one:
-
-\begin{quote}
-``$\alpha+\beta$ is the $\beta$-th ordinal greater than or equal to $\alpha$''
-\end{quote}
-
-The purpose of this section is to give a meaning to the construction
-``the $\alpha$-th element of $X$''  where $X$ is any non-empty subset of $\mathbb{O}$.
-We follow Schütte's approach, by defining the notion of \emph{ordering functions},
-a way to associate a unique ordinal to each element of $X$.
-Complete definitions and proofs can be found in Module
- \href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
-{\texttt{Schutte.Ordering\_Functions}} ).
-
-\subsection{Definitions}
-
-A \emph{segment} is a set $A$ of ordinals such that, whenever  $\alpha\in A$ and
-$\beta<\alpha$, then $\beta\in A$; a segment is  \emph{proper} if it strictly included in $\mathbb{O}$.
-
-\input{movies/snippets/Ordering_Functions/segmentDef}
-
-Let  $A$ be a segment, and $B$ a subset of $\mathbb{O}$ : an \emph{ordering function for $A$ and  $B$} is a strictly increasing bijection from $A$ to $B$.
-The set $B$ is said to be an \emph{ordering segment} of $A$.
-Our definition in \coq{} is a direct translation of the mathematical text of~\cite{schutte}.
-
-\index{maths}{Ordinal numbers!Ordering functions}
-\index{hydras}{Library Schutte!Predicates!ordering function@ordering\_function}
-
-\input{movies/snippets/Ordering_Functions/orderingFunctionDef}
-
-We are now able to associate with any subset $B$ of $\mathbb{O}$ its ordering segment and ordering function.
-
-\input{movies/snippets/Ordering_Functions/ordDef}
-
-Thus (\texttt{ord $B \;\alpha$}) is the $\alpha$-th element of $B$.
-Please note that the last definition uses the epsilon-based operator \texttt{some} and
-not \texttt{the}. This is due to the fact that we cannot prove the unicity (w.r.t. Leibniz' equality) of the ordering function of a given set. 
-By contrast, we admit the axiom  \texttt{Extensionality\_Ensembles}, from the library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Sets.Ensembles.html}{Coq.Sets.Ensembles}, so we use the operator \texttt{the} in the definition of
-\texttt{the\_ordering\_segment}.
-
-One of the main theorems of
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html\#ordering_function_ex}%
-{\texttt{Ordering\_Functions}} 
-associates a unique segment and a unique (up to extensionality) ordering function to every subset $B$ of $\mathbb{O}$.
-
-\input{movies/snippets/Ordering_Functions/orderingFunctionEx}
-
-Thus,  our function \texttt{ord}  which enumerates the elements of $B$ is defined in a non-ambiguous way.
-Let us quote the following theorems (see Library
-\href{../theories/html/hydras.Schutte.Ordering_Functions.html}%
-{\texttt{Schutte.Ordering\_Functions}} for more details).
-
-\input{movies/snippets/Ordering_Functions/orderingLe}
-\input{movies/snippets/Ordering_Functions/Th1352}
-
-\subsection{Ordinal addition}
-
-We are now ready to define and study addition on the type \texttt{Ord}.
-The following definitions and proofs can be consulted in Module
-\href{../theories/html/hydras.Schutte.Addition.html}%
-{\texttt{Schutte.Addition.v}}.
-
-\index{hydras}{Library Schutte!Functions!plus}
-
-\input{movies/snippets/Addition/additionDef}
-
-In other words,  $\alpha + \beta$ is the  $\beta$-th ordinal greater than or equal to $\alpha$. 
-Thanks to generic properties of ordering functions, we can show the following 
-properties of addition on $\mathbb{O}$. First, we prove a useful lemma:
-
-\input{movies/snippets/Addition/plusElim}
-
-As a use-case, let us prove that $0$ is a right neutral element of $+$.
-
-\input{movies/snippets/Addition/alphaPlusZero}
-
-The following lemmas are proved the same way.
-
-\input{movies/snippets/Addition/bunchOfLemmas}
-
-The following lemmas are not direct applications of \texttt{plus\_elim}.
-
-\input{movies/snippets/Addition/plusAssoc}
-\input{movies/snippets/Addition/finitePlusInfinite}
-
-It is interesting to compare the proof of these lemmas with the
-computational proofs of the corresponding statements in Module
-\href{../theories/html/hydras.Epsilon0.T1.html}%
-{\texttt{Epsilon0.T1}}. 
-For instance, the proof of the lemma 
-\texttt{one\_plus\_omega} uses the continuity of ordering functions (applied to  \texttt{(plus 1)}) and compares the limit of the $\omega$-sequences $i_{(i \in \mathbb{N})}$ and
-$(1+i)i_{(i \in \mathbb{N})}$, whereas in the library  \texttt{Epsilon0/T1}, the equality 
-$1+\omega=\omega$ is just proved with \texttt{reflexivity}!
-
-\subsubsection{Multiplication by a natural number}
-
-The multiplication of an ordinal by a natural number is defined in terms of addition.
-This operation is useful for the study of Cantor normal forms.
-
-\input{movies/snippets/Addition/multFin}
-
-\section{The exponential of basis \texorpdfstring{$\omega$}{omega}}
-
-In this section, we define the function which maps any $\alpha\in\mathbb{O}$ to
-the ordinal  $\omega^\alpha$, also written 
-$\phi_0(\alpha)$. 
-It is an opportunity to apply the definitions and results of the preceding section. 
-Indeed,  Schütte first defines a subset of $\mathbb{O}$: the set of additive principal ordinals, and $\phi_0$  is just defined as the ordering function of this set.
-
-\subsection{Additive principal ordinals}
-
-\index{maths}{Ordinal numbers!Additive principal ordinals}
-\index{hydras}{Library Schutte!Predicates!AP@AP (additive principal ordinals)}
-
-\begin{definition}
-A non-zero ordinal  $\alpha$ is said to be \emph{additive principal} if, for all  $\beta<\alpha$, $\beta+\alpha$ is equal to  $\alpha$.
-We call \texttt{AP} the set of additive principal ordinals.
-
-\end{definition}
-
-\noindent\emph{From Module \href{../theories/html/hydras.Schutte.AP.html}%
-{\texttt{Schutte.AP}}}
-
-\input{movies/snippets/AP/APDef}
-
-\subsection{The function \texttt{phi0}}
-
-Let us call  $\phi_0$ the ordering function of \texttt{AP}.
-In the mathematical text, we shall use indifferently the notations  $\omega^\alpha$ and$\phi_0(\alpha)$.
-
-\index{hydras}{Library Schutte!Functions!phi0}
-
-\input{movies/snippets/AP/phi0Def}
-
-\subsection{Omega-towers and the ordinal \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-Using $\phi_0$, we can define recursively the set of finite omega-towers.
-
-\input{movies/snippets/AP/omegaTower}
-
-\label{sect:epsilon0-as-limit}
-Then, the ordinal  $\epsilon_0$ is defined as the limit of the sequence of all finite towers (a kind of infinite tower).
-
-\input{movies/snippets/AP/epsilon0Def}
-
-The rest of our library \texttt{AP} is devoted to the proof of properties of additive principal ordinals, hence of the ordering function  $\phi0$ and the ordinal $\epsilon_0$ (which we could not express within the type \texttt{T1}).
-
-\subsection{Properties of the set  \texttt{AP}}
-
-The set of additive principal ordinals is not empty: it contains at least the ordinals  $1$ and  $\omega$.
-
-\inputsnippets{AP/APOne,AP/APOne_least_AP,AP/APOne_AP_omega,AP/APOne_omega_second_AP}
-
-The set  \texttt{AP} is  \emph{closed} under addition, and unbounded.
-\label{lemma:AP-plus-closed}
-
-\input{movies/snippets/AP/APPlusClosed}
-
-\input{movies/snippets/AP/APUnbounded}
-
-Finally, \texttt{AP} is (topologically) \emph{closed} and ordered by the segment of all countable ordinals.
-
-\index{hydras}{Library Schutte!Predicates!Closed}
-
-From Module \href{../theories/html/hydras.Schutte.Schutte_basics.html}%
-{\texttt{~Schutte.Schutte\_basics}}
-
-\input{movies/snippets/Schutte_basics/ClosedDef}
-\input{movies/snippets/AP/APClosed}
-
-\subsubsection{Properties of the function \texorpdfstring{$\phi_0$}{phi0}}
-
-The ordering function $\phi_0$ of the set \texttt{AP} is defined on the full set $\mathbb{O}$ and is continuous (Schütte calls such a  function  \emph{normal}).
-
-\begin{Coqsrc}
-Theorem normal_phi0 : normal phi0 AP.
-\end{Coqsrc}
-
-The following properties come from  the definition of $\phi_0$ as the ordering function of \texttt{AP}.
-It may be interesting to compare these proofs with the computational ones described in Chapter ~\ref{chap:T1}.
-
-\input{movies/snippets/AP/APPhi0}
-
-\subsection{A last example}
-\label{Ex42-schutte}
-
-Let us prove again the equality $\omega+42+\omega^2= \omega^2$.Let us recall that $\omega^2$ is an abbreviation of $\phi_0(2)$,
-\emph{i.e} the third  additive principal ordinal.
-
-\inputsnippets{Schutte/Ex42a}
-
-Our proof is very different from the computational proof of Sect~\vref{Ex42-E0}.
-By definition of additive principal ordinals, 
-it suffices to prove the inequality $\omega+42< \phi_0(2)$.
-
-\inputsnippets{Schutte/Ex42b}
-
-Since the set \texttt{AP} of additive principals  is closed under addition
-(by Lemma \texttt{AP\_plus\_closed}, page~\pageref{lemma:AP-plus-closed}) , it suffices to prove the inequalities $\omega<\phi_0(2)$ and $42<\phi_0(2)$.
-
-\inputsnippets{Schutte/Ex42d, Schutte/Ex42c, Schutte/Ex42e}
-
-\section{More about \texorpdfstring{$\epsilon_0$}{\texttt{epsilon0}}}
-
-Let us recall that the limit ordinal  $\epsilon_0$ cannot be written within the type \texttt{T1}.
-Since we are now considering the set of all countable ordinals, we can now prove some properties of this ordinal.
-
-We prove the inequality  $\alpha<\omega^\alpha$ whenever $\alpha < \epsilon_0$.
-\emph{Note that this condition was implicit in Module
-\href{../theories/html/hydras.Epsilon0.T1.html\#lt_phi0}{Epsilon0.T1}.}
-
-\input{movies/snippets/AP/ltPhi0}
-
-The proof is as follows:
-\begin{enumerate}
-\item Since $\alpha<\epsilon_0$, consider the least $i$ such that $\alpha$ is strictly less than the omega-tower of height $i$.
-\item
-  \begin{itemize}
-  \item If $i=0$, then the result is trivial (because $\alpha=0$)
- \item  Otherwise let $i=j+1$; 
-          $\alpha$ is greater than or equal to the omega-tower of height $j$.
-         By monotony,  $\phi_0(\alpha)$ is greater than or equal to 
-        the omega-tower of height $j+1$, thus strictly greater than $\alpha$
-  \end{itemize}
- \end{enumerate}
-
-Moreover,  $\epsilon_0$ is the least ordinal $\alpha$ that verifies the equality 
-$\alpha = \omega^\alpha$, in other words, the least fixpoint of the function  $\phi_0$.
-
-\input{movies/snippets/AP/epsilon0Lfp}
-
-\section{Critical ordinals}
-
-\index{maths}{Ordinal numbers!Critical ordinals}
-\index{hydras}{Library Schutte!Predicates!Cr@Cr (critical ordinals)}
-
-For any  (countable) ordinal $\alpha$, the set $\textit{Cr}(\alpha)$ is inductively defined 
-as follows by Schütte (p.81 of~\cite{schutte}).
-
-\begin{quote}
-  \begin{itemize}
-  \item $\textit{Cr}(0)$ is the set \textit{AP} of additive principal ordinals.
-  \item If $0<\alpha$, then $\textit{Cr}(\alpha)$ is the intersection of all the sets of fixpoints of the $\textit{Cr}(\beta)$ for $\beta<\alpha$.
-  \end{itemize}
-\end{quote}
-
-This definition is translated in \coq{} in 
-Module \href{../theories/html/hydras.Schutte.Critical.html}%
-{\texttt{Schutte.Critical}}, as the least fixpoint of a functional.
-
-\input{movies/snippets/Critical/CrDef}
-
-Lets us denote by $\phi_\alpha$ the ordering function of the set $\textit{Cr}(\alpha)$ and by $A_\alpha$ its ordering segment.
-
-\input{movies/snippets/Critical/phiDef}
-
-\label{sect:phi-schutte}
-
-For instance,  we prove that $\textit{Cr}(0)$ is the set of additive principals and that $\epsilon_0$
-belongs to $\textit{Cr}(1)$.
-
-\inputsnippets{Critical/CrZeroAP,
-  Critical/epsilon0Cr1}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
- Prove that $\epsilon_0$ is the least element of $\textit{Cr}(1)$.
-\end{exercise}
-
-\subsection{A flavor of infinity}
-
-The family of the $\textit{Cr}(\alpha)$s is made of infinitely many unbounded (hence infinite) sets.
-Let us quote Lemma 5, p. 82  of~\cite{schutte}:
-\begin{quote}
-  For all $\alpha$, the set $\textit{Cr}(\alpha)$ is closed (for the least upper bound of non-empty countable sets) and unbounded.
-\end{quote}
-
-We prove this result by a transfinite induction on $\alpha$ of the conjunction of  both properties.
-
-\index{maths}{Transfinite induction}
-
-\input{movies/snippets/Critical/Lemma5}
-
-\section{Cantor normal form}
-
-The notion of Cantor normal form is defined for all countable ordinals.
-Nevertheless, note that, contrary to the implementation based on type \texttt{T1},
-the Cantor normal form of an ordinal $\alpha$ may contain $\alpha$ as a 
-sub-term\footnote{This would prevent us from trying to represent Cantor normal forms as finite trees (like in Sect.~\ref{sec:T1-inductive-def})}.
-
-\index{maths}{Ordinal numbers!Cantor normal form}
-\index{hydras}{Library Schutte!Predicates!is\_cnf\_of@is\_cnf\_of (to be a Cantor normal form of}
-
-We represent  Cantor normal forms as lists of ordinals.
-A  list $l$ is a Cantor normal form of a given ordinal $\alpha$ if it satisfies two conditions:
-
-\begin{itemize}
-\item The list  $l$ is sorted (in decreasing order) w.r.t. the order $\leq$
-\item The sum of all the  $\omega^{\beta_i}$ where the $\beta_i$ are the terms of $l$ (in this order) is equal to $\alpha$.
-\end{itemize}
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Schutte.CNF.html\#cnf_t}%
-{\texttt{Schutte.CNF}}}
-
-\input{movies/snippets/CNF/Defs}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction on $\alpha$, we prove that every countable ordinal $\alpha$ 
- has at least a Cantor normal form.
-
-\input{movies/snippets/CNF/cnfExists}
-
-By structural induction on lists, we prove that this normal form is unique.
-
-\inputsnippets{CNF/cnfUnicity,CNF/cnfExUnique}
-
-Finally, the following two lemmas relate  $\epsilon_0$ with Cantor normal forms.
-
-If $\alpha<\epsilon_0$, then the Cantor normal form of $\alpha$ is made of ordinals strictly less than $\alpha$.
-
-\input{movies/snippets/CNF/cnfLtEpsilon0}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Please consider the following statement :
-
-\begin{Coqsrc}
-Lemma cnf_lt_epsilon0_iff : 
- forall l alpha, 
-   is_cnf_of alpha l ->  
-   (alpha < epsilon0 <->  Forall (fun beta =>  beta < alpha) l).
-\end{Coqsrc}
-
-Is it true ?
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/schutte_cnf_counter_example.v}{exercises/ordinals/schutte\_cnf\_counter\_example.v}.}
-\end{exercise}
-
-Finally, the Cantor normal form of $\epsilon_0$ is just $\omega^{\epsilon_0}$.
-
-\input{movies/snippets/CNF/cnfOfEpsilon0}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Implement pages 82 to 85 of~\cite{schutte} (critical, strongly critical, maximal critical ordinals, Feferman's ordinal $\Gamma_0$).
-\end{project}
-
-\begin{remark}
-The sub-directory
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/Gamma0}{theories/oridinals/Gamma0} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form.
-\end{remark}
-
-\section{An embedding of \texttt{T1} into \texttt{Ord}}
-
-Our library 
-\href{../theories/html/hydras.Schutte.Correctness_E0.html}%
-{\texttt{Schutte.Correctness\_E0}} establishes the link between two very different modelizations of ordinal numbers. In other words, it ``validates'' a data structure in terms of
-a classical mathematical discourse considered as a model. 
-First, we define a function from \texttt{T1} into  \texttt{Ord} by structural recursion.
-
-\input{movies/snippets/Correctness_E0/injectDef}
-
-This function enjoys good commutation properties with respect to the main operations which
-allow us to build Cantor normal forms.
-
-\input{movies/snippets/Correctness_E0/commutationLemmas}
-\input{movies/snippets/Correctness_E0/injectPlus}
-\input{movies/snippets/Correctness_E0/injectMultFinR}
-
-% \begin{Coqsrc}
-% Theorem inject_mono (beta gamma : T1) :
-%   T1.lt  beta gamma -> 
-%   T1.nf beta -> T1.nf gamma -> 
-%   inject beta < inject gamma.
-
-% Theorem inject_injective (beta gamma : T1) : nf beta -> nf gamma ->
-%   inject beta = inject gamma -> beta = gamma.
-% \end{Coqsrc}
-
-Finally, we prove that \texttt{inject} is a bijection from the set of all terms of \texttt{T1} in normal form to the set 
-\texttt{members epsilon0} of the elements of \texttt{Ord} strictly less than  $\epsilon_0$.
-
-\input{movies/snippets/Correctness_E0/injectLtEpsilon0}
-\input{movies/snippets/Correctness_E0/embedding}
-
-\subsection{Remarks}
-Let us recall that the library \href{../theories/html/hydras.Schutte.Schutte.html}%
-{\texttt{Schutte}} depends on five \emph{axioms} and lies explicitly in the  
-framework of classical logic with a weak version of the axiom of choice
-(please look at the documentation of
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.ChoiceFacts.html}{\texttt{Coq.Logic.ChoiceFacts}}).
-Nevertheless, the other modules:
-\href{../theories/html/hydras.Epsilon0.Epsilon0.html}%
-{\texttt{Epsilon0}},
-\href{../theories/html/hydras.Hydra.Hydra.html}%
-{\texttt{Hydra}}, et 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}%
-{\texttt{Gamma0}}
-do not import any axioms and are really constructive.
-
-\index{hydras}{Projects}
-\begin{project}
-There is no construction of ordinal multiplication in~\cite{schutte}. 
-It would be interesting to derive this operation from Schütte's axioms,
-and prove its consistence with multiplication in ordinal notations for 
-$\epsilon_0$ and $\Gamma_0$.
-\end{project}
-
-\section{Related work}
-
-In~\cite{grimm:hal-00911710}, José Grimm establishes the consistency between our ordinal notations \texttt{T1} and \texttt{T2} (Veblen normal form) and his implementation
-of ordinal numbers after Bourbaki's set theory.
-
-The Gaia project ~\url{https://github.com/coq-community/gaia} maintains Grimm's  theory of ordinals as part of coq-community on GitHub. Integration
-of the present ordinal theory with Gaia, i.e., relating the different notions of ordinals
-and transferring relevant results, is an interesting project.
-First experiments in that direction are developped in
-the \href{https://github.com/coq-community/hydra-battles/blob/master/theories/gaia/}{theories/gaia/} directory.
-
-\chapter{The Ordinal \texorpdfstring{$\Gamma_0$}{Gamma0} (first draft)}
-
-\emph{This chapter and the files it presents are still very incomplete, considering the impressive properties of $\Gamma_0$~\cite{Gallier91}.
-We hope to add new material soon, and accept contributions!}
-
-\begin{todo}
-  Build a T2Bridge! (note: Gamma0's wellfoundedness is missing in \gaia (?).
-\end{todo}
-
-\section{Introduction}
-We present a notation system for the ordinal $\Gamma_0$, following Chapter V, Section 14 of~\cite{schutte}: ``A notation system for the ordinals $<\Gamma_0$''.
-We try to be as close as possible to Schütte's text and usual practices of \coq{} developments.
-
-The ordinal $\Gamma_0$ is defined in Section 13 of ~\cite{schutte} as the least \emph{strongly critical ordinal}. It is widely known as the \emph{Feferman-Schütte ordinal}.
-
-Section V, 13 of~\cite{schutte} defines \emph{strongly critical} and
-\emph{maximal $\alpha$-critical} ordinals:
-
-\begin{itemize}
-\item $\alpha$ is strongly critical if
-$\alpha$ is $\alpha$-critical,
-\item $\gamma$ is maximal $\alpha$-critical if $\gamma$ is $\alpha$-critical, and, for all $\xi>\alpha$, $\gamma$ is not $\xi$-critical.
-
-\end{itemize}
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Schutte.Critical.html\#strongly_critical}%
-{\texttt{Schutte.Critical}}}
-
-\input{movies/snippets/Critical/Gamma0Def}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that a (countable)  ordinal $\alpha$ is strongly critical iff 
-$\phi_\alpha(0)=\alpha$ (Theorem 13.13 of~\cite{schutte} ). 
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that the set of strongly critical ordinals is unbounded and closed (Theorem 13.14 of~\cite{schutte} ).
-Thus this set is not empty,  hence has a least element.
-Otherwise, the definition of $\Gamma_0$ above would be useless.
-\end{project}
-
-In the present version of this development, we  only study $\Gamma_0$ as a notation system, much more powerful than the ordinal notation for $\epsilon_0$.
-
-%\index{Projects}
-%
-%\begin{project}
-% Schûtte's section 13 of~\cite{schutte} contain several definitions and lemmas which give another view on $\Gamma_0$.  We leave it as a project to implement them in \coq{}.  
-%\end{project}
-
-\section{The type \texttt{T2} of ordinal terms}
-
-The notation system for ordinals less than $\gamma_0$ comes from the following theorem of~\cite{schutte}, where $\psi\,\alpha$ is the ordering function 
-of the set of maximal $\alpha$-critical ordinals.
-
-\index{hydras}{Library Gamma0!Types!T2}
-
-\begin{quote}
-  Any ordinal $\not= 0$ which is not strongly critical can be expressed in terms of $+$ and $\psi$.
-\end{quote}
-
-\index{hydras}{Projects}
-\begin{project}
-This theorem is not formally proved in this development yet. It should be!
-\end{project}
-
-Like in Chapter~\ref{chap:T1}, we define an inductive type with two constructors, one for $0$, the other for the construction $\psi(\alpha,\beta)\times(n+1)+\gamma$, adapting a Manolios-Vroon-like notation~\cite{Manolios2005} to
-\emph{Veblen normal forms}.
-\label{types:T2}
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T2}%
-{\texttt{Gamma0.T2}}}
-
-\inputsnippets{T2/T2Def, T2/psiDef}
-
-Like in chapter~\ref{chap:T1}, we get familiar with the type \texttt{T2} by recognizing simple constructs like finite ordinals, $\omega$, etc., as inhabitants of \texttt{T2}.
-
-\input{movies/snippets/T2/Notations}
-
-\section{A strict order on T2}
-
-Let us define a strict order on type \texttt{T2}. The following definition is 
-an adaptation of Schütte's, taking into account the multiplications by a natural number (inspired by~\cite{Manolios2005}, and also present in \texttt{T1}).
-
-\label{sect:t2-lt-def}
-
-\input{movies/snippets/T2/ltDef}
-
-Seven constructors! In order to get accustomed with this definition, let us look at a small set of examples, covering all the constructors of \texttt{lt}.
-
-\input{movies/snippets/T2/ltExamples}
-
-\index{hydras}{Projects}
-\begin{project}
-Write a tactic that solves automatically goals of the form (\texttt{$\alpha$ t2< $\beta$}), where $\alpha$ and $\beta$ are closed terms of type \texttt{T2}.
-\end{project}
-
-\section{Veblen normal form}
-\begin{definition}
-  A term of the form $\psi(\alpha_1,\beta_1)\times n_1+ \psi(\alpha_2,\beta_2)\times n_2+\dots+\psi(\alpha_k,\beta_k)\times n_k$ is said to be in
- \emph{[Veblen] normal form} if for every $i<n$, $\psi(\alpha_i,\beta_i)<\psi(\alpha_{i+1},\beta_{i+1})$, all the $\alpha_i$ and $\beta_i$ are in normal form, and all the $n_i$ are strictly positive integers.
-\end{definition}
-
-\input{movies/snippets/T2/nfDef}
-
-Let us look at some positive examples (we have to prove some inversion lemmas before proving counter-examples).
-
-\input{movies/snippets/T2/nfExamples}
-
-\subsection{Length of a term}
-
-The notion of \emph{term length} is introduced by Schütte as a helper for proving (at least) the \emph{trichotomy} property and transitivity of the strict order \texttt{lt} on \texttt{T2}.
-These properties are proved by induction on length.
-
-\subsection{Trichotomy}
-
-\emph{Trichotomy} is another name for the well-known property of decidable total ordering (like Standard Library's \texttt{Compare\_dec.lt\_eq\_lt\_dec}).
-
-We first prove by induction on $l$ the following lemma:
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0\#tricho_aux}%
-{\texttt{Gamma0.Gamma0}}}
-
-\input{movies/snippets/Gamma0/tricho}
-
-\input{movies/snippets/Gamma0/compare}
-
-With the help of \texttt{compare}, we get a boolean version of \texttt{nf}
-(being in Veblen normal form).
-
-\input{movies/snippets/Gamma0/nfb}
-
-\section{Main functions on \texttt{T2}}
-
-\subsection{Successor}
-The successor function is defined by structural recursion.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#succ}%
-{\texttt{Gamma0.T2}}}
-
-\input{movies/snippets/T2/succDef}
-
-\subsection{Addition}
-
-Like for Cantor normal forms (see Sect.~\ref{sect:infix-plus-T1}),  the definition of addition in \texttt{T2}  requires comparison between ordinal terms.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0T2.html\#succ}%
-  {\texttt{Gamma0.Gamma0}}}
-
-\input{movies/snippets/Gamma0/plusDef}
-
-\subsection{The Veblen function \texorpdfstring{$\phi$}{\texttt{phi}}}
-
-The enumeration function of critical ordinals, presented in Sect.~\vref{sect:phi-schutte}, is recursively defined in type \texttt{T2}.
-
-\input{movies/snippets/Gamma0/phiDef}
-
-Despite its complexity, the function \texttt{phi} is well adapted to proofs by simplification or computation.
-
-The relation between the constructor $\psi$ and the function $\phi$ is
-studied in~\cite{schutte}, and partially implemented in this development.
-\emph{Please contribute!}
-
-For instance, the following theorem states that, if $\gamma$ is the sum of a limit ordinal $\beta$ and a finite ordinal $n$, and $\beta$ is a fixpoint of
-$\phi(\alpha)$, then $\psi(\alpha,\gamma)=\phi_\alpha(\gamma+1)$.
-
-\input{movies/snippets/Gamma0/phiPsi}
-\input{movies/snippets/Gamma0/Ex9}
-
-On the other hand, $\phi$ can be expressed in terms of $\psi$.
-
-\input{movies/snippets/Gamma0/phiOfPsi}
-\input{movies/snippets/Gamma0/Ex10}
-
-\index{hydras}{Projects}
-\begin{project}
-Please study a way to pretty print ordinal terms in Veblen normal form (see Section~\vref{sect:ppT1}).
-\end{project}
-
-\section{An ordinal notation for \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}}
-
-In order to consider type \texttt{T2} as an ordinal notation, we have to build an instance of class \texttt{ON} (See Definition page~\pageref{types:ON}).
-
-First, we define a type that contains only terms in Veblen normal form, and redefine \texttt{lt} and \texttt{compare} by delegation (see for comparison the construction of type \texttt{E0} in Sect.~\vref{sect:E0-def}).
-
-\input{movies/snippets/Gamma0/G0a}
-\input{movies/snippets/Gamma0/G0b}
-
-Then, we buils an instance of class \texttt{ON}. 
-function \texttt{compare} is correct.
-
-\input{movies/snippets/Gamma0/ltSto}
-\input{movies/snippets/Gamma0/ltWf}
-\input{movies/snippets/Gamma0/ONGamma0}
-\input{movies/snippets/Gamma0/G0z}
-
-\begin{remark}
-The proof of \texttt{lt\_wf} has been written by \'Evelyne Contejean, using her library on the recursive path ordering (see also remark~\vref{remark:a3pat}).
-\end{remark}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that \texttt{Epsilon0} (page~\pageref{instance-epsilon0})
-is a sub-notation system of \texttt{Gamma0}.
-
-Prove that the implementations of \texttt{succ}, \texttt{+}, $\phi_0$, etc.
-are compatible in both notation systems.
-
-Note that a function \texttt{T1\_inj} from \texttt{T1} to \texttt{T2} has already been defined. It may help to complete the task.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T1_to_T2}%
-{\texttt{Gamma0.T2}}}
-
-\input{movies/snippets/T2/T1ToT2}
-
-\end{project}
-
-\begin{project}
-Prove that the notation system \texttt{Gamma0} is a correct implementation 
-of the segment $[0,\Gamma_0)$ of the set of countable ordinals.
-\end{project}

### 2022-05-08 doc/schutte-chapter.tex
reorganization of tex files

--- 2022-04-11 doc/part-hydras.tex
@@
-\section*{Introduction}
-
-In this part, we present a development  for the \coq{} proof assistant, after the work of Kirby and Paris~\cite{KP82}. This formalization contains the following main parts:
-
-\begin{itemize}
-\item Representation in \coq{} of hydras and hydra battles.
-\item A proof that every battle is finite and won by Hercules.
-This proof is based on a \emph{variant} which maps any hydra to an ordinal strictly less than $\epsilon_0$ and is strictly decreasing along any battle.
-
-\item Using a combinatorial toolkit designed by J.~Ketonen and R.~Solovay~\cite{KS81}, we prove that, for any ordinal $\mu<\epsilon_0$, there exists no such variant mapping any hydra to an ordinal strictly less than $\mu$.
-Thus, the complexity of $\epsilon_0$ is really needed in the previous proof.
-
-\item We prove a relation between the length of a ``classic''  kind of  battles \footnote{This class is also called \emph{standard} in this document (text and proofs).
-The \emph{replication factor} of the hydra is exactly $i$ at the $i$-th round of the battle (see Sect~\vref{sect:replication-def}).}
-and the Wainer-Hardy hierarchy of ``rapidly growing functions'' $H_\alpha$~\cite{Wainer1970}.
-The considered class of battles, which we call \emph{standard},  is the most considered one in the scientific  literature (including popularization).
-\end{itemize}
-
-Simply put, this document tries to combine the scientific interest of two articles~\cite{KP82, KS81} and a book~\cite{schutte} with the playful activity of truly proving theorems.
-We hope  that such a work, besides exploring a nice piece of discrete mathematics,
-will show how \coq{} and its standard library are well fitted to help us to understand some non-trivial mathematical developments, and also to experiment the constructive parts of  the proof through functional programming.
-
- We also hope to provide a little clarification on infinity (both potential and actual) through the notions of function, computation, limit,
- type and proof.
-
- \subsection*{Compatibility with \gaia (in progress)}
- \label{sect:gaia-first-intro}
-The \gaia project~\cite{Gaia} by Jos\'e Grimm, Alban Quadrat,  and Carlos Simpson,
-aims  to formalize mathematics in \coq  in
-the style of Nicolas Bourbaki.
-It contains many definitions and results about ordinal numbers.
-In Chapter~\ref{gaia-chapter}, we present some modules which allow \HydrasLib' users to apply lemmas proven in \gaia, and vice versa.
-Remarks about compatibility with \gaia are signalled with the picture \gaiasign.
-
-%\section{Remarks}
-
-\subsection*{Difference from Kirby and Paris's work}
-In~\cite{KP82}, Kirby and Paris show  that there is no proof of termination of all hydra battles in Peano Arithmetic (PA).
-Since we are used to writing proofs in higher order logic, the restriction to PA was quite unnatural for us.
-So we chose to prove another statement without any reference to PA, by considering a class of proofs indexed by ordinal numbers up to $\epsilon_0$.
-
-\subsection*{State of the development}
-The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on  notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
-We added new material: axiomatic definitions of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, function derinition by equations, etc.
-
-We are now working in order to make clumsy proofs more readable, simplify definitions, and ``factorize'' proofs as much as possible. 
-Many possible improvements are suggested as ``todo''s or ``projects'' in this text.
-
-\section*{Future work (projects)}
-\index{hydras}{Projects}
-
-This document and the proof scripts are far from being complete.
-
-First, there must be a lot of typos to correct, references and index items to add. Many proofs are too complex and should be simplified, etc.
-
-The following extensions are planned, but help is needed:
-
-\begin{itemize}
-\item Semiautomatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
-%\item Extension to $\Gamma_0$ (in Veblen normal form)
-\item More lemmas about hierarchies of rapidly growing
-  functions, and their relationship 
-    with primitive recursive functions and provability in Peano arithmetic 
-(following~\cite{KS81, KP82}).
-\item From \coq's point of view, this development could be used as an illustration of the evolution of the software, every time new libraries and sets of tactics could help to simplify the proofs.
-\end{itemize}
-
-\subsection*{Main references}
-
-In our development, we adapt the definitions and prove many theorems which
-we found in the following articles. 
-\begin{itemize}
-\item ``Accessible independence results for Peano arithmetic''  by Laurie Kirby and Jeff Paris~\cite{KP82}
-\item ''Rapidly growing Ramsey Functions'' by Jussi Ketonen and Robert Solovay~\cite{KS81}
-\item ``The Termite and the Tower'', by Will Sladek~\cite{Sladek07thetermite}
-\item Chapter V of ``Proof Theory'' by Kurt Schütte~\cite{schutte}
-\end{itemize}
-
-\chapter{Hydras and hydra games}
-
-\label{sec:orgheadline91}
-\label{chapter:hydras}
-
-This chapter is dedicated to the representation of hydras and rules of the hydra game in \coq's specification language:~\gallina.
-
-Technically, a \emph{hydra} is just a finite ordered tree, each node of which 
-has any number of sons. Contrary to the computer science tradition, we display the hydras 
-with the heads up and the foot (i.e., the root of the tree) down.
-Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the 
-module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}). 
-\emph{For a less formal description of hydras, please see 
-\url{https://www.smbc-comics.com/comic/hydra}.}
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {$\bullet$};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H4) at (4,2) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N3) to [bend left= 10]  (H1) ;
-\draw (N3) to [bend right= 16] (H2);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (foot) to [bend right = 10] (H4) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-\caption{The hydra \texttt{Hy} \label{fig:Hy}}
-\end{figure}
-
-We use a specific vocabulary for talking about hydras. Table~\ref{tab:hyd2tree} shows the correspondence between our terminology and the usual vocabulary for trees in computer science.
-
-\begin{figure}[h]
-  \centering
-  \begin{tabular}{ll}
-Hydras & Finite rooted trees\\
-\hline
-foot & root\\
-head & leaf\\
-node & node\\
-segment  & (directed) edge \\
-sub-hydra & subtree\\
-daughter & immediate subtree\\
-\end{tabular}
-  \caption{Translation from hydras to trees}
-  \label{tab:hyd2tree}
-\end{figure}
-
-The hydra \texttt{Hy} has a \emph{foot} (below), five \emph{heads}, and eight \emph{segments}. 
-We leave it to the reader to define various parameters such as the height, the size, the highest arity (number of sons of a node) of a hydra.
-In our example, these parameters have the respective values $4$, $9$ and $3$.
-
-\subsection{The rules of the game}
-
-\label{sec:orgheadline44}
-\label{sect:replication-def}
-
-A \emph{hydra battle} is a fight between Hercules and the Hydra. 
-More formally, a  battle is a sequence of \emph{rounds}.
-At each round:
-\begin{itemize}
-\item If the hydra is composed of just one head, the battle is finished
-and  Hercules is the winner.
-\item Otherwise, Hercules chops off \emph{one} head of the hydra,
-
-\begin{itemize}
-\item If the head is at distance 1 from the foot, the head is just lost by the hydra, with no more reaction.
-\item Otherwise, let us denote by \(r\) the node that was at distance \(2\) from 
-the removed head in the direction of the foot,  and consider the  sub-hydra \(h'\) of \(h\), whose  root is \(r\) \footnote{$h'$ will be called ``the wounded part of the hydra'' in the subsequent text.
-In Figures~\vref{fig:Hy2} and ~\vref{fig:Hy4}, this sub-hydra  is displayed in red.}.
-Let $n$ be some natural number.
-Then $h'$ is replaced by  $n+1$ of copies of \(h'\) which share the same root $r$.
- The \emph{replication factor} $n$ may be different (and generally is)   at each round of the fight.
-It may be chosen by the hydra, according to its strategy, or imposed by some 
-particular rule.
-In many presentations of hydra battles, this number is increased by $1$ at each round.
-In the following presentation, we will also consider battles where the hydra is free to choose its ~replication factor at each round of the battle\footnote{Let us recall that, if the chopped-off head was at distance 1 from the foot, the replication factor is meaningless.}.
-\end{itemize}
-\end{itemize}
-
-Note that the description given in~\cite{KP82} of the replication process in hydra battles is also  semi-formal.
-
-\label{original-rules}
-
-\begin{quote}
-  ``From the node that used to be attached to the head which was just chopped off, traverse one 
-segment towards the root until the next node is reached. From this node sprout $n$ replicas of 
-that part of the hydra (after decapitation) which is ``above'' the segment just traversed, i.e., those 
-nodes and segments from which, in order to reach the root, this segment would have to be 
-traversed. If the head just chopped off had the root of its nodes, no new head is grown. ''
-\end{quote}
-
-Moreover, we note that this description is in \emph{imperative} terms.
-In order to formally study the properties of hydra battles, we prefer to use a mathematical vocabulary, i.e., graphs, relations, functions, etc.
-Thus, the replication process will be represented as a binary relation on a data type \texttt{Hydra},
-linking the state of the hydra \emph{before} and \emph{after} the transformation.
-A battle will thus be represented as a sequence of terms of type \texttt{Hydra}, respecting the rules of the game.
-
-\subsection{Example}
-Let us start a battle between Hercules and the hydra \texttt{Hy} of Fig.~\ref{fig:Hy}.
-
-At the first round, Hercules chooses to chop off the rightmost head of \texttt{Hy}.
-Since this head is near the floor, the hydra simply loses this head. Let us call 
- \texttt{Hy'} the resulting state of the hydra, represented in Fig.~\vref{fig:Hy-prime}.
-
-Next, assume Hercules chooses to chop off one of the two highest heads of \texttt{Hy'}, for instance the rightmost one.
-Fig.~\vref{fig:Hy2} represents the broken segment in dashed lines, and the part that will be replicated in red.
-Assume also that the hydra decides to add 4 copies of the red part\footnote{In other words, the replication factor at this round is equal to $4$.}.
-We obtain a new state \texttt{Hy''} depicted in Fig.~\ref{fig:Hy3}.
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {$\bullet$};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H2) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-%\node (H4) at (6,0) {$\Xey[2][lightgray]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3) ;
-\draw (N3) to [bend left= 10]  (H1) ;
-\draw (N3) to [bend right= 16] (H2);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-
-\caption{\texttt{Hy'}: the state  of \texttt{Hy} after one round \label{fig:Hy-prime}}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4)  {{\color{lightred}$\bullet$}};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H1) at (0,8) {$\Sey[2][lightred]$};
-%\node (H2) at (5,0) {$\Xey[2][lightgray]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\node (ex) at (5,8) {};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw  (N2) -- (N3)[draw=lightred];
-\draw (N3) to   [bend left= 10](H1) [draw=lightred];
-\draw [dashed] (N3) to [bend left= 10](ex);
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend right= 16] (H5);
-\end{tikzpicture}
-\caption{A second beheading}
-\label{fig:Hy2}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H1) at (0,8) {$\Smiley[2][vertfluo]$};
-\node (H11) at (2,8) {$\Smiley[2][vertfluo]$};
-\node (H12) at (4,8) {$\Smiley[2][vertfluo]$};
-\node (H13) at (6,8) {$\Smiley[2][vertfluo]$};
-\node (H14) at (8,8) {$\Smiley[2][vertfluo]$};
-
-\node (N3) at (1,6) {$\bullet$};
-\node (N31) at (2,6) {$\bullet$};
-\node (N32) at (3,6) {$\bullet$};
-\node (N33) at (4,6) {$\bullet$};
-\node (N34) at (5,6) {$\bullet$};
-
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-\node (H5) at (4,4) {$\Smiley[2][vertfluo]$};
-\draw (foot) -- (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N2) -- (N31);
-\draw (N2) -- (N32);
-\draw (N2) -- (N33);
-\draw (N2) -- (N34);
-\draw (N3) to   [bend left= 10](H1) ;
-\draw (N31) to   [bend left= 10](H11) ;
-\draw (N32) to   [bend left= 10](H12) ;
-\draw (N33) to   [bend left= 10](H13) ;
-\draw (N34) to   [bend left= 10](H14) ;
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw (N1) to [bend left= 10]  (H5) ;
-\end{tikzpicture}
-\caption{\texttt{Hy''}: the state of \texttt{Hy} after two rounds \label{fig:Hy3}}
-\end{figure}
-
-Figs.~\ref{fig:Hy4} and~\vref{fig:Hy5} represent a possible third round of the battle, with a replication factor equal to $2$. Let us call \texttt{Hy'''} the state of the hydra after that third round.
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (foot) at (2,0)  {{\color{lightred}$\bullet$}};
-\node (N1) at (2,2) {{\color{lightred}$\bullet$}};
-\node (N2) at (2,4) {{\color{lightred}$\bullet$}};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (exN4) at (4,4) {};
-\node (H1) at (0,8) {$\Sey[2][lightred]$};
-\node (H11) at (2,8) {$\Sey[2][lightred]$};
-\node (H12) at (4,8) {$\Sey[2][lightred]$};
-\node (H13) at (6,8) {$\Sey[2][lightred]$};
-\node (H14) at (8,8) {$\Sey[2][lightred]$};
-
-\node (N3) at (1,6) {{\color{lightred}$\bullet$}};
-\node (N31) at (2,6) {{\color{lightred}$\bullet$}};
-\node (N32) at (3,6) {{\color{lightred}$\bullet$}};
-\node (N33) at (4,6) {{\color{lightred}$\bullet$}};
-\node (N34) at (5,6) {{\color{lightred}$\bullet$}};
-
-\node (H0) at (0,2) {$\Smiley[2][vertfluo]$};
-%\node (H5) at (4,0) {$\Xey[2][lightgray]$};
-\draw (foot) -- (N1)[very thick,draw=lightred] ;
-\draw (N1) -- (N2)[draw=lightred];
-\draw (N2) -- (N3)[draw=lightred];
-\draw (N2) -- (N31)[draw=lightred];
-\draw (N2) -- (N32)[draw=lightred];
-\draw (N2) -- (N33)[draw=lightred];
-\draw (N2) -- (N34)[draw=lightred];
-\draw (N3) to   [bend left= 10](H1) [draw=lightred];
-\draw (N31) to   [bend left= 10](H11) [draw=lightred];
-\draw (N32) to   [bend left= 10](H12) [draw=lightred];
-\draw (N33) to   [bend left= 10](H13) [draw=lightred];
-\draw (N34) to   [bend left= 10](H14) [draw=lightred];
-\draw (foot) to [bend left= 10]  (H0) ;
-\draw [dashed] (N1) to  [bend left= 10](exN4);
-\end{tikzpicture}
-\caption{A third beheading (wounded part in red) \label{fig:Hy4}}
-\end{figure}
-
-\begin{figure}[hp]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-
-\node (foot) at (10,0) {$\bullet$};
-
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\bullet$};
-\node (N3) at (2,6) {{\color{lightred}$\bullet$}};
-\node (H1) at (0,8) {$\Smiley[1][vertfluo]$};
-\node (H11) at (2,8) {$\Smiley[1][vertfluo]$};
-\node (H12) at (4,8) {$\Smiley[1][vertfluo]$};
-\node (H13) at (6,8) {$\Smiley[1][vertfluo]$};
-\node (H14) at (8,8) {$\Smiley[1][vertfluo]$};
-
-\node (N3) at (1,6) {$\bullet$};
-\node (N31) at (2,6) {$\bullet$};
-\node (N32) at (3,6) {$\bullet$};
-\node (N33) at (4,6) {$\bullet$};
-\node (N34) at (5,6) {$\bullet$};
-
-\node (H0) at (-3,3) {$\Smiley[1][vertfluo]$};
-
-\draw (foot) to [bend left=10] (N1)[very thick] ;
-\draw (N1) -- (N2);
-\draw (N2) -- (N3);
-\draw (N2) -- (N31);
-\draw (N2) -- (N32);
-\draw (N2) -- (N33);
-\draw (N2) -- (N34);
-\draw (N3) to   [bend left= 10](H1) ;
-\draw (N31) to   [bend left= 10](H11) ;
-\draw (N32) to   [bend left= 10](H12) ;
-\draw (N33) to   [bend left= 10](H13) ;
-\draw (N34) to   [bend left= 10](H14) ;
-\draw (foot) to [bend left = 15]  (H0) ;
-
-% second copy 
-\node (N01) at (12,2) {$\bullet$};
-\node (N02) at (12,4) {$\bullet$};
-\node (N03) at (12,6) {{\color{lightred}$\bullet$}};
-\node (H001) at (10,8) {$\Smiley[1][vertfluo]$};
-\node (H0011) at (12,8) {$\Smiley[1][vertfluo]$};
-\node (H0012) at (14,8) {$\Smiley[1][vertfluo]$};
-\node (H0013) at (16,8) {$\Smiley[1][vertfluo]$};
-\node (H0014) at (18,8) {$\Smiley[1][vertfluo]$};
-
-\node (N03) at (11,6) {$\bullet$};
-\node (N031) at (12,6) {$\bullet$};
-\node (N032) at (13,6) {$\bullet$};
-\node (N033) at (14,6) {$\bullet$};
-\node (N034) at (15,6) {$\bullet$};
-
-\draw (foot) -- (N01)[very thick] ;
-\draw (N01) -- (N02);
-\draw (N02) -- (N03);
-\draw (N02) -- (N031);
-\draw (N02) -- (N032);
-\draw (N02) -- (N033);
-\draw (N02) -- (N034);
-\draw (N03) to   [bend left= 10](H001) ;
-\draw (N031) to   [bend left= 10](H0011) ;
-\draw (N032) to   [bend left= 10](H0012) ;
-\draw (N033) to   [bend left= 10](H0013) ;
-\draw (N034) to   [bend left= 10](H0014) ;
-
-% third copy 
-\node (N001) at (22,2) {$\bullet$};
-\node (N002) at (22,4) {$\bullet$};
-\node (N003) at (22,6) {{\color{lightred}$\bullet$}};
-\node (H001) at (20,8) {$\Smiley[1][vertfluo]$};
-\node (H0011) at (22,8) {$\Smiley[1][vertfluo]$};
-\node (H0012) at (24,8) {$\Smiley[1][vertfluo]$};
-\node (H0013) at (26,8) {$\Smiley[1][vertfluo]$};
-\node (H0014) at (28,8) {$\Smiley[1][vertfluo]$};
-
-\node (N003) at (21,6) {$\bullet$};
-\node (N0031) at (22,6) {$\bullet$};
-\node (N0032) at (23,6) {$\bullet$};
-\node (N0033) at (24,6) {$\bullet$};
-\node (N0034) at (25,6) {$\bullet$};
-
-\draw (foot) -- (N001)[very thick] ;
-\draw (N001) -- (N002);
-\draw (N002) -- (N003);
-\draw (N002) -- (N0031);
-\draw (N002) -- (N0032);
-\draw (N002) -- (N0033);
-\draw (N002) -- (N0034);
-\draw (N003) to   [bend left= 10](H001) ;
-\draw (N0031) to   [bend left= 10](H0011) ;
-\draw (N0032) to   [bend left= 10](H0012) ;
-\draw (N0033) to   [bend left= 10](H0013) ;
-\draw (N0034) to   [bend left= 10](H0014) ;
-\end{tikzpicture}
-\caption{The configuration \texttt{Hy'''} of \texttt{Hy} \label{fig:Hy5}}
-\end{figure}
-\FloatBarrier
-
-We leave it to the reader  to guess the following  rounds of the battle \dots
- % Please keep in mind that, in this 
-% the hydra is free to chose any number of replications at each time, whereas
-% Hercules chops only one head per round.
-
-% Let us precise that, in this game, Hercules wins if the hydra is eventually reduced 
-% to a single head. 
-% We know from~\cite{KP82} that, whichever the initial configuration of the
-% hydra, and the strategies of both players, Hercules eventually wins. The 
-% aforementionned paper shows also that there do not exist any \emph{simple} proof of this result.
-
-\section{Hydras and their representation in \emph{Coq}}
-\label{sec:orgheadline48}
-
-\index{hydras}{Library Hydra!Types!Hydra}
-\index{hydras}{Library Hydra!Types!Hydrae}
-
-In order to describe trees where each node can have an arbitrary (but finite) number of sons, it is usual to define a type where each node carries a \emph{forest}, \emph{i.e} a list of trees
-(see for instance Chapter 14, pages 400-406 of \cite{BC04}, also available as~\cite{BC04ch14}).
-
-For this purpose, we define two mutual \emph{ad-hoc}  inductive types, where \texttt{Hydra} is the main type, and \texttt{Hydrae} is a helper for describing finite sequences of hydras.
-\label{types:Hydra}
-\label{types:Hydrae}
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hydra}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/HydraDef}
-
-%\index{To do}
-\index{hydras}{Projects}
-
-\begin{project}
-Look for an existing library on trees with nodes of arbitrary arity, in order to replace  this ad-hoc type with something more generic.
-\end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-
- Another very similar representation could use the \texttt{list} type family instead of the specific 
-type \texttt{Hydrae}:
-
-\input{movies/snippets/Hydra_Definitions/HydraAlt}
-
-Using this representation, re-define all the constructions of this chapter.
-You will probably have to use patterns described for instance in~\cite{BC04} or the archives of the \coq communication channels (please consult~\url{https://coq.inria.fr/community.html}).
-
-\end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-The type \texttt{Hydra} above describes hydras as \emph{plane trees}, i.e., as drawn on a sheet of paper or computer screen. Thus, hydras are \emph{oriented},
-and it is appropriate to consider a \emph{leftmost} or \emph{rightmost} head of
-the beast. It could be interesting to consider another representation, in which
-every non-leaf node has a \emph{multi-set} -- not an ordered list -- of daughters.
-\end{project}
-
-\subsubsection{Abbreviations}
-
-We provide several notations for hydra patterns  which occur often in our developments.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#head}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/headsEtc}
-
-For instance, the hydra \texttt{Hy}  of Figure~\vref{fig:Hy} is defined in \emph{Gallina} as follows:
-
-\vspace{4mm}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html\#Hy}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/Hy}
-
-Hydras quite frequently contain  multiple adjacent  copies of the same subtree. The following functions
-will help us to describe and reason about replications in hydra battles.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#hcons_mult}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hconsMult}
-
-\vspace{4mm}
-
-For instance, the hydra \texttt{Hy''} of Fig~\vref{fig:Hy3}  can be defined in \coq{} as follows:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/HySecond}
-
-\subsubsection{Recursive functions on type \texttt{Hydra}}
-\label{sec:orgheadline41}
-\label{sec:hsize-def}
-
-In order to  define a recursive function over the type \texttt{Hydra}, one has to consider the three constructors 
-\texttt{node}, \texttt{hnil} and \texttt{hcons} of the mutually inductive types \texttt{Hydra} and \texttt{Hydrae}. 
-Let us define for instance the function which  computes the number of nodes of any hydra:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hsize}
-
-\input{movies/snippets/Hydra_Examples/HySize}
-
-Likewise, the \emph{height} (maximum distance between the foot and a head) 
-is defined by mutual recursion:
-
-\input{movies/snippets/Hydra_Definitions/height}
-
-\input{movies/snippets/Hydra_Examples/HyHeight}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Define a function \texttt{max\_degree: Hydra $\arrow$ nat} which  returns the highest degree of a node in any hydra.
-For instance, the evaluation of the term \texttt{(max\_degree Hy)} should return $3$.
-\end{exercise}
-
-\subsection{Induction principles for hydras}
-\label{sec:orgheadline42}
-
-In this section, we show how induction principles are used to prove properties on the type 
-\texttt{Hydra}. Let us consider for instance the following statement:
-\begin{quote}
-  `` The height of any hydra is strictly less than its size. ''
-\end{quote}
-
-\subsubsection{A failed attempt}
-
-One may try to use the default tactic of proof by induction, which corresponds to an application of the automatically  generated  induction principle for  type \texttt{Hydra}:
-
-\input{movies/snippets/Hydra_Examples/HydraInd}
-
-Let us start a simple proof by induction.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/BadInductiona}
-
-We might be tempted to do an induction on the sequence \texttt{s}:
-
-\input{movies/snippets/Hydra_Examples/BadInductionb}
-
-The first subgoal is trivial.
-
-\input{movies/snippets/Hydra_Examples/BadInductionc}
-
-Let us look at the second subgoal of the induction.
-
-\input{movies/snippets/Hydra_Examples/BadInductiond}
-
-We notice that this subgoal does not contain any hypothesis
-on the height and size of the hydra \texttt{h}. So, it looks hard to prove the conclusion. Let's stop.
-
-\input{movies/snippets/Hydra_Examples/BadInductione}
-
-\subsubsection{A Principle of mutual induction}
-In order to get an appropriate induction scheme for the types 
-\texttt{Hydra} and \texttt{Hydrae}, we can use  \coq{}'s  command \texttt{Scheme}.
-
-\index{coq}{Mutually inductive types}
-%\index{coq}{Commands!Scheme}
-
-\input{movies/snippets/Hydra_Definitions/HydraRect2}
-
-\input{movies/snippets/Hydra_Examples/HydraRect2Check}
-
-\subsubsection{A Correct proof}
-
-Let us now use \texttt{Hydra\_rect2} for proving that the height of any hydra is strictly less than its size.
-Using this scheme requires an auxiliary predicate, called \texttt{P0} in \texttt{Hydra\_rect2}'s statement.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/hForall}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-\input{movies/snippets/Hydra_Examples/heightLtSizea}
-
-The first subgoal is easily solvable, using some arithmetic.
-The second and third ones are
-almost trivial. We let the reader look at the source. 
-\input{movies/snippets/Hydra_Examples/heightLtSizez}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-It happens very often that, in the proof of  a proposition of the form 
-(\texttt{$\forall\,$ h:Hydra, $P$ h}), the predicate \texttt{P0}
-is  (\texttt{h\_forall $P$}).
-Design a tactic for induction on hydras that frees the user from binding explicitly \texttt{P0},  and solves trivial subgoals.
-Apply it for writing  a shorter proof of \texttt{height\_lt\_size}.
-\end{exercise}
-
-\section{Relational description of hydra battles}
-
-In this section, we represent the rules of hydra battles as a binary relation associated with
-a \emph{round}, i.e., an interaction composed of the two following actions:
-\begin{enumerate}
-\item Hercules chops off one head of the hydra.
-\item Then, the  hydra replicates the wounded part (if the head is at distance $\geq 2$ from the foot).
-\end{enumerate}
-The relation associated with each round of the battle is parameterized  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot,
-but present for consistency's sake).
-
-In our description,  we will apply the following naming convention: if $h$ represents the configuration of the hydra before a round, then the configuration of $h$ after this round will be called $h'$.
- Thus, we are going to define a proposition  (\texttt{round\_n $n\;h\;h'$})  whose intended meaning will be `` the hydra $h$  is transformed into $h'$  in a single round of a battle, with the expected replication factor $n$ ''.
-
-Since the replication of parts of the hydra depends on the distance of the chopped head from  the foot, we  decompose our description into two main  cases, under the form of a bunch of [mutually] inductive predicates over the types \texttt{Hydra} and \texttt{Hydrae}.
-
-The mutually exclusive cases we consider are the following:
-\begin{itemize}
-\item \textbf{R1}: The chopped off head was at distance 1 from the foot.
-\item \textbf{R2}: The chopped off head was at a distance greater than or equal to  $2$ from the foot.
-\end{itemize}
-
-\subsection{Chopping off a head at distance 1 from the foot (relation  R1)}
-
-If Hercules chops off a head near the floor, there is no replication at all. We use an auxiliary 
-predicate \texttt{S0}, associated with the removing of one head from a sequence of hydras.
-
-\vspace{4pt}\emph{From Module\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/S0Def}
-\input{movies/snippets/Hydra_Definitions/R1Def}
-
-\subsubsection{Example}
-\label{sec:orgheadline45}
-
-Let us represent in \coq{}   the transformation of the hydra of Fig.~\vref{fig:Hy} into
-the configuration represented in Fig.~\ref{fig:Hy-prime}.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}
-
-\input{movies/snippets/Hydra_Examples/Hy1}
-
-\subsection{Chopping off a head at distance \texorpdfstring{$\geq 2$}{>= 2} from the foot (relation R2) }
-
-Let us now consider beheadings  where the chopped-off head is at distance greater than or equal to $2$ from the foot. All the following relations are parameterized by the replication factor  $n$.
-
- Let $s$ be a sequence of hydras. 
-The proposition (\texttt{S1 n s s'}) holds if $s'$ is obtained by replacing some element $h$ of $s$ by 
-$n+1$ copies of $h'$, where  the proposition (\texttt{R1 h h'}) holds, in other words, $h'$ is just $h$, without the chopped-off  head.
-\texttt{S1} is an inductive relation with two constructors that allow us to choose the position in $s'$ of the wounded sub-hydra $h$.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#S1}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/S1Def}
-
-The rest of the definition is composed of two mutually inductive relations on hydras and sequences of hydras.
-The first constructor of \texttt{R2} describes the case where the chopped head is exactly at height $2$.
-The others constructors allow us to consider beheadings at height strictly greater than $2$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/R2Def}
-
-\subsubsection{Example}
-Let us prove the transformation of \texttt{Hy'} into \texttt{Hy''} (see Fig.~\vref{fig:Hy3}).
-We use an experimental set of tactics\footnote{See the \texttt{Ltac} definitions in~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions.}}  for specifying the place where the
-interaction between Hercules and the hydra holds.
-
-\vspace{4pt}\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}.
-
-\input{movies/snippets/Hydra_Examples/R2Example}
-
-The reader is encouraged to look at all the successive subgoals of this example.
-\emph{Please consider also exercise~\vref{exo:interactive-battle}.}
-
-\subsection{Binary relation associated with a round}
-
-We combine the two cases above into a single relation.
-First,  we define the  relation \texttt{(round\_n n h h')} where \texttt{n} is the expected number of  replications (irrelevant in the case of an \texttt{R1}-transformation).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#round_n}{Hydra.Hydra\_Definitions}}
-
-\index{hydras}{Library Hydra!Predicates!round\_n}
-
-\input{movies/snippets/Hydra_Definitions/roundNDef}
-
-By abstraction over \texttt{n}, we define a \emph{round} (small step) of a battle:
-
-\index{hydras}{Library Hydra!Predicates!round}
-\label{sect:infix-round}
-
-\input{movies/snippets/Hydra_Definitions/roundDef}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Give a direct translation of Kirby and Paris's description of hydra battles (quoted on page~\pageref{original-rules}) and prove that our relational description is consistent with theirs.
-\end{project}
-
-\subsection{Rounds and battles}
-
-Using library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}, we define \texttt{round\_plus},  the transitive closure of \texttt{round}, and \texttt{round\_star},  the reflexive and transitive closure of \texttt{round}.
-
-\label{sect:infix-rounds}
-
-\input{movies/snippets/Hydra_Definitions/roundPlus}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-  Prove that if \texttt{$h$ -+-> $h'$}, then
-  the height of $h'$ is less or equal than the height of $h$.
-
-\end{exercise}
-
-\begin{remark}
-\label{remark:transitive-closure}
-\coq's library \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Coq.Relations.Relation\_Operators} 
-contains three logically equivalent definitions of the transitive closure of a binary relation. This equivalence is proved in 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} .
-
-Why three definitions for a single mathematical concept?
-Each definition generates an associated induction principle. 
- According to the form of statement one would like to prove, there is a ``best choice'':
-
-\begin{itemize}
-\item For proving $\forall y, x\,R^+\,y \;\arrow\; P\,y$, prefer 
-\texttt{clos\_trans\_n1}
-\item For proving $\forall x,\,x\,R^+\,y \;\arrow\; P\,x$, prefer \texttt{clos\_trans\_1n}
-\item For proving $\forall x\,y, \,x\,R^+\,y \;\arrow\;P\,x\,y$,  
-prefer \texttt{clos\_trans},
-\end{itemize}
-But there is no ``wrong choice'' at all: the equivalence lemmas in \linebreak 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Operators_Properties.html}{Coq.Relations.Operators\_Properties} 
- allow the user
-to convert any one of the three closures into another one before applying the corresponding elimination tactic.
-The same remark also holds for reflexive and transitive closures. 
-\end{remark}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Define a restriction of \coqsimple{round},  where Hercules always chops off
-the leftmost among the lowest heads.
-
-Prove that, if $h$ is not a simple head, then there exists a unique $h'$ such that $h$  is transformed into $h'$ in one round, according to this restriction.
-
-\end{exercise}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}[Interactive battles]
-\label{exo:interactive-battle}
-Given a hydra \texttt{h}, the specification of a hydra battle for \texttt{h} is the type 
-\Verb@{h':Hydra | h -*-> h'}@. In order to avoid long sequences of \texttt{split}, \texttt{left}, and 
-\texttt{right}, design a set of dedicated tactics for the interactive building of a battle.
-Your tactics will have the following functionalities:
-\begin{itemize}
-\item  Chose to stop a battle, or continue
-\item Chose an expected number of replications
-\item Navigate in a hydra, looking for a head to chop off.
-\end{itemize}
-
-Use your tactics for simulating a small part of a hydra battle, for instance the rounds which lead from
-\texttt{Hy} to \texttt{Hy'''}  (Fig.~\vref{fig:Hy5}).
-
-\textbf{Hints:} 
-\begin{itemize}
-
-\item Please keep in mind that the last  configuration of your interactively built battle is known only at the end of the battle.
-Thus, you will have to create and solve subgoals with existential variables.
-For that purpose, the tactic \texttt{eexists}, applied to the
-goal \Verb@{h':Hydra | h -*-> h'}@ generates the subgoal \Verb|h -*-> ?h'|.
-\item You may use Gérard Huet's \emph{zipper} data structure~\cite{zipper} for writing tactics associated with Hercules's  interactive search for a head to chop off.
-\end{itemize}
-
-\end{exercise}
-
-\subsection{Classes of battles}
-\label{sect:battle-classes}
-
-In some presentations of hydra battles, e.g.~\cite{KP82, bauer2008}, the transformation associated with the $i$-th round may depend on $i$.
-For instance, in these articles, the replication factor at the $i$-th round is equal to $i$.
-In other examples, one can allow the hydra to apply any replication factor at any time.
-In order to be the most general as possible, we define the type of predicates which relate the state of the hydra before and after the $i$-th round of a battle.
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
-\label{types:Battle}
-\index{hydras}{Library Hydra!Type classes!Battle}
-
-\input{movies/snippets/Hydra_Definitions/BattleDef}
-
-The most general class of battles is \texttt{free}, which allows the hydra to chose any replication factor at every step:
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#free}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/freeDef}
-
-We chose to call \emph{standard}\footnote{This appellation is ours.
-If there is a better one, we will change it.} the kind of battles which appear  most often in the literature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$
-
-\vspace{6pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#standard}{Hydra.Hydra\_Definitions}}
-\input{movies/snippets/Hydra_Definitions/standardDef}
-
-\subsection{Big steps}
-
-Let $B$ be any instance of class \texttt{Battle}. It is easy to define inductively the relation between the $i$-th and the $j$-th steps of a battle of type $B$.
-
-\vspace{6pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/battleRelDef}
-
-The following property allows us to build battles by composition of smaller ones.
-
-%% TODO display subgoals when fixed
-
-\vspace{6pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-\noindent
-\input{movies/snippets/Hydra_Lemmas/battleTrans}
-
-% \begin{remark}
-%  The class \texttt{free} is strongly related with the transitive closure  \texttt{round\_plus}, as expressed by the following lemmas.
-
-% \vspace{4pt}
-% \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-
-%  \begin{Coqsrc}
-%  Lemma battle_free_equiv1 : forall i j h h',  
-%              battle free i h j h' ->   h -+-> h'.
-
-%  Lemma battle_free_equiv2 : forall h h',
-%      h -+-> h' ->
-%     forall i, exists j,  battle free i h j h'.
-%  \end{Coqsrc}
-
-% \end{remark}
-
-\section{A long battle}
-\label{sect:big-battle}
-
-In this section we consider a simple example of battle, starting with a small hydra,
-shown on figure~\vref{fig:hinit}, with a simple strategy for both players:
-
-\begin{itemize}
-\item At each round, Hercules chops off the rightmost head of the hydra.
-\item The battle is {standard}: at the round number $i$, the expected replication is $i$.
-\end{itemize}
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (6,0) {$\bullet$};
-\node (n1) at  (3,3) {$\bullet$};
-\node (h1) at  (1,6) {$\Smiley[1][green]$};
-\node (h2) at  (3,6) {$\Smiley[1][green]$};
-\node (h3) at  (6,6) {$\Smiley[1][green]$};
-\node (h4) at  (6,6) {$\Smiley[1][green]$};
-\node (h5) at  (6,3) {$\Smiley[1][green]$};
-\node (h6) at  (9,3) {$\Smiley[1][green]$};
-\draw (foot) -- (n1);
-\draw (n1) to   [bend left=20] (h1);
-\draw (n1) to   (h2);
-\draw (n1) to   [bend right=20] (h3);
-\draw (foot) -- (h5);
-\draw (foot) to  [bend right=20] (h6);
-\end{tikzpicture}
-
-  \caption{The hydra \texttt{hinit}}
-  \label{fig:hinit}
-\end{figure}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.BigBattle.html}{Hydra.BigBattle}}
-
-\input{movies/snippets/BigBattle/hinitDef}
-
-The lemma we would like to prove is ``The considered battle lasts exactly $N$ rounds'',
-with $N$ being a natural number we gave to guess.
-
-But the  battle is so long that no \emph{test} can give us any estimation of its length, and we do need the expressive power of logic to compute this length.
-However, in order to  guess this length, we made some experiments, computing with \gallina{}, \coq{}'s  functional programming language.
-Thus, we can consider this development as a collaboration of proof with computation.
-In the following lines, we show how we found experimentally the value of the number $N$.
-The complete proof is in file \url{../theories/html/hydras.Hydra.BigBattle.html}.
-
-\subsection{First rounds}
-During the two first rounds, our hydra loses its two rightmost heads.  Figure~\vref{fig:hinit-plus2} shows the state of the hydra   just before the third round.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (3,0) {$\bullet$};
-\node (n1) at  (3,3) {$\bullet$};
-\node (h1) at  (1,6) {$\Smiley[1][green]$};
-\node (h2) at  (3,6) {$\Smiley[1][green]$};
-\node (h3) at  (6,6) {$\Smiley[1][green]$};
-\node (h4) at  (6,6) {$\Smiley[1][green]$};
-\draw (foot) -- (n1);
-\draw (n1) to   [bend left=20] (h1);
-\draw (n1) to   (h2);
-\draw (n1) to   [bend right=20] (h3);
-\end{tikzpicture}
-
-  \caption{The hydra (\texttt{hyd1 h3})}
-  \label{fig:hinit-plus2}
-\end{figure}
-
-The following lemma  is a formal description of these first rounds, in terms of the
-\texttt{battle} predicate.
-
-\input{movies/snippets/BigBattle/L02}
-
-\subsection{Looking for regularities}
-
-A first study with pencil and paper suggested us that, after three rounds, the hydra always looks like in figure~\vref{fig:hinit-plusn} (with a variable number of 
-subtrees of height 1 or 0).
-Thus, we introduce a few handy abbreviations.
-
-\input{movies/snippets/BigBattle/Notations}
-
-For instance, the hydra shown in  Fig~\ref{fig:hinit-plusn} is  (\texttt{hyd 3 4 2}). The hydra (\texttt{hyd 0 0 0})  is the ``final'' hydra of any terminating battle, {i.e.},
-a tree whith exactly one node and no edge.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[thick, scale=0.30]
- \node (foot) at (15,0) {$\bullet$};
-\node (a) at  (3,4) {$\bullet$};
-\node (b) at  (6,4) {$\bullet$};
-\node (c) at  (9,4) {$\bullet$};
-\node (d) at  (13,4) {$\bullet$};
-\node (e) at  (16,4) {$\bullet$};
-\node (f) at  (19,4) {$\bullet$};
-\node (g) at  (22,4) {$\bullet$};
-\node (h) at  (25,4) {$\Smiley[1][green]$};
-\node (i) at  (28,4) {$\Smiley[1][green]$};
-\node (aa) at  (2.5,8) {$\Smiley[1][green]$};
-\node (ab) at  (3.5,8) {$\Smiley[1][green]$};
-\node (ba) at  (5.5,8) {$\Smiley[1][green]$};
-\node (bb) at  (6.5,8) {$\Smiley[1][green]$};
-\node (ca) at  (8.5,8) {$\Smiley[1][green]$};
-\node (cb) at  (9.5,8) {$\Smiley[1][green]$};
-\node (da) at  (13,8) {$\Smiley[1][green]$};
-\node (ea) at  (16,8) {$\Smiley[1][green]$};
-\node (fa) at  (19,8) {$\Smiley[1][green]$};
-\node (ga) at  (22,8) {$\Smiley[1][green]$};
-\draw (foot) -- (a);
-\draw (foot) -- (b);
-\draw (foot) -- (c);
-\draw (foot) -- (d);
-\draw (foot) -- (e);
-\draw (foot) -- (f);
-\draw (foot) -- (g);
-\draw (foot) -- (h);
-\draw (foot) -- (i);
-\draw (a) -- (aa);
-\draw (a) -- (ab);
-\draw (b) -- (ba);
-\draw (b) -- (bb);
-\draw (c) -- (ca);
-\draw (c) -- (cb);
-\draw (d) -- (da);
-\draw (e) -- (ea);
-\draw (f) -- (fa);
-\draw (g) -- (ga);
-\end{tikzpicture}
-
-  \caption{The hydra (\texttt{hyd 3 4 2})}
-  \label{fig:hinit-plusn}
-\end{figure}
-
-With these notations, we get a formal description of the first three rounds.
-
-\input{movies/snippets/BigBattle/L23L03}
-
-\subsection{Testing  \dots}
-\label{sect:testing}
-In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple data type for representing the states as tuples composed of
-the round number, and the respective number of daughters  \texttt{h2}, \texttt{h1}, and heads
-of the current hydra.
-
-\input{movies/snippets/BigBattle/stateDef}
-
-The following function returns the next configuration of the game.
-Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle only start with the lemma
-\texttt{step\_battle}, page~\pageref{lemma:step-battle}.
-
-\input{movies/snippets/BigBattle/nextDef}
-
-We can make bigger steps through iterations of \texttt{next}.
-The functional \texttt{iterate}, similar to Standard Library's \texttt{Nat.iter},
-is defined and studied in~\href{../theories/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.
-\index{hydras}{Library Prelude!iterate}
-
-\label{Functions:iterate}
-
-\input{movies/snippets/Iterates/iterateDef}
-
-The following function computes the state of the battle at the $n$-th round.
-
-\input{movies/snippets/BigBattle/testDefTests}
-
-The battle we are studying looks to be awfully long. Let us concentrate our
-tests on some particular events : the states where $\texttt{nh}=0$.
-From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} is equal to zero.
-
-\input{movies/snippets/BigBattle/smartTest}
-
-Thus, $(1 + 11)$ rounds later, the \texttt{n1} field is equal to $2$, and 
-\texttt{nh}   to $0$.
-
-\input{movies/snippets/BigBattle/smartTestb}
-
-Next round, we decrement \texttt{n2} and set \texttt{n1} to $95$.
-
-\input{movies/snippets/BigBattle/smartTestc}
-
-We now have some intuition of the sequence.
-It looks like the next ``\texttt{nh}=0'' event will happen at the $192=2(95+1)$-th round, then at the $2(192+1)$-th round, etc.
-
-\input{movies/snippets/BigBattle/doubleS}
-
-\subsection{Proving \dots}
-We are now able to reason about the sequence of transitions defined by our hydra battle. Instead of using the data-type \texttt{state} we study the relationship
-between different configurations of the battle.
-
-Let us define a binary relation associated with every round of the battle.
-In the following definition \texttt{i} is associated with the round number (or date, if we consider a discrete time), and \texttt{a}, \texttt{b}, \texttt{c} respectively associated with the number of occurrences of \texttt{h2}, \texttt{h1} and heads connected to the hydra's foot.
-
-\input{movies/snippets/BigBattle/oneStep}
-
-The relation between \texttt{one\_step} and the rules of hydra battles is asserted by the following lemma.
-
-\label{lemma:step-battle}
-
-\input{movies/snippets/BigBattle/stepBattle}
-
-\vspace{4pt}
-
-Next, we define ``big steps'' as the transitive closure of \texttt{one\_step},
-and reachability (from the initial configuration of figure~\ref{fig:hinit} at time $0$).
-
-\input{movies/snippets/BigBattle/steps}
-
-The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{battle}.
-
-\input{movies/snippets/BigBattle/stepsBattle}
-
-\vspace{4pt}
-
-Thus, any result about \texttt{steps} will be applicable to standard battles.
-Using the predicate \texttt{steps},  our study of the length of the considered battle
-can  be decomposed into three parts:
-
-\begin{enumerate}
-\item  Characterization and proofs of regularities of some events (inspired by our experiments of Sect.~\ref{sect:testing}).
-\item Study of the beginning of the battle
-\item Computing the exact length of the battle.
-\end{enumerate}
-
-First, we prove that, if at round $i$ the hydra is equal to
-(\texttt{hyd a (S b) 0}), then it will be equal to (\texttt{hyd a b 0}) at the $2(i+1)$-th round.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/LS}
-
-From now on, the lemma \texttt{reachable\_S} allows us to watch larger and larger steps of 
-the battle.
-
-\input{movies/snippets/BigBattle/L4}
-
-\input{movies/snippets/BigBattle/L10To95}
-
-\subsection{Giant steps}
-
-We are now able to make bigger steps in the simulation of the battle.
-First, we iterate the lemma \texttt{reachable\_S}.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/Bigstep}
-
-\vspace{4pt}
-
-Applying lemmas \texttt{BigStep} and \texttt{L95} we make a first jump.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/MDef}
-
-Figure~\ref{fig:HM}  represents the hydra at the $M$-th round.
-At the $(M+1)$-th round, it will look like in fig~\ref{fig:HM-plus1}.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:HM}}
-The state of the hydra after $M$ rounds.
-% The hydra \texttt{h} of the proof that \(\omega^2\) is too small for proving Hercules' victory
-
-\end{figure}
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (0,5) {$\bullet$};
-\node (N12) at (0,8) {$\Smiley[2][green]$};
-\node (N2) at (2,5) {$\bullet$};
-\node (N22) at (2,8) {$\Smiley[2][green]$};
-\node (N3) at (4,5) {$\bullet$};
-\node (N32) at (4,8) {$\Smiley[2][green]$};
-\node (N4) at (6,5) {$\bullet$};
-\node (N42) at (6,8) {$\Smiley[2][green]$};
-
-\node (Ndots) at (12,8) {\Huge $\dots$};
-\node (Ndots2) at (12,5) {\Huge $\dots$};
-
-\node (N8) at (18,5) {$\bullet$};
-\node (N82) at (18,8) {$\Smiley[2][green]$};
-\node (N9) at (20,5) {$\bullet$};
-\node (N92) at (20,8) {$\Smiley[2][green]$};
-
-\draw (foot) -- (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N8);
-\draw (foot) -- (N9);
-\draw (N1) to  (N12);
-\draw (N2) to  (N22);
-\draw (N3) to  (N32);
-\draw (N4) to  (N42);
-\draw (N8) to  (N82);
-\draw (N9) to  (N92);
-\end{tikzpicture}
-\caption{\label{fig:HM-plus1}}
-The state of the hydra after $M+1$ rounds (with $M+1$ heads).
-
-\end{figure}
-
-\input{movies/snippets/BigBattle/L295S}
-
-\vspace{4pt}
-
-Then, applying once more the lemma \texttt{BigStep}, we get the exact time when
-Hercules wins!
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/NDef}
-
-\vspace{4pt}
-
-We are now able to prove formally that the considered battle is 
-composed of $N$ steps.
-
-\vspace{4pt}
-
-\input{movies/snippets/BigBattle/Done}
-
-\subsection{A minoration lemma}
-
-Now, we would like to get an intuition of  how big the number $N$ is.
-For that purpose, we use a minoration of the function \texttt{doubleS} by the
-function (\texttt{fun n => 2 * n}).
-
-\vspace{4pt}
-
-\input{movies/snippets/Exp2/exp2Def}
-
-\vspace{4pt}
-
-Using a few facts (proven in 
-\href{../theories/html/hydras.Hydra.BigBattle.html}{hydras.Hydra.BigBattle}),we get several  minorations.
-
-\input{movies/snippets/BigBattle/minorationLemmas}
-
-\vspace{4pt}
-
-The number $N$ is greater than or  equal to $2^{2^{95}\times 95}.$ If we wrote $N$ in base $10$, $N$ would require at least $10^{30}$ digits!
-
-\section{Generic properties}
-
-The example we just studied shows that the termination of any battle may take a very long time. If we want to study hydra battles in general, we have to consider 
-any hydra and any strategy, both for Hercules and the hydra itself. So, we first  give some definitions, generally borrowed from transition systems vocabulary (see~\cite{tel_2000} for instance).
-
-\subsection{Liveliness}
-
-Let $B$ be an instance of \texttt{Battle}. We say that $B$ is \emph{alive} if
-for any configuration $(i,h)$, where $h$ is not a head, there exists a further step in class $B$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}
-
-\input{movies/snippets/Hydra_Definitions/AliveDef}
-
-The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
-\href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems} show that the classes \texttt{free} and \texttt{standard} satisfy this property.
-
-\input{movies/snippets/Hydra_Theorems/AliveThms}
-
-Both theorems are proved with the help of the  following strongly specified function:
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
-
-\input{movies/snippets/Hydra_Lemmas/nextRoundDec}
-
-\subsection{Termination}
-
-The termination of \emph{any}  battle is naturally expressed by the predicate \texttt{well\_founded} defined in the module \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}
- of the Standard Library.
-
-\index{hydras}{Library Hydra!Predicates!Termination}
-
-\input{movies/snippets/Hydra_Definitions/TerminationDef}
-
-Let $B$ be an instance of class \texttt{Battle}. A \emph{variant} for $B$ consists
-in a well-founded relation $<$  on some type \texttt{A}, and a function
-(also called a \emph{measure}) $m:\texttt{Hydra}\rightarrow A$ such that for any successive steps $(i,h)$ and $(1+i,h')$  of a battle in $B$, the inequality $m(h')<m(h)$ holds.
-
-\vspace{4pt}
-\noindent
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}
-
-\label{sect:hvariant-def}
-
-\index{hydras}{Library Hydra!Type classes!Hvariant}
-
-\input{movies/snippets/Hydra_Definitions/HvariantDef}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
- Prove that, if there exists some  instance of (\texttt{Hvariant Lt wf\_Lt $B$ $m$}), then there exists no infinite battle in  $B$.
-\end{exercise}
-
-\subsection{A  small proof of impossibility}
-%\index{coq}{Proofs of impossibility}
-
-\label{omega-case}
-
-When one wants to prove a termination theorem with the help of a variant, 
-one has to consider first a well-founded set $(A,<)$, then a strictly decreasing measure on this set.
-The following two lemmas show that if  the order structure $(A,<)$ is too simple, it is useless to look for a convenient measure, which simply no exists.
-Such kind of result is useful, because it saves you time and effort.
-
-The best known well-founded order is the natural order on the set $\mathbb{N}$ of natural numbers (the type \texttt{nat} of Standard library).
-It would be interesting to look for some measure $m:\texttt{nat}\arrow\texttt{nat}$ and prove it is a variant.
-
-Unfortunately, we can prove that 
-\emph{no} instance of class (\texttt{WfVariant round Peano.lt $m$}) can be built, where
-$m$ is \emph{any} function of type \texttt{Hydra $\arrow$ nat}.
-
-Let us present the main steps of that proof, the script of which  is in the module ~\href{../theories/html/hydras.Hydra.Omega_Small.html}{Hydra/Omega\_Small.v} \footnote{ The name of this file means ``the ordinal $\omega$ is too small for proving the termination of [free] hydra battles ''.
-In effect, the elements of $\omega$, considered as a set, are just the natural numbers (see next chapter for more details)}.
-
-%\subsubsection{Preliminaries}
-
-Let us assume there exists some variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
-    the  termination of all hydra battles.
-
-\input{movies/snippets/Omega_Small/omegaSmalla}
-
-We define an injection $\iota$ from the type \texttt{nat} into \texttt{Hydra}.
-For any natural number $i$, $\iota(i)$ is the hydra composed of a foot and
-$i+1$ heads at height $1$. For instance, Fig.~\ref{fig:flower} represents the hydra $\iota(3)$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (4,0) {$\bullet$};
-\node (N1) at (2,2) {$\Smiley[2][green]$};
-\node (N2) at (4,2) {$\Smiley[2][green]$};
-\node (N3) at (6,2) {$\Smiley[2][green]$};
-\node (N4) at (8,2) {$\Smiley[2][green]$};
-\draw (foot) to [bend left =25] (N1);
-\draw (foot) to [bend left =15] (N2);
-\draw (foot) to [bend right =15] (N3);
-\draw (foot) to [bend right =25] (N4);
-\end{tikzpicture}
-\caption{\label{fig:flower}
-The hydra $\iota(3)$}
-\end{figure}
-
-\input{movies/snippets/Omega_Small/iotaDef}
-
-Let us consider now some hydra \texttt{big\_h} out of the range of the injection $\iota$ (see Fig.~\vref{fig:h-omega-omega}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (2,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to  (N2);
-\end{tikzpicture}
-\caption{\label{fig:h-omega-omega}}
- The hydra \texttt{big\_h}.
-\end{figure}
-
-\input{movies/snippets/Omega_Small/bigHDef}
-
- Using the functions $m$ and $\iota$, we define a second hydra \texttt{small\_h}, and show
- there is a one-round battle that transforms \texttt{big\_h} into \texttt{small\_h}. Please note that,
-due to the hypothesis \texttt{Hvar}, we are interested in the termination of \emph{free} battles. 
-There is no problem to consider a round with (\texttt{m big\_h}) as the replication factor.
-
-\input{movies/snippets/Omega_Small/smallHDef}
-
-But, by hypothesis, $m$ is a variant. Hence, we infer the following inequality.
-
-\vspace{4pt}
-
-\input{movies/snippets/Omega_Small/mLt}
-
-In order to get a contradiction, it suffices to  prove the inequality
-$m(\texttt{big\_h}) \leq m(\texttt{small\_h})$ i.e.,  $m(\texttt{big\_h}\leq m(\iota (m(\texttt{big\_h})))$.
-
-\input{movies/snippets/Omega_Small/mGea}
-
-Intuitively, it means that, from any hydra of the form (\texttt{iota $i$}), the battle will 
-take (at least) $i$ rounds. Thus the associated measure cannot be less than $i$.
-Technically, we prove this lemma by Peano induction on $i$.
-
-\begin{itemize}
-\item The base case $i=0$ is trivial
-\item Otherwise, let $i$ be any natural number and assume  the inequality
-  $i \leq m(\iota(i))$.
-  \begin{enumerate}
-  \item  But the hydra $\iota(S(i))$ can be transformed in one round into
-    $\iota(i)$ (by losing its rightmost head, for instance)
-  \item Since $m$ is a variant, we have $m(\iota(i)) < m(\iota(S(i)))$,
-    hence  $i< m(\iota(S(i)))$, which implies  $S(i)\leq  m(\iota(S(i)))$.
-  \end{enumerate}
-\end{itemize}
-
-We are now ready to complete our impossibility proof.
-
-\vspace{4pt}
-
-\inputsnippets{Omega_Small/mGeb, Omega_Small/mGez,
-  Omega_Small/omegaSmallz}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\subsubsection{Conclusion}
-
-In order to build a variant for proving the termination of all hydra battles, we need to consider order structures more complex than the usual order on type \texttt{nat}. 
-The notion of \emph{ordinal number} provides a catalogue of well-founded order types.
-For a reasonably large bunch of ordinal numbers, \emph{ordinal notations} are data-types which allow the \coq{} user to define functions, to compute and prove some properties, for instance by reflection.
-
-The next chapter is dedicated to a generic formalization of ordinal notations, and chapter~\ref{chap:T1} to a proof of termination of all hydra battles with the help of an ordinal notation for the interval $[0,\epsilon_0)$\,\footnote{We use the mathematical notation $[a,b)$ for the interval $\{x|a\leq x < b\}$.}.
-\index{maths}{Notations!Interval}
-
-%--------------------------------------------------------------
-
-\chapter{Introduction to ordinal numbers and ordinal notations}
-\label{chap:ON}
-
-The proof of termination of all hydra battles presented in~\cite{KP82} is based
-on \emph{ordinal numbers}.
-From a mathematical point of view, an ordinal is a representative of an equivalence class for isomorphisms of  totally ordered well-founded sets.
-
-For the computer scientist, ordinals are tools for proving the totality of a given recursive function, or termination of a transition system. \emph{Ordinal arithmetic} 
-provides a set of functions whose properties, like \emph{monotony}, allow to define \emph{variants}, \emph{i.e.} strictly decreasing measures used in proofs of termination.
-
-\vspace{4pt}
-
-Let us have a look at Figure~\ref{fig:ordinal-sequence}. It presents a few items of a  sequence of ordinal numbers, which extends the sequence of natural numbers.
-
-\begin{figure}[h]
-  \centering
-\fbox{\Large
-  \begin{minipage}{1.0\linewidth}
-  \begin{align*}
-     &\textcolor{blue}{0},\,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,\ldots\\
-&\textcolor{red}{\omega},\,\omega+1,\omega+2,\omega+3,\ldots\\
-&\textcolor{red}{\omega\times 2},\,\omega\times 2+1,\ldots, \textcolor{red}{\omega\times 3},\,\omega\times 3+1,\ldots, \textcolor{red}{\omega\times 4},\ldots\\
-&\textcolor{red}{\omega^2},\ldots, \textcolor{red}{\omega^2\times 42},\ldots,\textcolor{red}{\omega^3},\ldots, \textcolor{red}{\omega^4},\omega^4+1,\ldots\\
-&\textcolor{red}{\omega^\omega},\ldots, \textcolor{red}{\omega^\omega+\omega^7\times 8},\ldots,\textcolor{red}{\omega^\omega\times 2},\omega^\omega\times 2+1, \ldots\\
-&\textcolor{red}{\omega^{\omega^\omega}},\ldots, \textcolor{red}{\omega^{\omega^\omega}+\omega^\omega\times 42+ \omega^{55}+\omega}, \ldots, \textcolor{red}{\omega^{\omega^{\omega+1}}}, \omega^{\omega^{\omega+1}}+1,\dots\\
-& \textcolor{red}{\epsilon_0 (= \omega^{\epsilon_0)}}, \epsilon_0+1, \epsilon_0+2, \epsilon_0+3, \ldots\\
-& \textcolor{red}{\epsilon_1}, \ldots, \textcolor{red}{\epsilon_2}, \ldots, \textcolor{red}{\epsilon_\omega},\ldots \\
-& \textcolor{red}{\Gamma_0}, \Gamma_0+1, \Gamma_0+2, \Gamma_0+3,\ldots, \textcolor{red}{\Gamma_0+\omega}, \ldots\\
-&\ldots
-  \end{align*}   
-  \end{minipage}}
-
-  \caption{A short overview of the sequence of ordinal numbers}
-  \label{fig:ordinal-sequence}
-\end{figure}
-
-Let us comment some features of this figure:
-
-\begin{itemize}
-\item The ordinals are listed in a strictly increasing order. 
-\item Dots : ``$\ldots$'' stand for  infinite sequences of ordinals, not shown for lack of space.
-For instance, the ordinal $42$ is not shown in the first line, but it exists, somewhere between $17$ and $\omega$.
-\item Each ordinal printed in black is the immediate successor of another ordinal.
-We call it a \emph{successor} ordinal.
-For instance, $12$ is the successor of $11$, and $\omega^4+1$ the successor of $\omega^4$.
-\item Ordinals (displayed in red)  that  follow immediately dots are called \emph{limit ordinals}.
-With respect to the order induced by this sequence, any limit ordinal $\alpha$ is the least upper bound of  the set $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
-\item
-For instance $\omega$ is the least upper bound of the set of all finite ordinals (in the first line). It is also the first limit ordinal, and the first infinite ordinal, in the sense that 
-the set $\mathbb{O}_\omega$ is infinite.
-\item The ordinal $\epsilon_0$ is the first number which is equal to its own exponential of base $\omega$.
-It plays an important role in proof theory, and is particularly studied in chapters~\ref{chap:T1} to \ref{chap:alpha-large}.
-\item Any ordinal is  either the ordinal \textcolor{blue}{$0$},
-a successor ordinal, or a \textcolor{red}{limit ordinal}.
-\end{itemize}
-
-\section{The mathematical point of view}
-
-\subsection{Well-ordered sets}
-Let us start with some definitions.
-A  \emph{well-ordered set} is a set provided with a binary relation $<$ which has the following properties.
-\begin{description}
-\item[irreflexivity] : $\forall x\in A, x\not< x$
-\item[transitivity] : $\forall x\,y\,z\in A, x<y \Rightarrow y<z \Rightarrow x<z$
-\item[trichotomy]: $\forall x\,y\in A, x<y \vee x = y \vee y < x$
-\item[well foundedness]: $<$ is well-founded (every element of $A$ is accessible)\footnote{In classical mathematics, we would say that there is no infinite sequence $a_1>a_2> \dots a_n> a_{n+1}\dots$ in $A$.
-In contrast, \coq's standard library contains
-an inductive definition of a predicate \texttt{Acc} which allows us to write 
-constructive proofs of accessibility (See \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}).}.
-\end{description}
-
-The best known examples of well-ordered sets are the set $\mathbb{N}$ of natural numbers (with the usual order $<$), as well as any finite segment $[0,i)=\{j\in\mathbb{N}\,|\,j<i\}$.
-The disjoint union of two copies of $\mathbb{N}$, \emph{i.e.} the set $\{0,1\}\times\mathbb{N}$ is also well-ordered,
-with respect to the order below:
-
-\begin{align*}
-(i,j) < (i,k) & \;\textit{\textbf{if} }\; j < k\\
-(0,k) < (1,l) & \;\textit{\textbf{for\,any}}\;k \;\textit{\textbf{and}} \; l
-\end{align*}
-
-\subsection{Ordinal numbers}
-
-\index{maths}{Ordinal numbers}
-
-Let $(A,<_A)$ and $(B,<_B)$ two well-ordered sets. $A$ and $B$ are said to have \emph{the same order type} if 
-there exists a strictly monotonous bijection $b$ from $A$ to $B$, \emph{i.e.} which verifies the proposition
-$\forall x\,y\in A,\, x <_A y \Rightarrow b(x) <_B  b(y)$.
-
-Having the same order type is an equivalence relation between well-ordered sets. Ordinal numbers (in short: \emph{ordinals})  are descriptions (\emph{names}) of the equivalence classes.
-For instance, the order type of $(\mathbb{N},<)$ is associated with the ordinal called  $\omega$, and the order we considered on 
-the disjoint union of $\mathbb{N}$ and itself is named $\omega+\omega$.
-
-In a set-theoretic framework, one can consider any ordinal $\alpha$ as a well-ordered set, whose  elements are just the ordinals strictly less than $\alpha$, \emph{i.e.} the \emph{segment} $\mathbb{O}_\alpha=[0, \alpha)$.
-So, one can speak about \emph{finite}, \emph{infinite}, \emph{countable}, etc., ordinals.
-Nevertheless, since we work within type theory,
-we do not identify ordinals as sets of ordinals, but consider the correspondence between ordinals and sets of ordinals as the function that maps $\alpha$ to $\mathbb{O}_\alpha$.
-For instance $\mathbb{O}_\omega=\mathbb{N}$, and $\mathbb{O}_7=\{0,1,2,3,4,5,6\}$.
-
-We cannot cite all the literature published on ordinals since Cantor's book
-\cite{cantorbook}, and 
-leave it to the reader to explore the bibliography.
-
-The introduction of Jos\'e Grimm's report~\cite{grimm:hal-00911710} contains a nice presentation of the main properties of ordinals.
-
-\section{Ordinal numbers in Coq}
-
-Two kinds of representation of ordinals are defined in our development.
-
-\begin{itemize}
-\item A ``mathematical'' representation of the set of countable ordinal numbers, after Kurt Schütte~\cite{schutte}.
-This representation uses several (hopefully harmless) axioms.
-We use it as a reference for proving the correctness of ordinal notations.
-  We are also progressively importing \gaia's~\cite{Gaia} definitions and theorems, based on Bourbaki's set theory~\cite{Grimm2016}.
-Chapter~\ref{gaia-chapter} is dedicated to the connexion between libraries \HydrasLib and \gaia.
-
-\item A family of \emph{ordinal notations} (also called \emph{[ordinal] notation systems}), \emph{i.e.} data types used to represent segments $[0,\mu)$, where $\mu$ is some countable ordinal.
-Each ordinal notation is defined inside the Calculus of Inductive Constructions (without axioms).
-Many functions are defined, allowing proofs by computation.
-Note that proofs of
-correctness of a given ordinal notation with respect to Schütte's model obviously use axioms.
-Please execute the \texttt{Print Assumptions} command in case of doubt.
-\end{itemize}
-
-% \section{Countable ordinals}
-
-% Chapter~\ref{chap:schutte} of this document presents an adaptation to \coq{} of an axiomatization in classical logic of the set of countable ordinals by K. Schütte~\cite{schutte}. 
-% That formalization is quite complex, technical and unshamedly non-constructive,  so we put its description  in the last chapter of this document.
-
-% Please note that Schütte considers the (uncountable) set $\mathbb{O}$ of all countable ordinals.
-This set is well ordered (which is one of Schütte's axioms), and associates to any ordinal $\alpha$ the segment $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
-
-% In our adaptation to \coq{}, we declare a type \texttt{Ord}, a binary relation \texttt{lt} (with infix notation \texttt{"\_<\_"}, and assume Schütte's axiom. In Chapter~\ref{chap:schutte},
-% we derive some interesting properties of countable ordinals from these axioms.
-
-It is interesting to compare proofs of a given property (for instance the associativity of addition) both in the computational framework of some ordinal notation, and in mathematical  models
-of Schütte or \gaia.
-
-\section{Ordinal Notations}
-
-Fortunately, the ordinals we need for  studying hydra battles are much simpler than Schütte's, and can be represented as quite simple data types in \gallina.
-
-Let $\alpha$ be some (countable) ordinal; 
-in \coq{} terms, we call \emph{ordinal notation for $\alpha$} a structure composed 
-of:
-\begin{itemize}
-\item A data type $A$ for representing all ordinals strictly below $\alpha$,
-\item A well founded order $<$ on $A$, 
-\item A correct function for comparing two ordinals. Note  that the reflexive closure of $<$ is thus a \emph{total order}.
-\end{itemize}
-
-Such a structure may be proved correct relatively to a bigger ordinal notation or to Schütte's or \gaia's model.
-
-\subsection{Classes for ordinal notations}
-
-From the \coq{} user's point of view, an ordinal notation is
-a structure allowing to compare two ordinals by computation, and proving by well-founded induction.
-
-\subsubsection{The \texttt{Comparison} and \texttt{Comparable} classes}
-
-We use the operational class \texttt{Comparison} of comparison functions to define the \texttt{Comparable} class, contributed by Jérémy Damour and Théo Zimmermann, which allows us to apply generic lemmas and tactics about decidable strict orders.
-The correctness of the comparison function is expressed through Stdlib's type 
-\texttt{Datatypes.CompareSpec} as specialized by \texttt{Datatypes.CompSpec}.
-
-\begin{Coqsrc}
-  Inductive CompareSpec (Peq Plt Pgt : Prop) :
-  comparison -> Prop :=
-    CompEq : Peq -> CompareSpec Peq Plt Pgt Eq
-  | CompLt : Plt -> CompareSpec Peq Plt Pgt Lt
-  | CompGt : Pgt -> CompareSpec Peq Plt Pgt Gt.
-
-Definition CompSpec {A} (eq lt : A -> A -> Prop) (x y : A) :=
- CompareSpec (eq x y) (lt x y) (lt y x).
-\end{Coqsrc}
-
-\emph{From Module~\href{../theories/html/hydras.Prelude/mparable.html\#Hvariant}{Prelude.Comparable}}
-
-\label{sect:comparable-def}
-
-\input{movies/snippets/Comparable/ComparableDef}
-
-\subsubsection{The \texttt{ON} class}
-
-The following class definition, parameterized with a type $A$, a binary relation \texttt{lt} on $A$, specifies that \texttt{lt} is a well-founded strict order, provided with a correct comparison function.
-
-\vspace{4pt}
-\noindent\emph{From
-Library~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}}
-
-\label{types:ON}
-\index{hydras}{Library OrdinalNotations!Type classes!ON}
-
-\input{movies/snippets/ON_Generic/ONDef}
-
-We give  also a few handy definitions and lemmas for any ordinal notation.
-
-\label{sect:on-lt-notation}
-\label{sect:on-le-notation}
-\label{sect:measure-ON}
-\label{sect:bigO-ON}
-
-\inputsnippets{ON_Generic/ONDefsa}
-\inputsnippets{ON_Generic/ONDefsb}
-
-\begin{remark}
-The infix notations \texttt{o<} and \texttt{o<=} are defined in order to make apparent the distinction between the various notation scopes that may co-exist in a same statement.
-So the infix \texttt{<} and \texttt{<=} are reserved to the natural numbers.
-In the mathematical formulas, however, we still use $<$ and $\leq$ for comparing ordinals.
-\end{remark}
-
-% \subsection{Ordinal notations and  termination measures}
-% \label{sect:measure-ON}
-
-% The following lemma (together with the type class mechnism) allows us to define termination measures over any ordinal notation.
-It is just an application of  the libraries \texttt{Coq.Wellfounded.Inverse\_Image}
-% and  \texttt{Coq.Wellfounded.Inclusion}.
-
-% \begin{Coqsrc}
-% Definition measure_lt {A:Type}{lt: relation A}
-%             {compare : A -> A -> comparison}
-%             {on : ON lt compare}
-%             {B : Type} (m : B -> A) : relation B :=
-%              fun x y => on_lt (m x) (m y).
-
-% Lemma wf_measure  {A:Type}(lt: relation A)
-%             {compare : A -> A -> comparison}
-%             {on : ON lt compare}
-%             {B : Type}
-%             (m : B -> A):  well_founded (measure_lt m). 
-% \end{Coqsrc}
-
-% A simple example of application is given in Sect.~\vref{sect:merge-example}.
-
-\section{Example: the ordinal \texorpdfstring{$\omega$}{omega}}
-
-The simplest example of ordinal notation is built over the type \texttt{nat} of \coq's standard library. We just have to apply already proven lemmas about Peano numbers.
-
-\vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_Omega.html}{OrdinalNotations.ON\_Omega}}
-
-\inputsnippets{ON_Omega/OmegaDefa, ON_Omega/OmegaDefb}
-
-\section{Sum of  two ordinal notations}
-
-Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective types \texttt{A} and \texttt{B}.
-
- We consider a new strict order
-on the disjoint sum of the associated types, by putting all elements of \texttt{A} before the elements of \texttt{B} (thanks to Standard Library's relation operator \texttt{le\_AsB}).
-
-\vspace{4pt}
-\noindent
-\emph{From Library~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}}.
-
-\begin{Coqanswer}
-Inductive
-le_AsB (A B : Type) (leA : A -> A -> Prop) (leB : B -> B -> Prop)
-  : A + B -> A + B -> Prop :=
-| le_aa : forall x y : A, leA x y -> le_AsB A B leA leB (inl x) (inl y)
-| le_ab : forall (x : A) (y : B), le_AsB A B leA leB (inl x) (inr y)
-| le_bb : forall x y : B, leB x y -> le_AsB A B leA leB (inr x) (inr y)
-\end{Coqanswer}
-
-%\pagebreak
-
-\vspace{4pt}
-\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.ON_plus.html}{OrdinalNotations.ON\_plus}}
-
-\input{movies/snippets/ON_plus/Defs}
-
-In order to build an instance of \texttt{Comparable}, we have to define a correct comparison function.
-
-\inputsnippets{ON_plus/compareDef, ON_plus/compareCorrect}
-\inputsnippets{ON_plus/plusComp}
-
-Standard library's lemma \texttt{Wellfounded.Disjoint\_Union.wf\_disjoint\_sum} 
-is applied  to prove that our order \texttt{lt} is well-founded, allowing us to build an instance of \texttt{ON}:
-
-\inputsnippets{ON_plus/ltWf, ON_plus/OnPlus}
-
-\subsection{The ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
-
-The ordinal $\omega+\omega$ (also known as $\omega\times 2$) may be represented as the concatenation 
-of two copies of $\omega$ (Figure~\ref{fig:omega-plus-omega}).
-It is also represented by the two first lines of Figure~\ref{fig:ordinal-sequence}.
-
-\begin{figure}[h]
-   \centering
-   \begin{tikzpicture}[very thick, scale=0.5]
-\begin{scope}[color=blue]
-\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
-\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
-\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
-\node (Adots) at (6,0) {$\ldots$};
-\node(An) at (8,0)[label=below:$n$]{$\bullet$};
-\node(A2) at (10,0)[label=below:$n+1$]{$\bullet$};
-\node (Adots1) at (12,0) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red]
-\node(B0) at (14,0)[label=below:$0$,label=above:\textcolor{red}{$\omega$}]{$\bullet$};
-\node(B1) at (16,0)[label=below:$1$, label=above:$\omega+1$]{$\bullet$};
-\node(B2) at (18,0)[label=below:$2$,label=above:$\omega+2$]{$\bullet$};
-\node (Bdots) at (20,0) {$\ldots$};
-\node (Bn) at (22,0) [label=below:$p$, label=above:$\omega+p$]{$\bullet$};
-\node (Bdots2) at (24,0) {$\ldots$};
-\end{scope}
-\end{tikzpicture}
-   \caption{\textcolor{blue}{$\omega+{\color{red}\omega}$}}
-   \label{fig:omega-plus-omega}
- \end{figure}
-
-We can define this notation in \coq{} as an instance of \texttt{ON\_plus}.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/OmegaPlusOmegaDef}
-
-\vspace{4pt}
-
-We can now define abbreviations.
-For instance, the finite ordinals are represented by terms built with  the constructor \texttt{inl}, and the first infinite ordinal $\omega$ by the term \texttt{(inr 0)}.
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Omega_plus_omega/finiteOmega}
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Omega_plus_omega/ltOmega}
-
-% \label{warning:coercions}
-% \index{Coq!Coercions} 
-% \begin{remark}
-% Beware of coercions and notation scopes!
-% Let us consider the following goal:
-
-% \begin{Coqsrc}
-%  Goal (6 o< 8).
-%  auto with arith.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% 1 subgoal (ID 9)
-
-%   ============================
-%   6 o< 8
-% \end{Coqanswer}
-
-% Please keep in mind that the current notation scope interprets the infix \texttt{``<''} as the predicate \texttt{Omega\_plus\_omega.lt} and not \texttt{Nat.lt}.
-More,  the coercion mechanism converts the terms \texttt{6:nat} [resp.
-\texttt{8:nat} ]
-% into \texttt{inl 6} [resp. \texttt{inl 8}].  So, the initial goal is correctly interpreted by \coq{}, but not as an inequality between two natural numbers.
-
-% \begin{Coqsrc}
-% Set Printing All.
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% 1 subgoal (ID 337)
-
-%   ============================
-%   @on_lt nat Peano.lt Nat.compare Omega (S (S (S (S (S (S O))))))
-%     (S (S (S (S (S (S (S (S O))))))))
-% \end{Coqanswer}
-
-% Anyway, the initial goal is provable, using \texttt{le\_AsB}'s first constructor.
-
-% \begin{Coqsrc}
-%   constructor; auto with arith.
-% Qed.
-% \end{Coqsrc}
-
-% \end{remark}
-%
-
-\section{Limits and successors}
-
-Let us look again at our implementation of $\omega+\omega$. We can classify its elements into three categories:
-
-\begin{itemize}
-\item The least ordinal, \texttt{(inl 0)}, also known as  \texttt{(fin 0)}.
-\item The first infinite ordinal $\omega$.
-\item The remaining ordinals, either of the form \texttt{(inl (S $i$))} or \texttt{(inr (S $i$))} (in black on Figure~\ref{fig:ordinal-sequence}), called \emph{successor ordinals}.
-\end{itemize}
-
-\subsection{Definitions}
-It would be interesting to specify at the most generic level, what is a zero, a successor or a limit ordinal. Let $<$ be a strict order on a type $A$.
-
-\begin{itemize}
-\item A \emph{least} element is a minorant (in the large sense) of the full set  on $A$,
-\item $y$ is a \emph{successor} of $x$ if $x<y$ and there is no element between $x$ and $y$. We will also say that $x$ is a \emph{predecessor} of $y$.
-\item $x$ is a \emph{limit} if $x$ is not a least element, and for any $y$ such that $yo<x$,
- there exists some $z$ such that $y<z<x$.
-\end{itemize}
-
-The following definitions are in Library \href{../theories/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.
-
-\input{movies/snippets/MoreOrders/Defs}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove, that, in any ordinal notation system, every ordinal has at most one predecessor, and at most one successor.
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/predSuccUnicity.v}{exercises/ordinals/predSuccUnicity.v}.}
-
-\end{exercise}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove, that, in any ordinal notation system, if $\beta$ is a successor of $\alpha$,
-then for any $\gamma$, $\gamma<\beta$ implies 
-$\gamma\leq\alpha$.
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/exercises/ordinals/lt_succ_le.v}{exercises/ordinals/lt\_succ\_le.v}.}
-\end{exercise}
-
-\subsection{Limits and successors in \texorpdfstring{$\omega+\omega$}{omega+omega}}
-
-Using the definitions above, we can prove the following lemma:
-
-\vspace{4pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/limitIff}
-
-\vspace{4pt}
-
-Regarding successors, let us define the following function and prove its correctness:
-
-\input{movies/snippets/ON_Omega_plus_omega/succDef}
-
-\input{movies/snippets/ON_Omega_plus_omega/succCorrect}
-
-\vspace{4pt}
-
-We can also check whether an ordinal is a successor by a simple computation:
-
-\input{movies/snippets/ON_Omega_plus_omega/succb}
-
-\vspace{4pt}
-
-Finally, the nature of any ordinal is decidable (inside this notation system) :
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/ZeroLimitSuccDec}
-
-\section{Product of ordinal notations}
-
-Let \texttt{NA} and \texttt{NB} be two ordinal notations, on the respective  ordered types \texttt{A} and \texttt{B}.
-The product of \texttt{NA} and \texttt{NB} is considered as the concatenation of $B$ copies of $A$, ordered by the lexicographic order on $B\times A$.
-
-In \coq{}, we build an instance of class \texttt{ON} through a sequence of steps as for the sum of ordinal notations.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_mult.html}{OrdinalNotations.ON\_mult}}
-
-\inputsnippets{ON_mult/Defs}
-\inputsnippets{ON_mult/multComp,  ON_mult/ONMult}
-\inputsnippets{ON_mult/endDefs}
-
-\section{The ordinal \texorpdfstring{$\omega^2$}{omega2}}
-
-The ordinal $\omega^2$ (also called $\phi_0(2)$, see Chap.~\ref{chap:schutte}), is an instance of the multiplication presented in the preceding section.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
-
-\inputsnippets{ON_Omega2/Omega2Def, ON_Omega2/Defs}
-\inputsnippets{ON_Omega2/Ex1}
-
-\subsection{Arithmetic of \texorpdfstring{$\omega^2$}{omega^2}}
-
-\subsubsection{Successor}
-
-The successor of any ordinal is defined by a simple pattern-matching.
-
-\input{movies/snippets/ON_Omega2/succ}
-
-This function is proved to be correct w.r.t. the \texttt{Successor} predicate.
-
-\input{movies/snippets/ON_Omega2/succOK}
-
-\input{movies/snippets/ON_Omega2/succLemmas}
-
-\subsubsection{Addition}
-
-We can define on \texttt{Omega2} an addition which extends the addition on \texttt{nat}. Please note that this operation is not commutative:
-
-\inputsnippets{ON_Omega2/plusDef, ON_Omega2/plusExamples}
-
-\subsubsection{Multiplication}
-
-The restriction of ordinal multiplication to the segment $[0,\omega^2)$ is not a total function.
-For instance $\omega\times\omega= \omega^2$ is outside the set of represented values.
-Nevertheless, we can define two operations mixing natural numbers and ordinals.
-
-\input{movies/snippets/ON_Omega2/multFinDef}
-\input{movies/snippets/ON_Omega2/multFinExamples}
-
-Multiplication with a finite ordinal and addition are related through the following lemma:
-
-\inputsnippets{ON_Omega2/uniqueDecompositiona,
-  ON_Omega2/uniqueDecompositionb,
-  ON_Omega2/uniqueDecompositionz}
-
-\subsection{A proof of termination using \texorpdfstring{$\omega^2$}{omega^2}} 
-\label{sect:merge-example}
-
-Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily a total function which merges two lists (example contributed by Pascal Manoury).
-
-%\index{coq}{Commands!Function}
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
-
-\inputsnippets{ON_Omega2/Merge, ON_Omega2/Mergeb}
-
-\subsection{Yet another  proof of impossibility}
-\label{omega2-case}
-
-In Sect.~\vref{omega-case}, we proved that there exists no variant from \texttt{Hydra} to \texttt{(nat,$<$)}
-(\emph{i.e.} the ordinal $\omega$) for proving the termination of all hydra battles.
-We  prove now that  the ordinal $\omega^2$ is also insufficient for this purpose.
-
-The proof we are going to comment has exactly the same structure as in Section~\ref{omega-case}.
- Nevertheless, the proof of technical  lemmas is a little more complex, due to 
- the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
-Consider for instance that there exists an infinite number of ordinals  between
-$\omega$ and $\omega\times 2$.
-
-The detailed  proof script is in the file
-\href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/Hydra/Omega2_Small.v}{theories/ordinals/Hydra/Omega2\_Small.v}.
-
-\subsubsection{Preliminaries}
-Let us assume there is a variant from \texttt{Hydra} into $\omega^2$  for proving the   termination of all hydra battles.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}
-
-\input{movies/snippets/Omega2_Small/Impossibility}
-\input{movies/snippets/Omega2_Small/Impossibilitya}
-
-We  follow the same pattern as in Sect.~\ref{omega-case}.
-First, we define an injection $\iota$ from type \texttt{ON\_Omega2.t} into \texttt{Hydra}, by
- associating to  each ordinal $\omega\times i+ j = (i,j)$ the hydra with $i$ branches of length $2$ and
-$j$ branches of length $1$.
-
-%% revenir ici
-
-\vspace{4pt}
-\emph{From Module ~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}
-
-\input{movies/snippets/Omega2_Small/Impossibilityc}
-
-For instance, Figure~\vref{fig:essai2} shows the hydra associated to the ordinal 
-$(3,5)$, a.k.a. $\omega\times 3 + 5$.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.4]
-\node (foot) at (6,0) {$\bullet$};
-\node (N1) at (1,3) {$\bullet$};
-\node (N2) at (3,3) {$\bullet$};
-\node (N3) at (5,3) {$\bullet$};
-\node (N4) at (8,3) {$\Smiley[2][green]$};
-\node (N5) at (11,3) {$\Smiley[2][green]$};
-\node (N6) at (14,3) {$\Smiley[2][green]$};
-\node (N7) at (17,3){$\Smiley[2][green]$};
-\node (N8) at (20,3){$\Smiley[2][green]$};
-\node  (N9) at (0,5) {$\Smiley[2][green]$};
-\node (N10) at (2,5) {$\Smiley[2][green]$};
-\node (N11) at (4,5) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (foot) -- (N5);
-\draw (foot) -- (N6);
-\draw (foot) to [bend right=10] (N7);
-\draw (foot) to [bend right=15] (N8);
-\draw (N1) to [bend left=10] (N9);
-\draw (N2) -- (N10);
-\draw (N3) -- (N11);
-\end{tikzpicture}
-\caption{\label{fig:essai2}
-The hydra $\iota(\omega\times 3+5)$}
-\end{figure}
-
-Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5]
-\node (foot) at (2,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (3,4) {$\Smiley[2][green]$};
-\node (N3) at (1,4) {$\Smiley[2][green]$};
-\draw (foot) -- (N1);
-\draw (N1) to [bend right =15] (N2);
-\draw (N1) to  [bend left=15](N3);
-\end{tikzpicture}
-\caption{\label{fig:h-omega2-small}}
- The hydra \texttt{big\_h}.
-\end{figure}
-
-\input{movies/snippets/Omega2_Small/Impossibilityb}
-
- In a second step, we build a ``smaller'' hydra\footnote{With respect to the measure $m$.}.
-
-\input{movies/snippets/Omega2_Small/Impossibilityd}
-
-\vspace{4pt}
-
-Like in Sect.~\ref{omega-case}, we prove the inequality $m\;\texttt{big\_h} \,\texttt{o<=}\, m\;\texttt{small\_h} \,\texttt{o<=}\, m\;\texttt{big\_h}$, which is obviously false.
-
-\subsubsection{Proof of the inequality \texttt{m small\_h o< m big\_h}}
-
-In order to prove the inequality  \texttt{m\_lt: m small\_h o< m big\_h}, it suffices to
-build a battle transforming \texttt{big\_h} into \texttt{small\_h}.
-
-First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h} as $(i,j)$.
-If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
-If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the following result.
-
-\input{movies/snippets/Omega2_Small/bigToSmall}
-
-Since $m$ is a variant, we infer the following inequality:
-
-\input{movies/snippets/Omega2_Small/mLt}
-
-\subsubsection{Proof of the inequality \texttt{m big\_h o<= m small\_h} }
-
-The proof of the inequality \texttt{m big\_h o<= m small\_h} is quite more complex than in Sect~\ref{omega-case}.
-If we consider any ordinal $\alpha=(i,j)$, where $i>0$, there exists an infinite number of
-ordinals strictly less than $\alpha$, and there exists an infinite number of battles that start from
-$\iota(\alpha)$.
-Indeed, at any configuration $\iota(k,0)$, where $k>0$, the hydra can freely choose any replication number.
-Intuitively, the measure of such a hydra must be large enough for taking into account
-all the possible battles issued from that hydra.
-Let us now give more technical details.
-
-The first steps of our proof start a well-founded induction over $\omega^2$.
-
-\input{movies/snippets/Omega2_Small/mGe}
-
-A case analysis on $i$ and $j$ allows us to
-consider three cases :
-
-\begin{itemize}
-\item $i=j=0$: the inequality is trivial.
-\item $i=1+l, j=0$ ($(i,j)$ is a limit ordinal): By the induction hypothesis \texttt{IHij},
-  $(l,k)\texttt{o<=} m(\iota(l,k))$ for any $k$. But (by the rules of the hydra game), $\iota(i,0)$ is transformed into any $\iota(l,k)$ in one round. Thus $m(\iota(l,k)) < m(\iota(i,0))$ for any $k$.
-  Therefore, $(l,k) <  m(\iota(i,0))$ for any $k$, thus
-  $(i,0)\,\texttt{o<=}\,m(\iota(i,0))$.
- \item $j= l+1$  ($(i,j)$ is a successor). We apply the induction hypothesis to the pair $(i,l)$.
- \end{itemize}
-
- Please look at the proof script for more details.
-
- \input{movies/snippets/Omega2_Small/mGeb}
-
-\subsubsection{End of the proof}
-From \texttt{m\_ge}, we get \texttt{$m$\, big\_h o<= $m$\, small\_h = $m$\,(iota (m big\_h)) }. 
-Since $<$ is a strict order (irreflexive  and transitive), this inequality is incompatible with the strict inequality  \texttt{m small\_h o< m big\_h} (lemma \texttt{m\_lt}).
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#Impossible}{Hydra.Omega2\_Small}}
-
-  \input{movies/snippets/Omega2_Small/Impossible}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that there exists no variant $m$ from \texttt{Hydra} into $\omega^2$ for proving
-    the  termination of all \emph{standard} battles.
-\end{exercise}
-
-\begin{remark}
- We  prove in Chapter~\ref{ks-chapter} a generalization of the impossibility lemmas of
-Sect.~\ref{omega-case} and this section, with the same proof structure, but with much more 
-complex technical details.
- \end{remark}
-
-% \index{Exercises}
-% \begin{exercise}
-
-% \label{sec:orgheadline63}
-% Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.
-
-% \begin{itemize}
-% \item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
-%   lexicographic ordering
-% \item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.
-
-% For instance, the following inequality holds:
-% \[\langle 4,3,3,3,3,3,3,2,2,2 \rangle\,<\,\langle 4,4,2 \rangle\]
-% \end{itemize}
-
-% \end{exercise}
-
-\section{A notation for finite ordinals}
-
-Let $n$ be some natural number. The segment associated with $n$ is the interval 
-$[0,n)\,=\,\{0,1,\dots,n-1\}$. 
-One may represent the ordinal $n$ by a sigma type.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
-
-\label{def:Finite-ord-type}
-
-\input{movies/snippets/ON_Finite/Defs}
-
-\subsubsection{Examples}
-
-For instance, let us build two elements of the segment $[0, 7)$, \emph{i.e.} two
-inhabitants of   type (\texttt{t 7}), and prove a simple  inequality (see Fig.~\ref{fig:O7}).
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[very thick, scale=0.6]
-
-\node (N0) at (0,0) {$\bullet$};
-\node (i0) at (0,1) {$0$};
-\node (N1) at (2,0) {$\bullet$};
-\node (i1) at (2,1) {$1$};
-\node (N2) at (4,0) {$\bullet$};
-\node (i2) at (4,1) {$2$};
-\node (N3) at (6,0) {$\bullet$};
-\node (i3) at (6,1) {$3$};
-\node (N4) at (8,0) {$\bullet$};
-\node (i4) at (8,1) {$4$};
-\node (N5) at (10,0) {$\bullet$};
-\node (i5) at (10,1) {$5$};
-\node (N6) at (12,0) {$\bullet$};
-\node (i6) at (12,1) {$6$};
-\node(alpha1) at (4,-1) {$\alpha_1$};
-\node(alpha2) at (10,-1) {$\beta_1$};
-\end{tikzpicture}
-
-\caption{The segment $\mathbb{O}_7$\label{fig:O7}}
-\end{figure}
-
-%\index{coq}{Commands!Program}
-
-\input{movies/snippets/ON_Finite/Example1}
-
-Note that the type (\texttt{t 0}) is empty, and that, for any natural number
- $n$, $n$ does not belong to (\texttt{t $n$}).
-
- \inputsnippets{ON_Finite/t0Empty, ON_Finite/bad10}
-
-\subsubsection{Comparison function}
-
-In order to build an instance of \texttt{ON}, we define a comparison function,  and prove its correctiness.
-
-\vspace{4pt}
-
-\input{movies/snippets/ON_Finite/compareDef}
-
-\begin{remark}
- The proof of \texttt{compare\_correct} uses a well-known pattern of \coq{}.
-Let us consider  the following subgoal.
-
-\input{movies/snippets/ON_Finite/compareCorrectb}
-
-Applying the tactic \texttt{f\_equal} generates a simpler subgoal.
-
-\input{movies/snippets/ON_Finite/compareCorrectc}
-
-We have now to prove that there exists at most one  proof of (\texttt{Nat.ltb x0 (S n)}). This is not obvious, but  a consequence of the following lemma of library 
-\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.Eqdep_dec.html}{Coq.Logic.Eqdep\_dec}.
-
-\index{Coq}{Unicity of equality proofs}
-\label{sect:eq-proof-unicity}
-
-\begin{Coqanswer}
-eq_proofs_unicity_on :
-forall (A : Type) (x : A),
-(forall y : A, x = y \/ x <> y) -> 
-forall (y : A) (p1 p2 : x = y), p1 = p2
-\end{Coqanswer}
-
-Thus unicity of proofs of \texttt{Nat.ltb x0 (S n)}  comes from the decidability of
-equality on type \texttt{bool}.
-This is why we used the boolean function \texttt{Nat.ltb} instead of the inductive predicate \texttt{Nat.lt} in the definition of type \texttt{t $n$} (see page~\pageref{def: Finite-ord-type}).
-For more information about this pattern, please look at the numerous mailing lists and 
-FAQs on \coq{}).
-
-\inputsnippets{ON_Finite/compareCorrectd}
-
-\end{remark}
-
-\begin{remark}
-
-Please note  that attempting to compare a term  of type (\texttt{t $n$}) with a term of
-type (\texttt{t $p$})  leads to an error if $n$ and $p$ are not convertible.
-
-\input{movies/snippets/ON_Finite/Example2}
-
-\end{remark}
-
-\subsubsection{Building an instance of \texttt{ON}}
-
-Applying lemmas of the libraries \texttt{Coq.Wellfounded.Inverse\_Image}, \linebreak
- \texttt{Coq.Wellfounded.Inclusion}, and \texttt{Coq.Arith.Wf\_nat}, we prove that our
-relation \texttt{lt} is well founded.
-
-\input{movies/snippets/ON_Finite/ltWf}
-
-Now we can build our instance of \texttt{OrdinalNotation}.
-
-\input{movies/snippets/ON_Finite/ONInstance}
-
-\begin{remark}
-It is important to keep in mind  that the integer $n$ is not an ``element'' of \texttt{FinOrd $n$}.
-In set-theoretic presentations of ordinals, the set associated with the ordinal $n$ is $\{0,1,\dots,n-1\}$.
-In our formalization, the interpretation of an ordinal as a set is realized by the function \texttt{bigO} (see Section\vref{sect:bigO-ON}).
-\end{remark}
-
-\begin{remark}
- There is no interesting arithmetic on finite ordinals, since functions like successor, addition, etc.,  cannot be represented in \coq{} as \emph{total} functions.
-\end{remark}
-
-\section{See also \dots}
-
-Finite ordinals are also formalized in \ssreflect/\mathcomp~\cite{MCB}.
-In Module~\href{../theories/html/gaia_hydras.ON_gfinite.html}{gaia\_hydras.ON\_gfinite}, we build an instance of class \texttt{ON}.
-
-\inputsnippets{ON_gfinite/finordLtDef, ON_gfinite/finordON}
-
-\inputsnippets{ON_gfinite/Examples, ON_gfinite/Examplesb}
-
-  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.
-
-%%%
-
-\section{Comparing two ordinal notations}
-
-It is sometimes useful to compare two ordinal notations with respect to expressive power
-(the segment of ordinals  they represent).
-
-The following class specifies a strict inclusion of segments.
-The notation \texttt{OA} describes a segment $[0,\alpha)$, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented in \texttt{OA}).
-We require also  that the comparison functions of the two notation systems are compatible.
-
-\begin{figure}[h]
-   \centering
-   \begin{tikzpicture}[very thick, scale=0.6]
-\begin{scope}[color=blue]
-\node (A) at (0,0) {$A$};
-\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
-\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
-\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
-\node (Adots) at (6,0) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red]
-\node (B) at (0,2) {$B$};
-\node(B0) at (2,2)[label=above:$0$]{$\bullet$};
-\node(B1) at (3,2)[label=above:$1$]{$\bullet$};
-\node(B2) at (4,2)[label=above:$2$]{$\bullet$};
-\node (Bdots) at (6,2) {$\ldots$};
-\node (b) at (8,2) [label=above:$\omega$]{$\bullet$};
-\node (bsucc) at (9,2) [label=above:$\omega+1$]{$\bullet$};
-\node (Bdots2) at (10,2) {$\ldots$};
-\end{scope}
-\begin{scope}[color=red!50!blue]
-\draw [->,thin] (A0) -- node [auto] {$\iota$} (B0);
-\draw [->,thin] (A1) -- node [auto] {$\iota$} (B1);
-\draw [->,thin] (A2) -- node [auto] {$\iota$} (B2);
-\draw [->,thin] (Adots) -- node [auto] {$\iota$} (Bdots);
-\end{scope}
-\end{tikzpicture}
-   \caption{\textcolor{blue}{$A$} is a sub-segment  of \textcolor{red}{$B$}}
-   \label{fig:subsegment}
- \end{figure}
-
-If \texttt{OB} is presumed to be correct, then we may consider that \texttt{OA} ``inherits'' its correctness from the bigger notation system \texttt{OB}.
-
-\label{types:SubON}
-\index{hydras}{Library OrdinalNotations!Type classes!SubON}
-
-following definition
-(in ~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{ON\_Generic}).
-
-\input{movies/snippets/ON_Generic/SubONDef}
-
-For instance, we prove that \texttt{Omega} is a sub-notation of
-\texttt{Omega\_plus\_Omega} (with $\omega$ as the first ``new'' ordinal, and \texttt{fin} as the injection).
-
-\vspace{10pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}
-
-\input{movies/snippets/ON_Omega_plus_omega/Incl}
-
-We can also show that, if $i<j$, then the segment $[0,i)$ is a ``sub-segment'' of
-$[0,j)$. Since the terms  ($t\;i$) and ($t\;j$) are not convertible, we consider a ``cast'' 
-function $\iota$ from ($t\;i$) into ($t\;j$), and prove that this function is  a monotonous bijection  from ($t\;i$) to
-the segment $[0,i)$ of ($t\;j$).
-
-%\index{coq}{Commands!Program}
-
-\vspace{10pt}
-
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}
-
-\inputsnippets{ON_Finite/InclIJ, ON_Finite/InclIJa,
-  ON_Finite/InclIJb,ON_Finite/InclIJc, ON_Finite/InclIJd,
-ON_Finite/InclIJz}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Prove that \texttt{Omega\_plus\_Omega} cannot be a sub-notation of \texttt{Omega}.
-\end{exercise}
-
-\index{hydras}{Projects}
-\begin{project}
-Adapt the definition of \texttt{Hvariant} (Sect.~\ref{sect:hvariant-def}) in order to
-have an ordinal notation as argument. Prove that if $O_A$ is a sub-notation of $O_B$, then any variant defined on  $O_A$ can be automatically transformed into 
-a variant on $O_B$.
-\end{project}
-
-\section{Comparing an ordinal notation with Schütte's model}
-
-Finally, it may be interesting to compare an ordinal notation with the more theoretical model from Schütte (well, at least with our formalization of that model).
-This would be a relative proof of correctness of the considered  ordinal  notation.
-
-The following class specifies that a notation \texttt{OA} describes a segment $[0,\alpha)$,
-where $\alpha$ is a countable ordinal \emph{à la}  Schütte.
-
-\label{types:ON-for}
-\index{hydras}{Library OrdinalNotations!Type classes!ON\_correct}
-
-\input{movies/snippets/ON_Generic/ONCorrect}
-
-For instance, the following theorem tells that \texttt{Epsilon0}, our notation system for the segment $[0,\epsilon0)$ is a correct implementation of the theoretically defined  ordinal $\epsilon_0$
-(see chapter~\ref{chap:schutte} for more details).
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
-
-\input{movies/snippets/Correctness_E0/Epsilon0Correct}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Same work, but replace Schütte's model with \gaia's.
-\end{project}
-
-\begin{project}
-  When you have read Chapter~\ref{chap:schutte}, prove that the sum of two ordinal notations \texttt{ON\_plus} implements the addition of ordinals.
-\end{project}
-
-\section{Isomorphism of ordinal notations}
-
-In some cases we want to show that two notation systems describe the same segment (for instance $[0,3+\omega)$ and $[0,\omega)$\;).
-For this purpose, one may prove that the two notation systems are order-isomorphic.
-
-\index{hydras}{Library OrdinalNotations!Type classes!ON\_Iso}
-
-\label{types:ON-iso}
-
-\input{movies/snippets/ON_Generic/ONIso}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-\label{exo:i-plus-omega}
-Let $i$ be some natural number. Prove that the notation systems 
-\texttt{Omega} and (\texttt{ON\_plus (OrdFin $i$) Omega}) are isomorphic.
-
-{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$, that we prove also in larger notation systems, as well as in Schütte's model.}
-This exercise is partially solved for $i=3$ (in ~\href{../theories/html/hydras.OrdinalNotations.Example_3PlusOmega.html}{OrdinalNotations.Example\_3PlusOmega}).
-
-\end{exercise}
-
-\index{hydras}{Projects}
-\label{exo:ON-mult}
-\begin{project}
-% Define in \coq{} the product of two ordinal notations $N_A$ and $N_B$.
-% If $A$ [resp. $B$] is the underlying type of $N_A$ [resp. $N_B$], the
-% product \texttt{ON\_mult $N_A$ $N_B$} is implemented over the cartesian product $B\times A$ (with the lexicographic ordering).
-
-This exercise is about the non-commutativity of the multiplication of ordinals, reflected in ordinal notations.
-
-For instance, the
-elements of the product (\texttt{ON\_mult Omega (FinOrd 3)}) are ordered as follows.
-\[(0,0),(0,1),(0,2),(0,3),(0,4),\dots,{\color{red}(1,0),} (1,1),(1,2),\dots, {\color{red}(2,0)},(2,1),(2,2),\dots\]
-
-Note that the elements of  (\texttt{ON\_mult (FinOrd 3) Omega}) are differently ordered (without limit ordinals):
-\[(0,0),(1,0),(2,0),(0,1),(1,1),(2,1),(0,2),(1,2),(2,2),(0,3),\dots\]
-
-Prove formally  that (\texttt{ON\_mult (FinOrd $i$) Omega}) is isomorphic to
-\texttt{Omega}  whilst
-\texttt{Omega}  is a sub-notation of (\texttt{ON\_mult Omega (FinOrd $i$)}),
-for any strictly positive $i$.
-
-\textbf{Note:} Like Exercise~\ref{exo:i-plus-omega}, this project corresponds to the [in]equalities $i+\omega=\omega<\omega+i$, for any natural number $i$.
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-Consider two isomorphic ordinal notations \texttt{OA} and \texttt{OB}.
-Prove that, if \texttt{OA} [resp. \texttt{OB}] is a correct implementation 
-of $\alpha$ [resp. $\beta$], then $\alpha=\beta$.
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-\label{project:succ-limit-dec}
-Add to the class \texttt{ON} the requirement that for any $\alpha$ it is decidable whether $\alpha$ is $0$, a successor or a limit ordinal.
-
-\textbf{Hint:}   Beware of the instances associated with sum and product of notations!
-  You may consider additional fields 
-to make the sum and product of notations ``compositional''.
-
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-\label{project:on-setoid}
-Reconsider the  class \texttt{ON}, with an equivalence instead of Leibniz equality.
-\end{project}
-
-%%%% ICI ICI
-
-\section{Other ordinal notations}
-%%% TODO : Fix the multiplication function in branch FixOmegaOmega
-
-% \index{hydras}{Projects}
-
-% \begin{project}
-% The directory \texttt{theories/OmegaOmega} contains an ad-hoc formalization of $\omega^\omega$, contributed by Pascal Manoury.
-Every ordinal $\alpha$ is represented by a list $l$ whose elements are the coefficients of $\omega$ in  the Cantor normal form of $\alpha$ (in reverse order).
-For instance, the ordinal
-% $\omega^{8}\times 5 + \omega^{6}\times 8 + \omega^2\times 10 + \omega + 7$ is represented by the list \texttt{[5;0;8;0;0.0;10,1,7]}.
-
-%  Develop this representation and compare it with the other ordinal notations.
-
-% \end{project}
-
-\index{hydras}{Projects}
-
-\begin{project}
-Let $N_A$ be a notation system for ordinals strictly less than $\alpha$, 
-with the strict order $(A,<_A)$. Please build the notation system
-\texttt{ON\_Expl $N_A$}, on the type of multisets of elements of $A$
-(or, if preferred, the type of non-increasing finite sequences on $A$,
-provided with the lexicographic ordering on lists).
-
-For instance, let us take $N_A=\texttt{Omega}$, and take $\alpha=\langle 4,4,2,1,0\rangle$,
- $\beta=\langle 4,3,3,3,3,3,2\rangle$, and $\gamma=\langle 5\rangle$. Then $\beta<\alpha<\gamma$.
-
-In contrast the list $\langle5,6,3,3\rangle$ is not non-increasing (\emph{i.e.} sorted w.r.t. $\geq$), so it is not to be considered.
-
-Note that if the notation $N_A$ implements the ordinal 
-$\alpha$,  the new notation $\omega^{N_A}$ must implement the ordinal $\phi_0(\alpha)$, a.k.a. $\omega^\alpha$ (see chapter~\ref{chap:schutte})
-
-\end{project}
-
-\begin{remark}
- The set of ordinal terms in Cantor normal form (see Chap.~\ref{chap:T1}) and 
-in Veblen normal form (see 
-\href{../theories/html/hydras.Gamma0.Gamma0.html}{Gamma0.Gamma0}) are shown to be ordinal notation systems, but there is a lot of work to be done in order to unify ad-hoc  definitions and proofs which were written before the definition of the \texttt{ON} type class.
-\end{remark}
-
-\section{Ordinal Notations and \gaia's ordinals}
-\gaiasign Module~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge} contains an instance of class \texttt{ON E0lt compare}, where \texttt{E0lt} is the order on the well-formed ordinal terms below $\epsilon_0$ defined in \gaiaHydras library (please see Chapter~\ref{gaia-chapter}).
-
-%------------------------------------------------------------------------
-
-\chapter[A proof of termination, using epsilon0]{A proof of termination, using ordinals below \texorpdfstring{$\epsilon_0$}{Epsilon0}}
-
-\label{cnf-math-def}
-\label{chap:T1}
-
-In this chapter, we adapt to \coq{} the well-known~\cite{KP82}  proof that Hercules eventually wins every battle, whichever the strategy  of each player.
-In other words, we present  a formal and self contained proof of termination  of all [free] hydra battles.
-First, we take from Manolios and Vroon~\cite{Manolios2005} a representation of the ordinal $\epsilon_0$ as terms in Cantor normal form.
-Then, we define a variant for hydra battles as a measure that maps any hydra to some ordinal strictly less than $\epsilon_0$.
-
-\section{The ordinal \texorpdfstring{\(\epsilon_0\)}{epsilon0}}
-\label{sec:epsilon0-intro}
-
-The ordinal \(\epsilon_0\) is the least ordinal number that satisfies 
-the equation \(\alpha = \omega^\alpha\), where \(\omega\) is 
-the least infinite ordinal.
-Thus, we can intuitively consider \(\epsilon_0\) as an
-\emph{infinite} \(\omega\)-tower.
-
-\subsection{Cantor normal form}
-\index{maths}{Cantor normal form}
-
-Any ordinal strictly less that \(\epsilon_0\) 
-can be finitely represented by a unique  \emph{Cantor normal form}, 
-that is, an expression  which is either  the ordinal \(0\) or 
-a sum  \(\omega^{\alpha_1} \times n_1 + \omega^{\alpha_2} \times n_2 + 
-  \dots + \omega^{\alpha_p} \times n_p\) where all the \(\alpha_i\) 
-are ordinals in Cantor  normal form, \(\alpha_1 > \alpha_2 > \alpha_p\), 
-and all the \(n_i\) are positive integers.
-
-An example of Cantor normal form is displayed in Fig \ref{fig:cnf-example}:
-Note that  any ordinal of
-the form \(\omega^0 \times i + 0\) is just written \(i\).
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
-\node[color=blue]{$\omega^{(\omega^\omega\,+\, \omega^2 \times 8 \,+\, \omega)}+ \omega^\omega + \omega^4+ 6$};
-\end{tikzpicture}
-\caption{\label{fig:cnf-example}
-An ordinal in Cantor normal form}
-\end{figure}
-
-In the rest of this section, we define an inductive type for representing in \texttt{Coq}
-all the ordinals strictly  less than  \(\epsilon_0\), then extend some arithmetic operations
-to this type, and finally prove that our representation fits well with 
-the expected mathematical properties: the order we define is a well order, 
-and the decomposition into Cantor normal form  is consistent 
-with the implementation of the arithmetic operations of exponentiation of base \(\omega\) 
-and addition.
-
-\paragraph*{Remark}
-\label{sec:orgheadline65}
-Unless explicitly mentioned, the term ``ordinal" will be used instead of
-``ordinal strictly less than \(\epsilon_0\)" (except in Chapter~\ref{chap:schutte} where it stands for ``countable ordinal'').
-
-\subsection{A data type for  ordinals in Cantor normal form}
-\label{sec:orgheadline72}
-\label{sec:T1-inductive-def}
-
-% Our user contribution~\cite{CantorContrib} represents 
-% the set of ordinals strictly less than $\epsilon_0$ in Cantor normal form as in~\cite{Manolios2005}, and also the set
-% of ordinals strictly  less than $\Gamma_0$ in Veblen normal form.
-
-    Let us define an inductive type whose 
-constructors are respectively associated
-with the ways to build Cantor normal forms:
-
-\begin{itemize}
-\item the ordinal \(0\)
-\item the construction \((\alpha,\, n,\,\beta)  \mapsto \omega^\alpha \times (n + 1)+ \beta \quad (n\in\mathbb{N})\)
-\end{itemize}
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#T1}{Epsilon0.T1}}
-
-\label{types:T1}
-%\index{Constants!zero:T1}
-\index{hydras}{Library Epsilon0!Types!T1}
-
-\input{movies/snippets/T1/T1Def}
-
-\subsubsection{Example}
-
-\label{alpha0-def}
-For instance, the ordinal  $\omega^\omega+\omega^3\times 5+2$ is represented by the following term of type \texttt{T1}:
-
-\input{movies/snippets/T1/alpha0}
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.5, level 1/.style={sibling distance=6cm},
-level 2/.style={sibling distance=35mm},  
-level 3/.style={sibling distance=17mm}]
-\node  {cons}
-  child {  node {cons}
-            child { node {cons} child {node {zero}} child {node{0}} child{node{zero}}}
-         child {node {0}}
-         child {node {zero}}}
-    child {node {0}}
-   child {node {cons} 
- child { node {cons} child {node {zero}} child {node{2}} child{node{zero}}}
-  child {node {4}}
-         child {node {cons} child {node {zero}} child {node{1}} child{node{zero}}}};
-
-\end{tikzpicture}
-
-\caption{The tree-like representation of the ordinal $\omega^\omega+\omega^3\times 5 +2$\label{fig:cnf-tree}}
-
-\end{figure}
-
-\paragraph{Remark}
-For simplicity's sake, we chose to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the construction (\texttt{cons $\alpha$ $n$ $\beta$}) is intended to represent the
-ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
-In a future version, we should replace  the type \texttt{nat} with \texttt{positive} in \texttt{T1}'s 
-definition. But this replacement would take a lot of time \dots{}
-
-\paragraph{Remark}
-The name \texttt{T1} we gave to this data-type  is proper to this development and refers
-to a hierarchy of ordinal notations.
-For instance, in Library \href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2},  the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{Gaia, schutte}).
-\noindent
-
-\input{movies/snippets/T2/T2Def}
-
-\subsection{About Gaia}
-\gaiasign Chapter~\vref{gaia-chapter} describes the present state of a
-project of making compatible the libraries \HydrasLib and \gaia. At present, both libraries contain their own version of the inductive type \texttt{T1}, with the same base name for the constructors.
-
-\inputsnippets{nfwfgaia/T1def}
-
-Thus, many examples of this and following chapters can be run
-in \gaia's context. We will signal any difference (notation, name) which may appear.
-A closer integration (same data type and functions) is still in project.
-
-\begin{remark}
-  \label{remark:hydra-gaia-renaming}
-  \textcolor{lookcolor}{In future releases, we plan to make some
-    \HydrasLib identifiers progressively  deprecated, in favour of \gaia's names.}
-\end{remark}
-
-\subsection{Abbreviations}
-
-Some abbreviations may help to write more concisely complex ordinal terms.
-
-\subsubsection{Finite ordinals}
-\label{sec:orgheadline67}
-
-For representing finite ordinals, \emph{i.e.} natural numbers, we first introduce a notation for terms of the form $n+1$, then define a coercion from type \texttt{nat} into \texttt{T1}.
-\label{sect:notation-FS}
-
-\label{sect:notation-F}
-
-\input{movies/snippets/T1/finiteOrds}
-
-\paragraph*{\gaiasign}
-  In \texttt{gaia.ssete9}, the $n$-th finite ordinal is also written
-  \texttt{$\backslash${F}\,$n$}. There is no coercion in \gaia from \texttt{nat} to \texttt{T1}.
-
-  \vspace{4pt}
-
-  From~\href{../theories/html/gaia_hydras.HydraGaia_Examples.html}{gaia\_hydras.HydraGaia\_Examples}.
-
-  \inputsnippets{HydraGaia_Examples/Demoa}
-
-% \index{Coq!Coercions}
-% \index{Functions!Coercions@Coercions (from nat to ordinal types)}
-% \begin{remark}
-% Please refer to the remark~\pageref{warning:coercions} about the use of coercions.
-% % The use of coercions like \texttt{fin} allow us to be close to the mathematical tradition where natural numbers are ordinals too.
-% % Nevertheless, it may happen that a goal like \texttt{3 < 5} could be 
-% % interpreted as \texttt{(lt (fin 3) (fin 5))},  depending on the current notation scope.  
-% % When this misinterpretation happens, tactics like \texttt{auto with arith}, \texttt{lia} do not work!
-% % Thus, it is useful to write \texttt{(3 < 5)\%nat}  an inequality between two natural numbers. 
-% \end{remark}
-
-\subsubsection{The ordinal \(\omega\)}
-\label{sec:orgheadline68}
-
-  Since \(\omega\)'s Cantor normal form is
-i.e. \(\omega^{\omega^0}\times 1+ 0\), we can define the following abbreviation:
-
-\label{sect:omega-notation2}
-
-\input{movies/snippets/T1/omegaDef}
-
-Note that \texttt{omega} is not an identifier in \HydrasLib, thus any tactic like \texttt{unfold omega} would fail.
-
-\paragraph*{\gaiasign}
-In \texttt{gaia.ssete9}, the ordinal $\omega$ is bound to the \emph{constant}  \texttt{T1omega}.
-
-\subsubsection{The ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
-\label{sect:notation-phi0}
-We provide also a notation for ordinals of the form $\omega^\alpha$.
-
-\index{hydras}{Library Epsilon0!Notations!phi0@phi0 (exponential of base omega)}
-
-\input{movies/snippets/T1/phi0Def}
-
-\index{maths}{Additive principal ordinals}
-
-\begin{remark}
-\label{sec:orgheadline69}
-The name \(\phi_0\)
-   comes from ordinal numbers theory. In~\cite{schutte}, Schütte defines 
-$\phi_0$  as the ordering (\emph{i.e.} enumerating) function of the set  of \emph{additive principal ordinals} \emph{i.e.} strictly positive ordinals $\alpha$ that verify $\forall \beta<\alpha, \beta+\alpha=\alpha$.
-For Schütte,  $\omega^\alpha$ is just a notation for $\phi_0(\alpha)$.
-See also Chapter~\ref{chap:schutte} of this document.
-\end{remark}
-
-\paragraph*{\gaiasign}
-In \texttt{gaia.ssete9} the identifier \texttt{phi0} is bound to a plain constant (not a notation).
-
-\subsubsection{The hierarchy of \(\omega\)-towers:}
-\label{sec:orgheadline71}
-
-The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximated by the sequence of $\omega$-towers (see also Sect~\vref{sect:epsilon0-as-limit} ).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/towerDef}
-
-For instance, Figure~\ref{fig:tower7} represents  the ordinal returned by the
- evaluation of the term \texttt{omega\_tower 7}.
-
-\begin{figure}[htb]
-\centering
-\begin{tikzpicture}[scale=2, every node/.style={transform shape}]
-\node[color=blue]{$\omega^{{{\omega}^{{{\omega}}^{{{\omega}}^{{\omega^{{\omega}^{\omega}}}}}}}}$};
-\end{tikzpicture}
-\caption{\label{fig:tower7}
-The $\omega$-tower of height 7}
-\end{figure}
-
-\subsection{Pretty-printing ordinals in Cantor normal form}
-\label{sect:ppT1}
-\index{hydras}{Library Epsilon0!Types!ppT1}
-
-Let us consider again the ordinal $\alpha_0$ defined in section~\vref{alpha0-def}. 
-If we ask \coq{} to print its  normal form, we get a hardly readable term of type \texttt{T1}.
-
-\input{movies/snippets/T1/alpha0Compute}
-
-The following data type defines a more readable abstract syntax for  ordinals terms in Cantor normal form:
-
-\label{types:ppT1}
-\index{hydras}{Library Epsilon0!Functions!pp@ pp (pretty printing terms in Cantor normal form)}
-
-\input{movies/snippets/T1/ppT1Def}
-
-The function \texttt{pp: T1 -> ppT1} converts any closed term of type \texttt{T1} into a human-readable expression. For instance, let us convert the term \texttt{alpha\_0}.
-
-\input{movies/snippets/T1/ppAlpha0}
-
-\paragraph*{\gaiasign}
-In ~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge},
-we define a variant of \texttt{pp} for pretty-printing terms of type \texttt{ssete9.CantorOrdinal.T1} (see Sect.~\vref{sect:gaia-ppT1}).
-
-\index{hydras}{Projects}
-\begin{project}
-Design  (in \ocaml?) a set of tools for systematically pretty printing ordinal terms in Cantor normal form.
-\end{project}
-
-\subsection{Comparison between ordinal terms}
-\label{sec:orgheadline73}
-
-% Our formalisation of Cantor Normal Form will take two steps:
-% 1 Definition of a strict order \texttt{o<} on the type \texttt{T1}, 
-% 2 Using \texttt{o<} for characterizing terms in normal form.
-
-In order to compare two terms of type \texttt{T1}, we define a recursive function \texttt{compare} that maps two ordinal tems $\alpha$ and $\beta$ to a value of type \texttt{comparison}.
-This type is defined in \coq's standard library
-\texttt{Init.Datatypes} and
-contains three constructors:  \texttt{Lt} (less than), \texttt{Eq} (equal), and
-\texttt{Gt} (greater than).
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/compareDef}
-
-\label{Predicates:lt-T1}
-Please note that this definition of \texttt{lt} makes it easy to write proofs by computation, as shown by the following examples.
-
-\vspace{4pt}
-
-\input{movies/snippets/T1/ltExamples}
-
-Links between the function \texttt{compare} and the relations
-\texttt{lt} and \texttt{eq} are established through the following lemmas (~\vref{sect:comparable-def}).
-\vspace{4pt}
-
-\input{movies/snippets/T1/Instances}
-
-\paragraph*{\gaiasign}
-In \gaia, the strict order \texttt{T1lt} on ordinal terms is directly defined as a boolean function (in \ssreflect's style).
-
-\inputsnippets{nfwfgaia/T1ltDef}
-
-In section~\vref{sect:lt-compat-gaia}, we show that both definitions are mutually equivalent.
-
-\subsubsection{A Predicate for Characterizing Normal Forms}
-\label{sect:t1-nf}
-
-\label{sec:orgheadline74}
-\label{sec:orgheadline75}
-Our data-type \texttt{T1} allows us to write expressions that
-are not properly in Cantor normal form as specified in Section \ref{sec:epsilon0-intro}.
-For instance, consider the following term of type  \texttt{T1}.
-
-\input{movies/snippets/T1/badTerm}
-
-This term would have been written \(\omega^1\times 2 + \omega^\omega \times 3\) in the usual mathematical notation.
-We note that the exponents of $\omega$ are not in the right (strictly decreasing) order.
-Nevertheless, with the help of the order \texttt{lt} on \texttt{T1}, we are now able to characterize
-the set of all well-formed ordinal terms:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#nf_b}{Epsilon0.T1}}
-
-\label{Predicates:nf-T1}
-
-\input{movies/snippets/T1/nfDef}
-
-\input{movies/snippets/T1/nfAlpha0}
-
-\input{movies/snippets/T1/nfBadTerm}
-
-\paragraph*{\gaiasign}
-In \gaia, the boolean function which characterizes ordinal terms in normal form is defined as follows:
-
-\inputsnippets{nfwfgaia/T1nfDef}
-
-In Sect.~\vref{nf-gaia-compat}, we show that \gaia's \texttt{T1nf} is extensionally equivalent with \HydrasLib'  \texttt{nf}.
-
-\subsection{Making normality implicit}
-  We would like to get rid of terms of type \texttt{T1} which are not in Cantor normal form.
-A simple way to do this is to consider statements of the form 
-\texttt{forall alpha: T1, nf alpha -> $P$ alpha}, where $P$ is a predicate over type \texttt{T1}, like in the following lemma \footnote{Ordinal addition is formally defined a little later (page~\ref{sect:infix-plus-T1})}.
-
-\input{movies/snippets/T1/plusIsZero}
-
-\vspace{4pt}
-
-But this style leads to clumsy statements, and generates too many subgoals in interactive proofs (although often solved with \texttt{auto} or \texttt{eauto}).
-
-One may encapsulate conditions of the form \texttt{(nf $\alpha$)} in
-the most used predicates. For instance, we introduce the restriction of \texttt{lt} to terms in normal form, and provide a handy notation for this restriction.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Prelude.Restriction.html}{hydras.Prelude.Restriction}}
-
-\input{movies/snippets/Restriction/restrictionDef}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#LT}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/LTDef}
-
-\label{Predicates:LT-T1}
-
-For instance, in the following lemma, the condition that $\alpha$ is in normal form is included in the condition $\alpha< 1$.
-
-\input{movies/snippets/T1/LTOne}
-
-\subsubsection{\texttt{E0}: a sigma-type for \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-As we noticed in Sect.~\ref{sect:t1-nf}, the type \texttt{T1} is not a correct ordinal notation, since it contains terms that are not in Cantor normal form.
-In certain contexts (for instance in Sections~\ref{sect:L-equations}, \ref{sect:hardy},
-and \ref{sect:wainer}),  we need to define total recursive functions on well-formed ordinal terms less  than $\epsilon_0$, using the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777}.
- In order to define a type whose inhabitants represent just ordinals, we build a type gathering a term of type \texttt{T1} and a proof that this term is in normal form.
-
-\label{sect:E0-def}
-\label{types:E0}
-\index{hydras}{Library Epsilon0!Types!E0}
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html\#E0}{Epsilon0.E0}}
-
-\input{movies/snippets/E0/E0Def}
-
-Many constructs : types, predicates, functions, notations, etc., on type \texttt{T1} are adapted to \texttt{E0}.
-
-First, we declare a notation scope for \texttt{E0}, then we redefine the predicates of comparison.
-
-\input{movies/snippets/E0/E0Scope}
-\label{Predicates:Lt-E0}
-\input{movies/snippets/E0/LtLeDef}
-
-Equality in \texttt{E0} is just Leibniz equality. Note that, since \texttt{nf} is
-defined by a Boolean function, for  any term $\alpha:\texttt{T1}$, there exists at most one proof of \texttt{nf $\alpha$}, thus two ordinals of type \texttt{E0} are
-equal if and only if their projection to \texttt{T1} are equal (see also Sect.~\vref{sect:eq-proof-unicity}).
-
-\vspace{4pt}
-
-\index{coq}{Unicity of equality proofs}
-
-\inputsnippets{E0/nfProofUnicity}
-\inputsnippets{E0/E0EqIff}
-
-In order to  upgrade constants and functions from type \texttt{T1} to \texttt{E0}, we have to prove that
-the term they build is in normal form.
-For instance, let us represent the ordinals $0$ and $\omega$   as instances of the class \texttt{E0}.
-
-\vspace{4pt}
-
-\label{sect:omega-T1}
-
-\input{movies/snippets/E0/ZeroOmega}
-
-%\index{Constants!zero:T1}
-
-\paragraph*{\gaiasign}
- The library~\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines a \gaia-compatible type \texttt{E0}.
-
-\subsection{Syntactic definition of limit and successor ordinals}
-
-Pattern matching and structural recursion allow us to define boolean characterizations  of successor and limit ordinals.
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/succbLimitb}
-
-The correctness of these definitions with respect to the mathematical notions of
-limit and successor ordinals is established through several lemmas. For instance,
-Lemma \texttt{canonS\_limit\_lub}, page~\pageref{lemma:canonS-limit}, shows that
-if $\alpha$ is (syntactically) a limit ordinal, then it is the least upper bound of
-a strictly increasing sequence of ordinals.
-
-\paragraph*{\gaiasign}
-In \gaia, the boolean functions associated with limit and successor ordinals are called \texttt{T1limit} and \texttt{T1is\_succ}.
-
- From~\href{../theories/html/gaia_hydras.HydraGaia_Examples.html}{gaia\_hydras.HydraGaia\_Examples}.
-
-\inputsnippets{HydraGaia_Examples/Demob}
-
-\subsection{Arithmetic on \texorpdfstring{$\epsilon_0$}{epsilon0}}
-\subsubsection{Successor}
-
-\index{hydras}{Library Epsilon0!Functions!succ}
-
-The successor of any ordinal $\alpha< \epsilon_0$ is defined by structural 
-recursion on its Cantor normal form.
-
-\label{Functions:succ-T1}
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}
-
-\input{movies/snippets/T1/succDef}
-
-The following lemma establishes the connection between the  function
-\texttt{succ} and the Boolean predicate \texttt{succb}.
-
-\vspace{4pt}
-
-\input{movies/snippets/T1/succbIff}
-
-\begin{exercise}[\gaiasign]
-  Look for the \gaia-theorem which corresponds to \texttt{succb\_iff}.
-\end{exercise}
-
-\subsubsection{Successor function on \texttt{E0}}
-
-The function \texttt{succ} on \texttt{T1} is extended to \texttt{E0} the following way:
-
- \emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
-
- \inputsnippets{E0/SuccOnE0}
-
-\index{hydras}{Exercises}
- \begin{exercise}
-Prove in \coq{} that for any ordinal $0< \alpha<\epsilon_0$, $\alpha$ is a limit if 
-and only if for all $\beta<\alpha$, the interval $[\beta,\alpha)$ is
-infinite.
-
-\emph{You may start this exercise with the file
-     \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/Limit_Infinity.v}{exercises/ordinals/Limit\_Infinity.v}.}
- \end{exercise}
-
-\subsubsection{Addition and multiplication}
-
-Ordinal addition and multiplication are also defined by structural recursion over the type \texttt{T1}. Please note that they use the \texttt{compare} function on some subterms of their arguments.
-
-\label{sect:infix-plus-T1}
-
-\input{movies/snippets/T1/plusDef}
-
-\input{movies/snippets/T1/multDef}
-
-\paragraph*{\gaiasign}
-We keep \gaia's base names for
-addition and multiplication of ordinal terms below $\epsilon_0$.
-Please refer to Sect.~\ref{sect:plus-mult-gaia-hydras} about compatibility of both arithmetics.
-
-\subsubsection{Examples}
-
-The following examples are instances of \emph{proofs by computation}. Please note that  addition and multiplication on \texttt{T1}
-are not commutative. Moreover,  both operations fail to be strictly monotonous in their first argument.
-
-\input{movies/snippets/T1/plusMultExamples}
-
-\input{movies/snippets/T1/notMono}
-
-The function \texttt{succ} is related with addition through the following lemma:
-
-\inputsnippets{T1/succIsPlusOne, T1/succIsPlusOnez}
-
-\subsubsection{Arithmetic on type \texttt{E0}}
-
- We define an addition in type \texttt{E0}, since the sum of two terms in normal form is in normal form too.
-
-\input{movies/snippets/T1/plusNf}
-
-\input{movies/snippets/E0/plusE0}
-\input{movies/snippets/E0/CheckPlus}
-
-\begin{remark}
-In all this development, two representations of ordinals co-exist: ordinal terms (type \texttt{T1}, notation scope \texttt{t1\_scope}, for reasoning on the tree-structure of Cantor normal forms), and ordinal terms \emph{known to be in normal form} (type \texttt{E0}, notation scope \texttt{E0\_scope}).
-Looking at the contexts displayed by \coq{} prevents you from any risk of confusion.
-\end{remark}
-
-%% To simplify !
-\index{hydras}{Exercises} 
-\begin{exercise}
-Prove that for any ordinal $\alpha:\texttt{E0}$, 
-$\omega\leq \alpha$ if and only if, for any natural number $i$,
-$i+\alpha=\alpha$.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/ge_omega_iff.v}{exercises/ordinals/ge\_omega\_iff.v}.}
-\end{exercise}
-
-\subsection{A proof by computation}
-\label{Ex42-E0}
-
-It is interesting to compare the following proof of the equality
-$\omega+42+\omega^2$ with the more theoretical proof in Sect~\vref{Ex42-schutte}.
-
-\inputsnippets{E0/Ex42}
-
-\section{Well-foundedness and transfinite induction}
-
-\index{maths}{Transfinite induction}
-
-\subsection{About  well-foundedness}
-\label{sec:T1Wf}
-\label{sec:orgheadline82}
-   In order to use \texttt{T1} for proving termination results,
-we need to prove that  our order \texttt{<} is well-founded. Then we will get \emph{transfinite induction} for free.
-
-The proof of well-foundedness of the strict order $<$ on Cantor normal forms is already 
-available in the Cantor contribution by Castéran and Contejean~\cite{CantorContrib}. That proof relies on a library on recursive path orderings written by
-E. Contejean. We present here  a direct proof of the same result, which does not require any knowledge on r.p.o.s.
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that the \emph{total} order \texttt{lt} on \texttt{T1} is not well-founded. 
-\textbf{Hint:}  You will have to build a counter-example with terms of type \texttt{T1}
-which are not in Cantor normal form.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/T1_ltNotWf.v}{exercises/ordinals/T1\_ltNotWf.v}.}
-\end{exercise}
-
-\subsubsection{A first attempt}
-\label{sec:orgheadline77}
-%\index{coq}{Well-founded induction}
-
-It is natural to try to prove by structural induction over \texttt{T1} 
-that every term in normal form is \texttt{LT}-accessible.
-
-Unfortunately, it won't work. Let us consider some well-formed term
- $\alpha=\texttt{cons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are \texttt{LT}-accessible.
-In order to prove the accessibility of $\alpha$, we have to consider
-any well formed term \(\delta\) such that \(\delta<\alpha\). 
-But nothing guarantees that \(\delta\)  is strictly  less than \(\beta\) nor \(\gamma\), and we cannot use the induction hypotheses on   \(\beta\) nor \(\gamma\).
-
-\input{movies/snippets/T1/wfLTBada}
-
-The problem comes from the too weak hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
-\(\gamma\);
-for instance \(\delta\) may be of the form
-\texttt{cons $\beta$ $p'$  $\gamma'$},
-where    \(p' < n\).
-Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.
-
-\input{movies/snippets/T1/wfLTBadz}
-
-\subsubsection{Using a stronger inductive predicate.}
-\label{sec:orgheadline78}
-\label{sec:strongly-accessible}
-
-  Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
-through \texttt{LT}, we propose to consider the following (stronger) predicate:
-
-\input{movies/snippets/T1/AccStrongDef}
-
-The following lemma is an application of the strict inequality 
-$\alpha < \omega ^\alpha$. If $\alpha$ is strongly accessible, then, by definition,
-$\omega^\alpha$ is accessible, thus $\alpha$ is \emph{a fortiori} accessible.
-
-\input{movies/snippets/T1/AccStrongStronger}
-
-Thus, it remains to prove that every ordinal strictly less than $\epsilon_0$
-is strongly accessible.
-
-% \subsubsection{Structure of the proof of well-foundedness of \texttt{LT}}
-
-\label{sec:orgheadline81}
-\label{proof-wf-epsilon0}
-\paragraph{A helper}
-\label{sec:orgheadline79}
-
-First, we prove that, for  any \texttt{LT}-accessible term $\alpha$, $\alpha$ is 
-strongly accessible too (\emph{i.e.} any well formed
-term (\texttt{cons $\alpha$ $n$ $\beta$})  is accessible).
-
-The proof is structured as an induction on $\alpha'$s accessibility. Let us consider
-any  accessible term $\alpha$.
-
-\input{movies/snippets/T1/AccImpAccStrong}
-
-Let \texttt{n:nat} and \texttt{beta:T1} such that (\texttt{cons alpha n beta}) is in normal form. 
-We prove first that \texttt{beta} is accessible,  which allows us to prove by well-founded induction on \texttt{beta}, 
-and natural induction on \texttt{n}, that (\texttt{cons alpha n beta}) is accessible.
-The proof, quite long, can be consulted in \href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}.
-
-\paragraph{Accessibility of any well-formed ordinal term}
-\label{sec:orgheadline80}
-
-Our goal is still to prove accessibility of any well formed ordinal term.
-Thanks to our previous lemmas, we are almost done (by a simple structural induction!).
-
-\input{movies/snippets/T1/nfAcc}
-
-\input{movies/snippets/T1/T1Wf}
-
-\index{maths}{Transfinite induction}
-
-Traditionnally, well-founded induction on ordinals is called \emph{transfinite induction}.
-
-\inputsnippets{T1/transfiniteRecursor, T1/transfiniteInduction}
-
-As a corollary, the order \texttt{Lt} on type \texttt{E0} is well-founded too.
-
-\inputsnippets{E0/E0LtWf}
-
-\begin{remark}
-\label{remark:a3pat}
-The alternate proof of well-foundedness using \'Evelyne Contejean's work on
-    recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the
-    library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0.Epsilon0rpo}.
- \end{remark}
-
-\subsection{An ordinal notation for  \texorpdfstring{$\epsilon_0$}{epsilon0}}
-
-We build an instance of \texttt{ON}, and prove its correctness w.r.t. Schutte's model (see Chapter~\ref{chap:schutte}).
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
-
-\input{movies/snippets/E0/InstanceEpsilon0}
-
-\label{instance-epsilon0}
-
-\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
-
-\input{movies/snippets/Correctness_E0/injectDef}
-
-\input{movies/snippets/Correctness_E0/Epsilon0Correct}
-
-\index{hydras}{Projects}
-\begin{project}
- \emph{This exercise is a continuation of Project~\vref{exo:ON-mult}.}
-Use \texttt{ON\_mult} to define an ordinal notation \texttt{Omega2} for $\omega^2=\omega\times\omega$.
-
-Prove that \texttt{Omega2} is a sub-notation of \texttt{Epsilon0}.
-
-Define on \texttt{Omega2} an addition compatible with the addition on \texttt{Epsilon0}.
-
-\textbf{Hint}. You may use the following definition (in 
-    \href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{OrdinalNotations.ON\_Generic}).
-
-\input{movies/snippets/ON_Generic/SubONSameOp}
-
-     \end{project}
-
-    \index{hydras}{Projects}
-    \begin{project}
-    The class \texttt{ON} of ordinal notations has been defined long after this 
-    chapter, and is not used in the development of the type \texttt{E0} yet.
-    A better integration of both notions should simplify the development on ordinals in Cantor normal form. This integration is planned for the future versions.
-
-    \end{project}
-
-    \section{A refinement of \texttt{E0} : an ordinal notation for \texorpdfstring{$\omega^\omega$}{omega\^omega}}
-
-    In Module   \href{https://github.com/coq-community/hydra-battles/blob/master/theories/
-      ordinals/OrdinalNotations/OmegaOmega.v}{theories/ordinals/OrdinalNotations/OmegaOmega.v},
-    we represent ordinals below $\omega^\omega$ by lists of pairs of natural numbers (with the same coefficient shift as in \texttt{T1}).
-    For instance, the ordinal $\omega^4\times 10 + \omega^3 + \omega+ 5$ is represented by the list \texttt{(4,9)::(3,0)::(1,0)::(0,4)::nil}.
-
-\inputsnippets{OmegaOmega/LODef}
-
-  The usual operations : \texttt{succ}, \texttt{+}, \texttt{*} are simple variants of the same operations in \texttt{T1}.
-
-  \inputsnippets{OmegaOmega/succPlusMult}
-
-    We establish this representation as a \emph{refinement} of the data types we used to represent ordinals less than $\epsilon_0$.
-Thus, many properties like well-foundedness of $<$ and associativity of $+$,  of this ordinal notations have very short proofs.
-
-    \inputsnippets{OmegaOmega/refineDef, OmegaOmega/phi0Ref,
-      OmegaOmega/succRef, OmegaOmega/plusRef,
-      OmegaOmega/multRef}
-
-    In order to make an ordinal notation for $\omega^\omega$, we follow the same steps as for $\epsilon_0$:
-    \begin{enumerate}
-    \item Define an order \texttt{lt}, which refines the order
-      \texttt{lt} on \texttt{T1}.
-      \inputsnippets{OmegaOmega/compareDef}
-      \inputsnippets{ OmegaOmega/ltDef, OmegaOmega/ltRef}
-
-    \item Define the predicate ``to be in normal form''.
-      \inputsnippets{OmegaOmega/nfDef}
-    \item Define a class \texttt{OO} of terms in normal form, and an embedding from \texttt{E0} into \texttt{OO}.
-      \inputsnippets{OmegaOmega/OODef}
-       \inputsnippets{OmegaOmega/embedDef}
-      \item Infer well-foundedness of the order on \texttt{OO}.
-    \inputsnippets{OmegaOmega/ltWf}
-
-    \end{enumerate}
-
-    \inputsnippets{OmegaOmega/ONOO, OmegaOmega/OOz}
-
-    Let us show a few examples.
-
-    \inputsnippets{OmegaOmega/OODemo}
-
-    \index{hydras}{Exercises}
-
-    \subsection{Related work}
-    The article~\cite{Manoury2016}  defines another representation of ordinals below $\omega^\omega$ based on lists of natural numbers.
-
-\begin{exercise}[not trivial]
-  It may be interesting to write a \emph{direct} proof  of well-foundedness of the order in $\omega^\omega$ (\emph{i.e.}
-  without using properties of $\epsilon_0$). This exercise may
-  help to understand better the proof structure of Sect.~\vref{sec:strongly-accessible}.
-    \end{exercise}
-%ici
-    \section{A variant for hydra battles}
-
-    In order to prove the termination of any hydra battle, we try to define a variant mapping hydras to ordinals strictly less than $\epsilon_0$.
-    In order to make such a variant easy to define (for instance by a structural recursion), we introduce a variant of addition, which, contrary to
-    $+$, is commutative and strictly monotonous in both of its arguments. This last property makes it possible to prove that our function is 
-    truly a variant for hydra battles (in Sect.~\vref{sect:variant-decr}).
-
-    \subsection{Natural sum (a.k.a. Hessenberg's  sum)}
-    \label{sec:orgheadline87}
-    \label{hydra-variant}
-    \label{sect:hessenberg-def}
-
-    Natural sum (Hessenberg's  sum) is a commutative and monotonous version of
-    addition. It is used as an auxiliary operation  for defining variants
-    for hydra battles, where Hercules is allowed to chop off any  head of the hydra.
-
-    In the literature, the natural sum of ordinals \(\alpha\) and \(\beta\)
-    is often denoted by \(\alpha \# \beta\)  or  \(\alpha \oplus  \beta\).
-    Thus we called \texttt{oplus} the associated \emph{Coq} function.
-
-    \subsubsection{Definition of \texttt{oplus}}
-    \label{sec:orgheadline84}
-    %\index{Functions!oplus @ oplus (Hessenberg commutative sum)}
-
-    The definition of \texttt{oplus} is recursive in both of its 
-    arguments and uses  the same pattern as for the \texttt{merge} function on lists of library
-    \texttt{Coq.Sorting.Mergesort}.
-
-    \begin{enumerate}
-    \item Define a nested recursive function, using the \texttt{Fix} 
-        construct
-
-    \item Build a principle of induction dedicated to \texttt{oplus}
-
-    \item Establish equations associated to each case of the definition.
-    \end{enumerate}
-
-    \paragraph{Nested recursive definition}
-    \label{sec:orgheadline83}
-
-    The following definition is composed of 
-    \begin{itemize}
-    \item A main function \texttt{oplus}, structurally recursive in its 
-    first argument \texttt{alpha}
-    \item An auxiliary function \texttt{oplus\_aux} within the scope of \texttt{alpha},
-    structurally recursive in its argument \texttt{beta};  \texttt{oplus\_aux beta} 
-       is supposed to compute  \texttt{oplus alpha beta}.
-    \end{itemize}
-
-%  \pagebreak 
-  \vspace{4pt}
-    \emph{From Module~\href{../theories/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}
-
-    \label{sect:infix-oplus}
-
-    \input{movies/snippets/Hessenberg/oplusDef}
-
-    The reader will note that each recursive call of the functions
-    \texttt{oplus} and \texttt{oplus\_aux} satisfies \emph{Coq}'s constraint
-    on recursive definitions. The function \texttt{oplus} is recursively called on a sub-term of its first argument,
-    and \texttt{oplus\_aux} on a sub-term of its unique argument.
-    Thus, \texttt{oplus}'s definition is accepted by \coq{} as a structurally recursive function.
-
-    \subsubsection{Rewriting lemmas}
-    \label{sec:orgheadline86}
-
-    \emph{Coq}'s constraints on recursive definitions result in 
-    the quite  complex form of \texttt{oplus}'s definition.
-    Proofs of properties of this function can be simpler if we
-     derive a few  rewriting lemmas that will help to simplify 
-    expressions of the form (\texttt{oplus $\alpha$ $\beta$}).
-
-    A first set of lemmas correspond to the various cases of \texttt{oplus}'s 
-    definition. They can be proved almost immediately. Here are a few examples.
-
-    \inputsnippets{Hessenberg/oplusNeutral,
-      Hessenberg/oplusCompareLt}
-
-    % \subsubsection{A hand-made induction principle}
-    % \label{sec:orgheadline85}
-
-    % \index{Coq!Commands!Functional Scheme}
-
-    % \emph{Coq} contains a command  \texttt{Functional Scheme} that 
-    % generates induction principles which correspond to recursive functions.
-    % Unfortunately, the current version ( \texttt{8.11.0} ) doesn't work on \texttt{oplus},
-    % probably because of the inner \texttt{Fix}.
-
-    % \begin{Coqsrc}
-    % Functional Scheme oplus_ind := Induction for oplus Sort Prop.
-    % \end{Coqsrc}
-
-    % \begin{Coqanswer}
-    % Error: Anomaly "todo." Please report at http://coq.inria.fr/bugs/.
-    % \end{Coqanswer}
-
-    % Fortunately, it's a good exercise for a semi-experienced user, to write
-    % her/him-self induction principles similar to the ones returned by
-    % \texttt{Functional Scheme}.
-
-    % \begin{itemize}
-    % \item First, we choose to write a version for sort \texttt{Type}, since versions
-    % for sorts \texttt{Prop} and \texttt{Set} can be easily derived from
-    % the former one. According to \emph{Coq}'s naming politics, we will call our 
-    % principle \texttt{oplus\_rect}
-
-    % \item The conclusion of \texttt{oplus\_rect} will be (\texttt{$P$ a b (oplus a b)}),
-    % where $P$ is an arbitrary function of type 
-    % \texttt{T1 -> T1 -> T1 -> Type}
-
-    % \item The premises of \texttt{oplus\_rect} will describe how to build an induction 
-    % on the graph of \texttt{oplus}.
-    % \end{itemize}
-
-    % We are now ready to state and prove \texttt{oplus\_rect}, and the reader
-    % will note that the statement is longer than the proof script itself,
-    % which is a standard proof by induction, simplification and case-analysis 
-    % that follows  \texttt{oplus}'s definition.
-
-    % We associate also a tactic to the application of \texttt{oplus\_rect}.
-
-    % \begin{Coqsrc}
-    %  Lemma oplus_rect:
-    %       forall P: T1 -> T1 -> T1 -> Type, 
-    %         (forall a:T1, P zero a a) ->
-    %         (forall a: T1, P a zero a) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Gt ->
-    %            P b1 (cons a2 n2 b2) o ->
-    %            P (cons a1 n1 b1) (cons a2 n2 b2)
-    %              (cons a1 n1 o)) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Lt ->
-    %            P (cons a1 n1 b1) b2 o ->
-    %            P (cons a1 n1 b1) (cons a2 n2 b2) 
-    %            (cons a2 n2 o)) ->
-    %         (forall a1 n1 b1 a2 n2 b2 o,
-    %            compare a1 a2 = Eq ->
-    %            P b1 b2 o ->
-    %           P (cons a1 n1 b1) (cons a2 n2 b2)
-    %             (cons a1 (S (n1 + n2)%nat) o)) ->
-    %          forall a b, P a b (oplus a b).
-    % Proof with auto.
-    %    induction a.
-    %    -    intro; simpl; destruct b;auto.
-    %    -   induction b.
-    %        + apply X0.
-    %        + case_eq (compare a1 b1).
-    %          * intro Comp; unfold oplus; rewrite Comp.
-    %            cbn; apply X3 ...
-    %          * intro Comp; cbn; rewrite Comp; apply X2...
-    %          * intro Comp; cbn; rewrite Comp ...
-    %  Defined.
-
-    % Ltac oplus_induction a b:= pattern (oplus a b); apply oplus_rect.
-    % \end{Coqsrc}
-
-    % \index{Exercises}
-
-    % \begin{exercise}
-    % The induction principle \texttt{oplus\_rect} is still unused in our development. 
-    % Please build some nice examples of application.
-    % \end{exercise}
-
-    \index{hydras}{Projects}
-    \begin{project}
-    Compare \texttt{oplus}'s definition (with inner fixpoint) with other possibilities
-    (\texttt{coq-equations}, \texttt{Function}, etc.).
-    \end{project}
-    \subsection{More theorems on Hessenberg's sum}
-
-    We need to prove some properties of $\oplus$, particularly about 
-    its relation with the order $<$ on \texttt{T1}.
-
-    \subsubsection{Commutativity, associativity}
-
-    We prove  the commutativity of $\oplus$ in two steps. 
-    First, we prove by transfinite induction on $\alpha$ that the restriction of $\oplus$ to the
-    interval $[0..\alpha)$ is commutative.
-
-    \index{maths}{Transfinite induction}
-
-    \input{movies/snippets/Hessenberg/oplusComm0}
-
-    Then, we infer  $\oplus$'s commutativity for any pair of ordinals:
-    Let $\alpha$ and $\beta$ be two ordinals strictly less than $\epsilon_0$. Both ordinals $\alpha$ and $\beta$ are
-    strictly less than $\textrm{max}(\alpha,\beta)+1$.
-        Thus, we have just to apply the lemma \coqsimple{oplus\_comm\_0}.
-
-\input{movies/snippets/Hessenberg/oplusComm}
-
-    Associativity of Hessenberg's sum is proved the same way.
-
-    \input{movies/snippets/Hessenberg/oplusAssoc0}
-    \input{movies/snippets/Hessenberg/oplusAssoc}
-
-    \subsubsection{Monotony}
-
-    At last, we prove that $\oplus$ is strictly monotonous in both of its arguments.
-
-    \input{movies/snippets/Hessenberg/oplusMono}
-
-      \index{hydras}{Projects}
-
-    \begin{project}
-    The library \texttt{Hessenberg} looks too long (proof scripts and compilation).
-    Please try to make it simpler and more efficient!
-    Thanks!
-    \end{project}
-
-    \paragraph*{\gaiasign} The module
-      \href{../theories/html/gaia_hydras.~GHessenberg.html}{gaia\_hydras.GHessenberg} defines a version of Hessenberg's sum
-      compatible with \gaia's type \texttt{T1}.
-
-    \subsection{A termination measure for hydra battles }
-
-    \label{sec:hydra-measure}
-
-    Let us define a measure from type \texttt{Hydra} into \texttt{T1}.
-
-    \vspace{4pt}
-    \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Termination.html\#m}{Hydra.Hydra\_Termination}}
-
-   \input{movies/snippets/Hydra_Termination/mDef}
-
-    First, we prove that the measure $m(h)$  of any hydra $h$ is a well-formed ordinal term of type \texttt{T1}.
-
-    \input{movies/snippets/Hydra_Termination/mNf}
-      \input{movies/snippets/Hydra_Termination/msNf}
-
-    For proving the termination of all hydra battles, we have to prove that
-    \texttt{m} is a variant. First, a few technical lemmas follow the decomposition of \texttt{round} into several relations. Then the lemma \texttt{round\_decr} gathers all the cases.
-
-    \label{sect:variant-decr}
-
-    \input{movies/snippets/Hydra_Termination/S0Decr}
-    \input{movies/snippets/Hydra_Termination/R1Decr}
-    \input{movies/snippets/Hydra_Termination/S1Decr}
-    \input{movies/snippets/Hydra_Termination/R2Decr}
-     \input{movies/snippets/Hydra_Termination/RoundDecr}
-
-      Finally, we prove termination of all (free) battles.
-
-    \label{thm:every-battle-terminates}
-
-    \input{movies/snippets/Hydra_Termination/FinalThm}
-
-    \paragraph*{\gaiasign}
-    The module~\href{../theories/html/gaia_hydras.GHydra.html}{gaia\_hydras.GHydra} contains a proof of \gaiaHydras version of termination of all battles.
-
-    \inputsnippets{GHydra/mDef, GHydra/mVariant, GHydra/Termination}
-
-    \section*{Conclusion}
-
-    Let us recall three results we have proved so far.
-    \begin{itemize}
-    \item There exists a strictly decreasing variant which maps \texttt{Hydra} into 
-    the segment $[0,\epsilon_0)$ for proving the termination of any hydra battle
-    \item There exists \emph{no} such variant from \texttt{Hydra} into 
-    $[0,\omega^2)$, \emph{a fortiori} into $[0,\omega)$.
-    \end{itemize}
-
-    So, a  natural question is `` Does there exist any strictly decreasing variant mapping
-    type \texttt{Hydra} into some interval $[0,\alpha)$ (where $\alpha <\epsilon_0$) for proving the termination of all hydra battles''.
-The next chapter is dedicated to a formal proof that there exists no such $\alpha$, even if we consider a restriction to the set of ``standard'' battles.
-
-    %\include{epsilon0}
-
-    %\include{impossibility-proofs}
-
-%-------------------------------------------------------------------
-
-\chapter[The Ketonen-Solovay machinery]{Accessibility inside \texorpdfstring{$\epsilon_0$}{Epsilon0}: The Ketonen-Solovay Machinery\label{ks-chapter}}
-\label{chap:ketonen}
-\index{maths}{Ordinal numbers!Ketonen-Solovay machinery}
-
-\section{Introduction}
-The reader may think that our proof of termination in the previous  chapter requires a lot of mathematical tools and may be too  complex. So, the question is ``is there  any  simpler proof'' ?
-
-In their article~\cite{KP82}, Kirby and Paris show that this result cannot be proved in Peano arithmetic.
-Their proof uses some knowledge about model theory and non-standard models of Peano arithmetic.
-In this chapter, we focus on a specific class of proofs of termination of hydra battles: construction of some variant mapping the type \texttt{Hydra} into a given initial  segment of ordinals.
-Our proof relies only on the Calculus of Inductive Constructions and is a natural complement of the results proven in the previous chapters.
-
-\begin{itemize}
-\item There is no variant mapping the type \texttt{Hydra} into the interval $[0,\omega^2)$ (section ~\vref{omega2-case}), and \emph{a fortiori} $[0,\omega)$ (section ~\vref{omega-case}).
-
-\item There exists a variant which maps the type \texttt{Hydra} into the
-interval $[0,\epsilon_0)$ (theorem \texttt{every\_battle\_terminates}, in section~\vref{thm:every-battle-terminates}).
-\end{itemize}
-
-Thus, a very natural question is the following one:
-\begin{quote}
-  `` Is there  any variant from
-\texttt{Hydra} into some interval $[0,\mu)$, where $\mu<\epsilon_0$, for proving the termination of all hydra battles ?''
-\end{quote}
-
-We prove in \coq{} the following result:
-
-\begin{quote}
-There is no variant for proving the termination of all hydra battles
-from \texttt{Hydra} into the interval $[0..\mu)$, where
-$\mu< \epsilon_0$.
-The same impossibility holds even if we consider only standard battles (with the successive replication factors $0,1,2,\dots,t,t+1,\dots$).
-\end{quote}
-
-Our proofs are  constructive and require no axioms: they are  closed terms of the CIC, and are mainly composed on function definitions and proofs of properties of these functions. 
-They  share much theoretical material with Kirby and Paris', although they do not use any knowledge about Peano arithmetic nor model  theory.  The combinatorial arguments we use and implement
-come from 
- an article by J.~Ketonen and R.~Solovay~\cite{KS81}, already  cited in the work
- by L.~Kirby and J.~Paris. % on the termination of Goodstein sequences and hydra battles~\cite{KP82}.
- Section $2$ of this article: ''A hierarchy of probably recursive functions'', contains a systematic study of \emph{canonical sequences}, which are closely related to
-rounds of hydra battles. 
-Nevertheless, they have the same global structure as the simple proofs described in
-sections~\vref{omega-case} and \vref{omega2-case}. 
-We invite the reader to compare the three proofs step by step, lemma by lemma.
-
-\section{Canonical Sequences}
-\label{ketonen-solovay-sect}
-\index{maths}{Ordinal numbers!Canonical sequences}
-
-Canonical sequences are functions that associate an ordinal $\canonseq{\alpha}{i}$ to every ordinal $\alpha<\epsilon_0$ and positive integer $i$. They satisfy several nice properties:
-
-\index{maths}{Transfinite induction}
-\begin{itemize}
-\item If $\alpha\not=0$, then $\canonseq{\alpha}{i}<\alpha$. Thus canonical sequences can be used for proofs by transfinite induction or function definition by transfinite recursion
-\item If $\lambda$ is a limit ordinal, then $\lambda$ is the least upper bound of the set 
-$\{\canonseq{\lambda}{i}\;|\,i\in\mathbb{N}_1\}$
-
-\item If $\beta<\alpha<\epsilon_0$, then there is a ``path'' from $\alpha$ to $\beta$, \emph{i.e.} a
-sequence $\alpha_0=\alpha, \alpha_1, \dots, \alpha_n=\beta$, where for every $k<n$, there exists some $i_k$ such that $\alpha_{k+1}=\canonseq{\alpha_k}{i_k}$
-\item Canonical sequences correspond tightly to rounds of hydra battles: if $\alpha\not=0$,
-then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1})$ in one round with
-the replication factor $i$ (Lemma \href{../theories/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
-\item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) battle from $\iota(\alpha)$ to $\iota(\beta)$.
-\end{itemize}
-
-\begin{remark}
-  In~\cite{KS81}, canonical sequences are defined for any ordinal $\alpha <\epsilon_0$,
-by stating that if $\alpha$ is a successor ordinal $\beta+1$,  the sequence associated with 
-$\alpha$ is simply the constant sequence whose terms are equal to $\beta$.
-Likewise, the canonical sequence of $0$ maps any natural number to $0$.
-
-This convention allows us to make total the function that maps any ordinal $\alpha$ and natural number $i$ to the ordinal $\canonseq{\alpha}{i}$.
-
-\end{remark}
-
-First, let us recall how canonical sequences are defined in~\cite{KS81}.
-For efficiency's sake, we decided not to implement directly K.\&S's definitions, but to define in \gallina{} simply typed structurally recursive functions which share the abstract properties which are used in the mathematical proofs\footnote{With a small difference: the $0$-th term of the canonical sequence is not the same in our development as in~\cite{KS81}.}.
-
-\subsubsection{Mathematical definition of canonical sequences}
-
-In~\cite{KS81} the definition of $\canonseq{\alpha}{i}$ is based on the following remark:
-\begin{quote}
-Any non-zero ordinal $\alpha$ can be decomposed in a unique way as the product
-$\omega^\beta\times (\gamma+1)$.
-\end{quote}
-
-Thus the $\canonseq{\alpha}{i}$\,s are defined in terms of this decomposition:
-\begin{definition}[Canonical sequences: mathematical definition]
-\label{def:canonseq-math}
-
-\end{definition}
-\begin{mathframe}
-  \begin{itemize}
-\item Let $\lambda<\epsilon_0$ be a limit ordinal
-
-\begin{itemize}
-\item If $\lambda=\omega^{\alpha+1}\times (\beta+1)$, then 
-$\canonseq{\lambda}{i}= \omega^{\alpha+1}\times\beta +  \omega^\alpha \times i$
-\item If $\lambda=\omega^{\gamma}\times (\beta+1)$, where $\gamma<\lambda$ is a limit ordinal, then 
-$\canonseq{\lambda}{i}=\omega^{\gamma}\times \beta + \omega^{\canonseq{\gamma}{i}}$
-\end{itemize}
-
-\item For successor ordinals, we have $\canonseq{\alpha+1}{i}= \alpha$
-
-\item Finally, $\canonseq{0}{i}= 0$.
-\end{itemize}
-\end{mathframe}
-
-\subsubsection{Canonical sequences in \coq}
-\index{hydras}{Library Epsilon0!Functions!canon}
-\index{hydras}{Library Epsilon0!Functions!canonS}
-
-Our definition may look more complex than the mathematical one, but
-uses plain structural recursion over the type \coqsimple{T1}. Thus, tactics like
-\coqsimple{cbn}, \coqsimple{simpl}, \coqsimple{compute}, etc., are applicable.
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canon}{Epsilon0.Canon}}
-
-\label{Functions:canonS}
-\label{Functions:canon}
-
-\input{movies/snippets/Canon/canonDef}
-
-\paragraph*{\gaiasign} Our library~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon} transposes the  function \texttt{canon} into \gaia's type \texttt{T1}
-(please see Sect.\~ref{sect:gcanon}).
-
-\paragraph*{Remark}
-In the present state of this library, the following specializations of \texttt{canon} are still used in some proofs or lemma statements. They are planned to be deprecated.
-
-\input{movies/snippets/Canon/CanonS0}
-
-For instance \coq's computing facilities allow us to verify the equalities\linebreak 
-\mathcolor{$\canonseq{\omega^\omega}{3} = \omega^3$} and
-\mathcolor{$\canonseq{\omega^\omega*3}{42} = \omega^\omega*2 + \omega^{42}$}.
-
-\input{movies/snippets/Canon/canonExamples}
-
-% \index{hydras}{Projects}
-% \begin{project}
-% Many lemmas presented in this chapter were stated and proved before the introduction of 
-% the type class \texttt{ON} of ordinal notations, and in particular its  instance \texttt{Epsilon0}.
-% Thus definitions and lemmas refer to the type \texttt{T1} of possibly not well-formed terms.
-% This should be fixed in  a future version.
-% \end{project}
-
-\subsection{Basic properties of canonical sequences}
-
-We did not  try to prove that our definition truly implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
-abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
-~\href{../theories/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}.
-For instance, the equality $\canonseq{\alpha+1}{i}=\alpha$  can be  proved by  structural induction on $\alpha$.
-
-\input{movies/snippets/Canon/canonSucc}
-
-\subsubsection{Canonical sequences and the order $<$}
-
-\index{maths}{Transfinite induction}
-
-We prove by transfinite induction over $\alpha$ that $\canonseq{\alpha}{i}$ is an ordinal strictly less than $\alpha$ (assuming $\alpha\not=0$).
-This property allows us to use the function \texttt{canonS} and its derivatives in function definitions by transfinite recursion.
-
-\label{lemma:canon_LT}
-
-\inputsnippets{Canon/canonLT}
-
-\paragraph*{\gaiasign} This lemma is also available in
-Library ~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon}:
-
-\inputsnippets{GCanon/gcanonLt}
-
-\subsubsection{Limit ordinals are truly limits}
-The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
-is the limit of the sequence $\canonseq{\lambda}{i}\;(1\le i)$.
-
-\vspace{4pt}
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}
-
-\input{movies/snippets/Canon/canonSLimitStrong}
-
-\label{lemma:canonS-limit}
-
-Note the use of \coq's \texttt{sig} type in the theorem's statement, which
-relates the boolean function \texttt{limitb} defined on the \texttt{T1} data-type with a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.
-We can also state directly that $\lambda$ is a (strict) least upper bound of the elements of its canonical sequence.
-
-\input{movies/snippets/Canon/canonSLimitLub}
-
-\paragraph*{\gaiasign}  In \gaiaHydras, the statement use a slightly different vocabulary:
-\inputsnippets{GCanon/gcanonLimitOf}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}\label{exo:simply-typed-canonseq}
-Instead of using the \texttt{sig} type, define a simply typed function that, given two ordinals $\alpha$ and $\beta$, returns a natural number $i$ such that, if $\alpha$ is a limit ordinal and $\beta<\alpha$, then $\beta< \canonseq{\alpha}{i+1}$.
-Of course, you will have to prove the correctness of your function.
-
-\textbf{Hint:} You may add to your function a third argument usually called \texttt{fuel} for allowing you to give a structurally 
-recursive function (\emph{cf} the post of Guillaume Melquiond on Coq-club (Dec 21, 2020)
-\url{https://sympa.inria.fr/sympa/arc/coq-club/2020-12/msg00069.html}).
-The type \texttt{fuel}, an alternative 
-to \texttt{nat} is available on \href{../theories/html/hydras.Prelude.Fuel.html}{Prelude.Fuel})
-.
-%\index{coq}{Giving fuel to a long computation}
-
-\end{exercise}
-
-\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0}\,: paths}
-\index{maths}{Ordinal numbers!Accessibility inside epsilon0}
-\label{sect:pathes-intro}
-
-Let us consider a kind of accessibility problem inside $\epsilon_0$: given two ordinals $\alpha$ and $\beta$, where $\beta<\alpha<\epsilon_0$, find a \emph{path} consisting of a finite sequence $\gamma_0=\alpha,\dots,\gamma_l=\beta$,
-where, for every $i<l$, $\gamma_i \not= 0$ \footnote{This condition allows us to ignore paths which end by a lot of useless $0$s.} and there exists some strictly positive integer $s_i$
-such that $\gamma_{i+1}=\canonseq{\gamma}{s_i}$.
-
-Let $s$ be the sequence $\langle s_0,s_1,\dots, s_{l-1} \rangle$. We describe the
-existence of such a path with the notation $\alpha\xrightarrow [s]{}\beta$.
-We say also that the considered path from $\alpha$ to $\beta$ \emph{starts at [index] $s_0$ and ends at $s_l$}.
-
-For instance, we have $\omega*2 \xrightarrow[2,2,2,4,5]{}3$, through the 
-path $\langle\omega\times 2, \omega+2,\omega+1,\omega,4,3\rangle$.
-
-\begin{remark}
-
-Note that, given $\alpha$ and $\beta$, where $\beta < \alpha$, the sequence $s$ which leads from $\alpha$ to $\beta$ is not unique.
-Indeed, if $\alpha$ is a limit ordinal, the first element of $s$ can be any integer $i$ such that $\beta<\canonseq{\alpha}{i}$, and if $\alpha$ is a successor ordinal,
-then the sequence $s$ can start with any positive integer.
-
-For instance, we have also 
-$\omega*2 \xrightarrow[3,4,5,6]{}\omega$. 
-Likewise,
-$\omega*2 \xrightarrow[1,2,1,4]{} 0$ and
-$\omega*2 \xrightarrow[3,3,3,3,3,3,3,3]{} 0$.
-\end{remark}
-
-\subsection{Formal definition}
-
-\label{path-to-definition}
-
-In \coq{}, the notion of path can be simply defined as an inductive predicate 
-parameterized by the destination $\beta$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\index{hydras}{Library Epsilon0!Predicates!path\_to}
-\label{sect:path-to-def}
-
-\input{movies/snippets/Paths/transitionDefs}
-
-\input{movies/snippets/Paths/pathDef}
-
-\begin{remark}
-In the present version of our library, we use a variant \texttt{path\_toS} of
-\texttt{path\_to}, where the proposition
-(\texttt{path\_toS $\beta$ $s$ $\alpha$}) is equivalent to
-(\texttt{path\_to $\beta$ (List.map S $s$) $\alpha$}). This variant is scheduled to be deprecated.
-\end{remark}
-
-\paragraph*{\gaiasign}
-The library~\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} transposes the  notion of path into \gaia's type \texttt{T1}
-(please see Sect.\~ref{sect:gpath}).
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Write a tactic for solving goals of the form (\texttt{path\_to $\beta$ $s$ $\alpha$})
-where $\alpha$, $\beta$ and $s$ are closed terms. 
-You should solve automatically the following goals:
-
-\inputsnippets{Paths/exPathTac} 
-\end{exercise}
-
-\subsection{Existence of a path}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction on $\alpha$, we prove that for any $\beta<\alpha$, 
-one can build a path from $\alpha$ to $\beta$ (in other terms, $\beta$ is accessible from $\alpha$).
-
-\input{movies/snippets/Paths/LTPathTo}
-
-\noindent 
-From the lemma \texttt{canon\_LT}~\vref{lemma:canon_LT}, we can convert any path into an inequality on ordinals (by induction on paths).
-
-\input{movies/snippets/Paths/pathToLT}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}[continuation of exercise~\vref{exo:simply-typed-canonseq}]
-Define a simply typed function for computing a path from $\alpha$ to $\beta$.
-\end{exercise}
-
-\subsection{Paths and hydra battles}
-\label{KS-o2h}
-
-In order to apply our knowledge about  ordinal numbers less than $\epsilon_0$ to the study of hydra battles, we define an injection
-from the interval $[0,\epsilon_0)$ into the type \texttt{Hydra}.
-
-\vspace{4pt}
-
-\emph{From Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}}
-
-\input{movies/snippets/O2H/iotaDef}
-
-For instance Fig.~\ref{fig:iota-example} shows the image by $\iota$ of the ordinal  \textcolor{black}{$\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$}
-
-  \begin{figure}[htb]
-\centering
-\begin{tikzpicture}[very thick, scale=0.3]
-\node (foot) at (10,0) {$\bullet$};
-\node (N1) at (2,2) {$\bullet$};
-\node (N2) at (10,2) {$\bullet$};
-\node (N22) at (7,2) {$\bullet$};
-\node (N3) at (14,2) {$\bullet$};
-\node (N4) at (18,2) {$\Smiley[2][green]$};
-\node (N5) at (0,4) {$\bullet$};
-\node (N6) at (2,5) {$\Smiley[2][green]$};
-\node (N7) at (4,6) {$\Smiley[2][green]$};
-\node (N88) at (7,4) {$\bullet$};
-\node (N8) at (10,4) {$\bullet$};
-\node (N9) at (14,6) {$\Smiley[2][green]$};
-\node (N10) at (0,8) {$\Smiley[2][green]$};
-\node (N11) at (10,7) {$\Smiley[2][green]$};
-\node (N111) at (7,7) {$\Smiley[2][green]$};
-\draw (foot) to [bend left=10] (N1);
-\draw (foot) -- (N2);
-\draw (foot) -- (N22);
-\draw (foot) -- (N3);
-\draw (foot) -- (N4);
-\draw (N1) to  (N5);
-\draw (N1) to   [bend left=10] (N6);
-\draw (N1) to   [bend right=20] (N7);
-\draw (N2) to  (N8);
-\draw (N22) to  (N88);
-\draw (N8) to  (N11);
-\draw (N88) to  (N111);
-\draw (N3) to  (N9);
-\draw (N5) to  (N10);
-\end{tikzpicture}
-\caption{The hydra $\iota(\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$) \label{fig:iota-example}}
-
-\end{figure}
-
-The following lemma (proved in ~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H.v}) maps  canonical sequences to rounds of hydra battles.
-
-\label{lemma:canonS-iota}
-
-\input{movies/snippets/O2H/canonSIota}
-
-The next step of our development extends this relationship to
-the order $<$ on $[0,\epsilon_0)$ on one side, and hydra battles on the other side.
-
-\input{movies/snippets/O2H/pathToRoundPlus}
-
-As a corollary, we are now able to transform any inequality $\beta<\alpha<\epsilon_0$ into a (free) battle.
-
-\input{movies/snippets/O2H/LTToRoundPlus}
-
-\section{A  proof of impossibility}
-
-We now have  the tools for proving that  there exists no variant bounded by some $\mu<\epsilon_0$ for proving the termination   of all battles.
-The proof we are going to show is a proof by contradiction.
-It  can
- be considered as a generalization of the
-proofs described in  sections~\vref{omega-case} and \vref{omega2-case}.
-
-In the module
-\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$.
-This part of the development is parameterized by some class $B$ of battles, which will be instantiated later to \texttt{free} or \texttt{standard}.
-
-\vspace{4pt}
-\noindent
-From~\href{../theories/html/hydras.Hydra.~Hydra_Definitions.html}{Hydra.Hydra\_Definitions}:
-
-\input{movies/snippets/Hydra_Definitions/BoundedVariant}
-
-Let us assume there exists such a variant:
-
-\input{movies/snippets/Epsilon0_Needed_Generic/theContext}
-
-\label{remark:m-decrease}
-\begin{remark}
-  The hypothesis \texttt{m\_decrease} is not provable  in general, but is satisfied by
-the  \texttt{free} and \texttt{standard} kinds of battles. This trick allows to 
-``factorize'' our proofs  of impossibility.
-\end{remark}
-
-\index{maths}{Transfinite induction}
-
-First, we prove that $m(\iota(\alpha))$ is always greater than or equal to $\alpha$, by  transfinite induction over $\alpha$.
-
-\input{movies/snippets/Epsilon0_Needed_Generic/mGe0}
-
-\begin{itemize}
-\item If $\alpha=0$, the inequality trivially holds
-\item If $\alpha$ is the successor of  some ordinal $\beta$, the inequality $\beta \leq m(\iota(\beta))$ holds (by induction hypothesis).
-But the hydra $\iota(\alpha)$ is transformed in one round into
-$\iota(\beta)$, thus $m(\iota(\beta))<m(\iota(\alpha))$. Hence $\beta<m(\iota(\alpha))$, which implies $\alpha \leq m(\iota(\alpha))$
-\item If $\alpha$ is a limit ordinal, then $\alpha$ is the least upper bound of the set
-of all  the $\canonseq{\alpha}{i}$.  Thus, we have just to prove that $\canonseq{\alpha}{i}< m(\iota(\alpha))$ for any $i$. 
-\begin{itemize}
-\item Let $i$ be some natural number.
-By the induction hypothesis, we have $\canonseq{\alpha}{i} \leq m(\iota(\canonseq{\alpha}{i}))$.
-But the hydra $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i})$ in one round, thus $m(\iota(\canonseq{\alpha}{i})) < m(\iota(\alpha))$, by our hypothesis \texttt{m\_decrease}.
-\end{itemize}
-\end{itemize}
-
-Please note that the impossibility proofs of 
-sections~\vref{omega-case} and \vref{omega2-case} contain a similar lemma, also called \texttt{m\_ge}.
-We are now able to build a counter-example.
-
-\input{movies/snippets/Epsilon0_Needed_Generic/mGeGeneric}
-
-The (big) rest of the proof is dedicated to prove formally the converse inequality 
-\texttt{m small\_h t1< m big\_h}.
-
-\subsection{The case of free battles}
-\label{sec:free-battles-case}
-Let us now consider that $B$ is instantiated to \texttt{free} (which means that we are considering proofs of termination of \emph{all} battles).
-The following lemmas are proved in Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
-The case $B=\texttt{standard}$ is studied in section~\vref{std-case}.
-
-\input{movies/snippets/Epsilon0_Needed_Free/theContext}
-
-\begin{enumerate}
-\item The following lemma is an application of \texttt{m\_ge\_generic}, since \texttt{free}
-satisfies trivially the hypothesis \texttt{m\_decrease} (see page~\pageref{remark:m-decrease}).
-
-\inputsnippets{Epsilon0_Needed_Free/mGe}
-
-\item From the hypothesis \texttt{Hy}, we have \texttt{m big\_h t1< mu}
-\item By Lemma \texttt{LT\_to\_round\_plus}, we get a (free) battle from
-\texttt{big\_h = iota mu} to \texttt{small\_h = iota (m big\_h)}.
-
-\input{movies/snippets/Epsilon0_Needed_Free/bigToSmall}
-
-\item From the hypotheses on $m$, we infer:
-\input{movies/snippets/Epsilon0_Needed_Free/mLt}
-
-\item From lemmas \texttt{m\_ge} and \texttt{m\_lt}, and the irreflexivity of $<$, we get a contradiction. 
-\input{movies/snippets/Epsilon0_Needed_Free/ImpossibilityFree}
-
-\end{enumerate}
-
-We have now proved there exists no bounded variant for the class of free battles.
-
-\input{movies/snippets/Epsilon0_Needed_Free/CheckDemo}
-
-\section{The case of standard battles}
-\label{sec:standard-intro}\label{std-case}
-One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
-Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}.
-
-\input{movies/snippets/O2H/LTToRoundPlus}
-
-This lemma builds a battle out of any inequality $\beta<\alpha$. 
-It is a straightforward application of \texttt{LT\_path\_to} of
-Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}:
-
-\input{movies/snippets/Paths/LTPathTo}
-
-The sequence $s$, used to build the sequence of replication factors of the battle, depends on 
-$\alpha$ and $\beta$, so we cannot be sure that the generated battle is a genuine standard battle.
-
-The solution to this issue comes  once again from Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
-$\alpha_0=\alpha,\alpha_1,\dots,\alpha_k=\beta$ where $\alpha_{j+1}$ is equal
-to $\canonseq{\alpha_j}{i_j}$ where $i_j$ is \emph{any} natural number, 
-we consider various constraints on these sequences.
-In particular, a path is called \emph{standard} if $i_{j+1} = i_j + 1$ for every $j<k$.
-It  corresponds to a ``segment'' of some standard battles. 
-Please note that the vocabulary on paths is ours, but all the concepts come really from~\cite{KS81}.
-
-In \coq{}, standard paths can be defined as follows.
-
-\vspace{4pt}
-
-\emph{From
-Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/standardPathR}
-
-In the mathematical text and figures, we shall use the notation 
-$\alpha \xrightarrow[i,j]{}\beta$ for the proposition 
-(\texttt{standard\_path $i$ $\alpha$ $j$ $\beta$}).
-In~\cite{KS81} the notation is
-$\alpha \xrightarrow[i]{*}\beta$
-for 
-the proposition  $\exists j, i<j \wedge \alpha \xrightarrow[i,j]{} \beta$.
-
-Our goal is now  to transform any inequality $\beta<\alpha<\epsilon_0$ into a standard path $\alpha \xrightarrow[i,j]{} \beta$ for some $i$ and $j$, then into a standard battle
-from $\iota(\alpha+i)$ to $\iota(\beta)$. 
-Following~\cite{KS81}, we proceed in two stages:
-\begin{enumerate}
-\item we simulate plain (free) paths from $\alpha$ to $\beta$ with
-paths made of steps $(\gamma,\canonseq{\gamma}{n})$, \emph{with the same $n$ all along the path}
-\item we simulate any such path by a standard path.
-\end{enumerate}
-
-\subsection{Paths with a constant index}
-
-First of all, paths with a constant index 
-enjoy nice properties. They are defined as paths where all the $i_j$ are equal to the same natural number $i$, for some $i>0$.
-
-Like in~\cite{KS81}, we shall use the notation $\alpha \xrightarrow[i]{} \beta$ for denoting such a path, also called an $i$-path.
-
-\input{movies/snippets/Paths/constPathDef}
-
-% Paths with a given index can be effectively computed.
-% Given $i$, $\alpha$ and $l$, the following function returns the ordinal $\beta$ such that there exists a path 
-% $\alpha \xrightarrow [i+1] {} \beta$ of length $l$.
-
-% \begin{Coqsrc}
-% Fixpoint const_funS (i:nat)(alpha : T1)(l:nat):  T1  :=
-%   match l
-%   with
-%   | 0 => alpha
-%   | S m => const_funS i (canonS i alpha) m
-%   end.
-% \end{Coqsrc}
-
-% The following computations show  applications of \texttt{constS\_fun} to the 
-% ordinal $\omega^\omega$, with various values of $i$ and $l$.
-
-% \begin{Coqsrc}
-% Compute  (const_funS 2 (omega ^omega)  55).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-%   = zero
-%      : T1 
-% \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (const_funS 2 (omega ^omega) 15).
-% \end{Coqsrc}
-
-%   \begin{Coqanswer}
-%  = (omega ^ 2 * 2)%pT1
-%      : ppT1   
-%   \end{Coqanswer}
-
-% \begin{Coqsrc}
-% Compute pp (const_funS 4 (omega^omega)  100).
-% \end{Coqsrc}
-
-% \begin{Coqanswer}
-% = (omega ^ 4 * 4 + omega ^ 3 * 4 + omega ^ 2 + omega * 4 + 4)%pT1
-%      : ppT1
-% \end{Coqanswer}
-
-A most interesting property of $i$-paths is that we can ``upgrade'' their index, as stated by K.\&S.'s Corollary 12.
-
-\index{maths}{Transfinite induction}
-
-\vspace{4pt}
-\noindent
-\emph{From
-  Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/Cor12}
-
-We  also use a version of \texttt{Cor12} with large inequalities.
-
-\input{movies/snippets/Paths/Cor121}
-
-\subsubsection{Sketch of proof of \texttt{Cor12}}
-\index{maths}{Transfinite induction}
-
-First, let us prove the following lemma (numbered \texttt{2.4} in Ketonen and Solovay's article).
-
-\input{movies/snippets/Paths/KSThm24}
-
-We prove this lemma by transfinite induction on $\alpha$.
-Let us consider a path $\alpha \xrightarrow [i]{} \beta$ $(i>0)$. Its first step is
-the pair $(\alpha,\canonseq{\alpha}{i})$, We have $\canonseq{\alpha}{i}<\alpha$ and
-$\canonseq{\alpha}{i} \xrightarrow [i]{} \beta$. 
-Let $n$ be any natural number such that $n>i$.
-By the induction hypothesis, there exists a path $\canonseq{\alpha}{n} \xrightarrow[i]{} \beta$.
-\begin{itemize}
-\item  If $\alpha$ is a successor ordinal $\gamma+1$, then $\canonseq{\alpha}{n} =
-\canonseq{\alpha}{i}=\gamma$. Thus we have a path 
-$\alpha  \xrightarrow [n]{}  \gamma \xrightarrow [n]{} \beta$
-\item If $\alpha$ is a limit ordinal, we apply Theorem \texttt{KS\_thm\_2\_4} (see above).
-
-%   \begin{theorem}
-% Let $\lambda$ be a limit ordinal, then for any pair of indices $0<i<j$, there is a path $\canonseq{\lambda}{j} \xrightarrow[1]{} \canonseq{\lambda}{i}$.    
-%   \end{theorem}
-
-The proof of the limit case, is decomposed into a sequence
-of path constructions leading to $\alpha \xrightarrow[n]{} \beta$.
-
- \begin{enumerate}
- \item $\alpha \xrightarrow[n]{} \canonseq{\alpha}{n}$ (single step path)
- \item $\canonseq{\alpha}{n} \xrightarrow[1]{} \canonseq{\alpha}{i}$ (by \texttt{Theorem\_2\_4}),
-\item $\canonseq{\alpha}{n} \xrightarrow[n]{} \canonseq{\alpha}{i}$ (applying the induction hypothesis to the preceding path);
-\item $\canonseq{\alpha}{i} \xrightarrow[n]{} \beta$ (applying the induction hypothesis)
-\item $\alpha \xrightarrow[n]{} \beta$ (by composition of 1, 3, and 4).
-
- \end{enumerate}
-
-\end{itemize}
-
-\begin{remark}
- \texttt{Cor12} ``casts'' $i$-paths into $n$-paths for any $n>i$.
-But the obtained $n$-path can be much longer than the original $i$-path.
-The following exercise will give an idea of this increase. 
-\end{remark}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-  Prove that  the length of the $i+1$-path from
-  $\omega^\omega$ to $\omega^i$ is $1 + (i+1)^{(i+1)}$, for any $i$. Note that the $i$-path from
-  $\omega^\omega$ to $\omega^i$ is only one step long.
- \end{exercise}
-
-Why is \texttt{Cor12} so useful? 
-Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
-we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying corollary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
-where $n$ is the maximum of the indices $i$ met in the induction.
-
- Lemma 1, Section 2.6 of~\cite{KS81} is naturally expressed in terms of \coq's
-\verb@sig@ construct.
-
-\label{lemma:L-2_6-1}
-%\index{coq}{Sigma types}
-
-\input{movies/snippets/Paths/Lemma261}
-
-Intuitively, Lemma~\texttt{2\_6\_1}  shows that if $\beta<\alpha<\epsilon_0$, then there exists  a battle from $\iota(\alpha)$ to $\iota(\beta)$ where the replication factor is constant, although large enough.
-
- \paragraph*{\gaiasign}
- Corollary \texttt{Cor12} and Lemma \texttt{Lemma2\_6\_1} are also available in
-\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} 
-(please see Sect.~\vref{sect:gpath}).
-
-\subsection{Casting paths with a constant index into a standard path}
-
-The article~\cite{KS81} contains 
-the following lemma, which allows us to simulate $i$-paths by $[i+1,j]$-paths, where $j$ is large enough.
-
-\input{movies/snippets/Paths/constantToStandard}
-
-\subsubsection{Sketch of proof of \texttt{constant\_to\_standard\_path}}
-
-Our proof follows the proof by Ketonen and Solovay, including its organization as a sequence of lemma.
-Since it is a non-trivial proof, we will comment its main steps below   (see Figure~\vref{fig:belle-preuve-1} to Figure~\vref{fig:fin-belle-preuve}).
-
-\subsubsection*{Preliminaries}
-
-Please note that, given an ordinal $\alpha:\texttt{T1}$, and two natural numbers $i$ and $l$, there exists at most a standard path $\alpha \xrightarrow [i,i+l]{*} \beta$.
-The following function computes $\beta$ from $\alpha$, $i$ and $l$.
-
-\input{movies/snippets/Paths/standardGnaw}
-
-\index{maths}{Transfinite induction}
-
-By transfinite induction over  $\alpha$, we prove that the ordinal $0$ is reachable from any ordinal $\alpha<\epsilon_0$ by some standard path.
-
-\input{movies/snippets/Paths/standardPathToZero}
-
-\paragraph*{}
-Now, let us consider two ordinals  $\beta<\alpha<\epsilon_0$.  Let $p$  be some $(n+1)$-path from $\alpha$ to $\beta$.
-
-\input{movies/snippets/Paths/ConstantToStandardProof}
-
-Applying \texttt{standard\_path\_to\_zero}, $0$ is reachable from $\alpha$ by some standard path.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \draw[->, very thick,blue] (alpha)-- node [below]{$n+1$} node [above] {$+$} (beta);
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-    \node (alpha3) at (20,5) {};
-  \node (alphalast) at (35,5) {};
-  \node (zero) at (45,0) {$0$};
-  \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate=40]{$n+1$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha2)-- node [below]{$n+3$}  (alpha3);
-
-  \node (dots) at (24,5) {$\dots$};
-  \draw [->, dashed, very thick,blue] (alphalast)-- node [below, rotate=-26]{$n+p+1$}  (zero);
-
-\end{tikzpicture}
-\caption{A nice proof (1)}
-  \label{fig:belle-preuve-1}
-\end{figure}
-
-\paragraph*{}
-
-Since comparison on \texttt{T1} is decidable, one can compute the last step $\gamma$ of the standard path from $(\alpha,n+1)$  such that $\beta\leq \gamma$.
-Let $l$ be the length of the path from $\alpha$ to $\gamma$.  
-This step of the proof is illustrated in figure~\vref{fig:belle-preuve-2}.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (17,5) {$\ldots$};
-    \node (alpha3) at (20,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (38,0) {$\delta$};
-    \draw [->, dashed,very thick,blue] (alpha)-- node [below,rotate=35]{$n+1$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{$n+2$}  (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below,rotate = -48]{\tiny $n+l$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{$n+l+1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{$n+1$} node [above] {$+$} (beta);
-   \draw[thick] (alpha)--  (gamma);
-   \draw[thick] (gamma)--  node [above] {$\geq$} (beta);
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-\end{tikzpicture}
-
-\caption{A nice proof (2)}
-  \label{fig:belle-preuve-2}
-\end{figure}
-
-\paragraph*{}
-
-\begin{itemize}
-\item If $\beta=\gamma$, it's OK! We have got a standard path
-from  
-$\alpha$ to $\beta$ with successive indices  $n+1, n+2, \dots, n+l+1$
-
-\item Otherwise,  $\beta < \gamma$.  Let us consider  $\delta=\canonseq{\gamma}{n+l+1}$.
-By applying several times lemma \texttt{Cor12},  one converts  every path of Fig~\ref{fig:belle-preuve-2} into
- a $n+l+1$-path  (see figure~\ref{fig:belle-preuve-3}).
-
-But $\gamma$ is on the $n+l+1$-path from $\alpha$ to $\beta$.
-As shown by figure~\vref{fig:fin-belle-preuve}, the ordinal $\delta$, reachable from
-$\gamma$ in one single step,  must be greater than or equal to $\beta$, which contradicts our  hypothesis $\beta < \gamma$.
-
-\begin{figure}[h]
-  \centering
-
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-    \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (17,5) {$\ldots$};
-    \node (alpha3) at (20,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (38,0) {$\delta$};
-     \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -48]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\tiny $n+l+1$} node [above]{\color{red} $1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
-    \draw[thick] (gamma)--   node [above]{\color{red} $>$}(beta);
-   \draw[thick] (alpha)--  (gamma);
-
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-
-\end{tikzpicture}
-
-\caption{A nice proof (3)}
-  \label{fig:belle-preuve-3}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-\begin{tikzpicture}[very thick, scale=0.25]
-\node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (32, 0){$\beta$};
-
-  \node (alpha1) at (5,5) {};
-  \node (alpha2) at (13,5) {};
-  \node (dots) at (15,5) {$\ldots$};
-    \node (alpha3) at (18,5) {};
-    \node (gamma) at (24,0) {$\gamma$};
-    \node (delta) at (42,0) {$\delta$};
-    \draw [->, dashed,very thick,blue] (alpha)-- node [below, rotate = 40] {\tiny $n+l+1$}  node [above, rotate = 40]{\tiny $+$}  (alpha1);
-  \draw [->, dashed,very thick,blue] (alpha1)-- node [below]{\tiny $n+l+1$} node [above]{\tiny $+$} (alpha2);
-   \draw [->, dashed,very thick,blue] (alpha3)-- node [below, rotate = -36]{\tiny $n+l+1$} node [above, rotate = -36]{\tiny $+$}  (gamma);
-   \draw  [->, dashed, blue] (gamma) to    [bend left=80] node [below]{\small $n+l+1$} node [above]{\color{red} $1$} (delta);
-   \draw[->, very thick,blue] (alpha) to [bend right=34] node [below]{\small $n+l+1$} node [above] {\tiny $+$} (beta);
-   \draw[thick] (alpha)--  (gamma);
-
-    \draw[thick] (gamma)--  node [below]{\tiny $n+l+1$} node [above]{\color{red} $+$}(beta);
-    \draw[thick] (beta)--  node [above] {$>$} (delta);
-
-\end{tikzpicture}
-
-\caption{A nice proof (4)}
-  \label{fig:fin-belle-preuve}
-\end{figure}
-
-\end{itemize}
- The only possible case is  thus $\beta=\gamma$, so we have got a standard path  from $\alpha$ to $\beta$.
-
- \input{movies/snippets/Paths/constantToStandard0}
- \input{movies/snippets/Paths/constantToStandardz}
-
-Here is the full statement of the conversion from constant to standard paths.
-
-\input{movies/snippets/Paths/constantToStandardPath}
-
-Applying \texttt{Lemma2\_6\_1} and \texttt{constant\_to\_standard\_path}, we get the following corollary.
-
-\input{movies/snippets/Paths/LTToStandardPath}
-
-\subsection{Back to hydras}
-\label{sec:standard-battles-cases}
-We are now able to complete our proof that there exists no bounded variant for proving the termination of standard hydra battles. This proof can
-be consulted in the module 
-\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}{Hydra.Epsilon0\_Needed\_Std}.
-Please note that it has the same global structure as in section~\ref{sec:free-battles-case} 
-% ICI !
-Applying the  lemmas  \texttt{Lemma2\_6\_1} of the module 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
-{Epsilon0.pathS}   and 
-\href{../theories/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
-{\texttt{constant\_to\_standard\_path}},
-we can convert any inequality $\beta<\alpha<\epsilon_0$ into a standard path from
-$\alpha$ to  $\beta$, then into a fragment of a standard battle from 
-$\iota(\alpha)$ to $\iota(\beta)$, hence the inequality $m(\iota(\beta))<m(\iota(\alpha))$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html\#LT_to_standard_battle}{Hydra.Epsilon0\_Needed\_Std}}
-
-\input{movies/snippets/Epsilon0_Needed_Std/LTToStandardBattle}
-
-Now, please consider the following context:
-
-\input{movies/snippets/Epsilon0_Needed_Std/theContext}
-
-In the same way as for free battles, we import a large inequality 
-from 
-the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Epsilon0\_Needed\_Generic}.
-
-\input{movies/snippets/Epsilon0_Needed_Std/mGe}
-
-\paragraph*{} If remains to prove the following strict inequality, in order to have a contradiction.
-
-\input{movies/snippets/Epsilon0_Needed_Std/mLt}
-
-\paragraph*{Sketch of proof:} Let us recall that $\texttt{big\_h} = \iota(\mu)$
- and $\texttt{small\_h} = \iota (m (\texttt{big\_h}))$.
-
-Since $m(\texttt{big\_h})< \mu$, there exists a standard path from $\mu$ to
-$m(\texttt{big\_h})$, hence a   standard battle from $\iota(\mu)$  to
-$\iota(m(\texttt{big\_h}))$,  i.e. from \texttt{big\_h} to \texttt{small\_h}.
-
-Since $m$ is assumed to be a variant for standard battles, we get the inequality  $m(\texttt{small\_h}) < m(\texttt{big\_h})$.
-
-\input{movies/snippets/Epsilon0_Needed_Std/endOfProof}
-
-\subsection{Remarks}
-
-We are grateful to 
- J. Ketonen and R. Solovay  for the high quality of their explanations and proof details.
-Our proof follows tightly the sequence of lemmas in their article, with a focus on 
-constructive aspects.
-Roughly speaking, our implementation \emph{builds}, out of a hypothetical
-  variant $m$, bounded by some ordinal $\mu<\epsilon_0$, a hydra \texttt{big\_h} which verifies the impossible inequality  $m(\texttt{big\_h})< m(\texttt{big\_h})$.
-
-On may ask whether the preceding results are too restrictive, since they 
-refer to a particular data type \texttt{T1}.
-In fact, our representation of ordinals strictly less than 
- $\epsilon_0$ is faithful to their mathematical definition, at least 
-Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
-(please see also the module
-\href{../theories/html/hydras.Schutte.Correctness_E0.html}{hydras.Schutte.Correctness\_E0}).
-
-Thus, we can infer that our theorems can be applied to any well order.
-
-\index{hydras}{Projects}
-\begin{project}
-Study a possible modification of the definition of a variant  (for  standard battles).
-
-\begin{itemize}
-\item The variant is assumed to be strictly decreasing \emph{on configurations 
-reachable from some initial configuration where the replication factor is equal to $0$}
-\item The variant may depend on the number of the current round.
-\end{itemize}
-
-In other words, its type should be \texttt{nat -> Hydra -> T1}, and it must 
-verify the inequality $m\, (S\,i)\, h' < m\,i\, h$ whenever the configuration 
-$(i,h)$ is reachable from some initial configuration $(0,h_0)$
-and \texttt{h} is transformed into \texttt{h'} in the considered round.
-Can we still prove the theorems of section~\ref{std-case} with this new definition?
-
-\end{project}
-
-%---------------------------------------------------------------------
-\chapter{Large sets and rapidly growing functions}\label{chap:alpha-large}
-
-\begin{remark}
-Some notations (mainly names of fast-growing functions) of our development may differ slightly from the literature.
-Although this fact does not affect our proofs, we are preparing a future version where the names $F\_alpha$, $f\_alpha$, $H\_alpha$, etc., are fully consistent with the cited articles.
-
-\end{remark}
-%\section{Introduction}
-
-In this chapter, we try to feel how long a standard battle can be.
-To be precise, for any ordinal $\alpha<\epsilon_0$ and any positive integer $k$,
-we give a minoration of the number of steps of a standard battle which
-starts with the hydra $\iota(\alpha)$ and the replication factor $k$.
-
-We express this number in terms of the Hardy hierarchy of fast-growing 
-functions~\cite{BW85, Wainer1970, KS81, Promel2013}.
- From the \coq{} user's point of view, such  functions are  very 
-attractive:  they are defined as functions  in \gallina{}, and we can apply them \emph{in theory}, but they are so complex that you will never be able to look at the result of the computation.
- Thus, our knowledge on these functions must rely on \emph{proofs}, not tests. In our development, we use often the rewriting rules generated by \coq's \texttt{Equations} plug-in.
-
-\section{Definitions}
-
-%\subsection{Definition}
-
-\begin{definition}
-Let $0<\alpha<\epsilon_0$ be any ordinal, and $s=\langle s_1, s_2, \dots, s_N\rangle$ a finite sequence of strictly positive natural numbers.
-
-We say that $s$ is \emph{$\alpha$-large} if the sequence $\langle \alpha_0=\alpha,\dots,\alpha_{i+1}=\canonseq{\alpha_i}{i+1},\dots \rangle$ leads to $0$. 
-We say also that $s$ is \emph{minimally $\alpha$-large} (in short:
-\emph{$\alpha$-mlarge}) if $s$ is $\alpha$-large 
- and every strict prefix of $s$ leads to a non-zero ordinal (\emph{cf} Sect.~\vref{sect:path-to-def}).
-
-\index{maths}{Ordinal numbers!Large sets}
-\index{maths}{Ordinal numbers!Minimal large sets}
-
-\end{definition}
-
-\begin{remark}
-  Ketonen and Solovay~\cite{KS81} consider  large finite \emph{sets} of natural numbers,  but they are mainly used as sequences. Thus, we chose to represent them explicitely as (sorted) lists. 
-\end{remark}
-
-The following function ``gnaws'' an ordinal $\alpha$, following a sequence of indices (ignoring the $0$s).
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}
-
-\input{movies/snippets/Paths/gnawDef}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_SetsPaths.html\#gnaw}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/largeDef}
-
-Minimal large sequences can be directly defined in terms of the
-predicate \texttt{path\_to} (\vref{sect:path-to-def}) which already prohibits paths containing non-final \texttt{zero}s.
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#mlarge}{Epsilon0.Large\_Sets}}
-
-\index{hydras}{Library Epsilon0!Predicates!mlarge@mlarge (minimal large sequences)}
-
-\input{movies/snippets/Large_Sets/mlargeDef}
-
-Let us consider two integers $k$ and $l$, such that $0<k<l$. In order to check whether the interval $[k,l]$ is minimally large for $\alpha$, it is enough to
-follow from $\alpha$ the path associated with the interval $[k,l)$ and verify that the last ordinal we obtain is equal to $1$.
-
-\subsection{Examples}
-
-For instance the interval $[6,70]$ leads $\omega^2$ to $\omega\times 2 + 56$. Thus this interval is not $\omega^2$-large.
-
-\noindent
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html\#mlarge}{Epsilon0.Large\_Sets\_Examples}}
-
-\input{movies/snippets/Large_Sets_Examples/gnawEx1}
-
-The interval $[6,700]$ is $\omega^2$-large, but not
-$\omega^2$-mlarge, since $[6,699]$ is also $\omega^2$-large.
-
-\input{movies/snippets/Large_Sets_Examples/gnawEx2}
-
-The following lemma relates minimal largeness with the
-function 
-\texttt{gnaw}.
-
-\input{movies/snippets/Large_Sets/mlargeIff}
-
-\vspace{4pt}
- \noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets_Examples.html}{Epsilon0.Large\_Sets\_Examples}}
-
-\input{movies/snippets/Large_Sets_Examples/Ex1Lemma}
-
-\section{Length of minimal large sequences}
-
-Now, consider any natural number $k>0$ any ordinal $0<\alpha<\epsilon_0$.  We would like to compute
-a number $l$ such that the interval $[k,l]$ is $\alpha$-mlarge. So, 
-the standard battle starting with $\iota(\alpha)$ and the replication factor $k$ will end after $(l-k+1)$ steps.
-
-First, we notice that this  number $l$ exists, since the segment $[0,\epsilon_0)$ is well-founded and $\canonseq{\alpha}{i}<\alpha$ for any $i$ and $\alpha>0$.
-Moreover, it is unique:
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/mlargeUnicity}
-
-Thus, we would like to define a function, parameterized by $\alpha$ which associates to any  strictly positive integer $k$ the number $l$ such that
-the interval $[k,l]$ is $\alpha$-mlarge. It would be fine to write in \gallina{} a definition like this:
-
-\begin{Coqbad}
-Function L_ (alpha: E0) (i:nat) :  nat := ...
-\end{Coqbad}
-
-But we do not know how to fill the dots yet \dots{}   In the next section, we will 
-use \coq{} to reason  about the \emph{specification} of \texttt{L},
-prove properties of any function which satisfies this specification.
-In Sect.~\ref{sect:L-equations}, we use the \texttt{coq-equations} plug-in
-to define a function \texttt{L\_}, and prove its correctness w.r.t. its specification.
-
-\subsection{Formal specification}
-
-Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider any  function which  maps  any strictly positive integer $k$ to the number $l$, where 
-the interval $[k,l)$ is $\alpha$-mlarge.
-
-\begin{remark}
-In~\cite{KS81} Ketonen and Solovay consider the least natural number $l$ where the interval $[k,l]$ ($l$ included) is $\alpha$-large, and call $H_\alpha$ the function which maps $k$ to $l$.
-We chose to consider intervals $[l,k)$ instead of $[l,k]$
-in order to simplify  some statements and proofs in composition lemmas associated with the ordinals of the form $\alpha\times i$ and 
-$\omega^\alpha\times i + \beta$.
-Clearly, both approaches are related through the equality
-$L_\alpha(k)=H_\alpha(k)+1$, for any non-null $\alpha$ and $k$.
-\end{remark}
-
-Our specification of the function \texttt{L} is as follows:
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/LSpecDef}
-
-\begin{todo}
- Check if the functions $L_\alpha$ are the same as
-\cite{KS81}' functions $f_\alpha$ (p. 297).
-\end{todo}
-
-Note that, for $\alpha\not=0$, the value of $f(0)$ is not specified.
-Nevertheless, the restriction of $f$ to the set of strictly positive integers is unique (up to extensionality).
-
-\input{movies/snippets/Large_Sets/LSpecUnicity}
-
-\subsection{Abstract properties}
-
-Let us now prove properties of any function $f$ (if any) which satisfies 
-\texttt{L\_spec}.
-We are looking for properties which could be used for writing \emph{equations} and prove the correctness of the function generated by the \texttt{coq-equations} plug-in.
-Moreover, they will give us some examples (for small values of $\alpha$).
-
-The properties we consider are defined in \href{../theories/html/hydras.Prelude.Iterates.html\#fun_le}{Prelude.Iterates}.
-
-\label{sect:abstract-arith-prop}
-\input{movies/snippets/Iterates/funLeDef}
-
-Our exploration of the $L_\alpha$\,s  considers the usual cases of a proof by transfinite induction: zero, successors and limit ordinals.
-The lemmas we are going to prove will be applied in a big proof by induction in Sect~\vref{sect:L-correct-proof}.
-
-\index{maths}{Transfinite induction}
-
-\subsubsection{The  ordinal zero}
-\label{sect:L-spec-zero}
-The base case is directly a consequence of the specification.
-
-\input{movies/snippets/Large_Sets/LZeroInv}
-
-\subsubsection{Successor ordinals}
-\label{sect:L-spec-succ}
-Let $\beta$ be some ordinal, and assume the arithmetic function $f$ satisfies 
-the specification $(\texttt{L\_spec}\;\beta)$.  Let $k$ be any natural number.
-Any path from $\texttt{succ}\,\beta$ to $0$ starting at $k+1$ can be decomposed into a first step from $\texttt{succ}\,\beta$ to $\beta$, then a path from
-$\beta$ at $k+2$ to $0$. 
-By hypothesis the interval $[k+2, f(k+2)-1]$ is $\beta$-mlarge.
-But the interval $[k+1, f(k+2)-1]$ is the concatenation of the singleton
-$\{k+1\}$ and the interval $[k+2, f(k+2)-1]$.
-So, the function $\lambda\,k.\,f(k+1)$ satisfies the specification $\texttt{L\_spec}\,\beta$.
-
-Note that our decomposition of intervals works only if the intervals we consider are not empty.
-In order to ensure this property, we assume that $f\;k$ is always greater than $k$, which we note \texttt{S <<= f}, or \texttt{(fun\_le S f)}.
-
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Large_Sets.html}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/SectionSucc}
-
-\subsubsection{Limit ordinals}
-\label{sect:L-spec-lim}
-
-Let $\lambda<\epsilon_0$ be any limit ordinal. In a similar way as for successors, we decompose any path from $\lambda$  into a first step to
-$\canonseq{\lambda}{k}$, followed by a path to $0$. In the following section, we assume that there exists a correct function computing  $L_{\canonseq{\lambda}{k}}$ for any strictly positive $k$.
-
-\input{movies/snippets/Large_Sets/SectionLim}
-
-\subsection{First results}
-
-Applying the previous lemmas on successors and limit ordinals, 
-we obtain a few  correct implementations of \texttt{(L\_spec $\alpha$)} for small values of $\alpha$.
-
-\subsubsection{Finite ordinals}
-
-By iterating the functional \texttt{L\_succ}, we get a realization of
-\texttt{(L\_spec $i$)} for any natural number $i$.
-
-\input{movies/snippets/Large_Sets/LFinDef}
-\vspace{-16pt}
-\input{movies/snippets/Large_Sets/LFinOk}
-
-\subsubsection{The first limit ordinal  \texorpdfstring{$\omega$}{omega}}
-
-The lemmas \texttt{L\_fin\_ok} and \texttt{L\_lim\_ok}   allow us to get 
-by diagonalization a correct implementation for 
-\texttt{L\_spec omega}.
-
-\input{movies/snippets/Large_Sets/LOmegaDef}
-\vspace{-16pt}
-\input{movies/snippets/Large_Sets/LOmegaOk}
-
-\subsubsection{Towards  \texorpdfstring{$\omega^2$}{omega*omega}}
-
-We would like to get exact formulas for the ordinal $\omega^2$, a.k.a.
-$\phi_0(2)$. This ordinal is the limit of the sequence $\omega\times i\;(i \in \mathbb{N})$. Thus, we have to study ordinals of this form, then use 
-our lemma on limits.
-
-The following lemma establishes a path from $\omega\times ( i+1)$ to
-$\omega \times i$.
-
-\input{movies/snippets/Large_Sets/pathToOmegaMult}
-
-Let us consider a path from  $\omega\times(i+1)$ to $0$ starting at $k+1$.
-A first ``big step'' will lead to $\omega\times i$ at $2(k+1)$. If $i>0$, the
-next jump leads to $\omega\times(i-1)$ at $2(2(k+1))+1$, etc.
-
-The following lemma expresses the length of the mlarge sequences associated with any finite multiple of $\omega$.
-
-\input{movies/snippets/Large_Sets/omegaMultMlarge0}
-
-\emph{From Module~ \href{../theories/html/hydras.Epsilon0.Large_Sets.html\#L_omega_mult}{Epsilon0.Large\_Sets}}
-
-\input{movies/snippets/Large_Sets/LOmegaMultDef}
-
-More generally, we prove the equality $L_{\omega\times i}(k)=2^i\times(k+1)-1$.
-
-\input{movies/snippets/Large_Sets/LOmegaMultEqn}
-
-Correctness of the function \texttt{L\_omega\_mult} is asserted through the following lemma.
-
-\input{movies/snippets/Large_Sets/LOmegaMultOk}
-
-By diagonalization, we obtain a simple formula for $L_{\omega^2}$.
-
-\input{movies/snippets/Large_Sets/LOmegaSquare}
-\input{movies/snippets/Large_Sets/LOmegaSquareEqn}
-\input{movies/snippets/Large_Sets/LOmegaSquareOk}
-
-%%%% ICI
-
-\subsubsection{Going further}
-Let us consider a last example, ``computing'' $L_{\omega^3}$.
-Let us study the canonical sequence associated with $\omega^3$, the elements of which are 
-$\omega^2\times i\;(i\in\mathbb{N}_1)$.
-
-To this end, we prove a generic lemma, which expresses $L_{\omega^\alpha\times i}$ as an iterate of $L_{\omega^\alpha}$.
-Note that in this lemma, we assume that the function associated with $\alpha$ is strictly monotonous and
-greater or equal than the successor function, and prove that $L_{\omega^\alpha\times i}$ satisfies  the same properties.
-
-\inputsnippets{Large_Sets/phi0Mult}
-
-\inputsnippets{Large_Sets/phi0MultOk,
-Large_Sets/Phi0MultSLe}
-
-Let us look now
-at the ordinal $\omega^2\times i$, using \texttt{L\_phi0\_mult}.
-
-\input{movies/snippets/Large_Sets/LOmegaSquareTimes}
-
-We are now ready to get an exact formula for $L_{\omega^3}$, by diagonalization over $L_{\omega^2\times i}$.
-
-\input{movies/snippets/Large_Sets/LOmegaCube}
-
-Thus, for instance, $L_{\omega^3}(3)=L_{\omega^2\times 4}(3)$.
-
-\input{movies/snippets/Large_Sets/LOmegaCube3Eq}
-
-This number is quite big. Using \texttt{Ocaml}'s \texttt{float} arithmetic,
-we can under-approximate it by $2^{3.8\times10^{30}}\times 3.8\times{10^{30}}$.
-
-\begin{Coqsrc}
-# let exp2 x = 2.0 ** x;;
-
-val exp2 : float -> float = <fun>
-#   exp2 95.0 *. 97.0 -. 1.0;;
-- : float = 3.84256588194182037e+30
-# let n = exp2 95.0 ;;
-# let p = n *. 97.0 -. 1.0;;
-val p : float = 3.84256588194182037e+30
-
-Estimation :
-2 ** (3.84 e+30) * 3.84 e+30.
-\end{Coqsrc}
-
-\subsection{Using \texttt{Equations}}
-\label{sect:L-equations}
-
-Note that we did not define any function $L_\alpha$ \emph{for any $\alpha<\epsilon_0$} yet.
-We have got no more than a collection of proved realizations of $\texttt{L\_spec}\;\alpha$ for a few values of $\alpha$.
-
-\index{coq}{Plug-ins!Equations}
-
-Using the \texttt{coq-equations} plug-in by 
-M. Sozeau~\cite{sozeau:hal-01671777}, we will now define a function \texttt{L\_} which maps any ordinal  $\alpha<\epsilon_0$ to a proven realization of 
-$\texttt{L\_spec}\;\alpha$.   
-To this end, we represent ordinals as inhabitants of the type 
-\texttt{E0} of well-formed ordinal terms (see Sect~\vref{sect:E0-def}). So, we define a total function \texttt{L\_} of type
-\texttt{E0 -> nat -> nat}, by transfinite recursion, considering the usual three cases : $\alpha=0$, $\alpha$ is a successor, $\alpha$ is a limit ordinal.
-
-\subsubsection{Definition}
-
-\vspace{4pt}
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#L_}{L\_alpha}).}
-
-\label{Functions:L-alpha}
- \index{hydras}{Library Epsilon0!Functions!L\_@L\_ (final step of a minimal path}
-
-\input{movies/snippets/L_alpha/LDef}
-
-This definition results in a bunch of automatically generated lemmas. For instance:
-
-\input{movies/snippets/L_alpha/AboutLEquation1}
-
-In most cases, it may be useful to write human-readable  paraphrases of these statements.
-
-\inputsnippets{L_alpha/Paraphrasesa}
-\inputsnippets{L_alpha/Paraphrasesb}
-\inputsnippets{L_alpha/Paraphrasesc}
-
-Using these three lemmas as rewrite rules, we can prove more properties of the functions \texttt{L\_$\alpha$}.
-
-\input{movies/snippets/L_alpha/LFiniteOmega}
-
-By  well-founded induction on $\alpha$, we prove the following properties:
-
-\input{movies/snippets/L_alpha/LGeS}
-
-\input{movies/snippets/L_alpha/LCorrect}
-
-\label{sect:L-correct-proof}
-
-Please note that the proof of \texttt{L\_correct} applies the lemmas proven in Sections~\ref{sect:L-spec-zero}, ~\ref{sect:L-spec-succ} and ~\ref{sect:L-spec-lim}.
-Our previous study of \texttt{L\_spec} allowed us to pave the way for the definition by \texttt{Equations} and the correctness proof.
-
-\paragraph*{\gaiasign}
-The module
-\href{../theories/html/gaia_hydras.GL_alpha.html}{gaia\_hydras.GL\_alpha} contains an adaptation of
-\href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0/L\_alpha}.
-
-\subsubsection{Back to hydra battles}
-\label{def:L-alpha}
-
-Lemma \texttt{battle\_length\_std } of
-Module~\href{../theories/html/hydras.Hydra.Battle_length.html}{Hydra.Battle\_length} relates the length of standard battles with the functions $L_\alpha$.
-
-\input{movies/snippets/Battle_length/battleLengthStd}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-Instead of considering standard paths and battles, consider battles where the replication factor is a constant $k$.
-Please use \texttt{Equations} in order to define the function that computes the length of the $k$-path which leads  from $\alpha$ to $0$.
-Prove a few  exact formulas and minoration lemmas.
-\end{exercise}
-
-\section{A variant of the Hardy hierarchy}
-
-\label{sect:hardy}
-
-In order to give a feeling on  the complexity of the functions  $L_\alpha$s, we compare them with a better known family of functions, the  \emph{Hardy hierarchy} of fast growing functions,
-presented for instance in~\cite{Promel2013}. 
-\index{maths}{Rapidly growing functions!Hardy Hierarchy}
-
-\begin{remark}
-  Indeed, the functions presented in this section are a \emph{variant} of the Hardy hierarchy of functions.
-In the future versions of this development, we will correct the references to the literature.
-For the time being, we call our functions $H'_\alpha$ in order to underline the difference from ``classic'' Hardy functions.
-\end{remark}
-
-For each ordinal $\alpha$ below $\epsilon_0$, $H'_\alpha$ is 
-a total arithmetic function, defined  by  transfinite recursion on $\alpha$, according to three cases:
-
-\index{maths}{Transfinite induction}
-
-\begin{itemize}
-\item If $\alpha=0$, then $H'_\alpha (k)= k$ for any natural number $k$.
-\item If $\alpha=\textrm{succ}(\beta)$, then 
-$H'_\alpha(k)=H'_\beta(k+1)$ for any $k \in \mathbb{N}$
-\item If $\alpha$ is a limit ordinal, then 
-$H'_\alpha(k) = H'_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
-\end{itemize}
-
-\begin{remark}
- The ``classic'' definition of the Hardy hierarchy differs in the third equation.
-
-\begin{itemize}
-\item If $\alpha=0$, then $H_\alpha (k)= k$ for any natural number $k$.
-\item If $\alpha=\textrm{succ}(\beta)$, then 
-$H_\alpha(k)=H_\beta(k+1)$ for any $k \in \mathbb{N}$
-\item If $\alpha$ is a limit ordinal, then 
-$H_\alpha(k) = H_{(\canonseq{\alpha}{k})}(k)$ for any $k\in \mathbb{N}$.
-\end{itemize}
-
-\end{remark}
-
-\subsection{Definition in \texttt{Coq}}
-
-We define a function \texttt{H'\_} of type \texttt{E0 -> nat -> nat} by transfinite induction over the type \texttt{E0} of the well formed ordinals below $\epsilon_0$.
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}}
-
-\index{hydras}{Library Epsilon0!Functions!H\_@H\_ (Hardy hierarchy (variant))}
-\index{coq}{Plug-ins!Equations}
-\label{Functions:Hprime-alpha}
-
-\input{movies/snippets/Hprime/HprimeDef}
-\input{movies/snippets/Hprime/HprimeDefb}
-
-\input{movies/snippets/Hprime/paraphrasesa}
-\input{movies/snippets/Hprime/paraphrasesb}
-\input{movies/snippets/Hprime/paraphrasesc}
-\input{movies/snippets/Hprime/paraphrasesd}
-
-\subsection{First  steps of the H' hierarchy}
-Using rewrite rules from \texttt{H'\_eq1} to \texttt{H'\_succ\_eqn}, we can explore the functions $H'_\alpha$ for  small values of $\alpha$.
-
-\paragraph*{\gaiasign}
-The module
-\href{../theories/html/gaia_hydras.GHprime.html}{gaia\_hydras.GHprime} contains an adaptation of
-\href{../theories/html/hydras.Epsilon0.Hprime.html}{Epsilon0.Hprime}.
-
-\subsubsection{Finite ordinals}
-
-By induction on $i$, we prove a simple expression of \texttt{H'\_ (Fin i)}, where 
-\texttt{Fin $i$}  is the $i$-th finite ordinal.
-
-\input{movies/snippets/Hprime/HprimeFin}
-
-\subsubsection{Multiples of \texorpdfstring{$\omega$}{omega}}
-
-Since the canonical sequence of $\omega$ is composed of finite ordinals, 
-it is easy to get the formula associated with $H'_\omega$.
-
-\input{movies/snippets/Hprime/HprimeOmega}
-
-Before going further, we prove a useful rewriting lemma:
-
-\input{movies/snippets/Hprime/HprimePlusFin}
-
-Then, we get easily formulas for $H'_{\omega+i}$, and $H'_{\omega\times i}$ for any natural number $i$.
-
-\input{movies/snippets/Hprime/HprimeExamplesa}
-\input{movies/snippets/Hprime/HprimeExamplesb}
-\input{movies/snippets/Hprime/HprimeExamplesc}
-\input{movies/snippets/Hprime/HprimeExamplesd}
-
-Since $\omega^2$ is a limit ordinal, we prove the following equality: 
-\[H'_{\omega^2} (k) = 2 ^ {k+1} \times (k+1) - 1\]
-
-\input{movies/snippets/Hprime/HprimeOmegaSqr}
-
-\subsubsection{New limits}
-
-Our next step would be to prove an exact formula for $H'_{\omega^\omega}(k)$.
-Since the canonical sequence of $\omega^\omega$ is composed of all the
-$\omega^i$, we first need to express $H'_{\omega^i}$ for any natural number $i$.
-
-Let $i$ and $k$ be two natural numbers. 
-The ordinal $\canonseq{\omega^(i+1)}{k}$ is the product
-$\omega^i \times k$, so we need also to consider ordinals of this form.
-
-\begin{enumerate}
-\item First,  we express $H'_{\omega^\alpha \times (i+2)}$ in terms of
-$H'_{\omega^\alpha \times (i+1)}$.
-
-\input{movies/snippets/Hprime/HprimeOmegaTerm1}
-
-\item
-Then, we prove by induction on $i$ that $H'_{\omega^\alpha \times (i+1)}$ is just the
-$(i+1)$-th iterate of $H'_{\omega^\alpha}$.
-
-\input{movies/snippets/Hprime/HprimeOmegaTerm}
-
-\item In particular, we derive a formula for $H'_{\omega^{i+1}}$.
-
-\input{movies/snippets/Hprime/HprimeSuccFun}
-\input{movies/snippets/Hprime/HprimePhi0SI}
-
-\item We get now a  formula for $H'_{\omega^3}$:
-
-\input{movies/snippets/Hprime/HprimeOmegaCube}
-\end{enumerate}
-
-\subsubsection{A numerical example}
-\label{sect:bignum}
-
-It is hard to capture the complexity of this function by looking only at this
-``exact'' formula. 
-Let us consider a simple example: the number $H'_{\omega^3}(3)$.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3a}
-
-Thus, the number $H_{\omega^3}(3)$ can be written as four nested applications of $f$.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3b}
-
-In a more classical writing, this number is displayed as follows:
-
-{\Large
-$$
-H'_{\omega^3}(3) =  2 ^ {(2 ^ {N + 1} \, (N+1) )}   \,  (2 ^ {N+1} \, ( N +1) ) - 1
-$$
-}
-
-In order to make this statement more readable, we introduce a local definition.
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3c}
-
-This number looks quite big; let us compute an approximation with \texttt{Ocaml}:
-
-\begin{Coqsrc}
-# (2.0 ** 64.0 *. 64.0 -. 1.0);; 
-\end{Coqsrc}
-
-\begin{Coqanswer}
-- : float = 1.1805916207174113e+21
-\end{Coqanswer}
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3d}
-
-We leave as an exercise to determine the best approximation as possible of
- the size of this number (for instance its number of digits).  For instance, if
-we do not take into account the multiplications in the formula above,
-we obtain that, in base $2$, the number $H'_{\omega^3}(3)$ has at least
-$2^{10^{21}}$  digits. But it is still an under-approximation !
-
-\input{movies/snippets/Hprime/HprimeOmegaCube3e}
-
-\subsubsection{A formula for \texorpdfstring{$H'_{\omega^\omega}$}{\texttt{H'(phi0(omega))}}}
-
-Now, we can get at last an exact formula for $H'_{\omega^\omega}$.
-
-\input{movies/snippets/Hprime/HprimePhi0Omega}
-
-Using extensionality of the functional \texttt{iterate}, we also get a closed formula.
-
-\input{movies/snippets/Hprime/HprimePhi0OmegaClosed}
-
-Note that this formula contains two occurrences of the functional \texttt{iterate}, the outer one is in fact a second-order iteration (on type \texttt{nat -> nat)}
-and the inner one  first-order (on type \texttt{nat}).
-
-\subsection{Abstract properties of  
-\texorpdfstring{$H'_\alpha$}{H'}}
-~\label{sect:H-alpha-prop}
-
-Since pure computation seems to be useless for dealing with expressions of the form $H'_\alpha(k)$, even for small values of $\alpha$ and $k$, we need to prove theorems for comparing $H'_\alpha(k)$ and $H'_\beta(l)$, in terms of comparison
-between $\alpha$ and $\beta$ on the one hand, $k$ and $l$ on the other hand.
-
-But beware of fake theorems! For instance, one could believe that $H'$ is monotonous in its first argument. The following proof shows this is false.
-
-\input{movies/snippets/Hprime/HprimeNonMono1}
-
-On the contrary, the functions of the $H'$ hierarchy have the following five properties~\cite{KS81}: for any $\alpha < \epsilon_0$,
-\begin{itemize}
-\item the function $H'_\alpha$ is strictly monotonous :
-      For all $n,p \in\mathbb{N}, n < p \Rightarrow H'_\alpha(n)< H'_\alpha(p)$.
-\item If $\alpha \not= 0$, then for every $n$, $n<H'_\alpha(n)$.
-\item The function $H'_\alpha$ is pointwise less or equal than $H'_{\alpha+1}$
-
-\item For any $n\geq 1$, $H'_\alpha(n)<H'_{\alpha+1}(n)$.
-\emph{We say that $H'_{\alpha+1}$ dominates $H'_\alpha$ from $1$}.
-\item For any $n$ and $\beta$, if $\alpha \xrightarrow[n]{} \beta$, then
-$H'_\beta(n)\leq H'_\alpha(n)$.
-\end{itemize}
-
-\index{maths}{Abstract properties of arithmetic functions}
-\index{hydras}{Abstract properties of arithmetic functions}
-
-\index{maths}{Transfinite induction}
-
-In \coq{}, we follow the  proof of~\cite{KS81}. This proof is mainly a single  proof by transfinite induction on $\alpha$ of the conjunction of the five properties.
-For each $\alpha$, the three cases : $\alpha=0$, $\alpha$ is a limit, and 
-$\alpha$ is a successor are considered. Inside each case, the five sub-properties are proved sequentially, using the abstract properties defined in Sect~\vref{sect:abstract-arith-prop}
-
-\input{movies/snippets/Hprime/PAP}
-
-Using a few lemmas \emph{à la} Ketonen-Solovay, we prove that
-if $\alpha<\beta$, then $H'_\beta$ eventually dominates
-$H'_\alpha$.
-We let the reader look at the proof (Section \texttt{Proof\_of\_H'\_mono\_l} of \href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}).
-
-\noindent
-\input{movies/snippets/Hprime/HprimeDom}
-
-%ICI
-\subsection{Comparison between \texttt{L\_} and \texttt{H'\_} }
-
-By well-founded induction on $\alpha$, we prove that
- $H'_\alpha(i) \leq L_\alpha(i+1)$ for any $i$.
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#H'_L_}{Epsilon0.L\_alpha}}
-
-\input{movies/snippets/L_alpha/HprimeL}
-
-\subsubsection{Back to hydras}
-
-The following theorem relates the length of (standard) battles with the the $H'$ family of fast growing functions.
-
-\vspace{4pt}
-
-\noindent
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems}}
-
-\input{movies/snippets/Hydra_Theorems/battleLengthStdHardy}
-
-Since $H'_{\omega^3}(3)= H'_{\omega^3+3}(0)$, the big number shown in Section~\vref{sect:bignum} is less or equal than 
-the number of steps of the battle with the initial hydra of
-Figure~\ref{fig:start}, and  $0$ as initial replication factor.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}[very thick, scale=0.3]
-  \node (h1) at (4,0){$\bullet$};
-  \node (h2) at (2,2){$\bullet$};
-  \node (h3) at (0,4){$\bullet$};
-  \node (h4) at (1,4){$\bullet$};
-  \node (h5) at (2,4){$\bullet$};
-  \node (h6) at (4.5,2){$\bullet$};
-  \node (h7) at (6.5,2){$\bullet$};
-  \node (h8) at (8.5,2){$\bullet$};
-    \draw (h1) -- (h2) ;
-    \draw (h2) -- (h3) ;
-    \draw (h2) -- (h4) ;
-    \draw (h2) -- (h5) ;
-    \draw (h1) -- (h6) ;
-    \draw (h1) -- (h7) ;
-    \draw (h1) -- (h8);
- \end{tikzpicture}
-
-  \caption{The hydra corresponding to the ordinal $\omega^3+3$}
-  \label{fig:start}
-\end{figure}
-
-\section{A variant of the Wainer hierarchy (functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
-\label{sect:wainer}
-
-\index{maths}{Rapidly growing functions!Wainer Hierarchy}
-
-Ketonen and Solovay introduce in~\cite{KS81} a ``trivial'' variant of the Wainer hierarchy~\cite{BW85, Wainer1970} of fast growing functions, indexed by ordinals below $\epsilon_0$.
-The functions $F_\alpha$ are defined by the following equations.
-
-\label{F_equations}
-\begin{itemize}
-\item $F_0(i)=i+1$
-\item $F_{\beta+1}(i)= (F_\beta)^{(i+1)}(i)$, where $f^{(i)}$ is the $i$-th iterate of $f$.
-\item $F_\alpha(i) = F_{\canonseq{\alpha}{i}} (i)$ if $\alpha$ is a limit ordinal.
-\end{itemize}
-
-\begin{remark}
-The difference with the ``classic'' Wainer hierarchy 
-$f_\alpha\;(\alpha<\epsilon_0)$ lies in the second equation:
-$f_{\beta+1}(i) = (f_\beta)^{(i)}(i)$ and not
-$f_{\beta+1}(i) = (f_\beta)^{(i+1)}(i)$.
-
-A module about 
-the classic Wainer hierarchy is in preparation.
-
-Note also that \cite{KS81} defines also $F_{\epsilon_0}$ (by the third equation).
-Since $\epsilon_0$ is not representable in type \texttt{E0}, our implementation in \coq{} does not take $F_{\epsilon_0}$ into account.
-
-\end{remark}
-
-A first attempt is to write a definition of $F_\alpha$ by equations, in the same way as for $H_\alpha$ (the functional \texttt{iterate} has already been used in Sect.\vref{Functions:iterate}).
-
-\index{hydras}{Library Prelude!iterate}
-
-\input{movies/snippets/Iterates/iterateDef}
-
-\emph{From\href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}~}.
-
-\index{coq}{Plug-ins!Equations}
-
- \input{movies/snippets/F_alpha/FailDemo}
-
-We presume that this error comes from the recursive call of \texttt{F\_} inside
-an application of \texttt{iterate}. The workaround we propose is to define first 
-the iteration of \texttt{F\_}  as an helper $F^*$, then to define the function $F$ as a ``iterating $F^*$ once''.
-
-\texttt{Equations} accepts the following definition, relying on  lexicographic ordering on pairs $(\alpha,n)$.
-
-\label{sect:F-equations}
-
-\index{coq}{Plug-ins!Equations}
-\label{Functions:F-alpha}
-\index{maths}{Rapidly growing functions}
-\index{hydras}{Library Epsilon0!Functions!F\_@F\_ (Wainer hierarchy)}
-
-\input{movies/snippets/F_alpha/goodDefa}
-\input{movies/snippets/F_alpha/goodDefb}
-
-It is quite easy to prove that our functional \texttt{F\_} satisfies the equations on page~\pageref{sect:F-equations}.
-\index{hydras}{Library Prelude!iterate}
-
-\input{movies/snippets/F_alpha/FEquations}
-
-As for the Hardy functions, we can use these equalities as rewrite rules for
-``computing'' some values of $F_\alpha(i)$, for small values of $\alpha$.
-
-\input{movies/snippets/F_alpha/FirstValues}
-
-Like in Sect~\ref{sect:H-alpha-prop}, we prove by induction the following properties (see~\cite{KS81}).
-
-\input{movies/snippets/F_alpha/FalphaThms}
-
-As a corollary, we prove that, if $\beta<\alpha$, then
-$F_\alpha$ dominates $F_\beta$ (p. 284 of~\cite{KS81}).
-
-\input{movies/snippets/F_alpha/FDomContext}
-\input{movies/snippets/F_alpha/FDom}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove the following property:
-
-\begin{Coqsrc}
-Lemma LF3 : dominates_from  2 (F_ 3) (fun  n => iterate exp2 n n).
-\end{Coqsrc}
-
-\emph{You may start this exercise with the file
-\href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
-\end{exercise}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-Prove that, for any $\alpha\geq 3$ and $n\geq 2$,
-$F_\alpha(1+n)\geq 2^{F_\alpha(n)}$.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_3.v}{exercises/ordinals/F\_3.v}.}
-\end{exercise}
-
-%ici
-\index{hydras}{Exercises}
-
-\begin{exercise}
-It is tempting to prove a simple property of monotony 
-of the function \texttt{F\_}.
-
-\begin{quote}
-   Let $\alpha\leq\beta<\epsilon_0$. For any $n\geq 2$,
-$F_\alpha(n)\leq F_\beta(n)$. 
-\end{quote}
-Prove or disprove this statement.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/is_F_monotonous.v}{exercises/ordinals/is\_F\_monotonous.v}.}
-\end{exercise}
-
-\index{hydras}{Exercises}
-\begin{exercise}
-
-Prove that for any $n\geq 2$, $\textrm{Ack}\,\,n\,n\leq  F_\omega(n)$, where \textrm{Ack} is the Ackermann function.
-Next, prove that $F_\alpha$ is not primitive recursive, for any $\alpha\geq\omega$  (please see Sect.~\vref{sect:ack-not-PR}).
-On the other hand, please show that for any natural number $n$, the function $F_n$ is primitive recursive.
-Thus $F_\alpha$ is primitive recursive if and only if $\alpha$ is finite.
-
-\emph{You may start this exercise with the file
-    \href{https://github.com/coq-community/hydra-battles/tree/master/exercises/ordinals/F_omega.v}{exercises/ordinals/F\_Omega.v}.
-Properties of the Ackermann function are studied in
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/Ack.v}{theories/ordinals/MoreAck/Ack.v} and
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/MoreAck/AckNotPR.v}{theories/ordinals/MoreAck/AckNotPR.v}
-.}
-\end{exercise}
-
-\paragraph*{\gaiasign} Library~\href{../theories/html/gaia_hydras.GF_alpha.html}{gaia\_hydras.GF\_alpha}  defines the functions $F_\alpha$ for \gaia's world.
-
-\section{More about rapidly growing functions}
-
-In Sect.~\vref{sect:battle-length-notPR}, we prove that the length of hydra-battles (for a given hydra, according to the initial replication factor) is not primitive recursive in general. 
-This proof uses properties of the Ackermann function, and the $H'_\alpha$, $F_\alpha$, $L_\alpha$ families of functions.
-
-% ---------------------------------------------------
-\input{gaia-chapter}
-
-%----------------------------------------------------------------------
 \chapter[Countable ordinals (after Sch\"{u}tte)]{Kurt Schütte's axiomatic definition of countable ordinals}
@@
 the \href{https://github.com/coq-community/hydra-battles/blob/master/theories/gaia/}{theories/gaia/} directory.
-
-\chapter{The Ordinal \texorpdfstring{$\Gamma_0$}{Gamma0} (first draft)}
-
-\emph{This chapter and the files it presents are still very incomplete, considering the impressive properties of $\Gamma_0$~\cite{Gallier91}.
-We hope to add new material soon, and accept contributions!}
-
-\begin{todo}
-  Build a T2Bridge! (note: Gamma0's wellfoundedness is missing in \gaia (?).
-\end{todo}
-
-\section{Introduction}
-We present a notation system for the ordinal $\Gamma_0$, following Chapter V, Section 14 of~\cite{schutte}: ``A notation system for the ordinals $<\Gamma_0$''.
-We try to be as close as possible to Schütte's text and usual practices of \coq{} developments.
-
-The ordinal $\Gamma_0$ is defined in Section 13 of ~\cite{schutte} as the least \emph{strongly critical ordinal}. It is widely known as the \emph{Feferman-Schütte ordinal}.
-
-Section V, 13 of~\cite{schutte} defines \emph{strongly critical} and
-\emph{maximal $\alpha$-critical} ordinals:
-
-\begin{itemize}
-\item $\alpha$ is strongly critical if
-$\alpha$ is $\alpha$-critical,
-\item $\gamma$ is maximal $\alpha$-critical if $\gamma$ is $\alpha$-critical, and, for all $\xi>\alpha$, $\gamma$ is not $\xi$-critical.
-
-\end{itemize}
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Schutte.Critical.html\#strongly_critical}%
-{\texttt{Schutte.Critical}}}
-
-\input{movies/snippets/Critical/Gamma0Def}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that a (countable)  ordinal $\alpha$ is strongly critical iff 
-$\phi_\alpha(0)=\alpha$ (Theorem 13.13 of~\cite{schutte} ). 
-\end{project}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that the set of strongly critical ordinals is unbounded and closed (Theorem 13.14 of~\cite{schutte} ).
-Thus this set is not empty,  hence has a least element.
-Otherwise, the definition of $\Gamma_0$ above would be useless.
-\end{project}
-
-In the present version of this development, we  only study $\Gamma_0$ as a notation system, much more powerful than the ordinal notation for $\epsilon_0$.
-
-%\index{Projects}
-%
-%\begin{project}
-% Schûtte's section 13 of~\cite{schutte} contain several definitions and lemmas which give another view on $\Gamma_0$.  We leave it as a project to implement them in \coq{}.  
-%\end{project}
-
-\section{The type \texttt{T2} of ordinal terms}
-
-The notation system for ordinals less than $\gamma_0$ comes from the following theorem of~\cite{schutte}, where $\psi\,\alpha$ is the ordering function 
-of the set of maximal $\alpha$-critical ordinals.
-
-\index{hydras}{Library Gamma0!Types!T2}
-
-\begin{quote}
-  Any ordinal $\not= 0$ which is not strongly critical can be expressed in terms of $+$ and $\psi$.
-\end{quote}
-
-\index{hydras}{Projects}
-\begin{project}
-This theorem is not formally proved in this development yet. It should be!
-\end{project}
-
-Like in Chapter~\ref{chap:T1}, we define an inductive type with two constructors, one for $0$, the other for the construction $\psi(\alpha,\beta)\times(n+1)+\gamma$, adapting a Manolios-Vroon-like notation~\cite{Manolios2005} to
-\emph{Veblen normal forms}.
-\label{types:T2}
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T2}%
-{\texttt{Gamma0.T2}}}
-
-\inputsnippets{T2/T2Def, T2/psiDef}
-
-Like in chapter~\ref{chap:T1}, we get familiar with the type \texttt{T2} by recognizing simple constructs like finite ordinals, $\omega$, etc., as inhabitants of \texttt{T2}.
-
-\input{movies/snippets/T2/Notations}
-
-\section{A strict order on T2}
-
-Let us define a strict order on type \texttt{T2}. The following definition is 
-an adaptation of Schütte's, taking into account the multiplications by a natural number (inspired by~\cite{Manolios2005}, and also present in \texttt{T1}).
-
-\label{sect:t2-lt-def}
-
-\input{movies/snippets/T2/ltDef}
-
-Seven constructors! In order to get accustomed with this definition, let us look at a small set of examples, covering all the constructors of \texttt{lt}.
-
-\input{movies/snippets/T2/ltExamples}
-
-\index{hydras}{Projects}
-\begin{project}
-Write a tactic that solves automatically goals of the form (\texttt{$\alpha$ t2< $\beta$}), where $\alpha$ and $\beta$ are closed terms of type \texttt{T2}.
-\end{project}
-
-\section{Veblen normal form}
-\begin{definition}
-  A term of the form $\psi(\alpha_1,\beta_1)\times n_1+ \psi(\alpha_2,\beta_2)\times n_2+\dots+\psi(\alpha_k,\beta_k)\times n_k$ is said to be in
- \emph{[Veblen] normal form} if for every $i<n$, $\psi(\alpha_i,\beta_i)<\psi(\alpha_{i+1},\beta_{i+1})$, all the $\alpha_i$ and $\beta_i$ are in normal form, and all the $n_i$ are strictly positive integers.
-\end{definition}
-
-\input{movies/snippets/T2/nfDef}
-
-Let us look at some positive examples (we have to prove some inversion lemmas before proving counter-examples).
-
-\input{movies/snippets/T2/nfExamples}
-
-\subsection{Length of a term}
-
-The notion of \emph{term length} is introduced by Schütte as a helper for proving (at least) the \emph{trichotomy} property and transitivity of the strict order \texttt{lt} on \texttt{T2}.
-These properties are proved by induction on length.
-
-\subsection{Trichotomy}
-
-\emph{Trichotomy} is another name for the well-known property of decidable total ordering (like Standard Library's \texttt{Compare\_dec.lt\_eq\_lt\_dec}).
-
-We first prove by induction on $l$ the following lemma:
-
-\vspace{4pt}
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0\#tricho_aux}%
-{\texttt{Gamma0.Gamma0}}}
-
-\input{movies/snippets/Gamma0/tricho}
-
-\input{movies/snippets/Gamma0/compare}
-
-With the help of \texttt{compare}, we get a boolean version of \texttt{nf}
-(being in Veblen normal form).
-
-\input{movies/snippets/Gamma0/nfb}
-
-\section{Main functions on \texttt{T2}}
-
-\subsection{Successor}
-The successor function is defined by structural recursion.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#succ}%
-{\texttt{Gamma0.T2}}}
-
-\input{movies/snippets/T2/succDef}
-
-\subsection{Addition}
-
-Like for Cantor normal forms (see Sect.~\ref{sect:infix-plus-T1}),  the definition of addition in \texttt{T2}  requires comparison between ordinal terms.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.Gamma0T2.html\#succ}%
-  {\texttt{Gamma0.Gamma0}}}
-
-\input{movies/snippets/Gamma0/plusDef}
-
-\subsection{The Veblen function \texorpdfstring{$\phi$}{\texttt{phi}}}
-
-The enumeration function of critical ordinals, presented in Sect.~\vref{sect:phi-schutte}, is recursively defined in type \texttt{T2}.
-
-\input{movies/snippets/Gamma0/phiDef}
-
-Despite its complexity, the function \texttt{phi} is well adapted to proofs by simplification or computation.
-
-The relation between the constructor $\psi$ and the function $\phi$ is
-studied in~\cite{schutte}, and partially implemented in this development.
-\emph{Please contribute!}
-
-For instance, the following theorem states that, if $\gamma$ is the sum of a limit ordinal $\beta$ and a finite ordinal $n$, and $\beta$ is a fixpoint of
-$\phi(\alpha)$, then $\psi(\alpha,\gamma)=\phi_\alpha(\gamma+1)$.
-
-\input{movies/snippets/Gamma0/phiPsi}
-\input{movies/snippets/Gamma0/Ex9}
-
-On the other hand, $\phi$ can be expressed in terms of $\psi$.
-
-\input{movies/snippets/Gamma0/phiOfPsi}
-\input{movies/snippets/Gamma0/Ex10}
-
-\index{hydras}{Projects}
-\begin{project}
-Please study a way to pretty print ordinal terms in Veblen normal form (see Section~\vref{sect:ppT1}).
-\end{project}
-
-\section{An ordinal notation for \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}}
-
-In order to consider type \texttt{T2} as an ordinal notation, we have to build an instance of class \texttt{ON} (See Definition page~\pageref{types:ON}).
-
-First, we define a type that contains only terms in Veblen normal form, and redefine \texttt{lt} and \texttt{compare} by delegation (see for comparison the construction of type \texttt{E0} in Sect.~\vref{sect:E0-def}).
-
-\input{movies/snippets/Gamma0/G0a}
-\input{movies/snippets/Gamma0/G0b}
-
-Then, we buils an instance of class \texttt{ON}. 
-function \texttt{compare} is correct.
-
-\input{movies/snippets/Gamma0/ltSto}
-\input{movies/snippets/Gamma0/ltWf}
-\input{movies/snippets/Gamma0/ONGamma0}
-\input{movies/snippets/Gamma0/G0z}
-
-\begin{remark}
-The proof of \texttt{lt\_wf} has been written by \'Evelyne Contejean, using her library on the recursive path ordering (see also remark~\vref{remark:a3pat}).
-\end{remark}
-
-\index{hydras}{Projects}
-\begin{project}
-Prove that \texttt{Epsilon0} (page~\pageref{instance-epsilon0})
-is a sub-notation system of \texttt{Gamma0}.
-
-Prove that the implementations of \texttt{succ}, \texttt{+}, $\phi_0$, etc.
-are compatible in both notation systems.
-
-Note that a function \texttt{T1\_inj} from \texttt{T1} to \texttt{T2} has already been defined. It may help to complete the task.
-
-\noindent\emph{From \href{../theories/html/hydras.Gamma0.T2.html\#T1_to_T2}%
-{\texttt{Gamma0.T2}}}
-
-\input{movies/snippets/T2/T1ToT2}
-
-\end{project}
-
-\begin{project}
-Prove that the notation system \texttt{Gamma0} is a correct implementation 
-of the segment $[0,\Gamma_0)$ of the set of countable ordinals.
-\end{project}

### 2022-05-08 doc/JFLA2022/Slides/Slides.tex
changes to JFLA slides

@@
 
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+  \begin{figure}[ht]
+\centering
+\scalebox{0.7}{
+\begin{tikzpicture}[thick]
+
+  \begin{scope}[yshift=2.0cm, xshift=1.7cm] 
+     \node[ellipse, dotted, fill=orange!20, minimum height=0.5cm,minimum width=1.5cm] (settheory) { Set theory\footnote{N. Bourbaki}};
+      \node[ellipse, dotted, fill=orange!20, minimum height=0.5cm,minimum width=1.5cm, left of=settheory, node distance=8.4cm] (prooftheory) { Proof theory\footnote{K.
+Schutte, J.
+Ketonen, R.
+Solovay, L.
+Kirby \& J.
+Paris}};
+\end{scope}
+
+\begin{scope}[yshift=1.0cm, xshift=-1.5cm]
+   \node[ellipse, dotted, draw= black!20,fill=green!20,minimum height=0.5cm,minimum width=1.5cm] (acl2) { [ACL2]\footnote{P. Manolios \&  D. Vroon}};   
+\end{scope}
+
+\begin{scope}[xshift=-6.1cm]
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel\footnote{R. O'Connor} };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (additions) { Additions };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (cantor) { Cantor\footnote{C. \& E. Contejean} };
+
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3.2cm] (cats) { Categories in ZFC };
+\end{scope}
+
+\begin{scope}[yshift=-1cm,xshift=0.45cm]
+  \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=2.5cm] (mathcomp) { MathComp };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=paramcoq, node distance=4.1cm] (equations) { Equations };
+\end{scope}
+
+\begin{scope}[yshift=-2cm,xshift=-4cm]
+\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
+\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,left of=hydras, node distance=3.5cm] (pockcc) { Pocklington };
+\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=hydras, node distance=2.5cm] (chains) { Addition-chains };
+\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,right of=chains, node distance=2.2cm] (gaia) { Gaia\footnote{J. Grimm} };
+\end{scope}
+
+\begin{scope}[yshift=-3cm,xshift=-6cm]
+ \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
+ \node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=3cm] (gaiahydras) { Gaia-hydras };
+\end{scope}
+
+\draw[lookcolor,->,dotted] (acl2.south) -- (cantor.north) ;
+\draw[lookcolor,->,dashed] (additions.south) -- (chains.north) ;
+\draw[lookcolor,->,dashed] (goedel) -- (goedelcc) ;
+\draw[lookcolor,->,dashed] (goedel) -- (hydras.north west) ;
+\draw[lookcolor,->,dashed] (cantor) -- (hydras.north east) ;
+\draw[lookcolor,->,dashed] (cantor.south) -- (gaia.north west) ;
+\draw[lookcolor,->,dashed] (pock) -- (pockcc) ;
+\draw[lookcolor,->,dashed] ([xshift=2.1cm]cats) -- (gaia) ;
+
+\draw[->] (hydras) -- (gaiahydras) ;
+\draw[->] (hydras.south west) -- (goedelcc) ;
+\draw[->] (gaia.south west) -- (gaiahydras.east) ;
+
+\draw[->] (equations) -- (hydras) ;
+\draw[->] (mathcomp) -- (gaia) ;
+\draw[->] (mathcomp) -- (chains) ;
+\draw[->] (paramcoq.south) -- (chains) ;
+\draw[->] (pockcc) -- (goedelcc) ;
+\draw[dotted, ->] (prooftheory) -- (hydras);
+\draw[dotted, ->] (settheory) -- (cats);
+
+\end{tikzpicture}
+}
+\caption{Genealogy and dependencies for \Hydras packages.}
+  \label{fig:genealogy}
+\end{figure}
+\end{frame}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{frame}
-  \frametitle{Yet another book on \coq?}
-  \begin{table}
-  \centering
-  \footnotesize
-  \begin{tabular}{lll}
-  \hline
-  \textbf{Title} & \textbf{Year} & \textbf{Category}\\
-    \hline
-   \coq reference manual &  $\geq 1989$ & \\
-  Coq'Art & 2004 & Traditional\\
-  Software Foundations & 2007 & Executable\\
-  Certified Programming with Dependent Types & 2011 &  Executable\\
-  Program Logics for Certified Compilers & 2014 & Traditional\\
-  Programs and Proofs & 2014 & Executable\\
-  Formal Reasoning About Programs& 2017 & Traditional\\
-  Computer Arithmetic and Formal Proofs & 2017 & Traditional\\
-    Mathematical Components & 2018 & Traditional\\
-  \hline
-  \end{tabular}
-  \caption{Coq book categorization.}
-  \label{tbl:books}
-\end{table}
+% \begin{frame}
+%   \frametitle{Yet another book on \coq?}
+%   \begin{table}
+%   \centering
+%   \footnotesize
+%   \begin{tabular}{lll}
+%   \hline
+%   \textbf{Title} & \textbf{Year} & \textbf{Category}\\
+%     \hline
+%    \coq reference manual &  $\geq 1989$ & \\
+%   Coq'Art & 2004 & Traditional\\
+%   Software Foundations & 2007 & Executable\\
+%   Certified Programming with Dependent Types & 2011 &  Executable\\
+%   Program Logics for Certified Compilers & 2014 & Traditional\\
+%   Programs and Proofs & 2014 & Executable\\
+%   Formal Reasoning About Programs& 2017 & Traditional\\
+%   Computer Arithmetic and Formal Proofs & 2017 & Traditional\\
+%     Mathematical Components & 2018 & Traditional\\
+%   \hline
+%   \end{tabular}
+%   \caption{Coq book categorization.}
+%   \label{tbl:books}
+% \end{table}
 
-\end{frame}
+% \end{frame}
 
@@
 \begin{frame}
-   \begin{block}{Example-driven books}
+   \begin{block}{An example-driven book}
     \begin{itemize}
@@
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-  \begin{figure}[ht]
-\centering
-\scalebox{0.7}{
-\begin{tikzpicture}[thick]
-
-  \begin{scope}[yshift=1.5cm, xshift=-1.5cm]
-\node[ellipse, dotted, draw= black!20,fill=green!20,minimum height=0.5cm,minimum width=1.5cm] (acl2) { [ACL2]\footnote{P. Manolios \&  D. Vroon}};
-\end{scope}
-
-\begin{scope}[xshift=-6cm]
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel\footnote{R. O'Connor} };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (additions) { Additions };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (cantor) { Cantor\footnote{C. \& E. Contejean} };
-
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3.2cm] (cats) { Categories in ZFC };
-\end{scope}
-
-\begin{scope}[yshift=-1cm,xshift=0.45cm]
-  \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=2.5cm] (mathcomp) { MathComp };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=paramcoq, node distance=4.5cm] (equations) { Equations };
-\end{scope}
-
-\begin{scope}[yshift=-2cm,xshift=-4cm]
-\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
-\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,left of=hydras, node distance=3.5cm] (pockcc) { Pocklington };
-\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=hydras, node distance=2.5cm] (chains) { Addition-chains };
-\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,right of=chains, node distance=2.2cm] (gaia) { Gaia\footnote{J. Grimm} };
-\end{scope}
-
-\begin{scope}[yshift=-3cm,xshift=-6cm]
- \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
- \node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=3cm] (gaiahydras) { Gaia-hydras };
-\end{scope}
-
-\draw[lookcolor,->,dotted] (acl2.south) -- (cantor.north) ;
-\draw[lookcolor,->,dashed] (additions.south) -- (chains.north) ;
-\draw[lookcolor,->,dashed] (goedel) -- (goedelcc) ;
-\draw[lookcolor,->,dashed] (goedel) -- (hydras.north west) ;
-\draw[lookcolor,->,dashed] (cantor) -- (hydras.north east) ;
-\draw[lookcolor,->,dashed] (cantor.south) -- (gaia.north west) ;
-\draw[lookcolor,->,dashed] (pock) -- (pockcc) ;
-\draw[lookcolor,->,dashed] ([xshift=2.1cm]cats) -- (gaia) ;
-
-\draw[->] (hydras) -- (gaiahydras) ;
-\draw[->] (hydras.south west) -- (goedelcc) ;
-\draw[->] (gaia.south west) -- (gaiahydras.east) ;
-
-\draw[->] (equations) -- (hydras) ;
-\draw[->] (mathcomp) -- (gaia) ;
-\draw[->] (mathcomp) -- (chains) ;
-\draw[->] (paramcoq.south) -- (chains) ;
-\draw[->] (pockcc) -- (goedelcc) ;
-\end{tikzpicture}
-}
-\caption{Genealogy and dependencies for \Hydras packages.}
-% Dotted boxes represent historical Coq contributions, while regular boxes represent maintained \coq packages.
-Orange packages are maintained in the \Hydras GitHub repository, while blue packages are maintained in other \community repositories.
-Dotted lines represent \coq code ancestry, while regular lines represent direct code dependencies.}
-
-  \label{fig:genealogy}
-\end{figure}
-\end{frame}
 

### 2022-05-08 doc/JFLA2022/Slides/Slides.tex
changes to JFLA slides

@@
         \end{itemize}
-           \item book in pdf (300p), generated  with \textcolor{lookcolor}{\alectr} and \LaTeX, continuously updated and deployed
+           \item A book in pdf (300p), generated  with \textcolor{lookcolor}{\alectr} and \LaTeX, continuously updated and deployed
     \end{itemize}
@@
 Paris}};
+  \node[ellipse, dotted, fill=orange!20, minimum height=0.5cm,minimum width=1.5cm, left of=settheory, node distance=4.8cm] (algo) { Algorithmics\footnote{S. Brlek \emph{et. al}}};
 \end{scope}
@@
 \begin{scope}[yshift=1.0cm, xshift=-1.5cm]
-   \node[ellipse, dotted, draw= black!20,fill=green!20,minimum height=0.5cm,minimum width=1.5cm] (acl2) { [ACL2]\footnote{P. Manolios \&  D. Vroon}};   
+   \node[rectangle, dotted, draw= black!20,fill=green!20,minimum height=0.5cm,minimum width=1.5cm] (acl2) { [ACL2]\footnote{P. Manolios \&  D. Vroon}};   
 \end{scope}
@@
 \draw[dotted, ->] (settheory) -- (cats);
-
+\draw[dotted, ->] (prooftheory) -- (goedel);
+\draw[dotted, ->] (algo) -- (additions);
 \end{tikzpicture}
@@
     \item No need to be complete (there is already a reference manual)
-    \item No need to introduce \coq (there are already a lot of initiation books and tutorials)
-    \item Allow to compare several different formalizations of a given notion, and several proofs of the same theorem.
+ %   \item No need to introduce \coq (there are already a lot of initiation books and tutorials)
+    \item Comparison of  different formalizations of a given notion, and several proofs of the same theorem.
       \item Show also interesting dead ends (formalization, tactics, proof structure).
@@
 
-\section{Example: Kirby \& Paris' hydras}
+\section{Kirby \& Paris' hydras}
 

### 2022-05-09 doc/JFLA2022/Slides/Slides.tex
corrections in tex files

@@
 \end{frame}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-  \begin{figure}[ht]
-\centering
-\scalebox{0.7}{
-\begin{tikzpicture}[thick]
-
-  \begin{scope}[yshift=2.0cm, xshift=1.7cm] 
-     \node[ellipse, dotted, fill=orange!20, minimum height=0.5cm,minimum width=1.5cm] (settheory) { Set theory\footnote{N. Bourbaki}};
-     \node[ellipse, dotted, fill=orange!20, minimum height=0.5cm,minimum width=1.5cm, left of=settheory, node distance=8.4cm] (prooftheory) { Proof theory\footnote{K.
-Schutte, J.
-Ketonen, R.
-Solovay, L.
-Kirby \& J.
-Paris}};
-  \node[ellipse, dotted, fill=orange!20, minimum height=0.5cm,minimum width=1.5cm, left of=settheory, node distance=4.8cm] (algo) { Algorithmics\footnote{S. Brlek \emph{et. al}}};
-\end{scope}
-
-\begin{scope}[yshift=1.0cm, xshift=-1.5cm]
-   \node[rectangle, dotted, draw= black!20,fill=green!20,minimum height=0.5cm,minimum width=1.5cm] (acl2) { [ACL2]\footnote{P. Manolios \&  D. Vroon}};   
-\end{scope}
-
-\begin{scope}[xshift=-6.1cm]
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel\footnote{R. O'Connor} };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (additions) { Additions };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (cantor) { Cantor\footnote{C. \& E. Contejean} };
-
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3.2cm] (cats) { Categories in ZFC };
-\end{scope}
-
-\begin{scope}[yshift=-1cm,xshift=0.45cm]
-  \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=2.5cm] (mathcomp) { MathComp };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=paramcoq, node distance=4.1cm] (equations) { Equations };
-\end{scope}
-
-\begin{scope}[yshift=-2cm,xshift=-4cm]
-\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
-\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,left of=hydras, node distance=3.5cm] (pockcc) { Pocklington };
-\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=hydras, node distance=2.5cm] (chains) { Addition-chains };
-\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,right of=chains, node distance=2.2cm] (gaia) { Gaia\footnote{J. Grimm} };
-\end{scope}
-
-\begin{scope}[yshift=-3cm,xshift=-6cm]
- \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
- \node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=3cm] (gaiahydras) { Gaia-hydras };
-\end{scope}
-
-\draw[lookcolor,->,dotted] (acl2.south) -- (cantor.north) ;
-\draw[lookcolor,->,dashed] (additions.south) -- (chains.north) ;
-\draw[lookcolor,->,dashed] (goedel) -- (goedelcc) ;
-\draw[lookcolor,->,dashed] (goedel) -- (hydras.north west) ;
-\draw[lookcolor,->,dashed] (cantor) -- (hydras.north east) ;
-\draw[lookcolor,->,dashed] (cantor.south) -- (gaia.north west) ;
-\draw[lookcolor,->,dashed] (pock) -- (pockcc) ;
-\draw[lookcolor,->,dashed] ([xshift=2.1cm]cats) -- (gaia) ;
-
-\draw[->] (hydras) -- (gaiahydras) ;
-\draw[->] (hydras.south west) -- (goedelcc) ;
-\draw[->] (gaia.south west) -- (gaiahydras.east) ;
-
-\draw[->] (equations) -- (hydras) ;
-\draw[->] (mathcomp) -- (gaia) ;
-\draw[->] (mathcomp) -- (chains) ;
-\draw[->] (paramcoq.south) -- (chains) ;
-\draw[->] (pockcc) -- (goedelcc) ;
-\draw[dotted, ->] (prooftheory) -- (hydras);
-\draw[dotted, ->] (settheory) -- (cats);
-\draw[dotted, ->] (prooftheory) -- (goedel);
-\draw[dotted, ->] (algo) -- (additions);
-\end{tikzpicture}
-}
-\caption{Genealogy and dependencies for \Hydras packages.}
-  \label{fig:genealogy}
-\end{figure}
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \begin{frame}
-%   \frametitle{Yet another book on \coq?}
-%   \begin{table}
-%   \centering
-%   \footnotesize
-%   \begin{tabular}{lll}
-%   \hline
-%   \textbf{Title} & \textbf{Year} & \textbf{Category}\\
-%     \hline
-%    \coq reference manual &  $\geq 1989$ & \\
-%   Coq'Art & 2004 & Traditional\\
-%   Software Foundations & 2007 & Executable\\
-%   Certified Programming with Dependent Types & 2011 &  Executable\\
-%   Program Logics for Certified Compilers & 2014 & Traditional\\
-%   Programs and Proofs & 2014 & Executable\\
-%   Formal Reasoning About Programs& 2017 & Traditional\\
-%   Computer Arithmetic and Formal Proofs & 2017 & Traditional\\
-%     Mathematical Components & 2018 & Traditional\\
-%   \hline
-%   \end{tabular}
-%   \caption{Coq book categorization.}
-%   \label{tbl:books}
-% \end{table}
-
-% \end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{frame}
-   \begin{block}{An example-driven book}
-    \begin{itemize}
-    \item The structure is determined by the presented topics (discrete math, algorithmics, etc.). \coq technology (formalization and proof patterns) is presented when needed.
-    \item No need to be complete (there is already a reference manual)
- %   \item No need to introduce \coq (there are already a lot of initiation books and tutorials)
-    \item Comparison of  different formalizations of a given notion, and several proofs of the same theorem.
-      \item Show also interesting dead ends (formalization, tactics, proof structure).
-    \end{itemize}
-  \end{block}
-  % \begin{block}{Examples}
-  %   \begin{itemize}
-  %   \item Tsukuba Coq Users's Group: regular languages, $\lambda$-calculus, arithmetic, ... (in Japanese + ssreflect)
-  %   \item Software Foundations (B. Pierce \emph{et al.})
-  %   \end{itemize}
-  % \end{block}
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \begin{frame}
@@
 
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\centering
+\scalebox{0.7}{
+\begin{tikzpicture}[thick]
+
+  %%%%%%%%%%%% First line %%%%%%%%%%%%%%%%
+  \begin{scope}[yshift=2.0cm, xshift=3cm]
+     \node[ellipse, dotted, fill=orange!20, minimum height=0.5cm,minimum width=1.5cm] (algo) { Algorithmics};
+     \node[ellipse, dotted, fill=orange!20, minimum height=0.5cm,minimum width=1.5cm, left of=algo, node distance=6cm] (settheory) { Set theory};
+     \node[ellipse, dotted, fill=orange!20, minimum height=0.5cm,minimum width=1.5cm, left of=settheory, node distance=6cm] (prooftheory) {Proof theory};
+
+\end{scope}
+
+%%%%%% ACL2   %%%%%%%
+
+\begin{scope}[yshift=1.0cm, xshift=-3.7cm]
+   \node[rectangle, dotted, draw= black!20,fill=green!20,minimum height=0.5cm,minimum width=1.5cm] (acl2) { [ACL2]\footnote{P. Manolios \&  D. Vroon}};   
+\end{scope}
+
+%%%%%%% Contribs %%%%%%%%%%%%%%%%%%%%
+
+\begin{scope}[xshift=-5.7cm]
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel\footnote{R. O'Connor} };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (cantor) { Cantor\footnote{C. \& E. Contejean}};
+
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3.6cm] (cats) { Categories in ZFC };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cats, node distance=3.2cm] (additions) {Additions};
+
+\end{scope}
+
+%%%%%%%%%%%%%%%%% Plug-ins %%%%%%%%%%%%%%
+
+\begin{scope}[yshift=-0.8cm,xshift=-1.45cm]
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm] (mathcomp) { MathComp };
+
+  \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm, right of=mathcomp, node distance=5.4cm] (paramcoq) { Paramcoq};
+
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=mathcomp, node distance=7.7cm] (equations) { Equations };
+\end{scope}
+
+\begin{scope}[yshift=-2cm,xshift=-4.2cm]
+\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
+\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,left of=hydras, node distance=3.5cm] (pockcc) { Pocklington };
+\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm, right of=hydras, node distance=3.5cm] (gaia) { Gaia\footnote{J. Grimm} };
+
+\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=gaia, node distance=3.5cm] (chains) { Addition-chains };
+
+\end{scope}
+
+%%%%%%%%%%%%%%%%% goedel, etc . %%%%%%%%%%%%%%%
+\begin{scope}[yshift=-3.7cm,xshift=-6cm]
+ \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
+ \node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=3cm] (gaiahydras) { Gaia-hydras };
+\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=gaiahydras, node distance=5.1cm] (alectryon) { Alectryon};
+
+\end{scope}
+
+\begin{scope}[yshift=-5cm,xshift=4cm]
+\node[rectangle, draw=black,fill=red!80,minimum height=0.5cm,minimum width=1.5cm] (book) { \Hydras};  
+\end{scope}
+
+\draw[lookcolor,->,dotted] (acl2.south) -- (cantor.north) ;
+\draw[lookcolor,->,dashed] (additions.south) -- (chains.north) ;
+\draw[lookcolor,->,dashed] (goedel) -- (goedelcc) ;
+\draw[lookcolor,->,dashed] (goedel) -- (hydras.north) ;
+\draw[lookcolor,->,dashed] (cantor) -- (hydras.north east) ;
+\draw[lookcolor,->,dashed] (cantor.south) -- (gaia.north west) ;
+\draw[lookcolor,->,dashed] (pock) -- (pockcc) ;
+\draw[lookcolor,->,dashed] ([xshift=2.1cm]cats) -- (gaia) ;
+
+\draw[->] (hydras.south) -- (gaiahydras) ;
+\draw[->] (hydras.south west) -- (goedelcc) ;
+\draw[->] (gaia.south west) -- (gaiahydras.north) ;
+
+\draw[->] (equations) -- (hydras) ;
+\draw[->] (mathcomp) -- (gaia) ;
+\draw[->] (mathcomp) -- (chains) ;
+\draw[->] (paramcoq.south) -- (chains) ;
+\draw[->] (pockcc) -- (goedelcc) ;
+\draw[dotted, ->] (prooftheory) -- (hydras);
+\draw[dotted, ->] (settheory) -- (cats);
+\draw[dotted, ->] (prooftheory) -- (goedel);
+\draw[dotted, ->] (algo) -- (additions);
+\draw[->, red] (alectryon) -- (book) ;
+\draw[->, red] (chains) -- (book) ;
+\draw[->, red] (gaiahydras) -- (book) ;
+\draw[->, red] (hydras.south) -- (book) ;
+
+\end{tikzpicture}
+}
+% \caption{Genealogy and dependencies for \Hydras.}
+%   \label{fig:genealogy}
+% \end{figure}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \begin{frame}
+%   \frametitle{Yet another book on \coq?}
+%   \begin{table}
+%   \centering
+%   \footnotesize
+%   \begin{tabular}{lll}
+%   \hline
+%   \textbf{Title} & \textbf{Year} & \textbf{Category}\\
+%     \hline
+%    \coq reference manual &  $\geq 1989$ & \\
+%   Coq'Art & 2004 & Traditional\\
+%   Software Foundations & 2007 & Executable\\
+%   Certified Programming with Dependent Types & 2011 &  Executable\\
+%   Program Logics for Certified Compilers & 2014 & Traditional\\
+%   Programs and Proofs & 2014 & Executable\\
+%   Formal Reasoning About Programs& 2017 & Traditional\\
+%   Computer Arithmetic and Formal Proofs & 2017 & Traditional\\
+%     Mathematical Components & 2018 & Traditional\\
+%   \hline
+%   \end{tabular}
+%   \caption{Coq book categorization.}
+%   \label{tbl:books}
+% \end{table}
+
+% \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \begin{frame}
+%    \begin{block}{An example-driven book}
+%     \begin{itemize}
+%     \item The structure is determined by the presented topics (discrete math, algorithmics, etc.). \coq technology (formalization and proof patterns) is presented when needed.
+%     \item No need to be complete (there is already a reference manual)
+%  %   \item No need to introduce \coq (there are already a lot of initiation books and tutorials)
+%     \item Comparison of  different formalizations of a given notion, and several proofs of the same theorem.
+%       \item Show also interesting dead ends (formalization, tactics, proof structure).
+%     \end{itemize}
+%   \end{block}
+%   % \begin{block}{Examples}
+%   %   \begin{itemize}
+%   %   \item Tsukuba Coq Users's Group: regular languages, $\lambda$-calculus, arithmetic, ... (in Japanese + ssreflect)
+%   %   \item Software Foundations (B. Pierce \emph{et al.})
+%   %   \end{itemize}
+%   % \end{block}
+% \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{frame}
-  \frametitle{Example: Kirby \& Paris' hydras}
+  \frametitle{Kirby \& Paris' hydras}
   \begin{figure}[h]
@@
 
-\begin{frame}
-  \begin{block}{}
-    Using  \texttt{PrimRec\_PrimRecs\_ind}, we prove the following lemma:
-
-Comparison of the length of hydra-battles (w.r.t to the initial replication factor) with the diagonalized Ackermann function proves the theorem.
-  \end{block}
-\end{frame}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-05-09 doc/gaia-chapter.tex
corrections in tex files

@@
 
-%\inputsnippets{T1Bridge/utf8try}
-
 \subsubsection{Unary functions}
@@
 The definition of the binary operations $+$ and $\times$
-requires the comparison of ordinal terms, which is not implemented the same way in both libraries. Thus, the proof of the following lemmas applies applies
+requires the comparison of ordinal terms, which is not implemented in \gaiaHydras  the same way as in \HydrasLib. Thus, the proof of the following lemmas applies applies
 compatibility lemmas like \texttt{compare\_ref} (see Section~\vref{sect:gaia-compare-ref}).
@@
 \label{sect:gaia-masking}
-We try to respect the following ``protocol''.
-Let us assume we adapt some definitions and properties
-from a given module \texttt{hydras.Epsilon0.Foo}, and a constant \texttt{bar} of this module, whose absolute name is
-thus \texttt{hydras.Epsilon0.Foo.bar}.
-\begin{enumerate}
-\item We create a module \texttt{gaia\_hydras.GFoo}
-  (``G'' for \gaia).
-\item Inside \texttt{GFoo}, the identifier \texttt{bar}
-  is reserved in priority to \gaia's version.
-\HydrasLib' constructions are available through qualified names, like \texttt{Epsilon0.Foo.bar}, \texttt{Foo.bar}, or even a short notation like \texttt{hbar}  (``h'' for \HydrasLib).
-\end{enumerate}
+We try to respect the following convention:
+Let us consider  some module \texttt{hydras.Epsilon0.Foo}.
+Its adaptation to \gaia will be a new module called \linebreak
+\texttt{gaia\_hydras.GFoo} (``G'' for \gaia).
+
+Let \texttt{bar} be a symbol defined in \texttt{Foo}.
+Inside \texttt{GFoo}, the identifier \texttt{bar}
+  is reserved in priority to \gaia's adaptation.
+  \HydrasLib' definition is still available through qualified names, like \texttt{Epsilon0.Foo.bar}, \texttt{Foo.bar}, or even a short notation like \texttt{hbar}  (``h'' for \HydrasLib).
 

### 2022-05-09 doc/gaia-chapter.tex
corrections in tex files

@@
 
-Some constructions of \HydrasLib were (to our knowledge) not implemented in \gaia yet.
-For instance, commutative (``Hessenberg'') addition, canonical sequences and Ketonen-Solovay machinery (Chapters~\ref{chap:T1}, \ref{chap:ketonen} and \ref{chap:alpha-large}).
+Some constructions of \HydrasLib were (to our knowledge) not implemented in \gaia yet, for instance, commutative (``Hessenberg'') addition, canonical sequences and Ketonen-Solovay machinery (Chapters~\ref{chap:T1}, \ref{chap:ketonen} and \ref{chap:alpha-large}).
 In order not to copy long proofs, we chose to

### 2022-05-11 doc/JFLA2022/Slides/Slides.tex
work on JFLA Slides

@@
 \begin{frame}
+  \frametitle{\Hydras: genealogy and dependencies}
 \centering
-\scalebox{0.7}{
+\scalebox{0.6}{
 \begin{tikzpicture}[thick]
@@
 \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel\footnote{R. O'Connor} };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington\footnote{O. Caprotti \& M. Oostdijk}};
 \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (cantor) { Cantor\footnote{C. \& E. Contejean}};
 
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3.6cm] (cats) { Categories in ZFC };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3.6cm] (cats) { Categories in ZFC\footnote{C. Simpson} };
 \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cats, node distance=3.2cm] (additions) {Additions};
@@
 
-\begin{scope}[yshift=-0.8cm,xshift=-1.45cm]
+\begin{scope}[yshift=-0.8cm,xshift=-1.2cm]
 \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm] (mathcomp) { MathComp };
@@
 \draw[lookcolor,->,dotted] (acl2.south) -- (cantor.north) ;
-\draw[lookcolor,->,dashed] (additions.south) -- (chains.north) ;
+\draw[lookcolor,->,dashed] (additions.south west) -- (chains.north) ;
 \draw[lookcolor,->,dashed] (goedel) -- (goedelcc) ;
@@
 
-\pause
+%\pause
 \textbf{Remark:} The proof is not trivial. The built path could miss $\beta$!
@@
 
-\begin{frame}[fragile]
+% \begin{frame}[fragile]
 
-      \begin{block}{}
-    The functions \textcolor{mathcolor}{$H'_\alpha, F_\alpha\;(\alpha< \epsilon_0)$} are implemented with \textcolor{plugincolor}{coq-equations}.
-    \vspace{5pt}
-    \begin{footnotesize}
-      \inputsnippets{../Hprime_HprimeDef}   
-    \end{footnotesize}
+%       \begin{block}{}
+%     The functions \textcolor{mathcolor}{$H'_\alpha, F_\alpha\;(\alpha< \epsilon_0)$} are implemented with \textcolor{plugincolor}{coq-equations}.
+%     \vspace{5pt}
+%     \begin{footnotesize}
+%       \inputsnippets{../Hprime_HprimeDef}   
+%     \end{footnotesize}
 
-  \end{block}
+%   \end{block}
 
-\end{frame}
+% \end{frame}
 
@@
 
-  \begin{block}{Theorem}
+  \begin{block}{Theorems}
+    \begin{itemize}
+    \item $F_\alpha$ is primitive recursive iff $\alpha<\omega$.
+     \item  
     The function which computes the length of hydra battles
-    (according to the initial index) is not primitive recursive
+    for $\alpha$ (according to the initial index) is not primitive recursive
     (for $\alpha\geq \omega^\omega$).
-    \vspace{4pt}
+
+    \end{itemize}
+
+  \begin{block}{Sketch of proof:}
+
+   We apply monotony properties and the famous lemma:
+    \vspace{5pt}
+
+       \begin{quote}
+      {\color{mathcolor}
+        For any $n$-ary primitive recursive  function $f$, 
+        there exists some $q$ such that the following inequality holds:
+         \[\forall\,x_1,\dots,x_n,  
+          f(x_1,\dots,\,x_n)\leq\textrm{Ack}(q,\textrm{max}(x_1,\dots,x_n))\]}
+    \end{quote}
+    (Proof by induction on any primitive recursive function of any arity.)
   \end{block}
 
-  \begin{block}{Sketch of proof}
-
-    We prove and apply a famous lemma:
-
-       \begin{quote}
-
-      {\color{mathcolor}
-        \[
-          \exists q, \forall x_1,\dots,x_n, 
-          f(x_1,\dots,\,x_n)\leq\textrm{Ack}(q,\textrm{max}(x_1,\dots,x_n))
-        \]}
-    \end{quote}
-
   \end{block}
-
 \end{frame}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \begin{frame}
-
-%   \begin{block}{}
-%     {\color{coqstylecolor}
-%       \begin{itemize}
-%       \item \emph{In order to prove this theorem, we chose to host and  maintain under \community Russel O'Connor's library \textcolor{plugincolor}{goedel}, which contains a formalization of primitive recursive functions.}
-
-%       \item    \emph{The interest of O'Connor's contribution led us to write a chapter on primitive recursive functions, with [counter-]examples and exercices.}
-
-%       \item \emph{Proving formally that a given function is not primitive recursive is a nice example of application of a complex (mutually recursive, dependently typed) induction principle.
-You have to take into account the set of PR functions of any arity.}
-
-%       \end{itemize}
-%     }
-%   \end{block}
-% \end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \section{Project Organization}
-\begin{frame}
-  \frametitle{Project  Organization}
-  \begin{itemize}
-   \item Compatibility with \gaia
-  \item Integration with \community
-  \item Documentation with \alectr
-  \item Package dependencies
-  \item Continuous integration
-  \end{itemize}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Project  Organization}
+%   \begin{itemize}
+%    \item Compatibility with \gaia
+%   \item Integration with \community
+%   \item Documentation with \alectr
+%   \item Package dependencies
+%   \item Continuous integration
+%   \end{itemize}
+% \end{frame}
 
@@
   \end{block}
+  \begin{block}{}
+    We experiment also the other direction (importing \texttt{ssete9} lemmas to \Hydras).
+  \end{block}
 \end{frame}
@@
   \begin{block}{}
-  {\color{lookcolor}  With \Hydras, we hope to demonstrate that projects with complex dependencies are feasible to manage using recent advances in build management and infrastructure.}
+    With \Hydras, we hope to demonstrate that projects with complex dependencies are feasible to manage using recent advances in build management and infrastructure.
   \end{block}
@@
 
-  \begin{block}{To do ?}
+  \begin{block}{To do}
     Define a language for
@@
 \end{frame}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \begin{frame}
-%   \frametitle{Package organization and dependencies}
-
-% \begin{figure}[ht]
-% \centering
-% \scalebox{0.7}{
-% \begin{tikzpicture}[thick]
-
-% \begin{scope}[xshift=-6cm]
-% \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel };
-% \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington };
-% \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (additions) { Additions };
-% \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (cantor) { Cantor };
-
-% \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3.2cm] (cats) { Categories in ZFC };
-% \end{scope}
-
-% \begin{scope}[yshift=-1cm,xshift=0.45cm]
-%   \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
-% \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=2.5cm] (mathcomp) { MathComp };
-% \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=paramcoq, node distance=4.5cm] (equations) { Equations };
-% \end{scope}
-
-% \begin{scope}[yshift=-2cm,xshift=-4cm]
-% \node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
-% \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,left of=hydras, node distance=3.5cm] (pockcc) { Pocklington };
-% \node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=hydras, node distance=2.5cm] (chains) { Addition-chains };
-% \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,right of=chains, node distance=2.2cm] (gaia) { Gaia };
-% \end{scope}
-
-% \begin{scope}[yshift=-3cm,xshift=-6cm]
-%  \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
-%  \node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=3cm] (gaiahydras) { Gaia-hydras };
-% \end{scope}
-
-% \draw[lookcolor,->,dashed] (additions.south) -- (chains.north) ;
-% \draw[lookcolor,->,dashed] (goedel) -- (goedelcc) ;
-% \draw[lookcolor,->,dashed] (goedel) -- (hydras.north west) ;
-% \draw[lookcolor,->,dashed] (cantor) -- (hydras.north east) ;
-% \draw[lookcolor,->,dashed] (cantor.south) -- (gaia.north west) ;
-% \draw[lookcolor,->,dashed] (pock) -- (pockcc) ;
-% \draw[lookcolor,->,dashed] ([xshift=2.1cm]cats) -- (gaia) ;
-
-% \draw[->] (hydras) -- (gaiahydras) ;
-% \draw[->] (hydras.south west) -- (goedelcc) ;
-% \draw[->] (gaia.south west) -- (gaiahydras.east) ;
-
-% \draw[->] (equations) -- (hydras) ;
-% \draw[->] (mathcomp) -- (gaia) ;
-% \draw[->] (mathcomp) -- (chains) ;
-% \draw[->] (paramcoq.south) -- (chains) ;
-% \draw[->] (pockcc) -- (goedelcc) ;
-% \end{tikzpicture}
-% }
-% \caption{\textcolor{lookcolor}{Genealogy} and dependencies for \Hydras packages.}
-% Dotted boxes represent historical Coq contributions, while regular boxes represent maintained \coq packages.
-Orange packages are maintained in the \Hydras GitHub repository, while blue packages are maintained in other \community repositories.
-Dotted lines represent \coq code ancestry, while regular lines represent direct code dependencies.}
-
-%   \label{fig:genealogy}
-% \end{figure}
-
-% %%%% TODO : Resize  !
-% {\scriptsize
-% \begin{table}[ht]
-% \centering
-% \scriptsize
-% \begin{tabular}{llllrr}
-% \hline
-% \textbf{Package name} &  \textbf{Version} & \textbf{Spec LOC} & \textbf{Proof LOC}\\
-% \hline
-% Pocklington & \texttt{8.12.0} & 825 & 3,798 \\
-% Goedel  & \texttt{8.13.0} & 2,799 & 10,762 \\
-% Gaia &  \texttt{1.12} & 28,850 & 124,839 \\
-% Hydra-battles & \texttt{0.5} & 14,368 & 48,385 \\
-% Addition-chains & \texttt{0.5} & 1,961 & 2,210 \\
-% Gaia-hydras &  \texttt{0.8} & 677 & 2,023 \\
-% \hline
-% \end{tabular}
-% \caption{\scriptsize Current numbers of lines of code for packages in the \Hydras galaxy.}
-% \label{tbl:loc}
-% \end{table}}
-%\end{frame}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-05-13 doc/hydra-chapter.tex
revision book (-> p. 28)

@@
 \begin{project}
-The type \texttt{Hydra} above describes hydras as \emph{plane trees}, i.e., as drawn on a sheet of paper or computer screen. Thus, hydras are \emph{oriented},
-and it is appropriate to consider a \emph{leftmost} or \emph{rightmost} head of
+  Our type \texttt{Hydra} describes hydras as
+  \emph{plane oriented trees}, i.e., as drawn on a sheet of paper or computer screen. Thus, it is appropriate to consider a \emph{leftmost} or \emph{rightmost} head of
 the beast. It could be interesting to consider another representation, in which
@@
 
-For instance, the hydra \texttt{Hy}  of Figure~\vref{fig:Hy} is defined in \emph{Gallina} as follows:
+For instance, the hydra \texttt{Hy}  of Figure~\vref{fig:Hy} is defined as follows:
 
@@
 
-For instance, the hydra \texttt{Hy''} of Fig~\vref{fig:Hy3}  can be defined in \coq{} as follows:
+Let us consider for instance the hydra \texttt{Hy''} of Fig~\vref{fig:Hy3}.
 
@@
 
-Let us look at the second subgoal of the induction.
+Let us look now at the second subgoal of the induction.
 
@@
 
-We notice that this subgoal does not contain any hypothesis
+We notice that this sub-goal does not contain any hypothesis
 on the height and size of the hydra \texttt{h}. So, it looks hard to prove the conclusion. Let's stop.

### 2022-05-13 doc/hydras.tex
minor changes in theories/gaia
reorganization of tex files
revision book (-> p. 28)

@@
 \begin{scope}[xshift=-6.5cm]
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel };
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) {Goedel };
 \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=1.9cm] (pock) { Pocklington };
@@
 \label{sec:orgheadline4}
-\begin{itemize}
-\item The present distribution has been checked with version 8.14.1 of the Coq proof assistant, with the plug-ins \texttt{coq-paramcoq}, \texttt{coq-equations}
-and \texttt{coq-mathcomp-algebra}.
-
-\item Please refer to \href{https://github.com/coq-community/hydra-battles#readme}{the README file of the project}
-\end{itemize}
+ The present distribution has been checked with version 8.14.1 of the Coq proof assistant, with a few plug-ins.
+\emph{Please refer to \href{https://github.com/coq-community/hydra-battle\#readme}{the README file of the project}.}
 
@@
 
-\include{part-hydras}
+\include{part-hydras-intro}
+\include{hydra-chapter}
+\include{ordinal-chapter}
+\include{epsilon0-chapter}
+\include{ks-chapter}
+\include{large-sets-chapter}
+\include{gaia-chapter}
+\include{schutte-chapter}
+\include{Gamma0-chapter}
 \include{chapter-primrec}
+
 \part{A few  certified algorithms}

### 2022-05-14 doc/hydra-chapter.tex
revision book (-> p. 32)

@@
 Design a tactic for induction on hydras that frees the user from binding explicitly \texttt{P0},  and solves trivial subgoals.
-Apply it for writing  a shorter proof of \texttt{height\_lt\_size}.
+Apply it for writing  a shorter proof script of \texttt{height\_lt\_size}.
 \end{exercise}
@@
 
-If Hercules chops off a head near the floor, there is no replication at all. We use an auxiliary 
+If Hercules chops off a head close to the root, there is no replication at all. We use an auxiliary 
 predicate \texttt{S0}, associated with the removing of one head from a sequence of hydras.
@@
 
-We combine the two cases above into a single relation.
+Let us merge \texttt{R1} and \texttt{R1}  into a single relation.
 First,  we define the  relation \texttt{(round\_n n h h')} where \texttt{n} is the expected number of  replications (irrelevant in the case of an \texttt{R1}-transformation).
+Then, we define a \emph{round} (small step) of a battle
+by abstraction over \texttt{n},
 
@@
 \index{hydras}{Library Hydra!Predicates!round\_n}
-
-\input{movies/snippets/Hydra_Definitions/roundNDef}
-
-By abstraction over \texttt{n}, we define a \emph{round} (small step) of a battle:
-
 \index{hydras}{Library Hydra!Predicates!round}
@@
 
-\input{movies/snippets/Hydra_Definitions/roundDef}
+\inputsnippets{Hydra_Definitions/roundNDef, Hydra_Definitions/roundDef}
 
@@
 \begin{itemize}
-\item For proving $\forall y, x\,R^+\,y \;\arrow\; P\,y$, prefer 
+\item To prove $\forall y, x\,R^+\,y \;\arrow\; P\,y$, prefer 
 \texttt{clos\_trans\_n1}
-\item For proving $\forall x,\,x\,R^+\,y \;\arrow\; P\,x$, prefer \texttt{clos\_trans\_1n}
-\item For proving $\forall x\,y, \,x\,R^+\,y \;\arrow\;P\,x\,y$,  
+\item To prove proving $\forall x,\,x\,R^+\,y \;\arrow\; P\,x$, prefer \texttt{clos\_trans\_1n}
+\item To prove $\forall x\,y, \,x\,R^+\,y \;\arrow\;P\,x\,y$,  
 prefer \texttt{clos\_trans},
@@
 \begin{itemize}
-\item  Chose to stop a battle, or continue
-\item Chose an expected number of replications
+\item  Choose to stop a battle, or continue
+\item Choose an expected number of replications
 \item Navigate in a hydra, looking for a head to chop off.
@@
 
-The most general class of battles is \texttt{free}, which allows the hydra to chose any replication factor at every step:
+The most general class of battles is \texttt{free}, which allows the hydra to choose any replication factor at every step:
 

### 2022-05-14 doc/hydra-chapter.tex
deprecate 'battle' -> 'rounds'

@@
 
-% \begin{remark}
-%  The class \texttt{free} is strongly related with the transitive closure  \texttt{round\_plus}, as expressed by the following lemmas.
-
-% \vspace{4pt}
-% \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}
-
-%  \begin{Coqsrc}
-%  Lemma battle_free_equiv1 : forall i j h h',  
-%              battle free i h j h' ->   h -+-> h'.
-
-%  Lemma battle_free_equiv2 : forall h h',
-%      h -+-> h' ->
-%     forall i, exists j,  battle free i h j h'.
-%  \end{Coqsrc}
-
-% \end{remark}
-
 \section{A long battle}
@@
 
-The following lemma  is a formal description of these first rounds, in terms of the
-\texttt{battle} predicate.
+The following lemma  is a formal description of these first rounds, in terms of the \texttt{rounds} predicate.
 
@@
 Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle only start with the lemma
-\texttt{step\_battle}, page~\pageref{lemma:step-battle}.
+\texttt{step\_rounds}, page~\pageref{lemma:step-battle}.
 

### 2022-05-15 doc/hydra-chapter.tex
revision book (-> p. 43)

@@
 
-The following property allows us to build battles by composition of smaller ones.
+The following property allows us to build battle rounds by composition of smaller ones.
 
@@
 \item At each round, Hercules chops off the rightmost head of the hydra.
-\item The battle is {standard}: at the round number $i$, the expected replication is $i$.
+\item The battle is {standard}: at the round number $i$, the expected replication factor is $i$.
 \end{itemize}
@@
 The lemma we would like to prove is ``The considered battle lasts exactly $N$ rounds'',
-with $N$ being a natural number we gave to guess.
-
-But the  battle is so long that no \emph{test} can give us any estimation of its length, and we do need the expressive power of logic to compute this length.
-However, in order to  guess this length, we made some experiments, computing with \gallina{}, \coq{}'s  functional programming language.
+with $N$ being a natural number we have to guess.
+
+But the  battle is so long that no \emph{test} can give us any estimation of its length.
+Nevertheless, in order to  guess this length, we made some experiments, computing with \gallina{}, \coq{}'s  functional programming language.
 Thus, we can consider this development as a collaboration of proof with computation.
-In the following lines, we show how we found experimentally the value of the number $N$.
+In the rest of this section, we show how we found experimentally the value of the number $N$.
 The complete proof is in file \url{../theories/html/hydras.Hydra.BigBattle.html}.
@@
 \subsection{Proving \dots}
-We are now able to reason about the sequence of transitions defined by our hydra battle. Instead of using the data-type \texttt{state} we study the relationship
-between different configurations of the battle.
+We are now able to reason about the sequence of transitions defined by our hydra battle.
 
@@
 In the following definition \texttt{i} is associated with the round number (or date, if we consider a discrete time), and \texttt{a}, \texttt{b}, \texttt{c} respectively associated with the number of occurrences of \texttt{h2}, \texttt{h1} and heads connected to the hydra's foot.
+For convenience, we do not use the type \texttt{state} of the preceeding section, but consider the round numbers and the number of hydras \texttt{h2}, \texttt{h1} and heads as separate arguments of the relation (which is no more ---formally--- ``binary'').
 

### 2022-05-16 doc/hydra-chapter.tex
revision book (chap 2, done)

@@
 
+\vspace{4pt}
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
+
+\input{movies/snippets/Hydra_Lemmas/nextRoundDec}
+
+\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Theorems}{Hydra.Hydra\_Theorems}}
+
 \input{movies/snippets/Hydra_Theorems/AliveThms}
-
-Both theorems are proved with the help of the  following strongly specified function:
-
-\vspace{4pt}
-\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}
-
-\input{movies/snippets/Hydra_Lemmas/nextRoundDec}
 
@@
 one has to consider first a well-founded set $(A,<)$, then a strictly decreasing measure on this set.
-The following two lemmas show that if  the order structure $(A,<)$ is too simple, it is useless to look for a convenient measure, which simply no exists.
+The following  lemma shows that,  if  the order structure $(A,<)$ is too simple, it is useless to look for a convenient measure, which simply no exists.
 Such kind of result is useful, because it saves you time and effort.
@@
 
-Let us assume there exists some variant $m$ from \texttt{Hydra} into \texttt{nat} for proving
+Let us assume there exists some variant $m$ from \texttt{Hydra} into (\texttt{nat},$<$)  for proving
     the  termination of all hydra battles.

### 2022-05-16 doc/ordinal-chapter.tex
revision book (<= p. 61)

@@
 
-Such a structure may be proved correct relatively to a bigger ordinal notation or to Schütte's or \gaia's model.
+Such a structure should be proved correct relatively to a bigger ordinal notation or to Schütte's or \gaia's model.
 
@@
 
-\subsubsection{The \texttt{Comparison} and \texttt{Comparable} classes}
-
-We use the operational class \texttt{Comparison} of comparison functions to define the \texttt{Comparable} class, contributed by Jérémy Damour and Théo Zimmermann, which allows us to apply generic lemmas and tactics about decidable strict orders.
+\subsubsection{The \texttt{Comparable} class}
+
+The \texttt{Comparable} class, contributed by Jérémy Damour and Théo Zimmermann, allows us to apply generic lemmas and tactics about decidable strict orders.
 The correctness of the comparison function is expressed through Stdlib's type 
@@
 So the infix \texttt{<} and \texttt{<=} are reserved to the natural numbers.
-In the mathematical formulas, however, we still use $<$ and $\leq$ for comparing ordinals.
+In mathematical formulas, however, we still use $<$ and $\leq$ for inequalities between ordinals.
 \end{remark}
@@
 
-\inputsnippets{ON_Omega/OmegaDefa, ON_Omega/OmegaDefb}
+\inputsnippets{ON_Omega/OmegaDefa}
+\inputsnippets{ON_Omega/OmegaDefb}
 
@@
 \item $y$ is a \emph{successor} of $x$ if $x<y$ and there is no element between $x$ and $y$. We will also say that $x$ is a \emph{predecessor} of $y$.
-\item $x$ is a \emph{limit} if $x$ is not a least element, and for any $y$ such that $yo<x$,
+\item $x$ is a \emph{limit} if $x$ is not a least element, and for any $y$ such that $y<x$,
  there exists some $z$ such that $y<z<x$.
@@
 
-\subsection{Arithmetic of \texorpdfstring{$\omega^2$}{omega^2}}
+\subsection{Arithmetic on \texorpdfstring{$\omega^2$}{omega^2}}
 
@@
 
-\subsubsection{Multiplication}
+\subsubsection{Finite multiplication}
 
@@
 
-A case analysis on $i$ and $j$ allows us to
+A case analysis on $i$ and $j$ leads  us to
 consider three cases :
@@
 
-In order to build an instance of \texttt{ON}, we define a comparison function,  and prove its correctiness.
+In order to build an instance of \texttt{ON}, we define a comparison function, then prove its correctness.
 
@@
 equality on type \texttt{bool}.
-This is why we used the boolean function \texttt{Nat.ltb} instead of the inductive predicate \texttt{Nat.lt} in the definition of type \texttt{t $n$} (see page~\pageref{def: Finite-ord-type}).
+This is why we used the boolean function \texttt{Nat.ltb} instead of the inductive predicate \texttt{Nat.lt} in the definition of type \texttt{t $n$} (see page~\pageref{def:Finite-ord-type}).
 For more information about this pattern, please look at the numerous mailing lists and 
@@
 
-Finite ordinals are also formalized in \ssreflect/\mathcomp~\cite{MCB}.
+\gaiasign Finite ordinals are also formalized in \ssreflect/\mathcomp~\cite{MCB}.
 In Module~\href{../theories/html/gaia_hydras.ON_gfinite.html}{gaia\_hydras.ON\_gfinite}, we build an instance of class \texttt{ON}.
@@
 The following class specifies a strict inclusion of segments.
-The notation \texttt{OA} describes a segment $[0,\alpha)$, and \texttt{OB} is a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented in \texttt{OA}).
-We require also  that the comparison functions of the two notation systems are compatible.
+The ordinal notation \texttt{OA} describes a segment $[0,\alpha)$, and \texttt{OB}  a larger segment (which contains a notation for $\alpha$, whilst $\alpha$ is not represented in \texttt{OA}) (like in Fig.~\ref{fig:subsegment}).
+We require also the comparison functions of the two notation systems to be compatible.
 
@@
 \begin{scope}[color=blue]
-\node (A) at (0,0) {$A$};
+\node (A) at (0,0) {$\omega$};
 \node(A0) at (2,0)[label=below:$0$]{$\bullet$};
@@
 \begin{scope}[color=red]
-\node (B) at (0,2) {$B$};
+\node (B)  at (0,2) {$\omega+\omega$};
 \node(B0) at (2,2)[label=above:$0$]{$\bullet$};
@@
 \end{tikzpicture}
-   \caption{\textcolor{blue}{$A$} is a sub-segment  of \textcolor{red}{$B$}}
+   \caption{\textcolor{blue}{$\omega$} is a sub-segment  of \textcolor{red}{$\omega+\omega$}}
    \label{fig:subsegment}

### 2022-05-17 doc/epsilon0-chapter.tex
Revision book to page 82. deprecate a few symbols

@@
 We note that the exponents of $\omega$ are not in the right (strictly decreasing) order.
-Nevertheless, with the help of the order \texttt{lt} on \texttt{T1}, we are now able to characterize
-the set of all well-formed ordinal terms:
+The following boolean function determines whether a given ordinal term is well formed.
 
@@
 \index{gaiabridge}{Type of well formed ordinal terms below $\epsilon_0$}
- The library~\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines a \gaia-compatible type \texttt{E0}.
+ Our library~\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines also a type \texttt{E0} (which doesn't exist in \gaia-\texttt{ssete9}).
 
@@
 \noindent
-\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}
+\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#T1is_succ}{Epsilon0.T1}}
 
@@
 \index{gaiabridge}{Limit and successor ordinals}
-In \gaia, the boolean functions associated with limit and successor ordinals are called \texttt{T1limit} and \texttt{T1is\_succ}.
+In \gaia, the boolean functions associated with limit and successor ordinals are also called \texttt{T1limit} and \texttt{T1is\_succ}.
 
@@
 The following lemma establishes the connection between the  function
-\texttt{succ} and the Boolean predicate \texttt{succb}.
+\texttt{succ} and the Boolean predicate \texttt{T1is\_succ}.
 
@@
 \begin{exercise}[\gaiasign]
-  Look for the \gaia-theorem which corresponds to \texttt{succb\_iff}.
+  Look for the \gaia-theorem which corresponds to \texttt{T1is\_succ\_iff}.
 \end{exercise}

### 2022-05-17 doc/ks-chapter.tex
Revision book to page 82. deprecate a few symbols

@@
 Note the use of \coq's \texttt{sig} type in the theorem's statement, which
-relates the boolean function \texttt{limitb} defined on the \texttt{T1} data-type with a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.
+relates the boolean function \texttt{T1limit} defined on the \texttt{T1} data-type with a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.
 We can also state directly that $\lambda$ is a (strict) least upper bound of the elements of its canonical sequence.

### 2022-05-17 doc/ordinal-chapter.tex
Revision book to page 82. deprecate a few symbols

@@
 
-If \texttt{OB} is presumed to be correct, then we may consider that \texttt{OA} ``inherits'' its correctness from the bigger notation system \texttt{OB}.
-
 \label{types:SubON}
@@
 
-following definition
-(in ~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{ON\_Generic}).
+\noindent
+From~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{ON\_Generic}.
 

### 2022-05-18 doc/epsilon0-chapter.tex
revision book (chapter 4)

@@
 \vspace{4pt}
+\noindent
 \emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}
@@
 
-\begin{exercise}[\gaiasign]
-  Look for the \gaia-theorem which corresponds to \texttt{T1is\_succ\_iff}.
+\begin{exercise}
+  \gaiasign Look for the \gaia-theorem which corresponds to \texttt{T1is\_succ\_iff}.
 \end{exercise}
@@
 
-\begin{remark}
+\begin{remark}[Related work]
 \label{remark:a3pat}
-The alternate proof of well-foundedness using \'Evelyne Contejean's work on
-    recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the
+A proof of well-foundedness using \'Evelyne Contejean's work on
+    recursive path ordering~\cite{DershowitzRPO, a3pat} is also available in the
     library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0.Epsilon0rpo}.
+
+    In~\cite{Manolios2005}, Manolios and Vroom prove the well-foundedness of ordinal terms below $\epsilon_0$ by reduction to the natural order on the set of natural numbers.
  \end{remark}
@@
 
-We build an instance of \texttt{ON}, and prove its correctness w.r.t. Schutte's model (see Chapter~\ref{chap:schutte}).
-
+We are now able to build an instance of \texttt{ON}.
+
+\vspace{4pt}
+\noindent
 \emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}
@@
 
+We prove also that this notation is correct w.r.t. Schutte's model (see Chapter~\ref{chap:schutte}).
+
+\vspace{4pt}
+\noindent
 \emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}
@@
 
-    \section{A refinement of \texttt{E0} : an ordinal notation for \texorpdfstring{$\omega^\omega$}{omega\^omega}}
+    \section{An ordinal notation for \texorpdfstring{$\omega^\omega$}{omega\^omega}}
 
@@
 
-\begin{exercise}[not trivial]
+\begin{exercise}
   It may be interesting to write a \emph{direct} proof  of well-foundedness of the order in $\omega^\omega$ (\emph{i.e.}
@@
     First, we prove by transfinite induction on $\alpha$ that the restriction of $\oplus$ to the
-    interval $[0..\alpha)$ is commutative.
+    interval $[0,\alpha)$ is commutative.
 
@@
     So, a  natural question is `` Does there exist any strictly decreasing variant mapping
-    type \texttt{Hydra} into some interval $[0,\alpha)$ (where $\alpha <\epsilon_0$) for proving the termination of all hydra battles''.
-The next chapter is dedicated to a formal proof that there exists no such $\alpha$, even if we consider a restriction to the set of ``standard'' battles.
+    type \texttt{Hydra} into some interval $[0,\mu)$ (where $\mu <\epsilon_0$) for proving the termination of all hydra battles''.
+The next chapter is dedicated to a formal proof that there exists no such $\mu$, even if we consider a restriction to the set of ``standard'' battles.
 

### 2022-05-19 doc/epsilon0-chapter.tex
book re-reading

@@
 
-    \section*{Conclusion}
-
-    Let us recall three results we have proved so far.
-    \begin{itemize}
-    \item There exists a strictly decreasing variant which maps \texttt{Hydra} into 
-    the segment $[0,\epsilon_0)$ for proving the termination of any hydra battle
-    \item There exists \emph{no} such variant from \texttt{Hydra} into 
-    $[0,\omega^2)$, \emph{a fortiori} into $[0,\omega)$.
-    \end{itemize}
-
-    So, a  natural question is `` Does there exist any strictly decreasing variant mapping
-    type \texttt{Hydra} into some interval $[0,\mu)$ (where $\mu <\epsilon_0$) for proving the termination of all hydra battles''.
+    % \section*{Conclusion}
+
+    % Let us recall three results we have proved so far.
+    % \begin{itemize}
+    % \item There exists a strictly decreasing variant which maps \texttt{Hydra} into 
+    % the segment $[0,\epsilon_0)$ for proving the termination of any hydra battle
+    % \item There exists \emph{no} such variant from \texttt{Hydra} into 
+    % $[0,\omega^2)$, \emph{a fortiori} into $[0,\omega)$.
+    % \end{itemize}
+
+    % So, a  natural question is `` Does there exist any strictly decreasing variant mapping
+    % type \texttt{Hydra} into some interval $[0,\mu)$ (where $\mu <\epsilon_0$) for proving the termination of all hydra battles''.
 The next chapter is dedicated to a formal proof that there exists no such $\mu$, even if we consider a restriction to the set of ``standard'' battles.

### 2022-05-19 doc/ks-chapter.tex
book re-reading

--- 2022-05-08 doc/ks-chapter.tex
@@
 Our proofs are  constructive and require no axioms: they are  closed terms of the CIC, and are mainly composed on function definitions and proofs of properties of these functions. 
-They  share much theoretical material with Kirby and Paris', although they do not use any knowledge about Peano arithmetic nor model  theory.  The combinatorial arguments we use and implement
+They  borrow much theoretical material from Kirby and Paris, although they do not use any knowledge about Peano arithmetic nor model  theory.  The combinatorial arguments we use and implement
 come from 
@@
 \begin{itemize}
-\item If $\alpha\not=0$, then $\canonseq{\alpha}{i}<\alpha$. Thus canonical sequences can be used for proofs by transfinite induction or function definition by transfinite recursion
+\item If $\alpha\not=0$, then $\canonseq{\alpha}{i}<\alpha$. Thus canonical sequences can be used in proofs by transfinite induction or function definition by transfinite recursion
 \item If $\lambda$ is a limit ordinal, then $\lambda$ is the least upper bound of the set 
@@
 sequence $\alpha_0=\alpha, \alpha_1, \dots, \alpha_n=\beta$, where for every $k<n$, there exists some $i_k$ such that $\alpha_{k+1}=\canonseq{\alpha_k}{i_k}$
-\item Canonical sequences correspond tightly to rounds of hydra battles: if $\alpha\not=0$,
-then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1})$ in one round with
-the replication factor $i$ (Lemma \href{../theories/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
-\item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) battle from $\iota(\alpha)$ to $\iota(\beta)$.
 \end{itemize}
@@
 \begin{remark}
-  In~\cite{KS81}, canonical sequences are defined for any ordinal $\alpha <\epsilon_0$,
+  Canonical sequences are defined for any ordinal $\alpha <\epsilon_0$,
 by stating that if $\alpha$ is a successor ordinal $\beta+1$,  the sequence associated with 
@@
 
-This convention allows us to make total the function that maps any ordinal $\alpha$ and natural number $i$ to the ordinal $\canonseq{\alpha}{i}$.
-
+Thus, the function that maps any ordinal $\alpha$ and natural number $i$ to the ordinal $\canonseq{\alpha}{i}$ is total.
 \end{remark}
 
+\subsection{Canonical sequences and hydra battles}
+
+Canonical sequences correspond tightly to rounds of hydra battles: if $\alpha\not=0$,
+then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1})$ in one round with
+the replication factor $i$ (Lemma \href{../theories/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
+Thus, whenever $\beta<\alpha<\epsilon_0$, there exists a (free) battle from $\iota(\alpha)$ to $\iota(\beta)$.
+
+\subsection{Definitions}
 First, let us recall how canonical sequences are defined in~\cite{KS81}.
-For efficiency's sake, we decided not to implement directly K.\&S's definitions, but to define in \gallina{} simply typed structurally recursive functions which share the abstract properties which are used in the mathematical proofs\footnote{With a small difference: the $0$-th term of the canonical sequence is not the same in our development as in~\cite{KS81}.}.
+For efficiency's sake, we decided not to implement directly K.\&S's definitions, but to define in \gallina{} simply typed structurally recursive functions which share the abstract properties which are used in the mathematical proofs.
 
@@
 
-\paragraph*{\gaiasign} Our library~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon} transposes the  function \texttt{canon} into \gaia's type \texttt{T1}
-(please see Sect.\~ref{sect:gcanon}).
+\paragraph*{\gaiasign}
+The translation of \texttt{canon} compatible with \gaia's data-types is
+defined in ~\href{../theories/html/gaia_hydras.GCanon.html\#canon}{gaia\_hydras.GCanon} (please see Sect.~\ref{sect:gcanon}).
+
 \index{gaiabridge}{Canonical sequences}
@@
 
-Note the use of \coq's \texttt{sig} type in the theorem's statement, which
-relates the boolean function \texttt{limitb} defined on the \texttt{T1} data-type with a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.
+Note the use of \coq's \texttt{sig} type in the theorem's statement, which expresses a constructive view of the limit of a sequence: for any $\beta<\lambda$, we can compute an item of the canonical sequence of $\lambda$ which is greater than $\beta$.
 We can also state directly that $\lambda$ is a (strict) least upper bound of the elements of its canonical sequence.
@@
 Let us consider a kind of accessibility problem inside $\epsilon_0$: given two ordinals $\alpha$ and $\beta$, where $\beta<\alpha<\epsilon_0$, find a \emph{path} consisting of a finite sequence $\gamma_0=\alpha,\dots,\gamma_l=\beta$,
-where, for every $i<l$, $\gamma_i \not= 0$ \footnote{This condition allows us to ignore paths which end by a lot of useless $0$s.} and there exists some strictly positive integer $s_i$
+where, for every $i<l$, $\gamma_i \not= 0$
+and there exists some strictly positive integer $s_i$
 such that $\gamma_{i+1}=\canonseq{\gamma}{s_i}$.
@@
 Note that, given $\alpha$ and $\beta$, where $\beta < \alpha$, the sequence $s$ which leads from $\alpha$ to $\beta$ is not unique.
-Indeed, if $\alpha$ is a limit ordinal, the first element of $s$ can be any integer $i$ such that $\beta<\canonseq{\alpha}{i}$, and if $\alpha$ is a successor ordinal,
-then the sequence $s$ can start with any positive integer.
 
-For instance, we have also 
-$\omega*2 \xrightarrow[3,4,5,6]{}\omega$. 
+For instance, we have 
+$\omega\times 2 \xrightarrow[2]{}\omega$ and
+$\omega\times 2 \xrightarrow[3,4,5,6]{}\omega$. 
 Likewise,
-$\omega*2 \xrightarrow[1,2,1,4]{} 0$ and
-$\omega*2 \xrightarrow[3,3,3,3,3,3,3,3]{} 0$.
+$\omega\times 2 \xrightarrow[1,2,1,4]{} 0$ and
+$\omega\times 2 \xrightarrow[3,3,3,3,3,3,3,3]{} 0$.
 \end{remark}
@@
 The library~\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} transposes the  notion of path into \gaia's type \texttt{T1}
-(please see Sect.\~ref{sect:gpath}).
+(please see Sect.~\ref{sect:gpath}).
 
@@
 \noindent 
-From the lemma \texttt{canon\_LT}~\vref{lemma:canon_LT}, we can convert any path into an inequality on ordinals (by induction on paths).
+By the lemma \texttt{canon\_LT}~(Sct.\vref{lemma:canon_LT}), we can convert any path into an inequality on ordinals (by induction on paths).
 
@@
 
-Our goal is now  to transform any inequality $\beta<\alpha<\epsilon_0$ into a standard path $\alpha \xrightarrow[i,j]{} \beta$ for some $i$ and $j$, then into a standard battle
-from $\iota(\alpha+i)$ to $\iota(\beta)$. 
+Our goal is now  to transform any inequality $\beta<\alpha<\epsilon_0$ into a standard path $\alpha \xrightarrow[i,j]{} \beta$ for some $i$ and $j$, which corresponds to a standard battle
+from $\iota(\alpha)$ (at time $i$)  to $\iota(\beta)$
+(at time $j$) .
 Following~\cite{KS81}, we proceed in two stages:
@@
 \item we simulate plain (free) paths from $\alpha$ to $\beta$ with
-paths made of steps $(\gamma,\canonseq{\gamma}{n})$, \emph{with the same $n$ all along the path}
+paths made of steps $(\gamma,\canonseq{\gamma}{n})$, \emph{with the same $n$ all along the path}.
 \item we simulate any such path by a standard path.

### 2022-05-19 doc/epsilon0-chapter.tex
Update doc/epsilon0-chapter.tex
Co-authored-by: Karl Palmskog <palmskog@gmail.com>

@@
 \index{gaiabridge}{Type of well formed ordinal terms below $\epsilon_0$}
- Our library~\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines also a type \texttt{E0} (which doesn't exist in \gaia-\texttt{ssete9}).
+ Our library~\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} also defines a type \texttt{E0} (which doesn't exist in \gaia-\texttt{ssete9}).
 

### 2022-05-20 doc/large-sets-chapter.tex

@@
 Some notations (mainly names of fast-growing functions) of our development may differ slightly from the literature.
-Although this fact does not affect our proofs, we are preparing a future version where the names $F\_alpha$, $f\_alpha$, $H\_alpha$, etc., are fully consistent with the cited articles.
+Although this fact does not affect our proofs, we are preparing a future version where the names $F_\alpha$, $f_\alpha$, $H_\alpha$, etc., are fully consistent with the cited articles.
 
@@
 
-Now, consider any natural number $k>0$ any ordinal $0<\alpha<\epsilon_0$.  We would like to compute
+Now, let us consider some natural number $k>0$ and an ordinal $0<\alpha<\epsilon_0$.  We would like to compute
 a number $l$ such that the interval $[k,l]$ is $\alpha$-mlarge. So, 
@@
 In Sect.~\ref{sect:L-equations}, we use the \texttt{coq-equations} plug-in
-to define a function \texttt{L\_}, and prove its correctness w.r.t. its specification.
+to define  \texttt{L\_}, then prove its correctness w.r.t. its specification.
 

### 2022-05-23 doc/hydras.tex
update bibliography

@@
 The Reference Manual and several tutorials are available on the \coq website~\cite{Coq}.
-The first chapters of textbooks like \emph{Interactive Theorem Proving and Program Development}~\cite{BC04}, \emph{Software Foundations}~\cite{SF} or  \emph{Certified Programming with Dependent Types} ~\cite{chlipalacpdt2011} will give you the right background.
+The first chapters of textbooks like \emph{Interactive Theorem Proving and Program Development}~\cite{BC04},
+\emph{Software Foundations}~\cite{SF},
+\emph{Programs and Proofs}~\cite{Sergey:PnP},
+or  \emph{Certified Programming with Dependent Types} ~\cite{chlipalacpdt2011} will give you the right background.
 

### 2022-05-24 doc/ks-chapter.tex
small corrections

@@
 Our proofs are  constructive and require no axioms: they are  closed terms of the CIC, and are mainly composed on function definitions and proofs of properties of these functions. 
-They  borrow much theoretical material from Kirby and Paris, although they do not use any knowledge about Peano arithmetic nor model  theory.  The combinatorial arguments we use and implement
+They  borrow much theoretical material from Kirby and Paris, although they do not use any knowledge about Peano arithmetic nor about model  theory.  The combinatorial arguments we use and implement
 come from 
@@
 
+\texttt{Cor12} is a consequence of the following theorem (numbered \texttt{2.4} in Ketonen and Solovay's article), proven by transfinite induction on $\alpha$.
+
+\input{movies/snippets/Paths/KSThm24}
+
+\vspace{4pt}
+
 \subsubsection{Sketch of proof of \texttt{Cor12}}
@@
 
-First, let us prove the following lemma (numbered \texttt{2.4} in Ketonen and Solovay's article).
+\texttt{Cor12} is also proved by transfinite induction on $\alpha$.  Let us give  a sketch of its proof \footnote{This proof sketch is a slight simplification of the formal proof script:
+  The strictly positive indexes $i$ and $n$ stand for the
+  terms \texttt{(S i)} and \texttt{(S n)}. We do not explicit  the (simpler) case where the considered path is made of only one step.}
 
-\input{movies/snippets/Paths/KSThm24}
-
-We prove this lemma by transfinite induction on $\alpha$.
 Let us consider a path $\alpha \xrightarrow [i]{} \beta$ $(i>0)$. Its first step is
@@
 Let us  consider two ordinals  $\beta<\alpha<\epsilon_0$. By induction on $\alpha$,
-we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some integer. Applying corollary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
+we decompose any inequality $\beta<\alpha$ into $\beta < \canonseq{\alpha}{i}< \alpha$, where $i$ is some natural number.
+Applying corollary \texttt{Cor12'} we build a $n$-path from $\beta$ to $\alpha$,
 where $n$ is the maximum of the indices $i$ met in the induction.
@@
 Let $l$ be the length of the path from $\alpha$ to $\gamma$.  
-This step of the proof is illustrated in figure~\vref{fig:belle-preuve-2}.
+This step of the proof is illustrated in figure~\vref{fig:belle-preuve-2} \footnote{Please note that in these figures, smaller ordinals are represented on the right side!}.
 

### 2022-05-24 doc/hydra-chapter.tex
small corrections

@@
 In the following definition \texttt{i} is associated with the round number (or date, if we consider a discrete time), and \texttt{a}, \texttt{b}, \texttt{c} respectively associated with the number of occurrences of \texttt{h2}, \texttt{h1} and heads connected to the hydra's foot.
-For convenience, we do not use the type \texttt{state} of the preceeding section, but consider the round numbers and the number of hydras \texttt{h2}, \texttt{h1} and heads as separate arguments of the relation (which is no more ---formally--- ``binary'').
+For convenience, we do not use the type \texttt{state} of the preceding section, but consider the round numbers and the number of hydras \texttt{h2}, \texttt{h1} and heads as separate arguments of the relation (which is no more ---formally--- ``binary'').
 

### 2022-05-24 doc/JFLA2022/Slides/Slides.tex
update JFLA Slides
work on JFLA slides

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\begin{frame}
-  \tableofcontents
-\end{frame}
+% \begin{frame}
+%   \tableofcontents
+% \end{frame}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=mathcomp, node distance=7.7cm] (equations) { Equations };
+
+   \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=equations, node distance=3cm] (libhyps) { LibHyps };
 \end{scope}
@@
 \draw[->] (equations) -- (hydras) ;
+\draw[->] (libhyps) -- (hydras) ;
+
 \draw[->] (mathcomp) -- (gaia) ;
@@
           allowing to present and compare various formalization and proving techniques.
+          \item It allows the development team to experiment various techniques of CI/CD and project organization and maintenance.
       \end{itemize}
@@
     \item Formalization by Minki Cho, using a different representation.
-{\color{darkorange}
+{\color{blue}
 \begin{verbatim}
@@
   \end{block}
-\end{frame}
 
-  \begin{frame}
-\frametitle{Call to contributors}
   \begin{block}{}
@@
       \item New chapters
-      \item Any correction or improvement (text and/or proofs)
+      \item Any correction or suggestion of improvement (text and/or proofs)
 
@@
 \end{document}
-
-\begin{frame}
-  \frametitle{Ordinals in \coq}
-
-  \begin{block}{}
-    Ordinals less than {\color{mathcolor}$\omega^\omega, \epsilon_0,  \Gamma_0$} are represented as \emph{ordinal notations}: comparison, canonical sequences, partition limit/successors/zero ordinals are implemented by certified functions, allowing the user to write proofs by computation.
-
-    \begin{footnotesize}
-      \inputsnippets{./ONDef}
-
-      \inputsnippets{../E0_Ex42}
-    \end{footnotesize}
-
-  \end{block}
-\end{frame}

### 2022-06-02 doc/JFLA2022/Slides/Slides.tex
small changes to JFLA slides

@@
     \frametitle{Paths and accessibility}
-    \begin{block}{Theorem (K. \& S., 1981) }
+    \begin{block}{ }
       % If $\beta<\alpha<\epsilon_0$, there exists $i$ and $l$ in $\mathbb{N}$ such that $ \alpha \rplus{i, i+1, i+2, \dots,l} \beta$.
-
-      \vspace{6pt}
 
@@
 \node (alpha) at (0,0) {$\alpha$};
-    \node (beta) at (40, 0){$\beta$};
+    \node (beta) at (35, 0){\textcolor{red}{$\beta$}};
+   \node (zero) at (40, 0){$0$};
 
   \draw[->, very thick,black] (alpha)--  node [above]{$>$} (beta);
-
+ \draw[->, very thick,black] (beta)--  node [above]{$>$} (zero);
   \node (alpha1) at (5,5) {};
@@
     \draw [->, dashed,very thick,red] (alpha3)-- node [below]{$i+3$}  (alpha4);
-  \node (dots) at (30,5) {$\dots$};
+    \only<1>{\node (dots) at (26,5) {$\;\dots\;?$};}
+     \only<2>{\node (dots) at (26,5) {\textcolor{red}{$\;\dots\;$}};}
   \node (search) at (32,5){};
- \draw [->, dashed, very thick,red] (search)-- node [below, rotate=-26]{$l$}  (beta);
+ \only<2>{\draw [->, dashed, very thick,red] (search)-- node [below, rotate=-26]{$l$}  (beta);}
 \end{tikzpicture}
 
+\only<2>{OK for large enough $i$ (K. \& S., 1981);}
+\vspace{6pt}
+
 %\pause
-\textbf{Remark:} The proof is not trivial. The built path could miss $\beta$!
+\only<2>{\textbf{Remark:} The proof is not trivial. The built path could miss $\beta$!}
 
@@
 
-\textbf{Cor:} There is no decreasing measure mapping hydras to  $[0, \mu)$ (for any $\mu<\epsilon_0$)  for proving the termination of the considered kind of hydra battles.
+\only<2>{\textbf{Cor:} There is no decreasing measure mapping hydras to  $[0, \mu)$ (for any $\mu<\epsilon_0$)  for proving the termination of the considered kind of hydra battles.}
 

### 2022-06-02 doc/JFLA2022/Slides/canonDef.tex
small changes to JFLA slides

@@
     \begin{input}
-      \PY{k+kn}{Definition}~\PY{n+nf}{canon}~\PY{o}{(}\PY{n+nv}{a}\PY{o}{:}~\PY{n}{T1}\PY{o}{)}~\PY{o}{(}\PY{n+nv}{i}\PY{o}{:}\PY{n}{nat}\PY{o}{)}~\PY{o}{:}~\PY{n}{T1}~\PY{o}{:=}\nl
-      ~~\PY{n}{h2g}~\PY{o}{(}\PY{n}{hcanon}~\PY{o}{(}\PY{n}{g2h}~\PY{n}{a}\PY{o}{)}~\PY{n}{i}\PY{o}{).}
+      \PY{o}{\PYZsh{}[}\PY{k+kn}{global}\PY{o}{]}~\PY{k+kn}{Notation}~\PY{n+nf}{hcanon}~\PY{o}{:=}~\PY{n}{Epsilon0}\PY{o}{.}\PY{n}{Canon}\PY{o}{.}\PY{n}{canon}\PY{o}{.}~\nl
+    \end{input}
+  \end{sentence}
+  \sep
+  \begin{txt}
+    \nl
+  \end{txt}
+  \sep
+  \begin{sentence}
+    \begin{input}
+      \PY{k+kn}{Definition}~\PY{n+nf}{canon}~\PY{o}{(}\PY{n+nv}{a}\PY{o}{:}~\PY{n}{T1}\PY{o}{)}~\PY{o}{(}\PY{n+nv}{i}\PY{o}{:}\PY{n}{nat}\PY{o}{)}~\PY{o}{:}~\PY{n}{T1}~\PY{o}{:=}~~\PY{n}{h2g}~\PY{o}{(}\PY{n}{hcanon}~\PY{o}{(}\PY{n}{g2h}~\PY{n}{a}\PY{o}{)}~\PY{n}{i}\PY{o}{).}
     \end{input}

### 2022-06-02 doc/JFLA2022/Slides/gcanonSLE.tex
small changes to JFLA slides

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Lemma}~\PY{n+nf}{canonS\PYZus{}LE}~\PY{n+nv}{a}~\PY{n+nv}{n}\PY{o}{:}\nl
      ~~~~\PY{n}{T1nf}~\PY{n}{a}~\PY{o}{\PYZhy{}\PYZgt{}}~\PY{n}{canon}~\PY{n}{a}~\PY{n}{n}\PY{o}{.+}\PY{l+m+mi}{1}~\PY{o}{\PYZlt{}=}~\PY{n}{canon}~\PY{n}{a}~\PY{n}{n}\PY{o}{.+}\PY{l+m+mi}{2}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Proof}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~\PY{n+nb}{rewrite}~\PY{o}{\PYZhy{}(}\PY{n}{h2g\PYZus{}g2hK}~\PY{n}{a}\PY{o}{)}~\PY{o}{\PYZhy{}}\PY{n}{nf\PYZus{}ref}~\PY{o}{/}\PY{n}{canon}~\PY{n}{g2h\PYZus{}h2gK}~\PY{o}{=\PYZgt{}}~\PY{n}{Hnf}\PY{o}{;}\nl
      ~~\PY{n+nb+bp}{by}~\PY{n+nb}{apply}~\PY{n}{le\PYZus{}ref}\PY{o}{,}~\PY{n}{canonS\PYZus{}LE}\PY{o}{.}~\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Qed}\PY{o}{.}
    \end{input}
  \end{sentence}
\end{alectryon}

### 2022-06-02 doc/JFLA2022/Slides/gcanonLimitOf.tex
small changes to JFLA slides

@@
     \begin{input}
-      ~\PY{k+kn}{Lemma}~\PY{n+nf}{gcanon\PYZus{}limit\PYZus{}of}~\PY{n+nv}{lambda}~\PY{o}{(}\PY{n+nv}{Hnf}~\PY{o}{:}~\PY{n}{T1nf}~\PY{n}{lambda}\PY{o}{)}~\PY{o}{(}\PY{n+nv}{Hlim}~\PY{o}{:}~\PY{n}{T1limit}~\PY{n}{lambda}\PY{o}{)}~\PY{o}{:}\nl
-      ~~~\PY{n}{limit\PYZus{}of}~\PY{o}{(}\PY{n}{canon}~\PY{n}{lambda}\PY{o}{)}~\PY{n}{lambda}\PY{o}{.}\nl
-    \end{input}
-  \end{sentence}
-  \sep
-  \begin{sentence}
-    \begin{input}
-      ~\PY{k+kn}{Proof}\PY{o}{.}\nl
-    \end{input}
-  \end{sentence}
-  \sep
-  \begin{sentence}
-    \begin{input}
-      ~~~\PY{n+nb}{split}~\PY{o}{=\PYZgt{}}~\PY{o}{//.}\nl
-    \end{input}
-  \end{sentence}
-  \sep
-  \begin{sentence}
-    \begin{input}
-      ~~~\PY{n+nb}{move}~\PY{o}{=\PYZgt{}}~\PY{n}{n}~\PY{n}{m}~\PY{n}{Hnm}\PY{o}{;}~\PY{n+nb}{apply}~\PY{n}{gcanon\PYZus{}limit\PYZus{}mono}~\PY{o}{=\PYZgt{}}~\PY{o}{//.}\nl
-    \end{input}
-  \end{sentence}
-  \sep
-  \begin{sentence}
-    \begin{input}
-      ~~~\PY{n+nb}{apply}~\PY{n}{gcanon\PYZus{}limit\PYZus{}v2}~\PY{o}{=\PYZgt{}}~\PY{o}{//.}\nl
-    \end{input}
-  \end{sentence}
-  \sep
-  \begin{sentence}
-    \begin{input}
-      ~\PY{k+kn}{Qed}\PY{o}{.}~~~~
+      ~\PY{k+kn}{Lemma}~\PY{n+nf}{canon\PYZus{}limit\PYZus{}of}~\PY{n+nv}{lambda}~\PY{o}{(}\PY{n+nv}{Hnf}~\PY{o}{:}~\PY{n}{T1nf}~\PY{n}{lambda}\PY{o}{)}~\PY{o}{(}\PY{n+nv}{Hlim}~\PY{o}{:}~\PY{n}{T1limit}~\PY{n}{lambda}\PY{o}{)}~\PY{o}{:}\nl
+      ~~~\PY{n}{limit\PYZus{}of}~\PY{o}{(}\PY{n}{canon}~\PY{n}{lambda}\PY{o}{)}~\PY{n}{lambda}\PY{o}{.}
     \end{input}

### 2022-06-02 doc/JFLA2022/Slides/Slides.tex
small changes to JFLA slides

@@
 \begin{scope}[yshift=-5cm,xshift=4cm]
-\node[rectangle, draw=black,fill=red!80,minimum height=0.5cm,minimum width=1.5cm] (book) { \Hydras};  
+  \node[rectangle, draw=black,fill=red!60,minimum height=0.5cm,minimum width=1.5cm] (book) {hydras.pdf};
+    \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm, left of=book, node distance=6cm] (latex) {hydras.tex};
 \end{scope}
@@
 \draw[lookcolor,->,dashed] (goedel) -- (hydras.north) ;
-\draw[lookcolor,->,dashed] (cantor) -- (hydras.north east) ;
+\draw[lookcolor,->,dashed] (cantor) -- (hydras.north) ;
 \draw[lookcolor,->,dashed] (cantor.south) -- (gaia.north west) ;
@@
 \draw[->, red] (hydras.south) -- (book) ;
-
+\draw[->, red] (latex) -- (book) ;
 \end{tikzpicture}
@@
       \inputsnippets{./canonDef}
+      \inputsnippets{./gcanonSLE}
       \inputsnippets{./gcanonLimitOf}}
+
   \end{block}

### 2022-06-02 doc/JFLA2022/Slides/Slides.tex
Micro changes to the JFLA slides.

@@
       \begin{itemize}
-      \item Created in 2018,  run by volunteer Coq users on GitHub
-      \item At present, over than 50 projects, maintained by over than 30 volunteers
+      \item Created in 2018,  run by volunteer Coq users on GitHub.
+      \item At present, over 50 projects, maintained by over 30 volunteers.
       \item Some packages were unmaintained before their integration to \community. According to their scientific or historical interest, we decided to revive them.
@@
    \item Experimental platform for the  development of \textcolor{lookcolor}{maintained} and \textcolor{lookcolor}{documented} libraries of formal proofs.
-    \item A project of \community, developped on Github
+    \item A project of \community, developped on GitHub
       \begin{itemize}
@@
         \end{itemize}
-           \item A book in pdf (300p), generated  with \textcolor{lookcolor}{\alectr} and \LaTeX, continuously updated and deployed
+           \item A book in PDF (300p), generated  with \textcolor{lookcolor}{\alectr} and \LaTeX, continuously updated and deployed
     \end{itemize}
@@
   \begin{block}{}
-    \Hydras presents a consistent set of examples, which allow us to formalize 
+    \Hydras presents a consistent set of examples, which allows us to formalize
     some \textcolor{mathcolor}{discrete math}, discuss

### 2022-06-06 doc/JFLA2022/Slides/Slides.tex
small changes to JFLA slides
micro-modifs

@@
       \item A book in PDF (300p), generated  with \textcolor{lookcolor}{\alectr} and \LaTeX, continuously updated and deployed
+  \item          When appropriate, we present various proofs of the same statement, and interesting false starts.
     \end{itemize}
@@
   \begin{block}{}
+
  \Hydras presents a consistent set of examples, which allows us to formalize
     some \textcolor{mathcolor}{discrete math}, discuss
-    \textcolor{coqstylecolor}{\coq style, specification and proof patterns}, and depend on \textcolor{plugincolor}{a few plug-ins}.
+    \textcolor{coqstylecolor}{\coq style, specification and proof patterns}.
+
+    %, and depend on \textcolor{plugincolor}{a few plug-ins}.
+
   \end{block}
@@
 \node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington\footnote{O. Caprotti \& M. Oostdijk}};
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (cantor) { Cantor\footnote{C. \& E. Contejean}};
+\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (cantor) { Cantor\footnote{C. \& E. Contejean
+    \textcolor{pauillac}{(Pauillac 2006)}}};
 
@@
 
-\only<2>{\textbf{Cor:} There is no decreasing measure mapping hydras to  $[0, \mu)$ (for any $\mu<\epsilon_0$)  for proving the termination of the considered kind of hydra battles.}
+\only<2>{\textbf{Cor:} There is no decreasing measure mapping hydras to  $[0, \mu)$ (for any $\mu<\epsilon_0$)  for proving the termination of the considered kind of hydra battles
+  (extends \textcolor{banyuls}{Banyuls 2018}).}
 
@@
           allowing to present and compare various formalization and proving techniques.
-          \item It allows the development team to experiment various techniques of CI/CD and project organization and maintenance.
+          \item It allows the developers of the project to experiment various techniques of CI/CD and project organization and maintenance.
       \end{itemize}

### 2022-06-11 doc/JFLA2022/Slides/Slides.tex
small changes to JFLA slides

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-  \frametitle{What hydras are talking about?}
-  \begin{block}{}
+% \begin{frame}
+%   \frametitle{What hydras are talking about?}
+%   \begin{block}{}
 
- \Hydras presents a consistent set of examples, which allows us to formalize
-    some \textcolor{mathcolor}{discrete math}, discuss
-    \textcolor{coqstylecolor}{\coq style, specification and proof patterns}.
+%  \Hydras presents a consistent set of examples, which allows us to formalize
+%     some \textcolor{mathcolor}{discrete math}, discuss
+%     \textcolor{coqstylecolor}{\coq style, specification and proof patterns}.
 
-    %, and depend on \textcolor{plugincolor}{a few plug-ins}.
+%     %, and depend on \textcolor{plugincolor}{a few plug-ins}.
 
-  \end{block}
-  \begin{block}{}
-    \begin{description}
-    \item[hydra battles:]
+%   \end{block}
+%   \begin{block}{}
+%     \begin{description}
+%     \item[hydra battles:]
 
-      \textcolor{mathcolor}{ordinals},
-      \textcolor{mathcolor}{rapidly growing functions},
-      \textcolor{mathcolor}{primitive recursive functions}, 
-      \textcolor{coqstylecolor}{dependently typed functions},
-      \textcolor{coqstylecolor}{operational type classes},
-      \textcolor{coqstylecolor}{proofs of well-foundedness},
-      \textcolor{coqstylecolor}{impossibility proofs},
-      \textcolor{coqstylecolor}{uniqueness of identity proofs (UIP)},
-      \textcolor{coqstylecolor}{indefinite description (Hilbert's epsilon operator)}, 
-      \textcolor{plugincolor}{equations},
-      \textcolor{plugincolor}{ackermann (goedel)},
-      \textcolor{plugincolor}{gaia}
+%       \textcolor{mathcolor}{ordinals},
+%       \textcolor{mathcolor}{rapidly growing functions},
+%       \textcolor{mathcolor}{primitive recursive functions}, 
+%       \textcolor{coqstylecolor}{dependently typed functions},
+%       \textcolor{coqstylecolor}{operational type classes},
+%       \textcolor{coqstylecolor}{proofs of well-foundedness},
+%       \textcolor{coqstylecolor}{impossibility proofs},
+%       \textcolor{coqstylecolor}{uniqueness of identity proofs (UIP)},
+%       \textcolor{coqstylecolor}{indefinite description (Hilbert's epsilon operator)}, 
+%       \textcolor{plugincolor}{equations},
+%       \textcolor{plugincolor}{ackermann/goedel},
+%       \textcolor{plugincolor}{gaia}
 
-    \item[addition chains (exponentiation algorithms):]
-      \textcolor{mathcolor}{algorithmics},
-      \textcolor{mathcolor}{arithmetic},
-      \textcolor{coqstylecolor}{operational type classes},
-      \textcolor{coqstylecolor}{monadic notations},
-      \textcolor{coqstylecolor}{PHOAS},
-      \textcolor{coqstylecolor}{parametricity}, 
-      \textcolor{plugincolor}{paramcoq}
-    \end{description}
-  \end{block}
-\end{frame}
+%     \item[addition chains (exponentiation algorithms):]
+%       \textcolor{mathcolor}{algorithmics},
+%       \textcolor{mathcolor}{arithmetic},
+%       \textcolor{coqstylecolor}{operational type classes},
+%       \textcolor{coqstylecolor}{monadic notations},
+%       \textcolor{coqstylecolor}{PHOAS},
+%       \textcolor{coqstylecolor}{parametricity}, 
+%       \textcolor{plugincolor}{paramcoq}
+%     \end{description}
+%   \end{block}
+% \end{frame}
 
@@
     \centering
-    \begin{tikzpicture}[very thick, scale=0.25]
+    \begin{tikzpicture}[very thick, scale=0.20]
       \node (h1) at (4,0){$\bullet$};
@@
     \centering
-    \begin{tikzpicture}[very thick, scale=0.25]
+    \begin{tikzpicture}[very thick, scale=0.2]
       \node (hn1) at (4,0){$\bullet$};
@@
   \end{figure}
+
+ Cf.  \textcolor{pauillac}{Pauillac 2006} and \textcolor{banyuls}{Banyuls 2018}. Today, let us focus on some combinatorial aspects (number of rounds of a battle).
 \end{frame}
@@
 
-\only<2>{OK for large enough $i$ (K. \& S., 1981);}
+\only<2>{OK for large enough $i$ (K. \& S., 1981)}
 \vspace{6pt}
 
-%\pause
-\only<2>{\textbf{Remark:} The proof is not trivial. The built path could miss $\beta$!}
+% \pause
+\only<1>{Goal: reach \textcolor{red}{$\beta$} from $\alpha$.}
+\only<2>{\textbf{Remark:} The proof is not trivial. The deterministic red path could miss $\beta$!}
 
@@
 \only<2>{\textbf{Cor:} There is no decreasing measure mapping hydras to  $[0, \mu)$ (for any $\mu<\epsilon_0$)  for proving the termination of the considered kind of hydra battles
-  (extends \textcolor{banyuls}{Banyuls 2018}).}
+  (stronger than \textcolor{banyuls}{Banyuls 2018}).}
 
@@
 \textbf{Wainer hierarchy (variant)}
-
-We use also the following family in our proofs.
 
@@
     \end{quote}
-    (Proof by induction on any primitive recursive function of any arity.)
+  %  (Proof by induction on any primitive recursive function of any arity.)
   \end{block}
@@
   \frametitle{Compatibility with Gaia}
-  \begin{block}{From \Hydras to \gaia}
+  \begin{block}{}
     {\footnotesize
@@
   \end{block}
-  \begin{block}{}
-    We experiment also the other direction (importing \texttt{ssete9} lemmas to \Hydras).
-  \end{block}
+  % \begin{block}{}
+  %   We experiment also the other direction (importing \texttt{ssete9} lemmas to \Hydras).
+  % \end{block}
 \end{frame}
@@
 \section{Presenting proofs with \alectr}
-\begin{frame}
+\begin{frame}[fragile]
   \frametitle{Presenting proofs with \alectr}
-  \begin{block}{}
-    \begin{itemize}
- \item     The continuously evolving version of the book is continuously deployed from the master branch to the \community website.
+ %  \begin{block}{}
+%     \begin{itemize}
+%  \item     The continuously evolving version of the book is continuously deployed from the master branch to the \community website.
 
-     \item
-      \coq libraries contain \emph{snippet declarations} which are converted during the compilation of the book into \LaTeX\xspace blocks which are inserted into the book.
+%      \item
+%       \coq libraries contain \emph{snippet declarations} which are converted during the compilation of the book into \LaTeX\xspace blocks which are inserted into the book.
 
-    \item The order in which \coq code, goals and answers are presented in the book is determined by the difficulty of the mathematical examples, or the assumed level in \coq of the reader, not by the structure of the libraries.
+%     \item The order in which \coq code, goals and answers are presented in the book is determined by the difficulty of the mathematical examples, or the assumed level in \coq of the reader, not by the structure of the libraries.
 
-    \end{itemize}
-  \end{block}
-\end{frame}
+%     \end{itemize}
+%   \end{block}
+% \end{frame}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{frame}[fragile]
+% \begin{frame}[fragile]
 \begin{alltt}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
   \frametitle{Perspectives}
-  \begin{block}{Library on ordinals}
+  \begin{block}{}
     \begin{itemize}
-  \item \gaiahydra:
-    \begin{itemize}
-    \item Remove definition duplicates
-     \item types \texttt{T2}, \texttt{T3}, connection between 
-    Schütte's model with \texttt{ssete10} (Bourbaki' set theory).
-    \end{itemize}
 
-    \item Formalization by Minki Cho, using a different representation.
+    \item  Links with other representations of ordinals:
+
+      \noindent
+         Example (Minki Cho):
+  \begin{scriptsize}
     {\color{blue}
@@
 Inductive Ord.t :=
-| Ord.build: 
-    forall (A: Type), (A -> Ord.t) -> Ord.t.
+| Ord.build: forall (A: Type), (A -> Ord.t) -> Ord.t.
 \end{verbatim}
       }
-    \end{itemize}
+ \end{scriptsize}
 
-  \end{block}
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{frame}
-  \begin{block}{}
-    \begin{itemize}
       \item More results about hyper-complexity and fast growing functions
@@
     \item More examples and exercises
-    \item etc.
      \end{itemize}

### 2022-06-11 doc/JFLA2022/Slides/Slides.tex
small changes to JFLA slides

@@
     \begin{itemize}
-   \item Experimental platform for the  development of \textcolor{lookcolor}{maintained} and \textcolor{lookcolor}{documented} libraries of formal proofs.
-    \item A project of \community, developped on GitHub
-      \begin{itemize}
-      \item CI/CD
-        \item Tools for collaboration (issues, PR, etc.)
-        \end{itemize}
-      \item A book in PDF (300p), generated  with \textcolor{lookcolor}{\alectr} and \LaTeX, continuously updated and deployed
-  \item          When appropriate, we present various proofs of the same statement, and interesting false starts.
+    \item A project of \community:  experimental platform for the  development of
+      \textcolor{lookcolor}{documented} and \textcolor{lookcolor}{maintained} libraries of formal proofs: 
+    CI/CD,  Tools for collaboration (issues, PR, etc.)
+
+  \item A book in PDF (300p), generated  with \textcolor{lookcolor}{\alectr} and \LaTeX, continuously updated and deployed.
+
+    This book is structured around a few examples (algorithmics, ordinal numbers).
+        When appropriate, various proofs of the same statement, and interesting false starts are presented.
     \end{itemize}
@@
 
- Cf.  \textcolor{pauillac}{Pauillac 2006} and \textcolor{banyuls}{Banyuls 2018}. Today, let us focus on some combinatorial aspects (number of rounds of a battle).
+ Cf.  \textcolor{pauillac}{Pauillac 2006} and \textcolor{banyuls}{Banyuls 2018}. Today, let us focus on some combinatorial aspects (\emph{e.g.} number of rounds of a battle).
 \end{frame}
@@
 
-\only<2>{OK for large enough $i$ (K. \& S., 1981)}
+\only<2>{{\color{lookcolor}OK for large enough $i$ (K. \& S., 1981)}}
 \vspace{6pt}
@@
 
- Let us define \textcolor{mathcolor}{$H'_\alpha(i)$} as the least $l$ such that \textcolor{mathcolor}{$\alpha_l=0$} (Figure page.~\pageref{fig:accessibility}). 
  {\color{mathcolor}
@@
 
-\vspace{4pt}
+\textcolor{mathcolor}{$H'_\alpha(i)$} is the least $l$ such that \textcolor{mathcolor}{$\alpha_l=0$} (cf. previous slide)).
+
+\vspace{6pt}
 
@@
 
-    \item  Links with other representations of ordinals:
-
-      \noindent
-         Example (Minki Cho):
-  \begin{scriptsize}
+    \item  Links with other representations of ordinals, \emph{e.g.}:
+  \begin{footnotesize}
     {\color{blue}
 \begin{verbatim}
+(* Minki Cho *)
 Inductive Ord.t :=
@@
       }
- \end{scriptsize}
+ \end{footnotesize}
 
@@
   \begin{block}{}
-  {\color{plugincolor}\Hydras} and  {\color{plugincolor}\community} are on GitHub. {\color{cyan}Please contribute, reuse!}
+  {\color{plugincolor}\Hydras} and  {\color{plugincolor}\community} are on GitHub. {\color{cyan}Please clone, fork, contribute!}
       \begin{itemize}
       \item New chapters
-      \item Any correction or suggestion of improvement (text and/or proofs)
+      \item Any correction or suggestion of improvement (text and/or proofs).
 

### 2022-06-15 doc/JFLA2022/Slides/Slides.tex
Slides (small changes)

@@
     \begin{block}{Formalization}
-
+      {\color{mathcolor}
       $$\omega^{\omega^2+1}+1 \rplus{1} \omega^{\omega^2+1}
@@
 
-      $$ \omega^{\omega^2+1}+1 \rplus{1,2,3,4,\dots,x} 0$$
+      $$ \omega^{\omega^2+1}+1 \rplus{1,2,3,4,\dots,x} 0$$}
     \end{block}
@@
       \begin{block}{Small steps}
-        At time $i$, 
-           $\alpha \rplus{i} \canonseq{\alpha}{i}$,
-        where $\canonseq{\alpha}{\_}$ is the \emph{canonical sequence} of $\alpha$.
+        At the $i$-th round, 
+           \textcolor{mathcolor}{$\alpha \rplus{i} \canonseq{\alpha}{i}$},
+           where \textcolor{mathcolor}{$\canonseq{\alpha}{\_}$} is the \emph{canonical sequence} of \textcolor{mathcolor}{$\alpha$}.
+
              \begin{itemize}
-        \item $\canonseq{0}{i}=0$
-        \item $\canonseq{\beta+1}{i}=\beta$
-        \item If $\lambda$ is a limit ordinal, then $\lambda$ is the $\omega$-limit of $\canonseq{\lambda}{\_}$.
+             \item  \textcolor{mathcolor}{$\canonseq{0}{i}=0$}
+             \item \textcolor{mathcolor}{$\canonseq{\beta+1}{i}=\beta$}
+             \item If \textcolor{mathcolor}{$\alpha$} is a limit ordinal, then \textcolor{mathcolor}{$\alpha$} is the limit of the sequence
+               \textcolor{mathcolor}{$\canonseq{\alpha}{i}_{ (i\in\mathbb{N})}$}.
              \end{itemize}
@@
 \begin{frame}[fragile]
-  \begin{block}{Examples}
 
-      {\color{mathcolor}
-
-    \begin{align*}
-      H'_{\omega^3}(k) \geq\, & \textit{hyper\_exp2}(k+1) \\
-      H'_{\omega^\omega}(k)=\,& \textbf{let}\,F\;f\;i := f^{(i+1)}(i)\,
-                      \textbf{in}\, F^{(k)}(\lambda k.\,2k+1)\,k\\
-    \end{align*}
-       }%mathcolor
-  \end{block}
-
-\begin{block}{Lemmas (after K.\& S.)}
+\begin{block}{Lemmas (after K.\,\&\,S.)}
    \begin{itemize} 
@@
        (Proof by mutual transfinite induction)
+ \end{block}
+
+ \begin{block}{Examples}
+
+      {\color{mathcolor}
+
+        \begin{align*}
+        H'_{\omega^\omega}(k)=\,& \textbf{let}\,F\;f\;i := f^{(i+1)}(i)\,
+                                  \textbf{in}\, F^{(k)}(\lambda k.\,2k+1)\,k\\
+      H'_{\omega^3}(k) \geq\, & \textit{hyper\_exp2}(k+1) \\
+      H'_{\omega^\alpha}(k)\geq\,& F_\alpha(k)\; (k>0)\\
+    \end{align*}
+       }%mathcolor
   \end{block}

### 2022-06-15 doc/JFLA2022/Slides/Slides.tex
Slides (small changes)

@@
 
+ \textbf{Definition:} \textcolor{mathcolor}{$H'_\alpha(i)$} is the least $l$ such that \textcolor{mathcolor}{$\alpha_l=0$} (cf. previous slide).
+
  {\color{mathcolor}
@@
   H'_{\beta+1}(i) &=H'_\beta(i+1) \\
-  H'_\lambda(i) &= H'_{(\canonseq{\lambda}{i+1})}(i)  \quad\textit{($\lambda$ limit ordinal)} 
+  H'_\alpha(i) &= H'_{(\canonseq{\alpha}{i+1})}(i)  \quad\textit{($\alpha$ limit ordinal)} 
 \end{align}}
-
-\textcolor{mathcolor}{$H'_\alpha(i)$} is the least $l$ such that \textcolor{mathcolor}{$\alpha_l=0$} (cf. previous slide)).
 
@@
 F_{\beta+1}(i)&= (F_\beta)^{(i+1)}(i)\\
-F_\lambda(i) &= F_{(\canonseq{\lambda}{i})} (i) \quad \textit{($\lambda$ limit ordinal)}
+F_\alpha(i) &= F_{(\canonseq{\alpha}{i})} (i) \quad \textit{($\alpha$ limit ordinal)}
 \end{align}}}
@@
     \begin{itemize}
-    \item $F_\alpha$ is primitive recursive iff $\alpha<\omega$.
+    \item \textcolor{mathcolor}{$F_\alpha$} is primitive recursive iff \textcolor{mathcolor}{$\alpha<\omega$}.
      \item  
     The function which computes the length of hydra battles
-    for $\alpha$ (according to the initial index) is not primitive recursive
-    (for $\alpha\geq \omega^\omega$).
-
+    for \textcolor{mathcolor}{$\alpha$} (according to the initial index) is not primitive recursive
+    (for $\textcolor{mathcolor}{\alpha\geq \omega^\omega}$).
     \end{itemize}

### 2022-06-19 doc/JFLA2022/Slides/Slides.tex
Slides (small changes)

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \begin{frame}
+\begin{frame}[label={accessibility}]
+
     \frametitle{Paths and accessibility}
@@
 \only<2>{\textbf{Cor:} There is no decreasing measure mapping hydras to  $[0, \mu)$ (for any $\mu<\epsilon_0$)  for proving the termination of the considered kind of hydra battles
-  (stronger than \textcolor{banyuls}{Banyuls 2018}).}
+  (stronger than \textcolor{banyuls}{Banyuls 2018}).
+}
 
@@
 
-\begin{frame}[fragile]
+\begin{frame}[fragile, label={fgflemmas}]
 
@@
 
-\begin{frame}
-
+\begin{frame}[label={fgftheorems}]
+\label{thms-fastgrowing}
   \begin{block}{Theorems}
@@
   \end{block}
+  The lemmas mentionned in previous slides (accessibility, fast growing functions) are ``ported'' this way from \texttt{Hydra-battles}   into \texttt{Gaia-Hydras}.
   % \begin{block}{}

### 2022-06-19 doc/JFLA2022/Slides/gcanonSLE.tex
Slides (small changes)

@@
     \begin{input}
-      \PY{k+kn}{Lemma}~\PY{n+nf}{canonS\PYZus{}LE}~\PY{n+nv}{a}~\PY{n+nv}{n}\PY{o}{:}\nl
-      ~~~~\PY{n}{T1nf}~\PY{n}{a}~\PY{o}{\PYZhy{}\PYZgt{}}~\PY{n}{canon}~\PY{n}{a}~\PY{n}{n}\PY{o}{.+}\PY{l+m+mi}{1}~\PY{o}{\PYZlt{}=}~\PY{n}{canon}~\PY{n}{a}~\PY{n}{n}\PY{o}{.+}\PY{l+m+mi}{2}\PY{o}{.}\nl
+      \PY{k+kn}{Lemma}~\PY{n+nf}{canonS\PYZus{}LE}~\PY{n+nv}{a}~\PY{n+nv}{n}\PY{o}{:}~\PY{n}{T1nf}~\PY{n}{a}~\PY{o}{\PYZhy{}\PYZgt{}}~\PY{n}{canon}~\PY{n}{a}~\PY{n}{n}\PY{o}{.+}\PY{l+m+mi}{1}~\PY{o}{\PYZlt{}=}~\PY{n}{canon}~\PY{n}{a}~\PY{n}{n}\PY{o}{.+}\PY{l+m+mi}{2}\PY{o}{.}\nl
     \end{input}

### 2022-06-22 doc/JFLA2022/Slides/Slides.tex
last changes in Slides ?

@@
          \input{../../movies/snippets/Canon/canonExamplesb}
-       \item Not the same definition as in K. \& S.'s article, but same abstract properties (monotony, limits, \dots).
+       \item Not the same definition as in K. \& S.'s article, but same abstract properties (monotony, limits, \dots). Sounds to be extentionally the same function.
        \end{itemize}
@@
 
-    \frametitle{Paths and accessibility}
+    \frametitle{Example: Paths and accessibility}
     \begin{block}{ }
@@
 
-\only<2>{{\color{lookcolor}OK for large enough $i$ (K. \& S., 1981)}}
+\only<2>{{\color{lookcolor}OK for large enough $i$ ! (K. \& S., 1981)}}
 \vspace{6pt}
@@
 
-\only<2>{\textbf{Cor:} There is no decreasing measure mapping hydras to  $[0, \mu)$ (for any $\mu<\epsilon_0$)  for proving the termination of the considered kind of hydra battles
+\only<2>{\textbf{Cor:} There is no decreasing measure mapping hydras to  $[0, \mu)$ (for any $\mu<\epsilon_0$)  for proving the termination of \textcolor{lookcolor}{\emph{the considered kind of}} hydra battles
   (stronger than \textcolor{banyuls}{Banyuls 2018}).
@@
   \frametitle{Perspectives}
+
   \begin{block}{}
+
     \begin{itemize}
 
+ \item Relation text/formalization,  new toolchain for \alectr
+ \item More illustrations of proof and specification patterns,  new exercises
+       \item New chapters (algorithmics, geometry ?)
+ \end{itemize}
+
+ \vspace{8pt}
+
+   \begin{itemize}
+ \item Develop the \texttt{Hydra-Gaia} part (more integration).
+ \item More results about hyper-complexity and fast growing functions
  \item  Links with other representations of ordinals, \emph{e.g.}:
@@
 \begin{verbatim}
-(* Minki Cho *)
+(** By Minki Cho **)
 Inductive Ord.t :=
@@
 
-      \item More results about hyper-complexity and fast growing functions
-    \item New toolchain for \alectr
-    \item More proof and specification patterns
-    \item More examples and exercises
      \end{itemize}
@@
   \begin{block}{}
-  {\color{plugincolor}\Hydras} and  {\color{plugincolor}\community} are on GitHub. {\color{cyan}Please clone, fork, contribute!}
-      \begin{itemize}
-      \item New chapters
-      \item Any correction or suggestion of improvement (text and/or proofs).
-
-    \end{itemize}
+  {\it \Hydras and  \community are on GitHub. Please clone, fork, comment, raise issues, contribute!}
 \end{block}

### 2022-06-22 doc/JFLA2022/Slides/Slides.tex
last changes in Slides ?

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+  \frametitle{Ordinals: a short reminder}
+
+  \begin{figure}[h]
+    \scalebox{0.6}{
+  \centering
+\fbox{\Large
+  \begin{minipage}{0.4\linewidth}
+  \begin{align*}
+     &\textcolor{blue}{0},\,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,\ldots\\
+&\textcolor{red}{\omega},\,\omega+1,\omega+2,\omega+3,\ldots\\
+&\textcolor{red}{\omega\times 2},\,\omega\times 2+1,\ldots, \textcolor{red}{\omega\times 3},\,\omega\times 3+1,\ldots, \textcolor{red}{\omega\times 4},\ldots\\
+&\textcolor{red}{\omega^2},\ldots, \textcolor{red}{\omega^2\times 42},\ldots,\textcolor{red}{\omega^3},\ldots, \textcolor{red}{\omega^4},\omega^4+1,\ldots\\
+&\textcolor{red}{\omega^\omega},\ldots, \textcolor{red}{\omega^\omega+\omega^7\times 8},\ldots,\textcolor{red}{\omega^\omega\times 2},\omega^\omega\times 2+1, \ldots\\
+&\textcolor{red}{\omega^{\omega^\omega}},\ldots, \textcolor{red}{\omega^{\omega^\omega}+\omega^\omega\times 42+ \omega^{55}+\omega}, \ldots, \textcolor{red}{\omega^{\omega^{\omega+1}}}, \omega^{\omega^{\omega+1}}+1,\dots\\
+     & \textcolor{red!50}{\epsilon_0 (= \omega^{\epsilon_0)}},
+       \textcolor{black!50}{\epsilon_0+1}, \textcolor{black!50}{\epsilon_0+2}, \textcolor{black!50}{\epsilon_0+3}, \ldots\\
+& \textcolor{red!50}{\epsilon_1}, \ldots, \textcolor{red!50}{\epsilon_2}, \ldots, \textcolor{red!50}{\epsilon_\omega},\ldots \\
+& \textcolor{red!50}{\Gamma_0}, \textcolor{black!50}{\Gamma_0+1}, \textcolor{black!50}{\Gamma_0+2}, \textcolor{black!50}{\Gamma_0+3},\ldots, \textcolor{red!50}{\Gamma_0+\omega}, \ldots\\
+&\ldots
+  \end{align*}   
+  \end{minipage}}}
+
+  \caption{A short overview of the sequence of ordinal numbers}
+  \label{fig:ordinal-sequence}
+\end{figure}
+\end{frame}
 

### 2022-06-24 doc/JFLA2022/Slides/Slides.tex
last changes in Slides ?

--- 2022-06-22 doc/JFLA2022/Slides/Slides.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Uncomment if there's time enough
 
-\section{Kirby \& Paris' hydras}
+% \section{Kirby \& Paris' hydras}
 
+% \begin{frame}
+%   \frametitle{Kirby \& Paris' hydras}
+%  \begin{figure}[h]
+%     \centering
+%     \begin{tikzpicture}[very thick, scale=0.20]
+%       \node (h1) at (4,0){$\bullet$};
+%       \node[blue] (h2) at (4,2){$\bullet$};
+%       \node[blue] (h3) at (2,4){$\bullet$};
+%       \node[blue] (h4) at (0,6){$\bullet$};
+%       \node[red!90!black] (h7) at (6,2){$\bullet$};
+%       \node[blue] (h5) at (4,6){$\bullet$};
+%       \node[blue] (h6) at (6,4){$\bullet$};
+%       \draw (h1) -- (h2) ;
+%       \draw (h2) -- (h3) ;
+%       \draw (h1) -- node[red!90!black,font=\small,sloped,
+%       shift={(0.01,-0.075)},rotate=90]{\textbf{\ding{34}}} (h7);
+%       \draw (h3) -- (h4) ;
+%       \draw (h3) -- (h5) ;
+%       \draw (h2) -- (h6);
+
+%       \node (h11) at (14,0){$\bullet$};
+%       \node[blue] (h12) at (14,2){$\bullet$};
+%       \node[blue] (h13) at (12,4){$\bullet$};
+%       \node[blue] (h14) at (10,6){$\bullet$};
+%       \node[blue] (h15) at (14,6){$\bullet$};
+%       \node[red!90!black] (h16) at (16,4){$\bullet$};
+%       \draw (h11) -- (h12) ;
+%       \draw (h12) -- (h13) ;
+%       \draw (h12) -- node[red!90!black,
+%       font=\small,sloped,shift={(0.01,-0.075)},
+%       rotate=90]{\textbf{\ding{34}}} (h16);
+%       \draw (h13) -- (h14) ;
+%       \draw (h13) -- (h15) ;
+
+%       \node (hn1) at (24,0){$\bullet$};
+%       \node[blue] (hn2) at (22,2) {$\bullet$};
+%       \node[blue] (hn3) at (22,4) {$\bullet$};
+%       \node[blue] (hn4) at (20,6){$\bullet$};
+%       \node[blue] (hn5) at (22,6){$\bullet$};
+%       \draw (hn1) -- (hn2) ;
+%       \draw (hn2) -- (hn3) ;
+%       \draw (hn3) -- (hn4) ;
+%       \draw (hn3) -- (hn5) ;
+%       \node [blue](hn2b) at (24,2) {$\bullet$};
+%       \node[blue] (hn3b) at (24,4) {$\bullet$};
+%       \node [blue](hn4b) at (23.4,6){$\bullet$};
+%       \node [blue](hn5b) at (24.6,6){$\bullet$};
+%       \draw (hn1) -- (hn2b) ;
+%       \draw (hn2b) -- (hn3b) ;
+%       \draw (hn3b) -- (hn4b) ;
+%       \draw (hn3b) -- (hn5b) ;
+%       \node [blue](hn2c) at (26,2) {$\bullet$};
+%       \node [blue](hn3c) at (26,4) {$\bullet$};
+%       \node [blue](hn4c) at (26,6){$\bullet$};
+%       \node [red!90!black] (hn5c) at (28,6){$\bullet$};
+%       \draw (hn1) -- (hn2c) ;
+%       \draw (hn2c) -- (hn3c) ;
+%       \draw (hn3c) -- (hn4c) ;
+%       \draw (hn3c) -- node[red!90!black,
+%       font=\small,sloped,shift={(0.01,-0.075)},
+%       rotate=90]{\textbf{\ding{34}}} (hn5c) ;
+%     \end{tikzpicture}
+
+%     \caption{Three successive states of a hydra in a battle,
+%       at time $t=1,2,3$\label{life1}}
+%  \label{fig:round}
+%  \end{figure}
+
+%   \begin{figure}[h]
+%     \centering
+%     \begin{tikzpicture}[very thick, scale=0.2]
+%       \node (hn1) at (4,0){$\bullet$};
+%       \node[blue] (hn2) at (2,2) {$\bullet$};
+%       \node[blue] (hn3) at (2,4) {$\bullet$};
+%       \node[blue] (hn4) at (0,6){$\bullet$};
+%       \node[blue] (hn5) at (2,6){$\bullet$};
+%       \draw (hn1) -- (hn2) ;
+%       \draw (hn2) -- (hn3) ;
+%       \draw (hn3) -- (hn4) ;
+%       \draw (hn3) -- (hn5) ;
+%       \node [blue](hn2b) at (4,2) {$\bullet$};
+%       \node [blue](hn3b) at (4,4) {$\bullet$};
+%       \node [blue](hn4b) at (3.4,6){$\bullet$};
+%       \node [blue] (hn5b) at (4.6,6){$\bullet$};
+%       \draw (hn1) -- (hn2b) ;
+%       \draw (hn2b) -- (hn3b) ;
+%       \draw (hn3b) -- (hn4b) ;
+%       \draw (hn3b) -- (hn5b) ;
+%       \node [blue](hn2c) at (6,2) {$\bullet$};
+%       \node [blue](hn3c) at (6,4) {$\bullet$};
+%       \node [blue](hn4c) at (6,6){$\bullet$};
+%       \draw (hn1) -- (hn2c) ;
+%       \draw (hn2c) -- (hn3c) ;
+%       \draw (hn3c) -- (hn4c) ;
+%       \node [blue](hn3d) at (8,4) {$\bullet$};
+%       \draw (hn2c) -- (hn3d) ;
+%       \node [blue](hn4d) at (8,6) {$\bullet$};
+%       \draw (hn3d) -- (hn4d) ;
+
+%       \node [blue](hn3e) at (10,4) {$\bullet$};
+%       \draw (hn2c) -- (hn3e) ;
+%       \node [blue](hn4e) at (10,6) {$\bullet$};
+%       \draw (hn3e) -- (hn4e) ;
+
+%       \node [blue](hn3f) at (12,4) {$\bullet$};
+%       \draw (hn2c) -- (hn3f) ;
+%       \node [red!90!black](hn4f) at (12,6) {$\bullet$};
+%       \draw (hn3f) -- node[red!90!black,
+%       font=\small,sloped,shift={(0.01,-0.075)},
+%       rotate=90]{\textbf{\ding{34}}} (hn4f) ;
+
+%       \node at (16,0) {$\dots$} ;
+
+%       \node [black](hn4h) at (21,0) {$\bullet$};
+%       \node [red!90!black](hn4g) at (21,2) {$\bullet$};
+%       \draw (hn4h) -- node[red!90!black,
+%       font=\small,sloped,shift={(0.01,-0.075)},
+%       rotate=90]{\textbf{\ding{34}}} (hn4g) ;
+
+%       \node [black](hn4f) at (24,0) {$\bullet$};
+
+%     \end{tikzpicture}
+%     \caption{At $t=4,\dots,\textcolor{red}{x}-1,\textcolor{red}{x=?}$
+%       \label{life2}}
+%   \end{figure}
+
+%  Cf.  \textcolor{pauillac}{Pauillac 2006} and \textcolor{banyuls}{Banyuls 2018}. Today, let us focus on some combinatorial aspects (\emph{e.g.} number of rounds of a battle).
+% \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}
-  \frametitle{Kirby \& Paris' hydras}
- \begin{figure}[h]
-    \centering
-    \begin{tikzpicture}[very thick, scale=0.20]
-      \node (h1) at (4,0){$\bullet$};
-      \node[blue] (h2) at (4,2){$\bullet$};
-      \node[blue] (h3) at (2,4){$\bullet$};
-      \node[blue] (h4) at (0,6){$\bullet$};
-      \node[red!90!black] (h7) at (6,2){$\bullet$};
-      \node[blue] (h5) at (4,6){$\bullet$};
-      \node[blue] (h6) at (6,4){$\bullet$};
-      \draw (h1) -- (h2) ;
-      \draw (h2) -- (h3) ;
-      \draw (h1) -- node[red!90!black,font=\small,sloped,
-      shift={(0.01,-0.075)},rotate=90]{\textbf{\ding{34}}} (h7);
-      \draw (h3) -- (h4) ;
-      \draw (h3) -- (h5) ;
-      \draw (h2) -- (h6);
-
-      \node (h11) at (14,0){$\bullet$};
-      \node[blue] (h12) at (14,2){$\bullet$};
-      \node[blue] (h13) at (12,4){$\bullet$};
-      \node[blue] (h14) at (10,6){$\bullet$};
-      \node[blue] (h15) at (14,6){$\bullet$};
-      \node[red!90!black] (h16) at (16,4){$\bullet$};
-      \draw (h11) -- (h12) ;
-      \draw (h12) -- (h13) ;
-      \draw (h12) -- node[red!90!black,
-      font=\small,sloped,shift={(0.01,-0.075)},
-      rotate=90]{\textbf{\ding{34}}} (h16);
-      \draw (h13) -- (h14) ;
-      \draw (h13) -- (h15) ;
-
-      \node (hn1) at (24,0){$\bullet$};
-      \node[blue] (hn2) at (22,2) {$\bullet$};
-      \node[blue] (hn3) at (22,4) {$\bullet$};
-      \node[blue] (hn4) at (20,6){$\bullet$};
-      \node[blue] (hn5) at (22,6){$\bullet$};
-      \draw (hn1) -- (hn2) ;
-      \draw (hn2) -- (hn3) ;
-      \draw (hn3) -- (hn4) ;
-      \draw (hn3) -- (hn5) ;
-      \node [blue](hn2b) at (24,2) {$\bullet$};
-      \node[blue] (hn3b) at (24,4) {$\bullet$};
-      \node [blue](hn4b) at (23.4,6){$\bullet$};
-      \node [blue](hn5b) at (24.6,6){$\bullet$};
-      \draw (hn1) -- (hn2b) ;
-      \draw (hn2b) -- (hn3b) ;
-      \draw (hn3b) -- (hn4b) ;
-      \draw (hn3b) -- (hn5b) ;
-      \node [blue](hn2c) at (26,2) {$\bullet$};
-      \node [blue](hn3c) at (26,4) {$\bullet$};
-      \node [blue](hn4c) at (26,6){$\bullet$};
-      \node [red!90!black] (hn5c) at (28,6){$\bullet$};
-      \draw (hn1) -- (hn2c) ;
-      \draw (hn2c) -- (hn3c) ;
-      \draw (hn3c) -- (hn4c) ;
-      \draw (hn3c) -- node[red!90!black,
-      font=\small,sloped,shift={(0.01,-0.075)},
-      rotate=90]{\textbf{\ding{34}}} (hn5c) ;
-    \end{tikzpicture}
-
-    \caption{Three successive states of a hydra in a battle,
-      at time $t=1,2,3$\label{life1}}
- \label{fig:round}
- \end{figure}
+  \frametitle{Ordinals: a short reminder}
 
   \begin{figure}[h]
+    \scalebox{0.6}{
   \centering
-    \begin{tikzpicture}[very thick, scale=0.2]
-      \node (hn1) at (4,0){$\bullet$};
-      \node[blue] (hn2) at (2,2) {$\bullet$};
-      \node[blue] (hn3) at (2,4) {$\bullet$};
-      \node[blue] (hn4) at (0,6){$\bullet$};
-      \node[blue] (hn5) at (2,6){$\bullet$};
-      \draw (hn1) -- (hn2) ;
-      \draw (hn2) -- (hn3) ;
-      \draw (hn3) -- (hn4) ;
-      \draw (hn3) -- (hn5) ;
-      \node [blue](hn2b) at (4,2) {$\bullet$};
-      \node [blue](hn3b) at (4,4) {$\bullet$};
-      \node [blue](hn4b) at (3.4,6){$\bullet$};
-      \node [blue] (hn5b) at (4.6,6){$\bullet$};
-      \draw (hn1) -- (hn2b) ;
-      \draw (hn2b) -- (hn3b) ;
-      \draw (hn3b) -- (hn4b) ;
-      \draw (hn3b) -- (hn5b) ;
-      \node [blue](hn2c) at (6,2) {$\bullet$};
-      \node [blue](hn3c) at (6,4) {$\bullet$};
-      \node [blue](hn4c) at (6,6){$\bullet$};
-      \draw (hn1) -- (hn2c) ;
-      \draw (hn2c) -- (hn3c) ;
-      \draw (hn3c) -- (hn4c) ;
-      \node [blue](hn3d) at (8,4) {$\bullet$};
-      \draw (hn2c) -- (hn3d) ;
-      \node [blue](hn4d) at (8,6) {$\bullet$};
-      \draw (hn3d) -- (hn4d) ;
+\fbox{\Large
+  \begin{minipage}{0.4\linewidth}
+  \begin{align*}
+     &\textcolor{blue}{0},\,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,\ldots\\
+&\textcolor{red}{\omega},\,\omega+1,\omega+2,\omega+3,\ldots\\
+&\textcolor{red}{\omega\times 2},\,\omega\times 2+1,\ldots, \textcolor{red}{\omega\times 3},\,\omega\times 3+1,\ldots, \textcolor{red}{\omega\times 4},\ldots\\
+&\textcolor{red}{\omega^2},\ldots, \textcolor{red}{\omega^2\times 42},\ldots,\textcolor{red}{\omega^3},\ldots, \textcolor{red}{\omega^4},\omega^4+1,\ldots\\
+&\textcolor{red}{\omega^\omega},\ldots, \textcolor{red}{\omega^\omega+\omega^7\times 8},\ldots,\textcolor{red}{\omega^\omega\times 2},\omega^\omega\times 2+1, \ldots\\
+&\textcolor{red}{\omega^{\omega^\omega}},\ldots, \textcolor{red}{\omega^{\omega^\omega}+\omega^\omega\times 42+ \omega^{55}+\omega}, \ldots, \textcolor{red}{\omega^{\omega^{\omega+1}}}, \omega^{\omega^{\omega+1}}+1,\dots\\
+     & \textcolor{red!55}{\epsilon_0 (= \omega^{\epsilon_0)}},
+       \textcolor{black!55}{\epsilon_0+1}, \textcolor{black!55}{\epsilon_0+2}, \textcolor{black!55}{\epsilon_0+3}, \ldots\\
+& \textcolor{red!45}{\epsilon_1}, \ldots, \textcolor{red!45}{\epsilon_2}, \ldots, \textcolor{red!45}{\epsilon_\omega},\ldots \\
+& \textcolor{red!40}{\Gamma_0}, \textcolor{black!40}{\Gamma_0+1}, \textcolor{black!40}{\Gamma_0+2}, \textcolor{black!40}{\Gamma_0+3},\ldots, \textcolor{red!40}{\Gamma_0+\omega}, \ldots\\
+&\ldots
+  \end{align*}   
+  \end{minipage}}}
 
-      \node [blue](hn3e) at (10,4) {$\bullet$};
-      \draw (hn2c) -- (hn3e) ;
-      \node [blue](hn4e) at (10,6) {$\bullet$};
-      \draw (hn3e) -- (hn4e) ;
-
-      \node [blue](hn3f) at (12,4) {$\bullet$};
-      \draw (hn2c) -- (hn3f) ;
-      \node [red!90!black](hn4f) at (12,6) {$\bullet$};
-      \draw (hn3f) -- node[red!90!black,
-      font=\small,sloped,shift={(0.01,-0.075)},
-      rotate=90]{\textbf{\ding{34}}} (hn4f) ;
-
-      \node at (16,0) {$\dots$} ;
-
-      \node [black](hn4h) at (21,0) {$\bullet$};
-      \node [red!90!black](hn4g) at (21,2) {$\bullet$};
-      \draw (hn4h) -- node[red!90!black,
-      font=\small,sloped,shift={(0.01,-0.075)},
-      rotate=90]{\textbf{\ding{34}}} (hn4g) ;
-
-      \node [black](hn4f) at (24,0) {$\bullet$};
-
-    \end{tikzpicture}
-    \caption{At $t=4,\dots,\textcolor{red}{x}-1,\textcolor{red}{x=?}$
-      \label{life2}}
+  %\caption{A short overview of the sequence of ordinal numbers}
+  \label{fig:ordinal-sequence}
 \end{figure}
-
- Cf.  \textcolor{pauillac}{Pauillac 2006} and \textcolor{banyuls}{Banyuls 2018}. Today, let us focus on some combinatorial aspects (\emph{e.g.} number of rounds of a battle).
 \end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
   \begin{block}{}
-       \begin{itemize}
-       \item In Coq, canonical sequences are defined by structural recursion on ordinal terms in Cantor normal form.
 
-   {\color{mathcolor}
-         $$\canonseq{\omega^{\omega^{\omega+1}+1}}{42}=
-         \omega^{\omega^{\omega+1}}\times 42$$
+    Canonical sequences are implemented by a structurally recursive function on terms in Cantor normal form.
+
+\vspace{5pt}
+
+    Example: $$\textcolor{mathcolor}{\canonseq{\omega^{\omega^{\omega+1}+1}}{42}}$$
+
+      {\footnotesize
+    \input{../../movies/snippets/Canon/canonExamplesb}
   }
 
-         \input{../../movies/snippets/Canon/canonExamplesb}
-       \item Not the same definition as in K. \& S.'s article, but same abstract properties (monotony, limits, \dots). Sounds to be extentionally the same function.
-       \end{itemize}
+   %   
+    For efficiency's sake, our definition in \coq is sligthly different from Ketonen and Solovay's article (based on the decomposition on any ordinal \textcolor{mathcolor}{$\alpha$} into  \textcolor{mathcolor}{$\beta(\gamma+1)$}).
+
+We \emph{believe} that both definitions are extentionally equivalent.
+
+     %  \end{itemize}
      \end{block}
@@
           \begin{tikzpicture}[very thick, scale=0.25]
+\node (epsilon0) at (-3,0) {\textcolor{black!60}{$(\epsilon_0>\;)$}};
 \node (alpha) at (0,0) {$\alpha$};
@@
 
-\only<2>{{\color{lookcolor}OK for large enough $i$ ! (K. \& S., 1981)}}
+\only<2>{{\color{lookcolor}If $i$ is large enough, we are sure to reach $\beta$ (K. \& S., 1981)}}
 \vspace{6pt}
@@
 % \pause
-\only<1>{Goal: reach \textcolor{red}{$\beta$} from $\alpha$.}
-\only<2>{\textbf{Remark:} The proof is not trivial. The deterministic red path could miss $\beta$!}
+\only<1>{Can we  reach \textcolor{red}{$\beta$} from $\alpha$? Under which condition? }
 
@@
 
-  {\small
-\textbf{Hardy Hierarchy (variant)}
+  {
 
@@
   H'_\alpha(i) &= H'_{(\canonseq{\alpha}{i+1})}(i)  \quad\textit{($\alpha$ limit ordinal)} 
-\end{align}}
+\end{align}}}
 
-\vspace{6pt}
+{$H'$ is a slight variant of the \emph{Hardy Hierarchy}. In \coq, we implement it with the \texttt{equations} plug-in.}
+%\vspace{6pt}
 
-\textbf{Wainer hierarchy (variant)}
+% \textbf{Wainer hierarchy (variant)}
 
-{\color{mathcolor}
-\begin{align}
-F_0(i)& =i+1\\
-F_{\beta+1}(i)&= (F_\beta)^{(i+1)}(i)\\
-F_\alpha(i) &= F_{(\canonseq{\alpha}{i})} (i) \quad \textit{($\alpha$ limit ordinal)}
-\end{align}}}
+% {\color{mathcolor}
+% \begin{align}
+% F_0(i)& =i+1\\
+% F_{\beta+1}(i)&= (F_\beta)^{(i+1)}(i)\\
+% F_\alpha(i) &= F_{(\canonseq{\alpha}{i})} (i) \quad \textit{($\alpha$ limit ordinal)}
+% \end{align}}}
 \end{block}
@@
       H'_{\omega^3}(k) \geq\, & \textit{hyper\_exp2}(k+1) \\
-      H'_{\omega^\alpha}(k)\geq\,& F_\alpha(k)\; (k>0)\\
+ %     H'_{\omega^\alpha}(k)\geq\,& F_\alpha(k)\; (k>0)\\
     \end{align*}
@@
 \label{thms-fastgrowing}
-  \begin{block}{Theorems}
-    \begin{itemize}
-    \item \textcolor{mathcolor}{$F_\alpha$} is primitive recursive iff \textcolor{mathcolor}{$\alpha<\omega$}.
-     \item  
+  \begin{block}{Theorem}
+    % \begin{itemize}
+    % \item \textcolor{mathcolor}{$F_\alpha$} is primitive recursive iff \textcolor{mathcolor}{$\alpha<\omega$}.
+    %  \item  
     The function which computes the length of hydra battles
@@
     (for $\textcolor{mathcolor}{\alpha\geq \omega^\omega}$).
-    \end{itemize}
+  %  \end{itemize}
 
@@
 
-\begin{frame}
-  \frametitle{}
-  \begin{scriptsize}
-    \inputsnippets{./SchemePrimRecInda}
-  \end{scriptsize}
-\end{frame}
+% \begin{frame}
+%   \frametitle{}
+%   \begin{scriptsize}
+%     \inputsnippets{./SchemePrimRecInda}
+%   \end{scriptsize}
+% \end{frame}
 

### 2022-06-25 doc/JFLA2022/Slides/Slides.tex
last changes in Slides ?

@@
 \begin{scope}[yshift=-5cm,xshift=4cm]
-  \node[rectangle, draw=black,fill=red!60,minimum height=0.5cm,minimum width=1.5cm] (book) {hydras.pdf};
-    \node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm, left of=book, node distance=6cm] (latex) {hydras.tex};
+  \node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm] (book) {hydras.pdf};
+    \node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm, left of=book, node distance=6cm] (latex) {hydras.tex};
 \end{scope}
@@
   \frametitle{The Ketonen-Solovay machinery\footnote{Rapidly Growing {R}amsey Functions (1981)}}
-    \begin{block}{Formalization}
+    \begin{block}{Paths associated with hydra battles}
       {\color{mathcolor}
@@
 
-      \begin{block}{Small steps}
+      \begin{block}{}
         At the $i$-th round, 
@@
 
-    Canonical sequences are implemented by a structurally recursive function on terms in Cantor normal form.
+    Canonical sequences are implemented by a structurally recursive function on ordinal terms in Cantor normal form.
 
@@
 
-       \item     If \textcolor{mathcolor}{$\alpha \rplus{n,n,\dots,n} \beta$} then \textcolor{mathcolor}{$H'_\beta(n)\leq H'_\alpha(n)$}
+       \item     If \textcolor{mathcolor}{$\alpha \rplus{i,i,\dots,i} \beta$} then \textcolor{mathcolor}{$H'_\beta(i)\leq H'_\alpha(i)$}.
 

### 2022-06-25 doc/JFLA2022/Slides/Slides.tex
last changes in Slides ?

@@
 
-\begin{block}{}
-
-  {
-
- \textbf{Definition:} \textcolor{mathcolor}{$H'_\alpha(i)$} is the least $l$ such that \textcolor{mathcolor}{$\alpha_l=0$} (cf. previous slide).
+\begin{block}{Definition:}
+  \textcolor{mathcolor}{$H'_\alpha(i)$} is the least $l$ such that \textcolor{mathcolor}{$\alpha_l=0$} (cf. previous slide).
 
@@
   H'_\alpha(i) &= H'_{(\canonseq{\alpha}{i+1})}(i)  \quad\textit{($\alpha$ limit ordinal)} 
-\end{align}}}
+\end{align}}
 
-{$H'$ is a slight variant of the \emph{Hardy Hierarchy}. In \coq, we implement it with the \texttt{equations} plug-in.}
+$H'$ is a slight variant of the \emph{Hardy Hierarchy}. In \coq, we implement it with the \texttt{equations} plug-in.
+\end{block}
+
+\begin{block}{Example:}
+  {\color{mathcolor}
+  $$      H'_{\omega^\omega}(i)=\, \textbf{let}\,F\;f\;k := f^{(k+1)}(k)\,
+                                  \textbf{in}\, F^{(i)}(\lambda k.\,2k+1)\,i$$
+}
+\end{block}
+
 %\vspace{6pt}
@@
 % \end{align}}}
-\end{block}
+
 \end{frame}
@@
 
-       \item     If \textcolor{mathcolor}{$\alpha \rplus{i,i,\dots,i} \beta$} then \textcolor{mathcolor}{$H'_\beta(i)\leq H'_\alpha(i)$}.
+       \item     If \textcolor{mathcolor}{$\alpha \rplus{i,i,\dots,i} \beta$} then \textcolor{mathcolor}{$H'_\alpha(i)\geq H'_\beta(i)$}.
 
@@
 
- \begin{block}{Examples}
+ \begin{block}{Examples:}
 
@@
         \begin{align*}
-        H'_{\omega^\omega}(k)=\,& \textbf{let}\,F\;f\;i := f^{(i+1)}(i)\,
-                                  \textbf{in}\, F^{(k)}(\lambda k.\,2k+1)\,k\\
-      H'_{\omega^3}(k) \geq\, & \textit{hyper\_exp2}(k+1) \\
- %     H'_{\omega^\alpha}(k)\geq\,& F_\alpha(k)\; (k>0)\\
+      H'_{\omega^3}(i) \geq\, & \textit{hyper\_exp2}(i+1) \\
+      H'_{\omega^\omega}(i)\geq\,& \textit{Ack}(i,i)\; (i\geq 2)\\
     \end{align*}
   }%mathcolor
+ Thus, for any \textcolor{mathcolor}{$\alpha\geq\omega^\omega$}, \textcolor{mathcolor}{$H'_\alpha$} is not primitive recursive.
   \end{block}
@@
 
-\begin{frame}[label={fgftheorems}]
-\label{thms-fastgrowing}
-  \begin{block}{Theorem}
-    % \begin{itemize}
-    % \item \textcolor{mathcolor}{$F_\alpha$} is primitive recursive iff \textcolor{mathcolor}{$\alpha<\omega$}.
-    %  \item  
-    The function which computes the length of hydra battles
-    for \textcolor{mathcolor}{$\alpha$} (according to the initial index) is not primitive recursive
-    (for $\textcolor{mathcolor}{\alpha\geq \omega^\omega}$).
-  %  \end{itemize}
+% \begin{frame}[label={fgftheorems}]
+% \label{thms-fastgrowing}
+%   \begin{block}{Theorem}
+%     % \begin{itemize}
+%     % \item \textcolor{mathcolor}{$F_\alpha$} is primitive recursive iff \textcolor{mathcolor}{$\alpha<\omega$}.
+%     %  \item  
+%     The function which computes the length of hydra battles
+%     for \textcolor{mathcolor}{$\alpha$} (according to the initial index) is not primitive recursive
+%     (for $\textcolor{mathcolor}{\alpha\geq \omega^\omega}$).
+%   %  \end{itemize}
 
-  \begin{block}{Sketch of proof:}
+%   \begin{block}{Sketch of proof:}
 
-   We apply monotony properties and the famous lemma:
-    \vspace{5pt}
+%    We apply monotony properties and the famous lemma:
+%     \vspace{5pt}
 
-       \begin{quote}
-      {\color{mathcolor}
-        For any $n$-ary primitive recursive  function $f$, 
-        there exists some $q$ such that the following inequality holds:
-         \[\forall\,x_1,\dots,x_n,  
-          f(x_1,\dots,\,x_n)\leq\textrm{Ack}(q,\textrm{max}(x_1,\dots,x_n))\]}
-    \end{quote}
-  %  (Proof by induction on any primitive recursive function of any arity.)
-  \end{block}
+%        \begin{quote}
+%       {\color{mathcolor}
+%         For any $n$-ary primitive recursive  function $f$, 
+%         there exists some $q$ such that the following inequality holds:
+%          \[\forall\,x_1,\dots,x_n,  
+%           f(x_1,\dots,\,x_n)\leq\textrm{Ack}(q,\textrm{max}(x_1,\dots,x_n))\]}
+%     \end{quote}
+%   %  (Proof by induction on any primitive recursive function of any arity.)
+%   \end{block}
 
-  \end{block}
-\end{frame}
+%   \end{block}
+% \end{frame}
 

### 2022-06-26 doc/JFLA2022/Slides/Slides.tex
last changes in Slides ?

@@
 
-\begin{block}{Definition:}
+\textbf{Definition: }
   \textcolor{mathcolor}{$H'_\alpha(i)$} is the least $l$ such that \textcolor{mathcolor}{$\alpha_l=0$} (cf. previous slide).
@@
 
-$H'$ is a slight variant of the \emph{Hardy Hierarchy}. In \coq, we implement it with the \texttt{equations} plug-in.
-\end{block}
+\textcolor{mathcolor}{$H'$} is a slight variant of the \emph{Hardy Hierarchy} (defined by transfinite recursion with \texttt{equations}).
+ We can prove exact, but not always very readable, formulas for small values of \textcolor{mathcolor}{$\alpha$} (\emph{e.g.} \textcolor{mathcolor}{$\omega^\omega$}).
 
-\begin{block}{Example:}
-  {\color{mathcolor}
-  $$      H'_{\omega^\omega}(i)=\, \textbf{let}\,F\;f\;k := f^{(k+1)}(k)\,
-                                  \textbf{in}\, F^{(i)}(\lambda k.\,2k+1)\,i$$
-}
-\end{block}
+\inputsnippets{./HprimePhi0OmegaClosed}
 

### 2022-07-02 doc/hydras.tex
add author address

@@
 
-\author{Pierre Castéran\\ Univ.
-Bordeaux, CNRS, Bordeaux INP, LaBRI, UMR 5800, F-33400 Talence, France \thanks{With contributions by Yves Bertot, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Karl Palmskog, Clément Pit-Claudel, and Théo Zimmermann.
+% \author{Pierre Castéran\\ Univ.
+Bordeaux, CNRS, Bordeaux INP, LaBRI, UMR 5800, F-33400 Talence, France\footnote{ email: pierre casteran at gmail dot com} \thanks{With contributions by Yves Bertot, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Karl Palmskog, Clément Pit-Claudel, and Théo Zimmermann.
 The formalization of primitive recursive functions was originally authored by
-Russel O'Connor\cite{OConnor05}.}}
-\date{\today}
-%\date{August 2021}
+% Russel O'Connor\cite{OConnor05}.}}
+% \date{\today}
 
@@
        \emph{Pierre Castéran, LaBRI, Univ.
-Bordeaux, CNRS UMR 5800, \\ With contributions by Yves Bertot, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Karl Palmskog, Clément Pit-Claudel, and Théo Zimmermann.
+Bordeaux, CNRS UMR 5800\\ \textbf{email:} pierre dot casteran arobas gmail dot com.
+\\ With contributions by Yves Bertot, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Karl Palmskog, Clément Pit-Claudel, and Théo Zimmermann.
 \\ The formalization of primitive recursive functions was originally authored by Russel O'Connor.}%\cite{OConnor05}.}

### 2022-07-02 doc/JFLA2022/Slides/HprimePhi0OmegaClosed.tex

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Lemma}~\PY{n+nf}{H\PYZsq{}\PYZus{}Phi0\PYZus{}omega\PYZus{}exact\PYZus{}formula}~\PY{n+nv}{k}~\PY{o}{:}\nl
      ~~\PY{n}{H\PYZsq{}\PYZus{}}~\PY{o}{(}\PY{n}{E0phi0}~\PY{n}{E0omega}\PY{o}{)}~\PY{n}{k}~\PY{o}{=}\nl
      ~~~~\PY{k+kr}{let}~\PY{n+nv}{F}~\PY{n+nv}{f}~\PY{n+nv}{i}~\PY{o}{:=}~\PY{n}{iterate}~\PY{n}{f}~\PY{o}{(}\PY{n}{S}~\PY{n}{i}\PY{o}{)}~\PY{n}{i}\nl
      ~~~~\PY{k+kr}{in}~\PY{k+kr}{let}~\PY{n+nv}{g}~\PY{n+nv}{k}~\PY{o}{:=}~\PY{n}{S}~\PY{o}{(}\PY{l+m+mi}{2}~\PY{o}{*}~\PY{n}{k}\PY{o}{)\PYZpc{}}\PY{n}{nat}\nl
      ~~~~~~~\PY{k+kr}{in}~\PY{n}{iterate}~\PY{n}{F}~\PY{n}{k}~\PY{n}{g}~\PY{n}{k}\PY{o}{.}~
    \end{input}
  \end{sentence}
\end{alectryon}

### 2022-07-27 doc/ordinal-chapter.tex
Remove the example of termination of list merge (can be solved with omega)
More compatibility 8.15 8.16 + new example on Ackermann function

@@
 
-\subsection{A proof of termination using \texorpdfstring{$\omega^2$}{omega^2}} 
-\label{sect:merge-example}
-
-Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily a total function which merges two lists (example contributed by Pascal Manoury).
-
-%\index{coq}{Commands!Function}
-
-\vspace{4pt}
-\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
-
-\inputsnippets{ON_Omega2/Merge, ON_Omega2/Mergeb}
+%%% removed (bad example for omega*omega )
+
+% \subsection{A proof of termination using \texorpdfstring{$\omega^2$}{omega^2}} 
+% \label{sect:merge-example}
+
+% Using the lemma of Sect.~\vref{sect:measure-ON}, we can define easily a total function which merges two lists (example contributed by Pascal Manoury).
+
+% %\index{coq}{Commands!Function}
+
+% \vspace{4pt}
+% \noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}}
+
+% \inputsnippets{ON_Omega2/Merge, ON_Omega2/Mergeb}
+
+\subsubsection{Example: Ackermann function}
+
+\index{maths}{Ackermann function}
+\index{coq}{Plug-ins!Equations}
+
+The famous Ackermann function can be defined with the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777}, with a measure towards $\omega^2$.
+
+\vspace{6pt}
+\noindent
+From~ \href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{OrdinalNotations.ON\_Omega2}.
+
+\inputsnippets{ON_Omega2/AckOmega2}
 

### 2022-07-28 doc/chapter-primrec.tex
small changes
JFLA Slides (draft)
revision book (-> p. 28)
work on Ackermann (proof scripts + book)

@@
 
-  Getting rid of the term \texttt{x-x}, we generate two easy-to-solve sub-goals.
+  Getting rid of the term \texttt{x-x}, we generate two easy-to-solve subgoals.
 
@@
 \index{primrec}{Predicates!isPR}
-\index{coq}{Extensionally equal functions}
+%\index{coq}{Extensionally equal functions}
 
@@
 \begin{remark}
-In the library \texttt{primRec}, all these lemmas are opaque (registered with \texttt{Qed}.
+In the library \texttt{primRec}, all these lemmas are opaque (registered with \texttt{Qed}).
 Thus they do not allow the user to look at the witness of a proof of a \texttt{isPR} statement.
@@
 \texttt{PrimRec} allows us to write trivial proofs of primitive recursivity. 
-Although  the following lemmas are already proven in 
-\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec},
-we wrote alternate proofs in 
+Some of   the following lemmas are already proven in 
+\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}. More examples are given in 
 \href{../theories/html/hydras.MoreAck.PrimRecExamples.html}%
@@
 
-\input{movies/snippets/PrimRecExamples/SuccIsPR}
+\input{movies/snippets/primRec/SuccIsPR}
 
@@
 
-\input{movies/snippets/primRec/SchemePrimRecInd}
+\inputsnippets{primRec/SchemePrimRecInd,
+primRec/SchemePrimRecInda}
 

### 2022-09-01 doc/hydras.tex
On gaia (#131)
* minor change

* experiments on a type of ordinal notation (draft)

* try to fix an issue (coq 8.14)

* try to fix an issue (coq 8.13)

* restrict to MathComp 1.13.0 and later in opam and ci

* small simplification

* slightly improve mathComp code

Co-authored-by: Karl Palmskog <palmskog@gmail.com>
new part in the book

@@
 \end{quote}
+\vspace{6pt}
 \mbox{}
+\label{watercolorgray}
+ {\centering
+      \includegraphics[width=12cm]{hydre-pierrette.jpg}
+      \centerline{\emph{``The hydra in black and white'': Watercolor  by Pierrette Cassou-Noguès}}
+}
 \newpage
@@
 \include{Gamma0-chapter}
+
+\part{Ackermann, G\"{o}del, Peano\,\dots}
+This part contains comments, example  and exercises about Russel O'Connor's work on G\"{o}del's first incompleteness theorem.
+This work is maintained by Coq community~\cite{CoqCommunity} volunteers: (projects Goedel~\cite{Goedel} and Hydra-battles~\cite{HydraBattles}).
+
+Some changes are made to the aforementionned libraries, mainly because of the recent evolution of coq and its libraries.
+Nevertheless, the definitions, lemmas and theorems of the original contribution have been kept in this new release.
+
+\emph{\color{red!80}This maintenance job has just started, ant will probably be long to complete. Help will be welcome!}
+
 \include{chapter-primrec}

### 2022-09-04 doc/chapter-encoding.tex
sart a new chapter on Goedel encoding (after Russel)

\chapter{G\"{o}del's Encoding (in construction)}
\label{chap:encoding}

### 2022-09-05 doc/chapter-encoding.tex
start chapter on goedel encodings

@@
 \label{chap:encoding}
+
+\section{Cantor pairing function}
+
+The library \href{../theories/html/hydras.Ackermann.cPair.html}{Ackermann.cPair} defines and study Cantor's bijection from
+$\mathbb{N}\times\mathbb{N}$ into $\mathbb{N}$.
+Indeed the \texttt{cPair} function used in this library
+ is slightly different 
+     from the "usual" Cantor pairing function shown in  a big part 
+     of the litterature (and Coq's standard library). 
+      Since both versions are equivalent upto a swap of the 
+      rguments [a] and [b], we still use  Russel O'Connors definitions and statements, mainly in order to not have to modify the order of sub-goals in long proofs.
+
+      \subsection{A helper function}
+
+      The following function computes the sum of all natural numbers between $1$ and $n$: $\Sigma_{i=1}^{i=n}\,i$.
+
+      \inputsnippets{cPair/sumToNDef}
+
+      \inputsnippets{cPair/sumToN1, cPair/sumToN1}
+
+      The tools presented in Chapter~\ref{chapter:primrec} allow us to prove that \texttt{cPair} is primitive recursive.
+
+      \inputsnippets{cPair/sumToNPR}
+
+      \subsection{Cantor's pairing function}
+
+      In the Ackermann/G\"{o}del projects, the Cantor pairing function is defined as below:
+
+      $$\textrm{cPair}\,a\,b = a+ \Sigma_{i=1}^{i=a+b}\,i$$
+
+      \inputsnippets{cPair/cPairDef}
+
+      Figure~\ref{fig:cpair} shows a few values of
+      \texttt{cPair\,$a$\,$b$ }, where $a$ is the line number and $b$ the column number.
+
+      \begin{figure}[h]
+        \centering
+
+         \[
+        \begin{array}[h]{c|cccccc}
+          &0&1&2&3&4&\dots \\
+          \hline 
+          0&0&1&3& 6 & 10&\dots \\
+          1&2&4&7&11& \dots& \\
+          2&5&8 & 12& \dots&& \\
+            3&9&13&\dots & && \\
+          4&14&\dots&&&&\\
+          \dots&\dots&&&&&\\
+        \end{array}
+        \]
+
+        \caption{Cantor pairing function (first values)\label{fig:cpair}}
+      \end{figure}
+
+      \begin{remark}
+        Compatibility with Standard lib's pairing function is stated
+        by the equality \texttt{cPair $a$ $b$ = Cantor.to\_nat $(b,a)$} for any $a$ and $b$.
+Obviously, this swap and uncurrying of $a$ and $b$ doesn't change the fundamental properties of Cantor's pairing function.
+      \end{remark}

### 2022-09-05 doc/chapter-encoding.tex
start chapter on goedel encodings

@@
      from the "usual" Cantor pairing function shown in  a big part 
-     of the litterature (and Coq's standard library). 
+     of the litterature , and Coq's standard library \footnote{In 
+\url{https://coq.inria.fr/distrib/current/stdlib/Coq.Arith.Cantor.html}}.
        Since both versions are equivalent upto a swap of the 
@@
       \begin{remark}
-        Compatibility with Standard lib's pairing function is stated
+        Compatibility with Standard lib's pairing function
+        is stated
         by the equality \texttt{cPair $a$ $b$ = Cantor.to\_nat $(b,a)$} for any $a$ and $b$.
-Obviously, this swap and uncurrying of $a$ and $b$ doesn't change the fundamental properties of Cantor's pairing function.
+Obviously, this swap and uncurrying of $a$ and $b$ doesn't change the fundamental properties of Cantor's pairing function (being a primitive recursive bijection, monotony properties).
       \end{remark}

### 2022-09-06 doc/chapter-encoding.tex
start chapter on goedel encodings

@@
       \end{remark}
+
+      \subsubsection{Main properties}
+
+      In order to prove that the function \texttt{cPair} is primitive recursive, we express it as a composition of already proven primitive recursive functions.
+
+      \inputsnippets{cPair/cPairIsPR}
+
+      \texttt{cPair}'s injectivity is stated by two lemmas:
+
+      \inputsnippets{cPair/cPairInj1, cPair/cPairInj2}
+
+      \subsection{Projections}
+
+      Let $a$ be some natural number, we look for two natural numbers $x$ and $y$ such that $\texttt{cPair}\,x\,y=a$.
+More we want to prove that the function which compute $x$ [resp.
+$y$] out of $a$ are primitive recursive.
+
+      Let us show on a small example how these projections are defined. Let's take for instance $a=11$. Please look at the diagram of Fig.~\ref{fig:cpair}.
+
+      The following function looks for the anti-diagonal $\{(x,y)|x+y=k\}$ the number $11$ belongs to, and returns $k$.
+
+      \inputsnippets{cPair/searchXY}
+
+      In our example, $11$ belongs to the anti-diagonal $\{(x,y)|x+y=4\}$, which contains values from $10$ to $14$.
+      Thus the line containing $11$ is the line $x=11-10=1$, and the column is $y=4-1= 3$. Finally, we get  $11=\texttt{cPair}\,1\,3$.
+
+       \inputsnippets{cPair/cPairPI12}
+
+       The following lemmas (still from \href{../theories/html/hydras.Ackermann.cPair.html}{Ackermann.cPair}) prove the correctness of these projections.
+
+       \inputsnippets{cPair/cPairProjections,
+         cPair/cPairProjections1, cPair/cPairProjections2,
+         cPair/cPairPi1IsPR, cPair/cPairPi2IsPR}
+
+       Finally, let us show a few inequalities.
+
+       \inputsnippets{cPair/cPairLe1, cPair/cPairLe1A,
+         cPair/cPairLe2, cPair/cPairLe2A, cPair/cPairLe3,
+         cPair/cPairLt1,cPair/cPairLt2}

### 2022-09-07 doc/chapter-encoding.tex
work on the  chapter on goedel encodings

@@
          cPair/cPairLt1,cPair/cPairLt2}
+
+       \subsection{List encoding}
+
+       The encoding of a list of natural numbers is based
+       on \texttt{cPair}, through a structural recursion.
+
+       \inputsnippets{cPair/codeListDef}
+
+       Let us look at the main step of the proof that \texttt{codeList} is injective.
+
+  \inputsnippets{cPair/codeListInj}
+  \inputsnippets{cPair/codeListInja}
+
+  Applying \texttt{cPair}'s injectivity  and the induction hypothesis allows us to complete the proof.
+
+  \subsubsection{Encoding the \texttt{nth} function}
+
+  The following function allows us to compute the $n$-th element of a list of natural numbers, \emph{directly on the encoding of the considered list}.
+
+  In the current version of the Ackermann Library, this function is defined \emph{via} an interactive proof. Its specification and
+  correctness are proved by  separate lemmas.
+
+  \inputsnippets{cPair/codeNthDef}
+  \inputsnippets{cPair/codeNthCorrect}
+   \inputsnippets{cPair/codeNthIsPR}
+
+  \begin{exercise}
+  Give a new equivalent definition of \texttt{codeNth}, without using tactics like \texttt{assert}.
+You may define an help function for this purpose, in which case, please specify what your helper computes.
+  \end{exercise}

### 2022-09-09 doc/chapter-encoding.tex
work on the chapter on goedel encodings; new exercise

@@
 
+   \index{hydras}{Exercises}
   \begin{exercise}
@@
   \end{exercise}
+
+     \index{hydras}{Exercises}
+  \begin{exercise}
+    Please consider the following definition:
+
+    \inputsnippets{OnCodeList/membersDef}
+
+    \noindent
+
+    Prove that the functions \texttt{codeList} and \texttt{members} are inverse of each other.
+  \end{exercise}
+
+  \subsection{Strong recursion}
+
+  List encoding helps us to define primitive recursive functions
+  where the computation of $f\;n$ may  depend of
+  [part of] the values $f\;0$, $f\;1$, \dots, $f\;(n-1)$.
+
+  A simple example is the \texttt{div2} function
+  defined by
+  \begin{align*}
+    \texttt{div2}\;0&=0\\
+    \texttt{div2}\;1&=1\\
+    \texttt{div2}\;(n+2)&=S(\texttt{div2}\;n)\\
+  \end{align*}
+
+  The trick is to define a helper $h$ where
+  $h\;n\;a$ is the natural number which represents
+  the sequence $\langle \texttt{div2}(n-1),
+  \texttt{div2}(n-2),\dots,\texttt{div2}\;1, \texttt{div2}\;0\rangle$.
+
+  In \coq, the helper associated with \texttt{div2} is defined in
+  \href{../theories/html/hydras.MoreAck/PrimRecExamples}{MoreAck.PrimRecExamples}.
+
+  \inputsnippets{PrimRecExamples/fdiv2Def}
+
+  The function \texttt{(evalStrongRec\;$n$\;$h$\;$c$)} computes
+  $(f\;c)$ if $h$ is the helper associated with $f$.
+
+  \inputsnippets{PrimRecExamples/fdiv2Examples}
+
+  This function is defined in
+  \href{../theories/html/hydras.Ackermann.cPair.html}{Ackermann.cPair}.
+   Definitions and related lemmas are quite tricky, but look easy to apply.
+
+   \inputsnippets{cPair/evalStrongRecDef}
+
+   \inputsnippets{cPair/evalStrongRecPR}
+    \inputsnippets{cPair/computeEvalStrongRecHelp}

### 2022-09-30 doc/chapter-encoding.tex
small changes

@@
   The trick is to define a helper $h$ where
-  $h\;n\;a$ is the natural number which represents
+  $h\;n\;a$ is the natural number which encodes
   the sequence $\langle \texttt{div2}(n-1),
@@
   $(f\;c)$ if $h$ is the helper associated with $f$.
+  This function is defined in
+  \href{../theories/html/hydras.Ackermann.cPair.html}{Ackermann.cPair}.
 
@@
 
-  This function is defined in
-  \href{../theories/html/hydras.Ackermann.cPair.html}{Ackermann.cPair}.
+  Trying to compute the half of $5$ this way resulted in unbearably long computation times \dots.
+
    Definitions and related lemmas are quite tricky, but look easy to apply.
@@
     \inputsnippets{cPair/computeEvalStrongRecHelp}
+
+  \index{ordinals}{Exercises}
+    \begin{exercise}
+      Prove formally that this implementation of \texttt{div2} is correct.
+    \end{exercise}
+
+   \index{ordinals}{Exercises}
+    \begin{exercise}
+     Define a function for computing the Fibonacci numbers by strong recursion.
+    \end{exercise}

### 2022-09-30 doc/hydras.tex
minor changes
sart a new chapter on Goedel encoding (after Russel)
update acknowledgments

@@
 Bordeaux, CNRS UMR 5800\\ \textbf{email:} pierre dot casteran arobas gmail dot com.
-\\ With contributions by Yves Bertot, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Karl Palmskog, Clément Pit-Claudel, and Théo Zimmermann.
+\\ With contributions by Yves Bertot, Ilm\={a}rs C\={i}rulis, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Karl Palmskog, Clément Pit-Claudel, and Théo Zimmermann.
 \\ The formalization of primitive recursive functions was originally authored by Russel O'Connor.}%\cite{OConnor05}.}
@@
 \label{sec:orgheadline5}
-    Many thanks to Yves Bertot, \'Evelyne Contejean, Jéremy Damour,   Florian Hatat,  David Ilcinkas, 
+    Many thanks to Yves Bertot, Ilm\={a}rs C\={i}rulis, \'Evelyne Contejean, Jéremy Damour,   Florian Hatat,  David Ilcinkas, 
 Pascal Manoury,  Karl Palmskog, Cl\'ement Pit-Claudel, Sylvain Salvati, Alan Schmitt and Théo Zimmermann for their help on the elaboration of this library and  document, and to the
@@
 
-\emph{\color{red!80}This maintenance job has just started, ant will probably be long to complete. Help will be welcome!}
+\emph{\color{red!80}This maintenance job has just started, and will probably be long to complete. Help is welcome!}
 
 \include{chapter-primrec}
+\include{chapter-encoding}
 

### 2022-10-07 doc/chapter-fol.tex
start a chapter on FOL

\chapter{First Order Logic (in construction)}
\label{chap:fol}

\section{Data types}

\subsection{Languages}

A \emph{language} is a structure composed of relation and function symbols, each symbol is given an \emph{arity} (number of arguments).

From~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol}
\inputsnippets{fol/LanguageDef}

In ~\href{../theories/html/hydras.Ackermann.Languages.html}{Ackermann.Languages}, we can find two examples of languages built with the usual symbols of arithmetic.

The first language, \texttt{LNT} (\emph{Language of Number Theory}) has just function symbols for $+$, $\times$, $0$ and successor.

The second language, LNN, is the language of NN (\emph{Natural Numbers}) and has
the same function symbols as LNT plus one relation symbol for less than, LT.

First, we declare two ranked alphabets. 
\inputsnippets{Languages/LNTDef1}

In a second time, we build \texttt{LNT} and \texttt{LNN} by filling \texttt{Language}'s \texttt{arity} field.

\inputsnippets{Languages/LNTDef2}

Let us show a few examples (from ~\href{../theories/html/hydras.MoreAck.fol_Examples.html}{MoreAck.fol\_Examples}).

\inputsnippets{fol_Examples/arityTest}

\subsection{Terms}

Given a language $L$, we define the set of \emph{terms} and
$n$-\emph{tuples} of terms.

\inputsnippets{fol/TermDef}

For instance the term $v_1+0$, where $v_1$ is a variable,
is represented by a term of type (\texttt{Term LNN}).

\inputsnippets{fol_Examples/v1Plus0}

### 2022-10-11 doc/chapter-fol.tex
start a chapter on FOL

@@
 \inputsnippets{fol_Examples/v1Plus0}
+
+\subsection{Formulas}
+
+The type of first order formulas over $L$ is defined 
+in~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} as an inductive data type.
+
+\inputsnippets{fol/FormulaDef}
+
+The full abstract syntax of first-order formulas is completed 
+with a few more notations. Note that we are studying \emph{classical} first-order logic, so the existential quantifier is derived from the universal one through negation.
+
+\inputsnippets{fol/FolFull, fol/folPlus}
+
+\subsection{Examples}
+
+In order to get more readable terms and formulas, we can define a few notations in ~\href{../theories/html/hydras.MoreAck.fol_Examples.html}{MoreAck.fol\_Examples}
+
+\inputsnippets{fol_Examples/instantiations}
+
+For instance, the term $v_1+0$ is written as below.
+
+\inputsnippets{fol_Examples/v1Plus01}
+
+Let us give two examples of first-order formulas.
+
+\begin{itemize}
+\item $\forall\,v_0\;v_1,\, v_0<v_1\, \longrightarrow\, v_0<v_1 + v_0$
+\inputsnippets{fol_Examples/f1Example}
+\item $\forall\,v_0,\,\exists\,v_1,\,v_0<v_1$
+\inputsnippets{fol_Examples/f2Example}
+
+\end{itemize}
+
+\subsection{Decidability of equality}
+
+Under the assumption that equality is decidable on 
+function and relation symbols of any language $L$,
+we get decidability of equality on terms and formulas.
+Because of dependent types, the proofs are quite long and technical. The reader may consult them in \href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol}
+
+\begin{todo}
+  Should we talk about the decomposition lemmas
+\texttt{nilTerms} and \texttt{consTerms} ?
+\end{todo}
+
+\inputsnippets{fol/formDec1, fol/formDec2,
+fol/formDec3,fol/formDec4,fol/formDec5}
+
+\subsection{Induction principles on formula depth}
+
+\subsubsection{Depth of a formula}
+
+\inputsnippets{fol/depthDef}
+
+\inputsnippets{fol_Examples/ltdepth1}
+
+\begin{todo}
+Motivate the induction principles based on depth. Compatible with term substitution and universal quantifier elimination.
+\end{todo}
+
+\begin{todo}
+ Look for the principles which are really used in Ackermann or/and Goedel libraries, and comment them.
+ Maybe skip the helpers (unused in other files)
+\end{todo}

### 2022-10-11 doc/hydras.tex
start a chapter on FOL

@@
       \centerline{\emph{``The blue hydra'': Watercolor  by Pierrette Cassou-Noguès}}
+}
+
+\newpage
+% \mbox{~}
+\thispagestyle{empty}
+\label{roosterhydra}
+ {\centering
+      \includegraphics[width=12cm]{rooster-hydra.png}\\
+      \centerline{\emph{``The incomplete rooster hydra'': AI art generated by Karl Palmskog}}
 }
@@
 
-Last but not least, many thanks to Pierrette Cassou-Noguès for the watercolor on page~\pageref{watercolor}.
+Last but not least, many thanks to Pierrette Cassou-Noguès for the watercolor on pages~\pageref{watercolorgray} and
+\pageref{watercolor}. Thanks to Karl Palmskog for his
+\emph{rooster hydra}, page~\pageref{roosterhydra}.
 
@@
 This work is maintained by Coq community~\cite{CoqCommunity} volunteers: (projects Goedel~\cite{Goedel} and Hydra-battles~\cite{HydraBattles}).
+The main reference to this work is  Russel O'Connor's article~\cite{OConnor05}, which we strongly encourage the reader to 
+consult regularly.
 
@@
 \include{chapter-primrec}
+\include{chapter-fol}
+
 \include{chapter-encoding}

### 2022-10-17 doc/chapter-fol.tex
more comments on FOL

@@
 
-Let us give two examples of first-order formulas.
+Let us give a few examples of first-order formulas.
 
-\begin{itemize}
-\item $\forall\,v_0\;v_1,\, v_0<v_1\, \longrightarrow\, v_0<v_1 + v_0$
+\label{fol:examplesf1f2f3}
+\begin{enumerate}
+\item $\forall\,v_0, v_0=0 \vee \exists\;v_1,\, v_0= S\;v_1$
+\item $\exists\,v_2,\, 0< v_2 \wedge 4=v_2+v_2$
+\item $ v_0=0 \vee \exists\;v_1,\, v_0= S\;v_1$
+\item $v_0= v_1 + v_1 \leftrightarrow v_0 = v_1 \times 2$
+\end{enumerate}
 \inputsnippets{fol_Examples/f1Example}
-\item $\forall\,v_0,\,\exists\,v_1,\,v_0<v_1$
 \inputsnippets{fol_Examples/f2Example}
 
-\end{itemize}
+\begin{todo}
+Comment on the use of boths list and sets of formulas.
+(explained in Russel's paper, p.7).
+
+\inputsnippets{fol/SystemDef}
+
+\end{todo}
 
@@
 
+The purpose of this section is to present ways of proving 
+properties of any formula $a$ (or defining functions over the type of formulas).
+
+Structural induction/recursion principles are generated by 
+\texttt{Formula}'s definition, for instance:
+
+\inputsnippets{fol_Examples/FormulaRect}
+
 \subsubsection{Depth of a formula}
+As explained in~\cite{OConnor05}, some functions on the \texttt{Formula} type cannot be defined by structural recursion in 
+O'Connors formalization. For instance, the substitution of a term to a free variable (because of quantifiers). 
+In such cases, we may define functions by well-founded recursion. In this section, we consider the well-founded strict-order associated with a measure: the \emph{depth} of a formula.
 
-\inputsnippets{fol/depthDef}
+\begin{todo}
+Forward link to its application.
+\end{todo}
 
-\inputsnippets{fol_Examples/ltdepth1}
+The function \texttt{depth} computes the \emph{depth} of the 
+ skeleton of any formula.
+
+\inputsnippets{fol/depthDef, fol_Examples/depthCompute}
+
+\subsubsection{Induction on depth}
+
+The function \texttt{depth} allows us to define a strict order on 
+\texttt{Formula $L$}.
+
+\inputsnippets{fol/ltDepth, fol_Examples/ltdepth1}
 
@@
 
+\inputsnippets{fol_examples/depthRecDemo}
+
 \begin{todo}
@@
 \end{todo}
+
+The library~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} contains several induction principles, applied 
+throughout \texttt{Ackermann} and \texttt{Goedel} projects.
+
+Let us for instance have a look  at \texttt{Formula\_depth\_ind2}. Its application in order to prove $P\;a$ generates 5 sub-goals.
+
+\inputsnippets{fol_Examples/depthRecDemo2}
+
+\begin{itemize}
+\item Goals $1$ to $4$ correspond to  usual proofs by structural induction.
+\item Goal $5$ is associated with a universal quantification $f=\forall\,v,a$.
+In this case, we have to prove that $P\;b$ holds for any formula $b$ which has a depth strictly less than $f$.
+Such a $b$ may for instance the result of replacing the free occurrences of $v$ in $a$ with any term $t$.
+  \begin{todo}
+   Make a link to an appropriate example.
+  \end{todo}
+\end{itemize}
+
+\subsection{Free variables and substitutions}
+
+Section \texttt{Free\_Variables} 
+of~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} defines the list of free variables which occur in a term (respectively in a formula).
+
+We advise the reader to look at these functions, and limit ourselves to give a few examples.
+
+Let us consider again the formulas \texttt{f3} and \texttt{f4} of page~\pageref{fol:examplesf1f2f3}.
+
+The list of free variable occurrences of a formula is given by the function \texttt{freeVarFormula}. The function \texttt{close}
+binds all the free variables of a formula with universal quantifiers.
+
+\inputsnippets{fol_Examples/freeVarExamples}

### 2022-10-28 doc/chapter-fol.tex
fix a latex error
fix a snippet error
Minor changes in Latex files

@@
 
-In order to get more readable terms and formulas, we can define a few notations in ~\href{../theories/html/hydras.MoreAck.fol_Examples.html}{MoreAck.fol\_Examples}
-
-\inputsnippets{fol_Examples/instantiations}
-
 For instance, the term $v_1+0$ is written as below.
 
-\inputsnippets{fol_Examples/v1Plus01}
+\inputsnippets{fol_Examples/v1Plus0}
 
@@
 \inputsnippets{fol_Examples/f1Example}
-\inputsnippets{fol_Examples/f2Example}
 
@@
 
-\inputsnippets{fol/depthDef, fol_Examples/depthCompute}
+\inputsnippets{fol/depthDef,  fol_Examples/DepthCompute}
 
@@
 
-\inputsnippets{fol/ltDepth, fol_Examples/ltdepth1}
+\inputsnippets{fol_Examples/ltDepth1}
 
@@
 
-\inputsnippets{fol_examples/depthRecDemo}
+\inputsnippets{fol_Examples/depthRecDemo}
 
@@
 \inputsnippets{fol_Examples/freeVarExamples}
+
+In order to get more readable terms and formulas, we can define a few notations in ~\href{../theories/html/hydras.Ackermann.LNN.html}{Ackermann.LNN\_Examples}
+
+\inputsnippets{LNN/instantiations}

### 2022-11-14 doc/chapter-fol.tex

@@
 
-In order to get more readable terms and formulas, we can define a few notations in ~\href{../theories/html/hydras.Ackermann.LNN.html}{Ackermann.LNN\_Examples}
+In order to get more readable terms and formulas, we can define a few notations in ~\href{../theories/html/hydras.Ackermann.LNN.html}{Ackermann.LNN}
 

### 2022-11-18 doc/chapter-fol.tex
experimental notations

@@
 
-In order to get more readable terms and formulas, we can define a few notations in ~\href{../theories/html/hydras.Ackermann.LNN.html}{Ackermann.LNN}
+\inputsnippets{LNN/instantiations}
 
-\inputsnippets{LNN/instantiations}
+\section{Notations for Formulas (experimental)}
+
+In order to get more readable terms and formulas, we can define a few notations in ~\href{../theories/html/hydras.MoreAck.FOL_notations.html}{MoreAck.FOL\_notations} and
+~\href{../theories/html/hydras.Ackermann.LNN_notations.html}{MoreAck.LNN\_notations}.
+
+\section{Proofs}
+
+\subsection{Contexts as lists}
+
+\begin{exercise}
+
+Prove the following lemma:
+\inputsnippets{fol_Examples/eqRefl}
+
+\end{exercise}
+
+\subsection{Contexts as sets}
+
+\begin{exercise}
+
+Prove the following lemma:
+\inputsnippets{fol_Examples/MPdiag}
+
+\end{exercise}

### 2022-11-21 doc/chapter-fol.tex
minor changes, bullets in folLogic

@@
 In order to get more readable terms and formulas, we can define a few notations in ~\href{../theories/html/hydras.MoreAck.FOL_notations.html}{MoreAck.FOL\_notations} and
-~\href{../theories/html/hydras.Ackermann.LNN_notations.html}{MoreAck.LNN\_notations}.
+~\href{../theories/html/hydras.MoreAck.LNN_notations.html}{MoreAck.LNN\_notations}.
+Please note that these notation scopes are experimental: We are going to use them in examples and exercises before using them in large original proof scripts (in the \texttt{ordinals/Ackermann/} sub-directory).
+
+We try to define notation scopes as close as possible to \coq's syntax for propositions.
+
+Let us take for instance the following proposition (in math form):
+
+$$\forall\, v_0,\, v_0=0\vee \exists\,v_1,\,v_1=1+v_0$$
+
+Here is a definition, using directly the \texttt{goedel/Ackermann}'s project syntax.
+
+\inputsnippets{LNN_Examples/uglyF0}
+
+Note that, because of redefinitions, the disjonction \texttt{orH}
+can be expanded in terms of  implication and negation (for instance when we use \texttt{Compute}).
+
+\inputsnippets{LNN_Examples/uglyF0a}
+
+\inputsnippets{LNN_Examples/CNNF0}
 

### 2022-11-28 doc/chapter-fol.tex
Example of folLogic

@@
 \end{exercise}
+
+\section{Derived rules and natural deduction}
+
+ The library 
+ \href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic} contains many derived rules which allow the user to build proofs in a natural deduction style (with introduction and elimination rules).
+
+\subsection{Example}
+
+For instance, let us prove Peirce's rule.
+ (in  ~\href{../theories/html/hydras.MoreAck.folPeirce.html}{MoreAck.folPeirce} ).
+
+\subsubsection{Prelude}
+
+\inputsnippets{folPeirce/prelude}
+
+\subsection{Proof of Pierce's law}
+
+\inputsnippets{folPeirce/PeirceProof}
+
+Let us try to do an implication introduction.
+
+\inputsnippets{folPeirce/step1}
+
+Now, we may use the law of excluded middle with the formula $P$. The only non-trivial case is about $\sim P$.
+
+\inputsnippets{folPeirce/step2}
+
+The rest of the proof is composed of basic proof steps, 
+and bookkeeping steps (about \texttt{Ensembles.In}).
+
+\inputsnippets{folPeirce/step3, folPeirce/step4}

### 2022-12-05 doc/hydras.tex
minor changes
Minor changes in Latex files
New notations (#139)
* experiment fol notations

* first attempts

* experiment FOL notations

* experiment FOL notations

* minor change

* minor change

* minor change

* experimental notations

* experimental notations

* minor changes, bullets in folLogic

* bulletizing folLogic.v (half done, to resume at forallE)

* bulletized folLogic.v

* Example of folLogic

* more examples

* drinkers theorem in FOL (clumsy draft)

* re-structuring prro scripts in folLogic2

* re-structuring proof scripts in folLogic[23]

@@
 \makeindex{additions}
-\makeindex{primrec}
+\makeindex{ackermann}
 \begin{document}
@@
 \\ With contributions by Yves Bertot, Ilm\={a}rs C\={i}rulis, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Karl Palmskog, Clément Pit-Claudel, and Théo Zimmermann.
-\\ The formalization of primitive recursive functions was originally authored by Russel O'Connor.}%\cite{OConnor05}.}
+\\ The formalization of primitive recursive functions and Peano Arithmetic was originally authored by Russel O'Connor.}%\cite{OConnor05}.}
 
@@
 \printindex{maths}{Mathematical notions and algorithmics}
-%\printindex{hydras}{Library hydras: Ordinals and hydra battles}
-%\printindex{primrec}{Library hydras.Ackermann: Primitive recursive functions}
-%\printindex{additions}{Library additions: Addition chains}
+\printindex{hydras}{Library hydras: Ordinals and hydra battles}
+\printindex{ackermann}{Library hydras.Ackermann: Primitive recursive functions, G\"{o}del encoding}
+\printindex{additions}{Library additions: Addition chains}
 

### 2022-12-21 doc/chapter-fol.tex
more examples
re-structuring prro scripts in folLogic2
documentation on FOL (draft)

@@
 \subsection{Contexts as lists}
+\begin{todo}
+Describe the \texttt{Prf} type (from \texttt{folProof.v}).
+
+Give examples in a toy language.
+
+Give an example which motivates \texttt{SysPrf}:
+
+$P,P , P\arrow P \arrow Q, \vdash Q$ ?
+\end{todo}
 
@@
 
-\subsection{Proof of Pierce's law}
+\subsection{Proof of Peirce's law}
 
@@
 \inputsnippets{folPeirce/step3, folPeirce/step4}
+
+\index{ackermann}{Exercises}
+\begin{exercise}
+Prove, using the rules described in 
+ \href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic}, the famous \emph{drinkers theorem}:
+
+$$\exists\,x,\; (D(x)\Longrightarrow \forall\,y,\; D(y))$$
+\end{exercise}
+
+where $D$ (for ``drinks'') is some predicate symbol of arity $1$.

### 2022-12-21 doc/hydras.tex
documentation on FOL (draft)

@@
 
-\emph{\color{red!80}This maintenance job has just started, and will probably be long to complete. Help is welcome!}
+\emph{\color{red!80}These maintenance and documentation 
+jobs have just started, and will probably be long to complete. Help is welcome!}
+
+\section{File contents}
+
+All Russel O'Connor's files have been stored in two directories, in order to simplify packages maintenance.
+
+\begin{itemize}
+\item \texttt{theories/goedel/}: Proofs which depend on 
+\texttt{CoqPrime} package.
+\item  \texttt{theories/ordinals/Ackermann/} : all the rest:
+ definition of primitive recursive functions, first-order  logic,
+Peano Arithmetics, G\"{o}del's encoding.
+\item Some additions we made: examples, exercices, new notations, etc.,  are stored in a specific directory \texttt{theories/ordinal/MoreAck/}.
+\end{itemize}
+
+\section{Warning}
+Russel O'Connors contribution contains more than 42 KLoc.
+Since its construction, \coq{},  its libraries and recommended style have evolved a lot.
+We have just started to ``modernize'' this code.
+We apologize for provisional inconsistencies of presentation (code and documentation).
 

### 2022-12-27 doc/chapter-fol.tex
documentation on FOL (draft)

@@
 \label{chap:fol}
+
+\section{Introduction}
 
@@
 \inputsnippets{fol/LanguageDef}
+
+\subsubsection{A toy language}
+In order to show a few simple examples of statements and proofs, we define a small language with very few symbols:
+two constant symbols: $a$ and $b$, a function symbol $f$ (of arity 1), three propositional symbols $A$, $B$ and $C$, a one-place predicate symbol $P$, and a binary relational symbol $R$.
+
+\inputsnippets{folExample2/ToyDef}
+
+\subsubsection{Languages for Arithmetic}
 
@@
 \inputsnippets{fol/TermDef}
+
+\begin{remark}[Variables]
+In O'Connor's formalization of first-order logic, the variables are \gallina terms of the form \texttt{var $L$ $i$}, where $L$ is
+ a language and $i$ a natural number. 
+Although other choices may be considered : PHOAS, de Bruijn indices, etc,  we   respect this representation, in order not to break long proof scipts which use this representation (please look at \cite{OConnor05} for a related discussion).
+
+\end{remark}
+\subsubsection{Examples}
+
+With our toy language, we can build the following terms, respectively written $a$, $f(a)$ and $f(f(v_1))$ in usual mathematical notation.
+
+Writing these terms in bare \gallina syntax would result in quite clumsy scripts. Thus, we may use a still experimental notation scope which allows us to write/print terms and formulas in a 
+notation similar to \coq's.
+
+We put both the associated \coq{} terms, in bare \gallina{} syntax, and with a still experimental and provisional notation, defined in
+~\href{../theories/html/hydras.MoreAck.FOL_notations.html}{MoreAck.FOL\_notations}).
+
+\inputsnippets{folExample2/toyNotation}
+
+\inputsnippets{folExample2/smallTerms}
 
@@
 \inputsnippets{fol_Examples/f1Example}
+
+\begin{remark}[bound variables]
+\inputsnippets{folExample2/boundVars}  
+\end{remark}
 

### 2023-02-16 doc/hydras.tex
documentation on FOL (draft)
fix an inconsistent section numbering

@@
 \part{Ackermann, G\"{o}del, Peano\,\dots}
+
+\chapter{Introduction}
+
 This part contains comments, example  and exercises about Russel O'Connor's work on G\"{o}del's first incompleteness theorem.
@@
 \end{itemize}
+
+\begin{todo}
+Add information on recent developments on formal 
+proofs of G\"odel  incompleteness theorems. Justify the decision of working on \emph{this} development.  
+\end{todo}
 

### 2023-02-16 doc/chap-intro-goedel.tex
small correction in latex document

--- 2023-02-16 doc/hydras.tex
@@
-\documentclass[twoside,a4paper]{book}
-
-%\usepackage{fontenc}
-\usepackage[utf8x]{inputenc}
-
-\usepackage{fontspec}
-%\setmainfont{DejaVu Serif}
-
-\setmonofont{DejaVu Sans Mono}[Scale=MatchLowercase]
-\usepackage{alltt}
-\usepackage{multind}
-\usepackage{hyperref}
-\usepackage{url}
-\usepackage{xcolor, fancyvrb, mdframed}
-\usepackage{amsmath}
-%\usepackage{fancyhdr}
-\usepackage{varioref}
-\usepackage{synttree, proof, centernot}
-%\usepackage[firstpage]{draftwatermark}
-\usepackage{verbatim}
-\usepackage[note]{marginote}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{mathtools}
-\usepackage{dsfont}
-\usepackage{float}
-\usepackage{threeparttable}
-\usepackage{fontspec}
-\usepackage{mathtools}
-\usepackage{xcolor}
-\usepackage{caption}
-\usepackage{placeins}
-\usepackage{tikz}
-\usepackage{tikzsymbols}
-\usetikzlibrary{arrows}
-\usepackage{graphicx}
-\usepackage{amsmath,mathdots,dsfont}
-\usepackage{amssymb}
-\usepackage{newunicodechar}
-\usepackage{theorem}
-\usepackage{xspace}
-\usepackage{texments}
-
-%%% for movies by alectryon
-\usepackage{movies/snippets/assets/alectryon}
-\usepackage{movies/snippets/assets/pygments}
-
-% Prevent breaks in the middle of syntactic units
-\let\OldPY\PY
-\def\PY#1#2{\OldPY{#1}{\mbox{#2}}}
-
-\setcounter{tocdepth}{1}
-\input{coqmacros}
-
-\newcommand{\gaiacompat}{\textcolor{blue}{Gaia}}
-\newcommand{\gaiasign}{\texorpdfstring{\mycircled[orange!40]{{\mbox{\color{blue!65}{\scriptsize{G}}}}}}{Gaia-Hydra}}
-\newtheorem{gaiahydra}{\gaiasign}
-
-%%% One hypothesis per line 
-\makeatletter
-\renewcommand{\alectryon@hyps@sep}{\alectryon@nl}
-\makeatother
-
-%%% \snippets{A,B,C,…} inputs a series of snippets as one block (with \itemsep
-%%% between them).  A, B, C should be paths to files in movies/snippets/.
-\usepackage{etoolbox}
-\makeatletter
-\newcounter{snippets}
-\newcommand{\inputsnippets}[1]
-  {{\setlength{\itemsep}{1pt}\setlength{\parsep}{0pt}% Adjust spacing
-    \alectryon@copymacros\begin{io}
-      \forcsvlist{\item\@inputsnippet}{#1}
-    \end{io}}}
-\let\input@old\input % Save definition of \input
-\newcommand{\@inputsnippet}[1]
-  {{\renewenvironment{alectryon}{}{}% Skip \begin{alectryon} included in snippet
-    \refstepcounter{snippets}%
-    \input@old{movies/snippets/#1}}}
-\AtEndDocument{\message{Snippets: \thesnippets}}
-\makeatother
-
-% \SetWatermarkLightness{0.8}
-\definecolor{darkblue}{rgb}{0,0.1,0.7}
-% \SetWatermarkScale{0.35}
-% \SetWatermarkText{Work Continuously in Progress}
-% \SetWatermarkColor{lightorange}
-
-% \author{Pierre Castéran\\ Univ.
-Bordeaux, CNRS, Bordeaux INP, LaBRI, UMR 5800, F-33400 Talence, France\footnote{ email: pierre casteran at gmail dot com} \thanks{With contributions by Yves Bertot, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Karl Palmskog, Clément Pit-Claudel, and Théo Zimmermann.
-The formalization of primitive recursive functions was originally authored by
-% Russel O'Connor\cite{OConnor05}.}}
-% \date{\today}
-
-\title{Hydras \& Co.}
-%\pagestyle{fancy}
-%\fancyhead{Hydras, Ordinals, \& Co.}
-\makeindex{gaiabridge}
-\makeindex{coq}
-\makeindex{maths}
-\makeindex{hydras}
-\makeindex{additions}
-\makeindex{ackermann}
-\begin{document}
-\begin{titlepage}
-   \begin{center}
-       \vspace*{1cm}
-
-       \textbf{\Huge Hydras \& Co.}
-
-       \vspace{0.5cm}
-        {\emph{\Large Formalized mathematics in \coq for inspiration and entertainement}}
-
-       \vspace{1.5cm}
-
-       \emph{Pierre Castéran, LaBRI, Univ.
-Bordeaux, CNRS UMR 5800\\ \textbf{email:} pierre dot casteran arobas gmail dot com.
-\\ With contributions by Yves Bertot, Ilm\={a}rs C\={i}rulis, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Karl Palmskog, Clément Pit-Claudel, and Théo Zimmermann.
-\\ The formalization of primitive recursive functions and Peano Arithmetic was originally authored by Russel O'Connor.}%\cite{OConnor05}.}
-
- \vfill
-        \today
-
-      \end{center}
-\vfill
-      \centering
-      \includegraphics[width=13cm]{epsilon0.jpg}
-      \centerline{\emph{\small {\color{darkblue}Ordinal numbers in Veblen normal form}}}
-\end{titlepage}
-
-\clearpage
-\newpage
-\thispagestyle{empty}
-\begin{quote}
- \emph{(Aprieta el bast\'on con las dos manos, se yergue un tanto,
-casi con entusiasmo)} ¡Caramba! Claro \dots{} los n\'umeros
-transfinitos, Kantor \dots{}
-
-[ Jorge Luis Borges]
-\end{quote}
-\vspace{17mm}
-\begin{quote}
-  Nessun senso percepisce l'infinito. Nessun senso permette di concludere ch'esso esista. L'infinito, in effetti, non puo' essere l'oggetto dei sensi.
-
-  [Giordano Bruno] \emph{De l'infinito, universo e mondi}
-\end{quote}
-
-\vspace{17mm}
-\begin{quote} 
-I start from one point and go as far as possible.
-
-[John Coltrane]
-\end{quote}
-\vspace{6pt}
-\mbox{}
-\label{watercolorgray}
- {\centering
-      \includegraphics[width=12cm]{hydre-pierrette.jpg}
-      \centerline{\emph{``The hydra in black and white'': Watercolor  by Pierrette Cassou-Noguès}}
-}
-\newpage
-\newpage
-% \mbox{~}
-\thispagestyle{empty}
-\label{watercolor}
- {\centering
-      \includegraphics[width=13cm]{blue-hydra.jpg}
-      \centerline{\emph{``The blue hydra'': Watercolor  by Pierrette Cassou-Noguès}}
-}
-
-\newpage
-% \mbox{~}
-\thispagestyle{empty}
-\label{roosterhydra}
- {\centering
-      \includegraphics[width=12cm]{rooster-hydra.png}\\
-      \centerline{\emph{``The incomplete rooster hydra'': AI art generated by Karl Palmskog}}
-}
-
- \tableofcontents
-
-%-------------------------------------------------------------------
-
-\chapter{Introduction}
-
-\vspace{16pt}
-
-\section{Generalities}
-
-Proof assistants are excellent tools for exploring the structure of mathematical proofs,
-studying  which hypotheses are really needed, and which proof patterns are useful and/or
-necessary. Since the development of a theory is represented as a bunch of computer files,
-everyone is able to read the proofs with an arbitrary level of detail, or to play with the theory by writing alternate proofs or definitions.
-
-Among all the theorems proved with the help of proof assistants like \coq{}~\cite{Coq,BC04}, \hol{}~\cite{HOL}, \isabelle{}~\cite{isabelle},  etc.,
-several statements and proofs  share some interesting features:
-\begin{itemize}
-\item Their statements are easy to understand, even by non-mathematicians
-\item Their proof requires some non-trivial mathematical tools
-\item Their mechanization on computer presents some methodological interest.
-\end{itemize}
-
-This is obviously the case of the four-color theorem~\cite{fourcolors}  and the Kepler conjecture~\cite{flyspeck2015}.
-We do not mention impressive works like the proof of the odd-order theorem ~\cite{oddorderthm}, since understanding its statement requires a quite good mathematical culture.
-
-In this document, we present two examples which seem to have the above properties.
-
-\begin{itemize}
-\item Hydra games (a.k.a. \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians:
-L. Kirby and J. Paris~\cite{KP82}: \emph{Accessible Independence Results for Peano Arithmetic}. 
-Although the mathematical contents of this 
-paper are quite advanced, the rules of hydra battles are very easy to understand\footnote{Let us underline the analogy between hydra battles and interactive theorem proving.
-Hercules is the user (you!), and hydra's heads are the subgoals: you may think that applying a tactic would solve a subgoal, but it results often in the multiplication of such tasks.}.
-There are now several sites on the Internet where you can find tutorials on hydra games, together with simulators you can play with.
-See, for instance, the blogpost and source code written by Andrej Bauer~\cite{bauer2008,BauerHydra}.
-
-Hydra battles, as well as Goodstein sequences~\cite{goodstein_1944, KP82}
-are a nice way to present complex termination problems.
-The article by Kirby and Paris presents a proof of termination
-based on ordinal numbers, as well as a proof that this termination is not
-provable in Peano arithmetic. In the book dedicated to 
-J.P. ~Jouannaud \cite{HommageJPJ}, N.~Dershowitz and G.~Moser  give a thorough survey on this topic~\cite{Dershowitz2007}.
-
-We present a (still partial, under continuous development) implementation in \coq of the various techniques shown in
-Kirby \& Paris' and Ketonen \& Solovay's~\cite{KS81} article.
-
-\index{gaiabridge}{Introduction}
-Our library \gaiaHydras is dedicated to make compatible our lemmas with José Grimm's \gaia project (designed for \ssreflect/\mathcomp) (please look at Sect.~\vref{sect:gaia-first-intro} and the paragraphs signalled with {\gaiasign}).
-
-\item In the second part, we are interested in computing $x^n$ with the least number of multiplications as possible.
-We use the notion of \emph{addition chains}~\cite{brauer1939,DBLP:journals/ipl/BerstelB87}, to generate efficient certified exponentiation functions.
-\end{itemize}
-
-\paragraph*{Warning:}
-
-This document is \emph{not} an introductory text for \coq, and there are many aspects of this proof assistant that are not covered. 
-The reader should already have some basic experience with the \coq system.
-The Reference Manual and several tutorials are available on the \coq website~\cite{Coq}.
-The first chapters of textbooks like \emph{Interactive Theorem Proving and Program Development}~\cite{BC04},
-\emph{Software Foundations}~\cite{SF},
-\emph{Programs and Proofs}~\cite{Sergey:PnP},
-or  \emph{Certified Programming with Dependent Types} ~\cite{chlipalacpdt2011} will give you the right background.
-
- \subsection{Structure of \Hydras}
-
- \Hydras is made of three main packages: \HydrasLib, \gaiaHydras, and \additions. Figure~\ref{fig:genealogy}  illustrates
- the complex relationships: inheritance from historical contributions to \coq, and dependency with other \coq packages.
- Many thanks to  Karl Palmskog and Théo Zimmermann for the CI/CD design of \Hydras and the automation of documentation maintenance.
- Please look for a more detailed description in~\cite{jfla2022}.
-
- \begin{figure}[ht]
-\centering
-{\footnotesize
-\begin{tikzpicture}[thick]
-
-\begin{scope}[xshift=-6.5cm]
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) {Goedel };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=1.9cm] (pock) { Pocklington };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=1.8cm] (additions) { Additions };
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=1.8cm] (cantor) { Cantor };
-
-\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=2.2cm] (cats) { Categories in ZFC };
-\end{scope}
-
-\begin{scope}[yshift=-1cm,xshift=-4.25cm]
-  \node[rectangle, draw=black,fill=blue!20,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=5.8cm] (mathcomp) { MathComp };
-\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=1.5cm,left of=paramcoq, node distance=4.5cm] (equations) { Equations };
-\end{scope}
-
-\begin{scope}[yshift=-2cm,xshift=-4cm]
-\node[rectangle, draw=black,fill=orange!30,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
-\node[rectangle, draw=black,fill=blue!20,minimum height=0.5cm,minimum width=1.5cm,left of=hydras, node distance=3.5cm] (pockcc) { Pocklington };
-\node[rectangle, draw=black,fill=orange!30,minimum height=0.5cm,minimum width=1.5cm,right of=hydras, node distance=2.5cm] (chains) { Addition-chains };
-\node[rectangle, draw=black,fill=blue!20,minimum height=0.5cm,minimum width=1.5cm,right of=chains, node distance=2.2cm] (gaia) { Gaia };
-\end{scope}
-
-\begin{scope}[yshift=-3cm,xshift=-6cm]
- \node[rectangle, draw=black,fill=blue!20,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
- \node[rectangle, draw=black,fill=orange!30,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=3cm] (gaiahydras) { Gaia-hydras };
-\end{scope}
-
-\draw[->,dotted] (additions.south) -- (chains.north) ;
-\draw[->,dotted] (goedel) -- (goedelcc) ;
-\draw[->,dotted] (goedel) -- (hydras.north west) ;
-\draw[->,dotted] (cantor) -- (hydras.north east) ;
-\draw[->,dotted] (cantor.south) -- (gaia.north west) ;
-\draw[->,dotted] (pock) -- (pockcc) ;
-\draw[->,dotted] ([xshift=2.1cm]cats) -- (gaia) ;
-
-\draw[->] (hydras) -- (gaiahydras) ;
-\draw[->] (hydras.south west) -- (goedelcc) ;
-\draw[->] (gaia.south west) -- (gaiahydras.east) ;
-
-\draw[->] (equations) -- (hydras) ;
-\draw[->] (mathcomp) -- (gaia) ;
-\draw[->] (mathcomp) -- (chains) ;
-\draw[->] (paramcoq.south) -- (chains) ;
-\draw[->] (pockcc) -- (goedelcc) ;
-\end{tikzpicture}
-}
-\caption{Genealogy and dependencies for \Hydras packages.
-Dotted boxes represent historical Coq contributions, while regular boxes represent maintained \coq packages.
-Orange packages are maintained in the \Hydras GitHub repository, while light blue packages are maintained in other \community repositories.
-Dotted lines represent \coq code ancestry, while regular lines represent direct code dependencies.}
-  \label{fig:genealogy}
-\end{figure}
-
-\subsection{Documenting theories with \alectr}
-
-Quotations of \coq source and answers are progressively replaced from copy-pasted \emph{verbatim} to automatically generated \emph{LaTeX} blocks, using Clément Pit-Claudel's \alectr  tool~\cite{alectryonpaper, alectryongithub}.
-Many thanks to Jérémy Damour, Clément Pit-Claudel  and Théo Zimmermann who designed tools for maintaining consistency between the always evolving \coq{} modules and documentation written in \emph{LaTeX}.
-
-Besides the guarantee of consistency between theories and documentation, we hope to give a corpus for experimenting new ways of documenting the implementation of non-trivial mathematics on a proof assistant.
-
-\subsection{Trust in our proofs}
-\label{sect:trust-in-proofs}
-
-Unlike mathematical literature, where definitions and proofs are spread out over many articles and books,
-the whole proof is now inside your computer. It is composed from the \texttt{.v} files you downloaded and
-parts of \coq's standard library.
-Thus, there is no ambiguity in our definitions and the premises of the theorems.
-Furthermore, you will be able to navigate through the development, using your favorite text editor or IDE, and some commands like \texttt{Search}, \texttt{Locate},  etc.
-
-\subsection{Assumed redundancy}
-
-It may often happen that several definitions of a given concept, or several proofs of a given theorem are possible.
-If all the versions present some interest, we will make them available, since each one may be of some methodological
-interest (by illustrating some tactic of proof pattern, for instance).
-We use \coq's module system to make several proofs of a given theorem co-exist in our libraries (see also Sect~\vref{sect:alt-proofs}).
-After some discussions of the pros and cons of each solution, we develop only one of them, leaving the others  as exercises or projects (i.e., big or difficult exercises).
-In order to discuss which assumptions are really needed for proving a theorem, we will also present 
-several aborted proofs.
-Of course, do not hesitate to contribute nice proofs or alternative definitions!
-
-It may also happen that some proof looks to be useless, because the proven theorem is a trivial consequence of another (also proven) result.
-For instance, let us consider the three following statements:
-\begin{enumerate}
-\item There is no measure into $\mathbb{N}$ for proving the termination of all hydra battles (Sect~\vref{omega-case}).
-\item There is no measure into the interval\,\footnote{We use the notation $[a,b)$ for denoting the set of ordinals greater or equal than $a$ and strictly less than $b$.}  $[0,\omega^2)$ for proving the termination of all hydra battles (Sect.~\vref{omega2-case}).
-\item There is no measure into $[0,\mu)$ for proving the termination of all hydra battles, for any $\mu<\epsilon_0$ (Sect.\vref{sec:free-battles-case}).
-\end{enumerate}
-
-Obviously, the third theorem implies the second one, which implies the first one.
-So, theoretically, a library would contain only a proof of $(3)$ and remarks for $(2)$ and $(1)$.
-But we found it interesting to make all the three proofs available, allowing the reader to compare their common structure and notice their technical differences.
-In particular, the proof of $(3)$ uses several non-trivial combinatorial properties of ordinal numbers up to $\epsilon_0$~\cite{KS81}, whilst $(1)$ and $(2)$ use simple properties of $\mathbb{N}$ and $\mathbb{N}^2$.
-
-\subsection{About logic}
-
-Most of the proofs we present are \emph{constructive}.
-Whenever possible, we provide the user with an associated function, which she or he can apply in \gallina{} or \ocaml{} in order to get a ``concrete'' feeling of the meaning of the considered theorem.
-For instance, in Chapter~\vref{chap:ketonen}, the notion of \emph{limit ordinal} is
-made more ``concrete'' thanks to a function \texttt{canon} which computes every item of a sequence which converges on a given limit ordinal $\alpha$.
-This simply typed function allows the user/reader to make her/his own experimentations.
-For instance, one can very easily compute the $42$-nd item of a sequence which converges towards $\omega^{\omega^\omega}$.
-
-Except in the \texttt{Schutte} library, dedicated to an axiomatic presentation of the set of countable ordinal numbers, all of our development is axiom-free, and respects the rules of intuitionistic logic.
-Note that we also use the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777} in the definition of  several rapidly growing hierarchy of functions, in Chap.~\ref{chap:alpha-large}.
-This plug-in imports several known-as-harmless  axioms.
-
-% \begin{Coqsrc}
-% FunctionalExtensionality.functional_extensionality_dep : 
-% forall (A : Type) (B : A -> Type) (f g : forall x : A, B x),
-% (forall x : A, f x = g x) -> f = g
-% \end{Coqsrc}
-
-%\index{coq}{Commands!Print Assumptions}
-
-At any place of our development, you may use the  \texttt{Print Assumptions {\it ident}} command in order to verify on which axiom the theorem {\it ident} may depend.
-The following example is extracted from
-Library~\href{../theories/html/hydras.Epsilon0.F_alpha.html}{hydras.Epsilon0.F\_alpha}, where we use the \texttt{coq-equations} plug-in (see Sect.~\vref{sect:wainer}).
-
-\input{movies/snippets/F_alpha/DemoAssumptions}
-
-\subsection{Typographical Conventions}
-
-\subsubsection{Using \alectr}
-Whenever possible, we use \alectr to display \coq code (definition, proof scripts) and answers.
-
-Here are two  examples from Chapters~\ref{chapter:primrec}
-and~\ref{chapter-powers}.
-
-\vspace{4pt}
-
-\input{movies/snippets/Ack/AckFixpointIterate.tex}
-
-\inputsnippets{Fib2/fibEuclDemo}
-
-\subsubsection{Verbatim quotations}
-
-Quotations of \coq{} source  from others libraries (\coq's standard library, borrowed plug-ins)  are displayed as follows.
-
-\begin{Coqsrc}
-  Inductive CompareSpec (Peq Plt Pgt : Prop) :
-  comparison -> Prop :=
-    CompEq : Peq -> CompareSpec Peq Plt Pgt Eq
-  | CompLt : Plt -> CompareSpec Peq Plt Pgt Lt
-  | CompGt : Pgt -> CompareSpec Peq Plt Pgt Gt.
-\end{Coqsrc}
-
-We use also verbatim code inclusions when the examples lead to very long computations.
-
-\begin{Coqbad}
-Example C87_ok_slow : chain_correct 87 C87.
-Proof.
-Time slow_chain_correct_tac.
-\end{Coqbad}
-
-\begin{Coqanswer}
-Finished transaction in 49.927 secs (49.445u,0.079s) (successful)
-\end{Coqanswer}
-
-\begin{Coqbad}
-Qed.
-\end{Coqbad}
-
- \subsection{Remark}
-
-In general, we do not include full proof scripts in this document.
-The only exceptions are very short proofs (\emph{e.g.}, proofs by computation, or by application of automatic tactics).
-Likewise, we may display only the important steps on a long interactive proof, for instance, in the following lemma (\vref{lemma:L-2_6-1}):
-
-\input{movies/snippets/Paths/Lemma261}
-
-The reader may consult the full proof scripts with Proof General or CoqIDE, for instance.
-
-\subsection{Active links}
-The  links which appear in this pdf  document lead are of three possible kinds of destination:
-\begin{itemize}
-\item Local links to the document itself,
-\item External links, mainly to \coq's website,
-\item Local links to pages generated by \texttt{coqdoc}. According to the current makefile (through the commands \texttt{make html} and \texttt{make pdf}), 
-  the pages generated by \texttt{coqdoc} are stored at 
-the relative address \texttt{../theories/html/*.html} (from the location of the pdf).
-Thus,  active links to our \coq{} modules may be incorrect if you did not get this \texttt{pdf} document by compiling the distribution available at
-\url{https://github.com/coq-community/hydra-battles}.
-
-\end{itemize}
-
-\subsection{Alternative or bad definitions}
-\label{sect:alt-proofs}
-Finally, we decided to include definitions or lemma statements, as well as tactics,  that lead to
-dead-ends or too complex developments, with the following coloring.
-Bad definitions 
- are ''masked'' inside modules called \texttt{Bad}, \texttt{Bad1}, etc.
-
-\input{movies/snippets/Schutte_basics/BadBottoma}
-\input{movies/snippets/Schutte_basics/trivialProps}
-\input{movies/snippets/Schutte_basics/Failure}
-
-Likewise, alternative, but still unexplored definitions will be presented in modules
-\texttt{Alt}, \texttt{Alt1}, etc. Using these definitions is left as an implicit exercise.
-
-\input{movies/snippets/Hydra_Definitions/HydraAlt}
-
-\section{How to install the libraries}
-\label{sec:orgheadline4}
- The present distribution has been checked with version 8.14.1 of the Coq proof assistant, with a few plug-ins.
-\emph{Please refer to \href{https://github.com/coq-community/hydra-battle\#readme}{the README file of the project}.}
-
-\section{Comments on exercises and projects}
-
-Although we do not plan to include complete solutions to the exercises, 
-we think it would be useful to include comments and hints, and questions/answers from the users. In contrast, ``projects'' are supposed, once completed, to be included in the repository.
-
-Please consult the sub-directory \texttt{exercises/} of the
- project (in construction).
-
-\section{Acknowledgements}
-\label{sec:orgheadline5}
-    Many thanks to Yves Bertot, Ilm\={a}rs C\={i}rulis, \'Evelyne Contejean, Jéremy Damour,   Florian Hatat,  David Ilcinkas, 
-Pascal Manoury,  Karl Palmskog, Cl\'ement Pit-Claudel, Sylvain Salvati, Alan Schmitt and Théo Zimmermann for their help on the elaboration of this library and  document, and to the
- members of the \emph{Formal Methods} team and the \emph{Coq working group} at laBRI for their helpful comments 
-on  oral presentations of this work.
-
-Many thanks also to the Coq development team and the members of the \emph{Coq Club} for interesting discussions about the \coq{} system and the Calculus of Inductive Constructions.
-
-The author of the present document wishes to express his gratitude to the late Patrick Dehornoy, whose talk  was determinant for our desire to work on this topic.
-I owe my interest in discrete mathematics and their relation to formal proofs and functional programming  to Srecko Brlek.
-Equally, there is W.
-H.
-Burge's book ``\emph{Recursive Programming Techniques}'' ~\cite{burge} which was a great  source of inspiration.
-
-Last but not least, many thanks to Pierrette Cassou-Noguès for the watercolor on pages~\pageref{watercolorgray} and
-\pageref{watercolor}. Thanks to Karl Palmskog for his
-\emph{rooster hydra}, page~\pageref{roosterhydra}.
-
-\subsection{Contributions}
-
-Yves Bertot made nice optimizations  to algorithms presented in Chapter~\ref{chapter-powers}.
-\'Evelyne Contejean contributed libraries on the recursive path ordering (\emph{rpo}) for proving the well-foundedness of our representation of $\epsilon_0$ and $\Gamma_0$.
-Florian Hatat proved many useful lemmas on countable sets, which we used in our adaptation of Schütte's formalization of countable ordinals.
-Pascal Manoury is integrating the ordinal $\omega^\omega$ into our hierarchy of ordinal notations.
-
-The formalization of primitive recursive functions was originally a part of  Russel O'Connor's work on G\"odel's incompleteness theorems~\cite{OConnor05}.
-
-\label{sec:orgheadline2}
-
-Any form of contribution  is welcome: correction of errors (typos and more serious mistakes), improvement of
-Coq scripts, proposition of inclusion of new chapters, and generally any
-comment or proposition that would help us. The text contains several \emph{projects} which, when completed, may improve the present work.
-Please do not hesitate to share your contributions, for instance using pull requests and issues on GitHub. Thank you in advance!
-
-% \subsubsection{Links to the Coq source}
-
-\part{Hydras and ordinals}
-
-\include{part-hydras-intro}
-\include{hydra-chapter}
-\include{ordinal-chapter}
-\include{epsilon0-chapter}
-\include{ks-chapter}
-\include{large-sets-chapter}
-\include{gaia-chapter}
-\include{schutte-chapter}
-\include{Gamma0-chapter}
-
-\part{Ackermann, G\"{o}del, Peano\,\dots}
-
 \chapter{Introduction}
@@
 We apologize for provisional inconsistencies of presentation (code and documentation).
-
-\include{chapter-primrec} 
-\include{chapter-fol}
-
-\include{chapter-encoding}
-
-\part{A few  certified algorithms}
-
-\include{chapter-powers}
-
-\part{Appendices}
-
-\bibliographystyle{alpha}
-\bibliography{thebib}
-
-\chapter{Index and tables}
-
-% \begin{todo}
-%  Still very incomplete!
-% \end{todo}
-{\Large \textbf{In progress} This index is currently under reorganization.  We apologize for its incompleteness! }
-
-\printindex{gaiabridge}{Links to \gaia Library}
-\label{gaia-index}
-\printindex{coq}{Coq, plug-ins and standard library}
-\printindex{maths}{Mathematical notions and algorithmics}
-\printindex{hydras}{Library hydras: Ordinals and hydra battles}
-\printindex{ackermann}{Library hydras.Ackermann: Primitive recursive functions, G\"{o}del encoding}
-\printindex{additions}{Library additions: Addition chains}
-
-% \section{Main notations}
-
-% \begin{table}[H]
-%   \centering
-%   \begin{threeparttable}
-%     \caption{Ordinals and ordinal notations}
-% \begin{tabular}{|r | c|c|c|c|l|}
-% \hline
-% Name & Gallina&Math& Description& Page \\\hline
-% \texttt{lt : T1->T1->Prop}& lt alpha beta & $\alpha < \beta$& strict order on type \texttt{T1} \tnote{1} & \pageref{Predicates:lt-T1}\\
-% \texttt{LT: T1->T1->Prop}& alpha o< beta & $\alpha < \beta$& strict order on type \texttt{T1}   \tnote{2} & \pageref{Predicates:LT-T1}\\
-% \texttt{Lt : E0->E0->Prop} & alpha o< beta & $\alpha < \beta$& strict order on type \texttt{E0} \tnote{3} & \pageref{Predicates:Lt-E0} \\
-% \texttt{nf: T1->Prop} & \texttt{nf alpha} && alpha is in Cantor normal form & \pageref{Predicates:nf-T1}\\
-%  \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality \tnote{4} & \pageref{sect:on-lt-notation}\\
-%  \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
-% \texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & ordinal addition & \pageref{sect:infix-plus-T1}, \dots\\
-% \texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\
-
-% F & \texttt{F $n$} & $n$ & The $n$-th finite ordinal &  
-% \pageref{sect:notation-F}, \pageref{sect:notation-F-sch}\\ 
-% FS & \texttt{FS $n$} & $n+1$ & The $n+1$-th finite ordinal  \tnote{5} &  
-% \pageref{sect:notation-FS}\\ 
-% omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega}, \pageref{sect:omega-T1}, \pageref{sect:omega-notation2}, \dots\\
-% phi0     & \texttt{phi0 alpha} & $\phi_0(\alpha),\; \omega^\alpha$&exponential of base $\omega$ & \pageref{sect:notation-phi0}\\
-
-% \hline
-% \end{tabular}
-% \begin{tablenotes}
-%   \item[1] This order is total, but not well-founded, because of not well formed terms.
-% \item[2] Restriction of \texttt{lt} to terms in normal form; this order is partial, but well-founded.
-% \item[3] This order is total \emph{and} well-founded.
-% \item [4]
-% Some notations may belong to several scopes. For instance, ``\texttt{o<}'' is
-% bound in \texttt{ON\_scope}, \texttt{E0\_scope}, \texttt{t1\_scope}, etc., and locally in several libraries.
-%   \item [5] Note that there exist also various coercions from \texttt{nat} to types of ordinal.
-Depending on the current scope and  \coq's syntactic analysis algorithm, \texttt{F} may be left implicit.
-% \end{tablenotes}
-% \end{threeparttable}
-% \end{table}
-
-% \vspace{4pt}
-
-% \begin{table}[H]
-
-%   \begin{threeparttable}
-%     \caption{hydra battles}
-% \begin{tabular}{|c|c|c|c|l|}
-% \hline
-% Name & Gallina&Math& Description& Page \\\hline
-% \texttt{round} & \texttt{h -1-> h'} & & one round of a battle & \pageref{sect:infix-round} \\
-% \texttt{rounds} & \texttt{h -+-> h'} & & one or more  rounds of a battle & \pageref{sect:infix-rounds} \\
-% \texttt{round\_star} & \texttt{h -*-> h'} & & any number of rounds of a battle & \pageref{sect:infix-rounds} \\
-% \hline
-% \end{tabular}
-
-% %\begin{tablenotes}
-% %\end{tablenotes}
-
-%   \end{threeparttable}
-
-% \end{table}
-
-% \begin{table}[H]
-%   \centering
-%   \begin{threeparttable}
-%     \caption{Addition chains}
-% \begin{tabular}{|c|c|c|c|l|}
-% \hline
-% Name & Gallina&Math& Description& Page \\\hline
-% \texttt{Mult} & \texttt{$z$ <--- $x$ times $y$} & & monadic notation & \pageref{monadic-mult} \\
-% \hline
-% \end{tabular}
-
-% %\begin{tablenotes}
-% %\end{tablenotes}
-
-%   \end{threeparttable}
-
-% \end{table}
-
-\end{document}

### 2023-02-20 doc/chap-intro-goedel.tex
small correction in latex document

@@
-\chapter{Introduction}
+\chapter{General presentation (draft)}
 
-This part contains comments, example  and exercises about Russel O'Connor's work on G\"{o}del's first incompleteness theorem.
-This work is maintained by Coq community~\cite{CoqCommunity} volunteers: (projects Goedel~\cite{Goedel} and Hydra-battles~\cite{HydraBattles}).
+\section{Introduction}
+
+This part contains comments, examples  and exercises about Russel O'Connor's work on G\"{o}del's first incompleteness theorem  \cite{Godel1986-GDECW}.
+O'Connor's work was published in 2005~\cite{OConnor05}, and released as a user-contribution of the \coq proof assistant.
+This work is now maintained by Coq community~\cite{CoqCommunity} volunteers and split into two projects: Goedel~\cite{Goedel} and Hydra-battles~\cite{HydraBattles}.
+
 The main reference to this work is  Russel O'Connor's article~\cite{OConnor05}, which we strongly encourage the reader to 
 consult regularly.
+
+% ici
+
+\section{The reason behind this work}
+
+  \begin{itemize}
+  \item Historical interest on G\"{o}del's proof
+  \item Many popularization publications, \emph{e.g.}
+    \cite{smullyan1992godel, Hofstadter1999Godel, GoedelCassou}, vs. technical complexity of the full proof .
+
+    \begin{quote}
+      The purpose of a proof is \emph{understanding}.
+      Efim Zelmanov in~\cite{mathproof}.  
+          \end{quote}
+        \item O'Connor's proof was written at the end of \coq V7, rewritten at the beginning of \coq V8. Since then, \coq and its ecosystem evolved a lot (new styles, tactics, documentation tools).
+          We think this evolution should benefit
+          to the original proof-scripts and make their understanding easier.
+        \end{itemize}
+
+\section{Historical context}
+Russel O'Connor's work is the first computer verified proof 
+the essential incompleteness of arithmetic.
+
+\begin{todo}
+Reference to previous/later implementations.
+\end{todo}
 

### 2023-02-26 doc/hydras.tex
Implicits (#143)
* fix an inconsistent section numbering

* small correction in latex document

* small correction in latex document

* implicit arguments for impHn (to do: iffH, existH, andH, etc.)

* implicit arguments for notH (to do: iffH, existH, andH, etc.)

* implicit arguments for forallH (to do: iffH, existH, andH, etc.)

* implicit arguments for orH (to do: iffH, existH, andH, etc.)

* implicit arguments for andH (to do: iffH, existH, etc.)

* implicit arguments for iffH (to do:  existH, etc.)

* implicit arguments for equal (to do:  existH, etc.)

* implicit arguments for existH (to

--- 2022-12-05 doc/hydras.tex
@@
 \part{Ackermann, G\"{o}del, Peano\,\dots}
-This part contains comments, example  and exercises about Russel O'Connor's work on G\"{o}del's first incompleteness theorem.
-This work is maintained by Coq community~\cite{CoqCommunity} volunteers: (projects Goedel~\cite{Goedel} and Hydra-battles~\cite{HydraBattles}).
-The main reference to this work is  Russel O'Connor's article~\cite{OConnor05}, which we strongly encourage the reader to 
-consult regularly.
 
-Some changes are made to the aforementionned libraries, mainly because of the recent evolution of coq and its libraries.
-Nevertheless, the definitions, lemmas and theorems of the original contribution have been kept in this new release.
-
-\emph{\color{red!80}This maintenance job has just started, and will probably be long to complete. Help is welcome!}
-
+\include{chap-intro-goedel}
 \include{chapter-primrec} 

### 2023-02-28 doc/chapter-fol.tex
First experiments on Fol notations

@@
 
-\inputsnippets{folExample2/ToyDef}
+\inputsnippets{FolExamples/ToyDef}
 
@@
 
-Let us show a few examples (from ~\href{../theories/html/hydras.MoreAck.fol_Examples.html}{MoreAck.fol\_Examples}).
+Let us show a few examples (from ~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}).
 
-\inputsnippets{fol_Examples/arityTest}
+\inputsnippets{FolExamples/arityTest}
 
@@
 
-\inputsnippets{folExample2/toyNotation}
+\inputsnippets{FolExamples/toyNotation}
 
-\inputsnippets{folExample2/smallTerms}
+\inputsnippets{FolExamples/smallTerms}
 
@@
 
-\inputsnippets{fol_Examples/v1Plus0}
+\inputsnippets{FolExamples/v1Plus0}
 
@@
 
-For instance, the term $v_1+0$ is written as below.
-
-\inputsnippets{fol_Examples/v1Plus0}
-
 Let us give a few examples of first-order formulas.
@@
 \end{enumerate}
-\inputsnippets{fol_Examples/f1Example}
+\inputsnippets{FolExamples/f1Example}
 
 \begin{remark}[bound variables]
-\inputsnippets{folExample2/boundVars}  
+\inputsnippets{FolExamples/boundVars}  
 \end{remark}
@@
 
-\inputsnippets{fol_Examples/FormulaRect}
+\inputsnippets{FolExamples/FormulaRect}
 
@@
 
-\inputsnippets{fol/depthDef,  fol_Examples/DepthCompute}
+\inputsnippets{fol/depthDef,  FolExamples/DepthCompute}
 
@@
 
-\inputsnippets{fol_Examples/ltDepth1}
+\inputsnippets{FolExamples/ltDepth1}
 
@@
 
-\inputsnippets{fol_Examples/depthRecDemo}
+\inputsnippets{FolExamples/depthRecDemo}
 
@@
 
-\inputsnippets{fol_Examples/depthRecDemo2}
+\inputsnippets{FolExamples/depthRecDemo2}
 
@@
 
-\inputsnippets{fol_Examples/freeVarExamples}
+\inputsnippets{FolExamples/freeVarExamples}
 
@@
 In order to get more readable terms and formulas, we can define a few notations in ~\href{../theories/html/hydras.MoreAck.FOL_notations.html}{MoreAck.FOL\_notations} and
-~\href{../theories/html/hydras.MoreAck.LNN_notations.html}{MoreAck.LNN\_notations}.
+~\href{../theories/html/hydras.MoreAck.LNN.html}{MoreAck.LNN}.
 Please note that these notation scopes are experimental: We are going to use them in examples and exercises before using them in large original proof scripts (in the \texttt{ordinals/Ackermann/} sub-directory).
@@
 Prove the following lemma:
-\inputsnippets{fol_Examples/eqRefl}
+\inputsnippets{FolExamples/eqRefl}
 
@@
 Prove the following lemma:
-\inputsnippets{fol_Examples/MPdiag}
+\inputsnippets{FolExamples/MPdiag}
 
@@
 For instance, let us prove Peirce's rule.
- (in  ~\href{../theories/html/hydras.MoreAck.folPeirce.html}{MoreAck.folPeirce} ).
+ (in  ~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples} ).
 
@@
 
-\inputsnippets{folPeirce/prelude}
+%\inputsnippets{folPeirce/prelude}
 
@@
 
-\inputsnippets{folPeirce/PeirceProof}
+\inputsnippets{FolExamples/PeirceProof}
 
@@
 
-\inputsnippets{folPeirce/step1}
+\inputsnippets{FolExamples/step1}
 
@@
 
-\inputsnippets{folPeirce/step2}
+\inputsnippets{FolExamples/step2}
 
@@
 
-\inputsnippets{folPeirce/step3, folPeirce/step4}
+\inputsnippets{FolExamples/step3, FolExamples/step4}
 

### 2023-04-27 doc/chapter-fol.tex
First experiments on Fol notations
Examples of new notations (doc / MoreAck)

@@
 A \emph{language} is a structure composed of relation and function symbols, each symbol is given an \emph{arity} (number of arguments).
+As suggested by Russel O'Connor in~\cite{OConnor05}, we consider two arity functions instead of a function defined on the sum type \texttt{Relations + Functions}.
 
@@
 In order to show a few simple examples of statements and proofs, we define a small language with very few symbols:
-two constant symbols: $a$ and $b$, a function symbol $f$ (of arity 1), three propositional symbols $A$, $B$ and $C$, a one-place predicate symbol $P$, and a binary relational symbol $R$.
+two constant symbols: $a$ and $b$, three function symbols $f$, $g$ and $h$  (of respective arity 1,1 and 2), three propositional symbols $A$, $B$ and $C$, a one-place predicate symbol $P$, and a binary relational symbol $R$.
+
+From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}.
 
@@
 
-\subsubsection{Languages for Arithmetic}
+\begin{remark}
+  The constructors of types \texttt{Rel} and \texttt{Fun} are suffixed by an underscore, in order to reserve the names \texttt{a}, \texttt{f}, \texttt{h}, \texttt{R}, etc.
+to functions which build terms and formulas
+(please look at next sections).
 
-In ~\href{../theories/html/hydras.Ackermann.Languages.html}{Ackermann.Languages}, we can find two examples of languages built with the usual symbols of arithmetic.
-
-The first language, \texttt{LNT} (\emph{Language of Number Theory}) has just function symbols for $+$, $\times$, $0$ and successor.
-
-The second language, LNN, is the language of NN (\emph{Natural Numbers}) and has
-the same function symbols as LNT plus one relation symbol for less than, LT.
-
-First, we declare two ranked alphabets. 
-\inputsnippets{Languages/LNTDef1}
-
-In a second time, we build \texttt{LNT} and \texttt{LNN} by filling \texttt{Language}'s \texttt{arity} field.
-
-\inputsnippets{Languages/LNTDef2}
-
-Let us show a few examples (from ~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}).
-
-\inputsnippets{FolExamples/arityTest}
+\end{remark}
 
@@
 
+\inputsnippets{FolExamples/TermExamples1}
+
 \begin{remark}[Variables]
@@
 \end{remark}
-\subsubsection{Examples}
 
-With our toy language, we can build the following terms, respectively written $a$, $f(a)$ and $f(f(v_1))$ in usual mathematical notation.
+\begin{remark}
+ We can use \coq's \texttt{Notation} features to print and parse terms in a more readable form.
 
-Writing these terms in bare \gallina syntax would result in quite clumsy scripts. Thus, we may use a still experimental notation scope which allows us to write/print terms and formulas in a 
-notation similar to \coq's.
+\inputsnippets{FolExamples/beforeAbbreviations, FolExamples/termAbbreviations, FolExamples/NowItsBetter}
 
-We put both the associated \coq{} terms, in bare \gallina{} syntax, and with a still experimental and provisional notation, defined in
-~\href{../theories/html/hydras.MoreAck.FOL_notations.html}{MoreAck.FOL\_notations}).
-
-\inputsnippets{FolExamples/toyNotation}
-
-\inputsnippets{FolExamples/smallTerms}
-
-For instance the term $v_1+0$, where $v_1$ is a variable,
-is represented by a term of type (\texttt{Term LNN}).
-
-\inputsnippets{FolExamples/v1Plus0}
+\end{remark}
 
@@
 
-Let us give a few examples of first-order formulas.
+Let us give a few examples of first-order formulas. on our toy language.
 
@@
 \begin{enumerate}
-\item $\forall\,v_0, v_0=0 \vee \exists\;v_1,\, v_0= S\;v_1$
-\item $\exists\,v_2,\, 0< v_2 \wedge 4=v_2+v_2$
-\item $ v_0=0 \vee \exists\;v_1,\, v_0= S\;v_1$
-\item $v_0= v_1 + v_1 \leftrightarrow v_0 = v_1 \times 2$
+\item $R\;a\;b$
+\item $\forall v_0 v_1, (R\;v_0\;v_1 \leftrightarrow R\;v_1\;v_0$
+\item $ \forall v_0, v_0=a \vee \exists\;v_1,\, v_0= f(v_1)$
 \end{enumerate}
-\inputsnippets{FolExamples/f1Example}
 
+\inputsnippets{FolExamples/FormExamples}
+
+If we define a few new notations, the notation scope \texttt{fol\_scope} allow us to parse and print formulas in a less clumsy way.
+
+From href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol}
+\inputsnippets{fol/folScope}
+
+\inputsnippets{FolExamples/toyNotationForm}
+
+\inputsnippets{FolExamples/toyNotationForm2}
 \begin{remark}[bound variables]
@@
 \end{remark}
+
+\subsubsection{Languages for Arithmetic}
+
+In ~\href{../theories/html/hydras.Ackermann.Languages.html}{Ackermann.Languages}, we can find two examples of languages built with the usual symbols of arithmetic.
+
+The first language, \texttt{LNT} (\emph{Language of Number Theory}) has just function symbols for $+$, $\times$, $0$ and successor.
+
+The second language, LNN, is the language of NN (\emph{Natural Numbers}) and has
+the same function symbols as LNT plus one relation symbol for less than, LT.
+
+First, we declare two ranked alphabets.
+
+\begin{remark}
+  We depart a little from \cite{Goedel}'s notations, where the 
+function and relation symbols are called \texttt{Plus}, 
+\texttt{Mult}, \texttt{LT}, etc. In our version, these type constructors are called \texttt{Plus\_}, 
+\texttt{Mult\_}, \texttt{LT\_}, etc., while the names without final underscores are bound to term building functions (\emph{e.g.}
+the function which takes two terms and builds the term representing their sum).
+\end{remark}
+
+\inputsnippets{Languages/LNTDef1}
+
+In a second time, we build \texttt{LNT} and \texttt{LNN} by filling \texttt{Language}'s \texttt{arity} field.
+
+\inputsnippets{Languages/LNTDef2}
+
+Let us show a few examples (from ~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}).
+
+\inputsnippets{FolExamples/arityTest}
+
+\begin{todo}
+Start with toy Language , then LNT and LNN.
+\end{todo}
+
+% \subsubsection{Examples}
+
+% With our toy language, we can build the following terms, respectively written $a$, $f(a)$ and $f(f(v_1))$ in usual mathematical notation.
+
+% Writing these terms in bare \gallina syntax would result in quite clumsy scripts. Thus, we may use a still experimental notation scope which allows us to write/print terms and formulas in a 
+% notation similar to \coq's.
+
+% We put both the associated \coq{} terms, in bare \gallina{} syntax, and with a still experimental and provisional notation, defined in
+% ~\href{../theories/html/hydras.Ackerm.FOL_notations.html}{MoreAck.FOL\_notations}).
+
+% \inputsnippets{FolExamples/toyNotation}
+
+% \inputsnippets{FolExamples/smallTerms}
+
+For instance the term $v_1+0$, where $v_1$ is a variable,
+is represented by the following \gallina term of type 
+(\texttt{fol.Term LNT}).
+
+\inputsnippets{FolExamples/v1Plus0}
 
@@
 
+\begin{todo}
+  Present the general issue about evaluation, and our provisional solution.
+\end{todo}
+
 \inputsnippets{LNN_Examples/uglyF0a}

### 2023-04-28 doc/chapter-fol.tex
Examples of new notations (doc / MoreAck)

@@
 
-If we define a few new notations, the notation scope \texttt{fol\_scope} allow us to parse and print formulas in a less clumsy way.
+In order to make the first order formulas more readable, we use a generic notation scope called \texttt{fol\_scope}, which uses mainly \coq's 
+notations for connectives and quantifiers. The \texttt{\%fol} delimiter 
+allows the user to distinguisf \texttt{FOL} connectives from their \coq equivalent.
 
@@
 
+In addition to \texttt{fol\_scope}, we introduce a few abbreviations specific to our toy language.
+
 \inputsnippets{FolExamples/toyNotationForm}
 
+The following interactions show how the formulas on $L$ may be 
+parsed or printed.
+
 \inputsnippets{FolExamples/toyNotationForm2}
+
 \begin{remark}[bound variables]

### 2023-05-08 doc/chap-intro-goedel.tex

@@
 \section{Introduction}
+
+It was the first computer verified proof 
+the essential incompleteness of arithmetic.
+The main reference to this work is  Russel O'Connor's article~\cite{OConnor05}, which we strongly encourage the reader to 
+consult regularly.
+
+Several reasons --- enumerated below --- led us to maintain and document this work in the framework of 
+Coq community~\cite{CoqCommunity}.
+
+ \begin{itemize}
+ \item Historical interest on G\"{o}del's proof and its mechanizations, as shown by the numerous litterature on this topic
+   (for instance \cite{smullyan1992godel, Hofstadter1999Godel, GoedelCassou}).
+
+  \item O'Connor's proof was written at the end of \coq V7, then rewritten at the beginning of \coq V8. Since then, \coq and its ecosystem evolved a lot (new styles, tactics, documentation tools).
+          We think this evolution should benefit
+          to the original proof-scripts and make their understanding easier.
+
+        \item Finally, let us quote  Efim Zelmanov in~\cite{mathproof}
+             \begin{quote}
+      The purpose of a proof is \emph{understanding}.
+
+    \end{quote}
+    We hope that underlying some points of the proof will
+     make easier to understand this large and technical work.
+
+        \end{itemize}
+
+        For technical reasons --- mainly in order to make easier the installation and use of its sub-libraries, we split the project
+into two main parts: Goedel~\cite{Goedel} and Hydra-battles~\cite{HydraBattles}.
+
+Some changes are made to the aforementionned libraries, mainly because of the recent evolution of coq and its libraries.
+Nevertheless, the definitions, lemmas and theorems of the original contribution have been kept in this new release.
+
+\emph{\color{red!80}These maintenance and documentation 
+jobs have just started, and will probably be long to complete. Help is welcome!}
 
@@
 consult regularly.
-
-% ici
-
-\section{The reason behind this work}
-
-  \begin{itemize}
-  \item Historical interest on G\"{o}del's proof
-  \item Many popularization publications, \emph{e.g.}
-    \cite{smullyan1992godel, Hofstadter1999Godel, GoedelCassou}, vs. technical complexity of the full proof .
-
-    \begin{quote}
-      The purpose of a proof is \emph{understanding}.
-      Efim Zelmanov in~\cite{mathproof}.  
-          \end{quote}
-        \item O'Connor's proof was written at the end of \coq V7, rewritten at the beginning of \coq V8. Since then, \coq and its ecosystem evolved a lot (new styles, tactics, documentation tools).
-          We think this evolution should benefit
-          to the original proof-scripts and make their understanding easier.
-        \end{itemize}
-
-\section{Historical context}
-Russel O'Connor's work is the first computer verified proof 
-the essential incompleteness of arithmetic.
-
-\begin{todo}
-Reference to previous/later implementations.
-\end{todo}
-
-Some changes are made to the aforementionned libraries, mainly because of the recent evolution of coq and its libraries.
-Nevertheless, the definitions, lemmas and theorems of the original contribution have been kept in this new release.
-
-\emph{\color{red!80}These maintenance and documentation 
-jobs have just started, and will probably be long to complete. Help is welcome!}
 

### 2023-05-08 doc/schutte-chapter.tex
missing space

@@
 \begin{itemize}
-\item It allows us to give a \emph{name} (using \texttt{Definition}) two witnesses 
+\item It allows us to give a \emph{name} (using \texttt{Definition}) to witnesses 
 of existential quantifiers (let us recall that, in classical logic, one may consider non-constructive proofs of existential statements)
@@
 Let us call  $\phi_0$ the ordering function of \texttt{AP}.
-In the mathematical text, we shall use indifferently the notations  $\omega^\alpha$ and$\phi_0(\alpha)$.
+In the mathematical text, we shall use indifferently the notations  $\omega^\alpha$ and $\phi_0(\alpha)$.
 

### 2023-05-09 doc/chapter-fol.tex
fix a few typos

@@
   The constructors of types \texttt{Rel} and \texttt{Fun} are suffixed by an underscore, in order to reserve the names \texttt{a}, \texttt{f}, \texttt{h}, \texttt{R}, etc.
-to functions which build terms and formulas
-(please look at next sections).
-
+to functions which build terms and formulas (please look at Sect~\ref{sec:fol-term-notations} and \ref{sec:fol-atomic-notations}).
 \end{remark}
@@
  a language and $i$ a natural number. 
-Although other choices may be considered : PHOAS, de Bruijn indices, etc,  we   respect this representation, in order not to break long proof scipts which use this representation (please look at \cite{OConnor05} for a related discussion).
+Although other choices may be considered : PHOAS, de Bruijn indices, etc,  we   respect this representation, in order not to break long proof scripts which use this representation (please look at \cite{OConnor05} for a related discussion).
 
@@
 
-\begin{remark}
+\subsubsection{Simplified notations for terms}
+\label{sec:fol-term-notations}
+
  We can use \coq's \texttt{Notation} features to print and parse terms in a more readable form.
@@
 \inputsnippets{FolExamples/beforeAbbreviations, FolExamples/termAbbreviations, FolExamples/NowItsBetter}
-
-\end{remark}
 
@@
 
+\subsection{Notations for first-order formulas}
+
 In order to make the first order formulas more readable, we use a generic notation scope called \texttt{fol\_scope}, which uses mainly \coq's 
@@
 
-From href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol}
+\vspace{6pt}
+
+\noindent From \href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol}
+
 \inputsnippets{fol/folScope}
+
+\subsubsection{Simplified notations for atomic formulas}
+\label{sec:fol-atomic-notations}
 
@@
 
+\noindent From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}
+
+\vspace{6pt}
+
 \inputsnippets{FolExamples/toyNotationForm2}
@@
 \begin{remark}[bound variables]
+
+  In~\cite{Goedel}, there is no De Bruijn encoding of bound variables (see also~\cite{OConnor05}).
+The following example shows two formulas which share the same structure, are logically equivalent, but are not Leibniz equal.
+
+\vspace{4pt}
+
+\noindent From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}
+
 \inputsnippets{FolExamples/boundVars}  

### 2023-05-09 doc/chapter-fol.tex
fix a few typos

@@
 \begin{remark}
+  \label{rem:underscores}
   The constructors of types \texttt{Rel} and \texttt{Fun} are suffixed by an underscore, in order to reserve the names \texttt{a}, \texttt{f}, \texttt{h}, \texttt{R}, etc.
@@
 \inputsnippets{fol/TermDef}
+
+\subsubsection{Examples}
+
+Let us give the \gallina terms of opur toy language, corresponding to the terms $a$,
+$f(a)$, and $h(f(a),a)$. The ``term'' $h(a)$ is not well formed, since the arity of $h$ is not respected.
 
@@
 The second language, LNN, is the language of NN (\emph{Natural Numbers}) and has
-the same function symbols as LNT plus one relation symbol for less than, LT.
+the same function symbols as LNT plus one relation symbol for the strict inequality $<$ : \texttt{LT} (less than).
 
@@
 \texttt{Mult\_}, \texttt{LT\_}, etc., while the names without final underscores are bound to term building functions (\emph{e.g.}
-the function which takes two terms and builds the term representing their sum).
+the function which takes two terms and builds the term representing their sum) (see Remark~\ref{rem:underscores}).
 \end{remark}
@@
 \inputsnippets{FolExamples/v1Plus0}
-
-\begin{todo}
-Comment on the use of boths list and sets of formulas.
-(explained in Russel's paper, p.7).
-
-\inputsnippets{fol/SystemDef}
-
-\end{todo}
 

### 2023-05-12 doc/chapter-fol.tex
work on chapter 12

@@
 
+This chapter is devoted to the presentation to  data structures for representing first order formulas over a ranked alphabet, and the basic functions and predicates over these types.
+Although all the following constructions come directly from~\cite{Goedel}, we introduced minor changes (mainly syntax) to
+take into account recent changes in \coq(new constructions, tactics, notations, etc.).
+
 \section{Data types}
@@
 
-A \emph{language} is a structure composed of relation and function symbols, each symbol is given an \emph{arity} (number of arguments).
-As suggested by Russel O'Connor in~\cite{OConnor05}, we consider two arity functions instead of a function defined on the sum type \texttt{Relations + Functions}.
+A \emph{language} is a structure composed of relation and function symbols, each symbol is given an \emph{arity} (number of arguments)\,\footnote{As suggested by Russel O'Connor in~\cite{OConnor05}, we consider two arity functions instead of a function defined on the sum type \texttt{Relations + Functions}}.
 
@@
 
-\subsubsection{A toy language}
+\subsubsection{Example: a toy language}
 In order to show a few simple examples of statements and proofs, we define a small language with very few symbols:
-two constant symbols: $a$ and $b$, three function symbols $f$, $g$ and $h$  (of respective arity 1,1 and 2), three propositional symbols $A$, $B$ and $C$, a one-place predicate symbol $P$, and a binary relational symbol $R$.
+two constant symbols: $a$ and $b$, three function symbols $f$, $g$ and $h$  (of respective arity 1, 1 and 2), three propositional symbols $A$, $B$ and $C$, a one-place predicate symbol $P$, and a binary relational symbol $R$.
 
@@
 
-Given a language $L$, we define the set of \emph{terms} and
-$n$-\emph{tuples} of terms.
+Given a language $L$, we define the type of \emph{terms} and
+$n$-\emph{tuples of terms} over $L$.
 
@@
 
-\subsubsection{Examples}
-
-Let us give the \gallina terms of opur toy language, corresponding to the terms $a$,
-$f(a)$, and $h(f(a),a)$. The ``term'' $h(a)$ is not well formed, since the arity of $h$ is not respected.
-
-\inputsnippets{FolExamples/TermExamples1}
-
 \begin{remark}[Variables]
-In O'Connor's formalization of first-order logic, the variables are \gallina terms of the form \texttt{var $L$ $i$}, where $L$ is
+In O'Connor's formalization of first-order logic,  variables are \gallina terms of the form (\texttt{var $L$ $i$}), where $L$ is
  a language and $i$ a natural number. 
@@
 
-\subsubsection{Simplified notations for terms}
-\label{sec:fol-term-notations}
+\begin{remark}
+This representation of terms uses mutually inductive data-types instead of list of terms. Please see also Remark~\ref{hydra:mutually-inductive-vs-lists}.
+\end{remark}
 
- We can use \coq's \texttt{Notation} features to print and parse terms in a more readable form.
+\subsubsection{Examples}
+\label{sect:folTermExamples}
 
-\inputsnippets{FolExamples/beforeAbbreviations, FolExamples/termAbbreviations, FolExamples/NowItsBetter}
+Let us define some \gallina terms over our toy language, respecively corresponding to the terms $a$,
+$f(a)$, and $h(f(a),a)$. The fourth ``term'' is not well formed, since the arity of $h$ is not respected\,\footnote{Strictly speaking, it's not a (well typed) term!}.
+
+\inputsnippets{FolExamples/TermExamples1}
 
@@
 
-The full abstract syntax of first-order formulas is completed 
-with a few more notations. Note that we are studying \emph{classical} first-order logic, so the existential quantifier is derived from the universal one through negation.
+The  abstract syntax of first-order formulas is completed 
+with a few abbreviations\,\footnote{Technically, these abbreviations take the form of \emph{global definitions}}.
+Note that we are studying \emph{classical} first-order logic, for instance the existential quantifier is derived from the universal one through negation.
 
@@
 \subsection{Examples}
+\label{sect:folFormExamples}
 
-Let us give a few examples of first-order formulas. on our toy language.
+Let us give a few examples of first-order formulas over $L$.
 
@@
 \item $R\;a\;b$
-\item $\forall v_0 v_1, (R\;v_0\;v_1 \leftrightarrow R\;v_1\;v_0$
+\item $\forall v_0\; v_1, R\;v_0\;v_1 \leftrightarrow R\;v_1\;v_0$
 \item $ \forall v_0, v_0=a \vee \exists\;v_1,\, v_0= f(v_1)$
@@
 
-\subsection{Notations for first-order formulas}
+\section{Simplified notations for terms and formulas}
+\label{sect:fol-notations}
 
-In order to make the first order formulas more readable, we use a generic notation scope called \texttt{fol\_scope}, which uses mainly \coq's 
+ We use \coq's \texttt{Notation} features to print and parse terms and formulas  in a more readable form.
+
+ \subsection{Parsing and printing terms}
+ \label{sec:fol-term-notations}
+
+Let us consider the term $t_2$ of Sect\ref{sect:folTermExamples}~. Its normal form is quite hard to read:
+
+\inputsnippets{FolExamples/beforeAbbreviations}
+
+We may use \coq's abbreviation feature  to print and parse $L$-terms with a more natural concrete syntax.
+
+\inputsnippets{FolExamples/termAbbreviations, FolExamples/NowItsBetter}
+
+\subsection{Parsing and printing first-order formulas}
+
+In order to make the first order formulas more readable, we define  a  notation scope called \texttt{fol\_scope}, which contains 
 notations for connectives and quantifiers. The \texttt{\%fol} delimiter 
-allows the user to distinguisf \texttt{FOL} connectives from their \coq equivalent.
+allows the user to distinguish \texttt{FOL} connectives from their \coq equivalent.
 
@@
 
-\inputsnippets{fol/folScope}
+\inputsnippets{fol/folScope1}
+
+The following notations are used (when printing) when  some computation 
+    expands a disjunction, conjuction, etc. 
+    in terms of implication and negation.
+
+\inputsnippets{fol/folScope2}
 
@@
 
-In addition to \texttt{fol\_scope}, we introduce a few abbreviations specific to our toy language.
+In addition to \texttt{fol\_scope}, we can introduce a few abbreviations specific to our toy language.
 
@@
 
-The following interactions show how the formulas on $L$ may be 
+The following interactions show how the formulas 
+of~\ref{sect:folFormExamples} may be 
 parsed or printed.
@@
 
-\begin{remark}[bound variables]
+\section{Computing and reasoning on first-order formulas}
+
+\subsection{Bound variables}
 
@@
 \inputsnippets{FolExamples/boundVars}
-\end{remark}
-
-\subsubsection{Languages for Arithmetic}
-
-In ~\href{../theories/html/hydras.Ackermann.Languages.html}{Ackermann.Languages}, we can find two examples of languages built with the usual symbols of arithmetic.
-
-The first language, \texttt{LNT} (\emph{Language of Number Theory}) has just function symbols for $+$, $\times$, $0$ and successor.
-
-The second language, LNN, is the language of NN (\emph{Natural Numbers}) and has
-the same function symbols as LNT plus one relation symbol for the strict inequality $<$ : \texttt{LT} (less than).
-
-First, we declare two ranked alphabets.
-
-\begin{remark}
-  We depart a little from \cite{Goedel}'s notations, where the 
-function and relation symbols are called \texttt{Plus}, 
-\texttt{Mult}, \texttt{LT}, etc. In our version, these type constructors are called \texttt{Plus\_}, 
-\texttt{Mult\_}, \texttt{LT\_}, etc., while the names without final underscores are bound to term building functions (\emph{e.g.}
-the function which takes two terms and builds the term representing their sum) (see Remark~\ref{rem:underscores}).
-\end{remark}
-
-\inputsnippets{Languages/LNTDef1}
-
-In a second time, we build \texttt{LNT} and \texttt{LNN} by filling \texttt{Language}'s \texttt{arity} field.
-
-\inputsnippets{Languages/LNTDef2}
-
-Let us show a few examples (from ~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}).
-
-\inputsnippets{FolExamples/arityTest}
 
   \begin{todo}
-Start with toy Language , then LNT and LNN.
+   Link to the lemmas which attest the equivalence of these formulas (properties of substitution, logical equivalence).
   \end{todo}
-
-% \subsubsection{Examples}
-
-% With our toy language, we can build the following terms, respectively written $a$, $f(a)$ and $f(f(v_1))$ in usual mathematical notation.
-
-% Writing these terms in bare \gallina syntax would result in quite clumsy scripts. Thus, we may use a still experimental notation scope which allows us to write/print terms and formulas in a 
-% notation similar to \coq's.
-
-% We put both the associated \coq{} terms, in bare \gallina{} syntax, and with a still experimental and provisional notation, defined in
-% ~\href{../theories/html/hydras.Ackerm.FOL_notations.html}{MoreAck.FOL\_notations}).
-
-% \inputsnippets{FolExamples/toyNotation}
-
-% \inputsnippets{FolExamples/smallTerms}
-
-For instance the term $v_1+0$, where $v_1$ is a variable,
-is represented by the following \gallina term of type 
-(\texttt{fol.Term LNT}).
-
-\inputsnippets{FolExamples/v1Plus0}
 
@@
 \inputsnippets{FolExamples/freeVarExamples}
+
+\section{Languages for Arithmetic}
+
+Two languages built with the usual symbols of arithmetic are 
+defined in ~\href{../theories/html/hydras.Ackermann.Languages.html}{Ackermann.Languages}.
+
+\begin{itemize}
+\item The first language: \texttt{LNT} (\emph{Language of Number Theory}) has just function symbols for $+$, $\times$, $0$ and successor.
+\item The second language: LNN (\emph{Language of Natural Numbers})  has
+the same function symbols as LNT plus one relation symbol for the strict inequality $<$ : \texttt{LT} (less than).
+\end{itemize}
+
+\subsubsection{Language of Number Theory (LNT)}
+
+First, we declare two alphabets.
+
+\inputsnippets{Languages/LNTDef1}
+
+In a second time, we build \texttt{LNT} and \texttt{LNN} by filling \texttt{Language}'s \texttt{arity} field.
+
+\inputsnippets{Languages/LNTDef2}
+
+\begin{remark}
+  We depart a little from \cite{Goedel}'s notations, where the 
+function and relation symbols are called \texttt{Plus}, 
+\texttt{Mult}, \texttt{LT}, etc. In our version, these type constructors are called \texttt{Plus\_}, 
+\texttt{Mult\_}, \texttt{LT\_}, etc., while the names without final underscores are bound to term building functions (\emph{e.g.}
+the function which takes two terms and builds the term representing their sum) (see Remark~\ref{rem:underscores}).
+\end{remark}
+
+\subsubsection{Language of Natural Numbers (LNN)}
+
+\texttt{LNN} is an extension of \texttt{LNT}, by the addition 
+of the $<$ relation symbol.
+
+\inputsnippets{Languages/LNNDef}
+
+\subsubsection{Examples}
+
+Let us show a few examples (from ~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}).
+
+\inputsnippets{FolExamples/arityTest}
+
+% \subsubsection{Examples}
+
+% With our toy language, we can build the following terms, respectively written $a$, $f(a)$ and $f(f(v_1))$ in usual mathematical notation.
+
+% Writing these terms in bare \gallina syntax would result in quite clumsy scripts. Thus, we may use a still experimental notation scope which allows us to write/print terms and formulas in a 
+% notation similar to \coq's.
+
+% We put both the associated \coq{} terms, in bare \gallina{} syntax, and with a still experimental and provisional notation, defined in
+% ~\href{../theories/html/hydras.Ackerm.FOL_notations.html}{MoreAck.FOL\_notations}).
+
+% \inputsnippets{FolExamples/toyNotation}
+
+% \inputsnippets{FolExamples/smallTerms}
+
+For instance the term $v_1+0$, where $v_1$ is a variable,
+is represented by the following \gallina term of type 
+(\texttt{fol.Term LNT}).
+
+\inputsnippets{FolExamples/v1Plus0}
 

### 2023-05-21 doc/chapter-fol.tex
change notations of fol-variables

@@
 
+\subsubsection{Bound variables}
+
+  In~\cite{Goedel}, there is no De Bruijn encoding of bound variables (see also~\cite{OConnor05}).
+The following example shows two formulas which share the same structure, are logically equivalent, but are not Leibniz equal.
+
+\vspace{4pt}
+
+\noindent From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}
+
+\inputsnippets{FolExamples/boundVars}
+
+  \begin{todo}
+   Link to the lemmas which attest the equivalence of these formulas (properties of substitution, logical equivalence).
+  \end{todo}
+
+  \index{ackermann}{Projects}
+  \begin{project}
+ Use a HOAS representation for FOL terms and formulas (without breaking proof scripts).   
+  \end{project}
+
 \subsubsection{Examples}
@@
 \subsection{Examples}
-\label{sect:folFormExamples}
 
@@
 \end{enumerate}
+
+\label{sect:folFormExamples}
 
@@
 
-\subsection{Bound variables}
-
-  In~\cite{Goedel}, there is no De Bruijn encoding of bound variables (see also~\cite{OConnor05}).
-The following example shows two formulas which share the same structure, are logically equivalent, but are not Leibniz equal.
-
-\vspace{4pt}
-
-\noindent From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}
-
-\inputsnippets{FolExamples/boundVars}
+\subsection{On variables}
 
 \begin{todo}
-   Link to the lemmas which attest the equivalence of these formulas (properties of substitution, logical equivalence).
+  Difference from HOAS. Free and bound occurrences.
 \end{todo}
@@
 \subsection{Decidability of equality}
+Let $L$ be a language, and let us assume that equality 
+of function and relation symbols of $L$ are decidable.
+Under this assumption, equality of terms and formulas over $L$ is decidable too.
 
-Under the assumption that equality is decidable on 
-function and relation symbols of any language $L$,
-we get decidability of equality on terms and formulas.
 Because of dependent types, the proofs are quite long and technical. The reader may consult them in \href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol}
@@
 \subsection{Induction principles on formula depth}
+
+\inputsnippets{FolExamples/substNotStructural}
 
@@
 
+\begin{todo}
+Link to folProp.v
+\end{todo}
+
 Section \texttt{Free\_Variables} 
@@
 \inputsnippets{FolExamples/freeVarExamples}
+
+\subsection{Multiple substitutions}
+
+\begin{todo}
+Link to subAll.v
+\end{todo}
 
@@
 
+\begin{todo}
+Fix the issue about the way formulas are printed (mix of different scopes).
+\end{todo}
 \inputsnippets{LNN_Examples/uglyF0a}

### 2023-05-23 doc/chapter-fol.tex
redaction of chapter 12 (in progress)

@@
 
-This chapter is devoted to the presentation to  data structures for representing first order formulas over a ranked alphabet, and the basic functions and predicates over these types.
+This chapter is devoted to the presentation to  data structures for representing first order formulas over a ranked alphabet, and the basic functions and predicates over these types:
+
+\begin{itemize}
+\item Abstract syntax of terms and formulas over a ranked alphabet composed of function and relation symbols.
+\item Induction principles over terms and formulas.
+\item Definition and main properties of substitution of terms to variables.
+\end{itemize}
+
 Although all the following constructions come directly from~\cite{Goedel}, we introduced minor changes (mainly syntax) to
@@
 
-A \emph{language} is a structure composed of relation and function symbols, each symbol is given an \emph{arity} (number of arguments)\,\footnote{As suggested by Russel O'Connor in~\cite{OConnor05}, we consider two arity functions instead of a function defined on the sum type \texttt{Relations + Functions}}.
+A \emph{language} is a structure composed of relation and function symbols, each symbol is given an \emph{arity} (number of arguments)\,\footnote{As suggested by Russel O'Connor in~\cite{OConnor05}, we consider two arity functions instead of a single function defined on the sum type \texttt{Relations + Functions}.}.
 
@@
 
+\begin{remark}
+This representation of terms uses mutually inductive data-types instead of list of terms. Please see also Remark~\ref{hydra:mutually-inductive-vs-lists}.
+\end{remark}
+
 \begin{remark}[Variables]
-In O'Connor's formalization of first-order logic,  variables are \gallina terms of the form (\texttt{var $L$ $i$}), where $L$ is
- a language and $i$ a natural number. 
-Although other choices may be considered : PHOAS, de Bruijn indices, etc,  we   respect this representation, in order not to break long proof scripts which use this representation (please look at \cite{OConnor05} for a related discussion).
+In O'Connor's formalization of first-order logic,  variables are 
+just natural numbers, and the \emph{coercion} from
+\texttt{nat} to \texttt{Term L} is the constructor (\texttt{@var $L$}).
+Although other choices may be considered : PHOAS, de Bruijn indices, etc,  we still the data structures of~\cite{Goedel}, in order not to break long proof scripts which use this representation (please look at \cite{OConnor05} for a related discussion).
 
@@
 
+\subsubsection{Examples}
+\label{sect:folTermExamples}
+
+Let us define some \gallina terms over our toy language, respecively corresponding to the terms $a$,
+$f(a)$, and $h(f(a),a)$.
+
+First, in order to make terms on $L$ more readable, we introduce a few abbreviations.
+
+From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}.
+
+\inputsnippets{FolExamples/termAbbreviations}
+
+\inputsnippets{FolExamples/TermExamples1}
+
 \begin{remark}
-This representation of terms uses mutually inductive data-types instead of list of terms. Please see also Remark~\ref{hydra:mutually-inductive-vs-lists}.
+The fourth ``term'' \texttt{t4} is not well formed, since the arity of $h$ is not respected\,\footnote{Strictly speaking, it's not a (well typed) term!}.
 \end{remark}
+
+\subsubsection{Other Languages}
+
+\begin{todo}
+Link to the chapter which presents \texttt{LNT} and \texttt{LNN}.
+\end{todo}
+
+\subsection{First-order formulas}
+
+The type of first order formulas over $L$ is defined 
+in~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} as an inductive data type.
+
+\inputsnippets{fol/FormulaDef}
+
+\begin{remark}
+The type \texttt{Formula} (over $L$) has only five constructors,
+respectively bound to term equalities $t_1=t_2$, atomic formulas of the form $(r\;t_1\;\dots\;t_n)$, implications $(F \arrow G)$, negations $({\sim}F)$, and universal quantifications $(\forall\,i, F)$.
+
+In~\cite{Goedel}, no \emph{constructors} are associated with
+disjunction, conjonction, logical equivalence and existential quantifier.
+These constructs are formalized through \emph{definitions}\footnote{Please keep in mind that we are considering a classical logic.}.
+
+\inputsnippets{fol/FolFull, fol/folPlus}
+
+This convention allows the user to reduce to 5 (instead of 10) the number of cases in ''\texttt{match F with \dots}'' constructs.
+On the other hand, some computation may expand a connective like $\vee$ or
+$\wedge$, or an existential quantification into a ``basic'' formula (see Sect.\vref{sect:fol-issue}).
+\end{remark}
+
+\subsection{Examples}
+
+Let us give a few examples of first-order formulas over $L$.
+\label{fol:examplesf1f2f3}
+\begin{enumerate}
+\item $R\;a\;b$
+\item $\forall v_0\; v_1, R\;v_0\;v_1 \arrow R\;v_1\;v_0$
+\item $ \forall v_0, v_0=a \vee \exists\;v_1,\, v_0= f(v_1)$
+\item $(\forall v_1, v_0 = v_1) \vee \exists\,v_0\,v_1, v_0 \not= v_1$
+\end{enumerate}
+
+First, we define abbreviations for atomic formulas over $L$.
+
+\noindent From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}
+
+\inputsnippets{FolExamples/toyNotationForm}
+
+\label{sec:fol-atomic-notations}
+\label{sect:folFormExamples}
+
+\inputsnippets{FolExamples/FormExamples}
+
+% The following interactions show how the formulas 
+% of~\ref{sect:folFormExamples} may be 
+% parsed or printed.
+
+% \noindent From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}
+
+% \vspace{6pt}
+
+% \inputsnippets{FolExamples/toyNotationForm2}
 
@@
 
+In the formula \texttt{F4} \vpageref[above]{sect:folFormExamples}, the  term \texttt{(var 0)} occurs freely and also inside the scope of a quantifier.
+
   \begin{todo}
@@
 
-\subsubsection{Examples}
-\label{sect:folTermExamples}
-
-Let us define some \gallina terms over our toy language, respecively corresponding to the terms $a$,
-$f(a)$, and $h(f(a),a)$. The fourth ``term'' is not well formed, since the arity of $h$ is not respected\,\footnote{Strictly speaking, it's not a (well typed) term!}.
-
-\inputsnippets{FolExamples/TermExamples1}
-
-\subsection{Formulas}
-
-The type of first order formulas over $L$ is defined 
-in~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} as an inductive data type.
-
-\inputsnippets{fol/FormulaDef}
-
-The  abstract syntax of first-order formulas is completed 
-with a few abbreviations\,\footnote{Technically, these abbreviations take the form of \emph{global definitions}}.
-Note that we are studying \emph{classical} first-order logic, for instance the existential quantifier is derived from the universal one through negation.
-
-\inputsnippets{fol/FolFull, fol/folPlus}
-
-\subsection{Examples}
-
-Let us give a few examples of first-order formulas over $L$.
-
-\label{fol:examplesf1f2f3}
-\begin{enumerate}
-\item $R\;a\;b$
-\item $\forall v_0\; v_1, R\;v_0\;v_1 \leftrightarrow R\;v_1\;v_0$
-\item $ \forall v_0, v_0=a \vee \exists\;v_1,\, v_0= f(v_1)$
-\end{enumerate}
-
-\label{sect:folFormExamples}
-
-\inputsnippets{FolExamples/FormExamples}
-
-\section{Simplified notations for terms and formulas}
+\section{A notation scope for first-order terms and formulas}
 \label{sect:fol-notations}
@@
  We use \coq's \texttt{Notation} features to print and parse terms and formulas  in a more readable form.
+To this purpose, we build \texttt{fol\_scope}, a  notation scope
+where the main connectives and quantifiers get a syntax close to \coq's.  Additionnally, a term of the form \texttt{(var $i$)} is
+just written \texttt{v\#$i$}.
 
- \subsection{Parsing and printing terms}
- \label{sec:fol-term-notations}
+\inputsnippets{fol/folScope1}
 
-Let us consider the term $t_2$ of Sect\ref{sect:folTermExamples}~. Its normal form is quite hard to read:
-
-\inputsnippets{FolExamples/beforeAbbreviations}
-
-We may use \coq's abbreviation feature  to print and parse $L$-terms with a more natural concrete syntax.
-
-\inputsnippets{FolExamples/termAbbreviations, FolExamples/NowItsBetter}
-
-\subsection{Parsing and printing first-order formulas}
-
-In order to make the first order formulas more readable, we define  a  notation scope called \texttt{fol\_scope}, which contains 
-notations for connectives and quantifiers. The \texttt{\%fol} delimiter 
+The \texttt{\%fol} delimiter 
 allows the user to distinguish \texttt{FOL} connectives from their \coq equivalent.
+\emph{We discourage the reader from \emph{opening} \texttt{fol\_scope} and similar scopes : \texttt{nn\_scope}, \texttt{nt\_scope}, which would make disappear the \texttt{\%fol} suffix from the first-order formulas}.
 
@@
 
-\inputsnippets{fol/folScope1}
+\inputsnippets{FolExamples/toyNotationForm2}
 
-The following notations are used (when printing) when  some computation 
-    expands a disjunction, conjuction, etc. 
-    in terms of implication and negation.
+\subsection{The issue with derived constructions}
+\label{sect:fol-issue}
+
+The connectives and quantifiers $\vee$, $\wedge$, $\exists$, etc. may raise an issue when printing computed formulas.
+For instance, a formula like $F \wedge G$ could be transformed into $\sim(\sim F \vee \sim G)$,  and even
+into $\sim(\sim\sim F \arrow \sim B)$, which would 
+cause serious problems of readability.
+
+In such a case, we propose to print such a formula as $F \wedge' G$, to make it syntactically distinct but very similar to $F \wedge G$.
 
@@
 
-\subsubsection{Simplified notations for atomic formulas}
-\label{sec:fol-atomic-notations}
+The following examples show how the primed connectors 
+and quantifiers behave with respect to convertibility and 
+input/output.
 
-In addition to \texttt{fol\_scope}, we can introduce a few abbreviations specific to our toy language.
-
-\inputsnippets{FolExamples/toyNotationForm}
-
-The following interactions show how the formulas 
-of~\ref{sect:folFormExamples} may be 
-parsed or printed.
-
-\noindent From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}
-
-\vspace{6pt}
-
-\inputsnippets{FolExamples/toyNotationForm2}
+\inputsnippets{FolExamples/toyNotationForm3}
 

### 2023-06-05 doc/chapter-fol.tex
Remove Export FolNotations
Removed  useless Exports
redaction of chapter 12 (in progress)

@@
 
+Structural induction/recursion principles are generated by 
+\texttt{Formula}'s definition, for instance:
+
+\inputsnippets{FolExamples/FormulaRect}
+
 \begin{remark}
 The type \texttt{Formula} (over $L$) has only five constructors,
-respectively bound to term equalities $t_1=t_2$, atomic formulas of the form $(r\;t_1\;\dots\;t_n)$, implications $(F \arrow G)$, negations $({\sim}F)$, and universal quantifications $(\forall\,i, F)$.
+respectively associated with term equalities $t_1=t_2$, atomic formulas of the form $(r\;t_1\;\dots\;t_n)$, implications $(F \arrow G)$, negations $({\sim}F)$, and universal quantifications $(\forall\,i, F)$.
 
-In~\cite{Goedel}, no \emph{constructors} are associated with
+In~\cite{Goedel}, no \emph{constructors} of type \texttt{(Formula L)} are associated with
 disjunction, conjonction, logical equivalence and existential quantifier.
@@
 \label{fol:examplesf1f2f3}
-\begin{enumerate}
-\item $R\;a\;b$
-\item $\forall v_0\; v_1, R\;v_0\;v_1 \arrow R\;v_1\;v_0$
-\item $ \forall v_0, v_0=a \vee \exists\;v_1,\, v_0= f(v_1)$
-\item $(\forall v_1, v_0 = v_1) \vee \exists\,v_0\,v_1, v_0 \not= v_1$
-\end{enumerate}
+\begin{description}
+\item[F1] $R\;a\;b$
+\item[F2] $\forall v_0\; v_1, R\;v_0\;v_1 \arrow R\;v_1\;v_0$
+\item[F3] $ \forall v_0, v_0=a \vee \exists\;v_1,\, v_0= f(v_1)$
+\item[F4] $(\forall v_1, v_0 = v_1) \vee \exists\,v_0\,v_1, v_0 \not= v_1$
+\item[F5] $v_0 =a \vee v_0 = f(v_1)$
+ \item[F6] $\forall v_0,\,\exists v_1, v_0= f(v_1)\wedge v_0\not= v_1$
+\end{description}
+
+Let us now define these formulas as terms of type
+\texttt{(Formula L)}.
 
@@
 In~\cite{Goedel}, there is no De Bruijn encoding of bound variables (see also~\cite{OConnor05}).
+
+For instance, the the  term \texttt{(var 0)}
+occurs freely and also inside the scope of a quantifier in the
+formula \texttt{F4} \vpageref[above]{sect:folFormExamples}.
+
 The following example shows two formulas which share the same structure, are logically equivalent, but are not Leibniz equal.
@@
 
-In the formula \texttt{F4} \vpageref[above]{sect:folFormExamples}, the  term \texttt{(var 0)} occurs freely and also inside the scope of a quantifier.
+I
 
@@
 To this purpose, we build \texttt{fol\_scope}, a  notation scope
-where the main connectives and quantifiers get a syntax close to \coq's.  Additionnally, a term of the form \texttt{(var $i$)} is
-just written \texttt{v\#$i$}.
+where the main connectives and quantifiers get a syntax close to \coq's.  Additionnally, a term of the form \texttt{(@var \_ $i$)} is
+just printed and parsed  \texttt{v\#$i$}.
 
@@
 
+\inputsnippets{FolExamples/toyNotationForm4}
+
 \section{Computing and reasoning on first-order formulas}
 
-\subsection{On variables}
+\subsection{Decidability of equality}
+% Deciding whether two formulas are equal is a part of the verification of a proof (for instance in the implementation of the \emph{assumption} rule).
 
-\begin{todo}
-  Difference from HOAS. Free and bound occurrences.
-\end{todo}
-
-\subsection{Decidability of equality}
 Let $L$ be a language, and let us assume that equality 
@@
 
-\begin{todo}
-  Should we talk about the decomposition lemmas
-\texttt{nilTerms} and \texttt{consTerms} ?
-\end{todo}
-
 \inputsnippets{fol/formDec1, fol/formDec2,
@@
 
-\subsection{Induction principles on formula depth}
+\begin{remark}
+Please note that \texttt{term\_dec}, \texttt{terms\_dec} and
+\texttt{formula\_dec} are \emph{opaque}.
 
-\inputsnippets{FolExamples/substNotStructural}
+The function \texttt{formula\_dec} is mainly used in
+\href{../theories/html/hydras.Ackermann.PA.html}{Ackermann.PA}, in order to check whether a given formula belongs to the axioms of Peano arithmetic.
+\begin{todo}
+  Look for the use of \texttt{open}  (in codePA)
+\end{todo}
 
-The purpose of this section is to present ways of proving 
-properties of any formula $a$ (or defining functions over the type of formulas).
+\end{remark}
 
-Structural induction/recursion principles are generated by 
-\texttt{Formula}'s definition, for instance:
+\subsection{Variables and substitutions}
 
-\inputsnippets{FolExamples/FormulaRect}
+Since free and bound occurrences of a variable $i$ are represented the same way, much care should be taken in programming the substitution of a term to a variable in order to avoid \emph{variable capture}.
+
+For instance let $F$ be the formula 
+$\exists v_2, v_1 \not= f(v_2)$. The result of the substitution 
+of the term $f(v_2)$ to the (free) occurrences of $v_1$ in $F$
+ \emph{is not} $\exists v_2, f(v_2) \not= f(v_2)$, but, for instance $\exists v_3, f(v_2) \not= f(v_3)$, where $v_3$ is a ``fresh'' variable.
+
+\inputsnippets{FolExamples/substExample1}
+\inputsnippets{FolExamples/substExample2}
+
+From this little example, we can guess that the function
+\texttt{substF} which is assumed to compute the substitution in a formula $F$ the substitution of a term $t$ to the free occurrences of a variable $v$ will not be implemented as a direct structural recursive function.
+For instance, the formula $G$, obtained through a variable renaming, is not a strict subterm of $F$.
+
+To fix this issue, we can associate  a measure (\emph{e.g.} a natural number) to each formula, such that the renaming of a strict subformula $F$ is strictly smaller that $F$.
 
 \subsubsection{Depth of a formula}
-As explained in~\cite{OConnor05}, some functions on the \texttt{Formula} type cannot be defined by structural recursion in 
-O'Connors formalization. For instance, the substitution of a term to a free variable (because of quantifiers). 
-In such cases, we may define functions by well-founded recursion. In this section, we consider the well-founded strict-order associated with a measure: the \emph{depth} of a formula.
-
-\begin{todo}
-Forward link to its application.
-\end{todo}
 
 The function \texttt{depth} computes the \emph{depth} of the 
- skeleton of any formula.
+ skeleton of any formula, and allows us to define a well-founded strict order on 
+\texttt{Formula $L$}.
 
@@
 \subsubsection{Induction on depth}
-
-The function \texttt{depth} allows us to define a strict order on 
-\texttt{Formula $L$}.
 
@@
 
+\begin{todo}
+Make a chapter!
+\end{todo}
 \section{Languages for Arithmetic}
@@
 
-\begin{todo}
-Fix the issue about the way formulas are printed (mix of different scopes).
-\end{todo}
-\inputsnippets{LNN_Examples/uglyF0a}
-
 \inputsnippets{LNN_Examples/CNNF0}

### 2023-06-07 doc/chapter-fol.tex
more abbreviations

@@
 
-\inputsnippets{fol/depthDef,  FolExamples/DepthCompute}
+\inputsnippets{fol/depthDef}
+
+\begin{remark}
+  The depth of a formula takes into account its abstract syntax tree \emph{with respect to the base connective and quantifiers : $\arrow$, $\sim$ and $\forall$}.
+    Formulas which contain $\vee$, $\wedge$, $\exists$, etc. are translated into
+    basic formulas before the computation of their depth. In the example below, 
+the conjunction is translated into a bigger term than the disjunction.
+
+\inputsnippets{FolExamples/DepthCompute}
+\end{remark}
 
 \subsubsection{Induction on depth}
-
-\inputsnippets{FolExamples/ltDepth1}
 

### 2023-06-11 doc/chapter-fol.tex
more abbreviations

--- 2023-05-23 doc/chapter-fol.tex
@@
 
-This chapter is devoted to the presentation to  data structures for representing first order formulas over a ranked alphabet, and the basic functions and predicates over these types:
-
-\begin{itemize}
-\item Abstract syntax of terms and formulas over a ranked alphabet composed of function and relation symbols.
-\item Induction principles over terms and formulas.
-\item Definition and main properties of substitution of terms to variables.
-\end{itemize}
-
+This chapter is devoted to the presentation to  data structures for representing first order formulas over a ranked alphabet, and the basic functions and predicates over these types.
 Although all the following constructions come directly from~\cite{Goedel}, we introduced minor changes (mainly syntax) to
@@
 
-A \emph{language} is a structure composed of relation and function symbols, each symbol is given an \emph{arity} (number of arguments)\,\footnote{As suggested by Russel O'Connor in~\cite{OConnor05}, we consider two arity functions instead of a single function defined on the sum type \texttt{Relations + Functions}.}.
+A \emph{language} is a structure composed of relation and function symbols, each symbol is given an \emph{arity} (number of arguments)\,\footnote{As suggested by Russel O'Connor in~\cite{OConnor05}, we consider two arity functions instead of a function defined on the sum type \texttt{Relations + Functions}}.
 
@@
 
+\begin{remark}[Variables]
+In O'Connor's formalization of first-order logic,  variables are \gallina terms of the form (\texttt{var $L$ $i$}), where $L$ is
+ a language and $i$ a natural number. 
+Although other choices may be considered : PHOAS, de Bruijn indices, etc,  we   respect this representation, in order not to break long proof scripts which use this representation (please look at \cite{OConnor05} for a related discussion).
+
+\end{remark}
+
 \begin{remark}
@@
 \end{remark}
-
-\begin{remark}[Variables]
-In O'Connor's formalization of first-order logic,  variables are 
-just natural numbers, and the \emph{coercion} from
-\texttt{nat} to \texttt{Term L} is the constructor (\texttt{@var $L$}).
-Although other choices may be considered : PHOAS, de Bruijn indices, etc,  we still the data structures of~\cite{Goedel}, in order not to break long proof scripts which use this representation (please look at \cite{OConnor05} for a related discussion).
-
-\end{remark}
-
-\subsubsection{Examples}
-\label{sect:folTermExamples}
-
-Let us define some \gallina terms over our toy language, respecively corresponding to the terms $a$,
-$f(a)$, and $h(f(a),a)$.
-
-First, in order to make terms on $L$ more readable, we introduce a few abbreviations.
-
-From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}.
-
-\inputsnippets{FolExamples/termAbbreviations}
-
-\inputsnippets{FolExamples/TermExamples1}
-
-\begin{remark}
-The fourth ``term'' \texttt{t4} is not well formed, since the arity of $h$ is not respected\,\footnote{Strictly speaking, it's not a (well typed) term!}.
-\end{remark}
-
-\subsubsection{Other Languages}
-
-\begin{todo}
-Link to the chapter which presents \texttt{LNT} and \texttt{LNN}.
-\end{todo}
-
-\subsection{First-order formulas}
-
-The type of first order formulas over $L$ is defined 
-in~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} as an inductive data type.
-
-\inputsnippets{fol/FormulaDef}
-
-\begin{remark}
-The type \texttt{Formula} (over $L$) has only five constructors,
-respectively bound to term equalities $t_1=t_2$, atomic formulas of the form $(r\;t_1\;\dots\;t_n)$, implications $(F \arrow G)$, negations $({\sim}F)$, and universal quantifications $(\forall\,i, F)$.
-
-In~\cite{Goedel}, no \emph{constructors} are associated with
-disjunction, conjonction, logical equivalence and existential quantifier.
-These constructs are formalized through \emph{definitions}\footnote{Please keep in mind that we are considering a classical logic.}.
-
-\inputsnippets{fol/FolFull, fol/folPlus}
-
-This convention allows the user to reduce to 5 (instead of 10) the number of cases in ''\texttt{match F with \dots}'' constructs.
-On the other hand, some computation may expand a connective like $\vee$ or
-$\wedge$, or an existential quantification into a ``basic'' formula (see Sect.\vref{sect:fol-issue}).
-\end{remark}
-
-\subsection{Examples}
-
-Let us give a few examples of first-order formulas over $L$.
-\label{fol:examplesf1f2f3}
-\begin{enumerate}
-\item $R\;a\;b$
-\item $\forall v_0\; v_1, R\;v_0\;v_1 \arrow R\;v_1\;v_0$
-\item $ \forall v_0, v_0=a \vee \exists\;v_1,\, v_0= f(v_1)$
-\item $(\forall v_1, v_0 = v_1) \vee \exists\,v_0\,v_1, v_0 \not= v_1$
-\end{enumerate}
-
-First, we define abbreviations for atomic formulas over $L$.
-
-\noindent From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}
-
-\inputsnippets{FolExamples/toyNotationForm}
-
-\label{sec:fol-atomic-notations}
-\label{sect:folFormExamples}
-
-\inputsnippets{FolExamples/FormExamples}
-
-% The following interactions show how the formulas 
-% of~\ref{sect:folFormExamples} may be 
-% parsed or printed.
-
-% \noindent From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}
-
-% \vspace{6pt}
-
-% \inputsnippets{FolExamples/toyNotationForm2}
 
@@
 
-In the formula \texttt{F4} \vpageref[above]{sect:folFormExamples}, the  term \texttt{(var 0)} occurs freely and also inside the scope of a quantifier.
-
   \begin{todo}
@@
 
-\section{A notation scope for first-order terms and formulas}
+\subsubsection{Examples}
+\label{sect:folTermExamples}
+
+Let us define some \gallina terms over our toy language, respecively corresponding to the terms $a$,
+$f(a)$, and $h(f(a),a)$. The fourth ``term'' is not well formed, since the arity of $h$ is not respected\,\footnote{Strictly speaking, it's not a (well typed) term!}.
+
+\inputsnippets{FolExamples/TermExamples1}
+
+\subsection{Formulas}
+
+The type of first order formulas over $L$ is defined 
+in~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} as an inductive data type.
+
+\inputsnippets{fol/FormulaDef}
+
+The  abstract syntax of first-order formulas is completed 
+with a few abbreviations\,\footnote{Technically, these abbreviations take the form of \emph{global definitions}}.
+Note that we are studying \emph{classical} first-order logic, for instance the existential quantifier is derived from the universal one through negation.
+
+\inputsnippets{fol/FolFull, fol/folPlus}
+
+\subsection{Examples}
+\label{sect:folFormExamples}
+
+Let us give a few examples of first-order formulas over $L$.
+
+\label{fol:examplesf1f2f3}
+\begin{enumerate}
+\item $R\;a\;b$
+\item $\forall v_0\; v_1, R\;v_0\;v_1 \leftrightarrow R\;v_1\;v_0$
+\item $ \forall v_0, v_0=a \vee \exists\;v_1,\, v_0= f(v_1)$
+\end{enumerate}
+
+\inputsnippets{FolExamples/FormExamples}
+
+\section{Simplified notations for terms and formulas}
 \label{sect:fol-notations}
@@
  We use \coq's \texttt{Notation} features to print and parse terms and formulas  in a more readable form.
-To this purpose, we build \texttt{fol\_scope}, a  notation scope
-where the main connectives and quantifiers get a syntax close to \coq's.  Additionnally, a term of the form \texttt{(var $i$)} is
-just written \texttt{v\#$i$}.
 
-\inputsnippets{fol/folScope1}
+ \subsection{Parsing and printing terms}
+ \label{sec:fol-term-notations}
 
-The \texttt{\%fol} delimiter 
+Let us consider the term $t_2$ of Sect\ref{sect:folTermExamples}~. Its normal form is quite hard to read:
+
+\inputsnippets{FolExamples/beforeAbbreviations}
+
+We may use \coq's abbreviation feature  to print and parse $L$-terms with a more natural concrete syntax.
+
+\inputsnippets{FolExamples/termAbbreviations, FolExamples/NowItsBetter}
+
+\subsection{Parsing and printing first-order formulas}
+
+In order to make the first order formulas more readable, we define  a  notation scope called \texttt{fol\_scope}, which contains 
+notations for connectives and quantifiers. The \texttt{\%fol} delimiter 
 allows the user to distinguish \texttt{FOL} connectives from their \coq equivalent.
-\emph{We discourage the reader from \emph{opening} \texttt{fol\_scope} and similar scopes : \texttt{nn\_scope}, \texttt{nt\_scope}, which would make disappear the \texttt{\%fol} suffix from the first-order formulas}.
 
@@
 
-\inputsnippets{FolExamples/toyNotationForm2}
+\inputsnippets{fol/folScope1}
 
-\subsection{The issue with derived constructions}
-\label{sect:fol-issue}
-
-The connectives and quantifiers $\vee$, $\wedge$, $\exists$, etc. may raise an issue when printing computed formulas.
-For instance, a formula like $F \wedge G$ could be transformed into $\sim(\sim F \vee \sim G)$,  and even
-into $\sim(\sim\sim F \arrow \sim B)$, which would 
-cause serious problems of readability.
-
-In such a case, we propose to print such a formula as $F \wedge' G$, to make it syntactically distinct but very similar to $F \wedge G$.
+The following notations are used (when printing) when  some computation 
+    expands a disjunction, conjuction, etc. 
+    in terms of implication and negation.
 
@@
 
-The following examples show how the primed connectors 
-and quantifiers behave with respect to convertibility and 
-input/output.
+\subsubsection{Simplified notations for atomic formulas}
+\label{sec:fol-atomic-notations}
 
-\inputsnippets{FolExamples/toyNotationForm3}
+In addition to \texttt{fol\_scope}, we can introduce a few abbreviations specific to our toy language.
+
+\inputsnippets{FolExamples/toyNotationForm}
+
+The following interactions show how the formulas 
+of~\ref{sect:folFormExamples} may be 
+parsed or printed.
+
+\noindent From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}
+
+\vspace{6pt}
+
+\inputsnippets{FolExamples/toyNotationForm2}
 
 \section{Computing and reasoning on first-order formulas}
-
-\subsection{On variables}
-
-\begin{todo}
-  Difference from HOAS. Free and bound occurrences.
-\end{todo}
 
@@
 \subsection{Induction principles on formula depth}
-
-\inputsnippets{FolExamples/substNotStructural}
 
@@
 
-\begin{todo}
-Link to folProp.v
-\end{todo}
-
 Section \texttt{Free\_Variables} 
@@
 \inputsnippets{FolExamples/freeVarExamples}
-
-\subsection{Multiple substitutions}
-
-\begin{todo}
-Link to subAll.v
-\end{todo}
 

### 2023-06-11 doc/hydra-chapter.tex

@@
 
-\index{hydras}{Projects}
-
-\begin{project}
-
+%\index{hydras}{Projects}
+
+\begin{remark}[Mutually inductive types vs lists of hydras]
+\label{hydra:mutually-inductive-vs-lists}
  Another very similar representation could use the \texttt{list} type family instead of the specific 
@@
 
-Using this representation, re-define all the constructions of this chapter.
+Using this representation, one can re-define all the constructions of this chapter, which is left as an exercise.
 You will probably have to use patterns described for instance in~\cite{BC04} or the archives of the \coq communication channels (please consult~\url{https://coq.inria.fr/community.html}).
 
-\end{project}
+\end{remark}
 

### 2023-06-13 doc/chapter-fol.tex

--- 2023-06-07 doc/chapter-fol.tex
@@
 
-I
-
   \begin{todo}
@@
 \section{Computing and reasoning on first-order formulas}
+
+\subsection{Structural recursion on formulas}
+
+\subsubsection{Free variables}
+The functions \texttt{freeVarT} [resp. \texttt{freeVarTs},  and
+\texttt{freeVarF}] compute the multiset (as a list with possible repetitions) of the free occurrences of variables in a term [resp. a vector of terms, a formula].
+
+\inputsnippets{folProp/freeVarT}
+
+\inputsnippets{folProp/freeVarF, FolExamples/freeVarEx}
+
+Note that \texttt{freeVarF} is defined by cases over the basic connectives. In the following example, the connective \texttt{<->} is expanded into a conjunction of two implications, thus the 
+list returned by \texttt{freeVarF} may contain redundancies.
+
+\inputsnippets{FolExamples/freeVarDup}
 
@@
 \subsection{Free variables and substitutions}
-
-\begin{todo}
-Link to folProp.v
-\end{todo}
-
-Section \texttt{Free\_Variables} 
-of~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} defines the list of free variables which occur in a term (respectively in a formula).
-
-We advise the reader to look at these functions, and limit ourselves to give a few examples.
-
-Let us consider again the formulas \texttt{f3} and \texttt{f4} of page~\pageref{fol:examplesf1f2f3}.
-
-The list of free variable occurrences of a formula is given by the function \texttt{freeVarFormula}. The function \texttt{close}
-binds all the free variables of a formula with universal quantifiers.
-
-\inputsnippets{FolExamples/freeVarExamples}
 

### 2023-06-15 doc/chapter-fol.tex
small changes in documentation

@@
 
-This chapter is devoted to the presentation to  data structures for representing first order formulas over a ranked alphabet, and the basic functions and predicates over these types:
+This chapter is devoted to the presentation of  data structures for representing terms and first order formulas over a ranked alphabet, and the basic functions and predicates over these types, more precisely:
 
@@
 
-\subsubsection{Example: a toy language}
+\subsubsection{Example: $L$, a toy language}
 In order to show a few simple examples of statements and proofs, we define a small language with very few symbols:
@@
   The constructors of types \texttt{Rel} and \texttt{Fun} are suffixed by an underscore, in order to reserve the names \texttt{a}, \texttt{f}, \texttt{h}, \texttt{R}, etc.
-to functions which build terms and formulas (please look at Sect~\ref{sec:fol-term-notations} and \ref{sec:fol-atomic-notations}).
+to the functions which build terms and formulas (please look at Sect~\ref{sec:fol-term-notations} and \ref{sec:fol-atomic-notations}).
 \end{remark}
@@
 \begin{remark}
-This representation of terms uses mutually inductive data-types instead of list of terms. Please see also Remark~\ref{hydra:mutually-inductive-vs-lists}.
+This representation of terms uses mutually inductive data-types instead of lists or vectors of terms. Please see also Remark~\ref{hydra:mutually-inductive-vs-lists}.
 \end{remark}
@@
 \texttt{nat} to \texttt{Term L} is the constructor (\texttt{@var $L$}).
-Although other choices may be considered : PHOAS, de Bruijn indices, etc,  we still the data structures of~\cite{Goedel}, in order not to break long proof scripts which use this representation (please look at \cite{OConnor05} for a related discussion).
+Although other choices may be considered : PHOAS, de Bruijn indices, etc,  we still the data structures of~\cite{Goedel}, in order not to break long proof scripts which use this representation (please look at Section 2 of~\cite{OConnor05} for a related discussion).
 
@@
 \label{sect:folTermExamples}
+\label{sec:fol-term-notations}
 
-Let us define some \gallina terms over our toy language, respecively corresponding to the terms $a$,
-$f(a)$, and $h(f(a),a)$.
+Let us build a few \gallina terms over our toy language, 
+respectively corresponding to the terms $a$,
+$f(a)$, $h(f(a),a)$, and $h(f(v_0),g(v_1))$.
 
@@
 
-\begin{remark}
-The fourth ``term'' \texttt{t4} is not well formed, since the arity of $h$ is not respected\,\footnote{Strictly speaking, it's not a (well typed) term!}.
-\end{remark}
+The following ``term'' \texttt{t4} is not well formed, since the arity of $h$ is not respected\,\footnote{Strictly speaking, it's not a (well typed) term!}.
+
+\inputsnippets{FolExamples/TermExamplesFail}
 
@@
 The type of first order formulas over $L$ is defined 
-in~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} as an inductive data type.
+in~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} as an inductive data type, with a limited set of basic constructions:
+\emph{term equalities} $t_1=t_2$,
+\emph{atomic propositions} $R\;t_1\;\dots\;t_n$, where $R$ is a relation symbol of arity $n$,
+\emph{implications} $A \rightarrow B$,
+\emph{negations} $\sim\,A$,
+and \emph{universal quantifications} $\forall\,i,\;A$.
+
+\vspace{6pt}
+
+\noindent From~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol}
 
@@
 \begin{remark}
-The type \texttt{Formula} (over $L$) has only five constructors,
-respectively associated with term equalities $t_1=t_2$, atomic formulas of the form $(r\;t_1\;\dots\;t_n)$, implications $(F \arrow G)$, negations $({\sim}F)$, and universal quantifications $(\forall\,i, F)$.
 
@@
 
-For instance, the the  term \texttt{(var 0)}
+For instance, the term \texttt{(var 0)}
 occurs freely and also inside the scope of a quantifier in the

### 2023-06-16 doc/chapter-fol.tex
small changes in documentation

@@
 \inputsnippets{fol/FormulaDef}
-
-Structural induction/recursion principles are generated by 
-\texttt{Formula}'s definition, for instance:
-
-\inputsnippets{FolExamples/FormulaRect}
 
@@
 
+Structural induction/recursion principles are generated by 
+\texttt{Formula}'s definition, for instance:
+
+\inputsnippets{FolExamples/FormulaRect}
+
 \subsubsection{Free variables}
@@
 
+\begin{remark}
   Note that \texttt{freeVarF} is defined by cases over the basic connectives. In the following example, the connective \texttt{<->} is expanded into a conjunction of two implications, thus the 
@@
 \inputsnippets{FolExamples/freeVarDup}
+
+\end{remark}
+
+\subsubsection{Closing a formula}
+Function \texttt{freeVarF} is used in 
+the  function \texttt{close}, which  universally quantifies  all the free variables of a formula.
+
+\noindent From \href{../theories/html/hydras.Ackermann.folProp.html}{Ackermann.folProp}
+
+\inputsnippets{folProp/closeDef}
+
+\noindent From \href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}
+
+\inputsnippets{FolExamples/closeExample}
+
+% \subsubsection{Computing a fresh variable}
+
+% It is easy to get a new variable, which
 
@@
 \inputsnippets{FolExamples/substExample1}
+
+Thus, the function (\texttt{fun f => substF L f i t}) doesn't commute with all \texttt{Formula}'s constructors.
+
 \inputsnippets{FolExamples/substExample2}
@@
 
-To fix this issue, we can associate  a measure (\emph{e.g.} a natural number) to each formula, such that the renaming of a strict subformula $F$ is strictly smaller that $F$.
+Thus, we cannot expect do define simply this operation by 
+structural recursion over formulas. 
+In~\cite{Goedel}, the approach is to define a \emph{measure}
+which maps every formula to a natural number, then define substitution in such a way that the substitution of a term to a free variable in a strict sub-formula  of a formula $F$ is smaller
+(w.r.t. this measure) than $F$. 
+Then substitution could be defined by well-founded recursion.
+
+Fortunately, such a measure : the \emph{depth} of a formula,
+can be easily defined by structural recursion.
+
+% To fix this issue, we can associate  a measure (\emph{e.g.} a natural number) to each formula, such that the renaming of a strict subformula $F$ is strictly smaller that $F$.
 
@@
 
-\begin{todo}
-Motivate the induction principles based on depth. Compatible with term substitution and universal quantifier elimination.
-\end{todo}
+% \begin{todo}
+% Motivate the induction principles based on depth. Compatible with term substitution and universal quantifier elimination.
+% \end{todo}
 

### 2023-06-20 doc/chapter-fol.tex
fix a latex error
small changes in documentation

@@
 
-Structural induction/recursion principles are generated by 
-\texttt{Formula}'s definition, for instance:
+Structural induction/recursion principles are generated from
+\texttt{Term}, \texttt{Terms} and \texttt{Formula}'s definition, for instance:
 
-\inputsnippets{FolExamples/FormulaRect}
+\inputsnippets{fol/TermSchemes, FolExamples/FormulaRect}
 
@@
 
+Concerning formulas, the treatement of binding is realized 
+whith the help of \texttt{List} library function \texttt{remove}.
 \inputsnippets{folProp/freeVarF, FolExamples/freeVarEx}
@@
 \begin{remark}
-  Note that \texttt{freeVarF} is defined by cases over the basic connectives. In the following example, the connective \texttt{<->} is expanded into a conjunction of two implications, thus the 
+  Note that \texttt{freeVarF} is defined by cases over the \emph{basic} connectives. Formulas with contain \texttt{iffH} or
+\texttt{ifThenElseH} are expanded before the application of
+\texttt{freevarF}, and the 
 list returned by \texttt{freeVarF} may contain redundancies.
@@
 Since free and bound occurrences of a variable $i$ are represented the same way, much care should be taken in programming the substitution of a term to a variable in order to avoid \emph{variable capture}.
+
+\begin{todo}
+  Present \texttt{substT} and \texttt{substTs} as structurally recursive functions. Then show \texttt{BadSubst.v} and motivate
+ depth induction.
+\end{todo}
+
+\inputsnippets{folProp/subsTdef}
+
+\inputsnippets{BadSubst/BadSubstFdef}
+
+\inputsnippets{BadSubst/BadSubstFexample}
 

### 2023-06-21 doc/chapter-fol.tex
small changes in documentation

@@
 
+If we want to get the \emph{set} of variables wich occur freely 
+in a formula $F$, we may use the function \texttt{List.nodup}.
+
 \inputsnippets{FolExamples/freeVarDup}
@@
 
+\begin{remark}
+  The function \texttt{close} applies \texttt{freeVarF} and
+  \texttt{List.nodup}  in order to add a sequence of universal
+  quantifications \texttt{allH $i_1$ \dots $i_k$}, in an order determined by the implementation of these functions.
+It may be interesting to check whether the proof of properties of \texttt{close} depend or not from this implementation.
+\end{remark}
 % \subsubsection{Computing a fresh variable}
@@
 \href{../theories/html/hydras.Ackermann.PA.html}{Ackermann.PA}, in order to check whether a given formula belongs to the axioms of Peano arithmetic.
+
 \begin{todo}
@@
 
+The substitution of a term to the free occurrences of a given variable $v$ is at the heart of the implementation of universal
+quantifier elimination.
+
 Since free and bound occurrences of a variable $i$ are represented the same way, much care should be taken in programming the substitution of a term to a variable in order to avoid \emph{variable capture}.
 
-\begin{todo}
-  Present \texttt{substT} and \texttt{substTs} as structurally recursive functions. Then show \texttt{BadSubst.v} and motivate
- depth induction.
-\end{todo}
+Substitution of a term $t$ to all the occurrences of a variable $x$ in a term or a vector of terms is easy to define as a pair of mutually structurally recursive functions.
 
 \inputsnippets{folProp/subsTdef}
+\inputsnippets{FolExamples/substTExample}
 
+Concerning formulas, it could be tempting to define substitution the same way.
 \inputsnippets{BadSubst/BadSubstFdef}
+
+Let us consider for instance the formula $F = \forall\,v_1, \,\exists\,v_2,\,v_1\not=f(v_2)$ (satisfiable if we take for instance $f$ to be interpreted as the successor function).
+
+If we eliminate the universal quantifier by substituting in
+the sub-formula $F_1=  \exists\,v_2,\,v_1\not=f(v_2)$ the free
+occurrences of $v_1$ with $f(v_2)$, our naive implementation of
+\texttt{substF} returns the absurd proposition $\exists\,v_2,\,f(v_2)\not=f(v_2)$.
+We  say that the free occurrence of $v_2$ in the term $f(v_2)$ has been \emph{captured} by the binding $\exists\,v_2,\dots$.
 
@@
 
-For instance let $F$ be the formula 
-$\exists v_2, v_1 \not= f(v_2)$. The result of the substitution 
-of the term $f(v_2)$ to the (free) occurrences of $v_1$ in $F$
- \emph{is not} $\exists v_2, f(v_2) \not= f(v_2)$, but, for instance $\exists v_3, f(v_2) \not= f(v_3)$, where $v_3$ is a ``fresh'' variable.
+In our example, we can obtain a correct result, if
+\begin{enumerate}
+\item We consider a \emph{fresh} variable, \emph{i.e.}  different from $v_1$ and $v_2$, say for instance $v_3$,
+\item we substitute $v_3$ to $v_2$ in $F_1$ which results in
+   $F_2= \exists\,v_3,\,v_1\not=f(v_3)$
+\item we substitute the term $f(v_2)$ to $v_1$ in $F_2$, which gives us 
+  $\exists\,v_3,\,f(v_2)\not=f(v_3)$.
+\end{enumerate}
+
+The notion of fresh variable is implemented through a function
+\texttt{newvar (l: list nat) : nat} which returns a number which doesn't belong to $l$.
+
+But the following attempt fails, because the renaming of a  variable in a 
+sub-formula of a formula $F$ is not structurally smaller than $F$.
+
+\inputsnippets{BadSubst/BadSubstF2}
+
+%ICI
 

### 2023-06-22 doc/chapter-fol.tex
small changes in documentation

@@
 
-\inputsnippets{FolExamples/substExample1}
+%\inputsnippets{FolExamples/substExample1}
 
-Thus, the function (\texttt{fun f => substF L f i t}) doesn't commute with all \texttt{Formula}'s constructors.
+% Thus, the function (\texttt{fun f => substF L f i t}) doesn't commute with all \texttt{Formula}'s constructors.
 
-\inputsnippets{FolExamples/substExample2}
+% \inputsnippets{FolExamples/substExample2}
 
-From this little example, we can guess that the function
-\texttt{substF} which is assumed to compute the substitution in a formula $F$ the substitution of a term $t$ to the free occurrences of a variable $v$ will not be implemented as a direct structural recursive function.
+% From this little example, we can guess that the function
+% \texttt{substF} which is assumed to compute the substitution in a formula $F$ the substitution of a term $t$ to the free occurrences of a variable $v$ will not be implemented as a direct structural recursive function.
 For instance, the formula $G$, obtained through a variable renaming, is not a strict subterm of $F$.
 
-Thus, we cannot expect do define simply this operation by 
-structural recursion over formulas. 
-In~\cite{Goedel}, the approach is to define a \emph{measure}
-which maps every formula to a natural number, then define substitution in such a way that the substitution of a term to a free variable in a strict sub-formula  of a formula $F$ is smaller
-(w.r.t. this measure) than $F$. 
-Then substitution could be defined by well-founded recursion.
+% Thus, we cannot expect do define simply this operation by 
+% structural recursion over formulas. 
+% In~\cite{Goedel}, the approach is to define a \emph{measure}
+% which maps every formula to a natural number, then define substitution in such a way that the substitution of a term to a free variable in a strict sub-formula  of a formula $F$ is smaller
+% (w.r.t. this measure) than $F$. 
+% Then substitution could be defined by well-founded recursion.
 
-Fortunately, such a measure : the \emph{depth} of a formula,
-can be easily defined by structural recursion.
+Fortunately, \coq allows us to define functions by well-founded recursion, and in particular with the help of a \emph{measure} mapping every formula to an already known well-founded type.
 
@@
 
-The function \texttt{depth} computes the \emph{depth} of the 
- skeleton of any formula, and allows us to define a well-founded strict order on 
-\texttt{Formula $L$}.
+The function \texttt{depth} computes the \emph{depth} of  any formula, \emph{i.e} the height of the sub-tree made
+by erasing all nodes but those nodes labelled with
+\texttt{allH}, \texttt{impH} and \texttt{notH}.
 
@@
 
+Lemma \texttt{fol.Formula\_depth\_rec} is the basic
+induction principle based on depth.
+
+\inputsnippets{FolExamples/depthRec}
+
 % \begin{todo}
@@
 
-The library~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} contains several induction principles, applied 
+The library~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} contains several derived induction principles, applied 
 throughout \texttt{Ackermann} and \texttt{Goedel} projects.

### 2023-06-26 doc/chapter-fol.tex
small changes in documentation

@@
   \texttt{List.nodup}  in order to add a sequence of universal
-  quantifications \texttt{allH $i_1$ \dots $i_k$}, in an order determined by the implementation of these functions.
+  quantifications \texttt{allH $i_1\, \dots\, i_k$}, in an order determined by the implementation of these functions.
 It may be interesting to check whether the proof of properties of \texttt{close} depend or not from this implementation.
@@
 
-Let us for instance have a look  at \texttt{Formula\_depth\_ind2}. Its application in order to prove $P\;a$ generates 5 sub-goals.
+Let us for instance have a look  at \texttt{Formula\_depth\_ind2} which helps  to prove a goal $(P\;a)$ by generating five sub-goals.
 
@@
 \begin{itemize}
-\item Goals $1$ to $4$ correspond to  usual proofs by structural induction.
+\item Goals $1$ to $4$ correspond to  usual proofs by structural induction (without refering to \texttt{depth}).
 \item Goal $5$ is associated with a universal quantification $f=\forall\,v,a$.
 In this case, we have to prove that $P\;b$ holds for any formula $b$ which has a depth strictly less than $f$.
-Such a $b$ may for instance the result of replacing the free occurrences of $v$ in $a$ with any term $t$.
+Such a $b$ may for instance be the result of replacing the free occurrences of $v$ in $a$ with any term $t$.
+
   \begin{todo}
@@
 
-\subsection{Free variables and substitutions}
+\subsection{A correct definition of substF}
+
+Substitution of a term to (free- occurrences of a variable 
+in a formula is defined in a section of \href{../theories/html/hydras.Ackermann.folProp.html}{Ackermann.folProp}.
+The definition itself takes 200 lines of \coq code, so we will only comment the \emph{structure} of this definition.
+
+Fortunately, the reader may skip a few complex definitions 
+(often because of dependent pattern matching), whose purpose
+is twofold:
+\begin{itemize}
+\item Ensure that the substitution in a formula $F$ of a term $t$ to a variable $x$ returns a formula of the same depth as $F$,
+  which allows to define a function by well-founded recursion on depth.
+  \inputsnippets{folProp/substFHelp, folProp/substFDef}
+\item Prove a few \emph{equations} which will be used in firther proofs on substitutions.
+
+\inputsnippets{folProp/subFormulaEqual}
+\inputsnippets{folProp/subFormulaRelation}
+\inputsnippets{folProp/subFormulaImp}
+\inputsnippets{folProp/subFormulaNot}
+\inputsnippets{folProp/subFormulaForall}
+
+\item Similar equations are also proved for derived connectors and quantifiers, for instance:
+
+\inputsnippets{folProp/subFormulaAnd}
+\inputsnippets{folProp/subFormulaExist}
+
+\end{itemize}
+
+\begin{project}
+  Is it possible to get a more readable definition of \texttt{substF} using the \emph{Equations} plug-in~\cite{sozeau:hal-01671777} ? 
+\end{project}
 

### 2023-06-26 doc/schutte-chapter.tex
small changes in documentation

@@
 Finally, we prove that \texttt{inject} is a bijection from the set of all terms of \texttt{T1} in normal form to the set 
-\texttt{members epsilon0} of the elements of \texttt{Ord} strictly less than  $\epsilon_0$.
+(\texttt{members epsilon0}) of the elements of \texttt{Ord} strictly less than  $\epsilon_0$.
 
@@
 \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Logic.ChoiceFacts.html}{\texttt{Coq.Logic.ChoiceFacts}}).
-Nevertheless, the other modules:
+
+On the other hand, the other libraries:
 \href{../theories/html/hydras.Epsilon0.Epsilon0.html}%

### 2023-06-30 doc/chapter-fol.tex
minor corrections
more documentation

@@
 
+\inputsnippets{FolExamples/substExample1}
+
 \begin{project}
@@
 
+The function \texttt{subFormula} defined in
+\href{../theories/html/hydras.Ackermann.subAll.html}{Ackermann.subAll} allows to substitute a term $t_i$ to ecah free occurrence of the variable $v_i$ in a formula $F$.
+The dependance of $t-i$ from $v_i$ is given through a function from \texttt{nat} to
+(\texttt{Term L}).
+
+\inputsnippets{FolExamples/subAllExample1}
+
+\section{Proofs}
+
+\subsection{Proof trees}
+
+Proof trees in first-order logic are the inhabitants of the \texttt{Prf $L$} 
+inductive type, defined in \href{../theories/html/hydras.Ackermann.folProof.html}{Ackermann.folProof}. Since this type has 14 constructors, we present
+them one by one, after quoting its full definition.
+
+Please note that the following constructions are parametrized with an
+arbitrary language $L$, decared in the \texttt{ProofH} section.
+
+\inputsnippets{folProof/prelude}
+
+\inputsnippets{folProof/PrfDef}
+
+\subsubsection{\texttt{Prf}'s type}
+
+We want to define a data type  \texttt{Prf $l$ $F$}  of ``proof trees of the formula $F$, where $l$ is the list of assumptions used in the proof, enumerated left-to-right (\emph{i.e.} the fringe of the proof tree)''.
+Please note that this representation
+
+\inputsnippets{folProof/PrfAbout}
+
+\subsubsection{The axiom rule}
+
+Let $A$ be a formula on $L$.  The \texttt{AXM} rule builds a proof-tree
+of $A$ which uses \emph{exactly} the singleton list \texttt{[$\,A\,$]}.
+
+\inputsnippets{folProof/AXM}
+
+\noindent From \href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}
+
+\inputsnippets{FolExamples/PrfEx1}
+
+\subsubsection{Modus Ponens}
+
+\inputsnippets{folProof/MP}
+
+\inputsnippets{FolExamples/PrfEx2}
+
+\inputsnippets{FolExamples/eat}
+
 \begin{todo}
-Link to subAll.v
+Give examples in a toy language.
+
+Give an example which motivates \texttt{SysPrf}:
+
+$P,P , P\arrow P \arrow Q, \vdash Q$ ?
 \end{todo}
+
+\begin{exercise}
+
+Prove the following lemma:
+\inputsnippets{FolExamples/eqRefl}
+
+\end{exercise}
+
+\subsection{Contexts as sets}
+
+\begin{exercise}
+
+Prove the following lemma:
+\inputsnippets{FolExamples/MPdiag}
+
+\end{exercise}
+
+\section{Derived rules and natural deduction}
+
+ The library 
+ \href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic} contains many derived rules which allow the user to build proofs in a natural deduction style (with introduction and elimination rules).
+
+\subsection{Example}
+
+For instance, let us prove Peirce's rule.
+ (in  ~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples} ).
+
+\subsubsection{Prelude}
+
+%\inputsnippets{folPeirce/prelude}
+
+\subsection{Proof of Peirce's law}
+
+\inputsnippets{FolExamples/PeirceProof}
+
+Let us try to do an implication introduction.
+
+\inputsnippets{FolExamples/step1}
+
+Now, we may use the law of excluded middle with the formula $P$. The only non-trivial case is about $\sim P$.
+
+\inputsnippets{FolExamples/step2}
+
+The rest of the proof is composed of basic proof steps, 
+and bookkeeping steps (about \texttt{Ensembles.In}).
+
+\inputsnippets{FolExamples/step3, FolExamples/step4}
+
+\index{ackermann}{Exercises}
+\begin{exercise}
+Prove, using the rules described in 
+ \href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic}, the famous \emph{drinkers theorem}:
+
+$$\exists\,x,\; (D(x)\Longrightarrow \forall\,y,\; D(y))$$
+\end{exercise}
+
+where $D$ (for ``drinks'') is some predicate symbol of arity $1$.
+
+\section{Languages for Arithmetic}
 
@@
 \end{todo}
-\section{Languages for Arithmetic}
 
@@
 \inputsnippets{LNN_Examples/CNNF0}
-
-\section{Proofs}
-
-\subsection{Contexts as lists}
-\begin{todo}
-Describe the \texttt{Prf} type (from \texttt{folProof.v}).
-
-Give examples in a toy language.
-
-Give an example which motivates \texttt{SysPrf}:
-
-$P,P , P\arrow P \arrow Q, \vdash Q$ ?
-\end{todo}
-
-\begin{exercise}
-
-Prove the following lemma:
-\inputsnippets{FolExamples/eqRefl}
-
-\end{exercise}
-
-\subsection{Contexts as sets}
-
-\begin{exercise}
-
-Prove the following lemma:
-\inputsnippets{FolExamples/MPdiag}
-
-\end{exercise}
-
-\section{Derived rules and natural deduction}
-
- The library 
- \href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic} contains many derived rules which allow the user to build proofs in a natural deduction style (with introduction and elimination rules).
-
-\subsection{Example}
-
-For instance, let us prove Peirce's rule.
- (in  ~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples} ).
-
-\subsubsection{Prelude}
-
-%\inputsnippets{folPeirce/prelude}
-
-\subsection{Proof of Peirce's law}
-
-\inputsnippets{FolExamples/PeirceProof}
-
-Let us try to do an implication introduction.
-
-\inputsnippets{FolExamples/step1}
-
-Now, we may use the law of excluded middle with the formula $P$. The only non-trivial case is about $\sim P$.
-
-\inputsnippets{FolExamples/step2}
-
-The rest of the proof is composed of basic proof steps, 
-and bookkeeping steps (about \texttt{Ensembles.In}).
-
-\inputsnippets{FolExamples/step3, FolExamples/step4}
-
-\index{ackermann}{Exercises}
-\begin{exercise}
-Prove, using the rules described in 
- \href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic}, the famous \emph{drinkers theorem}:
-
-$$\exists\,x,\; (D(x)\Longrightarrow \forall\,y,\; D(y))$$
-\end{exercise}
-
-where $D$ (for ``drinks'') is some predicate symbol of arity $1$.

### 2023-07-02 doc/chapter-fol.tex
more documentation on FOL

@@
 Please note that the following constructions are parametrized with an
-arbitrary language $L$, decared in the \texttt{ProofH} section.
+arbitrary language $L$, declared in the \texttt{ProofH} section.
 
@@
 We want to define a data type  \texttt{Prf $l$ $F$}  of ``proof trees of the formula $F$, where $l$ is the list of assumptions used in the proof, enumerated left-to-right (\emph{i.e.} the fringe of the proof tree)''.
-Please note that this representation
 
 \inputsnippets{folProof/PrfAbout}
+
+\subsubsection{Warning}
+
+We won't respect the order in which \texttt{Prf}'s constructor 
+are enumerated in \texttt{folProof.v}. We preferred to present 
+these constructors in an order inspired by a sequence
+ of simple examples. On the other hand, we didn't change this order in 
+\texttt{folProof.v}, in order not to break complex proofs by pattern-matching.
+
+See \href{https://en.wikipedia.org/wiki/List_of_Hilbert_systems}.
 
@@
 
+Let $A$ and $B$ be two formulas on $L$, $\textit{Axm1}$
+ and $\textit{Axm2}$ two sequences of formulas.
+If we have 
+two proof trees of respective types \texttt{Prf Axm1 $A\arrow B$} and  \texttt{Prf Axm2 $B$}, then we build a proof tree
+for $B$ whose fringe is the concatenation of $s_{AB}$ and $s_A$.
+
 \inputsnippets{folProof/MP}
+
+The following proof script is a quite naive application of 
+\texttt{AXM} and \texttt{MP}.
 
@@
 
-\inputsnippets{FolExamples/eat}
+We can make this proof script shorter, using existential variables.
+
+\inputsnippets{FolExamples/cutMP}
+
+\inputsnippets{FolExamples/cutMP2}
+
+\subsubsection{Hilbert's axioms for implication}
+
+\inputsnippets{folProof/IMP12}
+
+\inputsnippets{FolExamples/PrfEx3}
+
+\begin{remark}
+\label{sec:remark-on-hilbert-systems}
+One may think that this proof 
+is quite clumsy. \emph{Right}.
+But we must recall that \texttt{Prf} is a basic Hilbert-like
+proof system, which will make it easier to study.
+In the next chapter, we will consider a derived proof system 
+where the \emph{deduction theorem} will allow us to consider shorter and more natural proofs.
+\end{remark}
+
+\begin{exercise}[**]
+  Is it possible to build a term of type
+\texttt{Prf L [A] B -> Prf L [] (A->B)\%fol} ?
+\end{exercise}
+
+\subsubsection{The rule of contraposition}
+
+The only rule about the \texttt{notH} connective is
+the \emph{contraposition rule}.
+
+\inputsnippets{folProof/CP}
+
+The following script shows that \texttt{CP} entails the 
+derived rule of \emph{proof by contradiction}.
+
+\inputsnippets{FolExamples/PrfEx4}
+
+\begin{remark}
+Same remark as~\ref{sec:remark-on-hilbert-systems}.
+\end{remark}
+
+\begin{exercise}
+  Replay this proof with pen and paper!
+\end{exercise}
 
 \begin{todo}
-Give examples in a toy language.
-
-Give an example which motivates \texttt{SysPrf}:
+To do propositional connectives before FO (in the doc, but not 
+ in the vernacular file)
+To do : EQ1 before GEN
 
@@
 
+%%%%%%%%%%%%%% New chapter 
 \subsection{Contexts as sets}
@@
 where $D$ (for ``drinks'') is some predicate symbol of arity $1$.
+
+%%%%%%%%%%%%%%% New chapter!
 

### 2023-07-03 doc/chapter-fol.tex
improve chapter-fol.tex

@@
 
-Although all the following constructions come directly from~\cite{Goedel}, we introduced minor changes (mainly syntax) to
+Although all the following constructions come directly from Russel O'Connor's work~\cite{Goedel}, we introduced minor (mainly syntactic) changes  to
 take into account recent changes in \coq(new constructions, tactics, notations, etc.).
@@
 In order to show a few simple examples of statements and proofs, we define a small language with very few symbols:
-two constant symbols: $a$ and $b$, three function symbols $f$, $g$ and $h$  (of respective arity 1, 1 and 2), three propositional symbols $A$, $B$ and $C$, a one-place predicate symbol $P$, and a binary relational symbol $R$.
+two constant symbols: $a$ and $b$, three function symbols $f$, $g$ and $h$  (of respective arity 1, 1 and 2), three propositional symbols $A$, $B$ and $C$, two one-place predicates symbols $P$ and $Q$, and a binary relational symbol $R$.
 
@@
 \begin{remark}
-This representation of terms uses mutually inductive data-types instead of lists or vectors of terms. Please see also Remark~\ref{hydra:mutually-inductive-vs-lists}.
+This representation of terms uses mutually inductive data-types instead of lists or vectors of terms. Please see also Remark~\vref{hydra:mutually-inductive-vs-lists}.
 \end{remark}
@@
 In O'Connor's formalization of first-order logic,  variables are 
-just natural numbers, and the \emph{coercion} from
+just natural numbers, and the conversion from
 \texttt{nat} to \texttt{Term L} is the constructor (\texttt{@var $L$}).
@@
 disjunction, conjonction, logical equivalence and existential quantifier.
-These constructs are formalized through \emph{definitions}\footnote{Please keep in mind that we are considering a classical logic.}.
+These constructs are formalized through \emph{definitions} in terms of \texttt{impH}, \texttt{notH} and \texttt{forallH}\,\footnote{Please keep in mind that we are considering a classical logic.}.
 
@@
 
-This convention allows the user to reduce to 5 (instead of 10) the number of cases in ''\texttt{match F with \dots}'' constructs.
-On the other hand, some computation may expand a connective like $\vee$ or
+This convention allows the user to reduce to 5 (instead of 10) the number of cases in ''\texttt{match F with \dots}'' terms. On the other hand, some computation may expand a connective like $\vee$ or
 $\wedge$, or an existential quantification into a ``basic'' formula (see Sect.\vref{sect:fol-issue}).
@@
 For instance, the term \texttt{(var 0)}
-occurs freely and also inside the scope of a quantifier in the
+occurs both freely and  inside the scope of a quantifier in the
 formula \texttt{F4} \vpageref[above]{sect:folFormExamples}.
@@
   \begin{project}
- Use a HOAS representation for FOL terms and formulas (without breaking proof scripts).   
+ Define a [P]HOAS representation for FOL terms and formulas.
+Could we avoid  to  break some proof scripts?   
   \end{project}
@@
 
+\begin{remark}
+In some situations (like in the proof of \texttt{PrfEx4}~\vref{prfex4})
+the user may be puzzled when a formula [s]he typed explicitely
+\texttt{$\sim A$ ->$\sim B$}  will be printed 
+\texttt{$A$  \textbackslash/' $\sim B$}. We will try to fix this issue.
+\end{remark}
+
 \section{Computing and reasoning on first-order formulas}
@@
 Concerning formulas, the treatement of binding is realized 
-whith the help of \texttt{List} library function \texttt{remove}.
+whith the help of the \texttt{List} library function \texttt{remove}.
 \inputsnippets{folProp/freeVarF, FolExamples/freeVarEx}
@@
   \texttt{List.nodup}  in order to add a sequence of universal
-  quantifications \texttt{allH $i_1\, \dots\, i_k$}, in an order determined by the implementation of these functions.
+  quantifications (\texttt{allH $i_1\, \dots\, i_k$}), in an order determined by the actual implementation of these functions.
 It may be interesting to check whether the proof of properties of \texttt{close} depend or not from this implementation.
@@
 
-Let us consider for instance the formula $F = \forall\,v_1, \,\exists\,v_2,\,v_1\not=f(v_2)$ (satisfiable if we take for instance $f$ to be interpreted as the successor function).
+Let us consider for instance the formula $F = \forall\,v_1, \,\exists\,v_2,\,v_1\not=f(v_2)$ (satisfiable if we take for instance $f$ to be interpreted as the successor function on natural numbers).
 
@@
 
-In our example, we can obtain a correct result, if
+In this example, we could obtain a correct result, if
 \begin{enumerate}
@@
 in a formula is defined in a section of \href{../theories/html/hydras.Ackermann.folProp.html}{Ackermann.folProp}.
-The definition itself takes 200 lines of \coq code, so we will only comment the \emph{structure} of this definition.
+The definition itself takes 200 lines of \coq code, so we will only comment its structure.
 
@@
   \inputsnippets{folProp/substFHelp, folProp/substFDef}
-\item Prove a few \emph{equations} which will be used in firther proofs on substitutions.
+\item Prove a few \emph{equations} which will be used in further proofs.
 
@@
 \end{itemize}
+
+Let us look at a few examples. Despite the complexity of its definitoin, the function \texttt{substF} behaves well with respect with computations.
 
@@
 Proof trees in first-order logic are the inhabitants of the \texttt{Prf $L$} 
-inductive type, defined in \href{../theories/html/hydras.Ackermann.folProof.html}{Ackermann.folProof}. Since this type has 14 constructors, we present
-them one by one, after quoting its full definition.
+inductive type, defined in \href{../theories/html/hydras.Ackermann.folProof.html}{Ackermann.folProof} and displayed in Figure~\vref{fig:prf-def}.
 
@@
 
+\begin{figure}[p]
+  \centering
 \inputsnippets{folProof/PrfDef}  
+  \caption{Definition of the type \texttt{Prf} of proof trees}
+  \label{fig:prf-def}
+\end{figure}
 
@@
 
-We want to define a data type  \texttt{Prf $l$ $F$}  of ``proof trees of the formula $F$, where $l$ is the list of assumptions used in the proof, enumerated left-to-right (\emph{i.e.} the fringe of the proof tree)''.
+The type  \texttt{Prf $l$ $F$}  is the type of ``proof trees of the formula $F$, where $l$ is the list of assumptions used in the proof, enumerated left-to-right (\emph{i.e.} the fringe of the proof tree)''.
 
-\inputsnippets{folProof/PrfAbout}
+In the rest of this section, we comment  every one of 
+\texttt{Prf}'s 14 constructors,
+ and give simple examples of their application.
 
@@
 
-We won't respect the order in which \texttt{Prf}'s constructor 
-are enumerated in \texttt{folProof.v}. We preferred to present 
+We won't respect the order in which \texttt{Prf}'s constructors
+of type \texttt{folProof}  
+are enumerated in \texttt{folProof.v} (see Figure~\ref{fig:prf-def}). 
+Instead, we preferred to present 
 these constructors in an order inspired by a sequence
@@
 
-See \href{https://en.wikipedia.org/wiki/List_of_Hilbert_systems}.
+For more information on Hilbert proof system, you may consult
+\url{https://en.wikipedia.org/wiki/List_of_Hilbert_systems}.
 
-\subsubsection{The axiom rule}
+\subsubsection{The axiom rule: \texttt{AXM}}
 
@@
 of $A$ which uses \emph{exactly} the singleton list \texttt{[$\,A\,$]}.
-
-\inputsnippets{folProof/AXM}
 
@@
 
-\subsubsection{Modus Ponens}
+\subsubsection{Modus Ponens: \texttt{MP}}
 
@@
 for $B$ whose fringe is the concatenation of $s_{AB}$ and $s_A$.
-
-\inputsnippets{folProof/MP}
 
@@
 
-\subsubsection{Hilbert's axioms for implication}
-
-\inputsnippets{folProof/IMP12}
+\subsubsection{Hilbert's axioms for implication: \texttt{IMP1} and \texttt{IMP2}}
 
@@
 
-\subsubsection{The rule of contraposition}
+\subsubsection{The rule of contraposition: \texttt{CP}}
 
@@
 
-\inputsnippets{folProof/CP}
-
 The following script shows that \texttt{CP} entails the 
 derived rule of \emph{proof by contradiction}.
+
+\label{prfex4}
 
@@
 
-\begin{todo}
-To do propositional connectives before FO (in the doc, but not 
- in the vernacular file)
-To do : EQ1 before GEN
+\subsubsection{Rules about the universal quantifier: \texttt{FA1},
+\texttt{FA2} and\texttt{FA3}}
 
-$P,P , P\arrow P \arrow Q, \vdash Q$ ?
-\end{todo}
+\inputsnippets{FolExamples/PrfEx5}
 
-\begin{exercise}
+\inputsnippets{FolExamples/PrfEx6}
 
-Prove the following lemma:
+\inputsnippets{FolExamples/PrfContrex7}
+
+\inputsnippets{FolExamples/PrfEx8}
+
+\subsubsection{Axioms for Equality: \texttt{EQ1} to \texttt{EQ3}}
+
+The following proof applies \texttt{FA1} in order to build a proof
+of \texttt{Prf L [] (t=t)\%fol}  for any $t$\,\footnote{Please note that the quantification on $f$ is \coq's. Thus, \texttt{eq\_ref} is kind of scheme.}
+
 \inputsnippets{FolExamples/eqRefl}
 
-\end{exercise}
+\subsubsection{Axioms schemes for Equality: \texttt{EQ4} and \texttt{EQ5}}
 
-%%%%%%%%%%%%%% New chapter 
-\subsection{Contexts as sets}
+The constructors \texttt{EQ4}  and \texttt{EQ5} build an infinite number of axioms, parameterized by a  relation  or function symbol.
 
-\begin{exercise}
+The function \texttt{AxmEq4} generates a formula with $n$ pairs of free variables (with $n$ the arity of the considered relation symbol).
 
-Prove the following lemma:
-\inputsnippets{FolExamples/MPdiag}
+\inputsnippets{FolExamples/PrfEx910}
 
-\end{exercise}
+Please note that \texttt{EQ4} uses a \emph{sequence} of variables generated by \texttt{AxmEq4}. Any other sequence may cause 
+\texttt{EQ4} to fail.
 
-\section{Derived rules and natural deduction}
+\inputsnippets{FolExamples/PrfContrex9}
 
- The library 
- \href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic} contains many derived rules which allow the user to build proofs in a natural deduction style (with introduction and elimination rules).
+The following script shows that \texttt{EQ5} is to function symbols 
+what \texttt{EQ4} is to relation symbols.
 
-\subsection{Example}
+\inputsnippets{FolExamples/PrfEx11}
 
-For instance, let us prove Peirce's rule.
- (in  ~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples} ).
+\section{Concluding remarks}
 
-\subsubsection{Prelude}
+The type \texttt{Prf} is composed of very simple rules.
+Nevertheless, the examples presented in the previous section seem to show that proving even simple theorems is not trivial at all.
 
-%\inputsnippets{folPeirce/prelude}
-
-\subsection{Proof of Peirce's law}
-
-\inputsnippets{FolExamples/PeirceProof}
-
-Let us try to do an implication introduction.
-
-\inputsnippets{FolExamples/step1}
-
-Now, we may use the law of excluded middle with the formula $P$. The only non-trivial case is about $\sim P$.
-
-\inputsnippets{FolExamples/step2}
-
-The rest of the proof is composed of basic proof steps, 
-and bookkeeping steps (about \texttt{Ensembles.In}).
-
-\inputsnippets{FolExamples/step3, FolExamples/step4}
-
-\index{ackermann}{Exercises}
-\begin{exercise}
-Prove, using the rules described in 
- \href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic}, the famous \emph{drinkers theorem}:
-
-$$\exists\,x,\; (D(x)\Longrightarrow \forall\,y,\; D(y))$$
-\end{exercise}
-
-where $D$ (for ``drinks'') is some predicate symbol of arity $1$.
+Indeed, in the next chapter, we will consider a proof system
+\texttt{SysPrf}, based on \texttt{Prf}, the properties of which will 
+allow us to prove theorems in a very simpler way.
 
 %%%%%%%%%%%%%%% New chapter!
-
-\section{Languages for Arithmetic}
-
-\begin{todo}
-Make a chapter!
-\end{todo}
-
-Two languages built with the usual symbols of arithmetic are 
-defined in ~\href{../theories/html/hydras.Ackermann.Languages.html}{Ackermann.Languages}.
-
-\begin{itemize}
-\item The first language: \texttt{LNT} (\emph{Language of Number Theory}) has just function symbols for $+$, $\times$, $0$ and successor.
-\item The second language: LNN (\emph{Language of Natural Numbers})  has
-the same function symbols as LNT plus one relation symbol for the strict inequality $<$ : \texttt{LT} (less than).
-\end{itemize}
-
-\subsubsection{Language of Number Theory (LNT)}
-
-First, we declare two alphabets.
-
-\inputsnippets{Languages/LNTDef1}
-
-In a second time, we build \texttt{LNT} and \texttt{LNN} by filling \texttt{Language}'s \texttt{arity} field.
-
-\inputsnippets{Languages/LNTDef2}
-
-\begin{remark}
-  We depart a little from \cite{Goedel}'s notations, where the 
-function and relation symbols are called \texttt{Plus}, 
-\texttt{Mult}, \texttt{LT}, etc. In our version, these type constructors are called \texttt{Plus\_}, 
-\texttt{Mult\_}, \texttt{LT\_}, etc., while the names without final underscores are bound to term building functions (\emph{e.g.}
-the function which takes two terms and builds the term representing their sum) (see Remark~\ref{rem:underscores}).
-\end{remark}
-
-\subsubsection{Language of Natural Numbers (LNN)}
-
-\texttt{LNN} is an extension of \texttt{LNT}, by the addition 
-of the $<$ relation symbol.
-
-\inputsnippets{Languages/LNNDef}
-
-\subsubsection{Examples}
-
-Let us show a few examples (from ~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}).
-
-\inputsnippets{FolExamples/arityTest}
-
-% \subsubsection{Examples}
-
-% With our toy language, we can build the following terms, respectively written $a$, $f(a)$ and $f(f(v_1))$ in usual mathematical notation.
-
-% Writing these terms in bare \gallina syntax would result in quite clumsy scripts. Thus, we may use a still experimental notation scope which allows us to write/print terms and formulas in a 
-% notation similar to \coq's.
-
-% We put both the associated \coq{} terms, in bare \gallina{} syntax, and with a still experimental and provisional notation, defined in
-% ~\href{../theories/html/hydras.Ackerm.FOL_notations.html}{MoreAck.FOL\_notations}).
-
-% \inputsnippets{FolExamples/toyNotation}
-
-% \inputsnippets{FolExamples/smallTerms}
-
-For instance the term $v_1+0$, where $v_1$ is a variable,
-is represented by the following \gallina term of type 
-(\texttt{fol.Term LNT}).
-
-\inputsnippets{FolExamples/v1Plus0}
-
-\inputsnippets{LNN/instantiations}
-
-\section{Notations for Formulas (experimental)}
-
-In order to get more readable terms and formulas, we can define a few notations in ~\href{../theories/html/hydras.MoreAck.FOL_notations.html}{MoreAck.FOL\_notations} and
-~\href{../theories/html/hydras.MoreAck.LNN.html}{MoreAck.LNN}.
-Please note that these notation scopes are experimental: We are going to use them in examples and exercises before using them in large original proof scripts (in the \texttt{ordinals/Ackermann/} sub-directory).
-
-We try to define notation scopes as close as possible to \coq's syntax for propositions.
-
-Let us take for instance the following proposition (in math form):
-
-$$\forall\, v_0,\, v_0=0\vee \exists\,v_1,\,v_1=1+v_0$$
-
-Here is a definition, using directly the \texttt{goedel/Ackermann}'s project syntax.
-
-\inputsnippets{LNN_Examples/uglyF0}
-
-Note that, because of redefinitions, the disjonction \texttt{orH}
-can be expanded in terms of  implication and negation (for instance when we use \texttt{Compute}).
-
-\begin{todo}
-  Present the general issue about evaluation, and our provisional solution.
-\end{todo}
-
-\inputsnippets{LNN_Examples/CNNF0}

### 2023-07-03 doc/chapter-lnn-lnt.tex
improve chapter-fol.tex

\chapter{Languages for Arithmetic (in construction)}

\begin{todo}
Make a chapter!
\end{todo}

Two languages built with the usual symbols of arithmetic are 
defined in ~\href{../theories/html/hydras.Ackermann.Languages.html}{Ackermann.Languages}.

\begin{itemize}
\item The first language: \texttt{LNT} (\emph{Language of Number Theory}) has just function symbols for $+$, $\times$, $0$ and successor.
\item The second language: LNN (\emph{Language of Natural Numbers})  has
the same function symbols as LNT plus one relation symbol for the strict inequality $<$ : \texttt{LT} (less than).
\end{itemize}

\subsubsection{Language of Number Theory (LNT)}

First, we declare two alphabets.

\inputsnippets{Languages/LNTDef1}

In a second time, we build \texttt{LNT} and \texttt{LNN} by filling \texttt{Language}'s \texttt{arity} field.

\inputsnippets{Languages/LNTDef2}

\begin{remark}
  We depart a little from \cite{Goedel}'s notations, where the 
function and relation symbols are called \texttt{Plus}, 
\texttt{Mult}, \texttt{LT}, etc. In our version, these type constructors are called \texttt{Plus\_}, 
\texttt{Mult\_}, \texttt{LT\_}, etc., while the names without final underscores are bound to term building functions (\emph{e.g.}
the function which takes two terms and builds the term representing their sum) (see Remark~\ref{rem:underscores}).
\end{remark}

\subsubsection{Language of Natural Numbers (LNN)}

\texttt{LNN} is an extension of \texttt{LNT}, by the addition 
of the $<$ relation symbol.

\inputsnippets{Languages/LNNDef}

\subsubsection{Examples}

Let us show a few examples (from ~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}).

\inputsnippets{FolExamples/arityTest}

% \subsubsection{Examples}

% With our toy language, we can build the following terms, respectively written $a$, $f(a)$ and $f(f(v_1))$ in usual mathematical notation.

% Writing these terms in bare \gallina syntax would result in quite clumsy scripts. Thus, we may use a still experimental notation scope which allows us to write/print terms and formulas in a 
% notation similar to \coq's.

% We put both the associated \coq{} terms, in bare \gallina{} syntax, and with a still experimental and provisional notation, defined in
% ~\href{../theories/html/hydras.Ackerm.FOL_notations.html}{MoreAck.FOL\_notations}).

% \inputsnippets{FolExamples/toyNotation}

% \inputsnippets{FolExamples/smallTerms}

For instance the term $v_1+0$, where $v_1$ is a variable,
is represented by the following \gallina term of type 
(\texttt{fol.Term LNT}).

\inputsnippets{FolExamples/v1Plus0}

\inputsnippets{LNN/instantiations}

\section{Notations for Formulas (experimental)}

In order to get more readable terms and formulas, we can define a few notations in ~\href{../theories/html/hydras.MoreAck.FOL_notations.html}{MoreAck.FOL\_notations} and
~\href{../theories/html/hydras.MoreAck.LNN.html}{MoreAck.LNN}.
Please note that these notation scopes are experimental: We are going to use them in examples and exercises before using them in large original proof scripts (in the \texttt{ordinals/Ackermann/} sub-directory).

We try to define notation scopes as close as possible to \coq's syntax for propositions.

Let us take for instance the following proposition (in math form):

$$\forall\, v_0,\, v_0=0\vee \exists\,v_1,\,v_1=1+v_0$$

Here is a definition, using directly the \texttt{goedel/Ackermann}'s project syntax.

\inputsnippets{LNN_Examples/uglyF0}

Note that, because of redefinitions, the disjonction \texttt{orH}
can be expanded in terms of  implication and negation (for instance when we use \texttt{Compute}).

\begin{todo}
  Present the general issue about evaluation, and our provisional solution.
\end{todo}

\inputsnippets{LNN_Examples/CNNF0}

### 2023-07-03 doc/chapter-natural-deduction.tex
improve chapter-fol.tex

\chapter{Natural Deduction (in construction)}

%%%%%%%%%%%%%% New chapter 
\subsection{Contexts as sets}

\begin{exercise}

Prove the following lemma:
\inputsnippets{FolExamples/MPdiag}

\end{exercise}

\section{Derived rules and natural deduction}

 The library 
 \href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic} contains many derived rules which allow the user to build proofs in a natural deduction style (with introduction and elimination rules).

\subsection{Example}

For instance, let us prove Peirce's rule.
 (in  ~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples} ).

\subsubsection{Prelude}

%\inputsnippets{folPeirce/prelude}

\subsection{Proof of Peirce's law}

\inputsnippets{FolExamples/PeirceProof}

Let us try to do an implication introduction.

\inputsnippets{FolExamples/step1}

Now, we may use the law of excluded middle with the formula $P$. The only non-trivial case is about $\sim P$.

\inputsnippets{FolExamples/step2}

The rest of the proof is composed of basic proof steps, 
and bookkeeping steps (about \texttt{Ensembles.In}).

\inputsnippets{FolExamples/step3, FolExamples/step4}

\index{ackermann}{Exercises}
\begin{exercise}
Prove, using the rules described in 
 \href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic}, the famous \emph{drinkers theorem}:

$$\exists\,x,\; (D(x)\Longrightarrow \forall\,y,\; D(y))$$
\end{exercise}

where $D$ (for ``drinks'') is some predicate symbol of arity $1$.

### 2023-07-07 doc/chapter-natural-deduction.tex
use latex package turnstile

@@
 %%%%%%%%%%%%%% New chapter 
-\subsection{Contexts as sets}
+\section{Contexts as sets}
+
+Let us look again at the proof scripts shown in~\vref{proof:PrfEx2} and \vref{proof:PrfEx3}.
+\begin{itemize}
+\item The statement of \texttt{PrfEx2} contains a sequence
+of hypotheses with two occurrences of \texttt{A}. Moreover, the order in which the 4 hypotheses are listed is determined by the type of the constructor \texttt{MP}  (please look at
+Figure~\vref{fig:prf-def}). It would be better to replace this
+precise list of hypotheses with ``any list whose elements
+belong to the set $\{A, A \arrow B; A \arrow B \arrow C\}$''.
+
+\item In the proof of \texttt{PrfEx3} the \coq user would certainly ask ``How do we apply implication's introduction rule?''.
+\end{itemize}
+
+The answer to both questions in~\cite{Goedel}  is the definition of a proof system, derived from \texttt{Prf}, which considers
+\emph{sets of hypotheses} (called \emph{systems} in ~\cite{Goedel}) instead of
+\emph{list of hypotheses}, thus making abstraction of
+repetition and order of appearance of hypotheses in the context.
+
+The new system is simply defined as below:
+\inputsnippets{folProof/SysPrfDef}
+
+\subsubsection{Notations}
+
+\begin{itemize}
+\item  $\Gamma \sdash A$ for
+  (\texttt{SysPrf $L$ $\Gamma$ $A$}).
+\item $\Gamma, \Delta$ for $\Gamma\cup\Delta$
+\item $\Gamma, A,\dots B$ for $\Gamma\cup\{A \dots B\}$
+\end{itemize}
+
+\subsubsection{Examples}
+
+In the following script, we use \texttt{PrfEx2}
+as a witness for proving a set-based version of the 
+original proof term.
+
+$$A \sststile{\mathrm{S}}{} A$$
+
+\inputsnippets{folExamples/SysPrfEx2}
 
@@
 Prove the following lemma:
-\inputsnippets{FolExamples/MPdiag}
+\inputsnippets{FolExamples/MPSys}
 
 \end{exercise}
+
+\subsection{Using properties of sets}
+
+The following three lemmas, from 
+\href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic} are direct consequences of \texttt{SysPrf}'s definition.
+
+\inputsnippets{folLogic/Axm}
+
+\inputsnippets{folLogic/SysExtend}
+
+\section{The deduction theorem}
+
+\inputsnippets{Deduction/deductionThm}
 

### 2023-07-11 doc/chapter-fol.tex
use latex package turnstile
work on chapter-natural-deduction

@@
 
+We will also use (in the text) the notation $l \hdash{} F$ for the type \texttt{Prf $l$ $F$}.
+
 In the rest of this section, we comment  every one of 
@@
 \url{https://en.wikipedia.org/wiki/List_of_Hilbert_systems}.
+
+\subsubsection{Notation}
 
@@
 
+\label{proof:PrfEx2}
 \inputsnippets{FolExamples/PrfEx2}
@@
 
+\label{proof:PrfEx3}
 \inputsnippets{FolExamples/PrfEx3}
@@
 \texttt{SysPrf}, based on \texttt{Prf}, the properties of which will 
-allow us to prove theorems in a very simpler way.
+allow us to prove theorems in a much simpler way.
 

### 2023-07-11 doc/chapter-natural-deduction.tex
work on chapter-natural-deduction

@@
 \emph{list of hypotheses}, thus making abstraction of
-repetition and order of appearance of hypotheses in the context.
+the repetition and order of appearance of hypotheses in the context.
 
-The new system is simply defined as below:
+The new system is simply defined as below (for  a given language $L$)\,\footnote{In some shown snippets, arguments like $L$ may be or not be implicit (depending on the section they are extracted from).
+Please look at the \coq source.}.
+
 \inputsnippets{folProof/SysPrfDef}
@@
 
+In the text, we may use the abbreviation $\Gamma \sdash A$ for
+(\texttt{SysPrf $L$ $\Gamma$ $A$}) and
+$\sdash A$ for
+(\texttt{SysPrf $L$ Empty\_set $A$})
+
+We may also omit obvious braces in some set expressions:
+
 \begin{itemize}
-\item  $\Gamma \sdash A$ for
-  (\texttt{SysPrf $L$ $\Gamma$ $A$}).
-\item $\Gamma, \Delta$ for $\Gamma\cup\Delta$
+
+\item $\Gamma, \Delta$ for $\Gamma\cup\Delta$,
+\item $A, B, C$ for $\{A, B, C\}$,
 \item $\Gamma, A,\dots B$ for $\Gamma\cup\{A \dots B\}$
+\item \emph{etc.}
 \end{itemize}
@@
 as a witness for proving a set-based version of the 
-original proof term.
-
-$$A \sststile{\mathrm{S}}{} A$$
+original proof term, namely $A; A \arrow B; A \arrow B \arrow C \sdash C$.
 
@@
 
-Prove the following lemma:
+Prove the following lemma (without the handy lemmas from
+\href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic} and their corollaries!).
+
 \inputsnippets{FolExamples/MPSys}
@@
 
-\section{The deduction theorem}
+\section{The Deduction theorem}
+
+The deduction theorem (proved in \href{../theories/html/hydras.Ackermann.Deduction.html}{Ackermann.Deduction} is a handy 
+tool for proving a proposition $A\arrow B$ by pushing 
+the hypothesis $A$ into the context (it corresponds to 
+implication introduction in \coq).
 
 \inputsnippets{Deduction/deductionThm}
+
+\subsection{sketch of proof}
+
+We advise the reader to replay this proof on h.er.is computer
+ in order to better understand its structure.
+
+Let us give only a few comments.
+
+Let us assume the hypothesis \texttt{H: $T, g \sdash f$},
+meaning that there exists some list $F$ whose elements 
+belong to $T \cup \{g\}$, and a proof-tree $t$ of type
+$F \vdash h$.
+
+The heart of the proof is an induction on $t$ proving 
+$F \cap T \sdash (g\arrow f)$ \footnote{Please forgive the implicit coercion from lists to sets!}.
+
+Please note that the case $S=\texttt{nil}$ is common to 
+many constructors of \texttt{Prf}, thus the proof script starts
+with a study of this particular case, simply applied 11 times in the rest of the proof.
 
@@
 
-\subsection{Example}
+\subsection{Example: proof of Peirce's law}
 
-For instance, let us prove Peirce's rule.
- (in  ~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples} ).
+For instance, let us prove Peirce's rule, \emph{i.e.}
+$\sdash ((A \arrow B) \arrow A) \arrow A$ for any formulas $A$ and $B$.
 
-\subsubsection{Prelude}
-
-%\inputsnippets{folPeirce/prelude}
-
-\subsection{Proof of Peirce's law}
+ The proof in \coq is available in  ~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}.
 
@@
 
-Let us try to do an implication introduction.
+Let us start with  an implication introduction.
 
@@
 
-Now, we may use the law of excluded middle with the formula $P$. The only non-trivial case is about $\sim P$.
+Now, we may eliminate the disjunction 
+in the instance $\sim A \vee A$ of the law of excluded middle.
+The only non-trivial case is about $\sim A$.
 

### 2023-07-12 doc/chapter-natural-deduction.tex
fix a case error (typically MacOS)
work on chapter-natural-deduction

@@
 
-\inputsnippets{folExamples/SysPrfEx2}
+\inputsnippets{FolExamples/SysPrfEx2}
 
@@
 We advise the reader to replay this proof on h.er.is computer
- in order to better understand its structure.
-
-Let us give only a few comments.
+ in order to better understand its structure, which we will only comment briefly.
 
@@
 with a study of this particular case, simply applied 11 times in the rest of the proof.
+
+ \index{ackermann}{Projects}
+
+\begin{project}[**]
+  The type \texttt{SysPrf} has sort \texttt{Prop}, which prevent us from
+extracting the underlying proof tree and its fringe from an hypothesis 
+of the form $X\sdash f$. 
+It would be nice (\emph{e.g.} for a better understanding of the proof of the deduction theorem) to be able to compute the the proof-tree built
+by the proof of the deduction lemma. 
+\end{project}
 

### 2023-07-12 doc/chapter-natural-deduction.tex
work on chapter-natural-deduction

@@
 
-\subsection{sketch of proof}
+\subsection{Sketch of proof}
 
@@
 of the form $X\sdash f$. 
-It would be nice (\emph{e.g.} for a better understanding of the proof of the deduction theorem) to be able to compute the the proof-tree built
+It would be nice (\emph{e.g.} for a better understanding of the proof of the deduction theorem) to be able to compute the  proof-tree built
 by the proof of the deduction lemma. 
+On a fresh branch of the project, please change the definition of
+\texttt{SysPrf} and fix the errors this change could cause in the rest of the files. 
 \end{project}

### 2023-07-12 doc/chapter-natural-deduction.tex
work on chapter-natural-deduction

@@
 
+The \emph{rule of implication elimination} is derived from \texttt{Prf}'s 
+\emph{modus ponens} \texttt{MP}.  Since all elements of the fringe 
+\texttt{x} [resp. \texttt{x1}] of the proof tree \texttt{px} [resp. \texttt{px1}] belong to T, so are the elements of the fringe \texttt{x++x1} of
+(\texttt{MP \dots{}  px px1}).
+
+\inputsnippets{folLogic/impE}
+
 \section{The Deduction theorem}
@@
 \texttt{SysPrf} and fix the errors this change could cause in the rest of the files. 
+\emph{If ensuring compatibility with all the \texttt{Goedel} project is too long and/or difficult, you may just make the changes in separate modules with an ``informative'' \texttt{SysPrf} and limit the compatibility study  to the contents of basic modules like
+\texttt{Deduction}, \texttt{folLogic}, etc.
+Perhaps you will have to consider other implementation of finite sets of formulas (\emph{e.g.} lists).
+}
+
 \end{project}
@@
  \href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic} contains many derived rules which allow the user to build proofs in a natural deduction style (with introduction and elimination rules).
+
+%\subsection{Applications of the Deduction Theorem}
+
+The \emph{rule of implication introduction} is a trivial application of the deduction theorem.
+
+\inputsnippets{folLogic/impI}
 

### 2023-07-12 doc/chapter-natural-deduction.tex
work on chapter-natural-deduction

@@
 
+The following lemma (corresponding to the \texttt{CP} constructor of type \texttt{Prf}) is proven with the help of \texttt{impE}, \texttt{impI}, 
+\texttt{sysWeaken} and \texttt{CP}.
+
+\inputsnippets{folLogic/contradiction}
+
+We let the reader consult the proof of the following lemmas, or, much better, re-prove them as exercises.
+
+\inputsnippets{folLogic/nnE}
+
+\inputsnippets{ folLogic/nnI}
+
+\inputsnippets{folLogic/cp1}
+
+\inputsnippets{folLogic/cp2}
+
+\subsection{Rules for derived connectives and quantifiers}
+
+Let us keep in mind that the derived connectives: \texttt{orH}, \texttt{andH}, etc. and the existential quantifier \texttt{existH}  are defined
+in terms of \texttt{impH}, \texttt{notH}, \texttt{forallH}.
+By unfolding these definitions, we prove easily a few natural deduction rules for the derived symbols. 
+For instance, the law of excluded middle for $f$: $ T\sdash \sim f \vee f$
+for any $T$,  is just an abbreviation of $T \sdash \sim\sim f \arrow f$.
+
+\inputsnippets{folLogic/noMiddle}
+
+\inputsnippets{folLogic/orI1}
+
+\inputsnippets{folLogic/orI2}
+
 \subsection{Example: proof of Peirce's law}

### 2023-07-15 doc/chapter-natural-deduction.tex
work on chapter-natural-deduction

@@
 For instance, the law of excluded middle for $f$: $ T\sdash \sim f \vee f$
-for any $T$,  is just an abbreviation of $T \sdash \sim\sim f \arrow f$.
+for any $T$,  is just an abbreviation of $T \sdash \sim\sim{}f \arrow f$.
 
 \inputsnippets{folLogic/noMiddle}
+
+\begin{remark}
+  The lines marked ``optional'' are just here in order to
+temporarily deactivate the notation which print any formula 
+of the form $\sim{}\!A \arrow B$ as \texttt{$A$ \disj' $B$}.
+ These three lines can be safely removed.
+\end{remark}
+
+Let us go on \dots.
+We leave the completion of the proofs as an exercise.
+We encourage the reader to write h.er.is own proofs, then compare them with the scripts of \href{../theories/html/hydras.Ackermann.folLogic.html}{folLogic}.
 
@@
 \inputsnippets{folLogic/orI2}
+
+\inputsnippets{folLogic/orE}
+
+\inputsnippets{folLogic/orSys}
 

### 2023-07-15 doc/chapter-fol.tex
work on chapter-natural-deduction

@@
 
-We can make this proof script shorter, using existential variables.
+\begin{figure}[h]
+  \centering
+  \[
+  \infer{C}{ 
+    \infer{B \arrow C}
+    {\infer{ A \arrow B \arrow C}
+      { A \arrow B \arrow C}
+      &
+      \infer{A}{A}}
+    & \infer{B}{
+      \infer{A\arrow B}{A\arrow B} &
+      \infer{A}{A}}
+                     }
+\]
+  \caption{The proof tree of \texttt{PrfEx2}}
+  \label{fig:PrfEx2}
+\end{figure}
+
+Figure~\ref{fig:PrfEx2} shows the tree-like structure of 
+\texttt{PrfEx2}. The list of used hypotheses is the fringe of the tree. The unary nodes are applications of \texttt{AXM} and the binary nodes are associated with \texttt{MP}.
+
+\vspace{4pt}
+
+We can make our  proof script shorter, using existential variables.
 

### 2023-07-30 doc/chapter-natural-deduction.tex
corrections chap 12 and 13
corrections chap 13

@@
 
+\begin{remark}
+  The type \texttt{SysPrf} has sort \texttt{Prop}, which prevent us from
+extracting the underlying proof tree and its fringe from a proof of $X\sdash f$. We only know that such a proof exists, but cannot get it automatically through a \coq function.
+\end{remark}
+
+\index{ackermann}{Projects}
+
+\begin{project}
+It would be nice (\emph{e.g.} for a better understanding of the proof of the deduction theorem) to be able to compute the  proof-tree built
+by the proof of the deduction lemma. 
+On a fresh branch of the project, please change the definition of
+\texttt{SysPrf} and fix the errors this change could cause in the rest of the files. 
+\emph{If ensuring compatibility with all the \texttt{Goedel} project is too long and/or difficult, you may just make the changes in separate modules with an ``informative'' \texttt{SysPrf} and limit the compatibility study  to the contents of basic modules like
+\texttt{Deduction}, \texttt{folLogic}, etc.
+Perhaps you will have to consider other implementation of finite sets of formulas (\emph{e.g.} lists).
+}
+
+\end{project}
+
 \subsubsection{Examples}
@@
 The deduction theorem (proved in \href{../theories/html/hydras.Ackermann.Deduction.html}{Ackermann.Deduction} is a handy 
-tool for proving a proposition $A\arrow B$ by pushing 
-the hypothesis $A$ into the context (it corresponds to 
-implication introduction in \coq).
+tool for proving a proposition $f\arrow g$ by pushing 
+the hypothesis $f$ into the context (it corresponds roughly to 
+the implication introduction rule in \coq).
 
@@
 with a study of this particular case, simply applied 11 times in the rest of the proof.
-
- \index{ackermann}{Projects}
-
-\begin{project}[**]
-  The type \texttt{SysPrf} has sort \texttt{Prop}, which prevent us from
-extracting the underlying proof tree and its fringe from an hypothesis 
-of the form $X\sdash f$. 
-It would be nice (\emph{e.g.} for a better understanding of the proof of the deduction theorem) to be able to compute the  proof-tree built
-by the proof of the deduction lemma. 
-On a fresh branch of the project, please change the definition of
-\texttt{SysPrf} and fix the errors this change could cause in the rest of the files. 
-\emph{If ensuring compatibility with all the \texttt{Goedel} project is too long and/or difficult, you may just make the changes in separate modules with an ``informative'' \texttt{SysPrf} and limit the compatibility study  to the contents of basic modules like
-\texttt{Deduction}, \texttt{folLogic}, etc.
-Perhaps you will have to consider other implementation of finite sets of formulas (\emph{e.g.} lists).
-}
-
-\end{project}
 
@@
 
+\inputsnippets{folLogic/andI}
+
+\inputsnippets{folLogic/andE1, folLogic/andE2}
+
+\inputsnippets{FolExamples/SearchSysPrf}
+
 \subsection{Example: proof of Peirce's law}
@@
 
-Let us start with  an implication introduction.
+Let us start with  an implication introduction. The judgement to prove becomes  $(A \arrow B) \arrow A \sdash A$.
 
@@
 
+\vspace{4pt}
+
+Next goal is $(A \arrow B) \arrow A \sdash \sim\!A \arrow A$.
+Please keep in mind  that our current notation system interprets
+$\sim\!A \arrow A$ as a disjunction!
+
 \inputsnippets{FolExamples/step2}
 
-The rest of the proof is composed of basic proof steps, 
-and bookkeeping steps (about \texttt{Ensembles.In}).
+The rest of the proof is composed of basic proof steps.
 

### 2023-08-10 doc/chap-intro-goedel.tex
corrections chap 10

@@
 \section{Introduction}
+This part contains comments, examples  and exercises about Russel O'Connor's work on G\"{o}del's first incompleteness theorem  \cite{Godel1986-GDECW}.
+O'Connor's work was published in 2005~\cite{OConnor05}, and released as a user-contribution of the \coq proof assistant.
+This work is now maintained by Coq community~\cite{CoqCommunity} volunteers and split into two projects: Goedel~\cite{Goedel} and Hydra-battles~\cite{HydraBattles}.
+
+The main reference to this work is  Russel O'Connor's article~\cite{OConnor05}, which we strongly encourage the reader to 
+consult regularly.
 
@@
 
-  \item O'Connor's proof was written at the end of \coq V7, then rewritten at the beginning of \coq V8. Since then, \coq and its ecosystem evolved a lot (new styles, tactics, documentation tools).
+  \item O'Connor's proof was written at the end of \coq V7, then rewritten at the beginning of \coq V8.
+Since then, \coq and its ecosystem evolved a lot (new styles, tactics, documentation tools, etc.).
           We think this evolution should benefit
@@
 
-        \item Finally, let us quote  Efim Zelmanov in~\cite{mathproof}
+        \item Finally, let us quote  Efim Zelmanov.
              \begin{quote}
-      The purpose of a proof is \emph{understanding}.
+      ``The purpose of a proof is \emph{understanding}'' ~\cite{mathproof}
 
@@
 
-        For technical reasons --- mainly in order to make easier the installation and use of its sub-libraries, we split the project
+        For technical reasons --- mainly in order to make simpler the installation and use of its sub-libraries, we split the project
 into two main parts: Goedel~\cite{Goedel} and Hydra-battles~\cite{HydraBattles}.
@@
 jobs have just started, and will probably be long to complete. Help is welcome!}
-
-This part contains comments, examples  and exercises about Russel O'Connor's work on G\"{o}del's first incompleteness theorem  \cite{Godel1986-GDECW}.
-O'Connor's work was published in 2005~\cite{OConnor05}, and released as a user-contribution of the \coq proof assistant.
-This work is now maintained by Coq community~\cite{CoqCommunity} volunteers and split into two projects: Goedel~\cite{Goedel} and Hydra-battles~\cite{HydraBattles}.
-
-The main reference to this work is  Russel O'Connor's article~\cite{OConnor05}, which we strongly encourage the reader to 
-consult regularly.
 
@@
  definition of primitive recursive functions, first-order  logic,
-Peano Arithmetics, G\"{o}del's encoding.
+Peano Arithmetic, G\"{o}del's encoding.
 \item Some additions we made: examples, exercices, new notations, etc.,  are stored in a specific directory \texttt{theories/ordinal/MoreAck/}.
 \end{itemize}
+
+\subsection{The \texttt{Ackermann} sub-directory}
+
+The following main topics are studied in  \texttt{theories/ordinals/Ackermann/}: 
+The following list presents the main modules 
+in a dependance-compatible order.
+
+\begin{description}
+
+ \item[Primitive Recursive Functions]
+  \mbox{}
+
+ This topic is discussed in Chapter~\vref{chapter:primrec}.
+
+  \begin{itemize}
+\item
+\href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat}
+  \item 
+  \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}
+ \item 
+  \href{../theories/html/hydras.Ackermann.cPair.html}{Ackermann.cPair}
+
+\item 
+  \href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}  
+ \item 
+  \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack}
+
+\item
+  \href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR}
+
+Ackermann function is not primitive recursive
+
+  \end{itemize}
+
+\item[First Order Logic]
+  \mbox{}
+
+This part presents \coq definitions and basic properties of first-order languages and proofs.
+
+  \begin{itemize}
+  \item \href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol}
+ \item \href{../theories/html/hydras.Ackermann.folProp.html}{Ackermann.folProp}
+\item \href{../theories/html/hydras.Ackermann.folProof.html}{Ackermann.folProof}
+\item \href{../theories/html/hydras.Ackermann.model.html}{Ackermann.model}
+\item \href{../theories/html/hydras.Ackermann.code.html}{Ackermann.code}
+ \item 
+  \href{../theories/html/hydras.Ackermann.prLogic.html}{Ackermann.prLogic}
+
+\item \href{../theories/html/hydras.Ackermann.codeList.html}{Ackermann.codeList}
+\item \href{../theories/html/hydras.Ackermann.codeFreeVar.html}{Ackermann.codeFreeVar}
+
+\item \href{../theories/html/hydras.Ackermann.checkPrf.html}{Ackermann.checkPrf}
+\item \href{../theories/html/hydras.Ackermann.wellFormed.html}{Ackermann.wellFormed}
+\item \href{../theories/html/hydras.Ackermann.codeSubTerm.html}{Ackermann.codeSubTerm}
+\item \href{../theories/html/hydras.Ackermann.codeSubFormula.html}{Ackermann.codeSubFormula}
+
+  \end{itemize}
+
+\item[Natural Deduction]
+ \mbox{}
+Thanks to the \emph{Deduction Theorem}, we prove many 
+lemmas about provability in first-order-logic, many of them can be considered as natural deduction rules.
+
+\begin{itemize}
+\item \href{../theories/html/hydras.Ackermann.Deduction.html}{Ackermann.Deduction}
+\item \href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic}
+\item \href{../theories/html/hydras.Ackermann.folLogic2.html}{Ackermann.folLogic2}
+\item \href{../theories/html/hydras.Ackermann.folLogic3.html}{Ackermann.folLogic3}
+\item \href{../theories/html/hydras.Ackermann.folReplace.html}{Ackermann.folReplace}
+\item \href{../theories/html/hydras.Ackermann.subProp.html}{Ackermann.subProp}
+\item \href{../theories/html/hydras.Ackermann.subAll.html}{Ackermann.subAll}
+ \item  \href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}
+\end{itemize}
+
+\item [Languages of Arithmetic]
+\mbox{}
+
+\begin{itemize}
+\item \href{../theories/html/hydras.Ackermann.Languages.html}{Ackermann.Languages}
+\item \href{../theories/html/hydras.Ackermann.LNN.html}{Ackermann.LNN}
+\item \href{../theories/html/hydras.Ackermann.LNT.html}{Ackermann.LNT}
+\item \href{../theories/html/hydras.Ackermann.NN.html}{Ackermann.NN} Axioms for Natural Numbers and basic properties.
+\item \href{../theories/html/hydras.Ackermann.NNtheory.html}{Ackermann.NNtheory}
+\item \href{../theories/html/hydras.Ackermann.PA.html}{Ackermann.PA} Peano Arithmetic: axioms  and first properties.
+\item \href{../theories/html/hydras.Ackermann.LNN2LNT.html}{Ackermann.LNN2LNT}
+\item \href{../theories/html/hydras.Ackermann.NN2PA.html}{Ackermann.NN2PA}
+
+\item \href{../theories/html/hydras.Ackermann.PAtheory.html}{Ackermann.PAtheory}
+\item \href{../theories/html/hydras.Ackermann.PAconsistent.html}{Ackermann.PAconsistent}
+
+\item \href{../theories/html/hydras.Ackermann.codePA.html}{Ackermann.codePA}
+
+\item \href{../theories/html/hydras.Ackermann.codeNatToTerm.html}{Ackermann.codeNat2Term}
+\item \href{../theories/html/hydras.Ackermann.wConsistent.html}{Ackermann.wConsistent}
+\item \href{../theories/html/hydras.Ackermann.expressible.html}{Ackermann.expressible}
+
+\item \href{../theories/html/hydras.MoreAck.LNN_Examples.html}{MoreAck.LNN\_Examples}
+\end{itemize}
+
+\item[Modules dependent on \texttt{CoqPrime}]
+
+\mbox{}
+
+  \begin{itemize}
+  \item \href{../theories/html/Goedel.PRrepresentable.html}{Goedel.PRrepresentable}
+ \item \href{../theories/html/Goedel.fixPoint.html}{Goedel.fixPoint}
+ \item \href{../theories/html/Goedel.codeSysPrf.html}{Goedel.codeSysPrf}
+ \item \href{../theories/html/Goedel.rosser.html}{Goedel.rosser}
+ \item \href{../theories/html/Goedel.goedel1.html}{Goedel.goedel1}
+ \item \href{../theories/html/Goedel.rosserPA.html}{Goedel.rosserPA}
+ \item \href{../theories/html/Goedel.goedel2.html}{Goedel.goedel2}
+  \end{itemize}
+\end{description}
 

### 2023-08-28 doc/chapter-fol.tex

@@
 \subsubsection{Rules about the universal quantifier: \texttt{FA1},
-\texttt{FA2} and\texttt{FA3}}
+\texttt{FA2} and \texttt{FA3}}
 
@@
 The following proof applies \texttt{FA1} in order to build a proof
-of \texttt{Prf L [] (t=t)\%fol}  for any $t$\,\footnote{Please note that the quantification on $f$ is \coq's. Thus, \texttt{eq\_ref} is kind of scheme.}
+of \texttt{Prf L [] (t=t)\%fol}  for any $t$\,\footnote{Please note that the quantification on $t$ is at the meta-level (\coq's level;
+not \texttt{FOL}).}.
+
+Please look also at Figure~\vref{fig:fol-eq-refl}
 
 \inputsnippets{FolExamples/eqRefl}
+
+\begin{figure}[h]
+  \centering
+
+\[
+  \infer{t=t}{
+    \infer[FA1]{(\forall v_0, v_0 = v_0) \,\arrow\, t=t}{}
+    &
+    \infer[GEN]{\forall v_0, v_0 = v_0}{
+      \infer[EQ1]{v_0 = v_0}{}
+      }
+    }
+\]
+
+ \caption{Proof tree of \texttt{(eq\_refl $t$)}}
+\label{fig:fol-eq-refl}
+\end{figure}
 

### 2023-08-28 doc/chapter-natural-deduction.tex

@@
 
+In a few words, proving a statement \texttt{SysPrf \_ $T$ $A$} 
+is proving the existence of a proof-tree of type \texttt{Prf \textit{Hyp}  $A$}, where \textit{Hyp} is a list of hypotheses all elements of which belong to $T$.
+
 \subsubsection{Notations}
 
-In the text, we may use the abbreviation $\Gamma \sdash A$ for
-(\texttt{SysPrf $L$ $\Gamma$ $A$}) and
+In the text, we may use the abbreviation $T \sdash A$ for
+(\texttt{SysPrf $L$ $T$ $A$}) and
 $\sdash A$ for
@@
 
-\item $\Gamma, \Delta$ for $\Gamma\cup\Delta$,
+\item $T,U$ for $T \cup U$,
 \item $A, B, C$ for $\{A, B, C\}$,
-\item $\Gamma, A,\dots B$ for $\Gamma\cup\{A \dots B\}$
+\item $T, A,\dots B$ for $T \cup\{A \dots B\}$
 \item \emph{etc.}
@@
   The type \texttt{SysPrf} has sort \texttt{Prop}, which prevent us from
-extracting the underlying proof tree and its fringe from a proof of $X\sdash f$. We only know that such a proof exists, but cannot get it automatically through a \coq function.
+extracting the underlying proof tree and its fringe from a proof of $T\sdash f$. We only know that such a proof exists, but cannot get it automatically through a \coq function.
 \end{remark}
@@
 
-\subsubsection{Examples}
+\subsubsection{Example}
 
@@
  The library 
- \href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic} contains many derived rules which allow the user to build proofs in a natural deduction style (with introduction and elimination rules).
+ \href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic},
+\href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic2} and
+\href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic3} contain many derived rules which allow the user to build proofs in a natural deduction style (with introduction and elimination rules).
+
+We present here only a few examples of these rules, the reader
+may consult these libraries with \texttt{Search} or by looking 
+at the \texttt{coqdoc} generated files. A meta-exercise would be to re-prove a few of these lemmas and/or build an example of application.
 
@@
 
-Let us go on \dots.
-We leave the completion of the proofs as an exercise.
-We encourage the reader to write h.er.is own proofs, then compare them with the scripts of \href{../theories/html/hydras.Ackermann.folLogic.html}{folLogic}.
-
 \inputsnippets{folLogic/orI1}
-
-\inputsnippets{folLogic/orI2}
 
@@
 
-\inputsnippets{folLogic/andE1, folLogic/andE2}
+%\inputsnippets{folLogic/andE1, folLogic/andE2}
+
+Here are a few examples of looking for rules using \coq's \texttt{Search} command:
 

### 2023-09-04 doc/hydras.tex
fix a latex error
improve chapter-fol.tex
use latex package turnstile
work on chapter-natural-deduction
more acknowledgements

@@
 \usepackage{amsmath}
+\usepackage{turnstile}
 %\usepackage{fancyhdr}
@@
 \usepackage{texments}
+\usepackage{proof}
 
@@
 \newtheorem{gaiahydra}{\gaiasign}
+
+\newcommand {\sdash} {\sststile{\mathrm{S}}{}}
+\newcommand {\hdash} {\sststile{}{}}
 
@@
 \makeindex{ackermann}
+
 \begin{document}
@@
 Bordeaux, CNRS UMR 5800\\ \textbf{email:} pierre dot casteran arobas gmail dot com.
-\\ With contributions by Yves Bertot, Ilm\={a}rs C\={i}rulis, \'Evelyne Contejean,  Jérémy Damour, Florian Hatat, Pascal Manoury, Karl Palmskog, Clément Pit-Claudel, and Théo Zimmermann.
-\\ The formalization of primitive recursive functions and Peano Arithmetic was originally authored by Russel O'Connor.}%\cite{OConnor05}.}
+\\ With contributions by Yves Bertot, Ilm\={a}rs C\={i}rulis, \'Evelyne Contejean,  Jérémy Damour, St\'ephane Desarzens, Florian Hatat, Pascal Manoury, Karl Palmskog, Clément Pit-Claudel, and Théo Zimmermann.
+\\ The formalization of primitive recursive functions and Peano Arithmetic was originally authored by Russel O'Connor~\cite{OConnor05}.}
 
@@
 \label{sec:orgheadline5}
-    Many thanks to Yves Bertot, Ilm\={a}rs C\={i}rulis, \'Evelyne Contejean, Jéremy Damour,   Florian Hatat,  David Ilcinkas, 
+    Many thanks to Yves Bertot, Ilm\={a}rs C\={i}rulis, \'Evelyne Contejean, Jéremy Damour,   St\'ephane Desarzens, Florian Hatat,  David Ilcinkas, 
 Pascal Manoury,  Karl Palmskog, Cl\'ement Pit-Claudel, Sylvain Salvati, Alan Schmitt and Théo Zimmermann for their help on the elaboration of this library and  document, and to the
@@
 
-% \subsubsection{Links to the Coq source}
+%\subsubsection{Links to the Coq source}
 
@@
 \include{chapter-fol}
-
+\include{chapter-natural-deduction}
+\include{chapter-lnn-lnt}
 \include{chapter-encoding}

### 2023-09-13 doc/schutte-chapter.tex
fix a few Hint locality attributes
Document the use of zorns-lemma

@@
 
-We use Florian Hatat's library on countable sets, written as he was a student of  \emph{\'Ecole Normale Supérieure de Lyon}.
+Countable sets were
+
  A set $A$ is countable if there is an injective function from $A$ to $\mathbb{N}$ (see 
@@
 {\texttt{Schutte.Countable}}).
+
+This library was initially written by Florian Hatat, as he was a student of  \emph{\'Ecole Normale Supérieure de Lyon}.
+St\'ephane Desarzens adapted it in order to make it compatible with
+the \texttt{Topology/ZornsLemma} project~\cite{zornslemma}.
 

### 2023-10-13 doc/chap-intro-goedel.tex
Small wording fixes and improvements in Goedel intro chapter.

@@
 
-It was the first computer verified proof 
-the essential incompleteness of arithmetic.
+It was the first computer verified proof of
+the essential incompleteness\footnote{Todo: explain ``essential''} of Peano arithmetic.
 The main reference to this work is  Russel O'Connor's article~\cite{OConnor05}, which we strongly encourage the reader to 
@@
 
+\begin{todo}
+  Cite~\cite{Dowek2023}.
+\end{todo}
+
  \begin{itemize}
- \item Historical interest on G\"{o}del's proof and its mechanizations, as shown by the numerous litterature on this topic
+ \item Historical interest in G\"{o}del's proof and its mechanizations, as shown by the abundant litterature on this topic
    (for instance \cite{smullyan1992godel, Hofstadter1999Godel, GoedelCassou}).
@@
 
-        For technical reasons --- mainly in order to make simpler the installation and use of its sub-libraries, we split the project
+        For technical reasons --- mainly in order to simplify the installation and use of its sub-libraries, we split the project
 into two main parts: Goedel~\cite{Goedel} and Hydra-battles~\cite{HydraBattles}.
 
-Some changes are made to the aforementionned libraries, mainly because of the recent evolution of coq and its libraries.
-Nevertheless, the definitions, lemmas and theorems of the original contribution have been kept in this new release.
+Some changes are made to the aforementionned libraries, mainly because of the recent evolution of Coq and its libraries.
+Nevertheless, the definitions, lemmas and theorems of the original contribution have been preserved in this new release.
 

### 2023-10-13 doc/chapter-primrec.tex
sart a new chapter on Goedel encoding (after Russel)
work on the chapter on goedel encodings; new exercise
add a missing citation
New notations (#139)
* experiment fol notations

* first attempts

* experiment FOL notations

* experiment FOL notations

* minor change

* minor change

* minor change

* experimental notations

* experimental notations

* minor changes, bullets in folLogic

* bulletizing folLogic.v (half done, to resume at forallE)

* bulletized folLogic.v

* Example of folLogic

* more examples

* drinkers theorem in FOL (clumsy draft)

* re-structuring prro scripts in folLogic2

@@
 
-\begin{remark}
- The theory of primitive recursive function is now hosted in
-the \texttt{theories/ordinals/Ackermann} directory.
-The specific part on G\"odel's theorem,  is also on
-coq-community (\url{https://github.com/coq-community/goedel}) and requires the 
-\href{https://github.com/coq-community/pocklington}{Pocklington library} for lemmas on primality.
-\end{remark}
-
 This chapter contains some comments on Russel's library, as well as a few extensions.
@@
 
-\index{primrec}{Types!naryFunc}
+\index{ackermann}{Types!naryFunc}
 \input{movies/snippets/extEqualNat/naryFunc.tex}
@@
 
-\index{primrec}{Predicates!extEqual}
+\index{ackermann}{Predicates!extEqual}
 
@@
 
-\index{primrec}{Types!PrimRec}
-\index{primrec}{Types!PrimRecs}
+\index{ackermann}{Types!PrimRec}
+\index{ackermann}{Types!PrimRecs}
 \label{def:Primrec}
@@
 
-\index{primrec}{Functions!evalPrimRec}
-\index{primrec}{Functions!evalPrimRecs}
+\index{ackermann}{Functions!evalPrimRec}
+\index{ackermann}{Functions!evalPrimRecs}
 
@@
 
-\index{primrec}{Predicates!isPR}
+\index{ackermann}{Predicates!isPR}
 %\index{coq}{Extensionally equal functions}
@@
 
-\index{primrec}{Exercises}
+\index{ackermann}{Exercises}
 \begin{exercise}
@@
 
-\index{primrec}{Exercises}
+\index{ackermann}{Exercises}
 
@@
 
+\vspace{4pt}
+
+\emph{See also the chapter~\ref{chap:encoding} on G\"{o}del's encodings.}
+
 \end{exercise}
 
-\index{primrec}{Exercises}
+\index{ackermann}{Exercises}
 \begin{exercise}
@@
 \index{maths}{Ackermann function}
-\index{primrec}{Ackermann function}
+\index{ackermann}{Ackermann function}
 
@@
 
+\subsection{Related work}
+
+This proof is very close to the 1993 proof by Nora Szasz with the \texttt{Alf} proof assistant~\cite{Szasz93}. This proof has also been adapted
+by Lawrence C. Paulson to Isabelle/HOL~\cite{paulson_2021}.
+
 \section{The length of standard hydra battles}

### 2023-10-26 doc/chapter-primrec.tex
Nat2 term coercion (#166)
* minor changes

* Making isPR a type class (to resume)

* Converting isPR into a type class

* Making isPR a class (continued)

* IsPR is now a type class
Starting to improve the presentation of primitive recursive functions

@@
 \begin{description}
-  \item[zero] the constant function of value $0$ is primitive recursive.
+  \item[zero] the natural number $0$ is a primitive recursive function without arguments (in other words, a  \emph{constant}, or a \emph{nullary} function).
 \item[S] The successor function $S:\mathbb{N}\rightarrow\mathbb{N}$ is primitive recursive.
- \item[projections] For any pair $0< i\leq n$, the projection $\pi_{i,n}: \mathbb{N}^n\rightarrow\mathbb{N}$, defined by $\pi_{i,n}(x_1,x_2,\dots,x_{n})=x_i$, is primitive recursive.
+ \item[projections] For any $i$ and $n$, such that $0< i\leq n$, the projection $\pi_{i,n}: \mathbb{N}^n\rightarrow\mathbb{N}$, defined by $\pi_{i,n}(x_1,x_2,\dots,x_{n})=x_i$, is primitive recursive.
 \item[composition] For any $n$ and $m$, if $h: \mathbb{N}^m\rightarrow\mathbb{N}$, and
@@
 
+\subsection{A few (informal) examples}
+Before playing with primitive recursive functions in \coq{}, let us get familiar with the definition above, with the help of a few simple examples.
+\label{sect:primrec-examples1}
+
+\subsubsection{Projections}
+
+For instance, the projection $\pi{2,3}$ satisfies the equation
+$\pi{1,3}(x,y,z)=y$ for any natural numbers $x$, $y$ and $z$.
+
+\subsubsection{The constant function of value 0}
+\label{sect:k0}
+
+The \emph{unary} function which always returns $0$ may be defined through the \emph{composition} construction (with $n=1$, $m=0$, and $h=\texttt{zero}$).
+
+\subsection{Addition on natural natural numbers}
+
+Addition may be defined by primitive recursion:
+
+\begin{align}
+  0 + x_1 &:= x_1 \\
+  S x_0 + x_1 &:= S (x_0 + x_1)
+\end{align}
+
+Both equations can be rewritten as follows:
+
+\begin{align}
+0 + x_1 &:= g(x_1) \\
+S p + x_1 &:= h(p, p+x_1, x_1) \\
+\textbf{where}\; g(x_1)&:= x_1 \\
+\textbf{and}\; h(p,x, x_1) &:= S\;x 
+\end{align}
+
+It remains to show that $g$ and $h$ are primitive recursive, which is almost immediate:
+
+\begin{itemize}
+\item $g$ is the projection $\pi_{1,1}$,
+\item $h$ is the composition (with $n=3$ and $m=1$) of the function \textbf{S} and the projection $\pi_{2,3}$ .
+\end{itemize}
+
+\subsection{Multiplication on natural natural numbers}
+
+The following equations define the product of two natural numbers:
+
+\begin{align}
+  0 \times x_1 &:= 0 \\
+  S x_0 \times x_1 &:= (x_0 \times x_1) + x_1
+\end{align}
+
+This function is an instance of the primitive recursion scheme,
+with $n=1$, $g$ the constant unary function which returns $0$ (see Subsection~\vref{sect:k0}),
+ and $h$ the function defined by $h(p,x,n_1)=x+n_1$, which can be written as the composition of $+$ and the projections 
+$\pi_{2,3}$ and $\pi_{3,3}$ (The three of them  being  primitive recursive).
+
+\index{hydras}{Exercises}
+\begin{exercise}
+ Show that the factorial function is primitive recursive (by the same method as above). 
+\end{exercise}
+
+\begin{remark}
+
+The four examples above are defined in \coq in Section~\vref{sect:primrec-examples2}
+
+\end{remark}
   We present in this chapter a formalization of  primitive recursive functions, taken from  Russel O'Connor's formalization in \coq{} of
@@
 
-Another example?
-Let us consider the following term\footnote{Of course, we never typed this term \emph{verbatim}; we obtained it by an interactive proof the reader will be able to make after 
-reading Sect.\vref{sect:proofs-of-isPR}.}:
+\subsection{A few arithmetic functions}
+\label{sect:primrec-examples2}
+Let us show how the few functions presented in Section~\vref{sect:primrec-examples2} can be defined in \coq.
 
-\label{sect:bigfac}
+\subsubsection{Constant function of value 0}
 
-\input{movies/snippets/PrimRecExamples/bigPRa}
+\inputsnippets{PrimRecExamples/cst0}
 
-Let us now interpret this term as an arithmetic function.
+\subsubsection{Addition}
 
-\vspace{4pt}
-\noindent
+\inputsnippets{PrimRecExamples/addition}
 
-\input{movies/snippets/PrimRecExamples/bigPRb}
+\subsubsection{Multiplication}
 
-After this test, the term \texttt{bigPR} looks to be a primitive recursive definition of the factorial function, although we haven't proved this fact yet.
-Fortunately, we will see in the following sections simple ways to prove that a given function is primitive recursive, without building such an unreadable term.
+\inputsnippets{PrimRecExamples/multiplication}
+
+\subsubsection{The factorial function}
+
+\inputsnippets{PrimRecExamples/factorial}
 
@@
 
-\input{movies/snippets/PrimRecExamples/isPRExtEqualTrans}
+\input{movies/snippets/primRec/isPRExtEqual}
 

### 2023-11-02 doc/chapter-primrec.tex
New scopes (#167)
* UTF8 symbols for beta function

* experiment custom entries for fol propositions

* More experiments with notations (unstable)

* New notations for PR functions (to complete)

* New notations for PR functions and FOL formulas (experimental)
Minor changes (mainly notations)

@@
 
-Primitive recursive functions are total functions from $\mathbb{N}^n$ to
+\emph{Primitive recursive functions} are total functions from $\mathbb{N}^n$ to
 $\mathbb{N}$, for some $n\in\mathbb{N}$. Note that not all 
@@
 (see for instance Sect.~\vref{sect:ack-not-PR}).
+
+Primitive recursive \emph{relations} are boolean total functions whose \emph{characteristic function} --- obtained by mapping the returned value to $1$ (\texttt{true}) or $0$ (\texttt{false}) ---  is primitive recursive.
+
+\section{Mathematical definition}
+\label{sect:primrecmathdef}
 
@@
 
-For instance, the projection $\pi{2,3}$ satisfies the equation
-$\pi{1,3}(x,y,z)=y$ for any natural numbers $x$, $y$ and $z$.
+For instance, the projection $\pi_{2,3}$ satisfies the equation
+$\pi_{2,3}(x,y,z)=y$ for any natural numbers $x$, $y$ and $z$.
 
@@
 
-\subsection{Addition on natural natural numbers}
+\subsubsection{Addition on natural natural numbers}
 
@@
 
-\begin{align}
+\begin{align*}
   0 + x_1 &:= x_1 \\
   S x_0 + x_1 &:= S (x_0 + x_1)
-\end{align}
+\end{align*}
 
@@
 
-\begin{align}
-0 + x_1 &:= g(x_1) \\
-S p + x_1 &:= h(p, p+x_1, x_1) \\
-\textbf{where}\; g(x_1)&:= x_1 \\
-\textbf{and}\; h(p,x, x_1) &:= S\;x 
-\end{align}
+\begin{align*}
+&0 + x_1 := g(x_1) \\
+& S \; p +x_1 := h(p, p+x_1, x_1) \\
+&\textbf{where}\; g(x_1):= x_1 \\
+&\textbf{and}\; h(p,x, x_1) := S\;x 
+\end{align*}
 
@@
 
-\subsection{Multiplication on natural natural numbers}
+\subsubsection{Multiplication on natural natural numbers}
 
@@
 
-\begin{align}
+\begin{align*}
   0 \times x_1 &:= 0 \\
   S x_0 \times x_1 &:= (x_0 \times x_1) + x_1
-\end{align}
+\end{align*}
 
 This function is an instance of the primitive recursion scheme,
-with $n=1$, $g$ the constant unary function which returns $0$ (see Subsection~\vref{sect:k0}),
+with $n=1$, $g$ the constant unary function which returns $0$ (see subsection~\vref{sect:k0}),
  and $h$ the function defined by $h(p,x,n_1)=x+n_1$, which can be written as the composition of $+$ and the projections 
-$\pi_{2,3}$ and $\pi_{3,3}$ (The three of them  being  primitive recursive).
+$\pi_{2,3}$ and $\pi_{3,3}$ (the three of them  being  primitive recursive).
 
@@
 \begin{exercise}
- Show that the factorial function is primitive recursive (by the same method as above). 
+ Build a primitive recursive definition of the factorial function,
+using the constructions of  Section~\ref{sect:primrecmathdef},
+addition and multiplication.
 \end{exercise}
@@
 
-The four examples above are defined in \coq in Section~\vref{sect:primrec-examples2}
+The four examples above are defined as \coq terms in Section~\vref{sect:primrec-examples2}
 
@@
 O'Connor's library on Gödel's incompleteness theorems contains a little more 
-than 45K lines of scripts. The part dedicated to primitive recursive functions and Peano arithmetic is 32K lines long and is originally structured in 38 modules.
+than 45K lines of code. The part dedicated to primitive recursive functions and Peano arithmetic is 32K lines long and is originally structured in 38 modules.
 Thus, we propose a partial exploration of this library, through examples and exercises. Our additions to the original library --- mainly examples and counter-examples ---,
@@
 
+In order to make the terms of type \texttt{Primrec $n$} more readable, we introduce some notations, mainly inspired by \coq's standard library's notations for vectors.
+
+\inputsnippets{primRec/PRNotations}
+
+\subsection{A few arithmetic functions}
+\label{sect:primrec-examples2}
+Let us show how the few functions presented in Section~\vref{sect:primrec-examples2} can be defined in \coq.
+
+\noindent
+\emph{From \href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.primRecExamples}.}
+
+\inputsnippets{PrimRecExamples/FirstExamples}
+
 \subsection{A little bit of semantics} 
@@
 
+\subsubsection{A few tests}
+Let us consider again the functions defined in 
+previous sections.
+
 \vspace{4pt}
@@
 
-\input{movies/snippets/PrimRecExamples/evalPrimRecEx}
+\input{movies/snippets/PrimRecExamples/tests}
 
@@
 
-\subsection{A few arithmetic functions}
-\label{sect:primrec-examples2}
-Let us show how the few functions presented in Section~\vref{sect:primrec-examples2} can be defined in \coq.
+\subsubsection{A few proofs of correctness}
 
-\subsubsection{Constant function of value 0}
+It is tempting to \emph{prove} that our functions \texttt{cst0},
+\texttt{plus}, \texttt{mult} and \texttt{fact} are correct implementations in \texttt{PrimRec} of the mathematical functions we
+consider.
 
-\inputsnippets{PrimRecExamples/cst0}
-
-\subsubsection{Addition}
-
-\inputsnippets{PrimRecExamples/addition}
-
-\subsubsection{Multiplication}
-
-\inputsnippets{PrimRecExamples/multiplication}
-
-\subsubsection{The factorial function}
-
-\inputsnippets{PrimRecExamples/factorial}
+\input{movies/snippets/PrimRecExamples/correctness}
 
@@
 
-\begin{align}
+\begin{align*}
 A(0,n)&=n+1\\
@@
 A(m+1,n+1)&=A(m,A(m+1,n))
-\end{align}
+\end{align*}
 

### 2023-11-09 doc/chapter-primrec.tex
Remove the notation for empty vectors of PR functions
Replaced misc.inj_right_pair2 with stdlib's Eqdep_dec.inj_pair2_eq_dec

@@
 
+% If we consider any value of $n$, the same construction
+% builds the constant function of type $\mathbb{N}^n \arrow \mathbb{N}$ which returns $0$.
+
+\subsubsection{Constant functions}
+
+Let $k$ be some natural number; the unary constant function which always returns $k$ is built through $k$ nested compositions of the 
+successor function to the  unary constant function which returns $0$.
+
 \subsubsection{Addition on natural natural numbers}
@@
 This function is an instance of the primitive recursion scheme,
-with $n=1$, $g$ the constant unary function which returns $0$ (see subsection~\vref{sect:k0}),
+with $n=1$, $g$ is the constant unary function which returns $0$ (see subsection~\vref{sect:k0}),
  and $h$ the function defined by $h(p,x,n_1)=x+n_1$, which can be written as the composition of $+$ and the projections 

### 2023-11-15 doc/chapter-primrec.tex
Corrections to chapter 11 (coq and doc)

@@
 \section{Introduction}
-Primitive recursive functions are a small class of total functions, corresponding to the expressive power of a simple imperative programming language without \textbf{while} loops, in which every program execution terminates.
-
-\emph{Primitive recursive functions} are total functions from $\mathbb{N}^n$ to
-$\mathbb{N}$, for some $n\in\mathbb{N}$. Note that not all 
+\emph{Primitive recursive functions} are a small class of total arithmetic functions  from $\mathbb{N}^n$ to
+$\mathbb{N}$, for some $n\in\mathbb{N}$, corresponding to the expressive power of a simple imperative programming language without \textbf{while} loops, in which every program execution terminates.
+Please  note that not all 
 total $n$-ary recursive functions are primitive recursive
@@
 
-The traditional definition of the set of primitive recursive functions is structured as an inductive definition 
+The traditional definition of primitive recursive functions is structured as an inductive definition 
 in five rules: three base cases, and two recursive construction rules.
@@
 \subsection{A few (informal) examples}
-Before playing with primitive recursive functions in \coq{}, let us get familiar with the definition above, with the help of a few simple examples.
+\label{sect:PR-informal}
 \label{sect:primrec-examples1}
+Before playing with primitive recursive functions in \coq{}, let us get familiar with their mathematical definition, with the help of a few simple examples, which will be considered again
+as \coq terms in Section~\vref{sect:primrec-examples2}.
 
@@
 Let $k$ be some natural number; the unary constant function which always returns $k$ is built through $k$ nested compositions of the 
-successor function to the  unary constant function which returns $0$.
+successor function with the  unary constant function which returns $0$.
 
@@
   0 + x_1 &:= x_1 \\
-  S x_0 + x_1 &:= S (x_0 + x_1)
+  S\; x_0 + x_1 &:= S (x_0 + x_1)
 \end{align*}
@@
 
-\begin{remark}
+\section{First look at the Ackermann library}
 
-The four examples above are defined as \coq terms in Section~\vref{sect:primrec-examples2}
-
-\end{remark}
-  We present in this chapter a formalization of  primitive recursive functions, taken from  Russel O'Connor's formalization in \coq{} of
+  We present here a formalization of  primitive recursive functions, taken from  Russel O'Connor's formalization in \coq{} of
 G\"odel's incompleteness theorems~\cite{OConnor05}.
 
-This chapter contains some comments on Russel's library, as well as a few extensions.
-Contributions (under the form of comments, new examples or exercises) are welcome!.
-
-\section{First look at the Ackermann library}
+A few additions and/or small changes (mainly notations, and adaptation to the continuously evolving practices of \coq development) have been made to
+O'Connor's original library. 
+Contributions (under the form of comments, new examples or exercises) are welcome!
 
@@
 
-\section{Basic definitions}
+\section{Abstract syntax for primitive recursive functions}
 \index{maths}{Primitive recursive functions}
@@
 
-The  \texttt{primRec} library allows us to consider primitive functions on \texttt{nat}, with any number of arguments, in 
-curried form. This is made possible in 
-\href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat} by the following definition:
+The  \href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat}
+ library allows us to consider primitive functions on \texttt{nat}, with any number of arguments, in curried form. This is made possible in 
+by the following definition:
 
@@
 
-Likewise, arbitrary boolean predicates may have an arbitrary number of arguments. The dependent type
-(\texttt{naryRel $n$}), defined in the same way as \texttt{naryFunc}, is the type of $n$-ary functions from
-\texttt{nat} into \texttt{bool}.
+\subsection{Extensional equality}
 
-\input{movies/snippets/PrimRecExamples/naryRel2}
-
-The magic of dependent types makes it possible to define recursively extensional equality between functions of the same arity.
+Dependent types make it possible to define recursively extensional equality between functions of the same arity.
 
@@
 \index{ackermann}{Predicates!extEqual}
+\inputsnippets{extEqualNat/extEqualDef}
 
-\input{movies/snippets/extEqualNat/extEqualDef}
+Module \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec} defines and export  the notation ``\texttt{$f$ =x= $g$}'' for ``\texttt{extEqual $n$ $f$ $g$}'' \footnote{in parsing mode, provided $f$ has explicitely the type (\texttt{naryFunc $n$}).}
 
@@
 
+\subsection{Boolean predicates}
+
+Like arithmetic functions,  arbitrary boolean predicates may have an arbitrary number of arguments. The dependent type
+(\texttt{naryRel $n$}), defined in the same way as \texttt{naryFunc}, is the type of $n$-ary functions from
+\texttt{nat} into \texttt{bool}.
+
+\noindent
+\emph{From \href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat}}
+\inputsnippets{extEqualNat/naryRelDef}
+
+\inputsnippets{primRec/ltBoolDef, primRec/leBoolDef}
+
+\noindent
+\emph{From \href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat}}
+\input{movies/snippets/PrimRecExamples/naryRel2}
+
 \subsection{A Data-type for Primitive Recursive Functions}
@@
 
+\begin{remark}
+The \texttt{PrimRec}  type family is indeed a kind of 
+programming language for writing primitive recursive functions.
+The link to the mathematical notion of such functions will be
+established in Section~\ref{primrec-semantics} when we give a semantics which maps any term of type (\texttt{PrimRec $n$}) 
+to a function of type (\texttt{naryFunc $n$}).
+\end{remark}
 \index{coq}{Dependent types}
@@
 
-\subsection{A few arithmetic functions}
+\subsubsection{Examples}
 \label{sect:primrec-examples2}
-Let us show how the few functions presented in Section~\vref{sect:primrec-examples2} can be defined in \coq.
+
+Let us show how the functions described in \ref{sect:PR-informal} can be described by terms of type ``\texttt{PrimRec \_}''.
 
@@
 \subsection{A little bit of semantics} 
-Please note that inhabitants of type (\texttt{PrimRec $n$}) are not \coq{} functions like \texttt{Nat.mul}, or factorial, etc.
-The data-type (\texttt{PrimRec $n$}) is indeed an abstract syntax for the language of primitive recursive functions.
+\label{primrec-semantics}
+
+Inhabitants of type (\texttt{PrimRec $n$}) are not \coq{} functions like \texttt{Nat.mul}, or \texttt{Arith.Factorial.fact}, etc.
+but terms of an abstract syntax for the language of primitive recursive functions.
 The bridge between this language and the word of usual functions
@@
 (\texttt{evalprimRecS $n$ $m$}) of type 
-$\texttt{PrimRecs}\,n\,m \rightarrow  \texttt{Vector.t}\,(\texttt{naryFunc}\,n)\,m$.
+($\texttt{PrimRecs}\,n\,m \rightarrow  \texttt{Vector.t}\,(\texttt{naryFunc}\,n)\,m$).
 
@@
 Both functions are mutually defined through dependent pattern matching. We advise the readers who 
-would feel uneasy with dependent types to consult Adam Chlipala's \emph{cpdt}  book~\cite{chlipalacpdt2011}. We leave it to the reader  to look also at the helper functions in
-\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.
+would feel uneasy with dependent types to consult Adam Chlipala's \emph{cpdt}  book~\cite{chlipalacpdt2011}. We invite the 
+ reader  to look also at the helper functions in
+\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}, namely \texttt{evalConstFunc}, \texttt{evalProjFunc}, \texttt{evalComposeFunc}, and \texttt{evalPrimRecFunc}, etc.
 
@@
 
-\vspace{4pt}
+\subsubsection{A few tests}
 
-Looks complicated? The following examples show that, when
-the arity is fixed, these definitions behave well w.r.t. 
-\coq's reduction rules. Moreover, they make the interpretation functions more ``concrete''.
-
-\subsubsection{A few tests}
-Let us consider again the functions defined in 
-previous sections.
+The following examples show that the functions 
+\texttt{evalPrimRec} and \texttt{evalPrimRecs}
+ behave well w.r.t. 
+\coq's reduction rules. They can also be considered as 
+elementary tests of our definitions of \texttt{cst0}, \texttt{cst},
+\texttt{plus}, \texttt{mult} and \texttt{fact}.
 
@@
 
-\subsubsection{A few proofs of correctness}
+\subsubsection{Correctness proofs}
 
-It is tempting to \emph{prove} that our functions \texttt{cst0},
-\texttt{plus}, \texttt{mult} and \texttt{fact} are correct implementations in \texttt{PrimRec} of the mathematical functions we
+It is now time to \emph{prove} that our functions \texttt{cst0},
+\texttt{cst}, \texttt{plus}, \texttt{mult} and \texttt{fact} are correct implementations in \texttt{PrimRec} of the mathematical functions we
 consider.
@@
 
-\section{Proving that a given arithmetic function is primitive recursive}
+\section{Proving that a given \coq arithmetic function is primitive recursive}
 \label{sect:proofs-of-isPR}
@@
 The example in the preceding section clearly shows that, in order to prove that a given arithmetic function
-(defined in \gallina{} as usual) is primitive recursive, trying to give  by hand a term  of type (\texttt{PrimRec $n$}) is not a good method, since such terms may be huge and complex, even for simple arithmetic functions.
-The method proposed in Library \texttt{primRec} is the following one:
+(defined in \gallina{} as usual) is primitive recursive, trying to \emph{type} a term  of type (\texttt{PrimRec $n$}) is not a good method, since such terms may be too large, even for simple arithmetic functions.
+The method proposed in the library \texttt{Ackermann.primRec} is the following one:
 
 \begin{enumerate}
-\item Define a type corresponding to the statement "the function \texttt{$f$:naryFunc $n$} is primitive recursive ''.
-\item Prove handy lemmas which may help to prove that a given function is primitive recursive.
+\item Define a type corresponding to the statements of the form  "the $n$-ary function $f$ is primitive recursive ''.
+\item Prove handy lemmas which may help to prove that a given function is primitive recursive. These lemmas can be considered as a way to build \emph{silently} large terms of 
+type (\texttt{PrimRec $n$}) in intermediate steps of the proof.
+More, we can associate \emph{tactics} with these lemmas.
+
 \end{enumerate}
-
-Thus, the proof that a function, like \texttt{factorial}, is primitive recursive may be interactive, without having to type complex terms at any step of the development.
 
@@
 of the form (\texttt{isPR $n$ $f$}). We won't list all these lemmas here, but give a few examples of
-how they may be applied.
+how they may be searched, then applied.
 
 \begin{remark}
-In the library \texttt{primRec}, all these lemmas are opaque (registered with \texttt{Qed}).
-Thus they do not allow the user to look at the witness of a proof of a \texttt{isPR} statement.
-Our example of page~\pageref{sect:bigfac} was built using a  copy of \texttt{primRec.v} where many \texttt{Qed}s have been replaced with
-\texttt{Defined}s.
-
-If it does not cause compatibility problems (with 
-\href{https://github.com/coq-community/goedel}{goedel library} for instance), we plan to make all theses lemmas transparent.
+In the library \texttt{primRec}, all these lemmas are opaque (registered with \texttt{Qed}). Thus they do not allow the user to look at the witness of a proof of a \texttt{isPR} statement. 
+It may be useful to make transparent all the instances of
+\texttt{isPR} in the \texttt{Ackermann} and \texttt{goedel} libraries.
 \end{remark}
@@
 
-The constructors \texttt{zeroFunc}, \texttt{succFunc},  and \texttt{projFunc} of type
-\texttt{PrimRec} allows us to write trivial proofs of primitive recursivity. 
-Some of   the following lemmas are already proven in 
-\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}. More examples are given in 
+Simple proofs of statements (\texttt{isPR $n$ $f$}) may be just 
+applications of the constructor \texttt{is\_pr}, often
+thanks to the tactic call \texttt{exists $x$} where $x$ 
+is some (hopefully) correct term of type (\texttt{PrimRec $n$}).
+
+Let us show a few examples from
 \href{../theories/html/hydras.MoreAck.PrimRecExamples.html}%
-{Ackermann.MoreAck.PrimRecExamples.v}, in order to illustrate the main proof patterns.
+{Ackermann.MoreAck.PrimRecExamples.v}\footnote{Some of them are also in \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.}.
 
@@
 
-Likewise, we prove that the successor function on \texttt{nat} is primitive recursive too.
-
-\input{movies/snippets/primRec/SuccIsPR}
+\input{movies/snippets/primRecExamples/SuccIsPR}
 
@@
 Please note that the projection $\pi_{1,1}$ is just the identity on \texttt{nat}, and is realized by 
-(\texttt{projFunc 1 0}).
+(\texttt{projFunc 1 0}) (see Sect.~\vref{sect:primrec-examples2}).
 
@@
 
+\subsubsection{The predecessor (total) function}
+
+The predecessor function is defined by the following equations:
+
+\begin{align*}
+\texttt{pred}(0) &= 0 \\
+\texttt{pred}(1+n) &= n = \pi_{1,2}(n,\texttt{pred}(n))
+\end{align*}
+
+This definition is easy to translate into a term of type \texttt{PrimRec 1}.
+
+\inputsnippets{PrimRecExamples/xpred}
+
 \subsubsection{Using function composition}
@@
 Let us look at the proof that any constant $n$ of type \texttt{nat} has type (\texttt{PR 0})
-(lemma  \texttt{const1\_NIsPR} of \texttt{primRec}). We carry out a proof by induction on $n$, the base case of which is already proven.
-Now, let us assume $n$ is \texttt{PR $n$}, with $x:\texttt{PrimRec}\,0$ as a ``realizer''.
+(lemma  \texttt{const0\_NIsPR} of \texttt{primRec}). We carry out a proof by induction on $n$, the base case of which is already proven.
+Now, let us assume $n$ is \texttt{PR $0$}, with $x:\texttt{PrimRec}\,0$ as a ``realizer''.
 Thus we would like to compose this constant function with the unary successor function.
@@
 \vspace{4pt}
-Thus, we get a quite simple proof of \texttt{const1\_NIsPR}.
+Here is a quite simple proof of \texttt{const0\_NIsPR}.
 
@@
 
-\subsubsection{Proving that \texttt{plus} is primitive recursive}
+\subsubsection{Another proof that \texttt{Nat.add} is primitive recursive}
 
-The lemma \texttt{plusIsPR} is already proven in \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}. We present in 
-\href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}
-a commented version of this proof,
+We have already proven that \texttt{Nat.add} is primitive recursive. The following alternative proof, --- more detalled  ---, 
+shows how to search and apply lemmas from
+the \texttt{Ackermann} library.
 
-First, we look for lemmas which may help to prove that a given function obtained with the recursor \texttt{nat\_rec} is primitive recursive.
+Let us look for some lemma which could help to prove some 
+recursive arithmetic binary function is primitive recursive.
 
-\input{movies/snippets/PrimRecExamples/PrimRecExamplesSearch}
+\inputsnippets{PrimRecExamples/PRnatRecSearch}
 
-The following lemma shows it suffices to prove that
-Standard library's function \texttt{plus} is extensionally equal to a function defined with
-\texttt{nat\_rec}.
+Good! Let us express addition in terms of \texttt{nat\_rec}.
 
-\input{movies/snippets/primRec/isPRExtEqual}
+\inputsnippets{PrimRecExamples/plusIsPRa}
 
-Thus, let us define an helper and prove its equivalence with \texttt{plus}.
+The lemma \texttt{Ackermann.isPRextEqual} tells us that if a function $g$ is extensionally equal to a primitive recursive function, then $g$ is primitive recursive too.
 
-\input{movies/snippets/PrimRecExamples/plusAlt}
+\inputsnippets{PrimRecExamples/isPRextEqual}
+
+% The lemma \texttt{plusIsPR} is already proven in \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}. We present in 
+% \href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}
+% a commented version of this proof,
+
+% First, we look for lemmas which may help to prove that a given function obtained with the recursor \texttt{nat\_rec} is primitive recursive.
+
+% \input{movies/snippets/PrimRecExamples/PRnatRecSearch}
+
+% Thus, let us define an helper and prove its equivalence with \texttt{Nat.add}.
+
+% \input{movies/snippets/PrimRecExamples/plusIsPRa}
 
@@
 
-We are now able to complete the proof.
+Let us start our proof.
 
-\input{movies/snippets/PrimRecExamples/plusIsPRa}
+\inputsnippets{PrimRecExamples/plusIsPRa1,
+  PrimRecExamples/plusIsPRa2}
 
@@
 \vspace{4pt}
-\input{movies/snippets/PrimRecExamples/checkFilter0101IsPR}
+\input{movies/snippets/PrimRecExamples/plusIsPRa3}
 \vspace{4pt}
 
-Thus, our first subgoal is solved easily. The rest of the proof 
+Thus, our first subgoal is easily solved. The rest of the proof 
 is just an application of already proven lemmas.
@@
 
-\begin{todo}
-Comment more examples from   \href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}.
-\end{todo}
-
 \index{ackermann}{Exercises}
 \begin{exercise}
-There is a lot of lemmas similar to \texttt{filter010IsPR} in the \texttt{primRec} library, useful to control the arity of functions.
-Thus, the reader may look at them, and invent simple examples of application for each lemma.
+A few lemmas similar to \texttt{filter010IsPR}, also shown  in the \texttt{primRec} library help the user to control the arity of functions.
+Thus, the reader may look at them, and invent h.er.is simple examples of application.
 \end{exercise}
-
-\index{ackermann}{Exercises}
-\begin{exercise}
-Multiplication of natural number is already proven in the \texttt{primRec} library. Write a proof of your own, then compare to the library's version.
-\end{exercise}
-
-%\input{movies/snippets/PrimRecExamples/justATest}
 
@@
 
-\input{movies/snippets/MorePRExamples/factDef}
-
 \input{movies/snippets/MorePRExamples/expDef}
@@
 
-\index{ackermann}{Exercises}
-\begin{exercise}
-Show that the function \texttt{min: naryFunc\,2} is primitive
-recursive.
+\subsection{More advanced examples}
 
-\emph{You may start this exercise with
-    \href{https://github.com/coq-community/hydra-battles/blob/master/exercises/primrec/MinPR.v}{exercises/primrec/MinPR.v}.}
+\subsubsection{The minimum of two natural numbers}
 
-\end{exercise}
+Let $a$ and $b$ be two natural numbers. The \emph{minimum} of $a$ and $b$ is $a$ if $a \leq b$, otherwise $b$.
+
+Thus, we propose the following definition:
+
+\inputsnippets{MinPR/minAltDef}
+
+Here is a sketch of proof that standard library's \texttt{min} is
+primitive recursive. The reader is kindly invited to fill the missing steps.
+
+\inputsnippets{MinPR/minProof1, MinPR/minProof2, MinPR/minProof3}
 

### 2023-11-17 doc/chapter-primrec.tex
Fix a latex error
Corrections to chapter 11 (primitive recursive functions)

@@
 
-\input{movies/snippets/primRecExamples/SuccIsPR}
+\input{movies/snippets/PrimRecExamples/SuccIsPR}
 
@@
 \index{ackermann}{Exercises}
-\begin{exercise}
+\begin{exercise}[The integer square root]
 
-Prove the following lemmas (which may help to solve the next  exercise).
+\mbox{}
+\vspace{4pt}
+
+\noindent
+ \textbf{1)} 
+Please consider the following specification of the function \texttt{boundedSearch} defined in 
+\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.
+
+\inputsnippets{isqrt/boundedSearchSpec}
+
+Prove the following lemmas.
 
@@
 
-\textbf{Note:}  The function \texttt{boundedSearch} is defined
-in Library \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}. 
-\end{exercise}
+\noindent
+\textbf{2)} 
+Let us consider the following definition of the relation `` $r$ is the integer square root of $n$ ''.
 
-\index{ackermann}{Exercises}
+\inputsnippets{isqrt/isqrtSpec}
 
-\begin{exercise}
-Prove that the function which returns the  integer square root of any natural number  is primitive recursive.
+Prove that the function which returns the  integer square root of any natural number  is primitive recursive (you may use 
+the function \texttt{boundedSearch} for this purpose).
 
@@
 and to play with dependent types.
+
+With respect to proof methods, we will see a first \emph{proof by induction on the set of all primitive recursive functions}.
+First, we will prove a property shared by all primitive function, 
+then show that the Ackermann function doesn't satisfy this property.
 

### 2023-11-20 doc/chapter-primrec.tex
Changes to chapter 11; Renaming some symbols in Epsilon.E0

@@
 the associated constructor and projections are primitive recursive.
+
+Please find here some definitions and lemmas you may use 
+in order to solve this exercise (non-exhaustive list).
+
+\inputsnippets{FibonacciPR/AboutCPair}
+
 \emph{You may start this exercise with the file
@@
 
-\section{Proving that a given function is \emph{not} primitive recursive}
-\label{sect:ack-not-PR}
+\section{Proofs by induction over all primitive recursive functions}
 
-The best known example of a total recursive function which is not primitive recursive is the Ackermann function.
-We show how to adapt the classic proof (see for instance~\cite{planetmath}) to the constraints of \gallina.
-We hope this formal proof
+We may want to prove properties shared with all the set of
+primitive function, for instance that any primitive recursive function is representable by a predicate in Peano Arithmetics (see Chap~›\vref{chap:PRRep}).
+In this chapter, we will prove by induction that every primitive recursive function has a given property which is not satisfied by  Ackermann function, and deduce that this function is not PR.
+We show how to adapt the classic proof (see for instance~\cite{planetmath}) to the constraints of \gallina. We hope this formal proof 
  is a nice opportunity to explore
@@
 
-With respect to proof methods, we will see a first \emph{proof by induction on the set of all primitive recursive functions}.
-First, we will prove a property shared by all primitive function, 
-then show that the Ackermann function doesn't satisfy this property.
+First, let us define Ackermann function in \coq, then look at a few properties of this function.
 
@@
 
-\subsection{A proof by induction on all primitive recursive functions}
-
 In order to prove that \texttt{Ack} (considered as a function of two arguments) is not primitive recursive, the usual method consists in two steps:
@@
 
-First, we prove that any primitive function of two arguments is majorized by \texttt{Ack}.
+First, we have to prove that any primitive function of two arguments is majorized by \texttt{Ack}.
+Let us look at the induction principles generated for the types
+\texttt{PrimRec $n$}. % and \texttt{PrimRecs $n$ $m$}
+
 If we look at the inductive definition of primitive recursive functions, page~\pageref{def:Primrec}, it is obvious that a proof by induction on the construction of primitive recursive functions must consider functions of any arity.
-
-The following scheme allows us to write proofs by induction on the class of primitive recursive functions.
 
@@
 \href{../theories/html/hydras.Prelude.MoreVectors.html}{Prelude.MoreVectors}).
-So, (\texttt{t\,$A$\,$n$}) is the type of vectors of $n$ elements of type $A$, and the constants \texttt{cons}, \texttt{nil}, \texttt{map}, etc., refer to vectors and not to lists.
-Likewise, the notation \texttt{x::v} is an abbreviation for
-\texttt{VectorDef.cons x \_ v}.
+So, (\texttt{t $A$ $n$}) is the type of vectors of $n$ elements of type $A$, and the constants \texttt{cons}, \texttt{nil}, \texttt{map}, etc., refer to vectors and not to lists.
+Likewise, the notation \texttt{$x$::$v$} is an abbreviation for
+\texttt{VectorDef.cons $x$ \_ $v$}.
 
@@
 
-The remaining cases are proved within a mutual  induction.
+The remaining cases are proved within the main  mutual  induction.
 
@@
 
-Then, we write  a short proof by contradiction.
+Then, we write  a short proof by contradiction, using a 
+\emph{diagonalized} version of Ackermann function.
 

### 2023-11-23 doc/chapter-primrec.tex
minor corrections

@@
 
+\label{sect:AcknPR}
+
 Please note also that for any given $n$, the unary function
@@
 
+\begin{todo}
+Fix the display the local definition of \texttt{x} in the following proof.
+\end{todo}
+
 \input{movies/snippets/AckNotPR/AckNotPR}
@@
 \begin{remark}
-Note that the Ackermann function is a counter-example to the (false) statement:
-\begin{quote}
-{\color{red}
-  ``Let $f$ be a function of type \texttt{naryFunc\,2}. If, for any $n$, the function $f(n)$ is primitive recursive, then f is primitive recursive.''}
-\end{quote}
+  It may be interesting to compare the following statements:
+
+  \begin{itemize}
+  \item Ackermann function is not primitive recursive.
+  \item For any $n$, the function \texttt{Ack $n$ \_} is primitive recursive (see~\vref{sect:AcknPR}).
+  \end{itemize}
 \end{remark}

### 2023-11-29 doc/chapter-primrec.tex
Complete a remark in Chapter 11 (Alectryon output)

@@
 \begin{todo}
-Fix the display the local definition of \texttt{x} in the following proof.
+  Fix the display of the local definition of \texttt{x} in the following proof: \emph{Alectryon} prints a \emph{declaration} \texttt{x: nat}, but forgets to print the definition \texttt{x := Nat.max 2 m},  which is  important to complete the proof.
 \end{todo}

### 2023-12-04 doc/chapter-primrec.tex
Small correction (doc)

--- 2023-11-23 doc/chapter-primrec.tex
@@
 
-\begin{todo}
-Fix the display the local definition of \texttt{x} in the following proof.
-\end{todo}
+\begin{remark}
+In the following snippet,  \emph{Alectryon}'s \texttt{Latex} generator  prints the \emph{local definition} of $x$ (as the maximum of $2$ and $m$) as a simple \emph{declaration} \texttt{x: nat}.
+Thus the proof script is correct, but the three last sub-goals are not correctly displayed, since 
+they do not show how the inequalities $2\leq x$ and $m \leq x$ could be inferred by \texttt{lia}.
+\end{remark}
 

### 2023-12-08 doc/chapter-primrec.tex
Version for Coq <= 8.17.1  (heavily uses Coq.Sets and Coq.Vector)

@@
 
-We may want to prove properties shared with all the set of
-primitive function, for instance that any primitive recursive function is representable by a predicate in Peano Arithmetics (see Chap~›\vref{chap:PRRep}).
-In this chapter, we will prove by induction that every primitive recursive function has a given property which is not satisfied by  Ackermann function, and deduce that this function is not PR.
+Let us consider the following theorem (see for instance ~\cite{planetmath}).
+
+\begin{quote}
+   There exists at least a total arithmetic function, \emph{e.g.} the \emph{Ackermann function},  which is not primitive recursive.
+\end{quote}
+
+We can prove this theorem in three successive steps:
+
+\begin{itemize}
+\item Define Ackermann function in \gallina.
+\item Define and prove a property shared by any primitive recursive functions.
+\item Prove that  Ackermann function does not satisfy this property.
+\end{itemize}
+
 We show how to adapt the classic proof (see for instance~\cite{planetmath}) to the constraints of \gallina. We hope this formal proof 
@@
 and to play with dependent types.
-
-First, let us define Ackermann function in \coq, then look at a few properties of this function.
 
@@
 where (\texttt{iterate\,$f$\,$n$})
-is the $n$-th iteration of $f$\,\footnote{Please not confuse with \texttt{primRec.iterate}, which is monomorphic and does not share the same order of arguments.}.
-It makes it possible to apply a few lemmas proved in
+is the $n$-th iteration of $f$\,\footnote{Please do not mistake
+  \texttt{iterate} (\emph{i.e.} \texttt{Prelude.iterate}) for
+  the monomorphic \texttt{primRec.iterate}, which does not share the same order of arguments.}. It makes it possible to apply a few lemmas proved in 
 \href{../theories/html/hydras.Prelude.Iterates.html}{Prelude.Iterates}, for instance about the monotony of the $n$-th iterate of a given function.

### 2024-01-08 doc/chapter-primrec.tex
small corrections

@@
 \begin{remark}
-In the following snippet,  \emph{Alectryon}'s \texttt{Latex} generator  prints the \emph{local definition} of $x$ (as the maximum of $2$ and $m$) as a simple \emph{declaration} \texttt{x: nat}.
+In the following snippet,  some versions \emph{Alectryon}'s \texttt{Latex} generator  print the \emph{local definition} of $x$ (as the maximum of $2$ and $m$) as a simple \emph{declaration} \texttt{x: nat}.
 Thus the proof script is correct, but the three last sub-goals are not correctly displayed, since 
 they do not show how the inequalities $2\leq x$ and $m \leq x$ could be inferred by \texttt{lia}.
+
+A correct goal display can be obtained with
+\href{https://github.com/Casteran/alectryonFix}{this fork}.
 \end{remark}
@@
 \begin{remark}
-It is easy to prove that any unary function which dominates \texttt{fun n => Ack n n} fails to be primitive recursive.
+It is easy to prove that any unary function which dominates (\texttt{fun n => Ack n n}) fails to be primitive recursive.
 To this end, we use an instance of \texttt{majorAnyPR} dealing with unary functions.

### 2024-01-09 doc/chapter-primrec.tex
small corrections

@@
 
+\section{Ackermann function is not primitive recursive}
+
+\label{sect:ack-not-PR}
+
 In order to prove that \texttt{Ack} (considered as a function of two arguments) is not primitive recursive, the usual method consists in two steps:
@@
 \begin{remark}
-In the following snippet,  some versions \emph{Alectryon}'s \texttt{Latex} generator  print the \emph{local definition} of $x$ (as the maximum of $2$ and $m$) as a simple \emph{declaration} \texttt{x: nat}.
+In the following snippet,  some versions of \emph{Alectryon}'s \texttt{Latex} generator  print the \emph{local definition} of $x$ (as the maximum of $2$ and $m$) as a simple \emph{declaration} \texttt{x: nat}.
 Thus the proof script is correct, but the three last sub-goals are not correctly displayed, since 

### 2024-01-10 doc/hydras.tex
Changes to chapter 11; Renaming some symbols in Epsilon.E0
corrections to the general introduction

@@
 
+If a formal development is large (at least 10 KLOCs), we believe that a human-readable document containing explanations, diagrams, code snippets, examples, exercises, etc.) would be useful
+for a better understanding of both the mathematical contents
+and the formalization techniques used in the
+development~\cite{Chiplunkar23}.
+
+This document has been generated with \alectr (see Sect.~\vref{sect:alectryon-intro}), which ensures the \texttt{pdf} is consistent with the last compiled version of the \coq project.
+
 Among all the theorems proved with the help of proof assistants like \coq{}~\cite{Coq,BC04}, \hol{}~\cite{HOL}, \isabelle{}~\cite{isabelle},  etc.,
@@
 \subsection{Documenting theories with \alectr}
+\label{sect:alectryon-intro}
 
@@
 Unlike mathematical literature, where definitions and proofs are spread out over many articles and books,
-the whole proof is now inside your computer. It is composed from the \texttt{.v} files you downloaded and
-parts of \coq's standard library.
+the whole proof is now inside your computer. It is composed from the \texttt{.v} files you downloaded,
+parts of \coq's standard library, and required \coq packages (see Fig.~\vref{fig:genealogy}).
 Thus, there is no ambiguity in our definitions and the premises of the theorems.
@@
 
-It may also happen that some proof looks to be useless, because the proven theorem is a trivial consequence of another (also proven) result.
+It may also happen that some direct proof looks to be useless,
+because the proven theorem is a trivial consequence of another (also proven) result.
 For instance, let us consider the three following statements:
@@
 Whenever possible, we use \alectr to display \coq code (definition, proof scripts) and answers.
-
 Here are two  examples from Chapters~\ref{chapter:primrec}
@@
 
-Quotations of \coq{} source  from others libraries (\coq's standard library, borrowed plug-ins)  are displayed as follows.
+Quotations of \coq{} source  from other libraries (\coq's standard library, borrowed plug-ins)  are displayed as follows.
 
@@
 
-We use also verbatim code inclusions when the examples lead to very long computations.
+We use also verbatim code inclusions when the examples lead to too long computations.
 
@@
 \label{sec:orgheadline4}
- The present distribution has been checked with version 8.14.1 of the Coq proof assistant, with a few plug-ins.
+ The present distribution has been checked with versions up to 8.18 of the Coq proof assistant, with a few plug-ins.
 \emph{Please refer to \href{https://github.com/coq-community/hydra-battle\#readme}{the README file of the project}.}
@@
 \include{chapter-encoding}
+\include{chapter-expressible}
 

### 2024-01-15 doc/hydra-chapter.tex
Improve doc (#172)
* small corrections

* corrections to the general introduction

* corrections to the general introduction

* corrections chapter 2

@@
 A battle will thus be represented as a sequence of terms of type \texttt{Hydra}, respecting the rules of the game.
+ In other terms, we consider hydra battles as \emph{transition systems}.
 
@@
 
-\begin{figure}[hp]
+\begin{figure}[h]
 \centering
@@
 Using this representation, one can re-define all the constructions of this chapter, which is left as an exercise.
-You will probably have to use patterns described for instance in~\cite{BC04} or the archives of the \coq communication channels (please consult~\url{https://coq.inria.fr/community.html}).
-
+You will probably have to use patterns described for instance in~\cite{BC04, BC04ch14} or the archives of the \coq communication channels (please consult~\url{https://coq.inria.fr/community.html}).
 \end{remark}
@@
 
-We notice that this sub-goal does not contain any hypothesis
-on the height and size of the hydra \texttt{h}. So, it looks hard to prove the conclusion. Let's stop.
+We notice immediately that the context of this sub-goal does not allow to infer its conclusion.
+ Let's stop.
 
@@
 In this section, we represent the rules of hydra battles as a binary relation associated with
-a \emph{round}, i.e., an interaction composed of the two following actions:
+a \emph{round}\footnote{usually called a \emph{small step semantics}}, i.e., an interaction composed of the two following actions:
 \begin{enumerate}
@@
 
-If Hercules chops off a head close to the root, there is no replication at all. We use an auxiliary 
+If Hercules chops off a head next to the root, there is no replication at all. We use an auxiliary 
 predicate \texttt{S0}, associated with the removing of one head from a sequence of hydras.
@@
 Let us represent in \coq{}   the transformation of the hydra of Fig.~\vref{fig:Hy} into
-the configuration represented in Fig.~\ref{fig:Hy-prime}.
+the configuration represented in Fig.~\vref{fig:Hy-prime}.
 
@@
 
-Let us merge \texttt{R1} and \texttt{R1}  into a single relation.
+Let us merge \texttt{R1} and \texttt{R2}  into a single relation.
 First,  we define the  relation \texttt{(round\_n n h h')} where \texttt{n} is the expected number of  replications (irrelevant in the case of an \texttt{R1}-transformation).
@@
 \input{movies/snippets/Hydra_Definitions/roundPlus}
-
-\index{hydras}{Exercises}
-
-\begin{exercise}
-  Prove that if \texttt{$h$ -+-> $h'$}, then
-  the height of $h'$ is less or equal than the height of $h$.
-
-\end{exercise}
 
@@
 
+\index{hydras}{Exercises}
+\begin{exercise}
+  Prove that if \texttt{$h$ -+-> $h'$}, then
+  the height of $h'$ is less or equal than the height of $h$.
+
+\end{exercise}
 \index{hydras}{Exercises}
@@
 \label{sect:testing}
-In order to study \emph{experimentally} the different  configurations of the  battle, we will use a simple data type for representing the states as tuples composed of
-the round number, and the respective number of daughters  \texttt{h2}, \texttt{h1}, and heads
-of the current hydra.
+In order to make the study of this battle easier,  we will use a simple data type for representing a configuration 
+$(\textit{round}, \texttt{hyd}\;n_2\;n_1\; n_h)$ as the $4$-tuple
 
@@
 The battle we are studying looks to be awfully long. Let us concentrate our
-tests on some particular events : the states where $\texttt{nh}=0$.
+tests on some particular events : the states where $n_h=0$.
 From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} is equal to zero.
@@
 In the following definition \texttt{i} is associated with the round number (or date, if we consider a discrete time), and \texttt{a}, \texttt{b}, \texttt{c} respectively associated with the number of occurrences of \texttt{h2}, \texttt{h1} and heads connected to the hydra's foot.
-For convenience, we do not use the type \texttt{state} of the preceding section, but consider the round numbers and the number of hydras \texttt{h2}, \texttt{h1} and heads as separate arguments of the relation (which is no more ---formally--- ``binary'').
+For convenience\footnote{In a few words, the type \texttt{state} was designed for performing \emph{computations}, and \texttt{steps} for writing \emph{interactive proofs}, inspired by the aforementionned computations.},
+we do not use the type \texttt{state} of the preceding section, but consider the round numbers and the number of hydras \texttt{h2}, \texttt{h1} and heads as separate arguments of the relation (which is no more ---formally--- ``binary'').
 
@@
 
-The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{battle}.
+The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{rounds}.
 
@@
 
-The number $N$ is greater than or  equal to $2^{2^{95}\times 95}.$ If we wrote $N$ in base $10$, $N$ would require at least $10^{30}$ digits!
+The number $N$ is greater than or  equal to $2^{2^{95}\times 95}.$ If we write $N$ in base $10$, $N$ would require at least $10^{30}$ digits!
 
@@
 
-The best known well-founded order is the natural order on the set $\mathbb{N}$ of natural numbers (the type \texttt{nat} of Standard library).
+The best known well-founded order is the natural order on the set $\mathbb{N}$ of natural numbers (the type \texttt{nat} of Standard Library).
 It would be interesting to look for some measure $m:\texttt{nat}\arrow\texttt{nat}$ and prove it is a variant.

### 2024-01-15 doc/schutte-chapter.tex
Fix a typo
Update doc/schutte-chapter.tex
Co-authored-by: Columbus240 <8899730+Columbus240@users.noreply.github.com>
Update doc (#161)
* fix a few Hint locality attributes

* Document the use of zorns-lemma

* Fix a typo

* Update doc/thebib.bib

Co-authored-by: Columbus240 <8899730+Columbus240@users.noreply.github.com>

* Update theories/ordinals/Schutte/Countable.v

Co-authored-by: Columbus240 <8899730+Columbus240@users.noreply.github.com>

* Update doc/schutte-chapter.tex

Co-authored-by: Columbus240 <8899730+Columbus240@users.noreply.github.com>

* Update theories/ordinals/Schutte/Countabl

@@
 \label{types:Ord}
-
-Countable sets were
 
@@
 Is it true ?
-
 \emph{You may start this exercise with the file
@@
 The sub-directory
-    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/Gamma0}{theories/oridinals/Gamma0} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form.
+    \href{https://github.com/coq-community/hydra-battles/tree/master/theories/ordinals/Gamma0}{theories/ordinals/Gamma0} contains an (incomplete, still undocumented) implementation of the set of ordinals below $\Gamma_0$, represented in Veblen normal form.
 \end{remark}

### 2024-01-22 doc/ordinal-chapter.tex
Small corrections in Chapter 3

@@
 With respect to the order induced by this sequence, any limit ordinal $\alpha$ is the least upper bound of  the set $\mathbb{O}_\alpha$ of all ordinals strictly less than $\alpha$.
-\item
-For instance $\omega$ is the least upper bound of the set of all finite ordinals (in the first line). It is also the first limit ordinal, and the first infinite ordinal, in the sense that 
+For instance,$\omega$ is the least upper bound of the set of all finite ordinals (in the first line).
+It is also the first limit ordinal, and the first \emph{infinite ordinal number}, in the sense that
 the set $\mathbb{O}_\omega$ is infinite.
@@
 \section{The mathematical point of view}
+
+We cannot cite all the literature published on ordinals since Cantor's book
+\cite{cantorbook}, and 
+leave it to the reader to explore the bibliography.
+The introduction of Jos\'e Grimm's report~\cite{grimm:hal-00911710} contains also a nice presentation of the main properties of ordinals.
+
+For simplicity's sake, we will only give the definitions which are useful for understanding our \coq development.
 
@@
 
-Having the same order type is an equivalence relation between well-ordered sets. Ordinal numbers (in short: \emph{ordinals})  are descriptions (\emph{names}) of the equivalence classes.
+Having the same order type is an equivalence relation between well-ordered sets. Ordinal numbers (in short: \emph{ordinals})  are descriptions (\emph{names}) of the associated equivalence classes.
 For instance, the order type of $(\mathbb{N},<)$ is associated with the ordinal called  $\omega$, and the order we considered on 
-the disjoint union of $\mathbb{N}$ and itself is named $\omega+\omega$.
+the disjoint union of two copies of $\mathbb{N}$ is associated with  $\omega \times 2$ (a.k.a. $\omega+\omega$).
 
@@
 For instance $\mathbb{O}_\omega=\mathbb{N}$, and $\mathbb{O}_7=\{0,1,2,3,4,5,6\}$.
-
-We cannot cite all the literature published on ordinals since Cantor's book
-\cite{cantorbook}, and 
-leave it to the reader to explore the bibliography.
-
-The introduction of Jos\'e Grimm's report~\cite{grimm:hal-00911710} contains a nice presentation of the main properties of ordinals.
 
@@
 Let $\alpha$ be some (countable) ordinal; 
-in \coq{} terms, we call \emph{ordinal notation for $\alpha$} a structure composed 
+we call \emph{ordinal notation for $\alpha$} any structure composed 
 of:
@@
 The correctness of the comparison function is expressed through Stdlib's type 
-\texttt{Datatypes.CompareSpec} as specialized by \texttt{Datatypes.CompSpec}.
+\texttt{Datatypes.CompareSpec} and predicate \texttt{Datatypes.CompSpec}.
 
@@
 
-\inputsnippets{ON_plus/ltWf, ON_plus/OnPlus}
-
-\subsection{The ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
+\inputsnippets{ON_plus/ltWf}
+
+\inputsnippets{ON_plus/OnPlus}
+
+\subsection{Example: The ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}
 
@@
 It is also represented by the two first lines of Figure~\ref{fig:ordinal-sequence}.
+We define this notation in \coq{} as an instance of \texttt{ON\_plus}.
 
@@
 
-We can define this notation in \coq{} as an instance of \texttt{ON\_plus}.
-
 \vspace{4pt}
@@
 
-\inputsnippets{ON_Omega2/plusDef, ON_Omega2/plusExamples}
+\inputsnippets{ON_Omega2/plusDef}
+
+\inputsnippets{ON_Omega2/plusExamples}
 

### 2024-01-29 doc/epsilon0-chapter.tex

@@
 
-In this chapter, we adapt to \coq{} the well-known~\cite{KP82}  proof that Hercules eventually wins every battle, whichever the strategy  of each player.
-In other words, we present  a formal and self contained proof of termination  of all [free] hydra battles.
+In this chapter, we adapt to \coq{} the well-known proof~\cite{KP82} that Hercules eventually wins every battle, whichever the strategy  of each player.
+In other words, we present  a formal and self-contained proof of termination  of all [free] hydra battles.
 First, we take from Manolios and Vroon~\cite{Manolios2005} a representation of the ordinal $\epsilon_0$ as terms in Cantor normal form.
@@
 the equation \(\alpha = \omega^\alpha\), where \(\omega\) is 
-the least infinite ordinal.
+the least infinite ordinal\footnote{For a precise ---\,\emph{i.e.} mathematical\,--- definition of $\omega^\alpha$, please see Sect.~\vref{sect:AP-and-phi0}.} .
 Thus, we can intuitively consider \(\epsilon_0\) as an
@@
 can be finitely represented by a unique  \emph{Cantor normal form}, 
-that is, an expression  which is either  the ordinal \(0\) or 
+that is, an expression  which is 
 a sum  \(\omega^{\alpha_1} \times n_1 + \omega^{\alpha_2} \times n_2 + 
-  \dots + \omega^{\alpha_p} \times n_p\) where all the \(\alpha_i\) 
-are ordinals in Cantor  normal form, \(\alpha_1 > \alpha_2 > \alpha_p\), 
+  \dots + \omega^{\alpha_p} \times n_p\) where $p\in\mathbb{N}$, all the \(\alpha_i\) 
+are ordinals in Cantor  normal form, \(\alpha_1 > \alpha_2 > \alpha_p\)
 and all the \(n_i\) are positive integers.
@@
 and the decomposition into Cantor normal form  is consistent 
-with the implementation of the arithmetic operations of exponentiation of base \(\omega\) 
-and addition.
+with other usual definitions of ordinals, for instance in \gaia~\cite{Gaia}, Schütte's book~\cite{schutte}, or larger ordinal notations cf. Chapter~\vref{chap:gamma0}.
 
@@
 Unless explicitly mentioned, the term ``ordinal" will be used instead of
-``ordinal strictly less than \(\epsilon_0\)" (except in Chapter~\ref{chap:schutte} where it stands for ``countable ordinal'').
+``ordinal strictly less than \(\epsilon_0\)" (except in Chapter~\ref{chap:schutte} where it stands for ``countable ordinal'', and Chapter~\ref{chap:gamma0}.).
 
@@
 ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
-In a future version, we should replace  the type \texttt{nat} with \texttt{positive} in \texttt{T1}'s 
+In a future version, we would like to replace  the type \texttt{nat} with standard library's type \texttt{positive} in \texttt{T1}'s 
 definition. But this replacement would take a lot of time \dots{}
@@
 
-Note that \texttt{omega} is not an identifier in \HydrasLib, thus any tactic like \texttt{unfold omega} would fail.
+Note that \texttt{T1omega} is not an identifier in \HydrasLib, thus any tactic like \texttt{unfold T1omega} would fail.
 
@@
 \paragraph*{\gaiasign}
-In \texttt{gaia.ssete9}, the ordinal $\omega$ is bound to the \emph{constant}  \texttt{T1omega}.
+In \texttt{gaia.ssete9}, the ordinal $\omega$ is bound to the \emph{constant}  \texttt{T1omega} (not a notation).
 
@@
 \begin{project}
-Design  (in \ocaml?) a set of tools for systematically pretty printing ordinal terms in Cantor normal form.
+Design tools for systematically pretty printing ordinal terms in Cantor normal form.
 \end{project}
@@
 
+\input{movies/snippets/T1/Instances}
+
 \label{Predicates:lt-T1}
@@
 
-Links between the function \texttt{compare} and the relations
-\texttt{lt} and \texttt{eq} are established through the following lemmas (~\vref{sect:comparable-def}).
-\vspace{4pt}
-
-\input{movies/snippets/T1/Instances}
+% Links between the function \texttt{compare} and the relations
+% \texttt{lt} and \texttt{eq} are established through the following lemmas (~\vref{sect:comparable-def}).
+% \vspace{4pt}
 

### 2024-01-29 doc/gaia-chapter.tex
small improvements
minor corrections

@@
 Functions \texttt{h2g} and \texttt{g2h} allow us to define
-a notion of ``data-refinement''  for constants, functions, predicates and relations. The following definitions express that some
+a notion of ``data-refinement''  for constants, functions, predicates and relations. The following definitions express that a given
 constant, function, relation defined in \HydrasLib ``implements'' the same concept of \gaia.
+
+From~\href{../theories/html/gaia_hydras/T1Bridge.html}%
+{\texttt{gaia\_hydras.T1Bridge}}.
 
@@
 
-\inputsnippets{T1Bridge/refines1R, T1Bridge/refines2R}
+\inputsnippets{T1Bridge/refines1R}
+
+\inputsnippets{T1Bridge/refines2R}
 
@@
 The following command lists us `gaia's lemmas (thanks to
-\gaia's  \texttt{cantor\_scope}.
+\gaia's  \texttt{cantor\_scope}).
 \inputsnippets{T1Bridge/SearchDemoa}
 
-With \texttt{t1\_scope}:
+Within \texttt{t1\_scope}:
 \inputsnippets{T1Bridge/SearchDemob}
@@
 
-\section{Rapidly growing arithmetic functions}
-\index{gaiabridge}{Rapidly growing functions}
-
-In this section, we consider a few families of arithmetic functions, indexed by ordinal numbers below $\epsilon_0$.
-The type of such families is \texttt{E0 -> nat -> nat} where
-\texttt{E0} is the type of ordinal terms in Cantor normal form, strictly below $\epsilon_0$.
-In order to compare such functions, we define a few abstract properties in
-\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} (adapted from 
-\href{../theories/html/hydras.Prelude.Iterates.html}{hydras.Prelude.Iterates}).
-
-\inputsnippets{T1Bridge/MonoDef}
-
-In \HydrasLib, these functions are defined by transfinite induction, using the \texttt{coq-equations} plug-in~\cite{sozeau:hal-01671777}. For that purpose, we defined a sigma-type
-\texttt{hydras.Epsilon0.E0.E0} of ordinal-terms in Cantor normal form.
-
 \subsection{A type for well-formed ordinal terms}
 \label{sect:E0-gaia-hydras}
-For compatibility's sake, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}.
+For compatibility's sake, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}. This type is mainly used in
+the definition with \texttt{Equations} of rapidly growing functions indexed by ordinals (see Sect.~\vref{sect:gaia-rapidly-growing}).
 
@@
 \inputsnippets{T1Bridge/E0EqMixin, T1Bridge/E0EqType}
+
+\subsubsection{Definitions}
 
@@
 \inputsnippets{T1Bridge/gE0h2gG2h}
+
+\subsubsection{Main lemmas about \texttt{E0}}
 
@@
 \inputsnippets{T1Bridge/gEpsilon0}
+
+\section{Rapidly growing arithmetic functions}
+\index{gaiabridge}{Rapidly growing functions}
+\label{sect:gaia-rapidly-growing}
+
+In this section, we consider a few families of arithmetic functions, indexed by ordinal numbers below $\epsilon_0$.
+The type of such families is \texttt{E0 -> nat -> nat} where
+\texttt{E0} is the type of ordinal terms in Cantor normal form, strictly below $\epsilon_0$.
+In order to compare such functions, we define a few abstract properties in
+\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} (adapted from 
+\href{../theories/html/hydras.Prelude.Iterates.html}{hydras.Prelude.Iterates}).
+
+\inputsnippets{T1Bridge/MonoDef}
+
+In \HydrasLib, these functions are defined by transfinite induction, using the \texttt{coq-equations} plug-in~\cite{sozeau:hal-01671777}. For that purpose, we defined a sigma-type
+\texttt{hydras.Epsilon0.E0.E0} of ordinal-terms in Cantor normal form.
 

### 2024-01-29 doc/hydras.tex

@@
 \subsubsection{Verbatim quotations}
+In some situations, we replace \alectr snippets with
+verbatim blocks.
 
-Quotations of \coq{} source  from other libraries (\coq's standard library, borrowed plug-ins)  are displayed as follows.
-
+\begin{itemize}
+\item When the quoted source belongs to some library on which we do not have the write permission, we cannot include
+  directives for generating snippets.
+ For instance, the following code belongs to \coq's standard library.
   \begin{Coqsrc}
@@
 
-We use also verbatim code inclusions when the examples lead to too long computations.
+\item 
+We use also verbatim code inclusions when the examples would lead to too long computations during the compilation and the
+documentation generation.
 
@@
 \end{Coqbad}
-
+\end{itemize}
  \subsection{Remark}
@@
 
-\part{Ackermann, G\"{o}del, Peano\,\dots}
+%\part{Ackermann, G\"{o}del, Peano\,\dots}
 
-\include{chap-intro-goedel}
+%\include{chap-intro-goedel}
 \include{chapter-primrec} 
-\include{chapter-fol}
-\include{chapter-natural-deduction}
-\include{chapter-lnn-lnt}
-\include{chapter-encoding}
-\include{chapter-expressible}
+% \include{chapter-fol}
+% \include{chapter-natural-deduction}
+% \include{chapter-lnn-lnt}
+% \include{chapter-encoding}
+% \include{chapter-expressible}
 

### 2024-01-29 doc/epsilon0-chapter.tex
Corrections in doc (#177)
* minor change

* small changes in doc

@@
 and the decomposition into Cantor normal form  is consistent 
-with other usual definitions of ordinals, for instance in \gaia~\cite{Gaia}, Schütte's book~\cite{schutte}, or larger ordinal notations cf. Chapter~\vref{chap:gamma0}.
+with usual definition of ordinals, for instance in \gaia~\cite{Gaia}, Schütte's book~\cite{schutte}, or larger ordinal notations~\vref{chap:gamma0}.
 
@@
 Unless explicitly mentioned, the term ``ordinal" will be used instead of
-``ordinal strictly less than \(\epsilon_0\)" (except in Chapter~\ref{chap:schutte} where it stands for ``countable ordinal'', and Chapter~\ref{chap:gamma0}.).
+``ordinal strictly less than \(\epsilon_0\)" (except in Chapter~\ref{chap:schutte} where it stands for ``countable ordinal'').
 

### 2024-01-30 doc/epsilon0-chapter.tex
improve Alectryon output

@@
 First, we prove that, for  any \texttt{LT}-accessible term $\alpha$, $\alpha$ is 
-strongly accessible too (\emph{i.e.} any well formed
-term (\texttt{cons $\alpha$ $n$ $\beta$})  is accessible).
-
-The proof is structured as an induction on $\alpha'$s accessibility. Let us consider
+strongly accessible too.
+The following proof is structured as an induction on $\alpha'$s accessibility. Let us consider
 any  accessible term $\alpha$.
@@
 
-Let \texttt{n:nat} and \texttt{beta:T1} such that (\texttt{cons alpha n beta}) is in normal form. 
-We prove first that \texttt{beta} is accessible,  which allows us to prove by well-founded induction on \texttt{beta}, 
-and natural induction on \texttt{n}, that (\texttt{cons alpha n beta}) is accessible.
+First, we prove that, for any $n$ and $\beta$, if (\texttt{cons $\alpha$ $n$ $\beta$}) is in normal form, then
+ $\beta$ is accessible.
+
+\inputsnippets{T1/betaAcc}
+
+The lemma \texttt{betaAcc} allows us to prove by well-founded induction on $\beta$, 
+and natural induction on $n$ that (\texttt{cons $\alpha$ $n$ $\beta$}) is accessible.
+
+\inputsnippets{T1/useBetaAcc}
+
 The proof, quite long, can be consulted in \href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}.

### 2024-02-02 doc/epsilon0-chapter.tex
minor changes in pdf doc

@@
 any well formed term \(\delta\) such that \(\delta<\alpha\). 
-But nothing guarantees that \(\delta\)  is strictly  less than \(\beta\) nor \(\gamma\), and we cannot use the induction hypotheses on   \(\beta\) nor \(\gamma\).
+% But nothing guarantees that \(\delta\)  is strictly  less than \(\beta\) nor \(\gamma\), and we cannot use the induction hypotheses on   \(\beta\) nor \(\gamma\).
 
@@
 
-The lemma \texttt{betaAcc} allows us to prove by well-founded induction on $\beta$, 
+The new hypothesis  \texttt{beta\_Acc} allows us to prove by well-founded induction on $\beta$, 
 and natural induction on $n$ that (\texttt{cons $\alpha$ $n$ $\beta$}) is accessible.
@@
 
+\subsection{Transfinite induction}
 \index{maths}{Transfinite induction}

### 2024-02-07 doc/chapter-primrec.tex
minor changes in pdf doc

@@
 
-\subsubsection{The constant function of value 0}
+\subsubsection{Constant functions}
 \label{sect:k0}
 
-The \emph{unary} function which always returns $0$ may be defined through the \emph{composition} construction (with $n=1$, $m=0$, and $h=\texttt{zero}$).
+The \emph{nullary} constant function which returns $0$ is
+simply the \texttt{zero} construction.
 
-% If we consider any value of $n$, the same construction
-% builds the constant function of type $\mathbb{N}^n \arrow \mathbb{N}$ which returns $0$.
+If we want to consider the \emph{unary} function which
+maps any natural number $i$ to $0$, we may built it
+through the \emph{composition} construction, instanciated
+with $n=1$, $m=0$, and $h=\texttt{zero}$.
 
-\subsubsection{Constant functions}
+\index{hydras}{Exercises}
 
-Let $k$ be some natural number; the unary constant function which always returns $k$ is built through $k$ nested compositions of the 
-successor function with the  unary constant function which returns $0$.
+\begin{exercise}
+Let $m$ and $k$ be two natural numbers; please build the primitive recursive function which maps any
+tuple $t\in \mathbb{N}^m$ to $k$.
+\end{exercise}
 
@@
 
-Module \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec} defines and export  the notation ``\texttt{$f$ =x= $g$}'' for ``\texttt{extEqual $n$ $f$ $g$}'' \footnote{in parsing mode, provided $f$ has explicitely the type (\texttt{naryFunc $n$}).}
+Module \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec} defines and export  the notation ``\texttt{$f$ =x= $g$}'' for ``\texttt{extEqual $n$ $f$ $g$}'' \footnote{in parsing mode, the provided $f$ should be explicitely typed as  (\texttt{naryFunc $n$}).}
+
+\vspace{6pt}
+\noindent
+\emph{From \href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}}
 
@@
 
-Inhabitants of type (\texttt{PrimRec $n$}) are not \coq{} functions like \texttt{Nat.mul}, or \texttt{Arith.Factorial.fact}, etc.
+Inhabitants of type (\texttt{PrimRec $n$}) are not \coq{} functions like \texttt{Nat.mul}, \linebreak \texttt{Arith.Factorial.fact}, etc.
 but terms of an abstract syntax for the language of primitive recursive functions.
@@
 This function is defined by mutual recursion,  together with the  function 
-(\texttt{evalprimRecS $n$ $m$}) of type 
+(\texttt{evalprimRecS $n$ $m$}) of type \linebreak
 ($\texttt{PrimRecs}\,n\,m \rightarrow  \texttt{Vector.t}\,(\texttt{naryFunc}\,n)\,m$).

### 2024-02-13 doc/chapter-primrec.tex
minor changes in pdf doc

@@
 The example in the preceding section clearly shows that, in order to prove that a given arithmetic function
-(defined in \gallina{} as usual) is primitive recursive, trying to \emph{type} a term  of type (\texttt{PrimRec $n$}) is not a good method, since such terms may be too large, even for simple arithmetic functions.
-The method proposed in the library \texttt{Ackermann.primRec} is the following one:
+(defined in \gallina{} as usual) is primitive recursive, trying to \emph{type} a term  of type (\texttt{PrimRec $n$}) is not always a good method, since such terms may be too large, even for simple arithmetic functions.
+The interactive method proposed in the library \texttt{Ackermann.primRec} is the following one:
 
@@
 \item Prove handy lemmas which may help to prove that a given function is primitive recursive. These lemmas can be considered as a way to build \emph{silently} large terms of 
-type (\texttt{PrimRec $n$}) in intermediate steps of the proof.
-More, we can associate \emph{tactics} with these lemmas.
+type (\texttt{PrimRec $p$})\footnote{Beware that $p$ is not always equal to $n$.} in intermediate steps of the proof.
+More, we will be able to associate \emph{tactics} with these lemmas.
 
@@
 
-The function we consider is defined and proven correct in
-Module~\href{../theories/html/hydras.Hydra.Battle_length.html}{Hydra.Battle\_length}.
+The following lemma (from 
+Module~\href{../theories/html/hydras.Hydra.Battle_length.html}{Hydra.Battle\_length}) relates the length of standard hydra battles with the $L_\alpha$ rapidly growing hierarchy (please see Sect.~\vref{sect:lalpha-section}).
 
-\input{movies/snippets/Battle_length/BattleLength}
+\inputsnippets{Battle_length/BattleLengtha}
 

### 2024-02-19 doc/chapter-primrec.tex
small corrections

--- 2024-02-07 doc/chapter-primrec.tex
@@
 (lemma  \texttt{const0\_NIsPR} of \texttt{primRec}). We carry out a proof by induction on $n$, the base case of which is already proven.
-Now, let us assume $n$ is \texttt{PR $0$}, with $x:\texttt{PrimRec}\,0$ as a ``realizer''.
+Now, let us assume $n$ is \texttt{PR $0$}, and call $(x:\texttt{PrimRec}\,0)$ its ``realizer''.
 Thus we would like to compose this constant function with the unary successor function.
 
-This is exactly the role of the instance \texttt{composeFunc 0 1} of the dependently typed
-function \texttt{composeFunc}, as shown by the following lemma.
-
-\vspace{4pt}
-\input{movies/snippets/PrimRecExamples/compose01}
-
-\vspace{4pt}
-Here is a quite simple proof of \texttt{const0\_NIsPR}.
+This is exactly the role of the function \texttt{(composeFunc 0 1)}. Here is a quite simple proof of \texttt{const0\_NIsPR}.
 
@@
 
-We have already proven that \texttt{Nat.add} is primitive recursive. The following alternative proof, --- more detalled  ---, 
+We have already proven that \texttt{Nat.add} is primitive recursive. The following alternative proof, --- more detailed  ---, 
 shows how to search and apply lemmas from
@@
 
-Let us look for some lemma which could help to prove some 
+Let us look for some lemma which could help to prove a given  
 recursive arithmetic binary function is primitive recursive.
